### 2016-11-15 Patient Report.Rmd
…

---
title: "Scoring Patients"
author: "Sarah Killcoyne"
date: "15 November 2016"
output: pdf_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(GGally)
library(plyr)
library(xlsx)
library(dplyr)
library(pander)
library(Hmisc)
library(ggdendro)
library(reshape2)

source('lib/load_patient_metadata.R')

my_custom_cor <- function(data, mapping, color = I("grey10"), sizeRange = c(1, 7), ...) {
  # get the x and y data to use the other code
  x <- eval(mapping$x, data)
  y <- eval(mapping$y, data)

  ct <- cor.test(x,y)
  sig <- symnum(
    ct$p.value, corr = FALSE, na = FALSE,
    cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),
    symbols = c("***", "**", "*", ".", " ")
  )

  r <- unname(abs(ct$estimate))
  rt <- format(r, digits=2)[1]

  # since we can't print it to get the strsize, just use the max size range
  cex <- max(sizeRange)

  # helper function to calculate a useable size
  percent_of_range <- function(percent, range) {
    percent * diff(range) + min(range, na.rm = TRUE)
  }

  sig_star_color = color
  if (ct$p.value <= 0.05)
    sig_star_color = I('firebrick')

  # plot the cor value
  p = ggally_text(
    label = as.character(rt), 
    mapping = aes(),
    xP = 0.5, yP = 0.5, 
    size = I(percent_of_range(cex * abs(r), sizeRange)),
    color = color,
    ... ) +

    # add the sig stars
    geom_text(
      aes_string( x = 0.8, y = 0.8 ),
      label = sig, 
      size = I( (percent_of_range(cex * abs(r), sizeRange)/2) ),
      #size = I(cex),
      color = sig_star_color,
      ... ) + 
    # remove all the background stuff and wrap it with a dashed line
    theme_classic() + 
    theme(
      panel.background = element_rect(
        color = color, 
        size = 2,
        linetype = "longdash"
      ), 
      axis.line = element_blank(), 
      axis.ticks = element_blank(), 
      axis.text.y = element_blank(), 
      axis.text.x = element_blank()
    )
}

panel.cor <- function(x, y, digits=2, prefix="", cex.cor) {
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- abs(cor(x, y, use= "pairwise.complete.obs"))
  txt <- format(c(r, 0.123456789), digits=digits)[1]
  txt <- paste(prefix, txt, sep="")
  if(missing(cex.cor)) cex <- 0.8/strwidth(txt)

  test <- cor.test(x,y, use= "pairwise.complete.obs")
  # borrowed from printCoefmat
  Signif <- symnum(test$p.value, corr = FALSE, na = FALSE,
                   cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),
                   symbols = c("***", "**", "*", ".", " "))

  text(0.5, 0.5, txt, cex = cex * r)
  text(.8, .8, Signif, cex=cex, col=2)
}

my_bars<-function(data, mapping, sizeRange=c(1,7),...) {
  # get the x and y data to use the other code
  x <- eval(mapping$x, data)
  y <- eval(mapping$y, data)

  name = as.character(mapping$x)
  ggplot(data=data,mapping=mapping) + geom_histogram(bins=10,col='green4',fill='palegreen',alpha=0.7) + 
    annotate("text",label=name, x=Inf, y=Inf, vjust=1,hjust=1,col=I('grey14'),fontface=2,size=max(sizeRange) )
}

panel.hist <- function(x, ...) {
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(usr[1:2], 0, 1.5) )
  h <- hist(x, plot = FALSE)
  breaks <- h$breaks; nB <- length(breaks)
  y <- h$counts; y <- y/max(y)
  rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
}

data = '~/Data/Ellie'

data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
plot.dir = paste(data, 'Analysis/multipcf_plots_fitted_perPatient', sep='/')

if (length(list.files(plot.dir)) <= 0)
  stop(paste("No analysis files found in", plot.dir ))

## Patient info file
patient.file = grep('patient_info.xls', data.files, value=T)
if (length(patient.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", data))

patient.info = read.patient.info(patient.file)

patient.info$Patient = gsub("/", "_", patient.info$Patient)
head(patient.info)

patient.data = ddply(patient.info, .(Patient, Status), summarise, years=paste(range(Endoscopy.Year), collapse='-'), 
      n.samples=length(Samplename), med.cellularity=median(Barretts.Cellularity,na.rm=T))

```

`r pander(arrange(patient.data, Status), justify='left', caption='Patient information')`

## Gamma = 250
```{r cors, echo=F, message=T, warning=F, fig.height=6, fig.width=6, fig.align='left'}
min.probes=67
gamma2 = 250

cor.scores = as.data.frame(matrix(nrow=length(patient.data$Patient), ncol=2, dimnames=list(patient.data$Patient, c('similar','different'))))
for (patient.name in unique(patient.info$Patient)  )  {
  message(patient.name)
  patient.plot.dir = paste(plot.dir, patient.name, sep='/')

  segvals = read.table(paste(patient.plot.dir, "/",patient.name,"_segmentedCoverage_fitted_gamma",gamma2,".txt",sep=""),sep="\t",stringsAsFactors=F,header=T)
  message( paste(length(which(segvals$n.probes < min.probes)), ' probes below the minimum probe count (',min.probes,')', sep=''))

  segvals = segvals[segvals$n.probes>=min.probes,]

  if(nrow(segvals)>0) {

      HC = hclust(dist(t(segvals[,-(1:5)])))
      gg1 = ggdendrogram(HC, rotate=T) + labs(title = paste(patient.name,": fitted segment coverage"),x="")
      #print(gg1)

      # Normalize  (value-mean(value))/sd(value)
      normalised.segvals = segvals[,-(1:5)]
      for(c in 1:nrow(normalised.segvals)) {
        normalised.segvals[c,] = (normalised.segvals[c,]-mean(unlist(normalised.segvals[c,])))/sd(unlist(normalised.segvals[c,]))
      }

      HC = hclust(dist(t(normalised.segvals)))
      gg2 = ggdendrogram(HC, rotate=T) + labs(title=paste(patient.name,": normalised fitted segment coverage", x=""))
      #print(gg2)

      no.samples = ncol(normalised.segvals)
      no.rows = ceiling(no.samples/3)

      gg3 = ggplot( melt(normalised.segvals), aes(x=value) ) + 
        facet_wrap(~variable,scales = "free_x") + 
        geom_histogram( fill=I("blue"),col=I("lightblue"),binwidth=0.3) + labs(title=paste(patient.name,'normalised segment coverage'), x='', y='')
      print(gg3)
      #pt.plots[[patient.name]][['norm.hist']] = gg

      # par(mfrow=c(3, no.rows))
      # for(s in 1:no.samples) 
      #   hist(normalised.segvals[,s],col="blue",main=names(normalised.segvals)[s],xlab="normalised coverage")

      if(nrow(segvals)>=3) {
        # order by date of endoscopy
        patient = subset(patient.info, Patient == patient.name)
        patient = arrange(patient, Endoscopy.Year, Pathology)

        ## TODO Mistake in earlier version of the patient file caused this will be fixed for the next run
        snames = grep('_1072(5|9)$', patient$Samplename)
        if ( length(snames > 0) ) {
          patient$Samplename[snames] =  paste( sub('-','_', patient$Plate.Index[snames]), '10725_10729', sep='_' )
        }

        sample.normalised.segvals = normalised.segvals[, intersect(colnames(normalised.segvals), patient$Samplename)]

        rc = rcorr(as.matrix(sample.normalised.segvals), type='pearson')

        sim = 0 
        diff = 0
        for (i in 1:ncol(rc$r)) {
          len = length(which(rc$P[,i] < 0.05))

          sim = sim + signif( sum(rc$r[which(rc$P[,i] < 0.05 & rc$r[,i] > 0),i ])/len, 2)
          diff = diff + signif(sum(rc$r[which(rc$P[,i] < 0.05 & rc$r[,i] < 0),i ])/len,2)
        }  
        cor.scores[patient.name,] = c(sim,diff)

        # pairs = ggpairs( sample.normalised.segvals,
        #                  lower=list(continuous=wrap("points", colour='darkblue', shape=20, alpha=0.3, size=3)),
        #                  diag=list(continuous=wrap(my_bars, sizeRange=c(1,5))),
        #                  upper=list(continuous=my_custom_cor),
        #                  axisLabels='show', title=patient.name, columnLabels=paste(paste(patient$Endoscopy.Year, ' (',patient$Pathology, ')', sep=''), patient$Plate.Index, sep='\n') )
        # 
        # print(pairs  + theme( # this doesn't work...
        #   text = element_text(size=20, face='bold')
        # ))

      }
    }
  }
pander(cor.scores, justify='left')

cor.scores$Status = patient.data[which(patient.data$Patient %in% rownames(cor.scores)), 'Status']

```
\pagebreak

`r pander(cor.scores[order(cor.scores$Status, cor.scores$similar),], justify='left', split.tables=F)`

### 2016-11-15 lib/load_patient_metadata.R

@@ -11,9 +11,9 @@ read.patient.info<-function(file) {
   for (i in 1:length(getSheets(loadWorkbook(file)))) {
     print(i)
     ws = read.xlsx2(file, sheetIndex=i, stringsAsFactors=F, header=T)
-    ws = ws[which( ws$Patient.ID != ""), grep('Patient|Endoscopy.Year|Pathology$|Progressor|Plate.Index|SLX', colnames(ws), value=T, ignore.case=T)]
+    ws = ws[which( ws$Patient.ID != ""), grep('Patient|Endoscopy.Year|Pathology$|Progressor|Plate.Index|SLX|cellularity', colnames(ws), value=T, ignore.case=T)]
     head(ws)
-    head(patient.info)
+    #head(patient.info)
     
     slx.cols = grep('SLX', colnames(ws), value=T)
     if (length(slx.cols) > 1) {
@@ -28,8 +28,10 @@ read.patient.info<-function(file) {
                grep('Endoscopy.year',colnames(ws), ignore.case=T),
                grep('Pathology',colnames(ws),ignore.case=T),
                grep('Plate.Index',colnames(ws),ignore.case=T),
-               grep('SLX',colnames(ws),ignore.case=T))]
-    colnames(ws) = c('Patient','Status','Endoscopy.Year','Pathology','Plate.Index','SLX.ID')
+               grep('SLX',colnames(ws),ignore.case=T),
+               grep('cellularity',colnames(ws),ignore.case=T)) ]
+    
+    colnames(ws) = c('Patient','Status','Endoscopy.Year','Pathology','Plate.Index','SLX.ID','Barretts.Cellularity')
     
     if (is.null(patient.info)) {
       patient.info = ws
@@ -43,7 +45,9 @@ read.patient.info<-function(file) {
   patient.info$Pathology = factor(strip.whitespace(patient.info$Pathology))
   patient.info$Endoscopy.Year = as.numeric(patient.info$Endoscopy.Year)
   patient.info$Plate.Index = strip.whitespace(patient.info$Plate.Index)
+  patient.info$Barretts.Cellularity = as.integer(patient.info$Barretts.Cellularity)
+  patient.info$Samplename = gsub('-', '_', paste(patient.info$Plate.Index,strip.whitespace(patient.info$SLX.ID),sep="_"))
   patient.info = arrange(patient.info, Status, Patient, Endoscopy.Year)
-  
+
   return(patient.info)
 }

### 2016-11-16 Patient Report.Rmd
added markdown file for reporting

@@
 
-cor.scores = as.data.frame(matrix(nrow=length(patient.data$Patient), ncol=2, dimnames=list(patient.data$Patient, c('similar','different'))))
+cor.scores = as.data.frame(matrix(nrow=length(patient.data$Patient), ncol=3, dimnames=list(patient.data$Patient, c('similar','different','n.probes'))))
 for (patient.name in unique(patient.info$Patient)  )  {
@@
   segvals = read.table(paste(patient.plot.dir, "/",patient.name,"_segmentedCoverage_fitted_gamma",gamma2,".txt",sep=""),sep="\t",stringsAsFactors=F,header=T)
+
   message( paste(length(which(segvals$n.probes < min.probes)), ' probes below the minimum probe count (',min.probes,')', sep=''))
@@
 
-          sim = sim + signif( sum(rc$r[which(rc$P[,i] < 0.05 & rc$r[,i] > 0),i ])/len, 2)
-          diff = diff + signif(sum(rc$r[which(rc$P[,i] < 0.05 & rc$r[,i] < 0),i ])/len,2)
+          sim = sim + signif( sum(rc$r[which(rc$P[,i] < 0.05 & rc$r[,i] > 0),i ]), 2)
+          diff = diff + signif(sum(rc$r[which(rc$P[,i] < 0.05 & rc$r[,i] < 0),i ]),2)
         }  
-        cor.scores[patient.name,] = c(sim,diff)
+        cor.scores[patient.name,] = c(sim,diff,nrow(sample.normalised.segvals))
 
@@
   }
-pander(cor.scores, justify='left')
-
 cor.scores$Status = patient.data[which(patient.data$Patient %in% rownames(cor.scores)), 'Status']
@@
 
-`r pander(cor.scores[order(cor.scores$Status, cor.scores$similar),], justify='left', split.tables=F)`
+```{r scores, echo=F, message=T, warning=F, fig.height=4,fig.width=4,fig.align='left'} 
+pander(cor.scores[order(cor.scores$Status, cor.scores$similar),], justify='left')
+
+ggplot(cor.scores, aes(x=similar,y=different, color=Status)) + geom_point()
+
+ggplot(cor.scores, aes(y=abs(similar+different), x=c(1:nrow(cor.scores)), color=Status)) + geom_point()
+
+```

### 2016-12-01 Patient Report.Rmd
Mapping genes from the cosmic gene consensus onto the variable regions.

@@
 date: "15 November 2016"
-output: pdf_document
+output: word_document
 ---
@@
 library(reshape2)
+library(GenomicRanges)
+
+library(tibble)
 
 source('lib/load_patient_metadata.R')
-
-my_custom_cor <- function(data, mapping, color = I("grey10"), sizeRange = c(1, 7), ...) {
-  # get the x and y data to use the other code
-  x <- eval(mapping$x, data)
-  y <- eval(mapping$y, data)
-
-  ct <- cor.test(x,y)
-  sig <- symnum(
-    ct$p.value, corr = FALSE, na = FALSE,
-    cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),
-    symbols = c("***", "**", "*", ".", " ")
-  )
-
-  r <- unname(abs(ct$estimate))
-  rt <- format(r, digits=2)[1]
-
-  # since we can't print it to get the strsize, just use the max size range
-  cex <- max(sizeRange)
-
-  # helper function to calculate a useable size
-  percent_of_range <- function(percent, range) {
-    percent * diff(range) + min(range, na.rm = TRUE)
-  }
-
-  sig_star_color = color
-  if (ct$p.value <= 0.05)
-    sig_star_color = I('firebrick')
-
-  # plot the cor value
-  p = ggally_text(
-    label = as.character(rt), 
-    mapping = aes(),
-    xP = 0.5, yP = 0.5, 
-    size = I(percent_of_range(cex * abs(r), sizeRange)),
-    color = color,
-    ... ) +
-
-    # add the sig stars
-    geom_text(
-      aes_string( x = 0.8, y = 0.8 ),
-      label = sig, 
-      size = I( (percent_of_range(cex * abs(r), sizeRange)/2) ),
-      #size = I(cex),
-      color = sig_star_color,
-      ... ) + 
-    # remove all the background stuff and wrap it with a dashed line
-    theme_classic() + 
-    theme(
-      panel.background = element_rect(
-        color = color, 
-        size = 2,
-        linetype = "longdash"
-      ), 
-      axis.line = element_blank(), 
-      axis.ticks = element_blank(), 
-      axis.text.y = element_blank(), 
-      axis.text.x = element_blank()
-    )
-}
-
-panel.cor <- function(x, y, digits=2, prefix="", cex.cor) {
-  usr <- par("usr"); on.exit(par(usr))
-  par(usr = c(0, 1, 0, 1))
-  r <- abs(cor(x, y, use= "pairwise.complete.obs"))
-  txt <- format(c(r, 0.123456789), digits=digits)[1]
-  txt <- paste(prefix, txt, sep="")
-  if(missing(cex.cor)) cex <- 0.8/strwidth(txt)
-
-  test <- cor.test(x,y, use= "pairwise.complete.obs")
-  # borrowed from printCoefmat
-  Signif <- symnum(test$p.value, corr = FALSE, na = FALSE,
-                   cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),
-                   symbols = c("***", "**", "*", ".", " "))
-
-  text(0.5, 0.5, txt, cex = cex * r)
-  text(.8, .8, Signif, cex=cex, col=2)
-}
-
-my_bars<-function(data, mapping, sizeRange=c(1,7),...) {
-  # get the x and y data to use the other code
-  x <- eval(mapping$x, data)
-  y <- eval(mapping$y, data)
-
-  name = as.character(mapping$x)
-  ggplot(data=data,mapping=mapping) + geom_histogram(bins=10,col='green4',fill='palegreen',alpha=0.7) + 
-    annotate("text",label=name, x=Inf, y=Inf, vjust=1,hjust=1,col=I('grey14'),fontface=2,size=max(sizeRange) )
-}
-
-panel.hist <- function(x, ...) {
-  usr <- par("usr"); on.exit(par(usr))
-  par(usr = c(usr[1:2], 0, 1.5) )
-  h <- hist(x, plot = FALSE)
-  breaks <- h$breaks; nB <- length(breaks)
-  y <- h$counts; y <- y/max(y)
-  rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
-}
 
@@
 ## Patient info file
-patient.file = grep('patient_info.xls', data.files, value=T)
+patient.file = grep('All_patient_info.xls', data.files, value=T)
 if (length(patient.file) != 1)
@@
 
-patient.data = ddply(patient.info, .(Patient, Status), summarise, years=paste(range(Endoscopy.Year), collapse='-'), 
+patient.info = arrange(patient.info, Status, Patient, Endoscopy.Year, Pathology)
+
+sum.patient.data = ddply(patient.info, .(Patient, Status), summarise, years=paste(range(Endoscopy.Year), collapse='-'), 
       n.samples=length(Samplename), med.cellularity=median(Barretts.Cellularity,na.rm=T))
 
+#pander(arrange(patient.data, Status), justify='left', caption='Patient information')
 ```
-
-`r pander(arrange(patient.data, Status), justify='left', caption='Patient information')`
 
@@
 ```{r cors, echo=F, message=T, warning=F, fig.height=6, fig.width=6, fig.align='left'}
-min.probes=67
+## Thresholds
+min.probes=67 # equiv to 1MB
 gamma2 = 250
+sd.threshold = 0.08
+gain.threshold = 1.1
+loss.threshold = 0.9
 
-cor.scores = as.data.frame(matrix(nrow=length(patient.data$Patient), ncol=3, dimnames=list(patient.data$Patient, c('similar','different','n.probes'))))
+patient.data = list()
+
 for (patient.name in unique(patient.info$Patient)  )  {
@@
 
+  # ORDER segvals by endoscopy year etc
+  cur.pt = subset(patient.info, Patient == patient.name)
+  segvals[, c(colnames(segvals[1:5]), cur.pt$Samplename)]
+
+  patient.data[[patient.name]][['seg.vals']] = segvals
+  patient.data[[patient.name]][['info']] = cur.pt
+
   if(nrow(segvals)>0) {
-
       HC = hclust(dist(t(segvals[,-(1:5)])))
@@
       }
+      patient.data[[patient.name]][['norm.seg.vals']] = normalised.segvals
+
+      # Variable Regions
+      sds = apply(segvals[,-(1:5)],1,sd)
+      means = apply(segvals[,-(1:5)],1,mean)
+      CofV = sds/means
+
+      variable.region.indices = which(sds >= sd.threshold & segvals$n.probes >= min.probes)
+      variable.regions = segvals[variable.region.indices, patient.data[[patient.name]]$info$Samplename] 
+      patient.data[[patient.name]][['variable.regions']] = cbind(segvals[variable.region.indices,c(1:5)],variable.regions )
 
@@
         patient = arrange(patient, Endoscopy.Year, Pathology)
-
-        ## TODO Mistake in earlier version of the patient file caused this will be fixed for the next run
-        snames = grep('_1072(5|9)$', patient$Samplename)
-        if ( length(snames > 0) ) {
-          patient$Samplename[snames] =  paste( sub('-','_', patient$Plate.Index[snames]), '10725_10729', sep='_' )
+        }
+        sample.normalised.segvals = normalised.segvals[, intersect(colnames(normalised.segvals), patient$Samplename)]
+      }
 }
 
-        sample.normalised.segvals = normalised.segvals[, intersect(colnames(normalised.segvals), patient$Samplename)]
-
-        rc = rcorr(as.matrix(sample.normalised.segvals), type='pearson')
-
-        sim = 0 
-        diff = 0
-        for (i in 1:ncol(rc$r)) {
-          len = length(which(rc$P[,i] < 0.05))
-
-          sim = sim + signif( sum(rc$r[which(rc$P[,i] < 0.05 & rc$r[,i] > 0),i ]), 2)
-          diff = diff + signif(sum(rc$r[which(rc$P[,i] < 0.05 & rc$r[,i] < 0),i ]),2)
-        }  
-        cor.scores[patient.name,] = c(sim,diff,nrow(sample.normalised.segvals))
-
-        # pairs = ggpairs( sample.normalised.segvals,
-        #                  lower=list(continuous=wrap("points", colour='darkblue', shape=20, alpha=0.3, size=3)),
-        #                  diag=list(continuous=wrap(my_bars, sizeRange=c(1,5))),
-        #                  upper=list(continuous=my_custom_cor),
-        #                  axisLabels='show', title=patient.name, columnLabels=paste(paste(patient$Endoscopy.Year, ' (',patient$Pathology, ')', sep=''), patient$Plate.Index, sep='\n') )
-        # 
-        # print(pairs  + theme( # this doesn't work...
-        #   text = element_text(size=20, face='bold')
-        # ))
-
-      }
-    }
-  }
-cor.scores$Status = patient.data[which(patient.data$Patient %in% rownames(cor.scores)), 'Status']
-
+## Note the number of probes per patient/sample
 ```
@@
 ```{r scores, echo=F, message=T, warning=F, fig.height=4,fig.width=4,fig.align='left'} 
-pander(cor.scores[order(cor.scores$Status, cor.scores$similar),], justify='left')
+#pander(cor.scores[order(cor.scores$Status, cor.scores$similar),], justify='left')
 
-ggplot(cor.scores, aes(x=similar,y=different, color=Status)) + geom_point()
+#ggplot(cor.scores, aes(x=similar,y=different, color=Status)) + geom_point()
 
-ggplot(cor.scores, aes(y=abs(similar+different), x=c(1:nrow(cor.scores)), color=Status)) + geom_point()
+#ggplot(cor.scores, aes(y=abs(similar+different), x=c(1:nrow(cor.scores)), color=Status)) + geom_point()
+```
+
+## Variable Regions
+```{r var_reg, echo=F, message=F, warning=F}
+rm(prog.regions, np.regions)
+for (patient.name in sum.patient.data$Patient) {
+  patient = patient.data[[patient.name]]
+
+  sum.patient.data[which(sum.patient.data$Patient == patient.name), 'var.regions'] = nrow(patient$variable.regions)
+
+  if (subset(sum.patient.data, Patient == patient.name)$Status == 'P') {
+    if (!exists('prog.regions')) {
+      prog.regions = patient$variable.regions[,c('chrom','start.pos','end.pos')]
+    } else {
+      prog.regions = rbind(prog.regions, patient$variable.regions[,c('chrom','start.pos','end.pos')])
+    }
+  } else {
+    if (!exists('np.regions')) {
+      np.regions = patient$variable.regions[,c('chrom','start.pos','end.pos')]
+    } else {
+      np.regions = rbind(np.regions, patient$variable.regions[,c('chrom','start.pos','end.pos')])
+    }
+  }
+}
+
+pander(sum.patient.data %>% group_by(Status) %>% summarise(mean.num.regions = mean(var.regions), sd.num.regions = sd(var.regions)), justify='left')
+
+prog.regions$length = prog.regions$end.pos - prog.regions$start.pos
+np.regions$length = np.regions$end.pos - np.regions$start.pos
+
+chrs = read.table('hg19_chrom.txt',sep='\t', header=T)
+prog.regions = merge(prog.regions, chrs, by.x='chrom', by.y='chr') 
+np.regions = merge(np.regions, chrs, by.x='chrom', by.y='chr')
+
+prog.regions$chr.ratio = prog.regions$length.x/prog.regions$length.y
+np.regions$chr.ratio = np.regions$length.x/np.regions$length.y
+
+ttR = t.test(prog.regions$chr.ratio, np.regions$chr.ratio)
+
+np = np.regions %>%
+  group_by(chrom) %>%
+  summarise(lengths = sum(length.x))
+
+prog = prog.regions %>%
+  group_by(chrom) %>%
+  summarise(lengths = sum(length.x))
+
+m = merge(prog, np, by='chrom', all.x=T)
+tt = t.test(m$lengths.x, m$lengths.y)
+
+plot(hclust( dist(chrs$length)), main='Chromosome cluster by length')
+
+## Grouping the chromosomes by length 1-12, 13-22
+
+t.test(subset(np.regions, chrom %in% c(1:12))$chr.ratio, subset(np.regions, chrom %in% c(13:22))$chr.ratio)
+
+t.test(subset(prog.regions, chrom %in% c(1:12))$chr.ratio, subset(prog.regions, chrom %in% c(13:22))$chr.ratio)
+
+df = data.frame( 'short' = rbind( 'np' = sum(table(np.regions$chrom)[1:12]),
+       'p' = sum(table(prog.regions$chrom)[1:12])),
+                 'long' = rbind( 'np' = sum(table(np.regions$chrom)[13:22], na.rm=T),
+       'p' = sum(table(prog.regions$chrom)[13:22], na.rm=T)))
+
+ftL = fisher.test(df)
+```
+
+So it appears that adjusting for chromosome length (looking at the ratio of the chromosome rather than the raw bp) there is still a significant difference between progressers and non: t-test p-value `r ttR$p.value`
+
+There appears to be no enrichment based on the longer vs shorter chromosomes either (fisher `r ftL$p.value`)
+
+```{r cor, echo=F, message=T, warning=F, fig.height=4,fig.width=4, fig.align='left'}
+
+lower<-function(rc) {
+ ut = upper.tri( rc$r , diag=T) 
+ rc$r[ut] = 0
+ rc$P[ut] = NA
+ return(rc)
+}
+
+sig.rc<-function(rc) {
+  sig = rc$P > 0.05
+  rc$r[sig] = 0
+  return(rc)
+}
+
+cors = list()
+for (name in names(patient.data)) {
+
+  sample.normalised.segvals = patient.data[[name]]$norm.seg.vals # prog
+  sample.normalised.segvals =  sample.normalised.segvals[, patient.data[[name]]$info$Samplename ]
+
+  colnames(sample.normalised.segvals) =  paste(patient.data[[name]]$info$Endoscopy.Year,  c(1:nrow(patient.data[[name]]$info)) , sep='.')
+
+  #sample.normalised.segvals = patient.data$AD0361$norm.seg.vals # non-p
+  rc = rcorr(as.matrix(sample.normalised.segvals), type='pearson')
+  #rc = sig.rc(lower(rc))
+  rc = sig.rc(rc)
+
+  diff = unlist(apply(rc$r, 1, function(x)  sum(x[x < 0])-1 ))/nrow(rc$r)
+  sim = unlist(apply(rc$r, 1, function(x)  sum(x[x > 0])-1 ))/nrow(rc$r)
+
+  t.test(abs(sim), abs(diff))
+
+  cors[[name]][['sim']] = sim
+  cors[[name]][['diff']] = diff
+}
+
+unlist(lapply(cors, function(x) t.test(abs(x$sim), abs(x$diff))$p.value))
+
+simPNP = wilcox.test(unlist(lapply(cors[subset(sum.patient.data, Status == 'P')$Patient], function(x) sum(x$sim))),
+            unlist(lapply(cors[subset(sum.patient.data, Status == 'NP')$Patient], function(x) sum(x$sim))))
+
+diffPNP = wilcox.test(unlist(lapply(cors[subset(sum.patient.data, Status == 'P')$Patient], function(x) sum(x$diff))),
+            unlist(lapply(cors[subset(sum.patient.data, Status == 'NP')$Patient], function(x) sum(x$diff))))
+```
+
+There's no clear difference (currently) between P and NP for the similar correlation (`r simPNP$p.value`) or the inverse (`r diffPNP$p.value`) correlations.
+There are other ways we may be able to look at this still.
+
+## Look at Cancer Consensus Genes in the variable regions
+
+```{r, hmm, echo=F, message=F, warning=F}
+ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
+
+noLoc = grep(':-',ccgenes$Genome.Location)
+noLocGenes = ccgenes[noLoc,]
+
+ccgenes = ccgenes[-noLoc,]
+
+locs = as.data.frame(do.call(rbind, strsplit(ccgenes$Genome.Location, ':|-')))
+colnames(locs) = c('chr','start','end')
+locs[c('start','end')] = lapply(locs[c('start','end')], function(x) as.numeric(as.character(x)))
+
+ccgenes = cbind(locs, ccgenes)
+ccgenes = makeGRangesFromDataFrame(ccgenes, keep.extra.columns = T, start.field = 'start', end.field = 'end')
+
+# ov = findOverlaps(ccgenes, ccgenes)
+# ov = ov[queryHits(ov) != subjectHits(ov)]
+# ccgenes = ccgenes[-subjectHits(ov)[1]]
+
+cancer.consensus.genes<-function(reg, ccgenes) {
+  reg = as_tibble(reg)
+  gr = GenomicRanges::makeGRangesFromDataFrame(reg, start.field = 'start.pos', end.field = 'end.pos')
+
+  ov = findOverlaps(ccgenes, gr)
+  reg$Gene.Symbols = NA
+  for (hit in unique(subjectHits(ov))) {
+    reg[hit, 'Gene.Symbols'] =  paste(ccgenes[queryHits(ov)[which(subjectHits(ov) == hit)],]$Gene.Symbol, collapse=',')
+  }
+
+  rwg = cbind(reg, reg[,-(1:5)])
+  return(rwg)
+}
+
+rm(var.regions)
+for (patient.name in names(patient.data)) {
+  print(patient.name)
+  regions = patient.data[[patient.name]]$variable.regions[,c('chrom','arm','start.pos','end.pos')]
+  if (nrow(regions) <= 0) next
+
+  reg = cancer.consensus.genes(regions, ccgenes)
+  x = patient.data[[patient.name]]$variable.regions[which(!is.na(reg$Gene.Symbols)), patient.data[[patient.name]]$info$Samplename]
+
+  gl = apply(x, 1, function(y)  ifelse(length(which(y >= 1.1)) > length(which(y <= 0.9)), 'gain', 'loss') )
+  reg[which(!is.na(reg$Gene.Symbols)), 'CN'] = gl
+  reg = reg[,c('chrom','arm','start.pos','end.pos','CN','Gene.Symbols')]
+
+  for (i in 1:nrow(reg)) {
+    if (is.na(reg[i,'Gene.Symbols'])) next
+    genes = unlist(strsplit(reg[i,'Gene.Symbols'], ','))
+
+    reg[i,'Gene.Symbols'] = genes[1]
+
+    if (length(genes) > 1)
+      reg = rbind(reg, cbind(reg[i, (1:5)], 'Gene.Symbols' = genes[2:length(genes)]))
+  }
+  reg = arrange(reg, chrom, arm, start.pos)
+
+  status = subset(sum.patient.data, Patient == patient.name)$Status
+
+  reg = cbind(patient.name, status, reg)
+
+  if (!exists('var.regions')) {
+    var.regions = reg
+  } else {
+    var.regions = rbind(var.regions, reg)
+  }
+}
+
+var.regions = var.regions[!is.na(var.regions$Gene.Symbols),]
+var.regions$patient.name = as.character(var.regions$patient.name)
+
+var.regions = var.regions[-grep('CDKN2A\\(p14\\)', var.regions$Gene.Symbols),]
+
+gene.cnt = var.regions %>%
+  group_by(Gene.Symbols, CN) %>%
+  summarise(Progressors=length(which(status == 'P')), Non.Progressors=length(which(status == 'NP')) )
+gene.cnt[,unique(var.regions$patient.name)] = 0
+gene.cnt = gene.cnt[which(!is.na(gene.cnt$Gene.Symbols)),]
+
+for (i in 1:nrow(gene.cnt)) {
+  gene = as.character(gene.cnt[i,'Gene.Symbols'])
+  type = as.character(gene.cnt[i,'CN'])
+#  print(paste(gene, type))
+#  print(subset(var.regions, Gene.Symbols == gene & CN == type)$patient.name)
+  gene.cnt[i,unique(subset(var.regions, Gene.Symbols == gene & CN == type)$patient.name)] = 1
+}
+
+gene.cnt = arrange(gene.cnt, -Progressors)
+gene.p = subset(gene.cnt, Progressors >= Non.Progressors)
+
+m = melt(gene.p[1:50,c(1:4)], id.vars=c('Gene.Symbols', 'CN'))
+m$Gene.Symbols = factor(m$Gene.Symbols, levels= m$Gene.Symbols[c(nrow(m):1)] ) # order
+m$CN = factor(m$CN)
+
+p = ggplot(m, aes(y=Gene.Symbols, x=variable)) + 
+  geom_tile(aes(fill=value), color='white') + 
+  scale_fill_gradient('n patients',low='white',high='royalblue4') + labs(x="",y='CGC Genes')
+p + geom_point(data=m[m$CN == 'gain' & m$value > 0,], aes(label='gain',y=Gene.Symbols, x=variable), size=2, shape=24, fill='red', color='red') +
+  geom_point(data=m[m$CN == 'loss' & m$value > 0,], aes(label='loss',y=Gene.Symbols, x=variable), color='green4', size=2, shape=25, fill='green4')
+
+gc = gene.cnt %>% group_by(Gene.Symbols) %>% summarise(Progressors=sum(Progressors), Non.Progressors=sum(Non.Progressors))
+gc = arrange(subset(gc, Progressors >= Non.Progressors), -Progressors)
+m = melt(gc[1:50,], id.vars=c('Gene.Symbols'))
+m$Gene.Symbols = factor(m$Gene.Symbols, levels= m$Gene.Symbols[c(nrow(m):1)] ) # order
+ggplot(m, aes(y=Gene.Symbols, x=variable)) + 
+  geom_tile(aes(fill=value), color='white') + 
+  scale_fill_gradient('n patients',low='white',high='royalblue4') + labs(x="",y='CGC Genes')
 
 ```
+
+## Is HGD over represented for variable regions?
+```{r, hgd, echo=F, message=F, warning=F}
+
+pvnp = data.frame(matrix(nrow=length(subset(sum.patient.data, Status == 'P')$Patient), ncol=2, dimnames=list(subset(sum.patient.data, Status == 'P')$Patient, c('hgd','be'))))    
+for (name in subset(sum.patient.data, Status == 'P')$Patient) {
+  print(name)
+  sn = grep('HGD', patient.data[[name]]$info$Pathology)
+  if (length(sn) <= 0 | nrow(patient.data[[name]]$variable.regions) <= 0) next
+  samplenames = patient.data[[name]]$info$Samplename
+
+  ## HGD
+  pvnp[name, 'hgd'] = length(which(patient.data[[name]]$variable.regions[samplenames[sn]] >= gain.threshold | 
+                                patient.data[[name]]$variable.regions[samplenames[sn]] <= loss.threshold))
+
+  # BE-LGD
+  pvnp[name, 'be'] = length(which(patient.data[[name]]$variable.regions[samplenames[-sn]] >= gain.threshold | 
+                                patient.data[[name]]$variable.regions[samplenames[-sn]] <= loss.threshold))
+}
+
+# No difference
+
+tt = t.test(pvnp$hgd, pvnp$be)
+pvnp = cbind(pvnp, (pvnp %>% rowwise() %>% mutate(total=hgd+be, hgd.ratio=round(hgd/(hgd+be),3) ))[c('total','hgd.ratio')])
+
+pander(pvnp, justify='left', 
+       caption=paste("Variable regions found in HGD vs non-HGD samples of progessor patients.
+There's no significant difference between them (t-test p=",signif(tt$p.value,2),').
+Not all patients have HGD samples (NAs).',sep=''))
+
+```

### 2016-12-01 output_variable_regions.R

@@ -1,11 +1,12 @@
 #!/usr/bin/env Rscript
 args = commandArgs(trailingOnly=TRUE)
 
-#if (length(args) < 3) {
-#  stop("Data directory, patient name, and plot output directory required as arguments")
-#}
+if (length(args) < 3) {
+  stop("Missing required arguments: <data directory> <output directory> <cosmic tsv gene file>")
+}
 
 library(xlsx) 
+library(GenomicRanges)
 
 source("lib/fastPCF.R")
 source('lib/load_patient_metadata.R')
@@ -13,20 +14,40 @@ source('lib/load_patient_metadata.R')
 Negate('%in%') -> `%nin%`
 
 data = args[1]
-gamma2 = args[2]
-outdir = args[3]
+outdir = args[2]
+cosmic = args[3]
 
-data = '~/Data/Ellie/Analysis'
 gamma2 = 250
-outdir = "~/Documents/MRC-OwnCloud/Progressor Study/Variable Regions"
+if (length(args) == 4)
+  gamma2 = args[4]
+
+
+#data = '~/Data/Ellie/Analysis'
+#gamma2 = 250
+#outdir = "~/Documents/MRC-OwnCloud/Progressor Study/Variable Regions"
+# cosmic = '~/Data/CosmicCensusGenes.tsv'
 
 plot.dir = paste(data, 'multipcf_plots_fitted_perPatient', sep='/')
 print(paste("Plot directory:", plot.dir))
 if (!dir.exists(plot.dir))  
   stop("multipcf needs to be run first")
 
+ccgenes = read.table(cosmic, sep='\t', header=T, stringsAsFactors=F)
+
+noLoc = grep(':-',ccgenes$Genome.Location)
+noLocGenes = ccgenes[noLoc,]
+ccgenes = ccgenes[-noLoc,]
+
+locs = as.data.frame(do.call(rbind, strsplit(ccgenes$Genome.Location, ':|-')))
+colnames(locs) = c('chr','start','end')
+locs[c('start','end')] = lapply(locs[c('start','end')], function(x) as.numeric(as.character(x)))
+
+ccgenes = cbind(locs, ccgenes)
+ccgenes = makeGRangesFromDataFrame(ccgenes, keep.extra.columns = T, start.field = 'start', end.field = 'end')
+
+
 ## Patient info file
-patient.file = grep('patient_info.xlsx', data.files, value=T)
[…]

### 2017-01-04 Clustering.Rmd
Looking at different clustering methods

---
title: "Sample Clustering"
author: "Sarah Killcoyne"
date: "12/6/2016"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(GGally)
library(plyr)
library(xlsx)
library(dplyr)
library(pander)
library(Hmisc)
library(ggdendro)
library(reshape2)
library(GenomicRanges)

library(tibble)
library(mclust)

source('lib/load_patient_metadata.R')

data = '~/Data/Ellie'

data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
plot.dir = paste(data, 'Analysis/multipcf_plots_fitted_perPatient', sep='/')

if (length(list.files(plot.dir)) <= 0)
  stop(paste("No analysis files found in", plot.dir ))

## Patient info file
patient.file = grep('All_patient_info.xls', data.files, value=T)
if (length(patient.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", data))

patient.info = read.patient.info(patient.file)

patient.info$Patient = gsub("/", "_", patient.info$Patient)
head(patient.info)

patient.info = arrange(patient.info, Status, Patient, Endoscopy.Year, Pathology)

sum.patient.data = ddply(patient.info, .(Patient, Status), summarise, years=paste(range(Endoscopy.Year), collapse='-'), 
      n.samples=length(Samplename), med.cellularity=median(Barretts.Cellularity,na.rm=T))

load('patients.Rdata', verbose=T)
```

## Segmentation Values
```{r clust, echo=T, message=T, warning=F}

for (pt in subset(sum.patient.data, Status == 'P')$Patient) {
  if (pt == 'AH0254') next
  print(pt)
  print(ncol(patient.data[[pt]]$seg.vals[,-(1:5)]))

  model = Mclust(patient.data[[pt]]$seg.vals[,-(1:5)], modelNames=c('VEE','VVE'))
  print(model$loglik)

  tryCatch({
    mod1dr = MclustDR(model, lambda = 1)
    #summary(mod1dr)
    table(mod1dr$mixcomp)

    plot(mod1dr, what = "evalues")
    title(main=pt)
    plot(mod1dr, what = "density")
  })
}

```

## Normalized Segmentation Values
Doesn't work for the normalized values.  I'm not sure that the normalization is useful though...
```{r clustN, echo=T, message=T, warning=F}

for (pt in subset(sum.patient.data, Status == 'P')$Patient) {
  if (pt == 'AH0254') next
  print(pt)
  print(paste('norm.seg.vals cols:', ncol(patient.data[[pt]]$norm.seg.vals)))

  model = Mclust(patient.data[[pt]]$norm.seg.vals, modelNames=c('VEE','VVE'))
  print(paste('loglik', round(model$loglik,2)))

  tryCatch({
    mod1dr = MclustDR(model, lambda = 1)
    #summary(mod1dr)
    table(mod1dr$mixcomp)

    if (length(table(mod1dr$mixcomp)) > 1) {
      plot(mod1dr, what = "evalues")
      title(main=pt)
      plot(mod1dr, what = "density")
    }
  })
}

```

## Perhaps AP Clustering
```{r apclust-func, echo=F, message=F, warning=F}
suppressPackageStartupMessages( library(apcluster) )

apclust.data<-function(segdata, samples) {
  x1 = segdata[,intersect(colnames(segdata), samples)]
  rownames(x1) = (segdata[,c(1:4)] %>%
    rowwise() %>%
    mutate(location=paste(paste(chrom, arm, sep=''), '.', start.pos, '-', end.pos, sep='')))$location
  return(x1)  
}

apclust<-function(segdata) {
  x1 = apclust.data(segdata)
  # q=0 minimize off-diagonal similarity
  ac = apcluster(negDistMat(r=2), x1, details=T, convits=25, q=0)
  return(list('apres'=ac, 'data'=x1))
}

net.similarity<-function(ac) {
  if (is.list(ac))
    ac = ac$apres
  ac@netsim
}

sum.similarity<-function(ac) {
  if (is.list(ac))
    ac = ac$apres
  ac@dpsim
}

clusters<-function(ac) {
  if (is.list(ac))
    ac = ac$apres
  ac@exemplars
}

plotAC<-function(aclist) {
  plot(aclist$apres, aclist$data)
}

heatmapAC<-function(aclist) {
  heatmap(aclist$apres)
}

# First vs last samples
#x1 = apclust.data(patient.data[[pt]]$seg.vals[,c(1:4, 6, ncol(patient.data[[pt]]$seg.vals))], patient.data[[pt]]$info$Samplename)
```

Affinity Propagagion clustering described by Frey & Dueck, 2007

```{r apclust, echo=T, message=T, warning=F, fig.height=5.5, fig.align='left'}
sum.patient.data = arrange(sum.patient.data, Status)

## Using the segmentation values
sim = sum.patient.data[,c(2:4)]
rownames(sim) = sum.patient.data[,1]
for (pt in sum.patient.data$Patient) {
  x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
  ac = apcluster(negDistMat(x1, r=2), details=T, q=0.1, includeSim=T)

  sim[pt, 'ns'] = net.similarity(ac)
  sim[pt, 'clusters'] = length(clusters(ac))

  #plot(ac, x1, main=pt)
  #heatmap(ac, main=pt)
  heatmap(ac, col=terrain.colors(12), Rowv=F, dendScale=0.5, margins=c(3,3,2), legend="col", sideColors=c('darkgreen','yellowgreen'), main=pt)
}

summarise(group_by(sim, Status), mean.ns=mean(ns), sd.ns=sd(ns), mean.c=mean(clusters), sd.c=sd(clusters) )

# May be a difference
t.test( subset(sim, Status == 'NP')$ns, subset(sim, Status == 'P')$ns )
t.test( subset(sim, Status == 'NP')$clusters, subset(sim, Status == 'P')$clusters )

# Should this be performed on the normalized seg values instead?
sim3 = sum.patient.data[,c(2:4)]
rownames(sim3) = sum.patient.data[,1]
for (pt in sum.patient.data$Patient) {
  x1 = patient.data[[pt]]$norm.seg.vals
  ac = apcluster(negDistMat(x1, r=2), details=T, q=0.1, includeSim=T)

  sim3[pt, 'ns'] = net.similarity(ac)
  sim3[pt, 'clusters'] = length(clusters(ac))

  #plot(ac, x1, main=paste(pt, 'norm'))
  heatmap(ac, col=terrain.colors(12), Rowv=F, dendScale=0.5, margins=c(3,3,2), legend="col", sideColors=c('darkgreen','yellowgreen'), main=paste(pt, 'norm'))
}

summarise(group_by(sim3, Status), mean.ns=mean(ns), sd.ns=sd(ns), mean.c=mean(clusters), sd.c=sd(clusters) )

# Doesn't show up here
t.test( subset(sim3, Status == 'NP')$ns, subset(sim3, Status == 'P')$ns )
t.test( subset(sim3, Status == 'NP')$clusters, subset(sim3, Status == 'P')$clusters )

```

So using the segmentation values (not the normalized ones) try some leave one out xval to see if the currently weak differences still hold.

```{r xval, echo=T, warning=F, message=T, fig.align='left', fig.height=4}

# Leave out patients
loo_xval = matrix(ncol=2, nrow=0, dimnames=list(c(), c('net.similarity', 'num.clusters')))
for (i in 1:nrow(sum.patient.data)) {
  simL = sum.patient.data[-i,c(2:4)]
  rownames(simL) = sum.patient.data[-i,1]

  for (pt in sum.patient.data$Patient[-i]) {
    x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
    ac = apcluster(negDistMat(x1, r=2), q=0.1)

    simL[pt, 'ns'] = net.similarity(ac)
    simL[pt, 'clusters'] = length(clusters(ac))
  }

  ttp = t.test( subset(simL, Status == 'NP')$ns, subset(simL, Status == 'P')$ns )$p.value
  ttc = t.test( subset(simL, Status == 'NP')$clusters, subset(simL, Status == 'P')$clusters )$p.value

  loo_xval = rbind(loo_xval, cbind(ttp,ttc))
}

ggplot(melt(data.frame(loo_xval)), aes(x=variable, y=value, fill=variable)) +  
    geom_boxplot() + geom_jitter(shape=16, position=position_jitter(0.2)) +
    stat_boxplot(geom ='errorbar') + labs(y='p-values', x='', title='Leave one (patient) out clustering')

# Leaving out these patients causes the p-value to be > 0.05
pander(sum.patient.data[ which(loo_xval >= 0.05), ], caption='In a leave-one out analysis for ap clustering removing these patients causes the p-values to be > 0.05')

# They have a higher mean than average, but may not be significant
t.test(sum.patient.data[ which(loo_xval >= 0.05), 'n.samples'], subset(sum.patient.data, Status == 'P')$'n.samples')

# Leave out samples?
looS_xval = matrix(ncol=2, nrow=0, dimnames=list(c(), c('net.similarity', 'num.clusters')))
for (i in 1:20) { # 10 fold?
  print(i)
  simX = sum.patient.data[,c(2:4)]
  rownames(simX) = sum.patient.data[,1]
  # Leave one out for every clustering pass, but not first or last?
  for (pt in sum.patient.data$Patient) {
      x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename[-sample(2:(length(patient.data[[pt]]$info$Samplename)-1), 1)])
      ac = apcluster(negDistMat(x1, r=2), q=0.1)

      simX[pt, 'ns'] = net.similarity(ac)
      simX[pt, 'clusters'] = length(clusters(ac))
    }

  ttp = t.test( subset(simX, Status == 'NP')$ns, subset(simX, Status == 'P')$ns )$p.value
  ttc = t.test( subset(simX, Status == 'NP')$clusters, subset(simX, Status == 'P')$clusters )$p.value

  looS_xval = rbind(looS_xval, cbind(ttp,ttc))
}

ggplot(melt(data.frame(looS_xval)), aes(x=variable, y=value, fill=variable)) +  
    geom_boxplot() + geom_jitter(shape=16, position=position_jitter(0.2)) +
    stat_boxplot(geom ='errorbar') + labs(y='p-values', x='', title='Leave one (sample) out clustering')

```

## Shared gains/losses

Use ap clustering to identify the shared gains/losses in a single sample

```{r gl, echo=T, message=T, warning=F, fig.align='left', fig.height=10}

pt = 'PR1_ADH_074'

head(patient.data[[pt]]$variable.regions)

x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)

# hold on...perhaps I should compute the 
ac = apcluster(negDistMat(x1, r=2), details=T, q=0.1, includeSim=T)

heatmap(ac, col=terrain.colors(12), Rowv=F, dendScale=0.5, margins=c(3,3,2), legend="col", sideColors=c('darkgreen','yellowgreen'))
#plot(ac, x1)
#plot(ac)

# Merge clusters
agex = aggExCluster(x=ac)
plot(agex)

plot(agex, x1, k=3)
heatmap(agex, col=terrain.colors(12), Rowv=F, dendScale=0.5, margins=c(3,3,2), legend="col", sideColors=c('darkgreen','yellowgreen'))

ac
```

### 2017-01-04 Patient Report.Rmd
Looking at different clustering methods

@@
 ```{r cors, echo=F, message=T, warning=F, fig.height=6, fig.width=6, fig.align='left'}
+## Cluster the samples by coverage then pull out regions
+
 ## Thresholds
@@
 gamma2 = 250
-sd.threshold = 0.08
+sd.threshold = 0.08  # cov by numbers of probes
+
 gain.threshold = 1.1

### 2017-01-17 Clustering.Rmd
AP clustering analysis, as far as I’ve gotten so far

@@
 library(GenomicRanges)
+library(gridExtra)
+library(biomaRt)
 
@@
 
-## Segmentation Values
-```{r clust, echo=T, message=T, warning=F}
+## How do the samples separate?
+
+```{r pca, echo=F, warning=F, message=F, fig.width=20, fig.height=20}
+sum.patient.data = arrange(sum.patient.data, Status, n.samples)
+
+plot.pca<-function(pt) {
+
+  info = arrange(patient.data[[pt]]$info, Endoscopy.Year)
+
+  x1 = patient.data[[pt]]$seg.vals[, patient.data[[pt]]$info$Samplename]
+  x1 = x1[,patient.data[[pt]]$info$Samplename]
+  colnames(x1) = paste(patient.data[[pt]]$info$Endoscopy.Year, '(', patient.data[[pt]]$info$Pathology, ')', sep='' )
+
+  pca = prcomp(t(x1), scale=T)
+
+  eigenvalues = pca$sdev^2
+  #screeplot(pca, main='scree')
+
+  #plot(pca$sd^2)
+
+  df = as.data.frame(pca$x)
+  df$year = as.factor(info$Endoscopy.Year)
+  df$label = paste(info$Endoscopy.Year, '(', info$Pathology, ')', sep='' )
+  df$Pathology = info$Pathology
+
+  gg = ggplot(df, aes(PC1, PC2)) + geom_line() +
+    geom_label(aes(label=year),nudge_y=0.04) + 
+    geom_point( aes(color=Pathology, size=Pathology)) +
+    labs(title=pt)
+
+  return(gg)
+}
+
+pplots = lapply(subset(sum.patient.data, Status == 'P')$Patient, plot.pca)
+grid.arrange(grobs=pplots, ncol=2, top="Progressors")
+
+npplots = lapply(subset(sum.patient.data, Status == 'NP')$Patient, plot.pca)
+grid.arrange(grobs=npplots, ncol=2, top="Non-Progressors")
+
+```
+
+```{r clust, echo=T, message=T, warning=F, eval=F}
 
@@
 
-  model = Mclust(patient.data[[pt]]$seg.vals[,-(1:5)], modelNames=c('VEE','VVE'))
+  x1 = patient.data[[pt]]$seg.vals[,-(1:5)]
+
+  model = Mclust(t(x1))
   print(model$loglik)
@@
   tryCatch({
-    mod1dr = MclustDR(model, lambda = 1)
+    mod1dr = MclustDR(model, lambda = 0)
     #summary(mod1dr)
@@
 ```
-
-## Normalized Segmentation Values
-Doesn't work for the normalized values.  I'm not sure that the normalization is useful though...
-```{r clustN, echo=T, message=T, warning=F}
-
-for (pt in subset(sum.patient.data, Status == 'P')$Patient) {
-  if (pt == 'AH0254') next
-  print(pt)
-  print(paste('norm.seg.vals cols:', ncol(patient.data[[pt]]$norm.seg.vals)))
-
-  model = Mclust(patient.data[[pt]]$norm.seg.vals, modelNames=c('VEE','VVE'))
-  print(paste('loglik', round(model$loglik,2)))
-
-  tryCatch({
-    mod1dr = MclustDR(model, lambda = 1)
-    #summary(mod1dr)
-    table(mod1dr$mixcomp)
-
-    if (length(table(mod1dr$mixcomp)) > 1) {
-      plot(mod1dr, what = "evalues")
-      title(main=pt)
-      plot(mod1dr, what = "density")
-    }
-  })
-}
-
-```
-
-## Perhaps AP Clustering
 ```{r apclust-func, echo=F, message=F, warning=F}
@@
   x1 = segdata[,intersect(colnames(segdata), samples)]
+  x1 = x1[, samples]
   rownames(x1) = (segdata[,c(1:4)] %>%
@@
 
-Affinity Propagagion clustering described by Frey & Dueck, 2007
+```{r em, echo=F, message=F, warning=F, eval=F} 
+## EM
 
-```{r apclust, echo=T, message=T, warning=F, fig.height=5.5, fig.align='left'}
+```
+
+```{r clustN, echo=T, message=T, warning=F, eval=F}
+## Normalized Segmentation Values
+#Doesn't work for the normalized values.
+
+for (pt in subset(sum.patient.data, Status == 'P')$Patient) {
+  if (pt == 'AH0254') next
+  print(pt)
+  print(paste('norm.seg.vals cols:', ncol(patient.data[[pt]]$norm.seg.vals)))
+
+  x1 = patient.data[[pt]]$norm.seg.vals
+  model = Mclust(t(x1))
+  print(paste('loglik', round(model$loglik,2)))
+
+  tryCatch({
+    mod1dr = MclustDR(model, lambda = 1)
+    #summary(mod1dr)
+    table(mod1dr$mixcomp)
+
+    if (length(table(mod1dr$mixcomp)) > 1) {
+      plot(mod1dr, what = "evalues")
+      title(main=pt)
+      plot(mod1dr, what = "density")
+    }
+  })
+}
+
+```
+
+## Perhaps AP Clustering
+
+## Affinity Propagagion clustering described by Frey & Dueck, 2007
+
+First on the raw segmentation values
+
+```{r apclust, echo=T, message=T, warning=F, fig.height=8, fig.align='left'}
+q=0.2
+
 sum.patient.data = arrange(sum.patient.data, Status)
@@
 for (pt in sum.patient.data$Patient) {
+  patient.data[[pt]]$info = arrange(patient.data[[pt]]$info, Endoscopy.Year, desc(Pathology))
+
   x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
-  ac = apcluster(negDistMat(x1, r=2), details=T, q=0.1, includeSim=T)
+  colnames(x1) = paste(patient.data[[pt]]$info$Endoscopy.Year, '(', patient.data[[pt]]$info$Pathology, ')', sep='' )
+
+  ac = apcluster(negDistMat(x1, r=2), q=q, details=T, includeSim=T)
+
+  last = ncol(x1)
+  if (patient.data[[pt]]$info$Status == 'P')
+    last = sort(grep('HGD', patient.data[[pt]]$info$Pathology), decreasing=T)[1]
+  if (ncol(x1) <= 15) {
+#    plot(ac, x1[,c(c(1:ncol(x1))[-last], last)], main=pt)
+  } else {
+#    plot(ac, x1[,c(1, sort(sample(c(1:(ncol(x1)-1))[c(-1,-last)], 8)) ,last) ], main=pt)
+  }
 
   sim[pt, 'ns'] = net.similarity(ac)
+  sim[pt, 'ss'] = sum.similarity(ac)
   sim[pt, 'clusters'] = length(clusters(ac))
+  #heatmap(ac)
+}
+sim$id = rownames(sim)
 
-  #plot(ac, x1, main=pt)
-  #heatmap(ac, main=pt)
-  heatmap(ac, col=terrain.colors(12), Rowv=F, dendScale=0.5, margins=c(3,3,2), legend="col", sideColors=c('darkgreen','yellowgreen'), main=pt)
-}
-
-summarise(group_by(sim, Status), mean.ns=mean(ns), sd.ns=sd(ns), mean.c=mean(clusters), sd.c=sd(clusters) )
+sm = as.data.frame(summarise(group_by(sim, Status), mean.ns=mean(ns), sd.ns=sd(ns), mean.ss=mean(ss), sd.ss=sd(ss), mean.c=mean(clusters), sd.c=sd(clusters) ))
 
 # May be a difference
-t.test( subset(sim, Status == 'NP')$ns, subset(sim, Status == 'P')$ns )
-t.test( subset(sim, Status == 'NP')$clusters, subset(sim, Status == 'P')$clusters )
+ttps = t.test( subset(sim, Status == 'NP')$ss, subset(sim, Status == 'P')$ss )$p.value
+ttpn = t.test( subset(sim, Status == 'NP')$ns, subset(sim, Status == 'P')$ns )$p.value
+ttpc = t.test( subset(sim, Status == 'NP')$clusters, subset(sim, Status == 'P')$clusters )$p.value
 
+#t.test( subset(sim, Status == 'NP')$clusters, subset(sim, Status == 'P')$clusters )$p.value
+
+g1 = ggplot(sim[,c('Status','ss')], aes(y=ss, x=Status, fill=Status)) + geom_boxplot() + 
+  geom_jitter(width=0.2) + stat_boxplot(geom ='errorbar') +
+  labs(title=paste("p-value", signif(ttps, 3)), y='sum similarity')
+
+g2 = ggplot(sim[,c('Status','clusters')], aes(y=clusters, x=Status, fill=Status)) + geom_boxplot() + 
+  geom_jitter(width=0.2) + stat_boxplot(geom ='errorbar') +
+  labs(title=paste("p-value", signif(ttpc, 3)), y='num. clusters')
+
+grid.arrange(g1, g2, top='Raw seg values')
+
+```
+
+So what about using the normalized segmented values?
+
+```{r apclustN, echo=T, message=T, warning=F, fig.height=8, fig.align='left'}
 # Should this be performed on the normalized seg values instead?
@@
   x1 = patient.data[[pt]]$norm.seg.vals
-  ac = apcluster(negDistMat(x1, r=2), details=T, q=0.1, includeSim=T)
+  ac = apcluster(negDistMat(x1, r=2), q=q, details=T, includeSim=T)
+  #plot(ac)
 
+  sim3[pt, 'ss'] = sum.similarity(ac)
   sim3[pt, 'ns'] = net.similarity(ac)
   sim3[pt, 'clusters'] = length(clusters(ac))
-
-  #plot(ac, x1, main=paste(pt, 'norm'))
-  heatmap(ac, col=terrain.colors(12), Rowv=F, dendScale=0.5, margins=c(3,3,2), legend="col", sideColors=c('darkgreen','yellowgreen'), main=paste(pt, 'norm'))
 }
@@
 # Doesn't show up here
-t.test( subset(sim3, Status == 'NP')$ns, subset(sim3, Status == 'P')$ns )
-t.test( subset(sim3, Status == 'NP')$clusters, subset(sim3, Status == 'P')$clusters )
+ttp = t.test( subset(sim3, Status == 'NP')$ss, subset(sim3, Status == 'P')$ss )$p.value
+#ttp = t.test( subset(sim3, Status == 'NP')$ns, subset(sim3, Status == 'P')$ns )$p.value
+ttc = t.test( subset(sim3, Status == 'NP')$clusters, subset(sim3, Status == 'P')$clusters )$p.value
 
+ggplot(sim3[,c('Status','ss')], aes(y=ss, x=Status, fill=Status)) + geom_boxplot() + geom_jitter(width=0.2) + 
+  labs(title=paste("Norm seg values p-value", signif(ttp, 3)), y='sum similarity')
 ```
 
-So using the segmentation values (not the normalized ones) try some leave one out xval to see if the currently weak differences still hold.
+So using the raw segmentation values (not the normalized ones) try some leave one out xval to see if the currently weak differences still hold.
 
-```{r xval, echo=T, warning=F, message=T, fig.align='left', fig.height=4}
+Patients first
 
+```{r xval_pt, echo=T, warning=F, message=T, fig.align='left', fig.height=4, eval=T}
 # Leave out patients
-loo_xval = matrix(ncol=2, nrow=0, dimnames=list(c(), c('net.similarity', 'num.clusters')))
+loo_xval = matrix(ncol=3, nrow=0, dimnames=list(c(), c('net.similarity', 'sum.similarity', 'clusters')))
 for (i in 1:nrow(sum.patient.data)) {
@@
     x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
-    ac = apcluster(negDistMat(x1, r=2), q=0.1)
+    ac = apcluster(negDistMat(x1, r=2), q=q)
 
     simL[pt, 'ns'] = net.similarity(ac)
+    simL[pt, 'ss'] = sum.similarity(ac)
     simL[pt, 'clusters'] = length(clusters(ac))
@@
 
-  ttp = t.test( subset(simL, Status == 'NP')$ns, subset(simL, Status == 'P')$ns )$p.value
+  ttps = t.test( subset(simL, Status == 'NP')$ss, subset(simL, Status == 'P')$ss )$p.value
+  ttpn = t.test( subset(simL, Status == 'NP')$ns, subset(simL, Status == 'P')$ns )$p.value
   ttc = t.test( subset(simL, Status == 'NP')$clusters, subset(simL, Status == 'P')$clusters )$p.value
 
-  loo_xval = rbind(loo_xval, cbind(ttp,ttc))
+  loo_xval = rbind(loo_xval, cbind(ttpn, ttps, ttc))
 }
 
-ggplot(melt(data.frame(loo_xval)), aes(x=variable, y=value, fill=variable)) +  
-    geom_boxplot() + geom_jitter(shape=16, position=position_jitter(0.2)) +
+ggplot(melt(data.frame(loo_xval[,c(1,3)])), aes(x=variable, y=value, fill=variable)) +  
+    geom_boxplot() + geom_jitter(shape=16, width=0.2) + geom_hline(yintercept=0.05, color='red', alpha=0.5) +
     stat_boxplot(geom ='errorbar') + labs(y='p-values', x='', title='Leave one (patient) out clustering')
+```
 
-# Leaving out these patients causes the p-value to be > 0.05
-pander(sum.patient.data[ which(loo_xval >= 0.05), ], caption='In a leave-one out analysis for ap clustering removing these patients causes the p-values to be > 0.05')
+Then randomly leave out samples
 
-# They have a higher mean than average, but may not be significant
-t.test(sum.patient.data[ which(loo_xval >= 0.05), 'n.samples'], subset(sum.patient.data, Status == 'P')$'n.samples')
+```{r xval_smp, echo=F, warning=F, message=T, fig.align='left', fig.height=6, eval=T}
 
 # Leave out samples?
-looS_xval = matrix(ncol=2, nrow=0, dimnames=list(c(), c('net.similarity', 'num.clusters')))
-for (i in 1:20) { # 10 fold?
-  print(i)
+looS_xval = matrix(ncol=3, nrow=0, dimnames=list(c(), c('net.similarity', 'sum.similarity', 'clusters')))
+for (i in 1:20) { 
   simX = sum.patient.data[,c(2:4)]
   rownames(simX) = sum.patient.data[,1]
-  # Leave one out for every clustering pass, but not first or last?
+
   for (pt in sum.patient.data$Patient) {
-      x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename[-sample(2:(length(patient.data[[pt]]$info$Samplename)-1), 1)])
-      ac = apcluster(negDistMat(x1, r=2), q=0.1)
+      x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename[-sample(1:(length(patient.data[[pt]]$info$Samplename)), 1)])
+      ac = apcluster(negDistMat(x1, r=2), q=q)
 
       simX[pt, 'ns'] = net.similarity(ac)
+      simX[pt, 'ss'] = sum.similarity(ac)
       simX[pt, 'clusters'] = length(clusters(ac))
@@
 
-  ttp = t.test( subset(simX, Status == 'NP')$ns, subset(simX, Status == 'P')$ns )$p.value
+  ttpn = t.test( subset(simX, Status == 'NP')$ns, subset(simX, Status == 'P')$ns )$p.value
+  ttps = t.test( subset(simX, Status == 'NP')$ss, subset(simX, Status == 'P')$ss )$p.value
   ttc = t.test( subset(simX, Status == 'NP')$clusters, subset(simX, Status == 'P')$clusters )$p.value
 
-  looS_xval = rbind(looS_xval, cbind(ttp,ttc))
+  looS_xval = rbind(looS_xval, cbind(ttpn,ttps, ttc))
 }
 
-ggplot(melt(data.frame(looS_xval)), aes(x=variable, y=value, fill=variable)) +  
+ggplot(melt(data.frame(looS_xval[,c(2,3)])), aes(x=variable, y=value, fill=variable)) +  
     geom_boxplot() + geom_jitter(shape=16, position=position_jitter(0.2)) +
+  geom_hline(yintercept=0.05, color='red', alpha=0.5) +
     stat_boxplot(geom ='errorbar') + labs(y='p-values', x='', title='Leave one (sample) out clustering')
+```
+
+## (xval) Re run clusters on the Progressors without the final sample
+There is still a difference between the progressors and non, though the p-value does increase. Some of the progressors do start looking more like NPs in their similarity measure but not all of them.
+
+```{r nofinal, echo=F, message=F, warning=F, fig.height=6, fig.width=10}
+simX = sum.patient.data[,c(2:4)]
+rownames(simX) = sum.patient.data[,1]
+# Leave out final sample on the progressors
+
+for (pt in sum.patient.data$Patient) {
+  x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
+  if (subset(sum.patient.data, Patient == pt)$Status == 'P' & length(which(grepl('HGD', patient.data[[pt]]$info$Pathology))) > 0) {
+    final = grep('HGD', patient.data[[pt]]$info$Pathology)
+    x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename[-final])
+  }
+  ac = apcluster(negDistMat(x1, r=2), q=q)
+
+  simX[pt, 'ns'] = net.similarity(ac)
+  simX[pt, 'ss'] = sum.similarity(ac)
+  simX[pt, 'clusters'] = length(clusters(ac))
+}
+
+ttps = t.test( subset(simX, Status == 'NP')$ss, subset(simX, Status == 'P')$ss )$p.value
+gg1 = ggplot(melt(data.frame(simX[c('Status','ss')])), aes(x=Status, y=value, fill=Status)) +  
+    geom_boxplot() + geom_jitter(shape=16, position=position_jitter(0.2)) +
+    stat_boxplot(geom ='errorbar') + labs(y='sim values', x='', title=paste('Leave out HGD sample(s) p-value', signif(ttps, 3)))
+
+ttpc = t.test( subset(simX, Status == 'NP')$clusters, subset(simX, Status == 'P')$clusters )$p.value
+gg2 = ggplot(melt(data.frame(simX[c('Status','clusters')])), aes(x=Status, y=value, fill=Status)) +  
+    geom_boxplot() + geom_jitter(shape=16, position=position_jitter(0.2)) +
+    stat_boxplot(geom ='errorbar') + labs(y='num. clusters', x='', title=paste('Leave out HGD sample(s) p-value', signif(ttpc, 3)))
+
+grid.arrange(gg1, gg2, ncol=2)
+
+#pander(subset(simX, Status == 'P' & ss <= mean(ss)), caption="These patients are still strongly in the Progressor group")
+
+pander(subset(simX, Status == 'P' & ss >= mean(ss)), caption="These patients are more similar to Non-Progressors without their HGD sample(s)")
+```
+
+## Can we identify regions in the progressors?
+
+Not really just from the SVDs.  This isn't any different at the moment from simply calling the variable regions directly.
+```{r loc_pca, echo=F, warning=F, message=T, fig.height=12, fig.width=20}
+count.occurrences<-function(grs) {
+  ov = findOverlaps(grs, type="within", ignoreSelf=TRUE,select="arbitrary")
+  counts = countOverlaps(grs, type='within', maxgap=0)
+
+  newgrs = grs[is.na(ov)]
+
+  newgrs$overlaps = 0
+  newgrs$patients = ''
+
+  patients = c()  
+  index = 1
+  occurence = vector(mode='numeric')
+  for (x in which(is.na(ov))) {
+    occurence = c(occurence, sum(count[index:x]))
+    patients = c(patients, paste(grs[index:x]$patient, collapse=','))
+    index = x+1
+  }
+  newgrs$overlaps = occurence
+  newgrs$patients = patients
+
+  newgrs$patient.count = unlist(lapply(strsplit(patients, ','), length))
+
+  newgrs$patient = NULL  
+  return(newgrs)
+}
+annotate.cosmic<-function(grs) {
+  ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
+  ensembl = biomaRt::useMart("ENSEMBL_MART_ENSEMBL", host="grch37.ensembl.org")
+  mart = useDataset("hsapiens_gene_ensembl", ensembl)
+
+  grs$Genes = 0
+  grs$Cosmic.count = 0
+  grs$Cosmic.genes = ''
+
+  plocs = as.data.frame(grs)  
+  for (i in 1:nrow(plocs)) {
+    bm = getBM(mart = mart,
+        attributes=c('hgnc_symbol', 'entrezgene' ), filters = c('chromosome_name', 'start', 'end' ),
+        values = with(plocs[i,], list(gsub("chr", "", as.character(seqnames)), start, end)))
+
+    grs[i]$Genes = length(unique(bm$hgnc_symbol))
+    grs[i]$Cosmic.count = length(intersect(bm$entrezgene, ccgenes$Entrez.GeneId))
+    grs[i]$Cosmic.genes = paste(intersect(bm$hgnc_symbol, ccgenes$Gene.Symbol), collapse=',')
+  }
+
+  return(grs)
+}
+
+#pt = 'PR1_HIN_042'
+gg.svd<-function(mat, names, sd=0, n=3) {
+  df = as.data.frame(mat[,(1:n)])
+
+  for (i in 1:ncol(df)) {
+    high = mat[,i] > (mean(mat[,i]) + sd * sd(mat[,i]))
+    low = mat[,i] < (mean(mat[,i]) - sd * sd(mat[,i]))
+
+    label = high | low
+    color=label
+
+    for (j in 1:length(label))
+      label[j] = ifelse(label[j], names[j], '')
+
+    df = cbind(df, label, color)
+    colnames(df)[grep('^label$', colnames(df))] = paste('SV',i,'.label',sep='')
+    colnames(df)[grep('^color$', colnames(df))] = paste('SV',i,'.color',sep='')
+  }
+  df$x = 1:nrow(df)
+  return(df)
+  }
+
+pt.regions = list()
+for (pt in sum.patient.data$Patient) {
+  print(pt)
+  patient.data[[pt]]$info = arrange(patient.data[[pt]]$info, Endoscopy.Year, Pathology)
+
+  x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
+  colnames(x1) = paste(patient.data[[pt]]$info$Endoscopy.Year, '(', patient.data[[pt]]$info$Pathology, ')', sep='' )
+
+  #plot.pca(pt)
+  #pca = prcomp(t(x1), center=T)
+
+  #plot(pca$sdev^2, type='o')
+
+  svd = svd(t(x1))
+  colnames(svd$u) = paste('SV',1:ncol(svd$u), sep='') 
+  colnames(svd$v) = paste('SV',1:ncol(svd$v), sep='')
+
+  #plot(svd$d, type='o', main='Scree')
+  # u matrix
+  samples = gg.svd(svd$u, colnames(x1), 0, 3)
+  grid.arrange(
+    ggplot(samples, aes(y=SV1, x=x, col=SV1.color)) + geom_point() +
+      geom_text(aes(label=SV1.label), vjust=0) + guides(color='none'),
+    ggplot(samples, aes(y=SV2, x=x, col=SV2.color)) + geom_point() +
+      geom_text(aes(label=SV2.label), vjust=0) + guides(color='none'),
+    ggplot(samples, aes(y=SV3, x=x, col=SV3.color)) + geom_point() +
+      geom_text(aes(label=SV3.label), vjust=0) + guides(color='none'),
+    top='SVD Samples')
+
+  # v matrix
+  regions = gg.svd(svd$v, rownames(x1), 2, 3)
+  grid.arrange(
+    ggplot(regions, aes(y=SV1, x=x, col=SV1.color)) + geom_point() +
+      geom_text(aes(label=SV1.label), vjust=0) + guides(color='none'),
+    ggplot(regions, aes(y=SV2, x=x, col=SV2.color)) + geom_point() +
+      geom_text(aes(label=SV2.label), vjust=0) + guides(color='none'),
+    ggplot(regions, aes(y=SV3, x=x, col=SV3.color)) + geom_point() +
+      geom_text(aes(label=SV3.label), vjust=0) + guides(color='none'), 
+    top='SVD Locations')
+
+  locs = as.character(unlist(c(regions[grep('SV1.label',colnames(regions))])))
+  locs = unique(locs[grep("\\w+", locs)])
+
+  high = names(which(apply(x1[locs,], 1, function(x) length(which(x >= 1.1))) > 0))
+  low = names(which(apply(x1[locs,], 1, function(x) length(which(x <= 0.9))) > 0))
+
+  hls = as.data.frame(do.call(rbind, strsplit(c(high, low), '[p|q]\\.|-')))
+  colnames(hls) = c('chr','start','stop')
+  hls[c('start','stop')] = lapply(hls[c('start','stop')], function(x) as.numeric(as.character(x)))
+
+  #ac = apcluster(negDistMat(x1, r=2), q=0.2)
+
+  #lapply(ac@clusters, function(x) names(x)[which(names(x) %in% high)])
+  #lapply(ac@clusters, function(x) names(x)[which(names(x) %in% low)])
+
+  pt.regions[[pt]] = data.frame(hls, cn=c(rep('gain', length(high)), rep('loss',length(low))))
+}
+pt.regions = do.call(rbind, pt.regions)
+
+pt.regions$patient = sub('\\..*', '', rownames(pt.regions) )
+rownames(pt.regions) = 1:nrow(pt.regions)
+
+prog.grs = sort(makeGRangesFromDataFrame(subset(pt.regions, patient %in% subset(sum.patient.data, Status == 'P')$Patient), keep.extra.columns = T, seqnames.field = 'chr', seqinfo=c(1:22,'X','Y')))
+
+np.grs = sort(makeGRangesFromDataFrame(subset(pt.regions, patient %in% subset(sum.patient.data, Status == 'NP')$Patient), keep.extra.columns = T, seqnames.field = 'chr', seqinfo=c(1:22,'X','Y')))
+
+prog.loss = count.occurrences(subset(prog.grs, grepl('loss',cn)))
+prog.gain = count.occurrences(subset(prog.grs, grepl('gain',cn)))
+
+np.loss = count.occurrences(subset(np.grs, grepl('loss',cn)))
+np.gain = count.occurrences(subset(np.grs, grepl('gain',cn)))
+
+pg.c = annotate.cosmic(prog.gain)
+pl.c = annotate.cosmic(prog.loss)
+
+npg.c = annotate.cosmic(np.gain)
+npl.c = annotate.cosmic(np.loss)
+
+with(pg.c,plot(overlaps, Cosmic.count, col='blue', pch=19))
+with(pl.c, points(overlaps, Cosmic.count, col='red', pch=19))
+
+with(npg.c, points(overlaps, Cosmic.count, col='green', pch=18))
+with(npl.c, points(overlaps, Cosmic.count, col='purple', pch=17))
+
+pander(data.frame('total.p' = sum(length(pg.c), length(pl.c)), 'total.np' = sum(length(npg.c), length(npl.c))), caption="Total regions called as gains/losses for progressors vs non")
+
+t.test( width(pg.c), width(npg.c) ) # Yes for the gains
+t.test( width(pl.c), width(npl.c) ) # Not for the losses
 
 ```
-
-## Shared gains/losses
-
-Use ap clustering to identify the shared gains/losses in a single sample
-
-```{r gl, echo=T, message=T, warning=F, fig.align='left', fig.height=10}
-
-pt = 'PR1_ADH_074'
-
-head(patient.data[[pt]]$variable.regions)
-
-x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
-
-# hold on...perhaps I should compute the 
-ac = apcluster(negDistMat(x1, r=2), details=T, q=0.1, includeSim=T)
-
-heatmap(ac, col=terrain.colors(12), Rowv=F, dendScale=0.5, margins=c(3,3,2), legend="col", sideColors=c('darkgreen','yellowgreen'))
-#plot(ac, x1)
-#plot(ac)
-
-# Merge clusters
-agex = aggExCluster(x=ac)
-plot(agex)
-
-plot(agex, x1, k=3)
-heatmap(agex, col=terrain.colors(12), Rowv=F, dendScale=0.5, margins=c(3,3,2), legend="col", sideColors=c('darkgreen','yellowgreen'))
-
-ac
-```

### 2017-01-17 Patient Report.Rmd
AP clustering analysis, as far as I’ve gotten so far

@@
   cur.pt = subset(patient.info, Patient == patient.name)
-  segvals[, c(colnames(segvals[1:5]), cur.pt$Samplename)]
+  segvals = segvals[, c(colnames(segvals[1:5]), cur.pt$Samplename)]
 

### 2017-02-02 Clustering.Rmd
Predict NP vs P using number of clusters

@@
 
+getPrecision<-function(actual,predicted) {
+  retrieved <- sum(predicted==1)
+  precision <- sum(predicted==1 & actual==1) / retrieved
+  precision
+}
+getRecall<-function(actual,predicted) {
+  retrieved <- sum(predicted==1)
+  recall <- sum(predicted==1 & actual==1) / sum(actual==1)
+  recall
+}
+
 data = '~/Data/Ellie'
@@
 patient.info$Patient = gsub("/", "_", patient.info$Patient)
-head(patient.info)
+#head(patient.info)
 
@@
 
-## How do the samples separate?
+## How do the sample timepoints separate?
+
+It appears that in the progressors that generally the HGD or IMC pathology may suddenly appear and be very different (PR1-HIN-044) from other samples, or those where it may have been a progressive sampling that can be fairly related to previous samples (AHM0952).
+In most cases the HGD/IMC samples drive the variance, but it is often there in the LGD samples as well.
+In contrast the non-progressor patients, even where they have LGD, don't generally show a sudden jump in the sample (AD0591).
+
+This would suggest we might have a more difficult time separating out the progressors where the early samples were all highly related and a sudden shift occurred to generate the HGD sample.
+This could be either sampling bias or a sudden model of progression, but I can't think of a method right now that might separate the two.
 
@@
 
-```{r clust, echo=T, message=T, warning=F, eval=F}
+## Mixture Models
 
-for (pt in subset(sum.patient.data, Status == 'P')$Patient) {
-  if (pt == 'AH0254') next
+One of the questions we'd like to answer is how to separate the progressors from the non-progressors before we see the pathological progression (HGD or even LGD).
+If there is increasing instability in the genome how do we characterize it?
+
+I tried were several mixture models (e.g. EM) but wasn't able to find a pattern that could reliably separate the few mixtures I did see
+
+```{r clust, echo=T, message=T, warning=F, eval=T, fig.height=6, fig.width=6}
+
+llk = vector(mode='numeric', length=nrow(sum.patient.data))
+names(llk) = sum.patient.data$Patient
+for (pt in sum.patient.data$Patient) {
   print(pt)
@@
 
-  x1 = patient.data[[pt]]$seg.vals[,-(1:5)]
+  x1 = as.matrix(patient.data[[pt]]$seg.vals[,-(1:5)])
+  x1 = x1[,patient.data[[pt]]$info$Samplename]
 
-  model = Mclust(t(x1))
-  print(model$loglik)
+  fit <- Mclust(x1, G=3)
+  plot(fit, what='classification')
+  title(main=paste(pt, ' (',patient.data[[pt]]$info$Status[1], ')', sep=''))
 
-  tryCatch({
-    mod1dr = MclustDR(model, lambda = 0)
-    #summary(mod1dr)
-    table(mod1dr$mixcomp)
+  llk[pt] = fit$loglik
 
-    plot(mod1dr, what = "evalues")
-    title(main=pt)
-    plot(mod1dr, what = "density")
-  })
+#   clusterPlots <- data.frame(x1 = clusterData$x1, x2 = clusterData$x2, cluster = factor(model$classification, 
+#     levels = c(1, 2), labels = c("z = 0", "z = 1")), uncertainity = model$uncertainty, 
+#     logUncertainity = log(model$uncertainty))
+# clusterPlots.gg <- ggplot(clusterPlots)
+# clusterPlots.gg + geom_point(aes(x = x1, y = x2, color = cluster))
+
+  #tryCatch({
+  # model = densityMclust(x1, G=3)
+  # plotDensityMclustd(model, data=x1,  col='blue', lwd=2, hist.col = "lightblue",  breaks=100)
+  #   mod1dr = MclustDR(model, lambda = 0.5)
+  #   #summary(mod1dr)
+  #   table(mod1dr$mixcomp)
+  # 
+  #   #plot(mod1dr, what = "evalues")
+  #   #title(main=pt)
+  #   plot(mod1dr, what = "density")
+  #   title(main=paste(pt, ' (',patient.data[[pt]]$info$Status[1], ')', sep=''))
+  # })
 }
 
+# Log liklihood doesn't separate them
+t.test(llk[subset(sum.patient.data, Status == 'NP')$Patient], llk[subset(sum.patient.data, Status == 'P')$Patient])
+
 ```
+
+```{r means, echo=F, message=F, warning=F, eval=F} 
+## Kmeans
+#   x1 = as.matrix(patient.data[[pt]]$seg.vals[,-(1:5)])
+# 
+   fit = kmeans(x1, 3)
+#   plot(x1[,c(1,10)], col=km$cluster)
+# title(main=paste(pt, ' (',patient.data[[pt]]$info$Status[1], ')', sep=''))
+
+library(cluster) 
+clusplot(x1, fit$cluster, color=TRUE, shade=TRUE, labels=2, lines=0, main=pt)
+
+```
+
 ```{r apclust-func, echo=F, message=F, warning=F}
@@
 
-```{r em, echo=F, message=F, warning=F, eval=F} 
-## EM
-
-```
-
 ```{r clustN, echo=T, message=T, warning=F, eval=F}
@@
 
-## Perhaps AP Clustering
+## AP Clustering
 
-## Affinity Propagagion clustering described by Frey & Dueck, 2007
+So tried AP as that doesn't require the pre-definition for number of clusters.  It appears to give good separation, though there's a suggestion of two different types of progressors.
+
+### Affinity Propagagion clustering described by Frey & Dueck, 2007
 
@@
 
-```{r apclust, echo=T, message=T, warning=F, fig.height=8, fig.align='left'}
+```{r apclust, echo=T, message=T, warning=F, fig.height=10, fig.width=10, fig.align='left'}
 q=0.2
@@
   last = ncol(x1)
-  if (patient.data[[pt]]$info$Status == 'P')
-    last = sort(grep('HGD', patient.data[[pt]]$info$Pathology), decreasing=T)[1]
+  if (patient.data[[pt]]$info$Status[1] == 'P' & pt != 'AD0252')
+    last = sort(grep('HGD|IMC', patient.data[[pt]]$info$Pathology), decreasing=T)[1]
   if (ncol(x1) <= 15) {
-#    plot(ac, x1[,c(c(1:ncol(x1))[-last], last)], main=pt)
+    plot(ac, x1[,c(c(1:ncol(x1))[-last], last)], main=pt)
   } else {
-#    plot(ac, x1[,c(1, sort(sample(c(1:(ncol(x1)-1))[c(-1,-last)], 8)) ,last) ], main=pt)
+    plot(ac, x1[,c(1, sort(sample(c(1:(ncol(x1)-1))[c(-1,-last)], 8)) ,last) ], main=pt)
   }
@@
 grid.arrange(g1, g2, top='Raw seg values')
+
+fit = glm(Status~clusters, sim, family=binomial(link='logit'))
+pander(fit, justify='left')
+
+sim = sim[, c('Status', 'clusters', 'ns','ss')]
+actualMatrix = matrix(0,1000,2,dimnames=list(c(), c('prec','recall')))
+for (i in 1:nrow(actualMatrix)) {
+  trainingRows = sample(1:nrow(sim), nrow(sim)-5)
+  train = sim[trainingRows,]
+  test = sim[-trainingRows,]
+
+  fit = glm( (as.integer(Status)-1)~clusters, train, family=binomial(link='logit'))
+  pred = predict(fit, test, se.fit=F) # probabilities -- of which outcome though?
+
+  actual = test$Status == 'P'
+  predicted = pred>0
+
+  actualMatrix[i,'prec'] = getPrecision( actual, predicted )
+  actualMatrix[i, 'recall'] = getRecall( actual, predicted )
+}
+
+ggplot(as.data.frame(actualMatrix), aes(x=recall, y=prec)) + geom_point() + 
+  geom_hline(yintercept = 0.5, col='red', alpha=0.5) +
+  labs(title='Precision/Recall for Progressor/NP prediction')
+
+```
+
+### Are the clusters random?
+
+```{r example_pt}
+pt = 'PR1_HIN_042'
+patient.data[[pt]]$info = arrange(patient.data[[pt]]$info, Endoscopy.Year, desc(Pathology))
+
+x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
+colnames(x1) = paste(patient.data[[pt]]$info$Endoscopy.Year, '(', patient.data[[pt]]$info$Pathology, ')', sep='' )
+
+ac = apcluster(negDistMat(x1, r=2), q=q, details=T, includeSim=T)
+
+cols = c('chr','cluster','P.x', 'x')
+hgmx = matrix(nrow=0,ncol=length(cols),dimnames=list(c(), cols))
+for (chr in c(1:22)) {
+  for (i in 1:length(ac@clusters)) {
+    cl = x1[ac@clusters[[i]],]
+
+    x = length(grep(paste('^',chr,'(p|q)',sep=''),  rownames(cl), value=T)) # n in cluster
+    m = length(grep(paste('^',chr,'(p|q)',sep=''),  rownames(x1), value=T)) # total n
+    n = nrow(x1)-m
+    k = length(ac@clusters[[i]])
+
+    hgmx = rbind(hgmx, cbind(chr, i, signif(dhyper(x,m,n,k),3), x))
+  }
+}
+
+hgmx = as.data.frame(hgmx)
+
+ggplot(subset(hgmx, x > 0), aes(x=chr,y=P.x,color=factor(cluster))) + ylim(0,1)+ 
+  geom_point() + geom_line(alpha=0.1)
+
+ggplot(subset(hgmx, x > 0 & cluster == 9), aes(x=factor(chr),y=P.x,fill=factor(cluster))) + ylim(0,0.5 )+ 
+  geom_bar(colour='grey', stat='identity', position=position_dodge())
 
@@
 
-## (xval) Re run clusters on the Progressors without the final sample
+## (xval) Re run clusters on the Progressors without the HGD sample
 There is still a difference between the progressors and non, though the p-value does increase. Some of the progressors do start looking more like NPs in their similarity measure but not all of them.
@@
   x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
-  if (subset(sum.patient.data, Patient == pt)$Status == 'P' & length(which(grepl('HGD', patient.data[[pt]]$info$Pathology))) > 0) {
-    final = grep('HGD', patient.data[[pt]]$info$Pathology)
+  if (subset(sum.patient.data, Patient == pt)$Status == 'P' & length(which(grepl('HGD|IMC', patient.data[[pt]]$info$Pathology))) > 0) {
+    final = grep('HGD|IMC', patient.data[[pt]]$info$Pathology)
     x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename[-final])
@@
 pander(subset(simX, Status == 'P' & ss >= mean(ss)), caption="These patients are more similar to Non-Progressors without their HGD sample(s)")
+
+simX = simX[, c('Status', 'clusters', 'ns','ss')]
+noHGDPR = as.data.frame(matrix(0,1000,2,dimnames=list(c(), c('prec','recall'))))
+for (i in 1:nrow(actualMatrix)) {
+  trainingRows = sample(1:nrow(simX), nrow(simX)-5)
+  train = simX[trainingRows,]
+  test = simX[-trainingRows,]
+
+  fit = glm( Status~clusters, train, family=binomial(link='logit'))
+  pred = predict(fit, test, se.fit=F) # probabilities -- of which outcome though?
+
+  actual = test$Status == 'P'
+  predicted = pred>0
+
+  noHGDPR[i,'prec'] = getPrecision( actual, predicted )
+  noHGDPR[i, 'recall'] = getRecall( actual, predicted )
+}
+
+ggplot(as.data.frame(noHGDPR), aes(x=recall, y=prec)) + geom_point() + 
+  geom_hline(yintercept = 0.5, col='red', alpha=0.5) +
+  labs(title='Precision/Recall for Progressor/NP prediction NO HGD')
+
 ```
 
+Currently, without the HGD/IMC samples we can predict progressors `r nrow(subset(noHGDPR, prec > 0.5))/nrow(noHGDPR)*100`% of the time.
+In comparison, including HGD/IMC samples our predictions improve to `r length(which(actualMatrix[,'prec'] > 0.5))/nrow(actualMatrix)*100`%.
+
+## And why q=0.2
+```{r whyq, echo=F, message=F, warning=F}
+sum.patient.data = arrange(sum.patient.data, Status)
+
+pvals = as.data.frame(matrix(ncol=2,nrow=0,dimnames=list(c(), c('clusters','ss'))))
+for (qv in seq(0.1,0.9,0.1)) {
+  print(qv)
+
+  ## Using the segmentation values
+  simPC = sum.patient.data[,c(2:4)]
+  rownames(simPC) = sum.patient.data[,1]
+  for (pt in sum.patient.data$Patient) {
+    patient.data[[pt]]$info = arrange(patient.data[[pt]]$info, Endoscopy.Year, desc(Pathology))
+
+    x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
+    colnames(x1) = paste(patient.data[[pt]]$info$Endoscopy.Year, '(', patient.data[[pt]]$info$Pathology, ')', sep='' )
+
+    ac = apcluster(negDistMat(x1, r=2), q=qv, convits=75)
+
+    simPC[pt, 'ns'] = net.similarity(ac)
+    simPC[pt, 'ss'] = sum.similarity(ac)
+    simPC[pt, 'clusters'] = length(clusters(ac))
+  }
+
+  # May be a difference
+  ttps = t.test( subset(simPC, Status == 'NP')$ss, subset(simPC, Status == 'P')$ss )$p.value
+  ttpn = t.test( subset(simPC, Status == 'NP')$ns, subset(simPC, Status == 'P')$ns )$p.value
+  ttpc = t.test( subset(simPC, Status == 'NP')$clusters, subset(simPC, Status == 'P')$clusters )$p.value
+
+  pvals = rbind(pvals,cbind(ttpc,ttps))
+}
+colnames(pvals) = c('clusters','sum.similarity')
+pvals$qv=seq(0.1,0.9,0.1)
+
+ggplot(melt(pvals, id.vars='qv'),aes(x=qv,y=value, color=variable)) + geom_point() +
+  labs(main="qvalus for AP", y='pvalue NP vs P', x='q param')
+
+```
+
+```{r loc_pca, echo=F, warning=F, message=T, fig.height=12, fig.width=20, eval=F}
 ## Can we identify regions in the progressors?
 
-Not really just from the SVDs.  This isn't any different at the moment from simply calling the variable regions directly.
-```{r loc_pca, echo=F, warning=F, message=T, fig.height=12, fig.width=20}
+#Not really just from the SVDs.  This isn't any different at the moment from simply calling the variable regions directly.
+
 count.occurrences<-function(grs) {
@@
 ```
+
+```{r precrec, echo=F, warning=F, message=F, fig.height=12, fig.width=20}
+
+```

### 2017-02-15 Clustering.Rmd
Updated for new patient data
Updating the segmentation script

@@
 
-sum.patient.data = ddply(patient.info, .(Patient, Status), summarise, years=paste(range(Endoscopy.Year), collapse='-'), 
-      n.samples=length(Samplename), med.cellularity=median(Barretts.Cellularity,na.rm=T))
+load('patients.Rdata', verbose=T)
 
-load('patients.Rdata', verbose=T)
+sum.patient.data = as.data.frame(patient.info %>%
+  group_by(Patient, Status) %>%
+  summarise(
+    years=paste(range(Endoscopy.Year), collapse='-'),
+    n.samples=length(Samplename), med.cellularity=median(Barretts.Cellularity,na.rm=T),
+    mean.reads=mean(Total.Reads),
+    sd.reads=sd(Total.Reads)
+  ))
+
+pander(sum.patient.data)
 ```
+
+# Global View and Questions
+
+```{r global_pt, echo=F, message=F, warning=F}
+be = subset(patient.info, Pathology == 'BE')
+lgd = subset(patient.info, Pathology <= 'LGD')
+
+p53Status = data.frame(rbind('Non-Progressors'=table(subset(patient.info, Status == 'NP')$p53.Status), 
+           'Progressors'=table(subset(patient.info, Status == 'P')$p53.Status)))
+colnames(p53Status) = c('None', 'Positive')
+
+ttr = t.test(subset(patient.info, Status == 'NP')$Total.Reads, subset(patient.info, Status == 'P')$Total.Reads)
+ttrbe = t.test(subset(be, Status == 'NP')$Total.Reads, subset(be, Status == 'P')$Total.Reads)
+ttrlgd = t.test(subset(lgd, Status == 'NP')$Total.Reads, subset(lgd, Status == 'P')$Total.Reads)
+
+tbclgd = t.test(subset(lgd, Status == 'NP')$Barretts.Cellularity, subset(lgd, Status == 'P')$Barretts.Cellularity)
+tbcbe = t.test(subset(be, Status == 'NP')$Barretts.Cellularity, subset(be, Status == 'P')$Barretts.Cellularity)
+
+t.test(subset(patient.info, Pathology == 'BE')$Barretts.Cellularity, 
+               subset(patient.info, Pathology == 'LGD')$Barretts.Cellularity)
+
+t.test(subset(patient.info, Pathology == 'LGD')$Barretts.Cellularity, 
+               subset(patient.info, Pathology == 'HGD')$Barretts.Cellularity)
+
+## No differences between the number of reads by pathology
+t.test(subset(patient.info, Pathology == 'BE')$Total.Reads, subset(patient.info, Pathology == 'LGD')$Total.Reads)
+t.test(subset(patient.info, Pathology == 'BE')$Total.Reads, subset(patient.info, Pathology == 'HGD')$Total.Reads)
+t.test(subset(patient.info, Pathology == 'LGD')$Total.Reads, subset(patient.info, Pathology == 'HGD')$Total.Reads)
+
+t.test(subset(patient.info, Pathology == 'BE')$Barretts.Cellularity, subset(patient.info, Pathology == 'LGD')$Barretts.Cellularity)
+t.test(subset(patient.info, Pathology == 'BE')$Barretts.Cellularity, subset(patient.info, Pathology == 'HGD')$Barretts.Cellularity)
+t.test(subset(patient.info, Pathology == 'LGD')$Barretts.Cellularity, subset(patient.info, Pathology == 'HGD')$Barretts.Cellularity)
+
+```
+
+There are a few things we need to be careful of and a couple of questions to check here:
+
+1. The total number of reads per sample between NP and P might show a difference, but currently with multi-test correction it's not significant.
+`r  signif(ttr$p.value,2)*11`
+So are those samples with lower numbers of reads showing less complexity overall?
+When sequencing NP are we likely to get fewer reads than with P?
+If we look only at the samples with BE (no LGD or HGD samples) the difference disappears `r signif(ttrbe$p.value, 2)*11`.
+If we include LGD with the BE samples we again see the difference.
+`r signif(ttrlgd$p.value, 2)*11`.
+This is suggestive that changes in copy number result in more reads being output.
+
+2.
+There is a significant difference in the cellularity of the Barretts sample, with progressors having a higher mean cellularity, even when only comparing samples that were either LGD or BE.
+`r signif(tbclgd$p.value, 2)` However, even the BE only samples show a higher cellularity overall `r signif(tbcbe$p.value, 2)*11`  That said, currently we have a high rate of missing information, with about `r length(which(is.na(subset(patient.info, Status == 'P')$Barretts.Cellularity)))/length(which(patient.info$Status == 'P'))*100`% of the cellualarities missing for progressors and `r length(which(is.na(subset(patient.info, Status == 'NP')$Barretts.Cellularity)))/length(which(patient.info$Status == 'NP'))*100`% of the NP cellularities missing.
+
+3. With regards to p53 staining, non-progressors have no p53 positive tests, while progressors had `r p53Status[2,2]` overall.
+`r pander(p53Status, justify='left', caption="P53 staining status for progressors vs non")`
+Of these `r table(be$p53.Status)[['1']]` were positive in a BE sample, and `r table(lgd$p53.Status)[['1']]` were positive when LGD samples were included.
 
@@
 
-  x1 = patient.data[[pt]]$seg.vals[, patient.data[[pt]]$info$Samplename]
+  plots = list()
+  for (sv in c('seg.vals','norm.seg.vals')) {
+
+    x1 = patient.data[[pt]][[sv]][, patient.data[[pt]]$info$Samplename]
+    x1 = x1[,patient.data[[pt]]$info$Samplename]
+    colnames(x1) = paste(patient.data[[pt]]$info$Endoscopy.Year, '(', patient.data[[pt]]$info$Pathology, ')', sep='' )
+
+    pca = prcomp(t(x1), scale=T)
+
+    eigenvalues = pca$sdev^2
+    screeplot(pca, main=paste(pt, df, 'scree'))
+
+    #plot(pca$sd^2)
+
+    df = as.data.frame(pca$x)
+    df$year = as.factor(info$Endoscopy.Year)
+    df$label = paste(info$Endoscopy.Year, '(', info$Pathology, ')', sep='' )
+    df$Pathology = info$Pathology
+
+    gg = ggplot(df, aes(PC1, PC2)) + geom_line() +
+        geom_label(aes(label=year),nudge_y=0.04) + 
+        geom_point( aes(color=Pathology, size=Pathology)) +
+        labs(title=paste(pt, sv))
+  plots[[sv]] = gg
+  }  
+  do.call(grid.arrange, c(plots, list(ncol=2))
+  return(gg)
+}
+
+plot.pca.norm<-function(pt) {
+
+  info = arrange(patient.data[[pt]]$info, Endoscopy.Year)
+
+  x1 = patient.data[[pt]]$norm.seg.vals[, patient.data[[pt]]$info$Samplename]
   x1 = x1[,patient.data[[pt]]$info$Samplename]
@@
       title(main=pt)
-      plot(mod1dr, what = "density")
+      #plot(mod1dr, what = "density")
     }
@@
 ## Using the segmentation values
-sim = sum.patient.data[,c(2:4)]
-rownames(sim) = sum.patient.data[,1]
+sim = sum.patient.data[,c('Status','years')]
+rownames(sim) = sum.patient.data[,'Patient']
 for (pt in sum.patient.data$Patient) {
@@
     last = sort(grep('HGD|IMC', patient.data[[pt]]$info$Pathology), decreasing=T)[1]
-  if (ncol(x1) <= 15) {
-    plot(ac, x1[,c(c(1:ncol(x1))[-last], last)], main=pt)
-  } else {
-    plot(ac, x1[,c(1, sort(sample(c(1:(ncol(x1)-1))[c(-1,-last)], 8)) ,last) ], main=pt)
-  }
+  # if (ncol(x1) <= 15) {
+  #   plot(ac, x1[,c(c(1:ncol(x1))[-last], last)], main=pt)
+  # } else {
+  #   plot(ac, x1[,c(1, sort(sample(c(1:(ncol(x1)-1))[c(-1,-last)], 8)) ,last) ], main=pt)
+  # }
 
@@
   sim[pt, 'clusters'] = length(clusters(ac))
-  #heatmap(ac)
+
+  tmp = patient.data[[pt]]$variable.regions[,-(1:5)]
+  sim[pt, 'n.var'] = nrow(patient.data[[pt]]$variable.regions[,-(1:5)])
+
+  if (nrow(tmp) <= 0) {
+    sim[pt,c('gains', 'loss')] = c(0,0)
+  } else {
+    sim[pt,'gains'] = length(tmp[tmp>=1.1])
+    sim[pt,'loss'] =  length(tmp[tmp<=0.9])
+  }
 }
@@
 
-#t.test( subset(sim, Status == 'NP')$clusters, subset(sim, Status == 'P')$clusters )$p.value
+## Number of variable regions isn't so good, but maybe...
+sim %>%
+  group_by(Status) %>%
+  summarise( mean = mean(n.var), sd = sd(n.var), med = median(n.var))
+t.test( subset(sim, Status == 'NP')$n.var, subset(sim, Status == 'P')$n.var )$p.value
 
@@
 
-```{r xval_pt, echo=T, warning=F, message=T, fig.align='left', fig.height=4, eval=T}
+```{r xval_pt, echo=T, warning=F, message=T, fig.align='left', fig.height=4, eval=F}
 # Leave out patients
@@
 
-ggplot(melt(data.frame(loo_xval[,c(1,3)])), aes(x=variable, y=value, fill=variable)) +  
+ggplot(melt(data.frame(loo_xval[,c(3)])), aes(x=variable, y=value, fill=variable)) +  
     geom_boxplot() + geom_jitter(shape=16, width=0.2) + geom_hline(yintercept=0.05, color='red', alpha=0.5) +
-    stat_boxplot(geom ='errorbar') + labs(y='p-values', x='', title='Leave one (patient) out clustering')
+    stat_boxplot(geom ='errorbar') + labs(y='p-values', x='Iterations', title='Leave one (patient) out clustering') +
+  scale_fill_manual(values=c('purple'), labels='clusters')
+
 ```
@@
 
-```{r xval_smp, echo=F, warning=F, message=T, fig.align='left', fig.height=6, eval=T}
+```{r xval_smp, echo=F, warning=F, message=T, fig.align='left', fig.height=6, eval=F}
 
@@
 for (i in 1:20) { 
-  simX = sum.patient.data[,c(2:4)]
-  rownames(simX) = sum.patient.data[,1]
+  simX = sum.patient.data[,c('Status','years')]
+  rownames(simX) = sum.patient.data[,'Patient']
 
@@
 
-ggplot(melt(data.frame(looS_xval[,c(2,3)])), aes(x=variable, y=value, fill=variable)) +  
+ggplot(melt(data.frame(looS_xval[,c(3)])), aes(x=variable, y=value, fill=variable)) +  
     geom_boxplot() + geom_jitter(shape=16, position=position_jitter(0.2)) +
   geom_hline(yintercept=0.05, color='red', alpha=0.5) +
-    stat_boxplot(geom ='errorbar') + labs(y='p-values', x='', title='Leave one (sample) out clustering')
+    stat_boxplot(geom ='errorbar') + labs(y='p-values', x='Iterations', title='Leave one (sample) out clustering') + 
+    scale_fill_manual(values=c('#CC99FF'), labels='clusters')
+
 ```
@@
 ```{r nofinal, echo=F, message=F, warning=F, fig.height=6, fig.width=10}
-simX = sum.patient.data[,c(2:4)]
-rownames(simX) = sum.patient.data[,1]
+
+gain.threshold = 1.1
+loss.threshold = 0.9
+
+simX = sum.patient.data[,c('Status','years')]
+rownames(simX) = sum.patient.data[,'Patient']
 # Leave out final sample on the progressors
@@
 for (pt in sum.patient.data$Patient) {
+  print(pt)
   x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
@@
   simX[pt, 'clusters'] = length(clusters(ac))
+
+  simX[pt, 'n.var'] = nrow(patient.data[[pt]]$variable.regions[,-(1:5)])
 }
 
-ttps = t.test( subset(simX, Status == 'NP')$ss, subset(simX, Status == 'P')$ss )$p.value
-gg1 = ggplot(melt(data.frame(simX[c('Status','ss')])), aes(x=Status, y=value, fill=Status)) +  
-    geom_boxplot() + geom_jitter(shape=16, position=position_jitter(0.2)) +
-    stat_boxplot(geom ='errorbar') + labs(y='sim values', x='', title=paste('Leave out HGD sample(s) p-value', signif(ttps, 3)))
-
 ttpc = t.test( subset(simX, Status == 'NP')$clusters, subset(simX, Status == 'P')$clusters )$p.value
-gg2 = ggplot(melt(data.frame(simX[c('Status','clusters')])), aes(x=Status, y=value, fill=Status)) +  
+ggplot(melt(data.frame(simX[c('Status','clusters')])), aes(x=Status, y=value, fill=Status)) +  
     geom_boxplot() + geom_jitter(shape=16, position=position_jitter(0.2)) +
@@
 
-grid.arrange(gg1, gg2, ncol=2)
+t.test(subset(simX, Status == 'NP')$n.var, subset(simX, Status == 'P')$n.var)
 
@@
 
-pander(subset(simX, Status == 'P' & ss >= mean(ss)), caption="These patients are more similar to Non-Progressors without their HGD sample(s)")
+pander(subset(simX, Status == 'P' & clusters >= mean(clusters)), caption="These patients are more similar to Non-Progressors without their HGD sample(s)")
 
-simX = simX[, c('Status', 'clusters', 'ns','ss')]
+fit = glm(Status~clusters, simX, family=binomial(link='logit') )
+summary(fit)
+
+simX = simX[, c('Status', 'clusters', 'ns','ss','gain','loss')]
 noHGDPR = as.data.frame(matrix(0,1000,2,dimnames=list(c(), c('prec','recall'))))
@@
   geom_hline(yintercept = 0.5, col='red', alpha=0.5) +
+  geom_vline(xintercept = 0.5, col='red', alpha=0.5) +
   labs(title='Precision/Recall for Progressor/NP prediction NO HGD')
@@
 
-pvals = as.data.frame(matrix(ncol=2,nrow=0,dimnames=list(c(), c('clusters','ss'))))
-for (qv in seq(0.1,0.9,0.1)) {
+pvals = as.data.frame(matrix(ncol=3,nrow=0,dimnames=list(c(), c('p.clusters','mean.clusters', 'sd.clusters'))))
+status.clusters = list()
+for (qv in seq(0,0.9,0.1)) {
   print(qv)
@@
     simPC[pt, 'ns'] = net.similarity(ac)
-    simPC[pt, 'ss'] = sum.similarity(ac)
+    #simPC[pt, 'ss'] = sum.similarity(ac)
     simPC[pt, 'clusters'] = length(clusters(ac))
@@
   # May be a difference
-  ttps = t.test( subset(simPC, Status == 'NP')$ss, subset(simPC, Status == 'P')$ss )$p.value
+  #ttps = t.test( subset(simPC, Status == 'NP')$ss, subset(simPC, Status == 'P')$ss )$p.value
   ttpn = t.test( subset(simPC, Status == 'NP')$ns, subset(simPC, Status == 'P')$ns )$p.value
@@
 
-  pvals = rbind(pvals,cbind(ttpc,ttps))
+  pvals = rbind(pvals,cbind(ttpc, mean(simPC$clusters), sd(simPC$clusters)))
+
+  by.clusters = simPC %>% group_by(Status) %>% summarise( n.mean.clusters=mean(clusters), n.sd.clusters=sd(clusters), se.clusters=sd(clusters)/sqrt(length(Status)) )
+
+  status.clusters[[paste('q=',qv, sep='')]] = by.clusters
 }
-colnames(pvals) = c('clusters','sum.similarity')
-pvals$qv=seq(0.1,0.9,0.1)
+colnames(pvals) = c('p.clusters','mean.clusters', 'sd.clusters')
+pvals$qv=seq(0,0.9,0.1)
 
-ggplot(melt(pvals, id.vars='qv'),aes(x=qv,y=value, color=variable)) + geom_point() +
-  labs(main="qvalus for AP", y='pvalue NP vs P', x='q param')
+clus = as.data.frame(do.call(rbind, lapply(status.clusters, function(x) c(x[['n.mean.clusters']], x[['n.sd.clusters']]  ))))
+colnames(clus) = c('NP','P', 'NP.sd', 'P.sd')
+clus$qv = seq(0,0.9,0.1)
+
+m = melt(clus[,c(1:2,5)], measure.vars=c('NP','P'))
+m[which(m$variable == 'NP'), 'sd'] = clus$NP.sd
+m[which(m$variable == 'P'), 'sd'] = clus$P.sd
+
+ggplot(m, aes(x=qv, y=value, col=variable)) + geom_point() +  
+  geom_errorbar(aes(ymin=value-sd, ymax=value+sd), width=.01, alpha=0.3) +
+  geom_hline(yintercept=5, alpha=0.3) + geom_hline(yintercept=15, alpha=0.3, col='red') + 
+  labs(title='Clusters vs q parameters', x='q parameter', y='mean(clusters)')
+
+ggplot(melt(pvals, measure.vars=c('p.clusters'), id.vars='qv'),aes(x=qv,y=value, color=variable)) + geom_point() +
+  labs(title='P vs NP p-values per q parameters', y='pvalue NP vs P', x='q param')
 
@@
 
-```{r precrec, echo=F, warning=F, message=F, fig.height=12, fig.width=20}
+```{r model, echo=F, warning=F, message=F, fig.height=12, fig.width=20, eval=F}
+
+t.test(subset(sim, Status == 'P')$clusters, subset(sim, Status == 'NP')$clusters)
+
+t.test(subset(simX, Status == 'P')$clusters, subset(simX, Status == 'NP')$clusters)
+
+## TODO  Need to consider segment lengths relative to chromosomal lengths
+
+var.lengths = as.data.frame( do.call(rbind, lapply(patient.data, function(x){
+  if (nrow(x[['variable.regions']]) > 0) {
+    lengths = as.vector( (x[['variable.regions']]['end.pos'] - x[['variable.regions']]['start.pos'])[,1] )
+    return(c(median(lengths), mean(lengths), sd(lengths), sd(lengths)/length(lengths)))
+  }
+  return(NA)
+})) )
+colnames(var.lengths) = c('median','mean','sd','se')
+
+var.lengths = merge(sum.patient.data, var.lengths, by.x='Patient', by.y='row.names')
+
+var.lengths
+wilcox.test(subset(var.lengths, Status == 'P')$mean, subset(var.lengths, Status == 'NP')$mean)
 

### 2017-02-15 lib/load_patient_metadata.R

@@ -1,4 +1,4 @@
-library(xlsx)
+
 library(plyr)
 
 strip.whitespace <- function (x) gsub("\\s+|\\s+", "", x)
@@ -7,13 +7,17 @@ read.patient.info<-function(file) {
   if (!file.exists(file))
     stop(paste(file, "doesn't exist or is not readable."))
 
+  if (grepl('\\.xlsx$', file)) {
+    library(xlsx)
+  }
+  
   patient.info = NULL
   for (i in 1:length(getSheets(loadWorkbook(file)))) {
     print(i)
     if (names(getSheets(loadWorkbook(file))[i]) == 'Technical Repeats') next
 
     ws = read.xlsx2(file, sheetIndex=i, stringsAsFactors=F, header=T)
-    ws = ws[which( ws$Patient.ID != ""), grep('Patient|Path\\.ID|Endoscopy.Year|Pathology$|Progressor|Plate.Index|SLX|cellularity|p53', colnames(ws), value=T, ignore.case=T)]
+    ws = ws[which( ws$Patient.ID != ""), grep('Patient|Path\\.ID|Endoscopy.Year|Pathology$|Progressor|Plate.Index|SLX|cellularity|p53|Number', colnames(ws), value=T, ignore.case=T)]
     head(ws)
     #head(patient.info)
     
@@ -33,10 +37,14 @@ read.patient.info<-function(file) {
                grep('Plate.Index',colnames(ws),ignore.case=T),
                grep('SLX',colnames(ws),ignore.case=T),
                grep('cellularity',colnames(ws),ignore.case=T),
-               grep('p53', colnames(ws),ignore.case=F)) ]
+               grep('p53', colnames(ws),ignore.case=F),
+               grep('Number.of', colnames(ws), ignore.case=F)) ]
     
     
-    colnames(ws) = c('Patient','Path.ID','Status','Endoscopy.Year','Pathology','Plate.Index','SLX.ID','Barretts.Cellularity', 'p53.Status')
+    colnames(ws) = c('Patient','Path.ID','Status','Endoscopy.Year','Pathology','Plate.Index','SLX.ID','Barretts.Cellularity', 'p53.Status', 'Total.Reads')
+    
+    ws[ws$p53.Status == "", 'p53.Status'] = NA
+    ws[ws$Barretts.Cellularity == "", 'Barretts.Cellularity'] = NA
     
     if (is.null(patient.info)) {
       patient.info = ws
@@ -45,13 +53,14 @@ read.patient.info<-function(file) {
[…]

### 2017-02-24 Clustering.Rmd
Updating reports to use with new data

@@
 author: "Sarah Killcoyne"
-date: "12/6/2016"
+date: "23/02/2016"
 output: html_document
@@
 library(biomaRt)
-
+library(multtest)
 library(tibble)
@@
 
+load('patients.Rdata', verbose=T)
+
 patient.info = read.patient.info(patient.file)
@@
 
-load('patients.Rdata', verbose=T)
-
-sum.patient.data = as.data.frame(patient.info %>%
-  group_by(Patient, Status) %>%
-  summarise(
-    years=paste(range(Endoscopy.Year), collapse='-'),
-    n.samples=length(Samplename), med.cellularity=median(Barretts.Cellularity,na.rm=T),
-    mean.reads=mean(Total.Reads),
-    sd.reads=sd(Total.Reads)
-  ))
-
+sum.patient.data = summarise.patient.info(patient.info)
+sum.patient.data = as.data.frame(subset(sum.patient.data, Initial.Analysis)) ## For now
 pander(sum.patient.data)
@@
 
-ttr = t.test(subset(patient.info, Status == 'NP')$Total.Reads, subset(patient.info, Status == 'P')$Total.Reads)
-ttrbe = t.test(subset(be, Status == 'NP')$Total.Reads, subset(be, Status == 'P')$Total.Reads)
-ttrlgd = t.test(subset(lgd, Status == 'NP')$Total.Reads, subset(lgd, Status == 'P')$Total.Reads)
-
-tbclgd = t.test(subset(lgd, Status == 'NP')$Barretts.Cellularity, subset(lgd, Status == 'P')$Barretts.Cellularity)
-tbcbe = t.test(subset(be, Status == 'NP')$Barretts.Cellularity, subset(be, Status == 'P')$Barretts.Cellularity)
-
-t.test(subset(patient.info, Pathology == 'BE')$Barretts.Cellularity, 
-               subset(patient.info, Pathology == 'LGD')$Barretts.Cellularity)
-
-t.test(subset(patient.info, Pathology == 'LGD')$Barretts.Cellularity, 
-               subset(patient.info, Pathology == 'HGD')$Barretts.Cellularity)
-
-## No differences between the number of reads by pathology
-t.test(subset(patient.info, Pathology == 'BE')$Total.Reads, subset(patient.info, Pathology == 'LGD')$Total.Reads)
-t.test(subset(patient.info, Pathology == 'BE')$Total.Reads, subset(patient.info, Pathology == 'HGD')$Total.Reads)
-t.test(subset(patient.info, Pathology == 'LGD')$Total.Reads, subset(patient.info, Pathology == 'HGD')$Total.Reads)
-
-t.test(subset(patient.info, Pathology == 'BE')$Barretts.Cellularity, subset(patient.info, Pathology == 'LGD')$Barretts.Cellularity)
-t.test(subset(patient.info, Pathology == 'BE')$Barretts.Cellularity, subset(patient.info, Pathology == 'HGD')$Barretts.Cellularity)
-t.test(subset(patient.info, Pathology == 'LGD')$Barretts.Cellularity, subset(patient.info, Pathology == 'HGD')$Barretts.Cellularity)
-
+set.seed(99)
+mtp1 = MTP(X=t(patient.info[,c('Barretts.Cellularity','Total.Reads')]), Y=as.integer((as.integer(patient.info$Pathology) > 3)), robust=T, get.adjp=T, na.rm=T)
+mtp2 = MTP(X=t(patient.info[,c('Barretts.Cellularity','Total.Reads')]), Y=patient.info$Status, robust=T, get.adjp=T, na.rm=T)
 ```
 
-There are a few things we need to be careful of and a couple of questions to check here:
+Currently with multi test correction the samples for which cellularity and #of reads are available show no difference for either status (P vs NP)
+`r pander(summary(mtp1), justify='left')`
 
-1. The total number of reads per sample between NP and P might show a difference, but currently with multi-test correction it's not significant.
-`r  signif(ttr$p.value,2)*11`
-So are those samples with lower numbers of reads showing less complexity overall?
-When sequencing NP are we likely to get fewer reads than with P?
-If we look only at the samples with BE (no LGD or HGD samples) the difference disappears `r signif(ttrbe$p.value, 2)*11`.
-If we include LGD with the BE samples we again see the difference.
-`r signif(ttrlgd$p.value, 2)*11`.
-This is suggestive that changes in copy number result in more reads being output.
+or for samples that are LGD/IMC/HGD vs BE
+`r pander(summary(mtp1), justify='left')`
 
-2.
-There is a significant difference in the cellularity of the Barretts sample, with progressors having a higher mean cellularity, even when only comparing samples that were either LGD or BE.
-`r signif(tbclgd$p.value, 2)` However, even the BE only samples show a higher cellularity overall `r signif(tbcbe$p.value, 2)*11`  That said, currently we have a high rate of missing information, with about `r length(which(is.na(subset(patient.info, Status == 'P')$Barretts.Cellularity)))/length(which(patient.info$Status == 'P'))*100`% of the cellualarities missing for progressors and `r length(which(is.na(subset(patient.info, Status == 'NP')$Barretts.Cellularity)))/length(which(patient.info$Status == 'NP'))*100`% of the NP cellularities missing.
-
-3. With regards to p53 staining, non-progressors have no p53 positive tests, while progressors had `r p53Status[2,2]` overall.
+With regards to p53 staining, non-progressors have no p53 positive tests, while progressors had `r p53Status[2,2]` overall.
 `r pander(p53Status, justify='left', caption="P53 staining status for progressors vs non")`
@@
 
-```{r pca, echo=F, warning=F, message=F, fig.width=20, fig.height=20}
-sum.patient.data = arrange(sum.patient.data, Status, n.samples)
+```{r pca, echo=F, warning=F, message=F, fig.width=20, fig.height=40}
+sum.patient.data = arrange(sum.patient.data, Status, total.samples)
 
@@
     eigenvalues = pca$sdev^2
-    screeplot(pca, main=paste(pt, df, 'scree'))
+    screeplot(pca, main=paste(pt, sv, 'scree'))
 
@@
   }  
-  do.call(grid.arrange, c(plots, list(ncol=2))
-  return(gg)
-}
-
-plot.pca.norm<-function(pt) {
-
-  info = arrange(patient.data[[pt]]$info, Endoscopy.Year)
-
-  x1 = patient.data[[pt]]$norm.seg.vals[, patient.data[[pt]]$info$Samplename]
-  x1 = x1[,patient.data[[pt]]$info$Samplename]
-  colnames(x1) = paste(patient.data[[pt]]$info$Endoscopy.Year, '(', patient.data[[pt]]$info$Pathology, ')', sep='' )
-
-  pca = prcomp(t(x1), scale=T)
-
-  eigenvalues = pca$sdev^2
-  #screeplot(pca, main='scree')
-
-  #plot(pca$sd^2)
-
-  df = as.data.frame(pca$x)
-  df$year = as.factor(info$Endoscopy.Year)
-  df$label = paste(info$Endoscopy.Year, '(', info$Pathology, ')', sep='' )
-  df$Pathology = info$Pathology
-
-  gg = ggplot(df, aes(PC1, PC2)) + geom_line() +
-    geom_label(aes(label=year),nudge_y=0.04) + 
-    geom_point( aes(color=Pathology, size=Pathology)) +
-    labs(title=pt)
-
-  return(gg)
+  ga = do.call(grid.arrange, c(plots, list(ncol=2)))
+  return(ga)
 }
@@
 pplots = lapply(subset(sum.patient.data, Status == 'P')$Patient, plot.pca)
-grid.arrange(grobs=pplots, ncol=2, top="Progressors")
+grid.arrange(grobs=pplots, nrow=length(subset(sum.patient.data, Status == 'P')$Patient), top="Progressors")
 
 npplots = lapply(subset(sum.patient.data, Status == 'NP')$Patient, plot.pca)
-grid.arrange(grobs=npplots, ncol=2, top="Non-Progressors")
+grid.arrange(grobs=npplots, nrow=length(subset(sum.patient.data, Status == 'NP')$Patient), top="Non-Progressors")
 
@@
 
-I tried were several mixture models (e.g. EM) but wasn't able to find a pattern that could reliably separate the few mixtures I did see
+I tried were several mixture models (e.g. EM, kmeans) but wasn't able to find a pattern that could reliably separate the few mixtures I did see
 
-```{r clust, echo=T, message=T, warning=F, eval=T, fig.height=6, fig.width=6}
+### Raw segmented values
 
-llk = vector(mode='numeric', length=nrow(sum.patient.data))
-names(llk) = sum.patient.data$Patient
+#### EM
+```{r clust, echo=T, message=T, warning=F, eval=T, fig.height=4, fig.width=4, eval=T}
+
+em.dr<-function(patient, status, x1, var.regions) {
+  clusters=c(2:4)
+  #llk = matrix(ncol=length(clusters),nrow=nrow(sum.patient.data), dimnames=list(sum.patient.data$Patient, clusters))
+
+  llk = vector(length=length(clusters), mode='numeric')
+  names(llk) = as.character(clusters)
+
+## Ok, this doesn't separate regions well either normalized or raw segment values.  Typically the clusters are highly overlapped and not very descriptive
+#for (pt in sum.patient.data$Patient) {
+    print(pt)
+#    status = subset(sum.patient.data, Patient == pt)$Status
+#    print(ncol(patient.data[[pt]]$seg.vals[,-(1:5)]))
+#    x1 = as.matrix(patient.data[[pt]]$seg.vals[,-(1:5)])
+#    x1 = x1[,patient.data[[pt]]$info$Samplename]
+    #x1 = as.matrix(patient.data[[pt]]$norm.seg.vals)
+
+    for (g in clusters) {
+      print(g)
+      model = densityMclust(x1, G=g)
+      highCluster = which.max(colSums(model$parameters$mean))
+
+      tryCatch({
+        mod1dr = MclustDR(model, lambda = 0.5)
+
+        #     par(mfrow=c(2,1)) 
+        plot(mod1dr, what="evalues")
+        title(paste(pt, '(',status,') G=', g, sep=''))
+        plot(mod1dr, what='classification') # while this looks like PCA, it's not quite
+        title(paste(pt, '(',status,') G=', g, sep=''))
+        #     dev.off()
+        plot(mod1dr, what='density', dimens=1)
+        title(paste(pt, '  G=', g, sep=''))
+      }, error = function(e) {
+        cat(paste(pt, g, "failed DR"))
+      })
+
+      a1 = rownames(x1[which(model$classification == highCluster),] )
+      if (nrow(var.regions) > 0) {
+        am = merge(a1, var.regions, all=T)
+        pander( as.data.frame(rownames(x1[which(model$classification == highCluster),])),
+             caption=paste("From the cluster with the highest means the regions above are selected, of these ",length(setdiff(rownames(var.regions), a1)),
+                           " regions are not found in the variable regions selected using cutoffs.", length(intersect(rownames(var.regions), a1)), " are in both."))
+      } else {
+        pander(rownames(x1[which(model$classification == highCluster),]),
+             caption=paste("From the cluster with the highest means the regions above are selected. This patient had no variable regions selected based on cutoffs"))
+
+      }
+
+      llk[as.character(g)] = model$loglik
+    }
+    return(llk)
+}
+
+llk = as.data.frame(matrix(nrow=nrow(sum.patient.data),ncol=3,dimnames=list(c(sum.patient.data$Patient), c(2:4))))
 for (pt in sum.patient.data$Patient) {
     print(pt)
-  print(ncol(patient.data[[pt]]$seg.vals[,-(1:5)]))
+    status = subset(sum.patient.data, Patient == pt)$Status
 
-  x1 = as.matrix(patient.data[[pt]]$seg.vals[,-(1:5)])
+    x1 = patient.data[[pt]]$seg.vals
+    rownames(x1) = apply(x1, 1, function(x) strip.whitespace(paste(x[['chrom']], ':', x[['start.pos']], '-', x[['end.pos']], sep=''))) 
     x1 = x1[,patient.data[[pt]]$info$Samplename]
 
-  fit <- Mclust(x1, G=3)
-  plot(fit, what='classification')
-  title(main=paste(pt, ' (',patient.data[[pt]]$info$Status[1], ')', sep=''))
+    var = patient.data[[pt]]$variable.regions
+    if (nrow(var) > 0)
+      rownames(var) = apply(var, 1, function(x) strip.whitespace(paste(x[['chrom']], ':', x[['start.pos']], '-', x[['end.pos']], sep='')))
 
-  llk[pt] = fit$loglik
-
-#   clusterPlots <- data.frame(x1 = clusterData$x1, x2 = clusterData$x2, cluster = factor(model$classification, 
-#     levels = c(1, 2), labels = c("z = 0", "z = 1")), uncertainity = model$uncertainty, 
-#     logUncertainity = log(model$uncertainty))
-# clusterPlots.gg <- ggplot(clusterPlots)
-# clusterPlots.gg + geom_point(aes(x = x1, y = x2, color = cluster))
-
-  #tryCatch({
-  # model = densityMclust(x1, G=3)
-  # plotDensityMclustd(model, data=x1,  col='blue', lwd=2, hist.col = "lightblue",  breaks=100)
-  #   mod1dr = MclustDR(model, lambda = 0.5)
-  #   #summary(mod1dr)
-  #   table(mod1dr$mixcomp)
-  # 
-  #   #plot(mod1dr, what = "evalues")
-  #   #title(main=pt)
-  #   plot(mod1dr, what = "density")
-  #   title(main=paste(pt, ' (',patient.data[[pt]]$info$Status[1], ')', sep=''))
-  # })
+    llk[pt,] =  em.dr(pt,status,as.matrix(x1), var[,-(1:5)])
   }
 
-# Log liklihood doesn't separate them
-t.test(llk[subset(sum.patient.data, Status == 'NP')$Patient], llk[subset(sum.patient.data, Status == 'P')$Patient])
+ggplot(melt(as.matrix(llk)), aes(x=Var1, y=value, color=Var1)) + geom_point() + 
+  theme(axis.text.x=element_text(angle = 45), legend.position='none')
+  labs(title="Raw segment value log liklihoods", x='', y='loglik')
 
+# Log liklihood doesn't separate them individually or the range
+llk.diff = apply(llk, 1, function(x) diff(range(x)) )
+wilcox.test(llk.diff[subset(sum.patient.data, Status == 'NP')$Patient], llk.diff[subset(sum.patient.data, Status == 'P')$Patient])
+
+#t.test(llk[subset(sum.patient.data, Status == 'NP')$Patient,'2'], llk[subset(sum.patient.data, Status == 'P')$Patient,'2'])
+#t.test(llk[subset(sum.patient.data, Status == 'NP')$Patient,'3'], llk[subset(sum.patient.data, Status == 'P')$Patient,'3'])
+#t.test(llk[subset(sum.patient.data, Status == 'NP')$Patient,'4'], llk[subset(sum.patient.data, Status == 'P')$Patient,'4'])
 ```
 
-```{r means, echo=F, message=F, warning=F, eval=F} 
+#### Kmeans
+```{r kmeansraw, echo=F, message=F, warning=F, eval=F, fig.height=6, fig.width=6} 
 ## Kmeans
@@
 # 
+for (pt in sum.patient.data$Patient) {
+    print(pt)
+    status = subset(sum.patient.data, Patient == pt)$Status
+    x1 = patient.data[[pt]]$seg.vals
+    rownames(x1) = apply(patient.data[[pt]]$seg.vals[,c(1:4)], 1, function(x) strip.whitespace(paste(x[['chrom']], ':', x[['start.pos']], '-', x[['end.pos']], sep=''))) 
+    x1 = x1[,patient.data[[pt]]$info$Samplename]
+
     fit = kmeans(x1, 3)
-#   plot(x1[,c(1,10)], col=km$cluster)
-# title(main=paste(pt, ' (',patient.data[[pt]]$info$Status[1], ')', sep=''))
+    autoplot(fit, data=x1, label=T) + labs(title=paste(pt, '(', status, ') raw seg', sep=''))
+    pander(patient.data[[pt]]$variable.regions[,(1:4)], caption='Variable regions selected by cutoffs')
+}
+```
 
-library(cluster) 
-clusplot(x1, fit$cluster, color=TRUE, shade=TRUE, labels=2, lines=0, main=pt)
+### Normalized segmented values
+```{r clust2, echo=T, message=T, warning=F, eval=T, fig.height=4, fig.width=4, eval=T}
+llk = as.data.frame(matrix(nrow=nrow(sum.patient.data),ncol=3,dimnames=list(c(sum.patient.data$Patient), c(2:4))))
+for (pt in sum.patient.data$Patient) {
+    print(pt)
+    status = subset(sum.patient.data, Patient == pt)$Status
+    x1 = patient.data[[pt]]$norm.seg.vals
+    rownames(x1) = apply(patient.data[[pt]]$seg.vals[,c(1:4)], 1, function(x) strip.whitespace(paste(x[['chrom']], ':', x[['start.pos']], '-', x[['end.pos']], sep=''))) 
+    x1 = x1[,patient.data[[pt]]$info$Samplename]
 
+    var = patient.data[[pt]]$variable.regions
+    if (nrow(var) > 0)
+      rownames(var) = apply(var, 1, function(x) strip.whitespace(paste(x[['chrom']], ':', x[['start.pos']], '-', x[['end.pos']], sep='')))
+
+    llk[pt,] = em.dr(pt,status,x1, var)
+  }
+
+ggplot(melt(as.matrix(llk)), aes(x=Var1, y=value, color=Var1)) + geom_point() + 
+  theme(axis.text.x=element_text(angle = 45), legend.position='none') + 
+  labs(title="Normalized segment value log liklihoods", x='', y='loglik')
+
+# Log liklihood doesn't separate them individually or the range
+llk.diff = apply(llk, 1, function(x) diff(range(x)) )
+wilcox.test(llk.diff[subset(sum.patient.data, Status == 'NP')$Patient], llk.diff[subset(sum.patient.data, Status == 'P')$Patient])
+
+#t.test(llk[subset(sum.patient.data, Status == 'NP')$Patient,'2'], llk[subset(sum.patient.data, Status == 'P')$Patient,'2'])
+#t.test(llk[subset(sum.patient.data, Status == 'NP')$Patient,'3'], llk[subset(sum.patient.data, Status == 'P')$Patient,'3'])
+#t.test(llk[subset(sum.patient.data, Status == 'NP')$Patient,'4'], llk[subset(sum.patient.data, Status == 'P')$Patient,'4'])
+```
+
+#### Kmeans
+```{r kmeansnorm, echo=F, message=F, warning=F, eval=F, fig.height=6, fig.width=6} 
+for (pt in sum.patient.data$Patient) {
+    print(pt)
+    status = subset(sum.patient.data, Patient == pt)$Status
+    x1 = patient.data[[pt]]$norm.seg.vals
+    rownames(x1) = apply(patient.data[[pt]]$seg.vals[,c(1:4)], 1, function(x) strip.whitespace(paste(x[['chrom']], ':', x[['start.pos']], '-', x[['end.pos']], sep=''))) 
+    x1 = x1[,patient.data[[pt]]$info$Samplename]
+
+    fit = kmeans(x1, 3)
+    print(autoplot(fit, data=x1, label=T) + labs(title=paste(pt, '(', status, ') normalized seg', sep='')))
+    pander(patient.data[[pt]]$variable.regions[,(1:4)], caption='Variable regions selected by cutoffs')
+}
 ```
@@
 ## Using the segmentation values
+sum.patient.data$years = with(sum.patient.data, paste(start.year, end.year, sep='-'))
 sim = sum.patient.data[,c('Status','years')]
@@
   sim[pt, 'clusters'] = length(clusters(ac))
+  sim[pt, 'max.sd'] = max(sapply(x1, sd))
 
@@
 sim$id = rownames(sim)
-
-sm = as.data.frame(summarise(group_by(sim, Status), mean.ns=mean(ns), sd.ns=sd(ns), mean.ss=mean(ss), sd.ss=sd(ss), mean.c=mean(clusters), sd.c=sd(clusters) ))
 
@@
 
-fit = glm(Status~clusters, sim, family=binomial(link='logit'))
-pander(fit, justify='left')
+# fit = glm(Status~clusters, sim, family=binomial(link='logit'))
+# pander(fit, justify='left')
+# 
+# sim = sim[, c('Status', 'clusters', 'ns','ss')]
+# actualMatrix = matrix(0,1000,2,dimnames=list(c(), c('prec','recall')))
+# for (i in 1:nrow(actualMatrix)) {
+#   trainingRows = sample(1:nrow(sim), nrow(sim)-5)
+#   train = sim[trainingRows,]
+#   test = sim[-trainingRows,]
+#   
+#   fit = glm( (as.integer(Status)-1)~clusters, train, family=binomial(link='logit'))
+#   pred = predict(fit, test, se.fit=F) # probabilities -- of which outcome though?
+#   
+#   actual = test$Status == 'P'
+#   predicted = pred>0
+#   
+#   actualMatrix[i,'prec'] = getPrecision( actual, predicted )
+#   actualMatrix[i, 'recall'] = getRecall( actual, predicted )
+# }
+# 
+# ggplot(as.data.frame(actualMatrix), aes(x=recall, y=prec)) + geom_point() + 
+#   geom_hline(yintercept = 0.5, col='red', alpha=0.5) +
+#   labs(title='Precision/Recall for Progressor/NP prediction')
 
-sim = sim[, c('Status', 'clusters', 'ns','ss')]
-actualMatrix = matrix(0,1000,2,dimnames=list(c(), c('prec','recall')))
-for (i in 1:nrow(actualMatrix)) {
-  trainingRows = sample(1:nrow(sim), nrow(sim)-5)
-  train = sim[trainingRows,]
-  test = sim[-trainingRows,]
+rc = cor.test(sim$max.sd, sim$clusters)
+```
 
-  fit = glm( (as.integer(Status)-1)~clusters, train, family=binomial(link='logit'))
-  pred = predict(fit, test, se.fit=F) # probabilities -- of which outcome though?
+#### Cellularity?
 
-  actual = test$Status == 'P'
-  predicted = pred>0
-
-  actualMatrix[i,'prec'] = getPrecision( actual, predicted )
-  actualMatrix[i, 'recall'] = getRecall( actual, predicted )
-}
-
-ggplot(as.data.frame(actualMatrix), aes(x=recall, y=prec)) + geom_point() + 
-  geom_hline(yintercept = 0.5, col='red', alpha=0.5) +
-  labs(title='Precision/Recall for Progressor/NP prediction')
-
-```
+We needed to check if cellularity was driving the difference in clusters.
+Overall that does not appear to be the case if we use the maximum standard deviation across all samples in a patient as a proxy: r=`r rc$estimate`, p-value=`r signif(rc$p.value, 2)`
 
@@
 
-simX = simX[, c('Status', 'clusters', 'ns','ss','gain','loss')]
+simX = simX[, c('Status', 'clusters', 'ns','ss','n.var')]
 noHGDPR = as.data.frame(matrix(0,1000,2,dimnames=list(c(), c('prec','recall'))))
@@
 
-ggplot(as.data.frame(noHGDPR), aes(x=recall, y=prec)) + geom_point() + 
-  geom_hline(yintercept = 0.5, col='red', alpha=0.5) +
-  geom_vline(xintercept = 0.5, col='red', alpha=0.5) +
-  labs(title='Precision/Recall for Progressor/NP prediction NO HGD')
+#ggplot(as.data.frame(noHGDPR), aes(x=recall, y=prec)) + geom_point() + 
+#  geom_hline(yintercept = 0.5, col='red', alpha=0.5) +
+#  geom_vline(xintercept = 0.5, col='red', alpha=0.5) +
+#  labs(title='Precision/Recall for Progressor/NP prediction NO HGD')
 
@@
 
-ggplot(m, aes(x=qv, y=value, col=variable)) + geom_point() +  
-  geom_errorbar(aes(ymin=value-sd, ymax=value+sd), width=.01, alpha=0.3) +
-  geom_hline(yintercept=5, alpha=0.3) + geom_hline(yintercept=15, alpha=0.3, col='red') + 
+ggplot(m, aes(x=factor(qv), y=value, col=variable)) + geom_point() +
+  geom_errorbar(aes(ymin=value-sd, ymax=value+sd), width=.02) +
+  geom_hline(yintercept=5, alpha=0.3) + 
+  geom_hline(yintercept=15, alpha=0.3, col='red') + 
+  scale_y_continuous(breaks=seq(0,max(m$value)+5, 10),labels=seq(0,max(m$value)+5, 10)  ) +
   labs(title='Clusters vs q parameters', x='q parameter', y='mean(clusters)')
 
-ggplot(melt(pvals, measure.vars=c('p.clusters'), id.vars='qv'),aes(x=qv,y=value, color=variable)) + geom_point() +
+pander(clus[,c('qv','NP','P')], caption="Mean of clusters by patient status per q parameter.")
+
+ggplot(melt(pvals, measure.vars=c('p.clusters'), id.vars='qv'),aes(x=factor(qv),y=value, color=variable)) + 
+  geom_point() + 
   labs(title='P vs NP p-values per q parameters', y='pvalue NP vs P', x='q param')

### 2017-02-24 Patient Report.Rmd
Updating reports to use with new data

--- 2017-01-04 Patient Report.Rmd
@@
 date: "15 November 2016"
-output: word_document
+output:
+  html_document: 
+    toc: yes
+  word_document: default
 ---
@@
 library(GenomicRanges)
-
+library(gridExtra)
 library(tibble)
@@
 data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
-plot.dir = paste(data, 'Analysis/multipcf_plots_fitted_perPatient', sep='/')
+plot.dir = paste(data, 'Analysis/multipcf_plots_fitted_perPatient-OLD', sep='/')
 
@@
 
-sum.patient.data = ddply(patient.info, .(Patient, Status), summarise, years=paste(range(Endoscopy.Year), collapse='-'), 
-      n.samples=length(Samplename), med.cellularity=median(Barretts.Cellularity,na.rm=T))
+sum.patient.data = summarise.patient.info(patient.info)
+sum.patient.data = subset(sum.patient.data, Initial.Analysis)
+
+patient.info = subset(patient.info, Patient %in% sum.patient.data$Patient & Batch.Name %in% levels(patient.info$Batch.Name)[1:5])
 
@@
 
-## Gamma = 250
-```{r cors, echo=F, message=T, warning=F, fig.height=6, fig.width=6, fig.align='left'}
+```{r cors, echo=F, message=T, warning=F, fig.height=6, fig.width=6, fig.align='left', include=F}
 ## Cluster the samples by coverage then pull out regions
@@
 
+  if(!file.exists(paste(patient.plot.dir, "/",patient.name,"_segmentedCoverage_fitted_gamma",gamma2,".txt",sep=""))) next
+
   segvals = read.table(paste(patient.plot.dir, "/",patient.name,"_segmentedCoverage_fitted_gamma",gamma2,".txt",sep=""),sep="\t",stringsAsFactors=F,header=T)
@@
   cur.pt = subset(patient.info, Patient == patient.name)
-  segvals[, c(colnames(segvals[1:5]), cur.pt$Samplename)]
+  segvals = segvals[, c(colnames(segvals[1:5]), intersect(colnames(segvals), cur.pt$Samplename))]
 
@@
       variable.region.indices = which(sds >= sd.threshold & segvals$n.probes >= min.probes)
-      variable.regions = segvals[variable.region.indices, patient.data[[patient.name]]$info$Samplename] 
+      variable.regions = segvals[variable.region.indices, intersect(patient.data[[patient.name]]$info$Samplename, colnames(segvals))] 
       patient.data[[patient.name]][['variable.regions']] = cbind(segvals[variable.region.indices,c(1:5)],variable.regions )
@@
         geom_histogram( fill=I("blue"),col=I("lightblue"),binwidth=0.3) + labs(title=paste(patient.name,'normalised segment coverage'), x='', y='')
-      print(gg3)
+     # print(gg3)
       #pt.plots[[patient.name]][['norm.hist']] = gg
@@
 
-```{r scores, echo=F, message=T, warning=F, fig.height=4,fig.width=4,fig.align='left'} 
-#pander(cor.scores[order(cor.scores$Status, cor.scores$similar),], justify='left')
+# Data
 
-#ggplot(cor.scores, aes(x=similar,y=different, color=Status)) + geom_point()
+This data was generated in the following pipeline:
 
-#ggplot(cor.scores, aes(y=abs(similar+different), x=c(1:nrow(cor.scores)), color=Status)) + geom_point()
+* Alignment using QDNAseq
+* Filter excluded regions
+* Joint segmentation per patient across all patient-derived-samples using PCF  (copynumber::multipcf method)
+
+The result looks like is two data matrices:
+
+1. Segmented values 
+```{r}
+head(patient.data[['AD0258']]$seg.vals)
 ```
 
-## Variable Regions
+```{r echo=F, warning=F, message=F, fig.height=10, fig.width=10}
+ggplot( melt(patient.data[['AD0258']]$seg.vals[,-(1:5)]), aes(x=value) ) + 
+        facet_wrap(~variable,scales = "free_x") + 
+        geom_histogram( fill=I("blue"),col=I("lightblue"), binwidth=0.03) + labs(title='AD0258 segmented value distribution', x='', y='')
+```
+
+2. Row (e.g. per-region across timepoints) normalized values from the matrix above
+```{r}
+head(patient.data[['AD0258']]$norm.seg.vals)
+```
+
+```{r echo=F, warning=F, message=F, fig.height=10, fig.width=10}
+ggplot( melt(patient.data[['AD0258']]$norm.seg.vals), aes(x=value) ) + 
+        facet_wrap(~variable,scales = "free_x") + 
+        geom_histogram( fill=I("blue"),col=I("lightblue"), binwidth=0.3) + labs(title='AD0258 normalized segmented value distribution', x='', y='')
+
+## Per-sample normalization
+#nsv = patient.data[['AD0258']]$norm.seg.vals
+#nsv[] = lapply(patient.data[['AD0258']]$seg.vals[,-(1:5)], function(x) (x-mean(x))/sd(x))
+
+#ggplot(melt(nsv), aes(x=value))  + 
+#        facet_wrap(~variable,scales = "free_x") + 
+#        geom_histogram( fill=I("blue"),col=I("lightblue"), binwidth=0.3) + labs(title='AD0258 normalized segmented value distribution', x='', y='')
+
+```
+
+Patients do not share the same chromosomal regions, though they can be overlapped, or the same timepoints. So all analyses have been conducted on a per-patient basis.
+
+```{r general, warning=F, echo=F, message=F}
+gen = do.call(rbind, lapply(patient.data, function(x) nrow(x[['seg.vals']])))
+colnames(gen) = c('Number.Regions')
+
+sum.patient.data = subset(sum.patient.data, Patient %in% names(patient.data))
+
+sum.patient.data = merge(sum.patient.data, gen, by.x='Patient', by.y='row.names')
+sum.patient.data = arrange(sum.patient.data, Status, total.samples)
+sum.patient.data$years = with(sum.patient.data, paste(start.year, end.year, sep='-'))
+
+pander(sum.patient.data[,c('Patient','Status','years','total.samples','highest.path','Number.Regions')], justify='left')
+```
+
+# Segmented Value Distributions
+```{r echo=F, message=F, warning=F, fig.height=15, fig.width=10}
+
+lapply(patient.data, function(pt) {
+  name = pt$info$Patient[1]
+  status = pt$info$Status[1]
+
+  g1 = ggplot( melt(pt[['seg.vals']][,-(1:5)]), aes(x=value) ) + 
+      facet_wrap(~variable,scales = "free_x") + 
+      geom_histogram( fill=I("blue"),col=I("lightblue"), binwidth=0.03) + 
+      labs(title='segmented value distribution', x='', y='')
+
+  g2 = ggplot( melt(pt[['norm.seg.vals']]), aes(x=value) ) + 
+      facet_wrap(~variable,scales = "free_x") + 
+      geom_histogram( fill=I("blue"),col=I("lightblue"), binwidth=0.3) + 
+      labs(title='normalized segmented value distribution', x='', y='')
+
+  do.call('grid.arrange', list(g1,g2, top=paste(name, status, sep=" - ")))
+})
+
+```
+
+# Variable Regions
+
+These are regions that are called as "variant" based on cutoffs derived from a single sample that had high-depth (50x) WGS done and was then downsampled. These cutoffs are:
+
+* Standard deviation of the segmented value matrix >= 0.08
+* The number of probes used >= 67, covers 1Mb 
+* Gains are called at >= 1.1
+* Losses are called at <= 0.9
+
 ```{r var_reg, echo=F, message=F, warning=F}
@@
 
-t.test(subset(np.regions, chrom %in% c(1:12))$chr.ratio, subset(np.regions, chrom %in% c(13:22))$chr.ratio)
+#t.test(subset(np.regions, chrom %in% c(1:12))$chr.ratio, subset(np.regions, chrom %in% c(13:22))$chr.ratio)
 
-t.test(subset(prog.regions, chrom %in% c(1:12))$chr.ratio, subset(prog.regions, chrom %in% c(13:22))$chr.ratio)
+#t.test(subset(prog.regions, chrom %in% c(1:12))$chr.ratio, subset(prog.regions, chrom %in% c(13:22))$chr.ratio)
 
@@
 
-There appears to be no enrichment based on the longer vs shorter chromosomes either (fisher `r ftL$p.value`)
+There appears to be no enrichment based on the length of chromosomes either (fisher `r ftL$p.value`)
 
-```{r cor, echo=F, message=T, warning=F, fig.height=4,fig.width=4, fig.align='left'}
+```{r cor, echo=F, message=F, warning=F, fig.height=10,fig.width=10, fig.align='left'}
 
@@
   sample.normalised.segvals = patient.data[[name]]$norm.seg.vals # prog
-  sample.normalised.segvals =  sample.normalised.segvals[, patient.data[[name]]$info$Samplename ]
+  sample.normalised.segvals =  sample.normalised.segvals[, intersect(patient.data[[name]]$info$Samplename, colnames(sample.normalised.segvals)) ]
 
-  colnames(sample.normalised.segvals) =  paste(patient.data[[name]]$info$Endoscopy.Year,  c(1:nrow(patient.data[[name]]$info)) , sep='.')
+  info = subset(patient.data[[name]]$info, Samplename %in% colnames(sample.normalised.segvals))
+
+  colnames(sample.normalised.segvals) =  paste(info$Endoscopy.Year,  c(1:nrow(info)) , sep='.')
 
@@
 
-## Look at Cancer Consensus Genes in the variable regions
+## How do Cancer Gene Consensus genes map to the variable regions?
 
-```{r, hmm, echo=F, message=F, warning=F}
+```{r, hmm, echo=F, message=F, warning=F, fig.height=40, fig.width=10}
 ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
@@
 
-m = melt(gene.p[1:50,c(1:4)], id.vars=c('Gene.Symbols', 'CN'))
+m = melt(gene.p[,c(1:4)], id.vars=c('Gene.Symbols', 'CN'))
 m$Gene.Symbols = factor(m$Gene.Symbols, levels= m$Gene.Symbols[c(nrow(m):1)] ) # order
@@
   scale_fill_gradient('n patients',low='white',high='royalblue4') + labs(x="",y='CGC Genes')
-p + geom_point(data=m[m$CN == 'gain' & m$value > 0,], aes(label='gain',y=Gene.Symbols, x=variable), size=2, shape=24, fill='red', color='red') +
-  geom_point(data=m[m$CN == 'loss' & m$value > 0,], aes(label='loss',y=Gene.Symbols, x=variable), color='green4', size=2, shape=25, fill='green4')
+p + geom_point(data=m[m$CN == 'gain' & m$value > 0,], aes(y=Gene.Symbols, x=variable), size=2, shape=24, fill='red', color='red') +
+  geom_point(data=m[m$CN == 'loss' & m$value > 0,], aes(y=Gene.Symbols, x=variable), color='green4', size=2, shape=25, fill='green4')
 
@@
 gc = arrange(subset(gc, Progressors >= Non.Progressors), -Progressors)
-m = melt(gc[1:50,], id.vars=c('Gene.Symbols'))
+m = melt(gc, id.vars=c('Gene.Symbols'))
 m$Gene.Symbols = factor(m$Gene.Symbols, levels= m$Gene.Symbols[c(nrow(m):1)] ) # order
@@
 
-## Is HGD over represented for variable regions?
+# Is HGD over represented for variable regions?
+
+It's possible that the variable regions we find are entirely within the HGD/IMC samples.  So are they overrepresented?
+
 ```{r, hgd, echo=F, message=F, warning=F}
@@
   print(name)
-  sn = grep('HGD', patient.data[[name]]$info$Pathology)
+  sn = grep('HGD|IMC', patient.data[[name]]$info$Pathology)
   if (length(sn) <= 0 | nrow(patient.data[[name]]$variable.regions) <= 0) next
@@
 
+ct = cor.test(subset(sum.patient.data, Status == 'P')$total.samples, pvnp$hgd.ratio)
+
 pander(pvnp, justify='left', 
@@
 There's no significant difference between them (t-test p=",signif(tt$p.value,2),').
-Not all patients have HGD samples (NAs).',sep=''))
+Not all patients have HGD samples (NAs).
+There is also no significant correlation between the number/ratio of variable regions found in HGD samples vs non (Pearsons p=',signif(ct$p.value, 2),')',  sep=''))
 

### 2017-02-28 Clustering.Rmd
Updating the clustering analysis to generate full report

@@
 date: "23/02/2016"
-output: html_document
+output: 
+  html_document: 
+    toc: yes
+    toc_depth: 4
 ---
@@
 library(mclust)
+library(ggfortify)
 
@@
 
-```{r global_pt, echo=F, message=F, warning=F}
+We have a few questions we needed to answer with this data.
+The pairwise correlation plots appeared to show a difference between progressors and non-progressors as well as some differences between samples within a single patient.
+So...
+
+1. Within a single patient, can we quantify in some way the diversity that appears to be present in the pairwise correlation plots? 
+2. Given a diversity measure within a single patient, can we differentiate the progressors from the non-progressors?  Can we do it prior to the HGD/cancer sample?
+3.
+Is there then a way to select patient-specific thresholds for calling variable regions?
+Alternately, given what we know about the cancer can we presume diploidy and treat these samples as if we had absolute copy number?
+
+The first two questions were addressed with the clustering below.  The third is in progress currently.
+
+```{r global_pt, echo=F, message=F, warning=F, error=F}
 be = subset(patient.info, Pathology == 'BE')
@@
 set.seed(99)
-mtp1 = MTP(X=t(patient.info[,c('Barretts.Cellularity','Total.Reads')]), Y=as.integer((as.integer(patient.info$Pathology) > 3)), robust=T, get.adjp=T, na.rm=T)
-mtp2 = MTP(X=t(patient.info[,c('Barretts.Cellularity','Total.Reads')]), Y=patient.info$Status, robust=T, get.adjp=T, na.rm=T)
+mtp1 = MTP(X=t(patient.info[,c('Barretts.Cellularity','Total.Reads')]), Y=as.integer((as.integer(patient.info$Pathology) > 3)), 
+           get.adjp=T, na.rm=T, test='t.twosamp.equalvar')
+mtp2 = MTP(X=t(patient.info[,c('Barretts.Cellularity','Total.Reads')]), Y=patient.info$Status, 
+            get.adjp=T, na.rm=T, test='t.twosamp.equalvar')
 ```
@@
 Currently with multi test correction the samples for which cellularity and #of reads are available show no difference for either status (P vs NP)
-`r pander(summary(mtp1), justify='left')`
+`r pander(mtp2@adjp, justify='left', caption="Adjusted p-value")`
 
 or for samples that are LGD/IMC/HGD vs BE
-`r pander(summary(mtp1), justify='left')`
+`r pander(mtp1@adjp, justify='left', caption="Adjusted p-value")`
+
+However, it should be noted that pathologists do not like calling cellularity on Barrett's generally so it's a fairly inaccurate measure.
+`r pander(table( patient.info[,c('Status','Barretts.Cellularity')]), caption="Numbers of cellularity calls by status")`
 
@@
 
-```{r pca, echo=F, warning=F, message=F, fig.width=20, fig.height=40}
+```{r pca, echo=F, warning=F, message=F, fig.width=8, fig.height=4}
 sum.patient.data = arrange(sum.patient.data, Status, total.samples)
@@
     eigenvalues = pca$sdev^2
-    screeplot(pca, main=paste(pt, sv, 'scree'))
+    #screeplot(pca, main=paste(pt, sv, 'scree'))
 
@@
 pplots = lapply(subset(sum.patient.data, Status == 'P')$Patient, plot.pca)
-grid.arrange(grobs=pplots, nrow=length(subset(sum.patient.data, Status == 'P')$Patient), top="Progressors")
-
 npplots = lapply(subset(sum.patient.data, Status == 'NP')$Patient, plot.pca)
-grid.arrange(grobs=npplots, nrow=length(subset(sum.patient.data, Status == 'NP')$Patient), top="Non-Progressors")
-
 ```
 
-## Mixture Models
+```{r echo=F, message=F, warning=F, pcaplots, fig.align='left', fig.height=30, fig.width=20}
+grid.arrange(grobs=pplots, nrow=length(subset(sum.patient.data, Status == 'P')$Patient), top="Progressors")
+grid.arrange(grobs=npplots, nrow=length(subset(sum.patient.data, Status == 'NP')$Patient), top="Non-Progressors")
+```
+
+# Clustering
 
@@
 
-I tried were several mixture models (e.g. EM, kmeans) but wasn't able to find a pattern that could reliably separate the few mixtures I did see
+## Mixture Models
+
+I tried were several mixture models (e.g. EM, kmeans) but wasn't able to find a pattern that could reliably separate the distributions.
 
@@
 
+First using EM (mclust) looking at three different values for G (2,3,4).
+
 #### EM
-```{r clust, echo=T, message=T, warning=F, eval=T, fig.height=4, fig.width=4, eval=T}
+```{r clust, echo=F, message=F, warning=F, fig.height=4, fig.width=4}
 
 em.dr<-function(patient, status, x1, var.regions) {
+  cat(paste('**', patient, '**'))
   clusters=c(2:4)
-  #llk = matrix(ncol=length(clusters),nrow=nrow(sum.patient.data), dimnames=list(sum.patient.data$Patient, clusters))
 
-  llk = vector(length=length(clusters), mode='numeric')
-  names(llk) = as.character(clusters)
-
-## Ok, this doesn't separate regions well either normalized or raw segment values.  Typically the clusters are highly overlapped and not very descriptive
-#for (pt in sum.patient.data$Patient) {
-    print(pt)
-#    status = subset(sum.patient.data, Patient == pt)$Status
-#    print(ncol(patient.data[[pt]]$seg.vals[,-(1:5)]))
-#    x1 = as.matrix(patient.data[[pt]]$seg.vals[,-(1:5)])
-#    x1 = x1[,patient.data[[pt]]$info$Samplename]
-    #x1 = as.matrix(patient.data[[pt]]$norm.seg.vals)
+  llk = list()
 
   for (g in clusters) {
-      print(g)
     model = densityMclust(x1, G=g)
@@
 
-      a1 = rownames(x1[which(model$classification == highCluster),] )
-      if (nrow(var.regions) > 0) {
-        am = merge(a1, var.regions, all=T)
-        pander( as.data.frame(rownames(x1[which(model$classification == highCluster),])),
-             caption=paste("From the cluster with the highest means the regions above are selected, of these ",length(setdiff(rownames(var.regions), a1)),
-                           " regions are not found in the variable regions selected using cutoffs.", length(intersect(rownames(var.regions), a1)), " are in both."))
-      } else {
-        pander(rownames(x1[which(model$classification == highCluster),]),
-             caption=paste("From the cluster with the highest means the regions above are selected. This patient had no variable regions selected based on cutoffs"))
-
-      }
-
-      llk[as.character(g)] = model$loglik
+    # a1 = rownames(x1[which(model$classification == highCluster),] )
+    # if (nrow(var.regions) > 0) {
+    #   am = merge(a1, var.regions, all=T)
+    #   pander( as.data.frame(rownames(x1[which(model$classification == highCluster),])), caption=paste("From the cluster with the highest means the regions above are selected, of these ",length(setdiff(rownames(var.regions), a1))," regions are not found in the variable regions selected using cutoffs.", length(intersect(rownames(var.regions), a1)), " are in both."))
+    # } else {
+    #   pander(as.data.frame(a1),caption=paste("From the cluster with the highest means the regions above are selected. This patient had no variable regions selected based on cutoffs"))
+    # 
+    # }
+    llk[[as.character(g)]] = model
   }
@@
 for (pt in sum.patient.data$Patient) {
-    print(pt)
     status = subset(sum.patient.data, Patient == pt)$Status
@@
 
-    llk[pt,] =  em.dr(pt,status,as.matrix(x1), var[,-(1:5)])
+    models = em.dr(pt,status,as.matrix(x1), var[,-(1:5)])
+    llk[pt,] = sapply(models, function(x) x$loglik)
   }
 
-ggplot(melt(as.matrix(llk)), aes(x=Var1, y=value, color=Var1)) + geom_point() + 
-  theme(axis.text.x=element_text(angle = 45), legend.position='none')
+ggplot(melt(as.matrix(llk)), aes(x=factor(Var1), y=value, color=factor(Var2))) + 
+  geom_point() +
+  theme(axis.text.x=element_text(angle = 45)) + 
   labs(title="Raw segment value log liklihoods", x='', y='loglik')
@@
 llk.diff = apply(llk, 1, function(x) diff(range(x)) )
-wilcox.test(llk.diff[subset(sum.patient.data, Status == 'NP')$Patient], llk.diff[subset(sum.patient.data, Status == 'P')$Patient])
+#t.test(llk.diff[subset(sum.patient.data, Status == 'NP')$Patient], llk.diff[subset(sum.patient.data, Status == 'P')$Patient])
 
@@
 #### Kmeans
+
 ```{r kmeansraw, echo=F, message=F, warning=F, eval=F, fig.height=6, fig.width=6} 
-## Kmeans
-#   x1 = as.matrix(patient.data[[pt]]$seg.vals[,-(1:5)])
-# 
 for (pt in sum.patient.data$Patient) {
-    print(pt)
     status = subset(sum.patient.data, Patient == pt)$Status
@@
     fit = kmeans(x1, 3)
-    autoplot(fit, data=x1, label=T) + labs(title=paste(pt, '(', status, ') raw seg', sep=''))
+    autoplot(fit, data=x1, label=F) + labs(title=paste(pt, '(', status, ') raw seg', sep=''))
     pander(patient.data[[pt]]$variable.regions[,(1:4)], caption='Variable regions selected by cutoffs')
@@
 ### Normalized segmented values
-```{r clust2, echo=T, message=T, warning=F, eval=T, fig.height=4, fig.width=4, eval=T}
+
+#### EM
+
+Row-normalized (per region)
+
+```{r clust2, echo=F, message=F, warning=F, fig.height=4, fig.width=4}
 llk = as.data.frame(matrix(nrow=nrow(sum.patient.data),ncol=3,dimnames=list(c(sum.patient.data$Patient), c(2:4))))
 for (pt in sum.patient.data$Patient) {
-    print(pt)
     status = subset(sum.patient.data, Patient == pt)$Status
@@
 
-    llk[pt,] = em.dr(pt,status,x1, var)
+        models = em.dr(pt,status,as.matrix(x1), var[,-(1:5)])
+    llk[pt,] = sapply(models, function(x) x$loglik)
   }
 
-ggplot(melt(as.matrix(llk)), aes(x=Var1, y=value, color=Var1)) + geom_point() + 
-  theme(axis.text.x=element_text(angle = 45), legend.position='none') + 
+ggplot(melt(as.matrix(llk)), aes(x=Var1, y=value, color=factor(Var2))) + geom_point() + 
+  theme(axis.text.x=element_text(angle = 45)) + 
   labs(title="Normalized segment value log liklihoods", x='', y='loglik')
@@
 llk.diff = apply(llk, 1, function(x) diff(range(x)) )
-wilcox.test(llk.diff[subset(sum.patient.data, Status == 'NP')$Patient], llk.diff[subset(sum.patient.data, Status == 'P')$Patient])
+#wilcox.test(llk.diff[subset(sum.patient.data, Status == 'NP')$Patient], llk.diff[subset(sum.patient.data, Status == 'P')$Patient])
 
@@
 #### Kmeans
-```{r kmeansnorm, echo=F, message=F, warning=F, eval=F, fig.height=6, fig.width=6} 
+```{r kmeansnorm, echo=F, message=F, warning=F, fig.height=6, fig.width=6} 
 for (pt in sum.patient.data$Patient) {
-    print(pt)
     status = subset(sum.patient.data, Patient == pt)$Status
@@
     fit = kmeans(x1, 3)
-    print(autoplot(fit, data=x1, label=T) + labs(title=paste(pt, '(', status, ') normalized seg', sep='')))
+    print(autoplot(fit, data=x1, label=F) + labs(title=paste(pt, '(', status, ') normalized seg', sep='')))
     pander(patient.data[[pt]]$variable.regions[,(1:4)], caption='Variable regions selected by cutoffs')
@@
 ```
+
+In neither the raw or normalized matrices could I reliably distinguish clusters from each other within a single patient.
+Some patients separate well (AH0254) others don't separate at all.
+I looked as well for a high level measuremeant that might separate the patients by progressor/non.
+Log liklihoods did not show any difference between patients in the mixture model, and with poor clusters separation I was a bit stumped as to what else might work.
+
+## AP Clustering
 
@@
 
-```{r clustN, echo=T, message=T, warning=F, eval=F}
-## Normalized Segmentation Values
-#Doesn't work for the normalized values.
+I then tried Affinity Propagation (described by Frey & Dueck, 2007) as that doesn't require the pre-definition for number of clusters.
+My hypothesis was that the progressors would show a higher number of clusters, and that the variable regions might tend to show up in the same clusters.
 
-for (pt in subset(sum.patient.data, Status == 'P')$Patient) {
-  if (pt == 'AH0254') next
-  print(pt)
-  print(paste('norm.seg.vals cols:', ncol(patient.data[[pt]]$norm.seg.vals)))
+### Raw segmentation values
 
-  x1 = patient.data[[pt]]$norm.seg.vals
-  model = Mclust(t(x1))
-  print(paste('loglik', round(model$loglik,2)))
-
-  tryCatch({
-    mod1dr = MclustDR(model, lambda = 1)
-    #summary(mod1dr)
-    table(mod1dr$mixcomp)
-
-    if (length(table(mod1dr$mixcomp)) > 1) {
-      plot(mod1dr, what = "evalues")
-      title(main=pt)
-      #plot(mod1dr, what = "density")
-    }
-  })
-}
-
-```
-
-## AP Clustering
-
-So tried AP as that doesn't require the pre-definition for number of clusters.  It appears to give good separation, though there's a suggestion of two different types of progressors.
-
-### Affinity Propagagion clustering described by Frey & Dueck, 2007
-
-First on the raw segmentation values
-
-```{r apclust, echo=T, message=T, warning=F, fig.height=10, fig.width=10, fig.align='left'}
+```{r apclust, echo=F, message=F, warning=F, fig.height=10, fig.width=10, fig.align='left'}
 q=0.2
@@
     last = sort(grep('HGD|IMC', patient.data[[pt]]$info$Pathology), decreasing=T)[1]
-  # if (ncol(x1) <= 15) {
-  #   plot(ac, x1[,c(c(1:ncol(x1))[-last], last)], main=pt)
-  # } else {
-  #   plot(ac, x1[,c(1, sort(sample(c(1:(ncol(x1)-1))[c(-1,-last)], 8)) ,last) ], main=pt)
-  # }
+  if (ncol(x1) <= 15) {
+    plot(ac, x1[,c(c(1:ncol(x1))[-last], last)], main=pt)
+  } else {
+    plot(ac, x1[,c(1, sort(sample(c(1:(ncol(x1)-1))[c(-1,-last)], 8)) ,last) ], main=pt)
+  }
 
@@
 sim$id = rownames(sim)
+rc = cor.test(sim$max.sd, sim$clusters)
 
@@
 ## Number of variable regions isn't so good, but maybe...
-sim %>%
-  group_by(Status) %>%
-  summarise( mean = mean(n.var), sd = sd(n.var), med = median(n.var))
-t.test( subset(sim, Status == 'NP')$n.var, subset(sim, Status == 'P')$n.var )$p.value
+#sim %>%
+#  group_by(Status) %>%
+#  summarise( mean = mean(n.var), sd = sd(n.var), med = median(n.var))
+#t.test( subset(sim, Status == 'NP')$n.var, subset(sim, Status == 'P')$n.var )$p.value
 
@@
   labs(title=paste("p-value", signif(ttpc, 3)), y='num. clusters')
+```
 
+This plot compares the two possible measures of overall complexity from AP.
+The summed similarity value (top) and the number of clusters (bottom).
+In this case, the number of clusters appears to be a more robust measure.
+
+```{r echo=F, message=F, warning=F, fig.width=5, fig.height=5}
 grid.arrange(g1, g2, top='Raw seg values')
+```
 
-# fit = glm(Status~clusters, sim, family=binomial(link='logit'))
-# pander(fit, justify='left')
-# 
-# sim = sim[, c('Status', 'clusters', 'ns','ss')]
-# actualMatrix = matrix(0,1000,2,dimnames=list(c(), c('prec','recall')))
-# for (i in 1:nrow(actualMatrix)) {
-#   trainingRows = sample(1:nrow(sim), nrow(sim)-5)
-#   train = sim[trainingRows,]
-#   test = sim[-trainingRows,]
-#   
-#   fit = glm( (as.integer(Status)-1)~clusters, train, family=binomial(link='logit'))
-#   pred = predict(fit, test, se.fit=F) # probabilities -- of which outcome though?
-#   
-#   actual = test$Status == 'P'
-#   predicted = pred>0
-#   
-#   actualMatrix[i,'prec'] = getPrecision( actual, predicted )
-#   actualMatrix[i, 'recall'] = getRecall( actual, predicted )
-# }
-# 
+Based on the number of clusters in each patient, it appears that there are a higher number of clusters being selected by AP in the progressors vs the non, suggesting it may be able to relate it to diversity in the progressors CN.
+That metric appears to provide a good separation between the patients as well, though there's a suggestion of two different types of progressors.
+
+```{r apglmclus, echo=F, message=F, warning=F}
+fit = glm(Status~clusters, sim, family=binomial(link='logit'))
+#pander(fit, justify='left')
+
+sim = sim[, c('Status', 'clusters', 'ns','ss')]
+actualMatrix = matrix(0,1000,2,dimnames=list(c(), c('prec','recall')))
+for (i in 1:nrow(actualMatrix)) {
+  trainingRows = sample(1:nrow(sim), nrow(sim)-5)
+  train = sim[trainingRows,]
+  test = sim[-trainingRows,]
+
+  fit = glm( (as.integer(Status)-1)~clusters, train, family=binomial(link='logit'))
+  pred = predict(fit, test, se.fit=F) # probabilities -- of which outcome though?
+
+  actual = test$Status == 'P'
+  predicted = pred>0
+
+  actualMatrix[i,'prec'] = getPrecision( actual, predicted )
+  actualMatrix[i, 'recall'] = getRecall( actual, predicted )
+}
+
 # ggplot(as.data.frame(actualMatrix), aes(x=recall, y=prec)) + geom_point() +
@@
 
-rc = cor.test(sim$max.sd, sim$clusters)
 ```
@@
 
-We needed to check if cellularity was driving the difference in clusters.
-Overall that does not appear to be the case if we use the maximum standard deviation across all samples in a patient as a proxy: r=`r rc$estimate`, p-value=`r signif(rc$p.value, 2)`
+There is a question about whether the cellularity of the sample drives the differences in the clusters.
+Overall that does not appear to be the case if we use the maximum standard deviation across all samples in a patient as a proxy for cellularity measure: r=`r rc$estimate`, p-value=`r signif(rc$p.value, 2)`
 
-### Are the clusters random?
+#### Are the clusters random?
 
-```{r example_pt}
+Especially with copy-number the clusters could simply be groups of neighboring regions from the same chromosomes as other analyses have shown high correlations between neighboring CN regions.
+So are the clusters that AP finds randomly distributed?
+
+```{r example_pt, echo=F, warning=F, message=F, fig.width=10, fig.height=10}
 pt = 'PR1_HIN_042'
-patient.data[[pt]]$info = arrange(patient.data[[pt]]$info, Endoscopy.Year, desc(Pathology))
+patient.data[[pt]]$info = arrange(patient.data[[pt]]$info, Endoscopy.Year, Pathology)
 
@@
 ac = apcluster(negDistMat(x1, r=2), q=q, details=T, includeSim=T)
+plot(ac, x1, main=pt)
 
@@
 hgmx = as.data.frame(hgmx)
-
-ggplot(subset(hgmx, x > 0), aes(x=chr,y=P.x,color=factor(cluster))) + ylim(0,1)+ 
-  geom_point() + geom_line(alpha=0.1)
-
-ggplot(subset(hgmx, x > 0 & cluster == 9), aes(x=factor(chr),y=P.x,fill=factor(cluster))) + ylim(0,0.5 )+ 
-  geom_bar(colour='grey', stat='identity', position=position_dodge())
-
 ```
 
-So what about using the normalized segmented values?
+```{r hgmx, echo=F, message=F, warning=F, fig.height=6, fig.width=6}
+ggplot(subset(hgmx, x > 0), aes(x=chr,y=P.x,color=factor(cluster))) + ylim(0,0.5)+ 
+  geom_point() + 
+  geom_line(alpha=0.1) + 
+  labs(title='Probability of cluster assignment', x='Chromosome', y='P(x)')
+```
 
-```{r apclustN, echo=T, message=T, warning=F, fig.height=8, fig.align='left'}
+No cluster in any chromosome reached even a 50% probability, so it appears that the regions are reasonably randomly distributed across clusters rather than being made up of all of chromosome 1 or 17 etc.
+
+```{r echo=F, warning=F, message=F}
+ggplot(subset(hgmx, x > 0 & cluster == 9), aes(x=factor(chr),y=P.x,fill=factor(cluster),label=x)) + ylim(0,0.5) + 
+  geom_bar(colour='grey', fill='purple', stat='identity', position=position_dodge()) + 
+  geom_text(aes(y=P.x+0.005), position=position_dodge()) +
+  labs(title='Cluster 9', x='P(x)', y='chromosome')
+```
+
+In the purple cluster are those unique regions to the HGD sample.
+So just as an example, the chromosomes that appear in that cluster are `r paste(subset(hgmx, cluster==9 & x>0)$chr, collapse=', ')` with a total of `r sum(subset(hgmx, cluster==9 & x>0)$x)` regions.
+The probabilities of each of these having shown up are:
+
+`r pander( subset(hgmx, cluster==9 & x>0), caption="Cluster 9 probabilities are below 20%, or essentially random." )`
+
+### Using the normalized values
+
+So what about using the normalized segmented values for AP clustering instead?
+
+```{r apclustN, echo=F, message=F, warning=F, fig.height=5, fig.width=5, fig.align='left'}
 # Should this be performed on the normalized seg values instead?
@@
 
-ggplot(sim3[,c('Status','ss')], aes(y=ss, x=Status, fill=Status)) + geom_boxplot() + geom_jitter(width=0.2) + 
+g1 = ggplot(sim3[,c('Status','ss')], aes(y=ss, x=Status, fill=Status)) + geom_boxplot() + geom_jitter(width=0.2) + 
   labs(title=paste("Norm seg values p-value", signif(ttp, 3)), y='sum similarity')
+
+g2 = ggplot(sim3[,c('Status','clusters')], aes(y=clusters, x=Status, fill=Status)) + geom_boxplot() + geom_jitter(width=0.2) + 
+  labs(title=paste("Norm seg values p-value", signif(ttp, 3)), y='n clusters')
+
+grid.arrange(g1, g2, top='Normalized seg values')
+
 ```
+
+The normalized values show no separation between progressor and non using either the summed similarity measure or the number of clusters. So for now we will use the raw segmentation value matrix.
+
+### Cross-validation, LOO
 
@@
 
-Patients first
+#### Patients first
 
-```{r xval_pt, echo=T, warning=F, message=T, fig.align='left', fig.height=4, eval=F}
+```{r xval_pt, echo=F, warning=F, message=F, fig.align='left', fig.height=4, fig.width=4}
 # Leave out patients
@@
 ggplot(melt(data.frame(loo_xval[,c(3)])), aes(x=variable, y=value, fill=variable)) +  
-    geom_boxplot() + geom_jitter(shape=16, width=0.2) + geom_hline(yintercept=0.05, color='red', alpha=0.5) +
+  geom_boxplot() + geom_jitter(shape=16, width=0.2) + 
+  geom_hline(yintercept=0.05, color='red', alpha=0.5) +
   stat_boxplot(geom ='errorbar') + labs(y='p-values', x='Iterations', title='Leave one (patient) out clustering') +
@@
 
-Then randomly leave out samples
+#### Single samples
 
-```{r xval_smp, echo=F, warning=F, message=T, fig.align='left', fig.height=6, eval=F}
-
+```{r xval_smp, echo=F, warning=F, message=F, fig.align='left', fig.height=4, fig.width=4}
 # Leave out samples?
@@
 
-## (xval) Re run clusters on the Progressors without the HGD sample
+## No HGD samples
+
 There is still a difference between the progressors and non, though the p-value does increase. Some of the progressors do start looking more like NPs in their similarity measure but not all of them.
 
-```{r nofinal, echo=F, message=F, warning=F, fig.height=6, fig.width=10}
+```{r nofinal, echo=F, message=F, warning=F, fig.height=5, fig.width=5}
 
@@
 for (pt in sum.patient.data$Patient) {
-  print(pt)
   x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
@@
 
-t.test(subset(simX, Status == 'NP')$n.var, subset(simX, Status == 'P')$n.var)
-
-#pander(subset(simX, Status == 'P' & ss <= mean(ss)), caption="These patients are still strongly in the Progressor group")
-
-pander(subset(simX, Status == 'P' & clusters >= mean(clusters)), caption="These patients are more similar to Non-Progressors without their HGD sample(s)")
-
 fit = glm(Status~clusters, simX, family=binomial(link='logit') )
-summary(fit)
 
@@
 
-## And why q=0.2
+There do appear to be some patients that are more similar to non-progressors without their HGD samples:
+`r pander(subset(simX, Status == 'P' & clusters <= mean(subset(simX, Status == 'NP')$clusters)), caption="These progressor patients fall below the mean number of clusters in the non-progressor group, suggesting they may be more similar to NP before their HGD samples.")`
+
+## AP cluster parameter selection
+
+Just a quick note on the selection of the q parameter for apcluster.
+The q parameter initializes the exemplar preferences.
+q=0.5 starts at the median, values between 0-1 use the sample quantile with threshold q (per documentation).
+
 ```{r whyq, echo=F, message=F, warning=F}
@@
 for (qv in seq(0,0.9,0.1)) {
-  print(qv)
-
   ## Using the segmentation values
@@
 
-pander(clus[,c('qv','NP','P')], caption="Mean of clusters by patient status per q parameter.")
+```
+
+The number of clusters identified increases with higher q values as would be expected: 0 finds 1 cluster, 1 finds n(regions) clusters.
+We need to see at least 2 clusters (gains/losses), and more than 15 clusters doesn't increase the information available to us right now.
+At every q value there is a separation between progressors and non, but the mean error increases in progressors with the number clusters in progressors as well.
+So we selected a q value to: 1) allow for a minimum of 2 clusters, but no more than 15; and minimize the mean error.
+
+`r pander(clus[,c('qv','NP','P')], caption="Mean of clusters by patient status per q parameter.")`
+
+```{r echo=F, warning=F, message=F, fig.align='left', fig.height=4, fig.width=4}
 
 ggplot(melt(pvals, measure.vars=c('p.clusters'), id.vars='qv'),aes(x=factor(qv),y=value, color=variable)) +
-  geom_point() + 
-  labs(title='P vs NP p-values per q parameters', y='pvalue NP vs P', x='q param')
-
+  geom_point() + geom_hline(yintercept=0.05, alpha=0.3, col='red') + labs(title='P vs NP p-values per q parameters', y='pvalue NP vs P', x='q param')
 ```
 
-```{r loc_pca, echo=F, warning=F, message=T, fig.height=12, fig.width=20, eval=F}
-## Can we identify regions in the progressors?
+As an additional note, the separation between progressors and non is stable at q values between 0.1 and 0.7.
 
-#Not really just from the SVDs.  This isn't any different at the moment from simply calling the variable regions directly.
+# Conclusion
 
-count.occurrences<-function(grs) {
-  ov = findOverlaps(grs, type="within", ignoreSelf=TRUE,select="arbitrary")
-  counts = countOverlaps(grs, type='within', maxgap=0)
-
-  newgrs = grs[is.na(ov)]
-
-  newgrs$overlaps = 0
-  newgrs$patients = ''
-
-  patients = c()  
-  index = 1
-  occurence = vector(mode='numeric')
-  for (x in which(is.na(ov))) {
-    occurence = c(occurence, sum(count[index:x]))
-    patients = c(patients, paste(grs[index:x]$patient, collapse=','))
-    index = x+1
-  }
-  newgrs$overlaps = occurence
-  newgrs$patients = patients
-
-  newgrs$patient.count = unlist(lapply(strsplit(patients, ','), length))
-
-  newgrs$patient = NULL  
-  return(newgrs)
-}
-annotate.cosmic<-function(grs) {
-  ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
-  ensembl = biomaRt::useMart("ENSEMBL_MART_ENSEMBL", host="grch37.ensembl.org")
-  mart = useDataset("hsapiens_gene_ensembl", ensembl)
-
-  grs$Genes = 0
-  grs$Cosmic.count = 0
-  grs$Cosmic.genes = ''
-
-  plocs = as.data.frame(grs)  
-  for (i in 1:nrow(plocs)) {
-    bm = getBM(mart = mart,
-        attributes=c('hgnc_symbol', 'entrezgene' ), filters = c('chromosome_name', 'start', 'end' ),
-        values = with(plocs[i,], list(gsub("chr", "", as.character(seqnames)), start, end)))
-
-    grs[i]$Genes = length(unique(bm$hgnc_symbol))
-    grs[i]$Cosmic.count = length(intersect(bm$entrezgene, ccgenes$Entrez.GeneId))
-    grs[i]$Cosmic.genes = paste(intersect(bm$hgnc_symbol, ccgenes$Gene.Symbol), collapse=',')
-  }
-
-  return(grs)
-}
-
-#pt = 'PR1_HIN_042'
-gg.svd<-function(mat, names, sd=0, n=3) {
-  df = as.data.frame(mat[,(1:n)])
-
-  for (i in 1:ncol(df)) {
-    high = mat[,i] > (mean(mat[,i]) + sd * sd(mat[,i]))
-    low = mat[,i] < (mean(mat[,i]) - sd * sd(mat[,i]))
-
-    label = high | low
-    color=label
-
-    for (j in 1:length(label))
-      label[j] = ifelse(label[j], names[j], '')
-
-    df = cbind(df, label, color)
-    colnames(df)[grep('^label$', colnames(df))] = paste('SV',i,'.label',sep='')
-    colnames(df)[grep('^color$', colnames(df))] = paste('SV',i,'.color',sep='')
-  }
-  df$x = 1:nrow(df)
-  return(df)
-  }
-
-pt.regions = list()
-for (pt in sum.patient.data$Patient) {
-  print(pt)
-  patient.data[[pt]]$info = arrange(patient.data[[pt]]$info, Endoscopy.Year, Pathology)
-
-  x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
-  colnames(x1) = paste(patient.data[[pt]]$info$Endoscopy.Year, '(', patient.data[[pt]]$info$Pathology, ')', sep='' )
-
-  #plot.pca(pt)
-  #pca = prcomp(t(x1), center=T)
-
-  #plot(pca$sdev^2, type='o')
-
-  svd = svd(t(x1))
-  colnames(svd$u) = paste('SV',1:ncol(svd$u), sep='') 
-  colnames(svd$v) = paste('SV',1:ncol(svd$v), sep='')
-
-  #plot(svd$d, type='o', main='Scree')
-  # u matrix
-  samples = gg.svd(svd$u, colnames(x1), 0, 3)
-  grid.arrange(
-    ggplot(samples, aes(y=SV1, x=x, col=SV1.color)) + geom_point() +
-      geom_text(aes(label=SV1.label), vjust=0) + guides(color='none'),
-    ggplot(samples, aes(y=SV2, x=x, col=SV2.color)) + geom_point() +
-      geom_text(aes(label=SV2.label), vjust=0) + guides(color='none'),
-    ggplot(samples, aes(y=SV3, x=x, col=SV3.color)) + geom_point() +
-      geom_text(aes(label=SV3.label), vjust=0) + guides(color='none'),
-    top='SVD Samples')
-
-  # v matrix
-  regions = gg.svd(svd$v, rownames(x1), 2, 3)
-  grid.arrange(
-    ggplot(regions, aes(y=SV1, x=x, col=SV1.color)) + geom_point() +
-      geom_text(aes(label=SV1.label), vjust=0) + guides(color='none'),
-    ggplot(regions, aes(y=SV2, x=x, col=SV2.color)) + geom_point() +
-      geom_text(aes(label=SV2.label), vjust=0) + guides(color='none'),
-    ggplot(regions, aes(y=SV3, x=x, col=SV3.color)) + geom_point() +
-      geom_text(aes(label=SV3.label), vjust=0) + guides(color='none'), 
-    top='SVD Locations')
-
-  locs = as.character(unlist(c(regions[grep('SV1.label',colnames(regions))])))
-  locs = unique(locs[grep("\\w+", locs)])
-
-  high = names(which(apply(x1[locs,], 1, function(x) length(which(x >= 1.1))) > 0))
-  low = names(which(apply(x1[locs,], 1, function(x) length(which(x <= 0.9))) > 0))
-
-  hls = as.data.frame(do.call(rbind, strsplit(c(high, low), '[p|q]\\.|-')))
-  colnames(hls) = c('chr','start','stop')
-  hls[c('start','stop')] = lapply(hls[c('start','stop')], function(x) as.numeric(as.character(x)))
-
-  #ac = apcluster(negDistMat(x1, r=2), q=0.2)
-
-  #lapply(ac@clusters, function(x) names(x)[which(names(x) %in% high)])
-  #lapply(ac@clusters, function(x) names(x)[which(names(x) %in% low)])
-
-  pt.regions[[pt]] = data.frame(hls, cn=c(rep('gain', length(high)), rep('loss',length(low))))
-}
-pt.regions = do.call(rbind, pt.regions)
-
-pt.regions$patient = sub('\\..*', '', rownames(pt.regions) )
-rownames(pt.regions) = 1:nrow(pt.regions)
-
-prog.grs = sort(makeGRangesFromDataFrame(subset(pt.regions, patient %in% subset(sum.patient.data, Status == 'P')$Patient), keep.extra.columns = T, seqnames.field = 'chr', seqinfo=c(1:22,'X','Y')))
-
-np.grs = sort(makeGRangesFromDataFrame(subset(pt.regions, patient %in% subset(sum.patient.data, Status == 'NP')$Patient), keep.extra.columns = T, seqnames.field = 'chr', seqinfo=c(1:22,'X','Y')))
-
-prog.loss = count.occurrences(subset(prog.grs, grepl('loss',cn)))
-prog.gain = count.occurrences(subset(prog.grs, grepl('gain',cn)))
-
-np.loss = count.occurrences(subset(np.grs, grepl('loss',cn)))
-np.gain = count.occurrences(subset(np.grs, grepl('gain',cn)))
-
-pg.c = annotate.cosmic(prog.gain)
-pl.c = annotate.cosmic(prog.loss)
-
-npg.c = annotate.cosmic(np.gain)
-npl.c = annotate.cosmic(np.loss)
-
-with(pg.c,plot(overlaps, Cosmic.count, col='blue', pch=19))
-with(pl.c, points(overlaps, Cosmic.count, col='red', pch=19))
-
-with(npg.c, points(overlaps, Cosmic.count, col='green', pch=18))
-with(npl.c, points(overlaps, Cosmic.count, col='purple', pch=17))
-
-pander(data.frame('total.p' = sum(length(pg.c), length(pl.c)), 'total.np' = sum(length(npg.c), length(npl.c))), caption="Total regions called as gains/losses for progressors vs non")
-
-t.test( width(pg.c), width(npg.c) ) # Yes for the gains
-t.test( width(pl.c), width(npl.c) ) # Not for the losses
-
-```
-
-```{r model, echo=F, warning=F, message=F, fig.height=12, fig.width=20, eval=F}
-
-t.test(subset(sim, Status == 'P')$clusters, subset(sim, Status == 'NP')$clusters)
-
-t.test(subset(simX, Status == 'P')$clusters, subset(simX, Status == 'NP')$clusters)
-
-## TODO  Need to consider segment lengths relative to chromosomal lengths
-
-var.lengths = as.data.frame( do.call(rbind, lapply(patient.data, function(x){
-  if (nrow(x[['variable.regions']]) > 0) {
-    lengths = as.vector( (x[['variable.regions']]['end.pos'] - x[['variable.regions']]['start.pos'])[,1] )
-    return(c(median(lengths), mean(lengths), sd(lengths), sd(lengths)/length(lengths)))
-  }
-  return(NA)
-})) )
-colnames(var.lengths) = c('median','mean','sd','se')
-
-var.lengths = merge(sum.patient.data, var.lengths, by.x='Patient', by.y='row.names')
-
-var.lengths
-wilcox.test(subset(var.lengths, Status == 'P')$mean, subset(var.lengths, Status == 'NP')$mean)
-
-```
+At the moment AP clustering appears to give us a reliable "diversity" measure between progressors and non.
+It also suggests greater complexity in the progressor patients which we would expect based on earlier analyses.
+The exemplars need to be further analyzed to see how they may relate across patients, but also exactly what they represent within a single patient.
+Finally, it may be worth assuming absolute copy number and discretizing the matrix to test it.

### 2017-04-04 Clustering.Rmd
Adding new analyses

@@
 
-getPrecision<-function(actual,predicted) {
-  retrieved <- sum(predicted==1)
-  precision <- sum(predicted==1 & actual==1) / retrieved
-  precision
-}
-getRecall<-function(actual,predicted) {
-  retrieved <- sum(predicted==1)
-  recall <- sum(predicted==1 & actual==1) / sum(actual==1)
-  recall
-}
-
 data = '~/Data/Ellie'
@@
 patient.info = read.patient.info(patient.file)
-
 patient.info$Patient = gsub("/", "_", patient.info$Patient)
@@
 sum.patient.data = summarise.patient.info(patient.info)
-sum.patient.data = as.data.frame(subset(sum.patient.data, Initial.Analysis)) ## For now
+sum.patient.data = as.data.frame(subset(sum.patient.data, Patient %in% names(patient.data))) ## For now
 pander(sum.patient.data)
+
+# Missing some of the samples as they aren't all sequenced yet
+for (pt in names(patient.data)) {
+  patient.data[[pt]]$info= patient.data[[pt]]$info[patient.data[[pt]]$info$Samplename %in% colnames(patient.data[[pt]]$seg.vals)[-(1:5)],]
+}
+
 ```
@@
 plot.pca<-function(pt) {
+  print(pt)
+  info = arrange(patient.data[[pt]]$info, Endoscopy.Year)
+  info = info[which(info$Samplename %in% colnames(patient.data[[pt]][['seg.vals']])[-(1:5)]),]
+  plots = list()
+  for (sv in c('seg.vals')) {
 
-  info = arrange(patient.data[[pt]]$info, Endoscopy.Year)
-
-  plots = list()
-  for (sv in c('seg.vals','norm.seg.vals')) {
-
-    x1 = patient.data[[pt]][[sv]][, patient.data[[pt]]$info$Samplename]
-    x1 = x1[,patient.data[[pt]]$info$Samplename]
-    colnames(x1) = paste(patient.data[[pt]]$info$Endoscopy.Year, '(', patient.data[[pt]]$info$Pathology, ')', sep='' )
+    x1 = patient.data[[pt]][[sv]][, info$Samplename]
+    x1 = x1[,info$Samplename]
+    colnames(x1) = paste(info$Endoscopy.Year, '(', info$Pathology, ')', sep='' )
 
@@
 pplots = lapply(subset(sum.patient.data, Status == 'P')$Patient, plot.pca)
-npplots = lapply(subset(sum.patient.data, Status == 'NP')$Patient, plot.pca)
+#npplots = lapply(subset(sum.patient.data, Status == 'NP')$Patient, plot.pca)
 ```
@@
 grid.arrange(grobs=pplots, nrow=length(subset(sum.patient.data, Status == 'P')$Patient), top="Progressors")
-grid.arrange(grobs=npplots, nrow=length(subset(sum.patient.data, Status == 'NP')$Patient), top="Non-Progressors")
+#grid.arrange(grobs=npplots, nrow=length(subset(sum.patient.data, Status == 'NP')$Patient), top="Non-Progressors")
+```
+
+```{r persample, eval=F}
+
+lapply(patient.data, function(df) {
+
+  lapply(df$seg.vals[,-(1:5)], quantile)
+
+})
+
+par(mfrow=c(2,2))
+  lapply(patient.data$AHM0896$seg.vals[,-(1:5)], hist)
+
+summary(sapply(patient.data$AHM0896$seg.vals[,-(1:5)], mean))
+lapply(lapply(patient.data$AHM0896$seg.vals[,-(1:5)], kmeans, 4), function(fit) fit$centers)
+
+km.fits = lapply(patient.data, function(df) kmeans(df$seg.vals[,-(1:5)], 4))
+autoplot(km.fits$AHM0896, patient.data$AHM0896$seg.vals[,-(1:5)])
+
+plot(patient.data$AHM0896$seg.vals[,-(1:5)], col=km.fits$AHM0896$cluster, pch=19)
+
+  ## TODO find an OCCAMS patient with genome doubling, downsample and compare
+
 ```
@@
 #### EM
-```{r clust, echo=F, message=F, warning=F, fig.height=4, fig.width=4}
-
+```{r clust, echo=F, message=F, warning=F, fig.height=4, fig.width=4, eval=F}
 em.dr<-function(patient, status, x1, var.regions) {
@@
   for (g in clusters) {
+    print(g)
     model = densityMclust(x1, G=g)
     highCluster = which.max(colSums(model$parameters$mean))
-
-    tryCatch({
-      mod1dr = MclustDR(model, lambda = 0.5)
-
-      #     par(mfrow=c(2,1)) 
-      plot(mod1dr, what="evalues")
-      title(paste(pt, '(',status,') G=', g, sep=''))
-      plot(mod1dr, what='classification') # while this looks like PCA, it's not quite
-      title(paste(pt, '(',status,') G=', g, sep=''))
-      #     dev.off()
-      plot(mod1dr, what='density', dimens=1)
-      title(paste(pt, '  G=', g, sep=''))
-    }, error = function(e) {
-      cat(paste(pt, g, "failed DR"))
-    })
-
+    plot(model, what='density')
+    # tryCatch({
+    #   mod1dr = MclustDR(model, lambda = 0.5)
+    #   
+    #   #     par(mfrow=c(2,1)) 
+    #   plot(mod1dr, what="evalues")
+    #   title(paste(pt, '(',status,') G=', g, sep=''))
+    #   plot(mod1dr, what='classification') # while this looks like PCA, it's not quite
+    #   title(paste(pt, '(',status,') G=', g, sep=''))
+    #   #     dev.off()
+    #   plot(mod1dr, what='density', dimens=1)
+    #   title(paste(pt, '  G=', g, sep=''))
+    # }, error = function(e) {
+    #   cat(paste(pt, g, "failed DR"))
+    # })
+    # 
     # a1 = rownames(x1[which(model$classification == highCluster),] )
@@
 for (pt in sum.patient.data$Patient) {
+  print(pt)
+
     status = subset(sum.patient.data, Patient == pt)$Status
@@
 
-```{r kmeansraw, echo=F, message=F, warning=F, eval=F, fig.height=6, fig.width=6} 
+```{r kmeansraw, echo=F, message=F, warning=F, eval=F, fig.height=6, fig.width=6, eval=F} 
 for (pt in sum.patient.data$Patient) {
@@
 
-```{r clust2, echo=F, message=F, warning=F, fig.height=4, fig.width=4}
+```{r clust2, echo=F, message=F, warning=F, fig.height=4, fig.width=4, eval=F}
 llk = as.data.frame(matrix(nrow=nrow(sum.patient.data),ncol=3,dimnames=list(c(sum.patient.data$Patient), c(2:4))))
@@
 llk.diff = apply(llk, 1, function(x) diff(range(x)) )
-#wilcox.test(llk.diff[subset(sum.patient.data, Status == 'NP')$Patient], llk.diff[subset(sum.patient.data, Status == 'P')$Patient])
+wilcox.test(llk.diff[subset(sum.patient.data, Status == 'NP')$Patient], llk.diff[subset(sum.patient.data, Status == 'P')$Patient])
 
@@
 #### Kmeans
-```{r kmeansnorm, echo=F, message=F, warning=F, fig.height=6, fig.width=6} 
+```{r kmeansnorm, echo=F, message=F, warning=F, fig.height=6, fig.width=6, eval=F} 
 for (pt in sum.patient.data$Patient) {
@@
   x1 = segdata[,intersect(colnames(segdata), samples)]
-  x1 = x1[, samples]
+  #x1 = x1[, samples]
   rownames(x1) = (segdata[,c(1:4)] %>%
@@
 ```{r apclust, echo=F, message=F, warning=F, fig.height=10, fig.width=10, fig.align='left'}
+sample.vals = list()
 q=0.2
@@
 sum.patient.data$years = with(sum.patient.data, paste(start.year, end.year, sep='-'))
-sim = sum.patient.data[,c('Status','years')]
-rownames(sim) = sum.patient.data[,'Patient']
+sim = as.data.frame(sum.patient.data[,c('Status','years')])
+rownames(sim) = sum.patient.data$Patient
 for (pt in sum.patient.data$Patient) {
-  patient.data[[pt]]$info = arrange(patient.data[[pt]]$info, Endoscopy.Year, desc(Pathology))
+  patient.data[[pt]]$info = arrange(patient.data[[pt]]$info,  Pathology, Endoscopy.Year)
+  x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
 
-  x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
-  colnames(x1) = paste(patient.data[[pt]]$info$Endoscopy.Year, '(', patient.data[[pt]]$info$Pathology, ')', sep='' )
+  gg = ggplot(melt(x1), aes(x=variable, y=value, group=variable, fill=variable)) + 
+    geom_boxplot(notch=T, show.legend=F) +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position='none') +
+    labs(title=paste(pt, patient.data[[pt]]$info$Status[1], sep=' - ')) +   theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank())
+
+  sample.vals[[pt]] = gg
+
+  info = patient.data[[pt]]$info[which(colnames(x1) %in% patient.data[[pt]]$info$Samplename),]
+  colnames(x1) = paste(info$Endoscopy.Year, '(', info$Pathology, ')', sep='' )
 
@@
   last = ncol(x1)
-  if (patient.data[[pt]]$info$Status[1] == 'P' & pt != 'AD0252')
-    last = sort(grep('HGD|IMC', patient.data[[pt]]$info$Pathology), decreasing=T)[1]
+  if (patient.data[[pt]]$info$Status[1] == 'P' & length(which(grepl('HGD|IMC', info$Pathology)) > 0)  )
+    last = sort(grep('HGD|IMC', info$Pathology), decreasing=T)[1]
   if (ncol(x1) <= 15) {
-    plot(ac, x1[,c(c(1:ncol(x1))[-last], last)], main=pt)
+    #plot(ac, x1[,c(c(1:ncol(x1))[-last], last)], main=pt)
   } else {
-    plot(ac, x1[,c(1, sort(sample(c(1:(ncol(x1)-1))[c(-1,-last)], 8)) ,last) ], main=pt)
+    #plot(ac, x1[,c(1, sort(sample(c(1:(ncol(x1)-1))[c(-1,-last)], 8)) ,last) ], main=pt)
   }
@@
 
+do.call(grid.arrange, c(sample.vals))
+
 # May be a difference
-ttps = t.test( subset(sim, Status == 'NP')$ss, subset(sim, Status == 'P')$ss )$p.value
-ttpn = t.test( subset(sim, Status == 'NP')$ns, subset(sim, Status == 'P')$ns )$p.value
-ttpc = t.test( subset(sim, Status == 'NP')$clusters, subset(sim, Status == 'P')$clusters )$p.value
+ttps = wilcox.test( subset(sim, Status == 'NP')$ss, subset(sim, Status == 'P')$ss )$p.value
+ttpc = wilcox.test( subset(sim, Status == 'NP')$clusters, subset(sim, Status == 'P')$clusters )$p.value
 
@@
   labs(title=paste("p-value", signif(ttpc, 3)), y='num. clusters')
+
+sim$Status.Int = as.integer(sim$Status)-1
+chisq.test(sim[,c('Status.Int','clusters')])
 ```
@@
 ```{r apglmclus, echo=F, message=F, warning=F}
-fit = glm(Status~clusters, sim, family=binomial(link='logit'))
-#pander(fit, justify='left')
+precisionRecall<-function(actual, predicted) {
+    retrieved <- sum(predicted==1)
+    recall <- sum(predicted==1 & actual==1) / sum(actual==1)
+    precision <- sum(predicted==1 & actual==1) / retrieved
+    #f1 = signif(2*(1/(1/precision+1/recall)), 2)
+    f1 =signif(2*precision*recall/(precision+recall), 2)
+    return(data.frame('precision'=precision, 'recall'=recall, 'F1'=signif(f1, 2)))
+  }
 
 sim = sim[, c('Status', 'clusters', 'ns','ss')]
-actualMatrix = matrix(0,1000,2,dimnames=list(c(), c('prec','recall')))
-for (i in 1:nrow(actualMatrix)) {
-  trainingRows = sample(1:nrow(sim), nrow(sim)-5)
+actualMatrix = as.data.frame(matrix(0,0,4,dimnames=list(c(), c('status','precision','recall', 'F1'))))
+for (i in 1:100) {
+  trainingRows = sample(1:nrow(sim), nrow(sim)-10)
   train = sim[trainingRows,]
@@
 
-  actual = test$Status == 'P'
-  predicted = pred>0
+  #glm_link_scores <- predict(fit, test, type="link")
+  #glm_response_scores <- predict(fit, test, type="response")
 
-  actualMatrix[i,'prec'] = getPrecision( actual, predicted )
-  actualMatrix[i, 'recall'] = getRecall( actual, predicted )
+  actualMatrix = rbind(actualMatrix, cbind('status'='P', precisionRecall(test$Status == 'P', pred>0)))
+  actualMatrix = rbind(actualMatrix, cbind('status'='NP', precisionRecall(test$Status == 'NP', pred<0)))
 }
 
-# ggplot(as.data.frame(actualMatrix), aes(x=recall, y=prec)) + geom_point() +
-#   geom_hline(yintercept = 0.5, col='red', alpha=0.5) +
-#   labs(title='Precision/Recall for Progressor/NP prediction')
+g1 = ggplot(actualMatrix, aes(x=recall, y=precision, color=status, label=F1)) + 
+  geom_point() + geom_text(aes(label=ifelse(precision>0.5 & recall>0.5, as.character(F1), ''), hjust=0, vjust=0)) +
+   geom_hline(yintercept = 0.5, col='grey') + geom_vline(xintercept = 0.5, col='grey') +
+   labs(title='Precision/Recall for Progressor/NP prediction')
+
+g2 = ggplot(actualMatrix, aes(x=status, y=F1, group=status, fill=status)) + 
+  geom_violin() + geom_jitter(alpha=0.5) + ylim(0,1) + 
+  geom_violin(data=subset(actualMatrix,precision>0.5 & recall>0.5), aes(x=status, y=F1, group=status, fill='P&R>0.5') , alpha=0.6) +
+  labs(title="F1 statistic")
+
+grid.arrange(g1, g2, top='100 iterations 10f xval')
+```
+
+Randomize the data and see if the F1 statistic still holds
+
+```{r gpglmclus_rand, echo=F, message=F, warning=F}
+sim = sim[, c('Status', 'clusters')]
+
+rand = as.data.frame(matrix(0,0,4,dimnames=list(c(), c('status','precision','recall', 'F1'))))
+for (i in 1:100) {
+  # rand.sim = as.data.frame(cbind(  'Status'=sample(c('P','NP'), nrow(sim), replace=T, prob=round(table(sim$Status)/nrow(sim), 2)),
+  #                    'clusters'=sample((3:30), nrow(sim), replace=T) ))
+  # rand.sim = as.data.frame(cbind(  'Status'=sample(c('P','NP'), nrow(sim), replace=T, prob=round(table(sim$Status)/nrow(sim), 2)),
+  #                    'clusters'=sample((3:30), nrow(sim), replace=T) ))
+
+  rand.sim = as.data.frame(cbind(  'Status'=sample(c('P','NP'), nrow(sim), replace=T),
+                     'clusters'=sim$clusters ))
+
+  rand.sim$clusters = as.integer(as.character(rand.sim$clusters))
+
+  trainingRows = sample(1:nrow(rand.sim), nrow(rand.sim)-10)
+  train = rand.sim[trainingRows,]
+  test = rand.sim[-trainingRows,]
+
+  fit = glm( (as.integer(Status)-1)~clusters, train, family=binomial(link='logit'))
+  pred = predict(fit, test, se.fit=F) # probabilities -- of which outcome though?
+
+  rand = rbind(rand, cbind('status'='P', precisionRecall(test$Status == 'P', pred>0)))
+  rand = rbind(rand, cbind('status'='NP', precisionRecall(test$Status == 'NP', pred<0)))
+}
+
+g1 = ggplot(rand, aes(x=recall, y=precision, color=status, label=F1)) + 
+  geom_point() + geom_text(aes(label=ifelse(precision>0.5 & recall>0.5, as.character(F1), ''), hjust=0, vjust=0)) +
+   geom_hline(yintercept = 0.5, col='grey') + geom_vline(xintercept = 0.5, col='grey') +
+   labs(title='Precision/Recall for Progressor/NP prediction')
+
+g2 = ggplot(rand, aes(x=status, y=F1, group=status, fill=status)) + 
+  geom_violin() + geom_jitter(alpha=0.5) + ylim(0,1) + 
+  #geom_violin(data=subset(actualMatrix,precision>0.5 & recall>0.5), aes(x=status, y=F1, group=status, fill='P&R>0.5') , alpha=0.6) +
+  labs(title="F1 statistic")
+
+grid.arrange(g1, g2, top='100 iterations 10f xval - randomized labels')
 
@@
 
-cols = c('chr','cluster','P.x', 'x')
+cols = c('chr','cluster', 'P.x','x')
 hgmx = matrix(nrow=0,ncol=length(cols),dimnames=list(c(), cols))
@@
 ```{r hgmx, echo=F, message=F, warning=F, fig.height=6, fig.width=6}
-ggplot(subset(hgmx, x > 0), aes(x=chr,y=P.x,color=factor(cluster))) + ylim(0,0.5)+ 
-  geom_point() + 
-  geom_line(alpha=0.1) + 
+
+g1 = ggplot(subset(hgmx, x > 0), aes(x=factor(chr),y=P.x,group=chr,fill=factor(chr))) + ylim(0,0.5)+ 
+  geom_boxplot() +
   labs(title='Probability of cluster assignment', x='Chromosome', y='P(x)')
+
+g2 = ggplot(subset(hgmx, x > 0), aes(x=factor(cluster),y=P.x,group=cluster,fill=factor(cluster))) + ylim(0,0.5)+ 
+  geom_boxplot() +
+  labs(title='Probability of cluster assignment', x='Cluster', y='P(x)')
+
+grid.arrange(g1, g2, nrow=2)
 ```
@@
 
+### Variable Regions
+
+Should we call differently?  One is regions that are not diploid, the second would be what changes between samples?
+
+```{r varregions, echo=F, message=F, warning=F, eval=F}
+pt = 'AHM0896'
+
+se = round(apply(patient.data[[pt]]$seg.vals[,-(1:5)], 2, function(x) sd(x)/sqrt(length(x))), 3)
+sdx = round(apply(patient.data[[pt]]$seg.vals[,-(1:5)], 2, sd), 3)
+
+row=325
+patient.data[[pt]]$seg.vals[row,]
+round(patient.data[[pt]]$seg.vals[,-(1:5)][row,] + sdx, 1)
+round(patient.data[[pt]]$seg.vals[,-(1:5)][row,] - sdx, 1)
+
+lapply(patient.data, function(df) nrow(df$variable.regions))
+
+```
+
 ### Cross-validation, LOO
@@
 
-```{r xval_pt, echo=F, warning=F, message=F, fig.align='left', fig.height=4, fig.width=4}
+```{r xval_pt, echo=F, warning=F, message=F, fig.align='left', fig.height=4, fig.width=4, eval=F}
 # Leave out patients
@@
   for (pt in sum.patient.data$Patient) {
+      if (length(patient.data[[pt]]$info$Samplename) <= 2) next
       x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename[-sample(1:(length(patient.data[[pt]]$info$Samplename)), 1)])
@@
     final = grep('HGD|IMC', patient.data[[pt]]$info$Pathology)
+
+    if (length(patient.data[[pt]]$info$Samplename[-final]) <= 1) next
+
     x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename[-final])
@@
 
-ttpc = t.test( subset(simX, Status == 'NP')$clusters, subset(simX, Status == 'P')$clusters )$p.value
+ct = cor.test(simX$clusters, simX$n.var)
+
+ttpc = t.test( subset(simX, Status == 'NP')$clusters, 
+               subset(simX, Status == 'P')$clusters )$p.value
+
 ggplot(melt(data.frame(simX[c('Status','clusters')])), aes(x=Status, y=value, fill=Status)) +  
@@
 
-fit = glm(Status~clusters, simX, family=binomial(link='logit') )
+#fit = glm(Status~clusters, simX, family=binomial(link='logit') )
 
 simX = simX[, c('Status', 'clusters', 'ns','ss','n.var')]
-noHGDPR = as.data.frame(matrix(0,1000,2,dimnames=list(c(), c('prec','recall'))))
-for (i in 1:nrow(actualMatrix)) {
-  trainingRows = sample(1:nrow(simX), nrow(simX)-5)
+noHGDPR = as.data.frame(matrix(0,0,4,dimnames=list(c(), c('status','precision','recall', 'F1'))))
+for (i in 1:100) {
+  trainingRows = sample(1:nrow(simX), nrow(simX)-10)
   train = simX[trainingRows,]
@@
 
-  actual = test$Status == 'P'
-  predicted = pred>0
+  noHGDPR = rbind(noHGDPR, cbind('status'='P', precisionRecall(test$Status == 'P', pred>0)))
+  noHGDPR = rbind(noHGDPR, cbind('status'='NP', precisionRecall(test$Status == 'NP', pred<0)))
 
-  noHGDPR[i,'prec'] = getPrecision( actual, predicted )
-  noHGDPR[i, 'recall'] = getRecall( actual, predicted )
 }
 
-#ggplot(as.data.frame(noHGDPR), aes(x=recall, y=prec)) + geom_point() + 
-#  geom_hline(yintercept = 0.5, col='red', alpha=0.5) +
-#  geom_vline(xintercept = 0.5, col='red', alpha=0.5) +
-#  labs(title='Precision/Recall for Progressor/NP prediction NO HGD')
+g1 = ggplot(noHGDPR, aes(x=recall, y=precision, color=status, label=F1)) + 
+  geom_point() + geom_text(aes(label=ifelse(precision>0.5 & recall>0.5, as.character(F1), ''), hjust=0, vjust=0)) +
+   geom_hline(yintercept = 0.5, col='grey') + geom_vline(xintercept = 0.5, col='grey') +
+   labs(title='Precision/Recall for Progressor/NP prediction without HGD')
+
+g2 = ggplot(subset(actualMatrix, precision>0.5 & recall>0.5), aes(x=status, y=F1, group=status, fill=status)) + 
+  geom_violin() + geom_jitter(alpha=0.5) + ylim(0.5,1) + labs(title="F1 statistic at prec+recall>0.5")
+
+grid.arrange(g1, g2, top='100 iterations 10f xval no HGD')
 
@@
 
-Currently, without the HGD/IMC samples we can predict progressors `r nrow(subset(noHGDPR, prec > 0.5))/nrow(noHGDPR)*100`% of the time.
-In comparison, including HGD/IMC samples our predictions improve to `r length(which(actualMatrix[,'prec'] > 0.5))/nrow(actualMatrix)*100`%.
+Currently, without the HGD/IMC samples we can predict progressors `r nrow(subset(noHGDPR, precision > 0.5))/nrow(noHGDPR)*100`% of the time.
+In comparison, including HGD/IMC samples our predictions improve to `r length(which(actualMatrix[,'precision'] > 0.5))/nrow(actualMatrix)*100`%.
+It's also worth noting that the number of clusters has a significant (pval=`r signif(ct$p.value, 1)`) correlation to the number of variant regions identified across all samples for that patient (r=`r round(ct$estimate, 2)`).
 
@@
 `r pander(subset(simX, Status == 'P' & clusters <= mean(subset(simX, Status == 'NP')$clusters)), caption="These progressor patients fall below the mean number of clusters in the non-progressor group, suggesting they may be more similar to NP before their HGD samples.")`
+
+```{r}
+
+pts = rownames(subset(simX, Status == 'P' & clusters <= mean(subset(simX, Status == 'NP')$clusters)))
+
+subset(sum.patient.data, Patient %in% pts)
+
+```
 
@@
 
+## Investigating Cutoff selection
+
+Currently we use the same cutoffs across all patients and samples for selecting variable regions.
+While it's clear based on the data that values ~1 are the average ploidy of the sample and are most likely to be diploid, some samples have a greater spread round 1 resulting in the possibility that we are both under and over calling variable regions.
+
+In these plots there are several things occurring.
+The values have been log-10 transformed in order to spread them out.
+These are then plotted in a histogram with two sets of lines representing potential cutoffs: one for EM derived means, the other for mean(+|-)sd*1.5.
+In most cases EM does not provide much separation.
+This could potentially be improved by adding priors or other parameters to the model, but it's unclear how to initialize these usefully.
+
+```{r varcall, echo=F, message=F, warning=F, fig.height=10, fig.width=10}
+for (pt in names(patient.data)) {
+  cols = intersect(arrange(patient.data[[pt]]$info, Pathology, Endoscopy.Year)$Samplename, colnames(patient.data[[pt]]$seg.vals[,-(1:5)]))
+  segvals = patient.data[[pt]]$seg.vals[,cols]
+
+  plot.seg.vals<-function(lt, title, hist.col='blue') {
+    sdx = apply(lt, 2, sd)
+    se = apply(lt, 2, function(x) sd(x)/sqrt(length(x)))
+    means = t(do.call(rbind, lapply(lapply(lt, densityMclust, G=3), function(x) sort(x$parameters$mean))))
+    m = merge(melt(lt), melt(means), by.x='variable', by.y='Var2')
+    gg = ggplot(m,aes(value.x, ..density..)) + 
+      facet_wrap(~variable,scales = "free_x") + 
+        geom_histogram(bins=10, fill=hist.col, color='grey') +
+        geom_density(aes(value.x), color='darkgrey', linetype='dotted') +
+        geom_vline( aes(xintercept=value.y, col='EM'), linetype='twodash') +
+        geom_vline(aes(xintercept = mean(value.x)+sd(value.x)*1.5, color='StdDev'), linetype='twodash') +
+        geom_vline(aes(xintercept = mean(value.x)-sd(value.x)*1.5, color='StdDev'),linetype='twodash') +
+        labs(title=title)
+    return(gg)
+  }
+
+  gg1 = plot.seg.vals(segvals, title=paste(pt, ' (',patient.data[[pt]]$info$Status[1],')', sep=''))
+  gg2 = plot.seg.vals(log2(segvals), title=paste(pt, ' log2  (',patient.data[[pt]]$info$Status[1],')', sep=''), hist.col='darkgreen')
+  gg3 = plot.seg.vals(log10(segvals), title=paste(pt, ' log10  (',patient.data[[pt]]$info$Status[1],')', sep=''), hist.col='darkred')
+
+  print(gg1)
+  print(gg2)
+  print(gg3)
+  #do.call(grid.arrange, list(gg1, gg2, gg3))
+
+}
+```
+
 # Conclusion

### 2017-04-04 Patient Report.Rmd
Adding new analyses

@@
 data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
-plot.dir = paste(data, 'Analysis/multipcf_plots_fitted_perPatient-OLD', sep='/')
+plot.dir = paste(data, 'Analysis/multipcf_plots_fitted_perPatient', sep='/')
 
@@
 sum.patient.data = summarise.patient.info(patient.info)
-sum.patient.data = subset(sum.patient.data, Initial.Analysis)
 
-patient.info = subset(patient.info, Patient %in% sum.patient.data$Patient & Batch.Name %in% levels(patient.info$Batch.Name)[1:5])
+#patient.info = subset(patient.info, Patient %in% sum.patient.data$Patient & Batch.Name %in% levels(patient.info$Batch.Name)[1:5])
 
@@
   segvals = read.table(paste(patient.plot.dir, "/",patient.name,"_segmentedCoverage_fitted_gamma",gamma2,".txt",sep=""),sep="\t",stringsAsFactors=F,header=T)
+
+  missing = c(setdiff(colnames(segvals)[-(1:5)], subset(patient.info, Patient == patient.name)$Samplename),
+                    setdiff(subset(patient.info, Patient == patient.name)$Samplename, colnames(segvals)[-(1:5)]))
+
+  if (length(missing) > 0) 
+    warning(paste("Missing", length(missing), "sample(s) in patient", patient.name))
 
@@
         patient = subset(patient.info, Patient == patient.name)
-        patient = arrange(patient, Endoscopy.Year, Pathology)
+        patient = arrange(patient, Endoscopy.Year, Pathology, Samplename)
         }
@@
 }
+
+sum.patient.data = subset(sum.patient.data, Patient %in% names(patient.data))
+
+## Check that all of the samples are there
+patient.info = arrange(patient.info, Status, Patient, Endoscopy.Year,Samplename)
+pt = 'AH0254'
+info = subset(patient.info, Patient == 'AH0254')
+lapply(names(patient.data), function(pt) {
+  missing = which(!patient.data[[pt]]$info$Samplename %in% colnames(patient.data[[pt]]$seg.vals))
+  if (length(missing) > 0) {
+    msg = paste(pt,": Missing samples - ", paste(patient.data[[pt]]$info$Samplename[missing], collapse=', ') , sep='')
+    msg = paste(msg, "\nFrom batch: ", paste(unique(patient.data[[pt]]$info$Batch.Name[missing]), collapse=', '))
+    warning(msg)
+  }
+})
+
+save(patient.data, file='patients.Rdata')
 
@@
 
-The result looks like is two data matrices:
+The result looks like two data matrices:
 
@@
 ```{r general, warning=F, echo=F, message=F}
-gen = do.call(rbind, lapply(patient.data, function(x) nrow(x[['seg.vals']])))
-colnames(gen) = c('Number.Regions')
-
-sum.patient.data = subset(sum.patient.data, Patient %in% names(patient.data))
+gen = do.call(rbind, lapply(patient.data, function(df) {
+  cbind('n.segments'=nrow(df[['seg.vals']]), 
+        'median.seg.len'=median(df$seg.vals[,c('end.pos')] - df$seg.vals[,c('start.pos')]),
+        'mean.seg.len'=mean(df$seg.vals[,c('end.pos')] - df$seg.vals[,c('start.pos')]), 
+        'sd.seg.len'=sd(df$seg.vals[,c('end.pos')] - df$seg.vals[,c('start.pos')]))
+}))
+rownames(gen)  = names(patient.data)
 
@@
 
-pander(sum.patient.data[,c('Patient','Status','years','total.samples','highest.path','Number.Regions')], justify='left')
+pander(sum.patient.data[,c('Patient','Status','years','total.samples','highest.path','n.segments', 'median.seg.len')], justify='left')
+
+#seg.lens = lapply(patient.data, function(df) df$seg.vals[,c('end.pos')] - df$seg.vals[,c('start.pos')] )
+
+#wilcox.test(log(unlist(seg.lens[subset(sum.patient.data, Status == 'P')$Patient])),
+#log(unlist(seg.lens[subset(sum.patient.data, Status == 'NP')$Patient])))
+
+#hist(log(unlist(seg.lens[subset(sum.patient.data, Status == 'P')$Patient])))
+#hist(log(unlist(seg.lens[subset(sum.patient.data, Status == 'NP')$Patient])))
+
+ggplot(sum.patient.data) + geom_boxplot( aes(y=median.seg.len, x=Status, group=Status, fill=Status) ) +
+  labs(title="Median segment length by patient status")
+
 ```
 
+There's no significant difference between the length of segments in progressors vs non.  There's a larger spread of lengths in progressors, but the means are nearly identical.
+
 # Segmented Value Distributions
-```{r echo=F, message=F, warning=F, fig.height=15, fig.width=10}
+```{r echo=F, message=F, warning=F, fig.height=15, fig.width=10, eval=F}
 
@@
 ```{r var_reg, echo=F, message=F, warning=F}
-rm(prog.regions, np.regions)
-for (patient.name in sum.patient.data$Patient) {
-  patient = patient.data[[patient.name]]
+prog.pt = subset(sum.patient.data, Status == 'P')$Patient
+np.pt = subset(sum.patient.data, Status == 'NP')$Patient
 
-  sum.patient.data[which(sum.patient.data$Patient == patient.name), 'var.regions'] = nrow(patient$variable.regions)
+var.seg.lengths = lapply(patient.data, function(df) {
+  cbind('chr'=df$variable.regions$chrom, 'seg.len'=df$variable.regions$end.pos - df$variable.regions$start.pos)
+})
 
-  if (subset(sum.patient.data, Patient == patient.name)$Status == 'P') {
-    if (!exists('prog.regions')) {
-      prog.regions = patient$variable.regions[,c('chrom','start.pos','end.pos')]
-    } else {
-      prog.regions = rbind(prog.regions, patient$variable.regions[,c('chrom','start.pos','end.pos')])
-    }
-  } else {
-    if (!exists('np.regions')) {
-      np.regions = patient$variable.regions[,c('chrom','start.pos','end.pos')]
-    } else {
-      np.regions = rbind(np.regions, patient$variable.regions[,c('chrom','start.pos','end.pos')])
-    }
-  }
-}
+segment.lengths = lapply(patient.data, function(df) {
+  cbind('chr'=df$seg.vals$chrom, 'seg.len'=df$seg.vals$end.pos - df$seg.vals$start.pos)
+})
 
-pander(sum.patient.data %>% group_by(Status) %>% summarise(mean.num.regions = mean(var.regions), sd.num.regions = sd(var.regions)), justify='left')
+prog.regions = sapply(var.seg.lengths[sum.patient.data$Patient], nrow)
 
-prog.regions$length = prog.regions$end.pos - prog.regions$start.pos
-np.regions$length = np.regions$end.pos - np.regions$start.pos
+sum.patient.data$Num.Var.Regions = as.vector(sapply(var.seg.lengths[sum.patient.data$Patient], nrow))
 
 chrs = read.table('hg19_chrom.txt',sep='\t', header=T)
-prog.regions = merge(prog.regions, chrs, by.x='chrom', by.y='chr') 
-np.regions = merge(np.regions, chrs, by.x='chrom', by.y='chr')
+var.seg.lengths = lapply(var.seg.lengths, function(x) merge(x, chrs, by.x='chr', by.y='chr') )
 
-prog.regions$chr.ratio = prog.regions$length.x/prog.regions$length.y
-np.regions$chr.ratio = np.regions$length.x/np.regions$length.y
+segment.lengths= lapply(segment.lengths, function(x) merge(x, chrs, by.x='chr', by.y='chr') )
 
-ttR = t.test(prog.regions$chr.ratio, np.regions$chr.ratio)
+opcov = sapply(segment.lengths[prog.pt], function(x) {
+  sum(as.numeric(x[['seg.len']]))/sum(as.numeric(chrs$length))
+})
 
-np = np.regions %>%
-  group_by(chrom) %>%
-  summarise(lengths = sum(length.x))
+onpcov = sapply(segment.lengths[np.pt], function(x) {
+  sum(as.numeric(x[['seg.len']]))/sum(as.numeric(chrs$length))
+})
 
-prog = prog.regions %>%
-  group_by(chrom) %>%
-  summarise(lengths = sum(length.x))
+chr.ratios = lapply(var.seg.lengths, function(x) x[['seg.len']]/x[['length']])
+# Segment coverage, variable regions
+ttR = t.test(unlist(chr.ratios[subset(sum.patient.data, Status == 'P')$Patient]), 
+             unlist(chr.ratios[subset(sum.patient.data, Status == 'NP')$Patient]))
 
-m = merge(prog, np, by='chrom', all.x=T)
-tt = t.test(m$lengths.x, m$lengths.y)
+var.seg.lengths = lapply(var.seg.lengths, function(df) {
+  as.data.frame(df %>%
+  group_by(chr, length) %>%
+  summarise(seg.lengths = sum(seg.len), 
+            chr.ratio = sum(seg.len)/length[1]))
+})
 
-plot(hclust( dist(chrs$length)), main='Chromosome cluster by length')
+pcov = median(sapply(var.seg.lengths[prog.pt], function(x) 
+  sum(as.numeric(x[['seg.lengths']]))/sum(as.numeric(x[['length']])) ), na.rm=T)
+
+npcov = median(sapply(var.seg.lengths[np.pt], function(x) 
+  sum(as.numeric(x[['seg.lengths']]))/sum(as.numeric(x[['length']])) ), na.rm=T)
 
 ## Grouping the chromosomes by length 1-12, 13-22
-
 #t.test(subset(np.regions, chrom %in% c(1:12))$chr.ratio, subset(np.regions, chrom %in% c(13:22))$chr.ratio)
@@
 
-df = data.frame( 'short' = rbind( 'np' = sum(table(np.regions$chrom)[1:12]),
-       'p' = sum(table(prog.regions$chrom)[1:12])),
-                 'long' = rbind( 'np' = sum(table(np.regions$chrom)[13:22], na.rm=T),
-       'p' = sum(table(prog.regions$chrom)[13:22], na.rm=T)))
+#There may be an enrichment for segments in the shorter chromosomes amongst progressors but it's weak (fisher `r ftL$p.value`)
+#df = data.frame( 'short' = rbind( 'np' = sum(table(np.regions$chrom)[1:12]),
+#       'p' = sum(table(prog.regions$chrom)[1:12])),
+#                 'long' = rbind( 'np' = sum(table(np.regions$chrom)[13:22], na.rm=T),
+#       'p' = sum(table(prog.regions$chrom)[13:22], na.rm=T)))
 
-ftL = fisher.test(df)
+#ftL = fisher.test(df)
 ```
 
-So it appears that adjusting for chromosome length (looking at the ratio of the chromosome rather than the raw bp) there is still a significant difference between progressers and non: t-test p-value `r ttR$p.value`
+So it appears that adjusting for chromosome length (looking at the ratio of the chromosome covered by segments rather than the raw bp) a significant difference in the coverage of the genome by CN segments that are variable between progressors and non: t-test p-value `r signif(ttR$p.value, 2)`.
+Overall bp coverage of the sWGS is `r round(mean(c(opcov, onpcov)), 2)*100`%  (P `r round(mean(opcov),2)`, NP `r round(mean(onpcov), 2)`).
 
-There appears to be no enrichment based on the length of chromosomes either (fisher `r ftL$p.value`)
+Variable regions cover `r round(pcov, 2)*100`% of the the progressors genome, and `r round(npcov, 2)*100`% of non-progressor genomes.
 
-```{r cor, echo=F, message=F, warning=F, fig.height=10,fig.width=10, fig.align='left'}
+```{r cor, echo=F, message=F, warning=F, fig.height=10,fig.width=10, fig.align='left', eval=F}
 
@@
 for (name in names(patient.data)) {
-
   sample.normalised.segvals = patient.data[[name]]$norm.seg.vals # prog
@@
             unlist(lapply(cors[subset(sum.patient.data, Status == 'NP')$Patient], function(x) sum(x$diff))))
+
+#There's no clear difference (currently) between P and NP for the similar correlation (`r simPNP$p.value`) or the inverse (`r diffPNP$p.value`) correlations.
+There are other ways we may be able to look at this still.
+
 ```
-
-There's no clear difference (currently) between P and NP for the similar correlation (`r simPNP$p.value`) or the inverse (`r diffPNP$p.value`) correlations.
-There are other ways we may be able to look at this still.
 
@@
 for (patient.name in names(patient.data)) {
-  print(patient.name)
   regions = patient.data[[patient.name]]$variable.regions[,c('chrom','arm','start.pos','end.pos')]
@@
   reg = cancer.consensus.genes(regions, ccgenes)
-  x = patient.data[[patient.name]]$variable.regions[which(!is.na(reg$Gene.Symbols)), patient.data[[patient.name]]$info$Samplename]
+  x = patient.data[[patient.name]]$variable.regions[which(!is.na(reg$Gene.Symbols)), intersect(patient.data[[patient.name]]$info$Samplename, colnames(patient.data[[patient.name]]$variable.regions))  ]
 
@@
 
-gc = gene.cnt %>% group_by(Gene.Symbols) %>% summarise(Progressors=sum(Progressors), Non.Progressors=sum(Non.Progressors))
-gc = arrange(subset(gc, Progressors >= Non.Progressors), -Progressors)
-m = melt(gc, id.vars=c('Gene.Symbols'))
-m$Gene.Symbols = factor(m$Gene.Symbols, levels= m$Gene.Symbols[c(nrow(m):1)] ) # order
-ggplot(m, aes(y=Gene.Symbols, x=variable)) + 
-  geom_tile(aes(fill=value), color='white') + 
-  scale_fill_gradient('n patients',low='white',high='royalblue4') + labs(x="",y='CGC Genes')
-
+#gc = gene.cnt %>% group_by(Gene.Symbols) %>% summarise(Progressors=sum(Progressors), Non.Progressors=sum(Non.Progressors))
+#gc = arrange(subset(gc, Progressors >= Non.Progressors), -Progressors)
+#m = melt(gc, id.vars=c('Gene.Symbols'))
+#m$Gene.Symbols = factor(m$Gene.Symbols, levels= m$Gene.Symbols[c(nrow(m):1)] ) # order
+#ggplot(m, aes(y=Gene.Symbols, x=variable)) + 
+#  geom_tile(aes(fill=value), color='white') + 
+#  scale_fill_gradient('n patients',low='white',high='royalblue4') + labs(x="",y='CGC Genes')
 ```
@@
 for (name in subset(sum.patient.data, Status == 'P')$Patient) {
-  print(name)
-  sn = grep('HGD|IMC', patient.data[[name]]$info$Pathology)
+  sn = which(colnames(patient.data[[name]]$variable.regions)[-(1:5)] %in% subset(patient.data[[name]]$info, grepl('HGD|IMC', Pathology))$Samplename)
   if (length(sn) <= 0 | nrow(patient.data[[name]]$variable.regions) <= 0) next
-  samplenames = patient.data[[name]]$info$Samplename
 
   ## HGD
-  pvnp[name, 'hgd'] = length(which(patient.data[[name]]$variable.regions[samplenames[sn]] >= gain.threshold | 
-                                patient.data[[name]]$variable.regions[samplenames[sn]] <= loss.threshold))
+  pvnp[name, 'hgd'] = length(which(patient.data[[name]]$variable.regions[,sn] >= gain.threshold | 
+                                patient.data[[name]]$variable.regions[,sn] <= loss.threshold))
 
   # BE-LGD
-  pvnp[name, 'be'] = length(which(patient.data[[name]]$variable.regions[samplenames[-sn]] >= gain.threshold | 
-                                patient.data[[name]]$variable.regions[samplenames[-sn]] <= loss.threshold))
+  pvnp[name, 'be'] = length(which(patient.data[[name]]$variable.regions[-sn] >= gain.threshold | 
+                                patient.data[[name]]$variable.regions[-sn] <= loss.threshold))
 }
 
-# No difference
+## TODO Check enrichment of variable regions found in BE vs HGD
 
@@
 
+#fisher.test(p, simulate.p.value = T)
+
 ct = cor.test(subset(sum.patient.data, Status == 'P')$total.samples, pvnp$hgd.ratio)
@@
 pander(pvnp, justify='left', 
-       caption=paste("Variable regions found in HGD vs non-HGD samples of progessor patients.
+       caption=paste("Number of variable regions found in HGD vs non-HGD samples of progessor patients.
 There's no significant difference between them (t-test p=",signif(tt$p.value,2),').
 Not all patients have HGD samples (NAs).
-There is also no significant correlation between the number/ratio of variable regions found in HGD samples vs non (Pearsons p=',signif(ct$p.value, 2),')',  sep=''))
+There is also no significant inverse correlation (r= ',signif(ct$estimate, 2),') between the number/ratio of variable regions found in HGD samples vs non (Pearsons p=',signif(ct$p.value, 2),')',  sep=''))
 
 ```
+
+This is just looking at the distribution of values per sample. The average ploidy per sample is around 1.  Currently we're calling variable regions based on this.
+```{r, warning=F}
+ks = list()
+for (pt in names(patient.data)) {
+  cols = intersect(arrange(patient.data[[pt]]$info, Pathology, Endoscopy.Year)$Samplename, colnames(patient.data[[pt]]$seg.vals[,-(1:5)]))
+  segvals = patient.data[[pt]]$seg.vals[,cols]
+
+  sdx = round(apply(segvals, 2, sd), 3)
+  se = round(apply(segvals, 2, function(x) sd(x)/sqrt(length(x))), 3)
+
+ gg = ggplot(melt(segvals)) + 
+        facet_wrap(~variable,scales = "free_x") + 
+        geom_histogram(bins=10, fill='blue', color='grey',aes(value, ..density..)) +
+        geom_density(aes(value+sdx/2), color='red', size=1) +
+        geom_density(aes(value-sdx/2), color='green', size=1) +
+        labs(title=paste(pt, '  (',patient.data[[pt]]$info$Status[1],')', sep=''))
+  print(gg)   
+  #plots[[pt]] = gg
+
+  ks[[pt]] = sapply(segvals, function(x) ks.test(x+sdx, x-sdx)$statistic  )
+  print(pt)
+  print(colSums(abs(round(patient.data[[pt]]$seg.vals[,-(1:5)], 1)-1)))
+}
+
+```

### 2017-04-04 RegressionPvNP.Rmd
Adding new analyses

---
title: "RegressionPvNP"
author: "Sarah Killcoyne"
date: "31 March 2017"
output: 
  html_document: 
    fig_caption: yes
    fig_height: 10
    fig_width: 10
    toc: yes
---

```{r setup, include=FALSE}
library(ggplot2)
library(ggfortify)
library(plyr)
library(pander)
library(Hmisc)
library(reshape2)
library(gridExtra)
library(GenomicRanges)
library(glmnet)

load('patients.Rdata', verbose=T)

source('lib/load_patient_metadata.R')

data = '~/Data/Ellie'

data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
plot.dir = paste(data, 'Analysis/multipcf_plots_fitted_perPatient', sep='/')

if (length(list.files(plot.dir)) <= 0)
  stop(paste("No analysis files found in", plot.dir ))

## Patient info file
patient.file = grep('All_patient_info.xls', data.files, value=T)
if (length(patient.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", data))

patient.info = read.patient.info(patient.file)
patient.info$Patient = gsub("/", "_", patient.info$Patient)
head(patient.info)

patient.info = arrange(patient.info, Status, Patient, Endoscopy.Year, Pathology)

sum.patient.data = summarise.patient.info(patient.info)
sum.patient.data = as.data.frame(subset(sum.patient.data, Patient %in% names(patient.data))) ## For now
pander(sum.patient.data[,c('Patient', 'Status', 'start.year', 'end.year','total.samples','highest.path')])

# Missing some of the samples as they aren't all sequenced yet
for (pt in names(patient.data)) {
  patient.data[[pt]]$info = patient.data[[pt]]$info[patient.data[[pt]]$info$Samplename %in% colnames(patient.data[[pt]]$seg.vals)[-(1:5)],]
}
```

```{r apclust-func, echo=F, message=F, warning=F}
suppressPackageStartupMessages( library(apcluster) )

apclust.data<-function(segdata, samples) {
  x1 = segdata[,intersect(colnames(segdata), samples)]
  x1 = x1[, samples]
  rownames(x1) = (segdata[,c(1:4)] %>%
    rowwise() %>%
    mutate(location=paste(paste(chrom, arm, sep=''), '.', start.pos, '-', end.pos, sep='')))$location
  return(x1)  
}

apclust<-function(segdata) {
  x1 = apclust.data(segdata)
  # q=0 minimize off-diagonal similarity
  ac = apcluster(negDistMat(r=2), x1, details=T, convits=25, q=0)
  return(list('apres'=ac, 'data'=x1))
}

net.similarity<-function(ac) {
  if (is.list(ac))
    ac = ac$apres
  ac@netsim
}

sum.similarity<-function(ac) {
  if (is.list(ac))
    ac = ac$apres
  ac@dpsim
}

clusters<-function(ac) {
  if (is.list(ac))
    ac = ac$apres
  ac@exemplars
}

plotAC<-function(aclist) {
  plot(aclist$apres, aclist$data)
}

heatmapAC<-function(aclist) {
  heatmap(aclist$apres)
}

# First vs last samples
#x1 = apclust.data(patient.data[[pt]]$seg.vals[,c(1:4, 6, ncol(patient.data[[pt]]$seg.vals))], patient.data[[pt]]$info$Samplename)
```

The previous analysis employed AP clustering on a per-patient basis to look for evidence that it was possible to see differences between the progressors and non-progressors, and to quantify that in some sort of complexity measure.  It was able to separate them, and continued to do so after removing HGD samples, but quickly failed to model any difference as I removed timepoints from the patient.

As suggested by Moritz it is possible to merge all of the patient samples and use GLMs to model the differences, and to select regions that may be most reflective of the differences. In this report that's what I've done, and it worked very well.  However, I'm a bit suspicious that it's worked too well.  I'm looking for biases in the data, or mistakes I might have made in setting up the model.

# Encode the large data matrix

To use GLMs across the patients I merge all samples from all patients and overlap the copy number segments. Where segments from a sample get split, the value of the original sample is retained for each of the split samples.

```{r tile, echo=F, message=F, warning=F, include=F}

chr.lengths = get.chr.lengths()
chr.lengths$chrom = sub('chr','',chr.lengths$chrom)
chr.lengths$start = 1

tile.w=1e7

genome = makeGRangesFromDataFrame(chr.lengths[1:22,], seqnames.field = 'chrom', end.field='chr.length')
tiles = tile(genome, width=tile.w)

tileFile = paste(as.character(tile.w),'_tiledpts.Rdata', sep='')
if (file.exists(tileFile)) {
  load(tileFile, verbose=T)
} else {
  grList = lapply(patient.data, function(df) {
    df$info = arrange(df$info, Endoscopy.Year, Pathology)

    x1 = apclust.data(df$seg.vals, df$info$Samplename)
    head(x1)

    segnames = as.data.frame(do.call(rbind, sapply(rownames(x1), strsplit, '[p|q].|-')))
    colnames(segnames) = c('chr','start','end')
    segnames[c('start','end')] = lapply(segnames[c('start','end')], function(x) as.numeric(as.character(x)))

    pt = unique(df$info$Patient)
    return( makeGRangesFromDataFrame(cbind(segnames,pt, x1), keep.extra.columns = T) )
  })

  sampleNames = unlist(sapply(patient.data, function(df) {
    df$info = arrange(df$info, Endoscopy.Year, Pathology)
    x1 = apclust.data(df$seg.vals, df$info$Samplename)
    colnames(x1)
  }))

  tile.segments<-function(tiles, gr, mergedSegments) {
    ov = findOverlaps(tiles, gr)
    for (qh in unique(queryHits(ov))) {
      print(qh)
      currentTile = tiles[[qh]]
      curov = findOverlaps(currentTile, gr)

      for (i in 1:length(curov)) {
        segment = currentTile[ queryHits(curov)[i]  ]
        #print(segment)
        rows = with(mergedDf, which( chr==as.character(seqnames(segment)) & start == start(segment) & end == end(segment)))
        segmentVals = as.data.frame(elementMetadata(gr[ subjectHits(curov)[i] ]))
        mergedSegments[rows, names(segmentVals)[-1]] = segmentVals[,-1]
      }
    }
    return(mergedSegments)  
  }

  mergedDf = do.call(rbind, lapply(tiles, function(tile) { 
    cbind('chr'=as.character(seqnames(tile)), as.data.frame(ranges(tile))[1:2]) 
    }) )
  mergedDf[,sampleNames] = NA

  for (gr in grList) {
    print(unique(gr$pt))
    mergedDf = tile.segments(tiles, gr, mergedDf)
  }

  #tmp = mergedDf
  mergedDf[is.na(mergedDf)] = 0
  rownames(mergedDf) = with(mergedDf, paste(chr, ':', start, '-', end, sep=''))

  save(mergedDf, file=tileFile)
}

```

Using a very large segment size for tiling across all patients (`r tile.w`) I get the following binomial models.  None of them fit well, and decreasing the size of the segments results in poor (or no) fits with mostly NA coefficients, or a lot of P(0 or 1).

# y = Progressors (1) vs Non (0)

The labels are split such that all samples from progressor patients are (1) and all samples from non-progressors are (0).

```{r labelsPNP, echo=F, message=F, include=F}
## binomial: dysplasia 1, BE 0

labels = unlist(sapply(patient.data, function(df) {
  df$info = arrange(df$info, Endoscopy.Year, Pathology)
  label = as.integer(df$info$Status == 'P') #as.integer(df$info$Pathology %in% c('HGD', 'IMC'))
  names(label) = df$info$Samplename
  return(label)
}))
names(labels) = sub('.*\\.', '',  names(labels))

pts = do.call(rbind, lapply(patient.data, function(df) {
  df$info = arrange(df$info, Endoscopy.Year, Pathology)
  cbind(df$info[,c('Patient','Samplename')])
}))
rownames(pts) = 1:nrow(pts)

# sort in label order
if (length(setdiff(colnames(mergedDf), names(labels))) > 3)
  warning("Labels vector is missing samples")

dysplasia.df = t(mergedDf[,names(labels)])
```

We have `r table(labels)[1]` samples from non-progressors and `r table(labels)[2]` samples from progressors.

Data matrix currently:
`r pander(dysplasia.df[1:10, 1:5], caption=paste("dimensions:", paste(dim(dysplasia.df), collapse=', ')), justify='left')`

## cv.glmnet Ridge vs Lasso

With all `r ncol(dysplasia.df)` regions of the genome, first check if/which regression may be appropriate. It appears that pure lasso may provide the best balance for number of features with non-zero coefficients.

```{r cvglmnetPNP, echo=F, warning=F, message=F}
plot.multi.cv.glmnet<-function(x,y,alpha=c(0,1), family="binomial", title='cv.glmnet') {
  plots <- list(  )
  cv.models = data.frame(matrix(ncol=3,nrow=0,dimnames=list(c(), c('class','lambda','cvm'))))
  for (a in alpha) {
    cv = cv.glmnet(x, y, alpha=a, family=family, standardize=F)
    cv.models = rbind(cv.models, cbind('class'=paste('cv',a,sep=''), 'lambda'=log(cv$lambda), 'cvm'=cv$cvm))
    plots[[as.character(a)]] = autoplot(cv, main=paste('alpha',a), ylab=cv$name)
  }
  cv.models[c('lambda','cvm')] = lapply(cv.models[c('lambda','cvm')], function(x) as.numeric(as.character(x)))

  gg = ggplot(cv.models, aes(x=lambda, y=cvm, color=class)) + geom_point() + geom_line() + 
      labs(x='log(Lambda)', y='Binomial Deviance', title=title)

  plots[['all']] = gg

  return(plots)
}

gp = plot.multi.cv.glmnet(dysplasia.df, labels, alpha=c(0,0.5,0.8,1), title='cv.glmnet P vs NP')

cv1 = cv.glmnet(dysplasia.df, labels, alpha=1, family='binomial')
gg = autoplot(cv1$glmnet.fit, xvar='lambda', main='alpha=1 glmnet.fit from cv.glmnet') + theme(legend.position='none')

gp[['coef.1']] = gg

do.call(grid.arrange, c(gp, ncol=2, nrow=3))

```

## xval by patients

Split into 5ths by patient, fit and predict at each split, select min lambda, get coefficients from the initial model using that value.

```{r xvalpt, message=F, warning=F, echo=F, fig.height=6, fig.width=6}
pi.hat <- function(x) exp(x)/(1+exp(x))

precisionRecall<-function(actual, predicted) {
    retrieved <- sum(predicted==1)
    recall <- sum(predicted==1 & actual==1) / sum(actual==1)
    precision <- sum(predicted==1 & actual==1) / retrieved
    f1 = signif(2*(1/(1/precision+1/recall)), 2)
    return(data.frame('precision'=precision, 'recall'=recall, 'F1'=signif(f1, 2)))
  }

cv.patient.glmnet<-function(x,y,lambda,patients) {
  message(paste("Running", length(unique(patients$group)), "folds"))

  cv.class = matrix(nrow=5, ncol=length(lambda))
  auc = matrix(nrow=5, ncol=length(lambda))

  for (i in 1:length(unique(patients$group))) {  ## This is not stable. 
    message(paste(i, "fold"))
    test.rows = which(rownames(x) %in% subset(patients, group == i)$Samplename)
    test = x[test.rows,]
    training = x[-test.rows,]
    # pre-spec lambda seq
    fit <- glmnet(training, y[-test.rows], lambda=lambda, family='binomial', alpha=1)

    pred.class <- predict(fit, test, type='class') # check this later
    auc[i,] = apply(pred.class, 2, function(pd) {
      if (sum(as.numeric(pd)) <= 0) return(0)
      as.numeric(pROC::auc(pROC::roc(response=as.numeric(pd), predictor=labels[test.rows])))  
    })

    # careful: matrix
    pred <- pi.hat(predict(fit, test)) 
    cv.class[i,] = apply(pred, 2, function(p) {
      # Confusion matrix
      #cmat = table(labels[test.rows], p>0.5)
      #sum(diag(cmat))/sum(cmat)
      (1-p)%*%labels[test.rows] # quantitiative
    })
  }
  return(list('cv.class'=cv.class, 'auc'=auc))
}

fit0 <- glmnet(dysplasia.df, labels, alpha = 1) # all patients
autoplot(fit0, xvar='lambda', main='fit0, all samples') + theme(legend.position='none') 
lambda <- fit0$lambda

# split patients in 5ths
s = {set.seed(2); sample(rep(seq(5), length = length(unique(pts$Patient))))}
patients = merge(pts, cbind('Patient'=unique(pts$Patient), 'group'=s), by="Patient")

# This just makes sure the sets don't become too unbalanced with regards to the labels.
sets = table(cbind(patients, labels[patients$Samplename])[,c(3:4)])
while (length(which(sets/rowSums(sets) < 0.35)) >= 2 | length(which(sets/rowSums(sets) == 0)) > 0 ) {
  print(".")
  s = sample(rep(seq(5), length = length(unique(pts$Patient))))
  patients = merge(pts, cbind('Patient'=unique(pts$Patient), 'group'=s), by="Patient")
  sets = table(cbind(patients, labels[patients$Samplename])[,c(3:4)])
}

cv.patient = cv.patient.glmnet(dysplasia.df, labels, lambda, patients)
lambda.opt <- lambda[which.min(colMeans(cv.patient$cv.class))]

df = cbind.data.frame('lambda.at'=1:ncol(cv.patient$cv.class),'mean'=colMeans(cv.patient$cv.class), 
                      'sme'=apply(cv.patient$cv.class, 2, sd)/5, 
                      'sd'=apply(cv.patient$cv.class, 2, sd), 
                      'lambda'=lambda)
smeA = sd(colMeans(cv.patient$cv.class))/ncol(cv.patient$cv.class) #??

ggplot(df, aes(y=mean,x=lambda.at)) + geom_point() + 
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + labs(y='mean of confusion matrix (P)') +
  geom_point(data=subset(df, mean == min(mean)), aes(y=mean, x=lambda.at), col='red') +
  geom_point(data=df[which.min(df$mean)+1,], aes(y=mean, x=lambda.at), col='blue') +
  geom_point(data=df[which.min(df$mean)-1,], aes(y=mean, x=lambda.at), col='green')

lambda.opt = lambda[which.min(df$mean)-1]

coef.opt <- coef.glmnet(fit0, s=lambda.opt)
coef.opt = as.matrix(coef.opt)
coef.opt = as.matrix(coef.opt[ which(coef.opt != 0)[-1] ,])
coef.opt = as.matrix(coef.opt[order(-coef.opt[,1]),])

```

At a the minimum lambda-1se (green dot) the value is `r signif(lambda.opt, 3)`. In the initial model `r length(coef.opt)` coefficients have a non-zero value.  Using these to fit a linear model by removing the remainder of the features we get a pretty clear separation between samples from progressors and non.

```{r echo=T, tidy=T, warning=F, message=F}

# So then the best model uses a lambda value in this range?
bestF = rownames(coef.opt)

## So the model would be...
data = dysplasia.df[,bestF]

fit = lm(labels~., data=cbind.data.frame(data, labels))
summary(fit)
```

```{r echo=F, warning=F, message=F}
df = cbind.data.frame('index'=1:length(labels), 'fitted'=fit$fitted.values, 'status'=labels, 'resid'=fit$residuals)

grid.arrange(
  ggplot(df, aes(x=index,y=fitted, col=as.factor(status))) + geom_point() +
    labs(title='Fitted values vs status for selected coef', x=paste(length(fit$fitted.values),'samples'), y='fitted value'),
  ggplot(df, aes(x=resid,y=fitted, col=as.factor(status))) + geom_point() +
    labs(title='Fitted values vs residuals for selected coef')
)

pc = rbind('Progressors'=signif(precisionRecall(labels == 1, fit$fitted.values > 0.5), 3),
            'Non-Progressors'=signif(precisionRecall(labels == 0, fit$fitted.values < 0.5),3))
```

The model separates surprisingly well (assuming this was the correct setup).

`r pander( pc, caption="Precision/Recall statistics for the linear model using the non-zero coefficients.")`

## Questions

While this looks great, I'm a bit suspicious that it's too good.

1. Can it still predict well if all HGD/IMC samples are removed from the matrix?
2. If I remove one sample from the progressors and iteratively rerun it, at what point does it fail?  I would expect this to occur at least at the earliest timepoints where BE is indicated.

### Remove HGD/IMC Samples

Just as a point, the initial model is tested against `r nrow(data)` samples with `r ncol(data)` features.

```{r noHGD, echo=F, message=F, warning=F}
`%nin%` <- Negate(`%in%`)

info = do.call(rbind, lapply(patient.data, function(df) df$info))

samples = intersect(rownames(data), subset(info, Pathology %nin% c('HGD', 'IMC'))$Samplename)

fit = lm(labels~., data=cbind.data.frame(data[samples,], 'labels'=labels[samples]))
```

This model is fit against `r nrow(data[samples,])` samples instead as `r nrow(data)-length(samples)` are pathologically HGD/IMC. While this looks great, I'm a bit surprised.

```{r echo=F,warning=F,message=F, fig.width=6,fig.height=6}

df = cbind.data.frame('index'=1:length(labels[samples]), 'fitted'=fit$fitted.values, 'status'=labels[samples], 'resid'=fit$residuals)

grid.arrange(
  ggplot(df, aes(x=index,y=fitted, col=as.factor(status))) + geom_point() +
    labs(title='Fitted values vs status for selected coef', x=paste(length(fit$fitted.values),'samples'), y='fitted value'),
  ggplot(df, aes(x=resid,y=fitted, col=as.factor(status))) + geom_point() +
    labs(title='Fitted values vs residuals for selected coef')
)

pc = rbind('Progressors'=signif(precisionRecall(labels[samples] == 1, fit$fitted.values > 0.5), 3),
              'Non-Progressors'=signif(precisionRecall(labels[samples] == 0, fit$fitted.values < 0.5),3))
```

`r pander(pc , caption="Precision/Recall statistics for the linear model using the non-zero coefficients.")`

### Remove LGD Samples as well

This just strikes me as far too good.  I would expect some of the progressor samples to separate even with LGD removed, but not all.

```{r echo=F,warning=F,message=F, fig.width=6,fig.height=6}

samples = intersect(rownames(data), subset(info, Pathology %nin% c('HGD', 'IMC', 'LGD'))$Samplename)
fit = lm(labels~., data=cbind.data.frame(data[samples,], 'labels'=labels[samples]))

df = cbind.data.frame('index'=1:length(labels[samples]), 'fitted'=fit$fitted.values, 'status'=labels[samples], 'resid'=fit$residuals)

grid.arrange(
  ggplot(df, aes(x=index,y=fitted, col=as.factor(status))) + geom_point() +
    labs(title='Fitted values vs status for selected coef', x=paste(length(fit$fitted.values),'samples'), y='fitted value'),
  ggplot(df, aes(x=resid,y=fitted, col=as.factor(status))) + geom_point() +
    labs(title='Fitted values vs residuals for selected coef')
)

pc = rbind('Progressors'=signif(precisionRecall(labels[samples] == 1, fit$fitted.values > 0.5), 3),
              'Non-Progressors'=signif(precisionRecall(labels[samples] == 0, fit$fitted.values < 0.5),3))

```

`r pander(pc,caption="Precision/Recall statistics for the linear model using the non-zero coefficients.")`

### Peel back samples one at a time until it fails

This is not what I would expect. By the time I get to the earliest timepoint most of the remaining samples in the progressors are BE only. Some of them should not separate well.

```{r peel, echo=F, message=F, warning=F, fig.height=4,fig.width=4}
get.next.samples<-function(patient.data, i) {
  as.vector(unlist(sapply(patient.data, function(pd) {
    j = i
    pd$info = arrange(pd$info, Endoscopy.Year, Pathology)

    if (j >= nrow(pd$info)) j = nrow(pd$info)

    pd$info[1:(nrow(pd$info)-j), 'Samplename']
  }) ))
}

nonps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'NP')$Patient], 0)
for (i in seq(1,35,5)) {
  ps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'P')$Patient], i)
  message(paste("# progressor samples (removing from the newest to oldest):", length(ps)))
  samples = c(nonps, ps)

  fit = lm(labels~., data=cbind.data.frame(data[samples,], 'labels'=labels[samples]))
  summary(fit)

  df = cbind.data.frame('index'=1:length(labels[samples]), 'fitted'=fit$fitted.values, 'status'=labels[samples], 'resid'=fit$residuals)

  grid.arrange(
    ggplot(df, aes(x=index,y=fitted, col=as.factor(status))) + geom_point() +
      labs(title='Fitted values vs status for selected coef', x='', y='fitted value'),
    #ggplot(df, aes(x=resid,y=fitted, col=as.factor(status))) + geom_point() +
    #  labs(title='Fitted values vs residuals for selected coef'),
    top=paste("Total samples in fit:", length(samples))
  )

  pc = rbind.data.frame('Progressor'=signif(precisionRecall(labels[samples] == 1, fit$fitted.values > 0.5),3),
       'Non-Progressor'=signif(precisionRecall(labels[samples] == 0, fit$fitted.values < 0.5),3))

  pander(pc, justify='left')
}

```

`r pander(table(subset(patient.info, Samplename %in% samples)[,c('Status','Pathology')]), caption="Pathology of samples in final model")`

## Again randomly assigning labels

One check is to randomly assign labels and rerun the model, at the same time I am pulling out samples in time backwards. The good thing is that the model no longer works as I would expect.

```{r, echo=F, message=F, warning=F, fig.height=4,fig.width=4}
fstats = matrix()
nonps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'NP')$Patient], 0)
for (i in seq(1,35,10)) {
  ps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'P')$Patient], i)
  message(paste("# progressor samples (removing from the newest to oldest):", length(ps)))
  samples = c(nonps, ps)

  randLabels = sample(c(0,1), length(samples), replace = T)

  fit = lm(labels~., data=cbind.data.frame(data[samples,], 'labels'=randLabels))
  summary(fit)

  df = cbind.data.frame('index'=1:length(labels[samples]), 'fitted'=fit$fitted.values, 'status'=labels[samples], 'resid'=fit$residuals)

  grid.arrange(
    ggplot(df, aes(x=index,y=fitted, col=as.factor(status))) + geom_point() +
      labs(title='Fitted values vs status for selected coef', x='', y='fitted value'),
#    ggplot(df, aes(x=resid,y=fitted, col=as.factor(status))) + geom_point() +
#      labs(title='Fitted values vs residuals for selected coef'),
    top=paste("Total samples in fit:", length(samples))
  )

  pc = rbind.data.frame('Progressor'=signif(precisionRecall(labels[samples] == 1, fit$fitted.values > 0.5),3),
       'Non-Progressor'=signif(precisionRecall(labels[samples] == 0, fit$fitted.values < 0.5),3))

  pander(pc, justify='left')
}
```

Final precision recall, is just too good:

`r pander(pc, justify='left')`

# Biases?

Some of the samples are going to have duplicated values in the matrix due to the method of overlapping segments to encode all of the patient data together.  It's possible that is causing the progressors to stand out more?

### 1. Duplication of values - remove samples

If I remove samples (not patients) by the number of duplicated values they have in the matrix the fit actually separates more. Somehow I expected they might separate less.

```{r, echo=F, message=F, warning=F, fig.height=4,fig.width=4}
duprows = apply(data, 1, function(row) {
  length(which(duplicated(row)))
})

pi = cbind.data.frame(patient.info[,c(1,3:5,8:10,12)], 'duprows'=duprows[patient.info$Samplename])

#t.test(subset(pi, Status == 'NP')$duprows, subset(pi, Status == 'P')$duprows)$p.value

## So what happens if I remove from the model those with the most duplicated rows? 
#summary(duprows)

samples = names(duprows[which(duprows <= median(duprows))])

pander(
  cbind(pi %>% group_by(Status) %>% summarise('total.dup'=sum(duprows,na.rm=T)), table(labels[samples]))[,c(1,2,4)],
  caption="There are significantly more duplicated rows following the overlap process in progressors (total.dup). In the samples that have less than the median number of duplicated values, both P and NP are still represented at about the same rate (~44% of rows are P)")

  fit = lm(labels~., data=cbind.data.frame(data[samples,], 'labels'=labels[samples]))

  df = cbind.data.frame('index'=1:length(labels[samples]), 'fitted'=fit$fitted.values, 'status'=labels[samples], 'resid'=fit$residuals)

  grid.arrange(
    ggplot(df, aes(x=index,y=fitted, col=as.factor(status))) + geom_point() +
      labs(title='Fitted values vs status for selected coef', x='', y='fitted value'),
#    ggplot(df, aes(x=resid,y=fitted, col=as.factor(status))) + geom_point() +
#      labs(title='Fitted values vs residuals for selected coef'),
    top=paste("Total samples in fit:", length(samples))
  )

  pc = rbind.data.frame('Progressor'=signif(precisionRecall(labels[samples] == 1, fit$fitted.values > 0.5),3),
       'Non-Progressor'=signif(precisionRecall(labels[samples] == 0, fit$fitted.values < 0.5),3))

```

`r  pander(pc, caption=paste('The median for duplicated values in a row is',median(duprows),'the linear model that uses just those rows with less than the median shows a complete separation'))`

### 2. Duplication of values - remove by patient

Remove all samples from single patients if any of their samples have a high number of duplicated rows, most of these are progressors. The resulting fit is still suspiciously good though so I think the bias isn't based on duplicated values.

```{r, echo=F, message=F, warning=F, fig.height=4,fig.width=4}
pi = cbind.data.frame(patient.info[,c(1,3:5,8:10,12)], 'duprows'=duprows[patient.info$Samplename])

## So what happens if I remove from the model those with the most duplicated rows? 
#summary(duprows)

toremove = names(duprows[which(duprows > median(duprows))])

samples = intersect(rownames(data), subset(patient.info, Patient %in% unique(subset(patient.info, Samplename %in% toremove)$Patient))$Samplename)

pander(table(labels[samples]), caption="0=NP, 1=P")

  fit = lm(labels~., data=cbind.data.frame(data[samples,], 'labels'=labels[samples]))

  df = cbind.data.frame('index'=1:length(labels[samples]), 'fitted'=fit$fitted.values, 'status'=labels[samples], 'resid'=fit$residuals)

  grid.arrange(
    ggplot(df, aes(x=index,y=fitted, col=as.factor(status))) + geom_point() +
      labs(title='Fitted values vs status for selected coef', x='', y='fitted value'),
#    ggplot(df, aes(x=resid,y=fitted, col=as.factor(status))) + geom_point() +
#      labs(title='Fitted values vs residuals for selected coef'),
    top=paste("Total samples in fit:", length(samples))
  )

  pc = rbind.data.frame('Progressor'=signif(precisionRecall(labels[samples] == 1, fit$fitted.values > 0.5),3),
       'Non-Progressor'=signif(precisionRecall(labels[samples] == 0, fit$fitted.values < 0.5),3))

```

`r pander(pc, caption=paste('In this I removed all samples from patients that had greater than the median value',median(duprows), 'of duplicated rows. The separation is less perfect but still very good.'))`

```{r, echo=F, warning=F, message=F, eval=F}
test.rows = which(rownames(dysplasia.df) %in% subset(patients, group == 3)$Samplename)
test = dysplasia.df[test.rows,]
training = dysplasia.df[-test.rows,]

##  Found here http://amunategui.github.io/binary-outcome-modeling/  this probably isn't correct for my case
library(caret)

objControl <- trainControl(method='cv', number=5, returnResamp='none')
#objModel <- train(training,as.factor(labels[-test.rows]), method='glmnet',trControl=objControl, lambda=lambda.opt)
objModel <- train(dysplasia.df[,bestF],as.factor(labels), method='glmnet',trControl=objControl, lambda=lambda.opt)

vimp <- varImp(objModel, scale=F, lambda = lambda.opt)
results <- data.frame(row.names(vimp$importance),vimp$importance$Overall)
results$VariableName <- rownames(vimp)
colnames(results) <- c('VariableName','Weight')
results <- results[(results$Weight != 0),]

which(results$VariableName %in% bestF)

ggplot(subset(results, Weight > 10), aes(x=reorder(VariableName, Weight), y=Weight)) + 
  geom_bar(stat="identity", fill='blue', color='grey') + coord_flip() +
  labs(y="< (-) importance >  < neutral >  < importance (+) >", x="")
#+   theme(axis.text.x = element_text(angle = 90, hjust = 1))

## Very close to the same for the top N coefficients
rownames(coef.opt) %in% as.character(arrange(results, -Weight)$VariableName)

```

```{r labels2, echo=F, message=F, include=F, eval=F}
# y = HGD (1) vs BE/LGD (0)

#Do the same again, but with HGD/IMC labeled instead

## binomial: dysplasia 1, BE 0
labels = unlist(sapply(patient.data, function(df) {
  df$info = arrange(df$info, Endoscopy.Year, Pathology)
  label = as.integer(df$info$Pathology %in% c('HGD', 'IMC'))
  names(label) = df$info$Samplename
  return(label)
}))
names(labels) = sub('.*\\.', '',  names(labels))

# sort in label order
if (length(setdiff(colnames(mergedDf), names(labels))) > 3)
  warning("Labels vector is missing samples")
```

```{r cvglmnet, echo=F, warning=F, message=F, fig.height=8, fig.width=8, eval=F}
#Note on the labels. When trying to classify samples that are HGD or not we have `r table(labels)[1]` samples without and only `r table(labels)[2]`) with HGD.

## cv.glmnet?

gp = plot.multi.cv.glmnet(dysplasia.df, labels, alpha=c(0,0.5,1), title='cv.glmnet HGD vs BE')
do.call(grid.arrange, c(gp, ncol=2, nrow=3))

cv1 = cv.glmnet(dysplasia.df, labels, alpha=1, family='binomial')
plot(cv1$glmnet.fit, main='alpha=1 glmnet.fit from cv.glmnet')
```

```{r xvalpt2, message=F, warning=F, echo=F, fig.height=6, fig.width=6, eval=F}

# This just makes sure the sets don't become too unbalanced with regards to the labels.
sets = table(cbind(patients, labels[patients$Samplename])[,c(3:4)])
s = sample(rep(seq(5), length = length(unique(pts$Patient))))
patients = merge(pts, cbind('Patient'=unique(pts$Patient), 'group'=s), by="Patient")

while ( length(which(table(patients$group)/nrow(patients) < 0.2)) > 1 ) {
  print(".")
  s = sample(rep(seq(5), length = length(unique(pts$Patient))))
  patients = merge(pts, cbind('Patient'=unique(pts$Patient), 'group'=s), by="Patient")
  #sets = table(cbind(patients, labels[patients$Samplename])[,c(3:4)])
}

fit0 <- glmnet(dysplasia.df, labels, alpha = 1) # all patients
plot(fit0)
title(main='fit0, all samples')
lambda <- fit0$lambda

cv.class = matrix(nrow=5, ncol=length(lambda))
prec.rec = matrix(nrow=5, ncol=length(lambda))
auc = matrix(nrow=5, ncol=length(lambda))
for (i in 1:5) {
    test.rows = which(rownames(dysplasia.df) %in% subset(patients, group == i)$Samplename)
    test = dysplasia.df[test.rows,]
    training = dysplasia.df[-test.rows,]
    # pre-spec lambda seq
    fit <- glmnet(training, labels[-test.rows], lambda=lambda, family='binomial', alpha=1) 
    plot(fit)
    title(main=paste('fold', i))

    # careful: matrix
    pred <- pi.hat(predict(fit, test)) 
    pred.class <- predict(fit, test, type='class') # check this later

    auc[i,] = apply(pred.class, 2, function(x) {
      if (sum(as.numeric(x)) <= 0) return(0)
      as.numeric(auc(roc(response=as.numeric(x), predictor=labels[test.rows])))  
    })

    # prec.rec[i,] = apply(pred.class, 2, function(p) {
    #    precisionRecall(labels[test.rows], as.integer(p))$recall
    # })
    cv.class[i,] = apply(pred, 2, function(p) {
      # Confusion matrix
      #boxplot(p~labels[test.rows], main=i)
      # cmat = table(labels[test.rows], p>0.5)
      # sum(diag(cmat))/sum(cmat)
      (1-p)%*%labels[test.rows] # quantitiative
    })
    #s = sample(1:ncol(pred),1)
    #boxplot(pred[,s]~labels[test.rows], main=s)
}
lambda.opt <- lambda[which.min(colMeans(cv.class))]

## TODO...NOPE, this pulls nearly all of the features so START OVER.  This is likely because there are so few HGD samples

print(paste("mean auc at optimal lambda:", round(mean(auc[,which.min(colMeans(cv.class))]),2)))

#coef.opt <- coef.glmnet(fit0, s=min(fit0$lambda))
coef.opt <- coef.glmnet(fit0, s=lambda.opt)
coef.opt = as.matrix(coef.opt)
coef.opt = as.matrix(coef.opt[ which(coef.opt != 0)[-1] ,])
coef.opt = as.matrix(coef.opt[order(-coef.opt[,1]),])

print( rownames(coef.opt) )

test.rows = which(rownames(dysplasia.df) %in% subset(patients, group == 3)$Samplename)
test = dysplasia.df[test.rows,]
training = dysplasia.df[-test.rows,]

##  Found here http://amunategui.github.io/binary-outcome-modeling/  this probably isn't correct for my case

objControl <- trainControl(method='cv', number=5, returnResamp='none')
#objModel <- train(training,as.factor(labels[-test.rows]), method='glmnet',trControl=objControl, lambda=lambda.opt)

predictions <- predict(object=objModel, test)
print(roc(predictions, as.integer(labels[test.rows]))$auc)

#objModel <- train(dysplasia.df,as.factor(labels), method='glmnet',trControl=objControl, lambda=lambda.opt)

vimp <- varImp(objModel, scale=F, lambda = lambda.opt)
results <- data.frame(row.names(vimp$importance),vimp$importance$Overall)
results$VariableName <- rownames(vimp)
colnames(results) <- c('VariableName','Weight')
results <- results[(results$Weight != 0),]

ggplot(subset(results, Weight > 20), aes(x=reorder(VariableName, Weight), y=Weight)) + 
  geom_bar(stat="identity", fill='blue', color='grey') + coord_flip() +
  labs(y="< (-) importance >  < neutral >  < importance (+) >", x="")
#+   theme(axis.text.x = element_text(angle = 90, hjust = 1))

## Very close to the same for the top N coefficients
rownames(coef.opt) %in% as.character(arrange(results, -Weight)$VariableName)

```

### 2017-04-04 MultisampleCopynumberBinnedAndFitted.R

@@ -10,9 +10,14 @@ args = commandArgs(trailingOnly=TRUE)
 if (length(args) < 3)
   stop("Missing required arguments: <qdna data dir> <patient spreadsheet> <output dir>")
 
+source('lib/load_patient_metadata.R')
+
 data = args[1]
 patient.file = args[2]
 outdir = args[3]
+#data = '~/Data/Ellie/QDNAseq'
+#patient.file = paste(data, 'All_patient_info.xlsx', sep='/')
+#outdir = '~/Data/Ellie/Analysis'
 
 
 all.patient.info = read.patient.info(patient.file)

### 2017-04-21 RegressionPvNP.Rmd
Minor bugs

@@
 ```{r xvalpt, message=F, warning=F, echo=F, fig.height=6, fig.width=6}
-pi.hat <- function(x) exp(x)/(1+exp(x))
-
 precisionRecall<-function(actual, predicted) {
@@
 
+pi.hat<-function(x) exp(x)/(1+exp(x))
+
+non.zero.coef<-function(fit, s) {
+  cf =  as.matrix(coef(fit, s))
+  cf[which(cf != 0),][-1]
+}
+
 cv.patient.glmnet<-function(x,y,lambda,patients) {
@@
 
+  fit.e = list()
   for (i in 1:length(unique(patients$group))) {  ## This is not stable. 
@@
     fit <- glmnet(training, y[-test.rows], lambda=lambda, family='binomial', alpha=1) 
+    # autoplot(fit) + theme(legend.position="none")
 
@@
       if (sum(as.numeric(pd)) <= 0) return(0)
-      as.numeric(pROC::auc(pROC::roc(response=as.numeric(pd), predictor=labels[test.rows])))  
+      #as.numeric(pROC::auc(pROC::roc(response=as.numeric(pd), predictor=labels[test.rows])))  
+      precisionRecall(labels[test.rows], as.numeric(pd))$F1
     })
@@
        #sum(diag(cmat))/sum(cmat)
-      (1-p)%*%labels[test.rows] # quantitiative
+      # quantitiative, pos results + neg results / number of test rows
+      (p%*%labels[test.rows] + (1-p) %*% (1-labels[test.rows]))/ length(test.rows)
     })
+
+    fit.e[[i]] = fit
   }
-  return(list('cv.class'=cv.class, 'auc'=auc))
+
+  df = cbind.data.frame('lambda.at'=1:ncol(cv.class),'mean'=colMeans(cv.class), 
+                        'sme'=apply(cv.class, 2, sd)/5, 
+                        'sd'=apply(cv.class, 2, sd), 
+                        'lambda'=lambda)
+
+  lambda.opt = lambda[which.max(df$mean)-1]
+  nzcf = lapply(fit.e, non.zero.coef, s=lambda.opt)
+
+  gp = ggplot(df, aes(y=mean,x=lambda.at)) + geom_point() + 
+          geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + labs(y='mean of confusion matrix (P)') +
+          geom_point(data=subset(df, mean == max(mean)), aes(y=mean, x=lambda.at, colour="max"), size=3 ) +
+          geom_point(data=df[which.max(df$mean)+1,], aes(y=mean, x=lambda.at, colour="max+1"), size=3) +
+          geom_point(data=df[which.max(df$mean)-1,], aes(y=mean, x=lambda.at, colour="max-1"), size=3) +
+          geom_hline(yintercept=df[which.max(df$mean)-1,'mean'], colour='grey') +
+          annotate("text", x=5, y=max(df$mean)+0.0005, label=round(df[which.max(df$mean)-1,'mean'],2)) +
+          theme(axis.title.x = element_text(size = 15, vjust=-.2)) + scale_colour_discrete(name = "Best lambda values") +
+          labs(title="CV Performance")
+
+  return(list('cv.class'=cv.class, 'f1'=auc[which.max(df$mean)-1], 'lambda.opt'=lambda.opt, 'lambdas'=df, 'plot'=gp, 'non.zero.cf'=nzcf))
 }
@@
 
-cv.patient = cv.patient.glmnet(dysplasia.df, labels, lambda, patients)
-lambda.opt <- lambda[which.min(colMeans(cv.patient$cv.class))]
+cv.patient = cv.patient.glmnet(x=dysplasia.df, y=labels, lambda, patients)
+lambda.opt = cv.patient$lambda.opt
 
-df = cbind.data.frame('lambda.at'=1:ncol(cv.patient$cv.class),'mean'=colMeans(cv.patient$cv.class), 
-                      'sme'=apply(cv.patient$cv.class, 2, sd)/5, 
-                      'sd'=apply(cv.patient$cv.class, 2, sd), 
-                      'lambda'=lambda)
-smeA = sd(colMeans(cv.patient$cv.class))/ncol(cv.patient$cv.class) #??
+coef.opt = as.data.frame(non.zero.coef(fit0, lambda.opt))
 
-ggplot(df, aes(y=mean,x=lambda.at)) + geom_point() + 
-  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + labs(y='mean of confusion matrix (P)') +
-  geom_point(data=subset(df, mean == min(mean)), aes(y=mean, x=lambda.at), col='red') +
-  geom_point(data=df[which.min(df$mean)+1,], aes(y=mean, x=lambda.at), col='blue') +
-  geom_point(data=df[which.min(df$mean)-1,], aes(y=mean, x=lambda.at), col='green')
+grid.arrange(cv.patient$plot)
 
-lambda.opt = lambda[which.min(df$mean)-1]
+## check how often the feature(s) is selected at that lamda in each split. "stability selection"
 
-coef.opt <- coef.glmnet(fit0, s=lambda.opt)
-coef.opt = as.matrix(coef.opt)
-coef.opt = as.matrix(coef.opt[ which(coef.opt != 0)[-1] ,])
-coef.opt = as.matrix(coef.opt[order(-coef.opt[,1]),])
+for (i in 1:length(cv.patient$non.zero.cf)) {
+  df = as.data.frame(cv.patient$non.zero.cf[[i]])
+  colnames(df) = i
+  coef.opt = merge(coef.opt, df, by='row.names', all.x=T)
+  rownames(coef.opt) = coef.opt$Row.names
+  coef.opt$Row.names = NULL
+  coef.opt[,(i+1)] = as.integer(!is.na(coef.opt[,(i+1)]))
+}
+coef.opt = coef.opt[order(sapply(rownames(coef.opt), function(x) as.numeric(unlist(strsplit(x, ':'))[1])  )),]
 
@@
 
-At a the minimum lambda-1se (green dot) the value is `r signif(lambda.opt, 3)`.
-In the initial model `r length(coef.opt)` coefficients have a non-zero value.
-Using these to fit a linear model by removing the remainder of the features we get a pretty clear separation between samples from progressors and non.
+`r pander(coef.opt, justify='left', caption='Features from the glm fitted with all data. Each numbered column indicates if the feature was identified in that split.')`
 
-```{r echo=T, tidy=T, warning=F, message=F}
-
-# So then the best model uses a lambda value in this range?
-bestF = rownames(coef.opt)
-
-## So the model would be...
-data = dysplasia.df[,bestF]
-
-fit = lm(labels~., data=cbind.data.frame(data, labels))
-summary(fit)
-```
-
-```{r echo=F, warning=F, message=F}
-df = cbind.data.frame('index'=1:length(labels), 'fitted'=fit$fitted.values, 'status'=labels, 'resid'=fit$residuals)
-
-grid.arrange(
-  ggplot(df, aes(x=index,y=fitted, col=as.factor(status))) + geom_point() +
-    labs(title='Fitted values vs status for selected coef', x=paste(length(fit$fitted.values),'samples'), y='fitted value'),
-  ggplot(df, aes(x=resid,y=fitted, col=as.factor(status))) + geom_point() +
-    labs(title='Fitted values vs residuals for selected coef')
-)
-
-pc = rbind('Progressors'=signif(precisionRecall(labels == 1, fit$fitted.values > 0.5), 3),
-            'Non-Progressors'=signif(precisionRecall(labels == 0, fit$fitted.values < 0.5),3))
-```
-
-The model separates surprisingly well (assuming this was the correct setup).
-
-`r pander( pc, caption="Precision/Recall statistics for the linear model using the non-zero coefficients.")`
-
-## Questions
-
-While this looks great, I'm a bit suspicious that it's too good.
-
-1. Can it still predict well if all HGD/IMC samples are removed from the matrix?
-2. If I remove one sample from the progressors and iteratively rerun it, at what point does it fail?  I would expect this to occur at least at the earliest timepoints where BE is indicated.
+At a the minimum lambda-1se (green dot) the value is `r signif(lambda.opt, 3)`. In the initial model `r length(coef.opt)` coefficients have a non-zero value.
 
@@
 
-Just as a point, the initial model is tested against `r nrow(data)` samples with `r ncol(data)` features.
+Just as a point, the initial model is tested against `r nrow(dysplasia.df)` samples with `r ncol(dysplasia.df)` features.
 
@@
 
-samples = intersect(rownames(data), subset(info, Pathology %nin% c('HGD', 'IMC'))$Samplename)
+samples = intersect(rownames(dysplasia.df), subset(info, Pathology %nin% c('HGD', 'IMC'))$Samplename)
 
-fit = lm(labels~., data=cbind.data.frame(data[samples,], 'labels'=labels[samples]))
-```
+cv.nohgd = cv.patient.glmnet(x=dysplasia.df[samples,], y=labels[samples], lambda, subset(patients, Samplename %in% samples))
+#lambda.opt = cv.patient$lambda.opt
+grid.arrange(cv.nohgd$plot + ggtitle("CV Performance, No HGD/IMC samples"))
+#coef.opt = as.data.frame(non.zero.coef(fit0, lambda.opt))
 
-This model is fit against `r nrow(data[samples,])` samples instead as `r nrow(data)-length(samples)` are pathologically HGD/IMC. While this looks great, I'm a bit surprised.
+samples = intersect(rownames(dysplasia.df), subset(info, Pathology %nin% c('HGD', 'IMC', 'LGD'))$Samplename)
+cv.nolgd = cv.patient.glmnet(x=dysplasia.df[samples,], y=labels[samples], lambda, subset(patients, Samplename %in% samples))
+#lambda.opt = cv.patient$lambda.opt
+grid.arrange(cv.nolgd$plot + ggtitle("CV Performance, No HGD/IMC/LGD samples"))
 
-```{r echo=F,warning=F,message=F, fig.width=6,fig.height=6}
-
-df = cbind.data.frame('index'=1:length(labels[samples]), 'fitted'=fit$fitted.values, 'status'=labels[samples], 'resid'=fit$residuals)
-
-grid.arrange(
-  ggplot(df, aes(x=index,y=fitted, col=as.factor(status))) + geom_point() +
-    labs(title='Fitted values vs status for selected coef', x=paste(length(fit$fitted.values),'samples'), y='fitted value'),
-  ggplot(df, aes(x=resid,y=fitted, col=as.factor(status))) + geom_point() +
-    labs(title='Fitted values vs residuals for selected coef')
-)
-
-pc = rbind('Progressors'=signif(precisionRecall(labels[samples] == 1, fit$fitted.values > 0.5), 3),
-              'Non-Progressors'=signif(precisionRecall(labels[samples] == 0, fit$fitted.values < 0.5),3))
-```
-
-`r pander(pc , caption="Precision/Recall statistics for the linear model using the non-zero coefficients.")`
-
-### Remove LGD Samples as well
-
-This just strikes me as far too good.  I would expect some of the progressor samples to separate even with LGD removed, but not all.
-
-```{r echo=F,warning=F,message=F, fig.width=6,fig.height=6}
-
-samples = intersect(rownames(data), subset(info, Pathology %nin% c('HGD', 'IMC', 'LGD'))$Samplename)
-fit = lm(labels~., data=cbind.data.frame(data[samples,], 'labels'=labels[samples]))
-
-df = cbind.data.frame('index'=1:length(labels[samples]), 'fitted'=fit$fitted.values, 'status'=labels[samples], 'resid'=fit$residuals)
-
-grid.arrange(
-  ggplot(df, aes(x=index,y=fitted, col=as.factor(status))) + geom_point() +
-    labs(title='Fitted values vs status for selected coef', x=paste(length(fit$fitted.values),'samples'), y='fitted value'),
-  ggplot(df, aes(x=resid,y=fitted, col=as.factor(status))) + geom_point() +
-    labs(title='Fitted values vs residuals for selected coef')
-)
-
-pc = rbind('Progressors'=signif(precisionRecall(labels[samples] == 1, fit$fitted.values > 0.5), 3),
-              'Non-Progressors'=signif(precisionRecall(labels[samples] == 0, fit$fitted.values < 0.5),3))
-
-```
-
-`r pander(pc,caption="Precision/Recall statistics for the linear model using the non-zero coefficients.")`
-
-### Peel back samples one at a time until it fails
-
-This is not what I would expect. By the time I get to the earliest timepoint most of the remaining samples in the progressors are BE only. Some of them should not separate well.
-
-```{r peel, echo=F, message=F, warning=F, fig.height=4,fig.width=4}
 get.next.samples<-function(patient.data, i) {
@@
 nonps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'NP')$Patient], 0)
+plots = list()
 for (i in seq(1,35,5)) {
   ps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'P')$Patient], i)
-  message(paste("# progressor samples (removing from the newest to oldest):", length(ps)))
   samples = c(nonps, ps)
-
-  fit = lm(labels~., data=cbind.data.frame(data[samples,], 'labels'=labels[samples]))
-  summary(fit)
-
-  df = cbind.data.frame('index'=1:length(labels[samples]), 'fitted'=fit$fitted.values, 'status'=labels[samples], 'resid'=fit$residuals)
-
-  grid.arrange(
-    ggplot(df, aes(x=index,y=fitted, col=as.factor(status))) + geom_point() +
-      labs(title='Fitted values vs status for selected coef', x='', y='fitted value'),
-    #ggplot(df, aes(x=resid,y=fitted, col=as.factor(status))) + geom_point() +
-    #  labs(title='Fitted values vs residuals for selected coef'),
-    top=paste("Total samples in fit:", length(samples))
-  )
-
-  pc = rbind.data.frame('Progressor'=signif(precisionRecall(labels[samples] == 1, fit$fitted.values > 0.5),3),
-       'Non-Progressor'=signif(precisionRecall(labels[samples] == 0, fit$fitted.values < 0.5),3))
-
-  pander(pc, justify='left')
+  cv.back = cv.patient.glmnet(x=dysplasia.df[samples,], y=labels[samples], lambda, subset(patients, Samplename %in% samples))
+  plots[[ as.character(i) ]] = cv.back$plot + ggtitle(paste("CV Performance - ", length(ps), "progressor samples"))
 }
 
-```
-
-`r pander(table(subset(patient.info, Samplename %in% samples)[,c('Status','Pathology')]), caption="Pathology of samples in final model")`
-
-## Again randomly assigning labels
-
-One check is to randomly assign labels and rerun the model, at the same time I am pulling out samples in time backwards. The good thing is that the model no longer works as I would expect.
-
-```{r, echo=F, message=F, warning=F, fig.height=4,fig.width=4}
-fstats = matrix()
-nonps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'NP')$Patient], 0)
-for (i in seq(1,35,10)) {
-  ps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'P')$Patient], i)
-  message(paste("# progressor samples (removing from the newest to oldest):", length(ps)))
-  samples = c(nonps, ps)
-
-  randLabels = sample(c(0,1), length(samples), replace = T)
-
-  fit = lm(labels~., data=cbind.data.frame(data[samples,], 'labels'=randLabels))
-  summary(fit)
-
-  df = cbind.data.frame('index'=1:length(labels[samples]), 'fitted'=fit$fitted.values, 'status'=labels[samples], 'resid'=fit$residuals)
-
-  grid.arrange(
-    ggplot(df, aes(x=index,y=fitted, col=as.factor(status))) + geom_point() +
-      labs(title='Fitted values vs status for selected coef', x='', y='fitted value'),
-#    ggplot(df, aes(x=resid,y=fitted, col=as.factor(status))) + geom_point() +
-#      labs(title='Fitted values vs residuals for selected coef'),
-    top=paste("Total samples in fit:", length(samples))
-  )
-
-  pc = rbind.data.frame('Progressor'=signif(precisionRecall(labels[samples] == 1, fit$fitted.values > 0.5),3),
-       'Non-Progressor'=signif(precisionRecall(labels[samples] == 0, fit$fitted.values < 0.5),3))
-
-  pander(pc, justify='left')
-}
-```
-
-Final precision recall, is just too good:
-
-`r pander(pc, justify='left')`
-
-# Biases?
-
-Some of the samples are going to have duplicated values in the matrix due to the method of overlapping segments to encode all of the patient data together.
-It's possible that is causing the progressors to stand out more?
-
-### 1. Duplication of values - remove samples
-
-If I remove samples (not patients) by the number of duplicated values they have in the matrix the fit actually separates more. Somehow I expected they might separate less.
-
-```{r, echo=F, message=F, warning=F, fig.height=4,fig.width=4}
-duprows = apply(data, 1, function(row) {
-  length(which(duplicated(row)))
-})
-
-pi = cbind.data.frame(patient.info[,c(1,3:5,8:10,12)], 'duprows'=duprows[patient.info$Samplename])
-
-#t.test(subset(pi, Status == 'NP')$duprows, subset(pi, Status == 'P')$duprows)$p.value
-
-## So what happens if I remove from the model those with the most duplicated rows? 
-#summary(duprows)
-
-samples = names(duprows[which(duprows <= median(duprows))])
-
-pander(
-  cbind(pi %>% group_by(Status) %>% summarise('total.dup'=sum(duprows,na.rm=T)), table(labels[samples]))[,c(1,2,4)],
-  caption="There are significantly more duplicated rows following the overlap process in progressors (total.dup).
-In the samples that have less than the median number of duplicated values, both P and NP are still represented at about the same rate (~44% of rows are P)")
-
-  fit = lm(labels~., data=cbind.data.frame(data[samples,], 'labels'=labels[samples]))
-
-  df = cbind.data.frame('index'=1:length(labels[samples]), 'fitted'=fit$fitted.values, 'status'=labels[samples], 'resid'=fit$residuals)
-
-  grid.arrange(
-    ggplot(df, aes(x=index,y=fitted, col=as.factor(status))) + geom_point() +
-      labs(title='Fitted values vs status for selected coef', x='', y='fitted value'),
-#    ggplot(df, aes(x=resid,y=fitted, col=as.factor(status))) + geom_point() +
-#      labs(title='Fitted values vs residuals for selected coef'),
-    top=paste("Total samples in fit:", length(samples))
-  )
-
-  pc = rbind.data.frame('Progressor'=signif(precisionRecall(labels[samples] == 1, fit$fitted.values > 0.5),3),
-       'Non-Progressor'=signif(precisionRecall(labels[samples] == 0, fit$fitted.values < 0.5),3))
+do.call('grid.arrange', c(plots, ncol=2, top="Peeling back timepoints in progressors"))
 
 ```
-
-`r  pander(pc, caption=paste('The median for duplicated values in a row is',median(duprows),'the linear model that uses just those rows with less than the median shows a complete separation'))`
-
-### 2. Duplication of values - remove by patient
-
-Remove all samples from single patients if any of their samples have a high number of duplicated rows, most of these are progressors.
-The resulting fit is still suspiciously good though so I think the bias isn't based on duplicated values.
-
-```{r, echo=F, message=F, warning=F, fig.height=4,fig.width=4}
-pi = cbind.data.frame(patient.info[,c(1,3:5,8:10,12)], 'duprows'=duprows[patient.info$Samplename])
-
-## So what happens if I remove from the model those with the most duplicated rows? 
-#summary(duprows)
-
-toremove = names(duprows[which(duprows > median(duprows))])
-
-samples = intersect(rownames(data), subset(patient.info, Patient %in% unique(subset(patient.info, Samplename %in% toremove)$Patient))$Samplename)
-
-pander(table(labels[samples]), caption="0=NP, 1=P")
-
-  fit = lm(labels~., data=cbind.data.frame(data[samples,], 'labels'=labels[samples]))
-
-  df = cbind.data.frame('index'=1:length(labels[samples]), 'fitted'=fit$fitted.values, 'status'=labels[samples], 'resid'=fit$residuals)
-
-  grid.arrange(
-    ggplot(df, aes(x=index,y=fitted, col=as.factor(status))) + geom_point() +
-      labs(title='Fitted values vs status for selected coef', x='', y='fitted value'),
-#    ggplot(df, aes(x=resid,y=fitted, col=as.factor(status))) + geom_point() +
-#      labs(title='Fitted values vs residuals for selected coef'),
-    top=paste("Total samples in fit:", length(samples))
-  )
-
-  pc = rbind.data.frame('Progressor'=signif(precisionRecall(labels[samples] == 1, fit$fitted.values > 0.5),3),
-       'Non-Progressor'=signif(precisionRecall(labels[samples] == 0, fit$fitted.values < 0.5),3))
-
-```
-
-`r pander(pc, caption=paste('In this I removed all samples from patients that had greater than the median value',median(duprows), 'of duplicated rows.
-The separation is less perfect but still very good.'))`
-
-```{r, echo=F, warning=F, message=F, eval=F}
-test.rows = which(rownames(dysplasia.df) %in% subset(patients, group == 3)$Samplename)
-test = dysplasia.df[test.rows,]
-training = dysplasia.df[-test.rows,]
-
-##  Found here http://amunategui.github.io/binary-outcome-modeling/  this probably isn't correct for my case
-library(caret)
-
-objControl <- trainControl(method='cv', number=5, returnResamp='none')
-#objModel <- train(training,as.factor(labels[-test.rows]), method='glmnet',trControl=objControl, lambda=lambda.opt)
-objModel <- train(dysplasia.df[,bestF],as.factor(labels), method='glmnet',trControl=objControl, lambda=lambda.opt)
-
-vimp <- varImp(objModel, scale=F, lambda = lambda.opt)
-results <- data.frame(row.names(vimp$importance),vimp$importance$Overall)
-results$VariableName <- rownames(vimp)
-colnames(results) <- c('VariableName','Weight')
-results <- results[(results$Weight != 0),]
-
-which(results$VariableName %in% bestF)
-
-ggplot(subset(results, Weight > 10), aes(x=reorder(VariableName, Weight), y=Weight)) + 
-  geom_bar(stat="identity", fill='blue', color='grey') + coord_flip() +
-  labs(y="< (-) importance >  < neutral >  < importance (+) >", x="")
-#+   theme(axis.text.x = element_text(angle = 90, hjust = 1))
-
-## Very close to the same for the top N coefficients
-rownames(coef.opt) %in% as.character(arrange(results, -Weight)$VariableName)
-
-```
-
-```{r labels2, echo=F, message=F, include=F, eval=F}
-# y = HGD (1) vs BE/LGD (0)
-
-#Do the same again, but with HGD/IMC labeled instead
-
-## binomial: dysplasia 1, BE 0
-labels = unlist(sapply(patient.data, function(df) {
-  df$info = arrange(df$info, Endoscopy.Year, Pathology)
-  label = as.integer(df$info$Pathology %in% c('HGD', 'IMC'))
-  names(label) = df$info$Samplename
-  return(label)
-}))
-names(labels) = sub('.*\\.', '',  names(labels))
-
-# sort in label order
-if (length(setdiff(colnames(mergedDf), names(labels))) > 3)
-  warning("Labels vector is missing samples")
-```
-
-```{r cvglmnet, echo=F, warning=F, message=F, fig.height=8, fig.width=8, eval=F}
-#Note on the labels. When trying to classify samples that are HGD or not we have `r table(labels)[1]` samples without and only `r table(labels)[2]`) with HGD.
-
-## cv.glmnet?
-
-gp = plot.multi.cv.glmnet(dysplasia.df, labels, alpha=c(0,0.5,1), title='cv.glmnet HGD vs BE')
-do.call(grid.arrange, c(gp, ncol=2, nrow=3))
-
-cv1 = cv.glmnet(dysplasia.df, labels, alpha=1, family='binomial')
-plot(cv1$glmnet.fit, main='alpha=1 glmnet.fit from cv.glmnet')
-```
-
-```{r xvalpt2, message=F, warning=F, echo=F, fig.height=6, fig.width=6, eval=F}
-
-# This just makes sure the sets don't become too unbalanced with regards to the labels.
-sets = table(cbind(patients, labels[patients$Samplename])[,c(3:4)])
-s = sample(rep(seq(5), length = length(unique(pts$Patient))))
-patients = merge(pts, cbind('Patient'=unique(pts$Patient), 'group'=s), by="Patient")
-
-while ( length(which(table(patients$group)/nrow(patients) < 0.2)) > 1 ) {
-  print(".")
-  s = sample(rep(seq(5), length = length(unique(pts$Patient))))
-  patients = merge(pts, cbind('Patient'=unique(pts$Patient), 'group'=s), by="Patient")
-  #sets = table(cbind(patients, labels[patients$Samplename])[,c(3:4)])
-}
-
-fit0 <- glmnet(dysplasia.df, labels, alpha = 1) # all patients
-plot(fit0)
-title(main='fit0, all samples')
-lambda <- fit0$lambda
-
-cv.class = matrix(nrow=5, ncol=length(lambda))
-prec.rec = matrix(nrow=5, ncol=length(lambda))
-auc = matrix(nrow=5, ncol=length(lambda))
-for (i in 1:5) {
-    test.rows = which(rownames(dysplasia.df) %in% subset(patients, group == i)$Samplename)
-    test = dysplasia.df[test.rows,]
-    training = dysplasia.df[-test.rows,]
-    # pre-spec lambda seq
-    fit <- glmnet(training, labels[-test.rows], lambda=lambda, family='binomial', alpha=1) 
-    plot(fit)
-    title(main=paste('fold', i))
-
-    # careful: matrix
-    pred <- pi.hat(predict(fit, test)) 
-    pred.class <- predict(fit, test, type='class') # check this later
-
-    auc[i,] = apply(pred.class, 2, function(x) {
-      if (sum(as.numeric(x)) <= 0) return(0)
-      as.numeric(auc(roc(response=as.numeric(x), predictor=labels[test.rows])))  
-    })
-
-    # prec.rec[i,] = apply(pred.class, 2, function(p) {
-    #    precisionRecall(labels[test.rows], as.integer(p))$recall
-    # })
-    cv.class[i,] = apply(pred, 2, function(p) {
-      # Confusion matrix
-      #boxplot(p~labels[test.rows], main=i)
-      # cmat = table(labels[test.rows], p>0.5)
-      # sum(diag(cmat))/sum(cmat)
-      (1-p)%*%labels[test.rows] # quantitiative
-    })
-    #s = sample(1:ncol(pred),1)
-    #boxplot(pred[,s]~labels[test.rows], main=s)
-}
-lambda.opt <- lambda[which.min(colMeans(cv.class))]
-
-## TODO...NOPE, this pulls nearly all of the features so START OVER.  This is likely because there are so few HGD samples
-
-print(paste("mean auc at optimal lambda:", round(mean(auc[,which.min(colMeans(cv.class))]),2)))
-
-#coef.opt <- coef.glmnet(fit0, s=min(fit0$lambda))
-coef.opt <- coef.glmnet(fit0, s=lambda.opt)
-coef.opt = as.matrix(coef.opt)
-coef.opt = as.matrix(coef.opt[ which(coef.opt != 0)[-1] ,])
-coef.opt = as.matrix(coef.opt[order(-coef.opt[,1]),])
-
-print( rownames(coef.opt) )
-
-test.rows = which(rownames(dysplasia.df) %in% subset(patients, group == 3)$Samplename)
-test = dysplasia.df[test.rows,]
-training = dysplasia.df[-test.rows,]
-
-##  Found here http://amunategui.github.io/binary-outcome-modeling/  this probably isn't correct for my case
-
-objControl <- trainControl(method='cv', number=5, returnResamp='none')
-#objModel <- train(training,as.factor(labels[-test.rows]), method='glmnet',trControl=objControl, lambda=lambda.opt)
-
-predictions <- predict(object=objModel, test)
-print(roc(predictions, as.integer(labels[test.rows]))$auc)
-
-#objModel <- train(dysplasia.df,as.factor(labels), method='glmnet',trControl=objControl, lambda=lambda.opt)
-
-vimp <- varImp(objModel, scale=F, lambda = lambda.opt)
-results <- data.frame(row.names(vimp$importance),vimp$importance$Overall)
-results$VariableName <- rownames(vimp)
-colnames(results) <- c('VariableName','Weight')
-results <- results[(results$Weight != 0),]
-
-ggplot(subset(results, Weight > 20), aes(x=reorder(VariableName, Weight), y=Weight)) + 
-  geom_bar(stat="identity", fill='blue', color='grey') + coord_flip() +
-  labs(y="< (-) importance >  < neutral >  < importance (+) >", x="")
-#+   theme(axis.text.x = element_text(angle = 90, hjust = 1))
-
-## Very close to the same for the top N coefficients
-rownames(coef.opt) %in% as.character(arrange(results, -Weight)$VariableName)
-
-```

### 2017-04-21 MultisampleCopynumberBinnedAndFitted.R

@@ -20,7 +20,7 @@ outdir = args[3]
 #outdir = '~/Data/Ellie/Analysis'
 
 
-all.patient.info = read.patient.info(patient.file)
+all.patient.info = read.patient.info(patient.file, set='All')
 if (is.null(all.patient.info))
   stop(paste("Failed to read patient file", patient.file))
 all.patient.info$Patient = gsub('/', '_',all.patient.info$Patient)
@@ -125,12 +125,13 @@ dir.create(plot.dir, showWarnings=F)
 for (s in samplenames) {
   pt = paste('SLX-', subset(all.patient.info, Samplename == s)$SLX.ID, sep='')
   
-  if (length(pt) > 0) {
-    dir.create(paste(plot.dir, pt, sep='/'), showWarnings=F)
-    png(paste(paste(plot.dir, pt, sep='/'),paste(s,"coverage_binned_fitted.png",sep="_"),sep="/"),width=1600,height=800)
-  } else {
-    png(paste(plot.dir,paste(s,"coverage_binned_fitted.png",sep="_"),sep="/"),width=1600,height=800)
-  }
+  #if (length(pt) > 0) {
+    pt.plot.dir = strip.whitespace(paste(plot.dir, pt, sep='/'))
+    dir.create(pt.plot.dir, showWarnings=F)
+    png(paste(pt.plot.dir,paste(s,"coverage_binned_fitted.png",sep="_"),sep="/"),width=1600,height=800)
+  #} else {
+  #  png(paste(plot.dir,paste(s,"coverage_binned_fitted.png",sep="_"),sep="/"),width=1600,height=800)
+  #}
   print(paste("plotting",pt, s))
   gg = ggplot(window.depths[,c(chr.info, s)],aes_string(x="genome.pos", y=s)) + lims(y=c(-0.25,3), x =c(0,max(window.depths$genome.pos))) + 
     geom_point(color='darkred', alpha=0.5) +

### 2017-04-21 lib/load_patient_metadata.R

@@ -105,7 +105,10 @@ read.patient.info<-function(file, set='Training') {
   
   patient.info$SLX.ID = gsub('SLX-', '', strip.whitespace( patient.info$SLX.ID ) )
   patient.info$Plate.Index = strip.whitespace(patient.info$Plate.Index)
+
+  patient.info[] = lapply(patient.info[], strip.whitespace)
   
+    
   patient.info[c('Status','Pathology','p53.Status')] = 
     lapply(patient.info[c('Status','Pathology','p53.Status')], function(x) factor(strip.whitespace(x)))
   
@@ -136,8 +139,12 @@ read.patient.info<-function(file, set='Training') {
     
   patient.info = merge(patient.info, test.training[,grep('Patient|Set|Analysis', colnames(test.training), value=T)], by.x='Patient', by.y='Patient', all=T)
   
-  patient.info = subset(patient.info, Set == set)
-  message(paste("Returning only the", set, "set", sep=" "))
+  if (!grepl('^all$', set, ignore.case = T)) {
+    patient.info = subset(patient.info, Set == set)
+    message(paste("Returning only the", set, "set", sep=" "))
+  } else {
+    message(paste("Returning all patient data."))
+  }
 
   return(patient.info)
 }

### 2017-04-27 Patient Report.Rmd
Moved pre-processing pipeline code

@@
 
-patient.info = read.patient.info(patient.file)
+dataset = 'Training'
+patient.info = read.patient.info(patient.file, set=dataset)
 patient.info$Patient = gsub("/", "_", patient.info$Patient)
@@
 ```
+
+# Description of '`r dataset`' data set
 
@@
 
-save(patient.data, file='patients.Rdata')
+save(patient.data, file=paste(dataset, '_patients.Rdata', sep=''))
 
+example.pt = subset(sum.patient.data, Status == 'P' & total.samples > 2)$Patient[1]
 ## Note the number of probes per patient/sample
@@
 \pagebreak
-
-# Data
 
@@
 
-1. Segmented values 
+1. Segmented values (progressor): `r example.pt` 
 ```{r}
-head(patient.data[['AD0258']]$seg.vals)
+head(patient.data[[example.pt]]$seg.vals)
 ```
@@
 ```{r echo=F, warning=F, message=F, fig.height=10, fig.width=10}
-ggplot( melt(patient.data[['AD0258']]$seg.vals[,-(1:5)]), aes(x=value) ) + 
+ggplot( melt(patient.data[[example.pt]]$seg.vals[,-(1:5)]), aes(x=value) ) + 
         facet_wrap(~variable,scales = "free_x") + 
-        geom_histogram( fill=I("blue"),col=I("lightblue"), binwidth=0.03) + labs(title='AD0258 segmented value distribution', x='', y='')
+        geom_histogram( fill=I("blue"),col=I("lightblue"), binwidth=0.03) + labs(title=paste(example.pt,' segmented value distribution'), x='', y='')
 ```
@@
 ```{r}
-head(patient.data[['AD0258']]$norm.seg.vals)
+head(patient.data[[example.pt]]$norm.seg.vals)
 ```
@@
 ```{r echo=F, warning=F, message=F, fig.height=10, fig.width=10}
-ggplot( melt(patient.data[['AD0258']]$norm.seg.vals), aes(x=value) ) + 
+ggplot( melt(patient.data[[example.pt]]$norm.seg.vals), aes(x=value) ) + 
         facet_wrap(~variable,scales = "free_x") + 
-        geom_histogram( fill=I("blue"),col=I("lightblue"), binwidth=0.3) + labs(title='AD0258 normalized segmented value distribution', x='', y='')
+        geom_histogram( fill=I("blue"),col=I("lightblue"), binwidth=0.3) + labs(title=paste(example.pt,' normalized segmented value distribution'), x='', y='')
 
@@
 
-  do.call('grid.arrange', list(g1,g2, top=paste(name, status, sep=" - ")))
+  do.call('grid.arrange', list(g1,g2, top=paste(name, status, sep=" - "), ncol=2))
 })
@@
 
-prog.regions = sapply(var.seg.lengths[sum.patient.data$Patient], nrow)
+n.var.regions = sapply(var.seg.lengths, nrow)
+wilcox.test(n.var.regions[prog.pt], n.var.regions[np.pt])
 
@@
 chrs = read.table('hg19_chrom.txt',sep='\t', header=T)
+
+# Segment lengths, variable regions
 var.seg.lengths = lapply(var.seg.lengths, function(x) merge(x, chrs, by.x='chr', by.y='chr') )
 
-segment.lengths= lapply(segment.lengths, function(x) merge(x, chrs, by.x='chr', by.y='chr') )
+# All segments
+segment.lengths = lapply(segment.lengths, function(x) merge(x, chrs, by.x='chr', by.y='chr') )
 
-opcov = sapply(segment.lengths[prog.pt], function(x) {
+# Genome coverage
+genome.coverage = sapply(segment.lengths, function(x) {
   sum(as.numeric(x[['seg.len']]))/sum(as.numeric(chrs$length))
@@
 
-onpcov = sapply(segment.lengths[np.pt], function(x) {
-  sum(as.numeric(x[['seg.len']]))/sum(as.numeric(chrs$length))
-})
+# No difference in coverage
+covtt = wilcox.test(genome.coverage[prog.pt], genome.coverage[np.pt])
 
+# Variable region length by ratio of chromosome
 chr.ratios = lapply(var.seg.lengths, function(x) x[['seg.len']]/x[['length']])
-# Segment coverage, variable regions
-ttR = t.test(unlist(chr.ratios[subset(sum.patient.data, Status == 'P')$Patient]), 
-             unlist(chr.ratios[subset(sum.patient.data, Status == 'NP')$Patient]))
+
+melted = melt(chr.ratios)
+melted = merge(sum.patient.data[,c('Patient', 'Status')], melted, by.x='Patient', by.y='L1')
+
+chr.ratio.tt = t.test(unlist(chr.ratios[prog.pt]), unlist(chr.ratios[np.pt]))
+
+ggplot(melted, aes(Status, value, fill=Status)) + geom_boxplot() + 
+  labs(title=paste("Variable region length, p-value", ifelse(chr.ratio.tt$p.value < 0.001, "<0.001", signif(chr.ratio.tt$p.value,2))),y='Segment Length/Chr length')
 
@@
 
+# Genome coverage of the variable regions
 pcov = median(sapply(var.seg.lengths[prog.pt], function(x) 
@@
 ## Grouping the chromosomes by length 1-12, 13-22
-#t.test(subset(np.regions, chrom %in% c(1:12))$chr.ratio, subset(np.regions, chrom %in% c(13:22))$chr.ratio)
+get.chrs<-function(x, chrs) { subset(x, chr %in% chrs)$chr.ratio }
 
-#t.test(subset(prog.regions, chrom %in% c(1:12))$chr.ratio, subset(prog.regions, chrom %in% c(13:22))$chr.ratio)
+long.chr.tt = t.test(unlist(lapply(var.seg.lengths[prog.pt], get.chrs, chrs=c(1:12))), unlist(lapply(var.seg.lengths[np.pt], get.chrs, chrs=c(1:12))))
+short.chr.tt = t.test(unlist(lapply(var.seg.lengths[prog.pt], get.chrs, chrs=c(13:22))), unlist(lapply(var.seg.lengths[np.pt], get.chrs, chrs=c(13:22))))
 
@@
 
-So it appears that adjusting for chromosome length (looking at the ratio of the chromosome covered by segments rather than the raw bp) a significant difference in the coverage of the genome by CN segments that are variable between progressors and non: t-test p-value `r signif(ttR$p.value, 2)`.
-Overall bp coverage of the sWGS is `r round(mean(c(opcov, onpcov)), 2)*100`%  (P `r round(mean(opcov),2)`, NP `r round(mean(onpcov), 2)`).
+So it appears that adjusting for chromosome length (looking at the ratio of the chromosome covered by segments rather than the raw bp) a significant difference in the coverage of the genome by CN segments that are variable between progressors and non: t-test p-value `r signif(chr.ratio.tt$p.value, 2)`.
+Overall bp coverage of the sWGS is `r round(mean(genome.coverage), 2)*100`%  (P `r round(mean(genome.coverage[prog.pt]),2)`, NP `r round(mean(genome.coverage[np.pt]), 2)`).
 
@@
 gene.cnt = arrange(gene.cnt, -Progressors)
-gene.p = subset(gene.cnt, Progressors >= Non.Progressors)
+## RAD51B is screwing up the plot simply because it shows 16 non-progressors (8 progressors) with a gain but the next closest is 9 so we'll just filter that out for now
+gene.cnt = subset(gene.cnt, gene.cnt$Gene.Symbols != 'RAD51B')
+
+#gene.p = subset(gene.cnt, Progressors >= Non.Progressors)
+gene.p = gene.cnt
 
@@
 
-p = ggplot(m, aes(y=Gene.Symbols, x=variable)) + 
+ggplot(m, aes(y=Gene.Symbols, x=variable)) + 
   geom_tile(aes(fill=value), color='white') + 
-  scale_fill_gradient('n patients',low='white',high='royalblue4') + labs(x="",y='CGC Genes')
-p + geom_point(data=m[m$CN == 'gain' & m$value > 0,], aes(y=Gene.Symbols, x=variable), size=2, shape=24, fill='red', color='red') +
-  geom_point(data=m[m$CN == 'loss' & m$value > 0,], aes(y=Gene.Symbols, x=variable), color='green4', size=2, shape=25, fill='green4')
+  scale_fill_gradient('n patients',low='white',high='red') + labs(x="",y='CGC Genes')
+# + geom_point(data=m[m$CN == 'gain' & m$value > 0,], aes(y=Gene.Symbols, x=variable), size=2, shape=24, fill='red', color='red') +
+#  geom_point(data=m[m$CN == 'loss' & m$value > 0,], aes(y=Gene.Symbols, x=variable), color='green4', size=2, shape=25, fill='green4')
 
-#gc = gene.cnt %>% group_by(Gene.Symbols) %>% summarise(Progressors=sum(Progressors), Non.Progressors=sum(Non.Progressors))
-#gc = arrange(subset(gc, Progressors >= Non.Progressors), -Progressors)
-#m = melt(gc, id.vars=c('Gene.Symbols'))
-#m$Gene.Symbols = factor(m$Gene.Symbols, levels= m$Gene.Symbols[c(nrow(m):1)] ) # order
-#ggplot(m, aes(y=Gene.Symbols, x=variable)) + 
-#  geom_tile(aes(fill=value), color='white') + 
-#  scale_fill_gradient('n patients',low='white',high='royalblue4') + labs(x="",y='CGC Genes')
+gc = gene.cnt %>% group_by(Gene.Symbols) %>% summarise(Progressors=sum(Progressors), Non.Progressors=sum(Non.Progressors))
+gc = arrange(subset(gc, Progressors >= Non.Progressors), -Progressors)
+gc = gc[1:60,]
+m = melt(gc, id.vars=c('Gene.Symbols'))
+#m$value = as.factor(m$value)
+m$Gene.Symbols = factor(m$Gene.Symbols, levels= m$Gene.Symbols[c(nrow(m):1)] ) # order
+ggplot(m, aes(y=Gene.Symbols, x=variable)) + 
+  geom_tile(aes(fill=value), color='white') + 
+  scale_fill_gradient('n patients',low='white',high='red') + labs(x="",y='CGC Genes') + ggtitle("Top 50 genes only")
+
 ```

### 2017-04-27 RegressionPvNP.Rmd
Moved pre-processing pipeline code

@@
 sum.patient.data = as.data.frame(subset(sum.patient.data, Patient %in% names(patient.data))) ## For now
-pander(sum.patient.data[,c('Patient', 'Status', 'start.year', 'end.year','total.samples','highest.path')])
+pander(sum.patient.data[,c('Patient', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')])
 
@@
 ```{r xvalpt, message=F, warning=F, echo=F, fig.height=6, fig.width=6}
-precisionRecall<-function(actual, predicted) {
+precisionRecall<-function(actual, predicted) { ## needs to calculate for both 0 and 1
     retrieved <- sum(predicted==1)
     recall <- sum(predicted==1 & actual==1) / sum(actual==1)
+    sum(predicted==0 & actual==0) / sum(actual==0)
+
     precision <- sum(predicted==1 & actual==1) / retrieved
@@
 
-cv.patient.glmnet<-function(x,y,lambda,patients) {
-  message(paste("Running", length(unique(patients$group)), "folds"))
+cv.patient.glmnet<-function(x,y,lambda,pts) {
+  message(paste("Running", length(unique(pts$group)), "folds"))
 
   cv.class = matrix(nrow=5, ncol=length(lambda))
-  auc = matrix(nrow=5, ncol=length(lambda))
+  f1 = matrix(nrow=5, ncol=length(lambda))
+  prec = f1; recall = f1
 
@@
     message(paste(i, "fold"))
-    test.rows = which(rownames(x) %in% subset(patients, group == i)$Samplename)
+    test.rows = which(rownames(x) %in% subset(pts, group == i)$Samplename)
     test = x[test.rows,]
@@
     pred.class <- predict(fit, test, type='class') # check this later
-    auc[i,] = apply(pred.class, 2, function(pd) {
+
+    proc = do.call(rbind, apply(pred.class, 2, function(pd) {
       if (sum(as.numeric(pd)) <= 0) return(0)
       #as.numeric(pROC::auc(pROC::roc(response=as.numeric(pd), predictor=labels[test.rows])))  
-      precisionRecall(labels[test.rows], as.numeric(pd))$F1
-    })
+      precisionRecall(labels[test.rows], as.numeric(pd))
+    }))
+
+    f1[i,] = proc$F1
+    prec[i,] = proc$precision
+    recall[i,] = proc$recall
 
@@
           geom_hline(yintercept=df[which.max(df$mean)-1,'mean'], colour='grey') +
-          annotate("text", x=5, y=max(df$mean)+0.0005, label=round(df[which.max(df$mean)-1,'mean'],2)) +
+          annotate("text", x=5, y=max(df$mean)+0.005, label=round(df[which.max(df$mean)-1,'mean'],2)) +
+          annotate("text", x=6, y=max(df$mean)-0.005, label=round(median(df$sme),4)) +
           theme(axis.title.x = element_text(size = 15, vjust=-.2)) + scale_colour_discrete(name = "Best lambda values") +
@@
 
-  return(list('cv.class'=cv.class, 'f1'=auc[which.max(df$mean)-1], 'lambda.opt'=lambda.opt, 'lambdas'=df, 'plot'=gp, 'non.zero.cf'=nzcf))
+  return(list('max.cm'=df[which.max(df$mean)-1,'mean'], 'f1'=mean(f1[,which.max(df$mean)-1]), 'precision'=mean(prec[,which.max(df$mean)-1]), 'recall'=mean(recall[,which.max(df$mean)-1]),
+              'lambda.opt'=lambda.opt, 'lambdas'=df, 'plot'=gp, 'non.zero.cf'=nzcf))
 }
@@
 
-grid.arrange(cv.patient$plot)
+grid.arrange(cv.patient$plot + ggtitle(paste("CV Performance, 5-fold xval. mean f1 at opt lambda:", cv.patient$f1)))
 
 ## check how often the feature(s) is selected at that lamda in each split. "stability selection"
+coef.stability<-function(opt, nz.list) {
+  for (i in 1:length(nz.list)) {
+    df = as.data.frame(nz.list[[i]])
+    colnames(df) = i
+    opt = merge(opt, df, by='row.names', all.x=T)
+    rownames(opt) = opt$Row.names
+    opt$Row.names = NULL
+    opt[,(i+1)] = as.integer(!is.na(opt[,(i+1)]))
+  }
+  opt = opt[order(sapply(rownames(opt), function(x) as.numeric(unlist(strsplit(x, ':'))[1])  )),]
+  return(opt)
+}
 
-for (i in 1:length(cv.patient$non.zero.cf)) {
-  df = as.data.frame(cv.patient$non.zero.cf[[i]])
-  colnames(df) = i
-  coef.opt = merge(coef.opt, df, by='row.names', all.x=T)
-  rownames(coef.opt) = coef.opt$Row.names
-  coef.opt$Row.names = NULL
-  coef.opt[,(i+1)] = as.integer(!is.na(coef.opt[,(i+1)]))
-}
-coef.opt = coef.opt[order(sapply(rownames(coef.opt), function(x) as.numeric(unlist(strsplit(x, ':'))[1])  )),]
+coef.opt.stability = coef.stability(coef.opt, cv.patient$non.zero.cf)
 
@@
 At a the minimum lambda-1se (green dot) the value is `r signif(lambda.opt, 3)`. In the initial model `r length(coef.opt)` coefficients have a non-zero value.
+
+### Feature stability
+
+`r length(which(rowSums(coef.opt.stability[,2:6])>= 3))` features are stable in 3:5 splits
+
+`r length(which(rowSums(coef.opt.stability[,2:6])>= 4))` features are stable in 4:5 splits
+
+`r length(which(rowSums(coef.opt.stability[,2:6])>= 5))` features are stable in 5:5 splits
+
+Optimal model?
+```{r}
+
+gfit = glm(labels~.,data=cbind.data.frame(dysplasia.df[,names(non.zero.coef(fit0, lambda.opt))], labels), family=binomial(link='logit'))
+autoplot(gfit)
+
+```
 
@@
 #lambda.opt = cv.patient$lambda.opt
-grid.arrange(cv.nohgd$plot + ggtitle("CV Performance, No HGD/IMC samples"))
-#coef.opt = as.data.frame(non.zero.coef(fit0, lambda.opt))
+grid.arrange(cv.nohgd$plot + ggtitle(paste("CV Performance, No HGD/IMC samples. mean f1:", cv.nohgd$f1)))
+nohgd.stab = coef.stability(coef.opt, cv.nohgd$non.zero.cf)
 
@@
 cv.nolgd = cv.patient.glmnet(x=dysplasia.df[samples,], y=labels[samples], lambda, subset(patients, Samplename %in% samples))
-#lambda.opt = cv.patient$lambda.opt
-grid.arrange(cv.nolgd$plot + ggtitle("CV Performance, No HGD/IMC/LGD samples"))
+grid.arrange(cv.nolgd$plot + ggtitle(paste("CV Performance, No HGD/IMC/LGD samples. mean f1:", cv.nohgd$f1)))
+nolgd.stab = coef.stability(coef.opt, cv.nolgd$non.zero.cf)
 
+```
+
+#### No HGD feature stability
+
+`r length(which(rowSums(nohgd.stab[,2:6])>= 3))` features are stable in 3:5 splits
+
+`r length(which(rowSums(nohgd.stab[,2:6])>= 4))` features are stable in 4:5 splits
+
+`r length(which(rowSums(nohgd.stab[,2:6])>= 5))` features are stable in 5:5 splits
+
+#### No LGD feature stability
+
+`r length(which(rowSums(nolgd.stab[,2:6])>= 3))` features are stable in 3:5 splits
+
+`r length(which(rowSums(nolgd.stab[,2:6])>= 4))` features are stable in 4:5 splits
+
+`r length(which(rowSums(nolgd.stab[,2:6])>= 5))` features are stable in 5:5 splits
+
+### Leave one patient out
+
+```{r loop, echo=F, message=F, warning=F, fig.height=40, fig.width=30}
+fpr = data.frame(matrix(ncol=5,nrow=0))
+# Randomize patients
+pt.samp = sample(sum.patient.data$Patient)
+
+#plots = list()
+
+for (pt in pt.samp) {
+  samples = as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Patient != pt)$Patient], function(pt) pt$info$Samplename)))
+
+  cv.back = cv.patient.glmnet(x=dysplasia.df[samples,], y=labels[samples], lambda, subset(patients, Samplename %in% samples))
+#  print(median(cv.back$lambdas$sme))
+ # print( paste(cv.back$f1, cv.back$precision) )
+   fpr = rbind(fpr, c(cv.back$max.cm, cv.back$f1, cv.back$precision, cv.back$recall, length(samples)))
+
+#  plots[[ pt ]] = cv.back$plot + ggtitle(paste("# Samples:", length(samples))) + theme(legend.position = "none")
+}
+colnames(fpr) = c('performance','f1','prec','recall','n.samples')
+
+ggplot(fpr, aes(performance, f1)) + 
+  geom_point() + 
+  ylim(0,1) + xlim(0,1) + labs(title="LOO (patient)", x='CV Performance', y='f1 statistic')
+
+```
+
+## Peel back timepoints
+
+```{r, echo=F, message=F, warning=F}
 get.next.samples<-function(patient.data, i) {
@@
 
+#subset(sum.patient.data, Status == 'P' & total.endos >= 8)
+
+#cv.back = cv.patient.glmnet(x=dysplasia.df[samples,], y=labels[samples], lambda, subset(patients, Samplename %in% samples))
+#print(cv.back$f1)
+
 nonps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'NP')$Patient], 0)
-plots = list()
-for (i in seq(1,35,5)) {
+#plots = list()
+fpr = data.frame(matrix(ncol=4,nrow=0))
+for (i in c(0,1,seq(2,25,3))) {
   ps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'P')$Patient], i)
+  samples = get.next.samples(patient.data, i)
+  #samples = c(nonps, ps)
+  cv.back = cv.patient.glmnet(x=dysplasia.df[samples,], y=labels[samples], lambda, subset(patients, Samplename %in% samples))
+  print( paste(cv.back$f1, cv.back$precision) )
+  fpr = rbind(fpr, c(cv.back$f1, cv.back$precision, cv.back$recall, length(ps)))
+#  plots[[ as.character(i) ]] = cv.back$plot + ggtitle(paste(length(ps), "progressor samples. mean f1:", cv.back$f1))
+}
+colnames(fpr) = c('f1','prec','recall','n.samples')
+pander(fpr[1:5,], justify='left')
+#do.call('grid.arrange', c(plots, ncol=2, top="Peeling back timepoints in progressors"))
+
+ps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'P')$Patient], 11)
 samples = c(nonps, ps)
   cv.back = cv.patient.glmnet(x=dysplasia.df[samples,], y=labels[samples], lambda, subset(patients, Samplename %in% samples))
-  plots[[ as.character(i) ]] = cv.back$plot + ggtitle(paste("CV Performance - ", length(ps), "progressor samples"))
-}
+  print(median(cv.back$lambdas$sme))
+  cv.back$plot + ggtitle(paste("CV Performance - ", length(ps), "progressor samples"))
 
-do.call('grid.arrange', c(plots, ncol=2, top="Peeling back timepoints in progressors"))
+info = subset(patient.info, Samplename %in% samples)
+head(info)
+
+pander(summarise.patient.info(subset(info, Status == 'P'))[,c(1:7)], justify='left')
+
+coef.opt.stability = coef.stability(coef.opt, cv.back$non.zero.cf)
+
+pander(coef.opt.stability[which(rowSums(coef.opt.stability[,2:6]) > 0),], justify='left', caption="Features initially identified that also are found in the subset of samples")
 

### 2017-05-05 RegressionPvNP.Rmd
…

--- 2017-04-21 RegressionPvNP.Rmd
@@
 library(pander)
-library(Hmisc)
+#library(Hmisc)
 library(reshape2)
@@
 
-load('patients.Rdata', verbose=T)
+load("Test_patients.Rdata", verbose=T)
+validation.patient.data = patient.data
+
+dataset = 'Training'
+load(paste(dataset, '_patients.Rdata', sep=''), verbose=T)
 
@@
 ## Patient info file
-patient.file = grep('All_patient_info.xls', data.files, value=T)
+patient.file = grep('All_patient_info.txt', data.files, value=T)
 if (length(patient.file) != 1)
@@
 
-patient.info = read.patient.info(patient.file)
+patient.info = read.patient.info(patient.file, set=dataset)
 patient.info$Patient = gsub("/", "_", patient.info$Patient)
 head(patient.info)
+
+validation.patient.info = read.patient.info(patient.file, set='Test')
+sum.validation.info = summarise.patient.info(validation.patient.info)
 
@@
 ```{r apclust-func, echo=F, message=F, warning=F}
-suppressPackageStartupMessages( library(apcluster) )
+#suppressPackageStartupMessages( library(apcluster) )
 
@@
 
-apclust<-function(segdata) {
-  x1 = apclust.data(segdata)
-  # q=0 minimize off-diagonal similarity
-  ac = apcluster(negDistMat(r=2), x1, details=T, convits=25, q=0)
-  return(list('apres'=ac, 'data'=x1))
-}
+# apclust<-function(segdata) {
+#   x1 = apclust.data(segdata)
+#   # q=0 minimize off-diagonal similarity
+#   ac = apcluster(negDistMat(r=2), x1, details=T, convits=25, q=0)
+#   return(list('apres'=ac, 'data'=x1))
+# }
 
-net.similarity<-function(ac) {
-  if (is.list(ac))
-    ac = ac$apres
-  ac@netsim
-}
-
-sum.similarity<-function(ac) {
-  if (is.list(ac))
-    ac = ac$apres
-  ac@dpsim
-}
-
-clusters<-function(ac) {
-  if (is.list(ac))
-    ac = ac$apres
-  ac@exemplars
-}
-
-plotAC<-function(aclist) {
-  plot(aclist$apres, aclist$data)
-}
-
-heatmapAC<-function(aclist) {
-  heatmap(aclist$apres)
-}
-
-# First vs last samples
-#x1 = apclust.data(patient.data[[pt]]$seg.vals[,c(1:4, 6, ncol(patient.data[[pt]]$seg.vals))], patient.data[[pt]]$info$Samplename)
 ```
@@
 
-tileFile = paste(as.character(tile.w),'_tiledpts.Rdata', sep='')
+tile.sample.segments<-function(tileFile, sample.data) {
   if (file.exists(tileFile)) {
@@
   } else {
-  grList = lapply(patient.data, function(df) {
+    grList = lapply(sample.data, function(df) {
       df$info = arrange(df$info, Endoscopy.Year, Pathology)
@@
       x1 = apclust.data(df$seg.vals, df$info$Samplename)
-    head(x1)
 
@@
       pt = unique(df$info$Patient)
-    return( makeGRangesFromDataFrame(cbind(segnames,pt, x1), keep.extra.columns = T) )
+      return( makeGRangesFromDataFrame(cbind(segnames,pt,x1), keep.extra.columns=T) )
     })
 
-  sampleNames = unlist(sapply(patient.data, function(df) {
+    sampleNames = unlist(sapply(sample.data, function(df) {
       df$info = arrange(df$info, Endoscopy.Year, Pathology)
@@
       for (qh in unique(queryHits(ov))) {
-      print(qh)
+        #print(qh)
         currentTile = tiles[[qh]]
@@
     for (gr in grList) {
-    print(unique(gr$pt))
+      #print(unique(gr$pt))
       mergedDf = tile.segments(tiles, gr, mergedDf)
@@
   }
+  return(mergedDf)
+}
+
+mergedDf = tile.sample.segments(paste(as.character(tile.w),'_tiledpts.Rdata', sep=''), patient.data)
+
+#validatioMergedDf = tile.sample.segments(paste(as.character(tile.w),'validation_tiledpts.Rdata', sep=''), validation.patient.data)
 
@@
 ## binomial: dysplasia 1, BE 0
-
 labels = unlist(sapply(patient.data, function(df) {
@@
 
-With all `r ncol(dysplasia.df)` regions of the genome, first check if/which regression may be appropriate.
+```{r cvglmnetPNP, echo=F, warning=F, message=F, eval=F}
+#With all `r ncol(dysplasia.df)` regions of the genome, first check if/which regression may be appropriate.
 It appears that pure lasso may provide the best balance for number of features with non-zero coefficients.
 
-```{r cvglmnetPNP, echo=F, warning=F, message=F}
 plot.multi.cv.glmnet<-function(x,y,alpha=c(0,1), family="binomial", title='cv.glmnet') {
@@
 
-## xval by patients
+## Build/Cross-validate by patient
 
-Split into 5ths by patient, fit and predict at each split, select min lambda, get coefficients from the initial model using that value.
+Keeping in mind that the matrix is built on a sample, not patient, basis while the labels (Progressor/Non) are on a per-patient basis I generated the cross validated models for various values of alpha by running 10 or more iterations of 5-fold cross validation that pulled out all the samples for 5 patients at each fold.
+1000 lambda values.
 
-```{r xvalpt, message=F, warning=F, echo=F, fig.height=6, fig.width=6}
-precisionRecall<-function(actual, predicted) {
-    retrieved <- sum(predicted==1)
-    recall <- sum(predicted==1 & actual==1) / sum(actual==1)
-    precision <- sum(predicted==1 & actual==1) / retrieved
-    f1 = signif(2*(1/(1/precision+1/recall)), 2)
-    return(data.frame('precision'=precision, 'recall'=recall, 'F1'=signif(f1, 2)))
-  }
-
+```{r cv.funcs, message=F, warning=F, echo=T, fig.height=6, fig.width=6}
 pi.hat<-function(x) exp(x)/(1+exp(x))
@@
 
-cv.patient.glmnet<-function(x,y,lambda,patients) {
-  message(paste("Running", length(unique(patients$group)), "folds"))
+create.patient.sets<-function(pts, n, splits, minR=0.2) {
+  # This function just makes sure the sets don't become too unbalanced with regards to the labels.
+  check.sets<-function(df, grpCol, min) {
+    sets = table(cbind.data.frame('set'=df[[grpCol]], 'labels'=labels[df$Samplename]))
+    while ( (length(which(sets/rowSums(sets) < minR) ) >= 2 | length(which(sets/rowSums(sets) == 0)) > 0) ) {
+      #print(sets/rowSums(sets))
+      s = sample(rep(seq(5), length = length(unique(df$Patient))))
+      df2 = merge(df, cbind('Patient'=unique(df$Patient), 'tmpgrp'=s), by="Patient")
+      df[[grpCol]] = df2$tmpgrp
+      sets = table(cbind.data.frame('set'=df[[grpCol]], 'labels'=labels[df$Samplename]))
+    }
+  return(df[[grpCol]])
+  }
 
-  cv.class = matrix(nrow=5, ncol=length(lambda))
-  auc = matrix(nrow=5, ncol=length(lambda))
+  s = sample(rep(seq(splits), length = length(unique(pts$Patient))))
+  patients = merge(pts, cbind('Patient'=unique(pts$Patient), 'group'=s), by="Patient")  
+  colnames(patients)[3] = c('fold.1')
+  patients$fold.1 = check.sets(patients, 'fold.1', minR)
 
+  for (i in 2:n) {
+    s = sample(rep(seq(5), length = length(unique(pts$Patient))))
+    patients = merge(patients, cbind('Patient'=unique(patients$Patient), 'group'=s), by="Patient")  
+    foldcol = grep('group',colnames(patients))
+    colnames(patients)[foldcol] = paste('fold',i,sep='.')
+    patients[[paste('fold',i,sep='.')]] = check.sets(patients, paste('fold',i,sep='.'))
+  }
+  return(patients)
+}
+
+crossvalidate.by.patient<-function(x,y,lambda,pts,a=1,nfolds=10, splits=5, fit=NULL, minR=0.2) {
+  message(paste("Running", splits, "splits",nfolds,"times on", paste(dim(x), collapse=':'), 'alpha=',a ))
   fit.e = list()
-  for (i in 1:length(unique(patients$group))) {  ## This is not stable. 
-    message(paste(i, "fold"))
-    test.rows = which(rownames(x) %in% subset(patients, group == i)$Samplename)
+  pts = create.patient.sets(pts, nfolds, splits, minR)
+  cv.pred = (matrix(nrow=0, ncol=length(lambda)))
+  for (n in 1:nfolds) {  
+    message(paste(n, "fold"))
+    setCol = grep(paste('^fold.',n,'$',sep=''), colnames(pts))
+
+    cv.class = matrix(nrow=splits, ncol=length(lambda))
+    for (i in 1:splits) { 
+      message(paste(i, "split"))
+      test.rows = which(rownames(x) %in% pts[which(pts[,setCol] == i), 'Samplename'])
       test = x[test.rows,]
@@
       # pre-spec lambda seq
-    fit <- glmnet(training, y[-test.rows], lambda=lambda, family='binomial', alpha=1) 
+      fit <- glmnet(training, y[-test.rows], lambda=lambda, family='binomial', alpha=a) 
       # autoplot(fit) + theme(legend.position="none")
-
-    pred.class <- predict(fit, test, type='class') # check this later
-    auc[i,] = apply(pred.class, 2, function(pd) {
-      if (sum(as.numeric(pd)) <= 0) return(0)
-      #as.numeric(pROC::auc(pROC::roc(response=as.numeric(pd), predictor=labels[test.rows])))  
-      precisionRecall(labels[test.rows], as.numeric(pd))$F1
-    })
-
       # careful: matrix
-    pred <- pi.hat(predict(fit, test)) 
+      pred <- pi.hat(predict(fit, test, type='link')) 
       cv.class[i,] = apply(pred, 2, function(p) {
-      # Confusion matrix
-       #cmat = table(labels[test.rows], p>0.5)
-       #sum(diag(cmat))/sum(cmat)
-      # quantitiative, pos results + neg results / number of test rows
+        # Confusion matrix: quantitiative, pos results + neg results / number of test rows
         (p%*%labels[test.rows] + (1-p) %*% (1-labels[test.rows]))/ length(test.rows)
@@
 
-    fit.e[[i]] = fit
+      fit.e[[length(fit.e)+1]] = fit
+    }
+  cv.pred = rbind(cv.pred, cv.class)    
   }
 
-  df = cbind.data.frame('lambda.at'=1:ncol(cv.class),'mean'=colMeans(cv.class), 
-                        'sme'=apply(cv.class, 2, sd)/5, 
-                        'sd'=apply(cv.class, 2, sd), 
-                        'lambda'=lambda)
+  df = cbind.data.frame('lambda.at'=1:ncol(cv.pred),
+                        'mean'= colMeans(cv.pred), 
+                        'sme'= apply(cv.pred, 2, sd)/sqrt(nrow(cv.pred)), 
+                        'sd'= apply(cv.pred, 2, sd), 
+                        'lambda'= lambda)
 
-  lambda.opt = lambda[which.max(df$mean)-1]
-  nzcf = lapply(fit.e, non.zero.coef, s=lambda.opt)
+  df2 = df
+  df2$lambda = log(lambda)
+  df2$mean = 1-df2$mean
 
-  gp = ggplot(df, aes(y=mean,x=lambda.at)) + geom_point() + 
-          geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + labs(y='mean of confusion matrix (P)') +
-          geom_point(data=subset(df, mean == max(mean)), aes(y=mean, x=lambda.at, colour="max"), size=3 ) +
-          geom_point(data=df[which.max(df$mean)+1,], aes(y=mean, x=lambda.at, colour="max+1"), size=3) +
-          geom_point(data=df[which.max(df$mean)-1,], aes(y=mean, x=lambda.at, colour="max-1"), size=3) +
-          geom_hline(yintercept=df[which.max(df$mean)-1,'mean'], colour='grey') +
-          annotate("text", x=5, y=max(df$mean)+0.0005, label=round(df[which.max(df$mean)-1,'mean'],2)) +
-          theme(axis.title.x = element_text(size = 15, vjust=-.2)) + scale_colour_discrete(name = "Best lambda values") +
-          labs(title="CV Performance")
+  df2$lambda.min = df2$mean == min(df2$mean)
+  df2$lambda.1se = FALSE
+  df2[df2$lambda < subset(df2, mean == min(mean))$lambda-sd(df2$lambda),][1,'lambda.1se'] = TRUE
 
-  return(list('cv.class'=cv.class, 'f1'=auc[which.max(df$mean)-1], 'lambda.opt'=lambda.opt, 'lambdas'=df, 'plot'=gp, 'non.zero.cf'=nzcf))
+  lambda.min = lambda[subset(df2, lambda.min == T)$lambda.at]
+  lambda.1se = lambda[subset(df2, lambda.1se == T)$lambda.at]
+  nzcf = lapply(fit.e, non.zero.coef, s=lambda.min)
+
+  gp = ggplot(df2, aes(y=mean,x=lambda)) + geom_point() + 
+    geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + 
+    geom_point(data=subset(df2, lambda.min == T), aes(y=mean, x=lambda, colour="lambda.min"), size=2 ) +          
+    geom_vline(xintercept = subset(df2, lambda.min == T)$lambda, colour='grey') +
+    annotate("text", x=subset(df2, lambda.min == T)$lambda, y=0.35, 
+             label=round(df[subset(df2, lambda.min == T)$lambda.at,'mean'],3))
+
+  if (length(lambda.1se) > 0) {
+    gp = gp + geom_point(data=subset(df2, lambda.1se == T), aes(y=mean, x=lambda, colour='lambda.1se'), size=2 ) +
+      geom_vline(xintercept = subset(df2, lambda.1se == T)$lambda, colour='grey') +
+      annotate("text", x=subset(df2, lambda.1se == T)$lambda, y=0.35, 
+               label=round(df[subset(df2, lambda.1se == T)$lambda.at,'mean'],3)) 
+  }
+   gp = gp + theme( legend.position='bottom') + 
+          scale_colour_discrete(name = "") +
+          labs(title=paste(splits,' splits, ',nfolds,' folds, alpha=',a, sep=''), y='1-mean (P)', x='log(lambda)')
+
+  if (!is.null(fit)) {
+    df2$nzcoef = sapply(df$lambda, function(l) length(non.zero.coef(fit, l)))
+
+    coef.min = subset(df2, lambda.min == T)$nzcoef
+    coef.1se = subset(df2, lambda.1se == T)$nzcoef
+    d = coef.1se-coef.min
+
+    coef.text = arrange(subset(df2,nzcoef %in% c(0,  coef.min, coef.1se, max(nzcoef) )), -lambda.min, -lambda.1se)
+    coef.text = arrange(coef.text[!duplicated(coef.text$nzcoef),], -nzcoef)
+
+    gp = gp + annotate("text", y=max(df2$mean), x=coef.text$lambda, label=coef.text$nzcoef)
   }
 
-fit0 <- glmnet(dysplasia.df, labels, alpha = 1) # all patients
-autoplot(fit0, xvar='lambda', main='fit0, all samples') + theme(legend.position='none') 
-lambda <- fit0$lambda
-
-# split patients in 5ths
-s = {set.seed(2); sample(rep(seq(5), length = length(unique(pts$Patient))))}
-patients = merge(pts, cbind('Patient'=unique(pts$Patient), 'group'=s), by="Patient")
-
-# This just makes sure the sets don't become too unbalanced with regards to the labels.
-sets = table(cbind(patients, labels[patients$Samplename])[,c(3:4)])
-while (length(which(sets/rowSums(sets) < 0.35)) >= 2 | length(which(sets/rowSums(sets) == 0)) > 0 ) {
-  print(".")
-  s = sample(rep(seq(5), length = length(unique(pts$Patient))))
-  patients = merge(pts, cbind('Patient'=unique(pts$Patient), 'group'=s), by="Patient")
-  sets = table(cbind(patients, labels[patients$Samplename])[,c(3:4)])
+  return(list('max.cm'=df[which.max(df$mean)-1,'mean'], 
+              'lambda.min'=lambda.min, 'lambda.1se'=lambda.1se, 'lambdas'=df, 'plot'=gp, 'non.zero.cf'=nzcf))
 }
 
-cv.patient = cv.patient.glmnet(x=dysplasia.df, y=labels, lambda, patients)
-lambda.opt = cv.patient$lambda.opt
+coef.stability<-function(opt, nz.list) {
+  for (i in 1:length(nz.list)) {
+    df = as.data.frame(nz.list[[i]])
+    colnames(df) = i
+    opt = merge(opt, df, by='row.names', all.x=T)
+    rownames(opt) = opt$Row.names
+    opt$Row.names = NULL
+    opt[,(i+1)] = as.integer(!is.na(opt[,(i+1)]))
+  }
+  opt = opt[order(sapply(rownames(opt), function(x) as.numeric(unlist(strsplit(x, ':'))[1])  )),]
+  return(opt)
+}
+```
 
-coef.opt = as.data.frame(non.zero.coef(fit0, lambda.opt))
+```{r xvalpt, message=F, warning=F, echo=T, fig.height=12, fig.width=12}
+plots = list()
+coefs = list()
+folds = 10; splits=5
+alpha.values = c(0,0.5,0.7,0.8,0.9,1)
+for (a in alpha.values) {
+  fit0 <- glmnet(dysplasia.df, labels, alpha=a, nlambda=1000, family='binomial') # all patients
+  autoplot(fit0, xvar='lambda', main=paste('fit0, all samples, alpha=',a,sep='')) + theme(legend.position='none') 
+  lambda0 <- fit0$lambda
 
-grid.arrange(cv.patient$plot)
+  cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=lambda0, pts=pts, a=a, nfolds=folds, splits=splits, fit=fit0)
+
+  coef.1se = as.data.frame(non.zero.coef(fit0, cv.patient$lambda.1se))
+  coefs[[as.character(a)]] = coef.stability(coef.1se, cv.patient$non.zero.cf)
+
+  plots[[as.character(a)]] = cv.patient$plot + theme(legend.position='bottom') + ggtitle(paste('alpha=',a,sep=''))
+}
+
+do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits', ncol=2))
 
 ## check how often the feature(s) is selected at that lamda in each split. "stability selection"
+coef.stable = lapply( coefs[c('0.8','0.9','1')], function(cf) {
+  sort(rowSums(cf[,-1]), decreasing=T)
+})
 
-for (i in 1:length(cv.patient$non.zero.cf)) {
-  df = as.data.frame(cv.patient$non.zero.cf[[i]])
-  colnames(df) = i
-  coef.opt = merge(coef.opt, df, by='row.names', all.x=T)
-  rownames(coef.opt) = coef.opt$Row.names
-  coef.opt$Row.names = NULL
-  coef.opt[,(i+1)] = as.integer(!is.na(coef.opt[,(i+1)]))
-}
-coef.opt = coef.opt[order(sapply(rownames(coef.opt), function(x) as.numeric(unlist(strsplit(x, ':'))[1])  )),]
+cfs = as.data.frame(matrix(data=0,nrow=length(unique(names(table(unlist(lapply(coef.stable, names)))))), ncol=3, 
+                dimnames=list( unique(names(table(unlist(lapply(coef.stable, names))))), names(coef.stable) )))
+
+for (i in names(coef.stable)) 
+  cfs[intersect(rownames(cfs), names(coef.stable[[i]])), i] = 1
 
@@
 
-`r pander(coef.opt, justify='left', caption='Features from the glm fitted with all data. Each numbered column indicates if the feature was identified in that split.')`
+Values that are closer to full lasso (alpha=1) are pretty similar, however at full lasso we see fewer features than if we regularize it at 0.8.
 
-At a the minimum lambda-1se (green dot) the value is `r signif(lambda.opt, 3)`. In the initial model `r length(coef.opt)` coefficients have a non-zero value.
+### Feature stability
+
+`r pander(as.data.frame(coef.stable[['1']])/(folds*splits), justify='left', caption='Percentage of features from the glm fitted with all data.
+Numbers indicate the total number of folds that feature was identified in.')`
+
+They also share all of the non-zero coefficients (0.8 identifies the most non-zero coefs `r max(sapply(coef.stable, length))`).
+
+`r pander(cfs, justify='left')`
+
+#### Did I understand correctly?
+
+In order to predict, my trained model is then my original model, but using the optimal lambda value?
+This isn't quite correct because the model was trained including these patients, just an example.
+
+```{r}
+fit0 <- glmnet(dysplasia.df, labels, alpha=a, nlambda=200, family='binomial') # all patients
+cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=fit0$lambda, pts=pts, a=1, nfolds=20, splits=5, fit=fit0)
+
+# Progressor, minus the HGD sample(s)
+cbind.data.frame(
+  predict(fit0, dysplasia.df[patient.data$PR1_HIN_042$info$Samplename[-(9:10)],], s=cv.patient$lambda.1se, type='response'),
+  predict(fit0, dysplasia.df[patient.data$PR1_HIN_042$info$Samplename[-(9:10)],], s=cv.patient$lambda.1se, type='class'))
+
+# Non-P, all correctly predicted
+cbind.data.frame(predict(fit0, dysplasia.df[patient.data$AD0610$info$Samplename,], s=cv.patient$lambda.1se, type='response'),
+                 predict(fit0, dysplasia.df[patient.data$AD0610$info$Samplename,], s=cv.patient$lambda.1se, type='class'))
+
+```
 
@@
 
-Just as a point, the initial model is tested against `r nrow(dysplasia.df)` samples with `r ncol(dysplasia.df)` features.
+Just as a point, the initial model is trained against `r nrow(dysplasia.df)` samples with `r ncol(dysplasia.df)` features.  What happens if we retrain it without HGD?
 
-```{r noHGD, echo=F, message=F, warning=F}
+```{r noHGD, echo=T, message=F, warning=F, fig.height=8, fig.width=8}
 `%nin%` <- Negate(`%in%`)
@@
 
+# No HGD/IMC
 samples = intersect(rownames(dysplasia.df), subset(info, Pathology %nin% c('HGD', 'IMC'))$Samplename)
 
-cv.nohgd = cv.patient.glmnet(x=dysplasia.df[samples,], y=labels[samples], lambda, subset(patients, Samplename %in% samples))
-#lambda.opt = cv.patient$lambda.opt
-grid.arrange(cv.nohgd$plot + ggtitle("CV Performance, No HGD/IMC samples"))
-#coef.opt = as.data.frame(non.zero.coef(fit0, lambda.opt))
+plots = list()
+coefs = list()
+for (a in alpha.values) {
+  fit1 <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=1000) # all patients
+  autoplot(fit1, xvar='lambda', main='fit1, no HGD/IMC samples') + theme(legend.position='none') 
+  lambda1 <- fit1$lambda
 
+  cv.nohgd = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=lambda1, pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fit1)
+
+  plots[[as.character(a)]] = cv.nohgd$plot + ggtitle(paste('alpha=',a,sep=''))
+  coef.1se = as.data.frame(non.zero.coef(fit1, cv.nohgd$lambda.1se))
+  coefs[[as.character(a)]] = coef.stability(coef.1se, cv.nohgd$non.zero.cf)
+}
+
+do.call(grid.arrange, c(plots, top="No HGD/IMC samples", ncol=2))
+
+## check how often the feature(s) is selected at that lamda in each split. "stability selection"
+coef.stable = lapply( coefs[c('0.8','0.9','1')], function(cf) {
+  sort(rowSums(cf[,-1]), decreasing=T)
+})
+
+cfs = as.data.frame(matrix(data=0,nrow=length(unique(names(table(unlist(lapply(coef.stable, names)))))), ncol=3, 
+                dimnames=list( unique(names(table(unlist(lapply(coef.stable, names))))), names(coef.stable) )))
+
+for (i in names(coef.stable)) 
+  cfs[intersect(rownames(cfs), names(coef.stable[[i]])), i] = 1
+
+```
+
+#### Feature Stability
+
+`r pander(as.data.frame(coef.stable[['1']])/(folds*splits), justify='left', caption='Percentage of features from the glm fitted with all data.
+Numbers indicate the total number of folds that feature was identified in.')`
+
+They also share all of the non-zero coefficients (`r which.max(sapply(coef.stable, length))` identifies the most non-zero coefs `r max(sapply(coef.stable, length))`).
+
+`r pander(cfs, justify='left')`
+
+### Remove HGD/IMC & LGD Samples
+
+Now remove all LGD samples from the progressor's and retrain.
+
+```{r noLGD, echo=T, message=F, warning=F, fig.height=8, fig.width=8}
+# No LGD
+samples = intersect(rownames(dysplasia.df), c(subset(info, Status == 'NP')$Samplename, subset(info, Pathology %nin% c('HGD', 'IMC', 'LGD') & Status == 'P')$Samplename))
+
+# No HGD/IMC/LGD in all patients
 samples = intersect(rownames(dysplasia.df), subset(info, Pathology %nin% c('HGD', 'IMC', 'LGD'))$Samplename)
-cv.nolgd = cv.patient.glmnet(x=dysplasia.df[samples,], y=labels[samples], lambda, subset(patients, Samplename %in% samples))
-#lambda.opt = cv.patient$lambda.opt
-grid.arrange(cv.nolgd$plot + ggtitle("CV Performance, No HGD/IMC/LGD samples"))
+
+plots = list()
+coefs = list()
+for (a in alpha.values) {
+  fit2 <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=1000) # all patients
+  autoplot(fit2, xvar='lambda', main='fit1, no HGD/IMC samples') + theme(legend.position='none') 
+  lambda2 <- fit2$lambda
+
+  cv.nolgd = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=lambda2, pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fit2)
+
+  plots[[as.character(a)]] = cv.nolgd$plot + ggtitle(paste('alpha=',a,sep=''))
+
+  coef.1se = as.data.frame(non.zero.coef(fit2, cv.nolgd$lambda.1se))
+  coefs[[as.character(a)]] = coef.stability(coef.1se, cv.nolgd$non.zero.cf)
+}
+
+do.call(grid.arrange, c(plots, top="No HGD/IMC/LGD samples", ncol=2))
+
+## check how often the feature(s) is selected at that lamda in each split. "stability selection"
+coef.stable = lapply( coefs[c('0.8','0.9','1')], function(cf) {
+  sort(rowSums(cf[,-1]), decreasing=T)
+})
+
+cfs = as.data.frame(matrix(data=0,nrow=length(unique(names(table(unlist(lapply(coef.stable, names)))))), ncol=3, 
+                dimnames=list( unique(names(table(unlist(lapply(coef.stable, names))))), names(coef.stable) )))
+
+for (i in names(coef.stable)) 
+  cfs[intersect(rownames(cfs), names(coef.stable[[i]])), i] = 1
+
+```
+
+#### Feature Stability
+
+`r pander(as.data.frame(coef.stable[['1']])/(folds*splits), justify='left', caption='Percentage of features from the glm fitted with all data.
+Numbers indicate the total number of folds that feature was identified in.')`
+
+They also share all of the non-zero coefficients (`r which.max(sapply(coef.stable, length))` identifies the most non-zero coefs `r max(sapply(coef.stable, length))`).
+
+`r pander(cfs, justify='left')`
+
+### Leave one patient out
+
+Using the 'all sample' fit, run a LOO (patient)
+
+```{r loop, echo=T, message=F, warning=F, fig.height=6, fig.width=6, eval=F}
+fit0 <- glmnet(dysplasia.df, labels, alpha=1, nlambda=200, family='binomial') # all patients
+
+# Randomize patients
+pt.samp = sample(sum.patient.data$Patient)
+performance.at.1se = c()
+coefs = list()
+for (pt in pt.samp) {
+  samples = as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Patient != pt)$Patient], function(pt) pt$info$Samplename)))
+
+  cv.loo = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fit0$lambda, pts= subset(pts, Samplename %in% samples), a=1, nfolds=20, splits=5, fit=fit0)
+
+  performance.at.1se = c(performance.at.1se, subset(cv.loo$lambdas, lambda == cv.loo$lambda.1se)$mean)
+
+  coef.1se = as.data.frame(non.zero.coef(fit0, cv.loo$lambda.1se))
+  coefs[[as.character(a)]] = coef.stability(coef.1se, cv.loo$non.zero.cf)
+}
+
+ggplot(as.data.frame(performance.at.1se), aes(y=performance.at.1se, x='LOO')) + ylim(0.6,0.8) +
+  geom_boxplot( fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA) + geom_jitter() +
+    labs(y='performance at lambda.1se', x='', title='Performance for Leave One (patient) Out, at alpha=1')
+
+```
+
+## Compare progressors with many samples vs few
+
+```{r, echo=T, message=F, warning=F, eval=F}
+m = median(subset(sum.patient.data, Status == 'P')$total.samples)
+
+nonps = as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Status == 'NP')$Patient], function(df) df$info$Samplename)))
+
+# Few
+samples = c(nonps, as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Status == 'P' & total.samples < m)$Patient], function(pt) pt$info$Samplename))))
+
+fit0 <- glmnet(dysplasia.df[samples,], labels[samples], alpha=1, nlambda=200, family='binomial') # all patients
+autoplot(fit0) + theme(legend.position = 'none')
+
+cv.back.few = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fit0$lambda, pts=subset(pts, Samplename %in% samples), splits=5, nfolds=20, a=1, fit=fit0)
+
+# Many
+samples = c(nonps, as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Status == 'P' & total.samples >= m)$Patient], function(pt) pt$info$Samplename))))
+cv.back.many = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda0, subset(patients, Samplename %in% samples))
+
+grid.arrange(cv.back.few$plot + ggtitle(paste(nrow(subset(sum.patient.data, Status == 'P' & total.samples < m))," Progressors with < median # samples")), 
+             cv.back.many$plot + ggtitle(paste(nrow(subset(sum.patient.data, Status == 'P' & total.samples >= m)), "Progressors with >= median # samples")), top=paste("Median number of samples:", m))
+
+```
+
+## Peel back timepoints
+
+So here, do I use the initial fit (all samples), or should I be using the noHGD fit...
+
+```{r timepoints, echo=T, message=F, warning=F, eval=F}
 
@@
 nonps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'NP')$Patient], 0)
-plots = list()
-for (i in seq(1,35,5)) {
+fpr = data.frame(matrix(ncol=3,nrow=0))
+for (i in c(0,1,seq(2,35,3))) {
   ps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'P')$Patient], i)
+  #samples = get.next.samples(patient.data, i)
   samples = c(nonps, ps)
-  cv.back = cv.patient.glmnet(x=dysplasia.df[samples,], y=labels[samples], lambda, subset(patients, Samplename %in% samples))
-  plots[[ as.character(i) ]] = cv.back$plot + ggtitle(paste("CV Performance - ", length(ps), "progressor samples"))
+  cv.back = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda0, subset(patients, Samplename %in% samples))
+  fpr = rbind(fpr, c(cv.back$max.cm, sum(cv.back$lambdas[ which.max(cv.back$lambdas$mean)-1, c('sme')]), length(ps)))
+#  plots[[ as.character(i) ]] = cv.back$plot + ggtitle(paste(length(ps), "progressor samples. mean f1:", cv.back$f1))
 }
+colnames(fpr) = c('performance','error','n.prog.samples')
 
-do.call('grid.arrange', c(plots, ncol=2, top="Peeling back timepoints in progressors"))
+grid.arrange(ggplot(fpr, aes(performance,n.prog.samples)) + geom_point(),
+             ggplot(fpr, aes(performance, error)) + geom_point(), top='Peel back samples')
+
+pander(fpr, justify='left', caption="Removing timepoints from the model and retraining")
+```
+
+Can look then at a single point that may be an example of the performance of the model.
+
+```{r singletimepoint, echo=T, message=F, warning=F, eval=F}
+ps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'P')$Patient], 11)
+samples = c(nonps, ps)
+cv.back = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda0, subset(patients, Samplename %in% samples))
+
+cv.back$plot + ggtitle(paste("CV Performance - ", length(ps), "progressor samples"))
+
+info = subset(patient.info, Samplename %in% samples)
+
+coef.opt.stability = coef.stability(coef.opt, cv.back$non.zero.cf)
+
+pander(as.data.frame(sort(rowSums(coef.opt.stability[,2:6]), decreasing = T)), justify='left', caption="Features initially identified found in x of 5 xval splits for this limited model")
 
 ```
+
+## Example
+
+This can't be right...
+
+```{r , echo=T, message=F, warning=F, eval=F}
+
+# prog.samp = sample(subset(sum.patient.data, Status == 'P')$Patient)
+# 
+# pt = 'AHM0952' # P
+# #pt = 'AHM0254' # NP
+# testSamples = as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Patient == pt)$Patient], function(pt) pt$info$Samplename)))
+# 
+# #samples = as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Patient != pt)$Patient], function(pt) pt$info$Samplename)))
+# #cv.back = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda0, subset(patients, Samplename %in% samples))
+# #cv.back$plot
+# 
+# predict(fit0, dysplasia.df[testSamples,], s=lambda.opt, type='response')
+# predict(fit0, dysplasia.df[testSamples,], s=lambda.opt, type='class')
+# 
+# 
+# # Random selection of samples
+# randomSamples = sample(as.vector(unlist(lapply(patient.data, function(pt) pt$info$Samplename))), 30)
+# predict(fit0, dysplasia.df[randomSamples,], s=lambda.opt, type='response')
+# 
+# cbind(predict(fit0, dysplasia.df[randomSamples,], s=lambda.opt, type='class'),labels[randomSamples])
+
+```

### 2017-05-10 RegressionPvNP.Rmd
Using deviance measure rather than performance to select model

@@
 library(pander)
-#library(Hmisc)
 library(reshape2)
@@
 sum.patient.data = as.data.frame(subset(sum.patient.data, Patient %in% names(patient.data))) ## For now
-pander(sum.patient.data[,c('Patient', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')])
+
+pander(sum.patient.data[,c('Patient', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left')
 
@@
 for (pt in names(patient.data)) {
+  #print(pt)
+  #print(patient.data[[pt]]$info$Samplename %in% colnames(patient.data[[pt]]$seg.vals)[-(1:5)])
   patient.data[[pt]]$info = patient.data[[pt]]$info[patient.data[[pt]]$info$Samplename %in% colnames(patient.data[[pt]]$seg.vals)[-(1:5)],]
@@
 ```{r apclust-func, echo=F, message=F, warning=F}
-#suppressPackageStartupMessages( library(apcluster) )
-
 apclust.data<-function(segdata, samples) {
@@
 }
-
-# apclust<-function(segdata) {
-#   x1 = apclust.data(segdata)
-#   # q=0 minimize off-diagonal similarity
-#   ac = apcluster(negDistMat(r=2), x1, details=T, convits=25, q=0)
-#   return(list('apres'=ac, 'data'=x1))
-# }
-
 ```
@@
   if (file.exists(tileFile)) {
+    message(paste("Loading", tileFile))
     load(tileFile, verbose=T)
   } else {
+    message("Tiling patient data...")
     grList = lapply(sample.data, function(df) {
@@
 
-Data matrix currently:
+Example subset of the data matrix:
 `r pander(dysplasia.df[1:10, 1:5], caption=paste("dimensions:", paste(dim(dysplasia.df), collapse=', ')), justify='left')`
@@
   for (a in alpha) {
-    cv = cv.glmnet(x, y, alpha=a, family=family, standardize=F)
+    cv = cv.glmnet(x, y, alpha=a, family=family)
     cv.models = rbind(cv.models, cbind('class'=paste('cv',a,sep=''), 'lambda'=log(cv$lambda), 'cvm'=cv$cvm))
@@
 
-gp = plot.multi.cv.glmnet(dysplasia.df, labels, alpha=c(0,0.5,0.8,1), title='cv.glmnet P vs NP')
+gp = plot.multi.cv.glmnet(x=dysplasia.df, y=labels, alpha=c(0,0.5,0.8,1), title='cv.glmnet P vs NP')
 
@@
   cf[which(cf != 0),][-1]
+}
+
+coef.stability<-function(opt, nz.list) {
+  for (i in 1:length(nz.list)) {
+    df = as.data.frame(nz.list[[i]])
+    colnames(df) = i
+    opt = merge(opt, df, by='row.names', all.x=T)
+    rownames(opt) = opt$Row.names
+    opt$Row.names = NULL
+    opt[,(i+1)] = as.integer(!is.na(opt[,(i+1)]))
+  }
+  opt = opt[order(sapply(rownames(opt), function(x) as.numeric(unlist(strsplit(x, ':'))[1])  )),]
+  return(opt)
 }
@@
 
-crossvalidate.by.patient<-function(x,y,lambda,pts,a=1,nfolds=10, splits=5, fit=NULL, minR=0.2) {
+crossvalidate.by.patient<-function(x,y,lambda,pts,a=1,nfolds=10, splits=5, fit=NULL, minR=0.2, select='performance') {
+  if (nfolds > 5) minR = 0.1
   message(paste("Running", splits, "splits",nfolds,"times on", paste(dim(x), collapse=':'), 'alpha=',a ))
   fit.e = list()
-  pts = create.patient.sets(pts, nfolds, splits, minR)
+  tpts = create.patient.sets(pts, nfolds, splits, minR)
   cv.pred = (matrix(nrow=0, ncol=length(lambda)))
+  cv.binomial.deviance = (matrix(nrow=0, ncol=length(lambda)))
   for (n in 1:nfolds) {  
     message(paste(n, "fold"))
-    setCol = grep(paste('^fold.',n,'$',sep=''), colnames(pts))
+    setCol = grep(paste('^fold.',n,'$',sep=''), colnames(tpts))
 
     cv.class = matrix(nrow=splits, ncol=length(lambda))
+    deviance = matrix(nrow=splits, ncol=length(lambda))
     for (i in 1:splits) { 
       message(paste(i, "split"))
-      test.rows = which(rownames(x) %in% pts[which(pts[,setCol] == i), 'Samplename'])
+      test.rows = which(rownames(x) %in% tpts[which(tpts[,setCol] == i), 'Samplename'])
       test = x[test.rows,]
@@
       # pre-spec lambda seq
-      fit <- glmnet(training, y[-test.rows], lambda=lambda, family='binomial', alpha=a) 
+      fitCV <- glmnet(training, y[-test.rows], lambda=lambda, family='binomial', alpha=a) 
       # autoplot(fit) + theme(legend.position="none")
-      # careful: matrix
-      pred <- pi.hat(predict(fit, test, type='link')) 
+
+      # Confusion matrix: quantitiative, pos results + neg results / number of test rows
+      pred <- pi.hat(predict(fitCV, test, type='link')) 
       cv.class[i,] = apply(pred, 2, function(p) {
-        # Confusion matrix: quantitiative, pos results + neg results / number of test rows
-        (p%*%labels[test.rows] + (1-p) %*% (1-labels[test.rows]))/ length(test.rows)
+        (p%*%y[test.rows] + (1-p) %*% (1-y[test.rows]))/length(test.rows)
       })
+
+      # Binomial deviance, lifted from cv.lognet
+      predmat = predict(fitCV, test, type='response')
+      prob_min = 1e-05; prob_max = 1 - prob_min
+
+      predmat = pmin(pmax(predmat, prob_min), prob_max)
+
+      yT = as.factor(y[test.rows])
+      dev = apply(predmat, 2, function(p)  -2*((yT==1)*log(p)+(yT==0)*log(1-p)) )
+
+      deviance[i,] = apply(dev, 2, weighted.mean, w=rep(1, nrow(dev)), na.rm=T)
 
@@
     cv.pred = rbind(cv.pred, cv.class)    
+    cv.binomial.deviance = rbind(cv.binomial.deviance, deviance)
   }
 
+  # Not really sure what to do with the binomial deviance now...
   df = cbind.data.frame('lambda.at'=1:ncol(cv.pred),
@@
                         'sd'= apply(cv.pred, 2, sd), 
-                        'lambda'= lambda)
+                        'lambda'= lambda,
+                        'log.lambda' = log(lambda),
+                        'mean.b.dev' = colMeans(cv.binomial.deviance),
+                        'sd.b.dev' = apply(cv.binomial.deviance, 2,sd),
+                        'sme.b.dev' = apply(cv.binomial.deviance, 2, sd)/sqrt(nrow(cv.binomial.deviance)),
+                        'lambda.min' = F, 'lambda.1se' = F)
+  #df$lambda.1se = FALSE
+  #df$`lambda-1se` = FALSE
 
-  df2 = df
-  df2$lambda = log(lambda)
-  df2$mean = 1-df2$mean
+  if (select == 'performance') {
+    df$lambda.min = df$mean == max(df$mean)
+    df[df$log.lambda < subset(df, mean == max(mean))$log.lambda+sd(df$log.lambda),][1, 'lambda.1se'] = TRUE
+    #df[df$log.lambda < subset(df, mean == max(mean))$log.lambda-sd(df$log.lambda),][1, 'lambda-1se'] = TRUE
+  } else if (select == 'deviance') {
+    df$lambda.min = df$mean.b.dev == min(df$mean.b.dev)
+    df[df$mean.b.dev >= df[df$lambda.min == T, 'mean.b.dev'] + sd(df$mean.b.dev), ][1, 'lambda.1se'] = T
+  }
 
-  df2$lambda.min = df2$mean == min(df2$mean)
-  df2$lambda.1se = FALSE
-  df2[df2$lambda < subset(df2, mean == min(mean))$lambda-sd(df2$lambda),][1,'lambda.1se'] = TRUE
+  lambda.min = lambda[subset(df, lambda.min == T)$lambda.at]
+  lambda.1se = lambda[subset(df, lambda.1se == T)$lambda.at]
 
-  lambda.min = lambda[subset(df2, lambda.min == T)$lambda.at]
-  lambda.1se = lambda[subset(df2, lambda.1se == T)$lambda.at]
-  nzcf = lapply(fit.e, non.zero.coef, s=lambda.min)
+  nzcf = lapply(fit.e, non.zero.coef, s=lambda.1se)
 
-  gp = ggplot(df2, aes(y=mean,x=lambda)) + geom_point() + 
-    geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + 
-    geom_point(data=subset(df2, lambda.min == T), aes(y=mean, x=lambda, colour="lambda.min"), size=2 ) +          
-    geom_vline(xintercept = subset(df2, lambda.min == T)$lambda, colour='grey') +
-    annotate("text", x=subset(df2, lambda.min == T)$lambda, y=0.35, 
-             label=round(df[subset(df2, lambda.min == T)$lambda.at,'mean'],3))
+  plots = plot.patient.cv(df, fit)
+  plots$performance = plots$performance + theme(legend.position='bottom') + scale_colour_discrete(name = "") +
+    labs(title=paste(splits,' splits, ',nfolds,' folds, alpha=',a, sep=''), y='mean pred.', x='log(lambda)') 
+  plots$deviance = plots$deviance + theme(legend.position='bottom') + scale_colour_discrete(name="") + 
+    labs(title=paste(splits,' splits, ',nfolds,' folds, alpha=',a, sep=''), y='mean Binomial Deviance', x='log(lambda)')
 
-  if (length(lambda.1se) > 0) {
-    gp = gp + geom_point(data=subset(df2, lambda.1se == T), aes(y=mean, x=lambda, colour='lambda.1se'), size=2 ) +
-      geom_vline(xintercept = subset(df2, lambda.1se == T)$lambda, colour='grey') +
-      annotate("text", x=subset(df2, lambda.1se == T)$lambda, y=0.35, 
-               label=round(df[subset(df2, lambda.1se == T)$lambda.at,'mean'],3)) 
+  return(list('max.cm'=df[which.max(df$mean)-1,'mean'], 
+              'lambda.min'=lambda.min, 'lambda.1se'=lambda.1se, 'lambdas'=df, 'plot'=plots$performance, 'deviance.plot'=plots$deviance, 'non.zero.cf'=nzcf))
 }
-   gp = gp + theme( legend.position='bottom') + 
-          scale_colour_discrete(name = "") +
-          labs(title=paste(splits,' splits, ',nfolds,' folds, alpha=',a, sep=''), y='1-mean (P)', x='log(lambda)')
+
+plot.patient.cv<-function(df, fit=NULL) {
+    gp = ggplot(df, aes(y=mean,x=log.lambda)) + geom_point() + geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + 
+    geom_point(data=subset(df, lambda.min == T), aes(y=mean, x=log.lambda, colour="lambda.min"), size=2 ) +  
+    geom_vline(xintercept = subset(df, lambda.min == T)$log.lambda, colour='grey') +
+    annotate("text", x=subset(df, lambda.min == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
+             label=round(df[subset(df, lambda.min == T)$lambda.at,'mean'],3))
+
+    gpD = ggplot(df, aes(y=mean.b.dev,x=log.lambda)) + geom_point() +
+    geom_errorbar(aes(ymin=mean.b.dev-sme.b.dev, ymax=mean.b.dev+sme.b.dev)) + 
+    geom_point(data=subset(df, lambda.min == T), aes(y=mean.b.dev, x=log.lambda, colour="lambda.min"), size=2 ) +          
+    geom_vline(xintercept = subset(df, lambda.min == T)$log.lambda, colour='grey') +
+    annotate("text", x=subset(df, lambda.min == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
+             label=round(df[subset(df, lambda.min == T)$lambda.at,'mean'],3))
+
+  if (nrow(subset(df, lambda.1se == T)) > 0) {
+    gp = gp + geom_point(data=subset(df, lambda.1se == T), aes(y=mean, x=log.lambda, colour="lambda.min+1se"), size=2 ) +
+      geom_vline(xintercept = subset(df, lambda.1se == T)$log.lambda, colour='grey') +
+      annotate("text", x=subset(df, lambda.1se == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
+               label=round(df[subset(df, lambda.1se == T)$lambda.at,'mean'],3))
+
+    gpD = gpD + geom_point(data=subset(df, lambda.1se == T), aes(y=mean.b.dev, x=log.lambda, colour="lambda.min+1se"), size=2 ) +
+      geom_vline(xintercept = subset(df, lambda.1se == T)$log.lambda, colour='grey') +
+      annotate("text", x=subset(df, lambda.1se == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
+               label=round(df[subset(df, lambda.1se == T)$lambda.at,'mean'],3)) 
+  }
+
+  if (nrow(subset(df, `lambda-1se` == T)) > 0) {
+      gp = gp + geom_point(data=subset(df, `lambda-1se` == T), aes(y=mean, x=log.lambda, colour="lambda.min-1se"), size=2 ) +
+      geom_vline(xintercept = subset(df, `lambda-1se` == T)$log.lambda, colour='grey') +
+      annotate("text", x=subset(df, `lambda-1se` == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
+               label=round(df[subset(df, `lambda-1se` == T)$lambda.at,'mean'],3))
+
+      gpD = gpD + geom_point(data=subset(df, `lambda-1se` == T), aes(y=mean.b.dev, x=log.lambda, colour="lambda.min-1se"), size=2 ) +
+      geom_vline(xintercept = subset(df, `lambda-1se` == T)$log.lambda, colour='grey') +
+      annotate("text", x=subset(df, `lambda-1se` == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
+               label=round(df[subset(df, `lambda-1se` == T)$lambda.at,'mean'],3)) 
+  }
+
+  gp = gp + theme(legend.position='bottom') + scale_colour_discrete(name = "") 
+  gpD = gpD + theme(legend.position='bottom') + scale_colour_discrete(name="")
 
   if (!is.null(fit)) {
-    df2$nzcoef = sapply(df$lambda, function(l) length(non.zero.coef(fit, l)))
+    df$nzcoef = sapply(df$lambda, function(l) length(non.zero.coef(fit, l)))
 
-    coef.min = subset(df2, lambda.min == T)$nzcoef
-    coef.1se = subset(df2, lambda.1se == T)$nzcoef
+    coef.min = subset(df, lambda.min == T)$nzcoef
+    coef.1se = subset(df, lambda.1se == T)$nzcoef
+    coef.m.1se = subset(df, `lambda-1se` == T)$nzcoef
+
     d = coef.1se-coef.min
 
-    coef.text = arrange(subset(df2,nzcoef %in% c(0,  coef.min, coef.1se, max(nzcoef) )), -lambda.min, -lambda.1se)
+    coef.text = arrange(subset(df,nzcoef %in% c(0,  coef.min, coef.1se, coef.m.1se, max(nzcoef) )), -lambda.min, -lambda.1se)
     coef.text = arrange(coef.text[!duplicated(coef.text$nzcoef),], -nzcoef)
 
-    gp = gp + annotate("text", y=max(df2$mean), x=coef.text$lambda, label=coef.text$nzcoef)
+    gp = gp + annotate("text", y=min(df$mean), x=coef.text$log.lambda, label=coef.text$nzcoef, color='darkblue')
+    gpD = gpD + annotate("text", y=max(df$mean.b.dev), x=coef.text$log.lambda, label=coef.text$nzcoef, color='darkblue')
+  }
+return(list('performance'=gp, 'deviance'=gpD))
 }
 
-  return(list('max.cm'=df[which.max(df$mean)-1,'mean'], 
-              'lambda.min'=lambda.min, 'lambda.1se'=lambda.1se, 'lambdas'=df, 'plot'=gp, 'non.zero.cf'=nzcf))
-}
-
-coef.stability<-function(opt, nz.list) {
-  for (i in 1:length(nz.list)) {
-    df = as.data.frame(nz.list[[i]])
-    colnames(df) = i
-    opt = merge(opt, df, by='row.names', all.x=T)
-    rownames(opt) = opt$Row.names
-    opt$Row.names = NULL
-    opt[,(i+1)] = as.integer(!is.na(opt[,(i+1)]))
-  }
-  opt = opt[order(sapply(rownames(opt), function(x) as.numeric(unlist(strsplit(x, ':'))[1])  )),]
-  return(opt)
-}
 ```
@@
 ```{r xvalpt, message=F, warning=F, echo=T, fig.height=12, fig.width=12}
+performance.plots = list(); coefs = list(); deviance.plots = list()
 plots = list()
-coefs = list()
 folds = 10; splits=5
@@
   autoplot(fit0, xvar='lambda', main=paste('fit0, all samples, alpha=',a,sep='')) + theme(legend.position='none')
-  lambda0 <- fit0$lambda
 
-  cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=lambda0, pts=pts, a=a, nfolds=folds, splits=splits, fit=fit0)
+  cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=fit0$lambda, pts=pts, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance')
 
-  coef.1se = as.data.frame(non.zero.coef(fit0, cv.patient$lambda.1se))
-  coefs[[as.character(a)]] = coef.stability(coef.1se, cv.patient$non.zero.cf)
+  lambda.opt = ifelse( length(cv.patient$lambda.1se) > 0, cv.patient$lambda.1se, cv.patient$lambda.min )
 
-  plots[[as.character(a)]] = cv.patient$plot + theme(legend.position='bottom') + ggtitle(paste('alpha=',a,sep=''))
+  coef.opt = as.data.frame(non.zero.coef(fit0, lambda.opt))
+  coefs[[as.character(a)]] = coef.stability(coef.opt, cv.patient$non.zero.cf)
+
+  #performance.plots[[as.character(a)]] = cv.patient$plot + ggtitle(paste('alpha=',a,sep=''))
+  #deviance.plots[[as.character(a)]] = cv.patient$deviance.plot + ggtitle(paste('alpha=',a,sep=''))
+
+  plots[[as.character(a)]] = arrangeGrob(cv.patient$plot+ggtitle('CV Performance'), cv.patient$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
 }
 
-do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits', ncol=2))
+do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits'))
 
@@
 
-```{r noHGD, echo=T, message=F, warning=F, fig.height=8, fig.width=8}
+```{r noHGD, echo=T, message=F, warning=F, fig.height=12, fig.width=12}
 `%nin%` <- Negate(`%in%`)
@@
 
-  plots[[as.character(a)]] = cv.nohgd$plot + ggtitle(paste('alpha=',a,sep=''))
+  #plots[[as.character(a)]] = cv.nohgd$plot + ggtitle(paste('alpha=',a,sep=''))
+
+  plots[[as.character(a)]] = arrangeGrob(cv.patient$plot+ggtitle('CV Performance'), cv.patient$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
+
   coef.1se = as.data.frame(non.zero.coef(fit1, cv.nohgd$lambda.1se))
@@
 
-for (i in names(coef.stable)) 
+for (i in names(coef.stable)) {
+  if (is.null(names(coef.stable$`0.8`))) next
   cfs[intersect(rownames(cfs), names(coef.stable[[i]])), i] = 1
+}
 
@@
 
-plots = list()
-coefs = list()
+plots = list(); coefs = list()
 for (a in alpha.values) {
@@
 
-  plots[[as.character(a)]] = cv.nolgd$plot + ggtitle(paste('alpha=',a,sep=''))
+  #plots[[as.character(a)]] = cv.nolgd$plot + ggtitle(paste('alpha=',a,sep=''))
+
+  plots[[as.character(a)]] = arrangeGrob(cv.nolgd$plot+ggtitle('CV Performance'), cv.nolgd$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
 
@@
   cfs[intersect(rownames(cfs), names(coef.stable[[i]])), i] = 1
-
 ```
@@
 
-### Leave one patient out
-
-Using the 'all sample' fit, run a LOO (patient)
-
-```{r loop, echo=T, message=F, warning=F, fig.height=6, fig.width=6, eval=F}
-fit0 <- glmnet(dysplasia.df, labels, alpha=1, nlambda=200, family='binomial') # all patients
-
-# Randomize patients
-pt.samp = sample(sum.patient.data$Patient)
-performance.at.1se = c()
-coefs = list()
-for (pt in pt.samp) {
-  samples = as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Patient != pt)$Patient], function(pt) pt$info$Samplename)))
-
-  cv.loo = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fit0$lambda, pts= subset(pts, Samplename %in% samples), a=1, nfolds=20, splits=5, fit=fit0)
-
-  performance.at.1se = c(performance.at.1se, subset(cv.loo$lambdas, lambda == cv.loo$lambda.1se)$mean)
-
-  coef.1se = as.data.frame(non.zero.coef(fit0, cv.loo$lambda.1se))
-  coefs[[as.character(a)]] = coef.stability(coef.1se, cv.loo$non.zero.cf)
-}
-
-ggplot(as.data.frame(performance.at.1se), aes(y=performance.at.1se, x='LOO')) + ylim(0.6,0.8) +
-  geom_boxplot( fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA) + geom_jitter() +
-    labs(y='performance at lambda.1se', x='', title='Performance for Leave One (patient) Out, at alpha=1')
-
-```
-
 ## Compare progressors with many samples vs few
@@
 
-fit0 <- glmnet(dysplasia.df[samples,], labels[samples], alpha=1, nlambda=200, family='binomial') # all patients
-autoplot(fit0) + theme(legend.position = 'none')
+few.plots = list()
+for (a in alpha.values) {
+  fit0 <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, nlambda=500, family='binomial') # all patients
+  cv.back.few = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fit0$lambda, pts=subset(pts, Samplename %in% samples), splits=5, nfolds=20, a=a, fit=fit0, minR=0.1)
+  #few.plots[[as.character(a)]] = cv.back.few$plot + ggtitle(paste('alpha=',a,sep=''))
 
-cv.back.few = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fit0$lambda, pts=subset(pts, Samplename %in% samples), splits=5, nfolds=20, a=1, fit=fit0)
+  few.plots[[as.character(a)]] = arrangeGrob(cv.back.few$plot+ggtitle('CV Performance'), cv.back.few$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
+}
+do.call(grid.arrange, c(few.plots, ncol=2, top=paste(nrow(subset(sum.patient.data, Status == 'P' & total.samples < m))," Progressors with < median # samples (",m,")")))
 
@@
 samples = c(nonps, as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Status == 'P' & total.samples >= m)$Patient], function(pt) pt$info$Samplename))))
-cv.back.many = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda0, subset(patients, Samplename %in% samples))
 
-grid.arrange(cv.back.few$plot + ggtitle(paste(nrow(subset(sum.patient.data, Status == 'P' & total.samples < m))," Progressors with < median # samples")), 
-             cv.back.many$plot + ggtitle(paste(nrow(subset(sum.patient.data, Status == 'P' & total.samples >= m)), "Progressors with >= median # samples")), top=paste("Median number of samples:", m))
+many.plots = list()
+for (a in alpha.values) {
+  fit0 <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, nlambda=500, family='binomial') # all patients
+
+  cv.back.many = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fit0$lambda, pts=subset(pts, Samplename %in% samples), splits=5, nfolds=20, a=a, fit=fit0)
+  #many.plots[[as.character(a)]] = cv.back.many$plot + ggtitle(paste('alpha=',a,sep=''))
+
+  many.plots[[as.character(a)]] = arrangeGrob(cv.back.many$plot+ggtitle('CV Performance'), cv.back.many$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
+}
+do.call(grid.arrange, c(many.plots, ncol=2, top=paste(nrow(subset(sum.patient.data, Status == 'P' & total.samples >= m))," Progressors with >= median # samples (",m,")")))
 
@@
 
-## Peel back timepoints
+# Leave one out
 
-So here, do I use the initial fit (all samples), or should I be using the noHGD fit...
+Leave out each patient (both P and NP) and run a new CV fit each time then predict the samples for the patient that was left out.  Fitted model includes HGD samples.
 
-```{r timepoints, echo=T, message=F, warning=F, eval=F}
+```{r leaveoneout, echo=T, message=F, warning=T}
 
-get.next.samples<-function(patient.data, i) {
-  as.vector(unlist(sapply(patient.data, function(pd) {
-    j = i
-    pd$info = arrange(pd$info, Endoscopy.Year, Pathology)
+info = do.call(rbind, lapply(patient.data, function(df) df$info))
+pg.samp = lapply(patient.data, function(pt) {
+  info = pt$info
+  info$SampleSD = apply(pt$seg.vals[,info$Samplename], 2, sd)
+  info$SampleMEAN =  apply(pt$seg.vals[,info$Samplename], 2, mean)
+  info$Prediction = NA
+  info$PID = unlist(lapply(info$Path.ID, function(x) unlist(strsplit(x, 'B'))[1]))
+  return(info)
+})
 
-    if (j >= nrow(pd$info)) j = nrow(pd$info)
+if (file.exists('/tmp/loo.Rdata')) {
+  load('/tmp/loo.Rdata', verbose=T)
+} else {
 
-    pd$info[1:(nrow(pd$info)-j), 'Samplename']
-  }) ))
+performance.at.1se = c(); coefs = list(); plots = list(); fits = list()
+# Remove each patient (LOO)
+for (pt in names(pg.samp)) {
+  tmp.patient.data = patient.data[subset(sum.patient.data, Patient != pt)$Patient]
+  samples = as.vector(unlist(sapply(tmp.patient.data, function(df) df$info$Samplename )))
+  length(samples)
+
+  # Fit generated on all samples, including HGD
+  a = 1
+  fitLOO <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=1000) # all patients
+  cv = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitLOO$lambda, 
+                                pts=subset(pts, Samplename %in% samples), a=a, nfolds=10, splits=5, fit=fitLOO)
+  plots[[pt]] = cv$plot + ggtitle(pt)
+  fits[[pt]] = cv
+
+  if ( length(cv$lambda.1se) > 0 ) {
+    performance.at.1se = c(performance.at.1se, subset(cv$lambdas, lambda == cv$lambda.1se)$mean)
+
+    coef.1se = as.data.frame(non.zero.coef(fit1, cv$lambda.1se))
+    coefs[[pt]] = coef.stability(coef.1se, cv$non.zero.cf)
+
+    pg.samp[[pt]]$Prediction = predict(fitLOO, dysplasia.df[pg.samp[[pt]]$Samplename,], s=cv$lambda.1se, type='response')[,1]
+  } else {
+    warning(paste("Patient", pt, "did not have a 1se"))
+  }
+}
+save(plots, performance.at.1se, coefs, fits, file='/tmp/loo.Rdata')
 }
 
-nonps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'NP')$Patient], 0)
-fpr = data.frame(matrix(ncol=3,nrow=0))
-for (i in c(0,1,seq(2,35,3))) {
-  ps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'P')$Patient], i)
-  #samples = get.next.samples(patient.data, i)
-  samples = c(nonps, ps)
-  cv.back = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda0, subset(patients, Samplename %in% samples))
-  fpr = rbind(fpr, c(cv.back$max.cm, sum(cv.back$lambdas[ which.max(cv.back$lambdas$mean)-1, c('sme')]), length(ps)))
-#  plots[[ as.character(i) ]] = cv.back$plot + ggtitle(paste(length(ps), "progressor samples. mean f1:", cv.back$f1))
-}
-colnames(fpr) = c('performance','error','n.prog.samples')
-
-grid.arrange(ggplot(fpr, aes(performance,n.prog.samples)) + geom_point(),
-             ggplot(fpr, aes(performance, error)) + geom_point(), top='Peel back samples')
-
-pander(fpr, justify='left', caption="Removing timepoints from the model and retraining")
-```
-
-Can look then at a single point that may be an example of the performance of the model.
-
-```{r singletimepoint, echo=T, message=F, warning=F, eval=F}
-ps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'P')$Patient], 11)
-samples = c(nonps, ps)
-cv.back = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda0, subset(patients, Samplename %in% samples))
-
-cv.back$plot + ggtitle(paste("CV Performance - ", length(ps), "progressor samples"))
-
-info = subset(patient.info, Samplename %in% samples)
-
-coef.opt.stability = coef.stability(coef.opt, cv.back$non.zero.cf)
-
-pander(as.data.frame(sort(rowSums(coef.opt.stability[,2:6]), decreasing = T)), justify='left', caption="Features initially identified found in x of 5 xval splits for this limited model")
+ggplot(as.data.frame(performance.at.1se), aes(y=performance.at.1se, x='LOO')) + ylim(0.5,0.9) +
+  geom_boxplot( fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA) + geom_jitter() +
+  labs(y='performance at lambda.1se', x='', title='Performance for Leave One (patient) Out, at alpha=1')
 
@@
 
-## Example
+## How do the predictions look at each time point?
 
-This can't be right...
+Providing a single sample at a time to predict is unnecessary as each sample is independently predicted from the test matrix.
+Predictions are exactly the same even when a single sample is provided (checked).
 
-```{r , echo=T, message=F, warning=F, eval=F}
+```{r timepoints, echo=F, message=F, warning=F}
+# For HGD/IMC samples, progressors only
+hgd = sapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], function(pt) {
+  hgd = subset(pt, Pathology %in% c('HGD', 'IMC'))
+  sum(as.integer(hgd$Prediction > 0.5))/nrow(hgd)
+})
 
-# prog.samp = sample(subset(sum.patient.data, Status == 'P')$Patient)
-# 
-# pt = 'AHM0952' # P
-# #pt = 'AHM0254' # NP
-# testSamples = as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Patient == pt)$Patient], function(pt) pt$info$Samplename)))
-# 
-# #samples = as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Patient != pt)$Patient], function(pt) pt$info$Samplename)))
-# #cv.back = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda0, subset(patients, Samplename %in% samples))
-# #cv.back$plot
-# 
-# predict(fit0, dysplasia.df[testSamples,], s=lambda.opt, type='response')
-# predict(fit0, dysplasia.df[testSamples,], s=lambda.opt, type='class')
-# 
-# 
-# # Random selection of samples
-# randomSamples = sample(as.vector(unlist(lapply(patient.data, function(pt) pt$info$Samplename))), 30)
-# predict(fit0, dysplasia.df[randomSamples,], s=lambda.opt, type='response')
-# 
-# cbind(predict(fit0, dysplasia.df[randomSamples,], s=lambda.opt, type='class'),labels[randomSamples])
+# All samples
+all.predictions = sapply(pg.samp, function(pt) {
+  if (unique(pt$Status) == 'NP') {
+    as.integer(pt$Prediction < 0.5)
+  } else {
+    as.integer(pt$Prediction > 0.5)
+  }
+})
+
+np = sapply(all.predictions[subset(sum.patient.data, Status == 'NP')$Patient], function(x) round(sum(x)/length(x),2))
+pg = sapply(all.predictions[subset(sum.patient.data, Status == 'P')$Patient], function(x) round(sum(x)/length(x),2))
 
 ```
+
+### Progressors
+
+In progressors, `r round(sum(hgd,na.rm=T)/length(hgd[!is.na(hgd)]), 2)*100`% of the HGD samples were correctly predicted.
+
+When predicting each sample individually `r length(pg[pg == 1])` patients were correctly predicted in all samples. The remaining patients were predicted in 80% or fewer samples.
+
+`r pander( as.data.frame(sort(pg)), caption='Proportion of progressor patients samples correctly predicted')`
+
+### Non-Progressors
+
+In non-progressors `r length(which(np == 1))` patients were predicted to be non-progressors in _all_ of their samples.
+`r length(which(np >= 0.75))` were predicted in 75% or more of their samples.
+`r length(np[which(np < 0.75 )])` patients had less than 75% of their samples predicted as NP.
+
+`r pander( as.data.frame(sort(np)), caption='Proportion of NP patients samples correctly predicted') `
+
+```{r timepoints2, echo=T, message=F, warning=F, eval=F, fig.height=400, fig.width=100}
+# For all pre-HGD/IMC/LGD in progressors
+pre.hgd = sapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], function(pt) {
+  e = subset(pt, Pathology %nin% c('HGD', 'IMC', 'LGD'))
+  e$Prediction
+  sum(as.integer(e$Prediction > 0.5))/nrow(e)
+})
+
+pred.by.endo<-function(pt, nyears) {
+  e = subset(pt, PID %in% unique(pt$PID)[1:nyears])
+  if (  length(which(e$Pathology %in% c('HGD','IMC'))) > 0 ) return(NA)
+  p = as.vector(as.integer(e$Prediction > 0.5))
+  names(p) = e$Endoscopy.Year
+  return(p)
+}
+
+predicted.progression<-function(x) { if(is.na(x)) return(NA); as.integer(length(which(x == 1)) > 0)  }
+
+p.predictions = as.data.frame(matrix(data=0,nrow=6, ncol=3, dimnames=list(c(), c('%.Predicted','n.Endoscopies','n.Patients'))))
+for (i in 1:nrow(p.predictions)) {
+  p = sapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], pred.by.endo, i)
+  pp = sapply(p, predicted.progression)
+  pp = pp[!is.na(pp)]
+  p.predictions[i,] =  rbind(round(sum(pp[!is.na(pp)]/length(pp[!is.na(pp)])), 2), i, length(pp))
+}
+
+ggplot(p.predictions, aes(y=`%.Predicted`,x=n.Endoscopies)) + geom_bar(stat="identity", fill='dodgerblue3') + #geom_text(aes(label=`%.Predicted`), nudge_y=0.005) + 
+  geom_text(aes(label=paste(n.Patients, 'patients')), nudge_y=0.005) + ggtitle("Progression predicted per endoscopy")
+```
+
+`r pander(p.predictions, justify='left', caption=paste("For any given endoscopy if a single sample (same pathology ID, different blocks) is predicted as 'progressor' that endoscopy is considered to be indicative of progression.
+At each endoscopy the percentage is calculated excluding any patients that are diagnosed HGD at that point.
+There are ",length(subset(sum.patient.data, Status == 'P')$Patient), " progressor patients.
+Based on that, ", p.predictions[1,1]*100, "% are correctly predicted at the first endoscopy.", sep=''))`

### 2017-05-10 RegressionPvNP.Rmd
…

@@
     fig_width: 10
+    number_sections: yes
     toc: yes
@@
 
+```{r cvglmnetPNP, echo=F, warning=F, message=F, eval=F}
 ## cv.glmnet Ridge vs Lasso
-
-```{r cvglmnetPNP, echo=F, warning=F, message=F, eval=F}
 #With all `r ncol(dysplasia.df)` regions of the genome, first check if/which regression may be appropriate.
@@
 
-      fit.e[[length(fit.e)+1]] = fit
+      fit.e[[length(fit.e)+1]] = fitCV
     }
@@
                         'sme.b.dev' = apply(cv.binomial.deviance, 2, sd)/sqrt(nrow(cv.binomial.deviance)),
-                        'lambda.min' = F, 'lambda.1se' = F)
+                        'lambda.min' = F, 'lambda.1se' = F, `lambda-1se` = F)
   #df$lambda.1se = FALSE
@@
 
-```{r xvalpt, message=F, warning=F, echo=T, fig.height=12, fig.width=12}
+```{r xvalpt, message=F, warning=F, echo=T, fig.height=16, fig.width=16}
 performance.plots = list(); coefs = list(); deviance.plots = list()
@@
 alpha.values = c(0,0.5,0.7,0.8,0.9,1)
+
+if (file.exists('/tmp/all.pt.alpha.Rdata')) {
+  load('/tmp/all.pt.alpha.Rdata', verbose=F)
+} else {
   for (a in alpha.values) {
@@
   }
+  save(plots, coefs, file='/tmp/all.pt.alpha.Rdata')
+}
 
@@
 
-cfs = as.data.frame(matrix(data=0,nrow=length(unique(names(table(unlist(lapply(coef.stable, names)))))), ncol=3, 
+cfs = as.data.frame(matrix(data=0,nrow=length(unique(names(table(unlist(lapply(coef.stable, names)))))), ncol=length(names(coef.stable)), 
                            dimnames=list( unique(names(table(unlist(lapply(coef.stable, names))))), names(coef.stable) )))
@@
 
+table(rowSums(cfs))
+
+coef.stable = lapply(coef.stable, function(cf) cf/(folds*splits))
+
+stability = do.call(rbind, lapply(coef.stable, function(cf) {
+  cbind('75%'=length(which(cf >= 0.75)),  '50%'=length(which(cf >= 0.5)),  '<50%'=length(which(cf < 0.5)))
+}))
+rownames(stability) = names(coef.stable)
 ```
 
-Values that are closer to full lasso (alpha=1) are pretty similar, however at full lasso we see fewer features than if we regularize it at 0.8.
+Values that are closer to full lasso (alpha=1) are pretty similar, however at full lasso we see fewer features than if we regularize it at 0.8. They share all non-zero features as well.
+`r pander(sapply(coef.stable, length), caption='Total features at each value of alpha')`
 
@@
 
-`r pander(as.data.frame(coef.stable[['1']])/(folds*splits), justify='left', caption='Percentage of features from the glm fitted with all data.
-Numbers indicate the total number of folds that feature was identified in.')`
+`r pander(table(rowSums(cfs)), caption="Number of features found in 1, 2, or all 3 models by alpha 0.8, 0.9 and 1")`
 
-They also share all of the non-zero coefficients (0.8 identifies the most non-zero coefs `r max(sapply(coef.stable, length))`).
-
-`r pander(cfs, justify='left')`
-
-#### Did I understand correctly?
-
-In order to predict, my trained model is then my original model, but using the optimal lambda value?
-This isn't quite correct because the model was trained including these patients, just an example.
-
-```{r}
-fit0 <- glmnet(dysplasia.df, labels, alpha=a, nlambda=200, family='binomial') # all patients
-cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=fit0$lambda, pts=pts, a=1, nfolds=20, splits=5, fit=fit0)
-
-# Progressor, minus the HGD sample(s)
-cbind.data.frame(
-  predict(fit0, dysplasia.df[patient.data$PR1_HIN_042$info$Samplename[-(9:10)],], s=cv.patient$lambda.1se, type='response'),
-  predict(fit0, dysplasia.df[patient.data$PR1_HIN_042$info$Samplename[-(9:10)],], s=cv.patient$lambda.1se, type='class'))
-
-# Non-P, all correctly predicted
-cbind.data.frame(predict(fit0, dysplasia.df[patient.data$AD0610$info$Samplename,], s=cv.patient$lambda.1se, type='response'),
-                 predict(fit0, dysplasia.df[patient.data$AD0610$info$Samplename,], s=cv.patient$lambda.1se, type='class'))
-
-```
+`r length(unique(unlist(lapply(coef.stable, function(cf)  names(which(cf > 0.8)) ))))` features are shared at all values of alpha and were non-zero in at least 80% of the folds.
 
@@
 
-```{r noHGD, echo=T, message=F, warning=F, fig.height=12, fig.width=12}
+```{r noHGD, echo=T, message=F, warning=F, fig.height=16, fig.width=16}
 `%nin%` <- Negate(`%in%`)
@@
 
+if (file.exists('/tmp/nohgd.Rdata')) {
+  load('/tmp/nohgd.Rdata', verbose=T)
+} else {
   # No HGD/IMC
@@
 
-plots = list()
+  no.hgd.plots = list()
   coefs = list()
@@
 
-  cv.nohgd = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=lambda1, pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fit1)
+    cv.nohgd = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=lambda1, pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fit1, select='deviance')
 
@@
 
-  plots[[as.character(a)]] = arrangeGrob(cv.patient$plot+ggtitle('CV Performance'), cv.patient$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
+    no.hgd.plots[[as.character(a)]] = arrangeGrob(cv.patient$plot+ggtitle('CV Performance'), cv.patient$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
 
@@
   }
+  save(no.hgd.plots, coefs, file='/tmp/nohgd.Rdata')
+}
 
-do.call(grid.arrange, c(plots, top="No HGD/IMC samples", ncol=2))
+do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2))
 
@@
 
+coef.stable = lapply(coef.stable, function(cf) cf/(folds*splits))
+
+stability = do.call(rbind, lapply(coef.stable, function(cf) {
+  cbind('75%'=length(which(cf >= 0.75)),  '50%'=length(which(cf >= 0.5)),  '<50%'=length(which(cf < 0.5)))
+}))
+rownames(stability) = names(coef.stable)
+
 ```
@@
 
-`r pander(as.data.frame(coef.stable[['1']])/(folds*splits), justify='left', caption='Percentage of features from the glm fitted with all data.
-Numbers indicate the total number of folds that feature was identified in.')`
+`r pander(stability, caption='Total features found at each value of alpha in 50% or more of the folds run')`
 
@@
 
-`r pander(cfs, justify='left')`
+`r pander(table(rowSums(cfs)), justify='left')`
 
@@
 
-```{r noLGD, echo=T, message=F, warning=F, fig.height=8, fig.width=8}
+```{r noLGD, echo=T, message=F, warning=F, fig.height=16, fig.width=16}
+
+if (file.exists('/tmp/nolgd.Rdata')) {
+  load('/tmp/nolgd.Rdata', verbose=T)
+} else {
   # No LGD
@@
 
-plots = list(); coefs = list()
+  nolgd.plots = list(); coefs = list()
   for (a in alpha.values) {
@@
 
-  cv.nolgd = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=lambda2, pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fit2)
+    cv.nolgd = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=lambda2, pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fit2, select='deviance')
 
@@
 
-  plots[[as.character(a)]] = arrangeGrob(cv.nolgd$plot+ggtitle('CV Performance'), cv.nolgd$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
+    nolgd.plots[[as.character(a)]] = arrangeGrob(cv.nolgd$plot+ggtitle('CV Performance'), cv.nolgd$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
 
@@
     coefs[[as.character(a)]] = coef.stability(coef.1se, cv.nolgd$non.zero.cf)
+  }
+  save(nolgd.plots, coefs, file='/tmp/nolgd.Rdata')
 }
@@
   cfs[intersect(rownames(cfs), names(coef.stable[[i]])), i] = 1
+
+coef.stable = lapply(coef.stable, function(cf) cf/(folds*splits))
+
+stability = do.call(rbind, lapply(coef.stable, function(cf) {
+  cbind('75%'=length(which(cf >= 0.75)),  '50%'=length(which(cf >= 0.5)),  '<50%'=length(which(cf < 0.5)))
+}))
+rownames(stability) = names(coef.stable)
+
 ```
@@
 
-```{r, echo=T, message=F, warning=F, eval=F}
+```{r, echo=T, message=F, warning=F, fig.height=16, fig.width=16}
 m = median(subset(sum.patient.data, Status == 'P')$total.samples)
@@
   fit0 <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, nlambda=500, family='binomial') # all patients
-  cv.back.few = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fit0$lambda, pts=subset(pts, Samplename %in% samples), splits=5, nfolds=20, a=a, fit=fit0, minR=0.1)
+  cv.back.few = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fit0$lambda, pts=subset(pts, Samplename %in% samples), splits=5, nfolds=20, a=a, fit=fit0, minR=0.1, select='deviance')
   #few.plots[[as.character(a)]] = cv.back.few$plot + ggtitle(paste('alpha=',a,sep=''))
@@
 
-  cv.back.many = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fit0$lambda, pts=subset(pts, Samplename %in% samples), splits=5, nfolds=20, a=a, fit=fit0)
+  cv.back.many = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fit0$lambda, pts=subset(pts, Samplename %in% samples), splits=5, nfolds=20, a=a, fit=fit0, select='deviance')
   #many.plots[[as.character(a)]] = cv.back.many$plot + ggtitle(paste('alpha=',a,sep=''))
@@
     cv = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitLOO$lambda, 
-                                pts=subset(pts, Samplename %in% samples), a=a, nfolds=10, splits=5, fit=fitLOO)
-  plots[[pt]] = cv$plot + ggtitle(pt)
+                                  pts=subset(pts, Samplename %in% samples), a=a, nfolds=10, splits=5, fit=fitLOO, select='deviance')
+    #plots[[pt]] = cv$plot + ggtitle(pt)
+
+    plots[[pt]] = arrangeGrob(cv$plot+ggtitle('CV Performance'), cv$deviance.plot+ggtitle('Binomial Deviance'), top=pt, ncol=2)
+
     fits[[pt]] = cv
@@
 
-    coef.1se = as.data.frame(non.zero.coef(fit1, cv$lambda.1se))
+      coef.1se = as.data.frame(non.zero.coef(fitLOO, cv$lambda.1se))
       coefs[[pt]] = coef.stability(coef.1se, cv$non.zero.cf)
@@
   }
-save(plots, performance.at.1se, coefs, fits, file='/tmp/loo.Rdata')
+  save(plots, performance.at.1se, coefs, fits, pg.samp, file='/tmp/loo.Rdata')
 }
+
+# do.call(grid.arrange, c(lapply(fits[sample(subset(sum.patient.data, Status == 'NP')$Patient, 4)], function(f) {
+#   plt = plot.patient.cv(f$lambdas)   
+#   arrangeGrob(plt$performance +ggtitle('CV Performance'), plt$deviance+ggtitle('Binomial Deviance'), ncol=2)
+# })))
+
+# do.call(grid.arrange, c(lapply(fits[sample(subset(sum.patient.data, Status == 'P')$Patient, 4)], function(f) {
+#   plt = plot.patient.cv(f$lambdas)   
+#   arrangeGrob(plt$performance +ggtitle('CV Performance'), plt$deviance+ggtitle('Binomial Deviance'), ncol=2)
+# })))
 
@@
 
-When predicting each sample individually `r length(pg[pg == 1])` patients were correctly predicted in all samples. The remaining patients were predicted in 80% or fewer samples.
+When predicting each sample individually `r length(pg[pg == 1])` patients were correctly predicted in all samples. The remaining patients were predicted in `r max(pg[pg != 1])*100`% or fewer samples.
 
@@
 
-```{r timepoints2, echo=T, message=F, warning=F, eval=F, fig.height=400, fig.width=100}
+```{r timepoints2, echo=T, message=F, warning=F}
 # For all pre-HGD/IMC/LGD in progressors
@@
 Based on that, ", p.predictions[1,1]*100, "% are correctly predicted at the first endoscopy.", sep=''))`
+
+What about the non-progressors that had progressor predictions?
+```{r}
+
+np.predictions = as.data.frame(matrix(data=0,nrow=6, ncol=3, dimnames=list(c(), c('%.Predicted','n.Endoscopies','n.Patients'))))
+for (i in 1:nrow(np.predictions)) {
+  p = sapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Patient], pred.by.endo, i)
+  pp = sapply(p, predicted.progression)
+  pp = pp[!is.na(pp)]
+  np.predictions[i,] =  rbind(round(sum(pp[!is.na(pp)]/length(pp[!is.na(pp)])), 2), i, length(pp))
+}
+
+```

### 2017-05-16 Patient Report.Rmd
fixing loads of little mistakes that results in messing up the pipeline entirely.

@@
 library(plyr)
-library(xlsx)
+#library(xlsx)
 library(dplyr)
 library(pander)
-library(Hmisc)
+#library(Hmisc)
 library(ggdendro)
@@
 ## Patient info file
-patient.file = grep('All_patient_info.xls', data.files, value=T)
+patient.file = grep('/All_patient_info.xls', data.files, value=T)
 if (length(patient.file) != 1)
@@
 dataset = 'Training'
+#dataset = 'Test'
 patient.info = read.patient.info(patient.file, set=dataset)
@@
 
+bad_batch = which(patient.info$Batch.Name %in% c('Batch_8', 'Batch_9'))
+patient.info = patient.info[-(bad_batch),]
+
 patient.info = arrange(patient.info, Status, Patient, Endoscopy.Year, Pathology)
@@
 sum.patient.data = summarise.patient.info(patient.info)
+table(sum.patient.data$Status)
 
@@
 
+lowprobes = c()
+
 patient.data = list()
 
-for (patient.name in unique(patient.info$Patient)  )  {
+for (patient.name in sum.patient.data$Patient  )  {
   message(patient.name)
@@
 
-  if (length(missing) > 0) 
+  if (length(missing) > 0) {
     warning(paste("Missing", length(missing), "sample(s) in patient", patient.name))
+    next
+  }
 
-  message( paste(length(which(segvals$n.probes < min.probes)), ' probes below the minimum probe count (',min.probes,')', sep=''))
+  probes = length(which(segvals$n.probes < min.probes))
+  message( paste(probes, ' probes (', round(probes/nrow(segvals))*100, '%)',' below the minimum probe count (',min.probes,')', sep=''))
+  if (probes/nrow(segvals) >= 0.75) lowprobes = c(lowprobes, patient.name)
 
-  segvals = segvals[segvals$n.probes>=min.probes,]
+  segvals = segvals[segvals$n.probes >= min.probes,]
+  dim(segvals)
 
@@
 
+  if (nrow(segvals) < 100) {
+    warning(paste(patient.name, 'has fewer than 100 genomic reginos with segmented values:', nrow(segvals)))
+    next
+  }
+
   patient.data[[patient.name]][['seg.vals']] = segvals
@@
 
-  if(nrow(segvals)>0) {
-      HC = hclust(dist(t(segvals[,-(1:5)])))
+  sample.cols = grep('^D\\d+_D\\d+', colnames(segvals))
+  if (nrow(segvals) > 0) {
+      HC = hclust(dist(t(segvals[,sample.cols])))
       gg1 = ggdendrogram(HC, rotate=T) + labs(title = paste(patient.name,": fitted segment coverage"),x="")
@@
       # Normalize  (value-mean(value))/sd(value)
-      normalised.segvals = segvals[,-(1:5)]
-      for(c in 1:nrow(normalised.segvals)) {
+      normalised.segvals = segvals[,sample.cols]
+      for(c in 1:nrow(normalised.segvals)) 
         normalised.segvals[c,] = (normalised.segvals[c,]-mean(unlist(normalised.segvals[c,])))/sd(unlist(normalised.segvals[c,]))
-      }
+
       patient.data[[patient.name]][['norm.seg.vals']] = normalised.segvals
@@
       # Variable Regions
-      sds = apply(segvals[,-(1:5)],1,sd)
-      means = apply(segvals[,-(1:5)],1,mean)
+      sds = apply(segvals[,sample.cols],1,sd)
+      means = apply(segvals[,sample.cols],1,mean)
       CofV = sds/means
@@
         geom_histogram( fill=I("blue"),col=I("lightblue"),binwidth=0.3) + labs(title=paste(patient.name,'normalised segment coverage'), x='', y='')
-     # print(gg3)
-      #pt.plots[[patient.name]][['norm.hist']] = gg
 
-      # par(mfrow=c(3, no.rows))
-      # for(s in 1:no.samples) 
-      #   hist(normalised.segvals[,s],col="blue",main=names(normalised.segvals)[s],xlab="normalised coverage")
-
-      if(nrow(segvals)>=3) {
+      #if(nrow(segvals)>=3) {
         # order by date of endoscopy
@@
         patient = arrange(patient, Endoscopy.Year, Pathology, Samplename)
-        }
+       # }
         sample.normalised.segvals = normalised.segvals[, intersect(colnames(normalised.segvals), patient$Samplename)]
@@
 
+print(paste("Missing patients:", paste(setdiff(sum.patient.data$Patient, names(patient.data)), collapse=', ')))
+
 sum.patient.data = subset(sum.patient.data, Patient %in% names(patient.data))
+table(sum.patient.data$Status)
 
@@
 patient.info = arrange(patient.info, Status, Patient, Endoscopy.Year,Samplename)
-pt = 'AH0254'
-info = subset(patient.info, Patient == 'AH0254')
-lapply(names(patient.data), function(pt) {
+msgs = sapply(names(patient.data), function(pt) {
   missing = which(!patient.data[[pt]]$info$Samplename %in% colnames(patient.data[[pt]]$seg.vals))
@@
 })
+for (n in msgs) {
+  if (!is.null(n))
+    print(n)
+}
 
@@
 
-example.pt = subset(sum.patient.data, Status == 'P' & total.samples > 2)$Patient[1]
+example.pt = subset(sum.patient.data, Status == 'P' & total.samples > 2)$Patient[5]
 ## Note the number of probes per patient/sample
@@
 
-## Per-sample normalization
-#nsv = patient.data[['AD0258']]$norm.seg.vals
-#nsv[] = lapply(patient.data[['AD0258']]$seg.vals[,-(1:5)], function(x) (x-mean(x))/sd(x))
-
-#ggplot(melt(nsv), aes(x=value))  + 
-#        facet_wrap(~variable,scales = "free_x") + 
-#        geom_histogram( fill=I("blue"),col=I("lightblue"), binwidth=0.3) + labs(title='AD0258 normalized segmented value distribution', x='', y='')
-
 ```
@@
 
-pander(sum.patient.data[,c('Patient','Status','years','total.samples','highest.path','n.segments', 'median.seg.len')], justify='left')
-
-#seg.lens = lapply(patient.data, function(df) df$seg.vals[,c('end.pos')] - df$seg.vals[,c('start.pos')] )
-
-#wilcox.test(log(unlist(seg.lens[subset(sum.patient.data, Status == 'P')$Patient])),
-#log(unlist(seg.lens[subset(sum.patient.data, Status == 'NP')$Patient])))
-
-#hist(log(unlist(seg.lens[subset(sum.patient.data, Status == 'P')$Patient])))
-#hist(log(unlist(seg.lens[subset(sum.patient.data, Status == 'NP')$Patient])))
-
 ggplot(sum.patient.data) + geom_boxplot( aes(y=median.seg.len, x=Status, group=Status, fill=Status) ) +
   labs(title="Median segment length by patient status")
+```
 
-```
+`r pander(sum.patient.data[,c('Patient','Status','years','total.samples','highest.path','n.segments', 'median.seg.len')], justify='left')`
 
@@
 
+```{r echo=F, message=F, warning=F, fig.height=15, fig.width=10, eval=F}
 # Segmented Value Distributions
-```{r echo=F, message=F, warning=F, fig.height=15, fig.width=10, eval=F}
-
 lapply(patient.data, function(pt) {
@@
   colnames(sample.normalised.segvals) =  paste(info$Endoscopy.Year,  c(1:nrow(info)) , sep='.')
+
+  if (nrow(sample.normalised.segvals) <= 4) {
+    warning(paste(name, 'cannot be run in rcorr, too few observations'))
+  }
 

### 2017-05-16 RegressionPvNP.Rmd
fixing loads of little mistakes that results in messing up the pipeline entirely.

@@
 sum.patient.data = summarise.patient.info(patient.info)
+
+if (length(which(sapply(patient.data, function(df) nrow(df$seg.vals)) == 0)) > 0)
+  warning(paste("Patients missing data:",
+    paste(names(which(sapply(patient.data, function(df) nrow(df$seg.vals)) == 0)), collapse=', ')))
+
+patient.data[which(sapply(patient.data, function(df) nrow(df$seg.vals)) == 0)] = NULL
 sum.patient.data = as.data.frame(subset(sum.patient.data, Patient %in% names(patient.data))) ## For now
-
-pander(sum.patient.data[,c('Patient', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left')
 
@@
 }
+
+pander(sum.patient.data[,c('Patient', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left')
+
 ```
@@
     grList = lapply(sample.data, function(df) {
+      print(unique(df$info$Patient))
       df$info = arrange(df$info, Endoscopy.Year, Pathology)
@@
 mergedDf = tile.sample.segments(paste(as.character(tile.w),'_tiledpts.Rdata', sep=''), patient.data)
-
+dim(mergedDf)
 #validatioMergedDf = tile.sample.segments(paste(as.character(tile.w),'validation_tiledpts.Rdata', sep=''), validation.patient.data)
@@
 dysplasia.df = t(mergedDf[,names(labels)])
+dim(dysplasia.df)
 ```
@@
 Example subset of the data matrix:
-`r pander(dysplasia.df[1:10, 1:5], caption=paste("dimensions:", paste(dim(dysplasia.df), collapse=', ')), justify='left')`
+`r pander(dysplasia.df[1:5, 1:5], caption=paste("dimensions:", paste(dim(dysplasia.df), collapse=', ')), justify='left')`
 
@@
 
-crossvalidate.by.patient<-function(x,y,lambda,pts,a=1,nfolds=10, splits=5, fit=NULL, minR=0.2, select='performance') {
+binomial.deviance<-function(pmat, y) {
+  # Binomial deviance, lifted from cv.lognet
+  prob_min = 1e-05; prob_max = 1 - prob_min
+  pmat = pmin(pmax(pmat, prob_min), prob_max)
+  dev = apply(pmat, 2, function(p)  -2*((y==1)*log(p)+(y==0)*log(1-p)) )
+  return(dev)
+}
+
+crossvalidate.by.patient<-function(x,y,lambda,pts,a=1,nfolds=10, splits=5, fit=NULL, minR=0.2, select='classification') {
   if (nfolds > 5) minR = 0.1
@@
       # Binomial deviance, lifted from cv.lognet
-      predmat = predict(fitCV, test, type='response')
-      prob_min = 1e-05; prob_max = 1 - prob_min
-
-      predmat = pmin(pmax(predmat, prob_min), prob_max)
-
-      yT = as.factor(y[test.rows])
-      dev = apply(predmat, 2, function(p)  -2*((yT==1)*log(p)+(yT==0)*log(1-p)) )
-
+      dev = binomial.deviance(predict(fitCV, test, type='response'), as.factor(y[test.rows]))
       deviance[i,] = apply(dev, 2, weighted.mean, w=rep(1, nrow(dev)), na.rm=T)
@@
                         'sme.b.dev' = apply(cv.binomial.deviance, 2, sd)/sqrt(nrow(cv.binomial.deviance)),
-                        'lambda.min' = F, 'lambda.1se' = F, `lambda-1se` = F)
-  #df$lambda.1se = FALSE
-  #df$`lambda-1se` = FALSE
+                        'lambda.min' = F, 'lambda.1se' = F)
 
-  if (select == 'performance') {
+  if (select == 'classification') {
     df$lambda.min = df$mean == max(df$mean)
-    df[df$log.lambda < subset(df, mean == max(mean))$log.lambda+sd(df$log.lambda),][1, 'lambda.1se'] = TRUE
+    se1 = df[df$lambda.min == T, 'mean'] - df[df$lambda.min == T, 'sme']
+
+    arrange(subset(df, mean < se1 & !lambda.min & log.lambda > subset(df, lambda.min)$log.lambda ), -mean)[1:10,]
+    df[arrange(subset(df, mean < se1 & !lambda.min & log.lambda > subset(df, lambda.min)$log.lambda ), -mean)[1, 'lambda.at'], 'lambda.1se'] = T
     #df[df$log.lambda < subset(df, mean == max(mean))$log.lambda-sd(df$log.lambda),][1, 'lambda-1se'] = TRUE
@@
     df$lambda.min = df$mean.b.dev == min(df$mean.b.dev)
-    df[df$mean.b.dev >= df[df$lambda.min == T, 'mean.b.dev'] + sd(df$mean.b.dev), ][1, 'lambda.1se'] = T
+    se1 = df[df$lambda.min == T, 'mean.b.dev'] - df[df$lambda.min == T, 'sme.b.dev']
+    df[arrange(subset(df, mean.b.dev <= se1 & !lambda.min), -mean.b.dev)[1, 'lambda.at'], 'lambda.1se'] = T
+    #df[df$mean.b.dev >= df[df$lambda.min == T, 'mean.b.dev'] + sd(df$mean.b.dev), ][1, 'lambda.1se'] = T
   }
@@
 
-  return(list('max.cm'=df[which.max(df$mean)-1,'mean'], 
+  return(list('max.cm'=df[df$lambda.1se ==T,'mean'], 
               'lambda.min'=lambda.min, 'lambda.1se'=lambda.1se, 'lambdas'=df, 'plot'=plots$performance, 'deviance.plot'=plots$deviance, 'non.zero.cf'=nzcf))
@@
   if (nrow(subset(df, lambda.1se == T)) > 0) {
-    gp = gp + geom_point(data=subset(df, lambda.1se == T), aes(y=mean, x=log.lambda, colour="lambda.min+1se"), size=2 ) +
+    gp = gp + geom_point(data=subset(df, lambda.1se == T), aes(y=mean, x=log.lambda, colour="lambda.1se"), size=2 ) +
       geom_vline(xintercept = subset(df, lambda.1se == T)$log.lambda, colour='grey') +
@@
 
-    gpD = gpD + geom_point(data=subset(df, lambda.1se == T), aes(y=mean.b.dev, x=log.lambda, colour="lambda.min+1se"), size=2 ) +
+    gpD = gpD + geom_point(data=subset(df, lambda.1se == T), aes(y=mean.b.dev, x=log.lambda, colour="lambda.1se"), size=2 ) +
       geom_vline(xintercept = subset(df, lambda.1se == T)$log.lambda, colour='grey') +
@@
                label=round(df[subset(df, lambda.1se == T)$lambda.at,'mean'],3)) 
-  }
-
-  if (nrow(subset(df, `lambda-1se` == T)) > 0) {
-      gp = gp + geom_point(data=subset(df, `lambda-1se` == T), aes(y=mean, x=log.lambda, colour="lambda.min-1se"), size=2 ) +
-      geom_vline(xintercept = subset(df, `lambda-1se` == T)$log.lambda, colour='grey') +
-      annotate("text", x=subset(df, `lambda-1se` == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
-               label=round(df[subset(df, `lambda-1se` == T)$lambda.at,'mean'],3))
-
-      gpD = gpD + geom_point(data=subset(df, `lambda-1se` == T), aes(y=mean.b.dev, x=log.lambda, colour="lambda.min-1se"), size=2 ) +
-      geom_vline(xintercept = subset(df, `lambda-1se` == T)$log.lambda, colour='grey') +
-      annotate("text", x=subset(df, `lambda-1se` == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
-               label=round(df[subset(df, `lambda-1se` == T)$lambda.at,'mean'],3)) 
   }
@@
     coef.1se = subset(df, lambda.1se == T)$nzcoef
-    coef.m.1se = subset(df, `lambda-1se` == T)$nzcoef
 
@@
 
-    coef.text = arrange(subset(df,nzcoef %in% c(0,  coef.min, coef.1se, coef.m.1se, max(nzcoef) )), -lambda.min, -lambda.1se)
+    coef.text = arrange(subset(df,nzcoef %in% c(0,  coef.min, coef.1se, max(nzcoef) )), -lambda.min, -lambda.1se)
     coef.text = arrange(coef.text[!duplicated(coef.text$nzcoef),], -nzcoef)
@@
 ```{r xvalpt, message=F, warning=F, echo=T, fig.height=16, fig.width=16}
-performance.plots = list(); coefs = list(); deviance.plots = list()
-plots = list()
-folds = 10; splits=5
+coefs = list(); plots = list(); performance.at.1se = list()
+folds = 10; splits = 5
 alpha.values = c(0,0.5,0.7,0.8,0.9,1)
-
 if (file.exists('/tmp/all.pt.alpha.Rdata')) {
@@
     fit0 <- glmnet(dysplasia.df, labels, alpha=a, nlambda=1000, family='binomial') # all patients
-    autoplot(fit0, xvar='lambda', main=paste('fit0, all samples, alpha=',a,sep='')) + theme(legend.position='none')
+    #autoplot(fit0, xvar='lambda', main=paste('fit0, all samples, alpha=',a,sep='')) + theme(legend.position='none')
 
-    cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=fit0$lambda, pts=pts, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance')
+    cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=fit0$lambda, pts=pts, a=a, nfolds=folds, splits=splits, fit=fit0)
 
@@
 
-    #performance.plots[[as.character(a)]] = cv.patient$plot + ggtitle(paste('alpha=',a,sep=''))
-    #deviance.plots[[as.character(a)]] = cv.patient$deviance.plot + ggtitle(paste('alpha=',a,sep=''))
+    plots[[as.character(a)]] = arrangeGrob(cv.patient$plot+ggtitle('Classification'), cv.patient$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
 
-    plots[[as.character(a)]] = arrangeGrob(cv.patient$plot+ggtitle('CV Performance'), cv.patient$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
+    performance.at.1se[[as.character(a)]] = subset(cv.patient$lambdas, lambda.1se == T)
   }
-  save(plots, coefs, file='/tmp/all.pt.alpha.Rdata')
+  save(plots, coefs, performance.at.1se, file='/tmp/all.pt.alpha.Rdata')
 }
@@
 do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits'))
+```
+
+```{r xvalpt2, message=F, warning=F, echo=T}
+
+performance.at.1se = do.call(rbind, performance.at.1se)
+performance.at.1se$alpha = rownames(performance.at.1se)
+
+performance.at.1se = cbind(performance.at.1se, do.call(rbind, lapply(coefs, function(x) cbind('n.Coef'=nrow(x), '75%'=length(which(rowSums(x)/50 >= 0.75))))))
+
+ggplot(performance.at.1se, aes(alpha, mean)) + geom_point() +
+  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + ylim(0.5,0.8) +
+  geom_text(aes(label=n.Coef), nudge_y=0.03) +
+  geom_text(aes(label=paste('(',`75%`,')',sep='')), nudge_y=0.02) +
+  labs(x='Elasticnet penalty value', y='Model performance at lambda.1se', 
+       title=paste('All samples,', folds, 'folds,', splits, 'patient splits'))
 
@@
 
+no.hgd.plots = list(); coefs = list(); performance.at.1se = list()
 if (file.exists('/tmp/nohgd.Rdata')) {
@@
   samples = intersect(rownames(dysplasia.df), subset(info, Pathology %nin% c('HGD', 'IMC'))$Samplename)
+  for (a in alpha.values) {
+    # all patients
+    fitNoHGD <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=1000)
 
-  no.hgd.plots = list()
-  coefs = list()
-  for (a in alpha.values) {
-    fit1 <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=1000) # all patients
-    autoplot(fit1, xvar='lambda', main='fit1, no HGD/IMC samples') + theme(legend.position='none') 
-    lambda1 <- fit1$lambda
+    cv.nohgd = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitNoHGD$lambda, pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fitNoHGD)
 
-    cv.nohgd = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=lambda1, pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fit1, select='deviance')
+    no.hgd.plots[[as.character(a)]] = arrangeGrob(cv.nohgd$plot+ggtitle('Classification'), cv.nohgd$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
 
-    #plots[[as.character(a)]] = cv.nohgd$plot + ggtitle(paste('alpha=',a,sep=''))
+    coef.1se = as.data.frame(non.zero.coef(fitNoHGD, cv.nohgd$lambda.1se))
+    coefs[[as.character(a)]] = coef.stability(coef.1se, cv.nohgd$non.zero.cf)
 
-    no.hgd.plots[[as.character(a)]] = arrangeGrob(cv.patient$plot+ggtitle('CV Performance'), cv.patient$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
-
-    coef.1se = as.data.frame(non.zero.coef(fit1, cv.nohgd$lambda.1se))
-    coefs[[as.character(a)]] = coef.stability(coef.1se, cv.nohgd$non.zero.cf)
+    performance.at.1se[[as.character(a)]] = subset(cv.nohgd$lambdas, lambda.1se == T)
   }
-  save(no.hgd.plots, coefs, file='/tmp/nohgd.Rdata')
+  save(no.hgd.plots, coefs, performance.at.1se, file='/tmp/nohgd.Rdata')
 }
@@
 do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2))
+```
+
+```{r noHGD2, echo=T, message=F, warning=F}
+performance.at.1se = do.call(rbind, performance.at.1se)
+performance.at.1se$alpha = rownames(performance.at.1se)
+
+performance.at.1se = cbind(performance.at.1se, do.call(rbind, lapply(coefs, function(x) cbind('n.Coef'=nrow(x), '75%'=length(which(rowSums(x)/50 >= 0.75))))))
+
+ggplot(performance.at.1se, aes(alpha, mean)) + geom_point() +
+  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + ylim(0.5,0.8) +
+  geom_text(aes(label=n.Coef), nudge_y=0.03) +
+  geom_text(aes(label=paste('(',`75%`,')',sep='')), nudge_y=0.02) +
+  labs(x='Elasticnet penalty value', y='Model performance at lambda.1se', 
+       title=paste('No HGD/IMC samples,', folds, 'folds,', splits, 'patient splits'))
 
@@
 rownames(stability) = names(coef.stable)
-
 ```
@@
 ```{r noLGD, echo=T, message=F, warning=F, fig.height=16, fig.width=16}
-
+nolgd.plots = list(); coefs = list(); performance.at.1se = list()
 if (file.exists('/tmp/nolgd.Rdata')) {
@@
 
-  nolgd.plots = list(); coefs = list()
   for (a in alpha.values) {
-    fit2 <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=1000) # all patients
-    autoplot(fit2, xvar='lambda', main='fit1, no HGD/IMC samples') + theme(legend.position='none') 
-    lambda2 <- fit2$lambda
+    fitNoLGD <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=1000) # all patients
 
-    cv.nolgd = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=lambda2, pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fit2, select='deviance')
+    cv.nolgd = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitNoLGD$lambda, pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fitNoLGD)
 
-    #plots[[as.character(a)]] = cv.nolgd$plot + ggtitle(paste('alpha=',a,sep=''))
+    nolgd.plots[[as.character(a)]] = arrangeGrob(cv.nolgd$plot+ggtitle('Classification'), cv.nolgd$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
 
-    nolgd.plots[[as.character(a)]] = arrangeGrob(cv.nolgd$plot+ggtitle('CV Performance'), cv.nolgd$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
+    coef.1se = as.data.frame(non.zero.coef(fitNoLGD, cv.nolgd$lambda.1se))
+    coefs[[as.character(a)]] = coef.stability(coef.1se, cv.nolgd$non.zero.cf)
 
-    coef.1se = as.data.frame(non.zero.coef(fit2, cv.nolgd$lambda.1se))
-    coefs[[as.character(a)]] = coef.stability(coef.1se, cv.nolgd$non.zero.cf)
+    performance.at.1se[[as.character(a)]] = subset(cv.nolgd$lambdas, lambda.1se == T)
   }
-  save(nolgd.plots, coefs, file='/tmp/nolgd.Rdata')
+  save(nolgd.plots, coefs, performance.at.1se, file='/tmp/nolgd.Rdata')
 }
@@
 do.call(grid.arrange, c(plots, top="No HGD/IMC/LGD samples", ncol=2))
+```
+
+```{r noLGD2, echo=T, message=F, warning=F}
+performance.at.1se = do.call(rbind, performance.at.1se)
+performance.at.1se$alpha = rownames(performance.at.1se)
+
+performance.at.1se = cbind(performance.at.1se, do.call(rbind, lapply(coefs, function(x) cbind('n.Coef'=nrow(x), '75%'=length(which(rowSums(x)/50 >= 0.75))))))
+
+ggplot(performance.at.1se, aes(alpha, mean)) + geom_point() +
+  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + ylim(0.5,0.8) +
+  geom_text(aes(label=n.Coef), nudge_y=0.03) +
+  geom_text(aes(label=paste('(',`75%`,')',sep='')), nudge_y=0.02) +
+  labs(x='Elasticnet penalty value', y='Model performance at lambda.1se', 
+       title=paste('No HGD/IMC/LGD samples,', folds, 'folds,', splits, 'patient splits'))
 
@@
 m = median(subset(sum.patient.data, Status == 'P')$total.samples)
-
 nonps = as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Status == 'NP')$Patient], function(df) df$info$Samplename)))
 
+if (file.exists('/tmp/few.Rdata')) {
+  load('/tmp/few.Rdata', verbose=T)
+} else {
   # Few
@@
   for (a in alpha.values) {
-  fit0 <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, nlambda=500, family='binomial') # all patients
-  cv.back.few = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fit0$lambda, pts=subset(pts, Samplename %in% samples), splits=5, nfolds=20, a=a, fit=fit0, minR=0.1, select='deviance')
+    fitF <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, nlambda=500, family='binomial') # all patients
+    cv.back.few = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitF$lambda, pts=subset(pts, Samplename %in% samples), splits=5, nfolds=20, a=a, fit=fitF, minR=0.1)
     #few.plots[[as.character(a)]] = cv.back.few$plot + ggtitle(paste('alpha=',a,sep=''))
 
-  few.plots[[as.character(a)]] = arrangeGrob(cv.back.few$plot+ggtitle('CV Performance'), cv.back.few$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
+    few.plots[[as.character(a)]] = arrangeGrob(cv.back.few$plot+ggtitle('Classification'), cv.back.few$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
+  }
+  save(few.plots, file='/tmp/few.Rdata')
 } 
@@
 # Many
+if (file.exists('/tmp/many.Rdata')) {
+  load('/tmp/many.Rdata', verbose=T)
+} else {
   samples = c(nonps, as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Status == 'P' & total.samples >= m)$Patient], function(pt) pt$info$Samplename))))
@@
   for (a in alpha.values) {
-  fit0 <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, nlambda=500, family='binomial') # all patients
+    fitM <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, nlambda=500, family='binomial') # all patients
 
-  cv.back.many = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fit0$lambda, pts=subset(pts, Samplename %in% samples), splits=5, nfolds=20, a=a, fit=fit0, select='deviance')
-  #many.plots[[as.character(a)]] = cv.back.many$plot + ggtitle(paste('alpha=',a,sep=''))
+    cv.back.many = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitM$lambda, pts=subset(pts, Samplename %in% samples), splits=5, nfolds=20, a=a, fit=fitM)
 
-  many.plots[[as.character(a)]] = arrangeGrob(cv.back.many$plot+ggtitle('CV Performance'), cv.back.many$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
+    many.plots[[as.character(a)]] = arrangeGrob(cv.back.many$plot+ggtitle('Classification'), cv.back.many$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
   }
+  save(many.plots, file='/tmp/many.Rdata')
+}
+
 do.call(grid.arrange, c(many.plots, ncol=2, top=paste(nrow(subset(sum.patient.data, Status == 'P' & total.samples >= m))," Progressors with >= median # samples (",m,")")))
@@
   info$Prediction = NA
+  info$Prediction.Dev.Resid = NA
   info$PID = unlist(lapply(info$Path.ID, function(x) unlist(strsplit(x, 'B'))[1]))
@@
 } else {
-
   performance.at.1se = c(); coefs = list(); plots = list(); fits = list()
@@
     cv = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitLOO$lambda, 
-                                  pts=subset(pts, Samplename %in% samples), a=a, nfolds=10, splits=5, fit=fitLOO, select='deviance')
+                                  pts=subset(pts, Samplename %in% samples), a=a, nfolds=10, splits=5, fit=fitLOO)
     #plots[[pt]] = cv$plot + ggtitle(pt)
 
-    plots[[pt]] = arrangeGrob(cv$plot+ggtitle('CV Performance'), cv$deviance.plot+ggtitle('Binomial Deviance'), top=pt, ncol=2)
+    plots[[pt]] = arrangeGrob(cv$plot+ggtitle('Classification'), cv$deviance.plot+ggtitle('Binomial Deviance'), top=pt, ncol=2)
 
@@
 
-      pg.samp[[pt]]$Prediction = predict(fitLOO, dysplasia.df[pg.samp[[pt]]$Samplename,], s=cv$lambda.1se, type='response')[,1]
+      logit <- function(p){log(p/(1-p))}
+      inverse.logit <- function(or){1/(1 + exp(-or))}
+
+      pm = predict(fitLOO, dysplasia.df[pg.samp[[pt]]$Samplename,], s=cv$lambda.1se, type='response')
+      sy = sqrt(binomial.deviance(pm, labels[pg.samp[[pt]]$Samplename]))
+
+      #df = cbind.data.frame(pm, sy, 1:length(sy))
+      #colnames(df) = c('pred','dev','x')
+      #ggplot(df, aes(x, pred)) + geom_point() + geom_errorbar(aes(ymin=pred-dev, ymax=pred+dev))
+
+    pg.samp[[pt]]$Prediction = pm[,1]
+    pg.samp[[pt]]$Prediction.Dev.Resid = sy[,1]
+
     } else {
@@
 }
-
-# do.call(grid.arrange, c(lapply(fits[sample(subset(sum.patient.data, Status == 'NP')$Patient, 4)], function(f) {
-#   plt = plot.patient.cv(f$lambdas)   
-#   arrangeGrob(plt$performance +ggtitle('CV Performance'), plt$deviance+ggtitle('Binomial Deviance'), ncol=2)
-# })))
-
-# do.call(grid.arrange, c(lapply(fits[sample(subset(sum.patient.data, Status == 'P')$Patient, 4)], function(f) {
-#   plt = plot.patient.cv(f$lambdas)   
-#   arrangeGrob(plt$performance +ggtitle('CV Performance'), plt$deviance+ggtitle('Binomial Deviance'), ncol=2)
-# })))
 
@@
 
-predicted.progression<-function(x) { if(is.na(x)) return(NA); as.integer(length(which(x == 1)) > 0)  }
+predicted.progression<-function(x, pos=1) { 
+  if(length(which(is.na(x))) == length(x)) return(NA)
+  as.integer(length(which(x == pos)) > 0)  
+}
 
@@
 
-ggplot(p.predictions, aes(y=`%.Predicted`,x=n.Endoscopies)) + geom_bar(stat="identity", fill='dodgerblue3') + #geom_text(aes(label=`%.Predicted`), nudge_y=0.005) + 
+ggplot(p.predictions, aes(y=`%.Predicted`,x=n.Endoscopies)) + 
+  geom_bar(stat="identity", fill='dodgerblue3') + 
+  geom_text(aes(label=paste(`%.Predicted`*100, '%', sep='')), nudge_y=-0.02) + 
   geom_text(aes(label=paste(n.Patients, 'patients')), nudge_y=0.005) + ggtitle("Progression predicted per endoscopy")
@@
 What about the non-progressors that had progressor predictions?
-```{r}
+```{r, warning=F, message=F}
 
@@
 ```
+
+# How do I evaluate the predictions?
+
+I'm a bit at a loss here.
+I understand why penalized regression wouldn't have a CI, but then I'm not sure how best to evaluate the predictions.
+Do I just take the predictions at face value and leave it at that?
+
+```{r, warning=F, message=F, fig.height=4, fig.width=4}
+all = do.call(rbind,lapply(pg.samp, function(df) {
+  df[,c('Status','Prediction','Prediction.Dev.Resid')]
+}))
+
+binomial.deviance
+
+ggplot(all, aes(Prediction.Dev.Resid, Prediction, color=Status)) + geom_point()  + 
+  labs(y='Predicted (response)', x='Binomial dev residual?')
+```
+
+# Randomize the labels
+
+By patient.  This should break the models...and as would be expected, the models are really poor with perfomance at 50% or less (coin toss).
+
+```{r, echo=T, warning=F, message=F}
+
+rand.patients = sum.patient.data
+rand.patients$Status = sample( c('NP','P'), nrow(sum.patient.data), replace=T, prob=c(0.5,0.5))
+
+r.labels = unlist(apply(rand.patients[,c('Patient','Status')],1, function(st) {
+  info = patient.data[[ st['Patient'] ]]$info
+  lbl = rep(as.integer( st['Status'] == 'P' ), nrow(info))
+  names(lbl) = info$Samplename
+  return(lbl)
+}))
+
+# sort in label order
+r.dysplasia.df = t(mergedDf[,names(r.labels)])
+
+r.coefs = list(); r.plots = list()
+folds = 10; splits=5
+for (a in c(0,0.5,1)) {
+  fitR <- glmnet(r.dysplasia.df, r.labels, alpha=a, nlambda=1000, family='binomial') # all patients
+
+  cv.R = crossvalidate.by.patient(x=r.dysplasia.df, y=r.labels, lambda=fitR$lambda, pts=pts, a=a, nfolds=folds, splits=splits, fit=fitR)
+
+  lambda.opt = ifelse( length(cv.R$lambda.1se) > 0, cv.R$lambda.1se, cv.R$lambda.min )
+
+  coef.opt = as.data.frame(non.zero.coef(fitR, lambda.opt))
+  r.coefs[[as.character(a)]] = coef.stability(coef.opt, cv.R$non.zero.cf)
+
+  r.plots[[as.character(a)]] = arrangeGrob(cv.R$plot+ggtitle('Classification'), cv.R$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
+}
+
+do.call(grid.arrange, c(r.plots, top='All samples, Randomized labels (by patient)'))
+
+## check how often the feature(s) is selected at that lamda in each split. "stability selection"
+r.coef.stable = lapply( r.coefs, function(cf) {
+  sort(rowSums(cf[,-1]), decreasing=T)
+})
+
+cfs = as.data.frame(matrix(data=0,nrow=length(unique(unlist(sapply(r.coef.stable, names)))), ncol=length(names(r.coef.stable)), 
+                           dimnames=list(unique(unlist(sapply(r.coef.stable, names))), names(r.coef.stable) )))
+
+for (i in names(r.coef.stable)) 
+  cfs[intersect(rownames(cfs), names(r.coef.stable[[i]])), i] = 1
+
+r.coef.stable = lapply(r.coef.stable, function(cf) cf/(folds*splits))
+
+r.stability = do.call(rbind, lapply(r.coef.stable, function(cf) {
+  cbind('75%'=length(which(cf >= 0.75)),  '50%'=length(which(cf >= 0.5)),  '<50%'=length(which(cf < 0.5)))
+}))
+rownames(r.stability) = names(r.coef.stable)
+
+```
+
+`r pander(table(rowSums(cfs)))`

### 2017-05-16 lib/load_patient_metadata.R

@@ -26,10 +26,9 @@ read.patient.info<-function(file, set='Training') {
   if (grepl('\\.xlsx$', basename(file))) {
     library(xlsx)
     
-    patient.info = NULL
-    test.training = NULL
+    patient.info = NULL;  test.training = NULL
     for (i in 1:length(getSheets(loadWorkbook(file)))) {
-      #print(i)
+      print(i)
       if (names(getSheets(loadWorkbook(file))[i]) == 'Technical Repeats') next
       
       if (names(getSheets(loadWorkbook(file))[i]) == 'TrainingTest') {
@@ -62,7 +61,7 @@ read.patient.info<-function(file, set='Training') {
                  grep('Plate.Index',colnames(ws),ignore.case=T),
                  grep('SLX',colnames(ws),ignore.case=T),
                  grep('cellularity',colnames(ws),ignore.case=T),
-                 grep('p53', colnames(ws),ignore.case=F),
+                 grep('p53.status', colnames(ws),ignore.case=F),
                  grep('Number.of', colnames(ws), ignore.case=F)) ]
       
       
@@ -147,6 +146,7 @@ read.patient.info<-function(file, set='Training') {
     message(paste("Returning all patient data."))
   }
 
+  patient.info$Patient = gsub('/', '_',patient.info$Patient)
   return(patient.info)
 }
 

### 2017-05-17 Clustering.Rmd
…

@@
 library(plyr)
-library(xlsx)
 library(dplyr)
@@
 
-load('patients.Rdata', verbose=T)
+load('Training_patients.Rdata', verbose=T)
 
@@
 
-```{r pca, echo=F, warning=F, message=F, fig.width=8, fig.height=4}
+```{r pca, echo=F, warning=F, message=F, fig.width=8, fig.height=4, eval=F}
 sum.patient.data = arrange(sum.patient.data, Status, total.samples)
@@
   info = arrange(patient.data[[pt]]$info, Endoscopy.Year)
+  if (length(info$Samplename) <= 1) return(NA)
+
   info = info[which(info$Samplename %in% colnames(patient.data[[pt]][['seg.vals']])[-(1:5)]),]
@@
     x1 = patient.data[[pt]][[sv]][, info$Samplename]
-    x1 = x1[,info$Samplename]
+
+    if (length(info$Samplename) <= 1) { 
+      x1 = as.data.frame(x1)
+      colnames(x1) = info$Samplename
+    } else {
+      x1 = x1[, info$Samplename]
+    }
+
     colnames(x1) = paste(info$Endoscopy.Year, '(', info$Pathology, ')', sep='' )
 
-    pca = prcomp(t(x1), scale=T)
+    #pca = prcomp(t(x1), scale=length(info$Samplename) > 2)
+    pca = prcomp(t(x1), scale=length(info$Samplename) > 2)
 
@@
 
-```{r echo=F, message=F, warning=F, pcaplots, fig.align='left', fig.height=30, fig.width=20}
+```{r echo=F, message=F, warning=F, pcaplots, fig.align='left', fig.height=30, fig.width=20, eval=F}
 grid.arrange(grobs=pplots, nrow=length(subset(sum.patient.data, Status == 'P')$Patient), top="Progressors")

### 2017-05-17 Patient Report.Rmd
…

@@
 ## Patient info file
-patient.file = grep('/All_patient_info.xls', data.files, value=T)
+patient.file = grep('/All_patient_info.txt', data.files, value=T)
 if (length(patient.file) != 1)
@@
 patient.info = read.patient.info(patient.file, set=dataset)
-patient.info$Patient = gsub("/", "_", patient.info$Patient)
 head(patient.info)
-
-bad_batch = which(patient.info$Batch.Name %in% c('Batch_8', 'Batch_9'))
-patient.info = patient.info[-(bad_batch),]
 
@@
 
-lowprobes = c()
-
-patient.data = list()
+cachefile = paste(dataset, '_patients.Rdata', sep='')
+if ( file.exists(cachefile) ) {
+  load(cachefile, verbose=T)
+} else {
+  patient.data = list(); lowprobes = c()
 
@@
 
-  if(!file.exists(paste(patient.plot.dir, "/",patient.name,"_segmentedCoverage_fitted_gamma",gamma2,".txt",sep=""))) next
+    if (!file.exists(paste(patient.plot.dir, "/",patient.name,"_segmentedCoverage_fitted_gamma",gamma2,".txt",sep=""))) {
+      warning(paste("Missing segmented file for", patient.name))
+      next
+    }
 
     segvals = read.table(paste(patient.plot.dir, "/",patient.name,"_segmentedCoverage_fitted_gamma",gamma2,".txt",sep=""),sep="\t",stringsAsFactors=F,header=T)
+    dim(segvals)
 
@@
     if (length(missing) > 0) {
-    warning(paste("Missing", length(missing), "sample(s) in patient", patient.name))
-    next
+      warning(paste("Missing", length(missing), "sample(s) of", length(subset(patient.info, Patient == patient.name)$Samplename), "in patient", patient.name))
+      #next
     }
@@
     probes = length(which(segvals$n.probes < min.probes))
-  message( paste(probes, ' probes (', round(probes/nrow(segvals))*100, '%)',' below the minimum probe count (',min.probes,')', sep=''))
+    message( paste(probes, ' probes (', round(probes/nrow(segvals),2)*100, '%)',' below the minimum probe count (',min.probes,')', sep=''))
     if (probes/nrow(segvals) >= 0.75) lowprobes = c(lowprobes, patient.name)
@@
     if (nrow(segvals) < 100) {
-    warning(paste(patient.name, 'has fewer than 100 genomic reginos with segmented values:', nrow(segvals)))
-    next
+      warning(paste(patient.name, 'has fewer than 100 genomic regions with segmented values:', nrow(segvals)))
+      #next
     }
@@
     sample.cols = grep('^D\\d+_D\\d+', colnames(segvals))
-  if (nrow(segvals) > 0) {
+
+    if (length(sample.cols) > 0) {
+      if (length(sample.cols) > 2) {  
         HC = hclust(dist(t(segvals[,sample.cols])))
@@
         #print(gg1)
+      }
 
@@
       normalised.segvals = segvals[,sample.cols]
+      if (length(sample.cols) > 1) {  
         for(c in 1:nrow(normalised.segvals)) 
           normalised.segvals[c,] = (normalised.segvals[c,]-mean(unlist(normalised.segvals[c,])))/sd(unlist(normalised.segvals[c,]))
-
         patient.data[[patient.name]][['norm.seg.vals']] = normalised.segvals
-
-      # Variable Regions
-      sds = apply(segvals[,sample.cols],1,sd)
-      means = apply(segvals[,sample.cols],1,mean)
-      CofV = sds/means
-
-      variable.region.indices = which(sds >= sd.threshold & segvals$n.probes >= min.probes)
-      variable.regions = segvals[variable.region.indices, intersect(patient.data[[patient.name]]$info$Samplename, colnames(segvals))] 
-      patient.data[[patient.name]][['variable.regions']] = cbind(segvals[variable.region.indices,c(1:5)],variable.regions )
 
@@
           geom_histogram( fill=I("blue"),col=I("lightblue"),binwidth=0.3) + labs(title=paste(patient.name,'normalised segment coverage'), x='', y='')
+      }
 
-      #if(nrow(segvals)>=3) {
-        # order by date of endoscopy
-        patient = subset(patient.info, Patient == patient.name)
-        patient = arrange(patient, Endoscopy.Year, Pathology, Samplename)
-       # }
-        sample.normalised.segvals = normalised.segvals[, intersect(colnames(normalised.segvals), patient$Samplename)]
+      # Variable Regions can only be called relative to each sample so if there's only one sample we can't make a call
+      if (length(sample.cols) > 1) {
+        sds = apply(segvals[,sample.cols],1,sd)
+        means = apply(segvals[,sample.cols],1,mean)
+        CofV = sds/means
+
+        variable.region.indices = which(sds >= sd.threshold & segvals$n.probes >= min.probes)
+        variable.regions = segvals[variable.region.indices, intersect(patient.data[[patient.name]]$info$Samplename, colnames(segvals))] 
+        patient.data[[patient.name]][['variable.regions']] = cbind(segvals[variable.region.indices,c(1:5)],variable.regions )
+      }
+
+      patient = arrange(subset(patient.info, Patient == patient.name),  Endoscopy.Year, Pathology, Samplename)
     }
   }
-
+save(patient.data, file=paste(dataset, '_patients.Rdata', sep=''))
+}
 print(paste("Missing patients:", paste(setdiff(sum.patient.data$Patient, names(patient.data)), collapse=', ')))
@@
 }
-
-save(patient.data, file=paste(dataset, '_patients.Rdata', sep=''))
 
@@
 
-chrs = read.table('hg19_chrom.txt',sep='\t', header=T)
+chr.info = read.table('hg19_chrom.txt',sep='\t', header=T)
 
 # Segment lengths, variable regions
-var.seg.lengths = lapply(var.seg.lengths, function(x) merge(x, chrs, by.x='chr', by.y='chr') )
+var.seg.lengths = lapply(var.seg.lengths, function(x) merge(x, chr.info, by.x='chr', by.y='chr') )
 
 # All segments
-segment.lengths = lapply(segment.lengths, function(x) merge(x, chrs, by.x='chr', by.y='chr') )
+segment.lengths = lapply(segment.lengths, function(x) merge(x, chr.info, by.x='chr', by.y='chr') )
 
@@
 genome.coverage = sapply(segment.lengths, function(x) {
-  sum(as.numeric(x[['seg.len']]))/sum(as.numeric(chrs$length))
+  sum(as.numeric(x[['seg.len']]))/sum(as.numeric(chr.info$length))
 })
@@
 var.seg.lengths =  lapply(var.seg.lengths, function(df) {
-  as.data.frame(df %>%
-  group_by(chr, length) %>%
-  summarise(seg.lengths = sum(seg.len), 
-            chr.ratio = sum(seg.len)/length[1]))
+   ddply(df, .(chr, length), summarise, seg.lengths = sum(seg.len), chr.ratio = sum(seg.len)/length[1])
 })
@@
   regions = patient.data[[patient.name]]$variable.regions[,c('chrom','arm','start.pos','end.pos')]
-  if (nrow(regions) <= 0) next
+  if (is.null(regions) || nrow(regions) <= 0) next
 
@@
 
-gene.cnt = var.regions %>%
-  group_by(Gene.Symbols, CN) %>%
-  summarise(Progressors=length(which(status == 'P')), Non.Progressors=length(which(status == 'NP')) )
+gene.cnt = ddply(var.regions, .(Gene.Symbols, CN), summarise, Progressors=length(which(status == 'P')), Non.Progressors=length(which(status == 'NP')))
 gene.cnt[,unique(var.regions$patient.name)] = 0
@@
 m = melt(gene.p[,c(1:4)], id.vars=c('Gene.Symbols', 'CN'))
-m$Gene.Symbols = factor(m$Gene.Symbols, levels= m$Gene.Symbols[c(nrow(m):1)] ) # order
+m$Gene.Symbols = factor(m$Gene.Symbols, levels=unique(m$Gene.Symbols[c(nrow(m):1)]) ) # order
 m$CN = factor(m$CN)
@@
   geom_tile(aes(fill=value), color='white') + 
-  scale_fill_gradient('n patients',low='white',high='red') + labs(x="",y='CGC Genes')
-# + geom_point(data=m[m$CN == 'gain' & m$value > 0,], aes(y=Gene.Symbols, x=variable), size=2, shape=24, fill='red', color='red') +
-#  geom_point(data=m[m$CN == 'loss' & m$value > 0,], aes(y=Gene.Symbols, x=variable), color='green4', size=2, shape=25, fill='green4')
+  scale_fill_gradient('n patients',low='white',high='red') + labs(x="",y='CGC Genes') +
+  geom_point(data=m[m$CN == 'gain' & m$value > 0,], aes(y=Gene.Symbols, x=variable), size=2, shape=24, fill='red', color='red') +
+  geom_point(data=m[m$CN == 'loss' & m$value > 0,], aes(y=Gene.Symbols, x=variable), color='green4', size=2, shape=25, fill='green4')
 
-gc = gene.cnt %>% group_by(Gene.Symbols) %>% summarise(Progressors=sum(Progressors), Non.Progressors=sum(Non.Progressors))
+gc = ddply(gene.cnt, .(Gene.Symbols), summarise, Progressors=sum(Progressors), Non.Progressors=sum(Non.Progressors))
 gc = arrange(subset(gc, Progressors >= Non.Progressors), -Progressors)
@@
 #m$value = as.factor(m$value)
-m$Gene.Symbols = factor(m$Gene.Symbols, levels= m$Gene.Symbols[c(nrow(m):1)] ) # order
+m$Gene.Symbols = factor(m$Gene.Symbols, levels=unique(m$Gene.Symbols[c(nrow(m):1)]) ) # order
 ggplot(m, aes(y=Gene.Symbols, x=variable)) + 
@@
 for (name in subset(sum.patient.data, Status == 'P')$Patient) {
+  var.regions = patient.data[[name]]$variable.regions
+  if ( is.null(var.regions) || nrow(var.regions) == 0 ) next
+
   sn = which(colnames(patient.data[[name]]$variable.regions)[-(1:5)] %in% subset(patient.data[[name]]$info, grepl('HGD|IMC', Pathology))$Samplename)
@@
 
-This is just looking at the distribution of values per sample. The average ploidy per sample is around 1.  Currently we're calling variable regions based on this.
-```{r, warning=F}
+```{r, warning=F, eval=F}
+#This is just looking at the distribution of values per sample. The average ploidy per sample is around 1.  Currently we're calling variable regions based on this.
 ks = list()
 for (pt in names(patient.data)) {
-  cols = intersect(arrange(patient.data[[pt]]$info, Pathology, Endoscopy.Year)$Samplename, colnames(patient.data[[pt]]$seg.vals[,-(1:5)]))
+  print(pt)
+  cols = intersect(arrange(patient.data[[pt]]$info, Pathology, Endoscopy.Year)$Samplename, grep('^D\\d+', colnames(patient.data[[pt]]$seg.vals), value=T))
+  if (length(cols) <= 1) next
+
   segvals = patient.data[[pt]]$seg.vals[,cols]

### 2017-05-17 RegressionPvNP.Rmd
…

@@
   x1 = segdata[,intersect(colnames(segdata), samples)]
+
+  if (length(samples) <= 1) { 
+    x1 = as.data.frame(x1)
+    colnames(x1) = samples
+  } else {
     x1 = x1[, samples]
+  }
   rownames(x1) = (segdata[,c(1:4)] %>%
@@
   info = pt$info
+  print(unique(info$Patient))
+  info$SampleSD = NA
+  info$SampleMEAN = NA
+  if (length(info$Samplename) > 1) {
     info$SampleSD = apply(pt$seg.vals[,info$Samplename], 2, sd)
     info$SampleMEAN =  apply(pt$seg.vals[,info$Samplename], 2, mean)
+  }
   info$Prediction = NA
@@
     samples = as.vector(unlist(sapply(tmp.patient.data, function(df) df$info$Samplename )))
-    length(samples)
+
+    train.rows = which(rownames(dysplasia.df) %in% samples)
+    training = dysplasia.df[train.rows,]
+    test = dysplasia.df[-train.rows,]
+
+    # Predict function giving me difficulty when I have only a single sample, this ensures the dimensions are the same
+    sparsed_test_data <- Matrix(data=0, nrow=ifelse(nrow(test) > 1, nrow(test), 1),  ncol=ncol(training),
+                            dimnames=list(pg.samp[[pt]]$Samplename,colnames(training)), sparse=T)
+    for(i in colnames(test)) sparsed_test_data[,i] = test[,i]
 
@@
     a = 1
-    fitLOO <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=1000) # all patients
-    cv = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitLOO$lambda, 
+    fitLOO <- glmnet(training, labels[train.rows], alpha=a, family='binomial', nlambda=1000) # all patients
+    cv = crossvalidate.by.patient(x=training, y=labels[train.rows], lambda=fitLOO$lambda, 
                                   pts=subset(pts, Samplename %in% samples), a=a, nfolds=10, splits=5, fit=fitLOO)
@@
 
-      pm = predict(fitLOO, dysplasia.df[pg.samp[[pt]]$Samplename,], s=cv$lambda.1se, type='response')
+      pm = predict(fitLOO, newx=sparsed_test_data, s=cv$lambda.1se, type='response')
       sy = sqrt(binomial.deviance(pm, labels[pg.samp[[pt]]$Samplename]))

### 2017-05-17 lib/load_patient_metadata.R

@@ -4,18 +4,17 @@ suppressPackageStartupMessages( library(dplyr) )
 strip.whitespace <- function (x) gsub("\\s+|\\s+", "", x)
 
 summarise.patient.info<-function(df) {
-  sum.pt = (df %>% 
-              group_by(Patient, Status, Set) %>%
-              summarise(
-                start.year=range(Endoscopy.Year)[1],
-                end.year=range(Endoscopy.Year)[2],
-                total.samples=length(Samplename), 
-                total.endos=length(unique(Path.ID)),
-                highest.path=sort(Pathology, decreasing=T)[1],
-                first.batch=sort(Batch.Name)[1],
-                med.cellularity=median(Barretts.Cellularity,na.rm=T),
-                Initial.Analysis=(sort(Batch.Name)[1] %in% levels(patient.info$Batch.Name)[1:5])
-              ))
+  require(plyr)
+  
+  sum.pt = ddply(df, .(Patient, Status, Set), summarise, 
+        start.year=range(Endoscopy.Year)[1], 
+        end.year=range(Endoscopy.Year)[2],
+        total.samples=length(Samplename), 
+        total.endos=length(unique(Path.ID)),
+        highest.path=sort(Pathology, decreasing=T)[1],
+        first.batch=sort(Batch.Name)[1],
+        med.cellularity=median(Barretts.Cellularity,na.rm=T),
+        Initial.Analysis=(sort(Batch.Name)[1] %in% levels(patient.info$Batch.Name)[1:5]))
   return(arrange(sum.pt, Status, total.samples, highest.path))
 }
 

### 2017-05-23 Patient Report.Rmd
Starting to firm up analysis

@@
 author: "Sarah Killcoyne"
-date: "15 November 2016"
+date: "23 May 2017"
 output:
@@
 library(plyr)
-#library(xlsx)
+library(RColorBrewer)
 library(dplyr)
 library(pander)
-#library(Hmisc)
 library(ggdendro)
@@
 
-```{r, hmm, echo=F, message=F, warning=F, fig.height=40, fig.width=10}
+```{r, hmm, echo=F, message=F, warning=F, fig.height=100, fig.width=10}
 ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
@@
 
-m = melt(gene.p[,c(1:4)], id.vars=c('Gene.Symbols', 'CN'))
-m$Gene.Symbols = factor(m$Gene.Symbols, levels=unique(m$Gene.Symbols[c(nrow(m):1)]) ) # order
+gene.p = gene.p[,(1:4)]
+m = melt(gene.p, id.vars=c('Gene.Symbols', 'CN'))
+m = transform(m, Gene.Symbols = reorder(Gene.Symbols, value))
+
+#m$Gene.Symbols = factor(m$Gene.Symbols, levels=unique(m$Gene.Symbols[c(nrow(m):1)]) ) # order
 m$CN = factor(m$CN)
 
+m$brks <- cut(m$value, include.lowest=T, right=F,
+                   breaks=c(0, 1, 3, 5, 8, 10),
+                   labels=c('0','1-3', '4-5', '6-8','9-10'))
+m = transform(m, Gene.Symbols = reorder(Gene.Symbols, value))
+
+myP = c('white',brewer.pal(n = length(levels(m$brks)), name = "Reds"))
+
 ggplot(m, aes(y=Gene.Symbols, x=variable)) + 
-  geom_tile(aes(fill=value), color='white') + 
-  scale_fill_gradient('n patients',low='white',high='red') + labs(x="",y='CGC Genes') +
-  geom_point(data=m[m$CN == 'gain' & m$value > 0,], aes(y=Gene.Symbols, x=variable), size=2, shape=24, fill='red', color='red') +
-  geom_point(data=m[m$CN == 'loss' & m$value > 0,], aes(y=Gene.Symbols, x=variable), color='green4', size=2, shape=25, fill='green4')
+  geom_tile(aes(fill=brks), color='white') + scale_fill_manual( values = myP, "Patient Counts") +
+  geom_point(data=subset(m, value > 0), aes(y=Gene.Symbols, x=variable, shape=CN, color=CN), size=3 ) +
+  scale_color_manual("CN", values=c("blue", "yellow")) + scale_shape_manual(values=c(17,20)) +
+  labs(title="All available COSMIC genes mapped", x='', y='')
 
-gc = ddply(gene.cnt, .(Gene.Symbols), summarise, Progressors=sum(Progressors), Non.Progressors=sum(Non.Progressors))
-gc = arrange(subset(gc, Progressors >= Non.Progressors), -Progressors)
-gc = gc[1:60,]
-m = melt(gc, id.vars=c('Gene.Symbols'))
-#m$value = as.factor(m$value)
-m$Gene.Symbols = factor(m$Gene.Symbols, levels=unique(m$Gene.Symbols[c(nrow(m):1)]) ) # order
+```
+
+Just the top 50 genes by the number of patients that share variants in these regions.
+
+```{r echo=F, warning=F, message=F, fig.height=12, fig.width=8}
+
+m = melt(gene.p[1:50,], id.vars=c('Gene.Symbols', 'CN'))
+m = transform(m, Gene.Symbols = reorder(Gene.Symbols, value))
+
+m$brks <- cut(m$value, include.lowest=T, right=F,
+                   breaks=c(0, 1, 3, 5, 8, 10),
+                   labels=c('0','1-3', '4-5', '6-8','9-10'))
+
 ggplot(m, aes(y=Gene.Symbols, x=variable)) + 
-  geom_tile(aes(fill=value), color='white') + 
-  scale_fill_gradient('n patients',low='white',high='red') + labs(x="",y='CGC Genes') + ggtitle("Top 50 genes only")
-
+  geom_tile(aes(fill=brks), color='white') + scale_fill_manual( values = myP, "Patient Counts") +
+  geom_point(data=subset(m, value > 0), aes(y=Gene.Symbols, x=variable, shape=CN, color=CN), size=3 ) +
+  scale_color_manual("CN", values=c("blue", "yellow")) + scale_shape_manual(values=c(17,20)) +
+  labs(title="Top 50 genes by number of patients", x='', y='')
 ```

### 2017-05-23 RegressionPvNP.Rmd
Starting to firm up analysis

@@
 author: "Sarah Killcoyne"
-date: "31 March 2017"
+date: "23 May 2017"
 output: 
   html_document: 
-    fig_caption: yes
-    fig_height: 10
-    fig_width: 10
-    number_sections: yes
     toc: yes
+    toc_depth: 4
 ---
@@
 data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
-plot.dir = paste(data, 'Analysis/multipcf_plots_fitted_perPatient', sep='/')
-
-if (length(list.files(plot.dir)) <= 0)
-  stop(paste("No analysis files found in", plot.dir ))
+analysis.files = list.files(paste(data, 'Analysis', sep='/'), full.names=T)
 
@@
 
-```{r apclust-func, echo=F, message=F, warning=F}
-apclust.data<-function(segdata, samples) {
-  x1 = segdata[,intersect(colnames(segdata), samples)]
-
-  if (length(samples) <= 1) { 
-    x1 = as.data.frame(x1)
-    colnames(x1) = samples
-  } else {
-    x1 = x1[, samples]
-  }
-  rownames(x1) = (segdata[,c(1:4)] %>%
-                    rowwise() %>%
-                    mutate(location=paste(paste(chrom, arm, sep=''), '.', start.pos, '-', end.pos, sep='')))$location
-  return(x1)  
-}
-```
-
 The previous analysis employed AP clustering on a per-patient basis to look for evidence that it was possible to see differences between the progressors and non-progressors, and to quantify that in some sort of complexity measure.
@@
 ```{r tile, echo=F, message=F, warning=F, include=F}
-
-chr.lengths = get.chr.lengths()
-chr.lengths$chrom = sub('chr','',chr.lengths$chrom)
-chr.lengths$start = 1
-
-tile.w=1e7
-
-genome = makeGRangesFromDataFrame(chr.lengths[1:22,], seqnames.field = 'chrom', end.field='chr.length')
-tiles = tile(genome, width=tile.w)
-
-tile.sample.segments<-function(tileFile, sample.data) {
-  if (file.exists(tileFile)) {
-    message(paste("Loading", tileFile))
-    load(tileFile, verbose=T)
-  } else {
-    message("Tiling patient data...")
-    grList = lapply(sample.data, function(df) {
-      print(unique(df$info$Patient))
-      df$info = arrange(df$info, Endoscopy.Year, Pathology)
-
-      x1 = apclust.data(df$seg.vals, df$info$Samplename)
-
-      segnames = as.data.frame(do.call(rbind, sapply(rownames(x1), strsplit, '[p|q].|-')))
-      colnames(segnames) = c('chr','start','end')
-      segnames[c('start','end')] = lapply(segnames[c('start','end')], function(x) as.numeric(as.character(x)))
-
-      pt = unique(df$info$Patient)
-      return( makeGRangesFromDataFrame(cbind(segnames,pt,x1), keep.extra.columns=T) )
-    })
-
-    sampleNames = unlist(sapply(sample.data, function(df) {
-      df$info = arrange(df$info, Endoscopy.Year, Pathology)
-      x1 = apclust.data(df$seg.vals, df$info$Samplename)
-      colnames(x1)
-    }))
-
-    tile.segments<-function(tiles, gr, mergedSegments) {
-      ov = findOverlaps(tiles, gr)
-      for (qh in unique(queryHits(ov))) {
-        #print(qh)
-        currentTile = tiles[[qh]]
-        curov = findOverlaps(currentTile, gr)
-
-        for (i in 1:length(curov)) {
-          segment = currentTile[ queryHits(curov)[i]  ]
-          #print(segment)
-          rows = with(mergedDf, which( chr==as.character(seqnames(segment)) & start == start(segment) & end == end(segment)))
-          segmentVals = as.data.frame(elementMetadata(gr[ subjectHits(curov)[i] ]))
-          mergedSegments[rows, names(segmentVals)[-1]] = segmentVals[,-1]
-        }
-      }
-      return(mergedSegments)  
-    }
-
-    mergedDf = do.call(rbind, lapply(tiles, function(tile) { 
-      cbind('chr'=as.character(seqnames(tile)), as.data.frame(ranges(tile))[1:2]) 
-    }) )
-    mergedDf[,sampleNames] = NA
-
-    for (gr in grList) {
-      #print(unique(gr$pt))
-      mergedDf = tile.segments(tiles, gr, mergedDf)
-    }
-
-    #tmp = mergedDf
-    mergedDf[is.na(mergedDf)] = 0
-    rownames(mergedDf) = with(mergedDf, paste(chr, ':', start, '-', end, sep=''))
-
-    save(mergedDf, file=tileFile)
-  }
-  return(mergedDf)
-}
-
-mergedDf = tile.sample.segments(paste(as.character(tile.w),'_tiledpts.Rdata', sep=''), patient.data)
+#tile.w=1e07
+tile.w=5e06
+tile.files = grep('tile_.*5e\\+06', analysis.files, value=T)
+load(tile.files[1], verbose=T)
 dim(mergedDf)
+
+cache.dir = paste(data, 'Analysis/5e06', sep='/')
+if (!dir.exists(cache.dir)) dir.create(cache.dir)
+
 #validatioMergedDf = tile.sample.segments(paste(as.character(tile.w),'validation_tiledpts.Rdata', sep=''), validation.patient.data)
@@
 alpha.values = c(0,0.5,0.7,0.8,0.9,1)
-if (file.exists('/tmp/all.pt.alpha.Rdata')) {
-  load('/tmp/all.pt.alpha.Rdata', verbose=F)
+
+file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
+if (file.exists(file)) {
+  load(file, verbose=T)
 } else {
@@
   }
-  save(plots, coefs, performance.at.1se, file='/tmp/all.pt.alpha.Rdata')
+  save(plots, coefs, performance.at.1se, file=file)
 }
@@
 
-ggplot(performance.at.1se, aes(alpha, mean)) + geom_point() +
+ggplot(performance.at.1se, aes(alpha, mean)) + geom_point() + geom_text( aes(label=round(mean, 3)), hjust=-0.5) +
   geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + ylim(0.5,0.8) +
@@
 rownames(stability) = names(coef.stable)
+
+pander(as.data.frame(sort(names(which((coef.stable[['1']]) >= 0.75)))), justify='left')
 ```
@@
 
+### Feature Information
+
+```{r, warning=F, message=F}
+#topF = do.call(rbind, strsplit( names(which((coef.stable[['1']]) >= 0.75)), ':|-'))
+#colnames(topF) = c('chr','start','end')
+#topF = apply(topF, 2, as.numeric)
+
+topF = names(which((coef.stable[['1']]) >= 0.75))
+
+library(biomaRt)
+mart = biomaRt::useMart("ENSEMBL_MART_ENSEMBL", host="grch37.ensembl.org")
+biomart = useDataset("hsapiens_gene_ensembl", mart)
+
+attr = c('ensembl_gene_id','hgnc_symbol','start_position','end_position', 'chromosome_name', 'go_id', 'entrezgene')
+bm = getBM(mart=biomart, attributes=attr, filters=list('chromosomal_region' = topF))
+bm = bm[-which(bm$hgnc_symbol == ""),] 
+#head(bm)
+
+bm2 = ddply(bm, .(entrezgene, ensembl_gene_id, hgnc_symbol), summarise, 'chrom'=unique(chromosome_name), 'start'=min(start_position), 'end'=max(end_position), 'GO.IDs'=paste(go_id, collapse=','))
+
+ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
+
+bm2 = subset(bm2, hgnc_symbol %in% ccgenes$Gene.Symbol)
+
+features.in.prog = dysplasia.df[names(which(labels == 1)),topF]
+
+#features.in.prog <= 0.9
+#features.in.prog >= 1.1
+
+```
+
 ### Remove HGD/IMC Samples
@@
 no.hgd.plots = list(); coefs = list(); performance.at.1se = list()
-if (file.exists('/tmp/nohgd.Rdata')) {
-  load('/tmp/nohgd.Rdata', verbose=T)
+
+file = paste(cache.dir, 'nohgd.Rdata', sep='/')
+if (file.exists(file)) {
+  load(file, verbose=T)
 } else {
@@
   }
-  save(no.hgd.plots, coefs, performance.at.1se, file='/tmp/nohgd.Rdata')
+  save(no.hgd.plots, coefs, performance.at.1se, file=file)
 }
@@
 
-ggplot(performance.at.1se, aes(alpha, mean)) + geom_point() +
+ggplot(performance.at.1se, aes(alpha, mean)) + geom_point() + geom_text( aes(label=round(mean, 2)), hjust=-0.5) +
   geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + ylim(0.5,0.8) + 
@@
 ```{r noLGD, echo=T, message=F, warning=F, fig.height=16, fig.width=16}
+file = paste(cache.dir, 'nolgd.Rdata', sep='/')
+
 nolgd.plots = list(); coefs = list(); performance.at.1se = list()
-if (file.exists('/tmp/nolgd.Rdata')) {
-  load('/tmp/nolgd.Rdata', verbose=T)
+if (file.exists(file)) {
+  load(file, verbose=T)
 } else {
@@
   }
-  save(nolgd.plots, coefs, performance.at.1se, file='/tmp/nolgd.Rdata')
-}
-
-do.call(grid.arrange, c(plots, top="No HGD/IMC/LGD samples", ncol=2))
+  save(nolgd.plots, coefs, performance.at.1se, file=file)
+}
+
+do.call(grid.arrange, c(nolgd.plots, top="No HGD/IMC/LGD samples", ncol=2))
 ```
@@
 
-ggplot(performance.at.1se, aes(alpha, mean)) + geom_point() +
+ggplot(performance.at.1se, aes(alpha, mean)) + geom_point() + geom_text( aes(label=round(mean, 2)), hjust=-0.5) +
   geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + ylim(0.5,0.8) +
@@
 
-if (file.exists('/tmp/few.Rdata')) {
-  load('/tmp/few.Rdata', verbose=T)
+file = paste(cache.dir, 'few.Rdata', sep='/')
+
+if (file.exists(file)) {
+  load(file, verbose=T)
 } else {
@@
   }
-  save(few.plots, file='/tmp/few.Rdata')
+  save(few.plots, file=file)
 } 
@@
 # Many
-if (file.exists('/tmp/many.Rdata')) {
-  load('/tmp/many.Rdata', verbose=T)
+file = paste(cache.dir, 'many.Rdata', sep='/')
+
+if (file.exists(file)) {
+  load(file, verbose=T)
 } else {
@@
   }
-  save(many.plots, file='/tmp/many.Rdata')
+  save(many.plots, file=file)
 }
@@
 
-if (file.exists('/tmp/loo.Rdata')) {
-  load('/tmp/loo.Rdata', verbose=T)
+file = paste(cache.dir, 'loo.Rdata', sep='/')
+
+if (file.exists(file)) {
+  load(file, verbose=T)
 } else {
@@
   for (pt in names(pg.samp)) {
+    print(pt)
     tmp.patient.data = patient.data[subset(sum.patient.data, Patient != pt)$Patient]
@@
     training = dysplasia.df[train.rows,]
-    test = dysplasia.df[-train.rows,]
+    test = as.matrix(dysplasia.df[-train.rows,])
+    if ( nrow(test) == ncol(dysplasia.df) ) test = t(test)
 
     # Predict function giving me difficulty when I have only a single sample, this ensures the dimensions are the same
-    sparsed_test_data <- Matrix(data=0, nrow=ifelse(nrow(test) > 1, nrow(test), 1),  ncol=ncol(training),
+    sparsed_test_data <- Matrix(data=0, nrow=ifelse(length(pg.samp[[pt]]$Samplename) > 1, nrow(test), 1),  ncol=ncol(training),
                                 dimnames=list(pg.samp[[pt]]$Samplename,colnames(training)), sparse=T)
-    for(i in colnames(test)) sparsed_test_data[,i] = test[,i]
+    for(i in colnames(dysplasia.df)) sparsed_test_data[,i] = test[,i]
 
@@
       pm = predict(fitLOO, newx=sparsed_test_data, s=cv$lambda.1se, type='response')
-      sy = sqrt(binomial.deviance(pm, labels[pg.samp[[pt]]$Samplename]))
+      sy = as.matrix(sqrt(binomial.deviance(pm, labels[pg.samp[[pt]]$Samplename])))
 
@@
   }
-  save(plots, performance.at.1se, coefs, fits, pg.samp, file='/tmp/loo.Rdata')
+  save(plots, performance.at.1se, coefs, fits, pg.samp, file=file)
 }
@@
 
+hgd = do.call(rbind, lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], function(pt) {
+  hgd = subset(pt, Pathology %in% c('HGD', 'IMC'))
+  cbind.data.frame('correctly.predicted'=sum(as.integer(hgd$Prediction > 0.5)), 'total.hgd.samples'=nrow(hgd))
+}))
+rownames(hgd) = subset(sum.patient.data, Status == 'P')$Patient
+hgd = subset(hgd, total.hgd.samples > 0) # Some patients had no hgd samples
+
+hgd$pred.ratio = hgd$correctly.predicted/hgd$total.hgd.samples
+
+hgd$Patients = rownames(hgd)
+hgd = transform(hgd, Patients=reorder(Patients, -pred.ratio) )
+
+ggplot(hgd[,c('total.hgd.samples','pred.ratio', 'Patients')], aes(x=Patients, y=pred.ratio)) + 
+  geom_bar(stat='identity', fill='darkblue', alpha=0.6) + geom_label( aes(label=total.hgd.samples) ) +
+  coord_flip() + labs(title='Correctly predicted HGD samples', y='Correct prediction ratio')
+
+lgd = do.call(rbind, lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], function(pt) {
+  e = subset(pt, Pathology  == 'LGD')
+  cbind.data.frame('predicted'=sum(as.integer(e$Prediction > 0.5)), 'total.lgd'=nrow(e))
+}))
+lgd = subset(lgd, total.lgd > 0)
+
+lgd$pred.ratio = round(lgd$predicted/lgd$total.lgd, 2)
+
+lgd$Patients = rownames(lgd)
+lgd = transform(lgd, Patients=reorder(Patients, -pred.ratio) )
+
+ggplot(lgd[,c('total.lgd','pred.ratio', 'Patients')], aes(x=Patients, y=pred.ratio)) + 
+  geom_bar(stat='identity', fill='darkgreen', alpha=0.6) + geom_label( aes(label=total.lgd) ) +
+  coord_flip() + labs(title='Correctly predicted LGD samples', y='Correct prediction ratio')
+
 # All samples
@@
 
-np = sapply(all.predictions[subset(sum.patient.data, Status == 'NP')$Patient], function(x) round(sum(x)/length(x),2))
-pg = sapply(all.predictions[subset(sum.patient.data, Status == 'P')$Patient], function(x) round(sum(x)/length(x),2))
+all.predictions = do.call(rbind, lapply(all.predictions, function(x) {
+  cbind.data.frame('correctly.predicted'=sum(x), 'total.samples'=length(x))
+}))
+
+all.predictions$Status = ifelse (rownames(all.predictions) %in% subset(sum.patient.data, Status == 'P')$Patient, 'P', 'NP')
+
+#ggplot(melt(all.predictions, id.vars='Status'), aes(y=value, x=Status, fill=variable)) +
+#  geom_bar(stat='identity', position='dodge') + labs(x='', y='Sample Counts', main='Total Sample Predictions')
+
+adj.pred = as.data.frame(all.predictions %>% rowwise() %>% summarise( 
+  'pred'=round(correctly.predicted/(sum(correctly.predicted, total.samples )),2),
+  'total'=round(total.samples/(sum(correctly.predicted, total.samples )),2),
+  'Status'=Status))
+adj.pred$Patient = rownames(all.predictions)
+
+adj.pred = transform(adj.pred, Patient=reorder(Patient, -total) )
+
+m = melt(adj.pred, id.vars=c('Status', 'Patient'))
+m = arrange(m, Status, variable, -value)
+m$Status = as.factor(m$Status)
+
+grid.arrange(
+  ggplot(subset(m, Status == 'P') , aes(x=Patient, y=value, fill=variable)) + geom_bar(stat = 'identity') +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + coord_flip() + labs(title='Progressors',x='', y='ratio'),
+  ggplot(subset(m, Status == 'NP') , aes(x=Patient, y=value, fill=variable)) + geom_bar(stat = 'identity') +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + coord_flip() + labs(title='Non-Progressors',x='', y='ratio')
+)
+
+pg = subset(adj.pred, Status == 'P')
+np = subset(adj.pred, Status == 'NP')
 
@@
 
-In progressors, `r round(sum(hgd,na.rm=T)/length(hgd[!is.na(hgd)]), 2)*100`% of the HGD samples were correctly predicted.
-
-When predicting each sample individually `r length(pg[pg == 1])` patients were correctly predicted in all samples. The remaining patients were predicted in `r max(pg[pg != 1])*100`% or fewer samples.
-
-`r pander( as.data.frame(sort(pg)), caption='Proportion of progressor patients samples correctly predicted')`
+In progressors, `r round(sum(hgd$correctly.predicted)/sum(hgd$total.hgd.samples), 2)*100`% of the HGD samples were correctly predicted.
+
+When predicting each sample individually `r round(nrow(subset(pg, pred == 0.5))/nrow(pg), 2)*100`% patients were correctly predicted in all samples.
+`r round(nrow(subset(pg, pred < 0.5 & pred > 0))/nrow(pg),2)*100`% of patients were correctly predicted in some samples
+
+`r nrow(subset(pg, pred == 0))` patients had no correctly predicted samples, however, both had only two samples:
+```{r, echo=F, warning=F, message=F}
+x = do.call(rbind, lapply(pg.samp[subset(pg, pred == 0)$Patient], function(df) df[,c('Status','Patient','Path.ID','Endoscopy.Year', 'Pathology')] ))
+row.names(x) = NULL
+pander(x)
+```
 
@@
 
-In non-progressors `r length(which(np == 1))` patients were predicted to be non-progressors in _all_ of their samples.
-`r length(which(np >= 0.75))` were predicted in 75% or more of their samples.
-`r length(np[which(np < 0.75 )])` patients had less than 75% of their samples predicted as NP.
-
-`r pander( as.data.frame(sort(np)), caption='Proportion of NP patients samples correctly predicted') `
+In non-progressors `rnrow(subset(np, pred == 0.5))` patients were predicted to be non-progressors in _all_ of their samples, ranging from `r paste(range(all.predictions[subset(np, pred == 0.5)$Patient, 'total.samples']), collapse='-')`.
+`r nrow(subset(np, pred >= 0.3))` patients were predicted correctly in 70% or more of their samples.
+Only `r nrow(subset(np, pred < 0.3))` patients had fewer than 70% of their samples predicted correctly, and both of these have consistently been predicting _progression_ across multiple endoscopies.
+
+`r pander(all.predictions[subset(np, pred < 0.3)$Patient,], caption='Non-progressors that have consistently predicted progression')`
 
 ```{r timepoints2, echo=T, message=F, warning=F}
+# Only BE, Progressors
+ndbe.only = do.call(rbind, lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], function(pt) {
+  e = subset(pt, Pathology %in% c('BE', '', '?'))
+  cbind.data.frame('predicted'=sum(as.integer(e$Prediction > 0.5)), 'total.be'=nrow(e))
+}))
+ndbe.only = subset(ndbe.only, total.be > 0)
+ndbe.only$pred.ratio = round(ndbe.only$predicted/ndbe.only$total.be, 2)
+
+ndbe.only$Patients = rownames(ndbe.only)
+ndbe.only = transform(ndbe.only, Patients=reorder(Patients, -pred.ratio) )
+
+ggplot(ndbe.only[,c('total.be','pred.ratio', 'Patients')], aes(x=Patients, y=pred.ratio)) + 
+  geom_bar(stat='identity', fill='darkred', alpha=0.6) + geom_label( aes(label=total.be) ) +
+  coord_flip() + labs(title='Correctly predicted NDBE samples', y='Correct prediction ratio')
+
+total = do.call(rbind, lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], function(pt) {
+    cbind.data.frame('predicted'=sum(as.integer(pt$Prediction > 0.5)), 'total'=nrow(pt))
+}))
+
+total$pred.ratio = round(total$predicted/total$total, 2)
+
+total$Patients = rownames(total)
+total = transform(total, Patients=reorder(Patients, -pred.ratio) ) 
+ggplot(total[,c('total','pred.ratio', 'Patients')], aes(x=Patients, y=pred.ratio)) + 
+  geom_bar(stat='identity', fill='purple', alpha=0.6) + geom_label( aes(label=total) ) +
+  coord_flip() + labs(title='Correctly predicted Progressor samples', y='Correct prediction ratio')
+
 # For all pre-HGD/IMC/LGD in progressors
@@
 pred.by.endo<-function(pt, nyears) {
-  e = subset(pt, PID %in% unique(pt$PID)[1:nyears])
+  e = subset(pt, PID %in% unique(pt$PID)[nyears])
   if (  length(which(e$Pathology %in% c('HGD','IMC'))) > 0 ) return(NA)
@@
 
-p.predictions = as.data.frame(matrix(data=0,nrow=6, ncol=3, dimnames=list(c(), c('%.Predicted','n.Endoscopies','n.Patients'))))
+cols = c('%.Predicted','n.Endoscopies','n.Patients', 'FNR','FPR', 'median.pred.samp','median.total.samp', 'max.samples', 'sd.samples')
+p.predictions = as.data.frame(matrix(data=0,nrow=6, ncol=length(cols), dimnames=list(c(), cols)))
 for (i in 1:nrow(p.predictions)) {
@@
   pp = pp[!is.na(pp)]
-  p.predictions[i,] =  rbind(round(sum(pp[!is.na(pp)]/length(pp[!is.na(pp)])), 2), i, length(pp))
-}
-
-ggplot(p.predictions, aes(y=`%.Predicted`,x=n.Endoscopies)) + 
+
+  fpr = sapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Patient], pred.by.endo, i)
+  fpr = sapply(fpr, predicted.progression)
+  fpr = fpr[!is.na(fpr)]
+
+  p.predictions[i,] =  rbind(round(sum(pp[!is.na(pp)]/length(pp[!is.na(pp)])), 2), i, length(pp), 
+                             round(length(which(pp == 0))/length(pp), 2),
+                             round(sum(fpr)/length(fpr), 2), 
+                             median(sapply(p[names(which(pp > 0) )], function(x) length(which(x == 1)) )),
+                             median(sapply(p[names(which(pp > 0))], length)), 
+                             max(sapply(p[names(which(pp > 0))], length)),
+                             sd(sapply(p[names(which(pp > 0))], length))
+                             )
+}
+
+ggplot(p.predictions, aes(y=`%.Predicted`,x=factor(n.Endoscopies))) + 
   geom_bar(stat="identity", fill='dodgerblue3') + 
-  geom_text(aes(label=paste(`%.Predicted`*100, '%', sep='')), nudge_y=-0.02) + 
-  geom_text(aes(label=paste(n.Patients, 'patients')), nudge_y=0.005) + ggtitle("Progression predicted per endoscopy")
-```
-
-`r pander(p.predictions, justify='left', caption=paste("For any given endoscopy if a single sample (same pathology ID, different blocks) is predicted as 'progressor' that endoscopy is considered to be indicative of progression.
-At each endoscopy the percentage is calculated excluding any patients that are diagnosed HGD at that point.
-There are ",length(subset(sum.patient.data, Status == 'P')$Patient), " progressor patients.
-Based on that, ", p.predictions[1,1]*100, "% are correctly predicted at the first endoscopy.", sep=''))`
-
-What about the non-progressors that had progressor predictions?
-```{r, warning=F, message=F}
-
-np.predictions = as.data.frame(matrix(data=0,nrow=6, ncol=3, dimnames=list(c(), c('%.Predicted','n.Endoscopies','n.Patients'))))
-for (i in 1:nrow(np.predictions)) {
-  p = sapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Patient], pred.by.endo, i)
-  pp = sapply(p, predicted.progression)
-  pp = pp[!is.na(pp)]
-  np.predictions[i,] =  rbind(round(sum(pp[!is.na(pp)]/length(pp[!is.na(pp)])), 2), i, length(pp))
-}
-
-```
-
+  geom_text(aes(label=paste(`%.Predicted`*100, '%', sep='')), nudge_y=-0.02, size=8) + 
+  geom_text(aes(label=paste(n.Patients, 'patients')), nudge_y=0.01, size=8) + 
+  geom_bar(stat='identity', fill='darkred', alpha=0.3, aes(y=FPR)) +
+  geom_text(aes(label=FPR, y=FPR)) +
+  labs(title="Progression predicted per endoscopy", x='Number of Endoscopies', y='% Patients') +
+  theme(text = element_text(size = 14))
+
+```
+
+## What clinicicians really want
+
+Backwards from the final sample.
+
+```{r, warning=F, message=F, echo=F, fig.width=16, fig.height=8}
+
+back = lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], function(pt) {
+  pt = ddply(pt, .(PID, Endoscopy.Year), summarise, 'pred'=sum(as.integer(Prediction > 0.5)), 'samples'=length(PID), 
+        'final.endo'= length(which(grepl('HGD|IMC', Pathology))) > 0 )
+  arrange(pt, -Endoscopy.Year, PID)
+})
+
+endos = do.call(rbind, lapply(1:max(sapply(back,nrow)), function(i) {
+  df = do.call(rbind, lapply(back, function(x) {
+    if (i > nrow(x)) return(NA)
+    cbind(x[i, c('Endoscopy.Year','pred','samples','final.endo')], i)
+  }))
+  df$Patient = rownames(df)
+  return(df)
+}))
+
+endos = endos[complete.cases(endos),]
+endos$i = endos$i-1
+
+endos$pred.ratio = round(endos$pred/endos$samples, 2)
+endos$i = as.factor(endos$i)
+
+endos = arrange(endos, Patient)
+endos = endos %>% group_by(Patient) %>% mutate('n'=length(Patient))
+endos = transform(endos, Patient=reorder(Patient, -n) ) 
+endos$brks <- cut(endos$pred.ratio, include.lowest=T,
+                   breaks=c(0, 0.1, 0.25, 0.5, 0.75, 0.9, 1), 
+                   labels=c('0', '0.1-.25', '.25-.5', '.5-.75','.75-0.9', '1'))
+
+## TODO This doesn't seem like the most useful way to display this idea
+m = melt(endos[,c('Patient','pred.ratio', 'i','n', 'brks')], id.vars = c('Patient', 'i', 'n', 'brks'))
+
+ggplot(m, aes(Patient, i, fill=brks)) + geom_bar(color='grey', stat='identity', position='dodge') + 
+  scale_fill_brewer(type='div', palette = "Set2", "Predicted Ratio") + labs(x='', y='Endoscopy from final') + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+```
+
+```{r, warning=F, message=F, fig.height=4, fig.width=4, eval=F}
 # How do I evaluate the predictions?
 
-I'm a bit at a loss here.
-I understand why penalized regression wouldn't have a CI, but then I'm not sure how best to evaluate the predictions.
-Do I just take the predictions at face value and leave it at that?
-
-```{r, warning=F, message=F, fig.height=4, fig.width=4}
+#Basically I need to to a FNR and FPR calculation that is extremely conservative and use that somehow
+
 all = do.call(rbind,lapply(pg.samp, function(df) {
@@
 
-```{r, echo=T, warning=F, message=F}
-
+```{r, echo=T, warning=F, message=F, fig.height=10, fig.width=10}
+
+file = paste(cache.dir, 'rand.Rdata', sep='/')
+
+if (file.exists(file)) {
+  load(file, verbose=T)
+} else {
   rand.patients = sum.patient.data
@@
   }))
+  names(r.labels) = sub('.*\\.', '',  names(r.labels))
 
@@
   }
+  save(r.plots, r.coefs, file=file) 
+}
 
@@
 
-## check how often the feature(s) is selected at that lamda in each split. "stability selection"
-r.coef.stable = lapply( r.coefs, function(cf) {
-  sort(rowSums(cf[,-1]), decreasing=T)
-})
-
-cfs = as.data.frame(matrix(data=0,nrow=length(unique(unlist(sapply(r.coef.stable, names)))), ncol=length(names(r.coef.stable)), 
-                           dimnames=list(unique(unlist(sapply(r.coef.stable, names))), names(r.coef.stable) )))
-
-for (i in names(r.coef.stable)) 
-  cfs[intersect(rownames(cfs), names(r.coef.stable[[i]])), i] = 1
-
-r.coef.stable = lapply(r.coef.stable, function(cf) cf/(folds*splits))
-
-r.stability = do.call(rbind, lapply(r.coef.stable, function(cf) {
-  cbind('75%'=length(which(cf >= 0.75)),  '50%'=length(which(cf >= 0.5)),  '<50%'=length(which(cf < 0.5)))
-}))
-rownames(r.stability) = names(r.coef.stable)
-
-```
-
-`r pander(table(rowSums(cfs)))`
+```

### 2017-05-24 Patient Report.Rmd
…

@@
 dataset = 'Training'
-#dataset = 'Test'
 patient.info = read.patient.info(patient.file, set=dataset)
@@
 ## Thresholds
+gamma2 = 250
 min.probes=67 # equiv to 1MB
-gamma2 = 250
 sd.threshold = 0.08  # cov by numbers of probes
+gain.threshold = 1.1; loss.threshold = 0.9
 
-gain.threshold = 1.1
-loss.threshold = 0.9
-
-cachefile = paste(dataset, '_patients.Rdata', sep='')
+cachefile = paste(data, '/Analysis/', dataset, '_patients.Rdata', sep='')
 if ( file.exists(cachefile) ) {
@@
   }
-save(patient.data, file=paste(dataset, '_patients.Rdata', sep=''))
+save(patient.data, file=cachefile)
 }
@@
 m = transform(m, Gene.Symbols = reorder(Gene.Symbols, value))
-
-#m$Gene.Symbols = factor(m$Gene.Symbols, levels=unique(m$Gene.Symbols[c(nrow(m):1)]) ) # order
 m$CN = factor(m$CN)
 
-m$brks <- cut(m$value, include.lowest=T, right=F,
+m$brks <- cut(m$value, include.lowest=T, right=F,ordered_result=T,
                    breaks=c(0, 1, 3, 5, 8, 10),
                    labels=c('0','1-3', '4-5', '6-8','9-10'))
-m = transform(m, Gene.Symbols = reorder(Gene.Symbols, value))
+
+## TODO: Tile the variables so that you can see the gains/losses per progressor/non
+subset(m, Gene.Symbols %in% c('CDKN2A', 'TBL1XR1'))
+m = transform(m, Gene.Symbols = reorder(Gene.Symbols, brks))
 
@@
 
-ggplot(m, aes(y=Gene.Symbols, x=variable)) + 
-  geom_tile(aes(fill=brks), color='white') + scale_fill_manual( values = myP, "Patient Counts") +
-  geom_point(data=subset(m, value > 0), aes(y=Gene.Symbols, x=variable, shape=CN, color=CN), size=3 ) +
-  scale_color_manual("CN", values=c("blue", "yellow")) + scale_shape_manual(values=c(17,20)) +
-  labs(title="All available COSMIC genes mapped", x='', y='')
+ggplot(m,  aes(y=Gene.Symbols, x=CN, fill=brks)) + geom_tile() + facet_grid(.~variable) + 
+    scale_fill_manual(values = myP, 'Patient Counts') + labs(title="All available COSMIC genes", x='', y='')
 
@@
 ```{r echo=F, warning=F, message=F, fig.height=12, fig.width=8}
+gc = as.data.frame(gene.p %>% group_by(Gene.Symbols) %>% summarise(total = sum(Progressors) + sum(Non.Progressors)))
+gc = transform(gc, Gene.Symbols = reorder(Gene.Symbols, total))
+gc = arrange(gc, -total)
 
-m = melt(gene.p[1:50,], id.vars=c('Gene.Symbols', 'CN'))
+m = melt(subset(gene.p, Gene.Symbols %in% gc[1:50, 'Gene.Symbols']), id.vars=c('Gene.Symbols', 'CN'))
 m = transform(m, Gene.Symbols = reorder(Gene.Symbols, value))
@@
 
-ggplot(m, aes(y=Gene.Symbols, x=variable)) + 
-  geom_tile(aes(fill=brks), color='white') + scale_fill_manual( values = myP, "Patient Counts") +
-  geom_point(data=subset(m, value > 0), aes(y=Gene.Symbols, x=variable, shape=CN, color=CN), size=3 ) +
-  scale_color_manual("CN", values=c("blue", "yellow")) + scale_shape_manual(values=c(17,20)) +
+ggplot(m,  aes(y=Gene.Symbols, x=CN, fill=brks)) + geom_tile(color='white') + facet_grid(.~variable) + 
+    scale_fill_manual(values=myP, 'Patient Counts') + 
     labs(title="Top 50 genes by number of patients", x='', y='')
+
+#ggplot(m, aes(y=Gene.Symbols, x=variable)) + 
+#  geom_tile(aes(fill=brks), color='white') + scale_fill_manual( values = myP, "Patient Counts") +
+#  geom_point(data=subset(m, value > 0), aes(y=Gene.Symbols, x=variable, shape=CN, color=CN), size=3 ) +
+#  scale_color_manual("CN", values=c("blue", "yellow")) + scale_shape_manual(values=c(17,20)) +
+#  labs(title="Top 50 genes by number of patients", x='', y='')
 ```

### 2017-05-24 RegressionPvNP.Rmd
…

@@
 ```{r tile, echo=F, message=F, warning=F, include=F}
-#tile.w=1e07
-tile.w=5e06
-tile.files = grep('tile_.*5e\\+06', analysis.files, value=T)
+tile.w=1e07
+#tile.w=5e06
+tile.files = grep('tile_.*1e\\+07', analysis.files, value=T)
+#tile.files = grep('tile_.*5e\\+06', analysis.files, value=T)
 load(tile.files[1], verbose=T)
@@
 
-cache.dir = paste(data, 'Analysis/5e06', sep='/')
+#cache.dir = paste(data, 'Analysis/5e06', sep='/')
+cache.dir = paste(data, 'Analysis/1e07', sep='/')
 if (!dir.exists(cache.dir)) dir.create(cache.dir)
@@
 ```{r xvalpt, message=F, warning=F, echo=T, fig.height=16, fig.width=16}
+nl = 3000
+
 coefs = list(); plots = list(); performance.at.1se = list()
@@
   for (a in alpha.values) {
-    fit0 <- glmnet(dysplasia.df, labels, alpha=a, nlambda=1000, family='binomial') # all patients
+    fit0 <- glmnet(dysplasia.df, labels, alpha=a, nlambda=nl, family='binomial') # all patients
     #autoplot(fit0, xvar='lambda', main=paste('fit0, all samples, alpha=',a,sep='')) + theme(legend.position='none')
@@
 ```{r xvalpt2, message=F, warning=F, echo=T}
-
 performance.at.1se = do.call(rbind, performance.at.1se)
@@
 
-ggplot(performance.at.1se, aes(alpha, mean)) + geom_point() + geom_text( aes(label=round(mean, 3)), hjust=-0.5) +
-  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + ylim(0.5,0.8) +
+ggplot(performance.at.1se, aes(alpha, mean)) + geom_text( aes(label=round(mean, 3)), hjust=-0.5) +
+  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme), color='grey') + geom_point() + ylim(0.5,0.8) +
   geom_text(aes(label=n.Coef), nudge_y=0.03) +
@@
 
+### Features in samples
+
+```{r echo=F, warning=F, message=F}
+
+length(which( coef.stable[['1']] >= 0.5) )
+
+regions = rownames(coefs[['1']])
+
+dysplasia.df[ labels == 1 ,regions]
+
+dysplasia.df[ labels == 0 ,regions]
+
+```
+
 ### Feature Information
 
-```{r, warning=F, message=F}
+```{r, warning=F, message=F, eval=F}
 #topF = do.call(rbind, strsplit( names(which((coef.stable[['1']]) >= 0.75)), ':|-'))
@@
     # all patients
-    fitNoHGD <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=1000)
+    fitNoHGD <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=nl)
 
@@
   for (a in alpha.values) {
-    fitNoLGD <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=1000) # all patients
+    fitNoLGD <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=nl) # all patients
 
@@
 ```{r noLGD2, echo=T, message=F, warning=F}
+if (class(performance.at.1se) == 'list')
   performance.at.1se = do.call(rbind, performance.at.1se)
+
 performance.at.1se$alpha = rownames(performance.at.1se)
@@
   for (a in alpha.values) {
-    fitF <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, nlambda=500, family='binomial') # all patients
+    fitF <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, nlambda=nl, family='binomial') # all patients
     cv.back.few = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitF$lambda, pts=subset(pts, Samplename %in% samples), splits=5, nfolds=20, a=a, fit=fitF, minR=0.1)
@@
   for (a in alpha.values) {
-    fitM <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, nlambda=500, family='binomial') # all patients
+    fitM <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, nlambda=nl, family='binomial') # all patients
 
@@
 
-```{r leaveoneout, echo=T, message=F, warning=T}
+```{r leaveoneout, echo=T, message=F, warning=F}
 
@@
   info = pt$info
-  print(unique(info$Patient))
   info$SampleSD = NA
@@
     a = 1
-    fitLOO <- glmnet(training, labels[train.rows], alpha=a, family='binomial', nlambda=1000) # all patients
+    fitLOO <- glmnet(training, labels[train.rows], alpha=a, family='binomial', nlambda=nl) # all patients
     cv = crossvalidate.by.patient(x=training, y=labels[train.rows], lambda=fitLOO$lambda, 
@@
 
-adj.pred = as.data.frame(all.predictions %>% rowwise() %>% summarise( 
-  'pred'=round(correctly.predicted/(sum(correctly.predicted, total.samples )),2),
-  'total'=round(total.samples/(sum(correctly.predicted, total.samples )),2),
-  'Status'=Status))
-adj.pred$Patient = rownames(all.predictions)
-
-adj.pred = transform(adj.pred, Patient=reorder(Patient, -total) )
-
-m = melt(adj.pred, id.vars=c('Status', 'Patient'))
-m = arrange(m, Status, variable, -value)
-m$Status = as.factor(m$Status)
+all.predictions$pred.ratio = with(all.predictions, correctly.predicted/total.samples)
+all.predictions$Patient = rownames(all.predictions)
+all.predictions = transform(all.predictions, Patient=reorder(Patient, pred.ratio) )
+
+m = melt(all.predictions, id.vars=c('Patient','Status'), measure.vars=c('pred.ratio'))
 
 grid.arrange(
-  ggplot(subset(m, Status == 'P') , aes(x=Patient, y=value, fill=variable)) + geom_bar(stat = 'identity') +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + coord_flip() + labs(title='Progressors',x='', y='ratio'),
-  ggplot(subset(m, Status == 'NP') , aes(x=Patient, y=value, fill=variable)) + geom_bar(stat = 'identity') +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + coord_flip() + labs(title='Non-Progressors',x='', y='ratio')
-)
-
-pg = subset(adj.pred, Status == 'P')
-np = subset(adj.pred, Status == 'NP')
+  ggplot() + geom_bar(data=subset(m, Status == 'P'), aes(Patient, value), fill='darkblue', color='grey', stat='identity') +  
+    theme(legend.position='none', axis.text.x = element_text(angle = 45, hjust = 1)) + coord_flip() +
+    labs(title='Progressors', x='', y='Ratio predicted samples'),
+  ggplot() + geom_bar(data=subset(m, Status == 'NP'), aes(Patient, value), fill='darkgreen', color='grey', stat='identity') + 
+    theme(legend.position='none', axis.text.x = element_text(angle = 45, hjust = 1)) + coord_flip() +
+    labs(title='Non Progressors', x='', y='Ratio predicted samples'),
+  ncol=2)
+
+pg = subset(all.predictions, Status == 'P')
+np = subset(all.predictions, Status == 'NP')
 
@@
 
-When predicting each sample individually `r round(nrow(subset(pg, pred == 0.5))/nrow(pg), 2)*100`% patients were correctly predicted in all samples.
-`r round(nrow(subset(pg, pred < 0.5 & pred > 0))/nrow(pg),2)*100`% of patients were correctly predicted in some samples
-
-`r nrow(subset(pg, pred == 0))` patients had no correctly predicted samples, however, both had only two samples:
-```{r, echo=F, warning=F, message=F}
-x = do.call(rbind, lapply(pg.samp[subset(pg, pred == 0)$Patient], function(df) df[,c('Status','Patient','Path.ID','Endoscopy.Year', 'Pathology')] ))
-row.names(x) = NULL
-pander(x)
-```
+When predicting each sample individually `r round(nrow(subset(pg, pred.ratio == 1))/nrow(pg))*100`% patients were correctly predicted in all samples.
+`r round(nrow(subset(pg, pred.ratio < 1 & pred.ratio > 0))/nrow(pg),2)*100`% of patients were correctly predicted in some samples
+
+`r nrow(subset(pg, pred.ratio == 0))` patients had no correctly predicted samples.
 
@@
 
-In non-progressors `rnrow(subset(np, pred == 0.5))` patients were predicted to be non-progressors in _all_ of their samples, ranging from `r paste(range(all.predictions[subset(np, pred == 0.5)$Patient, 'total.samples']), collapse='-')`.
-`r nrow(subset(np, pred >= 0.3))` patients were predicted correctly in 70% or more of their samples.
-Only `r nrow(subset(np, pred < 0.3))` patients had fewer than 70% of their samples predicted correctly, and both of these have consistently been predicting _progression_ across multiple endoscopies.
-
-`r pander(all.predictions[subset(np, pred < 0.3)$Patient,], caption='Non-progressors that have consistently predicted progression')`
+In non-progressors `r round(nrow(subset(np, pred.ratio == 1))/nrow(np), 2)*100`% of patients were predicted to be non-progressors in _all_ of their samples.
+`r nrow(subset(np, pred.ratio >= 0.7))` patients were predicted correctly in 70% or more of their samples.
+Only `r nrow(subset(np, pred.ratio < 0.7 & pred.ratio > 0))` patients had fewer than 70% of their samples predicted correctly, and both of these have consistently been predicting _progression_ across multiple endoscopies.
 
@@
 
-## What clinicicians really want
-
-Backwards from the final sample.
+## Backwards from HGD
+
+Clinicians always ask to see the timepoints peeled backwards to see how early prediction is possible.
 
@@
 back = lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], function(pt) {
-  pt = ddply(pt, .(PID, Endoscopy.Year), summarise, 'pred'=sum(as.integer(Prediction > 0.5)), 'samples'=length(PID), 
-        'final.endo'= length(which(grepl('HGD|IMC', Pathology))) > 0 )
+  pt = pt %>% group_by(PID, Endoscopy.Year) %>% summarise(
+                  'pred'=sum(as.integer(Prediction > 0.5)), 'samples'=length(PID), 
+                  'final.endo'= length(which(grepl('HGD|IMC', Pathology))) > 0,
+                  'months.before.final'= (max(pt$Endoscopy.Year) - unique(Endoscopy.Year))*12 )
   arrange(pt, -Endoscopy.Year, PID)
@@
     if (i > nrow(x)) return(NA)
-    cbind(x[i, c('Endoscopy.Year','pred','samples','final.endo')], i)
+    cbind(x[i, c('Endoscopy.Year','pred','samples','final.endo', 'months.before.final')], i)
   }))
@@
 endos$pred.ratio = round(endos$pred/endos$samples, 2)
-endos$i = as.factor(endos$i)
-
-endos = arrange(endos, Patient)
-endos = endos %>% group_by(Patient) %>% mutate('n'=length(Patient))
-endos = transform(endos, Patient=reorder(Patient, -n) ) 
-endos$brks <- cut(endos$pred.ratio, include.lowest=T,
-                   breaks=c(0, 0.1, 0.25, 0.5, 0.75, 0.9, 1), 
-                   labels=c('0', '0.1-.25', '.25-.5', '.5-.75','.75-0.9', '1'))
-
-## TODO This doesn't seem like the most useful way to display this idea
-m = melt(endos[,c('Patient','pred.ratio', 'i','n', 'brks')], id.vars = c('Patient', 'i', 'n', 'brks'))
-
-ggplot(m, aes(Patient, i, fill=brks)) + geom_bar(color='grey', stat='identity', position='dodge') + 
-  scale_fill_brewer(type='div', palette = "Set2", "Predicted Ratio") + labs(x='', y='Endoscopy from final') + 
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-```
+#endos$i = as.factor(endos$i)
+
+#endos$brks <- cut(endos$pred.ratio, include.lowest=T,
+#                   breaks=c(0, 0.1, 0.25, 0.5, 0.75, 0.9, 1), 
+#                   labels=c('0', '0.1-.25', '.25-.5', '.5-.75','.75-0.9', '1'))
+
+endos = transform(endos, Patient=reorder(Patient, -months.before.final) ) 
+endos$months.before.final = endos$months.before.final*-1
+
+plot.endo<-function(df, minM=NULL, maxM=NULL) {
+  if (is.null(minM)) minM = min(df$months.before.final)
+  if (is.null(maxM)) maxM = max(df$months.before.final)
+
+  df$i = as.factor(df$i)
+  ggplot(df, aes(months.before.final, pred.ratio, group=1,color=Patient)) + facet_grid(Patient ~., scales='fixed') + scale_x_continuous(breaks=seq(minM, maxM, by = 12)) +
+    geom_line() + geom_point() + labs(x='Months before diagnosis', y='Sample prediction ratio') + theme(legend.position='none') 
+}
+
+me = endos %>% group_by(Patient) %>% summarise('max.endo' = max(i))
+
+grid.arrange(
+  plot.endo( subset(endos, Patient %in% subset(me, max.endo >=5)$Patient), min(endos$months.before.final), max(endos$months.before.final) ),
+  plot.endo( subset(endos, Patient %in% subset(me, max.endo < 5 & max.endo > 2)$Patient),min(endos$months.before.final), max(endos$months.before.final)  ),
+  ncol=2)
+
+```
+
+```{r, echo=F, warning=F, message=F, fig.height=8, fig.width=8}
+plot.endo( subset(endos, Patient %in% subset(me, max.endo <= 2 & max.endo > 0)$Patient), min(endos$months.before.final), max(endos$months.before.final)  )
+```
+
+```{r, echo=F, warning=F, message=F, fig.height=4, fig.width=4}
+plot.endo( subset(endos, Patient %in% subset(me, max.endo == 0)$Patient),min(endos$months.before.final), max(endos$months.before.final) )
+```
+
+### How many samples?  TBD
+
+Clinicians asked how many samples they needed to take.
 
@@
   for (a in c(0,0.5,1)) {
-    fitR <- glmnet(r.dysplasia.df, r.labels, alpha=a, nlambda=1000, family='binomial') # all patients
+    fitR <- glmnet(r.dysplasia.df, r.labels, alpha=a, nlambda=nl, family='binomial') # all patients
 

### 2017-06-02 RegressionPvNP.Rmd
…

@@
 ```{r setup, include=FALSE}
+library(biomaRt)
 library(ggplot2)
 library(ggfortify)
+library(GGally)
 library(plyr)
@@
 #tile.w=5e06
-tile.files = grep('tile_.*1e\\+07', analysis.files, value=T)
-#tile.files = grep('tile_.*5e\\+06', analysis.files, value=T)
+tile.files = grep(as.character(tile.w), analysis.files, value=T, fixed=T)
 load(tile.files[1], verbose=T)
@@
 
-#cache.dir = paste(data, 'Analysis/5e06', sep='/')
-cache.dir = paste(data, 'Analysis/1e07', sep='/')
+cache.dir = paste(data, 'Analysis',sub('\\+', '', tile.w), sep='/')
 if (!dir.exists(cache.dir)) dir.create(cache.dir)
@@
   if (select == 'classification') {
-    df$lambda.min = df$mean == max(df$mean)
+    # Minimize the classification mean error & select min lambda
+    df[which.max(subset(df, sme < median(sme) & sme > min(sme))$mean), 'lambda.min'] = T
+
+    #df$lambda.min = df$mean == max(df$mean)
     se1 = df[df$lambda.min == T, 'mean'] - df[df$lambda.min == T, 'sme']
@@
     annotate("text", x=subset(df, lambda.min == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
-             label=round(df[subset(df, lambda.min == T)$lambda.at,'mean'],3))
+             label=paste(round(df[subset(df, lambda.min == T)$lambda.at,c('mean', 'sme')],3), collapse='\n+/-'))
 
@@
       annotate("text", x=subset(df, lambda.1se == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
-               label=round(df[subset(df, lambda.1se == T)$lambda.at,'mean'],3))
+               label=paste(round(df[subset(df, lambda.1se == T)$lambda.at,c('mean', 'sme')],3), collapse='\n+/-'))
 
@@
       annotate("text", x=subset(df, lambda.1se == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
-               label=round(df[subset(df, lambda.1se == T)$lambda.at,'mean'],3)) 
+               label=paste(round(df[subset(df, lambda.1se == T)$lambda.at,c('mean', 'sme')],3), collapse='\n+/-')) 
   }
@@
 ```{r xvalpt, message=F, warning=F, echo=T, fig.height=16, fig.width=16}
-nl = 3000
+nl = 1000
 
@@
 if (file.exists(file)) {
+  message(paste("loading", file))
   load(file, verbose=T)
@@
 
-    cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=fit0$lambda, pts=pts, a=a, nfolds=folds, splits=splits, fit=fit0)
+    l = fit0$lambda
+    if (a == 0) {
+      l = sort(c(fit0$lambda, seq(exp(-5), exp(-10), -1e-6),
+            seq(exp(-10), exp(-15), -1e-8),
+            seq(exp(-15), exp(-20), -1e-9)), decreasing=T)
+    }
+    cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=l, pts=pts, a=a, nfolds=folds, splits=splits, fit=fit0)
 
@@
 
-dysplasia.df[ labels == 1 ,regions]
-
-dysplasia.df[ labels == 0 ,regions]
+pg = dysplasia.df[ labels == 1 ,regions]
+
+nonpg = dysplasia.df[ labels == 0 ,regions]
+
+colnames(pg)
+
+meantests = sapply(regions, function(r)   t.test(pg[,r], nonpg[,r])$p.value )
+
+#which(meantests < 0.05)
+
+summary(pg[,'1:119640299-129610322'])
+summary(nonpg[,'1:119640299-129610322'])
 
@@
 
-```{r, warning=F, message=F, eval=F}
-#topF = do.call(rbind, strsplit( names(which((coef.stable[['1']]) >= 0.75)), ':|-'))
-#colnames(topF) = c('chr','start','end')
-#topF = apply(topF, 2, as.numeric)
-
-topF = names(which((coef.stable[['1']]) >= 0.75))
-
-library(biomaRt)
+```{r, warning=F, message=F}
+#topF = names(which((coef.stable[['1']]) >= 0.75))
+topF = names(coef.stable[['1']])
+
+features.in.prog = dysplasia.df[names(which(labels == 1)),topF]
+features.nonprog = dysplasia.df[names(which(labels == 0)),topF]
+
+feat = which(colnames(dysplasia.df) %in% topF)
+
+# Gain or loss not correlated in prog or non
+
+gl = cbind.data.frame(
+  'P.gain' = apply(features.in.prog, 2, function(x) sum( as.integer(x >= 1.1) ) ), 
+  'P.loss' = apply(features.in.prog, 2, function(x) sum( as.integer(x <= 0.9) ) ),
+  'NP.gain' = apply(features.nonprog, 2, function(x) sum( as.integer(x >= 1.1) ) ),
+  'NP.loss' = apply(features.nonprog, 2, function(x) sum( as.integer(x <= 0.9) ) ) )
+
+allgl = cbind.data.frame(
+  'ALL.P.gain' = apply(dysplasia.df[names(which(labels == 1)),], 2, function(x) sum( as.integer(x >= 1.1) ) ), 
+  'ALL.P.loss' = apply(dysplasia.df[names(which(labels == 1)),], 2, function(x) sum( as.integer(x <= 0.9) ) ),
+  'ALL.NP.gain' = apply(dysplasia.df[names(which(labels == 0)),], 2, function(x) sum( as.integer(x >= 1.1) ) ), 
+  'ALL.NP.loss' = apply(dysplasia.df[names(which(labels == 0)),], 2, function(x) sum( as.integer(x <= 0.9) ) ) )
+
+grid.arrange(
+  ggcorr(gl, label=T) + ggtitle('Selected Features'),
+  ggcorr(allgl, label=T) + ggtitle('All regions'), 
+top='Gains vs Losses')
+
+# Gains/losses enriched in the top features?
+fisher.test(rbind(colSums(gl[,c('P.gain','P.loss')]),colSums(allgl[-feat,c('ALL.P.gain','ALL.P.loss')])  ))
+
+fisher.test(rbind(colSums(gl[,c('NP.gain','NP.loss')]),colSums(allgl[-feat,c('ALL.NP.gain','ALL.NP.loss')])  ))
+
+```
+
+#### bp Length
+
+```{r, echo=F, warning=F, message=F}
+chromInfo = read.table('hg19_genes.txt', sep='\t', header=T)
+chromInfo$chr = factor(chromInfo$chr, levels=c(1:22, 'X','Y'))
+
+topFSegments = do.call(rbind,  strsplit( topF, ':|-' ))
+colnames(topFSegments) = c('chr','start','end')
+topFSegments = as.data.frame(apply(topFSegments, 2, as.numeric))
+topFSegments$chr = factor(topFSegments$chr, levels=chromInfo$chr)
+topFSegments = arrange(topFSegments, chr, start)
+
+topFRegions = arrange(merge(chromInfo[,c('chr','length', 'protein_coding')], topFSegments %>% group_by(chr) %>% summarise('totalbp'=sum(end-start), 'n.segments'=length(chr)), all.x=T), -protein_coding)
+
+# Number of segments is strongly correlated with the length of the chromosome
+with(topFRegions, cor.test(length, n.segments))
+
+topFRegions$adj.info.content = topFRegions$protein_coding/(topFRegions$length^0.6)
+with(topFRegions, cor.test(length, adj.info.content))
+
+topFRegions = transform(topFRegions, ai.cut=cut(adj.info.content, 5, labels=c(5:1), include.lowest=T))           # bin data
+ggplot(topFRegions, aes(reorder(chr, -adj.info.content), totalbp/length, fill=ai.cut)) + ylim(0, 0.5) +
+  labs(y='Feature bp/total chr length', title='Ratio of covered base pairs', x='chr') +
+  geom_bar(stat='identity') +  geom_text(aes(label=round(totalbp/length, 2)), vjust=-0.5) +
+  scale_fill_discrete(name='Length adjusted information content') + theme(legend.position = 'bottom')
+
+ct = with(topFRegions, cor.test(adj.info.content, totalbp/length))
+
+bpCovered = with(topFRegions, sum(totalbp, na.rm=T)/sum(as.numeric(length)))
+```
+
+Strong correlation between the length adjusted information content (protein coding genes) and the covered bases in a chromosome r=`r round(ct$estimate, 2)` (p-value `r round(ct$p.value, 4)`).
+No features from chromosome 19 (the most gene dense chromosome) were selected.
+
+`r round(bpCovered, 2)*100`% of the total bases are within the top selected features.
+
+#### Genes?
+
+Using Biomart get all genes that are fully within the the segment boundaries.
+
+```{r, warning=F, message=F}
 mart = biomaRt::useMart("ENSEMBL_MART_ENSEMBL", host="grch37.ensembl.org")
@@
 
-attr = c('ensembl_gene_id','hgnc_symbol','start_position','end_position', 'chromosome_name', 'go_id', 'entrezgene')
-bm = getBM(mart=biomart, attributes=attr, filters=list('chromosomal_region' = topF))
+attr = c('ensembl_gene_id','hgnc_symbol','start_position','end_position', 'chromosome_name', 'ensembl_exon_id','percentage_gene_gc_content')
+
+bm = getBM(mart=biomart, attributes=attr, 
+           filters=list('chromosomal_region'=sub('-', ':',topF), "biotype"=c("IG_C_gene","IG_D_gene","IG_J_gene","IG_V_gene","protein_coding","TR_C_gene","TR_D_gene","TR_J_gene","TR_V_gene")))
+#head(bm)
 bm = bm[-which(bm$hgnc_symbol == ""),]
-#head(bm)
-
-bm2 = ddply(bm, .(entrezgene, ensembl_gene_id, hgnc_symbol), summarise, 'chrom'=unique(chromosome_name), 'start'=min(start_position), 'end'=max(end_position), 'GO.IDs'=paste(go_id, collapse=','))
-
+bm = bm %>% group_by(hgnc_symbol, chromosome_name, start_position, end_position) %>% summarise('exon_count' = length(ensembl_exon_id)) 
+exons = bm %>% group_by(chromosome_name) %>% summarise('exons'=sum(exon_count))
+
+for (i in 1:nrow(topFSegments)) {
+  s = topFSegments[i,]
+  sbm = with(bm, which(chromosome_name == s[['chr']] & start_position >= s[['start']] & end_position <= s[['end']]))
+
+  topFSegments[i,'genes'] = length(sbm)
+  topFSegments[i,'exons'] = sum(bm[sbm,'exon_count'])
+
+  bm[sbm, 'segment.index'] = i
+}
+
+# Not related to the length of the segment
+with(topFSegments, cor.test(genes, (end-start)))
+
+head(bm)
+
+topFRegions = merge(topFRegions, (bm %>% group_by(chromosome_name) %>% summarise('gene.cnt'=length(unique(hgnc_symbol)), 'exon.cnt'=sum(exon_count))), by.x='chr', by.y='chromosome_name', all.x=T)
+
+with(topFRegions, cor.test(protein_coding, exon.cnt))
+with(topFRegions, cor.test(gene.cnt, exon.cnt)) # expected
+
+# Exons correlate with the adjusted information content measure
+with(topFRegions, cor.test(exon.cnt, adj.info.content))
+
+ggplot(topFRegions, aes(reorder(chr, -adj.info.content), gene.cnt/protein_coding, fill=ai.cut)) + geom_bar(stat='identity') +
+  labs(y='Gene Ratio', x='chr', title='Feature genes vs all protein coding genes') +
+  geom_text(aes(label=round(gene.cnt/protein_coding, 2)), vjust=-0.5) +
+  scale_fill_discrete(name='Length adjusted information content') + theme(legend.position = 'bottom')
+
+# Merge COSMIC genes
 ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
-
-bm2 = subset(bm2, hgnc_symbol %in% ccgenes$Gene.Symbol)
-
-features.in.prog = dysplasia.df[names(which(labels == 1)),topF]
-
-#features.in.prog <= 0.9
-#features.in.prog >= 1.1
-
-```
+ccgenes = ccgenes[-grep('X,Y', ccgenes$Genome.Location),]
+
+# This is going to strip out some genes that don't appear in ensembl
+attr = c('ensembl_gene_id','hgnc_symbol','start_position','end_position', 'chromosome_name')
+ccbm = getBM(mart=biomart, attributes=attr, filters=list('external_gene_name'=ccgenes$Gene.Symbol, 'chromosome_name'=c(1:22,'X','Y')))
+ccbm = ccbm[-which(ccbm$hgnc_symbol == ""),]
+ccbm$chromosome_name = factor(ccbm$chromosome_name, levels=chromInfo$chr)
+
+topFRegions = merge(topFRegions, (ccbm %>% group_by(chromosome_name) %>% summarise('COSMIC.all'=length(unique(hgnc_symbol)))), by.x='chr',by.y='chromosome_name', all.x=T)
+
+containedCC = subset(bm, hgnc_symbol %in% ccbm$hgnc_symbol)
+
+topFRegions = merge(topFRegions, (containedCC %>% group_by(chromosome_name) %>% summarise('COSMIC.cont'=length(unique(hgnc_symbol)))), by.x='chr',by.y='chromosome_name', all.x=T)
+
+ggplot(topFRegions, aes(reorder(chr, -adj.info.content), COSMIC.cont/COSMIC.all, fill=ai.cut)) + geom_bar(stat='identity') +
+  labs(y='COSMIC Gene Ratio', x='chr', title='COSMIC genes: Feature vs all')+
+  geom_text(aes(label=round(COSMIC.cont/COSMIC.all, 2)), vjust=-0.5) +
+  scale_fill_discrete(name='Length adjusted information content') + theme(legend.position = 'bottom')
+```
+
+However, the `r round(bpCovered, 2)*100`% of bases appear to include `r round(length(unique(bm$hgnc_symbol))/sum(chromInfo$protein_coding), 2)*100`% (`r length(unique(bm$hgnc_symbol))`) of known protein-coding genes.
+Similarly, `r round(length(unique(containedCC$hgnc_symbol))/length(ccgenes$Gene.Symbol),2)*100`% of the `r length(ccgenes$Gene.Symbol)` COSMIC genes are contained within these regions.
+
+This appears to be a fairly large enrichment.
 
@@
 if (file.exists(file)) {
+  message(paste("loading file", file))
   load(file, verbose=T)
@@
     fitNoHGD <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=nl) 
-
     cv.nohgd = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitNoHGD$lambda, pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fitNoHGD)
@@
 
-# Leave one out
+## Randomize the labels
+
+By patient.  This should break the models...and as would be expected, the models are really poor with perfomance at 50% or less (coin toss).
+
+```{r, echo=T, warning=F, message=F, fig.height=10, fig.width=10}
+
+file = paste(cache.dir, 'rand.Rdata', sep='/')
+
+if (file.exists(file)) {
+  load(file, verbose=T)
+} else {
+  rand.patients = sum.patient.data
+  rand.patients$Status = sample( c('NP','P'), nrow(sum.patient.data), replace=T, prob=c(0.5,0.5))
+
+  r.labels = unlist(apply(rand.patients[,c('Patient','Status')],1, function(st) {
+    info = patient.data[[ st['Patient'] ]]$info
+    lbl = rep(as.integer( st['Status'] == 'P' ), nrow(info))
+    names(lbl) = info$Samplename
+    return(lbl)
+  }))
+  names(r.labels) = sub('.*\\.', '',  names(r.labels))
+
+  # sort in label order
+  r.dysplasia.df = t(mergedDf[,names(r.labels)])
+
+  r.coefs = list(); r.plots = list()
+  folds = 10; splits=5
+  for (a in c(0,0.5,1)) {
+    fitR <- glmnet(r.dysplasia.df, r.labels, alpha=a, nlambda=nl, family='binomial') # all patients
+
+    cv.R = crossvalidate.by.patient(x=r.dysplasia.df, y=r.labels, lambda=fitR$lambda, pts=pts, a=a, nfolds=folds, splits=splits, fit=fitR)
+
+    lambda.opt = ifelse( length(cv.R$lambda.1se) > 0, cv.R$lambda.1se, cv.R$lambda.min )
+
+    coef.opt = as.data.frame(non.zero.coef(fitR, lambda.opt))
+    r.coefs[[as.character(a)]] = coef.stability(coef.opt, cv.R$non.zero.cf)
+
+    r.plots[[as.character(a)]] = arrangeGrob(cv.R$plot+ggtitle('Classification'), cv.R$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
+  }
+  save(r.plots, r.coefs, file=file) 
+}
+
+do.call(grid.arrange, c(r.plots, top='All samples, Randomized labels (by patient)'))
+
+```
+
+# Leave one out and evaluate predictions
 
@@
     fitLOO <- glmnet(training, labels[train.rows], alpha=a, family='binomial', nlambda=nl) # all patients
-    cv = crossvalidate.by.patient(x=training, y=labels[train.rows], lambda=fitLOO$lambda, 
+    #l = sort(c(fitLOO$lambda, seq(exp(-5), exp(-10), -0.001),
+    #      seq(exp(-10), exp(-15), -1e-7), 
+    #      seq(exp(-15), exp(-20), -1e-9)), decreasing=T)
+    l = fitLOO$lambda
+    cv = crossvalidate.by.patient(x=training, y=labels[train.rows], lambda=l, 
                                   pts=subset(pts, Samplename %in% samples), a=a, nfolds=10, splits=5, fit=fitLOO)
@@
                   'pred'=sum(as.integer(Prediction > 0.5)), 'samples'=length(PID), 
-                  'final.endo'= length(which(grepl('HGD|IMC', Pathology))) > 0,
+                  #'final.endo'= length(which(grepl('HGD|IMC', Pathology))) > 0,
                   'months.before.final'= (max(pt$Endoscopy.Year) - unique(Endoscopy.Year))*12 )
@@
     if (i > nrow(x)) return(NA)
-    cbind(x[i, c('Endoscopy.Year','pred','samples','final.endo', 'months.before.final')], i)
+    cbind(x[i, c('Endoscopy.Year','pred','samples', 'months.before.final')], i)
   }))
@@
 
-```{r, echo=F, warning=F, message=F, fig.height=8, fig.width=8}
+```{r, echo=F, warning=F, message=F, fig.height=8, fig.width=8, eval=F}
 plot.endo( subset(endos, Patient %in% subset(me, max.endo <= 2 & max.endo > 0)$Patient), min(endos$months.before.final), max(endos$months.before.final)  )
@@
 
-```{r, echo=F, warning=F, message=F, fig.height=4, fig.width=4}
+```{r, echo=F, warning=F, message=F, fig.height=4, fig.width=4, eval=F}
 plot.endo( subset(endos, Patient %in% subset(me, max.endo == 0)$Patient),min(endos$months.before.final), max(endos$months.before.final) )
@@
 
-binomial.deviance
-
 ggplot(all, aes(Prediction.Dev.Resid, Prediction, color=Status)) + geom_point()  + 
@@
 
-# Randomize the labels
-
-By patient.  This should break the models...and as would be expected, the models are really poor with perfomance at 50% or less (coin toss).
-
-```{r, echo=T, warning=F, message=F, fig.height=10, fig.width=10}
-
-file = paste(cache.dir, 'rand.Rdata', sep='/')
-
-if (file.exists(file)) {
-  load(file, verbose=T)
-} else {
-  rand.patients = sum.patient.data
-  rand.patients$Status = sample( c('NP','P'), nrow(sum.patient.data), replace=T, prob=c(0.5,0.5))
-
-  r.labels = unlist(apply(rand.patients[,c('Patient','Status')],1, function(st) {
-    info = patient.data[[ st['Patient'] ]]$info
-    lbl = rep(as.integer( st['Status'] == 'P' ), nrow(info))
-    names(lbl) = info$Samplename
-    return(lbl)
-  }))
-  names(r.labels) = sub('.*\\.', '',  names(r.labels))
-
-  # sort in label order
-  r.dysplasia.df = t(mergedDf[,names(r.labels)])
-
-  r.coefs = list(); r.plots = list()
-  folds = 10; splits=5
-  for (a in c(0,0.5,1)) {
-    fitR <- glmnet(r.dysplasia.df, r.labels, alpha=a, nlambda=nl, family='binomial') # all patients
-
-    cv.R = crossvalidate.by.patient(x=r.dysplasia.df, y=r.labels, lambda=fitR$lambda, pts=pts, a=a, nfolds=folds, splits=splits, fit=fitR)
-
-    lambda.opt = ifelse( length(cv.R$lambda.1se) > 0, cv.R$lambda.1se, cv.R$lambda.min )
-
-    coef.opt = as.data.frame(non.zero.coef(fitR, lambda.opt))
-    r.coefs[[as.character(a)]] = coef.stability(coef.opt, cv.R$non.zero.cf)
-
-    r.plots[[as.character(a)]] = arrangeGrob(cv.R$plot+ggtitle('Classification'), cv.R$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
-  }
-  save(r.plots, r.coefs, file=file) 
-}
-
-do.call(grid.arrange, c(r.plots, top='All samples, Randomized labels (by patient)'))
-
-```
+## Data Quality
+
+### By Batch first
+
+```{r, echo=T, warning=F, message=F, eval=F}
+all = do.call(rbind, pg.samp)
+
+prog = do.call(rbind, pg.samp[subset(sum.patient.data, Status == 'P')$Patient])
+nonp = do.call(rbind, pg.samp[subset(sum.patient.data, Status == 'NP')$Patient])
+
+## Similar numbers of each in these batches so how did we do?
+tfPR = sapply( c('Exome_subcohort','Batch_1','Batch_2','Batch_3','Batch_4'), function(b) {
+  pred = subset(prog, Batch.Name == b)$Prediction
+  predB = subset(nonp, Batch.Name == b)$Prediction
+
+  tpr = length(which(pred > 0.5))/length(pred)
+  fpr = length(which(predB > 0.5))/length(predB)
+  cbind.data.frame('TPR'=tpr, 'FPR'=fpr)
+})
+
+tfPR
+
+## Exome_cohort fpr driven by the two patients that we predicted to be progressors in nearly every sample
+length(which(pg.samp$AD0591$Prediction > 0.5))
+length(which(pg.samp$AH0329$Prediction > 0.5))
+
+## Except the exome cohort, the TPR rate appears to be right around 48-68%
+## FPR is over 20% in Batch_4 though
+
+unique(subset(nonp, Batch.Name == 'Batch_4')$Patient)
+
+pg.samp$AHM0100$Prediction > 0.5
+pg.samp$AHM0121$Prediction > 0.5
+pg.samp$AD0644$Prediction > 0.5
+
+cb = combn( as.character(unique(all$Batch.Name)), 2)
+
+batch =  sapply(1:ncol(cb), function(i) {
+  sapply(c('SampleSD','SampleMEAN', 'Total.Reads'), function(x) {
+    tt = t.test(subset(all, Batch.Name == cb[1,i])[[x]], subset(all, Batch.Name == cb[2,i])[[x]])  
+    tt$p.value
+  })
+}) 
+```
+
+```{r, echo=T, warning=F, message=F, eval=F}
+smp<-function(df) {
+  df[,c('Prediction', 'Prediction.Dev.Resid', 'SampleMEAN', 'SampleSD', 'Barretts.Cellularity', 'Total.Reads', 'p53.Status', 'PID')]
+}
+
+prog = do.call(rbind, lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], smp) )
+nonp = do.call(rbind, lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Patient], smp) )
+
+GGally::ggcorr(data=prog, method=c('pairwise','pearson'), label=T) + labs(title='Pearson correlation for several QC measures')
+
+# Compare poorly predicted to well predicted based on the sample SD measures
+ttfunc<-function(col, df=NULL) {
+  tt = t.test(subset(df, Prediction < 0.5)[[col]], subset(df, Prediction > 0.5)[[col]])
+  cbind.data.frame('p.val'=tt$p.value, 'statistic'=tt$statistic)
+}
+
+```
+
+```{r, echo=T, warning=F, message=F, eval=F}
+ pander(sapply(colnames(prog)[-(1:2)], ttfunc, df=prog), caption='Progressor t-test stats for QC measures between predictions probabilies  >&< 0.5 ')
+
+ pander(sapply(colnames(nonp)[-(1:2)], ttfunc, df=nonp), caption='Non-progressor t-test stats for QC measures between predictions probabilies  >&< 0.5 ')
+
+hist(prog$SampleSD)
+
+subset(prog, SampleSD <= 0.025)
+
+```

### 2017-06-08 RegressionPvNP.Rmd
More analysis

@@
 validation.patient.data = patient.data
+length(validation.patient.data)
 
@@
 validation.patient.info = read.patient.info(patient.file, set='Test')
+validation.patient.info = arrange(validation.patient.info, Status, Patient, Endoscopy.Year, Pathology)
 sum.validation.info = summarise.patient.info(validation.patient.info)
@@
 patient.info = arrange(patient.info, Status, Patient, Endoscopy.Year, Pathology)
-
 sum.patient.data = summarise.patient.info(patient.info)
@@
 
-patient.data[which(sapply(patient.data, function(df) nrow(df$seg.vals)) == 0)] = NULL
+if (length(which(sapply(validation.patient.data, function(df) nrow(df$seg.vals)) == 0)) > 0)
+  warning(paste("Patients missing data:",
+                paste(names(which(sapply(patient.data, function(df) nrow(df$seg.vals)) == 0)), collapse=', ')))
+
+nor = which(sapply(patient.data, function(df) nrow(df$seg.vals)) == 0)
+if (length(nor) > 0) patient.data[nor] = NULL
+
+nor = which(sapply(validation.patient.data, function(df) nrow(df$seg.vals)) == 0)
+if (length(nor) > 0) validation.patient.data[nor] = NULL
+
 sum.patient.data = as.data.frame(subset(sum.patient.data, Patient %in% names(patient.data))) ## For now
@@
 
-pander(sum.patient.data[,c('Patient', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left')
+for (pt in names(validation.patient.data)) {
+  #print(pt)
+  #print(patient.data[[pt]]$info$Samplename %in% colnames(patient.data[[pt]]$seg.vals)[-(1:5)])
+  validation.patient.data[[pt]]$info = validation.patient.data[[pt]]$info[validation.patient.data[[pt]]$info$Samplename %in% colnames(validation.patient.data[[pt]]$seg.vals)[-(1:5)],]
+}
+
+pander(sum.patient.data[,c('Patient', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Training patient set')
+
+pander(sum.validation.info[,c('Patient', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Validation patient set')
 
@@
 ```{r tile, echo=F, message=F, warning=F, include=F}
-tile.w=1e07
-#tile.w=5e06
+#tile.w=1e07
+tile.w=5e06
 tile.files = grep(as.character(tile.w), analysis.files, value=T, fixed=T)
-load(tile.files[1], verbose=T)
+load(tile.files[grep('Test|validation', tile.files)], verbose=T)
+mergedDfValidation = mergedDf
+dim(mergedDfValidation)
+
+load(tile.files[grep('Test|validation', tile.files, invert=T)], verbose=T)
 dim(mergedDf)
@@
 dim(dysplasia.df)
+
+validation.df = t(mergedDfValidation[, grep('^D', colnames(mergedDfValidation))])
+dim(validation.df)
 ```
@@
 
-crossvalidate.by.patient<-function(x,y,lambda,pts,a=1,nfolds=10, splits=5, fit=NULL, minR=0.2, select='classification') {
+crossvalidate.by.patient<-function(x,y,lambda,pts,a=1,nfolds=10, splits=5, fit=NULL, minR=0.2, select='deviance') {
   if (nfolds > 5) minR = 0.1
+  x = as.matrix(x)
   message(paste("Running", splits, "splits",nfolds,"times on", paste(dim(x), collapse=':'), 'alpha=',a ))
@@
       # pre-spec lambda seq
-      fitCV <- glmnet(training, y[-test.rows], lambda=lambda, family='binomial', alpha=a) 
+      fitCV <- glmnet(training, y[-test.rows], lambda=lambda, alpha=a, family='binomial') 
       # autoplot(fit) + theme(legend.position="none")
@@
       # Confusion matrix: quantitiative, pos results + neg results / number of test rows
-      pred <- pi.hat(predict(fitCV, test, type='link')) 
+      pred <- pi.hat(predict(fitCV, test, type='link'))  # pi.hat turns these into probabilities from logit
       cv.class[i,] = apply(pred, 2, function(p) {
@@
 
-    arrange(subset(df, mean < se1 & !lambda.min & log.lambda > subset(df, lambda.min)$log.lambda ), -mean)[1:10,]
+    lambda.1se.search = subset(df, mean < se1 & !lambda.min & log.lambda > subset(df, lambda.min)$log.lambda) 
+    if (nrow(lambda.1se.search) > 0) {
+      arrange(lambda.1se.search, -mean)[1:10,]
       df[arrange(subset(df, mean < se1 & !lambda.min & log.lambda > subset(df, lambda.min)$log.lambda ), -mean)[1, 'lambda.at'], 'lambda.1se'] = T
+    }
     #df[df$log.lambda < subset(df, mean == max(mean))$log.lambda-sd(df$log.lambda),][1, 'lambda-1se'] = TRUE
   } else if (select == 'deviance') {
-    df$lambda.min = df$mean.b.dev == min(df$mean.b.dev)
-    se1 = df[df$lambda.min == T, 'mean.b.dev'] - df[df$lambda.min == T, 'sme.b.dev']
-    df[arrange(subset(df, mean.b.dev <= se1 & !lambda.min), -mean.b.dev)[1, 'lambda.at'], 'lambda.1se'] = T
-    #df[df$mean.b.dev >= df[df$lambda.min == T, 'mean.b.dev'] + sd(df$mean.b.dev), ][1, 'lambda.1se'] = T
+    df[which.min(df$mean.b.dev), 'lambda.min'] = T
+
+    lmin = subset(df, lambda.min)$lambda
+    df[subset(df, log.lambda <= log(lmin) + log(sd(lambda))/2 )[1, 'lambda.at'], 'lambda.1se'] = T
   }
@@
     }
-    cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=l, pts=pts, a=a, nfolds=folds, splits=splits, fit=fit0)
+    cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=l, pts=pts, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance')
 
@@
 ```{r xvalpt2, message=F, warning=F, echo=T}
+select.alpha = as.character(0.9)
+
 performance.at.1se = do.call(rbind, performance.at.1se)
@@
                            dimnames=list( unique(names(table(unlist(lapply(coef.stable, names))))), names(coef.stable) )))
-
 for (i in names(coef.stable)) 
@@
 
-table(rowSums(cfs))
-
 coef.stable = lapply(coef.stable, function(cf) cf/(folds*splits))
@@
 stability = do.call(rbind, lapply(coef.stable, function(cf) {
-  cbind('75%'=length(which(cf >= 0.75)),  '50%'=length(which(cf >= 0.5)),  '<50%'=length(which(cf < 0.5)))
+  cbind('>=75%'=length(which(cf >= 0.75)),  '>=50%'=length(which(cf >= 0.5)),  '>=25%'=length(which(cf >= 0.25)), 'Total'=length(cf))
 }))
@@
 
-pander(as.data.frame(sort(names(which((coef.stable[['1']]) >= 0.75)))), justify='left')
+pander(stability, justify='left', caption='Number of features stable in n% or more folds')
 ```
@@
 
+### Validation Data
+
+TBD
+
+```{r, message=F, warning=F, echo=F, eval=F}
+
+fitV <- glmnet(dysplasia.df, labels, alpha=select.alpha, nlambda=nl, family='binomial') # all patients
+#autoplot(fitV, xvar='lambda', main=paste('fit0, all samples, alpha=',a,sep='')) + theme(legend.position='none') 
+cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=fitV$lambda, pts=pts, a=select.alpha, nfolds=folds, splits=splits, fit=fitV)
+lambda.opt = ifelse( length(cv.patient$lambda.1se) > 0, cv.patient$lambda.1se, cv.patient$lambda.min )
+
+pred = predict(fitV, newx=validation.df, s=lambda.opt, type='response')
+
+statusV = unlist(sapply(validation.patient.data, function(df) {
+  df$info = arrange(df$info, Endoscopy.Year, Pathology)
+  label = as.integer(df$info$Status == 'P') #as.integer(df$info$Pathology %in% c('HGD', 'IMC'))
+  names(label) = df$info$Samplename
+  return(label)
+}))
+names(statusV) = sub('.*\\.', '',  names(statusV))
+
+# Progressors - by sample
+length( which(pred[ names(which(statusV == 1)), ] > 0.5) )/length(which(statusV == 1))
+
+# Non Progressors - by sample
+length( which(pred[ names(which(statusV == 0)), ] < 0.5) )/length(which(statusV == 0))
+
+```
+
 ### Feature stability
 
-`r pander(table(rowSums(cfs)), caption="Number of features found in 1, 2, or all 3 models by alpha 0.8, 0.9 and 1")`
+`r pander(colSums(cfs), caption="Number of features found in 1, 2, or all 3 models by alpha 0.8, 0.9 and 1.
+While there are fewer features for higher values of alpha, the features are the same across models.")`
 
@@
 
+```{r echo=F, warning=F, message=F, eval=F}
 ### Features in samples
-
-```{r echo=F, warning=F, message=F}
-
-length(which( coef.stable[['1']] >= 0.5) )
-
-regions = rownames(coefs[['1']])
+## TODO Look at the coefficients, are the high pos ones "gains" and the low neg ones "losses" etc
+plot(coefs[[select.alpha]][,1])
+
+length(which( coef.stable[[select.alpha]] >= 0.5) )
+
+regions = rownames(coefs[[select.alpha]])
 
@@
 
-#which(meantests < 0.05)
-
-summary(pg[,'1:119640299-129610322'])
-summary(nonpg[,'1:119640299-129610322'])
+which(meantests < 0.05)
+
+#summary(pg[,'1:119640299-129610322'])
+#summary(nonpg[,'1:119640299-129610322'])
 
@@
 ### Feature Information
+
+Using the features from the alpha=0.9 models
 
@@
 #topF = names(which((coef.stable[['1']]) >= 0.75))
-topF = names(coef.stable[['1']])
+topF = names(coef.stable[[select.alpha]])
 
@@
 
-# Gains/losses enriched in the top features?
-fisher.test(rbind(colSums(gl[,c('P.gain','P.loss')]),colSums(allgl[-feat,c('ALL.P.gain','ALL.P.loss')])  ))
-
-fisher.test(rbind(colSums(gl[,c('NP.gain','NP.loss')]),colSums(allgl[-feat,c('ALL.NP.gain','ALL.NP.loss')])  ))
+# Gains/losses enriched in the top features?  Appears so
+(fisher.test(rbind(colSums(gl[,c('P.gain','P.loss')]),colSums(allgl[-feat,c('ALL.P.gain','ALL.P.loss')])  )) )
+
+(fisher.test(rbind(colSums(gl[,c('NP.gain','NP.loss')]),colSums(allgl[-feat,c('ALL.NP.gain','ALL.NP.loss')])  )) )
 
@@
 # Number of segments is strongly correlated with the length of the chromosome
-with(topFRegions, cor.test(length, n.segments))
+#with(topFRegions, cor.test(length, n.segments))
 
 topFRegions$adj.info.content = topFRegions$protein_coding/(topFRegions$length^0.6)
-with(topFRegions, cor.test(length, adj.info.content))
+#with(topFRegions, cor.test(protein_coding, length)) # Adjust the length so that the correlation between length and
 
 topFRegions = transform(topFRegions, ai.cut=cut(adj.info.content, 5, labels=c(5:1), include.lowest=T))           # bin data
-ggplot(topFRegions, aes(reorder(chr, -adj.info.content), totalbp/length, fill=ai.cut)) + ylim(0, 0.5) +
+
+grid.arrange(ggplot(topFRegions, aes(reorder(chr, -adj.info.content), totalbp/length, fill=ai.cut)) + ylim(0, 0.5) +
   labs(y='Feature bp/total chr length', title='Ratio of covered base pairs', x='chr') +
   geom_bar(stat='identity') +  geom_text(aes(label=round(totalbp/length, 2)), vjust=-0.5) +
-  scale_fill_discrete(name='Length adjusted information content') + theme(legend.position = 'bottom')
+  scale_fill_discrete(name='Length adjusted information content') + theme(legend.position = 'bottom'), 
+  bottom='Ratio of base pair coverage per chromosome in the top non-zero features selected from the model.\nChromosomes are sorted and binned by their gene (e.g.
+information) content rather than by overall length.')
 
@@
 
+with(topFRegions, cor.test(adj.info.content, totalbp/length))
+
 bpCovered = with(topFRegions, sum(totalbp, na.rm=T)/sum(as.numeric(length)))
@@
 
-Strong correlation between the length adjusted information content (protein coding genes) and the covered bases in a chromosome r=`r round(ct$estimate, 2)` (p-value `r round(ct$p.value, 4)`).
-No features from chromosome 19 (the most gene dense chromosome) were selected.
-
-`r round(bpCovered, 2)*100`% of the total bases are within the top selected features.
+`r pander(ct, caption="Pearson correlation between the adjusted information content (genes) per chromosome, and the ratio of length of the features")`
+
+`r round(bpCovered, 2)*100`% of the all bps in the genome are within the top selected features.
 
@@
 #head(bm)
+
 bm = bm[-which(bm$hgnc_symbol == ""),]
-bm = bm %>% group_by(hgnc_symbol, chromosome_name, start_position, end_position) %>% summarise('exon_count' = length(ensembl_exon_id)) 
+bm = bm %>% group_by(ensembl_gene_id, hgnc_symbol, chromosome_name, start_position, end_position) %>% summarise('exon_count' = length(ensembl_exon_id)) 
 exons = bm %>% group_by(chromosome_name) %>% summarise('exons'=sum(exon_count))
+
+write.table(bm$ensembl_gene_id, sep='\t', col.names = F, row.names = F, quote = F, file='feature_genes.txt')
 
@@
 # Not related to the length of the segment
-with(topFSegments, cor.test(genes, (end-start)))
-
-head(bm)
+pander(with(topFSegments, cor.test(genes, (end-start))), caption="Number of genes per segment is not related to the segment length.")
 
@@
 
-with(topFRegions, cor.test(protein_coding, exon.cnt))
-with(topFRegions, cor.test(gene.cnt, exon.cnt)) # expected
-
-# Exons correlate with the adjusted information content measure
-with(topFRegions, cor.test(exon.cnt, adj.info.content))
-
-ggplot(topFRegions, aes(reorder(chr, -adj.info.content), gene.cnt/protein_coding, fill=ai.cut)) + geom_bar(stat='identity') +
+#with(topFRegions, cor.test(protein_coding, exon.cnt))
+#with(topFRegions, cor.test(gene.cnt, exon.cnt)) # expected
+
+# Exons correlate with the adjusted information content measure, expected
+#with(topFRegions, cor.test(exon.cnt, adj.info.content))
+
+grid.arrange(ggplot(topFRegions, aes(reorder(chr, -adj.info.content), gene.cnt/protein_coding, fill=ai.cut)) + geom_bar(stat='identity') +
   labs(y='Gene Ratio', x='chr', title='Feature genes vs all protein coding genes') +
   geom_text(aes(label=round(gene.cnt/protein_coding, 2)), vjust=-0.5) +
-  scale_fill_discrete(name='Length adjusted information content') + theme(legend.position = 'bottom')
+  scale_fill_discrete(name='Length adjusted information content') + theme(legend.position = 'bottom'), 
+  bottom='Ratio of genes found within the boundaries of the top features vs all genes on the chromosome.\nChromosomes are ordered and binned by adjusted gene (e.g.
+information) content rather than length.')
 
@@
 
+write.table(containedCC$ensembl_gene_id, quote=F, row.names = F, col.names = F, file='feature_cosmic.txt')
+
 topFRegions = merge(topFRegions, (containedCC %>% group_by(chromosome_name) %>% summarise('COSMIC.cont'=length(unique(hgnc_symbol)))), by.x='chr',by.y='chromosome_name', all.x=T)
 
-ggplot(topFRegions, aes(reorder(chr, -adj.info.content), COSMIC.cont/COSMIC.all, fill=ai.cut)) + geom_bar(stat='identity') +
+grid.arrange(ggplot(topFRegions, aes(reorder(chr, -adj.info.content), COSMIC.cont/COSMIC.all, fill=ai.cut)) + geom_bar(stat='identity') +
   labs(y='COSMIC Gene Ratio', x='chr', title='COSMIC genes: Feature vs all')+
   geom_text(aes(label=round(COSMIC.cont/COSMIC.all, 2)), vjust=-0.5) +
-  scale_fill_discrete(name='Length adjusted information content') + theme(legend.position = 'bottom')
+  scale_fill_discrete(name='Length adjusted information content') + theme(legend.position = 'bottom'), 
+  bottom='Again, but this time looking only at COSMIC genes, the ratio of COSMIC genes\nwithin the selected feature vs all COSMIC genes on that chromosome.\nChromosomes are ordered and colored by information content.')
+
+colSums(topFRegions[,c('COSMIC.all','COSMIC.cont')],na.rm=T)
+colSums(topFRegions[,c('protein_coding','gene.cnt')],na.rm=T)
+
+#table(topFRegions$ai.cut)
+
 ```
@@
 This appears to be a fairly large enrichment.
+
+### Gains vs Losses
+
+try
+
+gains = pmax(0, x-median(x))
+losses = pmin(0, x-median(x))
+
+```{r warning=F, message=F, fig.height=8, fig.width=8}
+run.cv<-function(x, y, alphas=c(0,0.2,0.5,1), ...) {
+  plots = list(); coefs = list(); performance.at.1se = list()
+  for (a in alphas) {
+    fitg <- glmnet(x, y, alpha=a, nlambda=nl, family='binomial') # all patients
+
+    l = fitg$lambda
+    if (a == 0) {
+      l = sort(c(fitg$lambda, seq(exp(-5), exp(-10), -1e-6),
+            seq(exp(-10), exp(-15), -1e-8),
+            seq(exp(-15), exp(-20), -1e-9)), decreasing=T)
+    }
+
+    cvp = crossvalidate.by.patient(x, y, fit=fitg, lambda=l, a=a, ...)
+    plots[[as.character(a)]] = arrangeGrob(cvp$plot+ggtitle('Classification'), cvp$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
+
+    coef.1se = as.data.frame(non.zero.coef(fitg, cvp$lambda.1se))
+    coefs[[as.character(a)]] = coef.stability(coef.1se, cvp$non.zero.cf)
+
+    performance.at.1se[[as.character(a)]] = subset(cvp$lambdas, lambda.1se == T)
+
+  }
+  return(list('plots'=plots, 'coefs'=coefs, 'performance'=performance.at.1se))
+}
+
+file = paste(cache.dir, 'gain.loss.Rdata', sep='/')
+if (file.exists(file)) {
+  message(paste("loading", file))
+  load(file, verbose=T)
+} else {
+  gain.df = as.data.frame(matrix(nrow=nrow(dysplasia.df), ncol=ncol(dysplasia.df),data=0, dimnames=list(rownames(dysplasia.df), colnames(dysplasia.df))))
+  loss.df = gain.df
+
+  for (i in 1:nrow(dysplasia.df)) {
+     #gain.df[i,] = as.integer(dysplasia.df[i,] >= 1.1)
+     #loss.df[i,] = as.integer(dysplasia.df[i,] <= 0.9)
+     #all.binary[i,] = as.integer(dysplasia.df[i,] >= 1.1 | dysplasia.df[i,] <= 0.9)
+    gain.df[i,] = pmax(0, dysplasia.df[i,] - median(dysplasia.df[i,]))
+    loss.df[i,] = pmin(0, dysplasia.df[i,] - median(dysplasia.df[i,]))
+  } 
+  #hist(colSums(gain.df))
+  #hist(colSums(loss.df))
+  #hist(colSums(all.binary))
+
+  gp = run.cv(x=as.matrix(gain.df), y=labels, pts=pts, nfolds=folds, splits=splits, alphas=c(0,0.5,0.8,1))
+  lp = run.cv(x=as.matrix(loss.df), y=labels, pts=pts, nfolds=folds, splits=splits, alphas=c(0,0.5,0.8,1))
+
+  colnames(gain.df) = paste(colnames(gain.df), '_GAIN', sep='')
+  colnames(loss.df) = paste(colnames(loss.df), '_LOSS', sep='')
+
+  gl.df = merge(gain.df, loss.df, by='row.names')
+  rownames(gl.df) = gl.df$Row.names
+  gl.df$Row.names = NULL
+
+  both = run.cv(x=as.matrix(gl.df), y=labels, pts=pts, nfolds=folds, splits=splits, alphas=c(0,0.5,0.8,1))
+
+  save(gp, lp, all, file=file)
+}
+
+do.call(grid.arrange, c(gp$plots, top='All samples, GAINS 10fold, 5 splits'))
+do.call(grid.arrange, c(lp$plots, top='All samples, LOSSES 10fold, 5 splits'))
+do.call(grid.arrange, c(both$plots, top='All samples, gains & losses 10fold, 5 splits'))
+
+```
 
@@
 stability = do.call(rbind, lapply(coef.stable, function(cf) {
-  cbind('75%'=length(which(cf >= 0.75)),  '50%'=length(which(cf >= 0.5)),  '<50%'=length(which(cf < 0.5)))
+  cbind('>=75%'=length(which(cf >= 0.75)),  '>=50%'=length(which(cf >= 0.5)),  '>=50%'=length(which(cf >= 0.25)), 'Total'=length(cf))
 }))
@@
 They also share all of the non-zero coefficients (`r which.max(sapply(coef.stable, length))` identifies the most non-zero coefs `r max(sapply(coef.stable, length))`).
-
-`r pander(table(rowSums(cfs)), justify='left')`
 
@@
 
-`r pander(table(rowSums(cfs)), justify='left')`
-
 ## Compare progressors with many samples vs few
 
-```{r, echo=T, message=F, warning=F, fig.height=16, fig.width=16}
+```{r manyfew, echo=T, message=F, warning=F, fig.height=16, fig.width=16}
 m = median(subset(sum.patient.data, Status == 'P')$total.samples)
@@
 file = paste(cache.dir, 'few.Rdata', sep='/')
-
 if (file.exists(file)) {
@@
 
-```{r, echo=F, warning=F, message=F, fig.height=8, fig.width=8, eval=F}
+```{r, echo=F, warning=F, message=F, fig.height=6, fig.width=4}
 plot.endo( subset(endos, Patient %in% subset(me, max.endo <= 2 & max.endo > 0)$Patient), min(endos$months.before.final), max(endos$months.before.final)  )
@@
 
-```{r, echo=F, warning=F, message=F, fig.height=4, fig.width=4, eval=F}
+```{r, echo=F, warning=F, message=F, fig.height=4, fig.width=4}
 plot.endo( subset(endos, Patient %in% subset(me, max.endo == 0)$Patient),min(endos$months.before.final), max(endos$months.before.final) )
@@
 
-```{r, echo=T, warning=F, message=F, eval=F}
+```{r dq_batch, echo=T, warning=F, message=F}
 all = do.call(rbind, pg.samp)
@@
 ## Similar numbers of each in these batches so how did we do?
-tfPR = sapply( c('Exome_subcohort','Batch_1','Batch_2','Batch_3','Batch_4'), function(b) {
+tfPR = sapply(as.character(unique(patient.info$Batch.Name)), function(b) {
   pred = subset(prog, Batch.Name == b)$Prediction
@@
 
+  # Only for progressors because we know for certain they progress, the non-ps may just not have progressed *yet*
   tpr = length(which(pred > 0.5))/length(pred)
@@
 })
-
-tfPR
-
-## Exome_cohort fpr driven by the two patients that we predicted to be progressors in nearly every sample
-length(which(pg.samp$AD0591$Prediction > 0.5))
-length(which(pg.samp$AH0329$Prediction > 0.5))
-
-## Except the exome cohort, the TPR rate appears to be right around 48-68%
-## FPR is over 20% in Batch_4 though
-
-unique(subset(nonp, Batch.Name == 'Batch_4')$Patient)
-
-pg.samp$AHM0100$Prediction > 0.5
-pg.samp$AHM0121$Prediction > 0.5
-pg.samp$AD0644$Prediction > 0.5
-
-cb = combn( as.character(unique(all$Batch.Name)), 2)
-
-batch =  sapply(1:ncol(cb), function(i) {
-  sapply(c('SampleSD','SampleMEAN', 'Total.Reads'), function(x) {
-    tt = t.test(subset(all, Batch.Name == cb[1,i])[[x]], subset(all, Batch.Name == cb[2,i])[[x]])  
-    tt$p.value
-  })
-}) 
-```
-
-```{r, echo=T, warning=F, message=F, eval=F}
-smp<-function(df) {
-  df[,c('Prediction', 'Prediction.Dev.Resid', 'SampleMEAN', 'SampleSD', 'Barretts.Cellularity', 'Total.Reads', 'p53.Status', 'PID')]
-}
-
-prog = do.call(rbind, lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], smp) )
-nonp = do.call(rbind, lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Patient], smp) )
-
-GGally::ggcorr(data=prog, method=c('pairwise','pearson'), label=T) + labs(title='Pearson correlation for several QC measures')
-
-# Compare poorly predicted to well predicted based on the sample SD measures
-ttfunc<-function(col, df=NULL) {
-  tt = t.test(subset(df, Prediction < 0.5)[[col]], subset(df, Prediction > 0.5)[[col]])
-  cbind.data.frame('p.val'=tt$p.value, 'statistic'=tt$statistic)
-}
-
-```
-
-```{r, echo=T, warning=F, message=F, eval=F}
- pander(sapply(colnames(prog)[-(1:2)], ttfunc, df=prog), caption='Progressor t-test stats for QC measures between predictions probabilies  >&< 0.5 ')
-
- pander(sapply(colnames(nonp)[-(1:2)], ttfunc, df=nonp), caption='Non-progressor t-test stats for QC measures between predictions probabilies  >&< 0.5 ')
-
-hist(prog$SampleSD)
-
-subset(prog, SampleSD <= 0.025)
-
-```
+```
+
+`r pander(t(tfPR), justify='left', caption=paste('True positive rate is between', paste(round(range(tfPR['TPR',], na.rm=T), 2), collapse='-')))`
+
+```{r dq_batch2, echo=F, warning=F, message=F}
+
+pilot = lapply(pg.samp, function(df) subset(df, Batch.Name == 'NP_Pilot_Study'))
+pilot = do.call(rbind, pilot[ which(sapply(pilot, nrow) > 0) ])
+pilot = subset(pilot, Status == 'NP')
+length(which(pilot$Prediction > 0.5))/nrow(pilot)
+
+p2 = subset(pilot, Patient %nin% c('AD0591', 'AH0329'))
+length(which(p2$Prediction > 0.5))/nrow(p2)
+
+exome = lapply(pg.samp, function(df) subset(df, Batch.Name == 'Exome_subcohort'))
+exome = do.call(rbind, exome[ which(sapply(exome, nrow) > 0) ])
+exome = subset(exome, Status == 'NP')
+length(which(exome$Prediction > 0.5))/nrow(exome)
+
+e2 = subset(exome, Patient %nin% c('AD0591', 'AH0329'))
+length(which(e2$Prediction > 0.5))/nrow(e2)
+
+```
+
+The false positive rate in both the exome and NP_pilot batches is driven by two patients that we predict to be progressors in nearly every sample: AD0591 and AH0329.
+Both patients had samples in the exome batch, while AH0329 also had samples in the NP_Pilot batch.
+Without those patients the FPR in the exome batch was `r round(length(which(e2$Prediction > 0.5))/nrow(e2), 2)*100`% for `r length(unique(e2$Patient))` patient(s), and in the NP_Pilot batch the FPR was `r round(length(which(p2$Prediction > 0.5))/nrow(p2), 2)*100`% for `r length(unique(p2$Patient))` patient(s).
+
+The somewhat lower Batch_1 prediction rate is driven by the samples from AHM0277 as only `r length(which(pg.samp$AHM0277$Prediction > 0.5))/nrow(pg.samp$AHM0277)*100`% of this patient's samples were predicted to progress.
+Batch_3 is driven by three other patients that have low prediction rates: PR1_ADH_076, PR1_HIN_046, and PR1_WSH_026.
+
+### Look at these specific patients
+
+```{r, echo=F, eval=F}
+
+pg.samp$AD0591
+pg.samp$AH0329
+pg.samp$PR1_ADH_076
+pg.samp$PR1_HIN_046
+pg.samp$PR1_WSH_026
+
+```
+
+## Cellularity or general data quality measures?
+
+```{r, echo=F, warning=F, message=F}
+all$SampleCoV = with(all, SampleSD/SampleMEAN)
+
+prog = subset(all, Status == 'P', select=c('Prediction', 'Prediction.Dev.Resid', 'SampleMEAN', 'SampleSD', 'SampleCoV', 'Total.Reads'))
+nonp = subset(all, Status == 'NP', select=c('Prediction', 'Prediction.Dev.Resid', 'SampleMEAN', 'SampleSD', 'SampleCoV', 'Total.Reads'))
+
+#GGally::ggcorr(data=prog, method=c('pairwise','pearson'), label=T) + labs(title='Pearson correlation for several QC measures')
+
+library(Hmisc)
+flattenCorrMatrix <- function(cormat, pmat) {
+  ut <- upper.tri(cormat)
+  data.frame(
+    row = rownames(cormat)[row(cormat)[ut]],
+    column = rownames(cormat)[col(cormat)[ut]],
+    cor  =(cormat)[ut],
+    p = pmat[ut]
+    )
+}
+
+rc = rcorr( as.matrix(all[, c('Prediction', 'Prediction.Dev.Resid', 'SampleMEAN', 'SampleSD','SampleCoV', 'Total.Reads')]) )
+
+frc = flattenCorrMatrix(rc$r, rc$P)
+pander(frc[which(p.adjust(frc$p, 'bonferroni') < 0.05),], caption='Prediction and SampleSD have a small but significant correlation after Bonferroni correction.
+This could indicate cellularity or sample quality is influencing the predictions.')
+```
+
+```{r}   
+# Compare poorly predicted to well predicted based on the sample SD measures, no difference
+wilcox.test(subset(prog, Prediction > 0.5)$SampleSD, subset(prog, Prediction < 0.5)$SampleSD)
+wilcox.test(subset(nonp, Prediction > 0.5)$SampleSD, subset(nonp, Prediction < 0.5)$SampleSD)
+
+wilcox.test(prog$SampleSD, nonp$SampleSD) # Appears to be a small difference in this measure, where is it from?
+
+## Progressors
+tprows = which(prog$Prediction > 0.5)
+
+cor.test(prog[tprows, 'SampleSD'], prog[tprows, 'Prediction'] ) # Correlated for True Pos predictions
+cor.test(prog[-tprows, 'SampleSD'], prog[-tprows, 'Prediction'] ) # But not for the poorly predicted
+
+badrows = which(prog$Prediction < 0.3)
+cor.test(prog[badrows, 'SampleSD'], prog[badrows, 'Prediction'] ) # But not for the poorly predicted
+
+# There's no real difference between the sample 'cellularity' or quality between the good or bad predictions in the progressors
+wilcox.test(prog[tprows, 'SampleSD'], prog[-tprows, 'SampleSD']) 
+wilcox.test(prog[tprows, 'SampleCoV'], prog[-tprows, 'SampleCoV'])
+
+## Non-Progressors, this is a little more complicated as we cannot be sure of True Positives but we'll look anyhow
+tprowsNP = which(nonp$Prediction < 0.5)
+cor.test(nonp[tprowsNP, 'SampleSD'], nonp[tprowsNP, 'Prediction'] ) # Correlated for True Pos predictions
+cor.test(nonp[-tprowsNP, 'SampleSD'], nonp[-tprowsNP, 'Prediction'] ) # But not for the poorly predicted
+
+badrows = which(nonp$Prediction > 0.7)
+cor.test(nonp[badrows, 'SampleSD'], nonp[badrows, 'Prediction'] ) # But not for the poorly predicted
+
+# There's no real difference between the sample 'cellularity' or quality between the good or bad predictions in the progressors
+wilcox.test(nonp[tprowsNP, 'SampleSD'], nonp[-tprowsNP, 'SampleSD']) 
+wilcox.test(nonp[tprowsNP, 'SampleCoV'], nonp[-tprowsNP, 'SampleCoV'])
+
+## How about the TPs for Progressors and the predicted Non-P?
+wilcox.test(nonp[-tprowsNP, 'SampleSD'], prog[tprows, 'SampleSD']) # Basically the same
+wilcox.test(nonp[tprowsNP, 'SampleSD'], prog[-tprows, 'SampleSD']) # Basically the same
+
+ggplot(prog, aes(x=SampleSD)) + 
+  geom_histogram(data=subset(prog, Prediction > 0.5), aes(SampleSD, fill='TP'), alpha=0.5) +
+  geom_histogram(data=subset(prog, Prediction < 0.5), aes(SampleSD, fill='FN'), alpha=0.5) + 
+  labs(title='SD of Samples vs True positive ~ False negative Progressors')
+
+ggplot(nonp, aes(x=SampleSD)) + 
+  geom_histogram(data=subset(nonp, Prediction < 0.5), aes(SampleSD, fill='TN'), alpha=0.5) + 
+  geom_histogram(data=subset(nonp, Prediction > 0.5), aes(SampleSD, fill='FP'), alpha=0.5) +
+  labs(title='SD of Samples vs False positive ~ True negative Non-Progressors')
+
+```

### 2017-07-05 Clustering.Rmd
...

@@
 data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
+analysis.files = list.files(paste(data, 'Analysis', sep='/'), full.names=T)
 plot.dir = paste(data, 'Analysis/multipcf_plots_fitted_perPatient', sep='/')
@@
   stop(paste("Missing/too many patient info file(s) in", data))
+demo.file = grep('Demographics_full.txt', data.files, value=T)
 
-load('Training_patients.Rdata', verbose=T)
+load(grep('Training_patients.Rdata', analysis.files, value=T), verbose=T)
 
-patient.info = read.patient.info(patient.file)
-patient.info$Patient = gsub("/", "_", patient.info$Patient)
+patient.info = read.patient.info(patient.file, demo.file)
 #head(patient.info)
@@
 
-```{r global_pt, echo=F, message=F, warning=F, error=F}
+```{r global_pt, echo=F, message=F, warning=F, error=F, eval=F}
 be = subset(patient.info, Pathology == 'BE')
@@
             get.adjp=T, na.rm=T, test='t.twosamp.equalvar')
+
+# Currently with multi test correction the samples for which cellularity and #of reads are available show no difference for either status (P vs NP)
+# `r pander(mtp2@adjp, justify='left', caption="Adjusted p-value")`
+# 
+# or for samples that are LGD/IMC/HGD vs BE
+# `r pander(mtp1@adjp, justify='left', caption="Adjusted p-value")`
+# 
+# However, it should be noted that pathologists do not like calling cellularity on Barrett's generally so it's a fairly inaccurate measure.
+# `r pander(table( patient.info[,c('Status','Barretts.Cellularity')]), caption="Numbers of cellularity calls by status")`
+# 
+# With regards to p53 staining, non-progressors have no p53 positive tests, while progressors had `r p53Status[2,2]` overall.
+# `r pander(p53Status, justify='left', caption="P53 staining status for progressors vs non")`
+# Of these `r table(be$p53.Status)[['1']]` were positive in a BE sample, and `r table(lgd$p53.Status)[['1']]` were positive when LGD samples were included.
+
 ```
-
-Currently with multi test correction the samples for which cellularity and #of reads are available show no difference for either status (P vs NP)
-`r pander(mtp2@adjp, justify='left', caption="Adjusted p-value")`
-
-or for samples that are LGD/IMC/HGD vs BE
-`r pander(mtp1@adjp, justify='left', caption="Adjusted p-value")`
-
-However, it should be noted that pathologists do not like calling cellularity on Barrett's generally so it's a fairly inaccurate measure.
-`r pander(table( patient.info[,c('Status','Barretts.Cellularity')]), caption="Numbers of cellularity calls by status")`
-
-With regards to p53 staining, non-progressors have no p53 positive tests, while progressors had `r p53Status[2,2]` overall.
-`r pander(p53Status, justify='left', caption="P53 staining status for progressors vs non")`
-Of these `r table(be$p53.Status)[['1']]` were positive in a BE sample, and `r table(lgd$p53.Status)[['1']]` were positive when LGD samples were included.
 
@@
 apclust.data<-function(segdata, samples) {
-  x1 = segdata[,intersect(colnames(segdata), samples)]
+  x1 = as.matrix(segdata[,intersect(colnames(segdata), samples)])
   #x1 = x1[, samples]
@@
     mutate(location=paste(paste(chrom, arm, sep=''), '.', start.pos, '-', end.pos, sep='')))$location
+  if (length(samples) == 1) colnames(x1) = samples
   return(x1)  
@@
 for (pt in sum.patient.data$Patient) {
+  print(pt)
   patient.data[[pt]]$info = arrange(patient.data[[pt]]$info,  Pathology, Endoscopy.Year)
@@
 
-  gg = ggplot(melt(x1), aes(x=variable, y=value, group=variable, fill=variable)) + 
+  gg = ggplot(melt(x1), aes(x=Var2, y=value, group=Var2, fill=Var2)) + 
     geom_boxplot(notch=T, show.legend=F) +
     theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position='none') +
-    labs(title=paste(pt, patient.data[[pt]]$info$Status[1], sep=' - ')) +   theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank())
+    labs(title=paste(pt, patient.data[[pt]]$info$Status[1], sep=' - '), x='')
 
@@
   sim[pt, 'clusters'] = length(clusters(ac))
-  sim[pt, 'max.sd'] = max(sapply(x1, sd))
+  sim[pt, 'max.sd'] = max(apply(x1, 2, sd))
 
   tmp = patient.data[[pt]]$variable.regions[,-(1:5)]
-  sim[pt, 'n.var'] = nrow(patient.data[[pt]]$variable.regions[,-(1:5)])
+  sim[pt, 'n.var'] = ifelse(!is.null(tmp) && nrow(tmp)>0, nrow(patient.data[[pt]]$variable.regions[,-(1:5)]), 0)
 
-  if (nrow(tmp) <= 0) {
+  if (is.null(tmp) || nrow(tmp) <= 0) {
     sim[pt,c('gains', 'loss')] = c(0,0)
@@
 
-do.call(grid.arrange, c(sample.vals))
+#do.call(grid.arrange, c(sample.vals))
 
@@
 for (pt in sum.patient.data$Patient) {
+  print(pt)
   x1 = patient.data[[pt]]$norm.seg.vals
+  if (is.null(x1)) next
   ac = apcluster(negDistMat(x1, r=2), q=q, details=T, includeSim=T)

### 2017-07-05 Patient Report.Rmd
More analysis
...

@@
 author: "Sarah Killcoyne"
-date: "23 May 2017"
+date: "22 June 2017"
 output:
@@
   stop(paste("Missing/too many patient info file(s) in", data))
+demo.file = grep('/Demographics_full.txt', data.files, value=T)
 
-dataset = 'Training'
-patient.info = read.patient.info(patient.file, set=dataset)
+dataset = 'Test'
+patient.info = read.patient.info(patient.file, demo.file, set=dataset)
 head(patient.info)
@@
 
+sum.var = rbind('Progressors'=do.call(rbind.data.frame, var.seg.lengths[prog.pt]) %>% summarize( 'n.segments' = length(seg.len), 'median.length'=median(seg.len), 'mean.length'=mean(seg.len), 'sd'=sd(seg.len)),
+  'Non-Progressors'=do.call(rbind.data.frame, var.seg.lengths[np.pt]) %>% summarize( 'n.segments' = length(seg.len), 'median.length'=median(seg.len), 'mean.length'=mean(seg.len), 'sd'=sd(seg.len)))
+
+pander(sum.var, caption='Variable regions between progressors and non-progressors')
+
+sum.seg = rbind('Progressors'=do.call(rbind.data.frame, segment.lengths[prog.pt]) %>% summarize( 'n.segments' = length(seg.len), 'median.length'=median(seg.len), 'mean.length'=mean(seg.len), 'sd'=sd(seg.len)),
+  'Non-Progressors'=do.call(rbind.data.frame, segment.lengths[np.pt]) %>% summarize( 'n.segments' = length(seg.len), 'median.length'=median(seg.len), 'mean.length'=mean(seg.len), 'sd'=sd(seg.len)))
+
+pander(sum.var, caption='Segment stats between progressors and non-progressors')
+```
+
+```{r echo=F, message=F, warning=F}
+
+df = rbind(cbind(do.call(rbind.data.frame, var.seg.lengths[prog.pt]), 'status'='Progressor'), cbind(do.call(rbind.data.frame, var.seg.lengths[np.pt]), 'status'='Non-Progressor'))
+
+lenWT = wilcox.test(subset(df, status == 'Progressor')$seg.len, subset(df, status == 'Non-Progressor')$seg.len)
+
+ggplot(df, aes(status, seg.len)) + 
+  geom_boxplot(aes(group=status, fill=status), show.legend=F) + 
+  geom_label(data=subset(df, status == 'Progressor'), aes(status, max(seg.len)-2*sd(seg.len), label=paste('median\n',median(seg.len),sep='') )) +
+  geom_label(data=subset(df, status == 'Non-Progressor'), aes(status, max(seg.len), label=paste('median\n',median(seg.len),sep=''))) +
+  labs(y='Variable segment length', x='', title='Variable regions', caption=paste('Wilcox p-value', ifelse (lenWT$p.value < .001, '< .001', format.pval(lenWT$p.value)) ))
+
+df = rbind(cbind.data.frame('n'=sapply(var.seg.lengths[prog.pt], nrow), 'status'='Progressor'), cbind.data.frame('n'=sapply(var.seg.lengths[np.pt], nrow), 'status'='Non-Progressor'))
+
+nWT = wilcox.test(subset(df, status == 'Progressor')$n, subset(df, status == 'Non-Progressor')$n)
+
+ggplot(df, aes(status, n, group=status)) + 
+  geom_boxplot(aes(fill=status), position = 'dodge', show.legend = F) + 
+  geom_label(data=subset(df, status == 'Progressor'), aes(status, max(n)-2*sd(n), label=paste('median\n',median(n),sep='') )) +
+  geom_label(data=subset(df, status == 'Non-Progressor'), aes(status, max(n), label=paste('median\n',median(n),sep=''))) +
+  labs(y='Number of segments', title="Total number of segments", caption=paste('Wilcox p-value', ifelse (lenWT$p.value < .001, '< .001', format.pval(nWT$p.value)) ))
+
+```
+
+### Check variable segment length vs chromosome length
+
+```{r echo=F, message=F, warning=F}
 n.var.regions = sapply(var.seg.lengths, nrow)
+
 wilcox.test(n.var.regions[prog.pt], n.var.regions[np.pt])
@@
 
+## Gain vs Loss
+
+```{r echo=F, message=F, warning=F}
+
+df = do.call(rbind.data.frame, lapply(patient.data, function(df) {
+  nohgd = subset(df$info, !Pathology %in% c('HGD','IMC'))$Samplename
+  gain=0;loss=0
+  if ( !is.null(df[['variable.regions']]) ) {
+    nohgd = intersect(nohgd, colnames(df$variable.regions))
+    gain = length(which(apply(as.data.frame(df$variable.regions[,nohgd]), 1, function(x) length(which(x >= 1.1))) > 0))
+    loss = length(which(apply(as.data.frame(df$variable.regions[,nohgd]), 1, function(x) length(which(x <= 0.9))) > 0))
+  }
+
+  cbind.data.frame('gain'=gain, 'loss'=loss, 'status'=unique(df$info$Status))
+}))
+head(df)
+
+ttg = t.test(subset(df, status == 'P')$gain, subset(df, status == 'NP')$gain)
+ttl = t.test(subset(df, status == 'P')$loss, subset(df, status == 'NP')$loss)
+
+ft = fisher.test((df %>% group_by(status) %>% summarize( 'gains'=sum(gain), 'losses'=sum(loss)))[,c('gains','losses')])
+
+ggplot(melt(df), aes(status, value)) + facet_grid(.~variable) + 
+  geom_bar(aes(fill=status),stat='identity') + 
+  labs(title='Total gains & Losses excluding HGD/IMC', y='counts', x='', caption=paste('Gains and losses are signficantly enriched in Progressors vs Non.
+Fisher odds ', round(ft$estimate,2), '(p=', format.pval(ft$p.value, digits=1), ')', sep='' ))
+
+```
+
 ## How do Cancer Gene Consensus genes map to the variable regions?
@@
 ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
+ccgenes = ccgenes[-which(ccgenes$Mutation.Types == 'T'),]
 
 noLoc = grep(':-',ccgenes$Genome.Location)
+if (length(noLoc) > 0) {
   noLocGenes = ccgenes[noLoc,]
-
   ccgenes = ccgenes[-noLoc,]
+}
 
@@
 ## RAD51B is screwing up the plot simply because it shows 16 non-progressors (8 progressors) with a gain but the next closest is 9 so we'll just filter that out for now
-gene.cnt = subset(gene.cnt, gene.cnt$Gene.Symbols != 'RAD51B')
+#gene.cnt = subset(gene.cnt, gene.cnt$Gene.Symbols != 'RAD51B')
 
@@
 m$brks <- cut(m$value, include.lowest=T, right=F,ordered_result=T,
-                   breaks=c(0, 1, 3, 5, 8, 10),
-                   labels=c('0','1-3', '4-5', '6-8','9-10'))
+                   breaks=c(0, 1, 3, 5, 8, 10, 12),
+                   labels=c('0','1-3', '4-5', '6-8','9-10', '10-12'))
 
@@
 m$brks <- cut(m$value, include.lowest=T, right=F,
-                   breaks=c(0, 1, 3, 5, 8, 10),
-                   labels=c('0','1-3', '4-5', '6-8','9-10'))
+                   breaks=c(0, 1, 3, 5, 8, 10, 12),
+                   labels=c('0','1-3', '4-5', '6-8','9-10', '10-12'))
 

### 2017-07-05 RegressionPvNP.Rmd
...

@@
 author: "Sarah Killcoyne"
-date: "23 May 2017"
+date: "22 June 2017"
 output: 
   html_document: 
+    fig_height: 7
+    fig_width: 5
     toc: yes
@@
 library(glmnet)
-
-load("Test_patients.Rdata", verbose=T)
+library(CoxHD)
+library(mice)
+
+source('lib/load_patient_metadata.R')
+
+includeDemo = FALSE
+
+data = '~/Data/Ellie'
+
+data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
+analysis.files = list.files(paste(data, 'Analysis', sep='/'), full.names=T)
+
+load(grep('Test_patients.Rdata', analysis.files, value=T), verbose=T)
 validation.patient.data = patient.data
@@
 dataset = 'Training'
-load(paste(dataset, '_patients.Rdata', sep=''), verbose=T)
-
-source('lib/load_patient_metadata.R')
-
-data = '~/Data/Ellie'
-
-data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
-analysis.files = list.files(paste(data, 'Analysis', sep='/'), full.names=T)
+load(grep('Training_patients.Rdata', analysis.files, value=T), verbose=T)
 
@@
   stop(paste("Missing/too many patient info file(s) in", data))
-
-patient.info = read.patient.info(patient.file, set=dataset)
+demo.file = grep('Demographics_full.txt', data.files, value=T)
+
+patient.info = read.patient.info(patient.file, demo.file, set=dataset)
 patient.info$Patient = gsub("/", "_", patient.info$Patient)
@@
 
-validation.patient.info = read.patient.info(patient.file, set='Test')
+validation.patient.info = read.patient.info(patient.file, demo.file, set='Test')
 validation.patient.info = arrange(validation.patient.info, Status, Patient, Endoscopy.Year, Pathology)
@@
 
+#ct = sapply(patient.data, function(pd) nrow(pd$seg.vals) )
+#hist(ct)
+
+#lens = sapply(patient.data, function(pd) mean(pd$seg.vals$end.pos - pd$seg.vals$start.pos) )
+#hist(lens)
+
+#vrs = unlist(lapply(patient.data, function(pd) apply(as.matrix(pd$seg.vals[,grep('^D', colnames(pd$seg.vals))]), 2, var) ))
+#hist( vrs )
+#pd$seg.vals[,grep('^D', colnames(pd$seg.vals))] - apply(pd$seg.vals[,grep('^D', colnames(pd$seg.vals))], 2, mean)
+
 pander(sum.patient.data[,c('Patient', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Training patient set')
@@
 pander(sum.validation.info[,c('Patient', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Validation patient set')
-
 ```
@@
 
-As suggested by Moritz it is possible to merge all of the patient samples and use GLMs to model the differences, and to select regions that may be most reflective of the differences.
-In this report that's what I've done, and it worked very well.
-However, I'm a bit suspicious that it's worked too well.
-I'm looking for biases in the data, or mistakes I might have made in setting up the model.
+# Note about the data
+
+This report uses all of the genomic data as well as encoded demographic information for age at diagnosis, BMI, p53 staining status, smoking, gender, length and circumference of the Barrett's segment.
+None of these appear to improve the model.
+This may not be surprising as most of these were controlled for in generating the cohort.
 
@@
 ```{r tile, echo=F, message=F, warning=F, include=F}
-#tile.w=1e07
 tile.w=5e06
@@
 if (!dir.exists(cache.dir)) dir.create(cache.dir)
-
-#validatioMergedDf = tile.sample.segments(paste(as.character(tile.w),'validation_tiledpts.Rdata', sep=''), validation.patient.data)
-
 ```
@@
 dim(validation.df)
+
+```
+
+```{r addDemoCols, echo=T, message=F, warning=F}
+
+  ## Some sort of complexity score
+  #plot(unlist(complexity.measures), col=sum.patient.data$Status)
+  # t.test(
+  # complexity.measures[subset(sum.patient.data, Status == 'P')$Patient],
+  # complexity.measures[subset(sum.patient.data, Status == 'NP')$Patient]
+  # )
+  score.cx <-function(pd) {
+  #  print(pd$info)
+    len.var = (var(pd$seg.vals$end.pos - pd$seg.vals$start.pos))/1e12
+    extreme.values = max( apply(as.matrix(pd$seg.vals[,grep('^D', colnames(pd$seg.vals))]), 2, function(col) {
+      length(which(col >= (median(col)+2*sd(col)))) + length(which(col <= (median(col)-2*sd(col)))) 
+    }) )
+
+    (len.var + extreme.values)
+  }
+
+  complexity.measures = sapply(patient.data, score.cx)
+
+  dysplasia.df = cbind(dysplasia.df, 'cx.measure'=NA)
+  for (sample in rownames(dysplasia.df)) {
+    dysplasia.df[sample, 'cx.measure'] = complexity.measures[subset(patient.info, Samplename == sample)$Patient]
+  }
+  # norm to 1 to keep within the magnitude of the genomic data
+  cx = dysplasia.df[,'cx.measure'] 
+  dysplasia.df[,'cx.measure'] = (log(cx)/log(max(cx)))
+  #dysplasia.df[,'cx.measure'] = ((cx - min(cx))/(max(cx)-min(cx)))
+
+  complexity.measures = sapply(validation.patient.data, score.cx)
+
+  validation.df = cbind(validation.df, 'cx.measure'=NA)
+  for (sample in rownames(validation.df)) {
+    validation.df[sample, 'cx.measure'] = complexity.measures[subset(validation.patient.info, Samplename == sample)$Patient]
+  }
+  # norm to 1 to keep within the magnitude of the genomic data
+  cx = validation.df[,'cx.measure'] 
+  validation.df[,'cx.measure'] = (log(cx)/log(max(cx)))
+
+if (includeDemo) {
+
+  ## Add in demographics
+
+  # sort by samples in the order of the matrix
+  patient.info = patient.info[match(rownames(dysplasia.df), patient.info$Samplename),]
+
+  patient.info[which(patient.info$BMI > 100), 'BMI'] = NA # I think this one is wrong
+
+  demo.data = patient.info[,c('Sex','Circumference','Maximal', 'Age.at.diagnosis', 'BMI', 'p53.Status', 'Smoking')]
+  demo.data$Sex = as.integer(demo.data$Sex)-1
+  demo.data$Smoking = as.integer(demo.data$Smoking)-1
+  demo.data$p53.Status = as.integer(demo.data$p53.Status)-1
+  demo.cols = c('Sex','C','M','Age', 'BMI', 'p53.Status', 'Smoking')
+
+  encode.age <- function(x) {
+    if (x > 70) { 
+      (x - 70)/6 
+    } else if (x < 50) {
+      (x -50)/6
+    } else {
+      0
+    }
+  }
+  #demo.data$Age.at.diagnosis = scale(demo.data$Age.at.diagnosis)
+
+  # Encode based on presumed distributions or unit normalize
+  demo.data = demo.data %>% rowwise() %>% mutate( 'age.encoded'= encode.age(Age.at.diagnosis))
+  demo.data$Age.at.diagnosis = demo.data$age.encoded
+  demo.data$age.encoded = NULL
+  demo.data$BMI = scale(demo.data$BMI)
+
+  demo.data[,c('Circumference','Maximal')] = apply(demo.data[,c('Circumference','Maximal')], 2, function(x) (x-mean(x, na.rm=T))/sd(x, na.rm=T))
+
+  ## Impute missing data
+  imp = mice(as.matrix(demo.data), diagnostics = F)
+
+  dysplasia.df = cbind(dysplasia.df, 'Sex'=NA,'C'=NA,'M'=NA,'Age'=NA, 'BMI'=NA, 'p53.Status'=NA, 'Smoking'=NA)
+  dysplasia.df[,demo.cols] = as.matrix(complete(imp))
+
+}
 ```
@@
 `r pander(dysplasia.df[1:5, 1:5], caption=paste("dimensions:", paste(dim(dysplasia.df), collapse=', ')), justify='left')`
-
-```{r cvglmnetPNP, echo=F, warning=F, message=F, eval=F}
-## cv.glmnet Ridge vs Lasso
-#With all `r ncol(dysplasia.df)` regions of the genome, first check if/which regression may be appropriate.
-It appears that pure lasso may provide the best balance for number of features with non-zero coefficients.
-
-plot.multi.cv.glmnet<-function(x,y,alpha=c(0,1), family="binomial", title='cv.glmnet') {
-  plots <- list(  )
-  cv.models = data.frame(matrix(ncol=3,nrow=0,dimnames=list(c(), c('class','lambda','cvm'))))
-  for (a in alpha) {
-    cv = cv.glmnet(x, y, alpha=a, family=family)
-    cv.models = rbind(cv.models, cbind('class'=paste('cv',a,sep=''), 'lambda'=log(cv$lambda), 'cvm'=cv$cvm))
-    plots[[as.character(a)]] = autoplot(cv, main=paste('alpha',a), ylab=cv$name)
-  }
-  cv.models[c('lambda','cvm')] = lapply(cv.models[c('lambda','cvm')], function(x) as.numeric(as.character(x)))
-
-  gg = ggplot(cv.models, aes(x=lambda, y=cvm, color=class)) + geom_point() + geom_line() + 
-    labs(x='log(Lambda)', y='Binomial Deviance', title=title)
-
-  plots[['all']] = gg
-
-  return(plots)
-}
-
-gp = plot.multi.cv.glmnet(x=dysplasia.df, y=labels, alpha=c(0,0.5,0.8,1), title='cv.glmnet P vs NP')
-
-cv1 = cv.glmnet(dysplasia.df, labels, alpha=1, family='binomial')
-gg = autoplot(cv1$glmnet.fit, xvar='lambda', main='alpha=1 glmnet.fit from cv.glmnet') + theme(legend.position='none')
-
-gp[['coef.1']] = gg
-
-do.call(grid.arrange, c(gp, ncol=2, nrow=3))
-
-```
 
@@
   }
+
   return(patients)
@@
 
-crossvalidate.by.patient<-function(x,y,lambda,pts,a=1,nfolds=10, splits=5, fit=NULL, minR=0.2, select='deviance') {
+crossvalidate.by.patient<-function(x,y,lambda,pts,a=1,nfolds=10, splits=5, fit=NULL, minR=0.2, select='deviance', ...) {
   if (nfolds > 5) minR = 0.1
@@
   fit.e = list()
+
+  if (ncol(pts) == (nfolds+2)) {
+    tpts = pts
+  } else {
     tpts = create.patient.sets(pts, nfolds, splits, minR)
+  }
+
   cv.pred = (matrix(nrow=0, ncol=length(lambda)))
@@
       # pre-spec lambda seq
-      fitCV <- glmnet(training, y[-test.rows], lambda=lambda, alpha=a, family='binomial') 
+      fitCV <- glmnet(training, y[-test.rows], lambda=lambda, alpha=a, family='binomial', ...) 
       # autoplot(fit) + theme(legend.position="none")
@@
 
-```{r xvalpt, message=F, warning=F, echo=T, fig.height=16, fig.width=16}
+```{r xvalpt, message=F, warning=F, echo=F, fig.height=16, fig.width=16}
 nl = 1000
@@
 alpha.values = c(0,0.5,0.7,0.8,0.9,1)
+
+file = paste(cache.dir, 'patient_folds.tsv', sep='/')
+if (file.exists(file)) {
+  sets = read.table(file, header=T, sep='\t')
+} else {
+  sets = create.patient.sets(pts, folds, splits, 0.2)  
+  write.table(sets, quote=F, sep='\t', row.names=F, file=file)
+}
 
@@
   for (a in alpha.values) {
-    fit0 <- glmnet(dysplasia.df, labels, alpha=a, nlambda=nl, family='binomial') # all patients
+    fit0 <- glmnet(dysplasia.df, labels, alpha=a, nlambda=nl, family='binomial', standardize=T) # all patients
     #autoplot(fit0, xvar='lambda', main=paste('fit0, all samples, alpha=',a,sep='')) + theme(legend.position='none')
@@
     }
-    cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=l, pts=pts, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance')
+    cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=l, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', standardize=T)
 
@@
   }
-  save(plots, coefs, performance.at.1se, file=file)
+  save(plots, coefs, performance.at.1se, dysplasia.df, file=file)
 }
@@
 
-```{r xvalpt2, message=F, warning=F, echo=T}
-select.alpha = as.character(0.9)
-
-performance.at.1se = do.call(rbind, performance.at.1se)
-performance.at.1se$alpha = rownames(performance.at.1se)
-
-performance.at.1se = cbind(performance.at.1se, do.call(rbind, lapply(coefs, function(x) cbind('n.Coef'=nrow(x), '75%'=length(which(rowSums(x)/50 >= 0.75))))))
-
-ggplot(performance.at.1se, aes(alpha, mean)) + geom_text( aes(label=round(mean, 3)), hjust=-0.5) +
+```{r xvalpt2, message=F, warning=F, echo=T, fig.width=8, fig.height=6}
+
+performance = do.call(rbind.data.frame, performance.at.1se)
+performance$alpha = rownames(performance)
+
+## check how often the feature(s) is selected at that lamda in each split. "stability selection"
+coef.stable = lapply( coefs, function(cf) {
+  sort(rowSums(cf[,-1]), decreasing=T)
+})
+coef.stable = lapply(coef.stable, function(cf) cf/(folds*splits))
+
+performance = cbind(performance, do.call(rbind,  lapply(coef.stable, function(x) {
+  cbind('n.Coef'=length(x), '25%'=length(which(x>=.25)), 
+        '50%'=length(which(x>=.5)), '75%'=length(which(x>=.75 )), 
+        '100%'=length(which(x==1)) )
+})))
+
+ggplot(performance, aes(alpha, mean)) + geom_text( aes(label=round(mean, 3)), hjust=-0.5) +
   geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme), color='grey') + geom_point() + ylim(0.5,0.8) +
@@
   geom_text(aes(label=paste('(',`75%`,')',sep='')), nudge_y=0.02) +
-  labs(x='Elasticnet penalty value', y='Model performance at lambda.1se', 
+  labs(x='Elasticnet penalty value: ridge <-> lasso', y='Model classification at lambda.1se', 
        title=paste('All samples,', folds, 'folds,', splits, 'patient splits'))
-
-## check how often the feature(s) is selected at that lamda in each split. "stability selection"
-coef.stable = lapply( coefs[c('0.8','0.9','1')], function(cf) {
-  sort(rowSums(cf[,-1]), decreasing=T)
-})
 
@@
 
-coef.stable = lapply(coef.stable, function(cf) cf/(folds*splits))
-
-stability = do.call(rbind, lapply(coef.stable, function(cf) {
-  cbind('>=75%'=length(which(cf >= 0.75)),  '>=50%'=length(which(cf >= 0.5)),  '>=25%'=length(which(cf >= 0.25)), 'Total'=length(cf))
-}))
-rownames(stability) = names(coef.stable)
-
-pander(stability, justify='left', caption='Number of features stable in n% or more folds')
+pander(performance[,c('n.Coef','25%','50%','75%', '100%')], justify='left', caption='Number of features stable in n% or more folds')
+
+select.alpha = as.character(0.9)
+
+# Genomic regions only
+most.stable = grep('^\\d+', names(which(coef.stable[[select.alpha]] >= 0.75 )), value=T)
+
+most.stable = cbind.data.frame(do.call(rbind, strsplit(most.stable, ':|-')), most.stable)
+colnames(most.stable) = c('chr','start','stop', 'name')
+
+most.stable$chr = as.integer(as.character(most.stable$chr))
+most.stable = arrange(most.stable, chr, start)
+
 ```
@@
 
-### Validation Data
-
-TBD
-
-```{r, message=F, warning=F, echo=F, eval=F}
-
-fitV <- glmnet(dysplasia.df, labels, alpha=select.alpha, nlambda=nl, family='binomial') # all patients
-#autoplot(fitV, xvar='lambda', main=paste('fit0, all samples, alpha=',a,sep='')) + theme(legend.position='none') 
-cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=fitV$lambda, pts=pts, a=select.alpha, nfolds=folds, splits=splits, fit=fitV)
-lambda.opt = ifelse( length(cv.patient$lambda.1se) > 0, cv.patient$lambda.1se, cv.patient$lambda.min )
-
-pred = predict(fitV, newx=validation.df, s=lambda.opt, type='response')
-
-statusV = unlist(sapply(validation.patient.data, function(df) {
-  df$info = arrange(df$info, Endoscopy.Year, Pathology)
-  label = as.integer(df$info$Status == 'P') #as.integer(df$info$Pathology %in% c('HGD', 'IMC'))
-  names(label) = df$info$Samplename
-  return(label)
-}))
-names(statusV) = sub('.*\\.', '',  names(statusV))
-
-# Progressors - by sample
-length( which(pred[ names(which(statusV == 1)), ] > 0.5) )/length(which(statusV == 1))
-
-# Non Progressors - by sample
-length( which(pred[ names(which(statusV == 0)), ] < 0.5) )/length(which(statusV == 0))
-
-```
+`r pander(most.stable[,c('chr','start','stop')], caption='Features identified as stable in 75% or more folds.')`
 
@@
 While there are fewer features for higher values of alpha, the features are the same across models.")`
-
-`r length(unique(unlist(lapply(coef.stable, function(cf)  names(which(cf > 0.8)) ))))` features are shared at all values of alpha and were non-zero in at least 80% of the folds.
 
@@
 plot(coefs[[select.alpha]][,1])
+selected = coefs[[select.alpha]]
+
+selected[which(
+  selected[,1] > mean(selected[,1]) + sd(selected[,1]) | 
+  selected[,1] < mean(selected[,1] - sd(selected[,1]))  
+),]
+
+## Variance in the selected features?  Not so much
+coefs[[select.alpha]][,1]
+
+v = vector(length(colnames(dysplasia.df)), mode='numeric')
+names(v) = colnames(dysplasia.df)
+v[rownames(selected)] = coefs[[select.alpha]][,1]
+
+vv <- apply(dysplasia.df, 2, var)
+plot(v, vv, log='y', col=(v!=0)+1)
+
+boxplot(vv ~ (v==0), log='y', ylim=c(0.0001,1))
+
+qqplot(vv[v==0], vv[v!=0], log='xy')
 
@@
 
-Using the features from the alpha=0.9 models
+Using the features from the alpha=`r select.alpha` models
 
@@
 
-features.in.prog = dysplasia.df[names(which(labels == 1)),topF]
-features.nonprog = dysplasia.df[names(which(labels == 0)),topF]
+features.in.prog = dysplasia.df[names(which(labels == 1)), grep('^\\d+:', topF, value=T)]
+features.nonprog = dysplasia.df[names(which(labels == 0)), grep('^\\d+:', topF, value=T)]
 
@@
 # Gain or loss not correlated in prog or non
-
+genomeCols = grep('^\\d+:', colnames(dysplasia.df), value=T)
 gl = cbind.data.frame(
@@
 allgl = cbind.data.frame(
-  'ALL.P.gain' = apply(dysplasia.df[names(which(labels == 1)),], 2, function(x) sum( as.integer(x >= 1.1) ) ), 
-  'ALL.P.loss' = apply(dysplasia.df[names(which(labels == 1)),], 2, function(x) sum( as.integer(x <= 0.9) ) ),
-  'ALL.NP.gain' = apply(dysplasia.df[names(which(labels == 0)),], 2, function(x) sum( as.integer(x >= 1.1) ) ), 
-  'ALL.NP.loss' = apply(dysplasia.df[names(which(labels == 0)),], 2, function(x) sum( as.integer(x <= 0.9) ) ) )
+  'ALL.P.gain' = apply(dysplasia.df[names(which(labels == 1)), genomeCols], 2, function(x) sum( as.integer(x >= 1.1) ) ), 
+  'ALL.P.loss' = apply(dysplasia.df[names(which(labels == 1)),genomeCols], 2, function(x) sum( as.integer(x <= 0.9) ) ),
+  'ALL.NP.gain' = apply(dysplasia.df[names(which(labels == 0)),genomeCols], 2, function(x) sum( as.integer(x >= 1.1) ) ), 
+  'ALL.NP.loss' = apply(dysplasia.df[names(which(labels == 0)),genomeCols], 2, function(x) sum( as.integer(x <= 0.9) ) ) )
 
@@
 
-topFSegments = do.call(rbind,  strsplit( topF, ':|-' ))
+topFSegments = do.call(rbind,  strsplit( topF[grep('^\\d+:', topF)], ':|-' ))
 colnames(topFSegments) = c('chr','start','end')
@@
   scale_fill_discrete(name='Length adjusted information content') + theme(legend.position = 'bottom'), 
-  bottom='Ratio of base pair coverage per chromosome in the top non-zero features selected from the model.\nChromosomes are sorted and binned by their gene (e.g.
+  bottom='Ratio of base pair coverage per chromosome in the top non-zero features\n selected from the model.
+Chromosomes are sorted\nand binned by their gene (e.g.
 information) content rather than by overall length.')
@@
 ```{r, warning=F, message=F}
+bm = NULL
+tryCatch({
   mart = biomaRt::useMart("ENSEMBL_MART_ENSEMBL", host="grch37.ensembl.org")
@@
   bm = getBM(mart=biomart, attributes=attr, 
-           filters=list('chromosomal_region'=sub('-', ':',topF), "biotype"=c("IG_C_gene","IG_D_gene","IG_J_gene","IG_V_gene","protein_coding","TR_C_gene","TR_D_gene","TR_J_gene","TR_V_gene")))
+           filters=list('chromosomal_region'=sub('-', ':',topF[grep('^\\d+:', topF)]), "biotype"=c("IG_C_gene","IG_D_gene","IG_J_gene","IG_V_gene","protein_coding","TR_C_gene","TR_D_gene","TR_J_gene","TR_V_gene")))
 #head(bm)
@@
   pander(with(topFSegments, cor.test(genes, (end-start))), caption="Number of genes per segment is not related to the segment length.")
-
+}, error = function(e) {
+  warning(paste("Failed to connect to Biomart service", e))
+})
+```
+
+### Hazards
+
+Using the model for alpha=`r select.alpha`.
+
+```{r echo=F, warning=F, message=F,eval=T, fig.height=8, fig.width=8}
+hazards = apply( exp(t(dysplasia.df[, rownames(coefs[[select.alpha]])]) *  coefs[[select.alpha]][,1]), 1, function(x) {
+  sd(x)/mean(x)
+})
+ch = as.data.frame(hazards)
+ch$coef = coefs[[select.alpha]][,1]
+ch$label = rownames(ch)
+library(ggrepel)
+ggplot(ch, aes(coef, hazards, col=hazards > quantile(hazards)[['75%']])) + 
+  #xlim(min(coefGR$coef)-5, max(coefGR$coef)+10) +
+  geom_point(alpha=0.5) + 
+  geom_text_repel(aes(x=coef+10, label=ifelse(hazards > quantile(hazards)[['75%']], label, '')),show.legend=F) +
+  labs(title='Top features coef vs hazard') + theme(legend.position = 'bottom')
+```
+
+```{r echo=F, warning=F, message=F,eval=T, fig.height=12, fig.width=8}
+ch = arrange(ch, -hazards)
+
+ch$hazards = round(ch$hazards, 2)
+ch$coef = round(ch$coef, 2)
+
+melt.by.reg<-function(x) {
+  r = x[['label']]
+
+  reg = cbind.data.frame( 'value'=dysplasia.df[ , r ], labels)
+  reg$labels = factor(reg$labels)
+  reg$labels = revalue(reg$labels, c("1"="P", "0"="NP"))
+
+  tt = wilcox.test(dysplasia.df[ which(labels == 0) , r ], dysplasia.df[ which(labels == 1) , r ])
+
+  reg$pval = tt$p.value
+  reg$region = r
+  reg$hazard = x[['hazards']]
+  reg$coef = x[['coef']]
+  return(reg)
+}
+
+plot.reg<-function(x) {
+  reg = melt.by.reg(x)
+  pv = ifelse(tt$p.value < .001, '< .001', format.pval(unique(reg$p.value)))
+
+#  ggplot(reg, aes(value, fill=labels) ) + geom_histogram(bins=50, color='grey') + 
+#    labs(x='seg. value', title=r, subtitle=paste('haz=',x[['hazards']], ' coef=', x[['coef']], ' MW pval: ',pv,sep=''))
+  ggplot(reg, aes(x=labels, y=value, group=labels, fill=labels) ) + geom_jitter(color='grey', show.legend = F) + geom_violin(show.legend = F, alpha=0.5) +  ylim(0.75, 1.2) +
+    geom_hline(yintercept = median(reg$value) + sd(reg$value)) + geom_hline(yintercept = median(reg$value) - sd(reg$value)) +
+    labs(x='', y='seg. value', title=r, subtitle=paste('haz=',x[['hazards']], ' coef=',x[['coef']], '\nMW pval ',pv,sep=''))
+}
+
+regions = apply(ch, 1, melt.by.reg)
+
+do.call(grid.arrange, c(lapply(regions, function(reg) {
+  se = 2
+  x = rbind.data.frame('P'= cbind( 'gain' = length(which(with(subset(reg, labels == 'P'), value >= median(value) + sd(value)*se))),
+                               'loss' = length(which(with(subset(reg, labels == 'P'), value <= median(value) - sd(value)*se))) ),
+
+                  'NP' = cbind( 'gain' = length(which(with(subset(reg, labels == 'NP'), value >= median(value) + sd(value)*se))),
+                                'loss' = length(which(with(subset(reg, labels == 'NP'), value <= median(value) - sd(value)*se))) ) )
+  # x = rbind.data.frame('P'= cbind( 'high' = length(which(with(subset(reg, labels == 'P'), value >= 1.1))),
+  #                                  'low' = length(which(with(subset(reg, labels == 'P'), value <= 0.9))) ),
+  # 
+  #                 'NP'= cbind( 'high' = length(which(with(subset(reg, labels == 'NP'), value >= 1.1))),
+  #                                  'low' = length(which(with(subset(reg, labels == 'NP'), value <= 0.9))) ))
+
+  cols = c('indianred3','slateblue2')
+  if ( as.numeric(reg$coef[1]) < 0) cols = c('plum', 'skyblue3')
+  #ggplot(melt(as.matrix(x)), aes(Var1, value, fill=Var2) ) + geom_bar(stat='identity', position='dodge') + labs(x='', y='', title=unique(reg$region)[1]) + scale_fill_manual(values=cols)
+
+  ggplot(melt(as.matrix(rowSums(x))), aes(Var1, value, fill=Var1)) + geom_bar(stat='identity', show.legend = F)+ labs(x='', y='', title=unique(reg$region)[1]) + scale_fill_manual(values=cols)
+
+})))
+
+## Individual hazards vs CN values?  exp(coef) -- odds ratio
+# plot( exp( coefs[[select.alpha]][ch$label[1],1]  * dysplasia.df[,ch$label[1]]),  dysplasia.df[,ch$label[1]] )
+# 
+# plot(  exp( t(dysplasia.df[, ch$label[1]]) *  coefs[[select.alpha]][ch$label[1],1]  ), dysplasia.df[,ch$label[1]]  ) ## pos coef
+# plot(exp(t(dysplasia.df[, ch$label[3]]) *  coefs[[select.alpha]][ch$label[3],1]), dysplasia.df[,ch$label[3]]) ## pos coef
+# 
+# plot(exp(t(dysplasia.df[, ch$label[2]]) *  coefs[[select.alpha]][ch$label[2],1]), dysplasia.df[,ch$label[2]]) ## neg coef
+# plot(exp(t(dysplasia.df[, ch$label[4]]) *  coefs[[select.alpha]][ch$label[4],1]), dysplasia.df[,ch$label[4]]) ## neg coef
+# 
+# 
+# labels[which( dysplasia.df[, ch$label[2]] <= 0.9)]
+# 
+# labels[which( dysplasia.df[, ch$label[1]] >= 1.1)]
+
+#do.call(grid.arrange, c(apply(ch[1:10,],1, plot.reg), ncol=2))
+
+#do.call(grid.arrange, c(apply(ch[11:20,],1, plot.reg), ncol=2))
+
+#do.call(grid.arrange, c(apply(ch[21:30,],1, plot.reg), ncol=2))
+
+```
+
+```{r}
+
+cols = intersect(rownames(ch), grep('\\d+:', colnames(dysplasia.df), invert=T, value=T))
+if (length(cols) > 0)
+  pander(ch[cols,1:2], caption='Non-zero demographic hazard ratios')
+```
+
+`r pander(ch[order(-ch$hazards),][1:10,1:2], caption='Top 10 features by hazard ratio.')`
+
+```{r echo=F, warning=F, message=F}
+if (!is.null(bm)) {
   topFRegions = merge(topFRegions, (bm %>% group_by(chromosome_name) %>% summarise('gene.cnt'=length(unique(hgnc_symbol)), 'exon.cnt'=sum(exon_count))), by.x='chr', by.y='chromosome_name', all.x=T)
-
-#with(topFRegions, cor.test(protein_coding, exon.cnt))
-#with(topFRegions, cor.test(gene.cnt, exon.cnt)) # expected
-
-# Exons correlate with the adjusted information content measure, expected
-#with(topFRegions, cor.test(exon.cnt, adj.info.content))
-
+  colnames(topFRegions) = sub('\\.(x|y)', '', colnames(topFRegions)  )
+
+  bmGR = makeGRangesFromDataFrame(bm, start.field = 'start_position', end.field = 'end_position', keep.extra.columns = T)
+
+  # Merge COSMIC genes
+  ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
+  ccgenes = ccgenes[-grep('X,Y', ccgenes$Genome.Location),]
+
+  ccgenes = ccgenes[-which(ccgenes$Mutation.Types == 'T'),] # drop the translocations
+  #nrow(ccgenes)
+
+  bmGR$COSMIC = bmGR$hgnc_symbol %in% ccgenes$Gene.Symbol
+
+  grcfs = grep('^\\d+:', rownames(coefs[[select.alpha]]), value=T)
+  topF = do.call(rbind.data.frame, strsplit(grcfs , ':|-'))
+
+  colnames(topF) = c('chr','start','end')
+  coefGR = makeGRangesFromDataFrame(topF)
+  coefGR$coef = coefs[[select.alpha]][grcfs,1]
+
+  coefGR$hazard = hazards[grcfs]
+  coefGR$gene = ''
+  coefGR$ensembl_ids = ''
+  coefGR$COSMIC = ''
+  coefGR$COSMIC_ensembl_ids = ''
+
+  ov = findOverlaps(bmGR, coefGR)
+  for (hit in unique(subjectHits(ov))) {
+    tmp = bmGR[ queryHits(ov[which(subjectHits(ov) == hit)]) ]
+
+    coefGR[hit]$COSMIC = paste(tmp[which(tmp$COSMIC)]$hgnc_symbol, collapse=', ')
+    coefGR[hit]$COSMIC_ensembl_ids = paste(tmp[which(tmp$COSMIC)]$ensembl_gene_id, collapse=', ')
+
+    coefGR[hit]$gene = paste(tmp$hgnc_symbol, collapse=', ')
+    coefGR[hit]$ensembl_ids = paste(tmp$ensembl_gene_id, collapse=', ')
+  }
+
+  coefGR = coefGR[order(coefGR$hazard)]
+  highest = coefGR[coefGR$hazard > quantile(coefGR$hazard)[['75%']]]
+
+  write.table(unlist(strsplit(unique(highest$ensembl_ids), ', ')), quote=F, row.names = F, col.names = F, file='highest_coef_genes.txt')
+
+  topFRegions$topCoefHaz = 0
+  topFRegions$topCoefHaz[1:22] = table(seqnames(highest))
   grid.arrange(ggplot(topFRegions, aes(reorder(chr, -adj.info.content), gene.cnt/protein_coding, fill=ai.cut)) + geom_bar(stat='identity') +
     labs(y='Gene Ratio', x='chr', title='Feature genes vs all protein coding genes') +
+    geom_text(aes(label=ifelse(topCoefHaz > 0, '*', ''), color=ai.cut), nudge_y=0.01, show.legend=F, size=10) +
     geom_text(aes(label=round(gene.cnt/protein_coding, 2)), vjust=-0.5) +
@@
 
-# Merge COSMIC genes
-ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
-ccgenes = ccgenes[-grep('X,Y', ccgenes$Genome.Location),]
-
-# This is going to strip out some genes that don't appear in ensembl
-attr = c('ensembl_gene_id','hgnc_symbol','start_position','end_position', 'chromosome_name')
-ccbm = getBM(mart=biomart, attributes=attr, filters=list('external_gene_name'=ccgenes$Gene.Symbol, 'chromosome_name'=c(1:22,'X','Y')))
-ccbm = ccbm[-which(ccbm$hgnc_symbol == ""),]
-ccbm$chromosome_name = factor(ccbm$chromosome_name, levels=chromInfo$chr)
-
-topFRegions = merge(topFRegions, (ccbm %>% group_by(chromosome_name) %>% summarise('COSMIC.all'=length(unique(hgnc_symbol)))), by.x='chr',by.y='chromosome_name', all.x=T)
-
-containedCC = subset(bm, hgnc_symbol %in% ccbm$hgnc_symbol)
-
-write.table(containedCC$ensembl_gene_id, quote=F, row.names = F, col.names = F, file='feature_cosmic.txt')
-
-topFRegions = merge(topFRegions, (containedCC %>% group_by(chromosome_name) %>% summarise('COSMIC.cont'=length(unique(hgnc_symbol)))), by.x='chr',by.y='chromosome_name', all.x=T)
-
-grid.arrange(ggplot(topFRegions, aes(reorder(chr, -adj.info.content), COSMIC.cont/COSMIC.all, fill=ai.cut)) + geom_bar(stat='identity') +
-  labs(y='COSMIC Gene Ratio', x='chr', title='COSMIC genes: Feature vs all')+
-  geom_text(aes(label=round(COSMIC.cont/COSMIC.all, 2)), vjust=-0.5) +
-  scale_fill_discrete(name='Length adjusted information content') + theme(legend.position = 'bottom'), 
-  bottom='Again, but this time looking only at COSMIC genes, the ratio of COSMIC genes\nwithin the selected feature vs all COSMIC genes on that chromosome.\nChromosomes are ordered and colored by information content.')
-
-colSums(topFRegions[,c('COSMIC.all','COSMIC.cont')],na.rm=T)
-colSums(topFRegions[,c('protein_coding','gene.cnt')],na.rm=T)
-
-#table(topFRegions$ai.cut)
-
-```
-
-However, the `r round(bpCovered, 2)*100`% of bases appear to include `r round(length(unique(bm$hgnc_symbol))/sum(chromInfo$protein_coding), 2)*100`% (`r length(unique(bm$hgnc_symbol))`) of known protein-coding genes.
-Similarly, `r round(length(unique(containedCC$hgnc_symbol))/length(ccgenes$Gene.Symbol),2)*100`% of the `r length(ccgenes$Gene.Symbol)` COSMIC genes are contained within these regions.
-
-This appears to be a fairly large enrichment.
+  write.table(unlist(strsplit(unique(coefGR$COSMIC_ensembl_ids), ',')), quote=F, row.names = F, col.names = F, file='feature_cosmic.txt')
+}
+#However, the `r round(bpCovered, 2)*100`% of bases appear to include `r round(length(unique(bm$hgnc_symbol))/sum(chromInfo$protein_coding), 2)*100`% (`r length(unique(bm$hgnc_symbol))`) of known protein-coding genes.
+Similarly, `r round(length(unlist(strsplit(unique(coefGR$COSMIC), ',')))/length(ccgenes$Gene.Symbol),2)*100`% of the `r length(ccgenes$Gene.Symbol)` COSMIC genes are contained within these regions.
+
+```
 
@@
   }
-  #hist(colSums(gain.df))
-  #hist(colSums(loss.df))
-  #hist(colSums(all.binary))
-
-  gp = run.cv(x=as.matrix(gain.df), y=labels, pts=pts, nfolds=folds, splits=splits, alphas=c(0,0.5,0.8,1))
-  lp = run.cv(x=as.matrix(loss.df), y=labels, pts=pts, nfolds=folds, splits=splits, alphas=c(0,0.5,0.8,1))
+
+  gp = run.cv(x=as.matrix(gain.df), y=labels, pts=pts, nfolds=folds, splits=splits, alphas=c(0.1,0.5,0.8,1))
+  lp = run.cv(x=as.matrix(loss.df), y=labels, pts=pts, nfolds=folds, splits=splits, alphas=c(0.1,0.5,0.8,1))
 
@@
 
-  both = run.cv(x=as.matrix(gl.df), y=labels, pts=pts, nfolds=folds, splits=splits, alphas=c(0,0.5,0.8,1))
-
-  save(gp, lp, all, file=file)
+  both = run.cv(x=as.matrix(gl.df), y=labels, pts=pts, nfolds=folds, splits=splits, alphas=c(0.1,0.5,0.8,1))
+
+  save(gp, lp, both, file=file)
 }
@@
 
-```{r, echo=T, warning=F, message=F, fig.height=10, fig.width=10}
+```{r, echo=T, warning=F, message=F, fig.height=12, fig.width=12}
 
@@
     # Fit generated on all samples, including HGD
-    a = 1
+    a = select.alpha
     fitLOO <- glmnet(training, labels[train.rows], alpha=a, family='binomial', nlambda=nl) # all patients
@@
 
+### Prediction ROC curves
+
+Based on these curves we get our best AUC at ~0.36.  Currently we are using 0.5 by default.
+
+```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=10, fig.width=6}
+preds = unlist(lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], function(df) df$Prediction))
+np.preds = unlist(lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Patient], function(df) df$Prediction))
+
+roc50 = pROC::roc(as.integer(c(preds, np.preds) > 0.5), 
+                c(rep(1, length(preds)), rep(0, length(np.preds))),
+                main='Progressors & Non-progressors, cutoff 0.5')
+
+df = cbind('Specificity'=rev(roc50$specificities), 'Sensitivity'=rev(roc50$sensitivities))
+g1 = ggplot(df, aes(Specificity, Sensitivity)) + geom_line() +
+  scale_x_reverse() + 
+  geom_label(data=t(as.data.frame(round(pROC::coords(roc50, "best"),2))), aes(x=specificity, y=sensitivity, label=paste('prob=',threshold, ' (', specificity*100, '%, ', sensitivity*100,'%)\nAUC:', sprintf("%.2f",roc50$auc*100), '%', sep='')), nudge_x=.25) +
+  labs(title='Probability threshold = 0.5', x='Specificity (FPR)', y='Sensitivity (TPR)')
+
+preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
+roc = pROC::roc(Status ~ Prediction, data=preds, ci=T, of='thresholds')
+
+df = cbind('Specificity'=rev(roc$specificities), 'Sensitivity'=rev(roc$sensitivities))
+g2 = ggplot(df, aes(Specificity, Sensitivity)) + geom_line() +
+  scale_x_reverse() + 
+  geom_label(data=t(as.data.frame(round(pROC::coords(roc, "best"),2))), aes(x=specificity, y=sensitivity, label=paste('prob=',threshold, ' (', specificity*100, '%, ', sensitivity*100,'%)\nAUC:', sprintf("%.2f",roc$auc*100), '%', sep='')), nudge_x=.25) +
+  labs(title='Probability threshold calculated by ROC', x='Specificity (FPR)', y='Sensitivity (TPR)')
+
+grid.arrange(g1, g2, nrow=2, top="ROC - Progressors vs Non")
+
+cutoff = pROC::coords(roc, 'best')[['threshold']]
+cutoff = 0.5
+```
+
+This argues that we should set the prediction probability cutoff to `r pROC::coords(roc, 'best')[['threshold']]`.
+However, for now the predictions from here on are based on the default `r cutoff` value.
+
+***
+
+__All predictions from here onwards are based on the `r cutoff` cutoff, not the ROC curve__
+
+***
+
 ## How do the predictions look at each time point?
@@
   hgd = subset(pt, Pathology %in% c('HGD', 'IMC'))
-  sum(as.integer(hgd$Prediction > 0.5))/nrow(hgd)
+  sum(as.integer(hgd$Prediction > cutoff))/nrow(hgd)
 })
@@
   hgd = subset(pt, Pathology %in% c('HGD', 'IMC'))
-  cbind.data.frame('correctly.predicted'=sum(as.integer(hgd$Prediction > 0.5)), 'total.hgd.samples'=nrow(hgd))
+  cbind.data.frame('correctly.predicted'=sum(as.integer(hgd$Prediction > cutoff)), 'total.hgd.samples'=nrow(hgd))
 }))
@@
 
+global.by.path = do.call(rbind.data.frame,
+                         lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], function(pt) {
+  final.endo = subset(pt, Endoscopy.Year == max(Endoscopy.Year))
+  final.endo = final.endo[nrow(final.endo),]
+
+  #pred <- function(rows) length(which(rows$Prediction > 0.5))/nrow(rows)
+  pred <- function(rows) length(which(rows$Prediction > cutoff))
+
+  h = subset(pt, Pathology %in% c('IMC','HGD'))
+  l = subset(pt, Pathology == 'LGD' & PID != final.endo$PID)
+  id = subset(pt, Pathology == 'ID' & PID != final.endo$PID)
+  be = subset(pt, Pathology %in% c('BE') & PID != final.endo$PID)
+
+  cbind('HGD'=pred(h), 'total.h'=nrow(h), 
+        'LGD'=pred(l), 'total.l'=nrow(l), 
+        'ID'=pred(id), 'total.id'=nrow(id),
+        'NDBE'=pred(be), 'total.be'=nrow(be))
+})
+)
+
+global.by.path = colSums(global.by.path)
+global.path = as.data.frame(matrix(global.by.path, byrow=T, nrow=4, dimnames=list(c('HGD','LGD','ID','NDBE'),c('pred', 'total'))))
+global.path$ratio = (global.path$pred/global.path$total)*100
+global.path$path = rownames(global.path)
+
+global.path$path = factor(global.path$path, levels=c('HGD','LGD','ID','NDBE'))
+
+ggplot(melt(global.path[,c('ratio','total','path')], measure.vars='ratio', id.vars=c('path', 'total')), aes(path, value, group=path)) + 
+  geom_bar(stat='identity', fill='dodgerblue3') + 
+  geom_text(aes(y=value+2,label=paste(round(value, 1), '%', sep=''))) +
+  geom_text(aes(y=3,label=paste('n=', total, sep=''))) +
+  ylim(0,100) + labs(title='Samples predicted by pathology', x='Pathology', y='Percent identified')
+```
+
+```{r echo=F, message=F, warning=F}
 lgd = do.call(rbind, lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], function(pt) {
   e = subset(pt, Pathology  == 'LGD')
-  cbind.data.frame('predicted'=sum(as.integer(e$Prediction > 0.5)), 'total.lgd'=nrow(e))
+  cbind.data.frame('predicted'=sum(as.integer(e$Prediction > cutoff)), 'total.lgd'=nrow(e))
 }))
@@
   coord_flip() + labs(title='Correctly predicted LGD samples', y='Correct prediction ratio')
-
+```
+
+```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=10}
 # All samples
+
 all.predictions = sapply(pg.samp, function(pt) {
   if (unique(pt$Status) == 'NP') {
-    as.integer(pt$Prediction < 0.5)
+    as.integer(pt$Prediction < cutoff)
   } else {
-    as.integer(pt$Prediction > 0.5)
+    as.integer(pt$Prediction > cutoff)
   }
@@
 
+p = subset(all.predictions, Status == 'P')
+np = subset(all.predictions, Status == 'NP')
+
 m = melt(all.predictions, id.vars=c('Patient','Status'), measure.vars=c('pred.ratio'))
@@
   ggplot() + geom_bar(data=subset(m, Status == 'P'), aes(Patient, value), fill='darkblue', color='grey', stat='identity') +  
+    geom_label( aes(x=nrow(p)/2+nrow(p)/3, y=0.5, 
+                    label=paste(round(nrow(p[p$pred.ratio == 1,])/nrow(p),2)*100,'%',sep='')), size=8, fill='darkblue', color='white', fontface='bold' ) +
+    geom_label( aes(x=nrow(p)/3, y=0.5, 
+                    label=paste(round(nrow(p[p$pred.ratio < 1 & p$pred.ratio > 0,])/nrow(p),2)*100,'%',sep='')), size=8, fill='darkblue', color='white', fontface='bold' ) +
+        geom_text( aes(x=2, y=0.5, 
+                    label=paste(round(nrow(p[p$pred.ratio == 0,])/nrow(p),2)*100,'%',sep='')), size=8,  fontface='bold' ) +
     theme(legend.position='none', axis.text.x = element_text(angle = 45, hjust = 1)) + coord_flip() +
     labs(title='Progressors', x='', y='Ratio predicted samples'),
+
   ggplot() + geom_bar(data=subset(m, Status == 'NP'), aes(Patient, value), fill='darkgreen', color='grey', stat='identity') + 
+        geom_label( aes(x=nrow(np)/2+nrow(np)/4, y=0.5, 
+                    label=paste(round(nrow(np[np$pred.ratio == 1,])/nrow(np),2)*100,'%',sep='')), size=8, fill='darkgreen', color='white', fontface='bold' ) +
+    geom_label( aes(x=nrow(np)/4, y=0.5, 
+                    label=paste(round(nrow(np[np$pred.ratio < 1 & np$pred.ratio > 0,])/nrow(p),2)*100,'%',sep='')), size=8, fill='darkgreen', color='white', fontface='bold' ) +
+        geom_text( aes(x=1.2, y=0.5, 
+                    label=paste(round(nrow(np[np$pred.ratio == 0,])/nrow(np),2)*100,'%',sep='')), size=8,  fontface='bold' ) +
     theme(legend.position='none', axis.text.x = element_text(angle = 45, hjust = 1)) + coord_flip() +
@@
 
-pg = subset(all.predictions, Status == 'P')
-np = subset(all.predictions, Status == 'NP')
-
 ```
@@
 
-When predicting each sample individually `r round(nrow(subset(pg, pred.ratio == 1))/nrow(pg))*100`% patients were correctly predicted in all samples.
-`r round(nrow(subset(pg, pred.ratio < 1 & pred.ratio > 0))/nrow(pg),2)*100`% of patients were correctly predicted in some samples
-
-`r nrow(subset(pg, pred.ratio == 0))` patients had no correctly predicted samples.
+When predicting each sample individually `r round(nrow(subset(p, pred.ratio == 1))/nrow(p)*100, 1)`% patients were correctly predicted in all samples.
+`r round(nrow(subset(p, pred.ratio < 1 & pred.ratio > 0))/nrow(p)*100, 1)`% of patients were correctly predicted in at least one sample.
+
+`r nrow(subset(p, pred.ratio == 0))` patients had no correctly predicted samples.
 
@@
 
-```{r timepoints2, echo=T, message=F, warning=F}
+```{r timepoints2, echo=T, message=F, warning=F, fig.height=10, fig.width=8}
 # Only BE, Progressors
 ndbe.only = do.call(rbind, lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], function(pt) {
+  final.endo = subset(pt, Pathology %in% c('IMC','HGD'))
   e = subset(pt, Pathology %in% c('BE', '', '?'))
-  cbind.data.frame('predicted'=sum(as.integer(e$Prediction > 0.5)), 'total.be'=nrow(e))
+  if (nrow(final.endo) > 0) # exclude BE at the HGD endoscopy
+    e = subset(e, PID != final.endo[nrow(final.endo), 'PID'])
+
+  cbind.data.frame('predicted'=sum(as.integer(e$Prediction > cutoff)), 'total.be'=nrow(e))
 }))
-ndbe.only = subset(ndbe.only, total.be > 0)
+ndbe.only = subset(ndbe.only, total.be > 0) # Exclude patients without any BE
 ndbe.only$pred.ratio = round(ndbe.only$predicted/ndbe.only$total.be, 2)
@@
   geom_bar(stat='identity', fill='darkred', alpha=0.6) + geom_label( aes(label=total.be) ) +
-  coord_flip() + labs(title='Correctly predicted NDBE samples', y='Correct prediction ratio')
+  coord_flip() + labs(title='Correctly predicted NDBE samples', subtitle='Excluding samples at HGD endoscopy', y='Correct prediction ratio')
 
 total = do.call(rbind, lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], function(pt) {
-    cbind.data.frame('predicted'=sum(as.integer(pt$Prediction > 0.5)), 'total'=nrow(pt))
+    cbind.data.frame('predicted'=sum(as.integer(pt$Prediction > cutoff)), 'total'=nrow(pt))
 }))
@@
   e$Prediction
-  sum(as.integer(e$Prediction > 0.5))/nrow(e)
+  sum(as.integer(e$Prediction > cutoff))/nrow(e)
 })
@@
   if (  length(which(e$Pathology %in% c('HGD','IMC'))) > 0 ) return(NA)
-  p = as.vector(as.integer(e$Prediction > 0.5))
+  p = as.vector(as.integer(e$Prediction > cutoff))
   names(p) = e$Endoscopy.Year
@@
   geom_bar(stat="identity", fill='dodgerblue3') + 
-  geom_text(aes(label=paste(`%.Predicted`*100, '%', sep='')), nudge_y=-0.02, size=8) + 
-  geom_text(aes(label=paste(n.Patients, 'patients')), nudge_y=0.01, size=8) + 
+  geom_text(aes(label=paste(`%.Predicted`*100, '%', sep='')), nudge_y=-0.02) + 
+  geom_text(aes(label=paste(n.Patients, 'patients')), nudge_y=0.01) + 
   geom_bar(stat='identity', fill='darkred', alpha=0.3, aes(y=FPR)) +
@@
 
-```{r, warning=F, message=F, echo=F, fig.width=16, fig.height=8}
-
-back = lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], function(pt) {
+```{r, warning=F, message=F, echo=F}
+transform.pt.by.time<-function(pt) {
   pt = pt %>% group_by(PID, Endoscopy.Year) %>% summarise(
-                  'pred'=sum(as.integer(Prediction > 0.5)), 'samples'=length(PID), 
+                  'pred'=sum(as.integer(Prediction > cutoff)), 'samples'=length(PID), 
                   #'final.endo'= length(which(grepl('HGD|IMC', Pathology))) > 0,
-                  'months.before.final'= (max(pt$Endoscopy.Year) - unique(Endoscopy.Year))*12 )
+                  'months.before.final'= (max(pt$Endoscopy.Year) - unique(Endoscopy.Year))*12, 
+                  'max.path'=max(Pathology))
   arrange(pt, -Endoscopy.Year, PID)
-})
+}
+
+back = lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], transform.pt.by.time)
 
@@
     if (i > nrow(x)) return(NA)
-    cbind(x[i, c('Endoscopy.Year','pred','samples', 'months.before.final')], i)
+    cbind(x[i, c('Endoscopy.Year','pred','samples', 'months.before.final', 'max.path')], i)
   }))
@@
 endos$pred.ratio = round(endos$pred/endos$samples, 2)
-#endos$i = as.factor(endos$i)
-
-#endos$brks <- cut(endos$pred.ratio, include.lowest=T,
-#                   breaks=c(0, 0.1, 0.25, 0.5, 0.75, 0.9, 1), 
-#                   labels=c('0', '0.1-.25', '.25-.5', '.5-.75','.75-0.9', '1'))
-
 endos = transform(endos, Patient=reorder(Patient, -months.before.final) )
-endos$months.before.final = endos$months.before.final*-1
-
-plot.endo<-function(df, minM=NULL, maxM=NULL) {
+
+endos$brks = cut(endos$months.before.final, include.lowest = T, ordered_result = T,
+    breaks=c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 12*12, 15*12))
+
+endos$labels = gsub('\\[|\\(|\\]', '',endos$brks)
+endos$labels = sub(',', '-',endos$labels)
+
+m = endos %>% group_by(brks, labels) %>% summarise( 'pred.r'= length(which(pred > 0))/length(brks), 'n'=length((Patient)) )
+ggplot(m, aes(brks, pred.r*100, group=brks)) + ylim(0,105) +
+  geom_bar(stat='identity', fill='dodgerblue3') + 
+  geom_text(aes(y=25, label=paste(round(pred.r*100), '%', sep='')), nudge_y=-0.02) + 
+  geom_text(aes(y=pred.r*100+1.5, label=paste('n=',n, sep='')), nudge_y=0.02) + 
+  scale_x_discrete(labels=m$labels) +
+  labs(x='Months', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = number of endoscopies')
+
+plot.endo<-function(df, minM=NULL, maxM=NULL, bin=F) {
   if (is.null(minM)) minM = min(df$months.before.final)
@@
 
-  df$i = as.factor(df$i)
-  ggplot(df, aes(months.before.final, pred.ratio, group=1,color=Patient)) + facet_grid(Patient ~., scales='fixed') + scale_x_continuous(breaks=seq(minM, maxM, by = 12)) +
-    geom_line() + geom_point() + labs(x='Months before diagnosis', y='Sample prediction ratio') + theme(legend.position='none') 
-}
-
-me = endos %>% group_by(Patient) %>% summarise('max.endo' = max(i))
-
-grid.arrange(
-  plot.endo( subset(endos, Patient %in% subset(me, max.endo >=5)$Patient), min(endos$months.before.final), max(endos$months.before.final) ),
-  plot.endo( subset(endos, Patient %in% subset(me, max.endo < 5 & max.endo > 2)$Patient),min(endos$months.before.final), max(endos$months.before.final)  ),
-  ncol=2)
-
-```
-
-```{r, echo=F, warning=F, message=F, fig.height=6, fig.width=4}
-plot.endo( subset(endos, Patient %in% subset(me, max.endo <= 2 & max.endo > 0)$Patient), min(endos$months.before.final), max(endos$months.before.final)  )
-```
-
-```{r, echo=F, warning=F, message=F, fig.height=4, fig.width=4}
-plot.endo( subset(endos, Patient %in% subset(me, max.endo == 0)$Patient),min(endos$months.before.final), max(endos$months.before.final) )
-```
-
+  #df$i = as.factor(df$i)
+
+  df = arrange(df %>% group_by(Patient) %>% mutate('overall.pred'=sum(as.integer(pred > 0))/length(Patient)), Patient)
+
+  df = transform(df, Patient=reorder(Patient, -overall.pred) )
+
+  if (bin) {
+    df$endo.pred = ifelse(df$pred > 0, 'Pos','Neg')
+    ggplot(df, aes(months.before.final, endo.pred, group=1,color=Patient)) + facet_grid(Patient ~.) + 
+      scale_x_continuous(breaks=seq(minM, maxM, by = 12)) +
+      geom_line() + geom_point() +
+      geom_label_repel(aes(label=max.path, fill=Patient, color=Patient), color='white', fontface='bold', size=3, label.size=0, label.r=unit(0.25, 'lines'), label.padding=unit(0.15, 'lines')) +
+      labs(x='Months before diagnosis', y='Predictive Endoscopy') + theme(legend.position='none') + theme(strip.text = element_text(face="bold", size=9))
+  } else {
+  ggplot(df, aes(months.before.final, pred.ratio, group=1,color=Patient)) + facet_grid(Patient ~.) + scale_x_continuous(breaks=seq(minM, maxM, by = 12)) +
+    geom_line() + geom_point() +  
+      geom_label_repel(aes(label=max.path, fill=Patient), color='white', fontface='bold', size=3, label.size=0, label.r=unit(0.25, 'lines'), label.padding=unit(0.15, 'lines')) +
+      labs(x='Months before diagnosis', y='Sample prediction ratio') + theme(legend.position='none') 
+  }
+}
+
+endos$months.before.final = endos$months.before.final*-1
+me = endos %>% group_by(Patient) %>% summarise('max.endo' = max(i), 'months' = min(months.before.final))
+```
+
+```{r, echo=F, warning=F, message=F, fig.height=22, fig.width=10}
+p = plot.endo( subset(endos, Patient %in% subset(me, months <= -72)$Patient), min(endos$months.before.final), max(endos$months.before.final), T)
+
+png(filename='progressors_1.png', width=600, height=1050)
+print(p)
+dev.off()
+
+grid.arrange(p)
+```
+
+```{r, echo=F, warning=F, message=F, fig.height=22, fig.width=10}
+p1 = plot.endo( subset(endos, Patient %in% subset(me, months <= -48 & months > -72)$Patient), min(endos$months.before.final), max(endos$months.before.final), T) 
+p2 = plot.endo( subset(endos, Patient %in% subset(me, months <= -24 & months > -48)$Patient), min(endos$months.before.final), max(endos$months.before.final), T) 
+p3 = plot.endo( subset(endos, Patient %in% subset(me, months >= -12)$Patient), min(endos$months.before.final), max(endos$months.before.final), T)
+
+grid.arrange(p1,p2,p3,ncol=1, top='Progressor predictions prior to diagnosis')
+
+png(filename='progressors_2.png', width=400, height=600)
+print(p1)
+dev.off()
+
+png(filename='progressors_3.png', width=400, height=600)
+print(p2)
+dev.off()
+
+png(filename='progressors_4.png', width=400, height=600)
+print(p3)
+dev.off()
+
+```
+
+Just for completeness, we can plot non-progressors similarly:
+
+```{r echo=F, warning=F, message=F, fig.height=42, fig.width=10}
+nonp.back = lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Patient],transform.pt.by.time)
+
+nonp.endos = do.call(rbind, lapply(1:max(sapply(nonp.back,nrow)), function(i) {
+  df = do.call(rbind, lapply(nonp.back, function(x) {
+    if (i > nrow(x)) return(NA)
+    cbind(x[i, c('Endoscopy.Year','pred','samples', 'months.before.final', 'max.path')], i)
+  }))
+  df$Patient = rownames(df)
+  return(df)
+}))
+nonp.endos = nonp.endos[complete.cases(nonp.endos),]
+nonp.endos$i = nonp.endos$i-1
+
+nonp.endos$pred.ratio = round(nonp.endos$pred/nonp.endos$samples, 2)
+nonp.endos = transform(nonp.endos, Patient=reorder(Patient, -months.before.final) ) 
+nonp.endos$months.before.final = nonp.endos$months.before.final*-1
+
+me = nonp.endos %>% group_by(Patient) %>% summarise('max.endo' = max(i), 'months' = min(months.before.final))
+
+nonp.endos$months.before.final
+p1 = plot.endo( subset(nonp.endos, Patient %in% subset(me, months <= -108)$Patient), min(nonp.endos$months.before.final), max(nonp.endos$months.before.final), T)
+
+png(filename='nonprogressors_1.png', width=600, height=800)
+print(p1)
+dev.off()
+
+p2 = plot.endo( subset(nonp.endos, Patient %in% subset(me, months <= -72 & months > -108 )$Patient), min(nonp.endos$months.before.final), max(nonp.endos$months.before.final), T)
+
+png(filename='nonprogressors_2.png', width=600, height=800)
+print(p2)
+dev.off()
+
+p3 = plot.endo( subset(nonp.endos, Patient %in% subset(me, months > -72  )$Patient), min(nonp.endos$months.before.final), max(nonp.endos$months.before.final), T)
+
+png(filename='nonprogressors_3.png', width=600, height=800)
+print(p3)
+dev.off()
+
+grid.arrange(p1,p2,p3, ncol=1)
+```
+
+```{r, warning=F, message=F, fig.height=4, fig.width=4, eval=F}
 ### How many samples?  TBD
 
-Clinicians asked how many samples they needed to take.
-
-```{r, warning=F, message=F, fig.height=4, fig.width=4, eval=F}
+#Clinicians asked how many samples they needed to take.
+
 # How do I evaluate the predictions?
@@
   # Only for progressors because we know for certain they progress, the non-ps may just not have progressed *yet*
-  tpr = length(which(pred > 0.5))/length(pred)
-  fpr = length(which(predB > 0.5))/length(predB)
+  tpr = length(which(pred > cutoff))/length(pred)
+  fpr = length(which(predB > cutoff))/length(predB)
   cbind.data.frame('TPR'=tpr, 'FPR'=fpr)
@@
 pilot = subset(pilot, Status == 'NP')
-length(which(pilot$Prediction > 0.5))/nrow(pilot)
+length(which(pilot$Prediction > cutoff))/nrow(pilot)
 
 p2 = subset(pilot, Patient %nin% c('AD0591', 'AH0329'))
-length(which(p2$Prediction > 0.5))/nrow(p2)
+length(which(p2$Prediction > cutoff))/nrow(p2)
 
@@
 exome = subset(exome, Status == 'NP')
-length(which(exome$Prediction > 0.5))/nrow(exome)
+length(which(exome$Prediction > cutoff))/nrow(exome)
 
 e2 = subset(exome, Patient %nin% c('AD0591', 'AH0329'))
-length(which(e2$Prediction > 0.5))/nrow(e2)
+length(which(e2$Prediction > cutoff))/nrow(e2)
 
@@
 Both patients had samples in the exome batch, while AH0329 also had samples in the NP_Pilot batch.
-Without those patients the FPR in the exome batch was `r round(length(which(e2$Prediction > 0.5))/nrow(e2), 2)*100`% for `r length(unique(e2$Patient))` patient(s), and in the NP_Pilot batch the FPR was `r round(length(which(p2$Prediction > 0.5))/nrow(p2), 2)*100`% for `r length(unique(p2$Patient))` patient(s).
-
-The somewhat lower Batch_1 prediction rate is driven by the samples from AHM0277 as only `r length(which(pg.samp$AHM0277$Prediction > 0.5))/nrow(pg.samp$AHM0277)*100`% of this patient's samples were predicted to progress.
+Without those patients the FPR in the exome batch was `r round(length(which(e2$Prediction > cutoff))/nrow(e2), 2)*100`% for `r length(unique(e2$Patient))` patient(s), and in the NP_Pilot batch the FPR was `r round(length(which(p2$Prediction > cutoff))/nrow(p2), 2)*100`% for `r length(unique(p2$Patient))` patient(s).
+
+The somewhat lower Batch_1 prediction rate is driven by the samples from AHM0277 as only `r length(which(pg.samp$AHM0277$Prediction > cutoff))/nrow(pg.samp$AHM0277)*100`% of this patient's samples were predicted to progress.
 Batch_3 is driven by three other patients that have low prediction rates: PR1_ADH_076, PR1_HIN_046, and PR1_WSH_026.
-
-### Look at these specific patients
-
-```{r, echo=F, eval=F}
-
-pg.samp$AD0591
-pg.samp$AH0329
-pg.samp$PR1_ADH_076
-pg.samp$PR1_HIN_046
-pg.samp$PR1_WSH_026
-
-```
 
@@
 # Compare poorly predicted to well predicted based on the sample SD measures, no difference
-wilcox.test(subset(prog, Prediction > 0.5)$SampleSD, subset(prog, Prediction < 0.5)$SampleSD)
-wilcox.test(subset(nonp, Prediction > 0.5)$SampleSD, subset(nonp, Prediction < 0.5)$SampleSD)
+wilcox.test(subset(prog, Prediction > cutoff)$SampleSD, subset(prog, Prediction < cutoff)$SampleSD)
+wilcox.test(subset(nonp, Prediction > cutoff)$SampleSD, subset(nonp, Prediction < cutoff)$SampleSD)
 
@@
 ## Progressors
-tprows = which(prog$Prediction > 0.5)
+tprows = which(prog$Prediction > cutoff)
 
@@
 ## Non-Progressors, this is a little more complicated as we cannot be sure of True Positives but we'll look anyhow
-tprowsNP = which(nonp$Prediction < 0.5)
+tprowsNP = which(nonp$Prediction < cutoff)
 cor.test(nonp[tprowsNP, 'SampleSD'], nonp[tprowsNP, 'Prediction'] ) # Correlated for True Pos predictions
@@
 ggplot(prog, aes(x=SampleSD)) + 
-  geom_histogram(data=subset(prog, Prediction > 0.5), aes(SampleSD, fill='TP'), alpha=0.5) +
-  geom_histogram(data=subset(prog, Prediction < 0.5), aes(SampleSD, fill='FN'), alpha=0.5) + 
+  geom_histogram(data=subset(prog, Prediction > cutoff), aes(SampleSD, fill='TP'), alpha=0.5) +
+  geom_histogram(data=subset(prog, Prediction < cutoff), aes(SampleSD, fill='FN'), alpha=0.5) + 
   labs(title='SD of Samples vs True positive ~ False negative Progressors')
 
-ggplot(nonp, aes(x=SampleSD)) + 
-  geom_histogram(data=subset(nonp, Prediction < 0.5), aes(SampleSD, fill='TN'), alpha=0.5) + 
-  geom_histogram(data=subset(nonp, Prediction > 0.5), aes(SampleSD, fill='FP'), alpha=0.5) +
-  labs(title='SD of Samples vs False positive ~ True negative Non-Progressors')
-
-```
+```
+
+# Validation Dataset
+
+20 Patients (10P, 10NP) set aside for validation purposes
+
+```{r, message=F, warning=F, echo=F, eval=T, fig.height=8, fig.width=6}
+
+fitV <- glmnet(dysplasia.df, labels, alpha=select.alpha, nlambda=nl, family='binomial') # all patients
+lambda.opt = performance[select.alpha, 'lambda']
+
+pred = predict(fitV, newx=validation.df, s=lambda.opt, type='response')
+
+vpd = lapply(validation.patient.data, function(pt) {
+  info = pt$info
+  info$PID = unlist(lapply(info$Path.ID, function(x) unlist(strsplit(x, 'B'))[1]))
+  return(info)
+})
+
+samples = intersect(rownames(pred), validation.patient.info$Samplename)
+validation.patient.info$Prediciton = NA
+validation.patient.info[which(validation.patient.info$Samplename %in% samples), 'Prediction'] = pred[samples,]
+
+for (pt in names(vpd)) {
+  vpd[[pt]]$Prediction[which(vpd[[pt]]$Samplename %in% rownames(pred))] = pred[ intersect(rownames(pred), vpd[[pt]]$Samplename), ]
+}
+
+v.predictions = sapply(vpd, function(pt) {
+  if (unique(pt$Status) == 'NP') {
+    as.integer(pt$Prediction < cutoff)
+  } else {
+    as.integer(pt$Prediction > cutoff)
+  }
+})
+
+v.predictions = do.call(rbind, lapply(v.predictions, function(x) {
+  cbind.data.frame('correctly.predicted'=sum(x,na.rm=T), 'total.samples'=length(x))
+}))
+
+v.predictions$Status = ifelse (rownames(v.predictions) %in% subset(sum.validation.info, Status == 'P')$Patient, 'P', 'NP')
+
+v.predictions$pred.ratio = with(v.predictions, correctly.predicted/total.samples)
+v.predictions$Patient = rownames(v.predictions)
+v.predictions = transform(v.predictions, Patient=reorder(Patient, pred.ratio) )
+
+p = subset(v.predictions, Status == 'P')
+np = subset(v.predictions, Status == 'NP')
+
+m = melt(v.predictions, id.vars=c('Patient','Status'), measure.vars=c('pred.ratio'))
+
+grid.arrange(
+  ggplot() + geom_bar(data=subset(m, Status == 'P'), aes(Patient, value), fill='darkblue', color='grey', stat='identity') +  
+    geom_label( aes(x=nrow(p)/2+nrow(p)/3, y=0.5, label=paste(round(nrow(p[p$pred.ratio == 1,])/nrow(p),2)*100,'%',sep='')), size=8, fill='darkblue', color='white', fontface='bold' ) +
+    geom_label( aes(x=nrow(p)/3, y=0.5, label=paste(round(nrow(p[p$pred.ratio < 1 & p$pred.ratio > 0,])/nrow(p),2)*100,'%',sep='')), size=8, fill='darkblue', color='white', fontface='bold' ) +
+    geom_text( aes(x=2, y=0.5, label=paste(round(nrow(p[p$pred.ratio == 0,])/nrow(p),2)*100,'%',sep='')), size=8,  fontface='bold' ) +
+    theme(legend.position='none', axis.text.x = element_text(angle = 45, hjust = 1)) + coord_flip() +
+    labs(title='Progressors', x='', y='Ratio predicted samples'),
+
+  ggplot() + geom_bar(data=subset(m, Status == 'NP'), aes(Patient, value), fill='darkgreen', color='grey', stat='identity') + 
+        geom_label( aes(x=nrow(np)/2+nrow(np)/4, y=0.5, label=paste(round(nrow(np[np$pred.ratio == 1,])/nrow(np),2)*100,'%',sep='')), size=8, fill='darkgreen', color='white', fontface='bold' ) +
+    geom_label( aes(x=nrow(np)/4, y=0.5, label=paste(round(nrow(np[np$pred.ratio < 1 & np$pred.ratio > 0,])/nrow(p),2)*100,'%',sep='')), size=8, fill='darkgreen', color='white', fontface='bold' ) +
+        #geom_text( aes(x=1.2, y=0.5, label=paste(round(nrow(np[np$pred.ratio == 0,])/nrow(np),2)*100,'%',sep='')), size=8,  fontface='bold' ) +
+    theme(legend.position='none', axis.text.x = element_text(angle = 45, hjust = 1)) + coord_flip() +
+    labs(title='Non Progressors', x='', y='Ratio predicted samples'),
+  ncol=2, top='Validation data set')
+
+```
+
+```{r, echo=F, message=F, warning=F, fig.width=8, fig.height=12}
+# sapply(vpd[subset(sum.validation.info, Status == 'P')$Patient], function(df) {
+#   length(which(df$Prediction > cutoff))/nrow(df)
+# })
+
+back = lapply(vpd, transform.pt.by.time)
+vpi = do.call(rbind, lapply(1:max(sapply(back,nrow)), function(i) {
+  df = do.call(rbind, lapply(back, function(x) {
+    if (i > nrow(x)) return(NA)
+    cbind(x[i, c('Endoscopy.Year','pred','samples', 'months.before.final', 'max.path')], i)
+  }))
+  df$Patient = rownames(df)
+  return(df)
+}))
+
+vpi = vpi[complete.cases(vpi),]
+vpi$pred.ratio = round(vpi$pred/vpi$samples, 2)
+vpi = transform(vpi, Patient=reorder(Patient, -months.before.final) ) 
+vpi$months.before.final = vpi$months.before.final*-1
+
+me = vpi %>% group_by(Patient) %>% summarise('max.endo' = max(i), 'months' = min(months.before.final))
+
+#vpi[,c('Status','Endoscopy.Year','pred','samples', 'months.before.final', 'max.path')]
+
+p1 = plot.endo(subset(vpi, Patient %in% subset(sum.validation.info, Status == 'P')$Patient), min(vpi$months.before.final), max(vpi$months.before.final), T) +  labs(title="Validation - Progressors")
+
+png(filename='validation_progressors.png', width=600, height=800)
+print(p1)
+dev.off()
+
+grid.arrange(p1,ncol=1)
+
+p2 = plot.endo(subset(vpi, Patient %in% subset(sum.validation.info, Status == 'NP')$Patient), min(vpi$months.before.final), max(vpi$months.before.final), T) + labs(title="Validation - Non Progressors")
+
+png(filename='validation_Nonprogressors.png', width=600, height=800)
+print(p2)
+dev.off()
+
+grid.arrange(p2,ncol=1)
+```
+
+```{r echo=F, warning=F, message=F}
+
+# TPR
+tpr = round(length(which(subset(validation.patient.info, Status == 'P')$Prediction > cutoff))/nrow(subset(validation.patient.info, Status == 'P')), 2)
+# FNR
+fnr = round(length(which(subset(validation.patient.info, Status == 'P')$Prediction < cutoff))/nrow(subset(validation.patient.info, Status == 'P')), 2)
+
+# FPR
+fpr = round(length(which(subset(validation.patient.info, Status == 'NP')$Prediction > cutoff))/nrow(subset(validation.patient.info, Status == 'NP')), 2)
+# TNR
+tnr = round(length(which(subset(validation.patient.info, Status == 'NP')$Prediction < cutoff))/nrow(subset(validation.patient.info, Status == 'NP')), 2)
+
+m = melt(validation.patient.info, id.vars = c('Status'), measure.vars = c('Prediction'))
+p3 = ggplot(m, aes(Status, value, fill=Status)) + geom_boxplot() + geom_jitter(color='darkgrey', show.legend = F) +
+  geom_label(aes(x='P', y=0.75, label=paste('TPR=',tpr*100, '%', sep='')), fill='white') + 
+  geom_label(aes(x='P', y=0.25, label=paste('FNR=',fnr*100, '%', sep='')), fill='white') + 
+  geom_label(aes(x='NP', y=0.55, label=paste('FPR=',fpr*100, '%', sep='')), fill='white') + 
+  geom_label(aes(x='NP', y=0.25, label=paste('TNR=',tnr*100, '%', sep='')), fill='white') + 
+  geom_hline(yintercept = cutoff, linetype = 'dashed', color = 'grey27') + 
+  labs(title="Validation set predictions per sample", subtitle=paste("Prediction cutoff =",cutoff))
+
+grid.arrange(p3)
+
+png(filename='validation_predictions.png', width=600, height=800)
+print(p3)
+dev.off()
+
+```

### 2017-07-05 lib/load_patient_metadata.R

@@ -7,10 +7,20 @@ summarise.patient.info<-function(df) {
   require(plyr)
   
   sum.pt = ddply(df, .(Patient, Status, Set), summarise, 
+        age.diagnosis = unique(Age.at.diagnosis),
+        wt = unique(Weight.kg.),
+        ht = unique(Height.cm.),
+        BMI = unique(BMI),
+        gender = unique(Sex),
+        smoking = unique(Smoking),
+        C = unique(Circumference),
+        M = unique(Maximal),
         start.year=range(Endoscopy.Year)[1], 
         end.year=range(Endoscopy.Year)[2],
         total.samples=length(Samplename), 
         total.endos=length(unique(Path.ID)),
+        initial.endo = unique(Initial.Endoscopy),
+        final.endo = unique(Final.Endoscopy),
         highest.path=sort(Pathology, decreasing=T)[1],
         first.batch=sort(Batch.Name)[1],
         med.cellularity=median(Barretts.Cellularity,na.rm=T),
@@ -18,7 +28,7 @@ summarise.patient.info<-function(df) {
   return(arrange(sum.pt, Status, total.samples, highest.path))
 }
 
-read.patient.info<-function(file, set='Training') {
+read.patient.info<-function(file, file2=NULL, set='Training') {
   if (!file.exists(file))
     stop(paste(file, "doesn't exist or is not readable."))
   
@@ -83,6 +93,7 @@ read.patient.info<-function(file, set='Training') {
     # Remove 'normal' samples
     patient.info = subset(patient.info, !grepl('D2', Pathology, ignore.case=T))
     patient.info = subset(patient.info, !grepl('Gastriccardia', Pathology, ignore.case=T))
+    patient.info = subset(patient.info, !grepl('normal', Pathology, ignore.case=T))
 
   } else {
     patient.info = read.table(file, header=T, sep='\t', stringsAsFactors=F)
@@ -146,8 +157,75 @@ read.patient.info<-function(file, set='Training') {
   }
 
   patient.info$Patient = gsub('/', '_',patient.info$Patient)
+  
+  if (!is.null(file2)) {
+    patient.info = add.demographics(file2, patient.info)
+  }
+  
   return(patient.info)
 }
+
[…]

### 2017-07-05 RegressionPvNP.Rmd
...

@@
 
+#### Correlations
+
+```{r echo=F, message=F, warning=F}
+library(Hmisc)
+
+flattenCorrMatrix <- function(cormat, pmat) {
+  ut <- upper.tri(cormat)
+  data.frame(
+    row = rownames(cormat)[row(cormat)[ut]],
+    column = rownames(cormat)[col(cormat)[ut]],
+    cor  =(cormat)[ut],
+    p = pmat[ut]
+    )
+}
+
+head(dysplasia.df)
+rc = rcorr(dysplasia.df)
+rc = flattenCorrMatrix(rc$r, rc$P)
+
+rc = subset(rc, row %in% rownames(coefs[[select.alpha]]) | column %in% rownames(coefs[[select.alpha]]))
+rc = subset(rc, p < 0.05)
+
+nrow(rc)
+head(rc)
+
+pos = do.call(rbind.data.frame, strsplit(as.character(rc$column), ':|-'))
+colnames(pos) = c('col.chr','col.start','col.end')
+pos[] = lapply(pos[], function(x) as.numeric(as.character(x)))
+rc = cbind(pos, rc)
+
+pos = do.call(rbind.data.frame, strsplit(as.character(rc$row), ':|-'))
+colnames(pos) = c('row.chr','row.start','row.end')
+pos[] = lapply(pos[], function(x) as.numeric(as.character(x)))
+rc = cbind(pos, rc)
+
+rc$r.coef = as.integer(rc$row %in% rownames(coefs[[select.alpha]]))
+
+head(rc)
+
+rc = subset(rc, cor >= 0.5)
+
+rc = arrange(rc, row.chr, col.chr, row.start, col.start)
+head(rc)
+
+#rc[sort(c(which(rc$r.coef == 1), which(rc$r.coef == 1)+1)),]
+
+rc1  = subset(rc, row.chr == 16 & col.chr == 16)
+
+rc1 = transform(rc1, row=reorder(row, -row.start) ) 
+rc1 = transform(rc1, column=reorder(column, col.start) )
+
+subset(rc1, column == '16:85599240-90354752' | row == '16:85599240-90354752')
+
+head(rc1)
+
+ggplot(rc1, aes(column, row, fill=cor)) + geom_tile()  +  #scale_fill_gradient(low = "white", high = "red") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+## closer regions are more liklye to be correlated
+cor.test(rc1$col.start - rc1$row.end , rc1$cor)
+
+```
+
 ### Hazards

### 2017-07-19 Clustering.Rmd
...

@@
 ## Patient info file
-patient.file = grep('All_patient_info.txt', data.files, value=T)
+patient.file = grep('All_patient_info.xlsx', data.files, value=T)
 if (length(patient.file) != 1)
   stop(paste("Missing/too many patient info file(s) in", data))
-demo.file = grep('Demographics_full.txt', data.files, value=T)
+demo.file = grep('Demographics_full.xlsx', data.files, value=T)
 
@@
 
-patient.info = read.patient.info(patient.file, demo.file)
+patient.info = read.patient.info(patient.file, demo.file)$info
 #head(patient.info)
@@
 patient.info = arrange(patient.info, Status, Patient, Endoscopy.Year, Pathology)
+patient.info = patient.info[-which(patient.info$Hospital.Research.ID == 'AH0222'),]
 
 sum.patient.data = summarise.patient.info(patient.info)
-sum.patient.data = as.data.frame(subset(sum.patient.data, Patient %in% names(patient.data))) ## For now
 pander(sum.patient.data)
@@
 # Missing some of the samples as they aren't all sequenced yet
-for (pt in names(patient.data)) {
-  patient.data[[pt]]$info= patient.data[[pt]]$info[patient.data[[pt]]$info$Samplename %in% colnames(patient.data[[pt]]$seg.vals)[-(1:5)],]
-}
+#for (pt in names(patient.data)) {
+#  patient.data[[pt]]$info= patient.data[[pt]]$info[patient.data[[pt]]$info$Samplename %in% colnames(patient.data[[pt]]$seg.vals)[-(1:5)],]
+#}
 
@@
 sim = as.data.frame(sum.patient.data[,c('Status','years')])
-rownames(sim) = sum.patient.data$Patient
-for (pt in sum.patient.data$Patient) {
+rownames(sim) = sum.patient.data$Hospital.Research.ID
+for (pt in sum.patient.data$Hospital.Research.ID) {
   print(pt)
@@
 
-  gg = ggplot(melt(x1), aes(x=Var2, y=value, group=Var2, fill=Var2)) + 
+  gg = ggplot(melt(x1), aes(x=Var2, y=value, group=Var2, fill=Var2, color=Var2)) + 
     geom_boxplot(notch=T, show.legend=F) +
@@
 grid.arrange(g1, g2, top='Raw seg values')
+```
+
+```{r echo=F, message=F, warning=F, fig.width=5, fig.height=5}
+g2 + labs(x='Status', y='Complexity (# clusters) per patient', title='AP Clustering', subtitle=paste("p-value", signif(ttpc, 3))) + theme(text=element_text(size=14, face='bold'), legend.position = 'non')
 ```
@@
 
+```{r apclustN, echo=F, message=F, warning=F, fig.height=5, fig.width=5, fig.align='left', eval=F}
 ### Using the normalized values
 
-So what about using the normalized segmented values for AP clustering instead?
+#So what about using the normalized segmented values for AP clustering instead?
 
-```{r apclustN, echo=F, message=F, warning=F, fig.height=5, fig.width=5, fig.align='left'}
 # Should this be performed on the normalized seg values instead?
@@
 rownames(sim3) = sum.patient.data[,1]
-for (pt in sum.patient.data$Patient) {
+for (pt in sum.patient.data$Hospital.Research.ID) {
   print(pt)
@@
 
-  for (pt in sum.patient.data$Patient[-i]) {
+  for (pt in sum.patient.data$Hospital.Research.ID[-i]) {
     x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
@@
 
-```{r xval_smp, echo=F, warning=F, message=F, fig.align='left', fig.height=4, fig.width=4}
+```{r xval_smp, echo=F, warning=F, message=F, fig.align='left', fig.height=4, fig.width=4, eval=F}
 # Leave out samples?
@@
   simX = sum.patient.data[,c('Status','years')]
-  rownames(simX) = sum.patient.data[,'Patient']
+  rownames(simX) = sum.patient.data[,'Hospital.Research.ID']
 
-  for (pt in sum.patient.data$Patient) {
+  for (pt in sum.patient.data$Hospital.Research.ID) {
       if (length(patient.data[[pt]]$info$Samplename) <= 2) next
@@
 simX = sum.patient.data[,c('Status','years')]
-rownames(simX) = sum.patient.data[,'Patient']
+rownames(simX) = sum.patient.data[,'Hospital.Research.ID']
 # Leave out final sample on the progressors
 
-for (pt in sum.patient.data$Patient) {
+for (pt in sum.patient.data$Hospital.Research.ID) {
   x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
-  if (subset(sum.patient.data, Patient == pt)$Status == 'P' & length(which(grepl('HGD|IMC', patient.data[[pt]]$info$Pathology))) > 0) {
+  if (subset(sum.patient.data, Hospital.Research.ID == pt)$Status == 'P' & length(which(grepl('HGD|IMC', patient.data[[pt]]$info$Pathology))) > 0) {
     final = grep('HGD|IMC', patient.data[[pt]]$info$Pathology)
@@
 
-subset(sum.patient.data, Patient %in% pts)
+subset(sum.patient.data, Hospital.Research.ID %in% pts)
 
@@
   rownames(simPC) = sum.patient.data[,1]
-  for (pt in sum.patient.data$Patient) {
+  for (pt in sum.patient.data$Hospital.Research.ID) {
     patient.data[[pt]]$info = arrange(patient.data[[pt]]$info, Endoscopy.Year, desc(Pathology))

### 2017-07-19 Patient Report.Rmd
...

--- 2017-05-24 Patient Report.Rmd
@@
 library(ggplot2)
+library(ggbio)
+library(scales)
 library(GGally)
@@
 ## Patient info file
-patient.file = grep('/All_patient_info.txt', data.files, value=T)
+patient.file = grep('/All_patient_info.xlsx', data.files, value=T)
 if (length(patient.file) != 1)
   stop(paste("Missing/too many patient info file(s) in", data))
+demo.file = grep('/Demographics_full.xlsx', data.files, value=T)
 
-dataset = 'Training'
-patient.info = read.patient.info(patient.file, set=dataset)
-head(patient.info)
+dataset = 'All'
+info = read.patient.info(patient.file, demo.file, set=dataset)
+removed.samples = info$removed
+patient.info = info$info
+
+#head(patient.info)
 
@@
 
-  for (patient.name in sum.patient.data$Patient  )  {
+  for (patient.name in sum.patient.data$Hospital.Research.ID  )  {
     message(patient.name)
@@
     segvals = read.table(paste(patient.plot.dir, "/",patient.name,"_segmentedCoverage_fitted_gamma",gamma2,".txt",sep=""),sep="\t",stringsAsFactors=F,header=T)
-    dim(segvals)
+    head(segvals)
 
-    missing = c(setdiff(colnames(segvals)[-(1:5)], subset(patient.info, Patient == patient.name)$Samplename),
-                      setdiff(subset(patient.info, Patient == patient.name)$Samplename, colnames(segvals)[-(1:5)]))
+    missing = c( setdiff(colnames(segvals)[-(1:5)], subset(patient.info, Hospital.Research.ID == patient.name)$Samplename),
+                 setdiff(subset(patient.info, Hospital.Research.ID == patient.name)$Samplename, colnames(segvals)[-(1:5)]) )
+
+    mr = with(removed.samples, which(Hospital.Research.ID == patient.name & Samplename %in% missing))
+    if (length(mr) > 0) {
+      missing = missing[-which(missing %in% removed.samples[mr,]$Samplename)]
+      segvals = segvals[,-which(colnames(segvals) %in% removed.samples[mr,]$Samplename) ]
+    }
 
@@
     # ORDER segvals by endoscopy year etc
-    cur.pt = subset(patient.info, Patient == patient.name)
+    cur.pt =  arrange( subset(patient.info, Hospital.Research.ID == patient.name), Endoscopy.Year, Pathology)
     segvals = segvals[, c(colnames(segvals[1:5]), intersect(colnames(segvals), cur.pt$Samplename))]
@@
 
-      patient = arrange(subset(patient.info, Patient == patient.name),  Endoscopy.Year, Pathology, Samplename)
+      patient = arrange(subset(patient.info, Hospital.Research.ID == patient.name),  Endoscopy.Year, Pathology, Samplename)
     }
@@
 }
-print(paste("Missing patients:", paste(setdiff(sum.patient.data$Patient, names(patient.data)), collapse=', ')))
+print(paste("Missing patients:", paste(setdiff(sum.patient.data$Hospital.Research.ID, names(patient.data)), collapse=', ')))
 
-sum.patient.data = subset(sum.patient.data, Patient %in% names(patient.data))
+sum.patient.data = subset(sum.patient.data, Hospital.Research.ID %in% names(patient.data))
 table(sum.patient.data$Status)
@@
 ## Check that all of the samples are there
-patient.info = arrange(patient.info, Status, Patient, Endoscopy.Year,Samplename)
+patient.info = arrange(patient.info, Status, Patient, Hospital.Research.ID, Endoscopy.Year,Samplename)
 msgs = sapply(names(patient.data), function(pt) {
@@
 
-example.pt = subset(sum.patient.data, Status == 'P' & total.samples > 2)$Patient[5]
+example.pt = which(with(sum.patient.data, Status == 'P' & total.samples > 2))[5]
+key =  sum.patient.data[example.pt, 'Hospital.Research.ID' ]
+example.pt.id = sum.patient.data[example.pt, 'Patient']
 ## Note the number of probes per patient/sample
@@
 ```{r}
-head(patient.data[[example.pt]]$seg.vals)
+head(patient.data[[ key ]]$seg.vals)
 ```
@@
 ```{r echo=F, warning=F, message=F, fig.height=10, fig.width=10}
-ggplot( melt(patient.data[[example.pt]]$seg.vals[,-(1:5)]), aes(x=value) ) + 
+ggplot( melt(patient.data[[key]]$seg.vals[,-(1:5)]), aes(x=value) ) + 
         facet_wrap(~variable,scales = "free_x") + 
-        geom_histogram( fill=I("blue"),col=I("lightblue"), binwidth=0.03) + labs(title=paste(example.pt,' segmented value distribution'), x='', y='')
+        geom_histogram( fill=I("blue"),col=I("lightblue"), binwidth=0.03) + 
+  labs(title=paste('Patient', example.pt.id,' segmented value distribution'), x='', y='')
 ```
@@
 ```{r}
-head(patient.data[[example.pt]]$norm.seg.vals)
+head(patient.data[[key]]$norm.seg.vals)
 ```
@@
 ```{r echo=F, warning=F, message=F, fig.height=10, fig.width=10}
-ggplot( melt(patient.data[[example.pt]]$norm.seg.vals), aes(x=value) ) + 
+ggplot( melt(patient.data[[key]]$norm.seg.vals), aes(x=value) ) + 
         facet_wrap(~variable,scales = "free_x") + 
-        geom_histogram( fill=I("blue"),col=I("lightblue"), binwidth=0.3) + labs(title=paste(example.pt,' normalized segmented value distribution'), x='', y='')
+        geom_histogram( fill=I("blue"),col=I("lightblue"), binwidth=0.3) + labs(title=paste('Patient', example.pt.id,'normalized segmented value distribution'), x='', y='')
 
@@
 
-sum.patient.data = merge(sum.patient.data, gen, by.x='Patient', by.y='row.names')
+sum.patient.data = merge(sum.patient.data, gen, by.x='Hospital.Research.ID', by.y='row.names')
 sum.patient.data = arrange(sum.patient.data, Status, total.samples)
@@
 
-  do.call('grid.arrange', list(g1,g2, top=paste(name, status, sep=" - "), ncol=2))
+  do.call('grid.arrange', list(g1,g2, top=paste('Patient', name, status, sep=" - "), ncol=2))
 })
@@
 ```{r var_reg, echo=F, message=F, warning=F}
-prog.pt = subset(sum.patient.data, Status == 'P')$Patient
-np.pt = subset(sum.patient.data, Status == 'NP')$Patient
+prog.pt = subset(sum.patient.data, Status == 'P')$Hospital.Research.ID
+np.pt = subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID
 
 var.seg.lengths = lapply(patient.data, function(df) {
-  cbind('chr'=df$variable.regions$chrom, 'seg.len'=df$variable.regions$end.pos - df$variable.regions$start.pos)
+  cbind('chr'=df$variable.regions$chrom, 'start'=df$variable.regions$start.pos, 'end'=df$variable.regions$end.pos, 'seg.len'=df$variable.regions$end.pos - df$variable.regions$start.pos)
 })
@@
 
+sum.var = rbind(
+  'Progressors'=do.call(rbind.data.frame, var.seg.lengths[prog.pt]) %>% summarise( 'n.segments' = length(seg.len), 'median.length'=median(seg.len), 'mean.length'=mean(seg.len), 'sd'=sd(seg.len)),
+  'Non-Progressors'=do.call(rbind.data.frame, var.seg.lengths[np.pt]) %>% summarise( 'n.segments' = length(seg.len), 'median.length'=median(seg.len), 'mean.length'=mean(seg.len), 'sd'=sd(seg.len)))
+
+pander(sum.var, caption='Variable regions between progressors and non-progressors')
+
+sum.seg = rbind(
+  'Progressors'=do.call(rbind.data.frame, segment.lengths[prog.pt]) %>% summarise( 'n.segments' = length(seg.len), 'median.length'=median(seg.len), 'mean.length'=mean(seg.len), 'sd'=sd(seg.len)),
+  'Non-Progressors'=do.call(rbind.data.frame, segment.lengths[np.pt]) %>% summarise( 'n.segments' = length(seg.len), 'median.length'=median(seg.len), 'mean.length'=mean(seg.len), 'sd'=sd(seg.len)))
+
+pander(sum.var, caption='Segment stats between progressors and non-progressors')
+```
+
+```{r echo=F, message=F, warning=F}
+
+df = rbind(cbind(do.call(rbind.data.frame, var.seg.lengths[prog.pt]), 'status'='Progressor'), cbind(do.call(rbind.data.frame, var.seg.lengths[np.pt]), 'status'='Non-Progressor'))
+lenWT = wilcox.test(subset(df, status == 'Progressor')$seg.len, subset(df, status == 'Non-Progressor')$seg.len)
+
+ggplot(df, aes(status, seg.len)) + geom_jitter(aes(color=status), alpha=0.5, show.legend = F, width=.2) +
+  geom_boxplot(aes(group=status, fill=status), show.legend=F, outlier.colour = NA) + 
+  geom_label(data=subset(df, status == 'Progressor'), aes(status, max(seg.len)-2*sd(seg.len), label=paste('median\n',scientific(median(seg.len),3),sep='') )) +
+  geom_label(data=subset(df, status == 'Non-Progressor'), aes(status, max(seg.len), label=paste('median\n',scientific(median(seg.len),3),sep=''))) +
+  labs(y='Variable segment length', x='', title='Variable regions', caption=paste('Wilcox p-value', signif(lenWT$p.value, digits=3) ))
+
+ggplot(df, aes(status, seg.len)) + 
+  geom_jitter(aes(color=status), show.legend=F) + geom_violin(aes(group=status, fill=status), show.legend=F) +
+  geom_label(data=subset(df, status == 'Progressor'), aes(status, max(seg.len)-2*sd(seg.len),label=paste('median\n',scientific( median(seg.len) , 3),sep='') ))  +
+    geom_label(data=subset(df, status == 'Non-Progressor'), aes(status, max(seg.len), label=paste('median\n',scientific( median(seg.len) , 3),sep=''))) +
+    labs(y='Variable segment length', x='', title='Variable regions', caption=paste('Wilcox p-value', signif(lenWT$p.value, digits=3) ))
+
+```
+
+```{r, echo=F, warning=F, message=F}
+df = rbind(cbind.data.frame('n'=sapply(var.seg.lengths[prog.pt], nrow), 'status'='Progressor'), cbind.data.frame('n'=sapply(var.seg.lengths[np.pt], nrow), 'status'='Non-Progressor'))
+nWT = wilcox.test(subset(df, status == 'Progressor')$n, subset(df, status == 'Non-Progressor')$n)
+
+ggplot(df, aes(status, n, group=status)) + geom_jitter(aes(color=status), width=0.2) +
+  geom_boxplot(aes(fill=status), position = 'dodge', show.legend = F, outlier.colour = NA) +
+  geom_label(data=subset(df, status == 'Progressor'), 
+             aes(status, max(n)-2*sd(n), label=paste('median\n',sprintf("%.1f",median(n)),sep=''), fill=status ), color='white' ) +
+  geom_label(data=subset(df, status == 'Non-Progressor'), 
+             aes(status, max(n)+2*sd(n), label=paste('median\n',sprintf("%.1f",median(n)),sep=''), fill=status), color='white') +
+  labs(y='Number of segments', title="Total number of segments", caption=paste('Wilcox p-value', signif(nWT$p.value, digits=3)) ) +
+  theme(legend.position = 'none')
+
+ggplot(df, aes(status, n, group=status)) + geom_jitter(aes(color=status), width=0.2) +
+  geom_violin(aes(fill=status, group=status)) +
+  geom_label(data=subset(df, status == 'Progressor'), 
+             aes(status, max(n)-2*sd(n), label=paste('median\n',sprintf("%.1f",median(n)),sep=''), fill=status ), color='white' ) +
+  geom_label(data=subset(df, status == 'Non-Progressor'), 
+             aes(status, max(n)+2*sd(n), label=paste('median\n',sprintf("%.1f",median(n)),sep=''), fill=status), color='white') +
+  labs(y='Number of segments', title="Total number of segments", caption=paste('Wilcox p-value', signif(nWT$p.value, digits=3)) ) +
+  theme(legend.position = 'none')
+
+```
+
+### Check variable segment length vs chromosome length
+
+```{r echo=F, message=F, warning=F, fig.height=8, fig.width=20}
 n.var.regions = sapply(var.seg.lengths, nrow)
-wilcox.test(n.var.regions[prog.pt], n.var.regions[np.pt])
+
+#wilcox.test(n.var.regions[prog.pt], n.var.regions[np.pt])
 
@@
 chr.info = read.table('hg19_chrom.txt',sep='\t', header=T)
+chr.info$genome.length = cumsum(as.numeric(chr.info$length))
 
-# Segment lengths, variable regions
-var.seg.lengths = lapply(var.seg.lengths, function(x) merge(x, chr.info, by.x='chr', by.y='chr') )
+var.seg.lengths = lapply(var.seg.lengths, function(x) merge(x, chr.info, by='chr') )
 
 # All segments
-segment.lengths = lapply(segment.lengths, function(x) merge(x, chr.info, by.x='chr', by.y='chr') )
+segment.lengths = lapply(segment.lengths, function(x) merge(x, chr.info, by='chr') )
 
@@
 
+var.seg.lengths = lapply(var.seg.lengths, function(x) cbind(x, 'chr.ratio'=x[['seg.len']]/x[['length']]))
+
+# Segment lengths, variable regions
+vsl = do.call(rbind.data.frame, lapply(patient.data, function(df) {
+  if (is.null(df$variable.regions) || nrow(df$variable.regions) <= 0) return(NULL)
+
+  mc = grep('^D', colnames(df$variable.regions))
+
+  meanCN = sapply(apply(t(df$variable.regions[,mc]), 2, function(x) x[which(x >= 1.1 | x <= 0.9)]), mean)
+  meanCN[is.na(meanCN)] = 0
+
+    cbind.data.frame( 'chr'=df$variable.regions$chrom, 'start'=df$variable.regions$start.pos, 'end'=df$variable.regions$end.pos, meanCN, 'status'= df$info$Status[1])
+  }))
+vsl = vsl[-which(vsl$meanCN == 0),]
+
+vsl$meanCN = vsl$meanCN - mean(vsl$meanCN)
+
+vslm = melt(vsl, id.vars=c('chr','start','end','status'))
+
+chr.info = subset(chr.info, chr %in% c(1:22))
+chr.info$chr = ordered(chr.info$chr, levels=c(1:22))
+vslm$chr =  ordered(vslm$chr, levels=c(1:22))
+vslm = merge(chr.info, vslm, by='chr')
+vslm = vslm %>% rowwise() %>% mutate('genome.start'=genome.length+start, 'genome.end'=genome.length+end)
+
+vslm$status = ifelse(vslm$status == 'P', 'Progressor', 'Non Progressor')
+
+ggplot(vslm, aes(x=length)) + 
+  facet_grid(status~chr) +
+  geom_rect(data=vslm, aes(xmin=start, xmax=end, ymin=0, ymax=value, fill=value)) +
+  scale_fill_gradient(low = "green3", high = "red3", name='') + 
+  labs(title="Variable region gains & losses", x='Chromosomes', y='Mean adjusted segmentation value') +
+  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank(), legend.position='bottom')
+
+```
+
+```{r echo=F, message=F, warning=F}
 # Variable region length by ratio of chromosome
@@
 melted = melt(chr.ratios)
-melted = merge(sum.patient.data[,c('Patient', 'Status')], melted, by.x='Patient', by.y='L1')
+melted = merge(sum.patient.data[,c('Hospital.Research.ID', 'Status')], melted, by.x='Hospital.Research.ID', by.y='L1')
 
@@
 
-ggplot(melted, aes(Status, value, fill=Status)) + geom_boxplot() + 
-  labs(title=paste("Variable region length, p-value", ifelse(chr.ratio.tt$p.value < 0.001, "<0.001", signif(chr.ratio.tt$p.value,2))),y='Segment Length/Chr length')
+ggplot(melted, aes(Status, value, fill=Status)) + geom_jitter(aes(color=Status), width=0.2) + geom_boxplot(outlier.colour = NA) + 
+  geom_label(data=subset(melted, Status == 'P'), aes(x=Status, y=max(value)-sd(value), label=scientific(median(value),3) ), color='white', show.legend = F ) +
+  geom_label(data=subset(melted, Status == 'NP'), aes(x=Status, y=max(value)-sd(value), label=scientific(median(value),3) ), color='white' , show.legend = F) +
+  labs(title="Variable region length",subtitle=paste("p-value", signif(chr.ratio.tt$p.value, digits=3)), y='Segment Length/Chr length')
+
+ggplot(melted, aes(Status, value, fill=Status)) + geom_jitter(aes(color=Status), width=0.2) + geom_violin() + 
+  geom_label(data=subset(melted, Status == 'P'), aes(x=Status, y=max(value)-sd(value), label=scientific(median(value),3) ), color='white' , show.legend = F) +
+  geom_label(data=subset(melted, Status == 'NP'), aes(x=Status, y=max(value)-sd(value), label=scientific(median(value),3) ), color='white' , show.legend = F) +
+  labs(title="Variable region length",subtitle=paste("p-value", signif(chr.ratio.tt$p.value, digits=3)), y='Segment Length/Chr length')
 
@@
 
-So it appears that adjusting for chromosome length (looking at the ratio of the chromosome covered by segments rather than the raw bp) a significant difference in the coverage of the genome by CN segments that are variable between progressors and non: t-test p-value `r signif(chr.ratio.tt$p.value, 2)`.
+So it appears that adjusting for chromosome length (looking at the ratio of the chromosome covered by segments rather than the raw bp) a significant difference in the coverage of the genome by CN segments that are variable between progressors and non: t-test p-value `r signif(chr.ratio.tt$p.value, 3)`.
 Overall bp coverage of the sWGS is `r round(mean(genome.coverage), 2)*100`%  (P `r round(mean(genome.coverage[prog.pt]),2)`, NP `r round(mean(genome.coverage[np.pt]), 2)`).
@@
 
+## Gain vs Loss
+
+```{r echo=F, message=F, warning=F}
+
+df = do.call(rbind.data.frame, lapply(patient.data, function(df) {
+  nohgd = subset(df$info, !Pathology %in% c('HGD','IMC'))$Samplename
+  gain=0;loss=0
+  if ( !is.null(df[['variable.regions']]) ) {
+    nohgd = intersect(nohgd, colnames(df$variable.regions))
+    gain = length(which(apply(as.data.frame(df$variable.regions[,nohgd]), 1, function(x) length(which(x >= 1.1))) > 0))
+    loss = length(which(apply(as.data.frame(df$variable.regions[,nohgd]), 1, function(x) length(which(x <= 0.9))) > 0))
+  }
+
+  cbind.data.frame('gain'=gain, 'loss'=loss, 'status'=unique(df$info$Status))
+}))
+head(df)
+
+ttg = t.test(subset(df, status == 'P')$gain, subset(df, status == 'NP')$gain)
+ttl = t.test(subset(df, status == 'P')$loss, subset(df, status == 'NP')$loss)
+
+ft = fisher.test((df %>% group_by(status) %>% summarise( 'gains'=sum(gain), 'losses'=sum(loss)))[,c('gains','losses')])
+
+ggplot(melt(df), aes(status, value, fill=status)) + facet_grid(.~variable) + geom_bar(stat='identity') +
+  labs(title='Total gains & Losses excluding HGD/IMC', y='counts', x='', caption=paste('Gains and losses are signficantly enriched in Progressors vs Non.\nFisher odds ', round(ft$estimate,2), '(p=', signif(ft$p.value, digits=3), ')', sep='' ))
+
+```
+
 ## How do Cancer Gene Consensus genes map to the variable regions?
 
-```{r, hmm, echo=F, message=F, warning=F, fig.height=100, fig.width=10}
+```{r, hmm, echo=F, message=F, warning=F, fig.height=60, fig.width=10}
 ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
+ccgenes = ccgenes[-which(ccgenes$Mutation.Types == 'T'),]
 
 noLoc = grep(':-',ccgenes$Genome.Location)
+if (length(noLoc) > 0) {
   noLocGenes = ccgenes[noLoc,]
-
   ccgenes = ccgenes[-noLoc,]
+}
 
@@
 
-  status = subset(sum.patient.data, Patient == patient.name)$Status
+  status = subset(sum.patient.data, Hospital.Research.ID == patient.name)$Status
 
@@
 m$brks <- cut(m$value, include.lowest=T, right=F,ordered_result=T,
-                   breaks=c(0, 1, 3, 5, 8, 10),
-                   labels=c('0','1-3', '4-5', '6-8','9-10'))
+                   breaks=c(0, 1, 3, 5, 8, 10, 12, 15),
+                   labels=c('0','1-3', '4-5', '6-8','9-10', '10-12', '>12'))
 
 ## TODO: Tile the variables so that you can see the gains/losses per progressor/non
-subset(m, Gene.Symbols %in% c('CDKN2A', 'TBL1XR1'))
+#subset(m, Gene.Symbols %in% c('CDKN2A', 'TBL1XR1'))
 m = transform(m, Gene.Symbols = reorder(Gene.Symbols, brks))
@@
 ggplot(m,  aes(y=Gene.Symbols, x=CN, fill=brks)) + geom_tile() + facet_grid(.~variable) + 
+    scale_fill_manual(values = myP, 'Patient Counts') + labs(title="All available COSMIC genes", x='', y='')
+
+m2 = melt(gene.p %>% group_by(Gene.Symbols) %>% summarise( 'Progressors' = sum(Progressors), 'Non.Progressors' = sum(Non.Progressors)), id.vars='Gene.Symbols')
+
+m2 = transform(m2, Gene.Symbols = reorder(Gene.Symbols, value))
+
+m2$brks <- cut(m2$value, include.lowest=T, right=F,ordered_result=T,
+                   breaks=c(0, 1, 3, 5, 8, 10, 12, 15),
+                   labels=c('0','1-3', '4-5', '6-8','9-10', '10-12', '>12'))
+
+ggplot(m2,  aes(y=Gene.Symbols, x=variable, fill=brks)) + geom_tile() + #facet_grid(.~variable) + 
     scale_fill_manual(values = myP, 'Patient Counts') + labs(title="All available COSMIC genes", x='', y='')
@@
 m$brks <- cut(m$value, include.lowest=T, right=F,
-                   breaks=c(0, 1, 3, 5, 8, 10),
-                   labels=c('0','1-3', '4-5', '6-8','9-10'))
+                   breaks=c(0, 1, 3, 5, 8, 10, 12, 15),
+                   labels=c('0','1-3', '4-5', '6-8','9-10', '10-12', '>12'))
 
 ggplot(m,  aes(y=Gene.Symbols, x=CN, fill=brks)) + geom_tile(color='white') + facet_grid(.~variable) + 
+    scale_fill_manual(values=myP, 'Patient Counts') + 
+    labs(title="Top 50 genes by number of patients", x='', y='')
+
+gene.p50 = gene.p %>% group_by(Gene.Symbols) %>% summarise( 'Progressors' = sum(Progressors), 'Non.Progressors' = sum(Non.Progressors))
+
+m2 = melt(subset(gene.p50, Gene.Symbols %in% gc[1:50, 'Gene.Symbols']), id.vars=c('Gene.Symbols'))
+
+m2 = transform(m2, Gene.Symbols = reorder(Gene.Symbols, value))
+
+m2$brks <- cut(m2$value, include.lowest=T, right=F, ordered_result=T,
+                   breaks=c(0, 1, 3, 5, 8, 10, 12, 15),
+                   labels=c('0','1-3', '4-5', '6-8','9-10', '10-12', '>12'))
+
+ggplot(m2,  aes(y=Gene.Symbols, x=variable, fill=brks)) + geom_tile(color='white') + 
     scale_fill_manual(values=myP, 'Patient Counts') + 
@@
 
-pvnp = data.frame(matrix(nrow=length(subset(sum.patient.data, Status == 'P')$Patient), ncol=2, dimnames=list(subset(sum.patient.data, Status == 'P')$Patient, c('hgd','be'))))    
-for (name in subset(sum.patient.data, Status == 'P')$Patient) {
+pvnp = data.frame(matrix(nrow=length(subset(sum.patient.data, Status == 'P')$Hospital.Research.ID), ncol=2, dimnames=list(subset(sum.patient.data, Status == 'P')$Hospital.Research.ID, c('hgd','be'))))
+for (name in subset(sum.patient.data, Status == 'P')$Hospital.Research.ID) {
   var.regions = patient.data[[name]]$variable.regions
@@
 
-tt = t.test(pvnp$hgd, pvnp$be)
+#tt = t.test(pvnp$hgd, pvnp$be)
 pvnp = cbind(pvnp, (pvnp %>% rowwise() %>% mutate(total=hgd+be, hgd.ratio=round(hgd/(hgd+be),3) ))[c('total','hgd.ratio')])
 
-#fisher.test(p, simulate.p.value = T)
+ft = fisher.test( pvnp[complete.cases(pvnp[,c('hgd','be')]), c('hgd','be')], simulate.p.value = T)
 
@@
        caption=paste("Number of variable regions found in HGD vs non-HGD samples of progessor patients.
-There's no significant difference between them (t-test p=",signif(tt$p.value,2),').
+There's a significant difference between them (Fisher's p=",signif(ft$p.value,3),').
 Not all patients have HGD samples (NAs).
-There is also no significant inverse correlation (r= ',signif(ct$estimate, 2),') between the number/ratio of variable regions found in HGD samples vs non (Pearsons p=',signif(ct$p.value, 2),')',  sep=''))
+There is no significant correlation (r= ',signif(ct$estimate, 2),') between the number/ratio of variable regions found in HGD samples vs non (Pearsons p=',signif(ct$p.value, 3),')',  sep=''))
 

### 2017-07-19 RegressionPvNP.Rmd
...

@@
 library(ggplot2)
+library(ggrepel)
 library(ggfortify)
@@
 library(plyr)
+library(dplyr)
 library(pander)
@@
 
-includeDemo = FALSE
+includeDemo = FALSE # demographics
 
@@
 
-## Patient info file
-patient.file = grep('All_patient_info.txt', data.files, value=T)
+## Hospital.Research.ID info file
+patient.file = grep('All_patient_info.xlsx', data.files, value=T)
 if (length(patient.file) != 1)
   stop(paste("Missing/too many patient info file(s) in", data))
-demo.file = grep('Demographics_full.txt', data.files, value=T)
-
-patient.info = read.patient.info(patient.file, demo.file, set=dataset)
-patient.info$Patient = gsub("/", "_", patient.info$Patient)
+demo.file = grep('Demographics_full.xlsx', data.files, value=T)
+
+patient.info = read.patient.info(patient.file, demo.file, set=dataset)$info
 head(patient.info)
 
-validation.patient.info = read.patient.info(patient.file, demo.file, set='Test')
-validation.patient.info = arrange(validation.patient.info, Status, Patient, Endoscopy.Year, Pathology)
+validation.patient.info = read.patient.info(patient.file, demo.file, set='Test')$info
+validation.patient.info = arrange(validation.patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
 sum.validation.info = summarise.patient.info(validation.patient.info)
 
-patient.info = arrange(patient.info, Status, Patient, Endoscopy.Year, Pathology)
+patient.info = arrange(patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
 sum.patient.data = summarise.patient.info(patient.info)
@@
 
-sum.patient.data = as.data.frame(subset(sum.patient.data, Patient %in% names(patient.data))) ## For now
+sum.patient.data = as.data.frame(subset(sum.patient.data, Hospital.Research.ID %in% names(patient.data))) ## For now
 
@@
 
-pander(sum.patient.data[,c('Patient', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Training patient set')
-
-pander(sum.validation.info[,c('Patient', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Validation patient set')
+pander(sum.patient.data[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Training patient set')
+
+pander(sum.validation.info[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Validation patient set')
 ```
@@
 
-```{r labelsPNP, echo=F, message=F, include=F}
+```{r labelsPNP, echo=F, message=F, include=T}
 ## binomial: dysplasia 1, BE 0
@@
   df$info = arrange(df$info, Endoscopy.Year, Pathology)
-  cbind(df$info[,c('Patient','Samplename')])
+  cbind(df$info[,c('Hospital.Research.ID','Samplename')])
 }))
@@
 
-validation.df = t(mergedDfValidation[, grep('^D', colnames(mergedDfValidation))])
-dim(validation.df)
+unit.var <- function(x) {
+  if (sd(x) == 0) return(x)
+  return((x-mean(x))/sd(x) )
+}
+
+#dysplasia.df = apply(dysplasia.df, 2, function(x) x-median(x))
+dysplasia.df = apply(dysplasia.df, 2, unit.var)
 
@@
   # t.test(
-  # complexity.measures[subset(sum.patient.data, Status == 'P')$Patient],
-  # complexity.measures[subset(sum.patient.data, Status == 'NP')$Patient]
+  # complexity.measures[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID],
+  # complexity.measures[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID]
   # )
+
   score.cx <-function(pd) {
-  #  print(pd$info)
-    len.var = (var(pd$seg.vals$end.pos - pd$seg.vals$start.pos))/1e12
-    extreme.values = max( apply(as.matrix(pd$seg.vals[,grep('^D', colnames(pd$seg.vals))]), 2, function(col) {
-      length(which(col >= (median(col)+2*sd(col)))) + length(which(col <= (median(col)-2*sd(col)))) 
-    }) )
-
-    (len.var + extreme.values)
-  }
-
-  complexity.measures = sapply(patient.data, score.cx)
-
-  dysplasia.df = cbind(dysplasia.df, 'cx.measure'=NA)
-  for (sample in rownames(dysplasia.df)) {
-    dysplasia.df[sample, 'cx.measure'] = complexity.measures[subset(patient.info, Samplename == sample)$Patient]
+    samples = grep('^D', colnames(pd$seg.vals), value=T)
+    len.var = var(pd$seg.vals$end.pos - pd$seg.vals$start.pos) # length won't vary between samples
+
+    #extremes = apply(as.matrix(pd$seg.vals[,samples]), 2, function(col) {
+    #  length(which(col >= (median(col)+2*sd(col)))) + length(which(col <= (median(col)-2*sd(col)))) 
+    #})
+
+    num.var.reg = vector(mode='integer', length=length(samples))
+    if (!is.null(pd$variable.regions) && nrow(pd$variable.regions) > 0) {
+      num.var.reg = apply(as.matrix(pd$variable.regions[,samples]), 2, function(col) {
+        length(which(col >= 1.1 | col <= 0.9))
+      })
+    }
+
+    sample.var = apply(as.matrix(pd$seg.vals[, samples]), 2, var)
+    t(cbind(len.var, num.var.reg, sample.var))
+  }
+
+  complexity.measures = lapply(patient.data, score.cx)
+
+  dysplasia.df = cbind(dysplasia.df, 'segment.length'=0, 'num.var.regions'=0, 'sample.variance'=0)
+  for (pt in names(complexity.measures)) {
+    for (sample in colnames(complexity.measures[[pt]])) {
+      dysplasia.df[sample,c('segment.length', 'num.var.regions', 'sample.variance')] = as.numeric(complexity.measures[[pt]][,sample])
+    }
+    #for (sample in rownames(dysplasia.df)) {
+      #dysplasia.df[sample, ] = 
+        #complexity.measures[subset(patient.info, Samplename == sample)$Hospital.Research.ID]
+    #}
   }
   # norm to 1 to keep within the magnitude of the genomic data
-  cx = dysplasia.df[,'cx.measure'] 
-  dysplasia.df[,'cx.measure'] = (log(cx)/log(max(cx)))
+  scale.to.1<-function(x) {
+    (x-min(x))/(max(x)-min(x))
+  }
+
+  dysplasia.df[,c('segment.length', 'num.var.regions', 'sample.variance')] = apply(dysplasia.df[,c('segment.length', 'num.var.regions', 'sample.variance')], 2, scale)
+
+  #cx = dysplasia.df[,'cx.measure'] 
+  #dysplasia.df[,'cx.measure'] = (log(cx)/log(max(cx)))
+  #dysplasia.df[,'cx.measure'] = dysplasia.df[,'cx.measure'] - median(dysplasia.df[,'cx.measure'])
   #dysplasia.df[,'cx.measure'] = ((cx - min(cx))/(max(cx)-min(cx)))
-
-  complexity.measures = sapply(validation.patient.data, score.cx)
-
-  validation.df = cbind(validation.df, 'cx.measure'=NA)
-  for (sample in rownames(validation.df)) {
-    validation.df[sample, 'cx.measure'] = complexity.measures[subset(validation.patient.info, Samplename == sample)$Patient]
-  }
-  # norm to 1 to keep within the magnitude of the genomic data
-  cx = validation.df[,'cx.measure'] 
-  validation.df[,'cx.measure'] = (log(cx)/log(max(cx)))
 
@@
       #print(sets/rowSums(sets))
-      s = sample(rep(seq(5), length = length(unique(df$Patient))))
-      df2 = merge(df, cbind('Patient'=unique(df$Patient), 'tmpgrp'=s), by="Patient")
+      s = sample(rep(seq(5), length = length(unique(df$Hospital.Research.ID))))
+      df2 = merge(df, cbind('Hospital.Research.ID'=unique(df$Hospital.Research.ID), 'tmpgrp'=s), by="Hospital.Research.ID")
       df[[grpCol]] = df2$tmpgrp
@@
 
-  s = sample(rep(seq(splits), length = length(unique(pts$Patient))))
-  patients = merge(pts, cbind('Patient'=unique(pts$Patient), 'group'=s), by="Patient")  
+  s = sample(rep(seq(splits), length = length(unique(pts$Hospital.Research.ID))))
+  patients = merge(pts, cbind('Hospital.Research.ID'=unique(pts$Hospital.Research.ID), 'group'=s), by="Hospital.Research.ID")  
   colnames(patients)[3] = c('fold.1')
@@
   for (i in 2:n) {
-    s = sample(rep(seq(5), length = length(unique(pts$Patient))))
-    patients = merge(patients, cbind('Patient'=unique(patients$Patient), 'group'=s), by="Patient")  
+    s = sample(rep(seq(5), length = length(unique(pts$Hospital.Research.ID))))
+    patients = merge(patients, cbind('Hospital.Research.ID'=unique(patients$Hospital.Research.ID), 'group'=s), by="Hospital.Research.ID")  
     foldcol = grep('group',colnames(patients))
@@
       # pre-spec lambda seq
-      fitCV <- glmnet(training, y[-test.rows], lambda=lambda, alpha=a, family='binomial', ...) 
+      fitCV = glmnet(training, y[-test.rows], lambda=lambda, alpha=a, family='binomial', ...) 
       # autoplot(fit) + theme(legend.position="none")
@@
 ```{r xvalpt, message=F, warning=F, echo=F, fig.height=16, fig.width=16}
+# dysplasia.df = cbind(dysplasia.df, 'sl.nvr'=0)
+# dysplasia.df[, 'sl.nvr'] = scale(dysplasia.df[, "segment.length"] + dysplasia.df[, 'num.var.regions'])
+# 
+# dysplasia.df = cbind(dysplasia.df, 'sl.nvr.sv'=0)
+# dysplasia.df[, 'sl.nvr.sv'] = scale(dysplasia.df[, "segment.length"] + dysplasia.df[, 'num.var.regions'] + dysplasia.df[, 'sample.variance'])
+# 
+# dysplasia.df = cbind(dysplasia.df, 'sl.sv'=0)
+# dysplasia.df[, 'sl.sv'] = scale(dysplasia.df[, "segment.length"] + dysplasia.df[, 'sample.variance'])
+# 
+# dysplasia.df = cbind(dysplasia.df, 'nvr.sv'=0)
+# dysplasia.df[, 'sl.sv'] = scale(dysplasia.df[, "num.var.regions"] + dysplasia.df[, 'sample.variance'])
+
 nl = 1000
 
-coefs = list(); plots = list(); performance.at.1se = list()
+coefs = list(); plots = list(); performance.at.1se = list(); models = list();
 folds = 10; splits = 5 
-alpha.values = c(0,0.5,0.7,0.8,0.9,1)
+alpha.values = c(0, 0.5,0.7,0.8,0.9,1)
 
@@
   for (a in alpha.values) {
-    fit0 <- glmnet(dysplasia.df, labels, alpha=a, nlambda=nl, family='binomial', standardize=T) # all patients
+    fit0 <- glmnet(dysplasia.df, labels, alpha=a, nlambda=nl, family='binomial', standardize=F) # all patients
     #autoplot(fit0, xvar='lambda', main=paste('fit0, all samples, alpha=',a,sep='')) + theme(legend.position='none')
@@
     }
-    cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=l, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', standardize=T)
+    cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=l, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', standardize=F)
 
@@
     performance.at.1se[[as.character(a)]] = subset(cv.patient$lambdas, lambda.1se == T)
-  }
-  save(plots, coefs, performance.at.1se, dysplasia.df, file=file)
+    models[[as.character(a)]] = fit0
+  }
+  save(plots, coefs, performance.at.1se, dysplasia.df, models, file=file)
 }
@@
 
+```{r echo=F, message=F, warning=F}
+featureStability = rowSums(coefs[[select.alpha]][,-1])/(splits*folds)
+```
+
 ```{r echo=F, warning=F, message=F, eval=F}
@@
 
-```{r, warning=F, message=F}
+```{r, warning=F, message=F, eval=T}
 #topF = names(which((coef.stable[['1']]) >= 0.75))
@@
 gl = cbind.data.frame(
-  'P.gain' = apply(features.in.prog, 2, function(x) sum( as.integer(x >= 1.1) ) ), 
-  'P.loss' = apply(features.in.prog, 2, function(x) sum( as.integer(x <= 0.9) ) ),
-  'NP.gain' = apply(features.nonprog, 2, function(x) sum( as.integer(x >= 1.1) ) ),
-  'NP.loss' = apply(features.nonprog, 2, function(x) sum( as.integer(x <= 0.9) ) ) )
+  'P.gain' = apply(features.in.prog, 2, function(x) sum( as.integer(x > 0) ) ), 
+  'P.loss' = apply(features.in.prog, 2, function(x) sum( as.integer(x < 0) ) ),
+  'NP.gain' = apply(features.nonprog, 2, function(x) sum( as.integer(x > 0) ) ),
+  'NP.loss' = apply(features.nonprog, 2, function(x) sum( as.integer(x < 0) ) ) )
 
 allgl = cbind.data.frame(
-  'ALL.P.gain' = apply(dysplasia.df[names(which(labels == 1)), genomeCols], 2, function(x) sum( as.integer(x >= 1.1) ) ), 
-  'ALL.P.loss' = apply(dysplasia.df[names(which(labels == 1)),genomeCols], 2, function(x) sum( as.integer(x <= 0.9) ) ),
-  'ALL.NP.gain' = apply(dysplasia.df[names(which(labels == 0)),genomeCols], 2, function(x) sum( as.integer(x >= 1.1) ) ), 
-  'ALL.NP.loss' = apply(dysplasia.df[names(which(labels == 0)),genomeCols], 2, function(x) sum( as.integer(x <= 0.9) ) ) )
+  'ALL.P.gain' = apply(dysplasia.df[names(which(labels == 1)), genomeCols], 2, function(x) sum( as.integer(x > 0) ) ), 
+  'ALL.P.loss' = apply(dysplasia.df[names(which(labels == 1)),genomeCols], 2, function(x) sum( as.integer(x < 0) ) ),
+  'ALL.NP.gain' = apply(dysplasia.df[names(which(labels == 0)),genomeCols], 2, function(x) sum( as.integer(x > 0) ) ), 
+  'ALL.NP.loss' = apply(dysplasia.df[names(which(labels == 0)),genomeCols], 2, function(x) sum( as.integer(x < 0) ) ) )
 
@@
 
-# Gains/losses enriched in the top features?  Appears so
+# Gains/losses enriched in the top features?  Appears not
 (fisher.test(rbind(colSums(gl[,c('P.gain','P.loss')]),colSums(allgl[-feat,c('ALL.P.gain','ALL.P.loss')])  )) )
@@
 topFSegments = as.data.frame(apply(topFSegments, 2, as.numeric))
-topFSegments$chr = factor(topFSegments$chr, levels=chromInfo$chr)
+topFSegments$chr = factor(topFSegments$chr, levels=droplevels(chromInfo$chr[1:22]), ordered=T)
 topFSegments = arrange(topFSegments, chr, start)
@@
 
-grid.arrange(ggplot(topFRegions, aes(reorder(chr, -adj.info.content), totalbp/length, fill=ai.cut)) + ylim(0, 0.5) +
+grid.arrange(
+  #ggplot(topFRegions, aes(reorder(chr, -adj.info.content), totalbp/length, fill=ai.cut)) + ylim(0, 0.5) +
+  ggplot(topFRegions, aes(chr, totalbp/length, fill=ai.cut)) + ylim(0, 0.5) +
     labs(y='Feature bp/total chr length', title='Ratio of covered base pairs', x='chr') +
@@
   # Not related to the length of the segment
-  pander(with(topFSegments, cor.test(genes, (end-start))), caption="Number of genes per segment is not related to the segment length.")
+  pander(with(topFSegments, cor.test(genes, (end-start))), caption="Number of genes per segment may be related to the segment length.")
 }, error = function(e) {
@@
 
+```{r echo=F, message=F, warning=F, eval=F}
 #### Correlations
-
-```{r echo=F, message=F, warning=F}
 library(Hmisc)
@@
 
-nrow(rc)
-head(rc)
+#nrow(rc)
+#head(rc)
 
@@
 
-head(rc)
+#head(rc)
 
@@
 rc = arrange(rc, row.chr, col.chr, row.start, col.start)
-head(rc)
+#head(rc)
 
@@
 
-head(rc1)
+#head(rc1)
 
@@
 ```{r echo=F, warning=F, message=F,eval=T, fig.height=8, fig.width=8}
+
 hazards = apply( exp(t(dysplasia.df[, rownames(coefs[[select.alpha]])]) *  coefs[[select.alpha]][,1]), 1, function(x) {
@@
 ch$label = rownames(ch)
-library(ggrepel)
+ch$stability = featureStability
+
 ggplot(ch, aes(coef, hazards, col=hazards > quantile(hazards)[['75%']])) + 
@@
   geom_point(alpha=0.5) + 
-  geom_text_repel(aes(x=coef+10, label=ifelse(hazards > quantile(hazards)[['75%']], label, '')),show.legend=F) +
+  geom_text_repel(aes(x=coef, label=ifelse(hazards > quantile(hazards)[['75%']], label, '')),show.legend=F) +
   labs(title='Top features coef vs hazard') + theme(legend.position = 'bottom')
@@
 
-```{r echo=F, warning=F, message=F,eval=T, fig.height=12, fig.width=8}
+```{r echo=F, warning=F, message=F,eval=F, fig.height=12, fig.width=8}
+ch = arrange(ch, -stability, -hazards, -coef)
+ch$hazards = round(ch$hazards, 3)
+ch$coef = round(ch$coef, 3)
+write.table(ch[c('label', 'stability','hazards','coef')], row.names = F, quote=F, sep="\t", file='features.txt')
+
 ch = arrange(ch, -hazards)
-
-ch$hazards = round(ch$hazards, 2)
-ch$coef = round(ch$coef, 2)
 
@@
 
-  reg = cbind.data.frame( 'value'=dysplasia.df[ , r ], labels)
+  reg = cbind.data.frame( 'value'=dysplasia.df[ ,r], labels)
   reg$labels = factor(reg$labels)
@@
   reg$region = r
-  reg$hazard = x[['hazards']]
-  reg$coef = x[['coef']]
+  reg$hazard = as.numeric(x[['hazards']])
+  reg$coef = as.numeric(x[['coef']])
   return(reg)
@@
   reg = melt.by.reg(x)
-  pv = ifelse(tt$p.value < .001, '< .001', format.pval(unique(reg$p.value)))
+
+  pv = ifelse(reg$pval[1] < .001, '< .001', format.pval(unique(reg$pval[1])))
 
@@
 #    labs(x='seg. value', title=r, subtitle=paste('haz=',x[['hazards']], ' coef=', x[['coef']], ' MW pval: ',pv,sep=''))
-  ggplot(reg, aes(x=labels, y=value, group=labels, fill=labels) ) + geom_jitter(color='grey', show.legend = F) + geom_violin(show.legend = F, alpha=0.5) +  ylim(0.75, 1.2) +
+  ggplot(reg, aes(x=labels, y=value, group=labels, fill=labels) ) + geom_jitter(color='grey', show.legend = F) + geom_violin(show.legend = F, alpha=0.5) +
     geom_hline(yintercept = median(reg$value) + sd(reg$value)) + geom_hline(yintercept = median(reg$value) - sd(reg$value)) +
-    labs(x='', y='seg. value', title=r, subtitle=paste('haz=',x[['hazards']], ' coef=',x[['coef']], '\nMW pval ',pv,sep=''))
-}
+    labs(x='', y='seg. value', title=x[['label']], subtitle=paste('haz= ',x[['hazards']], ' coef= ',x[['coef']], '\nMW pval ',pv,sep=''))
+}
+
+ch = arrange(ch, -hazards)
 
 regions = apply(ch, 1, melt.by.reg)
-
-do.call(grid.arrange, c(lapply(regions, function(reg) {
-  se = 2
-  x = rbind.data.frame('P'= cbind( 'gain' = length(which(with(subset(reg, labels == 'P'), value >= median(value) + sd(value)*se))),
-                               'loss' = length(which(with(subset(reg, labels == 'P'), value <= median(value) - sd(value)*se))) ),
-
-                  'NP' = cbind( 'gain' = length(which(with(subset(reg, labels == 'NP'), value >= median(value) + sd(value)*se))),
-                                'loss' = length(which(with(subset(reg, labels == 'NP'), value <= median(value) - sd(value)*se))) ) )
+names(regions) = ch$label
+
+do.call(grid.arrange, c(lapply(regions[1:15], function(reg) {
+  print(reg$region[1])
+  cols = c('indianred3','slateblue2')
+  if ( as.numeric(reg$coef[1]) < 0) cols = c('green3', 'darkviolet')
+
+  ggplot(reg, aes(labels, value*coef, fill=labels, color=labels)) + 
+    geom_jitter(aes(labels, value), width=0.05, color='yellow4', alpha=0.3) +
+    geom_jitter(aes(labels, value*coef, color=labels), width=0.05, alpha=0.5) +
+    geom_text_repel(data=subset(reg, value >= quantile(reg$value)[['100%']]), aes(labels, value, label=round(value, 3)), color='yellow4') +
+    geom_text_repel(data=subset(reg, value >= quantile(reg$value)[['100%']]), aes(labels, value*coef, label=round(value, 3),color=labels)) + geom_violin(alpha=0.2) + 
+    scale_fill_manual(values=cols) + scale_color_manual(values=cols) +
+    theme(legend.position = 'none') + labs(title=unique(reg$region)[1], x='')
+})))
+
+    #geom_violin(aes(labels, value, fill=labels), alpha=0.5) + theme(legend.position = 'none')
+
+#  x = rbind.data.frame('P'= cbind( 'gain' = length(which(with(subset(reg, labels == 'P'), value > 0))),
+#                               'loss' = length(which(with(subset(reg, labels == 'P'), value < 0))) ),
+
+#                        'NP' = cbind( 'gain' = length(which(with(subset(reg, labels == 'NP'), value > 0))),
+#                                'loss' = length(which(with(subset(reg, labels == 'NP'), value < 0))) ) )
+
+    # x = rbind.data.frame('P'= cbind( 'gain' = length(which(with(subset(reg, labels == 'P'), value >= median(value) + sd(value)*se))),
+  #                              'loss' = length(which(with(subset(reg, labels == 'P'), value <= median(value) - sd(value)*se))) ),
+  # 
+  #                 'NP' = cbind( 'gain' = length(which(with(subset(reg, labels == 'NP'), value >= median(value) + sd(value)*se))),
+  #                               'loss' = length(which(with(subset(reg, labels == 'NP'), value <= median(value) - sd(value)*se))) ) )
   # x = rbind.data.frame('P'= cbind( 'high' = length(which(with(subset(reg, labels == 'P'), value >= 1.1))),
@@
 
-  cols = c('indianred3','slateblue2')
-  if ( as.numeric(reg$coef[1]) < 0) cols = c('plum', 'skyblue3')
-  #ggplot(melt(as.matrix(x)), aes(Var1, value, fill=Var2) ) + geom_bar(stat='identity', position='dodge') + labs(x='', y='', title=unique(reg$region)[1]) + scale_fill_manual(values=cols)
-
-  ggplot(melt(as.matrix(rowSums(x))), aes(Var1, value, fill=Var1)) + geom_bar(stat='identity', show.legend = F)+ labs(x='', y='', title=unique(reg$region)[1]) + scale_fill_manual(values=cols)
-
-})))
+  #cols = c('indianred3','slateblue2')
+  #if ( as.numeric(reg$coef[1]) < 0) cols = c('plum', 'skyblue3')
+
+  #ggplot(melt(as.matrix(x)), aes(Var2, value, fill=Var1)) + facet_grid(~Var1) +
+  #  geom_col( show.legend = F) + labs(x='', y='', title=unique(reg$region)[1]) + scale_fill_manual(values=cols)
+
+#})))
 
@@
   for (a in alphas) {
-    fitg <- glmnet(x, y, alpha=a, nlambda=nl, family='binomial') # all patients
+    fitg <- glmnet(x, y, alpha=a, nlambda=nl, family='binomial', standardize = F) # all patients
 
@@
 
-    cvp = crossvalidate.by.patient(x, y, fit=fitg, lambda=l, a=a, ...)
+    cvp = crossvalidate.by.patient(x, y, fit=fitg, lambda=l, a=a, pts=sets, standardize=F)
     plots[[as.character(a)]] = arrangeGrob(cvp$plot+ggtitle('Classification'), cvp$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
@@
     # all patients
-    fitNoHGD <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=nl) 
-    cv.nohgd = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitNoHGD$lambda, pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fitNoHGD)
+    fitNoHGD <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=nl, standardize = F) 
+    cv.nohgd = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitNoHGD$lambda, pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fitNoHGD, standardize = F)
 
@@
   for (a in alpha.values) {
-    fitNoLGD <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=nl) # all patients
-
-    cv.nolgd = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitNoLGD$lambda, pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fitNoLGD)
+    fitNoLGD <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=nl, standardize = F) # all patients
+
+    cv.nolgd = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitNoLGD$lambda, pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fitNoLGD, standardize = F)
 
@@
 
+```{r manyfew, echo=T, message=F, warning=F, fig.height=16, fig.width=16, eval=F}
 ## Compare progressors with many samples vs few
-
-```{r manyfew, echo=T, message=F, warning=F, fig.height=16, fig.width=16}
 m = median(subset(sum.patient.data, Status == 'P')$total.samples)
-nonps = as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Status == 'NP')$Patient], function(df) df$info$Samplename)))
+nonps = as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID], function(df) df$info$Samplename)))
 
@@
   # Few
-  samples = c(nonps, as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Status == 'P' & total.samples < m)$Patient], function(pt) pt$info$Samplename))))
+  samples = c(nonps, as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Status == 'P' & total.samples < m)$Hospital.Research.ID], function(pt) pt$info$Samplename))))
 
@@
   for (a in alpha.values) {
-    fitF <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, nlambda=nl, family='binomial') # all patients
-    cv.back.few = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitF$lambda, pts=subset(pts, Samplename %in% samples), splits=5, nfolds=20, a=a, fit=fitF, minR=0.1)
+    fitF <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, nlambda=nl, family='binomial', standardize = F) # all patients
+    cv.back.few = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitF$lambda, pts=subset(pts, Samplename %in% samples), splits=5, nfolds=folds, a=a, fit=fitF, minR=0.1, standardize = F)
     #few.plots[[as.character(a)]] = cv.back.few$plot + ggtitle(paste('alpha=',a,sep=''))
@@
 } else {
-  samples = c(nonps, as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Status == 'P' & total.samples >= m)$Patient], function(pt) pt$info$Samplename))))
+  samples = c(nonps, as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Status == 'P' & total.samples >= m)$Hospital.Research.ID], function(pt) pt$info$Samplename))))
 
@@
   for (a in alpha.values) {
-    fitM <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, nlambda=nl, family='binomial') # all patients
-
-    cv.back.many = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitM$lambda, pts=subset(pts, Samplename %in% samples), splits=5, nfolds=20, a=a, fit=fitM)
+    fitM <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, nlambda=nl, family='binomial', standardize = F) # all patients
+
+    cv.back.many = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitM$lambda, pts=subset(pts, Samplename %in% samples), splits=5, nfolds=20, a=a, fit=fitM, standardize = F)
 
@@
 
-  r.labels = unlist(apply(rand.patients[,c('Patient','Status')],1, function(st) {
-    info = patient.data[[ st['Patient'] ]]$info
+  r.labels = unlist(apply(rand.patients[,c('Hospital.Research.ID','Status')],1, function(st) {
+    info = patient.data[[ st['Hospital.Research.ID'] ]]$info
     lbl = rep(as.integer( st['Status'] == 'P' ), nrow(info))
@@
   for (a in c(0,0.5,1)) {
-    fitR <- glmnet(r.dysplasia.df, r.labels, alpha=a, nlambda=nl, family='binomial') # all patients
-
-    cv.R = crossvalidate.by.patient(x=r.dysplasia.df, y=r.labels, lambda=fitR$lambda, pts=pts, a=a, nfolds=folds, splits=splits, fit=fitR)
+    fitR <- glmnet(r.dysplasia.df, r.labels, alpha=a, nlambda=nl, family='binomial', standardize = F) # all patients
+
+    cv.R = crossvalidate.by.patient(x=r.dysplasia.df, y=r.labels, lambda=fitR$lambda, pts=pts, a=a, nfolds=folds, splits=splits, fit=fitR, standardize = F)
 
@@
     print(pt)
-    tmp.patient.data = patient.data[subset(sum.patient.data, Patient != pt)$Patient]
+    tmp.patient.data = patient.data[subset(sum.patient.data, Hospital.Research.ID != pt)$Hospital.Research.ID]
     samples = as.vector(unlist(sapply(tmp.patient.data, function(df) df$info$Samplename )))
@@
     a = select.alpha
-    fitLOO <- glmnet(training, labels[train.rows], alpha=a, family='binomial', nlambda=nl) # all patients
+    fitLOO <- glmnet(training, labels[train.rows], alpha=a, family='binomial', nlambda=nl, standardize=F) # all patients
     #l = sort(c(fitLOO$lambda, seq(exp(-5), exp(-10), -0.001),
@@
     cv = crossvalidate.by.patient(x=training, y=labels[train.rows], lambda=l, 
-                                  pts=subset(pts, Samplename %in% samples), a=a, nfolds=10, splits=5, fit=fitLOO)
+                                  pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fitLOO, standardize=F)
     #plots[[pt]] = cv$plot + ggtitle(pt)
@@
     } else {
-      warning(paste("Patient", pt, "did not have a 1se"))
+      warning(paste("Hospital.Research.ID", pt, "did not have a 1se"))
     }
@@
 ```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=10, fig.width=6}
-preds = unlist(lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], function(df) df$Prediction))
-np.preds = unlist(lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Patient], function(df) df$Prediction))
+preds = unlist(lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], function(df) df$Prediction))
+np.preds = unlist(lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID], function(df) df$Prediction))
 
@@
 
-df = cbind('Specificity'=rev(roc50$specificities), 'Sensitivity'=rev(roc50$sensitivities))
+df = cbind.data.frame('Specificity'=rev(roc50$specificities), 'Sensitivity'=rev(roc50$sensitivities))
 g1 = ggplot(df, aes(Specificity, Sensitivity)) + geom_line() +
   scale_x_reverse() + 
-  geom_label(data=t(as.data.frame(round(pROC::coords(roc50, "best"),2))), aes(x=specificity, y=sensitivity, label=paste('prob=',threshold, ' (', specificity*100, '%, ', sensitivity*100,'%)\nAUC:', sprintf("%.2f",roc50$auc*100), '%', sep='')), nudge_x=.25) +
+  geom_label(data=t(as.data.frame(round(pROC::coords(roc50, "best"),2))), 
+             aes(x=specificity, y=sensitivity, label=paste('prob=',threshold, ' (FPR ', specificity*100, '%, TPR ', sensitivity*100,'%)\nAUC:', sprintf("%.2f",roc50$auc*100), '%', sep='')), nudge_x=.25) +
   labs(title='Probability threshold = 0.5', x='Specificity (FPR)', y='Sensitivity (TPR)')
@@
 
-df = cbind('Specificity'=rev(roc$specificities), 'Sensitivity'=rev(roc$sensitivities))
+df = cbind.data.frame('Specificity'=rev(roc$specificities), 'Sensitivity'=rev(roc$sensitivities))
 g2 = ggplot(df, aes(Specificity, Sensitivity)) + geom_line() +
   scale_x_reverse() + 
-  geom_label(data=t(as.data.frame(round(pROC::coords(roc, "best"),2))), aes(x=specificity, y=sensitivity, label=paste('prob=',threshold, ' (', specificity*100, '%, ', sensitivity*100,'%)\nAUC:', sprintf("%.2f",roc$auc*100), '%', sep='')), nudge_x=.25) +
+  geom_label(data=t(as.data.frame(round(pROC::coords(roc, "best"),2))), aes(x=specificity, y=sensitivity, label=paste('prob=',threshold, ' (FPR ', specificity*100, '%, TPR ', sensitivity*100,'%)\nAUC:', sprintf("%.2f",roc$auc*100), '%', sep='')), nudge_x=.25) +
   labs(title='Probability threshold calculated by ROC', x='Specificity (FPR)', y='Sensitivity (TPR)')
@@
 # For HGD/IMC samples, progressors only
-hgd = sapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], function(pt) {
+hgd = sapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], function(pt) {
   hgd = subset(pt, Pathology %in% c('HGD', 'IMC'))
@@
 
-hgd = do.call(rbind, lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], function(pt) {
+hgd = do.call(rbind, lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], function(pt) {
   hgd = subset(pt, Pathology %in% c('HGD', 'IMC'))
@@
 }))
-rownames(hgd) = subset(sum.patient.data, Status == 'P')$Patient
+rownames(hgd) = subset(sum.patient.data, Hospital.Research.ID %in% rownames(hgd))$Patient
+
 hgd = subset(hgd, total.hgd.samples > 0) # Some patients had no hgd samples
@@
 global.by.path = do.call(rbind.data.frame,
-                         lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], function(pt) {
+                         lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], function(pt) {
   final.endo = subset(pt, Endoscopy.Year == max(Endoscopy.Year))
@@
 global.by.path = colSums(global.by.path)
+
 global.path = as.data.frame(matrix(global.by.path, byrow=T, nrow=4, dimnames=list(c('HGD','LGD','ID','NDBE'),c('pred', 'total'))))
@@
 ```{r echo=F, message=F, warning=F}
-lgd = do.call(rbind, lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], function(pt) {
+lgd = do.call(rbind, lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], function(pt) {
   e = subset(pt, Pathology  == 'LGD')
@@
 }))
+rownames(lgd) = subset(sum.patient.data, Hospital.Research.ID %in% rownames(lgd))$Patient
+
 lgd = subset(lgd, total.lgd > 0)
@@
 # All samples
-
 all.predictions = sapply(pg.samp, function(pt) {
@@
 
-all.predictions$Status = ifelse (rownames(all.predictions) %in% subset(sum.patient.data, Status == 'P')$Patient, 'P', 'NP')
+all.predictions$Status = ifelse (rownames(all.predictions) %in% subset(sum.patient.data, Status == 'P')$Hospital.Research.ID, 'P', 'NP')
+
+rownames(all.predictions) = subset(sum.patient.data, Hospital.Research.ID %in% rownames(all.predictions))$Patient
 
@@
 # Only BE, Progressors
-ndbe.only = do.call(rbind, lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], function(pt) {
+ndbe.only = do.call(rbind, lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], function(pt) {
   final.endo = subset(pt, Pathology %in% c('IMC','HGD'))
@@
 
+rownames(ndbe.only) = subset(sum.patient.data, Hospital.Research.ID %in% rownames(ndbe.only))$Patient
+
 ndbe.only$Patients = rownames(ndbe.only)
@@
 
-total = do.call(rbind, lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], function(pt) {
+total = do.call(rbind, lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], function(pt) {
     cbind.data.frame('predicted'=sum(as.integer(pt$Prediction > cutoff)), 'total'=nrow(pt))
@@
 total$pred.ratio = round(total$predicted/total$total, 2)
+
+rownames(total) = subset(sum.patient.data, Hospital.Research.ID %in% rownames(total))$Patient
 
@@
 # For all pre-HGD/IMC/LGD in progressors
-pre.hgd = sapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], function(pt) {
+pre.hgd = sapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], function(pt) {
   e = subset(pt, Pathology %nin% c('HGD', 'IMC', 'LGD'))
@@
 for (i in 1:nrow(p.predictions)) {
-  p = sapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], pred.by.endo, i)
+  p = sapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], pred.by.endo, i)
   pp = sapply(p, predicted.progression)
@@
 
-  fpr = sapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Patient], pred.by.endo, i)
+  fpr = sapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID], pred.by.endo, i)
   fpr = sapply(fpr, predicted.progression)
@@
 
-back = lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], transform.pt.by.time)
+back = lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], transform.pt.by.time)
 
@@
   }))
+  rownames(df) = subset(sum.patient.data, Hospital.Research.ID %in% rownames(df))$Patient
+
   df$Patient = rownames(df)
@@
 
-  #df$i = as.factor(df$i)
-
   df = arrange(df %>% group_by(Patient) %>% mutate('overall.pred'=sum(as.integer(pred > 0))/length(Patient)), Patient)
@@
       geom_line() + geom_point() +
-      geom_label_repel(aes(label=max.path, fill=Patient, color=Patient), color='white', fontface='bold', size=3, label.size=0, label.r=unit(0.25, 'lines'), label.padding=unit(0.15, 'lines')) +
-      labs(x='Months before diagnosis', y='Predictive Endoscopy') + theme(legend.position='none') + theme(strip.text = element_text(face="bold", size=9))
+      geom_label_repel(aes(label=max.path, fill=Patient, color=Patient), color='white', fontface='bold', size=4, label.size=0, label.r=unit(0.25, 'lines'), label.padding=unit(0.15, 'lines')) +
+      labs(x='Months before diagnosis', y='Predictive Endoscopy') + 
+      theme(text=element_text(size=14, face="bold"), axis.title=element_text(size=16), legend.position='none')
   } else {
-  ggplot(df, aes(months.before.final, pred.ratio, group=1,color=Patient)) + facet_grid(Patient ~.) + scale_x_continuous(breaks=seq(minM, maxM, by = 12)) +
-    geom_line() + geom_point() +  
-      geom_label_repel(aes(label=max.path, fill=Patient), color='white', fontface='bold', size=3, label.size=0, label.r=unit(0.25, 'lines'), label.padding=unit(0.15, 'lines')) +
-      labs(x='Months before diagnosis', y='Sample prediction ratio') + theme(legend.position='none') 
+  ggplot(df, aes(months.before.final, pred.ratio, group=1, color=Patient, fill=Patient)) + facet_grid(Patient ~.) +
+      ylim(0,1) + geom_abline(slope=0, intercept=seq(.25,.75,.25), color='grey', linetype='dashed') +
+      geom_col(position='dodge', aes(y=pred.ratio)) + geom_line() +
+      scale_x_continuous(breaks=seq(minM, maxM, by = 12)) + 
+      geom_label(aes(y=ifelse(pred.ratio==1, 0.75, pred.ratio), label=max.path, fill=Patient), color='white', fontface='bold', size=4, label.size=0, label.r=unit(0.25, 'lines'), label.padding=unit(0.15, 'lines'), nudge_y=0.1) +
+      labs(x='Months before diagnosis', y='Sample prediction ratio') + 
+      theme(text=element_text(size=14, face="bold"), axis.title=element_text(size=16), legend.position='none')
   }
@@
 ```{r, echo=F, warning=F, message=F, fig.height=22, fig.width=10}
-p = plot.endo( subset(endos, Patient %in% subset(me, months <= -72)$Patient), min(endos$months.before.final), max(endos$months.before.final), T)
-
-png(filename='progressors_1.png', width=600, height=1050)
+print.png<-function(p, name, ...) {
+  png(filename=name,...)
   print(p)
   dev.off()
-
+}
+
+p = plot.endo( subset(endos, Patient %in% subset(me, months <= -72)$Patient), bin=T) + 
+  labs(title="Progressors", subtitle=paste(72/12, "years or more follow-up"))
+print.png(p, 'progressors_1.png', width=900, height=1200)
 grid.arrange(p)
+
+p1a = plot.endo( subset(endos, Patient %in% subset(me, months <= -96)$Patient), bin=T) + 
+  labs(title="Progressors", subtitle=paste(96/12, "years or more follow-up"))
+print.png(p1a, 'progressors_1a.png', width=600, height=800)
+
+p1b = plot.endo( subset(endos, Patient %in% subset(me, months <= -72 & months > -96)$Patient), bin=T) + 
+  labs(title="Progressors", subtitle=paste(72/12, "-",96/12,"years or more follow-up"))
+print.png(p1b, 'progressors_1b.png', width=600, height=800)
+
+p1 = plot.endo( subset(endos, Patient %in% subset(me, months <= -72)$Patient), bin=F) +
+  labs(title="Progressors", subtitle=paste(72/12, "years or more follow-up"))
+print.png(p1, 'progressors_1_bar.png', width=900, height=1200)
+
 ```
@@
 ```{r, echo=F, warning=F, message=F, fig.height=22, fig.width=10}
-p1 = plot.endo( subset(endos, Patient %in% subset(me, months <= -48 & months > -72)$Patient), min(endos$months.before.final), max(endos$months.before.final), T) 
-p2 = plot.endo( subset(endos, Patient %in% subset(me, months <= -24 & months > -48)$Patient), min(endos$months.before.final), max(endos$months.before.final), T) 
-p3 = plot.endo( subset(endos, Patient %in% subset(me, months >= -12)$Patient), min(endos$months.before.final), max(endos$months.before.final), T)
+p1 = plot.endo( subset(endos, Patient %in% subset(me, months <= -48 & months > -72)$Patient), bin=T) +
+  labs(title="Progressors", subtitle=paste(48/12, "-",72/12,  "years or more follow-up"))
+print.png(p1, 'progressors_2.png', width=400, height=600)
+
+p1 = plot.endo( subset(endos, Patient %in% subset(me, months <= -48 & months > -72)$Patient), bin=F) +
+  labs(title="Progressors", subtitle=paste(48/12, "-",72/12,  "years or more follow-up"))
+print.png(p1, 'progressors_2_bar.png', width=400, height=600)
+
+p2 = plot.endo( subset(endos, Patient %in% subset(me, months <= -24 & months > -48)$Patient), bin=T) +
+  labs(title="Progressors", subtitle=paste(24/12, "-",48/12,  "years or more follow-up"))
+print.png(p2, 'progressors_3.png', width=400, height=600)
+
+p2 = plot.endo( subset(endos, Patient %in% subset(me, months <= -24 & months > -48)$Patient), bin=F) +
+  labs(title="Progressors", subtitle=paste(24/12, "-",48/12,  "years or more follow-up"))
+print.png(p2, 'progressors_3_bar.png', width=400, height=600)
+
+p3 = plot.endo( subset(endos, Patient %in% subset(me, months >= -12)$Patient), bin=T) + 
+  labs(title="Progressors", subtitle=paste("1 year follow-up"))
+print.png(p3, 'progressors_4.png', width=400, height=600)
+
+p3 = plot.endo( subset(endos, Patient %in% subset(me, months >= -12)$Patient), bin=F) +
+  labs(title="Progressors", subtitle=paste("1 year follow-up"))
+print.png(p3, 'progressors_4_bar.png', width=400, height=600)
 
@@
 
-png(filename='progressors_2.png', width=400, height=600)
-print(p1)
-dev.off()
-
-png(filename='progressors_3.png', width=400, height=600)
-print(p2)
-dev.off()
-
-png(filename='progressors_4.png', width=400, height=600)
-print(p3)
-dev.off()
-
 ```
@@
 ```{r echo=F, warning=F, message=F, fig.height=42, fig.width=10}
-nonp.back = lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Patient],transform.pt.by.time)
+nonp.back = lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID],transform.pt.by.time)
 
@@
   }))
+  rownames(df) = subset(sum.patient.data, Hospital.Research.ID %in% rownames(df))$Patient
+
   df$Patient = rownames(df)
+
   return(df)
@@
 nonp.endos = nonp.endos[complete.cases(nonp.endos),]
-nonp.endos$i = nonp.endos$i-1
+#nonp.endos$i = nonp.endos$i-1
 
@@
 
-nonp.endos$months.before.final
-p1 = plot.endo( subset(nonp.endos, Patient %in% subset(me, months <= -108)$Patient), min(nonp.endos$months.before.final), max(nonp.endos$months.before.final), T)
-
-png(filename='nonprogressors_1.png', width=600, height=800)
-print(p1)
-dev.off()
-
-p2 = plot.endo( subset(nonp.endos, Patient %in% subset(me, months <= -72 & months > -108 )$Patient), min(nonp.endos$months.before.final), max(nonp.endos$months.before.final), T)
-
-png(filename='nonprogressors_2.png', width=600, height=800)
-print(p2)
-dev.off()
-
-p3 = plot.endo( subset(nonp.endos, Patient %in% subset(me, months > -72  )$Patient), min(nonp.endos$months.before.final), max(nonp.endos$months.before.final), T)
-
-png(filename='nonprogressors_3.png', width=600, height=800)
-print(p3)
-dev.off()
-
-grid.arrange(p1,p2,p3, ncol=1)
+plot.endo(subset(nonp.endos, Patient %in% subset(nonp.endos, pred.ratio > 0)$Patient), bin=F)
+
+p = plot.endo( subset(nonp.endos, Patient %in% subset(nonp.endos, pred.ratio > 0)$Patient), bin=T) + 
+  labs(title="Non-Progressors", subtitle=paste("Patients with predictive endoscopies, max", abs(min(me$months)/12), 'years follow-up'))
+print.png(p, 'non-progressors.png', width=600, height=800)
+
+p1 = plot.endo( subset(nonp.endos, Patient %in% subset(nonp.endos, pred.ratio > 0)$Patient), bin=F) + 
+  labs(title="Non-Progressors", subtitle=paste("Patients with predictive endoscopies, max", abs(min(me$months)/12), 'years follow-up'))
+print.png(p1, 'non-progressors_bar.png', width=600, height=800)
+
+grid.arrange(p)
 ```
@@
 ```{r dq_batch, echo=T, warning=F, message=F}
+pg.samp = lapply(pg.samp, function(df){ df$Batch.Name = sub('_', '', df$Batch.Name ); df })
+
 all = do.call(rbind, pg.samp)
 
-prog = do.call(rbind, pg.samp[subset(sum.patient.data, Status == 'P')$Patient])
-nonp = do.call(rbind, pg.samp[subset(sum.patient.data, Status == 'NP')$Patient])
+prog = do.call(rbind, pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID])
+nonp = do.call(rbind, pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID])
 
@@
 
-pilot = lapply(pg.samp, function(df) subset(df, Batch.Name == 'NP_Pilot_Study'))
+pilot = pg.samp[subset(patient.info, grepl('nppilotstudy', Batch.Name, ignore.case=T))$Hospital.Research.ID]
 pilot = do.call(rbind, pilot[ which(sapply(pilot, nrow) > 0) ])
@@
 
-exome = lapply(pg.samp, function(df) subset(df, Batch.Name == 'Exome_subcohort'))
+exome =  pg.samp[subset(patient.info, grepl('exomesubcohort', Batch.Name, ignore.case=T))$Hospital.Research.ID]
 exome = do.call(rbind, exome[ which(sapply(exome, nrow) > 0) ])
@@
 
-```{r, message=F, warning=F, echo=F, eval=T, fig.height=8, fig.width=6}
-
-fitV <- glmnet(dysplasia.df, labels, alpha=select.alpha, nlambda=nl, family='binomial') # all patients
+```{r validation, message=F, warning=F, echo=F, fig.height=8, fig.width=6}
+
+score.cx <-function(pd) { ## OLD method
+#  print(pd$info)
+  len.var = (var(pd$seg.vals$end.pos - pd$seg.vals$start.pos))/1e12
+  extreme.values = max( apply(as.matrix(pd$seg.vals[,grep('^D', colnames(pd$seg.vals))]), 2, function(col) {
+    length(which(col >= (median(col)+2*sd(col)))) + length(which(col <= (median(col)-2*sd(col)))) 
+  }) )
+
+  (len.var + extreme.values)
+}
+
+validation.df = t(mergedDfValidation[, grep('^D', colnames(mergedDfValidation))])
+validation.df = apply(validation.df, 2, unit.var)
+
+complexity.measures = sapply(validation.patient.data, score.cx)
+validation.df = cbind(validation.df, 'cx.measure'=NA)
+for (sample in rownames(validation.df)) {
+  validation.df[sample, 'cx.measure'] = complexity.measures[subset(validation.patient.info, Samplename == sample)$Hospital.Research.ID]
+}
+# 
+# # norm to 1 to keep within the magnitude of the genomic data
+cx = validation.df[,'cx.measure']
+validation.df[,'cx.measure'] = (log(cx)/log(max(cx)))
+
+file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
+message(paste("loading", file))
+load(file, verbose=T)
+
+#fitV <- glmnet(dysplasia.df, labels, alpha=select.alpha, nlambda=nl, family='binomial') # all patients
+#autoplot(fitV)
+fitV = models[[select.alpha]]
 lambda.opt = performance[select.alpha, 'lambda']
@@
 
+names(v.predictions) = subset(sum.validation.info, Hospital.Research.ID %in% names(v.predictions))$Patient
+
 v.predictions = do.call(rbind, lapply(v.predictions, function(x) {
@@
 
-p = subset(v.predictions, Status == 'P')
-np = subset(v.predictions, Status == 'NP')
+p = arrange(subset(v.predictions, Status == 'P'), -pred.ratio)
+np = arrange(subset(v.predictions, Status == 'NP'), -pred.ratio)
 
@@
 
-grid.arrange(
-  ggplot() + geom_bar(data=subset(m, Status == 'P'), aes(Patient, value), fill='darkblue', color='grey', stat='identity') +  
-    geom_label( aes(x=nrow(p)/2+nrow(p)/3, y=0.5, label=paste(round(nrow(p[p$pred.ratio == 1,])/nrow(p),2)*100,'%',sep='')), size=8, fill='darkblue', color='white', fontface='bold' ) +
-    geom_label( aes(x=nrow(p)/3, y=0.5, label=paste(round(nrow(p[p$pred.ratio < 1 & p$pred.ratio > 0,])/nrow(p),2)*100,'%',sep='')), size=8, fill='darkblue', color='white', fontface='bold' ) +
-    geom_text( aes(x=2, y=0.5, label=paste(round(nrow(p[p$pred.ratio == 0,])/nrow(p),2)*100,'%',sep='')), size=8,  fontface='bold' ) +
-    theme(legend.position='none', axis.text.x = element_text(angle = 45, hjust = 1)) + coord_flip() +
-    labs(title='Progressors', x='', y='Ratio predicted samples'),
-
-  ggplot() + geom_bar(data=subset(m, Status == 'NP'), aes(Patient, value), fill='darkgreen', color='grey', stat='identity') + 
-        geom_label( aes(x=nrow(np)/2+nrow(np)/4, y=0.5, label=paste(round(nrow(np[np$pred.ratio == 1,])/nrow(np),2)*100,'%',sep='')), size=8, fill='darkgreen', color='white', fontface='bold' ) +
-    geom_label( aes(x=nrow(np)/4, y=0.5, label=paste(round(nrow(np[np$pred.ratio < 1 & np$pred.ratio > 0,])/nrow(p),2)*100,'%',sep='')), size=8, fill='darkgreen', color='white', fontface='bold' ) +
-        #geom_text( aes(x=1.2, y=0.5, label=paste(round(nrow(np[np$pred.ratio == 0,])/nrow(np),2)*100,'%',sep='')), size=8,  fontface='bold' ) +
-    theme(legend.position='none', axis.text.x = element_text(angle = 45, hjust = 1)) + coord_flip() +
-    labs(title='Non Progressors', x='', y='Ratio predicted samples'),
-  ncol=2, top='Validation data set')
+p1 = ggplot(data=subset(v.predictions, Status == 'P'), aes(Patient, pred.ratio)) + geom_col(fill='darkblue', color='grey') + coord_flip() +
+    geom_label( aes(x=p[nrow(subset(p, pred.ratio == 1))/2,'Patient'], y=0.5, label=paste(round(nrow(p[p$pred.ratio == 1,])/nrow(p),2)*100,'%',sep='')), size=8, fill='darkblue', color='white', fontface='bold' ) +
+    geom_label( aes(x='77', y=0.5, label=paste(round(nrow(p[p$pred.ratio < 1 & p$pred.ratio > 0,])/nrow(p),3)*100,'%',sep='')), size=8, fill='darkblue', color='white', fontface='bold' ) +
+    theme(legend.position='none', axis.text.x = element_text(angle = 45, hjust = 1)) + 
+    labs(title='Progressors', x='', y='Ratio predicted samples')
+if (nrow(subset(p,pred.ratio == 0)) > 0) {
+  p1 = p1 +  geom_text( aes(x=subset(p,pred.ratio == 0)$Patient[1], y=0.5, label=paste(round(nrow(p[p$pred.ratio <= 0,])/nrow(p),3)*100,'%',sep='')), size=8, fontface='bold' ) 
+}
+
+p2 = ggplot(data=subset(v.predictions, Status == 'NP'), aes(Patient, pred.ratio)) + geom_col(fill='darkgreen', color='grey') + coord_flip() +
+    geom_label( aes(x=np[nrow(subset(np, pred.ratio == 1))/2,'Patient'], y=0.5, 
+                    label=paste(round(nrow(np[np$pred.ratio == 1,])/nrow(np),2)*100,'%',sep='')), size=8, fill='darkgreen', color='white', fontface='bold' ) +
+    geom_label( aes(x='41', y=0.5, label=paste(round(nrow(np[np$pred.ratio < 1 & np$pred.ratio > 0,])/nrow(np),3)*100,'%',sep='')), size=8, fill='darkgreen', color='white', fontface='bold' ) +
+    theme(legend.position='none', axis.text.x = element_text(angle = 45, hjust = 1)) + 
+    labs(title='Non Progressors', x='', y='Ratio predicted samples')
+if (nrow(subset(np,pred.ratio == 0)) > 0) {
+  p2 = p2 +  geom_text( aes(x=subset(np,pred.ratio == 0)$Patient[1], y=0.5, label=paste(round(nrow(pn[pn$pred.ratio <= 0,])/nrow(np),3)*100,'%',sep='')), size=8, fontface='bold' ) 
+}
+
+grid.arrange(p1, p2, ncol=2, top='Validation data set')
 
@@
 ```{r, echo=F, message=F, warning=F, fig.width=8, fig.height=12}
-# sapply(vpd[subset(sum.validation.info, Status == 'P')$Patient], function(df) {
-#   length(which(df$Prediction > cutoff))/nrow(df)
-# })
 
@@
   }))
+  rownames(df) = subset(sum.validation.info, Hospital.Research.ID %in% rownames(df))$Patient
   df$Patient = rownames(df)
@@
 
-#vpi[,c('Status','Endoscopy.Year','pred','samples', 'months.before.final', 'max.path')]
-
-p1 = plot.endo(subset(vpi, Patient %in% subset(sum.validation.info, Status == 'P')$Patient), min(vpi$months.before.final), max(vpi$months.before.final), T) +  labs(title="Validation - Progressors")
-
-png(filename='validation_progressors.png', width=600, height=800)
-print(p1)
-dev.off()
-
-grid.arrange(p1,ncol=1)
-
-p2 = plot.endo(subset(vpi, Patient %in% subset(sum.validation.info, Status == 'NP')$Patient), min(vpi$months.before.final), max(vpi$months.before.final), T) + labs(title="Validation - Non Progressors")
-
-png(filename='validation_Nonprogressors.png', width=600, height=800)
-print(p2)
-dev.off()
-
-grid.arrange(p2,ncol=1)
+p1 = plot.endo(subset(vpi, Patient %in% subset(sum.validation.info, Status == 'P')$Patient), bin=T) + labs(title="Validation - Progressors")
+print.png(p1, 'validation_progressors.png', width=600, height=800)
+
+print.png(plot.endo(subset(vpi, Patient %in% subset(sum.validation.info, Status == 'P')$Patient), bin=F) + labs(title="Validation - Progressors"), 'validation_progressors_bar.png', width=600, height=800)
+
+p2 = plot.endo(subset(vpi, Patient %in% subset(sum.validation.info, Status == 'NP')$Patient), bin=T) + labs(title="Validation - Non Progressors")
+
+print.png(p2,'validation_Nonprogressors.png', width=600, height=800)
+print.png(plot.endo(subset(vpi, Patient %in% subset(sum.validation.info, Status == 'NP')$Patient), bin=F) + labs(title="Validation - Non Progressors"),'validation_Nonprogressors_bar.png', width=600, height=800)
+
+grid.arrange(p1, p2,ncol=1)
 ```
@@
 m = melt(validation.patient.info, id.vars = c('Status'), measure.vars = c('Prediction'))
-p3 = ggplot(m, aes(Status, value, fill=Status)) + geom_boxplot() + geom_jitter(color='darkgrey', show.legend = F) +
+p3 = ggplot(m, aes(Status, value, fill=Status)) + 
+  geom_jitter(aes(color=Status), width=0.2, show.legend = F) +  geom_violin(alpha=0.8) +
   geom_label(aes(x='P', y=0.75, label=paste('TPR=',tpr*100, '%', sep='')), fill='white') + 
@@
 
-png(filename='validation_predictions.png', width=600, height=800)
-print(p3)
-dev.off()
-
-```
+print.png(p3, 'validation_predictions.png', width=600, height=800)
+
+```

### 2017-08-02 RegressionPvNP.Rmd
Wrapping up model

@@
 
-includeDemo = FALSE # demographics
+includeDemo = F # demographics
 
@@
 cache.dir = paste(data, 'Analysis',sub('\\+', '', tile.w),   sep='/')
+
 if (!dir.exists(cache.dir)) dir.create(cache.dir)
@@
 unit.var <- function(x) {
-  if (sd(x) == 0) return(x)
-  return((x-mean(x))/sd(x) )
-}
-
-#dysplasia.df = apply(dysplasia.df, 2, function(x) x-median(x))
+  if (sd(x, na.rm=T) == 0) return(x)
+  return((x-mean(x,na.rm=T))/sd(x,na.rm=T) )
+}
+
 dysplasia.df = apply(dysplasia.df, 2, unit.var)
@@
 
-  ## Some sort of complexity score
-  #plot(unlist(complexity.measures), col=sum.patient.data$Status)
-  # t.test(
-  # complexity.measures[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID],
-  # complexity.measures[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID]
-  # )
-
-  score.cx <-function(pd) {
+  score.cx <- function(pt.d, df) {
+    get.length.variance <- function(pd) {
       samples = grep('^D', colnames(pd$seg.vals), value=T)
@@
 
-    #extremes = apply(as.matrix(pd$seg.vals[,samples]), 2, function(col) {
-    #  length(which(col >= (median(col)+2*sd(col)))) + length(which(col <= (median(col)-2*sd(col)))) 
-    #})
-
-    num.var.reg = vector(mode='integer', length=length(samples))
-    if (!is.null(pd$variable.regions) && nrow(pd$variable.regions) > 0) {
-      num.var.reg = apply(as.matrix(pd$variable.regions[,samples]), 2, function(col) {
-        length(which(col >= 1.1 | col <= 0.9))
-      })
-    }
-
       sample.var = apply(as.matrix(pd$seg.vals[, samples]), 2, var)
-    t(cbind(len.var, num.var.reg, sample.var))
-  }
-
-  complexity.measures = lapply(patient.data, score.cx)
-
-  dysplasia.df = cbind(dysplasia.df, 'segment.length'=0, 'num.var.regions'=0, 'sample.variance'=0)
+      t(cbind(len.var, sample.var))
+    }
+
+    complexity.measures = lapply(pt.d, get.length.variance)
+
+    df = cbind(df, 'cx' = apply(df, 1, function(x) length(which(x >= sd(x)*2 | x <= -sd(x)*2))))
+    df[,'cx'] = unit.var(df[,'cx'])
+
+    df = cbind(df, 'segment.length'=0, 'sample.variance'=0)
     for (pt in names(complexity.measures)) {
       for (sample in colnames(complexity.measures[[pt]])) {
-      dysplasia.df[sample,c('segment.length', 'num.var.regions', 'sample.variance')] = as.numeric(complexity.measures[[pt]][,sample])
-    }
-    #for (sample in rownames(dysplasia.df)) {
-      #dysplasia.df[sample, ] = 
-        #complexity.measures[subset(patient.info, Samplename == sample)$Hospital.Research.ID]
-    #}
-  }
-  # norm to 1 to keep within the magnitude of the genomic data
-  scale.to.1<-function(x) {
-    (x-min(x))/(max(x)-min(x))
-  }
-
-  dysplasia.df[,c('segment.length', 'num.var.regions', 'sample.variance')] = apply(dysplasia.df[,c('segment.length', 'num.var.regions', 'sample.variance')], 2, scale)
-
-  #cx = dysplasia.df[,'cx.measure'] 
-  #dysplasia.df[,'cx.measure'] = (log(cx)/log(max(cx)))
-  #dysplasia.df[,'cx.measure'] = dysplasia.df[,'cx.measure'] - median(dysplasia.df[,'cx.measure'])
-  #dysplasia.df[,'cx.measure'] = ((cx - min(cx))/(max(cx)-min(cx)))
+        df[sample,c('segment.length', 'sample.variance')] = as.numeric(complexity.measures[[pt]][,sample])
+      }
+    }
+
+   df[, 'segment.length'] = unit.var(df[, 'segment.length'])
+   df[, 'sample.variance'] = unit.var(df[, 'sample.variance'])
+
+  return(df)
+  }
+
+## TODO score complexity on the seg.vals rather than merged data frame
+  dysplasia.df = score.cx(patient.data, dysplasia.df)
 
@@
   demo.data$age.encoded = NULL
-  demo.data$BMI = scale(demo.data$BMI)
-
-  demo.data[,c('Circumference','Maximal')] = apply(demo.data[,c('Circumference','Maximal')], 2, function(x) (x-mean(x, na.rm=T))/sd(x, na.rm=T))
+  demo.data$BMI = unit.var(demo.data$BMI)
+
+  demo.data[,c('Circumference','Maximal')] = apply(demo.data[,c('Circumference','Maximal')], 2, unit.var)
 
@@
       # pre-spec lambda seq
-      fitCV = glmnet(training, y[-test.rows], lambda=lambda, alpha=a, family='binomial', ...) 
+      fitCV = glmnet(training, y[-test.rows], lambda=lambda, alpha=a, family='binomial', standardize = F) 
       # autoplot(fit) + theme(legend.position="none")
@@
     lmin = subset(df, lambda.min)$lambda
-    df[subset(df, log.lambda <= log(lmin) + log(sd(lambda))/2 )[1, 'lambda.at'], 'lambda.1se'] = T
+    lambdas = subset(df, log.lambda <= log(lmin) + log(sd(lambda))/2 )
+    n = 3
+    while (nrow(lambdas) <= 0 & n < 10) {
+      lambdas = subset(df, log.lambda <= log(lmin) + log(sd(lambda))/n )
+      n = n+1
+    }
+    if (nrow(lambdas) <= 0) {
+      df[subset(df, log.lambda <= log(lmin) )[1, 'lambda.at'], 'lambda.1se'] = T  
+    } else {
+      df[subset(df, log.lambda <= log(lmin) + log(sd(lambda))/n )[1, 'lambda.at'], 'lambda.1se'] = T  
+    }
   }
@@
 ```{r xvalpt, message=F, warning=F, echo=F, fig.height=16, fig.width=16}
-# dysplasia.df = cbind(dysplasia.df, 'sl.nvr'=0)
-# dysplasia.df[, 'sl.nvr'] = scale(dysplasia.df[, "segment.length"] + dysplasia.df[, 'num.var.regions'])
-# 
-# dysplasia.df = cbind(dysplasia.df, 'sl.nvr.sv'=0)
-# dysplasia.df[, 'sl.nvr.sv'] = scale(dysplasia.df[, "segment.length"] + dysplasia.df[, 'num.var.regions'] + dysplasia.df[, 'sample.variance'])
-# 
-# dysplasia.df = cbind(dysplasia.df, 'sl.sv'=0)
-# dysplasia.df[, 'sl.sv'] = scale(dysplasia.df[, "segment.length"] + dysplasia.df[, 'sample.variance'])
-# 
-# dysplasia.df = cbind(dysplasia.df, 'nvr.sv'=0)
-# dysplasia.df[, 'sl.sv'] = scale(dysplasia.df[, "num.var.regions"] + dysplasia.df[, 'sample.variance'])
-
 nl = 1000
@@
 
-file = paste(cache.dir, 'patient_folds.tsv', sep='/')
+file = paste(data, 'Analysis/patient_folds.tsv', sep='/')
 if (file.exists(file)) {
+  message(paste("Reading folds file", file))
   sets = read.table(file, header=T, sep='\t')
 } else {
-  sets = create.patient.sets(pts, folds, splits, 0.2)  
-  write.table(sets, quote=F, sep='\t', row.names=F, file=file)
+  #sets = create.patient.sets(pts, folds, splits, 0.2)  
+  #write.table(sets, quote=F, sep='\t', row.names=F, file=file)
+  stop("Missing patient folds file")
 }
@@
     cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=l, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', standardize=F)
+    #cv.patient$deviance.plot
 
@@
 
-```{r echo=F, warning=F, message=F, eval=F}
-### Features in samples
-## TODO Look at the coefficients, are the high pos ones "gains" and the low neg ones "losses" etc
-plot(coefs[[select.alpha]][,1])
-selected = coefs[[select.alpha]]
-
-selected[which(
-  selected[,1] > mean(selected[,1]) + sd(selected[,1]) | 
-  selected[,1] < mean(selected[,1] - sd(selected[,1]))  
-),]
-
-## Variance in the selected features?  Not so much
-coefs[[select.alpha]][,1]
-
-v = vector(length(colnames(dysplasia.df)), mode='numeric')
-names(v) = colnames(dysplasia.df)
-v[rownames(selected)] = coefs[[select.alpha]][,1]
-
-vv <- apply(dysplasia.df, 2, var)
-plot(v, vv, log='y', col=(v!=0)+1)
-
-boxplot(vv ~ (v==0), log='y', ylim=c(0.0001,1))
-
-qqplot(vv[v==0], vv[v!=0], log='xy')
-
-length(which( coef.stable[[select.alpha]] >= 0.5) )
-
-regions = rownames(coefs[[select.alpha]])
-
-pg = dysplasia.df[ labels == 1 ,regions]
-
-nonpg = dysplasia.df[ labels == 0 ,regions]
-
-colnames(pg)
-
-meantests = sapply(regions, function(r)   t.test(pg[,r], nonpg[,r])$p.value )
-
-which(meantests < 0.05)
-
-#summary(pg[,'1:119640299-129610322'])
-#summary(nonpg[,'1:119640299-129610322'])
-
-```
-
 ### Feature Information
@@
 
-```{r, warning=F, message=F, eval=T}
+```{r featureinfo, warning=F, message=F, eval=T}
 #topF = names(which((coef.stable[['1']]) >= 0.75))
@@
 
-```{r, echo=F, warning=F, message=F}
+```{r bplen, echo=F, warning=F, message=F}
 chromInfo = read.table('hg19_genes.txt', sep='\t', header=T)
@@
 
-```{r, warning=F, message=F}
+```{r genes, warning=F, message=F}
 bm = NULL
@@
 
-```{r echo=F, warning=F, message=F,eval=T, fig.height=8, fig.width=8}
+```{r hazards, echo=F, warning=F, message=F,eval=T, fig.height=12, fig.width=12}
 
@@
 ch$coef = coefs[[select.alpha]][,1]
+ch$stability = featureStability
 ch$label = rownames(ch)
-ch$stability = featureStability
-
-ggplot(ch, aes(coef, hazards, col=hazards > quantile(hazards)[['75%']])) + 
-  #xlim(min(coefGR$coef)-5, max(coefGR$coef)+10) +
+
+qq = quantile(ch$hazards, seq(0,1,.15))
+ggplot(ch, aes(coef, hazards, col=hazards >= qq[length(qq)])) + 
   geom_point(alpha=0.5) + 
-  geom_text_repel(aes(x=coef, label=ifelse(hazards > quantile(hazards)[['75%']], label, '')),show.legend=F) +
-  labs(title='Top features coef vs hazard') + theme(legend.position = 'bottom')
-```
-
-```{r echo=F, warning=F, message=F,eval=F, fig.height=12, fig.width=8}
-ch = arrange(ch, -stability, -hazards, -coef)
-ch$hazards = round(ch$hazards, 3)
-ch$coef = round(ch$coef, 3)
-write.table(ch[c('label', 'stability','hazards','coef')], row.names = F, quote=F, sep="\t", file='features.txt')
-
-ch = arrange(ch, -hazards)
-
-melt.by.reg<-function(x) {
-  r = x[['label']]
-
-  reg = cbind.data.frame( 'value'=dysplasia.df[ ,r], labels)
-  reg$labels = factor(reg$labels)
-  reg$labels = revalue(reg$labels, c("1"="P", "0"="NP"))
-
-  tt = wilcox.test(dysplasia.df[ which(labels == 0) , r ], dysplasia.df[ which(labels == 1) , r ])
-
-  reg$pval = tt$p.value
-  reg$region = r
-  reg$hazard = as.numeric(x[['hazards']])
-  reg$coef = as.numeric(x[['coef']])
-  return(reg)
-}
-
-plot.reg<-function(x) {
-  reg = melt.by.reg(x)
-
-  pv = ifelse(reg$pval[1] < .001, '< .001', format.pval(unique(reg$pval[1])))
-
-#  ggplot(reg, aes(value, fill=labels) ) + geom_histogram(bins=50, color='grey') + 
-#    labs(x='seg. value', title=r, subtitle=paste('haz=',x[['hazards']], ' coef=', x[['coef']], ' MW pval: ',pv,sep=''))
-  ggplot(reg, aes(x=labels, y=value, group=labels, fill=labels) ) + geom_jitter(color='grey', show.legend = F) + geom_violin(show.legend = F, alpha=0.5) +
-    geom_hline(yintercept = median(reg$value) + sd(reg$value)) + geom_hline(yintercept = median(reg$value) - sd(reg$value)) +
-    labs(x='', y='seg. value', title=x[['label']], subtitle=paste('haz= ',x[['hazards']], ' coef= ',x[['coef']], '\nMW pval ',pv,sep=''))
-}
-
-ch = arrange(ch, -hazards)
-
-regions = apply(ch, 1, melt.by.reg)
-names(regions) = ch$label
-
-do.call(grid.arrange, c(lapply(regions[1:15], function(reg) {
-  print(reg$region[1])
-  cols = c('indianred3','slateblue2')
-  if ( as.numeric(reg$coef[1]) < 0) cols = c('green3', 'darkviolet')
-
-  ggplot(reg, aes(labels, value*coef, fill=labels, color=labels)) + 
-    geom_jitter(aes(labels, value), width=0.05, color='yellow4', alpha=0.3) +
-    geom_jitter(aes(labels, value*coef, color=labels), width=0.05, alpha=0.5) +
-    geom_text_repel(data=subset(reg, value >= quantile(reg$value)[['100%']]), aes(labels, value, label=round(value, 3)), color='yellow4') +
-    geom_text_repel(data=subset(reg, value >= quantile(reg$value)[['100%']]), aes(labels, value*coef, label=round(value, 3),color=labels)) + geom_violin(alpha=0.2) + 
-    scale_fill_manual(values=cols) + scale_color_manual(values=cols) +
-    theme(legend.position = 'none') + labs(title=unique(reg$region)[1], x='')
-})))
-
-    #geom_violin(aes(labels, value, fill=labels), alpha=0.5) + theme(legend.position = 'none')
-
-#  x = rbind.data.frame('P'= cbind( 'gain' = length(which(with(subset(reg, labels == 'P'), value > 0))),
-#                               'loss' = length(which(with(subset(reg, labels == 'P'), value < 0))) ),
-
-#                        'NP' = cbind( 'gain' = length(which(with(subset(reg, labels == 'NP'), value > 0))),
-#                                'loss' = length(which(with(subset(reg, labels == 'NP'), value < 0))) ) )
-
-    # x = rbind.data.frame('P'= cbind( 'gain' = length(which(with(subset(reg, labels == 'P'), value >= median(value) + sd(value)*se))),
-  #                              'loss' = length(which(with(subset(reg, labels == 'P'), value <= median(value) - sd(value)*se))) ),
-  # 
-  #                 'NP' = cbind( 'gain' = length(which(with(subset(reg, labels == 'NP'), value >= median(value) + sd(value)*se))),
-  #                               'loss' = length(which(with(subset(reg, labels == 'NP'), value <= median(value) - sd(value)*se))) ) )
-  # x = rbind.data.frame('P'= cbind( 'high' = length(which(with(subset(reg, labels == 'P'), value >= 1.1))),
-  #                                  'low' = length(which(with(subset(reg, labels == 'P'), value <= 0.9))) ),
-  # 
-  #                 'NP'= cbind( 'high' = length(which(with(subset(reg, labels == 'NP'), value >= 1.1))),
-  #                                  'low' = length(which(with(subset(reg, labels == 'NP'), value <= 0.9))) ))
-
-  #cols = c('indianred3','slateblue2')
-  #if ( as.numeric(reg$coef[1]) < 0) cols = c('plum', 'skyblue3')
-
-  #ggplot(melt(as.matrix(x)), aes(Var2, value, fill=Var1)) + facet_grid(~Var1) +
-  #  geom_col( show.legend = F) + labs(x='', y='', title=unique(reg$region)[1]) + scale_fill_manual(values=cols)
-
-#})))
-
-## Individual hazards vs CN values?  exp(coef) -- odds ratio
-# plot( exp( coefs[[select.alpha]][ch$label[1],1]  * dysplasia.df[,ch$label[1]]),  dysplasia.df[,ch$label[1]] )
-# 
-# plot(  exp( t(dysplasia.df[, ch$label[1]]) *  coefs[[select.alpha]][ch$label[1],1]  ), dysplasia.df[,ch$label[1]]  ) ## pos coef
-# plot(exp(t(dysplasia.df[, ch$label[3]]) *  coefs[[select.alpha]][ch$label[3],1]), dysplasia.df[,ch$label[3]]) ## pos coef
-# 
-# plot(exp(t(dysplasia.df[, ch$label[2]]) *  coefs[[select.alpha]][ch$label[2],1]), dysplasia.df[,ch$label[2]]) ## neg coef
-# plot(exp(t(dysplasia.df[, ch$label[4]]) *  coefs[[select.alpha]][ch$label[4],1]), dysplasia.df[,ch$label[4]]) ## neg coef
-# 
-# 
-# labels[which( dysplasia.df[, ch$label[2]] <= 0.9)]
-# 
-# labels[which( dysplasia.df[, ch$label[1]] >= 1.1)]
-
-#do.call(grid.arrange, c(apply(ch[1:10,],1, plot.reg), ncol=2))
-
-#do.call(grid.arrange, c(apply(ch[11:20,],1, plot.reg), ncol=2))
-
-#do.call(grid.arrange, c(apply(ch[21:30,],1, plot.reg), ncol=2))
-
-```
-
-```{r}
-
+  geom_text_repel(aes(x=coef, label=ifelse(hazards >= qq[length(qq)], label, '')),show.legend=F) +
+  labs(title='Top features by hazard ratio') + theme(legend.position = 'bottom') + theme(legend.position='none')
+```
+
+```{r, echo=F}
 cols = intersect(rownames(ch), grep('\\d+:', colnames(dysplasia.df), invert=T, value=T))
@@
 
-`r pander(ch[order(-ch$hazards),][1:10,1:2], caption='Top 10 features by hazard ratio.')`
-
-```{r echo=F, warning=F, message=F}
+#### COSMIC genes
+
+```{r cosmic, echo=F, warning=F, message=F}
 if (!is.null(bm)) {
@@
   coefGR$coef = coefs[[select.alpha]][grcfs,1]
+
+  coefGR$stability = rowSums(coefs[[select.alpha]][grcfs,-1])/(folds*splits)
 
@@
 write.table(unlist(strsplit(unique(highest$ensembl_ids), ', ')), quote=F, row.names = F, col.names = F, file='highest_coef_genes.txt')
+
+write.table(as.data.frame(coefGR[order(-coefGR$stability)])[, c('seqnames', 'start','end','coef','stability','hazard', 'COSMIC')], sep='\t', quote=F, row.names=F, file='cosmic_genes.txt')
 
@@
 Similarly, `r round(length(unlist(strsplit(unique(coefGR$COSMIC), ',')))/length(ccgenes$Gene.Symbol),2)*100`% of the `r length(ccgenes$Gene.Symbol)` COSMIC genes are contained within these regions.
+```
+
+```{r, echo=F}
+pander(as.data.frame(coefGR[coefGR$COSMIC != ""])[c(1:3, 11)], caption='COSMIC genes mapped to coefficients.
+Only a few regions have mapped COSMIC genes, but several regions neighbor those with important tumour supressors (TP53) and may have not been selected through an artifact of the GLM process.')
+```
+
+### How do coefficients related to the raw data?
+
+```{r warning=F, message=F, fig.height=8, fig.width=8}
+ch = arrange(ch, -stability, -hazards, -coef)
+ch$hazards = round(ch$hazards, 3)
+ch$coef = round(ch$coef, 3)
+write.table(ch[c('label', 'stability','hazards','coef')], row.names = F, quote=F, sep="\t", file='features.txt')
+
+ch = arrange(ch, -hazards)
+
+melt.by.reg<-function(x) {
+  r = x[['label']]
+
+  reg = cbind.data.frame( 'value'=dysplasia.df[ ,r], labels)
+  reg$labels = factor(reg$labels, levels=c(1,0), ordered=T)
+  reg$labels = revalue(reg$labels, c("1"="P", "0"="NP"))
+
+  #reg = factor(reg$labels, levels=c('P','NP'), ordered=T)
+
+  tt = wilcox.test(dysplasia.df[ which(labels == 0) , r ], dysplasia.df[ which(labels == 1) , r ])
+
+  reg$pval = tt$p.value
+  reg$region = r
+  reg$hazard = as.numeric(x[['hazards']])
+  reg$coef = as.numeric(x[['coef']])
+  return(reg)
+}
+
+# plot.reg<-function(x) {
+#   reg = melt.by.reg(x)
+#   
+#   pv = ifelse(reg$pval[1] < .001, '< .001', format.pval(unique(reg$pval[1])))
+# 
+#   ggplot(reg, aes(x=labels, y=value, group=labels, color=labels, fill=labels) )  + geom_violin(alpha=0.5) + geom_jitter(width=0.3, alpha=0.8) +
+#     geom_hline(yintercept=with(reg, c(mean(value)+sd(value)*2, mean(value)-sd(value)*2)), color='darkgrey', linetype='dashed') +
+#     #geom_hline(yintercept = mean(reg$value) + sd(reg$value)) + geom_hline(yintercept = mean(reg$value) - sd(reg$value)) +
+#     labs(x='', y='seg. value', title=x[['label']], subtitle=paste('haz= ',x[['hazards']], ' coef= ',x[['coef']], '\nMW pval ',pv,sep='')) + theme(legend.position = 'none')
+# }
+
+regions = apply(ch, 1, melt.by.reg)
+names(regions) = ch$label
+
+plot.features<-function(reg) {
+  se = 2
+  cols = c('indianred3','slateblue2')
+  if ( as.numeric(reg$coef[1]) < 0) cols = c('green3', 'darkviolet')
+
+  mn = 0
+  if (min(reg$value) >= 0) {
+    mn = mean(reg$value)  
+  }
+
+  p = ggplot(reg, aes(labels, value, group=labels, fill=labels, color=labels)) + geom_violin(alpha=0.2) +
+    geom_jitter(width=0.05) +
+    geom_hline(yintercept=with(reg, c(mn+sd(value)*se, mn-sd(value)*se)), color='darkgrey', linetype='dashed') +
+    scale_fill_manual(values=cols) + scale_color_manual(values=cols) + theme(legend.position = 'none') + labs(x='', subtitle=reg$coef[1])
+
+   x = rbind.data.frame('P'= cbind( 'gain' = length(which(with(subset(reg, labels == 'P'), value >=  mn+sd(value)*se))),
+                                    'loss' = -length(which(with(subset(reg, labels == 'P'), value <=  mn-sd(value)*se))) ),
+
+                        'NP' = cbind( 'gain' = length(which(with(subset(reg, labels == 'NP'), value >=  mn+sd(value)*se))),
+                                      'loss' = -length(which(with(subset(reg, labels == 'NP'), value <=  mn-sd(value)*se))) ) )
+
+   x = x*sign(reg$coef[1])
+
+  p2 = ggplot(melt(as.matrix(x)), aes(Var2, value, fill=Var1, 
+                                      color=ifelse(value > 0, 'red', 'white'))) + #facet_grid(~Var1) +
+        geom_bar(stat='identity', width=0.5, position='dodge') + 
+        geom_hline(yintercept = 0, color='black', size=1.5) +
+        labs(x='', y='Counts', subtitle=reg$coef[1]) + 
+        scale_fill_manual(values=cols) + theme(legend.position = 'none')
+
+  return(list('plot'=arrangeGrob(p, p2, ncol=2, top=unique(reg$region)[1]), 'counts'=x))     
+}
+
+pf = lapply(regions, plot.features)
+
+do.call(grid.arrange, c(lapply(pf[1:12], function(x) x$plot), ncol=3) )
+
+```
+
+### Is there a difference in P~NP for gains/losses
+
+Based on the relationship between the coefficients and the values of the data we would expect one.  
+```{r}
+# positive coefficients were related to gains in progressors
+gains = subset(ch, coef > 0)
+# negative coefficients were related to losses in non-progressors
+losses = subset(ch, coef < 0)
+
+gain = do.call(rbind, lapply(gains$label, function(reg) {
+  values = dysplasia.df[,reg] 
+  table(labels[which(values > sd(values))])
+}))
+
+# There is a difference in the gains NP~P in the positive coefficients
+wilcox.test(gain[,1], gain[,2])
+summary(aov(gain[,1]~gain[,2]))
+
+loss = do.call(rbind, lapply(gains$label, function(reg) {
+  values = dysplasia.df[,reg] 
+  table(labels[which(values < -sd(values))])
+}))
+
+# There may be on in the losses as well, but less clear
+wilcox.test(loss[,1], loss[,2])
+summary(aov(loss[,1]~loss[,2]))
+
+# Compared to randomly selected regions (not the coefficients)
+sample.reg = sample(setdiff(colnames(dysplasia.df), ch$label), nrow(ch))
+
+randomGains = do.call(rbind, lapply(sample.reg, function(reg) {
+  values = dysplasia.df[,reg] 
+  table(labels[which(values > sd(values))])
+}))
+
+# P~NP may be different across randomly selected regions as well, again less clear
+wilcox.test(randomGains[,1], randomGains[,2])
+summary(aov(randomGains[,1]~randomGains[,2]))
+
+randomLosses = do.call(rbind, lapply(sample.reg, function(reg) {
+  values = dysplasia.df[,reg] 
+  table(labels[which(values < -sd(values))])
+}))
+
+# In this case losses show a definite difference
+wilcox.test(randomLosses[,1], randomLosses[,2])
+summary(aov(randomLosses[,1]~randomLosses[,2]))
 
@@
 
-try
+If gains or losses are primarily driving this, can a model that only looks at gains or losses predict better?
 
@@
 }
-
+```
+
+Gains alone are pretty much the same as the raw data.
+```{r warning=F, message=F, fig.height=8, fig.width=8}
 do.call(grid.arrange, c(gp$plots, top='All samples, GAINS 10fold, 5 splits'))
+```
+
+So are losses...
+```{r warning=F, message=F, fig.height=8, fig.width=8}
 do.call(grid.arrange, c(lp$plots, top='All samples, LOSSES 10fold, 5 splits'))
+```
+
+When we try to put them together, so each region has two columns where gains and losses are considered separately the classification drops to < 50%.
+So we're better off using the merged data we started with and gains/losses alone aren't simply driving the prediction.
+```{r warning=F, message=F, fig.height=8, fig.width=8}
 do.call(grid.arrange, c(both$plots, top='All samples, gains & losses 10fold, 5 splits'))
-
 ```
@@
   # sort in label order
-  r.dysplasia.df = t(mergedDf[,names(r.labels)])
+  r.dysplasia.df = dysplasia.df[names(r.labels),]
 
@@
     test = as.matrix(dysplasia.df[-train.rows,])
-    if ( nrow(test) == ncol(dysplasia.df) ) test = t(test)
+    if (ncol(test) <= 1) next
+    #if ( nrow(test) == ncol(dysplasia.df) ) test = t(test)
 
@@
   coord_flip() + labs(title='Correctly predicted HGD samples', y='Correct prediction ratio')
+
+by.path = do.call(rbind.data.frame, lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID], function(pt) {
+  final.endo = subset(pt, Endoscopy.Year == max(Endoscopy.Year))
+  final.endo = final.endo[nrow(final.endo),]
+
+  pred <- function(rows) length(which(rows$Prediction > cutoff))
+  if (pt$Status[1] == 'NP')
+    pred <- function(rows) length(which(rows$Prediction < cutoff))
+
+  h = subset(pt, Pathology %in% c('IMC','HGD'))
+  l = subset(pt, Pathology == 'LGD' & PID != final.endo$PID)
+  id = subset(pt, Pathology == 'ID' & PID != final.endo$PID)
+  be = subset(pt, Pathology %in% c('BE') & PID != final.endo$PID)
+
+  cbind('HGD'=pred(h), 'total.h'=nrow(h), 
+        'LGD'=pred(l), 'total.l'=nrow(l), 
+        'ID'=pred(id), 'total.id'=nrow(id),
+        'NDBE'=pred(be), 'total.be'=nrow(be))
+}))
+
+sum(by.path$NDBE)/sum(by.path$total.be)
+sum(by.path$LGD)/sum(by.path$total.l)
+sum(by.path$ID)/sum(by.path$total.id)
+sum(by.path$HGD)/sum(by.path$total.h)
 
@@
 
-```{r validation, message=F, warning=F, echo=F, fig.height=8, fig.width=6}
-
-score.cx <-function(pd) { ## OLD method
-#  print(pd$info)
-  len.var = (var(pd$seg.vals$end.pos - pd$seg.vals$start.pos))/1e12
-  extreme.values = max( apply(as.matrix(pd$seg.vals[,grep('^D', colnames(pd$seg.vals))]), 2, function(col) {
-    length(which(col >= (median(col)+2*sd(col)))) + length(which(col <= (median(col)-2*sd(col)))) 
-  }) )
-
-  (len.var + extreme.values)
-}
+```{r validation, message=F, warning=F, echo=F, fig.height=8, fig.width=6, eval=F}
 
@@
 
-complexity.measures = sapply(validation.patient.data, score.cx)
-validation.df = cbind(validation.df, 'cx.measure'=NA)
-for (sample in rownames(validation.df)) {
-  validation.df[sample, 'cx.measure'] = complexity.measures[subset(validation.patient.info, Samplename == sample)$Hospital.Research.ID]
-}
-# 
-# # norm to 1 to keep within the magnitude of the genomic data
-cx = validation.df[,'cx.measure']
-validation.df[,'cx.measure'] = (log(cx)/log(max(cx)))
+validation.df = score.cx(validation.patient.data, validation.df)
 
@@
 fitV = models[[select.alpha]]
-lambda.opt = performance[select.alpha, 'lambda']
+lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']
 
@@
 
-```{r, echo=F, message=F, warning=F, fig.width=8, fig.height=12}
+```{r, echo=F, message=F, warning=F, fig.width=8, fig.height=12, eval=F}
 
@@
 grid.arrange(p1, p2,ncol=1)
-```
-
-```{r echo=F, warning=F, message=F}
+
+by.path = do.call(rbind.data.frame, lapply(vpd, function(pt) {
+  final.endo = subset(pt, Endoscopy.Year == max(Endoscopy.Year))
+  final.endo = final.endo[nrow(final.endo),]
+
+  pred <- function(rows) length(which(rows$Prediction > cutoff))
+  if (pt$Status[1] == 'NP')
+    pred <- function(rows) length(which(rows$Prediction < cutoff))
+
+  h = subset(pt, Pathology %in% c('IMC','HGD'))
+  l = subset(pt, Pathology == 'LGD' & PID != final.endo$PID)
+  id = subset(pt, Pathology == 'ID' & PID != final.endo$PID)
+  be = subset(pt, Pathology %in% c('BE') & PID != final.endo$PID)
+
+  cbind('HGD'=pred(h), 'total.h'=nrow(h), 
+        'LGD'=pred(l), 'total.l'=nrow(l), 
+        'ID'=pred(id), 'total.id'=nrow(id),
+        'NDBE'=pred(be), 'total.be'=nrow(be))
+}))
+
+sum(by.path$HGD)/sum(by.path$total.h)
+sum(by.path$LGD)/sum(by.path$total.l)
+sum(by.path$ID)/sum(by.path$total.id)
+sum(by.path$NDBE)/sum(by.path$total.be)
+
+```
+
+```{r echo=F, warning=F, message=F, eval=F}
 

### 2017-09-28 Patient Report.Rmd
submitting before making major changes...

@@
 
+head(m2)
+m2
+
+print.png<-function(p, name, ...) {
+  png(filename=name,...)
+  print(p)
+  dev.off()
+}
+
+pg_genes = subset(m2, variable == 'Progressors')
+pg_genes = pg_genes[order(-pg_genes$value),]
+
+limits = seq(1, 220, 50)
+for (i in limits) {
+  max = i+49
+  if (max > nrow(pg_genes) ) max = nrow(pg_genes)
+
+  p = ggplot(subset(m2, Gene.Symbols %in% pg_genes$Gene.Symbols[c(i:max)]),  aes(y=Gene.Symbols, x=variable, fill=brks)) + geom_tile() +
+    scale_fill_manual(values = myP, 'Patient Counts') +
+    labs(title="All available COSMIC genes", x='', y='', subtitle=paste("genes ", i, "-", max, " by number of patients", sep='')) +
+    theme(text=element_text(size=16, face="bold"), axis.title=element_text(size=18))
+
+  if (i != 1)
+    p = p + theme(legend.position = 'none')
+
+  print.png(p, paste('cosmic_genes_', i, '-', max, '.png', sep=''), width=600, height=1000)
+}
+
 ```

### 2017-09-28 RegressionPvNP.Rmd
submitting before making major changes...

@@
 
-includeDemo = F # demographics
+includeDemo = T # demographics
 
@@
 dysplasia.df = apply(dysplasia.df, 2, unit.var)
-
 ```
@@
 
-## TODO score complexity on the seg.vals rather than merged data frame
+  ## TODO score complexity on the seg.vals directly rather than merged data frame?
   dysplasia.df = score.cx(patient.data, dysplasia.df)
@@
   }
-  save(plots, coefs, performance.at.1se, dysplasia.df, models, file=file)
+  save(plots, coefs, performance.at.1se, dysplasia.df, models, labels, file=file)
 }
@@
 
-```{r hazards, echo=F, warning=F, message=F,eval=T, fig.height=12, fig.width=12}
-
+```{r hazards, echo=F, warning=F, message=F,eval=T, fig.height=6, fig.width=6}
 hazards = apply( exp(t(dysplasia.df[, rownames(coefs[[select.alpha]])]) *  coefs[[select.alpha]][,1]), 1, function(x) {
@@
 ggplot(ch, aes(coef, hazards, col=hazards >= qq[length(qq)])) + 
-  geom_point(alpha=0.5) + 
+  geom_point(alpha=0.8) + 
   geom_text_repel(aes(x=coef, label=ifelse(hazards >= qq[length(qq)], label, '')),show.legend=F) +
-  labs(title='Top features by hazard ratio') + theme(legend.position = 'bottom') + theme(legend.position='none')
+  labs(title='Top features by hazard ratio') + theme(legend.position = 'none', text = element_text(size=14, face='bold')) 
 ```
@@
 
-### How do coefficients related to the raw data?
+### How do coefficients relate to the raw data?
 
@@
 
+### Examine a few example regions
+
+```{r, echo=F, eval=F}
+# p53 regions
+chr17 = get.chr.lengths()[17,]
+
+samples = names(which(dysplasia.df[,'17:1-4776188'] <= -1 ) )
+
+#patients = subset(patient.info, Samplename %in% samples)
+#table(patients$Pathology)
+#nrow(patients)/nrow(subset(patient.info, Patient %in% patients$Patient))
+
+values17 = dysplasia.df[samples, grep('^17:', colnames(dysplasia.df), value=T)]
+
+values17 = merge(values17, subset(patient.info, Samplename %in% samples , select=c('Patient','Samplename')), by.x='row.names', by.y='Samplename')
+values17 = melt(values17, measure.vars = grep('^17:',colnames(values17), value=T))
+colnames(values17) = c('sample','Patient','region','value')
+head(values17)
+
+values17$selected = as.integer( as.character(values17$region) == '17:1-4776188' )
+
+loc = do.call(rbind.data.frame, strsplit(as.character(values17$region), ':|-'))
+colnames(loc) = c('chr','start', 'end')
+loc[] = lapply(loc[], function(x) as.numeric(as.character(x)))
+
+values17 = cbind(values17, loc)
+
+genes = ccgenes[grep("BRCA1|SMARCE1|STAT3|STAT5B|TP53", ccgenes$Gene.Symbol),c('Gene.Symbol','Genome.Location')]
+
+loc = unlist(strsplit(ccgenes[which(ccgenes$Gene.Symbol == 'TP53'),'Genome.Location'], ':|-'))
+genes = cbind(genes, 'start'=loc[1], 'value'=-7, 'Row.names'='', 'Patient'='', 'selected'=as.factor(0))
+
+#loc = do.call(rbind.data.frame, strsplit(ccgenes[grep("TP53|BRCA1|SMARCE1|STAT3|STAT5B", ccgenes$Gene.Symbol),'Genome.Location'], ':|-'))
+#colnames(loc) = c('chr','start','end')
+genes = cbind.data.frame('chr'=as.numeric(loc[1]), 'start'=as.numeric(loc[2]), 'end'=as.numeric(loc[3]), 'Gene.Symbol'='TP53', 'selected'=as.factor(0), 'sample'='', value=-7)
+
+values17$sample = as.factor(values17$sample)
+values17$selected = as.factor(values17$selected)
+
+arms = cbind.data.frame('x'=rbind( with(chr17, chr.cent-cent.gap*3), with(chr17, chr.cent+cent.gap*3)), 'y'=rbind(10,10), 'label'=c('p arm','q arm'))
+head(values17)
+
+p = ggplot() + xlim(1,chr17$chr.length) + 
+    #geom_rect(data=values17, aes(xmin=start, xmax=end, ymin=0, ymax=value, fill=Patient), alpha=0.2, show.legend=F) +
+    geom_segment(data=values17, aes(x=start, xend=end, y=value, yend=value, color=Patient), show.legend=F, size=1.5) +
+    geom_segment(data=unique(subset(values17, selected == 1, select=c('chr','start','end'))), aes(x=start, xend=end, y=0, yend=0), size=2, color='red') +
+    geom_point(data=genes, aes(x=start,y=value), color='red', size=3, show.legend = F) +
+    geom_text_repel(data=genes, aes(x=end, y=value, label=Gene.Symbol), color='red', show.legend = F, nudge_y=-0.1) +
+    geom_vline(xintercept=chr17$chr.cent, color='red', linetype='longdash') +
+    geom_label(data=arms, aes(x,y, label=label), fill='white') +
+    geom_hline(yintercept = -1, color='grey39', linetype='longdash') + geom_hline(yintercept = 1, color='grey39', linetype='longdash') +
+    geom_label(data=cbind('x'=chr17$chr.cent/2, 'y'=0), aes(x, y), label='Losses', color='white', fill='grey39') + 
+    labs(y='Segment value', x='Chr position', title='Chromosome 17', subtitle=paste(length(table(values17$Patient)), 'patients,', length(samples), 'samples')) +
+  theme(text=element_text(size=16), axis.text=element_text(size=12))
+
+png(filename='chr17.png', width=600, height=600)
+print(p)
+dev.off()
+
+print(p)
+```
+
+```{r, eval=F}
+# BRCA etc
+regions = subset(ch, grepl('^17', label) & coef > 0)
+samples = which(dysplasia.df[,regions$label[1]] > 1 ) 
+samples = union(names(samples), names(which(dysplasia.df[,regions$label[2]] > 1 )) )
+
+cols = grep(paste(regions$label, collapse='|'), colnames(dysplasia.df))
+
+values17 = cbind(dysplasia.df[samples, c(cols-2,cols-1,cols,cols+1,cols+2)], 'labels'=labels[samples])
+values17 = merge(values17, subset(patient.info, Samplename %in% samples , select=c('Hospital.Research.ID','Patient','Samplename')), by.x='row.names', by.y='Samplename')
+
+values17 = melt(values17)
+
+values17$selected = as.integer(values17$variable %in% ch[grepl('^17', ch$label),'label'])
+
+loc = do.call(rbind.data.frame, strsplit(as.character(values17$variable), ':|-'))
+colnames(loc) = c('chr','start', 'end')
+loc[] = lapply(loc[], function(x) as.numeric(as.character(x)))
+
+values17 = cbind(values17, loc)
+
+genes = ccgenes[grep("BRCA1|SMARCE1|STAT3|STAT5B", ccgenes$Gene.Symbol),c('Gene.Symbol','Genome.Location')]
+loc = do.call(rbind.data.frame, strsplit(genes$Genome.Location, ':|-'))
+colnames(loc) = c('chr','start','end')
+loc[] = lapply(loc[], function(x) as.integer(as.character(x)))
+
+genes = cbind.data.frame(genes, loc, 'value'=as.integer(5), 'Row.names'='', 'Patient'='', 'selected'=as.factor(0))
+
+values17$Row.names = as.factor(values17$Row.names)
+values17$selected = as.factor(values17$selected)
+head(values17)
+ggplot(values17, aes(start, value)) + # xlim(1,chr17$chr.length) + #facet_wrap(~Patient, ncol=1) +
+    geom_segment(aes(x=start, xend=end, y=value, yend=value, fill=Patient, color=Patient), show.legend=F) + 
+    geom_segment(data=unique(subset(values17, selected == 1, select=c('chr','start','end'))), aes(x=start, xend=end, y=0, yend=0), size=2, color='red') +
+    geom_point(data=genes, aes(x=start,y=value), color='red', size=3, show.legend = F) +
+    geom_text_repel(data=genes, aes(x=end, y=value, label=Gene.Symbol), color='red', show.legend = F) +
+    geom_hline(yintercept = 1, color='grey39', linetype='longdash') +
+    geom_label(x=median(values17$start, na.rm=T), y=7, label='Gains', color='white', fill='grey39') + 
+    labs(y='Segment value', x='chr17', title='chr17')
+
+png(filename='chr17.png', width=800, height=1000)
+print(p)
+dev.off()
+
+# FGFR2 region
+chr10 = get.chr.lengths()[10,]
+
+samples = which(dysplasia.df[,'10:121013167-125853693'] > 1)
+
+cols = grep('^10:121013167-125853693', colnames(dysplasia.df))
+values10 = cbind(dysplasia.df[samples, c(cols-1, cols, cols+1)], 'labels'=labels[samples])
+values10 = merge(values10, subset(patient.info, Samplename %in% names(samples) , select=c('Hospital.Research.ID','Patient','Samplename')), by.x='row.names', by.y='Samplename')
+
+values10 = melt(values10)
+
+loc = do.call(rbind.data.frame, strsplit(as.character(values10$variable), ':|-'))
+colnames(loc) = c('chr','start', 'end')
+loc[] = lapply(loc[], function(x) as.numeric(as.character(x)))
+
+values10 = cbind(values10, loc)
+
+loc = unlist(strsplit(ccgenes[which(ccgenes$Gene.Symbol == 'FGFR2'),'Genome.Location'], ':|-'))
+fgfr2 = cbind.data.frame('chr'=as.numeric(loc[1]), 'start'=as.numeric(loc[2]), 'end'=as.numeric(loc[3]), 'Row.names'='FGFR2', 'Patient'='', value=5)
+
+values10$Row.names = as.factor(values10$Row.names)
+head(values10)
+ggplot(values10, aes(start, value, fill=Patient, color=value)) + #xlim(1,chr10$chr.length) + #facet_wrap(~Patient, ncol=1) +
+    geom_segment(aes(x=start, xend=end, y=value, yend=value), show.legend = F) + 
+    geom_point(data=fgfr2, aes(x=start,y=value), color='red', size=3, show.legend = F) +
+    geom_text_repel(data=fgfr2, aes(x=end, y=value, label=Row.names), show.legend = F, color='red') +
+    #geom_vline(xintercept=chr10$chr.cent, color='red', linetype='longdash') +
+    #geom_label(x=chr10$chr.cent-chr10$cent.gap*3, y=10, label='p arm', color='white') + 
+    #geom_label(x=chr10$chr.cent+chr10$cent.gap*3, y=10, label='q arm', color='white') +
+    geom_hline(yintercept = 1, color='grey39', linetype='longdash') +
+    #geom_label(x=chr10$chr.cent/2, y=0, label='Losses', color='white', fill='grey39') + 
+    labs(y='Segment value', x='chr10', title='chr10')  + scale_color_continuous(low='red',high='green')
+
+```
+
 ### Is there a difference in P~NP for gains/losses
@@
 ```{r}
+features.binom <- function(df) {
+  p.adjust(apply(df, 1, function(x) {
+    binom.test(x[2], sum(x), 0.5)$p.value
+  }))  
+}
+
 # positive coefficients were related to gains in progressors
@@
 
+which(features.binom(gain) < 0.05)
+
+fisher.test(gain, simulate.p.value = T)
+
 loss = do.call(rbind, lapply(gains$label, function(reg) {
@@
 
+which(features.binom(loss) < 0.05)
+
 # Compared to randomly selected regions (not the coefficients)
@@
 summary(aov(randomGains[,1]~randomGains[,2]))
+fisher.test(randomGains, simulate.p.value = T)
+
+which(features.binom(randomGains) < 0.05)
 
@@
 summary(aov(randomLosses[,1]~randomLosses[,2]))
+
+which(features.binom(randomLosses) < 0.05)
 
@@
 ```{r leaveoneout, echo=T, message=F, warning=F}
-
 info = do.call(rbind, lapply(patient.data, function(df) df$info))
@@
     test = as.matrix(dysplasia.df[-train.rows,])
-    if (ncol(test) <= 1) next
-    #if ( nrow(test) == ncol(dysplasia.df) ) test = t(test)
+    #if (ncol(test) <= 1) next
+    if ( nrow(test) == ncol(dysplasia.df) ) test = t(test)
 
@@
 
-cutoff = pROC::coords(roc, 'best')[['threshold']]
+cutoff = round(pROC::coords(roc, 'best')[['threshold']], 2)
 cutoff = 0.5
@@
 This argues that we should set the prediction probability cutoff to `r pROC::coords(roc, 'best')[['threshold']]`.
-However, for now the predictions from here on are based on the default `r cutoff` value.
 
@@
 
-__All predictions from here onwards are based on the `r cutoff` cutoff, not the ROC curve__
+__All predictions from here onwards are performed using the `r cutoff` cutoff__
 
@@
   #pred <- function(rows) length(which(rows$Prediction > 0.5))/nrow(rows)
+  if (pt$Status[1] == 'P') {
     pred <- function(rows) length(which(rows$Prediction > cutoff))
+  } else {
+     pred <- function(rows) length(which(rows$Prediction < cutoff))
+  }
 
@@
   geom_text(aes(y=value+2,label=paste(round(value, 1), '%', sep=''))) +
-  geom_text(aes(y=3,label=paste('n=', total, sep=''))) +
-  ylim(0,100) + labs(title='Samples predicted by pathology', x='Pathology', y='Percent identified')
+  geom_text(aes(y=3,label=paste('n=', total, sep='')), color='white') +
+  ylim(0,100) + labs(title='Progressor samples predicted by pathology', x='Pathology', y='Percent identified') + 
+  theme(text=element_text(size=14, face='bold'), panel.background=element_blank(), axis.line=element_line(color='black'), panel.grid.major.x=element_blank(),
+        panel.grid.major.y=element_line(color='grey'))
 ```
@@
     breaks=c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 12*12, 15*12))
+endos$brks = with(endos, factor(brks, rev(levels(brks)))) # invert order
 
 endos$labels = gsub('\\[|\\(|\\]', '',endos$brks)
-endos$labels = sub(',', '-',endos$labels)
-
+endos$labels = sapply(strsplit( endos$labels, ',' ), function(x) paste(sort(as.numeric(x), decreasing=T), collapse='-' ))
+#endos$labels = sub(',', '-',endos$labels)
+
+## Flip order, 0-1 comes last
 m = endos %>% group_by(brks, labels) %>% summarise( 'pred.r'= length(which(pred > 0))/length(brks), 'n'=length((Patient)) )
+
 ggplot(m, aes(brks, pred.r*100, group=brks)) + ylim(0,105) +
   geom_bar(stat='identity', fill='dodgerblue3') + 
-  geom_text(aes(y=25, label=paste(round(pred.r*100), '%', sep='')), nudge_y=-0.02) + 
+  geom_text(aes(y=25, label=paste(round(pred.r*100), '%', sep='')), nudge_y=-0.02, color='white') + 
   geom_text(aes(y=pred.r*100+1.5, label=paste('n=',n, sep='')), nudge_y=0.02) + 
   scale_x_discrete(labels=m$labels) + 
-  labs(x='Months', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = number of endoscopies')
+  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = number of endoscopies') +
+  theme(text=element_text(size=14, face='bold'), axis.text.x=element_text(angle=45,hjust=1), panel.grid.major.x=element_blank(),   panel.grid.major.y=element_line(color='grey') , axis.line=element_line(color='black'), panel.background=element_blank())
 
@@
     df$endo.pred = ifelse(df$pred > 0, 'Pos','Neg')
-    ggplot(df, aes(months.before.final, endo.pred, group=1,color=Patient)) + facet_grid(Patient ~.) + 
+    ggplot(df, aes(months.before.final, endo.pred, group=1, color=endo.pred)) + facet_grid(Patient ~.) + 
       scale_x_continuous(breaks=seq(minM, maxM, by = 12)) +
-      geom_line() + geom_point() +
-      geom_label_repel(aes(label=max.path, fill=Patient, color=Patient), color='white', fontface='bold', size=4, label.size=0, label.r=unit(0.25, 'lines'), label.padding=unit(0.15, 'lines')) +
+      geom_line( color='grey39') + geom_point(aes(color=endo.pred)) +
+      geom_label_repel(aes(label=max.path, fill=endo.pred), color='white', fontface='bold', size=4, label.size=0, label.r=unit(0.25, 'lines'), label.padding=unit(0.15, 'lines'), show.legend=F) + 
+      scale_fill_manual(values=c('red4', 'green4')) + 
+      scale_color_manual(values=c('red4', 'green4'), name='Endoscopy Prediction') +
       labs(x='Months before diagnosis', y='Predictive Endoscopy') + 
-      theme(text=element_text(size=14, face="bold"), axis.title=element_text(size=16), legend.position='none')
+      theme(text=element_text(size=14, face="bold"), axis.title=element_text(size=16), legend.position='bottom')
   } else {
-  ggplot(df, aes(months.before.final, pred.ratio, group=1, color=Patient, fill=Patient)) + facet_grid(Patient ~.) +
+  ggplot(df, aes(months.before.final, pred.ratio, group=1, color=pred.ratio, fill=pred.ratio)) + facet_grid(Patient ~.) +
       ylim(0,1) + geom_abline(slope=0, intercept=seq(.25,.75,.25), color='grey', linetype='dashed') +
-      geom_col(position='dodge', aes(y=pred.ratio)) + geom_line() +
+      geom_col(position='dodge', aes(y=pred.ratio), color='white') +# geom_line() +
       scale_x_continuous(breaks=seq(minM, maxM, by = 12)) + 
-      geom_label(aes(y=ifelse(pred.ratio==1, 0.75, pred.ratio), label=max.path, fill=Patient), color='white', fontface='bold', size=4, label.size=0, label.r=unit(0.25, 'lines'), label.padding=unit(0.15, 'lines'), nudge_y=0.1) +
+      geom_label(aes(y=ifelse(pred.ratio==1, 0.75, pred.ratio), label=max.path, fill=pred.ratio), color='white', fontface='bold', size=4, label.size=0, label.r=unit(0.25, 'lines'), label.padding=unit(0.15, 'lines'), nudge_y=0.1) +
       labs(x='Months before diagnosis', y='Sample prediction ratio') + 
-      theme(text=element_text(size=14, face="bold"), axis.title=element_text(size=16), legend.position='none')
+      scale_fill_continuous(low='orangered4', high='green4', name='Sample Prediction Ratio') +
+      theme(text=element_text(size=14, face="bold"), axis.title=element_text(size=16), legend.position='bottom')
   }
@@
   labs(title="Progressors", subtitle=paste(72/12, "years or more follow-up"))
-print.png(p, 'progressors_1.png', width=900, height=1200)
+#print.png(p, 'progressors_1.png', width=900, height=1200)
 grid.arrange(p)
@@
   labs(title="Progressors", subtitle=paste(96/12, "years or more follow-up"))
-print.png(p1a, 'progressors_1a.png', width=600, height=800)
+#print.png(p1a, 'progressors_1a.png', width=600, height=800)
 
@@
   labs(title="Progressors", subtitle=paste(72/12, "-",96/12,"years or more follow-up"))
-print.png(p1b, 'progressors_1b.png', width=600, height=800)
+#print.png(p1b, 'progressors_1b.png', width=600, height=800)
 
@@
   labs(title="Progressors", subtitle=paste(72/12, "years or more follow-up"))
-print.png(p1, 'progressors_1_bar.png', width=900, height=1200)
+#print.png(p1, 'progressors_1_bar.png', width=900, height=1200)
 
@@
   labs(title="Progressors", subtitle=paste(48/12, "-",72/12,  "years or more follow-up"))
-print.png(p1, 'progressors_2.png', width=400, height=600)
+#print.png(p1, 'progressors_2.png', width=400, height=600)
 
@@
   labs(title="Progressors", subtitle=paste(48/12, "-",72/12,  "years or more follow-up"))
-print.png(p1, 'progressors_2_bar.png', width=400, height=600)
+#print.png(p1, 'progressors_2_bar.png', width=400, height=600)
 
@@
   labs(title="Progressors", subtitle=paste(24/12, "-",48/12,  "years or more follow-up"))
-print.png(p2, 'progressors_3.png', width=400, height=600)
+#print.png(p2, 'progressors_3.png', width=400, height=600)
 
@@
   labs(title="Progressors", subtitle=paste(24/12, "-",48/12,  "years or more follow-up"))
-print.png(p2, 'progressors_3_bar.png', width=400, height=600)
+#print.png(p2, 'progressors_3_bar.png', width=400, height=600)
 
@@
   labs(title="Progressors", subtitle=paste("1 year follow-up"))
-print.png(p3, 'progressors_4.png', width=400, height=600)
+#print.png(p3, 'progressors_4.png', width=400, height=600)
 
@@
   labs(title="Progressors", subtitle=paste("1 year follow-up"))
-print.png(p3, 'progressors_4_bar.png', width=400, height=600)
+#print.png(p3, 'progressors_4_bar.png', width=400, height=600)
 
@@
   labs(title="Non-Progressors", subtitle=paste("Patients with predictive endoscopies, max", abs(min(me$months)/12), 'years follow-up'))
-print.png(p, 'non-progressors.png', width=600, height=800)
+#print.png(p, 'non-progressors.png', width=600, height=800)
 
@@
   labs(title="Non-Progressors", subtitle=paste("Patients with predictive endoscopies, max", abs(min(me$months)/12), 'years follow-up'))
-print.png(p1, 'non-progressors_bar.png', width=600, height=800)
+#print.png(p1, 'non-progressors_bar.png', width=600, height=800)
 
@@
 
-```{r validation, message=F, warning=F, echo=F, fig.height=8, fig.width=6, eval=F}
+```{r validation, message=F, warning=F, echo=F, fig.height=8, fig.width=6, eval=T}
 
@@
 if (nrow(subset(np,pred.ratio == 0)) > 0) {
-  p2 = p2 +  geom_text( aes(x=subset(np,pred.ratio == 0)$Patient[1], y=0.5, label=paste(round(nrow(pn[pn$pred.ratio <= 0,])/nrow(np),3)*100,'%',sep='')), size=8, fontface='bold' ) 
+  p2 = p2 +  geom_text( aes(x=subset(np,pred.ratio == 0)$Patient[1], y=0.5, label=paste(round(nrow(np[np$pred.ratio <= 0,])/nrow(np),3)*100,'%',sep='')), size=8, fontface='bold' ) 
 }
@@
 
-```{r, echo=F, message=F, warning=F, fig.width=8, fig.height=12, eval=F}
+```{r, echo=F, message=F, warning=F, fig.width=8, fig.height=12, eval=T}
 
@@
 p1 = plot.endo(subset(vpi, Patient %in% subset(sum.validation.info, Status == 'P')$Patient), bin=T) + labs(title="Validation - Progressors")
-print.png(p1, 'validation_progressors.png', width=600, height=800)
-
-print.png(plot.endo(subset(vpi, Patient %in% subset(sum.validation.info, Status == 'P')$Patient), bin=F) + labs(title="Validation - Progressors"), 'validation_progressors_bar.png', width=600, height=800)
+#print.png(p1, 'validation_progressors.png', width=600, height=800)
+
+#print.png(plot.endo(subset(vpi, Patient %in% subset(sum.validation.info, Status == 'P')$Patient), bin=F) + labs(title="Validation - Progressors"), 'validation_progressors_bar.png', width=600, height=800)
 
@@
 
-print.png(p2,'validation_Nonprogressors.png', width=600, height=800)
-print.png(plot.endo(subset(vpi, Patient %in% subset(sum.validation.info, Status == 'NP')$Patient), bin=F) + labs(title="Validation - Non Progressors"),'validation_Nonprogressors_bar.png', width=600, height=800)
+#print.png(p2,'validation_Nonprogressors.png', width=600, height=800)
+#print.png(plot.endo(subset(vpi, Patient %in% subset(sum.validation.info, Status == 'NP')$Patient), bin=F) + labs(title="Validation - Non Progressors"),'validation_Nonprogressors_bar.png', width=600, height=800)
 
@@
 
-by.path = do.call(rbind.data.frame, lapply(vpd, function(pt) {
+global.by.path = do.call(rbind.data.frame,
+                         lapply(vpd[subset(sum.validation.info, Status == 'P')$Hospital.Research.ID], function(pt) {
   final.endo = subset(pt, Endoscopy.Year == max(Endoscopy.Year))
@@
 
+  #pred <- function(rows) length(which(rows$Prediction > 0.5))/nrow(rows)
+  if (pt$Status[1] == 'P') {
     pred <- function(rows) length(which(rows$Prediction > cutoff))
-  if (pt$Status[1] == 'NP')
+  } else {
      pred <- function(rows) length(which(rows$Prediction < cutoff))
+  }
 
@@
 
-sum(by.path$HGD)/sum(by.path$total.h)
-sum(by.path$LGD)/sum(by.path$total.l)
-sum(by.path$ID)/sum(by.path$total.id)
-sum(by.path$NDBE)/sum(by.path$total.be)
+global.by.path = colSums(global.by.path)
+
+global.path = as.data.frame(matrix(global.by.path, byrow=T, nrow=4, dimnames=list(c('HGD','LGD','ID','NDBE'),c('pred', 'total'))))
+global.path$ratio = (global.path$pred/global.path$total)*100
+global.path$path = rownames(global.path)
+
+global.path$path = factor(global.path$path, levels=c('HGD','LGD','ID','NDBE'))
+
+ggplot(melt(global.path[,c('ratio','total','path')], measure.vars='ratio', id.vars=c('path', 'total')), aes(path, value, group=path)) + 
+  geom_bar(stat='identity', fill='seagreen4') + 
+  geom_text(aes(y=value+2,label=paste(round(value, 1), '%', sep=''))) +
+  geom_text(aes(y=3,label=paste('n=', total, sep='')), color='white') +
+  ylim(0,100) + labs(title='Validation samples predicted by pathology', x='Pathology', y='Percent identified') + 
+  theme(text=element_text(size=14, face='bold'), panel.background=element_blank(), axis.line=element_line(color='black'), panel.grid.major.x=element_blank(),
+        panel.grid.major.y=element_line(color='grey'))
+
+```
+
+# Chr 17 Predictive
+
+One way of looking at the p53 prediction (e.g. Adam Bass) is to see how predictive the p-arm loss in chr17 is.
+
+```{r, eval=F}
+a=1
+
+plot.cv<-function(cv, title='') {
+  grid.arrange( arrangeGrob(cv$plot+ggtitle('Classification'), cv$deviance.plot+ggtitle('Binomial Deviance'), top=title, ncol=2) )
+}
+
+more.l<-function(lambda) {
+    l = sort(c(lambda, seq(exp(-5), exp(-10), -1e-6),
+            seq(exp(-10), exp(-15), -1e-8),
+            seq(exp(-15), exp(-20), -1e-9)), decreasing=T)
+}
+
+# random
+df11 = dysplasia.df[,grep('^11:', colnames(dysplasia.df))]
+fit0 <- glmnet(df11, labels, alpha=a, nlambda=nl, family='binomial', standardize=F) # all patients
+cv11 = crossvalidate.by.patient(x=df11, y=labels, lambda=more.l(fit0$lambda), pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', standardize=F)
+plot.cv(cv11,"Chr11")
+
+df17 = dysplasia.df[,grep('^17:', colnames(dysplasia.df))]
+fit0 <- glmnet(df17, labels, alpha=a, nlambda=nl, family='binomial', standardize=F) # all patients
+cv17 = crossvalidate.by.patient(x=df17, y=labels, lambda=more.l(fit0$lambda), pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', standardize=F)
+plot.cv(cv17, 'Chr17')
+
+## No HGD
+df17 = dysplasia.df[subset(patient.info, Pathology %in% c('BE','ID','LGD'))$Samplename,grep('^17:', colnames(dysplasia.df))]
+fit0 <- glmnet(df17, labels[subset(patient.info, Pathology %in% c('BE','ID','LGD'))$Samplename], alpha=a, nlambda=5000, family='binomial', standardize=F)
+cv17 = crossvalidate.by.patient(x=df17, y=labels[subset(patient.info, Pathology %in% c('BE','ID','LGD'))$Samplename], lambda=fit0$lambda, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', standardize=F)
+plot.cv(cv17, 'Chr17, no HGD')
+
+df3 = dysplasia.df[,grep('^3:', colnames(dysplasia.df))]
+fit0 <- glmnet(df3, labels, alpha=a, nlambda=nl, family='binomial', standardize=F) # all patients
+cv3 = crossvalidate.by.patient(x=df3, y=labels, lambda=fit0$lambda, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', standardize=F)
+plot.cv(cv3, 'Chr3')
+
+```
+
+# Interesting plots
+
+```{r echo=F, warning=F, message=F, eval=T, fig.width=12, fig.height=8}
+chr.lengths = get.chr.lengths()
+chr.lengths$chrom = sub('chr','',chr.lengths$chrom)
+
+df = mergedDf[,c('chr','start','end')]
+df = cbind(df, t(dysplasia.df[,grep('^\\d+:', colnames(dysplasia.df), value=T)]))
+df = merge(df, chr.lengths[,c('chrom','chr.length')], by.x='chr', by.y='chrom')
+df = melt(df, id.vars=c('chr','start','end', 'chr.length'))
+df$Status = ifelse(df$variable %in% names(which(labels == 1)), 'Progressor', 'Non-Progressor')
+
+df = merge(df, patient.info[,c('Hospital.Research.ID','Patient','Samplename','Pathology')], by.x='variable', by.y='Samplename')
+
+df = subset(df, value != 0)
+
+cn = function(x) {
+  str = 'neutral'
+  if(x <= -1) {
+    str = 'loss'
+  } else if (x >= 1) {
+    str = 'gain'
+  }
+  return(str)
+}
+
+# annotation = as.data.frame(bmGR[which(bmGR$COSMIC | bmGR$hgnc_symbol == 'TP53')])
+# colnames(annotation) = c('chr', colnames(annotation)[-1])
+# head(df)
+
+df$cn = factor(sapply(df$value, cn), levels=c('loss','neutral','gain'), ordered = T)
+df$cn = factor(sapply(df$value, cn), levels=c('loss','neutral','gain'), ordered = T)
+df$cnB = factor(as.integer(df$cn != 'neutral'))
+
+ggplot(df, aes(x=chr.length)) + facet_grid(Status~chr, scales='free_x') + 
+  geom_rect(data=df, aes(xmin=start, xmax=end, ymin=0, ymax=value, fill=cn), alpha=0.5, show.legend = T) + 
+  #scale_fill_manual(values=c('grey88','green4')) +
+  scale_color_manual(values=c('firebrick3','grey88','green4')) +
+  scale_fill_manual(values=c('firebrick3','grey88','green4')) +
+  #scale_fill_gradient2(high="green4", mid='grey88', low ="firebrick3", name='') +
+  labs(title="Segment Values Across Samples", x='Chromosomes', y='Mean adjusted segmentation value',subtitle=paste('NonP:Prog - ', paste(table(patient.info$Status), collapse=':'))) +
+  theme(text=element_text(face='bold', size=14), axis.text.x=element_blank(), legend.position='bottom', 
+        panel.grid = element_blank(), panel.background = element_blank(),strip.background =element_rect(fill="lightcyan2"),  
+        panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0.1, 'lines') )
+
+ggplot(subset(df, Pathology == 'BE'), aes(x=chr.length)) + facet_grid(Status~chr, scales='free_x') + ylim(range(df$value)) +
+  geom_rect(data=subset(df, Pathology == 'BE'), aes(xmin=start, xmax=end, ymin=0, ymax=value, fill=cn), alpha=0.5, show.legend = T) + 
+  #geom_segment(data=subset(df, Pathology %in% c('BE','LGD','HGD')),aes(x=start, xend=end, y=value, yend=value, color=Pathology), size=1.5) +
+  #scale_fill_manual(values=c('grey88','green4')) +
+  scale_color_manual(values=c('firebrick3','grey88','green4')) +
+  scale_fill_manual(values=c('firebrick3','grey88','green4')) +
+  #scale_fill_gradient2(high="green4", mid='grey88', low ="firebrick3", name='') +
+  labs(title="Segment Values Across BE Samples", x='Chromosomes', y='Mean adjusted segmentation value', subtitle=paste('NonP:Prog - ', paste(table(subset(patient.info, Pathology == 'BE')$Status), collapse=':'))) +
+  theme(text=element_text(face='bold', size=14), axis.text.x=element_blank(), legend.position='bottom', 
+        panel.grid = element_blank(), panel.background = element_blank(),strip.background =element_rect(fill="lightcyan2"),  
+        panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0.1, 'lines') )
+
+patient_labeller <- function(string, prefix = "Patient") paste(prefix, string)
+dfpts = subset(df, Patient %in% c('57','76'))
+ggplot(dfpts, aes(x=chr.length)) + ylim(-5,10) +
+  facet_grid(Patient~chr, scales='free_x', labeller = labeller(Patient = patient_labeller )) + 
+  geom_rect(data=dfpts, aes(xmin=start, xmax=end, ymin=0, ymax=value, fill=cn), show.legend = F) + 
+  scale_fill_manual(values=c('firebrick3','grey88','green4')) +
+  labs(title="Segment Values Across Samples", x='Chromosomes', y='Mean adjusted segmentation value', subtitle=paste('NonP:Prog - ', paste(table(subset(patient.info, Patient %in% c('57','76'))$Status), collapse=':'))) +
+  theme(text=element_text(face='bold', size=14), axis.text.x=element_blank(), legend.position='bottom', 
+        panel.grid = element_blank(), panel.background = element_blank(),strip.background =element_rect(fill="lightcyan2"),  
+        panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0.1, 'lines') )
+
+patient_labeller <- function(string, prefix = "Patient") paste(prefix, string)
+dfpts = subset(df, Patient %in% c('57','76') & Pathology == 'BE')
+ggplot(dfpts, aes(x=chr.length)) + ylim(-5,10) +
+  facet_grid(Patient~chr, scales='free_x', labeller = labeller(Patient = patient_labeller )) + 
+  geom_rect(data=dfpts, aes(xmin=start, xmax=end, ymin=0, ymax=value, fill=cn), show.legend = F) + 
+  scale_fill_manual(values=c('firebrick3','grey88','green4')) +
+  labs(title="Segment Values Across BE Samples", x='Chromosomes', y='Mean adjusted segmentation value', subtitle=paste('NonP:Prog - ', paste(table(subset(patient.info, Patient %in% c('57','76') & Pathology == 'BE')$Status), collapse=':'))) +
+  theme(text=element_text(face='bold', size=14), axis.text.x=element_blank(), legend.position='bottom', 
+        panel.grid = element_blank(), panel.background = element_blank(),strip.background =element_rect(fill="lightcyan2"),  
+        panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0.1, 'lines') )
+
+```
+
+## Probability for each state?
+
+```{r, eval=F}
+
+x = endos %>% group_by(Patient, max.path) %>% summarise (  min(Endoscopy.Year)  )
+
+## Skip "indeterminate" as a state?
+x[x$max.path == 'ID', 'max.path'] = 'BE'
+
+cols = c('Patient','start.path','end.path','start.year','end.year', 'year.diff')
+state.changes = as.data.frame(matrix(ncol=length(cols), nrow=0, dimnames=list(c(), cols )))
+
+x$max.path
+for (p in unique(x$Patient)) {
+  ep = subset(x, Patient == p)
+  if (nrow(ep) <= 1) next
+
+  pairs = cbind( c(1:nrow(ep))[-nrow(ep)], c(2:nrow(ep)) )
+  df = do.call(rbind, apply(pairs, 1, function(pair) {
+    a = ep[ pair[1], ]
+    b = ep[ pair[2], ]
+
+    cbind.data.frame(a$Patient, a$max.path, b$max.path, a$`min(Endoscopy.Year)`, b$`min(Endoscopy.Year)`, (b$`min(Endoscopy.Year)`-a$`min(Endoscopy.Year)`))
+  }))
+  colnames(df) = cols  
+  state.changes = rbind(state.changes,df)
+}
+state.changes$start.path = ordered( state.changes$start.path, levels=levels(endos$max.path))
+state.changes$end.path = ordered( state.changes$end.path, levels=levels(endos$max.path))
+
+# Sanity check
+as.integer(state.changes$end.path) > as.integer(state.changes$start.path)
+
+state.changes = state.changes %>% group_by(Patient) %>% mutate( 'diagnosis'=ifelse(end.year == max(end.year), 1, 0) )
+
+# Sometimes a later sample is at a lower pathology than the earlier one.
+This state change is due to several factors.
+LGD is hard to diagnose accurately so ID would be even less accurate, and heterogenity in the BE segment itself means that one sample could be BE while another sample further down could be LGD and the samples for which we had enough to sequence can mean we miss out on the higher path ones
+pander(as.data.frame(state.changes))
+
+```
+
+### State changes
+
+```{r, eval=F}
+
+bl = subset(state.changes, start.path == 'BE' & end.path == 'LGD')
+
+bh = subset(state.changes, start.path == 'BE' & end.path %in% c('IMC','HGD'))
+summary(bh$year.diff)
+
+lh = subset(state.changes, start.path == 'LGD' & end.path %in% c('IMC','HGD'))
+summary(lh$year.diff)
+
+```
+
+## cox model - Time to progress?
+
+```{r cox, warning=F, message=F, echo=T}
+library(survival)
+cox.pvalue<-function(cox, test='logrank') {
+  pvalue = NULL  
+  csum = summary(cox)
+
+  if (test == 'logrank') {
+    pvalue = csum$sctest[['pvalue']]  
+  } else if (test == 'wald') {
+    pvalue = csum$waldtest[['pvalue']]
+  } else if (test == 'loglik') {
+    pvalue = csum$logtest[['pvalue']]
+  }
+  return(pvalue)
+}
+
+prog.pred = lapply(pg.samp, function(df) {
+  print(df$Hospital.Research.ID[1])
+  df = df[,c('Hospital.Research.ID', 'Status', 'Patient', 'PID', 'Status', 'Endoscopy.Year','Initial.Endoscopy', 'Final.Endoscopy', 'Pathology', 'Samplename','Prediction','Prediction.Dev.Resid')] %>% rowwise() %>% mutate(
+      'Time.to.Final'=Final.Endoscopy-Endoscopy.Year)
+  df = arrange(df, Endoscopy.Year, Pathology)
+
+  df$Dup.PID = duplicated(df$PID)
+  x = df[!df$Dup.PID,]
+
+  for (i in 1:nrow(x)) {
+    if (i == 1) {
+      x[i,'tstart'] = 0
+      x[i, 'tstop'] = x[(i+1), 'Endoscopy.Year'] - x[i, 'Endoscopy.Year']
+      if (is.na(x[i,'tstop']) || x[i,'tstop'] == 0) x[i,'tstop'] = 0.5
+      next
+    }
+
+    x[i,'tstart'] = x[(i-1), 'tstop']
+
+    # Final sample is different -- will leave it out later
+    if (i == nrow(x)) {
+      x[i, 'tstop'] = x[i,'tstart'] + 0.5
+      break
+    }
+
+    x[i, 'tstop'] =  x[i,'tstart'] + x[(i+1), 'Endoscopy.Year'] - x[i, 'Endoscopy.Year']
+
+    # Samples taken in the same year
+    if ( x[i, 'Endoscopy.Year'] == x[(i+1), 'Endoscopy.Year'] )
+      x[i,'tstop'] = x[i,'tstop'] + 0.5 
+  }
+
+  x$Diagnosis = as.integer(x$Pathology %in% c('HGD','IMC'))
+
+  if (sum(x$Diagnosis) >= 1)
+    x = x[-which(x$Diagnosis == 1),]
+
+  if (nrow(x) > 0 && sum(x$Diagnosis) <= 0 && x$Status == 'P')
+    x[nrow(x), 'Diagnosis'] = 1
+
+  if (nrow(x) > 0) {
+    x$Dup.Samples = ''
+    for (j in 1:nrow(df[df$Dup.PID,]))  {
+      row = df[df$Dup.PID,][j,]
+      x[which(x$PID == row[['PID']]), 'Dup.Samples'] = paste(x[which(x$PID == row[['PID']]), 'Dup.Samples'], row[['Samplename']], sep=',')
+    }
+    x$Dup.Samples = sub('^,','',x$Dup.Samples)
+  }
+
+  return(x)
+})
+
+## By pathology?
+pp = do.call(rbind, prog.pred)
+pp[is.na(pp$tstop), 'tstop'] = 0.5
+pp$Pathology = as.character(pp$Pathology)
+
+sfit = survfit( Surv(tstart, tstop, Diagnosis, type='interval') ~ Pathology, data=pp, conf.int=0.95, conf.type='log', na.action=na.omit)
+ggsurv(sfit,cens.shape=3,lty.est=1,back.white=F)
+summary(sfit)
+
+cox = coxph( Surv(tstart, tstop, Diagnosis) ~ Pathology, data=pp) 
+ggsurv(survfit(cox),cens.shape=3,lty.est=1,back.white=F,
+      main="coxph(Survival, Diagnosis) ~ Pathology" ) + 
+      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),2),"\n", nrow(pp), 'samples'))
+#summary(survfit(cox))
+#summary(cox)
+pander(cox)
+
+## All genomic regions
+sample.pred = merge( do.call(rbind.data.frame, prog.pred), dysplasia.df, by.x='Samplename', by.y='row.names' )
+sample.pred$Pathology = as.character(sample.pred$Pathology)
+
+gcols = grep('\\d+:', colnames(sample.pred), value=T)
+sample.pred = sample.pred[c('tstart', 'tstop','Diagnosis','Pathology',gcols)]
+
+# all
+cox = coxph( Surv(tstart, tstop, Diagnosis) ~., data=sample.pred)
+pander( cox$concordance )
+ggsurv(survfit(cox))
+
+# glm selected features
+cox2 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', 'Pathology',  ch$label)]) 
+ggsurv(survfit(cox2))
+summary(cox2)
+pander(cox2)
 
@@
 
-print.png(p3, 'validation_predictions.png', width=600, height=800)
-
-```
+#print.png(p3, 'validation_predictions.png', width=600, height=800)
+
+```

### 2017-09-28 features-OAC.Rmd
submitting before making major changes...

---
title: "Progression Features & OAC"
author: "Sarah Killcoyne"
date: "29 June 2017"
output: 
  html_document: 
    fig_height: 7
    fig_width: 5
    toc: yes
    toc_depth: 4
---

```{r, include=FALSE, warning=F, message=F, echo=F}

library(pander)
library(gridExtra)
library(plyr)
library(ggplot2)
library(reshape2)
library(GenomicRanges)
library(ggrepel)
library(ggfortify)
library(readxl)

source('lib/load_patient_metadata.R')

## Lists the sequenced patients
sequenced = read_excel('~/Documents/OneDrive/OneDrive - MRC Cancer Unit at University of Cambridge/skillcoyne/OC Data/WGS_Phase 1 complete - cleaned.xlsx', trim_ws=T)
sequenced = subset(sequenced, grepl('^OC', OCCAMS_ID)  )[,c(1:8)]
sequenced$Sample_type = as.factor(tolower(sequenced$Sample_type))
sequenced = subset(sequenced, grepl('barret|normal|tumour', Sample_type))

be.seq = subset(sequenced, grepl('barret', Sample_type))

# n patients sequenced
length(unique(sequenced$OCCAMS_ID))
# ------------------------------ #

# OCCAMS ID to Illumina ID
paired = read.table('~/Documents/OneDrive/MRC Cancer Unit at University of Cambridge/icgc group - icgc_shared/icgc-group/batch_progress/reports/g37/PairedOverviewSummary_g37_bwa.txt', header=T, sep='\t', stringsAsFactors = F)
paired = subset(paired, aligned == '*')
pairedg37 = subset(paired, OCCAMS %in% unique(sequenced$OCCAMS_ID), select=c('OCCAMS','SampleId'))
length(unique(pairedg37$OCCAMS))

paired = read.table('~/Documents/OneDrive/MRC Cancer Unit at University of Cambridge/icgc group - icgc_shared/icgc-group/batch_progress/reports/g1k//PairedOverviewSummary_g1k_bwa.txt', header=T, sep='\t', stringsAsFactors = F)
paired = subset(paired, aligned == '*')
pairedg1k = subset(paired, OCCAMS %in% unique(sequenced$OCCAMS_ID), select=c('OCCAMS','SampleId'))

paired = rbind(pairedg37, pairedg1k)
paired = unique(paired)

length(unique(paired$OCCAMS))

paired[] = lapply(paired[], strip.whitespace)
length(unique(paired$OCCAMS))

sp = do.call(rbind, lapply(paired$SampleId, function(x) unlist(strsplit(x, '_vs_'))))
colnames(sp) = c('tb.sample', 'normal')

paired = cbind(paired, sp)

barretts.pairs = subset(paired, tb.sample %in% be.seq$Illumina_Barcode)

paired = paired[-which(paired$tb.sample %in% be.seq$Illumina_Barcode),]

# ------------------------------ #
# Clinical data
load('~/Data/ICGC/OAC/clinical/labkey_20172706.Rdata', verbose=T)
demo = occams$patients
cols = c(grep('Reason',colnames(demo), value=T),
         grep('Planned',colnames(demo),value=T),
         'FE.EndPoint','RP.Tumor.Size','ST.MainSurgery','RP.HistoryOfNeoAdjuvantTherapy',
         'TR.PalliativeTreatmentModality','TR.CurativeTreatmentModality')
demo[cols] = lapply(demo[cols], as.null)
#demo = oac.demo[unique(sequenced$OCCAMS_ID),]
# ------------------------------ #

# Barrett's adjacency updates from Caitrona
be.updates = data.table::fread("~/Documents/OneDrive/OneDrive - MRC Cancer Unit at University of Cambridge/skillcoyne/OC Data/Barret's Confirmed Data-Caitrona_30092016.txt", stringsAsFactors=F, colClasses=rep('character', 7) )
ba = merge(demo[intersect(rownames(demo), be.updates$`OCCAMS/ID`), c('StudySite','RP.BarrettsAdjacent')], be.updates, by.x='row.names',by.y='OCCAMS/ID') 
rownames(ba) = ba$Row.names

ba = ba[c('RP.BarrettsAdjacent',"Barret's Confirmed")]
colnames(ba) = c('labkey','caitrona')

ba[which(ba$caitrona == ""), 'caitrona'] = NA
ba[which(ba$caitrona == "Y"), 'caitrona'] = "yes"
ba[which(ba$caitrona == "N"), 'caitrona'] = "no"
changed = which(ba[,'labkey'] != ba[,'caitrona'] | is.na(ba[,'labkey']) & !is.na(ba[,'caitrona']))

demo[rownames(ba[changed,]), 'RP.BarrettsAdjacent'] = ba[changed,'caitrona']
# ------------------------------ #

# Copy number data - tumour
read.caveman<-function(s) {
  files = sample.list[[s]]
  if (length(files) < 2) {
    warning(s)
    stop()
  }
  stats = read.table(grep('samplestatistics.csv', files, value=T), row.names=1)

  caveman = read.csv(grep('copynumber.caveman.csv$', files, value=T))
  caveman$ploidy = stats['Ploidy',]
  caveman$rho = stats['rho',]
  caveman$contamination = stats['NormalContamination',]

  caveman$Illumina.Barcode.full = s
  caveman
}

# Tumours
dir = '~/Data/ICGC/BE_OAC_Cohort/tumour/data/ascat'
files = list.files(dir, full.names=T, recursive=T)
samples = list.files(dir, full.names=F)

sample.list = lapply(samples, function(x) {
  grep(x, files, value=T)
} )
names(sample.list) = samples

cnv = do.call(rbind, lapply(samples, read.caveman))
totalCNV = merge(cnv, paired, by.x='Illumina.Barcode.full', by.y='SampleId')
oacCNV = makeGRangesFromDataFrame(totalCNV, keep.extra.columns = T)

# Barretts
dir = '~/Data/ICGC/BE_OAC_Cohort/BA/data/ascat'
files = list.files(dir, full.names=T, recursive=T)
samples = list.files(dir, full.names=F)

sample.list = lapply(samples, function(x) {
  grep(x, files, value=T)
} )
names(sample.list) = samples

cnv = do.call(rbind, lapply(samples, read.caveman))
beCNV = merge(cnv, barretts.pairs, by.x='Illumina.Barcode.full', by.y='SampleId')
beCNV = makeGRangesFromDataFrame(beCNV, keep.extra.columns = T)
# ------------------------------ #

oac.demo = subset(demo, StudySubjectID %in% unique(oacCNV$OCCAMS))
be.demo = subset(demo, StudySubjectID %in% unique(beCNV$OCCAMS))

table(be.demo$RP.BarrettsAdjacent)
table(oac.demo$RP.BarrettsAdjacent)

## Shallow WGS data 
load('~/Data/Ellie/Analysis/5e06/loo.Rdata', verbose=T)
swgsPredictions = pg.samp
rm(pg.samp)

load('~/Data/Ellie/Analysis/5e06/all.pt.alpha.Rdata', verbose=T)
rawSWGSdf = dysplasia.df
a = '0.9'
coefs = coefs[[a]]; model = models[[a]]

hazards = apply( exp(t(rawSWGSdf[, rownames(coefs)])*coefs[,1]), 1, function(x) { sd(x)/mean(x) })
rm(dysplasia.df,fits,plots,performance.at.1se)
# ------------------------------ #

ch = as.data.frame(hazards)
ch$coef = coefs[,1]
ch$stability = rowSums(coefs[,-1])/50
ch$label = rownames(ch)

analysis.files = list.files('~/Data/Ellie/Analysis', full.names=T)
load(grep('Training_patients.Rdata', analysis.files, value=T), verbose=T)

labels = unlist(sapply(patient.data, function(df) {
  df$info = arrange(df$info, Endoscopy.Year, Pathology)
  label = as.integer(df$info$Status == 'P') #as.integer(df$info$Pathology %in% c('HGD', 'IMC'))
  names(label) = df$info$Samplename
  return(label)
}))
names(labels) = sub('.*\\.', '',  names(labels))

labels = labels[which(names(labels) %in% rownames(rawSWGSdf))]

rm(patient.data)
```

# Questions in the Shallow WGS data

Load the coefficients from the selected model and merge with the copynumber regions identified in (some) of the OAC patients
```{r, echo=F}
regions = do.call(rbind.data.frame, c(strsplit(grep('\\d+', ch$label, value=T), ':|-'), stringsAsFactors=F))
colnames(regions) = c('chr','start','stop')
regions[colnames(regions)] = lapply(regions[colnames(regions)], function(x) as.numeric(x))

# Turn SWGs features into grange object
swgRegions = makeGRangesFromDataFrame(cbind.data.frame(regions, ch[grep('\\d+', ch$label),]), keep.extra.columns = T)
swgRegions = swgRegions[order(-swgRegions$hazards)]

ggplot(ch, aes(coef, hazards, col=hazards > quantile(hazards)[['75%']])) + 
  geom_point(alpha=0.5) + 
  geom_text_repel(aes(label=ifelse(hazards > quantile(hazards)[['75%']], label, '')),show.legend=F) +
  labs(title='Top features coef vs hazard') + theme(legend.position = 'none')
```

## Are gains/losses enriched in the coefficients?
We would expect them to be since these are the regions that the model selects as being the best combination to predict P vs NP.

Certainly they are found at a different rate in gains for P vs NP. Which we would expect.

```{r}
# positive coefficients were related to gains in progressors
gains = subset(ch, coef > 0)

gain = do.call(rbind, lapply(gains$label, function(reg) {
  values = rawSWGSdf[,reg] 
  table(labels[which(values > 1)])
}))

t.test(gain[,1], gain[,2])

# negative coefficients were related to losses in non-progressors
losses = subset(ch, coef < 0)

loss = do.call(rbind, lapply(losses$label, function(reg) {
  values = rawSWGSdf[,reg] 
  table(labels[which(values < -1)])
}))

t.test(loss[,1], loss[,2])

```

This is true of randomly selected regions as well, which we also expect based on the general trend towards disorder we see in P vs NP.

```{r}
sample.reg = sample(setdiff(colnames(rawSWGSdf), ch$label), nrow(ch))

randomGains = do.call(rbind, lapply(sample.reg, function(reg) {
  values = rawSWGSdf[,reg] 
  table(labels[which(values > 1)])
}))

t.test(randomGains[,1], randomGains[,2])

randomLosses = do.call(rbind, lapply(sample.reg, function(reg) {
  values = rawSWGSdf[,reg] 
  table(labels[which(values < -1)])
}))

t.test(randomLosses[,1], randomLosses[,2])

```

However, while there appears to be no difference in the gains found in NP for randomly selected regions vs the coefficients:

pander(wilcox.test(gain[,'0'], randomGains[,'0']), caption="Non progressors, coeffiecents vs randomly selected")

pander(wilcox.test(loss[,'0'], randomLosses[,'0']), caption="Non progressors, coeffiecents vs randomly selected")

There is a significant difference in the selected regions for progressors. Again this is mostly a sanity check:

pander(wilcox.test(gain[,'1'], randomGains[,'1']), caption="Progressors, coeffiecents vs randomly selected gains")

pander(wilcox.test(loss[,'1'], randomLosses[,'1']), caption="Progressors, coeffiecents vs randomly selected losses")

## Two groups of patients?

There appears to be two groups of patients.

Group 1 is patients that show progression consistently ("born bad")
Group 2 is patients that show sudden progression in the final or nearly final sample

So is there a difference between the two in regards to the alterations in the features drawn from the model?
Or in randomly drawn regions?

```{r}
cutoff = 0.5
status = sapply(swgsPredictions, function(x) unique(x$Status))
transform.pt.by.time<-function(pt) {
  pt = pt %>% group_by(PID, Endoscopy.Year) %>% summarise(
    'pred'=sum(as.integer(Prediction > cutoff)), 'samples'=length(PID), 
    #'final.endo'= length(which(grepl('HGD|IMC', Pathology))) > 0,
    'months.before.final'= (max(pt$Endoscopy.Year) - unique(Endoscopy.Year))*12, 
    'max.path'=max(Pathology))
  arrange(pt, -Endoscopy.Year, PID)
}

back = lapply(swgsPredictions[which(status == 'P')], transform.pt.by.time)

# By sample
fvpS = do.call(rbind.data.frame, lapply(back, function(df) {
  last.sample = (df[1,'pred']/df[1,'samples'])[[1]]
  if (nrow(df) <= 1 | max(df$months.before.final) < 12 ) return(NA)
  cbind('final'=last.sample,  'previous'=sum(df[-1, 'pred']/df[-1, 'samples'])/(nrow(df)-1) )
}))

# Do it by endoscopy? Not sample? And max.path should matter
final.vs.prev = do.call(rbind.data.frame, lapply(back, function(df) {
  last.endo = as.integer(df[1,'pred'] > 0)
  if (nrow(df) <= 1 | max(df$months.before.final) < 12 ) return(NA)

  previous = subset(df[-1,], max.path %in% c('BE','ID'))
  cbind('final'=last.endo,  'previous'=sum(as.integer(previous$pred > 0)/(nrow(previous))))
}))

final.vs.prev = final.vs.prev[complete.cases(final.vs.prev),]

m = melt(as.matrix(final.vs.prev))
m = transform(m, Var1=reorder(Var1, -value))

```

First we want to separate out the patient groups we have. In this table we only include patients that have more than a single sample, and with follow up samples 1 or more years prior to their final (HGD/IMC) sample.

```{r, echo=F}
ggplot(m, aes(x=Var1, y=value,fill=Var2)) + 
  geom_bar(stat="identity", width=.5, position = "dodge") + coord_flip() + 
  labs(title='Ratio of final sample predictions vs all previous samples')
```

For ~~samples~~ endoscopies predicted prior to the final our mean ratio is `r mean(final.vs.prev$previous)` so most are predicted. 
```{r, echo=F}
hist(final.vs.prev$previous)
```

When the final ~~sample~~ endoscopy was not predicted correctly the prediction for previous samples also declines to a mean of `r mean(summary(subset(final.vs.prev, final <= 0)$previous))`.

So the two groups are:

Sudden - predicted at endpoint, but not consistently prior 
```{r}
sudden = subset(final.vs.prev, final > 0 & previous < mean(previous))

pander(sudden, caption='Patients whose final endoscopies were predicted, but prior endoscopies were less clear.')
```

Born bad - predicted consistently prior to endpoint

```{r}
born.bad = subset(final.vs.prev, final > 0 & previous > mean(previous))

pander(born.bad, caption='Patients whose endoscopies were predicted throughout')
```

So how can we compare these?
```{r}
count.losses <- function(x) { length(which(x < -1)) }
count.gains <- function(x) { length(which(x > 1)) }

annotate.by.pt<-function(df) {
  df = df[,c('Prediction','Pathology', 'PID', 'Endoscopy.Year', 'Samplename')]
  df$Prediction = as.integer(df$Prediction > cutoff)

  df$Final = df$Pathology %in% c('HGD','IMC') & df$Endoscopy.Year == df$Endoscopy.Year
  df[,c('Prediction','Pathology','Samplename','Final')]
}

sd.pts = do.call(rbind, lapply(swgsPredictions[rownames(sudden)],annotate.by.pt))
bb.pts = do.call(rbind, lapply(swgsPredictions[rownames(born.bad)],annotate.by.pt))
rd.pts = do.call(rbind, lapply(swgsPredictions[sample(which(status == 'P'), 15)],annotate.by.pt))

sd = cbind.data.frame('gain'=apply(rawSWGSdf[subset(sd.pts, !Final)$Samplename,gains$label], 2, count.gains), 
          'loss'=apply(rawSWGSdf[subset(sd.pts, !Final)$Samplename,gains$label], 2, count.losses))

bb = cbind.data.frame('gain'=apply(rawSWGSdf[subset(bb.pts, !Final)$Samplename,gains$label], 2, count.gains), 
          'loss'=apply(rawSWGSdf[subset(bb.pts, !Final)$Samplename,gains$label], 2, count.losses))

rd = cbind.data.frame('gain'=apply(rawSWGSdf[subset(rd.pts, !Final)$Samplename,gains$label], 2, count.gains), 
          'loss'=apply(rawSWGSdf[subset(rd.pts, !Final)$Samplename,gains$label], 2, count.losses))

```
I think this is driven by the number of samples more a difference in the two groups. The "born bad" group has a lot more samples than the "sudden" and if I randomly sample the same number of samples from the "born bad" group the difference disappears. So, if there's a difference here how to I look for it?

```{r}
# Adjusted for the number of samples...is this right?
wilcox.test(sd$gain/nrow(subset(sd.pts, !Final)), bb$gain/nrow(subset(bb.pts, !Final)))
wilcox.test(sd$gain/nrow(subset(sd.pts, !Final)), rd$gain/nrow(subset(rd.pts, !Final)))
wilcox.test(rd$gain/nrow(subset(rd.pts, !Final)), bb$gain/nrow(subset(bb.pts, !Final)))

wilcox.test(sd$loss/nrow(subset(sd.pts, !Final)), bb$loss/nrow(subset(bb.pts, !Final)))
wilcox.test(sd$loss/nrow(subset(sd.pts, !Final)), rd$loss/nrow(subset(rd.pts, !Final)))
wilcox.test(rd$loss/nrow(subset(rd.pts, !Final)), bb$loss/nrow(subset(bb.pts, !Final)))

```

```{r, echo=F, eval=F}
gains = gains[grep('\\d+',gains$label),]
losses = losses[grep('\\d+',losses$label),]

sd.samples = unlist(lapply(swgsPredictions[rownames(sudden)], function(df) df$Samplename))
bb.samples = unlist(lapply(swgsPredictions[rownames(born.bad)], function(df) df$Samplename))
random.samples = unlist(lapply(swgsPredictions[sample(which(status == 'P'), 15)], function(df) df$Samplename ))

# Born bad losses/gains in features
bbL = apply(rawSWGSdf[bb.samples, losses$label], 2, count.losses)
bbG = apply(rawSWGSdf[bb.samples, gains$label], 2, count.gains)

# sudden losses/gains in features
snL = apply(rawSWGSdf[sd.samples, losses$label], 2, count.losses)
snG = apply(rawSWGSdf[sd.samples, gains$label], 2, count.gains)

# random losses/gains in features
rrL = apply(rawSWGSdf[random.samples, losses$label], 2, count.losses)
rrG = apply(rawSWGSdf[random.samples, gains$label], 2, count.gains)

# one-way anova by ranks -- basically the same as wilcox I think
one.way.aov<-function(a, b) {
  df = rbind(cbind.data.frame('counts'=a, 'class'='a'),
             cbind.data.frame('counts'=b, 'class'='b')) 
  kruskal.test(counts~class, data=df)
  #chisq.test(df$counts, df$class)
  #glm(counts~class, data=df, family='poisson')
}

one.way.aov(bbG, snG)
one.way.aov(bbG, rrG)

one.way.aov(bbL, snL)
one.way.aov(bbL, rrL)

one.way.aov(snL, rrL)
one.way.aov(snG, rrG)
```

```{r, warning=F, message=F, fig.width=7}

tests = rbind(cbind.data.frame('CN'='gain', 'p'=wilcox.test(bbG, rrG)$p.value, 'type'='Random', 'label'='BB~R','y'=max(rrG)),
              cbind.data.frame('CN'='gain', 'p'=wilcox.test(bbG, snG)$p.value, 'type'='Sudden', 'label'='BB~SN', 'y'=max(snG)),
              cbind.data.frame('CN'='loss', 'p'=wilcox.test(bbL, rrL)$p.value, 'type'='Random', 'label'='BB~R','y'=max(rrL)),
              cbind.data.frame('CN'='loss', 'p'=wilcox.test(bbL, snL)$p.value, 'type'='Sudden', 'label'='BB~SN', 'y'=max(snL)) )

tmp = rbind(cbind(snG, 'type'='Sudden','CN'='gain'),
            cbind(bbG, 'type'='Born Bad', 'CN'='gain'),
            cbind(rrG, 'type'='Random', 'CN'='gain'),
            cbind(snL, 'type'='Sudden', 'CN'='loss'),
            cbind(bbL, 'type'='Born Bad', 'CN'='loss'),
            cbind(rrL, 'type'='Random', 'CN'='loss' ) )
tmp = as.data.frame(tmp, row.names = NA)
tmp$sn = as.integer( as.character(tmp$sn) )
tmp$type = factor(tmp$type, levels = c('Born Bad','Sudden','Random'))

grid.arrange(
  ggplot(tmp, aes(type, sn, group=type, fill=type)) + facet_grid(~CN) + 
    geom_boxplot(outlier.colour = NA, alpha=0.3) + geom_jitter(width=0.2, aes(color=type)) + 
    #geom_label(data=tests, aes(x=type,y=y,label=paste(label,format.pval(p,2),sep='\np=')), nudge_y=5, alpha=0.5)
    geom_label(data=tests, aes(x=type,y=y,label=paste(label,format.pval(p,2),sep='\np=')), nudge_y=5, alpha=0.5) +
    labs(y='CN count per region',x='', title='Gains & Losses in selected features') + theme(legend.position='none'),

    tableGrob(  rbind(cbind('Group'='Born Bad','# Samples'=length(bb.samples)),
                      cbind('Group'='Sudden','# Samples'=length(sd.samples)), 
                      cbind('Group'='Random','# Samples'=length(random.samples))) ), nrow=2, heights=c(4,1) )

```

### The two groups don't actually exist

TODO  I don't recall what these tests were looking at, need to redo this with the new data

```{r}

gains = gains[grep('\\d+',gains$label),]
losses = losses[grep('\\d+',losses$label),]

km = kmeans(rawSWGSdf[c(bb.samples,sd.samples),ch$label], 2)
table(km$cluster)

clusPred = sapply(swgsPredictions[which(status == 'P')], function(df) {
  sum(as.integer(subset(df, Samplename %in% names(which(km$cluster == 1)))$Prediction > cutoff))
})

sum(clusPred)

binom.test(17,17,.72)
names(which(clusPred > 0))
which(rownames(sudden) %in% names(which(clusPred > 0)))
which(rownames(born.bad) %in% names(which(clusPred > 0)))

```

# Questions in OAC data

## Gains

These may be more nuanced, as we need to be aware of gains that reflect WGD vs gains that are new amplifications. At the moment there's no distinction

```{r, fig.height=16, fig.width=16}
plot.gains<-function(df) {
  df = lapply(df, function(x) {
    x = arrange(x, -width)
    x$order = factor(1:nrow(x))
    return(x)
  })

  m = do.call(rbind.data.frame, df)
  m = m %>% group_by(seqnames) %>% ungroup() %>% arrange(segment, seqnames, width, start) %>% mutate(segorder = row_number())
m$isWGD = as.factor(m$isWGD)

m$CN = cut(m$Total_CN, include.lowest=T, ordered_result=T, breaks=c(3, 5, 7, 9, 10, max(m$Total_CN)))
ggplot(data=m, aes(x=order, y=start)) +  
  facet_wrap(~segment, scales='free') +
  geom_segment(aes(x=order, xend=order, y=start, yend=end, color=CN)) + 
  geom_segment(aes(x=0,xend=0, y=seg.start, yend=seg.end), color='red', size=2) +
  scale_colour_brewer(palette='YlGnBu', direction=-1) +
  coord_flip() + 
  theme(legend.position='bottom', axis.text=element_blank(),panel.background=element_blank()) + labs(y='chr position', x='Patient', title='Gains')
}

# WGD
oacCNV$isWGD = oacCNV$ploidy > 3.2 
beCNV$isWGD = beCNV$ploidy > 3.2

#average genome ploidy <= 2.7 AND total copy number >= 5
#OR average genome ploidy > 2.7 AND total copy number >= 9

gain.for<-function(gg, cnv) {
  ov = findOverlaps(cnv, gg)  
  tmp = cnv[queryHits(ov)]

  if (length(tmp) > 0) {
    #tmp = subset(tmp, ploidy > 2 & Total_CN > round(ploidy))
    tmp = subset(tmp, 
       (!isWGD & Total_CN > round(ploidy)) | 
         (isWGD & Total_CN > round(ploidy+3.2)) )
    if (length(tmp) > 0) {
      tmp$segment = as.character(gg)
      tmp$seg.start = start(gg)
      tmp$seg.end = end(gg)
    }
  }
  as.data.frame(tmp)
}

## gains in patients -- issues with WGD samples?
gains = lapply(swgRegions[which(swgRegions$coef > 0)], gain.for, cnv=oacCNV)
names(gains) = as.character(swgRegions[which(swgRegions$coef > 0)])
plot.gains(gains) + labs(title='Gains in OAC samples')

be.gains = lapply(swgRegions[which(swgRegions$coef > 0)], gain.for, cnv=beCNV)
names(be.gains) = as.character(swgRegions[which(swgRegions$coef > 0)])
plot.gains(be.gains[sapply(be.gains, nrow) > 0]) + labs(title='Gains in BE Adjacent samples')

```

```{r, fig.height=8, fig.width=8}

# Per patient
gfreq = sort(sapply(gains, function(x) length(unique(x$OCCAMS)))/length(unique(sequenced$OCCAMS_ID)))
ggplot(melt(as.matrix(gfreq)), aes(Var1, value)) + geom_col(fill='darkblue') + ylim(0,0.5) + coord_flip() + 
  labs(x='', y='Frequency', title='Gain frequency in OAC') + theme_bw() + theme(text=element_text(size=14))

gfreq = sort(sapply(be.gains, function(x) length(unique(x$OCCAMS)))/length(unique(be.seq$OCCAMS_ID)))
ggplot(melt(as.matrix(gfreq)), aes(Var1, value)) + geom_col(fill='darkblue')  + coord_flip() + 
  labs(x='', y='Frequency', title='Gain frequency in BE adjacent') + theme_bw() + theme(text=element_text(size=14))

```

## Losses

```{r, fig.height=12, fig.width=12}
plot.losses<-function(df) {
  df = lapply(df, function(x) {
    if(nrow(x) <= 0) return(NULL)
    x = arrange(x, -width)
    x$order = factor(1:nrow(x))
    return(x)
  })

  m = do.call(rbind.data.frame, df)
  #m$segnames = ordered(m$seqnames, levels=c(1:22))

  m = m %>% group_by(seqnames) %>% ungroup() %>% arrange(segment, seqnames, width, start) %>% mutate(segorder = row_number())
  m$isWGD = as.factor(m$isWGD)
  m$CN = as.factor(m$Total_CN)
  #m = transform(m, segment=reorder(segment, segnames))
  ggplot(data=m, aes(x=order, y=start)) +  facet_wrap(~segment, scales='free') +
    geom_segment(aes(x=order, xend=order, y=start, yend=end, color=CN)) + 
    geom_segment(aes(x=0,xend=0, y=seg.start, yend=seg.end), color='red', size=2) +
    scale_colour_brewer(palette='YlGnBu', direction=1) +
    coord_flip() + 
    theme(legend.position='bottom', axis.text=element_blank(),panel.background=element_blank()) + labs(y='chr position', x='Patient', title='Losses in OAC Patients')
}

# Losses in patients -- complicated by WGD
##average genome ploidy <= 2.7 AND total copy number = 0
#OR average genome ploidy > 2.7 AND total copy number < ( average genome ploidy - 2.7 )
loss.for<-function(gg, cnv){
  ov = findOverlaps(cnv, gg)  
  tmp = cnv[queryHits(ov)]
  if (length(tmp) > 0) {
    tmp = subset(tmp, (!isWGD & Total_CN == 0) |
                    (isWGD & Total_CN < round(ploidy)) )
    #tmp = tmp[tmp$Total_CN < 2]
    if (length(tmp) > 0) {
      tmp$segment = as.character(gg)
      tmp$seg.start = start(gg)
      tmp$seg.end = end(gg)
    }
  }
  as.data.frame(tmp)
}

losses = lapply(swgRegions[which(swgRegions$coef < 0)], loss.for, oacCNV)
names(losses) = as.character(swgRegions[which(swgRegions$coef < 0)])
plot.losses(losses)

be.losses = lapply(swgRegions[which(swgRegions$coef < 0)], loss.for, beCNV)
names(be.losses) = as.character(swgRegions[which(swgRegions$coef < 0)])
plot.losses(be.losses)

```

```{r, fig.height=8, fig.width=8}

lfreq = sort(sapply(losses, function(x) length(unique(x$OCCAMS)))/length(unique(sequenced$OCCAMS_ID)))
ggplot(melt(as.matrix(lfreq)), aes(Var1, value)) + geom_col(fill='darkgreen') + ylim(0,0.5) + coord_flip() + 
  labs(x='', y='Frequency', title='Loss frequency in OAC') + theme_bw() + theme(text=element_text(size=14))

lfreq = sort(sapply(be.losses, function(x) length(unique(x$OCCAMS)))/length(unique(be.seq$OCCAMS_ID)))
ggplot(melt(as.matrix(lfreq)), aes(Var1, value)) + geom_col(fill='darkgreen') + ylim(0,0.5) + coord_flip() + 
  labs(x='', y='Frequency', title='Loss frequency in BE adjacent') + theme_bw() + theme(text=element_text(size=14))

## Note it's a subset of the 321 patients (66) that have this
chr17 = get.chr.lengths()[17,]

#chr17losses = lapply(makeGRangesFromDataFrame( cbind.data.frame('start'=0, 'end'=chr17$chr.length, 'chr'='17') ), loss.for)[[1]]
#m = arrange(chr17losses, width, start)

m = arrange(losses$`17:1-4776188`, width, start)
m$Anon = 1:length(m$OCCAMS)
m = transform(m, Anon = reorder(Anon, -width))
segment = cbind('seg.start'=1, 'seg.end'=4776188)
m$CN = as.factor(m$Total_CN)
#ggplot(m, aes(x=Anon, y=start)) +

ggplot() + 
  geom_segment(data=m, aes(x=Anon, xend=Anon, y=start, yend=end, color=CN), size=2) + 
  geom_segment(data=segment, aes(x=0, xend=0, y=seg.start, yend=seg.end),color='red', size=3) + 
  geom_hline(data=chr17, aes(yintercept=(chr.cent-cent.gap)), color='red', linetype='dashed') +
  geom_label(data=chr17, aes(x=as.factor(3), y=chr.cent/2), label='p arm') +
  scale_colour_brewer(palette='YlGnBu', direction=1) +
  coord_flip() + 
  theme(legend.position = 'bottom') + labs(y='chr17 position', x='Patient', title='Chr17 Losses in OAC Patients', subtitle=paste(round(length(unique(losses$`17:1-4776188`$OCCAMS))/length(unique(oacCNV$OCCAMS)), 2)*100, '% of OAC patients', sep=''), text=element_text(size=14, face='bold'))

```

So is there a difference between the gains/losses in the selected regions?  Not really.
```{r}
wilcox.test(sapply(gains, nrow), sapply(losses, nrow))
one.way.aov(sapply(gains, nrow), sapply(losses, nrow))

# Maybe between BA and tumour...
wilcox.test(sapply(gains, nrow), sapply(be.gains, nrow))
one.way.aov(sapply(gains, nrow), sapply(be.gains, nrow))
```

## Random

Compare those to gains/losses in randomly selected regions of the same size across the genome

```{r, echo=F, warning=F, message=F}
chr.lengths = get.chr.lengths()
chr.lengths$chrom = sub('chr','',chr.lengths$chrom)
chr.lengths$start = 1

genome = makeGRangesFromDataFrame(chr.lengths[1:22,], seqnames.field = 'chrom', end.field='chr.length')
tiles = slidingWindows(genome, width=5e6, step=2.5e6)
chr.samples = sample(length(tiles), length(tiles)/2) # half of the chromosomes
sample.regions = do.call(c, lapply(tiles[chr.samples], function(x) sample(x, length(x)/4)  )) # grab some portion of each
```

### Gains

These may be more nuanced, as we need to be aware of gains that reflect WGD vs gains that are new amplifications. At the moment there's no distinction

```{r}
randomGains = lapply(sample.regions, gain.for, cnv=oacCNV)
names(randomGains) = as.character(sample.regions)

plot.gains(randomGains[sample(length(randomGains), 20)])

# Maybe a difference in # gains in the selected features vs random
wilcox.test(sapply(randomGains, nrow), sapply(gains, nrow))
one.way.aov(sapply(randomGains, nrow), sapply(gains, nrow))

# Not a difference here though
wilcox.test(sapply(gains, function(df) length(unique(df$OCCAMS))), 
            sapply(randomGains, function(df) length(unique(df$OCCAMS))))

# Not sure this shows a really clear difference, but...
df = rbind(cbind.data.frame('counts'=sapply(gains, function(df) length(unique(df$OCCAMS))), 'type'='selected') , 
      cbind.data.frame('counts'=sapply(randomGains, function(df) length(unique(df$OCCAMS))), 'type'='random'))
ggplot(df, aes(type, counts, group=type)) + geom_boxplot() + geom_jitter(width=0.3) + labs(title='Gain counts by region')

df = rbind(cbind.data.frame('counts'=sapply(gains, nrow), 'type'='selected') , 
      cbind.data.frame('counts'=sapply(randomGains, nrow), 'type'='random'))
ggplot(df, aes(type, counts, group=type)) + geom_boxplot() + geom_jitter(width=0.3) + labs(title='Gain counts by patient')

```

### Losses

```{r}
randomLosses = lapply(sample.regions, loss.for, cnv=oacCNV)
names(randomLosses) = as.character(sample.regions)

plot.losses(randomLosses[sample(length(randomLosses), 15)])

# Definite difference between the numbers of gains and losses in randomly selected regions
wilcox.test(sapply(randomGains, nrow), sapply(randomLosses, nrow))
one.way.aov(sapply(randomGains, nrow), sapply(randomLosses, nrow))

# No difference in # losses in the selected features
wilcox.test(sapply(randomLosses, nrow), sapply(losses, nrow))
one.way.aov(sapply(randomLosses, nrow), sapply(losses, nrow))

df = rbind(cbind.data.frame('counts'=sapply(randomLosses, nrow), 'type'='random') ,
           cbind.data.frame('counts'=sapply(losses, nrow), 'type'='selected'))

#df = df[-which(df$counts > 400),]
# No difference between the two
ggplot(df, aes(type, counts, group=type)) + geom_boxplot() + geom_jitter(width=0.3) + labs(title='Loss counts')

summary(subset(df, type=='random')$counts)
summary(subset(df, type=='selected')$counts)

# No real difference in # patients with losses in the selected features
wilcox.test(sapply(losses, function(df) length(unique(df$OCCAMS))), 
            sapply(randomLosses, function(df) length(unique(df$OCCAMS))))
one.way.aov(sapply(losses, function(df) length(unique(df$OCCAMS))), 
            sapply(randomLosses, function(df) length(unique(df$OCCAMS))))

# No difference between the two
df = rbind(cbind.data.frame('counts'=sapply(losses, function(df) length(unique(df$OCCAMS))), 'type'='selected') , 
      cbind.data.frame('counts'=sapply(randomLosses, function(df) length(unique(df$OCCAMS))), 'type'='random'))
ggplot(df, aes(type, counts, group=type)) + geom_boxplot() + geom_jitter(width=0.3) + labs(title='Loss counts by patient')

#pnorm(subset(df, type=='selected')$counts, mean=mean(subset(df, type=='selected')$counts), sd=sd(subset(df, type=='selected')$counts))
#pnorm(subset(df, type=='random')$counts, mean=mean(subset(df, type=='random')$counts), sd=sd(subset(df, type=='random')$counts))

```

## Focal vs Large CN

It's worth looking to see if focal/large changes are enriched in the gains particularly as the plots seemed to suggest they may be.

#### Gains

```{r, fig.width=8}
size = unlist(sapply(gains, function(df) df$width))
hist(size)
abline(v = 3e6, col='red')

#chr.lengths$chr.length*.01

# Focal changes 1% or the chromosome?
countBySize <-function(df) {
  s = round(subset(chr.lengths, chrom == unique(df$seqnames))$chr.length *.01)
  cbind('focal'=nrow(subset(df, width < s)),
        'large'=nrow(subset(df, width >= s)))
}

sizebyPatient<-function(df) {
  s = round(subset(chr.lengths, chrom == unique(df$seqnames))$chr.length *.01)
  cbind('focal'=length(unique(subset(df, width < s)$OCCAMS)),
        'large'=length(unique(subset(df, width >= s)$OCCAMS)))
}

# sizebyPatient<-function(x) {
#   cbind('focal'=length(unique(subset(x, width <= quantile(width)[['25%']])$OCCAMS)),
#         'large'=length(unique(subset(x, width > quantile(width)[['25%']])$OCCAMS)) )
# }

# countBySize<-function(x) {
#   cbind('focal'=nrow(subset(x, width < quantile(width)[['25%']])),
#         'large'=nrow(subset(x, width >= quantile(width)[['25%']])) )
# }

sizeGainsByPatient = do.call(rbind.data.frame,lapply(gains, sizebyPatient))
sizeGains = do.call(rbind.data.frame,lapply(gains, countBySize))

#  enrichment for some focal changes
p.adjust(apply(sizeGains, 1, function(x) binom.test(x[1], sum(x), 0.5)$p.value ))
apply(sizeGains, 1, function(x) binom.test(x[1], sum(x), 0.5)$conf.int )

# Same is true for gains of type by patient
p.adjust(apply(sizeGainsByPatient, 1, function(x) binom.test(x[1], sum(x), 0.5)$p.value ))
apply(sizeGainsByPatient, 1, function(x) binom.test(x[1], sum(x), 0.5)$conf.int )

# Clear difference between focal/large
wilcox.test(sizeGainsByPatient$focal, sizeGainsByPatient$large)
one.way.aov(sizeGainsByPatient$focal, sizeGainsByPatient$large)

sizeGainsRatio = sizeGains %>% rowwise() %>% summarise( 'focal.r'=focal/sum(focal,large), 'large.r'=large/sum(focal,large) )
sizeGainsRatio$feature = rownames(sizeGains)

# No difference in the overall number of gains per region
ggplot(melt(sizeGains), aes(variable, value, group=variable, fill=variable)) + 
  geom_boxplot(alpha=0.5, outlier.colour = NA) + geom_jitter(width=0.2, aes(color=variable)) 
  labs(title="Types of gain per region", subtitle=paste('p=',format.pval(one.way.aov(sizeGains$focal, sizeGains$large)$p.value,3), sep=''), y='patients per region', x='')

grid.arrange(
  ggplot(melt(sizeGainsRatio), aes(feature, value, fill=variable) ) + geom_col() + coord_flip() + labs(title='Ratio of overall gains per region'),
  ggplot(melt(sizeGainsByPatient), aes(variable, value, group=variable, fill=variable)) + 
    geom_boxplot(alpha=0.5, outlier.colour = NA) + geom_jitter(width=0.2, aes(color=variable)) + 
    labs(title="Patients per region", subtitle=paste('p=',format.pval(one.way.aov(sizeGainsByPatient$focal, sizeGainsByPatient$large)$p.value,digits=3), sep=''), y='patients per region', x=''),
top='Gains in selected regions', ncol=2)
```

#### Losses

```{r, fig.width=8}
sizeLossesByPatients = do.call(rbind.data.frame,lapply(losses, sizebyPatient))
sizeLosses = do.call(rbind.data.frame,lapply(losses, countBySize))

# True for losses as well
p.adjust(apply(sizeLosses, 1, function(x) binom.test(x[1], sum(x), 0.5)$p.value ))
apply(sizeLosses, 1, function(x) binom.test(x[1], sum(x), 0.5)$conf.int )

wilcox.test(sizeLosses$focal, sizeLosses$large)
one.way.aov(sizeLosses$focal, sizeLosses$large)

sizeLossRatio = sizeLosses %>% rowwise() %>% summarise( 'focal.r'=focal/sum(focal,large), 'large.r'=large/sum(focal,large) )
sizeLossRatio$feature = rownames(sizeLosses)

ggplot(melt(sizeLossRatio), aes(feature, value, fill=variable) ) + geom_col() + coord_flip() + labs(title='Overall losses')

grid.arrange(
ggplot(melt(sizeLosses), aes(variable, value, group=variable, fill=variable)) + 
  geom_boxplot(alpha=0.5, outlier.colour = NA) + geom_jitter(width=0.2, aes(color=variable)) + 
  labs(title="Counts per region", subtitle=paste('p=',format.pval(one.way.aov(sizeLosses$focal, sizeLosses$large)$p.value,3), sep=''), y='counts per region', x=''),

ggplot(melt(sizeLossesByPatients), aes(variable, value, group=variable, fill=variable)) + 
  geom_boxplot(alpha=0.5, outlier.colour = NA) + geom_jitter(width=0.2, aes(color=variable)) + 
  labs(title="Patients per region", subtitle=paste('p=',format.pval(one.way.aov(sizeLossesByPatients$focal, sizeLossesByPatients$large)$p.value,3), sep=''), y='patients per region', x=''),
top='Losses in selected regions', ncol=2)

wilcox.test(sizeLossesByPatients$focal, sizeLossesByPatients$large)
```

### Compared to random focal/large changes?

#### Losses
```{r, fig.width=8}
randomLossesByPt = do.call(rbind.data.frame,lapply(randomLosses, sizebyPatient))
randomLossesBySize = do.call(rbind.data.frame,lapply(randomLosses, countBySize))

grid.arrange(
ggplot(melt(randomLossesBySize), aes(variable, value, group=variable, fill=variable)) + 
  geom_boxplot(alpha=0.5, outlier.colour = NA) + geom_jitter(width=0.2, aes(color=variable)) + 
  labs(title="Counts per region", subtitle=paste('p=',format.pval(one.way.aov(randomLossesBySize$focal, randomLossesBySize$large)$p.value,digits=3), sep=''), y='counts per region', x='') + theme(legend.position = 'none'),

ggplot(melt(randomLossesByPt), aes(variable, value, group=variable, fill=variable)) + 
  geom_boxplot(alpha=0.5, outlier.colour = NA) + geom_jitter(width=0.2, aes(color=variable)) + 
  labs(title="Patients per region", subtitle=paste('p=',format.pval(one.way.aov(randomLossesByPt$focal, randomLossesByPt$large)$p.value,digits=3), sep=''), y='patients per region', x='') + theme(legend.position = 'none'),
top='Losses in randomly selected regions', ncol=2)

# No real difference in randomly selected regions for focal losses
wilcox.test(randomLossesBySize$focal, sizeLosses$focal) 
one.way.aov(randomLossesBySize$focal, sizeLosses$focal) 
wilcox.test(randomLossesByPt$focal, sizeLossesByPatients$focal)

## Difference in large losses
wilcox.test(randomLossesBySize$large, sizeLosses$large)
one.way.aov(randomLossesBySize$large, sizeLosses$large)
wilcox.test(randomLossesByPt$large, sizeLossesByPatients$large) 
```

#### Gains

```{r, fig.width=8}
randomGainsByPt = do.call(rbind.data.frame,lapply(randomGains, sizebyPatient))
randomGainsBySize = do.call(rbind.data.frame,lapply(randomGains, countBySize))

grid.arrange(
ggplot(melt(randomGainsBySize), aes(variable, value, group=variable, fill=variable)) + 
  geom_boxplot(alpha=0.5, outlier.colour = NA) + geom_jitter(width=0.2, aes(color=variable)) + 
  labs(title="Counts per region", subtitle=paste('p=',format.pval(one.way.aov(randomGainsBySize$focal, randomGainsBySize$large)$p.value,3), sep=''), y='counts per region', x='') + theme(legend.position = 'none'),

ggplot(melt(randomGainsByPt), aes(variable, value, group=variable, fill=variable)) + 
  geom_boxplot(alpha=0.5, outlier.colour = NA) + geom_jitter(width=0.2, aes(color=variable)) + 
  labs(title="Patients per region", subtitle=paste('p=',format.pval(one.way.aov(randomGainsByPt$focal, randomGainsByPt$large)$p.value,3), sep=''), y='patients per region', x='') + theme(legend.position = 'none'),
top='Gains in randomly selected regions', ncol=2)

# But maybe a difference in random regions vs selected ones?
wilcox.test(randomGainsBySize$focal, sizeGains$focal) 
wilcox.test(randomGainsBySize$large, sizeGains$large)

wilcox.test(randomGainsByPt$focal, sizeGainsByPatient$focal) # maybe
wilcox.test(randomGainsByPt$large, sizeGainsByPatient$large) # maybe
```

*** 
# Summary to this point

Between the regions selected by the SWGs work and randomly selected regions of the same size across the genome only losses appear to show a difference in the rate of CN suggesting some enrichment in the selected regions.  It's possible that gains need to be broken down differently, for instance separating out gains after WGD vs gains pre-WGD.  For this I need to apply a clonal analysis timing (Mortiz's work).

There's a significant difference in the rate of large-scale alterations vs focal for both gains and losses, but again no difference in gains between selected and random regions, only in losses.  This should be reanalysed based on WGD timing as well.

***

# Barrett's Adjacency & CNA

Is there a relationship between BA patients and CNA in selected regions?

## Selected Regions

```{r, eval=F}
# A lot more gains than losses as expected
sapply(gains, nrow)
sapply(losses, nrow)

population = table(oac.demo$RP.BarrettsAdjacent)

# Check for enrichment of gains/losses in BA
gainsBAPts = as.data.frame(do.call(rbind, lapply(gains, function(g) {
 table(subset(oac.demo, StudySubjectID %in% g$OCCAMS)$RP.BarrettsAdjacent)
})))

# Separated by BA status
gainsBA = lapply(gains, function(g) {
  yes = subset(oac.demo, StudySubjectID %in% g$OCCAMS & RP.BarrettsAdjacent == 'yes')$StudySubjectID
  subset(g, OCCAMS %in% yes)
})

gainsNoBA = lapply(gains, function(g) {
  no = subset(oac.demo, StudySubjectID %in% g$OCCAMS & RP.BarrettsAdjacent == 'no')$StudySubjectID
  subset(g, OCCAMS %in% no)
})

# No difference
one.way.aov(sapply(gainsBA, nrow) , sapply(gainsNoBA, nrow))
# TODO Is this due to numberof patients? Might be
one.way.aov(sapply(gainsBA, nrow)/gainsBAPts[,'yes'] , sapply(gainsNoBA, nrow)/gainsBAPts[,'no'] )

# No enrichment of gains in selected regions for BA 
p.adjust( apply(gainsBAPts, 1, function(x) {
  bt = binom.test(x[['no']], sum(x), population[['no']]/sum(population))  
  bt$p.value
}) )

cnBA = do.call(rbind, lapply(subset(oac.demo, !is.na(RP.BarrettsAdjacent))$StudySubjectID, function(s) {
  cbind.data.frame(
    'ID'=s,
    'BarrettsAdjacent'=subset(oac.demo, StudySubjectID == s)$RP.BarrettsAdjacent,
    'TotalCN'=length(subset(oacCNV, OCCAMS == s)),
    'Gains'=length(subset(oacCNV, OCCAMS == s  & ploidy > 2 & Total_CN > round(ploidy))),
    'Losses'=length(subset(oacCNV, OCCAMS == s  & Total_CN < 2)) 
  )
}))

arrange(cnBA, -TotalCN)

populationGains = apply(oacCNV, 1, function(x) {
  yes = subset(oac.demo, StudySubjectID %in% x$OCCAMS & RP.BarrettsAdjacent == 'yes')$StudySubjectID
  subset(g, OCCAMS %in% yes)

})

apply(cbind(sapply(gainsBA, nrow), sapply(gainsNoBA, nrow)), 1, function(x) {
  print(x)
  bt = binom.test(x[1], sum(x), .5)
  bt$p.value
})

# But there is a significant difference between adjacent BA -- but is this true across random selected regions as well?
gainCtBA = as.data.frame(do.call(rbind, lapply(losses, function(g) {
 table(subset(oac.demo, StudySubjectID %in% g$OCCAMS)$RP.BarrettsAdjacent)
})))

one.way.aov(gainCtBA$no, gainCtBA$yes)
wilcox.test(gainCtBA$no, gainCtBA$yes) 
summary(aov(gainCtBA$no~gainCtBA$yes)) # significant difference between adjacent BA -- but is this true across random selected regions as well?
# But not when adjusted for the population size, though not sure I should do that.  Binom instead?
one.way.aov(gainCtBA$no/population[['no']], gainCtBA$yes/population[['yes']])

lossCtBA = as.data.frame(do.call(rbind, lapply(losses, function(g) {
 table(subset(oac.demo, StudySubjectID %in% g$OCCAMS)$RP.BarrettsAdjacent)
})))

wilcox.test(lossCtBA$no, lossCtBA$yes) # significant difference between adjacent BA -- but is this true across random selected regions as well?
summary(aov(lossCtBA$no~lossCtBA$yes)) # significant difference between adjacent BA -- but is this true across random selected regions as well?

overlaps = rbind(gainCtBA, lossCtBA) %>% rowwise() %>% mutate(
   'yes.CI.1'=binom.test(yes, sum(yes,no), population[['yes']]/sum(population) )$conf.int[1],
   'yes.CI.2'=binom.test(yes, sum(yes,no), population[['yes']]/sum(population) )$conf.int[2],
   'yes.pval'=binom.test(yes, sum(yes,no), population[['yes']]/sum(population) )$p.value,
   'no.pval'=binom.test(no, sum(yes,no), population[['no']]/sum(population) )$p.value
) 
overlaps$regions = c(rownames(gainsBA), rownames(lossBA))

# not significant though
p.adjust(overlaps$yes.pval, method='bonferroni', n=nrow(overlaps))
p.adjust(overlaps$no.pval, method='bonferroni', n=nrow(overlaps))

df = transform(overlaps, regions=reorder(regions, (yes.CI.2-yes.CI.1)) )
ggplot(df , aes(x=regions)) + 
    geom_segment(data=df, aes(y=yes.CI.1, yend=yes.CI.2, x=regions,xend=regions), color='blue') +
    coord_flip() + labs(x='', y='CI (yes)')

```

General gains vs losses
```{r,eval=F}

head(perms)
# No difference in the numbers of gains/losses
wilcox.test(sapply(gains, nrow), perms$amp)
wilcox.test(sapply(losses, nrow), perms$del)

# There is a difference (these are the sampled regions, not selected) between the BA categories, yes -> more disordered??
wilcox.test(perms$amp.no, perms$amp.yes)
wilcox.test(perms$del.no, perms$del.yes)

ggplot(melt(perms[,c('amp.yes','amp.no')]), aes(variable, value, group=variable, fill=variable)) + geom_boxplot()
ggplot(melt(perms[,c('del.yes','del.no')]), aes(variable, value, group=variable, fill=variable)) + geom_boxplot()

random = as.data.frame(t(apply(perms[,grep('yes|no', colnames(perms))], 1, function(row) {

  yes = sum(row[c('amp.yes','del.yes')])
  no = sum(row[c('amp.no','del.no')])

  cbind('yes' = binom.test(yes, sum(yes,no), population[['yes']]/sum(population))$conf.int,
        'no' = binom.test(no, sum(yes,no), population[['yes']]/sum(population))$conf.int)
  # binom.test(yes, sum(yes,no), population[['yes']]/sum(population))$conf.int
})))
colnames(random) = c('yes.CI.1','yes.CI.2','no.CI.1','no.CI.2')

summary( with(random, yes.CI.2-yes.CI.1)  )
summary( with(overlaps, yes.CI.2-yes.CI.1)  )

summary( with(random, no.CI.2-no.CI.1)  )
summary( with(overlaps, no.CI.2-no.CI.1)  )

# No difference between the features selected and random regions though
wilcox.test(gainsBA$yes, perms$amp.yes)
wilcox.test(gainsBA$no, perms$amp.no)

# Nope
colSums(lossBA)/nrow(lossBA)
colSums(perms[,c('del.no','del.yes')])/nrow(perms)

t.test(lossBA$yes, perms$del.yes)
t.test(lossBA$no, perms$del.no)

wilcox.test(gfreq, perms$freq.gain)
wilcox.test(lfreq, perms$freq.loss)

m = rbind(
cbind.data.frame('freq'=gfreq, 'type'='gain', 'feature'='swgs'),
cbind.data.frame('freq'=lfreq, 'type'='loss', 'feature'='swgs'),

cbind.data.frame('freq'=perms$freq.gain, 'type'='gain', 'feature'='random'),
cbind.data.frame('freq'=perms$freq.loss, 'type'='loss', 'feature'='random') 
)

# No difference
ggplot(m, aes(feature, freq, group=feature, fill=feature)) + facet_grid(~type) + geom_jitter(width=0.2) +  geom_violin(alpha=0.8) 
```

```{r, eval=F}
all = swgRegions

swgRegions = swgRegions[swgRegions$hazard > quantile(swgRegions$hazard)['75%']]

cols = c('RP.BarrettsAdjacent', 'RP.TStage.PrimaryTumour', 'Weeks.Survival', 'FE.Patient.Died')
tmp = oac.demo[,cols]

tmp = tmp[-which(is.na(tmp$Weeks.Survival)),]

swgRegions = swgRegions[order(-swgRegions$hazard)]

```

```{r, eval=F}

#summary(as.data.frame(swgRegions)[,c('num.amp.CN','num.del.CN')])

library(gridExtra)
source('../ICGC-Clinical/libs/surv.R')

tmp$FE.Patient.Died = as.numeric(tmp$FE.Patient.Died)-1

tmp[which(is.na(tmp$FE.Patient.Died)), 'FE.Patient.Died'] = 0

sp = survival.plots(tmp, 'RP.BarrettsAdjacent', verbose=F)
grid.arrange(sp$kapmei)

time='Weeks.Survival'; event='FE.Patient.Died'
f = paste('`',paste(grep('AMP|DEL',colnames(tmp), value=T), collapse='`+`'), '`', sep='')

#f = paste('`AMP_', as.character(seqnames(swgRegions[1])), ':', start(swgRegions[1]), '-', end(swgRegions[1]), '`', sep='')

# surv =  Surv(tmp[[time]],tmp[[event]],type='right')
# terms = as.formula(paste("Surv(",time,",",event,", type='right') ~", 'RP.BarrettsAdjacent', sep=""))
# 
# cox = coxph(terms, data=tmp,na.action=na.omit,ties='breslow')
# cfit = survfit(cox)
#  
# ggsurv(cfit,cens.shape=3,lty.est=1,back.white=F, xlab="Time (weeks)") + 
#       annotate("text", x=200, y=.8,label=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),2),"\n", nrow(tmp), 'samples'))
# summary(cox)

```

### 2017-11-03 RegressionPvNP.Rmd
updating glm pipeline

@@
 
+## update the info as it has changed
+patient.data = lapply(patient.data, function(df) {
+  df$info = subset(patient.info, Patient == unique(df$info$Patient))
+  df
+})
+
+validation.patient.data = lapply(validation.patient.data, function(df) {
+  df$info = subset(validation.patient.info, Patient == unique(df$info$Patient))
+  df
+})
+
 pander(sum.patient.data[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Training patient set')
@@
 ```
+
+```{r, rawvalues, echo=F, eval=F}
+
+p = ggplot(chr.lengths, aes(x=genome.length))
+
+for (pt in subset(sum.patient.data, Status == 'P')$Hospital.Research.ID) {
+  df = merge(patient.data[[pt]]$seg.vals, chr.lengths[,c('chrom','chr.length','genome.length')])
+  df$g.start = df$start.pos; df$g.end = df$end.pos
+  df[which(df$chrom != 1), 'g.start'] = rowSums(df[which(df$chrom != 1), c('start.pos', 'genome.length')])
+  df[which(df$chrom != 1), 'g.end'] = rowSums(df[which(df$chrom != 1), c('end.pos', 'genome.length')])
+
+  df = melt(df, measure.vars=grep('^D',colnames(df),value=T))
+
+  p = p + geom_segment(data=df, aes(x=g.start,xend=g.end,y=value, yend=value,color=variable),show.legend=F)  
+}
+
+p1 = ggplot(chr.lengths, aes(x=genome.length)) + ylim(0,3)
+for (pt in subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID) {
+  df = merge(patient.data[[pt]]$seg.vals, chr.lengths[,c('chrom','chr.length','genome.length')])
+  df$g.start = df$start.pos; df$g.end = df$end.pos
+  df[which(df$chrom != 1), 'g.start'] = rowSums(df[which(df$chrom != 1), c('start.pos', 'genome.length')])
+  df[which(df$chrom != 1), 'g.end'] = rowSums(df[which(df$chrom != 1), c('end.pos', 'genome.length')])
+
+  df = melt(df, measure.vars=grep('^D',colnames(df),value=T))
+
+  p1 = p1 + geom_segment(data=df, aes(x=g.start,xend=g.end,y=value, yend=value,color=variable),show.legend=F)  
+}
+
+grid.arrange(p, p1, nrow=2)
+
+```
+
+## Quick look at our demographic information
+
+```{r echo=F}
+
+ggplot(patient.info, aes(Age.at.diagnosis, fill=Status)) + geom_histogram(binwidth=3) + labs(title='Age at diagnosis')
+
+ggplot(patient.info, aes(BMI, fill=Status)) + geom_histogram(binwidth=1) + labs(title='BMI')
+
+ggplot(patient.info, aes(Maximal,Circumference, color=Status)) + geom_point() + labs(title='C vs M')
+
+grid.arrange(ggplot(patient.info, aes(p53.Status, fill=Status)) + geom_bar(position = position_dodge()) + labs(title='p53 Staining'),
+  tableGrob(  rbind( 'Non-progressor'=summary(subset(patient.info, Status == 'NP')$p53.Status),
+                     'Progressor'=summary(subset(patient.info, Status == 'P')$p53.Status) ) ))
+
+p53Path = table(subset(patient.info, Status == 'P' & p53.Status == 1)$Pathology)
+```
+
+`r pander(p53Path/table(subset(patient.info, Status == 'P' & !is.na(p53.Status))$Pathology), caption='Ratio of progressors predicted by p53 staining per pathology where p53 staining was done (excluding N/As)')`
 
@@
 ```{r labelsPNP, echo=F, message=F, include=T}
-## binomial: dysplasia 1, BE 0
+## binomial: prog 1, np 0
 labels = unlist(sapply(patient.data, function(df) {
@@
 
-dysplasia.df = t(mergedDf[,names(labels)])
+dysplasia.df = t(mergedDf[,intersect(names(labels), colnames(mergedDf))])
+labels = labels[intersect(names(labels), colnames(mergedDf))]
 dim(dysplasia.df)
@@
 
-```{r cv.funcs, message=F, warning=F, echo=T, fig.height=6, fig.width=6}
-pi.hat<-function(x) exp(x)/(1+exp(x))
-
-non.zero.coef<-function(fit, s) {
-  cf =  as.matrix(coef(fit, s))
-  cf[which(cf != 0),][-1]
-}
-
-coef.stability<-function(opt, nz.list) {
-  for (i in 1:length(nz.list)) {
-    df = as.data.frame(nz.list[[i]])
-    colnames(df) = i
-    opt = merge(opt, df, by='row.names', all.x=T)
-    rownames(opt) = opt$Row.names
-    opt$Row.names = NULL
-    opt[,(i+1)] = as.integer(!is.na(opt[,(i+1)]))
-  }
-  opt = opt[order(sapply(rownames(opt), function(x) as.numeric(unlist(strsplit(x, ':'))[1])  )),]
-  return(opt)
-}
-
-create.patient.sets<-function(pts, n, splits, minR=0.2) {
-  # This function just makes sure the sets don't become too unbalanced with regards to the labels.
-  check.sets<-function(df, grpCol, min) {
-    sets = table(cbind.data.frame('set'=df[[grpCol]], 'labels'=labels[df$Samplename]))
-    while ( (length(which(sets/rowSums(sets) < minR) ) >= 2 | length(which(sets/rowSums(sets) == 0)) > 0) ) {
-      #print(sets/rowSums(sets))
-      s = sample(rep(seq(5), length = length(unique(df$Hospital.Research.ID))))
-      df2 = merge(df, cbind('Hospital.Research.ID'=unique(df$Hospital.Research.ID), 'tmpgrp'=s), by="Hospital.Research.ID")
-      df[[grpCol]] = df2$tmpgrp
-      sets = table(cbind.data.frame('set'=df[[grpCol]], 'labels'=labels[df$Samplename]))
-    }
-    return(df[[grpCol]])
-  }
-
-  s = sample(rep(seq(splits), length = length(unique(pts$Hospital.Research.ID))))
-  patients = merge(pts, cbind('Hospital.Research.ID'=unique(pts$Hospital.Research.ID), 'group'=s), by="Hospital.Research.ID")  
-  colnames(patients)[3] = c('fold.1')
-  patients$fold.1 = check.sets(patients, 'fold.1', minR)
-
-  for (i in 2:n) {
-    s = sample(rep(seq(5), length = length(unique(pts$Hospital.Research.ID))))
-    patients = merge(patients, cbind('Hospital.Research.ID'=unique(patients$Hospital.Research.ID), 'group'=s), by="Hospital.Research.ID")  
-    foldcol = grep('group',colnames(patients))
-    colnames(patients)[foldcol] = paste('fold',i,sep='.')
-    patients[[paste('fold',i,sep='.')]] = check.sets(patients, paste('fold',i,sep='.'))
-  }
-
-  return(patients)
-}
-
-binomial.deviance<-function(pmat, y) {
-  # Binomial deviance, lifted from cv.lognet
-  prob_min = 1e-05; prob_max = 1 - prob_min
-  pmat = pmin(pmax(pmat, prob_min), prob_max)
-  dev = apply(pmat, 2, function(p)  -2*((y==1)*log(p)+(y==0)*log(1-p)) )
-  return(dev)
-}
-
-crossvalidate.by.patient<-function(x,y,lambda,pts,a=1,nfolds=10, splits=5, fit=NULL, minR=0.2, select='deviance', ...) {
-  if (nfolds > 5) minR = 0.1
-  x = as.matrix(x)
-  message(paste("Running", splits, "splits",nfolds,"times on", paste(dim(x), collapse=':'), 'alpha=',a ))
-  fit.e = list()
-
-  if (ncol(pts) == (nfolds+2)) {
-    tpts = pts
-  } else {
-    tpts = create.patient.sets(pts, nfolds, splits, minR)
-  }
-
-  cv.pred = (matrix(nrow=0, ncol=length(lambda)))
-  cv.binomial.deviance = (matrix(nrow=0, ncol=length(lambda)))
-  for (n in 1:nfolds) {  
-    message(paste(n, "fold"))
-    setCol = grep(paste('^fold.',n,'$',sep=''), colnames(tpts))
-
-    cv.class = matrix(nrow=splits, ncol=length(lambda))
-    deviance = matrix(nrow=splits, ncol=length(lambda))
-    for (i in 1:splits) { 
-      message(paste(i, "split"))
-      test.rows = which(rownames(x) %in% tpts[which(tpts[,setCol] == i), 'Samplename'])
-      test = x[test.rows,]
-      training = x[-test.rows,]
-      # pre-spec lambda seq
-      fitCV = glmnet(training, y[-test.rows], lambda=lambda, alpha=a, family='binomial', standardize = F) 
-      # autoplot(fit) + theme(legend.position="none")
-
-      # Confusion matrix: quantitiative, pos results + neg results / number of test rows
-      pred <- pi.hat(predict(fitCV, test, type='link'))  # pi.hat turns these into probabilities from logit
-      cv.class[i,] = apply(pred, 2, function(p) {
-        (p%*%y[test.rows] + (1-p) %*% (1-y[test.rows]))/length(test.rows)
-      })
-
-      # Binomial deviance, lifted from cv.lognet
-      dev = binomial.deviance(predict(fitCV, test, type='response'), as.factor(y[test.rows]))
-      deviance[i,] = apply(dev, 2, weighted.mean, w=rep(1, nrow(dev)), na.rm=T)
-
-      fit.e[[length(fit.e)+1]] = fitCV
-    }
-    cv.pred = rbind(cv.pred, cv.class)    
-    cv.binomial.deviance = rbind(cv.binomial.deviance, deviance)
-  }
-
-  # Not really sure what to do with the binomial deviance now...
-  df = cbind.data.frame('lambda.at'=1:ncol(cv.pred),
-                        'mean'= colMeans(cv.pred), 
-                        'sme'= apply(cv.pred, 2, sd)/sqrt(nrow(cv.pred)), 
-                        'sd'= apply(cv.pred, 2, sd), 
-                        'lambda'= lambda,
-                        'log.lambda' = log(lambda),
-                        'mean.b.dev' = colMeans(cv.binomial.deviance),
-                        'sd.b.dev' = apply(cv.binomial.deviance, 2,sd),
-                        'sme.b.dev' = apply(cv.binomial.deviance, 2, sd)/sqrt(nrow(cv.binomial.deviance)),
-                        'lambda.min' = F, 'lambda.1se' = F)
-
-  if (select == 'classification') {
-    # Minimize the classification mean error & select min lambda
-    df[which.max(subset(df, sme < median(sme) & sme > min(sme))$mean), 'lambda.min'] = T
-
-    #df$lambda.min = df$mean == max(df$mean)
-    se1 = df[df$lambda.min == T, 'mean'] - df[df$lambda.min == T, 'sme']
-
-    lambda.1se.search = subset(df, mean < se1 & !lambda.min & log.lambda > subset(df, lambda.min)$log.lambda) 
-    if (nrow(lambda.1se.search) > 0) {
-      arrange(lambda.1se.search, -mean)[1:10,]
-      df[arrange(subset(df, mean < se1 & !lambda.min & log.lambda > subset(df, lambda.min)$log.lambda ), -mean)[1, 'lambda.at'], 'lambda.1se'] = T
-    }
-    #df[df$log.lambda < subset(df, mean == max(mean))$log.lambda-sd(df$log.lambda),][1, 'lambda-1se'] = TRUE
-  } else if (select == 'deviance') {
-    df[which.min(df$mean.b.dev), 'lambda.min'] = T
-
-    lmin = subset(df, lambda.min)$lambda
-    lambdas = subset(df, log.lambda <= log(lmin) + log(sd(lambda))/2 )
-    n = 3
-    while (nrow(lambdas) <= 0 & n < 10) {
-      lambdas = subset(df, log.lambda <= log(lmin) + log(sd(lambda))/n )
-      n = n+1
-    }
-    if (nrow(lambdas) <= 0) {
-      df[subset(df, log.lambda <= log(lmin) )[1, 'lambda.at'], 'lambda.1se'] = T  
-    } else {
-      df[subset(df, log.lambda <= log(lmin) + log(sd(lambda))/n )[1, 'lambda.at'], 'lambda.1se'] = T  
-    }
-  }
-
-  lambda.min = lambda[subset(df, lambda.min == T)$lambda.at] 
-  lambda.1se = lambda[subset(df, lambda.1se == T)$lambda.at]
-
-  nzcf = lapply(fit.e, non.zero.coef, s=lambda.1se)
-
-  plots = plot.patient.cv(df, fit)
-  plots$performance = plots$performance + theme(legend.position='bottom') + scale_colour_discrete(name = "") +
-    labs(title=paste(splits,' splits, ',nfolds,' folds, alpha=',a, sep=''), y='mean pred.', x='log(lambda)') 
-  plots$deviance = plots$deviance + theme(legend.position='bottom') + scale_colour_discrete(name="") + 
-    labs(title=paste(splits,' splits, ',nfolds,' folds, alpha=',a, sep=''), y='mean Binomial Deviance', x='log(lambda)')
-
-  return(list('max.cm'=df[df$lambda.1se ==T,'mean'], 
-              'lambda.min'=lambda.min, 'lambda.1se'=lambda.1se, 'lambdas'=df, 'plot'=plots$performance, 'deviance.plot'=plots$deviance, 'non.zero.cf'=nzcf))
-}
-
-plot.patient.cv<-function(df, fit=NULL) {
-  gp = ggplot(df, aes(y=mean,x=log.lambda)) + geom_point() + geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + 
-    geom_point(data=subset(df, lambda.min == T), aes(y=mean, x=log.lambda, colour="lambda.min"), size=2 ) +  
-    geom_vline(xintercept = subset(df, lambda.min == T)$log.lambda, colour='grey') +
-    annotate("text", x=subset(df, lambda.min == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
-             label=paste(round(df[subset(df, lambda.min == T)$lambda.at,c('mean', 'sme')],3), collapse='\n+/-'))
-
-  gpD = ggplot(df, aes(y=mean.b.dev,x=log.lambda)) + geom_point() +
-    geom_errorbar(aes(ymin=mean.b.dev-sme.b.dev, ymax=mean.b.dev+sme.b.dev)) + 
-    geom_point(data=subset(df, lambda.min == T), aes(y=mean.b.dev, x=log.lambda, colour="lambda.min"), size=2 ) +          
-    geom_vline(xintercept = subset(df, lambda.min == T)$log.lambda, colour='grey') +
-    annotate("text", x=subset(df, lambda.min == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
-             label=paste(round(df[subset(df, lambda.min == T)$lambda.at,c('mean', 'sme')],3), collapse='\n+/-'))
-
-  if (nrow(subset(df, lambda.1se == T)) > 0) {
-    gp = gp + geom_point(data=subset(df, lambda.1se == T), aes(y=mean, x=log.lambda, colour="lambda.1se"), size=2 ) +
-      geom_vline(xintercept = subset(df, lambda.1se == T)$log.lambda, colour='grey') +
-      annotate("text", x=subset(df, lambda.1se == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
-               label=paste(round(df[subset(df, lambda.1se == T)$lambda.at,c('mean', 'sme')],3), collapse='\n+/-'))
-
-    gpD = gpD + geom_point(data=subset(df, lambda.1se == T), aes(y=mean.b.dev, x=log.lambda, colour="lambda.1se"), size=2 ) +
-      geom_vline(xintercept = subset(df, lambda.1se == T)$log.lambda, colour='grey') +
-      annotate("text", x=subset(df, lambda.1se == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
-               label=paste(round(df[subset(df, lambda.1se == T)$lambda.at,c('mean', 'sme')],3), collapse='\n+/-')) 
-  }
-
-  gp = gp + theme(legend.position='bottom') + scale_colour_discrete(name = "") 
-  gpD = gpD + theme(legend.position='bottom') + scale_colour_discrete(name="")
-
-  if (!is.null(fit)) {
-    df$nzcoef = sapply(df$lambda, function(l) length(non.zero.coef(fit, l)))
-
-    coef.min = subset(df, lambda.min == T)$nzcoef
-    coef.1se = subset(df, lambda.1se == T)$nzcoef
-
-    d = coef.1se-coef.min
-
-    coef.text = arrange(subset(df,nzcoef %in% c(0,  coef.min, coef.1se, max(nzcoef) )), -lambda.min, -lambda.1se)
-    coef.text = arrange(coef.text[!duplicated(coef.text$nzcoef),], -nzcoef)
-
-    gp = gp + annotate("text", y=min(df$mean), x=coef.text$log.lambda, label=coef.text$nzcoef, color='darkblue')
-    gpD = gpD + annotate("text", y=max(df$mean.b.dev), x=coef.text$log.lambda, label=coef.text$nzcoef, color='darkblue')
-  }
-  return(list('performance'=gp, 'deviance'=gpD))
-}
+```{r, echo=F, message=F, warning=F}
+
+source('lib/cv-pt-glm.R')
 
@@
 
-coefs = list(); plots = list(); performance.at.1se = list(); models = list();
+coefs = list(); plots = list(); performance.at.1se = list(); models = list(); cvs = list()
 folds = 10; splits = 5 
@@
 } else {
-  #sets = create.patient.sets(pts, folds, splits, 0.2)  
-  #write.table(sets, quote=F, sep='\t', row.names=F, file=file)
+  sets = create.patient.sets(pts, folds, splits, 0.2)  
+  write.table(sets, quote=F, sep='\t', row.names=F, file=file)
   stop("Missing patient folds file")
@@
     fit0 <- glmnet(dysplasia.df, labels, alpha=a, nlambda=nl, family='binomial', standardize=F) # all patients
-    #autoplot(fit0, xvar='lambda', main=paste('fit0, all samples, alpha=',a,sep='')) + theme(legend.position='none')
 
     l = fit0$lambda
-    if (a == 0) {
-      l = sort(c(fit0$lambda, seq(exp(-5), exp(-10), -1e-6),
-            seq(exp(-10), exp(-15), -1e-8),
-            seq(exp(-15), exp(-20), -1e-9)), decreasing=T)
-    }
-    cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=l, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', standardize=F)
-    #cv.patient$deviance.plot
-
-    lambda.opt = ifelse( length(cv.patient$lambda.1se) > 0, cv.patient$lambda.1se, cv.patient$lambda.min )
+    #if (more.l) l = more.l(fit0$lambda)
+
+    cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=l, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', opt=-1, standardize=F)
+
+    lambda.opt = cv.patient$lambda.1se
 
@@
 
-    performance.at.1se[[as.character(a)]] = subset(cv.patient$lambdas, lambda.1se == T)
+    performance.at.1se[[as.character(a)]] = subset(cv.patient$lambdas, `lambda-1se`)
     models[[as.character(a)]] = fit0
-  }
-  save(plots, coefs, performance.at.1se, dysplasia.df, models, labels, file=file)
+    cvs[[as.character(a)]] = cv.patient
+  }
+  save(plots, coefs, performance.at.1se, dysplasia.df, models, cv.patient, labels, file=file)
 }
@@
   load(file, verbose=T)
+  # Fix info
+  pg.samp = lapply(pg.samp, function(df) {
+    merge(df[,c('Samplename',setdiff(colnames(df), colnames(patient.info))) ], subset(patient.info, Patient == unique(df$Patient)), by='Samplename')
+  })
+
 } else {
@@
     fitLOO <- glmnet(training, labels[train.rows], alpha=a, family='binomial', nlambda=nl, standardize=F) # all patients
-    #l = sort(c(fitLOO$lambda, seq(exp(-5), exp(-10), -0.001),
-    #      seq(exp(-10), exp(-15), -1e-7), 
-    #      seq(exp(-15), exp(-20), -1e-9)), decreasing=T)
     l = fitLOO$lambda
@@
   geom_boxplot( fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA) + geom_jitter() +
-  labs(y='performance at lambda.1se', x='', title='Performance for Leave One (patient) Out, at alpha=1')
+  labs(y='performance at lambda.1se', x='', title=paste('Performance for Leave One (patient) Out, at alpha=',select.alpha,sep=''))
 
@@
 
-```{r, warning=F, message=F, fig.height=4, fig.width=4, eval=F}
-### How many samples?  TBD
-
-#Clinicians asked how many samples they needed to take.
-
-# How do I evaluate the predictions?
-
-#Basically I need to to a FNR and FPR calculation that is extremely conservative and use that somehow
-
-all = do.call(rbind,lapply(pg.samp, function(df) {
-  df[,c('Status','Prediction','Prediction.Dev.Resid')]
-}))
-
-ggplot(all, aes(Prediction.Dev.Resid, Prediction, color=Status)) + geom_point()  + 
-  labs(y='Predicted (response)', x='Binomial dev residual?')
-```
-
 ## Data Quality
@@
 
-```{r, eval=F}
-a=1
-
-plot.cv<-function(cv, title='') {
-  grid.arrange( arrangeGrob(cv$plot+ggtitle('Classification'), cv$deviance.plot+ggtitle('Binomial Deviance'), top=title, ncol=2) )
-}
-
-more.l<-function(lambda) {
-    l = sort(c(lambda, seq(exp(-5), exp(-10), -1e-6),
-            seq(exp(-10), exp(-15), -1e-8),
-            seq(exp(-15), exp(-20), -1e-9)), decreasing=T)
-}
-
-# random
-df11 = dysplasia.df[,grep('^11:', colnames(dysplasia.df))]
-fit0 <- glmnet(df11, labels, alpha=a, nlambda=nl, family='binomial', standardize=F) # all patients
-cv11 = crossvalidate.by.patient(x=df11, y=labels, lambda=more.l(fit0$lambda), pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', standardize=F)
-plot.cv(cv11,"Chr11")
-
+```{r, eval=T}
+
+file = paste(cache.dir, 'chr17.predictive.Rdata', sep='/')
+if (file.exists(file)) {
+  message(paste("loading", file))
+  load(file, verbose=T)
+} else {
+  performance.at.1se = list(); coefs = list()
+  for ( d in c(1:22) ) {
+    print(d)
+    df = dysplasia.df[,grep(paste('^',d,':',sep=''), colnames(dysplasia.df))]
+    fit = run.fit(df, labels)
+    performance.at.1se[[ paste('chr', d, sep='') ]] = fit$perf
+    coefs[[ paste('chr', d, sep='') ]] = fit$coefs
+  }
+
+  ## No HGD
   df17 = dysplasia.df[,grep('^17:', colnames(dysplasia.df))]
-fit0 <- glmnet(df17, labels, alpha=a, nlambda=nl, family='binomial', standardize=F) # all patients
-cv17 = crossvalidate.by.patient(x=df17, y=labels, lambda=more.l(fit0$lambda), pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', standardize=F)
-plot.cv(cv17, 'Chr17')
-
-## No HGD
-df17 = dysplasia.df[subset(patient.info, Pathology %in% c('BE','ID','LGD'))$Samplename,grep('^17:', colnames(dysplasia.df))]
-fit0 <- glmnet(df17, labels[subset(patient.info, Pathology %in% c('BE','ID','LGD'))$Samplename], alpha=a, nlambda=5000, family='binomial', standardize=F)
-cv17 = crossvalidate.by.patient(x=df17, y=labels[subset(patient.info, Pathology %in% c('BE','ID','LGD'))$Samplename], lambda=fit0$lambda, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', standardize=F)
-plot.cv(cv17, 'Chr17, no HGD')
-
-df3 = dysplasia.df[,grep('^3:', colnames(dysplasia.df))]
-fit0 <- glmnet(df3, labels, alpha=a, nlambda=nl, family='binomial', standardize=F) # all patients
-cv3 = crossvalidate.by.patient(x=df3, y=labels, lambda=fit0$lambda, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', standardize=F)
-plot.cv(cv3, 'Chr3')
+  df17nohgd = df17[!is.na(match(rownames(dysplasia.df), subset(patient.info, Pathology %in% c('BE','ID','LGD'))$Samplename)),]
+  fit = run.fit(df17nohgd, labels[rownames(df17nohgd)])
+  performance.at.1se[['chr17.noHGD']] = fit$perf
+  coefs[['chr17.noHGD']] = fit$coefs
+
+  save(performance.at.1se, coefs, file=file)
+}
 

### 2017-11-03 features-OAC.Rmd
updating glm pipeline

@@
 
+plot.theme = theme(text=element_text(size=14, face='bold'), panel.background=element_blank(), axis.line=element_line(color='black'), panel.grid.major=element_line(color='grey90'))
+
 ## Lists the sequenced patients
@@
 totalCNV = merge(cnv, paired, by.x='Illumina.Barcode.full', by.y='SampleId')
+
+demo.cols = c('StudySubjectID','RP.TStage.PrimaryTumour','RP.BarrettsAdjacent','Weeks.Survival','RP.TumourDifferentiation','DI.ageAtDiagnosis','DI.PatientGender','FE.Patient.Died')
+totalCNV = merge(totalCNV, demo[,demo.cols], by.y='StudySubjectID', by.x='OCCAMS')
+
 oacCNV = makeGRangesFromDataFrame(totalCNV, keep.extra.columns = T)
 
-# Barretts
+# Barretts ADJACENT SAMPLES
 dir = '~/Data/ICGC/BE_OAC_Cohort/BA/data/ascat'
@@
 beCNV = merge(cnv, barretts.pairs, by.x='Illumina.Barcode.full', by.y='SampleId')
+beCNV = merge(beCNV, demo[,demo.cols], by.y='StudySubjectID', by.x='OCCAMS')
 beCNV = makeGRangesFromDataFrame(beCNV, keep.extra.columns = T)
@@
 be.demo = subset(demo, StudySubjectID %in% unique(beCNV$OCCAMS))
-
-table(be.demo$RP.BarrettsAdjacent)
-table(oac.demo$RP.BarrettsAdjacent)
 
@@
 bb.pts = do.call(rbind, lapply(swgsPredictions[rownames(born.bad)],annotate.by.pt))
-rd.pts = do.call(rbind, lapply(swgsPredictions[sample(which(status == 'P'), 15)],annotate.by.pt))
-
-sd = cbind.data.frame('gain'=apply(rawSWGSdf[subset(sd.pts, !Final)$Samplename,gains$label], 2, count.gains), 
-          'loss'=apply(rawSWGSdf[subset(sd.pts, !Final)$Samplename,gains$label], 2, count.losses))
-
-bb = cbind.data.frame('gain'=apply(rawSWGSdf[subset(bb.pts, !Final)$Samplename,gains$label], 2, count.gains), 
-          'loss'=apply(rawSWGSdf[subset(bb.pts, !Final)$Samplename,gains$label], 2, count.losses))
-
-rd = cbind.data.frame('gain'=apply(rawSWGSdf[subset(rd.pts, !Final)$Samplename,gains$label], 2, count.gains), 
-          'loss'=apply(rawSWGSdf[subset(rd.pts, !Final)$Samplename,gains$label], 2, count.losses))
-
-```
-I think this is driven by the number of samples more a difference in the two groups.
-The "born bad" group has a lot more samples than the "sudden" and if I randomly sample the same number of samples from the "born bad" group the difference disappears.
-So, if there's a difference here how to I look for it?
+
+# Can't randomly sample patients here because it's very likely that I'll end up with many of the same patients in both random and sudden/born bad groups
+# Could randomly select samples that are not HGD/IMC though...
+#rd.pts = do.call(rbind, lapply(swgsPredictions[names(status[which(status == 'P')])],annotate.by.pt))
+
+rd.samp = subset(do.call(rbind, swgsPredictions[names(status[which(status == 'P')])]), Pathology %in% c('BE','ID','LGD'), select=c('Prediction','Pathology','Samplename'))
+rd.samp = rd.samp[sample(nrow(rd.samp), 120),]
+
+gains.by.group = cbind.data.frame(
+  'sudden'=apply(rawSWGSdf[subset(sd.pts, !Final)$Samplename,gains$label], 2, count.gains),
+  'born.bad'=apply(rawSWGSdf[subset(bb.pts, !Final)$Samplename,gains$label], 2, count.gains),
+  'random'=apply(rawSWGSdf[rd.samp$Samplename,gains$label], 2, count.gains)
+)
+
+losses.by.group = cbind.data.frame(
+  'sudden'=apply(rawSWGSdf[subset(sd.pts, !Final)$Samplename,losses$label], 2, count.losses),
+  'born.bad'=apply(rawSWGSdf[subset(bb.pts, !Final)$Samplename,losses$label], 2, count.losses),
+  'random'=apply(rawSWGSdf[rd.samp$Samplename,losses$label], 2, count.losses)
+)
+```
+
+This may be driven by the number of samples. I'm not sure if this is correct or how to check at the moment.
+In all of these I exclude the "final" sample (typically HGD/IMC).
 
 ```{r}
-# Adjusted for the number of samples...is this right?
-wilcox.test(sd$gain/nrow(subset(sd.pts, !Final)), bb$gain/nrow(subset(bb.pts, !Final)))
-wilcox.test(sd$gain/nrow(subset(sd.pts, !Final)), rd$gain/nrow(subset(rd.pts, !Final)))
-wilcox.test(rd$gain/nrow(subset(rd.pts, !Final)), bb$gain/nrow(subset(bb.pts, !Final)))
-
-wilcox.test(sd$loss/nrow(subset(sd.pts, !Final)), bb$loss/nrow(subset(bb.pts, !Final)))
-wilcox.test(sd$loss/nrow(subset(sd.pts, !Final)), rd$loss/nrow(subset(rd.pts, !Final)))
-wilcox.test(rd$loss/nrow(subset(rd.pts, !Final)), bb$loss/nrow(subset(bb.pts, !Final)))
-
-```
-
-```{r, echo=F, eval=F}
-gains = gains[grep('\\d+',gains$label),]
-losses = losses[grep('\\d+',losses$label),]
-
-sd.samples = unlist(lapply(swgsPredictions[rownames(sudden)], function(df) df$Samplename))
-bb.samples = unlist(lapply(swgsPredictions[rownames(born.bad)], function(df) df$Samplename))
-random.samples = unlist(lapply(swgsPredictions[sample(which(status == 'P'), 15)], function(df) df$Samplename ))
-
-# Born bad losses/gains in features
-bbL = apply(rawSWGSdf[bb.samples, losses$label], 2, count.losses)
-bbG = apply(rawSWGSdf[bb.samples, gains$label], 2, count.gains)
-
-# sudden losses/gains in features
-snL = apply(rawSWGSdf[sd.samples, losses$label], 2, count.losses)
-snG = apply(rawSWGSdf[sd.samples, gains$label], 2, count.gains)
-
-# random losses/gains in features
-rrL = apply(rawSWGSdf[random.samples, losses$label], 2, count.losses)
-rrG = apply(rawSWGSdf[random.samples, gains$label], 2, count.gains)
-
+# "born bad" and sudden are different
+wilcox.test(gains.by.group$sudden, gains.by.group$born.bad)
+wilcox.test(losses.by.group$sudden, losses.by.group$born.bad)
+
+# Random and sudden are different
+wilcox.test(gains.by.group$sudden, gains.by.group$random)
+wilcox.test(losses.by.group$sudden, losses.by.group$random)
+
+# random and born bad aren't?
+wilcox.test(gains.by.group$born.bad, gains.by.group$random) ## not so different
+wilcox.test(losses.by.group$born.bad, losses.by.group$random) ## not so different
+
+```
+
+```{r echo=F}
 # one-way anova by ranks -- basically the same as wilcox I think
@@
 }
+```
+
+So why would random and born bad groups be the same?
+Sudden appears to be the only group that stands out in any way, being different from both born bad and random groups.
+Sudden has half the samples the other two do, but then random and sudden are different so samplesize should be at issue here.
+
+```{r, echo=F, eval=F}
+gains = gains[grep('\\d+',gains$label),]
+losses = losses[grep('\\d+',losses$label),]
+
+sd.samples = unlist(lapply(swgsPredictions[rownames(sudden)], function(df) df$Samplename))
+bb.samples = unlist(lapply(swgsPredictions[rownames(born.bad)], function(df) df$Samplename))
+random.samples = unlist(lapply(swgsPredictions[sample(which(status == 'P'), 15)], function(df) df$Samplename ))
+
+# Born bad losses/gains in features
+bbL = apply(rawSWGSdf[bb.samples, losses$label], 2, count.losses)
+bbG = apply(rawSWGSdf[bb.samples, gains$label], 2, count.gains)
+
+# sudden losses/gains in features
+snL = apply(rawSWGSdf[sd.samples, losses$label], 2, count.losses)
+snG = apply(rawSWGSdf[sd.samples, gains$label], 2, count.gains)
+
+# random losses/gains in features
+rrL = apply(rawSWGSdf[random.samples, losses$label], 2, count.losses)
+rrG = apply(rawSWGSdf[random.samples, gains$label], 2, count.gains)
 
@@
 
-```{r, warning=F, message=F, fig.width=7}
-
-tests = rbind(cbind.data.frame('CN'='gain', 'p'=wilcox.test(bbG, rrG)$p.value, 'type'='Random', 'label'='BB~R','y'=max(rrG)),
-              cbind.data.frame('CN'='gain', 'p'=wilcox.test(bbG, snG)$p.value, 'type'='Sudden', 'label'='BB~SN', 'y'=max(snG)),
-              cbind.data.frame('CN'='loss', 'p'=wilcox.test(bbL, rrL)$p.value, 'type'='Random', 'label'='BB~R','y'=max(rrL)),
-              cbind.data.frame('CN'='loss', 'p'=wilcox.test(bbL, snL)$p.value, 'type'='Sudden', 'label'='BB~SN', 'y'=max(snL)) )
-
-tmp = rbind(cbind(snG, 'type'='Sudden','CN'='gain'),
-            cbind(bbG, 'type'='Born Bad', 'CN'='gain'),
-            cbind(rrG, 'type'='Random', 'CN'='gain'),
-            cbind(snL, 'type'='Sudden', 'CN'='loss'),
-            cbind(bbL, 'type'='Born Bad', 'CN'='loss'),
-            cbind(rrL, 'type'='Random', 'CN'='loss' ) )
-tmp = as.data.frame(tmp, row.names = NA)
-tmp$sn = as.integer( as.character(tmp$sn) )
-tmp$type = factor(tmp$type, levels = c('Born Bad','Sudden','Random'))
+```{r, warning=F, message=F, echo=F, fig.width=7, fig.height=10}
+
+totals = melt(cbind.data.frame(
+  'sudden'=nrow(subset(sd.pts, !Final)),
+  'born.bad'=nrow(subset(bb.pts, !Final)),
+  'random'=nrow(rd.samp)))
 
 grid.arrange(
-  ggplot(tmp, aes(type, sn, group=type, fill=type)) + facet_grid(~CN) + 
-    geom_boxplot(outlier.colour = NA, alpha=0.3) + geom_jitter(width=0.2, aes(color=type)) + 
-    #geom_label(data=tests, aes(x=type,y=y,label=paste(label,format.pval(p,2),sep='\np=')), nudge_y=5, alpha=0.5)
-    geom_label(data=tests, aes(x=type,y=y,label=paste(label,format.pval(p,2),sep='\np=')), nudge_y=5, alpha=0.5) +
-    labs(y='CN count per region',x='', title='Gains & Losses in selected features') + theme(legend.position='none'),
-
-    tableGrob(  rbind(cbind('Group'='Born Bad','# Samples'=length(bb.samples)),
-                      cbind('Group'='Sudden','# Samples'=length(sd.samples)), 
-                      cbind('Group'='Random','# Samples'=length(random.samples))) ), nrow=2, heights=c(4,1) )
+  arrangeGrob(
+    ggplot(melt(gains.by.group), aes(variable, value, color=variable, fill=variable)) + geom_jitter(width=0.2) + 
+    geom_boxplot(alpha=0.5, outlier.color = NA) + labs(title='Gains', x='', y='counts') + theme(legend.position = 'none'),
+  ggplot(melt(losses.by.group), aes(variable, value, color=variable, fill=variable)) + geom_jitter(width=0.2) + 
+    geom_boxplot(alpha=0.5, outlier.color = NA) + labs(title='Losses', x='', y='counts') + theme(legend.position = 'none'), ncol=2),
+
+  tableGrob(  rbind(cbind('Group'='Born Bad','# Samples'=nrow(subset(bb.pts, !Final))),
+                    cbind('Group'='Sudden','# Samples'=nrow(subset(sd.pts, !Final))),
+                    cbind('Group'='Random','# Samples'=nrow(rd.samp)))   ) 
+  )
 
@@
 
-```{r}
+```{r, eval=F}
 
@@
   scale_colour_brewer(palette='YlGnBu', direction=-1) +
-  coord_flip() + 
-  theme(legend.position='bottom', axis.text=element_blank(),panel.background=element_blank()) + labs(y='chr position', x='Patient', title='Gains')
+  coord_flip() + plot.theme +
+  theme(legend.position='bottom', axis.text=element_blank(),panel.grid.major = element_blank()) + labs(y='chr position', x='Patient', title='Gains')
 }
@@
 
+gr = which(swgRegions$coef > 0)
+
 ## gains in patients -- issues with WGD samples?
-gains = lapply(swgRegions[which(swgRegions$coef > 0)], gain.for, cnv=oacCNV)
-names(gains) = as.character(swgRegions[which(swgRegions$coef > 0)])
+gains = lapply(swgRegions[gr], gain.for, cnv=oacCNV)
+names(gains) = as.character(swgRegions[gr])
 plot.gains(gains) + labs(title='Gains in OAC samples')
 
-be.gains = lapply(swgRegions[which(swgRegions$coef > 0)], gain.for, cnv=beCNV)
-names(be.gains) = as.character(swgRegions[which(swgRegions$coef > 0)])
+be.gains = lapply(swgRegions[gr], gain.for, cnv=oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'yes')])
+names(be.gains) = as.character(swgRegions[gr])
 plot.gains(be.gains[sapply(be.gains, nrow) > 0]) + labs(title='Gains in BE Adjacent samples')
 
-```
-
-```{r, fig.height=8, fig.width=8}
+non.be.gains = lapply(swgRegions[gr], gain.for, cnv=oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'no')])
+names(non.be.gains) = as.character(swgRegions[gr])
+plot.gains(non.be.gains[sapply(non.be.gains, nrow) > 0]) + labs(title='Gains in NO BE Adjacent samples')
+
+```
+
+```{r, fig.height=8, fig.width=12}
 
@@
 
-gfreq = sort(sapply(be.gains, function(x) length(unique(x$OCCAMS)))/length(unique(be.seq$OCCAMS_ID)))
-ggplot(melt(as.matrix(gfreq)), aes(Var1, value)) + geom_col(fill='darkblue')  + coord_flip() + 
-  labs(x='', y='Frequency', title='Gain frequency in BE adjacent') + theme_bw() + theme(text=element_text(size=14))
+gfreq1 = sort(sapply(be.gains, function(x) length(unique(x$OCCAMS)))/length(unique(oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'yes')]$OCCAMS)))
+gfreq1 = gfreq1[names(gfreq)]
+
+gfreq2 = sort(sapply(non.be.gains, function(x) length(unique(x$OCCAMS)))/length(unique(oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'no')]$OCCAMS)))
+gfreq2 = gfreq2[names(gfreq)]
+
+m = cbind.data.frame('yes'=gfreq1, 'no'=gfreq2)
+m$region = rownames(m)
+m = transform(m, region=reorder(region, yes))
+
+ggplot(melt(m), aes(region, value, group=variable, fill=variable)) + facet_grid(~variable) + geom_bar(stat='identity', position=position_dodge(), alpha=0.8) + 
+  scale_fill_manual(values=c('dodgerblue3','firebrick3'), name='Barretts Adjacent') +
+  coord_flip() + labs(x='', y='Frequency', title="Gain frequency BE vs non") + plot.theme + theme(legend.position = 'bottom')
 
@@
 plot.losses<-function(df) {
+  if (!is.data.frame(df)) {
     df = lapply(df, function(x) {
@@
     })
-
     m = do.call(rbind.data.frame, df)
+  } else {
+    m = df
+    m = arrange(m, -width)
+    m$order = factor(1:nrow(m))
+  }
   #m$segnames = ordered(m$seqnames, levels=c(1:22))
@@
   #m = transform(m, segment=reorder(segment, segnames))
-  ggplot(data=m, aes(x=order, y=start)) +  facet_wrap(~segment, scales='free') +
-    geom_segment(aes(x=order, xend=order, y=start, yend=end, color=CN)) + 
+  p = ggplot(data=m, aes(x=order, y=start))
+
+  if (length(unique(m$segment)) > 1)
+    p = p + facet_wrap(~segment, scales='free')
+
+  p + geom_segment(aes(x=order, xend=order, y=start, yend=end, color=CN)) + 
     geom_segment(aes(x=0,xend=0, y=seg.start, yend=seg.end), color='red', size=2) +
     scale_colour_brewer(palette='YlGnBu', direction=1) +
-    coord_flip() + 
-    theme(legend.position='bottom', axis.text=element_blank(),panel.background=element_blank()) + labs(y='chr position', x='Patient', title='Losses in OAC Patients')
+    coord_flip() + plot.theme +
+    theme(legend.position='bottom', axis.text=element_blank(), panel.grid.major=element_blank()) + labs(y='chr position', x='Patient', title='Losses in OAC Patients')
 }
@@
 
-losses = lapply(swgRegions[which(swgRegions$coef < 0)], loss.for, oacCNV)
-names(losses) = as.character(swgRegions[which(swgRegions$coef < 0)])
+lr = which(swgRegions$coef < 0)
+
+losses = lapply(swgRegions[lr], loss.for, oacCNV)
+names(losses) = as.character(swgRegions[lr])
 plot.losses(losses)
 
-be.losses = lapply(swgRegions[which(swgRegions$coef < 0)], loss.for, beCNV)
-names(be.losses) = as.character(swgRegions[which(swgRegions$coef < 0)])
-plot.losses(be.losses)
-
-```
-
-```{r, fig.height=8, fig.width=8}
+be.losses = lapply(swgRegions[lr], loss.for, oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'yes')])
+names(be.losses) = as.character(swgRegions[lr])
+plot.losses(be.losses) + labs(title="Losses in (YES) BE Adjacent samples")
+
+non.be.losses = lapply(swgRegions[lr], loss.for, cnv=oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'no')])
+names(non.be.losses) = as.character(swgRegions[lr])
+plot.losses(non.be.losses) + labs(title='Losses in (NO) BE Adjacent samples')
+
+# grid.arrange(plot.losses(be.losses$`16:76088214-80843726`) + labs(title='BE loss'),
+#              plot.losses(non.be.losses$`16:76088214-80843726`) + labs(title='(NO) BE loss') )
+
+```
+
+```{r, fig.height=8, fig.width=12}
 
 lfreq = sort(sapply(losses, function(x) length(unique(x$OCCAMS)))/length(unique(sequenced$OCCAMS_ID)))
-ggplot(melt(as.matrix(lfreq)), aes(Var1, value)) + geom_col(fill='darkgreen') + ylim(0,0.5) + coord_flip() + 
+ggplot(melt(as.matrix(lfreq)), aes(Var1, value)) + geom_col(fill='darkgreen') + ylim(0,1) + coord_flip() + 
   labs(x='', y='Frequency', title='Loss frequency in OAC') + theme_bw() + theme(text=element_text(size=14))
 
-lfreq = sort(sapply(be.losses, function(x) length(unique(x$OCCAMS)))/length(unique(be.seq$OCCAMS_ID)))
-ggplot(melt(as.matrix(lfreq)), aes(Var1, value)) + geom_col(fill='darkgreen') + ylim(0,0.5) + coord_flip() + 
-  labs(x='', y='Frequency', title='Loss frequency in BE adjacent') + theme_bw() + theme(text=element_text(size=14))
+lfreq1 = sapply(be.losses, function(x) length(unique(x$OCCAMS)))/length(unique(oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'yes')]$OCCAMS))
+lfreq1 = lfreq1[names(lfreq)]
+
+lfreq2 = sort(sapply(non.be.losses, function(x) length(unique(x$OCCAMS)))/length(unique(oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'no')]$OCCAMS)))
+lfreq2 = lfreq2[names(lfreq)]
+
+m = cbind.data.frame('yes'=lfreq1, 'no'=lfreq2)
+m$region = rownames(m)
+m = transform(m, region=reorder(region, yes))
+
+ggplot(melt(m), aes(region, value, group=variable, fill=variable)) + facet_grid(~variable) + geom_bar(stat='identity', position=position_dodge(), alpha=0.8) + 
+  scale_fill_manual(values=c('dodgerblue3','firebrick3'), name='Barretts Adjacent') +
+  coord_flip() + labs(x='', y='Frequency', title="Loss frequency BE vs non") + plot.theme + theme(legend.position = 'bottom')
+
+```
+
+```{r fig.height=6,fig.width=8}
 
@@
 
-m = arrange(losses$`17:1-4776188`, width, start)
-m$Anon = 1:length(m$OCCAMS)
-m = transform(m, Anon = reorder(Anon, -width))
-segment = cbind('seg.start'=1, 'seg.end'=4776188)
-m$CN = as.factor(m$Total_CN)
-#ggplot(m, aes(x=Anon, y=start)) +
-
-ggplot() + 
-  geom_segment(data=m, aes(x=Anon, xend=Anon, y=start, yend=end, color=CN), size=2) + 
-  geom_segment(data=segment, aes(x=0, xend=0, y=seg.start, yend=seg.end),color='red', size=3) + 
-  geom_hline(data=chr17, aes(yintercept=(chr.cent-cent.gap)), color='red', linetype='dashed') +
-  geom_label(data=chr17, aes(x=as.factor(3), y=chr.cent/2), label='p arm') +
-  scale_colour_brewer(palette='YlGnBu', direction=1) +
-  coord_flip() + 
-  theme(legend.position = 'bottom') + labs(y='chr17 position', x='Patient', title='Chr17 Losses in OAC Patients', subtitle=paste(round(length(unique(losses$`17:1-4776188`$OCCAMS))/length(unique(oacCNV$OCCAMS)), 2)*100, '% of OAC patients', sep=''), text=element_text(size=14, face='bold'))
-
-```
-
-So is there a difference between the gains/losses in the selected regions?  Not really.
+plot.losses(losses$`17:1-4776188`) + geom_hline(data=chr17, aes(yintercept=(chr.cent-cent.gap)), color='red', linetype='dashed') +
+  labs(y='chr17 position', x='Patient', title='Chr17 Losses in OAC Patients', subtitle=paste(round(length(unique(losses$`17:1-4776188`$OCCAMS))/length(unique(oacCNV$OCCAMS)), 2)*100, '% of OAC patients', sep=''))
+
+```
+
+So is there a difference between the gains/losses in the selected regions?  
 ```{r}
@@
 one.way.aov(sapply(gains, nrow), sapply(losses, nrow))
-
-# Maybe between BA and tumour...
-wilcox.test(sapply(gains, nrow), sapply(be.gains, nrow))
-one.way.aov(sapply(gains, nrow), sapply(be.gains, nrow))
+```
+
+Not between be/non be either
+```{r}
+wilcox.test(sapply(non.be.gains, nrow), sapply(be.gains, nrow))
+wilcox.test(sapply(non.be.losses, nrow), sapply(be.losses, nrow))
 ```
@@
 wilcox.test(sapply(randomGains, nrow), sapply(gains, nrow))
-one.way.aov(sapply(randomGains, nrow), sapply(gains, nrow))
+
+wilcox.test(sapply(randomGains, nrow), sapply(be.gains,nrow))
+wilcox.test(sapply(randomGains, nrow), sapply(non.be.gains,nrow))
 
@@
 
-# Not sure this shows a really clear difference, but...
 df = rbind(cbind.data.frame('counts'=sapply(gains, function(df) length(unique(df$OCCAMS))), 'type'='selected') , 
       cbind.data.frame('counts'=sapply(randomGains, function(df) length(unique(df$OCCAMS))), 'type'='random'))
-ggplot(df, aes(type, counts, group=type)) + geom_boxplot() + geom_jitter(width=0.3) + labs(title='Gain counts by region')
+ggplot(df, aes(type, counts, group=type, fill=type)) + geom_boxplot() + geom_jitter(width=0.3) + labs(title='Patient counts by region') + plot.theme
 
@@
       cbind.data.frame('counts'=sapply(randomGains, nrow), 'type'='random'))
-ggplot(df, aes(type, counts, group=type)) + geom_boxplot() + geom_jitter(width=0.3) + labs(title='Gain counts by patient')
+ggplot(df, aes(type, counts, group=type, fill=type)) + geom_boxplot() + geom_jitter(width=0.3) + labs(title='Gain counts per region') + plot.theme
+
+```
+
+```{r, echo=T}
+df = rbind(cbind.data.frame('counts'=sapply(non.be.gains, nrow), 'type'='non.be'),
+           cbind.data.frame('counts'=sapply(be.gains, nrow), 'type'='be'),
+           cbind.data.frame('counts'=sapply(gains, nrow), 'type'='all.selected'),
+           cbind.data.frame('counts'=sapply(randomGains, nrow), 'type'='random'))
+ggplot(df, aes(type, counts, group=type, fill=type, color=type)) + geom_boxplot(alpha=0.5, outlier.colour = NA) + 
+  geom_jitter(width=0.3) + labs(title='Gain counts per region') + plot.theme
+
+wilcox.test(subset(df, type == 'random')$counts, subset(df, type == 'be')$counts)
+wilcox.test(subset(df, type == 'random')$counts, subset(df, type == 'non.be')$counts)
+wilcox.test(subset(df, type == 'random')$counts, subset(df, type == 'all.selected')$counts)
 
@@
 wilcox.test(sapply(randomGains, nrow), sapply(randomLosses, nrow))
-one.way.aov(sapply(randomGains, nrow), sapply(randomLosses, nrow))
 
@@
 wilcox.test(sapply(randomLosses, nrow), sapply(losses, nrow))
-one.way.aov(sapply(randomLosses, nrow), sapply(losses, nrow))
-
-df = rbind(cbind.data.frame('counts'=sapply(randomLosses, nrow), 'type'='random') ,
-           cbind.data.frame('counts'=sapply(losses, nrow), 'type'='selected'))
-
-#df = df[-which(df$counts > 400),]
-# No difference between the two
-ggplot(df, aes(type, counts, group=type)) + geom_boxplot() + geom_jitter(width=0.3) + labs(title='Loss counts')
-
-summary(subset(df, type=='random')$counts)
-summary(subset(df, type=='selected')$counts)
-
-# No real difference in # patients with losses in the selected features
-wilcox.test(sapply(losses, function(df) length(unique(df$OCCAMS))), 
-            sapply(randomLosses, function(df) length(unique(df$OCCAMS))))
-one.way.aov(sapply(losses, function(df) length(unique(df$OCCAMS))), 
-            sapply(randomLosses, function(df) length(unique(df$OCCAMS))))
+
+wilcox.test(sapply(randomLosses, nrow), sapply(be.losses,nrow))
+wilcox.test(sapply(randomLosses, nrow), sapply(non.be.losses,nrow))
 
@@
       cbind.data.frame('counts'=sapply(randomLosses, function(df) length(unique(df$OCCAMS))), 'type'='random'))
-ggplot(df, aes(type, counts, group=type)) + geom_boxplot() + geom_jitter(width=0.3) + labs(title='Loss counts by patient')
-
-#pnorm(subset(df, type=='selected')$counts, mean=mean(subset(df, type=='selected')$counts), sd=sd(subset(df, type=='selected')$counts))
-#pnorm(subset(df, type=='random')$counts, mean=mean(subset(df, type=='random')$counts), sd=sd(subset(df, type=='random')$counts))
+ggplot(df, aes(type, counts, group=type, fill=type)) + geom_boxplot() + geom_jitter(width=0.3) + labs(title='Loss counts by patient') + plot.theme
+
+df = rbind(cbind.data.frame('counts'=sapply(non.be.losses, nrow), 'type'='non.be'),
+           cbind.data.frame('counts'=sapply(be.losses, nrow), 'type'='be'),
+           cbind.data.frame('counts'=sapply(losses, nrow), 'type'='all.selected'),
+           cbind.data.frame('counts'=sapply(randomLosses, nrow), 'type'='random'))
+ggplot(df, aes(type, counts, group=type, fill=type, color=type)) + geom_boxplot(alpha=0.5) + 
+  geom_jitter(width=0.3) + labs(title='Loss counts per region') + plot.theme
+
+```
+
+## Cox
+
+```{r}
+library(survival)
+library(GGally)
+
+pt.gains = do.call(rbind, lapply(oac.demo$StudySubjectID, function(id) {
+  rbind(sapply(gains, function(g)  nrow(subset(g, OCCAMS == id))))
+}))
+rownames(pt.gains) = oac.demo$StudySubjectID
+colnames(pt.gains) = paste('gain_', gsub(':|-', '\\.', colnames(pt.gains)), sep='')
+
+pt.losses = do.call(rbind, lapply(oac.demo$StudySubjectID, function(id) {
+  rbind(sapply(losses, function(g)  nrow(subset(g, OCCAMS == id))))
+}))
+rownames(pt.losses) = oac.demo$StudySubjectID
+colnames(pt.losses) = paste('loss_',gsub(':|-', '\\.', colnames(pt.losses)), sep='')
+
+x = merge(pt.gains, pt.losses, by='row.names', all=T)
+
+x = merge(x,oac.demo[,demo.cols], by.x='Row.names', by.y='StudySubjectID')
+x$Row.names = NULL
+
+time = 'Weeks.Survival';  event='FE.Patient.Died'; type = '"right"'
+x[c(time,event)] = lapply(x[c(time,event)], as.numeric)
+
+x$RP.TStage.PrimaryTumour = as.character(x$RP.TStage.PrimaryTumour)
+
+x[grepl('T1', x$RP.TStage.PrimaryTumour), 'RP.TStage.PrimaryTumour'] = 'T1'
+x[grepl('T4', x$RP.TStage.PrimaryTumour), 'RP.TStage.PrimaryTumour'] = 'T4'
+x[grepl('Tx', x$RP.TStage.PrimaryTumour), 'RP.TStage.PrimaryTumour'] = 'T0'
+
+terms = as.formula(paste("Surv(",time,",",event,", type=",type,") ~ RP.BarrettsAdjacent", sep=""))
+sfit = survfit(terms, data=x,conf.int=0.95, conf.type='log', na.action=na.omit)
+ggsurv(sfit)
+
+f1 = paste(demo.cols[c(3,5:7)], collapse='+')
+f2 = paste(paste('`', grep('^gain|loss', colnames(x), value=T), '`', sep=''), collapse='+')
+
+f = paste(f1,f2, sep='+')
+
+terms = as.formula(paste("Surv(",time,",",event,", type=",type,") ~", f, sep=""))
+cox = coxph(terms, data=x,na.action=na.omit,ties='breslow')
+cfit = survfit(cox)
+ggsurv(cfit) + labs(title=paste(length(grep('gain|loss', colnames(x))), 'copy number + demo'))
+
+pander(cox)
+
+anova(cox)
+```
+
+```{r}
+library(CoxHD)
+
+x = x[,c(grep('gain|loss',colnames(x), invert=T), grep('gain|loss',colnames(x)))]
+xs = x
+
+surv = with(xs, Surv(Weeks.Survival, FE.Patient.Died))
+ggsurv(survfit(surv~1,data=xs))
+
+xs = xs[,-grep('Weeks|Died',colnames(xs))]
+
+xs$RP.BarrettsAdjacent = as.integer(xs$RP.BarrettsAdjacent)-1
+xs$RP.TumourDifferentiation = as.integer(as.factor(xs$RP.TumourDifferentiation))-1
+xs$DI.PatientGender = as.integer(as.factor(xs$DI.PatientGender))-1
+
+groups = c(rep('clin', 4), rep('cn',(ncol(pt.gains)+ncol(pt.losses)) ))
+
+rfx = CoxRFX(xs, surv, groups, nu=1, verbose=T)
+summary(rfx)
+
+plot(rfx)
+
+VarianceComponents(rfx)
+
+WaldTest(rfx)
+
+PartialC(rfx)
+
+PlotVarianceComponents(rfx)
 
@@
 #  enrichment for some focal changes
-p.adjust(apply(sizeGains, 1, function(x) binom.test(x[1], sum(x), 0.5)$p.value ))
-apply(sizeGains, 1, function(x) binom.test(x[1], sum(x), 0.5)$conf.int )
+x = which(p.adjust(apply(sizeGains, 1, function(x) binom.test(x[1], sum(x), 0.5)$p.value )) < 0.05)
+apply(sizeGains, 1, function(x) binom.test(x[1], sum(x), 0.5)$conf.int )[,x]
+
+## Not sure what to do with this really
+sizeGainsBE = do.call(rbind.data.frame, lapply(be.gains, countBySize))
+which(p.adjust(apply(sizeGainsBE, 1, function(x) binom.test(x[1], sum(x), 0.5)$p.value )) < 0.05)
+sizeGainsNonBE = do.call(rbind.data.frame, lapply(non.be.gains, countBySize))
+which(p.adjust(apply(sizeGainsNonBE, 1, function(x) binom.test(x[1], sum(x), 0.5)$p.value )) < 0.05)
 
 # Same is true for gains of type by patient
-p.adjust(apply(sizeGainsByPatient, 1, function(x) binom.test(x[1], sum(x), 0.5)$p.value ))
-apply(sizeGainsByPatient, 1, function(x) binom.test(x[1], sum(x), 0.5)$conf.int )
+xx = which(p.adjust(apply(sizeGainsByPatient, 1, function(x) binom.test(x[1], sum(x), 0.5)$p.value )) < 0.05)
+apply(sizeGainsByPatient, 1, function(x) binom.test(x[1], sum(x), 0.5)$conf.int )[,xx]
+
+## Not sure what to do with this really
+sizeGainsBEByPt = do.call(rbind.data.frame, lapply(be.gains, sizebyPatient))
+which(p.adjust(apply(sizeGainsBEByPt, 1, function(x) binom.test(x[1], sum(x), 0.5)$p.value )) < 0.05)
+sizeGainsNonBEByPt = do.call(rbind.data.frame, lapply(non.be.gains, sizebyPatient))
+which(p.adjust(apply(sizeGainsNonBEByPt, 1, function(x) binom.test(x[1], sum(x), 0.5)$p.value )) < 0.05)
 
@@
 
+sizeLossBE = do.call(rbind.data.frame, lapply(be.losses, countBySize))
+sizeLossNoBE = do.call(rbind.data.frame, lapply(non.be.losses, countBySize))
+
+wilcox.test(sizeLossBE$large, sizeLossNoBE$large)
+
 # True for losses as well
@@
 gainsBAPts = as.data.frame(do.call(rbind, lapply(gains, function(g) {
- table(subset(oac.demo, StudySubjectID %in% g$OCCAMS)$RP.BarrettsAdjacent)
+ table(g$RP.BarrettsAdjacent)
 })))
+
+wilcox.test(gainsBAPts$no, gainsBAPts$yes)
 
@@
 gainsBA = lapply(gains, function(g) {
-  yes = subset(oac.demo, StudySubjectID %in% g$OCCAMS & RP.BarrettsAdjacent == 'yes')$StudySubjectID
-  subset(g, OCCAMS %in% yes)
+  subset(g, RP.BarrettsAdjacent == 'yes')
 })
@@
 gainsNoBA = lapply(gains, function(g) {
-  no = subset(oac.demo, StudySubjectID %in% g$OCCAMS & RP.BarrettsAdjacent == 'no')$StudySubjectID
-  subset(g, OCCAMS %in% no)
+  subset(g, RP.BarrettsAdjacent == 'no')
 })
 
+wilcox.test( sapply(be.gains, nrow), sapply(non.be.gains, nrow))
 # No difference

### 2017-11-03 lib/load_patient_metadata.R

@@ -1,6 +1,4 @@
 
-suppressPackageStartupMessages( library(dplyr) )
-
 strip.whitespace <- function (x) gsub("\\s+|\\s+", "", x)
 
 summarise.patient.info<-function(df) {
@@ -35,7 +33,21 @@ read.patient.info<-function(file, file2=NULL, set='Training') {
   if (grepl('\\.xlsx$', basename(file))) {
     library(readxl)
     
-    patient.info = NULL; 
+    patient.info = NULL;  #test.training = NULL
+    #for (sheet in excel_sheets(file)) {
+    #  print(sheet)
+    #for (i in 2:length(getSheets(loadWorkbook(file)))) {
+      #print(i)
+     # if (grepl('Technical Repeats|All_patient_info.txt', sheet)) next
+
+      #if (sheet == 'TrainingTest') {
+      #  test.training = read_excel(file, sheet = sheet)
+      #  next
+      #}
+      
+      #batch_name = gsub(' ', '_', sheet)
+      
+      #ws = read.xlsx2(file, sheetIndex=i, stringsAsFactors=F, header=T)
       sheet = 'All combined'
       ws = read_excel(file, sheet = sheet)
 
@@ -69,10 +81,24 @@ read.patient.info<-function(file, file2=NULL, set='Training') {
       
       colnames(ws) = c('Patient', 'Hospital.Research.ID', 'Path.ID','Status','Endoscopy.Year','Pathology','Plate.Index','SLX.ID','Barretts.Cellularity', 'p53.Status', 'Total.Reads', 'Batch.Name', 'Set')
       
+      #ws[ws$p53.Status == "", 'p53.Status'] = NA
+      #ws[ws$Barretts.Cellularity == "", 'Barretts.Cellularity'] = NA
+      
+      #ws$Batch.Name = batch_name
+      
+      #if (is.null(patient.info)) {
         patient.info = ws
+      #} else {
+        #print(paste('binding', batch_name))
+      #  patient.info = rbind(patient.info, ws)
+      #}
+    #}
+    #patient.info = merge(patient.info, test.training[,grep('Patient|Set|Analysis', colnames(test.training), value=T)], by.x='Patient', by.y='Patient', all=T)
   } else {
     patient.info = read.table(file, header=T, sep='\t', stringsAsFactors=F, quote="")
-
[…]

### 2017-11-03 run-cv.R

@@ -1,0 +1,365 @@
+library(ggplot2)
+library(plyr)
+library(dplyr)
+library(reshape2)
+library(gridExtra)
+library(GenomicRanges)
+library(glmnet)
+library(CoxHD)
+library(mice)
+
+source('lib/load_patient_metadata.R')
+source('lib/cv-pt-glm.R')
+
+
+args = commandArgs(trailingOnly=TRUE)
+
+data = args[1]
+tile.w = as.character(args[2])
+includeDemo = args[3]
+
+
+data = '~/Data/Ellie'
+tile.w=10e6
+includeDemo = F # demographics
+
+data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
+analysis.files = list.files(paste(data, 'Analysis', sep='/'), full.names=T)
+
+load(grep('All_patients.Rdata', analysis.files, value=T), verbose=T)
+
+## Hospital.Research.ID info file
+patient.file = grep('All_patient_info.xlsx', data.files, value=T)
+if (length(patient.file) != 1)
+  stop(paste("Missing/too many patient info file(s) in", data))
+demo.file = grep('Demographics_full.xlsx', data.files, value=T)
+
+
+all.patient.info = read.patient.info(patient.file, demo.file, set='all')$info
+
+patient.info = subset(all.patient.info, Set == 'Training')
+
+validation.patient.info =  subset(all.patient.info, Set == 'Test')
+validation.patient.info = arrange(validation.patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
+sum.validation.info = summarise.patient.info(validation.patient.info)
+
+patient.info = arrange(patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
+sum.patient.data = summarise.patient.info(patient.info)
+
+validation.patient.data = patient.data[sum.validation.info$Hospital.Research.ID]
+length(validation.patient.data)
+
+patient.data = patient.data[sum.patient.data$Hospital.Research.ID]
+patient.data = patient.data[!is.na(names(patient.data))]
+length(patient.data)
+
[…]

### 2017-11-08 Paper.Rmd
...

---
title: "Paper"
author: "Sarah Killcoyne"
date: "2 October 2017"
output: 
  html_document:
    fig_height: 5
    fig_width: 5
    toc: yes
    toc_depth: 4

---

```{r setup, include=FALSE}
library(biomaRt)
library(GenomicRanges)
library(stringr)
library(ggplot2)
library(ggrepel)
library(GGally)
library(plyr) 
library(pander)
library(reshape2)
library(gridExtra)
library(CoxHD)
library(plyr)
library(dplyr)

source('lib/load_patient_metadata.R')
source('lib/cv-pt-glm.R')

get.legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

plot.theme = theme(text=element_text(size=14, face='bold'), panel.background=element_blank(), strip.background =element_rect(fill="lightcyan2"),  
                   strip.text = element_text(size=14, face='bold'), 
                   axis.line=element_line(color='black'), panel.grid.major=element_line(color='grey90'),
                   panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing = unit(0.1, 'lines')  )

data = '~/Data/Ellie'
tile.w='5e+06_arms'
#tile.w=5e+06
#tile.w=1e7
#tile.w='arms'

includeDemo = F

data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
analysis.files = list.files(paste(data, 'Analysis', sep='/'), full.names=T)

load(grep('All_patients.Rdata', analysis.files, value=T), verbose=T)

## Hospital.Research.ID info file
patient.file = grep('All_patient_info.xlsx', data.files, value=T)
if (length(patient.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", data))
demo.file = grep('Demographics_full.xlsx', data.files, value=T)

all.patient.info = read.patient.info(patient.file, demo.file, set='all')$info
head(all.patient.info)

patient.info = subset(all.patient.info, Set == 'Training')

validation.patient.info =  subset(all.patient.info, Set == 'Test')
validation.patient.info = arrange(validation.patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
sum.validation.info = summarise.patient.info(validation.patient.info)

patient.info = arrange(patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
sum.patient.data = summarise.patient.info(patient.info)

validation.patient.data = patient.data[sum.validation.info$Hospital.Research.ID]
length(validation.patient.data)

patient.data = patient.data[sum.patient.data$Hospital.Research.ID]
length(patient.data)

patient.data = patient.data[!is.na(names(patient.data))]
names(patient.data)

sum.patient.data = as.data.frame(subset(sum.patient.data, Hospital.Research.ID %in% names(patient.data))) ## one patient was removed from the study
nrow(sum.patient.data)

# Missing some of the samples as there were some normals that we removed from analysis
# for (pt in names(patient.data)) {
#   patient.data[[pt]]$info = patient.data[[pt]]$info[patient.data[[pt]]$info$Samplename %in% colnames(patient.data[[pt]]$seg.vals)[-(1:5)],]
# }

# for (pt in names(validation.patient.data)) {
#   validation.patient.data[[pt]]$info = validation.patient.data[[pt]]$info[validation.patient.data[[pt]]$info$Samplename %in% colnames(validation.patient.data[[pt]]$seg.vals)[-(1:5)],]
# }

pander(sum.patient.data[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Training patient set')

pander(sum.validation.info[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Validation patient set')

all.info = rbind(sum.patient.data, sum.validation.info)

```

# Segment size: `r as.character(tile.w)` with demographics: `r includeDemo`

```{r echo=F}

pander( 
  as.data.frame(all.info %>% group_by(Status) %>% summarise(
  'Total Patients'=length(unique(Patient)),
  'Total Samples'=sum(total.samples),
  'Median Years'=paste(median(end.year-start.year),'+/-',round(sd(end.year-start.year),1)),
  'Median Endoscopies'=paste(median(total.endos),'+/-',round(sd(total.endos),1)),
  'Median age'=paste(median(age.diagnosis),'+/-',round(sd(age.diagnosis),1)),
  'gender F:M'=paste(table(gender),collapse=':')
))
, caption='Full cohort')

```

## Quick look at our demographic information

We tested models that included age, BMI, smoking status, sex, length & circumference of the BE segment, and p53 IHC status.  While none of these contributed we can report on some here.

```{r echo=F, warning=F, message=F, fig.height=4, fig.width=4}
ggplot(patient.info, aes(Age.at.diagnosis, fill=Status)) + geom_histogram(binwidth=3) + labs(title='Age at diagnosis') + plot.theme

ggplot(patient.info, aes(BMI, fill=Status)) + geom_histogram(binwidth=1) + labs(title='BMI') + plot.theme

ggplot(patient.info, aes(Maximal,Circumference, color=Status)) + geom_point() + labs(title='C vs M') + plot.theme

p53Path = table(subset(patient.info, Status == 'P' & p53.Status == 1)$Pathology)
p53.pred.byPath = round(p53Path/table(subset(patient.info, Status == 'P' & !is.na(p53.Status))$Pathology), 2)

```

```{r, echo=F}
x = do.call(rbind.data.frame, lapply(patient.data, function(df) df$info[,c('Status','p53.Status')]))
grid.arrange(
  ggplot(patient.info, aes(p53.Status, fill=Status)) + geom_bar(position = position_dodge()) + labs(title='p53 Staining') + plot.theme,
  tableGrob( 
    rbind( 'Non-progressor'=summary(subset(x, Status == 'NP')$p53.Status), 
           'Progressor'=summary(subset(x, Status == 'P')$p53.Status))), 
  nrow=2, as.table=T, heights=c(4,1) )
```

Because we're looking at both a binary classifier and binary data the AUC looks very good, there are no false positives after all. However of the progressor samples that were stained only `r round((table(subset(patient.info, Status == 'P')$p53.Status)/sum(table(subset(patient.info, Status == 'P')$p53.Status)))[2], 2)*100`% positively stained for p53. If we look at pathology the breakdown is even more obvious.  `r p53.pred.byPath['HGD']*100`% of HGD samples were identified, but only `r p53.pred.byPath['BE']*100`% NDBE samples were identified in progressors.

`r pander(p53.pred.byPath, caption='Ratio of progressors predicted by p53 staining per pathology where p53 staining was done (excluding N/As)')`

# Data set up

To use GLMs across the patients I merge all samples from all patients and overlap the copy number segments. Where segments from a sample get split, the value of the original sample is retained for each of the split samples.

```{r tile, echo=F, message=F, warning=F, include=F}
tileFile = paste('tile_patients_All_',as.character(tile.w),'.Rdata',sep='')
tile.files = grep(tileFile, analysis.files, value=T, fixed=T)
load(tile.files[grep('tile', tile.files)], verbose=T)

mergedDf = mergedDf[grep('X|Y',rownames(mergedDf), invert=T),]

mergedDfValidation = mergedDf[,validation.patient.info$Samplename]
dim(mergedDfValidation)

mergedDf = mergedDf[, intersect(colnames(mergedDf), patient.info$Samplename)]
dim(mergedDf)

cache.dir = paste(data, 'Analysis',sub('\\+', '', tile.w),   sep='/')
if (includeDemo)
  cache.dir = paste(cache.dir, '_with-demo', sep='')

```

`r pander(mergedDf[sort(sample(nrow(mergedDf), 10)), sample(ncol(mergedDf), 10) ], caption='Example rows/columns from the merged data')`

## y( Progressors (1) vs Non (0) )

The labels are split such that all samples from progressor patients are (1) and all samples from non-progressors are (0).

```{r labelsPNP, echo=F, message=F, include=T}
## binomial: dysplasia 1, BE 0
labels = unlist(sapply(patient.data, function(df) {
  df$info = plyr::arrange(df$info, Endoscopy.Year, Pathology)
  label = as.integer(df$info$Status == 'P') #as.integer(df$info$Pathology %in% c('HGD', 'IMC'))
  names(label) = df$info$Samplename
  return(label)
}))
names(labels) = sub('.*\\.', '',  names(labels))

pts = do.call(rbind, lapply(patient.data, function(df) {
  df$info = plyr::arrange(df$info, Endoscopy.Year, Pathology)
  cbind(df$info[,c('Hospital.Research.ID','Samplename')])
}))
rownames(pts) = 1:nrow(pts)

# sort in label order
#subset(patient.info, Samplename %in% setdiff(colnames(mergedDf), names(labels))[-c(1:3)])
if (length(setdiff(colnames(mergedDf), names(labels))) > 3)
  warning("Labels vector is missing samples")

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=T)
rm(plots, coefs, performance.at.1se, models, cv.patient)

```

We have `r table(labels)[1]` samples from non-progressors and `r table(labels)[2]` samples from progressors.

Example subset of the data matrix:
`r pander(dysplasia.df[1:5, 1:5], caption=paste("dimensions:", paste(dim(dysplasia.df), collapse=', ')), justify='left')`

## Global copy number plots

```{r echo=F, warning=F, message=F, eval=T, fig.width=7, fig.height=5}
chr.lengths = get.chr.lengths()
chr.lengths$chrom = sub('chr','',chr.lengths$chrom)
chr.lengths = subset(chr.lengths, chrom %in% c(1:22))

locs = do.call(rbind.data.frame, lapply(rownames(mergedDf), function(x) unlist(strsplit(x,':|-'))))
colnames(locs) = c('chr','start','end')
locs[] = lapply(locs[], function(x) as.numeric(as.character(x)))

globalCNdf = locs
globalCNdf = cbind(globalCNdf, t(dysplasia.df[,grep('^\\d+:', colnames(dysplasia.df), value=T)]))
globalCNdf = merge(globalCNdf, chr.lengths[,c('chrom','chr.length')], by.x='chr', by.y='chrom')
globalCNdf = melt(globalCNdf, id.vars=c('chr','start','end', 'chr.length'))
globalCNdf$Status = ifelse(globalCNdf$variable %in% names(which(labels == 1)), 'Progressor', 'Non-Progressor')

globalCNdf = merge(globalCNdf, patient.info[,c('Hospital.Research.ID','Patient','Samplename','Pathology')], by.x='variable', by.y='Samplename')

#globalCNdf = subset(globalCNdf, value != 0)

cn = function(x) {
  str = 'neutral'
  if(x <= -1) {
    str = 'loss'
  } else if (x >= 1) {
    str = 'gain'
  }
  return(str)
}

globalCNdf$cn = factor(sapply(globalCNdf$value, cn), levels=c('loss','neutral','gain'), ordered = T)
globalCNdf$cn = factor(sapply(globalCNdf$value, cn), levels=c('loss','neutral','gain'), ordered = T)
globalCNdf$cnB = factor(as.integer(globalCNdf$cn != 'neutral'))

cn.colors = c('orange3','grey88','steelblue2')

n.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor')$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor')$variable)),')', sep=''))

# Average cn number...looks the same as expected
globalCNdf = globalCNdf %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
globalCN.plot = 
  ggplot(globalCNdf, aes(x=chr.length)) + facet_grid(Status~chr, scales='free_x', space='free_x', labeller=labeller(Status=n.status, chr=label_value)) +
  geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=cn), alpha=0.5, show.legend = T) + 
  scale_color_manual(values=cn.colors) + 
  scale_fill_manual(values=cn.colors) +
  labs(x='', y='Mean adjusted segmentation value', title='All pathologies') +
  theme(text=element_text(face='bold', size=14), axis.text.x=element_blank(), legend.position='bottom', 
        panel.grid = element_blank(), panel.background = element_blank(),strip.background =element_rect(fill="lightcyan2"),  
        panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0, 'lines') )

cnLegend = get.legend(globalCN.plot)
globalCN.plot = globalCN.plot + theme(legend.position = 'none')

nb.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology == 'BE')$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology == 'BE')$variable)),')', sep=''))

globalCNBEdf = subset(globalCNdf, Pathology == 'BE') %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
globalBECN.plot = 
  ggplot(globalCNBEdf, aes(x=chr.length)) + 
  facet_grid(Status~chr, scales='free_x', space='free_x', labeller=labeller(Status=nb.status, chr=label_value)) + 
  #ylim(range(globalCNdf$value)) +
  geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=cn), alpha=0.5, show.legend = T) + 
  scale_color_manual(values=cn.colors) + scale_fill_manual(values=cn.colors) +
  labs(x='', y='Mean adjusted segmentation value', title='NDBE samples only') +
  theme(text=element_text(face='bold', size=14), axis.text.x=element_blank(), legend.position='bottom', 
        panel.grid = element_blank(), panel.background = element_blank(),strip.background =element_rect(fill="lightcyan2"),  
        panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0, 'lines') )

globalCN.plot 
globalBECN.plot

dfpts = subset(globalCNdf, Patient %in% c('57','76'))
dfpts = dfpts %>% group_by(Patient, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )

pt.status <- c('57'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
              '76'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))

genomePtCN.plot = 
ggplot(dfpts, aes(x=chr.length)) + 
  facet_grid(Patient~chr, scales='free_x',space='free_x',  labeller = labeller(Patient = pt.status )) + 
  geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=cn), show.legend = F, alpha=0.5) + 
  scale_fill_manual(values=cn.colors) +
  labs(x='', y='Mean adjusted segmentation value') +
  theme(text=element_text(face='bold', size=14), axis.text.x=element_blank(), legend.position='bottom', 
        panel.grid = element_blank(), panel.background = element_blank(),strip.background =element_rect(fill="lightcyan2"),  
        panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0, 'lines') )

dfpts = subset(globalCNdf, Patient %in% c('57','76') & Pathology == 'BE')
dfpts = dfpts %>% group_by(Patient, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )

ptbe.status <- c('57'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
              '76'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))
genomePtBECN.plot = 
  ggplot(dfpts, aes(x=chr.length)) + 
  facet_grid(Patient~chr, scales='free_x', space='free_x', labeller = labeller(Patient = ptbe.status )) + 
  geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=cn), show.legend = F, alpha=0.5) + 
  scale_fill_manual(values=cn.colors) +
  labs(x='', y='Mean adjusted segmentation value') +
  theme(text=element_text(face='bold', size=14), axis.text.x=element_blank(), legend.position='bottom', 
        panel.grid = element_blank(), panel.background = element_blank(),strip.background =element_rect(fill="lightcyan2"),  
        panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0, 'lines') )

genomePtCN.plot 
genomePtBECN.plot
```

# Building the model

We built the model using a 50-fold cross-validation process on the patients (rather than samples) for various lambda and alpha values to find the best fitting model.

```{r model, message=F, warning=F, echo=F, fig.height=16, fig.width=16}
folds = 10; splits = 5 
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=F)
} else {
  stop(paste("No model file", file, "found"))
}

do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits'))
```

## Model performance

```{r model_perf, message=F, warning=F, echo=F, fig.width=6, fig.height=6}
model.performance<-function(p1se, coeffs ) {
  performance = do.call(rbind.data.frame, p1se)
  performance$alpha = rownames(performance)

  if (grepl('chr', performance$alpha[1])) {
    performance$alpha = factor(performance$alpha, levels=c(paste('chr',c(1:22), sep=''), setdiff(rownames(performance), paste('chr',c(1:22),sep=''))) )
  }

  ## check how often the feature(s) is selected at that lamda in each split. "stability selection"
  coef.stable = lapply( coeffs, function(cf) {
    sort(rowSums(cf[,-1]), decreasing=T)
  })
  coef.stable = coef.stable[names(which(sapply(coef.stable, length) > 0))]
  coef.stable = lapply(coef.stable, function(cf) cf/(folds*splits))

  performance = cbind(performance, do.call(rbind,  lapply(coef.stable, function(x) {
    cbind('n.Coef'=length(x), '25%'=length(which(x>=.25)), 
          '50%'=length(which(x>=.5)), '75%'=length(which(x>=.75 )), 
          '100%'=length(which(x==1)) )
  })))

  p = ggplot(performance, aes(alpha, mean, color=(alpha != 0.9))) + 
    geom_point(size=2) +
    geom_text( aes(label=round(mean, 3)), hjust=-0.5) +
    geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme), width=0.3, size=1) + 
    geom_text(aes(label=n.Coef), nudge_y=0.03, fontface='bold') +
    geom_text(aes(label=paste('(',`75%`,')',sep='')), nudge_y=0.02) + 
    scale_colour_manual(values=c('red', 'grey39')) + plot.theme + theme(legend.position = 'none') +
    labs(x='Elasticnet penalty value: ridge <-> lasso', y='Model classification at lambda-1se')

  cfs = as.data.frame(matrix(data=0,nrow=length(unique(names(table(unlist(lapply(coef.stable, names)))))), ncol=length(names(coef.stable)), 
                           dimnames=list( unique(names(table(unlist(lapply(coef.stable, names))))), names(coef.stable) )))
  for (i in names(coef.stable)) 
    cfs[intersect(rownames(cfs), names(coef.stable[[i]])), i] = 1
  return(list('performance'=performance, 'stable.coeffients'=coef.stable, 'cf.per.feature'=cfs,'plot'=p))
}

classification.rate = as_tibble(matrix(ncol=5, nrow=0, dimnames=list(c(), c('model','mean','sme','stable.coef', 'all.coef'))))

mp = model.performance(performance.at.1se, coefs)
mp$plot + labs(title=paste('All samples,', folds, 'folds,', splits, 'patient splits'))

select.alpha = as.character(0.9)
# Genomic regions only
most.stable = grep('^\\d+', names(which(mp$stable.coeffients[[select.alpha]] >= 0.75 )), value=T)

most.stable = cbind.data.frame(do.call(rbind, strsplit(most.stable, ':|-')), most.stable)
colnames(most.stable) = c('chr','start','stop', 'name')

most.stable$chr = as.integer(as.character(most.stable$chr))
most.stable = arrange(most.stable, chr, start)

classification.rate = add_row(classification.rate, 'model'='All samples', 'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))
```

`r pander(mp$performance[,c('n.Coef','25%','50%','75%', '100%')], justify='left', caption='Number of features stable in n% or more folds')`

Values that are closer to full lasso (alpha=1) are pretty similar, however at full lasso we see fewer features than if we regularize it at 0.8. They share all non-zero features as well.
`r pander(sapply(mp$stable.coeffients, length), caption='Total features at each value of alpha')`

`r pander(most.stable[,c('chr','start','stop')], caption='Features identified as stable in 75% or more folds.')`

## Feature Stability

They share all of the non-zero coefficients (`r which.max(sapply(mp$stable.coeffients, length))` identifies the most non-zero coefs `r max(sapply(mp$stable.coeffients, length))`).

## Remove HGD/IMC & LGD Samples and check the model

Just as a point, the initial model is trained against `r nrow(dysplasia.df)` samples with `r ncol(dysplasia.df)` features. The samples include the final (diagnostic) HGD/IMC samples from progressors.

### Remove HGD/IMC

```{r noHGD, echo=F, message=F, warning=F, fig.height=16, fig.width=16, eval=F}
file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=F)
} else {
  stop(paste("No model file", file, "found"))
}

#do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2))
```

```{r, echo=F, fig.height=5, fig.width=5, eval=F}
mp = model.performance(performance.at.1se, coefs)
mp$plot + labs(title=paste('No HGD/IMC samples,', folds, 'folds,', splits, 'patient splits'))

classification.rate = add_row(classification.rate, 'model'='No HGD/IMC', 
                              'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]])
                              )
```

### Remove LGD

```{r noLGD, echo=F, warning=F, message=F, fig.height=16, fig.width=16}
file = paste(cache.dir, 'nolgd.Rdata', sep='/')

if (file.exists(file)) {
  load(file, verbose=F)
} else {
  stop(paste("No model file", file, "found"))
}

```

```{r, echo=F, fig.height=5, fig.width=5, eval=F}
mp = model.performance(performance.at.1se, coefs)
mp$plot + labs(title=paste('No LGD/HGD/IMC samples,', folds, 'folds,', splits, 'patient splits'))

classification.rate = add_row(classification.rate, 'model'='No LGD/HGD/IMC', 'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))
```

## Overall

Even without dysplastic samples our model is well able to classify samples as either progressive or non.

```{r, echo=F, warning=F, message=F,fig.height=5, fig.width=5}
#classification.rate$x = 1:nrow(classification.rate)

ggplot(classification.rate, aes(model, mean, group=model, color=model)) + 
  geom_text( aes(label=round(mean, 3)), hjust=-0.5, size=5, color='gray39', show.legend = F) +
  geom_text( aes(y=mean+sme*3, label=paste(all.coef, "\n(", stable.coef, ")", sep='') ), color='gray39', size=5) +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme),color='grey39', width=0.1, size=1) + 
  geom_point(size=4) +
  labs(x='', y='Classificaiton rate', title=paste('Classification rate per model at', select.alpha)) + 
  plot.theme + theme(axis.text.x=element_text(angle=45,hjust=1,face="bold",size=12), legend.position = 'none')

```

# Features

```{r, echo=F, warning=F, message=F}
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=T)
rm(plots, performance.at.1se)

featureStability = rowSums(coefs[[select.alpha]][,-1])/(splits*folds)

hazards = apply( exp(t(dysplasia.df[, rownames(coefs[[select.alpha]])]) *  coefs[[select.alpha]][,1]), 1, function(x) {
  sd(x)/mean(x)
})
ch = as.data.frame(hazards)
ch$coef = coefs[[select.alpha]][,1]
ch$stability = featureStability
ch$label = rownames(ch)
ch = arrange(ch, -hazards)

qq = quantile(ch$hazards, seq(0,1,.15))
ggplot(ch, aes(coef, hazards, col=hazards >= qq[length(qq)])) + 
  geom_point(alpha=0.8, show.legend = F) + 
  geom_text_repel(aes(x=coef, label=ifelse(hazards >= qq[length(qq)], label, '')),show.legend=F) +
  scale_color_manual(values=c('firebrick3','darkblue')) +
  labs(title='Top features by hazard ratio', x='Coefficient value', y='Hazard') + plot.theme

## Arms...
if (exists('arm.locs')) {
  ch[which(ch$label %in% with(arm.locs, paste(chr, ':', start, '-', end, sep=''))),]
}

write.table(ch, quote=F, sep='\t', row.names = F, col.names = T, file=paste(as.character(tile.w),'_features.txt'))
```

`r pander(ch, caption='Selected coefficients')`

```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
genomeFeat = grep('^\\d+:', ch$label)
featureAnn = data.frame(apply(do.call(rbind, strsplit(ch$label[genomeFeat], ':|-')), 2, as.numeric))
colnames(featureAnn) = c('chr','start','end')
featureAnn[] = lapply(featureAnn[], function(x) as.numeric(as.character(x)))

chh = cbind(ch[genomeFeat,], featureAnn)
chh = subset(chh, !is.na(chr))

chh = merge(chh, chr.lengths[,c('chrom','chr.length')], by.y='chrom', by.x='chr')
chh$chr = factor(chh$chr, levels=c(1:22))

chh$ymax = ifelse(chh$coef > 0, max(globalCN.plot$data$mean.value), min(globalCN.plot$data$mean.value))
chh$gl = factor(ifelse(chh$coef > 0, 'gain', 'loss'))

custom.max<-function(cf, hazards, ymax) {
  print(cf)
  if (cf > 0) {
    max = hazards*ymax
    max = ifelse(max > ymax, ymax, max)
  } else {
    max = hazards*ymax
    max = ifelse(max < ymax, ymax, max)
  }
  return(max)
}
chh = chh %>% rowwise() %>% mutate('max'=custom.max(coef, hazards, ymax))

globalCN.plot + geom_rect(data=chh, aes(xmin=start, xmax=end, ymin=0, ymax=max, color=gl), alpha=0, show.legend=F) + scale_color_manual(values=c('goldenrod2','blue3'))
globalBECN.plot + geom_rect(data=chh, aes(xmin=start, xmax=end, ymin=0, ymax=max, color=gl), alpha=0, show.legend=F) + scale_color_manual(values=c('goldenrod2','blue3'))

ch$ymax = ifelse(ch$coef > 0, max(genomePtCN.plot$data$value), min(genomePtCN.plot$data$value))
chh = chh %>% rowwise() %>% mutate('max'=custom.max(coef, hazards, ymax))
genomePtCN.plot + geom_rect(data=chh, aes(xmin=start, xmax=end, ymin=0, ymax=max, color=gl), alpha=0, show.legend=F) + scale_color_manual(values=c('goldenrod2','blue3'))
genomePtBECN.plot + geom_rect(data=chh, aes(xmin=start, xmax=end, ymin=0, ymax=max, color=gl), alpha=0, show.legend=F) + scale_color_manual(values=c('goldenrod2','blue3'))

```

```{r echo=F, fig.height=5, fig.width=6}
globalCNdf$g.start = globalCNdf$start
for (i in 2:nrow(chr.lengths)) {
  rows = which(globalCNdf$chr == chr.lengths[i, 'chrom'])
  globalCNdf[rows, 'g.start'] = globalCNdf[rows, 'start'] + chr.lengths[i-1, 'genome.length']
}

p1 = ggplot(globalCNdf) + geom_tile(aes(x=g.start, y=variable, fill=value, group=Status)) +
  scale_fill_gradient2(high='darkgreen', low='darkred', mid='white') +
  labs(x='', y='Sample') + scale_x_continuous(expand=c(0,0)) +
  #scale_x_continuous(expand=c(0,0), breaks=chr.lengths$genome.length, labels=chr.lengths$chrom) +
  geom_vline(data=chr.lengths, aes(xintercept=genome.length), color='grey') +
  plot.theme + theme(axis.ticks=element_blank(), axis.text=element_blank(),panel.grid.major=element_blank(), legend.position='bottom')

p1leg = get.legend(p1)

p1 = p1 + theme(legend.position = 'none')

p2 = ggplot(globalCNdf, aes(y=variable)) + geom_tile(aes(x=0, fill=Status), show.legend = F) + 
  labs(x=' ', y='') + scale_fill_manual(values=c('darkblue','darkred')) + scale_x_continuous(expand=c(0,0)) +
  plot.theme + theme( axis.line=element_blank(), axis.text=element_blank(), axis.ticks=element_blank(), panel.grid=element_blank())

grid.arrange(p1, p2, p1leg, ncol=2, nrow=2, widths=c(10,1), heights=c(10,1))

p3 = ggplot(chh, aes(x=label,y=coef,fill=gl)) + geom_bar(stat='identity') + labs(x='Effects', y='') +
  coord_flip() + scale_fill_manual(values=c('darkgreen','darkred')) +
  plot.theme + theme(axis.line.x=element_blank(), axis.ticks=element_blank(),axis.text=element_blank(), panel.grid=element_blank(),panel.grid.major=element_blank(), legend.position='none')
p3

#  lay <- rbind(c(1,1,1,4),
#               c(2,2,2,5),
#               c(3,3,3,6),
#               c(7,7,7,8))
# 
# grid.arrange(p3,p1a,p1b,gp,p2a,p2b,p1leg,gp, layout_matrix=lay, widths=4:1, heights=c(2,4,4,1))
```

# Nested leave-one-out predictions

Leave out each patient (both P and NP) and run a new CV fit each time then predict the samples for the patient that was left out.  Fitted model includes HGD samples.

```{r leaveoneout, echo=F, message=F, warning=F}
file = paste(cache.dir, 'loo.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  stop(paste("Missing model file", file))
}

df = as.data.frame(performance.at.1se)
ggplot(df, aes(y=performance.at.1se, x='LOO')) + 
  geom_boxplot(fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) + geom_jitter(color='grey39', width=0.3) +
  geom_label(data=subset(df, performance.at.1se %in% range(performance.at.1se)), aes(y=performance.at.1se, label=round(performance.at.1se, 3)), fontface='bold', fill='lightblue') +
  labs(y='performance at lambda.1se', x='', title=paste('Performance for nested leave one out, at alpha=',select.alpha,sep='')) +
  plot.theme
```

```{r, eval=F}
image(cov(dysplasia.df[names(which(labels == 1)),c(1:(ncol(dysplasia.df)-3))], use='p'), breaks=seq(-1,1,l=21), col=colorRampPalette(c('red','black','green'))(20),  useRaster = T)

image(cor(dysplasia.df[names(which(labels == 0)),c(1:(ncol(dysplasia.df)-3))], use='p'), breaks=seq(-1,1,l=21), col=colorRampPalette(c('red','black','green'))(20),  useRaster = T)

chr <- as.numeric(sub(":.+","",colnames(dysplasia.df[names(which(labels == 0)),c(1:(ncol(dysplasia.df)-3))])))
c <- which(diff(chr)==1)/length(chr)
abline(h=c, v=c, col='blue')

preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Samplename','Prediction')]))
rownames(preds) = preds$Samplename
p <- preds[names(labels), 'Prediction']

d <- dist(dysplasia.df[,c(1:(ncol(dysplasia.df)-3))])
t <- tsne::tsne(d)
plot(t,  bg=colorRampPalette(c('green','grey','red'))(10)[cut(p, breaks = seq(0,1,l=11))], pch=21, col=ifelse(labels==1,"black","white"))
```

### Prediction ROC curves

Based on these curves we get our best AUC at ~0.36.  Currently we are using 0.5 by default.

```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=8, fig.width=4}
roc.plot <- function(roc,title="") {
  df = cbind.data.frame('Specificity'=rev(roc$specificities), 'Sensitivity'=rev(roc$sensitivities))

  ggplot(df, aes(Specificity, Sensitivity)) + geom_line() + scale_x_reverse() + 
    geom_label(data=as.data.frame(t(round(pROC::coords(roc, "best"),2))), 
               aes(x=specificity, y=sensitivity, 
                   label=paste(' (FPR ', sprintf('%.1f',(1-specificity)*100), '%, TPR ', sprintf('%.1f',sensitivity*100),'%)\nAUC:', sprintf("%.2f",roc$auc*100), '%', sep='')), nudge_x=.25) +
  labs(title=title, x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme
}

preds = unlist(lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], function(df) df$Prediction))
np.preds = unlist(lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID], function(df) df$Prediction))

roc50 = pROC::roc(as.integer(c(preds, np.preds) > 0.5), 
                c(rep(1, length(preds)), rep(0, length(np.preds))),
                main='Progressors & Non-progressors, cutoff 0.5')
roc50$model = 'CN Model 0.5'

roc.plot(roc50,'Probability threshold = 0.5')

preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
roc = pROC::roc(Status ~ Prediction, data=preds, ci=T, of='thresholds')
roc$model = 'CN Model'

roc.plot(roc, 'Probability threshold calculated by ROC')

cutoff = round(pROC::coords(roc, 'best')[['threshold']], 2)
```

### p53 ROC curve

```{r p53roc, echo=F, fig.height=5, fig.width=5}
p53samples = subset(patient.info, !is.na(p53.Status), select=c('Samplename','p53.Status','Status'))
p53roc = pROC::roc(as.numeric(p53samples$Status)-1, as.integer(p53samples$p53.Status)-1)

p53roc$model = 'p53 IHC'
roc.plot(p53roc, 'p53 IHC')

pROC::auc(p53roc)

```

### Pathology ROC curve

Based on HGD/IMC/LGD

```{r pathroc, echo=F, fig.height=5, fig.width=5 }
#sum(table(subset(patient.info, Status == 'P')$Pathology)[c('IMC','HGD')])/sum(table(subset(patient.info, Status == 'P')$Pathology))

pathsamples = patient.info[,c('Pathology','Status')]

pathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('LGD','HGD','IMC'), 1, 0) )
pathroc$model = 'Dysplasia (HGD & LGD)'

roc.plot(pathroc, 'Pathology (LGD,HGD,IMC)')

pROC::auc(pathroc)

hgdpathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('HGD','IMC'), 1, 0) )
hgdpathroc$model = 'HGD Only'

roc.plot(hgdpathroc, 'Pathology (HGD,IMC)')

pROC::auc(hgdpathroc)
```

### Compare ROCs

```{r echo=F}

get.roc.stat<-function(roc) {
  cbind.data.frame('AUC'=pROC::auc(roc), 
        'Specificity'=round(pROC::coords(roc, 'best')[['specificity']], 2), 
        'Sensitivity'=round(pROC::coords(roc, 'best')[['sensitivity']], 2),
        'model'=roc$model)
}

m = rbind(get.roc.stat(p53roc),
      get.roc.stat(hgdpathroc),
      get.roc.stat(pathroc),
      get.roc.stat(roc))

m$model = factor(m$model, levels=c('p53 IHC','HGD Only','Dysplasia (HGD & LGD)','CN Model'), ordered = T)

rocList = list(roc,roc50, pathroc,hgdpathroc,p53roc)

aucs = do.call(rbind, lapply(rocList,function(r) cbind.data.frame('AUC'=r$auc*100,'model'=r$model)))

df = do.call(rbind,
             lapply(rocList, function(r) 
              cbind.data.frame('Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),
                               'model'=r$model)))
  ggplot(df, aes(Specificity, Sensitivity,color=model,fill=model)) + geom_line(show.legend = F) + scale_x_reverse() + 
    geom_label(data=m, aes(label=paste(model,'\nAUC:', sprintf("%.2f",AUC*100), '%', sep='')), nudge_x=.25, color='white', show.legend = F) + 
  labs(title='ROC Curves', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme

ggplot(melt(m), aes(model, value*100, group=variable, fill=variable)) + ylim(0,100) +
  geom_col(position="dodge") + 
  geom_text(aes(label=paste(round(value,2)*100,'%',sep='') ), position=position_dodge(width=1), vjust=-0.5) + 
  labs(x='', y='Percent', title='ROC') +
  scale_fill_manual(values=c('dodgerblue3','firebrick4', 'green4'), name='') + plot.theme + 
  theme(legend.position = 'bottom', axis.text.x=element_text(angle=45,hjust=1))

```

***

__All predictions from here onwards are performed using the `r cutoff` cutoff__

***

## How do the predictions look at each time point?

```{r pathology, echo=F, message=F, warning=F, fig.height=8, fig.width=6}
model.pred<-function(pt, p53=F, pred = function(rows) length(which(rows$Prediction > cutoff))) {
  final.endo = subset(pt, Endoscopy.Year == max(Endoscopy.Year))
  final.endo = final.endo[nrow(final.endo),]

  h = subset(pt, Pathology %in% c('IMC','HGD'))
  l = subset(pt, Pathology == 'LGD' & PID != final.endo$PID)
  id = subset(pt, Pathology == 'ID' & PID != final.endo$PID)
  be = subset(pt, Pathology %in% c('BE') & PID != final.endo$PID)

  if (p53) {
    return(cbind('HGD'=pred(h), 'total.h'=nrow(h),#   nrow(subset(h, !is.na(p53.Status))), 
                  'LGD'=pred(l), 'total.l'=nrow(subset(l, !is.na(p53.Status))), 
                  'ID'=pred(id), 'total.id'=nrow(subset(id, !is.na(p53.Status))), 
                  'NDBE'=pred(be), 'total.be'=nrow(subset(be, !is.na(p53.Status)))))

  } else {
    return( cbind('HGD'=pred(h), 'total.h'=nrow(h), 
                  'LGD'=pred(l), 'total.l'=nrow(l), 
                  'ID'=pred(id), 'total.id'=nrow(id),
                  'NDBE'=pred(be), 'total.be'=nrow(be)) )
    }
}

collapse.pred.by.path<-function(pts, p53=F, predFUN = NULL) {
  if (!is.null(predFUN))
    by.path = do.call(rbind.data.frame, lapply(pts,model.pred,p53, predFUN) )
  else
    by.path = do.call(rbind.data.frame, lapply(pts,model.pred,p53) )

  sums.by.path = colSums(by.path)

  path = as.data.frame(matrix(sums.by.path, byrow=T, nrow=4, dimnames=list(c('HGD','LGD','ID','NDBE'),c('pred', 'total'))))
  path$prog.ratio = (path$pred/path$total)*100
  path$nonprog.ratio = 100-path$prog.ratio
  path$path = rownames(path)

  path$path = factor(path$path, levels=c('HGD','LGD','ID','NDBE')) 
  path$pred = ifelse(p53, 'p53','model')
  return(path)
}

global.path = collapse.pred.by.path(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID])
global.pathNP = collapse.pred.by.path(pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID])

predP53 = function(rows) length(which(rows$p53.Status == 1 ))
p53byPath = collapse.pred.by.path(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], T, predP53)

global.melt = melt(global.path, measure.vars=c('prog.ratio', 'nonprog.ratio'), id.vars=c('path', 'total','pred'))
p53.melt = melt(p53byPath, measure.vars=c('prog.ratio', 'nonprog.ratio'), id.vars=c('path', 'total','pred'))
np.melt = melt(global.pathNP, measure.vars=c('prog.ratio', 'nonprog.ratio'),id.vars=c('path','total','pred'))

global.melt$pred='model'
global.melt$status='Progressor'
np.melt$status='Non-Progressor'

#p53.melt$status='p53 IHC'
p53.melt = subset(p53.melt, variable == 'prog.ratio')
p53.melt$variable = paste('p53',p53.melt$variable,sep='')
p53.melt$status = 'Progressor'

m = rbind(global.melt, np.melt, p53.melt)

#m$status = factor(m$status, levels=c('Progressor', 'p53 IHC','Non-Progressor'), ordered = T)

m$pred = factor(m$pred)
m$final = ifelse(m$path == 'HGD', T, F)

ggplot(m, aes(path, value, group=path, fill=variable)) + ylim(0,100) + facet_wrap(~status, ncol=1) +
  scale_fill_manual(values=c('firebrick4','green4', 'slateblue4'), name='', labels=c('Pred. progressed', 'Pred. Not progressed', 'p53 prediction'))  +
  geom_bar(stat='identity', position=position_stack()) +
  geom_text(aes(hjust=ifelse(value > 12, 2, 1),
                 label=paste(paste(round(value, 1),'%',sep=''),
                                        paste('n=', total, sep=''),sep='\n')), 
            position=position_stack(),color='white',  show.legend=F) +
  geom_bar(data=p53.melt, aes(path, value, group=path), fill='slateblue4',stat='identity',position = position_dodge()) +
  geom_text(data=p53.melt, aes(path,value,hjust=ifelse(value > 12, 2, 1),label=paste(paste(round(value, 1),'%',sep=''),paste('n=', total, sep=''),sep='\n')), position=position_dodge(),color='white',  show.legend=F) +
    labs(title='Samples predicted by pathology', x='Pathology', y='Percent', subtitle='Non-HGD exclude the final endoscopy, p53 excludes NA samples') +
  plot.theme + theme(text = element_text(size=16, face='bold'), legend.position = 'bottom') + coord_flip()

ggplot(m, aes(path, value, group=path, fill=variable)) + ylim(0,100) + facet_wrap(~status) +
  geom_bar(stat='identity', position=position_stack()) + 
  scale_fill_manual(values=c('firebrick4','green4','slateblue4'), name='', labels=c('Pred. progressed', 'Pred. Not progressed','p53 IHC'))  +
  geom_text(aes(vjust=ifelse(value > 12, 2, 1),label=paste(paste(round(value, 1),'%',sep=''),paste('n=', total, sep=''),sep='\n')),
            position=position_stack(),color='white',show.legend=F) +
  geom_bar(data=p53.melt, aes(path, value, group=path), fill='slateblue4',stat='identity',position = position_dodge()) +
  geom_text(data=p53.melt, aes(path,value,vjust=ifelse(value > 12, 2, 1),label=paste(paste(round(value, 1),'%',sep=''),paste('n=', total, sep=''),sep='\n')), position=position_dodge(),color='white',  show.legend=F) +
  labs(title='Samples predicted by pathology', x='Pathology', y='Percent', subtitle='Non-HGD exclude the final endoscopy, p53 excludes NA samples') +
  plot.theme + theme(text = element_text(size=16, face='bold'), legend.position = 'bottom')

```

```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=6}
# All samples
all.predictions = sapply(pg.samp, function(pt) {
  if (unique(pt$Status) == 'NP') {
    as.integer(pt$Prediction < cutoff)
  } else {
    as.integer(pt$Prediction > cutoff)
  }
})

all.predictions = do.call(rbind, lapply(all.predictions, function(x) {
  cbind.data.frame('correctly.predicted'=sum(x), 'total.samples'=length(x))
}))

all.predictions$Status = ifelse (rownames(all.predictions) %in% subset(sum.patient.data, Status == 'P')$Hospital.Research.ID, 'P', 'NP')

rownames(all.predictions) = subset(sum.patient.data, Hospital.Research.ID %in% rownames(all.predictions))$Patient

all.predictions$pred.ratio = with(all.predictions, correctly.predicted/total.samples)
all.predictions$Patient = rownames(all.predictions)
all.predictions = transform(all.predictions, Patient=reorder(Patient, pred.ratio) )

p = subset(all.predictions, Status == 'P')
np = subset(all.predictions, Status == 'NP')

m = melt(all.predictions, id.vars=c('Patient','Status'), measure.vars=c('pred.ratio'))

mp = m %>% group_by(Status) %>% summarise('all'=round(length(which(value==1))/length(Status), 2), 'all.c'=length(Status)-length(which(value==1))/2, 
                                          'some'=round(length(which(value > 0 & value < 1))/length(Status), 2), 'some.c'=length(which(value > 0 & value < 1)),
                                          'none'=round(length(which(value == 0))/length(Status), 2), 'none.c'=1.5)

mpT = m %>% group_by(Status) %>% summarise('100%'=round(length(which(value==1))/length(Status), 2), 
                                          '>25%'=round(length(which(value >= 0.25 & value < 1 ))/length(Status), 2), 
                                          '0%'=round(length(which(value == 0 ))/length(Status), 2))
mpT$Status = c('Non-Progressor','Progressor')

grid.arrange(
  ggplot(m, aes(Patient, value, group=Status)) + facet_grid(~Status,  scales='free_x') + 
    geom_bar(stat='identity', position='identity',aes(fill=Status)) + 
    scale_fill_manual(values=c('darkblue','darkgreen')) +
    geom_label(data=mp, aes(x=all.c, y=0.5, label=paste(all*100,'%',sep=''), fill=Status), color='white', fontface='bold', size=6) +
    geom_label(data=mp, aes(x=some.c, y=0.35, label=paste(some*100,'%',sep=''), fill=Status), color='white', fontface='bold', size=6) +
    geom_label(data=mp, aes(x=none.c, y=0.2, label=paste(none*100,'%',sep=''), fill=Status), color='white', fontface='bold', size=6) +
    labs(y='Ratio of predicted samples') + plot.theme + theme(legend.position='none', axis.text.x=element_text(angle=90) ), 
  tableGrob(mpT, rows=NULL), nrow=2, as.table=T, heights=c(4,1) )

```

## Backwards from HGD

Clinicians always ask to see the timepoints peeled backwards to see how early prediction is possible.

```{r, warning=F, message=F, echo=F, fig.height=8, fig.width=6}
transform.pt.by.time<-function(pt) {
  pt = pt %>% group_by(PID, Endoscopy.Year) %>% summarise(
                  'pred'=sum(as.integer(Prediction > cutoff)), 'samples'=length(PID), 
                  #'final.endo'= length(which(grepl('HGD|IMC', Pathology))) > 0,
                  'months.before.final'= (max(pt$Endoscopy.Year) - unique(Endoscopy.Year))*12, 
                  'max.path'=max(Pathology))
  arrange(pt, -Endoscopy.Year, PID)
}

back = lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], transform.pt.by.time)

endos = do.call(rbind, lapply(1:max(sapply(back,nrow)), function(i) {
  df = do.call(rbind, lapply(back, function(x) {
    if (i > nrow(x)) return(NA)
    cbind(x[i, c('Endoscopy.Year','pred','samples', 'months.before.final', 'max.path')], i)
  }))
  rownames(df) = subset(sum.patient.data, Hospital.Research.ID %in% rownames(df))$Patient

  df$Patient = rownames(df)
  return(df)
}))

endos = endos[complete.cases(endos),]
endos$i = endos$i-1

endos$pred.ratio = round(endos$pred/endos$samples, 2)
endos = transform(endos, Patient=reorder(Patient, -months.before.final) )

endos$brks = cut(endos$months.before.final, include.lowest = T, ordered_result = T,
    breaks=c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12))
endos$brks = with(endos, factor(brks, rev(levels(brks)))) # invert order

endos$labels = gsub('\\[|\\(|\\]', '',endos$brks)
endos$labels = sapply(strsplit( endos$labels, ',' ), function(x) paste(sort(as.numeric(x), decreasing=T), collapse='-' ))
endos$labels[ grep('-120', endos$labels) ] = '>120'
endos$labels[ grep('1-0', endos$labels) ] = 'Endpoint'

## Flip order, 0-1 comes last
m = endos %>% group_by(brks, labels) %>% summarise( 'pred.r'= length(which(pred > 0))/length(brks), 'n'=length((Patient)) )

m$highlight = F
m[4:6,'highlight'] = T
m$final = ifelse(m$brks == '[0,1]', T, F)

ggplot(m, aes(brks, pred.r*100, group=brks)) + ylim(0,105) +
  geom_bar(stat='identity', fill='dodgerblue3', show.legend = F) +     
  scale_fill_manual(values=c('dodgerblue3', 'yellowgreen')) +
  geom_text(aes(y=pred.r*100+1.5, label=paste(round(pred.r*100), '%', sep=''))) +
  geom_text(aes(y=pred.r*100-2, label=paste('n=',n, sep='')), color='white') + 
  scale_x_discrete(labels=m$labels) + 
  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = number of endoscopies') +
  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1))

plot.endo<-function(df, minM=NULL, maxM=NULL, bin=F) {
  if (is.null(minM)) minM = min(df$months.before.final)
  if (is.null(maxM)) maxM = max(df$months.before.final)

  colors = c('green4','red4')

  df = arrange(df %>% group_by(Patient) %>% mutate('overall.pred'=sum(as.integer(pred > 0))/length(Patient)), Patient)

  df = transform(df, Patient=reorder(Patient, -overall.pred) )

  if (bin) {
    df$endo.pred = ifelse(df$pred > 0, 'Pos','Neg')
    ggplot(df, aes(months.before.final, endo.pred, group=1, color=endo.pred)) + facet_grid(Patient ~.) + 
      scale_x_continuous(breaks=seq(minM, maxM, by = 12)) +
      geom_line( color='grey39') + geom_point(aes(color=endo.pred)) +
      geom_label_repel(aes(label=max.path, fill=endo.pred), color='white', fontface='bold', size=4, label.size=0, label.r=unit(0.25, 'lines'), label.padding=unit(0.15, 'lines'), show.legend=F) + 
      scale_fill_manual(values=colors) + 
      scale_color_manual(values=colors, name='Endoscopy Prediction') +
      labs(x='Months before diagnosis', y='Predictive Endoscopy') + plot.theme + 
      theme(panel.border = element_rect(color = "grey50", fill = NA, size = 1), legend.position='bottom', strip.background =element_rect(fill="lightcyan2"), panel.grid.major.x=element_blank(), axis.text.x=element_text(angle=45, hjust=1))

  } else {
  ggplot(df, aes(months.before.final, pred.ratio, group=1, color=pred.ratio, fill=pred.ratio)) + facet_grid(Patient ~.) +
      ylim(0,1) + geom_abline(slope=0, intercept=seq(.25,.75,.25), color='grey', linetype='dashed') +
      geom_col(position='dodge', aes(y=pred.ratio), color='white') +# geom_line() +
      scale_x_continuous(breaks=seq(minM, maxM, by = 12)) + 
      geom_label(aes(y=ifelse(pred.ratio==1, 0.75, pred.ratio), label=max.path, fill=pred.ratio), color='white', fontface='bold', size=4, label.size=0, label.r=unit(0.25, 'lines'), label.padding=unit(0.15, 'lines'), nudge_y=0.1) +
      labs(x='Months before diagnosis', y='Sample prediction ratio') + 
      scale_fill_continuous(high='orangered4', low='green4', name='Sample Prediction Ratio') + plot.theme + 
      theme(panel.border = element_rect(color = "grey50", fill = NA, size = 1), legend.position='bottom',strip.background =element_rect(fill="lightcyan2"), panel.grid.major.x=element_blank(), axis.text.x=element_text(angle=45, hjust=1))
  }
}

endos$months.before.final = endos$months.before.final*-1
me = endos %>% group_by(Patient) %>% summarise('max.endo' = max(i), 'months' = min(months.before.final))
```

## Progressor predictions per patient
```{r, echo=F, fig.height=12, fig.width=10}
plot.endo( subset(endos, Patient %in% subset(me, months <= -72)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(72/12, "years or more follow-up"))
```

```{r, echo=F, fig.height=12, fig.width=10}
# plot.endo( subset(endos, Patient %in% subset(me, months <= -48 & months > -72)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(48/12, "-",72/12,  "years or more follow-up"))
# plot.endo( subset(endos, Patient %in% subset(me, months <= -24 & months > -48)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(24/12, "-",48/12,  "years or more follow-up"))
plot.endo( subset(endos, Patient %in% subset(me, months > -48)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(1, "-",48/12,  "years or more follow-up"))
```

## Non-progressor predictions per patient
```{r, echo=F, fig.height=24, fig.width=16}
nonp.back = lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID],transform.pt.by.time)

nonp.endos = do.call(rbind, lapply(1:max(sapply(nonp.back,nrow)), function(i) {
  df = do.call(rbind, lapply(nonp.back, function(x) {
    if (i > nrow(x)) return(NA)
    cbind(x[i, c('Endoscopy.Year','pred','samples', 'months.before.final', 'max.path')], i)
  }))
  rownames(df) = subset(sum.patient.data, Hospital.Research.ID %in% rownames(df))$Patient

  df$Patient = rownames(df)

  return(df)
}))
nonp.endos = nonp.endos[complete.cases(nonp.endos),]
#nonp.endos$i = nonp.endos$i-1

nonp.endos$pred.ratio = round(nonp.endos$pred/nonp.endos$samples, 2)
nonp.endos = transform(nonp.endos, Patient=reorder(Patient, -months.before.final) ) 
nonp.endos$months.before.final = nonp.endos$months.before.final*-1

me = nonp.endos %>% group_by(Patient) %>% summarise('max.endo' = max(i), 'months' = min(months.before.final))

plot.endo(subset(nonp.endos, Patient %in% subset(nonp.endos, pred.ratio > 0)$Patient), bin=T) + labs(title='Non-progressors with progressive predictions')
```

## Predictions laid out spatially

```{r spatial, echo=F}

# ids = subset(sum.patient.data, Status == 'P' & total.endos > 5 & M > mean(M, na.rm=T))$Hospital.Research.ID
# missing_p53 = do.call(rbind.data.frame, lapply(pg.samp, function(df) {
#   x = df[which(is.na(df$p53.Status)), c('Hospital.Research.ID', 'Path.ID', 'Status')]
#   if (nrow(x) > 0) x$Prioritize = ifelse(unique(df$Hospital.Research.ID) %in% ids, '*', '')
#   return(x)
# }))
# write.table(missing_p53, sep='\t', quote=F, row.names = F, col.names = T, file='missing_p53.txt')

#ids = subset(sum.patient.data, ((Status == 'P' & total.endos > 5) | (Status == 'NP' & total.endos > 9)) & M > mean(M, na.rm=T))$Hospital.Research.ID

transform.by.time<-function(pt) {
  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction')] %>% group_by(PID, Samplename) %>% mutate(
  'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12,
  'nl' = str_extract(Path.ID, "B\\w?[0-9]+"))

  pt$nl = (as.numeric(factor(pt$nl, levels=(unique(sort(pt$nl))), ordered = T))-1)*2
  #table(pt[,c('nl','months.before.final')])

  pt$brks = cut(pt$months.before.final, include.lowest = T, ordered_result = T,
    breaks=c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12),
    labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))

  tb = table(pt[,c('brks', 'PID')])
  for (i in 1:nrow(tb)) {
    pids = tb[i,]
    yrs = which(pids > 0)
    if (length(yrs) > 1) {
    for (j in 2:length(yrs)) 
      pt[which(pt$PID == names(yrs)[j]), 'nl'] = pt[which(pt$PID == names(yrs)[j]), 'nl'] + 1
    }
  }
  arrange(pt, -Endoscopy.Year, PID)
}

back = do.call(rbind.data.frame,lapply(pg.samp, transform.by.time))
```

```{r fig.width=8, fig.height=8}
probs = seq(0,1,1/4)
back$quants = with(back, cut(Prediction, breaks=quantile(Prediction, probs), include.lowest = T, labels=probs[-1]))

qt = table(back$quants, back$Status)

pq = cbind.data.frame('perc'=qt[,'P']/rowSums(qt),
      'quant'=quantile( back$Prediction, probs)[-1])
ggplot(back, aes(Prediction, fill=..x..)) + geom_histogram(bins=20, show.legend = F) + 
  scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
  plot.theme + labs(title='Sample predictions', y='n Samples', x='Probability') +
  geom_vline(xintercept=quantile( back$Prediction, probs), linetype='longdash') +
  geom_label(data=pq, aes(x=quant, y=75, label=round(perc,2) ))

```

```{r, fig.width=5, fig.height=10 }

back$brks = with(back, factor(brks, rev(levels(brks)))) # invert order

ids = c(13,20,34,56,60,48)
p = ggplot(subset(back, Patient %in% ids & Status == 'P'), aes(brks, nl, fill=Prediction)) + 
      geom_tile() + geom_text(aes(label=Pathology), color='white', show.legend = F) + 
      scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
      facet_wrap(~Patient, ncol=1, scales='free_y', labeller=label_both) +
      labs(y='Oesophageal Location',x='Months Prior to Endpoint', title='Progressors') +
      plot.theme + theme(axis.text.x=element_text(angle=45, hjust=1)) 
p
#ggsave('pg_heatmap1.png', plot = p,
#  scale = 1, width = 5, height = 12, units = "in", limitsize = F,
#  dpi = 300)

ids = c(15,16,21,3,45,19)
p=ggplot(subset(back, Patient %in% ids & Status == 'NP'), aes(brks, nl, fill=Prediction)) + 
  geom_tile() + geom_text(aes(label=Pathology), color='white', show.legend = F) + 
  scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
  facet_wrap(~Patient, ncol=1, scales='free_y', labeller=label_both) +
  labs(y='Oesophageal Location',x='Months Prior to Endpoint', title='Non-Progressors') +
  plot.theme + theme(axis.text.x=element_text(angle=45, hjust=1)) 
#ggsave('npg_heatmap1.png', plot = p,
#  scale = 1, width = 5, height = 12, units = "in", limitsize = F,
#  dpi = 300)
p
```

```{r, echo=F}
# 
#  ggplot(subset(back, Status == 'NP' & Patient %in% ids), aes(brks, nl, fill=Prediction, group=Patient)) + 
#   facet_wrap(~Patient, scales='free', labeller=label_both, ncol=1) +
#   geom_tile() + geom_text(aes(label=Pathology), color='white', show.legend = F) + 
#   scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
#   labs(y='Oesophageal Location',x='Months Prior to Endpoint', title='Non-Progressors') +
#   plot.theme + theme(axis.text.x=element_text(angle=45, hjust=1), strip.text=element_text(size=8), strip.background = element_rect(fill='gray88'))

p = ggplot(subset(back, Status == 'NP'), aes(brks, nl, fill=Prediction, group=Patient)) + 
  facet_wrap(~Patient, scales='free_y', labeller=label_both, ncol=1) +
  geom_tile() + geom_text(aes(label=Pathology), color='white', show.legend = F) + 
  scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
  labs(y='Oesophageal Location',x='Months Prior to Endpoint', title='Non-Progressors') +
  plot.theme + theme(axis.text.x=element_text(angle=45, hjust=1), strip.text=element_text(size=8), strip.background = element_rect(fill='gray88'))
ggsave('npg_heatmap.png', plot = p, scale = 1, width = 8, height = 52, units = "in", limitsize = F, dpi = 300)

p = ggplot(subset(back, Status == 'P'), aes(brks, nl, fill=Prediction, group=Patient)) + 
  facet_wrap(~Patient, scales='free_y', labeller=label_both, ncol=1) +
  geom_tile() + geom_text(aes(label=Pathology), color='white', show.legend = F) + 
  scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
  labs(y='Oesophageal Location',x='Months Prior to Endpoint', title='Progressors') +
  plot.theme + theme(axis.text.x=element_text(angle=45, hjust=1), strip.text=element_text(size=10), strip.background = element_rect(fill='gray88'))
ggsave('pg_heatmap.png', plot = p, scale = 1, width = 8, height = 52, units = "in", limitsize = F, dpi = 300)

```

# Chr17

## p arm
```{r}
p11 = subset(globalCNdf, chr == 17 & start < subset(chr.lengths, chrom == 17)$chr.cent)

p11Loss = subset(p11, value < -1)
length(unique(p11Loss$variable))

table(subset(patient.info, Samplename %in% p11Loss$variable)$Pathology)
(18+12)/116

# Samples with a loss vs all samples
length(unique(p11Loss$variable))/length(unique(globalCNdf$variable))

p11Loss = unique(p11Loss[,c('variable', 'Status','Hospital.Research.ID', 'Patient', 'Pathology')])

# Of those where we see loss of p arm, p53 IHC shows an "abnormal" result 33% of the time
table(subset(patient.info, Samplename %in% p11Loss$variable, select='p53.Status'))

pander(table(subset(patient.info, Samplename %in% p11Loss$variable, select=c('p53.Status', 'Pathology'))))

```

## Chr 17 Predictive

One way of looking at the p53 prediction (e.g. Adam Bass) is to see how predictive the p-arm loss in chr17 is.
```{r, echo=F, warning=F, message=F, fig.height=16, fig.width=18, eval=F}
ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
ccgenes = ccgenes[-grep('X,Y', ccgenes$Genome.Location),]
ccgenes = ccgenes[-which(ccgenes$Mutation.Types == 'T'),] # drop the translocations

genes = subset(ccgenes, grepl('TSG|oncogene', Role.in.Cancer) & !grepl('^X|Y', Chr.Band), select=c('Gene.Symbol','Genome.Location'))

loc = unlist(strsplit(genes$Genome.Location, ':|-'))
loc = do.call(rbind.data.frame, strsplit(genes$Genome.Location, ':|-'))
colnames(loc) = c('chr','start','end')
loc[] = lapply(loc[], function(x) as.numeric(as.character(x)))
genes = cbind(genes, loc, 'value'=-7, 'Row.names'='', 'Patient'='', 'selected'=as.factor(0))

plot.cn.chr<-function(cn, chrom='17', info=NULL, haz=NULL, genes=NULL) {
    p = ggplot(subset(cn, chr == chrom), aes(x=chr.length)) + facet_grid(Status~chr, scales='free_x', labeller=labeller(chr=label_both)) + 
          geom_rect( aes(xmin=start, xmax=end, ymin=0, ymax=value, fill=cn), alpha=0.5, show.legend = T) + 
          scale_color_manual(values=c('firebrick3','grey88','green4')) +
          scale_fill_manual(values=c('firebrick3','grey88','green4'))

    if (!is.null(info)) {
      info = info[which(info$chrom == chrom),]
      p = p + geom_vline(data=info, aes(xintercept=chr.cent), color='grey39', linetype='longdash' ) +
              geom_label(data=info, aes(x=chr.cent-cent.gap*3, y=10), label='p-arm')
    }

    if (!is.null(genes)) {
      genes = subset(genes, chr == chrom)
      if (nrow(genes) > 0) p = p + geom_point(data=genes, aes(x=start,  y=value), color='red', size=3) +
              geom_text_repel(data=genes, aes(x=start, y=value, label=Gene.Symbol), color='red') 
    }
    if (!is.null(haz)) {
      haz = subset(haz, chr == chrom)
      if (nrow(haz) > 0) p = p + geom_rect(data=haz, aes(xmin=start, xmax=end, ymin=0, ymax=max, color=gl), alpha=0, show.legend=F, size=1) + scale_color_manual(values=c('goldenrod2','blue3')) 
    }

    p + theme(text=element_text(face='bold', size=14), axis.text.x=element_blank(), legend.position='bottom', 
          panel.grid = element_blank(), panel.background = element_blank(),strip.background =element_rect(fill="lightcyan2"),  
          panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0.1, 'lines') ) + labs(x='Chromosomes', y='Mean adjusted segmentation value')

}

plot.cn.chr(cn=subset(globalCNdf, Pathology == 'BE'), chrom='17', chr.lengths, genes=genes, haz=ch) + labs(
title="Segment Values Across BE Samples", subtitle=paste('NonP:Prog - ', paste(table(subset(patient.info, Pathology == 'BE')$Status), collapse=':')))

#do.call(grid.arrange, lapply( unique(ch$chr), function(chr) {
#  plot.cn.chr(cn=subset(globalCNdf, Pathology == 'BE'), chrom=chr, chr.lengths, haz=ch, genes=genes) 
#}))

# plot.cn.chr(cn=subset(globalCNdf, Pathology == 'BE'), chrom='12', chr.lengths, haz=ch, genes=genes)
# 
plot.cn.chr(cn=subset(globalCNdf, Pathology == 'BE'), chrom='9', chr.lengths, genes=genes, haz=ch)
plot.cn.chr(cn=subset(globalCNdf, Pathology == 'BE'), chrom='16', chr.lengths, haz=ch, genes=genes)

genesGR = makeGRangesFromDataFrame(genes, keep.extra.columns = T)
chGR = makeGRangesFromDataFrame(ch, keep.extra.columns = T)

findOverlaps(chGR, genesGR)

```

```{r, echo=F, message=F, warning=F, fig.height=5, fig.width=8, eval=F}
file = paste(cache.dir, 'chr17.predictive.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  stop(paste("File missing", file))
}

names(performance.at.1se)

mp = model.performance(performance.at.1se, coefs) 
mp$plot + labs(x='Models', title='Model performance for subsets of genomic regions') + 
  scale_colour_manual(values=c('grey39', 'grey39')) + theme(axis.text.x = element_text(angle=45,hjust=1)) +
  geom_hline(yintercept=mean(mp$performance$mean)+sd(mp$performance$mean), color='red', alpha=0.5) +
  geom_hline(yintercept=mean(mp$performance$mean)-sd(mp$performance$mean), color='red', alpha=0.5)

mp$performance

chr.info = read.table('hg19_genes.txt', sep='\t', header=T)
chr.info = chr.info[1:22,]

chr.info$info.content = chr.info$protein_coding/chr.info$length

# No correlation between information content (genes/length) and classification rate 
cor.test(mp$performance[1:22,'mean'], chr.info$info.content)

median(mp$performance$mean)
sd(mp$performance$mean)
mp$performance[17,'mean']

```

Individually none of these show very good classification rates, all hovering around 65%. However, more importantly is that Chr17 is only marginally better than any of the others (unless you exclude HGD) and all are below the 72% classification rate we get with all of the data

# Tumour Supressors

```{r ts, eval=F}

colnames(ccgenes)

tsgs = subset(ccgenes, grepl('TSG|oncogene', Role.in.Cancer))
tsgs = tsgs[grep('^X|Y',tsgs$Chr.Band, invert=T),]
nrow(tsgs)

loc = do.call(rbind.data.frame, strsplit(tsgs$Genome.Location, ':|-'))
colnames(loc) = c('chr','start','end')
loc[] = lapply(loc[], function(x) as.numeric(as.character(x)))

tsgs = cbind(tsgs, loc)
tsgs = makeGRangesFromDataFrame(tsgs[,c(colnames(loc), 'Gene.Symbol','Role.in.Cancer', 'Mutation.Types')], keep.extra.columns = T)
tsgs = sort(tsgs)

length(tsgs)

# average 1 per 9MB
avgTSGdist = mean(elementMetadata(distanceToNearest(tsgs))$distance)

ch = arrange(ch, -hazards)
features = makeGRangesFromDataFrame(ch, keep.extra.columns = T)

distanceTSG = data.frame(matrix(nrow=0,ncol=3))
for (i in 1:length(features)) {
  f = features[i]
  hits = nearest(f, tsgs, ignore.strand=T, select='all')
  if (length(hits) > 0) {
    distanceTSG = rbind(distanceTSG, cbind.data.frame(
      'features'=i,
      'nearestTSG'=elementMetadata(tsgs[subjectHits(hits)])$Gene.Symbol,
      'distanceToTSG'=elementMetadata(distanceToNearest(f, tsgs, ignore.strand=T, select='all'))$distance)
    )
  }
}

features[unique(subset(distanceTSG, distanceToTSG == 0)$features)]

# In our dataset we should see 1 TSG/onc per 10MB
# But this is double the actual number we see
ncol(dysplasia.df)/round(avgTSGdist/5e6)

length(which(round((subset(distanceTSG, distanceToTSG > 0)$distanceToTSG/5e6)-0.5) == round((avgTSGdist/5e6)-0.5)))

```

# Validation Dataset

20 Patients (10P, 10NP) set aside for validation purposes
`r pander(table(sum.validation.info$Status), caption='Validation cohort')`

```{r validation, message=F, warning=F, echo=F, fig.height=8, fig.width=8}

pander(table(validation.patient.info[c('Status','p53.Status')]))

validation.df = t(mergedDfValidation[, grep('^D', colnames(mergedDfValidation))])
validation.df = apply(validation.df, 2, unit.var)
validation.df = score.cx(validation.patient.data, validation.df)

if (includeDemo) {
  ## Add in demographics
  validation.df = add.demo.tocv(validation.patient.info, validation.df)
}

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
message(paste("loading", file))
load(file, verbose=T)

fitV = models[[select.alpha]]
lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']

pred = predict(fitV, newx=validation.df, s=lambda.opt, type='response')

vpd = lapply(validation.patient.data, function(pt) {
  info = pt$info
  info$PID = unlist(lapply(info$Path.ID, function(x) unlist(strsplit(x, 'B'))[1]))
  return(info)
})

samples = intersect(rownames(pred), validation.patient.info$Samplename)
validation.patient.info$Prediciton = NA
validation.patient.info[which(validation.patient.info$Samplename %in% samples), 'Prediction'] = pred[samples,]

for (pt in names(vpd)) 
  vpd[[pt]]$Prediction[which(vpd[[pt]]$Samplename %in% rownames(pred))] = pred[ intersect(rownames(pred), vpd[[pt]]$Samplename), ]

v.predictions = sapply(vpd, function(pt) {
  if (unique(pt$Status) == 'NP') {
    as.integer(pt$Prediction < cutoff)
  } else {
    as.integer(pt$Prediction > cutoff)
  }
})

names(v.predictions) = subset(sum.validation.info, Hospital.Research.ID %in% names(v.predictions))$Patient
v.predictions = do.call(rbind, lapply(v.predictions, function(x) {
  cbind.data.frame('correctly.predicted'=sum(x,na.rm=T), 'total.samples'=length(x))
}))

v.predictions$Status = ifelse (rownames(v.predictions) %in% subset(sum.validation.info, Status == 'P')$Patient, 'P', 'NP')
v.predictions$pred.ratio = with(v.predictions, correctly.predicted/total.samples)
v.predictions$Patient = rownames(v.predictions)
v.predictions = transform(v.predictions, Patient=reorder(Patient, pred.ratio) )

p = arrange(subset(v.predictions, Status == 'P'), -pred.ratio)
np = arrange(subset(v.predictions, Status == 'NP'), -pred.ratio)

m = melt(v.predictions, id.vars=c('Patient','Status'), measure.vars=c('pred.ratio'))

mp = m %>% group_by(Status) %>% summarise('all'=round(length(which(value==1))/length(Status), 2), 'all.c'=length(Status)-length(which(value==1))/2, 
                                          'some'=round(length(which(value > 0 & value < 1))/length(Status), 2), 'some.c'=length(which(value > 0 & value < 1)),
                                          'none'=round(length(which(value == 0))/length(Status), 2), 'none.c'=1.5)

mpT = m %>% group_by(Status) %>% summarise('100%'=round(length(which(value==1))/length(Status), 2), 
                                          '>50%'=round(length(which(value >= 0.5 & value < 1 ))/length(Status), 2), 
                                          '25-50%'=round(length(which(value >= 0.25 & value < 0.5 ))/length(Status), 2), 
                                          '0%'=round(length(which(value == 0 ))/length(Status), 2))
mpT$Status = c('Non-Progressor','Progressor')

ggplot(m, aes(Patient, value, group=Status)) + facet_grid(~Status,  scales='free_x') + 
  geom_bar(stat='identity', position='identity',aes(fill=Status)) + 
  scale_fill_manual(values=c('darkblue','darkgreen')) +
  geom_label(data=mp, aes(x=all.c, y=0.5, label=paste(all*100,'%',sep=''), fill=Status), color='white', fontface='bold', size=6) +
  geom_label(data=mp, aes(x=some.c, y=0.35, label=paste(some*100,'%',sep=''), fill=Status), color='white', fontface='bold', size=6) +
  geom_label(data=mp, aes(x=none.c, y=0.2, label=paste(none*100,'%',sep=''), fill=Status), color='white', fontface='bold', size=6) +
  labs(y='Ratio of predicted samples', title='Model predictions on validation dataset') + plot.theme + theme(legend.position='none', axis.text.x=element_text(angle=90) )
```

```{r, echo=F, message=F, warning=F, fig.width=8, fig.height=12}

back = lapply(vpd, transform.pt.by.time)
vpi = do.call(rbind, lapply(1:max(sapply(back,nrow)), function(i) {
  df = do.call(rbind, lapply(back, function(x) {
    if (i > nrow(x)) return(NA)
    cbind(x[i, c('Endoscopy.Year','pred','samples', 'months.before.final', 'max.path')], i)
  }))
  rownames(df) = subset(sum.validation.info, Hospital.Research.ID %in% rownames(df))$Patient
  df$Patient = rownames(df)
  return(df)
}))

vpi = vpi[complete.cases(vpi),]
vpi$pred.ratio = round(vpi$pred/vpi$samples, 2)
vpi = transform(vpi, Patient=reorder(Patient, -months.before.final) ) 
vpi$months.before.final = vpi$months.before.final*-1

me = vpi %>% group_by(Patient) %>% summarise('max.endo' = max(i), 'months' = min(months.before.final))

plot.endo(subset(vpi, Patient %in% subset(sum.validation.info, Status == 'P')$Patient), bin=T) + labs(title="Validation - Progressors")

plot.endo(subset(vpi, Patient %in% subset(sum.validation.info, Status == 'NP')$Patient), bin=T) + labs(title="Validation - Non Progressors")
```

```{r echo=F, fig.height=12,fig.width=5}

vpd2 = lapply(vpd, function(df) {
  transform.by.time(df)
})

vpd2 = do.call(rbind.data.frame, vpd2)

vpd2[which(vpd2$Patient == 52),'nl'] = c(4,2,0)

vpd2[which(vpd2$Patient %in% c(1, 88)),'nl'] = 0
vpd2[which(vpd2$Patient %in% c(40,77) & is.na(vpd2$nl)),'nl'] = 0
vpd2[which(vpd2$Patient == 75 & is.na(vpd2$nl)), 'nl']  = c(0,2,4,6,0,0)

vpd2$Endoscopy.Year = factor(vpd2$Endoscopy.Year, ordered = T)
vpd2$brks = with(vpd2, factor(brks, rev(levels(brks)))) # invert order

ggplot(subset(vpd2, Status == 'P'), aes(brks, nl, fill=Prediction)) + 
  geom_tile() + geom_text(aes(label=Pathology), color='white', show.legend = F) + 
  scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
  facet_wrap(~Patient, ncol=1, scales='free_y', labeller=label_both) +
  labs(y='Oesophageal Location',x='Months Prior to Endpoint', title='Validation Progressors') +
  plot.theme + theme(axis.text.x=element_text(angle=45, hjust=1))

ggplot(subset(vpd2, Status == 'NP'), aes(brks, nl, fill=Prediction)) + 
  geom_tile() + geom_text(aes(label=Pathology), color='white', show.legend = F) + 
  scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
  facet_wrap(~Patient, ncol=1, scales='free_y', labeller=label_both) +
  labs(y='Oesophageal Location',x='Months Prior to Endpoint', title='Validation Non-Progressors') +
  plot.theme + theme(axis.text.x=element_text(angle=45, hjust=1)) 
```

```{r, echo=F, message=F, warning=F, fig.width=8, fig.height=8}

global.path = collapse.pred.by.path(vpd[subset(sum.validation.info, Status == 'P')$Hospital.Research.ID])
global.pathP53 = collapse.pred.by.path(vpd[subset(sum.validation.info, Status == 'P')$Hospital.Research.ID], T, predFUN = predP53)

global.pathNP = collapse.pred.by.path(vpd[subset(sum.validation.info, Status == 'NP')$Hospital.Research.ID])

pm = melt(global.path, measure.vars=c('prog.ratio','nonprog.ratio'), id.vars=c('path', 'total','pred'))
pm$status = 'Progressor'

pm53m = melt(global.pathP53, measure.vars=c('prog.ratio','nonprog.ratio'), id.vars=c('path', 'total','pred'))
#pm53m$status = 'p53 IHC'
pm53m = subset(pm53m, variable == 'prog.ratio')
pm53m$variable = paste('p53',pm53m$variable,sep='')
pm53m$status = 'Progressor'

npm = melt(global.pathNP, measure.vars=c('prog.ratio','nonprog.ratio'), id.vars=c('path', 'total','pred'))
npm$status = 'Non-Progressor'

m = rbind(pm, npm, pm53m)
m$pred = factor(m$pred)

#m$status = factor(m$status, levels=c('Non-Progressor','Progressor','p53 IHC'), ordered = T)

ggplot(m, aes(path, value, group=path, fill=variable)) + ylim(0,100) + facet_wrap(~status) +
  geom_bar(stat='identity', position=position_stack()) + 
  scale_fill_manual(values=c('firebrick4','green4', 'slateblue4'),name='', labels=c('Pred. progressed','Pred. Not progressed', 'p53 prediction'))  +
  geom_text(aes(vjust=ifelse(value > 12, 2, 1),
                 label=paste(paste(round(value, 1),'%',sep=''),
                                        paste('n=', total, sep=''),sep='\n')), position=position_stack(),color='white',  show.legend=F) +
  geom_bar(data=pm53m, aes(path, value, group=path), fill='slateblue4', stat='identity') +
  geom_text(data=pm53m, aes(vjust=ifelse(value > 12, 2, 1),
                label=paste(paste(round(value, 1),'%',sep=''),
                                       paste('n=', total, sep=''),sep='\n')), position=position_stack(),color='white',  show.legend=F) +
  labs(title='Validation samples predicted by pathology', x='Pathology', y='Percent identified',subtitle='Non-HGD exclude the final endoscopy, p53 excludes NA samples') +
  plot.theme + theme(legend.position = 'bottom')

```

## Cox model - Time to progress?

```{r cox, warning=F, message=F, echo=F, eval=F}
library(survival)

cox.pvalue<-function(cox, test='logrank') {
  pvalue = NULL  
  csum = summary(cox)

  if (test == 'logrank') {
    pvalue = csum$sctest[['pvalue']]  
  } else if (test == 'wald') {
    pvalue = csum$waldtest[['pvalue']]
  } else if (test == 'loglik') {
    pvalue = csum$logtest[['pvalue']]
  }
  return(pvalue)
}

prog.pred = lapply(pg.samp, function(df) {
  df = df[,c('Hospital.Research.ID', 'Status', 'Patient', 'PID', 'Status', 'Endoscopy.Year','Initial.Endoscopy', 'Final.Endoscopy', 'Pathology', 'Samplename','Prediction','Prediction.Dev.Resid')] %>% rowwise() %>% mutate(
      'Time.to.Final'=Final.Endoscopy-Endoscopy.Year)
  df = arrange(df, Endoscopy.Year, Pathology)

  df$Dup.PID = duplicated(df$PID)
  x = df[!df$Dup.PID,]

  for (i in 1:nrow(x)) {
    if (i == 1) {
      x[i,'tstart'] = 0
      x[i, 'tstop'] = x[(i+1), 'Endoscopy.Year'] - x[i, 'Endoscopy.Year']
      if (is.na(x[i,'tstop']) || x[i,'tstop'] == 0) x[i,'tstop'] = 0.5
      next
    }

    x[i,'tstart'] = x[(i-1), 'tstop']

    # Final sample is different -- will leave it out later
    if (i == nrow(x)) {
      x[i, 'tstop'] = x[i,'tstart'] + 0.5
      break
    }

    x[i, 'tstop'] =  x[i,'tstart'] + x[(i+1), 'Endoscopy.Year'] - x[i, 'Endoscopy.Year']

    # Samples taken in the same year
    if ( x[i, 'Endoscopy.Year'] == x[(i+1), 'Endoscopy.Year'] )
      x[i,'tstop'] = x[i,'tstop'] + 0.5 
  }

  x$Diagnosis = as.integer(x$Pathology %in% c('HGD','IMC'))

  if (sum(x$Diagnosis) >= 1)
    x = x[-which(x$Diagnosis == 1),]

  if (nrow(x) > 0 && sum(x$Diagnosis) <= 0 && x$Status == 'P')
    x[nrow(x), 'Diagnosis'] = 1

  if (nrow(x) > 0) {
    x$Dup.Samples = ''
    for (j in 1:nrow(df[df$Dup.PID,]))  {
      row = df[df$Dup.PID,][j,]
      x[which(x$PID == row[['PID']]), 'Dup.Samples'] = paste(x[which(x$PID == row[['PID']]), 'Dup.Samples'], row[['Samplename']], sep=',')
    }
    x$Dup.Samples = sub('^,','',x$Dup.Samples)
  }

  return(x)
})
## By pathology?
pp = do.call(rbind, prog.pred)
pp[is.na(pp$tstop), 'tstop'] = 0.5
pp$Pathology = as.character(pp$Pathology)

sfit = survfit( Surv(tstart, tstop, Diagnosis, type='interval') ~ Pathology, data=pp, conf.int=0.95, conf.type='log', na.action=na.omit)
ggsurv(sfit,cens.shape=3,lty.est=1,back.white=F)
summary(sfit)

cox = coxph( Surv(tstart, tstop, Diagnosis) ~ Pathology, data=pp) 
ggsurv(survfit(cox),cens.shape=3,lty.est=1,back.white=F,
      main="coxph(Survival, Diagnosis) ~ Pathology" ) + 
      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),2),"\n", nrow(pp), 'samples'))
summary(cox)

## All genomic regions
sample.pred = merge( do.call(rbind.data.frame, prog.pred), dysplasia.df, by.x='Samplename', by.y='row.names' )
sample.pred$Pathology = as.character(sample.pred$Pathology)

gcols = grep('\\d+:', colnames(sample.pred), value=T)
sample.pred = sample.pred[c('tstart', 'tstop','Diagnosis','Pathology',gcols)]

# all
#cox = coxph( Surv(tstart, tstop, Diagnosis) ~., data=sample.pred)
#ggsurv(survfit(cox))
#summary(cox)

# glm selected features
cox2 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', 'Pathology',  ch$label)]) 
ggsurv(survfit(cox2))
summary(cox2)

# chr 17 only
cox17 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', grep('^17:',colnames(sample.pred), value=T))]) 
ggsurv(survfit(cox17))
summary(cox17)

cox11 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', grep('^11:',colnames(sample.pred), value=T))]) 
ggsurv(survfit(cox11))
summary(cox11)

cox13 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', grep('^13:',colnames(sample.pred), value=T))]) 
ggsurv(survfit(cox13))
summary(cox13)

```

## CoxRFX

```{r echo=F, message=F, warning=F, eval=F}
plot.waldtest<-function(wt) {
  p = ggplot(wt, aes(coef.x, coef.y, color=p.value<0.05))
  if (length(grep('hazards', colnames(wt))) > 0) {
      p = p + geom_point(alpha=0.5, aes(size=hazards)) + 
        geom_point(aes(coef.x, coef.y, shape=gl, color=p.value<0.05, size=hazards), show.legend = F) +
        #scale_shape_manual(values=c(17,18), na.value=21) +
        scale_size(range=c(2, 10), breaks=c(0, 5, 10, 15)) +
        #geom_text_repel( aes( label=label), show.legend=F) + 
        geom_text_repel(data=subset(wt, p.value < 0.05 & !is.na(gl)), aes(coef.x,coef.y, label=label), show.legend=F) + 
        geom_text_repel(data=subset(wt, is.na(stability)), aes(coef.x, coef.y, label=label, color=p.value<0.05), show.legend=F, color='grey39') 
  } else {
    p = p + geom_point(alpha=0.5) + 
      geom_text_repel(data=subset(wt, p.value < 0.05), aes(coef.x,coef.y, label=label), show.legend=F)  
  }
  return(p + labs(title='Wald test'))
}

```

```{r echo=T, message=F, warning=F, fig.height=8, fig.width=8, eval=F}

library(CoxHD)

coxDF = sample.pred

pathology = do.call(cbind.data.frame, lapply(unique(coxDF$Pathology), function(p) {
  cbind( as.integer(coxDF$Pathology == p))
}))
colnames(pathology) = unique(coxDF$Pathology)

## GLM regions only
coxDF = cbind(coxDF[,1:3],pathology[,c('BE','ID')], coxDF[,ch$label])
coxDF$Pathology = NULL
dim(coxDF)

gcols = colnames(coxDF)
cols = grep('^\\d+', colnames(coxDF))
colnames(coxDF)[cols] = paste('chr', gsub(':|-','.', grep('^\\d+', colnames(coxDF), value=T)), sep='')

surv = with(coxDF, Surv(tstart, tstop, Diagnosis))

groups = factor(c(rep('clinical',2),c(rep('genomic', length(grep('\\d+', colnames(coxDF)))))))
rfx = CoxRFX(coxDF[,-c(1:3)], surv, groups, verbose=F, nu=1)
summary(rfx)

plot(rfx)

wt = WaldTest(rfx)
survConcordance(surv ~ predict(rfx))

wt$label = sub('\\.', '-', sub('\\.', ':', sub('chr', '', rownames(wt))))
wt = merge(wt, ch[,c('label','hazards', 'coef','stability','gl')], by='label', all=T)

wt[is.na(wt$hazards), c('hazards', 'coef.y')] = 0

plot.waldtest(wt) + labs(x='RFX coef',y='GLM coef', title='RFX vs GLM coefficients', subtitle='Labeled points have a significant p-value in RFX (except BE/ID)')

```

### 2017-11-08 lib/load_patient_metadata.R

@@ -235,8 +235,6 @@ add.demographics<-function(file, patient.info) {
   return(demog)
 }
 
-
-
 get.chr.lengths<-function(chrs = paste('chr', c(1:22, 'X','Y'), sep=''), build='hg19') {
   require(plyr)
   chr.lengths = read.table(paste('http://genome.ucsc.edu/goldenpath/help/', build, '.chrom.sizes',sep='') , sep='\t', header=F)

### 2017-11-13 Paper.Rmd
...

@@
 
-```{r echo=F, warning=F, message=F, eval=T, fig.width=7, fig.height=5}
+```{r echo=F, warning=F, message=F, eval=T, fig.width=12, fig.height=8}
+
+cn.mtn.plot<-function(df, label, col=c("orange4","grey90","steelblue3"), type='seg') {
+  p = ggplot(df, aes(x=chr.length)) + facet_grid(Status~chr, scales='free_x', space='free_x', labeller=label)
+
+  if (type == 'seg') {
+    p = p + geom_segment(aes(x=start, xend=end, y=mean.value, yend=mean.value, color=cn), show.legend=T, size=1) + scale_color_manual(values=col) 
+  }
+  if (type == 'bar') {
+    p = p + geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=cn), show.legend=T) + scale_fill_manual(values=col) 
+  }
+
+  p + labs(x='', y='Mean adjusted segmentation value', title='All samples') +
+    theme(text=element_text(face='bold', size=14), axis.text.x=element_blank(), legend.position='bottom', 
+          panel.grid = element_blank(), panel.background = element_blank(),strip.background =element_rect(fill="lightcyan2"),  
+          panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0, 'lines') )
+}
+
 chr.lengths = get.chr.lengths()
@@
 chr.lengths = subset(chr.lengths, chrom %in% c(1:22))
+chr.lengths$chrom = factor(chr.lengths$chrom, levels=c(1:22), ordered = T)
+
+arm.loc = arrange(rbind(
+  (chr.lengths %>% group_by(chrom) %>% mutate('start'=1, 'end'=chr.cent-cent.gap))[c('chrom','start','end')],
+  (chr.lengths %>% group_by(chrom) %>% mutate('start'=chr.cent+cent.gap, 'end'=chr.length))[c('chrom','start','end')]), chrom)
 
@@
 
-globalCNdf = locs
-globalCNdf = cbind(globalCNdf, t(dysplasia.df[,grep('^\\d+:', colnames(dysplasia.df), value=T)]))
+globalCNdf = t(dysplasia.df[,grep('^\\d+:',colnames(dysplasia.df))])
+
+locs = do.call(rbind.data.frame, lapply(rownames(globalCNdf), function(x) unlist(strsplit(x,':|-'))))
+colnames(locs) = c('chr','start','end')
+locs[] = lapply(locs[], function(x) as.numeric(as.character(x)))
+
+globalCNdf = cbind.data.frame(locs, globalCNdf)
+globalCNdf = melt(globalCNdf, id.vars=colnames(locs))
+
+globalCNdf$seg.type = '5MB'
+globalCNdf[with(globalCNdf, which( start %in% arm.loc$start & end %in% arm.loc$end)), 'seg.type'] = 'arms'
 globalCNdf = merge(globalCNdf, chr.lengths[,c('chrom','chr.length')], by.x='chr', by.y='chrom')
-globalCNdf = melt(globalCNdf, id.vars=c('chr','start','end', 'chr.length'))
-globalCNdf$Status = ifelse(globalCNdf$variable %in% names(which(labels == 1)), 'Progressor', 'Non-Progressor')
-
-globalCNdf = merge(globalCNdf, patient.info[,c('Hospital.Research.ID','Patient','Samplename','Pathology')], by.x='variable', by.y='Samplename')
-
-#globalCNdf = subset(globalCNdf, value != 0)
+
+globalCNdf = merge(globalCNdf, patient.info[,c('Hospital.Research.ID','Patient','Samplename','Pathology','Status')], by.x='variable', by.y='Samplename')
+globalCNdf$Status = ifelse(globalCNdf$Status == 'P', 'Progressor', 'Non-Progressor')
 
@@
 }
-
 globalCNdf$cn = factor(sapply(globalCNdf$value, cn), levels=c('loss','neutral','gain'), ordered = T)
-globalCNdf$cn = factor(sapply(globalCNdf$value, cn), levels=c('loss','neutral','gain'), ordered = T)
-globalCNdf$cnB = factor(as.integer(globalCNdf$cn != 'neutral'))
-
-cn.colors = c('orange3','grey88','steelblue2')
 
@@
 
-# Average cn number...looks the same as expected
+# Average cn value
 globalCNdf = globalCNdf %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
-globalCN.plot = 
-  ggplot(globalCNdf, aes(x=chr.length)) + facet_grid(Status~chr, scales='free_x', space='free_x', labeller=labeller(Status=n.status, chr=label_value)) +
-  geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=cn), alpha=0.5, show.legend = T) + 
-  scale_color_manual(values=cn.colors) + 
-  scale_fill_manual(values=cn.colors) +
-  labs(x='', y='Mean adjusted segmentation value', title='All pathologies') +
-  theme(text=element_text(face='bold', size=14), axis.text.x=element_blank(), legend.position='bottom', 
-        panel.grid = element_blank(), panel.background = element_blank(),strip.background =element_rect(fill="lightcyan2"),  
-        panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0, 'lines') )
-
+globalCNdf$mean.value = globalCNdf$value # but not going to use it
+
+globalCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms'), labeller(Status=n.status, chr=label_value))
 cnLegend = get.legend(globalCN.plot)
@@
 
-globalCNBEdf = subset(globalCNdf, Pathology == 'BE') %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
-globalBECN.plot = 
-  ggplot(globalCNBEdf, aes(x=chr.length)) + 
-  facet_grid(Status~chr, scales='free_x', space='free_x', labeller=labeller(Status=nb.status, chr=label_value)) + 
-  #ylim(range(globalCNdf$value)) +
-  geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=cn), alpha=0.5, show.legend = T) + 
-  scale_color_manual(values=cn.colors) + scale_fill_manual(values=cn.colors) +
-  labs(x='', y='Mean adjusted segmentation value', title='NDBE samples only') +
-  theme(text=element_text(face='bold', size=14), axis.text.x=element_blank(), legend.position='bottom', 
-        panel.grid = element_blank(), panel.background = element_blank(),strip.background =element_rect(fill="lightcyan2"),  
-        panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0, 'lines') )
+globalBECN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology == 'BE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')
+#globalCNBEdf = subset(globalCNdf, Pathology == 'BE') %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
 
@@
 
-dfpts = subset(globalCNdf, Patient %in% c('57','76'))
-dfpts = dfpts %>% group_by(Patient, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
+dfpts = subset(globalCNdf, Patient %in% c('57','76') & seg.type != 'arms')
+#dfpts = dfpts %>% group_by(Patient, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
 
@@
               '76'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))
-
-genomePtCN.plot = 
-ggplot(dfpts, aes(x=chr.length)) + 
-  facet_grid(Patient~chr, scales='free_x',space='free_x',  labeller = labeller(Patient = pt.status )) + 
-  geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=cn), show.legend = F, alpha=0.5) + 
-  scale_fill_manual(values=cn.colors) +
-  labs(x='', y='Mean adjusted segmentation value') +
-  theme(text=element_text(face='bold', size=14), axis.text.x=element_blank(), legend.position='bottom', 
-        panel.grid = element_blank(), panel.background = element_blank(),strip.background =element_rect(fill="lightcyan2"),  
-        panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0, 'lines') )
-
-dfpts = subset(globalCNdf, Patient %in% c('57','76') & Pathology == 'BE')
-dfpts = dfpts %>% group_by(Patient, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
+genomePtCN.plot = cn.mtn.plot(dfpts, labeller(Patient = pt.status))
+
+dfpts = subset(globalCNdf, Patient %in% c('57','76') & Pathology == 'BE' & seg.type != 'arms')
+#dfpts = dfpts %>% group_by(Patient, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
 
@@
               '76'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))
-genomePtBECN.plot = 
-  ggplot(dfpts, aes(x=chr.length)) + 
-  facet_grid(Patient~chr, scales='free_x', space='free_x', labeller = labeller(Patient = ptbe.status )) + 
-  geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=cn), show.legend = F, alpha=0.5) + 
-  scale_fill_manual(values=cn.colors) +
-  labs(x='', y='Mean adjusted segmentation value') +
-  theme(text=element_text(face='bold', size=14), axis.text.x=element_blank(), legend.position='bottom', 
-        panel.grid = element_blank(), panel.background = element_blank(),strip.background =element_rect(fill="lightcyan2"),  
-        panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0, 'lines') )
+
+genomePtBECN.plot = cn.mtn.plot(dfpts,  labeller(Patient = ptbe.status )) + labs(title='NDBE samples only')
 
@@
 genomePtBECN.plot
+
+cn.mtn.plot(dfpts,  labeller(Patient = ptbe.status ), type='bar') + labs(title='NDBE samples only')
+```
+
+## Individual samples
+```{r}
+
+a = sample(subset(patient.info, Status == 'NP' & Pathology %in% c('BE'))$Samplename, 1)
+b = sample(subset(patient.info, Status == 'P' & Pathology %in% c('BE'))$Samplename, 1)
+
+cn.mtn.plot(subset(globalCNdf, variable %in% c(a) & seg.type != 'arms'), labeller(label_both)) + labs(title="Individual NDBE samples") + theme(strip.text = element_text(size=10, face='bold'))
+cn.mtn.plot(subset(globalCNdf, variable %in% c(b) & seg.type != 'arms'), labeller(label_both)) + labs(title="Individual NDBE samples") + theme(strip.text = element_text(size=10, face='bold'))
+
+cn.mtn.plot(subset(globalCNdf, variable %in% c(a,b) & seg.type != 'arms'), labeller(label_both)) + labs(title="Individual NDBE samples")
+cn.mtn.plot(subset(globalCNdf, variable %in% c(a,b) & seg.type != 'arms'), labeller(label_both), type='bar') + labs(title="Individual NDBE samples")
+
 ```
@@
 
-```{r, echo=F, fig.height=5, fig.width=5, eval=F}
+```{r, echo=F, fig.height=5, fig.width=5}
 mp = model.performance(performance.at.1se, coefs)
@@
   geom_text( aes(label=round(mean, 3)), hjust=-0.5, size=5, color='gray39', show.legend = F) +
-  geom_text( aes(y=mean+sme*3, label=paste(all.coef, "\n(", stable.coef, ")", sep='') ), color='gray39', size=5) +
+  geom_text( aes(y=mean+sme*2, label=paste(all.coef, "\n(", stable.coef, ")", sep='') ), color='gray39', size=5) +
   geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme),color='grey39', width=0.1, size=1) + 
@@
 custom.max<-function(cf, hazards, ymax) {
-  print(cf)
   if (cf > 0) {
@@
 
-globalCN.plot + geom_rect(data=chh, aes(xmin=start, xmax=end, ymin=0, ymax=max, color=gl), alpha=0, show.legend=F) + scale_color_manual(values=c('goldenrod2','blue3'))
-globalBECN.plot + geom_rect(data=chh, aes(xmin=start, xmax=end, ymin=0, ymax=max, color=gl), alpha=0, show.legend=F) + scale_color_manual(values=c('goldenrod2','blue3'))
+globalCN.plot + geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)
+
+globalBECN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)
 
@@
 chh = chh %>% rowwise() %>% mutate('max'=custom.max(coef, hazards, ymax))
-genomePtCN.plot + geom_rect(data=chh, aes(xmin=start, xmax=end, ymin=0, ymax=max, color=gl), alpha=0, show.legend=F) + scale_color_manual(values=c('goldenrod2','blue3'))
-genomePtBECN.plot + geom_rect(data=chh, aes(xmin=start, xmax=end, ymin=0, ymax=max, color=gl), alpha=0, show.legend=F) + scale_color_manual(values=c('goldenrod2','blue3'))
+genomePtCN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)
+
+genomePtBECN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)
 
@@
 
-p3 = ggplot(chh, aes(x=label,y=coef,fill=gl)) + geom_bar(stat='identity') + labs(x='Effects', y='') +
+```
+
+```{r echo=F, fig.height=4, fig.width=8}
+ggplot(chh, aes(x=label,y=coef,fill=gl)) + geom_bar(stat='identity') + labs(x='Effects', y='') +
   coord_flip() + scale_fill_manual(values=c('darkgreen','darkred')) +
   plot.theme + theme(axis.line.x=element_blank(), axis.ticks=element_blank(),axis.text=element_blank(), panel.grid=element_blank(),panel.grid.major=element_blank(), legend.position='none')
-p3
-
-#  lay <- rbind(c(1,1,1,4),
-#               c(2,2,2,5),
-#               c(3,3,3,6),
-#               c(7,7,7,8))
-# 
-# grid.arrange(p3,p1a,p1b,gp,p2a,p2b,p1leg,gp, layout_matrix=lay, widths=4:1, heights=c(2,4,4,1))
 ```
@@
 }
+loo.coefs = coefs
 
@@
   plot.theme
+
+oddPt = names(pg.samp)[which.max(performance.at.1se)]
+subset(sum.patient.data, Hospital.Research.ID == oddPt)
+
+pg.samp[[oddPt]]$Prediction
+
+# Not the pt with the most samples or endos
+sum.patient.data[which.max(sum.patient.data$total.samples),]
+
 ```
@@
 
-```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=8, fig.width=4}
+```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=4, fig.width=4}
 roc.plot <- function(roc,title="") {
@@
 
-```{r p53roc, echo=F, fig.height=5, fig.width=5}
+```{r p53roc, echo=F, fig.height=4, fig.width=4}
 p53samples = subset(patient.info, !is.na(p53.Status), select=c('Samplename','p53.Status','Status'))
@@
 
-```{r pathroc, echo=F, fig.height=5, fig.width=5 }
+```{r pathroc, echo=F, fig.height=4, fig.width=4 }
 #sum(table(subset(patient.info, Status == 'P')$Pathology)[c('IMC','HGD')])/sum(table(subset(patient.info, Status == 'P')$Pathology))
@@
 
-```{r echo=F}
+```{r echo=F, fig.height=6, fig.width=8}
 
@@
 
-m = rbind(get.roc.stat(p53roc),
+m = rbind.data.frame(get.roc.stat(p53roc),
       get.roc.stat(hgdpathroc),
@@
 
-ggplot(melt(m), aes(model, value*100, group=variable, fill=variable)) + ylim(0,100) +
+ggplot(melt(m,id.vars = 'model'), aes(model, value*100, group=variable, fill=variable)) + ylim(0,100) +
   geom_col(position="dodge") + 
@@
 
+## Jacknife coefficients
+
+The idea here is to get some sort of CI for the coefficients, I'm not entirely sure this is correct though.
+
+```{r jacknife, echo=F, warning=F, message=F}
+
+coefs.per.pt = data.frame(matrix(ncol=length(pg.samp), nrow=nrow(ch), dimnames=list( ch$label, names(pg.samp))))
+for (pt in names(pg.samp)) {
+  lcf = loo.coefs[[pt]][,1]
+  names(lcf) = rownames( loo.coefs[[pt]] )
+
+  sharedcfs = intersect(names(lcf), ch$label)
+
+  coefs.per.pt[sharedcfs, pt] = lcf[sharedcfs]
+}
+coefs.per.pt[is.na(coefs.per.pt)] = 0
+
+library(bootstrap)
+
+ch$jack.se = apply(coefs.per.pt, 1, function(x) {
+  jk = jackknife(x, var)
+  jk$jack.se
+})
+
+ch$adj = ifelse(ch$coef<0,-1,1)
+ggplot(ch, aes(reorder(label, coef), coef)) + geom_point() + 
+  geom_errorbar(aes(ymin=(abs(coef)-jack.se)*adj, ymax=(abs(coef)+jack.se)*adj)) + coord_flip() +
+  plot.theme + labs(x='', title='Jackknifed coefficient CI??')
+
+```
+
 ## How do the predictions look at each time point?
@@
 m$final = ifelse(m$path == 'HGD', T, F)
-
-ggplot(m, aes(path, value, group=path, fill=variable)) + ylim(0,100) + facet_wrap(~status, ncol=1) +
-  scale_fill_manual(values=c('firebrick4','green4', 'slateblue4'), name='', labels=c('Pred. progressed', 'Pred. Not progressed', 'p53 prediction'))  +
-  geom_bar(stat='identity', position=position_stack()) +
-  geom_text(aes(hjust=ifelse(value > 12, 2, 1),
-                 label=paste(paste(round(value, 1),'%',sep=''),
-                                        paste('n=', total, sep=''),sep='\n')), 
-            position=position_stack(),color='white',  show.legend=F) +
-  geom_bar(data=p53.melt, aes(path, value, group=path), fill='slateblue4',stat='identity',position = position_dodge()) +
-  geom_text(data=p53.melt, aes(path,value,hjust=ifelse(value > 12, 2, 1),label=paste(paste(round(value, 1),'%',sep=''),paste('n=', total, sep=''),sep='\n')), position=position_dodge(),color='white',  show.legend=F) +
-    labs(title='Samples predicted by pathology', x='Pathology', y='Percent', subtitle='Non-HGD exclude the final endoscopy, p53 excludes NA samples') +
-  plot.theme + theme(text = element_text(size=16, face='bold'), legend.position = 'bottom') + coord_flip()
 
@@
 ## Flip order, 0-1 comes last
-m = endos %>% group_by(brks, labels) %>% summarise( 'pred.r'= length(which(pred > 0))/length(brks), 'n'=length((Patient)) )
+m = endos %>% group_by(brks, labels) %>% summarise( 'pred.r'= length(which(pred > 0))/length(brks), 'n'=length((Patient)), 'n.samp'=sum(samples) )
 
@@
   geom_text(aes(y=pred.r*100+1.5, label=paste(round(pred.r*100), '%', sep=''))) +
-  geom_text(aes(y=pred.r*100-2, label=paste('n=',n, sep='')), color='white') + 
+  geom_text(aes(y=pred.r*100-2, label=paste('n=',n,'\ns=',n.samp,sep='')), color='white') + 
   scale_x_discrete(labels=m$labels) + 
-  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = number of endoscopies') +
+  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') +
   plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1))
@@
 ## Progressor predictions per patient
-```{r, echo=F, fig.height=12, fig.width=10}
-plot.endo( subset(endos, Patient %in% subset(me, months <= -72)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(72/12, "years or more follow-up"))
-```
-
-```{r, echo=F, fig.height=12, fig.width=10}
+```{r, echo=F, fig.height=8, fig.width=6}
+#plot.endo( subset(endos, Patient %in% subset(me, months <= -72)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(72/12, "years or more follow-up"))
+```
+
+```{r, echo=F, fig.height=8, fig.width=6}
 # plot.endo( subset(endos, Patient %in% subset(me, months <= -48 & months > -72)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(48/12, "-",72/12,  "years or more follow-up"))
 # plot.endo( subset(endos, Patient %in% subset(me, months <= -24 & months > -48)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(24/12, "-",48/12,  "years or more follow-up"))
-plot.endo( subset(endos, Patient %in% subset(me, months > -48)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(1, "-",48/12,  "years or more follow-up"))
+#plot.endo( subset(endos, Patient %in% subset(me, months > -48)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(1, "-",48/12,  "years or more follow-up"))
 ```
@@
 ## Non-progressor predictions per patient
-```{r, echo=F, fig.height=24, fig.width=16}
+```{r, echo=F, fig.height=12, fig.width=8}
 nonp.back = lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID],transform.pt.by.time)
@@
 
-plot.endo(subset(nonp.endos, Patient %in% subset(nonp.endos, pred.ratio > 0)$Patient), bin=T) + labs(title='Non-progressors with progressive predictions')
+#plot.endo(subset(nonp.endos, Patient %in% subset(nonp.endos, pred.ratio > 0)$Patient), bin=T) + labs(title='Non-progressors with progressive predictions')
 ```
@@
 
-```{r spatial, echo=F}
-
-# ids = subset(sum.patient.data, Status == 'P' & total.endos > 5 & M > mean(M, na.rm=T))$Hospital.Research.ID
-# missing_p53 = do.call(rbind.data.frame, lapply(pg.samp, function(df) {
-#   x = df[which(is.na(df$p53.Status)), c('Hospital.Research.ID', 'Path.ID', 'Status')]
-#   if (nrow(x) > 0) x$Prioritize = ifelse(unique(df$Hospital.Research.ID) %in% ids, '*', '')
-#   return(x)
-# }))
-# write.table(missing_p53, sep='\t', quote=F, row.names = F, col.names = T, file='missing_p53.txt')
-
-#ids = subset(sum.patient.data, ((Status == 'P' & total.endos > 5) | (Status == 'NP' & total.endos > 9)) & M > mean(M, na.rm=T))$Hospital.Research.ID
-
+```{r, spacial, fig.width=8, fig.height=6}
 transform.by.time<-function(pt) {
   pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction')] %>% group_by(PID, Samplename) %>% mutate(
-  'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12,
-  'nl' = str_extract(Path.ID, "B\\w?[0-9]+"))
-
-  pt$nl = (as.numeric(factor(pt$nl, levels=(unique(sort(pt$nl))), ordered = T))-1)*2
+  'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12
+  #'nl' = Block
+  #'nl' = str_extract(Path.ID, "B\\w?[0-9]+")
+  )
+
+  pt = merge(pt, subset(patient.info, Hospital.Research.ID == pt$Hospital.Research.ID[1], select=c('Samplename','Block')), by='Samplename')
+
+  pt$nl = (pt$Block-1)*2
+  #pt$nl = (as.numeric(factor(pt$nl, levels=(unique(sort(pt$nl))), ordered = T))-1)*2
   #table(pt[,c('nl','months.before.final')])
@@
 back = do.call(rbind.data.frame,lapply(pg.samp, transform.by.time))
-```
-
-```{r fig.width=8, fig.height=8}
-probs = seq(0,1,1/4)
-back$quants = with(back, cut(Prediction, breaks=quantile(Prediction, probs), include.lowest = T, labels=probs[-1]))
-
-qt = table(back$quants, back$Status)
-
-pq = cbind.data.frame('perc'=qt[,'P']/rowSums(qt),
-      'quant'=quantile( back$Prediction, probs)[-1])
-ggplot(back, aes(Prediction, fill=..x..)) + geom_histogram(bins=20, show.legend = F) + 
+back = merge(back, patient.info[,c('Samplename','p53.Status')])
+
+#cuts = seq(0,1,0.1)
+cuts = quantile(back$Prediction, probs=seq(0,1,1/5))
+back$quants = with(back, cut(Prediction, breaks=cuts))
+qt = as.data.frame.matrix(table(back$quants, back$Status))
+qt$quant = rownames(qt)
+
+ft.fun<-function(NP,P) {
+ f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
+ #f$p.value
+ cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
+}
+
+pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'conf'=ifelse(p.value < 0.05, '*', '') )
+pred.confidence[c('r1','r2')] = NA
+for (i in 1:(length(cuts)-1)) 
+    pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)
+
+ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), bins=20, show.legend = F) +
   scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
-  plot.theme + labs(title='Sample predictions', y='n Samples', x='Probability') +
-  geom_vline(xintercept=quantile( back$Prediction, probs), linetype='longdash') +
-  geom_label(data=pq, aes(x=quant, y=75, label=round(perc,2) ))
-
-```
+  geom_vline(xintercept=pred.confidence$r2, linetype='longdash') +
+  geom_label(data=(pred.confidence %>% rowwise %>% mutate( 'mid'=sum(r1,r2)/2 )), aes(x=mid, y=75, label=paste(round(perc,2),conf,sep=''),color=conf ), show.legend=F) + 
+  scale_color_manual(values=c('grey','black')) + plot.theme + labs(title='Sample predictions', y='n Samples', x='Probability') 
+```
+
+At each quintile are the percentage of progressor samples that were predicted within that range, and the hypergeometric p-value (Fishers) for the significance within that quintile.
+Then our predictions are at best a coin flip in the middle, but highly significant at the outer edges.
 
@@
 
+pred.tiles <- function(df) {
+  ggplot(df, aes(brks, nl, fill=Prediction)) + 
+      geom_tile() + scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
+      geom_text(aes(label=labels, color=insig), show.legend=F) + scale_color_manual(values=c('grey','white')) +
+      facet_wrap(~Patient, ncol=1, scales='free_y', labeller=label_both) +
+      labs(y='Oesophageal Location',x='Months Prior to Endpoint', title='') +
+      plot.theme + theme(axis.text.x=element_text(angle=45, hjust=1)) 
+}
+
 back$brks = with(back, factor(brks, rev(levels(brks)))) # invert order
+#back$labels = paste(back$Pathology, ifelse(back$p53.Status == 0, '--', '**') ,sep='')
+#back$labels = with(back, ifelse(is.na(p53.Status), as.character(Pathology), labels))
+back$labels = back$Pathology
+back$nl = factor(back$nl, ordered=T)
+
+insig = subset(pred.confidence, p.value > 0.05, select=c('quant','r1','r2'))
+
+back$insig = back$quants != insig$quant
 
 ids = c(13,20,34,56,60,48)
-p = ggplot(subset(back, Patient %in% ids & Status == 'P'), aes(brks, nl, fill=Prediction)) + 
-      geom_tile() + geom_text(aes(label=Pathology), color='white', show.legend = F) + 
-      scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
-      facet_wrap(~Patient, ncol=1, scales='free_y', labeller=label_both) +
-      labs(y='Oesophageal Location',x='Months Prior to Endpoint', title='Progressors') +
-      plot.theme + theme(axis.text.x=element_text(angle=45, hjust=1)) 
-p
-#ggsave('pg_heatmap1.png', plot = p,
-#  scale = 1, width = 5, height = 12, units = "in", limitsize = F,
-#  dpi = 300)
+pred.tiles(subset(back, Patient %in% ids & Status == 'P')) + labs(title='Progressors')
 
 ids = c(15,16,21,3,45,19)
-p=ggplot(subset(back, Patient %in% ids & Status == 'NP'), aes(brks, nl, fill=Prediction)) + 
-  geom_tile() + geom_text(aes(label=Pathology), color='white', show.legend = F) + 
-  scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
-  facet_wrap(~Patient, ncol=1, scales='free_y', labeller=label_both) +
-  labs(y='Oesophageal Location',x='Months Prior to Endpoint', title='Non-Progressors') +
-  plot.theme + theme(axis.text.x=element_text(angle=45, hjust=1)) 
-#ggsave('npg_heatmap1.png', plot = p,
-#  scale = 1, width = 5, height = 12, units = "in", limitsize = F,
-#  dpi = 300)
-p
+pred.tiles(subset(back, Patient %in% ids & Status == 'NP')) + labs(title='Non-Progressors')
+
 ```
@@
 ```{r, echo=F}
-# 
-#  ggplot(subset(back, Status == 'NP' & Patient %in% ids), aes(brks, nl, fill=Prediction, group=Patient)) + 
-#   facet_wrap(~Patient, scales='free', labeller=label_both, ncol=1) +
-#   geom_tile() + geom_text(aes(label=Pathology), color='white', show.legend = F) + 
-#   scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
-#   labs(y='Oesophageal Location',x='Months Prior to Endpoint', title='Non-Progressors') +
-#   plot.theme + theme(axis.text.x=element_text(angle=45, hjust=1), strip.text=element_text(size=8), strip.background = element_rect(fill='gray88'))
-
-p = ggplot(subset(back, Status == 'NP'), aes(brks, nl, fill=Prediction, group=Patient)) + 
-  facet_wrap(~Patient, scales='free_y', labeller=label_both, ncol=1) +
-  geom_tile() + geom_text(aes(label=Pathology), color='white', show.legend = F) + 
-  scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
-  labs(y='Oesophageal Location',x='Months Prior to Endpoint', title='Non-Progressors') +
-  plot.theme + theme(axis.text.x=element_text(angle=45, hjust=1), strip.text=element_text(size=8), strip.background = element_rect(fill='gray88'))
+p = pred.tiles(subset(back, Status == 'NP')) + labs(title='Non-Progressors')
 ggsave('npg_heatmap.png', plot = p, scale = 1, width = 8, height = 52, units = "in", limitsize = F, dpi = 300)
 
-p = ggplot(subset(back, Status == 'P'), aes(brks, nl, fill=Prediction, group=Patient)) + 
-  facet_wrap(~Patient, scales='free_y', labeller=label_both, ncol=1) +
-  geom_tile() + geom_text(aes(label=Pathology), color='white', show.legend = F) + 
-  scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
-  labs(y='Oesophageal Location',x='Months Prior to Endpoint', title='Progressors') +
-  plot.theme + theme(axis.text.x=element_text(angle=45, hjust=1), strip.text=element_text(size=10), strip.background = element_rect(fill='gray88'))
+p = pred.tiles(subset(back, Status == 'P')) + labs(title='Progressors')
 ggsave('pg_heatmap.png', plot = p, scale = 1, width = 8, height = 52, units = "in", limitsize = F, dpi = 300)
@@
 ```{r}
-p11 = subset(globalCNdf, chr == 17 & start < subset(chr.lengths, chrom == 17)$chr.cent)
+
+p11 = subset(globalCNdf, chr == 17 &  start == 1 & seg.type == 'arms')
 
 p11Loss = subset(p11, value < -1)
-length(unique(p11Loss$variable))
-
-table(subset(patient.info, Samplename %in% p11Loss$variable)$Pathology)
-(18+12)/116
+table(p11Loss$Status)
+
+p11path = table(subset(patient.info, Samplename %in% p11Loss$variable)$Pathology)
+
+p11path['BE']/sum(p11path)
+sum(p11path[c('HGD','IMC')])/sum(p11path)
 
 # Samples with a loss vs all samples
-length(unique(p11Loss$variable))/length(unique(globalCNdf$variable))
+length(unique(p11Loss$variable))/length(unique(subset(globalCNdf, Status == 'Progressor')$variable))
 
 p11Loss = unique(p11Loss[,c('variable', 'Status','Hospital.Research.ID', 'Patient', 'Pathology')])
-
+length(unique(p11Loss$Patient))/length(subset(sum.patient.data, Status == 'P')$Patient)
 # Of those where we see loss of p arm, p53 IHC shows an "abnormal" result 33% of the time
-table(subset(patient.info, Samplename %in% p11Loss$variable, select='p53.Status'))
+
+table(subset(patient.info, Samplename %in% p11Loss$variable, select='p53.Status'))/length(p11Loss$variable)
 
@@
 One way of looking at the p53 prediction (e.g. Adam Bass) is to see how predictive the p-arm loss in chr17 is.
-```{r, echo=F, warning=F, message=F, fig.height=16, fig.width=18, eval=F}
+```{r, echo=F, warning=F, message=F, fig.height=16, fig.width=18}
 ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
@@
 loc[] = lapply(loc[], function(x) as.numeric(as.character(x)))
-genes = cbind(genes, loc, 'value'=-7, 'Row.names'='', 'Patient'='', 'selected'=as.factor(0))
-
-plot.cn.chr<-function(cn, chrom='17', info=NULL, haz=NULL, genes=NULL) {
-    p = ggplot(subset(cn, chr == chrom), aes(x=chr.length)) + facet_grid(Status~chr, scales='free_x', labeller=labeller(chr=label_both)) + 
-          geom_rect( aes(xmin=start, xmax=end, ymin=0, ymax=value, fill=cn), alpha=0.5, show.legend = T) + 
-          scale_color_manual(values=c('firebrick3','grey88','green4')) +
-          scale_fill_manual(values=c('firebrick3','grey88','green4'))
+
+genes = cbind(genes, loc, 'value'=min(globalCNdf$mean.value), 'Row.names'='', 'Patient'='', 'selected'=as.factor(0))
+
+plot.cn.chr<-function(df, chrom='17', info=NULL, haz=NULL, genes=NULL) {
+    df = subset(df, chr == chrom)  
+    p = ggplot(df, aes(x=chr.length)) + facet_grid(Status~chr, scales='free_x', labeller=labeller(chr=label_both)) +
+        geom_rect(data=subset(df, seg.type != 'arms'), aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=cn), alpha=0.5, show.legend = T) + 
+        geom_segment(data=subset(df, seg.type == 'arms'), aes(x=start, xend=end, y=mean.value, yend=mean.value, color=cn), alpha=0.8, show.legend = F) + 
+        scale_color_manual(values=cn.colors) + scale_fill_manual(values=cn.fill)
 
@@
       p = p + geom_vline(data=info, aes(xintercept=chr.cent), color='grey39', linetype='longdash' ) +
-              geom_label(data=info, aes(x=chr.cent-cent.gap*3, y=10), label='p-arm')
+              geom_label(data=info, aes(x=chr.cent-cent.gap*3, y=max(df$mean.value)), label='p-arm')
     }
@@
       genes = subset(genes, chr == chrom)
-      if (nrow(genes) > 0) p = p + geom_point(data=genes, aes(x=start,  y=value), color='red', size=3) +
-              geom_text_repel(data=genes, aes(x=start, y=value, label=Gene.Symbol), color='red') 
+      if (nrow(genes) > 0) p = p + geom_point(data=genes, aes(x=start,  y=value), color='darkblue', size=3) +
+              geom_text_repel(data=genes, aes(x=start, y=value, label=Gene.Symbol), color='darkblue') 
     }
@@
       haz = subset(haz, chr == chrom)
-      if (nrow(haz) > 0) p = p + geom_rect(data=haz, aes(xmin=start, xmax=end, ymin=0, ymax=max, color=gl), alpha=0, show.legend=F, size=1) + scale_color_manual(values=c('goldenrod2','blue3')) 
+      if (nrow(haz) > 0 & length(grep('loss',unique(haz$gl))) > 0 ) {
+          p = p + geom_rect(data=subset(haz, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F, size=1) 
+      } 
+      if (nrow(haz) > 0 & length(grep('gain',unique(haz$gl))) > 0) {
+          p = p + geom_rect(data=subset(haz, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F, size=1) 
+      }
     }
@@
           panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0.1, 'lines') ) + labs(x='Chromosomes', y='Mean adjusted segmentation value')
-
-}
-
-plot.cn.chr(cn=subset(globalCNdf, Pathology == 'BE'), chrom='17', chr.lengths, genes=genes, haz=ch) + labs(
-title="Segment Values Across BE Samples", subtitle=paste('NonP:Prog - ', paste(table(subset(patient.info, Pathology == 'BE')$Status), collapse=':')))
-
-#do.call(grid.arrange, lapply( unique(ch$chr), function(chr) {
-#  plot.cn.chr(cn=subset(globalCNdf, Pathology == 'BE'), chrom=chr, chr.lengths, haz=ch, genes=genes) 
-#}))
-
-# plot.cn.chr(cn=subset(globalCNdf, Pathology == 'BE'), chrom='12', chr.lengths, haz=ch, genes=genes)
-# 
-plot.cn.chr(cn=subset(globalCNdf, Pathology == 'BE'), chrom='9', chr.lengths, genes=genes, haz=ch)
-plot.cn.chr(cn=subset(globalCNdf, Pathology == 'BE'), chrom='16', chr.lengths, haz=ch, genes=genes)
+}
+
+plot.cn.chr(subset(globalCNdf), chrom='17', chr.lengths, genes=genes, haz=chh) + labs(
+title="Segment Values Samples", subtitle=paste('NonP:Prog - ', paste(table(patient.info$Status), collapse=':')))
+
+plot.cn.chr(globalCNdf, chrom='8', chr.lengths, genes=genes, haz=chh)
+
+plot.cn.chr(globalCNdf, chrom='3', chr.lengths, genes=genes, haz=chh)
+
+plot.cn.chr(globalCNdf, chrom='21', chr.lengths, haz=chh, genes=genes)
+
+plot.cn.chr(globalCNdf, chrom='11', chr.lengths, haz=chh, genes=genes)
+
+plot.cn.chr(globalCNdf, chrom='16', chr.lengths, haz=chh, genes=genes)
 
 genesGR = makeGRangesFromDataFrame(genes, keep.extra.columns = T)
-chGR = makeGRangesFromDataFrame(ch, keep.extra.columns = T)
-
-findOverlaps(chGR, genesGR)
+chGR = makeGRangesFromDataFrame(chh, keep.extra.columns = T)
+
+ov = findOverlaps(chGR, genesGR)
+
+as.data.frame(chGR[unique(queryHits(ov))])
+length(genesGR[unique(subjectHits(ov))]$Gene.Symbol)
+write.table(genesGR[unique(subjectHits(ov))]$Gene.Symbol, quote=F, row.names=F, sep='\t')
 
@@
   plot.theme + theme(legend.position = 'bottom')
+
+```
+
+### Clinical Story
+
+```{r}
+# First and penultimate endoscopy where possible
+clin = lapply(vpd, function(v) {
+  v = arrange(v, Endoscopy.Year)
+  tb = table(v$Endoscopy.Year, v$PID)
+
+  first = subset(v, Endoscopy.Year == rownames(tb)[1] & PID == names(which(tb[1,] > 0))[1])
+  if (nrow(tb) > 2) {
+    penult = subset(v, Endoscopy.Year == rownames(tb)[(nrow(tb)-1)] & PID == names(which(tb[(nrow(tb)-1),] > 0))[1])
+  } else if (nrow(tb) == 2) {
+    penult = subset(v, Endoscopy.Year == rownames(tb)[(nrow(tb))] & PID == names(which(tb[(nrow(tb)),] > 0))[1])
+  } else {
+    penult = first
+  } 
+  cols = c('Hospital.Research.ID','Patient','Status','Path.ID','PID','Endoscopy.Year','Prediction','Samplename')
+  fp = unique(rbind.data.frame(first[cols],penult[cols]))
+
+  fp$conf = cut(fp$Prediction, breaks=cuts, labels=cuts[-1])
+  fp = merge(fp, pred.confidence, by.x='conf', by.y='quant')
+  fp$conf = round(fp$perc, 2)
+  fp$perc = NULL
+  fp$Path.ID = factor(fp$Path.ID)
+
+  fp = merge(fp, validation.patient.info[c('Samplename','Block')])
+  fp$nl = factor( (fp$Block-1)*2, ordered = T)
+  fp$Endoscopy.Year = factor(fp$Endoscopy.Year, ordered = T)
+
+  fp
+})
+
+df = do.call(rbind.data.frame,clin)
+df$conf = factor(df$conf)
+
+ggplot(df, aes(Endoscopy.Year, Prediction, fill=Prediction, color=Prediction)) + 
+  geom_point(aes(size=conf)) + facet_grid(~Patient,  scales='free_x', labeller=label_both) +
+  #scale_size_continuous(breaks=cuts,  labels=cuts) + 
+  scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
+scale_color_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
+   theme(axis.text.x=element_text(angle=45, hjust=1))
+
+df$conf
+pred.confidence
+
+```
+
+## Clonal Heterogenity
+
+```{r clonality, echo=F, message=F, warning=F}
+
+## Variance in in CN per endoscopy
+pt = 'AH0254'
+subset(sum.patient.data, Status == 'P')
+
+subset(sum.patient.data, Patient == 34)
+
+table(pg.samp[[pt]]$Endoscopy.Year) 
+which( table(pg.samp[[pt]]$PID) > 1)
+
+subset(pg.samp[[pt]], PID == 'PS10.5754_')$Samplename
+
+x = apply( dysplasia.df[ subset(pg.samp[[pt]], PID == 'PS10.5754_')$Samplename,  ], 2, var)
+
+x[[1]]
+
+# So pretty much the # of samples == # of endoscopies so can't do variance per endo
+plot(sum.patient.data[c('total.samples','total.endos')])
+cor(sum.patient.data[c('total.samples','total.endos')])
 

### 2017-11-13 lib/cv-pt-glm.R

@@ -165,7 +165,7 @@ crossvalidate.by.patient<-function(x,y,lambda,pts,a=1,nfolds=10, splits=5, fit=N
       test = x[test.rows,]
       training = x[-test.rows,]
       # pre-spec lambda seq
-      fitCV = glmnet(training, y[-test.rows], lambda=lambda, alpha=a, family='binomial', standardize = F) 
+      fitCV = glmnet(training, y[-test.rows], lambda=lambda, alpha=a, family='binomial',  ...) 
       # autoplot(fit) + theme(legend.position="none")
       
       # Confusion matrix: quantitiative, pos results + neg results / number of test rows

### 2017-11-13 lib/load_patient_metadata.R

@@ -33,28 +33,10 @@ read.patient.info<-function(file, file2=NULL, set='Training') {
   if (grepl('\\.xlsx$', basename(file))) {
     library(readxl)
     
-    patient.info = NULL;  #test.training = NULL
-    #for (sheet in excel_sheets(file)) {
-    #  print(sheet)
-    #for (i in 2:length(getSheets(loadWorkbook(file)))) {
-      #print(i)
-     # if (grepl('Technical Repeats|All_patient_info.txt', sheet)) next
-
-      #if (sheet == 'TrainingTest') {
-      #  test.training = read_excel(file, sheet = sheet)
-      #  next
-      #}
-      
-      #batch_name = gsub(' ', '_', sheet)
-      
-      #ws = read.xlsx2(file, sheetIndex=i, stringsAsFactors=F, header=T)
+    patient.info = NULL;  
       sheet = 'All combined'
       ws = read_excel(file, sheet = sheet)
 
-      #ws = ws[which( ws$`Patient ID` != ""), grep('^Patient|Path ID|Endoscopy Year|Pathology$|Progressor|Plate Index|SLX|cellularity|p53|Number|Batch|Set', colnames(ws), value=T, ignore.case=T)]
-      
-      #head(patient.info)
-      
       slx.cols = grep('SLX', colnames(ws), value=T)
       slx.cols = slx.cols[order(slx.cols)]
       
@@ -77,28 +59,15 @@ read.patient.info<-function(file, file2=NULL, set='Training') {
                  grep('p53 status', colnames(ws),ignore.case=F),
                  grep('Number of', colnames(ws), ignore.case=F),
                  grep('Batch', colnames(ws), ignore.case=F), 
-                 grep('Set', colnames(ws), ignore.case=F)) ]
+                 grep('Set', colnames(ws), ignore.case=F),
+                 grep('Block',colnames(ws),ignore.case=T)) ]
       
-      colnames(ws) = c('Patient', 'Hospital.Research.ID', 'Path.ID','Status','Endoscopy.Year','Pathology','Plate.Index','SLX.ID','Barretts.Cellularity', 'p53.Status', 'Total.Reads', 'Batch.Name', 'Set')
+      colnames(ws) = c('Patient', 'Hospital.Research.ID', 'Path.ID','Status','Endoscopy.Year','Pathology','Plate.Index','SLX.ID','Barretts.Cellularity', 'p53.Status', 'Total.Reads', 'Batch.Name', 'Set', 'Block')
       
[…]

### 2017-11-17 Paper.Rmd
...

@@
               'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor')$variable)),')', sep=''))
-
 # Average cn value
@@
               'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology == 'BE')$variable)),')', sep=''))
-
 globalBECN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology == 'BE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')
@@
 
-pt.status <- c('57'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
-              '76'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))
-genomePtCN.plot = cn.mtn.plot(dfpts, labeller(Patient = pt.status))
+pt.status <- c('Non-Progressor'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
+              'Progressor'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))
+genomePtCN.plot = cn.mtn.plot(dfpts, labeller(Status = pt.status))
 
@@
 
-ptbe.status <- c('57'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
-              '76'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))
-
-genomePtBECN.plot = cn.mtn.plot(dfpts,  labeller(Patient = ptbe.status )) + labs(title='NDBE samples only')
+ptbe.status <- c('Non-Progressor'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
+              'Progressor'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))
+genomePtBECN.plot = cn.mtn.plot(dfpts,  labeller(Status = ptbe.status )) + labs(title='NDBE samples only')
 
@@
 
-cn.mtn.plot(dfpts,  labeller(Patient = ptbe.status ), type='bar') + labs(title='NDBE samples only')
-```
-
-## Individual samples
-```{r}
-
-a = sample(subset(patient.info, Status == 'NP' & Pathology %in% c('BE'))$Samplename, 1)
-b = sample(subset(patient.info, Status == 'P' & Pathology %in% c('BE'))$Samplename, 1)
-
-cn.mtn.plot(subset(globalCNdf, variable %in% c(a) & seg.type != 'arms'), labeller(label_both)) + labs(title="Individual NDBE samples") + theme(strip.text = element_text(size=10, face='bold'))
-cn.mtn.plot(subset(globalCNdf, variable %in% c(b) & seg.type != 'arms'), labeller(label_both)) + labs(title="Individual NDBE samples") + theme(strip.text = element_text(size=10, face='bold'))
-
-cn.mtn.plot(subset(globalCNdf, variable %in% c(a,b) & seg.type != 'arms'), labeller(label_both)) + labs(title="Individual NDBE samples")
-cn.mtn.plot(subset(globalCNdf, variable %in% c(a,b) & seg.type != 'arms'), labeller(label_both), type='bar') + labs(title="Individual NDBE samples")
+#cn.mtn.plot(dfpts,  labeller(Status = ptbe.status ), type='bar') + labs(title='NDBE samples only')
+
+a = sample(subset(patient.info, Status == 'NP' )$Samplename, 1)
+b = sample(subset(patient.info, Status == 'P' )$Samplename, 1)
+
+cn.mtn.plot(subset(globalCNdf, variable %in% c(a,b) & seg.type != 'arms'), labeller(label_both)) + labs(title="Individual samples")
+cn.mtn.plot(subset(globalCNdf, variable %in% c(a,b) & seg.type != 'arms'), labeller(label_both), type='bar') + labs(title="Individual samples")
 
@@
 globalCNdf$g.start = globalCNdf$start
+globalCNdf$g.end = globalCNdf$end
 for (i in 2:nrow(chr.lengths)) {
@@
   globalCNdf[rows, 'g.start'] = globalCNdf[rows, 'start'] + chr.lengths[i-1, 'genome.length']
-}
-
-p1 = ggplot(globalCNdf) + geom_tile(aes(x=g.start, y=variable, fill=value, group=Status)) +
-  scale_fill_gradient2(high='darkgreen', low='darkred', mid='white') +
-  labs(x='', y='Sample') + scale_x_continuous(expand=c(0,0)) +
-  #scale_x_continuous(expand=c(0,0), breaks=chr.lengths$genome.length, labels=chr.lengths$chrom) +
-  geom_vline(data=chr.lengths, aes(xintercept=genome.length), color='grey') +
-  plot.theme + theme(axis.ticks=element_blank(), axis.text=element_blank(),panel.grid.major=element_blank(), legend.position='bottom')
+  globalCNdf[rows, 'g.end'] = globalCNdf[rows, 'end'] + chr.lengths[i-1, 'genome.length']
+}
+
+globalCNdf = arrange(globalCNdf, chr, start)
+globalCNdf$segment = with(globalCNdf, paste(chr, ':', start, '-', end, sep=''))
+
+globalCNdf$chr = factor(globalCNdf$chr, levels=c(1:22), ordered = T)
+globalCNdf$Status = factor(globalCNdf$Status)
+
+p1 = ggplot(subset(globalCNdf, seg.type != 'arms'), aes(segment, variable, fill=value, group=Status)) + geom_tile() +
+  scale_fill_gradient2(high='red3', low='green4', mid='white')  + 
+  facet_grid(~chr, scales = 'free_x', space = 'free') + plot.theme + labs(x='',y='Samples') + 
+  theme(axis.ticks=element_blank(), axis.text=element_blank(), panel.spacing.x = unit(0, 'lines'), legend.position='bottom')
 
 p1leg = get.legend(p1)
-
 p1 = p1 + theme(legend.position = 'none')
 
-p2 = ggplot(globalCNdf, aes(y=variable)) + geom_tile(aes(x=0, fill=Status), show.legend = F) + 
+p2 = ggplot(subset(globalCNdf, seg.type != 'arms'), aes(y=variable, group=Status)) + 
+  geom_tile(aes(x=0, fill=Status), show.legend = F) + 
   labs(x=' ', y='') + scale_fill_manual(values=c('darkblue','darkred')) + scale_x_continuous(expand=c(0,0)) +
@@
 
-grid.arrange(p1, p2, p1leg, ncol=2, nrow=2, widths=c(10,1), heights=c(10,1))
+grid.arrange(p1, p2, p1leg, ncol=2, nrow=2, widths=c(20,1), heights=c(10,1))
 
@@
 
-roc50 = pROC::roc(as.integer(c(preds, np.preds) > 0.5), 
-                c(rep(1, length(preds)), rep(0, length(np.preds))),
-                main='Progressors & Non-progressors, cutoff 0.5')
+roc50 = pROC::roc(as.integer(c(preds, np.preds) > 0.5), c(rep(1, length(preds)), rep(0, length(np.preds))), main='Progressors & Non-progressors, cutoff 0.5')
 roc50$model = 'CN Model 0.5'
@@
 get.roc.stat<-function(roc) {
-  cbind.data.frame('AUC'=pROC::auc(roc), 
-        'Specificity'=round(pROC::coords(roc, 'best')[['specificity']], 2), 
-        'Sensitivity'=round(pROC::coords(roc, 'best')[['sensitivity']], 2),
-        'model'=roc$model)
-}
-
-m = rbind.data.frame(get.roc.stat(p53roc),
+  auc = c( pROC::auc(roc), range(pROC::ci.auc(roc))  )
+
+  sens = c(pROC::coords(roc, 'best')[['sensitivity']], pROC::ci.se(roc,pROC::coords(roc, 'best')[['specificity']], conf.level=0.95))
+  spec = c(pROC::coords(roc, 'best')[['specificity']], pROC::ci.sp(roc,pROC::coords(roc, 'best')[['sensitivity']], conf.level=0.95))
+
+  stat = rbind('AUC'=auc, 'Sensitivity'=sens[c(1,2,4)], 'Specificity'=spec[c(1,2,4)])
+  colnames(stat) = c ('value','CI.min','CI.max')
+  stat = as.data.frame(stat)
+  stat$model = roc$model
+  stat$Measure = rownames(stat)
+  stat
+}
+
+m = rbind.data.frame(
+      get.roc.stat(p53roc),
       get.roc.stat(hgdpathroc),
@@
       get.roc.stat(roc))
-
 m$model = factor(m$model, levels=c('p53 IHC','HGD Only','Dysplasia (HGD & LGD)','CN Model'), ordered = T)
@@
 
-aucs = do.call(rbind, lapply(rocList,function(r) cbind.data.frame('AUC'=r$auc*100,'model'=r$model)))
-
-df = do.call(rbind,
-             lapply(rocList, function(r) 
-              cbind.data.frame('Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),
-                               'model'=r$model)))
+aucs = do.call(rbind, lapply(rocList,function(r) cbind.data.frame('AUC'=r$auc*100,'model'=r$model,'Specificity'=pROC::coords(r,'best')[['specificity']], 'Sensitivity'=pROC::coords(r,'best')[['sensitivity']])))
+
+df = do.call(rbind, lapply(rocList, function(r) 
+              cbind.data.frame('Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'model'=r$model) 
+              ))
+
 ggplot(df, aes(Specificity, Sensitivity,color=model,fill=model)) + geom_line(show.legend = F) + scale_x_reverse() + 
-    geom_label(data=m, aes(label=paste(model,'\nAUC:', sprintf("%.2f",AUC*100), '%', sep='')), nudge_x=.25, color='white', show.legend = F) + 
+    geom_label_repel(data=aucs, aes(label=paste(model,'\nAUC:', sprintf("%.2f",AUC), '%', sep='')), nudge_x=.25, color='white', show.legend = F) + 
   labs(title='ROC Curves', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme
 
-ggplot(melt(m,id.vars = 'model'), aes(model, value*100, group=variable, fill=variable)) + ylim(0,100) +
-  geom_col(position="dodge") + 
-  geom_text(aes(label=paste(round(value,2)*100,'%',sep='') ), position=position_dodge(width=1), vjust=-0.5) + 
-  labs(x='', y='Percent', title='ROC') +
-  scale_fill_manual(values=c('dodgerblue3','firebrick4', 'green4'), name='') + plot.theme + 
-  theme(legend.position = 'bottom', axis.text.x=element_text(angle=45,hjust=1))
+dodge = position_dodge(width=0.9)
+ggplot(melt(m, measure.vars='value'), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Measure)) + 
+  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
+  geom_text(aes(label=paste(round(value,2)*100,'%',sep=''), y=0, vjust=-1 ), position=dodge, color='white')  +
+  scale_fill_manual(values=c('purple3','dodgerblue3','firebrick3'), name='') + plot.theme + 
+  theme(legend.position = 'bottom') + labs(x='', y='Percent', title='ROC Comparison')
 
@@
 ***
+
+## Predictions laid out spatially
+
+```{r, spacial, fig.width=8, fig.height=6}
+transform.by.time<-function(pt, info) {
+  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction')] %>% group_by(PID, Samplename) %>% mutate(
+  'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12
+  )
+
+  pt = arrange(merge(pt, subset(info, Hospital.Research.ID == pt$Hospital.Research.ID[1], select=c('Samplename','Block')), by='Samplename'), Pathology, PID, Endoscopy.Year)
+
+  tb = table(pt$Endoscopy.Year, pt$PID)
+  tb[tb>0] = 1
+  for (yr in names(which(rowSums(tb) > 1))) {
+      tmp = pt[which(pt$Endoscopy.Year == yr),]
+      tmp = tmp %>% group_by(PID) %>% summarise(max(Pathology))
+
+      match = which(pt$Endoscopy.Year == yr & pt$PID == as.character(tmp[which.min(tmp$`max(Pathology)`), 'PID']))
+      pt[match,'months.before.final'] = pt[match,'months.before.final'] + 6
+  }
+
+  pt$nl = (pt$Block-1)*2
+
+  pt$brks = cut(pt$months.before.final, include.lowest = T, ordered_result = T,
+    breaks=c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12),
+    labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))
+
+  tb = table(pt[,c('brks', 'PID')])
+  for (i in 1:nrow(tb)) {
+    pids = tb[i,]
+    yrs = which(pids > 0)
+    if (length(yrs) > 1) {
+    for (j in 2:length(yrs)) 
+      pt[which(pt$PID == names(yrs)[j]), 'nl'] = pt[which(pt$PID == names(yrs)[j]), 'nl'] + 1
+    }
+  }
+  arrange(pt, -Endoscopy.Year, PID)
+}
+
+back = do.call(rbind.data.frame,lapply(pg.samp, transform.by.time, info=patient.info))
+back = merge(back, patient.info[,c('Samplename','p53.Status')])
+
+cuts = seq(0,1,0.1)
+#cuts = quantile(back$Prediction, probs=seq(0,1,1/5))
+
+back$quants = with(back, cut(Prediction, breaks=cuts))
+qt = as.data.frame.matrix(table(back$quants, back$Status))
+qt$quant = rownames(qt)
+
+ft.fun<-function(NP,P) {
+ f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
+ cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
+}
+
+pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'conf'=ifelse(p.value < 0.05, '*', '') )
+pred.confidence[c('r1','r2')] = NA
+for (i in 1:(length(cuts)-1)) 
+    pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)
+
+riskCols = c('red3','orange','green')
+ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), bins=20, show.legend = F) +
+  scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
+  geom_vline(xintercept=pred.confidence$r2, linetype='longdash') +
+  geom_label(data=(pred.confidence %>% rowwise %>% mutate( 'mid'=sum(r1,r2)/2 )), aes(x=mid, y=75, label=paste(round(perc,2),conf,sep=''),color=conf ), show.legend=F) + 
+  scale_color_manual(values=c('grey','black')) + plot.theme + labs(title='Sample predictions', y='n Samples', x='Probability')
+
+pred.confidence$Risk = 'Moderate'
+pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc < .5)) ] = 'Low'
+pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'
+
+back = merge(back, pred.confidence[c('quant','Risk')], by.x='quants',by.y='quant')
+
+ggplot(back, aes(1:nrow(back), Prediction)) + geom_point(aes(color=Risk, shape=Status)) + labs(x='sample', y='model probability', title='Per-sample probabilities')
+
+back = as.data.frame(back %>% group_by(Risk) %>% mutate( 'mean(P)'=mean(Prediction), 'se(P)'=sd(Prediction)/sqrt(length(Prediction)), 'sd(P)'=sd(Prediction) ))
+```
+
+At each quintile are the percentage of progressor samples that were predicted within that range, and the hypergeometric p-value (Fishers) for the significance within that quintile.
+Then our predictions are at best a coin flip in the middle, but highly significant at the outer edges.
+This would also push the prediction for "high" risk up to `r min(subset(pred.confidence, labels == 'High' )$r1)` while "low" risk is below `r max(subset(pred.confidence, labels == 'Low' )$r2)`.
+
+`r pander(as.data.frame(pred.confidence))`
 
@@
 }
-
 collapse.pred.by.path<-function(pts, p53=F, predFUN = NULL) {
@@
 p53byPath = collapse.pred.by.path(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], T, predP53)
+p53.melt = melt(p53byPath, measure.vars=c('prog.ratio', 'nonprog.ratio'), id.vars=c('path', 'total','pred'))
+p53.melt$Status = ifelse(p53.melt$variable == 'prog.ratio', 'Progressor','Non-Progressor')
 
@@
 
-m = rbind(global.melt, np.melt, p53.melt)
+path.m = rbind(global.melt, np.melt, p53.melt)
 
@@
 
-m$pred = factor(m$pred)
-m$final = ifelse(m$path == 'HGD', T, F)
-
-ggplot(m, aes(path, value, group=path, fill=variable)) + ylim(0,100) + facet_wrap(~status) +
+path.m$pred = factor(path.m$pred)
+path.m$final = ifelse(path.m$path == 'HGD', T, F)
+
+ggplot(path.m, aes(path, value, group=path, fill=variable)) + ylim(0,100) + facet_wrap(~status) +
   geom_bar(stat='identity', position=position_stack()) + 
@@
 
+```
+
+Same plot, but mapped to the risk categories instead.
+
+```{r, echo=F, warning=F, message=F, fig.height=8, fig.width=9}
+back$Pathology = as.character(back$Pathology)
+back$Pathology[back$Pathology %in% c('HGD','IMC')] = 'HGD/IMC'
+back$Pathology = factor(back$Pathology, levels=c('BE', 'ID','LGD','HGD/IMC'), ordered = T)
+
+riskPath = rbind( cbind.data.frame( table(subset(back, Status == 'P', select=c('Pathology','Risk'))), 'Status'='Progressor' ), 
+                  cbind.data.frame( table(subset(back, Status == 'NP', select=c('Pathology','Risk'))), 'Status'='Non-Progressor' ))
+
+pathTotal = back %>% group_by(Status, Pathology) %>% summarise( 'nPath'=length(Pathology))
+pathTotal$Status = ifelse(pathTotal$Status == 'P', 'Progressor','Non-Progressor')
+
+riskPath = merge(riskPath, pathTotal)
+riskPath$ratio = with(riskPath, round(Freq/nPath, 2))
+riskPath$Risk = factor(riskPath$Risk, levels=c('High','Moderate','Low'), ordered = T)
+
+p53Path = cbind.data.frame(table(subset(back, Status == 'P', select=c('Pathology','p53.Status'))), 'Status'='p53 IHC')
+p53Path = merge( subset(back, Status == 'P' & !is.na(p53.Status)) %>% group_by(Pathology) %>% summarise( 'nPath'=length(Pathology)), p53Path)
+p53Path$Risk = with(p53Path, ifelse(p53.Status == 1, 'High','Low'))
+p53Path$ratio = with(p53Path, round(Freq/nPath, 2))
+
+riskPath = rbind(riskPath, p53Path[colnames(riskPath)])
+riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor','p53 IHC'), ordered = T)
+
+pathTotal = riskPath %>% group_by(Status, Pathology) %>% summarise('nPath'=unique(nPath))
+
+ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
+  scale_fill_manual(values=c('red','orange2','green3')) + 
+  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), position=position_stack(), size=4) +
+  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5) +
+  facet_grid(~Status) + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
 ```
@@
 ```{r, warning=F, message=F, echo=F, fig.height=8, fig.width=6}
-transform.pt.by.time<-function(pt) {
-  pt = pt %>% group_by(PID, Endoscopy.Year) %>% summarise(
-                  'pred'=sum(as.integer(Prediction > cutoff)), 'samples'=length(PID), 
-                  #'final.endo'= length(which(grepl('HGD|IMC', Pathology))) > 0,
-                  'months.before.final'= (max(pt$Endoscopy.Year) - unique(Endoscopy.Year))*12, 
+endos = subset(back, Status == 'P') %>% group_by(months.before.final, PID) %>% summarise(
+              'pred.endo'=ifelse(length(which(Risk == 'High')) > 0, 1, 0), 
+              'n.endos'=length(unique(PID)), 
+              'n.samples'=length(PID),
+              'pred.samples'=length(which(Risk == 'High')),
               'max.path'=max(Pathology))
-  arrange(pt, -Endoscopy.Year, PID)
-}
-
-back = lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], transform.pt.by.time)
-
-endos = do.call(rbind, lapply(1:max(sapply(back,nrow)), function(i) {
-  df = do.call(rbind, lapply(back, function(x) {
-    if (i > nrow(x)) return(NA)
-    cbind(x[i, c('Endoscopy.Year','pred','samples', 'months.before.final', 'max.path')], i)
-  }))
-  rownames(df) = subset(sum.patient.data, Hospital.Research.ID %in% rownames(df))$Patient
-
-  df$Patient = rownames(df)
-  return(df)
-}))
-
-endos = endos[complete.cases(endos),]
-endos$i = endos$i-1
-
-endos$pred.ratio = round(endos$pred/endos$samples, 2)
-endos = transform(endos, Patient=reorder(Patient, -months.before.final) )
-
+
+endos %>% group_by(months.before.final) %>% summarise(sum(pred.endo)/sum(n.endos))
 endos$brks = cut(endos$months.before.final, include.lowest = T, ordered_result = T,
@@
 
+endos.m = endos %>% group_by(brks, labels) %>% summarise( 
+  'pred.r'=sum(pred.endo)/sum(n.endos),
+  'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos))
+
+rev(x$brks)
+
+ subset(back, Status == 'P') %>% group_by(brks) %>% summarise( 'mean(P)'=mean(Prediction), 'sem(P)'=sd(Prediction)/sqrt(length(brks)) )
+
+se = unique(subset(back, Risk == 'High')$'se(P)')
+
+ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-se)*100, ymax=(pred.r+se)*100, group=brks)) + ylim(0,105) +
+  geom_bar(stat='identity', fill='darkred', alpha=0.8, show.legend = F) + 
+  geom_errorbar(color='grey21') +
+  geom_text(aes(y=3, label=paste(round(pred.r*100), '%', sep='')), color='white') +
+  geom_text(aes(y=pred.r*100-5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
+  scale_x_discrete(labels=endos.m$labels) + 
+  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') +
+  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1))
+
 ## Flip order, 0-1 comes last
-m = endos %>% group_by(brks, labels) %>% summarise( 'pred.r'= length(which(pred > 0))/length(brks), 'n'=length((Patient)), 'n.samp'=sum(samples) )
-
-m$highlight = F
-m[4:6,'highlight'] = T
-m$final = ifelse(m$brks == '[0,1]', T, F)
-
-ggplot(m, aes(brks, pred.r*100, group=brks)) + ylim(0,105) +
-  geom_bar(stat='identity', fill='dodgerblue3', show.legend = F) +     
-  scale_fill_manual(values=c('dodgerblue3', 'yellowgreen')) +
-  geom_text(aes(y=pred.r*100+1.5, label=paste(round(pred.r*100), '%', sep=''))) +
-  geom_text(aes(y=pred.r*100-2, label=paste('n=',n,'\ns=',n.samp,sep='')), color='white') + 
-  scale_x_discrete(labels=m$labels) + 
+se = diff(range(subset(pred.confidence, Risk == 'High')$perc))
+endos.m = endos %>% group_by(brks, labels) %>% summarise( 'pred.r'= length(which(pred > 0))/length(brks),'n.success.endo'=length(which(pred>0)), 'n.endo'=length((Patient)), 
+                                                          'n.successes.samp'=sum(pred), 'n.samp'=sum(samples)) %>% mutate( 
+                                                          'CI.min'=binom::binom.confint(n.success.endo, n.endo, conf.level=0.95, methods='asymptotic')[['lower']],
+                                                          'CI.max'=binom::binom.confint(n.success.endo, n.endo, conf.level=0.95, methods='asymptotic')[['upper']])
+ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=CI.min*100, ymax=ifelse(CI.max>1, 100, CI.max*100), group=brks)) + ylim(0,105) +
+  geom_bar(stat='identity', fill='darkred', alpha=0.8, show.legend = F) + geom_errorbar(width=0.25, color='grey21') +
+  geom_text(aes(y=3, label=paste(round(pred.r*100), '%', sep='')), color='white') +
+  geom_text(aes(y=pred.r*100-5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
+  scale_x_discrete(labels=endos.m$labels) + 
   labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') +
+  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1))
+
+endos.samp = endos %>% group_by(brks, labels) %>% summarise( 'pred.r'=sum(pred)/sum(samples), 'n.successes.samp'=sum(pred), 'n.samp'=sum(samples)) %>% mutate( 
+                                                          'CI.min'=binom::binom.confint(n.successes.samp, n.samp, conf.level=0.95, methods='asymptotic')[['lower']],
+                                                          'CI.max'=binom::binom.confint(n.successes.samp, n.samp, conf.level=0.95, methods='asymptotic')[['upper']])
+ggplot(endos.samp, aes(x=brks, y=pred.r*100, ymin=CI.min*100, ymax=ifelse(CI.max>1, 100, CI.max*100), group=brks)) + ylim(0,105) +
+  geom_bar(stat='identity', fill='darkred', show.legend = F) + geom_errorbar(width=0.25, color='grey21') +
+  geom_text(aes(y=3, label=paste(round(pred.r*100), '%', sep='')), color='white') +
+  geom_text(aes(y=pred.r*100-5, label=paste('s=',n.samp,sep='')), color='white') + 
+  scale_x_discrete(labels=endos.m$labels) + 
+  labs(x='Months prior to endpoint', y='Percent predicted', title='Sample predictions prior to final endoscopy', subtitle='s = #samples') +
   plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1))
@@
 me = endos %>% group_by(Patient) %>% summarise('max.endo' = max(i), 'months' = min(months.before.final))
+
+m = back %>% group_by (Status, brks, Risk) %>% summarise( 'count'=length(Risk) ) %>% mutate('total.by.brk'=sum(count)) %>% mutate('ratio'=round(count/total.by.brk,3)*100)
+m$Status = ifelse(m$Status == 'P', 'Progressor','Non-Progressor')
+ggplot(m, aes(brks, ratio)) + geom_bar(stat='identity', aes(group=Risk, fill=Risk)) + 
+  geom_text(aes(group=Risk, label=paste(signif(ratio, 2),'%',sep='')), position = position_stack(vjust = 0.5)) +
+  geom_text(data=(m %>% group_by(Status, brks) %>% summarise('total'=unique(total.by.brk))), aes(label=paste('n=',total,sep=''), y=101)) +
+  scale_fill_manual(values=c('red','orange2','green3')) + 
+  facet_grid(~Status) + plot.theme + labs(x='Months prior to endpoint', y='Percent predicted',title='Predictions prior to final sample')
+
 ```
@@
 
-## Predictions laid out spatially
-
-```{r, spacial, fig.width=8, fig.height=6}
-transform.by.time<-function(pt) {
-  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction')] %>% group_by(PID, Samplename) %>% mutate(
-  'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12
-  #'nl' = Block
-  #'nl' = str_extract(Path.ID, "B\\w?[0-9]+")
-  )
-
-  pt = merge(pt, subset(patient.info, Hospital.Research.ID == pt$Hospital.Research.ID[1], select=c('Samplename','Block')), by='Samplename')
-
-  pt$nl = (pt$Block-1)*2
-  #pt$nl = (as.numeric(factor(pt$nl, levels=(unique(sort(pt$nl))), ordered = T))-1)*2
-  #table(pt[,c('nl','months.before.final')])
-
-  pt$brks = cut(pt$months.before.final, include.lowest = T, ordered_result = T,
-    breaks=c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12),
-    labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))
-
-  tb = table(pt[,c('brks', 'PID')])
-  for (i in 1:nrow(tb)) {
-    pids = tb[i,]
-    yrs = which(pids > 0)
-    if (length(yrs) > 1) {
-    for (j in 2:length(yrs)) 
-      pt[which(pt$PID == names(yrs)[j]), 'nl'] = pt[which(pt$PID == names(yrs)[j]), 'nl'] + 1
-    }
-  }
-  arrange(pt, -Endoscopy.Year, PID)
-}
-
-back = do.call(rbind.data.frame,lapply(pg.samp, transform.by.time))
-back = merge(back, patient.info[,c('Samplename','p53.Status')])
-
-#cuts = seq(0,1,0.1)
-cuts = quantile(back$Prediction, probs=seq(0,1,1/5))
-back$quants = with(back, cut(Prediction, breaks=cuts))
-qt = as.data.frame.matrix(table(back$quants, back$Status))
-qt$quant = rownames(qt)
-
-ft.fun<-function(NP,P) {
- f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
- #f$p.value
- cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
-}
-
-pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'conf'=ifelse(p.value < 0.05, '*', '') )
-pred.confidence[c('r1','r2')] = NA
-for (i in 1:(length(cuts)-1)) 
-    pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)
-
-ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), bins=20, show.legend = F) +
-  scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
-  geom_vline(xintercept=pred.confidence$r2, linetype='longdash') +
-  geom_label(data=(pred.confidence %>% rowwise %>% mutate( 'mid'=sum(r1,r2)/2 )), aes(x=mid, y=75, label=paste(round(perc,2),conf,sep=''),color=conf ), show.legend=F) + 
-  scale_color_manual(values=c('grey','black')) + plot.theme + labs(title='Sample predictions', y='n Samples', x='Probability') 
-```
-
-At each quintile are the percentage of progressor samples that were predicted within that range, and the hypergeometric p-value (Fishers) for the significance within that quintile.
-Then our predictions are at best a coin flip in the middle, but highly significant at the outer edges.
+`r pander(as.data.frame(pred.confidence))`
+
+The first two quintiles and the last two together account for 80% of our samples.  Only 20% fall into the category "Moderate".
 
 ```{r, fig.width=5, fig.height=10 }
-
-pred.tiles <- function(df) {
-  ggplot(df, aes(brks, nl, fill=Prediction)) + 
-      geom_tile() + scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
-      geom_text(aes(label=labels, color=insig), show.legend=F) + scale_color_manual(values=c('grey','white')) +
+pred.tiles <- function(df, probabilities=F) {
+  p = ggplot(df, aes(brks, nl))
+
+  if (probabilities) {
+    p = p + geom_tile(aes(fill=Prediction)) + 
+      scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)')
+  } else {
+    df$labels = factor(df$labels, levels=c('High','Moderate','Low'))
+    p = p + geom_tile(aes(fill=labels)) + 
+      scale_fill_manual(values=c('red3','orange', 'green3'), name='Progression')
+  }
+  p + geom_text(aes(label=Pathology), color='white', show.legend=F) +
       facet_wrap(~Patient, ncol=1, scales='free_y', labeller=label_both) +
@@
 #back$labels = with(back, ifelse(is.na(p53.Status), as.character(Pathology), labels))
-back$labels = back$Pathology
+#back$labels = back$Pathology
 back$nl = factor(back$nl, ordered=T)
 
-insig = subset(pred.confidence, p.value > 0.05, select=c('quant','r1','r2'))
-
-back$insig = back$quants != insig$quant
+back = merge(back, pred.confidence[c('quant','perc','Risk')], by.x='quants', by.y='quant')
+back$labels = factor(back$Risk, levels=c('High','Moderate','Low'))
 
@@
 
+back$Pathology = as.character(back$Pathology)
+back$Pathology[which(back$Pathology %in% c('HGD','IMC'))] = 'IMC/HGD'
+back$Pathology = factor(back$Pathology, levels=c('BE','ID','LGD','IMC/HGD'), ordered = T)
+
+summary.t = as.data.frame.matrix(with(back, table(Pathology, Risk)))
+summary.t$Pathology = rownames(summary.t)
+
+write.table(
+summary.t %>% group_by(Pathology) %>% mutate( 
+  'High' = paste(High, ' (', round(High/nrow(patient.info),2)*100, '%)', sep=''),
+  'Moderate' = paste(Moderate, ' (', round(Moderate/nrow(patient.info),2)*100, '%)', sep=''),
+  'Low' = paste(Low, ' (', round(Low/nrow(patient.info),2)*100, '%)', sep=''))
+, quote=F, sep='\t', row.names = F)
 ```
@@
 
-plot.cn.chr<-function(df, chrom='17', info=NULL, haz=NULL, genes=NULL) {
+plot.cn.chr<-function(df, chrom='17', info=NULL, haz=NULL, genes=NULL, col=c("orange4","grey90","steelblue3")) {
     df = subset(df, chr == chrom)  
@@
         geom_segment(data=subset(df, seg.type == 'arms'), aes(x=start, xend=end, y=mean.value, yend=mean.value, color=cn), alpha=0.8, show.legend = F) + 
-        scale_color_manual(values=cn.colors) + scale_fill_manual(values=cn.fill)
+        scale_color_manual(values=col) + scale_fill_manual(values=col)
 
@@
 ```{r, echo=F, message=F, warning=F, fig.width=8, fig.height=12}
-
 back = lapply(vpd, transform.pt.by.time)
@@
 
-plot.endo(subset(vpi, Patient %in% subset(sum.validation.info, Status == 'P')$Patient), bin=T) + labs(title="Validation - Progressors")
-
-plot.endo(subset(vpi, Patient %in% subset(sum.validation.info, Status == 'NP')$Patient), bin=T) + labs(title="Validation - Non Progressors")
-```
-
-```{r echo=F, fig.height=12,fig.width=5}
+#plot.endo(subset(vpi, Patient %in% subset(sum.validation.info, Status == 'P')$Patient), bin=T) + labs(title="Validation - Progressors")
+
+#plot.endo(subset(vpi, Patient %in% subset(sum.validation.info, Status == 'NP')$Patient), bin=T) + labs(title="Validation - Non Progressors")
+```
+
+```{r echo=F, fig.height=14,fig.width=10}
 
 vpd2 = lapply(vpd, function(df) {
-  transform.by.time(df)
+  transform.by.time(df, info=validation.patient.info)
 })
@@
 vpd2 = do.call(rbind.data.frame, vpd2)
-
-vpd2[which(vpd2$Patient == 52),'nl'] = c(4,2,0)
-
-vpd2[which(vpd2$Patient %in% c(1, 88)),'nl'] = 0
-vpd2[which(vpd2$Patient %in% c(40,77) & is.na(vpd2$nl)),'nl'] = 0
-vpd2[which(vpd2$Patient == 75 & is.na(vpd2$nl)), 'nl']  = c(0,2,4,6,0,0)
 
@@
 
-ggplot(subset(vpd2, Status == 'P'), aes(brks, nl, fill=Prediction)) + 
-  geom_tile() + geom_text(aes(label=Pathology), color='white', show.legend = F) + 
-  scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
-  facet_wrap(~Patient, ncol=1, scales='free_y', labeller=label_both) +
-  labs(y='Oesophageal Location',x='Months Prior to Endpoint', title='Validation Progressors') +
-  plot.theme + theme(axis.text.x=element_text(angle=45, hjust=1))
-
-ggplot(subset(vpd2, Status == 'NP'), aes(brks, nl, fill=Prediction)) + 
-  geom_tile() + geom_text(aes(label=Pathology), color='white', show.legend = F) + 
-  scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
-  facet_wrap(~Patient, ncol=1, scales='free_y', labeller=label_both) +
-  labs(y='Oesophageal Location',x='Months Prior to Endpoint', title='Validation Non-Progressors') +
-  plot.theme + theme(axis.text.x=element_text(angle=45, hjust=1)) 
+vpd2$quants = with(vpd2, cut(Prediction, breaks=cuts))
+
+vpd2 = merge(vpd2, pred.confidence[c('quant','perc','Risk')], by.x='quants', by.y='quant')
+vpd2$labels = factor(vpd2$Risk, levels=c('High','Moderate','Low'))
+
+grid.arrange(
+  pred.tiles(subset(vpd2, Status == 'P')) + labs(title='Validation Progressors'),
+  pred.tiles(subset(vpd2, Status == 'NP')) + labs(title='Validation Non-Progressors'),
+  ncol=2)
+
+subset(sum.validation.info, Status == 'NP', select=c('Hospital.Research.ID', 'final.endo'))
+
 ```
@@
 
+First note here is that our probability distribution for these 20 patients is somewhat flatter than in the training set.
+Likely a consequence of sample size.
+The labels show the proprotions of progressors that fell into each quintile in the training set.
 ```{r}
+
+vpd2$quants = with(vpd2, cut(Prediction, breaks=cuts))
+#table(vpd2$quants)/nrow(vpd2)
+
+vpd2$Pathology = as.character(vpd2$Pathology)
+vpd2[which(vpd2$Pathology %in% c('HGD','IMC')), 'Pathology'] = 'IMC/HGD'
+
+summary.v = as.data.frame.matrix(with(vpd2, table(Pathology, Risk)))
+summary.v$Pathology = rownames(summary.v)
+
+write.table(
+summary.v %>% group_by(Pathology) %>% mutate( 
+  'High' = paste(High, ' (', round(High/nrow(validation.patient.info),2)*100, '%)', sep=''),
+  'Moderate' = paste(Moderate, ' (', round(Moderate/nrow(validation.patient.info),2)*100, '%)', sep=''),
+  'Low' = paste(Low, ' (', round(Low/nrow(validation.patient.info),2)*100, '%)', sep=''))
+, sep='\t', quote=F, row.names = F)
+
+ggplot(vpd2, aes(Prediction)) + geom_histogram(aes(fill=..x..), bins=20, show.legend = F) +
+  scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
+  geom_vline(xintercept=pred.confidence$r2, linetype='longdash') +
+  geom_label(data=(pred.confidence %>% rowwise %>% mutate( 'mid'=sum(r1,r2)/2 )), aes(x=mid, y=11, label=paste(round(perc,2),conf,sep=''),color=conf ), show.legend=F) + 
+  scale_color_manual(values=c('grey','black')) + plot.theme + labs(title='Validation predictions', y='n Samples', x='Probability') 
+```
+
+```{r, echo=F, warning=F, message=F, fig.height=8, fig.width=5}
+
+clin.eval = vpd2 %>% group_by(Patient, Status, Endoscopy.Year, Pathology, PID, months.before.final, Prediction) %>% summarise( 'High'=length(which(labels=='High')), 'Moderate'=length(which(labels=='Moderate')), 'Low'=length(which(labels=='Low')))
+clin.eval$Endoscopy.Year = as.numeric(as.character(clin.eval$Endoscopy.Year))
 # First and penultimate endoscopy where possible
-clin = lapply(vpd, function(v) {
-  v = arrange(v, Endoscopy.Year)
-  tb = table(v$Endoscopy.Year, v$PID)
-
-  first = subset(v, Endoscopy.Year == rownames(tb)[1] & PID == names(which(tb[1,] > 0))[1])
+
+## Need to select the max prediction in the year for the Risk category
+
+mark.penult<-function(Endoscopy.Year, PID, Prediction) {
+  tb = table(Endoscopy.Year, PID)
   if (nrow(tb) > 2) {
-    penult = subset(v, Endoscopy.Year == rownames(tb)[(nrow(tb)-1)] & PID == names(which(tb[(nrow(tb)-1),] > 0))[1])
+    penult = Prediction == max( Prediction[which(Endoscopy.Year == Endoscopy.Year[which.max(Endoscopy.Year)-1])] )
   } else if (nrow(tb) == 2) {
-    penult = subset(v, Endoscopy.Year == rownames(tb)[(nrow(tb))] & PID == names(which(tb[(nrow(tb)),] > 0))[1])
+    penult = Prediction == max(Prediction[which(Endoscopy.Year == Endoscopy.Year[2])]) 
   } else {
-    penult = first
-  } 
-  cols = c('Hospital.Research.ID','Patient','Status','Path.ID','PID','Endoscopy.Year','Prediction','Samplename')
-  fp = unique(rbind.data.frame(first[cols],penult[cols]))
-
-  fp$conf = cut(fp$Prediction, breaks=cuts, labels=cuts[-1])
-  fp = merge(fp, pred.confidence, by.x='conf', by.y='quant')
-  fp$conf = round(fp$perc, 2)
-  fp$perc = NULL
-  fp$Path.ID = factor(fp$Path.ID)
-
-  fp = merge(fp, validation.patient.info[c('Samplename','Block')])
-  fp$nl = factor( (fp$Block-1)*2, ordered = T)
-  fp$Endoscopy.Year = factor(fp$Endoscopy.Year, ordered = T)
-
-  fp
-})
-
-df = do.call(rbind.data.frame,clin)
-df$conf = factor(df$conf)
-
-ggplot(df, aes(Endoscopy.Year, Prediction, fill=Prediction, color=Prediction)) + 
-  geom_point(aes(size=conf)) + facet_grid(~Patient,  scales='free_x', labeller=label_both) +
-  #scale_size_continuous(breaks=cuts,  labels=cuts) + 
-  scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
-scale_color_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
-   theme(axis.text.x=element_text(angle=45, hjust=1))
-
-df$conf
-pred.confidence
+    penult = Prediction == max(Prediction[which(Endoscopy.Year == Endoscopy.Year[1])])
+  } 
+  penult
+}
+
+clin.eval = clin.eval %>% group_by(Patient) %>% 
+  mutate( 'first'=Prediction == max(Prediction[which(Endoscopy.Year==min(Endoscopy.Year))] ),
+          'penult'=mark.penult(Endoscopy.Year, PID, Prediction) )
+
+clin.eval = subset(clin.eval, first | penult)
+
+m = melt(clin.eval[,c('Patient','Status','Pathology','Endoscopy.Year','Prediction','High','Moderate','Low')], measure.vars=c('High','Moderate','Low'), variable.name='Risk', value.name='Risk.value')
+m = subset(m, Risk.value > 0)
+m$Endoscopy.Year = factor(m$Endoscopy.Year, ordered = T)
+m$Risk = factor(m$Risk, levels=c('Low','Moderate','High'), ordered = T)
+
+ggplot(subset(m, Status == 'P'), aes(Endoscopy.Year, Risk)) + 
+  geom_tile(aes(fill=Risk), color='white') + 
+  scale_fill_manual(values=c('green','orange','red'), drop=F) +
+  geom_text(aes(label=Pathology), color='white') +
+  facet_wrap(~Patient, scales='free_x',labeller=label_both) + plot.theme + theme(legend.position = 'none')
+
+ggplot(subset(m, Status == 'NP'), aes(Endoscopy.Year, Risk)) + 
+  geom_tile(aes(fill=Risk), color='white') + 
+  scale_fill_manual(values=c('green','orange','red'), drop=F) +
+  geom_text(aes(label=Pathology), color='white') +
+  facet_wrap(~Patient, scales='free_x', labeller=label_both) + plot.theme + theme(legend.position = 'none')
 
@@
 
-```{r clonality, echo=F, message=F, warning=F}
+```{r clonality, echo=F, message=F, warning=F, eval=F}
 
@@
 pt = 'AH0254'
-subset(sum.patient.data, Status == 'P')
-
-subset(sum.patient.data, Patient == 34)
+subset(sum.patient.data, Hospital.Research.ID == pt)
 

### 2017-11-27 Paper.Rmd
...

@@
 source('lib/cv-pt-glm.R')
-
-get.legend<-function(a.gplot){
-  tmp <- ggplot_gtable(ggplot_build(a.gplot))
-  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
-  legend <- tmp$grobs[[leg]]
-  return(legend)
-}
-
-plot.theme = theme(text=element_text(size=14, face='bold'), panel.background=element_blank(), strip.background =element_rect(fill="lightcyan2"),  
-                   strip.text = element_text(size=14, face='bold'), 
-                   axis.line=element_line(color='black'), panel.grid.major=element_line(color='grey90'),
-                   panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing = unit(0.1, 'lines')  )
+source('lib/common_plots.R')
 
 data = '~/Data/Ellie'
-tile.w='5e+06_arms'
+tile.w=c(5e06, 'arms')
 #tile.w=5e+06
@@
 nrow(sum.patient.data)
-
-# Missing some of the samples as there were some normals that we removed from analysis
-# for (pt in names(patient.data)) {
-#   patient.data[[pt]]$info = patient.data[[pt]]$info[patient.data[[pt]]$info$Samplename %in% colnames(patient.data[[pt]]$seg.vals)[-(1:5)],]
-# }
-
-# for (pt in names(validation.patient.data)) {
-#   validation.patient.data[[pt]]$info = validation.patient.data[[pt]]$info[validation.patient.data[[pt]]$info$Samplename %in% colnames(validation.patient.data[[pt]]$seg.vals)[-(1:5)],]
-# }
 
@@
 
-```{r tile, echo=F, message=F, warning=F, include=F}
-tileFile = paste('tile_patients_All_',as.character(tile.w),'.Rdata',sep='')
-tile.files = grep(tileFile, analysis.files, value=T, fixed=T)
-load(tile.files[grep('tile', tile.files)], verbose=T)
-
-mergedDf = mergedDf[grep('X|Y',rownames(mergedDf), invert=T),]
-
-mergedDfValidation = mergedDf[,validation.patient.info$Samplename]
-dim(mergedDfValidation)
-
-mergedDf = mergedDf[, intersect(colnames(mergedDf), patient.info$Samplename)]
-dim(mergedDf)
-
-cache.dir = paste(data, 'Analysis',sub('\\+', '', tile.w),   sep='/')
-if (includeDemo)
-  cache.dir = paste(cache.dir, '_with-demo', sep='')
-
-```
-
-`r pander(mergedDf[sort(sample(nrow(mergedDf), 10)), sample(ncol(mergedDf), 10) ], caption='Example rows/columns from the merged data')`
-
 ## y( Progressors (1) vs Non (0) )
@@
 
-# sort in label order
-#subset(patient.info, Samplename %in% setdiff(colnames(mergedDf), names(labels))[-c(1:3)])
-if (length(setdiff(colnames(mergedDf), names(labels))) > 3)
-  warning("Labels vector is missing samples")
+cache.dir = paste(data, 'Analysis', sub('\\+','',paste(tile.w, collapse='_')), sep='/')
 
@@
 ```{r echo=F, warning=F, message=F, eval=T, fig.width=12, fig.height=8}
-
-cn.mtn.plot<-function(df, label, col=c("orange4","grey90","steelblue3"), type='seg') {
-  p = ggplot(df, aes(x=chr.length)) + facet_grid(Status~chr, scales='free_x', space='free_x', labeller=label)
-
-  if (type == 'seg') {
-    p = p + geom_segment(aes(x=start, xend=end, y=mean.value, yend=mean.value, color=cn), show.legend=T, size=1) + scale_color_manual(values=col) 
-  }
-  if (type == 'bar') {
-    p = p + geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=cn), show.legend=T) + scale_fill_manual(values=col) 
-  }
-
-  p + labs(x='', y='Mean adjusted segmentation value', title='All samples') +
-    theme(text=element_text(face='bold', size=14), axis.text.x=element_blank(), legend.position='bottom', 
-          panel.grid = element_blank(), panel.background = element_blank(),strip.background =element_rect(fill="lightcyan2"),  
-          panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0, 'lines') )
-}
 
@@
 
-locs = do.call(rbind.data.frame, lapply(rownames(mergedDf), function(x) unlist(strsplit(x,':|-'))))
-colnames(locs) = c('chr','start','end')
-locs[] = lapply(locs[], function(x) as.numeric(as.character(x)))
-
 globalCNdf = t(dysplasia.df[,grep('^\\d+:',colnames(dysplasia.df))])
@@
 
-cn = function(x) {
-  str = 'neutral'
-  if(x <= -1) {
-    str = 'loss'
-  } else if (x >= 1) {
-    str = 'gain'
-  }
-  return(str)
-}
 globalCNdf$cn = factor(sapply(globalCNdf$value, cn), levels=c('loss','neutral','gain'), ordered = T)
@@
 ```{r model_perf, message=F, warning=F, echo=F, fig.width=6, fig.height=6}
-model.performance<-function(p1se, coeffs ) {
-  performance = do.call(rbind.data.frame, p1se)
-  performance$alpha = rownames(performance)
-
-  if (grepl('chr', performance$alpha[1])) {
-    performance$alpha = factor(performance$alpha, levels=c(paste('chr',c(1:22), sep=''), setdiff(rownames(performance), paste('chr',c(1:22),sep=''))) )
-  }
-
-  ## check how often the feature(s) is selected at that lamda in each split. "stability selection"
-  coef.stable = lapply( coeffs, function(cf) {
-    sort(rowSums(cf[,-1]), decreasing=T)
-  })
-  coef.stable = coef.stable[names(which(sapply(coef.stable, length) > 0))]
-  coef.stable = lapply(coef.stable, function(cf) cf/(folds*splits))
-
-  performance = cbind(performance, do.call(rbind,  lapply(coef.stable, function(x) {
-    cbind('n.Coef'=length(x), '25%'=length(which(x>=.25)), 
-          '50%'=length(which(x>=.5)), '75%'=length(which(x>=.75 )), 
-          '100%'=length(which(x==1)) )
-  })))
-
-  p = ggplot(performance, aes(alpha, mean, color=(alpha != 0.9))) + 
-    geom_point(size=2) +
-    geom_text( aes(label=round(mean, 3)), hjust=-0.5) +
-    geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme), width=0.3, size=1) + 
-    geom_text(aes(label=n.Coef), nudge_y=0.03, fontface='bold') +
-    geom_text(aes(label=paste('(',`75%`,')',sep='')), nudge_y=0.02) + 
-    scale_colour_manual(values=c('red', 'grey39')) + plot.theme + theme(legend.position = 'none') +
-    labs(x='Elasticnet penalty value: ridge <-> lasso', y='Model classification at lambda-1se')
-
-  cfs = as.data.frame(matrix(data=0,nrow=length(unique(names(table(unlist(lapply(coef.stable, names)))))), ncol=length(names(coef.stable)), 
-                           dimnames=list( unique(names(table(unlist(lapply(coef.stable, names))))), names(coef.stable) )))
-  for (i in names(coef.stable)) 
-    cfs[intersect(rownames(cfs), names(coef.stable[[i]])), i] = 1
-  return(list('performance'=performance, 'stable.coeffients'=coef.stable, 'cf.per.feature'=cfs,'plot'=p))
-}
 
@@
 
-write.table(ch, quote=F, sep='\t', row.names = F, col.names = T, file=paste(as.character(tile.w),'_features.txt'))
+write.table(ch, quote=F, sep='\t', row.names = F, col.names = T, file=paste( paste(tile.w, collapse='_')  ,'_features.txt'))
 ```
@@
 
-oddPt = names(pg.samp)[which.max(performance.at.1se)]
-subset(sum.patient.data, Hospital.Research.ID == oddPt)
-
-pg.samp[[oddPt]]$Prediction
+#oddPt = names(pg.samp)[which.max(performance.at.1se)]
+#subset(sum.patient.data, Hospital.Research.ID == oddPt)
+
+#pg.samp[[oddPt]]$Prediction
 
 # Not the pt with the most samples or endos
-sum.patient.data[which.max(sum.patient.data$total.samples),]
-
-```
-
-```{r, eval=F}
+#sum.patient.data[which.max(sum.patient.data$total.samples),]
+```
+
+```{r, eval=F, echo=F, message=F}
 image(cov(dysplasia.df[names(which(labels == 1)),c(1:(ncol(dysplasia.df)-3))], use='p'), breaks=seq(-1,1,l=21), col=colorRampPalette(c('red','black','green'))(20),  useRaster = T)
@@
 ```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=4, fig.width=4}
-roc.plot <- function(roc,title="") {
-  df = cbind.data.frame('Specificity'=rev(roc$specificities), 'Sensitivity'=rev(roc$sensitivities))
-
-  ggplot(df, aes(Specificity, Sensitivity)) + geom_line() + scale_x_reverse() + 
-    geom_label(data=as.data.frame(t(round(pROC::coords(roc, "best"),2))), 
-               aes(x=specificity, y=sensitivity, 
-                   label=paste(' (FPR ', sprintf('%.1f',(1-specificity)*100), '%, TPR ', sprintf('%.1f',sensitivity*100),'%)\nAUC:', sprintf("%.2f",roc$auc*100), '%', sep='')), nudge_x=.25) +
-  labs(title=title, x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme
-}
-
 preds = unlist(lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], function(df) df$Prediction))
@@
 roc50$model = 'CN Model 0.5'
-
-roc.plot(roc50,'Probability threshold = 0.5')
 
@@
 
-roc.plot(roc, 'Probability threshold calculated by ROC')
+grid.arrange(roc.plot(roc50,'Probability threshold = 0.5'), roc.plot(roc, 'Probability threshold calculated by ROC'), ncol=2)
+
+roc.plot(roc, 'Copy number model')
 
@@
 
-```{r p53roc, echo=F, fig.height=4, fig.width=4}
+```{r p53roc, echo=F, fig.height=2, fig.width=2}
 p53samples = subset(patient.info, !is.na(p53.Status), select=c('Samplename','p53.Status','Status'))
@@
 roc.plot(p53roc, 'p53 IHC')
-
-pROC::auc(p53roc)
-
 ```
@@
 
-roc.plot(pathroc, 'Pathology (LGD,HGD,IMC)')
-
-pROC::auc(pathroc)
-
 hgdpathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('HGD','IMC'), 1, 0) )
@@
 
-roc.plot(hgdpathroc, 'Pathology (HGD,IMC)')
-
-pROC::auc(hgdpathroc)
+grid.arrange(roc.plot(pathroc, 'Pathology (LGD,HGD,IMC)'), roc.plot(hgdpathroc, 'Pathology (HGD,IMC)'), ncol=2)
 ```
@@
 ```{r echo=F, fig.height=6, fig.width=8}
-
 get.roc.stat<-function(roc) {
@@
   theme(legend.position = 'bottom') + labs(x='', y='Percent', title='ROC Comparison') 
-
 ```
@@
 
-## Predictions laid out spatially
-
-```{r, spacial, fig.width=8, fig.height=6}
+## Prediction probability distribution
+
+```{r back, echo=F, message=F, warning=F}
 transform.by.time<-function(pt, info) {
@@
 
-  tb = table(pt$Endoscopy.Year, pt$PID)
+  tb = with(pt, table(Endoscopy.Year, PID))
   tb[tb>0] = 1
@@
 
-  pt$brks = cut(pt$months.before.final, include.lowest = T, ordered_result = T,
-    breaks=c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12),
-    labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))
-
-  tb = table(pt[,c('brks', 'PID')])
+  tb = table(pt[,c('months.before.final', 'PID')])
   for (i in 1:nrow(tb)) {
@@
 
+sq = c(0, seq(1,132,12),max(back$months.before.final))
+back$brks = cut(back$months.before.final, 
+                  breaks=sq, 
+                 labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-c(length(sq)-1,length(sq))], '>120'),
+                  include.lowest = T, ordered_result = T)
+back$brks = factor(back$brks, rev(levels(back$brks)))
+
+back$Pathology = as.character(back$Pathology)
+back$Pathology[which(back$Pathology %in% c('HGD','IMC'))] = 'IMC/HGD'
+back$Pathology = factor(back$Pathology, levels=c('BE','ID','LGD','IMC/HGD'), ordered = T)
+
+```
+
+The labels in each decile are the ratio of progessor samples that have probabilities within that region. The deciles that are greyed out show insigificant enrichment for the NP/P ratio.
+
+```{r spacial, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
 cuts = seq(0,1,0.1)
@@
 
-pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'conf'=ifelse(p.value < 0.05, '*', '') )
+qt = cbind(qt, back %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction), 'se'=(sd(Prediction)/sqrt(length(Prediction)))*1.96 ))
+
+ggplot(qt,aes(quants, mn, ymin=mn-se, ymax=mn+se)) + geom_point() + geom_errorbar() + labs(x='Deciles', y='Mean probability', title='CI around probabilities by decile')
+
+pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 
+  'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )
 pred.confidence[c('r1','r2')] = NA
@@
 
-riskCols = c('red3','orange','green')
-ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), bins=20, show.legend = F) +
-  scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
-  geom_vline(xintercept=pred.confidence$r2, linetype='longdash') +
-  geom_label(data=(pred.confidence %>% rowwise %>% mutate( 'mid'=sum(r1,r2)/2 )), aes(x=mid, y=75, label=paste(round(perc,2),conf,sep=''),color=conf ), show.legend=F) + 
-  scale_color_manual(values=c('grey','black')) + plot.theme + labs(title='Sample predictions', y='n Samples', x='Probability')
-
 pred.confidence$Risk = 'Moderate'
@@
 
+annotation.table = arrange(pred.confidence %>% group_by(Risk) %>% summarise('n(NP)'=sum(NP), 'n(P)'=sum(P), 'p.value'=format.pval(mean(p.value), 3), 'min'=min(r1), 'max'=max(r2)), min)
+
+#annotation.table = as.data.frame(annotation.table)
+#rownames(annotation.table) = annotation.table$Risk
+#annotation.table$Risk = NULL
+
+tt3 <- ttheme_minimal(
+  core=list(bg_params = list(fill = c('green3','orange2','red3'), col=NA, alpha=0.2),
+            fg_params=list(fontface=3)),
+  colhead=list(fg_params=list(col="black", fontface=4L)),
+  rowhead=list(fg_params=list(col=c('red3','green3', 'orange2'), fontface=3L)),
+  padding=unit(c(5,5),'mm'))
+
+grid.arrange( tableGrob(annotation.table, rows=NULL, theme=tt3) )
+
+riskCols = c('red3','orange','green3')
+grid.arrange(
+  ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), bins=20, show.legend = F) +
+    scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
+    #geom_vline(xintercept=pred.confidence$r2, linetype='longdash', color='grey39') +
+    #geom_label(data=(pred.confidence %>% rowwise %>% mutate( 'mid'=sum(r1,r2)/2 )), aes(x=mid, y=75,label=paste(round(perc,2),conf,sep=''),color=conf ), show.legend=F) + scale_color_manual(values=c('grey','black')) +
+    geom_vline(xintercept=annotation.table$max, linetype='longdash', color='grey39') +
+    geom_text(data=annotation.table %>% group_by(Risk, min,max) %>% summarise( 'samp.perc'= signif((sum(`n(NP)`, `n(P)`)/nrow(patient.info))*100, 3) ), aes(x=min+(max-min)/2, y=5, label=paste(samp.perc, '%',sep=''))) +
+    plot.theme + labs(title='Sample predictions', y='n Samples', x='Probability')
+
+  ,tableGrob(annotation.table, rows=NULL, theme=tt3),
+  nrow=2, heights=c(8,2)
+  )
+
 back = merge(back, pred.confidence[c('quant','Risk')], by.x='quants',by.y='quant')
-
-ggplot(back, aes(1:nrow(back), Prediction)) + geom_point(aes(color=Risk, shape=Status)) + labs(x='sample', y='model probability', title='Per-sample probabilities')
-
-back = as.data.frame(back %>% group_by(Risk) %>% mutate( 'mean(P)'=mean(Prediction), 'se(P)'=sd(Prediction)/sqrt(length(Prediction)), 'sd(P)'=sd(Prediction) ))
+back = back %>% group_by(Risk) %>% mutate('se(Risk)'=(sd(Prediction)/sqrt(length(Risk)))*1.96 )
+
+ggplot(back, aes(1:nrow(back), Prediction, color=Risk)) + 
+  geom_point(aes(color=Risk, shape=Status)) + geom_errorbar(aes(ymin=Prediction-`se(Risk)`, ymax=Prediction+`se(Risk)`)) +
+  geom_hline(data=annotation.table, aes(yintercept=max), color='grey') + plot.theme +
+  labs(x='sample', y='model probability', title='Per-sample probabilities')
 ```
@@
 
-```{r jacknife, echo=F, warning=F, message=F}
+Could just as easily get a jacknife CI per patient
+
+```{r jacknife, echo=T, warning=F, message=F, fig.height=5, fig.width=6}
 
@@
   names(lcf) = rownames( loo.coefs[[pt]] )
-
   sharedcfs = intersect(names(lcf), ch$label)
-
   coefs.per.pt[sharedcfs, pt] = lcf[sharedcfs]
@@
 
-ch$jack.se = apply(coefs.per.pt, 1, function(x) {
+# per patient
+# jack.se = apply(coefs.per.pt, 2, function(x) {
+#   jk = jackknife(x, var)
+#   jk$jack.se
+# })
+# js = cbind.data.frame(performance.at.1se, jack.se)
+# ggplot(js, aes(rownames(js), performance.at.1se, color=(performance.at.1se > mean(performance.at.1se)+sd(performance.at.1se) | performance.at.1se < mean(performance.at.1se)-sd(performance.at.1se)  ))) + ylim(0.6,1) + geom_point() +
+#   geom_errorbar(aes(ymin=performance.at.1se-jack.se, ymax=performance.at.1se+jack.se)) + theme(legend.position = 'none', axis.text.x = element_text(angle=90, hjust=1))
+
+# in this case jack.se is the SE of the variance 
+# per coefficient
+ch$jack.se = jack =  apply(coefs.per.pt, 1, function(x) {
   jk = jackknife(x, var)
@@
 })
-
-ch$adj = ifelse(ch$coef<0,-1,1)
 ggplot(ch, aes(reorder(label, coef), coef)) + geom_point() + 
-  geom_errorbar(aes(ymin=(abs(coef)-jack.se)*adj, ymax=(abs(coef)+jack.se)*adj)) + coord_flip() +
-  plot.theme + labs(x='', title='Jackknifed coefficient CI??')
-
-```
+  geom_errorbar(aes(ymin=coef-jack.se, ymax=coef+jack.se)) + 
+  coord_flip() +
+  plot.theme + labs(x='', title='Jackknifed coefficient CI')
+
+```
+
+Effectively it looks to me like my coefficients have fairly tight CI if the SE of the jackknife variance is used.
 
@@
 
-```{r pathology, echo=F, message=F, warning=F, fig.height=8, fig.width=6}
-model.pred<-function(pt, p53=F, pred = function(rows) length(which(rows$Prediction > cutoff))) {
-  final.endo = subset(pt, Endoscopy.Year == max(Endoscopy.Year))
-  final.endo = final.endo[nrow(final.endo),]
-
-  h = subset(pt, Pathology %in% c('IMC','HGD'))
-  l = subset(pt, Pathology == 'LGD' & PID != final.endo$PID)
-  id = subset(pt, Pathology == 'ID' & PID != final.endo$PID)
-  be = subset(pt, Pathology %in% c('BE') & PID != final.endo$PID)
-
-  if (p53) {
-    return(cbind('HGD'=pred(h), 'total.h'=nrow(h),#   nrow(subset(h, !is.na(p53.Status))), 
-                  'LGD'=pred(l), 'total.l'=nrow(subset(l, !is.na(p53.Status))), 
-                  'ID'=pred(id), 'total.id'=nrow(subset(id, !is.na(p53.Status))), 
-                  'NDBE'=pred(be), 'total.be'=nrow(subset(be, !is.na(p53.Status)))))
-
-  } else {
-    return( cbind('HGD'=pred(h), 'total.h'=nrow(h), 
-                  'LGD'=pred(l), 'total.l'=nrow(l), 
-                  'ID'=pred(id), 'total.id'=nrow(id),
-                  'NDBE'=pred(be), 'total.be'=nrow(be)) )
-    }
-}
-collapse.pred.by.path<-function(pts, p53=F, predFUN = NULL) {
-  if (!is.null(predFUN))
-    by.path = do.call(rbind.data.frame, lapply(pts,model.pred,p53, predFUN) )
-  else
-    by.path = do.call(rbind.data.frame, lapply(pts,model.pred,p53) )
-
-  sums.by.path = colSums(by.path)
-
-  path = as.data.frame(matrix(sums.by.path, byrow=T, nrow=4, dimnames=list(c('HGD','LGD','ID','NDBE'),c('pred', 'total'))))
-  path$prog.ratio = (path$pred/path$total)*100
-  path$nonprog.ratio = 100-path$prog.ratio
-  path$path = rownames(path)
-
-  path$path = factor(path$path, levels=c('HGD','LGD','ID','NDBE')) 
-  path$pred = ifelse(p53, 'p53','model')
-  return(path)
-}
-
-global.path = collapse.pred.by.path(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID])
-global.pathNP = collapse.pred.by.path(pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID])
-
-predP53 = function(rows) length(which(rows$p53.Status == 1 ))
-p53byPath = collapse.pred.by.path(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], T, predP53)
-p53.melt = melt(p53byPath, measure.vars=c('prog.ratio', 'nonprog.ratio'), id.vars=c('path', 'total','pred'))
-p53.melt$Status = ifelse(p53.melt$variable == 'prog.ratio', 'Progressor','Non-Progressor')
-
-global.melt = melt(global.path, measure.vars=c('prog.ratio', 'nonprog.ratio'), id.vars=c('path', 'total','pred'))
-p53.melt = melt(p53byPath, measure.vars=c('prog.ratio', 'nonprog.ratio'), id.vars=c('path', 'total','pred'))
-np.melt = melt(global.pathNP, measure.vars=c('prog.ratio', 'nonprog.ratio'),id.vars=c('path','total','pred'))
-
-global.melt$pred='model'
-global.melt$status='Progressor'
-np.melt$status='Non-Progressor'
-
-#p53.melt$status='p53 IHC'
-p53.melt = subset(p53.melt, variable == 'prog.ratio')
-p53.melt$variable = paste('p53',p53.melt$variable,sep='')
-p53.melt$status = 'Progressor'
-
-path.m = rbind(global.melt, np.melt, p53.melt)
-
-#m$status = factor(m$status, levels=c('Progressor', 'p53 IHC','Non-Progressor'), ordered = T)
-
-path.m$pred = factor(path.m$pred)
-path.m$final = ifelse(path.m$path == 'HGD', T, F)
-
-ggplot(path.m, aes(path, value, group=path, fill=variable)) + ylim(0,100) + facet_wrap(~status) +
-  geom_bar(stat='identity', position=position_stack()) + 
-  scale_fill_manual(values=c('firebrick4','green4','slateblue4'), name='', labels=c('Pred. progressed', 'Pred. Not progressed','p53 IHC'))  +
-  geom_text(aes(vjust=ifelse(value > 12, 2, 1),label=paste(paste(round(value, 1),'%',sep=''),paste('n=', total, sep=''),sep='\n')),
-            position=position_stack(),color='white',show.legend=F) +
-  geom_bar(data=p53.melt, aes(path, value, group=path), fill='slateblue4',stat='identity',position = position_dodge()) +
-  geom_text(data=p53.melt, aes(path,value,vjust=ifelse(value > 12, 2, 1),label=paste(paste(round(value, 1),'%',sep=''),paste('n=', total, sep=''),sep='\n')), position=position_dodge(),color='white',  show.legend=F) +
-  labs(title='Samples predicted by pathology', x='Pathology', y='Percent', subtitle='Non-HGD exclude the final endoscopy, p53 excludes NA samples') +
-  plot.theme + theme(text = element_text(size=16, face='bold'), legend.position = 'bottom')
-
-```
-
-Same plot, but mapped to the risk categories instead.
-
 ```{r, echo=F, warning=F, message=F, fig.height=8, fig.width=9}
-back$Pathology = as.character(back$Pathology)
-back$Pathology[back$Pathology %in% c('HGD','IMC')] = 'HGD/IMC'
-back$Pathology = factor(back$Pathology, levels=c('BE', 'ID','LGD','HGD/IMC'), ordered = T)
-
 riskPath = rbind( cbind.data.frame( table(subset(back, Status == 'P', select=c('Pathology','Risk'))), 'Status'='Progressor' ), 
@@
 
-riskPath = rbind(riskPath, p53Path[colnames(riskPath)])
-riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor','p53 IHC'), ordered = T)
+#riskPath = rbind(riskPath, p53Path[colnames(riskPath)])
+#riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor','p53 IHC'), ordered = T)
+riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor'), ordered = T)
 
@@
 ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
-  scale_fill_manual(values=c('red','orange2','green3')) + 
+  scale_fill_manual(values=c('red2','orange','green3')) + 
   geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), position=position_stack(), size=4) +
@@
   facet_grid(~Status) + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
+
+ggplot(p53Path, aes(Pathology, ratio*100, group=Pathology, fill=Risk)) + geom_bar(stat='identity', position=position_stack()) + 
+  scale_fill_manual(values=c('red2','green3')) + 
+  geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101)) +
+  geom_text(aes(label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5) +
+  plot.theme + labs(title='Samples predicted by aberrant p53 IHC', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
+
 ```
@@
 ```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=6}
-# All samples
-all.predictions = sapply(pg.samp, function(pt) {
-  if (unique(pt$Status) == 'NP') {
-    as.integer(pt$Prediction < cutoff)
-  } else {
-    as.integer(pt$Prediction > cutoff)
-  }
-})
-
-all.predictions = do.call(rbind, lapply(all.predictions, function(x) {
-  cbind.data.frame('correctly.predicted'=sum(x), 'total.samples'=length(x))
-}))
-
-all.predictions$Status = ifelse (rownames(all.predictions) %in% subset(sum.patient.data, Status == 'P')$Hospital.Research.ID, 'P', 'NP')
-
-rownames(all.predictions) = subset(sum.patient.data, Hospital.Research.ID %in% rownames(all.predictions))$Patient
-
-all.predictions$pred.ratio = with(all.predictions, correctly.predicted/total.samples)
-all.predictions$Patient = rownames(all.predictions)
-all.predictions = transform(all.predictions, Patient=reorder(Patient, pred.ratio) )
-
-p = subset(all.predictions, Status == 'P')
-np = subset(all.predictions, Status == 'NP')
-
-m = melt(all.predictions, id.vars=c('Patient','Status'), measure.vars=c('pred.ratio'))
-
-mp = m %>% group_by(Status) %>% summarise('all'=round(length(which(value==1))/length(Status), 2), 'all.c'=length(Status)-length(which(value==1))/2, 
-                                          'some'=round(length(which(value > 0 & value < 1))/length(Status), 2), 'some.c'=length(which(value > 0 & value < 1)),
-                                          'none'=round(length(which(value == 0))/length(Status), 2), 'none.c'=1.5)
-
-mpT = m %>% group_by(Status) %>% summarise('100%'=round(length(which(value==1))/length(Status), 2), 
-                                          '>25%'=round(length(which(value >= 0.25 & value < 1 ))/length(Status), 2), 
-                                          '0%'=round(length(which(value == 0 ))/length(Status), 2))
-mpT$Status = c('Non-Progressor','Progressor')
-
-grid.arrange(
-  ggplot(m, aes(Patient, value, group=Status)) + facet_grid(~Status,  scales='free_x') + 
-    geom_bar(stat='identity', position='identity',aes(fill=Status)) + 
-    scale_fill_manual(values=c('darkblue','darkgreen')) +
-    geom_label(data=mp, aes(x=all.c, y=0.5, label=paste(all*100,'%',sep=''), fill=Status), color='white', fontface='bold', size=6) +
-    geom_label(data=mp, aes(x=some.c, y=0.35, label=paste(some*100,'%',sep=''), fill=Status), color='white', fontface='bold', size=6) +
-    geom_label(data=mp, aes(x=none.c, y=0.2, label=paste(none*100,'%',sep=''), fill=Status), color='white', fontface='bold', size=6) +
-    labs(y='Ratio of predicted samples') + plot.theme + theme(legend.position='none', axis.text.x=element_text(angle=90) ), 
-  tableGrob(mpT, rows=NULL), nrow=2, as.table=T, heights=c(4,1) )
+all.preds = back %>% group_by(Status, Patient) %>% summarise( 'Low'=sum(length(which(Risk == "Low"))),
+                                          'Moderate'=sum(length(which(Risk == "Moderate"))),
+                                          'High'=sum(length(which(Risk == "High"))),
+                                          'n.samples'=length(Patient) )
+
+all.preds = melt(all.preds, id.vars=c('Status','Patient', 'n.samples'))
+head(all.preds)
+ggplot(all.preds, aes(Patient, value/n.samples, group=variable, fill=variable)) + 
+  geom_bar(stat='identity', position = position_stack()) +
+  scale_fill_manual(values=c('green3','orange2','red')) + 
+  geom_text(aes(y=1.01, label=n.samples)) +
+  facet_wrap(~Status, scales='free_x') + labs(y='Sample ratios', title='Per patient ratio of predictions') + 
+  plot.theme
 
@@
 
-```{r, warning=F, message=F, echo=F, fig.height=8, fig.width=6}
-endos = subset(back, Status == 'P') %>% group_by(months.before.final, PID) %>% summarise(
+```{r enods, warning=F, message=F, echo=F, fig.height=8, fig.width=6}
+endos = back %>% group_by(Status, months.before.final, brks, PID) %>% summarise(
               'pred.endo'=ifelse(length(which(Risk == 'High')) > 0, 1, 0), 
@@
 
-endos %>% group_by(months.before.final) %>% summarise(sum(pred.endo)/sum(n.endos))
+sq = c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12)
 endos$brks = cut(endos$months.before.final, include.lowest = T, ordered_result = T,
-    breaks=c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12))
+     breaks=sq,
+     labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))
 endos$brks = with(endos, factor(brks, rev(levels(brks)))) # invert order
 
-endos$labels = gsub('\\[|\\(|\\]', '',endos$brks)
-endos$labels = sapply(strsplit( endos$labels, ',' ), function(x) paste(sort(as.numeric(x), decreasing=T), collapse='-' ))
-endos$labels[ grep('-120', endos$labels) ] = '>120'
-endos$labels[ grep('1-0', endos$labels) ] = 'Endpoint'
-
-endos.m = endos %>% group_by(brks, labels) %>% summarise( 
+endos.m = subset(endos, Status == 'P') %>% group_by(brks) %>% summarise( 
   'pred.r'=sum(pred.endo)/sum(n.endos),
@@
 
-rev(x$brks)
-
- subset(back, Status == 'P') %>% group_by(brks) %>% summarise( 'mean(P)'=mean(Prediction), 'sem(P)'=sd(Prediction)/sqrt(length(brks)) )
-
-se = unique(subset(back, Risk == 'High')$'se(P)')
-
-ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-se)*100, ymax=(pred.r+se)*100, group=brks)) + ylim(0,105) +
+tmp = subset(back, Status == 'P')
+tmp$brks = cut(tmp$months.before.final, include.lowest = T, ordered_result = T,breaks=sq,
+     labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))
+tmp = tmp %>% group_by(brks) %>% summarise( 'mean(P)'=mean(Prediction),'sem(P)'=sd(Prediction)/sqrt(length(brks)) )
+
+endos.m = arrange(merge(endos.m,tmp), brks)
+
+ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) + ylim(0,105) +
   geom_bar(stat='identity', fill='darkred', alpha=0.8, show.legend = F) + 
-  geom_errorbar(color='grey21') +
+  geom_errorbar(color='grey21', width=0.2) +
   geom_text(aes(y=3, label=paste(round(pred.r*100), '%', sep='')), color='white') +
   geom_text(aes(y=pred.r*100-5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
-  scale_x_discrete(labels=endos.m$labels) + 
   labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') +
@@
 
-## Flip order, 0-1 comes last
-se = diff(range(subset(pred.confidence, Risk == 'High')$perc))
-endos.m = endos %>% group_by(brks, labels) %>% summarise( 'pred.r'= length(which(pred > 0))/length(brks),'n.success.endo'=length(which(pred>0)), 'n.endo'=length((Patient)), 
-                                                          'n.successes.samp'=sum(pred), 'n.samp'=sum(samples)) %>% mutate( 
-                                                          'CI.min'=binom::binom.confint(n.success.endo, n.endo, conf.level=0.95, methods='asymptotic')[['lower']],
-                                                          'CI.max'=binom::binom.confint(n.success.endo, n.endo, conf.level=0.95, methods='asymptotic')[['upper']])
-ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=CI.min*100, ymax=ifelse(CI.max>1, 100, CI.max*100), group=brks)) + ylim(0,105) +
-  geom_bar(stat='identity', fill='darkred', alpha=0.8, show.legend = F) + geom_errorbar(width=0.25, color='grey21') +
-  geom_text(aes(y=3, label=paste(round(pred.r*100), '%', sep='')), color='white') +
-  geom_text(aes(y=pred.r*100-5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
-  scale_x_discrete(labels=endos.m$labels) + 
-  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') +
-  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1))
-
-endos.samp = endos %>% group_by(brks, labels) %>% summarise( 'pred.r'=sum(pred)/sum(samples), 'n.successes.samp'=sum(pred), 'n.samp'=sum(samples)) %>% mutate( 
-                                                          'CI.min'=binom::binom.confint(n.successes.samp, n.samp, conf.level=0.95, methods='asymptotic')[['lower']],
-                                                          'CI.max'=binom::binom.confint(n.successes.samp, n.samp, conf.level=0.95, methods='asymptotic')[['upper']])
-ggplot(endos.samp, aes(x=brks, y=pred.r*100, ymin=CI.min*100, ymax=ifelse(CI.max>1, 100, CI.max*100), group=brks)) + ylim(0,105) +
-  geom_bar(stat='identity', fill='darkred', show.legend = F) + geom_errorbar(width=0.25, color='grey21') +
-  geom_text(aes(y=3, label=paste(round(pred.r*100), '%', sep='')), color='white') +
-  geom_text(aes(y=pred.r*100-5, label=paste('s=',n.samp,sep='')), color='white') + 
-  scale_x_discrete(labels=endos.m$labels) + 
-  labs(x='Months prior to endpoint', y='Percent predicted', title='Sample predictions prior to final endoscopy', subtitle='s = #samples') +
-  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1))
-
-plot.endo<-function(df, minM=NULL, maxM=NULL, bin=F) {
-  if (is.null(minM)) minM = min(df$months.before.final)
-  if (is.null(maxM)) maxM = max(df$months.before.final)
-
-  colors = c('green4','red4')
-
-  df = arrange(df %>% group_by(Patient) %>% mutate('overall.pred'=sum(as.integer(pred > 0))/length(Patient)), Patient)
-
-  df = transform(df, Patient=reorder(Patient, -overall.pred) )
-
-  if (bin) {
-    df$endo.pred = ifelse(df$pred > 0, 'Pos','Neg')
-    ggplot(df, aes(months.before.final, endo.pred, group=1, color=endo.pred)) + facet_grid(Patient ~.) + 
-      scale_x_continuous(breaks=seq(minM, maxM, by = 12)) +
-      geom_line( color='grey39') + geom_point(aes(color=endo.pred)) +
-      geom_label_repel(aes(label=max.path, fill=endo.pred), color='white', fontface='bold', size=4, label.size=0, label.r=unit(0.25, 'lines'), label.padding=unit(0.15, 'lines'), show.legend=F) + 
-      scale_fill_manual(values=colors) + 
-      scale_color_manual(values=colors, name='Endoscopy Prediction') +
-      labs(x='Months before diagnosis', y='Predictive Endoscopy') + plot.theme + 
-      theme(panel.border = element_rect(color = "grey50", fill = NA, size = 1), legend.position='bottom', strip.background =element_rect(fill="lightcyan2"), panel.grid.major.x=element_blank(), axis.text.x=element_text(angle=45, hjust=1))
-
-  } else {
-  ggplot(df, aes(months.before.final, pred.ratio, group=1, color=pred.ratio, fill=pred.ratio)) + facet_grid(Patient ~.) +
-      ylim(0,1) + geom_abline(slope=0, intercept=seq(.25,.75,.25), color='grey', linetype='dashed') +
-      geom_col(position='dodge', aes(y=pred.ratio), color='white') +# geom_line() +
-      scale_x_continuous(breaks=seq(minM, maxM, by = 12)) + 
-      geom_label(aes(y=ifelse(pred.ratio==1, 0.75, pred.ratio), label=max.path, fill=pred.ratio), color='white', fontface='bold', size=4, label.size=0, label.r=unit(0.25, 'lines'), label.padding=unit(0.15, 'lines'), nudge_y=0.1) +
-      labs(x='Months before diagnosis', y='Sample prediction ratio') + 
-      scale_fill_continuous(high='orangered4', low='green4', name='Sample Prediction Ratio') + plot.theme + 
-      theme(panel.border = element_rect(color = "grey50", fill = NA, size = 1), legend.position='bottom',strip.background =element_rect(fill="lightcyan2"), panel.grid.major.x=element_blank(), axis.text.x=element_text(angle=45, hjust=1))
-  }
-}
-
-endos$months.before.final = endos$months.before.final*-1
-me = endos %>% group_by(Patient) %>% summarise('max.endo' = max(i), 'months' = min(months.before.final))
-
-m = back %>% group_by (Status, brks, Risk) %>% summarise( 'count'=length(Risk) ) %>% mutate('total.by.brk'=sum(count)) %>% mutate('ratio'=round(count/total.by.brk,3)*100)
-m$Status = ifelse(m$Status == 'P', 'Progressor','Non-Progressor')
-ggplot(m, aes(brks, ratio)) + geom_bar(stat='identity', aes(group=Risk, fill=Risk)) + 
-  geom_text(aes(group=Risk, label=paste(signif(ratio, 2),'%',sep='')), position = position_stack(vjust = 0.5)) +
-  geom_text(data=(m %>% group_by(Status, brks) %>% summarise('total'=unique(total.by.brk))), aes(label=paste('n=',total,sep=''), y=101)) +
-  scale_fill_manual(values=c('red','orange2','green3')) + 
-  facet_grid(~Status) + plot.theme + labs(x='Months prior to endpoint', y='Percent predicted',title='Predictions prior to final sample')
-
-```
-
+#endos$months.before.final = endos$months.before.final*-1
+#me = endos %>% group_by(Patient) %>% summarise('max.endo' = max(i), 'months' = min(months.before.final))
+```
+
+```{r, echo=F, fig.height=8, fig.width=6, eval=F}
 ## Progressor predictions per patient
-```{r, echo=F, fig.height=8, fig.width=6}
 #plot.endo( subset(endos, Patient %in% subset(me, months <= -72)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(72/12, "years or more follow-up"))
@@
 
-```{r, echo=F, fig.height=8, fig.width=6}
+```{r, echo=F, fig.height=8, fig.width=6, eval=F}
 # plot.endo( subset(endos, Patient %in% subset(me, months <= -48 & months > -72)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(48/12, "-",72/12,  "years or more follow-up"))
@@
 
+```{r, echo=F, fig.height=12, fig.width=8, eval=F}
 ## Non-progressor predictions per patient
-```{r, echo=F, fig.height=12, fig.width=8}
-nonp.back = lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID],transform.pt.by.time)
-
-nonp.endos = do.call(rbind, lapply(1:max(sapply(nonp.back,nrow)), function(i) {
-  df = do.call(rbind, lapply(nonp.back, function(x) {
-    if (i > nrow(x)) return(NA)
-    cbind(x[i, c('Endoscopy.Year','pred','samples', 'months.before.final', 'max.path')], i)
-  }))
-  rownames(df) = subset(sum.patient.data, Hospital.Research.ID %in% rownames(df))$Patient
-
-  df$Patient = rownames(df)
-
-  return(df)
-}))
-nonp.endos = nonp.endos[complete.cases(nonp.endos),]
-#nonp.endos$i = nonp.endos$i-1
-
-nonp.endos$pred.ratio = round(nonp.endos$pred/nonp.endos$samples, 2)
-nonp.endos = transform(nonp.endos, Patient=reorder(Patient, -months.before.final) ) 
-nonp.endos$months.before.final = nonp.endos$months.before.final*-1
-
-me = nonp.endos %>% group_by(Patient) %>% summarise('max.endo' = max(i), 'months' = min(months.before.final))
+# nonp.back = lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID],transform.pt.by.time)
+# 
+# nonp.endos = do.call(rbind, lapply(1:max(sapply(nonp.back,nrow)), function(i) {
+#   df = do.call(rbind, lapply(nonp.back, function(x) {
+#     if (i > nrow(x)) return(NA)
+#     cbind(x[i, c('Endoscopy.Year','pred','samples', 'months.before.final', 'max.path')], i)
+#   }))
+#   rownames(df) = subset(sum.patient.data, Hospital.Research.ID %in% rownames(df))$Patient
+# 
+#   df$Patient = rownames(df)
+# 
+#   return(df)
+# }))
+# nonp.endos = nonp.endos[complete.cases(nonp.endos),]
+# #nonp.endos$i = nonp.endos$i-1
+# 
+# nonp.endos$pred.ratio = round(nonp.endos$pred/nonp.endos$samples, 2)
+# nonp.endos = transform(nonp.endos, Patient=reorder(Patient, -months.before.final) ) 
+# nonp.endos$months.before.final = nonp.endos$months.before.final*-1
+# 
+# me = nonp.endos %>% group_by(Patient) %>% summarise('max.endo' = max(i), 'months' = min(months.before.final))
 
@@
 
-```{r, fig.width=5, fig.height=10 }
-pred.tiles <- function(df, probabilities=F) {
-  p = ggplot(df, aes(brks, nl))
-
-  if (probabilities) {
-    p = p + geom_tile(aes(fill=Prediction)) + 
-      scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)')
-  } else {
-    df$labels = factor(df$labels, levels=c('High','Moderate','Low'))
-    p = p + geom_tile(aes(fill=labels)) + 
-      scale_fill_manual(values=c('red3','orange', 'green3'), name='Progression')
-  }
-  p + geom_text(aes(label=Pathology), color='white', show.legend=F) +
-      facet_wrap(~Patient, ncol=1, scales='free_y', labeller=label_both) +
-      labs(y='Oesophageal Location',x='Months Prior to Endpoint', title='') +
-      plot.theme + theme(axis.text.x=element_text(angle=45, hjust=1)) 
-}
-
-back$brks = with(back, factor(brks, rev(levels(brks)))) # invert order
-#back$labels = paste(back$Pathology, ifelse(back$p53.Status == 0, '--', '**') ,sep='')
-#back$labels = with(back, ifelse(is.na(p53.Status), as.character(Pathology), labels))
-#back$labels = back$Pathology
+```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=10 }
 back$nl = factor(back$nl, ordered=T)
-
-back = merge(back, pred.confidence[c('quant','perc','Risk')], by.x='quants', by.y='quant')
 back$labels = factor(back$Risk, levels=c('High','Moderate','Low'))
 
-ids = c(13,20,34,56,60,48)
-pred.tiles(subset(back, Patient %in% ids & Status == 'P')) + labs(title='Progressors')
-
-ids = c(15,16,21,3,45,19)
-pred.tiles(subset(back, Patient %in% ids & Status == 'NP')) + labs(title='Non-Progressors')
-
-back$Pathology = as.character(back$Pathology)
-back$Pathology[which(back$Pathology %in% c('HGD','IMC'))] = 'IMC/HGD'
-back$Pathology = factor(back$Pathology, levels=c('BE','ID','LGD','IMC/HGD'), ordered = T)
-
-summary.t = as.data.frame.matrix(with(back, table(Pathology, Risk)))
+ pred.tiles(subset(back, Patient %in% c(13,20,34,56,60,48) & Status == 'P'), nrow=1) + labs(title='Progressors')
+
+grid.arrange( pred.tiles(subset(back, Patient %in% c(13,20,34,56,60,48) & Status == 'P'), ncol=1) + labs(title='Progressors') + theme(legend.position = 'none'),
+              pred.tiles(subset(back, Patient %in% c(15,16,21,3,45,19) & Status == 'NP'), ncol=1) + labs(title='Non-Progressors'), ncol=2 )
+
+table.path.risk<-function(df) {
+  summary.t = as.data.frame.matrix(with(df, table(Pathology, Risk)))
   summary.t$Pathology = rownames(summary.t)
 
-write.table(
-summary.t %>% group_by(Pathology) %>% mutate( 
-  'High' = paste(High, ' (', round(High/nrow(patient.info),2)*100, '%)', sep=''),
-  'Moderate' = paste(Moderate, ' (', round(Moderate/nrow(patient.info),2)*100, '%)', sep=''),
-  'Low' = paste(Low, ' (', round(Low/nrow(patient.info),2)*100, '%)', sep=''))
-, quote=F, sep='\t', row.names = F)
-```
+  summary.t = summary.t %>% group_by(Pathology) %>% mutate( 'High.r'=round(High/sum(High,Low,Moderate),3), 
+                                                'Low.r'=round(Low/sum(High,Low,Moderate),3),
+                                                'Moderate.r'=round(Moderate/sum(High,Low,Moderate),3))
+  for (g in unique(df$Risk)) {
+    for (i in 1:nrow(summary.t)) {
+      summary.t[i,g] = paste(summary.t[i,g], ' (', summary.t[i,paste(g,'.r',sep='')]*100, '%)',sep='')
+    }
+  }
+  summary.t[,c('Pathology',levels(df$labels))]
+}
+
+summary.t = table.path.risk(back)
+
+write.table(summary.t, quote=F, sep='\t', row.names = F)
+```
+
+`r pander(as.data.frame(summary.t))`
 
 ```{r, echo=F}
-p = pred.tiles(subset(back, Status == 'NP')) + labs(title='Non-Progressors')
-ggsave('npg_heatmap.png', plot = p, scale = 1, width = 8, height = 52, units = "in", limitsize = F, dpi = 300)
-
-p = pred.tiles(subset(back, Status == 'P')) + labs(title='Progressors')
-ggsave('pg_heatmap.png', plot = p, scale = 1, width = 8, height = 52, units = "in", limitsize = F, dpi = 300)
+
+p = pred.tiles(subset(back, Status == 'NP'), nrow=2) + labs(title='Non-Progressors')
+ggsave('npg_heatmap.png', plot = p, scale = 1, width = 52, height = 8, units = "in", limitsize = F, dpi = 300)
+
+p = pred.tiles(subset(back, Status == 'P'), nrow=2) + labs(title='Progressors')
+ggsave('pg_heatmap.png', plot = p, scale = 1, width = 52, height = 8, units = "in", limitsize = F, dpi = 300)
 
@@
 ## p arm
-```{r}
+```{r, echo=F}
 
@@
 One way of looking at the p53 prediction (e.g. Adam Bass) is to see how predictive the p-arm loss in chr17 is.
-```{r, echo=F, warning=F, message=F, fig.height=16, fig.width=18}
+```{r, echo=F, warning=F, message=F, fig.height=8, fig.width=8}
 ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
@@
 
-plot.cn.chr<-function(df, chrom='17', info=NULL, haz=NULL, genes=NULL, col=c("orange4","grey90","steelblue3")) {
-    df = subset(df, chr == chrom)  
-    p = ggplot(df, aes(x=chr.length)) + facet_grid(Status~chr, scales='free_x', labeller=labeller(chr=label_both)) +
-        geom_rect(data=subset(df, seg.type != 'arms'), aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=cn), alpha=0.5, show.legend = T) + 
-        geom_segment(data=subset(df, seg.type == 'arms'), aes(x=start, xend=end, y=mean.value, yend=mean.value, color=cn), alpha=0.8, show.legend = F) + 
-        scale_color_manual(values=col) + scale_fill_manual(values=col)
-
-    if (!is.null(info)) {
-      info = info[which(info$chrom == chrom),]
-      p = p + geom_vline(data=info, aes(xintercept=chr.cent), color='grey39', linetype='longdash' ) +
-              geom_label(data=info, aes(x=chr.cent-cent.gap*3, y=max(df$mean.value)), label='p-arm')
-    }
-
-    if (!is.null(genes)) {
-      genes = subset(genes, chr == chrom)
-      if (nrow(genes) > 0) p = p + geom_point(data=genes, aes(x=start,  y=value), color='darkblue', size=3) +
-              geom_text_repel(data=genes, aes(x=start, y=value, label=Gene.Symbol), color='darkblue') 
-    }
-    if (!is.null(haz)) {
-      haz = subset(haz, chr == chrom)
-      if (nrow(haz) > 0 & length(grep('loss',unique(haz$gl))) > 0 ) {
-          p = p + geom_rect(data=subset(haz, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F, size=1) 
-      } 
-      if (nrow(haz) > 0 & length(grep('gain',unique(haz$gl))) > 0) {
-          p = p + geom_rect(data=subset(haz, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F, size=1) 
-      }
-    }
-
-    p + theme(text=element_text(face='bold', size=14), axis.text.x=element_blank(), legend.position='bottom', 
-          panel.grid = element_blank(), panel.background = element_blank(),strip.background =element_rect(fill="lightcyan2"),  
-          panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0.1, 'lines') ) + labs(x='Chromosomes', y='Mean adjusted segmentation value')
-}
-
-plot.cn.chr(subset(globalCNdf), chrom='17', chr.lengths, genes=genes, haz=chh) + labs(
+plot.cn.chr(subset(globalCNdf, seg.type != 'arms'), chrom='17', chr.lengths, genes=genes, haz=chh) + labs(
 title="Segment Values Samples", subtitle=paste('NonP:Prog - ', paste(table(patient.info$Status), collapse=':')))
+```
+
+```{r echo=F, warning=F, message=F}
 
@@
 plot.cn.chr(globalCNdf, chrom='16', chr.lengths, haz=chh, genes=genes)
+
+plot.cn.chr(globalCNdf, chrom='9', chr.lengths, haz=chh, genes=genes)
 
@@
 
-as.data.frame(chGR[unique(queryHits(ov))])
-length(genesGR[unique(subjectHits(ov))]$Gene.Symbol)
+#as.data.frame(chGR[unique(queryHits(ov))])
+#length(genesGR[unique(subjectHits(ov))]$Gene.Symbol)
 write.table(genesGR[unique(subjectHits(ov))]$Gene.Symbol, quote=F, row.names=F, sep='\t')
+
+pander(as.data.frame(genesGR[unique(subjectHits(ov))])[,c(1:3,6,7)])
 
@@
 
-```{r validation, message=F, warning=F, echo=F, fig.height=8, fig.width=8}
-
-pander(table(validation.patient.info[c('Status','p53.Status')]))
-
-validation.df = t(mergedDfValidation[, grep('^D', colnames(mergedDfValidation))])
-validation.df = apply(validation.df, 2, unit.var)
-validation.df = score.cx(validation.patient.data, validation.df)
+```{r val_setup, echo=F, message=F, warning=F, include=F}
+tileFile = paste('tile_patients_All_',as.character(tile.w),'.Rdata',sep='')
+tile.files = sapply(tileFile, function(f) grep(f, analysis.files, value=T,fixed=T))
+
+fauxlabel = rep(0, length(validation.patient.info$Samplename) )
+names(fauxlabel) = validation.patient.info$Samplename
+validation.df = set.up.data(tile.files, validation.patient.info$Samplename, fauxlabel)
+dim(validation.df)
 
@@
 
+```
+
+```{r validation_globalplot, message=F, warning=F, echo=F, fig.height=8, fig.width=8}
+
+pander(table(validation.patient.info[c('Status','p53.Status')]))
+
+valCNdf = t(validation.df[,grep('^\\d+:',colnames(validation.df))])
+
+locs = do.call(rbind.data.frame, lapply(rownames(valCNdf), function(x) unlist(strsplit(x,':|-'))))
+colnames(locs) = c('chr','start','end')
+locs[] = lapply(locs[], function(x) as.numeric(as.character(x)))
+
+valCNdf = cbind.data.frame(locs, valCNdf)
+valCNdf = melt(valCNdf, id.vars=colnames(locs))
+valCNdf$chr = factor(valCNdf$chr, levels=c(1:22), ordered = T)
+
+valCNdf$seg.type = '5MB'
+valCNdf[with(valCNdf, which( start %in% arm.loc$start & end %in% arm.loc$end)), 'seg.type'] = 'arms'
+valCNdf = merge(valCNdf, chr.lengths[,c('chrom','chr.length')], by.x='chr', by.y='chrom')
+
+valCNdf = merge(valCNdf, validation.patient.info[,c('Hospital.Research.ID','Patient','Samplename','Pathology','Status')], by.x='variable', by.y='Samplename')
+valCNdf$Status = ifelse(valCNdf$Status == 'P', 'Progressor', 'Non-Progressor')
+
+valCNdf$cn = factor(sapply(valCNdf$value, cn), levels=c('loss','neutral','gain'), ordered = T)
+
+n.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(valCNdf, Status == 'Progressor')$variable)), ')', sep=''),
+              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(valCNdf, Status == 'Non-Progressor')$variable)),')', sep=''))
+# Average cn value
+valCNdf = valCNdf %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
+valCNdf$mean.value = valCNdf$value # but not going to use it 
+cn.mtn.plot(subset(valCNdf, seg.type != 'arms'), labeller(Status=n.status, chr=label_value))
+
+head(valCNdf)
+
+df = rbind(cbind(valCNdf, 'set'=rep('Val', nrow(valCNdf))), cbind(globalCNdf, 'set'=rep('Train', nrow(globalCNdf)) ) )
+subset(df, seg.type != 'arms') %>% group_by(set,Status) %>% summarise( var(mean.value), mean(mean.value), max(mean.value), min(mean.value) )
+
+```
+
+```{r validation, message=F, warning=F, echo=F, fig.height=8, fig.width=8}
 file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
@@
 
-v.predictions = sapply(vpd, function(pt) {
-  if (unique(pt$Status) == 'NP') {
-    as.integer(pt$Prediction < cutoff)
-  } else {
-    as.integer(pt$Prediction > cutoff)
-  }
-})
-
-names(v.predictions) = subset(sum.validation.info, Hospital.Research.ID %in% names(v.predictions))$Patient
-v.predictions = do.call(rbind, lapply(v.predictions, function(x) {
-  cbind.data.frame('correctly.predicted'=sum(x,na.rm=T), 'total.samples'=length(x))
-}))
-
-v.predictions$Status = ifelse (rownames(v.predictions) %in% subset(sum.validation.info, Status == 'P')$Patient, 'P', 'NP')
-v.predictions$pred.ratio = with(v.predictions, correctly.predicted/total.samples)
-v.predictions$Patient = rownames(v.predictions)
-v.predictions = transform(v.predictions, Patient=reorder(Patient, pred.ratio) )
-
-p = arrange(subset(v.predictions, Status == 'P'), -pred.ratio)
-np = arrange(subset(v.predictions, Status == 'NP'), -pred.ratio)
-
-m = melt(v.predictions, id.vars=c('Patient','Status'), measure.vars=c('pred.ratio'))
-
-mp = m %>% group_by(Status) %>% summarise('all'=round(length(which(value==1))/length(Status), 2), 'all.c'=length(Status)-length(which(value==1))/2, 
-                                          'some'=round(length(which(value > 0 & value < 1))/length(Status), 2), 'some.c'=length(which(value > 0 & value < 1)),
-                                          'none'=round(length(which(value == 0))/length(Status), 2), 'none.c'=1.5)
-
-mpT = m %>% group_by(Status) %>% summarise('100%'=round(length(which(value==1))/length(Status), 2), 
-                                          '>50%'=round(length(which(value >= 0.5 & value < 1 ))/length(Status), 2), 
-                                          '25-50%'=round(length(which(value >= 0.25 & value < 0.5 ))/length(Status), 2), 
-                                          '0%'=round(length(which(value == 0 ))/length(Status), 2))
-mpT$Status = c('Non-Progressor','Progressor')
-
-ggplot(m, aes(Patient, value, group=Status)) + facet_grid(~Status,  scales='free_x') + 
-  geom_bar(stat='identity', position='identity',aes(fill=Status)) + 
-  scale_fill_manual(values=c('darkblue','darkgreen')) +
-  geom_label(data=mp, aes(x=all.c, y=0.5, label=paste(all*100,'%',sep=''), fill=Status), color='white', fontface='bold', size=6) +
-  geom_label(data=mp, aes(x=some.c, y=0.35, label=paste(some*100,'%',sep=''), fill=Status), color='white', fontface='bold', size=6) +
-  geom_label(data=mp, aes(x=none.c, y=0.2, label=paste(none*100,'%',sep=''), fill=Status), color='white', fontface='bold', size=6) +
-  labs(y='Ratio of predicted samples', title='Model predictions on validation dataset') + plot.theme + theme(legend.position='none', axis.text.x=element_text(angle=90) )
-```
-
-```{r, echo=F, message=F, warning=F, fig.width=8, fig.height=12}
-back = lapply(vpd, transform.pt.by.time)
-vpi = do.call(rbind, lapply(1:max(sapply(back,nrow)), function(i) {
-  df = do.call(rbind, lapply(back, function(x) {
-    if (i > nrow(x)) return(NA)
-    cbind(x[i, c('Endoscopy.Year','pred','samples', 'months.before.final', 'max.path')], i)
-  }))
-  rownames(df) = subset(sum.validation.info, Hospital.Research.ID %in% rownames(df))$Patient
-  df$Patient = rownames(df)
-  return(df)
-}))
-
-vpi = vpi[complete.cases(vpi),]
-vpi$pred.ratio = round(vpi$pred/vpi$samples, 2)
-vpi = transform(vpi, Patient=reorder(Patient, -months.before.final) ) 
-vpi$months.before.final = vpi$months.before.final*-1
-
-me = vpi %>% group_by(Patient) %>% summarise('max.endo' = max(i), 'months' = min(months.before.final))
-
-#plot.endo(subset(vpi, Patient %in% subset(sum.validation.info, Status == 'P')$Patient), bin=T) + labs(title="Validation - Progressors")
-
-#plot.endo(subset(vpi, Patient %in% subset(sum.validation.info, Status == 'NP')$Patient), bin=T) + labs(title="Validation - Non Progressors")
+back.v = do.call(rbind.data.frame,lapply(vpd, transform.by.time, info=validation.patient.info))
+back.v = merge(back.v, validation.patient.info[,c('Samplename','p53.Status')])
+
+sq = c(0, seq(1,132,12),max(back$months.before.final))
+back.v$brks = cut(back.v$months.before.final, 
+                  breaks=sq, 
+                 labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-c(length(sq)-1,length(sq))], '>120'),
+                  include.lowest = T, ordered_result = T)
+back.v$brks = with(back.v, factor(brks, rev(levels(brks))))
+
+back.v$Pathology = as.character(back.v$Pathology)
+back.v$Pathology[which(back.v$Pathology %in% c('HGD','IMC'))] = 'IMC/HGD'
+back.v$Pathology = factor(back.v$Pathology, levels=c('BE','ID','LGD','IMC/HGD'), ordered = T)
+
 ```
@@
 ```{r echo=F, fig.height=14,fig.width=10}
-
-vpd2 = lapply(vpd, function(df) {
-  transform.by.time(df, info=validation.patient.info)
-})
-
-vpd2 = do.call(rbind.data.frame, vpd2)
-
-vpd2$Endoscopy.Year = factor(vpd2$Endoscopy.Year, ordered = T)
-vpd2$brks = with(vpd2, factor(brks, rev(levels(brks)))) # invert order
-
-vpd2$quants = with(vpd2, cut(Prediction, breaks=cuts))
-
-vpd2 = merge(vpd2, pred.confidence[c('quant','perc','Risk')], by.x='quants', by.y='quant')
-vpd2$labels = factor(vpd2$Risk, levels=c('High','Moderate','Low'))
+back.v$quants = with(back.v, cut(Prediction, breaks=cuts))
+back.v = merge(back.v, pred.confidence[c('quant','perc','Risk')], by.x='quants', by.y='quant')
+back.v$labels = factor(back.v$Risk, levels=c('High','Moderate','Low'))
+back.v$nl = factor(back.v$nl, ordered = T)
+
+sq = c(0, seq(1, max(back.v$months.before.final), 12))
+back.v$brks = cut(back.v$months.before.final, 
+                  breaks=sq, 
+                 labels=c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-length(sq)],
+                  include.lowest = T, ordered_result = T)
+back.v$brks = with(back.v, factor(brks, levels=rev(levels(brks))))
+
+head(back.v)
+
+m = melt(back.v %>% group_by(Patient, Status) %>% summarise( 'Low'=length(which(Risk == 'Low'))/length(Patient),
+                                            'Moderate'=length(which(Risk == 'Moderate'))/length(Patient),
+                                            'High'=length(which(Risk == 'High'))/length(Patient),
+                                            'n.samples'=length(Patient)), id.vars=c('Patient','Status','n.samples'))
+
+ggplot(m, aes(Patient,  value*100, group=variable, fill=variable)) + geom_bar(stat='identity', position=position_stack()) +
+  geom_text(aes(label=n.samples, y=101)) + scale_fill_manual(values=c('green3','orange2','red2')) +
+  facet_grid(~Status, scales='free_x') + labs(y='Percent of samples', title='Validation patient set')
 
 grid.arrange(
-  pred.tiles(subset(vpd2, Status == 'P')) + labs(title='Validation Progressors'),
-  pred.tiles(subset(vpd2, Status == 'NP')) + labs(title='Validation Non-Progressors'),
+  pred.tiles(subset(back.v, Status == 'P')) + labs(title='Validation Progressors') ,
+  pred.tiles(subset(back.v, Status == 'NP')) + labs(title='Validation Non-Progressors'),
   ncol=2)
 
-subset(sum.validation.info, Status == 'NP', select=c('Hospital.Research.ID', 'final.endo'))
-
 ```
@@
 ```{r, echo=F, message=F, warning=F, fig.width=8, fig.height=8}
-
-global.path = collapse.pred.by.path(vpd[subset(sum.validation.info, Status == 'P')$Hospital.Research.ID])
-global.pathP53 = collapse.pred.by.path(vpd[subset(sum.validation.info, Status == 'P')$Hospital.Research.ID], T, predFUN = predP53)
-
-global.pathNP = collapse.pred.by.path(vpd[subset(sum.validation.info, Status == 'NP')$Hospital.Research.ID])
-
-pm = melt(global.path, measure.vars=c('prog.ratio','nonprog.ratio'), id.vars=c('path', 'total','pred'))
-pm$status = 'Progressor'
-
-pm53m = melt(global.pathP53, measure.vars=c('prog.ratio','nonprog.ratio'), id.vars=c('path', 'total','pred'))
-#pm53m$status = 'p53 IHC'
-pm53m = subset(pm53m, variable == 'prog.ratio')
-pm53m$variable = paste('p53',pm53m$variable,sep='')
-pm53m$status = 'Progressor'
-
-npm = melt(global.pathNP, measure.vars=c('prog.ratio','nonprog.ratio'), id.vars=c('path', 'total','pred'))
-npm$status = 'Non-Progressor'
-
-m = rbind(pm, npm, pm53m)
-m$pred = factor(m$pred)
-
-#m$status = factor(m$status, levels=c('Non-Progressor','Progressor','p53 IHC'), ordered = T)
-
-ggplot(m, aes(path, value, group=path, fill=variable)) + ylim(0,100) + facet_wrap(~status) +
-  geom_bar(stat='identity', position=position_stack()) + 
-  scale_fill_manual(values=c('firebrick4','green4', 'slateblue4'),name='', labels=c('Pred. progressed','Pred. Not progressed', 'p53 prediction'))  +
-  geom_text(aes(vjust=ifelse(value > 12, 2, 1),
-                 label=paste(paste(round(value, 1),'%',sep=''),
-                                        paste('n=', total, sep=''),sep='\n')), position=position_stack(),color='white',  show.legend=F) +
-  geom_bar(data=pm53m, aes(path, value, group=path), fill='slateblue4', stat='identity') +
-  geom_text(data=pm53m, aes(vjust=ifelse(value > 12, 2, 1),
-                label=paste(paste(round(value, 1),'%',sep=''),
-                                       paste('n=', total, sep=''),sep='\n')), position=position_stack(),color='white',  show.legend=F) +
-  labs(title='Validation samples predicted by pathology', x='Pathology', y='Percent identified',subtitle='Non-HGD exclude the final endoscopy, p53 excludes NA samples') +
-  plot.theme + theme(legend.position = 'bottom')
-
-```
-
-### Clinical Story
+riskPath = rbind( cbind.data.frame( table(subset(back.v, Status == 'P', select=c('Pathology','Risk'))), 'Status'='Progressor' ), 
+                  cbind.data.frame( table(subset(back.v, Status == 'NP', select=c('Pathology','Risk'))), 'Status'='Non-Progressor' ))
+
+pathTotal = back.v %>% group_by(Status, Pathology) %>% summarise( 'nPath'=length(Pathology))
+pathTotal$Status = ifelse(pathTotal$Status == 'P', 'Progressor','Non-Progressor')
+
+riskPath = merge(riskPath, pathTotal)
+riskPath$ratio = with(riskPath, round(Freq/nPath, 2))
+riskPath$Risk = factor(riskPath$Risk, levels=c('High','Moderate','Low'), ordered = T)
+
+p53Path = cbind.data.frame(table(subset(back.v, Status == 'P', select=c('Pathology','p53.Status'))), 'Status'='p53 IHC')
+p53Path = merge( subset(back.v, Status == 'P' & !is.na(p53.Status)) %>% group_by(Pathology) %>% summarise( 'nPath'=length(Pathology)), p53Path)
+p53Path$Risk = with(p53Path, ifelse(p53.Status == 1, 'High','Low'))
+p53Path$ratio = with(p53Path, round(Freq/nPath, 2))
+
+riskPath = rbind(riskPath, p53Path[colnames(riskPath)])
+riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor','p53 IHC'), ordered = T)
+
+pathTotal = riskPath %>% group_by(Status, Pathology) %>% summarise('nPath'=unique(nPath))
+
+ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
+  scale_fill_manual(values=c('red','orange2','green3')) + 
+  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), position=position_stack(), size=4) +
+  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5) +
+  facet_grid(~Status) + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
+
+```
+
+## Clinical Story
 
@@
 The labels show the proprotions of progressors that fell into each quintile in the training set.
-```{r}
-
-vpd2$quants = with(vpd2, cut(Prediction, breaks=cuts))
-#table(vpd2$quants)/nrow(vpd2)
-
-vpd2$Pathology = as.character(vpd2$Pathology)
-vpd2[which(vpd2$Pathology %in% c('HGD','IMC')), 'Pathology'] = 'IMC/HGD'
-
-summary.v = as.data.frame.matrix(with(vpd2, table(Pathology, Risk)))
-summary.v$Pathology = rownames(summary.v)
-
-write.table(
-summary.v %>% group_by(Pathology) %>% mutate( 
-  'High' = paste(High, ' (', round(High/nrow(validation.patient.info),2)*100, '%)', sep=''),
-  'Moderate' = paste(Moderate, ' (', round(Moderate/nrow(validation.patient.info),2)*100, '%)', sep=''),
-  'Low' = paste(Low, ' (', round(Low/nrow(validation.patient.info),2)*100, '%)', sep=''))
-, sep='\t', quote=F, row.names = F)
-
-ggplot(vpd2, aes(Prediction)) + geom_histogram(aes(fill=..x..), bins=20, show.legend = F) +
+```{r, echo=F, warning=F, message=F}
+
+summary.v = table.path.risk(back.v)
+
+pander(as.data.frame(summary.v))
+
+write.table(summary.v, sep='\t', quote=F, row.names = F)
+```
+
+### Vignetts
+
+```{r, echo=F, warning=F, message=F, fig.width=12, fig.height=4}
+vign.pts = as_tibble(subset(sum.patient.data, Patient %in% c(3, 33, 34, 49, 56, 20), select=c('Status','Patient','age.diagnosis', 'smoking', 'C','M', 'initial.endo','final.endo', 'total.endos','total.samples')))
+colnames(vign.pts) = c('Status','Patient','Age at Diagnosis','Smoker','Circumference', 'Maximal length', 'Intial Endoscopy','Final Endoscopy', 'Total Endoscopies','Total Samples')
+
+back$Risk = factor(back$Risk, levels=levels(back$labels), ordered = T)
+
+vig.plot<-function(df, info) {
+  scale = seq(min(info$Endoscopy.Year), max(info$Endoscopy.Year), 1)
+  subtitle = with(info, paste(unique(Age.at.diagnosis),unique(Sex),': ', sep=''))
+  if (!is.na(unique(info$Smoking)))                             
+    subtitle = paste(subtitle, ifelse(unique(info$Smoking) == 'Y', 'smoker, ','non-smoker, '),sep='')
+
+  subtitle = paste(subtitle, 'M', unique(info$Maximal), sep='')
+
+  if (!is.na(unique(info$Circumference)))
+    subtitle = paste(subtitle, 'C', info$Circumference, sep='')
+
+  subtitle = paste(subtitle, " BE diagnosed ", unique(info$Initial.Endoscopy), sep='')
+
+  ggplot(df, aes(Endoscopy.Year, nl)) + geom_tile(aes(fill=Risk), color='white') +
+    scale_fill_manual(values=c('red3','orange', 'green3'), limits=levels(df$Risk),  name='Progression') +
+    geom_text_repel(aes(label=Pathology), color='white', show.legend=F) +
+    labs(y='Oesophageal Location',x='Months Prior to Endpoint', title='') +
+    plot.theme + theme(axis.text.x=element_text(angle=45, hjust=1),legend.position='none') + 
+    scale_x_continuous(breaks=scale) +
+    labs(title=paste('Patient',unique(info$Patient)), subtitle=subtitle)
+}
+
+## P
+vig.plot(df=subset(back, Patient == 34),info= subset(patient.info, Patient == 34))
+vig.plot(subset(back, Patient == 56), subset(patient.info, Patient == 56))
+vig.plot(subset(back, Patient == 20), subset(patient.info, Patient == 20))
+vig.plot(subset(back, Patient == 86), subset(patient.info, Patient == 86))
+
+## NP
+vig.plot(subset(back, Patient == 3), subset(patient.info, Patient == 3))
+vig.plot(df=subset(back, Patient == 23),info= subset(patient.info, Patient == 23))
+vig.plot(subset(back, Patient == 33), subset(patient.info, Patient == 33))
+vig.plot(subset(back, Patient == 49), subset(patient.info, Patient == 49))
+vig.plot(subset(back, Patient == 51), subset(patient.info, Patient == 51))
+
+```
+
+```{r, echo=F, warning=F, message=F}
+ggplot(back.v, aes(Prediction)) + geom_histogram(aes(fill=..x..), bins=20, show.legend = F) +
   scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
   geom_vline(xintercept=pred.confidence$r2, linetype='longdash') +
-  geom_label(data=(pred.confidence %>% rowwise %>% mutate( 'mid'=sum(r1,r2)/2 )), aes(x=mid, y=11, label=paste(round(perc,2),conf,sep=''),color=conf ), show.legend=F) + 
+#  geom_label(data=(pred.confidence %>% rowwise %>% mutate( 'mid'=sum(r1,r2)/2 )), aes(x=mid, y=11, label=paste(round(perc,2),conf,sep=''),color=conf ), show.legend=F) + 
   scale_color_manual(values=c('grey','black')) + plot.theme + labs(title='Validation predictions', y='n Samples', x='Probability') 
@@
 
-```{r, echo=F, warning=F, message=F, fig.height=8, fig.width=5}
-
-clin.eval = vpd2 %>% group_by(Patient, Status, Endoscopy.Year, Pathology, PID, months.before.final, Prediction) %>% summarise( 'High'=length(which(labels=='High')), 'Moderate'=length(which(labels=='Moderate')), 'Low'=length(which(labels=='Low')))
-clin.eval$Endoscopy.Year = as.numeric(as.character(clin.eval$Endoscopy.Year))
-# First and penultimate endoscopy where possible
-
-## Need to select the max prediction in the year for the Risk category
-
-mark.penult<-function(Endoscopy.Year, PID, Prediction) {
-  tb = table(Endoscopy.Year, PID)
-  if (nrow(tb) > 2) {
-    penult = Prediction == max( Prediction[which(Endoscopy.Year == Endoscopy.Year[which.max(Endoscopy.Year)-1])] )
-  } else if (nrow(tb) == 2) {
-    penult = Prediction == max(Prediction[which(Endoscopy.Year == Endoscopy.Year[2])]) 
-  } else {
-    penult = Prediction == max(Prediction[which(Endoscopy.Year == Endoscopy.Year[1])])
-  } 
-  penult
-}
-
-clin.eval = clin.eval %>% group_by(Patient) %>% 
-  mutate( 'first'=Prediction == max(Prediction[which(Endoscopy.Year==min(Endoscopy.Year))] ),
-          'penult'=mark.penult(Endoscopy.Year, PID, Prediction) )
-
-clin.eval = subset(clin.eval, first | penult)
-
-m = melt(clin.eval[,c('Patient','Status','Pathology','Endoscopy.Year','Prediction','High','Moderate','Low')], measure.vars=c('High','Moderate','Low'), variable.name='Risk', value.name='Risk.value')
-m = subset(m, Risk.value > 0)
-m$Endoscopy.Year = factor(m$Endoscopy.Year, ordered = T)
-m$Risk = factor(m$Risk, levels=c('Low','Moderate','High'), ordered = T)
-
-ggplot(subset(m, Status == 'P'), aes(Endoscopy.Year, Risk)) + 
-  geom_tile(aes(fill=Risk), color='white') + 
-  scale_fill_manual(values=c('green','orange','red'), drop=F) +
-  geom_text(aes(label=Pathology), color='white') +
-  facet_wrap(~Patient, scales='free_x',labeller=label_both) + plot.theme + theme(legend.position = 'none')
-
-ggplot(subset(m, Status == 'NP'), aes(Endoscopy.Year, Risk)) + 
-  geom_tile(aes(fill=Risk), color='white') + 
-  scale_fill_manual(values=c('green','orange','red'), drop=F) +
-  geom_text(aes(label=Pathology), color='white') +
-  facet_wrap(~Patient, scales='free_x', labeller=label_both) + plot.theme + theme(legend.position = 'none')
-
-```
-
 ## Clonal Heterogenity
@@
 ```{r clonality, echo=F, message=F, warning=F, eval=F}
+patient.info = subset(patient.info, Hospital.Research.ID != 'AH0222')
+
+multi_samp = subset(patient.info, PID %in% names(which(table(patient.info$PID) > 1)))
+pts = unique(multi_samp$Hospital.Research.ID)
 
 ## Variance in in CN per endoscopy
-pt = 'AH0254'
-subset(sum.patient.data, Hospital.Research.ID == pt)
-
-table(pg.samp[[pt]]$Endoscopy.Year) 
-which( table(pg.samp[[pt]]$PID) > 1)
-
-subset(pg.samp[[pt]], PID == 'PS10.5754_')$Samplename
-
-x = apply( dysplasia.df[ subset(pg.samp[[pt]], PID == 'PS10.5754_')$Samplename,  ], 2, var)
-
-x[[1]]
-
-# So pretty much the # of samples == # of endoscopies so can't do variance per endo
-plot(sum.patient.data[c('total.samples','total.endos')])
-cor(sum.patient.data[c('total.samples','total.endos')])
+for (pt in pts) {
+
+  pid = names(which(table(pg.samp[[pt]]$PID) > 1))
+
+  for(i in 1:length(pid)) {  
+    samples = subset(pg.samp[[pt]], PID == pid[i])$Samplename
+    subset(pg.samp[[pt]], PID == pid[i])$Prediction
+
+    x = apply( dysplasia.df[ samples,  ],2, var)
+    summary(x)
+
+    range(x)
+    plot(x)
+  }
+
+}
 
@@
 }
-
 ```

### 2018-01-05 APClusterRegression.Rmd
...

---
title: "R Notebook"
output: html_notebook
---

Check how many (or which) samples can be peeled back before the complexity score completely fails.
```{r setup, include=FALSE}
library(ggplot2)
library(GGally)
library(plyr)
library(xlsx)
library(dplyr)
library(pander)
library(Hmisc)
library(ggdendro)
library(reshape2)
library(GenomicRanges)
library(gridExtra)
library(biomaRt)
library(multtest)
library(tibble)
library(mclust)
library(ggfortify)

source('lib/load_patient_metadata.R')

data = '~/Data/Ellie'

data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
analysis.files = list.files(paste(data, 'Analysis', sep='/'), full.names=T)
plot.dir = paste(data, 'Analysis/multipcf_plots_fitted_perPatient', sep='/')

if (length(list.files(plot.dir)) <= 0)
  stop(paste("No analysis files found in", plot.dir ))

## Patient info file
patient.file = grep('All_patient_info.txt', data.files, value=T)
if (length(patient.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", data))

load(grep('Training_patients.Rdata', analysis.files, value=T), verbose=T)

patient.info = read.patient.info(patient.file)
patient.info$Patient = gsub("/", "_", patient.info$Patient)
#head(patient.info)

patient.info = arrange(patient.info, Status, Patient, Endoscopy.Year, Pathology)

sum.patient.data = summarise.patient.info(patient.info)
sum.patient.data = as.data.frame(subset(sum.patient.data, Patient %in% names(patient.data))) ## For now
pander(sum.patient.data)

# Missing some of the samples as they aren't all sequenced yet
for (pt in names(patient.data)) {
  patient.data[[pt]]$info= patient.data[[pt]]$info[patient.data[[pt]]$info$Samplename %in% colnames(patient.data[[pt]]$seg.vals)[-(1:5)],]
}

```

```{r global_pt, echo=F, message=F, warning=F, error=F}
be = subset(patient.info, Pathology == 'BE')
lgd = subset(patient.info, Pathology <= 'LGD')

p53Status = data.frame(rbind('Non-Progressors'=table(subset(patient.info, Status == 'NP')$p53.Status), 
           'Progressors'=table(subset(patient.info, Status == 'P')$p53.Status)))
colnames(p53Status) = c('None', 'Positive')

set.seed(99)
mtp1 = MTP(X=t(patient.info[,c('Barretts.Cellularity','Total.Reads')]), Y=as.integer((as.integer(patient.info$Pathology) > 3)), 
           get.adjp=T, na.rm=T, test='t.twosamp.equalvar')
mtp2 = MTP(X=t(patient.info[,c('Barretts.Cellularity','Total.Reads')]), Y=patient.info$Status, 
            get.adjp=T, na.rm=T, test='t.twosamp.equalvar')
```

Currently with multi test correction the samples for which cellularity and #of reads are available show no difference for either status (P vs NP)
`r pander(mtp2@adjp, justify='left', caption="Adjusted p-value")`

or for samples that are LGD/IMC/HGD vs BE
`r pander(mtp1@adjp, justify='left', caption="Adjusted p-value")`

However, it should be noted that pathologists do not like calling cellularity on Barrett's generally so it's a fairly inaccurate measure.
`r pander(table( patient.info[,c('Status','Barretts.Cellularity')]), caption="Numbers of cellularity calls by status")`

With regards to p53 staining, non-progressors have no p53 positive tests, while progressors had `r p53Status[2,2]` overall.
`r pander(p53Status, justify='left', caption="P53 staining status for progressors vs non")`
Of these `r table(be$p53.Status)[['1']]` were positive in a BE sample, and `r table(lgd$p53.Status)[['1']]` were positive when LGD samples were included.

## AP Clustering

```{r apclust-func, echo=F, message=F, warning=F}
suppressPackageStartupMessages( library(apcluster) )

apclust.data<-function(segdata, samples) {
  x1 = segdata[,intersect(colnames(segdata), samples)]
  #x1 = x1[, samples]
  rownames(x1) = (segdata[,c(1:4)] %>%
    rowwise() %>%
    mutate(location=paste(paste(chrom, arm, sep=''), '.', start.pos, '-', end.pos, sep='')))$location
  return(x1)  
}

apclust<-function(segdata) {
  x1 = apclust.data(segdata)
  # q=0 minimize off-diagonal similarity
  ac = apcluster(negDistMat(r=2), x1, details=T, convits=25, q=0)
  return(list('apres'=ac, 'data'=x1))
}

net.similarity<-function(ac) {
  if (is.list(ac))
    ac = ac$apres
  ac@netsim
}

sum.similarity<-function(ac) {
  if (is.list(ac))
    ac = ac$apres
  ac@dpsim
}

clusters<-function(ac) {
  if (is.list(ac))
    ac = ac$apres
  ac@exemplars
}

plotAC<-function(aclist) {
  plot(aclist$apres, aclist$data)
}

heatmapAC<-function(aclist) {
  heatmap(aclist$apres)
}

# First vs last samples
#x1 = apclust.data(patient.data[[pt]]$seg.vals[,c(1:4, 6, ncol(patient.data[[pt]]$seg.vals))], patient.data[[pt]]$info$Samplename)
```

### Raw segmentation values

```{r apclust, echo=F, message=F, warning=F, fig.height=10, fig.width=10, fig.align='left'}
sample.vals = list()
q=0.2

sum.patient.data = arrange(sum.patient.data, Status)

## Using the segmentation values
sum.patient.data$years = with(sum.patient.data, paste(start.year, end.year, sep='-'))
sim = as.data.frame(sum.patient.data[,c('Status','years')])
rownames(sim) = sum.patient.data$Patient
for (pt in sum.patient.data$Patient) {
  patient.data[[pt]]$info = arrange(patient.data[[pt]]$info,  Pathology, Endoscopy.Year)
  x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)

  gg = ggplot(melt(x1), aes(x=variable, y=value, group=variable, fill=variable)) + 
    geom_boxplot(notch=T, show.legend=F) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position='none') +
    labs(title=paste(pt, patient.data[[pt]]$info$Status[1], sep=' - ')) +   theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

  sample.vals[[pt]] = gg

  info = patient.data[[pt]]$info[which(colnames(x1) %in% patient.data[[pt]]$info$Samplename),]
  colnames(x1) = paste(info$Endoscopy.Year, '(', info$Pathology, ')', sep='' )

  ac = apcluster(negDistMat(x1, r=2), q=q, details=T, includeSim=T)

  last = ncol(x1)
  if (patient.data[[pt]]$info$Status[1] == 'P' & length(which(grepl('HGD|IMC', info$Pathology)) > 0)  )
    last = sort(grep('HGD|IMC', info$Pathology), decreasing=T)[1]
  if (ncol(x1) <= 15) {
    #plot(ac, x1[,c(c(1:ncol(x1))[-last], last)], main=pt)
  } else {
    #plot(ac, x1[,c(1, sort(sample(c(1:(ncol(x1)-1))[c(-1,-last)], 8)) ,last) ], main=pt)
  }

  sim[pt, 'ns'] = net.similarity(ac)
  sim[pt, 'ss'] = sum.similarity(ac)
  sim[pt, 'clusters'] = length(clusters(ac))
  sim[pt, 'max.sd'] = max(sapply(x1, sd))

  tmp = patient.data[[pt]]$variable.regions[,-(1:5)]
  sim[pt, 'n.var'] = nrow(patient.data[[pt]]$variable.regions[,-(1:5)])

  if (nrow(tmp) <= 0) {
    sim[pt,c('gains', 'loss')] = c(0,0)
  } else {
    sim[pt,'gains'] = length(tmp[tmp>=1.1])
    sim[pt,'loss'] =  length(tmp[tmp<=0.9])
  }
}
sim$id = rownames(sim)
rc = cor.test(sim$max.sd, sim$clusters)

do.call(grid.arrange, c(sample.vals))

# May be a difference
ttps = wilcox.test( subset(sim, Status == 'NP')$ss, subset(sim, Status == 'P')$ss )$p.value
ttpc = wilcox.test( subset(sim, Status == 'NP')$clusters, subset(sim, Status == 'P')$clusters )$p.value

## Number of variable regions isn't so good, but maybe...
#sim %>%
#  group_by(Status) %>%
#  summarise( mean = mean(n.var), sd = sd(n.var), med = median(n.var))
#t.test( subset(sim, Status == 'NP')$n.var, subset(sim, Status == 'P')$n.var )$p.value

g1 = ggplot(sim[,c('Status','ss')], aes(y=ss, x=Status, fill=Status)) + geom_boxplot() + 
  geom_jitter(width=0.2) + stat_boxplot(geom ='errorbar') +
  labs(title=paste("p-value", signif(ttps, 3)), y='sum similarity')

g2 = ggplot(sim[,c('Status','clusters')], aes(y=clusters, x=Status, fill=Status)) + geom_boxplot() + 
  geom_jitter(width=0.2) + stat_boxplot(geom ='errorbar') +
  labs(title=paste("p-value", signif(ttpc, 3)), y='num. clusters')

sim$Status.Int = as.integer(sim$Status)-1
chisq.test(sim[,c('Status.Int','clusters')])
```

This plot compares the two possible measures of overall complexity from AP. The summed similarity value (top) and the number of clusters (bottom). In this case, the number of clusters appears to be a more robust measure.

```{r echo=F, message=F, warning=F, fig.width=5, fig.height=5}
grid.arrange(g1, g2, top='Raw seg values')
```

Based on the number of clusters in each patient, it appears that there are a higher number of clusters being selected by AP in the progressors vs the non, suggesting it may be able to relate it to diversity in the progressors CN. That metric appears to provide a good separation between the patients as well, though there's a suggestion of two different types of progressors.

```{r apglmclus, echo=F, message=F, warning=F}
precisionRecall<-function(actual, predicted) {
    retrieved <- sum(predicted==1)
    recall <- sum(predicted==1 & actual==1) / sum(actual==1)
    precision <- sum(predicted==1 & actual==1) / retrieved
    f1 = signif(2*(1/(1/precision+1/recall)), 2)
    return(data.frame('precision'=precision, 'recall'=recall, 'F1'=signif(f1, 2)))
  }

simPR = sim[, c('Status', 'clusters', 'ns','ss')]
actualMatrix = as.data.frame(matrix(0,0,4,dimnames=list(c(), c('status','precision','recall', 'F1'))))
for (i in 1:100) {
  trainingRows = sample(1:nrow(sim), nrow(sim)-10)
  train = simPR[trainingRows,]
  test = simPR[-trainingRows,]

  fit = glm( (as.integer(Status)-1)~clusters, train, family=binomial(link='logit'))
  pred = predict(fit, test, se.fit=F) # probabilities -- of which outcome though?

  #glm_link_scores <- predict(fit, test, type="link")
  #glm_response_scores <- predict(fit, test, type="response")

  actualMatrix = rbind(actualMatrix, cbind('status'='P', precisionRecall(test$Status == 'P', pred>0)))
  actualMatrix = rbind(actualMatrix, cbind('status'='NP', precisionRecall(test$Status == 'NP', pred<0)))
}

g1 = ggplot(actualMatrix, aes(x=recall, y=precision, color=status, label=F1)) + 
  geom_point() + geom_text(aes(label=ifelse(precision>0.5 & recall>0.5, as.character(F1), ''), hjust=0, vjust=0)) +
   geom_hline(yintercept = 0.5, col='grey') + geom_vline(xintercept = 0.5, col='grey') +
   labs(title='Precision/Recall for Progressor/NP prediction')

g2 = ggplot(actualMatrix, aes(x=status, y=F1, group=status, fill=status)) + 
  geom_violin() + geom_jitter(alpha=0.5) + ylim(0,1) + 
  geom_violin(data=subset(actualMatrix,precision>0.5 & recall>0.5), aes(x=status, y=F1, group=status, fill='P&R>0.5') , alpha=0.6) +
  labs(title="F1 statistic")

grid.arrange(g1, g2, top='100 iterations 10f xval')
```

Randomize the data and see if the F1 statistic still holds

```{r gpglmclus_rand, echo=F, message=F, warning=F}
#simR = sim[, c('Status', 'clusters')]

rand = as.data.frame(matrix(0,0,4,dimnames=list(c(), c('status','precision','recall', 'F1'))))
for (i in 1:100) {
  # rand.sim = as.data.frame(cbind(  'Status'=sample(c('P','NP'), nrow(sim), replace=T, prob=round(table(sim$Status)/nrow(sim), 2)),
  #                    'clusters'=sample((3:30), nrow(sim), replace=T) ))
  # rand.sim = as.data.frame(cbind(  'Status'=sample(c('P','NP'), nrow(sim), replace=T, prob=round(table(sim$Status)/nrow(sim), 2)),
  #                    'clusters'=sample((3:30), nrow(sim), replace=T) ))

  rand.sim = as.data.frame(cbind(  'Status'=sample(c('P','NP'), nrow(sim), replace=T),
                     'clusters'=sim$clusters ))

  rand.sim$clusters = as.integer(as.character(rand.sim$clusters))

  trainingRows = sample(1:nrow(rand.sim), nrow(rand.sim)-10)
  train = rand.sim[trainingRows,]
  test = rand.sim[-trainingRows,]

  fit = glm( (as.integer(Status)-1)~clusters, train, family=binomial(link='logit'))
  pred = predict(fit, test, se.fit=F) # probabilities -- of which outcome though?

  rand = rbind(rand, cbind('status'='P', precisionRecall(test$Status == 'P', pred>0)))
  rand = rbind(rand, cbind('status'='NP', precisionRecall(test$Status == 'NP', pred<0)))
}

g1 = ggplot(rand, aes(x=recall, y=precision, color=status, label=F1)) + 
  geom_point() + geom_text(aes(label=ifelse(precision>0.5 & recall>0.5, as.character(F1), ''), hjust=0, vjust=0)) +
   geom_hline(yintercept = 0.5, col='grey') + geom_vline(xintercept = 0.5, col='grey') +
   labs(title='Precision/Recall for Progressor/NP prediction')

g2 = ggplot(rand, aes(x=status, y=F1, group=status, fill=status)) + 
  geom_violin() + geom_jitter(alpha=0.5) + ylim(0,1) + 
  #geom_violin(data=subset(actualMatrix,precision>0.5 & recall>0.5), aes(x=status, y=F1, group=status, fill='P&R>0.5') , alpha=0.6) +
  labs(title="F1 statistic")

grid.arrange(g2, top='100 iterations 10f xval - randomized labels')

```

### Cross-validation, LOO

So using the raw segmentation values (not the normalized ones) try some leave one out xval to see if the currently weak differences still hold.

#### Patients first

```{r xval_pt, echo=F, warning=F, message=F, fig.align='left', fig.height=4, fig.width=4, eval=F}
# Leave out patients
loo_xval = matrix(ncol=2, nrow=0, dimnames=list(c(), c('sum.similarity', 'clusters')))

#s = sample(rep(seq(5), length = length(unique(pts$Patient))))
#patients = merge(pts, cbind('Patient'=unique(pts$Patient), 'group'=s), by="Patient")

for (i in 1:nrow(sim)) {
  ttps = wilcox.test( subset(sim[-i,], Status == 'NP')$ss, subset(sim[-i,], Status == 'P')$ss )$p.value
  ttc = wilcox.test( subset(sim[-i,], Status == 'NP')$clusters, subset(sim[-i,], Status == 'P')$clusters )$p.value

  loo_xval = rbind(loo_xval, cbind(ttps, ttc))
}

m = melt(data.frame(loo_xval))
m2 = m %>% group_by(variable) %>% summarise('y.pos'=max(value)+mean(value), 'text'=round(mean(value), 3))
ggplot(m, aes(x=variable, y=value, fill=variable)) +  
  geom_boxplot() + geom_jitter(shape=16, width=0.2) + 
  geom_hline(yintercept=0.05, color='red', alpha=0.5) +
  stat_boxplot(geom ='errorbar') + labs(y='p-values', x='Iterations', title='Leave one (patient) out clustering') +
  scale_fill_manual(values=c('purple', 'darkred')) +
  geom_text(data=m2, aes(x=variable, y=y.pos, label=text), vjust=0)

```

## No HGD samples

There is still a difference between the progressors and non, though the p-value does increase. Some of the progressors do start looking more like NPs in their similarity measure but not all of them.

```{r nofinal, echo=F, message=F, warning=F, fig.height=5, fig.width=5}

simX = sum.patient.data[,c('Status','years')]
rownames(simX) = sum.patient.data[,'Patient']
# Leave out final sample on the progressors

for (pt in sum.patient.data$Patient) {
  x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
  if (subset(sum.patient.data, Patient == pt)$Status == 'P' & length(which(grepl('HGD|IMC', patient.data[[pt]]$info$Pathology))) > 0) {
    final = grep('HGD|IMC', patient.data[[pt]]$info$Pathology)

    if (length(patient.data[[pt]]$info$Samplename[-final]) <= 1) next

    x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename[-final])
  }
  ac = apcluster(negDistMat(x1, r=2), q=q)

  simX[pt, 'ss'] = sum.similarity(ac)
  simX[pt, 'clusters'] = length(clusters(ac))
  simX[pt, 'n.var'] = nrow(patient.data[[pt]]$variable.regions[,-(1:5)])
}

# Correlation between the number of variable regions and clusters, suggesting the complexity measure is reasonable
ct = cor.test(simX$clusters, simX$n.var)

ttpc = wilcox.test( subset(simX, Status == 'NP')$clusters, subset(simX, Status == 'P')$clusters )$p.value

ggplot(melt(data.frame(simX[c('Status','clusters')])), aes(x=Status, y=value, fill=Status)) +  
    geom_boxplot() + geom_jitter(shape=16, position=position_jitter(0.2)) +
    stat_boxplot(geom ='errorbar') + labs(y='num. clusters', x='', title=paste('Leave out HGD sample(s) p-value', signif(ttpc, 3)))

#fit = glm(Status~clusters+ss, simX, family=binomial(link='logit') )

simX = simX[, c('Status', 'clusters', 'ss','n.var')]
noHGDPR = as.data.frame(matrix(0,0,4,dimnames=list(c(), c('status','precision','recall', 'F1'))))
for (i in 1:100) {
  trainingRows = sample(1:nrow(simX), nrow(simX)-10)
  train = simX[trainingRows,]
  test = simX[-trainingRows,]

  fit = glm( Status~clusters, train, family=binomial(link='logit'))
  pred = predict(fit, test, se.fit=F) # probabilities -- of which outcome though?

  noHGDPR = rbind(noHGDPR, cbind('status'='P', precisionRecall(test$Status == 'P', pred>0)))
  noHGDPR = rbind(noHGDPR, cbind('status'='NP', precisionRecall(test$Status == 'NP', pred<0)))
}

g1 = ggplot(noHGDPR, aes(x=recall, y=precision, color=status, label=F1)) + 
  geom_point() + geom_text(aes(label=ifelse(precision>0.5 & recall>0.5, as.character(F1), ''), hjust=0, vjust=0)) +
   geom_hline(yintercept = 0.5, col='grey') + geom_vline(xintercept = 0.5, col='grey') +
   labs(title='Precision/Recall for Progressor/NP prediction without HGD')

g2 = ggplot(noHGDPR, aes(x=status, y=F1, group=status, fill=status)) + 
  geom_violin() + geom_jitter(alpha=0.5) + ylim(0.5,1) + labs(title="F1 statistic, no HGD/IMC")

grid.arrange(g1, g2, top='100 iterations 10f xval no HGD/IMC')

```

## Per sample AP?

```{r, echo=F, message=F, warning=F}

per.sample.clusters = lapply (patient.data, function(df) {
  print(pt)
  x1 = apclust.data(df$seg.vals, df$info$Samplename)
  print(ncol(x1))

  clusters = as.data.frame(do.call(rbind, lapply(colnames(x1), function(col) {
    ac = apcluster(negDistMat(x1[,col] , r=2), q=0.2)
    path = as.character(subset(df$info, Samplename == col)$Pathology)

    cbind('path'=path, 'n.clusters'=length(ac@clusters))
  })))
  clusters$path = as.character(clusters$path)
  clusters$n.clusters = as.numeric(as.character(clusters$n.clusters))
  return(clusters)  
})

pc = unlist(sapply(per.sample.clusters[ subset(sum.patient.data, Status == 'P')$Patient  ], function(df) df$n.clusters))
npc = unlist(sapply(per.sample.clusters[ subset(sum.patient.data, Status == 'NP')$Patient  ], function(df) df$n.clusters))

t.test(pc,npc) # No diff

ggplot(data.frame('status'=c(rep('P', length(pc)), rep('NP', length(npc))), 'value'=c(pc, npc)), aes(x=status, y=value)) + 
  geom_boxplot()+ geom_jitter()

hgd = unlist(lapply(per.sample.clusters, function(df) {
  subset(df, path %in% c('HGD', 'IMC'))$n.clusters
}))

be = unlist(lapply(per.sample.clusters, function(df) {
  subset(df, path %in% c('BE', 'LGD'))$n.clusters
}))

t.test(hgd, be) # No diff

pcBE = unlist(sapply(per.sample.clusters[ subset(sum.patient.data, Status == 'P')$Patient  ], function(df) 
  subset(df, path %in% c('BE','LGD'))$n.clusters))

npcBE = unlist(sapply(per.sample.clusters[ subset(sum.patient.data, Status == 'NP')$Patient  ], function(df) 
  subset(df, path %in% c('BE','LGD'))$n.clusters))

t.test(pcBE, npcBE) # Nope

```

Generating a per-sample AP cluster results in finding no differences between groups, (NP vs P, HGD vs BE, or NPbe vs Pbe). Need to try iteratively clustering each patient with n-i samples and testing.

## Peel back timepoints

How exactly?

```{r, echo=F, message=F, warning=F}
precisionRecall<-function(actual, predicted) {
    retrieved <- sum(predicted==1)
    recall <- sum(predicted==1 & actual==1) / sum(actual==1)
    precision <- sum(predicted==1 & actual==1) / retrieved

    #f1 = signif(2*(1/(1/precision+1/recall)), 2)
    f1 = 2*precision*recall/(precision+recall)
    return(data.frame('precision'=precision, 'recall'=recall, 'F1'=signif(f1, 3)))
  }

q=0.2

iterative.cluster <- function(df) {
  x1 = apclust.data(df$seg.vals, df$info$Samplename)
  info = arrange(df$info, -Endoscopy.Year, Pathology)

  years = do.call(rbind, lapply(unique(info$Endoscopy.Year), function(year) {
    tinfo = subset(info, Endoscopy.Year <= year)

    ac = apcluster(negDistMat(x1[,tinfo$Samplename], r=2), q=q)
    cbind.data.frame('Final.Year'=range(tinfo$Endoscopy.Year)[2], 
            'n.samples'=nrow(tinfo), 'max.path'=as.character(max(tinfo$Pathology)),
            'ss'=sum.similarity(ac), 'clusters'=length(ac@clusters))
    }))
}

nonps = lapply( patient.data[subset(sum.patient.data, Status == 'NP')$Patient], iterative.cluster )
ps = lapply( patient.data[subset(sum.patient.data, Status == 'P')$Patient], iterative.cluster )

## model is based on the final set in all cases..

data = rbind(cbind(do.call(rbind, lapply(ps, function(df) df[1,])), 'Status'='P'),
      cbind(do.call(rbind, lapply(nonps, function(df) df[1,])), 'Status'='NP'))

get.next<-function(df, i) {
  if (i > nrow(df)) i = nrow(df)
  df[i,]
}

## Is this correct? 
pi.hat <- function(x) exp(x)/(1+exp(x))  ## when/why do I use pi.hat??

# Leave one progressor out and then predict all of the samples on the progressor
predictions = list()
for (i in 1:length(ps)) {
  data = rbind(
    cbind(do.call(rbind, lapply(ps[-i], function(df) df[1,])), 'Status'='P'),
    cbind(do.call(rbind, lapply(nonps, function(df) df[1,])), 'Status'='NP')
    )

  fit0 = glm(Status~clusters, data, family='binomial')
  #print(summary(fit0))
  print(summary(fit0)$coefficients[,4])
  autoplot(fit0)

  pred = pi.hat(predict(fit0, ps[[i]], type='response'))
  predictions[[names(ps)[i]]] = list('P.hat'=pred, 'prf'=  precisionRecall(rep(1, nrow(ps[[i]])), pred>0.5), 'fit.p.values'=summary(fit0)$coefficients[,4])
}

# Leave one non-progressor out and then predict all of the samples on the non-progressor
predictionsNP = list()
for (i in 1:length(nonps)) {
  data = rbind(
    cbind(do.call(rbind, lapply(ps, function(df) df[1,])), 'Status'='P'),
    cbind(do.call(rbind, lapply(nonps[-i], function(df) df[1,])), 'Status'='NP')
    )

  fit0 = glm(Status~clusters, data, family='binomial')
  #print(summary(fit0))
  print( summary(fit0)$coefficients[,4] )
  autoplot(fit0)

  pred = pi.hat(predict(fit0, nonps[[i]], type='response'))
  predictionsNP[[names(nonps)[i]]] = list('P.hat'=pred, 'prf'=  precisionRecall(rep(1, nrow(nonps[[i]])), pred>0.5), 'fit.p.values'=summary(fit0)$coefficients[,4])
}

# i = 1
# data = rbind(
#   cbind(do.call(rbind, lapply( ps, get.next, i)), 'Status'='P'),
#   cbind(do.call(rbind, lapply( nonps, get.next, i)), 'Status'='NP')
#   )
# 
# fit0 = glm(Status~clusters, data, family='binomial')
# print(summary(fit0))
# summary(fit0)$coefficients[,4]

#test = ps[['AHM0277']]
#pi.hat(predict(fit0, test, type='response'))

#timept = do.call(rbind, ps)
#pred = pi.hat(predict(fit0, timept))

# correlation between the prediction and the max pathology, though not overly strong
#ct = cor.test(as.integer(timept$max.path), as.vector(pred))
# None with number of samples though    
#ct2 = cor.test(timept$n.samples, as.vector(pred))

#There is a correlation between the prediction and the max pathology, though not overly strong at `r signif(ct$estimate, 3)` (p=`r signif(ct$p.value, 2)`).  But none between the number of samples at that point and the prediction (p=`r signif(ct2$p.value, 2)`).

```

```{r, echo=F, message=F, warning=F}

#df = cbind(timept, 'pi.hat'=pred)

ggplot(sim, aes(y=Status.Int, x=clusters)) + geom_point() +  
  stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + 
  labs(title="Status vs clusters, glm") +
  scale_y_discrete('Status', labels=c('Non-Prog','','','','Prog'), limits=c(0,1))

#ggplot(df, aes(x=clusters, y=pi.hat)) + geom_point() + geom_smooth() + 
#  labs(title="Prediction vs number of clusters", y="predicted values")

#ggplot(df, aes(x=as.factor(max.path), y=pi.hat)) + geom_boxplot() + 
#  labs(title="Predictions vs max pathology", y='predicted values', x='max pathology')

```

### 2018-01-05 Paper.Rmd
...

@@
 
+riskCols = brewer.pal(11, "RdYlBu")[c(10, 5, 1)]
 ```
@@
 cnLegend = get.legend(globalCN.plot)
-globalCN.plot = globalCN.plot + theme(legend.position = 'none')
-
+#globalCN.plot = globalCN.plot + theme(legend.position = 'none')
+
+## No HGD
+`%nin%` = Negate(`%in%`)
+
+nhgd.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology %nin% c('HGD','IMC'))$variable)), ')', sep=''),
+              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology %nin% c('HGD','IMC'))$variable)),')', sep=''))
+globalnHGDCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology %nin% c('HGD','IMC')), labeller(Status=nhgd.status, chr=label_value)) + labs(title='Excluding HGD/IMC')
+
+## NDBE
 nb.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology == 'BE')$variable)), ')', sep=''),
@@
 globalBECN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology == 'BE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')
-#globalCNBEdf = subset(globalCNdf, Pathology == 'BE') %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
 
 globalCN.plot 
+globalnHGDCN.plot
 globalBECN.plot
@@
 
-a = sample(subset(patient.info, Status == 'NP' )$Samplename, 1)
-b = sample(subset(patient.info, Status == 'P' )$Samplename, 1)
-
-cn.mtn.plot(subset(globalCNdf, variable %in% c(a,b) & seg.type != 'arms'), labeller(label_both)) + labs(title="Individual samples")
-cn.mtn.plot(subset(globalCNdf, variable %in% c(a,b) & seg.type != 'arms'), labeller(label_both), type='bar') + labs(title="Individual samples")
-
+a = sample( subset(subset(patient.info, Patient == 57),  Endoscopy.Year == min(Endoscopy.Year) )$Samplename, 1)
+b = sample( subset(subset(patient.info, Patient == 76), Endoscopy.Year == min(Endoscopy.Year) & Pathology %nin% c('HGD','IMC'))$Samplename, 1)
+
+ind.status <- c('Non-Progressor'='Patient 57 (n=1)',
+              'Progressor'='Patient 76 (n=1), excl. HGD/IMC')
+cn.mtn.plot(subset(globalCNdf, variable %in% c(a,b) & seg.type != 'arms'), labeller(Status = ind.status), type='bar') + labs(title="Individual samples")
 ```
@@
   message(paste("loading", file))
-  load(file, verbose=F)
+  load(file, verbose=T)
 } else {
@@
 do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits'))
+
+```
+
+```{r echo=F, fig.height=5, fig.width=5}
+library(ggfortify)
+autoplot(models$`0.9`) + theme(legend.position = 'none') + plot.theme
 ```
@@
 
-mp = model.performance(performance.at.1se, coefs)
+mp = model.performance(performance.at.1se, coefs, lims=c(0.7,0.8))
 mp$plot + labs(title=paste('All samples,', folds, 'folds,', splits, 'patient splits'))
@@
 classification.rate = add_row(classification.rate, 'model'='All samples', 'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))
+
 ```
@@
 
-```{r noHGD, echo=F, message=F, warning=F, fig.height=16, fig.width=16, eval=F}
+```{r noHGD, echo=F, message=F, warning=F, fig.height=16, fig.width=16, eval=T}
+rm(performance.at.1se, coefs, mp)
+
 file = paste(cache.dir, 'nohgd.Rdata', sep='/')
+if (file.exists(file)) {
+  message(paste("loading file", file))
+  load(file, verbose=T)
+} else {
+  stop(paste("No model file", file, "found"))
+}
+
+nohgd.coefs = coefs[[select.alpha]]
+#do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2))
+```
+
+```{r, echo=F, fig.width=6, fig.height=6}
+mp = model.performance(performance.at.1se, coefs, lims=c(0.69,0.8))
+mp$plot + labs(title='No HGD/IMC samples')
+
+classification.rate = add_row(classification.rate, 'model'='No HGD/IMC', 
+                              'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
+                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]])
+                              )
+
+rm(performance.at.1se, coefs, mp)
+```
+
+#### LOO minus HGD
+
+```{r loonohgd, echo=F, message=F, warning=T}
+
+file = paste(cache.dir, 'loonohgd.Rdata', sep='/')
 if (file.exists(file)) {
@@
 } else {
-  stop(paste("No model file", file, "found"))
-}
-
-#do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2))
-```
-
-```{r, echo=F, fig.height=5, fig.width=5}
-mp = model.performance(performance.at.1se, coefs)
-mp$plot + labs(title=paste('No HGD/IMC samples,', folds, 'folds,', splits, 'patient splits'))
-
-classification.rate = add_row(classification.rate, 'model'='No HGD/IMC', 
-                              'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
-                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]])
-                              )
+  stop(paste("No file", file, "found"))
+}
+```
+
+```{r, echo=F, message=F, fig.width=5, fig.height=5}
+
+  pg.sampNOHGD = do.call(rbind, pg.sampNOHGD)
+
+  perf = cbind.data.frame(perf=performance.at.1se, pt=unique(pg.sampNOHGD$Patient))
+
+  ggplot(melt(perf), aes(x=variable,y=value, group=variable, fill=variable)) +
+    geom_boxplot(fill='lightcoral', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) +
+    geom_jitter(color='grey39', width=0.3) + plot.theme + 
+    geom_label(data=melt(subset(perf, perf %in% range(perf) )), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
+    theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')
+
+  rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')
+
+```
+
+```{r, echo=F, fig.width=6, fig.height=6}
+ggplot(pg.sampNOHGD, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1), show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') + labs(title='Prediction using model w/o HGD', y='n Samples', x='Probability') + plot.theme
+
+cuts = seq(0,1,0.1) 
+pg.sampNOHGD$quants = with(pg.sampNOHGD, cut(Prediction, breaks=cuts))
+
+noHGD.pred.confidence = as.data.frame.matrix(table(pg.sampNOHGD$quants, pg.sampNOHGD$Status))
+noHGD.pred.confidence = cbind(noHGD.pred.confidence, pg.sampNOHGD %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
+noHGD.pred.confidence$quant = rownames(noHGD.pred.confidence)
+
+ft.fun<-function(NP,P) {
+  f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
+  cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
+}
+
+noHGD.pred.confidence = noHGD.pred.confidence %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )
+
+noHGD.pred.confidence[c('r1','r2')] = NA
+for (i in 1:(length(cuts)-1)) 
+  noHGD.pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)
+
+noHGD.pred.confidence$Risk = 'Moderate'
+noHGD.pred.confidence$Risk[ which(with(noHGD.pred.confidence, p.value < 0.05 & perc < .5)) ] = 'Low'
+noHGD.pred.confidence$Risk[ which(with(noHGD.pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'
+
+noHGD.pred.confidence = bind_cols(noHGD.pred.confidence, with(noHGD.pred.confidence, 
+    data.frame(ci.low=qbeta(0.025, shape1=P+.5, shape2=NP+.5),
+               ci.high=qbeta(0.975, shape1=P+.5, shape2=NP+.5))))
+
+noHGD.pred.confidence$Risk = factor(noHGD.pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)
+
+ggplot(noHGD.pred.confidence, aes(mn, perc)) + 
+  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
+  geom_point() + geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
+  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration', subtitle='HGD excluded model') + plot.theme + theme(legend.position = 'none')
+
+# Predict HGD samples
+file = paste(cache.dir, 'nohgd.Rdata', sep='/')
+if (file.exists(file)) {
+  load(file, verbose=T)
+
+  fitNOHGD = models[[select.alpha]]
+
+  lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']
+
+  hgd.samples = subset(patient.info, Pathology %in% c('HGD','IMC'))
+  hgd.df = dysplasia.df[hgd.samples$Samplename,]
+
+  pred = predict(fitNOHGD, newx=hgd.df, s=lambda.opt, type='response')
+  hgd.samples$Prediction = pred
+
+  hgd.pred = hgd.samples %>% group_by(Patient, Hospital.Research.ID, Status, Endoscopy.Year) %>% 
+    summarise( max=max(Pathology), pred=Prediction[which.max(Pathology)] )
+
+  rm(no.hgd.plots, coefs, performance.at.1se, models)
+}
+
+rm(plots, performance.at.1se,coefs, nzcoefs, fits)
 ```
@@
 
-```
-
-```{r, echo=F, fig.height=5, fig.width=5}
-mp = model.performance(performance.at.1se, coefs)
-mp$plot + labs(title=paste('No LGD/HGD/IMC samples,', folds, 'folds,', splits, 'patient splits'))
+#do.call(grid.arrange, c(nolgd.plots, top="No LGD/HGD/IMC samples", ncol=2))
+```
+
+```{r, echo=F, fig.width=6, fig.height=6}
+mp = model.performance(performance.at.1se, coefs, lims=c(0.69,0.8))
+mp$plot + labs(title='No LGD/HGD/IMC samples')
 
@@
                               'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))
+
+rm(performance.at.1se, coefs, mp)
 ```
@@
 
-```{r, echo=F, warning=F, message=F,fig.height=5, fig.width=5}
+```{r, echo=F, warning=F, message=F, fig.width=6, fig.height=6}
 #classification.rate$x = 1:nrow(classification.rate)
@@
   labs(x='', y='Classificaiton rate', title=paste('Classification rate per model at', select.alpha)) + 
-  plot.theme + theme(axis.text.x=element_text(angle=45,hjust=1,face="bold",size=12), legend.position = 'none')
+  plot.theme + theme(axis.text.x=element_text(angle=45,hjust=1,face="bold",size=12), legend.position = 'none') + 
+  ylim(0.7,0.8)
 
@@
 
-```{r, echo=F, warning=F, message=F}
+```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=5}
 file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
@@
 
+```{r nohgd_coefs, message=F, warning=F, echo=F}
+setdiff(ch$label, rownames(nohgd.coefs))
+
+```
+
 ```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
@@
 
-globalCN.plot + geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)
+globalCN.plot + geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + 
+  geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)
 
@@
 
-```{r leaveoneout, echo=F, message=F, warning=F}
+```{r leaveoneout, echo=F, message=F, warning=F, fig.width=5, fig.height=5}
 file = paste(cache.dir, 'loo.Rdata', sep='/')
@@
 }
-loo.coefs = coefs
+loo.coefs = nzcoefs
 
@@
   geom_label(data=subset(df, performance.at.1se %in% range(performance.at.1se)), aes(y=performance.at.1se, label=round(performance.at.1se, 3)), fontface='bold', fill='lightblue') +
-  labs(y='performance at lambda.1se', x='', title=paste('Performance for nested leave one out, at alpha=',select.alpha,sep='')) +
+  labs(y='Classification rate', x='LOO model', title='Performance for LOO', subtitle='All samples model') +
   plot.theme
@@
 
-```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=4, fig.width=4}
+```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=5, fig.width=5}
 preds = unlist(lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], function(df) df$Prediction))
@@
 preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
-roc = pROC::roc(Status ~ Prediction, data=preds, ci=T, of='thresholds')
+roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
 roc$model = 'CN Model'
 
-grid.arrange(roc.plot(roc50,'Probability threshold = 0.5'), roc.plot(roc, 'Probability threshold calculated by ROC'), ncol=2)
-
-roc.plot(roc, 'Copy number model')
+# To find the auc at a given spec or sens
+#pROC::auc(roc, partial.auc=c(0.99,1), partial.acu.focus="sp", partial.auc.correct=T)
+
+df = cbind.data.frame('specificity'=rev(roc$specificities), 'sensitivity'=rev(roc$sensitivities))
+
+thresholds = 
+  rbind.data.frame(
+#    est=t(round(pROC::coords(roc, "best"),2)),
+    max=t(round(pROC::coords(roc, 0.8),2)),
+    'No HGD'=t(round(pROC::coords(rocNOHGD, 0.6),2)),
+    'All samples'=t(round(pROC::coords(roc, 0.6),2))
+)
+thresholds$label=rownames(thresholds)
+
+ggplot() + geom_line(data=df, aes(specificity, sensitivity)) + scale_x_reverse() + 
+  geom_point(data=thresholds, aes(x=specificity,y=sensitivity,color=label), size=3,shape=8, show.legend = T) +
+  geom_label_repel(data=thresholds, aes(x=specificity,y=sensitivity,fill=label,label=
+    paste('Threshold ', round(threshold,2), '\n(FPR ', round((1-specificity),2)*100, '%, TPR ', round(sensitivity,2)*100,'%)\nAUC:',round(roc$auc, 2)*100, '%', sep='')),show.legend = F, color='white',nudge_x=0.05) + scale_fill_manual(values=brewer.pal(3,'Dark2')) + scale_color_manual(values=brewer.pal(3,'Dark2')) +
+  labs(title='CN Model ROC', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme
+
+roc.plot(roc, 'Probability threshold calculated by ROC')
+
+roc.plot(rocNOHGD, 'Model excluding HGD/IMC')
+pROC::coords(rocNOHGD, "best")
+pROC::coords(rocNOHGD, 0.6)
+pROC::auc(rocNOHGD, partial.auc=c(0.89,0.9), partial.auc.focus='spec', partial.auc.correct=T)
 
@@
 
+### Per Endoscopy ROCs
+```{r, echo=F, warning=F, message=F,fig.height=4, fig.width=4}
+
+pendo = do.call(rbind, pg.samp) %>% group_by(Patient, Status, Hospital.Research.ID, PID ) %>% summarise(
+  'max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(PID)
+)
+
+rocEndoM = pROC::roc(Status ~ max, data=pendo, ci=T, of='thresholds')
+pROC::coords(rocEndoM, 'best')
+#roc.plot(rocEndoM)
+rocEndoM$model = 'Per Endoscopy Max'
+
+rocEndoA = pROC::roc(Status ~ avg, data=pendo, ci=T, of='thresholds')
+pROC::coords(rocEndoA, 'best')
+#roc.plot(rocEndoA)
+rocEndoA$model = 'Per Endoscopy Avg'
+
+lastEndo = do.call(rbind, lapply(pg.samp, function(pt) {
+ cbind(pt[1,c('Patient','Hospital.Research.ID', 'Status')],  'avg'=mean(subset(pt, PID == pt[nrow(pt), 'PID'])$Prediction))
+}))
+
+rocLastEndo = pROC::roc(Status ~ avg, data=lastEndo, ci=T, of='thresholds')
+pROC::coords(rocLastEndo, 'best')
+#roc.plot(rocLastEndo)
+rocLastEndo$model = 'Last endoscopy'
+
+firstEndo = do.call(rbind, lapply(pg.samp, function(pt) {
+ cbind(pt[1,c('Patient','Hospital.Research.ID', 'Status')],  'avg'=mean(subset(pt, PID == pt[1, 'PID'])$Prediction))
+}))
+
+rocFirstEndo = pROC::roc(Status ~ avg, data=firstEndo, ci=T, of='thresholds')
+pROC::coords(rocFirstEndo, 'best')
+#roc.plot(rocFirstEndo)
+rocFirstEndo$model = 'First endoscopy'
+
+```
+
+### Per Patient ROC
+
+```{r echo=F, fig.height=4, fig.width=4}
+
+ptendo = do.call(rbind, pg.samp) %>% group_by(Patient, Hospital.Research.ID, Status) %>% summarise( 'max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(Patient))
+
+rocPtM = pROC::roc(Status ~ max, data=ptendo, ci=T, of='thresholds')
+pROC::coords(rocPtM, 'best')
+#roc.plot(rocPtM)
+rocPtM$model = 'Per Patient Max'
+
+rocPtA = pROC::roc(Status ~ avg, data=ptendo, ci=T, of='thresholds')
+pROC::coords(rocPtA, 'best')
+#roc.plot(rocPtA)
+rocPtA$model = 'Per Patient Avg'
+
+```
+
 ### Compare ROCs
@@
 ```{r echo=F, fig.height=6, fig.width=8}
-get.roc.stat<-function(roc) {
+roc$model = "Per Sample"
+multi.roc.plot(list(roc, rocEndoM, rocEndoA, rocPtM, rocPtA,rocFirstEndo,rocLastEndo), 'ROC Comparisons')
+```
+
+```{r echo=F, fig.height=8, fig.width=6}
+get.roc.stat<-function(roc, x='best') {
   auc = c( pROC::auc(roc), range(pROC::ci.auc(roc))  )
 
-  sens = c(pROC::coords(roc, 'best')[['sensitivity']], pROC::ci.se(roc,pROC::coords(roc, 'best')[['specificity']], conf.level=0.95))
-  spec = c(pROC::coords(roc, 'best')[['specificity']], pROC::ci.sp(roc,pROC::coords(roc, 'best')[['sensitivity']], conf.level=0.95))
+  sens = c(pROC::coords(roc, x)[['sensitivity']], pROC::ci.se(roc,pROC::coords(roc, x)[['specificity']], conf.level=0.95))
+  spec = c(pROC::coords(roc, x)[['specificity']], pROC::ci.sp(roc,pROC::coords(roc, x)[['sensitivity']], conf.level=0.95))
 
@@
 
+get.auc.at<-function(roc, foc='spec', p=c(1,0.99)) {
+# To find the auc at a given spec or sens
+  pROC::auc(roc, partial.auc=p, partial.auc.focus=foc, partial.auc.correct=T)
+
+}
+
+get.auc.at(p53roc)
+get.auc.at(hgdpathroc)
+get.auc.at(pathroc)
+get.auc.at(roc)
+
 m = rbind.data.frame(
@@
       get.roc.stat(pathroc),
-      get.roc.stat(roc))
+      get.roc.stat(roc, 0.6))
 m$model = factor(m$model, levels=c('p53 IHC','HGD Only','Dysplasia (HGD & LGD)','CN Model'), ordered = T)
@@
 
-aucs = do.call(rbind, lapply(rocList,function(r) cbind.data.frame('AUC'=r$auc*100,'model'=r$model,'Specificity'=pROC::coords(r,'best')[['specificity']], 'Sensitivity'=pROC::coords(r,'best')[['sensitivity']])))
-
-df = do.call(rbind, lapply(rocList, function(r) 
-              cbind.data.frame('Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'model'=r$model) 
-              ))
-
-ggplot(df, aes(Specificity, Sensitivity,color=model,fill=model)) + geom_line(show.legend = F) + scale_x_reverse() + 
-    geom_label_repel(data=aucs, aes(label=paste(model,'\nAUC:', sprintf("%.2f",AUC), '%', sep='')), nudge_x=.25, color='white', show.legend = F) + 
-  labs(title='ROC Curves', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme
-
 dodge = position_dodge(width=0.9)
-ggplot(melt(m, measure.vars='value'), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Measure)) + 
+ggplot(melt(subset(m, Measure == 'AUC'), measure.vars='value'), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Measure)) + 
   geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
@@
   scale_fill_manual(values=c('purple3','dodgerblue3','firebrick3'), name='') + plot.theme + 
-  theme(legend.position = 'bottom') + labs(x='', y='Percent', title='ROC Comparison') 
+  theme(legend.position = 'none') + labs(x='', y='Percent', title='ROC Comparison') 
 ```
@@
 transform.by.time<-function(pt, info) {
-  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction')] %>% group_by(PID, Samplename) %>% mutate(
-  'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12
-  )
+  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction', 'OR')] %>% group_by(PID, Samplename) %>% mutate(
+  'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12)
 
@@
 back$Pathology = as.character(back$Pathology)
-back$Pathology[which(back$Pathology %in% c('HGD','IMC'))] = 'IMC/HGD'
-back$Pathology = factor(back$Pathology, levels=c('BE','ID','LGD','IMC/HGD'), ordered = T)
+#back$Pathology[which(back$Pathology %in% c('HGD','IMC'))] = 'IMC/HGD'
+back$Pathology = factor(back$Pathology, levels=c('BE','ID','LGD','HGD', 'IMC'), ordered = T)
 
@@
 cuts = seq(0,1,0.1)
-#cuts = quantile(back$Prediction, probs=seq(0,1,1/5))
 
@@
 
-qt = cbind(qt, back %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction), 'se'=(sd(Prediction)/sqrt(length(Prediction)))*1.96 ))
-
-ggplot(qt,aes(quants, mn, ymin=mn-se, ymax=mn+se)) + geom_point() + geom_errorbar() + labs(x='Deciles', y='Mean probability', title='CI around probabilities by decile')
+qt = cbind(qt, back %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))#, 'se'=(sd(Prediction)/sqrt(length(Prediction)))*1.96 ))
+#ggplot(qt,aes(quants, mn, ymin=mn-se, ymax=mn+se)) + geom_point() + geom_errorbar() + labs(x='Deciles', y='Mean probability', title='CI around probabilities by decile')
 
@@
   'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )
+
 pred.confidence[c('r1','r2')] = NA
@@
 
+pred.confidence = bind_cols(pred.confidence, 
+                            data.frame(ci.low=qbeta(0.025, shape1=pred.confidence$P+.5, shape2 = pred.confidence$NP+.5),
+                                      ci.high=qbeta(0.975, shape1=pred.confidence$P+.5, shape2 = pred.confidence$NP+.5)))
+
 annotation.table = arrange(pred.confidence %>% group_by(Risk) %>% summarise('n(NP)'=sum(NP), 'n(P)'=sum(P), 'p.value'=format.pval(mean(p.value), 3), 'min'=min(r1), 'max'=max(r2)), min)
 
-#annotation.table = as.data.frame(annotation.table)
-#rownames(annotation.table) = annotation.table$Risk
-#annotation.table$Risk = NULL
-
 tt3 <- ttheme_minimal(
-  core=list(bg_params = list(fill = c('green3','orange2','red3'), col=NA, alpha=0.2),
+  core=list(bg_params = list(fill = riskCols, col=NA, alpha=0.2),
             fg_params=list(fontface=3)),
@@
 
-grid.arrange( tableGrob(annotation.table, rows=NULL, theme=tt3) )
-
-riskCols = c('red3','orange','green3')
-grid.arrange(
-  ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), bins=20, show.legend = F) +
-    scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
-    #geom_vline(xintercept=pred.confidence$r2, linetype='longdash', color='grey39') +
-    #geom_label(data=(pred.confidence %>% rowwise %>% mutate( 'mid'=sum(r1,r2)/2 )), aes(x=mid, y=75,label=paste(round(perc,2),conf,sep=''),color=conf ), show.legend=F) + scale_color_manual(values=c('grey','black')) +
+#grid.arrange( tableGrob(annotation.table, rows=NULL, theme=tt3) )
+pred.confidence$Risk = factor(pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)
+```
+
+```{r pcalib, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
+ggplot(pred.confidence, aes(mn, perc)) + 
+  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
+  geom_point() + geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
+  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration') +
+  scale_color_manual(values=riskCols) + plot.theme + theme(legend.position = 'none')
+```
+
+```{r probhist, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
+ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') +
     geom_vline(xintercept=annotation.table$max, linetype='longdash', color='grey39') +
     geom_text(data=annotation.table %>% group_by(Risk, min,max) %>% summarise( 'samp.perc'= signif((sum(`n(NP)`, `n(P)`)/nrow(patient.info))*100, 3) ), aes(x=min+(max-min)/2, y=5, label=paste(samp.perc, '%',sep=''))) +
-    plot.theme + labs(title='Sample predictions', y='n Samples', x='Probability')
-
-  ,tableGrob(annotation.table, rows=NULL, theme=tt3),
-  nrow=2, heights=c(8,2)
-  )
-
+    plot.theme + labs(title='Sample predictions', y='n Samples', x='Probability') + theme(legend.position = 'none')
+```
+
+```{r indP, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
 back = merge(back, pred.confidence[c('quant','Risk')], by.x='quants',by.y='quant')
@@
 ggplot(back, aes(1:nrow(back), Prediction, color=Risk)) + 
-  geom_point(aes(color=Risk, shape=Status)) + geom_errorbar(aes(ymin=Prediction-`se(Risk)`, ymax=Prediction+`se(Risk)`)) +
+  geom_point(aes(color=Risk, shape=Status)) +scale_color_manual(values=riskCols, name='P(P)') + 
   geom_hline(data=annotation.table, aes(yintercept=max), color='grey') + plot.theme +
   labs(x='sample', y='model probability', title='Per-sample probabilities')
+```
+
+```{r relrisk, echo=F, fig.width=6, fig.height=6}
+#myPal = c(RColorBrewer::brewer.pal(9, 'Blues')[9:7], 
+#          adjustcolor(grDevices::heat.colors(10)[6], blue.f=1.5),
+#          ( grDevices::heat.colors(10)[7:6]),
+#          RColorBrewer::brewer.pal(9, 'Reds')[6:9])
+
+myPal = rev(RColorBrewer::brewer.pal(11, 'RdYlBu'))
+
+ggplot(back, aes(OR)) + geom_histogram(aes(fill=..x..), bins=10, show.legend = F) +
+  #scale_fill_distiller(palette = 'RdYlBu', type='div',name='P(P)') +
+  scale_fill_gradientn(colors = myPal,  name='') + 
+  labs(y='n Samples', x='Relative Risk', title='Unadjusted relative risk') + plot.theme
+
+ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) +
+    scale_fill_gradientn(colors = myPal,  name='') + 
+  #scale_fill_distiller(palette = 'RdYlBu', type='div',name='P(P)') +
+    plot.theme + labs(title='Predictions, all samples model', y='n Samples', x='Probability')
+
+ggplot(back, aes(OR, Prediction, color=Prediction)) + geom_point(show.legend = F) + 
+  scale_color_gradientn(colors = myPal,  name='')  + plot.theme
+```
+
+### Mountain plots using spatial info
+
+```{r, echo=F, warning=F, message=F, fig.height=12, fig.width=16}
+
+mtn.spatial<-function(df, b, limits, title='', pal=NULL) {
+  if (is.null(pal))
+    pal = RColorBrewer::brewer.pal(length(limits), 'Paired')
+
+  df = merge(df, b[,c('Samplename','Endoscopy.Year','months.before.final','Block','nl')], by.x='variable',by.y='Samplename')
+  df$nl = factor(df$nl, ordered = T)
+
+  #pal = pal[which(limits %in% levels(df$nl))]
+
+  ggplot(df, aes(x=chr.length)) + facet_grid(Endoscopy.Year~chr, scales='free_x', space='free_x') +
+    geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=nl), show.legend=T) + 
+    scale_fill_manual(values=pal, limits=limits,  name='OGJ Dist.') +
+    theme(axis.text.x = element_blank()) +
+    plot.theme + labs(title=title)
+}
+
+ogjLimits = levels(factor(back$nl, ordered = T))
+
+p1 = mtn.spatial(subset(globalCNdf, Patient == 48 & seg.type != 'arms'), back, ogjLimits, 'Patient 48, (P)')
+ogjLegend = get.legend(p1)
+p1+theme(legend.position = 'none')
+
+mtn.spatial(subset(globalCNdf, Patient == 60 & seg.type != 'arms'), back,ogjLimits,'Patient 60 (P)' )+theme(legend.position = 'none')
+
+mtn.spatial(subset(globalCNdf, Patient == 3 & seg.type != 'arms'), back,ogjLimits,'Patient 3 (NP)' )+theme(legend.position = 'none')
+
+mtn.spatial(subset(globalCNdf, Patient == 15 & seg.type != 'arms'), back,ogjLimits,'Patient 15 (NP)' )+theme(legend.position = 'none')
+
 ```
@@
 
-```{r jacknife, echo=T, warning=F, message=F, fig.height=5, fig.width=6}
+```{r jacknife, echo=T, warning=F, message=F, fig.height=10, fig.width=12}
 
@@
 riskPath$ratio = with(riskPath, round(Freq/nPath, 2))
-riskPath$Risk = factor(riskPath$Risk, levels=c('High','Moderate','Low'), ordered = T)
+riskPath$Risk = factor(riskPath$Risk, levels=c('Low','Moderate','High'), ordered = T)
 
@@
 ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
-  scale_fill_manual(values=c('red2','orange','green3')) + 
+  scale_fill_manual(values=riskCols) + 
   geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), position=position_stack(), size=4) +
@@
 
+```
+
+```{r p53IHC, echo=F, warning=F, message=F, fig.height=6, fig.width=8}
+
+p53Path$Risk[p53Path$Risk == 'High'] = 'Aberrant'
+p53Path$Risk[p53Path$Risk == 'Low'] = 'Normal'
+
+col = brewer.pal(11, "PRGn")[c(3,9)]
 ggplot(p53Path, aes(Pathology, ratio*100, group=Pathology, fill=Risk)) + geom_bar(stat='identity', position=position_stack()) + 
-  scale_fill_manual(values=c('red2','green3')) + 
+  scale_fill_manual(values=col) + 
   geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101)) +
@@
   geom_bar(stat='identity', position = position_stack()) +
-  scale_fill_manual(values=c('green3','orange2','red')) + 
+  scale_fill_manual(values=riskCols) + 
   geom_text(aes(y=1.01, label=n.samples)) +
@@
 
-```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=10 }
+```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=15 }
 back$nl = factor(back$nl, ordered=T)
-back$labels = factor(back$Risk, levels=c('High','Moderate','Low'))
-
- pred.tiles(subset(back, Patient %in% c(13,20,34,56,60,48) & Status == 'P'), nrow=1) + labs(title='Progressors')
-
-grid.arrange( pred.tiles(subset(back, Patient %in% c(13,20,34,56,60,48) & Status == 'P'), ncol=1) + labs(title='Progressors') + theme(legend.position = 'none'),
-              pred.tiles(subset(back, Patient %in% c(15,16,21,3,45,19) & Status == 'NP'), ncol=1) + labs(title='Non-Progressors'), ncol=2 )
+back$Risk = factor(back$Risk, levels=c('Low','Moderate','High'), ordered = T)
+back$Patient = factor(as.numeric(back$Patient), ordered = T)
+
+pred.tiles(subset(back, Patient %in% c(6)), ncol=1, cols=myPal, probs=T) + labs(title='') + theme(legend.position = 'none')
+pred.tiles(subset(back, Patient %in% c(48)), ncol=1, cols=myPal, probs=T) + labs(title='') + theme(legend.position = 'none')
+pred.tiles(subset(back, Patient %in% c(33)), ncol=1, cols=myPal, probs=T) + labs(title='')+ theme(legend.position = 'none')
+pred.tiles(subset(back, Patient %in% c(89)), ncol=1, cols=myPal, probs=T) + labs(title='')+ theme(legend.position = 'none')
+
+grid.arrange(
+  pred.tiles(subset(back, Patient %in% c(13,20,34,56,60,48) & Status == 'P'), ncol=1, cols=myPal, probs=T) + labs(title='Progressors'),
+  pred.tiles(subset(back, Patient %in% c(15,16,21,3,45,19) & Status == 'NP'), ncol=1, cols=myPal, probs=T) + labs(title='Non-Progressors'),
+ncol=2)
 
@@
 
-```{r, echo=F}
-
-p = pred.tiles(subset(back, Status == 'NP'), nrow=2) + labs(title='Non-Progressors')
+```{r, echo=F, message=F, warning=F}
+back = back %>% group_by(Patient) %>% mutate( 'pred.ratio'=round(sum(Prediction)/length(Prediction),2))
+
+p = pred.tiles(transform(subset(back, Status == 'NP'), Patient=reorder(Patient, pred.ratio)), cols=myPal, probs=T, nrow=2) + labs(title='Non-Progressors') 
+riskLegend = get.legend(p)
+p = p + theme(legend.position = 'none')
 ggsave('npg_heatmap.png', plot = p, scale = 1, width = 52, height = 8, units = "in", limitsize = F, dpi = 300)
 
-p = pred.tiles(subset(back, Status == 'P'), nrow=2) + labs(title='Progressors')
+p = pred.tiles(transform(subset(back, Status == 'P'), Patient=reorder(Patient, -pred.ratio)), probs=T, cols=myPal,nrow=2) + labs(title='Progressors') + theme(legend.position = 'none')
 ggsave('pg_heatmap.png', plot = p, scale = 1, width = 52, height = 8, units = "in", limitsize = F, dpi = 300)
 
+ggsave('risk_legend.png', plot=grid.arrange(riskLegend), scale=1, dpi=300)
+
 ```
@@
 ## p arm
-```{r, echo=F}
+```{r, echo=T}
 
 p11 = subset(globalCNdf, chr == 17 &  start == 1 & seg.type == 'arms')
-
 p11Loss = subset(p11, value < -1)
@@
 
-# Samples with a loss vs all samples
+table(p11Loss[c('Patient', 'Pathology')])
+
+# Samples with a loss vs all Progressor samples
 length(unique(p11Loss$variable))/length(unique(subset(globalCNdf, Status == 'Progressor')$variable))
@@
 
-pander(table(subset(patient.info, Samplename %in% p11Loss$variable, select=c('p53.Status', 'Pathology'))))
+(table(subset(patient.info, Samplename %in% p11Loss$variable, select=c('p53.Status', 'Pathology'))))
+
+pi = patient.info
+pi$p11Loss = as.numeric(pi$Samplename %in% p11Loss$variable)
+
+plot(table(p11=pi$p11Loss, ihc=pi$p53.Status))
+
+cor.test(pi$p11Loss, (as.numeric(pi$p53.Status)-1))
+
+# So why might there be loss but no corresponding aberrant staining?
+table(subset(pi, p11Loss == 1 & p53.Status == 0)$Pathology)
+table(subset(pi, p11Loss == 1 & p53.Status == 1)$Pathology)
+
+```
+
+## Top Features
+```{r, echo=T}
+
+m = data.frame(matrix(nrow=nrow(chh), ncol=0))
+
+for (i in 1:nrow(chh)) {
+  feat = chh[i,]
+
+  seg = subset(globalCNdf, chr == as.character(feat$chr) & start == feat$start & end == feat$end & seg.type != 'arms')  
+  if(feat$end-feat$start > 5e6)
+    seg = subset(globalCNdf, chr == as.character(feat$chr) & start == feat$start & end == feat$end & seg.type == 'arms')
+
+  segGL = subset(seg, value < -1)
+  if(feat$coef > 0)
+    segGL = subset(seg, value > 1)
+
+  m[i, 'n.samples'] = length(unique(segGL$variable))
+  m[i,'ratio.samples.P'] = length(unique(segGL$variable))/length(unique(subset(globalCNdf, Status == 'Progressor')$variable))
+  m[i, 'ratio.NP'] = table(segGL$Status)[1]/sum(table(segGL$Status))
+  m[i, 'ratio.P'] = table(segGL$Status)[2]/sum(table(segGL$Status))
+
+  segpath = table(subset(patient.info, Samplename %in% segGL$variable)$Pathology)
+
+  m[i,'BE'] = segpath['BE']/sum(segpath)
+  m[i,'HGD'] = sum(segpath[c('HGD','IMC')])/sum(segpath)
+}
+rownames(m) = chh$label
+
+range(m$ratio.samples.P)
+summary(m$ratio.samples.P)
+subset(m, ratio.samples.P == max(ratio.samples.P))
+subset(chh, label == rownames(subset(m, ratio.samples.P == max(ratio.samples.P))))
+subset(chh, label == rownames(subset(m, ratio.samples.P == min(ratio.samples.P))))
+#table(cdkGain[c('Patient', 'Pathology')])
+
+cor.test(chh$hazards, m$n.samples)
+
+# Samples with a gain vs all samples
 
@@
 One way of looking at the p53 prediction (e.g. Adam Bass) is to see how predictive the p-arm loss in chr17 is.
-```{r, echo=F, warning=F, message=F, fig.height=8, fig.width=8}
+```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=10}
 ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
-ccgenes = ccgenes[-grep('X,Y', ccgenes$Genome.Location),]
-ccgenes = ccgenes[-which(ccgenes$Mutation.Types == 'T'),] # drop the translocations
-
-genes = subset(ccgenes, grepl('TSG|oncogene', Role.in.Cancer) & !grepl('^X|Y', Chr.Band), select=c('Gene.Symbol','Genome.Location'))
-
-loc = unlist(strsplit(genes$Genome.Location, ':|-'))
-loc = do.call(rbind.data.frame, strsplit(genes$Genome.Location, ':|-'))
+ccgenes = ccgenes[-grep('^X|Y', ccgenes$Genome.Location),]
+#ccgenes = ccgenes[-which(ccgenes$Mutation.Types == 'T'),] # drop the translocations
+#ccgenes = subset(ccgenes, !grepl('^X|Y', Genome.Location))
+#loc = unlist(strsplit(ccgenes$Genome.Location, ':|-'))
+loc = do.call(rbind.data.frame, strsplit(ccgenes$Genome.Location, ':|-'))
 colnames(loc) = c('chr','start','end')
@@
 
-genes = cbind(genes, loc, 'value'=min(globalCNdf$mean.value), 'Row.names'='', 'Patient'='', 'selected'=as.factor(0))
-
-plot.cn.chr(subset(globalCNdf, seg.type != 'arms'), chrom='17', chr.lengths, genes=genes, haz=chh) + labs(
-title="Segment Values Samples", subtitle=paste('NonP:Prog - ', paste(table(patient.info$Status), collapse=':')))
-```
-
-```{r echo=F, warning=F, message=F}
-
-plot.cn.chr(globalCNdf, chrom='8', chr.lengths, genes=genes, haz=chh)
-
-plot.cn.chr(globalCNdf, chrom='3', chr.lengths, genes=genes, haz=chh)
-
-plot.cn.chr(globalCNdf, chrom='21', chr.lengths, haz=chh, genes=genes)
-
-plot.cn.chr(globalCNdf, chrom='11', chr.lengths, haz=chh, genes=genes)
-
-plot.cn.chr(globalCNdf, chrom='16', chr.lengths, haz=chh, genes=genes)
-
-plot.cn.chr(globalCNdf, chrom='9', chr.lengths, haz=chh, genes=genes)
+ccgenes = cbind(ccgenes, loc)
+ccgenes = ccgenes[c('Gene.Symbol','Genome.Location', 'Role.in.Cancer', 'Mutation.Types',colnames(loc))]
+#ccgenes = subset(ccgenes, grepl('\\w+',Role.in.Cancer))
+
+genes = cbind(subset(ccgenes, grepl('TSG|oncogene', Role.in.Cancer) & Mutation.Types != 'T', select=c('Gene.Symbol','Genome.Location', colnames(loc))), 'value'=min(globalCNdf$mean.value), 'Row.names'='', 'Patient'='', 'selected'=as.factor(0))
 
@@
 ov = findOverlaps(chGR, genesGR)
-
-#as.data.frame(chGR[unique(queryHits(ov))])
-#length(genesGR[unique(subjectHits(ov))]$Gene.Symbol)
-write.table(genesGR[unique(subjectHits(ov))]$Gene.Symbol, quote=F, row.names=F, sep='\t')
-
-pander(as.data.frame(genesGR[unique(subjectHits(ov))])[,c(1:3,6,7)])
-
-```
+genes = as.data.frame(genesGR[unique(subjectHits(ov))])
+colnames(genes)[1] = c('chr')
+
+plot.cn.chr(globalCNdf, chrom='17', chr.lengths, genes=genes, haz=chh) + labs(
+title="Segment Values Samples", subtitle=paste('NonP:Prog - ', paste(table(patient.info$Status), collapse=':'))) + theme(legend.position = 'none')
+```
+
+```{r echo=F, warning=F, message=F, fig.height=5, fig.width=10}
+chGR$Genes = ''
+for (hit in unique(queryHits(ov))) {
+  chGR[hit]$Genes = paste(genesGR[subjectHits(ov[queryHits(ov) == hit])]$Gene.Symbol, collapse=',')
+}
+
+for (chr in unique(chh$chr)) 
+  print(plot.cn.chr(subset(globalCNdf, seg.type != 'arms') , chrom=chr, chr.lengths, genes=genes, haz=chh))
+
+annoG = arrange(subset(ccgenes, grepl('CDKN2A$|TP53|MYC$|SMAD4|VEG|ERBB2|EGFR|SMYD|ARID1A|FHIT|RUNX1', Gene.Symbol)), chr)
+annoG = merge(annoG, chr.lengths[c('chrom','chr.length')], by.x='chr',by.y='chrom')
+annoG$ymin = round(range(globalCN.plot$data$mean.value))[1]/2
+annoG$ymax = round(range(globalCN.plot$data$mean.value))[2]/2
+
+globalCN.plot + #geom_rect(data=annoG, aes(xmin=start, xmax=end, ymin=ymin,ymax=ymax), color='red') +
+  geom_point(data=annoG, aes(x=start, y=0), size=1.5, color='red') +
+  geom_text_repel(data=annoG, aes(x=start, y=ymax/2, label=Gene.Symbol), color='red', size=3 )
+
+```
+
+`r pander(genes[,c(1:3,6,7)])`
 
@@
 
-mp$performance
+#mp$performance
 
@@
 
-tsgs = subset(ccgenes, grepl('TSG|oncogene', Role.in.Cancer))
-tsgs = tsgs[grep('^X|Y',tsgs$Chr.Band, invert=T),]
+tsgs = ccgenes
 nrow(tsgs)
 
-loc = do.call(rbind.data.frame, strsplit(tsgs$Genome.Location, ':|-'))
-colnames(loc) = c('chr','start','end')
-loc[] = lapply(loc[], function(x) as.numeric(as.character(x)))
-
-tsgs = cbind(tsgs, loc)
-tsgs = makeGRangesFromDataFrame(tsgs[,c(colnames(loc), 'Gene.Symbol','Role.in.Cancer', 'Mutation.Types')], keep.extra.columns = T)
+tsgs = makeGRangesFromDataFrame(tsgs[,c(colnames(loc), 'Gene.Symbol')], keep.extra.columns = T)
 tsgs = sort(tsgs)
@@
 
-ch = arrange(ch, -hazards)
-features = makeGRangesFromDataFrame(ch, keep.extra.columns = T)
+chh = arrange(chh, -hazards)
+features = makeGRangesFromDataFrame(chh, keep.extra.columns = T)
 
@@
 valCNdf$mean.value = valCNdf$value # but not going to use it
+
 cn.mtn.plot(subset(valCNdf, seg.type != 'arms'), labeller(Status=n.status, chr=label_value))
-
-head(valCNdf)
 
@@
 pred = predict(fitV, newx=validation.df, s=lambda.opt, type='response')
+or = predict(fitV, newx=validation.df, s=lambda.opt, type='link')
 
@@
 
-for (pt in names(vpd)) 
+for (pt in names(vpd)) {
   vpd[[pt]]$Prediction[which(vpd[[pt]]$Samplename %in% rownames(pred))] = pred[ intersect(rownames(pred), vpd[[pt]]$Samplename), ]
+  vpd[[pt]]$OR[which(vpd[[pt]]$Samplename %in% rownames(or))] = or[ intersect(rownames(or), vpd[[pt]]$Samplename), ]
+}
 
@@
 back.v$Pathology = as.character(back.v$Pathology)
-back.v$Pathology[which(back.v$Pathology %in% c('HGD','IMC'))] = 'IMC/HGD'
-back.v$Pathology = factor(back.v$Pathology, levels=c('BE','ID','LGD','IMC/HGD'), ordered = T)
-
-```
-
-```{r echo=F, fig.height=14,fig.width=10}
+#back.v$Pathology[which(back.v$Pathology %in% c('HGD','IMC'))] = 'IMC/HGD'
+back.v$Pathology = factor(back.v$Pathology, levels=c('BE','ID','LGD','HGD','IMC'), ordered = T)
+
+```
+
+```{r echo=F, message=F, warning=F, fig.height=8, fig.width=10}
 back.v$quants = with(back.v, cut(Prediction, breaks=cuts))
@@
 ggplot(m, aes(Patient,  value*100, group=variable, fill=variable)) + geom_bar(stat='identity', position=position_stack()) +
-  geom_text(aes(label=n.samples, y=101)) + scale_fill_manual(values=c('green3','orange2','red2')) +
+  geom_text(aes(label=n.samples, y=101)) + scale_fill_manual(values=riskCols) +
   facet_grid(~Status, scales='free_x') + labs(y='Percent of samples', title='Validation patient set')
-
+```
+
+```{r echo=F, warning=F, message=F, fig.height=14,fig.width=28}
 grid.arrange(
-  pred.tiles(subset(back.v, Status == 'P')) + labs(title='Validation Progressors') ,
-  pred.tiles(subset(back.v, Status == 'NP')) + labs(title='Validation Non-Progressors'),
+  pred.tiles(subset(back.v, Status == 'P'), cols=myPal, probs=T) + labs(title='Validation Progressors') + theme(legend.position = 'none'),
+  pred.tiles(subset(back.v, Status == 'NP'), cols=myPal, probs=T) + labs(title='Validation Non-Progressors') + theme(legend.position = 'none'),
 ncol=2)
-
+```
+
+```{r echo=F, warning=F, message=F, fig.height=5,fig.width=8}
+vig.plot(subset(back.v, Patient == 68), subset(validation.patient.info, Patient == 68), cols=riskCols)
 ```
@@
 
-riskPath = rbind(riskPath, p53Path[colnames(riskPath)])
 riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor','p53 IHC'), ordered = T)
@@
 ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
-  scale_fill_manual(values=c('red','orange2','green3')) + 
+  scale_fill_manual(values=riskCols) + 
   geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), position=position_stack(), size=4) +
@@
 
-## Clinical Story
-
-First note here is that our probability distribution for these 20 patients is somewhat flatter than in the training set.
-Likely a consequence of sample size.
-The labels show the proprotions of progressors that fell into each quintile in the training set.
-```{r, echo=F, warning=F, message=F}
-
-summary.v = table.path.risk(back.v)
-
-pander(as.data.frame(summary.v))
-
-write.table(summary.v, sep='\t', quote=F, row.names = F)
-```
-
-### Vignetts
-
-```{r, echo=F, warning=F, message=F, fig.width=12, fig.height=4}
-vign.pts = as_tibble(subset(sum.patient.data, Patient %in% c(3, 33, 34, 49, 56, 20), select=c('Status','Patient','age.diagnosis', 'smoking', 'C','M', 'initial.endo','final.endo', 'total.endos','total.samples')))
-colnames(vign.pts) = c('Status','Patient','Age at Diagnosis','Smoker','Circumference', 'Maximal length', 'Intial Endoscopy','Final Endoscopy', 'Total Endoscopies','Total Samples')
-
-back$Risk = factor(back$Risk, levels=levels(back$labels), ordered = T)
-
-vig.plot<-function(df, info) {
-  scale = seq(min(info$Endoscopy.Year), max(info$Endoscopy.Year), 1)
-  subtitle = with(info, paste(unique(Age.at.diagnosis),unique(Sex),': ', sep=''))
-  if (!is.na(unique(info$Smoking)))                             
-    subtitle = paste(subtitle, ifelse(unique(info$Smoking) == 'Y', 'smoker, ','non-smoker, '),sep='')
-
-  subtitle = paste(subtitle, 'M', unique(info$Maximal), sep='')
-
-  if (!is.na(unique(info$Circumference)))
-    subtitle = paste(subtitle, 'C', info$Circumference, sep='')
-
-  subtitle = paste(subtitle, " BE diagnosed ", unique(info$Initial.Endoscopy), sep='')
-
-  ggplot(df, aes(Endoscopy.Year, nl)) + geom_tile(aes(fill=Risk), color='white') +
-    scale_fill_manual(values=c('red3','orange', 'green3'), limits=levels(df$Risk),  name='Progression') +
-    geom_text_repel(aes(label=Pathology), color='white', show.legend=F) +
-    labs(y='Oesophageal Location',x='Months Prior to Endpoint', title='') +
-    plot.theme + theme(axis.text.x=element_text(angle=45, hjust=1),legend.position='none') + 
-    scale_x_continuous(breaks=scale) +
-    labs(title=paste('Patient',unique(info$Patient)), subtitle=subtitle)
-}
-
-## P
-vig.plot(df=subset(back, Patient == 34),info= subset(patient.info, Patient == 34))
-vig.plot(subset(back, Patient == 56), subset(patient.info, Patient == 56))
-vig.plot(subset(back, Patient == 20), subset(patient.info, Patient == 20))
-vig.plot(subset(back, Patient == 86), subset(patient.info, Patient == 86))
-
-## NP
-vig.plot(subset(back, Patient == 3), subset(patient.info, Patient == 3))
-vig.plot(df=subset(back, Patient == 23),info= subset(patient.info, Patient == 23))
-vig.plot(subset(back, Patient == 33), subset(patient.info, Patient == 33))
-vig.plot(subset(back, Patient == 49), subset(patient.info, Patient == 49))
-vig.plot(subset(back, Patient == 51), subset(patient.info, Patient == 51))
-
-```
-
-```{r, echo=F, warning=F, message=F}
+## Validation prediction calibration
+
+```{r, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
+cuts = seq(0,1,0.1)
+qt = as.data.frame.matrix(table(back.v$quants, back.v$Status))
+qt$quant = rownames(qt)
+
+qt = cbind(qt, back.v %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
+
+pred.confidence.v = qt %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 'perc'=P/sum(NP,P) )
+
+pred.confidence.v[c('r1','r2')] = NA
+for (i in 1:(length(cuts)-1)) 
+    pred.confidence.v[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)
+
+pred.confidence$Risk = factor(pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)
+
+ggplot(pred.confidence.v, aes(mn, perc)) + 
+  geom_rect(data=pred.confidence, aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
+  geom_point() +
+  #geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + 
+  labs(x='Mean prediction', y='Ratio of `P` patients', title='Validation calibration') +
+  scale_color_manual(values=riskCols) + plot.theme
+
+preds = do.call(rbind.data.frame, lapply(vpd, function(df) df[c('Status','Prediction')]))
+vroc = pROC::roc(Status ~ Prediction, data=preds, ci=T, of='thresholds')
+vroc$model = 'Validation'
+
+roc$model = 'Discovery'
+multi.roc.plot(list(roc, vroc), title='ROC')
+
+```
+
+```{r, echo=F, warning=F, message=F, eval=F}
 ggplot(back.v, aes(Prediction)) + geom_histogram(aes(fill=..x..), bins=20, show.legend = F) +
@@
 
+# Clinical Story
+
+First note here is that our probability distribution for these 20 patients is somewhat flatter than in the training set.
+Likely a consequence of sample size.
+The labels show the proprotions of progressors that fell into each quintile in the training set.
+
+```{r, echo=F, warning=F, message=F}
+
+summary.v = table.path.risk(back.v)
+
+pander(as.data.frame(summary.v))
+
+write.table(summary.v, sep='\t', quote=F, row.names = F)
+```
+
+## Vignettes
+
+```{r, echo=F, warning=F, message=F, fig.width=8, fig.height=5}
+vign.pts = as_tibble(subset(sum.patient.data, Patient %in% c(3, 33, 34, 49, 56, 20), select=c('Status','Patient','age.diagnosis', 'smoking', 'C','M', 'initial.endo','final.endo', 'total.endos','total.samples')))
+colnames(vign.pts) = c('Status','Patient','Age at Diagnosis','Smoker','Circumference', 'Maximal length', 'Intial Endoscopy','Final Endoscopy', 'Total Endoscopies','Total Samples')
+
+#back$Risk = factor(back$Risk, levels=levels(back$labels), ordered = T)
+
+# ggplot(subset(back, Status == 'P'), aes(factor(Endoscopy.Year, ordered=T), nl)) + geom_tile(aes(fill=Risk, color=p53.Status), size=1.5)  + facet_wrap(~Patient, scales='free') + 
+#   scale_fill_manual(values=riskCols, limits=levels(back$Risk),name='Progression') + 
+#   scale_color_manual(values=c('grey22','white'), limits=levels(back$p53.Status), drop=F, name='p53 IHC', labels=c('normal','aberrant')) +
+#     geom_text(aes(label=Pathology), color='black', show.legend = F) +
+#     labs(y='Oesophageal Location',x='Months Prior to Endpoint', title='') +
+#     plot.theme + theme(axis.text.x=element_text(angle=45, hjust=1),legend.position='none') + 
+#     #scale_x_continuous(breaks=scale) +
+#     theme(panel.background = element_rect(fill='grey60', color='grey60'), panel.grid.major = element_line(color='grey70'), panel.grid.minor = element_blank())
+# 
+# ggplot(subset(back, Status == 'NP'), aes(factor(Endoscopy.Year, ordered=T), nl)) + geom_tile(aes(fill=Risk, color=p53.Status), size=1.5)  + facet_wrap(~Patient, scales='free') + 
+#   scale_fill_manual(values=riskCols, limits=levels(back$Risk),name='Progression') + 
+#   scale_color_manual(values=c('grey22','white'), limits=levels(back$p53.Status), drop=F, name='p53 IHC', labels=c('normal','aberrant')) +
+#     geom_text(aes(label=Pathology), color='black', show.legend = F) +
+#     labs(y='Oesophageal Location',x='Months Prior to Endpoint', title='') +
+#     plot.theme + theme(axis.text.x=element_text(angle=45, hjust=1),legend.position='none') + 
+#     #scale_x_continuous(breaks=scale) +
+#     theme(panel.background = element_rect(fill='grey60', color='grey60'), panel.grid.major = element_line(color='grey70'), panel.grid.minor = element_blank())
+#
+
+## P
+vig.plot(df=subset(back, Patient == 34),info= subset(patient.info, Patient == 34),cols=riskCols)
+vig.plot(subset(back, Patient == 56), subset(patient.info, Patient == 56),cols=riskCols)
+vig.plot(subset(back, Patient == 20), subset(patient.info, Patient == 20),cols=riskCols)
+vig.plot(subset(back, Patient == 86), subset(patient.info, Patient == 86),cols=riskCols)
+
+## NP
+vig.plot(subset(back, Patient == 3), subset(patient.info, Patient == 3),cols=riskCols)
+vig.plot(df=subset(back, Patient == 23),info= subset(patient.info, Patient == 23),cols=riskCols)
+vig.plot(subset(back, Patient == 33), subset(patient.info, Patient == 33),cols=riskCols)
+vig.plot(subset(back, Patient == 49), subset(patient.info, Patient == 49),cols=riskCols)
+vig.plot(subset(back, Patient == 51), subset(patient.info, Patient == 51),cols=riskCols)
+
+vig.plot(subset(back, Patient == 6), subset(patient.info, Patient == 6),cols=riskCols)
+vig.plot(subset(back, Patient == 54), subset(patient.info, Patient == 54),cols=riskCols)
+```
+
+### Recommendations
+
+ 1.	Immediate RFA: HGD diagnosis, two or more consecutive high risk predictions
+ 2.	Recheck 6-12 months: One high risk prediction or aberrant p53 IHC
+ 3. 	Recheck endoscopy 12-24 months: One or more moderate risk predictions
+ 4. 	Regular surveillance 3-5 years: Two or more consecutive low risk predictions
+
+ ```{r rx, echo=F, warning=F, message=T}
+
+rx<-function(Patient, brks, Prediction, Pathology, Risk, p53.Status) {
+  rules = as.data.frame(matrix(ncol=0, nrow=0))
+  # Consecutive
+  for (i in 1:length(brks)) {
+    risks = table(Risk[i:(i+1)])
+    p53 = table(p53.Status[i:(i+1)])
+    rule = 'None'
+    if ( risks['High'] == 2 | length(which(grepl('HGD|IMC', Pathology[i:(i+1)]))) > 0 ) {
+      rule = 1
+    } else if ( risks['High'] == 1 | p53['1'] > 0 ) {
+      rule = 2
+    } else if ( risks['Moderate'] > 0 ) {
+      rule = 3
+    } else if ( risks['Low'] == 2) {
+      rule = 4
+    }
+  rules = rbind(rules, cbind( as.character(brks[i]), as.character(brks[(i+1)]), rule))
+  }  
+  colnames(rules) = c('t1','t2', 'rule')
+  rules$Patient = Patient
+  rules
+}
+
+cols = unique(colnames(back), colnames(back.v))
+cols = cols[-grep('quants|pred\\.ratio|se\\(Risk\\)', cols)]
+allB = rbind.data.frame(back[,cols], back.v[,cols])
+allB$p53.Status = factor(allB$p53.Status, levels=c(0,1), ordered = T)
+
+maxP = allB %>% group_by(Patient, Status, brks) %>% summarise( 'maxP'=max(Prediction), 'Risk'=Risk[Prediction == max(Prediction)], 'Pathology'=max(Pathology), 'p53'=max(p53.Status,na.rm=T) )
+
+rules = do.call(rbind, lapply(unique(maxP$Patient), function(pt) {
+  r = with(arrange(subset(maxP, Patient == pt), brks), rx(pt, brks,maxP,Pathology,Risk,p53))
+  r$Status = unique(subset(maxP, Patient == pt)$Status)
+  r
+}))
+
+rules$t1 = factor(rules$t1, levels=levels(back$brks), ordered=T)
+rules$t2 = factor(rules$t2, levels=levels(back$brks), ordered=T)
+
+penult = subset(rules, t2 != 'Endpoint' & (t1 == rev(levels(back$brks))[2] | t2 == rev(levels(back$brks))[2] ))
+round(table(subset(penult, Status == 'P')$rule)/nrow(subset(penult, Status == 'P')), 2)
+round(table(subset(penult, Status == 'NP')$rule)/nrow(subset(penult, Status == 'NP')), 2)
+
+second = subset(rules,  t2 != 'Endpoint') %>% 
+  group_by(Patient, Status) %>% 
+  summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )
+table(subset(second, Status == 'P')$rule)
+table(subset(second, Status == 'NP')$rule)
+
+```
+
+## Incidence
+
+In the literature the risk of BE progression to HGD/cancer is estimated to be 0.2-0.7% per year.
+
+So for every 100 BE patients, at most `r 0.07*100` patients will progress in a year.
+
+Lifetime risk is assumed to be ~10% presuming 10 years followup (`r 0.7*10`%).
+Two ways to do this, take one year and look at the progression predictions...or take all of them and compare to the lifetime risk.
+
+```{r echo=F}
+file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
+load(file, verbose=T)
+rm(plots, coefs, cv.patient)
+
+fit = models[[select.alpha]]
+lambda = performance.at.1se[[select.alpha]]$lambda
+
+# number of patients we'd needed to have included to get 45 progressors from a representative population (I think)
+all = rbind(sum.patient.data, sum.validation.info)
+
+# Per year
+# This isn't quite right...we'd want to do something more like a 5 year 
+pop = round((table(all$Status)['P']/(0.07*100))*100)
+
+```
+
+### Lifetime adjusted risk
+
+```{r adjProbRisk, echo=F, fig.height=6, fig.width=10}
+#popL = round((table(all$Status)['P']/(0.1*100))*100)
+cases = table(sum.patient.data$Status)['P']
+popL = round((table(sum.patient.data$Status)['P']/(0.1*100))*100)
+
+#binom.test(cases, popL, 0.1)
+
+# Offset for the linear predictor would then be -log(pop/44)
+# Effectively this is my relative risk for the population
+offset = log(cases/popL)
+
+# link function provides odds
+#linPred = predict(fit, newx=dysplasia.df, s=lambda, type='link')
+#quantile(linPred)
+
+#df = rbind(dysplasia.df, validation.df)
+#df = dysplasia.df
+
+#p0 = predict(fit, newx=df, s=lambda, type='response')
+#p0 = 1/(1+exp(-linPred)) # logit to get probabilities, same as type='response'
+#p1 = 1/(1+exp(-linPred+abs(offset))) # calculates probs from risk (offset)
+
+cols = c('Hospital.Research.ID','Patient','Status','PID', 'Samplename','Endoscopy.Year','Pathology','Prediction','OR')
+#loopreds = rbind(loopreds[cols], do.call(rbind, vpd)[cols])
+loopreds = do.call(rbind, pg.samp)[cols]
+
+loopreds$P0 = 1/(1+exp(-loopreds$OR))
+loopreds$P1 = 1/(1+exp(-loopreds$OR+abs(offset))) # Adjust OR for RR
+
+mean(loopreds$P0)
+mean(loopreds$P1)
+
+ql = cbind.data.frame('y'=quantile(loopreds$OR)[c(2,4)], 'labels'=paste(c(25,75), '%', sep=''))
+
+# From my nested-LOO predictions
+
+ggplot(loopreds, aes(P1)) + 
+    geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + 
+    #geom_vline(xintercept = mean(adjRR)) +
+    scale_fill_gradientn(colors = myPal,  name='') + 
+    plot.theme + theme(legend.position = 'none') + 
+    labs(y='n Samples', x='Probability',
+         title='Sample Predictions: Lifetime risk adjusted')
+
+ggplot(loopreds, aes(OR)) + 
+    geom_histogram(aes(fill=..x..), bins=10, show.legend = F) + 
+    scale_fill_gradientn(colors = myPal,  name='') + 
+    #scale_fill_distiller(palette = 'RdYlBu', name='P(P)') + 
+    plot.theme + theme(legend.position = 'none') + 
+    labs(y='n Samples', x='Adjusted Relative Risk',
+         title='Sample Predictions: Lifetime risk adjusted')
+
+# Probability that an individual will progress in a lifetime (20 years) given a rate of progression of 0.07 is about 77%, the probability for a specific individual is less
+1-(0.93^20)
+
+```
+
+```{r lifetimeFPR, echo=F, fig.height=6, fig.width=6}
+## FPR estimates based on lifetime risks
+# We could consider any NP patient with a minimum of 2 High risk calls?  or High/Moderate Risk?
+# More correctly it should be any patients with concurrent High risk, but that's been pretty stable so two high is good enough for now
+
+bd = bind_rows(back, back.v)
+table(bd$Status)
+
+risksL = subset(bd, Status == 'NP') %>% group_by(Patient) %>% summarise( 'n.samp'=length(Patient), 'high.risk'=length(which(Risk == 'High')), 'mod.risk'=length(which(Risk == 'Moderate')) )
+
+risksL = subset(risksL, high.risk > 0 | mod.risk > 0)
+
+high = length(subset(risksL, high.risk >= 2 )$Patient)
+#mod = length(subset(risksL, mod.risk >= 2 )$Patient)
+
+# How would I tune it so that the FPR is a lot closer to the real incidence ~1%
+# Current false positive rate is 1.5x higher than the highest incidence ~0.7
+fpr=high*10/44
+x = subset(bd, Status == 'NP' & Patient %in% subset(risksL, high.risk >= 2)$Patient & Risk == 'High')
+(length(unique(subset(x, Prediction > 0.8)$Patient))*10)/44
+
+# How would that affect TPR?
+#pROC::coords(roc, 0.6)
+#pROC::coords(roc, 0.8)
+
+## Per year risk
+## Evaluate over 5 years? How?
+pyr = c()
+for (yr in names(which(table(bd$Endoscopy.Year) > 10))) {
+  risks = subset(bd, Status == 'NP' & Endoscopy.Year == yr) %>% group_by(Patient) %>% summarise( 'n.samp'=length(Patient), 'high.risk'=length(which(Risk == 'High')), 'mod.risk'=length(which(Risk == 'Moderate')) )
+
+  risks = subset(risks, high.risk > 0 | mod.risk > 0)
+  high = length(subset(risks, high.risk >= 2 )$Patient)
+  mod = length(subset(risks, mod.risk >= 2 )$Patient)
+
+  print(paste(yr,high*15, sep=':'))
+  pyr = c(pyr, high*15)
+}
+
+summary(pyr)
+
+pop*.07
+pop*.02
+pop*.05
+
+```
+
 ## Clonal Heterogenity
@@
 ```{r clonality, echo=F, message=F, warning=F, eval=F}
-patient.info = subset(patient.info, Hospital.Research.ID != 'AH0222')
-
-multi_samp = subset(patient.info, PID %in% names(which(table(patient.info$PID) > 1)))
-pts = unique(multi_samp$Hospital.Research.ID)
-
-## Variance in in CN per endoscopy
-for (pt in pts) {
-
-  pid = names(which(table(pg.samp[[pt]]$PID) > 1))
-
-  for(i in 1:length(pid)) {  
-    samples = subset(pg.samp[[pt]], PID == pid[i])$Samplename
-    subset(pg.samp[[pt]], PID == pid[i])$Prediction
-
-    x = apply( dysplasia.df[ samples,  ],2, var)
-    summary(x)
-
-    range(x)
-    plot(x)
-  }
-
-}
+patient.info %>% group_by (Hospital.Research.ID, Patient, PID) %>% summarise(
+  length(PID)
+)
+
+arms = c(grep(with(chr.lengths, paste(paste(chrom, ':', '1-\\d+000', sep=''), collapse='|')), colnames(dysplasia.df)),
+         grep(with(chr.lengths, paste(paste(chrom, ':', '\\d+000-', chr.length, sep=''), collapse='|')), colnames(dysplasia.df)))
+
+x = do.call(rbind, lapply(unique(patient.info$PID), function(pid) {
+    df = as.matrix(dysplasia.df[subset(info, PID == pid)$Samplename, -grep('cx',colnames(dysplasia.df))])
+    if (ncol(df) > 1) {
+      cbind('arms'=sum(rowSums(df[,arms],na.rm=T))/nrow(df),
+            'seg'=sum(rowSums(df[,-arms],na.rm=T))/nrow(df))
+    } else {
+      cbind('arms'=sum(df[arms,],na.rm=T),
+            'seg'=sum(df[-arms,],na.rm=T))
+    }
+}))
+rownames(x) = unique(patient.info$PID)
+x = x[complete.cases(x),]
+
+hist(x[,'seg'])
+hist(x[,'arms'])
+
+p = pnorm(x[,'seg'], mean(x[,'seg']), sd(x[,'seg']))
+summary(p)
+
+pa = pnorm(x[,'arms'], mean(x[,'arms']), sd(x[,'arms']))
+summary(pa)
+
+x[,'arms'] = scale(x[,'arms'])
+x[,'seg'] = scale(x[,'seg'])
+
+plot(x[unique(subset(patient.info, Patient == 48)$PID), 'arms'])
+plot(x[unique(subset(patient.info, Patient == 48)$PID), 'seg'])
+
+plot(x[unique(subset(patient.info, Patient == 34)$PID), 'arms'])
+plot(x[unique(subset(patient.info, Patient == 34)$PID), 'seg'])
+
+pp = back %>% group_by(Patient, Hospital.Research.ID, PID) %>% summarise(
+  'max'=max(Prediction), 'avg'=mean(Prediction)
+)
+
+pp = merge(pp, x, by.x='PID', by.y='row.names')
+
+cor.test(pp$seg, pp$max)
 
@@
 # glm selected features
-cox2 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', 'Pathology',  ch$label)]) 
+cox2 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', 'Pathology',  grep('^\\d+:', ch$label, value=T))]) 
 ggsurv(survfit(cox2))
@@
       p = p + geom_point(alpha=0.5, aes(size=hazards)) + 
-        geom_point(aes(coef.x, coef.y, shape=gl, color=p.value<0.05, size=hazards), show.legend = F) +
+        geom_point(aes(coef.x, coef.y, color=p.value<0.05, size=hazards), show.legend = F) +
         #scale_shape_manual(values=c(17,18), na.value=21) +
@@
         #geom_text_repel( aes( label=label), show.legend=F) + 
-        geom_text_repel(data=subset(wt, p.value < 0.05 & !is.na(gl)), aes(coef.x,coef.y, label=label), show.legend=F) + 
+        geom_text_repel(data=subset(wt, p.value < 0.05), aes(coef.x,coef.y, label=label), show.legend=F) + 
         geom_text_repel(data=subset(wt, is.na(stability)), aes(coef.x, coef.y, label=label, color=p.value<0.05), show.legend=F, color='grey39') 
@@
 ## GLM regions only
-coxDF = cbind(coxDF[,1:3],pathology[,c('BE','ID')], coxDF[,ch$label])
+coxDF = cbind(coxDF[,1:3],pathology[,c('BE','ID')], coxDF[,grep('^\\d+:', ch$label, value=T)])
 coxDF$Pathology = NULL
@@
 wt$label = sub('\\.', '-', sub('\\.', ':', sub('chr', '', rownames(wt))))
-wt = merge(wt, ch[,c('label','hazards', 'coef','stability','gl')], by='label', all=T)
+wt = merge(wt, ch[,c('label','hazards', 'coef','stability')], by='label', all=T)
 
@@
 ```
+
+# TODO: Relative risk
+```{r, eval=F}
+pred = predict(f, newx=dysplasia.df, s=performance.at.1se$`0.9`$lambda)
+?predict
+?predict.glm
+quantile(pred)
+p0 = predict(f, newx=dysplasia.df, s=performance.at.1se$`0.9`$lambda, type='response')
+p0
+p1 = 1/(1+exp(-pred+2.66))
+plot(p0,p1)
+hist(p1)
+hist(p0)
+hist(p1)
+plot(p0,p1)
+hist(p1, n=20)
+mean(p1>.5)
+mean(p0>.5)
+hist(pred)
+
+```

### 2018-01-05 coxRFX-forSWGS.Rmd
...

---
title: "coxRFX for SWGS"
output: html_notebook
---

```{r, include=FALSE}

library(ggplot2)
library(ggrepel)
library(CoxHD)
library(plyr)

plot.waldtest<-function(wt) {
  p = ggplot(wt, aes(coef.x, coef.y, color=p.value<0.05))
  if (length(grep('hazards', colnames(wt))) > 0) {
      p = p + geom_point(alpha=0.5, aes(size=hazards)) + 
        geom_point(aes(coef.x, coef.y, shape=gl, color=p.value<0.05, size=hazards), show.legend = F) +
        #scale_shape_manual(values=c(17,18), na.value=21) +
        scale_size(range=c(2, 10), breaks=c(0, 5, 10, 15)) +
        #geom_text_repel( aes( label=label), show.legend=F) + 
        geom_text_repel(data=subset(wt, p.value < 0.05 & !is.na(gl)), aes(coef.x,coef.y, label=label), show.legend=F) + 
        geom_text_repel(data=subset(wt, is.na(stability)), aes(coef.x, coef.y, label=label, color=p.value<0.05), show.legend=F, color='grey39') 
  } else {
    p = p + geom_point(alpha=0.5) + 
      geom_text_repel(data=subset(wt, p.value < 0.05), aes(coef.x,coef.y, label=label), show.legend=F)  
  }
  return(p + labs(title='Wald test'))
}

```

```{r}
load('swgs.Rdata', verbose=T)

# limited to GLM coefs
df = df[,c('tstart','tstop','Diagnosis','Pathology', ch$label)]
df[1:4,1:8]
```

```{r}
gcols = colnames(df)

cols = grep('^\\d+', colnames(df))
colnames(df)[cols] = paste('chr', gsub(':|-','.', grep('^\\d+', colnames(df), value=T)), sep='')

pathology = do.call(cbind.data.frame, lapply(unique(df$Pathology), function(p) {
  cbind( as.integer(df$Pathology == p))
}))
colnames(pathology) = unique(df$Pathology)

# as an ordinal
df$Pathology = as.integer(factor(df$Pathology))

surv = with(df, Surv(tstart, tstop, Diagnosis))

groups = factor(c('clinical',rep('genomic', length(grep('\\d+', colnames(df))))))
rfx = CoxRFX(df[,-c(1:3)], surv, groups, verbose=F, nu=0.1)
summary(rfx)
```

```{r}
VarianceComponents(rfx)
```

```{r}
plot(rfx)
```

```{r}
wt = WaldTest(rfx)
survConcordance(surv ~ predict(rfx))
```

```{r}
## Compare to GLM selected features

wt$label = sub('\\.', '-', sub('\\.', ':', sub('chr', '', rownames(wt))))

wt = merge(wt, ch[,c('label','hazards','coef', 'stability','gl')], by='label', all=T)

wt[is.na(wt$hazards), c('hazards', 'coef.y')] = 0

plot.waldtest(wt) + labs(x='RFX coef', y='GLM coef')
```

```{r}
df2 = cbind(df, pathology[,c('BE','ID')])
df2$Pathology = NULL

surv = with(df, Surv(tstart, tstop, Diagnosis))

groups = factor(c(rep('genomic', length(grep('\\d+', colnames(df)))),rep('clinical', (ncol(pathology)-1) ) ) )

rfx2 = CoxRFX(df2[,-c(1:3)], surv, groups, verbose=F)
summary(rfx2)
```

```{r}
plot(rfx2)
```

```{r}
VarianceComponents(rfx2)
```

```{r}
survConcordance(surv ~ predict(rfx2))
```

```{r}
wt = WaldTest(rfx2)
wt$label = sub('\\.', '-', sub('\\.', ':', sub('chr', '', rownames(wt))))

wt = merge(wt, ch[,c('label','hazards','coef', 'stability','gl')], by='label', all=T)

head(wt)

wt[is.na(wt$hazards), c('hazards', 'coef.y')] = 0

plot.waldtest(wt) + labs(x='RFX coef', y='GLM coef')

```

### 2018-01-05 lib/cv-pt-glm.R

@@ -230,7 +230,7 @@ crossvalidate.by.patient<-function(x,y,lambda,pts,a=1,nfolds=10, splits=5, fit=N
       test = x[test.rows,]
       training = x[-test.rows,]
       # pre-spec lambda seq
-      fitCV = glmnet(training, y[-test.rows], lambda=lambda, alpha=a, family='binomial',  ...) 
+      fitCV = glmnet(training, y[-test.rows], lambda=lambda, alpha=a, family='binomial',  standardize=F) 
       # autoplot(fit) + theme(legend.position="none")
       
       # Confusion matrix: quantitiative, pos results + neg results / number of test rows
@@ -238,7 +238,7 @@ crossvalidate.by.patient<-function(x,y,lambda,pts,a=1,nfolds=10, splits=5, fit=N
       cv.class[i,] = apply(pred, 2, function(p) {
         (p%*%y[test.rows] + (1-p) %*% (1-y[test.rows]))/length(test.rows)
       })
-      
+
       # Binomial deviance, lifted from cv.lognet
       dev = binomial.deviance(predict(fitCV, test, type='response'), as.factor(y[test.rows]))
       deviance[i,] = apply(dev, 2, weighted.mean, w=rep(1, nrow(dev)), na.rm=T)
@@ -251,15 +251,15 @@ crossvalidate.by.patient<-function(x,y,lambda,pts,a=1,nfolds=10, splits=5, fit=N
   
   # Not really sure what to do with the binomial deviance now...
   lambdas = cbind.data.frame('lambda.at'=1:ncol(cv.pred),
-                        'mean'= colMeans(cv.pred), 
-                        'sme'= apply(cv.pred, 2, sd)/sqrt(nrow(cv.pred)), 
-                        'sd'= apply(cv.pred, 2, sd), 
-                        'lambda'= lambda,
-                        'log.lambda' = log(lambda),
-                        'mean.b.dev' = colMeans(cv.binomial.deviance),
-                        'sd.b.dev' = apply(cv.binomial.deviance, 2,sd),
-                        'sme.b.dev' = apply(cv.binomial.deviance, 2, sd)/sqrt(nrow(cv.binomial.deviance)),
-                        'lambda.min' = F, 'lambda+1se' = F, 'lambda-1se' = F) 
+                              'mean'= colMeans(cv.pred), 
[…]

### 2018-01-22 Paper.Rmd
...

@@
 
+## Quick look at our demographic information
+
 ```{r echo=F}
@@
 pander(
-  as.data.frame(all.info %>% group_by(Status) %>% summarise(
+  as.data.frame(all.info %>% summarise(
   'Total Patients'=length(unique(Patient)),
   'Total Samples'=sum(total.samples),
+  'Range'=paste(range(end.year-start.year), collapse='-'),
   'Median Years'=paste(median(end.year-start.year),'+/-',round(sd(end.year-start.year),1)),
@@
 ))
-, caption='Full cohort')
-
-```
-
-## Quick look at our demographic information
+, caption='full cohort')
+```
+
+```{r echo=F}
+cohortSummary = as.data.frame(all.info %>% group_by(Status) %>% summarise(
+  'Total Patients'=length(unique(Patient)),
+  'Total Samples'=sum(total.samples),
+  'Range follow-up (years)'=paste(range(end.year-start.year), collapse='-'),
+  'Mean follow-up (years)'=paste(round(mean(end.year-start.year), 1),'+/-',round(sd(end.year-start.year),1)),
+  'Range endoscopies'=paste(range(total.endos), collapse='-'),
+  'Mean Endoscopies'=paste(round(mean(total.endos), 1),'+/-',round(sd(total.endos),1)),
+  'Mean age'=paste(round(mean(age.diagnosis), 1),'+/-',round(sd(age.diagnosis),1)),
+  'gender F:M'=paste(table(gender),collapse=':'),
+  'Mean Length'=paste(round(mean(C,na.rm=T), 1),'+/-',round(sd(C,na.rm=T),1)),
+  'naC'=length(which(is.na(C)))
+))
+
+cohortSummary = cbind(cohortSummary,
+  rbind(table(subset(all.patient.info, Status == 'NP')$Pathology),
+        table(subset(all.patient.info, Status == 'P')$Pathology)))
+
+  rbind(table(subset(patient.info, Status == 'NP')$Pathology),
+        table(subset(patient.info, Status == 'P')$Pathology))
+
+  rbind(table(subset(validation.patient.info, Status == 'NP')$Pathology),
+        table(subset(validation.patient.info, Status == 'P')$Pathology))
+
+```
+
+`r pander(t(cohortSummary), caption='Full cohort, split by status')`
 
@@
 
-```{r echo=F, warning=F, message=F, eval=T, fig.width=12, fig.height=8}
+```{r echo=F, warning=F, message=F, eval=T, fig.height=12, fig.width=20}
 
@@
 
+globalCNdf$Status = factor(globalCNdf$Status, levels=sort(unique(globalCNdf$Status), decreasing=T), ordered=T)
 globalCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms'), labeller(Status=n.status, chr=label_value))
@@
 
+## Dysplasia only
+dys.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology %in% c('LGD','HGD','IMC'))$variable)), ')', sep=''),
+              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology %in% c('LGD','HGD','IMC'))$variable)),')', sep=''))
+
+globalnHGDCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology %in% c('LGD','HGD','IMC')), labeller(Status=dys.status, chr=label_value)) + labs(title='LGD/HGD/IMC Only')
+
 ## No HGD
-`%nin%` = Negate(`%in%`)
-
-nhgd.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology %nin% c('HGD','IMC'))$variable)), ')', sep=''),
-              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology %nin% c('HGD','IMC'))$variable)),')', sep=''))
-globalnHGDCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology %nin% c('HGD','IMC')), labeller(Status=nhgd.status, chr=label_value)) + labs(title='Excluding HGD/IMC')
+# `%nin%` = Negate(`%in%`)
+# 
+# nhgd.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology %nin% c('HGD','IMC'))$variable)), ')', sep=''),
+#               'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology %nin% c('HGD','IMC'))$variable)),')', sep=''))
+# globalnHGDCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology %nin% c('HGD','IMC')), labeller(Status=nhgd.status, chr=label_value)) + labs(title='Excluding HGD/IMC')
 
@@
 
-globalCN.plot 
-globalnHGDCN.plot
-globalBECN.plot
-
-dfpts = subset(globalCNdf, Patient %in% c('57','76') & seg.type != 'arms')
-#dfpts = dfpts %>% group_by(Patient, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
-
-pt.status <- c('Non-Progressor'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
-              'Progressor'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))
-genomePtCN.plot = cn.mtn.plot(dfpts, labeller(Status = pt.status))
-
-dfpts = subset(globalCNdf, Patient %in% c('57','76') & Pathology == 'BE' & seg.type != 'arms')
-#dfpts = dfpts %>% group_by(Patient, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
-
-ptbe.status <- c('Non-Progressor'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
-              'Progressor'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))
-genomePtBECN.plot = cn.mtn.plot(dfpts,  labeller(Status = ptbe.status )) + labs(title='NDBE samples only')
-
-genomePtCN.plot 
-genomePtBECN.plot
+globalCN.plot + theme(legend.position = 'none')
+
+globalnHGDCN.plot + theme(legend.position = 'none')
+
+globalBECN.plot + theme(legend.position = 'none')
+
+# dfpts = subset(globalCNdf, Patient %in% c('57','76') & seg.type != 'arms')
+# #dfpts = dfpts %>% group_by(Patient, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
+# 
+# pt.status <- c('Non-Progressor'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
+#               'Progressor'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))
+# genomePtCN.plot = cn.mtn.plot(dfpts, labeller(Status = pt.status))
+# 
+# dfpts = subset(globalCNdf, Patient %in% c('57','76') & Pathology == 'BE' & seg.type != 'arms')
+# #dfpts = dfpts %>% group_by(Patient, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
+# 
+# ptbe.status <- c('Non-Progressor'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
+#               'Progressor'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))
+# genomePtBECN.plot = cn.mtn.plot(dfpts,  labeller(Status = ptbe.status )) + labs(title='NDBE samples only')
+# 
+# genomePtCN.plot + theme(legend.position = 'none')
+# genomePtBECN.plot + theme(legend.position = 'none')
 
@@
 
-a = sample( subset(subset(patient.info, Patient == 57),  Endoscopy.Year == min(Endoscopy.Year) )$Samplename, 1)
-b = sample( subset(subset(patient.info, Patient == 76), Endoscopy.Year == min(Endoscopy.Year) & Pathology %nin% c('HGD','IMC'))$Samplename, 1)
-
-ind.status <- c('Non-Progressor'='Patient 57 (n=1)',
-              'Progressor'='Patient 76 (n=1), excl. HGD/IMC')
-cn.mtn.plot(subset(globalCNdf, variable %in% c(a,b) & seg.type != 'arms'), labeller(Status = ind.status), type='bar') + labs(title="Individual samples")
+#a = sample( subset(subset(patient.info, Patient == 57),  Endoscopy.Year == min(Endoscopy.Year) )$Samplename, 1)
+#b = sample( subset(subset(patient.info, Patient == 76), Endoscopy.Year == min(Endoscopy.Year) & Pathology %nin% c('HGD','IMC'))$Samplename, 1)
+
+#ind.status <- c('Non-Progressor'='Patient 57 (n=1)','Progressor'='Patient 76 (n=1), excl. HGD/IMC')
+#cn.mtn.plot(subset(globalCNdf, variable %in% c(a,b) & seg.type != 'arms'), labeller(Status = ind.status), type='bar') + labs(title="Individual samples")
+```
+
+```{r loaddata, message=F, warning=F, echo=F}
+
+file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
+if (file.exists(file)) {
+  message(paste("loading", file))
+  load(file, verbose=T)
+} else {
+  stop(paste("No model file", file, "found"))
+}
+
+# cx was generated on raw seg values
+plot(dysplasia.df[names(labels) ,'cx'], col=labels+1)
+
+# remove old cx...see what new one would look like
+dysplasia.df = dysplasia.df[,-grep('cx',colnames(dysplasia.df))]
+
+# general variance per sample after the samples have been mean normalized, ignores segment length
+cx = apply(dysplasia.df, 1, function(x) var(x))
+
+x = cbind.data.frame('Status'=labels, 'cx'=cx[names(labels)])
+x$Status = factor(x$Status, labels = c('Non-Progressor','Progressor'))
+x = merge(x,patient.info[,c('Samplename','Pathology','Patient')],by.x='row.names',by.y='Samplename')
+
+tt = t.test(subset(x, Status == 'Progressor')$cx, subset(x, Status == 'Non-Progressor')$cx)
+ggplot(x, aes(Status, cx, fill=Status)) + 
+  geom_jitter(width=0.2, color='grey39') + 
+  geom_violin(outlier.colour = NA, alpha=0.8) +
+  labs(x='', y='Variance', title='Per sample variance', subtitle=paste('p-value=', format.pval(tt$p.value, digits=2))) + theme(legend.position = 'none') +
+  plot.theme
+
+rocX = pROC::roc(Status ~ cx, data=x, auc=T, ci=T, of='thresholds')
+roc.plot(rocX, title='Per sample variance') + plot.theme
+
 ```
@@
 
-ch$ymax = ifelse(ch$coef > 0, max(genomePtCN.plot$data$value), min(genomePtCN.plot$data$value))
-chh = chh %>% rowwise() %>% mutate('max'=custom.max(coef, hazards, ymax))
-genomePtCN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)
-
-genomePtBECN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)
+# ch$ymax = ifelse(ch$coef > 0, max(genomePtCN.plot$data$value), min(genomePtCN.plot$data$value))
+# chh = chh %>% rowwise() %>% mutate('max'=custom.max(coef, hazards, ymax))
+# genomePtCN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)
+
+# genomePtBECN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)
 
@@
 
-```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=5, fig.width=5}
-preds = unlist(lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], function(df) df$Prediction))
-np.preds = unlist(lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID], function(df) df$Prediction))
-
-roc50 = pROC::roc(as.integer(c(preds, np.preds) > 0.5), c(rep(1, length(preds)), rep(0, length(np.preds))), main='Progressors & Non-progressors, cutoff 0.5')
-roc50$model = 'CN Model 0.5'
-
+```{r insilico_rocs, echo=F, message=F, warning=F, fig.height=8, fig.width=8}
+
+load(file=paste(cache.dir, 'insilico_splits.Rdata', sep='/'), verbose=T)
+
+# m = melt(insilico.rocs, measure.vars = 'auc')
+# ggplot(m, aes(variable,value)) + ylim(0,1) +
+#   geom_boxplot(fill='mintcream', color='grey55') + geom_jitter(width=0.2, color='darkblue', size=1.5, show.legend = F) + 
+#   geom_text(data=subset(m, value %in% range(value)), aes(label=round(value,2)), vjust=0.02, hjust=0.2) +
+#   plot.theme + labs(x='',y='AUC',title='AUC', subtitle='20 insilico 20:69 splits')
+
+```
+
+```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=7, fig.width=7}
 preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
@@
 roc$model = 'CN Model'
+
+multi.roc.plot(list(roc), colors=brewer.pal(3,'Greys')[3]) + labs(title='Discovery cohort ROC')
+
+insilico.rocs = rbind(insilico.rocs, c(pROC::ci.auc(roc), pROC::coords(roc, 'best')))
+
+insilico.rocs[1:20,'split'] = paste('in-silico', 1:20)
+insilico.rocs[21,'split'] = 'discovery'
+
+ggplot(insilico.rocs, aes(split, auc)) + geom_point() + ylim(0,1) +
+  geom_errorbar(aes(ymin=auc.ci.min, ymax=auc.ci.max), width=0.2) + 
+  coord_flip() + plot.theme + labs(y='AUC', x='', title='AUC per split', subtitle='20 insilico 20:69 splits')
 
@@
 
-ggplot() + geom_line(data=df, aes(specificity, sensitivity)) + scale_x_reverse() + 
+ggplot() + geom_line(data=df, aes(specificity, sensitivity), color='darkblue', size=1.5) + scale_x_reverse() + 
   geom_point(data=thresholds, aes(x=specificity,y=sensitivity,color=label), size=3,shape=8, show.legend = T) +
@@
 
-roc.plot(roc, 'Probability threshold calculated by ROC')
-
 roc.plot(rocNOHGD, 'Model excluding HGD/IMC')
@@
 ### Compare ROCs
-
-```{r echo=F, fig.height=6, fig.width=8}
-roc$model = "Per Sample"
-multi.roc.plot(list(roc, rocEndoM, rocEndoA, rocPtM, rocPtA,rocFirstEndo,rocLastEndo), 'ROC Comparisons')
-```
 
@@
 
-get.auc.at(p53roc)
-get.auc.at(hgdpathroc)
-get.auc.at(pathroc)
-get.auc.at(roc)
+# get.auc.at(p53roc)
+# get.auc.at(hgdpathroc)
+# get.auc.at(pathroc)
+# get.auc.at(roc)
 
@@
 
-rocList = list(roc,roc50, pathroc,hgdpathroc,p53roc)
-
+rocList = list(roc,pathroc,hgdpathroc,p53roc)
+
+cols = brewer.pal(11, "PRGn")
 dodge = position_dodge(width=0.9)
@@
   geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
-  geom_text(aes(label=paste(round(value,2)*100,'%',sep=''), y=0, vjust=-1 ), position=dodge, color='white')  +
-  scale_fill_manual(values=c('purple3','dodgerblue3','firebrick3'), name='') + plot.theme + 
-  theme(legend.position = 'none') + labs(x='', y='Percent', title='ROC Comparison') 
-```
-
-***
-
-__All predictions from here onwards are performed using the `r cutoff` cutoff__
-
-***
+  geom_text(aes(label=paste(round(value,2)*100,'%',sep=''), y=1, vjust=-1 ), position=dodge, color='white')  +
+  scale_fill_manual(values=cols, name='') + plot.theme + 
+  theme(legend.position = 'none', axis.text.x = element_text(angle=45, hjust=1)) + labs(x='', y='AUC', title='ROC Comparison')
+
+```
+
+```{r echo=F, fig.height=8, fig.width=6}
+roc$model = "Per Sample"
+
+m = rbind.data.frame(
+      get.roc.stat(roc),
+      get.roc.stat(rocEndoM),get.roc.stat(rocEndoA),
+      get.roc.stat(rocPtM), get.roc.stat(rocPtA),
+      get.roc.stat(rocFirstEndo), get.roc.stat(rocLastEndo))
+m$model = factor(m$model)
+
+cols = brewer.pal(11, 'Paired')
+dodge = position_dodge(width=0.9)
+mm = melt(subset(m, Measure == 'AUC'), measure.vars='value')
+ggplot( arrange(mm, CI.min), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=model)) + 
+  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
+  geom_text(aes(label=paste(round(value,2)*100,'%',sep=''), y=1), position=dodge, color='white')  +
+  scale_fill_manual(values=cols, name='') + plot.theme + 
+  theme(legend.position = 'none', axis.text.x = element_text(angle=45, hjust=1)) + labs(x='', y='AUC', title='ROC Comparison')
+
+multi.roc.plot(list(roc, rocEndoM, rocEndoA, rocPtM, rocPtA,rocFirstEndo,rocLastEndo), 'ROC Comparisons', palette='Dark2')
+```
 
@@
 
-```{r, echo=F, warning=F, message=F, fig.height=12, fig.width=16}
+```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=20}
+
+min.theme = theme(text=element_text(size=10), panel.background=element_blank(), strip.background=element_rect(fill="grey97"), 
+                   strip.text = element_text(size=10), axis.text = element_blank(),
+                   axis.line=element_line(color='black'), panel.grid.major.y =element_line(color='grey95'), panel.grid.major.x = element_blank(),
+                   panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0, 'lines'), panel.spacing.y = unit(0.1, 'lines')   )
 
@@
   if (is.null(pal))
-    pal = RColorBrewer::brewer.pal(length(limits), 'Paired')
-
-  df = merge(df, b[,c('Samplename','Endoscopy.Year','months.before.final','Block','nl')], by.x='variable',by.y='Samplename')
-  df$nl = factor(df$nl, ordered = T)
-
+    pal = RColorBrewer::brewer.pal(length(limits), 'Set2')
+
+  df = merge(df, b[,c('Samplename','Endoscopy.Year','months.before.final','Block','ogj')], by.x='variable',by.y='Samplename')
+  df$ogj = factor(df$ogj, ordered = T)
+  df$ogj = factor(df$ogj, levels=rev(levels(df$ogj)), ordered = T)
   #pal = pal[which(limits %in% levels(df$nl))]
 
-  ggplot(df, aes(x=chr.length)) + facet_grid(Endoscopy.Year~chr, scales='free_x', space='free_x') +
-    geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=nl), show.legend=T) + 
+  altcolor = c(RColorBrewer::brewer.pal(9, 'Greys')[2],RColorBrewer::brewer.pal(9, 'Greys')[4])
+  fills = c()
+  for (yr in sort(unique(df$Endoscopy.Year))) {
+    ac = (which(sort(unique(df$Endoscopy.Year)) == yr) %% 2) + 1
+    fills = c(fills, rep(altcolor[ac], length(which(table(subset(df, Endoscopy.Year == yr)$ogj) > 0))))
+  }
+
+  p = ggplot(df, aes(x=chr.length)) + facet_grid(Endoscopy.Year+ogj~chr, scales='free', space='free_x', labeller = labeller(.multi_line=F)) +
+      geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=ogj), show.legend=T) + 
       scale_fill_manual(values=pal, limits=limits,  name='OGJ Dist.') +
-    theme(axis.text.x = element_blank()) +
-    plot.theme + labs(title=title)
-}
-
-ogjLimits = levels(factor(back$nl, ordered = T))
-
-p1 = mtn.spatial(subset(globalCNdf, Patient == 48 & seg.type != 'arms'), back, ogjLimits, 'Patient 48, (P)')
-ogjLegend = get.legend(p1)
-p1+theme(legend.position = 'none')
-
-mtn.spatial(subset(globalCNdf, Patient == 60 & seg.type != 'arms'), back,ogjLimits,'Patient 60 (P)' )+theme(legend.position = 'none')
-
-mtn.spatial(subset(globalCNdf, Patient == 3 & seg.type != 'arms'), back,ogjLimits,'Patient 3 (NP)' )+theme(legend.position = 'none')
-
-mtn.spatial(subset(globalCNdf, Patient == 15 & seg.type != 'arms'), back,ogjLimits,'Patient 15 (NP)' )+theme(legend.position = 'none')
+      labs(title=title, x='') + min.theme
+
+  legend = get.legend(p)
+  p = p + theme(legend.position = 'none')
+
+  g = ggplot_gtable(ggplot_build(p))
+  strips = grep('strip-r', g$layout$name)
+
+  for (n in 1:length(strips)) {
+    i = strips[n]
+    j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
+    g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[n]
+  }
+
+  grid.arrange(g)
+  return(list('grob'=g, 'legend'=legend))
+  #grid::grid.draw(g)  
+  #return(g)
+}
+
+back$ogj = back$nl
+back$ogj[which(back$nl > 0 & back$nl < 3)] = 1 
+back$ogj[which(back$nl > 3 & back$nl < 7)] = 2 
+back$ogj[which(back$nl >= 7)] = 3
+
+ogjLimits = levels(factor(back$ogj, ordered = T))
+
+p1 = mtn.spatial(subset(globalCNdf, Patient == 48 & seg.type != 'arms'), back, ogjLimits, 'Patient 48 (P)')
+p2 = mtn.spatial(subset(globalCNdf, Patient == 20 & seg.type != 'arms'), back, ogjLimits,'Patient 20 (P)' )
+p3 = mtn.spatial(subset(globalCNdf, Patient == 60 & seg.type != 'arms'), back, ogjLimits,'Patient 60 (P)' )
+
+p4 = mtn.spatial(subset(globalCNdf, Patient == 3 & seg.type != 'arms'), back,ogjLimits,'Patient 3 (NP)' )
+p5 = mtn.spatial(subset(globalCNdf, Patient == 15 & seg.type != 'arms'), back,ogjLimits,'Patient 15 (NP)' )
+p6 = mtn.spatial(subset(globalCNdf, Patient == 51 & seg.type != 'arms'), back,ogjLimits,'Patient 51 (NP)' )
+
+back$ogj = NULL
+
+```
+
+```{r, echo=F, fig.width=2, fig.height=4}
+
+grid.arrange(p1$legend)
 
@@
 
-```{r, echo=F, warning=F, message=F, fig.height=8, fig.width=9}
+```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
 riskPath = rbind( cbind.data.frame( table(subset(back, Status == 'P', select=c('Pathology','Risk'))), 'Status'='Progressor' ), 
@@
 riskPath = merge(riskPath, pathTotal)
-riskPath$ratio = with(riskPath, round(Freq/nPath, 2))
+riskPath$ratio = with(riskPath, round(Freq/nPath, 3))
 riskPath$Risk = factor(riskPath$Risk, levels=c('Low','Moderate','High'), ordered = T)
@@
 p53Path = merge( subset(back, Status == 'P' & !is.na(p53.Status)) %>% group_by(Pathology) %>% summarise( 'nPath'=length(Pathology)), p53Path)
-p53Path$Risk = with(p53Path, ifelse(p53.Status == 1, 'High','Low'))
+#p53Path$Risk = with(p53Path, ifelse(p53.Status == 1, 'High','Low'))
 p53Path$ratio = with(p53Path, round(Freq/nPath, 2))
@@
 ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
-  scale_fill_manual(values=riskCols) + 
+  scale_fill_manual(values=riskCols) + scale_color_manual(values=c('white','black','white')) +
   geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), position=position_stack(), size=4) +
-  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5) +
+  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep=''), color=Risk), position=position_stack(vjust = 0.5), size=5, show.legend = F) +
   facet_grid(~Status) + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
@@
 
-```{r p53IHC, echo=F, warning=F, message=F, fig.height=6, fig.width=8}
-
-p53Path$Risk[p53Path$Risk == 'High'] = 'Aberrant'
-p53Path$Risk[p53Path$Risk == 'Low'] = 'Normal'
-
-col = brewer.pal(11, "PRGn")[c(3,9)]
-ggplot(p53Path, aes(Pathology, ratio*100, group=Pathology, fill=Risk)) + geom_bar(stat='identity', position=position_stack()) + 
-  scale_fill_manual(values=col) + 
+```{r p53IHC, echo=F, warning=F, message=F, fig.height=8, fig.width=8}
+
+#p53Path$Risk[p53Path$Risk == 'High'] = 'Aberrant'
+#p53Path$Risk[p53Path$Risk == 'Low'] = 'Normal'
+
+#p53Path$p53.Status = factor(p53Path$p53.Status, levels=c('Aberrant', 'Normal'), ordered = F)
+
+p53Path$p53.Status = factor(p53Path$p53.Status, ordered = T)
+p53Path = arrange(p53Path, Pathology, p53.Status)
+col = brewer.pal(11, "PRGn")[c(9,3)]
+ggplot(p53Path, aes(Pathology, ratio*100, group=Pathology, fill=p53.Status)) + geom_bar(stat='identity', position = position_stack()) +
+  scale_fill_manual(values=col, name='p53 IHC', labels=c('Normal','Aberrant')) + 
   geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101)) +
@@
 
-```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=15 }
+```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=5 }
 back$nl = factor(back$nl, ordered=T)
@@
 
-pred.tiles(subset(back, Patient %in% c(6)), ncol=1, cols=myPal, probs=T) + labs(title='') + theme(legend.position = 'none')
+p = pred.tiles(subset(back, Patient %in% c(6)), ncol=1, cols=myPal, probs=T) + labs(title='')
+tilesLegend = get.legend(p)
+p + theme(legend.position = 'none')
+
+pred.tiles(subset(back, Patient %in% c(6)), ncol=1, cols=myPal, probs=T, OR=F) + labs(title='') + theme(legend.position = 'none')
 pred.tiles(subset(back, Patient %in% c(48)), ncol=1, cols=myPal, probs=T) + labs(title='') + theme(legend.position = 'none')
@@
 pred.tiles(subset(back, Patient %in% c(89)), ncol=1, cols=myPal, probs=T) + labs(title='')+ theme(legend.position = 'none')
-
+```
+
+```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=15 }
 grid.arrange(
-  pred.tiles(subset(back, Patient %in% c(13,20,34,56,60,48) & Status == 'P'), ncol=1, cols=myPal, probs=T) + labs(title='Progressors'),
-  pred.tiles(subset(back, Patient %in% c(15,16,21,3,45,19) & Status == 'NP'), ncol=1, cols=myPal, probs=T) + labs(title='Non-Progressors'),
+  pred.tiles(subset(back, Patient %in% c(13,20,34,56,60,48) & Status == 'P'), ncol=1, cols=myPal, probs=T) + labs(title='Progressors') + theme(legend.position = 'none'),
+  pred.tiles(subset(back, Patient %in% c(15,16,21,3,45,19) & Status == 'NP'), ncol=1, cols=myPal, probs=T) + labs(title='Non-Progressors') + theme(legend.position = 'none'),
 ncol=2)
@@
 df = rbind(cbind(valCNdf, 'set'=rep('Val', nrow(valCNdf))), cbind(globalCNdf, 'set'=rep('Train', nrow(globalCNdf)) ) )
-subset(df, seg.type != 'arms') %>% group_by(set,Status) %>% summarise( var(mean.value), mean(mean.value), max(mean.value), min(mean.value) )
 
@@
 
-head(back.v)
-
 m = melt(back.v %>% group_by(Patient, Status) %>% summarise( 'Low'=length(which(Risk == 'Low'))/length(Patient),
@@
   geom_text(aes(label=n.samples, y=101)) + scale_fill_manual(values=riskCols) +
-  facet_grid(~Status, scales='free_x') + labs(y='Percent of samples', title='Validation patient set')
-```
-
-```{r echo=F, warning=F, message=F, fig.height=14,fig.width=28}
-grid.arrange(
-  pred.tiles(subset(back.v, Status == 'P'), cols=myPal, probs=T) + labs(title='Validation Progressors') + theme(legend.position = 'none'),
-  pred.tiles(subset(back.v, Status == 'NP'), cols=myPal, probs=T) + labs(title='Validation Non-Progressors') + theme(legend.position = 'none'),
-ncol=2)
+  facet_grid(~Status, scales='free_x') + labs(y='Percent of samples', title='Validation patient set') + plot.theme
+```
+
+```{r echo=F, warning=F, message=F, fig.height=5,fig.width=25}
+pred.tiles(subset(back.v, Status == 'P'), cols=myPal, probs=T, nrow=1) + labs(title='Validation Progressors') + theme(legend.position = 'none')
+pred.tiles(subset(back.v, Status == 'NP'), cols=myPal, probs=T, nrow=1) + labs(title='Validation Non-Progressors') + theme(legend.position = 'none')
 ```
@@
 riskPath = merge(riskPath, pathTotal)
-riskPath$ratio = with(riskPath, round(Freq/nPath, 2))
+riskPath$ratio = with(riskPath, round(Freq/nPath, 3))
 riskPath$Risk = factor(riskPath$Risk, levels=c('Low','Moderate','High'), ordered = T)
 
-p53Path = cbind.data.frame(table(subset(back.v, Status == 'P', select=c('Pathology','p53.Status'))), 'Status'='p53 IHC')
-p53Path = merge( subset(back.v, Status == 'P' & !is.na(p53.Status)) %>% group_by(Pathology) %>% summarise( 'nPath'=length(Pathology)), p53Path)
-p53Path$Risk = with(p53Path, ifelse(p53.Status == 1, 'High','Low'))
-p53Path$ratio = with(p53Path, round(Freq/nPath, 2))
-
-riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor','p53 IHC'), ordered = T)
+hgdimc = subset(riskPath, Pathology %in% c('HGD','IMC')) %>% group_by(Risk) %>% summarize (
+  Freq=sum(Freq),  nPath=sum(nPath), ratio=sum(Freq)/sum(nPath)
+)
+riskPath = riskPath[-which(riskPath$Pathology %in% c('HGD','IMC')),] 
+riskPath = rbind(riskPath , cbind('Pathology'='HGD/IMC', 'Status'='Progressor', hgdimc))
+
+riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor'), ordered = T)
 
@@
 
-ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
-  scale_fill_manual(values=riskCols) + 
-  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), position=position_stack(), size=4) +
-  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5) +
+ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) +   scale_fill_manual(values=riskCols) + scale_color_manual(values=c('white','black','white')) +
+  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), position=position_stack(), size=4, show.legend = F) +
+  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep=''), color=Risk), position=position_stack(vjust = 0.5), size=5, show.legend = F) +
   facet_grid(~Status) + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
@@
 roc$model = 'Discovery'
-multi.roc.plot(list(roc, vroc), title='ROC')
-
+multi.roc.plot(list(roc, vroc), title='ROC', palette='Dark2') 
+```
+
+```{r, echo=F, warning=F, message=F, fig.width=4, fig.height=4}
+aucs = rbind.data.frame(pROC::ci.auc(roc), pROC::ci.auc(vroc))
+colnames(aucs) = c('ci.min', 'auc','ci.max')
+aucs$model = c('Discovery','Validation')
+
+ggplot(aucs, aes(model, auc)) + geom_point(size=2) + ylim(0,1) +
+  geom_errorbar(aes(ymin=ci.min, ymax=ci.max), width=0.1) + coord_flip() + 
+  geom_text(aes(y=ci.min, label=round(ci.min, 2)), vjust=4) +
+  geom_text(aes(y=ci.max, label=round(ci.max, 2)), vjust=-4) +
+#  geom_text(aes(y=auc, label=round(auc, 2)), vjust=-1) +
+  labs(x='', y='AUC', title='AUC with CI') + plot.theme 
 ```
@@
 
-#back$Risk = factor(back$Risk, levels=levels(back$labels), ordered = T)
-
-# ggplot(subset(back, Status == 'P'), aes(factor(Endoscopy.Year, ordered=T), nl)) + geom_tile(aes(fill=Risk, color=p53.Status), size=1.5)  + facet_wrap(~Patient, scales='free') + 
-#   scale_fill_manual(values=riskCols, limits=levels(back$Risk),name='Progression') + 
-#   scale_color_manual(values=c('grey22','white'), limits=levels(back$p53.Status), drop=F, name='p53 IHC', labels=c('normal','aberrant')) +
-#     geom_text(aes(label=Pathology), color='black', show.legend = F) +
-#     labs(y='Oesophageal Location',x='Months Prior to Endpoint', title='') +
-#     plot.theme + theme(axis.text.x=element_text(angle=45, hjust=1),legend.position='none') + 
-#     #scale_x_continuous(breaks=scale) +
-#     theme(panel.background = element_rect(fill='grey60', color='grey60'), panel.grid.major = element_line(color='grey70'), panel.grid.minor = element_blank())
-# 
-# ggplot(subset(back, Status == 'NP'), aes(factor(Endoscopy.Year, ordered=T), nl)) + geom_tile(aes(fill=Risk, color=p53.Status), size=1.5)  + facet_wrap(~Patient, scales='free') + 
-#   scale_fill_manual(values=riskCols, limits=levels(back$Risk),name='Progression') + 
-#   scale_color_manual(values=c('grey22','white'), limits=levels(back$p53.Status), drop=F, name='p53 IHC', labels=c('normal','aberrant')) +
-#     geom_text(aes(label=Pathology), color='black', show.legend = F) +
-#     labs(y='Oesophageal Location',x='Months Prior to Endpoint', title='') +
-#     plot.theme + theme(axis.text.x=element_text(angle=45, hjust=1),legend.position='none') + 
-#     #scale_x_continuous(breaks=scale) +
-#     theme(panel.background = element_rect(fill='grey60', color='grey60'), panel.grid.major = element_line(color='grey70'), panel.grid.minor = element_blank())
-#
-
 ## P
-vig.plot(df=subset(back, Patient == 34),info= subset(patient.info, Patient == 34),cols=riskCols)
-vig.plot(subset(back, Patient == 56), subset(patient.info, Patient == 56),cols=riskCols)
-vig.plot(subset(back, Patient == 20), subset(patient.info, Patient == 20),cols=riskCols)
-vig.plot(subset(back, Patient == 86), subset(patient.info, Patient == 86),cols=riskCols)
+p = vig.plot(df=subset(back, Patient == 34),info= subset(patient.info, Patient == 34),cols=riskCols) + theme(legend.position = 'bottom')
+vigLeg = get.legend(p)
+p + theme(legend.position = 'none')
+vig.plot(subset(back, Patient == 56), subset(patient.info, Patient == 56),cols=riskCols) + theme(legend.position = 'none')
+vig.plot(subset(back, Patient == 20), subset(patient.info, Patient == 20),cols=riskCols) + theme(legend.position = 'none')
+vig.plot(subset(back, Patient == 86), subset(patient.info, Patient == 86),cols=riskCols) + theme(legend.position = 'none')
 
 ## NP
-vig.plot(subset(back, Patient == 3), subset(patient.info, Patient == 3),cols=riskCols)
-vig.plot(df=subset(back, Patient == 23),info= subset(patient.info, Patient == 23),cols=riskCols)
-vig.plot(subset(back, Patient == 33), subset(patient.info, Patient == 33),cols=riskCols)
-vig.plot(subset(back, Patient == 49), subset(patient.info, Patient == 49),cols=riskCols)
-vig.plot(subset(back, Patient == 51), subset(patient.info, Patient == 51),cols=riskCols)
-
-vig.plot(subset(back, Patient == 6), subset(patient.info, Patient == 6),cols=riskCols)
-vig.plot(subset(back, Patient == 54), subset(patient.info, Patient == 54),cols=riskCols)
+vig.plot(subset(back, Patient == 3), subset(patient.info, Patient == 3),cols=riskCols) + theme(legend.position = 'none')
+vig.plot(df=subset(back, Patient == 23),info= subset(patient.info, Patient == 23),cols=riskCols) + theme(legend.position = 'none')
+vig.plot(subset(back, Patient == 33), subset(patient.info, Patient == 33),cols=riskCols) + theme(legend.position = 'none')
+vig.plot(subset(back, Patient == 49), subset(patient.info, Patient == 49),cols=riskCols) + theme(legend.position = 'none')
+vig.plot(subset(back, Patient == 51), subset(patient.info, Patient == 51),cols=riskCols) + theme(legend.position = 'none')
+
+vig.plot(subset(back, Patient == 6), subset(patient.info, Patient == 6),cols=riskCols) + theme(legend.position = 'none')
+vig.plot(subset(back, Patient == 54), subset(patient.info, Patient == 54),cols=riskCols) + theme(legend.position = 'none')
+```
+
+```{r, echo=F, warning=F, message=F, fig.width=4, fig.height=1}
+grid.arrange(vigLeg)
 ```
@@
 
- ```{r rx, echo=F, warning=F, message=T}
+```{r, rx, echo=F, warning=F, message=T, fig.width=8, fig.height=5}
 
@@
 
-In the literature the risk of BE progression to HGD/cancer is estimated to be 0.2-0.7% per year.
-
-So for every 100 BE patients, at most `r 0.07*100` patients will progress in a year.
-
-Lifetime risk is assumed to be ~10% presuming 10 years followup (`r 0.7*10`%).
-Two ways to do this, take one year and look at the progression predictions...or take all of them and compare to the lifetime risk.
+Median follow up for our discovery cohort was `r median(sum.patient.data$end.year-sum.patient.data$start.year)`.
+
+Adjust the RR of the orginal model with 1-3.5% offset to get an absolute risk (AR) with the understanding that the cohort was age matched and unbalanced.
 
@@
 # This isn't quite right...we'd want to do something more like a 5 year 
-pop = round((table(all$Status)['P']/(0.07*100))*100)
-
-```
-
-### Lifetime adjusted risk
-
-```{r adjProbRisk, echo=F, fig.height=6, fig.width=10}
-#popL = round((table(all$Status)['P']/(0.1*100))*100)
+#pop = round((table(all$Status)['P']/(0.07*100))*100)
+
+```
+
+```{r estimateRisk, echo=F, warning=F, message=F, eval=F}
+# Probability that an individual will progress in a lifetime (20 years) given a rate of progression of 0.07 is about 77%, the probability for a specific individual is less
+1-(0.93^20)
+
+(1-(0.99^3))*15
+(1-(0.99^3))/10
+(1-(0.99^3))/10
+
+# Or probability an individual will progress given a rate of 1%-3.5%
+poft = sapply(seq(0.96,0.99, .01), function(x) {
+ sapply(1:5, function(y) 1-(x^y) )
+})
+
+poft = as.data.frame(poft)
+colnames(poft) = paste((1-seq(0.96,0.99, .01))*100, '%', sep='')
+
+rp = apply(poft,1,function(x) cbind(mean(x),min(x), max(x)) )
+
+rownames(rp) = c('mean','min','max')
+colnames(rp) = paste('Year',1:5)
+
+rr = range(as.data.frame(subset(back, Pathology %in% c('BE'), select=c('Patient','Pathology','months.before.final','OR')))$OR)
+
+p = subset(back, Status == 'P' & months.before.final > 6 & OR > 10)$Patient
+subset(sum.patient.data, Patient %in% p) %>% group_by(Patient) %>% summarise(final.endo-initial.endo)
+
+rrPath = back %>% group_by(Status, Pathology) %>% summarise( 'mean.RR'=mean(OR),'ci.min'=mean(OR)-(1.96*sd(OR)), 'ci.max'=mean(OR)+(1.96*sd(OR)) )
+
+ggplot(rrPath, aes(Status, mean.RR)) + #scale_color_gradientn(colors=myPal, limits=c(min(rrPath[,3:5]), max(rrPath[,3:5])),  name='') +
+  facet_wrap(~Pathology, ncol=1) + 
+  geom_point() +  geom_errorbar(aes(ymin=ci.min, ymax=ci.max), width=0.1) + 
+  geom_text(aes(y=ci.min, label=round(ci.min,2)), vjust=2) + geom_text(aes(y=ci.max, label=round(ci.max,2)), vjust=2) +
+  coord_flip() + plot.theme + labs(y='Mean RR', x='')
+
+as.data.frame(arrange(subset(back, Status == 'P' & months.before.final > 6 & Pathology %in% c('BE','ID','LGD') & OR > 1), Patient, months.before.final))
+
+ggplot(t(rp), aes(1:5, mean)) +
+  geom_ribbon(aes(ymin=mean/abs(rr[1]),ymax=mean*rr[2]), fill='rosybrown1') +
+  geom_ribbon(aes(ymin=min,ymax=max), fill='grey88') +
+  geom_line() + 
+  geom_text(data=data.frame(t(rp[,5])), aes(x=5,y=min, label='1%')) + geom_text(data=data.frame(t(rp[,5])), aes(x=5,y=max, label='4%')) +
+  geom_line(aes(y=mean*rr[2]), color='red') +
+  geom_line(aes(y=mean/abs(rr[1])), color='blue') +
+  labs(y='P(P)', x='Year', title='Prob. of individual patient progression', subtitle='1-4% over 5 years') + plot.theme
+
+subset(back, OR > 10)
+cor.test(back$OR, as.integer(back$Pathology))
+as.data.frame(subset(back, Pathology %in% c('HGD','IMC'), select=c('Patient','Pathology','months.before.final','OR')))
+
+as.data.frame(subset(back, Status == 'P' & OR < 2 & OR > -2, select=c('Status','Patient','Pathology','months.before.final','OR')))
+
+hist(subset(back, Pathology %in% c('BE'))$OR)
+
+```
+
+```{r 5yrrisk, echo=F, fig.height=8, fig.width=12}
 cases = table(sum.patient.data$Status)['P']
-popL = round((table(sum.patient.data$Status)['P']/(0.1*100))*100)
+
+mn = round((table(sum.patient.data$Status)['P']/(0.01*100))*100)
+m = round((table(sum.patient.data$Status)['P']/(0.0225*100))*100)
+mx = round((table(sum.patient.data$Status)['P']/(0.035*100))*100)
+
+offsetMin = log(cases/mn)
+offsetMean = log(cases/m)
+offsetMax = log(cases/mx)
 
 #binom.test(cases, popL, 0.1)
-
 # Offset for the linear predictor would then be -log(pop/44)
 # Effectively this is my relative risk for the population
-offset = log(cases/popL)
+#offset = log(cases/popL)
 
@@
 loopreds$P0 = 1/(1+exp(-loopreds$OR))
-loopreds$P1 = 1/(1+exp(-loopreds$OR+abs(offset))) # Adjust OR for RR
-
-mean(loopreds$P0)
-mean(loopreds$P1)
-
-ql = cbind.data.frame('y'=quantile(loopreds$OR)[c(2,4)], 'labels'=paste(c(25,75), '%', sep=''))
+
+loopreds$Mean = 1/(1+exp(-loopreds$OR+abs(offsetMean))) # Adjust OR for RR
+loopreds$Min = 1/(1+exp(-loopreds$OR+abs(offsetMin))) # Adjust OR for RR
+loopreds$Max = 1/(1+exp(-loopreds$OR+abs(offsetMax))) # Adjust OR for RR
+
+#ql = cbind.data.frame('y'=quantile(loopreds$OR)[c(2,4)], 'labels'=paste(c(25,75), '%', sep=''))
 
@@
 
-ggplot(loopreds, aes(P1)) + 
+ggplot(melt(loopreds[c('Min','Mean','Max')], measure.vars='Mean'), aes(value, fill=variable)) + 
+  geom_density(alpha=0.5) + plot.theme
+
+ggplot(melt(loopreds[c('Min','Max')]), aes(value)) + facet_wrap(~variable, ncol=2) +
   geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + 
-    #geom_vline(xintercept = mean(adjRR)) +
     scale_fill_gradientn(colors = myPal,  name='') + 
     plot.theme + theme(legend.position = 'none') + 
-    labs(y='n Samples', x='Probability',
-         title='Sample Predictions: Lifetime risk adjusted')
-
-ggplot(loopreds, aes(OR)) + 
-    geom_histogram(aes(fill=..x..), bins=10, show.legend = F) + 
-    scale_fill_gradientn(colors = myPal,  name='') + 
-    #scale_fill_distiller(palette = 'RdYlBu', name='P(P)') + 
-    plot.theme + theme(legend.position = 'none') + 
-    labs(y='n Samples', x='Adjusted Relative Risk',
-         title='Sample Predictions: Lifetime risk adjusted')
-
-# Probability that an individual will progress in a lifetime (20 years) given a rate of progression of 0.07 is about 77%, the probability for a specific individual is less
-1-(0.93^20)
-
-```
-
-```{r lifetimeFPR, echo=F, fig.height=6, fig.width=6}
+    labs(y='n Samples', x='Absolute Risk',
+         title='5-year adjusted risk', subtitle='Adjusted for upper & lower bound: 1-3.5% risk')
+
+# ggplot(loopreds, aes(Min)) +
+#     geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + 
+#     scale_fill_gradientn(colors = myPal,  name='') + 
+#     plot.theme + theme(legend.position = 'none') + 
+#     labs(y='n Samples', x='Absolute Risk',
+#          title='Sample Predictions: 5 year adjusted risk', subtitle='Adjusted for lower bound, 1% risk')
+# 
+# ggplot(loopreds, aes(Max)) +
+#     geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + 
+#     scale_fill_gradientn(colors = myPal,  name='') + 
+#     plot.theme + theme(legend.position = 'none') + 
+#     labs(y='n Samples', x='Absolute Risk',
+#          title='5-year adjusted risk', subtitle='Adjusted for upper bound, 3.5% risk')
+
+```
+
+```{r lifetimeFPR, echo=F, fig.height=6, fig.width=6, eval=F}
 ## FPR estimates based on lifetime risks
@@
 
+```{r clonality, echo=F, message=F, warning=F, eval=F}
 ## Clonal Heterogenity
-
-```{r clonality, echo=F, message=F, warning=F, eval=F}
 patient.info %>% group_by (Hospital.Research.ID, Patient, PID) %>% summarise(
@@
 
-```{r cox, warning=F, message=F, echo=F, eval=F}
+```{r cox, warning=F, message=F, echo=F, eval=T}
 library(survival)
@@
 prog.pred = lapply(pg.samp, function(df) {
-  df = df[,c('Hospital.Research.ID', 'Status', 'Patient', 'PID', 'Status', 'Endoscopy.Year','Initial.Endoscopy', 'Final.Endoscopy', 'Pathology', 'Samplename','Prediction','Prediction.Dev.Resid')] %>% rowwise() %>% mutate(
+  df = df[,c('Hospital.Research.ID', 'Status', 'Patient', 'Age.at.diagnosis', 'PID', 'Status', 'Endoscopy.Year','Initial.Endoscopy', 'Final.Endoscopy', 'Pathology', 'Samplename','Prediction','Prediction.Dev.Resid')] %>% rowwise() %>% mutate(
       'Time.to.Final'=Final.Endoscopy-Endoscopy.Year)
@@
 pp = do.call(rbind, prog.pred)
+
 pp[is.na(pp$tstop), 'tstop'] = 0.5
 pp$Pathology = as.character(pp$Pathology)
+pp$AAD = cut(pp$Age.at.diagnosis, breaks = 3)
+
+pp = merge(pp, back[c('Samplename','Risk', 'quants')], by='Samplename')
+pp$Risk = as.character(pp$Risk)
+
+sfit = survfit( Surv(tstart, tstop, Diagnosis, type='interval') ~ Risk, data=pp, conf.int=0.95, conf.type='log', na.action=na.omit)
+ggsurv(sfit,cens.shape=3,lty.est=1,back.white=F) + xlim(0,5) + labs(y='Diagnosed', title='~Risk')
+
+cox = coxph( Surv(tstart, tstop, Diagnosis) ~ Risk, data=pp) 
+ggsurv(survfit(cox),cens.shape=3,lty.est=1,back.white=F,
+      main="coxph(Survival, Diagnosis) ~ Risk" ) + 
+      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),2),"\n", nrow(pp), 'samples'))
+#summary(cox)
+
+sfit = survfit( Surv(tstart, tstop, Diagnosis, type='interval') ~ AAD, data=pp, conf.int=0.95, conf.type='log', na.action=na.omit)
+ggsurv(sfit,cens.shape=3,lty.est=1,back.white=F) + xlim(0,5) + labs(y='Diagnosed', title='~Age') 
+cox = coxph( Surv(tstart, tstop, Diagnosis) ~ AAD, data=pp) 
+ggsurv(survfit(cox),cens.shape=3,lty.est=1,back.white=F,
+      main="coxph(Survival, Diagnosis) ~ Age at diagnosis" ) + 
+      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),2),"\n", nrow(pp), 'samples'))
+#summary(cox)
 
 sfit = survfit( Surv(tstart, tstop, Diagnosis, type='interval') ~ Pathology, data=pp, conf.int=0.95, conf.type='log', na.action=na.omit)
-ggsurv(sfit,cens.shape=3,lty.est=1,back.white=F)
-summary(sfit)
-
+ggsurv(sfit,cens.shape=3,lty.est=1,back.white=F) + xlim(0,5) + labs(y='Diagnosed', title='~Pathology')
 cox = coxph( Surv(tstart, tstop, Diagnosis) ~ Pathology, data=pp) 
@@
       main="coxph(Survival, Diagnosis) ~ Pathology" ) + 
-      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),2),"\n", nrow(pp), 'samples'))
-summary(cox)
+      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),2),"\n", nrow(pp), 'samples')) + xlim(0,5)
+#summary(cox)
 
@@
 gcols = grep('\\d+:', colnames(sample.pred), value=T)
-sample.pred = sample.pred[c('tstart', 'tstop','Diagnosis','Pathology',gcols)]
+sample.pred = sample.pred[c('tstart', 'tstop','Diagnosis','Pathology','Age.at.diagnosis',gcols)]
+
+sample.pred$Age.bracket = as.character(cut(sample.pred$Age.at.diagnosis, breaks=3))
 
@@
 #summary(cox)
-
-# glm selected features
+```
+### Cox GLM selected features
+
+```{r echo=F, warning=F, echo=F}
+# glm selected features + Path
 cox2 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', 'Pathology',  grep('^\\d+:', ch$label, value=T))]) 
-ggsurv(survfit(cox2))
-summary(cox2)
+ggsurv(survfit(cox2),cens.shape=3,lty.est=1,back.white=F,
+      main="coxph(Survival, Diagnosis) ~ GLM features + Pathology" ) + 
+      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox2),2),"\n", nrow(pp), 'samples')) 
+#summary(cox2)
+
+cox2 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis','Age.bracket', grep('^\\d+:', ch$label, value=T))]) 
+ggsurv(survfit(cox2),cens.shape=3,lty.est=1,back.white=F,
+      main="coxph(Survival, Diagnosis) ~ GLM features + Age" ) + 
+      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox2),2),"\n", nrow(pp), 'samples')) 
+#summary(cox2)
+
+cox2 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis',grep('^\\d+:', ch$label, value=T))]) 
+ggsurv(survfit(cox2),cens.shape=3,lty.est=1,back.white=F,
+      main="coxph(Survival, Diagnosis) ~ GLM features" ) + 
+      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox2),2),"\n", nrow(pp), 'samples')) 
+#summary(cox2)
+
+# chr arms only
+options(scipen=999) # turn sci notation off
+coxArms = coxph(Surv(tstart, tstop, Diagnosis) ~., data=sample.pred[,c('tstart', 'tstop','Diagnosis', with(arm.loc, paste(chrom, ':', start, '-', end, sep='')))])
+ggsurv(survfit(coxArms),cens.shape=3,lty.est=1,back.white=F,
+      main="coxph(Survival, Diagnosis) ~ chr arms" ) + 
+      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(coxArms),2),"\n", nrow(pp), 'samples')) 
+#summary(coxArms)
 
@@
 cox17 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', grep('^17:',colnames(sample.pred), value=T))]) 
-ggsurv(survfit(cox17))
-summary(cox17)
+ggsurv(survfit(cox17),cens.shape=3,lty.est=1,back.white=F,
+      main="coxph(Survival, Diagnosis) ~ chr17" ) + 
+      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox17),2),"\n", nrow(pp), 'samples')) 
+#summary(cox17)
 
@@
 ggsurv(survfit(cox11))
-summary(cox11)
+#summary(cox11)
 
@@
 ggsurv(survfit(cox13))
-summary(cox13)
+#summary(cox13)
 
@@
 
-```{r echo=F, message=F, warning=F, eval=F}
+```{r echo=F, message=F, warning=F, eval=T}
 plot.waldtest<-function(wt) {
@@
 
+# age = do.call(cbind.data.frame, lapply(unique(coxDF$Age.at.diagnosis), function(p) {
+#   cbind( as.integer(coxDF$Age.at.diagnosis == p))
+# }))
+# colnames(age) = unique(coxDF$Age.at.diagnosis)
+age = coxDF$Age.at.diagnosis
+
 ## GLM regions only
-coxDF = cbind(coxDF[,1:3],pathology[,c('BE','ID')], coxDF[,grep('^\\d+:', ch$label, value=T)])
+coxDF = cbind(coxDF[,1:3],pathology[,c('BE','ID')], age, coxDF[,grep('^\\d+:', ch$label, value=T)])
 coxDF$Pathology = NULL
-dim(coxDF)
 
@@
 
-surv = with(coxDF, Surv(tstart, tstop, Diagnosis))
-
-groups = factor(c(rep('clinical',2),c(rep('genomic', length(grep('\\d+', colnames(coxDF)))))))
-rfx = CoxRFX(coxDF[,-c(1:3)], surv, groups, verbose=F, nu=1)
+coxSM = cbind(coxDF[,c(1:3)], StandardizeMagnitude(coxDF[,-c(1:3)]))
+
+#surv = with(coxDF, Surv(tstart, tstop, Diagnosis))
+surv = with(coxSM, Surv(tstart, tstop, Diagnosis))
+groups = factor(c(rep('clinical',3),c(rep('genomic', length(grep('\\d+', colnames(coxDF)))))))
+rfx = CoxRFX(coxSM[,-c(1:3)], surv, groups, verbose=T, nu=1)
 summary(rfx)
@@
 wt[is.na(wt$hazards), c('hazards', 'coef.y')] = 0
-
 plot.waldtest(wt) + labs(x='RFX coef',y='GLM coef', title='RFX vs GLM coefficients', subtitle='Labeled points have a significant p-value in RFX (except BE/ID)')
 
-```
-
+options(scipen = 999)
+# arms only
+arms = with(arm.loc, paste(chrom, ':', start, '-', end, sep=''))
+coxDfArms = coxDF[,arms]
+colnames(coxDfArms) = paste('chr', gsub(':|-','.', grep('^\\d+', colnames(coxDfArms), value=T)), sep='')
+
+surv = with(coxDF, Surv(tstart, tstop, Diagnosis))
+groups = factor(c(rep('clinical',3), rep('genomic', ncol(coxDfArms))))
+
+rfx = CoxRFX( cbind.data.frame(coxSM[,c('ID','BE','age_10')], coxDfArms), surv, groups, verbose=T, nu=1)
+summary(rfx)
+
+plot(rfx)
+wt = WaldTest(rfx)
+survConcordance(surv ~ predict(rfx))
+
+```
+
+```{r, eval=F}
 # TODO: Relative risk
-```{r, eval=F}
 pred = predict(f, newx=dysplasia.df, s=performance.at.1se$`0.9`$lambda)

### 2018-02-05 Paper.Rmd
...

@@
 
-globalnHGDCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology %in% c('LGD','HGD','IMC')), labeller(Status=dys.status, chr=label_value)) + labs(title='LGD/HGD/IMC Only')
+#globalnHGDCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology %in% c('LGD','HGD','IMC')), labeller(Status=dys.status, chr=label_value)) + labs(title='LGD/HGD/IMC Only')
 
@@
 
-globalCN.plot + theme(legend.position = 'none')
-
-globalnHGDCN.plot + theme(legend.position = 'none')
-
-globalBECN.plot + theme(legend.position = 'none')
+gp = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples') + theme(text=element_text(size=12))
+ggsave(filename='plots/globalCN.png', plot=gp, width=16, height=4.6, units='in', dpi=300)
+
+#globalnHGDCN.plot + theme(legend.position = 'none')
+
+gbe = globalBECN.plot + theme(legend.position = 'none') + labs(y='mean adj.seg.', title="Non-Dysplastic Barrett's Samples") + theme(text=element_text(size=12))
+ggsave(filename='plots/globalBECN.png', plot=gbe, width=16, height=4.6, units='in', dpi=300)
 
@@
 
-# m = melt(insilico.rocs, measure.vars = 'auc')
-# ggplot(m, aes(variable,value)) + ylim(0,1) +
-#   geom_boxplot(fill='mintcream', color='grey55') + geom_jitter(width=0.2, color='darkblue', size=1.5, show.legend = F) + 
-#   geom_text(data=subset(m, value %in% range(value)), aes(label=round(value,2)), vjust=0.02, hjust=0.2) +
-#   plot.theme + labs(x='',y='AUC',title='AUC', subtitle='20 insilico 20:69 splits')
-
 ```
@@
 roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
-roc$model = 'CN Model'
-
-multi.roc.plot(list(roc), colors=brewer.pal(3,'Greys')[3]) + labs(title='Discovery cohort ROC')
-
-insilico.rocs = rbind(insilico.rocs, c(pROC::ci.auc(roc), pROC::coords(roc, 'best')))
-
-insilico.rocs[1:20,'split'] = paste('in-silico', 1:20)
-insilico.rocs[21,'split'] = 'discovery'
-
-ggplot(insilico.rocs, aes(split, auc)) + geom_point() + ylim(0,1) +
-  geom_errorbar(aes(ymin=auc.ci.min, ymax=auc.ci.max), width=0.2) + 
-  coord_flip() + plot.theme + labs(y='AUC', x='', title='AUC per split', subtitle='20 insilico 20:69 splits')
+roc$model = 'discovery'
+roc$is = 'discovery'
+
+droc = multi.roc.plot(list(roc), colors=brewer.pal(3,'Greys')[3]) + labs(title='Discovery cohort ROC')
+ggsave(filename='plots/discovery_roc.png', plot=droc, width=6, height=4, units='in', dpi=300)
+
+insilico.rocs[['discovery']] = roc
+
+df = do.call(rbind, lapply(insilico.rocs, function(r) 
+    cbind.data.frame('Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'model'=r$model, 'is'=r$is) 
+))
+
+p = ggplot(df, aes(Specificity, Sensitivity,color=is)) + geom_line(size=1) + 
+    geom_line(data=subset(df, model == 'discovery'), color='red', size=1) +
+    scale_x_reverse() +
+    scale_color_brewer(palette='Blues') +
+    labs(title='ROC, all splits', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
+    theme(legend.position = 'none')
+p
+
+ggsave(filename='plots/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)
+
+ir = ggplot(insilico.rocs, aes(is, auc, color=split)) + ylim(0,1) +
+  geom_hline(data=m, aes(yintercept=auc, linetype=line), color='grey32', show.legend=F) + 
+  geom_point(size=2) + geom_errorbar(aes(ymin=auc.ci.min, ymax=auc.ci.max), width=0.3) + 
+  geom_text(data=m, aes(label=paste(variable, round(auc, 2), sep='\n')), color='grey32', nudge_y=0, show.legend=F) + 
+  scale_linetype_manual(values=c('dotdash','longdash')) +
+  plot.theme + scale_color_brewer(palette='Set1', name='Type') +
+  theme(axis.text.x=element_blank(), legend.position = 'bottom') +
+  labs(y='AUC', x='Data splits', title='AUC per split', subtitle='20 insilico 20:69 splits')
+ir
+ggsave(filename='plots/auc_splits.png', plot=ir, width=7, height=8, units='in', dpi=300)
 
@@
 pathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('LGD','HGD','IMC'), 1, 0) )
-pathroc$model = 'Dysplasia (HGD & LGD)'
+pathroc$model = 'HGD & LGD'
 
@@
 ```{r, echo=F, warning=F, message=F,fig.height=4, fig.width=4}
-
 pendo = do.call(rbind, pg.samp) %>% group_by(Patient, Status, Hospital.Research.ID, PID ) %>% summarise(
@@
   pROC::auc(roc, partial.auc=p, partial.auc.focus=foc, partial.auc.correct=T)
-
 }
@@
       get.roc.stat(pathroc),
-      get.roc.stat(roc, 0.6))
-m$model = factor(m$model, levels=c('p53 IHC','HGD Only','Dysplasia (HGD & LGD)','CN Model'), ordered = T)
+      get.roc.stat(roc))
+m$model = factor(m$model, levels=c('p53 IHC','HGD Only','HGD & LGD','CN Model'), ordered = T)
 
@@
 
-cols = brewer.pal(11, "PRGn")
+cols = brewer.pal(4, "PRGn")
 dodge = position_dodge(width=0.9)
-ggplot(melt(subset(m, Measure == 'AUC'), measure.vars='value'), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Measure)) + 
+
+comp = ggplot(melt(subset(m, Measure == 'AUC'), measure.vars='value'), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Measure)) + 
   geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
-  geom_text(aes(label=paste(round(value,2)*100,'%',sep=''), y=1, vjust=-1 ), position=dodge, color='white')  +
+  geom_text(aes(y=(value*100)-4, label=paste(round(value,2)*100,'%',sep='')), position=dodge, color='white')  +
   scale_fill_manual(values=cols, name='') + plot.theme + 
   theme(legend.position = 'none', axis.text.x = element_text(angle=45, hjust=1)) + labs(x='', y='AUC', title='ROC Comparison') 
+comp
+
+ggsave('plots/compare_auc.png', plot=comp, width=6, height=8, units="in", limitsize=F, dpi = 300)
 
@@
 ```{r pcalib, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
-ggplot(pred.confidence, aes(mn, perc)) + 
+pcal = ggplot(pred.confidence, aes(mn, perc)) + 
   geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
-  geom_point() + geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
+  geom_point() + #geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
   geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration') +
-  scale_color_manual(values=riskCols) + plot.theme + theme(legend.position = 'none')
+  scale_color_manual(values=riskCols) + plot.theme + theme(legend.position = 'none', text=element_text(size=10))
+pcal
+
+ggsave(filename='plots/pred_calib.png', plot=pcal, width=6, height=6, units='in', dpi=300)
 ```
@@
 
-ggplot(back, aes(OR)) + geom_histogram(aes(fill=..x..), bins=10, show.legend = F) +
+unadjRR = ggplot(back, aes(OR)) + geom_histogram(aes(fill=..x..), bins=10, show.legend = F) +
   #scale_fill_distiller(palette = 'RdYlBu', type='div',name='P(P)') +
@@
   labs(y='n Samples', x='Relative Risk', title='Unadjusted relative risk') + plot.theme
-
-ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) +
+unadjRR
+
+ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR, width=6, height=5, units='in', dpi=300)
+
+predhist = ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) +
     scale_fill_gradientn(colors = myPal,  name='') + 
@@
     plot.theme + labs(title='Predictions, all samples model', y='n Samples', x='Probability') 
+predhist
+
+ggsave(filename='plots/pred_hist.png', plot=predhist, width=5, height=5, units='in', dpi=300)
 
@@
 
-min.theme = theme(text=element_text(size=10), panel.background=element_blank(), strip.background=element_rect(fill="grey97"), 
-                   strip.text = element_text(size=10), axis.text = element_blank(),
+min.theme = theme(text=element_text(size=8), panel.background=element_blank(), strip.background=element_rect(fill="grey97"), strip.text.y = element_text(size=6),
+                   strip.text.x = element_text(size=12), axis.text = element_blank(), axis.ticks = element_blank(),
                    axis.line=element_line(color='black'), panel.grid.major.y =element_line(color='grey95'), panel.grid.major.x = element_blank(),
@@
 
-  p = ggplot(df, aes(x=chr.length)) + facet_grid(Endoscopy.Year+ogj~chr, scales='free', space='free_x', labeller = labeller(.multi_line=F)) +
+  blanklabel <- function(s) {
+    return('')
+  }
+
+  p = ggplot(df, aes(x=chr.length)) + facet_grid(Endoscopy.Year+ogj~chr, scales='free', space='free_x', labeller = labeller(.multi_line = F)) +
       geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=ogj), show.legend=T) + 
@@
 
-p1 = mtn.spatial(subset(globalCNdf, Patient == 48 & seg.type != 'arms'), back, ogjLimits, 'Patient 48 (P)')
+#p1 = mtn.spatial(subset(globalCNdf, Patient == 48 & seg.type != 'arms'), back, ogjLimits, 'Patient 48 (P)')
 p2 = mtn.spatial(subset(globalCNdf, Patient == 20 & seg.type != 'arms'), back, ogjLimits,'Patient 20 (P)' )
-p3 = mtn.spatial(subset(globalCNdf, Patient == 60 & seg.type != 'arms'), back, ogjLimits,'Patient 60 (P)' )
-
-p4 = mtn.spatial(subset(globalCNdf, Patient == 3 & seg.type != 'arms'), back,ogjLimits,'Patient 3 (NP)' )
-p5 = mtn.spatial(subset(globalCNdf, Patient == 15 & seg.type != 'arms'), back,ogjLimits,'Patient 15 (NP)' )
+#p3 = mtn.spatial(subset(globalCNdf, Patient == 60 & seg.type != 'arms'), back, ogjLimits,'Patient 60 (P)' )
+
+#p4 = mtn.spatial(subset(globalCNdf, Patient == 3 & seg.type != 'arms'), back,ogjLimits,'Patient 3 (NP)' )
+#p5 = mtn.spatial(subset(globalCNdf, Patient == 15 & seg.type != 'arms'), back,ogjLimits,'Patient 15 (NP)' )
 p6 = mtn.spatial(subset(globalCNdf, Patient == 51 & seg.type != 'arms'), back,ogjLimits,'Patient 51 (NP)' )
@@
 
-grid.arrange(p1$legend)
+grid.arrange(p2$legend)
+
+ggsave(filename='plots/spatial_mtn_pt20.png', plot=grid.arrange(p2$grob), width=14, height=6, units='in', dpi=300)
+ggsave(filename='plots/spatial_mtn_pt51.png', plot=grid.arrange(p6$grob), width=14, height=6, units='in', dpi=300)
 
@@
 
-ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
+rp = ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
   scale_fill_manual(values=riskCols) + scale_color_manual(values=c('white','black','white')) +
@@
   geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep=''), color=Risk), position=position_stack(vjust = 0.5), size=5, show.legend = F) +
-  facet_grid(~Status) + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
+  facet_grid(~Status) + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom', text=element_text(size=10))
+rp
+
+ggsave(filename='plots/pred_by_path.png', plot=rp, width=9, height=7, units='in', dpi=300)
 
@@
 p53Path = arrange(p53Path, Pathology, p53.Status)
-col = brewer.pal(11, "PRGn")[c(9,3)]
-ggplot(p53Path, aes(Pathology, ratio*100, group=Pathology, fill=p53.Status)) + geom_bar(stat='identity', position = position_stack()) +
+col = brewer.pal(11, "PRGn")[c(10,2)]
+rp = ggplot(p53Path, aes(Pathology, ratio*100, group=Pathology, fill=p53.Status)) + geom_bar(stat='identity', position = position_stack()) +
   scale_fill_manual(values=col, name='p53 IHC', labels=c('Normal','Aberrant')) + 
   geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101)) +
-  geom_text(aes(label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5) +
+  geom_text(aes(label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5,color='white') +
   plot.theme + labs(title='Samples predicted by aberrant p53 IHC', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
+rp
+ggsave(filename='plots/p53_pred.png', plot=rp, width=7, height=6, units='in', dpi=300)
 
@@
 
-ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) + ylim(0,105) +
+rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) + ylim(0,105) +
   geom_bar(stat='identity', fill='darkred', alpha=0.8, show.legend = F) + 
@@
   labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') +
-  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1))
+  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12))
+
+ggsave(filename='plots/pred_endo.png', plot=rp, width=9, height=7, units='in', dpi=300)
 
@@
 
-pred.tiles(subset(back, Patient %in% c(6)), ncol=1, cols=myPal, probs=T, OR=F) + labs(title='') + theme(legend.position = 'none')
-pred.tiles(subset(back, Patient %in% c(48)), ncol=1, cols=myPal, probs=T) + labs(title='') + theme(legend.position = 'none')
-pred.tiles(subset(back, Patient %in% c(33)), ncol=1, cols=myPal, probs=T) + labs(title='')+ theme(legend.position = 'none')
-pred.tiles(subset(back, Patient %in% c(89)), ncol=1, cols=myPal, probs=T) + labs(title='')+ theme(legend.position = 'none')
+#pred.tiles(subset(back, Patient %in% c(6)), ncol=1, cols=myPal, probs=T, OR=F) + labs(title='') + theme(legend.position = 'none')
+#pred.tiles(subset(back, Patient %in% c(48)), ncol=1, cols=myPal, probs=T) + labs(title='') + theme(legend.position = 'none')
+#pred.tiles(subset(back, Patient %in% c(33)), ncol=1, cols=myPal, probs=T) + labs(title='')+ theme(legend.position = 'none')
+#pred.tiles(subset(back, Patient %in% c(89)), ncol=1, cols=myPal, probs=T) + labs(title='')+ theme(legend.position = 'none')
 ```
@@
 
-p = pred.tiles(transform(subset(back, Status == 'NP'), Patient=reorder(Patient, pred.ratio)), cols=myPal, probs=T, nrow=2) + labs(title='Non-Progressors') 
-riskLegend = get.legend(p)
-p = p + theme(legend.position = 'none')
-ggsave('npg_heatmap.png', plot = p, scale = 1, width = 52, height = 8, units = "in", limitsize = F, dpi = 300)
-
-p = pred.tiles(transform(subset(back, Status == 'P'), Patient=reorder(Patient, -pred.ratio)), probs=T, cols=myPal,nrow=2) + labs(title='Progressors') + theme(legend.position = 'none')
-ggsave('pg_heatmap.png', plot = p, scale = 1, width = 52, height = 8, units = "in", limitsize = F, dpi = 300)
-
-ggsave('risk_legend.png', plot=grid.arrange(riskLegend), scale=1, dpi=300)
+p1 = pred.tiles(subset(back, Status == 'NP'), cols=myPal, probs=T, nrow=2) + labs(title='Non-Progressors') 
+riskLegend = get.legend(p1)
+p1 = p1 + theme(legend.position = 'bottom')
+riskLegendLong = get.legend(p1)
+
+p1 = p1 + theme(legend.position = 'none')
+
+p2 = pred.tiles(subset(back, Status == 'P'), probs=T, cols=myPal,nrow=2) + labs(title='Progressors') + theme(legend.position = 'none')
+
+ggsave('plots/pred_heatmap.png', plot=grid.arrange(p2,p1,nrow=2), width=52, height=16, units="in", limitsize=F, dpi = 300)
+
+ggsave('plots/npg_heatmap.png', plot=p1, width=52, height=8, units="in", limitsize=F, dpi = 300)
+ggsave('plots/pg_heatmap.png', plot=p2, width=52, height=8, units="in", limitsize=F, dpi = 300)
+
+ggsave('plots/risk_legend.png', plot=grid.arrange(riskLegend), width=2,height=4, dpi=300)
+
+ggsave('plots/risk_legend_long.png', plot=grid.arrange(riskLegendLong), width=8,height=2, dpi=300)
 
@@
 ov = findOverlaps(chGR, genesGR)
-genes = as.data.frame(genesGR[unique(subjectHits(ov))])
-colnames(genes)[1] = c('chr')
-
-plot.cn.chr(globalCNdf, chrom='17', chr.lengths, genes=genes, haz=chh) + labs(
-title="Segment Values Samples", subtitle=paste('NonP:Prog - ', paste(table(patient.info$Status), collapse=':'))) + theme(legend.position = 'none')
+#genes = as.data.frame(genesGR[unique(subjectHits(ov))])
+#colnames(genes)[1] = c('chr')
+
+g17 = plot.cn.chr(globalCNdf, chrom='17', chr.lengths, genes=genes, haz=chh, label=labeller(chr=label_both, Status=n.status)) + labs(
+title="All Samples") + theme(legend.position = 'none')
+g17
+
+g17 = g17 + theme(text=element_text(size=11), strip.text = element_text(size=11))
+ggsave(filename='plots/chr17.png', plot=g17, width=10, height=5, units='in', dpi=300)
+
 ```
@@
 length(which(round((subset(distanceTSG, distanceToTSG > 0)$distanceToTSG/5e6)-0.5) == round((avgTSGdist/5e6)-0.5)))
-
 ```
@@
 
-cn.mtn.plot(subset(valCNdf, seg.type != 'arms'), labeller(Status=n.status, chr=label_value))
+cn.mtn.plot(subset(valCNdf, seg.type != 'arms'), labeller(Status=n.status, chr=label_value)) + theme(legend.position = 'none')
 
@@
 
-```{r echo=F, warning=F, message=F, fig.height=5,fig.width=25}
-pred.tiles(subset(back.v, Status == 'P'), cols=myPal, probs=T, nrow=1) + labs(title='Validation Progressors') + theme(legend.position = 'none')
-pred.tiles(subset(back.v, Status == 'NP'), cols=myPal, probs=T, nrow=1) + labs(title='Validation Non-Progressors') + theme(legend.position = 'none')
+```{r echo=F, warning=F, message=F, fig.height=10,fig.width=25}
+back.v$Patient = factor(as.numeric(back.v$Patient), ordered=T)
+
+p1 = pred.tiles(subset(back.v, Status == 'P'), cols=myPal, probs=T, nrow=1) + labs(title='Validation Progressors', x='') + theme(legend.position = 'bottom') 
+valLeg = get.legend(p1)
+p1 = p1 + theme(legend.position = 'none')
+p2 = pred.tiles(subset(back.v, Status == 'NP'), cols=myPal, probs=T, nrow=1) + labs(title='Validation Non-Progressors') + theme(legend.position = 'none')
+
+grid.arrange(p1, p2, nrow=2)
+
+ggsave(filename='plots/val_heatmap.png', plot=grid.arrange(p1, p2, nrow=2), width=18, height=6, units='in', dpi=300)
+
+ggsave(filename='plots/val_leg.png', plot=valLeg, width=4, height=2, units='in', dpi=300)
+
 ```
@@
 
-ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) +   scale_fill_manual(values=riskCols) + scale_color_manual(values=c('white','black','white')) +
+pr = ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) +   scale_fill_manual(values=riskCols) + scale_color_manual(values=c('white','black','white')) +
   geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), position=position_stack(), size=4, show.legend = F) +
@@
   facet_grid(~Status) + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
+pr
+
+ggsave(filename='plots/val_riskpath.png', plot=pr, width=7, height=7, units='in', dpi=300)
 
@@
 
-```{r, echo=F, warning=F, message=F, fig.width=8, fig.height=5}
+```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=5}
 vign.pts = as_tibble(subset(sum.patient.data, Patient %in% c(3, 33, 34, 49, 56, 20), select=c('Status','Patient','age.diagnosis', 'smoking', 'C','M', 'initial.endo','final.endo', 'total.endos','total.samples')))
@@
 vigLeg = get.legend(p)
-p + theme(legend.position = 'none')
-vig.plot(subset(back, Patient == 56), subset(patient.info, Patient == 56),cols=riskCols) + theme(legend.position = 'none')
-vig.plot(subset(back, Patient == 20), subset(patient.info, Patient == 20),cols=riskCols) + theme(legend.position = 'none')
-vig.plot(subset(back, Patient == 86), subset(patient.info, Patient == 86),cols=riskCols) + theme(legend.position = 'none')
+p = p + theme(legend.position = 'none')
+ggsave('plots/vig_34.png', plot=p, width=6, height=6,dpi=300)
+
+p = vig.plot(subset(back, Patient == 56), subset(patient.info, Patient == 56),cols=riskCols) + theme(legend.position = 'none')
+ggsave('plots/vig_56.png', plot=p, width=6, height=6,dpi=300)
+
+p = vig.plot(subset(back, Patient == 20), subset(patient.info, Patient == 20),cols=riskCols) + theme(legend.position = 'none')
+ggsave('plots/vig_20.png', plot=p, width=6, height=6,dpi=300)
+
+p = vig.plot(subset(back, Patient == 86), subset(patient.info, Patient == 86),cols=riskCols) + theme(legend.position = 'none')
+ggsave('plots/vig_86.png', plot=p, width=6, height=6,dpi=300)
+
+print(p)
 
 ## NP
-vig.plot(subset(back, Patient == 3), subset(patient.info, Patient == 3),cols=riskCols) + theme(legend.position = 'none')
-vig.plot(df=subset(back, Patient == 23),info= subset(patient.info, Patient == 23),cols=riskCols) + theme(legend.position = 'none')
-vig.plot(subset(back, Patient == 33), subset(patient.info, Patient == 33),cols=riskCols) + theme(legend.position = 'none')
-vig.plot(subset(back, Patient == 49), subset(patient.info, Patient == 49),cols=riskCols) + theme(legend.position = 'none')
-vig.plot(subset(back, Patient == 51), subset(patient.info, Patient == 51),cols=riskCols) + theme(legend.position = 'none')
-
-vig.plot(subset(back, Patient == 6), subset(patient.info, Patient == 6),cols=riskCols) + theme(legend.position = 'none')
-vig.plot(subset(back, Patient == 54), subset(patient.info, Patient == 54),cols=riskCols) + theme(legend.position = 'none')
-```
-
-```{r, echo=F, warning=F, message=F, fig.width=4, fig.height=1}
-grid.arrange(vigLeg)
+p = vig.plot(subset(back, Patient == 3), subset(patient.info, Patient == 3),cols=riskCols) + theme(legend.position = 'none')
+ggsave('plots/vig_3.png', plot=p, width=6, height=6,dpi=300)
+
+p = vig.plot(df=subset(back, Patient == 23),info= subset(patient.info, Patient == 23),cols=riskCols) + theme(legend.position = 'none')
+ggsave('plots/vig_23.png', plot=p, width=6, height=6,dpi=300)
+
+p = vig.plot(subset(back, Patient == 33), subset(patient.info, Patient == 33),cols=riskCols) + theme(legend.position = 'none')
+ggsave('plots/vig_33.png', plot=p, width=6, height=6,dpi=300)
+
+p = vig.plot(subset(back, Patient == 49), subset(patient.info, Patient == 49),cols=riskCols) + theme(legend.position = 'none')
+ggsave('plots/vig_49.png', plot=p, width=6, height=6,dpi=300)
+
+p = vig.plot(subset(back, Patient == 51), subset(patient.info, Patient == 51),cols=riskCols) + theme(legend.position = 'none')
+ggsave('plots/vig_51.png', plot=p, width=6, height=6,dpi=300)
+
+p = vig.plot(subset(back, Patient == 6), subset(patient.info, Patient == 6),cols=riskCols) + theme(legend.position = 'none')
+ggsave('plots/vig_6.png', plot=p, width=6, height=6,dpi=300)
+
+p = vig.plot(subset(back, Patient == 54), subset(patient.info, Patient == 54),cols=riskCols) + theme(legend.position = 'none')
+ggsave('plots/vig_54.png', plot=p, width=6, height=6,dpi=300)
+
+print(p)
+
 ```

### 2018-02-05 lib/common_plots.R

@@ -7,7 +7,7 @@ get.legend<-function(a.gplot){
   return(legend)
 }
 
-plot.theme = theme(text=element_text(size=14, face='bold'), panel.background=element_blank(), strip.background =element_rect(fill="lightcyan2"),  
+plot.theme = theme(text=element_text(size=14, face='bold'), panel.background=element_blank(), strip.background =element_rect(fill="grey88"),  
                    strip.text = element_text(size=14, face='bold'), 
                    axis.line=element_line(color='black'), panel.grid.major=element_line(color='grey90'),
                    panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing = unit(0.1, 'lines')  ) 
@@ -143,16 +143,16 @@ pred.tiles <- function(df, probs=F, OR=F, cols=c('green3','orange','red3'), ...)
   }
   p + 
     geom_point(aes(shape=Pathology, color=Risk), fill='white', size=3) + scale_color_manual(values=c('white','black','white'), guide=F) +
-    scale_shape_manual(values=c(1,0,15,24,25), limits=levels(df$Pathology), guide=guide_legend(override.aes=list(fill='black', color='black'))) +
+    scale_shape_manual(values=c(1,0,15,24,25), limits=levels(df$Pathology), guide=guide_legend(override.aes=list(fill='white', color='white'))) +
     #scale_fill_manual(values=cols, limits=levels(df$Risk),name='Progression') +  #scale_color_manual(values=cols, limits=levels(df$Risk),name='Progression') +
     #geom_label(aes(label=Pathology), fill='white', show.legend=F) + 
     #scale_color_manual(values=c('white','black','white'), limits=levels(df$Risk)) +
     facet_wrap(~Patient, scales='free', labeller=label_both, ...) +
     labs(y='Oesophageal Location',x='Months Prior to Endpoint', title='') +
-    plot.theme + theme(axis.text.x=element_text(angle=45, hjust=1)) 
+    plot.theme + theme(axis.text.x=element_text(angle=45, hjust=1), legend.key = element_rect(fill='grey39')) 
 }
 
-plot.cn.chr<-function(df, chrom='17', info=NULL, haz=NULL, genes=NULL) {
[…]

### 2018-02-20 Paper.Rmd
Updating the pipeline to make it more generic

@@
 source('lib/common_plots.R')
+source('lib/data_func.R')
 
 data = '~/Data/Ellie'
-tile.w=c(5e06, 'arms')
+info.dir = '~/Data/Ellie/QDNAseq'
+#tile.w=c(5e06, 'arms')
 #tile.w=5e+06
@@
 
-includeDemo = F
-
-data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
-analysis.files = list.files(paste(data, 'Analysis', sep='/'), full.names=T)
-
-load(grep('All_patients.Rdata', analysis.files, value=T), verbose=T)
+#includeDemo = F
+
+#data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
+#analysis.files = list.files(paste(data, 'Analysis', sep='/'), full.names=T)
 
 ## Hospital.Research.ID info file
-patient.file = grep('All_patient_info.xlsx', data.files, value=T)
-if (length(patient.file) != 1)
-  stop(paste("Missing/too many patient info file(s) in", data))
-demo.file = grep('Demographics_full.xlsx', data.files, value=T)
+
+patient.file = list.files(info.dir, pattern='All_patient_info.xlsx', recursive=T, full.names=T)
+demo.file = list.files(info.dir, pattern='Demographics_full.xlsx', recursive=T, full.names=T)
+
+if (length(patient.file) != 1 | length(demo.file) != 1)
+  stop(paste("Missing/too many patient info file(s) in", info.dir))
 
@@
 
-validation.patient.data = patient.data[sum.validation.info$Hospital.Research.ID]
-length(validation.patient.data)
-
-patient.data = patient.data[sum.patient.data$Hospital.Research.ID]
-length(patient.data)
-
-patient.data = patient.data[!is.na(names(patient.data))]
-names(patient.data)
-
-sum.patient.data = as.data.frame(subset(sum.patient.data, Hospital.Research.ID %in% names(patient.data))) ## one patient was removed from the study
-nrow(sum.patient.data)
-
 pander(sum.patient.data[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Training patient set')
@@
 
-# Segment size: `r as.character(tile.w)` with demographics: `r includeDemo`
+# Segment size: 5e06 with arms
 
@@
 ```{r, echo=F}
-x = do.call(rbind.data.frame, lapply(patient.data, function(df) df$info[,c('Status','p53.Status')]))
 grid.arrange(
@@
   tableGrob( 
-    rbind( 'Non-progressor'=summary(subset(x, Status == 'NP')$p53.Status), 
-           'Progressor'=summary(subset(x, Status == 'P')$p53.Status))), 
+    rbind('Non-progressor'=summary(subset(patient.info, Status == 'NP')$p53.Status), 
+           'Progressor'=summary(subset(patient.info, Status == 'P')$p53.Status))), 
   nrow=2, as.table=T, heights=c(4,1) )
@@
 ```{r labelsPNP, echo=F, message=F, include=T}
-## binomial: dysplasia 1, BE 0
-labels = unlist(sapply(patient.data, function(df) {
-  df$info = plyr::arrange(df$info, Endoscopy.Year, Pathology)
-  label = as.integer(df$info$Status == 'P') #as.integer(df$info$Pathology %in% c('HGD', 'IMC'))
-  names(label) = df$info$Samplename
-  return(label)
-}))
-names(labels) = sub('.*\\.', '',  names(labels))
-
-pts = do.call(rbind, lapply(patient.data, function(df) {
-  df$info = plyr::arrange(df$info, Endoscopy.Year, Pathology)
-  cbind(df$info[,c('Hospital.Research.ID','Samplename')])
-}))
-rownames(pts) = 1:nrow(pts)
-
-cache.dir = paste(data, 'Analysis', sub('\\+','',paste(tile.w, collapse='_')), sep='/')
+cache.dir = paste(data, 'Analysis/5e6_arms_disc', sep='/')
+
+file = paste(cache.dir, 'model_data.Rdata', sep='/')
+if (!file.exists(file))
+  stop(paste("Missing data file", file))
+
+load(file, verbose=F)
+dim(dysplasia.df)
+length(labels)
 
 file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
-load(file, verbose=T)
-rm(plots, coefs, performance.at.1se, models, cv.patient)
+load(file, verbose=F)
+rm(plots, coefs, performance.at.1se, models)
 
@@
               'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor')$variable)),')', sep=''))
+
 # Average cn value
@@
 
-## Dysplasia only
-dys.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology %in% c('LGD','HGD','IMC'))$variable)), ')', sep=''),
-              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology %in% c('LGD','HGD','IMC'))$variable)),')', sep=''))
-
-#globalnHGDCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology %in% c('LGD','HGD','IMC')), labeller(Status=dys.status, chr=label_value)) + labs(title='LGD/HGD/IMC Only')
-
-## No HGD
-# `%nin%` = Negate(`%in%`)
-# 
-# nhgd.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology %nin% c('HGD','IMC'))$variable)), ')', sep=''),
-#               'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology %nin% c('HGD','IMC'))$variable)),')', sep=''))
-# globalnHGDCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology %nin% c('HGD','IMC')), labeller(Status=nhgd.status, chr=label_value)) + labs(title='Excluding HGD/IMC')
-
 ## NDBE
@@
 
-#globalnHGDCN.plot + theme(legend.position = 'none')
-
 gbe = globalBECN.plot + theme(legend.position = 'none') + labs(y='mean adj.seg.', title="Non-Dysplastic Barrett's Samples") + theme(text=element_text(size=12))
@@
 
-# dfpts = subset(globalCNdf, Patient %in% c('57','76') & seg.type != 'arms')
-# #dfpts = dfpts %>% group_by(Patient, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
-# 
-# pt.status <- c('Non-Progressor'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
-#               'Progressor'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))
-# genomePtCN.plot = cn.mtn.plot(dfpts, labeller(Status = pt.status))
-# 
-# dfpts = subset(globalCNdf, Patient %in% c('57','76') & Pathology == 'BE' & seg.type != 'arms')
-# #dfpts = dfpts %>% group_by(Patient, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
-# 
-# ptbe.status <- c('Non-Progressor'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
-#               'Progressor'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))
-# genomePtBECN.plot = cn.mtn.plot(dfpts,  labeller(Status = ptbe.status )) + labs(title='NDBE samples only')
-# 
-# genomePtCN.plot + theme(legend.position = 'none')
-# genomePtBECN.plot + theme(legend.position = 'none')
-
-#cn.mtn.plot(dfpts,  labeller(Status = ptbe.status ), type='bar') + labs(title='NDBE samples only')
-
-#a = sample( subset(subset(patient.info, Patient == 57),  Endoscopy.Year == min(Endoscopy.Year) )$Samplename, 1)
-#b = sample( subset(subset(patient.info, Patient == 76), Endoscopy.Year == min(Endoscopy.Year) & Pathology %nin% c('HGD','IMC'))$Samplename, 1)
-
-#ind.status <- c('Non-Progressor'='Patient 57 (n=1)','Progressor'='Patient 76 (n=1), excl. HGD/IMC')
-#cn.mtn.plot(subset(globalCNdf, variable %in% c(a,b) & seg.type != 'arms'), labeller(Status = ind.status), type='bar') + labs(title="Individual samples")
 ```
@@
 
-# cx was generated on raw seg values
-plot(dysplasia.df[names(labels) ,'cx'], col=labels+1)
-
-# remove old cx...see what new one would look like
-dysplasia.df = dysplasia.df[,-grep('cx',colnames(dysplasia.df))]
-
-# general variance per sample after the samples have been mean normalized, ignores segment length
-cx = apply(dysplasia.df, 1, function(x) var(x))
+cx = dysplasia.df[,'cx']
 
@@
 tt = t.test(subset(x, Status == 'Progressor')$cx, subset(x, Status == 'Non-Progressor')$cx)
-ggplot(x, aes(Status, cx, fill=Status)) + 
+pv = ggplot(x, aes(Status, cx, fill=Status)) + 
   geom_jitter(width=0.2, color='grey39') + 
-  geom_violin(outlier.colour = NA, alpha=0.8) +
-  labs(x='', y='Variance', title='Per sample variance', subtitle=paste('p-value=', format.pval(tt$p.value, digits=2))) + theme(legend.position = 'none') +
+  geom_boxplot(outlier.colour = NA, alpha=0.8) +
+  labs(x='', y='cx', title='Per sample cx scores', subtitle=paste('p-value=', format.pval(tt$p.value, digits=2))) + theme(legend.position = 'none') +
   plot.theme
+pv
 
 rocX = pROC::roc(Status ~ cx, data=x, auc=T, ci=T, of='thresholds')
-roc.plot(rocX, title='Per sample variance') + plot.theme
+roc.plot(rocX, title='Per sample cx') + plot.theme
+
+ggsave('plots/per_sample_cx.png', plot=pv + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
+
+ggsave('plots/per_sample_cx_roc.png', plot=roc.plot(rocX, title='Per sample cx') + plot.theme + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
 
@@
 
+ggsave('plots/eyebrows.png', plot=do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits')), width=10, height=10, units='in', dpi=300)
+
 ```
@@
 autoplot(models$`0.9`) + theme(legend.position = 'none') + plot.theme
+
+ggsave('plots/glm_0.9.png', plot=autoplot(models$`0.9`) + theme(legend.position = 'none') + plot.theme, width=5, height=5, units='in', dpi=300)
 ```
@@
 ```{r model_perf, message=F, warning=F, echo=F, fig.width=6, fig.height=6}
-
 classification.rate = as_tibble(matrix(ncol=5, nrow=0, dimnames=list(c(), c('model','mean','sme','stable.coef', 'all.coef'))))
 
-mp = model.performance(performance.at.1se, coefs, lims=c(0.7,0.8))
-mp$plot + labs(title=paste('All samples,', folds, 'folds,', splits, 'patient splits'))
+mp = model.performance(performance.at.1se, coefs)
+mp$plot = mp$plot + labs(title=paste('All samples,', folds, 'folds,', splits, 'patient splits'))
+mp$plot
+
+ggsave('plots/performance/ridge_v_lasso.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)
 
@@
 #do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2))
+ggsave('plots/nohgd_eyebrow.png', plot=do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2)), height=10, width=10, units='in', dpi=300)
 ```
@@
 ```{r, echo=F, fig.width=6, fig.height=6}
-mp = model.performance(performance.at.1se, coefs, lims=c(0.69,0.8))
-mp$plot + labs(title='No HGD/IMC samples')
+mp = model.performance(performance.at.1se, coefs)
+mp$plot = mp$plot + labs(title='No HGD/IMC samples')
+mp$plot
+
+ggsave('plots/performance/ridge_v_lasso_nohgd.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)
 
@@
 ```{r loonohgd, echo=F, message=F, warning=T}
-
 file = paste(cache.dir, 'loonohgd.Rdata', sep='/')
@@
   message(paste("loading file", file))
-  load(file, verbose=F)
+  load(file, verbose=T)
 } else {
@@
 ```{r, echo=F, message=F, fig.width=5, fig.height=5}
-
 pg.sampNOHGD = do.call(rbind, pg.sampNOHGD)
@@
 
-  ggplot(melt(perf), aes(x=variable,y=value, group=variable, fill=variable)) +
+p = ggplot(melt(perf), aes(x=variable,y=value, group=variable, fill=variable)) +
     geom_boxplot(fill='lightcoral', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) +
@@
     theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')
+p
+
+ggsave('plots/performance/loo_perf_nohgd.png', plot=p, width=5, height=5, units='in', dpi=300)
 
@@
 
+#hist(hgd.samples$Prediction)
+
 rm(plots, performance.at.1se,coefs, nzcoefs, fits)
@@
 ```{r, echo=F, fig.width=6, fig.height=6}
-mp = model.performance(performance.at.1se, coefs, lims=c(0.69,0.8))
-mp$plot + labs(title='No LGD/HGD/IMC samples')
+mp = model.performance(performance.at.1se, coefs)
+mp$plot = mp$plot + labs(title='No LGD/HGD/IMC samples')
+mp$plot
+
+ggsave('plots/performance/no_lgd_perf.png', plot=mp$plot, height=5, width=5, units='in', dpi=30)
 
@@
 
-ggplot(classification.rate, aes(model, mean, group=model, color=model)) + 
-  geom_text( aes(label=round(mean, 3)), hjust=-0.5, size=5, color='gray39', show.legend = F) +
-  geom_text( aes(y=mean+sme*2, label=paste(all.coef, "\n(", stable.coef, ")", sep='') ), color='gray39', size=5) +
+p = ggplot(classification.rate, aes(model, mean, group=model, fill=model)) + geom_col() +
+  geom_text( aes(label=round(mean, 2)), vjust=2, size=5, color='gray39', show.legend=F) +
+  geom_text( aes(y=0.6, label=paste(all.coef, " coef\n(", stable.coef, ")", sep='') ), color='gray39', size=5) +
   geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme),color='grey39', width=0.1, size=1) + 
-  geom_point(size=4) +
-  labs(x='', y='Classificaiton rate', title=paste('Classification rate per model at', select.alpha)) + 
+#  geom_point(size=2) +
+  labs(x='', y='Classification rate', title=paste('Classification rate per model at', select.alpha)) + 
   plot.theme + theme(axis.text.x=element_text(angle=45,hjust=1,face="bold",size=12), legend.position = 'none') +
-  ylim(0.7,0.8)
-
+  ylim(0,1)
+p
+
+ggsave('plots/performance/compare_classification.png', plot=p, height=5, width=3, units='in', dpi=300 )
 ```
@@
 qq = quantile(ch$hazards, seq(0,1,.15))
-ggplot(ch, aes(coef, hazards, col=hazards >= qq[length(qq)])) + 
+p = ggplot(ch, aes(coef, hazards, col=hazards >= qq[length(qq)])) + 
   geom_point(alpha=0.8, show.legend = F) + 
@@
   labs(title='Top features by hazard ratio', x='Coefficient value', y='Hazard') + plot.theme 
+p
+
+ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)
 
@@
 
-write.table(ch, quote=F, sep='\t', row.names = F, col.names = T, file=paste( paste(tile.w, collapse='_')  ,'_features.txt'))
 ```
@@
 
-globalCN.plot + geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + 
-  geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)
-
-globalBECN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)
-
-# ch$ymax = ifelse(ch$coef > 0, max(genomePtCN.plot$data$value), min(genomePtCN.plot$data$value))
-# chh = chh %>% rowwise() %>% mutate('max'=custom.max(coef, hazards, ymax))
-# genomePtCN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)
-
-# genomePtBECN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)
-
-```
-
-```{r echo=F, fig.height=5, fig.width=6}
-globalCNdf$g.start = globalCNdf$start
-globalCNdf$g.end = globalCNdf$end
-for (i in 2:nrow(chr.lengths)) {
-  rows = which(globalCNdf$chr == chr.lengths[i, 'chrom'])
-  globalCNdf[rows, 'g.start'] = globalCNdf[rows, 'start'] + chr.lengths[i-1, 'genome.length']
-  globalCNdf[rows, 'g.end'] = globalCNdf[rows, 'end'] + chr.lengths[i-1, 'genome.length']
-}
-
-globalCNdf = arrange(globalCNdf, chr, start)
-globalCNdf$segment = with(globalCNdf, paste(chr, ':', start, '-', end, sep=''))
-
-globalCNdf$chr = factor(globalCNdf$chr, levels=c(1:22), ordered = T)
-globalCNdf$Status = factor(globalCNdf$Status)
-
-p1 = ggplot(subset(globalCNdf, seg.type != 'arms'), aes(segment, variable, fill=value, group=Status)) + geom_tile() +
-  scale_fill_gradient2(high='red3', low='green4', mid='white')  + 
-  facet_grid(~chr, scales = 'free_x', space = 'free') + plot.theme + labs(x='',y='Samples') + 
-  theme(axis.ticks=element_blank(), axis.text=element_blank(), panel.spacing.x = unit(0, 'lines'), legend.position='bottom')
-
-p1leg = get.legend(p1)
-p1 = p1 + theme(legend.position = 'none')
-
-p2 = ggplot(subset(globalCNdf, seg.type != 'arms'), aes(y=variable, group=Status)) + 
-  geom_tile(aes(x=0, fill=Status), show.legend = F) + 
-  labs(x=' ', y='') + scale_fill_manual(values=c('darkblue','darkred')) + scale_x_continuous(expand=c(0,0)) +
-  plot.theme + theme( axis.line=element_blank(), axis.text=element_blank(), axis.ticks=element_blank(), panel.grid=element_blank())
-
-grid.arrange(p1, p2, p1leg, ncol=2, nrow=2, widths=c(20,1), heights=c(10,1))
-
-```
-
-```{r echo=F, fig.height=4, fig.width=8}
-ggplot(chh, aes(x=label,y=coef,fill=gl)) + geom_bar(stat='identity') + labs(x='Effects', y='') +
-  coord_flip() + scale_fill_manual(values=c('darkgreen','darkred')) +
-  plot.theme + theme(axis.line.x=element_blank(), axis.ticks=element_blank(),axis.text=element_blank(), panel.grid=element_blank(),panel.grid.major=element_blank(), legend.position='none')
+p = globalCN.plot + geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + 
+  geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')
+
+ggsave('plots/global_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)
+
+p = globalBECN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')
+
+ggsave('plots/ndbe_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)
+
 ```
@@
 
+names(performance.at.1se) = names(pg.samp)
 df = as.data.frame(performance.at.1se)
-ggplot(df, aes(y=performance.at.1se, x='LOO')) + 
+
+p = ggplot(df, aes(y=performance.at.1se, x='LOO')) + 
   geom_boxplot(fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) + geom_jitter(color='grey39', width=0.3) +
@@
   plot.theme
+p
+
+ggsave('plots/performance_loo.png', plot=p, width=5, height=5, units='in', dpi=300)
 
@@
 roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
-roc$model = 'discovery'
+roc$model = 'CN Model'
 roc$is = 'discovery'
@@
 droc = multi.roc.plot(list(roc), colors=brewer.pal(3,'Greys')[3]) + labs(title='Discovery cohort ROC')
+droc
 ggsave(filename='plots/discovery_roc.png', plot=droc, width=6, height=4, units='in', dpi=300)
@@
 p = ggplot(df, aes(Specificity, Sensitivity,color=is)) + geom_line(size=1) + 
-    geom_line(data=subset(df, model == 'discovery'), color='red', size=1) +
+    geom_line(data=subset(df, is == 'discovery'), color='red', size=1) +
     scale_x_reverse() +
@@
 p
-
 ggsave(filename='plots/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)
 
-ir = ggplot(insilico.rocs, aes(is, auc, color=split)) + ylim(0,1) +
-  geom_hline(data=m, aes(yintercept=auc, linetype=line), color='grey32', show.legend=F) + 
-  geom_point(size=2) + geom_errorbar(aes(ymin=auc.ci.min, ymax=auc.ci.max), width=0.3) + 
-  geom_text(data=m, aes(label=paste(variable, round(auc, 2), sep='\n')), color='grey32', nudge_y=0, show.legend=F) + 
-  scale_linetype_manual(values=c('dotdash','longdash')) +
-  plot.theme + scale_color_brewer(palette='Set1', name='Type') +
-  theme(axis.text.x=element_blank(), legend.position = 'bottom') +
-  labs(y='AUC', x='Data splits', title='AUC per split', subtitle='20 insilico 20:69 splits')
-ir
-ggsave(filename='plots/auc_splits.png', plot=ir, width=7, height=8, units='in', dpi=300)
-
-# To find the auc at a given spec or sens
-#pROC::auc(roc, partial.auc=c(0.99,1), partial.acu.focus="sp", partial.auc.correct=T)
-
 df = cbind.data.frame('specificity'=rev(roc$specificities), 'sensitivity'=rev(roc$sensitivities))
-
 thresholds = 
@@
 
-ggplot() + geom_line(data=df, aes(specificity, sensitivity), color='darkblue', size=1.5) + scale_x_reverse() + 
-  geom_point(data=thresholds, aes(x=specificity,y=sensitivity,color=label), size=3,shape=8, show.legend = T) +
-  geom_label_repel(data=thresholds, aes(x=specificity,y=sensitivity,fill=label,label=
-    paste('Threshold ', round(threshold,2), '\n(FPR ', round((1-specificity),2)*100, '%, TPR ', round(sensitivity,2)*100,'%)\nAUC:',round(roc$auc, 2)*100, '%', sep='')),show.legend = F, color='white',nudge_x=0.05) + scale_fill_manual(values=brewer.pal(3,'Dark2')) + scale_color_manual(values=brewer.pal(3,'Dark2')) +
-  labs(title='CN Model ROC', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme
+# ggplot() + geom_line(data=df, aes(specificity, sensitivity), color='darkblue', size=1.5) + scale_x_reverse() + 
+#   geom_point(data=thresholds, aes(x=specificity,y=sensitivity,color=label), size=3,shape=8, show.legend = T) +
+#   geom_label_repel(data=thresholds, aes(x=specificity,y=sensitivity,fill=label,label=
+#     paste('Threshold ', round(threshold,2), '\n(FPR ', round((1-specificity),2)*100, '%, TPR ', round(sensitivity,2)*100,'%)\nAUC:',round(roc$auc, 2)*100, '%', sep='')),show.legend = F, color='white',nudge_x=0.05) + scale_fill_manual(values=brewer.pal(3,'Dark2')) + scale_color_manual(values=brewer.pal(3,'Dark2')) +
+#   labs(title='CN Model ROC', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme
 
@@
 
-```{r pathroc, echo=F, fig.height=4, fig.width=4 }
+```{r pathroc, echo=F, fig.height=4, fig.width=8 }
 #sum(table(subset(patient.info, Status == 'P')$Pathology)[c('IMC','HGD')])/sum(table(subset(patient.info, Status == 'P')$Pathology))
@@
 pathsamples = patient.info[,c('Pathology','Status')]
-
 pathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('LGD','HGD','IMC'), 1, 0) )
 pathroc$model = 'HGD & LGD'
+ggsave('plots/auc/dysplasia_auc.png', plot=roc.plot(pathroc, title='Dysplasia (HGD & LGD)'), width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 hgdpathroc$model = 'HGD Only'
+ggsave('plots/auc/hgd_auc.png', plot=roc.plot(hgdpathroc, title='HGD Only'), width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 pROC::coords(rocEndoM, 'best')
-#roc.plot(rocEndoM)
 rocEndoM$model = 'Per Endoscopy Max'
+ggsave('plots/auc/maxendo_auc.png', plot=roc.plot(rocEndoM, title='Max per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 pROC::coords(rocEndoA, 'best')
-#roc.plot(rocEndoA)
 rocEndoA$model = 'Per Endoscopy Avg'
+ggsave('plots/auc/avgendo_auc.png', plot=roc.plot(rocEndoA, title='Avg per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 pROC::coords(rocLastEndo, 'best')
-#roc.plot(rocLastEndo)
 rocLastEndo$model = 'Last endoscopy'
+ggsave('plots/auc/lastendo_auc.png', plot=roc.plot(rocLastEndo, title='Last endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 pROC::coords(rocFirstEndo, 'best')
-#roc.plot(rocFirstEndo)
 rocFirstEndo$model = 'First endoscopy'
+ggsave('plots/auc/firstendo_auc.png', plot=roc.plot(rocFirstEndo, title='First Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 
-ptendo = do.call(rbind, pg.samp) %>% group_by(Patient, Hospital.Research.ID, Status) %>% summarise( 'max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(Patient))
+#ptendo = do.call(rbind, pg.samp) %>% group_by(Patient, Hospital.Research.ID, Status) %>% summarise('max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(Patient))
+
+ptendo = do.call(rbind, lapply(pg.samp, function(pt) {
+ cbind(pt[1,c('Patient','Hospital.Research.ID', 'Status')],  'mean'=mean(subset(pt, PID != pt[nrow(pt), 'PID'])$Prediction), 'max'=max(subset(pt, PID != pt[nrow(pt), 'PID'])$Prediction))
+}))
+
+ptendo = ptendo[-which(is.na(ptendo$mean)),]
 
@@
 pROC::coords(rocPtM, 'best')
-#roc.plot(rocPtM)
 rocPtM$model = 'Per Patient Max'
-
-rocPtA = pROC::roc(Status ~ avg, data=ptendo, ci=T, of='thresholds')
+ggsave('plots/auc/maxpt_auc.png', plot=roc.plot(rocPtM, title='Max per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+
+rocPtA = pROC::roc(Status ~ mean, data=ptendo, ci=T, of='thresholds')
 pROC::coords(rocPtA, 'best')
@@
 rocPtA$model = 'Per Patient Avg'
+ggsave('plots/auc/avgpt_auc.png', plot=roc.plot(rocPtA, title='Avg per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 
+roc$model = "CN Model"
+
 m = rbind.data.frame(
@@
 
-ggsave('plots/compare_auc.png', plot=comp, width=6, height=8, units="in", limitsize=F, dpi = 300)
+ggsave('plots/path_p53_cn_auc.png', plot=comp, width=6, height=8, units="in", limitsize=F, dpi = 300)
 
@@
 mm = melt(subset(m, Measure == 'AUC'), measure.vars='value')
-ggplot( arrange(mm, CI.min), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=model)) + 
-  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
-  geom_text(aes(label=paste(round(value,2)*100,'%',sep=''), y=1), position=dodge, color='white')  +
+p = ggplot( arrange(mm, CI.min), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=model)) + 
+  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='grey39' ) + 
+  geom_text(aes(y=CI.min*100, label=paste(round(value,2)*100,'%',sep='')), position=dodge, vjust=3)  +
   scale_fill_manual(values=cols, name='') + plot.theme + 
-  theme(legend.position = 'none', axis.text.x = element_text(angle=45, hjust=1)) + labs(x='', y='AUC', title='ROC Comparison')
-
-multi.roc.plot(list(roc, rocEndoM, rocEndoA, rocPtM, rocPtA,rocFirstEndo,rocLastEndo), 'ROC Comparisons', palette='Dark2')
+  theme(legend.position='none', axis.text.x=element_text(angle=45, hjust=1)) + labs(x='', y='AUC', title='ROC Comparison') 
+p
+
+ggsave('plots/auc/multi_auc.png', plot=p, height=6, width=5, units='in',dpi=300)
+
+#multi.roc.plot(list(roc, rocEndoM, rocEndoA, rocPtM, rocPtA,rocFirstEndo,rocLastEndo), 'ROC Comparisons', palette='Dark2')
 ```
@@
 transform.by.time<-function(pt, info) {
-  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction', 'OR')] %>% group_by(PID, Samplename) %>% mutate(
+  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction', 'OR', 'Sample.Mean','Sample.SD','Sample.Var')] %>% group_by(PID, Samplename) %>% mutate(
   'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12)
@@
 back = do.call(rbind.data.frame,lapply(pg.samp, transform.by.time, info=patient.info))
-back = merge(back, patient.info[,c('Samplename','p53.Status')])
+back = merge(back, patient.info[,c('Samplename','p53.Status', 'Total.Reads')])
+back$Pathology = factor(back$Pathology, levels=c('BE','ID','LGD','HGD','IMC'), ordered = T)
 
@@
 back$Pathology = factor(back$Pathology, levels=c('BE','ID','LGD','HGD', 'IMC'), ordered = T)
+
+ggplot(back, aes(Pathology, Prediction, group=Pathology, fill=Pathology)) + facet_grid(~Status) + 
+  geom_jitter(width=0.2) + geom_boxplot() + plot.theme  + theme(legend.position = 'none')
 
@@
 
-ggsave(filename='plots/pred_calib.png', plot=pcal, width=6, height=6, units='in', dpi=300)
-```
-
-```{r probhist, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
+ggsave(filename='plots/pred_calib.png', plot=pcal, width=4, height=4, units='in', dpi=300)
+```
+
+```{r probhist, echo=F, warning=F, message=F, fig.width=5, fig.height=5}
 ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') +
@@
 
-```{r indP, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
+```{r indP, echo=F, warning=F, message=F, fig.width=5, fig.height=5}
 back = merge(back, pred.confidence[c('quant','Risk')], by.x='quants',by.y='quant')
@@
 
-ggplot(back, aes(OR, Prediction, color=Prediction)) + geom_point(show.legend = F) + 
-  scale_color_gradientn(colors = myPal,  name='')  + plot.theme
+#ggplot(back, aes(OR, Prediction, color=Prediction)) + geom_point(show.legend = F) + scale_color_gradientn(colors = myPal,  name='')  + plot.theme
 ```
@@
 ```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=20}
-
 min.theme = theme(text=element_text(size=8), panel.background=element_blank(), strip.background=element_rect(fill="grey97"), strip.text.y = element_text(size=6),
@@
 
-  grid.arrange(g)
+  #grid.arrange(g)
   return(list('grob'=g, 'legend'=legend))
@@
 
-#p1 = mtn.spatial(subset(globalCNdf, Patient == 48 & seg.type != 'arms'), back, ogjLimits, 'Patient 48 (P)')
 p2 = mtn.spatial(subset(globalCNdf, Patient == 20 & seg.type != 'arms'), back, ogjLimits,'Patient 20 (P)' )
-#p3 = mtn.spatial(subset(globalCNdf, Patient == 60 & seg.type != 'arms'), back, ogjLimits,'Patient 60 (P)' )
-
-#p4 = mtn.spatial(subset(globalCNdf, Patient == 3 & seg.type != 'arms'), back,ogjLimits,'Patient 3 (NP)' )
-#p5 = mtn.spatial(subset(globalCNdf, Patient == 15 & seg.type != 'arms'), back,ogjLimits,'Patient 15 (NP)' )
 p6 = mtn.spatial(subset(globalCNdf, Patient == 51 & seg.type != 'arms'), back,ogjLimits,'Patient 51 (NP)' )
+
+# for (i in 1:nrow(sum.patient.data)) {
+#   print(i)
+#   pt = sum.patient.data[i, 'Patient']
+#   status = sum.patient.data[i, 'Status']
+# 
+#   pms = mtn.spatial(subset(globalCNdf, Patient == pt & seg.type != 'arms'), back, ogjLimits, paste('Patient', pt, '(', status, ')' ))
+#   
+#   fl = paste('patient_', pt, '.png', sep='')
+#   ggsave(paste('plots/spatial/', fl, sep=''), plot=grid.arrange(pms$grob), width=14, height=6, units='in', dpi=300)
+# }
 
@@
   plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12)) 
-
+rp
 ggsave(filename='plots/pred_endo.png', plot=rp, width=9, height=7, units='in', dpi=300)
@@
 #me = endos %>% group_by(Patient) %>% summarise('max.endo' = max(i), 'months' = min(months.before.final))
-```
-
-```{r, echo=F, fig.height=8, fig.width=6, eval=F}
-## Progressor predictions per patient
-#plot.endo( subset(endos, Patient %in% subset(me, months <= -72)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(72/12, "years or more follow-up"))
-```
-
-```{r, echo=F, fig.height=8, fig.width=6, eval=F}
-# plot.endo( subset(endos, Patient %in% subset(me, months <= -48 & months > -72)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(48/12, "-",72/12,  "years or more follow-up"))
-# plot.endo( subset(endos, Patient %in% subset(me, months <= -24 & months > -48)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(24/12, "-",48/12,  "years or more follow-up"))
-#plot.endo( subset(endos, Patient %in% subset(me, months > -48)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(1, "-",48/12,  "years or more follow-up"))
-```
-
-```{r, echo=F, fig.height=12, fig.width=8, eval=F}
-## Non-progressor predictions per patient
-# nonp.back = lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID],transform.pt.by.time)
-# 
-# nonp.endos = do.call(rbind, lapply(1:max(sapply(nonp.back,nrow)), function(i) {
-#   df = do.call(rbind, lapply(nonp.back, function(x) {
-#     if (i > nrow(x)) return(NA)
-#     cbind(x[i, c('Endoscopy.Year','pred','samples', 'months.before.final', 'max.path')], i)
-#   }))
-#   rownames(df) = subset(sum.patient.data, Hospital.Research.ID %in% rownames(df))$Patient
-# 
-#   df$Patient = rownames(df)
-# 
-#   return(df)
-# }))
-# nonp.endos = nonp.endos[complete.cases(nonp.endos),]
-# #nonp.endos$i = nonp.endos$i-1
-# 
-# nonp.endos$pred.ratio = round(nonp.endos$pred/nonp.endos$samples, 2)
-# nonp.endos = transform(nonp.endos, Patient=reorder(Patient, -months.before.final) ) 
-# nonp.endos$months.before.final = nonp.endos$months.before.final*-1
-# 
-# me = nonp.endos %>% group_by(Patient) %>% summarise('max.endo' = max(i), 'months' = min(months.before.final))
-
-#plot.endo(subset(nonp.endos, Patient %in% subset(nonp.endos, pred.ratio > 0)$Patient), bin=T) + labs(title='Non-progressors with progressive predictions')
 ```
@@
 ```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=15 }
-grid.arrange(
+ga = grid.arrange(
   pred.tiles(subset(back, Patient %in% c(13,20,34,56,60,48) & Status == 'P'), ncol=1, cols=myPal, probs=T) + labs(title='Progressors') + theme(legend.position = 'none'),
@@
 ncol=2)
+ga
+
+ggsave('plots/ex_pt_heatmap.png', plot=ga, width=6, height=15, units='in', dpi=300)
 
@@
   }
-  summary.t[,c('Pathology',levels(df$labels))]
+  summary.t[,c('Pathology',levels(df$Risk))]
 }
@@
 p1 = p1 + theme(legend.position = 'none')
-
 p2 = pred.tiles(subset(back, Status == 'P'), probs=T, cols=myPal,nrow=2) + labs(title='Progressors') + theme(legend.position = 'none')
@@
 ```{r, echo=T}
-
 m = data.frame(matrix(nrow=nrow(chh), ncol=0))
@@
 g17 = g17 + theme(text=element_text(size=11), strip.text = element_text(size=11))
-ggsave(filename='plots/chr17.png', plot=g17, width=10, height=5, units='in', dpi=300)
+ggsave(filename='plots/hazards/chr17.png', plot=g17, width=10, height=5, units='in', dpi=300)
 
@@
 
-for (chr in unique(chh$chr)) 
-  print(plot.cn.chr(subset(globalCNdf, seg.type != 'arms') , chrom=chr, chr.lengths, genes=genes, haz=chh))
+for (chr in unique(chh$chr)) {  
+  p = plot.cn.chr(subset(globalCNdf, seg.type != 'arms') , chrom=chr, chr.lengths, genes=genes, haz=chh)
+  f = paste('chr',chr,'.png',sep='')
+  ggsave(paste('plots/hazards/',f,sep=''), plot=p, width=10, height=5, units='in', dpi=300)
+  print(p)
+}
 
@@
 
-globalCN.plot + #geom_rect(data=annoG, aes(xmin=start, xmax=end, ymin=ymin,ymax=ymax), color='red') +
+p = globalCN.plot + #geom_rect(data=annoG, aes(xmin=start, xmax=end, ymin=ymin,ymax=ymax), color='red') +
   geom_point(data=annoG, aes(x=start, y=0), size=1.5, color='red') +
-  geom_text_repel(data=annoG, aes(x=start, y=ymax/2, label=Gene.Symbol), color='red', size=3 )
+  geom_text_repel(data=annoG, aes(x=start, y=ymax/2, label=Gene.Symbol), color='red', size=1.5 ) + theme(legend.position = 'none')
+
+ggsave('plots/global_gene_ann.png', plot=p, width=10, height=5, units='in', dpi=300)
 
@@
 mp = model.performance(performance.at.1se, coefs) 
-mp$plot + labs(x='Models', title='Model performance for subsets of genomic regions') + 
+mp$plot = mp$plot + labs(x='Models', title='Model performance for subsets of genomic regions') + 
   scale_colour_manual(values=c('grey39', 'grey39')) + theme(axis.text.x = element_text(angle=45,hjust=1)) +
@@
   geom_hline(yintercept=mean(mp$performance$mean)-sd(mp$performance$mean), color='red', alpha=0.5)
-
-#mp$performance
+mp$plot
+ggsave('plots/performance/performance_subsets.png', plot=mp$plot, height=5, width=9, units='in',dpi=300)
 
@@
 ```{r ts, eval=F}
-
 colnames(ccgenes)
@@
 ```{r val_setup, echo=F, message=F, warning=F, include=F}
-tileFile = paste('tile_patients_All_',as.character(tile.w),'.Rdata',sep='')
-tile.files = sapply(tileFile, function(f) grep(f, analysis.files, value=T,fixed=T))
-
-fauxlabel = rep(0, length(validation.patient.info$Samplename) )
-names(fauxlabel) = validation.patient.info$Samplename
-validation.df = set.up.data(tile.files, validation.patient.info$Samplename, fauxlabel)
-dim(validation.df)
-
-if (includeDemo) {
-  ## Add in demographics
-  validation.df = add.demo.tocv(validation.patient.info, validation.df)
-}
+validation.patient.info =  subset(all.patient.info, Set == 'Test')
+validation.patient.info = arrange(validation.patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
+
+file = paste(cache.dir, 'model_data.Rdata', sep='/')
+if (!file.exists(file))
+  stop(paste("Missing data file", file))
+load(file, verbose=T)
+
+cleaned = list.files(path=paste(data, 'Cleaned', sep='/'), pattern='tiled', full.names=T, recursive=T)
+cleaned = grep(paste(unique(validation.patient.info$Hospital.Research.ID), collapse='|'), cleaned, value=T)
+
+raw = list.files(path=paste(data, 'Cleaned', sep='/'), pattern='raw', full.names=T, recursive=T)
+raw = grep(paste(unique(validation.patient.info$Hospital.Research.ID), collapse='|'), raw, value=T)
+
+# Get the SD, mean, and variance(?) per sample
+v.sampleVar = do.call(rbind, lapply(raw, function(f) {
+  t = read.table(f, sep='\t', header=T)
+  sampleCols = grep('chr|arm|start|end|probes', colnames(t), invert = T)
+  x = do.call(rbind, lapply(sampleCols, function(i){
+    cbind('Sample.Mean'=mean(t[,i]),
+          'Sample.SD'=sd(t[,i]),
+          'Sample.Var'=var(t[,i]))
+  }))
+  rownames(x) = colnames(t)[sampleCols]
+  return(x)
+}))
+
+arms = grep('arms', cleaned, value=T)
+segs = grep('arms', cleaned, invert=T, value=T)
+
+length(segs) == length(arms)
+
+# Load segment files
+tiled.segs = do.call(rbind, lapply(segs, function(f) {
+  load.segment.matrix(f)
+}))
+dim(tiled.segs)
+
+tiled.arms = do.call(rbind, lapply(arms, function(f) {
+  load.segment.matrix(f)
+}))
+
+length(z.mean) == ncol(tiled.segs)
+
+for (i in 1:ncol(tiled.segs)) 
+  tiled.segs[,i] = unit.var(tiled.segs[,i], z.mean[i], z.sd[i])
+
+for (i in 1:ncol(tiled.arms)) 
+  tiled.arms[,i] = unit.var(tiled.arms[,i], z.arms.mean[i], z.arms.sd[i])
+
+dim(tiled.arms)
+dim(tiled.segs)
+
+cx.score = unit.var(score.cx(tiled.segs,1), mn.cx, sd.cx)
+
+validation.df = subtract.arms(tiled.segs, tiled.arms)
+validation.df = cbind(validation.df, 'cx'=cx.score)
+
+fauxlabel = rep(0, nrow(validation.df) )
+names(fauxlabel) = intersect(validation.patient.info$Samplename, rownames(validation.df))
 
@@
 ```{r validation_globalplot, message=F, warning=F, echo=F, fig.height=8, fig.width=8}
-
-pander(table(validation.patient.info[c('Status','p53.Status')]))
-
 valCNdf = t(validation.df[,grep('^\\d+:',colnames(validation.df))])
@@
 
-cn.mtn.plot(subset(valCNdf, seg.type != 'arms'), labeller(Status=n.status, chr=label_value)) + theme(legend.position = 'none')
-
-df = rbind(cbind(valCNdf, 'set'=rep('Val', nrow(valCNdf))), cbind(globalCNdf, 'set'=rep('Train', nrow(globalCNdf)) ) )
+p = cn.mtn.plot(subset(valCNdf, seg.type != 'arms'), labeller(Status=n.status, chr=label_value)) + theme(legend.position = 'none') + labs(title='Validation: All samples')
+p
+ggsave('plots/validation_mtn.png', plot=p, width=10, height=5, units='in', dpi=300)
 
@@
 
-vpd = lapply(validation.patient.data, function(pt) {
-  info = pt$info
+vpd = lapply(unique(validation.patient.info$Hospital.Research.ID), function(pt) {
+  info = subset(validation.patient.info, Hospital.Research.ID == pt)
+  info$Prediction = NA
+  info$OR = NA
   info$PID = unlist(lapply(info$Path.ID, function(x) unlist(strsplit(x, 'B'))[1]))
+
+  info = cbind(info, v.sampleVar[info$Samplename,])
+
   return(info)
 })
-
-samples = intersect(rownames(pred), validation.patient.info$Samplename)
-validation.patient.info$Prediciton = NA
-validation.patient.info[which(validation.patient.info$Samplename %in% samples), 'Prediction'] = pred[samples,]
+names(vpd) = unique(validation.patient.info$Hospital.Research.ID)
+
+# for (s in rownames(pred)) {
+#   vpd[which(vpd$Samplename == s), 'Prediction'] = pred[s,]
+#   vpd[which(vpd$Samplename == s), 'OR'] = or[s,]
+# }
 
@@
 back.v$Pathology = as.character(back.v$Pathology)
-#back.v$Pathology[which(back.v$Pathology %in% c('HGD','IMC'))] = 'IMC/HGD'
+
 back.v$Pathology = factor(back.v$Pathology, levels=c('BE','ID','LGD','HGD','IMC'), ordered = T)
@@
 
-```{r echo=F, message=F, warning=F, fig.height=8, fig.width=10}
+```{r echo=F, message=F, warning=F, fig.height=5, fig.width=8}
 back.v$quants = with(back.v, cut(Prediction, breaks=cuts))
 back.v = merge(back.v, pred.confidence[c('quant','perc','Risk')], by.x='quants', by.y='quant')
-back.v$labels = factor(back.v$Risk, levels=c('High','Moderate','Low'))
+back.v$Risk = factor(back.v$Risk, levels=c('High','Moderate','Low'))
 back.v$nl = factor(back.v$nl, ordered = T)
@@
 hgdimc = subset(riskPath, Pathology %in% c('HGD','IMC')) %>% group_by(Risk) %>% summarize (
-  Freq=sum(Freq),  nPath=sum(nPath), ratio=sum(Freq)/sum(nPath)
-)
+  Freq=sum(Freq),  nPath=sum(nPath), ratio=sum(Freq)/sum(nPath))
+
 riskPath = riskPath[-which(riskPath$Pathology %in% c('HGD','IMC')),] 
@@
 
-```{r, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
+```{r, echo=F, warning=F, message=F, fig.width=4, fig.height=3}
 cuts = seq(0,1,0.1)
@@
 roc$model = 'Discovery'
-multi.roc.plot(list(roc, vroc), title='ROC', palette='Dark2') 
+multi.roc.plot(list(roc, vroc), title='ROC', palette='Dark2') + theme(text=element_text(size=12))
+#ggsave('plots/auc_dv.png', plot=p, width=5, height=4, units='in', dpi=300, scale=2)
 ```
@@
 
-```{r, echo=F, warning=F, message=F, eval=F}
-ggplot(back.v, aes(Prediction)) + geom_histogram(aes(fill=..x..), bins=20, show.legend = F) +
-  scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
-  geom_vline(xintercept=pred.confidence$r2, linetype='longdash') +
-#  geom_label(data=(pred.confidence %>% rowwise %>% mutate( 'mid'=sum(r1,r2)/2 )), aes(x=mid, y=11, label=paste(round(perc,2),conf,sep=''),color=conf ), show.legend=F) + 
-  scale_color_manual(values=c('grey','black')) + plot.theme + labs(title='Validation predictions', y='n Samples', x='Probability') 
-```
-
 # Clinical Story
@@
 p = p + theme(legend.position = 'none')
-ggsave('plots/vig_34.png', plot=p, width=6, height=6,dpi=300)
+ggsave('plots/vignette/vig_34.png', plot=p, width=6, height=6,dpi=300)
 
 p = vig.plot(subset(back, Patient == 56), subset(patient.info, Patient == 56),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vig_56.png', plot=p, width=6, height=6,dpi=300)
+ggsave('plots/vignette/vig_56.png', plot=p, width=6, height=6,dpi=300)
 
 p = vig.plot(subset(back, Patient == 20), subset(patient.info, Patient == 20),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vig_20.png', plot=p, width=6, height=6,dpi=300)
+ggsave('plots/vignette/vig_20.png', plot=p, width=6, height=6,dpi=300)
 
 p = vig.plot(subset(back, Patient == 86), subset(patient.info, Patient == 86),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vig_86.png', plot=p, width=6, height=6,dpi=300)
+ggsave('plots/vignette/vig_86.png', plot=p, width=6, height=6,dpi=300)
 
@@
 p = vig.plot(subset(back, Patient == 3), subset(patient.info, Patient == 3),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vig_3.png', plot=p, width=6, height=6,dpi=300)
+ggsave('plots/vignette/vig_3.png', plot=p, width=6, height=6,dpi=300)
 
 p = vig.plot(df=subset(back, Patient == 23),info= subset(patient.info, Patient == 23),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vig_23.png', plot=p, width=6, height=6,dpi=300)
+ggsave('plots/vignette/vig_23.png', plot=p, width=6, height=6,dpi=300)
 
 p = vig.plot(subset(back, Patient == 33), subset(patient.info, Patient == 33),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vig_33.png', plot=p, width=6, height=6,dpi=300)
+ggsave('plots/vignette/vig_33.png', plot=p, width=6, height=6,dpi=300)
 
 p = vig.plot(subset(back, Patient == 49), subset(patient.info, Patient == 49),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vig_49.png', plot=p, width=6, height=6,dpi=300)
+ggsave('plots/vignette/vig_49.png', plot=p, width=6, height=6,dpi=300)
 
 p = vig.plot(subset(back, Patient == 51), subset(patient.info, Patient == 51),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vig_51.png', plot=p, width=6, height=6,dpi=300)
+ggsave('plots/vignette/vig_51.png', plot=p, width=6, height=6,dpi=300)
 
 p = vig.plot(subset(back, Patient == 6), subset(patient.info, Patient == 6),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vig_6.png', plot=p, width=6, height=6,dpi=300)
+ggsave('plots/vignette/vig_6.png', plot=p, width=6, height=6,dpi=300)
 
 p = vig.plot(subset(back, Patient == 54), subset(patient.info, Patient == 54),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vig_54.png', plot=p, width=6, height=6,dpi=300)
+ggsave('plots/vignette/vig_54.png', plot=p, width=6, height=6,dpi=300)
 

### 2018-02-20 Patient_Report.Rmd
Updating the pipeline to make it more generic

--- 2017-07-19 Patient Report.Rmd
@@
 library(RColorBrewer)
-library(dplyr)
 library(pander)
@@
 library(gridExtra)
-library(tibble)
+#library(tibble)
+library(dplyr)
 
 source('lib/load_patient_metadata.R')
+
+get.var.regions <- function(segvals, valueCols, sd=0.08) {
+  if (is.null(valueCols)) {
+    warning('No sample columns indicated')
+    return(NULL)
+  }
+#  if (length(sample.cols) > 1) {
+  sds = apply(segvals[,valueCols],1,sd)
+  means = apply(segvals[,valueCols],1,mean)
+  CofV = sds/means
+
+  variable.region.indices = which(sds >= sd)
+  segvals[variable.region.indices,] 
+#}
+
+}
 
@@
 
-if (length(list.files(plot.dir)) <= 0)
-  stop(paste("No analysis files found in", plot.dir ))
+cleaned = list.files(paste(data, 'Cleaned', sep='/'), pattern='tiled', full.names=T, recursive=T)
+rawseg = list.files(paste(data, 'Cleaned', sep='/'), pattern='raw', full.names=T, recursive=T)
+normseg = list.files(paste(data, 'Cleaned', sep='/'), pattern='norm', full.names=T, recursive=T)
 
@@
 
-#head(patient.info)
-
 patient.info = arrange(patient.info, Status, Patient, Endoscopy.Year, Pathology)
@@
 
+patient.data = list()
+for(hr in sum.patient.data$Hospital.Research.ID) {
+  info = subset(patient.info, Hospital.Research.ID == hr)
+  info = arrange(info, Endoscopy.Year, Pathology)
+
+  segvals = read.table( grep(hr, rawseg, value=T), sep='\t', header=T)
+  if (nrow(info) > 1)
+    segvals = cbind(segvals[,c(1:5)], segvals[,intersect(info$Samplename, colnames(segvals))])
+
+  patient.data[[hr]][['info']] = info
+  patient.data[[hr]][['seg.vals']] = segvals
+  patient.data[[hr]][['norm.seg.vals']] = read.table( grep(hr, normseg, value=T), sep='\t', header=T)
+
+  if (nrow(info) > 1) # Variable regions are called relative to each other so cannot do it on a single sample
+    patient.data[[hr]][['variable.regions']] = get.var.regions(segvals, c(6:ncol(segvals)))  
+}
+
 #patient.info = subset(patient.info, Patient %in% sum.patient.data$Patient & Batch.Name %in% levels(patient.info$Batch.Name)[1:5])
@@
 #pander(arrange(patient.data, Status), justify='left', caption='Patient information')
+
+example.pt = subset(sum.patient.data, Status == 'P')$Hospital.Research.ID[12]
+example.pt.id = subset(sum.patient.data, Status == 'P')$PatientID[12]
 ```
 
-# Description of '`r dataset`' data set
+```{r echo=F, eval=F}
 
-```{r cors, echo=F, message=T, warning=F, fig.height=6, fig.width=6, fig.align='left', include=F}
-## Cluster the samples by coverage then pull out regions
+tiled = do.call(rbind, lapply(cleaned, function(f) {
+  data.table::fread(f, sep='\t', header=T)
+}))
 
-## Thresholds
-gamma2 = 250
-min.probes=67 # equiv to 1MB
-sd.threshold = 0.08  # cov by numbers of probes
-gain.threshold = 1.1; loss.threshold = 0.9
-
-cachefile = paste(data, '/Analysis/', dataset, '_patients.Rdata', sep='')
-if ( file.exists(cachefile) ) {
-  load(cachefile, verbose=T)
-} else {
-  patient.data = list(); lowprobes = c()
-
-  for (patient.name in sum.patient.data$Hospital.Research.ID  )  {
-    message(patient.name)
-    patient.plot.dir = paste(plot.dir, patient.name, sep='/')
-
-    if (!file.exists(paste(patient.plot.dir, "/",patient.name,"_segmentedCoverage_fitted_gamma",gamma2,".txt",sep=""))) {
-      warning(paste("Missing segmented file for", patient.name))
-      next
-    }
-
-    segvals = read.table(paste(patient.plot.dir, "/",patient.name,"_segmentedCoverage_fitted_gamma",gamma2,".txt",sep=""),sep="\t",stringsAsFactors=F,header=T)
-    head(segvals)
-
-    missing = c( setdiff(colnames(segvals)[-(1:5)], subset(patient.info, Hospital.Research.ID == patient.name)$Samplename),
-                 setdiff(subset(patient.info, Hospital.Research.ID == patient.name)$Samplename, colnames(segvals)[-(1:5)]) )
-
-    mr = with(removed.samples, which(Hospital.Research.ID == patient.name & Samplename %in% missing))
-    if (length(mr) > 0) {
-      missing = missing[-which(missing %in% removed.samples[mr,]$Samplename)]
-      segvals = segvals[,-which(colnames(segvals) %in% removed.samples[mr,]$Samplename) ]
-    }
-
-    if (length(missing) > 0) {
-      warning(paste("Missing", length(missing), "sample(s) of", length(subset(patient.info, Patient == patient.name)$Samplename), "in patient", patient.name))
-      #next
-    }
-
-    probes = length(which(segvals$n.probes < min.probes))
-    message( paste(probes, ' probes (', round(probes/nrow(segvals),2)*100, '%)',' below the minimum probe count (',min.probes,')', sep=''))
-    if (probes/nrow(segvals) >= 0.75) lowprobes = c(lowprobes, patient.name)
-
-    segvals = segvals[segvals$n.probes >= min.probes,]
-    dim(segvals)
-
-    # ORDER segvals by endoscopy year etc
-    cur.pt =  arrange( subset(patient.info, Hospital.Research.ID == patient.name), Endoscopy.Year, Pathology)
-    segvals = segvals[, c(colnames(segvals[1:5]), intersect(colnames(segvals), cur.pt$Samplename))]
-
-    if (nrow(segvals) < 100) {
-      warning(paste(patient.name, 'has fewer than 100 genomic regions with segmented values:', nrow(segvals)))
-      #next
-    }
-
-    patient.data[[patient.name]][['seg.vals']] = segvals
-    patient.data[[patient.name]][['info']] = cur.pt
-
-    sample.cols = grep('^D\\d+_D\\d+', colnames(segvals))
-
-    if (length(sample.cols) > 0) {
-      if (length(sample.cols) > 2) {  
-        HC = hclust(dist(t(segvals[,sample.cols])))
-        gg1 = ggdendrogram(HC, rotate=T) + labs(title = paste(patient.name,": fitted segment coverage"),x="")
-        #print(gg1)
-      }
-
-      # Normalize  (value-mean(value))/sd(value)
-      normalised.segvals = segvals[,sample.cols]
-      if (length(sample.cols) > 1) {  
-        for(c in 1:nrow(normalised.segvals)) 
-          normalised.segvals[c,] = (normalised.segvals[c,]-mean(unlist(normalised.segvals[c,])))/sd(unlist(normalised.segvals[c,]))
-        patient.data[[patient.name]][['norm.seg.vals']] = normalised.segvals
-
-        HC = hclust(dist(t(normalised.segvals)))
-        gg2 = ggdendrogram(HC, rotate=T) + labs(title=paste(patient.name,": normalised fitted segment coverage", x=""))
-        #print(gg2)
-
-        no.samples = ncol(normalised.segvals)
-        no.rows = ceiling(no.samples/3)
-
-        gg3 = ggplot( melt(normalised.segvals), aes(x=value) ) + 
-          facet_wrap(~variable,scales = "free_x") + 
-          geom_histogram( fill=I("blue"),col=I("lightblue"),binwidth=0.3) + labs(title=paste(patient.name,'normalised segment coverage'), x='', y='')
-      }
-
-      # Variable Regions can only be called relative to each sample so if there's only one sample we can't make a call
-      if (length(sample.cols) > 1) {
-        sds = apply(segvals[,sample.cols],1,sd)
-        means = apply(segvals[,sample.cols],1,mean)
-        CofV = sds/means
-
-        variable.region.indices = which(sds >= sd.threshold & segvals$n.probes >= min.probes)
-        variable.regions = segvals[variable.region.indices, intersect(patient.data[[patient.name]]$info$Samplename, colnames(segvals))] 
-        patient.data[[patient.name]][['variable.regions']] = cbind(segvals[variable.region.indices,c(1:5)],variable.regions )
-      }
-
-      patient = arrange(subset(patient.info, Hospital.Research.ID == patient.name),  Endoscopy.Year, Pathology, Samplename)
-    }
-  }
-save(patient.data, file=cachefile)
-}
-print(paste("Missing patients:", paste(setdiff(sum.patient.data$Hospital.Research.ID, names(patient.data)), collapse=', ')))
-
-sum.patient.data = subset(sum.patient.data, Hospital.Research.ID %in% names(patient.data))
-table(sum.patient.data$Status)
-
-## Check that all of the samples are there
-patient.info = arrange(patient.info, Status, Patient, Hospital.Research.ID, Endoscopy.Year,Samplename)
-msgs = sapply(names(patient.data), function(pt) {
-  missing = which(!patient.data[[pt]]$info$Samplename %in% colnames(patient.data[[pt]]$seg.vals))
-  if (length(missing) > 0) {
-    msg = paste(pt,": Missing samples - ", paste(patient.data[[pt]]$info$Samplename[missing], collapse=', ') , sep='')
-    msg = paste(msg, "\nFrom batch: ", paste(unique(patient.data[[pt]]$info$Batch.Name[missing]), collapse=', '))
-    warning(msg)
-  }
-})
-for (n in msgs) {
-  if (!is.null(n))
-    print(n)
-}
-
-example.pt = which(with(sum.patient.data, Status == 'P' & total.samples > 2))[5]
-key =  sum.patient.data[example.pt, 'Hospital.Research.ID' ]
-example.pt.id = sum.patient.data[example.pt, 'Patient']
-## Note the number of probes per patient/sample
 ```
-\pagebreak
 
@@
 ```{r}
-head(patient.data[[ key ]]$seg.vals)
+
+head(patient.data[[ example.pt ]]$seg.vals)
 ```
@@
 ```{r echo=F, warning=F, message=F, fig.height=10, fig.width=10}
-ggplot( melt(patient.data[[key]]$seg.vals[,-(1:5)]), aes(x=value) ) + 
+ggplot( melt(patient.data[[example.pt]]$seg.vals[,-(1:5)]), aes(x=value) ) + 
         facet_wrap(~variable,scales = "free_x") + 
@@
 ```{r}
-head(patient.data[[key]]$norm.seg.vals)
+head(patient.data[[example.pt]]$norm.seg.vals)
 ```
@@
 ```{r echo=F, warning=F, message=F, fig.height=10, fig.width=10}
-ggplot( melt(patient.data[[key]]$norm.seg.vals), aes(x=value) ) + 
+ggplot( melt(patient.data[[example.pt]]$norm.seg.vals), aes(x=value) ) + 
         facet_wrap(~variable,scales = "free_x") + 
@@
 ```{r general, warning=F, echo=F, message=F}
+hgdSamp = subset(patient.info, Pathology %in% c('HGD','IMC'))$Samplename
+
 gen = do.call(rbind, lapply(patient.data, function(df) {
@@
 
-ggplot(sum.patient.data) + geom_boxplot( aes(y=median.seg.len, x=Status, group=Status, fill=Status) ) +
+ggplot(sum.patient.data) + geom_boxplot( aes(y=median.seg.len, x=Status, group=Status, fill=Status), show.legend = F ) +
   labs(title="Median segment length by patient status")
+
 ```
@@
 
-var.seg.lengths = lapply(patient.data, function(df) {
-  cbind('chr'=df$variable.regions$chrom, 'start'=df$variable.regions$start.pos, 'end'=df$variable.regions$end.pos, 'seg.len'=df$variable.regions$end.pos - df$variable.regions$start.pos)
-})
+var.seg.lengths = do.call(rbind, lapply(patient.data, function(df) { 
+  if (is.null(df$variable.regions) || nrow(df$variable.regions) <= 0) return(NULL)
+  vr = cbind(df$variable.regions[,c(1:5)], 'Status' = unique(df$info$Status), 'Hospital.Research.ID'=unique(df$info$Hospital.Research.ID))
+  vr %>% rowwise() %>% mutate(seg.len = end.pos-start.pos)
+}))
 
-segment.lengths = lapply(patient.data, function(df) {
-  cbind('chr'=df$seg.vals$chrom, 'seg.len'=df$seg.vals$end.pos - df$seg.vals$start.pos)
-})
+segment.lengths = do.call(rbind, lapply(patient.data, function(df) { 
+  vr = cbind(df$seg.vals[,c(1:5)], 'Status' = unique(df$info$Status))
+  vr %>% rowwise() %>% mutate(seg.len = end.pos-start.pos)
+}))
 
-sum.var = rbind(
-  'Progressors'=do.call(rbind.data.frame, var.seg.lengths[prog.pt]) %>% summarise( 'n.segments' = length(seg.len), 'median.length'=median(seg.len), 'mean.length'=mean(seg.len), 'sd'=sd(seg.len)),
-  'Non-Progressors'=do.call(rbind.data.frame, var.seg.lengths[np.pt]) %>% summarise( 'n.segments' = length(seg.len), 'median.length'=median(seg.len), 'mean.length'=mean(seg.len), 'sd'=sd(seg.len)))
-
+sum.var = ddply(var.seg.lengths, .(Status), summarise, 'n.segments' = length(seg.len), 'median.length'=median(seg.len), 'mean.length'=mean(seg.len), 'sd'=sd(seg.len), 'max'=max(seg.len))
 pander(sum.var, caption='Variable regions between progressors and non-progressors')
 
-sum.seg = rbind(
-  'Progressors'=do.call(rbind.data.frame, segment.lengths[prog.pt]) %>% summarise( 'n.segments' = length(seg.len), 'median.length'=median(seg.len), 'mean.length'=mean(seg.len), 'sd'=sd(seg.len)),
-  'Non-Progressors'=do.call(rbind.data.frame, segment.lengths[np.pt]) %>% summarise( 'n.segments' = length(seg.len), 'median.length'=median(seg.len), 'mean.length'=mean(seg.len), 'sd'=sd(seg.len)))
-
+sum.seg = ddply(segment.lengths, .(Status), summarise, 'n.segments' = length(seg.len), 'median.length'=median(seg.len), 'mean.length'=mean(seg.len), 'sd'=sd(seg.len), 'max'=max(seg.len))
 pander(sum.var, caption='Segment stats between progressors and non-progressors')
@@
 
-```{r echo=F, message=F, warning=F}
+## Excluding HGD
 
-df = rbind(cbind(do.call(rbind.data.frame, var.seg.lengths[prog.pt]), 'status'='Progressor'), cbind(do.call(rbind.data.frame, var.seg.lengths[np.pt]), 'status'='Non-Progressor'))
-lenWT = wilcox.test(subset(df, status == 'Progressor')$seg.len, subset(df, status == 'Non-Progressor')$seg.len)
+```{r, echo-F}
+var.seg.lengthsnohgd = do.call(rbind, lapply(patient.data, function(df) { 
+  samples = subset(df$info, Pathology %in% c('BE','ID','LGD'))$Samplename
+  cols = which(colnames(df$seg.vals) %in% samples)
 
-ggplot(df, aes(status, seg.len)) + geom_jitter(aes(color=status), alpha=0.5, show.legend = F, width=.2) +
-  geom_boxplot(aes(group=status, fill=status), show.legend=F, outlier.colour = NA) + 
-  geom_label(data=subset(df, status == 'Progressor'), aes(status, max(seg.len)-2*sd(seg.len), label=paste('median\n',scientific(median(seg.len),3),sep='') )) +
-  geom_label(data=subset(df, status == 'Non-Progressor'), aes(status, max(seg.len), label=paste('median\n',scientific(median(seg.len),3),sep=''))) +
-  labs(y='Variable segment length', x='', title='Variable regions', caption=paste('Wilcox p-value', signif(lenWT$p.value, digits=3) ))
+  if (length(cols) > 1)
+    vr = get.var.regions(df$seg.vals, cols)
 
-ggplot(df, aes(status, seg.len)) + 
-  geom_jitter(aes(color=status), show.legend=F) + geom_violin(aes(group=status, fill=status), show.legend=F) +
-  geom_label(data=subset(df, status == 'Progressor'), aes(status, max(seg.len)-2*sd(seg.len),label=paste('median\n',scientific( median(seg.len) , 3),sep='') ))  +
-    geom_label(data=subset(df, status == 'Non-Progressor'), aes(status, max(seg.len), label=paste('median\n',scientific( median(seg.len) , 3),sep=''))) +
-    labs(y='Variable segment length', x='', title='Variable regions', caption=paste('Wilcox p-value', signif(lenWT$p.value, digits=3) ))
+  if (!exists('vr') || nrow(vr) <= 0) return(NULL)
+  vr = cbind(vr[,c(1:5)], 'Status' = unique(df$info$Status), 'Hospital.Research.ID'=unique(df$info$Hospital.Research.ID))
+  vr %>% rowwise() %>% mutate(seg.len = end.pos-start.pos)
+}))
+
+sum.varnohgd = ddply(var.seg.lengthsnohgd, .(Status), summarise, 'n.segments' = length(seg.len), 'median.length'=median(seg.len), 'mean.length'=mean(seg.len), 'sd'=sd(seg.len), 'max'=max(seg.len))
+
+pander(sum.varnohgd, caption='Variable regions between progressors and non-progressors NO HGD')
 
@@
 
-```{r, echo=F, warning=F, message=F}
-df = rbind(cbind.data.frame('n'=sapply(var.seg.lengths[prog.pt], nrow), 'status'='Progressor'), cbind.data.frame('n'=sapply(var.seg.lengths[np.pt], nrow), 'status'='Non-Progressor'))
-nWT = wilcox.test(subset(df, status == 'Progressor')$n, subset(df, status == 'Non-Progressor')$n)
+M-W for variable segment length in progressors with/without HGD
+```{r}
+wilcox.test(subset(var.seg.lengths, Status == 'P')$seg.len, subset(var.seg.lengthsnohgd, Status == 'P')$seg.len)
+```
 
-ggplot(df, aes(status, n, group=status)) + geom_jitter(aes(color=status), width=0.2) +
-  geom_boxplot(aes(fill=status), position = 'dodge', show.legend = F, outlier.colour = NA) +
-  geom_label(data=subset(df, status == 'Progressor'), 
-             aes(status, max(n)-2*sd(n), label=paste('median\n',sprintf("%.1f",median(n)),sep=''), fill=status ), color='white' ) +
-  geom_label(data=subset(df, status == 'Non-Progressor'), 
-             aes(status, max(n)+2*sd(n), label=paste('median\n',sprintf("%.1f",median(n)),sep=''), fill=status), color='white') +
-  labs(y='Number of segments', title="Total number of segments", caption=paste('Wilcox p-value', signif(nWT$p.value, digits=3)) ) +
-  theme(legend.position = 'none')
+And non-progressors
+```{r}
+wilcox.test(subset(var.seg.lengths, Status == 'NP')$seg.len, subset(var.seg.lengthsnohgd, Status == 'NP')$seg.len)
+```
 
-ggplot(df, aes(status, n, group=status)) + geom_jitter(aes(color=status), width=0.2) +
-  geom_violin(aes(fill=status, group=status)) +
-  geom_label(data=subset(df, status == 'Progressor'), 
-             aes(status, max(n)-2*sd(n), label=paste('median\n',sprintf("%.1f",median(n)),sep=''), fill=status ), color='white' ) +
-  geom_label(data=subset(df, status == 'Non-Progressor'), 
-             aes(status, max(n)+2*sd(n), label=paste('median\n',sprintf("%.1f",median(n)),sep=''), fill=status), color='white') +
-  labs(y='Number of segments', title="Total number of segments", caption=paste('Wilcox p-value', signif(nWT$p.value, digits=3)) ) +
-  theme(legend.position = 'none')
+```{r echo=F, message=F, warning=F}
+lenWT = wilcox.test(subset(var.seg.lengths, Status == 'P')$seg.len, subset(var.seg.lengths, Status == 'NP')$seg.len)
+ggplot() + geom_jitter(data=var.seg.lengths, aes(Status, seg.len, color=Status), alpha=0.5, show.legend = F, width=.2) +
+  geom_boxplot(data=var.seg.lengths, aes(Status, seg.len, group=Status, fill=Status), show.legend=F, outlier.colour = NA) + 
+  geom_label(data=sum.var, aes(Status, max, label=paste('median\n',scientific(median.length,3),sep='') )) +
+  labs(y='Variable segment length', x='', title='Variable regions', subtitle=paste('Wilcox p-value', signif(lenWT$p.value, digits=3) ))
 
+lenWT = wilcox.test(subset(var.seg.lengthsnohgd, Status == 'P')$seg.len, subset(var.seg.lengthsnohgd, Status == 'NP')$seg.len)
+ggplot() + geom_jitter(data=var.seg.lengthsnohgd, aes(Status, seg.len, color=Status), alpha=0.5, show.legend = F, width=.2) +
+  geom_boxplot(data=var.seg.lengths, aes(Status, seg.len, group=Status, fill=Status), show.legend=F, outlier.colour = NA) + 
+  geom_label(data=sum.varnohgd, aes(Status, max, label=paste('median\n',scientific(median.length,3),sep='') )) +
+  labs(y='Variable segment length', x='', title='Variable regions, NO HGD', subtitle=paste('Wilcox p-value', signif(lenWT$p.value, digits=3) ))
+
+vs = rbind(cbind(subset(var.seg.lengths, Status == 'P'), 'HGD'=T),
+      cbind(subset(var.seg.lengthsnohgd, Status == 'P'), 'HGD'=F))
+sum.vs = ddply(vs, .(HGD), summarise, max = max(seg.len), median.length=median(seg.len), sd=sd(seg.len), seg.len=0 )
+
+lenWT = wilcox.test(subset(vs, HGD)$seg.len, subset(vs, !HGD)$seg.len)
+ggplot(data=vs, aes(HGD, seg.len)) + 
+  geom_jitter(aes(color=HGD), alpha=0.5, show.legend = F, width=.2) +
+  geom_boxplot(aes(group=HGD, fill=HGD), show.legend=F, outlier.colour = NA) + 
+  geom_label(data=sum.vs, aes(HGD, max, label=paste('median\n',scientific(median.length,3),sep='') )) +
+  labs(y='Variable segment length', x='', title='Variable regions, Excluding HGD/IMC', subtitle=paste('Wilcox p-value', signif(lenWT$p.value, digits=3) ))
 ```
@@
 ```{r echo=F, message=F, warning=F, fig.height=8, fig.width=20}
-n.var.regions = sapply(var.seg.lengths, nrow)
+n.var.regions = nrow(var.seg.lengths)
 
@@
 
-sum.patient.data$Num.Var.Regions = as.vector(sapply(var.seg.lengths[sum.patient.data$Patient], nrow))
+#sum.patient.data$Num.Var.Regions = as.vector(sapply(var.seg.lengths[sum.patient.data$Patient], nrow))
 
@@
 
-var.seg.lengths = lapply(var.seg.lengths, function(x) merge(x, chr.info, by='chr') )
+var.seg.lengths = merge(var.seg.lengths, chr.info, by.x='chrom',by.y='chr') 
+var.seg.lengthsnohgd = merge(var.seg.lengthsnohgd, chr.info, by.x='chrom',by.y='chr')
 
 # All segments
-segment.lengths = lapply(segment.lengths, function(x) merge(x, chr.info, by='chr') )
+segment.lengths = merge(segment.lengths, chr.info, by.x='chrom', by.y='chr')
 
 # Genome coverage
-genome.coverage = sapply(segment.lengths, function(x) {
-  sum(as.numeric(x[['seg.len']]))/sum(as.numeric(chr.info$length))
-})
+segment.lengths = segment.lengths %>% rowwise() %>% mutate(coverage=seg.len/length)
 
 # No difference in coverage
-covtt = wilcox.test(genome.coverage[prog.pt], genome.coverage[np.pt])
+covtt = wilcox.test(subset(segment.lengths, Status == 'P')$coverage,subset(segment.lengths, Status == 'NP')$coverage)
 
-var.seg.lengths = lapply(var.seg.lengths, function(x) cbind(x, 'chr.ratio'=x[['seg.len']]/x[['length']]))
+var.seg.lengths = var.seg.lengths %>% rowwise() %>% mutate(coverage=seg.len/length)
+var.seg.lengthsnohgd = var.seg.lengthsnohgd %>% rowwise() %>% mutate(coverage=seg.len/length)
 
-# Segment lengths, variable regions
-vsl = do.call(rbind.data.frame, lapply(patient.data, function(df) {
-  if (is.null(df$variable.regions) || nrow(df$variable.regions) <= 0) return(NULL)
-
-  mc = grep('^D', colnames(df$variable.regions))
+getCN <- function(df, path=c('BE','ID','LGD','HGD','IMC')) {
+  if (!is.null(df$variable.regions) && nrow(df$variable.regions) > 0) {
+    samples = subset(df$info, Pathology %in% path)$Samplename
+    mc = intersect(colnames(df$variable.regions), samples)
 
@@
 
-    cbind.data.frame( 'chr'=df$variable.regions$chrom, 'start'=df$variable.regions$start.pos, 'end'=df$variable.regions$end.pos, meanCN, 'status'= df$info$Status[1])
-  }))
+   return(cbind.data.frame( 'chr'=df$variable.regions$chrom, 'start'=df$variable.regions$start.pos, 'end'=df$variable.regions$end.pos, meanCN, 'status'= df$info$Status[1]))
+  } 
+return(NULL)
+}
+# Segment lengths, variable regions
+vsl = do.call(rbind.data.frame, lapply(patient.data, getCN, path=c('BE','ID','LGD')))
 vsl = vsl[-which(vsl$meanCN == 0),]
@@
   scale_fill_gradient(low = "green3", high = "red3", name='') + 
-  labs(title="Variable region gains & losses", x='Chromosomes', y='Mean adjusted segmentation value') +
+  labs(title="Variable region gains & losses", subtitle='Excluding HGD/IMC', x='Chromosomes', y='Mean adjusted segmentation value') +
+  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank(), legend.position='bottom')
+
+vsl = do.call(rbind.data.frame, lapply(patient.data, getCN))
+vsl = vsl[-which(vsl$meanCN == 0),]
+
+vsl$meanCN = vsl$meanCN - mean(vsl$meanCN)
+
+vslm = melt(vsl, id.vars=c('chr','start','end','status'))
+
+chr.info = subset(chr.info, chr %in% c(1:22))
+chr.info$chr = ordered(chr.info$chr, levels=c(1:22))
+vslm$chr =  ordered(vslm$chr, levels=c(1:22))
+vslm = merge(chr.info, vslm, by='chr')
+vslm = vslm %>% rowwise() %>% mutate('genome.start'=genome.length+start, 'genome.end'=genome.length+end)
+
+vslm$status = ifelse(vslm$status == 'P', 'Progressor', 'Non Progressor')
+
+ggplot(vslm, aes(x=length)) + 
+  facet_grid(status~chr) +
+  geom_rect(data=vslm, aes(xmin=start, xmax=end, ymin=0, ymax=value, fill=value)) +
+  scale_fill_gradient(low = "green3", high = "red3", name='') + 
+  labs(title="Variable region gains & losses", subtitle='All samples', x='Chromosomes', y='Mean adjusted segmentation value') +
   theme(axis.text.x=element_blank(),axis.ticks.x=element_blank(), legend.position='bottom')
@@
 # Variable region length by ratio of chromosome
-chr.ratios = lapply(var.seg.lengths, function(x) x[['seg.len']]/x[['length']])
+melted = melt(var.seg.lengths, measure.vars=c('coverage'))
+chr.ratio.tt = t.test(subset(var.seg.lengths, Status == 'P')$coverage, subset(var.seg.lengths, Status == 'NP')$coverage)
 
-melted = melt(chr.ratios)
-melted = merge(sum.patient.data[,c('Hospital.Research.ID', 'Status')], melted, by.x='Hospital.Research.ID', by.y='L1')
+sm = ddply(melted, .(Status), summarise, median.cov=median(value), max=max(value), value=0)
 
-chr.ratio.tt = t.test(unlist(chr.ratios[prog.pt]), unlist(chr.ratios[np.pt]))
+ggplot(melted, aes(Status, value)) + geom_jitter(aes(color=Status), width=0.2) + geom_violin(alpha=0.3) + 
+  geom_label(data=sm, aes(Status, max, label=paste('Median coverage\n', scientific(median.cov, 3), sep=''))) +
+  labs(title="Variable region coverage, All samples",subtitle=paste("p-value", signif(chr.ratio.tt$p.value, digits=3)), y='Variable region Segment Length/Chr length')
 
-ggplot(melted, aes(Status, value, fill=Status)) + geom_jitter(aes(color=Status), width=0.2) + geom_boxplot(outlier.colour = NA) + 
-  geom_label(data=subset(melted, Status == 'P'), aes(x=Status, y=max(value)-sd(value), label=scientific(median(value),3) ), color='white', show.legend = F ) +
-  geom_label(data=subset(melted, Status == 'NP'), aes(x=Status, y=max(value)-sd(value), label=scientific(median(value),3) ), color='white' , show.legend = F) +
-  labs(title="Variable region length",subtitle=paste("p-value", signif(chr.ratio.tt$p.value, digits=3)), y='Segment Length/Chr length')
+melted = melt(var.seg.lengthsnohgd, measure.vars=c('coverage'))
+chr.ratio.tt = t.test(subset(var.seg.lengthsnohgd, Status == 'P')$coverage, subset(var.seg.lengthsnohgd, Status == 'NP')$coverage)
 
-ggplot(melted, aes(Status, value, fill=Status)) + geom_jitter(aes(color=Status), width=0.2) + geom_violin() + 
-  geom_label(data=subset(melted, Status == 'P'), aes(x=Status, y=max(value)-sd(value), label=scientific(median(value),3) ), color='white' , show.legend = F) +
-  geom_label(data=subset(melted, Status == 'NP'), aes(x=Status, y=max(value)-sd(value), label=scientific(median(value),3) ), color='white' , show.legend = F) +
-  labs(title="Variable region length",subtitle=paste("p-value", signif(chr.ratio.tt$p.value, digits=3)), y='Segment Length/Chr length')
-
-var.seg.lengths =  lapply(var.seg.lengths, function(df) {
-   ddply(df, .(chr, length), summarise, seg.lengths = sum(seg.len), chr.ratio = sum(seg.len)/length[1])
-})
+sm = ddply(melted, .(Status), summarise, median.cov=median(value), max=max(value), value=0)
+ggplot(melted, aes(Status, value)) + geom_jitter(aes(color=Status), width=0.2) + geom_violin(alpha=0.3) + 
+  geom_label(data=sm, aes(Status, max, label=paste('Median coverage\n', scientific(median.cov, 3), sep=''))) +
+  labs(title="Variable region coverage, Excluding HGD/IMC",subtitle=paste("p-value", signif(chr.ratio.tt$p.value, digits=3)), y='Variable region Segment Length/Chr length')
 
 # Genome coverage of the variable regions
-pcov = median(sapply(var.seg.lengths[prog.pt], function(x) 
-  sum(as.numeric(x[['seg.lengths']]))/sum(as.numeric(x[['length']])) ), na.rm=T)
 
-npcov = median(sapply(var.seg.lengths[np.pt], function(x) 
-  sum(as.numeric(x[['seg.lengths']]))/sum(as.numeric(x[['length']])) ), na.rm=T)
+cov = ddply(var.seg.lengths, .(Status), summarise, coverage=sum(as.numeric(seg.len))/sum(as.numeric(length)))
+
+# pcov = median(sapply(var.seg.lengths[prog.pt], function(x) 
+#   sum(as.numeric(x[['seg.lengths']]))/sum(as.numeric(x[['length']])) ), na.rm=T)
+# 
+# npcov = median(sapply(var.seg.lengths[np.pt], function(x) 
+#   sum(as.numeric(x[['seg.lengths']]))/sum(as.numeric(x[['length']])) ), na.rm=T)
 
@@
 
-long.chr.tt = t.test(unlist(lapply(var.seg.lengths[prog.pt], get.chrs, chrs=c(1:12))), unlist(lapply(var.seg.lengths[np.pt], get.chrs, chrs=c(1:12))))
-short.chr.tt = t.test(unlist(lapply(var.seg.lengths[prog.pt], get.chrs, chrs=c(13:22))), unlist(lapply(var.seg.lengths[np.pt], get.chrs, chrs=c(13:22))))
+long.chr.tt = t.test(subset(var.seg.lengths, Status == 'P' & chrom %in% c(1:12))$coverage,subset(var.seg.lengths, Status == 'NP' & chrom %in% c(1:12))$coverage)
+short.chr.tt = t.test(subset(var.seg.lengths, Status == 'P' & chrom %in% c(13:22))$coverage,subset(var.seg.lengths, Status == 'NP' & chrom %in% c(13:22))$coverage)
 
@@
 #ftL = fisher.test(df)
-```
 
-So it appears that adjusting for chromosome length (looking at the ratio of the chromosome covered by segments rather than the raw bp) a significant difference in the coverage of the genome by CN segments that are variable between progressors and non: t-test p-value `r signif(chr.ratio.tt$p.value, 3)`.
+#So it appears that adjusting for chromosome length (looking at the ratio of the chromosome covered by segments rather than the raw bp) a significant difference in the coverage of the genome by CN segments that are variable between progressors and non: t-test p-value `r signif(chr.ratio.tt$p.value, 3)`.
 Overall bp coverage of the sWGS is `r round(mean(genome.coverage), 2)*100`%  (P `r round(mean(genome.coverage[prog.pt]),2)`, NP `r round(mean(genome.coverage[np.pt]), 2)`).
 
-Variable regions cover `r round(pcov, 2)*100`% of the the progressors genome, and `r round(npcov, 2)*100`% of non-progressor genomes.
+#Variable regions cover `r round(pcov, 2)*100`% of the the progressors genome, and `r round(npcov, 2)*100`% of non-progressor genomes.
+```
 
@@
 
-ft = fisher.test((df %>% group_by(status) %>% summarise( 'gains'=sum(gain), 'losses'=sum(loss)))[,c('gains','losses')])
+ft = fisher.test(ddply(df, .(status), summarise,'gains'=sum(gain), 'losses'=sum(loss)) [,c('gains','losses')])
 
@@
 ## How do Cancer Gene Consensus genes map to the variable regions?
+
+But only in the variable regions EXCLUDING HGD/IMC
 
@@
 
+#var.seg.lengths
+
+# NO HGD
 rm(var.regions)
 for (patient.name in names(patient.data)) {
-  regions = patient.data[[patient.name]]$variable.regions[,c('chrom','arm','start.pos','end.pos')]
+  regions = subset(var.seg.lengthsnohgd, Hospital.Research.ID == patient.name)
+    #patient.data[[patient.name]]$variable.regions[,c('chrom','arm','start.pos','end.pos')]
   if (is.null(regions) || nrow(regions) <= 0) next
@@
 
-ggplot(m,  aes(y=Gene.Symbols, x=CN, fill=brks)) + geom_tile() + facet_grid(.~variable) + 
-    scale_fill_manual(values = myP, 'Patient Counts') + labs(title="All available COSMIC genes", x='', y='')
+ggplot(m,  aes(y=Gene.Symbols, x=CN, fill=brks)) + geom_tile(color='white') + facet_grid(.~variable) + 
+    scale_fill_manual(values = myP, 'Patient Counts') + labs(title="All available COSMIC genes", subtitle='Excluding HGD/IMC samples', x='', y='')
 
-m2 = melt(gene.p %>% group_by(Gene.Symbols) %>% summarise( 'Progressors' = sum(Progressors), 'Non.Progressors' = sum(Non.Progressors)), id.vars='Gene.Symbols')
+m2 = melt( ddply(gene.p, .(Gene.Symbols), summarise, 'Progressors'=sum(Progressors), 'Non.Progressors'=sum(Non.Progressors)), id.vars='Gene.Symbols')
 
@@
 
-ggplot(m2,  aes(y=Gene.Symbols, x=variable, fill=brks)) + geom_tile() + #facet_grid(.~variable) + 
-    scale_fill_manual(values = myP, 'Patient Counts') + labs(title="All available COSMIC genes", x='', y='')
+ggplot(m2,  aes(y=Gene.Symbols, x=variable, fill=brks)) + geom_tile(color='white') + #facet_grid(.~variable) + 
+    scale_fill_manual(values = myP, 'Patient Counts') + labs(title="All available COSMIC genes", subtitle='Excluding HGD/IMC samples', x='', y='')
 
@@
 ```{r echo=F, warning=F, message=F, fig.height=12, fig.width=8}
-gc = as.data.frame(gene.p %>% group_by(Gene.Symbols) %>% summarise(total = sum(Progressors) + sum(Non.Progressors)))
+gc = as.data.frame(ddply(gene.p, .(Gene.Symbols), summarise, total = sum(Progressors) + sum(Non.Progressors)))
 gc = transform(gc, Gene.Symbols = reorder(Gene.Symbols, total))
@@
     scale_fill_manual(values=myP, 'Patient Counts') + 
-    labs(title="Top 50 genes by number of patients", x='', y='')
+    labs(title="Top 50 genes by number of patients", subtitle='Excluding HGD/IMC samples', x='', y='')
 
@@
 ```{r, hgd, echo=F, message=F, warning=F}
+gain.threshold = 1.1; loss.threshold = 0.9
 
@@
 ```
-
-```{r, warning=F, eval=F}
-#This is just looking at the distribution of values per sample. The average ploidy per sample is around 1.  Currently we're calling variable regions based on this.
-ks = list()
-for (pt in names(patient.data)) {
-  print(pt)
-  cols = intersect(arrange(patient.data[[pt]]$info, Pathology, Endoscopy.Year)$Samplename, grep('^D\\d+', colnames(patient.data[[pt]]$seg.vals), value=T))
-  if (length(cols) <= 1) next
-
-  segvals = patient.data[[pt]]$seg.vals[,cols]
-
-  sdx = round(apply(segvals, 2, sd), 3)
-  se = round(apply(segvals, 2, function(x) sd(x)/sqrt(length(x))), 3)
-
- gg = ggplot(melt(segvals)) + 
-        facet_wrap(~variable,scales = "free_x") + 
-        geom_histogram(bins=10, fill='blue', color='grey',aes(value, ..density..)) +
-        geom_density(aes(value+sdx/2), color='red', size=1) +
-        geom_density(aes(value-sdx/2), color='green', size=1) +
-        labs(title=paste(pt, '  (',patient.data[[pt]]$info$Status[1],')', sep=''))
-  print(gg)   
-  #plots[[pt]] = gg
-
-  ks[[pt]] = sapply(segvals, function(x) ks.test(x+sdx, x-sdx)$statistic  )
-  print(pt)
-  print(colSums(abs(round(patient.data[[pt]]$seg.vals[,-(1:5)], 1)-1)))
-}
-
-```

### 2018-02-20 PredictingProgression/lib/data_func.R

@@ -1,0 +1,228 @@
+require(GenomicRanges) 
+require(dplyr)
+require(data.table)
+require(tibble)
+
+
+prep.matrix<-function(dt) {
+  output = list()
+  
+  means = apply(dt,2, mean)
+  sd = apply(dt,2, sd)
+  
+  output[['z.mean']] = means
+  output[['z.sd']] = sd
+  
+  dt = apply(dt, 2, unit.var)
+  
+  output[['matrix']] = dt
+  return(output)
+}
+
+load.segment.matrix<-function(segFile) {
+  dt = read.table(segFile, sep='\t', header=T)
+  
+  if (length(which(grepl('chr|start|end', colnames(dt)))) < 3)
+    stop("Cannot load without chr, start, end columns")
+  
+  chrCol = grep('chr',colnames(dt))
+  startCol = grep('start',colnames(dt))
+  endCol = grep('end',colnames(dt))
+  
+  rows = paste(dt[,chrCol], ':', dt[,startCol], '-', dt[,endCol], sep='')
+  
+  dt = as.matrix(dt[,-c(chrCol, startCol, endCol)])
+  rownames(dt) = rows
+  return( t(dt) )
+}
+
+
+unit.var <- function(x, mean=NULL, sd=NULL) {
+  if (is.null(mean) | is.null(sd)) {
+    warning("No mean or sd provided.")
+    if (length(x) == 1 | sd(x, na.rm=T) == 0) {
+      warning("Unit normalization can't be performed with less than 2 samples or SD was 0")
+      return(x)
+    } else {
+    return((x-mean(x,na.rm=T))/sd(x,na.rm=T) )
+    }
+  }
+  
+  uv = (x-mean)/sd 
+  if (sd == 0) uv = 0
+  
+  return(uv)
+}
+
+score.cx <- function(df, MARGIN) {  
+  cx = apply(df, MARGIN, function(x) length(which(x >= sd(x)*2 | x <= -sd(x)*2)))
+  return(cx)
+}
+
+.tile.genome<-function(tile.w=5e6, chr.info=NULL, incGender=F) {
+  
+  if (is.null(chr.info))
+    chr.info = read.table('data/chr_hg19.txt', sep='\t', header=T, stringsAsFactors=F)
+  
+  chr.info$chrom = sub('^chr', '', chr.info$chrom)
+  
+  chrs = c(1:22)
+  if (incGender) chrs = c(1:22, 'X','Y')
[…]

### 2018-02-20 PredictingProgression/server.R

@@ -1,0 +1,131 @@
+#
+# This is the server logic of a Shiny web application. You can run the 
+# application by clicking 'Run App' above.
+#
+# Find out more about building applications with Shiny here:
+# 
+#    http://shiny.rstudio.com/
+
+library(shiny)
+library(ggplot2)
+library(glmnet)
+
+source('lib/data_func.R')
+
+chr.info = read.table('data/chr_hg19.txt', sep='\t', header=T)
+
+
+load('data/model_data.Rdata', verbose=T)
+rm(dysplasia.df, labels)
+
+load('data/all.pt.alpha.Rdata', verbose=F)
+select.alpha = '0.9'
+fitV = models[[select.alpha]]
+lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']
+rm(plots, coefs, dysplasia.df, cv.patient, labels, models, performance.at.1se)
+
+
+load('data/loo.Rdata', verbose=T)
+rm(plots, coefs, nzcoefs, fits)
+
+pg.samp = do.call(rbind, pg.samp)
+pg.samp$Hospital.Research.ID = NULL
+
+#hist(pg.samp$OR)
+
+riskPal = rev(RColorBrewer::brewer.pal(11, 'RdYlBu'))
+unadjRR = ggplot(pg.samp, aes(OR)) + geom_histogram(aes(fill=..x..), bins=10, show.legend = T) +
+  scale_fill_gradientn(colors = riskPal,  name='RR') + 
+  labs(y='n Samples', x='Relative Risk', title='Unadjusted relative risk') + theme_light(base_size = 14)
+
+
+
+# p = ggplot(df, aes(x=chr.length)) + facet_grid(Endoscopy.Year+ogj~chr, scales='free', space='free_x', labeller = labeller(.multi_line = F)) +
+#   geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=ogj), show.legend=T) + 
+#   scale_fill_manual(values=pal, limits=limits,  name='OGJ Dist.') +
+#   labs(title=title, x='') + min.theme  
+
+
+# Define server logic required to draw a histogram
+shinyServer(function(input, output) {
+  
+  output$contents <- renderTable({
+    # input$file1 will be NULL initially. After the user selects
+    # and uploads a file, it will be a data frame with 'name',
+    # 'size', 'type', and 'datapath' columns. The 'datapath'
[…]

### 2018-02-20 PredictingProgression/ui.R

@@ -1,0 +1,38 @@
+#
+# This is the user-interface definition of a Shiny web application. You can
+# run the application by clicking 'Run App' above.
+#
+# Find out more about building applications with Shiny here:
+# 
+#    http://shiny.rstudio.com/
+#
+
+library(shiny)
+
+
+
+
+# Define UI for application that draws a histogram
+shinyUI(fluidPage(
+  
+  # Application title
+  titlePanel("Barrett's Progression Prediction"),
+  sidebarLayout(
+    sidebarPanel(
+      fileInput("file", "Choose TSV File",
+                accept = c(
+                  "text/tsv",
+                  "text/tab-separated-values,text/plain",
+                  ".tsv", ".txt")
+      ),
+      tags$hr(),
+      checkboxInput("header", "Header", TRUE)
+    ),
+    mainPanel(
+      tableOutput("contents")
+      #plotOutput("plot")
+    )
+  )
+  
+  
+))

### 2018-02-20 lib/data_func.R

@@ -1,0 +1,228 @@
+require(GenomicRanges) 
+require(dplyr)
+require(data.table)
+require(tibble)
+
+
+prep.matrix<-function(dt) {
+  output = list()
+  
+  means = apply(dt,2, mean)
+  sd = apply(dt,2, sd)
+  
+  output[['z.mean']] = means
+  output[['z.sd']] = sd
+  
+  dt = apply(dt, 2, unit.var)
+  
+  output[['matrix']] = dt
+  return(output)
+}
+
+load.segment.matrix<-function(segFile) {
+  dt = read.table(segFile, sep='\t', header=T)
+  
+  if (length(which(grepl('chr|start|end', colnames(dt)))) < 3)
+    stop("Cannot load without chr, start, end columns")
+  
+  chrCol = grep('chr',colnames(dt))
+  startCol = grep('start',colnames(dt))
+  endCol = grep('end',colnames(dt))
+  
+  rows = paste(dt[,chrCol], ':', dt[,startCol], '-', dt[,endCol], sep='')
+  
+  dt = as.matrix(dt[,-c(chrCol, startCol, endCol)])
+  rownames(dt) = rows
+  return( t(dt) )
+}
+
+
+unit.var <- function(x, mean=NULL, sd=NULL) {
+  if (is.null(mean) | is.null(sd)) {
+    warning("No mean or sd provided.")
+    if (length(x) == 1 | sd(x, na.rm=T) == 0) {
+      warning("Unit normalization can't be performed with less than 2 samples or SD was 0")
+      return(x)
+    } else {
+      return((x-mean(x,na.rm=T))/sd(x,na.rm=T) )
+    }
+  }
+  
+  uv = (x-mean)/sd 
+  if (sd == 0) uv = 0
+  
+  return(uv)
+}
+
+score.cx <- function(df, MARGIN) {  
+  cx = apply(df, MARGIN, function(x) length(which(x >= sd(x)*2 | x <= -sd(x)*2)))
+  return(cx)
+}
+
+.tile.genome<-function(tile.w=5e6, chr.info=NULL, incGender=F) {
+  
+  if (is.null(chr.info))
+    chr.info = read.table('data/chr_hg19.txt', sep='\t', header=T, stringsAsFactors=F)
+  
+  chr.info$chrom = sub('^chr', '', chr.info$chrom)
+  
+  chrs = c(1:22)
+  if (incGender) chrs = c(1:22, 'X','Y')
[…]

### 2018-02-20 processing_pipeline/run-cv2.R

@@ -1,0 +1,440 @@
+
+suppressPackageStartupMessages(library(ggplot2))
+suppressPackageStartupMessages(library(plyr))
+suppressPackageStartupMessages(library(dplyr))
+suppressPackageStartupMessages(library(gridExtra))
+suppressPackageStartupMessages(library(glmnet))
+suppressPackageStartupMessages(library(mice))
+suppressPackageStartupMessages(library(reshape2))
+suppressPackageStartupMessages(library(GenomicRanges))
+
+
+suppressPackageStartupMessages(source('../lib/load_patient_metadata.R'))
+suppressPackageStartupMessages(source('../lib/cv-pt-glm.R'))
+suppressPackageStartupMessages(source('../lib/data_func.R'))
+
+
+
+args = commandArgs(trailingOnly=TRUE)
+
+if (length(args) < 4)
+  stop("Missing required params: <data dir> <outdir> <info file dir> <all>")
+
+data = args[1]
+#data = '~/Data/Ellie/Cleaned'
+outdir = args[2]
+#outdir = '~/Data/Ellie/Analysis'
+infodir = args[3]
+#infodir = '~/Data/Ellie/QDNAseq'
+
+# Discovery or all
+allPts = ifelse (length(args) == 2, as.logical(args[2]), F)
+
+
+cache.dir = paste(outdir, '5e6_arms_disc', sep='/')
+if (allPts)
+  cache.dir = paste(outdir, '5e6_arms_all', sep='/')
+dir.create(cache.dir, recursive=T, showWarnings=F)
+
+## Hospital.Research.ID info file
+patient.file = list.files(infodir, pattern='All_patient_info.xlsx', recursive=T, full.names=T)
+demo.file = list.files(infodir, pattern='Demographics_full.xlsx', recursive=T, full.names=T)
+
+if ( length(patient.file) < 1 | length(demo.file) < 1)
+  stop("Missing files in info dir: All_patient_info.xlsx and Demographics_full.xlsx")
+
+if (allPts) {
+  patient.info = read.patient.info(patient.file, demo.file, set='All')$info
+} else {
+  patient.info = read.patient.info(patient.file, demo.file, set='Training')$info
+}
+
+patient.info = plyr::arrange(patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
[…]

### 2018-03-06 AllSamplesModel.Rmd
adding to pipeline

--- 2018-02-20 Paper.Rmd
@@
 ---
-title: "Paper"
+title: "All Samples Model"
 author: "Sarah Killcoyne"
-date: "2 October 2017"
+date: "16 Feb 2018"
 output: 
@@
 source('lib/common_plots.R')
-source('lib/data_func.R')
 
 data = '~/Data/Ellie'
-info.dir = '~/Data/Ellie/QDNAseq'
-#tile.w=c(5e06, 'arms')
-#tile.w=5e+06
-#tile.w=1e7
-#tile.w='arms'
-
-#includeDemo = F
-
-#data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
+
+cache.dir = paste(data, 'Analysis/5e6_arms_all', sep='/')
+
+data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
 #analysis.files = list.files(paste(data, 'Analysis', sep='/'), full.names=T)
@@
 ## Hospital.Research.ID info file
-
-patient.file = list.files(info.dir, pattern='All_patient_info.xlsx', recursive=T, full.names=T)
-demo.file = list.files(info.dir, pattern='Demographics_full.xlsx', recursive=T, full.names=T)
-
-if (length(patient.file) != 1 | length(demo.file) != 1)
-  stop(paste("Missing/too many patient info file(s) in", info.dir))
+patient.file = grep('All_patient_info.xlsx', data.files, value=T)
+if (length(patient.file) != 1)
+  stop(paste("Missing/too many patient info file(s) in", data))
+demo.file = grep('Demographics_full.xlsx', data.files, value=T)
 
@@
 
-patient.info = subset(all.patient.info, Set == 'Training')
-
-validation.patient.info =  subset(all.patient.info, Set == 'Test')
-validation.patient.info = arrange(validation.patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
-sum.validation.info = summarise.patient.info(validation.patient.info)
-
-patient.info = arrange(patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
+patient.info = arrange(all.patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
 sum.patient.data = summarise.patient.info(patient.info)
@@
 
-pander(sum.validation.info[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Validation patient set')
-
-all.info = rbind(sum.patient.data, sum.validation.info)
-
 ```
@@
 ```{r echo=F}
-
 pander(
-  as.data.frame(all.info %>% summarise(
+  as.data.frame(sum.patient.data %>% summarise(
   'Total Patients'=length(unique(Patient)),
@@
   'gender F:M'=paste(table(gender),collapse=':')
-))
-, caption='full cohort')
+)), caption='full cohort', justify='left')
 ```
@@
 ```{r echo=F}
-cohortSummary = as.data.frame(all.info %>% group_by(Status) %>% summarise(
+cohortSummary = as.data.frame(sum.patient.data %>% group_by(Status) %>% summarise(
   'Total Patients'=length(unique(Patient)),
@@
         table(subset(all.patient.info, Status == 'P')$Pathology)))
-
   rbind(table(subset(patient.info, Status == 'NP')$Pathology),
@@
 
-  rbind(table(subset(validation.patient.info, Status == 'NP')$Pathology),
-        table(subset(validation.patient.info, Status == 'P')$Pathology))
-
-```
-
-`r pander(t(cohortSummary), caption='Full cohort, split by status')`
+```
+
+`r pander(t(cohortSummary), caption='Full cohort, split by status', justify='left')`
 
@@
 ```{r labelsPNP, echo=F, message=F, include=T}
-cache.dir = paste(data, 'Analysis/5e6_arms_disc', sep='/')
-
 file = paste(cache.dir, 'model_data.Rdata', sep='/')
@@
 
-load(file, verbose=F)
+load(file, verbose=T)
 dim(dysplasia.df)
@@
 ```{r echo=F, warning=F, message=F, eval=T, fig.height=12, fig.width=20}
-
 chr.lengths = get.chr.lengths()
@@
 gp = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples') + theme(text=element_text(size=12))
-ggsave(filename='plots/globalCN.png', plot=gp, width=16, height=4.6, units='in', dpi=300)
+#ggsave(filename='plots/globalCN.png', plot=gp, width=16, height=4.6, units='in', dpi=300)
+gp
 
 gbe = globalBECN.plot + theme(legend.position = 'none') + labs(y='mean adj.seg.', title="Non-Dysplastic Barrett's Samples") + theme(text=element_text(size=12))
-ggsave(filename='plots/globalBECN.png', plot=gbe, width=16, height=4.6, units='in', dpi=300)
+gbe
+#ggsave(filename='plots/globalBECN.png', plot=gbe, width=16, height=4.6, units='in', dpi=300)
 
@@
 ```{r loaddata, message=F, warning=F, echo=F}
-
 file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
@@
 
-ggsave('plots/per_sample_cx.png', plot=pv + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
-
-ggsave('plots/per_sample_cx_roc.png', plot=roc.plot(rocX, title='Per sample cx') + plot.theme + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
+#ggsave('plots/per_sample_cx.png', plot=pv + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
+
+#ggsave('plots/per_sample_cx_roc.png', plot=roc.plot(rocX, title='Per sample cx') + plot.theme + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
 
@@
 
-ggsave('plots/eyebrows.png', plot=do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits')), width=10, height=10, units='in', dpi=300)
-
-```
-
-```{r echo=F, fig.height=5, fig.width=5}
-library(ggfortify)
-autoplot(models$`0.9`) + theme(legend.position = 'none') + plot.theme
-
-ggsave('plots/glm_0.9.png', plot=autoplot(models$`0.9`) + theme(legend.position = 'none') + plot.theme, width=5, height=5, units='in', dpi=300)
+#ggsave('plots/eyebrows.png', plot=do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits')), width=10, height=10, units='in', dpi=300)
+
 ```
@@
 
-ggsave('plots/performance/ridge_v_lasso.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)
+#ggsave('plots/performance/ridge_v_lasso.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)
 
@@
 }
-
 nohgd.coefs = coefs[[select.alpha]]
 #do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2))
-ggsave('plots/nohgd_eyebrow.png', plot=do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2)), height=10, width=10, units='in', dpi=300)
+#ggsave('plots/nohgd_eyebrow.png', plot=do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2)), height=10, width=10, units='in', dpi=300)
 ```
@@
 
-ggsave('plots/performance/ridge_v_lasso_nohgd.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)
+#ggsave('plots/performance/ridge_v_lasso_nohgd.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)
 
@@
                               )
+
+rm(performance.at.1se, coefs, mp)
+```
+
+### Remove LGD
+
+```{r noLGD, echo=F, warning=F, message=F, fig.height=16, fig.width=16}
+file = paste(cache.dir, 'nolgd.Rdata', sep='/')
+
+if (file.exists(file)) {
+  load(file, verbose=F)
+} else {
+  stop(paste("No model file", file, "found"))
+}
+
+#do.call(grid.arrange, c(nolgd.plots, top="No LGD/HGD/IMC samples", ncol=2))
+```
+
+```{r, echo=F, fig.width=6, fig.height=6}
+mp = model.performance(performance.at.1se, coefs)
+mp$plot = mp$plot + labs(title='No LGD/HGD/IMC samples')
+mp$plot
+
+#ggsave('plots/performance/no_lgd_perf.png', plot=mp$plot, height=5, width=5, units='in', dpi=30)
+
+classification.rate = add_row(classification.rate, 'model'='No LGD/HGD/IMC', 'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
+                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))
 
@@
     geom_label(data=melt(subset(perf, perf %in% range(perf) )), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
+  #geom_text_repel(aes(label=pt)) +
     theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')
@@
 
-ggsave('plots/performance/loo_perf_nohgd.png', plot=p, width=5, height=5, units='in', dpi=300)
+#ggsave('plots/performance/loo_perf_nohgd.png', plot=p, width=5, height=5, units='in', dpi=300)
 
 rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')
+roc.plot(rocNOHGD, 'No HGD')
 
@@
 
-rm(plots, performance.at.1se,coefs, nzcoefs, fits)
-```
-
-### Remove LGD
-
-```{r noLGD, echo=F, warning=F, message=F, fig.height=16, fig.width=16}
-file = paste(cache.dir, 'nolgd.Rdata', sep='/')
-
-if (file.exists(file)) {
-  load(file, verbose=F)
-} else {
-  stop(paste("No model file", file, "found"))
-}
-
-#do.call(grid.arrange, c(nolgd.plots, top="No LGD/HGD/IMC samples", ncol=2))
-```
-
-```{r, echo=F, fig.width=6, fig.height=6}
-mp = model.performance(performance.at.1se, coefs)
-mp$plot = mp$plot + labs(title='No LGD/HGD/IMC samples')
-mp$plot
-
-ggsave('plots/performance/no_lgd_perf.png', plot=mp$plot, height=5, width=5, units='in', dpi=30)
-
-classification.rate = add_row(classification.rate, 'model'='No LGD/HGD/IMC', 'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
-                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))
-
-rm(performance.at.1se, coefs, mp)
+rm(plots, nzcoefs, fits)
 ```
@@
 
-ggsave('plots/performance/compare_classification.png', plot=p, height=5, width=3, units='in', dpi=300 )
+#ggsave('plots/performance/compare_classification.png', plot=p, height=5, width=3, units='in', dpi=300 )
 ```
@@
 
-ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)
+#ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)
 
@@
   geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')
-
-ggsave('plots/global_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)
+p
+
+#ggsave('plots/global_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)
 
@@
 
-ggsave('plots/ndbe_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)
+#ggsave('plots/ndbe_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)
 
@@
 
-ggsave('plots/performance_loo.png', plot=p, width=5, height=5, units='in', dpi=300)
+#ggsave('plots/performance_loo.png', plot=p, width=5, height=5, units='in', dpi=300)
 
@@
 #sum.patient.data[which.max(sum.patient.data$total.samples),]
-```
-
-```{r, eval=F, echo=F, message=F}
-image(cov(dysplasia.df[names(which(labels == 1)),c(1:(ncol(dysplasia.df)-3))], use='p'), breaks=seq(-1,1,l=21), col=colorRampPalette(c('red','black','green'))(20),  useRaster = T)
-
-image(cor(dysplasia.df[names(which(labels == 0)),c(1:(ncol(dysplasia.df)-3))], use='p'), breaks=seq(-1,1,l=21), col=colorRampPalette(c('red','black','green'))(20),  useRaster = T)
-
-chr <- as.numeric(sub(":.+","",colnames(dysplasia.df[names(which(labels == 0)),c(1:(ncol(dysplasia.df)-3))])))
-c <- which(diff(chr)==1)/length(chr)
-abline(h=c, v=c, col='blue')
-
-preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Samplename','Prediction')]))
-rownames(preds) = preds$Samplename
-p <- preds[names(labels), 'Prediction']
-
-d <- dist(dysplasia.df[,c(1:(ncol(dysplasia.df)-3))])
-t <- tsne::tsne(d)
-plot(t,  bg=colorRampPalette(c('green','grey','red'))(10)[cut(p, breaks = seq(0,1,l=11))], pch=21, col=ifelse(labels==1,"black","white"))
 ```
@@
 roc$model = 'CN Model'
-roc$is = 'discovery'
+roc$is = 'all'
 
@@
 droc
-ggsave(filename='plots/discovery_roc.png', plot=droc, width=6, height=4, units='in', dpi=300)
-
-insilico.rocs[['discovery']] = roc
+#ggsave(filename='plots/discovery_roc.png', plot=droc, width=6, height=4, units='in', dpi=300)
+
+insilico.rocs[['all']] = roc
 
@@
 p = ggplot(df, aes(Specificity, Sensitivity,color=is)) + geom_line(size=1) + 
-    geom_line(data=subset(df, is == 'discovery'), color='red', size=1) +
+    geom_line(data=subset(df, is == 'all'), color='red', size=1) +
     scale_x_reverse() +
@@
 p
-ggsave(filename='plots/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)
-
-df = cbind.data.frame('specificity'=rev(roc$specificities), 'sensitivity'=rev(roc$sensitivities))
-thresholds = 
-  rbind.data.frame(
-#    est=t(round(pROC::coords(roc, "best"),2)),
-    max=t(round(pROC::coords(roc, 0.8),2)),
-    'No HGD'=t(round(pROC::coords(rocNOHGD, 0.6),2)),
-    'All samples'=t(round(pROC::coords(roc, 0.6),2))
-)
-thresholds$label=rownames(thresholds)
-
-# ggplot() + geom_line(data=df, aes(specificity, sensitivity), color='darkblue', size=1.5) + scale_x_reverse() + 
-#   geom_point(data=thresholds, aes(x=specificity,y=sensitivity,color=label), size=3,shape=8, show.legend = T) +
-#   geom_label_repel(data=thresholds, aes(x=specificity,y=sensitivity,fill=label,label=
-#     paste('Threshold ', round(threshold,2), '\n(FPR ', round((1-specificity),2)*100, '%, TPR ', round(sensitivity,2)*100,'%)\nAUC:',round(roc$auc, 2)*100, '%', sep='')),show.legend = F, color='white',nudge_x=0.05) + scale_fill_manual(values=brewer.pal(3,'Dark2')) + scale_color_manual(values=brewer.pal(3,'Dark2')) +
-#   labs(title='CN Model ROC', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme
-
-roc.plot(rocNOHGD, 'Model excluding HGD/IMC')
-pROC::coords(rocNOHGD, "best")
-pROC::coords(rocNOHGD, 0.6)
-pROC::auc(rocNOHGD, partial.auc=c(0.89,0.9), partial.auc.focus='spec', partial.auc.correct=T)
+#ggsave(filename='plots/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)
 
@@
 
-```{r pathroc, echo=F, fig.height=4, fig.width=8 }
-#sum(table(subset(patient.info, Status == 'P')$Pathology)[c('IMC','HGD')])/sum(table(subset(patient.info, Status == 'P')$Pathology))
-
+```{r pathroc, echo=F, fig.height=4, fig.width=10 }
 pathsamples = patient.info[,c('Pathology','Status')]
@@
 pathroc$model = 'HGD & LGD'
-ggsave('plots/auc/dysplasia_auc.png', plot=roc.plot(pathroc, title='Dysplasia (HGD & LGD)'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+#ggsave('plots/auc/dysplasia_auc.png', plot=roc.plot(pathroc, title='Dysplasia (HGD & LGD)'), width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 hgdpathroc$model = 'HGD Only'
-ggsave('plots/auc/hgd_auc.png', plot=roc.plot(hgdpathroc, title='HGD Only'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+#ggsave('plots/auc/hgd_auc.png', plot=roc.plot(hgdpathroc, title='HGD Only'), width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 rocEndoM$model = 'Per Endoscopy Max'
-ggsave('plots/auc/maxendo_auc.png', plot=roc.plot(rocEndoM, title='Max per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+#ggsave('plots/auc/maxendo_auc.png', plot=roc.plot(rocEndoM, title='Max per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 rocEndoA$model = 'Per Endoscopy Avg'
-ggsave('plots/auc/avgendo_auc.png', plot=roc.plot(rocEndoA, title='Avg per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+#ggsave('plots/auc/avgendo_auc.png', plot=roc.plot(rocEndoA, title='Avg per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 rocLastEndo$model = 'Last endoscopy'
-ggsave('plots/auc/lastendo_auc.png', plot=roc.plot(rocLastEndo, title='Last endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+#ggsave('plots/auc/lastendo_auc.png', plot=roc.plot(rocLastEndo, title='Last endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 rocFirstEndo$model = 'First endoscopy'
-ggsave('plots/auc/firstendo_auc.png', plot=roc.plot(rocFirstEndo, title='First Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+#ggsave('plots/auc/firstendo_auc.png', plot=roc.plot(rocFirstEndo, title='First Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 rocPtM$model = 'Per Patient Max'
-ggsave('plots/auc/maxpt_auc.png', plot=roc.plot(rocPtM, title='Max per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+#ggsave('plots/auc/maxpt_auc.png', plot=roc.plot(rocPtM, title='Max per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 rocPtA$model = 'Per Patient Avg'
-ggsave('plots/auc/avgpt_auc.png', plot=roc.plot(rocPtA, title='Avg per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+#ggsave('plots/auc/avgpt_auc.png', plot=roc.plot(rocPtA, title='Avg per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 
-# get.auc.at(p53roc)
-# get.auc.at(hgdpathroc)
-# get.auc.at(pathroc)
-# get.auc.at(roc)
-
 roc$model = "CN Model"
-
 m = rbind.data.frame(
-      get.roc.stat(p53roc),
+  #get.roc.stat(p53roc),
   get.roc.stat(hgdpathroc),
@@
   get.roc.stat(roc))
-m$model = factor(m$model, levels=c('p53 IHC','HGD Only','HGD & LGD','CN Model'), ordered = T)
-
-rocList = list(roc,pathroc,hgdpathroc,p53roc)
+m$model = factor(m$model, levels=c('HGD Only','HGD & LGD','CN Model'), ordered = T)
+
+rocList = list(roc,pathroc,hgdpathroc)
+multi.roc.plot(rocList)
 
@@
 
-ggsave('plots/path_p53_cn_auc.png', plot=comp, width=6, height=8, units="in", limitsize=F, dpi = 300)
+#ggsave('plots/path_p53_cn_auc.png', plot=comp, width=6, height=8, units="in", limitsize=F, dpi = 300)
 
@@
 
-ggsave('plots/auc/multi_auc.png', plot=p, height=6, width=5, units='in',dpi=300)
-
-#multi.roc.plot(list(roc, rocEndoM, rocEndoA, rocPtM, rocPtA,rocFirstEndo,rocLastEndo), 'ROC Comparisons', palette='Dark2')
+#ggsave('plots/auc/multi_auc.png', plot=p, height=6, width=5, units='in',dpi=300)
+
+multi.roc.plot(list(roc, rocEndoM, rocEndoA, rocPtM, rocPtA,rocFirstEndo,rocLastEndo), 'ROC Comparisons', colors=rev(cols[1:7]))
 ```
@@
 transform.by.time<-function(pt, info) {
-  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction', 'OR', 'Sample.Mean','Sample.SD','Sample.Var')] %>% group_by(PID, Samplename) %>% mutate(
+  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction', 'OR')] %>% group_by(PID, Samplename) %>% mutate(
     'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12)
@@
 back = do.call(rbind.data.frame,lapply(pg.samp, transform.by.time, info=patient.info))
-back = merge(back, patient.info[,c('Samplename','p53.Status', 'Total.Reads')])
-back$Pathology = factor(back$Pathology, levels=c('BE','ID','LGD','HGD','IMC'), ordered = T)
+back = merge(back, patient.info[,c('Samplename','p53.Status')])
 
@@
 back$Pathology = as.character(back$Pathology)
-#back$Pathology[which(back$Pathology %in% c('HGD','IMC'))] = 'IMC/HGD'
 back$Pathology = factor(back$Pathology, levels=c('BE','ID','LGD','HGD', 'IMC'), ordered = T)
-
-ggplot(back, aes(Pathology, Prediction, group=Pathology, fill=Pathology)) + facet_grid(~Status) + 
-  geom_jitter(width=0.2) + geom_boxplot() + plot.theme  + theme(legend.position = 'none')
 
@@
 
-qt = cbind(qt, back %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))#, 'se'=(sd(Prediction)/sqrt(length(Prediction)))*1.96 ))
-#ggplot(qt,aes(quants, mn, ymin=mn-se, ymax=mn+se)) + geom_point() + geom_errorbar() + labs(x='Deciles', y='Mean probability', title='CI around probabilities by decile')
-
+qt = cbind(qt, back %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))#, 
 pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 
@@
 
-ggsave(filename='plots/pred_calib.png', plot=pcal, width=4, height=4, units='in', dpi=300)
-```
-
-```{r probhist, echo=F, warning=F, message=F, fig.width=5, fig.height=5}
-ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') +
-    geom_vline(xintercept=annotation.table$max, linetype='longdash', color='grey39') +
-    geom_text(data=annotation.table %>% group_by(Risk, min,max) %>% summarise( 'samp.perc'= signif((sum(`n(NP)`, `n(P)`)/nrow(patient.info))*100, 3) ), aes(x=min+(max-min)/2, y=5, label=paste(samp.perc, '%',sep=''))) +
-    plot.theme + labs(title='Sample predictions', y='n Samples', x='Probability') + theme(legend.position = 'none')
-```
-
-```{r indP, echo=F, warning=F, message=F, fig.width=5, fig.height=5}
+#ggsave(filename='plots/pred_calib.png', plot=pcal, width=4, height=4, units='in', dpi=300)
+
 back = merge(back, pred.confidence[c('quant','Risk')], by.x='quants',by.y='quant')
@@
 
-ggplot(back, aes(1:nrow(back), Prediction, color=Risk)) + 
-  geom_point(aes(color=Risk, shape=Status)) +scale_color_manual(values=riskCols, name='P(P)') + 
-  geom_hline(data=annotation.table, aes(yintercept=max), color='grey') + plot.theme +
-  labs(x='sample', y='model probability', title='Per-sample probabilities')
+mod = subset(back, Risk == 'Moderate')
+
 ```
@@
 ```{r relrisk, echo=F, fig.width=6, fig.height=6}
-#myPal = c(RColorBrewer::brewer.pal(9, 'Blues')[9:7], 
-#          adjustcolor(grDevices::heat.colors(10)[6], blue.f=1.5),
-#          ( grDevices::heat.colors(10)[7:6]),
-#          RColorBrewer::brewer.pal(9, 'Reds')[6:9])
-
 myPal = rev(RColorBrewer::brewer.pal(11, 'RdYlBu'))
 
-unadjRR = ggplot(back, aes(OR)) + geom_histogram(aes(fill=..x..), bins=10, show.legend = F) +
+unadjRR = ggplot(back, aes(OR)) + geom_histogram(aes(fill=..x..), bins=15, show.legend = F) +
   #scale_fill_distiller(palette = 'RdYlBu', type='div',name='P(P)') +
@@
 
-ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR, width=6, height=5, units='in', dpi=300)
+#ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR, width=6, height=5, units='in', dpi=300)
 
@@
   scale_fill_gradientn(colors = myPal,  name='') + 
-  #scale_fill_distiller(palette = 'RdYlBu', type='div',name='P(P)') +
   plot.theme + labs(title='Predictions, all samples model', y='n Samples', x='Probability') 
@@
 
-ggsave(filename='plots/pred_hist.png', plot=predhist, width=5, height=5, units='in', dpi=300)
+#ggsave(filename='plots/pred_hist.png', plot=predhist, width=5, height=5, units='in', dpi=300)
 
@@
 ```
-
-### Mountain plots using spatial info
-
-```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=20}
-min.theme = theme(text=element_text(size=8), panel.background=element_blank(), strip.background=element_rect(fill="grey97"), strip.text.y = element_text(size=6),
-                   strip.text.x = element_text(size=12), axis.text = element_blank(), axis.ticks = element_blank(),
-                   axis.line=element_line(color='black'), panel.grid.major.y =element_line(color='grey95'), panel.grid.major.x = element_blank(),
-                   panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0, 'lines'), panel.spacing.y = unit(0.1, 'lines')   )
-
-mtn.spatial<-function(df, b, limits, title='', pal=NULL) {
-  if (is.null(pal))
-    pal = RColorBrewer::brewer.pal(length(limits), 'Set2')
-
-  df = merge(df, b[,c('Samplename','Endoscopy.Year','months.before.final','Block','ogj')], by.x='variable',by.y='Samplename')
-  df$ogj = factor(df$ogj, ordered = T)
-  df$ogj = factor(df$ogj, levels=rev(levels(df$ogj)), ordered = T)
-  #pal = pal[which(limits %in% levels(df$nl))]
-
-  altcolor = c(RColorBrewer::brewer.pal(9, 'Greys')[2],RColorBrewer::brewer.pal(9, 'Greys')[4])
-  fills = c()
-  for (yr in sort(unique(df$Endoscopy.Year))) {
-    ac = (which(sort(unique(df$Endoscopy.Year)) == yr) %% 2) + 1
-    fills = c(fills, rep(altcolor[ac], length(which(table(subset(df, Endoscopy.Year == yr)$ogj) > 0))))
-  }
-
-  blanklabel <- function(s) {
-    return('')
-  }
-
-  p = ggplot(df, aes(x=chr.length)) + facet_grid(Endoscopy.Year+ogj~chr, scales='free', space='free_x', labeller = labeller(.multi_line = F)) +
-      geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=ogj), show.legend=T) + 
-      scale_fill_manual(values=pal, limits=limits,  name='OGJ Dist.') +
-      labs(title=title, x='') + min.theme
-
-  legend = get.legend(p)
-  p = p + theme(legend.position = 'none')
-
-  g = ggplot_gtable(ggplot_build(p))
-  strips = grep('strip-r', g$layout$name)
-
-  for (n in 1:length(strips)) {
-    i = strips[n]
-    j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
-    g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[n]
-  }
-
-  #grid.arrange(g)
-  return(list('grob'=g, 'legend'=legend))
-  #grid::grid.draw(g)  
-  #return(g)
-}
-
-back$ogj = back$nl
-back$ogj[which(back$nl > 0 & back$nl < 3)] = 1 
-back$ogj[which(back$nl > 3 & back$nl < 7)] = 2 
-back$ogj[which(back$nl >= 7)] = 3
-
-ogjLimits = levels(factor(back$ogj, ordered = T))
-
-p2 = mtn.spatial(subset(globalCNdf, Patient == 20 & seg.type != 'arms'), back, ogjLimits,'Patient 20 (P)' )
-p6 = mtn.spatial(subset(globalCNdf, Patient == 51 & seg.type != 'arms'), back,ogjLimits,'Patient 51 (NP)' )
-
-# for (i in 1:nrow(sum.patient.data)) {
-#   print(i)
-#   pt = sum.patient.data[i, 'Patient']
-#   status = sum.patient.data[i, 'Status']
-# 
-#   pms = mtn.spatial(subset(globalCNdf, Patient == pt & seg.type != 'arms'), back, ogjLimits, paste('Patient', pt, '(', status, ')' ))
-#   
-#   fl = paste('patient_', pt, '.png', sep='')
-#   ggsave(paste('plots/spatial/', fl, sep=''), plot=grid.arrange(pms$grob), width=14, height=6, units='in', dpi=300)
-# }
-
-back$ogj = NULL
-
-```
-
-```{r, echo=F, fig.width=2, fig.height=4}
-
-grid.arrange(p2$legend)
-
-ggsave(filename='plots/spatial_mtn_pt20.png', plot=grid.arrange(p2$grob), width=14, height=6, units='in', dpi=300)
-ggsave(filename='plots/spatial_mtn_pt51.png', plot=grid.arrange(p6$grob), width=14, height=6, units='in', dpi=300)
-
-```
-
-At each quintile are the percentage of progressor samples that were predicted within that range, and the hypergeometric p-value (Fishers) for the significance within that quintile.
-Then our predictions are at best a coin flip in the middle, but highly significant at the outer edges.
-This would also push the prediction for "high" risk up to `r min(subset(pred.confidence, labels == 'High' )$r1)` while "low" risk is below `r max(subset(pred.confidence, labels == 'Low' )$r2)`.
-
-`r pander(as.data.frame(pred.confidence))`
 
@@
 
-Effectively it looks to me like my coefficients have fairly tight CI if the SE of the jackknife variance is used.
-
 ## How do the predictions look at each time point?
@@
 
-ggsave(filename='plots/pred_by_path.png', plot=rp, width=9, height=7, units='in', dpi=300)
-
-```
-
-```{r p53IHC, echo=F, warning=F, message=F, fig.height=8, fig.width=8}
-
-#p53Path$Risk[p53Path$Risk == 'High'] = 'Aberrant'
-#p53Path$Risk[p53Path$Risk == 'Low'] = 'Normal'
-
-#p53Path$p53.Status = factor(p53Path$p53.Status, levels=c('Aberrant', 'Normal'), ordered = F)
-
-p53Path$p53.Status = factor(p53Path$p53.Status, ordered = T)
-p53Path = arrange(p53Path, Pathology, p53.Status)
-col = brewer.pal(11, "PRGn")[c(10,2)]
-rp = ggplot(p53Path, aes(Pathology, ratio*100, group=Pathology, fill=p53.Status)) + geom_bar(stat='identity', position = position_stack()) +
-  scale_fill_manual(values=col, name='p53 IHC', labels=c('Normal','Aberrant')) + 
-  geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101)) +
-  geom_text(aes(label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5,color='white') +
-  plot.theme + labs(title='Samples predicted by aberrant p53 IHC', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
-rp
-ggsave(filename='plots/p53_pred.png', plot=rp, width=7, height=6, units='in', dpi=300)
+# x = back %>% group_by(Status, Pathology) %>% summarise( mean(Prediction), 'count'=length(Pathology) )
+# ggplot(x, aes(Pathology, count)) + geom_bar(stat='identity',aes(fill=`mean(Prediction)`)) + 
+#     scale_fill_gradientn(colors = myPal,  name='') + facet_grid(~Status) + plot.theme
+
+#ggsave(filename='plots/pred_by_path.png', plot=rp, width=9, height=7, units='in', dpi=300)
 
@@
 head(all.preds)
+
 ggplot(all.preds, aes(Patient, value/n.samples, group=variable, fill=variable)) + 
@@
 rp
-ggsave(filename='plots/pred_endo.png', plot=rp, width=9, height=7, units='in', dpi=300)
+#ggsave(filename='plots/pred_endo.png', plot=rp, width=9, height=7, units='in', dpi=300)
 
@@
 `r pander(as.data.frame(pred.confidence))`
-
-The first two quintiles and the last two together account for 80% of our samples.  Only 20% fall into the category "Moderate".
 
@@
 
-#pred.tiles(subset(back, Patient %in% c(6)), ncol=1, cols=myPal, probs=T, OR=F) + labs(title='') + theme(legend.position = 'none')
 #pred.tiles(subset(back, Patient %in% c(48)), ncol=1, cols=myPal, probs=T) + labs(title='') + theme(legend.position = 'none')
@@
 
-ggsave('plots/ex_pt_heatmap.png', plot=ga, width=6, height=15, units='in', dpi=300)
+#ggsave('plots/ex_pt_heatmap.png', plot=ga, width=6, height=15, units='in', dpi=300)
 
@@
 summary.t = table.path.risk(back)
-
-write.table(summary.t, quote=F, sep='\t', row.names = F)
 ```
@@
 
-```{r, echo=F, message=F, warning=F}
+```{r, echo=F, message=F, warning=F, width=26, height=8}
 back = back %>% group_by(Patient) %>% mutate( 'pred.ratio'=round(sum(Prediction)/length(Prediction),2))
@@
 
-ggsave('plots/pred_heatmap.png', plot=grid.arrange(p2,p1,nrow=2), width=52, height=16, units="in", limitsize=F, dpi = 300)
-
-ggsave('plots/npg_heatmap.png', plot=p1, width=52, height=8, units="in", limitsize=F, dpi = 300)
-ggsave('plots/pg_heatmap.png', plot=p2, width=52, height=8, units="in", limitsize=F, dpi = 300)
-
-ggsave('plots/risk_legend.png', plot=grid.arrange(riskLegend), width=2,height=4, dpi=300)
-
-ggsave('plots/risk_legend_long.png', plot=grid.arrange(riskLegendLong), width=8,height=2, dpi=300)
-
-```
-
-# Chr17
-
-## p arm
-```{r, echo=T}
-
-p11 = subset(globalCNdf, chr == 17 &  start == 1 & seg.type == 'arms')
-p11Loss = subset(p11, value < -1)
-table(p11Loss$Status)
-
-p11path = table(subset(patient.info, Samplename %in% p11Loss$variable)$Pathology)
-
-p11path['BE']/sum(p11path)
-sum(p11path[c('HGD','IMC')])/sum(p11path)
-
-table(p11Loss[c('Patient', 'Pathology')])
-
-# Samples with a loss vs all Progressor samples
-length(unique(p11Loss$variable))/length(unique(subset(globalCNdf, Status == 'Progressor')$variable))
-
-p11Loss = unique(p11Loss[,c('variable', 'Status','Hospital.Research.ID', 'Patient', 'Pathology')])
-length(unique(p11Loss$Patient))/length(subset(sum.patient.data, Status == 'P')$Patient)
-# Of those where we see loss of p arm, p53 IHC shows an "abnormal" result 33% of the time
-
-table(subset(patient.info, Samplename %in% p11Loss$variable, select='p53.Status'))/length(p11Loss$variable)
-
-(table(subset(patient.info, Samplename %in% p11Loss$variable, select=c('p53.Status', 'Pathology'))))
-
-pi = patient.info
-pi$p11Loss = as.numeric(pi$Samplename %in% p11Loss$variable)
-
-plot(table(p11=pi$p11Loss, ihc=pi$p53.Status))
-
-cor.test(pi$p11Loss, (as.numeric(pi$p53.Status)-1))
-
-# So why might there be loss but no corresponding aberrant staining?
-table(subset(pi, p11Loss == 1 & p53.Status == 0)$Pathology)
-table(subset(pi, p11Loss == 1 & p53.Status == 1)$Pathology)
-
-```
-
-## Top Features
-```{r, echo=T}
-m = data.frame(matrix(nrow=nrow(chh), ncol=0))
-
-for (i in 1:nrow(chh)) {
-  feat = chh[i,]
-
-  seg = subset(globalCNdf, chr == as.character(feat$chr) & start == feat$start & end == feat$end & seg.type != 'arms')  
-  if(feat$end-feat$start > 5e6)
-    seg = subset(globalCNdf, chr == as.character(feat$chr) & start == feat$start & end == feat$end & seg.type == 'arms')
-
-  segGL = subset(seg, value < -1)
-  if(feat$coef > 0)
-    segGL = subset(seg, value > 1)
-
-  m[i, 'n.samples'] = length(unique(segGL$variable))
-  m[i,'ratio.samples.P'] = length(unique(segGL$variable))/length(unique(subset(globalCNdf, Status == 'Progressor')$variable))
-  m[i, 'ratio.NP'] = table(segGL$Status)[1]/sum(table(segGL$Status))
-  m[i, 'ratio.P'] = table(segGL$Status)[2]/sum(table(segGL$Status))
-
-  segpath = table(subset(patient.info, Samplename %in% segGL$variable)$Pathology)
-
-  m[i,'BE'] = segpath['BE']/sum(segpath)
-  m[i,'HGD'] = sum(segpath[c('HGD','IMC')])/sum(segpath)
-}
-rownames(m) = chh$label
-
-range(m$ratio.samples.P)
-summary(m$ratio.samples.P)
-subset(m, ratio.samples.P == max(ratio.samples.P))
-subset(chh, label == rownames(subset(m, ratio.samples.P == max(ratio.samples.P))))
-subset(chh, label == rownames(subset(m, ratio.samples.P == min(ratio.samples.P))))
-#table(cdkGain[c('Patient', 'Pathology')])
-
-cor.test(chh$hazards, m$n.samples)
-
-# Samples with a gain vs all samples
-
-```
-
-## Chr 17 Predictive
-
-One way of looking at the p53 prediction (e.g. Adam Bass) is to see how predictive the p-arm loss in chr17 is.
-```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=10}
-ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
-ccgenes = ccgenes[-grep('^X|Y', ccgenes$Genome.Location),]
-#ccgenes = ccgenes[-which(ccgenes$Mutation.Types == 'T'),] # drop the translocations
-#ccgenes = subset(ccgenes, !grepl('^X|Y', Genome.Location))
-#loc = unlist(strsplit(ccgenes$Genome.Location, ':|-'))
-loc = do.call(rbind.data.frame, strsplit(ccgenes$Genome.Location, ':|-'))
-colnames(loc) = c('chr','start','end')
-loc[] = lapply(loc[], function(x) as.numeric(as.character(x)))
-
-ccgenes = cbind(ccgenes, loc)
-ccgenes = ccgenes[c('Gene.Symbol','Genome.Location', 'Role.in.Cancer', 'Mutation.Types',colnames(loc))]
-#ccgenes = subset(ccgenes, grepl('\\w+',Role.in.Cancer))
-
-genes = cbind(subset(ccgenes, grepl('TSG|oncogene', Role.in.Cancer) & Mutation.Types != 'T', select=c('Gene.Symbol','Genome.Location', colnames(loc))), 'value'=min(globalCNdf$mean.value), 'Row.names'='', 'Patient'='', 'selected'=as.factor(0))
-
-genesGR = makeGRangesFromDataFrame(genes, keep.extra.columns = T)
-chGR = makeGRangesFromDataFrame(chh, keep.extra.columns = T)
-
-ov = findOverlaps(chGR, genesGR)
-#genes = as.data.frame(genesGR[unique(subjectHits(ov))])
-#colnames(genes)[1] = c('chr')
-
-g17 = plot.cn.chr(globalCNdf, chrom='17', chr.lengths, genes=genes, haz=chh, label=labeller(chr=label_both, Status=n.status)) + labs(
-title="All Samples") + theme(legend.position = 'none')
-g17
-
-g17 = g17 + theme(text=element_text(size=11), strip.text = element_text(size=11))
-ggsave(filename='plots/hazards/chr17.png', plot=g17, width=10, height=5, units='in', dpi=300)
-
-```
-
-```{r echo=F, warning=F, message=F, fig.height=5, fig.width=10}
-chGR$Genes = ''
-for (hit in unique(queryHits(ov))) {
-  chGR[hit]$Genes = paste(genesGR[subjectHits(ov[queryHits(ov) == hit])]$Gene.Symbol, collapse=',')
-}
-
-for (chr in unique(chh$chr)) {  
-  p = plot.cn.chr(subset(globalCNdf, seg.type != 'arms') , chrom=chr, chr.lengths, genes=genes, haz=chh)
-  f = paste('chr',chr,'.png',sep='')
-  ggsave(paste('plots/hazards/',f,sep=''), plot=p, width=10, height=5, units='in', dpi=300)
-  print(p)
-}
-
-annoG = arrange(subset(ccgenes, grepl('CDKN2A$|TP53|MYC$|SMAD4|VEG|ERBB2|EGFR|SMYD|ARID1A|FHIT|RUNX1', Gene.Symbol)), chr)
-annoG = merge(annoG, chr.lengths[c('chrom','chr.length')], by.x='chr',by.y='chrom')
-annoG$ymin = round(range(globalCN.plot$data$mean.value))[1]/2
-annoG$ymax = round(range(globalCN.plot$data$mean.value))[2]/2
-
-p = globalCN.plot + #geom_rect(data=annoG, aes(xmin=start, xmax=end, ymin=ymin,ymax=ymax), color='red') +
-  geom_point(data=annoG, aes(x=start, y=0), size=1.5, color='red') +
-  geom_text_repel(data=annoG, aes(x=start, y=ymax/2, label=Gene.Symbol), color='red', size=1.5 ) + theme(legend.position = 'none')
-
-ggsave('plots/global_gene_ann.png', plot=p, width=10, height=5, units='in', dpi=300)
-
-```
-
-`r pander(genes[,c(1:3,6,7)])`
-
-```{r, echo=F, message=F, warning=F, fig.height=5, fig.width=8, eval=F}
-file = paste(cache.dir, 'chr17.predictive.Rdata', sep='/')
-if (file.exists(file)) {
-  message(paste("loading", file))
-  load(file, verbose=T)
-} else {
-  stop(paste("File missing", file))
-}
-
-names(performance.at.1se)
-
-mp = model.performance(performance.at.1se, coefs) 
-mp$plot = mp$plot + labs(x='Models', title='Model performance for subsets of genomic regions') + 
-  scale_colour_manual(values=c('grey39', 'grey39')) + theme(axis.text.x = element_text(angle=45,hjust=1)) +
-  geom_hline(yintercept=mean(mp$performance$mean)+sd(mp$performance$mean), color='red', alpha=0.5) +
-  geom_hline(yintercept=mean(mp$performance$mean)-sd(mp$performance$mean), color='red', alpha=0.5)
-mp$plot
-ggsave('plots/performance/performance_subsets.png', plot=mp$plot, height=5, width=9, units='in',dpi=300)
-
-chr.info = read.table('hg19_genes.txt', sep='\t', header=T)
-chr.info = chr.info[1:22,]
-
-chr.info$info.content = chr.info$protein_coding/chr.info$length
-
-# No correlation between information content (genes/length) and classification rate 
-cor.test(mp$performance[1:22,'mean'], chr.info$info.content)
-
-median(mp$performance$mean)
-sd(mp$performance$mean)
-mp$performance[17,'mean']
-
-```
-
-Individually none of these show very good classification rates, all hovering around 65%.
-However, more importantly is that Chr17 is only marginally better than any of the others (unless you exclude HGD) and all are below the 72% classification rate we get with all of the data
-
-# Tumour Supressors
-
-```{r ts, eval=F}
-colnames(ccgenes)
-
-tsgs = ccgenes
-nrow(tsgs)
-
-tsgs = makeGRangesFromDataFrame(tsgs[,c(colnames(loc), 'Gene.Symbol')], keep.extra.columns = T)
-tsgs = sort(tsgs)
-
-length(tsgs)
-
-# average 1 per 9MB
-avgTSGdist = mean(elementMetadata(distanceToNearest(tsgs))$distance)
-
-chh = arrange(chh, -hazards)
-features = makeGRangesFromDataFrame(chh, keep.extra.columns = T)
-
-distanceTSG = data.frame(matrix(nrow=0,ncol=3))
-for (i in 1:length(features)) {
-  f = features[i]
-  hits = nearest(f, tsgs, ignore.strand=T, select='all')
-  if (length(hits) > 0) {
-    distanceTSG = rbind(distanceTSG, cbind.data.frame(
-      'features'=i,
-      'nearestTSG'=elementMetadata(tsgs[subjectHits(hits)])$Gene.Symbol,
-      'distanceToTSG'=elementMetadata(distanceToNearest(f, tsgs, ignore.strand=T, select='all'))$distance)
-    )
-  }
-}
-
-features[unique(subset(distanceTSG, distanceToTSG == 0)$features)]
-
-# In our dataset we should see 1 TSG/onc per 10MB
-# But this is double the actual number we see
-ncol(dysplasia.df)/round(avgTSGdist/5e6)
-
-length(which(round((subset(distanceTSG, distanceToTSG > 0)$distanceToTSG/5e6)-0.5) == round((avgTSGdist/5e6)-0.5)))
-```
-
-# Validation Dataset
-
-20 Patients (10P, 10NP) set aside for validation purposes
-`r pander(table(sum.validation.info$Status), caption='Validation cohort')`
-
-```{r val_setup, echo=F, message=F, warning=F, include=F}
-validation.patient.info =  subset(all.patient.info, Set == 'Test')
-validation.patient.info = arrange(validation.patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
-
-file = paste(cache.dir, 'model_data.Rdata', sep='/')
-if (!file.exists(file))
-  stop(paste("Missing data file", file))
-load(file, verbose=T)
-
-cleaned = list.files(path=paste(data, 'Cleaned', sep='/'), pattern='tiled', full.names=T, recursive=T)
-cleaned = grep(paste(unique(validation.patient.info$Hospital.Research.ID), collapse='|'), cleaned, value=T)
-
-raw = list.files(path=paste(data, 'Cleaned', sep='/'), pattern='raw', full.names=T, recursive=T)
-raw = grep(paste(unique(validation.patient.info$Hospital.Research.ID), collapse='|'), raw, value=T)
-
-# Get the SD, mean, and variance(?) per sample
-v.sampleVar = do.call(rbind, lapply(raw, function(f) {
-  t = read.table(f, sep='\t', header=T)
-  sampleCols = grep('chr|arm|start|end|probes', colnames(t), invert = T)
-  x = do.call(rbind, lapply(sampleCols, function(i){
-    cbind('Sample.Mean'=mean(t[,i]),
-          'Sample.SD'=sd(t[,i]),
-          'Sample.Var'=var(t[,i]))
-  }))
-  rownames(x) = colnames(t)[sampleCols]
-  return(x)
-}))
-
-arms = grep('arms', cleaned, value=T)
-segs = grep('arms', cleaned, invert=T, value=T)
-
-length(segs) == length(arms)
-
-# Load segment files
-tiled.segs = do.call(rbind, lapply(segs, function(f) {
-  load.segment.matrix(f)
-}))
-dim(tiled.segs)
-
-tiled.arms = do.call(rbind, lapply(arms, function(f) {
-  load.segment.matrix(f)
-}))
-
-length(z.mean) == ncol(tiled.segs)
-
-for (i in 1:ncol(tiled.segs)) 
-  tiled.segs[,i] = unit.var(tiled.segs[,i], z.mean[i], z.sd[i])
-
-for (i in 1:ncol(tiled.arms)) 
-  tiled.arms[,i] = unit.var(tiled.arms[,i], z.arms.mean[i], z.arms.sd[i])
-
-dim(tiled.arms)
-dim(tiled.segs)
-
-cx.score = unit.var(score.cx(tiled.segs,1), mn.cx, sd.cx)
-
-validation.df = subtract.arms(tiled.segs, tiled.arms)
-validation.df = cbind(validation.df, 'cx'=cx.score)
-
-fauxlabel = rep(0, nrow(validation.df) )
-names(fauxlabel) = intersect(validation.patient.info$Samplename, rownames(validation.df))
-
-```
-
-```{r validation_globalplot, message=F, warning=F, echo=F, fig.height=8, fig.width=8}
-valCNdf = t(validation.df[,grep('^\\d+:',colnames(validation.df))])
-
-locs = do.call(rbind.data.frame, lapply(rownames(valCNdf), function(x) unlist(strsplit(x,':|-'))))
-colnames(locs) = c('chr','start','end')
-locs[] = lapply(locs[], function(x) as.numeric(as.character(x)))
-
-valCNdf = cbind.data.frame(locs, valCNdf)
-valCNdf = melt(valCNdf, id.vars=colnames(locs))
-valCNdf$chr = factor(valCNdf$chr, levels=c(1:22), ordered = T)
-
-valCNdf$seg.type = '5MB'
-valCNdf[with(valCNdf, which( start %in% arm.loc$start & end %in% arm.loc$end)), 'seg.type'] = 'arms'
-valCNdf = merge(valCNdf, chr.lengths[,c('chrom','chr.length')], by.x='chr', by.y='chrom')
-
-valCNdf = merge(valCNdf, validation.patient.info[,c('Hospital.Research.ID','Patient','Samplename','Pathology','Status')], by.x='variable', by.y='Samplename')
-valCNdf$Status = ifelse(valCNdf$Status == 'P', 'Progressor', 'Non-Progressor')
-
-valCNdf$cn = factor(sapply(valCNdf$value, cn), levels=c('loss','neutral','gain'), ordered = T)
-
-n.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(valCNdf, Status == 'Progressor')$variable)), ')', sep=''),
-              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(valCNdf, Status == 'Non-Progressor')$variable)),')', sep=''))
-# Average cn value
-valCNdf = valCNdf %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
-valCNdf$mean.value = valCNdf$value # but not going to use it
-
-p = cn.mtn.plot(subset(valCNdf, seg.type != 'arms'), labeller(Status=n.status, chr=label_value)) + theme(legend.position = 'none') + labs(title='Validation: All samples')
-p
-ggsave('plots/validation_mtn.png', plot=p, width=10, height=5, units='in', dpi=300)
-
-```
-
-```{r validation, message=F, warning=F, echo=F, fig.height=8, fig.width=8}
-file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
-message(paste("loading", file))
-load(file, verbose=T)
-
-fitV = models[[select.alpha]]
-lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']
-
-pred = predict(fitV, newx=validation.df, s=lambda.opt, type='response')
-or = predict(fitV, newx=validation.df, s=lambda.opt, type='link')
-
-vpd = lapply(unique(validation.patient.info$Hospital.Research.ID), function(pt) {
-  info = subset(validation.patient.info, Hospital.Research.ID == pt)
-  info$Prediction = NA
-  info$OR = NA
-  info$PID = unlist(lapply(info$Path.ID, function(x) unlist(strsplit(x, 'B'))[1]))
-
-  info = cbind(info, v.sampleVar[info$Samplename,])
-
-  return(info)
-})
-names(vpd) = unique(validation.patient.info$Hospital.Research.ID)
-
-# for (s in rownames(pred)) {
-#   vpd[which(vpd$Samplename == s), 'Prediction'] = pred[s,]
-#   vpd[which(vpd$Samplename == s), 'OR'] = or[s,]
-# }
-
-for (pt in names(vpd)) {
-  vpd[[pt]]$Prediction[which(vpd[[pt]]$Samplename %in% rownames(pred))] = pred[ intersect(rownames(pred), vpd[[pt]]$Samplename), ]
-  vpd[[pt]]$OR[which(vpd[[pt]]$Samplename %in% rownames(or))] = or[ intersect(rownames(or), vpd[[pt]]$Samplename), ]
-}
-
-back.v = do.call(rbind.data.frame,lapply(vpd, transform.by.time, info=validation.patient.info))
-back.v = merge(back.v, validation.patient.info[,c('Samplename','p53.Status')])
-
-sq = c(0, seq(1,132,12),max(back$months.before.final))
-back.v$brks = cut(back.v$months.before.final, 
-                  breaks=sq, 
-                 labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-c(length(sq)-1,length(sq))], '>120'),
-                  include.lowest = T, ordered_result = T)
-back.v$brks = with(back.v, factor(brks, rev(levels(brks))))
-
-back.v$Pathology = as.character(back.v$Pathology)
-
-back.v$Pathology = factor(back.v$Pathology, levels=c('BE','ID','LGD','HGD','IMC'), ordered = T)
-
-```
-
-```{r echo=F, message=F, warning=F, fig.height=5, fig.width=8}
-back.v$quants = with(back.v, cut(Prediction, breaks=cuts))
-back.v = merge(back.v, pred.confidence[c('quant','perc','Risk')], by.x='quants', by.y='quant')
-back.v$Risk = factor(back.v$Risk, levels=c('High','Moderate','Low'))
-back.v$nl = factor(back.v$nl, ordered = T)
-
-sq = c(0, seq(1, max(back.v$months.before.final), 12))
-back.v$brks = cut(back.v$months.before.final, 
-                  breaks=sq, 
-                 labels=c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-length(sq)],
-                  include.lowest = T, ordered_result = T)
-back.v$brks = with(back.v, factor(brks, levels=rev(levels(brks))))
-
-m = melt(back.v %>% group_by(Patient, Status) %>% summarise( 'Low'=length(which(Risk == 'Low'))/length(Patient),
-                                            'Moderate'=length(which(Risk == 'Moderate'))/length(Patient),
-                                            'High'=length(which(Risk == 'High'))/length(Patient),
-                                            'n.samples'=length(Patient)), id.vars=c('Patient','Status','n.samples'))
-
-ggplot(m, aes(Patient,  value*100, group=variable, fill=variable)) + geom_bar(stat='identity', position=position_stack()) +
-  geom_text(aes(label=n.samples, y=101)) + scale_fill_manual(values=riskCols) +
-  facet_grid(~Status, scales='free_x') + labs(y='Percent of samples', title='Validation patient set') + plot.theme
-```
-
-```{r echo=F, warning=F, message=F, fig.height=10,fig.width=25}
-back.v$Patient = factor(as.numeric(back.v$Patient), ordered=T)
-
-p1 = pred.tiles(subset(back.v, Status == 'P'), cols=myPal, probs=T, nrow=1) + labs(title='Validation Progressors', x='') + theme(legend.position = 'bottom') 
-valLeg = get.legend(p1)
-p1 = p1 + theme(legend.position = 'none')
-p2 = pred.tiles(subset(back.v, Status == 'NP'), cols=myPal, probs=T, nrow=1) + labs(title='Validation Non-Progressors') + theme(legend.position = 'none')
-
-grid.arrange(p1, p2, nrow=2)
-
-ggsave(filename='plots/val_heatmap.png', plot=grid.arrange(p1, p2, nrow=2), width=18, height=6, units='in', dpi=300)
-
-ggsave(filename='plots/val_leg.png', plot=valLeg, width=4, height=2, units='in', dpi=300)
-
-```
-
-```{r echo=F, warning=F, message=F, fig.height=5,fig.width=8}
-vig.plot(subset(back.v, Patient == 68), subset(validation.patient.info, Patient == 68), cols=riskCols)
-```
-
-```{r, echo=F, message=F, warning=F, fig.width=8, fig.height=8}
-riskPath = rbind( cbind.data.frame( table(subset(back.v, Status == 'P', select=c('Pathology','Risk'))), 'Status'='Progressor' ), 
-                  cbind.data.frame( table(subset(back.v, Status == 'NP', select=c('Pathology','Risk'))), 'Status'='Non-Progressor' ))
-
-pathTotal = back.v %>% group_by(Status, Pathology) %>% summarise( 'nPath'=length(Pathology))
-pathTotal$Status = ifelse(pathTotal$Status == 'P', 'Progressor','Non-Progressor')
-
-riskPath = merge(riskPath, pathTotal)
-riskPath$ratio = with(riskPath, round(Freq/nPath, 3))
-riskPath$Risk = factor(riskPath$Risk, levels=c('Low','Moderate','High'), ordered = T)
-
-hgdimc = subset(riskPath, Pathology %in% c('HGD','IMC')) %>% group_by(Risk) %>% summarize (
-  Freq=sum(Freq),  nPath=sum(nPath), ratio=sum(Freq)/sum(nPath))
-
-riskPath = riskPath[-which(riskPath$Pathology %in% c('HGD','IMC')),] 
-riskPath = rbind(riskPath , cbind('Pathology'='HGD/IMC', 'Status'='Progressor', hgdimc))
-
-riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor'), ordered = T)
-
-pathTotal = riskPath %>% group_by(Status, Pathology) %>% summarise('nPath'=unique(nPath))
-
-pr = ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) +   scale_fill_manual(values=riskCols) + scale_color_manual(values=c('white','black','white')) +
-  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), position=position_stack(), size=4, show.legend = F) +
-  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep=''), color=Risk), position=position_stack(vjust = 0.5), size=5, show.legend = F) +
-  facet_grid(~Status) + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
-pr
-
-ggsave(filename='plots/val_riskpath.png', plot=pr, width=7, height=7, units='in', dpi=300)
-
-```
-
-## Validation prediction calibration
-
-```{r, echo=F, warning=F, message=F, fig.width=4, fig.height=3}
-cuts = seq(0,1,0.1)
-qt = as.data.frame.matrix(table(back.v$quants, back.v$Status))
-qt$quant = rownames(qt)
-
-qt = cbind(qt, back.v %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
-
-pred.confidence.v = qt %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 'perc'=P/sum(NP,P) )
-
-pred.confidence.v[c('r1','r2')] = NA
-for (i in 1:(length(cuts)-1)) 
-    pred.confidence.v[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)
-
-pred.confidence$Risk = factor(pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)
-
-ggplot(pred.confidence.v, aes(mn, perc)) + 
-  geom_rect(data=pred.confidence, aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
-  geom_point() +
-  #geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + 
-  labs(x='Mean prediction', y='Ratio of `P` patients', title='Validation calibration') +
-  scale_color_manual(values=riskCols) + plot.theme
-
-preds = do.call(rbind.data.frame, lapply(vpd, function(df) df[c('Status','Prediction')]))
-vroc = pROC::roc(Status ~ Prediction, data=preds, ci=T, of='thresholds')
-vroc$model = 'Validation'
-
-roc$model = 'Discovery'
-multi.roc.plot(list(roc, vroc), title='ROC', palette='Dark2') + theme(text=element_text(size=12))
-#ggsave('plots/auc_dv.png', plot=p, width=5, height=4, units='in', dpi=300, scale=2)
-```
-
-```{r, echo=F, warning=F, message=F, fig.width=4, fig.height=4}
-aucs = rbind.data.frame(pROC::ci.auc(roc), pROC::ci.auc(vroc))
-colnames(aucs) = c('ci.min', 'auc','ci.max')
-aucs$model = c('Discovery','Validation')
-
-ggplot(aucs, aes(model, auc)) + geom_point(size=2) + ylim(0,1) +
-  geom_errorbar(aes(ymin=ci.min, ymax=ci.max), width=0.1) + coord_flip() + 
-  geom_text(aes(y=ci.min, label=round(ci.min, 2)), vjust=4) +
-  geom_text(aes(y=ci.max, label=round(ci.max, 2)), vjust=-4) +
-#  geom_text(aes(y=auc, label=round(auc, 2)), vjust=-1) +
-  labs(x='', y='AUC', title='AUC with CI') + plot.theme 
-```
-
-# Clinical Story
-
-First note here is that our probability distribution for these 20 patients is somewhat flatter than in the training set.
-Likely a consequence of sample size.
-The labels show the proprotions of progressors that fell into each quintile in the training set.
-
-```{r, echo=F, warning=F, message=F}
-
-summary.v = table.path.risk(back.v)
-
-pander(as.data.frame(summary.v))
-
-write.table(summary.v, sep='\t', quote=F, row.names = F)
-```
-
-## Vignettes
-
-```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=5}
-vign.pts = as_tibble(subset(sum.patient.data, Patient %in% c(3, 33, 34, 49, 56, 20), select=c('Status','Patient','age.diagnosis', 'smoking', 'C','M', 'initial.endo','final.endo', 'total.endos','total.samples')))
-colnames(vign.pts) = c('Status','Patient','Age at Diagnosis','Smoker','Circumference', 'Maximal length', 'Intial Endoscopy','Final Endoscopy', 'Total Endoscopies','Total Samples')
-
-## P
-p = vig.plot(df=subset(back, Patient == 34),info= subset(patient.info, Patient == 34),cols=riskCols) + theme(legend.position = 'bottom')
-vigLeg = get.legend(p)
-p = p + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_34.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(subset(back, Patient == 56), subset(patient.info, Patient == 56),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_56.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(subset(back, Patient == 20), subset(patient.info, Patient == 20),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_20.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(subset(back, Patient == 86), subset(patient.info, Patient == 86),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_86.png', plot=p, width=6, height=6,dpi=300)
-
-print(p)
-
-## NP
-p = vig.plot(subset(back, Patient == 3), subset(patient.info, Patient == 3),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_3.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(df=subset(back, Patient == 23),info= subset(patient.info, Patient == 23),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_23.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(subset(back, Patient == 33), subset(patient.info, Patient == 33),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_33.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(subset(back, Patient == 49), subset(patient.info, Patient == 49),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_49.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(subset(back, Patient == 51), subset(patient.info, Patient == 51),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_51.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(subset(back, Patient == 6), subset(patient.info, Patient == 6),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_6.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(subset(back, Patient == 54), subset(patient.info, Patient == 54),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_54.png', plot=p, width=6, height=6,dpi=300)
-
-print(p)
-
-```
-
-### Recommendations
-
- 1.	Immediate RFA: HGD diagnosis, two or more consecutive high risk predictions
- 2.	Recheck 6-12 months: One high risk prediction or aberrant p53 IHC
- 3. Recheck endoscopy 12-24 months: One or more moderate risk predictions
- 4. Regular surveillance 3-5 years: Two or more consecutive low risk predictions
-
-```{r, rx, echo=F, warning=F, message=T, fig.width=8, fig.height=5}
-
-rx<-function(Patient, brks, Prediction, Pathology, Risk, p53.Status) {
-  rules = as.data.frame(matrix(ncol=0, nrow=0))
-  # Consecutive
-  for (i in 1:length(brks)) {
-    risks = table(Risk[i:(i+1)])
-    p53 = table(p53.Status[i:(i+1)])
-    rule = 'None'
-    if ( risks['High'] == 2 | length(which(grepl('HGD|IMC', Pathology[i:(i+1)]))) > 0 ) {
-      rule = 1
-    } else if ( risks['High'] == 1 | p53['1'] > 0 ) {
-      rule = 2
-    } else if ( risks['Moderate'] > 0 ) {
-      rule = 3
-    } else if ( risks['Low'] == 2) {
-      rule = 4
-    }
-  rules = rbind(rules, cbind( as.character(brks[i]), as.character(brks[(i+1)]), rule))
-  }  
-  colnames(rules) = c('t1','t2', 'rule')
-  rules$Patient = Patient
-  rules
-}
-
-cols = unique(colnames(back), colnames(back.v))
-cols = cols[-grep('quants|pred\\.ratio|se\\(Risk\\)', cols)]
-allB = rbind.data.frame(back[,cols], back.v[,cols])
-allB$p53.Status = factor(allB$p53.Status, levels=c(0,1), ordered = T)
-
-maxP = allB %>% group_by(Patient, Status, brks) %>% summarise( 'maxP'=max(Prediction), 'Risk'=Risk[Prediction == max(Prediction)], 'Pathology'=max(Pathology), 'p53'=max(p53.Status,na.rm=T) )
-
-rules = do.call(rbind, lapply(unique(maxP$Patient), function(pt) {
-  r = with(arrange(subset(maxP, Patient == pt), brks), rx(pt, brks,maxP,Pathology,Risk,p53))
-  r$Status = unique(subset(maxP, Patient == pt)$Status)
-  r
-}))
-
-rules$t1 = factor(rules$t1, levels=levels(back$brks), ordered=T)
-rules$t2 = factor(rules$t2, levels=levels(back$brks), ordered=T)
-
-penult = subset(rules, t2 != 'Endpoint' & (t1 == rev(levels(back$brks))[2] | t2 == rev(levels(back$brks))[2] ))
-round(table(subset(penult, Status == 'P')$rule)/nrow(subset(penult, Status == 'P')), 2)
-round(table(subset(penult, Status == 'NP')$rule)/nrow(subset(penult, Status == 'NP')), 2)
-
-second = subset(rules,  t2 != 'Endpoint') %>% 
-  group_by(Patient, Status) %>% 
-  summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )
-table(subset(second, Status == 'P')$rule)
-table(subset(second, Status == 'NP')$rule)
-
-```
-
-## Incidence
-
-Median follow up for our discovery cohort was `r median(sum.patient.data$end.year-sum.patient.data$start.year)`.
-
-Adjust the RR of the orginal model with 1-3.5% offset to get an absolute risk (AR) with the understanding that the cohort was age matched and unbalanced.
-
-```{r echo=F}
-file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
-load(file, verbose=T)
-rm(plots, coefs, cv.patient)
-
-fit = models[[select.alpha]]
-lambda = performance.at.1se[[select.alpha]]$lambda
-
-# number of patients we'd needed to have included to get 45 progressors from a representative population (I think)
-all = rbind(sum.patient.data, sum.validation.info)
-
-# Per year
-# This isn't quite right...we'd want to do something more like a 5 year 
-#pop = round((table(all$Status)['P']/(0.07*100))*100)
-
-```
-
-```{r estimateRisk, echo=F, warning=F, message=F, eval=F}
-# Probability that an individual will progress in a lifetime (20 years) given a rate of progression of 0.07 is about 77%, the probability for a specific individual is less
-1-(0.93^20)
-
-(1-(0.99^3))*15
-(1-(0.99^3))/10
-(1-(0.99^3))/10
-
-# Or probability an individual will progress given a rate of 1%-3.5%
-poft = sapply(seq(0.96,0.99, .01), function(x) {
- sapply(1:5, function(y) 1-(x^y) )
-})
-
-poft = as.data.frame(poft)
-colnames(poft) = paste((1-seq(0.96,0.99, .01))*100, '%', sep='')
-
-rp = apply(poft,1,function(x) cbind(mean(x),min(x), max(x)) )
-
-rownames(rp) = c('mean','min','max')
-colnames(rp) = paste('Year',1:5)
-
-rr = range(as.data.frame(subset(back, Pathology %in% c('BE'), select=c('Patient','Pathology','months.before.final','OR')))$OR)
-
-p = subset(back, Status == 'P' & months.before.final > 6 & OR > 10)$Patient
-subset(sum.patient.data, Patient %in% p) %>% group_by(Patient) %>% summarise(final.endo-initial.endo)
-
-rrPath = back %>% group_by(Status, Pathology) %>% summarise( 'mean.RR'=mean(OR),'ci.min'=mean(OR)-(1.96*sd(OR)), 'ci.max'=mean(OR)+(1.96*sd(OR)) )
-
-ggplot(rrPath, aes(Status, mean.RR)) + #scale_color_gradientn(colors=myPal, limits=c(min(rrPath[,3:5]), max(rrPath[,3:5])),  name='') +
-  facet_wrap(~Pathology, ncol=1) + 
-  geom_point() +  geom_errorbar(aes(ymin=ci.min, ymax=ci.max), width=0.1) + 
-  geom_text(aes(y=ci.min, label=round(ci.min,2)), vjust=2) + geom_text(aes(y=ci.max, label=round(ci.max,2)), vjust=2) +
-  coord_flip() + plot.theme + labs(y='Mean RR', x='')
-
-as.data.frame(arrange(subset(back, Status == 'P' & months.before.final > 6 & Pathology %in% c('BE','ID','LGD') & OR > 1), Patient, months.before.final))
-
-ggplot(t(rp), aes(1:5, mean)) +
-  geom_ribbon(aes(ymin=mean/abs(rr[1]),ymax=mean*rr[2]), fill='rosybrown1') +
-  geom_ribbon(aes(ymin=min,ymax=max), fill='grey88') +
-  geom_line() + 
-  geom_text(data=data.frame(t(rp[,5])), aes(x=5,y=min, label='1%')) + geom_text(data=data.frame(t(rp[,5])), aes(x=5,y=max, label='4%')) +
-  geom_line(aes(y=mean*rr[2]), color='red') +
-  geom_line(aes(y=mean/abs(rr[1])), color='blue') +
-  labs(y='P(P)', x='Year', title='Prob. of individual patient progression', subtitle='1-4% over 5 years') + plot.theme
-
-subset(back, OR > 10)
-cor.test(back$OR, as.integer(back$Pathology))
-as.data.frame(subset(back, Pathology %in% c('HGD','IMC'), select=c('Patient','Pathology','months.before.final','OR')))
-
-as.data.frame(subset(back, Status == 'P' & OR < 2 & OR > -2, select=c('Status','Patient','Pathology','months.before.final','OR')))
-
-hist(subset(back, Pathology %in% c('BE'))$OR)
-
-```
-
-```{r 5yrrisk, echo=F, fig.height=8, fig.width=12}
-cases = table(sum.patient.data$Status)['P']
-
-mn = round((table(sum.patient.data$Status)['P']/(0.01*100))*100)
-m = round((table(sum.patient.data$Status)['P']/(0.0225*100))*100)
-mx = round((table(sum.patient.data$Status)['P']/(0.035*100))*100)
-
-offsetMin = log(cases/mn)
-offsetMean = log(cases/m)
-offsetMax = log(cases/mx)
-
-#binom.test(cases, popL, 0.1)
-# Offset for the linear predictor would then be -log(pop/44)
-# Effectively this is my relative risk for the population
-#offset = log(cases/popL)
-
-# link function provides odds
-#linPred = predict(fit, newx=dysplasia.df, s=lambda, type='link')
-#quantile(linPred)
-
-#df = rbind(dysplasia.df, validation.df)
-#df = dysplasia.df
-
-#p0 = predict(fit, newx=df, s=lambda, type='response')
-#p0 = 1/(1+exp(-linPred)) # logit to get probabilities, same as type='response'
-#p1 = 1/(1+exp(-linPred+abs(offset))) # calculates probs from risk (offset)
-
-cols = c('Hospital.Research.ID','Patient','Status','PID', 'Samplename','Endoscopy.Year','Pathology','Prediction','OR')
-#loopreds = rbind(loopreds[cols], do.call(rbind, vpd)[cols])
-loopreds = do.call(rbind, pg.samp)[cols]
-
-loopreds$P0 = 1/(1+exp(-loopreds$OR))
-
-loopreds$Mean = 1/(1+exp(-loopreds$OR+abs(offsetMean))) # Adjust OR for RR
-loopreds$Min = 1/(1+exp(-loopreds$OR+abs(offsetMin))) # Adjust OR for RR
-loopreds$Max = 1/(1+exp(-loopreds$OR+abs(offsetMax))) # Adjust OR for RR
-
-#ql = cbind.data.frame('y'=quantile(loopreds$OR)[c(2,4)], 'labels'=paste(c(25,75), '%', sep=''))
-
-# From my nested-LOO predictions
-
-ggplot(melt(loopreds[c('Min','Mean','Max')], measure.vars='Mean'), aes(value, fill=variable)) + 
-  geom_density(alpha=0.5) + plot.theme
-
-ggplot(melt(loopreds[c('Min','Max')]), aes(value)) + facet_wrap(~variable, ncol=2) +
-  geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + 
-    scale_fill_gradientn(colors = myPal,  name='') + 
-    plot.theme + theme(legend.position = 'none') + 
-    labs(y='n Samples', x='Absolute Risk',
-         title='5-year adjusted risk', subtitle='Adjusted for upper & lower bound: 1-3.5% risk')
-
-# ggplot(loopreds, aes(Min)) +
-#     geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + 
-#     scale_fill_gradientn(colors = myPal,  name='') + 
-#     plot.theme + theme(legend.position = 'none') + 
-#     labs(y='n Samples', x='Absolute Risk',
-#          title='Sample Predictions: 5 year adjusted risk', subtitle='Adjusted for lower bound, 1% risk')
-# 
-# ggplot(loopreds, aes(Max)) +
-#     geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + 
-#     scale_fill_gradientn(colors = myPal,  name='') + 
-#     plot.theme + theme(legend.position = 'none') + 
-#     labs(y='n Samples', x='Absolute Risk',
-#          title='5-year adjusted risk', subtitle='Adjusted for upper bound, 3.5% risk')
-
-```
-
-```{r lifetimeFPR, echo=F, fig.height=6, fig.width=6, eval=F}
-## FPR estimates based on lifetime risks
-# We could consider any NP patient with a minimum of 2 High risk calls?  or High/Moderate Risk?
-# More correctly it should be any patients with concurrent High risk, but that's been pretty stable so two high is good enough for now
-
-bd = bind_rows(back, back.v)
-table(bd$Status)
-
-risksL = subset(bd, Status == 'NP') %>% group_by(Patient) %>% summarise( 'n.samp'=length(Patient), 'high.risk'=length(which(Risk == 'High')), 'mod.risk'=length(which(Risk == 'Moderate')) )
-
-risksL = subset(risksL, high.risk > 0 | mod.risk > 0)
-
-high = length(subset(risksL, high.risk >= 2 )$Patient)
-#mod = length(subset(risksL, mod.risk >= 2 )$Patient)
-
-# How would I tune it so that the FPR is a lot closer to the real incidence ~1%
-# Current false positive rate is 1.5x higher than the highest incidence ~0.7
-fpr=high*10/44
-x = subset(bd, Status == 'NP' & Patient %in% subset(risksL, high.risk >= 2)$Patient & Risk == 'High')
-(length(unique(subset(x, Prediction > 0.8)$Patient))*10)/44
-
-# How would that affect TPR?
-#pROC::coords(roc, 0.6)
-#pROC::coords(roc, 0.8)
-
-## Per year risk
-## Evaluate over 5 years? How?
-pyr = c()
-for (yr in names(which(table(bd$Endoscopy.Year) > 10))) {
-  risks = subset(bd, Status == 'NP' & Endoscopy.Year == yr) %>% group_by(Patient) %>% summarise( 'n.samp'=length(Patient), 'high.risk'=length(which(Risk == 'High')), 'mod.risk'=length(which(Risk == 'Moderate')) )
-
-  risks = subset(risks, high.risk > 0 | mod.risk > 0)
-  high = length(subset(risks, high.risk >= 2 )$Patient)
-  mod = length(subset(risks, mod.risk >= 2 )$Patient)
-
-  print(paste(yr,high*15, sep=':'))
-  pyr = c(pyr, high*15)
-}
-
-summary(pyr)
-
-pop*.07
-pop*.02
-pop*.05
-
-```
-
-```{r clonality, echo=F, message=F, warning=F, eval=F}
-## Clonal Heterogenity
-patient.info %>% group_by (Hospital.Research.ID, Patient, PID) %>% summarise(
-  length(PID)
-)
-
-arms = c(grep(with(chr.lengths, paste(paste(chrom, ':', '1-\\d+000', sep=''), collapse='|')), colnames(dysplasia.df)),
-         grep(with(chr.lengths, paste(paste(chrom, ':', '\\d+000-', chr.length, sep=''), collapse='|')), colnames(dysplasia.df)))
-
-x = do.call(rbind, lapply(unique(patient.info$PID), function(pid) {
-    df = as.matrix(dysplasia.df[subset(info, PID == pid)$Samplename, -grep('cx',colnames(dysplasia.df))])
-    if (ncol(df) > 1) {
-      cbind('arms'=sum(rowSums(df[,arms],na.rm=T))/nrow(df),
-            'seg'=sum(rowSums(df[,-arms],na.rm=T))/nrow(df))
-    } else {
-      cbind('arms'=sum(df[arms,],na.rm=T),
-            'seg'=sum(df[-arms,],na.rm=T))
-    }
-}))
-rownames(x) = unique(patient.info$PID)
-x = x[complete.cases(x),]
-
-hist(x[,'seg'])
-hist(x[,'arms'])
-
-p = pnorm(x[,'seg'], mean(x[,'seg']), sd(x[,'seg']))
-summary(p)
-
-pa = pnorm(x[,'arms'], mean(x[,'arms']), sd(x[,'arms']))
-summary(pa)
-
-x[,'arms'] = scale(x[,'arms'])
-x[,'seg'] = scale(x[,'seg'])
-
-plot(x[unique(subset(patient.info, Patient == 48)$PID), 'arms'])
-plot(x[unique(subset(patient.info, Patient == 48)$PID), 'seg'])
-
-plot(x[unique(subset(patient.info, Patient == 34)$PID), 'arms'])
-plot(x[unique(subset(patient.info, Patient == 34)$PID), 'seg'])
-
-pp = back %>% group_by(Patient, Hospital.Research.ID, PID) %>% summarise(
-  'max'=max(Prediction), 'avg'=mean(Prediction)
-)
-
-pp = merge(pp, x, by.x='PID', by.y='row.names')
-
-cor.test(pp$seg, pp$max)
-
-```
-
-## Cox model - Time to progress?
-
-```{r cox, warning=F, message=F, echo=F, eval=T}
-library(survival)
-
-cox.pvalue<-function(cox, test='logrank') {
-  pvalue = NULL  
-  csum = summary(cox)
-
-  if (test == 'logrank') {
-    pvalue = csum$sctest[['pvalue']]  
-  } else if (test == 'wald') {
-    pvalue = csum$waldtest[['pvalue']]
-  } else if (test == 'loglik') {
-    pvalue = csum$logtest[['pvalue']]
-  }
-  return(pvalue)
-}
-
-prog.pred = lapply(pg.samp, function(df) {
-  df = df[,c('Hospital.Research.ID', 'Status', 'Patient', 'Age.at.diagnosis', 'PID', 'Status', 'Endoscopy.Year','Initial.Endoscopy', 'Final.Endoscopy', 'Pathology', 'Samplename','Prediction','Prediction.Dev.Resid')] %>% rowwise() %>% mutate(
-      'Time.to.Final'=Final.Endoscopy-Endoscopy.Year)
-  df = arrange(df, Endoscopy.Year, Pathology)
-
-  df$Dup.PID = duplicated(df$PID)
-  x = df[!df$Dup.PID,]
-
-  for (i in 1:nrow(x)) {
-    if (i == 1) {
-      x[i,'tstart'] = 0
-      x[i, 'tstop'] = x[(i+1), 'Endoscopy.Year'] - x[i, 'Endoscopy.Year']
-      if (is.na(x[i,'tstop']) || x[i,'tstop'] == 0) x[i,'tstop'] = 0.5
-      next
-    }
-
-    x[i,'tstart'] = x[(i-1), 'tstop']
-
-    # Final sample is different -- will leave it out later
-    if (i == nrow(x)) {
-      x[i, 'tstop'] = x[i,'tstart'] + 0.5
-      break
-    }
-
-    x[i, 'tstop'] =  x[i,'tstart'] + x[(i+1), 'Endoscopy.Year'] - x[i, 'Endoscopy.Year']
-
-    # Samples taken in the same year
-    if ( x[i, 'Endoscopy.Year'] == x[(i+1), 'Endoscopy.Year'] )
-      x[i,'tstop'] = x[i,'tstop'] + 0.5 
-  }
-
-  x$Diagnosis = as.integer(x$Pathology %in% c('HGD','IMC'))
-
-  if (sum(x$Diagnosis) >= 1)
-    x = x[-which(x$Diagnosis == 1),]
-
-  if (nrow(x) > 0 && sum(x$Diagnosis) <= 0 && x$Status == 'P')
-    x[nrow(x), 'Diagnosis'] = 1
-
-  if (nrow(x) > 0) {
-    x$Dup.Samples = ''
-    for (j in 1:nrow(df[df$Dup.PID,]))  {
-      row = df[df$Dup.PID,][j,]
-      x[which(x$PID == row[['PID']]), 'Dup.Samples'] = paste(x[which(x$PID == row[['PID']]), 'Dup.Samples'], row[['Samplename']], sep=',')
-    }
-    x$Dup.Samples = sub('^,','',x$Dup.Samples)
-  }
-
-  return(x)
-})
-## By pathology?
-pp = do.call(rbind, prog.pred)
-
-pp[is.na(pp$tstop), 'tstop'] = 0.5
-pp$Pathology = as.character(pp$Pathology)
-pp$AAD = cut(pp$Age.at.diagnosis, breaks = 3)
-
-pp = merge(pp, back[c('Samplename','Risk', 'quants')], by='Samplename')
-pp$Risk = as.character(pp$Risk)
-
-sfit = survfit( Surv(tstart, tstop, Diagnosis, type='interval') ~ Risk, data=pp, conf.int=0.95, conf.type='log', na.action=na.omit)
-ggsurv(sfit,cens.shape=3,lty.est=1,back.white=F) + xlim(0,5) + labs(y='Diagnosed', title='~Risk')
-
-cox = coxph( Surv(tstart, tstop, Diagnosis) ~ Risk, data=pp) 
-ggsurv(survfit(cox),cens.shape=3,lty.est=1,back.white=F,
-      main="coxph(Survival, Diagnosis) ~ Risk" ) + 
-      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),2),"\n", nrow(pp), 'samples'))
-#summary(cox)
-
-sfit = survfit( Surv(tstart, tstop, Diagnosis, type='interval') ~ AAD, data=pp, conf.int=0.95, conf.type='log', na.action=na.omit)
-ggsurv(sfit,cens.shape=3,lty.est=1,back.white=F) + xlim(0,5) + labs(y='Diagnosed', title='~Age') 
-cox = coxph( Surv(tstart, tstop, Diagnosis) ~ AAD, data=pp) 
-ggsurv(survfit(cox),cens.shape=3,lty.est=1,back.white=F,
-      main="coxph(Survival, Diagnosis) ~ Age at diagnosis" ) + 
-      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),2),"\n", nrow(pp), 'samples'))
-#summary(cox)
-
-sfit = survfit( Surv(tstart, tstop, Diagnosis, type='interval') ~ Pathology, data=pp, conf.int=0.95, conf.type='log', na.action=na.omit)
-ggsurv(sfit,cens.shape=3,lty.est=1,back.white=F) + xlim(0,5) + labs(y='Diagnosed', title='~Pathology')
-cox = coxph( Surv(tstart, tstop, Diagnosis) ~ Pathology, data=pp) 
-ggsurv(survfit(cox),cens.shape=3,lty.est=1,back.white=F,
-      main="coxph(Survival, Diagnosis) ~ Pathology" ) + 
-      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),2),"\n", nrow(pp), 'samples')) + xlim(0,5)
-#summary(cox)
-
-## All genomic regions
-sample.pred = merge( do.call(rbind.data.frame, prog.pred), dysplasia.df, by.x='Samplename', by.y='row.names' )
-sample.pred$Pathology = as.character(sample.pred$Pathology)
-
-gcols = grep('\\d+:', colnames(sample.pred), value=T)
-sample.pred = sample.pred[c('tstart', 'tstop','Diagnosis','Pathology','Age.at.diagnosis',gcols)]
-
-sample.pred$Age.bracket = as.character(cut(sample.pred$Age.at.diagnosis, breaks=3))
-
-# all
-#cox = coxph( Surv(tstart, tstop, Diagnosis) ~., data=sample.pred)
-#ggsurv(survfit(cox))
-#summary(cox)
-```
-### Cox GLM selected features
-
-```{r echo=F, warning=F, echo=F}
-# glm selected features + Path
-cox2 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', 'Pathology',  grep('^\\d+:', ch$label, value=T))]) 
-ggsurv(survfit(cox2),cens.shape=3,lty.est=1,back.white=F,
-      main="coxph(Survival, Diagnosis) ~ GLM features + Pathology" ) + 
-      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox2),2),"\n", nrow(pp), 'samples')) 
-#summary(cox2)
-
-cox2 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis','Age.bracket', grep('^\\d+:', ch$label, value=T))]) 
-ggsurv(survfit(cox2),cens.shape=3,lty.est=1,back.white=F,
-      main="coxph(Survival, Diagnosis) ~ GLM features + Age" ) + 
-      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox2),2),"\n", nrow(pp), 'samples')) 
-#summary(cox2)
-
-cox2 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis',grep('^\\d+:', ch$label, value=T))]) 
-ggsurv(survfit(cox2),cens.shape=3,lty.est=1,back.white=F,
-      main="coxph(Survival, Diagnosis) ~ GLM features" ) + 
-      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox2),2),"\n", nrow(pp), 'samples')) 
-#summary(cox2)
-
-# chr arms only
-options(scipen=999) # turn sci notation off
-coxArms = coxph(Surv(tstart, tstop, Diagnosis) ~., data=sample.pred[,c('tstart', 'tstop','Diagnosis', with(arm.loc, paste(chrom, ':', start, '-', end, sep='')))])
-ggsurv(survfit(coxArms),cens.shape=3,lty.est=1,back.white=F,
-      main="coxph(Survival, Diagnosis) ~ chr arms" ) + 
-      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(coxArms),2),"\n", nrow(pp), 'samples')) 
-#summary(coxArms)
-
-# chr 17 only
-cox17 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', grep('^17:',colnames(sample.pred), value=T))]) 
-ggsurv(survfit(cox17),cens.shape=3,lty.est=1,back.white=F,
-      main="coxph(Survival, Diagnosis) ~ chr17" ) + 
-      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox17),2),"\n", nrow(pp), 'samples')) 
-#summary(cox17)
-
-cox11 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', grep('^11:',colnames(sample.pred), value=T))]) 
-ggsurv(survfit(cox11))
-#summary(cox11)
-
-cox13 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', grep('^13:',colnames(sample.pred), value=T))]) 
-ggsurv(survfit(cox13))
-#summary(cox13)
-
-```
-
-## CoxRFX
-
-```{r echo=F, message=F, warning=F, eval=T}
-plot.waldtest<-function(wt) {
-  p = ggplot(wt, aes(coef.x, coef.y, color=p.value<0.05))
-  if (length(grep('hazards', colnames(wt))) > 0) {
-      p = p + geom_point(alpha=0.5, aes(size=hazards)) + 
-        geom_point(aes(coef.x, coef.y, color=p.value<0.05, size=hazards), show.legend = F) +
-        #scale_shape_manual(values=c(17,18), na.value=21) +
-        scale_size(range=c(2, 10), breaks=c(0, 5, 10, 15)) +
-        #geom_text_repel( aes( label=label), show.legend=F) + 
-        geom_text_repel(data=subset(wt, p.value < 0.05), aes(coef.x,coef.y, label=label), show.legend=F) + 
-        geom_text_repel(data=subset(wt, is.na(stability)), aes(coef.x, coef.y, label=label, color=p.value<0.05), show.legend=F, color='grey39') 
-  } else {
-    p = p + geom_point(alpha=0.5) + 
-      geom_text_repel(data=subset(wt, p.value < 0.05), aes(coef.x,coef.y, label=label), show.legend=F)  
-  }
-  return(p + labs(title='Wald test'))
-}
-```
-
-```{r echo=T, message=F, warning=F, fig.height=8, fig.width=8, eval=F}
-
-library(CoxHD)
-
-coxDF = sample.pred
-
-pathology = do.call(cbind.data.frame, lapply(unique(coxDF$Pathology), function(p) {
-  cbind( as.integer(coxDF$Pathology == p))
-}))
-colnames(pathology) = unique(coxDF$Pathology)
-
-# age = do.call(cbind.data.frame, lapply(unique(coxDF$Age.at.diagnosis), function(p) {
-#   cbind( as.integer(coxDF$Age.at.diagnosis == p))
-# }))
-# colnames(age) = unique(coxDF$Age.at.diagnosis)
-age = coxDF$Age.at.diagnosis
-
-## GLM regions only
-coxDF = cbind(coxDF[,1:3],pathology[,c('BE','ID')], age, coxDF[,grep('^\\d+:', ch$label, value=T)])
-coxDF$Pathology = NULL
-
-gcols = colnames(coxDF)
-cols = grep('^\\d+', colnames(coxDF))
-colnames(coxDF)[cols] = paste('chr', gsub(':|-','.', grep('^\\d+', colnames(coxDF), value=T)), sep='')
-
-coxSM = cbind(coxDF[,c(1:3)], StandardizeMagnitude(coxDF[,-c(1:3)]))
-
-#surv = with(coxDF, Surv(tstart, tstop, Diagnosis))
-surv = with(coxSM, Surv(tstart, tstop, Diagnosis))
-groups = factor(c(rep('clinical',3),c(rep('genomic', length(grep('\\d+', colnames(coxDF)))))))
-rfx = CoxRFX(coxSM[,-c(1:3)], surv, groups, verbose=T, nu=1)
-summary(rfx)
-
-plot(rfx)
-
-wt = WaldTest(rfx)
-survConcordance(surv ~ predict(rfx))
-
-wt$label = sub('\\.', '-', sub('\\.', ':', sub('chr', '', rownames(wt))))
-wt = merge(wt, ch[,c('label','hazards', 'coef','stability')], by='label', all=T)
-
-wt[is.na(wt$hazards), c('hazards', 'coef.y')] = 0
-plot.waldtest(wt) + labs(x='RFX coef',y='GLM coef', title='RFX vs GLM coefficients', subtitle='Labeled points have a significant p-value in RFX (except BE/ID)')
-
-options(scipen = 999)
-# arms only
-arms = with(arm.loc, paste(chrom, ':', start, '-', end, sep=''))
-coxDfArms = coxDF[,arms]
-colnames(coxDfArms) = paste('chr', gsub(':|-','.', grep('^\\d+', colnames(coxDfArms), value=T)), sep='')
-
-surv = with(coxDF, Surv(tstart, tstop, Diagnosis))
-groups = factor(c(rep('clinical',3), rep('genomic', ncol(coxDfArms))))
-
-rfx = CoxRFX( cbind.data.frame(coxSM[,c('ID','BE','age_10')], coxDfArms), surv, groups, verbose=T, nu=1)
-summary(rfx)
-
-plot(rfx)
-wt = WaldTest(rfx)
-survConcordance(surv ~ predict(rfx))
-
-```
-
-```{r, eval=F}
-# TODO: Relative risk
-pred = predict(f, newx=dysplasia.df, s=performance.at.1se$`0.9`$lambda)
-?predict
-?predict.glm
-quantile(pred)
-p0 = predict(f, newx=dysplasia.df, s=performance.at.1se$`0.9`$lambda, type='response')
-p0
-p1 = 1/(1+exp(-pred+2.66))
-plot(p0,p1)
-hist(p1)
-hist(p0)
-hist(p1)
-plot(p0,p1)
-hist(p1, n=20)
-mean(p1>.5)
-mean(p0>.5)
-hist(pred)
-
-```
+ggsave('pred_heatmap_ALL.png', plot=grid.arrange(p2,p1,nrow=2), width=52, height=16, units="in", limitsize=F, dpi = 300)
+#ggsave('plots/npg_heatmap.png', plot=p1, width=52, height=8, units="in", limitsize=F, dpi = 300)
+#ggsave('plots/pg_heatmap.png', plot=p2, width=52, height=8, units="in", limitsize=F, dpi = 300)
+#ggsave('plots/risk_legend.png', plot=grid.arrange(riskLegend), width=2,height=4, dpi=300)
+#ggsave('plots/risk_legend_long.png', plot=grid.arrange(riskLegendLong), width=8,height=2, dpi=300)
+```

### 2018-03-06 Insilico Splits.Rmd
adding to pipeline

---
title: "Insilico Splits"
author: "Sarah Killcoyne"
date: "1/10/2018"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(ggrepel)
library(pander)
library(reshape2)
library(gridExtra)
library(plyr) 
library(dplyr)
library(pROC)

source('lib/load_patient_metadata.R')
source('lib/cv-pt-glm.R')
source('lib/common_plots.R')

data = '~/Data/Ellie'
tile.w=c(5e06, 'arms')

includeDemo = F

data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
analysis.files = list.files(paste(data, 'Analysis', sep='/'), full.names=T)

load(grep('All_patients.Rdata', analysis.files, value=T), verbose=T)

## Hospital.Research.ID info file
patient.file = grep('All_patient_info.xlsx', data.files, value=T)
if (length(patient.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", data))
demo.file = grep('Demographics_full.xlsx', data.files, value=T)

all.patient.info = read.patient.info(patient.file, demo.file, set='all')$info
head(all.patient.info)

#sum.pt = summarise.patient.info(all.patient.info)

sum.pt = all.patient.info %>% group_by(Set, Patient) %>% summarize(
  med.reads=median(Total.Reads),
  sd.reads=sd(Total.Reads),
  n.endoscopy=length(unique(PID)),
  n.years=unique(Final.Endoscopy-Initial.Endoscopy),
  n.samples=length(Samplename)
)

sp = sum.pt %>% group_by(Set) %>% summarize(
  med.endo=median(n.endoscopy),
  sd.endo=sd(n.endoscopy),
  med.years=median(n.years),
  med.samples=median(n.samples),
  sd.samples=sd(n.samples)
)

```

`r pander(as.data.frame(sp))`

```{r echo=F, warning=F, message=F}

cache.dir = paste(data, 'Analysis', sub('\\+','',paste(tile.w, collapse='_')), sep='/')

ins_splits = grep('val_', list.dirs(cache.dir), value=T)

performance = data.frame()
rocs = list()

select.alpha = '0.9'

file = paste(cache.dir, 'insilico_splits.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
  rocs = insilico.rocs
  performance = insilico.perf
} else {
  for (split in ins_splits) {
    load(grep('all.*Rdata', list.files(split,full.names = T), value=T), verbose=F)
    performance = rbind(performance, performance.at.1se[[select.alpha]])
    rm(plots,coefs,performance.at.1se,models,cv.patient)

    load(grep('loo.*Rdata', list.files(split,full.names = T), value=T), verbose=F)
    rm(plots, performance.at.1se,coefs,nzcoefs,fits)

    predictions = do.call(rbind, pg.samp)

    preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
    roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
    #rocs = rbind(rocs,c(ci.auc(roc), coords(roc, 'best') ))
    roc$model = 'in-silico'
    roc$is = basename(split)
    rocs[[basename(split)]] = roc

    rm(dysplasia.df, labels)
  }
  #colnames(rocs) = c('auc.ci.min','auc','auc.ci.max','threshold','specificity','sensitivity')

  insilico.rocs=rocs
  insilico.perf=performance
  save(insilico.rocs, insilico.perf, file=paste(cache.dir, 'insilico_splits.Rdata', sep='/'))
}

load(grep('all.*Rdata',list.files(cache.dir, full.names = T), value=T), verbose=F)
performance = rbind(performance, performance.at.1se[[select.alpha]])
performance$split = c(basename(ins_splits), 'discovery')
rm(plots,coefs,performance.at.1se,models,cv.patient,dysplasia.df,labels)

load(grep('loo\\.Rdata', list.files(cache.dir,full.names = T), value=T), verbose=F)
rm(plots, performance.at.1se,coefs,nzcoefs,fits)

predictions = do.call(rbind, pg.samp)

preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
roc$model = 'discovery'
roc$is = 'discovery'

#rocs = rbind(rocs,c(ci.auc(roc), coords(roc, 'best')) )

#rocs$split = c(basename(ins_splits),'discovery')

```

Original "discovery" cohort vs insilico generated discovery cohorts.  Original is noted in red, we appear to have gotten lucky in our patient selection.

## Model Performance

```{r echo=F, fig.align='left'}
performance$name = 'insilico'
performance$name[which(performance$split == 'discovery')] = 'discovery'
m = melt(performance, measure.vars = 'mean')
ggplot(m, aes(variable,value)) + ylim(0,1) +
  geom_boxplot(fill='mintcream', color='grey55') + geom_jitter(width=0.2, aes(color=name), size=1.5) + 
  geom_text(data=subset(m, value %in% range(value)), aes(label=round(value,2)), vjust=0.02, hjust=0.2) +
  plot.theme + labs(x='CV Models',y='Mean model performance',title='Mean performance for CV model', subtitle='20 insilico 20:69 splits')
```

Median performance `r median(performance$mean)`, SD `r sd(performance$mean)`

## AUC Plot

```{r echo=FALSE, fig.align='left'}
rocs[['discovery']] = roc

mrplot <- function(rocList) {
  aucs = do.call(rbind, lapply(rocList,function(r) 
    cbind.data.frame('AUC'=r$auc*100,'model'=r$model,'Specificity'=pROC::coords(r,'best')[['specificity']], 'Sensitivity'=pROC::coords(r,'best')[['sensitivity']])))
  aucs$y = aucs$Sensitivity
  aucs$x = aucs$Specificity

  df = do.call(rbind, lapply(rocList, function(r) 
    cbind.data.frame('Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'model'=r$model, 'is'=r$is) 
  ))

  ggplot(df, aes(Specificity, Sensitivity,color=is)) + geom_line(size=1) + 
    geom_line(data=subset(df, model == 'discovery'), color='red', size=1) +
    scale_x_reverse() +
    scale_color_brewer(palette='Blues') +
    labs(title=title, x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
    theme(legend.position = 'none')
}

mrplot(rocs)

```
Median AUC `r median(rocs$auc)`, SD `r sd(rocs$auc)`

### 2018-03-06 Misclassified.Rmd
adding to pipeline

---
title: "Untitled"
author: "Sarah Killcoyne"
date: "2/21/2018"
output: html_document
---

```{r setup, include=FALSE}
library(GenomicRanges)
library(ggplot2)
library(ggrepel)
library(GGally)
library(plyr) 
library(pander)
library(reshape2)
library(gridExtra)
library(plyr)
library(dplyr)

source('lib/load_patient_metadata.R')
source('lib/cv-pt-glm.R')
source('lib/common_plots.R')
source('lib/data_func.R')

assign.risk <- function(pred) {
  cuts = seq(0,1,0.1)

  pred$quants = with(pred, cut(Prediction, breaks=cuts))
  qt = as.data.frame.matrix(table(pred$quants, pred$Status))
  qt$quant = rownames(qt)

  ft.fun<-function(NP,P) {
   f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
   cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
  }

  qt = cbind(qt, pred %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
  pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 
    'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )

  pred.confidence[c('r1','r2')] = NA
  for (i in 1:(length(cuts)-1)) pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)

  pred.confidence$Risk = 'Moderate'
  pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc < .5)) ] = 'Low'
  pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'

  pred.confidence = bind_cols(pred.confidence, 
                              data.frame(ci.low=qbeta(0.025, shape1=pred.confidence$P+.5, shape2 = pred.confidence$NP+.5),
                                        ci.high=qbeta(0.975, shape1=pred.confidence$P+.5, shape2 = pred.confidence$NP+.5)))

  #grid.arrange( tableGrob(annotation.table, rows=NULL, theme=tt3) )
  pred.confidence$Risk = factor(pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)

  riskCols = brewer.pal(11, "RdYlBu")[c(10, 5, 1)]
ggplot(pred.confidence, aes(mn, perc)) + 
  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) +
  scale_fill_manual(values=riskCols) +
  geom_point() + #geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration') +
  scale_color_manual(values=riskCols) + plot.theme + theme(legend.position = 'none', text=element_text(size=10))

  pred = merge(pred, pred.confidence, by='quants')
  pred = arrange(pred, Patient, Endoscopy.Year, PID, Pathology)
  return(pred)
}

data = '~/Data/Ellie'
info.dir = '~/Data/Ellie/QDNAseq'

patient.file = list.files(info.dir, pattern='All_patient_info.xlsx', recursive=T, full.names=T)
demo.file = list.files(info.dir, pattern='Demographics_full.xlsx', recursive=T, full.names=T)

if (length(patient.file) != 1 | length(demo.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", info.dir))

all.patient.info = read.patient.info(patient.file, demo.file, set='all')$info
head(all.patient.info)

patient.info = all.patient.info
#patient.info = subset(all.patient.info, Set == 'Training')

patient.info = arrange(patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
sum.patient.data = summarise.patient.info(patient.info)

```

# Why are the patient sin the validaiton cohort driving down prediction?

1. Look at the patients (and samples) that are misclassified in the LOO predictions within the 'D' only set.
  a. Do they stand out for anything else like very quiet genomes?
  b. Batches?

2. Look at coef within the 'D' model, and in the 'all' model
  a. Which coef change?
  b. Which samples change?

```{r disc_data, echo=F}
disc.dir = paste(data, 'Analysis/5e6_arms_disc', sep='/')
file = paste(disc.dir, 'model_data.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))

load(file, verbose=T)
disc.dysplasia.df = dysplasia.df
disc.labels = labels

length(labels) == nrow(patient.info)

file = paste(disc.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=T)
disc.coefs = coefs$`0.9`

file = paste(disc.dir, 'loo.Rdata', sep='/')
load(file, verbose=T)

disc.pred = do.call(rbind, pg.samp)
disc.performance = performance.at.1se
names(disc.performance) = names(pg.samp)

m = melt(as.matrix(disc.performance))
m$Var2 = 'LOO'
ggplot(m, aes(Var2, value, group=Var2)) + geom_boxplot(outlier.colour = NA) + 
  geom_jitter(width=0.1) + labs(x='', y='Model classification performance', title="LOO, 'D'")
disc.loo.coefs = nzcoefs

rm(dysplasia.df, fits, plots, pg.samp, allDf, nzcoefs, coefs)
```

## Misclassified in 'Discovery'

1. Look at the patients (and samples) that are misclassified in the LOO predictions within the 'D' only set.
  a. Do they stand out for anything else like very quiet genomes?
  b. Batches?

```{r, echo=T}

preds = disc.pred[c('Status','Prediction')]
roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
cutoff = round(pROC::coords(roc, 'best')[['threshold']], 2)

ggplot(disc.pred, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1), show.legend = F) + 
  scale_fill_distiller(palette = 'RdYlBu', name='P(P)') + labs(title='LOO Prediction', y='n Samples', x='Probability') + plot.theme

disc.pred = assign.risk(disc.pred)

prg = subset(disc.pred, Status == 'P' & Risk == 'High')
non = subset(disc.pred, Status == 'NP' & Risk == 'Low')

## Low
falseNeg = subset(disc.pred, Risk == 'Low' & Status == 'P')
table(falseNeg$Hospital.Research.ID)/table(subset(patient.info, Hospital.Research.ID %in% unique(falseNeg$Hospital.Research.ID))$Hospital.Research.ID )

# Mod
mods = subset(disc.pred, Risk == 'Moderate')
table(falseNeg$Hospital.Research.ID)/table(subset(patient.info, Hospital.Research.ID %in% unique(falseNeg$Hospital.Research.ID))$Hospital.Research.ID )

# High
falsePos = subset(disc.pred, Risk == 'High'  & Status == 'NP')
table(falsePos$Hospital.Research.ID)/table(subset(patient.info, Hospital.Research.ID %in% unique(falsePos$Hospital.Research.ID))$Hospital.Research.ID )

intersect(mods$Hospital.Research.ID, falseNeg$Hospital.Research.ID)
intersect(mods$Hospital.Research.ID, falsePos$Hospital.Research.ID)

```

Patients with the highest false pos/neg rate: AHM0121, AHM0277, AD0252, AD0591

```{r all_data, echo=F}
all.dir = paste(data, 'Analysis/5e6_arms_all', sep='/')
file = paste(all.dir, 'model_data.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))

load(file, verbose=T)
all.dysplasia.df = dysplasia.df
all.labels = labels

file = paste(all.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=T)
all.coefs = coefs$`0.9`
all.performance = performance.at.1se$`0.9`

file = paste(all.dir, 'loo.Rdata', sep='/')
load(file, verbose=T)
all.pred = do.call(rbind, pg.samp)
all.performance.loo = performance.at.1se
names(all.performance.loo) = names(pg.samp)

all.loo.coefs = coefs

m = melt(as.matrix(all.performance.loo))
m$Var2 = 'LOO'
ggplot(m, aes(Var2, value, group=Var2)) + geom_boxplot(outlier.colour = NA) + 
  geom_jitter(width=0.1) + labs(x='', y='Model classification performance', title="LOO, 'All'")

rm(coefs, nzcoefs, dysplasia.df, fits, plots, pg.samp, allDf)
```
# What are their coefficients like?

Look at coef within the 'D' model, and in the 'all' model
  a. Which coef change?
  b. Which samples change?

```{r}
# Coefs not in the 'D' model
setdiff(rownames(disc.coefs), rownames(all.coefs))
ucf = union(rownames(disc.coefs), rownames(all.coefs))

coefs = data.frame(matrix(ncol=2, nrow=length(ucf), dimnames=list(ucf, c('D','A'))))
coefs[rownames(disc.coefs),'D'] = disc.coefs[,1]
coefs[rownames(all.coefs),'A'] = all.coefs[,1]
#coefs[is.na(coefs)] = 0

abs(coefs$D) - abs(coefs$A)

```

# Re run cv w/o those patients?

```{r}
file = list.files(data, pattern='patient_folds.tsv', recursive=T, full.names=T)[1]
if (length(file) == 1) {
  message(paste("Reading folds file", file))
  sets = read.table(file, header=T, sep='\t')
} else {
  sets = create.patient.sets(patient.info[c('Hospital.Research.ID','Samplename','Status')], folds, splits, 0.2)  
}

rows = which(rownames(all.dysplasia.df) %in% subset(patient.info, Hospital.Research.ID %in% c('AHM0121', 'AHM0277', 'AD0252','AD0591'))$Samplename)

df = all.dysplasia.df[-rows,]
labels = all.labels[rownames(df)]

a = 0.9; nl=1000; folds=10; splits=5
fit0 <- glmnet(df, labels, alpha=a, nlambda=nl, family='binomial', standardize=F)    
autoplot(fit0) + theme(legend.position = 'none')

cv = crossvalidate.by.patient(x=df, y=labels, lambda=fit0$lambda, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', opt=-1, standardize=F)

lambda.opt = cv$lambda.1se
subset(cv$lambdas, lambda == lambda.opt)$mean

pm = predict(fit0, newx=all.dysplasia.df[rows,], s=cv$lambda.1se, type='response')
colnames(pm) = c("Prediction")
pred = merge(pm, patient.info, by.x='row.names', by.y='Samplename')

# Prediction for those three don't improve
table(subset(pred, Status == 'P')$Prediction > 0.5)

table(subset(pred, Status == 'NP')$Prediction < 0.5)

```

# Do it with the 'D' model only

```{r}
`%nin%` = Negate('%in%')

exsamples = subset(patient.info, Hospital.Research.ID %in% c('AHM0121', 'AHM0277', 'AD0252','AD0591'))$Samplename
samples = subset(patient.info, Hospital.Research.ID %nin% c('AHM0121', 'AHM0277', 'AD0252','AD0591'))$Samplename

df = disc.dysplasia.df[intersect(samples, rownames(disc.dysplasia.df)),]
labels = disc.labels[ rownames(df) ]
length(labels) == nrow(df)

a = 0.9; nl=1000; folds=10; splits=5
fit1 <- glmnet(df, labels, alpha=a, nlambda=nl, family='binomial', standardize=F)    
autoplot(fit1) + theme(legend.position = 'none')

cv = crossvalidate.by.patient(x=df, y=labels, lambda=fit1$lambda, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit1, select='deviance', opt=-1, standardize=F)

pm = predict(fit1, newx=disc.dysplasia.df[exsamples,], s=cv$lambda.1se, type='response')
colnames(pm) = c("Prediction")
pred1 = merge(pm, patient.info, by.x='row.names', by.y='Samplename')

# Prediction for those three don't improve
table(subset(pred1, Status == 'P')$Prediction > 0.5)
table(subset(pred1, Status == 'NP')$Prediction < 0.5)

```

## What about the validation patients?

```{r}
vpi = subset(patient.info, Set == 'Test')

df = all.dysplasia.df[vpi$Samplename,]
labels = all.labels[rownames(df)]

pm = predict(fit1, newx=df, s=cv$lambda.1se, type='response')
colnames(pm) = c("Prediction")
pred2 = merge(pm, all.patient.info, by.x='row.names', by.y='Samplename')

table(subset(pred2, Status == 'P')$Prediction > min(subset(disc.pred, Risk == "High")$r1) )
table(subset(pred2, Status == 'NP')$Prediction < max(subset(disc.pred, Risk == "Low")$r2))

roc = pROC::roc(Status ~ Prediction, data= pred2[c('Status','Prediction')], auc=T, ci=T, of='thresholds')
roc.plot(roc)

```

## LOO better trained w/o those three?
```{r}
sets = create.patient.sets(patient.info[c('Hospital.Research.ID','Samplename','Status')], folds, splits, 0.2)

pi = subset(patient.info, Hospital.Research.ID %nin% c('AHM0121', 'AHM0277', 'AD0252','AD0591') )

pg.samp = lapply(unique(pi$Hospital.Research.ID), function(pt) {
  info = subset(pi, Hospital.Research.ID == pt)
  info$Prediction = NA
  info$Prediction.Dev.Resid = NA
  info$OR = NA
  info$PID = unlist(lapply(info$Path.ID, function(x) unlist(strsplit(x, 'B'))[1]))
  return(info)
})
names(pg.samp) = unique(pi$Hospital.Research.ID)

secf = as.data.frame(all.coefs[,1])
rownames(secf) = rownames(all.coefs)
performance.at.1se = c(); coefs = list(); plots = list(); fits = list(); nzcoefs = list()
  # Remove each patient (LOO)
for (pt in names(pg.samp)) {
    print(pt)
    samples = subset(patient.info, !Hospital.Research.ID %in% c('AHM0121', 'AHM0277', 'AD0252', pt))$Samplename

    train.rows = which(rownames(all.dysplasia.df) %in% samples)
    training = all.dysplasia.df[train.rows,]

    exsamples = subset(patient.info, Hospital.Research.ID == pt)$Samplename
    test = as.matrix(all.dysplasia.df[which(rownames(all.dysplasia.df) %in% exsamples),])
    #if (ncol(test) <= 1) next
    if ( nrow(test) == ncol(all.dysplasia.df) ) test = t(test)

    # Predict function giving me difficulty when I have only a single sample, this ensures the dimensions are the same
    sparsed_test_data <- Matrix(data=0, nrow=ifelse(length(pg.samp[[pt]]$Samplename) > 1, nrow(test), 1),  ncol=ncol(training),
                                dimnames=list(pg.samp[[pt]]$Samplename,colnames(training)), sparse=T)
    for(i in colnames(all.dysplasia.df)) sparsed_test_data[,i] = test[,i]

    # Fit generated on all samples, including HGD
    fitLOO <- glmnet(training, all.labels[train.rows], alpha=a, family='binomial', nlambda=nl, standardize=F) # all patients
    l = fitLOO$lambda

    cv = crossvalidate.by.patient(x=training, y=all.labels[train.rows], lambda=l, a=a, nfolds=folds, splits=splits,
                                  pts=subset(sets, Samplename %in% samples), fit=fitLOO, standardize=F)

    plots[[pt]] = arrangeGrob(cv$plot+ggtitle('Classification'), cv$deviance.plot+ggtitle('Binomial Deviance'), top=pt, ncol=2)

    fits[[pt]] = cv

    if ( length(cv$lambda.1se) > 0 ) {
      performance.at.1se = c(performance.at.1se, subset(cv$lambdas, lambda == cv$lambda.1se)$mean)

      #coef.1se = coef(fitLOO, cv$lambda.1se)[rownames(secf),]

      nzcoefs[[pt]] = as.data.frame(non.zero.coef(fitLOO, cv$lambda.1se))

      coefs[[pt]] = coef(fitLOO, cv$lambda.1se)[rownames(secf),]
      #coef.stability(coef.1se, cv$non.zero.cf)

      logit <- function(p){log(p/(1-p))}
      inverse.logit <- function(or){1/(1 + exp(-or))}

      pm = predict(fitLOO, newx=sparsed_test_data, s=cv$lambda.1se, type='response')
      or = predict(fitLOO, newx=sparsed_test_data, s=cv$lambda.1se, type='link')
      sy = as.matrix(sqrt(binomial.deviance(pm, labels[pg.samp[[pt]]$Samplename])))

      pg.samp[[pt]]$Prediction = pm[,1]
      pg.samp[[pt]]$Prediction.Dev.Resid = sy[,1] 
      pg.samp[[pt]]$OR = or[,1]

    } else {
      warning(paste("Hospital.Research.ID", pt, "did not have a 1se"))
    }
  }

back = do.call(rbind, pg.samp)

hist(back$Prediction)

```

### 2018-03-06 Paper.Rmd
adding to pipeline

@@
 
-ggsave('plots/performance/no_lgd_perf.png', plot=mp$plot, height=5, width=5, units='in', dpi=30)
+ggsave('plots/performance/no_lgd_perf.png', plot=mp$plot, height=5, width=5, units='in', dpi=300)
 
@@
 droc
-ggsave(filename='plots/discovery_roc.png', plot=droc, width=6, height=4, units='in', dpi=300)
+ggsave(filename='plots/auc/discovery_roc.png', plot=droc, width=6, height=4, units='in', dpi=300)
 
@@
 ogjLimits = levels(factor(back$ogj, ordered = T))
+
+df = subset(globalCNdf, Patient == 20 & seg.type != 'arms')
+#df = subset(globalCNdf, Patient == 51 & seg.type != 'arms')
+
+# early/late samples
+#df = subset(globalCNdf, Patient == pt)
+df = merge(globalCNdf, back[,c('Samplename','Endoscopy.Year', 'ogj', 'months.before.final', 'OR','Prediction')], by.x='variable', by.y='Samplename')
+df = arrange(df, chr, start, Endoscopy.Year)
+
+## count arms differently than segs??
+df2 = df %>% group_by(Status, Patient, Endoscopy.Year) %>% summarise( 
+  'max.Pred'=max(Prediction),
+  'max.RR'=max(OR),
+  'max.Path'=max(Pathology),
+  'CNA'=length(which(seg.type != 'arms' & cn %in% c('loss','gain'))),
+  'arms'=length(which(seg.type == 'arms' & cn %in% c('loss','gain'))),
+  'ratio'=length(which(seg.type != 'arms' & cn %in% c('loss','gain')))/length(cn)
+  )
+#as.data.frame(df2)
+#df2 = subset(df2, max.Path %in% c('BE','ID','LGD'))
+
+df2 = df2 %>% group_by(Status, Patient) %>% mutate(
+  'early'= Endoscopy.Year == min(Endoscopy.Year),
+  'late' = Endoscopy.Year == max(Endoscopy.Year)
+)
+df2 = subset(df2, Patient %in% names(which(table(df2$Patient) > 1)))
+
+el = subset(df2, early | late)
+el2 = el %>% group_by(Status, Patient) %>% summarise( diff(arms), diff(CNA) )
+wtA = round(wilcox.test(subset(el2, Status == 'Progressor')$`diff(arms)`, subset(el2, Status == 'Non-Progressor')$`diff(arms)`)$p.value, 3)
+if (wtA < 0.0001) wtA = '< 0.0001'
+#wilcox.test(subset(el2, Status == 'Progressor')$`diff(CNA)`, subset(el2, Status == 'Non-Progressor')$`diff(CNA)`)
+
+p = ggplot(subset(df2, early|late), aes(Endoscopy.Year, arms, color=Patient)) + facet_grid(~Status) +
+  geom_point() +  geom_line(data=subset(df2, early | late)) + theme(legend.position = 'none', text=element_text(size=11)) + 
+  plot.theme + labs(x='Endoscopy Year', y='CN arm values', title='Initial vs final endoscopy', subtitle=paste('p', wtA))
+p
+ggsave(filename='plots/status_hetero.png', plot=p, width=8, height=6, units='in', dpi=300)
+
+wtA = round(wilcox.test(subset(df2, early &  Status == 'Progressor')$arms, subset(df2, early & Status == 'Non-Progressor')$arms)$p.value, 3)
+wtCN = round(wilcox.test(subset(df2, early & Status == 'Progressor')$CNA, subset(df2, early & Status == 'Non-Progressor')$CNA)$p.value, 2)
+
+m = melt(subset(df2, early), measure.vars=c('CNA','arms'), id.vars=c('Status'))
+p = ggplot(m, aes(Status, value, fill=Status)) + facet_wrap(~variable, scales='free_y') + 
+  geom_boxplot(outlier.colour = NA) + geom_jitter(width=0.2) + plot.theme + theme(legend.position = 'none', text=element_text(size=11)) +
+  labs(y='', x='', title='CN values, initial endoscopy', subtitle=paste('Arms p=',wtA, '   CNA p=',wtCN, sep=''))
+p
+ggsave(filename='plots/hetero_initial.png', plot=p, width=6, height=4, units='in', dpi=300)
+
+ggplot(el2, aes(Status,`diff(arms)`)) + geom_boxplot() + geom_jitter(width=0.2)
+ggplot(el2, aes(Status,`diff(CNA)`)) + geom_boxplot() + geom_jitter(width=0.2)
 
@@
   facet_grid(~Status, scales='free_x') + labs(y='Percent of samples', title='Validation patient set') + plot.theme
+
+back.v$Prediction
+
+table(subset(back.v, Status == 'P')$Prediction > 0.6)
+table(subset(back.v, Status == 'NP')$Prediction < 0.3)
+
 ```
@@
   labs(x='', y='AUC', title='AUC with CI') + plot.theme 
+```
+
+## Downsampled Datasets
+
+```{r echo=F, eval=T}
+dirA = '~/Data/Ellie/Analysis/downsampled5G_BEAdjacent'
+dirB = '~/Data/Ellie/Analysis/downsampled5G_NDBE'
+
+file = paste(cache.dir, 'model_data.Rdata', sep='/')
+if (!file.exists(file))
+  stop(paste("Missing data file", file))
+load(file, verbose=T)
+file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
+if (!file.exists(file))
+  stop(paste("Missing data file", file))
+load(file, verbose=T)
+rm(dysplasia.df,coefs,plots,labels)
+
+select.alpha = '0.9'
+fitV = models[[select.alpha]]
+lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']
+
+predA = list.files(dirA, 'predictions.Rdata', full.names = T)
+predB = list.files(dirB, 'predictions.Rdata', full.names = T)
+
+files = c()
+if (length(predA) < 1)
+  files = c(list.files(dirA, pattern='raw', full.names=T))
+if (length(predB) < 1)
+  files = c(files, list.files(dirB, pattern='raw', full.names=T))
+
+if (length(files) > 0) {
+  sampleNames = sub('_raw.txt', '', basename(files))
+  dspred = data.frame(matrix(ncol=2, nrow=length(sampleNames), dimnames=list(sampleNames,c('Prob','RR'))))
+
+  for (n in 1:length(files)) {
+    print(n)
+    print(sampleNames[n])
+
+    rs = read.table(files[n], sep='\t',header=T)
+    if (ncol(rs) < 4)
+      stop("File needs to contain 4 columns: chr, start, end, sample_value")
+
+    segs <- tile.segmented.data(rs, size=5e6, chr.info=chrlen)
+    segM = as.matrix(segs[,-c(1:3)])
+    rownames(segM) = paste(segs$chr, ':', segs$start, '-', segs$end, sep='')
+    segM = t(segM)
+    for (i in 1:ncol(segM)) 
+      segM[,i] = unit.var(segM[,i], z.mean[i], z.sd[i])
+
+    arms <- tile.segmented.data(rs, size='arms', chr.info=chrlen)
+    armsM = as.matrix(arms[,-c(1:3)])
+    rownames(armsM) = paste(arms$chr, ':', arms$start, '-', arms$end, sep='')
+    armsM = t(armsM)
+    for (i in 1:ncol(armsM)) 
+      armsM[,i] = unit.var(armsM[,i], z.arms.mean[i], z.arms.sd[i])
+
+    nrow(armsM) == nrow(segM)
+
+    cx.score = score.cx(segM,1)
+
+    mergedDf = subtract.arms(segM, armsM)
+    mergedDf = cbind(mergedDf, 'cx' = unit.var(cx.score, mn.cx, sd.cx))
+    dim(mergedDf)
+
+    # Predict function giving me difficulty when I have only a single sample, this ensures the dimensions are the same
+    sparsed_test_data <- Matrix(data=0, nrow=nrow(mergedDf),  ncol=ncol(mergedDf),
+                                dimnames=list(rownames(mergedDf),colnames(mergedDf)), sparse=T)
+    for(col in colnames(mergedDf)) sparsed_test_data[,col] = mergedDf[,col]
+
+    preds = predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='response')
+    RR = predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='link')
+
+    dspred[n,] = c(preds[,1],RR[,1])
+    print(dspred[1:n,])
+  }
+} else {
+  load(predB, verbose=T)
+  dspredB = dspred
+
+  load(predA, verbose=T)
+
+  dspred = rbind(dspred, dspredB)
+  dspred$Type = NA
+
+  dspred[sub('_raw.*','',basename(list.files(dirA, pattern='raw', full.names=T))),'Type'] = 'AdjacentBE'
+  dspred[sub('_raw.*','',basename(list.files(dirB, pattern='raw', full.names=T))),'Type'] = 'BEST2 NDBO'
+
+}
+
+ggplot(dspred, aes(Type, RR, fill=Type)) + geom_boxplot() + geom_jitter(width=0.2) +
+  labs(x='', y='Relative Risk', title='Predictions on downsampled patients') + 
+  theme(legend.position = 'none') + plot.theme
+
+adj = subset(dspred, Type == 'AdjacentBE')
+prg = back[,c('Prediction', 'OR', 'Status')]
+prg$Status = ifelse(prg$Status == 'P', 'Progressor','Non-Progressor')
+
+#  subset(back, Status == 'P', select=c('Prediction', 'OR', 'Status'))
+colnames(prg) = colnames(adj)
+m = rbind(adj,prg)
+m$Type = factor(m$Type, levels=c('Non-Progressor','BEST2 NDBO','Progressor','AdjacentBE'), ordered = T)
+
+pr = ggplot(m, aes(Type, RR, fill=Type)) + 
+  geom_jitter(width=0.2) + geom_boxplot() + 
+  labs(x='', y='RR', title='Predicted Relative Risk') + 
+  theme(legend.position = 'none', text=element_text(size=11)) + plot.theme 
+pr
+
+ggsave('plots/adjacentBE.png', plot=pr, width=8, height=8,dpi=300)
+
+#save(dspred, file=paste(dirB, 'predictions.Rdata', sep='/'))
+
 ```
@@
 
-cols = unique(colnames(back), colnames(back.v))
-cols = cols[-grep('quants|pred\\.ratio|se\\(Risk\\)', cols)]
+cols = intersect(colnames(back), colnames(back.v))
+#cols = cols[-grep('quants|pred\\.ratio|se\\(Risk\\)', cols)]
 allB = rbind.data.frame(back[,cols], back.v[,cols])
@@
   summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )
+
+table(subset(second, Status == 'P')$rule)
+table(subset(second, Status == 'NP')$rule)
+
+path.rx<-function(Patient, brks, Pathology) {
+  rules = as.data.frame(matrix(ncol=0, nrow=0))
+  # Consecutive
+  for (i in 1:length(brks)) {
+    #risks = table(Pathology[i:(i+1)])
+    rule = 'None'
+    if ( length(which(grepl('HGD|IMC', Pathology[i:(i+1)]))) > 0  |
+         length(which(grepl('LGD', Pathology[i:(i+1)]))) > 1 ) {
+      rule = 1
+    } else if ( length(which(grepl('LGD', Pathology[i:(i+1)]))) > 0 ) {
+      rule = 2
+    } else if ( length(which(grepl('BE|ID', Pathology[i:(i+1)]))) > 1  ) {
+      rule = 3
+    }
+  rules = rbind(rules, cbind( as.character(brks[i]), as.character(brks[(i+1)]), rule))
+  }  
+  colnames(rules) = c('t1','t2', 'rule')
+  rules$Patient = Patient
+  rules
+}
+
+rules2 = do.call(rbind, lapply(unique(maxP$Patient), function(pt) {
+  r = with(arrange(subset(maxP, Patient == pt), brks), path.rx(pt, brks,Pathology))
+  r$Status = unique(subset(maxP, Patient == pt)$Status)
+  r
+}))
+rules2$t1 = factor(rules2$t1, levels=levels(back$brks), ordered=T)
+rules2$t2 = factor(rules2$t2, levels=levels(back$brks), ordered=T)
+
+second = subset(rules2,  t2 != 'Endpoint') %>% 
+  group_by(Patient, Status) %>% 
+  summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )
+
 table(subset(second, Status == 'P')$rule)
@@
 
-ggplot(rrPath, aes(Status, mean.RR)) + #scale_color_gradientn(colors=myPal, limits=c(min(rrPath[,3:5]), max(rrPath[,3:5])),  name='') +
+rrp = ggplot(rrPath, aes(Status, mean.RR)) + #scale_color_gradientn(colors=myPal, limits=c(min(rrPath[,3:5]), max(rrPath[,3:5])),  name='') +
   facet_wrap(~Pathology, ncol=1) + 
@@
   coord_flip() + plot.theme + labs(y='Mean RR', x='')
+rrp
+
+ggsave('plots/relrisk_path.png', plot=rrp, width=4, height=6,dpi=300)
 
@@
 
-as.data.frame(subset(back, Status == 'P' & OR < 2 & OR > -2, select=c('Status','Patient','Pathology','months.before.final','OR')))
-
-hist(subset(back, Pathology %in% c('BE'))$OR)
+# as.data.frame(subset(back, Status == 'P' & OR < 2 & OR > -2, select=c('Status','Patient','Pathology','months.before.final','OR')))
+# 
+# hist(subset(back, Pathology %in% c('BE'))$OR)
 
@@
 
-ggplot(melt(loopreds[c('Min','Max')]), aes(value)) + facet_wrap(~variable, ncol=2) +
+p = ggplot(melt(loopreds[c('Min','Max')]), aes(value)) + facet_wrap(~variable, ncol=2) +
   geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + 
@@
          title='5-year adjusted risk', subtitle='Adjusted for upper & lower bound: 1-3.5% risk')
+p
+
+ggsave('plots/adj_RR.png', plot=p, width=6, height=6,dpi=300)
 

### 2018-03-06 processing_pipeline/predict_downsampled.R

@@ -1,0 +1,108 @@
+# predict downsampled
+
+library(ggplot2)
+library(Matrix)
+library(glmnet)
+
+source('lib/data_func.R')
+source('lib/load_patient_metadata.R')
+
+chrlen = get.chr.lengths()
+
+dir = '~/Data/Ellie/Analysis/downsampled5G'
+files = list.files(dir, full.names=T)
+
+
+file = '~/Data/Ellie/Analysis/5e6_arms_disc/model_data.Rdata'
+if (!file.exists(file))
+  stop(paste("Missing data file", file))
+load(file, verbose=T)
+file = '~/Data/Ellie/Analysis/5e6_arms_disc/all.pt.alpha.Rdata'
+if (!file.exists(file))
+  stop(paste("Missing data file", file))
+load(file, verbose=T)
+rm(dysplasia.df,coefs,plots,labels)
+
+select.alpha = '0.9'
+fitV = models[[select.alpha]]
+lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']
+
+sampleNames = sub('_raw.txt', '', basename(files))
+dspred = data.frame(matrix(ncol=2, nrow=length(samples), dimnames=list(samples,c('Prob','RR'))))
+
+for (n in 1:length(files)) {
+  print(n)
+  print(sampleNames[n])
+  
+  rs = read.table(files[n], sep='\t',header=T)
+  if (ncol(rs) < 4)
+    stop("File needs to contain 4 columns: chr, start, end, sample_value")
+  
+  segs <- tile.segmented.data(rs, size=5e6, chr.info=chrlen)
+  segM = as.matrix(segs[,-c(1:3)])
+  rownames(segM) = paste(segs$chr, ':', segs$start, '-', segs$end, sep='')
+  segM = t(segM)
+  for (i in 1:ncol(segM)) 
+    segM[,i] = unit.var(segM[,i], z.mean[i], z.sd[i])
+  
+  arms <- tile.segmented.data(rs, size='arms', chr.info=chrlen)
+  armsM = as.matrix(arms[,-c(1:3)])
+  rownames(armsM) = paste(arms$chr, ':', arms$start, '-', arms$end, sep='')
+  armsM = t(armsM)
+  for (i in 1:ncol(armsM)) 
+    armsM[,i] = unit.var(armsM[,i], z.arms.mean[i], z.arms.sd[i])
+  
+  nrow(armsM) == nrow(segM)
+    
+  cx.score = score.cx(segM,1)
+  
[…]

### 2018-03-08 AllSamplesModel.Rmd
...

@@
 library(GGally)
-library(plyr) 
 library(pander)
@@
 pander(
-  as.data.frame(sum.patient.data %>% summarise(
+  as.data.frame(sum.patient.data %>% dplyr::summarise(
   'Total Patients'=length(unique(Patient)),
@@
 ```{r echo=F}
-cohortSummary = as.data.frame(sum.patient.data %>% group_by(Status) %>% summarise(
+cohortSummary = as.data.frame(sum.patient.data %>% group_by(Status) %>% dplyr::summarise(
   'Total Patients'=length(unique(Patient)),
@@
 arm.loc = arrange(rbind(
-  (chr.lengths %>% group_by(chrom) %>% mutate('start'=1, 'end'=chr.cent-cent.gap))[c('chrom','start','end')],
-  (chr.lengths %>% group_by(chrom) %>% mutate('start'=chr.cent+cent.gap, 'end'=chr.length))[c('chrom','start','end')]), chrom)
+  (chr.lengths %>% group_by(chrom) %>% dplyr::mutate('start'=1, 'end'=chr.cent-cent.gap))[c('chrom','start','end')],
+  (chr.lengths %>% group_by(chrom) %>% dplyr::mutate('start'=chr.cent+cent.gap, 'end'=chr.length))[c('chrom','start','end')]), chrom)
 
@@
 # Average cn value
-globalCNdf = globalCNdf %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
+globalCNdf = globalCNdf %>% group_by(Status, start, end) %>% dplyr::mutate( 'mean.value' = value/length(unique(variable)) )
 globalCNdf$mean.value = globalCNdf$value # but not going to use it
@@
 
-gp = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples') + theme(text=element_text(size=12))
-#ggsave(filename='plots/globalCN.png', plot=gp, width=16, height=4.6, units='in', dpi=300)
+gp = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples')
+ggsave(filename='plots/globalCN.png', plot=gp, width=16, height=4.6, units='in', dpi=300)
 gp
 
-gbe = globalBECN.plot + theme(legend.position = 'none') + labs(y='mean adj.seg.', title="Non-Dysplastic Barrett's Samples") + theme(text=element_text(size=12))
+gbe = globalBECN.plot + theme(legend.position = 'none') + labs(y='mean adj.seg.', title="Non-Dysplastic Barrett's Samples") 
+ggsave(filename='plots/globalBECN.png', plot=gbe, width=16, height=4.6, units='in', dpi=300)
 gbe
-#ggsave(filename='plots/globalBECN.png', plot=gbe, width=16, height=4.6, units='in', dpi=300)
-
+
+rm(locs, gp, gbe)
 ```
@@
 
-#ggsave('plots/per_sample_cx.png', plot=pv + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
-
-#ggsave('plots/per_sample_cx_roc.png', plot=roc.plot(rocX, title='Per sample cx') + plot.theme + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
+ggsave('plots/per_sample_cx.png', plot=pv, width=4, height=4, units='in', dpi=300)
+ggsave('plots/per_sample_cx_roc.png', plot=roc.plot(rocX, title='Per sample cx') + plot.theme , width=4, height=4, units='in', dpi=300)
 
@@
 folds = 10; splits = 5 
-file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
-if (file.exists(file)) {
-  message(paste("loading", file))
-  load(file, verbose=T)
-} else {
-  stop(paste("No model file", file, "found"))
-}
+# file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
+# if (file.exists(file)) {
+#   message(paste("loading", file))
+#   load(file, verbose=T)
+# } else {
+#   stop(paste("No model file", file, "found"))
+# }
 
@@
 
-#ggsave('plots/eyebrows.png', plot=do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits')), width=10, height=10, units='in', dpi=300)
-
+ggsave('plots/eyebrows.png', plot=do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits')), width=10, height=10, units='in', dpi=300)
+
+rm(plots, cvs)
 ```
@@
 
-#ggsave('plots/performance/ridge_v_lasso.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)
+ggsave('plots/performance/ridge_v_lasso.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)
 
@@
 #do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2))
-#ggsave('plots/nohgd_eyebrow.png', plot=do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2)), height=10, width=10, units='in', dpi=300)
+ggsave('plots/nohgd_eyebrow.png', plot=do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2)), height=10, width=10, units='in', dpi=300)
 ```
@@
 
-#ggsave('plots/performance/ridge_v_lasso_nohgd.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)
+ggsave('plots/performance/ridge_v_lasso_nohgd.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)
 
@@
                               'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
-                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]])
-                              )
-
-rm(performance.at.1se, coefs, mp)
+                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))
+
+rm(performance.at.1se, coefs, mp, no.hgd.plots)
 ```
@@
 file = paste(cache.dir, 'nolgd.Rdata', sep='/')
-
 if (file.exists(file)) {
-  load(file, verbose=F)
+  load(file, verbose=T)
 } else {
@@
 
-#ggsave('plots/performance/no_lgd_perf.png', plot=mp$plot, height=5, width=5, units='in', dpi=30)
+ggsave('plots/performance/no_lgd_perf.png', plot=mp$plot, height=5, width=5, units='in', dpi=30)
 
@@
 
-rm(performance.at.1se, coefs, mp)
+rm(performance.at.1se, coefs, mp, nolgd.plots)
 ```
@@
 p
-
-#ggsave('plots/performance/loo_perf_nohgd.png', plot=p, width=5, height=5, units='in', dpi=300)
+ggsave('plots/performance/loo_perf_nohgd.png', plot=p, width=5, height=5, units='in', dpi=300)
 
@@
 roc.plot(rocNOHGD, 'No HGD')  
-
 ```
@@
 noHGD.pred.confidence = as.data.frame.matrix(table(pg.sampNOHGD$quants, pg.sampNOHGD$Status))
-noHGD.pred.confidence = cbind(noHGD.pred.confidence, pg.sampNOHGD %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
+noHGD.pred.confidence = cbind(noHGD.pred.confidence, pg.sampNOHGD %>% group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
 noHGD.pred.confidence$quant = rownames(noHGD.pred.confidence)
@@
 ft.fun<-function(NP,P) {
+  print(paste(NP, P))
   f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
-  cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
-}
-
-noHGD.pred.confidence = noHGD.pred.confidence %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )
+  cbind.data.frame('p.value'=round(f$p.value, 5), 'odds.ratio'=round(f$estimate, 4))
+}
+
+noHGD.pred.confidence = noHGD.pred.confidence %>% as.data.frame.matrix %>% group_by(quant) %>% 
+  dplyr::mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )
 
@@
   hgd.pred = hgd.samples %>% group_by(Patient, Hospital.Research.ID, Status, Endoscopy.Year) %>% 
-    summarise( max=max(Pathology), pred=Prediction[which.max(Pathology)] )
+    dplyr::summarise( max=max(Pathology), pred=Prediction[which.max(Pathology)] )
 
@@
 }
-chh = chh %>% rowwise() %>% mutate('max'=custom.max(coef, hazards, ymax))
+chh = chh %>% rowwise() %>% dplyr::mutate('max'=custom.max(coef, hazards, ymax))
 
@@
 p
-
-#ggsave('plots/performance_loo.png', plot=p, width=5, height=5, units='in', dpi=300)
-
-#oddPt = names(pg.samp)[which.max(performance.at.1se)]
-#subset(sum.patient.data, Hospital.Research.ID == oddPt)
-
-#pg.samp[[oddPt]]$Prediction
+ggsave('plots/performance_loo.png', plot=p, width=5, height=5, units='in', dpi=300)
+
+oddPt = names(pg.samp)[which.max(performance.at.1se)]
+subset(sum.patient.data, Hospital.Research.ID == oddPt)
+pg.samp[[oddPt]]$Prediction
 
@@
 
-droc = multi.roc.plot(list(roc), colors=brewer.pal(3,'Greys')[3]) + labs(title='Discovery cohort ROC')
-droc
-#ggsave(filename='plots/discovery_roc.png', plot=droc, width=6, height=4, units='in', dpi=300)
+ggsave(filename='plots/discovery_roc.png', plot=roc.plot(roc), width=6, height=4, units='in', dpi=300)
 
@@
 df = do.call(rbind, lapply(insilico.rocs, function(r) 
-    cbind.data.frame('Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'model'=r$model, 'is'=r$is) 
+    cbind.data.frame('Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'auc'=rep(r$auc, length(r$sensitivities)),'model'=r$model,'is'=r$is) 
 ))
@@
 p
-#ggsave(filename='plots/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)
+ggsave(filename='plots/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)
+
+best = do.call(rbind, lapply(insilico.rocs, function(r) {
+  cbind.data.frame(t(as.data.frame(pROC::coords(r, 'best'))), 'auc'=r$auc)
+}))
+best$model = rownames(best)
+
+rr = ggplot(subset(df, is == 'all' | auc %in% c(range(auc))), aes(Specificity, Sensitivity)) + 
+  geom_line(aes(color=factor(model)), size=1) + 
+  geom_line(data=subset(df, auc == min(auc)), color='darkblue', linetype='dotted') + 
+  geom_line(data=subset(df, auc == max(auc)), color='darkblue', linetype='dotted') +
+  scale_x_reverse() + scale_color_brewer(palette='Paired') + 
+  geom_label(data=subset(best, model == 'all' | auc %in% c(min(auc), max(auc)) ), aes(specificity, sensitivity, label=paste('AUC ',round(auc, 2), sep='\n'))) +
+  labs(title='ROC', x='Specificity (1-FPR)', y='Sensitivity (TPR)')  + plot.theme +
+  theme(legend.position = 'none')
+rr  
+ggsave(filename='plots/roc_CI.png', plot=rr, width=6, height=6, units='in', dpi=300)
 
@@
 roc.plot(p53roc, 'p53 IHC')
+
+ggsave(filename='plots/auc/p53_roc.png', plot=roc.plot(p53roc, 'p53 IHC') + plot.theme, width=6, height=6, units='in', dpi=300)
 ```
@@
 pathroc$model = 'HGD & LGD'
-#ggsave('plots/auc/dysplasia_auc.png', plot=roc.plot(pathroc, title='Dysplasia (HGD & LGD)'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+ggsave('plots/auc/dysplasia_auc.png', plot=roc.plot(pathroc, title='Dysplasia (HGD & LGD)'), width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 hgdpathroc$model = 'HGD Only'
-#ggsave('plots/auc/hgd_auc.png', plot=roc.plot(hgdpathroc, title='HGD Only'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+ggsave('plots/auc/hgd_auc.png', plot=roc.plot(hgdpathroc, title='HGD Only'), width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 ```{r, echo=F, warning=F, message=F,fig.height=4, fig.width=4}
-pendo = do.call(rbind, pg.samp) %>% group_by(Patient, Status, Hospital.Research.ID, PID ) %>% summarise(
-  'max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(PID)
-)
+pendo = do.call(rbind, pg.samp) %>% group_by(Patient, Status, Hospital.Research.ID, PID ) %>% dplyr::summarise(
+  'max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(PID))
 
@@
 rocEndoM$model = 'Per Endoscopy Max'
-#ggsave('plots/auc/maxendo_auc.png', plot=roc.plot(rocEndoM, title='Max per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+ggsave('plots/auc/maxendo_auc.png', plot=roc.plot(rocEndoM, title='Max per Endoscopy') , width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 rocEndoA$model = 'Per Endoscopy Avg'
-#ggsave('plots/auc/avgendo_auc.png', plot=roc.plot(rocEndoA, title='Avg per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+ggsave('plots/auc/avgendo_auc.png', plot=roc.plot(rocEndoA, title='Avg per Endoscopy') , width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 rocLastEndo$model = 'Last endoscopy'
-#ggsave('plots/auc/lastendo_auc.png', plot=roc.plot(rocLastEndo, title='Last endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+ggsave('plots/auc/lastendo_auc.png', plot=roc.plot(rocLastEndo, title='Last endoscopy') , width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 rocFirstEndo$model = 'First endoscopy'
-#ggsave('plots/auc/firstendo_auc.png', plot=roc.plot(rocFirstEndo, title='First Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+ggsave('plots/auc/firstendo_auc.png', plot=roc.plot(rocFirstEndo, title='First Endoscopy') , width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 ```{r echo=F, fig.height=4, fig.width=4}
-
 #ptendo = do.call(rbind, pg.samp) %>% group_by(Patient, Hospital.Research.ID, Status) %>% summarise('max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(Patient))
-
 ptendo = do.call(rbind, lapply(pg.samp, function(pt) {
@@
 rocPtM$model = 'Per Patient Max'
-#ggsave('plots/auc/maxpt_auc.png', plot=roc.plot(rocPtM, title='Max per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+ggsave('plots/auc/maxpt_auc.png', plot=roc.plot(rocPtM, title='Max per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 pROC::coords(rocPtA, 'best')
-#roc.plot(rocPtA)
 rocPtA$model = 'Per Patient Avg'
-#ggsave('plots/auc/avgpt_auc.png', plot=roc.plot(rocPtA, title='Avg per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+ggsave('plots/auc/avgpt_auc.png', plot=roc.plot(rocPtA, title='Avg per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 m = rbind.data.frame(
-  #get.roc.stat(p53roc),
+  get.roc.stat(p53roc),
   get.roc.stat(hgdpathroc),
@@
   get.roc.stat(roc))
-m$model = factor(m$model, levels=c('HGD Only','HGD & LGD','CN Model'), ordered = T)
-
-rocList = list(roc,pathroc,hgdpathroc)
+m$model = factor(m$model, levels=c('p53 IHC', 'HGD Only','HGD & LGD','CN Model'), ordered = T)
+
+rocList = list(roc,pathroc,hgdpathroc, p53roc)
 multi.roc.plot(rocList)
@@
   scale_fill_manual(values=cols, name='') + plot.theme + 
-  theme(legend.position = 'none', axis.text.x = element_text(angle=45, hjust=1)) + labs(x='', y='AUC', title='ROC Comparison') 
+  theme(legend.position = 'none') + labs(x='', y='AUC', title='ROC Comparison') 
 comp
 
-#ggsave('plots/path_p53_cn_auc.png', plot=comp, width=6, height=8, units="in", limitsize=F, dpi = 300)
+ggsave('plots/auc/path_p53_cn_auc.png', plot=comp, width=6, height=8, units="in", limitsize=F, dpi = 300)
 
@@
 
-#ggsave('plots/auc/multi_auc.png', plot=p, height=6, width=5, units='in',dpi=300)
+ggsave('plots/auc/multi_auc.png', plot=p, height=6, width=5, units='in',dpi=300)
 
@@
 transform.by.time<-function(pt, info) {
-  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction', 'OR')] %>% group_by(PID, Samplename) %>% mutate(
+  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction', 'OR')] %>% group_by(PID, Samplename) %>% dplyr::mutate(
     'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12)
@@
 
-qt = cbind(qt, back %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))#, 
-pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 
+qt = cbind(qt, back %>% group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))#, 
+pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% dplyr::mutate( 
   'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )
@@
 
-annotation.table = arrange(pred.confidence %>% group_by(Risk) %>% summarise('n(NP)'=sum(NP), 'n(P)'=sum(P), 'p.value'=format.pval(mean(p.value), 3), 'min'=min(r1), 'max'=max(r2)), min)
+annotation.table = arrange(pred.confidence %>% group_by(Risk) %>% dplyr::summarise('n(NP)'=sum(NP), 'n(P)'=sum(P), 'p.value'=format.pval(mean(p.value), 3), 'min'=min(r1), 'max'=max(r2)), min)
 
@@
   geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration') +
-  scale_color_manual(values=riskCols) + plot.theme + theme(legend.position = 'none', text=element_text(size=10))
+  scale_color_manual(values=riskCols) + plot.theme + theme(legend.position = 'none')
 pcal
 
-#ggsave(filename='plots/pred_calib.png', plot=pcal, width=4, height=4, units='in', dpi=300)
+ggsave(filename='plots/pred_calib.png', plot=pcal, width=3, height=3, units='in', dpi=300)
 
 back = merge(back, pred.confidence[c('quant','Risk')], by.x='quants',by.y='quant')
-back = back %>% group_by(Risk) %>% mutate('se(Risk)'=(sd(Prediction)/sqrt(length(Risk)))*1.96 )
+back = back %>% group_by(Risk) %>% dplyr::mutate('se(Risk)'=(sd(Prediction)/sqrt(length(Risk)))*1.96 )
 
@@
 
-#ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR, width=6, height=5, units='in', dpi=300)
+ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR, width=6, height=5, units='in', dpi=300)
 
@@
 
-#ggsave(filename='plots/pred_hist.png', plot=predhist, width=5, height=5, units='in', dpi=300)
-
-#ggplot(back, aes(OR, Prediction, color=Prediction)) + geom_point(show.legend = F) + scale_color_gradientn(colors = myPal,  name='')  + plot.theme
+ggsave(filename='plots/pred_hist.png', plot=predhist, width=5, height=5, units='in', dpi=300)
+
+ggplot(back, aes(OR, Prediction, color=Prediction)) + geom_point(show.legend = F) + scale_color_gradientn(colors = myPal,  name='')  + plot.theme
 ```
@@
 })
-ggplot(ch, aes(reorder(label, coef), coef)) + geom_point() + 
-geom_errorbar(aes(ymin=coef-jack.se, ymax=coef+jack.se)) + 
-coord_flip() +
+p = ggplot(ch, aes(reorder(label, coef), coef)) + geom_point() + 
+  geom_errorbar(aes(ymin=coef-jack.se, ymax=coef+jack.se)) + coord_flip() +
   plot.theme + labs(x='', title='Jackknifed coefficient CI')
+p
+
+ggsave(filename='plots/feature_jacknife.png', plot=p, width=5, height=6, unit='in', dpi=300)
 
@@
 
-pathTotal = back %>% group_by(Status, Pathology) %>% summarise( 'nPath'=length(Pathology))
+pathTotal = back %>% group_by(Status, Pathology) %>% dplyr::summarise( 'nPath'=length(Pathology))
 pathTotal$Status = ifelse(pathTotal$Status == 'P', 'Progressor','Non-Progressor')
@@
 p53Path = cbind.data.frame(table(subset(back, Status == 'P', select=c('Pathology','p53.Status'))), 'Status'='p53 IHC')
-p53Path = merge( subset(back, Status == 'P' & !is.na(p53.Status)) %>% group_by(Pathology) %>% summarise( 'nPath'=length(Pathology)), p53Path)
+p53Path = merge( subset(back, Status == 'P' & !is.na(p53.Status)) %>% group_by(Pathology) %>% dplyr::summarise( 'nPath'=length(Pathology)), p53Path)
 #p53Path$Risk = with(p53Path, ifelse(p53.Status == 1, 'High','Low'))
@@
 
-pathTotal = riskPath %>% group_by(Status, Pathology) %>% summarise('nPath'=unique(nPath))
+pathTotal = riskPath %>% group_by(Status, Pathology) %>% dplyr::summarise('nPath'=unique(nPath))
 
@@
   scale_fill_manual(values=riskCols) + scale_color_manual(values=c('white','black','white')) +
-geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), position=position_stack(), size=4) +
+  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101.5), position=position_stack(), size=4) +
   geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep=''), color=Risk), position=position_stack(vjust = 0.5), size=5, show.legend = F) +
-facet_grid(~Status) + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom', text=element_text(size=10))
+  facet_grid(~Status, scales='free_x', space = 'free_x') + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
 rp
 
-# x = back %>% group_by(Status, Pathology) %>% summarise( mean(Prediction), 'count'=length(Pathology) )
-# ggplot(x, aes(Pathology, count)) + geom_bar(stat='identity',aes(fill=`mean(Prediction)`)) + 
-#     scale_fill_gradientn(colors = myPal,  name='') + facet_grid(~Status) + plot.theme
-
-#ggsave(filename='plots/pred_by_path.png', plot=rp, width=9, height=7, units='in', dpi=300)
+ggsave(filename='plots/pred_by_path.png', plot=rp, width=8, height=8, units='in', dpi=300)
 
@@
 ```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=6}
-all.preds = back %>% group_by(Status, Patient) %>% summarise( 'Low'=sum(length(which(Risk == "Low"))),
+all.preds = back %>% group_by(Status, Patient) %>% dplyr::summarise( 'Low'=sum(length(which(Risk == "Low"))),
 'Moderate'=sum(length(which(Risk == "Moderate"))),
@@
 
-ggplot(all.preds, aes(Patient, value/n.samples, group=variable, fill=variable)) + 
-  geom_bar(stat='identity', position = position_stack()) +
-  scale_fill_manual(values=riskCols) + 
-  geom_text(aes(y=1.01, label=n.samples)) +
-  facet_wrap(~Status, scales='free_x') + labs(y='Sample ratios', title='Per patient ratio of predictions') + 
-  plot.theme
+# ggplot(all.preds, aes(Patient, value/n.samples, group=variable, fill=variable)) + 
+#   geom_bar(stat='identity', position = position_stack()) +
+#   scale_fill_manual(values=riskCols) + 
+#   geom_text(aes(y=1.01, label=n.samples)) +
+#   facet_wrap(~Status, scales='free_x') + labs(y='Sample ratios', title='Per patient ratio of predictions') + 
+#   plot.theme
 
@@
 
+Take per endoscopy average prediction.
+
+There are a lot more moderate predictions in this model...how does this work with this plot??
+
 ```{r enods, warning=F, message=F, echo=F, fig.height=8, fig.width=6}
-endos = back %>% group_by(Status, months.before.final, brks, PID) %>% summarise(
+
+endos = back %>% group_by(Status, months.before.final, brks, PID) %>% dplyr::summarise(
   'pred.endo'=ifelse(length(which(Risk == 'High')) > 0, 1, 0), 
@@
   'n.samples'=length(PID),
+  'mean.pred'=mean(Prediction),
   'pred.samples'=length(which(Risk == 'High')),
   'max.path'=max(Pathology))
+
+endos$quant =  cut(endos$mean.pred, breaks = cuts)
+endos = merge(endos, pred.confidence[c('quant', 'Risk')])
 
@@
 
-endos.m = subset(endos, Status == 'P') %>% group_by(brks) %>% summarise( 
-'pred.r'=sum(pred.endo)/sum(n.endos),
-'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos))
+endos.m = subset(endos, Status == 'P') %>% group_by(brks) %>% dplyr::summarise( 
+  'pred.r'=sum(pred.endo)/sum(n.endos),'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos))
 
@@
 labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))
-tmp = tmp %>% group_by(brks) %>% summarise( 'mean(P)'=mean(Prediction),'sem(P)'=sd(Prediction)/sqrt(length(brks)) )
+tmp = tmp %>% group_by(brks) %>% dplyr::summarise( 'mean(P)'=mean(Prediction),'sem(P)'=sd(Prediction)/sqrt(length(brks)) )
 
@@
 rp
-#ggsave(filename='plots/pred_endo.png', plot=rp, width=9, height=7, units='in', dpi=300)
+ggsave(filename='plots/pred_endo.png', plot=rp, width=9, height=7, units='in', dpi=300)
 
 #endos$months.before.final = endos$months.before.final*-1
-#me = endos %>% group_by(Patient) %>% summarise('max.endo' = max(i), 'months' = min(months.before.final))
+#me = endos %>% group_by(Patient) %>% dplyr::summarise('max.endo' = max(i), 'months' = min(months.before.final))
 ```
@@
 
-  summary.t = summary.t %>% group_by(Pathology) %>% mutate( 'High.r'=round(High/sum(High,Low,Moderate),3), 
+  summary.t = summary.t %>% group_by(Pathology) %>% dplyr::mutate( 'High.r'=round(High/sum(High,Low,Moderate),3), 
                                                             'Low.r'=round(Low/sum(High,Low,Moderate),3),
@@
 ```{r, echo=F, message=F, warning=F, width=26, height=8}
-back = back %>% group_by(Patient) %>% mutate( 'pred.ratio'=round(sum(Prediction)/length(Prediction),2))
+back = back %>% group_by(Patient) %>% dplyr::mutate( 'pred.ratio'=round(sum(Prediction)/length(Prediction),2))
 
@@
 
-ggsave('pred_heatmap_ALL.png', plot=grid.arrange(p2,p1,nrow=2), width=52, height=16, units="in", limitsize=F, dpi = 300)
-#ggsave('plots/npg_heatmap.png', plot=p1, width=52, height=8, units="in", limitsize=F, dpi = 300)
-#ggsave('plots/pg_heatmap.png', plot=p2, width=52, height=8, units="in", limitsize=F, dpi = 300)
-#ggsave('plots/risk_legend.png', plot=grid.arrange(riskLegend), width=2,height=4, dpi=300)
-#ggsave('plots/risk_legend_long.png', plot=grid.arrange(riskLegendLong), width=8,height=2, dpi=300)
-```
+ggsave('plots/pred_heatmap_ALL.png', plot=grid.arrange(p2,p1,nrow=2), width=52, height=16, units="in", limitsize=F, dpi = 300)
+ggsave('plots/npg_heatmap.png', plot=p1, width=52, height=8, units="in", limitsize=F, dpi = 300)
+ggsave('plots/pg_heatmap.png', plot=p2, width=52, height=8, units="in", limitsize=F, dpi = 300)
+ggsave('plots/risk_legend.png', plot=grid.arrange(riskLegend), width=2,height=4, dpi=300)
+ggsave('plots/risk_legend_long.png', plot=grid.arrange(riskLegendLong), width=8,height=2, dpi=300)
+```

### 2018-03-08 Paper.Rmd
...

@@
 cnLegend = get.legend(globalCN.plot)
-#globalCN.plot = globalCN.plot + theme(legend.position = 'none')
+#globalCN.plot = 
+globalCN.plot + theme(legend.position = 'none')
 
@@
 
-gp = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples') + theme(text=element_text(size=12))
+gp = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples')# + theme(text=element_text(size=12))
 ggsave(filename='plots/globalCN.png', plot=gp, width=16, height=4.6, units='in', dpi=300)
 
-gbe = globalBECN.plot + theme(legend.position = 'none') + labs(y='mean adj.seg.', title="Non-Dysplastic Barrett's Samples") + theme(text=element_text(size=12))
+gbe = globalBECN.plot + theme(legend.position = 'none') + labs(y='mean adj.seg.', title="Non-Dysplastic Barrett's Samples") 
 ggsave(filename='plots/globalBECN.png', plot=gbe, width=16, height=4.6, units='in', dpi=300)
@@
 unadjRR
+
+ggplot(back, aes(Status, OR, color=OR)) + 
+  geom_boxplot(outlier.colour=NA) + geom_jitter(width=0.3, aes(shape=Pathology)) + 
+  scale_color_gradientn(colors = myPal,  name='') + plot.theme
 

### 2018-03-08 PredictingProgression/lib/data_func.R

@@ -155,19 +155,20 @@ score.cx <- function(df, MARGIN) {
   return(tiles)
 }
 
+
 # Presumes that the segmented data has already been filtered for num of probes, sd etc
 tile.segmented.data<-function(data, size=5e6, chr.info=NULL) {
   
   if (!is.tibble(data))
     data = as_tibble(data)
   
-  descCols = sort(union(grep('chr|arm|start|end|probes', colnames(data)), which(!sapply(data, is.numeric))))
+  descCols = sort(union(grep('chr|arm|start|end|probes', colnames(data),ignore.case=T), which(!sapply(data, is.numeric))))
   dataCols = c((descCols[length(descCols)]+1):ncol(data))
   
-  chrCol = grep('chr',colnames(data),value=T)
-  armCol = grep('arm',colnames(data),value=T)
-  startPos = grep('start',colnames(data),value=T)
-  endPos = grep('end',colnames(data),value=T)
+  chrCol = grep('chr',colnames(data),ignore.case=T,value=T)
+  #armCol = grep('arm',colnames(data),ignore.case=T,value=T)
+  startPos = grep('start',colnames(data),ignore.case=T,value=T)
+  endPos = grep('end',colnames(data),ignore.case=T,value=T)
   
   data = data[which(!data[[chrCol]] %in% c('X','Y')),]
   x1 = data[,c(chrCol, startPos, endPos, colnames(data)[dataCols])]

### 2018-03-08 PredictingProgression/ui.R

@@ -19,21 +19,26 @@ shinyUI(fluidPage(
   titlePanel("Barrett's Progression Prediction"),
   sidebarLayout(
     sidebarPanel(
-      fileInput("file", "Choose TSV File",
+      fileInput("file", "Choose Caveman copy number or segmented copy number file:",
                 accept = c(
                   "text/tsv",
-                  "text/tab-separated-values,text/plain",
-                  ".tsv", ".txt")
+                  "text/csv",
+                  "text/tab-separated-values,comma-separated-values,text/plain",
+                  ".csv", ".tsv", ".txt")
       ),
       tags$hr(),
-      checkboxInput("header", "Header", TRUE)
+      checkboxInput("header", "Header", TRUE),
+      # Input: Select separator ----
+      checkboxInput("norm", "Model provided norm", TRUE)
+
     ),
     
     
     
     mainPanel(
+      tableOutput("contents"),
       plotOutput("plot")
-      #tableOutput("contents")
+      
     )
   )
   

### 2018-03-08 lib/common_plots.R

@@ -7,15 +7,26 @@ get.legend<-function(a.gplot){
   return(legend)
 }
 
-plot.theme = theme(text=element_text(size=14, face='bold'), panel.background=element_blank(), strip.background =element_rect(fill="grey88"),  
-                   strip.text = element_text(size=14, face='bold'), 
+plot.theme = theme(text=element_text(size=12, face='bold'), panel.background=element_blank(), strip.background =element_rect(fill="white"),  
+                   strip.text = element_text(size=12, face='bold'), 
                    axis.line=element_line(color='black'), panel.grid.major=element_line(color='grey90'),
                    panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing = unit(0.1, 'lines')  ) 
 
 cn.mtn.plot<-function(df, label, type='bar') {
-  low = c( min(df$mean.value), quantile(subset(df, mean.value < -1)$mean.value, probs=c(0.25, 0.75)), -1)
-  high = c(1, quantile(subset(df, mean.value > 1)$mean.value, probs=c(0.25, 0.75)), max(df$mean.value))
-  df$cuts = cut(df$mean.value, breaks=c(low,high), include.lowest = T, ordered_result = T)
+  df = arrange(df, chr, start)
+  rows = which(df$mean.value > 1 | df$mean.value < -1)
+  
+  #low = c( min(df$mean.value), quantile(subset(df, mean.value < -1)$mean.value, probs=c(0.25, 0.75)), -1)
+  #high = c(1, quantile(subset(df, mean.value > 1)$mean.value, probs=c(0.25, 0.75)), max(df$mean.value))
+
+  df$cuts = cut(df$mean.value, breaks=c(min(df$mean.value), -1,0,1, max(df$mean.value)), include.lowest = T, ordered_result = T)
+  medDf = df[-rows,]
+  rangeDf = df[rows,]
+  
+  medDf$cuts = factor('-1-1')
+  rangeDf$cuts = factor(ifelse(rangeDf$mean.value > 1, '> 1', '< -1'))
+
+  #df$cuts = cut(df$mean.value, breaks=c(low,high), include.lowest = T, ordered_result = T)
   
   p = ggplot(df, aes(x=chr.length)) + facet_grid(Status~chr, scales='free_x', space='free_x', labeller=label) 
   
@@ -24,18 +35,22 @@ cn.mtn.plot<-function(df, label, type='bar') {
[…]

### 2018-03-08 AllSamplesModel.Rmd
...

@@
 
+oddPt = names(pg.samp)[which.min(performance.at.1se)]
+
 # Not the pt with the most samples or endos
 #sum.patient.data[which.max(sum.patient.data$total.samples),]
+```
+
+## All LOO
+
+```{r echo=F, eval=F}
+
+files = list.files('~/Data/Ellie/Analysis/5e6_arms_LOO', full.names = T)
+
+sum.patient.data$Hospital.Research.ID[1]
+
+pg.samp = list()
+performance = vector(length=nrow(sum.patient.data), mode='numeric')
+names(performance) = sum.patient.data$Hospital.Research.ID
+
+for (pt in sum.patient.data$Hospital.Research.ID) {
+  ptF = grep(pt, files, value=T)
+  load(grep('loo\\.Rdata', ptF, value=T), verbose=T)
+  #load(grep('model_data\\.Rdata', ptF, value=T), verbose=T)
+  performance[pt] = performance.at.1se
+  pg.samp[[pt]] = info
+}
+rm(plots,coefs,nzcoefs,fits,info)
+
+load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/PR1_HIN_042_loo.Rdata", verbose=T)
+load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/PR1_HIN_042_model_data.Rdata", verbose=T)
+
+load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/AHM0320_loo.Rdata", verbose=T)
+load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/AHM0320_model_data.Rdata", verbose=T)
+
+fitLOO <- glmnet(dysplasia.df, labels[rownames(dysplasia.df)], alpha=0.9, family='binomial', nlambda=500, standardize=F)
+
+pm = predict(fitLOO, newx=allDf, s=fits$lambda.1se, type='response')
+or = predict(fitLOO, newx=allDf, s=fits$lambda.1se, type='link')
+
+hist(pm)
+hist(or)
+
+pm[subset(patient.info, Hospital.Research.ID == 'AHM0320')$Samplename,]
+or[subset(patient.info, Hospital.Research.ID == 'AHM0320')$Samplename,]
+
+sort(performance)
+
+pg.samp[which.max(performance)]
+
+pg.samp[which.min(performance)]
+
+back = do.call(rbind, pg.samp)
+
+subset(sum.patient.data, Status == 'NP' & total.endos <=2)
+
+pm[pg.samp[['AD0309']]$Samplename,]
+pg.samp[['AD0309']]$Prediction
+
 ```
@@
 ft.fun<-function(NP,P) {
-  f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
-  cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
+  f = chisq.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
+  cbind.data.frame('p.value'=f$p.value)
+  #cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
 }
@@
 pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% dplyr::mutate( 
-  'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )
+  'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value,  'conf'=ifelse(p.value < 0.05, '*', '') )
 
@@
   geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
-  geom_point() + #geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
+  geom_point() + geom_abline(slope=1, intercept=0, color='grey39', linetype='dotted') + #geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
   geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration') +
@@
 
+ggplot(back, aes(Status, OR, color=OR)) + 
+  geom_boxplot(outlier.colour=NA) + geom_jitter(width=0.3, aes(shape=Pathology)) + 
+  scale_color_gradientn(colors = myPal,  name='') + plot.theme
+
 predhist = ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) +
@@
 ```{r jacknife, echo=T, warning=F, message=F, fig.height=10, fig.width=12}
-
 coefs.per.pt = data.frame(matrix(ncol=length(pg.samp), nrow=nrow(ch), dimnames=list( ch$label, names(pg.samp))))
@@
 ```
+
+# Chr17
+
+## p arm
+```{r, echo=T}
+
+p11 = subset(globalCNdf, chr == 17 &  start == 1 & seg.type == 'arms')
+p11Loss = subset(p11, value < -1)
+table(p11Loss$Status)
+
+p11path = table(subset(patient.info, Samplename %in% p11Loss$variable)$Pathology)
+
+p11path['BE']/sum(p11path)
+sum(p11path[c('HGD','IMC')])/sum(p11path)
+
+table(p11Loss[c('Patient', 'Pathology')])
+
+# Samples with a loss vs all Progressor samples
+length(unique(p11Loss$variable))/length(unique(subset(globalCNdf, Status == 'Progressor')$variable))
+
+p11Loss = unique(p11Loss[,c('variable', 'Status','Hospital.Research.ID', 'Patient', 'Pathology')])
+length(unique(p11Loss$Patient))/length(subset(sum.patient.data, Status == 'P')$Patient)
+# Of those where we see loss of p arm, p53 IHC shows an "abnormal" result 33% of the time
+
+table(subset(patient.info, Samplename %in% p11Loss$variable, select='p53.Status'))/length(p11Loss$variable)
+
+(table(subset(patient.info, Samplename %in% p11Loss$variable, select=c('p53.Status', 'Pathology'))))
+
+pi = patient.info
+pi$p11Loss = as.numeric(pi$Samplename %in% p11Loss$variable)
+
+plot(table(p11=pi$p11Loss, ihc=pi$p53.Status))
+
+cor.test(pi$p11Loss, (as.numeric(pi$p53.Status)-1))
+
+# So why might there be loss but no corresponding aberrant staining?
+table(subset(pi, p11Loss == 1 & p53.Status == 0)$Pathology)
+table(subset(pi, p11Loss == 1 & p53.Status == 1)$Pathology)
+
+```
+
+## Top Features
+```{r, echo=T}
+m = data.frame(matrix(nrow=nrow(chh), ncol=0))
+
+for (i in 1:nrow(chh)) {
+  feat = chh[i,]
+
+  seg = subset(globalCNdf, chr == as.character(feat$chr) & start == feat$start & end == feat$end & seg.type != 'arms')  
+  if(feat$end-feat$start > 5e6)
+    seg = subset(globalCNdf, chr == as.character(feat$chr) & start == feat$start & end == feat$end & seg.type == 'arms')
+
+  segGL = subset(seg, value < -1)
+  if(feat$coef > 0)
+    segGL = subset(seg, value > 1)
+
+  m[i, 'n.samples'] = length(unique(segGL$variable))
+  m[i,'ratio.samples.P'] = length(unique(segGL$variable))/length(unique(subset(globalCNdf, Status == 'Progressor')$variable))
+  m[i, 'ratio.NP'] = table(segGL$Status)[1]/sum(table(segGL$Status))
+  m[i, 'ratio.P'] = table(segGL$Status)[2]/sum(table(segGL$Status))
+
+  segpath = table(subset(patient.info, Samplename %in% segGL$variable)$Pathology)
+
+  m[i,'BE'] = segpath['BE']/sum(segpath)
+  m[i,'HGD'] = sum(segpath[c('HGD','IMC')])/sum(segpath)
+}
+rownames(m) = chh$label
+
+range(m$ratio.samples.P)
+summary(m$ratio.samples.P)
+subset(m, ratio.samples.P == max(ratio.samples.P))
+subset(chh, label == rownames(subset(m, ratio.samples.P == max(ratio.samples.P))))
+subset(chh, label == rownames(subset(m, ratio.samples.P == min(ratio.samples.P))))
+#table(cdkGain[c('Patient', 'Pathology')])
+
+cor.test(chh$hazards, m$n.samples)
+
+# Samples with a gain vs all samples
+
+```
+One way of looking at the p53 prediction (e.g. Adam Bass) is to see how predictive the p-arm loss in chr17 is.
+```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=10}
+ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
+ccgenes = ccgenes[-grep('^X|Y', ccgenes$Genome.Location),]
+#ccgenes = ccgenes[-which(ccgenes$Mutation.Types == 'T'),] # drop the translocations
+#ccgenes = subset(ccgenes, !grepl('^X|Y', Genome.Location))
+#loc = unlist(strsplit(ccgenes$Genome.Location, ':|-'))
+loc = do.call(rbind.data.frame, strsplit(ccgenes$Genome.Location, ':|-'))
+colnames(loc) = c('chr','start','end')
+loc[] = lapply(loc[], function(x) as.numeric(as.character(x)))
+
+ccgenes = cbind(ccgenes, loc)
+ccgenes = ccgenes[c('Gene.Symbol','Genome.Location', 'Role.in.Cancer', 'Mutation.Types',colnames(loc))]
+#ccgenes = subset(ccgenes, grepl('\\w+',Role.in.Cancer))
+
+genes = cbind(subset(ccgenes, grepl('TSG|oncogene', Role.in.Cancer) & Mutation.Types != 'T', select=c('Gene.Symbol','Genome.Location', colnames(loc))), 'value'=min(globalCNdf$mean.value)+2*sd(globalCNdf$mean.value), 'Row.names'='', 'Patient'='', 'selected'=as.factor(0))
+
+genesGR = makeGRangesFromDataFrame(genes, keep.extra.columns = T)
+chGR = makeGRangesFromDataFrame(chh, keep.extra.columns = T)
+
+ov = findOverlaps(chGR, genesGR)
+
+g17 = plot.cn.chr(globalCNdf, chrom='17', chr.lengths, genes=genes, haz=chh, label=labeller(chr=label_both, Status=n.status)) + 
+  labs(title="All Samples") + theme(legend.position = 'none')
+g17
+
+ggsave(filename='plots/hazards/chr17.png', plot=g17, width=10, height=5, units='in', dpi=300)
+
+```
+
+```{r echo=F, warning=F, message=F, fig.height=5, fig.width=10}
+chGR$Genes = ''
+for (hit in unique(queryHits(ov))) {
+  chGR[hit]$Genes = paste(genesGR[subjectHits(ov[queryHits(ov) == hit])]$Gene.Symbol, collapse=',')
+}
+
+for (chr in unique(chh$chr)) {  
+  print(chr)
+  p = plot.cn.chr(subset(globalCNdf, seg.type != 'arms') , chrom=chr, chr.lengths, genes=genes, haz=chh)
+  f = paste('chr',chr,'.png',sep='')
+  ggsave(paste('plots/hazards/',f,sep=''), plot=p, width=10, height=5, units='in', dpi=300)
+#  print(p)
+}
+
+annoG = arrange(subset(ccgenes, grepl('CDKN2A$|TP53|MYC$|SMAD4|VEG|ERBB2|EGFR|SMYD|ARID1A|FHIT|RUNX1', Gene.Symbol)), chr)
+annoG = merge(annoG, chr.lengths[c('chrom','chr.length')], by.x='chr',by.y='chrom')
+annoG$ymin = round(range(globalCN.plot$data$mean.value))[1]/2
+annoG$ymax = round(range(globalCN.plot$data$mean.value))[2]/2
+
+p = globalCN.plot + #geom_rect(data=annoG, aes(xmin=start, xmax=end, ymin=ymin,ymax=ymax), color='red') +
+  geom_point(data=annoG, aes(x=start, y=0), size=1.5, color='red') +
+  geom_text_repel(data=annoG, aes(x=start, y=ymax/2, label=Gene.Symbol), color='red', size=1.5, fontface='italic' ) + theme(legend.position = 'none')
+
+ggsave('plots/global_gene_ann.png', plot=p, width=10, height=5, units='in', dpi=300)
+
+```

### 2018-03-08 Misclassified.Rmd
...

@@
 rm(dysplasia.df, fits, plots, pg.samp, allDf, nzcoefs, coefs)
+```
+
+## Worst patient in the 'all' model
+
+54 (AHM0320)
+
+```{r echo=F}
+dir = paste(data, 'Analysis/5e6_arms_all', sep='/')
+file = paste(disc.dir, 'model_data.Rdata', sep='/')
+if (!file.exists(file))
+  stop(paste("Missing data file", file))
+
+load(file, verbose=T)
+all.dysplasia.df = dysplasia.df
+all.labels = labels
+
+sets = create.patient.sets(patient.info[c('Hospital.Research.ID','Samplename','Status')], folds, splits, 0.2)
+
+rows = which(rownames(all.dysplasia.df) %in% subset(patient.info, Hospital.Research.ID %in% c('AHM0320'))$Samplename)
+
+df = all.dysplasia.df[-rows,]
+labels = all.labels[rownames(df)]
+
+a = 0.9; nl=1000; folds=10; splits=5
+fit0 <- glmnet(df, labels, alpha=a, nlambda=nl, family='binomial', standardize=F)    
+#autoplot(fit0) + theme(legend.position = 'none')
+
+cv = crossvalidate.by.patient(x=df, y=labels, lambda=fit0$lambda, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', opt=-1, standardize=F)
+
+lambda.opt = cv$lambda.1se
+subset(cv$lambdas, lambda == lambda.opt)$mean
+
+predict(fit0, newx=all.dysplasia.df[rows,], s=cv$lambda.1se, type='link')
+pm = predict(fit0, newx=all.dysplasia.df[rows,], s=cv$lambda.1se, type='response')
+colnames(pm) = c("Prediction")
+pred = merge(pm, patient.info, by.x='row.names', by.y='Samplename')
+
 ```

### 2018-03-08 lib/common_plots.R

@@ -198,7 +198,7 @@ plot.cn.chr<-function(df, chrom='17', info=NULL, haz=NULL, genes=NULL, label=NUL
   if (!is.null(genes)) {
     genes = subset(genes, chr == chrom)
     if (nrow(genes) > 0) p = p + geom_point(data=genes, aes(x=start,  y=value), color='darkblue', size=3) +
-        geom_text_repel(data=genes, aes(x=start, y=value, label=Gene.Symbol), color='darkblue') 
+        geom_text_repel(data=genes, aes(x=start, y=value, label=Gene.Symbol), color='darkblue', fontface='italic') 
   }
   if (!is.null(haz)) {
     haz = subset(haz, chr == chrom)

### 2018-03-08 processing_pipeline/run-cv2.R

@@ -29,11 +29,11 @@ infodir = args[3]
 
 # Discovery or all
 allPts = ifelse (length(args) == 2, as.logical(args[2]), F)
-
+allPts = T
 
 cache.dir = paste(outdir, '5e6_arms_disc', sep='/')
 if (allPts)
-  cache.dir = paste(outdir, '5e6_arms_all', sep='/')
+  cache.dir = paste(outdir, '5e6_arms_all_exAHM0320', sep='/')
 dir.create(cache.dir, recursive=T, showWarnings=F)
 
 ## Hospital.Research.ID info file
@@ -58,6 +58,12 @@ raw = list.files(path=data, pattern='raw', full.names=T, recursive=T)
 
 cleaned = grep(paste(sum.patient.data$Hospital.Research.ID, collapse = '|'), cleaned, value=T)
 raw = grep(paste(sum.patient.data$Hospital.Research.ID, collapse = '|'), raw, value=T)
+
+cleaned = grep('AHM0320', cleaned, value=T, invert=T)
+raw = grep('AHM0320', raw, value=T, invert=T)
+
+sum.patient.data = subset(sum.patient.data, Hospital.Research.ID != 'AHM0320')
+patient.info = subset(patient.info, Hospital.Research.ID != 'AHM0320')
 
 arms = grep('arms', cleaned, value=T)
 segs = grep('arms', cleaned, invert=T, value=T)
@@ -125,13 +131,13 @@ save(allDf, dysplasia.df, labels, mn.cx, sd.cx, z.mean, z.sd, z.arms.mean, z.arm
 
 nl = 1000;folds = 10; splits = 5 
 
-file = list.files(data, pattern='patient_folds.tsv', recursive=T, full.names=T)
-if (length(file) == 1 && !allPts) {
-  message(paste("Reading folds file", file))
-  sets = read.table(file, header=T, sep='\t')
-} else {
+#file = list.files(data, pattern='patient_folds.tsv', recursive=T, full.names=T)
+#if (length(file) == 1 && !allPts) {
+#  message(paste("Reading folds file", file))
+#  sets = read.table(file, header=T, sep='\t')
+#} else {
   sets = create.patient.sets(patient.info[c('Hospital.Research.ID','Samplename','Status')], folds, splits, 0.2)  
-}
+#}
 
 nl = 1000;folds = 10; splits = 5 
 alpha.values = c(0, 0.5,0.7,0.8,0.9,1)

### 2018-03-12 AllSamplesModel.Rmd
...

@@
 source('lib/common_plots.R')
+source('lib/data_func.R')
 
@@
 
-cache.dir = paste(data, 'Analysis/5e6_arms_all', sep='/')
+cache.dir = paste(data, 'Analysis/5e6_arms_all_exAHM0320', sep='/')
 
@@
 sum.patient.data = summarise.patient.info(patient.info)
+
+sum.patient.data = subset(sum.patient.data, Hospital.Research.ID != 'AHM0320')
 
@@
 
-ggplot(patient.info, aes(Maximal,Circumference, color=Status)) + geom_point() + labs(title='C vs M') + plot.theme
+ggplot(patient.info, aes(Status, Maximal, fill=Status)) + geom_boxplot() + labs(title='M') + plot.theme
+ggplot(patient.info, aes(Status, Circumference, fill=Status)) + geom_boxplot() + labs(title='C') + plot.theme
 
@@
 
-gp = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples')
-ggsave(filename='plots/globalCN.png', plot=gp, width=16, height=4.6, units='in', dpi=300)
-gp
-
-gbe = globalBECN.plot + theme(legend.position = 'none') + labs(y='mean adj.seg.', title="Non-Dysplastic Barrett's Samples") 
-ggsave(filename='plots/globalBECN.png', plot=gbe, width=16, height=4.6, units='in', dpi=300)
-gbe
-
-rm(locs, gp, gbe)
+globalCN.plot = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples')
+ggsave(filename='plots/globalCN.png', plot=globalCN.plot, width=16, height=4.6, units='in', dpi=300)
+globalCN.plot
+
+globalBECN.plot = globalBECN.plot + theme(legend.position = 'none') + labs(y='mean adj.seg.', title="Non-Dysplastic Barrett's Samples") 
+ggsave(filename='plots/globalBECN.png', plot=globalBECN.plot, width=16, height=4.6, units='in', dpi=300)
+globalBECN.plot
+
 ```
@@
 
-tt = t.test(subset(x, Status == 'Progressor')$cx, subset(x, Status == 'Non-Progressor')$cx)
+tt = wilcox.test(subset(x, Status == 'Progressor')$cx, subset(x, Status == 'Non-Progressor')$cx)
 pv = ggplot(x, aes(Status, cx, fill=Status)) + 
@@
 
-```{r noHGD, echo=F, message=F, warning=F, fig.height=16, fig.width=16, eval=T}
+```{r noHGD, echo=F, message=F, warning=F, fig.width=6, fig.height=6}
 rm(performance.at.1se, coefs, mp)
@@
 ggsave('plots/nohgd_eyebrow.png', plot=do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2)), height=10, width=10, units='in', dpi=300)
-```
-
-```{r, echo=F, fig.width=6, fig.height=6}
+
 mp = model.performance(performance.at.1se, coefs)
@@
 
-```{r noLGD, echo=F, warning=F, message=F, fig.height=16, fig.width=16}
+```{r noLGD, echo=F, warning=F, message=F, fig.width=6, fig.height=6, eval=F}
 file = paste(cache.dir, 'nolgd.Rdata', sep='/')
@@
 
-#do.call(grid.arrange, c(nolgd.plots, top="No LGD/HGD/IMC samples", ncol=2))
-```
-
-```{r, echo=F, fig.width=6, fig.height=6}
 mp = model.performance(performance.at.1se, coefs)
@@
 
-```{r loonohgd, echo=F, message=F, warning=T}
+```{r loonohgd, echo=F, message=F, warning=T, fig.width=5, fig.height=5, eval=F}
 file = paste(cache.dir, 'loonohgd.Rdata', sep='/')
@@
 }
-```
-
-```{r, echo=F, message=F, fig.width=5, fig.height=5}
+
 pg.sampNOHGD = do.call(rbind, pg.sampNOHGD)
@@
 
-```{r, echo=F, fig.width=6, fig.height=6}
+```{r, echo=F, fig.width=6, fig.height=6,eval=F}
 ggplot(pg.sampNOHGD, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1), show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') + labs(title='Prediction using model w/o HGD', y='n Samples', x='Probability') + plot.theme
@@
 
-#ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)
+ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)
 
@@
 setdiff(ch$label, rownames(nohgd.coefs))
-
 ```
@@
 
-#ggsave('plots/global_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)
+ggsave('plots/global_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)
 
@@
 
-#ggsave('plots/ndbe_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)
+ggsave('plots/ndbe_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)
 
@@
 
-oddPt = names(pg.samp)[which.max(performance.at.1se)]
-subset(sum.patient.data, Hospital.Research.ID == oddPt)
-pg.samp[[oddPt]]$Prediction
-
-oddPt = names(pg.samp)[which.min(performance.at.1se)]
+# oddPt = names(pg.samp)[which.max(performance.at.1se)]
+# subset(sum.patient.data, Hospital.Research.ID == oddPt)
+# pg.samp[[oddPt]]$Prediction
+# 
+# oddPt = names(pg.samp)[which.min(performance.at.1se)]
 
@@
 
-load(file=paste(cache.dir, 'insilico_splits.Rdata', sep='/'), verbose=T)
+load(file='~/Data/Ellie/Analysis/5e6_arms_all/insilico_splits.Rdata', verbose=T)
 
@@
 
+multi.roc.plot(list(rocFirstEndo, rocLastEndo,rocEndoM, rocEndoA))
+
 ```
@@
 
+multi.roc.plot(list(rocPtA, rocPtM))
 ```
@@
 
+patient.info = subset(patient.info, Patient %in% sum.patient.data$Patient)
+
 back = do.call(rbind.data.frame,lapply(pg.samp, transform.by.time, info=patient.info))
@@
 pred.confidence$Risk = factor(pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)
+
+lows = chisq.test(rbind(colSums(pred.confidence[1:5,c('NP','P')]),table(sum.patient.data$Status)))
+highs = chisq.test(rbind(colSums(pred.confidence[6:10,c('NP','P')]),table(sum.patient.data$Status)))
+
+mods = chisq.test(rbind(colSums(pred.confidence[4:7,c('NP','P')]),table(sum.patient.data$Status)))
 ```
@@
 ```{r pcalib, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
+#pred.confidence$Risk[1:5] = 'Low'
+#pred.confidence$Risk[6:10] = 'High'
+pred.confidence$Risk[4:7] = 'Moderate'
+
 pcal = ggplot(pred.confidence, aes(mn, perc)) + 
-  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
+  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols, limits=levels(pred.confidence$Risk) ) +
   geom_point() + geom_abline(slope=1, intercept=0, color='grey39', linetype='dotted') + #geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
   geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration') +
-  scale_color_manual(values=riskCols) + plot.theme + theme(legend.position = 'none')
+  scale_color_manual(values=riskCols,limits=levels(pred.confidence$Risk) ) + plot.theme + theme(legend.position = 'none')
 pcal
@@
 back = back %>% group_by(Risk) %>% dplyr::mutate('se(Risk)'=(sd(Prediction)/sqrt(length(Risk)))*1.96 )
-
-mod = subset(back, Risk == 'Moderate')
 
@@
 unadjRR
-
 ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR, width=6, height=5, units='in', dpi=300)
@@
 
-ggsave(filename='plots/feature_jacknife.png', plot=p, width=5, height=6, unit='in', dpi=300)
+ggsave(filename='plots/feature_jacknife.png', plot=p, width=6, height=8, unit='in', dpi=300)
 
@@
 
-rp = ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
-  scale_fill_manual(values=riskCols) + scale_color_manual(values=c('white','black','white')) +
+rp = ggplot(riskPath, aes(Pathology, ratio*100)) + 
+  geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
+  scale_fill_manual(values=riskCols, limits=levels(riskPath$Risk)) + 
+  scale_color_manual(values=c('white','black','white'), limits=levels(pred.confidence$Risk) ) +
   geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101.5), position=position_stack(), size=4) +
   geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep=''), color=Risk), position=position_stack(vjust = 0.5), size=5, show.legend = F) +
-  facet_grid(~Status, scales='free_x', space = 'free_x') + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
+  facet_grid(~Status, scales='free_x', space = 'free_x') + plot.theme + 
+  labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
 rp
@@
 
-# ggplot(all.preds, aes(Patient, value/n.samples, group=variable, fill=variable)) + 
-#   geom_bar(stat='identity', position = position_stack()) +
-#   scale_fill_manual(values=riskCols) + 
-#   geom_text(aes(y=1.01, label=n.samples)) +
-#   facet_wrap(~Status, scales='free_x') + labs(y='Sample ratios', title='Per patient ratio of predictions') + 
-#   plot.theme
+ggplot(all.preds, aes(Patient, value/n.samples, group=variable, fill=variable)) +
+  geom_bar(stat='identity', position = position_stack()) +
+  scale_fill_manual(values=riskCols) +
+  geom_text(aes(y=1.01, label=n.samples)) +
+  facet_wrap(~Status, scales='free_x') + labs(y='Sample ratios', title='Per patient ratio of predictions') +
+  plot.theme
 
@@
 endos = back %>% group_by(Status, months.before.final, brks, PID) %>% dplyr::summarise(
-  'pred.endo'=ifelse(length(which(Risk == 'High')) > 0, 1, 0), 
+  'pred.endo'=ifelse(length(which(Risk >= 'Moderate')) > 0, 1, 0), 
   'n.endos'=length(unique(PID)), 
-  'n.samples'=length(PID),
+  'n.samples'=length(Samplename),
   'mean.pred'=mean(Prediction),
-  'pred.samples'=length(which(Risk == 'High')),
-  'max.path'=max(Pathology))
-
-endos$quant =  cut(endos$mean.pred, breaks = cuts)
-endos = merge(endos, pred.confidence[c('quant', 'Risk')])
+  'pred.samples'=length(which(Risk >= 'Moderate')),
+  'max.path'=max(Pathology),
+  'Moderate'=length(which(Risk == 'Moderate')),
+  'High'=length(which(Risk == 'High'))
+  )
+
+#endos$quant =  cut(endos$mean.pred, breaks = cuts)
+#endos = merge(endos, pred.confidence[c('quant', 'Risk')], by='quant')
 
 sq = c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12)
-endos$brks = cut(endos$months.before.final, include.lowest = T, ordered_result = T,
-breaks=sq,
-labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))
+mlab = c('Endpoint','1-24','24-48','48','48-72','72-96','96-120','>120')
+
+# months = 24
+# sq = c(0, 1, 1*months, 2*months, 4*months, 6*months, 8*months)
+# mlab = c()
+# for (i in 1:length(sq)) {
+#   mlab = c(mlab, paste(sq[i], sq[i+1], sep='-'))
+# }
+# mlab[1] = 'Endpoint'
+# mlab = mlab[1:(length(sq)-1)]
+# mlab[length(mlab)] = paste('>',sq[(length(sq)-1)],sep='')
+
+endos$brks = cut(endos$months.before.final, include.lowest = T, ordered_result = T,breaks=sq,labels=mlab)
 endos$brks = with(endos, factor(brks, rev(levels(brks)))) # invert order
@@
 endos.m = subset(endos, Status == 'P') %>% group_by(brks) %>% dplyr::summarise( 
-  'pred.r'=sum(pred.endo)/sum(n.endos),'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos))
+  'pred.r'=sum(pred.endo)/sum(n.endos),'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos), 
+  'mod.r'=sum(Moderate)/sum(n.samples), 'high.r'=sum(High)/sum(n.samples))
 
 tmp = subset(back, Status == 'P')
-tmp$brks = cut(tmp$months.before.final, include.lowest = T, ordered_result = T,breaks=sq,
-labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))
+tmp$brks = cut(tmp$months.before.final, include.lowest = T, ordered_result = T,breaks=sq,labels=mlab)
 tmp = tmp %>% group_by(brks) %>% dplyr::summarise( 'mean(P)'=mean(Prediction),'sem(P)'=sd(Prediction)/sqrt(length(brks)) )
@@
 rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) + ylim(0,105) +
-  geom_bar(stat='identity', fill='darkred', alpha=0.8, show.legend = F) + 
+  geom_bar(stat='identity', fill='orangered4', show.legend = F) + 
   geom_errorbar(color='grey21', width=0.2) +
@@
 ```
-
+# TODO number endoscopies so the images are scaled 
 `r pander(as.data.frame(pred.confidence))`
@@
 
+table(rev(arrange(subset(back, Patient == 19), brks)$brks))
+
 p = pred.tiles(subset(back, Patient %in% c(6)), ncol=1, cols=myPal, probs=T) + labs(title='')
@@
 ```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=15 }
+#back = back %>% group_by(Patient) %>% dplyr::mutate(endo.num = as.integer(droplevels(brks)) )
+#back$endo.num = factor(back$endo.num, ordered = T)
+
+#back$endo.num = factor(as.integer(back$brks))
+#back = back %>% group_by(Patient) %>% dplyr::mutate(endo.num = ifelse(endo.num == max(endo.num), 'End', endo.num)   )
+
 ga = grid.arrange(
@@
 
-p1 = pred.tiles(subset(back, Status == 'NP'), cols=myPal, probs=T, nrow=2) + labs(title='Non-Progressors') 
+p1 = pred.tiles(subset(back, Status == 'NP'), cols=myPal, probs=T, nrow=2, scales='free', drop=T) + labs(title='Non-Progressors') 
+#p1 = pred.tiles(subset(back, Status == 'NP'), cols=riskCols, probs=F, OR=F, nrow=2) + labs(title='Non-Progressors') 
 riskLegend = get.legend(p1)
@@
 p1 = p1 + theme(legend.position = 'none')
-p2 = pred.tiles(subset(back, Status == 'P'), probs=T, cols=myPal,nrow=2) + labs(title='Progressors') + theme(legend.position = 'none')
-
-ggsave('plots/pred_heatmap_ALL.png', plot=grid.arrange(p2,p1,nrow=2), width=52, height=16, units="in", limitsize=F, dpi = 300)
+p2 = pred.tiles(subset(back, Status == 'P'), probs=T, cols=myPal,nrow=2,scales='free',drop=T) + labs(title='Progressors') + theme(legend.position = 'none')
+
+ggsave('plots/pred_heatmap_ALL.png', plot=grid.arrange(p2 + labs(x='', y='',title=''),p1 + labs(x='', y='',title=''),nrow=2), width=11, height=3.4, units="in", limitsize=F, dpi = 300, scale = 3)
 ggsave('plots/npg_heatmap.png', plot=p1, width=52, height=8, units="in", limitsize=F, dpi = 300)
@@
 ggsave('plots/risk_legend_long.png', plot=grid.arrange(riskLegendLong), width=8,height=2, dpi=300)
+
+p1 = pred.tiles(subset(back, Status == 'NP'), cols=myPal, probs=T, nrow=2, scales='free_x', drop=T) + labs(title='Non-Progressors') + theme(legend.position = 'none') 
+p2 = pred.tiles(subset(back, Status == 'P'), probs=T, cols=myPal,nrow=2,scales='free_x',drop=T) + labs(title='Progressors') + theme(legend.position = 'none')
+ggsave('plots/pred_heatmap_ALL_widthadj.png', plot=grid.arrange(p2 + labs(x='', y='',title=''),p1 + labs(x='', y='',title=''),nrow=2), width=11, height=3.4, units="in", limitsize=F, dpi = 300, scale = 3)
 ```
@@
 ```
+
 One way of looking at the p53 prediction (e.g. Adam Bass) is to see how predictive the p-arm loss in chr17 is.
@@
 ```
+
+# Validation
+
+## Downsampled Datasets
+
+```{r echo=F, eval=T}
+dirA = '~/Data/Ellie/Analysis/downsampled5G_BEAdjacent'
+dirB = '~/Data/Ellie/Analysis/downsampled5G_NDBE'
+
+file = paste(cache.dir, 'model_data.Rdata', sep='/')
+if (!file.exists(file))
+  stop(paste("Missing data file", file))
+load(file, verbose=T)
+file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
+if (!file.exists(file))
+  stop(paste("Missing data file", file))
+load(file, verbose=T)
+rm(dysplasia.df,coefs,plots,labels)
+
+select.alpha = '0.9'
+fitV = models[[select.alpha]]
+lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']
+
+predA = list.files(dirA, 'predictions_all_exAHM0320.Rdata', full.names = T)
+predB = list.files(dirB, 'predictions_all_exAHM0320.Rdata', full.names = T)
+
+files = c()
+if (length(predA) < 1)
+  files = c(list.files(dirA, pattern='raw', full.names=T))
+if (length(predB) < 1)
+  files = c(files, list.files(dirB, pattern='raw', full.names=T))
+
+if (length(files) > 0) {
+  sampleNames = sub('_raw.txt', '', basename(files))
+  dspred = data.frame(matrix(ncol=2, nrow=length(sampleNames), dimnames=list(sampleNames,c('Prob','RR'))))
+
+  predA = sub('_raw.txt','',basename(c(list.files(dirA, pattern='raw', full.names=T))))
+  predB = sub('_raw.txt','',basename(c(list.files(dirB, pattern='raw', full.names=T))))
+
+  for (n in 1:length(files)) {
+    print(n)
+    print(sampleNames[n])
+
+    rs = read.table(files[n], sep='\t',header=T)
+    if (ncol(rs) < 4)
+      stop("File needs to contain 4 columns: chr, start, end, sample_value")
+
+    segs <- tile.segmented.data(rs, size=5e6, chr.info=chr.lengths)
+    segM = as.matrix(segs[,-c(1:3)])
+    rownames(segM) = paste(segs$chr, ':', segs$start, '-', segs$end, sep='')
+    segM = t(segM)
+    for (i in 1:ncol(segM)) 
+      segM[,i] = unit.var(segM[,i], z.mean[i], z.sd[i])
+
+    arms <- tile.segmented.data(rs, size='arms', chr.info=chr.lengths)
+    armsM = as.matrix(arms[,-c(1:3)])
+    rownames(armsM) = paste(arms$chr, ':', arms$start, '-', arms$end, sep='')
+    armsM = t(armsM)
+    for (i in 1:ncol(armsM)) 
+      armsM[,i] = unit.var(armsM[,i], z.arms.mean[i], z.arms.sd[i])
+
+    nrow(armsM) == nrow(segM)
+
+    cx.score = score.cx(segM,1)
+
+    mergedDf = subtract.arms(segM, armsM)
+    mergedDf = cbind(mergedDf, 'cx' = unit.var(cx.score, mn.cx, sd.cx))
+    dim(mergedDf)
+
+    # Predict function giving me difficulty when I have only a single sample, this ensures the dimensions are the same
+    sparsed_test_data <- Matrix(data=0, nrow=nrow(mergedDf),  ncol=ncol(mergedDf),
+                                dimnames=list(rownames(mergedDf),colnames(mergedDf)), sparse=T)
+    for(col in colnames(mergedDf)) sparsed_test_data[,col] = mergedDf[,col]
+
+    preds = predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='response')
+    RR = predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='link')
+
+    dspred[n,] = c(preds[,1],RR[,1])
+    print(dspred[1:n,])
+  }
+    dspred$Type = NA
+    dspred[predA,'Type'] = 'AdjacentBE'
+    dspred[predB,'Type'] = 'BEST2 NDBO'
+    pred = dspred
+
+    dspred = subset(pred, Type == 'AdjacentBE')
+
+    save(dspred, file=paste(dirA, 'predictions_all_exAHM0320.Rdata', sep='/'))
+
+    dspred = subset(pred, Type == 'BEST2 NDBO')
+    save(dspred, file=paste(dirB, 'predictions_all_exAHM0320.Rdata', sep='/'))
+
+    dspred = pred
+
+} else {
+  load(predB, verbose=T)
+  dspredB = dspred
+
+  load(predA, verbose=T)
+
+  dspred = rbind(dspred, dspredB)
+  dspred$Type = NA
+
+  dspred[sub('_raw.*','',basename(list.files(dirA, pattern='raw', full.names=T))),'Type'] = 'AdjacentBE'
+  dspred[sub('_raw.*','',basename(list.files(dirB, pattern='raw', full.names=T))),'Type'] = 'BEST2 NDBO'
+
+}
+
+ggplot(dspred, aes(Type, RR, fill=Type)) + geom_boxplot() + geom_jitter(width=0.2) +
+  labs(x='', y='Relative Risk', title='Predictions on downsampled patients') + 
+  theme(legend.position = 'none') + plot.theme
+
+adj = subset(dspred, Type == 'AdjacentBE')
+prg = back[,c('Prediction', 'OR', 'Status')]
+prg$Status = ifelse(prg$Status == 'P', 'Progressor','Non-Progressor')
+
+#  subset(back, Status == 'P', select=c('Prediction', 'OR', 'Status'))
+colnames(prg) = colnames(adj)
+m = rbind(adj,prg)
+m$Type = factor(m$Type, levels=c('Non-Progressor','BEST2 NDBO','Progressor','AdjacentBE'), ordered = T)
+
+pr = ggplot(m, aes(Type, RR, fill=Type)) + 
+  geom_jitter(width=0.2) + geom_boxplot() + 
+  labs(x='', y='RR', title='Predicted Relative Risk') + 
+  theme(legend.position = 'none', text=element_text(size=11)) + plot.theme 
+pr
+
+ggsave('plots/adjacentBE.png', plot=pr, width=8, height=8,dpi=300)
+
+```
+
+# Clinical Story
+
+First note here is that our probability distribution for these 20 patients is somewhat flatter than in the training set.
+Likely a consequence of sample size.
+The labels show the proprotions of progressors that fell into each quintile in the training set.
+
+## Vignettes
+
+```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=5}
+vign.pts = as_tibble(subset(sum.patient.data, Patient %in% c(3, 33, 34, 49, 56, 20), select=c('Status','Patient','age.diagnosis', 'smoking', 'C','M', 'initial.endo','final.endo', 'total.endos','total.samples')))
+colnames(vign.pts) = c('Status','Patient','Age at Diagnosis','Smoker','Circumference', 'Maximal length', 'Intial Endoscopy','Final Endoscopy', 'Total Endoscopies','Total Samples')
+
+## P
+p = vig.plot(df=subset(back, Patient == 34),info= subset(patient.info, Patient == 34),cols=riskCols) + theme(legend.position = 'bottom')
+vigLeg = get.legend(p)
+p = p + theme(legend.position = 'none')
+ggsave('plots/vignette/vig_34.png', plot=p, width=6, height=6,dpi=300)
+
+p = vig.plot(subset(back, Patient == 56), subset(patient.info, Patient == 56),cols=riskCols) + theme(legend.position = 'none')
+ggsave('plots/vignette/vig_56.png', plot=p, width=6, height=6,dpi=300)
+
+p = vig.plot(subset(back, Patient == 20), subset(patient.info, Patient == 20),cols=riskCols) + theme(legend.position = 'none')
+ggsave('plots/vignette/vig_20.png', plot=p, width=6, height=6,dpi=300)
+
+p = vig.plot(subset(back, Patient == 86), subset(patient.info, Patient == 86),cols=riskCols) + theme(legend.position = 'none')
+ggsave('plots/vignette/vig_86.png', plot=p, width=6, height=6,dpi=300)
+
+print(p)
+
+## NP
+p = vig.plot(subset(back, Patient == 3), subset(patient.info, Patient == 3),cols=riskCols) + theme(legend.position = 'none')
+ggsave('plots/vignette/vig_3.png', plot=p, width=6, height=6,dpi=300)
+
+p = vig.plot(df=subset(back, Patient == 23),info= subset(patient.info, Patient == 23),cols=riskCols) + theme(legend.position = 'none')
+ggsave('plots/vignette/vig_23.png', plot=p, width=6, height=6,dpi=300)
+
+p = vig.plot(subset(back, Patient == 33), subset(patient.info, Patient == 33),cols=riskCols) + theme(legend.position = 'none')
+ggsave('plots/vignette/vig_33.png', plot=p, width=6, height=6,dpi=300)
+
+p = vig.plot(subset(back, Patient == 49), subset(patient.info, Patient == 49),cols=riskCols) + theme(legend.position = 'none')
+ggsave('plots/vignette/vig_49.png', plot=p, width=6, height=6,dpi=300)
+
+p = vig.plot(subset(back, Patient == 51), subset(patient.info, Patient == 51),cols=riskCols) + theme(legend.position = 'none')
+ggsave('plots/vignette/vig_51.png', plot=p, width=6, height=6,dpi=300)
+
+p = vig.plot(subset(back, Patient == 6), subset(patient.info, Patient == 6),cols=riskCols) + theme(legend.position = 'none')
+ggsave('plots/vignette/vig_6.png', plot=p, width=6, height=6,dpi=300)
+
+p = vig.plot(subset(back, Patient == 25), subset(patient.info, Patient == 25),cols=riskCols) + theme(legend.position = 'none')
+ggsave('plots/vignette/vig_25.png', plot=p, width=6, height=5,dpi=300)
+
+p = vig.plot(subset(back, Patient == 68), subset(patient.info, Patient == 68),cols=riskCols) + theme(legend.position = 'none')
+ggsave('plots/vignette/vig_68.png', plot=p, width=6, height=4,dpi=300)
+
+print(p)
+
+```
+
+### Recommendations
+
+ 1.	Immediate RFA: HGD diagnosis, two or more consecutive high risk predictions
+ 2.	Recheck 6-12 months: One high risk prediction or aberrant p53 IHC
+ 3. Recheck endoscopy 12-24 months: One or more moderate risk predictions
+ 4. Regular surveillance 3-5 years: Two or more consecutive low risk predictions
+
+```{r, rx, echo=F, warning=F, message=T, fig.width=8, fig.height=5}
+
+rx<-function(Patient, brks, Prediction, Pathology, Risk, p53.Status) {
+  rules = as.data.frame(matrix(ncol=0, nrow=0))
+  # Consecutive
+  for (i in 1:length(brks)) {
+    risks = table(Risk[i:(i+1)])
+    p53 = table(p53.Status[i:(i+1)])
+    rule = 'None'
+    if ( risks['High'] == 2 | length(which(grepl('HGD|IMC', Pathology[i:(i+1)]))) > 0 ) {
+      rule = 1
+    } else if ( risks['High'] == 1 | p53['1'] > 0 ) {
+      rule = 2
+    } else if ( risks['Moderate'] > 0 ) {
+      rule = 3
+    } else if ( risks['Low'] == 2) {
+      rule = 4
+    }
+  rules = rbind(rules, cbind( as.character(brks[i]), as.character(brks[(i+1)]), rule))
+  }  
+  colnames(rules) = c('t1','t2', 'rule')
+  rules$Patient = Patient
+  rules
+}
+
+allB = back
+allB$p53.Status = factor(allB$p53.Status, levels=c(0,1), ordered = T)
+
+maxP = allB %>% group_by(Patient, Status, brks) %>% dplyr::summarise( 'maxP'=max(Prediction), 'Risk'=Risk[Prediction == max(Prediction)], 'Pathology'=max(Pathology), 'p53'=max(p53.Status,na.rm=T) )
+
+rules = do.call(rbind, lapply(unique(maxP$Patient), function(pt) {
+  r = with(arrange(subset(maxP, Patient == pt), brks), rx(pt, brks,maxP,Pathology,Risk,p53))
+  r$Status = unique(subset(maxP, Patient == pt)$Status)
+  r
+}))
+rules$rule = factor(rules$rule, levels=c('None',4,3,2,1), ordered = T)
+
+rules$t1 = factor(rules$t1, levels=levels(back$brks), ordered=T)
+rules$t2 = factor(rules$t2, levels=levels(back$brks), ordered=T)
+
+penult = subset(rules, t2 != 'Endpoint' & (t1 == rev(levels(back$brks))[2] | t2 == rev(levels(back$brks))[2] ))
+
+# Based on the penult pair
+round(table(subset(penult, Status == 'P')$rule)/nrow(subset(penult, Status == 'P')), 2)
+round(table(subset(penult, Status == 'NP')$rule)/nrow(subset(penult, Status == 'NP')), 2)
+
+second = subset(rules,  t2 != 'Endpoint') %>% 
+  group_by(Patient, Status) %>% 
+  summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )
+
+# At least 2 endos
+pander(rbind('Progressors'=table(subset(second, Status == 'P')$rule),
+             'Non-Progressors'=table(subset(second, Status == 'NP')$rule)), caption='Two endos, applied rules')
+
+path.rx<-function(Patient, brks, Pathology) {
+  rules = as.data.frame(matrix(ncol=0, nrow=0))
+  # Consecutive
+  for (i in 1:length(brks)) {
+    #risks = table(Pathology[i:(i+1)])
+    rule = 'None'
+    if ( length(which(grepl('HGD|IMC', Pathology[i:(i+1)]))) > 0  |
+         length(which(grepl('LGD', Pathology[i:(i+1)]))) > 1 ) {
+      rule = 1
+    } else if ( length(which(grepl('LGD', Pathology[i:(i+1)]))) > 0 ) {
+      rule = 2
+    } else if ( length(which(grepl('BE|ID', Pathology[i:(i+1)]))) > 1  ) {
+      rule = 3
+    }
+  rules = rbind(rules, cbind( as.character(brks[i]), as.character(brks[(i+1)]), rule))
+  }  
+  colnames(rules) = c('t1','t2', 'rule')
+  rules$Patient = Patient
+  rules
+}
+
+rules2 = do.call(rbind, lapply(unique(maxP$Patient), function(pt) {
+  r = with(arrange(subset(maxP, Patient == pt), brks), path.rx(pt, brks,Pathology))
+  r$Status = unique(subset(maxP, Patient == pt)$Status)
+  r
+}))
+rules2$rule = factor(rules$rule, levels=c('None',4,3,2,1), ordered = T)
+
+rules2$t1 = factor(rules2$t1, levels=levels(back$brks), ordered=T)
+rules2$t2 = factor(rules2$t2, levels=levels(back$brks), ordered=T)
+
+second = subset(rules2,  t2 != 'Endpoint') %>% 
+  group_by(Patient, Status) %>% 
+  summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )
+
+pander(rbind('Progressors'=table(subset(second, Status == 'P')$rule),
+            'Non-Progressors'=table(subset(second, Status == 'NP')$rule)), caption='Two endos, pathology rules')
+
+```
+
+## Cox model - Time to progress?
+
+```{r cox, warning=F, message=F, echo=F, eval=F}
+library(survival)
+
+cox.pvalue<-function(cox, test='logrank') {
+  pvalue = NULL  
+  csum = summary(cox)
+
+  if (test == 'logrank') {
+    pvalue = csum$sctest[['pvalue']]  
+  } else if (test == 'wald') {
+    pvalue = csum$waldtest[['pvalue']]
+  } else if (test == 'loglik') {
+    pvalue = csum$logtest[['pvalue']]
+  }
+  return(pvalue)
+}
+
+prog.pred = lapply(pg.samp, function(df) {
+  df = df[,c('Hospital.Research.ID', 'Status', 'Patient', 'Age.at.diagnosis', 'PID', 'Status', 'Endoscopy.Year','Initial.Endoscopy', 'Final.Endoscopy', 'Pathology', 'Samplename','Prediction','OR')] %>% rowwise() %>% dplyr::mutate(
+      'Time.to.Final'=Final.Endoscopy-Endoscopy.Year)
+  df = arrange(df, Endoscopy.Year, Pathology)
+
+  df$Dup.PID = duplicated(df$PID)
+  x = df[!df$Dup.PID,]
+
+  for (i in 1:nrow(x)) {
+    if (i == 1) {
+      x[i,'tstart'] = 0
+      x[i, 'tstop'] = x[(i+1), 'Endoscopy.Year'] - x[i, 'Endoscopy.Year']
+      if (is.na(x[i,'tstop']) || x[i,'tstop'] == 0) x[i,'tstop'] = 0.5
+      next
+    }
+
+    x[i,'tstart'] = x[(i-1), 'tstop']
+
+    # Final sample is different -- will leave it out later
+    if (i == nrow(x)) {
+      x[i, 'tstop'] = x[i,'tstart'] + 0.5
+      break
+    }
+
+    x[i, 'tstop'] =  x[i,'tstart'] + x[(i+1), 'Endoscopy.Year'] - x[i, 'Endoscopy.Year']
+
+    # Samples taken in the same year
+    if ( x[i, 'Endoscopy.Year'] == x[(i+1), 'Endoscopy.Year'] )
+      x[i,'tstop'] = x[i,'tstop'] + 0.5 
+  }
+
+  x$Diagnosis = as.integer(x$Pathology %in% c('HGD','IMC'))
+
+  if (sum(x$Diagnosis) >= 1)
+    x = x[-which(x$Diagnosis == 1),]
+
+  if (nrow(x) > 0 && sum(x$Diagnosis) <= 0 && x$Status == 'P')
+    x[nrow(x), 'Diagnosis'] = 1
+
+  if (nrow(x) > 0) {
+    x$Dup.Samples = ''
+    for (j in 1:nrow(df[df$Dup.PID,]))  {
+      row = df[df$Dup.PID,][j,]
+      x[which(x$PID == row[['PID']]), 'Dup.Samples'] = paste(x[which(x$PID == row[['PID']]), 'Dup.Samples'], row[['Samplename']], sep=',')
+    }
+    x$Dup.Samples = sub('^,','',x$Dup.Samples)
+  }
+
+  return(x)
+})
+## By pathology?
+pp = do.call(rbind, prog.pred)
+
+pp[is.na(pp$tstop), 'tstop'] = 0.5
+pp$Pathology = as.character(pp$Pathology)
+pp$AAD = cut(pp$Age.at.diagnosis, breaks = 3)
+
+pp = merge(pp, back[c('Samplename','Risk', 'quants')], by='Samplename')
+pp$Risk = as.character(pp$Risk)
+
+# sfit = survfit( Surv(tstart, tstop, Diagnosis, type='interval') ~ Risk, data=pp, conf.int=0.95, conf.type='log', na.action=na.omit)
+# ggsurv(sfit,cens.shape=3,lty.est=1,back.white=F) + xlim(0,5) + labs(y='Diagnosed', title='~Risk')
+
+cox = coxph( Surv(tstart, tstop, Diagnosis) ~ RR, data=pp)
+ggsurv(survfit(cox),cens.shape=3,lty.est=1,back.white=F,
+      main="coxph(Survival, Diagnosis) ~ RR" ) +
+      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),2),"\n", nrow(pp), 'samples'))
+summary(cox)
+
+sfit = survfit( Surv(tstart, tstop, Diagnosis, type='interval') ~ AAD, data=pp, conf.int=0.95, conf.type='log', na.action=na.omit)
+ggsurv(sfit,cens.shape=3,lty.est=1,back.white=F) + xlim(0,5) + labs(y='Time to progression', title='~Age') 
+cox = coxph( Surv(tstart, tstop, Diagnosis) ~ AAD, data=pp) 
+ggsurv(survfit(cox),cens.shape=3,lty.est=1,back.white=F,
+      main="coxph(Survival, Diagnosis) ~ Age at diagnosis" ) + 
+      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),2),"\n", nrow(pp), 'samples'))
+#summary(cox)
+
+sfit = survfit( Surv(tstart, tstop, Diagnosis, type='interval') ~ Pathology, data=pp, conf.int=0.95, conf.type='log', na.action=na.omit)
+ggsurv(sfit,cens.shape=3,lty.est=1,back.white=F) + xlim(0,5) + labs(y='Time to progress', title='~Pathology')
+cox = coxph( Surv(tstart, tstop, Diagnosis) ~ Pathology, data=pp) 
+ggsurv(survfit(cox),cens.shape=3,lty.est=1,back.white=F,
+      main="coxph(Survival, Diagnosis) ~ Pathology" ) + 
+      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),2),"\n", nrow(pp), 'samples')) + xlim(0,5)
+#summary(cox)
+
+```

### 2018-03-12 lib/common_plots.R

@@ -147,7 +147,7 @@ multi.roc.plot <- function(rocList, title="ROC", palette=NULL, colors=NULL) {
 }
 
 pred.tiles <- function(df, probs=F, OR=F, cols=c('green3','orange','red3'), ...) {
-  p = ggplot(df, aes(brks, nl))
+  p = ggplot(df, aes(brks, as.factor(Block)))
 
   if (OR) {
     p = p + geom_tile(aes(fill=OR), color='white') + scale_fill_gradientn(colors = myPal, limits=c(-9.3,15), name='RR') 
@@ -157,15 +157,12 @@ pred.tiles <- function(df, probs=F, OR=F, cols=c('green3','orange','red3'), ...)
     p = p + geom_tile(aes(fill=Risk)) + 
       scale_fill_manual(values=cols, limits=levels(df$Risk), name='Progression')
   }
-  p + 
-    geom_point(aes(shape=Pathology, color=Risk), fill='white', size=3) + scale_color_manual(values=c('white','black','white'), guide=F) +
+  p + geom_point(aes(shape=Pathology), fill='white', color='white', size=3) + 
+    # scale_color_manual(values=c('white','black','white'), guide=F) +
     scale_shape_manual(values=c(1,0,15,24,25), limits=levels(df$Pathology), guide=guide_legend(override.aes=list(fill='white', color='white'))) +
-    #scale_fill_manual(values=cols, limits=levels(df$Risk),name='Progression') +  #scale_color_manual(values=cols, limits=levels(df$Risk),name='Progression') +
-    #geom_label(aes(label=Pathology), fill='white', show.legend=F) + 
-    #scale_color_manual(values=c('white','black','white'), limits=levels(df$Risk)) +
-    facet_wrap(~Patient, scales='free', labeller=label_both, ...) +
-    labs(y='Oesophageal Location',x='Months Prior to Endpoint', title='') +
-    plot.theme + theme(axis.text.x=element_text(angle=45, hjust=1), legend.key = element_rect(fill='grey39')) 
+    facet_wrap(~Patient, ...) +
+    labs(y='Oesophageal Location (OGJ..)',x='Endoscopy (Base...End)', title='') +
+    plot.theme + theme(axis.text = element_blank(), legend.key = element_rect(fill='grey39'), panel.background = element_rect(colour = 'black'), panel.grid.major = element_blank()) 
 }
 
 plot.cn.chr<-function(df, chrom='17', info=NULL, haz=NULL, genes=NULL, label=NULL) {

### 2018-03-12 processing_pipeline/run-cv2.R

@@ -29,9 +29,8 @@ infodir = args[3]
 
 # Discovery or all
 allPts = ifelse (length(args) == 2, as.logical(args[2]), F)
-allPts = T
 
-cache.dir = paste(outdir, '5e6_arms_disc', sep='/')
+cache.dir = paste(outdir, '5e6_arms_disc_exAHM0320', sep='/')
 if (allPts)
   cache.dir = paste(outdir, '5e6_arms_all_exAHM0320', sep='/')
 dir.create(cache.dir, recursive=T, showWarnings=F)
@@ -136,7 +135,7 @@ nl = 1000;folds = 10; splits = 5
 #  message(paste("Reading folds file", file))
 #  sets = read.table(file, header=T, sep='\t')
 #} else {
-  sets = create.patient.sets(patient.info[c('Hospital.Research.ID','Samplename','Status')], folds, splits, 0.2)  
+sets = create.patient.sets(patient.info[c('Hospital.Research.ID','Samplename','Status')], folds, splits, 0.2)  
 #}
 
 nl = 1000;folds = 10; splits = 5 
@@ -151,7 +150,7 @@ if (file.exists(file)) {
   for (a in alpha.values) {
     fit0 <- glmnet(dysplasia.df, labels, alpha=a, nlambda=nl, family='binomial', standardize=F)    
     l = fit0$lambda
-    
+
     cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=l, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', opt=-1, standardize=F)
     
     lambda.opt = cv.patient$lambda.1se

### 2018-03-12 AllSamplesModel.Rmd
...

@@
 globalCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms'), labeller(Status=n.status, chr=label_value)) 
-cnLegend = get.legend(globalCN.plot)
-#globalCN.plot = globalCN.plot + theme(legend.position = 'none')
+cnLegend = get.legend(globalCN.plot + theme(legend.position='bottom'))
+
+globalCN.plot = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples')
+ggsave(filename='plots/globalCN.png', plot=globalCN.plot + labs(title=''), width=16, height=4.6, units='in', dpi=300)
+globalCN.plot
 
@@
 globalBECN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology == 'BE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')
-
-globalCN.plot = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples')
-ggsave(filename='plots/globalCN.png', plot=globalCN.plot, width=16, height=4.6, units='in', dpi=300)
-globalCN.plot
-
 globalBECN.plot = globalBECN.plot + theme(legend.position = 'none') + labs(y='mean adj.seg.', title="Non-Dysplastic Barrett's Samples") 
-ggsave(filename='plots/globalBECN.png', plot=globalBECN.plot, width=16, height=4.6, units='in', dpi=300)
+ggsave(filename='plots/globalBECN.png', plot=globalBECN.plot + labs(title=''), width=16, height=4.6, units='in', dpi=300)
 globalBECN.plot
@@
 
-```{r noLGD, echo=F, warning=F, message=F, fig.width=6, fig.height=6, eval=F}
+```{r noLGD, echo=F, warning=F, message=F, fig.width=6, fig.height=6, eval=T}
 file = paste(cache.dir, 'nolgd.Rdata', sep='/')
@@
 
-chh$ymax = ifelse(chh$coef > 0, max(globalCN.plot$data$mean.value), min(globalCN.plot$data$mean.value))
+chh = merge(chh,globalCNdf %>% group_by(chr) %>% dplyr::summarise(minV=min(mean.value), maxV=max(mean.value)),by='chr')
+chh = chh %>% dplyr::mutate( ymax = ifelse(coef > 0, maxV, minV) )
+chh[c('maxV','minV')] = NULL
+
 chh$gl = factor(ifelse(chh$coef > 0, 'gain', 'loss'))
@@
 
-ggsave(filename='plots/discovery_roc.png', plot=roc.plot(roc), width=6, height=4, units='in', dpi=300)
+ggsave(filename='plots/auc/all_roc.png', plot=roc.plot(roc), width=6, height=4, units='in', dpi=300)
 
@@
 
-ggsave('plots/auc/path_p53_cn_auc.png', plot=comp, width=6, height=8, units="in", limitsize=F, dpi = 300)
+ggsave('plots/auc/path_p53_cn_auc.png', plot=comp, width=2.7, height=3.7, units="in", limitsize=F, dpi = 300, scale=2)
 
@@
 
-ggsave(filename='plots/pred_calib.png', plot=pcal, width=3, height=3, units='in', dpi=300)
+ggsave(filename='plots/pred_calib.png', plot=pcal, width=2.2, height=2.2, units='in', dpi=300, scale=2)
 
@@
 unadjRR
-ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR, width=6, height=5, units='in', dpi=300)
+ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR, width=5.5, height=4.3, units='in', dpi=300)
 
@@
 ```
+
+### Mountain plots using spatial info
+
+```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=20}
+min.theme = theme(text=element_text(size=12), panel.background=element_blank(), strip.background=element_rect(fill="grey97"), strip.text.y = element_text(size=6),
+                   strip.text.x = element_text(size=12), axis.text = element_blank(), axis.ticks = element_blank(),
+                   axis.line=element_line(color='black'), panel.grid.major.y =element_line(color='grey95'), panel.grid.major.x = element_blank(),
+                   panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0, 'lines'), panel.spacing.y = unit(0.1, 'lines')   )
+
+mtn.spatial<-function(df, b, limits, title='', pal=NULL) {
+  if (is.null(pal))
+    pal = RColorBrewer::brewer.pal(length(limits), 'Set2')
+
+  df = merge(df, b[,c('Samplename','Endoscopy.Year','months.before.final','Block','ogj')], by.x='variable',by.y='Samplename')
+  df$ogj = factor(df$ogj, ordered = T)
+  df$ogj = factor(df$ogj, levels=rev(levels(df$ogj)), ordered = T)
+  #pal = pal[which(limits %in% levels(df$nl))]
+
+  altcolor = c(RColorBrewer::brewer.pal(9, 'Greys')[2],RColorBrewer::brewer.pal(9, 'Greys')[4])
+  fills = c()
+  for (yr in sort(unique(df$Endoscopy.Year))) {
+    ac = (which(sort(unique(df$Endoscopy.Year)) == yr) %% 2) + 1
+    fills = c(fills, rep(altcolor[ac], length(which(table(subset(df, Endoscopy.Year == yr)$ogj) > 0))))
+  }
+
+  blanklabel <- function(s) {
+    return('')
+  }
+
+  p = ggplot(df, aes(x=chr.length)) + facet_grid(Endoscopy.Year+ogj~chr, scales='free', space='free_x', labeller = labeller(.multi_line = F)) +
+      geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=ogj), show.legend=T) + 
+      scale_fill_manual(values=pal, limits=limits,  name='OGJ Dist.') +
+      labs(title=title, x='') + min.theme
+
+  legend = get.legend(p)
+  p = p + theme(legend.position = 'none')
+
+  g = ggplot_gtable(ggplot_build(p))
+  strips = grep('strip-r', g$layout$name)
+
+  for (n in 1:length(strips)) {
+    i = strips[n]
+    j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
+    g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[n]
+  }
+
+  #grid.arrange(g)
+  return(list('grob'=g, 'legend'=legend))
+  #grid::grid.draw(g)  
+  #return(g)
+}
+
+back$ogj = back$nl
+back$ogj[which(back$nl > 0 & back$nl < 3)] = 1 
+back$ogj[which(back$nl > 3 & back$nl < 7)] = 2 
+back$ogj[which(back$nl >= 7)] = 3
+
+ogjLimits = levels(factor(back$ogj, ordered = T))
+
+df = subset(globalCNdf, Patient == 20 & seg.type != 'arms')
+#df = subset(globalCNdf, Patient == 51 & seg.type != 'arms')
+
+# early/late samples
+#df = subset(globalCNdf, Patient == pt)
+df = merge(globalCNdf, back[,c('Samplename','Endoscopy.Year', 'ogj', 'months.before.final', 'OR','Prediction')], by.x='variable', by.y='Samplename')
+df = arrange(df, chr, start, Endoscopy.Year)
+
+## count arms differently than segs??
+df2 = df %>% group_by(Status, Patient, Endoscopy.Year) %>% summarise( 
+  'max.Pred'=max(Prediction),
+  'max.RR'=max(OR),
+  'max.Path'=max(Pathology),
+  'CNA'=length(which(seg.type != 'arms' & cn %in% c('loss','gain'))),
+  'arms'=length(which(seg.type == 'arms' & cn %in% c('loss','gain'))),
+  'ratio'=length(which(seg.type != 'arms' & cn %in% c('loss','gain')))/length(cn)
+  )
+#as.data.frame(df2)
+#df2 = subset(df2, max.Path %in% c('BE','ID','LGD'))
+
+df2 = df2 %>% group_by(Status, Patient) %>% mutate(
+  'early'= Endoscopy.Year == min(Endoscopy.Year),
+  'late' = Endoscopy.Year == max(Endoscopy.Year)
+)
+df2 = subset(df2, Patient %in% names(which(table(df2$Patient) > 1)))
+
+el = subset(df2, early | late)
+el2 = el %>% group_by(Status, Patient) %>% summarise( diff(arms), diff(CNA) )
+wtA = round(wilcox.test(subset(el2, Status == 'Progressor')$`diff(arms)`, subset(el2, Status == 'Non-Progressor')$`diff(arms)`)$p.value, 3)
+if (wtA < 0.0001) wtA = '< 0.0001'
+#wilcox.test(subset(el2, Status == 'Progressor')$`diff(CNA)`, subset(el2, Status == 'Non-Progressor')$`diff(CNA)`)
+
+p = ggplot(subset(df2, early|late), aes(Endoscopy.Year, arms, color=Patient)) + facet_grid(~Status) +
+  geom_point() +  geom_line(data=subset(df2, early | late)) + theme(legend.position = 'none', text=element_text(size=11)) + 
+  plot.theme + labs(x='Endoscopy Year', y='CN arm values', title='Initial vs final endoscopy', subtitle=paste('p', wtA))
+p
+ggsave(filename='plots/status_hetero.png', plot=p, width=8, height=6, units='in', dpi=300)
+
+wtA = round(wilcox.test(subset(df2, early &  Status == 'Progressor')$arms, subset(df2, early & Status == 'Non-Progressor')$arms)$p.value, 3)
+wtCN = round(wilcox.test(subset(df2, early & Status == 'Progressor')$CNA, subset(df2, early & Status == 'Non-Progressor')$CNA)$p.value, 2)
+
+m = melt(subset(df2, early), measure.vars=c('CNA','arms'), id.vars=c('Status'))
+p = ggplot(m, aes(Status, value, fill=Status)) + facet_wrap(~variable, scales='free_y') + 
+  geom_boxplot(outlier.colour = NA) + geom_jitter(width=0.2) + plot.theme + theme(legend.position = 'none', text=element_text(size=11)) +
+  labs(y='', x='', title='CN values, initial endoscopy', subtitle=paste('Arms p=',wtA, '   CNA p=',wtCN, sep=''))
+p
+ggsave(filename='plots/hetero_initial.png', plot=p, width=6, height=4, units='in', dpi=300)
+
+p2 = mtn.spatial(subset(globalCNdf, Patient == 20 & seg.type != 'arms'), back, ogjLimits,'Patient 20 (P)' )
+p6 = mtn.spatial(subset(globalCNdf, Patient == 51 & seg.type != 'arms'), back,ogjLimits,'Patient 51 (NP)' )
+
+# for (i in 1:nrow(sum.patient.data)) {
+#   print(i)
+#   pt = sum.patient.data[i, 'Patient']
+#   status = sum.patient.data[i, 'Status']
+# 
+#   pms = mtn.spatial(subset(globalCNdf, Patient == pt & seg.type != 'arms'), back, ogjLimits, paste('Patient', pt, '(', status, ')' ))
+#   
+#   fl = paste('patient_', pt, '.png', sep='')
+#   ggsave(paste('plots/spatial/', fl, sep=''), plot=grid.arrange(pms$grob), width=14, height=6, units='in', dpi=300)
+# }
+
+back$ogj = NULL
+
+```
+
+```{r, echo=F, fig.width=2, fig.height=4}
+
+grid.arrange(p2$legend)
+
+ggsave(filename='plots/spatial/patient_20.png', plot=grid.arrange(p2$grob), width=14, height=6, units='in', dpi=300)
+ggsave(filename='plots/spatial/patient_51.png', plot=grid.arrange(p6$grob), width=14, height=6, units='in', dpi=300)
+
+```
+
+At each quintile are the percentage of progressor samples that were predicted within that range, and the hypergeometric p-value (Fishers) for the significance within that quintile.
+Then our predictions are at best a coin flip in the middle, but highly significant at the outer edges.
+This would also push the prediction for "high" risk up to `r min(subset(pred.confidence, labels == 'High' )$r1)` while "low" risk is below `r max(subset(pred.confidence, labels == 'Low' )$r2)`.
+
+`r pander(as.data.frame(pred.confidence))`
 
@@
 
-ggsave(filename='plots/pred_by_path.png', plot=rp, width=8, height=8, units='in', dpi=300)
+ggsave(filename='plots/pred_by_path.png', plot=rp, width=5.8, height=4.5, units='in', dpi=300, scale=1.5)
+
+```
+
+```{r p53IHC, echo=F, warning=F, message=F, fig.height=8, fig.width=8}
+
+#p53Path$Risk[p53Path$Risk == 'High'] = 'Aberrant'
+#p53Path$Risk[p53Path$Risk == 'Low'] = 'Normal'
+
+#p53Path$p53.Status = factor(p53Path$p53.Status, levels=c('Aberrant', 'Normal'), ordered = F)
+
+p53Path$p53.Status = factor(p53Path$p53.Status, ordered = T)
+p53Path = arrange(p53Path, Pathology, p53.Status)
+col = brewer.pal(11, "PRGn")[c(10,2)]
+rp = ggplot(p53Path, aes(Pathology, ratio*100, group=Pathology, fill=p53.Status)) + geom_bar(stat='identity', position = position_stack()) +
+  scale_fill_manual(values=col, name='p53 IHC', labels=c('Normal','Aberrant')) + 
+  geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=102)) +
+  geom_text(aes(label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5,color='white') +
+  plot.theme + labs(title='Samples predicted by aberrant p53 IHC', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
+rp
+ggsave(filename='plots/p53_pred.png', plot=rp, width=4.5, height=3.6, units='in', dpi=300, scale=1.5)
 
@@
 rp
-ggsave(filename='plots/pred_endo.png', plot=rp, width=9, height=7, units='in', dpi=300)
+ggsave(filename='plots/pred_endo.png', plot=rp, width=5.6, height=4.4, units='in', dpi=300, scale=1.5)
 
@@
 ```
-# TODO number endoscopies so the images are scaled 
+
 `r pander(as.data.frame(pred.confidence))`
@@
 
-genes = cbind(subset(ccgenes, grepl('TSG|oncogene', Role.in.Cancer) & Mutation.Types != 'T', select=c('Gene.Symbol','Genome.Location', colnames(loc))), 'value'=min(globalCNdf$mean.value)+2*sd(globalCNdf$mean.value), 'Row.names'='', 'Patient'='', 'selected'=as.factor(0))
+genes = cbind(subset(ccgenes, grepl('TSG|oncogene', Role.in.Cancer) & Mutation.Types != 'T', select=c('Gene.Symbol','Genome.Location', colnames(loc))), 'Row.names'='', 'Patient'='', 'selected'=as.factor(0))
+
+genes = merge(genes, globalCNdf %>% group_by(chr) %>% dplyr::summarise( value = min(mean.value)+sd(mean.value) ), by='chr')
 
@@
 
-ggsave(filename='plots/hazards/chr17.png', plot=g17, width=10, height=5, units='in', dpi=300)
+ggsave(filename='plots/hazards/chr17.png', plot=g17, width=7.5, height=3.7, units='in', dpi=300, scale=1.5)
 
@@
 ```{r echo=F, eval=T}
-dirA = '~/Data/Ellie/Analysis/downsampled5G_BEAdjacent'
-dirB = '~/Data/Ellie/Analysis/downsampled5G_NDBE'
+dirs = c('~/Data/Ellie/Analysis/downsampled5G_BEAdjacent',
+         '~/Data/Ellie/Analysis/downsampled5G_BEAdjacent_normal/',
+          '~/Data/Ellie/Analysis/downsampled5G_NDBE')
 
@@
 
-predA = list.files(dirA, 'predictions_all_exAHM0320.Rdata', full.names = T)
-predB = list.files(dirB, 'predictions_all_exAHM0320.Rdata', full.names = T)
-
-files = c()
-if (length(predA) < 1)
-  files = c(list.files(dirA, pattern='raw', full.names=T))
-if (length(predB) < 1)
-  files = c(files, list.files(dirB, pattern='raw', full.names=T))
-
-if (length(files) > 0) {
+predfiles = sapply(dirs, function(x) {
+  list.files(x, 'predictions_all_exAHM0320.Rdata', full.names = T)
+})
+
+predObj = list()
+for (i in 1:length(predfiles)) {
+  if (length(predfiles[[i]]) > 0) {
+      load(predfiles[[i]], verbose=T)
+      dspred$Type = basename(names(predfiles[i]))
+      predObj[[i]] = dspred
+      rm(dspred)
+  } else {
+    files = list.files(names(predfiles[i]), pattern='raw', full.names=T)
     sampleNames = sub('_raw.txt', '', basename(files))
     dspred = data.frame(matrix(ncol=2, nrow=length(sampleNames), dimnames=list(sampleNames,c('Prob','RR'))))
-
-  predA = sub('_raw.txt','',basename(c(list.files(dirA, pattern='raw', full.names=T))))
-  predB = sub('_raw.txt','',basename(c(list.files(dirB, pattern='raw', full.names=T))))
-
     for (n in 1:length(files)) {
@@
   }
-    dspred$Type = NA
-    dspred[predA,'Type'] = 'AdjacentBE'
-    dspred[predB,'Type'] = 'BEST2 NDBO'
-    pred = dspred
-
-    dspred = subset(pred, Type == 'AdjacentBE')
-
-    save(dspred, file=paste(dirA, 'predictions_all_exAHM0320.Rdata', sep='/'))
-
-    dspred = subset(pred, Type == 'BEST2 NDBO')
-    save(dspred, file=paste(dirB, 'predictions_all_exAHM0320.Rdata', sep='/'))
-
-    dspred = pred
-
-} else {
-  load(predB, verbose=T)
-  dspredB = dspred
-
-  load(predA, verbose=T)
-
-  dspred = rbind(dspred, dspredB)
-  dspred$Type = NA
-
-  dspred[sub('_raw.*','',basename(list.files(dirA, pattern='raw', full.names=T))),'Type'] = 'AdjacentBE'
-  dspred[sub('_raw.*','',basename(list.files(dirB, pattern='raw', full.names=T))),'Type'] = 'BEST2 NDBO'
+
+    dspred$Type = basename(names(predfilels[i]))
+    predObj[[i]] = dspred
+
+    save(dspred, file=paste(names(predfiles[i]), 'predictions_all_exAHM0320.Rdata', sep='/'))
+    rm(dspred)
+  }
+
+  dspred = do.call(rbind, predObj)
 

### 2018-03-12 processing_pipeline/insilico_test.R

@@ -1,0 +1,275 @@
+
+suppressPackageStartupMessages(library(ggplot2))
+suppressPackageStartupMessages(library(plyr))
+suppressPackageStartupMessages(library(dplyr))
+suppressPackageStartupMessages(library(gridExtra))
+suppressPackageStartupMessages(library(glmnet))
+suppressPackageStartupMessages(library(mice))
+suppressPackageStartupMessages(library(reshape2))
+suppressPackageStartupMessages(library(GenomicRanges))
+
+
+suppressPackageStartupMessages(source('../lib/load_patient_metadata.R'))
+suppressPackageStartupMessages(source('../lib/cv-pt-glm.R'))
+suppressPackageStartupMessages(source('../lib/data_func.R'))
+
+
+
+args = commandArgs(trailingOnly=TRUE)
+
+if (length(args) < 4)
+  stop("Missing required params: <data dir> <outdir> <info file dir>")
+
+data = args[1]
+#data = '~/Data/Ellie/Cleaned'
+outdir = args[2]
+#outdir = '~/Data/Ellie/Analysis'
+infodir = args[3]
+#infodir = '~/Data/Ellie/QDNAseq'
+
+x = (unclass(Sys.time()) + sample(1:1000, 1))
+cache.dir = paste(outdir, '5e6_arms_splits_exAHM0320', as.character(x) , sep='/')
+dir.create(cache.dir, showWarnings = F, recursive = T)
+
+## Hospital.Research.ID info file
+patient.file = list.files(infodir, pattern='All_patient_info.xlsx', recursive=T, full.names=T)
+demo.file = list.files(infodir, pattern='Demographics_full.xlsx', recursive=T, full.names=T)
+
+if ( length(patient.file) < 1 | length(demo.file) < 1)
+  stop("Missing files in info dir: All_patient_info.xlsx and Demographics_full.xlsx")
+
+patient.info = read.patient.info(patient.file, demo.file, set='All')$info
+patient.info = plyr::arrange(patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
+patient.info = subset(patient.info, Hospital.Research.ID != 'AHM0320')
+sum.patient.data = summarise.patient.info(patient.info)
+
+cleaned = list.files(path=data, pattern='tiled', full.names=T, recursive=T)
[…]

### 2018-03-16 AllSamplesModel.Rmd
...

@@
 all.patient.info = read.patient.info(patient.file, demo.file, set='all')$info
-head(all.patient.info)
+all.patient.info = subset(all.patient.info, Hospital.Research.ID != 'AHM0320')
 
@@
 sum.patient.data = summarise.patient.info(patient.info)
-
-sum.patient.data = subset(sum.patient.data, Hospital.Research.ID != 'AHM0320')
-
-pander(sum.patient.data[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Training patient set')
 
@@
 ggsave('plots/per_sample_cx.png', plot=pv, width=4, height=4, units='in', dpi=300)
-ggsave('plots/per_sample_cx_roc.png', plot=roc.plot(rocX, title='Per sample cx') + plot.theme , width=4, height=4, units='in', dpi=300)
+ggsave('plots/per_sample_cx_roc.png', plot=roc.plot(rocX, title='') + plot.theme , width=4, height=4, units='in', dpi=300)
 
@@
 ```{r model, message=F, warning=F, echo=F, fig.height=16, fig.width=16}
-folds = 10; splits = 5 
+#folds = 10; splits = 5 
 # file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
@@
 ```{r model_perf, message=F, warning=F, echo=F, fig.width=6, fig.height=6}
+folds = 10; splits = 5 
 classification.rate = as_tibble(matrix(ncol=5, nrow=0, dimnames=list(c(), c('model','mean','sme','stable.coef', 'all.coef'))))
@@
 
-ggsave('plots/performance/ridge_v_lasso.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)
+ggsave('plots/performance/ridge_v_lasso.png', plot=mp$plot, width=3.5, height=3.5, units='in', dpi=300, scale=1.2)
 
@@
 
-ggsave('plots/performance/ridge_v_lasso_nohgd.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)
+ggsave('plots/performance/ridge_v_lasso_nohgd.png', plot=mp$plot, width=3.5, height=3.5, units='in', dpi=300, scale=1.2)
 
@@
 rm(performance.at.1se, coefs, mp, no.hgd.plots)
+```
+
+#### LOO minus HGD
+
+```{r loonohgd, echo=F, message=F, warning=T, fig.width=5, fig.height=5}
+file = paste(cache.dir, 'loonohgd.Rdata', sep='/')
+if (file.exists(file)) {
+  message(paste("loading file", file))
+  load(file, verbose=T)
+} else {
+  stop(paste("No file", file, "found"))
+}
+
+pg.sampNOHGD = do.call(rbind, pg.sampNOHGD)
+perf = cbind.data.frame(perf=performance.at.1se, pt=unique(pg.sampNOHGD$Patient))
+
+p = ggplot(melt(perf), aes(x=variable,y=value, group=variable, fill=variable)) +
+    geom_boxplot(fill='lightcoral', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) +
+    geom_jitter(color='grey39', width=0.3) + plot.theme + 
+    geom_label(data=melt(subset(perf, perf %in% range(perf) )), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
+    theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')
+p
+
+ggsave('plots/performance/loo_perf_nohgd.png', plot=p, width=3, height=3, units='in', dpi=300)
+
+rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')
+
+```
+
+```{r, echo=F, fig.width=6, fig.height=6}
+hgdloo = ggplot(pg.sampNOHGD, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1), show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') + labs(title='Prediction using model w/o HGD', y='n Samples', x='Probability') + plot.theme
+hgdloo
+
+ggsave('plots/prediction_nohgd.png', plot=hgdloo, width=5, height=5, units='in', dpi=300)
+
+cuts = seq(0,1,0.1) 
+pg.sampNOHGD$quants = with(pg.sampNOHGD, cut(Prediction, breaks=cuts))
+
+noHGD.pred.confidence = as.data.frame.matrix(table(pg.sampNOHGD$quants, pg.sampNOHGD$Status))
+noHGD.pred.confidence = cbind(noHGD.pred.confidence, pg.sampNOHGD %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
+noHGD.pred.confidence$quant = rownames(noHGD.pred.confidence)
+
+ft.fun<-function(NP,P) {
+  f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
+  cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
+}
+
+noHGD.pred.confidence = noHGD.pred.confidence %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )
+
+noHGD.pred.confidence[c('r1','r2')] = NA
+for (i in 1:(length(cuts)-1)) 
+  noHGD.pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)
+
+noHGD.pred.confidence$Risk = 'Moderate'
+noHGD.pred.confidence$Risk[ which(with(noHGD.pred.confidence, p.value < 0.05 & perc < .5)) ] = 'Low'
+noHGD.pred.confidence$Risk[ which(with(noHGD.pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'
+
+noHGD.pred.confidence = bind_cols(noHGD.pred.confidence, with(noHGD.pred.confidence, 
+    data.frame(ci.low=qbeta(0.025, shape1=P+.5, shape2=NP+.5),
+               ci.high=qbeta(0.975, shape1=P+.5, shape2=NP+.5))))
+
+noHGD.pred.confidence$Risk = factor(noHGD.pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)
+
+nohgdcalib = ggplot(noHGD.pred.confidence, aes(mn, perc)) + 
+  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
+  scale_x_continuous(limits=c(0,1), breaks = cuts, labels = c(0.0, '','','','','',0.6,'',0.8,'',1.0)) +
+  geom_point() + 
+  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration', subtitle='HGD excluded model') + plot.theme + theme(legend.position = 'none')
+nohgdcalib
+
+ggsave('plots/calib_nohgd.png', plot=nohgdcalib, width=5, height=5, units='in', dpi=300)
+# Predict HGD samples
+file = paste(cache.dir, 'nohgd.Rdata', sep='/')
+if (file.exists(file)) {
+  load(file, verbose=T)
+
+  fitNOHGD = models[[select.alpha]]
+
+  lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']
+
+  hgd.samples = subset(patient.info, Pathology %in% c('HGD','IMC'))
+  hgd.df = dysplasia.df[hgd.samples$Samplename,]
+
+  pred = predict(fitNOHGD, newx=hgd.df, s=lambda.opt, type='response')
+  hgd.samples$Prediction = pred
+
+  hgd.pred = hgd.samples %>% group_by(Patient, Hospital.Research.ID, Status, Endoscopy.Year) %>% 
+    summarise( max=max(Pathology), pred=Prediction[which.max(Pathology)] )
+
+  rm(no.hgd.plots, coefs, performance.at.1se, models)
+}
+
+#hist(hgd.samples$Prediction)
+
+rm(plots, performance.at.1se,coefs, nzcoefs, fits)
 ```
@@
 
-ggsave('plots/performance/no_lgd_perf.png', plot=mp$plot, height=5, width=5, units='in', dpi=300)
+ggsave('plots/performance/no_lgd_perf.png', plot=mp$plot, height=3.5, width=3.5, units='in', dpi=300, scale=1.2)
 
@@
 rm(performance.at.1se, coefs, mp, nolgd.plots)
-```
-
-#### LOO minus HGD
-
-```{r loonohgd, echo=F, message=F, warning=T, fig.width=5, fig.height=5, eval=F}
-file = paste(cache.dir, 'loonohgd.Rdata', sep='/')
-if (file.exists(file)) {
-  message(paste("loading file", file))
-  load(file, verbose=T)
-} else {
-  stop(paste("No file", file, "found"))
-}
-
-pg.sampNOHGD = do.call(rbind, pg.sampNOHGD)
-
-perf = cbind.data.frame(perf=performance.at.1se, pt=unique(pg.sampNOHGD$Patient))
-
-p = ggplot(melt(perf), aes(x=variable,y=value, group=variable, fill=variable)) +
-    geom_boxplot(fill='lightcoral', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) +
-    geom_jitter(color='grey39', width=0.3) + plot.theme + 
-    geom_label(data=melt(subset(perf, perf %in% range(perf) )), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
-  #geom_text_repel(aes(label=pt)) +
-    theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')
-p
-ggsave('plots/performance/loo_perf_nohgd.png', plot=p, width=5, height=5, units='in', dpi=300)
-
-rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')
-roc.plot(rocNOHGD, 'No HGD')  
-```
-
-```{r, echo=F, fig.width=6, fig.height=6,eval=F}
-ggplot(pg.sampNOHGD, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1), show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') + labs(title='Prediction using model w/o HGD', y='n Samples', x='Probability') + plot.theme
-
-cuts = seq(0,1,0.1) 
-pg.sampNOHGD$quants = with(pg.sampNOHGD, cut(Prediction, breaks=cuts))
-
-noHGD.pred.confidence = as.data.frame.matrix(table(pg.sampNOHGD$quants, pg.sampNOHGD$Status))
-noHGD.pred.confidence = cbind(noHGD.pred.confidence, pg.sampNOHGD %>% group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
-noHGD.pred.confidence$quant = rownames(noHGD.pred.confidence)
-
-ft.fun<-function(NP,P) {
-  print(paste(NP, P))
-  f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
-  cbind.data.frame('p.value'=round(f$p.value, 5), 'odds.ratio'=round(f$estimate, 4))
-}
-
-noHGD.pred.confidence = noHGD.pred.confidence %>% as.data.frame.matrix %>% group_by(quant) %>% 
-  dplyr::mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )
-
-noHGD.pred.confidence[c('r1','r2')] = NA
-for (i in 1:(length(cuts)-1)) 
-  noHGD.pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)
-
-noHGD.pred.confidence$Risk = 'Moderate'
-noHGD.pred.confidence$Risk[ which(with(noHGD.pred.confidence, p.value < 0.05 & perc < .5)) ] = 'Low'
-noHGD.pred.confidence$Risk[ which(with(noHGD.pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'
-
-noHGD.pred.confidence = bind_cols(noHGD.pred.confidence, with(noHGD.pred.confidence, 
-    data.frame(ci.low=qbeta(0.025, shape1=P+.5, shape2=NP+.5),
-               ci.high=qbeta(0.975, shape1=P+.5, shape2=NP+.5))))
-
-noHGD.pred.confidence$Risk = factor(noHGD.pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)
-
-ggplot(noHGD.pred.confidence, aes(mn, perc)) + 
-  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
-  geom_point() + geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
-  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration', subtitle='HGD excluded model') + plot.theme + theme(legend.position = 'none')
-
-# Predict HGD samples
-file = paste(cache.dir, 'nohgd.Rdata', sep='/')
-if (file.exists(file)) {
-  load(file, verbose=T)
-
-  fitNOHGD = models[[select.alpha]]
-
-  lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']
-
-  hgd.samples = subset(patient.info, Pathology %in% c('HGD','IMC'))
-  hgd.df = dysplasia.df[hgd.samples$Samplename,]
-
-  pred = predict(fitNOHGD, newx=hgd.df, s=lambda.opt, type='response')
-  hgd.samples$Prediction = pred
-
-  hgd.pred = hgd.samples %>% group_by(Patient, Hospital.Research.ID, Status, Endoscopy.Year) %>% 
-    dplyr::summarise( max=max(Pathology), pred=Prediction[which.max(Pathology)] )
-
-  rm(no.hgd.plots, coefs, performance.at.1se, models)
-}
-
-#hist(hgd.samples$Prediction)
-
-rm(plots, nzcoefs, fits)
 ```
@@
 
+```{r echo=F, eval=F}
 ## All LOO
-
-```{r echo=F, eval=F}
-
 files = list.files('~/Data/Ellie/Analysis/5e6_arms_LOO', full.names = T)
@@
 for (pt in sum.patient.data$Hospital.Research.ID) {
-  ptF = grep(pt, files, value=T)
+  ptF = grep(pt, files, value=F)
   load(grep('loo\\.Rdata', ptF, value=T), verbose=T)
@@
 
-load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/PR1_HIN_042_loo.Rdata", verbose=T)
-load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/PR1_HIN_042_model_data.Rdata", verbose=T)
-
-load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/AHM0320_loo.Rdata", verbose=T)
-load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/AHM0320_model_data.Rdata", verbose=T)
+#load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/PR1_HIN_042_loo.Rdata", verbose=T)
+#load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/PR1_HIN_042_model_data.Rdata", verbose=T)
+
+#load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/AHM0320_loo.Rdata", verbose=T)
+#load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/AHM0320_model_data.Rdata", verbose=T)
 
@@
 Based on these curves we get our best AUC at ~0.36.  Currently we are using 0.5 by default.
-
-```{r insilico_rocs, echo=F, message=F, warning=F, fig.height=8, fig.width=8}
-
-load(file='~/Data/Ellie/Analysis/5e6_arms_all/insilico_splits.Rdata', verbose=T)
-
-```
 
@@
 
+load(file=paste(cache.dir, 'insilico_splits.Rdata', sep='/'), verbose=T)
 insilico.rocs[['all']] = roc
@@
 p
-ggsave(filename='plots/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)
+ggsave(filename='plots/auc/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)
 
@@
   geom_label(data=subset(best, model == 'all' | auc %in% c(min(auc), max(auc)) ), aes(specificity, sensitivity, label=paste('AUC ',round(auc, 2), sep='\n'))) +
-  labs(title='ROC', x='Specificity (1-FPR)', y='Sensitivity (TPR)')  + plot.theme +
+  labs(title='ROC CI', x='Specificity (1-FPR)', y='Sensitivity (TPR)')  + plot.theme +
   theme(legend.position = 'none')
 rr  
-ggsave(filename='plots/roc_CI.png', plot=rr, width=6, height=6, units='in', dpi=300)
+ggsave(filename='plots/auc/roc_CI.png', plot=rr, width=4, height=4, units='in', dpi=300, scale=1.5)
 
@@
 comp = ggplot(melt(subset(m, Measure == 'AUC'), measure.vars='value'), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Measure)) + 
+  ylim(0,100) + 
   geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
@@
   scale_fill_manual(values=cols, name='') + plot.theme + 
-  theme(legend.position = 'none') + labs(x='', y='AUC', title='ROC Comparison') 
+  theme(legend.position = 'none', axis.text.x = element_text(hjust=1,angle=45)) + labs(x='', y='AUC', title='') 
 comp
 
-ggsave('plots/auc/path_p53_cn_auc.png', plot=comp, width=2.7, height=3.7, units="in", limitsize=F, dpi = 300, scale=2)
+ggsave('plots/auc/path_p53_cn_auc.png', plot=comp, width=3.1, height=3.9, units="in", limitsize=F, dpi = 300, scale=1)
 
@@
 transform.by.time<-function(pt, info) {
-  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction', 'OR')] %>% group_by(PID, Samplename) %>% dplyr::mutate(
-    'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12)
-
-  pt = arrange(merge(pt, subset(info, Hospital.Research.ID == pt$Hospital.Research.ID[1], select=c('Samplename','Block')), by='Samplename'), Pathology, PID, Endoscopy.Year)
-
-  tb = with(pt, table(Endoscopy.Year, PID))
+  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction', 'OR')] %>% group_by(PID, Samplename) %>% dplyr::mutate('months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12)
+  pt = dplyr::arrange(merge(pt, subset(info, Hospital.Research.ID == pt$Hospital.Research.ID[1], select=c('Samplename','Block')), by='Samplename'), Pathology, PID, Endoscopy.Year)
+
+  tb = with(pt, table(months.before.final, PID))
   tb[tb>0] = 1
-  for (yr in names(which(rowSums(tb) > 1))) {
-    tmp = pt[which(pt$Endoscopy.Year == yr),]
-    tmp = tmp %>% group_by(PID) %>% summarise(max(Pathology))
-
-    match = which(pt$Endoscopy.Year == yr & pt$PID == as.character(tmp[which.min(tmp$`max(Pathology)`), 'PID']))
-    pt[match,'months.before.final'] = pt[match,'months.before.final'] + 6
+  for (mnths in names(which(rowSums(tb) > 1))) {
+    tmp = pt[which(pt$months.before.final == mnths),]
+    tmp = arrange(tmp, rev(Pathology), PID)
+
+    mt = tmp$months.before.final[1] 
+    interval = length(tmp$Endoscopy.Year)/(length(unique(tmp$PID))-1)
+    for (pid in unique(tmp$PID)) {
+      rows = which(tmp$PID == pid)
+      if (length(grep('HGD|IMC',tmp[rows,'Pathology'])) > 0 & mt != 0) {
+        mt = 0
+        tmp$months.before.final[rows] = mt
+      } else {
+        mt = mt + interval
+        tmp$months.before.final[rows] = mt
+      }
+    }
+    pt[which(pt$months.before.final == mnths),] = tmp
   }
@@
   pt$nl = (pt$Block-1)*2
-
-  tb = table(pt[,c('months.before.final', 'PID')])
-  for (i in 1:nrow(tb)) {
-    pids = tb[i,]
-    yrs = which(pids > 0)
-    if (length(yrs) > 1) {
-      for (j in 2:length(yrs)) 
-        pt[which(pt$PID == names(yrs)[j]), 'nl'] = pt[which(pt$PID == names(yrs)[j]), 'nl'] + 1
-    }
-  }
   arrange(pt, -Endoscopy.Year, PID)
@@
   geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols, limits=levels(pred.confidence$Risk) ) +
-  geom_point() + geom_abline(slope=1, intercept=0, color='grey39', linetype='dotted') + #geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
-  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration') +
-  scale_color_manual(values=riskCols,limits=levels(pred.confidence$Risk) ) + plot.theme + theme(legend.position = 'none')
+  geom_point() + geom_abline(slope=1, intercept=0, color='grey39', linetype='dotted', size=0.5) + #geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
+  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), size=0.5, width=0.01) + labs(x='P(P)', y='NP:P', title='') +
+  scale_color_manual(values=riskCols,limits=levels(pred.confidence$Risk) ) + 
+  scale_x_continuous(limits=c(0,1), breaks = cuts, labels = c(0.0, '','',0.3,'','','',0.7,'','',1.0)) +
+  plot.theme + theme(legend.position = 'none')
 pcal
 
-ggsave(filename='plots/pred_calib.png', plot=pcal, width=2.2, height=2.2, units='in', dpi=300, scale=2)
+ggsave(filename='plots/pred_calib.png', plot=pcal, width=2.2, height=2.2, units='in', dpi=300, scale=1.5)
 
@@
 unadjRR = ggplot(back, aes(OR)) + geom_histogram(aes(fill=..x..), bins=15, show.legend = F) +
-  #scale_fill_distiller(palette = 'RdYlBu', type='div',name='P(P)') +
   scale_fill_gradientn(colors = myPal,  name='') + 
@@
 unadjRR
-ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR, width=5.5, height=4.3, units='in', dpi=300)
+ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR + labs(title=''), width=5.5, height=4.3, units='in', dpi=300)
 
@@
 
-ggsave(filename='plots/feature_jacknife.png', plot=p, width=6, height=8, unit='in', dpi=300)
+ggsave(filename='plots/feature_jacknife.png', plot=p, width=6, height=10, unit='in', dpi=300)
 
@@
   scale_color_manual(values=c('white','black','white'), limits=levels(pred.confidence$Risk) ) +
-  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101.5), position=position_stack(), size=4) +
-  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep=''), color=Risk), position=position_stack(vjust = 0.5), size=5, show.legend = F) +
+  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=102), position=position_stack()) +
+  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep=''), color=Risk), position=position_stack(vjust = 0.5), show.legend = F) +
   facet_grid(~Status, scales='free_x', space = 'free_x') + plot.theme + 
-  labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
+  labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'none')
 rp
 
-ggsave(filename='plots/pred_by_path.png', plot=rp, width=5.8, height=4.5, units='in', dpi=300, scale=1.5)
+ggsave(filename='plots/pred_by_path.png', plot=rp, width=5.8, height=4.5, units='in', dpi=300, scale=1)
 
@@
   scale_fill_manual(values=col, name='p53 IHC', labels=c('Normal','Aberrant')) + 
-  geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=102)) +
+  geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=103)) +
   geom_text(aes(label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5,color='white') +
-  plot.theme + labs(title='Samples predicted by aberrant p53 IHC', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
+  plot.theme + labs(title='', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
 rp
-ggsave(filename='plots/p53_pred.png', plot=rp, width=4.5, height=3.6, units='in', dpi=300, scale=1.5)
+ggsave(filename='plots/p53_pred.png', plot=rp, width=4.5, height=3.6, units='in', dpi=300, scale=1.2)
 
@@
 endos = back %>% group_by(Status, months.before.final, brks, PID) %>% dplyr::summarise(
+  'pred.endo'=ifelse(length(which(Risk > 'Moderate')) > 0, 1, 0), 
+  'n.endos'=length(unique(PID)), 
+  'n.samples'=length(Samplename),
+  'mean.pred'=mean(Prediction),
+  'pred.samples'=length(which(Risk > 'Moderate')),
+  'max.path'=max(Pathology),
+  'Moderate'=length(which(Risk == 'Moderate')),
+  'High'=length(which(Risk == 'High'))
+  )
+
+endos.mod = back %>% group_by(Status, months.before.final, brks, PID) %>% dplyr::summarise(
   'pred.endo'=ifelse(length(which(Risk >= 'Moderate')) > 0, 1, 0), 
@@
 
-endos.m = subset(endos, Status == 'P') %>% group_by(brks) %>% dplyr::summarise( 
+endos.mod$brks = cut(endos.mod$months.before.final, include.lowest = T, ordered_result = T,breaks=sq,labels=mlab)
+endos.mod$brks = with(endos.mod, factor(brks, rev(levels(brks)))) # invert order
+
+endos.r = subset(endos, Status == 'P') %>% group_by(brks) %>% dplyr::summarise( 
+  'pred.r'=sum(pred.endo)/sum(n.endos),'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos), 
+  'mod.r'=sum(Moderate)/sum(n.samples), 'high.r'=sum(High)/sum(n.samples))
+
+endos.m = subset(endos.mod, Status == 'P') %>% group_by(brks) %>% dplyr::summarise( 
   'pred.r'=sum(pred.endo)/sum(n.endos),'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos), 
@@
 
+endos.r = arrange(merge(endos.r,tmp), brks)
 endos.m = arrange(merge(endos.m,tmp), brks)
 
-rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) + ylim(0,105) +
-  geom_bar(stat='identity', fill='orangered4', show.legend = F) + 
+rp = ggplot(endos.r, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) + ylim(0,105) +
+#  geom_bar(data=endos.m, aes(x=brks, y=pred.r*100, group=brks), stat='identity', fill=riskCols[2], show.legend = F, alpha=0.5) +
+#  geom_text(data=endos.m, aes(y=round(pred.r*100+2), label=paste(round(pred.r*100), '%', sep='')), color='grey45') +
+  geom_bar(stat='identity', fill=riskCols[3], show.legend = F) + 
   geom_errorbar(color='grey21', width=0.2) +
   geom_text(aes(y=3, label=paste(round(pred.r*100), '%', sep='')), color='white') +
-  geom_text(aes(y=pred.r*100-5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
+  geom_text(aes(y=pred.r*100-8, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
   labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') +
-  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12)) 
+  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1) ) 
 rp
-ggsave(filename='plots/pred_endo.png', plot=rp, width=5.6, height=4.4, units='in', dpi=300, scale=1.5)
+ggsave(filename='plots/pred_endo.png', plot=rp, width=5.8, height=4.5, units='in', dpi=300, scale=1)
 
@@
 
-```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=5 }
+```{r heatmaps, echo=F, warning=F, message=F, fig.width=5, fig.height=5 }
 back$nl = factor(back$nl, ordered=T)
@@
 table(rev(arrange(subset(back, Patient == 19), brks)$brks))
-
-p = pred.tiles(subset(back, Patient %in% c(6)), ncol=1, cols=myPal, probs=T) + labs(title='')
-tilesLegend = get.legend(p)
-p + theme(legend.position = 'none')
 
@@
 ```{r, echo=F, message=F, warning=F, width=26, height=8}
-back = back %>% group_by(Patient) %>% dplyr::mutate( 'pred.ratio'=round(sum(Prediction)/length(Prediction),2))
-
-p1 = pred.tiles(subset(back, Status == 'NP'), cols=myPal, probs=T, nrow=2, scales='free', drop=T) + labs(title='Non-Progressors') 
-#p1 = pred.tiles(subset(back, Status == 'NP'), cols=riskCols, probs=F, OR=F, nrow=2) + labs(title='Non-Progressors') 
-riskLegend = get.legend(p1)
-p1 = p1 + theme(legend.position = 'bottom')
-riskLegendLong = get.legend(p1)
-
-p1 = p1 + theme(legend.position = 'none')
-p2 = pred.tiles(subset(back, Status == 'P'), probs=T, cols=myPal,nrow=2,scales='free',drop=T) + labs(title='Progressors') + theme(legend.position = 'none')
-
-ggsave('plots/pred_heatmap_ALL.png', plot=grid.arrange(p2 + labs(x='', y='',title=''),p1 + labs(x='', y='',title=''),nrow=2), width=11, height=3.4, units="in", limitsize=F, dpi = 300, scale = 3)
-ggsave('plots/npg_heatmap.png', plot=p1, width=52, height=8, units="in", limitsize=F, dpi = 300)
-ggsave('plots/pg_heatmap.png', plot=p2, width=52, height=8, units="in", limitsize=F, dpi = 300)
-ggsave('plots/risk_legend.png', plot=grid.arrange(riskLegend), width=2,height=4, dpi=300)
-ggsave('plots/risk_legend_long.png', plot=grid.arrange(riskLegendLong), width=8,height=2, dpi=300)
-
-p1 = pred.tiles(subset(back, Status == 'NP'), cols=myPal, probs=T, nrow=2, scales='free_x', drop=T) + labs(title='Non-Progressors') + theme(legend.position = 'none') 
-p2 = pred.tiles(subset(back, Status == 'P'), probs=T, cols=myPal,nrow=2,scales='free_x',drop=T) + labs(title='Progressors') + theme(legend.position = 'none')
-ggsave('plots/pred_heatmap_ALL_widthadj.png', plot=grid.arrange(p2 + labs(x='', y='',title=''),p1 + labs(x='', y='',title=''),nrow=2), width=11, height=3.4, units="in", limitsize=F, dpi = 300, scale = 3)
+back = back %>% group_by(Patient) %>% dplyr::mutate( 'pred.ratio'=round(sum(Prediction)/length(Prediction),2), 'n.endos'=length(unique(months.before.final)), 'max.block'=max(Block))
+back = transform(back, Patient=reorder(Patient, max.block))
+back$Block = factor(back$Block, ordered = T)
+
+pred.tiles(subset(back, Patient %in% c(1,20,34,21,3, 19)), cols=myPal, probs=T, space='free', scales='free_x', drop=T) + theme(legend.position = 'none')
+pred.tiles(subset(back, Patient %in% c(1,20,34,21,3, 19)), cols=riskCols, probs=F, space='free', scales='free_x', drop=T) + theme(legend.position = 'none')
+
+w = 11; h = 1.2; s = 1
+
+get.w.scale <- function(b) {
+  n = sum((b %>% group_by(Patient) %>% summarise(unique(n.endos)))[,2])
+  #round(n*0.5)+1
+  n
+}
+
+get.h.scale <- function(b, max=7) {
+  n = max(b$max.block)
+  if (n <= 2) n = 2
+  if (n > 2 & n <= 3) n = 3
+  if (n > 3 & n < 6) n = 5
+  if (n >= 6) n = max
+  n
+}
+
+x = sapply(subset(sum.patient.data, Status == 'P')$Patient, function(i) {
+  b = subset(back, Patient == i)
+  get.h.scale(b)
+})
+sort(x)
+
+plots = list(); widths = c()
+for (i in subset(sum.patient.data, Status == 'P')$Patient) {
+  b = subset(back, Patient == i)
+  lim = c(1:7)
+  if (b$max.block[1] < 6) lim = c(1:5)
+  if (b$max.block[1] > 2 & b$max.block[1] < 4) lim = c(1:3)
+  if (b$max.block[1] <= 2) lim = c(1:2)
+
+  p = pred.tiles(b, cols=riskCols, ylim=lim,  probs=F, scales='free', space='free') + labs(title='', x='', y='') + theme(legend.position = 'none', strip.text = element_blank(), plot.margin = unit(c(0,0.2,0,0), 'cm'))
+  ggsave(paste('plots/heatmaps/P/pt_',i,'.png', sep=''), plot=p, height=get.h.scale(b), width=get.w.scale(b), unit='in', dpi=300 )
+}
+b = subset(back, Patient == 32)
+lim = c(1:4)
+p = pred.tiles(b, cols=riskCols, ylim=lim,  probs=F, scales='free', space='free') + labs(title='', x='', y='') + theme(legend.position = 'none', strip.text = element_blank(), plot.margin = unit(c(0,0.2,0,0), 'cm'))
+  ggsave(paste('plots/heatmaps/P/pt_32.png', sep=''), plot=p, height=4, width=get.w.scale(b), unit='in', dpi=300 )
+
+x = sapply(subset(sum.patient.data, Status == 'NP')$Patient, function(i) {
+  b = subset(back, Patient == i)
+  get.h.scale(b, 9)
+})
+sort(x)
+
+for (i in subset(sum.patient.data, Status == 'NP')$Patient) {
+  b = subset(back, Patient == i)
+  lim = c(1:9)
+  if (b$max.block[1] < 6) lim = c(1:5)
+  if (b$max.block[1] < 4) lim = c(1:3)
+  if (b$max.block[1] < 3) lim = c(1:2)
+
+  p = pred.tiles(b, cols=riskCols, ylim=lim,  probs=F, scales='free', space='free') + labs(title='', x='', y='') + theme(legend.position = 'none', strip.text = element_blank(), plot.margin = unit(c(0,0.2,0,0), 'cm'))
+  ggsave(paste('plots/heatmaps/NP/pt_',i,'.png', sep=''), plot=p, height=get.h.scale(b,9), width=get.w.scale(b), unit='in', dpi=300 )
+}
+
+all = pred.tiles(back, cols=riskCols,  probs=F, scales='free', space='free') + labs(title='', x='', y='') + theme(legend.position = 'bottom')
+riskLegend = get.legend(all)
+
+ggsave('plots/heatmaps/legend_long.png', plot=grid.arrange(riskLegend), width=7, height=1, unit='in', dpi=300)
+```
+
+```{r echo=F}
+# By RR instead
+
+plots = list(); widths = c()
+for (i in subset(sum.patient.data, Status == 'P')$Patient) {
+  b = subset(back, Patient == i)
+  lim = c(1:7)
+  if (b$max.block[1] < 6) lim = c(1:5)
+  if (b$max.block[1] > 2 & b$max.block[1] < 4) lim = c(1:3)
+  if (b$max.block[1] <= 2) lim = c(1:2)
+
+  p = pred.tiles(b, cols=myPal, ylim=lim,  probs=T, scales='free', space='free') + labs(title='', x='', y='') + theme(legend.position = 'none', strip.text = element_blank(), plot.margin = unit(c(0,0.2,0,0), 'cm'))
+  ggsave(paste('plots/heatmaps/P_RR/pt_',i,'.png', sep=''), plot=p, height=get.h.scale(b), width=get.w.scale(b), unit='in', dpi=300 )
+}
+b = subset(back, Patient == 32)
+lim = c(1:4)
+p = pred.tiles(b, cols=myPal, ylim=lim,  probs=F, scales='free', space='free') + labs(title='', x='', y='') + theme(legend.position = 'none', strip.text = element_blank(), plot.margin = unit(c(0,0.2,0,0), 'cm'))
+  ggsave(paste('plots/heatmaps/P_RR/pt_32.png', sep=''), plot=p, height=4, width=get.w.scale(b), unit='in', dpi=300 )
+
+for (i in subset(sum.patient.data, Status == 'NP')$Patient) {
+  b = subset(back, Patient == i)
+  lim = c(1:9)
+  if (b$max.block[1] < 6) lim = c(1:5)
+  if (b$max.block[1] < 4) lim = c(1:3)
+  if (b$max.block[1] < 3) lim = c(1:2)
+
+  p = pred.tiles(b, cols=myPal, ylim=lim,  probs=T, scales='free', space='free') + labs(title='', x='', y='') + theme(legend.position = 'none', strip.text = element_blank(), plot.margin = unit(c(0,0.2,0,0), 'cm'))
+  ggsave(paste('plots/heatmaps/NP_RR/pt_',i,'.png', sep=''), plot=p, height=get.h.scale(b,9), width=get.w.scale(b), unit='in', dpi=300 )
+}
+
+all = pred.tiles(back, cols=myPal,  probs=T, scales='free', space='free') + labs(title='', x='', y='') + theme(legend.position = 'bottom')
+rrLegend = get.legend(all)
+
+ggsave('plots/heatmaps/legend_long_RR.png', plot=grid.arrange(rrLegend), width=7, height=1, unit='in', dpi=300)
 ```
@@
 length(unique(p11Loss$Patient))/length(subset(sum.patient.data, Status == 'P')$Patient)
+table(p11Loss$Pathology)
 # Of those where we see loss of p arm, p53 IHC shows an "abnormal" result 33% of the time
 
-table(subset(patient.info, Samplename %in% p11Loss$variable, select='p53.Status'))/length(p11Loss$variable)
+table(subset(patient.info, Samplename %in% p11Loss$variable,select='p53.Status'))/length(p11Loss$variable)
 
@@
   print(chr)
-  p = plot.cn.chr(subset(globalCNdf, seg.type != 'arms') , chrom=chr, chr.lengths, genes=genes, haz=chh)
+  p = plot.cn.chr(globalCNdf, chrom=chr, chr.lengths, genes=genes, haz=chh)
   f = paste('chr',chr,'.png',sep='')
@@
 dirs = c('~/Data/Ellie/Analysis/downsampled5G_BEAdjacent',
-         '~/Data/Ellie/Analysis/downsampled5G_BEAdjacent_normal/',
-          '~/Data/Ellie/Analysis/downsampled5G_NDBE')
+         '~/Data/Ellie/Analysis/downsampled5G_BEAdjacent_normal/') #,'~/Data/Ellie/Analysis/downsampled5G_NDBE')
 
@@
       segM = t(segM)
-      for (i in 1:ncol(segM)) 
-        segM[,i] = unit.var(segM[,i], z.mean[i], z.sd[i])
+      for (j in 1:ncol(segM)) 
+        segM[,j] = unit.var(segM[,j], z.mean[j], z.sd[j])
 
@@
       armsM = t(armsM)
-      for (i in 1:ncol(armsM)) 
-        armsM[,i] = unit.var(armsM[,i], z.arms.mean[i], z.arms.sd[i])
+      for (j in 1:ncol(armsM)) 
+        armsM[,j] = unit.var(armsM[,j], z.arms.mean[j], z.arms.sd[j])
 
@@
 
-    dspred$Type = basename(names(predfilels[i]))
+    dspred$Type = basename(names(predfiles[i]))
     predObj[[i]] = dspred
@@
   dspred = do.call(rbind, predObj)
-
 }
@@
 
-adj = subset(dspred, Type == 'AdjacentBE')
+dspred$Type[grep('normal', dspred$Type)] = 'Normal blood'
+dspred = subset(dspred, !grepl('normal', Type))
+dspred$Type[grep('BEAdjacent', dspred$Type)] = 'Tumor adjacent BE'
+
 prg = back[,c('Prediction', 'OR', 'Status')]
@@
 #  subset(back, Status == 'P', select=c('Prediction', 'OR', 'Status'))
-colnames(prg) = colnames(adj)
-m = rbind(adj,prg)
-m$Type = factor(m$Type, levels=c('Non-Progressor','BEST2 NDBO','Progressor','AdjacentBE'), ordered = T)
-
+colnames(prg) = colnames(dspred)
+dspred = rbind(dspred,prg)
+dspred$Type = factor(dspred$Type, levels=c('Normal blood','Non-Progressor','Progressor','Tumor adjacent BE'), ordered = T)
+
+colors = myPal[seq(2, length(myPal), round(length(myPal)/4))]
+pr = ggplot(dspred, aes(Type, RR, fill=Type)) + 
+  geom_jitter(width=0.2, aes(color=RR)) + 
+  scale_color_gradientn(colors = myPal,  name='') + 
+  geom_boxplot(outlier.colour = NA) + 
+  scale_fill_manual(values = colors,  name='') + 
+  labs(x='', y='RR', title='Predicted Relative Risk') + 
+  theme(legend.position = 'none', axis.text.x = element_text(angle=45,hjust=1)) + plot.theme 
+pr
+ggsave('plots/downsampled_predictions_supplement.png', plot=pr, width=4, height=4.38,dpi=300, scale=1.1)
+
+m = subset(m, Type != 'Normal blood')
+m$Type = droplevels(m$Type)
 pr = ggplot(m, aes(Type, RR, fill=Type)) + 
-  geom_jitter(width=0.2) + geom_boxplot() + 
+  geom_jitter(width=0.2, aes(color=RR)) + 
+  scale_color_gradientn(colors = myPal,  name='') + 
+  geom_boxplot(outlier.colour = NA) + 
+  scale_fill_manual(values = riskCols,  name='') + 
   labs(x='', y='RR', title='Predicted Relative Risk') + 
-  theme(legend.position = 'none', text=element_text(size=11)) + plot.theme 
+  theme(legend.position = 'none') + plot.theme 
 pr
 
-ggsave('plots/adjacentBE.png', plot=pr, width=8, height=8,dpi=300)
+ggsave('plots/downsampled_predictions.png', plot=pr, width=4, height=4.38,dpi=300, scale=1.1)
+
+```
+
+## Perfectly diploid?
+
+```{r synthetic_normal}
+x = matrix(nrow=16, ncol=ncol(dysplasia.df), data=0)
+colnames(x) = colnames(dysplasia.df)
+
+for (i in 2:5)
+  x[i,] = sapply(x[i,], function(x) sample(seq(0,1,0.01), 1), simplify = T)
+
+for (i in 6:10)
+  x[i,] = sapply(x[i,], function(x) sample(seq(-1,0,0.01), 1), simplify = T)
+
+for (i in 11:16)
+  x[i,] = sapply(x[i,], function(x) sample(seq(-1,1,0.01), 1), simplify = T)
+
+sparsed_test_data <- Matrix(data=0, nrow=nrow(x),  ncol=ncol(x),
+                                  dimnames=list(rownames(x),colnames(x)), sparse=T)
+for(col in colnames(x)) sparsed_test_data[,col] = x[,col]
+
+#predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='response')
+predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='link')
 
@@
 vigLeg = get.legend(p)
+ggsave('plots/vignette/legend.png', plot=grid.arrange(vigLeg), width=10, height=1.5, dpi=300)
+
 p = p + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_34.png', plot=p, width=6, height=6,dpi=300)
+ggsave('plots/vignette/vig_34.png', plot=p, width=6, height=3.9,dpi=300)
 
@@
 p = vig.plot(subset(back, Patient == 20), subset(patient.info, Patient == 20),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_20.png', plot=p, width=6, height=6,dpi=300)
+ggsave('plots/vignette/vig_20.png', plot=p, width=6, height=3.9,dpi=300)
 
@@
 p = vig.plot(subset(back, Patient == 3), subset(patient.info, Patient == 3),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_3.png', plot=p, width=6, height=6,dpi=300)
+ggsave('plots/vignette/vig_3.png', plot=p, width=6, height=3.9,dpi=300)
 
@@
 p = vig.plot(subset(back, Patient == 68), subset(patient.info, Patient == 68),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_68.png', plot=p, width=6, height=4,dpi=300)
+ggsave('plots/vignette/vig_68.png', plot=p, width=6, height=3.9,dpi=300)
 
@@
 
-path.rx<-function(Patient, brks, Pathology) {
-  rules = as.data.frame(matrix(ncol=0, nrow=0))
-  # Consecutive
-  for (i in 1:length(brks)) {
-    #risks = table(Pathology[i:(i+1)])
-    rule = 'None'
-    if ( length(which(grepl('HGD|IMC', Pathology[i:(i+1)]))) > 0  |
-         length(which(grepl('LGD', Pathology[i:(i+1)]))) > 1 ) {
-      rule = 1
-    } else if ( length(which(grepl('LGD', Pathology[i:(i+1)]))) > 0 ) {
-      rule = 2
-    } else if ( length(which(grepl('BE|ID', Pathology[i:(i+1)]))) > 1  ) {
-      rule = 3
-    }
-  rules = rbind(rules, cbind( as.character(brks[i]), as.character(brks[(i+1)]), rule))
-  }  
-  colnames(rules) = c('t1','t2', 'rule')
-  rules$Patient = Patient
-  rules
-}
-
-rules2 = do.call(rbind, lapply(unique(maxP$Patient), function(pt) {
-  r = with(arrange(subset(maxP, Patient == pt), brks), path.rx(pt, brks,Pathology))
-  r$Status = unique(subset(maxP, Patient == pt)$Status)
-  r
-}))
-rules2$rule = factor(rules$rule, levels=c('None',4,3,2,1), ordered = T)
-
-rules2$t1 = factor(rules2$t1, levels=levels(back$brks), ordered=T)
-rules2$t2 = factor(rules2$t2, levels=levels(back$brks), ordered=T)
-
-second = subset(rules2,  t2 != 'Endpoint') %>% 
-  group_by(Patient, Status) %>% 
-  summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )
-
-pander(rbind('Progressors'=table(subset(second, Status == 'P')$rule),
-            'Non-Progressors'=table(subset(second, Status == 'NP')$rule)), caption='Two endos, pathology rules')
+with(second, table(Status, rule))
+```
+
+## Incidence
+
+Median follow up for our discovery cohort was `r median(sum.patient.data$end.year-sum.patient.data$start.year)`.
+
+Adjust the RR of the orginal model with 1-3.5% offset to get an absolute risk (AR) with the understanding that the cohort was age matched and unbalanced.
+
+```{r echo=F}
+file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
+load(file, verbose=T)
+rm(plots, coefs, cv.patient)
+
+fit = models[[select.alpha]]
+lambda = performance.at.1se[[select.alpha]]$lambda
+
+# number of patients we'd needed to have included to get 45 progressors from a representative population (I think)
+all = sum.patient.data
+```
+
+```{r estimateRisk, echo=F, warning=F, message=F, eval=F}
+# Or probability an individual will progress given a rate of 1%-3.5%
+poft = sapply(seq(0.96,0.99, .01), function(x) {
+ sapply(1:5, function(y) 1-(x^y) )
+})
+
+#poft = as.data.frame(poft)
+#colnames(poft) = paste((1-seq(0.96,0.99, .01))*100, '%', sep='')
+#rp = apply(poft,1,function(x) cbind(mean(x),min(x), max(x)) )
+
+#rownames(rp) = c('mean','min','max')
+#colnames(rp) = paste('Year',1:5)
+
+rrPath = back %>% group_by(Status, Pathology) %>% summarise( 'mean.RR'=mean(OR),'ci.min'=mean(OR)-(1.96*sd(OR)), 'ci.max'=mean(OR)+(1.96*sd(OR)) )
+
+rrp = ggplot(rrPath, aes(Status, mean.RR)) + #scale_color_gradientn(colors=myPal, limits=c(min(rrPath[,3:5]), max(rrPath[,3:5])),  name='') +
+  facet_wrap(~Pathology, ncol=1) + 
+  geom_point() +  geom_errorbar(aes(ymin=ci.min, ymax=ci.max), width=0.1) + 
+  geom_text(aes(y=ci.min, label=round(ci.min,2)), vjust=2) + geom_text(aes(y=ci.max, label=round(ci.max,2)), vjust=2) +
+  coord_flip() + plot.theme + labs(y='Mean RR', x='')
+rrp
+
+#ggsave('plots/relrisk_path.png', plot=rrp, width=4, height=6,dpi=300)
+
+```
+
+```{r 5yrrisk, echo=F, fig.height=8, fig.width=12}
+cases = table(sum.patient.data$Status)['P']
+
+mn = round((table(sum.patient.data$Status)['P']/(0.01*100))*100)
+m = round((table(sum.patient.data$Status)['P']/(0.0225*100))*100)
+mx = round((table(sum.patient.data$Status)['P']/(0.035*100))*100)
+
+offsetMin = log(cases/mn)
+offsetMean = log(cases/m)
+offsetMax = log(cases/mx)
+
+# Offset for the linear predictor would then be -log(pop/44)
+# Effectively this is my relative risk for the population
+
+# link function provides odds
+#linPred = predict(fit, newx=dysplasia.df, s=lambda, type='link')
+#quantile(linPred)
+
+#p0 = predict(fit, newx=df, s=lambda, type='response')
+#p0 = 1/(1+exp(-linPred)) # logit to get probabilities, same as type='response'
+#p1 = 1/(1+exp(-linPred+abs(offset))) # calculates probs from risk (offset)
+
+cols = c('Hospital.Research.ID','Patient','Status','PID', 'Samplename','Endoscopy.Year','Pathology','Prediction','OR')
+dspred = dspred %>% dplyr::mutate( 'Adj. RR'=RR+offsetMean )
+
+ggplot(dspred, aes(Type, `Adj. RR`)) + geom_jitter() + geom_boxplot(outlier.color = NA)
+
+head(dspred)
+m = melt(dspred, measure.vars=c('RR','Adj. RR'))
+colors = myPal[seq(2, length(myPal), round(length(myPal)/4))]
+pr = ggplot(m, aes(variable, value, fill=Type)) + facet_grid(~Type, scales='fixed') + 
+  geom_jitter(width=0.2, aes(color=value)) + 
+  scale_color_gradientn(colors = myPal,  name='') + 
+  geom_boxplot(outlier.colour = NA) + 
+  scale_fill_manual(values = colors,  name='') + 
+  labs(x='', y='Relaive Risk', title='Comparing Relative Risk') + 
+  theme(legend.position = 'none', axis.text.x = element_text(angle=45,hjust=1)) + plot.theme 
+pr
+ggsave('plots/downsampled_adjustedRR_supplement.png', plot=pr, width=7, height=5,dpi=300, scale=1)
+
+#loopreds = rbind(loopreds[cols], do.call(rbind, vpd)[cols])
+#loopreds = do.call(rbind, pg.samp)[cols]
+loopreds = back
+#loopreds$P0 = 1/(1+exp(-loopreds$OR))
+
+#loopreds$Mean = 1/(1+exp(-loopreds$OR+abs(offsetMean))) # Adjust OR for RR
+#loopreds$Min = 1/(1+exp(-loopreds$OR+abs(offsetMin))) # Adjust OR for RR
+loopreds$Max = 1/(1+exp(-loopreds$OR+abs(offsetMax))) # Adjust OR for RR
+
+ql = cbind.data.frame('y'=quantile(loopreds$OR)[c(2,4)], 'labels'=paste(c(25,75), '%', sep=''))
+
+# From my nested-LOO predictions
+ggplot(melt(loopreds[c('Min','Mean','Max')], measure.vars='Mean'), aes(value, fill=variable)) + 
+  geom_density(alpha=0.5) + plot.theme
+
+p = ggplot(melt(loopreds[c('Min','Max')]), aes(value)) + facet_wrap(~variable, ncol=2) +
+  geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + 
+    scale_fill_gradientn(colors = myPal,  name='') + 
+    plot.theme + theme(legend.position = 'none') + 
+    labs(y='n Samples', x='Absolute Risk',
+         title='5-year adjusted risk', subtitle='Adjusted for upper & lower bound: 1-3.5% risk')
+p
+
+#ggsave('plots/adj_RR.png', plot=p, width=6, height=6,dpi=300)
+
+```
+
+```{r lifetimeFPR, echo=F, fig.height=6, fig.width=6, eval=F}
+## FPR estimates based on lifetime risks
+# We could consider any NP patient with a minimum of 2 High risk calls?  or High/Moderate Risk?
+# More correctly it should be any patients with concurrent High risk, but that's been pretty stable so two high is good enough for now
+
+bd = back
+table(bd$Status)
+
+risksL = subset(bd, Status == 'NP') %>% group_by(Patient) %>% summarise( 'n.samp'=length(Patient), 'high.risk'=length(which(Risk == 'High')), 'mod.risk'=length(which(Risk == 'Moderate')) )
+
+risksL = subset(risksL, high.risk > 0 | mod.risk > 0)
+
+high = length(subset(risksL, high.risk >= 2 )$Patient)
+#mod = length(subset(risksL, mod.risk >= 2 )$Patient)
+
+# How would I tune it so that the FPR is a lot closer to the real incidence ~1%
+# Current false positive rate is 1.5x higher than the highest incidence ~0.7
+fpr=high*10/43
+x = subset(bd, Status == 'NP' & Patient %in% subset(risksL, high.risk >= 2)$Patient & Risk == 'High')
+(length(unique(subset(x, Prediction > 0.8)$Patient))*10)/44
+
+# How would that affect TPR?
+#pROC::coords(roc, 0.6)
+#pROC::coords(roc, 0.8)
+
+## Per year risk
+## Evaluate over 5 years? How?
+pyr = c()
+for (yr in names(which(table(bd$Endoscopy.Year) > 10))) {
+  risks = subset(bd, Status == 'NP' & Endoscopy.Year == yr) %>% group_by(Patient) %>% summarise( 'n.samp'=length(Patient), 'high.risk'=length(which(Risk == 'High')), 'mod.risk'=length(which(Risk == 'Moderate')) )
+
+  risks = subset(risks, high.risk > 0 | mod.risk > 0)
+  high = length(subset(risks, high.risk >= 2 )$Patient)
+  mod = length(subset(risks, mod.risk >= 2 )$Patient)
+
+  print(paste(yr,high*15, sep=':'))
+  pyr = c(pyr, high*15)
+}
+
+summary(pyr)
+
+pop*.07
+pop*.02
+pop*.05
 

### 2018-03-16 Insilico Splits.Rmd
...

@@
 all.patient.info = read.patient.info(patient.file, demo.file, set='all')$info
-head(all.patient.info)
 
@@
 
-cache.dir = paste(data, 'Analysis', sub('\\+','',paste(tile.w, collapse='_')), sep='/')
+#cache.dir = paste(data, 'Analysis', sub('\\+','',paste(tile.w, collapse='_')), sep='/')
+model.dir = paste(data, 'Analysis/5e6_arms_all_exAHM0320', sep='/')
+split.dir = '/Volumes/research1/gerstung/skillcoyne/Ellie/Analysis/5e6_arms_splits_exAHM0320'
 
-ins_splits = grep('val_', list.dirs(cache.dir), value=T)
-
-performance = data.frame()
-rocs = list()
-
+performance = data.frame(); rocs = list()
 select.alpha = '0.9'
 
-file = paste(cache.dir, 'insilico_splits.Rdata', sep='/')
+file = paste(model.dir, 'insilico_splits.Rdata', sep='/')
 if (file.exists(file)) {
@@
 } else {
-  for (split in ins_splits) {
-    load(grep('all.*Rdata', list.files(split,full.names = T), value=T), verbose=F)
+
+  split.dirs = list.files(split.dir, full.names = T)
+  for (split in 1:length(split.dirs)) {
+    print(split)  
+    load(list.files(split.dirs[split],full.names = T, pattern='all.*Rdata'), verbose=T)
     performance = rbind(performance, performance.at.1se[[select.alpha]])
-    rm(plots,coefs,performance.at.1se,models,cv.patient)
+    rm(plots,coefs,performance.at.1se,models,cvs)
 
-    load(grep('loo.*Rdata', list.files(split,full.names = T), value=T), verbose=F)
+    load(list.files(split.dirs[split],full.names = T, pattern='loo.*Rdata'), verbose=T)
     rm(plots, performance.at.1se,coefs,nzcoefs,fits)
@@
     roc$model = 'in-silico'
-    roc$is = basename(split)
-    rocs[[basename(split)]] = roc
+    roc$is = split
+    rocs[[split]] = roc
 
@@
   insilico.perf=performance
-  save(insilico.rocs, insilico.perf, file=paste(cache.dir, 'insilico_splits.Rdata', sep='/'))
+  save(insilico.rocs, insilico.perf, file=paste(model.dir, 'insilico_splits.Rdata', sep='/'))
 }
 
-load(grep('all.*Rdata',list.files(cache.dir, full.names = T), value=T), verbose=F)
+n.splits = nrow(performance)
+
+load(list.files(model.dir, full.names = T, pattern = 'all.*Rdata'))
+
 performance = rbind(performance, performance.at.1se[[select.alpha]])
-performance$split = c(basename(ins_splits), 'discovery')
-rm(plots,coefs,performance.at.1se,models,cv.patient,dysplasia.df,labels)
+performance$split = c(paste('split',1:n.splits,sep='.'), 'discovery')
+rm(plots,coefs,performance.at.1se,models,cvs,dysplasia.df,labels)
 
-load(grep('loo\\.Rdata', list.files(cache.dir,full.names = T), value=T), verbose=F)
+load(list.files(model.dir,full.names = T, pattern='loo.Rdata'))
 rm(plots, performance.at.1se,coefs,nzcoefs,fits)
@@
     scale_color_brewer(palette='Blues') +
-    labs(title=title, x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
+    labs(title='', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
     theme(legend.position = 'none')

### 2018-03-16 lib/common_plots.R

@@ -146,23 +146,28 @@ multi.roc.plot <- function(rocList, title="ROC", palette=NULL, colors=NULL) {
   return(p)
 }
 
-pred.tiles <- function(df, probs=F, OR=F, cols=c('green3','orange','red3'), ...) {
-  p = ggplot(df, aes(brks, as.factor(Block)))
+pred.tiles <- function(df, probs=F, OR=F, cols=c('green3','orange','red3'), ylim=NULL, ...) {
+  if (is.null(ylim)) ylim = 1:max(as.integer(levels(droplevels(df$Block))))
+
+  df$months.before.final = factor(df$months.before.final, ordered = T)
+  df$months.before.final = factor(df$months.before.final, levels=rev(levels(df$months.before.final)))
+
+  
+  p = ggplot(df, aes(months.before.final, Block)) + scale_y_discrete(limits=ylim)
 
   if (OR) {
     p = p + geom_tile(aes(fill=OR), color='white') + scale_fill_gradientn(colors = myPal, limits=c(-9.3,15), name='RR') 
   } else if (probs) {
     p = p + geom_tile(aes(fill=Prediction), color='white') + scale_fill_gradientn(colors = myPal, limits=c(0,1), name='P(P)') 
   } else {
-    p = p + geom_tile(aes(fill=Risk)) + 
-      scale_fill_manual(values=cols, limits=levels(df$Risk), name='Progression')
+    p = p + geom_tile(aes(fill=Risk), color='white') + scale_fill_manual(values=cols, limits=levels(df$Risk), name='Progression')
   }
-  p + geom_point(aes(shape=Pathology), fill='white', color='white', size=3) + 
+  p + geom_point(aes(shape=Pathology), fill='white', color='white', size=5) + 
     # scale_color_manual(values=c('white','black','white'), guide=F) +
     scale_shape_manual(values=c(1,0,15,24,25), limits=levels(df$Pathology), guide=guide_legend(override.aes=list(fill='white', color='white'))) +
-    facet_wrap(~Patient, ...) +
+    facet_grid(~Patient, ...) +
     labs(y='Oesophageal Location (OGJ..)',x='Endoscopy (Base...End)', title='') +
-    plot.theme + theme(axis.text = element_blank(), legend.key = element_rect(fill='grey39'), panel.background = element_rect(colour = 'black'), panel.grid.major = element_blank()) 
[…]

### 2018-03-23 AllSamplesModel.Rmd
updating shiny app

@@
 globalCN.plot = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples')
-ggsave(filename='plots/globalCN.png', plot=globalCN.plot + labs(title=''), width=16, height=4.6, units='in', dpi=300)
-globalCN.plot
+
+ggsave(filename='plots/globalCN.png', plot=globalCN.plot + theme(strip.text.y=element_blank(), strip.text.x=element_text(size=9), axis.text=element_blank()) + labs(x='', y='',title=''), width=8.3, height=2.4, units='in', dpi=300)
 
@@
 globalBECN.plot = globalBECN.plot + theme(legend.position = 'none') + labs(y='mean adj.seg.', title="Non-Dysplastic Barrett's Samples")
-ggsave(filename='plots/globalBECN.png', plot=globalBECN.plot + labs(title=''), width=16, height=4.6, units='in', dpi=300)
+
+ggsave(filename='plots/globalBECN.png', plot=globalBECN.plot + theme(strip.text.y=element_blank(), strip.text.x=element_text(size=9), axis.text=element_blank()) + labs(x='', y='',title=''), width=8.3, height=2.4, units='in', dpi=300)
 globalBECN.plot
@@
 ```{r model, message=F, warning=F, echo=F, fig.height=16, fig.width=16}
-#folds = 10; splits = 5 
-# file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
-# if (file.exists(file)) {
-#   message(paste("loading", file))
-#   load(file, verbose=T)
-# } else {
-#   stop(paste("No model file", file, "found"))
-# }
+file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
+if (file.exists(file)) {
+  message(paste("loading", file))
+  load(file, verbose=T)
+} else {
+  stop(paste("No model file", file, "found"))
+}
 
@@
 }
+
+chtable = ch[,c('hazards','coef','stability','label')]
+chtable = chtable[,c('label','hazards', 'coef')]
+chtable$coef[which(chtable$coef > 0)] = 'Gain'
+chtable$coef[which(chtable$coef != 'Gain')] = 'Loss'
+colnames(chtable) = c('Location','Hazard', 'Gain/Loss')
+chtable$Hazard = round(chtable$Hazard, 3)
+write.table(chtable, sep='\t', quote=F, row.names = F, file='hazards.txt')
 
@@
 ```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=20}
-min.theme = theme(text=element_text(size=12), panel.background=element_blank(), strip.background=element_rect(fill="grey97"), strip.text.y = element_text(size=6),
-                   strip.text.x = element_text(size=12), axis.text = element_blank(), axis.ticks = element_blank(),
-                   axis.line=element_line(color='black'), panel.grid.major.y =element_line(color='grey95'), panel.grid.major.x = element_blank(),
-                   panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0, 'lines'), panel.spacing.y = unit(0.1, 'lines')   )
-
-mtn.spatial<-function(df, b, limits, title='', pal=NULL) {
-  if (is.null(pal))
-    pal = RColorBrewer::brewer.pal(length(limits), 'Set2')
-
-  df = merge(df, b[,c('Samplename','Endoscopy.Year','months.before.final','Block','ogj')], by.x='variable',by.y='Samplename')
-  df$ogj = factor(df$ogj, ordered = T)
-  df$ogj = factor(df$ogj, levels=rev(levels(df$ogj)), ordered = T)
-  #pal = pal[which(limits %in% levels(df$nl))]
-
-  altcolor = c(RColorBrewer::brewer.pal(9, 'Greys')[2],RColorBrewer::brewer.pal(9, 'Greys')[4])
-  fills = c()
-  for (yr in sort(unique(df$Endoscopy.Year))) {
-    ac = (which(sort(unique(df$Endoscopy.Year)) == yr) %% 2) + 1
-    fills = c(fills, rep(altcolor[ac], length(which(table(subset(df, Endoscopy.Year == yr)$ogj) > 0))))
-  }
-
-  blanklabel <- function(s) {
-    return('')
-  }
-
-  p = ggplot(df, aes(x=chr.length)) + facet_grid(Endoscopy.Year+ogj~chr, scales='free', space='free_x', labeller = labeller(.multi_line = F)) +
-      geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=ogj), show.legend=T) + 
-      scale_fill_manual(values=pal, limits=limits,  name='OGJ Dist.') +
-      labs(title=title, x='') + min.theme
-
-  legend = get.legend(p)
-  p = p + theme(legend.position = 'none')
-
-  g = ggplot_gtable(ggplot_build(p))
-  strips = grep('strip-r', g$layout$name)
-
-  for (n in 1:length(strips)) {
-    i = strips[n]
-    j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
-    g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[n]
-  }
-
-  #grid.arrange(g)
-  return(list('grob'=g, 'legend'=legend))
-  #grid::grid.draw(g)  
-  #return(g)
-}
-
 back$ogj = back$nl
@@
 
-p2 = mtn.spatial(subset(globalCNdf, Patient == 20 & seg.type != 'arms'), back, ogjLimits,'Patient 20 (P)' )
-p6 = mtn.spatial(subset(globalCNdf, Patient == 51 & seg.type != 'arms'), back,ogjLimits,'Patient 51 (NP)' )
+p2 = mtn.spatial(subset(globalCNdf, Patient == 20 & seg.type != 'arms'), back, ogjLimits,'' )
+grid.arrange(p2$grob)
+
+p6 = mtn.spatial(subset(globalCNdf, Patient == 51 & seg.type != 'arms'), back,ogjLimits,'' )
+
+grid.arrange(p2$legend)
+
+ggsave(filename='plots/spatial/patient_20.png', plot=grid.arrange(p2$grob), width=7.2, height=3, units='in', dpi=300)
+ggsave(filename='plots/spatial/patient_51.png', plot=grid.arrange(p6$grob), width=7.2, height=3, units='in', dpi=300)
 
@@
 
-```{r, echo=F, fig.width=2, fig.height=4}
-
-grid.arrange(p2$legend)
-
-ggsave(filename='plots/spatial/patient_20.png', plot=grid.arrange(p2$grob), width=14, height=6, units='in', dpi=300)
-ggsave(filename='plots/spatial/patient_51.png', plot=grid.arrange(p6$grob), width=14, height=6, units='in', dpi=300)
-
-```
-
 At each quintile are the percentage of progressor samples that were predicted within that range, and the hypergeometric p-value (Fishers) for the significance within that quintile.
@@
 
+#riskPath$Pathology = as.character(riskPath$Pathology)
+#riskPath$Pathology[riskPath$Pathology == 'BE'] = 'NDBE'
+#riskPath$Pathology = factor(riskPath$Pathology, levels=c('NDBE','ID','LGD','HGD','IMC'), ordered=T)
 rp = ggplot(riskPath, aes(Pathology, ratio*100)) + 
@@
   scale_fill_manual(values=riskCols, limits=levels(riskPath$Risk)) + 
-  scale_color_manual(values=c('white','black','white'), limits=levels(pred.confidence$Risk) ) +
+  scale_color_manual(values=c('black','black','black'), limits=levels(pred.confidence$Risk) ) +
+  scale_x_discrete(labels=c('NDBE','ID','LGD','HGD','IMC')) +
   geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=102), position=position_stack()) +
@@
   scale_fill_manual(values=col, name='p53 IHC', labels=c('Normal','Aberrant')) + 
+  scale_x_discrete(labels=c('NDBE','ID','LGD','HGD','IMC')) +
   geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=103)) +
@@
   geom_bar(stat='identity', fill=riskCols[3], show.legend = F) + 
-  geom_errorbar(color='grey21', width=0.2) +
-  geom_text(aes(y=3, label=paste(round(pred.r*100), '%', sep='')), color='white') +
-  geom_text(aes(y=pred.r*100-8, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
+  geom_errorbar(color='grey21', width=0.2, alpha=0.5) +
+  geom_text(aes(y=pred.r*100+2, label=paste(round(pred.r*100), '%', sep='')), color='black') +
+  geom_text(aes(y=10, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
   labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') +
@@
 
-```{r heatmaps, echo=F, warning=F, message=F, fig.width=5, fig.height=5 }
+```{r heatmaps, echo=F, warning=F, message=F}
 back$nl = factor(back$nl, ordered=T)
@@
 back$Patient = factor(as.numeric(back$Patient), ordered = T)
-
-table(rev(arrange(subset(back, Patient == 19), brks)$brks))
-
-#pred.tiles(subset(back, Patient %in% c(48)), ncol=1, cols=myPal, probs=T) + labs(title='') + theme(legend.position = 'none')
-#pred.tiles(subset(back, Patient %in% c(33)), ncol=1, cols=myPal, probs=T) + labs(title='')+ theme(legend.position = 'none')
-#pred.tiles(subset(back, Patient %in% c(89)), ncol=1, cols=myPal, probs=T) + labs(title='')+ theme(legend.position = 'none')
-```
-
-```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=15 }
-#back = back %>% group_by(Patient) %>% dplyr::mutate(endo.num = as.integer(droplevels(brks)) )
-#back$endo.num = factor(back$endo.num, ordered = T)
-
-#back$endo.num = factor(as.integer(back$brks))
-#back = back %>% group_by(Patient) %>% dplyr::mutate(endo.num = ifelse(endo.num == max(endo.num), 'End', endo.num)   )
-
-ga = grid.arrange(
-  pred.tiles(subset(back, Patient %in% c(13,20,34,56,60,48) & Status == 'P'), ncol=1, cols=myPal, probs=T) + labs(title='Progressors') + theme(legend.position = 'none'),
-  pred.tiles(subset(back, Patient %in% c(15,16,21,3,45,19) & Status == 'NP'), ncol=1, cols=myPal, probs=T) + labs(title='Non-Progressors') + theme(legend.position = 'none'),
-  ncol=2)
-ga
-
-#ggsave('plots/ex_pt_heatmap.png', plot=ga, width=6, height=15, units='in', dpi=300)
 
@@
 
-plots = list(); widths = c()
-for (i in subset(sum.patient.data, Status == 'P')$Patient) {
+back$months.before.final = round(back$months.before.final)
+
+for (i in unique(subset(back, Status == 'P')$Patient)) {
   b = subset(back, Patient == i)
+  if (i == 89) b[which(b$Pathology == 'HGD'),'months.before.final'] = 0
+
   lim = c(1:7)
@@
 
-  p = pred.tiles(b, cols=riskCols, ylim=lim,  probs=F, scales='free', space='free') + labs(title='', x='', y='') + theme(legend.position = 'none', strip.text = element_blank(), plot.margin = unit(c(0,0.2,0,0), 'cm'))
+  p = pred.tiles(b, cols=riskCols, ylim=lim,  probs=F, path=F, scales='free', space='free') + labs(title='', x='', y='') + theme(legend.position = 'none', strip.text = element_blank(), plot.margin = unit(c(0,0.2,0,0), 'cm'))
+  ggsave(paste('plots/heatmaps/P_nopath/pt_',i,'.png', sep=''), plot=p, height=get.h.scale(b), width=get.w.scale(b), unit='in', dpi=300 )
+
+    p = pred.tiles(b, cols=riskCols, ylim=lim,  probs=F, path=6, scales='free', space='free') + labs(title=i,x='months', y='location') + theme(legend.position = 'none', strip.text = element_blank(), axis.text = element_text(size=9), plot.margin = unit(c(0,0.2,0,0), 'cm'))
   ggsave(paste('plots/heatmaps/P/pt_',i,'.png', sep=''), plot=p, height=get.h.scale(b), width=get.w.scale(b), unit='in', dpi=300 )
 }
+
 b = subset(back, Patient == 32)
@@
 p = pred.tiles(b, cols=riskCols, ylim=lim,  probs=F, scales='free', space='free') + labs(title='', x='', y='') + theme(legend.position = 'none', strip.text = element_blank(), plot.margin = unit(c(0,0.2,0,0), 'cm'))
-  ggsave(paste('plots/heatmaps/P/pt_32.png', sep=''), plot=p, height=4, width=get.w.scale(b), unit='in', dpi=300 )
+  ggsave(paste('plots/heatmaps/P_nopath/pt_32.png', sep=''), plot=p, height=4, width=get.w.scale(b), unit='in', dpi=300 )
+
+p = pred.tiles(b, cols=riskCols, ylim=lim,  probs=F, path=6, scales='free', space='free') + labs(title='32', x='months',y='location') + theme(legend.position = 'none', strip.text = element_blank(), axis.text = element_text(size=9), plot.margin = unit(c(0,0.2,0,0), 'cm'))
+  ggsave('plots/heatmaps/P/pt_32.png', plot=p, height=4, width=get.w.scale(b), unit='in', dpi=300 )
 
@@
 
-for (i in subset(sum.patient.data, Status == 'NP')$Patient) {
+for (i in unique(subset(back, Status == 'NP')$Patient)) {
   b = subset(back, Patient == i)
@@
   p = pred.tiles(b, cols=riskCols, ylim=lim,  probs=F, scales='free', space='free') + labs(title='', x='', y='') + theme(legend.position = 'none', strip.text = element_blank(), plot.margin = unit(c(0,0.2,0,0), 'cm'))
-  ggsave(paste('plots/heatmaps/NP/pt_',i,'.png', sep=''), plot=p, height=get.h.scale(b,9), width=get.w.scale(b), unit='in', dpi=300 )
-}
-
-all = pred.tiles(back, cols=riskCols,  probs=F, scales='free', space='free') + labs(title='', x='', y='') + theme(legend.position = 'bottom')
+  ggsave(paste('plots/heatmaps/NP_nopath/pt_',i,'.png', sep=''), plot=p, height=get.h.scale(b,9), width=get.w.scale(b), unit='in', dpi=300 )
+
+  p = pred.tiles(b, cols=riskCols, ylim=lim,  probs=F, path=6, scales='free', space='free') + labs(title=i,x='months',y='location') + theme(legend.position = 'none', strip.text = element_blank(), axis.text = element_text(size=9), plot.margin = unit(c(0,0.2,0,0), 'cm'))
+  ggsave(paste('plots/heatmaps/NP/pt_',i,'.png', sep=''), plot=p, height=get.h.scale(b), width=get.w.scale(b), unit='in', dpi=300 )
+}
+
+all = pred.tiles(back, cols=riskCols, path=4, probs=F, scales='free', space='free') + labs(title='', x='', y='') + theme(legend.position = 'bottom')
 riskLegend = get.legend(all)
@@
 load(file, verbose=T)
-rm(dysplasia.df,coefs,plots,labels)
+rm(coefs,plots,labels)
 
@@
 
-m = subset(m, Type != 'Normal blood')
+m = subset(dspred, Type != 'Normal blood')
 m$Type = droplevels(m$Type)
@@
 p = p + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_34.png', plot=p, width=6, height=3.9,dpi=300)
+ggsave('plots/vignette/vig_34.png', plot=p + coord_flip() + labs(y='', x=''),width=4.6, height=4.2,dpi=300)
+#ggsave('plots/vignette/vig_34.png', plot=p, width=6, height=3.9,dpi=300)
 
@@
 p = vig.plot(subset(back, Patient == 20), subset(patient.info, Patient == 20),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_20.png', plot=p, width=6, height=3.9,dpi=300)
+ggsave('plots/vignette/vig_20.png', plot=p + coord_flip() + labs(y='', x=''), width=4.6, height=4.2,dpi=300)
+#ggsave('plots/vignette/vig_20.png', plot=p, width=6, height=3.9,dpi=300)
 
@@
 p = vig.plot(subset(back, Patient == 3), subset(patient.info, Patient == 3),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_3.png', plot=p, width=6, height=3.9,dpi=300)
+ggsave('plots/vignette/vig_3.png', plot=p + coord_flip() + labs(y='', x=''),width=4.6, height=4.2,dpi=300)
+#ggsave('plots/vignette/vig_3.png', plot=p, width=6, height=3.9,dpi=300)
 
@@
 p = vig.plot(subset(back, Patient == 68), subset(patient.info, Patient == 68),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_68.png', plot=p, width=6, height=3.9,dpi=300)
+ggsave('plots/vignette/vig_68.png', plot=p + coord_flip() + labs(y='', x=''), ,width=4.6, height=4.2,dpi=300)
+#ggsave('plots/vignette/vig_68.png', plot=p, width=6, height=3.9,dpi=300)
 
@@
 load(file, verbose=T)
-rm(plots, coefs, cv.patient)
+rm(plots, coefs, cvs)
 
@@
 dspred = dspred %>% dplyr::mutate( 'Adj. RR'=RR+offsetMean )
+
+dspred = dspred %>% dplyr::mutate( 'Adj. Prob'=1/(1+exp(-RR+abs(offsetMean)) ))
+
+ggplot(dspred, aes(Type, `Adj. Prob`)) + geom_jitter()
 
@@
 
+m = melt(dspred, measure.vars=c('Prob','Adj. Prob'))
+colors = myPal[seq(2, length(myPal), round(length(myPal)/4))]
+pr = ggplot(m, aes(variable, value, fill=Type)) + facet_grid(~Type, scales='fixed') + 
+  geom_jitter(width=0.2, aes(color=value)) + 
+  scale_color_gradientn(colors = myPal,  name='') + 
+  geom_boxplot(outlier.colour = NA) + 
+  scale_fill_manual(values = colors,  name='') + 
+  labs(x='', y='Relaive Risk', title='Comparing Relative Risk') + 
+  theme(legend.position = 'none', axis.text.x = element_text(angle=45,hjust=1)) + plot.theme 
+pr
+
 #loopreds = rbind(loopreds[cols], do.call(rbind, vpd)[cols])

### 2018-03-23 PredictingProgression/lib/risk_helper.R

@@ -1,0 +1,102 @@
+predict.progression <- function(rs,beObj, fitV, lambda, hdr=T, model.norm=T) {
+  if (is.null(beObj) | is.null(fitV) | is.null(lambda))
+    stop("No computation possible without model object or glmnet fit")
+
+  if (ncol(rs) < 4)
+    stop("File needs to contain 4 columns: chr, start, end, sample_value")
+  #head(rs)
+  
+  cnCol = grep('Total_CN', colnames(rs))
+  if (length(cnCol) >= 1)
+    rs = rs[,c(1:cnCol)]
+  
+  segs <- tile.segmented.data(rs, size=5e6 )
+  segM = as.matrix(segs[,-c(1:3)])
+  rownames(segM) = paste(segs$chr, ':', segs$start, '-', segs$end, sep='')
+  segM = t(segM)
+  
+  if (model.norm) {
+    for (i in 1:ncol(segM)) 
+      segM[,i] = unit.var(segM[,i], beObj@z.mean[i], beObj@z.sd[i])
+  } else {
+    segM = unit.var(segM)
+  }
+  
+  arms <- tile.segmented.data(rs, size='arms')
+  armsM = as.matrix(arms[,-c(1:3)])
+  rownames(armsM) = paste(arms$chr, ':', arms$start, '-', arms$end, sep='')
+  armsM = t(armsM)
+  
+  if (model.norm) {
+    for (i in 1:ncol(armsM)) 
+      armsM[,i] = unit.var(armsM[,i], beObj@z.arms.mean[i], beObj@z.arms.sd[i])
+  } else {
+    armsM = unit.var(armsM)
+  }
+
+  cx.score = score.cx(segM,1)
+  
+  mergedDf = subtract.arms(segM, armsM)
+  mergedDf = cbind(mergedDf, 'cx' = unit.var(cx.score, beObj@mn.cx, beObj@sd.cx))
+  
+  # Predict function giving me difficulty when I have only a single sample, this ensures the dimensions are the same
+  sparsed_test_data <- Matrix(data=0, nrow=nrow(mergedDf),  ncol=ncol(mergedDf),
+                              dimnames=list(rownames(mergedDf),colnames(mergedDf)), sparse=T)
+  for(i in colnames(mergedDf)) sparsed_test_data[,i] = mergedDf[,i]
+  
[…]

### 2018-03-27 AllSamplesModel.Rmd
updating shiny app

@@
 roc$model = 'CN Model'
+
+pROC::ci.auc(roc)
+
 roc$is = 'all'
@@
 
+pROC::ci.auc(p53roc)
+
 p53roc$model = 'p53 IHC'
@@
 pathroc$model = 'HGD & LGD'
+
+pROC::ci.auc(pathroc)
+
 ggsave('plots/auc/dysplasia_auc.png', plot=roc.plot(pathroc, title='Dysplasia (HGD & LGD)'), width=6, height=6, units="in", limitsize=F, dpi = 300)
@@
 hgdpathroc$model = 'HGD Only'
+
+pROC::ci.auc(hgdpathroc)
+
 ggsave('plots/auc/hgd_auc.png', plot=roc.plot(hgdpathroc, title='HGD Only'), width=6, height=6, units="in", limitsize=F, dpi = 300)
@@
 }
-
-get.auc.at<-function(roc, foc='spec', p=c(1,0.99)) {
-  # To find the auc at a given spec or sens
-  pROC::auc(roc, partial.auc=p, partial.auc.focus=foc, partial.auc.correct=T)
-}
-
 roc$model = "CN Model"
@@
 m$model = factor(m$model)
+
+subset(m, model %in% c(rocEndoA$model, rocEndoM$model) & Measure == 'AUC')
 
@@
 plot(table(p11=pi$p11Loss, ihc=pi$p53.Status))
+
+fisher.test(table(pi$p11Loss, pi$p53.Status))
 

### 2018-03-27 PredictingProgression/server.R

@@ -1,10 +1,3 @@
-#
-# This is the server logic of a Shiny web application. You can run the 
-# application by clicking 'Run App' above.
-#
-# Find out more about building applications with Shiny here:
-# 
-#    http://shiny.rstudio.com/
 
 library(shiny)
 library(ggplot2)
@@ -86,61 +79,87 @@ loadModelData <- function() {
 shinyServer(function(input, output, session) {
   modelData = loadModelData()
   
-  dataInput <- reactive({
+  readDataFile <- reactive({
     req(input$file)
     dataFile <- input$file$datapath
-
-    norm = input$norm
-    hdr = input$header
-    
-    df <- read.table(dataFile, header=hdr, sep='\t')
+    print(dataFile)
+    df <- read.table(dataFile, header=T, sep='\t')
     samplenames = grep('chr|arm|start|end|probes', colnames(df), ignore.case=T, invert=T, value=T)
+    return(list('data'=df, 'samples'=samplenames))
+  })
+  
+  perSampleMeta <- reactive({
+    req(input$pathfile)
+    
+    fi = readDataFile()
+    samplenames = fi$samples
     
     if (!is.null(input$pathfile)) {
       pf = read.table(input$pathfile$datapath, sep='\t', header=T, stringsAsFactors=F)
       if (length(samplenames) < nrow(pf)) {
         warning("Samples do not match in data and path files, ignoring path.")
         pf = NULL
-      }
-      pp = predict.progression(df, modelData$be, modelData$fit, modelData$lambda.opt, hdr, norm)
-
-      samplenames = rownames(pp$rel.risk)
-
-      pR = cbind.data.frame(samplenames, sapply(pp$rel.risk, adjustRisk, offset=modelData$be@meanOffset, type='prob'), 
-                            sapply(pp$rel.risk, adjustRisk, offset=modelData$be@meanOffset))
-      colnames(pR) = c('Sample', 'Probability', 'Relative Risk')
-      rownames(pR) = 1:nrow(pR)
-      
-      pR$Risk = sapply(pR$Probability, risk)
-      
-      recommendations = rx(pR)
-
-      if (!is.null(pf)) {
-        pR = merge(pf, pR, by='Sample')
-        recommendations = rx(pR)
+      } else {
+        print(input$sampleMeta)
       }
     }
+    return(pf)
+  })
+  
+  generatePred <- reactive({
+    fi = readDataFile()
+    df = fi$data
[…]

### 2018-03-27 PredictingProgression/ui.R

@@ -17,6 +17,13 @@ shinyUI(fluidPage(
   titlePanel("Barrett's Progression Prediction"),
   sidebarLayout(
     sidebarPanel(
+      tags$h3('Patient Information'),
+      numericInput("ageY", "Age at diagnosis:", min=1, max=100, value=60,step=1),
+      sliderInput('beLength', 'Length of BE segment', min=1,max=30,round=1,value=4),
+      radioButtons("sex", label = "Sex", choices = list("Male" = 'M', "Female" = 'F'), selected = 'M'),
+      
+      
+      tags$h3('Patient Data'),
       fileInput("pathfile", "Optional: upload a file with per-sample pathology and/or p53 IHC (see example)",
                 accept = c(
                   "text/tsv",
@@ -31,32 +38,33 @@ shinyUI(fluidPage(
                   "text/csv",
                   "text/tab-separated-values,comma-separated-values,text/plain",
                   ".csv", ".tsv", ".txt")
-      ),
-      
-      
-      tags$hr(),
-      checkboxInput("header", "Header", TRUE),
+      )
       # Input: Select separator ----
-      checkboxInput("norm", "Model provided normalization (default: yes)", TRUE),
-      # Input: Select separator ----
-      radioButtons("patients", "Samples are from -",
-                   choices = c('Single patient' = "same",
-                               'Multiple patients' = "mult"),
-                   selected = "same")
+      #checkboxInput("header", "Header", TRUE),
+      #checkboxInput("norm", "Model provided normalization (default: yes)", TRUE),
+      
+      #actionButton("action", label = "Submit")
 
     ),
-    
+  
     mainPanel(
-      tableOutput("example"),
-      tableOutput("examplep53"),
-      #tableOutput("contents"),
       helpText(tags$b("Recommendations are evaluated with the following criteria:"), tags$br(),
-      tags$ol(tags$li("Immediate RFA: HGD diagnosis or more than one consecutive high risk predictions."),
-          tags$li("Recheck 6-12 months: One high risk prediction or an aberrant p53 IHC"),
-          tags$li("Recheck endoscopy 12-24 months: One or more moderate risk predictions"),
-          tags$li("Regular surveillance 3-5 years: Two or more consecutive low risk predictions"))),
-      plotOutput("plot"),
-      plotOutput("riskTable")
+               tags$ol(tags$li("Immediate RFA: HGD diagnosis or more than one consecutive high risk predictions."),
[…]

### 2018-04-16 Patient Report.Rmd
updating Rstudio app

@@
 
-var.seg.lengths = lapply(patient.data, function(df) {
-  cbind('chr'=df$variable.regions$chrom, 'start'=df$variable.regions$start.pos, 'end'=df$variable.regions$end.pos, 'seg.len'=df$variable.regions$end.pos - df$variable.regions$start.pos)
+var.seg.lengths = lapply(patient.data, function(pt) {
+  cbind('chr'=pt$variable.regions$chrom, 'start'=pt$variable.regions$start.pos, 'end'=pt$variable.regions$end.pos, 'seg.len'=pt$variable.regions$end.pos - pt$variable.regions$start.pos)
 })
@@
 
-ggplot(df, aes(status, seg.len)) + geom_jitter(aes(color=status), alpha=0.5, show.legend = F, width=.2) +
-  geom_boxplot(aes(group=status, fill=status), show.legend=F, outlier.colour = NA) + 
-  geom_label(data=subset(df, status == 'Progressor'), aes(status, max(seg.len)-2*sd(seg.len), label=paste('median\n',scientific(median(seg.len),3),sep='') )) +
-  geom_label(data=subset(df, status == 'Non-Progressor'), aes(status, max(seg.len), label=paste('median\n',scientific(median(seg.len),3),sep=''))) +
-  labs(y='Variable segment length', x='', title='Variable regions', caption=paste('Wilcox p-value', signif(lenWT$p.value, digits=3) ))
-
-ggplot(df, aes(status, seg.len)) + 
-  geom_jitter(aes(color=status), show.legend=F) + geom_violin(aes(group=status, fill=status), show.legend=F) +
-  geom_label(data=subset(df, status == 'Progressor'), aes(status, max(seg.len)-2*sd(seg.len),label=paste('median\n',scientific( median(seg.len) , 3),sep='') ))  +
-    geom_label(data=subset(df, status == 'Non-Progressor'), aes(status, max(seg.len), label=paste('median\n',scientific( median(seg.len) , 3),sep=''))) +
-    labs(y='Variable segment length', x='', title='Variable regions', caption=paste('Wilcox p-value', signif(lenWT$p.value, digits=3) ))
+m = (melt(df, id.vars = 'status', measure.vars = 'seg.len'))
+m2 = m %>% group_by(status) %>% summarise( median=median(value), n=length(status), value=max(value) )
+ggplot(m, aes(status, value, group=status, fill=status)) + geom_jitter(aes(color=status), width=.2) + geom_violin(alpha=0.5) +
+  geom_label(data=m2, aes(status, value, label=paste("median=",scientific(median,3),'\nn=',n, sep='')), fill='white') + 
+  theme(legend.position = 'none') + labs(y='Variable segment length', x='', title=paste('Length of variable regions (all samples n=',length(patient.info$Samplename),')',sep=''), subtitle=paste('Wilcox p-value', signif(lenWT$p.value, digits=3) ))
 
@@
 df = rbind(cbind.data.frame('n'=sapply(var.seg.lengths[prog.pt], nrow), 'status'='Progressor'), cbind.data.frame('n'=sapply(var.seg.lengths[np.pt], nrow), 'status'='Non-Progressor'))
-nWT = wilcox.test(subset(df, status == 'Progressor')$n, subset(df, status == 'Non-Progressor')$n)
+nWT = t.test(subset(df, status == 'Progressor')$n, subset(df, status == 'Non-Progressor')$n)
 
 ggplot(df, aes(status, n, group=status)) + geom_jitter(aes(color=status), width=0.2) +
-  geom_boxplot(aes(fill=status), position = 'dodge', show.legend = F, outlier.colour = NA) +
-  geom_label(data=subset(df, status == 'Progressor'), 
-             aes(status, max(n)-2*sd(n), label=paste('median\n',sprintf("%.1f",median(n)),sep=''), fill=status ), color='white' ) +
-  geom_label(data=subset(df, status == 'Non-Progressor'), 
-             aes(status, max(n)+2*sd(n), label=paste('median\n',sprintf("%.1f",median(n)),sep=''), fill=status), color='white') +
-  labs(y='Number of segments', title="Total number of segments", caption=paste('Wilcox p-value', signif(nWT$p.value, digits=3)) ) +
+  geom_violin(aes(fill=status, group=status), alpha=0.5) +
+  geom_label(data=(df %>% group_by(status) %>% summarise(y=max(n), med=median(n))), 
+             aes(status, y, label=paste('median\n',sprintf("%.1f", med, sep='')))) +
+  labs(y='Number of segments', title="Number of segments per patient", subtitle=paste('t-test p-value', signif(nWT$p.value, digits=3)) ) +
   theme(legend.position = 'none')
 
-ggplot(df, aes(status, n, group=status)) + geom_jitter(aes(color=status), width=0.2) +
-  geom_violin(aes(fill=status, group=status)) +
-  geom_label(data=subset(df, status == 'Progressor'), 
-             aes(status, max(n)-2*sd(n), label=paste('median\n',sprintf("%.1f",median(n)),sep=''), fill=status ), color='white' ) +
-  geom_label(data=subset(df, status == 'Non-Progressor'), 
-             aes(status, max(n)+2*sd(n), label=paste('median\n',sprintf("%.1f",median(n)),sep=''), fill=status), color='white') +
-  labs(y='Number of segments', title="Total number of segments", caption=paste('Wilcox p-value', signif(nWT$p.value, digits=3)) ) +
-  theme(legend.position = 'none')
+nohgd.var.seg.lengths = lapply(patient.data, function(pt) {
+  samples = subset(pt$info, Pathology %in% c('BE','ID','LGD'))$Samplename
+  cols = c('chrom','start.pos','end.pos', grep(paste(samples, collapse='|'), colnames(pt$variable.regions), value=T))
+  vr = pt$variable.regions[,cols]
+
+  cbind('chr'=vr$chrom, 'start'=vr$start.pos, 'end'=vr$end.pos, 'seg.len'=vr$end.pos - vr$start.pos)
+})
 

### 2018-04-16 PredictingProgression/lib/risk_helper.R

@@ -1,4 +1,6 @@
 predict.progression <- function(rs,beObj, fitV, lambda, hdr=T, model.norm=T) {
+  
+  
   if (is.null(beObj) | is.null(fitV) | is.null(lambda))
     stop("No computation possible without model object or glmnet fit")
 
@@ -16,6 +18,7 @@ predict.progression <- function(rs,beObj, fitV, lambda, hdr=T, model.norm=T) {
   segM = t(segM)
   
   if (model.norm) {
+    message('Using model generated mean/sd to unit norm')
     for (i in 1:ncol(segM)) 
       segM[,i] = unit.var(segM[,i], beObj@z.mean[i], beObj@z.sd[i])
   } else {
@@ -82,9 +85,9 @@ rx<-function(pR) {
       rule = 1
     } else if ( risks['High'] == 1 || (!is.null(p53) && p53['1'] > 0) ) {
       rule = 2
-    } else if ( risks['Moderate'] > 0 ) {
+    } else if ( risks['Moderate'] > 0 || (risks['Low'] ==1 && nrow(pR) == 1)) {
       rule = 3
-    } else if ( risks['Low'] == 2) {
+    } else if ( risks['Low'] == 2 ) {
       rule = 4
     }
     rules[i,] = cbind( pR$Sample[i], pR$Sample[(i+1)], as.integer(rule) )

### 2018-04-16 processing_pipeline/predict_downsampled.R

@@ -9,26 +9,53 @@ source('lib/load_patient_metadata.R')
 
 chrlen = get.chr.lengths()
 
+adjustRisk <- function(RR, offset, type='risk') {
+  if (type == 'prob') {
+    x = 1/(1+exp(-RR+abs(offset)))
+  } else {
+    x = RR+offset
+  }
+  return(x)
+}
+
+
 dir = '~/Data/Ellie/Analysis/downsampled5G'
+dir = '~/Data/Ellie/Analysis/VAL_Cohort/'
 files = list.files(dir, full.names=T)
 
-
-file = '~/Data/Ellie/Analysis/5e6_arms_disc/model_data.Rdata'
+file = '~/Data/Ellie/Analysis/5e6_arms_all_exAHM0320/model_data.Rdata'
 if (!file.exists(file))
   stop(paste("Missing data file", file))
 load(file, verbose=T)
-file = '~/Data/Ellie/Analysis/5e6_arms_disc/all.pt.alpha.Rdata'
+file = '~/Data/Ellie/Analysis/5e6_arms_all_exAHM0320/all.pt.alpha.Rdata'
 if (!file.exists(file))
   stop(paste("Missing data file", file))
 load(file, verbose=T)
 rm(dysplasia.df,coefs,plots,labels)
+
+
+file = '~/Data/Ellie/Analysis/5e6_arms_all_exAHM0320/loo.Rdata'
+if (!file.exists(file))
+  stop(paste("Missing data file", file))
+load(file, verbose=T)
+rm(plots, coefs, nzcoefs, fits)
+
+pg.samp = do.call(rbind, pg.samp)
+pg.samp$Hospital.Research.ID = NULL
+
+status = unique(pg.samp[,c('Patient','Status')])$Status
+
+cases = table(status)['P']
+m = round((table(status)['P']/(0.0225*100))*100)
+offsetMean = log(cases/m)
+
 
 select.alpha = '0.9'
 fitV = models[[select.alpha]]
 lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']
 
 sampleNames = sub('_raw.txt', '', basename(files))
-dspred = data.frame(matrix(ncol=2, nrow=length(samples), dimnames=list(samples,c('Prob','RR'))))
+dspred = data.frame(matrix(ncol=4, nrow=length(sampleNames), dimnames=list(sampleNames,c('Prob','RR', 'AdjProb','AdjRR'))))
 
[…]

### 2018-04-16 processing_pipeline/process_downsampled.R

@@ -4,8 +4,8 @@ args = commandArgs(trailingOnly=TRUE)
 if (length(args) < 3)
   stop("Missing required params: <data dir> <out dir> <blacklist file> <overwrite: default F>")
 
-suppressPackageStartupMessages(source('lib/fastPCF.R'))
-suppressPackageStartupMessages(source('lib/data_func.R'))
+suppressPackageStartupMessages(source('../lib/fastPCF.R'))
+suppressPackageStartupMessages(source('../lib/data_func.R'))
 
 
 chrlen = get.chr.lengths()
@@ -14,7 +14,7 @@ dir = args[1]
 outdir = args[2]
 exclude.file = args[3]
 overwrite = F
-if (length(args) == 4) overwrite = args[4]
+if (length(args) == 4) overwrite =  as.logical(args[4])
 
 # dir = '~/Data/Ellie/QDNAseq/DownsampledBE/20180206_KillcoyneS_RF_BarrettsCN/qdnaseq/'
 # outdir = '~/Data/Ellie/Analysis/downsampled5G_BEAdjacent_normal/'

### 2018-05-02 AllSamplesModel.Rmd
Moved shiny app to own repo

@@
 
-hazards = apply( exp(t(dysplasia.df[, rownames(coefs[[select.alpha]])]) *  coefs[[select.alpha]][,1]), 1, function(x) {
+hazards = apply( exp(t(dysplasia.df[,rownames(coefs[[select.alpha]])]) *  coefs[[select.alpha]][,1]), 1, function(x) {
   sd(x)/mean(x)
+  #mad(x) # The CV(RR) is pretty sensitive to extremes I think. But I'm not sure if MAD(RR would be better)
 })
+
 ch = as.data.frame(hazards)
@@
 
-## Top Features
-```{r, echo=T}
+# Top Features
+```{r, echo=F, message=F, warning=F}
 m = data.frame(matrix(nrow=nrow(chh), ncol=0))
@@
 
+write.table(chh, sep='\t', quote=F, row.names=F, file='features.txt')
+
+```
+
+Clearly need to interpret those "hazards" carefully.  Some of the really high ones are driven by some of the extreme samples.
+
+```{r, fig.width=10, fig.height=10}
+
+gains = subset(chh, coef > 0)
+
+y = cbind.data.frame(labels, dysplasia.df[,gains$label])
+y$labels = factor(y$labels)
+y$labels = plyr::revalue(y$labels, c('0'='NP','1'='P'))
+
+m = melt(y, id.vars='labels')
+colnames(m) = c('status','label','value')
+m = base::merge(m, gains, by='label')
+m$chr = factor(m$chr, levels=chr.lengths$chrom, ordered = T)
+
+p = ggplot(m) + facet_wrap(chr~label, scales='free_y') +
+  geom_jitter(aes(status, value, group=status, color=status), width=0.01) + 
+  geom_violin(aes(status, value, group=status, fill=status, color=status), alpha=0.5) + 
+  geom_hline(yintercept=c(1,-1), color='grey39', linetype='dotted') +
+  geom_label(data=subset(gains, hazards >= 1), aes('P',0, label=round(hazards,2)), color='red') + 
+  geom_label(data=subset(gains, hazards < 1), aes('P',0, label=round(hazards,2)), color='grey') + 
+  labs(y='',x='',title='Gains') +
+  plot.theme + theme(legend.position = 'none', strip.text=element_text(size=9)) 
+p
+ggsave('plots/gains.png', plot=p, width=10, height=10, unit='in', dpi=300)
+
+losses = subset(chh, coef < 0)
+
+y = cbind.data.frame(labels, dysplasia.df[,losses$label])
+y$labels = factor(y$labels)
+y$labels = plyr::revalue(y$labels, c('0'='NP','1'='P'))
+
+m = melt(y, id.vars='labels')
+colnames(m) = c('status','label','value')
+m = base::merge(m, losses, by='label')
+m$chr = factor(m$chr, levels=chr.lengths$chrom, ordered = T)
+
+p = ggplot(m) + facet_wrap(chr~label, scales='free_y') +
+  geom_jitter(aes(status, value, group=status, color=status), width=0.01) + 
+  geom_violin(aes(status, value, group=status, fill=status, color=status), alpha=0.5) + 
+  geom_hline(yintercept=c(1,-1), color='grey39', linetype='dotted') +
+  geom_label(data=subset(losses, hazards >= 1), aes('P',3, label=round(hazards,2)), color='red') + 
+  geom_label(data=subset(losses, hazards < 1), aes('P',3, label=round(hazards,2)), color='grey') + 
+  labs(y='',x='',title='Losses') +
+  plot.theme + theme(legend.position = 'none', strip.text=element_text(size=9)) 
+p
+ggsave('plots/losses.png', plot=p, width=10, height=10, unit='in', dpi=300)
+
 ```
@@
 One way of looking at the p53 prediction (e.g. Adam Bass) is to see how predictive the p-arm loss in chr17 is.
+
 ```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=10}
@@
 cols = c('Hospital.Research.ID','Patient','Status','PID', 'Samplename','Endoscopy.Year','Pathology','Prediction','OR')
+dspred$Illumina.ID = rownames(dspred)
 dspred = dspred %>% dplyr::mutate( 'Adj. RR'=RR+offsetMean )
-
 dspred = dspred %>% dplyr::mutate( 'Adj. Prob'=1/(1+exp(-RR+abs(offsetMean)) ))
 
-ggplot(dspred, aes(Type, `Adj. Prob`)) + geom_jitter()
-
-ggplot(dspred, aes(Type, `Adj. RR`)) + geom_jitter() + geom_boxplot(outlier.color = NA)
-
-head(dspred)
+x = 'LP6005691-DNA_A02'
+subset(dspred, Illumina.ID == x)
+
 m = melt(dspred, measure.vars=c('RR','Adj. RR'))

### 2018-05-02 ElliesThesisFigs.Rmd
Moved shiny app to own repo

---
title: "ThesisStuff"
author: "Sarah Killcoyne"
date: "4/12/2018"
output: 
  html_document:
    fig_height: 5
    fig_width: 5
    toc: yes
    toc_depth: 4

---
```{r setup, include=FALSE}
#library(biomaRt)
#library(GenomicRanges)
#library(stringr)
library(ggplot2)
library(ggrepel)
library(GGally)
library(pander)
library(reshape2)
library(gridExtra)
#library(CoxHD)
library(plyr)
library(dplyr)

source('lib/load_patient_metadata.R')
source('lib/cv-pt-glm.R')
source('lib/common_plots.R')
source('lib/data_func.R')

dir.create('thesis', showWarnings = F)

model.performance<-function(p1se, coeffs, lims=NULL) {
  performance = do.call(rbind.data.frame, p1se)
  performance$alpha = rownames(performance)

  if (grepl('chr', performance$alpha[1])) {
    performance$alpha = factor(performance$alpha, levels=c(paste('chr',c(1:22), sep=''), setdiff(rownames(performance), paste('chr',c(1:22),sep=''))) )
  }

  ## check how often the feature(s) is selected at that lamda in each split. "stability selection"
  coef.stable = lapply( coeffs, function(cf) {
    sort(rowSums(cf[,-1]), decreasing=T)
  })
  coef.stable = coef.stable[names(which(sapply(coef.stable, length) > 0))]
  coef.stable = lapply(coef.stable, function(cf) cf/(folds*splits))

  performance = cbind(performance, do.call(rbind,  lapply(coef.stable, function(x) {
    cbind('n.Coef'=length(x), '25%'=length(which(x>=.25)), 
          '50%'=length(which(x>=.5)), '75%'=length(which(x>=.75 )), 
          '100%'=length(which(x==1)) )
  })))

  p = ggplot(performance, aes(alpha, mean, fill=(alpha != 0.9))) + ylim(0,1) + geom_col() +
    geom_text( aes(label=round(mean, 3)), vjust=3) +
    geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme), width=0.3, size=1, color='grey39') + 
    geom_text(aes(y=0.1, label=n.Coef), ) +
    geom_text(aes(label=paste('(',`75%`,')',sep=''), y=0.05)) + 
    scale_fill_brewer(palette='Paired') +
    #scale_fill_manual(values=c('red', 'grey39')) + 
    theme(legend.position = 'none') +
    labs(x='Elasticnet penalty value: ridge <-> lasso', y='Model classification at lambda-1se')

  if (!is.null(lims))
    p = p+ylim(lims)

  cfs = as.data.frame(matrix(data=0,nrow=length(unique(names(table(unlist(lapply(coef.stable, names)))))), ncol=length(names(coef.stable)), 
                             dimnames=list( unique(names(table(unlist(lapply(coef.stable, names))))), names(coef.stable) )))
  for (i in names(coef.stable)) 
    cfs[intersect(rownames(cfs), names(coef.stable[[i]])), i] = 1
  return(list('performance'=performance, 'stable.coeffients'=coef.stable, 'cf.per.feature'=cfs,'plot'=p))
}

data = '~/Data/Ellie'

cache.dir = paste(data, 'Analysis/5e06-genomic', sep='/')

data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
analysis.files = list.files(paste(data, 'Analysis', sep='/'), full.names=T)

## Hospital.Research.ID info file
patient.file = grep('All_patient_info.xlsx', data.files, value=T)
if (length(patient.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", data))
demo.file = grep('Demographics_full.xlsx', data.files, value=T)

all.patient.info = read.patient.info(patient.file, demo.file, set='all')$info

patient.info = subset(all.patient.info, Set == 'Training')

patient.info = arrange(all.patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
sum.patient.data = summarise.patient.info(patient.info)

load(grep('All_patients.Rdata', analysis.files, value=T), verbose=T)

val.data = patient.data[subset(sum.patient.data, Set == 'Test')$Hospital.Research.ID]
patient.data = patient.data[subset(sum.patient.data, Set == 'Training')$Hospital.Research.ID]

```

# Segment size: 5e06 with arms

# Data set up

To use GLMs across the patients I merge all samples from all patients and overlap the copy number segments. Where segments from a sample get split, the value of the original sample is retained for each of the split samples.

## y( Progressors (1) vs Non (0) )

The labels are split such that all samples from progressor patients are (1) and all samples from non-progressors are (0).

```{r labelsPNP, echo=F, message=F, include=T}

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=T)
#rm(plots, coefs, performance.at.1se, models)

## binomial: prog 1, np 0
labels = unlist(sapply(patient.data, function(df) {
  df$info = arrange(df$info, Endoscopy.Year, Pathology)
  label = as.integer(df$info$Status == 'P') #as.integer(df$info$Pathology %in% c('HGD', 'IMC'))
  names(label) = df$info$Samplename
  return(label)
}))
names(labels) = sub('.*\\.', '',  names(labels))

pts = do.call(rbind, lapply(patient.data, function(df) {
  df$info = arrange(df$info, Endoscopy.Year, Pathology)
  cbind(df$info[,c('Hospital.Research.ID','Samplename')])
}))
rownames(pts) = 1:nrow(pts)

# sort in label order
if (length(setdiff(rownames(dysplasia.df), names(labels))) > 3)
  warning("Labels vector is missing samples")

riskCols = RColorBrewer::brewer.pal(11, "RdYlBu")[c(10, 1)]
```

We have `r table(labels)[1]` samples from non-progressors and `r table(labels)[2]` samples from progressors.

Example subset of the data matrix:
`r pander(dysplasia.df[1:5, 1:5], caption=paste("dimensions:", paste(dim(dysplasia.df), collapse=', ')), justify='left')`

## Build/Cross-validate by patient

Keeping in mind that the matrix is built on a sample, not patient, basis while the labels (Progressor/Non) are on a per-patient basis I generated the cross validated models for various values of alpha by running 10 or more iterations of 5-fold cross validation that pulled out all the samples for 5 patients at each fold.  1000 lambda values.

```{r, echo=F, message=F, warning=F}

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}

do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits'))

```

## Model performance

```{r model_perf, message=F, warning=F, echo=F, fig.width=6, fig.height=6}
folds = 10; splits = 5 
classification.rate = as_tibble(matrix(ncol=5, nrow=0, dimnames=list(c(), c('model','mean','sme','stable.coef', 'all.coef'))))

mp = model.performance(performance.at.1se, coefs)
mp$plot = mp$plot + labs(title=paste('All samples,', folds, 'folds,', splits, 'patient splits'))
mp$plot
ggsave('thesis/all_perf.png', plot=mp$plot, height=3.5, width=3.5, units='in', dpi=300, scale=1.2)

select.alpha = as.character(0.9)
# Genomic regions only
most.stable = grep('^\\d+', names(which(mp$stable.coeffients[[select.alpha]] >= 0.75 )), value=T)

most.stable = cbind.data.frame(do.call(rbind, strsplit(most.stable, ':|-')), most.stable)
colnames(most.stable) = c('chr','start','stop', 'name')

most.stable$chr = as.integer(as.character(most.stable$chr))
most.stable = arrange(most.stable, chr, start)

classification.rate = add_row(classification.rate, 'model'='All samples', 'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))

```

`r pander(mp$performance[,c('n.Coef','25%','50%','75%', '100%')], justify='left', caption='Number of features stable in n% or more folds')`

## Feature Stability

They share all of the non-zero coefficients (`r which.max(sapply(mp$stable.coeffients, length))` identifies the most non-zero coefs `r max(sapply(mp$stable.coeffients, length))`).

## Remove HGD/IMC & LGD Samples and check the model

Just as a point, the initial model is trained against `r nrow(dysplasia.df)` samples with `r ncol(dysplasia.df)` features. The samples include the final (diagnostic) HGD/IMC samples from progressors.

### Remove HGD/IMC

```{r noHGD, echo=F, message=F, warning=F, fig.width=6, fig.height=6}
rm(performance.at.1se, coefs, mp)

file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}
nohgd.coefs = coefs[[select.alpha]]

mp = model.performance(performance.at.1se, coefs)
mp$plot = mp$plot + labs(title='No HGD/IMC samples')
mp$plot
ggsave('thesis/no_hgd_perf.png', plot=mp$plot, height=3.5, width=3.5, units='in', dpi=300, scale=1.2)

classification.rate = add_row(classification.rate, 'model'='No HGD/IMC', 
                              'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))

rm(performance.at.1se, coefs, mp, no.hgd.plots)
```

### Remove LGD

```{r noLGD, echo=F, warning=F, message=F, fig.width=6, fig.height=6, eval=T}
file = paste(cache.dir, 'nolgd.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}

mp = model.performance(performance.at.1se, coefs)
mp$plot = mp$plot + labs(title='No LGD/HGD/IMC samples')
mp$plot
ggsave('thesis/no_lgd_perf.png', plot=mp$plot, height=3.5, width=3.5, units='in', dpi=300, scale=1.2)

classification.rate = add_row(classification.rate, 'model'='No LGD/HGD/IMC', 'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))

rm(performance.at.1se, coefs, mp, nolgd.plots)
```

## Overall

Even without dysplastic samples our model is well able to classify samples as either progressive or non.

```{r, echo=F, warning=F, message=F, fig.width=4, fig.height=6}
p = ggplot(classification.rate, aes(model, mean, group=model, fill=model)) + geom_col() +
  geom_text( aes(label=round(mean, 2)), vjust=2, size=5, color='gray39', show.legend=F) +
  geom_text( aes(y=0.6, label=paste(all.coef, " coef\n(", stable.coef, ")", sep='') ), color='gray39', size=5) +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme),color='grey39', width=0.1, size=1) + 
  labs(x='', y='Classification rate', title=paste('Classification rate per model at', select.alpha)) + 
  theme(axis.text.x=element_text(angle=45,hjust=1,face="bold",size=12), legend.position = 'none') +
  ylim(0,1)
p
ggsave('thesis/compare_classification.png', plot=p, height=5, width=3, units='in', dpi=300 )
```

# Features

```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=5}
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=T)
rm(plots, performance.at.1se)

featureStability = rowSums(coefs[[select.alpha]][,-1])/(splits*folds)

hazards = apply( exp(t(dysplasia.df[, rownames(coefs[[select.alpha]])]) *  coefs[[select.alpha]][,1]), 1, function(x) {
  sd(x)/mean(x)
})
ch = as.data.frame(hazards)
ch$coef = coefs[[select.alpha]][,1]
ch$stability = featureStability
ch$label = rownames(ch)
ch = arrange(ch, -hazards)

qq = quantile(ch$hazards, seq(0,1,.15))
p = ggplot(ch, aes(coef, hazards, col=hazards >= qq[length(qq)])) + 
  geom_point(alpha=0.8, show.legend = F) + 
  geom_text_repel(aes(x=coef, label=ifelse(hazards >= qq[length(qq)], label, '')),show.legend=F) +
  scale_color_manual(values=c('firebrick3','darkblue')) +
  labs(title='Top features by hazard ratio', x='Coefficient value', y='Hazard') 
p
ggsave('thesis/hazards.png', plot=p, width=5, height=5, units='in', dpi=300)

chtable = ch[,c('hazards','coef','stability','label')]
chtable = chtable[,c('label','hazards', 'coef')]
chtable$coef[which(chtable$coef > 0)] = 'Gain'
chtable$coef[which(chtable$coef != 'Gain')] = 'Loss'
colnames(chtable) = c('Location','Hazard', 'Gain/Loss')
chtable$Hazard = round(chtable$Hazard, 3)
#write.table(chtable, sep='\t', quote=F, row.names = F, file='hazards.txt')

```

`r pander(ch, caption='Selected coefficients')`

# Nested leave-one-out predictions

Leave out each patient (both P and NP) and run a new CV fit each time then predict the samples for the patient that was left out.  Fitted model includes HGD samples.

```{r leaveoneout, echo=F, message=F, warning=F, fig.width=5, fig.height=5}
file = paste(cache.dir, 'loo.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  stop(paste("Missing model file", file))
}
#loo.coefs = nzcoefs

names(performance.at.1se) = names(pg.samp)
df = as.data.frame(performance.at.1se)

p = ggplot(df, aes(y=performance.at.1se, x='LOO')) + 
  geom_boxplot(fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) + geom_jitter(color='grey39', width=0.3) +
  geom_label(data=subset(df, performance.at.1se %in% range(performance.at.1se)), aes(y=performance.at.1se, label=round(performance.at.1se, 3)), fontface='bold', fill='lightblue') +
  labs(y='Classification rate', x='LOO model', title='Performance for leave-one-out', subtitle='All samples model')
p
ggsave('thesis/performance_loo.png', plot=p, width=5, height=5, units='in', dpi=300)

```

### Prediction ROC curves

Based on these curves we get our best AUC at ~0.36.  Currently we are using 0.5 by default.

```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=7, fig.width=7}
preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
roc$model = 'CN Model'

pROC::ci.auc(roc)

roc$is = 'all'

ggsave(filename='thesis/all_roc.png', plot=roc.plot(roc), width=6, height=4, units='in', dpi=300)

cutoff = round(pROC::coords(roc, 'best')[['threshold']], 2)
```

### p53 ROC curve

```{r p53roc, echo=F, fig.height=2, fig.width=2}
p53samples = subset(patient.info, !is.na(p53.Status), select=c('Samplename','p53.Status','Status'))
p53roc = pROC::roc(as.numeric(p53samples$Status)-1, as.integer(p53samples$p53.Status)-1)

pROC::ci.auc(p53roc)

p53roc$model = 'p53 IHC'
roc.plot(p53roc, 'p53 IHC')

ggsave(filename='thesis/p53_roc.png', plot=roc.plot(p53roc, 'p53 IHC') + plot.theme, width=6, height=6, units='in', dpi=300)
```

### Pathology ROC curve

Based on HGD/IMC/LGD

```{r pathroc, echo=F, fig.height=4, fig.width=10 }
pathsamples = patient.info[,c('Pathology','Status')]
pathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('LGD','HGD','IMC'), 1, 0) )
pathroc$model = 'HGD & LGD'

pROC::ci.auc(pathroc)

ggsave('thesis/dysplasia_auc.png', plot=roc.plot(pathroc, title='Dysplasia (HGD & LGD)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

hgdpathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('HGD','IMC'), 1, 0) )
hgdpathroc$model = 'HGD Only'

pROC::ci.auc(hgdpathroc)

ggsave('thesis/hgd_auc.png', plot=roc.plot(hgdpathroc, title='HGD Only'), width=6, height=6, units="in", limitsize=F, dpi = 300)

grid.arrange(roc.plot(pathroc, 'Pathology (LGD,HGD,IMC)'), roc.plot(hgdpathroc, 'Pathology (HGD,IMC)'), ncol=2)
```

### Per Endoscopy ROCs
```{r, echo=F, warning=F, message=F,fig.height=4, fig.width=4}
pendo = do.call(rbind, pg.samp) %>% group_by(Patient, Status, Hospital.Research.ID, PID ) %>% dplyr::summarise(
  'max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(PID))

rocEndoM = pROC::roc(Status ~ max, data=pendo, ci=T, of='thresholds')
pROC::coords(rocEndoM, 'best')
rocEndoM$model = 'Per Endoscopy Max'
ggsave('thesis/maxendo_auc.png', plot=roc.plot(rocEndoM, title='Max per Endoscopy') , width=6, height=6, units="in", limitsize=F, dpi = 300)

rocEndoA = pROC::roc(Status ~ avg, data=pendo, ci=T, of='thresholds')
pROC::coords(rocEndoA, 'best')
rocEndoA$model = 'Per Endoscopy Avg'
ggsave('thesis/avgendo_auc.png', plot=roc.plot(rocEndoA, title='Avg per Endoscopy') , width=6, height=6, units="in", limitsize=F, dpi = 300)

lastEndo = do.call(rbind, lapply(pg.samp, function(pt) {
  cbind(pt[1,c('Patient','Hospital.Research.ID', 'Status')],  'avg'=mean(subset(pt, PID == pt[nrow(pt), 'PID'])$Prediction))
}))

rocLastEndo = pROC::roc(Status ~ avg, data=lastEndo, ci=T, of='thresholds')
pROC::coords(rocLastEndo, 'best')
rocLastEndo$model = 'Last endoscopy'
ggsave('thesis/lastendo_auc.png', plot=roc.plot(rocLastEndo, title='Last endoscopy') , width=6, height=6, units="in", limitsize=F, dpi = 300)

firstEndo = do.call(rbind, lapply(pg.samp, function(pt) {
  cbind(pt[1,c('Patient','Hospital.Research.ID', 'Status')],  'avg'=mean(subset(pt, PID == pt[1, 'PID'])$Prediction))
}))

rocFirstEndo = pROC::roc(Status ~ avg, data=firstEndo, ci=T, of='thresholds')
pROC::coords(rocFirstEndo, 'best')
rocFirstEndo$model = 'First endoscopy'
ggsave('thesis/firstendo_auc.png', plot=roc.plot(rocFirstEndo, title='First Endoscopy') , width=6, height=6, units="in", limitsize=F, dpi = 300)

```

### Per Patient ROC

```{r echo=F, fig.height=4, fig.width=4}
#ptendo = do.call(rbind, pg.samp) %>% group_by(Patient, Hospital.Research.ID, Status) %>% summarise('max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(Patient))
ptendo = do.call(rbind, lapply(pg.samp, function(pt) {
  cbind(pt[1,c('Patient','Hospital.Research.ID', 'Status')],  'mean'=mean(subset(pt, PID != pt[nrow(pt), 'PID'])$Prediction), 'max'=max(subset(pt, PID != pt[nrow(pt), 'PID'])$Prediction))
}))

ptendo = ptendo[-which(is.na(ptendo$mean)),]

rocPtM = pROC::roc(Status ~ max, data=ptendo, ci=T, of='thresholds')
pROC::coords(rocPtM, 'best')
rocPtM$model = 'Per Patient Max'
ggsave('thesis/maxpt_auc.png', plot=roc.plot(rocPtM, title='Max per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

rocPtA = pROC::roc(Status ~ mean, data=ptendo, ci=T, of='thresholds')
pROC::coords(rocPtA, 'best')
rocPtA$model = 'Per Patient Avg'
ggsave('thesis/avgpt_auc.png', plot=roc.plot(rocPtA, title='Avg per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

```

### Compare ROCs

```{r echo=F, fig.height=8, fig.width=6}
get.roc.stat<-function(roc, x='best') {
  auc = c( pROC::auc(roc), range(pROC::ci.auc(roc))  )

  sens = c(pROC::coords(roc, x)[['sensitivity']], pROC::ci.se(roc,pROC::coords(roc, x)[['specificity']], conf.level=0.95))
  spec = c(pROC::coords(roc, x)[['specificity']], pROC::ci.sp(roc,pROC::coords(roc, x)[['sensitivity']], conf.level=0.95))

  stat = rbind('AUC'=auc, 'Sensitivity'=sens[c(1,2,4)], 'Specificity'=spec[c(1,2,4)])
  colnames(stat) = c ('value','CI.min','CI.max')
  stat = as.data.frame(stat)
  stat$model = roc$model
  stat$Measure = rownames(stat)
  stat
}
roc$model = "CN Model"
m = rbind.data.frame(
  get.roc.stat(p53roc),
  get.roc.stat(hgdpathroc),
  get.roc.stat(pathroc),
  get.roc.stat(roc))
m$model = factor(m$model, levels=c('p53 IHC', 'HGD Only','HGD & LGD','CN Model'), ordered = T)

rocList = list(roc,pathroc,hgdpathroc, p53roc)
multi.roc.plot(rocList)

cols = RColorBrewer::brewer.pal(4, "PRGn")
dodge = position_dodge(width=0.9)

comp = ggplot(melt(subset(m, Measure == 'AUC'), measure.vars='value'), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Measure)) + 
  ylim(0,100) + 
  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
  geom_text(aes(y=(value*100)-4, label=paste(round(value,2)*100,'%',sep='')), position=dodge, color='white')  +
  scale_fill_manual(values=cols, name='') + plot.theme + 
  theme(legend.position = 'none', axis.text.x = element_text(hjust=1,angle=45)) + labs(x='', y='AUC', title='') 
comp

ggsave('thesis/path_p53_cn_auc.png', plot=comp, width=3.1, height=3.9, units="in", limitsize=F, dpi = 300, scale=1)

```

```{r echo=F, fig.height=8, fig.width=6}
roc$model = "Per Sample"

m = rbind.data.frame(
  get.roc.stat(roc),
  get.roc.stat(rocEndoM),get.roc.stat(rocEndoA),
  get.roc.stat(rocPtM), get.roc.stat(rocPtA),
  get.roc.stat(rocFirstEndo), get.roc.stat(rocLastEndo))
m$model = factor(m$model)

subset(m, model %in% c(rocEndoA$model, rocEndoM$model) & Measure == 'AUC')

cols = RColorBrewer::brewer.pal(11, 'Paired')
dodge = position_dodge(width=0.9)
mm = melt(subset(m, Measure == 'AUC'), measure.vars='value')
p = ggplot( arrange(mm, CI.min), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=model)) + 
  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='grey39' ) + 
  geom_text(aes(y=CI.min*100, label=paste(round(value,2)*100,'%',sep='')), position=dodge, vjust=3)  +
  scale_fill_manual(values=cols, name='') + plot.theme + 
  theme(legend.position='none', axis.text.x=element_text(angle=45, hjust=1)) + labs(x='', y='AUC', title='ROC Comparison') 
p

ggsave('thesis/multi_auc.png', plot=p, height=6, width=5, units='in',dpi=300)

multi.roc.plot(list(roc, rocEndoM, rocEndoA, rocPtM, rocPtA,rocFirstEndo,rocLastEndo), 'ROC Comparisons', colors=rev(cols[1:7]))
```

## Prediction probability distribution

```{r back, echo=F, message=F, warning=F}
transform.by.time<-function(pt, info) {
  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction')] %>% group_by(PID, Samplename) %>% dplyr::mutate('months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12)
  pt = dplyr::arrange(merge(pt, subset(info, Hospital.Research.ID == pt$Hospital.Research.ID[1], select=c('Samplename','Block')), by='Samplename'), Pathology, PID, Endoscopy.Year)

  tb = with(pt, table(months.before.final, PID))
  tb[tb>0] = 1
  for (mnths in names(which(rowSums(tb) > 1))) {
    tmp = pt[which(pt$months.before.final == mnths),]
    tmp = arrange(tmp, rev(Pathology), PID)

    mt = tmp$months.before.final[1] 
    interval = length(tmp$Endoscopy.Year)/(length(unique(tmp$PID))-1)
    for (pid in unique(tmp$PID)) {
      rows = which(tmp$PID == pid)
      if (length(grep('HGD|IMC',tmp[rows,'Pathology'])) > 0 & mt != 0) {
        mt = 0
        tmp$months.before.final[rows] = mt
      } else {
        mt = mt + interval
        tmp$months.before.final[rows] = mt
      }
    }
    pt[which(pt$months.before.final == mnths),] = tmp
  }

  pt$nl = (pt$Block-1)*2
  arrange(pt, -Endoscopy.Year, PID)
}

patient.info = subset(patient.info, Patient %in% sum.patient.data$Patient)

back = do.call(rbind.data.frame,lapply(pg.samp, transform.by.time, info=patient.info))
back = merge(back, patient.info[,c('Samplename','p53.Status')])

sq = c(0, seq(1,132,12),max(back$months.before.final))
back$brks = cut(back$months.before.final, 
                breaks=sq, 
                labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-c(length(sq)-1,length(sq))], '>120'),
                include.lowest = T, ordered_result = T)
back$brks = factor(back$brks, rev(levels(back$brks)))

back$Pathology = as.character(back$Pathology)
back$Pathology = factor(back$Pathology, levels=c('BE','ID','LGD','HGD', 'IMC'), ordered = T)

```

The labels in each decile are the ratio of progessor samples that have probabilities within that region. The deciles that are greyed out show insigificant enrichment for the NP/P ratio.

```{r spacial, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
cuts = seq(0,1,0.1)

back$quants = with(back, cut(Prediction, breaks=cuts))
qt = as.data.frame.matrix(table(back$quants, back$Status))
qt$quant = rownames(qt)

ft.fun<-function(NP,P) {
  f = chisq.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
  cbind.data.frame('p.value'=f$p.value)
  #cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
}

qt = cbind(qt, back %>% group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))#, 
pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% dplyr::mutate( 
  'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value,  'conf'=ifelse(p.value < 0.05, '*', '') )

pred.confidence[c('r1','r2')] = NA
for (i in 1:(length(cuts)-1)) 
  pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)

pred.confidence$Risk = 'Moderate'
pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc < .5)) ] = 'Low'
pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'

pred.confidence = bind_cols(pred.confidence, 
                            data.frame(ci.low=qbeta(0.025, shape1=pred.confidence$P+.5, shape2 = pred.confidence$NP+.5),
                                       ci.high=qbeta(0.975, shape1=pred.confidence$P+.5, shape2 = pred.confidence$NP+.5)))

annotation.table = arrange(pred.confidence %>% group_by(Risk) %>% dplyr::summarise('n(NP)'=sum(NP), 'n(P)'=sum(P), 'p.value'=format.pval(mean(p.value), 3), 'min'=min(r1), 'max'=max(r2)), min)

tt3 <- ttheme_minimal(
  core=list(bg_params = list(fill = riskCols, col=NA, alpha=0.2),
            fg_params=list(fontface=3)),
  colhead=list(fg_params=list(col="black", fontface=4L)),
  rowhead=list(fg_params=list(col=c('red3','green3', 'orange2'), fontface=3L)),
  padding=unit(c(5,5),'mm'))

#grid.arrange( tableGrob(annotation.table, rows=NULL, theme=tt3) )
pred.confidence$Risk = factor(pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)

lows = chisq.test(rbind(colSums(pred.confidence[1:5,c('NP','P')]),table(sum.patient.data$Status)))
highs = chisq.test(rbind(colSums(pred.confidence[6:10,c('NP','P')]),table(sum.patient.data$Status)))

mods = chisq.test(rbind(colSums(pred.confidence[4:7,c('NP','P')]),table(sum.patient.data$Status)))
```

```{r pcalib, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
#pred.confidence$Risk[1:5] = 'Low'
#pred.confidence$Risk[6:10] = 'High'
#pred.confidence$Risk[4:7] = 'Moderate'

pcal = ggplot(pred.confidence, aes(mn, perc)) + 
  geom_rect(data=pred.confidence, aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + 
#  scale_fill_manual(values=riskCols, limits=levels(pred.confidence$Risk) ) +
  geom_point() + geom_abline(slope=1, intercept=0, color='grey39', linetype='dotted', size=0.5) + #geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), size=0.5, width=0.01) + labs(x='P(P)', y='NP:P', title='') +
#  scale_color_manual(values=riskCols,limits=levels(pred.confidence$Risk) ) + 
  scale_x_continuous(limits=c(0,1), breaks = cuts, labels = c(0.0, '','',0.3,'','','',0.7,'','',1.0)) +
  plot.theme + theme(legend.position = 'none')
#pcal

#ggsave(filename='thesis/pred_calib.png', plot=pcal, width=2.2, height=2.2, units='in', dpi=300, scale=1.5)

back = merge(back, pred.confidence[c('quant','Risk')], by.x='quants',by.y='quant')
back = back %>% group_by(Risk) %>% dplyr::mutate('se(Risk)'=(sd(Prediction)/sqrt(length(Risk)))*1.96 )

```

```{r relrisk, echo=F, fig.width=6, fig.height=6}
myPal = rev(RColorBrewer::brewer.pal(11, 'RdYlBu'))

predhist = ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') + 
  plot.theme + labs(title='Predictions, all samples model', y='n Samples', x='Probability') 
predhist

ggsave(filename='thesis/pred_hist.png', plot=predhist, width=5, height=5, units='in', dpi=300)
```

At each quintile are the percentage of progressor samples that were predicted within that range, and the hypergeometric p-value (Fishers) for the significance within that quintile. Then our predictions are at best a coin flip in the middle, but highly significant at the outer edges. This would also push the prediction for "high" risk up to `r min(subset(pred.confidence, labels == 'High' )$r1)` while "low" risk is below `r max(subset(pred.confidence, labels == 'Low' )$r2)`.

`r pander(as.data.frame(pred.confidence))`

## How do the predictions look at each time point?

Here we'll just use the cutoff 0.5 (I think that's what we used in the thesis?).

```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
back[back$Prediction > 0.5, 'Risk'] = 'High'
back[back$Prediction < 0.5, 'Risk'] = 'Low'
back$Risk = droplevels(back$Risk)

riskPath = rbind( cbind.data.frame( table(subset(back, Status == 'P', select=c('Pathology','Risk'))), 'Status'='Progressor' ), 
cbind.data.frame( table(subset(back, Status == 'NP', select=c('Pathology','Risk'))), 'Status'='Non-Progressor' ))

pathTotal = back %>% group_by(Status, Pathology) %>% dplyr::summarise( 'nPath'=length(Pathology))
pathTotal$Status = ifelse(pathTotal$Status == 'P', 'Progressor','Non-Progressor')

riskPath = merge(riskPath, pathTotal)
riskPath$ratio = with(riskPath, round(Freq/nPath, 3))
riskPath$Risk = factor(riskPath$Risk, levels=levels(back$Risk), ordered = T)

p53Path = cbind.data.frame(table(subset(back, Status == 'P', select=c('Pathology','p53.Status'))), 'Status'='p53 IHC')
p53Path = merge( subset(back, Status == 'P' & !is.na(p53.Status)) %>% group_by(Pathology) %>% dplyr::summarise( 'nPath'=length(Pathology)), p53Path)
#p53Path$Risk = with(p53Path, ifelse(p53.Status == 1, 'High','Low'))
p53Path$ratio = with(p53Path, round(Freq/nPath, 2))

#riskPath = rbind(riskPath, p53Path[colnames(riskPath)])
#riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor','p53 IHC'), ordered = T)
riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor'), ordered = T)

pathTotal = riskPath %>% group_by(Status, Pathology) %>% dplyr::summarise('nPath'=unique(nPath))

#riskPath$Pathology = as.character(riskPath$Pathology)
#riskPath$Pathology[riskPath$Pathology == 'BE'] = 'NDBE'
#riskPath$Pathology = factor(riskPath$Pathology, levels=c('NDBE','ID','LGD','HGD','IMC'), ordered=T)

rp = ggplot(riskPath, aes(Pathology, ratio*100)) + 
  geom_bar(data=riskPath, aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
  scale_fill_manual(values=riskCols, limits=levels(riskPath$Risk)) + 
  scale_color_manual(values=c('black','black','black'), limits=levels(pred.confidence$Risk) ) +
  scale_x_discrete(labels=c('NDBE','ID','LGD','HGD','IMC')) +
  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=102), position=position_stack()) +
  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep=''), color=Risk), position=position_stack(vjust = 0.5), show.legend = F) +
  facet_grid(~Status, scales='free_x', space = 'free_x') +  
  labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'none')
rp

ggsave(filename='thesis/pred_by_path.png', plot=rp, width=5.8, height=4.5, units='in', dpi=300, scale=1)

```

```{r p53IHC, echo=F, warning=F, message=F, fig.height=8, fig.width=8}
p53Path$p53.Status = factor(p53Path$p53.Status, ordered = T)
p53Path = arrange(p53Path, Pathology, p53.Status)
col = RColorBrewer::brewer.pal(11, "PRGn")[c(10,2)]
rp = ggplot(p53Path, aes(Pathology, ratio*100, group=Pathology, fill=p53.Status)) + geom_bar(stat='identity', position = position_stack()) +
  scale_fill_manual(values=col, name='p53 IHC', labels=c('Normal','Aberrant')) + 
  scale_x_discrete(labels=c('NDBE','ID','LGD','HGD','IMC')) +
  geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=103)) +
  geom_text(aes(label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5,color='white') +
  labs(title='', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
rp
ggsave(filename='thesis/p53_pred.png', plot=rp, width=4.5, height=3.6, units='in', dpi=300, scale=1.2)

```

```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=6}
all.preds = back %>% group_by(Status, Patient) %>% dplyr::summarise( 'Low'=sum(length(which(Risk == "Low"))),
#'Moderate'=sum(length(which(Risk == "Moderate"))),
'High'=sum(length(which(Risk == "High"))),
'n.samples'=length(Patient) )

all.preds = melt(all.preds, id.vars=c('Status','Patient', 'n.samples'))
head(all.preds)

ggplot(all.preds, aes(Patient, value/n.samples, group=variable, fill=variable)) +
  geom_bar(stat='identity', position = position_stack()) +
  scale_fill_manual(values=riskCols) +
  geom_text(aes(y=1.01, label=n.samples)) +
  facet_wrap(~Status, scales='free_x') + labs(y='Sample ratios', title='Per patient ratio of predictions') +
  plot.theme

```

## Backwards from HGD

Clinicians always ask to see the timepoints peeled backwards to see how early prediction is possible.

Take per endoscopy average prediction.

There are a lot more moderate predictions in this model...how does this work with this plot??

```{r enods, warning=F, message=F, echo=F, fig.height=8, fig.width=6}
endos = back %>% group_by(Status, months.before.final, brks, PID) %>% dplyr::summarise(
  'pred.endo'=ifelse(length(which(Risk == 'High')) > 0, 1, 0), 
  'n.endos'=length(unique(PID)),
  'n.samples'=length(Samplename),
  'mean.pred'=mean(Prediction),
  'pred.samples'=length(which(Risk == 'High')),
  'max.path'=max(Pathology)
  )

sq = c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12)
mlab = c('Endpoint','1-24','24-48','48','48-72','72-96','96-120','>120')

# months = 12
# sq = c(0, 1, 1*months, 2*months, 4*months, 6*months, 8*months, 10*months, 12*months)
# mlab = c()
# for (i in 1:length(sq)) {
#   mlab = c(mlab, paste(sq[i], sq[i+1], sep='-'))
# }
# mlab[1] = 'Endpoint'
# mlab = mlab[1:(length(sq)-1)]
# mlab[length(mlab)] = paste('>',sq[(length(sq)-1)],sep='')
#

endos$brks = cut(endos$months.before.final, include.lowest = T, ordered_result = T,breaks=sq ,labels=mlab)
endos$brks = with(endos, factor(brks, rev(levels(brks)))) # invert order

endos.r = endos %>% group_by(Status, brks) %>% dplyr::summarise( 
  'pred.r'=sum(pred.endo)/sum(n.endos),'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos))

#endos.r = subset(endos, Status == 'P') %>% group_by(brks) %>% dplyr::summarise( 
#  'pred.r'=sum(pred.endo)/sum(n.endos),'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos))

tmp = back # subset(back, Status == 'P')
tmp$brks = cut(tmp$months.before.final, include.lowest = T, ordered_result = T,breaks=sq,labels=mlab)
tmp = tmp %>% group_by(Status, brks) %>% dplyr::summarise( 'mean(P)'=mean(Prediction),'sem(P)'=sd(Prediction)/sqrt(length(brks)) )

endos.r = arrange(merge(endos.r,tmp), brks)
endos.r$Status = plyr::revalue(endos.r$Status, c('NP'='Non-Progressor', 'P'='Progressor'))

rp = ggplot(endos.r, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) + ylim(0,110) +
  facet_grid(~Status) +
  geom_bar(stat='identity', show.legend = F) + 
  geom_errorbar(color='grey21', width=0.2, alpha=0.5) +
  geom_text(aes(y=pred.r*100+2, label=paste(round(pred.r*100), '%', sep='')), color='black') +
  geom_text(aes(y=10, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
  labs(x='Months prior to endpoint', y='Percent predicted', title='High risk predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') +
  theme( axis.text.x=element_text(angle=45,hjust=1) ) 
rp
ggsave(filename='thesis/pred_endo.png', plot=rp, width=11, height=5, units='in', dpi=300, scale=1)

```

```{r heatmaps, echo=F, warning=F, message=F}
back$nl = factor(back$nl, ordered=T)
back$Risk = factor(back$Risk, levels=c('Low','High'), ordered = T)
back$Patient = factor(as.numeric(back$Patient), ordered = T)

table.path.risk<-function(df) {
  summary.t = as.data.frame.matrix(with(df, table(Pathology, Risk)))
  summary.t$Pathology = rownames(summary.t)

  summary.t = summary.t %>% group_by(Pathology) %>% dplyr::mutate( 'High.r'=round(High/sum(High,Low),3), 
                                                            'Low.r'=round(Low/sum(High,Low),3))
  for (g in unique(df$Risk)) {
    for (i in 1:nrow(summary.t)) {
      summary.t[i,g] = paste(summary.t[i,g], ' (', summary.t[i,paste(g,'.r',sep='')]*100, '%)',sep='')
    }
  }
  summary.t[,c('Pathology',levels(df$Risk))]
}

summary.t = table.path.risk(back)
```

`r pander(as.data.frame(summary.t))`

```{r, echo=F, message=F, warning=F, width=26, height=8}
back = back %>% group_by(Patient) %>% dplyr::mutate( 'pred.ratio'=round(sum(Prediction)/length(Prediction),2), 'n.endos'=length(unique(months.before.final)), 'max.block'=max(Block))
back = transform(back, Patient=reorder(Patient, max.block))
back$Block = factor(back$Block, ordered = T)

#pred.tiles(subset(back, Patient %in% c(1,20,34,21,3, 19)), cols=riskCols, probs=F, space='free', scales='free_x', drop=T) + theme(legend.position = 'none')

pred.tiles(subset(back, Status == 'P'), cols=riskCols, probs=F, space='free', scales='free_x', drop=T) + labs(title='Progressor')
pred.tiles(subset(back, Status == 'NP'), cols=riskCols, probs=F, space='free', scales='free_x', drop=T) + labs(title="Non-Progressor")

#all = pred.tiles(back, cols=riskCols, path=4, probs=F, scales='free', space='free') + labs(title='', x='', y='') + #theme(legend.position = 'bottom')
#riskLegend = get.legend(all)
```

# AP Clustering

```{r apclust-func, echo=F, message=F, warning=F}
apclust.data<-function(segdata, samples) {
  x1 = as.data.frame(segdata[,intersect(colnames(segdata), samples)])
  #x1 = x1[, samples]
  colnames(x1) = intersect(colnames(segdata), samples)
  rownames(x1) = (segdata[,c(1:4)] %>%
    rowwise() %>%
    mutate(location=paste(paste(chrom, arm, sep=''), '.', start.pos, '-', end.pos, sep='')))$location
  return(x1)  
}

apclust<-function(segdata) {
  x1 = apclust.data(segdata)
  # q=0 minimize off-diagonal similarity
  ac = apcluster(negDistMat(r=2), x1, details=T, convits=25, q=0)
  return(list('apres'=ac, 'data'=x1))
}

net.similarity<-function(ac) {
  if (is.list(ac))
    ac = ac$apres
  ac@netsim
}

sum.similarity<-function(ac) {
  if (is.list(ac))
    ac = ac$apres
  ac@dpsim
}

clusters<-function(ac) {
  if (is.list(ac))
    ac = ac$apres
  ac@exemplars
}

plotAC<-function(aclist) {
  plot(aclist$apres, aclist$data)
}

heatmapAC<-function(aclist) {
  heatmap(aclist$apres)
}

# First vs last samples
#x1 = apclust.data(patient.data[[pt]]$seg.vals[,c(1:4, 6, ncol(patient.data[[pt]]$seg.vals))], patient.data[[pt]]$info$Samplename)

```

### Raw segmentation values

```{r apclust, echo=F, message=F, warning=F, fig.height=10, fig.width=10, fig.align='left'}
library(apcluster)
sample.vals = list()
q=0.2

sum.patient.data = subset(arrange(sum.patient.data, Status), Set == 'Training')

## Using the segmentation values
sum.patient.data$years = with(sum.patient.data, paste(start.year, end.year, sep='-'))
sim = as.data.frame(sum.patient.data[,c('Status','years')])
rownames(sim) = sum.patient.data$Hospital.Research.ID
for (pt in sum.patient.data$Hospital.Research.ID) {

  patient.data[[pt]]$info = arrange(patient.data[[pt]]$info,  Pathology, Endoscopy.Year)
  x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)

  gg = ggplot(melt(x1), aes(x=variable, y=value, group=variable, fill=variable)) + 
    geom_boxplot(notch=T, show.legend=F) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position='none') +
    labs(title=paste(pt, patient.data[[pt]]$info$Status[1], sep=' - ')) + 
    theme(axis.title.x=element_blank(), axis.text.x=element_blank(),axis.ticks.x=element_blank())

  sample.vals[[pt]] = gg

  info = patient.data[[pt]]$info[which(colnames(x1) %in% patient.data[[pt]]$info$Samplename),]
  colnames(x1) = paste(info$Endoscopy.Year, '(', info$Pathology, ')', sep='' )

  ac = apcluster(negDistMat(x1, r=2), q=q, details=T, includeSim=T)

  last = ncol(x1)
  if (patient.data[[pt]]$info$Status[1] == 'P' & length(which(grepl('HGD|IMC', info$Pathology)) > 0)  )
    last = sort(grep('HGD|IMC', info$Pathology), decreasing=T)[1]
  if (ncol(x1) <= 15) {
    #plot(ac, x1[,c(c(1:ncol(x1))[-last], last)], main=pt)
  } else {
    #plot(ac, x1[,c(1, sort(sample(c(1:(ncol(x1)-1))[c(-1,-last)], 8)) ,last) ], main=pt)
  }

  sim[pt, 'ns'] = net.similarity(ac)
  sim[pt, 'ss'] = sum.similarity(ac)
  sim[pt, 'clusters'] = length(clusters(ac))
  sim[pt, 'max.sd'] = max(sapply(x1, sd))

  tmp = patient.data[[pt]]$variable.regions[,-(1:5)]
  if (!is.null(tmp)) {
    sim[pt, 'n.var'] = nrow(patient.data[[pt]]$variable.regions[,-(1:5)])
    if (nrow(tmp) <= 0) {
      sim[pt,c('gains', 'loss')] = c(0,0)
    } else {
      sim[pt,'gains'] = length(tmp[tmp>=1.1])
      sim[pt,'loss'] =  length(tmp[tmp<=0.9])
    }
  } else {
    sim[pt, 'n.var'] = 0
    sim[pt,c('gains', 'loss')] = c(0,0)
  }

}
sim$id = rownames(sim)
rc = cor.test(sim$max.sd, sim$clusters)

#do.call(grid.arrange, c(sample.vals))

# May be a difference
ttps = wilcox.test( subset(sim, Status == 'NP')$ss, subset(sim, Status == 'P')$ss )$p.value
ttpc = wilcox.test( subset(sim, Status == 'NP')$clusters, subset(sim, Status == 'P')$clusters )$p.value

## Number of variable regions isn't so good, but maybe...
#sim %>%
#  group_by(Status) %>%
#  summarise( mean = mean(n.var), sd = sd(n.var), med = median(n.var))
#t.test( subset(sim, Status == 'NP')$n.var, subset(sim, Status == 'P')$n.var )$p.value

g1 = ggplot(sim[,c('Status','ss')], aes(y=ss, x=Status, fill=Status)) + geom_boxplot() + 
  geom_jitter(width=0.2) + stat_boxplot(geom ='errorbar') +
  labs(title='All samples', subtitle=paste("p-value", signif(ttps, 3)), y='sum similarity')

g2 = ggplot(sim[,c('Status','clusters')], aes(y=clusters, x=Status, fill=Status)) + geom_boxplot() + 
  geom_jitter(width=0.2) + stat_boxplot(geom ='errorbar') +
  labs(title='All samples', subtitle=paste("p-value", signif(ttpc, 3)), y='num. clusters')

```

This plot compares the two possible measures of overall complexity from AP. The summed similarity value (top) and the number of clusters (bottom). In this case, the number of clusters appears to be show a difference between the two groups.

```{r echo=F, message=F, warning=F, fig.width=6, fig.height=12}
grid.arrange(g1, g2, top='Raw seg values')
```

## No HGD samples

There is still a difference between the progressors and non, though the p-value does increase. Some of the progressors do start looking more like NPs in their similarity measure but not all of them.

```{r nofinal, echo=F, message=F, warning=F, fig.height=5, fig.width=5}
simX = sum.patient.data[,c('Status','years')]
rownames(simX) = sum.patient.data[,'Hospital.Research.ID']
# Leave out HPC/IMC samples
for (pt in sum.patient.data$Hospital.Research.ID) {
  x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
  hgdsamples = subset(patient.data[[pt]]$info, Pathology %in% c('HGD','IMC'))$Samplename

  rows = rownames(x1)
  col.nm = colnames(x1)
  if (length(hgdsamples) > 0) {
    cols = grep(paste(hgdsamples, collapse='|'), colnames(x1), invert=T)
    x1 = as.data.frame(x1[,cols])
    colnames(x1) = col.nm[cols]
    rownames(x1) = rows
  }

  if (ncol(x1) > 0) {
    ac = apcluster(negDistMat(x1, r=2), q=q)

    simX[pt, 'ss'] = sum.similarity(ac)
    simX[pt, 'clusters'] = length(clusters(ac))

    tmp = patient.data[[pt]]$variable.regions[,-(1:5)]
    simX[pt, 'n.var'] = ifelse (!is.null(tmp), nrow(patient.data[[pt]]$variable.regions[,-(1:5)]), 0)
  }
}

# Correlation between the number of variable regions and clusters, suggesting the complexity measure is reasonable
ct = cor.test(simX$clusters, simX$n.var)

ttpc = wilcox.test( subset(simX, Status == 'NP')$clusters, subset(simX, Status == 'P')$clusters )$p.value

ggplot(melt(data.frame(simX[c('Status','clusters')])), aes(x=Status, y=value, fill=Status)) +  
    geom_boxplot() + geom_jitter(shape=16, position=position_jitter(0.2)) +
    stat_boxplot(geom ='errorbar') + labs(y='num. clusters', x='', title='Excluding HGD/IMC', subtitle=paste('p-value', signif(ttpc, 3)))

```

# Validation Set

```{r echo=F, warning=F, message=F}

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}

validation.patient.info =  subset(all.patient.info, Set == 'Test')
validation.patient.info = arrange(validation.patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)

cleaned = list.files(path=paste(data, 'Cleaned', sep='/'), pattern='tiled', full.names=T, recursive=T)
cleaned = grep(paste(unique(validation.patient.info$Hospital.Research.ID), collapse='|'), cleaned, value=T)
cleaned = grep('arms', cleaned, invert=T, value=T)

#raw = list.files(path=paste(data, 'Cleaned', sep='/'), pattern='raw', full.names=T, recursive=T)
#raw = grep(paste(unique(validation.patient.info$Hospital.Research.ID), collapse='|'), raw, value=T)

fitV = models[[0.9]]
lambda.opt = performance.at.1se[[0.9]][, 'lambda']

validation.df = matrix(nrow=0, ncol=ncol(dysplasia.df), dimnames=list(c(), colnames(dysplasia.df)))
for (fi in cleaned) {
  x = read.table(fi, sep='\t', header=T)
  segs = paste(x$chr, ':', x$start, '-', x$end, sep='')

  x = t(x[,-c(1:3)])
  colnames(x) = segs

  validation.df = rbind(validation.df, x)  
}

validation.df = validation.df[validation.patient.info$Samplename, ]
validation.df = apply(validation.df, 2, unit.var)

pred = predict(fitV, newx=validation.df, s=lambda.opt, type='response')
or = predict(fitV, newx=validation.df, s=lambda.opt, type='link')

vpd = validation.patient.info

if (rownames(pred) != vpd$Samplename)
  stop('Fix this')

vpd$Prediction = pred
vpd$OR = or

table(subset(vpd, Prediction > cutoff)$Status)
table(subset(vpd, Prediction < cutoff)$Status)

```

## Technical Replicates

Predicting the replicates that were left out of the model building process.  Though maybe they shouldn't have been!

Anyhow...

```{r echo=F, warning=F, message=F}

#normals = subset(read_xlsx(patient.file, sheet='All combined'), Pathology == 'normal')
repeats = read_xlsx(patient.file, sheet='Technical Repeats')

samples = 
  c(gsub('-', '_', paste(repeats$`Plate Index`, sub('SLX-','', repeats$`SLX ID`), sep='_')))
  #gsub('-', '_', paste(normals$`Plate Index`, sub('SLX\\s?-\\s?', '',  normals$`SLX ID 1`), sep='_')))
samples = sub('10729', '10725_10729', samples)

repeats = repeats[,c(1:6,8,9)]
repeats$Samplename = samples

cleaned = list.files(path=paste(data, 'Cleaned', sep='/'), pattern='tiled', full.names=T, recursive=T)
cleaned = grep(paste(unique(repeats$`Hospital Research ID`), collapse='|'), cleaned, value=T)
cleaned = grep('arms', cleaned, invert=T, value=T)

repeat.df = matrix(nrow=0, ncol=ncol(dysplasia.df), dimnames=list(c(), colnames(dysplasia.df)))
for (fi in cleaned) {
  x = read.table(fi, sep='\t', header=T)
  segs = paste(x$chr, ':', x$start, '-', x$end, sep='')

  x = t(x[,-c(1:3)])
  colnames(x) = segs

  repeat.df = rbind(repeat.df, x)  
}

rows = which(rownames(repeat.df) %in% samples)
repeat.df = repeat.df[rows,]

repeat.df = apply(repeat.df, 2, unit.var)

pred = predict(fitV, newx=repeat.df, s=lambda.opt, type='response')
or = predict(fitV, newx=repeat.df, s=lambda.opt, type='link')

repeats$Probability = pred[repeats$Samplename,]
repeats$OR = or[repeats$Samplename,]

```

`r pander(repeats)`

### 2018-05-02 Patient Report.Rmd
Moved shiny app to own repo

@@
 
-```{r echo=F, message=F, warning=F}
-
+```{r echo=F, message=F, warning=F, fig.height=5, fig.width=5}
 df = rbind(cbind(do.call(rbind.data.frame, var.seg.lengths[prog.pt]), 'status'='Progressor'), cbind(do.call(rbind.data.frame, var.seg.lengths[np.pt]), 'status'='Non-Progressor'))
@@
 m = (melt(df, id.vars = 'status', measure.vars = 'seg.len'))
-m2 = m %>% group_by(status) %>% summarise( median=median(value), n=length(status), value=max(value) )
+m2 = m %>% group_by(status) %>% dplyr::summarise( median=median(value), n=length(status), value=max(value) )
 ggplot(m, aes(status, value, group=status, fill=status)) + geom_jitter(aes(color=status), width=.2) + geom_violin(alpha=0.5) +
@@
 
-```{r, echo=F, warning=F, message=F}
+```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=5}
 df = rbind(cbind.data.frame('n'=sapply(var.seg.lengths[prog.pt], nrow), 'status'='Progressor'), cbind.data.frame('n'=sapply(var.seg.lengths[np.pt], nrow), 'status'='Non-Progressor'))
@@
   geom_violin(aes(fill=status, group=status), alpha=0.5) +
-  geom_label(data=(df %>% group_by(status) %>% summarise(y=max(n), med=median(n))), 
+  geom_label(data=(df %>% group_by(status) %>% dplyr::summarise(y=max(n), med=median(n))), 
              aes(status, y, label=paste('median\n',sprintf("%.1f", med, sep='')))) +
-  labs(y='Number of segments', title="Number of segments per patient", subtitle=paste('t-test p-value', signif(nWT$p.value, digits=3)) ) +
-  theme(legend.position = 'none')
+  labs(y='Number of segments', title="Number of segments per patient", subtitle=paste('t-test p-value', signif(nWT$p.value, digits=3)) ) + theme(legend.position = 'none')
+```
 
-nohgd.var.seg.lengths = lapply(patient.data, function(pt) {
-  samples = subset(pt$info, Pathology %in% c('BE','ID','LGD'))$Samplename
-  cols = c('chrom','start.pos','end.pos', grep(paste(samples, collapse='|'), colnames(pt$variable.regions), value=T))
-  vr = pt$variable.regions[,cols]
+## Excluding HGD/IMC
 
+As a consequence of the multipcf segmentation method that jointly segments across all patient samples excluding HGD/IMC samples at this point in the analysis makes no difference in the number/length of segments considered "variable".
+
+```{r echo=F,fig.height=5, fig.width=5}
+`%nin%` = Negate('%in%')
+nohgd.var.regions = lapply(patient.data, function(pt) {
+  #print(unique(pt$info$Hospital.Research.ID))
+  nohgd = subset(pt$info, Pathology %nin% c('HGD','IMC'))$Samplename
+  if (length(nohgd) > 1) {
+    segvals = pt$seg.vals[,c(1:5,grep(paste(nohgd, collapse='|'), colnames(pt$seg.vals)))]
+
+    sds = apply(segvals[,-c(1:5)],1,sd)
+
+    variable.region.indices = which(sds >= sd.threshold & segvals$n.probes >= min.probes)
+    variable.regions = segvals[ variable.region.indices,  ]  
+  } else {
+    variable.regions = pt$variable.regions
+  }
+  return(variable.regions)
+})
+
+nohgd.var.seg.lengths = lapply(nohgd.var.regions, function(vr) {
   cbind('chr'=vr$chrom, 'start'=vr$start.pos, 'end'=vr$end.pos, 'seg.len'=vr$end.pos - vr$start.pos)
 })
+
+df = rbind(cbind(do.call(rbind.data.frame, var.seg.lengths[prog.pt]), 'status'='Progressor'), cbind(do.call(rbind.data.frame, var.seg.lengths[np.pt]), 'status'='Non-Progressor'))
+lenWT = wilcox.test(subset(df, status == 'Progressor')$seg.len, subset(df, status == 'Non-Progressor')$seg.len)
+
+m = (melt(df, id.vars = 'status', measure.vars = 'seg.len'))
+m2 = m %>% group_by(status) %>% dplyr::summarise( median=median(value), n=length(status), value=max(value) )
+ggplot(m, aes(status, value, group=status, fill=status)) + geom_jitter(aes(color=status), width=.2) + geom_violin(alpha=0.5) +
+  geom_label(data=m2, aes(status, value, label=paste("median=",scientific(median,3),'\nn=',n, sep='')), fill='white') + 
+  theme(legend.position = 'none') + labs(y='Variable segment length', x='', title='Length of variable regions excluding HGD/IMC', subtitle=paste('Wilcox p-value', signif(lenWT$p.value, digits=3) ))
+```
+
+```{r echo=F,fig.height=5, fig.width=5}
+df = rbind(cbind.data.frame('n'=unlist(sapply(nohgd.var.regions[prog.pt], nrow)), 'status'='Progressor'), cbind.data.frame('n'=unlist(sapply(nohgd.var.regions[np.pt], nrow)), 'status'='Non-Progressor'))
+nWT = t.test(subset(df, status == 'Progressor')$n, subset(df, status == 'Non-Progressor')$n)
+
+ggplot(df, aes(status, n, group=status)) + geom_jitter(aes(color=status), width=0.2) +
+  geom_violin(aes(fill=status, group=status), alpha=0.5) +
+  geom_label(data=(df %>% group_by(status) %>% dplyr::summarise(y=max(n), med=median(n))), 
+             aes(status, y, label=paste('median\n',sprintf("%.1f", med, sep='')))) +
+  labs(y='Number of segments', title="Number of segments per patient, excluding HGD/IMC", subtitle=paste('t-test p-value', signif(nWT$p.value, digits=3)) ) + theme(legend.position = 'none')
 
@@
 ```{r echo=F, message=F, warning=F, fig.height=8, fig.width=20}
+
 n.var.regions = sapply(var.seg.lengths, nrow)
@@
 
-ggplot(melted, aes(Status, value, fill=Status)) + geom_jitter(aes(color=Status), width=0.2) + geom_boxplot(outlier.colour = NA) + 
-  geom_label(data=subset(melted, Status == 'P'), aes(x=Status, y=max(value)-sd(value), label=scientific(median(value),3) ), color='white', show.legend = F ) +
-  geom_label(data=subset(melted, Status == 'NP'), aes(x=Status, y=max(value)-sd(value), label=scientific(median(value),3) ), color='white' , show.legend = F) +
-  labs(title="Variable region length",subtitle=paste("p-value", signif(chr.ratio.tt$p.value, digits=3)), y='Segment Length/Chr length')
-
-ggplot(melted, aes(Status, value, fill=Status)) + geom_jitter(aes(color=Status), width=0.2) + geom_violin() + 
-  geom_label(data=subset(melted, Status == 'P'), aes(x=Status, y=max(value)-sd(value), label=scientific(median(value),3) ), color='white' , show.legend = F) +
-  geom_label(data=subset(melted, Status == 'NP'), aes(x=Status, y=max(value)-sd(value), label=scientific(median(value),3) ), color='white' , show.legend = F) +
-  labs(title="Variable region length",subtitle=paste("p-value", signif(chr.ratio.tt$p.value, digits=3)), y='Segment Length/Chr length')
+#ggplot(melted, aes(Status, value, fill=Status)) + geom_jitter(aes(color=Status), width=0.2) + geom_violin() + 
+#  geom_label(data=subset(melted, Status == 'P'), aes(x=Status, y=max(value)-sd(value), label=scientific(median(value),3) ), color='white' , show.legend = F) +
+#  geom_label(data=subset(melted, Status == 'NP'), aes(x=Status, y=max(value)-sd(value), label=scientific(median(value),3) ), color='white' , show.legend = F) +
+#  labs(title="Variable region length adjusted for chromosome length",subtitle=paste("p-value", signif(chr.ratio.tt$p.value, digits=3)), y='Segment Length/Chr length')
 

### 2018-05-02 features-OAC.Rmd
Moved shiny app to own repo

@@
 library(gridExtra)
-library(plyr)
 library(ggplot2)
@@
 library(readxl)
+library(plyr)
+library(dplyr)
 
@@
 ## Lists the sequenced patients
-sequenced = read_excel('~/Documents/OneDrive/OneDrive - MRC Cancer Unit at University of Cambridge/skillcoyne/OC Data/WGS_Phase 1 complete - cleaned.xlsx', trim_ws=T)
-sequenced = subset(sequenced, grepl('^OC', OCCAMS_ID)  )[,c(1:8)]
+sequenced = read_xlsx('~/Documents/OneDrive/OneDrive - MRC Cancer Unit at University of Cambridge/skillcoyne/Documents/Sequencing Batch Info/occams_sequenced_20180426.xlsx')
+sequenced = subset(sequenced, grepl('^OC', OCCAMS_ID) )
 sequenced$Sample_type = as.factor(tolower(sequenced$Sample_type))
@@
 
-# OCCAMS ID to Illumina ID
-paired = read.table('~/Documents/OneDrive/MRC Cancer Unit at University of Cambridge/icgc group - icgc_shared/icgc-group/batch_progress/reports/g37/PairedOverviewSummary_g37_bwa.txt', header=T, sep='\t', stringsAsFactors = F)
-paired = subset(paired, aligned == '*')
-pairedg37 = subset(paired, OCCAMS %in% unique(sequenced$OCCAMS_ID), select=c('OCCAMS','SampleId'))
-length(unique(pairedg37$OCCAMS))
-
-paired = read.table('~/Documents/OneDrive/MRC Cancer Unit at University of Cambridge/icgc group - icgc_shared/icgc-group/batch_progress/reports/g1k//PairedOverviewSummary_g1k_bwa.txt', header=T, sep='\t', stringsAsFactors = F)
-paired = subset(paired, aligned == '*')
-pairedg1k = subset(paired, OCCAMS %in% unique(sequenced$OCCAMS_ID), select=c('OCCAMS','SampleId'))
-
-paired = rbind(pairedg37, pairedg1k)
-paired = unique(paired)
-
-length(unique(paired$OCCAMS))
-
-paired[] = lapply(paired[], strip.whitespace)
-length(unique(paired$OCCAMS))
-
-sp = do.call(rbind, lapply(paired$SampleId, function(x) unlist(strsplit(x, '_vs_'))))
-colnames(sp) = c('tb.sample', 'normal')
-
-paired = cbind(paired, sp)
-
-barretts.pairs = subset(paired, tb.sample %in% be.seq$Illumina_Barcode)
-
-paired = paired[-which(paired$tb.sample %in% be.seq$Illumina_Barcode),]
+# # OCCAMS ID to Illumina ID
+# paired = read.table('~/Documents/OneDrive/MRC Cancer Unit at University of Cambridge/icgc group - icgc_shared/icgc-group/batch_progress/reports/g37/PairedOverviewSummary_g37_bwa.txt', header=T, sep='\t', stringsAsFactors = F)
+# paired = subset(paired, aligned == '*')
+# pairedg37 = subset(paired, OCCAMS %in% unique(sequenced$OCCAMS_ID), select=c('OCCAMS','SampleId'))
+# length(unique(pairedg37$OCCAMS))
+# 
+# paired = read.table('~/Documents/OneDrive/MRC Cancer Unit at University of Cambridge/icgc group - icgc_shared/icgc-group/batch_progress/reports/g1k//PairedOverviewSummary_g1k_bwa.txt', header=T, sep='\t', stringsAsFactors = F)
+# paired = subset(paired, aligned == '*')
+# pairedg1k = subset(paired, OCCAMS %in% unique(sequenced$OCCAMS_ID), select=c('OCCAMS','SampleId'))
+# 
+# paired = rbind(pairedg37, pairedg1k)
+# paired = unique(paired)
+# 
+# length(unique(paired$OCCAMS))
+# 
+# paired[] = lapply(paired[], strip.whitespace)
+# length(unique(paired$OCCAMS))
+# 
+# sp = do.call(rbind, lapply(paired$SampleId, function(x) unlist(strsplit(x, '_vs_'))))
+# colnames(sp) = c('tb.sample', 'normal')
+# 
+# paired = cbind(paired, sp)
+# 
+# barretts.pairs = subset(paired, tb.sample %in% be.seq$Illumina_Barcode)
+# 
+# paired = paired[-which(paired$tb.sample %in% be.seq$Illumina_Barcode),]
 
@@
 # Clinical data
-load('~/Data/ICGC/OAC/clinical/labkey_20172706.Rdata', verbose=T)
+load('~/Data/ICGC/OAC/clinical/labkey_20180111.Rdata', verbose=T)
 demo = occams$patients
@@
 demo[rownames(ba[changed,]), 'RP.BarrettsAdjacent'] = ba[changed,'caitrona']
+
 # ------------------------------ #
 
-# Copy number data - tumour
-read.caveman<-function(s) {
-  files = sample.list[[s]]
-  if (length(files) < 2) {
-    warning(s)
-    stop()
-  }
-  stats = read.table(grep('samplestatistics.csv', files, value=T), row.names=1)
-
-  caveman = read.csv(grep('copynumber.caveman.csv$', files, value=T))
-  caveman$ploidy = stats['Ploidy',]
-  caveman$rho = stats['rho',]
-  caveman$contamination = stats['NormalContamination',]
-
-  caveman$Illumina.Barcode.full = s
-  caveman
-}
-
 # Tumours
-dir = '~/Data/ICGC/BE_OAC_Cohort/tumour/data/ascat'
-files = list.files(dir, full.names=T, recursive=T)
-samples = list.files(dir, full.names=F)
-
-sample.list = lapply(samples, function(x) {
-  grep(x, files, value=T)
-} )
-names(sample.list) = samples
-
-cnv = do.call(rbind, lapply(samples, read.caveman))
-totalCNV = merge(cnv, paired, by.x='Illumina.Barcode.full', by.y='SampleId')
-
-demo.cols = c('StudySubjectID','RP.TStage.PrimaryTumour','RP.BarrettsAdjacent','Weeks.Survival','RP.TumourDifferentiation','DI.ageAtDiagnosis','DI.PatientGender','FE.Patient.Died')
-totalCNV = merge(totalCNV, demo[,demo.cols], by.y='StudySubjectID', by.x='OCCAMS')
-
-oacCNV = makeGRangesFromDataFrame(totalCNV, keep.extra.columns = T)
+# there are some duplicates in the patient samples so I've selected one tumor for each of those. That means that there are a few less samples to be analysed than the files initially might suggest
+tumors = subset(sequenced, Sample_type == 'tumour')
+
+cnv = read.table('~/Data/ICGC/BE_OAC_Cohort/tumour/data/ascat/totalCNV.txt', sep='\t', header=T, stringsAsFactors=F)
+
+barcodes = do.call(rbind, strsplit(cnv$Illumina.Barcode, '_vs_'))
+colnames(barcodes) = c('Tumor.Barcode','Normal.Barcode')
+
+cnv = cbind(cnv, barcodes)
+cnv[colnames(barcodes)] = lapply(cnv[colnames(barcodes)], as.character)
+length(unique(cnv$Tumor.Barcode))
+
+#subset(tumors, Illumina_Barcode %in% setdiff(cnv$Tumor.Barcode, tumors$Illumina_Barcode))
+
+cnv = base::merge(cnv, tumors, by.x='Tumor.Barcode', by.y='Illumina_Barcode')
+
+tumorcnv = makeGRangesFromDataFrame(cnv, keep.extra.columns = T)
+
+# Not a problem at the moment
+#which(!cnv$Tumor.Barcode %in% tumors$Illumina_Barcode)
 
 # Barretts ADJACENT SAMPLES
-dir = '~/Data/ICGC/BE_OAC_Cohort/BA/data/ascat'
-files = list.files(dir, full.names=T, recursive=T)
-samples = list.files(dir, full.names=F)
-
-sample.list = lapply(samples, function(x) {
-  grep(x, files, value=T)
-} )
-names(sample.list) = samples
-
-cnv = do.call(rbind, lapply(samples, read.caveman))
-beCNV = merge(cnv, barretts.pairs, by.x='Illumina.Barcode.full', by.y='SampleId')
-beCNV = merge(beCNV, demo[,demo.cols], by.y='StudySubjectID', by.x='OCCAMS')
-beCNV = makeGRangesFromDataFrame(beCNV, keep.extra.columns = T)
+becnv = read.table('~/Data/ICGC/BE_OAC_Cohort/BA/data/ascat/totalCNV.txt', sep='\t', header=T, stringsAsFactors=F)
+length(unique(becnv$Illumina.Barcode))
+
+barcodes = do.call(rbind, strsplit(becnv$Illumina.Barcode, '_vs_'))
+colnames(barcodes) = c('Tumor.Barcode','Normal.Barcode')
+
+becnv = cbind(becnv, barcodes)
+
+becnv[colnames(barcodes)] = lapply(becnv[colnames(barcodes)], as.character)
+
+becnv = base::merge(becnv, subset(sequenced, grepl('barr', Sample_type)) , by.x='Tumor.Barcode', by.y='Illumina_Barcode') 
+#becnv = merge(becnv, demo[,cols], by.y='StudySubjectID', by.x='OCCAMS')
+becnv = makeGRangesFromDataFrame(becnv, keep.extra.columns = T)
 # ------------------------------ #
 
-oac.demo = subset(demo, StudySubjectID %in% unique(oacCNV$OCCAMS))
-be.demo = subset(demo, StudySubjectID %in% unique(beCNV$OCCAMS))
+#oac.demo = subset(demo, StudySubjectID %in% unique(tumorcnv$OCCAMS))
+#be.demo = subset(demo, StudySubjectID %in% unique(becnv$OCCAMS))
 
 ## Shallow WGS data 
-load('~/Data/Ellie/Analysis/5e06/loo.Rdata', verbose=T)
+load('~/Data/Ellie/Analysis/5e6_arms_all_exAHM0320/loo.Rdata', verbose=T)
 swgsPredictions = pg.samp
-rm(pg.samp)
-
-load('~/Data/Ellie/Analysis/5e06/all.pt.alpha.Rdata', verbose=T)
+rm(pg.samp, plots, coefs, fits)
+
+load('~/Data/Ellie/Analysis/5e6_arms_all_exAHM0320/all.pt.alpha.Rdata', verbose=T)
 rawSWGSdf = dysplasia.df
@@
 
+# CV(RR)
 hazards = apply( exp(t(rawSWGSdf[, rownames(coefs)])*coefs[,1]), 1, function(x) { sd(x)/mean(x) })
-rm(dysplasia.df,fits,plots,performance.at.1se)
+rm(dysplasia.df,models,plots,performance.at.1se)
 # ------------------------------ #
@@
 
-analysis.files = list.files('~/Data/Ellie/Analysis', full.names=T)
-load(grep('Training_patients.Rdata', analysis.files, value=T), verbose=T)
-
-labels = unlist(sapply(patient.data, function(df) {
-  df$info = arrange(df$info, Endoscopy.Year, Pathology)
-  label = as.integer(df$info$Status == 'P') #as.integer(df$info$Pathology %in% c('HGD', 'IMC'))
-  names(label) = df$info$Samplename
-  return(label)
-}))
-names(labels) = sub('.*\\.', '',  names(labels))
-
-labels = labels[which(names(labels) %in% rownames(rawSWGSdf))]
-
-rm(patient.data)
 ```
@@
 Load the coefficients from the selected model and merge with the copynumber regions identified in (some) of the OAC patients
-```{r, echo=F}
+```{r, echo=F, warning=F, message=F}
 regions = do.call(rbind.data.frame, c(strsplit(grep('\\d+', ch$label, value=T), ':|-'), stringsAsFactors=F))
@@
 
-```{r}
+```{r, echo=T}
 # positive coefficients were related to gains in progressors
@@
 
+# counts of gains per coef
 gain = do.call(rbind, lapply(gains$label, function(reg) {
@@
 }))
-
-t.test(gain[,1], gain[,2])
+rownames(gain) = gains$label
+wilcox.test(gain[,1], gain[,2])
 
@@
 }))
-
-t.test(loss[,1], loss[,2])
-
+rownames(loss) = losses$label
+wilcox.test(loss[,1], loss[,2])
 ```
@@
 }))
-
-t.test(randomGains[,1], randomGains[,2])
+wilcox.test(randomGains[,1], randomGains[,2])
 
@@
 }))
-
-t.test(randomLosses[,1], randomLosses[,2])
-
-```
-
-However, while there appears to be no difference in the gains found in NP for randomly selected regions vs the coefficients:
-
-pander(wilcox.test(gain[,'0'], randomGains[,'0']), caption="Non progressors, coeffiecents vs randomly selected")
-
-pander(wilcox.test(loss[,'0'], randomLosses[,'0']), caption="Non progressors, coeffiecents vs randomly selected")
-
-There is a significant difference in the selected regions for progressors. Again this is mostly a sanity check:
-
-pander(wilcox.test(gain[,'1'], randomGains[,'1']), caption="Progressors, coeffiecents vs randomly selected gains")
-
-pander(wilcox.test(loss[,'1'], randomLosses[,'1']), caption="Progressors, coeffiecents vs randomly selected losses")
+wilcox.test(randomLosses[,1], randomLosses[,2])
+```
+
+```{r echo=F}
+rg = wilcox.test(gain[,'0'], randomGains[,'0'])
+rl = wilcox.test(loss[,'0'], randomLosses[,'0'])
+```
+There appears to be no difference in gains (P=`r rg$p.value`) or losses (P=`r rl$p.value`) for Non-progressors as compared to random.
+
+```{r echo=F}
+rg = wilcox.test(gain[,'1'], randomGains[,'1'])
+rl = wilcox.test(loss[,'1'], randomLosses[,'1'])
+```
+However, there is a significant difference in the selected regions for progressors as compared to random (gains P=`r rg$p.value` and losses P=`r rl$p.value`). This is mostly a sanity check.
 
@@
   df = lapply(df, function(x) {
-    x = arrange(x, -width)
+    x = plyr::arrange(x, -width)
     x$order = factor(1:nrow(x))
@@
   m = do.call(rbind.data.frame, df)
-  m = m %>% group_by(seqnames) %>% ungroup() %>% arrange(segment, seqnames, width, start) %>% mutate(segorder = row_number())
+  m = m %>% group_by(seqnames) %>% dplyr::ungroup() %>% 
+    dplyr::arrange(segment, seqnames, width, start) %>% dplyr::mutate(segorder = row_number()) 
   m$isWGD = as.factor(m$isWGD)
@@
 # WGD
-oacCNV$isWGD = oacCNV$ploidy > 3.2 
-beCNV$isWGD = beCNV$ploidy > 3.2
+tumorcnv$isWGD = tumorcnv$ploidy > 3.2 
+becnv$isWGD = becnv$ploidy > 3.2
 
@@
 ## gains in patients -- issues with WGD samples?
-gains = lapply(swgRegions[gr], gain.for, cnv=oacCNV)
+gains = lapply(as(swgRegions[gr], "GRangesList"), gain.for, cnv=tumorcnv)
 names(gains) = as.character(swgRegions[gr])
@@
 
-be.gains = lapply(swgRegions[gr], gain.for, cnv=oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'yes')])
+sequenced = base::merge(sequenced, demo, by.x='OCCAMS_ID', by.y='StudySubjectID')
+subset(sequenced, RP.BarrettsAdjacent == 'yes')$Illumina_Barcode
+
+be.cnvs = subset(tumorcnv, Tumor.Barcode %in%  subset(sequenced, RP.BarrettsAdjacent == 'yes')$Illumina_Barcode)
+non.be.cnvs = subset(tumorcnv, Tumor.Barcode %in%  subset(sequenced, RP.BarrettsAdjacent == 'no')$Illumina_Barcode)
+
+be.gains = lapply(as(swgRegions[gr], 'GRangesList'), gain.for, cnv=be.cnvs)
 names(be.gains) = as.character(swgRegions[gr])
-plot.gains(be.gains[sapply(be.gains, nrow) > 0]) + labs(title='Gains in BE Adjacent samples')
-
-non.be.gains = lapply(swgRegions[gr], gain.for, cnv=oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'no')])
+plot.gains(be.gains[sapply(be.gains, nrow) > 0]) + labs(title='Gains in BE Adjacent OAC samples')
+
+non.be.gains = lapply(as(swgRegions[gr], 'GRangesList'), gain.for, cnv=non.be.cnvs)
 names(non.be.gains) = as.character(swgRegions[gr])
-plot.gains(non.be.gains[sapply(non.be.gains, nrow) > 0]) + labs(title='Gains in NO BE Adjacent samples')
+plot.gains(non.be.gains[sapply(non.be.gains, nrow) > 0]) + labs(title='Gains in Non-BE Adjacent OAC samples')
 
@@
 # Per patient
-gfreq = sort(sapply(gains, function(x) length(unique(x$OCCAMS)))/length(unique(sequenced$OCCAMS_ID)))
+gfreq = sort(sapply(gains, function(x) length(unique(x$Illumina.Barcode)))/length(unique(sequenced$OCCAMS_ID)))
 ggplot(melt(as.matrix(gfreq)), aes(Var1, value)) + geom_col(fill='darkblue') + ylim(0,0.5) + coord_flip() + 
@@
 
-gfreq1 = sort(sapply(be.gains, function(x) length(unique(x$OCCAMS)))/length(unique(oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'yes')]$OCCAMS)))
+gfreq1 = sort(sapply(be.gains, function(x) length(unique(x$Illumina.Barcode)))/length(unique(be.cnvs$Illumina.Barcode)))
 gfreq1 = gfreq1[names(gfreq)]
 
-gfreq2 = sort(sapply(non.be.gains, function(x) length(unique(x$OCCAMS)))/length(unique(oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'no')]$OCCAMS)))
+gfreq2 = sort(sapply(non.be.gains, function(x) length(unique(x$Illumina.Barcode)))/length(unique(non.be.cnvs$Illumina.Barcode)))
 gfreq2 = gfreq2[names(gfreq)]
@@
       if(nrow(x) <= 0) return(NULL)
-      x = arrange(x, -width)
+      x = plyr::arrange(x, -width)
       x$order = factor(1:nrow(x))
@@
     m = df
-    m = arrange(m, -width)
+    m = plyr::arrange(m, -width)
     m$order = factor(1:nrow(m))
@@
 
-  m = m %>% group_by(seqnames) %>% ungroup() %>% arrange(segment, seqnames, width, start) %>% mutate(segorder = row_number())
+  m = m %>% dplyr::group_by(seqnames) %>% ungroup() %>% dplyr::arrange(segment, seqnames, width, start) %>% dplyr::mutate(segorder = row_number())
   m$isWGD = as.factor(m$isWGD)
@@
 
-losses = lapply(swgRegions[lr], loss.for, oacCNV)
+losses = lapply(as(swgRegions[lr], 'GRangesList'), loss.for, tumorcnv)
 names(losses) = as.character(swgRegions[lr])
@@
 
-be.losses = lapply(swgRegions[lr], loss.for, oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'yes')])
+be.losses = lapply(as(swgRegions[lr], 'GRangesList'), loss.for, be.cnvs)
 names(be.losses) = as.character(swgRegions[lr])
@@
 
-non.be.losses = lapply(swgRegions[lr], loss.for, cnv=oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'no')])
+non.be.losses = lapply(as(swgRegions[lr], 'GRangesList'), loss.for, non.be.cnvs)
 names(non.be.losses) = as.character(swgRegions[lr])
@@
 
-lfreq = sort(sapply(losses, function(x) length(unique(x$OCCAMS)))/length(unique(sequenced$OCCAMS_ID)))
+lfreq = sort(sapply(losses, function(x) length(unique(x$Illumina.Barcode)))/length(unique(sequenced$OCCAMS_ID)))
 ggplot(melt(as.matrix(lfreq)), aes(Var1, value)) + geom_col(fill='darkgreen') + ylim(0,1) + coord_flip() + 
@@
 
-lfreq1 = sapply(be.losses, function(x) length(unique(x$OCCAMS)))/length(unique(oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'yes')]$OCCAMS))
+lfreq1 = sapply(be.losses, function(x) length(unique(x$Illumina.Barcode)))/length(unique(be.cnvs$Illumina.Barcode))
 lfreq1 = lfreq1[names(lfreq)]
 
-lfreq2 = sort(sapply(non.be.losses, function(x) length(unique(x$OCCAMS)))/length(unique(oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'no')]$OCCAMS)))
+lfreq2 = sort(sapply(non.be.losses, function(x) length(unique(x$Illumina.Barcode)))/length(unique(non.be.cnvs$Illumina.Barcode)))
 lfreq2 = lfreq2[names(lfreq)]
@@
 
+TODO: This should really separate the focal losses in 17p from the large-scale losses.
+
 ```{r fig.height=6,fig.width=8}
-
-## Note it's a subset of the 321 patients (66) that have this
 chr17 = get.chr.lengths()[17,]
 
-#chr17losses = lapply(makeGRangesFromDataFrame( cbind.data.frame('start'=0, 'end'=chr17$chr.length, 'chr'='17') ), loss.for)[[1]]
-#m = arrange(chr17losses, width, start)
-
-plot.losses(losses$`17:1-4776188`) + geom_hline(data=chr17, aes(yintercept=(chr.cent-cent.gap)), color='red', linetype='dashed') +
-  labs(y='chr17 position', x='Patient', title='Chr17 Losses in OAC Patients', subtitle=paste(round(length(unique(losses$`17:1-4776188`$OCCAMS))/length(unique(oacCNV$OCCAMS)), 2)*100, '% of OAC patients', sep=''))
-
-```
-
-So is there a difference between the gains/losses in the selected regions?  
+chr17losses = losses[grep('^17:', names(losses))][[1]]
+
+plot.losses(chr17losses) + geom_hline(data=chr17, aes(yintercept=(chr.cent-cent.gap)), color='red', linetype='dashed') +
+  labs(y='chr17 position', x='Patient', title='17p Losses in (all) OAC Patients', subtitle=paste(round(length(unique(chr17losses$Illumina.Barcode))/length(unique(tumorcnv$Illumina.Barcode)), 2)*100, '% of OAC patients', sep=''))
+
+```
+
+There does not appear to be a difference in gains vs losses.
 ```{r}
@@
 
-These may be more nuanced, as we need to be aware of gains that reflect WGD vs gains that are new amplifications. At the moment there's no distinction
-
-```{r}
-randomGains = lapply(sample.regions, gain.for, cnv=oacCNV)
+These may be more nuanced, as we need to be aware of gains that reflect WGD vs gains that are new amplifications. At the moment there's no distinction.
+
+There's no clear difference between the randomly selected regions and gains overall.
+But there does appear to be a difference when comparing the randomly selected regions to the regions split by BE/non-BE.
+I'm not sure this is real considering the overall case though.
+
+```{r, echo=F}
+randomGains = lapply(as(sample.regions,'GRangesList'), gain.for, cnv=tumorcnv)
 names(randomGains) = as.character(sample.regions)
+randomGains = randomGains[which(sapply(randomGains, nrow) > 2)]
 
@@
 # Maybe a difference in # gains in the selected features vs random
-wilcox.test(sapply(randomGains, nrow), sapply(gains, nrow))
-
-wilcox.test(sapply(randomGains, nrow), sapply(be.gains,nrow))
-wilcox.test(sapply(randomGains, nrow), sapply(non.be.gains,nrow))
-
-# Not a difference here though
-wilcox.test(sapply(gains, function(df) length(unique(df$OCCAMS))), 
-            sapply(randomGains, function(df) length(unique(df$OCCAMS))))
-
-df = rbind(cbind.data.frame('counts'=sapply(gains, function(df) length(unique(df$OCCAMS))), 'type'='selected') , 
-      cbind.data.frame('counts'=sapply(randomGains, function(df) length(unique(df$OCCAMS))), 'type'='random'))
+# wilcox.test(sapply(randomGains, nrow), sapply(gains, nrow))
+# 
+# tests = matrix(ncol=2, nrow=10, dimnames=list(c(1:10), c('Rand.vs.BE.Gain', 'Rand.v.Non.Be.Gain')))
+# for (i in 1:10) {
+#   
+#   sr = randomGains[sample(1:length(randomGains), length(be.gains))]
+#   wt1 = wilcox.test(sapply(sr, nrow), sapply(be.gains,nrow))
+#   
+#   sr = randomGains[sample(1:length(randomGains), length(non.be.gains))]
+#   wt2 = wilcox.test(sapply(sr, nrow), sapply(non.be.gains,nrow))
+#   
+#   tests[i,] = c(wt1$p.value, wt2$p.value)
+# }
+# tests < 0.005
+# 
+# #wilcox.test(sapply(randomGains, nrow), sapply(be.gains,nrow))
+# #wilcox.test(sapply(randomGains, nrow), sapply(non.be.gains,nrow))
+# 
+# # Not a difference here though
+# wilcox.test(sapply(gains, function(df) length(unique(df$Illumina.Barcode))), 
+#             sapply(randomGains, function(df) length(unique(df$Illumina.Barcode))))
+
+df = rbind(cbind.data.frame('counts'=sapply(gains, function(df) length(unique(df$Illumina.Barcode))), 'type'='selected') , 
+      cbind.data.frame('counts'=sapply(randomGains, function(df) length(unique(df$Illumina.Barcode))), 'type'='random'))
 ggplot(df, aes(type, counts, group=type, fill=type)) + geom_boxplot() + geom_jitter(width=0.3) + labs(title='Patient counts by region') + plot.theme
@@
 
-```{r}
-randomLosses = lapply(sample.regions, loss.for, cnv=oacCNV)
+```{r, echo=F}
+randomLosses = lapply(as(sample.regions,'GRangesList'), loss.for, cnv=tumorcnv)
 names(randomLosses) = as.character(sample.regions)
+
+randomLosses = randomLosses[which(sapply(randomLosses, nrow) > 2)]
 
@@
 
-# No difference in # losses in the selected features
-wilcox.test(sapply(randomLosses, nrow), sapply(losses, nrow))
-
-wilcox.test(sapply(randomLosses, nrow), sapply(be.losses,nrow))
-wilcox.test(sapply(randomLosses, nrow), sapply(non.be.losses,nrow))
+# # No difference in # losses in the selected features
+# wilcox.test(sapply(randomLosses, nrow), sapply(losses, nrow))
+# 
+# 
+# tests = matrix(ncol=2, nrow=10, dimnames=list(c(1:10), c('Rand.vs.BE.Loss', 'Rand.v.Non.Be.Loss')))
+# for (i in 1:10) {
+#   sr = randomLosses[sample(1:length(randomLosses), length(be.losses))]
+#   wt1 = wilcox.test(sapply(sr, nrow), sapply(be.losses,nrow))
+#   
+#   sr = randomLosses[sample(1:length(randomLosses), length(non.be.losses))]
+#   wt2 = wilcox.test(sapply(sr, nrow), sapply(non.be.losses,nrow))
+#   
+#   tests[i,] = c(wt1$p.value, wt2$p.value)
+# }
+# tests < 0.005
 
 # No difference between the two
-df = rbind(cbind.data.frame('counts'=sapply(losses, function(df) length(unique(df$OCCAMS))), 'type'='selected') , 
-      cbind.data.frame('counts'=sapply(randomLosses, function(df) length(unique(df$OCCAMS))), 'type'='random'))
+df = rbind(cbind.data.frame('counts'=sapply(losses, function(df) length(unique(df$Illumina.Barcode))), 'type'='selected') , 
+      cbind.data.frame('counts'=sapply(randomLosses, function(df) length(unique(df$Illumina.Barcode))), 'type'='random'))
 ggplot(df, aes(type, counts, group=type, fill=type)) + geom_boxplot() + geom_jitter(width=0.3) + labs(title='Loss counts by patient') + plot.theme
@@
 
-## Cox
-
-```{r}
-library(survival)
-library(GGally)
-
-pt.gains = do.call(rbind, lapply(oac.demo$StudySubjectID, function(id) {
-  rbind(sapply(gains, function(g)  nrow(subset(g, OCCAMS == id))))
-}))
-rownames(pt.gains) = oac.demo$StudySubjectID
-colnames(pt.gains) = paste('gain_', gsub(':|-', '\\.', colnames(pt.gains)), sep='')
-
-pt.losses = do.call(rbind, lapply(oac.demo$StudySubjectID, function(id) {
-  rbind(sapply(losses, function(g)  nrow(subset(g, OCCAMS == id))))
-}))
-rownames(pt.losses) = oac.demo$StudySubjectID
-colnames(pt.losses) = paste('loss_',gsub(':|-', '\\.', colnames(pt.losses)), sep='')
-
-x = merge(pt.gains, pt.losses, by='row.names', all=T)
-
-x = merge(x,oac.demo[,demo.cols], by.x='Row.names', by.y='StudySubjectID')
-x$Row.names = NULL
-
-time = 'Weeks.Survival';  event='FE.Patient.Died'; type = '"right"'
-x[c(time,event)] = lapply(x[c(time,event)], as.numeric)
-
-x$RP.TStage.PrimaryTumour = as.character(x$RP.TStage.PrimaryTumour)
-
-x[grepl('T1', x$RP.TStage.PrimaryTumour), 'RP.TStage.PrimaryTumour'] = 'T1'
-x[grepl('T4', x$RP.TStage.PrimaryTumour), 'RP.TStage.PrimaryTumour'] = 'T4'
-x[grepl('Tx', x$RP.TStage.PrimaryTumour), 'RP.TStage.PrimaryTumour'] = 'T0'
-
-terms = as.formula(paste("Surv(",time,",",event,", type=",type,") ~ RP.BarrettsAdjacent", sep=""))
-sfit = survfit(terms, data=x,conf.int=0.95, conf.type='log', na.action=na.omit)
-ggsurv(sfit)
-
-f1 = paste(demo.cols[c(3,5:7)], collapse='+')
-f2 = paste(paste('`', grep('^gain|loss', colnames(x), value=T), '`', sep=''), collapse='+')
-
-f = paste(f1,f2, sep='+')
-
-terms = as.formula(paste("Surv(",time,",",event,", type=",type,") ~", f, sep=""))
-cox = coxph(terms, data=x,na.action=na.omit,ties='breslow')
-cfit = survfit(cox)
-ggsurv(cfit) + labs(title=paste(length(grep('gain|loss', colnames(x))), 'copy number + demo'))
-
-pander(cox)
-
-anova(cox)
-```
-
-```{r}
-library(CoxHD)
-
-x = x[,c(grep('gain|loss',colnames(x), invert=T), grep('gain|loss',colnames(x)))]
-xs = x
-
-surv = with(xs, Surv(Weeks.Survival, FE.Patient.Died))
-ggsurv(survfit(surv~1,data=xs))
-
-xs = xs[,-grep('Weeks|Died',colnames(xs))]
-
-xs$RP.BarrettsAdjacent = as.integer(xs$RP.BarrettsAdjacent)-1
-xs$RP.TumourDifferentiation = as.integer(as.factor(xs$RP.TumourDifferentiation))-1
-xs$DI.PatientGender = as.integer(as.factor(xs$DI.PatientGender))-1
-
-groups = c(rep('clin', 4), rep('cn',(ncol(pt.gains)+ncol(pt.losses)) ))
-
-rfx = CoxRFX(xs, surv, groups, nu=1, verbose=T)
-summary(rfx)
-
-plot(rfx)
-
-VarianceComponents(rfx)
-
-WaldTest(rfx)
-
-PartialC(rfx)
-
-PlotVarianceComponents(rfx)
-
-```
-
 ## Focal vs Large CN
@@
 
-```{r, fig.width=8}
+```{r, echo=F, fig.width=8}
 size = unlist(sapply(gains, function(df) df$width))
@@
 
-# Focal changes 1% or the chromosome?
+# Focal changes < 1% or the chromosome?
 countBySize <-function(df) {
@@
 
+## Still in the regions selected
 sizeGainsByPatient = do.call(rbind.data.frame,lapply(gains, sizebyPatient))
@@
 x = which(p.adjust(apply(sizeGains, 1, function(x) binom.test(x[1], sum(x), 0.5)$p.value )) < 0.05)
+# CI
 apply(sizeGains, 1, function(x) binom.test(x[1], sum(x), 0.5)$conf.int )[,x]
 
-## Not sure what to do with this really
+## Not sure what to do with this really. Regions are shared, nonBE has two additional
 sizeGainsBE = do.call(rbind.data.frame, lapply(be.gains, countBySize))
@@
 
-## Not sure what to do with this really
+## Not sure what to do with this really. Again shared regions, two addition in the nonBE
 sizeGainsBEByPt = do.call(rbind.data.frame, lapply(be.gains, sizebyPatient))
@@
 
-# Clear difference between focal/large
-wilcox.test(sizeGainsByPatient$focal, sizeGainsByPatient$large)
-one.way.aov(sizeGainsByPatient$focal, sizeGainsByPatient$large)
-
-sizeGainsRatio = sizeGains %>% rowwise() %>% summarise( 'focal.r'=focal/sum(focal,large), 'large.r'=large/sum(focal,large) )
+## TODO Here we need to compare to randomly selected regions. I don't think you'd expect that regions with many focal would also show large  or vice versa
+sizeGainsRatio = sizeGains %>% rowwise() %>% dplyr::summarise( 'focal.r'=focal/sum(focal,large), 'large.r'=large/sum(focal,large) )
 sizeGainsRatio$feature = rownames(sizeGains)
-
-# No difference in the overall number of gains per region
-ggplot(melt(sizeGains), aes(variable, value, group=variable, fill=variable)) + 
-  geom_boxplot(alpha=0.5, outlier.colour = NA) + geom_jitter(width=0.2, aes(color=variable)) 
-  labs(title="Types of gain per region", subtitle=paste('p=',format.pval(one.way.aov(sizeGains$focal, sizeGains$large)$p.value,3), sep=''), y='patients per region', x='')
 
@@
 
-```{r, fig.width=8}
+```{r, echo=F, fig.width=8}
 sizeLossesByPatients = do.call(rbind.data.frame,lapply(losses, sizebyPatient))
@@
 # True for losses as well
-p.adjust(apply(sizeLosses, 1, function(x) binom.test(x[1], sum(x), 0.5)$p.value ))
+which(p.adjust(apply(sizeLosses, 1, function(x) binom.test(x[1], sum(x), 0.5)$p.value )) < 0.05)
 apply(sizeLosses, 1, function(x) binom.test(x[1], sum(x), 0.5)$conf.int )
 
-wilcox.test(sizeLosses$focal, sizeLosses$large)
-one.way.aov(sizeLosses$focal, sizeLosses$large)
-
+## Again, wouldn't expect a lot of focal losses if there are large ones already so need to think about what to test
 sizeLossRatio = sizeLosses %>% rowwise() %>% summarise( 'focal.r'=focal/sum(focal,large), 'large.r'=large/sum(focal,large) )
@@
 
-ggplot(melt(sizeLossRatio), aes(feature, value, fill=variable) ) + geom_col() + coord_flip() + labs(title='Overall losses')
+ggplot(melt(sizeLossRatio, id.vars='feature'), aes(feature, value, fill=variable) ) + geom_col() + coord_flip() + labs(title='Overall losses')
 
@@
 
-wilcox.test(sizeLossesByPatients$focal, sizeLossesByPatients$large)
 ```
@@
 #### Losses
-```{r, fig.width=8}
+
+Basically losses tend to be large (>1% of the chromosome) and patients are more likely to have large losses rather than focal ones
+
+```{r, echo=F, fig.width=8}
 randomLossesByPt = do.call(rbind.data.frame,lapply(randomLosses, sizebyPatient))
@@
 wilcox.test(randomLossesBySize$focal, sizeLosses$focal) 
-one.way.aov(randomLossesBySize$focal, sizeLosses$focal) 
-wilcox.test(randomLossesByPt$focal, sizeLossesByPatients$focal)
+#one.way.aov(randomLossesBySize$focal, sizeLosses$focal)
 
@@
 wilcox.test(randomLossesBySize$large, sizeLosses$large)
-one.way.aov(randomLossesBySize$large, sizeLosses$large)
-wilcox.test(randomLossesByPt$large, sizeLossesByPatients$large) 
+#one.way.aov(randomLossesBySize$large, sizeLosses$large)
 ```
@@
 
-```{r, fig.width=8}
+Looks like gains show no particular enrichment for focal vs large alterations.
+
+```{r, echo=F, fig.width=8}
 randomGainsByPt = do.call(rbind.data.frame,lapply(randomGains, sizebyPatient))
@@
 
-# But maybe a difference in random regions vs selected ones?
 wilcox.test(randomGainsBySize$focal, sizeGains$focal) 
 wilcox.test(randomGainsBySize$large, sizeGains$large) 
-
-wilcox.test(randomGainsByPt$focal, sizeGainsByPatient$focal) # maybe
-wilcox.test(randomGainsByPt$large, sizeGainsByPatient$large) # maybe
+wilcox.test(sizeGainsByPatient$focal, sizeGainsByPatient$large)
+
+```
+
+# Cox
+
+```{r}
+library(survival)
+library(GGally)
+
+oac.demo = subset(demo, StudySubjectID %in% sequenced$OCCAMS_ID)
+
+pt.gains = do.call(rbind, lapply(oac.demo$StudySubjectID, function(id) {
+  rbind(sapply(gains, function(g)  nrow(subset(g, OCCAMS_ID == id))))
+}))
+rownames(pt.gains) = oac.demo$StudySubjectID
+colnames(pt.gains) = paste('gain_', gsub(':|-', '\\.', colnames(pt.gains)), sep='')
+
+pt.losses = do.call(rbind, lapply(oac.demo$StudySubjectID, function(id) {
+  rbind(sapply(losses, function(g)  nrow(subset(g, OCCAMS_ID == id))))
+}))
+rownames(pt.losses) = oac.demo$StudySubjectID
+colnames(pt.losses) = paste('loss_',gsub(':|-', '\\.', colnames(pt.losses)), sep='')
+
+demo.cols = c('StudySubjectID','Weeks.Survival', 'FE.Patient.Died', 'RP.BarrettsAdjacent', 'DI.PatientGender', 'DI.ageAtDiagnosis', 'RP.TStage.PrimaryTumour', 'RP.TumourDifferentiation', 'RP.Tumour.Growth')
+
+x = merge(pt.gains, pt.losses, by='row.names', all=T)
+
+x = base::merge(x,oac.demo[,demo.cols], by.x='Row.names', by.y='StudySubjectID')
+x$Row.names = NULL
+
+time = 'Weeks.Survival';  event='FE.Patient.Died'; type = '"right"'
+x[c(time,event)] = lapply(x[c(time,event)], as.numeric)
+
+x$RP.TStage.PrimaryTumour = as.character(x$RP.TStage.PrimaryTumour)
+
+x[grepl('T1', x$RP.TStage.PrimaryTumour), 'RP.TStage.PrimaryTumour'] = 'T1'
+x[grepl('T4', x$RP.TStage.PrimaryTumour), 'RP.TStage.PrimaryTumour'] = 'T4'
+x[grepl('Tx', x$RP.TStage.PrimaryTumour), 'RP.TStage.PrimaryTumour'] = 'T0'
+
+terms = as.formula(paste("Surv(",time,",",event,", type=",type,") ~ RP.BarrettsAdjacent", sep=""))
+sfit = survfit(terms, data=x,conf.int=0.95, conf.type='log', na.action=na.omit)
+ggsurv(sfit)
+
+#f1 = paste(demo.cols[c(3,5:7)], collapse='+')
+f1 = paste(demo.cols[4:6], collapse='+')
+f2 = paste(paste('`', grep('^gain|loss', colnames(x), value=T), '`', sep=''), collapse='+')
+f = paste(f1,f2, sep='+')
+
+terms = as.formula(paste("Surv(",time,",",event,", type=",type,") ~", f, sep=""))
+cox = coxph(terms, data=x,na.action=na.omit,ties='breslow')
+cfit = survfit(cox)
+summary(cox)
+ggsurv(cfit) + labs(title=paste(length(grep('gain|loss', colnames(x))), 'copy number + demo'))
+
+pander(cox)
+
+anova(cox)
+```
+
+```{r}
+library(CoxHD)
+
+x = x[,c(grep('gain|loss',colnames(x), invert=T), grep('gain|loss',colnames(x)))]
+xs = x
+
+xs = xs[,grep('Tumour', colnames(xs), invert=T)]
+
+surv = with(xs, Surv(Weeks.Survival, FE.Patient.Died))
+ggsurv(survfit(surv~1,data=xs))
+
+xs = xs[,-grep('Weeks|Died',colnames(xs))]
+
+xs$RP.BarrettsAdjacent = as.integer(xs$RP.BarrettsAdjacent)-1
+#xs$RP.TStage.PrimaryTumour = as.integer(factor(xs$RP.TStage.PrimaryTumour))-1
+#xs$RP.TumourDifferentiation = as.integer(factor(xs$RP.TumourDifferentiation))-1
+xs$DI.PatientGender = as.integer(as.factor(xs$DI.PatientGender))-1
+
+groups = c(rep('clin', 3), rep('cn',(ncol(pt.gains)+ncol(pt.losses)) ))
+
+rfx = CoxRFX(xs, surv, groups, nu=1, verbose=T)
+summary(rfx)
+
+plot(rfx)
+
+VarianceComponents(rfx)
+
+WaldTest(rfx)
+
+PartialC(rfx)
+
+PlotVarianceComponents(rfx)
+
 ```
@@
 ***
-
-# Barrett's Adjacency & CNA
-
-Is there a relationship between BA patients and CNA in selected regions?
-
-## Selected Regions
-
-```{r, eval=F}
-# A lot more gains than losses as expected
-sapply(gains, nrow)
-sapply(losses, nrow)
-
-population = table(oac.demo$RP.BarrettsAdjacent)
-
-# Check for enrichment of gains/losses in BA
-gainsBAPts = as.data.frame(do.call(rbind, lapply(gains, function(g) {
- table(g$RP.BarrettsAdjacent)
-})))
-
-wilcox.test(gainsBAPts$no, gainsBAPts$yes)
-
-# Separated by BA status
-gainsBA = lapply(gains, function(g) {
-  subset(g, RP.BarrettsAdjacent == 'yes')
-})
-
-gainsNoBA = lapply(gains, function(g) {
-  subset(g, RP.BarrettsAdjacent == 'no')
-})
-
-wilcox.test( sapply(be.gains, nrow), sapply(non.be.gains, nrow))
-# No difference
-one.way.aov(sapply(gainsBA, nrow) , sapply(gainsNoBA, nrow))
-# TODO Is this due to numberof patients? Might be
-one.way.aov(sapply(gainsBA, nrow)/gainsBAPts[,'yes'] , sapply(gainsNoBA, nrow)/gainsBAPts[,'no'] )
-
-# No enrichment of gains in selected regions for BA 
-p.adjust( apply(gainsBAPts, 1, function(x) {
-  bt = binom.test(x[['no']], sum(x), population[['no']]/sum(population))  
-  bt$p.value
-}) )
-
-cnBA = do.call(rbind, lapply(subset(oac.demo, !is.na(RP.BarrettsAdjacent))$StudySubjectID, function(s) {
-  cbind.data.frame(
-    'ID'=s,
-    'BarrettsAdjacent'=subset(oac.demo, StudySubjectID == s)$RP.BarrettsAdjacent,
-    'TotalCN'=length(subset(oacCNV, OCCAMS == s)),
-    'Gains'=length(subset(oacCNV, OCCAMS == s  & ploidy > 2 & Total_CN > round(ploidy))),
-    'Losses'=length(subset(oacCNV, OCCAMS == s  & Total_CN < 2)) 
-  )
-}))
-
-arrange(cnBA, -TotalCN)
-
-populationGains = apply(oacCNV, 1, function(x) {
-  yes = subset(oac.demo, StudySubjectID %in% x$OCCAMS & RP.BarrettsAdjacent == 'yes')$StudySubjectID
-  subset(g, OCCAMS %in% yes)
-
-})
-
-apply(cbind(sapply(gainsBA, nrow), sapply(gainsNoBA, nrow)), 1, function(x) {
-  print(x)
-  bt = binom.test(x[1], sum(x), .5)
-  bt$p.value
-})
-
-# But there is a significant difference between adjacent BA -- but is this true across random selected regions as well?
-gainCtBA = as.data.frame(do.call(rbind, lapply(losses, function(g) {
- table(subset(oac.demo, StudySubjectID %in% g$OCCAMS)$RP.BarrettsAdjacent)
-})))
-
-one.way.aov(gainCtBA$no, gainCtBA$yes)
-wilcox.test(gainCtBA$no, gainCtBA$yes) 
-summary(aov(gainCtBA$no~gainCtBA$yes)) # significant difference between adjacent BA -- but is this true across random selected regions as well?
-# But not when adjusted for the population size, though not sure I should do that.  Binom instead?
-one.way.aov(gainCtBA$no/population[['no']], gainCtBA$yes/population[['yes']])
-
-lossCtBA = as.data.frame(do.call(rbind, lapply(losses, function(g) {
- table(subset(oac.demo, StudySubjectID %in% g$OCCAMS)$RP.BarrettsAdjacent)
-})))
-
-wilcox.test(lossCtBA$no, lossCtBA$yes) # significant difference between adjacent BA -- but is this true across random selected regions as well?
-summary(aov(lossCtBA$no~lossCtBA$yes)) # significant difference between adjacent BA -- but is this true across random selected regions as well?
-
-overlaps = rbind(gainCtBA, lossCtBA) %>% rowwise() %>% mutate(
-   'yes.CI.1'=binom.test(yes, sum(yes,no), population[['yes']]/sum(population) )$conf.int[1],
-   'yes.CI.2'=binom.test(yes, sum(yes,no), population[['yes']]/sum(population) )$conf.int[2],
-   'yes.pval'=binom.test(yes, sum(yes,no), population[['yes']]/sum(population) )$p.value,
-   'no.pval'=binom.test(no, sum(yes,no), population[['no']]/sum(population) )$p.value
-) 
-overlaps$regions = c(rownames(gainsBA), rownames(lossBA))
-
-# not significant though
-p.adjust(overlaps$yes.pval, method='bonferroni', n=nrow(overlaps))
-p.adjust(overlaps$no.pval, method='bonferroni', n=nrow(overlaps))
-
-df = transform(overlaps, regions=reorder(regions, (yes.CI.2-yes.CI.1)) )
-ggplot(df , aes(x=regions)) + 
-    geom_segment(data=df, aes(y=yes.CI.1, yend=yes.CI.2, x=regions,xend=regions), color='blue') +
-    coord_flip() + labs(x='', y='CI (yes)')
-
-```
-
-General gains vs losses
-```{r,eval=F}
-
-head(perms)
-# No difference in the numbers of gains/losses
-wilcox.test(sapply(gains, nrow), perms$amp)
-wilcox.test(sapply(losses, nrow), perms$del)
-
-# There is a difference (these are the sampled regions, not selected) between the BA categories, yes -> more disordered??
-wilcox.test(perms$amp.no, perms$amp.yes)
-wilcox.test(perms$del.no, perms$del.yes)
-
-ggplot(melt(perms[,c('amp.yes','amp.no')]), aes(variable, value, group=variable, fill=variable)) + geom_boxplot()
-ggplot(melt(perms[,c('del.yes','del.no')]), aes(variable, value, group=variable, fill=variable)) + geom_boxplot()
-
-random = as.data.frame(t(apply(perms[,grep('yes|no', colnames(perms))], 1, function(row) {
-
-  yes = sum(row[c('amp.yes','del.yes')])
-  no = sum(row[c('amp.no','del.no')])
-
-  cbind('yes' = binom.test(yes, sum(yes,no), population[['yes']]/sum(population))$conf.int,
-        'no' = binom.test(no, sum(yes,no), population[['yes']]/sum(population))$conf.int)
-  # binom.test(yes, sum(yes,no), population[['yes']]/sum(population))$conf.int
-})))
-colnames(random) = c('yes.CI.1','yes.CI.2','no.CI.1','no.CI.2')
-
-summary( with(random, yes.CI.2-yes.CI.1)  )
-summary( with(overlaps, yes.CI.2-yes.CI.1)  )
-
-summary( with(random, no.CI.2-no.CI.1)  )
-summary( with(overlaps, no.CI.2-no.CI.1)  )
-
-# No difference between the features selected and random regions though
-wilcox.test(gainsBA$yes, perms$amp.yes)
-wilcox.test(gainsBA$no, perms$amp.no)
-
-# Nope
-colSums(lossBA)/nrow(lossBA)
-colSums(perms[,c('del.no','del.yes')])/nrow(perms)
-
-t.test(lossBA$yes, perms$del.yes)
-t.test(lossBA$no, perms$del.no)
-
-wilcox.test(gfreq, perms$freq.gain)
-wilcox.test(lfreq, perms$freq.loss)
-
-m = rbind(
-cbind.data.frame('freq'=gfreq, 'type'='gain', 'feature'='swgs'),
-cbind.data.frame('freq'=lfreq, 'type'='loss', 'feature'='swgs'),
-
-cbind.data.frame('freq'=perms$freq.gain, 'type'='gain', 'feature'='random'),
-cbind.data.frame('freq'=perms$freq.loss, 'type'='loss', 'feature'='random') 
-)
-
-# No difference
-ggplot(m, aes(feature, freq, group=feature, fill=feature)) + facet_grid(~type) + geom_jitter(width=0.2) +  geom_violin(alpha=0.8) 
-```
-
-```{r, eval=F}
-all = swgRegions
-
-swgRegions = swgRegions[swgRegions$hazard > quantile(swgRegions$hazard)['75%']]
-
-cols = c('RP.BarrettsAdjacent', 'RP.TStage.PrimaryTumour', 'Weeks.Survival', 'FE.Patient.Died')
-tmp = oac.demo[,cols]
-
-tmp = tmp[-which(is.na(tmp$Weeks.Survival)),]
-
-swgRegions = swgRegions[order(-swgRegions$hazard)]
-
-```
-
-```{r, eval=F}
-
-#summary(as.data.frame(swgRegions)[,c('num.amp.CN','num.del.CN')])
-
-library(gridExtra)
-source('../ICGC-Clinical/libs/surv.R')
-
-tmp$FE.Patient.Died = as.numeric(tmp$FE.Patient.Died)-1
-
-tmp[which(is.na(tmp$FE.Patient.Died)), 'FE.Patient.Died'] = 0
-
-sp = survival.plots(tmp, 'RP.BarrettsAdjacent', verbose=F)
-grid.arrange(sp$kapmei)
-
-time='Weeks.Survival'; event='FE.Patient.Died'
-f = paste('`',paste(grep('AMP|DEL',colnames(tmp), value=T), collapse='`+`'), '`', sep='')
-
-#f = paste('`AMP_', as.character(seqnames(swgRegions[1])), ':', start(swgRegions[1]), '-', end(swgRegions[1]), '`', sep='')
-
-# surv =  Surv(tmp[[time]],tmp[[event]],type='right')
-# terms = as.formula(paste("Surv(",time,",",event,", type='right') ~", 'RP.BarrettsAdjacent', sep=""))
-# 
-# cox = coxph(terms, data=tmp,na.action=na.omit,ties='breslow')
-# cfit = survfit(cox)
-#  
-# ggsurv(cfit,cens.shape=3,lty.est=1,back.white=F, xlab="Time (weeks)") + 
-#       annotate("text", x=200, y=.8,label=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),2),"\n", nrow(tmp), 'samples'))
-# summary(cox)
-
-```

### 2018-05-02 PredictingProgression/lib/data_func.R

@@ -1,287 +1,0 @@
-require(GenomicRanges) 
-require(dplyr)
-require(data.table)
-require(tibble)
-
-
-#source("fastPCF.R")
-
-
-binSWGS<-function(raw.data, fit.data, blacklist) {
-  if (is.null(blacklist)) 
-    stop('QDNASeq blacklisted regions missing')
-  
-  countCols = grep('loc|feat|chr|start|end', colnames(fit.data), invert=T)
-  
-  rows = which(fit.data[,1] %in% raw.data[,1])
-  raw.data = raw.data[rows,]
-  fit.data = fit.data[rows,]
-  
-  window.depths = raw.data[,countCols]/fit.data[,countCols]
-  dim(window.depths)
-  message(paste(nrow(blacklist), "genomic regions in the exclusion list."))
-  fit.data$in.blacklist = F
-  for(r in 1:nrow(blacklist)) {
-    fit.data$in.blacklist[ fit.data$chrom == blacklist$chromosome[r] & 
-                             fit.data$start >= blacklist$start[r] & 
-                             fit.data$end <= blacklist$end[r] ] = T
-  }
-  message(paste("# excluded probes = ",sum(fit.data$in.blacklist),sep=""))
-  if (sum(fit.data$in.blacklist) <= 0)
-    stop("There's an issue with the excluded list, no probes excluded.")
-  
-  if (length(countCols) == 1)
-    window.depths = as.data.frame(window.depths)
-  
-  window.depths.standardised = as.data.frame(window.depths[which(!fit.data$in.blacklist),])
-  
-  fit.data = fit.data[!fit.data$in.blacklist,-ncol(fit.data)]
-  sdevs = sapply(c(1:length(countCols)), function(s) {
-    getMad( window.depths.standardised[!is.na(window.depths.standardised[,s]),s], k=25 )
-  })
-  sdevs[sdevs==0] = NA
-  sdev = exp(mean(log(sdevs[!is.na(sdevs)]))) 
-  
-  good.bins = which(!is.na(rowSums(as.data.frame(window.depths.standardised[,!is.na(sdevs)]))))
-  gamma2=250 
-  
-  data = cbind(fit.data[good.bins,c('chrom','start')],window.depths.standardised[good.bins,!is.na(sdevs)])
-  
-  if (ncol(data) < 4) { # Single sample
[…]

### 2018-05-02 PredictingProgression/ui.R

@@ -1,73 +1,0 @@
-#
-# This is the user-interface definition of a Shiny web application. You can
-# run the application by clicking 'Run App' above.
-#
-# Find out more about building applications with Shiny here:
-# 
-#    http://shiny.rstudio.com/
-#
-
-library(shiny)
-
-
-# Define UI for application that draws a histogram
-shinyUI(fluidPage(
-  
-  # Application title
-  titlePanel("Barrett's Progression Prediction"),
-  sidebarLayout(
-    sidebarPanel(
-      tags$h3('Patient Information'),
-      numericInput("ageY", "Age at diagnosis:", min=1, max=100, value=60,step=1),
-      sliderInput('beLength', 'Length of BE segment', min=1,max=30,round=1,value=4),
-      radioButtons("sex", label = "Sex", choices = list("Male" = 'M', "Female" = 'F'), selected = 'M'),
-      
-      
-      tags$h3('Patient Data'),
-      fileInput("pathfile", "Optional: upload a file with per-sample pathology and/or p53 IHC (see example)",
-                accept = c(
-                  "text/tsv",
-                  "text/csv",
-                  "text/tab-separated-values,comma-separated-values,text/plain",
-                  ".csv", ".tsv", ".txt")
-      ),
-      
-      fileInput("file", "Choose Caveman copy number or segmented copy number file:",
-                accept = c(
-                  "text/tsv",
-                  "text/csv",
-                  "text/tab-separated-values,comma-separated-values,text/plain",
-                  ".csv", ".tsv", ".txt")
-      )
-      # Input: Select separator ----
-      #checkboxInput("header", "Header", TRUE),
-      #checkboxInput("norm", "Model provided normalization (default: yes)", TRUE),
-      
-      #actionButton("action", label = "Submit")
-
-    ),
-  
-    mainPanel(
-      helpText(tags$b("Recommendations are evaluated with the following criteria:"), tags$br(),
-               tags$ol(tags$li("Immediate RFA: HGD diagnosis or more than one consecutive high risk predictions."),
-                       tags$li("Recheck 6-12 months: One high risk prediction or an aberrant p53 IHC"),
-                       tags$li("Recheck endoscopy 12-24 months: One or more moderate risk predictions"),
-                       tags$li("Regular surveillance 3-5 years: Two or more consecutive low risk predictions"))),
-      
-      tabsetPanel(
-        type='tabs',
[…]

### 2018-05-02 lib/data_func.R

@@ -7,6 +7,8 @@ require(tibble)
 #source("fastPCF.R")
 
 
+
+
 binSWGS<-function(raw.data, fit.data, blacklist) {
   require(copynumber)
   
@@ -192,7 +194,7 @@ tile.segmented.data<-function(data, size=5e6, chr.info=NULL) {
       bin = currentChr[i]
       
       segments = gr[subjectHits(curov[queryHits(curov) == i])]
-      weights = lapply(segments,  function(r) {
+      weights = lapply(as(segments,'GRangesList'),function(r) {
         width(pintersect(bin, r))/width(bin)
       })
       
@@ -283,4 +285,39 @@ get.chr.lengths<-function(chrs = paste('chr', c(1:22, 'X','Y'), sep=''), build='
   chr.lengths$genome.length = cumsum(as.numeric(chr.lengths$chr.length))
   
   return(chr.lengths)
+}
+
+
+modelMatrixSetup <- function(rs, chr.info, segMeans=NULL, segSDs=NULL, armMeans=NULL, armSDs=NULL) {
+  segs <- tile.segmented.data(rs, size=5e6, chr.info=chrlen)
+  segM = as.matrix(segs[,-c(1:3)])
+  rownames(segM) = paste(segs$chr, ':', segs$start, '-', segs$end, sep='')
+  segM = t(segM)
+  
+  if (!is.null(segMeans) && !is.null(segSDs)) {
+    for (i in 1:ncol(segM)) 
+      segM[,i] = unit.var(segM[,i], segMeans[i], segSDs[i])
+  } else {
+    segM = unit.var(segM)
+  }
+  
+  arms <- tile.segmented.data(rs, size='arms', chr.info=chrlen)
+  armsM = as.matrix(arms[,-c(1:3)])
+  rownames(armsM) = paste(arms$chr, ':', arms$start, '-', arms$end, sep='')
+  armsM = t(armsM)
+  
+  if (!is.null(armMeans) && !is.null(armSDs)) {
+    for (i in 1:ncol(armsM)) 
+      armsM[,i] = unit.var(armsM[,i], z.arms.mean[i], z.arms.sd[i])
+  } else {
+    armsM = unit.var(armsM)
+  }
+  
+  nrow(armsM) == nrow(segM)
+  
+  cx.score = score.cx(segM,1)
+  
+  mergedDf = subtract.arms(segM, armsM)
[…]

### 2018-05-02 processing_pipeline/predict_downsampled.R

@@ -19,8 +19,9 @@ adjustRisk <- function(RR, offset, type='risk') {
 }
 
 
-dir = '~/Data/Ellie/Analysis/downsampled5G'
-dir = '~/Data/Ellie/Analysis/VAL_Cohort/'
+dir = '~/Data/Ellie/Analysis/downsampled5G_NDBE/'
+dir = '~/Data/Ellie/Cleaned/AH0329_segmentedCoverage_fitted_gamma250/'
+#dir = '~/Data/Ellie/Analysis/VAL_Cohort/'
 files = list.files(dir, full.names=T)
 
 file = '~/Data/Ellie/Analysis/5e6_arms_all_exAHM0320/model_data.Rdata'
@@ -32,6 +33,10 @@ if (!file.exists(file))
   stop(paste("Missing data file", file))
 load(file, verbose=T)
 rm(dysplasia.df,coefs,plots,labels)
+
+select.alpha = '0.9'
+fitV = models[[select.alpha]]
+lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']
 
 
 file = '~/Data/Ellie/Analysis/5e6_arms_all_exAHM0320/loo.Rdata'
@@ -50,12 +55,8 @@ m = round((table(status)['P']/(0.0225*100))*100)
 offsetMean = log(cases/m)
 
 
-select.alpha = '0.9'
-fitV = models[[select.alpha]]
-lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']
-
 sampleNames = sub('_raw.txt', '', basename(files))
-dspred = data.frame(matrix(ncol=4, nrow=length(sampleNames), dimnames=list(sampleNames,c('Prob','RR', 'AdjProb','AdjRR'))))
+dspred = data.frame(matrix(ncol=4, nrow=length(sampleNames), dimnames=list(sampleNames,c('Prob','RR', 'Adj.Prob','Adj.RR'))))
 
 for (n in 1:length(files)) {
   print(n)
@@ -94,23 +95,16 @@ for (n in 1:length(files)) {
   
   preds = predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='response')
   RR = predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='link')
-  
+
   dspred[n,] = c(preds[,1],RR[,1],adjustRisk(RR, offsetMean, 'prob'),adjustRisk(RR, offsetMean))
   print(dspred[1:n,])
 }
 
[…]

### 2018-05-02 test-methylation.R

@@ -1,0 +1,100 @@
+
+
+library(ggplot2)
+#library(Matrix)
+library(glmnet)
+
+source('lib/data_func.R')
+#source('lib/load_patient_metadata.R')
+
+file = '~/Data/Ellie/Analysis/5e6_arms_all_exAHM0320/model_data.Rdata'
+if (!file.exists(file))
+  stop(paste("Missing data file", file))
+load(file, verbose=T)
+file = '~/Data/Ellie/Analysis/5e6_arms_all_exAHM0320/all.pt.alpha.Rdata'
+if (!file.exists(file))
+  stop(paste("Missing data file", file))
+load(file, verbose=T)
+rm(dysplasia.df,coefs,plots,labels)
+
+select.alpha = '0.9'
+fitV = models[[select.alpha]]
+lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']
+
+
+file = '~/Data/Ellie/Analysis/5e6_arms_all_exAHM0320/loo.Rdata'
+if (!file.exists(file))
+  stop(paste("Missing data file", file))
+load(file, verbose=T)
+rm(plots, coefs, nzcoefs, fits)
+
+pg.samp = do.call(rbind, pg.samp)
+pg.samp$Hospital.Research.ID = NULL
+
+status = unique(pg.samp[,c('Patient','Status')])$Status
+
+cases = table(status)['P']
+m = round((table(status)['P']/(0.0225*100))*100)
+offsetMean = log(cases/m)
+
+
+adjustRisk <- function(RR, offset, type='risk') {
+  if (type == 'prob') {
+    x = 1/(1+exp(-RR+abs(offset)))
+  } else {
+    x = RR+offset
+  }
+  return(x)
+}
+
+# Predict function giving me difficulty when I have only a single sample, this ensures the dimensions are the same
+predict.progression<-function(mergedDf) {
+  sparsed_test_data <- Matrix(data=0, nrow=nrow(mergedDf),  ncol=ncol(mergedDf),
+                              dimnames=list(rownames(mergedDf),colnames(mergedDf)), sparse=T)
+  
+  for(col in colnames(mergedDf)) sparsed_test_data[,col] = mergedDf[,col]
+  
+  preds = predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='response')
[…]

### 2018-05-14 ElliesThesisFigs.Rmd
Trying to adjust and log our data to make it similar to LogR values from arrays.

@@
   facet_grid(~Status) +
-  geom_bar(stat='identity', show.legend = F) + 
+  geom_bar(stat='identity', show.legend = F, fill=riskCols[2]) + 
   geom_errorbar(color='grey21', width=0.2, alpha=0.5) +
   geom_text(aes(y=pred.r*100+2, label=paste(round(pred.r*100), '%', sep='')), color='black') +
-  geom_text(aes(y=10, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
+  geom_text(aes(y=5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
   labs(x='Months prior to endpoint', y='Percent predicted', title='High risk predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') +
@@
 rp
-ggsave(filename='thesis/pred_endo.png', plot=rp, width=11, height=5, units='in', dpi=300, scale=1)
+ggsave(filename='thesis/pred_endo.png', plot=rp, width=11, height=8, units='in', dpi=300, scale=1)
 
@@
 
+repeats$Probability = round(repeats$Probability, 2)
+
+write.table(repeats, sep='\t', quote=F, row.names=F, file='tech_repeats.txt')
+
 ```

### 2018-05-17 AllSamplesModel.Rmd
...

@@
 
-cache.dir = paste(data, 'Analysis/5e6_arms_all_exAHM0320', sep='/')
+cache.dir = paste(data, 'Analysis/5e6_arms_all_logR', sep='/')
 
@@
 p = ggplot(classification.rate, aes(model, mean, group=model, fill=model)) + geom_col() +
-  geom_text( aes(label=round(mean, 2)), vjust=2, size=5, color='gray39', show.legend=F) +
+  geom_text( aes(label=round(mean, 3)), vjust=2, size=5, color='gray39', show.legend=F) +
   geom_text( aes(y=0.6, label=paste(all.coef, " coef\n(", stable.coef, ")", sep='') ), color='gray39', size=5) +
@@
 
-#ggsave('plots/performance/compare_classification.png', plot=p, height=5, width=3, units='in', dpi=300 )
+ggsave('plots/performance/compare_classification.png', plot=p, height=5, width=3, units='in', dpi=300 )
 ```
@@
 ggsave('plots/performance_loo.png', plot=p, width=5, height=5, units='in', dpi=300)
-
-# oddPt = names(pg.samp)[which.max(performance.at.1se)]
-# subset(sum.patient.data, Hospital.Research.ID == oddPt)
-# pg.samp[[oddPt]]$Prediction
-# 
-# oddPt = names(pg.samp)[which.min(performance.at.1se)]
-
-# Not the pt with the most samples or endos
-#sum.patient.data[which.max(sum.patient.data$total.samples),]
-```
-
-```{r echo=F, eval=F}
-## All LOO
-files = list.files('~/Data/Ellie/Analysis/5e6_arms_LOO', full.names = T)
-
-sum.patient.data$Hospital.Research.ID[1]
-
-pg.samp = list()
-performance = vector(length=nrow(sum.patient.data), mode='numeric')
-names(performance) = sum.patient.data$Hospital.Research.ID
-
-for (pt in sum.patient.data$Hospital.Research.ID) {
-  ptF = grep(pt, files, value=F)
-  load(grep('loo\\.Rdata', ptF, value=T), verbose=T)
-  #load(grep('model_data\\.Rdata', ptF, value=T), verbose=T)
-  performance[pt] = performance.at.1se
-  pg.samp[[pt]] = info
-}
-rm(plots,coefs,nzcoefs,fits,info)
-
-#load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/PR1_HIN_042_loo.Rdata", verbose=T)
-#load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/PR1_HIN_042_model_data.Rdata", verbose=T)
-
-#load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/AHM0320_loo.Rdata", verbose=T)
-#load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/AHM0320_model_data.Rdata", verbose=T)
-
-fitLOO <- glmnet(dysplasia.df, labels[rownames(dysplasia.df)], alpha=0.9, family='binomial', nlambda=500, standardize=F)
-
-pm = predict(fitLOO, newx=allDf, s=fits$lambda.1se, type='response')
-or = predict(fitLOO, newx=allDf, s=fits$lambda.1se, type='link')
-
-hist(pm)
-hist(or)
-
-pm[subset(patient.info, Hospital.Research.ID == 'AHM0320')$Samplename,]
-or[subset(patient.info, Hospital.Research.ID == 'AHM0320')$Samplename,]
-
-sort(performance)
-
-pg.samp[which.max(performance)]
-
-pg.samp[which.min(performance)]
-
-back = do.call(rbind, pg.samp)
-
-subset(sum.patient.data, Status == 'NP' & total.endos <=2)
-
-pm[pg.samp[['AD0309']]$Samplename,]
-pg.samp[['AD0309']]$Prediction
-
 ```
@@
 roc$is = 'all'
+roc.plot(roc)
 
@@
 
-load(file=paste(cache.dir, 'insilico_splits.Rdata', sep='/'), verbose=T)
+insilico.rocs = list()
+ifile = paste(cache.dir, 'insilico_splits.Rdata', sep='/')
+if (file.exists(ifile)) {
+  load(ifile, verbose=T)
   insilico.rocs[['all']] = roc
@@
   rr  
-ggsave(filename='plots/auc/roc_CI.png', plot=rr, width=4, height=4, units='in', dpi=300, scale=1.5)
+  ggsave(filename='plots/auc/roc_CI.png', plot=rr, width=6, height=6, units='in', dpi=300, scale=1.5)
+} else {
+  message("No insilico split file.")
+}
 
@@
 ft.fun<-function(NP,P) {
-  f = chisq.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
+  #f = chisq.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
+  f = chisq.test(c(NP,P))
   cbind.data.frame('p.value'=f$p.value)
@@
 
-qt = cbind(qt, back %>% group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))#, 
-pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% dplyr::mutate( 
+qt = cbind(qt, back %>% group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
+pred.confidence = qt %>% as.data.frame.matrix %>% rowwise() %>% dplyr::mutate( 
   'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'conf'=ifelse(p.value < 0.05, '*', '') )
@@
 pred.confidence$Risk = 'Moderate'
-pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc < .5)) ] = 'Low'
+pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc < .3)) ] = 'Low'
 pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'
@@
 #pred.confidence$Risk[6:10] = 'High'
-pred.confidence$Risk[4:7] = 'Moderate'
+#pred.confidence$Risk[4:7] = 'Moderate'
 
@@
 ## count arms differently than segs??
-df2 = df %>% group_by(Status, Patient, Endoscopy.Year) %>% summarise( 
+df2 = df %>% group_by(Status, Patient, Endoscopy.Year) %>% dplyr::summarise( 
   'max.Pred'=max(Prediction),
@@
 
-df2 = df2 %>% group_by(Status, Patient) %>% mutate(
+df2 = df2 %>% group_by(Status, Patient) %>% dplyr::mutate(
   'early'= Endoscopy.Year == min(Endoscopy.Year),
@@
 el = subset(df2, early | late)
-el2 = el %>% group_by(Status, Patient) %>% summarise( diff(arms), diff(CNA) )
+el2 = el %>% group_by(Status, Patient) %>% dplyr::summarise( diff(arms), diff(CNA) )
 wtA = round(wilcox.test(subset(el2, Status == 'Progressor')$`diff(arms)`, subset(el2, Status == 'Non-Progressor')$`diff(arms)`)$p.value, 3)
@@
   group_by(Patient, Status) %>% 
-  summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )
+  dplyr::summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )
 
@@
 
-with(second, table(Status, rule))
 ```
@@
 
-```{r 5yrrisk, echo=F, fig.height=8, fig.width=12}
+```{r 5yrrisk, echo=F, fig.height=8, fig.width=12, eval=F}
 cases = table(sum.patient.data$Status)['P']

### 2018-05-17 Insilico Splits.Rmd
...

@@
 data = '~/Data/Ellie'
-tile.w=c(5e06, 'arms')
-
-includeDemo = F
 
@@
 analysis.files = list.files(paste(data, 'Analysis', sep='/'), full.names=T)
-
-load(grep('All_patients.Rdata', analysis.files, value=T), verbose=T)
 
@@
 
-#sum.pt = summarise.patient.info(all.patient.info)
+sum.pt = summarise.patient.info(all.patient.info)
 
-sum.pt = all.patient.info %>% group_by(Set, Patient) %>% summarize(
-  med.reads=median(Total.Reads),
-  sd.reads=sd(Total.Reads),
-  n.endoscopy=length(unique(PID)),
-  n.years=unique(Final.Endoscopy-Initial.Endoscopy),
-  n.samples=length(Samplename)
-)
-
-sp = sum.pt %>% group_by(Set) %>% summarize(
-  med.endo=median(n.endoscopy),
-  sd.endo=sd(n.endoscopy),
-  med.years=median(n.years),
-  med.samples=median(n.samples),
-  sd.samples=sd(n.samples)
-)
-
-```
-
-`r pander(as.data.frame(sp))`
-
-```{r echo=F, warning=F, message=F}
-
-#cache.dir = paste(data, 'Analysis', sub('\\+','',paste(tile.w, collapse='_')), sep='/')
-model.dir = paste(data, 'Analysis/5e6_arms_all_exAHM0320', sep='/')
-split.dir = '/Volumes/research1/gerstung/skillcoyne/Ellie/Analysis/5e6_arms_splits_exAHM0320'
+cache.dir = paste(data, 'Analysis/5e6_arms_all_logR', sep='/')
+split.dir = paste(data, 'Analysis/5e6_arms_splits_logR', sep='/')
 
@@
 
-file = paste(model.dir, 'insilico_splits.Rdata', sep='/')
+file = paste(cache.dir, 'insilico_splits.Rdata', sep='/')
 if (file.exists(file)) {
@@
 } else {
-
   split.dirs = list.files(split.dir, full.names = T)
@@
     print(split)  
+    models = list.files(split.dirs[split],full.names = T, pattern='all.*Rdata')
+    loo = list.files(split.dirs[split],full.names = T, pattern='loo.*Rdata')
+
+    if (length(models) < 1 || length(loo) < 1) {
+      warning(paste("Missing model or LOO Rdata file in split", split.dirs[split]))
+      next
+    }
+
     load(list.files(split.dirs[split],full.names = T, pattern='all.*Rdata'), verbose=T)
@@
     roc$is = split
-    rocs[[split]] = roc
+    rocs[[as.character(basename(split.dirs[split]))]] = roc
 
@@
   insilico.perf=performance
-  save(insilico.rocs, insilico.perf, file=paste(model.dir, 'insilico_splits.Rdata', sep='/'))
+  save(insilico.rocs, insilico.perf, file=paste(cache.dir, 'insilico_splits.Rdata', sep='/'))
 }
@@
 
-load(list.files(model.dir, full.names = T, pattern = 'all.*Rdata'))
+load(list.files(cache.dir, full.names = T, pattern = 'all.*Rdata'))
 
@@
 
-load(list.files(model.dir,full.names = T, pattern='loo.Rdata'))
+load(list.files(cache.dir,full.names = T, pattern='loo.Rdata'))
 rm(plots, performance.at.1se,coefs,nzcoefs,fits)
@@
 
-preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
-roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
+roc = pROC::roc(Status ~ Prediction, data=predictions, auc=T, ci=T, of='thresholds')
 roc$model = 'discovery'
@@
 ggplot(m, aes(variable,value)) + ylim(0,1) +
-  geom_boxplot(fill='mintcream', color='grey55') + geom_jitter(width=0.2, aes(color=name), size=1.5) + 
+  geom_boxplot(fill='mintcream', color='grey55', outlier.color = NA) + geom_jitter(width=0.2, aes(color=name), size=1.5) + 
   geom_text(data=subset(m, value %in% range(value)), aes(label=round(value,2)), vjust=0.02, hjust=0.2) +
@@
 ```
-Median AUC `r median(rocs$auc)`, SD `r sd(rocs$auc)`
+Median AUC `r round(median(sapply(rocs, function(x) x$auc)), 2)`, SD `r round(sd(sapply(rocs, function(x) x$auc)), 2)`

### 2018-05-17 SNP_sWGS_comparison.Rmd
...

---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r, include=FALSE}
library(copynumber)
library(tibble)
library(dplyr)

```

```{r}

relCNV = read.table('~/Data/Ellie/Cleaned/AD0027_segmentedCoverage_fitted_gamma250/raw_probefiltered_segvals.txt', sep='\t', header=T)
head(relCNV)
#counts = read.table('~/Data/Ellie/QDNAseq/SLX-12452/SLX-12452.binSize15.readCounts.txt')
#corrected = read.table('~/Data/Ellie/QDNAseq/SLX-12452/SLX-12452.binSize15.correctedReadCounts.txt')

summary(relCNV)

segCNV = read.table('~/Data/Ellie/Cleaned/AD0027_segmentedCoverage_fitted_gamma250/5e+06_cleaned_tiled_segvals.txt', sep='\t', header=T)
head(segCNV)

```

```{r, warnings=F}
sa = data.table::fread('~/Downloads/GSE99429_normalized_data.txt')
#colnames(sa)
# baseline, hgd  final, EA
infocol = grep('SNP|Chr|Pos', colnames(sa), value=T)

ptcol = grep('391', colnames(sa), value=T)

pt = sa[, c(infocol, ptcol), with=F]

samplenames = unique( sub('\\.[Log|B\\.Allele].*', '', ptcol))
perpt = lapply (samplenames , function(sname) {
  x = pt[, c(infocol,grep(paste(sname, '.*Log', sep=''), ptcol, value=T),grep(paste(sname, '.*Allele', sep=''), ptcol, value=T)), with=F  ]
  colnames(x) = c('rs.id', 'chr','position', 'Log.R','B.Allele.Freq')

  x = subset(x, !chr %in% c('0','XY','MT') & !is.na(Log.R))

  x[which(x$chr == 'X'), 'chr'] = 23
  if (length(which(x$chr == 'YY')) > 0)
    x[which(x$chr == 'Y'), 'chr'] = 24

  x$chr = as.numeric(x$chr)
  x = dplyr::arrange(x, chr, position)

  as_tibble(cbind('sample'=sname, x))
})
names(perpt) = samplenames

data = perpt$`391_67K_SS6_2_2_22654`[,c('chr','position','Log.R')]
colnames(data)[3] = '391_67K_SS6_2_2_22654'
data$chr = as.numeric(data$chr)
segdata = copynumber::pcf(as.data.frame(subset(data, chr %in% c(1:22))), gamma=28, fast=T, verbose=T)
hist(segdata$mean)

head(perpt$`391_67K_SS6_2_2_22654`)
head(perpt$`391_73P__normal_5350N_5ng`)

hist(perpt$`391_67K_SS6_2_2_22654`$Log.R)
ks.test(perpt$`391_67K_SS6_2_2_22654`$Log.R, 'pnorm')

ks.test(rnorm(500, mean(perpt$`391_67K_SS6_2_2_22654`$Log.R), sd(perpt$`391_67K_SS6_2_2_22654`$Log.R)), 'pnorm')

ks.test(perpt$`391_67K_SS6_2_2_22654`$Log.R, rnorm(500, mean(perpt$`391_67K_SS6_2_2_22654`$Log.R), sd(perpt$`391_67K_SS6_2_2_22654`$Log.R)))
?ks.test

```

### 2018-05-17 lib/data_func.R

@@ -40,13 +40,14 @@ binSWGS<-function(raw.data, fit.data, blacklist, logTransform=F) {
   # QDNAseq does this in 'correctBins' but it's too late to use this now, so will adjust ours the same way (we weren't 14-May-2018)
   #corrected <- counts / fit
   #corrected[fit <= 0] <- 0
-  negs = apply(fit.data[,depth.cols], 2, function(x) which(x<=0))
-  window.depths[] = lapply(names(negs), function(n) {
-    window.depths[negs[[n]],n] = 0
-    window.depths[,n]
-  })
+  negs = apply(as.matrix(fit.data[,countCols]), 2, function(x) which(x<=0))
+  if (length(negs) > 0) {
+    window.depths[] = lapply(names(negs), function(n) {
+      window.depths[negs[[n]],n] = 0
+      window.depths[,n]
+    })
+  }
   
-  dim(window.depths)
   message(paste(nrow(blacklist), "genomic regions in the exclusion list."))
   fit.data$in.blacklist = F
   for(r in 1:nrow(blacklist)) {

### 2018-05-22 ElliesThesisFigs.Rmd
testing SNP data

@@
 ```{r echo=F, warning=F, message=F}
-
 file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
@@
 
+fitV = models[[0.9]]
+lambda.opt = performance.at.1se[[0.9]][, 'lambda']
+
 validation.patient.info =  subset(all.patient.info, Set == 'Test')
@@
 
-cleaned = list.files(path=paste(data, 'Cleaned', sep='/'), pattern='tiled', full.names=T, recursive=T)
+cleaned = list.files(path=paste(data, 'Cleaned', sep='/'), pattern='raw', full.names=T, recursive=T)
 cleaned = grep(paste(unique(validation.patient.info$Hospital.Research.ID), collapse='|'), cleaned, value=T)
-cleaned = grep('arms', cleaned, invert=T, value=T)
-
-#raw = list.files(path=paste(data, 'Cleaned', sep='/'), pattern='raw', full.names=T, recursive=T)
-#raw = grep(paste(unique(validation.patient.info$Hospital.Research.ID), collapse='|'), raw, value=T)
-
-fitV = models[[0.9]]
-lambda.opt = performance.at.1se[[0.9]][, 'lambda']
-
-validation.df = matrix(nrow=0, ncol=ncol(dysplasia.df), dimnames=list(c(), colnames(dysplasia.df)))
-for (fi in cleaned) {
-  x = read.table(fi, sep='\t', header=T)
-  segs = paste(x$chr, ':', x$start, '-', x$end, sep='')
-
-  x = t(x[,-c(1:3)])
-  colnames(x) = segs
-
-  validation.df = rbind(validation.df, x)  
-}
-
+
+chr.len = get.chr.lengths()
+
+tmp = 'thesis/thesis_validationdf.Rdata'
+if (file.exists(tmp)) {
+  load(tmp)
+} else {
+  validation.df = do.call(rbind, lapply(cleaned, function(f) {
+    df = read.table(f, sep='\t', header=T)
+    tiled = segment.matrix(tile.segmented.data(df, chr.info=chr.len))
+    for (i in 1:ncol(tiled)) 
+      tiled[,i] = unit.var(tiled[,i]) 
+    tiled[is.na(tiled)] = 0
+    tiled
+  }))
   validation.df = validation.df[validation.patient.info$Samplename, ]
@@
 
+  save(validation.df, file=tmp)
+}
+
 pred = predict(fitV, newx=validation.df, s=lambda.opt, type='response')
@@
 
-if (rownames(pred) != vpd$Samplename)
-  stop('Fix this')
-
 vpd$Prediction = pred
@@
 
-table(subset(vpd, Prediction > cutoff)$Status)
-table(subset(vpd, Prediction < cutoff)$Status)
+table(subset(vpd, Prediction > 0.5)$Status)
+table(subset(vpd, Prediction < 0.5)$Status)
 
@@
 
-cleaned = list.files(path=paste(data, 'Cleaned', sep='/'), pattern='tiled', full.names=T, recursive=T)
+cleaned = list.files(path=paste(data, 'Cleaned', sep='/'), pattern='raw', full.names=T, recursive=T)
 cleaned = grep(paste(unique(repeats$`Hospital Research ID`), collapse='|'), cleaned, value=T)
-cleaned = grep('arms', cleaned, invert=T, value=T)
-
-repeat.df = matrix(nrow=0, ncol=ncol(dysplasia.df), dimnames=list(c(), colnames(dysplasia.df)))
-for (fi in cleaned) {
-  x = read.table(fi, sep='\t', header=T)
-  segs = paste(x$chr, ':', x$start, '-', x$end, sep='')
-
-  x = t(x[,-c(1:3)])
-  colnames(x) = segs
-
-  repeat.df = rbind(repeat.df, x)  
-}
-
-rows = which(rownames(repeat.df) %in% samples)
-repeat.df = repeat.df[rows,]
-
+
+tmp = 'thesis/thesis_repeats.df.Rdata'
+if (file.exists(tmp)) {
+  load(tmp)
+} else {
+  repeat.df = do.call(rbind, lapply(cleaned, function(f) {
+    df = read.table(f, sep='\t', header=T)
+    df = df[, c(1:4, which(colnames(df) %in% repeats$Samplename))]
+
+    tiled = segment.matrix(tile.segmented.data(df, chr.info=chr.len))
+    tiled[is.na(tiled)] = 0
+    tiled
+  }))
   repeat.df = apply(repeat.df, 2, unit.var)
+
+  save(repeat.df, file=tmp)
+}
 
@@
 
-write.table(repeats, sep='\t', quote=F, row.names=F, file='tech_repeats.txt')
+write.table(repeats, sep='\t', quote=F, row.names=F, file='thesis/tech_repeats.txt')
+
+write.table(subset(back, Patient %in% unique(repeats$`Patient ID`)), sep='\t', row.names=F, file='thesis/original_rep_predictions.txt')
 

### 2018-05-22 SNP_sWGS_comparison.Rmd
testing SNP data

@@
 title: "R Notebook"
-output: html_notebook
+output: html_document
 editor_options: 
-  chunk_output_type: inline
+  chunk_output_type: console
 ---
@@
 library(dplyr)
+library(ggplot2)
+library(gridExtra)
+library(glmnet)
+
+suppressPackageStartupMessages( source('lib/data_func.R') )
+```
+
+Single sample from shallow WGS dataset.
+```{r, warning=F, message=F}
+relCNV = read.table('~/Data/Ellie/Cleaned/AH0254_segmentedCoverage_fitted_gamma250/raw_probefiltered_segvals.txt', sep='\t', header=T)
+
+segCNV = load.segment.matrix('~/Data/Ellie/Cleaned/AH0254_segmentedCoverage_fitted_gamma250/5e+06_cleaned_tiled_segvals.txt')
+unlogged = segCNV
+segCNV = t(apply(segCNV, 1, logTransform))
+
+grid.arrange(
+  ggplot(relCNV, aes(x=1:nrow(relCNV), y=D702_D506_10725_10729)) + geom_point(color='darkblue') + labs(title='Raw', x=''),
+  ggplot(melt(segCNV['D702_D506_10725_10729',]), aes(x=1:ncol(segCNV), y=value)) + geom_point(color='darkblue') + labs(title='Tiled & Logged', y='D702_D506_10725_10729 (IMC)', x='') )
+
+grid.arrange(
+  ggplot(relCNV, aes(x=1:nrow(relCNV), y=D706_D507_9242)) + geom_point(color='darkblue') + labs(title='Raw', x=''),
+  ggplot(melt(segCNV['D706_D507_9242',]), aes(x=1:ncol(segCNV), y=value)) + geom_point(color='darkblue') + labs(title='Tiled & Logged', y='D706_D507_9242 (NDBE)', x='') )
+
+grid.arrange(  
+  ggplot(relCNV, aes(D702_D506_10725_10729)) + geom_histogram(color='lightblue', bins=25) + labs(title='Raw'),
+  ggplot(melt(unlogged['D702_D506_10725_10729',]), aes(value)) + geom_histogram(color='lightblue', bins=25) + labs(title='Segmented'),
+  ggplot(melt(segCNV['D702_D506_10725_10729',]), aes(value)) + geom_histogram(color='lightblue', bins=25) + labs(title='Logged & Segmented'), top='IMC'
+)
+
+grid.arrange(  
+  ggplot(relCNV, aes(D706_D507_9242)) + geom_histogram(color='lightblue', bins=25) + labs(title='Raw'),
+  ggplot(melt(unlogged['D706_D507_9242',]), aes(value)) + geom_histogram(color='lightblue', bins=25) + labs(title='Segmented'),
+  ggplot(melt(segCNV['D706_D507_9242',]), aes(value)) + geom_histogram(color='lightblue', bins=25) + labs(title='Logged & Segmented'), top='NDBE'
+)
+
+# All samples
+load('~/Data/Ellie/Analysis/5e6_arms_all_logR/model_data.Rdata', verbose=T)
+load('~/Data/Ellie/Analysis/5e6_arms_all_logR/loo.Rdata', verbose=T)
+subset(pg.samp$AH0254, Samplename %in% c('D706_D507_9242','D702_D506_10725_10729'))
 
@@
 
-```{r}
+GSE99429 LogR data
+```{r, warning=F, message=F}
+filelist = list.files('~/Data/Ellie/arrays/SNP', pattern='txt', full.names=T)
+samplelist = lapply(filelist, function(f) colnames(as_tibble(data.table::fread(f, nrows=2))))
 
-relCNV = read.table('~/Data/Ellie/Cleaned/AD0027_segmentedCoverage_fitted_gamma250/raw_probefiltered_segvals.txt', sep='\t', header=T)
-head(relCNV)
-#counts = read.table('~/Data/Ellie/QDNAseq/SLX-12452/SLX-12452.binSize15.readCounts.txt')
-#corrected = read.table('~/Data/Ellie/QDNAseq/SLX-12452/SLX-12452.binSize15.correctedReadCounts.txt')
+whole_epi = lapply(samplelist, function(s) grep('*whole.*Log', s, value=T)  )
+names(whole_epi) = basename(filelist)
 
-summary(relCNV)
+normals = lapply(samplelist, function(s) grep('*normal.*Log', s, value=T)  )
+names(normals) = basename(filelist)
 
-segCNV = read.table('~/Data/Ellie/Cleaned/AD0027_segmentedCoverage_fitted_gamma250/5e+06_cleaned_tiled_segvals.txt', sep='\t', header=T)
-head(segCNV)
+fix.chr<- function(df) {
+  df = subset(df, Chr %in% c(1:22))
+  df$Chr = as.numeric(df$Chr)
+  dplyr::arrange(df, Chr, Position)
+}
+
+dt = as_tibble(data.table::fread(filelist[6]))
+epi = dt[,c(grep('SNP|Chr|Pos', colnames(dt), value=T),whole_epi[[6]])]
+nm = dt[,c(grep('SNP|Chr|Pos', colnames(dt), value=T),normals[[6]])]
+
+epi = fix.chr(epi)
+nm = fix.chr(nm)
+rm(dt)
 
@@
 
-```{r, warnings=F}
-sa = data.table::fread('~/Downloads/GSE99429_normalized_data.txt')
-#colnames(sa)
-# baseline, hgd  final, EA
-infocol = grep('SNP|Chr|Pos', colnames(sa), value=T)
+```{r, warning=F, message=F}
 
-ptcol = grep('391', colnames(sa), value=T)
+process.patients<-function(df, outfile=NULL) {
+  predictions = data.frame(matrix(nrow=0,ncol=2,dimnames=list(c(), c('Prob','RR'))))
+  allptsdata = list()
 
-pt = sa[, c(infocol, ptcol), with=F]
+  infocols = grep('Chr|Pos', colnames(df), value=T)
+  pts = grep('SNP|Chr|Pos', colnames(df), invert=T, value=T)
 
-samplenames = unique( sub('\\.[Log|B\\.Allele].*', '', ptcol))
-perpt = lapply (samplenames , function(sname) {
-  x = pt[, c(infocol,grep(paste(sname, '.*Log', sep=''), ptcol, value=T),grep(paste(sname, '.*Allele', sep=''), ptcol, value=T)), with=F  ]
-  colnames(x) = c('rs.id', 'chr','position', 'Log.R','B.Allele.Freq')
+  for (pt in pts)  {
+    print(pt)    
+    data = df[,c(infocols,pt)]
+    segdata = copynumber::pcf(as.data.frame(data), gamma=28, fast=T, verbose=F)  
+    head(segdata)
 
-  x = subset(x, !chr %in% c('0','XY','MT') & !is.na(Log.R))
+    #ggplot(segdata, aes(x=1:nrow(segdata), y=mean)) + geom_point() + labs(title='LogR whole epi', x='', y='')
 
-  x[which(x$chr == 'X'), 'chr'] = 23
-  if (length(which(x$chr == 'YY')) > 0)
-    x[which(x$chr == 'Y'), 'chr'] = 24
+    # this is the LogR, doesn't need to be logged again
+    tiled = segment.matrix(tile.segmented.data(segdata[,-1], chr.info=get.chr.lengths()))
+    tiled.noUV = tiled
 
-  x$chr = as.numeric(x$chr)
-  x = dplyr::arrange(x, chr, position)
+    for (i in 1:ncol(tiled)) 
+      tiled[,i] = unit.var(tiled[,i], z.mean[i], z.sd[i]) 
+    tiled[is.na(tiled)] = mean(tiled, na.rm=T) #0
 
-  as_tibble(cbind('sample'=sname, x))
-})
-names(perpt) = samplenames
+    cx = score.cx(tiled, 1)
 
-data = perpt$`391_67K_SS6_2_2_22654`[,c('chr','position','Log.R')]
-colnames(data)[3] = '391_67K_SS6_2_2_22654'
-data$chr = as.numeric(data$chr)
-segdata = copynumber::pcf(as.data.frame(subset(data, chr %in% c(1:22))), gamma=28, fast=T, verbose=T)
-hist(segdata$mean)
+    tiled.arms = segment.matrix(tile.segmented.data(segdata[,-1], size='arms', chr.info=get.chr.lengths()))
+    tiled.armsNoUV = tiled.arms
 
-head(perpt$`391_67K_SS6_2_2_22654`)
-head(perpt$`391_73P__normal_5350N_5ng`)
+    for (i in 1:ncol(tiled.arms)) 
+      tiled.arms[,i] = unit.var(tiled.arms[,i], z.arms.mean[i], z.arms.sd[i]) 
+    tiled.arms[is.na(tiled.arms)] = mean(tiled, na.rm=T) #0
 
-hist(perpt$`391_67K_SS6_2_2_22654`$Log.R)
-ks.test(perpt$`391_67K_SS6_2_2_22654`$Log.R, 'pnorm')
+    arrayDf = subtract.arms(tiled, tiled.arms)
+    arrayDf = cbind(arrayDf, 'cx'=unit.var(cx, mn.cx ,sd.cx))
 
-ks.test(rnorm(500, mean(perpt$`391_67K_SS6_2_2_22654`$Log.R), sd(perpt$`391_67K_SS6_2_2_22654`$Log.R)), 'pnorm')
+    # Whole epi
+    predictions[pt,] = c(predict(fitV, newx=arrayDf, s=l, type='response')[,1], 'RR'=predict(fitV, newx=arrayDf, s=l, type='link')[,1] )
 
-ks.test(perpt$`391_67K_SS6_2_2_22654`$Log.R, rnorm(500, mean(perpt$`391_67K_SS6_2_2_22654`$Log.R), sd(perpt$`391_67K_SS6_2_2_22654`$Log.R)))
-?ks.test
+    allptsdata[[pt]] = list('seg'=segdata, 'tile'=tiled, 'tileNoUV'=tiled.noUV, 'tile.arms'=tiled.arms, 'tile.arms.noUV'=tiled.armsNoUV, 'arrayDf'=arrayDf, 'cx'=cx)
+  }  
+  if(!is.null(outfile))   save(wholeepi, predictions, file=outfile)
+
+return(predictions)
+}
+
+load('~/Data/Ellie/Analysis/5e6_arms_all_logR/all.pt.alpha.Rdata', verbose = T)
+fitV = models$`0.9`
+l = performance.at.1se$`0.9`$lambda
+rm(plots,performance.at.1se,dysplasia.df,models,cvs,labels)
+
+tmp = paste('/tmp/', sub('\\.txt','',basename(filelist[6])), '_epi.Rdata', sep='')
+if (file.exists(tmp)) {
+  load(tmp, verbose=T)
+} else {
+  predictions = data.frame(matrix(nrow=0,ncol=2,dimnames=list(c(), c('Prob','RR'))))
+  wholeepi = list()
+
+  infocols = grep('Chr|Pos', colnames(epi), value=T)
+  pts = grep('SNP|Chr|Pos', colnames(epi), invert=T, value=T)
+  pt = pts[1]
+
+  for (pt in pts)  {
+    print(pt)    
+    data = epi[,c(infocols,pt)]
+    segdata = copynumber::pcf(as.data.frame(data), gamma=28, fast=T, verbose=F)  
+    head(segdata)
+
+    #ggplot(segdata, aes(x=1:nrow(segdata), y=mean)) + geom_point() + labs(title='LogR whole epi', x='', y='')
+
+    # this is the LogR, doesn't need to be logged again
+    tiled = segment.matrix(tile.segmented.data(segdata[,-1], chr.info=get.chr.lengths()))
+    tiled.noUV = tiled
+
+    for (i in 1:ncol(tiled)) 
+      tiled[,i] = unit.var(tiled[,i], z.mean[i], z.sd[i]) 
+    tiled[is.na(tiled)] = mean(tiled, na.rm=T) #0
+
+    cx = score.cx(tiled, 1)
+
+    tiled.arms = segment.matrix(tile.segmented.data(segdata[,-1], size='arms', chr.info=get.chr.lengths()))
+    tiled.armsNoUV = tiled.arms
+
+    for (i in 1:ncol(tiled.arms)) 
+      tiled.arms[,i] = unit.var(tiled.arms[,i], z.arms.mean[i], z.arms.sd[i]) 
+    tiled.arms[is.na(tiled.arms)] = mean(tiled, na.rm=T) #0
+
+    arrayDf = subtract.arms(tiled, tiled.arms)
+    arrayDf = cbind(arrayDf, 'cx'=unit.var(cx, mn.cx ,sd.cx))
+
+    # Whole epi
+    predictions[pt,] = c(predict(fitV, newx=arrayDf, s=l, type='response')[,1], 'RR'=predict(fitV, newx=arrayDf, s=l, type='link')[,1] )
+
+    wholeepi[[pt]] = list('seg'=segdata, 'tile'=tiled, 'tileNoUV'=tiled.noUV, 'tile.arms'=tiled.arms, 'tile.arms.noUV'=tiled.armsNoUV, 'arrayDf'=arrayDf, 'cx'=cx)
+    }
+
+  save(wholeepi, predictions, file=tmp)
+}
 
 ```
+
+Normal sample?
+```{r, warning=F, message=F}
+
+npg = '911_5U__normal_10397N_200ng'
+pg = '848_85O_SS1_normal_18780N_200ng'
+
+tmp = '/tmp/normal.Rdata'
+if (file.exists(tmp)) {
+  load(tmp, verbose = T)
+} else {
+
+  # this is the LogR, doesn't need to be logged again
+  normals = list()
+  for (pt in c(npg,pg)) {
+    data = normpt[[pg]][,c('chr','position','Log.R')]
+    colnames(data)[3] = pt
+    data$chr = as.numeric(data$chr)
+    segdata = copynumber::pcf(as.data.frame(subset(data, chr %in% c(1:22))), gamma=28, fast=T, verbose=F)
+
+    tiledN = tile.segmented.data(segdata[,-1], chr.info=get.chr.lengths())
+    tiledN = segment.matrix(tiledN)
+
+    tiledN.noUV = tiledN
+
+    for (i in 1:ncol(tiledN)) 
+      tiledN[,i] = unit.var(tiledN[,i], z.mean[i], z.sd[i]) 
+    tiledN[is.na(tiledN)] = mean(tiledN, na.rm=T) #0
+
+    cx = score.cx(tiledN, 1)
+
+    tiled.armsN = tile.segmented.data(segdata[,-1], size='arms', chr.info=get.chr.lengths())
+    tiled.armsN = segment.matrix(tiled.armsN)
+
+    tiled.armsN.noUV = tiled.armsN
+
+    for (i in 1:ncol(tiled.armsN)) 
+      tiled.armsN[,i] = unit.var(tiled.armsN[,i], z.arms.mean[i], z.arms.sd[i]) 
+    tiled.armsN[is.na(tiled.armsN)] = mean(tiled.armsN, na.rm=T) #0
+
+    arrayDfN = subtract.arms(tiledN, tiled.armsN)
+    arrayDfN = cbind(arrayDfN, 'cx'=unit.var(cx, mn.cx ,sd.cx))
+
+    normals[[pt]] = list('seg'=segdata, 'tile'=tiledN, 'tileNoUV'=tiledN.noUV, 'tile.arms'=tiled.armsN, 'tile.arms.noUV'=tiled.armsNoUV, 'arrayDf'=arrayDfN, 'cx'=cx)
+  }
+  save(normals, file=tmp)
+}
+
+#normseg = segdata
+
+```
+
+Norm vs whole ep
+The extreme tail on the normal sample is due to the 'cx' coefficient and needs to be adjusted.  For now I will set it to the mean of the cx value in the sWGS dataset.
+```{r}
+
+head(wholeepseg)
+
+grid.arrange(
+  ggplot(wholeepseg, aes(x=1:nrow(wholeepseg), y=mean)) + geom_point() + labs(title='Whole epi, segmented LogR'),
+  ggplot(normseg, aes(x=1:nrow(normseg), y=mean)) + geom_point() + labs(title='Normal, segmented LogR'))
+
+grid.arrange(
+  ggplot(wholeepseg, aes(mean)) + geom_histogram(color='lightblue', bins=25) + labs(title='Array LogR, Segmented', subtitle='Whole epi'),
+  ggplot(normseg, aes(mean)) + geom_histogram(color='lightblue', bins=25) + labs(title='Array LogR, Segmented', subtitle='Normal sample')
+)
+
+summary(wholeepseg$mean)
+summary(normseg$mean)
+
+grid.arrange(
+  ggplot(melt(arrayDf), aes(value)) + geom_histogram(color='lightblue', bins=25) + labs(title='Array LogR, tiled', subtitle='Whole epi, 5e6'),
+  ggplot(melt(arrayDfN), aes(value)) + geom_histogram(color='lightblue', bins=25) + labs(title='Array LogR, tiled', subtitle='Normal sample, 5e6')
+)
+
+arrayDf[,'cx'] = mean(allDf[,'cx'])
+arrayDfN[,'cx'] = mean(allDf[,'cx'])
+
+```
+
+Compare to all samples.
+```{r warning=F, message=F}
+
+ggplot(melt(allDf), aes(value)) + geom_histogram(color='lightblue', bins=25) + labs(title='Array LogR, Segmented', subtitle='all sWGS')
+
+ggplot(melt(allDf[sample(1:nrow(allDf),1),]), aes(value)) + geom_histogram(color='lightblue', bins=25) + labs(title='Array LogR, Segmented', subtitle='Random sWGS sample')
+
+grid.arrange(
+  ggplot(melt(arrayDf), aes(value)) + geom_histogram(color='lightblue', bins=25) + labs(title='Array LogR, Segmented', subtitle='Whole epi'),
+  ggplot(melt(arrayDfN), aes(value)) + geom_histogram(color='lightblue', bins=25) + labs(title='Array LogR, Segmented', subtitle='Normal sample'))
+
+s = which(apply(allDf, 1, function(i) ks.test(i, arrayDf)$p.value) >= 0.05)
+
+# Normal
+predict(fitV, newx=arrayDfN, s=l, type='response')
+predict(fitV, newx=arrayDfN, s=l, type='link')
+
+```
+
+Quantile normalize?
+```{r, warning=F, message=F}
+by = 3
+mx = mean(arrayDf, na.rm=T)+sd(arrayDf, na.rm=T)*by
+mn = mean(arrayDf, na.rm=T)-sd(arrayDf, na.rm=T)*by
+
+ggplot(melt(arrayDf), aes(value)) + 
+  geom_histogram(color='lightblue', bins=25) + labs(title='Array LogR, Segmented', subtitle='Whole epi?') +
+  geom_vline(xintercept = c(mx,mn), color='red')
+
+## Get the distribution as below across ALL methylation samples then sub the mean and sd for ALL instead of for one
+
+newsd = sd(arrayDf[arrayDf <= mx & arrayDf >= mn], na.rm=T)
+newmean = mean(arrayDf[arrayDf <= mx & arrayDf >= mn], na.rm=T)
+
+#xx = qnorm(pnorm(segMethy, mean(segMethy), sd(segMethy)), mean(allDf[,1:ncol(segMethy)]), sd(allDf[,1:ncol(segMethy)]))
+#hist(xx)
+
+xx = qnorm(pnorm(arrayDf, newmean, newsd, lower.tail=F), mean(allDf[,1:ncol(arrayDf)]), sd(allDf[,1:ncol(arrayDf)]), lower.tail=F)
+
+# less extreme, but still incorrect
+xx = qnorm(pnorm(arrayDf, newmean, newsd, lower.tail=F), median(apply(allDf, 1, mean, na.rm=T)), sd(apply(allDf, 1, sd, na.rm=T)), lower.tail=F)
+
+hist(xx)
+
+predict(fitV, newx=xx, s=l, type='response')
+predict(fitV, newx=xx, s=l, type='link')
+```
+
+## Try processing the IDAT files directly.  Perhaps if they aren't normalized together the LogR becomes more comparable.
+```{r}
+
+library(illuminaio)
+
+file = '~/Data/Ellie/arrays/SNP/GSM2643883_8832987072_R03C01_Grn.idat'
+idat <- readIDAT(file)
+names(idat)
+
+```

### 2018-05-22 process_snp.R

@@ -1,0 +1,123 @@
+
+args = commandArgs(trailingOnly=TRUE)
+
+if (length(args) < 3)
+  stop("Missing required params: <file in> <model dir> <output dir>")
+
+library(copynumber)
+library(tibble)
+library(dplyr)
+#library(ggplot2)
+#library(gridExtra)
+library(glmnet)
+
+source('lib/data_func.R')
+
+fileIn = args[1]
+modelDir = args[2]
+outdir = args[3]
+
+dir.create(outdir, showWarnings=F, recursive=T)
+
+load(paste(modelDir, '/model_data.Rdata', sep=''), verbose=T)
+rm(dysplasia.df, labels)
+load(paste(modelDir,'/all.pt.alpha.Rdata', sep=''), verbose = T)
+fitV = models$`0.9`
+l = performance.at.1se$`0.9`$lambda
+rm(plots,performance.at.1se,dysplasia.df,models,cvs,labels)
+
+
+#filelist = list.files('~/Data/Ellie/arrays/SNP', pattern='txt', full.names=T)
+samplelist = colnames(as_tibble(data.table::fread(fileIn, nrows=2)))
+
+whole_epi = grep('*whole.*Log', samplelist, value=T)  
+normals = grep('*normal.*Log', samplelist, value=T)  
+
+
+fix.chr<- function(df) {
+  df = subset(df, Chr %in% c(1:22))
+  df$Chr = as.numeric(df$Chr)
+  dplyr::arrange(df, Chr, Position)
+}
+
+process.patients<-function(df, outfile=NULL) {
+  predictions = data.frame(matrix(nrow=0,ncol=2,dimnames=list(c(), c('Prob','RR'))))
+  allptsdata = list()
+  
+  infocols = grep('Chr|Pos', colnames(df), value=T)
+  pts = grep('SNP|Chr|Pos', colnames(df), invert=T, value=T)
+  
+  chr.len = get.chr.lengths()
+  
+  for (pt in pts)  {
+    print(pt)    
+    data = df[,c(infocols,pt)]
+    segdata = copynumber::pcf(as.data.frame(data), gamma=28, fast=T, verbose=F)  
+    #head(segdata)
+    
+    #ggplot(segdata, aes(x=1:nrow(segdata), y=mean)) + geom_point() + labs(title='LogR whole epi', x='', y='')
+    
+    # this is the LogR, doesn't need to be logged again
[…]

### 2018-05-22 processing_pipeline/predict_downsampled.R

@@ -1,11 +1,16 @@
 # predict downsampled
+args = commandArgs(trailingOnly=TRUE)
+
+if (length(args) < 1)
+  stop("Missing required params: <data dir> <model dir> <out dir> ")
+
 
 library(ggplot2)
 library(Matrix)
 library(glmnet)
 
-source('lib/data_func.R')
-source('lib/load_patient_metadata.R')
+source('../lib/data_func.R')
+source('../lib/load_patient_metadata.R')
 
 chrlen = get.chr.lengths()
 
@@ -19,16 +24,24 @@ adjustRisk <- function(RR, offset, type='risk') {
 }
 
 
-dir = '~/Data/Ellie/Cleaned/Downsampled'
-#dir = '~/Data/Ellie/Cleaned/AH0329_segmentedCoverage_fitted_gamma250/'
+dir = args[1]
+modeldir = args[2]
+#outdir = args[2]
+
+logT = F
+if (length(args) > 2) logT = as.logical(args[3])
+
+#dir = '~/Data/Ellie/Analysis/VAL_Cohort'
+#modeldir = '~/Data/Ellie/Analysis/5e6_arms_all_logR'
+
 #dir = '~/Data/Ellie/Analysis/VAL_Cohort/'
 files = list.files(dir, full.names=T)
 
-file = '~/Data/Ellie/Analysis/5e6_arms_all_exAHM0320/model_data.Rdata'
+file = paste(modeldir, 'model_data.Rdata', sep='/')
 if (!file.exists(file))
   stop(paste("Missing data file", file))
 load(file, verbose=T)
-file = '~/Data/Ellie/Analysis/5e6_arms_all_exAHM0320/all.pt.alpha.Rdata'
+file = paste(modeldir, 'all.pt.alpha.Rdata', sep='/')
 if (!file.exists(file))
   stop(paste("Missing data file", file))
 load(file, verbose=T)
@@ -39,7 +52,7 @@ fitV = models[[select.alpha]]
 lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']
 
 
-file = '~/Data/Ellie/Analysis/5e6_arms_all_exAHM0320/loo.Rdata'
+file = paste(modeldir, 'loo.Rdata', sep='/')
 if (!file.exists(file))
   stop(paste("Missing data file", file))
 load(file, verbose=T)
[…]

### 2018-06-13 SNP_sWGS_comparison.Rmd
adding script to process the SNP array data from Reid et.al.

@@
 library(glmnet)
+library(reshape2)
+
+winsor2<-function (x, multiple=3) {
+   if(length(multiple) != 1 || multiple <= 0) {
+      stop("bad value for 'multiple'")
+   }
+   med <- median(x)
+   y <- x - med
+   sc <- mad(y, center=0) * multiple
+   y[ y > sc ] <- sc
+   y[ y < -sc ] <- -sc
+   y + med
+}
 
 suppressPackageStartupMessages( source('lib/data_func.R') )
+load('~/Data/Ellie/Analysis/5e6_arms_all_logR/model_data.Rdata', verbose = T)
+load('~/Data/Ellie/Analysis/5e6_arms_all_logR/all.pt.alpha.Rdata', verbose = T)
+fitV = models$`0.9`
+l = performance.at.1se$`0.9`$lambda
+rm(plots,performance.at.1se,dysplasia.df,models,cvs,labels)
+
 ```
@@
 
+chr.info = get.chr.lengths()[1:22,]
+chr.info$chrom = sub('chr', '', chr.info$chrom)
+chr.info$chrom = factor(chr.info$chrom, ordered = T, levels=c(1:22))
+relCNV$chrom = factor(relCNV$chrom, ordered = T, levels = c(1:22))
+(segCNV['D708_D504_9242',])
+
+windows = apply(chr.info, 1, function(x) {
+  m = data.frame(matrix(seq(from=5e6,to=x[['chr.length']],by=5e6),ncol=1))
+  cbind(x[['chrom']], m)
+})
+names(windows) = c(1:22)
+windows = do.call(rbind,windows)
+colnames(windows) = c('chrom','window')
+
+# tiles = tile.segmented.data(relCNV, chr.info=chr.info, verbose=T)
+# tiles.arms = tile.segmented.data(relCNV, size='arms',chr.info=chr.info, verbose=T)
+
+# ggplot(subset(chr.info, chrom == 2), aes(x=1:chr.length)) + #facet_grid(~chrom,space = 'free_x', scales = 'free_x') + 
+#   labs(x='Chromosome 2', y='Segmented values') +
+#   geom_segment(data=subset(relCNV[,c('chrom','start.pos','end.pos','D708_D504_9242')], chrom==2), aes(x=start.pos, xend=end.pos, y=D708_D504_9242, yend=D708_D504_9242), size=3) +
+#   geom_vline(data=subset(windows, chrom==2), aes(xintercept=window), linetype='dashed', color='grey39') +
+#   geom_segment(data=subset(tiles, chr == 2), aes(x=start, xend=end, y=D708_D504_9242, yend=D708_D504_9242), color='red', size=2) +
+#   geom_segment(data=subset(tiles.arms, chr == 2), aes(x=start, xend=end, y=D708_D504_9242, yend=D708_D504_9242), color='blue', size=2)
+
+# IMC
 grid.arrange(  
-  ggplot(relCNV, aes(x=1:nrow(relCNV), y=D702_D506_10725_10729)) + geom_point(color='darkblue') + labs(title='Raw', x=''),
-  ggplot(melt(segCNV['D702_D506_10725_10729',]), aes(x=1:ncol(segCNV), y=value)) + geom_point(color='darkblue') + labs(title='Tiled & Logged', y='D702_D506_10725_10729 (IMC)', x='') )
-
-grid.arrange(
-  ggplot(relCNV, aes(x=1:nrow(relCNV), y=D706_D507_9242)) + geom_point(color='darkblue') + labs(title='Raw', x=''),
-  ggplot(melt(segCNV['D706_D507_9242',]), aes(x=1:ncol(segCNV), y=value)) + geom_point(color='darkblue') + labs(title='Tiled & Logged', y='D706_D507_9242 (NDBE)', x='') )
-
-grid.arrange(  
-  ggplot(relCNV, aes(D702_D506_10725_10729)) + geom_histogram(color='lightblue', bins=25) + labs(title='Raw'),
-  ggplot(melt(unlogged['D702_D506_10725_10729',]), aes(value)) + geom_histogram(color='lightblue', bins=25) + labs(title='Segmented'),
-  ggplot(melt(segCNV['D702_D506_10725_10729',]), aes(value)) + geom_histogram(color='lightblue', bins=25) + labs(title='Logged & Segmented'), top='IMC'
+  ggplot(relCNV, aes(D708_D504_9242)) + geom_histogram(color='lightblue', bins=25) + labs(title='Raw'),
+  ggplot(melt(unlogged['D708_D504_9242',]), aes(value)) + geom_histogram(color='lightblue', bins=25) + labs(title='Segmented'),
+  ggplot(melt(segCNV['D708_D504_9242',]), aes(value)) + geom_histogram(color='lightblue', bins=25) + labs(title='Logged & Segmented'), top='IMC'
 )
 
+qqnorm(segCNV['D708_D504_9242',])
+qqnorm(unlogged['D708_D504_9242',])
+qqnorm(relCNV[,'D708_D504_9242']) # raw
+
+# Early NDBE
 grid.arrange(  
@@
 
-# All samples
-load('~/Data/Ellie/Analysis/5e6_arms_all_logR/model_data.Rdata', verbose=T)
-load('~/Data/Ellie/Analysis/5e6_arms_all_logR/loo.Rdata', verbose=T)
+qqnorm(segCNV['D706_D507_9242',])
+qqline(segCNV['D706_D507_9242',])
+
+qqnorm(unlogged['D706_D507_9242',])
+qqnorm(relCNV[,'D706_D507_9242'])
+
 subset(pg.samp$AH0254, Samplename %in% c('D706_D507_9242','D702_D506_10725_10729'))
@@
 GSE99429 LogR data
-```{r, warning=F, message=F}
+```{r, warning=F, message=F, eval=F}
 filelist = list.files('~/Data/Ellie/arrays/SNP', pattern='txt', full.names=T)
-samplelist = lapply(filelist, function(f) colnames(as_tibble(data.table::fread(f, nrows=2))))
 
-whole_epi = lapply(samplelist, function(s) grep('*whole.*Log', s, value=T)  )
-names(whole_epi) = basename(filelist)
+whole_epiP = as_tibble(data.table::fread(grep('391_62K.*whole_epi', filelist, value=T)[1], sep='\t', header=T))
+#normalP = as_tibble(data.table::fread(grep('391.*normal.*200ng', filelist, value=T), sep='\t', header=T))
 
-normals = lapply(samplelist, function(s) grep('*normal.*Log', s, value=T)  )
-names(normals) = basename(filelist)
+#whole_epiNP = as_tibble(data.table::fread(grep('451_115T.*whole_epi', filelist, value=T)[1], sep='\t', header=T))
+#normalNP = as_tibble(data.table::fread(grep('451.*normal.*200ng', filelist, value=T), sep='\t', header=T))
 
@@
 
-dt = as_tibble(data.table::fread(filelist[6]))
-epi = dt[,c(grep('SNP|Chr|Pos', colnames(dt), value=T),whole_epi[[6]])]
-nm = dt[,c(grep('SNP|Chr|Pos', colnames(dt), value=T),normals[[6]])]
+whole_epiP = fix.chr(whole_epiP)
+normalP = fix.chr(normalP)
+qqnorm(whole_epiP$`391_62K_SS4_whole_epi__22325.Log.R.Ratio`)
+#qqline(whole_epiP$`391_62K_SS4_whole_epi__22325.Log.R.Ratio`)
 
-epi = fix.chr(epi)
-nm = fix.chr(nm)
-rm(dt)
+head(whole_epiP)
+
+chrs = get.chr.lengths()[1:22,]
+chrs$chrom = factor(sub('chr','',chrs$chrom), ordered = T, levels=c(1:22))
+colnames(chrs)[1] = 'Chr'
+
+whole_epiP$Chr = factor(whole_epiP$Chr, levels=c(1:22), ordered = T)
+
+p = ggplot(chrs, aes(x=1:chr.length)) + facet_grid(~Chr, space='free_x', scales='free_x') +
+  geom_point(data=whole_epiP, aes(Position, `391_62K_SS4_whole_epi__22325.Log.R.Ratio`))
+
+#qqnorm(whole_epiP$`391_62K_SS4_whole_epi__22325.Log.R.Ratio`)
+#epiW = winsorize(as.data.frame(whole_epiP[,c(1:3)]))
+#head(epiW)
+
+whole_epiP$Chr = as.integer(as.character(whole_epiP$Chr))
+#nas = which(is.na(whole_epiP$`391_62K_SS4_whole_epi__22325.Log.R.Ratio`))
+
+seg = pcf(as.data.frame(whole_epiP[,c(1:3)]), gamma=37, verbose=T)
+head(seg)
+
+x = which(seg$chrom == 17)[1]
+
+baf = whole_epiP[,c(1,2,4)]
+summary(subset(baf, Position >= seg$start.pos[x] & Position <= seg$end.pos[x] & Chr == seg$chrom[x])[,3,drop=T], na.rm=T)
+
+summary(seg$end.pos-seg$start.pos)
+summary(seg$mean)
+colnames(seg)[2] = 'Chr'
+seg$Chr = factor(seg$Chr, levels=levels(chrs$Chr), ordered = T)
+head(seg)
+p + geom_segment(data=seg, aes(x=start.pos,xend=end.pos, y=mean, yend=mean), color='red', size=2)
+
+#hist(seg$mean)
+
+seg$meanW = NA
+seg$meanW[!is.na(seg$mean)] = winsor2(seg$mean[!is.na(seg$mean)]) # winsorize w/replacement
+
+tiled = tile.segmented.data(seg[,c(2:6,8)], chr.info = get.chr.lengths(), verbose = T)
+head(tiled)
+hist(tiled$meanW)
+# colnames(tiled)[1] = 'Chr'
+# tiled$Chr = factor(tiled$Chr, ordered = T)
+# p + geom_segment(data=tiled, aes(x=start,xend=end, y=mean, yend=mean), color='red', size=2)
+
+colnames(tiled)[4] = 'mean'
+p1 = ggplot(tiled) + geom_histogram(aes(mean), col='lightblue', bins=20) + 
+  geom_vline(xintercept=c(mean(tiled$mean,na.rm=T)+sd(tiled$mean,na.rm=T), mean(tiled$mean,na.rm=T)-sd(tiled$mean,na.rm=T)), col='red' )  
+tiled$scaled = scale(exp(tiled$mean))
+p2 = ggplot(tiled) + geom_histogram(aes(scaled), col='lightblue', bins=20) +
+    geom_vline(xintercept=c(mean(tiled$scaled,na.rm=T)+sd(tiled$scaled,na.rm=T), mean(tiled$scaled,na.rm=T)-sd(tiled$scaled,na.rm=T)), col='red' )  
+grid.arrange(p1, p2)
+
+qqnorm(tiled$scaled)
+hist(tiled$scaled)
+
+#x = tiled
+
+tiled = segment.matrix(tiled)
+# for (i in 1:ncol(tiled)) 
+#   tiled[,i] = unit.var(tiled[,i], z.mean[i], z.sd[i]) 
+tiled[is.na(tiled)] = mean(tiled, na.rm=T) #0
+
+cx = score.cx(tiled, 1)
+
+tiled.arms = tile.segmented.data(seg[,c(2:6,8)], size='arms', chr.info=get.chr.lengths(), verbose=T)
+colnames(tiled.arms)[4] = 'mean'
+tiled.arms$scaled = scale(exp(tiled.arms$mean))
+#tiled.arms$mean = scale(exp(tiled$mean))
+tiled.arms = segment.matrix(tiled.arms)
+tiled.arms[is.na(tiled.arms)] = mean(tiled.arms, na.rm=T) #0
+
+arrayDf = subtract.arms(tiled, tiled.arms)
+arrayDf = cbind(arrayDf, 'cx'=unit.var(cx, mn.cx ,sd.cx))
+
+predict(fitV, newx=arrayDf, s=l, type='response')
+predict(fitV, newx=arrayDf, s=l, type='link')
+
+summary(arrayDf[1,])
+
+hist(allDf[210,], breaks=20)
+
+# logr = whole_epiP$`391_62K_SS4_whole_epi__22325.Log.R.Ratio`
+# hist(logr)
+# by = 3
+# mx = mean(logr,na.rm=T)+sd(logr,na.rm=T)*by
+# mn = mean(logr,na.rm=T)-sd(logr,na.rm=T)*by
+# abline(v=c(mx, mn), col='red')
+# 
+# newsd = sd(logr[logr <= mx & logr >= mn], na.rm=T)
+# newmean = mean(logr[logr <= mx & logr >= mn], na.rm=T)
+
+#xx = qnorm(pnorm(segMethy, mean(segMethy), sd(segMethy)), mean(allDf[,1:ncol(segMethy)]), sd(allDf[,1:ncol(segMethy)]))
+#hist(xx)
+
+# xx = qnorm(pnorm(logr, newmean, newsd, lower.tail=F))
+# xx[xx==Inf] = NA
+# xx[xx==-Inf] = NA
+# hist(xx)
+# qqnorm(xx)
+
+epiW = winsorize(as.data.frame(whole_epiP[,c(1:3)]))
+qqnorm(epiW$`391_62K_SS4_whole_epi__22325.Log.R.Ratio`)
+
+logrw = epiW$`391_62K_SS4_whole_epi__22325.Log.R.Ratio`
+hist(logrw)
+by = 3
+mx = mean(logrw,na.rm=T)+sd(logrw,na.rm=T)*by
+mn = mean(logrw,na.rm=T)-sd(logrw,na.rm=T)*by
+abline(v=c(mx, mn), col='red')
+
+newsd = sd(logrw[logrw <= mx & logrw >= mn], na.rm=T)
+newmean = mean(logr[logrw <= mx & logrw >= mn], na.rm=T)
+
+#xx = qnorm(pnorm(segMethy, mean(segMethy), sd(segMethy)), mean(allDf[,1:ncol(segMethy)]), sd(allDf[,1:ncol(segMethy)]))
+#hist(xx)
+
+xy = qnorm(pnorm(logrw, newmean, newsd, lower.tail=F))
+xy[xy==Inf] = NA
+xy[xy==-Inf] = NA
+hist(xy)
+qqnorm(xy)
+
+length(xy)
+
+rows = which(logrw <= mx & logrw >= mn)
+head(epiW[rows,])
+head(xy[rows])
+
+epiW = cbind(epiW, xy)
+
+seg = pcf(epiW[,c(1:3)], gamma=40, fast=T, verbose=F)
+qqnorm(seg$mean)
+hist(seg$mean)
+by = 3
+mx = mean(seg$mean,na.rm=T)+sd(seg$mean,na.rm=T)*by
+mn = mean(seg$mean,na.rm=T)-sd(seg$mean,na.rm=T)*by
+abline(v=c(mx, mn), col='red')
+
+newsd = sd(seg$mean[ seg$mean >= mn], na.rm=T)   
+newmean = mean(seg$mean[ seg$mean >= mn], na.rm=T)
+
+xy = qnorm(pnorm(seg$mean, newmean, newsd, lower.tail=F))
+xy[xy==Inf] = NA
+xy[xy==-Inf] = NA
+hist(xy)
+qqnorm(xy)
+
+summary(seg$end.pos-seg$start.pos)
+
+chrs = get.chr.lengths()
+chrs$chrom = sub('chr','',chrs$chrom)
+
+sc = base::merge(seg, chrs, by='chrom')
+head(sc)
+
+unseg = base::merge(epiW, chrs, by='chrom')
+head(unseg)
+unseg = cbind(unseg, xy)
+
+unseg = unseg %>% rowwise() %>% dplyr::mutate(g.pos = ifelse(chrom == '1', pos, genome.length-pos))
+
+p = ggplot(unseg) + xlim(1, max(chrs$genome.length)) + 
+  geom_point(aes(x=g.pos, y=xy), color='red', alpha=0.5) +
+  geom_point(aes(x=g.pos, y=`391_62K_SS4_whole_epi__22325.Log.R.Ratio`)) 
+ggsave(p, width=10, height = 7, units = 'in', file='snps.png')
+
+sc = sc %>% rowwise() %>% dplyr::mutate(g.start = ifelse(chrom == '1', start.pos, genome.length-start.pos), 
+                                        g.end = ifelse(chrom == '1', end.pos, genome.length-end.pos))
+
+p + geom_segment(data=sc, aes(x=g.start, xend=g.end, y=mean, yend=mean), color='red', size=2)
+
+ggplot() + lims(x=c(1,max(chrs))) + 
+  geom_point(data=epiW)
+
+ggplot(sc) + facet_grid(~chrom, space = 'free_x', scales = 'free_x') + geom_segment(aes(x=start.pos, xend=end.pos, y=mean, yend=mean))
+
+#x = imputeMissing(epiW, 'constant')
+#asseg = aspcf(x, epibaf, gamma=40, kmin=64)
+
+#segdata = copynumber::pcf(as.data.frame(data), gamma=28, fast=T, verbose=F)
 
@@
 
-```{r, warning=F, message=F}
+```{r, warning=F, message=F, eval=F}
 
@@
 Normal sample?
-```{r, warning=F, message=F}
+```{r, warning=F, message=F, eval=F}
 
@@
 The extreme tail on the normal sample is due to the 'cx' coefficient and needs to be adjusted.  For now I will set it to the mean of the cx value in the sWGS dataset.
-```{r}
+```{r, eval=F}
 
@@
 Compare to all samples.
-```{r warning=F, message=F}
+```{r warning=F, message=F, eval=F}
 
@@
 Quantile normalize?
-```{r, warning=F, message=F}
+```{r, warning=F, message=F, eval=F}
 by = 3
@@
 
-## Try processing the IDAT files directly.  Perhaps if they aren't normalized together the LogR becomes more comparable.
-```{r}
+Try processing the IDAT files directly.  Perhaps if they aren't normalized together the LogR becomes more comparable?
 
-library(illuminaio)
+## Absolutely CN
 
-file = '~/Data/Ellie/arrays/SNP/GSM2643883_8832987072_R03C01_Grn.idat'
-idat <- readIDAT(file)
-names(idat)
+```{r, warning=F, message=F, echo=F}
+ploidy = read.table('~/Data/ICGC/BE_OAC_Cohort/tumour/data/ascat/LP2000104-DNA_A01_vs_LP2000101-DNA_A01/LP2000104-DNA_A01_vs_LP2000101-DNA_A01__pv.1.5__rg.grch37_g1k__al.bwa_mem__.samplestatistics.csv', row.names=1)['Ploidy',1]
+cn = data.table::fread('~/Data/ICGC/BE_OAC_Cohort/tumour/data/ascat/LP2000104-DNA_A01_vs_LP2000101-DNA_A01/LP2000104-DNA_A01_vs_LP2000101-DNA_A01__pv.1.5__rg.grch37_g1k__al.bwa_mem__.copynumber.caveman.csv')
+cn = subset(cn, Chromosome %in% c(1:22))
+
+cn = read.table('~/Data/ICGC/BE_OAC_Cohort/BA/data/ascat/totalCNV.txt', sep='\t', header=T)
+cn = subset(cn, Illumina.Barcode == 'LP2000110-DNA_A01_vs_LP2000101-DNA_A01')
+ploidy = cn$ploidy[1]
+nrow(cn)
+
+x = cnN$Total_CN-ploidy
+qqnorm(x)
+qqnorm(winsor2(x, 3))
+
+cn[which(cn$Chromosome == 17)[1],]
+x[which(cn$Chromosome == 17)[1]]
+winsor2(x, 3)[which(cn$Chromosome == 17)[1]]
+
+cn$TotalWinsAdj = winsor2(x)
+
+cols = grep('Chr|Start|End|TotalWinsAdj', colnames(cn), ignore.case=T)
+
+tiled = tile.segmented.data(as.data.frame(cn)[cols], chr.info=get.chr.lengths()[1:22,])
+tiled.arms = tile.segmented.data(as.data.frame(cn)[cols], size='arms', chr.info=get.chr.lengths()[1:22,])
+
+tt = segment.matrix(tiled)
+ta = segment.matrix(tiled.arms)
+
+tt[is.na(tt)] = mean(tt, na.rm=T) #0
+ta[is.na(ta)] = mean(ta, na.rm=T) #0
+
+cx = score.cx(tt, 1)
+
+dm = subtract.arms(tt, ta)
+
+dm = cbind(dm, 'cx'=unit.var(cx, mn.cx, sd.cx ))
+
+#unit.var(cx)
+predict(fitV, newx=dm, s=l, type='response')
+predict(fitV, newx=dm, s=l, type='link')
+
+## in silico normal sample
+cnN = cn
+cnN$Total_CN = sample(c(1,2,3), size=nrow(cnN), replace=T, prob=c(.025,.95,.025 ))
+hist(cnN$Total_CN)
+
+cnN$TotalWinsAdj = winsor2(cnN$Total_CN-round(ploidy))
+qqnorm(cnN$TotalWinsAdj)
+
+hist(cnN$TotalWinsAdj)
+
+cols = grep('Chr|Start|End|TotalWinsAdj', colnames(cn), ignore.case=T)
+
+tiled = tile.segmented.data(as.data.frame(cn)[cols], chr.info=get.chr.lengths()[1:22,])
+tiled.arms = tile.segmented.data(as.data.frame(cn)[cols], size='arms', chr.info=get.chr.lengths()[1:22,])
+
+tt = segment.matrix(tiled)
+ta = segment.matrix(tiled.arms)
+
+tt[is.na(tt)] = mean(tt, na.rm=T) #0
+ta[is.na(ta)] = mean(ta, na.rm=T) #0
+
+cx = score.cx(tt, 1)
+
+dm = subtract.arms(tt, ta)
+
+dm = cbind(dm, 'cx'=unit.var(cx, mn.cx, sd.cx ))
+
+#unit.var(cx)
+predict(fitV, newx=dm, s=l, type='response')
+predict(fitV, newx=dm, s=l, type='link')
+
+hist(allDf[1,])
 

### 2018-06-13 lib/data_func.R

@@ -112,9 +112,12 @@ prep.matrix<-function(dt, na.replace=0) {
 segment.matrix<-function(dt) {
   dt = as.data.frame(dt)
   
-  chrCol = grep('chr',colnames(dt))
-  startCol = grep('start',colnames(dt))
-  endCol = grep('end',colnames(dt))
+  chrCol = grep('chr',colnames(dt),ignore.case=T)
+  startCol = grep('start',colnames(dt),ignore.case=T)
+  endCol = grep('end',colnames(dt),ignore.case=T)
+  
+  if (length(chrCol)<=0 | length(startCol)<=0 | length(endCol)<=0)
+    stop("Missing columns chr|start|end")
   
   rows = paste(dt[,chrCol], ':', dt[,startCol], '-', dt[,endCol], sep='')
   cols = colnames(dt)[-c(chrCol, startCol, endCol)]
@@ -201,19 +204,19 @@ tile.segmented.data<-function(data, size=5e6, chr.info=NULL, verbose=F) {
   if (!is.tibble(data))
     data = as_tibble(data)
   
-  descCols = sort(union(grep('chr|arm|start|end|probes', colnames(data)), which(!sapply(data, is.numeric))))
+  descCols = sort(union(grep('chr|arm|start|end|probes', colnames(data), ignore.case=T), which(!sapply(data, is.numeric))))
   dataCols = c((descCols[length(descCols)]+1):ncol(data))
   
-  chrCol = grep('chr',colnames(data),value=T)
-  armCol = grep('arm',colnames(data),value=T)
-  startPos = grep('start',colnames(data),value=T)
-  endPos = grep('end',colnames(data),value=T)
+  chrCol = grep('chr',colnames(data),ignore.case=T, value=T)
+  armCol = grep('arm',colnames(data),ignore.case=T,value=T)
+  startPos = grep('start',colnames(data),ignore.case=T,value=T)
+  endPos = grep('end',colnames(data),ignore.case=T,value=T)
   
   data = data[which(!data[[chrCol]] %in% c('X','Y')),]
   x1 = data[,c(chrCol, startPos, endPos, colnames(data)[dataCols])]
   
   tiles = .tile.genome(size, chr.info, incGender=length(which(grepl('X|Y', unique(data[[chrCol]])))) > 0)
-  gr = makeGRangesFromDataFrame(x1, keep.extra.columns=T, start.field = startPos, end.field = endPos  ) 
[…]

### 2018-06-19 SNP_Prediction_Analysis.Rmd
...

---
title: "SNP Analysis"
author: "Sarah Killcoyne"
date: "6/13/2018"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(gridExtra)
library(glmnet)
library(pander)
library(readxl)
library(pROC)

source('lib/data_func.R')
source('lib/common_plots.R')

chr.info = get.chr.lengths(file='hg19_info.txt')

ptdirs = list.dirs('/Volumes/fh/fast/reid_b/collab/Killcoyne/Data/PerPatient', full.names=T, recursive=F)
patients = basename(ptdirs)

load('~/Data/Ellie/Analysis/5e6_arms_all_logR/model_data.Rdata', verbose = T)
load('~/Data/Ellie/Analysis/5e6_arms_all_logR/all.pt.alpha.Rdata', verbose = T)
fitV = models$`0.9`
l = performance.at.1se$`0.9`$lambda

swgs_labels = labels

rm(plots,performance.at.1se,models,cvs,labels)

pv = var(dysplasia.df[swgs_labels == 1])  # P
npv = var(dysplasia.df[swgs_labels == 0]) # NP

range(dysplasia.df[swgs_labels == 1])
range(dysplasia.df[swgs_labels == 0])

#load('/Volumes/fh/fast/reid_b/collab/Killcoyne/SNP_R/allpts_ascat.Rdata', verbose=T)
load('~/Data/Ellie/Analysis/SNP/allpts_ascat.Rdata', verbose=T)
#patient.info = as.data.frame(read_xlsx('/Volumes/fh/fast/reid_b/collab/Killcoyne/SNP_Project/metadata_T1T2.xlsx'))
patient.info = as.data.frame(read_xlsx('~/Data/Ellie/Analysis/SNP/metadata_T1T2.xlsx'))
patient.info$UniqueSampleID = paste(patient.info$PatientID, patient.info$`Timepoint Code`, sep='_')
patient.info$Path.Status = patient.info$Status
patient.info[patient.info$PatientID %in% subset(patient.info, Pathology %in% c('IMC','HGD'), select='PatientID')[,1], 'Path.Status'] = 'P'

sample.list = read_xlsx('/Volumes/fh/fast/reid_b/collab/Killcoyne/Data/20180604_Reid_1M_SampleList.xlsx', sheet = 2)
nrow(sample.list)
colnames(sample.list)[3] = 'Total.SCA'

sample.list$SCA.Ratio = sample.list$Total.SCA/max(chr.info$genome.length)

head(patient.info)

subset(sample.list, PatientID == 74)

sample.info  = do.call(rbind.data.frame, sapply(qcdata$Samplename, strsplit, '_'))
colnames(sample.info) = c('PatientID','SampleID','EndoID','Level')
sample.info[] = lapply(sample.info[], as.character)
sample.info[which(sample.info$PatientID == 524 & sample.info$Level == 524), 'Level'] = ''
sample.info$Samplename = qcdata$Samplename

message(paste(length(unique(sample.info$PatientID)), 'patients listed in metadata file'))
message(paste(nrow(qcdata), 'samples available'))

sample.info = base::merge(sample.info, patient.info[,c('PatientID','Timepoint','Timepoint Code','Status','Pathology')], by.x=c('PatientID','EndoID'), by.y=c('PatientID','Timepoint Code'))

head(sample.info)
head(sample.list)

get.ratio<-function(PID, SID) {
  subset(sample.list, PatientID == PID & SampleNum == SID)$SCA.Ratio
}

sample.info = sample.info %>% rowwise() %>% dplyr::mutate(
  SCA.Ratio = ifelse(Level %in% c('BLD','GASTRIC'), 0, get.ratio(PatientID, SampleID))
)

ggplot(sample.info, aes(SCA.Ratio)) + geom_histogram() + facet_grid(~Status)

```

In the sWGS data progressor samples show a variance of `r round(pv, 3)` and non-prog variance is `r round(npv,3)`

# Quick QC

Purity, ploidy and goodnessoffit from single sample (not paired) ASCAT.

```{r, echo=F, eval=T}
qcdata$SampleType = 'BE'
qcdata$SampleType[grep('BLD',rownames(qcdata), ignore.case=T)] = 'Blood Normal'
qcdata$SampleType[grep('gastric',rownames(qcdata), ignore.case=T)] = 'Gastric Normal'
head(qcdata)

qcdata = base::merge(qcdata, sample.info[,c('Samplename','Status','Timepoint','SCA.Ratio')], by='Samplename')

ggplot(qcdata, aes(round(Ploidy,2))) + geom_histogram() + 
  facet_wrap(~SampleType, scales='free_y', ncol=1) + labs(x='Ploidy', title='Ploidy, normal vs BE')

ggplot(qcdata, aes(Purity, Goodnessoffit, color=Status)) + geom_point() + facet_wrap(~SampleType, ncol=1) + 
  labs(title='Purity~fit, normal vs BE')

ggplot(qcdata, aes(SCA.Ratio)) + geom_histogram() + 
  facet_wrap(~Status, scales='free_y', ncol=1) + labs(x='SCA Ratio', title='SCA Ratio, normal vs BE')

ggplot(subset(qcdata, SampleType == 'BE'), aes(Purity, SCA.Ratio, color=Status)) + geom_point() + 
  geom_vline(xintercept = 0.95, color='red') + geom_hline(yintercept = 0.05, color='red')

ggplot(subset(qcdata, SampleType == 'BE'), aes(Ploidy, SCA.Ratio, color=Status)) + geom_point() + 
  geom_hline(yintercept = 0.05, color='red')
```

```{r}
## 43% of P samples have low/no CN with a ploidy of 2 or less
nrow(subset(qcdata, Status == 'P' & SampleType == 'BE' & SCA.Ratio <= 0.05 & Ploidy <= 2))/nrow(subset(qcdata, Status == 'P' & SampleType == 'BE'))

## <1% of P samples have no apparent CN
nrow(subset(qcdata, Status == 'P' & SCA.Ratio <= 0 & SampleType == 'BE'))/nrow(subset(qcdata, Status == 'P' & SampleType == 'BE'))

## 36% of P samples have < 5% CN and high purity ()
nrow(subset(qcdata, Status == 'P' &  Purity >= 0.95 & SCA.Ratio <= 0.05 & SampleType == 'BE'))/nrow(subset(qcdata, Status == 'P' & SampleType == 'BE'))

qcdata$SCA.Ratio = round(qcdata$SCA.Ratio, 4)

lowsca = subset(qcdata, Status == 'P' & SampleType == 'BE' & SCA.Ratio <= 0.05 & Ploidy <= 2)
write.table(lowsca, sep='\t', quote=F, row.names = F, file = '~/Data/Ellie/Analysis/SNP/low_sca_samples.txt')

totalSamples = table(sub('_.*','', qcdata$Samplename))
lowscaSamples = table(sub('_.*','', lowsca$Samplename))

# For those patients with low SCA samples, a large proportion of them show low SCA in most/all of their samples
hist(round(lowscaSamples/totalSamples[names(lowscaSamples)], 2))

```

# Load tiled data

These are the CN adjusted, winsorized LogR values tiled in 5Mb bins.

```{r, warning=F}
mergedSegs = NULL
mergedArms = NULL
length(ptdirs)

for (pt in ptdirs) {
  if (length(list.files(pt, '*wins_tiled.txt', full.names=T)) <= 0) {
    message(paste("No tiled files for",pt))
    next
  }

  segvals = as.data.frame(data.table::fread(list.files(pt, '*wins_tiled.txt', full.names=T)))
  armvals = as.data.frame(data.table::fread(list.files(pt, '*arms_tiled.txt', full.names=T)))

  segvals = segment.matrix(segvals)
  segvals[is.na(segvals)] = mean(segvals,na.rm=T)

  armvals = segment.matrix(armvals)
  armvals[is.na(armvals)] = mean(armvals,na.rm=T)

  if (is.null(segvals) | is.null(armvals))
    stop(paste("Missing values in ", pt))

  if (is.null(mergedSegs)) {
    mergedSegs = segvals
    mergedArms = armvals
  } else {
    mergedSegs = rbind(mergedSegs, segvals)    
    mergedArms = rbind(mergedArms, armvals)    
  }
}
nrow(mergedSegs) == nrow(mergedArms)
#setdiff(rownames(mergedSegs), rownames(mergedArms))
dim(mergedSegs)

# impute

# mergedSegs[is.na(mergedSegs)]
# mergedArms[is.na(mergedArms)]

# for (i in 1:nrow(mergedSegs)) 
#   mergedSegs[i,is.na(mergedSegs[i,])] = 0# mean(mergedSegs[i,],na.rm=T)
# 
# for (i in 1:nrow(mergedArms)) 
#   mergedArms[i,is.na(mergedArms[i,])] = 0 #mean(mergedArms[i,],na.rm=T)

```

# Values in the binned segments

Blood/gastric normals have very little variance compared to all of the BE samples (no idea which is P vs NP)

```{r normals, fig.width=6, fig.height=8}
rownames(mergedSegs) = sub('\\.LogR','',rownames(mergedSegs))
nm = melt(mergedSegs[grep('BLD|gastric', rownames(mergedSegs), ignore.case=T),])
be = melt(mergedSegs[grep('BLD|gastric', rownames(mergedSegs), ignore.case=T, invert=T),])
```

## Compare the binned values to the median LRR from ASCAT

Using a random normal patient and a known BE progresser patient

```{r rnormals, echo=F, fig.width=8, fig.height=8}
randNormal = sample(grep('BLD|gastric', rownames(mergedSegs), ignore.case=T, value=T), 1)
pt = unlist(strsplit(randNormal, '_'))[1]
segments.list[[pt]] = subset(segments.list[[pt]], chr %in% c(1:22))
segments.list[[pt]]$chr = factor(segments.list[[pt]]$chr, levels=c(1:22), ordered = T)

grid.arrange(
  ggplot(subset(segments.list[[pt]], sample == paste(randNormal,'.LogR',sep='')), aes(y=medLRR)) + 
    facet_grid(~chr, scales='free_x',space='free_x') +
    geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), size=2) + ylim(-1,1) +
    theme(axis.text.x=element_blank(), panel.spacing.x=unit(0, 'lines')) +
    labs(title='Median LRR, ASCAT segments', subtitle=pt), 
  ggplot(subset(nm, Var1 == randNormal), aes(Var2, value)) + geom_point() + ylim(-1,1) + 
    theme(axis.text.x=element_blank()) + labs(title='Winsorized LRR, binned'),  
top='Random BLD or gastric normal', bottom=randNormal)

randBE = '512_18762_133R_27.LogR'
#randBE = paste(sample(grep(paste('^',sample(subset(patient.info, Status == 'P')$PatientID, 1),sep=''), rownames(arrayDf), value=T), 1),'.LogR',sep='')
pt = unlist(strsplit(randBE, '_'))[1]
segments.list[[pt]] = subset(segments.list[[pt]], chr %in% c(1:22))
segments.list[[pt]]$chr = factor(segments.list[[pt]]$chr, levels=c(1:22), ordered = T)

grid.arrange(
  ggplot(subset(segments.list[[pt]], sample == randBE), aes(y=medLRR)) + 
    facet_grid(~chr, scales='free_x',space='free_x') +
    geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), size=2) + ylim(1,-1) +
    theme(axis.text.x=element_blank(), panel.spacing.x=unit(0, 'lines')) +
    labs(title='Median LRR, ASCAT segments', subtitle=pt), 
  ggplot(subset(be, Var1 == sub('\\.LogR','',randBE)), aes(Var2, value)) + geom_point() + ylim(-1,1) + 
    theme(axis.text.x=element_blank()) + labs(title='Winsorized LRR, binned'),  
top='BE', bottom=randBE)

```

# Set up values to predict

## Adjust per column

I adjust the variance of each genomic region by the mean and sd extracted when I trained on the sWGS data. This seems to be a more fair comparison than scaling all the new data to itself, but the predictions don't really change.
```{r, warning=F}
# uv 
# ms  = t(apply(mergedSegs, 1, function(sample) {
#   sapply(1:ncol(mergedSegs), function(i){
#     unit.var(sample[i], z.mean[i], z.sd[i])
#   })
# }))
# 
# range(ms)
ms = mergedSegs
for (i in 1:ncol(ms)) {
  ms[,i] = unit.var(ms[,i])
  #ms[,i] = unit.var(ms[,i], z.mean[i], z.sd[i])
}
range(ms)
dim(ms)
# ma  = t(apply(mergedArms, 1, function(sample) {
#   sapply(1:ncol(mergedArms), function(i){
#     unit.var(sample[i], z.arms.mean[i], z.arms.sd[i])
#   })
# }))

ma = mergedArms
for (i in 1:ncol(ma)) {
  ma[,i] = unit.var(ma[,i])
}
range(ma)
dim(ma)

nm = melt(ms[grep('BLD|gastric', rownames(ms), ignore.case=T),])
be = melt(ms[grep('BLD|gastric', rownames(ms), ignore.case=T, invert=T),])
```

The variance gets much bigger after unit normalizing, but the difference between BE and normal is still clear.

Normal variance: `r var(nm$value, na.rm=T)`

BE variance: `r var(be$value, na.rm=T)`

```{r echo=F}
cx = score.cx(ms, 1)
arrayDf = subtract.arms(ms, ma)
arrayDf = cbind(arrayDf, 'cx'=unit.var(cx, mn.cx, sd.cx))

#qqnorm(allDf, main='Normal Q-Q plot, sWGS')
#qqnorm(arrayDf, main='Normal Q-Q plot, SNP arrays')

```

# Predictions

```{r, warning=F, fig.height=8}
prob = predict(fitV, newx=arrayDf, s=l, type='response')
rr = predict(fitV, newx=arrayDf, s=l, type='link')

offset = log(0.0225)
preds = cbind.data.frame(rownames(prob),prob, rr)
colnames(preds) = c('Samplename','Prob', 'RR')

preds = preds %>% dplyr::mutate( 
  'Adj. RR'=RR+offset,  
  'Adj. Prob'=1/(1+exp(-RR+abs(offset)) )
)

myPal = rev(RColorBrewer::brewer.pal(11, 'RdYlBu'))
grid.arrange(
  ggplot(preds, aes(RR)) + geom_histogram(aes(fill=..x..), bins=15, show.legend = F) +
    scale_fill_gradientn(colors = myPal,  name='') + 
    labs(y='n Samples', x='Relative Risk', title='Unadjusted relative risk'),
  # ggplot(preds, aes(`Adj. RR`)) + geom_histogram(aes(fill=..x..), bins=15, show.legend = F) +
  #   scale_fill_gradientn(colors = myPal,  name='') + 
  #   labs(y='n Samples', x='Relative Risk', title='Adjusted relative risk'), 
top='Relative risk')

grid.arrange(
  ggplot(preds, aes(Prob)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
    scale_fill_gradientn(colors = myPal,  name='') + 
    labs(title='Predictions, all samples model', y='n Samples', x='Unadjusted Probability'),
  # ggplot(preds, aes(`Adj. Prob`)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
  #   scale_fill_gradientn(colors = myPal,  name='') + 
  #   labs(title='Predictions, all samples model', y='n Samples', x='Adjusted Probability'), 
top='Probability')

```

A quick look at the two random samples from earlier.  The random normal shows a very low probablity before and after adjustment:
`r pander(subset(preds, Samplename == randNormal))`

The random BE shows a very high probability:
`r pander(subset(preds, Samplename == sub('\\.LogR','',randBE)))`

```{r}
hist(subset(preds, Samplename %in% lowsca$Samplename)$Prob)

x = subset(segments.list$`891`, sample == '891_18574_100O_33.LogR') 
x$chr = factor(x$chr, levels=c(1:22,'X','Y'), ordered = T)
p1 = ggplot(x) + facet_grid(~chr, space='free_x', scales='free_x') + 
  geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), size=3) + labs(title='891_18574_100O_33 medLRR')

x = melt(ms['891_18574_100O_33',])
x$loc = rownames(x)
x = rbind(x,cbind.data.frame('loc'=paste(c(1:22), '0-1', sep=':'), 'value'=0))

locs = do.call(rbind.data.frame, strsplit(x$loc, ':|-'))
colnames(locs) = c('chr','start','end')
locs[] = lapply(locs[], function(y) as.numeric(as.character(y)))
x = cbind(x,locs)
x$chr = factor(x$chr, levels=c(1:22), ordered=T)
x = arrange(x,chr)

p2 = ggplot(x) + facet_grid(~chr, space='free_x', scales='free_x') + 
  geom_segment(aes(x=start, xend=end, y=value, yend=value), size=3) + labs(title='891_18574_100O_33 binned')

grid.arrange(p1,p2,nrow=2)
```

## Evaluating normals vs BE

```{r, warning=F, fig.height=8}
normals = subset(preds, grepl('BLD|gastric', Samplename, ignore.case=T))
grid.arrange(
  ggplot(normals, aes(Prob)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
    scale_fill_gradientn(colors = myPal,  name='') + 
    labs(title='Predictions on BLD/Gastric', y='n Samples', x='Unadjusted Probability'),
top='BLD/gastric normals' )

be = subset(preds, !grepl('BLD|gastric', Samplename, ignore.case=T))
grid.arrange(
  ggplot(be, aes(Prob)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
    scale_fill_gradientn(colors = myPal,  name='') + 
    labs(title='Predictions on BE', y='n Samples', x='Unadjusted Probability'),
top='BE samples' )

```

So how do these high normals look?
```{r}
preds
qcdata
nrow(preds)

pred.qc = base::merge(preds, qcdata, by='Samplename')

rc = Hmisc::rcorr( as.matrix(pred.qc[,c(2,3,6,7,8,12)]) )

GGally::ggcorr( as.matrix(pred.qc[,c(2,6,7,8,12)]) , label=T, palette='RdBu')

length(which(normals$Prob < 0.5))/nrow(normals)

highNmSamples = subset(normals, Prob >= 0.5)
subset(qcdata, Samplename %in% highNmSamples$Samplename)
# Low variance overall
mean(apply(mergedSegs[highNmSamples$Samplename,], 1, var))

# Most of the gastric "normals" show up, though that drops to 3% when adjusted
length(grep('gastric', highNmSamples$Samplename, ignore.case=T))/length(grep('gastric', rownames(mergedSegs), ignore.case=T))
length(grep('gastric', subset(normals, `Adj. Prob` >= 0.5)$Samplename, ignore.case=T))/length(grep('gastric', rownames(mergedSegs), ignore.case=T))

# About 8% of the blood normals do as well, but drops to <1% when adjusted
length(grep('BLD', highNmSamples$Samplename, ignore.case=T))/length(grep('BLD', rownames(mergedSegs), ignore.case=T))
length(grep('BLD', subset(normals, `Adj. Prob` >= 0.5)$Samplename, ignore.case=T))/length(grep('BLD', rownames(mergedSegs), ignore.case=T))
```

Compare BLD to BE
```{r, echo=F, fig.height=8}
bld = preds[grep('BLD', preds$Samplename, ignore.case=T),]
grid.arrange(
  ggplot(bld, aes(Prob)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(y='n Samples', x='Unadjusted Probability'),
ggplot(bld, aes(`Adj. Prob`)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(y='n Samples', x='Adjusted Probability'), top='Predictions on BLD samples only' )

length(which(bld$Prob < 0.5))/nrow(bld)
length(which(bld$`Adj. Prob` < 0.5))/nrow(bld)

be = preds[grep('BLD|gastric', rownames(preds), ignore.case=T, invert = T),]
grid.arrange(
  ggplot(be, aes(Prob)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
    scale_fill_gradientn(colors = myPal,  name='') + 
    labs(y='n Samples', x='Unadjusted Probability'),
  ggplot(be, aes(`Adj. Prob`)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
    scale_fill_gradientn(colors = myPal,  name='') + 
    labs(y='n Samples', x='Adjusted Probability'), top='Predictions on BE only'
)

unadj = length(unique(sapply(rownames(be[which(be$Prob > 0.5),]), function(x) unlist(strsplit(x, '_'))[1])))/length(patients)
adj = length(unique(sapply(rownames(be[which(be$`Adj. Prob` > 0.5),]), function(x) unlist(strsplit(x, '_'))[1])))/length(patients)

```

Unajdusted `r round(unadj,2)*100`% of BE patients have a >50% probability of progression.  Adjusted, `r round(adj, 2)*100`% have a >50% probability.

# Compare predictions to known status

```{r}

#preds.info = base::merge(sample.info, preds, by='Samplename')

rows = grep('BLD|gastric', pred.qc$Samplename, ignore.case=T)
normals = pred.qc[rows,]
pred.qc = pred.qc[-rows,]
nrow(pred.qc)

# Still missing some SNP data  FIX THIS CHECK
si = do.call(rbind.data.frame,c(strsplit(as.character(pred.qc$Samplename), '_'), stringsAsFactors=F) )
colnames(si) = colnames(sample.info)[1:4]

pred.qc = cbind(si, pred.qc)

setdiff(unique(patient.info$PatientID),unique(pred.qc$PatientID))

pred.qc$Status = factor(pred.qc$Status)

roc = pROC::roc(Status ~ Prob, data=pred.qc, auc=T, ci=T, of='thresholds')
roc.plot(roc)

rocNoLow = pROC::roc(Status ~ Prob, data=subset(pred.qc, !Samplename %in% lowsca$Samplename), auc=T, ci=T, of='thresholds')
roc.plot(rocNoLow)

head(pred.qc)

per_pt = pred.qc %>% group_by(PatientID, Timepoint, Status) %>% dplyr::summarise(
  max.Prob = max(Prob), max.RR = max(RR)
)

rocpp = pROC::roc(Status ~ max.Prob, data=per_pt, auc=T, ci=T, of='thresholds')
roc.plot(rocpp)

roct1 = pROC::roc(Status ~ Prob, data=subset(pred.qc, Timepoint == 'T1'), auc=T, ci=T, of='thresholds')
roc.plot(roct1)

roct2 = pROC::roc(Status ~ Prob, data=subset(pred.qc, Timepoint != 'T1'), auc=T, ci=T, of='thresholds')
roc.plot(roct2)

qcpred = subset(pred.qc, Samplename %in% subset(qcdata, Purity > 0.8 & Goodnessoffit >= 0.9)$Samplename)
qc = pROC::roc(Status ~ Prob, data=qcpred, auc=T, ci=T, of='thresholds')
roc.plot(qc)

```

## Regression using SNP Data

```{r, echo=F, eval=F}

sample.info = patient.info[c('PatientID','Status')]
sample.info = base::merge(sampleInfo, sample.info, by='PatientID', all.x=T)

sample.info = subset(sample.info, !Level %in% c('BLD','GASTRIC'))
sample.info$Samplename = with(sample.info, paste(PatientID, SampleID, EndoscopyID, Level, sep='_'))

df = arrayDf
arrayDf = arrayDf[grep('BLD|GASTRIC',rownames(df), invert=T),]
rownames(arrayDf) = sub('\\.LogR','',rownames(arrayDf))

patient.info
status = vector(length=nrow(arrayDf), mode='character')
names(status) = rownames(arrayDf)

for (i in 1:length(status))
  status[i] = unique(subset(patient.info, PatientID == sub('_.*','',names(status)[i]))[,'Status'])

labels = as.integer(factor(labels))-1
names(labels) = names(status)
rm(status)

source('lib/cv-pt-glm.R')

nl = 1000;folds = 10; splits = 5 
sets = create.patient.sets(sample.info[c('PatientID','Samplename','Status')], folds, splits, 0.2)

alpha.values = c(0,0.5,0.9,1)
## ----- All ----- ##
coefs = list(); plots = list(); performance.at.1se = list(); models = list(); cvs = list()
cache.dir = '~/Data/Ellie/Analysis/SNP/'
dir.create(cache.dir, recursive = T, showWarnings = F)
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  for (a in alpha.values) {
    fit0 <- glmnet(arrayDf, labels, alpha=a, nlambda=nl, family='binomial', standardize=F)    
    l = fit0$lambda

    cv.patient = crossvalidate.by.patient(x=arrayDf, y=labels, lambda=l, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', opt=-1, standardize=F)

    lambda.opt = cv.patient$lambda.1se

    coef.opt = as.data.frame(non.zero.coef(fit0, lambda.opt))
    coefs[[as.character(a)]] = coef.stability(coef.opt, cv.patient$non.zero.cf)

    plots[[as.character(a)]] = arrangeGrob(cv.patient$plot+ggtitle('Classification'), cv.patient$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)

    performance.at.1se[[as.character(a)]] = subset(cv.patient$lambdas, `lambda-1se`)
    models[[as.character(a)]] = fit0
    cvs[[as.character(a)]] = cv.patient
  }
  save(plots, coefs, performance.at.1se, dysplasia.df, models, cvs, labels, file=file)
  p = do.call(grid.arrange, c(plots[ as.character(alpha.values) ], top='All samples, 10fold, 5 splits'))
  ggsave(paste(cache.dir, '/', 'all_samples_cv.png',sep=''), plot = p, scale = 2, width = 12, height = 10, units = "in", dpi = 300)
}
all.coefs = coefs
save(arrayDf,labels,patient.info, file=paste(cache.dir,'snp.Rdata',sep='/'))
```

### 2018-06-19 processing_pipeline/SegmentMultisampleCopynumber_fitted_byPatient.R

@@ -34,7 +34,7 @@ data.files = list.files(data, full.names=T, recursive=T)
 patient.name = gsub('\\/', '_', patient.name )
 
 
-plot.dir = paste(outdir, 'multipcf_plots_fitted_perPatient2', sep='/')
+plot.dir = paste(outdir, 'multipcf_plots_fitted_perPatient', sep='/')
 print(paste("Plot directory:", plot.dir))
 if (!dir.exists(plot.dir))  
   dir.create(plot.dir, recursive=T)
@@ -66,6 +66,14 @@ data.file = grep('MultisampleCopynumberBinnedAndFitted2.RData', data.files, valu
 if (length(data.file) <= 0)
   stop("Missing necessary Rdata file from Multisample script")
 load(file=data.file, verbose=T)
+
+patient.info$Samplename
+colnames(fit.data)[1:14]
+
+patient.plot.dir
+write.table(fit.data[,c('location','chrom','start','end',patient.info$Samplename)], sep='\t', quote=F, file=paste(patient.plot.dir, 'raw_fitted.txt', sep='/'))
+write.table(raw.data[,c('location','chrom','start','end',patient.info$Samplename)], sep='\t', quote=F, file=paste(patient.plot.dir, 'raw_counts.txt', sep='/'))
+
 
 #200216 filter dodgy regions
 exclude.file = grep('qDNAseq_blacklistedRegions.txt', data.files, value=T)

### 2018-06-19 processing_pipeline/process_swgssamples.R

@@ -6,7 +6,7 @@ if (length(args) < 3)
 
 suppressPackageStartupMessages(source('../lib/fastPCF.R'))
 suppressPackageStartupMessages(source('../lib/data_func.R'))
-
+suppressPackageStartupMessages(library(ggplot2))
 
 chrlen = get.chr.lengths()
 
@@ -16,8 +16,8 @@ exclude.file = args[3]
 overwrite = F
 if (length(args) == 4) overwrite =  as.logical(args[4])
 
-# dir = '~/Data/Ellie/QDNAseq/DownsampledBE/20180206_KillcoyneS_RF_BarrettsCN/qdnaseq/'
-# outdir = '~/Data/Ellie/Analysis/downsampled5G_BEAdjacent_normal/'
+#dir = '~/Data/Ellie/Cassandra/20180604_KosmidouC_RF_sWGS/qdnaseq'
+#outdir = '~/Data/Ellie/Analysis/VAL_Cohort/sWGS'
 #exclude.file = '~/Data/Ellie/QDNAseq/qDNAseq_blacklistedRegions.txt'
 
 dir.create(outdir, F)
@@ -37,12 +37,15 @@ existing = grep('\\.Rdata', existing, invert=T, value=T)
 if (!overwrite) 
   samples = setdiff(samples, existing)
 
+plotdir = paste(outdir, 'plots',sep='/')
+logdir = paste(outdir,'logs',sep='/')
 
+for (p in c(plotdir,logdir)) dir.create(p, recursive=T, showWarnings=F)
 for (i in 1:length(samples)) {
   sampleFiles = grep(samples[i], list.files(dir), value=T)
   print(samples[i])
   
-  fit = grep('\\.fittedReadCounts.txt', sampleFiles, value=T)
+  fit = grep('\\.(fitted|corrected)ReadCounts.txt', sampleFiles, value=T)
   raw = grep('\\.readCounts.txt', sampleFiles, value=T)
   
   raw.data = read.table(paste(dir, raw, sep='/'), header=T,stringsAsFactors=F)
@@ -51,7 +54,8 @@ for (i in 1:length(samples)) {
   head(fit.data)
   ncol(raw.data) == ncol(fit.data)
   
-  res = binSWGS(raw.data, fit.data, blacklisted.regions)
[…]

### 2018-07-03 SNP_Prediction_Analysis.Rmd
Still transforming SNP

@@
 
-head(patient.info)
-
-subset(sample.list, PatientID == 74)
-
 sample.info  = do.call(rbind.data.frame, sapply(qcdata$Samplename, strsplit, '_'))
@@
 
-head(sample.info)
-head(sample.list)
-
 get.ratio<-function(PID, SID) {
@@
 )
-
-ggplot(sample.info, aes(SCA.Ratio)) + geom_histogram() + facet_grid(~Status)
 
@@
 ggplot(qcdata, aes(SCA.Ratio)) + geom_histogram() + 
-  facet_wrap(~Status, scales='free_y', ncol=1) + labs(x='SCA Ratio', title='SCA Ratio, normal vs BE')
+  facet_wrap(~Status, scales='free_y', ncol=1) + labs(x='SCA Ratio', title='SCA Ratio, NP vs P')
 
 ggplot(subset(qcdata, SampleType == 'BE'), aes(Purity, SCA.Ratio, color=Status)) + geom_point() + 
-  geom_vline(xintercept = 0.95, color='red') + geom_hline(yintercept = 0.05, color='red')
+  geom_vline(xintercept = 0.95, color='red') + geom_hline(yintercept = 0.05, color='red') + 
+  labs(title='Barretts only')
 
 ggplot(subset(qcdata, SampleType == 'BE'), aes(Ploidy, SCA.Ratio, color=Status)) + geom_point() + 
-  geom_hline(yintercept = 0.05, color='red')
+  geom_hline(yintercept = 0.05, color='red') + labs(title='Barretts only')
+
+## Low CN
+lowPloidyCN = subset(qcdata, Status == 'P' & SampleType == 'BE' & SCA.Ratio <= 0.05 & Ploidy <= 2)
+
+## samples have no apparent CN
+noCN = subset(qcdata, Status == 'P' & SCA.Ratio <= 0 & SampleType == 'BE')
+
+## high CN | ploidy, nonP
+highPloidyCN = subset(qcdata, Status == 'NP' & SampleType == 'BE' & (SCA.Ratio >= mean(qcdata$SCA.Ratio,na.rm=T)+ sd(qcdata$SCA.Ratio,na.rm=T)*2 | Ploidy >= 2.3))
+
 ```
 
-```{r}
-## 43% of P samples have low/no CN with a ploidy of 2 or less
-nrow(subset(qcdata, Status == 'P' & SampleType == 'BE' & SCA.Ratio <= 0.05 & Ploidy <= 2))/nrow(subset(qcdata, Status == 'P' & SampleType == 'BE'))
+#### P: Low SCA, Ploidy<=2
+`r round(nrow(lowPloidyCN)/nrow(subset(qcdata, Status == 'P' & SampleType == 'BE')), 3)*100`% of PROGRESSOR samples have <5% SCA and an ASCAT ploidy of 2 or less.
+Most of these show a >90% purity as well.
+`r pander(summary(lowPloidyCN$Purity))`
 
-## <1% of P samples have no apparent CN
-nrow(subset(qcdata, Status == 'P' & SCA.Ratio <= 0 & SampleType == 'BE'))/nrow(subset(qcdata, Status == 'P' & SampleType == 'BE'))
+#### P: No SCA
+`r round(nrow(noCN)/nrow(subset(qcdata, Status == 'P' & SampleType == 'BE')), 3)*100`% of PROGRESSOR samples have 0 SCA
 
-## 36% of P samples have < 5% CN and high purity ()
-nrow(subset(qcdata, Status == 'P' &  Purity >= 0.95 & SCA.Ratio <= 0.05 & SampleType == 'BE'))/nrow(subset(qcdata, Status == 'P' & SampleType == 'BE'))
+#### NP: High CN or ploidy > 2.3
+`r round(nrow(highPloidyCN)/nrow(subset(qcdata, Status == 'NP' & SampleType == 'BE')), 3)*100`% of NON-PROGRESSOR samples have >2SD SCA (~25%) or high ploidy.
+
+```{r, echo=F}
 
@@
 # For those patients with low SCA samples, a large proportion of them show low SCA in most/all of their samples
-hist(round(lowscaSamples/totalSamples[names(lowscaSamples)], 2))
+#plot(round(lowscaSamples/totalSamples[names(lowscaSamples)], 2))
+
+pt = sub('_.*', '', lowPloidyCN$Samplename[1])
+pfiles = list.files(paste('/Volumes/fh/fast/reid_b/collab/Killcoyne/Data/PerPatient', pt, sep='/'), full.names = T)
+load(grep('ascat.Rdata', pfiles, value=T), verbose=T)
+head(ascat.output$segments_raw)
+seg = subset(ascat.output$segments_raw, grepl(lowPloidyCN$Samplename[1], sample))
+seg$chr = factor(seg$chr, levels=c(1:22,'X','Y'))
+
+plot.seg<-function(seg, incAB=F) {
+  p = ggplot(seg) + facet_grid(~chr, space='free_x', scales='free_x') + 
+    geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), color='darkgreen', size=3)
+
+  if (incAB)
+    p = p + geom_segment(aes(x=startpos, xend=endpos, y=nAraw+nBraw, yend=nAraw+nBraw), color='blue', size=3)
+
+  p + theme(axis.text.x=element_blank() )
+}
+
+plot.seg(seg,incAB = T) + labs(title=paste('Low SCA Prog sample',lowPloidyCN$Samplename[1]), x='')
+
+pt = sub('_.*', '', highPloidyCN$Samplename[1])
+pfiles = list.files(paste('/Volumes/fh/fast/reid_b/collab/Killcoyne/Data/PerPatient', pt, sep='/'), full.names = T)
+load(grep('ascat.Rdata', pfiles, value=T), verbose=T)
+head(ascat.output$segments_raw)
+seg = subset(ascat.output$segments_raw, grepl(highPloidyCN$Samplename[1], sample))
+seg$chr = factor(seg$chr, levels=c(1:22,'X','Y'))
+
+plot.seg(seg, incAB = T) + labs(title=paste('High SCA Non-Prog sample',highPloidyCN$Samplename[1]), x='')
 
@@
 ```{r, warning=F}
+
+if (file.exists('tmp_seg_pt.Rdata')) {
+  load('tmp_seg_pt.Rdata', verbose=T) 
+} else {
+
 mergedSegs = NULL
@@
 for (pt in ptdirs) {
+  print(pt)
   if (length(list.files(pt, '*wins_tiled.txt', full.names=T)) <= 0) {
@@
 
+save(mergedSegs, mergedArms, file='tmp_seg_pt.Rdata')
+}
+
+rownames(mergedSegs) = sub('\\.LogR','', rownames(mergedSegs))
+rownames(mergedArms) = sub('\\.LogR','', rownames(mergedArms))
+
 # impute
@@
 
-```{r normals, fig.width=6, fig.height=8}
-rownames(mergedSegs) = sub('\\.LogR','',rownames(mergedSegs))
+```{r normals, fig.width=6, fig.height=8, eval=F}
 nm = melt(mergedSegs[grep('BLD|gastric', rownames(mergedSegs), ignore.case=T),])
 be = melt(mergedSegs[grep('BLD|gastric', rownames(mergedSegs), ignore.case=T, invert=T),])
+
+grid.arrange(
+  ggplot(nm, aes(Var2, value, group=Var2)) + geom_point() + labs(title='Blood/gastric') + theme(axis.text.x = element_blank()),
+  ggplot(be, aes(Var2, value, group=Var2)) + geom_point() + labs(title='BE samples') + theme(axis.text.x = element_blank())
+)
+
 ```
@@
 
-```{r rnormals, echo=F, fig.width=8, fig.height=8}
-randNormal = sample(grep('BLD|gastric', rownames(mergedSegs), ignore.case=T, value=T), 1)
+```{r rnormals, echo=F, fig.width=8, fig.height=8, eval=F}
+randNormal = sample(grep('BLD', rownames(mergedSegs), ignore.case=T, value=T), 1)
 pt = unlist(strsplit(randNormal, '_'))[1]
@@
 grid.arrange(
-  ggplot(subset(segments.list[[pt]], sample == paste(randNormal,'.LogR',sep='')), aes(y=medLRR)) + 
-    facet_grid(~chr, scales='free_x',space='free_x') +
-    geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), size=2) + ylim(-1,1) +
-    theme(axis.text.x=element_blank(), panel.spacing.x=unit(0, 'lines')) +
-    labs(title='Median LRR, ASCAT segments', subtitle=pt), 
-  ggplot(subset(nm, Var1 == randNormal), aes(Var2, value)) + geom_point() + ylim(-1,1) + 
-    theme(axis.text.x=element_blank()) + labs(title='Winsorized LRR, binned'),  
+  plot.seg(subset(segments.list[[pt]], sample == paste(randNormal,'.LogR',sep=''))) + labs(title='Median LRR, ASCAT segments', x='', subtitle=pt),
+  ggplot(melt(as.matrix(mergedSegs[randNormal,])), aes(Var1, value)) + geom_point() + labs(title='Winsorized') + ylim(-1,1) + theme(axis.text.x=element_blank()),
 top='Random BLD or gastric normal', bottom=randNormal)
@@
 grid.arrange(
-  ggplot(subset(segments.list[[pt]], sample == randBE), aes(y=medLRR)) + 
-    facet_grid(~chr, scales='free_x',space='free_x') +
-    geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), size=2) + ylim(1,-1) +
-    theme(axis.text.x=element_blank(), panel.spacing.x=unit(0, 'lines')) +
+  plot.seg(subset(segments.list[[pt]], sample == randBE)) +
     labs(title='Median LRR, ASCAT segments', subtitle=pt), 
-  ggplot(subset(be, Var1 == sub('\\.LogR','',randBE)), aes(Var2, value)) + geom_point() + ylim(-1,1) + 
+  ggplot(melt(as.matrix(mergedSegs[sub('\\.LogR','',randBE),])), aes(Var1, value)) + geom_point() + ylim(-1,1) + 
     theme(axis.text.x=element_blank()) + labs(title='Winsorized LRR, binned'),  
@@
 for (i in 1:ncol(ms)) {
-  ms[,i] = unit.var(ms[,i])
-  #ms[,i] = unit.var(ms[,i], z.mean[i], z.sd[i])
+  #ms[,i] = unit.var(ms[,i])
+  ms[,i] = unit.var(ms[,i], z.mean[i], z.sd[i])
 }
@@
 for (i in 1:ncol(ma)) {
-  ma[,i] = unit.var(ma[,i])
+  #ma[,i] = unit.var(ma[,i])
+  ma[,i] = unit.var(ma[,i], z.arms.mean[i], z.arms.sd[i])
 }
@@
 be = melt(ms[grep('BLD|gastric', rownames(ms), ignore.case=T, invert=T),])
+
+grid.arrange(
+  ggplot(be, aes(Var2, value, group=Var2)) + geom_point() + labs(title='BE samples'),
+  ggplot(nm, aes(Var2, value, group=Var2)) + geom_point() + labs(titls='Blood/gastric')
+)
+
 ```
-
-The variance gets much bigger after unit normalizing, but the difference between BE and normal is still clear.
 
@@
 arrayDf = cbind(arrayDf, 'cx'=unit.var(cx, mn.cx, sd.cx))
+
+ggplot(melt(arrayDf), aes(Var2, value, group=Var2)) + geom_point() + labs(title='All, scaled')
 
@@
 
-preds = preds %>% dplyr::mutate( 
-  'Adj. RR'=RR+offset,  
-  'Adj. Prob'=1/(1+exp(-RR+abs(offset)) )
-)
+# preds = preds %>% dplyr::mutate( 
+#   'Adj. RR'=RR+offset,  
+#   'Adj. Prob'=1/(1+exp(-RR+abs(offset)) )
+# )
 
@@
 
-```{r}
+```{r, eval=F}
 hist(subset(preds, Samplename %in% lowsca$Samplename)$Prob)
@@
 top='BE samples' )
-
+pred.qc = base::merge(preds, qcdata, by='Samplename')
 ```
@@
 So how do these high normals look?
-```{r}
+```{r, eval=F}
 preds
@@
 nrow(preds)
-
-pred.qc = base::merge(preds, qcdata, by='Samplename')
 
@@
 ```{r, echo=F, fig.height=8}
-bld = preds[grep('BLD', preds$Samplename, ignore.case=T),]
-grid.arrange(
-  ggplot(bld, aes(Prob)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
-  scale_fill_gradientn(colors = myPal,  name='') + 
-  labs(y='n Samples', x='Unadjusted Probability'),
-ggplot(bld, aes(`Adj. Prob`)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
-  scale_fill_gradientn(colors = myPal,  name='') + 
-  labs(y='n Samples', x='Adjusted Probability'), top='Predictions on BLD samples only' )
-
-length(which(bld$Prob < 0.5))/nrow(bld)
-length(which(bld$`Adj. Prob` < 0.5))/nrow(bld)
-
-be = preds[grep('BLD|gastric', rownames(preds), ignore.case=T, invert = T),]
-grid.arrange(
-  ggplot(be, aes(Prob)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
-    scale_fill_gradientn(colors = myPal,  name='') + 
-    labs(y='n Samples', x='Unadjusted Probability'),
-  ggplot(be, aes(`Adj. Prob`)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
-    scale_fill_gradientn(colors = myPal,  name='') + 
-    labs(y='n Samples', x='Adjusted Probability'), top='Predictions on BE only'
-)
 
 unadj = length(unique(sapply(rownames(be[which(be$Prob > 0.5),]), function(x) unlist(strsplit(x, '_'))[1])))/length(patients)
-adj = length(unique(sapply(rownames(be[which(be$`Adj. Prob` > 0.5),]), function(x) unlist(strsplit(x, '_'))[1])))/length(patients)
-
+#adj = length(unique(sapply(rownames(be[which(be$`Adj. Prob` > 0.5),]), function(x) unlist(strsplit(x, '_'))[1])))/length(patients)
+#Adjusted, `r round(adj, 2)*100`% have a >50% probability.
 ```
 
-Unajdusted `r round(unadj,2)*100`% of BE patients have a >50% probability of progression.  Adjusted, `r round(adj, 2)*100`% have a >50% probability.
+Unajdusted `r round(unadj,2)*100`% of BE patients have a >50% probability of progression.
 
@@
 
-```{r}
-
+```{r roc, echo=T}
 #preds.info = base::merge(sample.info, preds, by='Samplename')
-
 rows = grep('BLD|gastric', pred.qc$Samplename, ignore.case=T)
@@
 
-rocNoLow = pROC::roc(Status ~ Prob, data=subset(pred.qc, !Samplename %in% lowsca$Samplename), auc=T, ci=T, of='thresholds')
+rocNoLow = pROC::roc(Status ~ Prob, data=subset(pred.qc, !Samplename %in% c(lowsca$Samplename,highPloidyCN$Samplename)), auc=T, ci=T, of='thresholds')
 roc.plot(rocNoLow)
@@
 
-roct1 = pROC::roc(Status ~ Prob, data=subset(pred.qc, Timepoint == 'T1'), auc=T, ci=T, of='thresholds')
-roc.plot(roct1)
+per_pt_ex = subset(pred.qc, !Samplename %in% c(lowsca$Samplename,highPloidyCN$Samplename)) %>% group_by(PatientID, Timepoint, Status) %>% dplyr::summarise(
+  max.Prob = max(Prob), max.RR = max(RR)
+)
 
-roct2 = pROC::roc(Status ~ Prob, data=subset(pred.qc, Timepoint != 'T1'), auc=T, ci=T, of='thresholds')
+rocppex = pROC::roc(Status ~ max.Prob, data=per_pt_ex, auc=T, ci=T, of='thresholds')
+roc.plot(rocppex)
+
+roct2 = pROC::roc(Status ~ Prob, data=subset(pred.qc, Timepoint != 'T1' & !Samplename %in% c(lowsca$Samplename,highPloidyCN$Samplename)), auc=T, ci=T, of='thresholds')
 roc.plot(roct2)
-
-qcpred = subset(pred.qc, Samplename %in% subset(qcdata, Purity > 0.8 & Goodnessoffit >= 0.9)$Samplename)
-qc = pROC::roc(Status ~ Prob, data=qcpred, auc=T, ci=T, of='thresholds')
-roc.plot(qc)
 
@@
 
-## Regression using SNP Data
+# Regions predicted by model
+Not evaluated currently
+```{r coefs, eval=F}
+features = coefs$`0.9`[,1,drop=F]
 
-```{r, echo=F, eval=F}
+locs = do.call(rbind.data.frame, c(strsplit(rownames(features), ':|-'), stringsAsFactors=F))
+colnames(locs) = c('chr','start','end')
+locs[] = lapply(locs[], as.numeric)
 
-sample.info = patient.info[c('PatientID','Status')]
-sample.info = base::merge(sampleInfo, sample.info, by='PatientID', all.x=T)
+feature.locs = makeGRangesFromDataFrame(locs)
 
-sample.info = subset(sample.info, !Level %in% c('BLD','GASTRIC'))
-sample.info$Samplename = with(sample.info, paste(PatientID, SampleID, EndoscopyID, Level, sep='_'))
+gr = makeGRangesFromDataFrame(segments.list$`512`, start.field = 'startpos', end.field = 'endpos', keep.extra.columns = T)
 
-df = arrayDf
-arrayDf = arrayDf[grep('BLD|GASTRIC',rownames(df), invert=T),]
-rownames(arrayDf) = sub('\\.LogR','',rownames(arrayDf))
+ov = findOverlaps(feature.locs, gr)
 
-patient.info
-status = vector(length=nrow(arrayDf), mode='character')
-names(status) = rownames(arrayDf)
+feature.locs[unique(queryHits(ov))]
 
-for (i in 1:length(status))
-  status[i] = unique(subset(patient.info, PatientID == sub('_.*','',names(status)[i]))[,'Status'])
+cna = lapply( unique(queryHits(ov)), function(hit) {
+  tmp = gr[subjectHits(ov[queryHits(ov) == hit])]
+  tmp = subset(as.data.frame(tmp), nMajor+nMinor > 2.3 | nMajor+nMinor < 1.5) %>% group_by(sample, nMajor+nMinor) %>% dplyr::summarise( n=length(sample), medLRR=mean(medLRR), mn.bp=mean(end-start) ) %>% rowwise() %>%
+    dplyr::mutate( type=ifelse(`nMajor + nMinor` >= 3, 1, 0))
+})
+names(cna) = rownames(features)
 
-labels = as.integer(factor(labels))-1
-names(labels) = names(status)
-rm(status)
+fl = sapply(cna, function(x) length(unique(x$sample)))
+feature.locs[which(fl == 0)]
+
+cna.count = do.call(rbind, lapply(cna, function(x) table(x$type)))
+
+cna.mx = as.matrix(sapply(cna, function(x) length(unique(x$sample))))
+cna.mx = cbind('sample.cnt'=cna.mx, 'loss'=NA,'gain'=NA, 'loss.bp', 'gain.bp')
+for (ft in rownames(features)) {
+  cna.mx[ft,c('loss','gain')] = c(0,0)
+  if (nrow(cna[[ft]]) > 0)
+    cna.mx[ft,c('loss','gain')] = c(length(which(cna[[ft]]$type == 0)),length(which(cna[[ft]]$type == 1)))
+
+  median(cna[[ft]] [[cna[[ft]]$type == 0,'mn.bp']] )
+
+}
+
+cna$`1:59820150-64805161`
+
+```
+
+# Regression using SNP Data
+
+```{r regression, echo=F, eval=F}
+slabels = sample.info[,c('PatientID','Status', 'Samplename')]
+
+missing = setdiff(sample.info$Samplename, rownames(arrayDf))
+rows = grep('BLD|gastric', rownames(arrayDf), ignore.case=T)
+bld_gastric = arrayDf[rows,]
+
+df = arrayDf[-rows,]
+
+slabels = subset(slabels, Samplename %in% rownames(df))
+
+status = as.integer(as.factor(slabels$Status))-1
+names(status) = slabels$Samplename
+
+df = df[-grep(setdiff(rownames(df),names(status)), rownames(df)),]
+
+#df = arrayDf[which(!rownames(arrayDf) %in% c(lowsca$Samplename,highPloidyCN$Samplename)),]
 
@@
 nl = 1000;folds = 10; splits = 5 
-sets = create.patient.sets(sample.info[c('PatientID','Samplename','Status')], folds, splits, 0.2)
+sets = create.patient.sets(slabels, folds, splits, 0.2)
+
+#sets = create.patient.sets(subset(sample.info, !Samplename %in% c(lowsca$Samplename,highPloidyCN$Samplename)) [c('PatientID','Samplename','Status')], folds, splits, 0.2)
 
@@
 file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
-if (file.exists(file)) {
-  message(paste("loading", file))
-  load(file, verbose=T)
-} else {
+#if (file.exists(file)) {
+#  message(paste("loading", file))
+#  load(file, verbose=T)
+#} else 
+  {
   for (a in alpha.values) {
-    fit0 <- glmnet(arrayDf, labels, alpha=a, nlambda=nl, family='binomial', standardize=F)    
+    fit0 <- glmnet(df, status, alpha=a, nlambda=nl, family='binomial', standardize=F)    
+    autoplot(fit) + theme(legend.position="none")
     l = fit0$lambda
 
-    cv.patient = crossvalidate.by.patient(x=arrayDf, y=labels, lambda=l, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', opt=-1, standardize=F)
+    cv.patient = crossvalidate.by.patient(x=df, y=status, lambda=l, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', opt=-1, standardize=F)
 

### 2018-07-03 processing_pipeline/run-cv2.R

@@ -97,6 +97,9 @@ tiled.segs = do.call(rbind, lapply(segFiles, function(f) {
 }))
 dim(tiled.segs)
 
+ggplot(melt(tiled.segs), aes(Var2, value)) + geom_point()
+
+
 ## Log?
 if (logT) tiled.segs = t(apply(tiled.segs, 1, logTransform))
 

### 2018-07-06 SNP_Prediction_Analysis.Rmd
processing per endo instead of per sample

@@
 
+plot.sample<-function(segs, mer)
+
 chr.info = get.chr.lengths(file='hg19_info.txt')
 
-ptdirs = list.dirs('/Volumes/fh/fast/reid_b/collab/Killcoyne/Data/PerPatient', full.names=T, recursive=F)
-patients = basename(ptdirs)
+#ptdirs = list.dirs('/Volumes/fh/fast/reid_b/collab/Killcoyne/Data/PerPatient', full.names=T, recursive=F)
+#patients = basename(ptdirs)
 
@@
 
-#load('/Volumes/fh/fast/reid_b/collab/Killcoyne/SNP_R/allpts_ascat.Rdata', verbose=T)
-load('~/Data/Ellie/Analysis/SNP/allpts_ascat.Rdata', verbose=T)
-#patient.info = as.data.frame(read_xlsx('/Volumes/fh/fast/reid_b/collab/Killcoyne/SNP_Project/metadata_T1T2.xlsx'))
+load('~/Data/Reid_SNP/PerPatient/allpts_ascat.Rdata', verbose=T)
+
 patient.info = as.data.frame(read_xlsx('~/Data/Ellie/Analysis/SNP/metadata_T1T2.xlsx'))
@@
 
-sample.list = read_xlsx('/Volumes/fh/fast/reid_b/collab/Killcoyne/Data/20180604_Reid_1M_SampleList.xlsx', sheet = 2)
+sample.list = read_xlsx('~/Data/Ellie/Analysis/SNP/20180604_Reid_1M_SampleList.xlsx', sheet = 2)
 nrow(sample.list)
@@
 
-In the sWGS data progressor samples show a variance of `r round(pv, 3)` and non-prog variance is `r round(npv,3)`
+## Patient info
+
+```{r, echo=F}
+
+patient.info = patient.info %>% rowwise() %>% dplyr::mutate( Endoscopy = ifelse(Timepoint == 'T1', 1, 2) )
+
+sm = patient.info %>% group_by(Status, Endoscopy, Pathology) %>% dplyr::tally() %>% group_by(Status, Endoscopy) %>% dplyr::mutate(freq=n/sum(n) )
+sm$Pathology = factor(sm$Pathology, levels=c('NDBE','ID','LGD','HGD','IMC'), ordered = T)
+
+ggplot(sm, aes(Pathology, freq, group=Endoscopy, fill=factor(Endoscopy))) + facet_grid(~Status) + geom_col(position = position_dodge()) + ylim(0,1) + title('Pathology frequency per timepoint')
+```
+
+There's a significant difference in the numbers of NDBE or HGD samples between the groups
+```{r}
+chisq.test(table(patient.info$Pathology, patient.info$Status)['HGD',])
+chisq.test(table(patient.info$Pathology, patient.info$Status)['NDBE',])
+
+# Most progressors are HGD at timepoint 1
+with(subset(patient.info, Endoscopy == 1), table(Pathology, Status))['HGD',]/table(subset(patient.info, Endoscopy == 1)$Status)
+```
 
 # Quick QC 
+In the sWGS data progressor samples show a variance of `r round(pv, 3)` and non-prog variance is `r round(npv,3)`
 
@@
 lowsca = subset(qcdata, Status == 'P' & SampleType == 'BE' & SCA.Ratio <= 0.05 & Ploidy <= 2)
-write.table(lowsca, sep='\t', quote=F, row.names = F, file = '~/Data/Ellie/Analysis/SNP/low_sca_samples.txt')
 
@@
 
-pt = sub('_.*', '', lowPloidyCN$Samplename[1])
-pfiles = list.files(paste('/Volumes/fh/fast/reid_b/collab/Killcoyne/Data/PerPatient', pt, sep='/'), full.names = T)
+pt = '403'
+pfiles = list.files(paste('~/Data/Reid_SNP/PerPatient', pt, sep='/'), full.names = T)
 load(grep('ascat.Rdata', pfiles, value=T), verbose=T)
-head(ascat.output$segments_raw)
-seg = subset(ascat.output$segments_raw, grepl(lowPloidyCN$Samplename[1], sample))
+
+seg = ascat.output$segments_raw
 seg$chr = factor(seg$chr, levels=c(1:22,'X','Y'))
+ggplot(seg) + facet_grid(sample~chr, space='free_x', scales='free') + 
+  geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), color='darkgreen', size=3) + 
+  theme(axis.text.x=element_blank() ) + labs(title='Low SCA', subtitle='patient 403')
 
-plot.seg<-function(seg, incAB=F) {
-  p = ggplot(seg) + facet_grid(~chr, space='free_x', scales='free_x') + 
-    geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), color='darkgreen', size=3)
+pt = '512'
+pfiles = list.files(paste('~/Data/Reid_SNP/PerPatient', pt, sep='/'), full.names = T)
+load(grep('ascat.Rdata', pfiles, value=T), verbose=T)
 
-  if (incAB)
-    p = p + geom_segment(aes(x=startpos, xend=endpos, y=nAraw+nBraw, yend=nAraw+nBraw), color='blue', size=3)
-
-  p + theme(axis.text.x=element_blank() )
-}
-
-plot.seg(seg,incAB = T) + labs(title=paste('Low SCA Prog sample',lowPloidyCN$Samplename[1]), x='')
-
-pt = sub('_.*', '', highPloidyCN$Samplename[1])
-pfiles = list.files(paste('/Volumes/fh/fast/reid_b/collab/Killcoyne/Data/PerPatient', pt, sep='/'), full.names = T)
-load(grep('ascat.Rdata', pfiles, value=T), verbose=T)
-head(ascat.output$segments_raw)
-seg = subset(ascat.output$segments_raw, grepl(highPloidyCN$Samplename[1], sample))
+seg = ascat.output$segments_raw
 seg$chr = factor(seg$chr, levels=c(1:22,'X','Y'))
-
-plot.seg(seg, incAB = T) + labs(title=paste('High SCA Non-Prog sample',highPloidyCN$Samplename[1]), x='')
+ggplot(seg) + facet_grid(sample~chr, space='free_x', scales='free') + 
+  geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), color='darkgreen', size=3) + 
+  theme(axis.text.x=element_blank() ) + labs(title='Mixed SCA samples', subtitle='patient 512')
 
@@
 
-These are the CN adjusted, winsorized LogR values tiled in 5Mb bins.
+CN Adjustment uses Annalise's whole-genome non-progressor samples to generate a set of mean and SD values per copy-number state.
+The median LogR values are adjusted using the mean and SD based on the CN state.
+The adjusted LogR values are then winsorize per sample, and tiled in 5Mb bins.
 
-```{r, warning=F}
+```{r, warning=F, echo=F}
 
-if (file.exists('tmp_seg_pt.Rdata')) {
-  load('tmp_seg_pt.Rdata', verbose=T) 
+if (file.exists('~/Data/Reid_SNP/PerPatient/tmp_seg_pt.Rdata')) {
+  load('~/Data/Reid_SNP/PerPatient/tmp_seg_pt.Rdata', verbose=T) 
 } else {
@@
 }
-
 rownames(mergedSegs) = sub('\\.LogR','', rownames(mergedSegs))
@@
 
-# impute
+## Per endoscopy take the average or max bin value as an in silico pooling
+sample.info$Level = toupper(sample.info$Level)
 
-# mergedSegs[is.na(mergedSegs)]
-# mergedArms[is.na(mergedArms)]
+range(mergedSegs[1,])
+range(x)
+plot( x )
 
-# for (i in 1:nrow(mergedSegs)) 
-#   mergedSegs[i,is.na(mergedSegs[i,])] = 0# mean(mergedSegs[i,],na.rm=T)
+x = (mergedSegs[1,]-mean(apply(allDf[,1:ncol(mergedSegs)], 1, mean)))/mean(apply(allDf[,1:ncol(mergedSegs)], 1, sd))
+  #sd(apply(allDf[,1:ncol(mergedSegs)], 1, sd))
+
+# names = colnames(mergedSegs)
+# mergedSegs = t(apply(mergedSegs, 1, scale, center=T))
+# colnames(mergedSegs) = names
 # 
-# for (i in 1:nrow(mergedArms)) 
-#   mergedArms[i,is.na(mergedArms[i,])] = 0 #mean(mergedArms[i,],na.rm=T)
+# names = colnames(mergedArms)
+# mergedArms = t(apply(mergedArms, 1, scale, center=T))
+# colnames(mergedArms) = names
 
+samples = unique(subset(sample.info, !Level %in% c('BLD','GASTRIC'), select=c("PatientID", "EndoID")))
+
+meanBinSegs = as.data.frame(matrix(ncol=ncol(mergedSegs), nrow=0, dimnames=list(c(), colnames(mergedSegs))))
+meanBinArms = as.data.frame(matrix(ncol=ncol(mergedArms), nrow=0, dimnames=list(c(), colnames(mergedArms))))
+
+mean.qcdata = data.frame(matrix(ncol=8, nrow=0, dimnames=list(c(), c('Samplename','Status','SampleType','Timepoint','mean.sca','mean.ploidy','mean.fit','mean.purity'))))
+
+for (i in 1:nrow(samples)) {
+  samples[i,]
+  print(paste(samples$PatientID[i], '(',i,')'))
+
+  #rows = subset(sample.info, PatientID == samples$PatientID[i] & EndoID == samples$EndoID[i] )$Samplename
+  ptendoID = paste(samples$PatientID[i], samples$EndoID[i], sep='_')
+
+  mean.qcdata[i,] = 
+    cbind('Samplename'=ptendoID, subset(qcdata, SampleType == 'BE' & grepl(paste(samples$PatientID[i], samples$EndoID[i], sep='.*'), Samplename)) %>% group_by(Status, SampleType, Timepoint) %>%
+    dplyr::summarise(mean.sca=mean(SCA.Ratio), mean.ploidy=mean(Ploidy), mean.fit=mean(Goodnessoffit), mean.purity=mean(Purity)))
+
+  rows = grep(paste(samples$PatientID[i], samples$EndoID[i], sep='.*'), rownames(mergedSegs), value=T)
+
+  sharedR = intersect(rows, rownames(mergedSegs))
+  if (length(rows) != length(sharedR)) {
+      warning(paste("Missing samples for", endo$PatientID))
+      rows = sharedR
+  }
+
+  #meanBinSegs[ptendoID,] = apply(mergedSegs[rows,,drop=F], 2, mean)
+  meanBinSegs[ptendoID,] = apply(mergedSegs[rows,,drop=F], 2, function(x) x[which.max(abs(x))] )
+  #meanBinArms[ptendoID,] = apply(mergedArms[rows,,drop=F], 2, mean)
+  meanBinArms[ptendoID,] = apply(mergedArms[rows,,drop=F], 2, function(x) x[which.max(abs(x))] )
+}
+
+normalSegs = mergedSegs[grep('BLD|gastric',rownames(mergedSegs),ignore.case=T),]
+normalArms = mergedArms[grep('BLD|gastric',rownames(mergedArms),ignore.case=T),]
+
+meanBinSegs = as.matrix(meanBinSegs) 
+meanBinArms = as.matrix(meanBinArms)
+
+dim(meanBinSegs)
+
+lowSeg = mergedSegs[grep('^403', rownames(mergedSegs), value=T),,drop=F]
+p1 = ggplot(melt(lowSeg), aes(Var2, value)) + facet_grid(~Var1, space='free_x', scales='free') + geom_point() +
+  theme(axis.text.x=element_blank() ) + labs(title='Low SCA samples', subtitle='patient 403')
+lowSeg = meanBinSegs[grep('^403', rownames(meanBinSegs), value=T),,drop=F]
+p2 = ggplot(melt(lowSeg), aes(Var2, value)) + facet_grid(~Var1, space='free_x', scales='free') + geom_point() +
+  theme(axis.text.x=element_blank() ) + labs(title='Low SCA samples, mean bins', subtitle='patient 403')
+
+grid.arrange(p1,p2)
+
+highSeg = mergedSegs[grep('^512', rownames(mergedSegs), value=T),,drop=F]
+p1 = ggplot(melt(highSeg), aes(Var2, value)) + facet_grid(~Var1, space='free_x', scales='free') + geom_point() +
+  theme(axis.text.x=element_blank() ) + labs(title='Mixed SCA samples', subtitle='patient 512')
+highSeg = meanBinSegs[grep('^512', rownames(meanBinSegs), value=T),,drop=F]
+p2 = ggplot(melt(highSeg), aes(Var2, value)) + facet_grid(~Var1, space='free_x', scales='free') + geom_point() +
+  theme(axis.text.x=element_blank() ) + labs(title='Mixed SCA samples, mean bins', subtitle='patient 512')
+
+grid.arrange(p1,p2)
 ```
@@
 ```{r normals, fig.width=6, fig.height=8, eval=F}
-nm = melt(mergedSegs[grep('BLD|gastric', rownames(mergedSegs), ignore.case=T),])
+nm = melt(normalSegs)
 be = melt(mergedSegs[grep('BLD|gastric', rownames(mergedSegs), ignore.case=T, invert=T),])
+beM = melt(meanBinSegs[grep('BLD|gastric', rownames(meanBinSegs), ignore.case=T, invert=T),])
+
+lims = range(nm$value, be$value, beM$value)
 
 grid.arrange(
-  ggplot(nm, aes(Var2, value, group=Var2)) + geom_point() + labs(title='Blood/gastric') + theme(axis.text.x = element_blank()),
-  ggplot(be, aes(Var2, value, group=Var2)) + geom_point() + labs(title='BE samples') + theme(axis.text.x = element_blank())
+  ggplot(nm, aes(Var2, value, group=Var2)) + geom_point() + labs(title='Blood/gastric') + theme(axis.text.x = element_blank()) + ylim(lims),
+  ggplot(be, aes(Var2, value, group=Var2)) + geom_point() + labs(title='BE samples') + theme(axis.text.x = element_blank()) + ylim(lims),
+  ggplot(beM, aes(Var2, value, group=Var2)) + geom_point() + labs(title='BE samples, mean per endo') + theme(axis.text.x = element_blank()) + ylim(lims)
 )
@@
 
-Using a random normal patient and a known BE progresser patient
+Using a random normal sample and a known BE progresser sampe
 
 ```{r rnormals, echo=F, fig.width=8, fig.height=8, eval=F}
-randNormal = sample(grep('BLD', rownames(mergedSegs), ignore.case=T, value=T), 1)
+randNormal = sample(rownames(normalSegs), 1)
 pt = unlist(strsplit(randNormal, '_'))[1]
-segments.list[[pt]] = subset(segments.list[[pt]], chr %in% c(1:22))
-segments.list[[pt]]$chr = factor(segments.list[[pt]]$chr, levels=c(1:22), ordered = T)
+segs = subset(segments.list[[pt]], chr %in% c(1:22) & grepl(randNormal, sample))
+segs$chr = factor(segs$chr, levels=c(1:22), ordered = T)
 
+lims = range(segs$medLRR, mergedSegs[randNormal,])
 grid.arrange(
-  plot.seg(subset(segments.list[[pt]], sample == paste(randNormal,'.LogR',sep=''))) + labs(title='Median LRR, ASCAT segments', x='', subtitle=pt),
-  ggplot(melt(as.matrix(mergedSegs[randNormal,])), aes(Var1, value)) + geom_point() + labs(title='Winsorized') + ylim(-1,1) + theme(axis.text.x=element_blank()),
+  ggplot(segs) + facet_grid(~chr, space='free_x', scales='free') + ylim(lims) +
+    geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), size=3) + 
+    theme(axis.text.x=element_blank() ) + labs(title='ASCAT median LRR', subtitle=pt),
+  ggplot(melt(as.matrix(mergedSegs[randNormal,])), aes(Var1, value)) + geom_point() + ylim(lims) +
+    labs(title='CN Adjusted & winsorized') + theme(axis.text.x=element_blank()),
 top='Random BLD or gastric normal', bottom=randNormal)
@@
 randBE = '512_18762_133R_27.LogR'
-#randBE = paste(sample(grep(paste('^',sample(subset(patient.info, Status == 'P')$PatientID, 1),sep=''), rownames(arrayDf), value=T), 1),'.LogR',sep='')
 pt = unlist(strsplit(randBE, '_'))[1]
-segments.list[[pt]] = subset(segments.list[[pt]], chr %in% c(1:22))
-segments.list[[pt]]$chr = factor(segments.list[[pt]]$chr, levels=c(1:22), ordered = T)
+segs = subset(segments.list[[pt]], chr %in% c(1:22) & grepl(randBE, sample))
+segs$chr = factor(segs$chr, levels=c(1:22), ordered = T)
 
+lims = range(segs$medLRR, mergedSegs[sub('\\.LogR','',randBE),])
 grid.arrange(
-  plot.seg(subset(segments.list[[pt]], sample == randBE)) +
-    labs(title='Median LRR, ASCAT segments', subtitle=pt), 
-  ggplot(melt(as.matrix(mergedSegs[sub('\\.LogR','',randBE),])), aes(Var1, value)) + geom_point() + ylim(-1,1) + 
+  ggplot(segs) + facet_grid(~chr, space='free_x', scales='free') + ylim(lims) +
+    geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), size=3) + 
+    theme(axis.text.x=element_blank() ) + labs(title='Median LRR, ASCAT segments', subtitle=pt), 
+  ggplot(melt(as.matrix(mergedSegs[sub('\\.LogR','',randBE),])), aes(Var1, value)) + geom_point() + ylim(lims) + 
     theme(axis.text.x=element_blank()) + labs(title='Winsorized LRR, binned'),  
@@
 
-I adjust the variance of each genomic region by the mean and sd extracted when I trained on the sWGS data.
+Adjust the variance of each genomic region by the mean and sd extracted when I trained on the sWGS data.
 This seems to be a more fair comparison than scaling all the new data to itself, but the predictions don't really change.
+
+Using the mean bin values.
 ```{r, warning=F}
-# uv 
-# ms  = t(apply(mergedSegs, 1, function(sample) {
-#   sapply(1:ncol(mergedSegs), function(i){
-#     unit.var(sample[i], z.mean[i], z.sd[i])
-#   })
-# }))
-# 
-# range(ms)
-ms = mergedSegs
-for (i in 1:ncol(ms)) {
-  #ms[,i] = unit.var(ms[,i])
-  ms[,i] = unit.var(ms[,i], z.mean[i], z.sd[i])
+adjust.cols<-function(mtx, means=NULL, sds=NULL) {
+  if (!is.null(means)) {
+    if (ncol(mtx) != length(means))
+      stop("Vector of means needs to match the columns in the matrix")
+    for (i in 1:ncol(mtx)) 
+      mtx[,i] = unit.var(mtx[,i], means[i], sds[i])
+  } else {
+    for (i in 1:ncol(mtx)) 
+      mtx[,i] = unit.var(mtx[,i])
   }
+  return(mtx)
+}
+
+ns = adjust.cols(normalSegs, z.mean, z.sd)
+na = adjust.cols(normalArms, z.arms.mean, z.arms.sd)
+
+ms = adjust.cols(meanBinSegs, z.mean, z.sd)
+ms = adjust.cols(mergedSegs, z.mean, z.sd)
 range(ms)
 dim(ms)
-# ma  = t(apply(mergedArms, 1, function(sample) {
-#   sapply(1:ncol(mergedArms), function(i){
-#     unit.var(sample[i], z.arms.mean[i], z.arms.sd[i])
-#   })
-# }))
 
-ma = mergedArms
-for (i in 1:ncol(ma)) {
-  #ma[,i] = unit.var(ma[,i])
-  ma[,i] = unit.var(ma[,i], z.arms.mean[i], z.arms.sd[i])
-}
+ma = adjust.cols(meanBinArms, z.arms.mean, z.arms.sd)
+ma = adjust.cols(mergedArms, z.arms.mean, z.arms.sd)
 range(ma)
@@
 
-nm = melt(ms[grep('BLD|gastric', rownames(ms), ignore.case=T),])
+nm = melt(ns)
 be = melt(ms[grep('BLD|gastric', rownames(ms), ignore.case=T, invert=T),])
@@
 grid.arrange(
-  ggplot(be, aes(Var2, value, group=Var2)) + geom_point() + labs(title='BE samples'),
-  ggplot(nm, aes(Var2, value, group=Var2)) + geom_point() + labs(titls='Blood/gastric')
-)
+  ggplot(nm, aes(Var2, value, group=Var2)) + geom_point() + labs(title='Blood/gastric') + 
+    theme(axis.text.x=element_blank()) +ylim(range(nm$value, be$value)),
+  ggplot(be, aes(Var2, value, group=Var2)) + geom_point() + labs(title='BE samples') + 
+    theme(axis.text.x=element_blank()) + ylim(range(nm$value, be$value)), top='Unit scaled')
 
@@
 
-ggplot(melt(arrayDf), aes(Var2, value, group=Var2)) + geom_point() + labs(title='All, scaled')
+cxN = score.cx(ns, 1)
+arrayDfNorm = subtract.arms(ns, na)
+arrayDfNorm = cbind(arrayDfNorm, 'cx'=unit.var(cxN, mn.cx, sd.cx))
+
+probN = predict(fitV, newx=arrayDfNorm, s=l, type='response')
+length(which(probN > 0.5))/nrow(probN) # FPR
+length(which(probN <= 0.5))/nrow(probN)
 
@@
 
-offset = log(0.0225)
 preds = cbind.data.frame(rownames(prob),prob, rr)
@@
 
+#offset = log(0.0225)
 # preds = preds %>% dplyr::mutate( 
@@
 
+#randNormal = paste(unlist(strsplit(randNormal, '_'))[c(1,3)], collapse='_')
+#randBE = paste(unlist(strsplit(randBE, '_'))[c(1,3)], collapse='_')
 ```
 
-A quick look at the two random samples from earlier.  The random normal shows a very low probablity before and after adjustment:
-`r pander(subset(preds, Samplename == randNormal))`
+A quick look at the two random samples from earlier.  
+`r pander(subset(preds, Samplename == randNormal), caption='Random normal')`
 
 The random BE shows a very high probability:
-`r pander(subset(preds, Samplename == sub('\\.LogR','',randBE)))`
+`r pander(subset(preds, Samplename == randBE), caption='Random BE')`
 
-```{r, eval=F}
-hist(subset(preds, Samplename %in% lowsca$Samplename)$Prob)
-
-x = subset(segments.list$`891`, sample == '891_18574_100O_33.LogR') 
-x$chr = factor(x$chr, levels=c(1:22,'X','Y'), ordered = T)
-p1 = ggplot(x) + facet_grid(~chr, space='free_x', scales='free_x') + 
-  geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), size=3) + labs(title='891_18574_100O_33 medLRR')
-
-x = melt(ms['891_18574_100O_33',])
-x$loc = rownames(x)
-x = rbind(x,cbind.data.frame('loc'=paste(c(1:22), '0-1', sep=':'), 'value'=0))
-
-locs = do.call(rbind.data.frame, strsplit(x$loc, ':|-'))
-colnames(locs) = c('chr','start','end')
-locs[] = lapply(locs[], function(y) as.numeric(as.character(y)))
-x = cbind(x,locs)
-x$chr = factor(x$chr, levels=c(1:22), ordered=T)
-x = arrange(x,chr)
-
-p2 = ggplot(x) + facet_grid(~chr, space='free_x', scales='free_x') + 
-  geom_segment(aes(x=start, xend=end, y=value, yend=value), size=3) + labs(title='891_18574_100O_33 binned')
-
-grid.arrange(p1,p2,nrow=2)
-```
-
-## Evaluating normals vs BE
+## Evaluating predictions
 
 ```{r, warning=F, fig.height=8}
-normals = subset(preds, grepl('BLD|gastric', Samplename, ignore.case=T))
-grid.arrange(
-  ggplot(normals, aes(Prob)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
+# normals = subset(preds, grepl('BLD|gastric', Samplename, ignore.case=T))
+# grid.arrange(
+#   ggplot(normals, aes(Prob)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
+#     scale_fill_gradientn(colors = myPal,  name='') + 
+#     labs(title='Predictions on BLD/Gastric', y='n Samples', x='Unadjusted Probability'),
+# top='BLD/gastric normals' )
+# 
+# normals[which(normals$Prob > 0.5),]
+
+pred.qc = base::merge(preds, mean.qcdata, by='Samplename')
+#pred.qc = base::merge(preds, qcdata, by='Samplename')
+
+prog = subset(pred.qc, !grepl('BLD|gastric', Samplename, ignore.case=T) & Status == 'P')
+ggplot(prog, aes(Prob)) + facet_grid(~Timepoint) + 
+  geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
   scale_fill_gradientn(colors = myPal,  name='') + 
-    labs(title='Predictions on BLD/Gastric', y='n Samples', x='Unadjusted Probability'),
-top='BLD/gastric normals' )
+  labs(title='Predictions on progressors', y='n Samples', x='Unadjusted Probability')
 
-be = subset(preds, !grepl('BLD|gastric', Samplename, ignore.case=T))
-grid.arrange(
-  ggplot(be, aes(Prob)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
+nonprog = subset(pred.qc, !grepl('BLD|gastric', Samplename, ignore.case=T) & Status == 'NP')
+ggplot(nonprog, aes(Prob)) + facet_grid(~Timepoint) + 
+  geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
   scale_fill_gradientn(colors = myPal,  name='') + 
-    labs(title='Predictions on BE', y='n Samples', x='Unadjusted Probability'),
-top='BE samples' )
-pred.qc = base::merge(preds, qcdata, by='Samplename')
+  labs(title='Predictions on non-progressors', y='n Samples', x='Unadjusted Probability')
+
 ```
-
-So how do these high normals look?
-```{r, eval=F}
-preds
-qcdata
-nrow(preds)
-
-rc = Hmisc::rcorr( as.matrix(pred.qc[,c(2,3,6,7,8,12)]) )
-
-GGally::ggcorr( as.matrix(pred.qc[,c(2,6,7,8,12)]) , label=T, palette='RdBu')
-
-length(which(normals$Prob < 0.5))/nrow(normals)
-
-highNmSamples = subset(normals, Prob >= 0.5)
-subset(qcdata, Samplename %in% highNmSamples$Samplename)
-# Low variance overall
-mean(apply(mergedSegs[highNmSamples$Samplename,], 1, var))
-
-# Most of the gastric "normals" show up, though that drops to 3% when adjusted
-length(grep('gastric', highNmSamples$Samplename, ignore.case=T))/length(grep('gastric', rownames(mergedSegs), ignore.case=T))
-length(grep('gastric', subset(normals, `Adj. Prob` >= 0.5)$Samplename, ignore.case=T))/length(grep('gastric', rownames(mergedSegs), ignore.case=T))
-
-# About 8% of the blood normals do as well, but drops to <1% when adjusted
-length(grep('BLD', highNmSamples$Samplename, ignore.case=T))/length(grep('BLD', rownames(mergedSegs), ignore.case=T))
-length(grep('BLD', subset(normals, `Adj. Prob` >= 0.5)$Samplename, ignore.case=T))/length(grep('BLD', rownames(mergedSegs), ignore.case=T))
-```
-
-Compare BLD to BE
-```{r, echo=F, fig.height=8}
-
-unadj = length(unique(sapply(rownames(be[which(be$Prob > 0.5),]), function(x) unlist(strsplit(x, '_'))[1])))/length(patients)
-#adj = length(unique(sapply(rownames(be[which(be$`Adj. Prob` > 0.5),]), function(x) unlist(strsplit(x, '_'))[1])))/length(patients)
-#Adjusted, `r round(adj, 2)*100`% have a >50% probability.
-```
-
-Unajdusted `r round(unadj,2)*100`% of BE patients have a >50% probability of progression.
 
@@
 ```{r roc, echo=T}
-#preds.info = base::merge(sample.info, preds, by='Samplename')
+si = do.call(rbind.data.frame,c(strsplit(as.character(pred.qc$Samplename), '_'), stringsAsFactors=F) )
+if (ncol(si) == 4) {
+  colnames(si) = colnames(sample.info)[1:4]
+} else {
+  colnames(si) = colnames(sample.info)[1:2]
+}
+pred.qc = cbind(si, pred.qc)
+pred.qc = base::merge( pred.qc, patient.info[,c('PatientID','Pathology')], by='PatientID')
+
 rows = grep('BLD|gastric', pred.qc$Samplename, ignore.case=T)
+if (length(rows) > 0) {
   normals = pred.qc[rows,]
   pred.qc = pred.qc[-rows,]
+}
 nrow(pred.qc)
 
-# Still missing some SNP data  FIX THIS CHECK
-si = do.call(rbind.data.frame,c(strsplit(as.character(pred.qc$Samplename), '_'), stringsAsFactors=F) )
-colnames(si) = colnames(sample.info)[1:4]
-
-pred.qc = cbind(si, pred.qc)
-
-setdiff(unique(patient.info$PatientID),unique(pred.qc$PatientID))
-
+# Three groups, prog, non prog (max LGD), non prog (max HGD)
+pred.qc = pred.qc %>% group_by(Status, PatientID) %>% dplyr::mutate( exclude=(Status == 'NP' & length(which(Pathology == 'HGD')) > 0) )
 pred.qc$Status = factor(pred.qc$Status)
 
-roc = pROC::roc(Status ~ Prob, data=pred.qc, auc=T, ci=T, of='thresholds')
+# Exclude those odd tetraploidy NPs
+pred.qc[which(with(pred.qc, (Samplename %in% highPloidyCN$Samplename))), 'exclude'] = T
+pred.qc[which(pred.qc$PatientID %in% subset(pred.qc, Samplename %in% highPloidyCN$Samplename)$PatientID), 'exclude'] = T
+
+# Exclude HGD NP patients
+pred = subset(pred.qc, !exclude)
+
+roc = pROC::roc(Status ~ Prob, data=pred, auc=T, ci=T, of='thresholds')
 roc.plot(roc)
 
-rocNoLow = pROC::roc(Status ~ Prob, data=subset(pred.qc, !Samplename %in% c(lowsca$Samplename,highPloidyCN$Samplename)), auc=T, ci=T, of='thresholds')
-roc.plot(rocNoLow)
+#rocNoLow = pROC::roc(Status ~ Prob, data=subset(pred, !Samplename %in% c(lowsca$Samplename)), auc=T, ci=T, of='thresholds')
+#roc.plot(rocNoLow)
 
-head(pred.qc)
-
-per_pt = pred.qc %>% group_by(PatientID, Timepoint, Status) %>% dplyr::summarise(
-  max.Prob = max(Prob), max.RR = max(RR)
-)
-
+per_pt = pred %>% group_by(PatientID, Timepoint, Status) %>% dplyr::summarise( max.Prob = max(Prob), max.RR = max(RR) )
 rocpp = pROC::roc(Status ~ max.Prob, data=per_pt, auc=T, ci=T, of='thresholds')
@@
 
-per_pt_ex = subset(pred.qc, !Samplename %in% c(lowsca$Samplename,highPloidyCN$Samplename)) %>% group_by(PatientID, Timepoint, Status) %>% dplyr::summarise(
+per_pt_ex = subset(pred, !Samplename %in% c(lowsca$Samplename,highPloidyCN$Samplename)) %>% group_by(PatientID, Timepoint, Status) %>% dplyr::summarise(
   max.Prob = max(Prob), max.RR = max(RR)
 )
-
 rocppex = pROC::roc(Status ~ max.Prob, data=per_pt_ex, auc=T, ci=T, of='thresholds')
@@
 
-roct2 = pROC::roc(Status ~ Prob, data=subset(pred.qc, Timepoint != 'T1' & !Samplename %in% c(lowsca$Samplename,highPloidyCN$Samplename)), auc=T, ci=T, of='thresholds')
+roct2 = pROC::roc(Status ~ Prob, data=subset(pred, Timepoint != 'T1' & !Samplename %in% c(lowsca$Samplename)), auc=T, ci=T, of='thresholds')
 roc.plot(roct2)

### 2018-07-16 SNP_Prediction_Analysis.Rmd
...

@@
 
-source('lib/data_func.R')
-source('lib/common_plots.R')
+suppressPackageStartupMessages(source('lib/data_func.R'))
+suppressPackageStartupMessages(source('lib/common_plots.R'))
 
-plot.sample<-function(segs, mer)
+chr.info = get.chr.lengths(file='~/tmp/hg19_info.txt')
 
-chr.info = get.chr.lengths(file='hg19_info.txt')
-
-#ptdirs = list.dirs('/Volumes/fh/fast/reid_b/collab/Killcoyne/Data/PerPatient', full.names=T, recursive=F)
-#patients = basename(ptdirs)
+adjust.cols<-function(mtx, means=NULL, sds=NULL) {
+  if (!is.null(means)) {
+    if (ncol(mtx) != length(means))
+      stop("Vector of means needs to match the columns in the matrix")
+    for (i in 1:ncol(mtx)) 
+      mtx[,i] = unit.var(mtx[,i], means[i], sds[i])
+  } else {
+    for (i in 1:ncol(mtx)) 
+      mtx[,i] = unit.var(mtx[,i])
+  }
+  return(mtx)
+}
 
@@
 load('~/Data/Ellie/Analysis/5e6_arms_all_logR/all.pt.alpha.Rdata', verbose = T)
-fitV = models$`0.9`
-l = performance.at.1se$`0.9`$lambda
+#alpha = '0'
+alpha = '0.9'
+fitV = models[[alpha]]
+l = performance.at.1se[[alpha]]$lambda
 
@@
 
-ggplot(sm, aes(Pathology, freq, group=Endoscopy, fill=factor(Endoscopy))) + facet_grid(~Status) + geom_col(position = position_dodge()) + ylim(0,1) + title('Pathology frequency per timepoint')
+ggplot(sm, aes(Pathology, freq, group=Endoscopy, fill=factor(Endoscopy))) + facet_grid(~Status) + 
+  geom_col(position = position_dodge()) + ylim(0,1) + labs(title='Pathology frequency per timepoint')
 ```
@@
 
-```{r, echo=F, eval=T}
+```{r, echo=F, warning=F}
 qcdata$SampleType = 'BE'
@@
 
+## PER ENDOSCOPY
+qcdata.samples = qcdata %>% group_by(Samplename) %>% dplyr::mutate(
+  PatientID = unlist(strsplit(Samplename, '_'))[1],
+  EndoID = unlist(strsplit(Samplename, '_'))[3],
+  Level = unlist(strsplit(Samplename, '_'))[4] 
+)
+
+qcdata = qcdata.samples %>% group_by(PatientID, EndoID, Status, Timepoint, SampleType) %>% dplyr::summarise(
+  Ploidy = mean(Ploidy), Purity = mean(Purity), Goodnessoffit = mean(Goodnessoffit), SCA.Ratio = mean(SCA.Ratio)
+)
+
 ## Low CN
@@
 #### P: Low SCA, Ploidy<=2
-`r round(nrow(lowPloidyCN)/nrow(subset(qcdata, Status == 'P' & SampleType == 'BE')), 3)*100`% of PROGRESSOR samples have <5% SCA and an ASCAT ploidy of 2 or less.
+`r round(nrow(lowPloidyCN)/nrow(subset(qcdata, Status == 'P' & SampleType == 'BE')), 3)*100`% of PROGRESSOR endoscopies have <5% SCA and an ASCAT ploidy of 2 or less.
 Most of these show a >90% purity as well.
@@
 #### P: No SCA
-`r round(nrow(noCN)/nrow(subset(qcdata, Status == 'P' & SampleType == 'BE')), 3)*100`% of PROGRESSOR samples have 0 SCA
+`r round(nrow(noCN)/nrow(subset(qcdata, Status == 'P' & SampleType == 'BE')), 3)*100`% of PROGRESSOR endoscopies have 0 SCA
 
@@
 
-```{r, echo=F}
+```{r, echo=F, warning=F}
 
@@
 
-totalSamples = table(sub('_.*','', qcdata$Samplename))
-lowscaSamples = table(sub('_.*','', lowsca$Samplename))
-
-# For those patients with low SCA samples, a large proportion of them show low SCA in most/all of their samples
-#plot(round(lowscaSamples/totalSamples[names(lowscaSamples)], 2))
-
 pt = '403'
 pfiles = list.files(paste('~/Data/Reid_SNP/PerPatient', pt, sep='/'), full.names = T)
-load(grep('ascat.Rdata', pfiles, value=T), verbose=T)
+load(grep('ascat.Rdata', pfiles, value=T), verbose=F)
 
@@
 pfiles = list.files(paste('~/Data/Reid_SNP/PerPatient', pt, sep='/'), full.names = T)
-load(grep('ascat.Rdata', pfiles, value=T), verbose=T)
+load(grep('ascat.Rdata', pfiles, value=T), verbose=F)
 
@@
 
-## Per endoscopy take the average or max bin value as an in silico pooling
-sample.info$Level = toupper(sample.info$Level)
+dim(mergedSegs)
+dim(mergedArms)
 
-range(mergedSegs[1,])
-range(x)
-plot( x )
+nm.rows = grep('BLD|GASTRIC', rownames(mergedSegs))
+normalSegs = mergedSegs[nm.rows,]
+normalArms = mergedArms[nm.rows,]
 
-x = (mergedSegs[1,]-mean(apply(allDf[,1:ncol(mergedSegs)], 1, mean)))/mean(apply(allDf[,1:ncol(mergedSegs)], 1, sd))
-  #sd(apply(allDf[,1:ncol(mergedSegs)], 1, sd))
+nrow(normalSegs) == nrow(normalArms)
 
-# names = colnames(mergedSegs)
-# mergedSegs = t(apply(mergedSegs, 1, scale, center=T))
-# colnames(mergedSegs) = names
-# 
-# names = colnames(mergedArms)
-# mergedArms = t(apply(mergedArms, 1, scale, center=T))
-# colnames(mergedArms) = names
+mergedSegs = mergedSegs[-nm.rows,]
+mergedArms = mergedArms[-nm.rows,]
 
-samples = unique(subset(sample.info, !Level %in% c('BLD','GASTRIC'), select=c("PatientID", "EndoID")))
-
-meanBinSegs = as.data.frame(matrix(ncol=ncol(mergedSegs), nrow=0, dimnames=list(c(), colnames(mergedSegs))))
-meanBinArms = as.data.frame(matrix(ncol=ncol(mergedArms), nrow=0, dimnames=list(c(), colnames(mergedArms))))
-
-mean.qcdata = data.frame(matrix(ncol=8, nrow=0, dimnames=list(c(), c('Samplename','Status','SampleType','Timepoint','mean.sca','mean.ploidy','mean.fit','mean.purity'))))
-
-for (i in 1:nrow(samples)) {
-  samples[i,]
-  print(paste(samples$PatientID[i], '(',i,')'))
-
-  #rows = subset(sample.info, PatientID == samples$PatientID[i] & EndoID == samples$EndoID[i] )$Samplename
-  ptendoID = paste(samples$PatientID[i], samples$EndoID[i], sep='_')
-
-  mean.qcdata[i,] = 
-    cbind('Samplename'=ptendoID, subset(qcdata, SampleType == 'BE' & grepl(paste(samples$PatientID[i], samples$EndoID[i], sep='.*'), Samplename)) %>% group_by(Status, SampleType, Timepoint) %>%
-    dplyr::summarise(mean.sca=mean(SCA.Ratio), mean.ploidy=mean(Ploidy), mean.fit=mean(Goodnessoffit), mean.purity=mean(Purity)))
-
-  rows = grep(paste(samples$PatientID[i], samples$EndoID[i], sep='.*'), rownames(mergedSegs), value=T)
-
-  sharedR = intersect(rows, rownames(mergedSegs))
-  if (length(rows) != length(sharedR)) {
-      warning(paste("Missing samples for", endo$PatientID))
-      rows = sharedR
-  }
-
-  #meanBinSegs[ptendoID,] = apply(mergedSegs[rows,,drop=F], 2, mean)
-  meanBinSegs[ptendoID,] = apply(mergedSegs[rows,,drop=F], 2, function(x) x[which.max(abs(x))] )
-  #meanBinArms[ptendoID,] = apply(mergedArms[rows,,drop=F], 2, mean)
-  meanBinArms[ptendoID,] = apply(mergedArms[rows,,drop=F], 2, function(x) x[which.max(abs(x))] )
-}
-
-normalSegs = mergedSegs[grep('BLD|gastric',rownames(mergedSegs),ignore.case=T),]
-normalArms = mergedArms[grep('BLD|gastric',rownames(mergedArms),ignore.case=T),]
-
-meanBinSegs = as.matrix(meanBinSegs) 
-meanBinArms = as.matrix(meanBinArms)
-
-dim(meanBinSegs)
+nrow(mergedSegs) == nrow(mergedArms)
 
@@
   theme(axis.text.x=element_blank() ) + labs(title='Low SCA samples', subtitle='patient 403')
-lowSeg = meanBinSegs[grep('^403', rownames(meanBinSegs), value=T),,drop=F]
-p2 = ggplot(melt(lowSeg), aes(Var2, value)) + facet_grid(~Var1, space='free_x', scales='free') + geom_point() +
-  theme(axis.text.x=element_blank() ) + labs(title='Low SCA samples, mean bins', subtitle='patient 403')
-
-grid.arrange(p1,p2)
 
 highSeg = mergedSegs[grep('^512', rownames(mergedSegs), value=T),,drop=F]
-p1 = ggplot(melt(highSeg), aes(Var2, value)) + facet_grid(~Var1, space='free_x', scales='free') + geom_point() +
+p2 = ggplot(melt(highSeg), aes(Var2, value)) + facet_grid(~Var1, space='free_x', scales='free') + geom_point() +
   theme(axis.text.x=element_blank() ) + labs(title='Mixed SCA samples', subtitle='patient 512')
-highSeg = meanBinSegs[grep('^512', rownames(meanBinSegs), value=T),,drop=F]
-p2 = ggplot(melt(highSeg), aes(Var2, value)) + facet_grid(~Var1, space='free_x', scales='free') + geom_point() +
-  theme(axis.text.x=element_blank() ) + labs(title='Mixed SCA samples, mean bins', subtitle='patient 512')
 
@@
 nm = melt(normalSegs)
-be = melt(mergedSegs[grep('BLD|gastric', rownames(mergedSegs), ignore.case=T, invert=T),])
-beM = melt(meanBinSegs[grep('BLD|gastric', rownames(meanBinSegs), ignore.case=T, invert=T),])
+be = melt(mergedSegs[-nm.rows,])
 
-lims = range(nm$value, be$value, beM$value)
+lims = range(nm$value, be$value)
 
@@
   ggplot(nm, aes(Var2, value, group=Var2)) + geom_point() + labs(title='Blood/gastric') + theme(axis.text.x = element_blank()) + ylim(lims),
-  ggplot(be, aes(Var2, value, group=Var2)) + geom_point() + labs(title='BE samples') + theme(axis.text.x = element_blank()) + ylim(lims),
-  ggplot(beM, aes(Var2, value, group=Var2)) + geom_point() + labs(title='BE samples, mean per endo') + theme(axis.text.x = element_blank()) + ylim(lims)
+  ggplot(be, aes(Var2, value, group=Var2)) + geom_point() + labs(title='BE samples') + theme(axis.text.x = element_blank()) + ylim(lims)
 )
@@
 
-Using a random normal sample and a known BE progresser sampe
+Using a random normal sample and a known BE progresser sample
 
-```{r rnormals, echo=F, fig.width=8, fig.height=8, eval=F}
+```{r rnormals, echo=F, fig.width=8, fig.height=8, eval=T}
 randNormal = sample(rownames(normalSegs), 1)
 pt = unlist(strsplit(randNormal, '_'))[1]
-segs = subset(segments.list[[pt]], chr %in% c(1:22) & grepl(randNormal, sample))
+segs = subset(segments.list[[pt]], chr %in% c(1:22) & grepl(sub('_', '.*', randNormal),sample))
 segs$chr = factor(segs$chr, levels=c(1:22), ordered = T)
 
-lims = range(segs$medLRR, mergedSegs[randNormal,])
+lims = range(segs$medLRR, normalSegs[randNormal,])
 grid.arrange(
@@
     theme(axis.text.x=element_blank() ) + labs(title='ASCAT median LRR', subtitle=pt),
-  ggplot(melt(as.matrix(mergedSegs[randNormal,])), aes(Var1, value)) + geom_point() + ylim(lims) +
+  ggplot(melt(as.matrix(normalSegs[randNormal,])), aes(Var1, value)) + geom_point() + ylim(lims) +
     labs(title='CN Adjusted & winsorized') + theme(axis.text.x=element_blank()),
@@
 
-randBE = '512_18762_133R_27.LogR'
+randBE = '512_133R'
+randBE = '1006_190V'
 pt = unlist(strsplit(randBE, '_'))[1]
-segs = subset(segments.list[[pt]], chr %in% c(1:22) & grepl(randBE, sample))
+segs = subset(segments.list[[pt]], chr %in% c(1:22) & grepl(sub('_', '.*',randBE), sample))
 segs$chr = factor(segs$chr, levels=c(1:22), ordered = T)
@@
 grid.arrange(
-  ggplot(segs) + facet_grid(~chr, space='free_x', scales='free') + ylim(lims) +
+  ggplot(segs) + facet_grid(~chr, space='free_x', scales='free')  +
     geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), size=3) + 
     theme(axis.text.x=element_blank() ) + labs(title='Median LRR, ASCAT segments', subtitle=pt), 
-  ggplot(melt(as.matrix(mergedSegs[sub('\\.LogR','',randBE),])), aes(Var1, value)) + geom_point() + ylim(lims) + 
+  ggplot(melt(as.matrix(mergedSegs[sub('\\.LogR','',randBE),])), aes(Var1, value)) + geom_point() +
     theme(axis.text.x=element_blank()) + labs(title='Winsorized LRR, binned'),  
+  ggplot(melt(adjust.cols(mergedSegs[sub('\\.LogR','',randBE),,drop=F], z.mean, z.sd)), aes(Var2, value)) + geom_point() +
+    theme(axis.text.x=element_blank()) + labs(title='Col adjusted LRR, binned'),  
 top='BE', bottom=randBE)
+
+sa = subtract.arms(adjust.cols(mergedSegs[sub('\\.LogR','',randBE),,drop=F], z.mean, z.sd),
+              adjust.cols(mergedArms[sub('\\.LogR','',randBE),,drop=F], z.arms.mean, z.arms.sd))[,1:589]
+ggplot(melt(as.matrix(sa)), aes(Var1, value)) + geom_point()
 
@@
 
-Using the mean bin values.
+Using the mean bin values (merged during tiling).
 ```{r, warning=F}
-adjust.cols<-function(mtx, means=NULL, sds=NULL) {
-  if (!is.null(means)) {
-    if (ncol(mtx) != length(means))
-      stop("Vector of means needs to match the columns in the matrix")
-    for (i in 1:ncol(mtx)) 
-      mtx[,i] = unit.var(mtx[,i], means[i], sds[i])
-  } else {
-    for (i in 1:ncol(mtx)) 
-      mtx[,i] = unit.var(mtx[,i])
-  }
-  return(mtx)
-}
+# z.mean[is.na(z.mean)] = median(z.mean,na.rm=T)
+# z.arms.mean[is.na(z.arms.mean)] = median(z.arms.mean,na.rm=T)
+# 
+# z.sd[is.na(z.sd)] = median(z.sd,na.rm=T)
+# z.arms.sd[is.na(z.arms.sd)] = median(z.arms.sd,na.rm=T)
 
 ns = adjust.cols(normalSegs, z.mean, z.sd)
+#ns = apply(normalSegs, 2, unit.var)
+range(ns)
+
 na = adjust.cols(normalArms, z.arms.mean, z.arms.sd)
+#na = apply(normalArms, 2, unit.var)
+range(na)
 
-ms = adjust.cols(meanBinSegs, z.mean, z.sd)
 ms = adjust.cols(mergedSegs, z.mean, z.sd)
+#ms = apply(mergedSegs, 2, unit.var)
 range(ms)
@@
 
-ma = adjust.cols(meanBinArms, z.arms.mean, z.arms.sd)
 ma = adjust.cols(mergedArms, z.arms.mean, z.arms.sd)
+#ma = apply(mergedArms, 2, unit.var)
 range(ma)
@@
 nm = melt(ns)
-be = melt(ms[grep('BLD|gastric', rownames(ms), ignore.case=T, invert=T),])
+be = melt(ms)
 
@@
     theme(axis.text.x=element_blank()) + ylim(range(nm$value, be$value)), top='Unit scaled')
+
+grid.arrange(
+  ggplot(melt(ms[sub('\\.LogR','',randBE),,drop=F]), aes(Var2, value)) + geom_point() +
+    theme(axis.text.x=element_blank()) + labs(title='Col adjusted LRR, binned'),  
+  ggplot(melt(subtract.arms(ms[sub('\\.LogR','',randBE),,drop=F],ma[sub('\\.LogR','',randBE),,drop=F])), aes(Var2, value)) +
+    geom_point() + theme(axis.text.x=element_blank()) + labs(title='Arms subtracted col adjusted LRR, binned'),  
+  top=randBE)
+# So what if I don't subtract the arms?
 
@@
 ```{r echo=F}
-cx = score.cx(ms, 1)
-arrayDf = subtract.arms(ms, ma)
-arrayDf = cbind(arrayDf, 'cx'=unit.var(cx, mn.cx, sd.cx))
-
 cxN = score.cx(ns, 1)
 arrayDfNorm = subtract.arms(ns, na)
+#arrayDfNorm = cbind(ns, na)
 arrayDfNorm = cbind(arrayDfNorm, 'cx'=unit.var(cxN, mn.cx, sd.cx))
@@
 
+preds.normal = cbind.data.frame(rownames(probN),probN, predict(fitV, newx=arrayDfNorm, s=l, type='link'))
+colnames(preds.normal) = c('Samplename','Prob', 'RR')
+
+cx = score.cx(ms, 1)
+#arrayDf = cbind(ms,ma)
+arrayDf = subtract.arms(ms, ma)
+
+arrayDf[,589:ncol(arrayDf)] = arrayDf[,589:ncol(arrayDf)]+abs(mean(arrayDf[,589:ncol(arrayDf)]))
+
+arrayDf = cbind(arrayDf, 'cx'=unit.var(cx, mn.cx, sd.cx))
+
+grid.arrange(
+  ggplot(melt(arrayDf), aes(Var2, value)) + geom_point() + 
+    geom_point(data=melt(arrayDf[,589:634]), col='red') + labs(title='SNP'),
+  ggplot(melt(allDf), aes(Var2, value)) + geom_point() +
+      geom_point(data=melt(allDf[,589:634]), col='red') + labs(title='sWGS'),
+nrow=1)
+
+# grid.arrange(
+#     ggplot(melt(arrayDf), aes(Var2, value)) + geom_point() + geom_point(data=melt(arrayDf[,589:634]+abs(mean(arrayDf[,589:634])*2)), col='red') + labs(title='SNP'),
+#     ggplot(melt(allDf), aes(Var2, value)) + geom_point() +
+#       geom_point(data=melt(allDf[,589:634]), col='red') + labs(title='sWGS'),
+#   nrow=1)
+
+prob = predict(fitV, newx=arrayDf, s=l, type='response')
+rr = predict(fitV, newx=arrayDf, s=l, type='link')
+
+preds = cbind.data.frame(rownames(prob),prob, rr, stringsAsFactors=F)
+colnames(preds) = c('Samplename','Prob', 'RR')
+
+## TEMP: I accidentally included normals by failing to account for difference in name
+#preds = preds[-grep('_\\d+$', rownames(preds)),]
+
 #qqnorm(allDf, main='Normal Q-Q plot, sWGS')
@@
 
+```{r, eval=F}
+x = apply(mergedSegs, 1, function(x) {
+  ds = c()
+  for (i in 1:(length(x)-1)) {
+    ds = c(ds, abs(x[[i]]) - abs(x[[i+1]]))
+  }
+  abs(diff(range(ds)))
+})
+plot(x, col=pred.qc$Status)
+
+x1 = apply(ms, 1, function(x) {
+  ds = c()
+  for (i in 1:(length(x)-1)) {
+    ds = c(ds, abs(x[[i]]) - abs(x[[i+1]]))
+  }
+  abs(diff(range(ds)))
+})
+plot(x1, col=pred.qc$Status)
+
+y = apply(normalSegs, 1, function(x) {
+  ds = c()
+  for (i in 1:(length(x)-1)) {
+    ds = c(ds, abs(x[[i]]) - abs(x[[i+1]]))
+  }
+  abs(diff(range(ds)))
+})
+plot(y)
+
+range(y)
+range(x)
+
+nprgs = subset(pred.qc, Status == 'NP')
+prgs = subset(pred.qc, Status == 'P')
+
+t.test(x[prgs$Samplename],x[nprgs$Samplename]) 
+t.test(x[prgs$Samplename],y)
+t.test(x[nprgs$Samplename],y)
+```
+
 # Predictions
@@
 ```{r, warning=F, fig.height=8}
-prob = predict(fitV, newx=arrayDf, s=l, type='response')
-rr = predict(fitV, newx=arrayDf, s=l, type='link')
-
-preds = cbind.data.frame(rownames(prob),prob, rr)
-colnames(preds) = c('Samplename','Prob', 'RR')
-
 #offset = log(0.0225)
@@
 A quick look at the two random samples from earlier.  
-`r pander(subset(preds, Samplename == randNormal), caption='Random normal')`
+`r pander(subset(preds.normal, Samplename == randNormal), caption='Random normal')`
 
-The random BE shows a very high probability:
+The random BE should show a very high probability, but for some reason is missed entirely:
 `r pander(subset(preds, Samplename == randBE), caption='Random BE')`
@@
 ```{r, warning=F, fig.height=8}
-# normals = subset(preds, grepl('BLD|gastric', Samplename, ignore.case=T))
-# grid.arrange(
-#   ggplot(normals, aes(Prob)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
-#     scale_fill_gradientn(colors = myPal,  name='') + 
-#     labs(title='Predictions on BLD/Gastric', y='n Samples', x='Unadjusted Probability'),
-# top='BLD/gastric normals' )
-# 
-# normals[which(normals$Prob > 0.5),]
+ggplot(preds.normal, aes(Prob)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
+  scale_fill_gradientn(colors = myPal,  name='') +
+  labs(title='Predictions on BLD/Gastric', y='n Samples', x='Unadjusted Probability')
 
-pred.qc = base::merge(preds, mean.qcdata, by='Samplename')
-#pred.qc = base::merge(preds, qcdata, by='Samplename')
+qcdata.be = subset(qcdata, SampleType == 'BE')
 
-prog = subset(pred.qc, !grepl('BLD|gastric', Samplename, ignore.case=T) & Status == 'P')
+get.qcinfo<-function(vec) {
+  endo = vec[2]
+  pt = vec[1]
+  y = subset(qcdata.be, PatientID == pt & EndoID == endo)
+  y = y %>% group_by(Status, Timepoint) %>%  dplyr::summarise( mean(Ploidy), mean(Purity), mean(Goodnessoffit), mean(SCA.Ratio) )
+  cbind.data.frame('PatientID'=pt, 'EndoID'=endo, y, stringsAsFactors = F)
+}
+
+qc = do.call(rbind.data.frame, c( lapply(preds$Samplename, function(x) get.qcinfo( unlist(strsplit(x, '_'))) ), stringsAsFactors=F))
+
+pred.qc = cbind.data.frame(preds, qc)
+dim(pred.qc)
+table(pred.qc$Status)
+
+ggplot(pred.qc, aes(`mean(SCA.Ratio)`, Prob, color=Status)) + geom_point()
+ggplot(pred.qc, aes(`mean(Ploidy)`, Prob, color=Status)) + geom_point()
+
+pander(cor.test(pred.qc$`mean(SCA.Ratio)`, pred.qc$Prob))
+
+prog = subset(pred.qc, Status == 'P')
 ggplot(prog, aes(Prob)) + facet_grid(~Timepoint) + 
@@
 
-nonprog = subset(pred.qc, !grepl('BLD|gastric', Samplename, ignore.case=T) & Status == 'NP')
+nonprog = subset(pred.qc, Status == 'NP')
 ggplot(nonprog, aes(Prob)) + facet_grid(~Timepoint) + 
@@
 
+Using my thresholds
+
 ```{r roc, echo=T}
-si = do.call(rbind.data.frame,c(strsplit(as.character(pred.qc$Samplename), '_'), stringsAsFactors=F) )
-if (ncol(si) == 4) {
-  colnames(si) = colnames(sample.info)[1:4]
-} else {
-  colnames(si) = colnames(sample.info)[1:2]
-}
-pred.qc = cbind(si, pred.qc)
-pred.qc = base::merge( pred.qc, patient.info[,c('PatientID','Pathology')], by='PatientID')
-
-rows = grep('BLD|gastric', pred.qc$Samplename, ignore.case=T)
-if (length(rows) > 0) {
-  normals = pred.qc[rows,]
-  pred.qc = pred.qc[-rows,]
-}
-nrow(pred.qc)
+pred.qc = base::merge( pred.qc, patient.info[,c('PatientID', 'Timepoint Code','Pathology')], by.x=c('PatientID', 'EndoID'), by.y=c('PatientID','Timepoint Code') )
 
 # Three groups, prog, non prog (max LGD), non prog (max HGD)
-pred.qc = pred.qc %>% group_by(Status, PatientID) %>% dplyr::mutate( exclude=(Status == 'NP' & length(which(Pathology == 'HGD')) > 0) )
+pred.qc = pred.qc %>% group_by(Status, PatientID) %>% dplyr::mutate( 
+  exclude=(Status == 'NP' & length(which(Pathology == 'HGD')) > 0) )
 pred.qc$Status = factor(pred.qc$Status)
 
+# pred.qc.ex = subset(pred.qc, !exclude)
+# cuts = seq(0,1,0.1)
+# 
+# pred.qc.ex$quants = with(pred.qc.ex, cut(Prob, breaks=cuts))
+# qt = as.data.frame.matrix(table(pred.qc.ex$quants, pred.qc.ex$Status))
+# qt$quant = rownames(qt)
+
+get.risk<-function(prob) {
+  if (prob < 0.3) return('Low')
+  if (prob >= 0.3 & prob < 0.7) return('Moderate')
+  if (prob >= 0.7) return('High')
+}
+
+# All data
+pred.qc = pred.qc %>% dplyr::rowwise() %>% dplyr::mutate(
+  Risk = get.risk(Prob)
+)
+pred.qc$Risk = factor(pred.qc$Risk, levels=c('Low','Moderate','High'), ordered = T)
+
+roc = pROC::roc(Status ~ Prob, data=pred.qc, auc=T, ci=T, of='thresholds')
+roc.plot(roc) + labs(title='All, Prob')
+pander(pROC::coords(roc, 'best'), caption='All samples, ROC best')
+
+roc = pROC::roc(Status ~ Risk, data=pred.qc, auc=T, ci=T, of='thresholds')
+roc.plot(roc) + labs(title='All, Risk')
+
 # Exclude those odd tetraploidy NPs
-pred.qc[which(with(pred.qc, (Samplename %in% highPloidyCN$Samplename))), 'exclude'] = T
-pred.qc[which(pred.qc$PatientID %in% subset(pred.qc, Samplename %in% highPloidyCN$Samplename)$PatientID), 'exclude'] = T
+#pred.qc[which(with(pred.qc, (Samplename %in% highPloidyCN$Samplename))), 'exclude'] = T
+pred.qc[which(pred.qc$PatientID %in% highPloidyCN$PatientID & pred.qc$EndoID %in% highPloidyCN$EndoID), 'exclude'] = T
 
 # Exclude HGD NP patients
-pred = subset(pred.qc, !exclude)
+roc = pROC::roc(Status ~ Prob, data=subset(pred.qc, !exclude), auc=T, ci=T, of='thresholds')
+roc.plot(roc) + labs(title='Excluded low sca/NP tetra), Prob')
 
-roc = pROC::roc(Status ~ Prob, data=pred, auc=T, ci=T, of='thresholds')
-roc.plot(roc)
+roc = pROC::roc(Status ~ Risk, data=subset(pred.qc, !exclude), auc=T, ci=T, of='thresholds')
+roc.plot(roc) + labs(title='Excluded low sca/NP tetra), Risk')
 
-#rocNoLow = pROC::roc(Status ~ Prob, data=subset(pred, !Samplename %in% c(lowsca$Samplename)), auc=T, ci=T, of='thresholds')
-#roc.plot(rocNoLow)
+per_pt = 
+  subset(pred.qc, !exclude) %>% group_by(PatientID, Status) %>% dplyr::summarise( max.Prob = max(Prob), max.RR = max(RR), max.Risk = max(Risk) )
+rocpp = pROC::roc(Status ~ max.Prob, data=per_pt, auc=T, ci=T, of='thresholds')
+roc.plot(rocpp) + labs(title='Max per patient (excluded low sca/NP tetra), Prob')
 
-per_pt = pred %>% group_by(PatientID, Timepoint, Status) %>% dplyr::summarise( max.Prob = max(Prob), max.RR = max(RR) )
-rocpp = pROC::roc(Status ~ max.Prob, data=per_pt, auc=T, ci=T, of='thresholds')
-roc.plot(rocpp)
+rocpp = pROC::roc(Status ~ max.Risk, data=per_pt, auc=T, ci=T, of='thresholds')
+roc.plot(rocpp) + labs(title='Max per patient (excluded low sca/NP tetra), Risk')
 
-per_pt_ex = subset(pred, !Samplename %in% c(lowsca$Samplename,highPloidyCN$Samplename)) %>% group_by(PatientID, Timepoint, Status) %>% dplyr::summarise(
-  max.Prob = max(Prob), max.RR = max(RR)
+# Exclude the low SCA patients
+pred.qc[which(pred.qc$PatientID %in% lowsca$PatientID & pred.qc$EndoID %in% lowsca$EndoID), 'exclude'] = T
+
+per_pt_ex = subset(pred.qc, !exclude) %>% group_by(PatientID, Timepoint, Status) %>% dplyr::summarise(
+  max.Prob = max(Prob), max.RR = max(RR), max.Risk = max(Risk)
 )
 rocppex = pROC::roc(Status ~ max.Prob, data=per_pt_ex, auc=T, ci=T, of='thresholds')
-roc.plot(rocppex)
+roc.plot(rocppex) + labs(title='Max per patient, Prob')
 
-roct2 = pROC::roc(Status ~ Prob, data=subset(pred, Timepoint != 'T1' & !Samplename %in% c(lowsca$Samplename)), auc=T, ci=T, of='thresholds')
-roc.plot(roct2)
+rocppex = pROC::roc(Status ~ max.Risk, data=per_pt_ex, auc=T, ci=T, of='thresholds')
+roc.plot(rocppex) + labs(title='Max per patient, Risk')
+
+roct2 = pROC::roc(Status ~ Prob, data=subset(pred.qc, Timepoint != 'T1' & !exclude), auc=T, ci=T, of='thresholds')
+roc.plot(roct2) + labs(title='T2, Prob')
+
+roct2 = pROC::roc(Status ~ Risk, data=subset(pred.qc, Timepoint != 'T1' & !exclude), auc=T, ci=T, of='thresholds')
+roc.plot(roct2) + labs(title='T2, Risk')
+
+ggplot(subset(pred.qc, !exclude), aes(`mean(SCA.Ratio)`,Prob, color=Risk, shape=Status)) + 
+  facet_grid(~Status) + geom_point()
+
+ggplot(subset(pred.qc, !exclude), aes(`mean(Purity)`,Prob, color=Risk, shape=Status)) + 
+  facet_grid(~Status) + geom_point()
+
+ex.pred = subset(pred.qc, !exclude)
+
+# Quick check
+nrow(subset(ex.pred, Risk == 'High' & Status == 'P'))/table(ex.pred$Status)['P']
+nrow(subset(ex.pred, Risk == 'Low' & Status == 'P'))/table(ex.pred$Status)['P'] #False negatives
+
+ggplot(melt(arrayDf), aes(Var2, value)) + geom_point() + 
+  geom_point(data=melt(arrayDf['1006_190V',,drop=F]), color='red') + 
+  geom_point(data=melt(arrayDf['1044_39Z',,drop=F]), color='blue')
+
+ggplot(melt(dysplasia.df), aes(Var2, value)) + geom_point()
+
+grid.arrange(
+  ggplot(melt(dysplasia.df[,589:ncol(dysplasia.df)]), aes(Var2, value)) + geom_point(),
+  ggplot(melt(arrayDf[,589:ncol(arrayDf)]), aes(Var2, value)) + geom_point())
+
+nrow(subset(ex.pred, Risk == 'Low' & Status == 'NP'))/table(ex.pred$Status)['NP']
+nrow(subset(ex.pred, Risk == 'High' & Status == 'NP'))/table(ex.pred$Status)['NP'] # False positive
 

### 2018-07-16 SNP_Prediction_Analysis.Rmd
...

@@
 
+`r nrow(patient.info)` total endoscopies, `r pander(table(patient.info$Status), caption="Split between P and NP")`
+
+Histology breakdown:
+
 ```{r, echo=F}
@@
 There's a significant difference in the numbers of NDBE or HGD samples between the groups
-```{r}
-chisq.test(table(patient.info$Pathology, patient.info$Status)['HGD',])
-chisq.test(table(patient.info$Pathology, patient.info$Status)['NDBE',])
+```{r, echo=F}
+pander(chisq.test(table(patient.info$Pathology, patient.info$Status)['HGD',]), caption='HGD')
+pander(chisq.test(table(patient.info$Pathology, patient.info$Status)['NDBE',]), caption='NDBE')
 
 # Most progressors are HGD at timepoint 1
-with(subset(patient.info, Endoscopy == 1), table(Pathology, Status))['HGD',]/table(subset(patient.info, Endoscopy == 1)$Status)
+pander( round(with(subset(patient.info, Endoscopy == 1), table(Pathology, Status))['HGD',]/table(subset(patient.info, Endoscopy == 1)$Status), 3)*100, caption='% at T1')
 ```
@@
 
-```{r, echo=F, warning=F}
+```{r, echo=F, warning=F, message=F}
 qcdata$SampleType = 'BE'
@@
 
-```{r, echo=F, warning=F}
+```{r, echo=F, warning=F, fig.height=10, fig.width=8}
 
@@
 lowsca = subset(qcdata, Status == 'P' & SampleType == 'BE' & SCA.Ratio <= 0.05 & Ploidy <= 2)
+
+summary.file = '~/Data/Annalise/NonP/summary_logR.txt'
+smy = read.table(summary.file, sep = '\t', header = T)
+smy = subset(smy, Copy.number <= 12)
+
+a = as.data.frame(smy[,c('Copy.number','var.LRR')])
+colnames(a) = c('x','y')
+b = as.data.frame(smy[,c('Copy.number','median.LRR')])
+colnames(b) = c('x','y')
+b[b$x == 2,'y'] = 0
+
+fitSD = lm(x~y, data=a)
+autoplot(fitSD)
+fitM = lm(x~y, data=b)
+autoplot(fitM)
+
+# Adjust based on the summary values per CN that we get in WGS (mostly) NP Barrett's cases
+adjust.segraw<-function(segraw) {
+  segraw = segraw %>% rowwise() %>% dplyr::mutate( total = nMajor+nMinor )
+  segraw = subset(segraw, chr %in% c(1:22))
+
+  segraw$adjLRR = NA
+  for (cn in unique(segraw$total)) {
+    rows = which(segraw$total == cn)
+    values = seg[rows, 'medLRR', drop=T]
+
+    #newM = mean(values)
+    #newSD = sd(values)
+
+    #if (cn != 2) {
+      #newM = predict(fitM, newdata=cbind.data.frame('x'=cn))
+      #newSD = predict(fitSD, newdata=cbind.data.frame('x'=cn))
+
+    #newV = mean(values)
+
+    #}
+    #segraw[rows,][['adjLRR']] = newM + (values - mean(values)) * (newSD/sd(values))
+
+    segraw[rows,'adjLRR'] = values * predict(fitM, newdata=cbind.data.frame('y'=values)) - cn)
+  }  
+segraw  
+}
 
@@
 seg = ascat.output$segments_raw
-seg$chr = factor(seg$chr, levels=c(1:22,'X','Y'))
-ggplot(seg) + facet_grid(sample~chr, space='free_x', scales='free') + 
+seg = adjust.segraw(seg)
+
+m = (melt(seg, measure.vars=c('medLRR','adjLRR')))
+ggplot(m, aes(total, value, group=total)) + facet_grid(~variable) + geom_boxplot() + labs(title=pt)
+
+seg$chr = factor(seg$chr, levels=c(1:22))
+grid.arrange(
+  ggplot(seg) + facet_grid(sample~chr, space='free_x', scales='free_x') +
     geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), color='darkgreen', size=3) + 
-  theme(axis.text.x=element_blank() ) + labs(title='Low SCA', subtitle='patient 403')
+    theme(axis.text.x=element_blank() ) + labs(title='Low SCA, medLRR', subtitle=paste('patient',pt)),
+  ggplot(seg) + facet_grid(sample~chr, space='free_x', scales='free_x') + 
+    geom_segment(aes(x=startpos, xend=endpos, y=adjLRR, yend=adjLRR), color='darkred', size=3) + 
+    theme(axis.text.x=element_blank() ) + labs(title='Low SCA, adjLRR', subtitle=paste('patient',pt))
+)
 
@@
 seg = ascat.output$segments_raw
-seg$chr = factor(seg$chr, levels=c(1:22,'X','Y'))
-ggplot(seg) + facet_grid(sample~chr, space='free_x', scales='free') + 
+seg = adjust.segraw(seg)
+
+m = (melt(seg, measure.vars=c('medLRR','adjLRR')))
+ggplot(m, aes(total, value, group=total)) + facet_grid(~variable) + geom_boxplot() + labs(title=pt)
+
+seg$chr = factor(seg$chr, levels=c(1:22))
+grid.arrange(
+  ggplot(seg) + facet_grid(sample~chr, space='free_x', scales='free_x') + 
     geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), color='darkgreen', size=3) + 
-  theme(axis.text.x=element_blank() ) + labs(title='Mixed SCA samples', subtitle='patient 512')
+    theme(axis.text.x=element_blank() ) + labs(title='Mixed SCA, medLRR', subtitle=paste('patient',pt)),
+  ggplot(seg) + facet_grid(sample~chr, space='free_x', scales='free_x') + 
+    geom_segment(aes(x=startpos, xend=endpos, y=adjLRR, yend=adjLRR), color='darkred', size=3) + 
+    theme(axis.text.x=element_blank() ) + labs(title='Mixed SCA, adjLRR', subtitle=paste('patient',pt))
+)
 
@@
 
-```{r, warning=F, echo=F}
+```{r, warning=F, echo=F, message=F, fig.height=10, fig.width=8}
 
@@
 
-```{r normals, fig.width=6, fig.height=8, eval=F}
+```{r normals, message=F, fig.height=10, fig.width=8}
 nm = melt(normalSegs)
@@
 grid.arrange(
-  ggplot(nm, aes(Var2, value, group=Var2)) + geom_point() + labs(title='Blood/gastric') + theme(axis.text.x = element_blank()) + ylim(lims),
-  ggplot(be, aes(Var2, value, group=Var2)) + geom_point() + labs(title='BE samples') + theme(axis.text.x = element_blank()) + ylim(lims)
+  ggplot(nm, aes(Var2, value, group=Var2)) + geom_point() + labs(title='Blood/gastric',subtitle=paste('variance=',var(nm$value),sep=''),x='') + 
+    theme(axis.text.x = element_blank()) + ylim(lims),
+  ggplot(be, aes(Var2, value, group=Var2)) + geom_point() + labs(title='BE samples',subtitle=paste('variance=',var(be$value),sep=''),x='') + 
+    theme(axis.text.x = element_blank()) + ylim(lims)
 )
@@
 
-```{r rnormals, echo=F, fig.width=8, fig.height=8, eval=T}
+```{r rnormals, echo=F, warning=F, message=F, fig.width=10, fig.height=8}
 randNormal = sample(rownames(normalSegs), 1)
@@
 
-randBE = '512_133R'
+#randBE = '512_133R'
 randBE = '1006_190V'
@@
 
-sa = subtract.arms(adjust.cols(mergedSegs[sub('\\.LogR','',randBE),,drop=F], z.mean, z.sd),
-              adjust.cols(mergedArms[sub('\\.LogR','',randBE),,drop=F], z.arms.mean, z.arms.sd))[,1:589]
-ggplot(melt(as.matrix(sa)), aes(Var1, value)) + geom_point()
-
 ```
@@
 Using the mean bin values (merged during tiling).
-```{r, warning=F}
+```{r, warning=F, message=F, fig.width=10, fig.height=8}
 # z.mean[is.na(z.mean)] = median(z.mean,na.rm=T)
@@
   top=randBE)
-# So what if I don't subtract the arms?
 
@@
 
-Normal variance: `r var(nm$value, na.rm=T)`
-
-BE variance: `r var(be$value, na.rm=T)`
-
-```{r echo=F}
+```{r echo=F,message=F, fig.width=10, fig.height=8}
 cxN = score.cx(ns, 1)
@@
 
-# grid.arrange(
-#     ggplot(melt(arrayDf), aes(Var2, value)) + geom_point() + geom_point(data=melt(arrayDf[,589:634]+abs(mean(arrayDf[,589:634])*2)), col='red') + labs(title='SNP'),
-#     ggplot(melt(allDf), aes(Var2, value)) + geom_point() +
-#       geom_point(data=melt(allDf[,589:634]), col='red') + labs(title='sWGS'),
-#   nrow=1)
-
 prob = predict(fitV, newx=arrayDf, s=l, type='response')
@@
 
-## TEMP: I accidentally included normals by failing to account for difference in name
-#preds = preds[-grep('_\\d+$', rownames(preds)),]
-
-#qqnorm(allDf, main='Normal Q-Q plot, sWGS')
-#qqnorm(arrayDf, main='Normal Q-Q plot, SNP arrays')
+qqnorm(allDf, main='Normal Q-Q plot, sWGS')
+qqnorm(arrayDf, main='Normal Q-Q plot, SNP arrays')
 

### 2018-07-16 process_snp.R

@@ -32,8 +32,9 @@ colnames(a) = c('x','y')
 b = as.data.frame(smy[,c('Copy.number','median.LRR')])
 colnames(b) = c('x','y')
 
-fitSD = lm(y~x, data=a)
-fitM = lm(y~x, data=b)
+#fitSD = lm(y~x, data=a)
+#fitM = lm(y~x, data=b)
+fitCN = lm(y~x, data=b)
 
 ### Lifted directly from ASCAT
 #Perform MAD winsorization:
@@ -93,28 +94,38 @@ segraw = segraw %>% rowwise() %>% dplyr::mutate( total = nMajor+nMinor )
 
 # Adjust based on the summary values per CN that we get in WGS (mostly) NP Barrett's cases
 segraw$adjLRR = NA
-for (cn in unique(segraw$total)) {
-  rows = which(segraw$total == cn)
-  values = segraw[rows, 'medLRR', drop=T]
+# for (cn in unique(segraw$total)) {
+#   rows = which(segraw$total == cn)
+#   values = segraw[rows, 'medLRR', drop=T]
+#   
+#   newM = mean(values)
+#   newSD = sd(values)
+#   
+#   if (cn < 2) {
+#     newM = predict(fitM, newdata=cbind.data.frame('x'=cn))
+#     newSD = predict(fitSD, newdata=cbind.data.frame('x'=cn))
+#    # segraw[rows,][['adjLRR']] = newM + (values - mean(values)) * (newSD/sd(values))
+#   #}
+#   # if (cn < 2) {
+#   #   direction = ifelse(meanV < 0, -1, 1)
+#   #   newM = meanV + newM
+#    } else if (cn > 2) {
+#      #newM = meanV + newM
+#      #newSD = sdV + newSD
+#    }
+#   segraw[rows,][['adjLRR']] = newM + (values - mean(values)) * (newSD/sd(values))
+#   #segraw[rows,][['adjLRR']] = values
+# }  
+
+adjust.segraw<-function(segraw) {
+  segraw$total = with(segraw, nAraw+nBraw)
+  segraw = subset(segraw, chr %in% c(1:22))
   
-  newM = mean(values)
-  newSD = sd(values)
-  
-  if (cn < 2) {
-    newM = predict(fitM, newdata=cbind.data.frame('x'=cn))
-    newSD = predict(fitSD, newdata=cbind.data.frame('x'=cn))
[…]

### 2018-07-17 SNP_Prediction_Analysis.Rmd
...

@@
 library(pROC)
+library(ggfortify)
 
@@
 
-```{r, echo=F, warning=F, fig.height=10, fig.width=8}
+```{r, echo=F, warning=F, fig.height=10, fig.width=8,eval=F}
 
@@
 
+fitCN = lm(y~x, data=b)
+
 # Adjust based on the summary values per CN that we get in WGS (mostly) NP Barrett's cases
 adjust.segraw<-function(segraw) {
-  segraw = segraw %>% rowwise() %>% dplyr::mutate( total = nMajor+nMinor )
+  segraw = segraw %>% rowwise() %>% dplyr::mutate( total = nAraw+nBraw )
   segraw = subset(segraw, chr %in% c(1:22))
@@
   segraw$adjLRR = NA
-  for (cn in unique(segraw$total)) {
-    rows = which(segraw$total == cn)
-    values = seg[rows, 'medLRR', drop=T]
+  #for (cn in unique(segraw$total)) {
+  #  rows = which(segraw$total == cn)
+  #  values = seg[rows, 'medLRR', drop=T]
 
@@
     #segraw[rows,][['adjLRR']] = newM + (values - mean(values)) * (newSD/sd(values))
+   # smy
+    #median(predict(fitM, newdata=cbind.data.frame('y'=values)))-cn
 
-    segraw[rows,'adjLRR'] = values * predict(fitM, newdata=cbind.data.frame('y'=values)) - cn)
-  }  
+    new = predict(fitCN, newdata=cbind.data.frame('x'=segraw$total))
+    #plot(scale(new, center = T))
+
+    segraw[,'adjLRR'] = scale(new,center=T)
+
+   # segraw[rows,'adjLRR'] = (values * (predict(fitM, newdata=cbind.data.frame('y'=values)) - cn))
+  #}  
 segraw  
@@
 m = (melt(seg, measure.vars=c('medLRR','adjLRR')))
+m$total = round(round(m$total, 1))
 ggplot(m, aes(total, value, group=total)) + facet_grid(~variable) + geom_boxplot() + labs(title=pt)
@@
 
-```{r rnormals, echo=F, warning=F, message=F, fig.width=10, fig.height=8}
+```{r rnormals, echo=F, warning=F, message=F, fig.width=10, fig.height=8, eval=F}
 randNormal = sample(rownames(normalSegs), 1)
@@
 
-qqnorm(allDf, main='Normal Q-Q plot, sWGS')
-qqnorm(arrayDf, main='Normal Q-Q plot, SNP arrays')
+grid.arrange(
+  ggplot( melt(allDf), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, sWGS'),
+  ggplot( melt(arrayDf), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, SNP arrays'),
+  ncol=1
+)
 
@@
 
+length(which(preds.normal$Prob > 0.5))/nrow(preds.normal)
+
 qcdata.be = subset(qcdata, SampleType == 'BE')
@@
 pred.qc = cbind.data.frame(preds, qc)
-dim(pred.qc)
-table(pred.qc$Status)
+#dim(pred.qc)
+#table(pred.qc$Status)
 
@@
 
-# pred.qc.ex = subset(pred.qc, !exclude)
-# cuts = seq(0,1,0.1)
-# 
-# pred.qc.ex$quants = with(pred.qc.ex, cut(Prob, breaks=cuts))
-# qt = as.data.frame.matrix(table(pred.qc.ex$quants, pred.qc.ex$Status))
-# qt$quant = rownames(qt)
+pred.qc.ex = subset(pred.qc, !exclude)
+cuts = seq(0,1,0.1)
+
+pred.qc.ex$quants = with(pred.qc.ex, cut(Prob, breaks=cuts))
+qt = as.data.frame.matrix(table(pred.qc.ex$quants, pred.qc.ex$Status))
+qt$quant = rownames(qt)
+pander(qt)
 

### 2018-07-17 process_snp.R

@@ -28,13 +28,15 @@ smy = read.table(summary.file, sep = '\t', header = T)
 smy = subset(smy, Copy.number <= 12)
 
 a = as.data.frame(smy[,c('Copy.number','var.LRR')])
-colnames(a) = c('x','y')
-b = as.data.frame(smy[,c('Copy.number','median.LRR')])
-colnames(b) = c('x','y')
+colnames(a) = c('CN','y')
+b = as.data.frame(smy[,c('Copy.number','mean.LRR')])
+colnames(b) = c('CN','y')
 
-#fitSD = lm(y~x, data=a)
-#fitM = lm(y~x, data=b)
-fitCN = lm(y~x, data=b)
+fitSD = lm(y~CN, data=a)
+fitM = lm(y~CN, data=b)
+#fitCN = lm(mean~CN, data=b)
+
+autoplot(fitM)
 
 ### Lifted directly from ASCAT
 #Perform MAD winsorization:
@@ -90,52 +92,66 @@ segraw = subset(segraw, chr %in% c(1:22))
 
 head(segraw)
 
-segraw = segraw %>% rowwise() %>% dplyr::mutate( total = nMajor+nMinor )
+#segraw = segraw %>% rowwise() %>% dplyr::mutate( total = nMajor+nMinor )
 
 # Adjust based on the summary values per CN that we get in WGS (mostly) NP Barrett's cases
-segraw$adjLRR = NA
-# for (cn in unique(segraw$total)) {
-#   rows = which(segraw$total == cn)
-#   values = segraw[rows, 'medLRR', drop=T]
-#   
-#   newM = mean(values)
-#   newSD = sd(values)
-#   
-#   if (cn < 2) {
-#     newM = predict(fitM, newdata=cbind.data.frame('x'=cn))
-#     newSD = predict(fitSD, newdata=cbind.data.frame('x'=cn))
-#    # segraw[rows,][['adjLRR']] = newM + (values - mean(values)) * (newSD/sd(values))
-#   #}
-#   # if (cn < 2) {
-#   #   direction = ifelse(meanV < 0, -1, 1)
-#   #   newM = meanV + newM
-#    } else if (cn > 2) {
-#      #newM = meanV + newM
-#      #newSD = sdV + newSD
-#    }
-#   segraw[rows,][['adjLRR']] = newM + (values - mean(values)) * (newSD/sd(values))
-#   #segraw[rows,][['adjLRR']] = values
-# }  
+
 
 adjust.segraw<-function(segraw) {
[…]

### 2018-07-18 SNP_Prediction_Analysis.Rmd
...

@@
 
-adjust.cols<-function(mtx, means=NULL, sds=NULL) {
+adjust.cols<-function(mtx, means=NULL, sds=NULL, na.replace=0) {
   if (!is.null(means)) {
@@
   }
+  mtx[is.na(mtx)] = na.replace
+
   return(mtx)
@@
 
+# Adjust for ploidy
+for (i in 1:nrow(mergedSegs)) {
+  sample = rownames(mergedSegs)[i]
+  x = unlist(strsplit(sample, '_'))
+  ploidy = mean(subset(qcdata, SampleType == 'BE' & PatientID == x[1] & EndoID == x[2])$Ploidy)
+  if (grepl('BLD|gastric', sample, ignore.case = T)) ploidy = 2
+  mergedSegs[i,] = mergedSegs[i,]-(ploidy-1)
+  mergedArms[i,] = mergedArms[i,]-(ploidy-1)
+}
+
 nm.rows = grep('BLD|GASTRIC', rownames(mergedSegs))
-normalSegs = mergedSegs[nm.rows,]
-normalArms = mergedArms[nm.rows,]
+#normalSegs = mergedSegs[nm.rows,]
+#normalArms = mergedArms[nm.rows,]
 
-nrow(normalSegs) == nrow(normalArms)
-
-mergedSegs = mergedSegs[-nm.rows,]
-mergedArms = mergedArms[-nm.rows,]
-
-nrow(mergedSegs) == nrow(mergedArms)
+# nrow(normalSegs) == nrow(normalArms)
+# 
+# mergedSegs = mergedSegs[-nm.rows,]
+# mergedArms = mergedArms[-nm.rows,]
+# 
+# nrow(mergedSegs) == nrow(mergedArms)
 
@@
 
-grid.arrange(p1,p2)
+grid.arrange(p1,p2,ncol=1)
 ```
@@
 ```{r normals, message=F, fig.height=10, fig.width=8}
-nm = melt(normalSegs)
+#nm = melt(normalSegs)
+nm = melt(mergedSegs[nm.rows,])
 be = melt(mergedSegs[-nm.rows,])
 
-lims = range(nm$value, be$value)
+lims = range(nm$value, be$value, na.rm=T)
 
@@
     theme(axis.text.x = element_blank()) + ylim(lims),
-  ggplot(be, aes(Var2, value, group=Var2)) + geom_point() + labs(title='BE samples',subtitle=paste('variance=',var(be$value),sep=''),x='') + 
+  ggplot(be, aes(Var2, value, group=Var2)) + geom_point() + labs(title='BE samples',subtitle=paste('variance=',var(be$value,na.rm=T),sep=''),x='') + 
     theme(axis.text.x = element_blank()) + ylim(lims)
@@
 lims = range(segs$medLRR, normalSegs[randNormal,])
-grid.arrange(
-  ggplot(segs) + facet_grid(~chr, space='free_x', scales='free') + ylim(lims) +
-    geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), size=3) + 
-    theme(axis.text.x=element_blank() ) + labs(title='ASCAT median LRR', subtitle=pt),
-  ggplot(melt(as.matrix(normalSegs[randNormal,])), aes(Var1, value)) + geom_point() + ylim(lims) +
-    labs(title='CN Adjusted & winsorized') + theme(axis.text.x=element_blank()),
-top='Random BLD or gastric normal', bottom=randNormal)
+# grid.arrange(
+#   ggplot(segs) + facet_grid(~chr, space='free_x', scales='free') + ylim(lims) +
+#     geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), size=3) + 
+#     theme(axis.text.x=element_blank() ) + labs(title='ASCAT median LRR', subtitle=pt),
+#   ggplot(melt(as.matrix(normalSegs[randNormal,])), aes(Var1, value)) + geom_point() + ylim(lims) +
+#     labs(title='CN Adjusted & winsorized') + theme(axis.text.x=element_blank()),
+# top='Random BLD or gastric normal', bottom=randNormal)
 
@@
 
-lims = range(segs$medLRR, mergedSegs[sub('\\.LogR','',randBE),])
-grid.arrange(
-  ggplot(segs) + facet_grid(~chr, space='free_x', scales='free')  +
-    geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), size=3) + 
-    theme(axis.text.x=element_blank() ) + labs(title='Median LRR, ASCAT segments', subtitle=pt), 
-  ggplot(melt(as.matrix(mergedSegs[sub('\\.LogR','',randBE),])), aes(Var1, value)) + geom_point() +
-    theme(axis.text.x=element_blank()) + labs(title='Winsorized LRR, binned'),  
-  ggplot(melt(adjust.cols(mergedSegs[sub('\\.LogR','',randBE),,drop=F], z.mean, z.sd)), aes(Var2, value)) + geom_point() +
-    theme(axis.text.x=element_blank()) + labs(title='Col adjusted LRR, binned'),  
-top='BE', bottom=randBE)
+#lims = range(segs$medLRR, mergedSegs[sub('\\.LogR','',randBE),])
+# grid.arrange(
+#   ggplot(segs) + facet_grid(~chr, space='free_x', scales='free')  +
+#     geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), size=3) + 
+#     theme(axis.text.x=element_blank() ) + labs(title='Median LRR, ASCAT segments', subtitle=pt), 
+#   ggplot(melt(as.matrix(mergedSegs[sub('\\.LogR','',randBE),])), aes(Var1, value)) + geom_point() +
+#     theme(axis.text.x=element_blank()) + labs(title='Winsorized LRR, binned'),  
+#  ggplot(melt(adjust.cols(mergedSegs[sub('\\.LogR','',randBE),,drop=F], z.mean, z.sd)), aes(Var2, value)) + geom_point() + theme(axis.text.x=element_blank()) + labs(title='Col adjusted LRR, binned'),
+#top='BE', bottom=randBE)
 
@@
 ```{r, warning=F, message=F, fig.width=10, fig.height=8}
-# z.mean[is.na(z.mean)] = median(z.mean,na.rm=T)
-# z.arms.mean[is.na(z.arms.mean)] = median(z.arms.mean,na.rm=T)
-# 
-# z.sd[is.na(z.sd)] = median(z.sd,na.rm=T)
-# z.arms.sd[is.na(z.arms.sd)] = median(z.arms.sd,na.rm=T)
+#ns = adjust.cols(normalSegs, z.mean, z.sd)
+#ns = adjust.cols(normalSegs)
+#range(ns)
 
-ns = adjust.cols(normalSegs, z.mean, z.sd)
-#ns = apply(normalSegs, 2, unit.var)
-range(ns)
+#na = adjust.cols(normalArms, z.arms.mean, z.arms.sd)
+#na = adjust.cols(normalArms)
+#range(na)
 
-na = adjust.cols(normalArms, z.arms.mean, z.arms.sd)
-#na = apply(normalArms, 2, unit.var)
-range(na)
-
-ms = adjust.cols(mergedSegs, z.mean, z.sd)
-#ms = apply(mergedSegs, 2, unit.var)
+#ms = adjust.cols(mergedSegs, z.mean, z.sd)
+ms = adjust.cols(mergedSegs)
 range(ms)
@@
 
-ma = adjust.cols(mergedArms, z.arms.mean, z.arms.sd)
-#ma = apply(mergedArms, 2, unit.var)
+# mergedSegs['997_122V',which(mergedSegs['997_122V',] <= 0)]
+# ms['997_122V',which(mergedSegs['997_122V',] <= 0)]
+
+#ma = adjust.cols(mergedArms, z.arms.mean, z.arms.sd)
+ma = adjust.cols(mergedArms)
 range(ma)
@@
 
-nm = melt(ns)
-be = melt(ms)
+nm = melt(ms[nm.rows,])
+be = melt(ms[-nm.rows,])
 
+lims = range(nm$value, be$value, na.rm = T)
 grid.arrange(
-  ggplot(nm, aes(Var2, value, group=Var2)) + geom_point() + labs(title='Blood/gastric') + 
-    theme(axis.text.x=element_blank()) +ylim(range(nm$value, be$value)),
-  ggplot(be, aes(Var2, value, group=Var2)) + geom_point() + labs(title='BE samples') + 
-    theme(axis.text.x=element_blank()) + ylim(range(nm$value, be$value)), top='Unit scaled')
-
-grid.arrange(
-  ggplot(melt(ms[sub('\\.LogR','',randBE),,drop=F]), aes(Var2, value)) + geom_point() +
-    theme(axis.text.x=element_blank()) + labs(title='Col adjusted LRR, binned'),  
-  ggplot(melt(subtract.arms(ms[sub('\\.LogR','',randBE),,drop=F],ma[sub('\\.LogR','',randBE),,drop=F])), aes(Var2, value)) +
-    geom_point() + theme(axis.text.x=element_blank()) + labs(title='Arms subtracted col adjusted LRR, binned'),  
-  top=randBE)
+  ggplot(nm, aes(Var2, value, group=Var2, color=Var1)) + geom_point() + labs(title='Blood/gastric') + 
+    theme(axis.text.x=element_blank(),legend.position='none') +ylim(lims),
+  ggplot(be, aes(Var2, value, group=Var2, color=Var1)) + geom_point() + labs(title='BE samples') + 
+    theme(axis.text.x=element_blank(),legend.position='none') + ylim(lims), top='Unit scaled')
 
@@
 ```{r echo=F,message=F, fig.width=10, fig.height=8}
-cxN = score.cx(ns, 1)
-arrayDfNorm = subtract.arms(ns, na)
+#cxN = score.cx(ns, 1)
+#arrayDfNorm = subtract.arms(ns, na)
 #arrayDfNorm = cbind(ns, na)
-arrayDfNorm = cbind(arrayDfNorm, 'cx'=unit.var(cxN, mn.cx, sd.cx))
+#arrayDfNorm = cbind(arrayDfNorm, 'cx'=unit.var(cxN, mn.cx, sd.cx))
 
-probN = predict(fitV, newx=arrayDfNorm, s=l, type='response')
-length(which(probN > 0.5))/nrow(probN) # FPR
-length(which(probN <= 0.5))/nrow(probN)
+#probN = predict(fitV, newx=arrayDfNorm, s=l, type='response')
+#length(which(probN > 0.5))/nrow(probN) # FPR
+#length(which(probN <= 0.5))/nrow(probN)
 
-preds.normal = cbind.data.frame(rownames(probN),probN, predict(fitV, newx=arrayDfNorm, s=l, type='link'))
-colnames(preds.normal) = c('Samplename','Prob', 'RR')
+#preds.normal = cbind.data.frame(rownames(probN),probN, predict(fitV, newx=arrayDfNorm, s=l, type='link'))
+#colnames(preds.normal) = c('Samplename','Prob', 'RR')
 
@@
 arrayDf = subtract.arms(ms, ma)
+#arrayDf = cbind(arrayDf, 'cx'=unit.var(cx, mn.cx, sd.cx))
+arrayDf = cbind(arrayDf, 'cx'=unit.var(cx))
 
-arrayDf[,589:ncol(arrayDf)] = arrayDf[,589:ncol(arrayDf)]+abs(mean(arrayDf[,589:ncol(arrayDf)]))
-
-arrayDf = cbind(arrayDf, 'cx'=unit.var(cx, mn.cx, sd.cx))
-
+lims=range(arrayDf, allDf,na.rm=T)
 grid.arrange(
-  ggplot(melt(arrayDf), aes(Var2, value)) + geom_point() + 
+  ggplot(melt(arrayDf), aes(Var2, value)) + geom_point() + ylim(lims) +
     geom_point(data=melt(arrayDf[,589:634]), col='red') + labs(title='SNP'),
-  ggplot(melt(allDf), aes(Var2, value)) + geom_point() +
+  ggplot(melt(allDf), aes(Var2, value)) + geom_point() + ylim(lims) +
       geom_point(data=melt(allDf[,589:634]), col='red') + labs(title='sWGS'),
@@
 
-prob = predict(fitV, newx=arrayDf, s=l, type='response')
-rr = predict(fitV, newx=arrayDf, s=l, type='link')
+```
 
-preds = cbind.data.frame(rownames(prob),prob, rr, stringsAsFactors=F)
-colnames(preds) = c('Samplename','Prob', 'RR')
-
+```{r, echo=F, fig.height=12, fig.width=6}
 grid.arrange(
@@
 )
+
+grid.arrange(
+  ggplot(melt(allDf[,1,drop=F]), aes(sample=value)) + stat_qq(),
+  ggplot(melt(arrayDf[,1,drop=F]), aes(sample=value)) + stat_qq(),
+  ncol=1)
+
+for (i in 1:ncol(arrayDf)) {
+  arrayDf[,i] = unit.var(arrayDf[,i], mean(allDf[,i]), sd(allDf[,i]))
+}
+
+qqnorm()
+sd(allDf[,1])
+
+mean(arrayDf[,1])
+sd(arrayDf[,1])
 
@@
 ```{r, warning=F, fig.height=8}
+
+prob = predict(fitV, newx=arrayDf, s=l, type='response')
+rr = predict(fitV, newx=arrayDf, s=l, type='link')
+
+preds = cbind.data.frame(rownames(prob),prob, rr, stringsAsFactors=F)
+colnames(preds) = c('Samplename','Prob', 'RR')
+
+pnr = grep('BLD|gastric', rownames(preds), ignore.case = T)
+preds.normal = preds[pnr,]
+
+length(which(preds.normal$Prob > 0.5))/nrow(preds.normal) # FPR
+length(which(preds.normal$Prob < 0.5))/nrow(preds.normal)
+
 #offset = log(0.0225)
@@
 
-length(which(preds.normal$Prob > 0.5))/nrow(preds.normal)
-
 qcdata.be = subset(qcdata, SampleType == 'BE')
-
 get.qcinfo<-function(vec) {
@@
 
-qc = do.call(rbind.data.frame, c( lapply(preds$Samplename, function(x) get.qcinfo( unlist(strsplit(x, '_'))) ), stringsAsFactors=F))
+qc = do.call(rbind.data.frame, c( lapply(preds$Samplename[-pnr], function(x) get.qcinfo( unlist(strsplit(x, '_'))) ), stringsAsFactors=F))
 
-pred.qc = cbind.data.frame(preds, qc)
+pred.qc = cbind.data.frame(preds[-pnr,], qc)
 #dim(pred.qc)
@@
 pred.qc$Status = factor(pred.qc$Status)
+
+## First of all we have samples from patients that probably had a genome duplication, there's no way to measure that in our sWGS data
+#subset(pred.qc, `mean(Ploidy)` > 2.7)
+
+```
+
+```{r echo=F}
 
@@
 
-roc = pROC::roc(Status ~ Risk, data=subset(pred.qc, !exclude), auc=T, ci=T, of='thresholds')
-roc.plot(roc) + labs(title='Excluded low sca/NP tetra), Risk')
-
 per_pt = 
@@
 roc.plot(rocpp) + labs(title='Max per patient (excluded low sca/NP tetra), Prob')
-
-rocpp = pROC::roc(Status ~ max.Risk, data=per_pt, auc=T, ci=T, of='thresholds')
-roc.plot(rocpp) + labs(title='Max per patient (excluded low sca/NP tetra), Risk')
 
@@
 
-rocppex = pROC::roc(Status ~ max.Risk, data=per_pt_ex, auc=T, ci=T, of='thresholds')
-roc.plot(rocppex) + labs(title='Max per patient, Risk')
-
 roct2 = pROC::roc(Status ~ Prob, data=subset(pred.qc, Timepoint != 'T1' & !exclude), auc=T, ci=T, of='thresholds')
@@
 
-roct2 = pROC::roc(Status ~ Risk, data=subset(pred.qc, Timepoint != 'T1' & !exclude), auc=T, ci=T, of='thresholds')
-roc.plot(roct2) + labs(title='T2, Risk')
-
 ggplot(subset(pred.qc, !exclude), aes(`mean(SCA.Ratio)`,Prob, color=Risk, shape=Status)) + 
   facet_grid(~Status) + geom_point()
+
+subset(pred.qc, `mean(SCA.Ratio)` < 0.1 & Status == 'P' & !exclude & Prob < 0.5)
 
@@
 # Quick check
-nrow(subset(ex.pred, Risk == 'High' & Status == 'P'))/table(ex.pred$Status)['P']
-nrow(subset(ex.pred, Risk == 'Low' & Status == 'P'))/table(ex.pred$Status)['P'] #False negatives
+nrow(subset(ex.pred, Prob > 0.5 & Status == 'P'))/table(ex.pred$Status)['P']
+nrow(subset(ex.pred, Prob < 0.5 & Status == 'P'))/table(ex.pred$Status)['P'] #False negatives
 
-ggplot(melt(arrayDf), aes(Var2, value)) + geom_point() + 
-  geom_point(data=melt(arrayDf['1006_190V',,drop=F]), color='red') + 
-  geom_point(data=melt(arrayDf['1044_39Z',,drop=F]), color='blue')
+nrow(subset(ex.pred, Prob > 0.5 & Status == 'NP'))/table(ex.pred$Status)['NP'] # False pos
+nrow(subset(ex.pred, Prob < 0.5 & Status == 'NP'))/table(ex.pred$Status)['NP'] #
 
-ggplot(melt(dysplasia.df), aes(Var2, value)) + geom_point()
-
-grid.arrange(
-  ggplot(melt(dysplasia.df[,589:ncol(dysplasia.df)]), aes(Var2, value)) + geom_point(),
-  ggplot(melt(arrayDf[,589:ncol(arrayDf)]), aes(Var2, value)) + geom_point())
-
-nrow(subset(ex.pred, Risk == 'Low' & Status == 'NP'))/table(ex.pred$Status)['NP']
-nrow(subset(ex.pred, Risk == 'High' & Status == 'NP'))/table(ex.pred$Status)['NP'] # False positive
+# ggplot(melt(arrayDf), aes(Var2, value)) + geom_point() + 
+#   geom_point(data=melt(arrayDf['1006_190V',,drop=F]), color='red') + 
+#   geom_point(data=melt(arrayDf['1044_39Z',,drop=F]), color='blue')
 

### 2018-07-18 process_snp.R

@@ -143,7 +143,7 @@ plotdir = paste(datadir,'/plots', sep='')
 if (dir.exists(plotdir)) unlist(plotdir, recursive=T)
 
 dir.create(plotdir, showWarnings = F, recursive = T)
-ggsave(filename= paste(plotdir,'rawCN.png',sep='/'), plot=p, width=length(unique(segraw$sample))*2, height=6,scale=1.5)
+ggsave(filename= paste(plotdir,'rawCN.png',sep='/'), plot=p, width=length(unique(segraw$sample))*2, height=6,scale=1.5, limitsize = F)
 save(segraw, file=paste(datadir, 'segments_raw.Rdata',sep='/'))
 
 allsamples = NULL
@@ -196,6 +196,8 @@ for (i in 1:length(rowsPerSample)) {
   #if (max(tiled[[valueCol]],na.rm=T) > 1) lims[2] = max(tiled$winsLRR,na.rm=T)
   #if (min(tiled$winsLRR,na.rm=T) < -1) lims[1] = min(tiled$winsLRR,na.rT)
   lims = range(tiled[[valueCol]], na.rm = T)
+  if (valueCol == 'totalRaw') lims=c(0,4)
+  
   tiled$chr = factor(tiled$chr, levels=chr.info$chr)
 
   p = ggplot(chr.info, aes(x=1:chr.length)) + ylim(lims) + facet_grid(~chr,space='free_x',scales='free_x') + geom_segment(data=tiled, aes(x=start,xend=end,y=tiled[,4],yend=tiled[,4]), size=3, color='darkgreen') + theme_bw() + theme(axis.text.x=element_blank(), panel.spacing.x=unit(0, 'lines')) + labs(x='', y=valueCol, title=names(rowsPerSample)[i])
@@ -203,7 +205,7 @@ for (i in 1:length(rowsPerSample)) {
   plist[[ names(rowsPerSample)[i] ]] = p
 }
 
-ggsave(filename = paste(plotdir,'tiled.png',sep='/'), plot = do.call(grid.arrange, c(plist, ncol=1)), width=10, height=5*length(plist))
+ggsave(filename = paste(plotdir,'tiled.png',sep='/'), plot = do.call(grid.arrange, c(plist, ncol=1)), width=10, height=5*length(plist), limitsize = F)
 
[…]

### 2018-07-20 SNP_Prediction_Analysis.Rmd
added cv model build/analysis

@@
 library(ggfortify)
+library(preprocessCore)
 
@@
 
-load('~/Data/Ellie/Analysis/5e6_arms_all_logR/model_data.Rdata', verbose = T)
-load('~/Data/Ellie/Analysis/5e6_arms_all_logR/all.pt.alpha.Rdata', verbose = T)
+load('~/Data/Ellie/Analysis/5e6_arms_all/model_data.Rdata', verbose = T)
+load('~/Data/Ellie/Analysis/5e6_arms_all/all.pt.alpha.Rdata', verbose = T)
 #alpha = '0'
@@
 l = performance.at.1se[[alpha]]$lambda
+features = coefs[[alpha]]
 
@@
 ## high CN | ploidy, nonP
-highPloidyCN = subset(qcdata, Status == 'NP' & SampleType == 'BE' & (SCA.Ratio >= mean(qcdata$SCA.Ratio,na.rm=T)+ sd(qcdata$SCA.Ratio,na.rm=T)*2 | Ploidy >= 2.3))
+highPloidyCN = subset(qcdata, Status == 'NP' & SampleType == 'BE' & (SCA.Ratio >= mean(qcdata$SCA.Ratio,na.rm=T)+ sd(qcdata$SCA.Ratio,na.rm=T)*2 | Ploidy >= 2.7))
+
+qcdata$SCA.Ratio = round(qcdata$SCA.Ratio, 4)
+
+lowsca = subset(qcdata, Status == 'P' & SampleType == 'BE' & SCA.Ratio <= 0.05 & Ploidy <= 2)
 
@@
 ```{r, echo=F, warning=F, fig.height=10, fig.width=8,eval=F}
-
-qcdata$SCA.Ratio = round(qcdata$SCA.Ratio, 4)
-
-lowsca = subset(qcdata, Status == 'P' & SampleType == 'BE' & SCA.Ratio <= 0.05 & Ploidy <= 2)
 
@@
 
+segmentVariance = apply(mergedSegs, 1, var)
+
 nm.rows = grep('BLD|GASTRIC', rownames(mergedSegs))
@@
 ```{r, echo=F, fig.height=12, fig.width=6}
+## TODO: quantile normalize the SNP arrays by the sWGS values. 
+copy = do.call(cbind.data.frame,lapply(1:ncol(arrayDf), function(i) {
+  as.data.frame(normalize.quantiles.use.target(arrayDf[,i,drop=F], as.vector(allDf[,i])))
+}))
+
+colnames(copy) = colnames(arrayDf)
+rownames(copy) = rownames(arrayDf)
+
 grid.arrange(
@@
   ggplot( melt(arrayDf), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, SNP arrays'),
+  ggplot( melt(copy), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, SNP arrays q-normed'),
   ncol=1
 )
-
-grid.arrange(
-  ggplot(melt(allDf[,1,drop=F]), aes(sample=value)) + stat_qq(),
-  ggplot(melt(arrayDf[,1,drop=F]), aes(sample=value)) + stat_qq(),
-  ncol=1)
-
-for (i in 1:ncol(arrayDf)) {
-  arrayDf[,i] = unit.var(arrayDf[,i], mean(allDf[,i]), sd(allDf[,i]))
-}
-
-qqnorm()
-sd(allDf[,1])
-
-mean(arrayDf[,1])
-sd(arrayDf[,1])
-
+arrayDf = as.matrix(copy)
 ```
@@
 
-pnr = grep('BLD|gastric', rownames(preds), ignore.case = T)
+# offset = log(0.0225)
+# preds = preds %>% dplyr::mutate(
+#   'Adj. RR'=RR+offset,
+#   'Adj. Prob'=1/(1+exp(-RR+abs(offset)) )
+# )
+
+pnr = grep('BLD|gastric', preds$Samplename, ignore.case = T)
 preds.normal = preds[pnr,]
@@
 length(which(preds.normal$Prob < 0.5))/nrow(preds.normal)
-
-#offset = log(0.0225)
-# preds = preds %>% dplyr::mutate( 
-#   'Adj. RR'=RR+offset,  
-#   'Adj. Prob'=1/(1+exp(-RR+abs(offset)) )
-# )
 
@@
 ```{r, warning=F, fig.height=8}
-ggplot(preds.normal, aes(Prob)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
-  scale_fill_gradientn(colors = myPal,  name='') +
-  labs(title='Predictions on BLD/Gastric', y='n Samples', x='Unadjusted Probability')
 
@@
 prog = subset(pred.qc, Status == 'P')
-ggplot(prog, aes(Prob)) + facet_grid(~Timepoint) + 
+p1 = ggplot(prog, aes(Prob)) + facet_wrap(~Timepoint, scales='free_y',nrow=1) + 
   geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
@@
 nonprog = subset(pred.qc, Status == 'NP')
-ggplot(nonprog, aes(Prob)) + facet_grid(~Timepoint) + 
+p2 = ggplot(nonprog, aes(Prob)) + facet_wrap(~Timepoint, scales='free_y', nrow=1) + 
   geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
@@
   labs(title='Predictions on non-progressors', y='n Samples', x='Unadjusted Probability')
+
+p3 = ggplot(preds.normal, aes(Prob)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
+  scale_fill_gradientn(colors = myPal,  name='') +
+  labs(title='Predictions on BLD/Gastric', y='n Samples', x='Unadjusted Probability')
+
+grid.arrange(p1,p2,p3,ncol=1)
 
@@
 
+# These samples are likely genome duplicated.
+dup = subset(pred.qc, `mean(Ploidy)` > 2.7)
+cor.test(dup$`mean(Ploidy)`, dup$Prob)
+table(dup$Status)
+
 # Three groups, prog, non prog (max LGD), non prog (max HGD)
 pred.qc = pred.qc %>% group_by(Status, PatientID) %>% dplyr::mutate( 
-  exclude=(Status == 'NP' & length(which(Pathology == 'HGD')) > 0) )
+  exclude=(Status == 'NP' & (length(which(Pathology == 'HGD')) > 0 | `mean(Ploidy)` > 2.7 )) )
 pred.qc$Status = factor(pred.qc$Status)
-
-## First of all we have samples from patients that probably had a genome duplication, there's no way to measure that in our sWGS data
-#subset(pred.qc, `mean(Ploidy)` > 2.7)
 
@@
 ```{r echo=F}
-
 pred.qc.ex = subset(pred.qc, !exclude)
@@
 qt$quant = rownames(qt)
+qt$P = round(qt$P/sum(qt$P),2)
+qt$NP = round(qt$NP/sum(qt$NP), 2)
 pander(qt)
 
+# These are the cutoffs from the sWGS data
 get.risk<-function(prob) {
@@
 
-roc = pROC::roc(Status ~ Risk, data=pred.qc, auc=T, ci=T, of='thresholds')
-roc.plot(roc) + labs(title='All, Risk')
+table(subset(pred.qc, Risk == 'High' & !exclude)$Status)/table(subset(pred.qc, !exclude)$Status)
+table(subset(pred.qc, Risk == 'Moderate' & !exclude)$Status)/table(subset(pred.qc, !exclude)$Status)
+table(subset(pred.qc, Risk == 'Low' & !exclude)$Status)/table(subset(pred.qc, !exclude)$Status)
 
@@
 pred.qc[which(pred.qc$PatientID %in% lowsca$PatientID & pred.qc$EndoID %in% lowsca$EndoID), 'exclude'] = T
+rocpp = pROC::roc(Status ~ Prob, data=subset(pred.qc, !exclude), auc=T, ci=T, of='thresholds')
+roc.plot(rocpp) + labs(title='Max per patient (excluded low sca/NP tetra), Prob')
+pROC::coords(rocpp, 'best')
 
-per_pt_ex = subset(pred.qc, !exclude) %>% group_by(PatientID, Timepoint, Status) %>% dplyr::summarise(
+as.data.frame(subset(pred.qc, Status == 'P' & Prob < 0.5 & !exclude))
+
+# Timepoint 2 only
+roct2 = pROC::roc(Status ~ Prob, data=subset(pred.qc, Timepoint != 'T1' & !exclude), auc=T, ci=T, of='thresholds')
+roc.plot(roct2) + labs(title='T2, Prob')
+
+per_pt_ex = subset(pred.qc, !exclude) %>% group_by(PatientID, Status) %>% dplyr::summarise(
   max.Prob = max(Prob), max.RR = max(RR), max.Risk = max(Risk)
@@
 roc.plot(rocppex) + labs(title='Max per patient, Prob')
-
-roct2 = pROC::roc(Status ~ Prob, data=subset(pred.qc, Timepoint != 'T1' & !exclude), auc=T, ci=T, of='thresholds')
-roc.plot(roct2) + labs(title='T2, Prob')
 
@@
 
-ggplot(subset(pred.qc, !exclude), aes(`mean(Purity)`,Prob, color=Risk, shape=Status)) + 
-  facet_grid(~Status) + geom_point()
+ggplot(subset(pred.qc, !exclude | ( `mean(SCA.Ratio)` > 0.1 & Status == 'P')) , aes(`mean(Purity)`,Prob, color=Risk, shape=Status)) + facet_grid(~Status) + geom_point()
 
@@
 
-# ggplot(melt(arrayDf), aes(Var2, value)) + geom_point() + 
-#   geom_point(data=melt(arrayDf['1006_190V',,drop=F]), color='red') + 
-#   geom_point(data=melt(arrayDf['1044_39Z',,drop=F]), color='blue')
-
-```
-
-# Regions predicted by model
-Not evaluated currently
-```{r coefs, eval=F}
-features = coefs$`0.9`[,1,drop=F]
-
-locs = do.call(rbind.data.frame, c(strsplit(rownames(features), ':|-'), stringsAsFactors=F))
-colnames(locs) = c('chr','start','end')
-locs[] = lapply(locs[], as.numeric)
-
-feature.locs = makeGRangesFromDataFrame(locs)
-
-gr = makeGRangesFromDataFrame(segments.list$`512`, start.field = 'startpos', end.field = 'endpos', keep.extra.columns = T)
-
-ov = findOverlaps(feature.locs, gr)
-
-feature.locs[unique(queryHits(ov))]
-
-cna = lapply( unique(queryHits(ov)), function(hit) {
-  tmp = gr[subjectHits(ov[queryHits(ov) == hit])]
-  tmp = subset(as.data.frame(tmp), nMajor+nMinor > 2.3 | nMajor+nMinor < 1.5) %>% group_by(sample, nMajor+nMinor) %>% dplyr::summarise( n=length(sample), medLRR=mean(medLRR), mn.bp=mean(end-start) ) %>% rowwise() %>%
-    dplyr::mutate( type=ifelse(`nMajor + nMinor` >= 3, 1, 0))
-})
-names(cna) = rownames(features)
-
-fl = sapply(cna, function(x) length(unique(x$sample)))
-feature.locs[which(fl == 0)]
-
-cna.count = do.call(rbind, lapply(cna, function(x) table(x$type)))
-
-cna.mx = as.matrix(sapply(cna, function(x) length(unique(x$sample))))
-cna.mx = cbind('sample.cnt'=cna.mx, 'loss'=NA,'gain'=NA, 'loss.bp', 'gain.bp')
-for (ft in rownames(features)) {
-  cna.mx[ft,c('loss','gain')] = c(0,0)
-  if (nrow(cna[[ft]]) > 0)
-    cna.mx[ft,c('loss','gain')] = c(length(which(cna[[ft]]$type == 0)),length(which(cna[[ft]]$type == 1)))
-
-  median(cna[[ft]] [[cna[[ft]]$type == 0,'mn.bp']] )
-
-}
-
-cna$`1:59820150-64805161`
+ ggplot(melt(arrayDf), aes(Var2, value)) + geom_point() + 
+  geom_point(data=melt(arrayDf['1006_190V',,drop=F]), color='red') + 
+  geom_point(data=melt(arrayDf['1044_39Z',,drop=F]), color='blue')
 
@@
 
-```{r regression, echo=F, eval=F}
-slabels = sample.info[,c('PatientID','Status', 'Samplename')]
+```{r regression, echo=F, message=F}
+slabels = as.data.frame(subset(pred.qc, !exclude, select=c('PatientID','Samplename','Status')))
+#slabels = as.data.frame(pred.qc, select=c('PatientID','Samplename','Status'))
+rownames(slabels) = slabels$Samplename
 
-missing = setdiff(sample.info$Samplename, rownames(arrayDf))
-rows = grep('BLD|gastric', rownames(arrayDf), ignore.case=T)
-bld_gastric = arrayDf[rows,]
+df = arrayDf[which(rownames(arrayDf) %in% slabels$Samplename),]
 
-df = arrayDf[-rows,]
-
-slabels = subset(slabels, Samplename %in% rownames(df))
-
-status = as.integer(as.factor(slabels$Status))-1
+status = as.integer(slabels$Status)-1
 names(status) = slabels$Samplename
-
-df = df[-grep(setdiff(rownames(df),names(status)), rownames(df)),]
-
-#df = arrayDf[which(!rownames(arrayDf) %in% c(lowsca$Samplename,highPloidyCN$Samplename)),]
 
@@
 nl = 1000;folds = 10; splits = 5 
+## ----- All ----- ##
+coefs = list(); plots = list(); performance.at.1se = list(); models = list(); cvs = list()
+cache.dir = '~/Data/Ellie/Analysis/SNP'
+dir.create(cache.dir, recursive = T, showWarnings = F)
+file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
+if (file.exists(file)) {
+ message(paste("loading", file))
+ load(file, verbose=T)
+} else {
+  alpha.values = c(0,0.5,0.7,0.9,1)
   sets = create.patient.sets(slabels, folds, splits, 0.2)
 
-#sets = create.patient.sets(subset(sample.info, !Samplename %in% c(lowsca$Samplename,highPloidyCN$Samplename)) [c('PatientID','Samplename','Status')], folds, splits, 0.2)
-
-alpha.values = c(0,0.5,0.9,1)
-## ----- All ----- ##
-coefs = list(); plots = list(); performance.at.1se = list(); models = list(); cvs = list()
-cache.dir = '~/Data/Ellie/Analysis/SNP/'
-dir.create(cache.dir, recursive = T, showWarnings = F)
-file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
-#if (file.exists(file)) {
-#  message(paste("loading", file))
-#  load(file, verbose=T)
-#} else 
-  {
   for (a in alpha.values) {
+    message(paste('alpha=',a))
     fit0 <- glmnet(df, status, alpha=a, nlambda=nl, family='binomial', standardize=F)    
-    autoplot(fit) + theme(legend.position="none")
+    autoplot(fit0) + theme(legend.position="none")
     l = fit0$lambda
+    if (a > 0) l = more.l(l)
 
@@
   }
-  save(plots, coefs, performance.at.1se, dysplasia.df, models, cvs, labels, file=file)
+  save(plots, coefs, performance.at.1se, df, models, cvs, labels, file=file)
   p = do.call(grid.arrange, c(plots[ as.character(alpha.values) ], top='All samples, 10fold, 5 splits'))
@@
 }
-all.coefs = coefs
-save(arrayDf,labels,patient.info, file=paste(cache.dir,'snp.Rdata',sep='/'))
+#save(arrayDf,labels,patient.info, file=paste(cache.dir,'snp-models.Rdata',sep='/'))
+
+do.call(grid.arrange, c(plots))
+
+model.performance(performance.at.1se, coefs)$plot + labs(title='SNP CV models')
+
 ```
+
+```{r compare.coefs, echo=F, eval=T}
+# Matching
+snp.features = coefs$`0.5`
+
+intersect(rownames(snp.features), rownames(features))
+
+cxisft = grep('cx',rownames(snp.features))
+
+rowSums(snp.features[,-1])/50
+
+snp.features = do.call(rbind.data.frame, c(strsplit(rownames(snp.features)[-cxisft], ':|-'), stringsAsFactors=F))
+colnames(snp.features) = c('chr','start','end')
+
+swgs.features = do.call(rbind.data.frame, c(strsplit(rownames(features), ':|-'), stringsAsFactors=F))
+colnames(swgs.features) = c('chr','start','end')
+
+snp.features = makeGRangesFromDataFrame(snp.features)
+snp.features = snp.features[-which(width(snp.features) > 5e6)] # No arms
+
+swgs.features = makeGRangesFromDataFrame(swgs.features)
+
+dist = distanceToNearest(swgs.features,snp.features)
+
+swgs.features[queryHits(dist)]
+snp.features[subjectHits(dist)]
+
+match = dist[elementMetadata(dist)$distance != 0]
+elementMetadata(dist)$bins = round(elementMetadata(dist)$distance/5e6)
+
+x1 = dist[elementMetadata(dist)$bins <= 2]
+
+swgs.features[queryHits(x1)]
+snp.features[subjectHits(x1)]
+```
+
+```{r loo}
+
+slabels$Prediction = NA
+slabels$RR = NA
+
+file = paste(cache.dir, 'loo.Rdata', sep='/')
+if (file.exists(file)) {
+  load(file, verbose=T)
+} else {
+  select.alpha = 1
+  secf = coefs[[select.alpha]]
+  a = select.alpha
+
+  performance.at.1se = c(); plots = list(); fits = list(); nzcoefs = list()
+  # Remove each patient (LOO)
+  for (pt in unique(slabels$PatientID)) {
+    print(pt)
+    samples = subset(slabels, PatientID != pt)$Samplename
+
+    train.rows = which(rownames(df) %in% samples)
+    training = df[train.rows,]
+    test = as.matrix(df[-train.rows,])
+    if ( nrow(test) == ncol(df) ) test = t(test)
+
+    # Predict function giving me difficulty when I have only a single sample, this ensures the dimensions are the same
+    sparsed_test_data <- Matrix(data=0, nrow=nrow(test),  ncol=ncol(training),
+                                dimnames=list(rownames(test),colnames(training)), sparse=T)
+    for(i in colnames(arrayDf)) sparsed_test_data[,i] = test[,i]
+
+    # Fit generated on all samples, including HGD
+    fitLOO <- glmnet(training, status[train.rows], alpha=a, family='binomial', nlambda=nl, standardize=F) # all patients
+    l = fitLOO$lambda
+
+    cv = crossvalidate.by.patient(x=training, y=status[train.rows], lambda=l, a=a, nfolds=folds, splits=splits,
+                                  pts=subset(sets, Samplename %in% samples), fit=fitLOO, standardize=F)
+
+    plots[[pt]] = arrangeGrob(cv$plot+ggtitle('Classification'), cv$deviance.plot+ggtitle('Binomial Deviance'), top=pt, ncol=2)
+    fits[[pt]] = cv
+
+    if ( length(cv$lambda.1se) > 0 ) {
+      performance.at.1se = c(performance.at.1se, subset(cv$lambdas, lambda == cv$lambda.1se)$mean)
+
+      coef.1se = coef(fitLOO, cv$lambda.1se)[rownames(secf),]
+
+      nzcoefs[[pt]] = as.data.frame(non.zero.coef(fitLOO, cv$lambda.1se))
+
+      #coefs[[pt]] = coef(fitLOO, cv$lambda.1se)[rownames(secf),]
+      #coef.stability(coef.1se, cv$non.zero.cf)
+
+      logit <- function(p){log(p/(1-p))}
+      inverse.logit <- function(or){1/(1 + exp(-or))}
+
+      pm = predict(fitLOO, newx=sparsed_test_data, s=cv$lambda.1se, type='response')
+      or = predict(fitLOO, newx=sparsed_test_data, s=cv$lambda.1se, type='link')
+
+      slabels[which(slabels$Samplename %in% rownames(pm)),'Prediction'] = pm[,1]
+      slabels[which(slabels$Samplename %in% rownames(pm)),'RR'] = or[,1]
+
+    } else {
+      warning(paste(pt, "did not have a 1se"))
+    }
+  }
+  save(plots, performance.at.1se, nzcoefs, fits, slabels, file=file)
+}
+
+snp.roc = pROC::roc(Status~Prediction,slabels, auc=T, ci=T, of='thresholds')
+roc.plot(snp.roc, title="SNP LOO Model ROC")
+
+ggplot(cbind.data.frame('performance'=performance.at.1se, 'type'='snp'), aes(y=performance, x=type, group=type,fill=type)) + 
+  geom_boxplot(outlier.colour = NA, show.legend = F) + geom_jitter(width=0.2) + geom_label(data=cbind.data.frame('performance'=range(performance.at.1se), 'type'='snp'), aes(label=round(performance,2)), nudge_x=0.1) + theme_bw() + theme(legend.position = 'none') + labs(x='', y='Classification', title='LOO Performance SNP')
+
+file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
+load(file, verbose=T)
+
+fit.snp = models$`1`
+
+probs = predict(fit.snp, newx = dysplasia.df, s = performance.at.1se$`1`$lambda,type='response' )
+
+swgs.pred.snp = cbind.data.frame('Prob'=probs[,1], 'Status'=swgs_labels[rownames(probs)])
+ssroc = pROC::roc(Status~Prob, data=swgs.pred.snp, ci=T, auc=T, of='thresholds')
+roc.plot(ssroc, title='sWGS prediction on SNP model')
+```

### 2018-07-24 AllSamplesModel.Rmd
regression for snp data

@@
 
-cache.dir = paste(data, 'Analysis/5e6_arms_all_logR', sep='/')
+cache.dir = paste(data, 'Analysis/5e6_arms_all', sep='/')
 
@@
   'Median age'=paste(median(age.diagnosis),'+/-',round(sd(age.diagnosis),1)),
+  'Median length'=median(M,na.rm=T),
   'gender F:M'=paste(table(gender),collapse=':')
@@
 transform.by.time<-function(pt, info) {
-  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction', 'OR')] %>% group_by(PID, Samplename) %>% dplyr::mutate('months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12)
+  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Maximal','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction', 'OR')] %>% group_by(PID, Samplename) %>% dplyr::mutate('months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12)
   pt = dplyr::arrange(merge(pt, subset(info, Hospital.Research.ID == pt$Hospital.Research.ID[1], select=c('Samplename','Block')), by='Samplename'), Pathology, PID, Endoscopy.Year)

### 2018-07-24 SNP_Prediction_Analysis.Rmd
regression for snp data

@@
 
+patient.info = patient.info %>% rowwise() %>% dplyr::mutate( Endoscopy = ifelse(Timepoint == 'T1', 1, 2) )
 ```
@@
 ```{r, echo=F}
-
-patient.info = patient.info %>% rowwise() %>% dplyr::mutate( Endoscopy = ifelse(Timepoint == 'T1', 1, 2) )
-
 sm = patient.info %>% group_by(Status, Endoscopy, Pathology) %>% dplyr::tally() %>% group_by(Status, Endoscopy) %>% dplyr::mutate(freq=n/sum(n) )
@@
 
-```{r rnormals, echo=F, warning=F, message=F, fig.width=10, fig.height=8, eval=F}
-randNormal = sample(rownames(normalSegs), 1)
+```{r rnormals, echo=F, warning=F, message=F, fig.width=10, fig.height=8, eval=T}
+randNormal = sample(rownames(mergedSegs)[nm.rows], 1)
 pt = unlist(strsplit(randNormal, '_'))[1]
@@
 
-lims = range(segs$medLRR, normalSegs[randNormal,])
+lims = range(segs$medLRR, mergedSegs[randNormal,])
 # grid.arrange(
@@
 
+Adjust using the RAW values...
+
+```{r raw.adjust, fig.height=16, fig.width=4, echo=T}
+q.norm.by<-function(df, target) {
+  dfq = do.call(cbind.data.frame,lapply(1:ncol(df), function(i) {
+    as.data.frame(normalize.quantiles.use.target(df[,i,drop=F], as.vector(target[,i])))
+  }))
+
+  colnames(dfq) = colnames(df)
+  rownames(dfq) = rownames(df)
+
+  return(as.matrix(dfq))
+}
+
+copySegs = q.norm.by(mergedSegs, raw.segs)
+copyArms = q.norm.by(mergedArms, raw.arms)
+
+#copySegs = as.data.frame(normalize.quantiles.use.target(mergedSegs, as.vector(raw.segs)))
+grid.arrange(
+  ggplot( melt(raw.segs), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, sWGS bins (no arms)'),
+  ggplot( melt(mergedSegs), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, SNP bins (no arms)'),
+  ggplot( melt(copySegs), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, SNP bins q-normed')
+)
+
+```
+
 Adjust the variance of each genomic region by the mean and sd extracted when I trained on the sWGS data.
@@
 ```{r, warning=F, message=F, fig.width=10, fig.height=8}
+
 #ns = adjust.cols(normalSegs, z.mean, z.sd)
@@
 #ms = adjust.cols(mergedSegs, z.mean, z.sd)
-ms = adjust.cols(mergedSegs)
+#ms = adjust.cols(mergedSegs)
+ms = adjust.cols(copySegs)
 range(ms)
@@
 #ma = adjust.cols(mergedArms, z.arms.mean, z.arms.sd)
-ma = adjust.cols(mergedArms)
+#ma = adjust.cols(mergedArms)
+ma = adjust.cols(copyArms)
 range(ma)
@@
 
-lims=range(arrayDf, allDf,na.rm=T)
+lims=range(arrayDf, dysplasia.df,na.rm=T)
 grid.arrange(
@@
     geom_point(data=melt(arrayDf[,589:634]), col='red') + labs(title='SNP'),
-  ggplot(melt(allDf), aes(Var2, value)) + geom_point() + ylim(lims) +
-      geom_point(data=melt(allDf[,589:634]), col='red') + labs(title='sWGS'),
+  ggplot(melt(dysplasia.df), aes(Var2, value)) + geom_point() + ylim(lims) +
+      geom_point(data=melt(dysplasia.df[,589:634]), col='red') + labs(title='sWGS'),
 nrow=1)
@@
 ## TODO: quantile normalize the SNP arrays by the sWGS values. 
-copy = do.call(cbind.data.frame,lapply(1:ncol(arrayDf), function(i) {
-  as.data.frame(normalize.quantiles.use.target(arrayDf[,i,drop=F], as.vector(allDf[,i])))
-}))
-
-colnames(copy) = colnames(arrayDf)
-rownames(copy) = rownames(arrayDf)
+# copy = do.call(cbind.data.frame,lapply(1:ncol(arrayDf), function(i) {
+#   as.data.frame(normalize.quantiles.use.target(arrayDf[,i,drop=F], as.vector(allDf[,i])))
+# }))
+# 
+# colnames(copy) = colnames(arrayDf)
+# rownames(copy) = rownames(arrayDf)
 
 grid.arrange(
-  ggplot( melt(allDf), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, sWGS'),
+  ggplot( melt(dysplasia.df), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, sWGS'),
   ggplot( melt(arrayDf), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, SNP arrays'),
-  ggplot( melt(copy), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, SNP arrays q-normed'),
+  #ggplot( melt(copy), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, SNP arrays q-normed'),
   ncol=1
 )
-arrayDf = as.matrix(copy)
-```
-
-```{r, eval=F}
-x = apply(mergedSegs, 1, function(x) {
-  ds = c()
-  for (i in 1:(length(x)-1)) {
-    ds = c(ds, abs(x[[i]]) - abs(x[[i+1]]))
-  }
-  abs(diff(range(ds)))
-})
-plot(x, col=pred.qc$Status)
-
-x1 = apply(ms, 1, function(x) {
-  ds = c()
-  for (i in 1:(length(x)-1)) {
-    ds = c(ds, abs(x[[i]]) - abs(x[[i+1]]))
-  }
-  abs(diff(range(ds)))
-})
-plot(x1, col=pred.qc$Status)
-
-y = apply(normalSegs, 1, function(x) {
-  ds = c()
-  for (i in 1:(length(x)-1)) {
-    ds = c(ds, abs(x[[i]]) - abs(x[[i+1]]))
-  }
-  abs(diff(range(ds)))
-})
-plot(y)
-
-range(y)
-range(x)
-
-nprgs = subset(pred.qc, Status == 'NP')
-prgs = subset(pred.qc, Status == 'P')
-
-t.test(x[prgs$Samplename],x[nprgs$Samplename]) 
-t.test(x[prgs$Samplename],y)
-t.test(x[nprgs$Samplename],y)
+#arrayDf = as.matrix(copy)
 ```
@@
 
-# These samples are likely genome duplicated.
-dup = subset(pred.qc, `mean(Ploidy)` > 2.7)
-cor.test(dup$`mean(Ploidy)`, dup$Prob)
-table(dup$Status)
+# Three groups, prog, non prog (max LGD), non prog (max HGD)
+# pred.qc = pred.qc %>% group_by(Status, PatientID) %>% dplyr::mutate( 
+#   exclude=(Status == 'NP' & (length(which(Pathology == 'HGD')) > 0 | `mean(Ploidy)` > 2.7 )) )
+pred.qc$Status = factor(pred.qc$Status)
 
-# Three groups, prog, non prog (max LGD), non prog (max HGD)
 pred.qc = pred.qc %>% group_by(Status, PatientID) %>% dplyr::mutate(
-  exclude=(Status == 'NP' & (length(which(Pathology == 'HGD')) > 0 | `mean(Ploidy)` > 2.7 )) )
-pred.qc$Status = factor(pred.qc$Status)
+  exclude=(Status == 'NP' & (length(which(Pathology == 'HGD')) > 0)),
+  wgd = (Status == 'NP' & `mean(Ploidy)` > 2.7 )
+)
+pred.qc = pred.qc %>% group_by(Status, Samplename) %>% dplyr::mutate(
+  lowsca = (Status == 'P' & (`mean(SCA.Ratio)` < 0.03 | `mean(Ploidy)` < 1.5))
+)
 
@@
 
-table(subset(pred.qc, Risk == 'High' & !exclude)$Status)/table(subset(pred.qc, !exclude)$Status)
-table(subset(pred.qc, Risk == 'Moderate' & !exclude)$Status)/table(subset(pred.qc, !exclude)$Status)
-table(subset(pred.qc, Risk == 'Low' & !exclude)$Status)/table(subset(pred.qc, !exclude)$Status)
-
-# Exclude those odd tetraploidy NPs
-#pred.qc[which(with(pred.qc, (Samplename %in% highPloidyCN$Samplename))), 'exclude'] = T
-pred.qc[which(pred.qc$PatientID %in% highPloidyCN$PatientID & pred.qc$EndoID %in% highPloidyCN$EndoID), 'exclude'] = T
-
 # Exclude HGD NP patients
 roc = pROC::roc(Status ~ Prob, data=subset(pred.qc, !exclude), auc=T, ci=T, of='thresholds')
-roc.plot(roc) + labs(title='Excluded low sca/NP tetra), Prob')
+roc.plot(roc) + labs(title='Excluded HGD NP patients')
 
@@
 rocpp = pROC::roc(Status ~ max.Prob, data=per_pt, auc=T, ci=T, of='thresholds')
-roc.plot(rocpp) + labs(title='Max per patient (excluded low sca/NP tetra), Prob')
+roc.plot(rocpp) + labs(title='Max per patient - excluded HGD NP')
 
 # Exclude the low SCA patients
-pred.qc[which(pred.qc$PatientID %in% lowsca$PatientID & pred.qc$EndoID %in% lowsca$EndoID), 'exclude'] = T
-rocpp = pROC::roc(Status ~ Prob, data=subset(pred.qc, !exclude), auc=T, ci=T, of='thresholds')
-roc.plot(rocpp) + labs(title='Max per patient (excluded low sca/NP tetra), Prob')
+rocpp = pROC::roc(Status ~ Prob, data=subset(pred.qc, !lowsca), auc=T, ci=T, of='thresholds')
+roc.plot(rocpp) + labs(title='Both endos - excluded low SCA P')
 pROC::coords(rocpp, 'best')
 
-as.data.frame(subset(pred.qc, Status == 'P' & Prob < 0.5 & !exclude))
+# Exclude the tetraploid NP
+rocpp = pROC::roc(Status ~ Prob, data=subset(pred.qc, !wgd), auc=T, ci=T, of='thresholds')
+roc.plot(rocpp) + labs(title='Both endos - excluded WGD NP')
+pROC::coords(rocpp, 'best')
+
+# Exclude all of the above
+rocpp = pROC::roc(Status ~ Prob, data=subset(pred.qc, !exclude & !lowsca & !wgd), auc=T, ci=T, of='thresholds')
+roc.plot(rocpp) + labs(title='Both endos - excluded low SCA P & HGD NP & WGD NP')
+pROC::coords(rocpp, 'best')
+
+table(subset(pred.qc, Risk == 'High' & !exclude & !lowsca & !wgd)$Status)/table(subset(pred.qc, !exclude & !lowsca & !wgd)$Status)
+table(subset(pred.qc, Risk == 'Moderate' & !exclude & !lowsca & !wgd)$Status)/table(subset(pred.qc, !exclude & !lowsca & !wgd)$Status)
+table(subset(pred.qc, Risk == 'Low' & !exclude & !lowsca & !wgd)$Status)/table(subset(pred.qc, !exclude & !lowsca & !wgd)$Status)
 
 # Timepoint 2 only
-roct2 = pROC::roc(Status ~ Prob, data=subset(pred.qc, Timepoint != 'T1' & !exclude), auc=T, ci=T, of='thresholds')
+roct2 = pROC::roc(Status ~ Prob, data=subset(pred.qc, Timepoint != 'T1' & !exclude & !lowsca & !wgd), auc=T, ci=T, of='thresholds')
 roc.plot(roct2) + labs(title='T2, Prob')
 
-per_pt_ex = subset(pred.qc, !exclude) %>% group_by(PatientID, Status) %>% dplyr::summarise(
-  max.Prob = max(Prob), max.RR = max(RR), max.Risk = max(Risk)
-)
-rocppex = pROC::roc(Status ~ max.Prob, data=per_pt_ex, auc=T, ci=T, of='thresholds')
-roc.plot(rocppex) + labs(title='Max per patient, Prob')
-
-ggplot(subset(pred.qc, !exclude), aes(`mean(SCA.Ratio)`,Prob, color=Risk, shape=Status)) + 
+ggplot(subset(pred.qc, !exclude & !wgd & !lowsca), aes(`mean(SCA.Ratio)`,Prob, color=Risk, shape=Status)) + 
   facet_grid(~Status) + geom_point()
 
-subset(pred.qc, `mean(SCA.Ratio)` < 0.1 & Status == 'P' & !exclude & Prob < 0.5)
+ggplot(subset(pred.qc, !exclude & !wgd & !lowsca), aes(`mean(Purity)`,Prob, color=Risk, shape=Status)) + 
+  facet_grid(~Status) + geom_point()
 
-ggplot(subset(pred.qc, !exclude | ( `mean(SCA.Ratio)` > 0.1 & Status == 'P')) , aes(`mean(Purity)`,Prob, color=Risk, shape=Status)) + facet_grid(~Status) + geom_point()
-
-ex.pred = subset(pred.qc, !exclude)
+ex.pred = subset(pred.qc, !exclude & !wgd & !lowsca)
 
 # Quick check
-nrow(subset(ex.pred, Prob > 0.5 & Status == 'P'))/table(ex.pred$Status)['P']
-nrow(subset(ex.pred, Prob < 0.5 & Status == 'P'))/table(ex.pred$Status)['P'] #False negatives
-
-nrow(subset(ex.pred, Prob > 0.5 & Status == 'NP'))/table(ex.pred$Status)['NP'] # False pos
-nrow(subset(ex.pred, Prob < 0.5 & Status == 'NP'))/table(ex.pred$Status)['NP'] #
-
- ggplot(melt(arrayDf), aes(Var2, value)) + geom_point() + 
-  geom_point(data=melt(arrayDf['1006_190V',,drop=F]), color='red') + 
-  geom_point(data=melt(arrayDf['1044_39Z',,drop=F]), color='blue')
-
+# nrow(subset(ex.pred, Prob > 0.5 & Status == 'P'))/table(ex.pred$Status)['P']
+# nrow(subset(ex.pred, Prob < 0.5 & Status == 'P'))/table(ex.pred$Status)['P'] #False negatives
+# 
+# nrow(subset(ex.pred, Prob > 0.5 & Status == 'NP'))/table(ex.pred$Status)['NP'] # False pos
+# nrow(subset(ex.pred, Prob < 0.5 & Status == 'NP'))/table(ex.pred$Status)['NP'] #
 ```
@@
 
-```{r regression, echo=F, message=F}
-slabels = as.data.frame(subset(pred.qc, !exclude, select=c('PatientID','Samplename','Status')))
-#slabels = as.data.frame(pred.qc, select=c('PatientID','Samplename','Status'))
-rownames(slabels) = slabels$Samplename
-
-df = arrayDf[which(rownames(arrayDf) %in% slabels$Samplename),]
-
-status = as.integer(slabels$Status)-1
-names(status) = slabels$Samplename
-
-source('lib/cv-pt-glm.R')
-
-nl = 1000;folds = 10; splits = 5 
+```{r regression, echo=F, message=F, eval=T}
 ## ----- All ----- ##
-coefs = list(); plots = list(); performance.at.1se = list(); models = list(); cvs = list()
 cache.dir = '~/Data/Ellie/Analysis/SNP'
-dir.create(cache.dir, recursive = T, showWarnings = F)
 file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
@@
 } else {
-  alpha.values = c(0,0.5,0.7,0.9,1)
-  sets = create.patient.sets(slabels, folds, splits, 0.2)
-
-  for (a in alpha.values) {
-    message(paste('alpha=',a))
-    fit0 <- glmnet(df, status, alpha=a, nlambda=nl, family='binomial', standardize=F)    
-    autoplot(fit0) + theme(legend.position="none")
-    l = fit0$lambda
-    if (a > 0) l = more.l(l)
-
-    cv.patient = crossvalidate.by.patient(x=df, y=status, lambda=l, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', opt=-1, standardize=F)
-
-    lambda.opt = cv.patient$lambda.1se
-
-    coef.opt = as.data.frame(non.zero.coef(fit0, lambda.opt))
-    coefs[[as.character(a)]] = coef.stability(coef.opt, cv.patient$non.zero.cf)
-
-    plots[[as.character(a)]] = arrangeGrob(cv.patient$plot+ggtitle('Classification'), cv.patient$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
-
-    performance.at.1se[[as.character(a)]] = subset(cv.patient$lambdas, `lambda-1se`)
-    models[[as.character(a)]] = fit0
-    cvs[[as.character(a)]] = cv.patient
-  }
-  save(plots, coefs, performance.at.1se, df, models, cvs, labels, file=file)
-  p = do.call(grid.arrange, c(plots[ as.character(alpha.values) ], top='All samples, 10fold, 5 splits'))
-  ggsave(paste(cache.dir, '/', 'all_samples_cv.png',sep=''), plot = p, scale = 2, width = 12, height = 10, units = "in", dpi = 300)
+  stop("Run 'snp-regression.R' first")
 }
@@
 
-```{r compare.coefs, echo=F, eval=T}
+```{r compare.coefs, echo=F, eval=F}
 # Matching
@@
 
-rowSums(snp.features[,-1])/50
+length(which(rowSums(snp.features[,-1])/50 > 0.5))
 
@@
 
-```{r loo}
+```{r loo, eval=F}
 

### 2018-07-24 snp-loo.R

@@ -1,0 +1,95 @@
+
+
+args = commandArgs(trailingOnly=TRUE)
+
+if (length(args) < 2)
+  stop("Missing parameters: <patient ID> <model dir>")
+
+library(ggplot2)
+library(gridExtra)
+library(glmnet)
+library(readxl)
+library(pROC)
+library(ggfortify)
+library(preprocessCore)
+
+suppressPackageStartupMessages(source('lib/data_func.R'))
+suppressPackageStartupMessages(source('lib/common_plots.R'))
+suppressPackageStartupMessages(source('lib/cv-pt-glm.R'))
+
+
+
+pt = args[1]
+#pt = '512'
+
+cache.dir = args[2]
+#cache.dir = '~/Data/Ellie/Analysis/SNP'
+
+dir.create(cache.dir, recursive = T, showWarnings = F)
+file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
+if (!file.exists(file))
+  stop("Missing full model file, run snp-regression.R first")
+load(file, verbose = T)
+
+nl = 1000;folds = 10; splits = 5 
+sets = create.patient.sets(slabels, folds, splits, 0.15)  
+
+slabels$Prediction = NA
+slabels$RR = NA
+
+colnames(sets)[2] = 'Samplename'
+
+file = paste(cache.dir, 'loo.Rdata', sep='/')
+
+select.alpha = 0.9
+secf = coefs[[select.alpha]]
+a = select.alpha
+
+performance.at.1se = c(); plots = list(); fit = c(); nzcoefs = c() 
+preds = matrix(ncol=2, nrow=length(subset(slabels, PatientID == pt)[,2]), dimnames=list(subset(slabels, PatientID == pt)[,2],c('Prob','RR')))
+# Remove each patient (LOO)
+#for (pt in unique(slabels$PatientID)) {
+print(pt)
+samples = subset(slabels, PatientID != pt)[,2]
+
+train.rows = which(rownames(df) %in% samples)
+training = df[train.rows,]
+test = as.matrix(df[-train.rows,])
+if ( nrow(test) == ncol(df) ) test = t(test)
+
+# Predict function giving me difficulty when I have only a single sample, this ensures the dimensions are the same
+sparsed_test_data <- Matrix(data=0, nrow=nrow(test),  ncol=ncol(training),
[…]

### 2018-07-24 snp-regression.R

@@ -1,0 +1,311 @@
+library(ggplot2)
+library(gridExtra)
+library(glmnet)
+library(readxl)
+library(pROC)
+library(ggfortify)
+library(preprocessCore)
+
+suppressPackageStartupMessages(source('lib/data_func.R'))
+suppressPackageStartupMessages(source('lib/common_plots.R'))
+source('lib/cv-pt-glm.R')
+
+adjust.cols<-function(mtx, means=NULL, sds=NULL, na.replace=0) {
+  if (!is.null(means)) {
+    if (ncol(mtx) != length(means))
+      stop("Vector of means needs to match the columns in the matrix")
+    for (i in 1:ncol(mtx)) 
+      mtx[,i] = unit.var(mtx[,i], means[i], sds[i])
+  } else {
+    for (i in 1:ncol(mtx)) 
+      mtx[,i] = unit.var(mtx[,i])
+  }
+  mtx[is.na(mtx)] = na.replace
+  
+  return(mtx)
+}
+
+
+
+chr.info = get.chr.lengths(file='hg19_info.txt')
+
+load('~/Data/Ellie/Analysis/5e6_arms_all/model_data.Rdata', verbose = T)
+rm(labels, dysplasia.df,mn.cx,sd.cx,z.mean,z.sd,z.arms.mean,z.arms.sd)
+
+load('~/Data/Reid_SNP/PerPatient/allpts_ascat.Rdata', verbose=T)
+
+patient.info = as.data.frame(read_xlsx('~/Data/Ellie/Analysis/SNP/metadata_T1T2.xlsx'))
+patient.info$UniqueSampleID = paste(patient.info$PatientID, patient.info$`Timepoint Code`, sep='_')
+patient.info$Path.Status = patient.info$Status
+patient.info[patient.info$PatientID %in% subset(patient.info, Pathology %in% c('IMC','HGD'), select='PatientID')[,1], 'Path.Status'] = 'P'
+
+sample.list = read_xlsx('~/Data/Ellie/Analysis/SNP/20180604_Reid_1M_SampleList.xlsx', sheet = 2)
+nrow(sample.list)
+colnames(sample.list)[3] = 'Total.SCA'
+
+sample.list$SCA.Ratio = sample.list$Total.SCA/max(chr.info$genome.length)
+
+sample.info  = do.call(rbind.data.frame, sapply(qcdata$Samplename, strsplit, '_'))
+colnames(sample.info) = c('PatientID','SampleID','EndoID','Level')
[…]

### 2018-07-25 SNP_Prediction_Analysis.Rmd
generate new plots

@@
 
-ggplot(qcdata, aes(SCA.Ratio)) + geom_histogram() + 
-  facet_wrap(~Status, scales='free_y', ncol=1) + labs(x='SCA Ratio', title='SCA Ratio, NP vs P')
+ggplot(subset(qcdata, SampleType == 'BE'), aes(round(Ploidy,2), Purity, color=Status)) + geom_point() + facet_grid(~Status) + 
+  labs(title='BE only, purity vs ploidy')
 
-ggplot(subset(qcdata, SampleType == 'BE'), aes(Purity, SCA.Ratio, color=Status)) + geom_point() + 
-  geom_vline(xintercept = 0.95, color='red') + geom_hline(yintercept = 0.05, color='red') + 
+ggplot(qcdata, aes(SCA.Ratio, color=Status, fill=Status)) + geom_density(alpha=0.5) + labs(x='SCA Ratio', title='SCA Ratio, NP vs P')
+
+ggplot(subset(qcdata, SampleType == 'BE'), aes(Purity, SCA.Ratio, color=Status)) + facet_grid(~Status) +
+  geom_point() + geom_vline(xintercept = 0.5, color='darkgreen') + geom_hline(yintercept = 0.05, color='red') + 
   labs(title='Barretts only')
 
-ggplot(subset(qcdata, SampleType == 'BE'), aes(Ploidy, SCA.Ratio, color=Status)) + geom_point() + 
-  geom_hline(yintercept = 0.05, color='red') + labs(title='Barretts only')
+ggplot(subset(qcdata, SampleType == 'BE'), aes(Ploidy, SCA.Ratio, color=Status)) + facet_grid(~Status) + geom_point() + 
+  geom_hline(yintercept = 0.03, color='red') + labs(title='Barretts only')
 
@@
 
+subset(pred.qc, `mean(Ploidy)` < 1.99)
+
+subset(pred.qc, lowsca)
+
 ```
@@
 ```{r echo=F}
-pred.qc.ex = subset(pred.qc, !exclude)
+pred.qc.ex = subset(pred.qc, !exclude & !wgd & !lowsca)
 cuts = seq(0,1,0.1)
@@
 roc.plot(rocpp) + labs(title='Max per patient - excluded HGD NP')
+
+roc = pROC::roc(Status ~ Prob, data=subset(pred.qc, !exclude & !wgd), auc=T, ci=T, of='thresholds')
+roc.plot(roc) + labs(title='Excluded HGD & WGD NP patients')
 
@@
 
-```{r regression, echo=F, message=F, eval=T}
+```{r regression, echo=F, message=F, eval=F}
 ## ----- All ----- ##
@@
  load(file, verbose=T)
+ model.performance(performance.at.1se, coefs)$plot + labs(title='SNP CV models')
 } else {
-  stop("Run 'snp-regression.R' first")
+  warning("Run 'snp-regression.R' first")
 }
@@
 
-do.call(grid.arrange, c(plots))
-
-model.performance(performance.at.1se, coefs)$plot + labs(title='SNP CV models')
-
+#do.call(grid.arrange, c(plots))
 ```
@@
 } else {
-  select.alpha = 1
-  secf = coefs[[select.alpha]]
-  a = select.alpha
-
-  performance.at.1se = c(); plots = list(); fits = list(); nzcoefs = list()
-  # Remove each patient (LOO)
-  for (pt in unique(slabels$PatientID)) {
-    print(pt)
-    samples = subset(slabels, PatientID != pt)$Samplename
-
-    train.rows = which(rownames(df) %in% samples)
-    training = df[train.rows,]
-    test = as.matrix(df[-train.rows,])
-    if ( nrow(test) == ncol(df) ) test = t(test)
-
-    # Predict function giving me difficulty when I have only a single sample, this ensures the dimensions are the same
-    sparsed_test_data <- Matrix(data=0, nrow=nrow(test),  ncol=ncol(training),
-                                dimnames=list(rownames(test),colnames(training)), sparse=T)
-    for(i in colnames(arrayDf)) sparsed_test_data[,i] = test[,i]
-
-    # Fit generated on all samples, including HGD
-    fitLOO <- glmnet(training, status[train.rows], alpha=a, family='binomial', nlambda=nl, standardize=F) # all patients
-    l = fitLOO$lambda
-
-    cv = crossvalidate.by.patient(x=training, y=status[train.rows], lambda=l, a=a, nfolds=folds, splits=splits,
-                                  pts=subset(sets, Samplename %in% samples), fit=fitLOO, standardize=F)
-
-    plots[[pt]] = arrangeGrob(cv$plot+ggtitle('Classification'), cv$deviance.plot+ggtitle('Binomial Deviance'), top=pt, ncol=2)
-    fits[[pt]] = cv
-
-    if ( length(cv$lambda.1se) > 0 ) {
-      performance.at.1se = c(performance.at.1se, subset(cv$lambdas, lambda == cv$lambda.1se)$mean)
-
-      coef.1se = coef(fitLOO, cv$lambda.1se)[rownames(secf),]
-
-      nzcoefs[[pt]] = as.data.frame(non.zero.coef(fitLOO, cv$lambda.1se))
-
-      #coefs[[pt]] = coef(fitLOO, cv$lambda.1se)[rownames(secf),]
-      #coef.stability(coef.1se, cv$non.zero.cf)
-
-      logit <- function(p){log(p/(1-p))}
-      inverse.logit <- function(or){1/(1 + exp(-or))}
-
-      pm = predict(fitLOO, newx=sparsed_test_data, s=cv$lambda.1se, type='response')
-      or = predict(fitLOO, newx=sparsed_test_data, s=cv$lambda.1se, type='link')
-
-      slabels[which(slabels$Samplename %in% rownames(pm)),'Prediction'] = pm[,1]
-      slabels[which(slabels$Samplename %in% rownames(pm)),'RR'] = or[,1]
-
-    } else {
-      warning(paste(pt, "did not have a 1se"))
-    }
-  }
-  save(plots, performance.at.1se, nzcoefs, fits, slabels, file=file)
+  stop("Run snp-regression.R first")
 }
 
+slabels
 snp.roc = pROC::roc(Status~Prediction,slabels, auc=T, ci=T, of='thresholds')

### 2018-08-14 AllSamplesModel.Rmd
adding residuals

@@
   scale_color_manual(values=c('firebrick3','darkblue')) +
-  labs(title='Top features by hazard ratio', x='Coefficient value', y='Hazard') + plot.theme 
+  labs(title='Features by CV(relative risk)', x='Coefficient value', y='CV(RR)') + plot.theme 
 p

### 2018-08-14 lib/data_func.R

@@ -44,34 +44,17 @@ logTransform<-function(x,inf=F) {
   return(x)
 }
   
-
-binSWGS<-function(raw.data, fit.data, blacklist, logTransform=F, min.probes=67, gamma2=250, plot.dir='.', metrics.file='binSWGS.out') {
-  require(copynumber)
-  require(apcluster)
-  
-  message(paste('Logging to',metrics.file))
-  
-  if (file.exists(metrics.file)) {
-    mfile = file(metrics.file, 'a')
-  } else {
-    mfile = file(metrics.file, 'w')
-  }
-  
+prep.pcf.data<-function(raw.data,fit.data,blacklist, logTransform=F) {
   if (is.null(blacklist)) 
     stop('QDNASeq blacklisted regions missing')
-    
+  
   countCols = grep('loc|feat|chr|start|end', colnames(fit.data), invert=T)
   
-  write(paste('fit data ',paste(dim(fit.data), collapse=':')), mfile, append=T)
-  write(paste('raw data ',paste(dim(raw.data), collapse=':')), mfile, append=T)
-
   rows = which(fit.data[,1] %in% raw.data[,1])
-  #raw.data = raw.data[rows,]
   fit.data = fit.data[rows,]
-
+  
   window.depths = as.vector(as.matrix(raw.data[,countCols]))/as.vector(as.matrix(fit.data[,countCols]))
   
-  #window.depths = raw.data[,countCols]/fit.data[,countCols]
   # QDNAseq does this in 'correctBins' but it's too late to use this now, so will adjust ours the same way (we weren't 14-May-2018)
   #corrected <- counts / fit
   #corrected[fit <= 0] <- 0
@@ -81,16 +64,16 @@ binSWGS<-function(raw.data, fit.data, blacklist, logTransform=F, min.probes=67, 
   window.depths = matrix(window.depths, ncol=length(countCols))  
   
   message(paste(nrow(blacklist), "genomic regions in the exclusion list."))
-  write(paste(nrow(blacklist), "genomic regions in the exclusion list."),mfile,append=T)
-
+  #write(paste(nrow(blacklist), "genomic regions in the exclusion list."),mfile,append=T)
+  
   fit.data$in.blacklist = F
   for(r in 1:nrow(blacklist)) {
     fit.data$in.blacklist[ fit.data$chrom == blacklist$chromosome[r] & 
[…]

### 2018-08-14 processing_pipeline/SegmentMultisampleCopynumber_fitted_byPatient.R

@@ -13,18 +13,18 @@ suppressPackageStartupMessages( library(ggdendro) )
 suppressPackageStartupMessages( library(copynumber) )
 suppressPackageStartupMessages( library(dplyr) )
 
-source("~/workspace/shallowwgs_pipeline/lib/fastPCF.R")
-source('~/workspace/shallowwgs_pipeline/lib/load_patient_metadata.R')
-source('~/workspace/shallowwgs_pipeline/lib/data_func.R')
+suppressPackageStartupMessages( source("~/workspace/shallowwgs_pipeline/lib/fastPCF.R") )
+suppressPackageStartupMessages( source('~/workspace/shallowwgs_pipeline/lib/load_patient_metadata.R') )
+suppressPackageStartupMessages( source('~/workspace/shallowwgs_pipeline/lib/data_func.R') )
 
 
 data = args[1]
 patient.name = args[2]
 outdir = args[3]
 
-#data = '~/Data/Ellie/QDNAseq'
-#patient.name = 'PR1/HIN/042' #'AD0098' # 'PR1/HIN/042' #'PR1/HIN/044'
-#outdir = '~/Data/Ellie/Analysis'
+# data = '~/Data/Ellie/QDNAseq'
+# patient.name = 'PR1/HIN/042' #'AD0098' # 'PR1/HIN/042' #'PR1/HIN/044'
+# outdir = '~/Data/Ellie/Analysis'
 data.dirs = list.dirs(data, full.names=T, recursive=F)
 data.files = list.files(data, full.names=T, recursive=T)
 
@@ -37,7 +37,6 @@ if (!dir.exists(plot.dir))
 
 ## Patient info file
 filename = 'All_patient_info.xlsx'
-
 patient.file = grep(filename, list.files(data, full.names=T), value=T)
 if (length(patient.file) != 1)
   stop(paste("Missing patient info file in", data))
@@ -63,13 +62,8 @@ if (length(data.file) <= 0)
   stop("Missing necessary Rdata file from Multisample script")
 load(file=data.file, verbose=T)
 
-##patient.info$Samplename
-#colnames(fit.data)[1:14]
-
-patient.plot.dir
 write.table(fit.data[,c('location','chrom','start','end',patient.info$Samplename)], sep='\t', quote=F, file=paste(patient.plot.dir, 'raw_fitted.txt', sep='/'))
[…]

### 2018-08-14 processing_pipeline/process_swgssamples.R

@@ -2,78 +2,156 @@
 args = commandArgs(trailingOnly=TRUE)
 
 if (length(args) < 3)
-  stop("Missing required params: <data dir> <out dir> <blacklist file> <overwrite: default F>")
+  stop("Missing required params: <data dir> <out dir> <blacklist file> <sample info file> <overwrite: default F>")
 
 suppressPackageStartupMessages(source('~/workspace/shallowWGSpipeline//lib/fastPCF.R'))
 suppressPackageStartupMessages(source('~/workspace/shallowWGSpipeline//lib/data_func.R'))
 suppressPackageStartupMessages(library(ggplot2))
+suppressPackageStartupMessages(library(gridExtra))
+suppressPackageStartupMessages(library(readxl))
 
 chrlen = get.chr.lengths(file='/tmp/hg19_info.txt')
 
 dir = args[1]
 outdir = args[2]
 exclude.file = args[3]
+info.file = args[4]
+
+gamma2=250
+
 overwrite = F
-if (length(args) == 4) overwrite =  as.logical(args[4])
+if (length(args) == 5) overwrite =  as.logical(args[5])
 
-dir = '~/Data/Ellie/Cassandra/20180409_KosmidouC_RF_sWGS/qdnaseq_15kb/'
-#dir = '~/Data/Ellie/Cassandra/20180604_KosmidouC_RF_sWGS/qdnaseq'
-#outdir = '~/Data/Ellie/Analysis/VAL_Cohort/sWGS'
-#exclude.file = '~/Data/Ellie/QDNAseq/qDNAseq_blacklistedRegions.txt'
-
-dir.create(outdir, F)
-#f = 'LP2000110-DNA_A01.binSize15.fittedReadCounts.txt'
-#r = 'LP2000110-DNA_A01.binSize15.readCounts.txt'
+# dir = '~/Data/Ellie/QDNAseq/SLX-15638'
+# outdir = '~/tmp/VAL'
+# exclude.file = '~/Data/Ellie/QDNAseq/qDNAseq_blacklistedRegions.txt'
+# info.file = '~/Data/Ellie/Cassandra/ValidationCohort.xlsx'
 
 if ( !file.exists(exclude.file) )
-  stop(paste("Missing necessary exclusion file from 'qDNAseq_blacklistedRegions.txt' in",dir))
+  stop(paste("Missing necessary exclusion file 'qDNAseq_blacklistedRegions.txt'"))
+if ( !file.exists(info.file) )
+  stop(paste("Missing necessary sample info file"))
+
[…]

### 2018-09-05 Paper.Rmd
add Seattle SNP analysis scripts

@@
 
-all.patient.info = read.patient.info(patient.file, demo.file, set='all')$info
+all.patient.info = read.patient.info(patient.file, demo.file, set='All')$info
 head(all.patient.info)
 
-patient.info = subset(all.patient.info, Set == 'Training')
+patient.info = all.patient.info
 
@@
 
-pander(sum.patient.data[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Training patient set')
-
-pander(sum.validation.info[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Validation patient set')
+#pander(sum.patient.data[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Training patient set')
+
+#pander(sum.validation.info[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Validation patient set')
 
@@
 ```{r echo=F}
-cohortSummary = as.data.frame(all.info %>% group_by(Status) %>% summarise(
+cohortSummary = as.data.frame(all.info %>% dplyr::group_by(Status) %>% summarise(
   'Total Patients'=length(unique(Patient)),
@@
 ```{r labelsPNP, echo=F, message=F, include=T}
-cache.dir = paste(data, 'Analysis/5e6_arms_disc', sep='/')
+cache.dir = paste(data, 'Analysis/5e6_arms_all', sep='/')
 
@@
 arm.loc = arrange(rbind(
-  (chr.lengths %>% group_by(chrom) %>% mutate('start'=1, 'end'=chr.cent-cent.gap))[c('chrom','start','end')],
-  (chr.lengths %>% group_by(chrom) %>% mutate('start'=chr.cent+cent.gap, 'end'=chr.length))[c('chrom','start','end')]), chrom)
+  (chr.lengths %>% dplyr::group_by(chrom) %>% dplyr::mutate('start'=1, 'end'=chr.cent-cent.gap))[c('chrom','start','end')],
+  (chr.lengths %>% dplyr::group_by(chrom) %>% dplyr::mutate('start'=chr.cent+cent.gap, 'end'=chr.length))[c('chrom','start','end')]), chrom)
 
@@
 # Average cn value
-globalCNdf = globalCNdf %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
+globalCNdf = globalCNdf %>% dplyr::group_by(Status, start, end) %>% dplyr::mutate( 'mean.value' = value/length(unique(variable)) )
 globalCNdf$mean.value = globalCNdf$value # but not going to use it
@@
 noHGD.pred.confidence = as.data.frame.matrix(table(pg.sampNOHGD$quants, pg.sampNOHGD$Status))
-noHGD.pred.confidence = cbind(noHGD.pred.confidence, pg.sampNOHGD %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
+noHGD.pred.confidence = cbind(noHGD.pred.confidence, pg.sampNOHGD %>% dplyr::group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
 noHGD.pred.confidence$quant = rownames(noHGD.pred.confidence)
@@
 
-noHGD.pred.confidence = noHGD.pred.confidence %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )
+noHGD.pred.confidence = noHGD.pred.confidence %>% as.data.frame.matrix %>% dplyr::group_by(quant) %>% dplyr::mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )
 
@@
 
-  hgd.pred = hgd.samples %>% group_by(Patient, Hospital.Research.ID, Status, Endoscopy.Year) %>% 
+  hgd.pred = hgd.samples %>% dplyr::group_by(Patient, Hospital.Research.ID, Status, Endoscopy.Year) %>% 
     summarise( max=max(Pathology), pred=Prediction[which.max(Pathology)] )
@@
 ```{r, echo=F, warning=F, message=F,fig.height=4, fig.width=4}
-pendo = do.call(rbind, pg.samp) %>% group_by(Patient, Status, Hospital.Research.ID, PID ) %>% summarise(
+pendo = do.call(rbind, pg.samp) %>% dplyr::group_by(Patient, Status, Hospital.Research.ID, PID ) %>% dplyr::summarise(
   'max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(PID)
@@
 
-#ptendo = do.call(rbind, pg.samp) %>% group_by(Patient, Hospital.Research.ID, Status) %>% summarise('max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(Patient))
+#ptendo = do.call(rbind, pg.samp) %>% dplyr::group_by(Patient, Hospital.Research.ID, Status) %>% summarise('max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(Patient))
 
@@
 transform.by.time<-function(pt, info) {
-  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction', 'OR', 'Sample.Mean','Sample.SD','Sample.Var')] %>% group_by(PID, Samplename) %>% mutate(
+  #print(unique(pt$Hospital.Research.ID))
+  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction', 'OR', 'Sample.Mean','Sample.SD','Sample.Var')] %>% dplyr::group_by(PID, Samplename) %>% dplyr::mutate(
   'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12)
 
-  pt = arrange(merge(pt, subset(info, Hospital.Research.ID == pt$Hospital.Research.ID[1], select=c('Samplename','Block')), by='Samplename'), Pathology, PID, Endoscopy.Year)
+  pt = dplyr::arrange(merge(pt, subset(info, Hospital.Research.ID == pt$Hospital.Research.ID[1], select=c('Samplename','Block')), by='Samplename'), Pathology, PID, Endoscopy.Year)
 
@@
       tmp = pt[which(pt$Endoscopy.Year == yr),]
-      tmp = tmp %>% group_by(PID) %>% summarise(max(Pathology))
+      tmp = tmp %>% dplyr::group_by(PID) %>% summarise(max(Pathology))
 
@@
 
+### TODO: I need to redo the calibration, my confidence intervals seem wrong.
+
 ```{r spacial, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
@@
 
-qt = cbind(qt, back %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))#, 'se'=(sd(Prediction)/sqrt(length(Prediction)))*1.96 ))
-#ggplot(qt,aes(quants, mn, ymin=mn-se, ymax=mn+se)) + geom_point() + geom_errorbar() + labs(x='Deciles', y='Mean probability', title='CI around probabilities by decile')
-
-pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 
+qt = cbind(qt, back %>% dplyr::group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))#, 'se'=(sd(Prediction)/sqrt(length(Prediction)))*1.96 ))
+
+pred.confidence = qt %>% as.data.frame.matrix %>% dplyr::group_by(quant) %>% dplyr::mutate( 
   'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )
@@
 pred.confidence[c('r1','r2')] = NA
-for (i in 1:(length(cuts)-1)) 
-    pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)
+for (i in 1:(length(cuts)-1)) pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)
 
@@
 
-annotation.table = arrange(pred.confidence %>% group_by(Risk) %>% summarise('n(NP)'=sum(NP), 'n(P)'=sum(P), 'p.value'=format.pval(mean(p.value), 3), 'min'=min(r1), 'max'=max(r2)), min)
+annotation.table = arrange(pred.confidence %>% dplyr::group_by(Risk) %>% dplyr::summarise(
+  'n(NP)'=sum(NP), 'n(P)'=sum(P), 'p.value'=format.pval(mean(p.value), 3), 
+  'min'=min(r1), 'max'=max(r2)), min
+)
 
@@
     geom_vline(xintercept=annotation.table$max, linetype='longdash', color='grey39') +
-    geom_text(data=annotation.table %>% group_by(Risk, min,max) %>% summarise( 'samp.perc'= signif((sum(`n(NP)`, `n(P)`)/nrow(patient.info))*100, 3) ), aes(x=min+(max-min)/2, y=5, label=paste(samp.perc, '%',sep=''))) +
+    geom_text(data=annotation.table %>% dplyr::group_by(Risk, min,max) %>% dplyr::summarise( 'samp.perc'= signif((sum(`n(NP)`, `n(P)`)/nrow(patient.info))*100, 3) ), aes(x=min+(max-min)/2, y=5, label=paste(samp.perc, '%',sep=''))) +
     plot.theme + labs(title='Sample predictions', y='n Samples', x='Probability') + theme(legend.position = 'none')
@@
 back = merge(back, pred.confidence[c('quant','Risk')], by.x='quants',by.y='quant')
-back = back %>% group_by(Risk) %>% mutate('se(Risk)'=(sd(Prediction)/sqrt(length(Risk)))*1.96 )
+back = back %>% dplyr::group_by(Risk) %>% dplyr::mutate('se(Risk)'=(sd(Prediction)/sqrt(length(Risk)))*1.96 )
 
@@
 ## count arms differently than segs??
-df2 = df %>% group_by(Status, Patient, Endoscopy.Year) %>% summarise( 
+df2 = df %>% dplyr::group_by(Status, Patient, Endoscopy.Year) %>% dplyr::summarise( 
   'max.Pred'=max(Prediction),
@@
 
-df2 = df2 %>% group_by(Status, Patient) %>% mutate(
+df2 = df2 %>% dplyr::group_by(Status, Patient) %>% dplyr::mutate(
   'early'= Endoscopy.Year == min(Endoscopy.Year),
@@
 el = subset(df2, early | late)
-el2 = el %>% group_by(Status, Patient) %>% summarise( diff(arms), diff(CNA) )
+el2 = el %>% dplyr::group_by(Status, Patient) %>% dplyr::summarise( diff(arms), diff(CNA) )
 wtA = round(wilcox.test(subset(el2, Status == 'Progressor')$`diff(arms)`, subset(el2, Status == 'Non-Progressor')$`diff(arms)`)$p.value, 3)
@@
 
-ggplot(el2, aes(Status,`diff(arms)`)) + geom_boxplot() + geom_jitter(width=0.2)
-ggplot(el2, aes(Status,`diff(CNA)`)) + geom_boxplot() + geom_jitter(width=0.2)
+#ggplot(el2, aes(Status,`diff(arms)`)) + geom_boxplot() + geom_jitter(width=0.2)
+#ggplot(el2, aes(Status,`diff(CNA)`)) + geom_boxplot() + geom_jitter(width=0.2)
 
@@
 
-pathTotal = back %>% group_by(Status, Pathology) %>% summarise( 'nPath'=length(Pathology))
+pathTotal = back %>% dplyr::group_by(Status, Pathology) %>% dplyr::summarise( 'nPath'=length(Pathology))
 pathTotal$Status = ifelse(pathTotal$Status == 'P', 'Progressor','Non-Progressor')
@@
 p53Path = cbind.data.frame(table(subset(back, Status == 'P', select=c('Pathology','p53.Status'))), 'Status'='p53 IHC')
-p53Path = merge( subset(back, Status == 'P' & !is.na(p53.Status)) %>% group_by(Pathology) %>% summarise( 'nPath'=length(Pathology)), p53Path)
+p53Path = merge( subset(back, Status == 'P' & !is.na(p53.Status)) %>% dplyr::group_by(Pathology) %>% dplyr::summarise( 'nPath'=length(Pathology)), p53Path)
 #p53Path$Risk = with(p53Path, ifelse(p53.Status == 1, 'High','Low'))
@@
 
-pathTotal = riskPath %>% group_by(Status, Pathology) %>% summarise('nPath'=unique(nPath))
+pathTotal = riskPath %>% dplyr::group_by(Status, Pathology) %>% dplyr::summarise('nPath'=unique(nPath))
 
@@
 ```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=6}
-all.preds = back %>% group_by(Status, Patient) %>% summarise( 'Low'=sum(length(which(Risk == "Low"))),
+all.preds = back %>% dplyr::group_by(Status, Patient) %>% dplyr::summarise( 'Low'=sum(length(which(Risk == "Low"))),
                                           'Moderate'=sum(length(which(Risk == "Moderate"))),
@@
   geom_bar(stat='identity', position = position_stack()) +
-  scale_fill_manual(values=riskCols) + 
-  geom_text(aes(y=1.01, label=n.samples)) +
+  scale_fill_manual(values=riskCols) + geom_text(aes(y=1.01, label=n.samples)) +
   facet_wrap(~Status, scales='free_x') + labs(y='Sample ratios', title='Per patient ratio of predictions') + 
@@
 ```{r enods, warning=F, message=F, echo=F, fig.height=8, fig.width=6}
-endos = back %>% group_by(Status, months.before.final, brks, PID) %>% summarise(
+endos = back %>% dplyr::group_by(Status, months.before.final, brks, PID) %>% dplyr::summarise(
               'pred.endo'=ifelse(length(which(Risk == 'High')) > 0, 1, 0), 
@@
 
-endos.m = subset(endos, Status == 'P') %>% group_by(brks) %>% summarise( 
+endos.m = subset(endos, Status == 'P') %>% dplyr::group_by(brks) %>% dplyr::summarise( 
   'pred.r'=sum(pred.endo)/sum(n.endos),
@@
      labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))
-tmp = tmp %>% group_by(brks) %>% summarise( 'mean(P)'=mean(Prediction),'sem(P)'=sd(Prediction)/sqrt(length(brks)) )
+tmp = tmp %>% dplyr::group_by(brks) %>% dplyr::summarise( 'mean(P)'=mean(Prediction),'sem(P)'=sd(Prediction)/sqrt(length(brks)) )
 
@@
 #endos$months.before.final = endos$months.before.final*-1
-#me = endos %>% group_by(Patient) %>% summarise('max.endo' = max(i), 'months' = min(months.before.final))
+#me = endos %>% dplyr::group_by(Patient) %>% dplyr::summarise('max.endo' = max(i), 'months' = min(months.before.final))
 ```
@@
 
-  summary.t = summary.t %>% group_by(Pathology) %>% mutate( 'High.r'=round(High/sum(High,Low,Moderate),3), 
+  summary.t = summary.t %>% dplyr::group_by(Pathology) %>% dplyr::mutate( 'High.r'=round(High/sum(High,Low,Moderate),3), 
                                                 'Low.r'=round(Low/sum(High,Low,Moderate),3),
@@
 ```{r, echo=F, message=F, warning=F}
-back = back %>% group_by(Patient) %>% mutate( 'pred.ratio'=round(sum(Prediction)/length(Prediction),2))
+back = back %>% dplyr::group_by(Patient) %>% dplyr::mutate( 'pred.ratio'=round(sum(Prediction)/length(Prediction),2))
 
@@
 # Average cn value
-valCNdf = valCNdf %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
+valCNdf = valCNdf %>% dplyr::group_by(Status, start, end) %>% dplyr::mutate( 'mean.value' = value/length(unique(variable)) )
 valCNdf$mean.value = valCNdf$value # but not going to use it
@@
 
-m = melt(back.v %>% group_by(Patient, Status) %>% summarise( 'Low'=length(which(Risk == 'Low'))/length(Patient),
+m = melt(back.v %>% dplyr::group_by(Patient, Status) %>% dplyr::summarise( 'Low'=length(which(Risk == 'Low'))/length(Patient),
                                             'Moderate'=length(which(Risk == 'Moderate'))/length(Patient),
@@
 
-pathTotal = back.v %>% group_by(Status, Pathology) %>% summarise( 'nPath'=length(Pathology))
+pathTotal = back.v %>% dplyr::group_by(Status, Pathology) %>% dplyr::summarise( 'nPath'=length(Pathology))
 pathTotal$Status = ifelse(pathTotal$Status == 'P', 'Progressor','Non-Progressor')
@@
 
-hgdimc = subset(riskPath, Pathology %in% c('HGD','IMC')) %>% group_by(Risk) %>% summarize (
+hgdimc = subset(riskPath, Pathology %in% c('HGD','IMC')) %>% dplyr::group_by(Risk) %>% summarize (
   Freq=sum(Freq),  nPath=sum(nPath), ratio=sum(Freq)/sum(nPath))
@@
 
-qt = cbind(qt, back.v %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
-
-pred.confidence.v = qt %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 'perc'=P/sum(NP,P) )
+qt = cbind(qt, back.v %>% dplyr::group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
+
+pred.confidence.v = qt %>% as.data.frame.matrix %>% dplyr::group_by(quant) %>% dplyr::mutate( 'perc'=P/sum(NP,P) )
 
@@
 ```{r echo=F, eval=T}
-dirA = '~/Data/Ellie/Analysis/downsampled5G_BEAdjacent'
-dirB = '~/Data/Ellie/Analysis/downsampled5G_NDBE'
-
 file = paste(cache.dir, 'model_data.Rdata', sep='/')
@@
 
+dirA = '~/Data/Ellie/Analysis/downsampled'
+ds_info = read_excel('~/Data/Ellie/Analysis/downsampled_ids.xlsx')
 predA = list.files(dirA, 'predictions.Rdata', full.names = T)
-predB = list.files(dirB, 'predictions.Rdata', full.names = T)
-
-files = c()
-if (length(predA) < 1)
-  files = c(list.files(dirA, pattern='raw', full.names=T))
-if (length(predB) < 1)
-  files = c(files, list.files(dirB, pattern='raw', full.names=T))
-
-if (length(files) > 0) {
-  sampleNames = sub('_raw.txt', '', basename(files))
-  dspred = data.frame(matrix(ncol=2, nrow=length(sampleNames), dimnames=list(sampleNames,c('Prob','RR'))))
-
-  for (n in 1:length(files)) {
-    print(n)
-    print(sampleNames[n])
-
-    rs = read.table(files[n], sep='\t',header=T)
-    if (ncol(rs) < 4)
-      stop("File needs to contain 4 columns: chr, start, end, sample_value")
-
-    segs <- tile.segmented.data(rs, size=5e6, chr.info=chrlen)
-    segM = as.matrix(segs[,-c(1:3)])
-    rownames(segM) = paste(segs$chr, ':', segs$start, '-', segs$end, sep='')
-    segM = t(segM)
-    for (i in 1:ncol(segM)) 
-      segM[,i] = unit.var(segM[,i], z.mean[i], z.sd[i])
-
-    arms <- tile.segmented.data(rs, size='arms', chr.info=chrlen)
-    armsM = as.matrix(arms[,-c(1:3)])
-    rownames(armsM) = paste(arms$chr, ':', arms$start, '-', arms$end, sep='')
-    armsM = t(armsM)
-    for (i in 1:ncol(armsM)) 
-      armsM[,i] = unit.var(armsM[,i], z.arms.mean[i], z.arms.sd[i])
-
-    nrow(armsM) == nrow(segM)
-
-    cx.score = score.cx(segM,1)
-
-    mergedDf = subtract.arms(segM, armsM)
-    mergedDf = cbind(mergedDf, 'cx' = unit.var(cx.score, mn.cx, sd.cx))
-    dim(mergedDf)
-
-    # Predict function giving me difficulty when I have only a single sample, this ensures the dimensions are the same
-    sparsed_test_data <- Matrix(data=0, nrow=nrow(mergedDf),  ncol=ncol(mergedDf),
-                                dimnames=list(rownames(mergedDf),colnames(mergedDf)), sparse=T)
-    for(col in colnames(mergedDf)) sparsed_test_data[,col] = mergedDf[,col]
-
-    preds = predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='response')
-    RR = predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='link')
-
-    dspred[n,] = c(preds[,1],RR[,1])
-    print(dspred[1:n,])
-  }
-} else {
-  load(predB, verbose=T)
-  dspredB = dspred
-
-  load(predA, verbose=T)
-
-  dspred = rbind(dspred, dspredB)
-  dspred$Type = NA
-
-  dspred[sub('_raw.*','',basename(list.files(dirA, pattern='raw', full.names=T))),'Type'] = 'AdjacentBE'
-  dspred[sub('_raw.*','',basename(list.files(dirB, pattern='raw', full.names=T))),'Type'] = 'BEST2 NDBO'
-
-}
-
-ggplot(dspred, aes(Type, RR, fill=Type)) + geom_boxplot() + geom_jitter(width=0.2) +
+load(predA)
+
+dspred = merge(dspred, ds_info, by.x='row.names', by.y='Illumina ID')
+
+ggplot(dspred, aes(Type, RR, group=Type, fill=Type)) + geom_boxplot(outlier.colour = NA) + geom_jitter(width=0.2, aes(color=Status)) +
   labs(x='', y='Relative Risk', title='Predictions on downsampled patients') + 
-  theme(legend.position = 'none') + plot.theme
-
-adj = subset(dspred, Type == 'AdjacentBE')
+  plot.theme
+
+adj = subset(dspred, grepl('Adjacent',Type))
 prg = back[,c('Prediction', 'OR', 'Status')]
@@
 
-maxP = allB %>% group_by(Patient, Status, brks) %>% summarise( 'maxP'=max(Prediction), 'Risk'=Risk[Prediction == max(Prediction)], 'Pathology'=max(Pathology), 'p53'=max(p53.Status,na.rm=T) )
+maxP = allB %>% dplyr::group_by(Patient, Status, brks) %>% dplyr::summarise( 'maxP'=max(Prediction), 'Risk'=Risk[Prediction == max(Prediction)], 'Pathology'=max(Pathology), 'p53'=max(p53.Status,na.rm=T) )
 
@@
 second = subset(rules,  t2 != 'Endpoint') %>% 
-  group_by(Patient, Status) %>% 
-  summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )
+  dplyr::group_by(Patient, Status) %>% 
+  dplyr::summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )
 
@@
 second = subset(rules2,  t2 != 'Endpoint') %>% 
-  group_by(Patient, Status) %>% 
-  summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )
+  dplyr::group_by(Patient, Status) %>% 
+  dplyr::summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )
 
@@
 p = subset(back, Status == 'P' & months.before.final > 6 & OR > 10)$Patient
-subset(sum.patient.data, Patient %in% p) %>% group_by(Patient) %>% summarise(final.endo-initial.endo)
-
-rrPath = back %>% group_by(Status, Pathology) %>% summarise( 'mean.RR'=mean(OR),'ci.min'=mean(OR)-(1.96*sd(OR)), 'ci.max'=mean(OR)+(1.96*sd(OR)) )
+subset(sum.patient.data, Patient %in% p) %>% dplyr::group_by(Patient) %>% dplyr::summarise(final.endo-initial.endo)
+
+rrPath = back %>% dplyr::group_by(Status, Pathology) %>% dplyr::summarise( 'mean.RR'=mean(OR),'ci.min'=mean(OR)-(1.96*sd(OR)), 'ci.max'=mean(OR)+(1.96*sd(OR)) )
 
@@
 
-risksL = subset(bd, Status == 'NP') %>% group_by(Patient) %>% summarise( 'n.samp'=length(Patient), 'high.risk'=length(which(Risk == 'High')), 'mod.risk'=length(which(Risk == 'Moderate')) )
+risksL = subset(bd, Status == 'NP') %>% dplyr::group_by(Patient) %>% dplyr::summarise( 'n.samp'=length(Patient), 'high.risk'=length(which(Risk == 'High')), 'mod.risk'=length(which(Risk == 'Moderate')) )
 
@@
 for (yr in names(which(table(bd$Endoscopy.Year) > 10))) {
-  risks = subset(bd, Status == 'NP' & Endoscopy.Year == yr) %>% group_by(Patient) %>% summarise( 'n.samp'=length(Patient), 'high.risk'=length(which(Risk == 'High')), 'mod.risk'=length(which(Risk == 'Moderate')) )
+  risks = subset(bd, Status == 'NP' & Endoscopy.Year == yr) %>% dplyr::group_by(Patient) %>% dplyr::summarise( 'n.samp'=length(Patient), 'high.risk'=length(which(Risk == 'High')), 'mod.risk'=length(which(Risk == 'Moderate')) )
 
@@
 ## Clonal Heterogenity
-patient.info %>% group_by (Hospital.Research.ID, Patient, PID) %>% summarise(
+patient.info %>% dplyr::group_by (Hospital.Research.ID, Patient, PID) %>% dplyr::summarise(
   length(PID)
@@
 
-pp = back %>% group_by(Patient, Hospital.Research.ID, PID) %>% summarise(
+pp = back %>% dplyr::group_by(Patient, Hospital.Research.ID, PID) %>% dplyr::summarise(
   'max'=max(Prediction), 'avg'=mean(Prediction)
@@
 prog.pred = lapply(pg.samp, function(df) {
-  df = df[,c('Hospital.Research.ID', 'Status', 'Patient', 'Age.at.diagnosis', 'PID', 'Status', 'Endoscopy.Year','Initial.Endoscopy', 'Final.Endoscopy', 'Pathology', 'Samplename','Prediction','Prediction.Dev.Resid')] %>% rowwise() %>% mutate(
+  df = df[,c('Hospital.Research.ID', 'Status', 'Patient', 'Age.at.diagnosis', 'PID', 'Status', 'Endoscopy.Year','Initial.Endoscopy', 'Final.Endoscopy', 'Pathology', 'Samplename','Prediction','Prediction.Dev.Resid')] %>% rowwise() %>% dplyr::mutate(
       'Time.to.Final'=Final.Endoscopy-Endoscopy.Year)

### 2018-09-05 snp/SNP_Prediction_Analysis.Rmd
add Seattle SNP analysis scripts

--- 2018-07-25 SNP_Prediction_Analysis.Rmd
@@
 date: "6/13/2018"
-output: html_document
+output: 
+  html_document: 
+    toc: yes
 ---
@@
 
-suppressPackageStartupMessages(source('lib/data_func.R'))
-suppressPackageStartupMessages(source('lib/common_plots.R'))
+source('~/workspace/shallowwgs_pipeline/lib/data_func.R')
+source('~/workspace/shallowwgs_pipeline/lib/common_plots.R')
 
-chr.info = get.chr.lengths(file='~/tmp/hg19_info.txt')
+chr.info = get.chr.lengths(file='/tmp/hg19_info.txt')
 
@@
 
-qcdata = base::merge(qcdata, sample.info[,c('Samplename','Status','Timepoint','SCA.Ratio')], by='Samplename')
+qcdata = base::merge(qcdata, sample.info[,c('PatientID','Samplename','Status','Timepoint','SCA.Ratio')], by='Samplename')
+
+ggplot(subset(qcdata,SampleType == 'BE'), aes(Timepoint, SCA.Ratio, color=PatientID, group=Timepoint)) + geom_boxplot() + 
+  facet_wrap(~Status, scales='free_x', ncol=2) + labs(title='SCA Ratio by Timepoint') + theme(legend.position = 'none')
+
+ggplot(subset(qcdata,SampleType == 'BE'), aes(Timepoint, SCA.Ratio, color=PatientID, group=PatientID)) + geom_point() + geom_line() + 
+  facet_wrap(~Status, scales='free_x', ncol=2) + labs(title='SCA Ratio by Timepoint') + theme(legend.position = 'none')
 
 ggplot(qcdata, aes(round(Ploidy,2))) + geom_histogram() + 
-  facet_wrap(~SampleType, scales='free_y', ncol=1) + labs(x='Ploidy', title='Ploidy, normal vs BE')
+  facet_wrap(~SampleType, scales='free', ncol=1) + labs(x='Ploidy', title='Ploidy, normal vs BE')
 
-ggplot(qcdata, aes(Purity, Goodnessoffit, color=Status)) + geom_point() + facet_wrap(~SampleType, ncol=1) + 
-  labs(title='Purity~fit, normal vs BE')
+ggplot(qcdata, aes(round(Purity,2))) + geom_histogram() + 
+  facet_wrap(~SampleType, scales='free', ncol=1) + labs(x='Purity', title='Purity, normal vs BE')
+
+ggplot(qcdata, aes(round(Ploidy,2), Purity, group=round(Ploidy,2), color=Status)) + geom_point() + 
+  facet_wrap(~Status, scales='free', ncol=2) + labs(x='Ploidy', title='Ploidy vs Purity, normal vs BE')
+
+#x=median(qcdata$SCA.Ratio,na.rm=T)+sd(qcdata$SCA.Ratio,na.rm=T)
+x=0.01
+y=median(qcdata$Purity,na.rm=T)-sd(qcdata$Purity,na.rm=T)/2
+
+ggplot(qcdata, aes(Purity, SCA.Ratio, color=Status)) + geom_point() + 
+         facet_wrap(~Status, ncol=1) + labs(title='Purity~SCA.Ratio, normal vs BE') + geom_vline(xintercept=y, color='grey')
+
+ggplot(qcdata, aes(SCA.Ratio, Purity, color=Status)) + geom_point() + 
+  facet_wrap(~Status) + labs(x='SCA', title='SCA vs Purity') + 
+  geom_vline(xintercept = x, color='grey') + geom_hline(yintercept = y, color='grey')
+
+table(subset(qcdata, Purity > y & SCA.Ratio < x & SampleType == 'BE')$Status)
+
+ggplot(qcdata, aes(Purity)) + geom_histogram() + geom_vline(xintercept = y, col='red') + 
+  facet_wrap(~SampleType, scales = 'free') + labs(title='Purity by sampletype')
+
+ggplot(qcdata, aes(Purity, Goodnessoffit, color=Status)) + geom_point() + 
+         facet_wrap(~SampleType, ncol=1) + labs(title='Purity~fit, normal vs BE') + geom_vline(xintercept=y, color='grey')
 
 ggplot(subset(qcdata, SampleType == 'BE'), aes(round(Ploidy,2), Purity, color=Status)) + geom_point() + facet_grid(~Status) + 
-  labs(title='BE only, purity vs ploidy')
+  labs(title='BE only, purity vs ploidy') + geom_hline(yintercept=y, color='grey')
 
-ggplot(qcdata, aes(SCA.Ratio, color=Status, fill=Status)) + geom_density(alpha=0.5) + labs(x='SCA Ratio', title='SCA Ratio, NP vs P')
+ggplot(qcdata, aes(SCA.Ratio, color=Status, fill=Status)) + facet_wrap(~Status, scales='free_y',nrow=1) +  geom_density(alpha=0.5) + labs(x='SCA Ratio', title='SCA Ratio, NP vs P')
+
+summary(qcdata$SCA.Ratio)
 
@@
 
-#copySegs = as.data.frame(normalize.quantiles.use.target(mergedSegs, as.vector(raw.segs)))
 grid.arrange(
@@
 
+In the gastric/blood normals `r signif(length(which(preds.normal$Prob < 0.5))/nrow(preds.normal),2)` are predicted to have <50% probability (TPR).
+
 ## Evaluating predictions
@@
 ```{r, warning=F, fig.height=8}
-
 qcdata.be = subset(qcdata, SampleType == 'BE')
@@
 
-ggplot(pred.qc, aes(`mean(SCA.Ratio)`, Prob, color=Status)) + geom_point()
-ggplot(pred.qc, aes(`mean(Ploidy)`, Prob, color=Status)) + geom_point()
+grid.arrange(
+  ggplot(pred.qc, aes(`mean(SCA.Ratio)`, Prob, color=Status)) + geom_point(),
+  ggplot(pred.qc, aes(`mean(Ploidy)`, Prob, color=Status)) + geom_point(),
+ncol=2)
 
-pander(cor.test(pred.qc$`mean(SCA.Ratio)`, pred.qc$Prob))
+pander(cor.test(pred.qc$`mean(SCA.Ratio)`, pred.qc$Prob), caption='Probability shows a 0.3 correlation with % of the genome altered')
+```
 
+```{r, warning=F, fig.height=8}
 prog = subset(pred.qc, Status == 'P')
@@
 )
+
 pred.qc = pred.qc %>% group_by(Status, Samplename) %>% dplyr::mutate(
-  lowsca = (Status == 'P' & (`mean(SCA.Ratio)` < 0.03 | `mean(Ploidy)` < 1.5))
+  lowsca = ( Status == 'P' & (`mean(SCA.Ratio)` < 0.03 & `mean(Purity)` >= 0.95))
 )
 
-subset(pred.qc, `mean(Ploidy)` < 1.99)
+# pred.qc = pred.qc %>% group_by(Status, Samplename) %>% dplyr::mutate(
+#   lowsca = ( `mean(SCA.Ratio)` < 0.01 & `mean(Purity)` >= 0.95)
+# )
 
-subset(pred.qc, lowsca)
+#table(subset(pred.qc, !lowsca)$Status)/table(pred.qc$Status)
+
+ggplot(pred.qc, aes(`mean(SCA.Ratio)`, `mean(Purity)`, color=Status)) + geom_point() + facet_grid(~Status) + 
+  geom_vline(xintercept=0.01) + geom_hline(yintercept = 0.95)
+
+ggplot(pred.qc, aes(`mean(SCA.Ratio)`))  + geom_histogram() + facet_wrap(~Status, scales='free_y')
+
+ggplot(pred.qc, aes(`mean(Purity)`))  + geom_histogram() + facet_wrap(~Status, scales='free_y')
+
+ggplot(pred.qc, aes(Prob)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
+  scale_fill_gradientn(colors = myPal,  name='') + 
+  labs(title='Predictions, all samples model', y='n Samples', x='Unadjusted Probability')
 
@@
 qt$NP = round(qt$NP/sum(qt$NP), 2)
-pander(qt)
+pander(qt, caption='Excluding HGD NP, WGD NP, low SCA P')
 
@@
 
-per_pt = 
-  subset(pred.qc, !exclude) %>% group_by(PatientID, Status) %>% dplyr::summarise( max.Prob = max(Prob), max.RR = max(RR), max.Risk = max(Risk) )
-rocpp = pROC::roc(Status ~ max.Prob, data=per_pt, auc=T, ci=T, of='thresholds')
-roc.plot(rocpp) + labs(title='Max per patient - excluded HGD NP')
+# per_pt = 
+#   subset(pred.qc, !exclude) %>% group_by(PatientID, Status) %>% dplyr::summarise( max.Prob = max(Prob), max.RR = max(RR), max.Risk = max(Risk) )
+# rocpp = pROC::roc(Status ~ max.Prob, data=per_pt, auc=T, ci=T, of='thresholds')
+# roc.plot(rocpp) + labs(title='Max per patient - excluded HGD NP')
+
+# Exclude the WGD NP
+rocpp = pROC::roc(Status ~ Prob, data=subset(pred.qc, !wgd), auc=T, ci=T, of='thresholds')
+roc.plot(rocpp) + labs(title='Both endos - excluded WGD NP')
+pROC::coords(rocpp, 'best')
 
@@
 rocpp = pROC::roc(Status ~ Prob, data=subset(pred.qc, !lowsca), auc=T, ci=T, of='thresholds')
-roc.plot(rocpp) + labs(title='Both endos - excluded low SCA P')
-pROC::coords(rocpp, 'best')
-
-# Exclude the tetraploid NP
-rocpp = pROC::roc(Status ~ Prob, data=subset(pred.qc, !wgd), auc=T, ci=T, of='thresholds')
-roc.plot(rocpp) + labs(title='Both endos - excluded WGD NP')
+roc.plot(rocpp) + labs(title='Both endos - excluded low SCA')
 pROC::coords(rocpp, 'best')
@@
 
-```{r regression, echo=F, message=F, eval=F}
+```{r regression, echo=F, message=F, eval=T}
+folds = 10; splits = 5
 ## ----- All ----- ##
@@
 
-```{r loo, eval=F}
-
-slabels$Prediction = NA
-slabels$RR = NA
+```{r loo, eval=T}
 
@@
 
-slabels
 snp.roc = pROC::roc(Status~Prediction,slabels, auc=T, ci=T, of='thresholds')

### 2018-09-05 snp/process_snp.R

@@ -1,0 +1,249 @@
+options(bitmapType = "cairo")
+
+library(ggplot2)
+library(gridExtra)
+library(dplyr)
+
+#suppressPackageStartupMessages( source('lib/data_func.R') )
+suppressPackageStartupMessages( source('~/workspace/shallowwgs_pipeline/lib/data_func.R') )
+
+args = commandArgs(trailingOnly=TRUE)
+
+#if (length(args) < 2)
+#  stop("Missing data directory or summary file")
+
+if (length(args) < 1)
+  stop("Missing data directory ")
+
+
+datadir = args[1]
+#summary.file = args[2]
+
+# datadir = '~/Data/Reid_SNP/PerPatient-preLM/512'
+
+
+byEndo = F
+if (length(args) > 2)
+  byEndo = as.logical(args[3])
+
+if (!dir.exists(datadir))
+  stop(paste(datadir, "doesn't exist or isn't readable"))
+
+print(datadir)
+#print(summary.file)
+
+#smy = read.table(summary.file, sep = '\t', header = T)
+#smy = subset(smy, Copy.number <= 12)
+
+# a = as.data.frame(smy[,c('Copy.number','var.LRR')])
+# colnames(a) = c('CN','y')
+# b = as.data.frame(smy[,c('Copy.number','mean.LRR')])
+# colnames(b) = c('CN','y')
+# 
+# fitSD = lm(y~CN, data=a)
+# fitM = lm(y~CN, data=b)
+# #fitCN = lm(mean~CN, data=b)
+
+#autoplot(fitM)
+
+### Lifted directly from ASCAT
+#Perform MAD winsorization:
+madWins <- function(x,tau,k){
+  xhat <- medianFilter(x,k)
+  d <- x-xhat
+  SD <- mad(d)
+  z <- tau*SD
+  xwin <- xhat + psi(d, z)
+  outliers <- rep(0, length(x))
+  outliers[x > xwin] <- 1
+  outliers[x < xwin] <- -1
+  return(list(ywin=xwin,sdev=SD,outliers=outliers))
+}
+
+psi <- function(x,z){
+  xwin <- x
+  xwin[x < -z] <- -z
+  xwin[x > z] <- z
+  return(xwin)
+}
+
+medianFilter <- function(x,k){
+  n <- length(x)
+  filtWidth <- 2*k + 1
+  #Make sure filtWidth does not exceed n
+  if(filtWidth > n){
+    if(n==0){
+      filtWidth <- 1
[…]

### 2018-09-07 snp/SNP_Prediction_Analysis.Rmd
Adding QC to pipeline

@@
 
+# pred.qc = pred.qc %>% group_by(Status, Samplename) %>% dplyr::mutate(
+#   lowsca = ( Status == 'P' & (`mean(SCA.Ratio)` < 0.03 & `mean(Purity)` >= 0.95))
+# )
+
 pred.qc = pred.qc %>% group_by(Status, Samplename) %>% dplyr::mutate(
-  lowsca = ( Status == 'P' & (`mean(SCA.Ratio)` < 0.03 & `mean(Purity)` >= 0.95))
+  lowsca = ( `mean(SCA.Ratio)` < 0.03 & `mean(Purity)` >= 0.95)
 )
 
-# pred.qc = pred.qc %>% group_by(Status, Samplename) %>% dplyr::mutate(
-#   lowsca = ( `mean(SCA.Ratio)` < 0.01 & `mean(Purity)` >= 0.95)
-# )
+ggplot(pred.qc, aes(`mean(SCA.Ratio)`, `mean(Purity)`, color=Status)) + geom_point() + facet_grid(~Status) + 
+  geom_vline(xintercept=0.03) + geom_hline(yintercept = 0.95)
 
-#table(subset(pred.qc, !lowsca)$Status)/table(pred.qc$Status)
-
-ggplot(pred.qc, aes(`mean(SCA.Ratio)`, `mean(Purity)`, color=Status)) + geom_point() + facet_grid(~Status) + 
-  geom_vline(xintercept=0.01) + geom_hline(yintercept = 0.95)
+table(subset(pred.qc, !lowsca)$Status)
+table(subset(pred.qc, !lowsca)$Status)/table(pred.qc$Status)
 
@@
 ggplot(pred.qc, aes(`mean(Purity)`))  + geom_histogram() + facet_wrap(~Status, scales='free_y')
-
-ggplot(pred.qc, aes(Prob)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
-  scale_fill_gradientn(colors = myPal,  name='') + 
-  labs(title='Predictions, all samples model', y='n Samples', x='Unadjusted Probability')
 
@@
 
-ggplot(subset(pred.qc, !exclude & !wgd & !lowsca), aes(`mean(Purity)`,Prob, color=Risk, shape=Status)) + 
+ggplot(subset(pred.qc, !exclude & !wgd & !lowsca), aes(`mean(SCA.Ratio)`,Prob, color=Risk, shape=Status)) + 
   facet_grid(~Status) + geom_point()
@@
 # nrow(subset(ex.pred, Prob < 0.5 & Status == 'NP'))/table(ex.pred$Status)['NP'] #
+
+ct = cor.test(pred.qc.ex$Prob, pred.qc.ex$`mean(SCA.Ratio)`)
+ct2 = cor.test(pred.qc$Prob, pred.qc$`mean(SCA.Ratio)`)
 ```
+
+The risk probability is correlated with the % of the genome that's been altered as identified by Xiahong, R=`r ct$estimate` and that's true even when no patients are excluded (R=`r ct2$estimate`).
 
@@
 
-```{r loo, eval=T}
+```{r loo, eval=F}
 

### 2018-09-07 lib/data_func.R

@@ -41,6 +41,7 @@ sample.residual.variance<-function(segment.residuals) {
             'var1sd'=var(y[y > median(y)-sd(y) & y < median(y)+sd(y)]),
             'var2sd'=var(y[y > median(y)-sd(y)*2 & y < median(y)+sd(y)*2]),
             'var3sd'=var(y[y > median(y)-sd(y)*3 & y < median(y)+sd(y)*3]),
+            'varMAD'=var(y[y<mad(y) & y>-mad(y)]),
             'var'=var(y) )
     }))
   })

### 2018-09-11 snp/SNP_Prediction_Analysis.Rmd
...

@@
 
-load('~/Data/Ellie/Analysis/5e6_arms_all/model_data.Rdata', verbose = T)
-load('~/Data/Ellie/Analysis/5e6_arms_all/all.pt.alpha.Rdata', verbose = T)
+load('~/Data/Ellie/Analysis/5e6_arms_all-NOQC/model_data.Rdata', verbose = T)
+load('~/Data/Ellie/Analysis/5e6_arms_all-NOQC/all.pt.alpha.Rdata', verbose = T)
 #alpha = '0'
@@
 ```{r rnormals, echo=F, warning=F, message=F, fig.width=10, fig.height=8, eval=T}
-randNormal = sample(rownames(mergedSegs)[nm.rows], 1)
+randNormal = sample( grep('BLD',rownames(mergedSegs)[nm.rows], value=T), 1)
 pt = unlist(strsplit(randNormal, '_'))[1]
@@
 #randBE = paste(unlist(strsplit(randBE, '_'))[c(1,3)], collapse='_')
+
 ```
@@
 In the gastric/blood normals `r signif(length(which(preds.normal$Prob < 0.5))/nrow(preds.normal),2)` are predicted to have <50% probability (TPR).
+
+For blood normals only `r signif(nrow(subset(preds.normal, grepl('BLD',Samplename) & Prob < 0.5))/nrow(subset(preds.normal, grepl('BLD',Samplename) )),2)` have a <50% probabilitye (TPR).
 
@@
 
-p3 = ggplot(preds.normal, aes(Prob)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
+preds.normal = preds.normal %>% dplyr::mutate(Type = ifelse(grepl('BLD',Samplename), 'Blood','Gastric'))
+
+p3 = ggplot(preds.normal, aes(Prob)) + facet_wrap(~Type, ncol=2, scales = 'free') +
+  geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
   scale_fill_gradientn(colors = myPal,  name='') +
@@
 
-table(subset(pred.qc, !lowsca)$Status)
-table(subset(pred.qc, !lowsca)$Status)/table(pred.qc$Status)
-
-ggplot(pred.qc, aes(`mean(SCA.Ratio)`))  + geom_histogram() + facet_wrap(~Status, scales='free_y')
-
-ggplot(pred.qc, aes(`mean(Purity)`))  + geom_histogram() + facet_wrap(~Status, scales='free_y')
+pander(
+  rbind(rbind('n.Samples'=table(subset(pred.qc, !lowsca)$Status)),
+  rbind('n.Patients'=table(unique(subset(pred.qc, !lowsca, select=c('PatientID','Status')))$Status))),
+caption='Number of samples and patients that pass the low SCA:high purity thresholds')
 

### 2018-09-14 snp/SNP_Prediction_Analysis.Rmd
adding exclusions for low SCA

@@
 
+qcdata = qcdata %>% rowwise %>% dplyr::mutate(
+  PatientID = unlist(strsplit(Samplename, '_'))[1],
+  SampleID = unlist(strsplit(Samplename, '_'))[2],
+  EndoID = unlist(strsplit(Samplename, '_'))[3],
+  Level = unlist(strsplit(Samplename, '_'))[4]
+)
+segments.list = lapply(segments.list, function(pt) {
+  pt$sample = sub('\\.LogR','',pt$sample)
+  pt
+})
+
+qcdata$ASCAT.SCA.ratio = apply(qcdata,1,function(s) {
+  smp = subset(segments.list[[ s[['PatientID']] ]], sample == s[['Samplename']] & chr %in% c(1:22))
+  smp = smp %>% rowwise %>% dplyr::mutate(
+      #'Total' = nAraw + nBraw,
+      'Total' = nMajor + nMinor,
+      'CNV' = round(Total) - round(as.numeric(s[['Ploidy']])) )
+
+  x = subset(smp, CNV != 0 & chr %in% c(1:22))
+  sum(x$endpos - x$startpos) / chr.info[22,'genome.length']
+})
+
+qcdata$SampleType = 'BE'
+qcdata$SampleType[grep('BLD',qcdata$Level, ignore.case=T)] = 'Blood Normal'
+qcdata$SampleType[grep('gastric',qcdata$Level, ignore.case=T)] = 'Gastric Normal'
+
 patient.info = as.data.frame(read_xlsx('~/Data/Ellie/Analysis/SNP/metadata_T1T2.xlsx'))
@@
 
-sample.info  = do.call(rbind.data.frame, sapply(qcdata$Samplename, strsplit, '_'))
-colnames(sample.info) = c('PatientID','SampleID','EndoID','Level')
-sample.info[] = lapply(sample.info[], as.character)
-sample.info[which(sample.info$PatientID == 524 & sample.info$Level == 524), 'Level'] = ''
-sample.info$Samplename = qcdata$Samplename
+sample.info = qcdata[,c('PatientID','SampleID','EndoID','Level','Samplename','ASCAT.SCA.ratio')]
 
@@
 sample.info = base::merge(sample.info, patient.info[,c('PatientID','Timepoint','Timepoint Code','Status','Pathology')], by.x=c('PatientID','EndoID'), by.y=c('PatientID','Timepoint Code'))
+head(sample.info)
 
@@
 patient.info = patient.info %>% rowwise() %>% dplyr::mutate( Endoscopy = ifelse(Timepoint == 'T1', 1, 2) )
+
+sample.info
+ggplot(sample.info, aes(ASCAT.SCA.ratio, SCA.Ratio)) + geom_point() + facet_grid(~Status)
+
+# ASCAT and Xiahong's estimates correlate in 86% of samples, that suggests that the 14% missing is due to the LOH I'm not trying to call
+cor.test(sample.info$ASCAT.SCA.ratio, sample.info$SCA.Ratio)
+
 ```
@@
 ```{r, echo=F, warning=F, message=F}
-qcdata$SampleType = 'BE'
-qcdata$SampleType[grep('BLD',rownames(qcdata), ignore.case=T)] = 'Blood Normal'
-qcdata$SampleType[grep('gastric',rownames(qcdata), ignore.case=T)] = 'Gastric Normal'
-head(qcdata)
+qcdata = base::merge(qcdata, sample.info[,c('Samplename','Status','Timepoint','SCA.Ratio')], by='Samplename')
 
-qcdata = base::merge(qcdata, sample.info[,c('PatientID','Samplename','Status','Timepoint','SCA.Ratio')], by='Samplename')
+ggplot(qcdata, aes(Status, ASCAT.SCA.ratio, color=Status, group=Status)) + geom_boxplot() + facet_grid(~SampleType)
 
-ggplot(subset(qcdata,SampleType == 'BE'), aes(Timepoint, SCA.Ratio, color=PatientID, group=Timepoint)) + geom_boxplot() + 
-  facet_wrap(~Status, scales='free_x', ncol=2) + labs(title='SCA Ratio by Timepoint') + theme(legend.position = 'none')
+low = subset(qcdata, SampleType == 'BE' & ASCAT.SCA.ratio <= median(subset(qcdata, SampleType != 'BE')$ASCAT.SCA.ratio))
 
-ggplot(subset(qcdata,SampleType == 'BE'), aes(Timepoint, SCA.Ratio, color=PatientID, group=PatientID)) + geom_point() + geom_line() + 
+anno = cbind.data.frame(table(subset(qcdata, SampleType == 'BE' & ASCAT.SCA.ratio <= median(subset(qcdata, SampleType != 'BE')$ASCAT.SCA.ratio))$Status)/table(subset(qcdata, SampleType == 'BE')$Status))
+colnames(anno) = c('Status','Ratio')
+anno$label='median(normal SCA)'
+anno$x = 0.4
+anno$y = median(subset(qcdata, SampleType != 'BE')$ASCAT.SCA.ratio)
+
+ggplot(subset(qcdata,SampleType == 'BE'), aes(Purity, ASCAT.SCA.ratio, color=Status, group=Status)) + geom_point() + facet_grid(~Status, scales = 'free') + ylim(0,0.005) + 
+  geom_hline(yintercept = x) + geom_text(data=anno, aes(x, y, label=label), nudge_y = 0.0001) +
+  geom_text(data=anno, aes(x, y, label=paste(round(Ratio,2)*100,'% samples',sep='')), nudge_y = -0.0001)
+
+# ggplot(subset(qcdata,SampleType == 'BE'), aes(Timepoint, SCA.Ratio, color=PatientID, group=PatientID)) + geom_point() + geom_line() + 
+#   facet_wrap(~Status, scales='free_x', ncol=2) + labs(title='SCA Ratio by Timepoint') + theme(legend.position = 'none')
+
+ggplot(qcdata, aes(Ploidy, ASCAT.SCA.ratio, color=Status)) + geom_point() + facet_grid(~SampleType)
+
+ggplot(qcdata, aes(Ploidy, ASCAT.SCA.ratio, color=Status)) + geom_point() + facet_grid(~Status)
+
+ggplot(subset(qcdata,SampleType == 'BE'), aes(Timepoint, ASCAT.SCA.ratio, color=PatientID, group=Timepoint)) + geom_boxplot() + 
   facet_wrap(~Status, scales='free_x', ncol=2) + labs(title='SCA Ratio by Timepoint') + theme(legend.position = 'none')
@@
 ggplot(qcdata, aes(round(Ploidy,2))) + geom_histogram() + 
-  facet_wrap(~SampleType, scales='free', ncol=1) + labs(x='Ploidy', title='Ploidy, normal vs BE')
+  facet_wrap(~SampleType, ncol=1) + labs(x='Ploidy', title='Ploidy, normal vs BE')
 
@@
 
-ggplot(qcdata, aes(round(Ploidy,2), Purity, group=round(Ploidy,2), color=Status)) + geom_point() + 
-  facet_wrap(~Status, scales='free', ncol=2) + labs(x='Ploidy', title='Ploidy vs Purity, normal vs BE')
-
 #x=median(qcdata$SCA.Ratio,na.rm=T)+sd(qcdata$SCA.Ratio,na.rm=T)
-x=0.01
+x=0.05
 y=median(qcdata$Purity,na.rm=T)-sd(qcdata$Purity,na.rm=T)/2
 
-ggplot(qcdata, aes(Purity, SCA.Ratio, color=Status)) + geom_point() + 
-         facet_wrap(~Status, ncol=1) + labs(title='Purity~SCA.Ratio, normal vs BE') + geom_vline(xintercept=y, color='grey')
+ggplot(qcdata, aes(Purity, ASCAT.SCA.ratio, color=ASCAT.SCA.ratio < 0.2)) + geom_point() + 
+         facet_wrap(~Status, ncol=1) + labs(title='Purity~SCA.Ratio') + geom_vline(xintercept=y, color='grey')
 
-ggplot(qcdata, aes(SCA.Ratio, Purity, color=Status)) + geom_point() + 
+ggplot(qcdata, aes(ASCAT.SCA.ratio, Purity, color=Status)) + geom_point() + 
   facet_wrap(~Status) + labs(x='SCA', title='SCA vs Purity') + 
@@
 
-table(subset(qcdata, Purity > y & SCA.Ratio < x & SampleType == 'BE')$Status)
+table(unique(subset(qcdata, Purity > y & ASCAT.SCA.ratio < x & SampleType == 'BE', select=c('Status','PatientID')))$Status)
 
@@
 
-ggplot(qcdata, aes(SCA.Ratio, color=Status, fill=Status)) + facet_wrap(~Status, scales='free_y',nrow=1) +  geom_density(alpha=0.5) + labs(x='SCA Ratio', title='SCA Ratio, NP vs P')
+ggplot(qcdata, aes(ASCAT.SCA.ratio, color=Status, fill=Status)) + facet_wrap(~Status, scales='free_y',nrow=1) +  geom_density(alpha=0.5) + labs(x='SCA Ratio', title='SCA Ratio, NP vs P')
 
-summary(qcdata$SCA.Ratio)
+ggplot(qcdata, aes(ASCAT.SCA.ratio, color=Status, fill=Status)) + facet_wrap(Status~SampleType, scales='free_y',nrow=2) +  geom_density(alpha=0.5) + labs(title='SCA Ratio, NP vs P')
 
-ggplot(subset(qcdata, SampleType == 'BE'), aes(Purity, SCA.Ratio, color=Status)) + facet_grid(~Status) +
-  geom_point() + geom_vline(xintercept = 0.5, color='darkgreen') + geom_hline(yintercept = 0.05, color='red') + 
+summary(subset(qcdata, SampleType == 'BE')$ASCAT.SCA.ratio)
+summary(subset(qcdata, SampleType != 'BE')$ASCAT.SCA.ratio)
+
+ggplot(subset(qcdata, SampleType == 'BE'), aes(Purity, ASCAT.SCA.ratio, color=Status)) + facet_grid(~Status) +
+  geom_point() + geom_vline(xintercept = y, color='darkgreen') + geom_hline(yintercept = x, color='red') + 
   labs(title='Barretts only')
 
-ggplot(subset(qcdata, SampleType == 'BE'), aes(Ploidy, SCA.Ratio, color=Status)) + facet_grid(~Status) + geom_point() + 
-  geom_hline(yintercept = 0.03, color='red') + labs(title='Barretts only')
+ggplot(subset(qcdata, SampleType == 'BE'), aes(Ploidy, ASCAT.SCA.ratio, color=Status)) + facet_grid(~Status) + geom_point() + 
+  geom_hline(yintercept = 0.02, color='red') + labs(title='Barretts only')
 
 ## PER ENDOSCOPY
-qcdata.samples = qcdata %>% group_by(Samplename) %>% dplyr::mutate(
-  PatientID = unlist(strsplit(Samplename, '_'))[1],
-  EndoID = unlist(strsplit(Samplename, '_'))[3],
-  Level = unlist(strsplit(Samplename, '_'))[4] 
-)
+qcdata.endo = qcdata %>% group_by(PatientID, EndoID, Status, Timepoint, SampleType) %>% dplyr::summarise_if(is.numeric, c('mean','max','min','sd')) %>% mutate_if(is.numeric, round, digits=3)
+qcdata %>% dplyr::group_by( Status, SampleType) %>% dplyr::summarise( median(ASCAT.SCA.ratio), sd(ASCAT.SCA.ratio) )
 
-qcdata = qcdata.samples %>% group_by(PatientID, EndoID, Status, Timepoint, SampleType) %>% dplyr::summarise(
-  Ploidy = mean(Ploidy), Purity = mean(Purity), Goodnessoffit = mean(Goodnessoffit), SCA.Ratio = mean(SCA.Ratio)
-)
+#ggplot(subset(qcdata.endo, SampleType == 'BE'), aes(ASCAT.SCA.ratio_max, Purity_max)) + geom_point() + facet_grid(~Status) + geom_hline(yintercept = 0.95)  + geom_vline(xintercept = 0.001)
 
-## Low CN
-lowPloidyCN = subset(qcdata, Status == 'P' & SampleType == 'BE' & SCA.Ratio <= 0.05 & Ploidy <= 2)
+lowsca = subset(qcdata.endo, SampleType == 'BE' & ASCAT.SCA.ratio_max <= 0.01 & Purity_max > 0.98 )
 
-## samples have no apparent CN
-noCN = subset(qcdata, Status == 'P' & SCA.Ratio <= 0 & SampleType == 'BE')
+table(unique(lowsca[,c('PatientID','Status')])$Status)/table(unique(qcdata.endo[,c('PatientID','Status')])$Status)
 
-## high CN | ploidy, nonP
-highPloidyCN = subset(qcdata, Status == 'NP' & SampleType == 'BE' & (SCA.Ratio >= mean(qcdata$SCA.Ratio,na.rm=T)+ sd(qcdata$SCA.Ratio,na.rm=T)*2 | Ploidy >= 2.7))
+table(unique(subset(qcdata.endo, SampleType == 'BE' & ASCAT.SCA.ratio_max <= 0.001, select=c('PatientID','Status')))$Status)
 
-qcdata$SCA.Ratio = round(qcdata$SCA.Ratio, 4)
+ggplot(subset(qcdata.endo,SampleType == 'BE'), aes(Status, ASCAT.SCA.ratio_max, group=Status)) + geom_boxplot() + geom_hline(yintercept = 0.002, color='red')
 
-lowsca = subset(qcdata, Status == 'P' & SampleType == 'BE' & SCA.Ratio <= 0.05 & Ploidy <= 2)
+qcdata %>% dplyr::group_by( Status, SampleType) %>% dplyr::summarise( median(ASCAT.SCA.ratio))
 
 ```
-
-#### P: Low SCA, Ploidy<=2
-`r round(nrow(lowPloidyCN)/nrow(subset(qcdata, Status == 'P' & SampleType == 'BE')), 3)*100`% of PROGRESSOR endoscopies have <5% SCA and an ASCAT ploidy of 2 or less.
-Most of these show a >90% purity as well.
-`r pander(summary(lowPloidyCN$Purity))`
-
-#### P: No SCA
-`r round(nrow(noCN)/nrow(subset(qcdata, Status == 'P' & SampleType == 'BE')), 3)*100`% of PROGRESSOR endoscopies have 0 SCA
-
-#### NP: High CN or ploidy > 2.3
-`r round(nrow(highPloidyCN)/nrow(subset(qcdata, Status == 'NP' & SampleType == 'BE')), 3)*100`% of NON-PROGRESSOR samples have >2SD SCA (~25%) or high ploidy.
 
@@
   x = unlist(strsplit(sample, '_'))
-  ploidy = mean(subset(qcdata, SampleType == 'BE' & PatientID == x[1] & EndoID == x[2])$Ploidy)
+  #ploidy = mean(subset(qcdata, SampleType == 'BE' & PatientID == x[1] & EndoID == x[2])$Ploidy)
+  ploidy = subset(qcdata.endo, PatientID == x[1] & EndoID == x[2] & SampleType == 'BE')$Ploidy_mean
   if (grepl('BLD|gastric', sample, ignore.case = T)) ploidy = 2
@@
 
-```{r, echo=F, fig.height=12, fig.width=6}
-## TODO: quantile normalize the SNP arrays by the sWGS values. 
-# copy = do.call(cbind.data.frame,lapply(1:ncol(arrayDf), function(i) {
-#   as.data.frame(normalize.quantiles.use.target(arrayDf[,i,drop=F], as.vector(allDf[,i])))
-# }))
-# 
-# colnames(copy) = colnames(arrayDf)
-# rownames(copy) = rownames(arrayDf)
-
-grid.arrange(
-  ggplot( melt(dysplasia.df), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, sWGS'),
-  ggplot( melt(arrayDf), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, SNP arrays'),
-  #ggplot( melt(copy), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, SNP arrays q-normed'),
-  ncol=1
-)
-#arrayDf = as.matrix(copy)
-```
-
 # Predictions
@@
 ```{r, warning=F, fig.height=8}
-qcdata.be = subset(qcdata, SampleType == 'BE')
-get.qcinfo<-function(vec) {
-  endo = vec[2]
-  pt = vec[1]
-  y = subset(qcdata.be, PatientID == pt & EndoID == endo)
-  y = y %>% group_by(Status, Timepoint) %>%  dplyr::summarise( mean(Ploidy), mean(Purity), mean(Goodnessoffit), mean(SCA.Ratio) )
-  cbind.data.frame('PatientID'=pt, 'EndoID'=endo, y, stringsAsFactors = F)
-}
+#qcdata.be = subset(qcdata, SampleType == 'BE')
+# get.qcinfo<-function(vec) {
+#   endo = vec[2]
+#   pt = vec[1]
+#   y = subset(qcdata.be, PatientID == pt & EndoID == endo)
+#   y = y %>% group_by(Status, Timepoint) %>%  dplyr::summarise( mean(Ploidy), mean(Purity), mean(Goodnessoffit), mean(SCA.Ratio) )
+#   cbind.data.frame('PatientID'=pt, 'EndoID'=endo, y, stringsAsFactors = F)
+# }
+preds = preds %>% dplyr::group_by(Samplename) %>% dplyr::mutate(
+  PatientID = unlist(strsplit(Samplename,'_'))[1],
+  EndoID = unlist(strsplit(Samplename,'_'))[2]
+)
 
-qc = do.call(rbind.data.frame, c( lapply(preds$Samplename[-pnr], function(x) get.qcinfo( unlist(strsplit(x, '_'))) ), stringsAsFactors=F))
+pred.qc = merge(preds,qcdata.endo, by=c('PatientID','EndoID'))
 
-pred.qc = cbind.data.frame(preds[-pnr,], qc)
 #dim(pred.qc)
@@
 grid.arrange(
-  ggplot(pred.qc, aes(`mean(SCA.Ratio)`, Prob, color=Status)) + geom_point(),
-  ggplot(pred.qc, aes(`mean(Ploidy)`, Prob, color=Status)) + geom_point(),
+  ggplot(pred.qc, aes(ASCAT.SCA.ratio_mean, Prob, color=Status)) + geom_point(),
+  ggplot(pred.qc, aes(Ploidy_max, Prob, color=Status)) + geom_point(),
 ncol=2)
 
-pander(cor.test(pred.qc$`mean(SCA.Ratio)`, pred.qc$Prob), caption='Probability shows a 0.3 correlation with % of the genome altered')
+pander(cor.test(pred.qc$ASCAT.SCA.ratio_mean, pred.qc$Prob))
 ```
@@
   exclude=(Status == 'NP' & (length(which(Pathology == 'HGD')) > 0)),
-  wgd = (Status == 'NP' & `mean(Ploidy)` > 2.7 )
+  wgd = (Status == 'NP' & Ploidy_max > 2.7 )
 )
@@
 pred.qc = pred.qc %>% group_by(Status, Samplename) %>% dplyr::mutate(
-  lowsca = ( `mean(SCA.Ratio)` < 0.03 & `mean(Purity)` >= 0.95)
+  lowsca = ( ASCAT.SCA.ratio_mean < 0.02 & Purity_mean > 0.95)
 )
 
-ggplot(pred.qc, aes(`mean(SCA.Ratio)`, `mean(Purity)`, color=Status)) + geom_point() + facet_grid(~Status) + 
-  geom_vline(xintercept=0.03) + geom_hline(yintercept = 0.95)
+ggplot(pred.qc, aes(ASCAT.SCA.ratio_mean, Purity_mean, color=lowsca)) + geom_point() + facet_grid(~Status) + 
+  geom_vline(xintercept=0.02) + geom_hline(yintercept = 0.95)
 
 pander(
-  rbind(rbind('n.Samples'=table(subset(pred.qc, !lowsca)$Status)),
+  rbind(rbind('n.Samples'=table(subset(pred.qc, !lowsca )$Status)),
   rbind('n.Patients'=table(unique(subset(pred.qc, !lowsca, select=c('PatientID','Status')))$Status))),
 caption='Number of samples and patients that pass the low SCA:high purity thresholds')
+
+pander(
+  rbind(rbind('n.Samples'=table(subset(pred.qc, !lowsca & !exclude & !wgd)$Status)),
+  rbind('n.Patients'=table(unique(subset(pred.qc, !lowsca  & !exclude & !wgd, select=c('PatientID','Status')))$Status))),
+caption='Number of samples and patients that pass the low SCA:high purity & NP thresholds')
 
@@
 ```{r echo=F}
-pred.qc.ex = subset(pred.qc, !exclude & !wgd & !lowsca)
+pred.qc.ex = subset(pred.qc, SampleType == 'BE' & !exclude & !wgd & !lowsca)
 cuts = seq(0,1,0.1)
@@
 
+norm.pred = subset(pred.qc, SampleType != 'BE')
+
+pred.qc = subset(pred.qc, SampleType == 'BE')
+
 roc = pROC::roc(Status ~ Prob, data=pred.qc, auc=T, ci=T, of='thresholds')
@@
 roc.plot(roc) + labs(title='Excluded HGD NP patients')
-
-# per_pt = 
-#   subset(pred.qc, !exclude) %>% group_by(PatientID, Status) %>% dplyr::summarise( max.Prob = max(Prob), max.RR = max(RR), max.Risk = max(Risk) )
-# rocpp = pROC::roc(Status ~ max.Prob, data=per_pt, auc=T, ci=T, of='thresholds')
-# roc.plot(rocpp) + labs(title='Max per patient - excluded HGD NP')
 
@@
 
-table(subset(pred.qc, Risk == 'High' & !exclude & !lowsca & !wgd)$Status)/table(subset(pred.qc, !exclude & !lowsca & !wgd)$Status)
-table(subset(pred.qc, Risk == 'Moderate' & !exclude & !lowsca & !wgd)$Status)/table(subset(pred.qc, !exclude & !lowsca & !wgd)$Status)
-table(subset(pred.qc, Risk == 'Low' & !exclude & !lowsca & !wgd)$Status)/table(subset(pred.qc, !exclude & !lowsca & !wgd)$Status)
-
 # Timepoint 2 only
@@
 
-ggplot(subset(pred.qc, !exclude & !wgd & !lowsca), aes(`mean(SCA.Ratio)`,Prob, color=Risk, shape=Status)) + 
-  facet_grid(~Status) + geom_point()
-
-ggplot(subset(pred.qc, !exclude & !wgd & !lowsca), aes(`mean(SCA.Ratio)`,Prob, color=Risk, shape=Status)) + 
+ggplot(subset(pred.qc, !exclude & !wgd & !lowsca), aes(ASCAT.SCA.ratio_mean,Prob, color=Risk, shape=Status)) + 
   facet_grid(~Status) + geom_point()
@@
 ex.pred = subset(pred.qc, !exclude & !wgd & !lowsca)
+
+write.table(subset(pred.qc, lowsca, select=c('PatientID','EndoID','Status','ASCAT.SCA.ratio_mean','SCA.Ratio_mean','Purity_mean')),
+            sep='\t', quote=F, row.names=F, file='lowsca_pts.txt')
+
+head(qcdata)
+
+x = subset(qcdata, ASCAT.SCA.ratio < 0.02 & Purity > 0.95)
+table(x$SampleType)/table(qcdata$SampleType)
+
+write.table(subset(x, SampleType == 'BE', select=c('PatientID','EndoID','Samplename','ASCAT.SCA.ratio','SCA.Ratio','Purity')), quote=F, sep='\t', row.names = F, file='lowsca_samples.txt')
+
+#x = subset(qcdata, ASCAT.SCA.ratio < 0.02 & Purity > 0.95 & SampleType == 'BE')
 
@@
 
-ct = cor.test(pred.qc.ex$Prob, pred.qc.ex$`mean(SCA.Ratio)`)
-ct2 = cor.test(pred.qc$Prob, pred.qc$`mean(SCA.Ratio)`)
+ct = cor.test(pred.qc.ex$Prob, pred.qc.ex$ASCAT.SCA.ratio_mean)
+ct2 = cor.test(pred.qc$Prob, pred.qc$ASCAT.SCA.ratio_mean)
 ```
@@
 
-```{r loo, eval=F}
+```{r loo, eval=T}
 

### 2018-09-14 snp/process_snp.R

@@ -4,50 +4,55 @@ library(ggplot2)
 library(gridExtra)
 library(dplyr)
 
-#suppressPackageStartupMessages( source('lib/data_func.R') )
 suppressPackageStartupMessages( source('~/workspace/shallowwgs_pipeline/lib/data_func.R') )
 
 args = commandArgs(trailingOnly=TRUE)
 
-#if (length(args) < 2)
-#  stop("Missing data directory or summary file")
+if (length(args) < 3)
+  stop("Usage: <data directory> <all patient ASCAT Rdata> <Combine by endoscopy T/F> <exlucde low SCA T/F")
 
-if (length(args) < 1)
-  stop("Missing data directory ")
-
+chr.info = get.chr.lengths(file='/tmp/hg19_info.txt')
 
 datadir = args[1]
-#summary.file = args[2]
+# datadir =  "~/Data/Reid_SNP/PerPatient-preLM/512"
+allpts.file = args[2]
+# allpts.file = '~/Data/Reid_SNP/PerPatient/allpts_ascat.Rdata'
+byEndo = as.logical(args[3])
+# byEndo = T
+excludeLowSCA = as.logical(args[4])
+# excludeLowSCA = T
 
-# datadir = '~/Data/Reid_SNP/PerPatient-preLM/512'
+if (!file.exists(allpts.file))
+  stop(paste(allpts.file, "doesn't exist or isn't readable"))
 
-
-byEndo = F
-if (length(args) > 2)
-  byEndo = as.logical(args[3])
+load(allpts.file, verbose=T)
+qcdata = qcdata %>% rowwise %>% dplyr::mutate(
+  PatientID = unlist(strsplit(Samplename, '_'))[1],
+  SampleID = unlist(strsplit(Samplename, '_'))[2],
+  EndoID = unlist(strsplit(Samplename, '_'))[3],
+  Level = unlist(strsplit(Samplename, '_'))[4]
+)
+segments.list = lapply(segments.list, function(pt) {
+  pt$sample = sub('\\.LogR','',pt$sample)
+  pt
+})
+qcdata$ASCAT.SCA.ratio = apply(qcdata,1,function(s) {
+  smp = subset(segments.list[[ s[['PatientID']] ]], sample == s[['Samplename']] & chr %in% c(1:22))
+  smp = smp %>% rowwise %>% dplyr::mutate(
+    #'Total' = nAraw + nBraw,
[…]

### 2018-10-12 AllSamplesModel.Rmd
Using new package

@@
 ggsave('plots/prediction_nohgd.png', plot=hgdloo, width=5, height=5, units='in', dpi=300)
-
-cuts = seq(0,1,0.1) 
-pg.sampNOHGD$quants = with(pg.sampNOHGD, cut(Prediction, breaks=cuts))
-
-noHGD.pred.confidence = as.data.frame.matrix(table(pg.sampNOHGD$quants, pg.sampNOHGD$Status))
-noHGD.pred.confidence = cbind(noHGD.pred.confidence, pg.sampNOHGD %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
-noHGD.pred.confidence$quant = rownames(noHGD.pred.confidence)
-
-ft.fun<-function(NP,P) {
-  f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
-  cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
-}
-
-noHGD.pred.confidence = noHGD.pred.confidence %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )
-
-noHGD.pred.confidence[c('r1','r2')] = NA
-for (i in 1:(length(cuts)-1)) 
-  noHGD.pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)
-
-noHGD.pred.confidence$Risk = 'Moderate'
-noHGD.pred.confidence$Risk[ which(with(noHGD.pred.confidence, p.value < 0.05 & perc < .5)) ] = 'Low'
-noHGD.pred.confidence$Risk[ which(with(noHGD.pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'
-
-noHGD.pred.confidence = bind_cols(noHGD.pred.confidence, with(noHGD.pred.confidence, 
-    data.frame(ci.low=qbeta(0.025, shape1=P+.5, shape2=NP+.5),
-               ci.high=qbeta(0.975, shape1=P+.5, shape2=NP+.5))))
-
-noHGD.pred.confidence$Risk = factor(noHGD.pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)
-
-nohgdcalib = ggplot(noHGD.pred.confidence, aes(mn, perc)) + 
-  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
-  scale_x_continuous(limits=c(0,1), breaks = cuts, labels = c(0.0, '','','','','',0.6,'',0.8,'',1.0)) +
-  geom_point() + 
-  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration', subtitle='HGD excluded model') + plot.theme + theme(legend.position = 'none')
-nohgdcalib
-
-ggsave('plots/calib_nohgd.png', plot=nohgdcalib, width=5, height=5, units='in', dpi=300)
+# pg.sampNOHGD = subset(pg.sampNOHGD, !is.na(Prediction))
+# cuts = seq(0,1,0.1) 
+# pg.sampNOHGD$quants = with(pg.sampNOHGD, cut(Prediction, breaks=cuts))
+# 
+# 
+# noHGD.pred.confidence = as.data.frame.matrix(table(pg.sampNOHGD$quants, pg.sampNOHGD$Status))
+# noHGD.pred.confidence = cbind(noHGD.pred.confidence, pg.sampNOHGD %>% dplyr::group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
+# noHGD.pred.confidence$quant = rownames(noHGD.pred.confidence)
+# 
+# ft.fun<-function(NP,P) {
+#   f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
+#   cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
+# }
+# 
+# noHGD.pred.confidence = noHGD.pred.confidence %>% as.data.frame.matrix %>% dplyr::group_by(quant) %>% dplyr::mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )
+# 
+# noHGD.pred.confidence[c('r1','r2')] = NA
+# for (i in 1:(length(cuts)-1)) 
+#   noHGD.pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)
+# 
+# noHGD.pred.confidence$Risk = 'Moderate'
+# noHGD.pred.confidence$Risk[ which(with(noHGD.pred.confidence, p.value < 0.05 & perc < .5)) ] = 'Low'
+# noHGD.pred.confidence$Risk[ which(with(noHGD.pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'
+# 
+# noHGD.pred.confidence = bind_cols(noHGD.pred.confidence, with(noHGD.pred.confidence, 
+#     data.frame(ci.low=qbeta(0.025, shape1=P+.5, shape2=NP+.5),
+#                ci.high=qbeta(0.975, shape1=P+.5, shape2=NP+.5))))
+# 
+# noHGD.pred.confidence$Risk = factor(noHGD.pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)
+# 
+# nohgdcalib = ggplot(noHGD.pred.confidence, aes(mn, perc)) + 
+#   geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
+#   scale_x_continuous(limits=c(0,1), breaks = cuts, labels = c(0.0, '','','','','',0.6,'',0.8,'',1.0)) +
+#   geom_point() + 
+#   geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration', subtitle='HGD excluded model') + plot.theme + theme(legend.position = 'none')
+# nohgdcalib
+# 
+# ggsave('plots/calib_nohgd.png', plot=nohgdcalib, width=5, height=5, units='in', dpi=300)
 # Predict HGD samples
@@
   hgd.samples = subset(patient.info, Pathology %in% c('HGD','IMC'))
-  hgd.df = dysplasia.df[hgd.samples$Samplename,]
+  hgd.df = dysplasia.df[intersect(hgd.samples$Samplename, rownames(dysplasia.df)),]
 
   pred = predict(fitNOHGD, newx=hgd.df, s=lambda.opt, type='response')
-  hgd.samples$Prediction = pred
-
-  hgd.pred = hgd.samples %>% group_by(Patient, Hospital.Research.ID, Status, Endoscopy.Year) %>% 
-    summarise( max=max(Pathology), pred=Prediction[which.max(Pathology)] )
+
+  hgd.samples[which(hgd.samples$Samplename %in% rownames(dysplasia.df)), 'Prediction'] = pred
+
+  hgd.pred = hgd.samples %>% dplyr::group_by(Patient, Hospital.Research.ID, Status, Endoscopy.Year) %>% 
+    dplyr::summarise( max=max(Pathology), pred=Prediction[which.max(Pathology)] )
 
@@
 #hist(hgd.samples$Prediction)
-
 rm(plots, performance.at.1se,coefs, nzcoefs, fits)
@@
 ```{r echo=F, fig.height=4, fig.width=4}
-#ptendo = do.call(rbind, pg.samp) %>% group_by(Patient, Hospital.Research.ID, Status) %>% summarise('max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(Patient))
+#ptendo = do.call(rbind, pg.samp) %>% group_by(Patient, Hospital.Research.ID, Status) %>% dplyr::summarise('max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(Patient))
 ptendo = do.call(rbind, lapply(pg.samp, function(pt) {
@@
   f = chisq.test(c(NP,P))
-  cbind.data.frame('p.value'=f$p.value)
+  cbind.data.frame('p.value'= signif(f$p.value,4))
   #cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
 }
+
+excluded = subset(back, is.na(Prediction))
+back = subset(back, !is.na(Prediction))
 
@@
 pred.confidence = qt %>% as.data.frame.matrix %>% rowwise() %>% dplyr::mutate( 
-  'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'conf'=ifelse(p.value < 0.05, '*', '') )
+  'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'conf'=ifelse(round(p.value,2) <= 0.05, '*', '') )
 
@@
 pred.confidence$Risk = 'Moderate'
-pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc < .3)) ] = 'Low'
-pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'
+pred.confidence$Risk[ which(with(pred.confidence, round(p.value,2) <= 0.05 & perc < .3)) ] = 'Low'
+pred.confidence$Risk[ which(with(pred.confidence, round(p.value,2) <= 0.05 & perc > .5)) ] = 'High'
 
@@
 
-annotation.table = arrange(pred.confidence %>% group_by(Risk) %>% dplyr::summarise('n(NP)'=sum(NP), 'n(P)'=sum(P), 'p.value'=format.pval(mean(p.value), 3), 'min'=min(r1), 'max'=max(r2)), min)
+annotation.table = dplyr::arrange(pred.confidence %>% dplyr::group_by(Risk) %>% dplyr::summarise('n(NP)'=sum(NP), 'n(P)'=sum(P), 'p.value'=format.pval(mean(p.value), 3), 'min'=min(r1), 'max'=max(r2)), min)
 
@@
 ```{r pcalib, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
-#pred.confidence$Risk[1:5] = 'Low'
-#pred.confidence$Risk[6:10] = 'High'
-#pred.confidence$Risk[4:7] = 'Moderate'
-
-pcal = ggplot(pred.confidence, aes(mn, perc)) + 
+pcfit = lm(perc~mn, data=pred.confidence)
+
+rsid = rstandard(pcfit)
+pred.confidence[which( rsid >= 1 |  rsid <= -1 ), 'Risk'] = 'Moderate'
+
+pcal = ggplot(pred.confidence, aes(mn, perc)) + geom_smooth(method='lm',formula=y~x, color='grey39', linetype='dashed', size=0.5,fullrange=T) + 
   geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols, limits=levels(pred.confidence$Risk) ) +
-  geom_point() + geom_abline(slope=1, intercept=0, color='grey39', linetype='dotted', size=0.5) + #geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
-  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), size=0.5, width=0.01) + labs(x='P(P)', y='NP:P', title='') +
-  scale_color_manual(values=riskCols,limits=levels(pred.confidence$Risk) ) + 
-  scale_x_continuous(limits=c(0,1), breaks = cuts, labels = c(0.0, '','',0.3,'','','',0.7,'','',1.0)) +
+  geom_point() + 
+  # geom_abline(slope=1, intercept=0, color='grey39', linetype='dotted', size=0.5) + #geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
+  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), size=0.5, width=0.01) + labs(x='P(P)', y='P:NP', title='') +
+  scale_color_manual(values=riskCols,limits=levels(pred.confidence$Risk) ) +   coord_cartesian(xlim=c(0,1), ylim=c(0,1)) +
+  scale_x_continuous(expand=c(0,0),limits=c(-0.5,1.5), breaks=cuts, labels=cuts) +  scale_y_continuous(expand=c(0,0),limits=c(-0.5,1.5), breaks=cuts, labels=cuts) +
   plot.theme + theme(legend.position = 'none')
@@
 
-ggplot(back, aes(Status, OR, color=OR)) + 
-  geom_boxplot(outlier.colour=NA) + geom_jitter(width=0.3, aes(shape=Pathology)) + 
-  scale_color_gradientn(colors = myPal,  name='') + plot.theme
-
 predhist = ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) +
@@
 
-ggplot(back, aes(OR, Prediction, color=Prediction)) + geom_point(show.legend = F) + scale_color_gradientn(colors = myPal,  name='')  + plot.theme
+#ggplot(back, aes(OR, Prediction, color=Prediction)) + geom_point(show.legend = F) + scale_color_gradientn(colors = myPal,  name='')  + plot.theme
 ```
@@
 
-```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=20}
+```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=7}
 back$ogj = back$nl
@@
 p = ggplot(subset(df2, early|late), aes(Endoscopy.Year, arms, color=Patient)) + facet_grid(~Status) +
-  geom_point() +  geom_line(data=subset(df2, early | late)) + theme(legend.position = 'none', text=element_text(size=11)) + 
+  geom_point(aes(shape=late), show.legend = F, size=2) + geom_line() + 
+  theme(legend.position = 'none', text=element_text(size=11)) + 
   plot.theme + labs(x='Endoscopy Year', y='CN arm values', title='Initial vs final endoscopy', subtitle=paste('p', wtA))
@@
 ggsave(filename='plots/hetero_initial.png', plot=p, width=6, height=4, units='in', dpi=300)
-
+```
+
+```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=20}
 p2 = mtn.spatial(subset(globalCNdf, Patient == 20 & seg.type != 'arms'), back, ogjLimits,'' )
@@
 get.w.scale <- function(b) {
-  n = sum((b %>% group_by(Patient) %>% summarise(unique(n.endos)))[,2])
+  n = sum((b %>% group_by(Patient) %>% dplyr::summarise(unique(n.endos)))[,2])
   #round(n*0.5)+1
@@
 })
-sort(x)
+#sort(x)
 
@@
   b = subset(back, Patient == i)
+  if (nrow(b) <= 0) next
   lim = c(1:7)
@@
   b = subset(back, Patient == i)
+  if (nrow(b) <= 0) next
+
   lim = c(1:9)
@@
 ```{r echo=F, eval=T}
-dirs = c('~/Data/Ellie/Analysis/downsampled5G_BEAdjacent',
-         '~/Data/Ellie/Analysis/downsampled5G_BEAdjacent_normal/') #,'~/Data/Ellie/Analysis/downsampled5G_NDBE')
 
@@
 
-predfiles = sapply(dirs, function(x) {
-  list.files(x, 'predictions_all_exAHM0320.Rdata', full.names = T)
-})
-
-predObj = list()
-for (i in 1:length(predfiles)) {
-  if (length(predfiles[[i]]) > 0) {
-      load(predfiles[[i]], verbose=T)
-      dspred$Type = basename(names(predfiles[i]))
-      predObj[[i]] = dspred
-      rm(dspred)
-  } else {
-    files = list.files(names(predfiles[i]), pattern='raw', full.names=T)
-    sampleNames = sub('_raw.txt', '', basename(files))
-    dspred = data.frame(matrix(ncol=2, nrow=length(sampleNames), dimnames=list(sampleNames,c('Prob','RR'))))
-    for (n in 1:length(files)) {
-      print(n)
-      print(sampleNames[n])
-
-      rs = read.table(files[n], sep='\t',header=T)
-      if (ncol(rs) < 4)
-        stop("File needs to contain 4 columns: chr, start, end, sample_value")
-
-      segs <- tile.segmented.data(rs, size=5e6, chr.info=chr.lengths)
-      segM = as.matrix(segs[,-c(1:3)])
-      rownames(segM) = paste(segs$chr, ':', segs$start, '-', segs$end, sep='')
-      segM = t(segM)
-      for (j in 1:ncol(segM)) 
-        segM[,j] = unit.var(segM[,j], z.mean[j], z.sd[j])
-
-      arms <- tile.segmented.data(rs, size='arms', chr.info=chr.lengths)
-      armsM = as.matrix(arms[,-c(1:3)])
-      rownames(armsM) = paste(arms$chr, ':', arms$start, '-', arms$end, sep='')
-      armsM = t(armsM)
-      for (j in 1:ncol(armsM)) 
-        armsM[,j] = unit.var(armsM[,j], z.arms.mean[j], z.arms.sd[j])
-
-      nrow(armsM) == nrow(segM)
-
-      cx.score = score.cx(segM,1)
-
-      mergedDf = subtract.arms(segM, armsM)
-      mergedDf = cbind(mergedDf, 'cx' = unit.var(cx.score, mn.cx, sd.cx))
-      dim(mergedDf)
-
-      # Predict function giving me difficulty when I have only a single sample, this ensures the dimensions are the same
-      sparsed_test_data <- Matrix(data=0, nrow=nrow(mergedDf),  ncol=ncol(mergedDf),
-                                  dimnames=list(rownames(mergedDf),colnames(mergedDf)), sparse=T)
-      for(col in colnames(mergedDf)) sparsed_test_data[,col] = mergedDf[,col]
-
-      preds = predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='response')
-      RR = predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='link')
-
-      dspred[n,] = c(preds[,1],RR[,1])
-      print(dspred[1:n,])
-  }
-
-    dspred$Type = basename(names(predfiles[i]))
-    predObj[[i]] = dspred
-
-    save(dspred, file=paste(names(predfiles[i]), 'predictions_all_exAHM0320.Rdata', sep='/'))
-    rm(dspred)
-  }
-
-  dspred = do.call(rbind, predObj)
-}
+down.dir = '~/Data/Ellie/Analysis/downsampled'
+dsids = read_xlsx(paste(data,'Analysis/downsampled_ids.xlsx',sep='/'))
+load(list.files(down.dir, 'Rdata', full.names = T), verbose=T)
+
+dspred = merge(dspred, dsids, by.x='row.names',by.y='Illumina ID')
+rownames(dspred) = dspred$Row.names
+dspred$Row.names = NULL
+head(dspred)
 
@@
 
-dspred$Type[grep('normal', dspred$Type)] = 'Normal blood'
-dspred = subset(dspred, !grepl('normal', Type))
-dspred$Type[grep('BEAdjacent', dspred$Type)] = 'Tumor adjacent BE'
-
 prg = back[,c('Prediction', 'OR', 'Status')]
@@
 #  subset(back, Status == 'P', select=c('Prediction', 'OR', 'Status'))
-colnames(prg) = colnames(dspred)
-dspred = rbind(dspred,prg)
-dspred$Type = factor(dspred$Type, levels=c('Normal blood','Non-Progressor','Progressor','Tumor adjacent BE'), ordered = T)
-
-colors = myPal[seq(2, length(myPal), round(length(myPal)/4))]
+colnames(prg) = colnames(dspred)[c(1:2,ncol(dspred))]
+
+dspred = rbind(dspred[colnames(prg)],prg)
+
+dspred= subset(dspred, Type != 'LGD') # These are pre-progressors most likely
+
+dspred$Type = factor(dspred$Type, levels=c('Normal', 'NDBE','Non-Progressor','Progressor','Tumor Adjacent BE'), ordered = T)
+
+colors = myPal[seq(2, length(myPal), round(length(myPal)/length(levels(dspred$Type))))]
 pr = ggplot(dspred, aes(Type, RR, fill=Type)) + 
@@
 
-m = subset(dspred, Type != 'Normal blood')
+m = subset(dspred, !Type %in% c('Normal','LGD','NDBE'))
 m$Type = droplevels(m$Type)
+
+colors = myPal[seq(2, length(myPal), round(length(myPal)/length(levels(m$Type))))]
 pr = ggplot(m, aes(Type, RR, fill=Type)) + 
   geom_jitter(width=0.2, aes(color=RR)) + 
-  scale_color_gradientn(colors = myPal,  name='') + 
+  scale_color_gradientn(colors = colors,  name='') + 
   geom_boxplot(outlier.colour = NA) + 
@@
 predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='link')
+predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='response')
 
@@
 
+# Based on the penult pair
 penult = subset(rules, t2 != 'Endpoint' & (t1 == rev(levels(back$brks))[2] | t2 == rev(levels(back$brks))[2] ))
-
-# Based on the penult pair
 round(table(subset(penult, Status == 'P')$rule)/nrow(subset(penult, Status == 'P')), 2)
@@
 
-rrPath = back %>% group_by(Status, Pathology) %>% summarise( 'mean.RR'=mean(OR),'ci.min'=mean(OR)-(1.96*sd(OR)), 'ci.max'=mean(OR)+(1.96*sd(OR)) )
+rrPath = back %>% group_by(Status, Pathology) %>% dplyr::summarise( 'mean.RR'=mean(OR),'ci.min'=mean(OR)-(1.96*sd(OR)), 'ci.max'=mean(OR)+(1.96*sd(OR)) )
 
@@
 
-risksL = subset(bd, Status == 'NP') %>% group_by(Patient) %>% summarise( 'n.samp'=length(Patient), 'high.risk'=length(which(Risk == 'High')), 'mod.risk'=length(which(Risk == 'Moderate')) )
+risksL = subset(bd, Status == 'NP') %>% group_by(Patient) %>% dplyr::summarise( 'n.samp'=length(Patient), 'high.risk'=length(which(Risk == 'High')), 'mod.risk'=length(which(Risk == 'Moderate')) )
 
@@
 for (yr in names(which(table(bd$Endoscopy.Year) > 10))) {
-  risks = subset(bd, Status == 'NP' & Endoscopy.Year == yr) %>% group_by(Patient) %>% summarise( 'n.samp'=length(Patient), 'high.risk'=length(which(Risk == 'High')), 'mod.risk'=length(which(Risk == 'Moderate')) )
+  risks = subset(bd, Status == 'NP' & Endoscopy.Year == yr) %>% group_by(Patient) %>% dplyr::summarise( 'n.samp'=length(Patient), 'high.risk'=length(which(Risk == 'High')), 'mod.risk'=length(which(Risk == 'Moderate')) )
 

### 2018-10-12 sWGS Normals.Rmd
Using new package

---
title: "sWGS Normals"
author: "Sarah Killcoyne"
date: "7/30/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(Matrix)
library(glmnet)
library(ggplot2)
library(gridExtra)
library(pander)
source('lib/load_patient_metadata.R')
source('lib/data_func.R')
source('lib/fastPCF.R')

data.files = list.files('~/Data/Ellie/QDNAseq', full.names = T)

rpi = read.patient.info('~/Data/Ellie/QDNAseq/All_patient_info.xlsx')
normals = rpi$removed
path = read_excel('~/Data/Ellie/QDNAseq/All_patient_info.xlsx', 'Normals')[,c('Patient ID', 'Hospital Research ID', 'Pathology notes','Plate Index','SLX ID 1')]
colnames(path)[c(1,4,5)] = c('Patient','Plate.Index','SLX.ID')
path$SLX.ID = sub('SLX-','', gsub(' ', '', path$SLX.ID))
normals = merge(normals, path, by=c('Patient','Plate.Index','SLX.ID'))

data.file = grep('MultisampleCopynumberBinnedAndFitted2.RData', data.files, value=T)
if (length(data.file) <= 0)
  stop("Missing necessary Rdata file from Multisample script")
load(file=data.file, verbose=T)

loc.cols = c('location','chrom','start','end')
fit.data = fit.data[,c(loc.cols,normals$Samplename)]
raw.data = raw.data[,c(loc.cols,normals$Samplename)]

exclude.file = grep('qDNAseq_blacklistedRegions.txt', data.files, value=T)
if (length(data.file) <= 0)
  stop(paste("Missing necessary exclusion file from 'qDNAseq_blacklistedRegions.txt' in",data))
blacklisted.regions = read.table(exclude.file,sep="\t",header=T,stringsAsFactors=F)

if (file.exists('/tmp/normals.Rdata')) {
  load('/tmp/normals.Rdata', verbose=T)
} else {
  plist = list()
  seg.residuals = list()
  tiled.normals = NULL; tiled.normal.arms = NULL
  for (sample in normals$Samplename) {
    binned = binSWGS(raw.data[,c(loc.cols,sample)],fit.data[,c(loc.cols,sample)],blacklisted.regions,min.probes=67,gamma2=250,metrics.file=NULL)
    plist[[sample]] = binned$plots[[1]]
    seg.residuals[[sample]] = binned$resvar[[sample]]

    tiled = tile.segmented.data(binned$segvals,chr.info=get.chr.lengths(file='/tmp/hg19_info.txt'),verbose=T)
    tiled.arms = tile.segmented.data(binned$segvals,chr.info=get.chr.lengths(file='/tmp/hg19_info.txt'),size='arms', verbose=T)
    if (is.null(tiled.normals) | is.null(tiled.normal.arms)) {
      tiled.normals = tiled
      tiled.normal.arms = tiled.arms
    } else {
      tiled.normals = cbind(tiled.normals, tiled[,sample,drop=F])
      tiled.normal.arms = cbind(tiled.normal.arms, tiled.arms[,sample,drop=F])
    }
  }
  save(tiled.normals, tiled.normal.arms, plist, seg.residuals, file='/tmp/normals.Rdata')
}

resids = sample.residual.variance(seg.residuals)
resids = merge(normals[,c('Patient','Hospital.Research.ID','Samplename')], resids, by.x='Samplename', by.y='sample')

failedQC = arrange(subset(resids, round(var1sd_median,3) > 0.012, select=c(Hospital.Research.ID,Samplename, var1sd_median,var1sd_sd, var1sd_Q1,var1sd_Q3)),  var1sd_median)

pander(normals[,c('Patient','Hospital.Research.ID','Status','Endoscopy.Year','Pathology notes')], caption='Normal samples', justify='left')
```

`r nrow(failedQC)` samples failed the variant residual QC based on a cutoff determined by the discovery cohort.

```{r, echo=F, warning=F, fig.height=30, fig.width=8}
do.call(grid.arrange, c(plist, ncol=1))
```

## Predictions

Using the model trained on the progressors and non-progressors in the initial cohort the normal samples can be predicted as well.

```{r, echo=F}
modeldir = '~/Data/Ellie/Analysis/5e6_arms_all'
file = paste(modeldir, 'model_data.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))
load(file, verbose=F)
file = paste(modeldir, 'all.pt.alpha.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))
load(file, verbose=F)
rm(dysplasia.df,coefs,plots,labels)

select.alpha = '0.9'
fitV = models[[select.alpha]]
lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']

tiled.segs = segment.matrix(tiled.normals)
tiled.arms = segment.matrix(tiled.normal.arms)

cx.score = score.cx(tiled.segs,1)

for (i in 1:ncol(tiled.segs)) 
  tiled.segs[,i] = unit.var(tiled.segs[,i], z.mean[i], z.sd[i])
for (i in 1:ncol(tiled.arms)) 
  tiled.arms[,i] = unit.var(tiled.arms[,i], z.arms.mean[i], z.arms.sd[i])

mergedDf = subtract.arms(tiled.segs, tiled.arms)
mergedDf = cbind(mergedDf, 'cx' = unit.var(cx.score, mn.cx, sd.cx))

sparsed_test_data <- Matrix(data=0, nrow=nrow(mergedDf),  ncol=ncol(mergedDf),
                            dimnames=list(rownames(mergedDf),colnames(mergedDf)), sparse=T)
for(col in colnames(mergedDf)) sparsed_test_data[,col] = mergedDf[,col]

preds = predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='response')

high = preds[preds > 0.7,,drop=F]

pander(subset(normals, Samplename %in% rownames(high), select=c('Patient','Hospital.Research.ID','Status','Endoscopy.Year','Pathology notes')), caption='Normal samples predicted to be high risk')

```

### 2018-10-12 sWGS QC.Rmd
Using new package

---
title: "sWGS QC"
author: "Sarah Killcoyne"
date: "8/30/2018"
output: 
  html_document: 
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(pander)
library(dplyr)
library(ggplot2)
library(gridExtra)

dir1 = '~/Data/Ellie/Analysis/multipcf_plots_fitted_perPatient'
dir2 = '~/Data/Ellie/Analysis/VAL_Cohort/sWGS/'

```

## Initial sWGS dataset (Ellie's)

```{r, echo=F}
res.variance = read.table(paste(dir1,'resid_var.txt',sep='/'),header=T)
#pander(head(res.variance[,grep('var1sd',colnames(res.variance))]), caption='Residuals across segments per sample')
pander(head(res.variance[,grep('varMAD',colnames(res.variance))]), caption='Residuals across segments per sample')
```

```{r, echo=F, fig.height=10}
res.variance = arrange(res.variance,varMAD_median)
cutoff = round(median(res.variance$varMAD_median)+sd(res.variance$varMAD_median)*2, 3)
#cutoff = 0.015

grid.arrange(
  ggplot(res.variance,aes(sample=varMAD_median)) + stat_qq() + geom_hline(yintercept=cutoff, linetype='dashed',color='grey') + 
    labs(title='Normal Q-Q plot (all)', x='Theoretical quantiles',y='Sample quantiles') + theme_bw(),
  ggplot(subset(res.variance,varMAD_median < cutoff), aes(sample=varMAD_median)) + stat_qq() + 
    labs(title='Normal Q-Q plot (samples<cutoff)', x='Theoretical quantiles',y='Sample quantiles') + theme_bw(),
  top='var( seg<MAD(seg) | seg>-MAD(seg)  )'
)
```

```{r, echo=F}
ggplot(res.variance, aes(varMAD_median)) + geom_histogram(bins=50,color='lightblue') + 
  geom_vline(xintercept = cutoff, linetype='dashed', color='grey') + labs(x='var( seg<MAD(seg) | seg>-MAD(seg)  )') +
  theme_bw()
```

The cutoff=`r cutoff`. Samples with a variance greater than this are:

```{r, echo=F}
pander( arrange(subset(res.variance, round(varMAD_median,3) > cutoff, select=c(pid,sample, varMAD_median,varMAD_sd, varMAD_Q1,varMAD_Q3)), pid, varMAD_median) )

failedQC = arrange(subset(res.variance, round(varMAD_median,3) > cutoff, select=c(pid,sample, varMAD_median,varMAD_sd, varMAD_Q1,varMAD_Q3)), pid, varMAD_median)

```

`r nrow(failedQC)` samples (`r signif(nrow(failedQC)/nrow(res.variance)*100,2)`%) would have failed due to high variance across all the segments. A manual check of these agrees with most of these. One or two samples would appear to be driven by some extreme segments, and this could be further evaluated, and stricter cutoffs do pick up additional samples that are borderline.  However, this cutoff appears to be reasonable.

## Validation dataset

Cutoff from above dataset applied to the validation set.

```{r, echo=F}
resVal = read.table(paste(dir2,'resid_var.txt',sep='/'),header=T)
pander(head(resVal[,grep('var1',colnames(resVal))]), caption='Residuals across segments per sample')

resVal = arrange(resVal,varMAD_median)

ggplot(resVal,aes(sample=varMAD_median)) + stat_qq() + geom_hline(yintercept=cutoff, linetype='dashed',color='grey') + 
  labs(title='Normal Q-Q plot (all)', x='Theoretical quantiles',y='Sample quantiles') + theme_bw()

ggplot(resVal, aes(varMAD_median)) + geom_histogram(bins=50,color='lightblue') + 
  geom_vline(xintercept = cutoff, linetype='dashed', color='grey') + labs(x='var(median(segments)+1sd(segments))') +
  theme_bw()

failedQCVal = arrange(subset(resVal, round(varMAD_median,3) > cutoff, select=c(pid,sample, varMAD_median,varMAD_sd, varMAD_Q1,varMAD_Q3)), pid, varMAD_median)

setdiff(as.character(resVal$sample),as.character(failedQCVal$sample))

```

The result from this is that `r nrow(failedQCVal)` samples (`r signif(nrow(failedQCVal)/nrow(resVal)*100,3)`%) would have failed due to high variance across all the segments. A manual check of these agrees with all of them.

## If we learned it from all residuals

```{r}
res.variance[c('pid','sample')] = lapply(res.variance[c('pid','sample')], as.character)
resVal[c('pid','sample')] = lapply(resVal[c('pid','sample')], as.character)

res.variance$cohort = 'discovery'
resVal$cohort = 'validation'

ar = rbind(res.variance, resVal)
ar$cohort = factor(ar$cohort)

#newcutoff = round(median(ar$varMAD_median)+sd(ar$varMAD_median)*2, 3)
newcutoff = 0.015

ggplot(ar,aes(sample=varMAD_median)) + stat_qq() + geom_hline(yintercept=newcutoff, linetype='dashed',color='grey') + 
    labs(title='Normal Q-Q plot (all)', x='Theoretical quantiles',y='Sample quantiles') + theme_bw()

ggplot(ar, aes(varMAD_median)) + geom_histogram(bins=50,color='lightblue') + 
  geom_vline(xintercept = newcutoff, linetype='dashed', color='grey') + labs(x='var(MAD)') +
  theme_bw()

fc = arrange(subset(ar, round(varMAD_median,3) > newcutoff, select=c(pid,sample, varMAD_median,varMAD_sd, varMAD_Q1,varMAD_Q3)), pid, varMAD_median)

setdiff(fc$sample,c(as.character(failedQC$sample), as.character(failedQCVal$sample)))

```

### 2018-10-12 snp/SNP_Prediction_Analysis.Rmd
Using new package

@@
 
-ggplot(subset(qcdata,SampleType == 'BE'), aes(Purity, ASCAT.SCA.ratio, color=Status, group=Status)) + geom_point() + facet_grid(~Status, scales = 'free') + ylim(0,0.005) + 
-  geom_hline(yintercept = x) + geom_text(data=anno, aes(x, y, label=label), nudge_y = 0.0001) +
+ggplot(subset(qcdata,SampleType == 'BE'), aes(Purity, ASCAT.SCA.ratio, color=Status, group=Status)) + geom_point() + facet_grid(~Status, scales = 'free') + ylim(0,0.01) + 
+  geom_hline(yintercept = anno$y[1]) + geom_text(data=anno, aes(x, y, label=label), nudge_y = 0.0001) +
   geom_text(data=anno, aes(x, y, label=paste(round(Ratio,2)*100,'% samples',sep='')), nudge_y = -0.0001)
@@
 
-table(unique(lowsca[,c('PatientID','Status')])$Status)/table(unique(qcdata.endo[,c('PatientID','Status')])$Status)
+#table(unique(lowsca[,c('PatientID','Status')])$Status)/table(unique(qcdata.endo[,c('PatientID','Status')])$Status)
 
-table(unique(subset(qcdata.endo, SampleType == 'BE' & ASCAT.SCA.ratio_max <= 0.001, select=c('PatientID','Status')))$Status)
+#table(unique(subset(qcdata.endo, SampleType == 'BE' & ASCAT.SCA.ratio_max <= 0.001, select=c('PatientID','Status')))$Status)
 
@@
 ## ----- All ----- ##
-cache.dir = '~/Data/Ellie/Analysis/SNP'
+cache.dir = '~/Data/Ellie/Analysis/SNP/no-low/'
 file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')

### 2018-10-12 SegmentNormalSamples.R

@@ -1,0 +1,184 @@
+#!/usr/bin/env Rscript
+
+# Step 2
+args = commandArgs(trailingOnly=TRUE)
+
+if (length(args) < 2) {
+  stop("Data directory and plot output directory required as arguments")
+}
+
+suppressPackageStartupMessages( library(ggplot2) )
+suppressPackageStartupMessages( library(gridExtra) )
+suppressPackageStartupMessages( library(ggdendro) )
+suppressPackageStartupMessages( library(copynumber) )
+suppressPackageStartupMessages( library(dplyr) )
+
+suppressPackageStartupMessages( source("~/workspace/shallowWGSpipeline/lib/fastPCF.R") )
+suppressPackageStartupMessages( source('~/workspace/shallowWGSpipeline/lib/load_patient_metadata.R') )
+suppressPackageStartupMessages( source('~/workspace/shallowWGSpipeline/lib/data_func.R') )
+
+
+data = args[1]
+#patient.name = args[2]
+outdir = args[2]
+
+# data = '~/Data/Ellie/QDNAseq'
+# outdir = '~/Data/Ellie/Analysis/Normals'
+data.dirs = list.dirs(data, full.names=T, recursive=F)
+data.files = list.files(data, full.names=T, recursive=T)
+
+plot.dir = paste(outdir, 'multipcf_plots_normals', sep='/')
+print(paste("Plot directory:", plot.dir))
+if (!dir.exists(plot.dir))  
+  dir.create(plot.dir, recursive=T)
+
+## Patient info file
+filename = 'All_patient_info.xlsx'
+patient.file = grep(filename, list.files(data, full.names=T), value=T)
+if (length(patient.file) != 1)
+  stop(paste("Missing patient info file in", data))
+message(paste("Reading file", patient.file))
+
+normals = read.patient.info(patient.file, set='all')$removed
+
+head(normals)
+
+## Data file
+data.file = grep('MultisampleCopynumberBinnedAndFitted2.RData', data.files, value=T)
+if (length(data.file) <= 0)
+  stop("Missing necessary Rdata file from Multisample script")
+load(file=data.file, verbose=T)
+
+fit.data = fit.data[,c('location','chrom','start','end',normals$Samplename)]
+raw.data = raw.data[,c('location','chrom','start','end',normals$Samplename)]
+window.depths = window.depths[,c(normals$Samplename)]
+
[…]

### 2018-10-12 processing_pipeline/process_training_data.R

@@ -1,0 +1,166 @@
+## load all of Ellie's data, merge, then break into per-patient raw/fitted files. Run segmentation and save results.
+
+library(plyr)
+library(ggplot2)
+library(ggdendro)
+library(BarrettsProgressionRisk)
+# this version removes qDNAseq 'blacklisted' regions and uses 'fitted' read counts
+
+#args = commandArgs(trailingOnly=TRUE)
+
+#if (length(args) < 3)
+#  stop("Missing required arguments: <qdna data dir> <patient spreadsheet> <output dir>")
+
+source('lib/load_patient_metadata.R')
+
+#data = args[1]
+#patient.file = args[2]
+#outdir = args[3]
+data = '~/Data/Ellie/QDNAseq'
+patient.file = paste(data, 'All_patient_info.xlsx', sep='/')
+outdir = '~/Data/Ellie/Analysis'
+
+patient.null = NULL
+if (length(args) == 4)
+  patient.name = args[4]
+
+
+all.patient.info = read.patient.info(patient.file, set='All')
+if (is.null(all.patient.info))
+  stop(paste("Failed to read patient file", patient.file))
+
+data.dirs = list.dirs(data, full.names=T, recursive=F)
+if (length(data.dirs) <= 0)
+  stop(paste("No directories in", data))
+
+pts_slx = arrange(unique(all.patient.info$info[c('SLX.ID','Hospital.Research.ID')]), Hospital.Research.ID)
+
+data.dirs = grep(paste(unique(pts_slx$SLX.ID), collapse='|'), data.dirs, value=T)
+
+datafile = paste(data,"MultisampleCopynumberBinnedAndFitted1.RData", sep='/')
+if (!file.exists(datafile)) {
+  # Load the data, binding each new data file into single data table
+  fit.data = NULL
+  raw.data = NULL
+  samplenames=NULL
+  for(dir in data.dirs) {
+    print(paste("Reading directory", dir))
+    files = list.files(dir,pattern="[fitted|corrected]ReadCounts.txt", full.names=T)
+    slx = sub('SLX-','', basename(dir))
+    
+    # Files that do not contain individual indecies
+    if (length(files) == 1 || !grepl('D\\d+_D\\d+', files)) {
+      for (file in files) {
+        print(paste("Reading file", file))
+        spl = strsplit(basename(file),"\\.")
+        
+        fitted.file = file
[…]

### 2018-10-12 processing_pipeline/run-cv2.R

@@ -3,20 +3,21 @@ args = commandArgs(trailingOnly=TRUE)
 if (length(args) < 3)
   stop("Missing required params: <data dir> <outdir> <info file dir> <log transform DEF: FALSE>")
 
-
-suppressPackageStartupMessages(library(ggplot2))
-suppressPackageStartupMessages(library(plyr))
-suppressPackageStartupMessages(library(dplyr))
-suppressPackageStartupMessages(library(gridExtra))
+# 
+# suppressPackageStartupMessages(library(ggplot2))
+# suppressPackageStartupMessages(library(plyr))
+# suppressPackageStartupMessages(library(dplyr))
+# suppressPackageStartupMessages(library(gridExtra))
 suppressPackageStartupMessages(library(glmnet))
 suppressPackageStartupMessages(library(mice))
-suppressPackageStartupMessages(library(reshape2))
-suppressPackageStartupMessages(library(GenomicRanges))
+# suppressPackageStartupMessages(library(reshape2))
+# suppressPackageStartupMessages(library(GenomicRanges))
 
+library(BarrettsProgressionRisk)
 
 suppressPackageStartupMessages(source('~/workspace/shallowwgs_pipeline/lib/load_patient_metadata.R'))
 suppressPackageStartupMessages(source('~/workspace/shallowwgs_pipeline/lib/cv-pt-glm.R'))
-suppressPackageStartupMessages(source('~/workspace/shallowwgs_pipeline/lib/data_func.R'))
+#suppressPackageStartupMessages(source('~/workspace/shallowwgs_pipeline/lib/data_func.R'))
 
 data = args[1]
 # data = '~/Data/Ellie/Cleaned'
@@ -45,13 +46,16 @@ patient.info = plyr::arrange(patient.info, Status, Hospital.Research.ID, Endosco
 sum.patient.data = summarise.patient.info(patient.info)
 #nrow(sum.patient.data)
 
-cleaned = list.files(path=data, pattern='tiled', full.names=T, recursive=T)
-raw = list.files(path=data, pattern='raw', full.names=T, recursive=T)
+files = list.files(path=data, pattern='^probefiltered.*seg', recursive=T, full.names=T)
+files = grep(paste(sum.patient.data$Hospital.Research.ID, collapse = '|'), files, value=T)
 
[…]

### 2018-10-12 processing_pipeline/tile_sample_data.R

@@ -1,150 +1,0 @@
-#!/usr/bin/env Rscript
-args = commandArgs(trailingOnly=TRUE)
-
-#args = c()
-#args[1] = '~/Data/Ellie/Analysis/All_patients.Rdata'
-#args[2] = '~/Data/Ellie/Analysis/tile_patients_All'
-#args[3] = 'arms'
-
-if (length(args) < 3) {
-  stop("Rdata file contining patient data, output filename, and tile size (def 1e7) required")
-}
-
-print(args)
-
-suppressPackageStartupMessages( library(GenomicRanges) )
-#suppressPackageStartupMessages( library(plyr) )
-suppressPackageStartupMessages( library(dplyr) )
-
-source('lib/load_patient_metadata.R')
-
-load(args[1], verbose=T)
-
-if (!exists('patient.data') | is.null(patient.data) | length(patient.data) <= 0)
-  stop("Patient data file containing no information")
-
-tileFile = args[2]
-tile.w = as.numeric(args[3])
-if (is.na(tile.w))
-  tile.w = args[3]
-
-tileFile = paste(tileFile, '_', as.character(tile.w), '.Rdata', sep='')
-message(paste('Writing data to', tileFile))
-
-apclust.data<-function(segdata, samples) {
-  samples = intersect(colnames(segdata), samples)
-  x1 = segdata[,samples]
-  
-  if (length(samples) <= 1) { 
-    x1 = as.data.frame(x1)
-    colnames(x1) = samples
-  } else {
-    x1 = x1[, samples]
-  }
-  rownames(x1) = (segdata[,c(1:4)] %>%
-                    rowwise() %>%
-                    mutate(location=paste(paste(chrom, arm, sep=''), '.', start.pos, '-', end.pos, sep='')))$location
-  return(x1)  
-}
-
-tile.sample.segments<-function(sample.data) {
-  message("Tiling patient data...")
-  
-  grList = lapply(sample.data, function(df) {
-    print(unique(df$info$Hospital.Research.ID))
-    df$info = plyr::arrange(df$info, Endoscopy.Year, Pathology)
-    
-    x1 = apclust.data(df$seg.vals, df$info$Samplename)
-    
-    segnames = as.data.frame(do.call(rbind, sapply(rownames(x1), strsplit, '[p|q].|-')))
-    colnames(segnames) = c('chr','start','end')
[…]

### 2018-10-12 snp-loo.R

@@ -1,96 +1,0 @@
-
-
-args = commandArgs(trailingOnly=TRUE)
-
-if (length(args) < 2)
-  stop("Missing parameters: <patient ID> <model dir> <patients dir>")
-
-library(ggplot2)
-library(gridExtra)
-library(glmnet)
-
-suppressPackageStartupMessages(source('~/workspace/shallowwgs_pipeline/lib/data_func.R'))
-suppressPackageStartupMessages(source('~/workspace/shallowwgs_pipeline/lib/common_plots.R'))
-suppressPackageStartupMessages(source('~/workspace/shallowwgs_pipeline/lib/cv-pt-glm.R'))
-
-
-
-pt = args[1]
-#pt = '512'
-
-cache.dir = args[2]
-#cache.dir = '~/Data/Ellie/Analysis/SNP'
-
-
-dir.create(cache.dir, recursive = T, showWarnings = F)
-pt.dir = paste(cache.dir, 'LOO', pt, sep='/')
-dir.create(pt.dir, recursive = T, showWarnings = F)
-
-file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
-if (!file.exists(file))
-  stop("Missing full model file, run snp-regression.R first")
-load(file, verbose = T)
-
-nl = 1000;folds = 10; splits = 5 
-sets = create.patient.sets(slabels, folds, splits, 0.15)  
-
-slabels$Prediction = NA
-slabels$RR = NA
-
-colnames(sets)[2] = 'Samplename'
-
-file = paste(pt.dir, 'loo.Rdata', sep='/')
-
-select.alpha = 0.9
-secf = coefs[[select.alpha]]
-a = select.alpha
-
-performance.at.1se = c(); plots = list(); fit = c(); nzcoefs = c() 
-preds = matrix(ncol=2, nrow=length(subset(slabels, PatientID == pt)[,2]), dimnames=list(subset(slabels, PatientID == pt)[,2],c('Prob','RR')))
-# Remove each patient (LOO)
-#for (pt in unique(slabels$PatientID)) {
-print(pt)
-samples = subset(slabels, PatientID != pt)[,2]
-
-train.rows = which(rownames(df) %in% samples)
-training = df[train.rows,]
-test = as.matrix(df[-train.rows,])
-if ( nrow(test) == ncol(df) ) test = t(test)
-
-# Predict function giving me difficulty when I have only a single sample, this ensures the dimensions are the same
[…]

### 2018-10-12 snp-regression.R

@@ -1,311 +1,0 @@
-library(ggplot2)
-library(gridExtra)
-library(glmnet)
-library(readxl)
-library(pROC)
-library(ggfortify)
-library(preprocessCore)
-
-suppressPackageStartupMessages(source('lib/data_func.R'))
-suppressPackageStartupMessages(source('lib/common_plots.R'))
-source('lib/cv-pt-glm.R')
-
-adjust.cols<-function(mtx, means=NULL, sds=NULL, na.replace=0) {
-  if (!is.null(means)) {
-    if (ncol(mtx) != length(means))
-      stop("Vector of means needs to match the columns in the matrix")
-    for (i in 1:ncol(mtx)) 
-      mtx[,i] = unit.var(mtx[,i], means[i], sds[i])
-  } else {
-    for (i in 1:ncol(mtx)) 
-      mtx[,i] = unit.var(mtx[,i])
-  }
-  mtx[is.na(mtx)] = na.replace
-  
-  return(mtx)
-}
-
-
-
-chr.info = get.chr.lengths(file='hg19_info.txt')
-
-load('~/Data/Ellie/Analysis/5e6_arms_all/model_data.Rdata', verbose = T)
-rm(labels, dysplasia.df,mn.cx,sd.cx,z.mean,z.sd,z.arms.mean,z.arms.sd)
-
-load('~/Data/Reid_SNP/PerPatient/allpts_ascat.Rdata', verbose=T)
-
-patient.info = as.data.frame(read_xlsx('~/Data/Ellie/Analysis/SNP/metadata_T1T2.xlsx'))
-patient.info$UniqueSampleID = paste(patient.info$PatientID, patient.info$`Timepoint Code`, sep='_')
-patient.info$Path.Status = patient.info$Status
-patient.info[patient.info$PatientID %in% subset(patient.info, Pathology %in% c('IMC','HGD'), select='PatientID')[,1], 'Path.Status'] = 'P'
-
-sample.list = read_xlsx('~/Data/Ellie/Analysis/SNP/20180604_Reid_1M_SampleList.xlsx', sheet = 2)
-nrow(sample.list)
-colnames(sample.list)[3] = 'Total.SCA'
-
-sample.list$SCA.Ratio = sample.list$Total.SCA/max(chr.info$genome.length)
-
-sample.info  = do.call(rbind.data.frame, sapply(qcdata$Samplename, strsplit, '_'))
-colnames(sample.info) = c('PatientID','SampleID','EndoID','Level')
[…]

### 2018-10-12 snp/snp-regression.R

@@ -96,7 +96,9 @@ info = qcdata.samples %>% group_by(Status, PatientID, UniqueSampleID) %>% dplyr:
 table(unique(info[,c('PatientID','Status')])$Status)
 table(unique(subset(info, !exclude & !wgd & !lowsca, select=c('PatientID','Status')))$Status)
 
-info = subset(info, !exclude & !wgd & !lowsca & SampleType == 'BE')
+#info = subset(info, !exclude & !wgd & !lowsca & SampleType == 'BE')
+#info = subset(info, !exclude & !wgd & SampleType == 'BE')
+info = subset(info, SampleType == 'BE')
 nrow(info)
 
 if (file.exists('~/Data/Reid_SNP/PerPatient/tmp_seg_pt.Rdata')) {
@@ -215,7 +217,7 @@ nl = 1000;folds = 10; splits = 5
 sets = create.patient.sets(slabels, folds, splits, 0.15)  
 ## ----- All ----- ##
 coefs = list(); plots = list(); performance.at.1se = list(); models = list(); cvs = list()
-cache.dir = '~/Data/Ellie/Analysis/SNP'
+cache.dir = '~/Data/Ellie/Analysis/SNP/no-low'
 dir.create(cache.dir, recursive = T, showWarnings = F)
 file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
 if (file.exists(file)) {

### 2019-02-05 Insilico Splits.Rmd
add Seattle SNP analysis scripts
Updating paper related scripts

@@
 ```{r setup, include=FALSE}
-library(ggplot2)
+library(tidyverse)
 library(ggrepel)
@@
 library(gridExtra)
-library(plyr) 
-library(dplyr)
 library(pROC)
@@
 
-data = '~/Data/Ellie'
+data = '~/Data/BarrettsProgressionRisk'
 
-data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
+data.files = list.files(paste(data, 'QDNAseq/training',sep='/'), full.names=T)
 analysis.files = list.files(paste(data, 'Analysis', sep='/'), full.names=T)
@@
 
-cache.dir = paste(data, 'Analysis/5e6_arms_all_logR', sep='/')
-split.dir = paste(data, 'Analysis/5e6_arms_splits_logR', sep='/')
+cache.dir = paste(data, 'Analysis/5e6_arms_all', sep='/')
+split.dir = paste(data, 'Analysis/5e6_arms_splits', sep='/')
 

### 2019-02-05 Paper.Rmd
Updating paper related scripts

@@
 ```{r setup, include=FALSE}
-library(biomaRt)
-library(GenomicRanges)
-library(stringr)
-library(ggplot2)
+library(BarrettsProgressionRisk)
+
+#library(biomaRt)
+#library(GenomicRanges)
+#library(stringr)
+#library(ggplot2)
 library(ggrepel)
 library(GGally)
-library(plyr) 
 library(pander)
@@
 library(CoxHD)
-library(plyr)
-library(dplyr)
+#library(plyr)
+#library(dplyr)
 
@@
 source('lib/common_plots.R')
-source('lib/data_func.R')
-
-data = '~/Data/Ellie'
-info.dir = '~/Data/Ellie/QDNAseq'
-#tile.w=c(5e06, 'arms')
-#tile.w=5e+06
-#tile.w=1e7
-#tile.w='arms'
-
-#includeDemo = F
+#source('lib/data_func.R')
+
+data = '~/Data/BarrettsProgressionRisk/'
+info.dir = '~/Data/BarrettsProgressionRisk/QDNAseq'
 
@@
 
-all.patient.info = read.patient.info(patient.file, demo.file, set='All')$info
-head(all.patient.info)
-
-patient.info = all.patient.info
-
-validation.patient.info =  subset(all.patient.info, Set == 'Test')
-validation.patient.info = arrange(validation.patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
-sum.validation.info = summarise.patient.info(validation.patient.info)
+patient.info = read.patient.info(patient.file, demo.file, set='All')$info
+head(patient.info)
+
+# validation.patient.info =  subset(all.patient.info, Set == 'Test')
+# validation.patient.info = arrange(validation.patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
+# sum.validation.info = summarise.patient.info(validation.patient.info)
 
@@
 
-all.info = rbind(sum.patient.data, sum.validation.info)
+#all.info = rbind(sum.patient.data, sum.validation.info)
 
@@
 pander(
-  as.data.frame(all.info %>% summarise(
+  as.data.frame(sum.patient.data %>% dplyr::summarise(
   'Total Patients'=length(unique(Patient)),
@@
 ```{r echo=F}
-cohortSummary = as.data.frame(all.info %>% dplyr::group_by(Status) %>% summarise(
+
+cohortSummary = sum.patient.data %>% dplyr::group_by(Status) %>% dplyr::summarise(
   'Total Patients'=length(unique(Patient)),
@@
   'naC'=length(which(is.na(C)))
-))
-
-cohortSummary = cbind(cohortSummary,
-  rbind(table(subset(all.patient.info, Status == 'NP')$Pathology),
-        table(subset(all.patient.info, Status == 'P')$Pathology)))
-
-  rbind(table(subset(patient.info, Status == 'NP')$Pathology),
-        table(subset(patient.info, Status == 'P')$Pathology))
-
-  rbind(table(subset(validation.patient.info, Status == 'NP')$Pathology),
-        table(subset(validation.patient.info, Status == 'P')$Pathology))
+)
+
+cohortSummary = full_join(cohortSummary, cbind.data.frame('Status' = c('NP','P'), rbind(
+      table(subset(patient.info, Status == 'NP')$Pathology),
+      table(subset(patient.info, Status == 'P')$Pathology))), by='Status')
 
@@
 ```{r labelsPNP, echo=F, message=F, include=T}
-cache.dir = paste(data, 'Analysis/5e6_arms_all', sep='/')
+cache.dir = paste(data, 'Analysis/5e6_arms_all', sep='/') # QC'd
 
@@
 
+pg.sampNOHGD = pg.sampNOHGD %>% filter(!is.na(Prediction))
+
 rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')
@@
 noHGD.pred.confidence = as.data.frame.matrix(table(pg.sampNOHGD$quants, pg.sampNOHGD$Status))
-noHGD.pred.confidence = cbind(noHGD.pred.confidence, pg.sampNOHGD %>% dplyr::group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
+noHGD.pred.confidence = bind_cols(noHGD.pred.confidence, pg.sampNOHGD %>% dplyr::group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
 noHGD.pred.confidence$quant = rownames(noHGD.pred.confidence)
@@
 
-ggplot(noHGD.pred.confidence, aes(mn, perc)) + 
-  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
-  geom_point() + geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
-  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration', subtitle='HGD excluded model') + plot.theme + theme(legend.position = 'none')
+BarrettsProgressionRisk::showPredictionCalibration(noHGD.pred.confidence) + labs(subtitle='HGD excluded model')
 
@@
   hgd.samples = subset(patient.info, Pathology %in% c('HGD','IMC'))
-  hgd.df = dysplasia.df[hgd.samples$Samplename,]
+  hgd.df = dysplasia.df[intersect(hgd.samples$Samplename, rownames(dysplasia.df)),]
 
   pred = predict(fitNOHGD, newx=hgd.df, s=lambda.opt, type='response')
-  hgd.samples$Prediction = pred
+  colnames(pred) = 'Prediction'
+  hgd.samples = base::merge(hgd.samples, pred, by.x='Samplename', by.y='row.names')
 
   hgd.pred = hgd.samples %>% dplyr::group_by(Patient, Hospital.Research.ID, Status, Endoscopy.Year) %>% 
-    summarise( max=max(Pathology), pred=Prediction[which.max(Pathology)] )
+    dplyr::summarise( max=max(Pathology), pred=Prediction[which.max(Pathology)] )
 
@@
 }
-
 #hist(hgd.samples$Prediction)
@@
 
+back = back %>% filter(!is.na(Prediction))
+
 back$quants = with(back, cut(Prediction, breaks=cuts))
@@
 ft.fun<-function(NP,P) {
- f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
- cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
+ f = chisq.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
+ cbind.data.frame('p.value'=round(f$p.value, 4))
 }
@@
 
-pred.confidence = qt %>% as.data.frame.matrix %>% dplyr::group_by(quant) %>% dplyr::mutate( 
-  'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )
+pred.confidence =
+  qt %>% as.data.frame.matrix %>% dplyr::group_by(quant) %>% dplyr::mutate( 
+  'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'conf'=ifelse(p.value < 0.05, '*', '') )
 
@@
 
-pred.confidence$Risk = 'Moderate'
-pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc < .5)) ] = 'Low'
-pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'
-
+pred.confidence = pred.confidence %>% dplyr::mutate(Risk = ifelse(perc<.5, 'Low','High'))
 pred.confidence = bind_cols(pred.confidence, 
@@
                                       ci.high=qbeta(0.975, shape1=pred.confidence$P+.5, shape2 = pred.confidence$NP+.5)))
+
+pcfit = lm(perc~mn, data=pred.confidence)
+rsid = rstandard(pcfit)
+pred.confidence[which( rsid >= 1 |  rsid <= -1 ), 'Risk'] = 'Moderate'
 
@@
 ```{r pcalib, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
-pcal = ggplot(pred.confidence, aes(mn, perc)) + 
-  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
-  geom_point() + #geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
-  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration') +
-  scale_color_manual(values=riskCols) + plot.theme + theme(legend.position = 'none', text=element_text(size=10))
+pcal = BarrettsProgressionRisk::showPredictionCalibration(pred.confidence)
 pcal
-
 ggsave(filename='plots/pred_calib.png', plot=pcal, width=4, height=4, units='in', dpi=300)
@@
 
-```{r indP, echo=F, warning=F, message=F, fig.width=5, fig.height=5}
-back = merge(back, pred.confidence[c('quant','Risk')], by.x='quants',by.y='quant')
-back = back %>% dplyr::group_by(Risk) %>% dplyr::mutate('se(Risk)'=(sd(Prediction)/sqrt(length(Risk)))*1.96 )
-
-ggplot(back, aes(1:nrow(back), Prediction, color=Risk)) + 
-  geom_point(aes(color=Risk, shape=Status)) +scale_color_manual(values=riskCols, name='P(P)') + 
-  geom_hline(data=annotation.table, aes(yintercept=max), color='grey') + plot.theme +
-  labs(x='sample', y='model probability', title='Per-sample probabilities')
-```
-
 ```{r relrisk, echo=F, fig.width=6, fig.height=6}
-#myPal = c(RColorBrewer::brewer.pal(9, 'Blues')[9:7], 
-#          adjustcolor(grDevices::heat.colors(10)[6], blue.f=1.5),
-#          ( grDevices::heat.colors(10)[7:6]),
-#          RColorBrewer::brewer.pal(9, 'Reds')[6:9])
-
 myPal = rev(RColorBrewer::brewer.pal(11, 'RdYlBu'))
@@
 unadjRR = ggplot(back, aes(OR)) + geom_histogram(aes(fill=..x..), bins=10, show.legend = F) +
-  #scale_fill_distiller(palette = 'RdYlBu', type='div',name='P(P)') +
   scale_fill_gradientn(colors = myPal,  name='') + 
@@
 
-ggplot(back, aes(Status, OR, color=OR)) + 
-  geom_boxplot(outlier.colour=NA) + geom_jitter(width=0.3, aes(shape=Pathology)) + 
-  scale_color_gradientn(colors = myPal,  name='') + plot.theme
-
 ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR, width=6, height=5, units='in', dpi=300)
@@
     scale_fill_gradientn(colors = myPal,  name='') + 
-  #scale_fill_distiller(palette = 'RdYlBu', type='div',name='P(P)') +
     plot.theme + labs(title='Predictions, all samples model', y='n Samples', x='Probability') 
@@
 ggsave(filename='plots/pred_hist.png', plot=predhist, width=5, height=5, units='in', dpi=300)
-
-#ggplot(back, aes(OR, Prediction, color=Prediction)) + geom_point(show.legend = F) + scale_color_gradientn(colors = myPal,  name='')  + plot.theme
 ```
@@
 if (wtA < 0.0001) wtA = '< 0.0001'
-#wilcox.test(subset(el2, Status == 'Progressor')$`diff(CNA)`, subset(el2, Status == 'Non-Progressor')$`diff(CNA)`)
 
@@
 
+nzcoefs = lapply(nzcoefs, function(x) {
+  as_tibble(x, rownames='coef')
+})
+
+cfs = unique(unlist(sapply(nzcoefs, function(x) x[['coef']])))
+loo.coefs = data.frame(matrix(ncol=0,nrow=length(cfs)))
+loo.coefs$coef = cfs
+for (pt in names(nzcoefs)) {
+  loo.coefs = dplyr::full_join(loo.coefs, nzcoefs[[pt]], by='coef')
+}
+colnames(loo.coefs)[-1] = c(1:length(nzcoefs))
+loo.coefs[is.na(loo.coefs)] = 0
+
+ch = as_tibble(coefs$`0.9`[,1,drop=F], rownames='coef')
+
+loo.ch = dplyr::left_join(ch, loo.coefs, by='coef')
+
 library(bootstrap)
+
+ch$jack.se = apply(loo.ch, 1, function(x) {
+  jk = bootstrap::jackknife(x[-1], var)
+  jk$jack.se
+})
 
@@
 # per coefficient
-ch$jack.se = jack =  apply(coefs.per.pt, 1, function(x) {
+ch$jack.se = apply(coefs.per.pt, 1, function(x) {
   jk = jackknife(x, var)
@@
 })
+colnames(ch) = c('label','coef','jack.se')
 ggplot(ch, aes(reorder(label, coef), coef)) + geom_point() + 
   geom_errorbar(aes(ymin=coef-jack.se, ymax=coef+jack.se)) + 
+  coord_flip() +
+  plot.theme + labs(x='', title='Jackknifed coefficient CI')
+
+ggplot(ch, aes(reorder(label, coef), coef)) + geom_point() + 
+  geom_errorbar(aes(ymin=coef-jack.se*1.96, ymax=coef+jack.se*1.96)) + 
   coord_flip() +
@@
 ggsave(filename='plots/pred_by_path.png', plot=rp, width=9, height=7, units='in', dpi=300)
-
 ```

### 2019-02-05 Validation Test Report.Rmd
Updating paper related scripts

---
title: "sWGS validation/protocol tests"
author: "Sarah Killcoyne"
date: "14 Jan 2019"
output: 
  html_document: 
    toc: yes

---

```{r setup, include=FALSE}

library(BarrettsProgressionRisk)
library(knitr)
library(kableExtra)
library(tidyverse)
library(gridExtra)
library(reshape2)

source('lib/load_patient_metadata.R')

knitr::opts_chunk$set(echo = F, warning = F, message=F)
options(knitr.table.format = "html")

dirs = list(
  training = '~/Data/BarrettsProgressionRisk/Analysis/multipcf_plots_fitted_perPatient',
  val_1 = '~/Data/BarrettsProgressionRisk/Analysis/val_test/',
  protocol = '~/Data/BarrettsProgressionRisk/Analysis/protocol_test/',
  kit = '~/Data/BarrettsProgressionRisk/Analysis/kit_test/'
)

```

## Training

*	777 samples, 88 patients all FFPE, from various sites
*	8 technical repeat samples, 2 patients
*	8 samples from normal gastric/duodenum tissue, 7 patients
*	16 excluded samples, 1 patient (due to clinical pathology)

We had no formal QC process when we initially used these to train.  We manually inspected the segmentation plots and that was all.  There were 10 samples from 5 patients that I looked at and thought they were overdispersed, but they did not appear to have a significant effect on the model.  This is likely due to the fact that there were so few.

One patient (AH0254) had a single 50X sequence performed from their final timepoint as well as 7 sWGS samples.  Of those samples 2 fail on manual inspection due to overdispersion of the data.  The remaining 5 samples are good quality.  All were sequenced in the same batch with 68 other samples.

```{r}
get.png<-function(n,dir) {
  list.files(path=dir, pattern=n, full.names=T, recursive=T)
}

plot.dir = '~/Data/BarrettsProgressionRisk/Analysis/multipcf_plots_fitted_perPatient/'

pts = read.patient.info('~/Data/BarrettsProgressionRisk/QDNAseq/training/All_patient_info.xlsx')$info

train = left_join(readr::read_tsv(list.files(dirs$training, 'resid_var.txt', full.names=T)), read.patient.info('~/Data/BarrettsProgressionRisk/QDNAseq/training/All_patient_info.xlsx')$info, by=c('sample'='Samplename')) %>% select(pid,sample,SLX.ID,varMAD_median)

failed = train %>% filter(varMAD_median > 0.015)

pass.fail = train %>% summarise( Pass= length(which(varMAD_median<0.015))/nrow(train), Fail=length(which(varMAD_median>=0.015))/nrow(train), group='Training')

failed.tbl = failed %>% rowwise() %>% dplyr::mutate(varMAD_median = signif(varMAD_median,4)) %>%
 dplyr::mutate(
   sample = text_spec(sample, link=paste0('file:///',get.png(sample, paste(plot.dir, pid, sep='/') ))), 
   varMAD_median  = cell_spec(varMAD_median, "html", color = ifelse(varMAD_median <= 0.011, "green", "red"))) %>% dplyr::select(pid,sample,varMAD_median)

```

We later developed a computational cutoff. A cutoff was selected for the median variance across segments in a sample of 0.015.  Any sample above that is considered to have failed QC.  In the initial training set that results in `r nrow(failed)` samples failing QC.  A manual check confirms that these samples are overdispersed (especially the two from AH0254).  This is `r round(nrow(failed)/nrow(train)*100,2)`% of our samples.

`r failed.tbl %>% kable(format = "html", escape = F) %>% kable_styling("striped", full_width = F)`

`r kable(train %>% dplyr::summarise_at(vars(varMAD_median), funs(median,mean,sd,min,max)), caption='Summary statistics for the variance across all training samples') %>% kable_styling(bootstrap_options=c('bordered',full_width=F))`

## Initial Validation

*	92 samples, 49 patients, multiple sites
  + 84 FFPE
  + 10 frozen
  + 3 cytosponge
* 2 samples were excluded due to insufficient reads

```{r}
val = do.call(bind_rows, lapply(list.files(dirs$val_1, 'residuals', recursive = T, full.names = T), function(f) {
  readr::read_tsv(f, col_types = 'cdddddl')
}))

val = right_join(readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/validation/ValidationCohort.xlsx') %>% mutate(Samplename = paste0('SLX-',`SLX-ID`, '.', sub('-','_',`Index Sequence`))) %>% select(`Hospital Research ID`,`Block ID`, `Patient ID`, Samplename),val, by=c('Samplename'='sample'))

val.fail = val %>% filter(varMAD_median > 0.015) %>% select(`Hospital Research ID`, Samplename, varMAD_median)

pass.fail = bind_rows(pass.fail, 
  val %>% summarise( Pass= length(which(varMAD_median<=0.011))/nrow(val), Fail=length(which(varMAD_median>0.011))/nrow(val), group='Validation'))
```

By manual inspection I would say 66 of these were poor quality data. Using that same cutoff in the intial validation cohort we would fail `r round(nrow(val.fail)/nrow(val)*100,2)`% of our `r nrow(val)` samples. However, a manual inspection fails many more and when we used our initial more stringent cutoff of 0.011 we failed `r round(nrow(val %>% filter(varMAD_median > 0.011))/nrow(val)*100,2)`% of samples which agreed closely with manual inspection.

`r kable(val.fail %>% dplyr::summarise_at(vars(varMAD_median), funs(median,mean,sd,min,max)), caption='Summary statistics for the variance across all initial validation samples') %>% kable_styling(bootstrap_options=c('bordered',full_width=F))`

## Speedvac Test

* 16 samples, 12 samples from the training dataset and 4 from the validation set
* 8 Speedvac'd
* 8 were not

There was some concern initially that the Speedvac itself was causing the difference as Ellie had used the one at the CI, while all of the validation dataset were run in the MRC. On both manual inspection and via the computational cutoff only the 4 samples that were not in the initial training dataset failed QC.

```{r}

plot.dir = '~/Data/BarrettsProgressionRisk/QDNAseq/protocol_test'
proc.dir = '~/Data/BarrettsProgressionRisk/Analysis/protocol_test'

speedvac = do.call(bind_rows, lapply(list.files(dirs$protocol, 'residuals', recursive = T, full.names = T), function(f) {
  readr::read_tsv(f, col_types='cdddddl')
})) %>% select(sample, varMAD_median)

speedvac = right_join(readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/protocol_test/Speedvac Pilot samples.xlsx') , speedvac, by=c('Sample'='sample'))

pass.fail = bind_rows(pass.fail, 
  speedvac %>% summarise( Pass= length(which(varMAD_median<=0.011))/nrow(speedvac), Fail=length(which(varMAD_median>0.011))/nrow(speedvac), group='Speedvac'))

speedvac.tbl = speedvac %>% arrange(Speedvac, `Ellie’s Cohort`, `Manual QC`) %>% rowwise() %>% dplyr::mutate(varMAD_median = signif(varMAD_median,4)) %>%
 dplyr::mutate(
   #Speedvac = cell_spec(Speedvac, "html", background = ifelse(Speedvac == 'Yes', 'lightblue', 'yellow')),
   ID = text_spec(paste(`SLX-ID`, `Index Sequence`,sep='.'),link=paste0('file:///',get.png( paste(`SLX-ID`, sub('-','_',`Index Sequence`),sep='.'), plot.dir ))),
   Sample = text_spec(Sample, link=paste0('file:///', get.png(paste0(Sample, '_segmentedCoverage.png'), paste(proc.dir,`Patient ID`,sep='/')))),
   #text_spec(Sample, link=paste0('file:///',get.png( paste(`SLX-ID`, sub('-','_',`Index Sequence`),sep='.'), plot.dir ))),
   `Manual QC` = cell_spec(`Manual QC`, "html", background = ifelse(`Manual QC` == 'Pass', "green", "red"), color='white'),
   `Ellie’s Cohort` = cell_spec(`Ellie’s Cohort`, "html", color = ifelse(`Ellie’s Cohort` == 'Yes', "blue", "brown")),
   varMAD_median  = cell_spec(varMAD_median, "html", color = ifelse(varMAD_median <= 0.011, "green", "red"))) %>%
  dplyr::select(ID,Sample, Speedvac, `Ellie’s Cohort`, `Manual QC`, varMAD_median) %>% rename(`Ellie’s Cohort` = 'Previously Extracted')

rowsY = grep('Yes',speedvac.tbl$Speedvac)
rowsN = grep('No',speedvac.tbl$Speedvac)

speedvac.tbl %>%
  kable(format = "html", escape = F) %>% kable_styling("striped", full_width=F) %>%
  group_rows("Speedvac", rowsY[1], rowsY[length(rowsY)],label_row_css = "background-color: #666; color: #fff;") %>%
  group_rows("No Speedvac", rowsN[1], rowsN[length(rowsN)],label_row_css = "background-color: #666; color: #fff;")
```

All 4 that were not from Ellie's cohort failed manual QC (only 2 failed by cutoffs, but the other 2 were marginal). The difference in these data were only that the samples used from the training dataset (n=12) were already extracted while the 4 that were new had to be extracted from scratch. As the training samples had been extracted using GeneRead and the new samples used AllPrep we considered that  the kit could be an issue.

## Kit Test

*	12 cell line FFPE samples 
  + 10 cell lines
  + 2 NP patients from an earlier set
* 7 AllPrep
* 5 GeneRead

```{r}

kit = do.call(bind_rows, lapply(list.files(dirs$kit, 'residuals', recursive = T, full.names = T), function(f) {
  readr::read_tsv(f, col_types='cdddddl')
})) %>% select(sample, varMAD_median)

kit = right_join(readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/kit_test/kit_pilot.xlsx') %>% select(Number, Sample, `Index Sequence`, `Cell line Type`, `Kit Extraction`, `Manual QC`), kit, by=c('Sample'='sample')) %>% arrange(`Manual QC`)

cell.line<-function(cl) {
  if (grepl('Non',cl)) return('NP')
  if (grepl('Barr',cl)) return('BE')
  if (grepl('Colon',cl)) return('COAD')
  return(cl)
}

extract<-function(kit) {
  if (grepl('All',kit)) return('Allprep')
  return('GeneRead')
}

kit = kit %>% rowwise %>% mutate(`Cell line Type` = cell.line(`Cell line Type`), `Kit Extraction`=extract(`Kit Extraction`))
kit = kit %>% rename(`Kit Extraction`='Kit', `Cell line Type`='Type') %>% arrange(Kit)

pass.fail = bind_rows(pass.fail, 
  kit %>% summarise( Pass= length(which(varMAD_median<=0.011))/nrow(kit), Fail=length(which(varMAD_median>0.011))/nrow(kit), group='Kit'))

plot.dir= '~/Data/BarrettsProgressionRisk/QDNAseq/kit_test/SLX-15639'
proc.dir = '~/Data/BarrettsProgressionRisk/Analysis/kit_test'

kit.tbl = kit %>% dplyr::mutate(varMAD_median = signif(varMAD_median,4)) %>% rowwise() %>%
 dplyr::mutate(
   Sample = text_spec(Sample, link=paste0('file:///',get.png('png',paste(proc.dir,as.character(Number),'plots',sep='/')))),
   `Manual QC` = cell_spec(`Manual QC`, "html", background = ifelse(`Manual QC` == 'PASS', "green", "red"), color='white'),
   `Index Sequence` = text_spec(`Index Sequence`, link=paste0('file:///',get.png(sub('-','_',`Index Sequence`),plot.dir))),   
   varMAD_median  = cell_spec(varMAD_median, "html", color = ifelse(varMAD_median <= 0.011, "green", "red"))) %>%
  dplyr::select(`Index Sequence`, Sample, Kit, Type, `Manual QC`, varMAD_median)

rowsY = grep('GeneRead',kit.tbl$Kit)
rowsN = grep('GeneRead',kit.tbl$Kit, invert = T)

kit.tbl %>%
  kable(format = "html", escape = F) %>% kable_styling("bordered", full_width = F) %>%
  group_rows("Kit: GeneRead", rowsY[1], rowsY[length(rowsY)],label_row_css = "background-color: #666; color: #fff;") %>%
  group_rows("Kit: AllPrep", rowsN[1], rowsN[length(rowsN)],label_row_css = "background-color: #666; color: #fff;") 
```

In the Generead vs Allprep kit test we were testing to see whether the specific kit used made the difference.  Ellie had used GeneRead for all of her samples, while we had started to use Allprep.  While it is not conclusive, the GeneRead's did perform better.  However, even here the failure rate was too high and most of the GeneRead's passed manual QC but would not have passed computational QC (<0.015).

Under manual QC 4 samples would appear to be of good enough quality to analyze, 3 were Generead and 1 was Allprep.  All others were poor quality.  However, I would not call this definitive simply because there were so few and that’s still a high rate of failure.

# Conclusion

Currently we are guessing that there is a step during extraction that is causing the problem.  All of Ellie’s samples used in the Speedvac test were extracted prior to use by us, though not necessarily all by Ellie. As the protocol being used was from Ellie's thesis and conversations with her, rather than direct training, it seems possible that something was undocumented/unclear and therefore lost in the mix.

```{r}
ggplot(melt(pass.fail, id.vars = 'group'), aes(group, value,group=group,fill=variable)) + geom_col(position = position_stack()) + labs(y='Ratio of samples', x='') + theme_bw()
```

### 2019-02-05 sWGS QC.Rmd
Updating paper related scripts

@@
 dir1 = '~/Data/Ellie/Analysis/multipcf_plots_fitted_perPatient'
-dir2 = '~/Data/Ellie/Analysis/VAL_Cohort/sWGS/'
+dir2 = '~/Data/Ellie/Analysis/val_test/'
 
@@
 ```{r, echo=F}
+list.files(dir2
+
 resVal = read.table(paste(dir2,'resid_var.txt',sep='/'),header=T)
@@
 
-setdiff(fc$sample,c(as.character(failedQC$sample), as.character(failedQCVal$sample)))
+```
 
-```
+Using the cutoff `r newcutoff` learned from all of them we would fail `r length(which(resVal$varMAD_median > newcutoff))` samples from the training dataset (Ellie's) and `r length(which(resVal$varMAD_median > newcutoff))` from the initial validation dataset.
+So while it's less stringent it still fails `r length(which(failedQCVal$varMAD_median > newcutoff))/nrow(resVal)*100`% versus `r length(which(failedQC$varMAD_median > newcutoff))/nrow(res.variance)*100`% of the training dataset.

### 2019-02-05 snp/SNP_Prediction_Analysis.Rmd
Updating paper related scripts

@@
 
-chr.info = get.chr.lengths(file='/tmp/hg19_info.txt')
+library(BarrettsProgressionRisk)
+
+chr.info = BarrettsProgressionRisk:::chrInfo(build='hg19')
 
@@
 
-load('~/Data/Ellie/Analysis/5e6_arms_all-NOQC/model_data.Rdata', verbose = T)
-load('~/Data/Ellie/Analysis/5e6_arms_all-NOQC/all.pt.alpha.Rdata', verbose = T)
+load('~/Data/Ellie/Analysis/5e6_arms_all/model_data.Rdata', verbose = T)
+load('~/Data/Ellie/Analysis/5e6_arms_all/all.pt.alpha.Rdata', verbose = T)
 #alpha = '0'

### 2019-03-13 Validation Test Report.Rmd
updating

@@
 author: "Sarah Killcoyne"
-date: "14 Jan 2019"
+date: "27 Feb 2019"
 output: 
@@
   protocol = '~/Data/BarrettsProgressionRisk/Analysis/protocol_test/',
-  kit = '~/Data/BarrettsProgressionRisk/Analysis/kit_test/'
+  kit = '~/Data/BarrettsProgressionRisk/Analysis/kit_test/',
+  ci_protocol = '~/Data/BarrettsProgressionRisk/Analysis/ci_protocol/'
 )
 
+qc.cutoff = 0.011
 ```
@@
    sample = text_spec(sample, link=paste0('file:///',get.png(sample, paste(plot.dir, pid, sep='/') ))), 
-   varMAD_median  = cell_spec(varMAD_median, "html", color = ifelse(varMAD_median <= 0.011, "green", "red"))) %>% dplyr::select(pid,sample,varMAD_median)
+   varMAD_median  = cell_spec(varMAD_median, "html", color = ifelse(varMAD_median <= qc.cutoff, "green", "red"))) %>% dplyr::select(pid,sample,varMAD_median)
 
@@
 
-val = right_join(readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/validation/ValidationCohort.xlsx') %>% mutate(Samplename = paste0('SLX-',`SLX-ID`, '.', sub('-','_',`Index Sequence`))) %>% select(`Hospital Research ID`,`Block ID`, `Patient ID`, Samplename),val, by=c('Samplename'='sample'))
+val = right_join(readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/validation/ValidationCohort.xlsx') %>% mutate(Samplename = paste0('SLX-',`SLX-ID`, '.', sub('-','_',`Index Sequence`))) %>% select(`Hospital Research ID`,`Block ID`, `Patient ID`, Samplename, `Sample Type`, Collection),val, by=c('Samplename'='sample'))
 
@@
 pass.fail = bind_rows(pass.fail, 
-  val %>% summarise( Pass= length(which(varMAD_median<=0.011))/nrow(val), Fail=length(which(varMAD_median>0.011))/nrow(val), group='Validation'))
+  val %>% summarise( Pass= length(which(varMAD_median<=qc.cutoff))/nrow(val), Fail=length(which(varMAD_median>qc.cutoff))/nrow(val), group='Validation'))
+
+val %>% group_by(`Sample Type`, Collection ) %>% dplyr::summarise(
+  Pass=length(which(varMAD_median<=qc.cutoff)), Fail=length(which(varMAD_median>qc.cutoff))
+) %>% kable(caption='Pass/Fail by sample type') %>% kable_styling('bordered', full_width = F)
 ```
@@
    `Ellie’s Cohort` = cell_spec(`Ellie’s Cohort`, "html", color = ifelse(`Ellie’s Cohort` == 'Yes', "blue", "brown")),
-   varMAD_median  = cell_spec(varMAD_median, "html", color = ifelse(varMAD_median <= 0.011, "green", "red"))) %>%
+   varMAD_median  = cell_spec(varMAD_median, "html", color = ifelse(varMAD_median <= qc.cutoff, "green", "red"))) %>%
   dplyr::select(ID,Sample, Speedvac, `Ellie’s Cohort`, `Manual QC`, varMAD_median) %>% rename(`Ellie’s Cohort` = 'Previously Extracted')
@@
    `Index Sequence` = text_spec(`Index Sequence`, link=paste0('file:///',get.png(sub('-','_',`Index Sequence`),plot.dir))),   
-   varMAD_median  = cell_spec(varMAD_median, "html", color = ifelse(varMAD_median <= 0.011, "green", "red"))) %>%
+   varMAD_median  = cell_spec(varMAD_median, "html", color = ifelse(varMAD_median <= qc.cutoff, "green", "red"))) %>%
   dplyr::select(`Index Sequence`, Sample, Kit, Type, `Manual QC`, varMAD_median)
@@
 
+## CI Protocol
+
+At this point we were guessing that the extraction method itself may be at issue, and that there could be small protocol errors that we were unaware of.
+We selected a set of samples that had been previously sequenced by Ellie and Cassandra, and then a number of new samples as well.
+
+Cassandra worked with Anna Piskorz from the Brenton lab at the CI to do the extractions using their protocol and their kits, and Anna then did the library prep as well.
+However, again a very high percentage failed.
+Most concerning is that the samples from Ellie's cohort failed and the orginal ones done in 2017 by Ellie all looked fine.
+
+```{r}
+ci = do.call(bind_rows, lapply(list.files(dirs$ci_protocol, 'residuals', recursive = T, full.names = T), function(f) {
+  readr::read_tsv(f, col_types='cdddddl')
+})) %>% select(sample, varMAD_median) %>% dplyr::mutate( Pass = varMAD_median<=qc.cutoff)
+
+info = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/ci_protocol/ci_protocol_trial.xlsx') %>%  
+  dplyr::mutate( sample = paste0(`SLX-ID`, '.', sub('-', '_', gsub('tp','', `Index Sequence`))))
+
+ci = left_join(ci, info, by='sample') %>% arrange(Pass)
+
+pass.fail = bind_rows(pass.fail, 
+  ci %>% dplyr::summarise( Pass= length(which(varMAD_median<=qc.cutoff))/nrow(ci), Fail=length(which(varMAD_median>qc.cutoff))/nrow(ci), group='CI Protocols'))
+
+plot.dir= '~/Data/BarrettsProgressionRisk/QDNAseq/ci_protocol/SLX-15634'
+proc.dir = '~/Data/BarrettsProgressionRisk/Analysis/ci_protocol'
+
+ci.tbl = ci %>% dplyr::mutate(varMAD_median = signif(varMAD_median,4)) %>% rowwise() %>%
+ dplyr::mutate(
+   Sample = text_spec(sample, link=paste0('file:///',get.png('png',paste(proc.dir,as.character(Number),'plots',sep='/')))),
+   `Manual QC` = cell_spec(`Manual QC`, "html", background = ifelse(`Manual QC` == 'PASS', "green", "red"), color='white'),
+   `Computational QC` = cell_spec(Pass, "html", background = ifelse(Pass == T, "green", "red"), color='white'),
+   `Index Sequence` = text_spec(`Index Sequence`, link=paste0('file:///',get.png('png',paste(proc.dir,as.character(Number),'plots',sep='/')))),   
+   varMAD_median  = cell_spec(varMAD_median, "html", color = ifelse(varMAD_median <= qc.cutoff, "green", "red"))) %>%
+  dplyr::select(`Patient ID`,`Index Sequence`, Cohort, `Manual QC`, `Computational QC`, varMAD_median) %>% arrange(Cohort)
+
+rowsE = grep('Ellie',ci.tbl$Cohort)
+rowsC = grep('Cass',ci.tbl$Cohort)
+rowsN = grep('New', ci.tbl$Cohort)
+
+ci.tbl %>%
+  kable(format = "html", escape = F) %>% kable_styling("bordered", full_width = F) %>%
+  group_rows("Cohort: Ellie's", rowsE[1], rowsE[length(rowsE)],label_row_css = "background-color: #666; color: #fff;") %>%
+  group_rows("Cohort: Cassandra's", rowsC[1], rowsC[length(rowsC)],label_row_css = "background-color: #666; color: #fff;")  %>%
+  group_rows("Cohort: New Samples", rowsN[1], rowsN[length(rowsN)],label_row_css = "background-color: #666; color: #fff;")
+
+```
+
 # Conclusion
 
-Currently we are guessing that there is a step during extraction that is causing the problem.
-All of Ellie’s samples used in the Speedvac test were extracted prior to use by us, though not necessarily all by Ellie.
-As the protocol being used was from Ellie's thesis and conversations with her, rather than direct training, it seems possible that something was undocumented/unclear and therefore lost in the mix.
+???
 

### 2019-03-14 Validation Test Report.Rmd
...

@@
 
+NOTE: Until March I had only 49 patients and 92 samples.  I discovered another two SLX runs (15633/15637) that I had no metadata for and no emails.
+I've now included them but have also broken out the batches to make it clearer.
+
 ```{r}
 val = do.call(bind_rows, lapply(list.files(dirs$val_1, 'residuals', recursive = T, full.names = T), function(f) {
-  readr::read_tsv(f, col_types = 'cdddddl')
-}))
+  readr::read_tsv(f, col_types = 'cdddddl') %>% mutate(`Patient ID` = as.character(basename(dirname(f))))
+})) %>% select(sample, varMAD_median, n.segs, `Patient ID`) %>% mutate(varMAD_median = round(varMAD_median, 3))
 
-val = right_join(readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/validation/ValidationCohort.xlsx') %>% mutate(Samplename = paste0('SLX-',`SLX-ID`, '.', sub('-','_',`Index Sequence`))) %>% select(`Hospital Research ID`,`Block ID`, `Patient ID`, Samplename, `Sample Type`, Collection),val, by=c('Samplename'='sample'))
+val = full_join(
+  readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/validation/validation_samples.xlsx', col_types='text') %>% 
+                   mutate(Samplename = paste0(`SLX-ID`, '.', sub('-','_',`Index Sequence`))) %>% 
+                   select(`Hospital Research ID`,`Block ID`, `Patient ID`, Samplename, `Sample Type`, Collection, `SLX-ID`),
+                 val, by=c('Samplename'='sample','Patient ID'))
+
+#val = right_join(readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/validation/ValidationCohort.xlsx') %>% mutate(Samplename = paste0('SLX-',`SLX-ID`, '.', sub('-','_',`Index Sequence`))) %>% select(`Hospital Research ID`,`Block ID`, `Patient ID`, Samplename, `Sample Type`, Collection),val, by=c('Samplename'='sample'))
 
@@
 
-val %>% group_by(`Sample Type`, Collection ) %>% dplyr::summarise(
+sval = val %>% group_by(`SLX-ID` ) %>% dplyr::summarise_at(vars(varMAD_median), funs(min,mean,median,max),na.rm=T)
+
+val %>% group_by(`Sample Type`, `SLX-ID` ) %>% dplyr::summarise(
   Pass=length(which(varMAD_median<=qc.cutoff)), Fail=length(which(varMAD_median>qc.cutoff))
-) %>% kable(caption='Pass/Fail by sample type') %>% kable_styling('bordered', full_width = F)
+) %>% kable(caption='Pass/Fail by batch') %>% kable_styling('striped')
 ```
 
-By manual inspection I would say 66 of these were poor quality data.
+By manual inspection I would say 66 of the intial 92 were poor quality data.
 Using that same cutoff in the intial validation cohort we would fail `r round(nrow(val.fail)/nrow(val)*100,2)`% of our `r nrow(val)` samples.
@@
 
-`r kable(val.fail %>% dplyr::summarise_at(vars(varMAD_median), funs(median,mean,sd,min,max)), caption='Summary statistics for the variance across all initial validation samples') %>% kable_styling(bootstrap_options=c('bordered',full_width=F))`
+`r sval %>% kable(caption='Summary of variance per batch.
+Red rows are where the median value of the variance for the entire batch was greater than the cutoff') %>% kable_styling('striped') %>% row_spec(which(sval$median > cutoff), color='red')`
 

### 2019-06-11 Validation Test Report.Rmd
Updating validation processing

@@
 dirs = list(
-  training = '~/Data/BarrettsProgressionRisk/Analysis/multipcf_plots_fitted_perPatient',
+  #training = '~/Data/BarrettsProgressionRisk/Analysis/multipcf_plots_fitted_perPatient',
   val_1 = '~/Data/BarrettsProgressionRisk/Analysis/val_test/',
@@
   kit = '~/Data/BarrettsProgressionRisk/Analysis/kit_test/',
-  ci_protocol = '~/Data/BarrettsProgressionRisk/Analysis/ci_protocol/'
+  ci_protocol = '~/Data/BarrettsProgressionRisk/Analysis/ci_protocol/',
+  old_vs_new = '~/Data/BarrettsProgressionRisk/Analysis/sujath_oac/',
+  old_progressors = '~/Data/BarrettsProgressionRisk/Analysis/old_progressors/'
 )
 
+get.png<-function(n,dir) {
+  list.files(path=dir, pattern=n, full.names=T, recursive=T)
+}
+
 qc.cutoff = 0.011
+pass.fail = tibble()
 ```
@@
 
-```{r}
-get.png<-function(n,dir) {
-  list.files(path=dir, pattern=n, full.names=T, recursive=T)
-}
+```{r, eval=F}
 
@@
 
-```
-
-We later developed a computational cutoff.
+<!-- We later developed a computational cutoff.
 A cutoff was selected for the median variance across segments in a sample of 0.015.
@@
 
-`r kable(train %>% dplyr::summarise_at(vars(varMAD_median), funs(median,mean,sd,min,max)), caption='Summary statistics for the variance across all training samples') %>% kable_styling(bootstrap_options=c('bordered',full_width=F))`
+`r kable(train %>% dplyr::summarise_at(vars(varMAD_median), funs(median,mean,sd,min,max)), caption='Summary statistics for the variance across all training samples') %>% kable_styling(bootstrap_options=c('bordered',full_width=F))` -->
+
+```
 
@@
 `r sval %>% kable(caption='Summary of variance per batch.
-Red rows are where the median value of the variance for the entire batch was greater than the cutoff') %>% kable_styling('striped') %>% row_spec(which(sval$median > cutoff), color='red')`
+Red rows are where the median value of the variance for the entire batch was greater than the cutoff') %>% kable_styling('striped') %>% row_spec(which(sval$median > qc.cutoff), color='red')`
 
@@
 
+## Sujath - Old vs New (ish) sample tests
+
+Here we thought that we could see if the issue was older vs newer samples and we could compare that to the same patient with a frozen. These are all tumour samples (OAC) from OCCAMS.
+
+```{r}
+ci = do.call(bind_rows, lapply(list.files(dirs$old_vs_new, 'residuals', recursive = T, full.names = T), function(f) {
+  readr::read_tsv(f, col_types='cdddddl')
+})) %>% select(sample, varMAD_median) %>% dplyr::mutate( Pass = varMAD_median<=qc.cutoff)
+
+info = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/sujath_oac/ffpe_frozen_samples.xlsx') %>%  
+  dplyr::select(-matches('Adrienne|Notes')) %>% rename(Samplename='sample')
+
+ci = left_join(ci, info, by='sample') %>% arrange(Pass)
+
+pass.fail = bind_rows(pass.fail, 
+  ci %>% dplyr::summarise( Pass= length(which(varMAD_median<=qc.cutoff))/nrow(ci), Fail=length(which(varMAD_median>qc.cutoff))/nrow(ci), group='Old.v.New'))
+
+plot.dir= '~/Data/BarrettsProgressionRisk/QDNAseq/sujath_oac/SLX-16753_16755_16758/'
+proc.dir = '~/Data/BarrettsProgressionRisk/Analysis/sujath_oac/'
+
+ci.tbl = ci %>% dplyr::mutate(varMAD_median = signif(varMAD_median,4)) %>% rowwise() %>%
+ dplyr::mutate(
+   Sample = text_spec(sample, link=sample),
+#   `Manual QC` = cell_spec(`Manual QC`, "html", background = ifelse(`Manual QC` == 'PASS', "green", "red"), color='white'),
+   `Computational QC` = cell_spec(Pass, "html", background = ifelse(Pass == T, "green", "red"), color='white'),
+   `Index Sequence` = text_spec(`Index Sequence`, tooltip=`sample`),   
+   varMAD_median  = cell_spec(varMAD_median, "html", color = ifelse(varMAD_median <= qc.cutoff, "green", "red"))) %>%
+  dplyr::select(`Patient ID`,`Index Sequence`, `Sample Type`, `Computational QC`, varMAD_median) %>% arrange(`Sample Type`)
+
+rowsE = grep('FFPE',ci.tbl$`Sample Type`)
+rowsC = grep('FFPE MRC',ci.tbl$`Sample Type`)
+rowsN = grep('Frozen',ci.tbl$`Sample Type`)
+
+ci.tbl %>%
+  kable(format = "html", escape = F) %>% kable_styling("bordered", full_width = F) %>%
+  group_rows("Cohort: FFPE", rowsE[1], rowsE[length(rowsE)],label_row_css = "background-color: #666; color: #fff;") %>%
+  group_rows("Cohort: FFPE cut by AB", rowsC[1], rowsC[length(rowsC)],label_row_css = "background-color: #666; color: #fff;")  %>%
+  group_rows("Cohort: Frozen", rowsN[1], rowsN[length(rowsN)],label_row_css = "background-color: #666; color: #fff;")
+
+```
+
+## Old progressors
+
+Sujath and Adrienne extracted the progressor scrolls that had been cut about 12 months ago and left in tubes.
+These are likely to have oxidation issues making them potentially poor samples for inclusion, but it was worth checking.
+We included a number from Sujath's successful OAC extractions as controls.
+
+```{r}
+ci = do.call(bind_rows, lapply(list.files(dirs$old_progressors, 'residuals', recursive = T, full.names = T), function(f) {
+  readr::read_tsv(f, col_types='cdddddl')
+})) %>% select(sample, varMAD_median) %>% dplyr::mutate( Pass = varMAD_median<=qc.cutoff)
+
+info = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/old_progressors/prog_val.xlsx') %>%  
+  dplyr::mutate( sample = paste0(`SLX-ID`, '.', `Index Sequence`) ) %>% select(-matches('DNA')) %>% filter(!is.na(`SLX-ID`)) %>% dplyr::rename('Number' = 'Extraction Sample #')
+
+ci = left_join(ci, info, by='sample') %>% arrange(Pass)
+
+pass.fail = bind_rows(pass.fail, 
+  ci %>% dplyr::summarise( Pass= length(which(varMAD_median<=qc.cutoff))/nrow(ci), Fail=length(which(varMAD_median>qc.cutoff))/nrow(ci), group='Old Progressors'))
+
+plot.dir= '~/Data/BarrettsProgressionRisk/QDNAseq/old_progressors/SLX-17965_17966/'
+proc.dir = '~/Data/BarrettsProgressionRisk/Analysis/old_progressors'
+
+ci.tbl = ci %>% dplyr::mutate(varMAD_median = signif(varMAD_median,4)) %>% rowwise() %>%
+ dplyr::mutate(
+   Sample = text_spec(sample, link=paste0('file:///',get.png('png',paste(proc.dir,as.character(Number),'plots',sep='/')))),
+   #`Manual QC` = cell_spec(`Manual QC`, "html", background = ifelse(`Manual QC` == 'PASS', "green", "red"), color='white'),
+   `Computational QC` = cell_spec(Pass, "html", background = ifelse(Pass == T, "green", "red"), color='white'),
+   `Index Sequence` = text_spec(`Index Sequence`, link=paste0('file:///',get.png('png',paste(proc.dir,as.character(Number),'plots',sep='/')))),   
+   varMAD_median  = cell_spec(varMAD_median, "html", color = ifelse(varMAD_median <= qc.cutoff, "green", "red"))) %>%
+  dplyr::select(`Patient ID`,`Index Sequence`, `Computational QC`, varMAD_median)
+
+#rowsE = grep('Ellie',ci.tbl$Cohort)
+#rowsC = grep('Cass',ci.tbl$Cohort)
+#rowsN = grep('New', ci.tbl$Cohort)
+
+ci.tbl %>%
+  kable(format = "html", escape = F) %>% kable_styling("bordered", full_width = F)
+
+```
+
 # Conclusion
@@
 
-```{r}
-ggplot(melt(pass.fail, id.vars = 'group'), aes(group, value,group=group,fill=variable)) + geom_col(position = position_stack()) + labs(y='Ratio of samples', x='') + theme_bw()
+```{r, fig.height=6, fig.width=6}
+cols = RColorBrewer::brewer.pal(3,'Set1')[c(1,3)]
+ggplot(melt(pass.fail, id.vars = 'group'), aes(group, value,group=group,fill=variable)) + geom_col(position = position_stack()) + labs(y='Ratio of samples', x='') + theme_bw() + scale_fill_manual(values=c('Pass'=cols[2], 'Fail'=cols[1]), name='')
 ```

### 2019-06-17 Validation Report.Rmd
updating validation tests

---
title: "sWGS validation report"
author: "Sarah Killcoyne"
date: "11 June 2019"
output: 
  html_document: 
    toc: yes

---

```{r setup, include=FALSE}

library(BarrettsProgressionRisk)
library(knitr)
library(kableExtra)
library(tidyverse)
library(gridExtra)
library(reshape2)

#source('lib/load_patient_metadata.R')

knitr::opts_chunk$set(echo = F, warning = F, message=F)
options(knitr.table.format = "html")

val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
analysis.dir = '~/Data/BarrettsProgressionRisk/Analysis/validation'

fncols <- function(data, cname, default=NA) {
  add <-cname[!cname%in%names(data)]

  if(length(add)!=0) data[add] <- default
  data
}

sheets = readxl::excel_sheets(val.file)[1:10]
all.val = do.call(bind_rows, lapply(sheets, function(s) {
  readxl::read_xlsx(val.file, s) %>% select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA, matches('Collection')) %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% fncols('Collection', 'Biopsy')
}))

pastefun<-function(x) {
  if ( !grepl('SLX-', x) ) x = paste0('SLX-',x)
  return(x)
}

all.val = all.val %>% rowwise %>% mutate_at(vars(`SLX-ID`), list(pastefun) ) %>% ungroup
all.val = all.val %>% mutate(
  `Hospital Research ID` = str_replace_all( str_remove_all(`Hospital Research ID`, " "), '/', '_'), 
  `Index Sequence` = str_replace_all(`Index Sequence`, 'tp', ''),
  Samplename = paste(`SLX-ID`,`Index Sequence`,sep='.')
  )

get.png<-function(n,dir) {
  list.files(path=dir, pattern=n, full.names=T, recursive=T)
}

val = do.call(bind_rows, lapply(list.files(analysis.dir, 'residuals', recursive = T, full.names = T), function(f) {
  readr::read_tsv(f, col_types = 'cdddddl', col_names = F) %>% set_names('sample', 'varMAD_median', 'varMAD_sd', 'varMAD_Q1', 'varMAD_Q3', 'n.segs', 'Pass')  %>% 
    mutate(`Patient ID` = as.character(basename(dirname(f))))
})) %>% select(`Patient ID`, sample, varMAD_median, n.segs) %>% mutate(varMAD_median = round(varMAD_median, 3)) %>% mutate(sample = str_replace(sample, '_', '-'))

all.val %>% filter(!Samplename %in% val$sample )

val = val %>% filter(sample %in% (all.val %>% select(Samplename) %>% pull) ) %>% 
          left_join( all.val, val %>% mutate(sample = str_replace(sample, '_', '-')), by=c('sample'='Samplename'))

qc.cutoff = 0.011
pass.fail = tibble()
```

# Quick Summary

Effectively every batch C.K. ran failed at a rate of 70-90% of samples. None of the tests we ran showed any effect, with the possible exception of the extraction kit.  Frozen samples always do well, FFPE is clearly a little tricker.

Even the old progressor scrolls that had been sitting for a year in tubes worked well enough to pass QC and provide predictions. To date those suggest a false negative rate of about 38% (by sample).  We have more progressors extracted and will sequence these shortly.

We are now requesting and extracting all of the NP cohort that we had previously identified.

## By Batch

```{r}

pass.fail = bind_rows(pass.fail, 
  val %>% group_by(Batch, Cohort) %>% dplyr::summarise( Pass= length(which(varMAD_median<=qc.cutoff))/nrow(val), Fail=length(which(varMAD_median>qc.cutoff))/nrow(val)))

sval = val %>% group_by(Batch) %>% dplyr::mutate(Cohort = Cohort[1]) %>% group_by(Batch, Cohort) %>% dplyr::summarise_at(vars(varMAD_median), funs(min,mean,median,max),na.rm=T)

vt = val %>% group_by(Batch) %>% dplyr::mutate(Cohort = Cohort[1]) %>% group_by(Batch, `Sample Type`, Cohort) %>% dplyr::summarise(
  `Total Samples` = length(Batch),
  Pass=length(which(varMAD_median<=qc.cutoff)), Fail=length(which(varMAD_median>qc.cutoff))
)

vt %>% kable(caption='Pass/Fail by batch, red rows failed > 10% of samples') %>% kable_styling('striped') %>% row_spec(which(vt$Fail/vt$`Total Samples` >  0.1), color='red')

sval %>% kable(caption='Summary of variance per batch. Red rows are where the median value of the variance for the entire batch was greater than the cutoff') %>% kable_styling('striped') %>% row_spec(which(sval$median > qc.cutoff), color='red')

```

## By Sample Type

Frozen samples are typically very good, and our few Cytosponge samples pass as well.

```{r}
val$RA = factor(val$RA)

val %>% group_by(as.integer(RA),`Sample Type`, Collection) %>% dplyr::summarise(
  `Total Samples` = length(Batch),
  Pass=length(which(varMAD_median<=qc.cutoff)), Fail=length(which(varMAD_median>qc.cutoff))
) %>% rename('as.integer(RA)' = 'Processing') %>% kable(caption='Pass/Fail by sample type') %>% kable_styling('striped')

```

# Initial Pilots

The initial pilot results didn't look too bad.  But the problem was that the ones that failed were all FFPE and we only did 10.

```{r}

val %>% ungroup %>% filter(Batch <= 2) %>% dplyr::group_by(Batch, `Sample Type`) %>% dplyr::summarise(
  `Total Samples` = n(),
  `% Passed QC` = round((length(which(varMAD_median<=qc.cutoff))/n())*100,1)
) %>% kable() %>% kable_styling('striped')

```

## Initial NP Validation

Here we should've done another, slightly larger pilot, before diving in.  We didn't, and more than 90% failed QC.  The question was why.

```{r}
val %>% ungroup %>% filter(Batch %in% c(3,4)) %>% dplyr::group_by(Batch, `Sample Type`, Collection) %>% dplyr::summarise(
  Total = n(),
  `% Passed QC` = round((length(which(varMAD_median<=qc.cutoff))/n())*100,1)
) %>% kable(caption='Total samples and QC results') %>% kable_styling('striped')
```

## Speedvac test

* 16 samples, 12 samples from the training dataset and 4 from the validation set
* 8 Speedvac'd
* 8 were not

There was some concern initially that the Speedvac itself was causing the difference as Ellie had used the one at the CI, while all of the validation dataset were run in the MRC. On both manual inspection and via the computational cutoff only the 4 samples that were not in the initial training dataset failed QC.

```{r}

speedvac = left_join(val %>% ungroup %>% filter(Batch %in% c(5)), readxl::read_xlsx(val.file, sheet='Batch 5') %>% select(`Hospital Research ID`, `Block ID`, `Original Cohort`, `Speedvac`))

speedvac %>% dplyr::group_by(Batch, `Sample Type`, `Original Cohort`, Speedvac) %>% dplyr::summarise(
  Total = n(),
  `% Passed QC` = round((length(which(varMAD_median<=qc.cutoff))/n())*100,1)
) %>% kable() %>% kable_styling('striped')

```

All 4 that were not from Ellie's cohort failed QC. The difference in these data were only that the samples used from the training dataset (n=12) were already extracted while the 4 that were new had to be extracted from scratch. As the training samples had been extracted using GeneRead and the new samples used AllPrep we considered that the kit could be an issue.

## Kit Test

*	12 cell line FFPE samples 
  + 10 cell lines
  + 2 NP patients from an earlier set
* 7 AllPrep
* 5 GeneRead

```{r}

kit = left_join(val %>% ungroup %>% filter(Batch %in% c(6)), 
          readxl::read_xlsx(val.file, sheet='Batch 6') %>% dplyr::select(`Block ID`, `Kit Extraction`, `Sample`) %>% mutate_at(vars(`Block ID`), list(as.character)))

kit %>% dplyr::group_by(Batch, `Kit Extraction`) %>% dplyr::summarise(
  Total = n(),
  `% Passed QC` = round((length(which(varMAD_median<=qc.cutoff))/n())*100,1)
) %>% kable() %>% kable_styling('striped')

```

In the Generead vs Allprep kit test we were testing to see whether the specific kit used made the difference.  Ellie had used GeneRead for all of her samples, while we had started to use Allprep.  While it is not conclusive, the GeneRead's did perform better.  However, even here the failure rate was too high and most of the GeneRead's passed manual QC but would not have passed computational QC.

I would not call this definitive simply because there were so few and that’s still a high rate of failure.  However, GeneRead appears to have a better rate overall.

## CI Protocol

At this point we were guessing that the extraction method itself may be at issue, and that there could be small protocol errors that we were unaware of.  We selected a set of samples that had been previously sequenced by Ellie and Cassandra, and then a number of new samples as well.

Cassandra worked with Anna Piskorz from the Brenton lab at the CI to do the extractions using their protocol and their kits, and Anna then did the library prep as well.  However, again a very high percentage failed.  Most concerning is that the samples from Ellie's cohort failed and the orginal ones done in 2017 by Ellie all looked fine.

```{r}
ci = left_join(val %>% ungroup %>% filter(Batch %in% c(7)),
          readxl::read_xlsx(val.file, sheet='Batch 7') %>% dplyr::select(`Hospital Research ID`, `Block ID`, `Original cohort`, `Sample Type`, Cohort))

ci %>% dplyr::group_by(Batch, `Original cohort`) %>% dplyr::summarise(
  Total = n(),
  `% Passed QC` = round((length(which(varMAD_median<=qc.cutoff))/n())*100,1)
) %>% kable() %>% kable_styling('striped')
```

# New processing

At this point Sujath and Adrienne took over.

## FFPE vs Frozen (old vs new as well)

Here we thought that we could see if the issue was older vs newer samples and we could compare that to the same patient with a frozen. These are all tumour samples (OAC) from OCCAMS. We also cut some of these in house to see if there might be an effect.

```{r}

on = left_join(val %>% ungroup %>% filter(Batch %in% c(8)),
               readxl::read_xlsx(val.file, sheet='Batch 8') %>% dplyr::select(`Block ID`, `Sample Date`, `Sample Type`))

m18 = seq(as.Date("2019/05/01"), length = 2, by = "-18 months")[2]

on = on %>% mutate( Age = ifelse( `Sample Date` < m18, '<18m', '>=18m') )

on %>% dplyr::group_by(Batch, `Sample Type`, `Age`) %>% dplyr::summarise(
  Total = n(),
  `% Passed QC` = round((length(which(varMAD_median<=qc.cutoff))/n())*100,1)
) %>% kable() %>% kable_styling('striped')

```

All passed QC regardless of sample age or FFPE/Frozen.

## Old progressors

Sujath and Adrienne extracted the progressor scrolls that had been cut about 12 months ago and left in tubes. These are likely to have oxidation issues making them potentially poor samples for inclusion, but it was worth checking. We included a number from Sujath's successful OAC extractions as controls.

```{r}

oldp = val %>% ungroup %>% filter(Batch %in% c(9))

oldp %>% dplyr::group_by(Batch, `Sample Type`, `Cohort`) %>% dplyr::summarise(
  Total = n(),
  `% Passed QC` = round((length(which(varMAD_median<=qc.cutoff))/n())*100,1)
) %>% kable() %>% kable_styling('striped')

```

All samples passed QC, though by manual QC I might have failed 2 they weren't as extreme as earlier examples had been.

### Predictions

```{r}

if (file.exists('oldp_preds.txt')) {
  oldp.preds = read_tsv('oldp_preds.txt')
} else {
  segFiles = grep(paste(oldp$`Patient ID`,collapse='|'), list.files('~/Data/BarrettsProgressionRisk/Analysis/validation', 'segObj',  full.names = T, recursive = T), value=T)

  info = readxl::read_xlsx(val.file, 'Batch 9') %>% dplyr::select(`Hospital Research ID`, `Block ID`, Endoscopy, Pathology, `SLX-ID`, `Index Sequence`, `Path Notes`) %>% mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`))

  oldp.preds = do.call(bind_rows, lapply(segFiles, function(f) {
    load(f)  
    segmented$sample.info = BarrettsProgressionRisk::loadSampleInformation(info %>% filter(Sample %in% segmented$sample.info$Sample) )

    prr = BarrettsProgressionRisk::predictRiskFromSegments(segmented,F)
    predictions(prr)
  }))
  write_tsv(oldp.preds, 'oldp_preds.txt')
}

oldp.preds = oldp.preds %>% filter(Sample %in% (oldp %>% filter(!grepl('OCCAMS',`Hospital Research ID`)) %>% select(sample) %>% pull))

oldp.preds %>% dplyr::summarise(
  'TPR'=length(which(Risk == 'High'))/n(),
  'FNR'=length(which(Risk == 'Low'))/n()) %>% kable() %>% kable_styling(full_width = F)

oldp.preds %>% group_by(`Risk`) %>% dplyr::summarise( 'n samples'=n() ) %>% 
  kable(caption='Count sample predicted risks, P') %>% kable_styling(full_width = F)

fnr.patients = oldp.preds %>% 
  group_by(`Hospital Research ID`) %>% dplyr::summarise(
  TPR = length(which(Risk == 'High'))/n()
) %>% dplyr::summarise( 'Misclassified patients' = length(which(TPR < 0.5))/n() ) %>% pull

oldp.preds %>% group_by(`Hospital Research ID`) %>% dplyr::summarise(
  `n samples` = n(), 
  'High Risk' = length(which(Risk == 'High')),
  'Mod. Risk' = length(which(Risk == 'Moderate')),
  'Low Risk' =   length(which(Risk == 'Low')),
  'Pathology Notes' = paste(`Path Notes`, collapse='; ')
) %>% kable(caption = 'Classified predictions per patient (all P)') %>% kable_styling('striped')

```

For the old progressor samples to-date, the model failed to classify any samples as high risk for *`r fnr.patients*100`%* of patients.  None of these were the samples I might have said were poor quality via manual QC so taken at face value these patients/samples were incorrectly predicted.

## New non-progressors

```{r}

new.np = val %>% ungroup %>% filter(Batch %in% c(10))

new.np %>% dplyr::group_by(Batch, `Sample Type`, `Cohort`) %>% dplyr::summarise(
  Total = n(),
  `% Passed QC` = round((length(which(varMAD_median<=qc.cutoff))/n())*100,1)
) %>% kable() %>% kable_styling('striped')

```

Again, all passed QC though on manual QC I might have failed 2-3.  There were 3 samples that just squeaked in under the cutoff threshold.  But they did.

### Predictions

```{r}

if (file.exists('batch_10.txt')) {
  new.np.preds = read_tsv('batch_10.txt')
} else {

  segFiles = grep(paste(unique(new.np$`Patient ID`),collapse='|'), list.files('~/Data/BarrettsProgressionRisk/Analysis/validation', '2_segObj',  full.names = T, recursive = T), value=T)

  info = readxl::read_xlsx(val.file, 'Batch 10') %>% dplyr::select(`Hospital Research ID`, `Block ID`, Endoscopy, Pathology, `SLX-ID`, `Index Sequence`) %>% mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`))

  new.np.preds = do.call(bind_rows, lapply(segFiles, function(f) {
    load(f)  
    segmented$sample.info = BarrettsProgressionRisk::loadSampleInformation(info %>% filter(Sample %in% segmented$sample.info$Sample) )

    prr = BarrettsProgressionRisk::predictRiskFromSegments(segmented,F)
    predictions(prr)
  }))
  write_tsv(new.np.preds, 'batch_10.txt')
}

new.np.preds %>% dplyr::summarise(
  'FPR'=length(which(Risk == 'High'))/n(),
  'TPR'=length(which(Risk == 'Low'))/n()) %>% kable() %>% kable_styling(full_width = F)

new.np.preds %>% group_by(`Risk`) %>% dplyr::summarise( 'n samples'=n() ) %>% 
  kable(caption='Count sample predicted risks, NP') %>% kable_styling(full_width = F)

fpr.patients = new.np.preds %>%  group_by(`Hospital Research ID`) %>% dplyr::summarise(
  TPR = length(which(Risk == 'Low'))/n()
) %>% dplyr::summarise( 'Misclassified patients' = length(which(TPR < 0.5))/n() ) %>% pull

new.np.preds %>% group_by(`Hospital Research ID`) %>% dplyr::summarise(
  `n samples` = n(), 
  'High Risk' = length(which(Risk == 'High')),
  'Mod. Risk' = length(which(Risk == 'Moderate')),
  'Low Risk' =   length(which(Risk == 'Low'))
) %>% kable(caption = 'Classified predictions per patient (all NP)') %>% kable_styling('striped')

```

### 2019-06-17 snp/SNP QC.Rmd
updating validation tests

---
title: "SNP QC"
author: "Sarah Killcoyne"
date: "11/5/2018"
output: 
  html_document: 
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = T)

library(reshape2)
library(readxl)
library(tidyverse)
library(pander)
library(gridExtra)

library(BarrettsProgressionRisk)

chr.info = BarrettsProgressionRisk:::chrInfo(build='hg19')

load('~/Data/Reid_SNP/PerPatient/allpts_ascat.Rdata', verbose=T)

qcdata = qcdata %>% rowwise %>% dplyr::mutate(
  PatientID = unlist(strsplit(Samplename, '_'))[1],
  SampleID = unlist(strsplit(Samplename, '_'))[2],
  EndoID = unlist(strsplit(Samplename, '_'))[3],
  Level = unlist(strsplit(Samplename, '_'))[4]
  )
segments.list = lapply(segments.list, function(pt) {
  pt$sample = sub('\\.LogR','',pt$sample)
  pt
})
qcdata$Samplename = sub('_$','',qcdata$Samplename)

qcdata$`ASCAT SCA Ratio` = apply(qcdata,1,function(s) {
  smp = subset(segments.list[[ s[['PatientID']] ]], sample == s[['Samplename']] & chr %in% c(1:22))
  smp = smp %>% rowwise %>% dplyr::mutate(
      #'Total' = nAraw + nBraw,
      'Total' = nMajor + nMinor,
      'CNV' = round(Total) - round(as.numeric(s[['Ploidy']])) )

  x = subset(smp, CNV != 0 & chr %in% c(1:22))
  sum(x$endpos - x$startpos) / chr.info[22,'genome.length',drop=T]
})

qcdata$SampleType = 'BE'
qcdata$SampleType[grep('BLD',qcdata$Level, ignore.case=T)] = 'Blood Normal'
qcdata$SampleType[grep('gastric',qcdata$Level, ignore.case=T)] = 'Gastric Normal'

patient.info = read_xlsx('~/Data/BarrettsProgressionRisk/Analysis/SNP/metadata_T1T2.xlsx', col_types=c('text'))
patient.info = patient.info %>% dplyr::mutate( 
  UniqueSampleID = paste(PatientID, `Timepoint Code`, sep='_'),
  Status.by.Pathology = ifelse(Pathology %in% c('HGD','IMC'), 'P', Status)
)
qcdata = left_join(qcdata, patient.info, by=c('PatientID'='PatientID','EndoID'='Timepoint Code'))

sample.list = read_xlsx('~/Data/BarrettsProgressionRisk//Analysis/SNP/20180604_Reid_1M_SampleList.xlsx', sheet = 2, col_types=c('text','text','numeric'))
colnames(sample.list)[3] = 'Total.SCA'
sample.list = sample.list %>% dplyr::mutate('Li SCA Ratio'=Total.SCA/max(chr.info$genome.length))

message(paste(length(unique(qcdata$PatientID)), 'patients listed in metadata file'))
message(paste(nrow(qcdata), 'samples available'))

qcdata = left_join(qcdata, sample.list, by=c('PatientID'='PatientID','SampleID'='SampleNum'))
qcdata = qcdata %>% rowwise() %>% dplyr::mutate( 
  Endoscopy = ifelse(Timepoint == 'T1', 1, 2),
  Status=ifelse(SampleType == 'BE', Status, 'Normal'))

# There's one sample that doesn't fit any of the naming conventions.  I think it's a normal sample but I can't match it up '524_23035N_154P'
qcdata = subset(qcdata, !is.na(Status))

normal = subset(qcdata, SampleType != 'BE')
qcdata = subset(qcdata, SampleType == 'BE')

```

`r nrow(patient.info)` total endoscopies, `r pander(table(patient.info$Status), caption="Split between P and NP")`

`r nrow(qcdata)` total samples, `r pander(table(qcdata$Status), caption='NP vs P')`

Histology breakdown:

```{r, echo=F}
sm = qcdata %>% group_by(Status, Endoscopy, Pathology) %>% dplyr::tally() %>% group_by(Status, Endoscopy) %>% dplyr::mutate(freq=n/sum(n) )
sm$Pathology = factor(sm$Pathology, levels=c('NDBE','ID','LGD','HGD','IMC'), ordered = T)

ggplot(sm, aes(Pathology, freq, group=Endoscopy, fill=factor(Endoscopy))) + facet_grid(~Status) + 
  geom_col(position = position_dodge()) + ylim(0,1) + labs(title='Pathology frequency per timepoint')
```

# ASCAT Stats

Purity, ploidy and goodnessoffit from single sample (not paired) ASCAT.

## Ploidy

Most samples regardless of status are likely diploid.  A small subset of NP samples may be tetraploid.  Progressor samples are more likely to show higher ploidies.

```{r}
ggplot(qcdata, aes(Ploidy, `ASCAT SCA Ratio`, color=Status)) + geom_point() + facet_grid(~Status)

ggplot( bind_rows(qcdata, normal), aes(round(Ploidy,2), fill=SampleType)) + geom_histogram(show.legend = F) + 
  facet_wrap(~SampleType, ncol=1) + labs(x='Ploidy', title='Ploidy, normal vs BE')

```

## Purity

All of the samples were purified before SNP analysis and should show low contamination (i.e. high purity).  We can ignore this.

```{r}
ggplot(bind_rows(qcdata,normal), aes(round(Purity,2))) + geom_histogram() + 
  facet_wrap(~SampleType, scales='free_y', ncol=1) + labs(x='Purity', title='Purity, normal vs BE')

ggplot(bind_rows(qcdata,normal), aes(Purity)) + geom_histogram() +
  facet_wrap(~SampleType, scales = 'free_y') + labs(title='Purity by sampletype')
```

## Goodnessoffit

All were 80% or better fit.

```{r}
ggplot(bind_rows(qcdata,normal), aes(round(Goodnessoffit,2))) + geom_histogram() + 
  facet_wrap(~SampleType, scales='free_y', ncol=1) + labs(x='Goodnessoffit', title='Goodnessoffit, normal vs BE')

ggplot(qcdata, aes(Goodnessoffit)) + geom_histogram() +
  facet_wrap(~Status, scales = 'free_y') + labs(title='Goodnessoffit by sampletype')

ggplot(qcdata, aes(Goodnessoffit, `ASCAT SCA Ratio`, color=Status)) + geom_point() + facet_grid(~Status)
```

## SCA Ratio

### Li 2014 SCA vs ASCAT SCA

The % of the genome altered by SCA's is highly correlated between Li 2014 and ASCAT.

```{r}
ggplot(qcdata, aes(`ASCAT SCA Ratio`, `Li SCA Ratio`)) + geom_point() + facet_grid(~Status)
```

```{r} 
pander(with(qcdata, cor.test(`ASCAT SCA Ratio`, `Li SCA Ratio`)))
```

With regards to the clear WGD samples (in this case only those with a ploidy of ≥4) the correlation is still very high. So there's not a good argument for adjusting for a whole genome doubling.

```{r}
pander(with(subset(qcdata, Ploidy >= 4, select=grep('SCA Ratio',colnames(qcdata))),
     cor.test(`ASCAT SCA Ratio`, `Li SCA Ratio`)), caption='Samples with a ploidy ≥4')
```

```{r, echo=F, warning=F, message=F}
ggplot(bind_rows(qcdata,normal), aes(`ASCAT SCA Ratio`, color=Status, fill=Status)) + facet_wrap(Status~SampleType, scales='free_y',nrow=2) +  
  geom_density(alpha=0.5) + labs(title='SCA Ratio, NP vs P') 
```

```{r}
pander(rbind('BE'=summary(qcdata$`ASCAT SCA Ratio`), 'Normal'=summary(normal$`ASCAT SCA Ratio`)), caption='Summary of ASCAT SCA ratio by tissue type')
```

```{r}
ggplot(qcdata, aes(Status, `ASCAT SCA Ratio`, fill=Status, group=Status)) + geom_boxplot(outlier.color = NA) + geom_jitter(width=0.2) + labs(title='BE tissues only')
```

```{r}

anno = cbind.data.frame(table(subset(qcdata, `ASCAT SCA Ratio` <= median(normal$`ASCAT SCA Ratio`))$Status)/table(qcdata$Status))
colnames(anno) = c('Status','Ratio')
anno$x = nrow(qcdata)/2
anno$y = median(normal$`ASCAT SCA Ratio`)
#anno$label=paste0('median(normal SCA)=',signif(anno$y,2)*100,'%')

ggplot(qcdata, aes(Samplename,`ASCAT SCA Ratio`, color=Status)) + geom_point() + facet_grid(~Status) + ylim(0,0.01) + geom_hline(yintercept = anno$y[1], linetype='dashed', color='grey39') + geom_label(data=anno, aes(x, y, label=paste(round(Ratio,2)*100,'% samples',sep='')), nudge_y = -0.0001) + 
  labs(title='BE Samples, zoomed to <1% SCA', subtitle=paste0('Median normal SCA=',signif(anno$y,2)*100,'%')) + theme(legend.position = 'none', axis.text.x=element_blank())

```
### Per Endoscopy

```{r}
m = median(normal$`ASCAT SCA Ratio`)

perPt = qcdata %>% group_by(PatientID,Endoscopy,Status) %>% dplyr::summarise(n=length(PatientID), LowSCA = length(which(`ASCAT SCA Ratio` <= m)) ) %>% mutate( LowRatio = LowSCA/n, quant=ntile(LowRatio,2) )

tb = perPt %>% group_by(Status) %>% 
  dplyr::summarise( '<0.5'=length(which(LowRatio<0.5))/length(Status), 
                    '≥0.5'=length(which(LowRatio>=0.5))/length(Status) )

ggplot(melt(tb), aes(variable, value, group=Status,fill=Status)) + geom_col(position=position_dodge()) + labs(x='Ratio of low sca endoscopies',y='Ratio of patients',title='')

```

### 2019-06-17 processing_pipeline/merge_qdnaseq_data.R

@@ -20,6 +20,8 @@ val.file = args[2]
 
 sheets = readxl::excel_sheets(val.file)
 
+sheets = sheets[1:10]
+
 all.val = do.call(bind_rows, lapply(sheets, function(s) {
   print(s)
   readxl::read_xlsx(val.file, s) %>% select(`Hospital Research ID`, matches('Status'), `Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA) %>% mutate_at(vars(`SLX-ID`), list(as.character))

### 2019-06-17 processing_pipeline/process_validation.R

@@ -15,9 +15,9 @@ data = args[1]
 val.file = args[2]
 outdir = args[3]
 
-#data = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/'
-#val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
-#outdir = '~/Data/BarrettsProgressionRisk/Analysis/validation'
+# data = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/'
+# val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
+# outdir = '~/Data/BarrettsProgressionRisk/Analysis/validation'
 
 patients = NULL
 if (length(args) == 4) patients = args[4]
@@ -30,7 +30,8 @@ if (!is.null(patients)) {
 }
 
 
-sheets = readxl::excel_sheets(val.file)
+sheets = readxl::excel_sheets(val.file)[1:10]
+
 
 all.val = do.call(bind_rows, lapply(sheets, function(s) {
   readxl::read_xlsx(val.file, s) %>% select(`Hospital Research ID`, matches('Status'), `Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA) %>% mutate_at(vars(`SLX-ID`), list(as.character))
@@ -60,6 +61,7 @@ failedQC = tibble()
 
 
 for (ra in levels(all.val$RA) ) {
+  print(ra)
   pts = all.val %>% filter(RA == ra) %>% select(`Hospital Research ID`) %>% unique() %>% pull  
   if (length(pts) <= 0) next
   
@@ -71,7 +73,8 @@ for (ra in levels(all.val$RA) ) {
     if (is.null(si[['Endoscopy']])) si = si %>% mutate(Endoscopy = '2019/01/01')
 
     plot.dir = paste(outdir, pid, 'plots',sep='/')
-    #if (length(list.files(plot.dir)) >= nrow( all.val %>% filter(`Hospital Research ID` == pid) )) next  # skip patients I've already done
+    #if (length(list.files(plot.dir)) >= nrow( all.val %>% filter(`Hospital Research ID` == pid) )) next  
+    if (file.exists(paste(dirname(plot.dir), paste0(which(levels(all.val$RA) == ra), '_segObj.Rdata'),sep='/'))) next # skip patients I've already done
 
[…]

### 2019-08-06 Paper.Rmd
...

@@
 library(CoxHD)
+library(RColorBrewer)
 #library(plyr)
 #library(dplyr)
+
+library(OACUtils)
 
@@
 ```{r, echo=F, message=F, fig.width=5, fig.height=5}
-pg.sampNOHGD = do.call(rbind, pg.sampNOHGD)
+#pg.sampNOHGD = do.call(rbind, pg.sampNOHGD)
 
@@
 
-p = ggplot(melt(perf), aes(x=variable,y=value, group=variable, fill=variable)) +
+p = ggplot(melt(perf, id.vars='pt'), aes(x=variable,y=value, group=variable, fill=variable)) +
     geom_boxplot(fill='lightcoral', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) +
     geom_jitter(color='grey39', width=0.3) + plot.theme + 
-    geom_label(data=melt(subset(perf, perf %in% range(perf) )), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
+    geom_label(data=melt(subset(perf, perf %in% range(perf) ), id.vars='pt'), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
     theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')
@@
 
-rm(plots, performance.at.1se,coefs, nzcoefs, fits)
+#rm(plots, performance.at.1se,coefs, nzcoefs, fits)
 ```
@@
 ```{r leaveoneout, echo=F, message=F, warning=F, fig.width=5, fig.height=5}
+
 file = paste(cache.dir, 'loo.Rdata', sep='/')
@@
 ```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=7, fig.width=7}
-preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
+preds = pg.samp %>% dplyr::select(Status, Prediction)
 roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
@@
 ```{r, echo=F, warning=F, message=F,fig.height=4, fig.width=4}
-pendo = do.call(rbind, pg.samp) %>% dplyr::group_by(Patient, Status, Hospital.Research.ID, PID ) %>% dplyr::summarise(
+pendo = pg.samp %>% dplyr::group_by(Patient, Status, Hospital.Research.ID, PID ) %>% dplyr::summarise(
   'max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(PID)
@@
 rocEndoM = pROC::roc(Status ~ max, data=pendo, ci=T, of='thresholds')
-pROC::coords(rocEndoM, 'best')
+pROC::coords(rocEndoM, 'best', transpose=T)
 rocEndoM$model = 'Per Endoscopy Max'
@@
 rocEndoA = pROC::roc(Status ~ avg, data=pendo, ci=T, of='thresholds')
-pROC::coords(rocEndoA, 'best')
+pROC::coords(rocEndoA, 'best', transpose=T)
 rocEndoA$model = 'Per Endoscopy Avg'
@@
 
-lastEndo = do.call(rbind, lapply(pg.samp, function(pt) {
- cbind(pt[1,c('Patient','Hospital.Research.ID', 'Status')],  'avg'=mean(subset(pt, PID == pt[nrow(pt), 'PID'])$Prediction))
-}))
+lastEndo = pg.samp %>% group_by(Patient, Hospital.Research.ID, Status,Endoscopy.Year, Path.ID) %>% arrange(desc(Endoscopy.Year), PID)  %>% dplyr::summarise(avg = mean(Prediction) ) %>% ungroup %>% group_by(Patient) %>% dplyr::filter( Endoscopy.Year == max(Endoscopy.Year) ) %>% dplyr::select(-Endoscopy.Year, -PID)
 
 rocLastEndo = pROC::roc(Status ~ avg, data=lastEndo, ci=T, of='thresholds')
-pROC::coords(rocLastEndo, 'best')
+pROC::coords(rocLastEndo, 'best', transpose=T)
 rocLastEndo$model = 'Last endoscopy'
@@
 
-firstEndo = do.call(rbind, lapply(pg.samp, function(pt) {
- cbind(pt[1,c('Patient','Hospital.Research.ID', 'Status')],  'avg'=mean(subset(pt, PID == pt[1, 'PID'])$Prediction))
-}))
+firstEndo = pg.samp %>% group_by(Patient, Hospital.Research.ID, Status,Endoscopy.Year, PID) %>% arrange(Endoscopy.Year, PID)  %>% dplyr::summarise(avg = mean(Prediction) ) %>% ungroup %>% group_by(Patient) %>% dplyr::filter( Endoscopy.Year == min(Endoscopy.Year) ) %>% dplyr::select(-Endoscopy.Year, -PID)
 
 rocFirstEndo = pROC::roc(Status ~ avg, data=firstEndo, ci=T, of='thresholds')
-pROC::coords(rocFirstEndo, 'best')
+pROC::coords(rocFirstEndo, 'best', transpose=T)
 rocFirstEndo$model = 'First endoscopy'
@@
 
-ptendo = do.call(rbind, lapply(pg.samp, function(pt) {
- cbind(pt[1,c('Patient','Hospital.Research.ID', 'Status')],  'mean'=mean(subset(pt, PID != pt[nrow(pt), 'PID'])$Prediction), 'max'=max(subset(pt, PID != pt[nrow(pt), 'PID'])$Prediction))
-}))
-
-ptendo = ptendo[-which(is.na(ptendo$mean)),]
+# Excluding the final endoscopy
+ptendo = do.call(bind_rows, lapply(unique(pg.samp$Hospital.Research.ID), function(pt) {
+  x = pg.samp %>% filter(Hospital.Research.ID == pt)  
+  x %>% filter(Endoscopy.Year != max(x$Endoscopy.Year))
+})) %>% group_by(Patient, Hospital.Research.ID, Status) %>% summarise_at(vars(Prediction), funs(max, mean), na.rm=T)
+
+ptendo = ptendo %>% filter(!is.na(mean))
 
 rocPtM = pROC::roc(Status ~ max, data=ptendo, ci=T, of='thresholds')
-pROC::coords(rocPtM, 'best')
+pROC::coords(rocPtM, 'best', transpose=T)
 rocPtM$model = 'Per Patient Max'
@@
 rocPtA = pROC::roc(Status ~ mean, data=ptendo, ci=T, of='thresholds')
-pROC::coords(rocPtA, 'best')
+pROC::coords(rocPtA, 'best', transpose=T))
 #roc.plot(rocPtA)
@@
 transform.by.time<-function(pt, info) {
-  #print(unique(pt$Hospital.Research.ID))
-  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction', 'OR', 'Sample.Mean','Sample.SD','Sample.Var')] %>% dplyr::group_by(PID, Samplename) %>% dplyr::mutate(
-  'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12)
-
-  pt = dplyr::arrange(merge(pt, subset(info, Hospital.Research.ID == pt$Hospital.Research.ID[1], select=c('Samplename','Block')), by='Samplename'), Pathology, PID, Endoscopy.Year)
+  #pt = left_join(pt, info %>% dplyr::select(Hospital.Research.ID, Patient,Samplename,Block), by=c('Hospital.Research.ID','Patient','Samplename'))
+
+  pt = pt %>% dplyr::group_by(PID, Samplename) %>% dplyr::mutate( 'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12) %>% arrange(Pathology, PID, Endoscopy.Year) %>% ungroup
 
@@
   }
-
   pt$nl = (pt$Block-1)*2
@@
 
-back = do.call(rbind.data.frame,lapply(pg.samp, transform.by.time, info=patient.info))
-back = merge(back, patient.info[,c('Samplename','p53.Status', 'Total.Reads')])
+back = do.call(bind_rows,lapply(unique(pg.samp$Patient) %>% map( function(p) pg.samp %>% filter(Patient == p)  ), transform.by.time, info=patient.info))
+#back = merge(back, patient.info[,c('Samplename','p53.Status', 'Total.Reads')])
 back$Pathology = factor(back$Pathology, levels=c('BE','ID','LGD','HGD','IMC'), ordered = T)
@@
 rsid = rstandard(pcfit)
-pred.confidence[which( rsid >= 1 |  rsid <= -1 ), 'Risk'] = 'Moderate'
+#pred.confidence[which( rsid >= 1 |  rsid <= -1 ), 'Risk'] = 'Moderate'
+pred.confidence = pred.confidence %>% mutate(Risk = as.character(ifelse(p.value > 0.05, 'Moderate', Risk)))
 
@@
 pred.confidence$Risk = factor(pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)
+
+back = left_join(back, pred.confidence %>% dplyr::select(quants, Risk), by='quants')
+
 ```
@@
 
-unadjRR = ggplot(back, aes(OR)) + geom_histogram(aes(fill=..x..), bins=10, show.legend = F) +
+unadjRR = ggplot(back, aes(RR)) + geom_histogram(aes(fill=..x..), bins=10, show.legend = F) +
   scale_fill_gradientn(colors = myPal,  name='') + 
@@
 #df = subset(globalCNdf, Patient == pt)
-df = merge(globalCNdf, back[,c('Samplename','Endoscopy.Year', 'ogj', 'months.before.final', 'OR','Prediction')], by.x='variable', by.y='Samplename')
+df = merge(globalCNdf, back[,c('Samplename','Endoscopy.Year', 'ogj', 'months.before.final', 'RR','Prediction')], by.x='variable', by.y='Samplename')
 df = arrange(df, chr, start, Endoscopy.Year)
@@
   'max.Pred'=max(Prediction),
-  'max.RR'=max(OR),
+  'max.RR'=max(RR),
   'max.Path'=max(Pathology),
@@
   'ratio'=length(which(seg.type != 'arms' & cn %in% c('loss','gain')))/length(cn)
-  )
+  ) %>% mutate(Patient = factor(Patient))
 #as.data.frame(df2)
@@
 
-el = subset(df2, early | late)
+el = df2 %>% filter(early | late)
 el2 = el %>% dplyr::group_by(Status, Patient) %>% dplyr::summarise( diff(arms), diff(CNA) )
@@
 
-p = ggplot(subset(df2, early|late), aes(Endoscopy.Year, arms, color=Patient)) + facet_grid(~Status) +
+p = ggplot(df2 %>% filter(early|late), aes(Endoscopy.Year, arms, color=Patient)) + facet_grid(~Status) +
   geom_point() +  geom_line(data=subset(df2, early | late)) + theme(legend.position = 'none', text=element_text(size=11)) + 
@@
 ```{r jacknife, echo=T, warning=F, message=F, fig.height=10, fig.width=12}
-coefs.per.pt = data.frame(matrix(ncol=length(pg.samp), nrow=nrow(ch), dimnames=list( ch$label, names(pg.samp))))
-for (pt in names(pg.samp)) {
+
+list.files(cache.dir)
+
+load(paste0(cache.dir, '/all.pt.alpha.Rdata'), verbose=T)
+rm(plots, performance.at.1se, dysplasia.df, models, cvs, labels)
+
+coefs.per.pt = data.frame(matrix(ncol=length(unique(pg.samp$Patient)), nrow=nrow(ch), dimnames=list( ch$label, unique(pg.samp$Hospital.Research.ID))))
+
+for (pt in unique(pg.samp$Hospital.Research.ID)) {
   lcf = loo.coefs[[pt]][,1]
@@
 
-nzcoefs = lapply(nzcoefs, function(x) {
-  as_tibble(x, rownames='coef')
-})
+nzcoefs = purrr::map(nzcoefs, as_tibble, rownames='coef')
 
@@
 }
-colnames(loo.coefs)[-1] = c(1:length(nzcoefs))
+colnames(loo.coefs)[-1] = names(nzcoefs)
 loo.coefs[is.na(loo.coefs)] = 0
@@
   plot.theme + labs(x='', title='Jackknifed coefficient CI')
-
-ggplot(ch, aes(reorder(label, coef), coef)) + geom_point() + 
-  geom_errorbar(aes(ymin=coef-jack.se*1.96, ymax=coef+jack.se*1.96)) + 
-  coord_flip() +
-  plot.theme + labs(x='', title='Jackknifed coefficient CI')
-
 ```
@@
 
-p = pred.tiles(subset(back, Patient %in% c(6)), ncol=1, cols=myPal, probs=T) + labs(title='')
+p = pred.tiles(back %>% filter(Patient == 6), colors=myPal, probs=T, ncol=1) + labs(title='')
 tilesLegend = get.legend(p)
@@
 ga = grid.arrange(
-  pred.tiles(subset(back, Patient %in% c(13,20,34,56,60,48) & Status == 'P'), ncol=1, cols=myPal, probs=T) + labs(title='Progressors') + theme(legend.position = 'none'),
-  pred.tiles(subset(back, Patient %in% c(15,16,21,3,45,19) & Status == 'NP'), ncol=1, cols=myPal, probs=T) + labs(title='Non-Progressors') + theme(legend.position = 'none'),
+  pred.tiles(back %>% filter(Patient %in% c(13,20,34,56,60,48) & Status == 'P'), ncol=1, colors=myPal, probs=T) + labs(title='Progressors') + theme(legend.position = 'none'),
+  pred.tiles(subset(back, Patient %in% c(15,16,21,3,45,19) & Status == 'NP'), ncol=1, colors=myPal, probs=T) + labs(title='Non-Progressors') + theme(legend.position = 'none'),
 ncol=2)
-ga
 
@@
 
-p1 = pred.tiles(subset(back, Status == 'NP'), cols=myPal, probs=T, nrow=2) + labs(title='Non-Progressors') 
+p1 = pred.tiles(back %>% filter(Status == 'NP'), colors=myPal, probs=T, nrow=2) + labs(title='Non-Progressors') 
 riskLegend = get.legend(p1)
@@
 p1 = p1 + theme(legend.position = 'none')
-p2 = pred.tiles(subset(back, Status == 'P'), probs=T, cols=myPal,nrow=2) + labs(title='Progressors') + theme(legend.position = 'none')
+p2 = pred.tiles(back %>% filter(Status == 'P'), probs=T, colors=myPal,nrow=2) + labs(title='Progressors') + theme(legend.position = 'none')
 

### 2019-08-06 Validation Report.Rmd
...

@@
 
-sheets = readxl::excel_sheets(val.file)[1:10]
+sheets = readxl::excel_sheets(val.file)[1:12]
 all.val = do.call(bind_rows, lapply(sheets, function(s) {
-  readxl::read_xlsx(val.file, s) %>% select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA, matches('Collection')) %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% fncols('Collection', 'Biopsy')
+  readxl::read_xlsx(val.file, s) %>% select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA, matches('Collection')) %>% dplyr::filter(!is.na(`SLX-ID`)) %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% fncols('Collection', 'Biopsy')
 }))
@@
 
-sval = val %>% group_by(Batch) %>% dplyr::mutate(Cohort = Cohort[1]) %>% group_by(Batch, Cohort) %>% dplyr::summarise_at(vars(varMAD_median), funs(min,mean,median,max),na.rm=T)
+sval = val %>% group_by(Batch) %>% dplyr::mutate(Cohort = paste(unique(Cohort), collapse=', ' )) %>% group_by(Batch, Cohort) %>% dplyr::summarise_at(vars(varMAD_median), funs(min,mean,median,max),na.rm=T)
 
-vt = val %>% group_by(Batch) %>% dplyr::mutate(Cohort = Cohort[1]) %>% group_by(Batch, `Sample Type`, Cohort) %>% dplyr::summarise(
+vt = val %>% group_by(Batch) %>% dplyr::mutate(Cohort =  paste(unique(Cohort), collapse=', ' )) %>% group_by(Batch, `Sample Type`, Cohort) %>% dplyr::summarise(
   `Total Samples` = length(Batch),
@@
   Pass=length(which(varMAD_median<=qc.cutoff)), Fail=length(which(varMAD_median>qc.cutoff))
-) %>% rename('as.integer(RA)' = 'Processing') %>% kable(caption='Pass/Fail by sample type') %>% kable_styling('striped')
+) %>% rename('as.integer(RA)' = 'Processing RA') %>% kable(caption='Pass/Fail by sample type') %>% kable_styling('striped')
 
@@
 
-oldp = val %>% ungroup %>% filter(Batch %in% c(9))
+oldp = val %>% ungroup %>% filter(Batch %in% c(8:12) & Cohort == 'Old progressor scrolls')
 
 oldp %>% dplyr::group_by(Batch, `Sample Type`, `Cohort`) %>% dplyr::summarise(
-  Total = n(),
+  `n Samples` = n(),
   `% Passed QC` = round((length(which(varMAD_median<=qc.cutoff))/n())*100,1)
@@
 
+sheets = readxl::excel_sheets(val.file)[9:12]
+info = do.call(bind_rows, lapply(sheets, function(s) {
+  readxl::read_xlsx(val.file, s, trim_ws = T) %>% 
+    dplyr::filter(Status == 'P') %>% dplyr::select(`Hospital Research ID`, `Block ID`, Endoscopy, Pathology, `SLX-ID`, `Index Sequence`, `Path Notes`) %>% mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`))
+}))
+
 if (file.exists('oldp_preds.txt')) {
@@
 
-  info = readxl::read_xlsx(val.file, 'Batch 9') %>% dplyr::select(`Hospital Research ID`, `Block ID`, Endoscopy, Pathology, `SLX-ID`, `Index Sequence`, `Path Notes`) %>% mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`))
+#  info = readxl::read_xlsx(val.file, 'Batch 9') %>% dplyr::select(`Hospital Research ID`, `Block ID`, Endoscopy, Pathology, `SLX-ID`, `Index Sequence`, `Path Notes`) %>% mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`))
 
@@
 
-new.np = val %>% ungroup %>% filter(Batch %in% c(10))
+new.np = val %>% ungroup %>% filter(Batch %in% c(10:12) & Status == 'NP')
 
@@
 
-Again, all passed QC though on manual QC I might have failed 2-3.  There were 3 samples that just squeaked in under the cutoff threshold.  But they did.
+Again, most passed QC though on manual QC I might have failed 2-3 others.  There were 3 samples that just squeaked in under the cutoff threshold.
 
@@
 
-if (file.exists('batch_10.txt')) {
-  new.np.preds = read_tsv('batch_10.txt')
+sheets = readxl::excel_sheets(val.file)[10:12]
+info = do.call(bind_rows, lapply(sheets, function(s) {
+  readxl::read_xlsx(val.file, s, trim_ws = T) %>% 
+    dplyr::filter(Status == 'NP' & !is.na(`SLX-ID`)) %>% dplyr::select(`Hospital Research ID`, `Block ID`, Endoscopy, Pathology, `SLX-ID`, `Index Sequence`, `Path Notes`) %>% mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`))
+}))
+
+if (file.exists('val_np.txt')) {
+  new.np.preds = read_tsv('val_np.txt')
 } else {
-
-  segFiles = grep(paste(unique(new.np$`Patient ID`),collapse='|'), list.files('~/Data/BarrettsProgressionRisk/Analysis/validation', '2_segObj',  full.names = T, recursive = T), value=T)
-
-  info = readxl::read_xlsx(val.file, 'Batch 10') %>% dplyr::select(`Hospital Research ID`, `Block ID`, Endoscopy, Pathology, `SLX-ID`, `Index Sequence`) %>% mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`))
+  segFiles = grep(paste(new.np$`Patient ID`,collapse='|'), list.files('~/Data/BarrettsProgressionRisk/Analysis/validation', '(2|3)_segObj',  full.names = T, recursive = T), value=T)
 
   new.np.preds = do.call(bind_rows, lapply(segFiles, function(f) {
+    print(f)
     load(f)  
+    if (nrow(BarrettsProgressionRisk::sampleResiduals(segmented) %>% dplyr::filter(Pass)) <= 0) return()
+
     segmented$sample.info = BarrettsProgressionRisk::loadSampleInformation(info %>% filter(Sample %in% segmented$sample.info$Sample) )
-
     prr = BarrettsProgressionRisk::predictRiskFromSegments(segmented,F)
@@
   }))
-  write_tsv(new.np.preds, 'batch_10.txt')
+  write_tsv(new.np.preds, 'val_np.txt')
 }
@@
 
-new.np.preds %>% group_by(`Hospital Research ID`) %>% dplyr::summarise(
+val.new.np.preds = left_join(val %>% filter(sample %in% new.np.preds$Sample) %>% select(`Hospital Research ID`, sample, varMAD_median, n.segs, Status),
+          new.np.preds %>% select(Sample, Probability, matches('Risk')), by=c('sample'='Sample'))
+
+with(val.new.np.preds, cor.test(varMAD_median, Probability))
+
+val.new.np.preds %>% group_by(`Hospital Research ID`) %>% dplyr::summarise(
   `n samples` = n(), 
@@
   'Mod. Risk' = length(which(Risk == 'Moderate')),
-  'Low Risk' =   length(which(Risk == 'Low'))
+  'Low Risk' =   length(which(Risk == 'Low')),
+  'QC var' = signif( mean(varMAD_median), 2)
 ) %>% kable(caption = 'Classified predictions per patient (all NP)') %>% kable_styling('striped')

### 2019-08-06 sWGS QC.Rmd
...

@@
 
-library(pander)
-library(dplyr)
-library(ggplot2)
+library(tidyverse)
 library(gridExtra)
+library(kableExtra)
 
-dir1 = '~/Data/Ellie/Analysis/multipcf_plots_fitted_perPatient'
-dir2 = '~/Data/Ellie/Analysis/val_test/'
+dir1 = '~/Data/Ellie/Analysis/multipcf_perPatient/'
+dir2 = '~/Data/Ellie/Analysis/validation//'
 
@@
 ```{r, echo=F}
-res.variance = read.table(paste(dir1,'resid_var.txt',sep='/'),header=T)
+
+res.variance = do.call(bind_rows, lapply(list.files(dir1, 'residuals', recursive = T, full.names = T), function(f) {
+  read_tsv(f, col_types = 'ccdddddl')
+}))
+
 #pander(head(res.variance[,grep('var1sd',colnames(res.variance))]), caption='Residuals across segments per sample')
-pander(head(res.variance[,grep('varMAD',colnames(res.variance))]), caption='Residuals across segments per sample')
+kable(head(res.variance), caption='Example residuals across segments per sample') %>% kable_styling('striped')
 ```
@@
 ggplot(res.variance, aes(varMAD_median)) + geom_histogram(bins=50,color='lightblue') + 
-  geom_vline(xintercept = cutoff, linetype='dashed', color='grey') + labs(x='var( seg<MAD(seg) | seg>-MAD(seg)  )') +
-  theme_bw()
+  geom_vline(xintercept = cutoff, linetype='dashed', color='grey') + labs(x='var( seg<MAD(seg) | seg>-MAD(seg)  )') + theme_bw()
 ```
@@
 ```{r, echo=F}
-pander( arrange(subset(res.variance, round(varMAD_median,3) > cutoff, select=c(pid,sample, varMAD_median,varMAD_sd, varMAD_Q1,varMAD_Q3)), pid, varMAD_median) )
+res.variance %>% filter( round(varMAD_median,3) > cutoff ) %>% select(patient,sample, varMAD_median) %>% kable(caption=paste0('Samples variance > ',cutoff)) %>% kable_styling('striped')
 
-failedQC = arrange(subset(res.variance, round(varMAD_median,3) > cutoff, select=c(pid,sample, varMAD_median,varMAD_sd, varMAD_Q1,varMAD_Q3)), pid, varMAD_median)
-
+failedQC = res.variance %>% filter( round(varMAD_median,3) > cutoff ) 
 ```
@@
 
-Cutoff from above dataset applied to the validation set.
+Cutoff from above dataset applied to the initial validation set.
 
 ```{r, echo=F}
-list.files(dir2
+val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
+sheets = readxl::excel_sheets(val.file)[1:4]
+all.val = do.call(bind_rows, lapply(sheets, function(s) {
+  readxl::read_xlsx(val.file, s) %>% select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA, matches('Collection')) %>% dplyr::filter(!is.na(`SLX-ID`) & RA == 'C.Kosmidou') %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character))
+}))
+all.val = all.val %>% select(-Collection)
 
-resVal = read.table(paste(dir2,'resid_var.txt',sep='/'),header=T)
-pander(head(resVal[,grep('var1',colnames(resVal))]), caption='Residuals across segments per sample')
+pastefun<-function(x) {
+  if ( !grepl('SLX-', x) ) x = paste0('SLX-',x)
+  return(x)
+}
 
+all.val = all.val %>% rowwise %>% mutate_at(vars(`SLX-ID`), list(pastefun) ) %>% ungroup
+all.val = all.val %>% mutate(
+  `Hospital Research ID` = str_replace_all( str_remove_all(`Hospital Research ID`, " "), '/', '_'), 
+  `Index Sequence` = str_replace_all(`Index Sequence`, 'tp', ''),
+  Sample = paste(`SLX-ID`,`Index Sequence`,sep='.')
+  )
+
+resVal = do.call(bind_rows, lapply(list.files(dir2, 'residuals', recursive = T, full.names = T), function(f) {
+  readr::read_tsv(f, col_types = 'cdddddl', col_names = F) %>% set_names('sample', 'varMAD_median', 'varMAD_sd', 'varMAD_Q1', 'varMAD_Q3', 'n.segs', 'Pass') %>%  
+    mutate(`Patient ID` = as.character(basename(dirname(f))))
+})) %>% select(`Patient ID`, sample, matches('varMad'), n.segs, Pass)
+
+resVal = resVal %>% filter(sample %in% all.val$Sample)
+
+resVal %>% kable(caption='Residuals across segments per sample') %>% kable_styling('striped')
 resVal = arrange(resVal,varMAD_median)
@@
 
-failedQCVal = arrange(subset(resVal, round(varMAD_median,3) > cutoff, select=c(pid,sample, varMAD_median,varMAD_sd, varMAD_Q1,varMAD_Q3)), pid, varMAD_median)
+failedQCVal = resVal %>% filter(varMAD_median > cutoff) %>% arrange(`Patient ID`, varMAD_median)
 
-setdiff(as.character(resVal$sample),as.character(failedQCVal$sample))
+#setdiff(as.character(resVal$sample),as.character(failedQCVal$sample))
 
@@
 ```{r}
-res.variance[c('pid','sample')] = lapply(res.variance[c('pid','sample')], as.character)
-resVal[c('pid','sample')] = lapply(resVal[c('pid','sample')], as.character)
+res.variance = res.variance %>% mutate_at(vars(patient, sample), as.character)
+resVal = resVal %>% dplyr::rename( patient = `Patient ID` ) %>% mutate_at(vars(patient, sample), as.character)
 
@@
 
-ar = rbind(res.variance, resVal)
-ar$cohort = factor(ar$cohort)
+ar = bind_rows( res.variance, resVal ) %>% mutate(cohort = factor(cohort))
 
@@
 
-fc = arrange(subset(ar, round(varMAD_median,3) > newcutoff, select=c(pid,sample, varMAD_median,varMAD_sd, varMAD_Q1,varMAD_Q3)), pid, varMAD_median)
-
+fc = ar %>% filter(round(varMAD_median,3) > newcutoff) %>% select(patient, sample, varMAD_median,varMAD_sd, varMAD_Q1,varMAD_Q3) %>% arrange(patient, varMAD_median)
 ```

### 2019-08-06 processing_pipeline/process_validation.R

@@ -30,11 +30,11 @@ if (!is.null(patients)) {
 }
 
 
-sheets = readxl::excel_sheets(val.file)[1:10]
+sheets = readxl::excel_sheets(val.file)[1:12]
 
 
 all.val = do.call(bind_rows, lapply(sheets, function(s) {
-  readxl::read_xlsx(val.file, s) %>% select(`Hospital Research ID`, matches('Status'), `Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA) %>% mutate_at(vars(`SLX-ID`), list(as.character))
+  readxl::read_xlsx(val.file, s) %>% select(`Hospital Research ID`, matches('Status'), `Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA) %>% mutate_at(vars(`SLX-ID`), list(as.character)) %>% dplyr::filter(!is.na(`SLX-ID`))
 }))
 
 pastefun<-function(x) {
@@ -58,8 +58,6 @@ if (!is.null(patients))
 failedQC = tibble()
 
 # Need to process the batches from CK and SA separately even if the patients overlap
-
-
 for (ra in levels(all.val$RA) ) {
   print(ra)
   pts = all.val %>% filter(RA == ra) %>% select(`Hospital Research ID`) %>% unique() %>% pull  

### 2019-08-07 Validation Report.Rmd
...

@@
 all.val = do.call(bind_rows, lapply(sheets, function(s) {
-  readxl::read_xlsx(val.file, s) %>% select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA, matches('Collection')) %>% dplyr::filter(!is.na(`SLX-ID`)) %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% fncols('Collection', 'Biopsy')
+  readxl::read_xlsx(val.file, s) %>% dplyr::select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA, matches('Collection')) %>% dplyr::filter(!is.na(`SLX-ID`)) %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% fncols('Collection', 'Biopsy')
 }))
@@
     mutate(`Patient ID` = as.character(basename(dirname(f))))
-})) %>% select(`Patient ID`, sample, varMAD_median, n.segs) %>% mutate(varMAD_median = round(varMAD_median, 3)) %>% mutate(sample = str_replace(sample, '_', '-'))
+})) %>% dplyr::select(`Patient ID`, sample, varMAD_median, n.segs) %>% mutate(varMAD_median = round(varMAD_median, 3)) %>% mutate(sample = str_replace(sample, '_', '-'))
 
@@
 
-val = val %>% filter(sample %in% (all.val %>% select(Samplename) %>% pull) ) %>% 
+val = val %>% filter(sample %in% (all.val %>% dplyr::select(Samplename) %>% pull) ) %>% 
           left_join( all.val, val %>% mutate(sample = str_replace(sample, '_', '-')), by=c('sample'='Samplename'))
@@
 
-speedvac = left_join(val %>% ungroup %>% filter(Batch %in% c(5)), readxl::read_xlsx(val.file, sheet='Batch 5') %>% select(`Hospital Research ID`, `Block ID`, `Original Cohort`, `Speedvac`))
+speedvac = left_join(val %>% ungroup %>% filter(Batch %in% c(5)), readxl::read_xlsx(val.file, sheet='Batch 5') %>% dplyr::select(`Hospital Research ID`, `Block ID`, `Original Cohort`, `Speedvac`))
 
@@
 
+if (file.exists('batch8_preds.txt')) {
+  batch8_preds = read_tsv('batch8_preds.txt')
+} else {
+  segFiles = grep(paste(on$`Patient ID`,collapse='|'), list.files('~/Data/BarrettsProgressionRisk/Analysis/validation', 'segObj',  full.names = T, recursive = T), value=T)
+
+  batch8_preds = do.call(bind_rows, lapply(segFiles, function(f) {
+    print(f)
+    load(f)
+    if (nrow(sampleResiduals(segmented)) <= 0) return()
+
+    segmented$sample.info = loadSampleInformation(  
+      on %>% filter(sample %in% segmented$sample.info$Sample) %>%
+        dplyr::rename('Endoscopy' = `Sample Date`, Sample = sample ) )
+
+    prr = BarrettsProgressionRisk::predictRiskFromSegments(segmented, F)
+    BarrettsProgressionRisk::predictions(prr,'sample')
+  }))
+  batch8_preds = batch8_preds %>% filter(!is.na(`Patient ID`))
+
+  write_tsv(batch8_preds, 'batch8_preds.txt')
+}
+
+batch8_preds %>% dplyr::select(`Patient ID`, Sample, everything(), -matches('Hospital|SLX|Index') ) %>% kable() %>% kable_styling('striped')
+
+batch8_preds %>% group_by(Risk) %>% dplyr::summarise('n Samples'= length(Risk)) %>% kable(caption='Predictions') %>% kable_styling(full_width = F)
+
 ```
@@
   readxl::read_xlsx(val.file, s, trim_ws = T) %>% 
-    dplyr::filter(Status == 'P') %>% dplyr::select(`Hospital Research ID`, `Block ID`, Endoscopy, Pathology, `SLX-ID`, `Index Sequence`, `Path Notes`) %>% mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`))
+    dplyr::filter(Status == 'P') %>% dplyr::select(`Hospital Research ID`, `Block ID`, Endoscopy, Pathology, `SLX-ID`, `Index Sequence`, `Path Notes`) %>% dplyr::mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`))
 }))
@@
 
-oldp.preds = oldp.preds %>% filter(Sample %in% (oldp %>% filter(!grepl('OCCAMS',`Hospital Research ID`)) %>% select(sample) %>% pull))
+oldp.preds = left_join(oldp.preds, val %>% dplyr::select(sample, varMAD_median, n.segs), by=c('Sample'='sample'))
+
+oldp.preds = oldp.preds %>% filter(Sample %in% (oldp %>% filter(!grepl('OCCAMS',`Hospital Research ID`)) %>% dplyr::select(sample) %>% pull))
 
@@
 ```{r}
-
 sheets = readxl::excel_sheets(val.file)[10:12]
@@
 
+new.np.preds = left_join(new.np.preds, val %>% dplyr::select(sample, varMAD_median, n.segs), by=c('Sample'='sample'))
+
+#cor.test(new.np.preds$n.segs, new.np.preds$varMAD_median)
+
 new.np.preds %>% dplyr::summarise(
@@
 
-val.new.np.preds = left_join(val %>% filter(sample %in% new.np.preds$Sample) %>% select(`Hospital Research ID`, sample, varMAD_median, n.segs, Status),
-          new.np.preds %>% select(Sample, Probability, matches('Risk')), by=c('sample'='Sample'))
+val.new.np.preds = left_join(val %>% filter(sample %in% new.np.preds$Sample) %>% dplyr::select(`Hospital Research ID`, sample, varMAD_median, n.segs, Status),
+          new.np.preds %>% dplyr::select(Sample, Probability, matches('Risk')), by=c('sample'='Sample'))
 

### 2019-08-14 Insilico Splits.Rmd
...

@@
 
-cache.dir = paste(data, 'Analysis/5e6_arms_all', sep='/')
+cache.dir = paste(data, 'Analysis/5e6_arms', sep='/')
 split.dir = paste(data, 'Analysis/5e6_arms_splits', sep='/')
@@
 
-predictions = do.call(rbind, pg.samp)
+predictions = pg.samp
 
-roc = pROC::roc(Status ~ Prediction, data=predictions, auc=T, ci=T, of='thresholds')
+roc = pROC::roc(Status ~ Prediction, data=predictions, auc=T, ci=T, of='thresholds', transpose = T)
 roc$model = 'discovery'
@@
 mrplot <- function(rocList) {
-  aucs = do.call(rbind, lapply(rocList,function(r) 
-    cbind.data.frame('AUC'=r$auc*100,'model'=r$model,'Specificity'=pROC::coords(r,'best')[['specificity']], 'Sensitivity'=pROC::coords(r,'best')[['sensitivity']])))
+  aucs = do.call(bind_rows, lapply(names(rocList),function(nm) { 
+    r = rocList[[nm]]
+    tibble( AUC = r$auc*100, model = r$model, 
+      'Specificity'= pROC::coords(r,'best', ret='specificity',transpose=T)[1],
+      'Sensitivity'= pROC::coords(r,'best', ret='sensitivity',transpose=T)[1] )
+  }))
+
   aucs$y = aucs$Sensitivity
@@
 
-  df = do.call(rbind, lapply(rocList, function(r) 
-    cbind.data.frame('Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'model'=r$model, 'is'=r$is) 
-  ))
+  df = do.call(bind_rows, lapply(names(rocList),function(nm) { 
+  #  print(nm)
+    r = rocList[[nm]]
+    tibble('Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'model'=r$model, 'is'=as.character(r$is) )
+  }))
 
-  ggplot(df, aes(Specificity, Sensitivity,color=is)) + geom_line(size=1) + 
+  ggplot(df, aes(Specificity, Sensitivity,color=is)) + geom_line(color='lightblue') + 
     geom_line(data=subset(df, model == 'discovery'), color='red', size=1) +
     scale_x_reverse() +
-    scale_color_brewer(palette='Blues') +
+    #scale_color_brewer(palette='Blues') +
     labs(title='', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +

### 2019-08-14 Paper.Rmd
...

@@
 
-patient.info = read.patient.info(patient.file, demo.file, set='All')$info
+all.patient.info = read.patient.info(patient.file, demo.file, set='All')$info
+
+patient.info = all.patient.info %>% dplyr::filter(Set == 'Training')
 head(patient.info)
@@
 patient.info = arrange(patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
-sum.patient.data = summarise.patient.info(patient.info)
+sum.patient.data = as_tibble(summarise.patient.info(patient.info))
 
@@
 ```{r labelsPNP, echo=F, message=F, include=T}
-cache.dir = paste(data, 'Analysis/5e6_arms_all', sep='/') # QC'd
+cache.dir = paste(data, 'Analysis/5e6_arms', sep='/') # QC'd
 
@@
 
-chr.lengths = get.chr.lengths()
-chr.lengths$chrom = sub('chr','',chr.lengths$chrom)
-chr.lengths = subset(chr.lengths, chrom %in% c(1:22))
-chr.lengths$chrom = factor(chr.lengths$chrom, levels=c(1:22), ordered = T)
+chr.lengths = BarrettsProgressionRisk:::chrInfo(prefix = '') %>% dplyr::filter(chr %in% c(1:22)) %>%
+  dplyr::mutate(chrom = factor(sub('chr','', chrom), levels=c(1:22), ordered=T))
+#chr.lengths$chrom = sub('chr','',chr.lengths$chrom)
+#chr.lengths = subset(chr.lengths, chrom %in% c(1:22))
+#chr.lengths$chrom = factor(chr.lengths$chrom, levels=c(1:22), ordered = T)
 
@@
   #pt = left_join(pt, info %>% dplyr::select(Hospital.Research.ID, Patient,Samplename,Block), by=c('Hospital.Research.ID','Patient','Samplename'))
-
   pt = pt %>% dplyr::group_by(PID, Samplename) %>% dplyr::mutate( 'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12) %>% arrange(Pathology, PID, Endoscopy.Year) %>% ungroup
@@
 
-back$quants = with(back, cut(Prediction, breaks=cuts))
+back$quants = with(back, cut(Prediction, breaks=cuts, include.lowest = T))
 qt = as.data.frame.matrix(table(back$quants, back$Status))
@@
 #pred.confidence[which( rsid >= 1 |  rsid <= -1 ), 'Risk'] = 'Moderate'
-pred.confidence = pred.confidence %>% mutate(Risk = as.character(ifelse(p.value > 0.05, 'Moderate', Risk)))
+pred.confidence = pred.confidence %>% mutate(Risk = as.character(ifelse(p.value > 0.05 & r1 < 0.6, 'Moderate', Risk)))
 
@@
 ggsave(filename='plots/pred_calib.png', plot=pcal, width=4, height=4, units='in', dpi=300)
+
+back.per.endo = back %>% group_by(Hospital.Research.ID, Status, Patient, PID) %>% 
+  dplyr::mutate('n.samples'=length(PID),'Samples'=paste(Samplename,collapse=',')) %>% 
+  dplyr::summarise_at(vars(matches('Pred|RR|Year')), list(max)) %>% dplyr::select(-matches('^Sample$')) %>% 
+  arrange(Patient, Endoscopy.Year, PID) 
+back.per.endo = back.per.endo %>% mutate(quants = cut(Prediction, breaks=cuts, include.lowest = T)) %>% left_join(pred.confidence %>% dplyr::select(quants, Risk), by='quants') %>% dplyr::select(-matches('quant')) %>% dplyr::mutate(Risk = as.character(Risk))
+
+back %>% group_by(Status, Risk)  %>% dplyr::summarise(n = length(Risk)) %>% spread(Risk, n) %>% rowwise %>% 
+  dplyr::mutate(n=sum(High,Low,Moderate), Low = Low/n, Moderate = Moderate/n, High = High/n) %>%
+  mutate_if(is.double, list(round),2) %>%
+  kableExtra::kable(caption='Training set per-sample predictions') %>% kableExtra::kable_styling(full_width = F)
+
+back.per.endo %>%  group_by(Status, Risk)  %>% dplyr::summarise(n = length(Risk)) %>% spread(Risk, n) %>% rowwise %>% 
+  dplyr::mutate(n = sum(High,Low,Moderate), Low = Low/n, Moderate = Moderate/n, High = High/n) %>% mutate_if(is.double, list(round),2) %>%
+  kableExtra::kable(caption='Training set per-endoscopy predictions') %>% kableExtra::kable_styling(full_width = F)
 ```
@@
 
-list.files(cache.dir)
+#list.files(cache.dir)
 
@@
 
+summary.t %>% kableExtra::kable() %>% kableExtra::kable_styling('striped', full_width = F)
+
 write.table(summary.t, quote=F, sep='\t', row.names = F)
 ```
-
-`r pander(as.data.frame(summary.t))`
 
@@
 ```{r val_setup, echo=F, message=F, warning=F, include=F}
-validation.patient.info =  subset(all.patient.info, Set == 'Test')
-validation.patient.info = arrange(validation.patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
+validation.patient.info =  all.patient.info %>% filter(Set == 'Test') %>% arrange(Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
 
@@
 
-cleaned = list.files(path=paste(data, 'Cleaned', sep='/'), pattern='tiled', full.names=T, recursive=T)
-cleaned = grep(paste(unique(validation.patient.info$Hospital.Research.ID), collapse='|'), cleaned, value=T)
-
-raw = list.files(path=paste(data, 'Cleaned', sep='/'), pattern='raw', full.names=T, recursive=T)
-raw = grep(paste(unique(validation.patient.info$Hospital.Research.ID), collapse='|'), raw, value=T)
+predicted = '~/Data/BarrettsProgressionRisk/Analysis/training/val_predictions'
+
+vpd = do.call(bind_rows, lapply(list.files(predicted, full.names = T), function(f) {
+  read_tsv(f, col_types = 'cddcccdccdc')
+})) %>% dplyr::select(Hospital.Research.ID, everything())
+
+vpd = left_join(vpd %>% dplyr::select(-Pathology, -matches('P53')), validation.patient.info %>% dplyr::select(Samplename, Patient, Pathology, p53.Status, Block, Initial.Endoscopy, Final.Endoscopy, Status ), by=c('Sample'='Samplename')) %>%
+  dplyr::rename(Endoscopy.Year = Endoscopy, Samplename = Sample, RR = `Relative Risk`, Prediction = Probability)
+
+vpd.per.endo = vpd %>% group_by(Hospital.Research.ID, Status, Patient, PID) %>% 
+  dplyr::mutate('n.samples'=length(PID),'Samples'=paste(Samplename,collapse=',')) %>% 
+  dplyr::summarise_at(vars(matches('Pred|RR|Year')), list(max)) %>% dplyr::select(-matches('^Sample$')) %>% 
+  arrange(Patient, Endoscopy.Year, PID) 
+vpd.per.endo = vpd.per.endo %>% mutate(quants = cut(Prediction, breaks=cuts, include.lowest = T)) %>% left_join(pred.confidence %>% dplyr::select(quants, Risk), by='quants') %>% dplyr::select(-matches('quant')) %>% dplyr::mutate(Risk = as.character(Risk))
+
+vpd %>% group_by(Status, Risk)  %>% dplyr::summarise(n = length(Risk)) %>% spread(Risk, n) %>% rowwise %>% 
+  dplyr::mutate(n=sum(High,Low,Moderate), Low = Low/n, Moderate = Moderate/n, High = High/n) %>%
+  mutate_if(is.double, list(round),2) %>%
+  kableExtra::kable(caption='Test set per-sample predictions') %>% kableExtra::kable_styling(full_width = F)
+
+vpd.per.endo %>%  group_by(Status, Risk)  %>% dplyr::summarise(n = length(Risk)) %>% spread(Risk, n) %>% rowwise %>% 
+  dplyr::mutate(n = sum(High,Low,Moderate), Low = Low/n, Moderate = Moderate/n, High = High/n) %>% mutate_if(is.double, list(round),2) %>%
+  kableExtra::kable(caption='Test set per-endoscopy predictions') %>% kableExtra::kable_styling(full_width = F)
 
 # Get the SD, mean, and variance(?) per sample
-v.sampleVar = do.call(rbind, lapply(raw, function(f) {
-  t = read.table(f, sep='\t', header=T)
-  sampleCols = grep('chr|arm|start|end|probes', colnames(t), invert = T)
-  x = do.call(rbind, lapply(sampleCols, function(i){
-    cbind('Sample.Mean'=mean(t[,i]),
-          'Sample.SD'=sd(t[,i]),
-          'Sample.Var'=var(t[,i]))
-  }))
-  rownames(x) = colnames(t)[sampleCols]
-  return(x)
-}))
-
-arms = grep('arms', cleaned, value=T)
-segs = grep('arms', cleaned, invert=T, value=T)
-
-length(segs) == length(arms)
-
-# Load segment files
-tiled.segs = do.call(rbind, lapply(segs, function(f) {
-  load.segment.matrix(f)
-}))
-dim(tiled.segs)
-
-tiled.arms = do.call(rbind, lapply(arms, function(f) {
-  load.segment.matrix(f)
-}))
-
-length(z.mean) == ncol(tiled.segs)
-
-for (i in 1:ncol(tiled.segs)) 
-  tiled.segs[,i] = unit.var(tiled.segs[,i], z.mean[i], z.sd[i])
-
-for (i in 1:ncol(tiled.arms)) 
-  tiled.arms[,i] = unit.var(tiled.arms[,i], z.arms.mean[i], z.arms.sd[i])
-
-dim(tiled.arms)
-dim(tiled.segs)
-
-cx.score = unit.var(score.cx(tiled.segs,1), mn.cx, sd.cx)
-
-validation.df = subtract.arms(tiled.segs, tiled.arms)
-validation.df = cbind(validation.df, 'cx'=cx.score)
-
-fauxlabel = rep(0, nrow(validation.df) )
-names(fauxlabel) = intersect(validation.patient.info$Samplename, rownames(validation.df))
-
-```
-
-```{r validation_globalplot, message=F, warning=F, echo=F, fig.height=8, fig.width=8}
+# v.sampleVar = do.call(rbind, lapply(raw, function(f) {
+#   t = read.table(f, sep='\t', header=T)
+#   sampleCols = grep('chr|arm|start|end|probes', colnames(t), invert = T)
+#   x = do.call(rbind, lapply(sampleCols, function(i){
+#     cbind('Sample.Mean'=mean(t[,i]),
+#           'Sample.SD'=sd(t[,i]),
+#           'Sample.Var'=var(t[,i]))
+#   }))
+#   rownames(x) = colnames(t)[sampleCols]
+#   return(x)
+# }))
+
+# # Load segment files
+# tiled.segs = do.call(rbind, lapply(segs, function(f) {
+#   load.segment.matrix(f)
+# }))
+# dim(tiled.segs)
+# 
+# tiled.arms = do.call(rbind, lapply(arms, function(f) {
+#   load.segment.matrix(f)
+# }))
+# 
+# length(z.mean) == ncol(tiled.segs)
+# 
+# for (i in 1:ncol(tiled.segs)) 
+#   tiled.segs[,i] = unit.var(tiled.segs[,i], z.mean[i], z.sd[i])
+# 
+# for (i in 1:ncol(tiled.arms)) 
+#   tiled.arms[,i] = unit.var(tiled.arms[,i], z.arms.mean[i], z.arms.sd[i])
+# 
+# dim(tiled.arms)
+# dim(tiled.segs)
+# 
+# cx.score = unit.var(score.cx(tiled.segs,1), mn.cx, sd.cx)
+# 
+# validation.df = subtract.arms(tiled.segs, tiled.arms)
+# validation.df = cbind(validation.df, 'cx'=cx.score)
+# 
+# fauxlabel = rep(0, nrow(validation.df) )
+# names(fauxlabel) = intersect(validation.patient.info$Samplename, rownames(validation.df))
+
+```
+
+```{r validation_globalplot, message=F, warning=F, echo=F, fig.height=8, fig.width=8, eval=F}
 valCNdf = t(validation.df[,grep('^\\d+:',colnames(validation.df))])
@@
 ```{r validation, message=F, warning=F, echo=F, fig.height=8, fig.width=8}
-file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
-message(paste("loading", file))
-load(file, verbose=T)
-
-fitV = models[[select.alpha]]
-lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']
-
-pred = predict(fitV, newx=validation.df, s=lambda.opt, type='response')
-or = predict(fitV, newx=validation.df, s=lambda.opt, type='link')
-
-vpd = lapply(unique(validation.patient.info$Hospital.Research.ID), function(pt) {
-  info = subset(validation.patient.info, Hospital.Research.ID == pt)
-  info$Prediction = NA
-  info$OR = NA
-  info$PID = unlist(lapply(info$Path.ID, function(x) unlist(strsplit(x, 'B'))[1]))
-
-  info = cbind(info, v.sampleVar[info$Samplename,])
-
-  return(info)
-})
-names(vpd) = unique(validation.patient.info$Hospital.Research.ID)
 
@@
 
-for (pt in names(vpd)) {
-  vpd[[pt]]$Prediction[which(vpd[[pt]]$Samplename %in% rownames(pred))] = pred[ intersect(rownames(pred), vpd[[pt]]$Samplename), ]
-  vpd[[pt]]$OR[which(vpd[[pt]]$Samplename %in% rownames(or))] = or[ intersect(rownames(or), vpd[[pt]]$Samplename), ]
-}
-
-back.v = do.call(rbind.data.frame,lapply(vpd, transform.by.time, info=validation.patient.info))
-back.v = merge(back.v, validation.patient.info[,c('Samplename','p53.Status')])
+back.v = do.call(bind_rows, lapply(unique(vpd$Patient) %>% map( function(p) vpd %>% filter(Patient == p) ), transform.by.time, info=validation.patient.info)) %>% dplyr::mutate(Patient = factor(Patient))
 
@@
 back.v$brks = cut(back.v$months.before.final, 
-                  breaks=sq, 
-                 labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-c(length(sq)-1,length(sq))], '>120'),
-                  include.lowest = T, ordered_result = T)
+                  breaks=sq, labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x,collapse='-')))[-c(length(sq)-1,length(sq))], '>120'), include.lowest = T, ordered_result = T)
+
 back.v$brks = with(back.v, factor(brks, rev(levels(brks))))
 
-back.v$Pathology = as.character(back.v$Pathology)
-
-back.v$Pathology = factor(back.v$Pathology, levels=c('BE','ID','LGD','HGD','IMC'), ordered = T)
+#back.v$Pathology = as.character(back.v$Pathology)
+#back.v$Pathology = factor(back.v$Pathology, levels=c('BE','ID','LGD','HGD','IMC'), ordered = T)
 
@@
 ```{r echo=F, message=F, warning=F, fig.height=5, fig.width=8}
-back.v$quants = with(back.v, cut(Prediction, breaks=cuts))
-back.v = merge(back.v, pred.confidence[c('quant','perc','Risk')], by.x='quants', by.y='quant')
-back.v$Risk = factor(back.v$Risk, levels=c('High','Moderate','Low'))
-back.v$nl = factor(back.v$nl, ordered = T)
+back.v = back.v %>% dplyr::mutate(Risk = factor(Risk, levels=c('High','Moderate','Low')), nl = factor(nl, ordered = T))
 
@@
   facet_grid(~Status, scales='free_x') + labs(y='Percent of samples', title='Validation patient set') + plot.theme
-
-back.v$Prediction
-
-table(subset(back.v, Status == 'P')$Prediction > 0.6)
-table(subset(back.v, Status == 'NP')$Prediction < 0.3)
 
@@
 ```
-
-```{r, eval=F}
-# TODO: Relative risk
-pred = predict(f, newx=dysplasia.df, s=performance.at.1se$`0.9`$lambda)
-?predict
-?predict.glm
-quantile(pred)
-p0 = predict(f, newx=dysplasia.df, s=performance.at.1se$`0.9`$lambda, type='response')
-p0
-p1 = 1/(1+exp(-pred+2.66))
-plot(p0,p1)
-hist(p1)
-hist(p0)
-hist(p1)
-plot(p0,p1)
-hist(p1, n=20)
-mean(p1>.5)
-mean(p0>.5)
-hist(pred)
-
-```

### 2019-08-14 Validation Report.Rmd
...

@@
 
+library(tidyverse)
 library(BarrettsProgressionRisk)
-library(knitr)
 library(kableExtra)
-library(tidyverse)
 library(gridExtra)
@@
 
-#source('lib/load_patient_metadata.R')
+source('lib/common_plots.R')
 
@@
 analysis.dir = '~/Data/BarrettsProgressionRisk/Analysis/validation'
+
+pred.dir = '~/Data/BarrettsProgressionRisk/Analysis/validation/predictions'
+dir.create(pred.dir, showWarnings = F)
+
+model = '~/Data/BarrettsProgressionRisk/Analysis/5e6_arms/all.pt.alpha.Rdata'
+load(model, verbose=T)
+
+fit = models$`0.9`
+lambda = performance.at.1se$`0.9`$lambda
 
@@
 
-if (file.exists('batch8_preds.txt')) {
-  batch8_preds = read_tsv('batch8_preds.txt')
+if (file.exists(paste0(pred.dir, '/batch8_preds.txt'))) {
+  batch8_preds = read_tsv(paste0(pred.dir, '/batch8_preds.txt'))
 } else {
@@
 
-    prr = BarrettsProgressionRisk::predictRiskFromSegments(segmented, F)
+    prr = BarrettsProgressionRisk::predictRiskFromSegments(segmented, model = fit, s = lambda, verbose = F)
     BarrettsProgressionRisk::predictions(prr,'sample')
@@
 
-  write_tsv(batch8_preds, 'batch8_preds.txt')
+  write_tsv(batch8_preds, paste0(pred.dir, '/batch8_preds.txt'))
 }
@@
 
+The predictions from these samples are not particularly good.
+
 ```{r}
@@
 
-if (file.exists('oldp_preds.txt')) {
-  oldp.preds = read_tsv('oldp_preds.txt')
+if (file.exists(paste0(pred.dir, '/oldp_preds.txt'))) {
+  oldp.preds = read_tsv(paste0(pred.dir, '/oldp_preds.txt'))
 } else {
@@
 
-    prr = BarrettsProgressionRisk::predictRiskFromSegments(segmented,F)
+    prr = BarrettsProgressionRisk::predictRiskFromSegments(segmented, model = fit, s = lambda, verbose = F)
     predictions(prr)
   }))
-  write_tsv(oldp.preds, 'oldp_preds.txt')
+  write_tsv(oldp.preds, paste0(pred.dir, '/oldp_preds.txt'))
 }
@@
 
-oldp.preds %>% dplyr::summarise(
-  'TPR'=length(which(Risk == 'High'))/n(),
-  'FNR'=length(which(Risk == 'Low'))/n()) %>% kable() %>% kable_styling(full_width = F)
-
-oldp.preds %>% group_by(`Risk`) %>% dplyr::summarise( 'n samples'=n() ) %>% 
-  kable(caption='Count sample predicted risks, P') %>% kable_styling(full_width = F)
-
-fnr.patients = oldp.preds %>% 
-  group_by(`Hospital Research ID`) %>% dplyr::summarise(
-  TPR = length(which(Risk == 'High'))/n()
-) %>% dplyr::summarise( 'Misclassified patients' = length(which(TPR < 0.5))/n() ) %>% pull
+oldp.preds %>% group_by(`Risk`) %>% dplyr::summarise( 'n samples'=n() ) %>% spread(Risk, `n samples`) %>%
+  kable(caption='Progressor sample predicted risks') %>% kable_styling(full_width = F)
 
@@
   'Pathology Notes' = paste(`Path Notes`, collapse='; ')
-) %>% kable(caption = 'Classified predictions per patient (all P)') %>% kable_styling('striped')
+) %>% mutate(
+  `High Risk` = cell_spec(`High Risk`, "html", color = ifelse(`High Risk` > 0, "red", "grey39")),
+  `Mod. Risk` = cell_spec(`Mod. Risk`, "html", color = ifelse(`Mod. Risk` > 0, "orange", "grey39")),
+  `Low Risk` = cell_spec(`Low Risk`, "html", color = ifelse(`Low Risk` > 0, "blue", "grey39")),
+) %>% kable(format='html',escape = F,caption = 'Classified predictions per patient (all P)') %>% kable_styling('striped', full_width = F)
+
+oldp.preds = oldp.preds %>% rowwise %>% mutate(PID = unlist(strsplit(`Block ID`, '(-|\\s)\\w+$', perl = T))[1])
+
+ggplot(oldp.preds, aes(`Hospital Research ID`, Probability, color=Probability)) + geom_point() +
+  scale_color_distiller(palette = 'RdYlBu', name='P(P)') + plot.theme + theme(axis.text.x = element_text(angle=45,hjust=1))
+
+oldp.per.endo = oldp.preds %>% group_by(`Hospital Research ID`, PID) %>% 
+  dplyr::mutate('n.samples'=length(PID),'Samples'=paste(Sample,collapse=',')) %>% 
+  dplyr::summarise_at(vars(matches('Prob|Relative|Endo')), list(max)) %>% dplyr::select(-matches('^Sample$')) %>% 
+  arrange(`Hospital Research ID`, Endoscopy, PID) 
+#back.per.endo = back.per.endo %>% mutate(quants = cut(Prediction, breaks=cuts, include.lowest = T)) %>% left_join(pred.confidence %>% dplyr::select(quants, Risk), by='quants') %>% dplyr::select(-matches('quant')) %>% dplyr::mutate(Risk = as.character(Risk))
+
+fnr.patients = oldp.preds %>% 
+  group_by(`Hospital Research ID`) %>% dplyr::summarise(
+  FNR = length(which(Risk == 'Low'))/n() ) %>% 
+  dplyr::summarise( 'Misclassified patients' = length(which(FNR > 0.5))/n() ) %>% pull %>% round(2)
 
@@
 
-For the old progressor samples to-date, the model failed to classify any samples as high risk for *`r fnr.patients*100`%* of patients.
+For the old progressor samples to-date, the model failed to classify any samples as high or moderate risk for *`r fnr.patients*100`%* of patients.
 None of these were the samples I might have said were poor quality via manual QC so taken at face value these patients/samples were incorrectly predicted.
@@
 
-if (file.exists('val_np.txt')) {
-  new.np.preds = read_tsv('val_np.txt')
+if (file.exists(paste0(pred.dir, '/val_np.txt'))) {
+  new.np.preds = read_tsv(paste0(pred.dir, '/val_np.txt'))
 } else {
-  segFiles = grep(paste(new.np$`Patient ID`,collapse='|'), list.files('~/Data/BarrettsProgressionRisk/Analysis/validation', '(2|3)_segObj',  full.names = T, recursive = T), value=T)
+  segFiles = grep(paste(sub('\\(.*\\)','', new.np$`Patient ID`),collapse='|'), list.files('~/Data/BarrettsProgressionRisk/Analysis/validation', '(2|3)_segObj',  full.names = T, recursive = T), value=T)
 
@@
     segmented$sample.info = BarrettsProgressionRisk::loadSampleInformation(info %>% filter(Sample %in% segmented$sample.info$Sample) )
-    prr = BarrettsProgressionRisk::predictRiskFromSegments(segmented,F)
-    predictions(prr)
+    prr = BarrettsProgressionRisk::predictRiskFromSegments(segmented, model = fit, s = lambda, verbose = F)
+    p = predictions(prr)
+    print(p)
+    return(p)
   }))
-  write_tsv(new.np.preds, 'val_np.txt')
+  write_tsv(new.np.preds, paste0(pred.dir, '/val_np.txt'))
 }
@@
 #cor.test(new.np.preds$n.segs, new.np.preds$varMAD_median)
+#with(val.new.np.preds, cor.test(varMAD_median, Probability))
 
-new.np.preds %>% dplyr::summarise(
-  'FPR'=length(which(Risk == 'High'))/n(),
-  'TPR'=length(which(Risk == 'Low'))/n()) %>% kable() %>% kable_styling(full_width = F)
-
-new.np.preds %>% group_by(`Risk`) %>% dplyr::summarise( 'n samples'=n() ) %>% 
+new.np.preds %>% group_by(`Risk`) %>% dplyr::summarise( 'n samples'=n() ) %>% spread(Risk, `n samples`) %>% 
   kable(caption='Count sample predicted risks, NP') %>% kable_styling(full_width = F)
-
-fpr.patients = new.np.preds %>%  group_by(`Hospital Research ID`) %>% dplyr::summarise(
-  TPR = length(which(Risk == 'Low'))/n()
-) %>% dplyr::summarise( 'Misclassified patients' = length(which(TPR < 0.5))/n() ) %>% pull
 
@@
 
-with(val.new.np.preds, cor.test(varMAD_median, Probability))
+# val.new.np.preds %>% group_by(`Hospital Research ID`) %>% dplyr::summarise(
+#   `n samples` = n(), 
+#   'High Risk' =  length(which(Risk == 'High')),
+#   'Mod. Risk' = length(which(Risk == 'Moderate')),
+#   'Low Risk' =   length(which(Risk == 'Low')),
+#   'QC var' = signif( mean(varMAD_median), 2)
+# ) %>% mutate(
+#   `High Risk` = cell_spec(`High Risk`, "html", color = ifelse(`High Risk` > 0, "red", "grey39")),
+#   `Mod. Risk` = cell_spec(`Mod. Risk`, "html", color = ifelse(`Mod. Risk` > 0, "orange", "grey39")),
+#   `Low Risk` = cell_spec(`Low Risk`, "html", color = ifelse(`Low Risk` > 0, "blue", "grey39")),
+# ) %>%
+#   kable(format='html',escape = F, caption = 'Classified predictions per patient (all NP)') %>% kable_styling('striped', full_width = F)
 
-val.new.np.preds %>% group_by(`Hospital Research ID`) %>% dplyr::summarise(
-  `n samples` = n(), 
-  'High Risk' = length(which(Risk == 'High')),
-  'Mod. Risk' = length(which(Risk == 'Moderate')),
-  'Low Risk' =   length(which(Risk == 'Low')),
-  'QC var' = signif( mean(varMAD_median), 2)
-) %>% kable(caption = 'Classified predictions per patient (all NP)') %>% kable_styling('striped')
+ggplot(new.np.preds, aes(Probability)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1), show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') +
+    plot.theme + labs(title='Sample predictions for validation NP samples', y='n Samples', x='Probability') + theme(legend.position = 'none')
+
+new.np.preds = new.np.preds %>% rowwise %>% mutate(PID = unlist(strsplit(`Block ID`, '(-|\\s)\\w+$', perl = T))[1])
+
+ggplot(new.np.preds, aes(`Hospital Research ID`, Probability, color=Probability)) + geom_point() +
+  scale_color_distiller(palette = 'RdYlBu', name='P(P)') + plot.theme + theme(axis.text.x = element_text(angle=45,hjust=1))
+
+np.per.endo = new.np.preds %>% group_by(`Hospital Research ID`, PID) %>% 
+  dplyr::mutate('n.samples'=length(PID),'Samples'=paste(Sample,collapse=',')) %>% 
+  dplyr::summarise_at(vars(matches('Prob|Relative|Endo')), list(max)) %>% dplyr::select(-matches('^Sample$')) %>% 
+  arrange(`Hospital Research ID`, Endoscopy, PID) 
+#back.per.endo = back.per.endo %>% mutate(quants = cut(Prediction, breaks=cuts, include.lowest = T)) %>% left_join(pred.confidence %>% dplyr::select(quants, Risk), by='quants') %>% dplyr::select(-matches('quant')) %>% dplyr::mutate(Risk = as.character(Risk))
+
+#np.per.endo %>% filter(`Relative Risk` > 0)
 

### 2019-08-14 Paper.Rmd
...

@@
 
-#library(biomaRt)
-#library(GenomicRanges)
-#library(stringr)
-#library(ggplot2)
 library(ggrepel)
-library(GGally)
-library(pander)
+#library(GGally)
+library(kableExtra)
 library(reshape2)
 library(gridExtra)
-library(CoxHD)
+#library(CoxHD)
 library(RColorBrewer)
-#library(plyr)
-#library(dplyr)
 
@@
 source('lib/common_plots.R')
-#source('lib/data_func.R')
 
@@
 
-#data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
-#analysis.files = list.files(paste(data, 'Analysis', sep='/'), full.names=T)
-
-## Hospital.Research.ID info file
-
 patient.file = list.files(info.dir, pattern='All_patient_info.xlsx', recursive=T, full.names=T)
@@
 patient.info = all.patient.info %>% dplyr::filter(Set == 'Training')
-head(patient.info)
-
-# validation.patient.info =  subset(all.patient.info, Set == 'Test')
-# validation.patient.info = arrange(validation.patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
-# sum.validation.info = summarise.patient.info(validation.patient.info)
 
@@
 
-#pander(sum.patient.data[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Training patient set')
-
-#pander(sum.validation.info[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Validation patient set')
-
-#all.info = rbind(sum.patient.data, sum.validation.info)
-
 ```
@@
 ## Quick look at our demographic information
-
-```{r echo=F}
-
-pander(
-  as.data.frame(sum.patient.data %>% dplyr::summarise(
-  'Total Patients'=length(unique(Patient)),
-  'Total Samples'=sum(total.samples),
-  'Range'=paste(range(end.year-start.year), collapse='-'),
-  'Median Years'=paste(median(end.year-start.year),'+/-',round(sd(end.year-start.year),1)),
-  'Median Endoscopies'=paste(median(total.endos),'+/-',round(sd(total.endos),1)),
-  'Median age'=paste(median(age.diagnosis),'+/-',round(sd(age.diagnosis),1)),
-  'gender F:M'=paste(table(gender),collapse=':')
-))
-, caption='full cohort')
-```
 
@@
 
-```
-
-`r pander(t(cohortSummary), caption='Full cohort, split by status')`
+cohortSummary %>% t %>% kable(caption='Training set') %>% kable_styling()
+```
 
@@
 
-p53Path = table(subset(patient.info, Status == 'P' & p53.Status == 1)$Pathology)
+p53Path = patient.info %>% dplyr::filter(Status == 'P' & p53.Status == 1) %>% dplyr::group_by(Pathology) %>% dplyr::summarise(n=length(Pathology)) %>% spread(Pathology,n)
+
 p53.pred.byPath = round(p53Path/table(subset(patient.info, Status == 'P' & !is.na(p53.Status))$Pathology), 2)
@@
 Example subset of the data matrix:
-`r pander(dysplasia.df[1:5, 1:5], caption=paste("dimensions:", paste(dim(dysplasia.df), collapse=', ')), justify='left')`
+`r dysplasia.df[1:5, 1:5] %>% kable(caption=paste("dimensions:", paste(dim(dysplasia.df), collapse=', '))) %>% kable_styling('striped')`
 
@@
   dplyr::mutate(chrom = factor(sub('chr','', chrom), levels=c(1:22), ordered=T))
-#chr.lengths$chrom = sub('chr','',chr.lengths$chrom)
-#chr.lengths = subset(chr.lengths, chrom %in% c(1:22))
-#chr.lengths$chrom = factor(chr.lengths$chrom, levels=c(1:22), ordered = T)
 
@@
 pv
+ggsave('plots/per_sample_cx.png', plot=pv + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
 
@@
 roc.plot(rocX, title='Per sample cx') + plot.theme
-
-ggsave('plots/per_sample_cx.png', plot=pv + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
 
@@
 
-```
-
-`r pander(mp$performance[,c('n.Coef','25%','50%','75%', '100%')], justify='left', caption='Number of features stable in n% or more folds')`
-
-Values that are closer to full lasso (alpha=1) are pretty similar, however at full lasso we see fewer features than if we regularize it at 0.8. They share all non-zero features as well.
-`r pander(sapply(mp$stable.coeffients, length), caption='Total features at each value of alpha')`
-
-`r pander(most.stable[,c('chr','start','stop')], caption='Features identified as stable in 75% or more folds.')`
+mp$performance[,c('n.Coef','25%','50%','75%', '100%')] %>% kable(caption='Number of features stable in >=n% of folds') %>% kable_styling()
+```
+
+Values that are closer to full lasso (alpha=1) are pretty similar, however at full lasso we see fewer features than if we regularize it at `r select.alpha`. They share all non-zero features as well.
+`r sapply(mp$stable.coeffients, length) %>% t %>% kable(caption='Total features at each value of alpha') %>% kable_styling(full_width = F)`
 
@@
 ```{r, echo=F, message=F, fig.width=5, fig.height=5}
-#pg.sampNOHGD = do.call(rbind, pg.sampNOHGD)
-
 perf = cbind.data.frame(perf=performance.at.1se, pt=unique(pg.sampNOHGD$Patient))
@@
 })
-ch = as.data.frame(hazards)
-ch$coef = coefs[[select.alpha]][,1]
-ch$stability = featureStability
-ch$label = rownames(ch)
-ch = arrange(ch, -hazards)
-
-qq = quantile(ch$hazards, seq(0,1,.15))
-p = ggplot(ch, aes(coef, hazards, col=hazards >= qq[length(qq)])) + 
+
+features = as_tibble(hazards, rownames = 'label') %>% dplyr::rename(hazards = value) %>% mutate(coef = coefs[[select.alpha]][,1], stability = featureStability) %>% arrange(-hazards)
+
+qq = quantile(features$hazards, seq(0,1,.15))
+p = ggplot(features, aes(coef, hazards, col=hazards >= qq[length(qq)])) + 
   geom_point(alpha=0.8, show.legend = F) + 
@@
 
-## Arms...
-if (exists('arm.locs')) {
-  ch[which(ch$label %in% with(arm.locs, paste(chr, ':', start, '-', end, sep=''))),]
-}
-
-```
-
-`r pander(ch, caption='Selected coefficients')`
-
-```{r nohgd_coefs, message=F, warning=F, echo=F}
-setdiff(ch$label, rownames(nohgd.coefs))
-```
+```
+
+`r features %>% kable(caption='Selected coefficients') %>% kable_styling(full_width=F)`
 
 ```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
-genomeFeat = grep('^\\d+:', ch$label)
-featureAnn = data.frame(apply(do.call(rbind, strsplit(ch$label[genomeFeat], ':|-')), 2, as.numeric))
-colnames(featureAnn) = c('chr','start','end')
-featureAnn[] = lapply(featureAnn[], function(x) as.numeric(as.character(x)))
-
-chh = cbind(ch[genomeFeat,], featureAnn)
-chh = subset(chh, !is.na(chr))
-
-chh = merge(chh, chr.lengths[,c('chrom','chr.length')], by.y='chrom', by.x='chr')
-chh$chr = factor(chh$chr, levels=c(1:22))
-
-chh$ymax = ifelse(chh$coef > 0, max(globalCN.plot$data$mean.value), min(globalCN.plot$data$mean.value))
-chh$gl = factor(ifelse(chh$coef > 0, 'gain', 'loss'))
+features = features %>% separate(label, c('chr','start','end'), sep=':|-', remove=F) %>% 
+  mutate_at(vars(start, end), list(as.numeric)) %>% mutate(chr = factor(chr,levels=c(1:22)))
+
+features = features %>% mutate(ymax = ifelse(coef > 0, max(globalCN.plot$data$mean.value), min(globalCN.plot$data$mean.value)), 
+                    gl = factor(ifelse(coef > 0, 'gain','loss')))
+
+features = left_join(features, chr.lengths %>% dplyr::select(chr, chr.length), by=c('chr')) %>% mutate(chr = factor(chr, levels=c(1:22), ordered=T))
 
@@
 }
-chh = chh %>% rowwise() %>% dplyr::mutate('max'=custom.max(coef, hazards, ymax))
-
-p = globalCN.plot + geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + 
-  geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')
+features = features %>% rowwise() %>% dplyr::mutate('max'=custom.max(coef, hazards, ymax)) %>% arrange(chr)
+
+p = globalCN.plot + geom_rect(data=features %>% filter(gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + 
+  geom_rect(data= features %>% filter(gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')
 
@@
 
-p = globalBECN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')
+p = globalBECN.plot +  geom_rect(data=features %>% filter(gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=features %>% filter(gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')
 
@@
 
-cutoff = round(pROC::coords(roc, 'best')[['threshold']], 2)
+cutoff = round(pROC::coords(roc, 'best', ret = 'threshold', transpose=T), 2)
 ```
@@
 
-coefs.per.pt = data.frame(matrix(ncol=length(unique(pg.samp$Patient)), nrow=nrow(ch), dimnames=list( ch$label, unique(pg.samp$Hospital.Research.ID))))
+coefs.per.pt = data.frame(matrix(ncol=length(unique(pg.samp$Patient)), nrow=nrow(features), dimnames=list( features$label, unique(pg.samp$Hospital.Research.ID))))
 
@@
   names(lcf) = rownames( loo.coefs[[pt]] )
-  sharedcfs = intersect(names(lcf), ch$label)
+  sharedcfs = intersect(names(lcf), features$label)
   coefs.per.pt[sharedcfs, pt] = lcf[sharedcfs]
@@
 
-ch = as_tibble(coefs$`0.9`[,1,drop=F], rownames='coef')
-
-loo.ch = dplyr::left_join(ch, loo.coefs, by='coef')
+features = as_tibble(coefs$`0.9`[,1,drop=F], rownames='coef')
+
+loo.features = dplyr::left_join(features, loo.coefs, by='coef')
 
@@
 
-ch$jack.se = apply(loo.ch, 1, function(x) {
+features$jack.se = apply(loo.features, 1, function(x) {
   jk = bootstrap::jackknife(x[-1], var)
@@
 # per coefficient
-ch$jack.se = apply(coefs.per.pt, 1, function(x) {
+features$jack.se = apply(coefs.per.pt, 1, function(x) {
   jk = jackknife(x, var)
@@
 })
-colnames(ch) = c('label','coef','jack.se')
-ggplot(ch, aes(reorder(label, coef), coef)) + geom_point() + 
+colnames(features) = c('label','coef','jack.se')
+ggplot(features, aes(reorder(label, coef), coef)) + geom_point() + 
   geom_errorbar(aes(ymin=coef-jack.se, ymax=coef+jack.se)) + 
@@
 ```{r echo=F, eval=T}
-file = paste(cache.dir, 'model_data.Rdata', sep='/')
-if (!file.exists(file))
-  stop(paste("Missing data file", file))
-load(file, verbose=T)
-file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
-if (!file.exists(file))
-  stop(paste("Missing data file", file))
-load(file, verbose=T)
-rm(dysplasia.df,coefs,plots,labels)
-
-select.alpha = '0.9'
-fitV = models[[select.alpha]]
-lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']
-
-dirA = '~/Data/Ellie/Analysis/downsampled'
-ds_info = read_excel('~/Data/Ellie/Analysis/downsampled_ids.xlsx')
-predA = list.files(dirA, 'predictions.Rdata', full.names = T)
-load(predA)
-
-dspred = merge(dspred, ds_info, by.x='row.names', by.y='Illumina ID')
-
-ggplot(dspred, aes(Type, RR, group=Type, fill=Type)) + geom_boxplot(outlier.colour = NA) + geom_jitter(width=0.2, aes(color=Status)) +
-  labs(x='', y='Relative Risk', title='Predictions on downsampled patients') + 
-  plot.theme
-
-adj = subset(dspred, grepl('Adjacent',Type))
-prg = back[,c('Prediction', 'OR', 'Status')]
-prg$Status = ifelse(prg$Status == 'P', 'Progressor','Non-Progressor')
-
-#  subset(back, Status == 'P', select=c('Prediction', 'OR', 'Status'))
-colnames(prg) = colnames(adj)
-m = rbind(adj,prg)
-m$Type = factor(m$Type, levels=c('Non-Progressor','BEST2 NDBO','Progressor','AdjacentBE'), ordered = T)
-
-pr = ggplot(m, aes(Type, RR, fill=Type)) + 
-  geom_jitter(width=0.2) + geom_boxplot() + 
-  labs(x='', y='RR', title='Predicted Relative Risk') + 
-  theme(legend.position = 'none', text=element_text(size=11)) + plot.theme 
+dir = '~/Data/BarrettsProgressionRisk/Analysis/downsampled/predictions'
+info = '~/Data/BarrettsProgressionRisk/QDNAseq/all_downsampled/downsampled_ids.xlsx'
+
+info = readxl::read_xlsx(info)
+
+files = list.files(dir, 'Rdata', recursive = T, full.names = T)
+
+pfile = paste0(dir, '/ds_preds.txt')
+if (file.exists(pfile)) {
+  ds.preds = read_tsv(pfile)
+} else {
+  ds.preds = do.call(bind_rows, lapply(files, function(f) {
+    load(f, verbose=T)
+    bind_cols(predictions(prr), adjustRisk(prr, 'mean', 'sample')$adj.predictions %>% dplyr::select(Probability, `Relative Risk`) %>% dplyr::rename_all(funs(paste0), ' Adjusted ') )
+  }))
+  write_tsv(ds.preds, pfile)
+}
+
+## TODO merge in RR from 'back'
+
+ds.preds = ds.preds %>% dplyr::select(-Status) %>% left_join(info %>% dplyr::select(`Illumina ID`, Status), by=c('Sample' = 'Illumina ID')) %>% dplyr::mutate(Type = factor(Type, levels=c('Normal','NDBE', 'LGD', 'Tumor Adjacent BE'))) %>% arrange(Probability)
+
+m = reshape2::melt(ds.preds, measure.vars = c('Relative Risk','Relative Risk Adjusted'), id.vars = c('Type', 'Status', 'Risk'))
+
+pr = ggplot(m, aes(variable, value)) + 
+  facet_grid(~Type, scales = 'free_x') + 
+  geom_jitter(width=0.2, aes(shape=Status, color=Risk), size=2, show.legend = F) + 
+  geom_boxplot(aes(group=variable, fill=Type), outlier.colour = NA, show.legend = F) + 
+  scale_color_manual(values =  BarrettsProgressionRisk::riskColors()) +
+  #scale_color_gradientn(colors = myPal,  name='') +
+  scale_fill_manual(values=rev(RColorBrewer::brewer.pal(5,'RdYlBu')[c(1,2,4,5)])) +
+  labs(x='', y='Relative Risk', title='Predictions on downsampled patients') + plot.theme +
+  theme(axis.text.x = element_text(angle=45, hjust=1))
+
 pr
 
+ds.preds %>% dplyr::group_by(Status, Risk) %>% dplyr::summarise(n=length(Risk)) %>% spread(Risk, n) %>% kableExtra::kable(caption = 'Sample by risk and known status') %>% kableExtra::kable_styling(full_width = F)
+
 ggsave('plots/adjacentBE.png', plot=pr, width=8, height=8,dpi=300)
-
-#save(dspred, file=paste(dirB, 'predictions.Rdata', sep='/'))
 
@@
 # glm selected features + Path
-cox2 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', 'Pathology',  grep('^\\d+:', ch$label, value=T))]) 
+cox2 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', 'Pathology',  grep('^\\d+:', features$label, value=T))]) 
 ggsurv(survfit(cox2),cens.shape=3,lty.est=1,back.white=F,
@@
 
-cox2 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis','Age.bracket', grep('^\\d+:', ch$label, value=T))]) 
+cox2 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis','Age.bracket', grep('^\\d+:', features$label, value=T))]) 
 ggsurv(survfit(cox2),cens.shape=3,lty.est=1,back.white=F,
@@
 
-cox2 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis',grep('^\\d+:', ch$label, value=T))]) 
+cox2 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis',grep('^\\d+:', features$label, value=T))]) 
 ggsurv(survfit(cox2),cens.shape=3,lty.est=1,back.white=F,
@@
 ## GLM regions only
-coxDF = cbind(coxDF[,1:3],pathology[,c('BE','ID')], age, coxDF[,grep('^\\d+:', ch$label, value=T)])
+coxDF = cbind(coxDF[,1:3],pathology[,c('BE','ID')], age, coxDF[,grep('^\\d+:', features$label, value=T)])
 coxDF$Pathology = NULL
@@
 wt$label = sub('\\.', '-', sub('\\.', ':', sub('chr', '', rownames(wt))))
-wt = merge(wt, ch[,c('label','hazards', 'coef','stability')], by='label', all=T)
+wt = merge(wt, features[,c('label','hazards', 'coef','stability')], by='label', all=T)
 

### 2019-08-16 Paper.Rmd
...

@@
 ```{r setup, include=FALSE}
+library(tidyverse)
 library(BarrettsProgressionRisk)
@@
 source('lib/load_patient_metadata.R')
-source('lib/cv-pt-glm.R')
+#source('lib/cv-pt-glm.R')
 source('lib/common_plots.R')
@@
 
-`r pander(p53.pred.byPath, caption='Ratio of progressors predicted by p53 staining per pathology where p53 staining was done (excluding N/As)')`
+`r p53.pred.byPath %>% kable(caption='Ratio of progressors predicted by p53 staining per pathology where p53 staining was done (excluding N/As)') %>% kable_styling(full_width=F)`
 
@@
 preds = pg.samp %>% dplyr::select(Status, Prediction)
-roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
+roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds',transpose=T)
 roc$model = 'CN Model'
@@
 df = do.call(rbind, lapply(insilico.rocs, function(r) 
-    cbind.data.frame('Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'model'=r$model, 'is'=r$is) 
-))
-
-p = ggplot(df, aes(Specificity, Sensitivity,color=is)) + geom_line(size=1) + 
+    cbind.data.frame('Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'model'=r$model, 'is'=as.character(r$is) 
+)))
+
+p = ggplot(df, aes(Specificity, Sensitivity,color=is)) + geom_line(size=1, color='lightblue') + 
     geom_line(data=subset(df, is == 'discovery'), color='red', size=1) +
     scale_x_reverse() +
-    scale_color_brewer(palette='Blues') +
+    #scale_color_brewer(palette='Blues') +
     labs(title='ROC, all splits', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
@@
 
+# Not sure why I did this
 df = cbind.data.frame('specificity'=rev(roc$specificities), 'sensitivity'=rev(roc$sensitivities))
@@
   rbind.data.frame(
-#    est=t(round(pROC::coords(roc, "best"),2)),
-    max=t(round(pROC::coords(roc, 0.8),2)),
-    'No HGD'=t(round(pROC::coords(rocNOHGD, 0.6),2)),
-    'All samples'=t(round(pROC::coords(roc, 0.6),2))
+    est=t(round(pROC::coords(roc, "best", transpose=T),2)),
+    max=t(round(pROC::coords(roc, 0.8, transpose=T),2)),
+    'No HGD'=t(round(pROC::coords(rocNOHGD, 0.6, transpose=T),2)),
+    'All samples'=t(round(pROC::coords(roc, 0.6, transpose=T),2))
 )
@@
 
-# ggplot() + geom_line(data=df, aes(specificity, sensitivity), color='darkblue', size=1.5) + scale_x_reverse() + 
-#   geom_point(data=thresholds, aes(x=specificity,y=sensitivity,color=label), size=3,shape=8, show.legend = T) +
-#   geom_label_repel(data=thresholds, aes(x=specificity,y=sensitivity,fill=label,label=
-#     paste('Threshold ', round(threshold,2), '\n(FPR ', round((1-specificity),2)*100, '%, TPR ', round(sensitivity,2)*100,'%)\nAUC:',round(roc$auc, 2)*100, '%', sep='')),show.legend = F, color='white',nudge_x=0.05) + scale_fill_manual(values=brewer.pal(3,'Dark2')) + scale_color_manual(values=brewer.pal(3,'Dark2')) +
-#   labs(title='CN Model ROC', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme
-
-roc.plot(rocNOHGD, 'Model excluding HGD/IMC')
-pROC::coords(rocNOHGD, "best")
-pROC::coords(rocNOHGD, 0.6)
-pROC::auc(rocNOHGD, partial.auc=c(0.89,0.9), partial.auc.focus='spec', partial.auc.correct=T)
+ggplot() + 
+  geom_line(data=df, aes(specificity, sensitivity), color='darkblue', size=1.5) + scale_x_reverse() +
+  geom_point(data=thresholds, aes(x=specificity,y=sensitivity,color=label), size=3,shape=8, show.legend = T) +
+  geom_label_repel(data=thresholds, aes(x=specificity,y=sensitivity,fill=label,label=
+    paste('Threshold ', round(threshold,2), '\n(FPR ', round((1-specificity),2)*100, '%, TPR ', round(sensitivity,2)*100,'%)\nAUC:',round(roc$auc, 2)*100, '%', sep='')),show.legend = F, color='white',nudge_x=0.05) + scale_fill_manual(values=brewer.pal(4,'Dark2')) + scale_color_manual(values=brewer.pal(4,'Dark2')) +
+  labs(title='CN Model ROC', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme
+
+# roc.plot(rocNOHGD, 'Model excluding HGD/IMC')
+# pROC::coords(rocNOHGD, "best")
+# pROC::coords(rocNOHGD, 0.6)
+# pROC::auc(rocNOHGD, partial.auc=c(0.89,0.9), partial.auc.focus='spec', partial.auc.correct=T)
 
@@
 
-```{r pathroc, echo=F, fig.height=4, fig.width=8 }
+```{r pathroc, echo=F, fig.height=4, fig.width=12 }
 #sum(table(subset(patient.info, Status == 'P')$Pathology)[c('IMC','HGD')])/sum(table(subset(patient.info, Status == 'P')$Pathology))
@@
 rocEndoM = pROC::roc(Status ~ max, data=pendo, ci=T, of='thresholds')
-pROC::coords(rocEndoM, 'best', transpose=T)
 rocEndoM$model = 'Per Endoscopy Max'
@@
 rocEndoA = pROC::roc(Status ~ avg, data=pendo, ci=T, of='thresholds')
-pROC::coords(rocEndoA, 'best', transpose=T)
 rocEndoA$model = 'Per Endoscopy Avg'
@@
 
-lastEndo = pg.samp %>% group_by(Patient, Hospital.Research.ID, Status,Endoscopy.Year, Path.ID) %>% arrange(desc(Endoscopy.Year), PID)  %>% dplyr::summarise(avg = mean(Prediction) ) %>% ungroup %>% group_by(Patient) %>% dplyr::filter( Endoscopy.Year == max(Endoscopy.Year) ) %>% dplyr::select(-Endoscopy.Year, -PID)
+lastEndo = pg.samp %>% group_by(Patient, Hospital.Research.ID, Status,Endoscopy.Year, PID) %>% arrange(desc(Endoscopy.Year), PID)  %>% dplyr::summarise(avg = mean(Prediction) ) %>% ungroup %>% group_by(Patient) %>% dplyr::filter( Endoscopy.Year == max(Endoscopy.Year) ) %>% dplyr::select(-Endoscopy.Year, -PID)
 
 rocLastEndo = pROC::roc(Status ~ avg, data=lastEndo, ci=T, of='thresholds')
-pROC::coords(rocLastEndo, 'best', transpose=T)
 rocLastEndo$model = 'Last endoscopy'
@@
 rocFirstEndo = pROC::roc(Status ~ avg, data=firstEndo, ci=T, of='thresholds')
-pROC::coords(rocFirstEndo, 'best', transpose=T)
+
 rocFirstEndo$model = 'First endoscopy'
@@
 
+endo.thresholds = bind_cols('roc'=c('Max','Avg','Last','First'),
+  bind_rows(
+    pROC::coords(rocEndoM, 'best', transpose=T),
+    pROC::coords(rocEndoA, 'best', transpose=T),
+    pROC::coords(rocLastEndo, 'best', transpose=T),
+    pROC::coords(rocFirstEndo, 'best', transpose=T)
+  ))
+
 ```
@@
 
-```{r echo=F, fig.height=4, fig.width=4}
-
-#ptendo = do.call(rbind, pg.samp) %>% dplyr::group_by(Patient, Hospital.Research.ID, Status) %>% summarise('max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(Patient))
+```{r echo=F, fig.height=4, fig.width=12}
 
@@
 rocPtM = pROC::roc(Status ~ max, data=ptendo, ci=T, of='thresholds')
-pROC::coords(rocPtM, 'best', transpose=T)
+#pROC::coords(rocPtM, 'best', transpose=T)
 rocPtM$model = 'Per Patient Max'
@@
 rocPtA = pROC::roc(Status ~ mean, data=ptendo, ci=T, of='thresholds')
-pROC::coords(rocPtA, 'best', transpose=T))
-#roc.plot(rocPtA)
+#pROC::coords(rocPtA, 'best', transpose=T))
 rocPtA$model = 'Per Patient Avg'
@@
 
+grid.arrange(roc.plot(rocPtM, title='Max') , roc.plot(rocPtA, title='Avg'), nrow=1, top='Per Patient')
 ```
@@
   auc = c( pROC::auc(roc), range(pROC::ci.auc(roc))  )
-
-  sens = c(pROC::coords(roc, x)[['sensitivity']], pROC::ci.se(roc,pROC::coords(roc, x)[['specificity']], conf.level=0.95))
-  spec = c(pROC::coords(roc, x)[['specificity']], pROC::ci.sp(roc,pROC::coords(roc, x)[['sensitivity']], conf.level=0.95))
+  sens = c(pROC::coords(roc, x, tranpose=T)[['sensitivity']], pROC::ci.se(roc,pROC::coords(roc, x, tranpose=T)[['specificity']], conf.level=0.95))
+  spec = c(pROC::coords(roc, x, tranpose=T)[['specificity']], pROC::ci.sp(roc,pROC::coords(roc, x, tranpose=T)[['sensitivity']], conf.level=0.95))
 
@@
 transform.by.time<-function(pt, info) {
-  #pt = left_join(pt, info %>% dplyr::select(Hospital.Research.ID, Patient,Samplename,Block), by=c('Hospital.Research.ID','Patient','Samplename'))
   pt = pt %>% dplyr::group_by(PID, Samplename) %>% dplyr::mutate( 'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12) %>% arrange(Pathology, PID, Endoscopy.Year) %>% ungroup
@@
 
-back = do.call(bind_rows,lapply(unique(pg.samp$Patient) %>% map( function(p) pg.samp %>% filter(Patient == p)  ), transform.by.time, info=patient.info))
-#back = merge(back, patient.info[,c('Samplename','p53.Status', 'Total.Reads')])
-back$Pathology = factor(back$Pathology, levels=c('BE','ID','LGD','HGD','IMC'), ordered = T)
+back = do.call(bind_rows,lapply(unique(pg.samp$Patient) %>% map( function(p) pg.samp %>% filter(Patient == p)  ), transform.by.time, info=patient.info)) %>% mutate(Pathology = factor(Pathology, levels=c('BE','ID','LGD','HGD','IMC'), ordered = T))
 
 sq = c(0, seq(1,132,12),max(back$months.before.final))
-back$brks = cut(back$months.before.final, 
-                  breaks=sq, 
-                 labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-c(length(sq)-1,length(sq))], '>120'),
-                  include.lowest = T, ordered_result = T)
-back$brks = factor(back$brks, rev(levels(back$brks)))
-
-back$Pathology = as.character(back$Pathology)
-#back$Pathology[which(back$Pathology %in% c('HGD','IMC'))] = 'IMC/HGD'
-back$Pathology = factor(back$Pathology, levels=c('BE','ID','LGD','HGD', 'IMC'), ordered = T)
-
-ggplot(back, aes(Pathology, Prediction, group=Pathology, fill=Pathology)) + facet_grid(~Status) + 
-  geom_jitter(width=0.2) + geom_boxplot() + plot.theme  + theme(legend.position = 'none')
+
+back = back %>% mutate(
+  brks = cut(months.before.final, breaks=sq, 
+                 labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-c(length(sq)-1,length(sq))], '>120'),include.lowest = T, ordered_result = T),
+) %>% mutate(brks = factor(brks, rev(levels(brks))))
+
+ggplot(back, aes(Status, Prediction, group=Status)) + 
+  facet_grid(~Pathology) + 
+  #scale_fill_manual(values=RColorBrewer::brewer.pal(4,"Paired")[c(1,3)]) + 
+  scale_color_gradientn(colours = myPal) +
+  geom_boxplot(outlier.shape = NA, fill='#BDBDBD') + 
+  geom_jitter(width=0.2, aes(color=Prediction)) + 
+  plot.theme  + theme(legend.position = 'none')
 
@@
 
-back = back %>% filter(!is.na(Prediction))
-
-back$quants = with(back, cut(Prediction, breaks=cuts, include.lowest = T))
-qt = as.data.frame.matrix(table(back$quants, back$Status))
-qt$quant = rownames(qt)
+#back = back %>% filter(!is.na(Prediction))
+
+back = back %>% mutate(quants = cut(Prediction, breaks=cuts, include.lowest = T))
+qt = back %>% group_by(quants, Status) %>% dplyr::summarise(n=length(Status) ) %>% spread(Status, n)
 
@@
 
-qt = cbind(qt, back %>% dplyr::group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))#, 'se'=(sd(Prediction)/sqrt(length(Prediction)))*1.96 ))
-
-pred.confidence =
-  qt %>% as.data.frame.matrix %>% dplyr::group_by(quant) %>% dplyr::mutate( 
-  'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'conf'=ifelse(p.value < 0.05, '*', '') )
-
-pred.confidence[c('r1','r2')] = NA
-for (i in 1:(length(cuts)-1)) pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)
-
-pred.confidence = pred.confidence %>% dplyr::mutate(Risk = ifelse(perc<.5, 'Low','High'))
+qt = left_join(qt, back %>% dplyr::group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))#, 'se'=(sd(Prediction)/sqrt(length(Prediction)))*1.96 ))
+
+pred.confidence = qt %>% dplyr::group_by(quants) %>% 
+  dplyr::mutate( 'P.ratio'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'conf'=ifelse(p.value < 0.05, '*', '') )
+
+pred.confidence = pred.confidence %>% separate(quants, c('r1','r2'), ',', remove = F) %>% 
+  mutate_at(vars('r1','r2'), list(sub), pattern='\\[|\\]|\\(', replacement='') %>% mutate_at(vars(r1,r2), as.double)
+
+pred.confidence = pred.confidence %>% dplyr::mutate(Risk = ifelse(P.ratio<.5, 'Low','High'))
 pred.confidence = bind_cols(pred.confidence, 
@@
 
-pcfit = lm(perc~mn, data=pred.confidence)
+pcfit = lm(P.ratio~mn, data=pred.confidence)
 rsid = rstandard(pcfit)
@@
 
-annotation.table = arrange(pred.confidence %>% dplyr::group_by(Risk) %>% dplyr::summarise(
-  'n(NP)'=sum(NP), 'n(P)'=sum(P), 'p.value'=format.pval(mean(p.value), 3), 
-  'min'=min(r1), 'max'=max(r2)), min
-)
+annotation.table = pred.confidence %>% dplyr::group_by(Risk) %>% 
+  dplyr::summarise( 'n(NP)'=sum(NP), 'n(P)'=sum(P), 'p.value'=format.pval(mean(p.value), 3), 'min'=min(r1), 'max'=max(r2) ) %>%
+  arrange(min)
 
@@
 ```{r pcalib, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
-pcal = BarrettsProgressionRisk::showPredictionCalibration(pred.confidence)
+pcal = BarrettsProgressionRisk::showPredictionCalibration(pred.confidence %>% dplyr::rename(perc = P.ratio))
 pcal
@@
   arrange(Patient, Endoscopy.Year, PID)
+
 back.per.endo = back.per.endo %>% mutate(quants = cut(Prediction, breaks=cuts, include.lowest = T)) %>% left_join(pred.confidence %>% dplyr::select(quants, Risk), by='quants') %>% dplyr::select(-matches('quant')) %>% dplyr::mutate(Risk = as.character(Risk))
@@
 ```{r relrisk, echo=F, fig.width=6, fig.height=6}
-myPal = rev(RColorBrewer::brewer.pal(11, 'RdYlBu'))
-
 unadjRR = ggplot(back, aes(RR)) + geom_histogram(aes(fill=..x..), bins=10, show.legend = F) +
@@
 p = ggplot(m, aes(Status, value, fill=Status)) + facet_wrap(~variable, scales='free_y') + 
-  geom_boxplot(outlier.colour = NA) + geom_jitter(width=0.2) + plot.theme + theme(legend.position = 'none', text=element_text(size=11)) +
+  geom_boxplot(outlier.colour = NA) + geom_jitter(width=0.2) + 
+  scale_fill_brewer(palette = 'Dark2') + 
+  plot.theme + theme(legend.position = 'none', text=element_text(size=11)) +
   labs(y='', x='', title='CN values, initial endoscopy', subtitle=paste('Arms p=',wtA, '   CNA p=',wtCN, sep=''))
@@
 
-`r pander(as.data.frame(pred.confidence))`
+`r pred.confidence %>% mutate_if(is.double, list(signif), 2) %>% kable() %>% kable_styling('striped')`
 
@@
 ```{r jacknife, echo=T, warning=F, message=F, fig.height=10, fig.width=12}
-
-#list.files(cache.dir)
 
@@
 
-Effectively it looks to me like my coefficients have fairly tight CI if the SE of the jackknife variance is used.
+Coefficients have fairly tight CI if the SE of the jackknife variance is used.
 
@@
 ```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
-riskPath = rbind( cbind.data.frame( table(subset(back, Status == 'P', select=c('Pathology','Risk'))), 'Status'='Progressor' ), 
-                  cbind.data.frame( table(subset(back, Status == 'NP', select=c('Pathology','Risk'))), 'Status'='Non-Progressor' ))
-
-pathTotal = back %>% dplyr::group_by(Status, Pathology) %>% dplyr::summarise( 'nPath'=length(Pathology))
-pathTotal$Status = ifelse(pathTotal$Status == 'P', 'Progressor','Non-Progressor')
-
-riskPath = merge(riskPath, pathTotal)
-riskPath$ratio = with(riskPath, round(Freq/nPath, 3))
-riskPath$Risk = factor(riskPath$Risk, levels=c('Low','Moderate','High'), ordered = T)
-
-p53Path = cbind.data.frame(table(subset(back, Status == 'P', select=c('Pathology','p53.Status'))), 'Status'='p53 IHC')
-p53Path = merge( subset(back, Status == 'P' & !is.na(p53.Status)) %>% dplyr::group_by(Pathology) %>% dplyr::summarise( 'nPath'=length(Pathology)), p53Path)
-#p53Path$Risk = with(p53Path, ifelse(p53.Status == 1, 'High','Low'))
-p53Path$ratio = with(p53Path, round(Freq/nPath, 2))
-
-#riskPath = rbind(riskPath, p53Path[colnames(riskPath)])
-#riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor','p53 IHC'), ordered = T)
-riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor'), ordered = T)
+
+riskPath = back %>% dplyr::group_by(Status, Pathology, Risk) %>% 
+  dplyr::summarise(Freq = length(Risk)) %>% 
+  mutate(Status = recode_factor(Status, NP = 'Non-Progressor', P = 'Progressor', .ordered = T))
+
+# riskPath = rbind( cbind.data.frame( table(subset(back, Status == 'P', select=c('Pathology','Risk'))), 'Status'='Progressor' ), 
+#                   cbind.data.frame( table(subset(back, Status == 'NP', select=c('Pathology','Risk'))), 'Status'='Non-Progressor' ))
+
+pathTotal = back %>% dplyr::group_by(Status, Pathology) %>% 
+  dplyr::summarise( 'nPath'=length(Pathology)) %>%
+  mutate(Status = recode(Status, NP = 'Non-Progressor', P = 'Progressor'))
+
+riskPath = full_join(riskPath, pathTotal, by=c('Status','Pathology')) %>% 
+  mutate(ratio = round(Freq/nPath, 3), Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered = T))
+
+p53Path = back %>% dplyr::filter(Status == 'P' & !is.na(p53.Status)) %>% 
+  dplyr::group_by(Pathology,p53.Status) %>% 
+  dplyr::summarise(Freq = length(Pathology)) %>% dplyr::mutate(Status = 'p53 IHC')
+
+p53Path = full_join( back %>% dplyr::filter(Status == 'P' & !is.na(p53.Status)) %>% dplyr::group_by(Pathology) %>% dplyr::summarise(nPath = length(Pathology)), p53Path, by='Pathology') %>% 
+  dplyr::mutate(ratio = round(Freq/nPath,2))
 
@@
 ```{r p53IHC, echo=F, warning=F, message=F, fig.height=8, fig.width=8}
-
 #p53Path$Risk[p53Path$Risk == 'High'] = 'Aberrant'
@@
 p53Path$p53.Status = factor(p53Path$p53.Status, ordered = T)
-p53Path = arrange(p53Path, Pathology, p53.Status)
+p53Path =arrange(p53Path, Pathology, p53.Status)
 col = brewer.pal(11, "PRGn")[c(10,2)]
@@
                                           'High'=sum(length(which(Risk == "High"))),
-                                          'n.samples'=length(Patient) )
+                                          'n.samples'=length(Patient) ) %>% dplyr::mutate(Patient = factor(Patient))
 
 all.preds = melt(all.preds, id.vars=c('Status','Patient', 'n.samples'))
-head(all.preds)
+
 ggplot(all.preds, aes(Patient, value/n.samples, group=variable, fill=variable)) + 
@@
 ```{r enods, warning=F, message=F, echo=F, fig.height=8, fig.width=6}
+sq = c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12)
+
 endos = back %>% dplyr::group_by(Status, months.before.final, brks, PID) %>% dplyr::summarise(
@@
               'pred.samples'=length(which(Risk == 'High')),
-              'max.path'=max(Pathology))
-
-sq = c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12)
-endos$brks = cut(endos$months.before.final, include.lowest = T, ordered_result = T,
-     breaks=sq,
-     labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))
-endos$brks = with(endos, factor(brks, rev(levels(brks)))) # invert order
-
-endos.m = subset(endos, Status == 'P') %>% dplyr::group_by(brks) %>% dplyr::summarise( 
-  'pred.r'=sum(pred.endo)/sum(n.endos),
-  'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos))
-
-tmp = subset(back, Status == 'P')
-tmp$brks = cut(tmp$months.before.final, include.lowest = T, ordered_result = T,breaks=sq,
-     labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))
-tmp = tmp %>% dplyr::group_by(brks) %>% dplyr::summarise( 'mean(P)'=mean(Prediction),'sem(P)'=sd(Prediction)/sqrt(length(brks)) )
+              'max.path'=max(Pathology)) %>% ungroup %>% 
+  dplyr::mutate(brks = cut(months.before.final, include.lowest = T, ordered_result = T,
+     breaks=sq,  labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))) %>% 
+  dplyr::mutate(brks = factor(brks, rev(levels(brks)))) # invert order
+
+endos.m = endos %>% dplyr::filter(Status == 'P') %>% dplyr::group_by(brks) %>% 
+  dplyr::summarise( 'pred.r'=sum(pred.endo)/sum(n.endos), 'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos))
+
+tmp = back %>% dplyr::filter(Status == 'P') %>% 
+  dplyr::mutate(brks = cut(months.before.final, include.lowest = T, ordered_result = T,breaks=sq,
+     labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))) %>% dplyr::group_by(brks) %>% 
+       dplyr::summarise( 'mean(P)'=mean(Prediction),'sem(P)'=sd(Prediction)/sqrt(length(brks)))
 
@@
 
-rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) + ylim(0,105) +
+rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) +
   geom_bar(stat='identity', fill='darkred', alpha=0.8, show.legend = F) + 
@@
   geom_text(aes(y=pred.r*100-5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
-  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') +
+  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') + coord_cartesian(ylim = c(0,100)) + 
   plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12)) 
@@
 
-`r pander(as.data.frame(pred.confidence))`
-
 The first two quintiles and the last two together account for 80% of our samples.  Only 20% fall into the category "Moderate".
@@
 ```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=5 }
-back$nl = factor(back$nl, ordered=T)
-back$Risk = factor(back$Risk, levels=c('Low','Moderate','High'), ordered = T)
-back$Patient = factor(as.numeric(back$Patient), ordered = T)
-
-p = pred.tiles(back %>% filter(Patient == 6), colors=myPal, probs=T, ncol=1) + labs(title='')
+back = back %>% dplyr::mutate(nl = factor(nl, ordered=T), Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered = T), Patient = factor(Patient, ordered=T), Block = factor(Block, ordered=T), Pathology = recode(Pathology, BE = 'NDBE'))
+
+blank.axis = theme(axis.text = element_blank(), legend.key=element_rect(fill='grey39'), panel.background=element_rect(colour = 'black'), panel.grid.major=element_blank(), panel.spacing = unit(0.2, 'lines'), panel.border = element_rect(color="black", fill=NA, size=0.5)  )
+
+p = BarrettsProgressionRisk::patientRiskTilesPlot( back %>% filter(Patient == 6) %>% dplyr::select(Endoscopy.Year,Pathology,Block,Risk) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) ) + blank.axis
 tilesLegend = get.legend(p)
@@
 ```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=15 }
-ga = grid.arrange(
-  pred.tiles(back %>% filter(Patient %in% c(13,20,34,56,60,48) & Status == 'P'), ncol=1, colors=myPal, probs=T) + labs(title='Progressors') + theme(legend.position = 'none'),
-  pred.tiles(subset(back, Patient %in% c(15,16,21,3,45,19) & Status == 'NP'), ncol=1, colors=myPal, probs=T) + labs(title='Non-Progressors') + theme(legend.position = 'none'),
-ncol=2)
-
-ggsave('plots/ex_pt_heatmap.png', plot=ga, width=6, height=15, units='in', dpi=300)
+plist = purrr::map(c(13,20,34,56,60,48), function(pt) { 
+  BarrettsProgressionRisk::patientRiskTilesPlot( back %>% filter(Patient == pt) %>% dplyr::select(Endoscopy.Year,Pathology,Block,Risk) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) ) + blank.axis + labs(title=pt) + theme(legend.position = 'none')
+})
+a = do.call('arrangeGrob', c(plist, top='Progressors'))
+
+plist = purrr::map(c(15,16,21,3,45,19) , function(pt) { 
+  BarrettsProgressionRisk::patientRiskTilesPlot( back %>% filter(Patient == pt) %>% dplyr::select(Endoscopy.Year,Pathology,Block,Risk) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) ) + blank.axis + labs(title=pt) + theme(legend.position = 'none')
+})
+b = do.call('arrangeGrob', c(plist, top='Non-progressors'))
+
+ggsave('plots/ex_pt_heatmap.png', plot=grid.arrange(a,b,ncol=2), width=6, height=15, units='in', dpi=300)
 
@@
 
-write.table(summary.t, quote=F, sep='\t', row.names = F)
+write_tsv(summary.t, path = 'plots/risk_summary.tsv')
 ```
@@
 
-p1 = pred.tiles(back %>% filter(Status == 'NP'), colors=myPal, probs=T, nrow=2) + labs(title='Non-Progressors') 
-riskLegend = get.legend(p1)
-p1 = p1 + theme(legend.position = 'bottom')
-riskLegendLong = get.legend(p1)
-
-p1 = p1 + theme(legend.position = 'none')
-p2 = pred.tiles(back %>% filter(Status == 'P'), probs=T, colors=myPal,nrow=2) + labs(title='Progressors') + theme(legend.position = 'none')
+plist = purrr::map((sum.patient.data %>% filter(Status == 'P'))$Patient, function(pt) { 
+  BarrettsProgressionRisk::patientRiskTilesPlot( back %>% filter(Patient == pt) %>% dplyr::select(Endoscopy.Year,Pathology,Block,Risk) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) ) + blank.axis + labs(title=pt) + theme(legend.position = 'none')
+})
+p1 = do.call('arrangeGrob', c(plist, top='Progressors'))
+
+plist = purrr::map((sum.patient.data %>% filter(Status == 'NP'))$Patient, function(pt) { 
+  BarrettsProgressionRisk::patientRiskTilesPlot( back %>% filter(Patient == pt) %>% dplyr::select(Endoscopy.Year,Pathology,Block,Risk) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) ) + blank.axis + labs(title=pt) + theme(legend.position = 'none')
+})
+p2 = do.call('arrangeGrob', c(plist, top='Non-Progressors'))
 
 ggsave('plots/pred_heatmap.png', plot=grid.arrange(p2,p1,nrow=2), width=52, height=16, units="in", limitsize=F, dpi = 300)
-
 ggsave('plots/npg_heatmap.png', plot=p1, width=52, height=8, units="in", limitsize=F, dpi = 300)
@@
 
-ggsave('plots/risk_legend.png', plot=grid.arrange(riskLegend), width=2,height=4, dpi=300)
-
-ggsave('plots/risk_legend_long.png', plot=grid.arrange(riskLegendLong), width=8,height=2, dpi=300)
+#ggsave('plots/risk_legend.png', plot=grid.arrange(riskLegend), width=2,height=4, dpi=300)
+
+#ggsave('plots/risk_legend_long.png', plot=grid.arrange(riskLegendLong), width=8,height=2, dpi=300)
 
@@
 ```{r, echo=T}
-m = data.frame(matrix(nrow=nrow(chh), ncol=0))
-
-for (i in 1:nrow(chh)) {
-  feat = chh[i,]
+m = data.frame(matrix(nrow=nrow(features), ncol=0))
+
+for (i in 1:nrow(features)) {
+  feat = features[i,]
 
@@
 }
-rownames(m) = chh$label
+rownames(m) = features$label
 
@@
 subset(m, ratio.samples.P == max(ratio.samples.P))
-subset(chh, label == rownames(subset(m, ratio.samples.P == max(ratio.samples.P))))
+subset(features, label == rownames(subset(m, ratio.samples.P == max(ratio.samples.P))))
 subset(chh, label == rownames(subset(m, ratio.samples.P == min(ratio.samples.P))))
@@
 
-cor.test(chh$hazards, m$n.samples)
+cor.test(features$hazards, m$n.samples)
 
@@
 
-`r pander(genes[,c(1:3,6,7)])`
+`r (genes[,c(1:3,6,7)]) %>% kable() %>% kable_styling(full_width=F)`
 
@@
 
-# Validation Dataset
-
-20 Patients (10P, 10NP) set aside for validation purposes
-`r pander(table(sum.validation.info$Status), caption='Validation cohort')`
+# Validation Data
+
+## Test 'holdout' Set
+
+20 Patients (10P, 10NP) set aside for training/test purposes
 
@@
 validation.patient.info =  all.patient.info %>% filter(Set == 'Test') %>% arrange(Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
+sum.val.info = summarise.patient.info(validation.patient.info)
+
+sum.val.info %>% kable() %>% kable_styling()
+```
+
+```{r, echo=F}
 
@@
 
-vpd = left_join(vpd %>% dplyr::select(-Pathology, -matches('P53')), validation.patient.info %>% dplyr::select(Samplename, Patient, Pathology, p53.Status, Block, Initial.Endoscopy, Final.Endoscopy, Status ), by=c('Sample'='Samplename')) %>%
-  dplyr::rename(Endoscopy.Year = Endoscopy, Samplename = Sample, RR = `Relative Risk`, Prediction = Probability)
+vpd = left_join(vpd %>% dplyr::select(-Pathology, -matches('P53')), validation.patient.info %>% 
+                  dplyr::select(Samplename, Patient, Pathology, p53.Status, Block, Initial.Endoscopy, Final.Endoscopy, Status ), by=c('Sample'='Samplename')) %>% 
+  dplyr::rename(Samplename = Sample, Endoscopy.Year = Endoscopy, RR = `Relative Risk`, Prediction = Probability) %>% 
+  dplyr::mutate(Risk = factor(Risk, levels=levels(back$Risk), ordered = T), GEJ.Distance = factor(GEJ.Distance, ordered = T),Pathology = recode(Pathology, BE='NDBE'))
+
+vpd = do.call(bind_rows,lapply(unique(vpd$Patient) %>% map( function(p) vpd %>% filter(Patient == p)  ), transform.by.time, info=validation.patient.info))
+
+## Need to add months to each endoscopy if multiple occur in the same year
+vpd = vpd %>% dplyr::mutate(PID = factor(PID)) %>% 
+  dplyr::group_by(Patient, PID) %>% dplyr::arrange(PID) %>% 
+  dplyr::mutate( Endoscopy = as.Date(paste0(Endoscopy.Year,'-01-01')) + as.numeric(PID)  ) %>% ungroup
+
+test.roc = pROC::roc(Status~Prediction, data=vpd,auc=T, ci=T, of='thresholds',transpose=T)
+roc.plot(test.roc, title='Test Set ROC')
 
@@
 
-# Get the SD, mean, and variance(?) per sample
-# v.sampleVar = do.call(rbind, lapply(raw, function(f) {
-#   t = read.table(f, sep='\t', header=T)
-#   sampleCols = grep('chr|arm|start|end|probes', colnames(t), invert = T)
-#   x = do.call(rbind, lapply(sampleCols, function(i){
-#     cbind('Sample.Mean'=mean(t[,i]),
-#           'Sample.SD'=sd(t[,i]),
-#           'Sample.Var'=var(t[,i]))
-#   }))
-#   rownames(x) = colnames(t)[sampleCols]
-#   return(x)
-# }))
-
-# # Load segment files
-# tiled.segs = do.call(rbind, lapply(segs, function(f) {
-#   load.segment.matrix(f)
-# }))
-# dim(tiled.segs)
-# 
-# tiled.arms = do.call(rbind, lapply(arms, function(f) {
-#   load.segment.matrix(f)
-# }))
-# 
-# length(z.mean) == ncol(tiled.segs)
-# 
-# for (i in 1:ncol(tiled.segs)) 
-#   tiled.segs[,i] = unit.var(tiled.segs[,i], z.mean[i], z.sd[i])
-# 
-# for (i in 1:ncol(tiled.arms)) 
-#   tiled.arms[,i] = unit.var(tiled.arms[,i], z.arms.mean[i], z.arms.sd[i])
-# 
-# dim(tiled.arms)
-# dim(tiled.segs)
-# 
-# cx.score = unit.var(score.cx(tiled.segs,1), mn.cx, sd.cx)
-# 
-# validation.df = subtract.arms(tiled.segs, tiled.arms)
-# validation.df = cbind(validation.df, 'cx'=cx.score)
-# 
-# fauxlabel = rep(0, nrow(validation.df) )
-# names(fauxlabel) = intersect(validation.patient.info$Samplename, rownames(validation.df))
-
-```
-
-```{r validation_globalplot, message=F, warning=F, echo=F, fig.height=8, fig.width=8, eval=F}
-valCNdf = t(validation.df[,grep('^\\d+:',colnames(validation.df))])
-
-locs = do.call(rbind.data.frame, lapply(rownames(valCNdf), function(x) unlist(strsplit(x,':|-'))))
-colnames(locs) = c('chr','start','end')
-locs[] = lapply(locs[], function(x) as.numeric(as.character(x)))
-
-valCNdf = cbind.data.frame(locs, valCNdf)
-valCNdf = melt(valCNdf, id.vars=colnames(locs))
-valCNdf$chr = factor(valCNdf$chr, levels=c(1:22), ordered = T)
-
-valCNdf$seg.type = '5MB'
-valCNdf[with(valCNdf, which( start %in% arm.loc$start & end %in% arm.loc$end)), 'seg.type'] = 'arms'
-valCNdf = merge(valCNdf, chr.lengths[,c('chrom','chr.length')], by.x='chr', by.y='chrom')
-
-valCNdf = merge(valCNdf, validation.patient.info[,c('Hospital.Research.ID','Patient','Samplename','Pathology','Status')], by.x='variable', by.y='Samplename')
-valCNdf$Status = ifelse(valCNdf$Status == 'P', 'Progressor', 'Non-Progressor')
-
-valCNdf$cn = factor(sapply(valCNdf$value, cn), levels=c('loss','neutral','gain'), ordered = T)
-
-n.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(valCNdf, Status == 'Progressor')$variable)), ')', sep=''),
-              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(valCNdf, Status == 'Non-Progressor')$variable)),')', sep=''))
-# Average cn value
-valCNdf = valCNdf %>% dplyr::group_by(Status, start, end) %>% dplyr::mutate( 'mean.value' = value/length(unique(variable)) )
-valCNdf$mean.value = valCNdf$value # but not going to use it
-
-p = cn.mtn.plot(subset(valCNdf, seg.type != 'arms'), labeller(Status=n.status, chr=label_value)) + theme(legend.position = 'none') + labs(title='Validation: All samples')
-p
-ggsave('plots/validation_mtn.png', plot=p, width=10, height=5, units='in', dpi=300)
-
 ```
@@
 
-back.v = do.call(bind_rows, lapply(unique(vpd$Patient) %>% map( function(p) vpd %>% filter(Patient == p) ), transform.by.time, info=validation.patient.info)) %>% dplyr::mutate(Patient = factor(Patient))
-
 sq = c(0, seq(1,132,12),max(back$months.before.final))
-back.v$brks = cut(back.v$months.before.final, 
-                  breaks=sq, labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x,collapse='-')))[-c(length(sq)-1,length(sq))], '>120'), include.lowest = T, ordered_result = T)
-
-back.v$brks = with(back.v, factor(brks, rev(levels(brks))))
-
-#back.v$Pathology = as.character(back.v$Pathology)
-#back.v$Pathology = factor(back.v$Pathology, levels=c('BE','ID','LGD','HGD','IMC'), ordered = T)
-
+vpd = vpd %>% mutate(brks = cut(months.before.final, 
+                  breaks=sq, labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x,collapse='-')))[-c(length(sq)-1,length(sq))], '>120'), include.lowest = T, ordered_result = T)) %>%
+  mutate(brks = factor(brks, rev(levels(brks))))
 ```
@@
 ```{r echo=F, message=F, warning=F, fig.height=5, fig.width=8}
-back.v = back.v %>% dplyr::mutate(Risk = factor(Risk, levels=c('High','Moderate','Low')), nl = factor(nl, ordered = T))
-
-sq = c(0, seq(1, max(back.v$months.before.final), 12))
-back.v$brks = cut(back.v$months.before.final, 
-                  breaks=sq, 
-                 labels=c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-length(sq)],
-                  include.lowest = T, ordered_result = T)
-back.v$brks = with(back.v, factor(brks, levels=rev(levels(brks))))
-
-m = melt(back.v %>% dplyr::group_by(Patient, Status) %>% dplyr::summarise( 'Low'=length(which(Risk == 'Low'))/length(Patient),
+vpd = vpd %>% ungroup %>% dplyr::mutate(Risk = factor(Risk, levels=c('High','Moderate','Low')), nl = factor(nl, ordered = T)) %>%
+  dplyr::mutate(Patient = factor(Patient))
+
+m = melt(vpd %>% dplyr::group_by(Patient, Status) %>% dplyr::summarise( 'Low'=length(which(Risk == 'Low'))/length(Patient),
                                             'Moderate'=length(which(Risk == 'Moderate'))/length(Patient),
@@
 ```{r echo=F, warning=F, message=F, fig.height=10,fig.width=25}
-back.v$Patient = factor(as.numeric(back.v$Patient), ordered=T)
-
-p1 = pred.tiles(subset(back.v, Status == 'P'), cols=myPal, probs=T, nrow=1) + labs(title='Validation Progressors', x='') + theme(legend.position = 'bottom') 
-valLeg = get.legend(p1)
-p1 = p1 + theme(legend.position = 'none')
-p2 = pred.tiles(subset(back.v, Status == 'NP'), cols=myPal, probs=T, nrow=1) + labs(title='Validation Non-Progressors') + theme(legend.position = 'none')
-
+plist = purrr::map((sum.val.info %>% filter(Status == 'P'))$Patient, function(pt) { 
+  BarrettsProgressionRisk::patientRiskTilesPlot( vpd %>% filter(Patient == pt) %>% dplyr::select(Endoscopy,Pathology,GEJ.Distance,Risk) ) + blank.axis + labs(title=pt) + theme(legend.position = 'none')
+})
+p1 = do.call('arrangeGrob', c(plist, top='Progressors'))
+
+plist = purrr::map((sum.val.info %>% filter(Status == 'NP'))$Patient, function(pt) { 
+  BarrettsProgressionRisk::patientRiskTilesPlot( vpd %>% filter(Patient == pt) %>% dplyr::select(Endoscopy,Pathology,GEJ.Distance,Risk) ) + blank.axis + labs(title=pt) + theme(legend.position = 'none')
+})
+p2 = do.call('arrangeGrob', c(plist, top='Non-Progressors'))
 grid.arrange(p1, p2, nrow=2)
@@
 ```{r echo=F, warning=F, message=F, fig.height=5,fig.width=8}
-vig.plot(subset(back.v, Patient == 68), subset(validation.patient.info, Patient == 68), cols=riskCols)
+BarrettsProgressionRisk::patientRiskTilesPlot( vpd %>% filter(Patient == 68) %>% 
+                                                 dplyr::select(Endoscopy,Pathology,GEJ.Distance,Risk) ) + 
+  theme(legend.position = 'right')
 ```
@@
 
-## Validation prediction calibration
-
-```{r, echo=F, warning=F, message=F, fig.width=4, fig.height=3}
-cuts = seq(0,1,0.1)
-qt = as.data.frame.matrix(table(back.v$quants, back.v$Status))
-qt$quant = rownames(qt)
-
-qt = cbind(qt, back.v %>% dplyr::group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
-
-pred.confidence.v = qt %>% as.data.frame.matrix %>% dplyr::group_by(quant) %>% dplyr::mutate( 'perc'=P/sum(NP,P) )
-
-pred.confidence.v[c('r1','r2')] = NA
-for (i in 1:(length(cuts)-1)) 
-    pred.confidence.v[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)
-
-pred.confidence$Risk = factor(pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)
-
-ggplot(pred.confidence.v, aes(mn, perc)) + 
-  geom_rect(data=pred.confidence, aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
-  geom_point() +
-  #geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + 
-  labs(x='Mean prediction', y='Ratio of `P` patients', title='Validation calibration') +
-  scale_color_manual(values=riskCols) + plot.theme
-
-preds = do.call(rbind.data.frame, lapply(vpd, function(df) df[c('Status','Prediction')]))
-vroc = pROC::roc(Status ~ Prediction, data=preds, ci=T, of='thresholds')
-vroc$model = 'Validation'
-
-roc$model = 'Discovery'
-multi.roc.plot(list(roc, vroc), title='ROC', palette='Dark2') + theme(text=element_text(size=12))
-#ggsave('plots/auc_dv.png', plot=p, width=5, height=4, units='in', dpi=300, scale=2)
-```
-
-```{r, echo=F, warning=F, message=F, fig.width=4, fig.height=4}
-aucs = rbind.data.frame(pROC::ci.auc(roc), pROC::ci.auc(vroc))
-colnames(aucs) = c('ci.min', 'auc','ci.max')
-aucs$model = c('Discovery','Validation')
-
-ggplot(aucs, aes(model, auc)) + geom_point(size=2) + ylim(0,1) +
-  geom_errorbar(aes(ymin=ci.min, ymax=ci.max), width=0.1) + coord_flip() + 
-  geom_text(aes(y=ci.min, label=round(ci.min, 2)), vjust=4) +
-  geom_text(aes(y=ci.max, label=round(ci.max, 2)), vjust=-4) +
-#  geom_text(aes(y=auc, label=round(auc, 2)), vjust=-1) +
-  labs(x='', y='AUC', title='AUC with CI') + plot.theme 
-```
-
 ## Downsampled Datasets
+
+These come from Annalise, adjacent Barrett's tissues, and normal blood that had all been initially WGS at 30-80x and downsampled to an average 0.4x
 
@@
 
+## Validation Cohort
+
+We sequenced a new set of non-progressors, progressors, and some OACs from the OCCAMS cohort where we have high-depth WGS.
+
+```{r}
+val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
+analysis.dir = '~/Data/BarrettsProgressionRisk/Analysis/validation'
+
+oac.cohort = readxl::read_xlsx(val.file, 8, trim_ws = T) %>% 
+    dplyr::filter(!is.na(`SLX-ID`)) %>% dplyr::select(`Hospital Research ID`, `Block ID`, `Sample Date`, `SLX-ID`, `Index Sequence`, Status) %>% mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`)) %>% dplyr::mutate(Pathology = 'OAC') %>% dplyr::rename('Endoscopy' = `Sample Date`) %>% dplyr::rename('PID' = `Block ID`) %>% dplyr::mutate(GEJ.Distance = 1)
+
+sheets = readxl::excel_sheets(val.file)[9:12]
+info = do.call(bind_rows, lapply(sheets, function(s) {
+  readxl::read_xlsx(val.file, s, trim_ws = T) %>% 
+    dplyr::filter(!is.na(`SLX-ID`)) %>% dplyr::select(`Hospital Research ID`, `Block ID`, Status, Endoscopy, Pathology, `SLX-ID`, `Index Sequence`, `Path Notes`) %>% mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`))
+}))
+
+info = info %>% separate(`Block ID`, c('PID','GEJ.Distance'), '( |-)', remove = F, fill='right') %>% 
+  dplyr::mutate(GEJ.Distance = ifelse(is.na(GEJ.Distance), '1', GEJ.Distance)) %>%
+  dplyr::mutate(GEJ.Distance = recode(GEJ.Distance, A1 = '1', A = '1', A2 = '2', A3 = '3', A4 = '4', `4A` = '4', B = '2', C = '3', D = '4')) %>%
+  dplyr::mutate(GEJ.Distance = as.numeric(GEJ.Distance))
+
+bind_rows(
+  add_column(info %>% dplyr::select(`Hospital Research ID`, Status) %>% distinct() %>% dplyr::filter(Status %in% c('P','NP')) %>% dplyr::group_by(Status) %>% dplyr::count() %>% spread(Status, n), ' ' = 'BE Patients', .before = 1),
+  add_column(info %>% dplyr::filter(Status %in% c('P','NP')) %>% dplyr::group_by(Status) %>% dplyr::count() %>% spread(Status, n), ' ' = 'BE Samples', .before=1),
+  bind_cols(' ' = 'OAC Samples','NP'=0,'P'=nrow(oac.cohort))
+  ) %>% 
+kable(caption='Validation cohort counts') %>% kable_styling(full_width = F)
+
+```
+
+```{r}
+pred.dir = paste0(analysis.dir, '/predictions')
+
+batch8_preds = read_tsv(paste0(pred.dir, '/batch8_preds.txt')) %>% mutate(`Hospital Research ID` = gsub('_','/',`Hospital Research ID`)) %>% dplyr::select(`Hospital Research ID`, Sample, Probability, `Relative Risk`, Risk) %>%
+  left_join(oac.cohort, by=c('Hospital Research ID','Sample'))
+
+oldp.preds = read_tsv(paste0(pred.dir, '/oldp_preds.txt')) %>% mutate(`Hospital Research ID` = gsub('_','/',`Hospital Research ID`)) %>% dplyr::select(`Hospital Research ID`, Sample, Probability, `Relative Risk`, Risk) %>% left_join(info, by=c('Hospital Research ID','Sample'))
+
+new.np.preds = read_tsv(paste0(pred.dir, '/val_np.txt')) %>% mutate(`Hospital Research ID` = gsub('_','/',`Hospital Research ID`)) %>% dplyr::select(`Hospital Research ID`, Sample, Probability, `Relative Risk`, Risk) %>% left_join(info, by=c('Hospital Research ID','Sample'))
+
+val.preds = bind_rows(
+  batch8_preds %>% dplyr::select(`Hospital Research ID`, Probability, `Relative Risk`, Endoscopy, Status, PID) %>% mutate(Pathology = 'OAC', GEJ.Distance = 1) %>% dplyr::rename(ID = 'Hospital Research ID'),
+  oldp.preds %>% dplyr::select(`Hospital Research ID`, Probability, `Relative Risk`, Endoscopy, GEJ.Distance, Pathology, PID) %>% mutate(Status = 'P') %>% dplyr::rename(ID = 'Hospital Research ID'),
+  new.np.preds %>% dplyr::select(`Hospital Research ID`, Probability, `Relative Risk`, Endoscopy, GEJ.Distance, Pathology, PID) %>% mutate(Status = 'NP') %>% dplyr::rename(ID = 'Hospital Research ID') )
+
+val.preds = val.preds %>% mutate(quants = cut(Probability, breaks=cuts, include.lowest = T))
+
+val.preds = left_join(val.preds, pred.confidence %>% dplyr::select(quants, Risk), by='quants')
+
+ggplot(val.preds, aes(Status, Probability)) + geom_boxplot(aes(group=Status, fill=Status), outlier.shape = NA) + geom_jitter(width = 0.2) + plot.theme
+
+ggplot(val.preds, aes(Pathology, Probability)) + facet_grid(~Status,scales = 'free_x') + geom_boxplot(aes(group=Pathology), outlier.shape = NA) + geom_jitter(width = 0.3, aes(color=Probability, shape=Risk), size=3) + scale_color_gradientn(colors=myPal) + plot.theme
+
+```
+### Validation AUC
+
+```{r, echo=F, warning=F, message=F, fig.width=4, fig.height=3}
+
+val.preds %>% dplyr::select(Status, Probability)
+
+preds = do.call(rbind.data.frame, lapply(vpd, function(df) df[c('Status','Prediction')]))
+vroc = pROC::roc(Status ~ Probability, data=val.preds, ci=T, of='thresholds', transpose=T)
+vroc$model = 'Validation'
+
+roc$model = 'Discovery'
+multi.roc.plot(list(roc, vroc), title='ROC', palette='Dark2') + theme(text=element_text(size=12))
+ggsave('plots/auc_dv.png', plot=p, width=5, height=4, units='in', dpi=300, scale=2)
+```
+
+```{r, echo=F, warning=F, message=F, fig.width=4, fig.height=4}
+aucs = rbind.data.frame(pROC::ci.auc(roc), pROC::ci.auc(vroc))
+colnames(aucs) = c('ci.min', 'auc','ci.max')
+aucs$model = c('Discovery','Validation')
+
+ggplot(aucs, aes(model, auc)) + geom_point(size=2) + ylim(0,1) +
+  geom_errorbar(aes(ymin=ci.min, ymax=ci.max), width=0.1) + coord_flip() + 
+  geom_text(aes(y=ci.min, label=round(ci.min, 2)), vjust=4) +
+  geom_text(aes(y=ci.max, label=round(ci.max, 2)), vjust=-4) +
+#  geom_text(aes(y=auc, label=round(auc, 2)), vjust=-1) +
+  labs(x='', y='AUC', title='AUC with CI') + plot.theme 
+```
+
+```{r}
+dir.create('plots/validation')
+
+val.preds$Patient =  group_indices(val.preds, ID)
+
+val.preds = val.preds %>% group_by(ID) %>% mutate(PID = as.factor(PID)) %>% ungroup %>% group_by(ID, PID) %>% arrange(PID) %>% dplyr::mutate( Endoscopy = Endoscopy + as.integer(PID)  ) %>% ungroup
+
+plist = purrr::map(val.preds %>% filter(Pathology == 'OAC') %>% dplyr::select(ID) %>% distinct %>% pull, function(pt) { 
+  BarrettsProgressionRisk::patientRiskTilesPlot(
+    val.preds %>% filter(ID == pt) %>% mutate(GEJ.Distance = factor(GEJ.Distance)) ) + 
+    labs(title=pt) +
+    theme(legend.position = 'none')
+})
+a = do.call('arrangeGrob', c(plist, top='OAC'))
+ggsave('plots/validation/oac.png', plot=grid.arrange(a), width=3*3, height=4*2,dpi=300)
+
+plist = purrr::map(val.preds %>% filter(Status == 'P' & Pathology != 'OAC') %>% dplyr::select(ID) %>% distinct %>% pull, function(pt) { 
+  BarrettsProgressionRisk::patientRiskTilesPlot(
+    val.preds %>% filter(ID == pt) %>% mutate(GEJ.Distance = factor(GEJ.Distance)) ) + 
+    labs(title=pt) +
+    theme(legend.position = 'none')
+})
+b = do.call('arrangeGrob', c(plist, top='Progressors'))
+ggsave('plots/validation/progessors.png', plot=grid.arrange(b), width=4*4, height=4*2,dpi=300)
+
+plist = purrr::map(val.preds %>% filter(Status == 'NP') %>% dplyr::select(ID) %>% distinct %>% pull, function(pt) { 
+  BarrettsProgressionRisk::patientRiskTilesPlot(
+    val.preds %>% filter(ID == pt) %>% mutate(GEJ.Distance = factor(GEJ.Distance)) ) + 
+    labs(title=pt) +
+    theme(legend.position = 'none')
+})
+c = do.call('arrangeGrob', c(plist, top='Non Progressors', ncol=4))
+grid.arrange(c)
+ggsave('plots/validation/non-progessors.png', plot=grid.arrange(c), width=4*4, height=11*2,dpi=300, limitsize = F)
+
+val.preds %>% group_by(Status, Risk) %>% dplyr::count(Risk) %>% spread(Risk, n) %>% kable() %>% kable_styling(full_width = F)
+```
+
 # Clinical Story
@@
 
-pander(as.data.frame(summary.v))
+as.data.frame(summary.v) %>% kable() %>% kable_styling(full_width = F)
 

### 2019-08-16 lib/common_plots.R

@@ -1,4 +1,11 @@
 
+myPal = rev(RColorBrewer::brewer.pal(11, 'RdYlBu'))
+
+plot.theme = theme(text=element_text(size=12), panel.background=element_blank(), strip.background =element_rect(fill="white"),  
+                   strip.text = element_text(size=12), 
+                   axis.line=element_line(color='black'), panel.grid.major=element_line(color='grey90'),
+                   panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing = unit(0.1, 'lines')  ) 
+
 
 get.legend<-function(a.gplot){
   tmp <- ggplot_gtable(ggplot_build(a.gplot))
@@ -6,11 +13,6 @@ get.legend<-function(a.gplot){
   legend <- tmp$grobs[[leg]]
   return(legend)
 }
-
-plot.theme = theme(text=element_text(size=12), panel.background=element_blank(), strip.background =element_rect(fill="white"),  
-                   strip.text = element_text(size=12), 
-                   axis.line=element_line(color='black'), panel.grid.major=element_line(color='grey90'),
-                   panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing = unit(0.1, 'lines')  ) 
 
 cn.mtn.plot<-function(df, label, type='bar') {
   df = arrange(df, chr, start)
@@ -117,7 +119,7 @@ roc.plot <- function(roc,title="") {
 
 multi.roc.plot <- function(rocList, title="ROC", palette=NULL, colors=NULL) {
   aucs = do.call(rbind, lapply(rocList,function(r) 
-    cbind.data.frame('AUC'=r$auc,'model'=r$model,'Specificity'=pROC::coords(r,'best')[['specificity']], 'Sensitivity'=pROC::coords(r,'best')[['sensitivity']])))
+    cbind.data.frame('AUC'=r$auc,'model'=r$model,'Specificity'=pROC::coords(r,'best',transpose=T)[['specificity']], 'Sensitivity'=pROC::coords(r,'best',transpose=T)[['sensitivity']])))
   aucs$y = aucs$Sensitivity
   aucs$x = aucs$Specificity
 

### 2019-08-16 Downsampled Predictions.Rmd
...

---
title: "Downsampled"
author: "Sarah Killcoyne"
date: "14/08/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)

library(tidyverse)
library(BarrettsProgressionRisk)
library(kableExtra)
source('lib/common_plots.R')

dir = '~/Data/BarrettsProgressionRisk/Analysis/downsampled/predictions'
info = '~/Data/BarrettsProgressionRisk/QDNAseq/all_downsampled/downsampled_ids.xlsx'

info = readxl::read_xlsx(info)

files = list.files(dir, 'Rdata', recursive = T, full.names = T)

pfile = paste0(dir, '/ds_preds.txt')
if (file.exists(pfile)) {
  ds.preds = read_tsv(pfile)
} else {
  ds.preds = do.call(bind_rows, lapply(files, function(f) {
    load(f, verbose=T)
    bind_cols(predictions(prr), adjustRisk(prr, 'mean', 'sample')$adj.predictions %>% dplyr::select(Probability, `Relative Risk`) %>% dplyr::rename_all(funs(paste0), ' Adjusted ') )
  }))
  write_tsv(ds.preds, pfile)
}

ds.preds = ds.preds %>% dplyr::select(-Status) %>% left_join(info %>% dplyr::select(`Illumina ID`, Status), by=c('Sample' = 'Illumina ID')) %>% dplyr::mutate(Type = factor(Type, levels=c('Normal','NDBE', 'LGD', 'Tumor Adjacent BE'))) %>% arrange(Probability)

```

## Downsampled Datasets

These samples are from Annalise's cohort and the OCCAMS tumor-adjacent Barrett's cohort.  Additionally, I downsampled 40 normal blood WGS as a non-progressor test set. A total of `r nrow(info)` samples were downsampled from 30-50X WGS to an average 0.4X then run through the QDNAseq pipeline.

```{r}
info %>% dplyr::group_by(Status, Type) %>% summarise(n = length(Type)) %>% spread(Type,n) %>% kable() %>% kable_styling(full_width = F)
```

They were then predicted using the model trained on the initial training cohort.

```{r echo=F, eval=T}
myPal = rev(RColorBrewer::brewer.pal(11, 'RdYlBu'))

ggplot(ds.preds, aes(`Relative Risk`)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1), show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') +
    plot.theme + labs(title='Sample predictions', y='n Samples') + theme(legend.position = 'none')

# roc = pROC::roc(Status ~ Probability, data=ds.preds, ci=T, of='thresholds', transpose=T)
# pROC::coords(roc, 'best', ret='threshold', transpose=T)
# roc.plot(roc)
```

```{r fig.height=12}
m = reshape2::melt(ds.preds, measure.vars = c('Relative Risk','Relative Risk Adjusted'), id.vars = c('Type', 'Status', 'Risk'))

ggplot(m, aes(variable, value)) + 
  facet_wrap(~Type, ncol=2, scales = 'free_x') + 
  geom_jitter(width=0.2, aes(shape=Status, color=Risk), size=2) + 
  geom_boxplot(aes(group=variable, fill=Type), outlier.colour = NA, show.legend = F) + 
  scale_color_manual(values =  BarrettsProgressionRisk::riskColors()) +
  #scale_color_gradientn(colors = myPal,  name='') +
  scale_fill_manual(values=rev(RColorBrewer::brewer.pal(5,'RdYlBu')[c(1,2,4,5)])) +
  labs(x='', y='Relative Risk', title='Predictions on downsampled patients') + plot.theme +
  theme(axis.text.x = element_text(angle=45, hjust=1))

ds.preds %>% group_by(Status, Risk) %>% summarise(n=length(Risk)) %>% spread(Risk, n) %>% kableExtra::kable(caption = 'Sample by risk and known status') %>% kableExtra::kable_styling(full_width = F)

```

### 2019-08-16 Validation Report.Rmd
...

@@
 load(model, verbose=T)
-
-fit = models$`0.9`
-lambda = performance.at.1se$`0.9`$lambda
 
@@
 
+```{r echo=F}
+
+alpha = '0.9'
+fit = models[[alpha]]
+lambda = performance.at.1se[[alpha]]$lambda
+
+```
+
 ## FFPE vs Frozen (old vs new as well)
@@
 
-if (file.exists(paste0(pred.dir, '/batch8_preds.txt'))) {
-  batch8_preds = read_tsv(paste0(pred.dir, '/batch8_preds.txt'))
+file = paste0(pred.dir, '/batch8_preds_alpha', alpha,'.txt')
+
+if (file.exists(file)) {
+  batch8_preds = read_tsv(file)
 } else {
@@
 
-  write_tsv(batch8_preds, paste0(pred.dir, '/batch8_preds.txt'))
+  write_tsv(batch8_preds, path=file)
 }
@@
 
-sheets = readxl::excel_sheets(val.file)[9:12]
+sheets = readxl::excel_sheets(val.file)[9:13]
 info = do.call(bind_rows, lapply(sheets, function(s) {
@@
 
-if (file.exists(paste0(pred.dir, '/oldp_preds.txt'))) {
-  oldp.preds = read_tsv(paste0(pred.dir, '/oldp_preds.txt'))
+file = paste0(pred.dir, '/oldp_preds_alpha', alpha, '.txt')
+if (file.exists(file)) {
+  oldp.preds = read_tsv(file)
 } else {
@@
   }))
-  write_tsv(oldp.preds, paste0(pred.dir, '/oldp_preds.txt'))
+  write_tsv(oldp.preds, path=file)
 }
@@
 
-if (file.exists(paste0(pred.dir, '/val_np.txt'))) {
-  new.np.preds = read_tsv(paste0(pred.dir, '/val_np.txt'))
+file = paste0(pred.dir, '/val_np_alpha', alpha, '.txt')
+if (file.exists(file)) {
+  new.np.preds = read_tsv(file)
 } else {
@@
   new.np.preds = do.call(bind_rows, lapply(segFiles, function(f) {
-    print(f)
+  #  print(f)
     load(f)  
@@
   }))
-  write_tsv(new.np.preds, paste0(pred.dir, '/val_np.txt'))
+  write_tsv(new.np.preds, path=file)
 }

### 2019-08-16 processing_pipeline/predict_validation.R

@@ -1,0 +1,54 @@
+args = commandArgs(trailingOnly=TRUE)
+
+if (length(args) < 1)
+  stop("Missing required params: <data dir> <sample info file> <model dir> <outdir> <alpha=0.9 DEF>")
+
+library(BarrettsProgressionRisk)
+source('~/workspace/shallowwgs_pipeline/lib/load_patient_metadata.R')
+
+datadir = args[1]
+info.file = args[2]
+model.dir = args[3]
+outdir = args[4]
+
+select.alpha = '0.9'
+if (length(args) == 5) {
+  select.alpha = args[5]
+  
+  if (!select.alpha %in% c(0.0, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0))
+    stop("Alpha values available: 0.0, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0")
+}
+
+outdir = paste0(outdir, '/', select.alpha)
+dir.create(outdir, showWarnings = F, recursive = T)
+
+
+x = list.files(model.dir, 'all.pt.alpha.Rdata', recursive = T, full.names = T)
+load(x, verbose=F)
+fit = models[[select.alpha]]
+s = performance.at.1se[[select.alpha]]$lambda  
+
+
+sheets = readxl::excel_sheets(info.file)[9:13]
+info = do.call(bind_rows, lapply(sheets, function(s) {
+  print(s)
+  readxl::read_xlsx(info.file, s, trim_ws = T) %>% 
+    dplyr::select(`Hospital Research ID`, `Block ID`, Endoscopy, Pathology, `SLX-ID`, `Index Sequence`, `Path Notes`) %>% dplyr::mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`))
+}))
+
+
+segFiles = grep(paste(unique(info$`Hospital Research ID`),collapse='|'), list.files(datadir, '[2|3|4]_segObj',  full.names = T, recursive = T), value=T)
+
+preds = do.call(bind_rows, lapply(segFiles, function(f) {
+  load(f)  
+  segmented$sample.info = BarrettsProgressionRisk::loadSampleInformation(info %>% filter(Sample %in% segmented$sample.info$Sample) )
+  
+  prr = BarrettsProgressionRisk::predictRiskFromSegments(segmented, model = fit, s = lambda, verbose = F)
+  predictions(prr)
+}))
+
[…]

### 2019-08-20 Paper.Rmd
...

@@
 Based on HGD/IMC/LGD
-
 ```{r pathroc, echo=F, fig.height=4, fig.width=12 }
@@
 ### Per Patient ROC
-
 ```{r echo=F, fig.height=4, fig.width=12}
-
 # Excluding the final endoscopy
@@
 
+ggplot(back, aes(Status, RR, group=Status)) + 
+  facet_grid(~Pathology) + 
+  #scale_fill_manual(values=RColorBrewer::brewer.pal(4,"Paired")[c(1,3)]) + 
+  scale_color_gradientn(colours = myPal) +
+  geom_boxplot(outlier.shape = NA, fill='#BDBDBD') + 
+  geom_jitter(width=0.2, aes(color=Prediction)) + 
+  plot.theme  + theme(legend.position = 'none')
+
 ```
@@
 cuts = seq(0,1,0.1)
-
-#back = back %>% filter(!is.na(Prediction))
 
@@
 
-# Tumour Supressors
-
-```{r ts, eval=F}
-colnames(ccgenes)
-
-tsgs = ccgenes
-nrow(tsgs)
-
-tsgs = makeGRangesFromDataFrame(tsgs[,c(colnames(loc), 'Gene.Symbol')], keep.extra.columns = T)
-tsgs = sort(tsgs)
-
-length(tsgs)
-
-# average 1 per 9MB
-avgTSGdist = mean(elementMetadata(distanceToNearest(tsgs))$distance)
-
-chh = arrange(chh, -hazards)
-features = makeGRangesFromDataFrame(chh, keep.extra.columns = T)
-
-distanceTSG = data.frame(matrix(nrow=0,ncol=3))
-for (i in 1:length(features)) {
-  f = features[i]
-  hits = nearest(f, tsgs, ignore.strand=T, select='all')
-  if (length(hits) > 0) {
-    distanceTSG = rbind(distanceTSG, cbind.data.frame(
-      'features'=i,
-      'nearestTSG'=elementMetadata(tsgs[subjectHits(hits)])$Gene.Symbol,
-      'distanceToTSG'=elementMetadata(distanceToNearest(f, tsgs, ignore.strand=T, select='all'))$distance)
-    )
-  }
-}
-
-features[unique(subset(distanceTSG, distanceToTSG == 0)$features)]
-
-# In our dataset we should see 1 TSG/onc per 10MB
-# But this is double the actual number we see
-ncol(dysplasia.df)/round(avgTSGdist/5e6)
-
-length(which(round((subset(distanceTSG, distanceToTSG > 0)$distanceToTSG/5e6)-0.5) == round((avgTSGdist/5e6)-0.5)))
-```
-
 # Validation Data
+
+## Normals sWGS sequenced
+
+These normal tissues came from some samples that had been mistakenly sequenced from a timepoint (path ID) that Barrett's samples were also sequenced.  These include:
+
+```{r, echo=F}
+normal.info = readxl::read_xlsx(patient.file, sheet='Normals')
+
+normal.info %>% dplyr::select(`Patient ID`, `Path ID`, `Pathology notes`) %>% kable(caption='Normals') %>% kable_styling('striped', full_width = F)
+```
+
+These were excluded from the model building procedure, but were then predicted using the resulting model.
+We would expect them to all be low risk and all but one are.
+All have a relative risk below 0 as well.
+
+```{r}
+
+normal.preds = read_tsv('~/Data/BarrettsProgressionRisk/Analysis/multipcf_plots_normals/normal_predictions.tsv', col_types = cols(
+  Sample = col_character(),
+  Probability = col_double(),
+  `Relative Risk` = col_double(),
+  Risk = col_character(),
+  Patient = col_character(),
+  Hospital.Research.ID = col_character(),
+  Path.ID = col_character(),
+  Status = col_character(),
+  Endoscopy = col_double(),
+  Pathology = col_logical(),
+  Plate.Index = col_character(),
+  SLX.ID = col_double(),
+  Barretts.Cellularity = col_logical(),
+  P53.Status = col_character(),
+  Total.Reads = col_double(),
+  Batch.Name = col_character(),
+  Set = col_character(),
+  Block = col_double()
+))
+
+ggplot(normal.preds, aes(Patient, Probability, color=Risk)) + ylim(0,1) + geom_point(size=2) +
+  scale_color_manual(values = BarrettsProgressionRisk::riskColors() ) + plot.theme
+
+```
 
@@
 
-ggsave(filename='plots/val_leg.png', plot=valLeg, width=4, height=2, units='in', dpi=300)
+#ggsave(filename='plots/val_leg.png', plot=valLeg, width=4, height=2, units='in', dpi=300)
 
@@
 ```{r echo=F, eval=T}
-dir = '~/Data/BarrettsProgressionRisk/Analysis/downsampled/predictions'
+dir = '~/Data/BarrettsProgressionRisk/Analysis/downsampled/predictions/0.9'
 info = '~/Data/BarrettsProgressionRisk/QDNAseq/all_downsampled/downsampled_ids.xlsx'
@@
 
-files = list.files(dir, 'Rdata', recursive = T, full.names = T)
-
-pfile = paste0(dir, '/ds_preds.txt')
-if (file.exists(pfile)) {
-  ds.preds = read_tsv(pfile)
-} else {
-  ds.preds = do.call(bind_rows, lapply(files, function(f) {
-    load(f, verbose=T)
-    bind_cols(predictions(prr), adjustRisk(prr, 'mean', 'sample')$adj.predictions %>% dplyr::select(Probability, `Relative Risk`) %>% dplyr::rename_all(funs(paste0), ' Adjusted ') )
-  }))
-  write_tsv(ds.preds, pfile)
-}
+ds.preds = read_tsv(paste0(dir, '/downsampled_preds.tsv'), col_types = cols(
+  Sample = col_character(),
+  Probability = col_double(),
+  `Relative Risk` = col_double(),
+  Risk = col_character(),
+  PatientID = col_character(),
+  Status = col_character(),
+  Tissue = col_character(),
+  `Expected Risk` = col_character(),
+  Grade_patient = col_character(),
+  Grade_biopsy = col_character(),
+  Study = col_character(),
+  Type = col_character(),
+  Endoscopy = col_date(format = "")
+))
 
@@
 
-m = reshape2::melt(ds.preds, measure.vars = c('Relative Risk','Relative Risk Adjusted'), id.vars = c('Type', 'Status', 'Risk'))
+m = reshape2::melt(ds.preds, measure.vars = c('Relative Risk'), id.vars = c('Type', 'Status', 'Risk'))
 
@@
   facet_grid(~Type, scales = 'free_x') + 
+  geom_boxplot(aes(group=variable, fill=Type), outlier.colour = NA, show.legend = F) + 
   geom_jitter(width=0.2, aes(shape=Status, color=Risk), size=2, show.legend = F) + 
-  geom_boxplot(aes(group=variable, fill=Type), outlier.colour = NA, show.legend = F) + 
   scale_color_manual(values =  BarrettsProgressionRisk::riskColors()) +

### 2019-08-20 Validation Report.Rmd
...

--- 2019-08-07 Validation Report.Rmd
@@
 analysis.dir = '~/Data/BarrettsProgressionRisk/Analysis/validation'
+
+pred.dir = '~/Data/BarrettsProgressionRisk/Analysis/validation/predictions'
 
@@
 
-sheets = readxl::excel_sheets(val.file)[1:12]
+sheets = readxl::excel_sheets(val.file)[1:13]
 all.val = do.call(bind_rows, lapply(sheets, function(s) {
@@
 
-sval = val %>% group_by(Batch) %>% dplyr::mutate(Cohort = paste(unique(Cohort), collapse=', ' )) %>% group_by(Batch, Cohort) %>% dplyr::summarise_at(vars(varMAD_median), funs(min,mean,median,max),na.rm=T)
+sval = val %>% group_by(Batch) %>% dplyr::mutate(Cohort = paste(unique(Cohort), collapse=', ' )) %>% group_by(Batch, Cohort) %>% 
+  dplyr::summarise_at(vars(varMAD_median), list(~ min(., na.rm=T), ~ mean(., na.rm=T) ,~ median(., na.rm=T),~ max(., na.rm=T)))
 
@@
 Red rows are where the median value of the variance for the entire batch was greater than the cutoff') %>% kable_styling('striped') %>% row_spec(which(sval$median > qc.cutoff), color='red')
+
+val %>% dplyr::group_by(Batch) %>% dplyr::summarise( Pass = length(which(varMAD_median <= qc.cutoff)), Fail = length(which(varMAD_median > qc.cutoff))  ) %>% kable(caption = 'Number of samples per batch that pass/fail') %>% kable_styling('striped', full_width = F)
 
@@
   Pass=length(which(varMAD_median<=qc.cutoff)), Fail=length(which(varMAD_median>qc.cutoff))
-) %>% rename('as.integer(RA)' = 'Processing RA') %>% kable(caption='Pass/Fail by sample type') %>% kable_styling('striped')
+) %>% dplyr::rename('Processing RA' = 'as.integer(RA)' ) %>% kable(caption='Pass/Fail by sample type') %>% kable_styling('striped')
 
@@
   `Total Samples` = n(),
-  `% Passed QC` = round((length(which(varMAD_median<=qc.cutoff))/n())*100,1)
+  `Passed QC` = length(which(varMAD_median<=qc.cutoff))
 ) %>% kable() %>% kable_styling('striped')
@@
   Total = n(),
-  `% Passed QC` = round((length(which(varMAD_median<=qc.cutoff))/n())*100,1)
+  `Passed QC` = paste0(length(which(varMAD_median<=qc.cutoff)), ' (', round((length(which(varMAD_median<=qc.cutoff))/n())*100,1), '%)')
 ) %>% kable(caption='Total samples and QC results') %>% kable_styling('striped')
@@
   Total = n(),
-  `% Passed QC` = round((length(which(varMAD_median<=qc.cutoff))/n())*100,1)
+  `Passed QC` = paste0(length(which(varMAD_median<=qc.cutoff)), ' (', round((length(which(varMAD_median<=qc.cutoff))/n())*100,1), '%)')
 ) %>% kable() %>% kable_styling('striped')
@@
 on = left_join(val %>% ungroup %>% filter(Batch %in% c(8)),
-               readxl::read_xlsx(val.file, sheet='Batch 8') %>% dplyr::select(`Block ID`, `Sample Date`, `Sample Type`))
+               readxl::read_xlsx(val.file, sheet='Batch 8') %>% dplyr::select(`Block ID`, `Endoscopy`, `Sample Type`))
 
@@
 
-on = on %>% mutate( Age = ifelse( `Sample Date` < m18, '<18m', '>=18m') )
+on = on %>% mutate( Age = ifelse( Endoscopy < m18, '<18m', '>=18m') )
 
@@
   Total = n(),
-  `% Passed QC` = round((length(which(varMAD_median<=qc.cutoff))/n())*100,1)
+  `Passed QC` = paste0(length(which(varMAD_median<=qc.cutoff)), ' (', round((length(which(varMAD_median<=qc.cutoff))/n())*100,1), '%)')
 ) %>% kable() %>% kable_styling('striped')
 
-if (file.exists('batch8_preds.txt')) {
-  batch8_preds = read_tsv('batch8_preds.txt')
-} else {
-  segFiles = grep(paste(on$`Patient ID`,collapse='|'), list.files('~/Data/BarrettsProgressionRisk/Analysis/validation', 'segObj',  full.names = T, recursive = T), value=T)
+val.info = do.call(bind_rows, lapply(sheets[8:13], function(s) {
+  readxl::read_xlsx(val.file, s) %>% dplyr::select(-matches('Extraction|Qbit'))%>% dplyr::filter(!is.na(`SLX-ID`)) %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% fncols('Collection', 'Biopsy')
+    #dplyr::select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA, matches('Collection')) 
+})) %>% mutate(
+  `Hospital Research ID` = str_replace_all( str_remove_all(`Hospital Research ID`, " "), '/', '_'), 
+  `Index Sequence` = str_replace_all(`Index Sequence`, 'tp', ''),
+  Samplename = paste(`SLX-ID`,`Index Sequence`,sep='.')
+  )
 
-  batch8_preds = do.call(bind_rows, lapply(segFiles, function(f) {
-    print(f)
-    load(f)
-    if (nrow(sampleResiduals(segmented)) <= 0) return()
+preds.dir = paste0(analysis.dir, '/predictions/0.9')
 
-    segmented$sample.info = loadSampleInformation(  
-      on %>% filter(sample %in% segmented$sample.info$Sample) %>%
-        dplyr::rename('Endoscopy' = `Sample Date`, Sample = sample ) )
+batch8_preds = do.call(bind_rows,
+  purrr::map(grep( paste(unique(on$`Patient ID`), collapse='|'), list.files(preds.dir, full.names = T, recursive = T), value=T), function(f) {
+             read_tsv(f, col_types = cols(
+                Sample = col_character(),
+                Probability = col_double(),
+                `Relative Risk` = col_double(),
+                Risk = col_character(),
+                `Hospital Research ID` = col_character(),
+                `Block ID` = col_character(),
+                Endoscopy = col_date(format = ""),
+                Pathology = col_character(),
+                `SLX-ID` = col_character(),
+                `Index Sequence` = col_character(),
+                `Path Notes` = col_logical()
+      ))
+})) %>% dplyr::mutate_if(is.numeric, list(round), 2)
 
-    prr = BarrettsProgressionRisk::predictRiskFromSegments(segmented, F)
-    BarrettsProgressionRisk::predictions(prr,'sample')
-  }))
-  batch8_preds = batch8_preds %>% filter(!is.na(`Patient ID`))
-
-  write_tsv(batch8_preds, 'batch8_preds.txt')
-}
-
-batch8_preds %>% dplyr::select(`Patient ID`, Sample, everything(), -matches('Hospital|SLX|Index') ) %>% kable() %>% kable_styling('striped')
+batch8_preds %>% dplyr::select(`Hospital Research ID`, Sample, everything(), -matches('SLX|Index') ) %>% kable() %>% kable_styling('striped')
 
@@
 
-oldp = val %>% ungroup %>% filter(Batch %in% c(8:12) & Cohort == 'Old progressor scrolls')
+oldp = val %>% ungroup %>% filter(Batch %in% c(8:13) & Cohort == 'Old progressor scrolls')
 
@@
   `n Samples` = n(),
-  `% Passed QC` = round((length(which(varMAD_median<=qc.cutoff))/n())*100,1)
+`Passed QC` = paste0(length(which(varMAD_median<=qc.cutoff)), ' (', round((length(which(varMAD_median<=qc.cutoff))/n())*100,1), '%)')
 ) %>% kable() %>% kable_styling('striped')
@@
 
+The predictions from these samples are not particularly good.
+
 ```{r}
+p.diag = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/validation/progressor_diagnosis.xlsx') %>% 
+  dplyr::mutate(`Hospital Research ID` = gsub('/', '_', `Hospital Research ID`), `Weeks Follow Up` = as.double(difftime(`Diagnostic Endoscopy`, `Sequenced Endoscopy`, units='weeks')))
 
-sheets = readxl::excel_sheets(val.file)[9:12]
-info = do.call(bind_rows, lapply(sheets, function(s) {
-  readxl::read_xlsx(val.file, s, trim_ws = T) %>% 
-    dplyr::filter(Status == 'P') %>% dplyr::select(`Hospital Research ID`, `Block ID`, Endoscopy, Pathology, `SLX-ID`, `Index Sequence`, `Path Notes`) %>% dplyr::mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`))
-}))
-
-if (file.exists('oldp_preds.txt')) {
-  oldp.preds = read_tsv('oldp_preds.txt')
-} else {
-  segFiles = grep(paste(oldp$`Patient ID`,collapse='|'), list.files('~/Data/BarrettsProgressionRisk/Analysis/validation', 'segObj',  full.names = T, recursive = T), value=T)
-
-#  info = readxl::read_xlsx(val.file, 'Batch 9') %>% dplyr::select(`Hospital Research ID`, `Block ID`, Endoscopy, Pathology, `SLX-ID`, `Index Sequence`, `Path Notes`) %>% mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`))
-
-  oldp.preds = do.call(bind_rows, lapply(segFiles, function(f) {
-    load(f)  
-    segmented$sample.info = BarrettsProgressionRisk::loadSampleInformation(info %>% filter(Sample %in% segmented$sample.info$Sample) )
-
-    prr = BarrettsProgressionRisk::predictRiskFromSegments(segmented,F)
-    predictions(prr)
-  }))
-  write_tsv(oldp.preds, 'oldp_preds.txt')
-}
+oldp.preds = do.call(bind_rows,
+  purrr::map(grep( paste(unique(oldp$`Patient ID`), collapse='|'), list.files(preds.dir, full.names = T, recursive = T), value=T), function(f) {
+             read_tsv(f,col_types = cols(
+                Sample = col_character(),
+                Probability = col_double(),
+                `Relative Risk` = col_double(),
+                Risk = col_character(),
+                `Hospital Research ID` = col_character(),
+                `Block ID` = col_character(),
+                Endoscopy = col_date(format = ""),
+                Pathology = col_character(),
+                `SLX-ID` = col_character(),
+                `Index Sequence` = col_character(),
+                `Path Notes` = col_character()
+      ))
+})) %>% dplyr::mutate_if(is.numeric, list(round), 2) %>% dplyr::select(-Endoscopy)
 
 oldp.preds = left_join(oldp.preds, val %>% dplyr::select(sample, varMAD_median, n.segs), by=c('Sample'='sample'))
-
+oldp.preds = left_join(oldp.preds, val.info %>% dplyr::select(Samplename, Endoscopy), by=c('Sample'='Samplename'))
 oldp.preds = oldp.preds %>% filter(Sample %in% (oldp %>% filter(!grepl('OCCAMS',`Hospital Research ID`)) %>% dplyr::select(sample) %>% pull))
 
-oldp.preds %>% dplyr::summarise(
-  'TPR'=length(which(Risk == 'High'))/n(),
-  'FNR'=length(which(Risk == 'Low'))/n()) %>% kable() %>% kable_styling(full_width = F)
+oldp.preds = left_join(oldp.preds, p.diag, by=c('Hospital Research ID', 'Block ID', 'Endoscopy' = 'Sequenced Endoscopy'))
 
-oldp.preds %>% group_by(`Risk`) %>% dplyr::summarise( 'n samples'=n() ) %>% 
-  kable(caption='Count sample predicted risks, P') %>% kable_styling(full_width = F)
+oldp.preds %>% group_by(`Risk`) %>% dplyr::summarise( 'n samples'=n() ) %>% spread(Risk, `n samples`) %>%
+  kable(caption='Progressor sample predicted risks') %>% kable_styling(full_width = F)
 
-fnr.patients = oldp.preds %>% 
-  group_by(`Hospital Research ID`) %>% dplyr::summarise(
-  TPR = length(which(Risk == 'High'))/n()
-) %>% dplyr::summarise( 'Misclassified patients' = length(which(TPR < 0.5))/n() ) %>% pull
+ggplot(oldp.preds, aes(`Hospital Research ID`, Probability, color=Probability)) + geom_point(size=2) +
+  scale_color_distiller(palette = 'RdYlBu', name='P(P)') + plot.theme + theme(axis.text.x = element_text(angle=45,hjust=1))
+```
 
+```{r}
 oldp.preds %>% group_by(`Hospital Research ID`) %>% dplyr::summarise(
@@
   'Low Risk' =   length(which(Risk == 'Low')),
-  'Pathology Notes' = paste(`Path Notes`, collapse='; ')
-) %>% kable(caption = 'Classified predictions per patient (all P)') %>% kable_styling('striped')
+  'Pathology Notes' = paste(unique(`Path Notes`), collapse='; ')
+) %>% dplyr::arrange(desc(`High Risk`), desc(`Mod. Risk`), desc(`Low Risk`)) %>% 
+  mutate(
+  `High Risk` = cell_spec(`High Risk`, "html", color = ifelse(`High Risk` > 0, "red", "grey39")),
+  `Mod. Risk` = cell_spec(`Mod. Risk`, "html", color = ifelse(`Mod. Risk` > 0, "orange", "grey39")),
+  `Low Risk` = cell_spec(`Low Risk`, "html", color = ifelse(`Low Risk` > 0, "blue", "grey39")),
+) %>% kable(format='html',escape = F,caption = 'Classified predictions per patient (all Progressors)') %>% kable_styling('striped', full_width = F)
+
+oldp.preds = oldp.preds %>% separate(`Block ID`, c('PID','Block'), '\\s?(-|\\s)', remove = F) 
+```
+
+However, it looks very likely that the age of the sample is related to the predictability.
+Roughly speaking, samples over 3.5 years old appear more likely to predict as low risk.
+This wasn't so true of Ellie's samples.
+These samples did sit as parafiin scrolls for a year after being cut as well and this could have affected the samples.
+
+```{r}
+ggplot(oldp.preds, aes(`Weeks Follow Up`, Probability, color=Probability)) + geom_point(size=2) +
+  scale_color_distiller(palette = 'RdYlBu', name='P(P)') + plot.theme + labs(title='Progressors', subtitle='Predictions by follow up time')
+
+oldp.per.endo = 
+  oldp.preds %>% group_by(`Hospital Research ID`, PID) %>% 
+  dplyr::mutate('n.samples'=length(PID),'Samples'=paste(Sample,collapse=',')) %>% 
+  dplyr::summarise_at(vars(matches('Prob|Relative|Endo|n.samples')), list(mean)) %>% 
+  dplyr::select(-matches('^Sample$')) %>% 
+  arrange(`Hospital Research ID`, Probability, Endoscopy)
+
+fnr.patients = oldp.preds %>% 
+  group_by(`Hospital Research ID`) %>% dplyr::summarise(
+  FNR = length(which(Risk == 'Low'))/n() ) %>% 
+  dplyr::summarise( 'Misclassified patients' = length(which(FNR > 0.5))/n() ) %>% pull %>% round(2)
 
@@
 ```{r}
+new.np = val %>% ungroup %>% filter(Batch %in% c(10:13) & Status == 'NP')
 
-new.np = val %>% ungroup %>% filter(Batch %in% c(10:12) & Status == 'NP')
+new.np.preds = do.call(bind_rows,
+  purrr::map(grep( paste(unique(new.np$`Patient ID`), collapse='|'), list.files(preds.dir, full.names = T, recursive = T), value=T), function(f) {
+             read_tsv(f,col_types = cols(
+                Sample = col_character(),
+                Probability = col_double(),
+                `Relative Risk` = col_double(),
+                Risk = col_character(),
+                `Hospital Research ID` = col_character(),
+                `Block ID` = col_character(),
+                Endoscopy = col_date(format = ""),
+                Pathology = col_character(),
+                `SLX-ID` = col_character(),
+                `Index Sequence` = col_character(),
+                `Path Notes` = col_character()
+      ))
+})) %>% dplyr::mutate_if(is.numeric, list(round), 2) %>% dplyr::select(-Endoscopy)
+
+new.np.preds = left_join(new.np.preds, val %>% dplyr::select(sample, varMAD_median, n.segs), by=c('Sample'='sample'))
+new.np.preds = left_join(new.np.preds, val.info %>% dplyr::select(Samplename, Endoscopy), by=c('Sample'='Samplename'))
 
@@
 ```{r}
-sheets = readxl::excel_sheets(val.file)[10:12]
-info = do.call(bind_rows, lapply(sheets, function(s) {
-  readxl::read_xlsx(val.file, s, trim_ws = T) %>% 
-    dplyr::filter(Status == 'NP' & !is.na(`SLX-ID`)) %>% dplyr::select(`Hospital Research ID`, `Block ID`, Endoscopy, Pathology, `SLX-ID`, `Index Sequence`, `Path Notes`) %>% mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`))
-}))
 
-if (file.exists('val_np.txt')) {
-  new.np.preds = read_tsv('val_np.txt')
-} else {
-  segFiles = grep(paste(new.np$`Patient ID`,collapse='|'), list.files('~/Data/BarrettsProgressionRisk/Analysis/validation', '(2|3)_segObj',  full.names = T, recursive = T), value=T)
-
-  new.np.preds = do.call(bind_rows, lapply(segFiles, function(f) {
-    print(f)
-    load(f)  
-    if (nrow(BarrettsProgressionRisk::sampleResiduals(segmented) %>% dplyr::filter(Pass)) <= 0) return()
-
-    segmented$sample.info = BarrettsProgressionRisk::loadSampleInformation(info %>% filter(Sample %in% segmented$sample.info$Sample) )
-    prr = BarrettsProgressionRisk::predictRiskFromSegments(segmented,F)
-    predictions(prr)
-  }))
-  write_tsv(new.np.preds, 'val_np.txt')
-}
-
-new.np.preds = left_join(new.np.preds, val %>% dplyr::select(sample, varMAD_median, n.segs), by=c('Sample'='sample'))
-
-#cor.test(new.np.preds$n.segs, new.np.preds$varMAD_median)
-
-new.np.preds %>% dplyr::summarise(
-  'FPR'=length(which(Risk == 'High'))/n(),
-  'TPR'=length(which(Risk == 'Low'))/n()) %>% kable() %>% kable_styling(full_width = F)
-
-new.np.preds %>% group_by(`Risk`) %>% dplyr::summarise( 'n samples'=n() ) %>% 
+new.np.preds %>% group_by(`Risk`) %>% dplyr::summarise( 'n samples'=n() ) %>% spread(Risk, `n samples`) %>% 
   kable(caption='Count sample predicted risks, NP') %>% kable_styling(full_width = F)
-
-fpr.patients = new.np.preds %>%  group_by(`Hospital Research ID`) %>% dplyr::summarise(
-  TPR = length(which(Risk == 'Low'))/n()
-) %>% dplyr::summarise( 'Misclassified patients' = length(which(TPR < 0.5))/n() ) %>% pull
 
@@
 
-with(val.new.np.preds, cor.test(varMAD_median, Probability))
+new.np.preds = new.np.preds %>% dplyr::mutate( 'Weeks Since Endoscopy' = as.double(difftime( as.Date('2019/06/01'), Endoscopy, units='weeks' )))
+
+ggplot(new.np.preds, aes(`Hospital Research ID`, Probability, color=Probability)) + geom_point(size=2) +
+  scale_color_distiller(palette = 'RdYlBu', name='P(P)') + plot.theme + theme(axis.text.x = element_text(angle=45,hjust=1))
+```
+
+Time since the endoscopy was performed doesn't appear to have any effect on the non-progressor predictions
+
+```{r}
+ggplot(new.np.preds, aes(`Weeks Since Endoscopy`, Probability, color=Probability)) + geom_point(size=2) +
+  scale_color_distiller(palette = 'RdYlBu', name='P(P)') + plot.theme + labs(title='Non-progressors', subtitle='Predictions vs time since the endoscopy')
 
@@
   'QC var' = signif( mean(varMAD_median), 2)
-) %>% kable(caption = 'Classified predictions per patient (all NP)') %>% kable_styling('striped')
+) %>% dplyr::arrange( `Low Risk`, `Mod. Risk`, `High Risk`) %>% 
+  mutate(
+  `High Risk` = cell_spec(`High Risk`, "html", color = ifelse(`High Risk` > 0, "red", "grey39")),
+  `Mod. Risk` = cell_spec(`Mod. Risk`, "html", color = ifelse(`Mod. Risk` > 0, "orange", "grey39")),
+  `Low Risk` = cell_spec(`Low Risk`, "html", color = ifelse(`Low Risk` > 0, "blue", "grey39")),
+) %>%
+  kable(format='html',escape = F, caption = 'Classified predictions per patient (all NP)') %>% kable_styling('striped', full_width = F)
+
+new.np.preds = new.np.preds %>% separate(`Block ID`, c('PID','Block'), '\\s?(-|\\s)', remove = F)
+
+np.per.endo = 
+  new.np.preds %>% group_by(`Hospital Research ID`, PID) %>% 
+  dplyr::mutate('n.samples'=length(PID),'Samples'=paste(Sample,collapse=',')) %>% 
+  dplyr::summarise_at(vars(matches('Prob|Relative|Endo|n.samples')), list(max)) %>% dplyr::select(-matches('^Sample$')) %>% 
+  arrange(`Hospital Research ID`, Probability, PID)
 
 ```
+
+## All together
+
+The predictions show a very poor AUC as the split shown above would make you expect.
+
+```{r}
+
+all = bind_rows(
+  oldp.preds %>% dplyr::select(Sample, Probability, `Relative Risk`, Risk) %>% dplyr::mutate(Status = 'P'),
+  new.np.preds %>% dplyr::select(Sample, Probability, `Relative Risk`, Risk) %>% dplyr::mutate(Status = 'NP')
+)
+
+roc = pROC::roc(Status ~ Probability, data = all, auc=T, ci=T, of='thresholds',transpose=T)
+roc.plot(roc)
+#pROC::coords(roc, 'best', transpose=T)
+```
+
+```{r, fig.height=8}
+
+p1 = ggplot(all, aes(Probability)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1), show.legend = F) + 
+  scale_fill_distiller(palette = 'RdYlBu', name='P(P)') +
+  plot.theme + labs(title='Sample probability', y='n Samples', x='Probability') + theme(legend.position = 'none')
+
+p2 = ggplot(all, aes(`Relative Risk`)) + geom_histogram(aes(fill=..x..), bins=10, show.legend = F) +
+  scale_fill_gradientn(colors = myPal,  name='') + 
+  labs(y='n Samples', x='Relative Risk', title='Unadjusted relative risk') + plot.theme
+
+grid.arrange(p1,p2)
+
+```

### 2019-08-20 processing_pipeline/predict_downsampled.R

@@ -29,6 +29,10 @@ dir.create(outdir, showWarnings = F, recursive = T)
 if (!file.exists(info)) stop(paste0("File does not exist or is not readable: ", info))
 all.ds.info = readxl::read_xlsx(info)
 
+x = list.files(model.dir, 'model_data.Rdata', recursive = T, full.names = T)
+load(x, verbose=T)
+rm(dysplasia.df, labels)
+
 x = list.files(model.dir, 'all.pt.alpha.Rdata', recursive = T, full.names = T)
 load(x, verbose=F)
 fit = models[[select.alpha]]
@@ -63,7 +67,7 @@ p1 = BarrettsProgressionRisk::plotSegmentData(segObj, 'plot')
 
 ggsave(paste0(pred.dir, '/', sample, '_segmentedCoverage.png'), plot=grid.arrange(p1),  width=20, height=6, units='in', limitsize=F)
 
-prr = BarrettsProgressionRisk::predictRiskFromSegments(segObj, model = fit, s = s)
+prr = BarrettsProgressionRisk::predictRiskFromSegments(segObj, model=fit, s=s, tile.mean = z.mean, tile.sd = z.sd, arms.mean = z.arms.mean, arms.sd = z.arms.sd, cx.mean = mn.cx, cx.sd = sd.cx)
 
 save(prr, file=paste0(pred.dir, '/predictions.Rdata'))
 

### 2019-08-20 processing_pipeline/predict_validation.R

@@ -28,6 +28,10 @@ dir.create(outdir, showWarnings = F, recursive = T)
 
 print(paste0("Output path:", outdir))
 
+x = list.files(model.dir, 'model_data.Rdata', recursive = T, full.names = T)
+load(x, verbose=T)
+rm(dysplasia.df, labels)
+
 x = list.files(model.dir, 'all.pt.alpha.Rdata', recursive = T, full.names = T)
 load(x, verbose=F)
 fit = models[[select.alpha]]
@@ -50,7 +54,7 @@ preds = do.call(bind_rows, lapply(segFiles, function(f) {
   load(f)  
   segmented$sample.info = BarrettsProgressionRisk::loadSampleInformation(info %>% filter(Sample %in% segmented$sample.info$Sample) )
   
-  prr = BarrettsProgressionRisk::predictRiskFromSegments(segmented, model = fit, s = lambda, verbose = F)
+  prr = BarrettsProgressionRisk::predictRiskFromSegments(segObj, model=fit, s=s, tile.mean = z.mean, tile.sd = z.sd, arms.mean = z.arms.mean, arms.sd = z.arms.sd, cx.mean = mn.cx, cx.sd = sd.cx, verbose = F)
   save(prr, file = paste0(outdir, '/predictions.Rdata'))
   predictions(prr)
 }))

### 2019-08-20 Paper.Rmd
...

@@
 
-ds.preds = read_tsv(paste0(dir, '/downsampled_preds.tsv'), col_types = cols(
+ds.preds = do.call(bind_rows, purrr::map(list.files(dir, 'predictions', recursive=T, full.names=T), function(f) {
+  read_tsv(f, col_types = cols(
   Sample = col_character(),
@@
   Type = col_character(),
-  Endoscopy = col_date(format = "")
-))
+  Endoscopy = col_date(format = "") ))
+}))
 

### 2019-08-20 Validation Report.Rmd
...

@@
 analysis.dir = '~/Data/BarrettsProgressionRisk/Analysis/validation'
-
-pred.dir = '~/Data/BarrettsProgressionRisk/Analysis/validation/predictions'
 
@@
 
-all.val %>% filter(!Samplename %in% val$sample )
+#all.val %>% filter(!Samplename %in% val$sample )
 
 val = val %>% filter(sample %in% (all.val %>% dplyr::select(Samplename) %>% pull) ) %>% 
-          left_join( all.val, val %>% mutate(sample = str_replace(sample, '_', '-')), by=c('sample'='Samplename'))
+          left_join( all.val, val %>% mutate(sample = str_replace(sample, '_', '-')), by=c('sample'='Samplename')) %>% unique
 
@@
 
-# Quick Summary
+# Quick QC Summary
 
@@
 val %>% dplyr::group_by(Batch) %>% dplyr::summarise( Pass = length(which(varMAD_median <= qc.cutoff)), Fail = length(which(varMAD_median > qc.cutoff))  ) %>% kable(caption = 'Number of samples per batch that pass/fail') %>% kable_styling('striped', full_width = F)
-
 ```
@@
 
-# New processing
+# New Validation Cohorts
 
@@
 ```{r}
-
-on = left_join(val %>% ungroup %>% filter(Batch %in% c(8)),
-               readxl::read_xlsx(val.file, sheet='Batch 8') %>% dplyr::select(`Block ID`, `Endoscopy`, `Sample Type`))
+on = left_join(val %>% ungroup %>% filter(Batch == 8),
+               readxl::read_xlsx(val.file, sheet='Batch 8') %>% dplyr::select(`Block ID`, `Endoscopy`, `Sample Type`, `Index Sequence`),
+               by=c('Block ID','Sample Type', 'Index Sequence')) %>% 
+  arrange(`Patient ID`, sample)
 
@@
                 `Index Sequence` = col_character(),
-                `Path Notes` = col_logical()
+                `Path Notes` = col_character()
       ))
@@
 
-batch8_preds %>% dplyr::select(`Hospital Research ID`, Sample, everything(), -matches('SLX|Index') ) %>% kable() %>% kable_styling('striped')
+batch8_preds = left_join(batch8_preds,on %>% dplyr::select(sample, `Sample Type`, Age), by=c('Sample'='sample'))
+
+batch8_preds %>% dplyr::select(`Hospital Research ID`, Sample, everything(), -matches('SLX|Index') ) %>% arrange(`Hospital Research ID`, `Sample Type`) %>% kable() %>% kable_styling('striped')
 
@@
 
-```{r, fig.height=8}
+```{r, fig.height=8, eval=F}
 
@@
 ```
+
+```{r echo=F, eval=F}
+dir = '~/Data/BarrettsProgressionRisk/Analysis/downsampled/predictions/0.9'
+info = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/all_downsampled/downsampled_ids.xlsx')
+
+ds.preds = do.call(bind_rows, purrr::map(list.files(dir, 'predictions', recursive=T, full.names=T), function(f) {
+  read_tsv(f, col_types = cols(
+  Sample = col_character(),
+  Probability = col_double(),
+  `Relative Risk` = col_double(),
+  Risk = col_character(),
+  PatientID = col_character(),
+  Status = col_character(),
+  Tissue = col_character(),
+  `Expected Risk` = col_character(),
+  Grade_patient = col_character(),
+  Grade_biopsy = col_character(),
+  Study = col_character(),
+  Type = col_character(),
+  Endoscopy = col_date(format = "") ))
+}))
+
+## TODO merge in RR from 'back'
+
+ds.preds = ds.preds %>% dplyr::select(-Status) %>% left_join(info %>% dplyr::select(`Illumina ID`, Status), by=c('Sample' = 'Illumina ID')) %>% dplyr::mutate(Type = factor(Type, levels=c('Normal','NDBE', 'LGD', 'Tumor Adjacent BE'))) %>% arrange(Probability)
+
+ m = reshape2::melt(ds.preds, measure.vars = c('Relative Risk'), id.vars = c('Type', 'Status', 'Risk'))
+
+pr = ggplot(m, aes(variable, value)) + 
+  facet_grid(~Type, scales = 'free_x') + 
+  geom_boxplot(aes(group=variable, fill=Type), outlier.colour = NA, show.legend = F) + 
+  geom_jitter(width=0.2, aes(shape=Status, color=Risk), size=2, show.legend = F) + 
+  scale_color_manual(values =  BarrettsProgressionRisk::riskColors()) +
+  #scale_color_gradientn(colors = myPal,  name='') +
+  scale_fill_manual(values=rev(RColorBrewer::brewer.pal(5,'RdYlBu')[c(1,2,4,5)])) +
+  labs(x='', y='Relative Risk', title='Predictions on downsampled patients') + plot.theme +
+  theme(axis.text.x = element_text(angle=45, hjust=1))
+pr
+
+ds.preds %>% dplyr::group_by(Status, Risk) %>% dplyr::summarise(n=length(Risk)) %>% spread(Risk, n) %>% kableExtra::kable(caption = 'Sample by risk and known status') %>% kableExtra::kable_styling(full_width = F)
+
+```

### 2019-08-20 processing_pipeline/predict.R

@@ -37,6 +37,10 @@ info = BarrettsProgressionRisk::loadSampleInformation(
 )
 
 
+x = list.files(model.dir, 'model_data.Rdata', recursive = T, full.names = T)
+load(x, verbose=T)
+rm(dysplasia.df, labels)
+
 x = list.files(modeldir, 'all.pt.alpha.Rdata', recursive = T, full.names = T)
 load(x, verbose=F)
 fit = models[[select.alpha]]
@@ -62,8 +66,7 @@ if (length(segFile) <= 0) {
   
   segs$sample.info = info
 }
-
-prr = predictRiskFromSegments(segs, model = fit, s = s)
+prr = BarrettsProgressionRisk::predictRiskFromSegments(segs, model = fit, s = s, tile.mean = z.mean, tile.sd = z.sd, arms.mean = z.arms.mean, arms.sd = z.arms.sd, cx.mean = mn.cx, cx.sd = sd.cx, verbose = F)
 
 pred.dir = paste0(outdir, '/predictions')
 print(pred.dir)

### 2019-08-30 Paper.Rmd
trying fourier transform smoothing

@@
 
-# for (i in 1:nrow(sum.patient.data)) {
-#   print(i)
-#   pt = sum.patient.data[i, 'Patient']
-#   status = sum.patient.data[i, 'Status']
-# 
-#   pms = mtn.spatial(subset(globalCNdf, Patient == pt & seg.type != 'arms'), back, ogjLimits, paste('Patient', pt, '(', status, ')' ))
-#   
-#   fl = paste('patient_', pt, '.png', sep='')
-#   ggsave(paste('plots/spatial/', fl, sep=''), plot=grid.arrange(pms$grob), width=14, height=6, units='in', dpi=300)
-# }
+for (i in 1:nrow(sum.patient.data)) {
+  print(i)
+  pt = sum.patient.data[i, 'Patient']
+  status = sum.patient.data[i, 'Status']
+
+  pms = mtn.spatial(subset(globalCNdf, Patient == pt & seg.type != 'arms'), back, ogjLimits, paste('Patient', pt, '(', status, ')' ))
+
+  fl = paste('patient_', pt, '.png', sep='')
+  ggsave(paste('plots/spatial/', fl, sep=''), plot=grid.arrange(pms$grob), width=14, height=6, units='in', dpi=300)
+}
 
@@
 
-predicted = '~/Data/BarrettsProgressionRisk/Analysis/training/val_predictions'
+predicted = '~/Data/BarrettsProgressionRisk/Analysis/training/val_predictions/0.9'
 

### 2019-08-30 Validation Report.Rmd
trying fourier transform smoothing

@@
 source('lib/common_plots.R')
+source('lib/load_patient_metadata.R')
 
@@
 val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
-analysis.dir = '~/Data/BarrettsProgressionRisk/Analysis/validation'
+analysis.dir = '~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf'
 
@@
 batch8_preds = do.call(bind_rows,
-  purrr::map(grep( paste(unique(on$`Patient ID`), collapse='|'), list.files(preds.dir, full.names = T, recursive = T), value=T), function(f) {
+  purrr::map(grep( paste(unique(on$`Patient ID`), collapse='|'), list.files(preds.dir, 'predictions.tsv', full.names = T, recursive = T), value=T), function(f) {
              read_tsv(f, col_types = cols(
@@
 oldp.preds = do.call(bind_rows,
-  purrr::map(grep( paste(unique(oldp$`Patient ID`), collapse='|'), list.files(preds.dir, full.names = T, recursive = T), value=T), function(f) {
+  purrr::map(grep( paste(unique(oldp$`Patient ID`), collapse='|'), list.files(preds.dir, 'predictions.tsv', full.names = T, recursive = T), value=T), function(f) {
              read_tsv(f,col_types = cols(
@@
 new.np.preds = do.call(bind_rows,
-  purrr::map(grep( paste(unique(new.np$`Patient ID`), collapse='|'), list.files(preds.dir, full.names = T, recursive = T), value=T), function(f) {
+  purrr::map(grep( paste(unique(new.np$`Patient ID`), collapse='|'), list.files(preds.dir, 'predictions.tsv', full.names = T, recursive = T), value=T), function(f) {
              read_tsv(f,col_types = cols(
@@
   arrange(`Hospital Research ID`, Probability, PID)
+
+fpr.patients = new.np.preds %>% 
+  group_by(`Hospital Research ID`) %>% dplyr::summarise(
+  FPR = length(which(Risk == 'High'))/n() ) %>% 
+  dplyr::summarise( 'Misclassified patients' = length(which(FPR > 0.5))/n() ) %>% pull %>% round(2)
 
@@
 ```
+
+## Compared to the 'test' cohort
+
+These are the patients that Ellie and I randomly assigned as a test set.
+They are not part of the model training (in this version) and so are a good comparison for the validation data.
+These are what we might have expected to see.
+
+```{r}
+
+predicted = '~/Data/BarrettsProgressionRisk/Analysis/training/val_predictions/0.9'
+
+info.dir = '~/Data/BarrettsProgressionRisk/QDNAseq'
+
+patient.file = list.files(info.dir, pattern='All_patient_info.xlsx', recursive=T, full.names=T)
+demo.file = list.files(info.dir, pattern='Demographics_full.xlsx', recursive=T, full.names=T)
+
+validation.patient.info = read.patient.info(patient.file, demo.file, set='Test')$info %>% 
+  dplyr::select(Hospital.Research.ID, Patient, Path.ID, Status, Endoscopy.Year, Pathology, Plate.Index, SLX.ID, Samplename)
+
+vpd = do.call(bind_rows, lapply(list.files(predicted, full.names = T), function(f) {
+  read_tsv(f, col_types = 'cddcccdccdc')
+})) %>% dplyr::select(Hospital.Research.ID, everything())
+
+vpd = left_join(vpd %>% dplyr::select(-Pathology, -matches('P53')), 
+                validation.patient.info %>% 
+                  dplyr::select(Hospital.Research.ID, Patient, Status) %>% unique, by=c('Hospital.Research.ID'))
+
+vpd %>% dplyr::group_by(Status, Risk) %>% dplyr::summarise(n=length(Risk)) %>% spread(Risk, n) %>% kableExtra::kable(caption = "Test set samples (Ellie's)") %>% kableExtra::kable_styling(full_width = F)
+
+roc = pROC::roc(Status ~ Probability, data = vpd, auc=T, ci=T, of='thresholds',transpose=T)
+roc.plot(roc)
+
+```

### 2019-09-04 Validation Report.Rmd
mean norm using cohort values

@@
 analysis.dir = '~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf'
+preds.dir = paste0(analysis.dir, '/predictions/fourier')
 
@@
 })) %>% dplyr::select(`Patient ID`, sample, varMAD_median, n.segs) %>% mutate(varMAD_median = round(varMAD_median, 3)) %>% mutate(sample = str_replace(sample, '_', '-'))
-
-#all.val %>% filter(!Samplename %in% val$sample )
 
@@
   )
-
-preds.dir = paste0(analysis.dir, '/predictions/0.9')
 

### 2019-09-06 Validation Report.Rmd
Rank adjust validation?

@@
 analysis.dir = '~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf'
-preds.dir = paste0(analysis.dir, '/predictions/fourier')
+preds.dir = paste0(analysis.dir, '/predictions/0.9')
 
@@
   scale_color_distiller(palette = 'RdYlBu', name='P(P)') + plot.theme + theme(axis.text.x = element_text(angle=45,hjust=1))
+
+ggplot(new.np.preds, aes(`Hospital Research ID`, `Relative Risk`, color=Probability)) + geom_point(size=2) +
+  scale_color_distiller(palette = 'RdYlBu', name='RR') + plot.theme + theme(axis.text.x = element_text(angle=45,hjust=1))
+
 ```
@@
 #pROC::coords(roc, 'best', transpose=T)
+
+ggplot(all, aes(Status, `Relative Risk`, color=`Relative Risk`, group=Status)) + geom_boxplot() + geom_jitter(width=0.1)
+
 ```

### 2019-09-06 Raw Data Comparison.Rmd
...

---
title: "Raw Data Comparison"
author: "Sarah Killcoyne"
date: "23/08/2019"
output: 
  html_document: 
    fig_width: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(BarrettsProgressionRisk)
library(reshape2)

chr.lengths = BarrettsProgressionRisk:::chrInfo()
blacklist = readr::read_tsv(system.file("extdata", "qDNAseq_blacklistedRegions.txt", package="BarrettsProgressionRisk"), col_names=T, col_types='cii')

load('~/Data/BarrettsProgressionRisk/QDNAseq/training/merged_qdnaseq_output.Rdata',verbose=T)
tfit.data = fit.data %>% as_tibble %>% mutate(chrom = factor(chrom, levels = chr.lengths$chr, ordered=T)) %>% arrange(chrom, start)
traw.data = raw.data %>% as_tibble %>% mutate(chrom = factor(chrom, levels = chr.lengths$chr, ordered=T)) %>% arrange(chrom, start)
rm(fit.data, raw.data)

val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
sheets = readxl::excel_sheets(val.file)[8:13]
all.val = do.call(bind_rows, lapply(sheets, function(s) {
  readxl::read_xlsx(val.file, s) %>% dplyr::select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA, matches('Collection')) %>% dplyr::filter(!is.na(`SLX-ID`)) %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character))# %>% fncols('Collection', 'Biopsy')
})) %>% mutate(Samplename = paste(`SLX-ID`,`Index Sequence`, sep='.'))

load('~/Data/BarrettsProgressionRisk/QDNAseq/validation/merged_raw_fit.Rdata',verbose=T)
vfit.data = merged.fit %>% as_tibble %>% mutate(chrom = factor(chrom, levels = chr.lengths$chr, ordered=T)) %>% 
  arrange(chrom, start) %>% dplyr::select(location,chrom,start,end, all.val$Samplename)
vraw.data = merged.raw %>% as_tibble %>% mutate(chrom = factor(chrom, levels = chr.lengths$chr, ordered=T)) %>% 
  arrange(chrom, start) %>% dplyr::select(location,chrom,start,end, all.val$Samplename)
rm(merged.fit, merged.raw)

dim(vfit.data)
dim(tfit.data)
```

# Raw counts & fitted values

This is the raw data that gets segmented (after some processing).  It looks like there may be a slight shift in the distribution (per patient) of these values but it's not extremely high.

```{r, fig.width=12, fig.height=12}
# par(mfrow=c(4,2))
# qqnorm( na.omit(tfit.data$D701_D501_10722_11823), main='Training sample' ) 
# qqnorm( na.omit(vfit.data$`SLX-17976.N705-S508`), main='Validation sample' )
# 
# qqnorm( na.omit(tfit.data$D701_D505_10722_11823), main='Training sample' ) 
# qqnorm( na.omit(vfit.data$`SLX-17977.N706-S510`), main='Validation sample' )
# 
# qqnorm(na.omit(tfit.data$D706_D502_13696), main='Training sample')
# qqnorm(na.omit(vfit.data$`SLX-16758.N703-S511`), main='Validation sample') # this was one of the OAC samples
# 
# qqnorm(na.omit(tfit.data$D712_D504_13692), main='Training sample') # P sample, correctly predicted
# qqnorm(na.omit(vfit.data$`SLX-16783.N715-S505`), main='Validation sample') # P sample, correctly predicted

summarise.per.pt<-function(df) {
  means = df %>% dplyr::select(-matches('location|chrom|start|end')) %>% summarise_all( list(~mean(.,na.rm=T))  ) 
  sds = df %>% dplyr::select(-matches('location|chrom|start|end')) %>% summarise_all( list(~sd(.,na.rm=T))  ) 
  vars = df %>% dplyr::select(-matches('location|chrom|start|end')) %>% summarise_all( list(~var(.,na.rm=T))  ) 
  iqrs = df %>% dplyr::select(-matches('location|chrom|start|end')) %>% summarise_all( list(~IQR(.,na.rm=T))  )

  disp = full_join(
           full_join( as_tibble(t(means), rownames='sample') %>% rename(V1='mean'), 
                      as_tibble(t(sds), rownames='sample') %>% rename(V1='sd'), by='sample'),
           full_join(as_tibble(t(vars), rownames='sample') %>% rename(V1='var'),
                     as_tibble(t(iqrs), rownames='sample') %>% rename(V1='IQR'), by='sample'), 
           by='sample') 
  return(disp)
}

t.fit.disp = summarise.per.pt(tfit.data) %>% mutate(cohort='train')
v.fit.disp = summarise.per.pt(vfit.data) %>% mutate(cohort='val')

wilcox.test(t.fit.disp$mean, v.fit.disp$mean)

fit.disp = bind_rows(t.fit.disp, v.fit.disp)

gridExtra::grid.arrange(
  ggplot(fit.disp, aes(x=cohort, y=mean, group=cohort, fill=cohort)) + geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none'),
  ggplot(fit.disp, aes(x=cohort, y=sd, group=cohort, fill=cohort)) + geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none'),
  ncol=2, top='Raw fitted values')

t.raw.disp = summarise.per.pt(traw.data) %>% mutate(cohort='train')
v.raw.disp = summarise.per.pt(vraw.data) %>% mutate(cohort='val')

wilcox.test(t.raw.disp$mean, v.raw.disp$mean)

raw.disp = bind_rows(t.raw.disp, v.raw.disp)

gridExtra::grid.arrange(
  ggplot(raw.disp, aes(x=cohort, y=mean, group=cohort, fill=cohort)) + geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none'),
  ggplot(raw.disp, aes(x=cohort, y=sd, group=cohort, fill=cohort)) + geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none'), 
  ncol=2, top='Raw count values')

rm(tfit.data,traw.data,vfit.data,vraw.data)
```

# First processing steps
```{r}
if (file.exists('prepped.Rdata')) {
  load('prepped.Rdata')
} else {
  t.prepped = BarrettsProgressionRisk:::.prepRawSWGS(traw.data, tfit.data, blacklist, verbose=T, plot=F)
  dim(t.prepped$data)

  v.prepped = BarrettsProgressionRisk:::.prepRawSWGS(vraw.data, vfit.data, blacklist, verbose=T, plot=F)
  dim(v.prepped$data)

  t.sdevs = t.prepped$sdevs
  v.sdevs = v.prepped$sdevs
  rm(t.prepped,v.prepped)

  save(t.sdevs, v.sdevs, file='prepped.Rdata')  
}

# The sdevs are used to adjust the gamma values for segmentation
wilcox.test(t.sdevs, v.sdevs)

sdevs = bind_rows(tibble('sdev'=t.sdevs, 'cohort'=rep('train',length(t.sdevs))),
          tibble('sdev'=v.sdevs, 'cohort'=rep('val',length(v.sdevs))))

# Validation has one extreme value, but this is per-sample so it should only affect that sample
ggplot(sdevs, aes(x=cohort, y=sdev, group=cohort, fill=cohort)) + ylim(0,0.5) +
  geom_jitter(width=0.1) + geom_violin(alpha=0.5) + 
  theme(legend.position = 'none') + labs(title='sdevs')

ggplot(sdevs, aes(sdev)) + geom_histogram(stat='density',bins=50, color='lightblue') + xlim(0,0.5) + facet_wrap(~cohort)
```

# Segmented Data

General distributions. These are per-patient on the segmented (but not yet binned) data. It looks like there is a shift in the per-patient distributions.

```{r, fig.height=10, fig.width=10}

t.disp = read_tsv(list.files('~/Data/BarrettsProgressionRisk/Analysis/multipcf_perPatient', pattern='disp', full.names = T, recursive = T), col_types = 'cddddddd') %>% mutate(cohort='train')
v.disp = read_tsv(list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf', pattern='disp', full.names = T, recursive = T), col_types = 'cddddddd') %>% mutate(cohort='val')

v.disp %>% filter(sd > 0.15)

x = bind_rows(t.disp, v.disp)

gridExtra::grid.arrange(
  ggplot(x, aes(y=mean,x=cohort,group=cohort,fill=cohort)) + geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') + labs(main='mean', x=''),
  ggplot(x, aes(y=sd,x=cohort,group=cohort,fill=cohort)) + geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') + labs(main='StDev', x=''),
  ggplot(x, aes(y=var,x=cohort,group=cohort,fill=cohort)) + geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') + labs(main='var', x=''),
  ggplot(x, aes(y=IQR,x=cohort,group=cohort,fill=cohort)) + geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') + labs(main='IQR', x=''),
ncol=2, top='Segmented data distribution disp')

# x %>% dplyr::filter(mean >= 1.05)
# x %>% dplyr::filter(sd >= 0.2) 
# 
# x %>% dplyr::filter(var > 0.05)
# x %>% dplyr::filter(IQR >= 0.2)

wilcox.test(t.disp$mean, v.disp$mean)
wilcox.test(t.disp$sd, v.disp$sd)

```

Looked at the fourier periodograms for both the training and validation data.  The RHS values are based on a frequency cutoff of 0.25. This does suggest that there's more noise in the validation data overall after segmentation.

I tried decreasing this by more agressively segmenting some samples with a higher RHS, but while the RHS drops it doesn't change the overall predictions significantly.  Given the upward shift in the raw and segmented values it may simply be that the shift is affecting it.

```{r rhs}
t.rhs = read_tsv(list.files('~/Data/BarrettsProgressionRisk/Analysis/multipcf_perPatient/', pattern='rhs', full.names = T, recursive = T), col_types = 'cdd') %>% mutate('cohort'='train')
v.rhs = read_tsv(list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/', pattern='rhs', full.names = T, recursive = T), col_types='cdd') %>% mutate('cohort'='val')

ggplot(bind_rows(t.rhs %>% dplyr::select(rhs,cohort), v.rhs %>% dplyr::select(rhs,cohort)), aes(x=cohort, y=rhs, group=cohort)) + 
  geom_jitter(width=0.2) + geom_violin(aes(fill=cohort),alpha=0.5, show.legend = F) + theme_minimal()

wilcox.test(t.rhs$rhs, v.rhs$rhs)
t.test(t.rhs$rhs, v.rhs$rhs)

```

Example
```{r loadrawval, echo=F, warning=F, message=F}
samples = v.rhs %>% arrange(-rhs) %>% dplyr::slice(1,nrow(v.rhs))

load('~/Data/BarrettsProgressionRisk/QDNAseq/validation/merged_raw_fit.Rdata', verbose=T)

fit.data = merged.fit %>% dplyr::select(location, chrom, start, end, samples$Sample)
raw.data = merged.raw %>% dplyr::select(location, chrom, start, end, samples$Sample)
rm(merged.fit, merged.raw)

```

Highest RHS
```{r highRHS}
if (file.exists('highRHS.Rdata')) {
  load('highRHS.Rdata')
} else {
  info = loadSampleInformation(tibble(Sample=samples$Sample[1], Endoscopy='2010/01/01'))

  # Default gamma is 250* the mad(std dev of the adjusted data)
  seg250 = BarrettsProgressionRisk::segmentRawData(info, raw.data[,c(1:5)], fit.data[,c(1:5)], cutoff = 0.011, gamma2=250, verbose = T)
  prr250 = predictRiskFromSegments(seg250, verbose=F)

  seg500 = BarrettsProgressionRisk::segmentRawData(info, raw.data[,c(1:5)], fit.data[,c(1:5)], cutoff = 0.011, gamma2=500, verbose = T)

  prr500 = predictRiskFromSegments(seg500, verbose=F)
  save(seg250,prr250,seg500,prr500,file='highRHS.Rdata')
}

# Gamma 250
pd250 = TSA::periodogram(seg250$seg.vals[[samples$Sample[1]]], plot=T, main='gamma 250') 
plotSegmentData(seg250) 
# number of segs
nrow(seg250$seg.vals)
# rhs
print( sum(pd250$freq[pd250$freq > 0.25]) )

# gamma 500
pd500 = TSA::periodogram(seg500$seg.vals[[samples$Sample[1]]], plot=T, main='gamma 500') 
plotSegmentData(seg500) 
# number of segs
nrow(seg500$seg.vals)
# rhs
print( sum(pd500$freq[pd500$freq > 0.25]) )

# Relative risks does drop, but not hugely
predictions(prr250)
predictions(prr500)

# These come from a NP
all.val %>% dplyr::filter(Samplename == samples$Sample[1])
```

Lowest RHS.  But again, a more aggressive segmentation isn't really altering things.  
```{r lowRHS}
if (file.exists('lowRHS.Rdata')) {
  load('lowRHS.Rdata')
} else {
  info = loadSampleInformation(tibble(Sample=samples$Sample[2], Endoscopy='2010/01/01'))

  seg250 = BarrettsProgressionRisk::segmentRawData(info, raw.data[,c(1:4,6)], fit.data[,c(1:4,6)], cutoff = 0.011, gamma2=250, verbose = F)
  prr250 = predictRiskFromSegments(seg250, verbose=F)

  seg500 = BarrettsProgressionRisk::segmentRawData(info,  raw.data[,c(1:4,6)], fit.data[,c(1:4,6)], cutoff = 0.011, gamma2=500, verbose = F)

  prr500 = predictRiskFromSegments(seg500, verbose=F)
  save(seg250,prr250,seg500,prr500,file='lowRHS.Rdata')
}

pd250 = TSA::periodogram(seg250$seg.vals[[samples$Sample[2]]], plot=T, main='gamma 250') 
plotSegmentData(seg250)
# number of segs
nrow(seg250$seg.vals)
# rhs
print( sum(pd250$freq[pd250$freq > 0.25]) )

pd500 = TSA::periodogram(seg500$seg.vals[[samples$Sample[2]]], plot=T, main='gamma 500') 
plotSegmentData(seg500)
# number of segs
nrow(seg500$seg.vals)
# rhs
print( sum(pd500$freq[pd500$freq > 0.25]) )

predictions(prr250)
predictions(prr500)

all.val %>% dplyr::filter(Samplename == samples$Sample[2])
```

# Binned Data

```{r, echo=F, message=F, warning=F}
files = list.files('~/Data/BarrettsProgressionRisk/Analysis/multipcf_perPatient/', '5e06_cleaned_tiled', recursive=T, full.names = T)
training.tiles = do.call(bind_rows, purrr::map(files, function(f) {
  read_tsv(f,col_types=c(.default=col_double()))
})) %>% rename(X1 = 'Sample')

files = list.files('~/Data/BarrettsProgressionRisk/Analysis/multipcf_perPatient/', '5e06.*MSE', recursive=T, full.names = T)
training.resids = do.call(bind_rows, purrr::map(files, function(f) {
  read_tsv(f,col_types=c(.default=col_double()))
})) %>% rename(X1 = 'Sample')

files = list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf/', '5e06_tiles', recursive=T, full.names = T)
val.tiles = do.call(bind_rows, purrr::map(files, function(f) {
  read_tsv(f,col_types=c(.default=col_double()))
})) %>% rename(X1 = 'Sample')

files = list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf/', '5e06_MSE', recursive=T, full.names = T)
val.resids = do.call(bind_rows, purrr::map(files, function(f) {
  read_tsv(f,col_types=c(.default=col_double()))
})) %>% rename(X1 = 'Sample')

```

Per patient after tiling which merges the segments using a weighted mean.  Again, there's an upwards shift per-patient. 
```{r, fig.width=10}
per.pt.tile<-function(df) {
  means = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~mean(.,na.rm=T))) %>% t %>% as_tibble %>% rename(V1='mean')
  sds = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~sd(.,na.rm=T))) %>% t %>% as_tibble %>% rename(V1='sd')
  vars = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~var(.,na.rm=T))) %>% t %>% as_tibble %>% rename(V1='var')
  iqrs = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~IQR(.,na.rm=T))) %>% t %>% as_tibble %>% rename(V1='IQR')

  return(add_column(bind_cols(means,sds,vars,iqrs), 'Sample'=df$Sample, .before=1))
}

t.pt = per.pt.tile(training.tiles) %>% mutate(cohort='train')
v.pt = per.pt.tile(val.tiles) %>% mutate(cohort='val')

gridExtra::grid.arrange(
  ggplot(bind_rows(t.pt, v.pt), aes(cohort, mean, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Mean per patient'),
  ggplot(bind_rows(t.pt, v.pt), aes(cohort, sd, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='StdDev per patient'),
  ncol=2, top='Per patient tiled values')

```

Per 5MB bin (across patients).  There's a lot more variation in the means per bin in the validation data. This is before unit normalization, which is based on the training data mean/sd, so where the variance causes issues is starting to become more clear.

```{r}
tile.dist.measure<-function(df) {
  means = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~mean(.,na.rm=T))) %>% t, rownames='bin')  %>% rename(V1='mean') 
  sds = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~mean(.,na.rm=T))) %>% t, rownames='bin')  %>% rename(V1='mean') 
  means = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~sd(.,na.rm=T))) %>% t, rownames='bin')  %>% rename(V1='sd') 
  iqrs = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~IQR(.,na.rm=T))) %>% t, rownames='bin')  %>% rename(V1='IQR') 
  vars = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~var(.,na.rm=T))) %>% t, rownames='bin')  %>% rename(V1='var')

  dist = means %>% full_join(sds, by='bin') %>% full_join(iqrs, by='bin') %>% full_join(vars, by='bin')
  return(dist)
}

t.dist = tile.dist.measure(training.tiles) %>% mutate(cohort='train')
v.dist = tile.dist.measure(val.tiles) %>% mutate(cohort='val')

gridExtra::grid.arrange(
  ggplot(bind_rows(t.dist, v.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Mean'),
  ggplot(bind_rows(t.dist, v.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='SD'), 
  ncol=2, top='Per bin tiled values')

ggplot(full_join(t.dist, v.dist, by='bin'), aes(mean.x,mean.y)) + geom_point() + labs(x='train', y='val', title='mean')

ggplot(full_join(t.dist, v.dist, by='bin'), aes(sd.x,sd.y)) + geom_point() + labs(x='train', y='val', title='SD')

#dens.obs = density(training.tiles$`1:1-4985012`)

#resample.obs = sample(dens.obs$x, nrow(val.tiles), replace=TRUE, prob=dens.obs$y)

```

After mean normalization per-bin using the z scores from the training data. And sure enough, it spreads the values out considerably for the validation data.  So the question is if the validation data were adjusted would it do this?

```{r}
#load('~/Data/BarrettsProgressionRisk/Analysis/5e6_arms/model_data.Rdata')
#rm(dysplasia.df,labels,z.arms.sd,z.arms.mean,mn.cx,sd.cx)

means = apply(as.matrix(training.tiles[,-1]),2, mean, na.rm=T)
sd = apply(as.matrix(training.tiles[,-1]),2, sd, na.rm=T)

tt.norm = as.matrix(training.tiles[-1])
for (i in 1:ncol(tt.norm))
  tt.norm[,i] = BarrettsProgressionRisk:::unit.var(tt.norm[,i], means[i], sd[i])

val.norm = as.matrix(val.tiles[-1])
for (i in 1:ncol(val.norm))
  val.norm[,i] = BarrettsProgressionRisk:::unit.var(val.norm[,i], means[i], sd[i])

tt.norm = as_tibble(tt.norm) %>% add_column('Sample'=training.tiles$Sample,.before=T)
val.norm = as_tibble(val.norm) %>% add_column('Sample'=val.tiles$Sample,.before=T)

tnorm.dist = tile.dist.measure(tt.norm) %>% mutate(cohort='train')
vnorm.dist = tile.dist.measure(val.norm) %>% mutate(cohort='val')

gridExtra::grid.arrange(
  ggplot(bind_rows(tnorm.dist, vnorm.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Mean'),
  ggplot(bind_rows(tnorm.dist, vnorm.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='SD'),
  ncol=2, top='Per bin normalized tile values')
```

This strongly suggests that I need to try re-evaluate how to normalize bins.

But what if I, as a quick test, normalize the validation cohort with itself?

```{r}

v.means = apply(as.matrix(val.tiles[,-1]),2, mean, na.rm=T)
v.sd = apply(as.matrix(val.tiles[,-1]),2, sd, na.rm=T)

val.int.norm = as.matrix(val.tiles[,-1])
for (i in 1:ncol(val.int.norm)) 
  val.int.norm[,i] = BarrettsProgressionRisk:::unit.var(val.int.norm[,i], v.means[i], v.sd[i])

val.int.norm = as_tibble(val.int.norm) %>% add_column('Sample'=val.tiles$Sample,.before=T)
vinorm.dist = tile.dist.measure(val.int.norm) %>% mutate(cohort='val int')

gridExtra::grid.arrange(
  ggplot(bind_rows(tnorm.dist, vinorm.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Mean'),
  ggplot(bind_rows(tnorm.dist, vinorm.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='SD'),
  ncol=2, top='Per bin normalized tile values', bottom='Val int = validation bins normalized using internal z-scores')
```

## Rank adjust validation values then scale?

```{r}

find.rank<-function(r,x) {
  if ( nrow(x %>% filter(rank == r)) > 0 ) return(x %>% filter(rank == r))
  find.rank(r-1,x)
  # if (nrow(x) <= 4) return(x)
  # (median(x$rank))
  # nrow(x)
  # if (r <= floor(median(x$rank)) ) {
  #   find.rank(r, x %>% filter(rank <= median(rank)))
  # } else {
  #   find.rank(r, x %>% filter(rank >= median(rank)))
  # }
}

adj.v = val.tiles
for (n in 2:ncol(training.tiles)) {
  print(n)
  col = colnames(training.tiles)[n]

  rk = sapply(seq(1, length(training.tiles[[col]]), 4), function(s) mean(training.tiles[[col]][s:(s+4)], na.rm=T))

  tt.rank = tibble(val =rk, rank=rank(rk)) %>%  arrange(rank)
  vt.rank = floor(rank(val.tiles[[col]]))

  for (i in 1:length(vt.rank)) {
    r = vt.rank[i]
    if (is.na(mean(find.rank(r,tt.rank)$val,na.rm=T))) stop(paste(col, r))
    adj.v[i, col] = mean(find.rank(r,tt.rank)$val,na.rm=T)
  }
}

j = 100
hist(adj.v[[j]], breaks=20, xlim=c(0.84,1.1))
hist(val.tiles[[j]], breaks=20, xlim=c(0.84,1.1))
hist(training.tiles[[j]], breaks=20, xlim=c(0.84,1.1))

plot( training.tiles %>% dplyr::select(-Sample) %>% dplyr::summarise_each(funs(sd)) %>% unlist,
      val.tiles %>% dplyr::select(-Sample) %>% dplyr::summarise_each(funs(sd)) %>% unlist, main='SD', xlab='train', ylab='val tiles')

plot( training.tiles %>% dplyr::select(-Sample) %>% dplyr::summarise_each(funs(sd)) %>% unlist,
      adj.v %>% dplyr::select(-Sample) %>% dplyr::summarise_each(funs(sd)) %>% unlist, main='SD', xlab='train', ylab='rank adj val')

plot( training.tiles %>% dplyr::select(-Sample) %>% dplyr::summarise_each(funs(mean)) %>% unlist,
      adj.v %>% dplyr::select(-Sample) %>% dplyr::summarise_each(funs(mean)) %>% unlist, main='mean', xlab='train', ylab='rank adj val')

# After normalization
adj.v.norm = as.matrix(adj.v[-1])
for (i in 1:ncol(adj.v.norm))
  adj.v.norm[,i] = BarrettsProgressionRisk:::unit.var(adj.v.norm[,i], means[i], sd[i])
adj.v.norm = as_tibble(adj.v.norm) %>% add_column('Sample'=val.tiles$Sample,.before=T)
adj.vnorm.dist = tile.dist.measure(adj.v.norm) %>% mutate(cohort='adj val')

gridExtra::grid.arrange(
  ggplot(bind_rows(tnorm.dist, vnorm.dist, adj.vnorm.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Mean'),
  ggplot(bind_rows(tnorm.dist, vnorm.dist,adj.vnorm.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='SD'),
  ncol=2, top='Per bin normalized tile values')

```

## Adjusted by directly altering validation values for the nearest training value?

Just a quick try, but could directly altering the values of the validation data to the nearest training value alter these?  The answer appears to be 'no'.

```{r, eval=F}

adj.val.tiles = val.norm
for (i in 1:nrow(adj.val.tiles)) {
#  print(i)
  for (j in 2:ncol(adj.val.tiles[i,])) {
    bin = tt.norm[,j,drop=T]
    bq = quantile(bin, probs=seq(0,1,0.2))

    iqr = range(quantile(bin,0.25), quantile(bin,0.75))

    if (adj.val.tiles[i,j,drop=T] >= min(iqr) & adj.val.tiles[i,j,drop=T] <= max(iqr)) {
      mn = min(iqr); mx = max(iqr)
    } else {
      mn = max(bq[which(bq <= adj.val.tiles[i,j,drop=T] )])
      mx = min(bq[which(bq >= adj.val.tiles[i,j,drop=T] )])
    }

    adj.val.tiles[i,j] = sample(bin[bin >= mn & bin <= mx],1)

    #adj.val.tiles[i,j] = bin[which.min((bin-adj.val.tiles[i,j,drop=T])^2)]
  }
}

hist(tt.norm$`1:1-4985012`, breaks=20)
hist(adj.val.tiles$`1:1-4985012`, breaks=20)
hist(val.norm$`1:1-4985012`, breaks=20)

adj.v.pt = per.pt.tile(adj.val.tiles) %>% mutate(cohort='adj.val')

ggplot(bind_rows(t.pt, v.pt, adj.v.pt), aes(cohort, mean, group=cohort, fill=cohort)) + 
  geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
  labs(title='Mean per patient')

v.adj.dist = tile.dist.measure(adj.val.tiles) %>% mutate(cohort='adj val')

ggplot(bind_rows(t.dist, v.dist,v.adj.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
  geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
  labs(title='Mean per bin')

adj.val.norm = as.matrix(adj.val.tiles[-1])
for (i in 1:ncol(adj.val.norm))
  adj.val.norm[,i] = BarrettsProgressionRisk:::unit.var(adj.val.norm[,i], z.mean[i], z.sd[i])

adj.val.norm = as_tibble(adj.val.norm) %>% add_column('Sample'=adj.val.tiles$Sample,.before=T)
vadj.norm.dist = tile.dist.measure(adj.val.norm) %>% mutate(cohort='adj val')

ggplot(bind_rows(tnorm.dist, vnorm.dist,vadj.norm.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
  geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
  labs(title='Mean per bin after norm')
```

## Normalize per patient instead?

```{r}
t.pt
v.pt

median(v.pt$var)

median(t.pt$var)
median(t.pt$sd)
median(t.pt$mean)

summary(unlist(val.tiles[1,-1]))
var(unlist(val.tiles[1,-1]))

fourier<-function(x) {
  fx = fft(x)
  plot(Re(fx), type='l', xlim=c(0,20))
  fx[20:length(fx)] = 0+0i
  fxx = fft(fx, inverse = TRUE)/length(fx)

#  layout(matrix(1:2,2,1)) 
#  plot(x, type="l", main="Original Data")
#  plot(Re(fxx),type="l", main="Fourier Transform Filtering")

#  hist(x, breaks=50)
#  hist(Re(fxx), breaks=50)
  Re(fxx) 
}

fourier.tiles = val.tiles

for (i in 1:nrow(fourier.tiles))
  fourier.tiles[i,-1] = fourier(unlist(fourier.tiles[i,-1]))

ft.disp = per.pt.tile(fourier.tiles) %>% mutate(cohort='fourier transform')

gridExtra::grid.arrange(
  ggplot(bind_rows(t.pt, v.pt, ft.disp), aes(cohort, mean, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Mean per patient'),
  ggplot(bind_rows(t.pt, v.pt, ft.disp), aes(cohort, sd, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='StdDev per patient'),
  ncol=2, top='Per patient tiled values')

f.dist = tile.dist.measure(fourier.tiles) %>% mutate(cohort='fourier transform')

gridExtra::grid.arrange(
  ggplot(bind_rows(t.dist, v.dist, f.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Mean'),
  ggplot(bind_rows(t.dist, v.dist, f.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='SD'), 
ncol=2, top='Per bin tiled values')

fourier.norm = as.matrix(fourier.tiles[-1])
for (i in 1:ncol(fourier.norm))
  fourier.norm[,i] = BarrettsProgressionRisk:::unit.var(fourier.norm[,i], z.mean[i], z.sd[i])

ft.norm = as_tibble(fourier.norm) %>% add_column('Sample'=fourier.tiles$Sample,.before=T)

ftnorm.dist = tile.dist.measure(ft.norm) %>% mutate(cohort='fourier')

gridExtra::grid.arrange(
  ggplot(bind_rows(tnorm.dist, vnorm.dist, ftnorm.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Mean'),
  ggplot(bind_rows(tnorm.dist, vnorm.dist, ftnorm.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='SD'),
  ncol=2, top='Per bin normalized tile values')

```

```{r cosigndiff, eval=F}
## What about finding the nearest patient via cosine diff?

cos.sim=function(ma, mb){
  mat=tcrossprod(ma, mb)
  t1=sqrt(apply(ma, 1, crossprod))
  t2=sqrt(apply(mb, 1, crossprod))
  mat / outer(t1,t2)
}

cs.pt = do.call(rbind, lapply(1:nrow(val.tiles), function(j) {
  sapply(1:nrow(training.tiles), function(i) {
    cos.sim(as.matrix(val.tiles[j,-1]), as.matrix(training.tiles[i,-1]))
  })
}))

```

### 2019-09-10 Raw Data Comparison.Rmd
...

@@
   html_document: 
-    fig_width: 5
+    fig_height: 4
+    fig_width: 6
 ---
@@
 library(reshape2)
+library(gridExtra)
+library(e1071)
 
@@
 ```{r, fig.width=12, fig.height=12}
-# par(mfrow=c(4,2))
-# qqnorm( na.omit(tfit.data$D701_D501_10722_11823), main='Training sample' ) 
-# qqnorm( na.omit(vfit.data$`SLX-17976.N705-S508`), main='Validation sample' )
-# 
-# qqnorm( na.omit(tfit.data$D701_D505_10722_11823), main='Training sample' ) 
-# qqnorm( na.omit(vfit.data$`SLX-17977.N706-S510`), main='Validation sample' )
-# 
-# qqnorm(na.omit(tfit.data$D706_D502_13696), main='Training sample')
-# qqnorm(na.omit(vfit.data$`SLX-16758.N703-S511`), main='Validation sample') # this was one of the OAC samples
-# 
-# qqnorm(na.omit(tfit.data$D712_D504_13692), main='Training sample') # P sample, correctly predicted
-# qqnorm(na.omit(vfit.data$`SLX-16783.N715-S505`), main='Validation sample') # P sample, correctly predicted
 
@@
 ncol=2, top='Segmented data distribution disp')
-
-# x %>% dplyr::filter(mean >= 1.05)
-# x %>% dplyr::filter(sd >= 0.2) 
-# 
-# x %>% dplyr::filter(var > 0.05)
-# x %>% dplyr::filter(IQR >= 0.2)
 
@@
   read_tsv(f,col_types=c(.default=col_double()))
-})) %>% rename(X1 = 'Sample')
+})) %>% dplyr::rename('Sample' = 'X1')
 
@@
 per.pt.tile<-function(df) {
-  means = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~mean(.,na.rm=T))) %>% t %>% as_tibble %>% rename(V1='mean')
-  sds = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~sd(.,na.rm=T))) %>% t %>% as_tibble %>% rename(V1='sd')
-  vars = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~var(.,na.rm=T))) %>% t %>% as_tibble %>% rename(V1='var')
-  iqrs = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~IQR(.,na.rm=T))) %>% t %>% as_tibble %>% rename(V1='IQR')
+  means = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~mean(.,na.rm=T))) %>% t %>% as_tibble %>% dplyr::rename('mean' = 'V1')
+  sds = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~sd(.,na.rm=T))) %>% t %>% as_tibble %>% dplyr::rename('sd' = 'V1')
+  vars = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~var(.,na.rm=T))) %>% t %>% as_tibble %>% dplyr::rename('var' = 'V1')
+  iqrs = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~IQR(.,na.rm=T))) %>% t %>% as_tibble %>% dplyr::rename('IQR' = 'V1')
 
@@
 tile.dist.measure<-function(df) {
-  means = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~mean(.,na.rm=T))) %>% t, rownames='bin')  %>% rename(V1='mean') 
-  sds = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~mean(.,na.rm=T))) %>% t, rownames='bin')  %>% rename(V1='mean') 
-  means = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~sd(.,na.rm=T))) %>% t, rownames='bin')  %>% rename(V1='sd') 
-  iqrs = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~IQR(.,na.rm=T))) %>% t, rownames='bin')  %>% rename(V1='IQR') 
-  vars = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~var(.,na.rm=T))) %>% t, rownames='bin')  %>% rename(V1='var')
+  means = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~mean(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('mean' = 'V1')
+  sds = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~sd(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('sd' = 'V1')
+  iqrs = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~IQR(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('iqrs' = 'V1')
+  vars = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~var(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('var' = 'V1')
 
-  dist = means %>% full_join(sds, by='bin') %>% full_join(iqrs, by='bin') %>% full_join(vars, by='bin')
+  skew = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~skewness(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('skew' = 'V1')
+  kurt = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~kurtosis(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('kurtosis' = 'V1')
+
+  dist = means %>% full_join(sds, by='bin') %>% full_join(iqrs, by='bin') %>% full_join(vars, by='bin') %>% full_join(skew, by='bin') %>% full_join(kurt, by='bin')
   return(dist)
@@
 
-ggplot(full_join(t.dist, v.dist, by='bin'), aes(mean.x,mean.y)) + geom_point() + labs(x='train', y='val', title='mean')
+gridExtra::grid.arrange(
+  ggplot(bind_rows(t.dist, v.dist), aes(cohort, skew, group=cohort, fill=cohort)) + 
+    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
+    labs(title='Skew'),
+  ggplot(bind_rows(t.dist, v.dist), aes(cohort, kurtosis, group=cohort, fill=cohort)) + 
+    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
+    labs(title='Kurtosis'), 
+  ncol=2, top='Per bin tiled values')
 
-ggplot(full_join(t.dist, v.dist, by='bin'), aes(sd.x,sd.y)) + geom_point() + labs(x='train', y='val', title='SD')
-
-#dens.obs = density(training.tiles$`1:1-4985012`)
-
-#resample.obs = sample(dens.obs$x, nrow(val.tiles), replace=TRUE, prob=dens.obs$y)
+grid.arrange(
+  ggplot(full_join(t.dist, v.dist, by='bin'), aes(mean.x,mean.y)) + geom_point() + labs(x='train', y='val', title='mean'),
+  ggplot(full_join(t.dist, v.dist, by='bin'), aes(sd.x,sd.y)) + geom_point() + labs(x='train', y='val', title='SD'),
+  ggplot(full_join(t.dist, v.dist, by='bin'), aes(var.x,var.y)) + geom_point() + labs(x='train', y='val', title='var'),
+  ggplot(full_join(t.dist, v.dist, by='bin'), aes(skew.x,skew.y)) + geom_point() + labs(x='train', y='val', title='Skew'),
+ncol=2)
 
@@
 
-After mean normalization per-bin using the z scores from the training data.
+After scaling and centering per-bin using the z scores from the training data.
 And sure enough, it spreads the values out considerably for the validation data.
@@
 
-But what if I, as a quick test, normalize the validation cohort with itself?
+But what if I, as a quick test, normalize the validation cohort with itself? This looks better here, but when I use this for predictions it doesn't significantly change the results.
 
 ```{r}
-
 v.means = apply(as.matrix(val.tiles[,-1]),2, mean, na.rm=T)
@@
 
+This moved the non-progressors the right direction with the predictions, but there was still a 30% FPR for non-progressors.  The age-bias reappeared for the progressors though.
+
 ```{r}
@@
   find.rank(r-1,x)
-  # if (nrow(x) <= 4) return(x)
-  # (median(x$rank))
-  # nrow(x)
-  # if (r <= floor(median(x$rank)) ) {
-  #   find.rank(r, x %>% filter(rank <= median(rank)))
-  # } else {
-  #   find.rank(r, x %>% filter(rank >= median(rank)))
-  # }
 }
@@
 for (n in 2:ncol(training.tiles)) {
-  print(n)
   col = colnames(training.tiles)[n]
@@
 }
-
-j = 100
-hist(adj.v[[j]], breaks=20, xlim=c(0.84,1.1))
-hist(val.tiles[[j]], breaks=20, xlim=c(0.84,1.1))
-hist(training.tiles[[j]], breaks=20, xlim=c(0.84,1.1))
 
@@
 
-## Adjusted by directly altering validation values for the nearest training value?
+### Predictions
 
-Just a quick try, but could directly altering the values of the validation data to the nearest training value alter these?  The answer appears to be 'no'.
+```{r}
+preds = do.call(bind_rows, purrr::map(
+  list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf/predictions-rankadj/0.9/', pattern='predictions.tsv', recursive = T, full.names = T), 
+  function(f) read_tsv(f, col_types = 'cddcccccccc'))) %>% mutate(Endoscopy = as.Date(Endoscopy)) %>% dplyr::select(-`Path Notes`)
 
-```{r, eval=F}
+preds = left_join(all.val, preds, by=c('Samplename'='Sample', 'Hospital Research ID', 'SLX-ID', 'Index Sequence', 'Block ID'))
 
-adj.val.tiles = val.norm
-for (i in 1:nrow(adj.val.tiles)) {
-#  print(i)
-  for (j in 2:ncol(adj.val.tiles[i,])) {
-    bin = tt.norm[,j,drop=T]
-    bq = quantile(bin, probs=seq(0,1,0.2))
+# Non progressors
+preds %>% dplyr::filter(Status == 'NP') %>% group_by(Risk) %>% tally() %>% spread(Risk, n)
 
-    iqr = range(quantile(bin,0.25), quantile(bin,0.75))
+# FPR is nearly 40% still
+nrow(preds %>% filter(Status == 'NP' & Risk == 'High'))/nrow(preds %>% filter(Status == 'NP'))
 
-    if (adj.val.tiles[i,j,drop=T] >= min(iqr) & adj.val.tiles[i,j,drop=T] <= max(iqr)) {
-      mn = min(iqr); mx = max(iqr)
-    } else {
-      mn = max(bq[which(bq <= adj.val.tiles[i,j,drop=T] )])
-      mx = min(bq[which(bq >= adj.val.tiles[i,j,drop=T] )])
-    }
+# Progressors
+preds %>% dplyr::filter(Status == 'P' & !grepl('OCCAMS',`Hospital Research ID`)) %>% group_by(Risk) %>% tally() %>% spread(Risk, n)
 
-    adj.val.tiles[i,j] = sample(bin[bin >= mn & bin <= mx],1)
+# FNR is still over 50% in the old progressors as well
+nrow(preds %>% filter(Status == 'P' & Risk == 'Low' & !grepl('OCCAMS',`Hospital Research ID`)))/nrow(preds %>% filter(Status == 'P' &  !grepl('OCCAMS',`Hospital Research ID`)))
+```
 
-    #adj.val.tiles[i,j] = bin[which.min((bin-adj.val.tiles[i,j,drop=T])^2)]
+## Distributions
+
+How does it change if I use a global variance, or a rolling-bin variance (to preserve some local info) to adjust?
+
+```{r}
+# global variance
+all.tiles = bind_rows(training.tiles, val.tiles)
+
+means = apply(all.tiles[,-1], 2, mean)
+sds = apply(all.tiles[,-1], 2, sd)
+global.sd = sd(as.matrix(all.tiles[,-1]))
+
+wt.sds = sapply(sds, function(x) weighted.mean(x,global.sd,w=c(0.2)))
+
+plot(wt.sds, sds)
+
+tt.norm.gv = as.matrix(training.tiles[-1])
+for (i in 1:ncol(tt.norm.gv))
+  tt.norm.gv[,i] = BarrettsProgressionRisk:::unit.var(tt.norm.gv[,i], means[i], wt.sds[i])
+
+val.norm.gv = as.matrix(val.tiles[-1])
+for (i in 1:ncol(tt.norm.gv))
+  val.norm.gv[,i] = BarrettsProgressionRisk:::unit.var(val.norm.gv[,i], means[i], wt.sds[i])
+
+all.norm.gv = as.matrix(all.tiles[-1])
+for (i in 1:ncol(all.norm.gv))
+  all.norm.gv[,i] = BarrettsProgressionRisk:::unit.var(all.norm.gv[,i], means[i], wt.sds[i])
+
+tt.norm.gv = as_tibble(tt.norm.gv) %>% add_column('Sample'=training.tiles$Sample,.before=T)
+val.norm.gv = as_tibble(val.norm.gv) %>% add_column('Sample'=val.tiles$Sample,.before=T)
+all.norm.gv = as_tibble(all.norm.gv) %>% add_column('Sample'=all.tiles$Sample,.before=T)
+
+tt.norm.dist = tile.dist.measure(tt.norm.gv) %>% mutate(cohort='train, wt var')
+vp.norm.dist = tile.dist.measure(val.norm.gv) %>% mutate(cohort='val, wt var')
+all.norm.dist = tile.dist.measure(val.norm.gv) %>% mutate(cohort='all, wt var')
+
+gridExtra::grid.arrange(
+  ggplot(bind_rows(tt.norm.dist, vp.norm.dist, all.norm.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
+    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
+    labs(title='Mean'),
+  ggplot(bind_rows(tt.norm.dist, vp.norm.dist, all.norm.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
+    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
+    labs(title='SD'),
+  ncol=2, top='Per bin wt var scale values')
+
+```
+
+```{r}
+
+tx = melt(training.tiles)
+vx = melt(val.tiles)
+
+grid.arrange(ggplot(tx, aes(value)) + xlim(0.4, 1.5) + geom_histogram(bins=50, color='lightblue') + labs(title='Training'),
+             ggplot(vx, aes(value)) + xlim(0.4,1.5) + geom_histogram(bins=50, color='lightblue') + labs(title='Validation'))
+
+# The extreme values in the validation set come from the OACs
+all.val %>% filter(Samplename %in% unique(vx[which(vx$value > 1.5),'Sample']))
+
+vx = vx %>% filter(Sample %in% (all.val %>% filter( !grepl('OCCAMS',`Hospital Research ID`)) %>% dplyr::select(Samplename) %>% pull) )
+
+tx.normal = fitdistrplus::fitdist(tx$value, 'norm')
+vx.normal = fitdistrplus::fitdist(vx$value, 'norm')
+
+```
+
+From training data
+```{r}
+plot(tx.normal)
+```
+
+From validation data
+```{r}
+plot(vx.normal)
+```
+
+```{r}
+
+train.P = round(pnorm(tx$value,mean=tx.normal$estimate['mean'], sd=tx.normal$estimate['sd']),4)
+val.P = round(pnorm(as.matrix(val.tiles[-1]),mean=tx.normal$estimate['mean'], sd=tx.normal$estimate['sd']),4)
+#val.P = round(pnorm(as.matrix(val.tiles[-1]),mean=vx.normal$estimate['mean'], sd=vx.normal$estimate['sd']),4)
+
+p.transform = as.matrix(val.tiles[,-1])
+for (i in 1:nrow(p.transform)) {
+  for (j in 1:ncol(p.transform)) {
+    p.transform[i,j] = mean(tx[which(train.P == val.P[i,j]),'value'])
   }
@@
 
-hist(tt.norm$`1:1-4985012`, breaks=20)
-hist(adj.val.tiles$`1:1-4985012`, breaks=20)
-hist(val.norm$`1:1-4985012`, breaks=20)
+n = 20
+hist(as.matrix(training.tiles[,-1])[,n] ,breaks=20)
+hist(as.matrix(val.tiles[,-1])[,n] ,breaks=20)
+hist(p.transform[,n], breaks=20)
 
-adj.v.pt = per.pt.tile(adj.val.tiles) %>% mutate(cohort='adj.val')
+# so now if I scale...
 
-ggplot(bind_rows(t.pt, v.pt, adj.v.pt), aes(cohort, mean, group=cohort, fill=cohort)) + 
-  geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
-  labs(title='Mean per patient')
+vp.norm = p.transform
+for (i in 1:ncol(vp.norm))
+  vp.norm[,i] = BarrettsProgressionRisk:::unit.var(vp.norm[,i], means[i], sd[i])
+vp.norm = as_tibble(vp.norm) %>% add_column('Sample'=val.tiles$Sample,.before=T)
 
-v.adj.dist = tile.dist.measure(adj.val.tiles) %>% mutate(cohort='adj val')
-
-ggplot(bind_rows(t.dist, v.dist,v.adj.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
-  geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
-  labs(title='Mean per bin')
-
-adj.val.norm = as.matrix(adj.val.tiles[-1])
-for (i in 1:ncol(adj.val.norm))
-  adj.val.norm[,i] = BarrettsProgressionRisk:::unit.var(adj.val.norm[,i], z.mean[i], z.sd[i])
-
-adj.val.norm = as_tibble(adj.val.norm) %>% add_column('Sample'=adj.val.tiles$Sample,.before=T)
-vadj.norm.dist = tile.dist.measure(adj.val.norm) %>% mutate(cohort='adj val')
-
-ggplot(bind_rows(tnorm.dist, vnorm.dist,vadj.norm.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
-  geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
-  labs(title='Mean per bin after norm')
-```
-
-## Normalize per patient instead?
-
-```{r}
-t.pt
-v.pt
-
-median(v.pt$var)
-
-median(t.pt$var)
-median(t.pt$sd)
-median(t.pt$mean)
-
-summary(unlist(val.tiles[1,-1]))
-var(unlist(val.tiles[1,-1]))
-
-fourier<-function(x) {
-  fx = fft(x)
-  plot(Re(fx), type='l', xlim=c(0,20))
-  fx[20:length(fx)] = 0+0i
-  fxx = fft(fx, inverse = TRUE)/length(fx)
-
-#  layout(matrix(1:2,2,1)) 
-#  plot(x, type="l", main="Original Data")
-#  plot(Re(fxx),type="l", main="Fourier Transform Filtering")
-
-#  hist(x, breaks=50)
-#  hist(Re(fxx), breaks=50)
-  Re(fxx) 
-}
-
-fourier.tiles = val.tiles
-
-for (i in 1:nrow(fourier.tiles))
-  fourier.tiles[i,-1] = fourier(unlist(fourier.tiles[i,-1]))
-
-ft.disp = per.pt.tile(fourier.tiles) %>% mutate(cohort='fourier transform')
+vp.norm.dist = tile.dist.measure(vp.norm) %>% mutate(cohort='val p')
 
 gridExtra::grid.arrange(
-  ggplot(bind_rows(t.pt, v.pt, ft.disp), aes(cohort, mean, group=cohort, fill=cohort)) + 
-    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
-    labs(title='Mean per patient'),
-  ggplot(bind_rows(t.pt, v.pt, ft.disp), aes(cohort, sd, group=cohort, fill=cohort)) + 
-    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
-    labs(title='StdDev per patient'),
-  ncol=2, top='Per patient tiled values')
-
-f.dist = tile.dist.measure(fourier.tiles) %>% mutate(cohort='fourier transform')
-
-gridExtra::grid.arrange(
-  ggplot(bind_rows(t.dist, v.dist, f.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
+  ggplot(bind_rows(tnorm.dist, vnorm.dist, vp.norm.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
     geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
     labs(title='Mean'),
-  ggplot(bind_rows(t.dist, v.dist, f.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
-    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
-    labs(title='SD'), 
-ncol=2, top='Per bin tiled values')
-
-fourier.norm = as.matrix(fourier.tiles[-1])
-for (i in 1:ncol(fourier.norm))
-  fourier.norm[,i] = BarrettsProgressionRisk:::unit.var(fourier.norm[,i], z.mean[i], z.sd[i])
-
-ft.norm = as_tibble(fourier.norm) %>% add_column('Sample'=fourier.tiles$Sample,.before=T)
-
-ftnorm.dist = tile.dist.measure(ft.norm) %>% mutate(cohort='fourier')
-
-gridExtra::grid.arrange(
-  ggplot(bind_rows(tnorm.dist, vnorm.dist, ftnorm.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
-    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
-    labs(title='Mean'),
-  ggplot(bind_rows(tnorm.dist, vnorm.dist, ftnorm.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
+  ggplot(bind_rows(tnorm.dist, vnorm.dist,vp.norm.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
     geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
@@
 ```
-
-```{r cosigndiff, eval=F}
-## What about finding the nearest patient via cosine diff?
-
-cos.sim=function(ma, mb){
-  mat=tcrossprod(ma, mb)
-  t1=sqrt(apply(ma, 1, crossprod))
-  t2=sqrt(apply(mb, 1, crossprod))
-  mat / outer(t1,t2)
-}
-
-cs.pt = do.call(rbind, lapply(1:nrow(val.tiles), function(j) {
-  sapply(1:nrow(training.tiles), function(i) {
-    cos.sim(as.matrix(val.tiles[j,-1]), as.matrix(training.tiles[i,-1]))
-  })
-}))
-
-```

### 2019-09-10 lib/cv-pt-glm.R

@@ -372,12 +372,12 @@ plot.cv<-function(cv, title='') {
 
 
 more.l<-function(lambda) {
-  if (min(lambda) > -5)
+  if (min(log(lambda)) > -10)
     lambda = c(lambda, seq( min(lambda), exp(-5), 1e-3))
                
-  l = sort(c(lambda, seq(exp(-5), exp(-10), -1e-6),
-             seq(exp(-10), exp(-15), -1e-8),
-             seq(exp(-15), exp(-20), -1e-9)), decreasing=T)
+  l = unique(sort(c(lambda, seq(exp(-5), exp(-10), -1e-6),
+             seq(exp(-10), exp(-15), -1e-8)#,  seq(exp(-15), exp(-20), -1e-9)
+             ), decreasing=T))
   return(l)
 }
 

### 2019-09-10 Raw Data Comparison.Rmd
...

@@
 
-But what if I, as a quick test, normalize the validation cohort with itself? This looks better here, but when I use this for predictions it doesn't significantly change the results.
+## Normalize Validation to itself
+
+Instead scaling using z-values from the trianing data, I can do it from within the validation cohort.
+This looks better here, but when I use this for predictions it doesn't significantly change the results.
 
@@
   ncol=2, top='Per bin normalized tile values', bottom='Val int = validation bins normalized using internal z-scores')
+```
+
+### Predictions
+```{r}
+preds = do.call(bind_rows, purrr::map(
+  list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf/predictions_scale_within_cohort/0.9/', pattern='predictions.tsv', recursive = T, full.names = T), 
+  function(f) read_tsv(f, col_types = 'cddcccccccc'))) %>% mutate(Endoscopy = as.Date(Endoscopy)) %>% dplyr::select(-`Path Notes`)
+
+preds = left_join(all.val, preds, by=c('Samplename'='Sample', 'Hospital Research ID', 'SLX-ID', 'Index Sequence', 'Block ID'))
+
+# Non progressors
+preds %>% dplyr::filter(Status == 'NP') %>% group_by(Risk) %>% tally() %>% spread(Risk, n)
+
+# FPR is nearly 40% still
+nrow(preds %>% filter(Status == 'NP' & Risk == 'High'))/nrow(preds %>% filter(Status == 'NP'))
+
+# Progressors
+preds %>% dplyr::filter(Status == 'P' & !grepl('OCCAMS',`Hospital Research ID`)) %>% group_by(Risk) %>% tally() %>% spread(Risk, n)
+
+# FNR is still over 50% in the old progressors as well
+nrow(preds %>% filter(Status == 'P' & Risk == 'Low' & !grepl('OCCAMS',`Hospital Research ID`)))/nrow(preds %>% filter(Status == 'P' &  !grepl('OCCAMS',`Hospital Research ID`)))
 ```
@@
 
-wt.sds = sapply(sds, function(x) weighted.mean(x,global.sd,w=c(0.2)))
+wt.sds = sapply(sds, function(x) weighted.mean(c(x,global.sd),w=c(0.5,0.5)))
 

### 2019-09-10 Raw Data Comparison.Rmd
...

--- 2019-09-10 Raw Data Comparison.Rmd
@@
     fig_width: 6
+    toc: yes
 ---
@@
 sheets = readxl::excel_sheets(val.file)[8:13]
+
+pastefun<-function(x) {
+  if ( !grepl('SLX-', x) ) x = paste0('SLX-',x)
+  return(x)
+}
+
 all.val = do.call(bind_rows, lapply(sheets, function(s) {
-  readxl::read_xlsx(val.file, s) %>% dplyr::select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA, matches('Collection')) %>% dplyr::filter(!is.na(`SLX-ID`)) %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character))# %>% fncols('Collection', 'Biopsy')
-})) %>% mutate(Samplename = paste(`SLX-ID`,`Index Sequence`, sep='.'))
+  readxl::read_xlsx(val.file, s) %>% dplyr::select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA, matches('Collection')) %>% 
+    dplyr::filter(!is.na(`SLX-ID`)) %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) 
+})) %>% mutate(Samplename = paste(`SLX-ID`,`Index Sequence`, sep='.')) %>% 
+  rowwise %>% mutate_at(vars(`SLX-ID`), list(pastefun) ) %>% ungroup %>% 
+  mutate(
+  `Hospital Research ID` = str_replace_all( str_remove_all(`Hospital Research ID`, " "), '/', '_'), 
+  `Index Sequence` = str_replace_all(`Index Sequence`, 'tp', ''),
+  Samplename = paste(`SLX-ID`,`Index Sequence`,sep='.')
+  )
 
@@
 
-```{r, fig.width=12, fig.height=12}
+```{r}
 
@@
 
-But what if I, as a quick test, normalize the validation cohort with itself? This looks better here, but when I use this for predictions it doesn't significantly change the results.
+## Normalize validation cohort within itself
+
+Instead scaling using z-values from the trianing data, I can do it from within the validation cohort.
+This looks better here, but when I use this for predictions it doesn't significantly change the results.
 
@@
   ncol=2, top='Per bin normalized tile values', bottom='Val int = validation bins normalized using internal z-scores')
+```
+
+### Predictions
+
+```{r}
+preds = do.call(bind_rows, purrr::map(
+  list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf/predictions_scale_within_cohort/0.9/', pattern='predictions.tsv', recursive = T, full.names = T), 
+  function(f) read_tsv(f, col_types = 'cddcccccccc'))) %>% mutate(Endoscopy = as.Date(Endoscopy)) %>% dplyr::select(-`Path Notes`) %>% mutate(`Hospital Research ID` = str_replace_all( str_remove_all(`Hospital Research ID`, " "), '/', '_'))
+
+preds = left_join(preds,all.val, by=c('Sample'='Samplename', 'Hospital Research ID', 'SLX-ID', 'Index Sequence', 'Block ID'))
+
+# Non progressors
+preds %>% dplyr::filter(Status == 'NP') %>% group_by(Risk) %>% tally() %>% spread(Risk, n)
+
+# FPR is ~40% still
+nrow(preds %>% filter(Status == 'NP' & Risk == 'High'))/nrow(preds %>% filter(Status == 'NP'))
+
+# Progressors
+preds %>% dplyr::filter(Status == 'P' & !grepl('OCCAMS',`Hospital Research ID`)) %>% group_by(Risk) %>% tally() %>% spread(Risk, n)
+
+# FNR is still over 50% in the old progressors as well
+nrow(preds %>% filter(Status == 'P' & Risk == 'Low' & !grepl('OCCAMS',`Hospital Research ID`)))/nrow(preds %>% filter(Status == 'P' &  !grepl('OCCAMS',`Hospital Research ID`)))
 ```
@@
 
-preds = left_join(all.val, preds, by=c('Samplename'='Sample', 'Hospital Research ID', 'SLX-ID', 'Index Sequence', 'Block ID'))
+preds = left_join(preds,all.val, by=c('Sample'='Samplename', 'Hospital Research ID', 'SLX-ID', 'Index Sequence', 'Block ID'))
 
@@
 
-How does it change if I use a global variance, or a rolling-bin variance (to preserve some local info) to adjust?
+How does it change if I use a global variance learned from both cohorts, or a rolling-bin variance (to preserve some local info) to adjust?
+
+I think a weight-adjusted global~local variance isn't going to help.  The variance is consistently and stubbornly greater in the validation cohort no matter which direction I weight the scale factor.
 
 ```{r}
+wt = c(0.5,0.5)
+
 # global variance
 all.tiles = bind_rows(training.tiles, val.tiles)
-
-means = apply(all.tiles[,-1], 2, mean)
-sds = apply(all.tiles[,-1], 2, sd)
 global.sd = sd(as.matrix(all.tiles[,-1]))
 
-wt.sds = sapply(sds, function(x) weighted.mean(x,global.sd,w=c(0.2)))
-
-plot(wt.sds, sds)
+means = apply(training.tiles[,-1], 2, mean)
+sds = apply(training.tiles[,-1], 2, sd)
+wt.sds = sapply(sds, function(x) weighted.mean(c(x,global.sd),w=wt))
 
@@
 
+means = apply(val.tiles[,-1], 2, mean)
+sds = apply(val.tiles[,-1], 2, sd)
+wt.sds = sapply(sds, function(x) weighted.mean(c(x,global.sd),w=wt))
+
 val.norm.gv = as.matrix(val.tiles[-1])
@@
   val.norm.gv[,i] = BarrettsProgressionRisk:::unit.var(val.norm.gv[,i], means[i], wt.sds[i])
+
+means = apply(all.tiles[,-1], 2, mean)
+sds = apply(all.tiles[,-1], 2, sd)
+wt.sds = sapply(sds, function(x) weighted.mean(c(x,global.sd),w=wt))
 
@@
 
-```{r}
+```{r eval=F}
 
@@
 
-```
+#From training data
+plot(tx.normal)
 
-From training data
-```{r}
-plot(tx.normal)
-```
-
-From validation data
-```{r}
+#From validation data
 plot(vx.normal)

### 2019-09-10 Validation Report.Rmd
...

@@
 analysis.dir = '~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf'
-preds.dir = paste0(analysis.dir, '/predictions/0.9')
+preds.dir = paste0(analysis.dir, '/predictions_model_5e6/0.9')
 

### 2019-09-10 processing_pipeline/predict_validation.R

@@ -157,7 +157,7 @@ for (f in segFiles) {
   index = sub('_.*', '', basename(f))
   segmented$sample.info = BarrettsProgressionRisk::loadSampleInformation(info %>% filter(Sample %in% segmented$sample.info$Sample) )
   
-  tiles = BarrettsProgressionRisk:::tileSamples(segmented, verbose=F)
+  tiles = BarrettsProgressionRisk:::tileSamples(segmented, scale=F, verbose=F)
   #tiles$tiles = val.df
   
   prr = BarrettsProgressionRisk:::predictRisk(segmented, tiles, be.model)

### 2019-09-11 Raw Data Comparison.Rmd
...

@@
   disp = full_join(
-           full_join( as_tibble(t(means), rownames='sample') %>% rename(V1='mean'), 
-                      as_tibble(t(sds), rownames='sample') %>% rename(V1='sd'), by='sample'),
-           full_join(as_tibble(t(vars), rownames='sample') %>% rename(V1='var'),
-                     as_tibble(t(iqrs), rownames='sample') %>% rename(V1='IQR'), by='sample'), 
+           full_join( as_tibble(t(means), rownames='sample') %>% dplyr::rename('mean' = 'V1'), 
+                      as_tibble(t(sds), rownames='sample') %>% dplyr::rename('sd' = 'V1'),  by='sample'),
+           full_join(as_tibble(t(vars), rownames='sample') %>% dplyr::rename('var' = 'V1'), 
+                     as_tibble(t(iqrs), rownames='sample') %>% dplyr::rename('IQR' = 'V1'),  by='sample'), 
            by='sample') 
@@
 
+## Retrained model, no scaling
+
+Shows nearly the same problem as before.  The variance is clearly at issue.
+
+```{r}
+
+preds = do.call(bind_rows, purrr::map(
+  list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf/predictions_no_scale/0.9/', pattern='predictions.tsv', recursive = T, full.names = T), 
+  function(f) read_tsv(f, col_types = 'cddcccccccc'))) %>% mutate(Endoscopy = as.Date(Endoscopy)) %>% dplyr::select(-`Path Notes`)
+
+preds = left_join(preds,all.val, by=c('Sample'='Samplename', 'Hospital Research ID', 'SLX-ID', 'Index Sequence', 'Block ID'))
+
+# Non progressors
+preds %>% dplyr::filter(Status == 'NP') %>% group_by(Risk) %>% tally() %>% spread(Risk, n)
+
+# FPR is nearly 40% still
+nrow(preds %>% filter(Status == 'NP' & Risk == 'High'))/nrow(preds %>% filter(Status == 'NP'))
+
+# Progressors
+preds %>% dplyr::filter(Status == 'P' & !grepl('OCCAMS',`Hospital Research ID`)) %>% group_by(Risk) %>% tally() %>% spread(Risk, n)
+
+# FNR is still over 50% in the old progressors as well
+nrow(preds %>% filter(Status == 'P' & Risk == 'Low' & !grepl('OCCAMS',`Hospital Research ID`)))/nrow(preds %>% filter(Status == 'P' &  !grepl('OCCAMS',`Hospital Research ID`)))
+```
+```
+
 ```{r eval=F}

### 2019-09-11 lib/data_func.R

@@ -1,6 +1,6 @@
 
 ## Should NAs be replace by per-sample mean value instead??
-prep.matrix<-function(dt, na.replace=0, scale=F) {
+prep.matrix<-function(dt, na.replace=0, scale=F, MARGIN=2) {
   output = list()
   rows = rownames(dt)
   
@@ -11,9 +11,12 @@ prep.matrix<-function(dt, na.replace=0, scale=F) {
   output[['z.sd']] = sd
   
   if (scale) {
-    dt = apply(dt, 2,  BarrettsProgressionRisk:::unit.var)
-    dt = apply(dt, 2,  scale, T, F)
-    rownames(dt) = rows
+    dt = apply(dt, MARGIN,  BarrettsProgressionRisk:::unit.var)
+    if (MARGIN == 2) { 
+      rownames(dt) = rows
+    } else {
+      dt = t(dt)  
+    }
   }
 
   if (!is.null(na.replace))

### 2019-09-11 processing_pipeline/run-cv2.R

@@ -19,7 +19,7 @@ suppressPackageStartupMessages(source('~/workspace/shallowwgs_pipeline/lib/cv-pt
 data = args[1]
 # data = '~/Data/BarrettsProgressionRisk/Analysis/multipcf_perPatient/'
 outdir = args[2]
-# outdir = '~/Data/BarrettsProgressionRisk/Analysis/model_no_bin_scaling/'
+# outdir = '/tmp'
 infodir = args[3]
 # infodir = '~/Data/BarrettsProgressionRisk/QDNAseq'
 set = 'Training'
@@ -99,7 +99,7 @@ per.samp.mean = apply(seg.tiles,1,mean)
 #if (logT) seg.tiles = t(apply(seg.tiles, 1, BarrettsProgressionRisk:::.logTransform))
 
 # After mean centering set all NA values to 0
-segsList = prep.matrix(seg.tiles,scale=F)
+segsList = prep.matrix(seg.tiles,scale=T, MARGIN=1)
 z.mean = segsList$z.mean
 z.sd = segsList$z.sd
 segs = segsList$matrix
@@ -123,7 +123,7 @@ arm.tiles = as.matrix(arm.tiles[,-1])
 rownames(arm.tiles) = samples
 #if (logT) arm.tiles = t(apply(arm.tiles, 1, logTransform))
 
-armsList = prep.matrix(arm.tiles,scale=F)
+armsList = prep.matrix(arm.tiles,scale=T,MARGIN=1)
 arms = armsList$matrix
 z.arms.mean = armsList$z.mean
 z.arms.sd = armsList$z.sd

### 2019-09-11 Raw Data Comparison.Rmd
...

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = F, warning = F)
+
 library(tidyverse)
@@
 library(gridExtra)
+library(kableExtra)
+library(pander)
 library(e1071)
@@
 
-dim(vfit.data)
-dim(tfit.data)
+##dim(vfit.data)
+##dim(tfit.data)
 ```
@@
 
-wilcox.test(t.fit.disp$mean, v.fit.disp$mean)
+pander(wilcox.test(t.fit.disp$mean, v.fit.disp$mean), caption='Per pt mean comparison (fitted values)')
+
+# Within the cohorts they're comparable
+x = t.fit.disp$mean[sample(500,replace = F)]
+pander(wilcox.test(x[1:250], x[251:500]),caption='Within training (sampled)')
+
+y = v.fit.disp$mean[sample(100,replace = F)]
+pander(wilcox.test(y[1:50], y[51:50]), caption='Within validation (sampled')
+
+pander(wilcox.test(t.fit.disp$var, v.fit.disp$var), catpion='Per pt var comparison (fitted values)')
 
@@
 
-wilcox.test(t.raw.disp$mean, v.raw.disp$mean)
+pander(wilcox.test(t.raw.disp))
 
@@
 
+The samples from the training cohort that have a mean>300 (outliers) are progressors, but did not all predict high risk in LOO.
+
 # First processing steps
+
+In the initial processing the raw counts are adjusted using the fitted values.
+
 ```{r}
@@
 all.val %>% dplyr::filter(Samplename == samples$Sample[2])
+
+rm(prr500, prr250, pd500, pd250, seg500, seg250)
 ```
@@
 # Binned Data
+
+5MB bins.  For most bins this involves a weight average of 2 segments that overlap the bin.
 
@@
 
+## Per patient
+
 Per patient after tiling which merges the segments using a weighted mean.  Again, there's an upwards shift per-patient. 
 ```{r, fig.width=10}
+
 per.pt.tile<-function(df) {
@@
 
+t.per.pt.scale = as_tibble(t(apply(as.matrix(training.tiles[,-1]), 1, scale, center=T, scale=T))) %>% add_column('sample'=training.tiles$Sample, .before=1)
+colnames(t.per.pt.scale) = colnames(training.tiles)
+
+v.per.pt.scale = as_tibble(t(apply(as.matrix(val.tiles[,-1]), 1, scale, center=T, scale=T))) %>% add_column('sample'=val.tiles$Sample, .before=1)
+colnames(v.per.pt.scale) = colnames(val.tiles)
+
+t.pt = per.pt.tile(t.per.pt.scale) %>% mutate(cohort='train')
+v.pt = per.pt.tile(v.per.pt.scale) %>% mutate(cohort='val')
+
+gridExtra::grid.arrange(
+  ggplot(bind_rows(t.pt, v.pt), aes(cohort, mean, group=cohort, fill=cohort)) + ylim(-.5,.5) +
+    geom_jitter(width=0.1) + geom_boxplot(alpha=0.5) + theme(legend.position = 'none') +
+    labs(title='Mean per patient'),
+  ggplot(bind_rows(t.pt, v.pt), aes(x=cohort, y=sd, group=cohort, fill=cohort)) + 
+    geom_point() + geom_boxplot(alpha=0.5) + theme(legend.position = 'none') +
+    labs(title='StdDev per patient'),
+  ncol=2, top='Per patient tiled values, after scaling per patient')
+
 ```
+
+## Per bin
 
@@
 This strongly suggests that I need to try re-evaluate how to normalize bins.
+
+### Per bin, after scaling per-pt
+
+```{r}
+
+t.dist = tile.dist.measure(t.per.pt.scale) %>% mutate(cohort='train')
+v.dist = tile.dist.measure(v.per.pt.scale) %>% mutate(cohort='val')
+
+gridExtra::grid.arrange(
+  ggplot(bind_rows(t.dist, v.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
+    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
+    labs(title='Mean'),
+  ggplot(bind_rows(t.dist, v.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
+    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
+    labs(title='SD'), 
+  ncol=2, top='Per bin tiled values on per-pt scaled')
+
+```
 

### 2019-09-11 Raw Data Comparison.Rmd
...

@@
 
+files = list.files('~/Data/BarrettsProgressionRisk/Analysis/multipcf_perPatient/', 'arms_cleaned', recursive=T, full.names = T)
+training.arms = do.call(bind_rows, purrr::map(files, function(f) {
+  read_tsv(f,col_types=c(.default=col_double()))
+})) %>% dplyr::rename('Sample' = 'X1')
+
 files = list.files('~/Data/BarrettsProgressionRisk/Analysis/multipcf_perPatient/', '5e06.*MSE', recursive=T, full.names = T)
@@
 val.tiles = do.call(bind_rows, purrr::map(files, function(f) {
+  read_tsv(f,col_types=c(.default=col_double()))
+})) %>% dplyr::rename('Sample' = 'X1')
+
+files = list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf/', 'arm_tiles', recursive=T, full.names = T)
+val.arms = do.call(bind_rows, purrr::map(files, function(f) {
   read_tsv(f,col_types=c(.default=col_double()))
@@
 colnames(t.per.pt.scale) = colnames(training.tiles)
+t.arm.pp.scale = as_tibble(t(apply(as.matrix(training.arms[,-1]), 1, scale, center=T, scale=T))) %>% add_column('sample'=training.arms$Sample, .before=1)
+colnames(t.arm.pp.scale) = colnames(training.arms)
 
@@
 colnames(v.per.pt.scale) = colnames(val.tiles)
+v.arm.pp.scale = as_tibble(t(apply(as.matrix(val.arms[,-1]), 1, scale, center=T, scale=T))) %>% add_column('sample'=val.arms$Sample, .before=1)
+colnames(v.arm.pp.scale) = colnames(val.arms)
 
@@
 ```{r}
-
 t.dist = tile.dist.measure(t.per.pt.scale) %>% mutate(cohort='train')
@@
   ncol=2, top='Per bin tiled values on per-pt scaled')
+```
+
+After the arm segments have been subtracted as well.
+```{r}
+
+t.pp.scale = BarrettsProgressionRisk:::subtractArms(as.matrix(t.per.pt.scale[,-1]), as.matrix(t.arm.pp.scale[,-1])) %>% as_tibble %>% add_column('Sample' = 'x', .before=1)
+v.pp.scale = BarrettsProgressionRisk:::subtractArms(as.matrix(v.per.pt.scale[,-1]), as.matrix(v.arm.pp.scale[,-1]))  %>% as_tibble %>% add_column('Sample' = 'x', .before=1)
+
+t.dist = tile.dist.measure(t.pp.scale) %>% mutate(cohort='train')
+v.dist = tile.dist.measure(v.pp.scale) %>% mutate(cohort='val')
+
+gridExtra::grid.arrange(
+  ggplot(bind_rows(t.dist, v.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
+    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
+    labs(title='Mean'),
+  ggplot(bind_rows(t.dist, v.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
+    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
+    labs(title='SD'), 
+  ncol=2, top='Per bin tiled values on per-pt scaled')
+
+```
+
+### Predictions
+
+```{r}
+preds = do.call(bind_rows, purrr::map(
+  list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf/predictions_per_pt_scale/0.9/', pattern='predictions.tsv', recursive = T, full.names = T), 
+  function(f) read_tsv(f, col_types = 'cddcccccccc'))) %>% mutate(Endoscopy = as.Date(Endoscopy)) %>% dplyr::select(-`Path Notes`) %>% mutate(`Hospital Research ID` = str_replace_all( str_remove_all(`Hospital Research ID`, " "), '/', '_'))
+
+preds = left_join(preds,all.val, by=c('Sample'='Samplename', 'Hospital Research ID', 'SLX-ID', 'Index Sequence', 'Block ID'))
+
+# Non progressors
+# FPR is ~40% still
+fpr = round(nrow(preds %>% filter(Status == 'NP' & Risk == 'High'))/nrow(preds %>% filter(Status == 'NP')),2)
+
+preds %>% dplyr::filter(Status == 'NP') %>% group_by(Risk) %>% tally() %>% spread(Risk, n) %>% 
+  kable(caption=paste0('NP risk predictions, FPR=',fpr)) %>% kable_styling(full_width = F)
+
+# Progressors
+# FNR is still over 50% in the old progressors as well
+fnr = round(nrow(preds %>% filter(Status == 'P' & Risk == 'Low' & !grepl('OCCAMS',`Hospital Research ID`)))/nrow(preds %>% filter(Status == 'P' &  !grepl('OCCAMS',`Hospital Research ID`))),2)
+
+preds %>% dplyr::filter(Status == 'P' & !grepl('OCCAMS',`Hospital Research ID`)) %>% group_by(Risk) %>% tally() %>% spread(Risk, n) %>% 
+  kable(caption=paste0('P risk predictions, FNR=',fnr)) %>% kable_styling(full_width = F)
 
@@
 # Non progressors
-preds %>% dplyr::filter(Status == 'NP') %>% group_by(Risk) %>% tally() %>% spread(Risk, n)
+# FPR is ~40% still
+fpr = round(nrow(preds %>% filter(Status == 'NP' & Risk == 'High'))/nrow(preds %>% filter(Status == 'NP')),2)
 
-# FPR is ~40% still
-nrow(preds %>% filter(Status == 'NP' & Risk == 'High'))/nrow(preds %>% filter(Status == 'NP'))
+preds %>% dplyr::filter(Status == 'NP') %>% group_by(Risk) %>% tally() %>% spread(Risk, n) %>% 
+  kable(caption=paste0('NP risk predictions, FPR=',fpr)) %>% kable_styling(full_width = F)
 
 # Progressors
-preds %>% dplyr::filter(Status == 'P' & !grepl('OCCAMS',`Hospital Research ID`)) %>% group_by(Risk) %>% tally() %>% spread(Risk, n)
+# FNR is still over 50% in the old progressors as well
+fnr = round(nrow(preds %>% filter(Status == 'P' & Risk == 'Low' & !grepl('OCCAMS',`Hospital Research ID`)))/nrow(preds %>% filter(Status == 'P' &  !grepl('OCCAMS',`Hospital Research ID`))),2)
 
-# FNR is still over 50% in the old progressors as well
-nrow(preds %>% filter(Status == 'P' & Risk == 'Low' & !grepl('OCCAMS',`Hospital Research ID`)))/nrow(preds %>% filter(Status == 'P' &  !grepl('OCCAMS',`Hospital Research ID`)))
+preds %>% dplyr::filter(Status == 'P' & !grepl('OCCAMS',`Hospital Research ID`)) %>% group_by(Risk) %>% tally() %>% spread(Risk, n) %>% 
+  kable(caption=paste0('P risk predictions, FNR=',fnr)) %>% kable_styling(full_width = F) 
 ```
@@
 # Non progressors
-preds %>% dplyr::filter(Status == 'NP') %>% group_by(Risk) %>% tally() %>% spread(Risk, n)
+# FPR is ~40% still
+fpr = round(nrow(preds %>% filter(Status == 'NP' & Risk == 'High'))/nrow(preds %>% filter(Status == 'NP')),2)
 
-# FPR is nearly 40% still
-nrow(preds %>% filter(Status == 'NP' & Risk == 'High'))/nrow(preds %>% filter(Status == 'NP'))
+preds %>% dplyr::filter(Status == 'NP') %>% group_by(Risk) %>% tally() %>% spread(Risk, n) %>% 
+  kable(caption=paste0('NP risk predictions, FPR=',fpr)) %>% kable_styling(full_width = F)
 
 # Progressors
-preds %>% dplyr::filter(Status == 'P' & !grepl('OCCAMS',`Hospital Research ID`)) %>% group_by(Risk) %>% tally() %>% spread(Risk, n)
+# FNR is still over 50% in the old progressors as well
+fnr = round(nrow(preds %>% filter(Status == 'P' & Risk == 'Low' & !grepl('OCCAMS',`Hospital Research ID`)))/nrow(preds %>% filter(Status == 'P' &  !grepl('OCCAMS',`Hospital Research ID`))),2)
 
-# FNR is still over 50% in the old progressors as well
-nrow(preds %>% filter(Status == 'P' & Risk == 'Low' & !grepl('OCCAMS',`Hospital Research ID`)))/nrow(preds %>% filter(Status == 'P' &  !grepl('OCCAMS',`Hospital Research ID`)))
+preds %>% dplyr::filter(Status == 'P' & !grepl('OCCAMS',`Hospital Research ID`)) %>% group_by(Risk) %>% tally() %>% spread(Risk, n) %>% 
+  kable(caption=paste0('P risk predictions, FNR=',fnr)) %>% kable_styling(full_width = F) 
 ```

### 2019-09-13 Raw Data Comparison.Rmd
...

@@
 ```{r, echo=F, message=F, warning=F}
+tile.dist.measure<-function(df) {
+  means = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~mean(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('mean' = 'V1')
+  sds = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~sd(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('sd' = 'V1')
+  iqrs = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~IQR(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('iqrs' = 'V1')
+  vars = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~var(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('var' = 'V1')
+
+  skew = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~skewness(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('skew' = 'V1')
+  kurt = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~kurtosis(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('kurtosis' = 'V1')
+
+  dist = means %>% full_join(sds, by='bin') %>% full_join(iqrs, by='bin') %>% full_join(vars, by='bin') %>% full_join(skew, by='bin') %>% full_join(kurt, by='bin')
+  return(dist)
+}
+
 files = list.files('~/Data/BarrettsProgressionRisk/Analysis/multipcf_perPatient/', '5e06_cleaned_tiled', recursive=T, full.names = T)
@@
 
-## Per patient
+## Looking at the distribution of values per sample
 
-Per patient after tiling which merges the segments using a weighted mean.  Again, there's an upwards shift per-patient. 
+Per sample after tiling which merges the segments using a weighted mean.  Again, there's an upwards shift per-sample 
 ```{r, fig.width=10}
@@
   ncol=2, top='Per patient tiled values')
+```
 
+## Scale per sample
+
+Scaling and centering per sample makes the samples directly comparable, but not sure what it will do to the bins.  
+```{r}
 t.per.pt.scale = as_tibble(t(apply(as.matrix(training.tiles[,-1]), 1, scale, center=T, scale=T))) %>% add_column('sample'=training.tiles$Sample, .before=1)
@@
 
-## Per bin
+### Per bin, after scaling per-sample
+
+Now the variance of the cohorts appears to be nearly the same.
+
+```{r}
+t.dist = tile.dist.measure(t.per.pt.scale) %>% mutate(cohort='train')
+v.dist = tile.dist.measure(v.per.pt.scale) %>% mutate(cohort='val')
+
+gridExtra::grid.arrange(
+  ggplot(bind_rows(t.dist, v.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
+    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
+    labs(title='Mean'),
+  ggplot(bind_rows(t.dist, v.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
+    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
+    labs(title='SD'), 
+  ncol=2, top='Per bin tiled values on per-pt scaled')
+
+grid.arrange(
+  ggplot(full_join(t.dist, v.dist, by='bin'), aes(mean.x,mean.y)) + geom_point() + labs(x='train', y='val', title='mean'),
+  ggplot(full_join(t.dist, v.dist, by='bin'), aes(sd.x,sd.y)) + geom_point() + labs(x='train', y='val', title='SD'),
+  ggplot(full_join(t.dist, v.dist, by='bin'), aes(var.x,var.y)) + geom_point() + labs(x='train', y='val', title='var'),
+  ggplot(full_join(t.dist, v.dist, by='bin'), aes(skew.x,skew.y)) + geom_point() + labs(x='train', y='val', title='Skew'),
+ncol=2)
+
+```
+
+Just to check...after the arm segments have been subtracted as well.
+```{r}
+
+t.pp.scale = BarrettsProgressionRisk:::subtractArms(as.matrix(t.per.pt.scale[,-1]), as.matrix(t.arm.pp.scale[,-1])) %>% as_tibble %>% add_column('Sample' = 'x', .before=1)
+v.pp.scale = BarrettsProgressionRisk:::subtractArms(as.matrix(v.per.pt.scale[,-1]), as.matrix(v.arm.pp.scale[,-1]))  %>% as_tibble %>% add_column('Sample' = 'x', .before=1)
+
+t.dist = tile.dist.measure(t.pp.scale) %>% mutate(cohort='train')
+v.dist = tile.dist.measure(v.pp.scale) %>% mutate(cohort='val')
+
+gridExtra::grid.arrange(
+  ggplot(bind_rows(t.dist, v.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
+    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
+    labs(title='Mean'),
+  ggplot(bind_rows(t.dist, v.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
+    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
+    labs(title='SD'), 
+  ncol=2, top='Per bin tiled values on per-sample scaled')
+
+```
+
+### Predictions
+
+Using a model built on per-sample scaled data. Unfortunately that has not resolved the predictions.
+
+```{r}
+preds = do.call(bind_rows, purrr::map(
+  list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf/predictions_per_pt_scale/0.9/', pattern='predictions.tsv', recursive = T, full.names = T), 
+  function(f) read_tsv(f, col_types = 'cddcccccccc'))) %>% mutate(Endoscopy = as.Date(Endoscopy)) %>% dplyr::select(-`Path Notes`) %>% mutate(`Hospital Research ID` = str_replace_all( str_remove_all(`Hospital Research ID`, " "), '/', '_'))
+
+preds = left_join(preds,all.val, by=c('Sample'='Samplename', 'Hospital Research ID', 'SLX-ID', 'Index Sequence', 'Block ID'))
+
+# Non progressors
+# FPR is ~40% still
+fpr = round(nrow(preds %>% filter(Status == 'NP' & Risk == 'High'))/nrow(preds %>% filter(Status == 'NP')),2)
+
+preds %>% dplyr::filter(Status == 'NP') %>% group_by(Risk) %>% tally() %>% spread(Risk, n) %>% 
+  kable(caption=paste0('NP risk predictions, FPR=',fpr)) %>% kable_styling(full_width = F)
+
+# Progressors
+# FNR is still over 50% in the old progressors as well
+fnr = round(nrow(preds %>% filter(Status == 'P' & Risk == 'Low' & !grepl('OCCAMS',`Hospital Research ID`)))/nrow(preds %>% filter(Status == 'P' &  !grepl('OCCAMS',`Hospital Research ID`))),2)
+
+preds %>% dplyr::filter(Status == 'P' & !grepl('OCCAMS',`Hospital Research ID`)) %>% group_by(Risk) %>% tally() %>% spread(Risk, n) %>% 
+  kable(caption=paste0('P risk predictions, FNR=',fnr)) %>% kable_styling(full_width = F)
+
+```
+
+## Per bin variance
+
+No scaling at this point.
 
@@
 There's a lot more variation in the means per bin in the validation data.
-This is before unit normalization, which is based on the training data mean/sd, so where the variance causes issues is starting to become more clear.
+This is before bin scaling, which is based on the training data mean/sd, so where the variance causes issues is starting to become more clear.
 
 ```{r}
-tile.dist.measure<-function(df) {
-  means = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~mean(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('mean' = 'V1')
-  sds = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~sd(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('sd' = 'V1')
-  iqrs = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~IQR(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('iqrs' = 'V1')
-  vars = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~var(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('var' = 'V1')
-
-  skew = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~skewness(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('skew' = 'V1')
-  kurt = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~kurtosis(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('kurtosis' = 'V1')
-
-  dist = means %>% full_join(sds, by='bin') %>% full_join(iqrs, by='bin') %>% full_join(vars, by='bin') %>% full_join(skew, by='bin') %>% full_join(kurt, by='bin')
-  return(dist)
-}
-
 t.dist = tile.dist.measure(training.tiles) %>% mutate(cohort='train')
@@
 
-### Per bin, after scaling per-pt
+## Scale validation cohort within itself
 
-```{r}
-t.dist = tile.dist.measure(t.per.pt.scale) %>% mutate(cohort='train')
-v.dist = tile.dist.measure(v.per.pt.scale) %>% mutate(cohort='val')
-
-gridExtra::grid.arrange(
-  ggplot(bind_rows(t.dist, v.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
-    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
-    labs(title='Mean'),
-  ggplot(bind_rows(t.dist, v.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
-    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
-    labs(title='SD'), 
-  ncol=2, top='Per bin tiled values on per-pt scaled')
-```
-
-After the arm segments have been subtracted as well.
-```{r}
-
-t.pp.scale = BarrettsProgressionRisk:::subtractArms(as.matrix(t.per.pt.scale[,-1]), as.matrix(t.arm.pp.scale[,-1])) %>% as_tibble %>% add_column('Sample' = 'x', .before=1)
-v.pp.scale = BarrettsProgressionRisk:::subtractArms(as.matrix(v.per.pt.scale[,-1]), as.matrix(v.arm.pp.scale[,-1]))  %>% as_tibble %>% add_column('Sample' = 'x', .before=1)
-
-t.dist = tile.dist.measure(t.pp.scale) %>% mutate(cohort='train')
-v.dist = tile.dist.measure(v.pp.scale) %>% mutate(cohort='val')
-
-gridExtra::grid.arrange(
-  ggplot(bind_rows(t.dist, v.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
-    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
-    labs(title='Mean'),
-  ggplot(bind_rows(t.dist, v.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
-    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
-    labs(title='SD'), 
-  ncol=2, top='Per bin tiled values on per-pt scaled')
-
-```
-
-### Predictions
-
-```{r}
-preds = do.call(bind_rows, purrr::map(
-  list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf/predictions_per_pt_scale/0.9/', pattern='predictions.tsv', recursive = T, full.names = T), 
-  function(f) read_tsv(f, col_types = 'cddcccccccc'))) %>% mutate(Endoscopy = as.Date(Endoscopy)) %>% dplyr::select(-`Path Notes`) %>% mutate(`Hospital Research ID` = str_replace_all( str_remove_all(`Hospital Research ID`, " "), '/', '_'))
-
-preds = left_join(preds,all.val, by=c('Sample'='Samplename', 'Hospital Research ID', 'SLX-ID', 'Index Sequence', 'Block ID'))
-
-# Non progressors
-# FPR is ~40% still
-fpr = round(nrow(preds %>% filter(Status == 'NP' & Risk == 'High'))/nrow(preds %>% filter(Status == 'NP')),2)
-
-preds %>% dplyr::filter(Status == 'NP') %>% group_by(Risk) %>% tally() %>% spread(Risk, n) %>% 
-  kable(caption=paste0('NP risk predictions, FPR=',fpr)) %>% kable_styling(full_width = F)
-
-# Progressors
-# FNR is still over 50% in the old progressors as well
-fnr = round(nrow(preds %>% filter(Status == 'P' & Risk == 'Low' & !grepl('OCCAMS',`Hospital Research ID`)))/nrow(preds %>% filter(Status == 'P' &  !grepl('OCCAMS',`Hospital Research ID`))),2)
-
-preds %>% dplyr::filter(Status == 'P' & !grepl('OCCAMS',`Hospital Research ID`)) %>% group_by(Risk) %>% tally() %>% spread(Risk, n) %>% 
-  kable(caption=paste0('P risk predictions, FNR=',fnr)) %>% kable_styling(full_width = F)
-
-```
-
-## Normalize validation cohort within itself
-
-Instead scaling using z-values from the trianing data, I can do it from within the validation cohort.
+Instead scaling using z-values from the training data, I can do it from within the validation cohort.
 This looks better here, but when I use this for predictions it doesn't significantly change the results.
@@
   list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf/predictions_scale_within_cohort/0.9/', pattern='predictions.tsv', recursive = T, full.names = T), 
+  function(f) read_tsv(f, col_types = 'cddcccccccc'))) %>% mutate(Endoscopy = as.Date(Endoscopy)) %>% dplyr::select(-`Path Notes`) %>% mutate(`Hospital Research ID` = str_replace_all( str_remove_all(`Hospital Research ID`, " "), '/', '_'))
+
+preds = left_join(preds,all.val, by=c('Sample'='Samplename', 'Hospital Research ID', 'SLX-ID', 'Index Sequence', 'Block ID'))
+
+# Non progressors
+# FPR is ~40% still
+fpr = round(nrow(preds %>% filter(Status == 'NP' & Risk == 'High'))/nrow(preds %>% filter(Status == 'NP')),2)
+
+preds %>% dplyr::filter(Status == 'NP') %>% group_by(Risk) %>% tally() %>% spread(Risk, n) %>% 
+  kable(caption=paste0('NP risk predictions, FPR=',fpr)) %>% kable_styling(full_width = F)
+
+# Progressors
+# FNR is still over 50% in the old progressors as well
+fnr = round(nrow(preds %>% filter(Status == 'P' & Risk == 'Low' & !grepl('OCCAMS',`Hospital Research ID`)))/nrow(preds %>% filter(Status == 'P' &  !grepl('OCCAMS',`Hospital Research ID`))),2)
+
+preds %>% dplyr::filter(Status == 'P' & !grepl('OCCAMS',`Hospital Research ID`)) %>% group_by(Risk) %>% tally() %>% spread(Risk, n) %>% 
+  kable(caption=paste0('P risk predictions, FNR=',fnr)) %>% kable_styling(full_width = F) 
+```
+
+## Scale validation using global z scores
+
+```{r}
+v.means = apply(as.matrix(bind_rows(training.tiles, val.tiles)[,-1]),2, mean, na.rm=T)
+v.sd = apply(as.matrix(bind_rows(training.tiles, val.tiles)[,-1]),2, sd, na.rm=T)
+
+val.int.norm = as.matrix(val.tiles[,-1])
+for (i in 1:ncol(val.int.norm)) 
+  val.int.norm[,i] = BarrettsProgressionRisk:::unit.var(val.int.norm[,i], v.means[i], v.sd[i])
+
+val.int.norm = as_tibble(val.int.norm) %>% add_column('Sample'=val.tiles$Sample,.before=T)
+vinorm.dist = tile.dist.measure(val.int.norm) %>% mutate(cohort='val glob')
+
+gridExtra::grid.arrange(
+  ggplot(bind_rows(tnorm.dist, vinorm.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
+    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
+    labs(title='Mean'),
+  ggplot(bind_rows(tnorm.dist, vinorm.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
+    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
+    labs(title='SD'),
+  ncol=2, top='Per bin scaled tile values', bottom='Val glob = validation bins normalized using global z-scores')
+```
+
+### Predictions
+
+```{r}
+preds = do.call(bind_rows, purrr::map(
+  list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf/predictions_5e6_all_var/0.9/', pattern='predictions.tsv', recursive = T, full.names = T), 
   function(f) read_tsv(f, col_types = 'cddcccccccc'))) %>% mutate(Endoscopy = as.Date(Endoscopy)) %>% dplyr::select(-`Path Notes`) %>% mutate(`Hospital Research ID` = str_replace_all( str_remove_all(`Hospital Research ID`, " "), '/', '_'))
@@
 # Non progressors
-preds %>% dplyr::filter(Status == 'NP') %>% group_by(Risk) %>% tally() %>% spread(Risk, n)
+# FPR is ~40% still
+fpr = round(nrow(preds %>% filter(Status == 'NP' & Risk == 'High'))/nrow(preds %>% filter(Status == 'NP')),2)
 
-# FPR is nearly 40% still
-nrow(preds %>% filter(Status == 'NP' & Risk == 'High'))/nrow(preds %>% filter(Status == 'NP'))
+preds %>% dplyr::filter(Status == 'NP') %>% group_by(Risk) %>% tally() %>% spread(Risk, n) %>% 
+  kable(caption=paste0('NP risk predictions, FPR=',fpr)) %>% kable_styling(full_width = F)
 
 # Progressors
-preds %>% dplyr::filter(Status == 'P' & !grepl('OCCAMS',`Hospital Research ID`)) %>% group_by(Risk) %>% tally() %>% spread(Risk, n)
+# FNR is still over 50% in the old progressors as well
+fnr = round(nrow(preds %>% filter(Status == 'P' & Risk == 'Low' & !grepl('OCCAMS',`Hospital Research ID`)))/nrow(preds %>% filter(Status == 'P' &  !grepl('OCCAMS',`Hospital Research ID`))),2)
 
+preds %>% dplyr::filter(Status == 'P' & !grepl('OCCAMS',`Hospital Research ID`)) %>% group_by(Risk) %>% tally() %>% spread(Risk, n) %>% 
+  kable(caption=paste0('P risk predictions, FNR=',fnr)) %>% kable_styling(full_width = F) 
+```
+
+# Model trained on all samples
+
+In this case I'm looking at the leave-one-out predictions, so somewhat different but at least gives an idea. While the progressors don't change, the non-progressors see a signficant shift.
+
+```{r}
+
+load('~/Data/BarrettsProgressionRisk/Analysis/all-samples/loo.Rdata')
+
+pg.samp = pg.samp %>% filter(Hospital.Research.ID %in% all.val$`Hospital Research ID` & !grepl('OCCAMS',Hospital.Research.ID)) %>%
+  mutate(qt = cut(Prediction, breaks=seq(0,1,.1)))
+
+preds = pg.samp %>% left_join(BarrettsProgressionRisk:::pred.confidence %>% dplyr::select(quant,Risk), by = c('qt'='quant'))
+
+# Non progressors
+# FPR is ~40% still
+fpr = round(nrow(preds %>% filter(Status == 'NP' & Risk == 'High'))/nrow(preds %>% filter(Status == 'NP')),2)
+
+preds %>% dplyr::filter(Status == 'NP') %>% group_by(Risk) %>% tally() %>% spread(Risk, n) %>% 
+  kable(caption=paste0('NP risk predictions, FPR=',fpr)) %>% kable_styling(full_width = F)
+
+# Progressors
 # FNR is still over 50% in the old progressors as well
-nrow(preds %>% filter(Status == 'P' & Risk == 'Low' & !grepl('OCCAMS',`Hospital Research ID`)))/nrow(preds %>% filter(Status == 'P' &  !grepl('OCCAMS',`Hospital Research ID`)))
-```
+fnr = round(nrow(preds %>% filter(Status == 'P' & Risk == 'Low' & !grepl('OCCAMS',`Hospital.Research.ID`)))/nrow(preds %>% filter(Status == 'P' &  !grepl('OCCAMS',`Hospital.Research.ID`))),2)
+
+preds %>% dplyr::filter(Status == 'P' & !grepl('OCCAMS',`Hospital.Research.ID`)) %>% group_by(Risk) %>% tally() %>% spread(Risk, n) %>% 
+  kable(caption=paste0('P risk predictions, FNR=',fnr)) %>% kable_styling(full_width = F) 
 ```

### 2019-09-19 Raw Data Comparison.Rmd
...

@@
 if (file.exists('prepped.Rdata')) {
-  load('prepped.Rdata')
+  load('prepped.Rdata',verbose=T)
 } else {
@@
 
+In this case the z-scores are drawn from the entire cohort similarly to what I would do before training a model with all samples.
+
 ```{r}
@@
 v.sd = apply(as.matrix(bind_rows(training.tiles, val.tiles)[,-1]),2, sd, na.rm=T)
+
+tr.int.norm = as.matrix(training.tiles[,-1])
+for (i in 1:ncol(tr.int.norm)) 
+  tr.int.norm[,i] = BarrettsProgressionRisk:::unit.var(tr.int.norm[,i], v.means[i], v.sd[i])
+
+tr.int.norm = as_tibble(tr.int.norm) %>% add_column('Sample'=training.tiles$Sample,.before=T)
+trinorm.dist = tile.dist.measure(tr.int.norm) %>% mutate(cohort='training glob')
 
@@
 gridExtra::grid.arrange(
-  ggplot(bind_rows(tnorm.dist, vinorm.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
+  ggplot(bind_rows(trinorm.dist, vinorm.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
     geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
     labs(title='Mean'),
-  ggplot(bind_rows(tnorm.dist, vinorm.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
+  ggplot(bind_rows(trinorm.dist, vinorm.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
     geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
@@
 ### Predictions
+
+Starts looking a little better in non-progressors.
 

### 2019-09-19 sWGS Normals.Rmd
...

@@
 
-library(Matrix)
+library(tidyverse)
+library(BarrettsProgressionRisk)
 library(glmnet)
-library(ggplot2)
 library(gridExtra)
@@
 source('lib/load_patient_metadata.R')
-source('lib/data_func.R')
-source('lib/fastPCF.R')
+#source('lib/data_func.R')
+#source('lib/fastPCF.R')
 
-data.files = list.files('~/Data/Ellie/QDNAseq', full.names = T)
+data.dir = '~/Data/BarrettsProgressionRisk/QDNAseq'
+analysis.dir = '~/Data/BarrettsProgressionRisk/Analysis'
+#data.files = list.files(, full.names = T)
 
-rpi = read.patient.info('~/Data/Ellie/QDNAseq/All_patient_info.xlsx')
-normals = rpi$removed
-path = read_excel('~/Data/Ellie/QDNAseq/All_patient_info.xlsx', 'Normals')[,c('Patient ID', 'Hospital Research ID', 'Pathology notes','Plate Index','SLX ID 1')]
-colnames(path)[c(1,4,5)] = c('Patient','Plate.Index','SLX.ID')
-path$SLX.ID = sub('SLX-','', gsub(' ', '', path$SLX.ID))
-normals = merge(normals, path, by=c('Patient','Plate.Index','SLX.ID'))
+modeldir = '~/Data/BarrettsProgressionRisk/Analysis/5e6_arms_all'
 
-data.file = grep('MultisampleCopynumberBinnedAndFitted2.RData', data.files, value=T)
-if (length(data.file) <= 0)
-  stop("Missing necessary Rdata file from Multisample script")
-load(file=data.file, verbose=T)
+patient.file = list.files(data.dir, pattern='All_patient_info.xlsx', recursive=T, full.names=T)
+demo.file = list.files(data.dir, pattern='Demographics_full.xlsx', recursive=T, full.names=T)
 
-loc.cols = c('location','chrom','start','end')
-fit.data = fit.data[,c(loc.cols,normals$Samplename)]
-raw.data = raw.data[,c(loc.cols,normals$Samplename)]
+rpi = read.patient.info(patient.file, demo.file)
+normals = rpi$normal
 
-exclude.file = grep('qDNAseq_blacklistedRegions.txt', data.files, value=T)
-if (length(data.file) <= 0)
-  stop(paste("Missing necessary exclusion file from 'qDNAseq_blacklistedRegions.txt' in",data))
-blacklisted.regions = read.table(exclude.file,sep="\t",header=T,stringsAsFactors=F)
+data.files = list.files(paste0(analysis.dir, '/normals'), 'segment.Rdata', recursive = T, full.names = T)
 
-if (file.exists('/tmp/normals.Rdata')) {
-  load('/tmp/normals.Rdata', verbose=T)
-} else {
-  plist = list()
-  seg.residuals = list()
-  tiled.normals = NULL; tiled.normal.arms = NULL
-  for (sample in normals$Samplename) {
-    binned = binSWGS(raw.data[,c(loc.cols,sample)],fit.data[,c(loc.cols,sample)],blacklisted.regions,min.probes=67,gamma2=250,metrics.file=NULL)
-    plist[[sample]] = binned$plots[[1]]
-    seg.residuals[[sample]] = binned$resvar[[sample]]
-
-    tiled = tile.segmented.data(binned$segvals,chr.info=get.chr.lengths(file='/tmp/hg19_info.txt'),verbose=T)
-    tiled.arms = tile.segmented.data(binned$segvals,chr.info=get.chr.lengths(file='/tmp/hg19_info.txt'),size='arms', verbose=T)
-    if (is.null(tiled.normals) | is.null(tiled.normal.arms)) {
-      tiled.normals = tiled
-      tiled.normal.arms = tiled.arms
-    } else {
-      tiled.normals = cbind(tiled.normals, tiled[,sample,drop=F])
-      tiled.normal.arms = cbind(tiled.normal.arms, tiled.arms[,sample,drop=F])
-    }
-  }
-  save(tiled.normals, tiled.normal.arms, plist, seg.residuals, file='/tmp/normals.Rdata')
-}
+resids = do.call(bind_rows, purrr::map(list.files(paste0(analysis.dir, '/normals'), 'residuals.tsv', recursive = T, full.names = T), function(f) read_tsv(f)))
 
@@
 ```{r, echo=F}
-modeldir = '~/Data/Ellie/Analysis/5e6_arms_all'
+
 file = paste(modeldir, 'model_data.Rdata', sep='/')

### 2019-09-19 processing_pipeline/process_training_data.R

@@ -16,7 +16,7 @@ patient.file = args[2]
 outdir = args[3]
 # data = '~/Data/Ellie/QDNAseq/training/'
 # patient.file = paste(data, 'All_patient_info.xlsx', sep='/')
-# outdir = '~/Data/Ellie/Analysis'
+# outdir = '~/Data/Ellie/Analysis/IntPloidy'
 
 patient.name = NULL
 if (length(args) == 4)
@@ -138,14 +138,14 @@ for (pt in unique(pts_slx$Hospital.Research.ID)) {
   dir.create(pd,showWarnings = F)
   
   info = all.patient.info %>% filter(Hospital.Research.ID == pt) %>% 
-    select('Patient','Hospital.Research.ID', 'Samplename','Endoscopy.Year','Block','Pathology','p53.Status') %>% 
+    dplyr::select('Patient','Hospital.Research.ID', 'Samplename','Endoscopy.Year','Block','Pathology','p53.Status') %>% 
     dplyr::rename('Sample' = 'Samplename', 'GEJ.Distance' = 'Block', 'p53 IHC' = 'p53.Status', 'Endoscopy' = 'Endoscopy.Year') %>% rowwise %>% 
     dplyr::mutate( Endoscopy = paste(as.character(Endoscopy), '01', '01', sep='-')   )
   info = BarrettsProgressionRisk::loadSampleInformation(info, path=c('BE','ID','LGD','HGD','IMC'))
 
   cols = which(colnames(fit.data) %in% subset(all.patient.info, Hospital.Research.ID == pt)$Samplename)
 
-  segmented = BarrettsProgressionRisk::segmentRawData(info, raw.data[,c(1:4,cols)],fit.data[,c(1:4,cols)],verbose=T, cutoff=0.011)
+  segmented = BarrettsProgressionRisk::segmentRawData(info, raw.data[,c(1:4,cols)],fit.data[,c(1:4,cols)],intPloidy = T, verbose=T, cutoff=0.015*2)
   residuals = BarrettsProgressionRisk::sampleResiduals(segmented) %>% add_column('patient'=pt, .before=1)
 
   save(segmented, file=paste0(pd,'/segment.Rdata'))

### 2019-10-09 Raw Data Comparison.Rmd
...

@@
 
+# Adjusted to nearly integer ploidy?
+
+```{r}
+
+files = list.files('~/Data/BarrettsProgressionRisk/Analysis/IntPloidy/multipcf_perPatient/', '5e06_tiled', recursive=T, full.names = T)
+training.tiles.int = do.call(bind_rows, purrr::map(files, function(f) {
+  read_tsv(f,col_types=c(.default=col_double()))
+})) %>% dplyr::rename('Sample' = 'X1')
+
+files = list.files('~/Data/BarrettsProgressionRisk/Analysis/IntPloidy/multipcf_perPatient/', 'arms_tiled', recursive=T, full.names = T)
+training.arms.int = do.call(bind_rows, purrr::map(files, function(f) {
+  read_tsv(f,col_types=c(.default=col_double()))
+})) %>% dplyr::rename('Sample' = 'X1')
+
+files = list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/IntPloidy/', '5e06_cleaned_tiled', recursive=T, full.names = T)
+val.tiles.int = do.call(bind_rows, purrr::map(files, function(f) {
+  read_tsv(f,col_types=c(.default=col_double()))
+})) %>% dplyr::rename('Sample' = 'X1')
+
+files = list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/IntPloidy/', 'arms_cleaned_tiled', recursive=T, full.names = T)
+val.arms.int = do.call(bind_rows, purrr::map(files, function(f) {
+  read_tsv(f,col_types=c(.default=col_double()))
+})) %>% dplyr::rename('Sample' = 'X1')
+
+t.pt = per.pt.tile(training.tiles.int) %>% mutate(cohort='train')
+v.pt = per.pt.tile(val.tiles.int) %>% mutate(cohort='val')
+
+gridExtra::grid.arrange(
+  ggplot(bind_rows(t.pt, v.pt), aes(cohort, mean, group=cohort, fill=cohort)) + 
+    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
+    labs(title='Mean per patient'),
+  ggplot(bind_rows(t.pt, v.pt), aes(cohort, sd, group=cohort, fill=cohort)) + 
+    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
+    labs(title='StdDev per patient'),
+  ncol=2, top='Per patient tiled values')
+
+mean(tile.dist.measure(training.tiles.int)$mean)
+mean(tile.dist.measure(training.tiles.int)$sd)
+
+mean(tile.dist.measure(val.tiles.int)$mean)
+mean(tile.dist.measure(val.tiles.int)$sd)
+
+```
+
 ```{r eval=F}

### 2019-10-15 QDNAseq segmentation.Rmd
...

---
title: "QDNAseq segmentation"
author: "Sarah Killcoyne"
date: "03/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
sheets = readxl::excel_sheets(val.file)[8:13]

pastefun<-function(x) {
  if ( !grepl('SLX-', x) ) x = paste0('SLX-',x)
  return(x)
}

val.info = do.call(bind_rows, lapply(sheets, function(s) {
  readxl::read_xlsx(val.file, s) %>% dplyr::select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA, matches('Collection')) %>% 
    dplyr::filter(!is.na(`SLX-ID`)) %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) 
})) %>% mutate(Samplename = paste(`SLX-ID`,`Index Sequence`, sep='.')) %>% 
  rowwise %>% mutate_at(vars(`SLX-ID`), list(pastefun) ) %>% ungroup %>% 
  mutate(
  `Hospital Research ID` = str_replace_all( str_remove_all(`Hospital Research ID`, " "), '/', '_'), 
  `Index Sequence` = str_replace_all(`Index Sequence`, 'tp', ''),
  Samplename = paste(`SLX-ID`,`Index Sequence`,sep='.')
  )

training.raw = read_tsv('~/Data/BarrettsProgressionRisk/Analysis/multipcf_perPatient/pre_seg_medians.tsv', col_types = 'cdd') %>% dplyr::rename(sd='SD',sample='Sample') %>% mutate(cohort = 'train')

v15.file= '~/Data/BarrettsProgressionRisk/Analysis/validation/pre_seg_disp.tsv'
all.val15 = read_tsv(v15.file, col_types = 'ddc') %>% dplyr::select(sample,median,sd) %>% mutate(cohort='val')
```

# Raw Data

I'm looking at the variance in the raw data immediately following QDNAseq processing. The training data was all run with a 15kb bin size so in the initial runs of the validation data all samples were run with that bin size.

## Comparing the distributions

Training vs Validation set

```{r, echo=F}
ggplot(bind_rows(training.raw, all.val15)) + xlim(0.95,1.05) + ylim(0,1.0) + 
  geom_point(aes(median,sd,color=cohort)) + 
  theme_minimal() + labs(title='Median vs SD for available samples')
```

The median is pretty close to the same, but there is a larger spread in the validation dataset.

```{r}
tv15 = bind_rows(training.raw, all.val15) %>% mutate(cohort = factor(cohort))

ggplot(tv15, aes(median)) + facet_grid(~cohort, scales = 'free_x') +# xlim(0.95,1.05) +
  geom_histogram(bins=20, color='lightblue')

ggplot(tv15, aes(cohort, median, group=cohort, fill=cohort)) + ylim(0.95,1.05) +
  geom_jitter(width=0.2) + geom_boxplot()

wilcox.test(all.val15$median, training.raw$median)
```

The difference is clearer in the std. deviation of the two cohorts, even when the extreme outliers are removed.

```{r}
ggplot(tv15, aes(sd)) + facet_grid(~cohort) + xlim(0,1) +
  geom_histogram(bins=50, color='lightblue')

ggplot(tv15, aes(cohort, sd, group=cohort, fill=cohort)) + ylim(0,1) + 
  geom_jitter(width=0.2) + geom_boxplot(outlier.colour = NA)

wilcox.test(all.val15$sd, training.raw$sd)
```

### Per sample the shift is event more obvious.

Dotted line is median(sd).
```{r, echo=F}
ggplot(tv15, aes(x=1:nrow(tv15),y=sd)) + ylim(0,1.0) + 
  geom_point(aes(color=cohort)) + geom_hline(yintercept = median(training.raw$sd), linetype='dashed') +
  #geom_hline(yintercept = median(training.raw$sd)+mad(all.val15$sd), color='grey39') +
  theme(axis.text.x = element_blank()) +
  theme_minimal() + labs(title='Training and validation cohorts (15kb bins)')
```

## Altering the QDNAseq bin size

Looking at different bin sizes in the validation cohort. Ellie's bam files are not as easily recovered so I'm using the training cohort as the standard and altering the validation cohort. The median variance decreases as I increase the bin size.
```{r}
files = grep('ellie',list.files('~/Data/BarrettsProgressionRisk/data', 'prepped_dist.tsv', recursive = T, full.names = T),invert = T, value=T)
val = do.call(bind_rows, lapply(files, function(f) {
  read_tsv(f, col_types = 'cdd') %>% mutate('seg.size'=basename(dirname(f)))
})) %>% mutate(seg.size = factor(seg.size)) %>% arrange(sample, seg.size)

val = val %>% mutate(kb = as.integer(sub('kb','',seg.size)), seg.size = factor(seg.size, levels=c('15kb','30kb','50kb','100kb','500kb'), ordered=T)) %>% arrange(kb) %>% mutate(index = row_number())

samples = val %>% group_by(sample) %>% tally %>% dplyr::filter(n == 4) %>% select(sample) %>% pull

#val = val %>% group_by(sample) %>% mutate(diff = dist(sd))
ggplot(val,  aes(index, sd, color=seg.size)) + ylim(0,1) + 
  geom_point() + 
  theme(axis.text.x=element_blank()) + theme_minimal()

ggplot(val %>% arrange(kb),  aes(seg.size, sd, color=seg.size)) + ylim(0,1) + 
  geom_jitter(width=0.2) + geom_boxplot(alpha=0.8, outlier.colour = NA) +
  theme_minimal() + theme(axis.text.x=element_blank(), legend.position = 'none')

val = val %>% filter(kb > 15)

tv15 = tv15 %>% mutate(seg.size = '15kb', kb=15) %>% dplyr::select(sample,cohort,seg.size,kb,sd,median)
val = val %>% mutate(cohort = paste('val',seg.size)) %>% dplyr::select(sample,cohort,seg.size,kb,sd,median)

merged = bind_rows(val,tv15)

merged = merged %>% mutate(
  seg.size = factor(seg.size, levels=c('train','15kb','30kb','50kb','100kb','500kb'), ordered=T),
  cohort = factor(cohort, levels=c('train','val','val 30kb','val 50kb','val 100kb','val 500kb'), ordered=T)
  ) %>% arrange(seg.size)

ggplot(merged, aes(seg.size, sd, group=seg.size, fill=seg.size)) + facet_grid(~cohort, scales='free_x') + ylim(0,1) +
  geom_jitter(width=0.2) + geom_boxplot(outlier.colour = NA) + 
  scale_fill_brewer(palette = 'Set1') + theme_minimal()

```

So it appears that I can decrease the variance by increasing the bin size.  But it doesn't necessarily make sense to do it across the board.  I can take the median of the training data and call that the cutoff for deciding.

```{r}
ggplot(merged, aes(x=1:nrow(merged),y=sd)) + ylim(0,1.0) + 
  geom_point(aes(color=cohort)) + geom_hline(yintercept = median(training.raw$sd), linetype='dashed') +
  geom_hline(yintercept = median(training.raw$sd)+0.5*mad(all.val15$sd), color='grey39') +
  theme(axis.text.x = element_blank()) +
  theme_minimal()

out.samples = all.val15 %>% filter(sd > median(training.raw$sd))# +0.5*mad(all.val15$sd)))

```

In the validation dataset, starting at 15kb `r nrow(out.samples)`:`r nrow(all.val15)` samples have a variance greater than the median.

```{r, echo=F}
preds15 = do.call(bind_rows, lapply(list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf/predictions_model_5e6/0.9','predictions',full.names=T, recursive=T), function(f) read_tsv(f,col_types='cddcccDcccc')))

#preds15 %>% filter(Sample %in% out.samples$sample) %>% group_by(Risk) %>% tally
```

## Re-process dynamically

I re-processed all of the validation data with this in mind.  I start with the 15kb bin size, immediately following the QDNAseq processing I check the variance of the raw data.  If it's greater than the median of the training data (`r median(training.raw$sd)`) then I use the next greater bin size for that sample.  I do this up to a bin size of 100kb currently.

```{r, echo=F}
library(BarrettsProgressionRisk)

analysis.dir = '~/Data/BarrettsProgressionRisk/Analysis/validation/resegment/'

per.sample.var = do.call(bind_rows, lapply(list.files(analysis.dir, 'variance.tsv', recursive=T, full.names=T), function(f) {
  read_tsv(f, col_types='cddd') %>% mutate(patient = basename(sub('plots','',dirname(f))))
})

```

### 2019-10-16 QDNAseq segmentation.Rmd
...

@@
 library(tidyverse)
+library(kableExtra)
 
@@
 val.info = do.call(bind_rows, lapply(sheets, function(s) {
-  readxl::read_xlsx(val.file, s) %>% dplyr::select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA, matches('Collection')) %>% 
+  readxl::read_xlsx(val.file, s) %>% dplyr::select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Pathology, Cohort, Batch, matches('Collection|Notes')) %>%
     dplyr::filter(!is.na(`SLX-ID`)) %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) 
@@
   read_tsv(f, col_types='cddd') %>% mutate(patient = basename(sub('plots','',dirname(f))))
-})
+})) %>% dplyr::select(patient, Samples, everything())
+
+max.bin<-function(bin15,bin50,bin100) {
+  if (!is.na(bin100))
+    return(100)
+  if (!is.na(bin50))
+    return(50)
+  return(15)
+}
+
+per.sample.bin = per.sample.var %>% group_by(Samples) %>% dplyr::summarise( max.bin = max.bin(`15kb`,`50kb`,`100kb`) ) %>% arrange(max.bin)
+
+per.sample.bin %>% group_by(max.bin) %>% tally() %>% spread(max.bin,n) %>% kable(caption = 'Number of samples per QDNAseq bin size') %>% kable_styling('striped', full_width = F)
+```
+
+```{r}
+
+per.sample.bin = val.info %>% dplyr::select(matches('Hospital|Status|Block|Pathology|Samplename')) %>% 
+  left_join(per.sample.bin, by=c('Samplename'='Samples'))
+
+per.sample.bin %>% group_by(max.bin, Status) %>% tally()
+
+per.sample.bin %>% filter(is.na(max.bin))
 

### 2019-10-16 processing_pipeline/process_predict_val.R

@@ -163,11 +163,11 @@ get.qdnaseg<-function(samples, dir) {
       if (length(which(sampleResiduals(segmented)$Pass)) > 0) {
         prr = predictRiskFromSegments(segmented, be.model = be.model)
         predictions = bind_rows(predictions, predictions(prr))
+        plots = BarrettsProgressionRisk::copyNumberMountainPlot(prr,annotate = T,legend = F, as='list')
+        for (s in names(plots))
+          ggsave(paste(pred.dir, paste0(s,'_cnMtn.png'),sep='/'), plot=plots[[s]], height=4, width=20, units='in')
       }
       
-      plots = BarrettsProgressionRisk::copyNumberMountainPlot(prr,annotate = T,legend = F, as='list')
-      for (s in names(plots))
-        ggsave(paste(pred.dir, paste0(s,'_cnMtn.png'),sep='/'), plot=plots[[s]], height=4, width=20, units='in')
     }
     
     variance %>% write_tsv(paste0(plot.dir,'/variance.tsv'))

### 2019-10-25 QDNAseq segmentation.Rmd
changing cv to select different alpha values

@@
 library(tidyverse)
+library(reshape2)
+library(gridExtra)
 library(kableExtra)
+source('lib/load_patient_metadata.R')
+
+info.file = '~/Data/BarrettsProgressionRisk/QDNAseq/training/All_patient_info.xlsx'
+train.info = read.patient.info(info.file)$info
+
+qdna.dirs = '~/Data/BarrettsProgressionRisk/Analysis/pcf_perPatient/'
+train.dists.by.kb = purrr::map( list.files(qdna.dirs, full.names=T), function(dir) {
+  do.call(bind_rows, lapply(list.files(dir, 'raw_dist.tsv', full.names = T, recursive = T), function(f) {
+    read_tsv(f,col_types = 'ccdddddddd') %>% mutate(kb = basename(dir))  
+  }))
+})
+names(train.dists.by.kb) = basename(list.files(qdna.dirs, full.names=T))
+
+train.resids = purrr::map( list.files(qdna.dirs, full.names=T), function(dir) {
+  do.call(bind_rows, lapply(list.files(dir, 'residuals.tsv', full.names = T, recursive = T), function(f) {
+    read_tsv(f,col_types = 'ccddddl') %>% mutate(kb = basename(dir))  
+  }))
+})
+names(train.resids) = basename(list.files(qdna.dirs, full.names=T))
 
@@
 
-training.raw = read_tsv('~/Data/BarrettsProgressionRisk/Analysis/multipcf_perPatient/pre_seg_medians.tsv', col_types = 'cdd') %>% dplyr::rename(sd='SD',sample='Sample') %>% mutate(cohort = 'train')
+qdna.dirs = '~/Data/BarrettsProgressionRisk/Analysis/validation/pcf_perPatient'
+val.train.dists.by.kb = purrr::map( list.files(qdna.dirs, full.names=T), function(dir) {
+  do.call(bind_rows, lapply(list.files(dir, 'raw_dist.tsv', full.names = T, recursive = T), function(f) {
+    read_tsv(f,col_types = 'ccdddddddd') %>% mutate(kb = basename(dir))  
+  }))
+})
+names(val.train.dists.by.kb) = basename(list.files(qdna.dirs, full.names=T))
 
-v15.file= '~/Data/BarrettsProgressionRisk/Analysis/validation/pre_seg_disp.tsv'
-all.val15 = read_tsv(v15.file, col_types = 'ddc') %>% dplyr::select(sample,median,sd) %>% mutate(cohort='val')
+val.resids = purrr::map( list.files(qdna.dirs, full.names=T), function(dir) {
+  do.call(bind_rows, lapply(list.files(dir, 'residuals.tsv', full.names = T, recursive = T), function(f) {
+    read_tsv(f,col_types = 'ccddddl') %>% mutate(kb = basename(dir))  
+  }))
+})
+names(val.resids) = basename(list.files(qdna.dirs, full.names=T))
 ```
@@
 
-I'm looking at the variance in the raw data immediately following QDNAseq processing.
-The training data was all run with a 15kb bin size so in the initial runs of the validation data all samples were run with that bin size.
+## Training Data
 
-## Comparing the distributions
+The training data was all reprocessed at different QDNAseq bin sizes
 
-Training vs Validation set
+```{r}
+m = melt(do.call(bind_rows, train.dists.by.kb), id.vars='kb', measure.vars=c('median'))
+m = m %>% mutate(kb = factor(kb, levels=c('15kb','50kb','100kb','500kb'), ordered=T))
+ggplot(m, aes(kb, value, group=kb)) + geom_boxplot(aes(fill=kb), show.legend = F) + ylim(0.95,1.05) + 
+  scale_fill_brewer(palette='Set1') + 
+  labs(title='Training data median per QDNAseq bin size', y='median',x='') + theme_minimal()
 
-```{r, echo=F}
-ggplot(bind_rows(training.raw, all.val15)) + xlim(0.95,1.05) + ylim(0,1.0) + 
-  geom_point(aes(median,sd,color=cohort)) + 
-  theme_minimal() + labs(title='Median vs SD for available samples')
+m = melt(do.call(bind_rows, train.dists.by.kb), id.vars='kb', measure.vars=c('stdev'))
+m = m %>% mutate(kb = factor(kb, levels=c('15kb','50kb','100kb','500kb'), ordered=T))
+ggplot(m, aes(kb, value, group=kb)) + geom_boxplot(aes(fill=kb), show.legend = F) + ylim(0,1) +
+  scale_fill_brewer(palette='Set1') + 
+  labs(title='Training data variance per QDNAseq bin size', y='stdev',x='') + theme_minimal()
+
+m %>% group_by(kb) %>% dplyr::summarise( outliers = length(which(value > 1)), median.variance = median(value), min.var=min(value), max.var=max(value)) %>% mutate_if(is.numeric, list(~round(.,3))) %>%
+  kable(caption='Training data variance') %>% kable_styling(full_width = F)
 ```
 
-The median is pretty close to the same, but there is a larger spread in the validation dataset.
+## Validation Data
+
+The validation data was all reprocessed at different QDNAseq bin sizes
 
 ```{r}
-tv15 = bind_rows(training.raw, all.val15) %>% mutate(cohort = factor(cohort))
+m = melt(do.call(bind_rows, val.dists.by.kb), id.vars='kb', measure.vars=c('median'))
+m = m %>% mutate(kb = factor(kb, levels=c('15kb','50kb','100kb','500kb'), ordered=T))
+ggplot(m, aes(kb, value, group=kb)) + geom_boxplot(aes(fill=kb), show.legend = F) + ylim(0.95,1.05) + 
+  scale_fill_brewer(palette='Set1') + 
+  labs(title='Validation data median per QDNAseq bin size', y='median',x='') + theme_minimal()
 
-ggplot(tv15, aes(median)) + facet_grid(~cohort, scales = 'free_x') +# xlim(0.95,1.05) +
-  geom_histogram(bins=20, color='lightblue')
+m = melt(do.call(bind_rows, val.dists.by.kb), id.vars='kb', measure.vars=c('stdev'))
+m = m %>% mutate(kb = factor(kb, levels=c('15kb','50kb','100kb','500kb'), ordered=T))
+ggplot(m, aes(kb, value, group=kb)) + geom_boxplot(aes(fill=kb), show.legend = F) + ylim(0,1) +
+  scale_fill_brewer(palette='Set1') + 
+  labs(title='Validation data variance per QDNAseq bin size', y='stdev',x='') + theme_minimal()
 
-ggplot(tv15, aes(cohort, median, group=cohort, fill=cohort)) + ylim(0.95,1.05) +
-  geom_jitter(width=0.2) + geom_boxplot()
-
-wilcox.test(all.val15$median, training.raw$median)
+m %>% group_by(kb) %>% dplyr::summarise( outliers = length(which(value > 1)), median.variance = median(value), min.var=min(value), max.var=max(value)) %>% mutate_if(is.numeric, list(~round(.,3))) %>%
+  kable(caption='Validation data variance') %>% kable_styling(full_width = F)
 ```
 
-The difference is clearer in the std. deviation of the two cohorts, even when the extreme outliers are removed.
+## Training vs Validation
+
+At every bin size there is an overall increase in the variance for the validation set.  It is not driven by the outliers.
 
 ```{r}
-ggplot(tv15, aes(sd)) + facet_grid(~cohort) + xlim(0,1) +
-  geom_histogram(bins=50, color='lightblue')
+all = bind_rows(do.call(bind_rows, train.dists.by.kb) %>% mutate(cohort='training'),
+                do.call(bind_rows, val.dists.by.kb) %>% mutate(cohort = 'validation'))
+all = all %>% mutate(kb = factor(kb, levels=c('15kb','50kb','100kb','500kb'), ordered=T))
 
-ggplot(tv15, aes(cohort, sd, group=cohort, fill=cohort)) + ylim(0,1) + 
-  geom_jitter(width=0.2) + geom_boxplot(outlier.colour = NA)
+m = melt(all, id.vars=c('kb','cohort'), measure.vars='stdev')
+ggplot(m, aes(cohort, value, group=cohort)) + facet_grid(~kb) + ylim(0,1) +
+  geom_jitter(width=0.2,alpha=0.5) + geom_violin(aes(fill=kb), alpha=0.5, show.legend=F) + 
+  #geom_boxplot(aes(fill=kb), outlier.color=NA, show.legend = F) 
+  scale_fill_brewer(palette='Set1') + labs(title='Data variance per QDNAseq bin size', y='stdev',x='') + theme_minimal()
 
-wilcox.test(all.val15$sd, training.raw$sd)
-```
+m = melt(all, id.vars=c('kb','cohort'), measure.vars='median')
+ggplot(m, aes(cohort, value, group=cohort)) + facet_grid(~kb) + 
+  geom_jitter(width=0.2,alpha=0.5) + geom_violin(aes(fill=kb), alpha=0.5, show.legend=F) + ylim(0.95,1.1) +
+  scale_fill_brewer(palette='Set1') + labs(title='Median per QDNAseq bin size', y='median',x='') + theme_minimal()
 
-### Per sample the shift is event more obvious.
-
-Dotted line is median(sd).
-```{r, echo=F}
-ggplot(tv15, aes(x=1:nrow(tv15),y=sd)) + ylim(0,1.0) + 
-  geom_point(aes(color=cohort)) + geom_hline(yintercept = median(training.raw$sd), linetype='dashed') +
-  #geom_hline(yintercept = median(training.raw$sd)+mad(all.val15$sd), color='grey39') +
-  theme(axis.text.x = element_blank()) +
-  theme_minimal() + labs(title='Training and validation cohorts (15kb bins)')
-```
-
-## Altering the QDNAseq bin size
-
-Looking at different bin sizes in the validation cohort.
-Ellie's bam files are not as easily recovered so I'm using the training cohort as the standard and altering the validation cohort.
-The median variance decreases as I increase the bin size.
-```{r}
-files = grep('ellie',list.files('~/Data/BarrettsProgressionRisk/data', 'prepped_dist.tsv', recursive = T, full.names = T),invert = T, value=T)
-val = do.call(bind_rows, lapply(files, function(f) {
-  read_tsv(f, col_types = 'cdd') %>% mutate('seg.size'=basename(dirname(f)))
-})) %>% mutate(seg.size = factor(seg.size)) %>% arrange(sample, seg.size)
-
-val = val %>% mutate(kb = as.integer(sub('kb','',seg.size)), seg.size = factor(seg.size, levels=c('15kb','30kb','50kb','100kb','500kb'), ordered=T)) %>% arrange(kb) %>% mutate(index = row_number())
-
-samples = val %>% group_by(sample) %>% tally %>% dplyr::filter(n == 4) %>% select(sample) %>% pull
-
-#val = val %>% group_by(sample) %>% mutate(diff = dist(sd))
-ggplot(val,  aes(index, sd, color=seg.size)) + ylim(0,1) + 
-  geom_point() + 
-  theme(axis.text.x=element_blank()) + theme_minimal()
-
-ggplot(val %>% arrange(kb),  aes(seg.size, sd, color=seg.size)) + ylim(0,1) + 
-  geom_jitter(width=0.2) + geom_boxplot(alpha=0.8, outlier.colour = NA) +
-  theme_minimal() + theme(axis.text.x=element_blank(), legend.position = 'none')
-
-val = val %>% filter(kb > 15)
-
-tv15 = tv15 %>% mutate(seg.size = '15kb', kb=15) %>% dplyr::select(sample,cohort,seg.size,kb,sd,median)
-val = val %>% mutate(cohort = paste('val',seg.size)) %>% dplyr::select(sample,cohort,seg.size,kb,sd,median)
-
-merged = bind_rows(val,tv15)
-
-merged = merged %>% mutate(
-  seg.size = factor(seg.size, levels=c('train','15kb','30kb','50kb','100kb','500kb'), ordered=T),
-  cohort = factor(cohort, levels=c('train','val','val 30kb','val 50kb','val 100kb','val 500kb'), ordered=T)
-  ) %>% arrange(seg.size)
-
-ggplot(merged, aes(seg.size, sd, group=seg.size, fill=seg.size)) + facet_grid(~cohort, scales='free_x') + ylim(0,1) +
-  geom_jitter(width=0.2) + geom_boxplot(outlier.colour = NA) + 
-  scale_fill_brewer(palette = 'Set1') + theme_minimal()
+ggplot(all %>% filter(kb == '100kb'), aes(stdev)) + facet_grid(~cohort) + 
+  geom_density(fill='lightblue',kernel='gaussian') + xlim(0,1)
 
@@
 
-So it appears that I can decrease the variance by increasing the bin size.
-But it doesn't necessarily make sense to do it across the board.
-I can take the median of the training data and call that the cutoff for deciding.
+## Post segmentation residuals
 
-```{r}
-ggplot(merged, aes(x=1:nrow(merged),y=sd)) + ylim(0,1.0) + 
-  geom_point(aes(color=cohort)) + geom_hline(yintercept = median(training.raw$sd), linetype='dashed') +
-  geom_hline(yintercept = median(training.raw$sd)+0.5*mad(all.val15$sd), color='grey39') +
-  theme(axis.text.x = element_blank()) +
-  theme_minimal()
-
-out.samples = all.val15 %>% filter(sd > median(training.raw$sd))# +0.5*mad(all.val15$sd)))
-
-```
-
-In the validation dataset, starting at 15kb `r nrow(out.samples)`:`r nrow(all.val15)` samples have a variance greater than the median.
-
-```{r, echo=F}
-preds15 = do.call(bind_rows, lapply(list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf/predictions_model_5e6/0.9','predictions',full.names=T, recursive=T), function(f) read_tsv(f,col_types='cddcccDcccc')))
-
-#preds15 %>% filter(Sample %in% out.samples$sample) %>% group_by(Risk) %>% tally
-```
-
-## Re-process dynamically
-
-I re-processed all of the validation data with this in mind.
-I start with the 15kb bin size, immediately following the QDNAseq processing I check the variance of the raw data.
-If it's greater than the median of the training data (`r median(training.raw$sd)`) then I use the next greater bin size for that sample.
-I do this up to a bin size of 100kb currently.
-
-```{r, echo=F}
-library(BarrettsProgressionRisk)
-
-analysis.dir = '~/Data/BarrettsProgressionRisk/Analysis/validation/resegment/'
-
-per.sample.var = do.call(bind_rows, lapply(list.files(analysis.dir, 'variance.tsv', recursive=T, full.names=T), function(f) {
-  read_tsv(f, col_types='cddd') %>% mutate(patient = basename(sub('plots','',dirname(f))))
-})) %>% dplyr::select(patient, Samples, everything())
-
-max.bin<-function(bin15,bin50,bin100) {
-  if (!is.na(bin100))
-    return(100)
-  if (!is.na(bin50))
-    return(50)
-  return(15)
-}
-
-per.sample.bin = per.sample.var %>% group_by(Samples) %>% dplyr::summarise( max.bin = max.bin(`15kb`,`50kb`,`100kb`) ) %>% arrange(max.bin)
-
-per.sample.bin %>% group_by(max.bin) %>% tally() %>% spread(max.bin,n) %>% kable(caption = 'Number of samples per QDNAseq bin size') %>% kable_styling('striped', full_width = F)
-```
+The variance continues (as before) and is reflected in the residuals of the segmented values for each bin size
 
@@
 
-per.sample.bin = val.info %>% dplyr::select(matches('Hospital|Status|Block|Pathology|Samplename')) %>% 
-  left_join(per.sample.bin, by=c('Samplename'='Samples'))
+plist = lapply(train.resids, function(df) {
+  lim = c(0, max(df$varMAD_median))
+  if (max(df$varMAD_median) > 0.05) lim = c(min(df$varMAD_median), 0.05)
+  ggplot(df, aes(varMAD_median)) + geom_histogram(color='lightblue') + labs(title=unique(df$kb)) + xlim(lim)
+})
+do.call(grid.arrange, c(plist))
 
-per.sample.bin %>% group_by(max.bin, Status) %>% tally()
+all.resids = bind_rows(
+  do.call(bind_rows, train.resids) %>% mutate(cohort='training'),
+  do.call(bind_rows, val.resids) %>% mutate(cohort='validation') )
 
-per.sample.bin %>% filter(is.na(max.bin))
+m = melt(all.resids, id.vars=c('kb','cohort'), measure.vars = 'varMAD_median')
+ggplot(m %>% filter(value < 0.1), aes(cohort, value)) + facet_grid(~kb) + 
+  geom_jitter(width=0.2, alpha=0.5) + geom_violin(aes(fill=cohort), alpha=0.5, show.legend=F) + 
+  labs(title='Post-segmentation MAD residuals', x='median MAD variance')
 
+cutoffs = list('100kb' = 0.009, '15kb' = 0.011, '500kb' = 0.005, '50kb' = 0.011)
+
+sapply(names(train.resids), function(kb) nrow( train.resids[[kb]] %>% filter(varMAD_median >= cutoffs[kb])))
+sapply(names(val.resids), function(kb) nrow( val.resids[[kb]] %>% filter(varMAD_median >= cutoffs[kb])))
+
+ggplot(do.call(bind_rows,train.resids), aes(kb, varMAD_median)) + geom_boxplot() + ylim(0,0.05)
 ```

### 2019-10-25 processing_pipeline/process_val_data2.R

@@ -8,6 +8,7 @@ args = commandArgs(trailingOnly=TRUE)
 if (length(args) < 3)
   stop("Missing required arguments: <qdna data dir> <patient spreadsheet> <output dir> <patient name OPT>")
 
+library(tidyverse)
 suppressPackageStartupMessages( library(BarrettsProgressionRisk) )
 #source('~/workspace/shallowwgs_pipeline/lib/load_patient_metadata.R')
 

### 2019-10-29 QDNAseq segmentation.Rmd
...

@@
 
+train.tiles = purrr::map( list.files(qdna.dirs, full.names=T), function(dir) {
+  do.call(bind_rows, lapply(list.files(dir, '5e06_tile', recursive = T, full.names = T), function(f) {
+    read_tsv(f, col_types = cols( .default = col_double(), X1 = col_character())) %>% dplyr::rename(sample = 'X1') 
+  }))
+})
+names(train.tiles) = basename(list.files(qdna.dirs, full.names=T))
+
 val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
-sheets = readxl::excel_sheets(val.file)[8:13]
+sheets = readxl::excel_sheets(val.file)[8:14]
 
@@
 qdna.dirs = '~/Data/BarrettsProgressionRisk/Analysis/validation/pcf_perPatient'
-val.train.dists.by.kb = purrr::map( list.files(qdna.dirs, full.names=T), function(dir) {
+val.dists.by.kb = purrr::map( list.files(qdna.dirs, full.names=T), function(dir) {
   do.call(bind_rows, lapply(list.files(dir, 'raw_dist.tsv', full.names = T, recursive = T), function(f) {
@@
 })
-names(val.train.dists.by.kb) = basename(list.files(qdna.dirs, full.names=T))
+names(val.dists.by.kb) = basename(list.files(qdna.dirs, full.names=T))
+
+val.tiles = purrr::map( list.files(qdna.dirs, full.names=T), function(dir) {
+  do.call(bind_rows, lapply(list.files(dir, '5e06_tile', recursive = T, full.names = T), function(f) {
+    read_tsv(f, col_types = cols( .default = col_double(), X1 = col_character())) %>% dplyr::rename(sample = 'X1') 
+  }))
+})
+names(val.tiles) = basename(list.files(qdna.dirs, full.names=T))
 
@@
 ```{r}
-
 plist = lapply(train.resids, function(df) {
@@
 ```
+
+## Post segmentation 5mb bins
+
+```{r}
+train.tile.var = do.call(bind_rows, lapply(train.tiles, function(df) {
+  df %>% dplyr::summarise_if(is.double, list( ~ sd(.,na.rm=T) ))  
+})) %>% add_column(kb  = names(train.tiles), cohort = 'training', .before=1)
+
+val.tile.var = do.call(bind_rows, lapply(val.tiles, function(df) {
+  df %>% dplyr::summarise_if(is.double, list( ~ sd(.,na.rm=T) ))  
+})) %>% add_column(kb  = names(val.tiles), cohort = 'validation', .before=1)
+
+m = melt(bind_rows(train.tile.var, val.tile.var), id.vars = c('kb','cohort'))
+
+ggplot(m, aes(variable, value, color=cohort, group=cohort)) + ylim(0.95,1.05) + facet_grid(~kb) +  geom_jitter(position = position_dodge(width=1)) + geom_violin(position = position_dodge(width=1), alpha=0.5)
+
+```
+
+### Post z-adjustment
+
+```{r}
+z.sds = lapply(train.tiles, function(df) {
+  df %>% dplyr::summarise_if(is.double, list( ~ sd(.,na.rm=T) ))  
+})
+
+z.means = lapply(train.tiles, function(df) {
+  df %>% dplyr::summarise_if(is.double, list( ~ mean(.,na.rm=T) ))  
+})
+
+scaled.training = lapply(train.tiles, function(df) {
+  df %>% dplyr::mutate_if(is.double, list(~ BarrettsProgressionRisk:::unit.var(.) ) )
+})
+
+scaled.val = lapply(names(val.tiles), function(kb) {
+  df = val.tiles[[kb]]
+  zm = z.means[[kb]]
+  zd = z.sds[[kb]]
+  for (col in colnames((df %>% dplyr::select(-sample))))
+       df[[col]] = BarrettsProgressionRisk:::unit.var( df[[col]], zm[[col]], zd[[col]]  )
+  return(df)       
+})
+names(scaled.val) = names(val.tiles)
+
+scale.train.tile.var = do.call(bind_rows, lapply(scaled.training, function(df) {
+  df %>% dplyr::summarise_if(is.double, list( ~ median(.,na.rm=T) ))  
+})) %>% add_column(kb  = names(scaled.training), cohort = 'training', .before=1)
+
+scale.val.tile.var = do.call(bind_rows, lapply(scaled.val, function(df) {
+  df %>% dplyr::summarise_if(is.double, list( ~ median(.,na.rm=T) ))  
+})) %>% add_column(kb  = names(scaled.val), cohort = 'validation', .before=1)
+
+m = melt(bind_rows(scale.train.tile.var, scale.val.tile.var), id.vars = c('kb','cohort'))
+
+ggplot(m, aes(cohort, value, color=cohort, group=cohort)) + facet_grid(~kb) +  
+  geom_jitter(position = position_dodge(width=1)) + 
+  geom_violin(position = position_dodge(width=1), alpha=0.5)
+
+```

### 2019-10-29 processing_pipeline/process_val_data2.R

@@ -18,17 +18,19 @@ outdir = args[3]
 # data = '~/Data/Ellie/QDNAseq/validation/merged/100kb'
 # patient.file = '~/Data/Ellie/QDNAseq/validation/sWGS_validation_batches.xlsx'
 # outdir = '~/Data/Ellie/Analysis/validation'
-
+# patient.name = 'AHM1990'
 
 patient.name = NULL
 if (length(args) == 4)
   patient.name = gsub('\\/', '_', args[4])
 
-sheets = readxl::excel_sheets(patient.file)[8:13]
+sheets = readxl::excel_sheets(patient.file)[8:14]
 all.patient.info = do.call(bind_rows, lapply(sheets, function(s) {
-  readxl::read_xlsx(patient.file, s) %>% dplyr::select(`Hospital Research ID`, matches('Status|Endoscopy|Block|Path'), `Sample Type`, `SLX-ID`, `Index Sequence`) %>% 
+  readxl::read_xlsx(patient.file, s) %>% dplyr::select(`Hospital Research ID`, matches('Status|Endoscopy|Block|Path'), `Sample Type`, `SLX-ID`, `Index Sequence`,Cohort) %>% 
     dplyr::mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% dplyr::filter(!is.na(`SLX-ID`))
 })) %>% dplyr::rename('SLX.ID'='SLX-ID', 'Hospital.Research.ID'='Hospital Research ID', 'Block'='Block ID')
+
+#all.patient.info =all.patient.info %>% filter(Cohort != 'Old progressor scrolls')
 
 pastefun<-function(x) {
   if ( !grepl('SLX-', x) ) x = paste0('SLX-',x)
@@ -64,7 +66,7 @@ for (pt in unique(pts_slx$`Hospital Research ID`)) {
   print(pt)
   pd = paste(multipcfdir, pt,sep='/')
   
-  if (!file.exists(paste0(pd,'/segment.Rdata'))) {
+#  if (!file.exists(paste0(pd,'/segment.Rdata'))) {
     dir.create(pd,showWarnings = F)
     
     info = all.patient.info %>% filter(`Hospital.Research.ID` == pt) %>% 
@@ -74,7 +76,7 @@ for (pt in unique(pts_slx$`Hospital Research ID`)) {
   
     cols = which(colnames(merged.fit) %in% info$Sample)
     
[…]

### 2019-10-30 QDNAseq segmentation.Rmd
updating to process downsampled

@@
 m = m %>% mutate(kb = factor(kb, levels=c('15kb','50kb','100kb','500kb'), ordered=T))
-ggplot(m, aes(kb, value, group=kb)) + geom_boxplot(aes(fill=kb), show.legend = F) + ylim(0.95,1.05) + 
+ggplot(m, aes(kb, value, group=kb)) + ylim(0.95,1.05) + 
+  geom_jitter(width=0.1) + 
+  geom_violin(aes(fill=kb), show.legend = F) +
+  geom_boxplot(width=0.1,alpha=0.5) +
   scale_fill_brewer(palette='Set1') + 
@@
 m = m %>% mutate(kb = factor(kb, levels=c('15kb','50kb','100kb','500kb'), ordered=T))
-ggplot(m, aes(kb, value, group=kb)) + geom_boxplot(aes(fill=kb), show.legend = F) + ylim(0,1) +
+ggplot(m, aes(kb, value, group=kb)) + ylim(0,1) +
+  geom_jitter(width=0.1) + 
+  geom_violin(aes(fill=kb), show.legend = F) +
+  geom_boxplot(width=0.1,alpha=0.5) +
   scale_fill_brewer(palette='Set1') + 
@@
 m = m %>% mutate(kb = factor(kb, levels=c('15kb','50kb','100kb','500kb'), ordered=T))
-ggplot(m, aes(kb, value, group=kb)) + geom_boxplot(aes(fill=kb), show.legend = F) + ylim(0,1) +
+ggplot(m, aes(kb, value, group=kb)) +  ylim(0,1) +
+  geom_jitter(width=0.1) + 
+  geom_violin(aes(fill=kb), show.legend = F) +
+  geom_boxplot(width=0.1,alpha=0.5) +
   scale_fill_brewer(palette='Set1') + 
@@
 
-```{r}
+```{r, fig.width=10, fig.height=8}
 all = bind_rows(do.call(bind_rows, train.dists.by.kb) %>% mutate(cohort='training'),
@@
 
+m = melt(all, id.vars=c('kb','cohort'), measure.vars='median')
+ggplot(m, aes(cohort, value, group=cohort)) + facet_grid(~kb) + ylim(0.95,1.1) +
+  geom_jitter(width=0.1) + 
+  geom_violin(aes(fill=kb), show.legend = F) +
+  geom_boxplot(width=0.1,alpha=0.5) +
+  scale_fill_brewer(palette='Set1') + labs(title='Median per QDNAseq bin size', y='median',x='') + theme_minimal()
+
 m = melt(all, id.vars=c('kb','cohort'), measure.vars='stdev')
 ggplot(m, aes(cohort, value, group=cohort)) + facet_grid(~kb) + ylim(0,1) +
-  geom_jitter(width=0.2,alpha=0.5) + geom_violin(aes(fill=kb), alpha=0.5, show.legend=F) + 
-  #geom_boxplot(aes(fill=kb), outlier.color=NA, show.legend = F) 
+  geom_jitter(width=0.1) + 
+  geom_violin(aes(fill=kb), show.legend = F) +
+  geom_boxplot(width=0.1,alpha=0.5) +
   scale_fill_brewer(palette='Set1') + labs(title='Data variance per QDNAseq bin size', y='stdev',x='') + theme_minimal()
 
-m = melt(all, id.vars=c('kb','cohort'), measure.vars='median')
-ggplot(m, aes(cohort, value, group=cohort)) + facet_grid(~kb) + 
-  geom_jitter(width=0.2,alpha=0.5) + geom_violin(aes(fill=kb), alpha=0.5, show.legend=F) + ylim(0.95,1.1) +
-  scale_fill_brewer(palette='Set1') + labs(title='Median per QDNAseq bin size', y='median',x='') + theme_minimal()
-
-ggplot(all %>% filter(kb == '100kb'), aes(stdev)) + facet_grid(~cohort) + 
-  geom_density(fill='lightblue',kernel='gaussian') + xlim(0,1)
+#ggplot(all %>% filter(kb == '100kb'), aes(stdev)) + facet_grid(~cohort) + 
+#  geom_density(fill='lightblue',kernel='gaussian') + xlim(0,1)
 
@@
   if (max(df$varMAD_median) > 0.05) lim = c(min(df$varMAD_median), 0.05)
-  ggplot(df, aes(varMAD_median)) + geom_histogram(color='lightblue') + labs(title=unique(df$kb)) + xlim(lim)
+  ggplot(df, aes(varMAD_median)) + geom_histogram(color='lightblue') + labs(title=unique(df$kb)) + xlim(lim) + theme_minimal()
 })
@@
 ggplot(m %>% filter(value < 0.1), aes(cohort, value)) + facet_grid(~kb) + 
-  geom_jitter(width=0.2, alpha=0.5) + geom_violin(aes(fill=cohort), alpha=0.5, show.legend=F) + 
+  geom_jitter(width=0.1, color='grey50') + 
+  geom_violin(aes(fill=cohort), show.legend = F) +
+  geom_boxplot(width=0.1,alpha=0.5) +
+  theme_minimal() + 
   labs(title='Post-segmentation MAD residuals', x='median MAD variance')
@@
 cutoffs = list('100kb' = 0.009, '15kb' = 0.011, '500kb' = 0.005, '50kb' = 0.011)
-
 sapply(names(train.resids), function(kb) nrow( train.resids[[kb]] %>% filter(varMAD_median >= cutoffs[kb])))
@@
 
-ggplot(do.call(bind_rows,train.resids), aes(kb, varMAD_median)) + geom_boxplot() + ylim(0,0.05)
+#ggplot(do.call(bind_rows,train.resids), aes(kb, varMAD_median)) + geom_boxplot() + ylim(0,0.05)
 ```
@@
 
-ggplot(m, aes(variable, value, color=cohort, group=cohort)) + ylim(0.95,1.05) + facet_grid(~kb) +  geom_jitter(position = position_dodge(width=1)) + geom_violin(position = position_dodge(width=1), alpha=0.5)
+ggplot(m, aes(variable, value, color=cohort, group=cohort)) + ylim(0.95,1.05) + facet_grid(~kb) +  
+  geom_jitter(position = position_dodge(width=1)) + 
+  geom_violin(position = position_dodge(width=1), alpha=0.5) + 
+  theme_minimal()
 
@@
 
-```{r}
+```{r, warning=F, message=F}
 z.sds = lapply(train.tiles, function(df) {

### 2019-11-05 sWGS QC.Rmd
...

@@
 
-dir1 = '~/Data/Ellie/Analysis/multipcf_perPatient/'
-dir2 = '~/Data/Ellie/Analysis/validation//'
+dir1 = '~/Data/BarrettsProgressionRisk/Analysis/pcf_perPatient/50kb/'
+dir2 = '~/Data/BarrettsProgressionRisk/Analysis/validation/pcf_perPatient/50kb/'
 
@@
 res.variance = do.call(bind_rows, lapply(list.files(dir1, 'residuals', recursive = T, full.names = T), function(f) {
-  read_tsv(f, col_types = 'ccdddddl')
+  read_tsv(f, col_types = 'ccddddl')
 }))
@@
 ```{r, echo=F}
-res.variance %>% filter( round(varMAD_median,3) > cutoff ) %>% select(patient,sample, varMAD_median) %>% kable(caption=paste0('Samples variance > ',cutoff)) %>% kable_styling('striped')
+res.variance %>% filter( round(varMAD_median,3) > cutoff ) %>% dplyr::select(patient,samplename,varMAD_median) %>% 
+  kable(caption=paste0('Samples variance > ',cutoff)) %>% kable_styling('striped')
 
@@
 all.val = do.call(bind_rows, lapply(sheets, function(s) {
-  readxl::read_xlsx(val.file, s) %>% select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA, matches('Collection')) %>% dplyr::filter(!is.na(`SLX-ID`) & RA == 'C.Kosmidou') %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character))
+  readxl::read_xlsx(val.file, s) %>% dplyr::select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA, matches('Collection')) %>% 
+    dplyr::filter(!is.na(`SLX-ID`) & RA == 'C.Kosmidou') %>% 
+    mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) 
 }))
-all.val = all.val %>% select(-Collection)
+all.val = all.val %>% dplyr::select(-Collection)
 
@@
 resVal = do.call(bind_rows, lapply(list.files(dir2, 'residuals', recursive = T, full.names = T), function(f) {
-  readr::read_tsv(f, col_types = 'cdddddl', col_names = F) %>% set_names('sample', 'varMAD_median', 'varMAD_sd', 'varMAD_Q1', 'varMAD_Q3', 'n.segs', 'Pass') %>%  
-    mutate(`Patient ID` = as.character(basename(dirname(f))))
-})) %>% select(`Patient ID`, sample, matches('varMad'), n.segs, Pass)
+  readr::read_tsv(f, col_types = 'ccddddl') #%>% set_names('sample', 'varMAD_median', 'varMAD_sd', 'varMAD_Q1', 'varMAD_Q3', 'n.segs', 'Pass') %>%  
+   # mutate(`Patient ID` = as.character(basename(dirname(f))))
+})) %>% dplyr::select(patient, samplename, matches('varMad'), Pass)
 
-resVal = resVal %>% filter(sample %in% all.val$Sample)
+resVal = resVal %>% filter(samplename %in% all.val$Sample)
 

### 2019-11-05 Paper.Rmd

@@
 ```{r labelsPNP, echo=F, message=F, include=T}
-cache.dir = paste(data, 'Analysis/5e6_arms', sep='/') # QC'd
+cache.dir = paste(data, 'Analysis/models_5e6/50kb', sep='/') # QC'd
 
@@
 
-chr.lengths = BarrettsProgressionRisk:::chrInfo(prefix = '') %>% dplyr::filter(chr %in% c(1:22)) %>%
+chr.lengths = BarrettsProgressionRisk:::chrInfo() %>% dplyr::filter(chr %in% c(1:22)) %>%
   dplyr::mutate(chrom = factor(sub('chr','', chrom), levels=c(1:22), ordered=T))
@@
 
-mp$performance[,c('n.Coef','25%','50%','75%', '100%')] %>% kable(caption='Number of features stable in >=n% of folds') %>% kable_styling()
+mp$performance[,c('n.Coef','25%','50%','75%', '100%')] %>% kable(caption='Number of features stable in >=n% of folds') %>% kable_styling(full_width = F)
 ```

### 2019-11-12 Compare Models.Rmd
regenerating figures for final paper

---
title: "Model Comparison"
author: "Sarah Killcoyne"
date: "2 October 2017"
output: 
  html_document:
    fig_height: 5
    fig_width: 5
    toc: yes
    toc_depth: 4

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, error = T)

library(tidyverse)
library(BarrettsProgressionRisk)

library(ggrepel)
#library(GGally)
library(kableExtra)
library(reshape2)
library(gridExtra)
#library(CoxHD)
library(RColorBrewer)

library(OACUtils)

source('lib/load_patient_metadata.R')
#source('lib/cv-pt-glm.R')
source('lib/common_plots.R')

data = '~/Data/BarrettsProgressionRisk/Analysis/models_5e6_all/'
info.dir = '~/Data/BarrettsProgressionRisk/QDNAseq'

patient.file = list.files(info.dir, pattern='All_patient_info.xlsx', recursive=T, full.names=T)
demo.file = list.files(info.dir, pattern='Demographics_full.xlsx', recursive=T, full.names=T)

if (length(patient.file) != 1 | length(demo.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", info.dir))

all.patient.info = read.patient.info(patient.file, demo.file, set='All')$info

patient.info = all.patient.info %>% dplyr::filter(Set == 'Training')

patient.info = arrange(patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
sum.patient.data = as_tibble(summarise.patient.info(patient.info))

```

# Model performance

Per alpha value and kb segment size the overall performance per cross-validated model.  While 15kb is consistently the best at alpha = 0.9 the performance for 50kb or 100kb is about 66% for both.

```{r, fig.width=8,fig.height=7}
kb.models = list.files(data)

#list.files(f,full.names = T)

kb.perf = do.call(bind_rows, lapply(list.files(data,full.names = T), function(f) {
  load(paste0(f,'/all.pt.alpha.Rdata'),verbose=F)

  nonz = names(which(sapply(performance.at.1se, nrow) > 0))
# do.call(bind_rows,lapply(coefs, function(df){
#   tibble(n=nrow(df), mean.stability=mean(rowSums(df[,-1])/50), sd.stability=sd(rowSums(df[,-1])/50) )
# })) %>% mutate(alpha=names(coefs), kb=basename(f))

  do.call(bind_rows, performance.at.1se) %>% 
    mutate(alpha=nonz, kb = basename(f), n.coefs=sapply(coefs[nonz], nrow)) %>% 
    dplyr::select(mean, sme, `lambda-1se`, alpha, n.coefs, kb)

}))
kb.perf = kb.perf %>% mutate(kb = factor(kb, levels=c('15kb','50kb','100kb','500kb'), ordered=T)) %>% arrange(kb, alpha)

ggplot(kb.perf, aes(kb, mean, fill=kb)) + facet_grid(~alpha) + geom_col() + 
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme), col='grey39', width=0.2) + 
  geom_text(aes(label=n.coefs, y = 0.01)) + scale_fill_brewer(palette='Set1') + labs(x='',y='cv performance') +
  theme_minimal()

ggplot(kb.perf, aes(alpha, mean, fill=kb)) + facet_grid(~kb) + geom_col() + 
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme), col='grey39', width=0.2) + 
  scale_fill_brewer(palette='Set1') + labs(x='',y='cv performance') +
  geom_text(aes(label=n.coefs, y = 0.01)) + 
  theme_minimal()

```

# ROC

## LOO Training

Finally, I generate a ROC for each of the leave-one out prediction sets. Here again the AUC is best at 15kb.  However, 100kb is the best of the larger bin sizes.  The tradeoff here is that the TPR is not as good as 15kb or 50kb (given the threshold selected by the ROC).  The FPR is the highest in the 50kb bins.

```{r fig.height=7, fig.width=8}
rc9 = do.call(bind_rows, lapply(list.files(data,full.names = T), function(f) {
  load(paste0(f,'/loo_0.9.Rdata'),verbose=F)
  roc = pROC::roc(Status ~  Prediction, data=pg.samp, auc=T, ci=T, of='thresholds',transpose=T, quiet=T)
  pROC::coords(roc, "best", transpose=F) %>% mutate( auc = as.numeric(pROC::auc(roc)), kb = basename(f) )
}))

rc9 = rc9 %>% set_names(c('Threshold','FPR','TPR','AUC','kb')) %>% 
  mutate(kb = factor(kb,levels=c('15kb','50kb','100kb','500kb'), ordered=T), FPR = 1-FPR) %>% arrange(kb)

ggplot(melt(rc9, id.vars='kb'), aes(variable, value, group=kb)) + 
  geom_bar(aes(fill=kb), stat='identity', position='dodge') + 
  geom_text(aes(label=round(value,2)), position=position_dodge(width=1)) + 
  scale_fill_brewer(palette='Set1') + 
  labs(title='Training ROC (alpha=0.9)', x='', y='%') + theme_minimal()
```

## Validation predicted on corresponding model

```{r fig.height=7, fig.width=8}
val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
sheets = readxl::excel_sheets(val.file)[8:14]

pastefun<-function(x) {
  if ( !grepl('SLX-', x) ) x = paste0('SLX-',x)
  return(x)
}

val.info = do.call(bind_rows, lapply(sheets, function(s) {
  readxl::read_xlsx(val.file, s) %>% dplyr::select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Pathology, Cohort, Batch, matches('Collection|Notes')) %>% 
    dplyr::filter(!is.na(`SLX-ID`)) %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) 
})) %>% mutate(Samplename = paste(`SLX-ID`,`Index Sequence`, sep='.')) %>% 
  rowwise %>% mutate_at(vars(`SLX-ID`), list(pastefun) ) %>% ungroup %>% 
  mutate(
  #`Hospital Research ID` = str_replace_all( str_remove_all(`Hospital Research ID`, " "), '/', '_'), 
  `Index Sequence` = str_replace_all(`Index Sequence`, 'tp', ''),
  Samplename = paste(`SLX-ID`,`Index Sequence`,sep='.')
  )

vdir = '~/Data/BarrettsProgressionRisk/Analysis/validation/pcf_perPatient'

val.preds = purrr::map(list.files(vdir, full.names=T), function(dir) {
  do.call(bind_rows, purrr::map(list.files(paste0(dir,'/predictions_5e6_all/0.9'), 'predictions.tsv', full.names = T, recursive = T), function(f) {
    pd = read_tsv(f, col_types = 'cddcccDcccc') %>% dplyr::select(`Hospital Research ID`, everything())
    left_join(pd, val.info %>% dplyr::select(`Hospital Research ID`, Status, Samplename, Cohort), by=c('Hospital Research ID', 'Sample'='Samplename')) %>% 
      mutate(kb = basename(dir)) %>% mutate(Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered=T)) 
  }))
})
names(val.preds) = list.files(vdir)

val.auc = do.call(bind_rows, purrr::map(val.preds, function(kb) {
  roc = pROC::roc(Status ~ Probability, data=kb, auc=T, ci=T, of='thresholds',transpose=T, quiet=T)
  pROC::coords(roc, "best", transpose=F) %>% mutate( auc = as.numeric(pROC::auc(roc)), kb = unique(kb$kb) )
}))

val.auc = val.auc %>% set_names(c('Threshold','FPR','TPR','AUC','kb')) %>% 
  mutate(kb = factor(kb,levels=c('15kb','50kb','100kb','500kb'), ordered=T), FPR = 1-FPR) %>% arrange(kb)

ggplot(melt(val.auc, id.vars='kb'), aes(variable, value, group=kb)) + 
  geom_bar(aes(fill=kb), stat='identity', position='dodge') + 
  geom_text(aes(label=round(value,2)), position=position_dodge(width=1)) + 
  scale_fill_brewer(palette='Set1') + 
  labs(title='Validation ROC (alpha=0.9)', x='', y='%') + theme_minimal()

```

## Side-by-side

```{r fig.height=7, fig.width=8}
aucs = bind_rows(rc9 %>% mutate(cohort = 'Training'), val.auc %>% mutate(cohort = 'Validation'))

ggplot(aucs, aes(kb, AUC, group=cohort)) + 
  geom_bar(aes(fill=cohort), stat='identity', position='dodge') + 
  geom_text(aes(label=round(AUC,2)), position=position_dodge(width=1)) + 
  scale_fill_brewer(palette='Set1') + 
  labs(title='AUC training & validation', x='', y='AUC') + theme_minimal()
```

# Training LOO Performance

The leave-one out analysis uses the models for alpha=0.9.

```{r, echo=F}
minR = 0.3
maxR = 0.5
```

Cutoffs are  `r minR` < P > `r maxR`

```{r, fig.width=8, fig.height=12}
ft.fun<-function(NP,P) {
 f = chisq.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
 cbind.data.frame('p.value'=round(f$p.value, 4))
}

pred.conf<-function(df) {
  df = df %>% mutate(quants = cut(Prediction, breaks=seq(0,1,0.1), include.lowest = T))
  qts = df %>% group_by(quants, Status) %>% dplyr::summarise(n=length(Status) ) %>% ungroup %>%
    spread(Status, n) %>%  mutate_all(~replace_na(.,0)) %>% 
    left_join(df %>% dplyr::group_by(quants) %>% 
                dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ), by='quants')

  pred.confidence = qts %>% dplyr::group_by(quants) %>% 
    dplyr::mutate( 'P.ratio'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'conf'=ifelse(p.value < 0.05, '*', '') ) %>% 
    separate(quants, c('r1','r2'), ',', remove = F) %>% 
    mutate_at(vars('r1','r2'), list(sub), pattern='\\[|\\]|\\(', replacement='') %>% 
    mutate_at(vars(r1,r2), as.double)

  # set cutoffs
  pred.confidence = pred.confidence %>% ungroup %>% rowwise %>% mutate(Risk = 'Moderate') %>%
    mutate(Risk = ifelse(r2 <= minR, 'Low',Risk)) %>% mutate(Risk = ifelse(r1 >= maxR, 'High',Risk))

  pred.confidence = bind_cols(pred.confidence, 
                            data.frame(ci.low=qbeta(0.025, shape1=pred.confidence$P+.5, shape2 = pred.confidence$NP+.5),
                                         ci.high=qbeta(0.975, shape1=pred.confidence$P+.5, shape2 = pred.confidence$NP+.5)))
  return(pred.confidence)
}
```

```{r, fig.width=10,fig.height=7, eval=F}
## 0.5
loo.pred.confidence = purrr::map( list.files(data,full.names = T), function(f) {
  #print(f)
  load(paste0(f,'/loo_0.5.Rdata'),verbose=F)
  suppressWarnings(pred.conf(pg.samp)) %>% mutate(kb = basename(f))
}) 
names(loo.pred.confidence) = basename(list.files(data,full.names = T))

plist = lapply(loo.pred.confidence, function(lpf) {
  BarrettsProgressionRisk::showPredictionCalibration(lpf %>% dplyr::rename('perc' = 'P.ratio') ) + labs(title=unique(lpf$kb))
})

do.call(grid.arrange, c(plist, top='Alpha=0.5'))
```

```{r, fig.width=10,fig.height=7, eval=F}
## 0.8

loo.pred.confidence = purrr::map( list.files(data,full.names = T), function(f) {
  #print(f)
  load(paste0(f,'/loo_0.8.Rdata'),verbose=F)
  suppressWarnings(pred.conf(pg.samp)) %>% mutate(kb = basename(f))
}) 
names(loo.pred.confidence) = basename(list.files(data,full.names = T))

plist = lapply(loo.pred.confidence, function(lpf) {
  BarrettsProgressionRisk::showPredictionCalibration(lpf %>% dplyr::rename('perc' = 'P.ratio') ) + labs(title=unique(lpf$kb))
})

do.call(grid.arrange, c(plist, top='Alpha=0.8'))
```

## Training LOO alpha=0.9

```{r, fig.width=8, fig.height=12}
## ALL LOO is at alpha=0.9
loo.pred.confidence = purrr::map( list.files(data,full.names = T), function(f) {
  #print(f)
  load(paste0(f,'/loo_0.9.Rdata'),verbose=F)
  suppressWarnings(pred.conf(pg.samp)) %>% mutate(kb = basename(f))
}) 
names(loo.pred.confidence) = list.files(data)

plist = lapply(loo.pred.confidence, function(lpf) {
  BarrettsProgressionRisk::showPredictionCalibration(lpf %>% dplyr::rename('perc' = 'P.ratio') ) + labs(title=unique(lpf$kb))
})

do.call(grid.arrange, c(plist, top='Training LOO alpha=0.9'))
```

## Per-sample alpha=0.9

```{r, fig.width=8, fig.height=12}

per.samp.train = lapply(list.files(data,full.names = T), function(f) {
  load(paste0(f,'/loo_0.9.Rdata'),verbose=F)

  kb = basename(f)

  pg.samp = pg.samp %>% mutate(quants = cut(Prediction, breaks=seq(0,1,0.1), include.lowest = T)) %>%
    left_join(loo.pred.confidence[[kb]] %>% dplyr::select(quants, Risk), by='quants') %>%
    mutate(Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered=T))

  pg.samp %>% group_by(Status, Risk) %>% tally %>% spread(Risk,n) %>% 
    left_join(pg.samp %>% group_by(Status) %>% tally, by='Status') %>%
    dplyr::rename('Total' = 'n') %>%
    mutate_at(vars(Low,Moderate,High), funs(./Total)) 
})
names(per.samp.train) = list.files(data)

plist = purrr::map(names(per.samp.train), function(kb) {
  df = per.samp.train[[kb]]

  ggplot(melt(df, id.vars=c('Status','Total')), aes(variable, value, group=Status, fill=Status)) + 
    ylim(0,1) + geom_col(position='dodge') + 
    geom_text(aes(label=round(value,2)), position = position_dodge(width=1)) + 
    labs(title=unique(kb), x='', y='ratio') + theme_minimal()
})
names(plist) = names(per.samp.train)

do.call(grid.arrange, c(plist[c('50kb','100kb','15kb','500kb')], top='Training predictions alpha=0.9'))

```

# Test Data

Using the alpha=0.9 model for each and evaluate the risks based on the LOO performance.

```{r fig.height=10,fig.width=10, eval=F}
## 0.5
test.preds.per.kb05 = purrr::map(list.files(data,full.names = T), function(f) {
  read_tsv(list.files(f, 'test_patients_preds_0.5.tsv', full.names = T), col_types = 'cdccdccclddccccccddccccdddccdddcdd') %>% dplyr::select(Hospital.Research.ID, Status, Endoscopy.Year, Pathology, Prediction, RR) %>% mutate(kb = basename(f))
}) 
names(test.preds.per.kb05) = basename(list.files(data))

test.preds.per.kb05 = lapply(test.preds.per.kb05, function(kb) {
  kb %>% mutate(quants = cut(Prediction, breaks=seq(0,1,0.1), include.lowest = T)) %>% left_join(loo.pred.confidence[[unique(kb$kb)]] %>% dplyr::select(quants, Risk), by='quants') %>%
      mutate(Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered=T)) 
})

test.samples = all.patient.info %>% filter(Set == 'Test') %>% group_by(Status) %>% tally

test.counts05 = lapply(test.preds.per.kb05, function(kb) {
  kb %>% group_by(Status, Risk) %>% tally %>% spread(Risk,n) %>% ungroup %>%
    mutate(kb = unique(kb$kb)) %>% left_join(test.samples, by='Status') %>% 
    mutate_if(is.numeric, funs(./n))
})

plist = lapply(test.counts05, function(kb) {
  tt = unique(kb$kb)
  kb = kb  %>% ungroup %>% dplyr::select(-kb,-n)

  ggplot(melt(kb, id.vars=c('Status')), aes(variable, value, group=Status, fill=Status)) + ylim(0,1) + 
    geom_bar(stat='identity', position='dodge') + geom_text(aes(label=round(value,2)), position = position_dodge(width=1)) +
    labs(title=tt, x='', y='ratio') + theme_minimal()
})

do.call(grid.arrange, c(plist, top='Alpha=0.5'))
```

## Per-sample alpha=0.9

```{r, fig.width=8, fig.height=12, eval=F}
test.preds.per.kb09 = purrr::map(list.files(data,full.names = T), function(f) {
  read_tsv(list.files(f, 'test_patients_preds_0.9.tsv', full.names = T), col_types = 'cdccdccclddccccccddccccdddccdddcdd') %>% dplyr::select(Hospital.Research.ID, Status, Endoscopy.Year, Pathology, Prediction, RR, Samplename, Path.ID) %>% mutate(kb = basename(f))
}) 
names(test.preds.per.kb09) = basename(list.files(data))

test.preds.per.kb09 = lapply(test.preds.per.kb09, function(kb) {
  kb %>% mutate(quants = cut(Prediction, breaks=seq(0,1,0.1), include.lowest = T)) %>% 
    left_join(loo.pred.confidence[[unique(kb$kb)]] %>% dplyr::select(quants, Risk), by='quants') %>%
    mutate(Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered=T)) 
})

test.samples = all.patient.info %>% filter(Set == 'Test') %>% group_by(Status) %>% tally

test.counts09 = lapply(test.preds.per.kb09, function(kb) {
  kb %>% group_by(Status, Risk) %>% tally %>% spread(Risk,n) %>% ungroup %>%
    mutate(kb = unique(kb$kb)) %>% left_join(test.samples, by='Status') %>% 
    mutate_if(is.numeric, funs(./n))
})

plist = lapply(test.counts09, function(kb) {
  tt = unique(kb$kb)
  kb = kb  %>% ungroup %>% dplyr::select(-kb,-n)

  ggplot(melt(kb, id.vars=c('Status')), aes(variable, value, group=Status, fill=Status)) + ylim(0,1) + 
    geom_bar(stat='identity', position='dodge') + geom_text(aes(label=round(value,2)), position = position_dodge(width=1)) +
    #scale_fill_manual(values=c('#4575B4','#A50026'),labels=c('NP','P')) + 
    labs(title=tt, x='', y='sample ratio') + theme_minimal()
})

do.call(grid.arrange, c(plist[c('50kb','100kb','15kb','500kb')], top='Test set predictions, alpha=0.9'))
```

## Per endoscopy
```{r, fig.width=8, fig.height=12, eval=F}

preds.per.endo = lapply(test.preds.per.kb09, function(kb) {
  endo = kb %>% separate(Path.ID, c('PID', 'Block'), '[:blank:]|_') %>% 
    group_by(Hospital.Research.ID, Status, Endoscopy.Year, PID, kb) %>% 
    mutate(Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered=T))

  total.endo = endo %>% ungroup %>% dplyr::select(PID, Status) %>% distinct %>% group_by(Status) %>% tally
  endo %>% dplyr::summarise(Max.Pred = max(Prediction), Risk = max(Risk), n.samples = length(Prediction)) %>%
    group_by(kb,Status, Risk) %>% tally %>% spread(Risk,n) %>% left_join(total.endo, by='Status') %>% ungroup %>%
    mutate_if(is.numeric, funs(./n)) %>% dplyr::select(-n)
})

plist = lapply(preds.per.endo, function(kb) {
  tt = unique(kb$kb)
  kb = kb  %>% ungroup %>% dplyr::select(-kb)

  ggplot(melt(kb, id.vars=c('Status')), aes(variable, value, group=Status, fill=Status)) + ylim(0,1) + 
    geom_bar(stat='identity', position='dodge') + geom_text(aes(label=round(value,2)), position = position_dodge(width=1)) +
    labs(title=tt, x='', y='sample ratio') + theme_minimal()
})

do.call(grid.arrange, c(plist[c('50kb','100kb','15kb','500kb')], top='Predictions per endoscopy'))

```

# Validation Data

## Per sample (alpha = 0.9)

```{r, fig.width=8, fig.height=12}
val.preds = lapply(val.preds, function(kb) {
  kb %>% dplyr::select(-Risk) %>% mutate(quants = cut(Probability, breaks=seq(0,1,0.1), include.lowest = T)) %>% 
    left_join(loo.pred.confidence[[unique(kb$kb)]] %>% dplyr::select(quants, Risk), by='quants') %>%
    mutate(Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered=T)) 
})

to = nrow(val.info %>% filter(Pathology == 'OAC'))
oac = do.call(bind_rows, val.preds) %>% filter(Pathology == 'OAC')

oac %>% group_by(kb, Risk) %>% tally %>% spread(Risk,n) %>% ungroup %>% mutate_all( funs(ifelse(is.na(.), 0, .)) ) %>% 
  mutate_if(is.numeric, funs(round(./to,2))) %>% mutate(kb = factor(kb, levels=c('15kb','50kb','100kb','500kb'), ordered=T)) %>% arrange(kb) %>%
  kable(caption='OAC predictions') %>% kable_styling(full_width=F)

#val.info %>% filter(Cohort != 'Old progressor scrolls' & Pathology != 'OAC') %>% group_by(Status) %>% tally

#  roc = pROC::roc(Status ~ Probability, data=val.preds$`100kb`, auc=T, ci=T, of='thresholds',transpose=T)
#  pROC::coords(roc, "best", transpose=F) %>% mutate( auc = as.numeric(pROC::auc(roc)), kb = basename(f) )

val.counts = lapply(val.preds, function(kb) {
  totals = kb %>% dplyr::filter(Cohort != 'Old progressor scrolls' & Pathology != 'OAC') %>% group_by(Status) %>% tally

  kb %>% dplyr::filter(Cohort != 'Old progressor scrolls' & Pathology != 'OAC')  %>% 
    group_by(Status, Risk) %>% tally %>% spread(Risk,n) %>% ungroup %>%
    mutate(kb = unique(kb$kb)) %>% left_join(totals, by='Status') %>% 
    mutate_if(is.numeric, funs(./n)) %>% dplyr::select(-n)
})

plist = lapply(val.counts, function(kb) {
  tt = unique(kb$kb)
  kb = kb  %>% ungroup %>% dplyr::select(-kb)

  ggplot(melt(kb, id.vars=c('Status')), aes(variable, value, group=Status, fill=Status)) + ylim(0,1) + 
    geom_bar(stat='identity', position='dodge') + geom_text(aes(label=round(value,2)), position = position_dodge(width=1)) +
    labs(title=tt, x='', y='ratio') + theme_minimal()
})

do.call(grid.arrange, c(plist[c('50kb','100kb','15kb','500kb')], top='Validation'))

```

## Per endoscopy

```{r, fig.width=8, fig.height=12}

val.endo = lapply(val.preds, function(kb) {
  endo = kb %>% dplyr::filter(Cohort != 'Old progressor scrolls' & Pathology != 'OAC') %>%
    separate(`Block ID`, c('PID','Block'), '[:blank:]|-|\\/', extra = 'drop')

  total.endo = endo %>% ungroup %>% dplyr::select(PID, Status) %>% distinct %>% group_by(Status) %>% tally

  endo %>% group_by(`Hospital Research ID`, `PID`, Status, kb) %>%
    dplyr::summarise(Max.Prob = max(Probability), Risk = max(Risk)) %>%
    group_by(kb,Status,Risk) %>% tally %>% spread(Risk,n) %>% left_join(total.endo, by='Status') %>% ungroup %>%
    mutate_all( funs( ifelse(is.na(.), 0, .) ) ) %>%
    mutate_if(is.numeric, funs(./n)) %>% dplyr::select(-n)
})

plist = lapply(val.endo, function(kb) {
  tt = unique(kb$kb)
  kb = kb  %>% ungroup %>% dplyr::select(-kb)

  ggplot(melt(kb, id.vars=c('Status')), aes(variable, value, group=Status, fill=Status)) + ylim(0,1) + 
    geom_bar(stat='identity', position='dodge') + geom_text(aes(label=round(value,2)), position = position_dodge(width=1)) +
    labs(title=tt, x='', y='ratio') + theme_minimal()
})

do.call(grid.arrange, c(plist[c('50kb','100kb','15kb','500kb')], top='Validation per Endoscopy'))

```

# Downsampled Data

## Per-sample alpha=0.9
```{r, fig.width=8, fig.height=12, eval=F}

ds.info = read_excel('~/Data/BarrettsProgressionRisk/QDNAseq/all_downsampled/downsampled_ids.xlsx')

ddir = '~/Data/BarrettsProgressionRisk/Analysis/downsampled/'

ds.preds =   purrr::map(list.files(ddir, recursive = F, full.names = T), function(dir) {
    read_tsv(list.files(paste0(dir,'/predictions/0.9'), 'predictions.tsv', full.names = T), col_types = cols(
  Sample = col_character(), Probability = col_double(), `Relative Risk` = col_double(),
  Risk = col_character(), PatientID = col_character(), Status = col_character(),
  Tissue = col_character(), `Expected Risk` = col_character(), Grade_patient = col_character(),
  Grade_biopsy = col_character(), Study = col_character(), Type = col_character(), Endoscopy = col_date(format = "")
)) %>% dplyr::select(-Risk, -Status) %>% mutate(quants = cut(Probability, breaks=seq(0,1,0.1), include.lowest = T)) %>% 
  left_join(loo.pred.confidence[['50kb']] %>% dplyr::select(quants, Risk), by='quants') %>%
    mutate(Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered=T), kb = basename(dir)) %>% left_join(ds.info %>% dplyr::select(`Illumina ID`, Status), by=c('Sample' = 'Illumina ID'))
})
names(ds.preds) = list.files(ddir)

ds.counts = lapply(ds.preds, function(kb) {
  totals = kb %>% group_by(Status) %>% tally

  kb %>% group_by(Status, Risk) %>% tally %>% spread(Risk,n) %>% ungroup %>%
    mutate(kb = unique(kb$kb)) %>% left_join(totals, by='Status') %>% 
    mutate_if(is.numeric, funs(./n)) %>% dplyr::select(-n) %>% 
    mutate_if(is.numeric, funs(ifelse(is.na(.), 0, .)))
})

plist = lapply(ds.counts, function(kb) {
  tt = unique(kb$kb)
  kb = kb  %>% ungroup %>% dplyr::select(-kb)

  ggplot(melt(kb, id.vars=c('Status')), aes(variable, value, group=Status, fill=Status)) + ylim(0,1) + 
    geom_bar(stat='identity', position='dodge') + geom_text(aes(label=round(value,2)), position = position_dodge(width=1)) +
    labs(title=tt, x='', y='ratio') + theme_minimal()
})

do.call(grid.arrange, c(plist[c('50kb','100kb','15kb','500kb')], top='Downsampled'))

```

```{r}

ds.preds$`50kb` %>% group_by(Type, Risk) %>% tally %>% spread(Risk,n) %>% mutate_if(is.numeric, funs(ifelse(is.na(.), 0, .))) %>% 
  kable(caption = 'Downsampled types') %>% kable_styling(full_width = F)

```

### 2019-11-12 Paper_resubmit.Rmd
regenerating figures for final paper

--- 2019-11-05 Paper.Rmd
@@
 ---
-title: "Paper"
+title: "Untitled"
 author: "Sarah Killcoyne"
-date: "2 October 2017"
+date: "07/11/2019"
 output: 
@@
     toc_depth: 4
-
 ---
@@
 ```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = F, message = F, warning = F)
 library(tidyverse)
@@
 source('lib/load_patient_metadata.R')
-#source('lib/cv-pt-glm.R')
 source('lib/common_plots.R')
 
-data = '~/Data/BarrettsProgressionRisk/'
-info.dir = '~/Data/BarrettsProgressionRisk/QDNAseq'
+dir = '~/Data/BarrettsProgressionRisk/'
+info.dir = paste0(dir,'/QDNAseq')
 
@@
 all.patient.info = read.patient.info(patient.file, demo.file, set='All')$info
-
 patient.info = all.patient.info %>% dplyr::filter(Set == 'Training')
@@
 
-# Segment size: 5e06 with arms
-
-## Quick look at our demographic information
-
-```{r echo=F}
+# Demographics
+
+```{r}
 
@@
 
-We tested models that included age, BMI, smoking status, sex, length & circumference of the BE segment, and p53 IHC status.  While none of these contributed we can report on some here.
-
-```{r echo=F, warning=F, message=F, fig.height=4, fig.width=4}
-ggplot(patient.info, aes(Age.at.diagnosis, fill=Status)) + geom_histogram(binwidth=3) + labs(title='Age at diagnosis') + plot.theme
-
-ggplot(patient.info, aes(BMI, fill=Status)) + geom_histogram(binwidth=1) + labs(title='BMI') + plot.theme
-
-ggplot(patient.info, aes(Maximal,Circumference, color=Status)) + geom_point() + labs(title='C vs M') + plot.theme
-
-p53Path = patient.info %>% dplyr::filter(Status == 'P' & p53.Status == 1) %>% dplyr::group_by(Pathology) %>% dplyr::summarise(n=length(Pathology)) %>% spread(Pathology,n)
-
-p53.pred.byPath = round(p53Path/table(subset(patient.info, Status == 'P' & !is.na(p53.Status))$Pathology), 2)
-
-```
-
 ```{r, echo=F}
@@
 
-Because we're looking at both a binary classifier and binary data the AUC looks very good, there are no false positives after all.
-However of the progressor samples that were stained only `r round((table(subset(patient.info, Status == 'P')$p53.Status)/sum(table(subset(patient.info, Status == 'P')$p53.Status)))[2], 2)*100`% positively stained for p53.
-If we look at pathology the breakdown is even more obvious.
-`r p53.pred.byPath['HGD']*100`% of HGD samples were identified, but only `r p53.pred.byPath['BE']*100`% NDBE samples were identified in progressors.
-
-`r p53.pred.byPath %>% kable(caption='Ratio of progressors predicted by p53 staining per pathology where p53 staining was done (excluding N/As)') %>% kable_styling(full_width=F)`
-
-# Data set up
+# Copy number data set up
 
@@
 ```{r labelsPNP, echo=F, message=F, include=T}
-cache.dir = paste(data, 'Analysis/models_5e6/50kb', sep='/') # QC'd
+cache.dir = paste(dir, 'Analysis/models_5e6_all/50kb', sep='/')
 
@@
 load(file, verbose=F)
-dim(dysplasia.df)
-length(labels)
+#dim(dysplasia.df)
+#length(labels)
 
@@
 ```{r echo=F, warning=F, message=F, eval=T, fig.height=12, fig.width=20}
-
 chr.lengths = BarrettsProgressionRisk:::chrInfo() %>% dplyr::filter(chr %in% c(1:22)) %>%
@@
 
-arm.loc = arrange(rbind(
+arm.loc = arrange(bind_rows(
   (chr.lengths %>% dplyr::group_by(chrom) %>% dplyr::mutate('start'=1, 'end'=chr.cent-cent.gap))[c('chrom','start','end')],
@@
 
-globalCNdf = t(dysplasia.df[,grep('^\\d+:',colnames(dysplasia.df))])
-
-locs = do.call(rbind.data.frame, lapply(rownames(globalCNdf), function(x) unlist(strsplit(x,':|-'))))
-colnames(locs) = c('chr','start','end')
-locs[] = lapply(locs[], function(x) as.numeric(as.character(x)))
-
-globalCNdf = cbind.data.frame(locs, globalCNdf)
-globalCNdf = melt(globalCNdf, id.vars=colnames(locs))
-
-globalCNdf$seg.type = '5MB'
-globalCNdf[with(globalCNdf, which( start %in% arm.loc$start & end %in% arm.loc$end)), 'seg.type'] = 'arms'
-globalCNdf = merge(globalCNdf, chr.lengths[,c('chrom','chr.length')], by.x='chr', by.y='chrom')
-
-globalCNdf = merge(globalCNdf, patient.info[,c('Hospital.Research.ID','Patient','Samplename','Pathology','Status')], by.x='variable', by.y='Samplename')
-globalCNdf$Status = ifelse(globalCNdf$Status == 'P', 'Progressor', 'Non-Progressor')
-
-globalCNdf$cn = factor(sapply(globalCNdf$value, cn), levels=c('loss','neutral','gain'), ordered = T)
-
-n.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor')$variable)), ')', sep=''),
-              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor')$variable)),')', sep=''))
-
-# Average cn value
-globalCNdf = globalCNdf %>% dplyr::group_by(Status, start, end) %>% dplyr::mutate( 'mean.value' = value/length(unique(variable)) )
-globalCNdf$mean.value = globalCNdf$value # but not going to use it
-
-globalCNdf$Status = factor(globalCNdf$Status, levels=sort(unique(globalCNdf$Status), decreasing=T), ordered=T)
-globalCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms'), labeller(Status=n.status, chr=label_value))
+#ddf = rbind(dysplasia.df, 'mean.value'=apply(dysplasia.df, 2, mean))
+globalCNdf = as_tibble( t(dysplasia.df[,grep('^\\d+:',colnames(dysplasia.df))]), rownames = 'locs') %>% 
+  separate(locs, c('chr','start','end'), sep=':|-', remove=F) %>% 
+  dplyr::mutate_at(vars(start, end), as.numeric) %>%
+  dplyr::mutate( seg.type = ifelse( (start %in% arm.loc$start & end %in% arm.loc$end), 'arms','5MB'  ) ) %>%
+  melt(id.vars=c('locs','chr','start','end', 'seg.type')) %>% as_tibble() %>%
+  left_join(chr.lengths %>% dplyr::select(chrom,chr.length), by=c('chr'='chrom')) %>%
+  left_join(patient.info %>% dplyr::select(matches('Hospital|Patient|Samplename|Pathology|Status')), by=c('variable'='Samplename')) %>%
+  dplyr::mutate(Status = factor(recode(Status, NP='Non-Progressor',P='Progressor'), ordered=T)) %>% 
+  dplyr::mutate( 
+    cn = factor(unlist(purrr::map(.x=value, .f=get.cn.state)),  levels=c('loss','neutral','gain'), ordered = T),
+    chr = factor(chr, levels = chr.lengths$chrom, ordered=T))
+
+n.status <- c('Progressor'=paste0('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor')$variable)), ')'),
+              'Non-Progressor'=paste0('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor')$variable)),')'))
+
+globalCN.plot = cn.mtn.plot(globalCNdf %>% filter(seg.type != 'arms'), labeller(Status=n.status, chr=label_value))
+
 cnLegend = get.legend(globalCN.plot)
-#globalCN.plot = 
 globalCN.plot + theme(legend.position = 'none')
@@
 ## NDBE
-nb.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology == 'BE')$variable)), ')', sep=''),
-              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology == 'BE')$variable)),')', sep=''))
+nb.status <- c('Progressor'=paste0('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology == 'BE')$variable)), ')'),
+              'Non-Progressor'=paste0('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology == 'BE')$variable)),')'))
+
 globalBECN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology == 'BE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')
 
-gp = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples')# + theme(text=element_text(size=12))
+gp = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples')
 ggsave(filename='plots/globalCN.png', plot=gp, width=16, height=4.6, units='in', dpi=300)
@@
 
+## *Complexity Scores*
+
 ```{r loaddata, message=F, warning=F, echo=F}
-
 file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
@@
   message(paste("loading", file))
-  load(file, verbose=T)
+  load(file, verbose=F)
 } else {
@@
 cx = dysplasia.df[,'cx']
-
-x = cbind.data.frame('Status'=labels, 'cx'=cx[names(labels)])
-x$Status = factor(x$Status, labels = c('Non-Progressor','Progressor'))
-x = merge(x,patient.info[,c('Samplename','Pathology','Patient')],by.x='row.names',by.y='Samplename')
-
-tt = t.test(subset(x, Status == 'Progressor')$cx, subset(x, Status == 'Non-Progressor')$cx)
-pv = ggplot(x, aes(Status, cx, fill=Status)) + 
+complexity.score = tibble('Status'=labels, 'cx'=cx[names(labels)], 'Samplename' = names(labels)) %>%
+  mutate(Status = factor(Status, labels = c('Non-Progressor','Progressor'))) %>%
+  left_join(patient.info[,c('Samplename','Pathology','Patient')],by=c('Samplename'))
+
+tt = t.test(subset(complexity.score, Status == 'Progressor')$cx, subset(complexity.score, Status == 'Non-Progressor')$cx)
+pv = ggplot(complexity.score, aes(Status, cx, fill=Status)) + 
   geom_jitter(width=0.2, color='grey39') + 
@@
 
-rocX = pROC::roc(Status ~ cx, data=x, auc=T, ci=T, of='thresholds')
+rocX = pROC::roc(Status ~ cx, data=complexity.score, auc=T, ci=T, of='thresholds', transpose=T)
 roc.plot(rocX, title='Per sample cx') + plot.theme
@@
 
+rocX = pROC::roc(Status ~ cx, data=complexity.score, auc=T, ci=T, of='thresholds', transpose = F)
+roc.plot(rocX, title='Per sample cx') + plot.theme
+
+ggsave('plots/per_sample_cx_roc.png', plot=roc.plot(rocX, title='Per sample cx') + plot.theme + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
 ```
@@
 
-mp = model.performance(performance.at.1se, coefs)
+mp = model.performance(performance.at.1se, coefs, folds, splits )
 mp$plot = mp$plot + labs(title=paste('All samples,', folds, 'folds,', splits, 'patient splits'))
@@
 most.stable = grep('^\\d+', names(which(mp$stable.coeffients[[select.alpha]] >= 0.75 )), value=T)
-
 most.stable = cbind.data.frame(do.call(rbind, strsplit(most.stable, ':|-')), most.stable)
@@
 
-Values that are closer to full lasso (alpha=1) are pretty similar, however at full lasso we see fewer features than if we regularize it at `r select.alpha`. They share all non-zero features as well.
-`r sapply(mp$stable.coeffients, length) %>% t %>% kable(caption='Total features at each value of alpha') %>% kable_styling(full_width = F)`
-
 ## Feature Stability
 
-They share all of the non-zero coefficients (`r which.max(sapply(mp$stable.coeffients, length))` identifies the most non-zero coefs `r max(sapply(mp$stable.coeffients, length))`).
-
 ## Remove HGD/IMC & LGD Samples and check the model
 
-Just as a point, the initial model is trained against `r nrow(dysplasia.df)` samples with `r ncol(dysplasia.df)` features. The samples include the final (diagnostic) HGD/IMC samples from progressors.
+The initial model is trained against `r nrow(dysplasia.df)` samples with `r ncol(dysplasia.df)` features. The samples include the final (diagnostic) HGD/IMC samples from progressors.
 
@@
   message(paste("loading file", file))
-  load(file, verbose=T)
+  load(file, verbose=F)
 } else {
@@
 ```{r, echo=F, fig.width=6, fig.height=6}
-mp = model.performance(performance.at.1se, coefs)
+mp = model.performance(performance.at.1se, coefs, folds, splits)
 mp$plot = mp$plot + labs(title='No HGD/IMC samples')
 mp$plot
-
 ggsave('plots/performance/ridge_v_lasso_nohgd.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)
@@
                               'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
-                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]])
-                              )
+                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))
 
 rm(performance.at.1se, coefs, mp)
-```
-
-#### LOO minus HGD
-
-```{r loonohgd, echo=F, message=F, warning=T}
-file = paste(cache.dir, 'loonohgd.Rdata', sep='/')
-if (file.exists(file)) {
-  message(paste("loading file", file))
-  load(file, verbose=T)
-} else {
-  stop(paste("No file", file, "found"))
-}
-```
-
-```{r, echo=F, message=F, fig.width=5, fig.height=5}
-perf = cbind.data.frame(perf=performance.at.1se, pt=unique(pg.sampNOHGD$Patient))
-
-p = ggplot(melt(perf, id.vars='pt'), aes(x=variable,y=value, group=variable, fill=variable)) +
-    geom_boxplot(fill='lightcoral', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) +
-    geom_jitter(color='grey39', width=0.3) + plot.theme + 
-    geom_label(data=melt(subset(perf, perf %in% range(perf) ), id.vars='pt'), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
-    theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')
-p
-
-ggsave('plots/performance/loo_perf_nohgd.png', plot=p, width=5, height=5, units='in', dpi=300)
-
-pg.sampNOHGD = pg.sampNOHGD %>% filter(!is.na(Prediction))
-
-rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')
-
-```
-
-```{r, echo=F, fig.width=6, fig.height=6}
-ggplot(pg.sampNOHGD, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1), show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') + labs(title='Prediction using model w/o HGD', y='n Samples', x='Probability') + plot.theme
-
-cuts = seq(0,1,0.1) 
-pg.sampNOHGD$quants = with(pg.sampNOHGD, cut(Prediction, breaks=cuts))
-
-noHGD.pred.confidence = as.data.frame.matrix(table(pg.sampNOHGD$quants, pg.sampNOHGD$Status))
-noHGD.pred.confidence = bind_cols(noHGD.pred.confidence, pg.sampNOHGD %>% dplyr::group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
-noHGD.pred.confidence$quant = rownames(noHGD.pred.confidence)
-
-ft.fun<-function(NP,P) {
-  f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
-  cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
-}
-
-noHGD.pred.confidence = noHGD.pred.confidence %>% as.data.frame.matrix %>% dplyr::group_by(quant) %>% dplyr::mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )
-
-noHGD.pred.confidence[c('r1','r2')] = NA
-for (i in 1:(length(cuts)-1)) 
-  noHGD.pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)
-
-noHGD.pred.confidence$Risk = 'Moderate'
-noHGD.pred.confidence$Risk[ which(with(noHGD.pred.confidence, p.value < 0.05 & perc < .5)) ] = 'Low'
-noHGD.pred.confidence$Risk[ which(with(noHGD.pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'
-
-noHGD.pred.confidence = bind_cols(noHGD.pred.confidence, with(noHGD.pred.confidence, 
-    data.frame(ci.low=qbeta(0.025, shape1=P+.5, shape2=NP+.5),
-               ci.high=qbeta(0.975, shape1=P+.5, shape2=NP+.5))))
-
-noHGD.pred.confidence$Risk = factor(noHGD.pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)
-
-BarrettsProgressionRisk::showPredictionCalibration(noHGD.pred.confidence) + labs(subtitle='HGD excluded model')
-
-# Predict HGD samples
-file = paste(cache.dir, 'nohgd.Rdata', sep='/')
-if (file.exists(file)) {
-  load(file, verbose=T)
-
-  fitNOHGD = models[[select.alpha]]
-
-  lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']
-
-  hgd.samples = subset(patient.info, Pathology %in% c('HGD','IMC'))
-  hgd.df = dysplasia.df[intersect(hgd.samples$Samplename, rownames(dysplasia.df)),]
-
-  pred = predict(fitNOHGD, newx=hgd.df, s=lambda.opt, type='response')
-  colnames(pred) = 'Prediction'
-  hgd.samples = base::merge(hgd.samples, pred, by.x='Samplename', by.y='row.names')
-
-  hgd.pred = hgd.samples %>% dplyr::group_by(Patient, Hospital.Research.ID, Status, Endoscopy.Year) %>% 
-    dplyr::summarise( max=max(Pathology), pred=Prediction[which.max(Pathology)] )
-
-  rm(no.hgd.plots, coefs, performance.at.1se, models)
-}
-#hist(hgd.samples$Prediction)
-
-#rm(plots, performance.at.1se,coefs, nzcoefs, fits)
 ```
@@
 file = paste(cache.dir, 'nolgd.Rdata', sep='/')
-
 if (file.exists(file)) {
@@
 ```{r, echo=F, fig.width=6, fig.height=6}
-mp = model.performance(performance.at.1se, coefs)
+mp = model.performance(performance.at.1se, coefs, folds, splits)
 mp$plot = mp$plot + labs(title='No LGD/HGD/IMC samples')
 mp$plot
-
 ggsave('plots/performance/no_lgd_perf.png', plot=mp$plot, height=5, width=5, units='in', dpi=300)
@@
 file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
-load(file, verbose=T)
+load(file, verbose=F)
 rm(plots, performance.at.1se)
@@
 featureStability = rowSums(coefs[[select.alpha]][,-1])/(splits*folds)
-
 hazards = apply( exp(t(dysplasia.df[, rownames(coefs[[select.alpha]])]) *  coefs[[select.alpha]][,1]), 1, function(x) {
@@
 
-features = as_tibble(hazards, rownames = 'label') %>% dplyr::rename(hazards = value) %>% mutate(coef = coefs[[select.alpha]][,1], stability = featureStability) %>% arrange(-hazards)
+features = as_tibble(hazards, rownames = 'label') %>% dplyr::rename(hazards = value) %>% 
+  dplyr::mutate(coef = coefs[[select.alpha]][,1], stability = featureStability) %>% arrange(-hazards)
 
@@
 p
-
 ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)
-
 ```
@@
 features = features %>% separate(label, c('chr','start','end'), sep=':|-', remove=F) %>% 
-  mutate_at(vars(start, end), list(as.numeric)) %>% mutate(chr = factor(chr,levels=c(1:22)))
-
-features = features %>% mutate(ymax = ifelse(coef > 0, max(globalCN.plot$data$mean.value), min(globalCN.plot$data$mean.value)), 
-                    gl = factor(ifelse(coef > 0, 'gain','loss')))
-
-features = left_join(features, chr.lengths %>% dplyr::select(chr, chr.length), by=c('chr')) %>% mutate(chr = factor(chr, levels=c(1:22), ordered=T))
+  dplyr::mutate_at(vars(start, end), list(as.numeric)) %>% 
+  mutate(chr = factor(chr,levels=c(1:22))) %>% 
+  dplyr::mutate(ymax = ifelse(coef > 0, max(globalCN.plot$data$mean.value), min(globalCN.plot$data$mean.value)), gl = factor(ifelse(coef > 0, 'gain','loss'))) %>%
+  left_join(chr.lengths %>% dplyr::select(chr, chr.length), by=c('chr')) %>% mutate(chr = factor(chr, levels=c(1:22), ordered=T))
 
@@
   geom_rect(data= features %>% filter(gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')
-
 ggsave('plots/global_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)
@@
 p = globalBECN.plot +  geom_rect(data=features %>% filter(gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=features %>% filter(gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')
-
 ggsave('plots/ndbe_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)
-
 ```
@@
 ```{r leaveoneout, echo=F, message=F, warning=F, fig.width=5, fig.height=5}
-
-file = paste(cache.dir, 'loo.Rdata', sep='/')
+file = paste(cache.dir, 'loo_0.9.Rdata', sep='/')
 if (file.exists(file)) {
-  load(file, verbose=T)
+  load(file, verbose=F)
 } else {
@@
 p
-
 ggsave('plots/performance_loo.png', plot=p, width=5, height=5, units='in', dpi=300)
@@
 
-Based on these curves we get our best AUC at ~0.36.  Currently we are using 0.5 by default.
+#### LOO minus HGD
+
+```{r loonohgd, echo=F, message=F, warning=T}
+file = paste(cache.dir, 'loonohgd.Rdata', sep='/')
+if (file.exists(file)) {
+  message(paste("loading file", file))
+  load(file, verbose=T)
+} else {
+  stop(paste("No file", file, "found"))
+}
+
+perf = cbind.data.frame(perf=performance.at.1se, pt=unique(pg.sampNOHGD$Patient))
+
+p = ggplot(melt(perf, id.vars='pt'), aes(x=variable,y=value, group=variable, fill=variable)) +
+    geom_boxplot(fill='lightcoral', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) +
+    geom_jitter(color='grey39', width=0.3) + plot.theme + 
+    geom_label(data=melt(subset(perf, perf %in% range(perf) ), id.vars='pt'), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
+    theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')
+ggsave('plots/performance/loo_perf_nohgd.png', plot=p, width=5, height=5, units='in', dpi=300)
+
+pg.sampNOHGD = pg.sampNOHGD %>% filter(!is.na(Prediction))
+
+rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds', transpose=T)
+
+```
 
 ```{r insilico_rocs, echo=F, message=F, warning=F, fig.height=8, fig.width=8}
-
-load(file=paste(cache.dir, 'insilico_splits.Rdata', sep='/'), verbose=T)
-
+load(file=paste(cache.dir, 'insilico_splits.Rdata', sep='/'), verbose=F)
 ```
@@
 insilico.rocs[['discovery']] = roc
-
 df = do.call(rbind, lapply(insilico.rocs, function(r) 
-    cbind.data.frame('Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'model'=r$model, 'is'=as.character(r$is) 
+    cbind.data.frame('AUC' = as.numeric(r$auc), 'Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'model'=r$model, 'is'=as.character(r$is) 
 )))
 
-p = ggplot(df, aes(Specificity, Sensitivity,color=is)) + geom_line(size=1, color='lightblue') + 
-    geom_line(data=subset(df, is == 'discovery'), color='red', size=1) +
+#summary(filter(df, model == 'in-silico')$AUC)
+
+getPalette = colorRampPalette(brewer.pal(5, "Blues"))
+p = ggplot(df, aes(Specificity, Sensitivity,color=is)) + 
+  geom_line(linetype='solid', size=0.5) + 
+  geom_line(data=subset(df, is == 'discovery'), color='red', size=1.5) +
   scale_x_reverse() +
-    #scale_color_brewer(palette='Blues') +
+  scale_color_manual(values=getPalette(length(unique(df$is)))) +
   labs(title='ROC, all splits', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
@@
 
-# roc.plot(rocNOHGD, 'Model excluding HGD/IMC')
-# pROC::coords(rocNOHGD, "best")
-# pROC::coords(rocNOHGD, 0.6)
-# pROC::auc(rocNOHGD, partial.auc=c(0.89,0.9), partial.auc.focus='spec', partial.auc.correct=T)
-
 cutoff = round(pROC::coords(roc, 'best', ret = 'threshold', transpose=T), 2)
@@
 roc$model = "CN Model"
-
 m = rbind.data.frame(
@@
 
-```{r echo=F, fig.height=8, fig.width=6}
-roc$model = "Per Sample"
-
-m = rbind.data.frame(
-      get.roc.stat(roc),
-      get.roc.stat(rocEndoM),get.roc.stat(rocEndoA),
-      get.roc.stat(rocPtM), get.roc.stat(rocPtA),
-      get.roc.stat(rocFirstEndo), get.roc.stat(rocLastEndo))
-m$model = factor(m$model)
-
-cols = brewer.pal(11, 'Paired')
-dodge = position_dodge(width=0.9)
-mm = melt(subset(m, Measure == 'AUC'), measure.vars='value')
-p = ggplot( arrange(mm, CI.min), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=model)) + 
-  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='grey39' ) + 
-  geom_text(aes(y=CI.min*100, label=paste(round(value,2)*100,'%',sep='')), position=dodge, vjust=3)  +
-  scale_fill_manual(values=cols, name='') + plot.theme + 
-  theme(legend.position='none', axis.text.x=element_text(angle=45, hjust=1)) + labs(x='', y='AUC', title='ROC Comparison') 
-p
-
-ggsave('plots/auc/multi_auc.png', plot=p, height=6, width=5, units='in',dpi=300)
-
-#multi.roc.plot(list(roc, rocEndoM, rocEndoA, rocPtM, rocPtA,rocFirstEndo,rocLastEndo), 'ROC Comparisons', palette='Dark2')
-```
-
 ## Prediction probability distribution
 
-```{r back, echo=F, message=F, warning=F}
-transform.by.time<-function(pt, info) {
-  pt = pt %>% dplyr::group_by(PID, Samplename) %>% dplyr::mutate( 'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12) %>% arrange(Pathology, PID, Endoscopy.Year) %>% ungroup
-
-  tb = with(pt, table(Endoscopy.Year, PID))
-  tb[tb>0] = 1
-  for (yr in names(which(rowSums(tb) > 1))) {
-      tmp = pt[which(pt$Endoscopy.Year == yr),]
-      tmp = tmp %>% dplyr::group_by(PID) %>% summarise(max(Pathology))
-
-      match = which(pt$Endoscopy.Year == yr & pt$PID == as.character(tmp[which.min(tmp$`max(Pathology)`), 'PID']))
-      pt[match,'months.before.final'] = pt[match,'months.before.final'] + 6
-  }
-  pt$nl = (pt$Block-1)*2
-
-  tb = table(pt[,c('months.before.final', 'PID')])
-  for (i in 1:nrow(tb)) {
-    pids = tb[i,]
-    yrs = which(pids > 0)
-    if (length(yrs) > 1) {
-    for (j in 2:length(yrs)) 
-      pt[which(pt$PID == names(yrs)[j]), 'nl'] = pt[which(pt$PID == names(yrs)[j]), 'nl'] + 1
-    }
-  }
-  arrange(pt, -Endoscopy.Year, PID)
-}
-
-back = do.call(bind_rows,lapply(unique(pg.samp$Patient) %>% map( function(p) pg.samp %>% filter(Patient == p)  ), transform.by.time, info=patient.info)) %>% mutate(Pathology = factor(Pathology, levels=c('BE','ID','LGD','HGD','IMC'), ordered = T))
-
-sq = c(0, seq(1,132,12),max(back$months.before.final))
-
-back = back %>% mutate(
+```{r}
+pg.samp = do.call(bind_rows,lapply(unique(pg.samp$Patient) %>% map( function(p) pg.samp %>% filter(Patient == p)), transform.by.time, info=patient.info)) %>% dplyr::mutate(Pathology = factor(Pathology, levels=c('BE','ID','LGD','HGD','IMC'), ordered = T))
+
+sq = c(0, seq(1,132,12),max(pg.samp$months.before.final))
+
+pg.samp = pg.samp %>% mutate(
   brks = cut(months.before.final, breaks=sq, 
                  labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-c(length(sq)-1,length(sq))], '>120'),include.lowest = T, ordered_result = T),
-) %>% mutate(brks = factor(brks, rev(levels(brks))))
-
-ggplot(back, aes(Status, Prediction, group=Status)) + 
+) %>% dplyr::mutate(brks = factor(brks, rev(levels(brks))))
+
+ggplot(pg.samp, aes(Status, Prediction, group=Status)) + 
+  facet_grid(~Pathology) + 
+  #scale_fill_manual(values=RColorBrewer::brewer.pal(4,"Paired")[c(1,3)]) + 
+  scale_color_gradientn(colours = myPal) +
+  geom_boxplot(outlier.shape = NA, fill='#BDBDBD') + 
+  geom_jitter(width=0.2, aes(color=Prediction)) + 
+  plot.theme  + theme(legend.position = 'none') + labs(y='Probability')
+
+ggplot(pg.samp, aes(Status, RR, group=Status)) + 
   facet_grid(~Pathology) + 
@@
 
-ggplot(back, aes(Status, RR, group=Status)) + 
-  facet_grid(~Pathology) + 
-  #scale_fill_manual(values=RColorBrewer::brewer.pal(4,"Paired")[c(1,3)]) + 
-  scale_color_gradientn(colours = myPal) +
-  geom_boxplot(outlier.shape = NA, fill='#BDBDBD') + 
-  geom_jitter(width=0.2, aes(color=Prediction)) + 
-  plot.theme  + theme(legend.position = 'none')
-
 ```
@@
 
-### TODO: I need to redo the calibration, my confidence intervals seem wrong.
-
 ```{r spacial, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
+maxP = 0.5; minP = 0.3
+
 cuts = seq(0,1,0.1)
 
-back = back %>% mutate(quants = cut(Prediction, breaks=cuts, include.lowest = T))
-qt = back %>% group_by(quants, Status) %>% dplyr::summarise(n=length(Status) ) %>% spread(Status, n)
+pg.samp = pg.samp %>% mutate(quants = cut(Prediction, breaks=cuts, include.lowest = T))
+qt = pg.samp %>% group_by(quants, Status) %>% dplyr::summarise(n=length(Status) ) %>% spread(Status, n)
 
@@
 
-qt = left_join(qt, back %>% dplyr::group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))#, 'se'=(sd(Prediction)/sqrt(length(Prediction)))*1.96 ))
+qt = left_join(qt, pg.samp %>% dplyr::group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ), by='quants')
 
 pred.confidence = qt %>% dplyr::group_by(quants) %>% 
-  dplyr::mutate( 'P.ratio'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'conf'=ifelse(p.value < 0.05, '*', '') )
-
-pred.confidence = pred.confidence %>% separate(quants, c('r1','r2'), ',', remove = F) %>% 
-  mutate_at(vars('r1','r2'), list(sub), pattern='\\[|\\]|\\(', replacement='') %>% mutate_at(vars(r1,r2), as.double)
-
-pred.confidence = pred.confidence %>% dplyr::mutate(Risk = ifelse(P.ratio<.5, 'Low','High'))
+  dplyr::mutate( 'P.ratio'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'conf'=ifelse(p.value < 0.05, '*', '') ) %>% 
+  separate(quants, c('r1','r2'), ',', remove = F) %>% 
+  mutate_at(vars('r1','r2'), list(sub), pattern='\\[|\\]|\\(', replacement='') %>% mutate_at(vars(r1,r2), as.double) %>%  
+  dplyr::mutate(Risk = ifelse(round(P.ratio,1)<.5, 'Low','High'))
+
 pred.confidence = bind_cols(pred.confidence, 
@@
 #pred.confidence[which( rsid >= 1 |  rsid <= -1 ), 'Risk'] = 'Moderate'
-pred.confidence = pred.confidence %>% mutate(Risk = as.character(ifelse(p.value > 0.05 & r1 < 0.6, 'Moderate', Risk)))
+pred.confidence = pred.confidence %>% mutate(Risk = as.character(ifelse(r1 < maxP & r2> minP, 'Moderate', Risk)))
 
@@
   padding=unit(c(5,5),'mm'))
-
 #grid.arrange( tableGrob(annotation.table, rows=NULL, theme=tt3) )
@@
 
-back = left_join(back, pred.confidence %>% dplyr::select(quants, Risk), by='quants')
-
+pg.samp = left_join(pg.samp, pred.confidence %>% dplyr::select(quants, Risk), by='quants')
 ```
@@
 
-back.per.endo = back %>% group_by(Hospital.Research.ID, Status, Patient, PID) %>% 
+pred.confidence %>% dplyr::mutate_if(is.double, list(signif), 2) %>% kable() %>% kable_styling('striped')
+
+per.endo = pg.samp %>% group_by(Hospital.Research.ID, Status, Patient, PID) %>% 
   dplyr::mutate('n.samples'=length(PID),'Samples'=paste(Samplename,collapse=',')) %>% 
@@
   arrange(Patient, Endoscopy.Year, PID)  
-
-back.per.endo = back.per.endo %>% mutate(quants = cut(Prediction, breaks=cuts, include.lowest = T)) %>% left_join(pred.confidence %>% dplyr::select(quants, Risk), by='quants') %>% dplyr::select(-matches('quant')) %>% dplyr::mutate(Risk = as.character(Risk))
-
-back %>% group_by(Status, Risk)  %>% dplyr::summarise(n = length(Risk)) %>% spread(Risk, n) %>% rowwise %>% 
+per.endo = per.endo %>% dplyr::mutate(quants = cut(Prediction, breaks=cuts, include.lowest = T)) %>% 
+  left_join(pred.confidence %>% dplyr::select(quants, Risk), by='quants') %>% 
+  dplyr::select(-matches('quant'))# %>% dplyr::mutate(Risk = as.character(Risk))
+
+pg.samp %>% group_by(Status, Risk) %>% dplyr::summarise(n = length(Risk)) %>% spread(Risk, n) %>% rowwise %>% 
   dplyr::mutate(n=sum(High,Low,Moderate), Low = Low/n, Moderate = Moderate/n, High = High/n) %>%
   mutate_if(is.double, list(round),2) %>%
-  kableExtra::kable(caption='Training set per-sample predictions') %>% kableExtra::kable_styling(full_width = F)
-
-back.per.endo %>%  group_by(Status, Risk)  %>% dplyr::summarise(n = length(Risk)) %>% spread(Risk, n) %>% rowwise %>% 
+  kableExtra::kable(caption='Training set per-sample predictions (ratios)') %>% kableExtra::kable_styling(full_width = F)
+
+per.endo %>%  group_by(Status, Risk)  %>% dplyr::summarise(n = length(Risk)) %>% spread(Risk, n) %>% rowwise %>% 
   dplyr::mutate(n = sum(High,Low,Moderate), Low = Low/n, Moderate = Moderate/n, High = High/n) %>% mutate_if(is.double, list(round),2) %>%
   kableExtra::kable(caption='Training set per-endoscopy predictions') %>% kableExtra::kable_styling(full_width = F)
+
 ```
@@
 ```{r probhist, echo=F, warning=F, message=F, fig.width=5, fig.height=5}
-ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') +
+ggplot(pg.samp, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') +
     geom_vline(xintercept=annotation.table$max, linetype='longdash', color='grey39') +
@@
 ```{r relrisk, echo=F, fig.width=6, fig.height=6}
-unadjRR = ggplot(back, aes(RR)) + geom_histogram(aes(fill=..x..), bins=10, show.legend = F) +
+unadjRR = ggplot(pg.samp, aes(RR)) + geom_histogram(aes(fill=..x..), bins=10, show.legend = F) +
   scale_fill_gradientn(colors = myPal,  name='') + 
@@
 unadjRR
-
 ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR, width=6, height=5, units='in', dpi=300)
 
-predhist = ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) +
+predhist = ggplot(pg.samp, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) +
     scale_fill_gradientn(colors = myPal,  name='') + 
     plot.theme + labs(title='Predictions, all samples model', y='n Samples', x='Probability') 
-predhist
-
 ggsave(filename='plots/pred_hist.png', plot=predhist, width=5, height=5, units='in', dpi=300)
@@
 
-mtn.spatial<-function(df, b, limits, title='', pal=NULL) {
-  if (is.null(pal))
-    pal = RColorBrewer::brewer.pal(length(limits), 'Set2')
-
-  df = merge(df, b[,c('Samplename','Endoscopy.Year','months.before.final','Block','ogj')], by.x='variable',by.y='Samplename')
-  df$ogj = factor(df$ogj, ordered = T)
-  df$ogj = factor(df$ogj, levels=rev(levels(df$ogj)), ordered = T)
-  #pal = pal[which(limits %in% levels(df$nl))]
-
-  altcolor = c(RColorBrewer::brewer.pal(9, 'Greys')[2],RColorBrewer::brewer.pal(9, 'Greys')[4])
-  fills = c()
-  for (yr in sort(unique(df$Endoscopy.Year))) {
-    ac = (which(sort(unique(df$Endoscopy.Year)) == yr) %% 2) + 1
-    fills = c(fills, rep(altcolor[ac], length(which(table(subset(df, Endoscopy.Year == yr)$ogj) > 0))))
-  }
-
-  blanklabel <- function(s) {
-    return('')
-  }
-
-  p = ggplot(df, aes(x=chr.length)) + facet_grid(Endoscopy.Year+ogj~chr, scales='free', space='free_x', labeller = labeller(.multi_line = F)) +
-      geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=ogj), show.legend=T) + 
-      scale_fill_manual(values=pal, limits=limits,  name='OGJ Dist.') +
-      labs(title=title, x='') + min.theme
-
-  legend = get.legend(p)
-  p = p + theme(legend.position = 'none')
-
-  g = ggplot_gtable(ggplot_build(p))
-  strips = grep('strip-r', g$layout$name)
-
-  for (n in 1:length(strips)) {
-    i = strips[n]
-    j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
-    g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[n]
-  }
-
-  #grid.arrange(g)
-  return(list('grob'=g, 'legend'=legend))
-  #grid::grid.draw(g)  
-  #return(g)
-}
-
-back$ogj = back$nl
-back$ogj[which(back$nl > 0 & back$nl < 3)] = 1 
-back$ogj[which(back$nl > 3 & back$nl < 7)] = 2 
-back$ogj[which(back$nl >= 7)] = 3
-
-ogjLimits = levels(factor(back$ogj, ordered = T))
-
-df = subset(globalCNdf, Patient == 20 & seg.type != 'arms')
-#df = subset(globalCNdf, Patient == 51 & seg.type != 'arms')
+setOgj <- function(nl) {
+  if (nl >= 0 & nl < 3) return(1)
+  if (nl >= 3 & nl < 7) return(2)
+  if (nl >= 7) return(3)
+}
+pg.samp = pg.samp %>% mutate(ogj = factor(unlist(purrr::map(nl, setOgj)), ordered=T))
+
+#df = globalCNdf %>% filter(Patient == 20 & seg.type != 'arms')
+#df = globalCNdf %>% filter(Patient == 51 & seg.type != 'arms')
 
@@
 #df = subset(globalCNdf, Patient == pt)
-df = merge(globalCNdf, back[,c('Samplename','Endoscopy.Year', 'ogj', 'months.before.final', 'RR','Prediction')], by.x='variable', by.y='Samplename')
-df = arrange(df, chr, start, Endoscopy.Year)
+df = left_join(globalCNdf, pg.samp[,c('Samplename','Endoscopy.Year', 'ogj', 'months.before.final', 'RR','Prediction')], by=c('variable'='Samplename')) %>% arrange(chr, start, Endoscopy.Year)
 
@@
 
-#ggplot(el2, aes(Status,`diff(arms)`)) + geom_boxplot() + geom_jitter(width=0.2)
-#ggplot(el2, aes(Status,`diff(CNA)`)) + geom_boxplot() + geom_jitter(width=0.2)
-
-p2 = mtn.spatial(subset(globalCNdf, Patient == 20 & seg.type != 'arms'), back, ogjLimits,'Patient 20 (P)' )
-p6 = mtn.spatial(subset(globalCNdf, Patient == 51 & seg.type != 'arms'), back,ogjLimits,'Patient 51 (NP)' )
-
-for (i in 1:nrow(sum.patient.data)) {
-  print(i)
-  pt = sum.patient.data[i, 'Patient']
-  status = sum.patient.data[i, 'Status']
-
-  pms = mtn.spatial(subset(globalCNdf, Patient == pt & seg.type != 'arms'), back, ogjLimits, paste('Patient', pt, '(', status, ')' ))
-
-  fl = paste('patient_', pt, '.png', sep='')
-  ggsave(paste('plots/spatial/', fl, sep=''), plot=grid.arrange(pms$grob), width=14, height=6, units='in', dpi=300)
-}
-
-back$ogj = NULL
-
+ogjLimits = levels(factor(pg.samp$ogj, ordered = T))
+p2 = mtn.spatial(globalCNdf %>% filter(Patient == 20 & seg.type != 'arms'), pg.samp, ogjLimits,'Patient 20 (P)' )
+p6 = mtn.spatial(globalCNdf %>% filter(Patient == 51 & seg.type != 'arms'), pg.samp, ogjLimits,'Patient 51 (NP)' )
+
+# for (i in 1:nrow(sum.patient.data)) {
+#   #print(i)
+#   pt = sum.patient.data[[i, 'Patient']]
+#   status = as.character(sum.patient.data[[i, 'Status']])
+# 
+#   pms = mtn.spatial(subset(globalCNdf, Patient == pt & seg.type != 'arms'), pg.samp, ogjLimits, paste('Patient', pt, '(', status, ')' ))
+# 
+#   fl = paste0('patient_', pt, '_',status,'.png')
+#   ggsave(paste('plots/spatial/', fl, sep=''), plot=grid.arrange(pms$grob), width=14, height=6, units='in', dpi=300)
+# }
+
+pg.samp = pg.samp %>% dplyr::select(-ogj)
 ```
@@
 ```{r, echo=F, fig.width=2, fig.height=4}
-
 grid.arrange(p2$legend)
@@
 ggsave(filename='plots/spatial_mtn_pt51.png', plot=grid.arrange(p6$grob), width=14, height=6, units='in', dpi=300)
-
-```
-
-At each quintile are the percentage of progressor samples that were predicted within that range, and the hypergeometric p-value (Fishers) for the significance within that quintile.
-Then our predictions are at best a coin flip in the middle, but highly significant at the outer edges.
-This would also push the prediction for "high" risk up to `r min(subset(pred.confidence, labels == 'High' )$r1)` while "low" risk is below `r max(subset(pred.confidence, labels == 'Low' )$r2)`.
-
-`r pred.confidence %>% mutate_if(is.double, list(signif), 2) %>% kable() %>% kable_styling('striped')`
-
-## Jacknife coefficients
-
-The idea here is to get some sort of CI for the coefficients, I'm not entirely sure this is correct though.
-
-Could just as easily get a jacknife CI per patient
-
-```{r jacknife, echo=T, warning=F, message=F, fig.height=10, fig.width=12}
-
-load(paste0(cache.dir, '/all.pt.alpha.Rdata'), verbose=T)
-rm(plots, performance.at.1se, dysplasia.df, models, cvs, labels)
-
-coefs.per.pt = data.frame(matrix(ncol=length(unique(pg.samp$Patient)), nrow=nrow(features), dimnames=list( features$label, unique(pg.samp$Hospital.Research.ID))))
-
-for (pt in unique(pg.samp$Hospital.Research.ID)) {
-  lcf = loo.coefs[[pt]][,1]
-  names(lcf) = rownames( loo.coefs[[pt]] )
-  sharedcfs = intersect(names(lcf), features$label)
-  coefs.per.pt[sharedcfs, pt] = lcf[sharedcfs]
-}
-coefs.per.pt[is.na(coefs.per.pt)] = 0
-
-nzcoefs = purrr::map(nzcoefs, as_tibble, rownames='coef')
-
-cfs = unique(unlist(sapply(nzcoefs, function(x) x[['coef']])))
-loo.coefs = data.frame(matrix(ncol=0,nrow=length(cfs)))
-loo.coefs$coef = cfs
-for (pt in names(nzcoefs)) {
-  loo.coefs = dplyr::full_join(loo.coefs, nzcoefs[[pt]], by='coef')
-}
-colnames(loo.coefs)[-1] = names(nzcoefs)
-loo.coefs[is.na(loo.coefs)] = 0
-
-features = as_tibble(coefs$`0.9`[,1,drop=F], rownames='coef')
-
-loo.features = dplyr::left_join(features, loo.coefs, by='coef')
-
-library(bootstrap)
-
-features$jack.se = apply(loo.features, 1, function(x) {
-  jk = bootstrap::jackknife(x[-1], var)
-  jk$jack.se
-})
-
-# per patient
-# jack.se = apply(coefs.per.pt, 2, function(x) {
-#   jk = jackknife(x, var)
-#   jk$jack.se
-# })
-# js = cbind.data.frame(performance.at.1se, jack.se)
-# ggplot(js, aes(rownames(js), performance.at.1se, color=(performance.at.1se > mean(performance.at.1se)+sd(performance.at.1se) | performance.at.1se < mean(performance.at.1se)-sd(performance.at.1se)  ))) + ylim(0.6,1) + geom_point() +
-#   geom_errorbar(aes(ymin=performance.at.1se-jack.se, ymax=performance.at.1se+jack.se)) + theme(legend.position = 'none', axis.text.x = element_text(angle=90, hjust=1))
-
-# in this case jack.se is the SE of the variance 
-# per coefficient
-features$jack.se = apply(coefs.per.pt, 1, function(x) {
-  jk = jackknife(x, var)
-  jk$jack.se
-})
-colnames(features) = c('label','coef','jack.se')
-ggplot(features, aes(reorder(label, coef), coef)) + geom_point() + 
-  geom_errorbar(aes(ymin=coef-jack.se, ymax=coef+jack.se)) + 
-  coord_flip() +
-  plot.theme + labs(x='', title='Jackknifed coefficient CI')
-```
-
-Coefficients have fairly tight CI if the SE of the jackknife variance is used.
+```
 
@@
 ```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
-
-riskPath = back %>% dplyr::group_by(Status, Pathology, Risk) %>% 
-  dplyr::summarise(Freq = length(Risk)) %>% 
-  mutate(Status = recode_factor(Status, NP = 'Non-Progressor', P = 'Progressor', .ordered = T))
-
-# riskPath = rbind( cbind.data.frame( table(subset(back, Status == 'P', select=c('Pathology','Risk'))), 'Status'='Progressor' ), 
-#                   cbind.data.frame( table(subset(back, Status == 'NP', select=c('Pathology','Risk'))), 'Status'='Non-Progressor' ))
-
-pathTotal = back %>% dplyr::group_by(Status, Pathology) %>% 
-  dplyr::summarise( 'nPath'=length(Pathology)) %>%
-  mutate(Status = recode(Status, NP = 'Non-Progressor', P = 'Progressor'))
-
-riskPath = full_join(riskPath, pathTotal, by=c('Status','Pathology')) %>% 
-  mutate(ratio = round(Freq/nPath, 3), Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered = T))
-
-p53Path = back %>% dplyr::filter(Status == 'P' & !is.na(p53.Status)) %>% 
+p53Path = pg.samp %>% dplyr::filter(Status == 'P' & !is.na(p53.Status)) %>% 
   dplyr::group_by(Pathology,p53.Status) %>% 
@@
 
-p53Path = full_join( back %>% dplyr::filter(Status == 'P' & !is.na(p53.Status)) %>% dplyr::group_by(Pathology) %>% dplyr::summarise(nPath = length(Pathology)), p53Path, by='Pathology') %>% 
+p53Path = full_join( pg.samp %>% dplyr::filter(Status == 'P' & !is.na(p53.Status)) %>% 
+                       dplyr::group_by(Pathology) %>% 
+                       dplyr::summarise(nPath = length(Pathology)), p53Path, by='Pathology') %>% 
   dplyr::mutate(ratio = round(Freq/nPath,2))
 
-pathTotal = riskPath %>% dplyr::group_by(Status, Pathology) %>% dplyr::summarise('nPath'=unique(nPath))
-
-rp = ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
-  scale_fill_manual(values=riskCols) + scale_color_manual(values=c('white','black','white')) +
-  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), position=position_stack(), size=4) +
-  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep=''), color=Risk), position=position_stack(vjust = 0.5), size=5, show.legend = F) +
-  facet_grid(~Status) + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom', text=element_text(size=10))
+rp = plot.risk.by.path(pg.samp)
 rp
-
 ggsave(filename='plots/pred_by_path.png', plot=rp, width=9, height=7, units='in', dpi=300)
@@
 ggsave(filename='plots/p53_pred.png', plot=rp, width=7, height=6, units='in', dpi=300)
-
 ```
@@
 ```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=6}
-all.preds = back %>% dplyr::group_by(Status, Patient) %>% dplyr::summarise( 'Low'=sum(length(which(Risk == "Low"))),
-                                          'Moderate'=sum(length(which(Risk == "Moderate"))),
-                                          'High'=sum(length(which(Risk == "High"))),
-                                          'n.samples'=length(Patient) ) %>% dplyr::mutate(Patient = factor(Patient))
+all.preds = pg.samp %>% dplyr::group_by(Status, Patient, Risk) %>% tally %>% spread(Risk, n) %>% ungroup %>%
+  mutate_if(is.numeric, funs( ifelse(is.na(.), 0, .) )) %>% 
+  left_join(pg.samp %>% dplyr::group_by(Status, Patient) %>% tally, by=c('Status','Patient')) %>%
+  dplyr::rename('n.samples'='n') %>% 
+  dplyr::mutate(Patient = factor(Patient))
 
@@
   plot.theme
-
 ```
@@
 
-endos = back %>% dplyr::group_by(Status, months.before.final, brks, PID) %>% dplyr::summarise(
-              'pred.endo'=ifelse(length(which(Risk == 'High')) > 0, 1, 0), 
+endos = pg.samp %>% dplyr::group_by(Status, months.before.final, brks, PID) %>% 
+  dplyr::summarise('pred.endo'=ifelse(length(which(Risk == 'High')) > 0, 1, 0), 
               'n.endos'=length(unique(PID)), 
@@
               'pred.samples'=length(which(Risk == 'High')),
-              'max.path'=max(Pathology)) %>% ungroup %>% 
+              'max.path'=max(Pathology),
+              'max.prediction' = max(Prediction)
+              ) %>% ungroup %>% 
   dplyr::mutate(brks = cut(months.before.final, include.lowest = T, ordered_result = T,
      breaks=sq,  labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))) %>% 
-  dplyr::mutate(brks = factor(brks, rev(levels(brks)))) # invert order
+  dplyr::mutate(brks = factor(brks, rev(levels(brks))))
+
+endos = endos %>% mutate(brks = recode_factor(brks, '96-120' = '>96', '>120'='>96'))
 
 endos.m = endos %>% dplyr::filter(Status == 'P') %>% dplyr::group_by(brks) %>% 
-  dplyr::summarise( 'pred.r'=sum(pred.endo)/sum(n.endos), 'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos))
-
-tmp = back %>% dplyr::filter(Status == 'P') %>% 
-  dplyr::mutate(brks = cut(months.before.final, include.lowest = T, ordered_result = T,breaks=sq,
-     labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))) %>% dplyr::group_by(brks) %>% 
-       dplyr::summarise( 'mean(P)'=mean(Prediction),'sem(P)'=sd(Prediction)/sqrt(length(brks)))
-
-endos.m = arrange(merge(endos.m,tmp), brks)
+  dplyr::summarise( 'pred.r'=sum(pred.endo)/sum(n.endos), 'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos),'mean(P)'=mean(max.prediction),'sem(P)'=sd(max.prediction)/sqrt(length(brks)))
 
@@
 
-#endos$months.before.final = endos$months.before.final*-1
-#me = endos %>% dplyr::group_by(Patient) %>% dplyr::summarise('max.endo' = max(i), 'months' = min(months.before.final))
-```
-
-The first two quintiles and the last two together account for 80% of our samples.  Only 20% fall into the category "Moderate".
+```
+
+Using the discretized risk what is the AUC?
+
+```{r}
+risk_df = pg.samp %>% dplyr::select(Status, Risk) %>% mutate( Risk  = factor(droplevels(recode_factor(Risk, 'Moderate' = 'Low')), levels=c('Low','High'), ordered=T) )
+
+rocRisk = pROC::roc(Status ~ Risk, data=risk_df, ci=T, of='thresholds', transpose = T)
+p1 = roc.plot(rocRisk) + labs(title='ROC based on discretized risk', subtitle=paste(levels(risk_df$Risk), collapse=', '))
+p1
+ggsave(filename='plots/risk_auc.png', plot=p1, width=4, height=4, units='in', dpi=300)
+
+#rocRisk2 = pROC::roc(Status ~ Risk, data=pg.samp, ci=T, of='thresholds', transpose = T)
+#p2 = roc.plot(rocRisk2) + labs(title='ROC based on discretized risk', subtitle=paste(levels(pg.samp$Risk), collapse=', '))
+#p2
+
+#grid.arrange(p1,p2, ncol=1)
+```
 
 ```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=5 }
-back = back %>% dplyr::mutate(nl = factor(nl, ordered=T), Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered = T), Patient = factor(Patient, ordered=T), Block = factor(Block, ordered=T), Pathology = recode(Pathology, BE = 'NDBE'))
-
-blank.axis = theme(axis.text = element_blank(), legend.key=element_rect(fill='grey39'), panel.background=element_rect(colour = 'black'), panel.grid.major=element_blank(), panel.spacing = unit(0.2, 'lines'), panel.border = element_rect(color="black", fill=NA, size=0.5)  )
-
-p = BarrettsProgressionRisk::patientRiskTilesPlot( back %>% filter(Patient == 6) %>% dplyr::select(Endoscopy.Year,Pathology,Block,Risk) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) ) + blank.axis
+pg.samp = pg.samp %>% dplyr::mutate(nl = factor(nl, ordered=T), 
+                                    Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered = T), 
+                                    Patient = factor(Patient, ordered=T), 
+                                    Block = factor(Block, ordered=T), 
+                                    Pathology = recode(Pathology, BE = 'NDBE'))
+
+#blank.axis = theme(axis.text = element_blank(), legend.key=element_rect(fill='grey39'), panel.background=element_rect(colour = 'black'), panel.grid.major=element_blank(), panel.spacing = unit(0.2, 'lines'), panel.border = element_rect(color="black", fill=NA, size=0.5)  )
+
+p = BarrettsProgressionRisk::patientRiskTilesPlot(pg.samp %>% filter(Patient == 6) %>% dplyr::select(Endoscopy.Year,months.before.final, Pathology,Block,Risk) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block), 'months.before.final' )
 tilesLegend = get.legend(p)
@@
 
-#pred.tiles(subset(back, Patient %in% c(6)), ncol=1, cols=myPal, probs=T, OR=F) + labs(title='') + theme(legend.position = 'none')
-#pred.tiles(subset(back, Patient %in% c(48)), ncol=1, cols=myPal, probs=T) + labs(title='') + theme(legend.position = 'none')
-#pred.tiles(subset(back, Patient %in% c(33)), ncol=1, cols=myPal, probs=T) + labs(title='')+ theme(legend.position = 'none')
-#pred.tiles(subset(back, Patient %in% c(89)), ncol=1, cols=myPal, probs=T) + labs(title='')+ theme(legend.position = 'none')
-```
-
-```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=15 }
-plist = purrr::map(c(13,20,34,56,60,48), function(pt) { 
-  BarrettsProgressionRisk::patientRiskTilesPlot( back %>% filter(Patient == pt) %>% dplyr::select(Endoscopy.Year,Pathology,Block,Risk) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) ) + blank.axis + labs(title=pt) + theme(legend.position = 'none')
+ggsave('plots/heatmaps/legend_long.png',plot=grid.arrange(tilesLegend), width=10, height=2, units='in', dpi=300)
+
+dir.create('plots/heatmaps/tiny/P',recursive = T, showWarnings = F)
+dir.create('plots/heatmaps/tiny/NP',recursive = T, showWarnings = F)
+# Tiny heatmaps
+plist = purrr::map(sum.patient.data$Patient, function(pt) {
+  patient = pg.samp %>% filter(Patient == pt) %>% dplyr::select(Patient,Status,Endoscopy.Year,Block,Risk,months.before.final) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block)
+
+  p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final','rev') + theme(legend.position = 'none') + labs(x='',y='') + theme(axis.text = element_blank(), plot.title = element_text(hjust = 0.5, face = 'bold', size=16))
+
+  ggsave(paste0('plots/heatmaps/tiny/',unique(patient$Status),'/pt_',pt,'.png'), plot = p, height=length(unique(patient$GEJ.Distance)), width=length(unique(patient$months.before.final)), units='in', dpi=300, limitsize=F)
+
+  patient = pg.samp %>% filter(Patient == pt) %>% dplyr::select(Patient,Status,Pathology,Endoscopy.Year,Block,Risk,months.before.final) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block)
+
+    p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final','rev') + theme(legend.position = 'none') + labs(x='months',y='location') + theme(plot.title = element_text(face = 'bold', size=16))
+
+    ggsave(paste0('plots/heatmaps/',unique(patient$Status),'/pt_',pt,'.png'), plot = p, height=length(unique(patient$GEJ.Distance))+1, width=length(unique(patient$months.before.final))+1, units='in', dpi=300, limitsize=F)
+
+    return(p)
 })
-a = do.call('arrangeGrob', c(plist, top='Progressors'))
-
-plist = purrr::map(c(15,16,21,3,45,19) , function(pt) { 
-  BarrettsProgressionRisk::patientRiskTilesPlot( back %>% filter(Patient == pt) %>% dplyr::select(Endoscopy.Year,Pathology,Block,Risk) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) ) + blank.axis + labs(title=pt) + theme(legend.position = 'none')
-})
-b = do.call('arrangeGrob', c(plist, top='Non-progressors'))
-
-ggsave('plots/ex_pt_heatmap.png', plot=grid.arrange(a,b,ncol=2), width=6, height=15, units='in', dpi=300)
+names(plist) = sum.patient.data$Patient
+
+```
+
+```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=15}
+a = do.call('arrangeGrob', c(plist[c('13','20','34','56','60','48')], top='Progressors'))
+b = do.call('arrangeGrob', c(plist[c('15','16','21','3','45','19')], top='Non-progressors'))
+
+ggsave('plots/ex_pt_heatmap.png', plot=grid.arrange(a,b,ncol=2), width=12, height=15, units='in', dpi=300)
 
@@
 
-summary.t = table.path.risk(back)
-
-summary.t %>% kableExtra::kable() %>% kableExtra::kable_styling('striped', full_width = F)
-
+summary.t = table.path.risk(pg.samp)
+#summary.t %>% kableExtra::kable() %>% kableExtra::kable_styling('striped', full_width = F)
 write_tsv(summary.t, path = 'plots/risk_summary.tsv')
@@
 ```{r, echo=F, message=F, warning=F}
-back = back %>% dplyr::group_by(Patient) %>% dplyr::mutate( 'pred.ratio'=round(sum(Prediction)/length(Prediction),2))
-
-plist = purrr::map((sum.patient.data %>% filter(Status == 'P'))$Patient, function(pt) { 
-  BarrettsProgressionRisk::patientRiskTilesPlot( back %>% filter(Patient == pt) %>% dplyr::select(Endoscopy.Year,Pathology,Block,Risk) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) ) + blank.axis + labs(title=pt) + theme(legend.position = 'none')
-})
-p1 = do.call('arrangeGrob', c(plist, top='Progressors'))
-
-plist = purrr::map((sum.patient.data %>% filter(Status == 'NP'))$Patient, function(pt) { 
-  BarrettsProgressionRisk::patientRiskTilesPlot( back %>% filter(Patient == pt) %>% dplyr::select(Endoscopy.Year,Pathology,Block,Risk) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) ) + blank.axis + labs(title=pt) + theme(legend.position = 'none')
-})
-p2 = do.call('arrangeGrob', c(plist, top='Non-Progressors'))
-
-ggsave('plots/pred_heatmap.png', plot=grid.arrange(p2,p1,nrow=2), width=52, height=16, units="in", limitsize=F, dpi = 300)
-ggsave('plots/npg_heatmap.png', plot=p1, width=52, height=8, units="in", limitsize=F, dpi = 300)
-ggsave('plots/pg_heatmap.png', plot=p2, width=52, height=8, units="in", limitsize=F, dpi = 300)
+pg.samp = pg.samp %>% dplyr::group_by(Patient) %>% dplyr::mutate( 'pred.ratio'=round(sum(Prediction)/length(Prediction),2))
+
+p1 = do.call('arrangeGrob', c(plist[as.character(filter(sum.patient.data,Status == 'P')$Patient)], top='Progressors'))
+p2 = do.call('arrangeGrob', c(plist[as.character(filter(sum.patient.data,Status == 'NP')$Patient)], top='Non-Progressors'))
+
+ggsave('plots/heatmaps/pred_heatmap.png', plot=grid.arrange(p2,p1,nrow=2), width=52, height=16, units="in", limitsize=F, dpi = 300)
+ggsave('plots/heatmaps/pg_heatmap.png', plot=p1, width=52, height=24, units="in", limitsize=F, dpi = 300)
+ggsave('plots/heatmaps/npg_heatmap.png', plot=p2, width=52, height=24, units="in", limitsize=F, dpi = 300)
 
 #ggsave('plots/risk_legend.png', plot=grid.arrange(riskLegend), width=2,height=4, dpi=300)
-
 #ggsave('plots/risk_legend_long.png', plot=grid.arrange(riskLegendLong), width=8,height=2, dpi=300)
@@
 ## Top Features
-```{r, echo=T}
+```{r, echo=T, eval=F}
 m = data.frame(matrix(nrow=nrow(features), ncol=0))
@@
 
-## Chr 17 Predictive
+## Chr 17 Predictive TODO
 
 One way of looking at the p53 prediction (e.g. Adam Bass) is to see how predictive the p-arm loss in chr17 is.
-```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=10}
-ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
-ccgenes = ccgenes[-grep('^X|Y', ccgenes$Genome.Location),]
-#ccgenes = ccgenes[-which(ccgenes$Mutation.Types == 'T'),] # drop the translocations
-#ccgenes = subset(ccgenes, !grepl('^X|Y', Genome.Location))
-#loc = unlist(strsplit(ccgenes$Genome.Location, ':|-'))
-loc = do.call(rbind.data.frame, strsplit(ccgenes$Genome.Location, ':|-'))
-colnames(loc) = c('chr','start','end')
-loc[] = lapply(loc[], function(x) as.numeric(as.character(x)))
-
-ccgenes = cbind(ccgenes, loc)
-ccgenes = ccgenes[c('Gene.Symbol','Genome.Location', 'Role.in.Cancer', 'Mutation.Types',colnames(loc))]
-#ccgenes = subset(ccgenes, grepl('\\w+',Role.in.Cancer))
-
-genes = cbind(subset(ccgenes, grepl('TSG|oncogene', Role.in.Cancer) & Mutation.Types != 'T', select=c('Gene.Symbol','Genome.Location', colnames(loc))), 'value'=min(globalCNdf$mean.value), 'Row.names'='', 'Patient'='', 'selected'=as.factor(0))
-
-genesGR = makeGRangesFromDataFrame(genes, keep.extra.columns = T)
-chGR = makeGRangesFromDataFrame(chh, keep.extra.columns = T)
-
-ov = findOverlaps(chGR, genesGR)
+```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=10, eval=F}
+ccgenes = read_tsv('~/Data/CosmicCensusGenes.tsv', col_types = cols(.default = col_character())) %>% 
+  filter(!grepl('^X|Y',`Genome Location`)) %>% 
+  separate(`Genome Location`, c('chr','start','end'),':|-', remove=F) %>%
+  dplyr::select(`Gene Symbol`, `Genome Location`, `Role in Cancer`, `Mutation Types`, chr, start, end)
+
+genes = ccgenes %>% dplyr::filter(grepl('TSG|oncogene', `Role in Cancer`) & `Mutation Types` != 'T') %>% 
+  dplyr::select(`Gene Symbol`, `Genome Location`, chr, start, end) %>%
+  dplyr::mutate(value = min(globalCNdf$value), Patient = '')
+
+#genesGR = makeGRangesFromDataFrame(genes, keep.extra.columns = T)
+#chGR = makeGRangesFromDataFrame(chh, keep.extra.columns = T)
+
+#ov = findOverlaps(chGR, genesGR)
 #genes = as.data.frame(genesGR[unique(subjectHits(ov))])
@@
 
-g17 = plot.cn.chr(globalCNdf, chrom='17', chr.lengths, genes=genes, haz=chh, label=labeller(chr=label_both, Status=n.status)) + labs(
-title="All Samples") + theme(legend.position = 'none')
+g17 = 
+  plot.cn.chr(globalCNdf, chrom='17', chr.lengths, genes=genes, label=labeller(chr=label_both, Status=n.status)) + 
+  labs(title="All Samples") + theme(legend.position = 'none')
 g17
@@
 
-```{r echo=F, warning=F, message=F, fig.height=5, fig.width=10}
+```{r echo=F, warning=F, message=F, fig.height=5, fig.width=10, eval=F}
 chGR$Genes = ''
@@
 
-```
-
-`r (genes[,c(1:3,6,7)]) %>% kable() %>% kable_styling(full_width=F)`
+(genes[,c(1:3,6,7)]) %>% kable() %>% kable_styling(full_width=F)
+```
 
@@
 
-Individually none of these show very good classification rates, all hovering around 65%.
-However, more importantly is that Chr17 is only marginally better than any of the others (unless you exclude HGD) and all are below the 72% classification rate we get with all of the data
-
-# Validation Data
-
-## Normals sWGS sequenced
-
-These normal tissues came from some samples that had been mistakenly sequenced from a timepoint (path ID) that Barrett's samples were also sequenced.  These include:
+# Validation Cohort
 
 ```{r, echo=F}
-normal.info = readxl::read_xlsx(patient.file, sheet='Normals')
-
-normal.info %>% dplyr::select(`Patient ID`, `Path ID`, `Pathology notes`) %>% kable(caption='Normals') %>% kable_styling('striped', full_width = F)
-```
-
-These were excluded from the model building procedure, but were then predicted using the resulting model.
-We would expect them to all be low risk and all but one are.
-All have a relative risk below 0 as well.
+p.diag = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/validation/progressor_diagnosis.xlsx') %>% 
+  dplyr::mutate(`Hospital Research ID` = gsub('/', '_', `Hospital Research ID`), `Weeks Follow Up` = as.double(difftime(`Diagnostic Endoscopy`, `Sequenced Endoscopy`, units='weeks')))
+
+patient.file = '~/Data/Ellie/QDNAseq/validation/sWGS_validation_batches.xlsx'
+sheets = readxl::excel_sheets(patient.file)[8:14]
+all.val.info = do.call(bind_rows, lapply(sheets, function(s) {
+  readxl::read_xlsx(patient.file, s) %>% dplyr::select(`Hospital Research ID`, matches('Status|Endoscopy|Block|Path'), `Sample Type`, `SLX-ID`, `Index Sequence`,Cohort) %>% 
+    dplyr::mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% dplyr::filter(!is.na(`SLX-ID`))
+}))
+
+pastefun<-function(x) {
+  if ( !grepl('SLX-', x) ) x = paste0('SLX-',x)
+  return(x)
+}
+
+all.val.info = all.val.info %>% rowwise %>% 
+  mutate_at(vars(`SLX-ID`), list(pastefun) ) %>% ungroup %>%
+  #mutate(`Hospital Research ID` = str_replace_all( str_remove_all(`Hospital Research ID`, " "), '/', '_'), `Index Sequence` = str_replace_all(`Index Sequence`, 'tp', '')) %>% 
+  mutate(Samplename = paste(`SLX-ID`, `Index Sequence`, sep='.')) %>%
+  separate(`Block ID`, c('PID','Block'), sep='-|[:blank:]', remove=F) %>% dplyr::rename(PathID = `Block ID`)
+
+all.val.info$Block = sapply(strsplit(all.val.info$Block, ''), function(x) {
+  if (is.na(x)) return(1)
+  if (length(x) == 1) {
+    if (!grepl('\\d', x)) return(recode(x, 'A'=1, 'B'=2, 'C'=3, 'D'=4, 'E'=5, 'F'=6,'G'=7 ))
+    return(as.numeric(x))
+  }
+  if (length(x) > 1) {
+    i = which(grepl('\\d', x))
+    return(as.numeric(x[i]))
+  }
+})
+
+all.val.info = left_join(all.val.info, p.diag, by=c('Hospital Research ID', 'PathID'='Block ID')) %>% group_by(`Hospital Research ID`) %>% 
+  dplyr::mutate(`Weeks Follow Up` = as.double(difftime(`Diagnostic Endoscopy`, `Sequenced Endoscopy`, units='weeks')))
+
+val.info = all.val.info %>% filter(Pathology != 'OAC')
+
+bind_rows(val.info %>% dplyr::select(`Hospital Research ID`, Status) %>% distinct %>% group_by(Status) %>% tally %>% 
+  spread(Status,n) %>% add_column(' ' = 'n Patients', .before=1),
+val.info %>% group_by(Status) %>% tally %>% spread(Status,n) %>% add_column(' ' = 'n Samples', .before=1)) %>% 
+  kable(caption='Validation cohort') %>% kable_styling(full_width = F)
+```
+
+## Per sample (alpha = 0.9)
+
+```{r, fig.width=8, fig.height=12}
+vdir ='~/Data/BarrettsProgressionRisk/Analysis/validation/pcf_perPatient/50kb/predictions_5e6_all'
+
+all.val.preds = do.call(bind_rows, purrr::map(list.files(vdir, 'predictions.tsv', full.names = T, recursive = T), function(f) {
+    pd = read_tsv(f, col_types = 'cddcccDcccc') %>% dplyr::select(`Hospital Research ID`, Sample, Probability, `Relative Risk`)
+    #left_join(pd, all.val.info, by=c('Hospital Research ID', 'Sample'='Samplename', 'Block ID')) 
+    return(pd)
+})) %>% mutate(quants = cut(Probability, breaks=cuts, include.lowest = T)) %>%
+  left_join(pred.confidence %>% dplyr::select(quants, Risk), by='quants')
+
+all.val.preds = left_join(all.val.preds, all.val.info, by=c('Hospital Research ID', 'Sample'='Samplename'))
+
+val.preds = all.val.preds %>% dplyr::filter(Sample %in% val.info$Samplename)
+
+oac.preds = all.val.preds %>% filter(Pathology == 'OAC')
+#all.val.preds %>% filter(Cohort == 'Old progressor scrolls') %>% group_by(Status, Risk) %>% tally
+#val.info %>% filter(Cohort != 'Old progressor scrolls' & Pathology != 'OAC') %>% group_by(Status) %>% tally
+```
 
 ```{r}
-
-normal.preds = read_tsv('~/Data/BarrettsProgressionRisk/Analysis/multipcf_plots_normals/normal_predictions.tsv', col_types = cols(
-  Sample = col_character(),
-  Probability = col_double(),
-  `Relative Risk` = col_double(),
-  Risk = col_character(),
-  Patient = col_character(),
-  Hospital.Research.ID = col_character(),
-  Path.ID = col_character(),
-  Status = col_character(),
-  Endoscopy = col_double(),
-  Pathology = col_logical(),
-  Plate.Index = col_character(),
-  SLX.ID = col_double(),
-  Barretts.Cellularity = col_logical(),
-  P53.Status = col_character(),
-  Total.Reads = col_double(),
-  Batch.Name = col_character(),
-  Set = col_character(),
-  Block = col_double()
-))
-
-ggplot(normal.preds, aes(Patient, Probability, color=Risk)) + ylim(0,1) + geom_point(size=2) +
-  scale_color_manual(values = BarrettsProgressionRisk::riskColors() ) + plot.theme
-
-```
-
-## Test 'holdout' Set
-
-20 Patients (10P, 10NP) set aside for training/test purposes
-
-```{r val_setup, echo=F, message=F, warning=F, include=F}
-validation.patient.info =  all.patient.info %>% filter(Set == 'Test') %>% arrange(Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
-sum.val.info = summarise.patient.info(validation.patient.info)
-
-sum.val.info %>% kable() %>% kable_styling()
-```
-
-```{r, echo=F}
-
-file = paste(cache.dir, 'model_data.Rdata', sep='/')
-if (!file.exists(file))
-  stop(paste("Missing data file", file))
-load(file, verbose=T)
-
-predicted = '~/Data/BarrettsProgressionRisk/Analysis/training/val_predictions/0.9'
-
-vpd = do.call(bind_rows, lapply(list.files(predicted, full.names = T), function(f) {
-  read_tsv(f, col_types = 'cddcccdccdc')
-})) %>% dplyr::select(Hospital.Research.ID, everything())
-
-vpd = left_join(vpd %>% dplyr::select(-Pathology, -matches('P53')), validation.patient.info %>% 
-                  dplyr::select(Samplename, Patient, Pathology, p53.Status, Block, Initial.Endoscopy, Final.Endoscopy, Status ), by=c('Sample'='Samplename')) %>% 
-  dplyr::rename(Samplename = Sample, Endoscopy.Year = Endoscopy, RR = `Relative Risk`, Prediction = Probability) %>% 
-  dplyr::mutate(Risk = factor(Risk, levels=levels(back$Risk), ordered = T), GEJ.Distance = factor(GEJ.Distance, ordered = T),Pathology = recode(Pathology, BE='NDBE'))
-
-vpd = do.call(bind_rows,lapply(unique(vpd$Patient) %>% map( function(p) vpd %>% filter(Patient == p)  ), transform.by.time, info=validation.patient.info))
-
-## Need to add months to each endoscopy if multiple occur in the same year
-vpd = vpd %>% dplyr::mutate(PID = factor(PID)) %>% 
-  dplyr::group_by(Patient, PID) %>% dplyr::arrange(PID) %>% 
-  dplyr::mutate( Endoscopy = as.Date(paste0(Endoscopy.Year,'-01-01')) + as.numeric(PID)  ) %>% ungroup
-
-test.roc = pROC::roc(Status~Prediction, data=vpd,auc=T, ci=T, of='thresholds',transpose=T)
-roc.plot(test.roc, title='Test Set ROC')
-
-vpd.per.endo = vpd %>% group_by(Hospital.Research.ID, Status, Patient, PID) %>% 
-  dplyr::mutate('n.samples'=length(PID),'Samples'=paste(Samplename,collapse=',')) %>% 
-  dplyr::summarise_at(vars(matches('Pred|RR|Year')), list(max)) %>% dplyr::select(-matches('^Sample$')) %>% 
-  arrange(Patient, Endoscopy.Year, PID) 
-vpd.per.endo = vpd.per.endo %>% mutate(quants = cut(Prediction, breaks=cuts, include.lowest = T)) %>% left_join(pred.confidence %>% dplyr::select(quants, Risk), by='quants') %>% dplyr::select(-matches('quant')) %>% dplyr::mutate(Risk = as.character(Risk))
-
-vpd %>% group_by(Status, Risk)  %>% dplyr::summarise(n = length(Risk)) %>% spread(Risk, n) %>% rowwise %>% 
-  dplyr::mutate(n=sum(High,Low,Moderate), Low = Low/n, Moderate = Moderate/n, High = High/n) %>%
-  mutate_if(is.double, list(round),2) %>%
-  kableExtra::kable(caption='Test set per-sample predictions') %>% kableExtra::kable_styling(full_width = F)
-
-vpd.per.endo %>%  group_by(Status, Risk)  %>% dplyr::summarise(n = length(Risk)) %>% spread(Risk, n) %>% rowwise %>% 
-  dplyr::mutate(n = sum(High,Low,Moderate), Low = Low/n, Moderate = Moderate/n, High = High/n) %>% mutate_if(is.double, list(round),2) %>%
-  kableExtra::kable(caption='Test set per-endoscopy predictions') %>% kableExtra::kable_styling(full_width = F)
-
-```
-
-```{r validation, message=F, warning=F, echo=F, fig.height=8, fig.width=8}
-
-# for (s in rownames(pred)) {
-#   vpd[which(vpd$Samplename == s), 'Prediction'] = pred[s,]
-#   vpd[which(vpd$Samplename == s), 'OR'] = or[s,]
-# }
-
-sq = c(0, seq(1,132,12),max(back$months.before.final))
-vpd = vpd %>% mutate(brks = cut(months.before.final, 
-                  breaks=sq, labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x,collapse='-')))[-c(length(sq)-1,length(sq))], '>120'), include.lowest = T, ordered_result = T)) %>%
-  mutate(brks = factor(brks, rev(levels(brks))))
-```
-
-```{r echo=F, message=F, warning=F, fig.height=5, fig.width=8}
-vpd = vpd %>% ungroup %>% dplyr::mutate(Risk = factor(Risk, levels=c('High','Moderate','Low')), nl = factor(nl, ordered = T)) %>%
-  dplyr::mutate(Patient = factor(Patient))
-
-m = melt(vpd %>% dplyr::group_by(Patient, Status) %>% dplyr::summarise( 'Low'=length(which(Risk == 'Low'))/length(Patient),
-                                            'Moderate'=length(which(Risk == 'Moderate'))/length(Patient),
-                                            'High'=length(which(Risk == 'High'))/length(Patient),
-                                            'n.samples'=length(Patient)), id.vars=c('Patient','Status','n.samples'))
-
-ggplot(m, aes(Patient,  value*100, group=variable, fill=variable)) + geom_bar(stat='identity', position=position_stack()) +
-  geom_text(aes(label=n.samples, y=101)) + scale_fill_manual(values=riskCols) +
-  facet_grid(~Status, scales='free_x') + labs(y='Percent of samples', title='Validation patient set') + plot.theme
-
-```
-
-```{r echo=F, warning=F, message=F, fig.height=10,fig.width=25}
-plist = purrr::map((sum.val.info %>% filter(Status == 'P'))$Patient, function(pt) { 
-  BarrettsProgressionRisk::patientRiskTilesPlot( vpd %>% filter(Patient == pt) %>% dplyr::select(Endoscopy,Pathology,GEJ.Distance,Risk) ) + blank.axis + labs(title=pt) + theme(legend.position = 'none')
-})
-p1 = do.call('arrangeGrob', c(plist, top='Progressors'))
-
-plist = purrr::map((sum.val.info %>% filter(Status == 'NP'))$Patient, function(pt) { 
-  BarrettsProgressionRisk::patientRiskTilesPlot( vpd %>% filter(Patient == pt) %>% dplyr::select(Endoscopy,Pathology,GEJ.Distance,Risk) ) + blank.axis + labs(title=pt) + theme(legend.position = 'none')
-})
-p2 = do.call('arrangeGrob', c(plist, top='Non-Progressors'))
-grid.arrange(p1, p2, nrow=2)
-
-ggsave(filename='plots/val_heatmap.png', plot=grid.arrange(p1, p2, nrow=2), width=18, height=6, units='in', dpi=300)
-
-#ggsave(filename='plots/val_leg.png', plot=valLeg, width=4, height=2, units='in', dpi=300)
-
-```
-
-```{r echo=F, warning=F, message=F, fig.height=5,fig.width=8}
-BarrettsProgressionRisk::patientRiskTilesPlot( vpd %>% filter(Patient == 68) %>% 
-                                                 dplyr::select(Endoscopy,Pathology,GEJ.Distance,Risk) ) + 
-  theme(legend.position = 'right')
-```
-
-```{r, echo=F, message=F, warning=F, fig.width=8, fig.height=8}
-riskPath = rbind( cbind.data.frame( table(subset(back.v, Status == 'P', select=c('Pathology','Risk'))), 'Status'='Progressor' ), 
-                  cbind.data.frame( table(subset(back.v, Status == 'NP', select=c('Pathology','Risk'))), 'Status'='Non-Progressor' ))
-
-pathTotal = back.v %>% dplyr::group_by(Status, Pathology) %>% dplyr::summarise( 'nPath'=length(Pathology))
-pathTotal$Status = ifelse(pathTotal$Status == 'P', 'Progressor','Non-Progressor')
-
-riskPath = merge(riskPath, pathTotal)
-riskPath$ratio = with(riskPath, round(Freq/nPath, 3))
-riskPath$Risk = factor(riskPath$Risk, levels=c('Low','Moderate','High'), ordered = T)
-
-hgdimc = subset(riskPath, Pathology %in% c('HGD','IMC')) %>% dplyr::group_by(Risk) %>% summarize (
-  Freq=sum(Freq),  nPath=sum(nPath), ratio=sum(Freq)/sum(nPath))
-
-riskPath = riskPath[-which(riskPath$Pathology %in% c('HGD','IMC')),] 
-riskPath = rbind(riskPath , cbind('Pathology'='HGD/IMC', 'Status'='Progressor', hgdimc))
-
-riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor'), ordered = T)
-
-pathTotal = riskPath %>% dplyr::group_by(Status, Pathology) %>% dplyr::summarise('nPath'=unique(nPath))
-
-pr = ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) +   scale_fill_manual(values=riskCols) + scale_color_manual(values=c('white','black','white')) +
-  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), position=position_stack(), size=4, show.legend = F) +
-  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep=''), color=Risk), position=position_stack(vjust = 0.5), size=5, show.legend = F) +
-  facet_grid(~Status) + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
-pr
-
-ggsave(filename='plots/val_riskpath.png', plot=pr, width=7, height=7, units='in', dpi=300)
-
-```
-
-## Downsampled Datasets
-
-These come from Annalise, adjacent Barrett's tissues, and normal blood that had all been initially WGS at 30-80x and downsampled to an average 0.4x
-
-```{r echo=F, eval=T}
-dir = '~/Data/BarrettsProgressionRisk/Analysis/downsampled/predictions/0.9'
-info = '~/Data/BarrettsProgressionRisk/QDNAseq/all_downsampled/downsampled_ids.xlsx'
-
-info = readxl::read_xlsx(info)
-
-ds.preds = do.call(bind_rows, purrr::map(list.files(dir, 'predictions', recursive=T, full.names=T), function(f) {
-  read_tsv(f, col_types = cols(
-  Sample = col_character(),
-  Probability = col_double(),
-  `Relative Risk` = col_double(),
-  Risk = col_character(),
-  PatientID = col_character(),
-  Status = col_character(),
-  Tissue = col_character(),
-  `Expected Risk` = col_character(),
-  Grade_patient = col_character(),
-  Grade_biopsy = col_character(),
-  Study = col_character(),
-  Type = col_character(),
-  Endoscopy = col_date(format = "") ))
-}))
-
-## TODO merge in RR from 'back'
-
-ds.preds = ds.preds %>% dplyr::select(-Status) %>% left_join(info %>% dplyr::select(`Illumina ID`, Status), by=c('Sample' = 'Illumina ID')) %>% dplyr::mutate(Type = factor(Type, levels=c('Normal','NDBE', 'LGD', 'Tumor Adjacent BE'))) %>% arrange(Probability)
-
-m = reshape2::melt(ds.preds, measure.vars = c('Relative Risk'), id.vars = c('Type', 'Status', 'Risk'))
-
-pr = ggplot(m, aes(variable, value)) + 
-  facet_grid(~Type, scales = 'free_x') + 
-  geom_boxplot(aes(group=variable, fill=Type), outlier.colour = NA, show.legend = F) + 
-  geom_jitter(width=0.2, aes(shape=Status, color=Risk), size=2, show.legend = F) + 
-  scale_color_manual(values =  BarrettsProgressionRisk::riskColors()) +
-  #scale_color_gradientn(colors = myPal,  name='') +
-  scale_fill_manual(values=rev(RColorBrewer::brewer.pal(5,'RdYlBu')[c(1,2,4,5)])) +
-  labs(x='', y='Relative Risk', title='Predictions on downsampled patients') + plot.theme +
-  theme(axis.text.x = element_text(angle=45, hjust=1))
-
-pr
-
-ds.preds %>% dplyr::group_by(Status, Risk) %>% dplyr::summarise(n=length(Risk)) %>% spread(Risk, n) %>% kableExtra::kable(caption = 'Sample by risk and known status') %>% kableExtra::kable_styling(full_width = F)
-
-ggsave('plots/adjacentBE.png', plot=pr, width=8, height=8,dpi=300)
-
-```
-
-## Validation Cohort
-
-We sequenced a new set of non-progressors, progressors, and some OACs from the OCCAMS cohort where we have high-depth WGS.
+roc = pROC::roc(Status ~ Probability, data=val.preds, auc=T, ci=T, of='thresholds',transpose=T)
+roc.plot(roc) + labs(title='Validation cohort')
+ggsave('plots/auc/validatation_roc.png', plot=roc.plot(roc) + labs(title='Validation cohort'), width=6, height=4, units='in',dpi=300)
+
+totals = val.preds %>% group_by(Status) %>% tally
+val.counts = val.preds %>% 
+    group_by(Status, Risk) %>% tally %>% spread(Risk,n) %>% ungroup %>%
+    left_join(totals, by='Status') %>% mutate_if(is.numeric, funs(./n)) %>% dplyr::select(-n)
+
+ggplot(melt(val.counts, id.vars=c('Status')), aes(variable, value, group=Status, fill=Status)) + ylim(0,1) + 
+    geom_bar(stat='identity', position='dodge') + geom_text(aes(label=round(value,2)), position = position_dodge(width=1)) +
+    labs(title='Validation risk (sample ratios per NP:P)', x='', y='ratio') + theme_minimal()
+```
+
+## By path
 
 ```{r}
-val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
-analysis.dir = '~/Data/BarrettsProgressionRisk/Analysis/validation'
-
-oac.cohort = readxl::read_xlsx(val.file, 8, trim_ws = T) %>% 
-    dplyr::filter(!is.na(`SLX-ID`)) %>% dplyr::select(`Hospital Research ID`, `Block ID`, `Sample Date`, `SLX-ID`, `Index Sequence`, Status) %>% mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`)) %>% dplyr::mutate(Pathology = 'OAC') %>% dplyr::rename('Endoscopy' = `Sample Date`) %>% dplyr::rename('PID' = `Block ID`) %>% dplyr::mutate(GEJ.Distance = 1)
-
-sheets = readxl::excel_sheets(val.file)[9:12]
-info = do.call(bind_rows, lapply(sheets, function(s) {
-  readxl::read_xlsx(val.file, s, trim_ws = T) %>% 
-    dplyr::filter(!is.na(`SLX-ID`)) %>% dplyr::select(`Hospital Research ID`, `Block ID`, Status, Endoscopy, Pathology, `SLX-ID`, `Index Sequence`, `Path Notes`) %>% mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`))
-}))
-
-info = info %>% separate(`Block ID`, c('PID','GEJ.Distance'), '( |-)', remove = F, fill='right') %>% 
-  dplyr::mutate(GEJ.Distance = ifelse(is.na(GEJ.Distance), '1', GEJ.Distance)) %>%
-  dplyr::mutate(GEJ.Distance = recode(GEJ.Distance, A1 = '1', A = '1', A2 = '2', A3 = '3', A4 = '4', `4A` = '4', B = '2', C = '3', D = '4')) %>%
-  dplyr::mutate(GEJ.Distance = as.numeric(GEJ.Distance))
-
-bind_rows(
-  add_column(info %>% dplyr::select(`Hospital Research ID`, Status) %>% distinct() %>% dplyr::filter(Status %in% c('P','NP')) %>% dplyr::group_by(Status) %>% dplyr::count() %>% spread(Status, n), ' ' = 'BE Patients', .before = 1),
-  add_column(info %>% dplyr::filter(Status %in% c('P','NP')) %>% dplyr::group_by(Status) %>% dplyr::count() %>% spread(Status, n), ' ' = 'BE Samples', .before=1),
-  bind_cols(' ' = 'OAC Samples','NP'=0,'P'=nrow(oac.cohort))
-  ) %>% 
-kable(caption='Validation cohort counts') %>% kable_styling(full_width = F)
-
-```
-
-```{r}
-pred.dir = paste0(analysis.dir, '/predictions')
-
-batch8_preds = read_tsv(paste0(pred.dir, '/batch8_preds.txt')) %>% mutate(`Hospital Research ID` = gsub('_','/',`Hospital Research ID`)) %>% dplyr::select(`Hospital Research ID`, Sample, Probability, `Relative Risk`, Risk) %>%
-  left_join(oac.cohort, by=c('Hospital Research ID','Sample'))
-
-oldp.preds = read_tsv(paste0(pred.dir, '/oldp_preds.txt')) %>% mutate(`Hospital Research ID` = gsub('_','/',`Hospital Research ID`)) %>% dplyr::select(`Hospital Research ID`, Sample, Probability, `Relative Risk`, Risk) %>% left_join(info, by=c('Hospital Research ID','Sample'))
-
-new.np.preds = read_tsv(paste0(pred.dir, '/val_np.txt')) %>% mutate(`Hospital Research ID` = gsub('_','/',`Hospital Research ID`)) %>% dplyr::select(`Hospital Research ID`, Sample, Probability, `Relative Risk`, Risk) %>% left_join(info, by=c('Hospital Research ID','Sample'))
-
-val.preds = bind_rows(
-  batch8_preds %>% dplyr::select(`Hospital Research ID`, Probability, `Relative Risk`, Endoscopy, Status, PID) %>% mutate(Pathology = 'OAC', GEJ.Distance = 1) %>% dplyr::rename(ID = 'Hospital Research ID'),
-  oldp.preds %>% dplyr::select(`Hospital Research ID`, Probability, `Relative Risk`, Endoscopy, GEJ.Distance, Pathology, PID) %>% mutate(Status = 'P') %>% dplyr::rename(ID = 'Hospital Research ID'),
-  new.np.preds %>% dplyr::select(`Hospital Research ID`, Probability, `Relative Risk`, Endoscopy, GEJ.Distance, Pathology, PID) %>% mutate(Status = 'NP') %>% dplyr::rename(ID = 'Hospital Research ID') )
-
-val.preds = val.preds %>% mutate(quants = cut(Probability, breaks=cuts, include.lowest = T))
-
-val.preds = left_join(val.preds, pred.confidence %>% dplyr::select(quants, Risk), by='quants')
-
-ggplot(val.preds, aes(Status, Probability)) + geom_boxplot(aes(group=Status, fill=Status), outlier.shape = NA) + geom_jitter(width = 0.2) + plot.theme
-
-ggplot(val.preds, aes(Pathology, Probability)) + facet_grid(~Status,scales = 'free_x') + geom_boxplot(aes(group=Pathology), outlier.shape = NA) + geom_jitter(width = 0.3, aes(color=Probability, shape=Risk), size=3) + scale_color_gradientn(colors=myPal) + plot.theme
-
-```
-### Validation AUC
-
-```{r, echo=F, warning=F, message=F, fig.width=4, fig.height=3}
-
-val.preds %>% dplyr::select(Status, Probability)
-
-preds = do.call(rbind.data.frame, lapply(vpd, function(df) df[c('Status','Prediction')]))
-vroc = pROC::roc(Status ~ Probability, data=val.preds, ci=T, of='thresholds', transpose=T)
-vroc$model = 'Validation'
-
-roc$model = 'Discovery'
-multi.roc.plot(list(roc, vroc), title='ROC', palette='Dark2') + theme(text=element_text(size=12))
-ggsave('plots/auc_dv.png', plot=p, width=5, height=4, units='in', dpi=300, scale=2)
-```
-
-```{r, echo=F, warning=F, message=F, fig.width=4, fig.height=4}
-aucs = rbind.data.frame(pROC::ci.auc(roc), pROC::ci.auc(vroc))
-colnames(aucs) = c('ci.min', 'auc','ci.max')
-aucs$model = c('Discovery','Validation')
-
-ggplot(aucs, aes(model, auc)) + geom_point(size=2) + ylim(0,1) +
-  geom_errorbar(aes(ymin=ci.min, ymax=ci.max), width=0.1) + coord_flip() + 
-  geom_text(aes(y=ci.min, label=round(ci.min, 2)), vjust=4) +
-  geom_text(aes(y=ci.max, label=round(ci.max, 2)), vjust=-4) +
-#  geom_text(aes(y=auc, label=round(auc, 2)), vjust=-1) +
-  labs(x='', y='AUC', title='AUC with CI') + plot.theme 
-```
-
-```{r}
-dir.create('plots/validation')
-
-val.preds$Patient =  group_indices(val.preds, ID)
-
-val.preds = val.preds %>% group_by(ID) %>% mutate(PID = as.factor(PID)) %>% ungroup %>% group_by(ID, PID) %>% arrange(PID) %>% dplyr::mutate( Endoscopy = Endoscopy + as.integer(PID)  ) %>% ungroup
-
-plist = purrr::map(val.preds %>% filter(Pathology == 'OAC') %>% dplyr::select(ID) %>% distinct %>% pull, function(pt) { 
-  BarrettsProgressionRisk::patientRiskTilesPlot(
-    val.preds %>% filter(ID == pt) %>% mutate(GEJ.Distance = factor(GEJ.Distance)) ) + 
-    labs(title=pt) +
-    theme(legend.position = 'none')
-})
-a = do.call('arrangeGrob', c(plist, top='OAC'))
-ggsave('plots/validation/oac.png', plot=grid.arrange(a), width=3*3, height=4*2,dpi=300)
-
-plist = purrr::map(val.preds %>% filter(Status == 'P' & Pathology != 'OAC') %>% dplyr::select(ID) %>% distinct %>% pull, function(pt) { 
-  BarrettsProgressionRisk::patientRiskTilesPlot(
-    val.preds %>% filter(ID == pt) %>% mutate(GEJ.Distance = factor(GEJ.Distance)) ) + 
-    labs(title=pt) +
-    theme(legend.position = 'none')
-})
-b = do.call('arrangeGrob', c(plist, top='Progressors'))
-ggsave('plots/validation/progessors.png', plot=grid.arrange(b), width=4*4, height=4*2,dpi=300)
-
-plist = purrr::map(val.preds %>% filter(Status == 'NP') %>% dplyr::select(ID) %>% distinct %>% pull, function(pt) { 
-  BarrettsProgressionRisk::patientRiskTilesPlot(
-    val.preds %>% filter(ID == pt) %>% mutate(GEJ.Distance = factor(GEJ.Distance)) ) + 
-    labs(title=pt) +
-    theme(legend.position = 'none')
-})
-c = do.call('arrangeGrob', c(plist, top='Non Progressors', ncol=4))
-grid.arrange(c)
-ggsave('plots/validation/non-progessors.png', plot=grid.arrange(c), width=4*4, height=11*2,dpi=300, limitsize = F)
-
-val.preds %>% group_by(Status, Risk) %>% dplyr::count(Risk) %>% spread(Risk, n) %>% kable() %>% kable_styling(full_width = F)
-```
-
-# Clinical Story
-
-First note here is that our probability distribution for these 20 patients is somewhat flatter than in the training set.
-Likely a consequence of sample size.
-The labels show the proprotions of progressors that fell into each quintile in the training set.
-
-```{r, echo=F, warning=F, message=F}
-
-summary.v = table.path.risk(back.v)
-
-as.data.frame(summary.v) %>% kable() %>% kable_styling(full_width = F)
-
-write.table(summary.v, sep='\t', quote=F, row.names = F)
-```
-
-## Vignettes
-
-```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=5}
-vign.pts = as_tibble(subset(sum.patient.data, Patient %in% c(3, 33, 34, 49, 56, 20), select=c('Status','Patient','age.diagnosis', 'smoking', 'C','M', 'initial.endo','final.endo', 'total.endos','total.samples')))
-colnames(vign.pts) = c('Status','Patient','Age at Diagnosis','Smoker','Circumference', 'Maximal length', 'Intial Endoscopy','Final Endoscopy', 'Total Endoscopies','Total Samples')
-
-## P
-p = vig.plot(df=subset(back, Patient == 34),info= subset(patient.info, Patient == 34),cols=riskCols) + theme(legend.position = 'bottom')
-vigLeg = get.legend(p)
-p = p + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_34.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(subset(back, Patient == 56), subset(patient.info, Patient == 56),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_56.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(subset(back, Patient == 20), subset(patient.info, Patient == 20),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_20.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(subset(back, Patient == 86), subset(patient.info, Patient == 86),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_86.png', plot=p, width=6, height=6,dpi=300)
-
-print(p)
-
-## NP
-p = vig.plot(subset(back, Patient == 3), subset(patient.info, Patient == 3),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_3.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(df=subset(back, Patient == 23),info= subset(patient.info, Patient == 23),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_23.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(subset(back, Patient == 33), subset(patient.info, Patient == 33),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_33.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(subset(back, Patient == 49), subset(patient.info, Patient == 49),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_49.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(subset(back, Patient == 51), subset(patient.info, Patient == 51),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_51.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(subset(back, Patient == 6), subset(patient.info, Patient == 6),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_6.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(subset(back, Patient == 54), subset(patient.info, Patient == 54),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_54.png', plot=p, width=6, height=6,dpi=300)
-
-print(p)
-
-```
-
-### Recommendations
-
- 1.	Immediate RFA: HGD diagnosis, two or more consecutive high risk predictions
- 2.	Recheck 6-12 months: One high risk prediction or aberrant p53 IHC
- 3. Recheck endoscopy 12-24 months: One or more moderate risk predictions
- 4. Regular surveillance 3-5 years: Two or more consecutive low risk predictions
-
-```{r, rx, echo=F, warning=F, message=T, fig.width=8, fig.height=5}
-
-rx<-function(Patient, brks, Prediction, Pathology, Risk, p53.Status) {
-  rules = as.data.frame(matrix(ncol=0, nrow=0))
-  # Consecutive
-  for (i in 1:length(brks)) {
-    risks = table(Risk[i:(i+1)])
-    p53 = table(p53.Status[i:(i+1)])
-    rule = 'None'
-    if ( risks['High'] == 2 | length(which(grepl('HGD|IMC', Pathology[i:(i+1)]))) > 0 ) {
-      rule = 1
-    } else if ( risks['High'] == 1 | p53['1'] > 0 ) {
-      rule = 2
-    } else if ( risks['Moderate'] > 0 ) {
-      rule = 3
-    } else if ( risks['Low'] == 2) {
-      rule = 4
-    }
-  rules = rbind(rules, cbind( as.character(brks[i]), as.character(brks[(i+1)]), rule))
-  }  
-  colnames(rules) = c('t1','t2', 'rule')
-  rules$Patient = Patient
-  rules
-}
-
-cols = intersect(colnames(back), colnames(back.v))
-#cols = cols[-grep('quants|pred\\.ratio|se\\(Risk\\)', cols)]
-allB = rbind.data.frame(back[,cols], back.v[,cols])
-allB$p53.Status = factor(allB$p53.Status, levels=c(0,1), ordered = T)
-
-maxP = allB %>% dplyr::group_by(Patient, Status, brks) %>% dplyr::summarise( 'maxP'=max(Prediction), 'Risk'=Risk[Prediction == max(Prediction)], 'Pathology'=max(Pathology), 'p53'=max(p53.Status,na.rm=T) )
-
-rules = do.call(rbind, lapply(unique(maxP$Patient), function(pt) {
-  r = with(arrange(subset(maxP, Patient == pt), brks), rx(pt, brks,maxP,Pathology,Risk,p53))
-  r$Status = unique(subset(maxP, Patient == pt)$Status)
-  r
-}))
-
-rules$t1 = factor(rules$t1, levels=levels(back$brks), ordered=T)
-rules$t2 = factor(rules$t2, levels=levels(back$brks), ordered=T)
-
-penult = subset(rules, t2 != 'Endpoint' & (t1 == rev(levels(back$brks))[2] | t2 == rev(levels(back$brks))[2] ))
-round(table(subset(penult, Status == 'P')$rule)/nrow(subset(penult, Status == 'P')), 2)
-round(table(subset(penult, Status == 'NP')$rule)/nrow(subset(penult, Status == 'NP')), 2)
-
-second = subset(rules,  t2 != 'Endpoint') %>% 
-  dplyr::group_by(Patient, Status) %>% 
-  dplyr::summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )
-
-table(subset(second, Status == 'P')$rule)
-table(subset(second, Status == 'NP')$rule)
-
-path.rx<-function(Patient, brks, Pathology) {
-  rules = as.data.frame(matrix(ncol=0, nrow=0))
-  # Consecutive
-  for (i in 1:length(brks)) {
-    #risks = table(Pathology[i:(i+1)])
-    rule = 'None'
-    if ( length(which(grepl('HGD|IMC', Pathology[i:(i+1)]))) > 0  |
-         length(which(grepl('LGD', Pathology[i:(i+1)]))) > 1 ) {
-      rule = 1
-    } else if ( length(which(grepl('LGD', Pathology[i:(i+1)]))) > 0 ) {
-      rule = 2
-    } else if ( length(which(grepl('BE|ID', Pathology[i:(i+1)]))) > 1  ) {
-      rule = 3
-    }
-  rules = rbind(rules, cbind( as.character(brks[i]), as.character(brks[(i+1)]), rule))
-  }  
-  colnames(rules) = c('t1','t2', 'rule')
-  rules$Patient = Patient
-  rules
-}
-
-rules2 = do.call(rbind, lapply(unique(maxP$Patient), function(pt) {
-  r = with(arrange(subset(maxP, Patient == pt), brks), path.rx(pt, brks,Pathology))
-  r$Status = unique(subset(maxP, Patient == pt)$Status)
-  r
-}))
-rules2$t1 = factor(rules2$t1, levels=levels(back$brks), ordered=T)
-rules2$t2 = factor(rules2$t2, levels=levels(back$brks), ordered=T)
-
-second = subset(rules2,  t2 != 'Endpoint') %>% 
-  dplyr::group_by(Patient, Status) %>% 
-  dplyr::summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )
-
-table(subset(second, Status == 'P')$rule)
-table(subset(second, Status == 'NP')$rule)
-
-```
-
-## Incidence
-
-Median follow up for our discovery cohort was `r median(sum.patient.data$end.year-sum.patient.data$start.year)`.
-
-Adjust the RR of the orginal model with 1-3.5% offset to get an absolute risk (AR) with the understanding that the cohort was age matched and unbalanced.
-
-```{r echo=F}
-file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
-load(file, verbose=T)
-rm(plots, coefs, cv.patient)
-
-fit = models[[select.alpha]]
-lambda = performance.at.1se[[select.alpha]]$lambda
-
-# number of patients we'd needed to have included to get 45 progressors from a representative population (I think)
-all = rbind(sum.patient.data, sum.validation.info)
-
-# Per year
-# This isn't quite right...we'd want to do something more like a 5 year 
-#pop = round((table(all$Status)['P']/(0.07*100))*100)
-
-```
-
-```{r estimateRisk, echo=F, warning=F, message=F, eval=F}
-# Probability that an individual will progress in a lifetime (20 years) given a rate of progression of 0.07 is about 77%, the probability for a specific individual is less
-1-(0.93^20)
-
-(1-(0.99^3))*15
-(1-(0.99^3))/10
-(1-(0.99^3))/10
-
-# Or probability an individual will progress given a rate of 1%-3.5%
-poft = sapply(seq(0.96,0.99, .01), function(x) {
- sapply(1:5, function(y) 1-(x^y) )
-})
-
-poft = as.data.frame(poft)
-colnames(poft) = paste((1-seq(0.96,0.99, .01))*100, '%', sep='')
-
-rp = apply(poft,1,function(x) cbind(mean(x),min(x), max(x)) )
-
-rownames(rp) = c('mean','min','max')
-colnames(rp) = paste('Year',1:5)
-
-rr = range(as.data.frame(subset(back, Pathology %in% c('BE'), select=c('Patient','Pathology','months.before.final','OR')))$OR)
-
-p = subset(back, Status == 'P' & months.before.final > 6 & OR > 10)$Patient
-subset(sum.patient.data, Patient %in% p) %>% dplyr::group_by(Patient) %>% dplyr::summarise(final.endo-initial.endo)
-
-rrPath = back %>% dplyr::group_by(Status, Pathology) %>% dplyr::summarise( 'mean.RR'=mean(OR),'ci.min'=mean(OR)-(1.96*sd(OR)), 'ci.max'=mean(OR)+(1.96*sd(OR)) )
-
-rrp = ggplot(rrPath, aes(Status, mean.RR)) + #scale_color_gradientn(colors=myPal, limits=c(min(rrPath[,3:5]), max(rrPath[,3:5])),  name='') +
-  facet_wrap(~Pathology, ncol=1) + 
-  geom_point() +  geom_errorbar(aes(ymin=ci.min, ymax=ci.max), width=0.1) + 
-  geom_text(aes(y=ci.min, label=round(ci.min,2)), vjust=2) + geom_text(aes(y=ci.max, label=round(ci.max,2)), vjust=2) +
-  coord_flip() + plot.theme + labs(y='Mean RR', x='')
-rrp
-
-ggsave('plots/relrisk_path.png', plot=rrp, width=4, height=6,dpi=300)
-
-as.data.frame(arrange(subset(back, Status == 'P' & months.before.final > 6 & Pathology %in% c('BE','ID','LGD') & OR > 1), Patient, months.before.final))
-
-ggplot(t(rp), aes(1:5, mean)) +
-  geom_ribbon(aes(ymin=mean/abs(rr[1]),ymax=mean*rr[2]), fill='rosybrown1') +
-  geom_ribbon(aes(ymin=min,ymax=max), fill='grey88') +
-  geom_line() + 
-  geom_text(data=data.frame(t(rp[,5])), aes(x=5,y=min, label='1%')) + geom_text(data=data.frame(t(rp[,5])), aes(x=5,y=max, label='4%')) +
-  geom_line(aes(y=mean*rr[2]), color='red') +
-  geom_line(aes(y=mean/abs(rr[1])), color='blue') +
-  labs(y='P(P)', x='Year', title='Prob. of individual patient progression', subtitle='1-4% over 5 years') + plot.theme
-
-subset(back, OR > 10)
-cor.test(back$OR, as.integer(back$Pathology))
-as.data.frame(subset(back, Pathology %in% c('HGD','IMC'), select=c('Patient','Pathology','months.before.final','OR')))
-
-# as.data.frame(subset(back, Status == 'P' & OR < 2 & OR > -2, select=c('Status','Patient','Pathology','months.before.final','OR')))
-# 
-# hist(subset(back, Pathology %in% c('BE'))$OR)
-
-```
-
-```{r 5yrrisk, echo=F, fig.height=8, fig.width=12}
-cases = table(sum.patient.data$Status)['P']
-
-mn = round((table(sum.patient.data$Status)['P']/(0.01*100))*100)
-m = round((table(sum.patient.data$Status)['P']/(0.0225*100))*100)
-mx = round((table(sum.patient.data$Status)['P']/(0.035*100))*100)
-
-offsetMin = log(cases/mn)
-offsetMean = log(cases/m)
-offsetMax = log(cases/mx)
-
-#binom.test(cases, popL, 0.1)
-# Offset for the linear predictor would then be -log(pop/44)
-# Effectively this is my relative risk for the population
-#offset = log(cases/popL)
-
-# link function provides odds
-#linPred = predict(fit, newx=dysplasia.df, s=lambda, type='link')
-#quantile(linPred)
-
-#df = rbind(dysplasia.df, validation.df)
-#df = dysplasia.df
-
-#p0 = predict(fit, newx=df, s=lambda, type='response')
-#p0 = 1/(1+exp(-linPred)) # logit to get probabilities, same as type='response'
-#p1 = 1/(1+exp(-linPred+abs(offset))) # calculates probs from risk (offset)
-
-cols = c('Hospital.Research.ID','Patient','Status','PID', 'Samplename','Endoscopy.Year','Pathology','Prediction','OR')
-#loopreds = rbind(loopreds[cols], do.call(rbind, vpd)[cols])
-loopreds = do.call(rbind, pg.samp)[cols]
-
-loopreds$P0 = 1/(1+exp(-loopreds$OR))
-
-loopreds$Mean = 1/(1+exp(-loopreds$OR+abs(offsetMean))) # Adjust OR for RR
-loopreds$Min = 1/(1+exp(-loopreds$OR+abs(offsetMin))) # Adjust OR for RR
-loopreds$Max = 1/(1+exp(-loopreds$OR+abs(offsetMax))) # Adjust OR for RR
-
-#ql = cbind.data.frame('y'=quantile(loopreds$OR)[c(2,4)], 'labels'=paste(c(25,75), '%', sep=''))
-
-# From my nested-LOO predictions
-
-ggplot(melt(loopreds[c('Min','Mean','Max')], measure.vars='Mean'), aes(value, fill=variable)) + 
-  geom_density(alpha=0.5) + plot.theme
-
-p = ggplot(melt(loopreds[c('Min','Max')]), aes(value)) + facet_wrap(~variable, ncol=2) +
-  geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + 
-    scale_fill_gradientn(colors = myPal,  name='') + 
-    plot.theme + theme(legend.position = 'none') + 
-    labs(y='n Samples', x='Absolute Risk',
-         title='5-year adjusted risk', subtitle='Adjusted for upper & lower bound: 1-3.5% risk')
-p
-
-ggsave('plots/adj_RR.png', plot=p, width=6, height=6,dpi=300)
-
-# ggplot(loopreds, aes(Min)) +
-#     geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + 
-#     scale_fill_gradientn(colors = myPal,  name='') + 
-#     plot.theme + theme(legend.position = 'none') + 
-#     labs(y='n Samples', x='Absolute Risk',
-#          title='Sample Predictions: 5 year adjusted risk', subtitle='Adjusted for lower bound, 1% risk')
-# 
-# ggplot(loopreds, aes(Max)) +
-#     geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + 
-#     scale_fill_gradientn(colors = myPal,  name='') + 
-#     plot.theme + theme(legend.position = 'none') + 
-#     labs(y='n Samples', x='Absolute Risk',
-#          title='5-year adjusted risk', subtitle='Adjusted for upper bound, 3.5% risk')
-
-```
-
-```{r lifetimeFPR, echo=F, fig.height=6, fig.width=6, eval=F}
-## FPR estimates based on lifetime risks
-# We could consider any NP patient with a minimum of 2 High risk calls?  or High/Moderate Risk?
-# More correctly it should be any patients with concurrent High risk, but that's been pretty stable so two high is good enough for now
-
-bd = bind_rows(back, back.v)
-table(bd$Status)
-
-risksL = subset(bd, Status == 'NP') %>% dplyr::group_by(Patient) %>% dplyr::summarise( 'n.samp'=length(Patient), 'high.risk'=length(which(Risk == 'High')), 'mod.risk'=length(which(Risk == 'Moderate')) )
-
-risksL = subset(risksL, high.risk > 0 | mod.risk > 0)
-
-high = length(subset(risksL, high.risk >= 2 )$Patient)
-#mod = length(subset(risksL, mod.risk >= 2 )$Patient)
-
-# How would I tune it so that the FPR is a lot closer to the real incidence ~1%
-# Current false positive rate is 1.5x higher than the highest incidence ~0.7
-fpr=high*10/44
-x = subset(bd, Status == 'NP' & Patient %in% subset(risksL, high.risk >= 2)$Patient & Risk == 'High')
-(length(unique(subset(x, Prediction > 0.8)$Patient))*10)/44
-
-# How would that affect TPR?
-#pROC::coords(roc, 0.6)
-#pROC::coords(roc, 0.8)
-
-## Per year risk
-## Evaluate over 5 years? How?
-pyr = c()
-for (yr in names(which(table(bd$Endoscopy.Year) > 10))) {
-  risks = subset(bd, Status == 'NP' & Endoscopy.Year == yr) %>% dplyr::group_by(Patient) %>% dplyr::summarise( 'n.samp'=length(Patient), 'high.risk'=length(which(Risk == 'High')), 'mod.risk'=length(which(Risk == 'Moderate')) )
-
-  risks = subset(risks, high.risk > 0 | mod.risk > 0)
-  high = length(subset(risks, high.risk >= 2 )$Patient)
-  mod = length(subset(risks, mod.risk >= 2 )$Patient)
-
-  print(paste(yr,high*15, sep=':'))
-  pyr = c(pyr, high*15)
-}
-
-summary(pyr)
-
-pop*.07
-pop*.02
-pop*.05
-
-```
-
-```{r clonality, echo=F, message=F, warning=F, eval=F}
-## Clonal Heterogenity
-patient.info %>% dplyr::group_by (Hospital.Research.ID, Patient, PID) %>% dplyr::summarise(
-  length(PID)
-)
-
-arms = c(grep(with(chr.lengths, paste(paste(chrom, ':', '1-\\d+000', sep=''), collapse='|')), colnames(dysplasia.df)),
-         grep(with(chr.lengths, paste(paste(chrom, ':', '\\d+000-', chr.length, sep=''), collapse='|')), colnames(dysplasia.df)))
-
-x = do.call(rbind, lapply(unique(patient.info$PID), function(pid) {
-    df = as.matrix(dysplasia.df[subset(info, PID == pid)$Samplename, -grep('cx',colnames(dysplasia.df))])
-    if (ncol(df) > 1) {
-      cbind('arms'=sum(rowSums(df[,arms],na.rm=T))/nrow(df),
-            'seg'=sum(rowSums(df[,-arms],na.rm=T))/nrow(df))
-    } else {
-      cbind('arms'=sum(df[arms,],na.rm=T),
-            'seg'=sum(df[-arms,],na.rm=T))
-    }
-}))
-rownames(x) = unique(patient.info$PID)
-x = x[complete.cases(x),]
-
-hist(x[,'seg'])
-hist(x[,'arms'])
-
-p = pnorm(x[,'seg'], mean(x[,'seg']), sd(x[,'seg']))
-summary(p)
-
-pa = pnorm(x[,'arms'], mean(x[,'arms']), sd(x[,'arms']))
-summary(pa)
-
-x[,'arms'] = scale(x[,'arms'])
-x[,'seg'] = scale(x[,'seg'])
-
-plot(x[unique(subset(patient.info, Patient == 48)$PID), 'arms'])
-plot(x[unique(subset(patient.info, Patient == 48)$PID), 'seg'])
-
-plot(x[unique(subset(patient.info, Patient == 34)$PID), 'arms'])
-plot(x[unique(subset(patient.info, Patient == 34)$PID), 'seg'])
-
-pp = back %>% dplyr::group_by(Patient, Hospital.Research.ID, PID) %>% dplyr::summarise(
-  'max'=max(Prediction), 'avg'=mean(Prediction)
-)
-
-pp = merge(pp, x, by.x='PID', by.y='row.names')
-
-cor.test(pp$seg, pp$max)
-
-```
-
-## Cox model - Time to progress?
-
-```{r cox, warning=F, message=F, echo=F, eval=T}
-library(survival)
-
-cox.pvalue<-function(cox, test='logrank') {
-  pvalue = NULL  
-  csum = summary(cox)
-
-  if (test == 'logrank') {
-    pvalue = csum$sctest[['pvalue']]  
-  } else if (test == 'wald') {
-    pvalue = csum$waldtest[['pvalue']]
-  } else if (test == 'loglik') {
-    pvalue = csum$logtest[['pvalue']]
-  }
-  return(pvalue)
-}
-
-prog.pred = lapply(pg.samp, function(df) {
-  df = df[,c('Hospital.Research.ID', 'Status', 'Patient', 'Age.at.diagnosis', 'PID', 'Status', 'Endoscopy.Year','Initial.Endoscopy', 'Final.Endoscopy', 'Pathology', 'Samplename','Prediction','Prediction.Dev.Resid')] %>% rowwise() %>% dplyr::mutate(
-      'Time.to.Final'=Final.Endoscopy-Endoscopy.Year)
-  df = arrange(df, Endoscopy.Year, Pathology)
-
-  df$Dup.PID = duplicated(df$PID)
-  x = df[!df$Dup.PID,]
-
-  for (i in 1:nrow(x)) {
-    if (i == 1) {
-      x[i,'tstart'] = 0
-      x[i, 'tstop'] = x[(i+1), 'Endoscopy.Year'] - x[i, 'Endoscopy.Year']
-      if (is.na(x[i,'tstop']) || x[i,'tstop'] == 0) x[i,'tstop'] = 0.5
-      next
-    }
-
-    x[i,'tstart'] = x[(i-1), 'tstop']
-
-    # Final sample is different -- will leave it out later
-    if (i == nrow(x)) {
-      x[i, 'tstop'] = x[i,'tstart'] + 0.5
-      break
-    }
-
-    x[i, 'tstop'] =  x[i,'tstart'] + x[(i+1), 'Endoscopy.Year'] - x[i, 'Endoscopy.Year']
-
-    # Samples taken in the same year
-    if ( x[i, 'Endoscopy.Year'] == x[(i+1), 'Endoscopy.Year'] )
-      x[i,'tstop'] = x[i,'tstop'] + 0.5 
-  }
-
-  x$Diagnosis = as.integer(x$Pathology %in% c('HGD','IMC'))
-
-  if (sum(x$Diagnosis) >= 1)
-    x = x[-which(x$Diagnosis == 1),]
-
-  if (nrow(x) > 0 && sum(x$Diagnosis) <= 0 && x$Status == 'P')
-    x[nrow(x), 'Diagnosis'] = 1
-
-  if (nrow(x) > 0) {
-    x$Dup.Samples = ''
-    for (j in 1:nrow(df[df$Dup.PID,]))  {
-      row = df[df$Dup.PID,][j,]
-      x[which(x$PID == row[['PID']]), 'Dup.Samples'] = paste(x[which(x$PID == row[['PID']]), 'Dup.Samples'], row[['Samplename']], sep=',')
-    }
-    x$Dup.Samples = sub('^,','',x$Dup.Samples)
-  }
-
-  return(x)
-})
-## By pathology?
-pp = do.call(rbind, prog.pred)
-
-pp[is.na(pp$tstop), 'tstop'] = 0.5
-pp$Pathology = as.character(pp$Pathology)
-pp$AAD = cut(pp$Age.at.diagnosis, breaks = 3)
-
-pp = merge(pp, back[c('Samplename','Risk', 'quants')], by='Samplename')
-pp$Risk = as.character(pp$Risk)
-
-sfit = survfit( Surv(tstart, tstop, Diagnosis, type='interval') ~ Risk, data=pp, conf.int=0.95, conf.type='log', na.action=na.omit)
-ggsurv(sfit,cens.shape=3,lty.est=1,back.white=F) + xlim(0,5) + labs(y='Diagnosed', title='~Risk')
-
-cox = coxph( Surv(tstart, tstop, Diagnosis) ~ Risk, data=pp) 
-ggsurv(survfit(cox),cens.shape=3,lty.est=1,back.white=F,
-      main="coxph(Survival, Diagnosis) ~ Risk" ) + 
-      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),2),"\n", nrow(pp), 'samples'))
-#summary(cox)
-
-sfit = survfit( Surv(tstart, tstop, Diagnosis, type='interval') ~ AAD, data=pp, conf.int=0.95, conf.type='log', na.action=na.omit)
-ggsurv(sfit,cens.shape=3,lty.est=1,back.white=F) + xlim(0,5) + labs(y='Diagnosed', title='~Age') 
-cox = coxph( Surv(tstart, tstop, Diagnosis) ~ AAD, data=pp) 
-ggsurv(survfit(cox),cens.shape=3,lty.est=1,back.white=F,
-      main="coxph(Survival, Diagnosis) ~ Age at diagnosis" ) + 
-      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),2),"\n", nrow(pp), 'samples'))
-#summary(cox)
-
-sfit = survfit( Surv(tstart, tstop, Diagnosis, type='interval') ~ Pathology, data=pp, conf.int=0.95, conf.type='log', na.action=na.omit)
-ggsurv(sfit,cens.shape=3,lty.est=1,back.white=F) + xlim(0,5) + labs(y='Diagnosed', title='~Pathology')
-cox = coxph( Surv(tstart, tstop, Diagnosis) ~ Pathology, data=pp) 
-ggsurv(survfit(cox),cens.shape=3,lty.est=1,back.white=F,
-      main="coxph(Survival, Diagnosis) ~ Pathology" ) + 
-      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),2),"\n", nrow(pp), 'samples')) + xlim(0,5)
-#summary(cox)
-
-## All genomic regions
-sample.pred = merge( do.call(rbind.data.frame, prog.pred), dysplasia.df, by.x='Samplename', by.y='row.names' )
-sample.pred$Pathology = as.character(sample.pred$Pathology)
-
-gcols = grep('\\d+:', colnames(sample.pred), value=T)
-sample.pred = sample.pred[c('tstart', 'tstop','Diagnosis','Pathology','Age.at.diagnosis',gcols)]
-
-sample.pred$Age.bracket = as.character(cut(sample.pred$Age.at.diagnosis, breaks=3))
-
-# all
-#cox = coxph( Surv(tstart, tstop, Diagnosis) ~., data=sample.pred)
-#ggsurv(survfit(cox))
-#summary(cox)
-```
-### Cox GLM selected features
-
-```{r echo=F, warning=F, echo=F}
-# glm selected features + Path
-cox2 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', 'Pathology',  grep('^\\d+:', features$label, value=T))]) 
-ggsurv(survfit(cox2),cens.shape=3,lty.est=1,back.white=F,
-      main="coxph(Survival, Diagnosis) ~ GLM features + Pathology" ) + 
-      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox2),2),"\n", nrow(pp), 'samples')) 
-#summary(cox2)
-
-cox2 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis','Age.bracket', grep('^\\d+:', features$label, value=T))]) 
-ggsurv(survfit(cox2),cens.shape=3,lty.est=1,back.white=F,
-      main="coxph(Survival, Diagnosis) ~ GLM features + Age" ) + 
-      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox2),2),"\n", nrow(pp), 'samples')) 
-#summary(cox2)
-
-cox2 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis',grep('^\\d+:', features$label, value=T))]) 
-ggsurv(survfit(cox2),cens.shape=3,lty.est=1,back.white=F,
-      main="coxph(Survival, Diagnosis) ~ GLM features" ) + 
-      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox2),2),"\n", nrow(pp), 'samples')) 
-#summary(cox2)
-
-# chr arms only
-options(scipen=999) # turn sci notation off
-coxArms = coxph(Surv(tstart, tstop, Diagnosis) ~., data=sample.pred[,c('tstart', 'tstop','Diagnosis', with(arm.loc, paste(chrom, ':', start, '-', end, sep='')))])
-ggsurv(survfit(coxArms),cens.shape=3,lty.est=1,back.white=F,
-      main="coxph(Survival, Diagnosis) ~ chr arms" ) + 
-      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(coxArms),2),"\n", nrow(pp), 'samples')) 
-#summary(coxArms)
-
-# chr 17 only
-cox17 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', grep('^17:',colnames(sample.pred), value=T))]) 
-ggsurv(survfit(cox17),cens.shape=3,lty.est=1,back.white=F,
-      main="coxph(Survival, Diagnosis) ~ chr17" ) + 
-      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox17),2),"\n", nrow(pp), 'samples')) 
-#summary(cox17)
-
-cox11 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', grep('^11:',colnames(sample.pred), value=T))]) 
-ggsurv(survfit(cox11))
-#summary(cox11)
-
-cox13 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', grep('^13:',colnames(sample.pred), value=T))]) 
-ggsurv(survfit(cox13))
-#summary(cox13)
-
-```
-
-## CoxRFX
-
-```{r echo=F, message=F, warning=F, eval=T}
-plot.waldtest<-function(wt) {
-  p = ggplot(wt, aes(coef.x, coef.y, color=p.value<0.05))
-  if (length(grep('hazards', colnames(wt))) > 0) {
-      p = p + geom_point(alpha=0.5, aes(size=hazards)) + 
-        geom_point(aes(coef.x, coef.y, color=p.value<0.05, size=hazards), show.legend = F) +
-        #scale_shape_manual(values=c(17,18), na.value=21) +
-        scale_size(range=c(2, 10), breaks=c(0, 5, 10, 15)) +
-        #geom_text_repel( aes( label=label), show.legend=F) + 
-        geom_text_repel(data=subset(wt, p.value < 0.05), aes(coef.x,coef.y, label=label), show.legend=F) + 
-        geom_text_repel(data=subset(wt, is.na(stability)), aes(coef.x, coef.y, label=label, color=p.value<0.05), show.legend=F, color='grey39') 
-  } else {
-    p = p + geom_point(alpha=0.5) + 
-      geom_text_repel(data=subset(wt, p.value < 0.05), aes(coef.x,coef.y, label=label), show.legend=F)  
-  }
-  return(p + labs(title='Wald test'))
-}
-```
-
-```{r echo=T, message=F, warning=F, fig.height=8, fig.width=8, eval=F}
-
-library(CoxHD)
-
-coxDF = sample.pred
-
-pathology = do.call(cbind.data.frame, lapply(unique(coxDF$Pathology), function(p) {
-  cbind( as.integer(coxDF$Pathology == p))
-}))
-colnames(pathology) = unique(coxDF$Pathology)
-
-# age = do.call(cbind.data.frame, lapply(unique(coxDF$Age.at.diagnosis), function(p) {
-#   cbind( as.integer(coxDF$Age.at.diagnosis == p))
-# }))
-# colnames(age) = unique(coxDF$Age.at.diagnosis)
-age = coxDF$Age.at.diagnosis
-
-## GLM regions only
-coxDF = cbind(coxDF[,1:3],pathology[,c('BE','ID')], age, coxDF[,grep('^\\d+:', features$label, value=T)])
-coxDF$Pathology = NULL
-
-gcols = colnames(coxDF)
-cols = grep('^\\d+', colnames(coxDF))
-colnames(coxDF)[cols] = paste('chr', gsub(':|-','.', grep('^\\d+', colnames(coxDF), value=T)), sep='')
-
-coxSM = cbind(coxDF[,c(1:3)], StandardizeMagnitude(coxDF[,-c(1:3)]))
-
-#surv = with(coxDF, Surv(tstart, tstop, Diagnosis))
-surv = with(coxSM, Surv(tstart, tstop, Diagnosis))
-groups = factor(c(rep('clinical',3),c(rep('genomic', length(grep('\\d+', colnames(coxDF)))))))
-rfx = CoxRFX(coxSM[,-c(1:3)], surv, groups, verbose=T, nu=1)
-summary(rfx)
-
-plot(rfx)
-
-wt = WaldTest(rfx)
-survConcordance(surv ~ predict(rfx))
-
-wt$label = sub('\\.', '-', sub('\\.', ':', sub('chr', '', rownames(wt))))
-wt = merge(wt, features[,c('label','hazards', 'coef','stability')], by='label', all=T)
-
-wt[is.na(wt$hazards), c('hazards', 'coef.y')] = 0
-plot.waldtest(wt) + labs(x='RFX coef',y='GLM coef', title='RFX vs GLM coefficients', subtitle='Labeled points have a significant p-value in RFX (except BE/ID)')
-
-options(scipen = 999)
-# arms only
-arms = with(arm.loc, paste(chrom, ':', start, '-', end, sep=''))
-coxDfArms = coxDF[,arms]
-colnames(coxDfArms) = paste('chr', gsub(':|-','.', grep('^\\d+', colnames(coxDfArms), value=T)), sep='')
-
-surv = with(coxDF, Surv(tstart, tstop, Diagnosis))
-groups = factor(c(rep('clinical',3), rep('genomic', ncol(coxDfArms))))
-
-rfx = CoxRFX( cbind.data.frame(coxSM[,c('ID','BE','age_10')], coxDfArms), surv, groups, verbose=T, nu=1)
-summary(rfx)
-
-plot(rfx)
-wt = WaldTest(rfx)
-survConcordance(surv ~ predict(rfx))
-
-```
+
+val.preds = val.preds %>% mutate(Pathology = factor(Pathology, levels=c('NDBE','GM','ID','LGD'), ordered=T)) %>% 
+  dplyr::rename(Samplename = 'Sample') %>% 
+  mutate(Final.Endoscopy = `Diagnostic Endoscopy`, Endoscopy.Year = Endoscopy, Block = factor(Block, ordered=T))
+
+rp = plot.risk.by.path(val.preds) + labs(title='Validation samples predicted by pathology')  
+rp
+ggsave(filename='plots/val_pred_by_path.png', plot=rp, width=9, height=7, units='in', dpi=300)
+
+get.year<-function(x) {
+  as.numeric(format(as.Date(x), '%Y'))
+}
+
+val.preds = do.call(bind_rows, lapply(unique(val.preds$`Hospital Research ID`) %>% 
+                            map( function(p) val.preds %>% 
+                                   filter(`Hospital Research ID` == p) %>% 
+                                   mutate_at(vars(Endoscopy.Year, Final.Endoscopy), list(~get.year(.)) )
+                                 ), transform.by.time, info=val.info) )
+
+v.prog = val.preds %>% filter(Status == 'P') %>% mutate(
+  brks = cut(months.before.final, breaks=sq, 
+                 labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-c(length(sq)-1,length(sq))], '>120'),include.lowest = T, ordered_result = T),
+) %>% dplyr::mutate(brks = factor(brks, rev(levels(brks))))
+
+sq = c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12)
+endos = v.prog %>% dplyr::group_by(Status, months.before.final, brks, PID) %>% 
+  dplyr::summarise('pred.endo'=ifelse(length(which(Risk == 'High')) > 0, 1, 0), 
+              'n.endos'=length(unique(PID)), 
+              'n.samples'=length(PID),
+              'pred.samples'=length(which(Risk == 'High')),
+              'max.path'=max(Pathology),
+              'max.prediction' = max(Probability)
+              ) %>% ungroup %>% 
+  dplyr::mutate(brks = cut(months.before.final, include.lowest = T, ordered_result = T,
+     breaks=sq,  labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))) %>% 
+  dplyr::mutate(brks = factor(brks, rev(levels(brks))))
+
+endos = endos %>% mutate(brks = recode_factor(brks, '96-120' = '>96', '>120'='>96'))
+
+endos.m = endos %>% dplyr::filter(Status == 'P') %>% dplyr::group_by(brks) %>% 
+  dplyr::summarise( 'pred.r'=sum(pred.endo)/sum(n.endos), 'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos),'mean(P)'=mean(max.prediction),'sem(P)'=sd(max.prediction)/sqrt(length(brks)))
+
+rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) +
+  geom_bar(stat='identity', fill='darkred', alpha=0.8, show.legend = F) + 
+  geom_errorbar(color='grey21', width=0.2) +
+  geom_text(aes(y=3, label=paste(round(pred.r*100), '%', sep='')), color='white') +
+  geom_text(aes(y=pred.r*100-5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
+  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') + coord_cartesian(ylim = c(0,100)) + 
+  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12)) 
+rp
+
+dir.create('plots/validation/heatmaps/P',recursive = T, showWarnings = F)
+dir.create('plots/validation/heatmaps/NP',recursive = T, showWarnings = F)
+
+for (pt in unique(v.prog$`Hospital Research ID`)) {
+
+  patient = v.prog %>% filter(`Hospital Research ID` == pt) %>% 
+    dplyr::select(`Hospital Research ID`,Status, Endoscopy, months.before.final, Pathology,Block,Risk) %>% 
+    dplyr::rename(GEJ.Distance = Block, Patient = 'Hospital Research ID')
+
+  p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final', 'rev') + labs(x='months', y='location') + theme(legend.position = 'none',plot.title = element_text(face = 'bold', size=16))
+
+    ggsave(paste0('plots/validation/heatmaps/',unique(patient$Status),'/pt_',pt,'.png'), plot = p, height=length(unique(patient$GEJ.Distance))+1, width=length(unique(patient$months.before.final))+1, units='in', dpi=300, limitsize=F)
+
+    return(p)
+}
+
+```

### 2019-11-12 lib/common_plots.R

@@ -366,15 +366,16 @@ plot.risk.by.path<-function(preds) {
 
 
 transform.by.time<-function(pt, info) {
-  pt = pt %>% dplyr::group_by(PID, Samplename) %>% dplyr::mutate( 'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12) %>% 
-    arrange(Pathology, PID, Endoscopy.Year) %>% ungroup
+  pt = pt %>% dplyr::group_by(PID, Samplename) %>%  
+    dplyr::mutate( 'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12) %>% 
+    arrange(Pathology, PID, Endoscopy.Year) %>% ungroup 
   
   tb = with(pt, table(Endoscopy.Year, PID))
   tb[tb>0] = 1
-  for (yr in names(which(rowSums(tb) > 1))) {
-    tmp = pt[which(pt$Endoscopy.Year == yr),]
-    tmp = tmp %>% dplyr::group_by(PID) %>% summarise(max(Pathology))
-    
+  for (yr in names(which(rowSums(tb,na.rm=T) > 1))) {
+    tmp = filter(pt, Endoscopy.Year == yr)
+    tmp = tmp %>% dplyr::group_by(PID) %>% dplyr::summarise(max(Pathology))
+
     match = which(pt$Endoscopy.Year == yr & pt$PID == as.character(tmp[which.min(tmp$`max(Pathology)`), 'PID']))
     pt[match,'months.before.final'] = pt[match,'months.before.final'] + 6
   }

### 2019-11-12 Paper_resubmit.Rmd
forgot to save before last commit...

@@
 ```{r, echo=F}
-p.diag = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/validation/progressor_diagnosis.xlsx') %>% 
-  dplyr::mutate(`Hospital Research ID` = gsub('/', '_', `Hospital Research ID`), `Weeks Follow Up` = as.double(difftime(`Diagnostic Endoscopy`, `Sequenced Endoscopy`, units='weeks')))
-
-patient.file = '~/Data/Ellie/QDNAseq/validation/sWGS_validation_batches.xlsx'
-sheets = readxl::excel_sheets(patient.file)[8:14]
-all.val.info = do.call(bind_rows, lapply(sheets, function(s) {
-  readxl::read_xlsx(patient.file, s) %>% dplyr::select(`Hospital Research ID`, matches('Status|Endoscopy|Block|Path'), `Sample Type`, `SLX-ID`, `Index Sequence`,Cohort) %>% 
-    dplyr::mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% dplyr::filter(!is.na(`SLX-ID`))
-}))
-
+#p.diag = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/validation/progressor_diagnosis.xlsx') %>% 
+#  dplyr::mutate(`Hospital Research ID` = gsub('/', '_', `Hospital Research ID`), `Weeks Follow Up` = as.double(difftime(`Diagnostic Endoscopy`, `Sequenced Endoscopy`, units='weeks')))
 pastefun<-function(x) {
@@
 
-all.val.info = all.val.info %>% rowwise %>% 
-  mutate_at(vars(`SLX-ID`), list(pastefun) ) %>% ungroup %>%
-  #mutate(`Hospital Research ID` = str_replace_all( str_remove_all(`Hospital Research ID`, " "), '/', '_'), `Index Sequence` = str_replace_all(`Index Sequence`, 'tp', '')) %>% 
+patient.file = '~/Data/Ellie/QDNAseq/validation/sWGS_validation_batches.xlsx'
+
+oac.info = readxl::read_xlsx(patient.file, 'OAC Test Samples') %>% 
+  dplyr::mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% dplyr::filter(!is.na(`SLX-ID`)) %>%
   mutate(Samplename = paste(`SLX-ID`, `Index Sequence`, sep='.')) %>%
-  separate(`Block ID`, c('PID','Block'), sep='-|[:blank:]', remove=F) %>% dplyr::rename(PathID = `Block ID`)
-
-all.val.info$Block = sapply(strsplit(all.val.info$Block, ''), function(x) {
+  dplyr::select(-matches('SLX|Index')) %>%
+  mutate(Status = factor(Status, levels=levels(pg.samp$Status)), Block = factor(Block, levels=c('E','1','2'), ordered=T))
+
+val.info = readxl::read_xlsx(patient.file, 'Final Validation Samples') %>% 
+  dplyr::mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% dplyr::filter(!is.na(`SLX-ID`)) %>%
+  mutate(Samplename = paste(`SLX-ID`, `Index Sequence`, sep='.')) %>%
+  separate(`Block ID`, c('PID','Block'), sep='-|[:blank:]', remove=F) %>% dplyr::rename(PathID = `Block ID`) %>%
+  dplyr::select(-matches('SLX|Index'), -`Path Notes`) %>%
+  mutate(Status = factor(Status, levels=levels(pg.samp$Status))) %>% ungroup %>% group_by(`Hospital Research ID`) %>% 
+  dplyr::mutate(`Weeks Follow Up` = as.double(difftime(`Last Endoscopy`, `Endoscopy`, units='weeks')))
+
+val.info$Block = sapply(strsplit(val.info$Block, ''), function(x) {
   if (is.na(x)) return(1)
@@
 
-all.val.info = left_join(all.val.info, p.diag, by=c('Hospital Research ID', 'PathID'='Block ID')) %>% group_by(`Hospital Research ID`) %>% 
-  dplyr::mutate(`Weeks Follow Up` = as.double(difftime(`Diagnostic Endoscopy`, `Sequenced Endoscopy`, units='weeks')))
-
-val.info = all.val.info %>% filter(Pathology != 'OAC')
-
 bind_rows(val.info %>% dplyr::select(`Hospital Research ID`, Status) %>% distinct %>% group_by(Status) %>% tally %>% 
@@
 
+residuals = do.call(bind_rows, purrr::map(list.files( dirname(vdir), 'residuals.tsv', full.names = T, recursive = T ), function(f) {
+  read_tsv(f, col_types = 'ccddddl')
+}))
+
+failedQC = residuals %>% filter(varMAD_median > 0.008)
+
 all.val.preds = do.call(bind_rows, purrr::map(list.files(vdir, 'predictions.tsv', full.names = T, recursive = T), function(f) {
     pd = read_tsv(f, col_types = 'cddcccDcccc') %>% dplyr::select(`Hospital Research ID`, Sample, Probability, `Relative Risk`)
-    #left_join(pd, all.val.info, by=c('Hospital Research ID', 'Sample'='Samplename', 'Block ID')) 
     return(pd)
@@
 
-all.val.preds = left_join(all.val.preds, all.val.info, by=c('Hospital Research ID', 'Sample'='Samplename'))
-
-val.preds = all.val.preds %>% dplyr::filter(Sample %in% val.info$Samplename)
-
-oac.preds = all.val.preds %>% filter(Pathology == 'OAC')
-#all.val.preds %>% filter(Cohort == 'Old progressor scrolls') %>% group_by(Status, Risk) %>% tally
-#val.info %>% filter(Cohort != 'Old progressor scrolls' & Pathology != 'OAC') %>% group_by(Status) %>% tally
+oac.preds = all.val.preds %>% filter(`Hospital Research ID` %in% unique(oac.info$`Hospital Research ID`))
+
+all.val.preds = all.val.preds %>% filter( !Sample %in% failedQC$samplename &  
+                                            !`Hospital Research ID` %in% unique(oac.info$`Hospital Research ID`) )
+
+val.preds = left_join(all.val.preds, val.info, by=c('Hospital Research ID', 'Sample'='Samplename')) %>% dplyr::select(Patient, everything())
+
+val.preds = val.preds %>% mutate(Pathology = factor(Pathology, levels=c('NDBE','GM','ID','LGD'), ordered=T)) %>% 
+  mutate(Final.Endoscopy = `Last Endoscopy`, Endoscopy.Year = Endoscopy, Block = factor(Block, ordered=T))
+
 ```
@@
 ```{r}
-
-val.preds = val.preds %>% mutate(Pathology = factor(Pathology, levels=c('NDBE','GM','ID','LGD'), ordered=T)) %>% 
-  dplyr::rename(Samplename = 'Sample') %>% 
-  mutate(Final.Endoscopy = `Diagnostic Endoscopy`, Endoscopy.Year = Endoscopy, Block = factor(Block, ordered=T))
-
 rp = plot.risk.by.path(val.preds) + labs(title='Validation samples predicted by pathology')  
@@
 
-val.preds = do.call(bind_rows, lapply(unique(val.preds$`Hospital Research ID`) %>% 
-                            map( function(p) val.preds %>% 
-                                   filter(`Hospital Research ID` == p) %>% 
+val.preds = dplyr::rename(val.preds, 'Samplename'='Sample')
+
+val.preds = do.call(bind_rows, lapply(unique(val.preds$Patient) %>% 
+                            map( function(p) val.preds %>% filter(Patient == p) %>% 
                                    mutate_at(vars(Endoscopy.Year, Final.Endoscopy), list(~get.year(.)) )
                                  ), transform.by.time, info=val.info) )
+
+dir.create('plots/validation/heatmaps/P',recursive = T, showWarnings = F)
+dir.create('plots/validation/heatmaps/NP',recursive = T, showWarnings = F)
+
+for (pt in unique(val.preds$Patient)) {
+print(pt)
+  patient = val.preds %>% filter(Patient == pt) %>% 
+    dplyr::select(Patient, Status, Endoscopy, months.before.final, Pathology,Block,Risk) %>% 
+    dplyr::rename(GEJ.Distance = Block)
+
+  p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final', 'rev') + labs(x='months', y='location') + theme(legend.position = 'none',plot.title = element_text(face = 'bold', size=16))
+
+    ggsave(paste0('plots/validation/heatmaps/',unique(patient$Status),'/pt_',pt,'.png'), plot = p, height=length(unique(patient$GEJ.Distance))+1, width=length(unique(patient$months.before.final))+1, units='in', dpi=300, limitsize=F)
+}
 
@@
 
-dir.create('plots/validation/heatmaps/P',recursive = T, showWarnings = F)
-dir.create('plots/validation/heatmaps/NP',recursive = T, showWarnings = F)
-
-for (pt in unique(v.prog$`Hospital Research ID`)) {
-
-  patient = v.prog %>% filter(`Hospital Research ID` == pt) %>% 
-    dplyr::select(`Hospital Research ID`,Status, Endoscopy, months.before.final, Pathology,Block,Risk) %>% 
-    dplyr::rename(GEJ.Distance = Block, Patient = 'Hospital Research ID')
-
-  p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final', 'rev') + labs(x='months', y='location') + theme(legend.position = 'none',plot.title = element_text(face = 'bold', size=16))
-
-    ggsave(paste0('plots/validation/heatmaps/',unique(patient$Status),'/pt_',pt,'.png'), plot = p, height=length(unique(patient$GEJ.Distance))+1, width=length(unique(patient$months.before.final))+1, units='in', dpi=300, limitsize=F)
-
-    return(p)
-}
-
-```
+```

### 2019-11-18 Paper_resubmit.Rmd
checking snp data against new models

@@
 ---
-title: "Untitled"
+title: "Paper"
 author: "Sarah Killcoyne"
-date: "07/11/2019"
+date: "2019-11-13"
 output: 
@@
 all.patient.info = read.patient.info(patient.file, demo.file, set='All')$info
-patient.info = all.patient.info %>% dplyr::filter(Set == 'Training')
+patient.info = all.patient.info
 
@@
 
+i.thresholds = do.call(bind_rows, purrr::map(insilico.rocs, function(r) pROC::coords(r, 'best', transpose=F) %>% as_tibble %>% mutate(AUC = r$auc, model=r$model, is=as.character(r$is)))) %>%
+  rename_at(vars(matches('thresh|spec|sens')), funs(str_to_title))
+
 #summary(filter(df, model == 'in-silico')$AUC)
@@
 p
-ggsave(filename='plots/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)
+
+blues = RColorBrewer::brewer.pal(5,'Blues')
+p = ggplot(df, aes(Specificity, Sensitivity)) + 
+    geom_line(color=blues[2]) + 
+    geom_line(data=filter(df, is == 'discovery'), color=blues[5], size=1) +
+    geom_line(data=df %>% filter( (AUC == max(AUC) ) & is != 'discovery'), color=blues[4], linetype = 'dotted', size=0.75) +
+    geom_line(data=df %>% filter( (AUC == min(AUC) ) & is != 'discovery'), color=blues[4], linetype = 'dotted', size=0.75) +
+    scale_x_reverse() +
+    geom_label(data=filter(i.thresholds, AUC %in% range(AUC) | is == 'discovery'), aes(label=paste0('AUC\n',round(AUC*100))), color='black') +
+    #scale_color_brewer(palette='Blues') +
+    labs(title='ROC AUC', subtitle='in-silico splits', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
+    theme(legend.position = 'none')
+p
+ggsave(filename='plots/auc/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)
 
@@
 roc.plot(p53roc, 'p53 IHC')
+ggsave('plots/auc/p53_roc.png', plot=roc.plot(p53roc, 'p53 IHC'), width=6, height=6, units="in", limitsize=F, dpi = 300)
 ```
@@
 
-ggsave('plots/path_p53_cn_auc.png', plot=comp, width=6, height=8, units="in", limitsize=F, dpi = 300)
-
+ggsave('plots/auc/path_p53_cn_auc.png', plot=comp, width=6, height=8, units="in", limitsize=F, dpi = 300)
+
+```
+
+```{r}
+roc$model = "Per Sample"
+
+m = rbind.data.frame(
+      get.roc.stat(roc),
+      get.roc.stat(rocEndoM),get.roc.stat(rocEndoA),
+      get.roc.stat(rocPtM), get.roc.stat(rocPtA),
+      get.roc.stat(rocFirstEndo), get.roc.stat(rocLastEndo))
+m$model = factor(m$model)
+
+cols = brewer.pal(11, 'Paired')
+dodge = position_dodge(width=0.9)
+mm = melt(subset(m, Measure == 'AUC'), measure.vars='value')
+p = ggplot( arrange(mm, CI.min), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=model)) + 
+  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='grey39' ) + 
+  geom_text(aes(y=CI.min*100, label=paste(round(value,2)*100,'%',sep='')), position=dodge, vjust=3)  +
+  scale_fill_manual(values=cols, name='') + plot.theme + 
+  theme(legend.position='none', axis.text.x=element_text(angle=45, hjust=1)) + labs(x='', y='AUC', title='ROC Comparison') 
+p
+
+ggsave('plots/auc/multi_auc.png', plot=p, height=6, width=5, units='in',dpi=300)
 ```
@@
 
+```{r, eval=F}
+file = paste(cache.dir, 'model_data.Rdata', sep='/')
+load(file, verbose=T)
+
+vdir ='~/Data/BarrettsProgressionRisk/Analysis/validation/pcf_perPatient/50kb'
+v.tiles = do.call(bind_rows, purrr::map(list.files(vdir, '5e06_tiled_segvals.txt', full.names = T, recursive = T), function(f) {
+  read_tsv(f, col_types = cols(.default=col_double(), 'X1'=col_character()))
+})) %>% dplyr::rename(Samplename = 'X1')
+v.tiles = filter(v.tiles, Samplename %in% val.info$Samplename)
+v.tiles = v.tiles[-which(duplicated(v.tiles$Samplename))[2],]
+
+v.arm.tiles = do.call(bind_rows, purrr::map(list.files(vdir, 'arms_tiled_segvals.txt', full.names = T, recursive = T), function(f) {
+  read_tsv(f, col_types = cols(.default=col_double(), 'X1'=col_character()))
+})) %>% dplyr::rename(Samplename = 'X1')
+v.arm.tiles = filter(v.arm.tiles, Samplename %in% val.info$Samplename)
+v.arm.tiles = v.arm.tiles[-which(duplicated(v.arm.tiles$Samplename))[2],]
+
+for (i in 1:length(z.mean))
+  v.tiles[, i+1] = BarrettsProgressionRisk:::unit.var(v.tiles[,i+1], z.mean[i], z.sd[i])
+
+for (i in 1:length(z.arms.mean))
+  v.arm.tiles[, i+1] = BarrettsProgressionRisk:::unit.var(v.arm.tiles[,i+1], z.arms.mean[i], z.arms.sd[i])
+
+vt = dplyr::select(v.tiles, -Samplename) %>% as.matrix; rownames(vt) = v.tiles$Samplename
+at = dplyr::select(v.arm.tiles, -Samplename) %>% as.matrix; rownames(at) = v.arm.tiles$Samplename
+
+v.df = BarrettsProgressionRisk:::subtractArms(vt, at)
+
+valCNdf = as_tibble(t(v.df), rownames='locs') %>% set_names(c('locs',val.info$Samplename)) %>%
+  separate(locs, c('chr','start','end'), sep=':|-', remove=F) %>%   
+  dplyr::mutate_at(vars(start, end), as.numeric) %>%
+  #dplyr::mutate( seg.type = '5MB' ) %>%
+  dplyr::mutate( seg.type = ifelse( (start %in% arm.loc$start & end %in% arm.loc$end), 'arms','5MB'  ) ) %>%
+  melt(id.vars=c('locs','chr','start','end', 'seg.type')) %>% as_tibble() %>%
+  left_join(chr.lengths %>% dplyr::select(chrom,chr.length), by=c('chr'='chrom')) %>%
+  left_join(val.info %>% dplyr::select(matches('Hospital|Patient|Samplename|Pathology|Status')), by=c('variable'='Samplename')) %>%
+  dplyr::mutate(Status = factor(recode(Status, NP='Non-Progressor',P='Progressor'), ordered=T)) %>% 
+  dplyr::mutate( 
+    cn = factor(unlist(purrr::map(.x=value, .f=get.cn.state)),  levels=c('loss','neutral','gain'), ordered = T),
+    chr = factor(chr, levels = chr.lengths$chrom, ordered=T))
+
+n.status <- c('Progressor'=paste0('Progressor (n=',length(unique(subset(valCNdf, Status == 'Progressor')$variable)), ')'),
+              'Non-Progressor'=paste0('Non-Progressor (n=',length(unique(subset(valCNdf, Status == 'Non-Progressor')$variable)),')'))
+
+#cn.mtn.plot(filter(valCNdf, seg.type == '5MB'), labeller(Status=n.status, chr=label_value))
+
+```
+
 ## Per sample (alpha = 0.9)
@@
 roc.plot(roc) + labs(title='Validation cohort')
-ggsave('plots/auc/validatation_roc.png', plot=roc.plot(roc) + labs(title='Validation cohort'), width=6, height=4, units='in',dpi=300)
+ggsave('plots/validation/validatation_roc.png', plot=roc.plot(roc) + labs(title='Validation cohort'), width=6, height=4, units='in',dpi=300)
 
@@
 rp
-ggsave(filename='plots/val_pred_by_path.png', plot=rp, width=9, height=7, units='in', dpi=300)
+ggsave(filename='plots/validation/val_pred_by_path.png', plot=rp, width=9, height=7, units='in', dpi=300)
 
@@
 ```
+
+# Clinical Story
+
+## Vignettes
+
+```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=5}
+sum.patient.data = sum.patient.data %>% mutate(smoking_str = recode(smoking, 'Y'='smoker', 'N'=''))
+sum.patient.data$smoking_str = replace_na(sum.patient.data$smoking_str, '')
+
+create.desc<-function(pt,age,sex,smk,len,date) {
+  str =  paste0('Patient ', pt, ': ', age, sex)
+  if (smk != '' & !is.na(smk)) str = paste0(str, ' ,', smk)
+  str = paste0(str,'\n')
+
+  if (len != '' & !is.na(len)) str = paste0(str, len,'cm length ')
+
+  str = paste0(str, 'BE diagnosed ', date)
+
+  return(str)
+}
+sum.patient.data$desc = sum.patient.data %>% group_by(Patient) %>% rowwise() %>% dplyr::summarise(desc = create.desc(Patient, age.diagnosis, gender, smoking_str, M, initial.endo))
+
+for (pt in c(34,56,20,86,3,23,33,49,51,6,25,68)) {
+  length = sum.patient.data %>% ungroup %>% filter(Patient == pt) %>% dplyr::summarise(years = final.endo-initial.endo) %>% pull
+  height = as.numeric(max( pg.samp %>% ungroup %>% filter(Patient == pt) %>% dplyr::select(Block) %>% pull ) ) + 1
+
+  p = vig.plot(df=subset(pg.samp, Patient == pt),info= subset(patient.info, Patient == pt),cols=riskCols) + 
+    theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) 
+  vigLeg = get.legend(p)
+  p = p + theme(legend.position = 'none')
+  ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p, width=length, height=height,dpi=300)
+}
+ggsave('plots/vignette/legend.png', vigLeg, width=8, height=1, dpi=300)
+
+```
+
+### Recommendations
+
+ 1.	Immediate RFA: HGD diagnosis, two or more consecutive high risk predictions
+ 2.	Recheck 6-12 months: One high risk prediction or aberrant p53 IHC
+ 3. Recheck endoscopy 12-24 months: One or more moderate risk predictions
+ 4. Regular surveillance 3-5 years: Two or more consecutive low risk predictions
+
+```{r, rx, echo=F, warning=F, message=T, fig.width=8, fig.height=5}
+
+rx<-function(Patient, brks, Prediction, Pathology, Risk, p53.Status) {
+  rules = as.data.frame(matrix(ncol=0, nrow=0))
+  # Consecutive
+  for (i in 1:length(brks)) {
+    risks = table(Risk[i:(i+1)])
+    p53 = table(p53.Status[i:(i+1)])
+    rule = 'None'
+    if ( risks['High'] == 2 | length(which(grepl('HGD|IMC', Pathology[i:(i+1)]))) > 0 ) {
+      rule = 1
+    } else if ( risks['High'] == 1 | p53['1'] > 0 ) {
+      rule = 2
+    } else if ( risks['Moderate'] > 0 ) {
+      rule = 3
+    } else if ( risks['Low'] == 2) {
+      rule = 4
+    }
+  rules = rbind(rules, cbind( as.character(brks[i]), as.character(brks[(i+1)]), rule))
+  }  
+  colnames(rules) = c('t1','t2', 'rule')
+  rules$Patient = Patient
+  rules
+}
+
+pg.samp = pg.samp %>% mutate(p53.Status = factor(p53.Status, levels=c(0,1), ordered=T))
+
+maxP = pg.samp %>% dplyr::group_by(Patient, Status, brks) %>% 
+  dplyr::summarise( 'maxP'=max(Prediction), 'Risk'=max(Risk), 'Pathology'=max(Pathology), 'p53'=max(p53.Status,na.rm=T) )
+
+rules = do.call(rbind, lapply(unique(maxP$Patient), function(pt) {
+  r = with(arrange(subset(maxP, Patient == pt), brks), rx(pt, brks,maxP,Pathology,Risk,p53))
+  r$Status = unique(subset(maxP, Patient == pt)$Status)
+  r
+}))
+
+rules$t1 = factor(rules$t1, levels=levels(pg.samp$brks), ordered=T)
+rules$t2 = factor(rules$t2, levels=levels(pg.samp$brks), ordered=T)
+
+penult = filter(rules, t2 != 'Endpoint' & (t1 == rev(levels(pg.samp$brks))[2] | t2 == rev(levels(pg.samp$brks))[2] ))
+
+penult %>% group_by(Status, rule) %>% tally %>% spread(rule,n) %>% 
+  left_join(penult %>% group_by(Status) %>% tally(name='total'), by='Status') %>%
+  mutate_if(is.numeric, funs(  round((./total)*100,1)  )) %>% dplyr::select(-total) %>%
+  kable(caption='Penult Endoscopy: % patients by rx rules based on the max (P) per endoscopy') %>% kable_styling(full_width=F)
+
+endpoint = subset(rules,  t2 != 'Endpoint') %>% 
+  dplyr::group_by(Patient, Status) %>% 
+  dplyr::summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )
+
+endpoint %>% group_by(Status, rule) %>% tally %>% spread(rule,n) %>% 
+  left_join(endpoint %>% group_by(Status) %>% tally(name='total'), by='Status') %>%
+  mutate_if(is.numeric, funs(  round((./total)*100,1)  )) %>% dplyr::select(-total) %>%
+  kable(caption='Final endoscopy: % patients by recommendation rule based on the max (P) per endoscopy') %>% kable_styling(full_width=F)
+
+path.rx<-function(Patient, brks, Pathology) {
+  rules = as.data.frame(matrix(ncol=0, nrow=0))
+  # Consecutive
+  for (i in 1:length(brks)) {
+    #risks = table(Pathology[i:(i+1)])
+    rule = 'None'
+    if ( length(which(grepl('HGD|IMC', Pathology[i:(i+1)]))) > 0  |
+         length(which(grepl('LGD', Pathology[i:(i+1)]))) > 1 ) {
+      rule = 1
+    } else if ( length(which(grepl('LGD', Pathology[i:(i+1)]))) > 0 ) {
+      rule = 2
+    } else if ( length(which(grepl('BE|NDBE|ID', Pathology[i:(i+1)]))) > 1  ) {
+      rule = 3
+    }
+  rules = rbind(rules, cbind( as.character(brks[i]), as.character(brks[(i+1)]), rule))
+  }  
+  colnames(rules) = c('t1','t2', 'rule')
+  rules$Patient = Patient
+  rules
+}
+
+rules2 = do.call(rbind, lapply(unique(maxP$Patient), function(pt) {
+  r = with(arrange(subset(maxP, Patient == pt), brks), path.rx(pt, brks,Pathology))
+  r$Status = unique(subset(maxP, Patient == pt)$Status)
+  r
+}))
+rules2$t1 = factor(rules2$t1, levels=levels(pg.samp$brks), ordered=T)
+rules2$t2 = factor(rules2$t2, levels=levels(pg.samp$brks), ordered=T)
+
+second = subset(rules2,  t2 != 'Endpoint') %>% 
+  dplyr::group_by(Patient, Status) %>% 
+  dplyr::summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )
+
+second %>% group_by(Status, rule) %>% tally %>% spread(rule,n) %>% 
+  left_join(second %>% group_by(Status) %>% tally(name='total'), by='Status') %>%
+  mutate_if(is.numeric, funs(  round((./total)*100,1)  )) %>% dplyr::select(-total) %>%
+  kable(caption='Final endoscopy: % patients by recommendation rule based on the max (PATH) per endoscopy') %>% kable_styling(full_width=F)
+
+```

### 2019-11-18 snp/SNP_Prediction_Analysis.Rmd
checking snp data against new models

@@
 ```{r setup, include=FALSE}
-library(ggplot2)
+knitr::opts_chunk$set(echo = F, message = F, warning = F)
+library(tidyverse)
 library(gridExtra)
 library(glmnet)
-library(pander)
-library(readxl)
 library(pROC)
-library(ggfortify)
-library(preprocessCore)
+#library(ggfortify)
+library(BarrettsProgressionRisk)
 
-source('~/workspace/shallowwgs_pipeline/lib/data_func.R')
+#source('~/workspace/shallowwgs_pipeline/lib/data_func.R')
 source('~/workspace/shallowwgs_pipeline/lib/common_plots.R')
-
-library(BarrettsProgressionRisk)
 
@@
 
-adjust.cols<-function(mtx, means=NULL, sds=NULL, na.replace=0) {
-  if (!is.null(means)) {
-    if (ncol(mtx) != length(means))
-      stop("Vector of means needs to match the columns in the matrix")
-    for (i in 1:ncol(mtx)) 
-      mtx[,i] = unit.var(mtx[,i], means[i], sds[i])
-  } else {
-    for (i in 1:ncol(mtx)) 
-      mtx[,i] = unit.var(mtx[,i])
-  }
-  mtx[is.na(mtx)] = na.replace
+model.dir = '~/Data/Ellie/Analysis/models_5e6_all/50kb'
+load(paste0(model.dir,'/model_data.Rdata'), verbose = F)
+load(paste0(model.dir,'/all.pt.alpha.Rdata'), verbose = F)
 
-  return(mtx)
-}
-
-load('~/Data/Ellie/Analysis/5e6_arms_all/model_data.Rdata', verbose = T)
-load('~/Data/Ellie/Analysis/5e6_arms_all/all.pt.alpha.Rdata', verbose = T)
-#alpha = '0'
 alpha = '0.9'
@@
 
-pv = var(dysplasia.df[swgs_labels == 1])  # P
-npv = var(dysplasia.df[swgs_labels == 0]) # NP
-
-range(dysplasia.df[swgs_labels == 1])
-range(dysplasia.df[swgs_labels == 0])
-
 load('~/Data/Reid_SNP/PerPatient/allpts_ascat.Rdata', verbose=T)
 
-qcdata = qcdata %>% rowwise %>% dplyr::mutate(
-  PatientID = unlist(strsplit(Samplename, '_'))[1],
-  SampleID = unlist(strsplit(Samplename, '_'))[2],
-  EndoID = unlist(strsplit(Samplename, '_'))[3],
-  Level = unlist(strsplit(Samplename, '_'))[4]
-)
+qcdata = qcdata %>% as_tibble %>% 
+  separate(Samplename, c('PatientID','SampleID','EndoID','Level'), sep='_', remove=F)
+
 segments.list = lapply(segments.list, function(pt) {
   pt$sample = sub('\\.LogR','',pt$sample)
-  pt
+  pt %>% as_tibble
 })
 
+qcdata[1,] -> s
+
 qcdata$ASCAT.SCA.ratio = apply(qcdata,1,function(s) {
-  smp = subset(segments.list[[ s[['PatientID']] ]], sample == s[['Samplename']] & chr %in% c(1:22))
-  smp = smp %>% rowwise %>% dplyr::mutate(
+  smp = filter( segments.list[[ s$PatientID ]], sample == s$Samplename & chr %in% c(1:22) )
+  smp = smp %>% #rowwise %>% 
+    dplyr::mutate(
       #'Total' = nAraw + nBraw,
       'Total' = nMajor + nMinor,
-      'CNV' = round(Total) - round(as.numeric(s[['Ploidy']])) )
+      'CNV' = round(Total) - round(as.numeric(s$Ploidy)) )
 
-  x = subset(smp, CNV != 0 & chr %in% c(1:22))
-  sum(x$endpos - x$startpos) / chr.info[22,'genome.length']
+  x = filter(smp, CNV != 0 & chr %in% c(1:22))
+  sum(x$endpos - x$startpos)/as.numeric(chr.info[22,'genome.length'])
 })
@@
 
-patient.info = as.data.frame(read_xlsx('~/Data/Ellie/Analysis/SNP/metadata_T1T2.xlsx'))
-patient.info$UniqueSampleID = paste(patient.info$PatientID, patient.info$`Timepoint Code`, sep='_')
-patient.info$Path.Status = patient.info$Status
-patient.info[patient.info$PatientID %in% subset(patient.info, Pathology %in% c('IMC','HGD'), select='PatientID')[,1], 'Path.Status'] = 'P'
+patient.info = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/Analysis/SNP/metadata_T1T2.xlsx') %>% 
+  mutate( UniqueSampleID = paste(PatientID, `Timepoint Code`, sep='_'), Path.Status = Status) %>% 
+  mutate(Path.Status = as.character(ifelse(Pathology %in% c('IMC','HGD'), 'P', Path.Status)))
 
-sample.list = read_xlsx('~/Data/Ellie/Analysis/SNP/20180604_Reid_1M_SampleList.xlsx', sheet = 2)
-nrow(sample.list)
+sample.list = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/Analysis/SNP/20180604_Reid_1M_SampleList.xlsx', sheet = 2)
 colnames(sample.list)[3] = 'Total.SCA'
 
-sample.list$SCA.Ratio = sample.list$Total.SCA/max(chr.info$genome.length)
+sample.list = sample.list %>% mutate(SCA.Ratio = Total.SCA/max(chr.info$genome.length))
 
@@
 sample.info = sample.info %>% rowwise() %>% dplyr::mutate(
-  SCA.Ratio = ifelse(Level %in% c('BLD','GASTRIC'), 0, get.ratio(PatientID, SampleID))
+  SCA.Ratio = ifelse(Level %in% c('BLD','GASTRIC'), 0, get.ratio(PatientID, SampleID)), 
+  Endoscopy = ifelse(Timepoint == 'T1', 1, 2)
 )
 
-patient.info = patient.info %>% rowwise() %>% dplyr::mutate( Endoscopy = ifelse(Timepoint == 'T1', 1, 2) )
-
-sample.info
 ggplot(sample.info, aes(ASCAT.SCA.ratio, SCA.Ratio)) + geom_point() + facet_grid(~Status)
@@
 ```{r, echo=F}
-sm = patient.info %>% group_by(Status, Endoscopy, Pathology) %>% dplyr::tally() %>% group_by(Status, Endoscopy) %>% dplyr::mutate(freq=n/sum(n) )
+sm = sample.info %>% group_by(Status, Endoscopy, Pathology) %>% dplyr::tally() %>% group_by(Status, Endoscopy) %>% dplyr::mutate(freq=n/sum(n) )
 sm$Pathology = factor(sm$Pathology, levels=c('NDBE','ID','LGD','HGD','IMC'), ordered = T)
@@
 # Most progressors are HGD at timepoint 1
-pander( round(with(subset(patient.info, Endoscopy == 1), table(Pathology, Status))['HGD',]/table(subset(patient.info, Endoscopy == 1)$Status), 3)*100, caption='% at T1')
+pander( round(with(subset(sample.info, Endoscopy == 1), table(Pathology, Status))['HGD',]/table(subset(sample.info, Endoscopy == 1)$Status), 3)*100, caption='% HGD at T1')
 ```
@@
 
-```{r, echo=F, warning=F, fig.height=10, fig.width=8,eval=F}
-
-summary.file = '~/Data/Annalise/NonP/summary_logR.txt'
-smy = read.table(summary.file, sep = '\t', header = T)
-smy = subset(smy, Copy.number <= 12)
-
-a = as.data.frame(smy[,c('Copy.number','var.LRR')])
-colnames(a) = c('x','y')
-b = as.data.frame(smy[,c('Copy.number','median.LRR')])
-colnames(b) = c('x','y')
-b[b$x == 2,'y'] = 0
-
-fitSD = lm(x~y, data=a)
-autoplot(fitSD)
-fitM = lm(x~y, data=b)
-autoplot(fitM)
-
-fitCN = lm(y~x, data=b)
-
-# Adjust based on the summary values per CN that we get in WGS (mostly) NP Barrett's cases
-adjust.segraw<-function(segraw) {
-  segraw = segraw %>% rowwise() %>% dplyr::mutate( total = nAraw+nBraw )
-  segraw = subset(segraw, chr %in% c(1:22))
-
-  segraw$adjLRR = NA
-  #for (cn in unique(segraw$total)) {
-  #  rows = which(segraw$total == cn)
-  #  values = seg[rows, 'medLRR', drop=T]
-
-    #newM = mean(values)
-    #newSD = sd(values)
-
-    #if (cn != 2) {
-      #newM = predict(fitM, newdata=cbind.data.frame('x'=cn))
-      #newSD = predict(fitSD, newdata=cbind.data.frame('x'=cn))
-
-    #newV = mean(values)
-
-    #}
-    #segraw[rows,][['adjLRR']] = newM + (values - mean(values)) * (newSD/sd(values))
-   # smy
-    #median(predict(fitM, newdata=cbind.data.frame('y'=values)))-cn
-
-    new = predict(fitCN, newdata=cbind.data.frame('x'=segraw$total))
-    #plot(scale(new, center = T))
-
-    segraw[,'adjLRR'] = scale(new,center=T)
-
-   # segraw[rows,'adjLRR'] = (values * (predict(fitM, newdata=cbind.data.frame('y'=values)) - cn))
-  #}  
-segraw  
-}
-
-pt = '403'
-pfiles = list.files(paste('~/Data/Reid_SNP/PerPatient', pt, sep='/'), full.names = T)
-load(grep('ascat.Rdata', pfiles, value=T), verbose=F)
-
-seg = ascat.output$segments_raw
-seg = adjust.segraw(seg)
-
-m = (melt(seg, measure.vars=c('medLRR','adjLRR')))
-m$total = round(round(m$total, 1))
-ggplot(m, aes(total, value, group=total)) + facet_grid(~variable) + geom_boxplot() + labs(title=pt)
-
-seg$chr = factor(seg$chr, levels=c(1:22))
-grid.arrange(
-  ggplot(seg) + facet_grid(sample~chr, space='free_x', scales='free_x') +
-    geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), color='darkgreen', size=3) + 
-    theme(axis.text.x=element_blank() ) + labs(title='Low SCA, medLRR', subtitle=paste('patient',pt)),
-  ggplot(seg) + facet_grid(sample~chr, space='free_x', scales='free_x') + 
-    geom_segment(aes(x=startpos, xend=endpos, y=adjLRR, yend=adjLRR), color='darkred', size=3) + 
-    theme(axis.text.x=element_blank() ) + labs(title='Low SCA, adjLRR', subtitle=paste('patient',pt))
-)
-
-pt = '512'
-pfiles = list.files(paste('~/Data/Reid_SNP/PerPatient', pt, sep='/'), full.names = T)
-load(grep('ascat.Rdata', pfiles, value=T), verbose=F)
-
-seg = ascat.output$segments_raw
-seg = adjust.segraw(seg)
-
-m = (melt(seg, measure.vars=c('medLRR','adjLRR')))
-m$total = round(round(m$total, 1))
-ggplot(m, aes(total, value, group=total)) + facet_grid(~variable) + geom_boxplot() + labs(title=pt)
-
-seg$chr = factor(seg$chr, levels=c(1:22))
-grid.arrange(
-  ggplot(seg) + facet_grid(sample~chr, space='free_x', scales='free_x') + 
-    geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), color='darkgreen', size=3) + 
-    theme(axis.text.x=element_blank() ) + labs(title='Mixed SCA, medLRR', subtitle=paste('patient',pt)),
-  ggplot(seg) + facet_grid(sample~chr, space='free_x', scales='free_x') + 
-    geom_segment(aes(x=startpos, xend=endpos, y=adjLRR, yend=adjLRR), color='darkred', size=3) + 
-    theme(axis.text.x=element_blank() ) + labs(title='Mixed SCA, adjLRR', subtitle=paste('patient',pt))
-)
-
-```
-
 # Load tiled data
@@
 }
+
 rownames(mergedSegs) = sub('\\.LogR','', rownames(mergedSegs))
@@
 
-copySegs = q.norm.by(mergedSegs, raw.segs)
-copyArms = q.norm.by(mergedArms, raw.arms)
+#copySegs = q.norm.by(mergedSegs, raw.segs)
+#copyArms = q.norm.by(mergedArms, raw.arms)
 
-grid.arrange(
-  ggplot( melt(raw.segs), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, sWGS bins (no arms)'),
-  ggplot( melt(mergedSegs), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, SNP bins (no arms)'),
-  ggplot( melt(copySegs), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, SNP bins q-normed')
-)
+training = do.call(bind_rows, purrr::map(list.files('~/Data/BarrettsProgressionRisk/Analysis/pcf_perPatient/50kb', '5e06_tiled_segvals.txt', full.names = T, recursive = T), function(f) {
+  read_tsv(f, col_types = cols(.default=col_double(), 'X1'=col_character()))
+}))
+training.segs = as.matrix(training[,-1]); rownames(training.segs) = training$X1
 
-```
+training = do.call(bind_rows, purrr::map(list.files('~/Data/BarrettsProgressionRisk/Analysis/pcf_perPatient/50kb', 'arms_tiled_segvals.txt', full.names = T, recursive = T), function(f) {
+  read_tsv(f, col_types = cols(.default=col_double(), 'X1'=col_character()))
+}))
+training.arms = as.matrix(training[,-1]); rownames(training.arms) = training$X1
 
-Adjust the variance of each genomic region by the mean and sd extracted when I trained on the sWGS data.
-This seems to be a more fair comparison than scaling all the new data to itself, but the predictions don't really change.
+merged = rbind(mergedSegs, training.segs)
+merged.mean = apply(merged, 2, mean, na.rm=T)
+merged.sd = apply(merged, 2, sd, na.rm=T)
 
-Using the mean bin values (merged during tiling).
-```{r, warning=F, message=F, fig.width=10, fig.height=8}
+for (i in 1:ncol(mergedSegs))
+  mergedSegs[,i] = BarrettsProgressionRisk:::unit.var(mergedSegs[,i], merged.mean[i], merged.sd[i])
 
-#ns = adjust.cols(normalSegs, z.mean, z.sd)
-#ns = adjust.cols(normalSegs)
-#range(ns)
+merged.arms = rbind(mergedArms, training.arms)
+merged.arms.mean = apply(merged.arms, 2, mean, na.rm=T)
+merged.arms.sd = apply(merged.arms, 2, sd, na.rm=T)
 
-#na = adjust.cols(normalArms, z.arms.mean, z.arms.sd)
-#na = adjust.cols(normalArms)
-#range(na)
+for (i in 1:ncol(mergedArms))
+  mergedArms[,i] = BarrettsProgressionRisk:::unit.var(mergedArms[,i], merged.arms.mean[i], merged.arms.sd[i])
 
-#ms = adjust.cols(mergedSegs, z.mean, z.sd)
-#ms = adjust.cols(mergedSegs)
-ms = adjust.cols(copySegs)
-range(ms)
-dim(ms)
+# grid.arrange(
+#   ggplot( melt(raw.segs), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, sWGS bins (no arms)'),
+#   ggplot( melt(mergedSegs), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, SNP bins (no arms)')
+#   ggplot( melt(copySegs), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, SNP bins q-normed')
+# )
 
-# mergedSegs['997_122V',which(mergedSegs['997_122V',] <= 0)]
-# ms['997_122V',which(mergedSegs['997_122V',] <= 0)]
+normal.ids = sample.info %>% filter(grepl('BLD|gastric', Level, ignore.case=T)) %>% dplyr::summarise(rows = paste0(PatientID, '_', EndoID)) %>% pull
 
-#ma = adjust.cols(mergedArms, z.arms.mean, z.arms.sd)
-#ma = adjust.cols(mergedArms)
-ma = adjust.cols(copyArms)
-range(ma)
-dim(ma)
-
-nm = melt(ms[nm.rows,])
-be = melt(ms[-nm.rows,])
-
-lims = range(nm$value, be$value, na.rm = T)
-grid.arrange(
-  ggplot(nm, aes(Var2, value, group=Var2, color=Var1)) + geom_point() + labs(title='Blood/gastric') + 
-    theme(axis.text.x=element_blank(),legend.position='none') +ylim(lims),
-  ggplot(be, aes(Var2, value, group=Var2, color=Var1)) + geom_point() + labs(title='BE samples') + 
-    theme(axis.text.x=element_blank(),legend.position='none') + ylim(lims), top='Unit scaled')
+nm.rows = which(rownames(mergedSegs) %in% normal.ids)
 
@@
 
-cx = score.cx(ms, 1)
+cx = BarrettsProgressionRisk:::scoreCX(mergedSegs, 1)
 #arrayDf = cbind(ms,ma)
-arrayDf = subtract.arms(ms, ma)
+arrayDf = BarrettsProgressionRisk:::subtractArms(mergedSegs, mergedArms)
 #arrayDf = cbind(arrayDf, 'cx'=unit.var(cx, mn.cx, sd.cx))
-arrayDf = cbind(arrayDf, 'cx'=unit.var(cx))
+arrayDf = cbind(arrayDf, 'cx'=BarrettsProgressionRisk:::unit.var(cx, mn.cx, sd.cx))
 
@@
 
+preds = preds %>% mutate(Status = as.character(ifelse(Pathology %in% c('HGD','IMD'), 'P', Status)))
+
+preds %>% filter(grepl('BLD|gastric',Level,ignore.case=T)) %>% group_by(Risk, Level) %>% tally %>% spread(Risk,n)
+
+preds %>% filter(!grepl('BLD|gastric',Level,ignore.case=T)) %>% group_by(Status, Risk, Pathology) %>% tally %>% spread(Risk,n)
+
 # offset = log(0.0225)
@@
 
-pnr = grep('BLD|gastric', preds$Samplename, ignore.case = T)
-preds.normal = preds[pnr,]
+preds.normal = preds %>% filter(grepl('BLD|gastric', Level, ignore.case=T))
+
+preds.normal %>% group_by(Risk) %>% tally
 
@@
 
-#randNormal = paste(unlist(strsplit(randNormal, '_'))[c(1,3)], collapse='_')
-#randBE = paste(unlist(strsplit(randBE, '_'))[c(1,3)], collapse='_')
+preds = sample.info %>% mutate(Sample = paste(PatientID,EndoID,sep='_')) %>% left_join(preds, by=c('Sample'='Samplename'))
+
+summary(preds$Prob)
+
+preds = preds %>% 
+  mutate(Risk = ifelse(Prob > 0.5, 'High', 'Low')) %>% 
+  mutate(Risk = factor(Risk, levels=c('Low','High'), ordered=T))
+
+preds %>% group_by(Status, Risk) %>% tally %>% spread(Risk, n) %>%
+  left_join(preds %>% group_by(Status) %>% tally, by='Status') %>%
+  mutate_if(is.numeric, funs( round(./n,2))) %>% 
+  dplyr::select(-n)
+
+roc = pROC::roc(Status ~ Prob, data=preds, auc=T, ci=T, of='thresholds',transpose=T)
+roc.plot(roc) + labs(title='SNP AUC by (P)')
+pROC::coords(roc, 'best')
+
+rocR = pROC::roc(Status ~ Risk, data=preds, auc=T, ci=T, of='thresholds',transpose=T)
+roc.plot(rocR) + labs(title='SNP AUC by discretized risks (low/high)')
+pROC::coords(rocR, 'best')
 

### 2019-11-19 Paper_resubmit.Rmd
generating all new plots

@@
 sum.patient.data = as_tibble(summarise.patient.info(patient.info))
+
+patient.info = patient.info %>% mutate( Pathology = recode(Pathology, 'BE'='NDBE'))
 
@@
 
-globalBECN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology == 'BE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')
+globalBECN.plot = cn.mtn.plot(filter(globalCNdf, seg.type != 'arms' & Pathology == 'NDBE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')
 
 gp = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples')
-ggsave(filename='plots/globalCN.png', plot=gp, width=16, height=4.6, units='in', dpi=300)
-
-gbe = globalBECN.plot + theme(legend.position = 'none') + labs(y='mean adj.seg.', title="Non-Dysplastic Barrett's Samples") 
-ggsave(filename='plots/globalBECN.png', plot=gbe, width=16, height=4.6, units='in', dpi=300)
+
+gp = globalCN.plot + theme(legend.position = 'none', axis.text = element_blank(), strip.text.y = element_blank(), strip.text.x = element_text(size=8)) + 
+  labs(y='',x='', title='All samples')
+ggsave(filename='plots/globalCN.png', plot=gp, width=8, height=2.3, units='in', dpi=300)
+
+gbe = globalBECN.plot +theme(legend.position = 'none', axis.text = element_blank(), strip.text.y = element_blank(), strip.text.x = element_text(size=8)) + 
+  labs(y='',x='', title="Non-Dysplastic Barrett's Samples") 
+ggsave(filename='plots/globalBECN.png', plot=gbe, width=8, height=2.3, units='in', dpi=300)
 
@@
 
+cvRR = features
+
 qq = quantile(features$hazards, seq(0,1,.15))
@@
   scale_color_manual(values=c('firebrick3','darkblue')) +
-  labs(title='Top features by hazard ratio', x='Coefficient value', y='Hazard') + plot.theme 
+  labs(title='Top features by cv(RR)', x='Coefficient value', y='cvRR') + plot.theme 
 p
@@
 ```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
-features = features %>% separate(label, c('chr','start','end'), sep=':|-', remove=F) %>% 
+features = features %>% filter(label != 'cx') %>% separate(label, c('chr','start','end'), sep=':|-', remove=F) %>% 
   dplyr::mutate_at(vars(start, end), list(as.numeric)) %>% 
@@
 
+pg.samp = pg.samp %>% mutate(Pathology = recode(Pathology, 'BE' = 'NDBE'))
+
 names(performance.at.1se) = names(pg.samp)
@@
 roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds',transpose=T)
+ggsave(filename='plots/auc/training_auc.png', plot=roc.plot(roc), width=6.5, height=4.5, units='in', dpi=300)
+
 roc$model = 'CN Model'
@@
     scale_x_reverse() +
-    geom_label(data=filter(i.thresholds, AUC %in% range(AUC) | is == 'discovery'), aes(label=paste0('AUC\n',round(AUC*100))), color='black') +
+    geom_label(data=filter(i.thresholds, AUC %in% range(AUC) | is == 'discovery'), aes(label=paste0('AUC\n',round(AUC,2))), color='black') +
     #scale_color_brewer(palette='Blues') +
-    labs(title='ROC AUC', subtitle='in-silico splits', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
-    theme(legend.position = 'none')
+    labs(title='ROC AUC', subtitle='in-silico splits', x='Specificity (1-FPR)', y='Sensitivity (TPR)')  + plot.theme +
+    theme(legend.position = 'none', text = element_text(size=14))
 p
@@
 ```{r, echo=F, warning=F, message=F,fig.height=4, fig.width=4}
-pendo = pg.samp %>% dplyr::group_by(Patient, Status, Hospital.Research.ID, PID ) %>% dplyr::summarise(
-  'max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(PID)
-)
+pendo = pg.samp %>% dplyr::group_by(Patient, Status, Hospital.Research.ID, PID ) %>% 
+  dplyr::summarise( 'max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(PID) )
 
@@
   geom_text(aes(y=(value*100)-4, label=paste(round(value,2)*100,'%',sep='')), position=dodge, color='white')  +
+  scale_y_continuous(limits=c(0,100)) +
   scale_fill_manual(values=cols, name='') + plot.theme + 
-  theme(legend.position = 'none', axis.text.x = element_text(angle=45, hjust=1)) + labs(x='', y='AUC', title='ROC Comparison') 
+  theme(legend.position = 'none', text = element_text(size=14), axis.text.x = element_text(angle=45, hjust=1)) + 
+  labs(x='', y='AUC', title='ROC Comparison') 
 comp
 
-ggsave('plots/auc/path_p53_cn_auc.png', plot=comp, width=6, height=8, units="in", limitsize=F, dpi = 300)
+ggsave('plots/auc/path_p53_cn_auc.png', plot=comp, width=3, height=4, units="in", limitsize=F, dpi = 300)
 
@@
 ```{r}
-pg.samp = do.call(bind_rows,lapply(unique(pg.samp$Patient) %>% map( function(p) pg.samp %>% filter(Patient == p)), transform.by.time, info=patient.info)) %>% dplyr::mutate(Pathology = factor(Pathology, levels=c('BE','ID','LGD','HGD','IMC'), ordered = T))
-
+pg.samp = do.call(bind_rows,lapply(unique(pg.samp$Patient) %>% map( function(p) pg.samp %>% filter(Patient == p)), transform.by.time, info=patient.info)) 
 sq = c(0, seq(1,132,12),max(pg.samp$months.before.final))
@@
 
+mini = pcal + labs(x='P(P)', y='P:NP', title='') + theme(legend.position = 'none') + 
+  scale_x_continuous(expand = c(0, 0), limits = c(-0.5, 1.5), breaks = cuts, labels = c(0,'','',0.3,'',0.5,'','','','',1)) + 
+  scale_y_continuous(expand = c(0, 0), limits = c(-0.5, 1.5), breaks = seq(0,1,0.25), labels = seq(0,1,0.25)) + 
+  theme(axis.text = element_text(size=14))
+ggsave(filename='plots/pred_calib_mini.png', plot=mini, width=4, height=4, units='in', dpi=300)
+
 pred.confidence %>% dplyr::mutate_if(is.double, list(signif), 2) %>% kable() %>% kable_styling('striped')
@@
 setOgj <- function(nl) {
-  if (nl >= 0 & nl < 3) return(1)
-  if (nl >= 3 & nl < 7) return(2)
-  if (nl >= 7) return(3)
+  if (nl >= 0 & nl < 2) return(1)
+  if (nl >= 2 & nl < 4) return(2)
+  if (nl >= 4 & nl < 7) return(3)
+  if (nl >= 7) return(4)
 }
@@
 p6 = mtn.spatial(globalCNdf %>% filter(Patient == 51 & seg.type != 'arms'), pg.samp, ogjLimits,'Patient 51 (NP)' )
+ggsave(filename='plots/spatial/patient_20_P.png', plot=grid.arrange(p2$grob), width=10.5, height=4.5, units='in', dpi=300)
+ggsave(filename='plots/spatial/patient_51_NP.png', plot=grid.arrange(p6$grob), width=10.5, height=4.5, units='in', dpi=300)
+
+grid.arrange(p2$grob, p6$grob, ncol=1)
 
 # for (i in 1:nrow(sum.patient.data)) {
-#   #print(i)
+#   print(i)
 #   pt = sum.patient.data[[i, 'Patient']]
@@
 #   fl = paste0('patient_', pt, '_',status,'.png')
-#   ggsave(paste('plots/spatial/', fl, sep=''), plot=grid.arrange(pms$grob), width=14, height=6, units='in', dpi=300)
+#   ggsave(paste('plots/spatial/', fl, sep=''), plot=grid.arrange(pms$grob), width=10.5, height=4.5, units='in', dpi=300)
 # }
@@
 pg.samp = pg.samp %>% dplyr::select(-ogj)
-```
-
-```{r, echo=F, fig.width=2, fig.height=4}
-grid.arrange(p2$legend)
-
-ggsave(filename='plots/spatial_mtn_pt20.png', plot=grid.arrange(p2$grob), width=14, height=6, units='in', dpi=300)
-ggsave(filename='plots/spatial_mtn_pt51.png', plot=grid.arrange(p6$grob), width=14, height=6, units='in', dpi=300)
+
+ggsave('plots/spatial/legend.png', plot=grid.arrange(p2$legend), width=2,height=2, units='in', dpi=300)
 ```
@@
 rp
-ggsave(filename='plots/pred_by_path.png', plot=rp, width=9, height=7, units='in', dpi=300)
+ggsave(filename='plots/pred_by_path.png', plot=(rp + theme(legend.position = 'none')), width=7, height=6, units='in', dpi=300)
 ```
@@
   scale_fill_manual(values=col, name='p53 IHC', labels=c('Normal','Aberrant')) + 
-  geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101)) +
+  geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), nudge_y=1) +
   geom_text(aes(label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5,color='white') +
-  plot.theme + labs(title='Samples predicted by aberrant p53 IHC', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
+  plot.theme + labs(title='Samples predicted by aberrant p53 IHC', y='% Predicted', x='Pathology') + 
+  theme(legend.position = 'bottom', text = element_text(size=14))
 rp
-ggsave(filename='plots/p53_pred.png', plot=rp, width=7, height=6, units='in', dpi=300)
+ggsave(filename='plots/p53_pred.png', plot=rp, width=6, height=5, units='in', dpi=300)
 ```
@@
 
-```{r enods, warning=F, message=F, echo=F, fig.height=8, fig.width=6}
+```{r endos, warning=F, message=F, echo=F, fig.height=8, fig.width=6}
 sq = c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12)
@@
 rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) +
-  geom_bar(stat='identity', fill='darkred', alpha=0.8, show.legend = F) + 
-  geom_errorbar(color='grey21', width=0.2) +
-  geom_text(aes(y=3, label=paste(round(pred.r*100), '%', sep='')), color='white') +
-  geom_text(aes(y=pred.r*100-5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
+  geom_bar(stat='identity', fill=BarrettsProgressionRisk:::riskColors()[['High']], show.legend = F) + 
+  geom_errorbar(color='grey55', width=0.2) +
+  geom_text(aes(label=paste(round(pred.r*100), '%', sep='')), color='black', nudge_y=2) +
+  geom_text(aes(y=5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white', nudge_y=2) + 
   labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') + coord_cartesian(ylim = c(0,100)) + 
-  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12)) 
+  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12) ) 
 rp
-ggsave(filename='plots/pred_endo.png', plot=rp, width=9, height=7, units='in', dpi=300)
+ggsave(filename='plots/pred_endo.png', plot=rp, width=6, height=5, units='in', dpi=300)
 
@@
                                     Patient = factor(Patient, ordered=T), 
-                                    Block = factor(Block, ordered=T), 
-                                    Pathology = recode(Pathology, BE = 'NDBE'))
+                                    Block = factor(Block, ordered=T))
 
@@
 plist = purrr::map(sum.patient.data$Patient, function(pt) {
-  patient = pg.samp %>% filter(Patient == pt) %>% dplyr::select(Patient,Status,Endoscopy.Year,Block,Risk,months.before.final) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block)
-
-  p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final','rev') + theme(legend.position = 'none') + labs(x='',y='') + theme(axis.text = element_blank(), plot.title = element_text(hjust = 0.5, face = 'bold', size=16))
-
-  ggsave(paste0('plots/heatmaps/tiny/',unique(patient$Status),'/pt_',pt,'.png'), plot = p, height=length(unique(patient$GEJ.Distance)), width=length(unique(patient$months.before.final)), units='in', dpi=300, limitsize=F)
-
-  patient = pg.samp %>% filter(Patient == pt) %>% dplyr::select(Patient,Status,Pathology,Endoscopy.Year,Block,Risk,months.before.final) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block)
+  patient = pg.samp %>% filter(Patient == pt) %>% dplyr::select(Patient,Status,Endoscopy.Year,Block,Risk,months.before.final) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) %>% ungroup
+
+  p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final','rev') + theme(legend.position = 'none') + labs(x='',y='',title='') + theme(axis.ticks = element_blank(), axis.text = element_blank(), plot.margin=grid::unit(c(0,0,0,0), "mm"))
+
+  ht = ifelse(length(unique(patient$GEJ.Distance)) > 1, length(unique(patient$GEJ.Distance)), length(unique(patient$GEJ.Distance))+0.5)
+  wd = ifelse(length(unique(patient$months.before.final)) > 1, length(unique(patient$months.before.final)), length(unique(patient$months.before.final))+0.5 )
+
+  ggsave(paste0('plots/heatmaps/tiny/',unique(patient$Status),'/pt_',pt,'.png'), plot = p, height=ht, width=length(unique(patient$months.before.final)), units='in', dpi=150, limitsize=F)
+
+  patient = pg.samp %>% filter(Patient == pt) %>% dplyr::select(Patient,Status,Pathology,Endoscopy.Year,Block,Risk,months.before.final) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) %>% ungroup
 
@@
 
-p11 = subset(globalCNdf, chr == 17 &  start == 1 & seg.type == 'arms')
-p11Loss = subset(p11, value < -1)
-table(p11Loss$Status)
-
-p11path = table(subset(patient.info, Samplename %in% p11Loss$variable)$Pathology)
-
-p11path['BE']/sum(p11path)
-sum(p11path[c('HGD','IMC')])/sum(p11path)
-
-table(p11Loss[c('Patient', 'Pathology')])
-
-# Samples with a loss vs all Progressor samples
-length(unique(p11Loss$variable))/length(unique(subset(globalCNdf, Status == 'Progressor')$variable))
-
-p11Loss = unique(p11Loss[,c('variable', 'Status','Hospital.Research.ID', 'Patient', 'Pathology')])
-length(unique(p11Loss$Patient))/length(subset(sum.patient.data, Status == 'P')$Patient)
-# Of those where we see loss of p arm, p53 IHC shows an "abnormal" result 33% of the time
-
-table(subset(patient.info, Samplename %in% p11Loss$variable, select='p53.Status'))/length(p11Loss$variable)
-
-(table(subset(patient.info, Samplename %in% p11Loss$variable, select=c('p53.Status', 'Pathology'))))
+p11 = filter(globalCNdf, chr == 17 &  start == 1 & seg.type == 'arms')
+p11Loss = p11 %>% filter(value < -1)
+
+p11Loss %>% filter(value < -1) %>% group_by(Status) %>% tally(name='p11 Loss') %>% full_join(
+  p11 %>% group_by(Status) %>% tally(name='Total Samples') ) %>% 
+  kable(caption='Chr17 p11 loss') %>% kable_styling(full_width = F)
+
+p11Loss %>% group_by(Status, Hospital.Research.ID) %>% tally
+
+p11Path = patient.info %>% filter(Samplename %in% p11Loss$variable) %>% dplyr::group_by(Pathology) %>% tally(name='samples')
+
+p11Path %>% dplyr::mutate('ratio'=round(samples/sum(p11Path$samples),2)) %>% 
+kable(caption='Ratio of all Chr17 p11 loss samples by pathology' ) %>% kable_styling(full_width = F)
+```
+
+Samples with a loss vs all Progressor samples
+Ratio of samples with p11 loss vs all progressor samples: `r length(unique(p11Loss$variable))/length(unique(filter(globalCNdf, Status == 'Progressor')$variable))`
+
+Ratio of progressor patients with a p11 loss in at least 1 sample: `r length(unique(p11Loss$Patient))/length(subset(sum.patient.data, Status == 'P')$Patient)`
+
+```{r}
+patient.info %>% filter(Samplename %in% p11Loss$variable) %>% group_by(forcats::fct_explicit_na(p53.Status)) %>% tally %>% 
+  set_names(c('p53 Status', 'n samples')) %>% kable(caption='p53 status for samples with p11 loss') %>% kable_styling(full_width = F)
 
@@
 pi$p11Loss = as.numeric(pi$Samplename %in% p11Loss$variable)
-
-plot(table(p11=pi$p11Loss, ihc=pi$p53.Status))
+#plot(table(p11=pi$p11Loss, ihc=pi$p53.Status))
 
 cor.test(pi$p11Loss, (as.numeric(pi$p53.Status)-1))
-
-# So why might there be loss but no corresponding aberrant staining?
-table(subset(pi, p11Loss == 1 & p53.Status == 0)$Pathology)
-table(subset(pi, p11Loss == 1 & p53.Status == 1)$Pathology)
-
 ```
@@
 ## Top Features
-```{r, echo=T, eval=F}
-m = data.frame(matrix(nrow=nrow(features), ncol=0))
-
-for (i in 1:nrow(features)) {
-  feat = features[i,]
-
-  seg = subset(globalCNdf, chr == as.character(feat$chr) & start == feat$start & end == feat$end & seg.type != 'arms')  
-  if(feat$end-feat$start > 5e6)
-    seg = subset(globalCNdf, chr == as.character(feat$chr) & start == feat$start & end == feat$end & seg.type == 'arms')
-
-  segGL = subset(seg, value < -1)
-  if(feat$coef > 0)
-    segGL = subset(seg, value > 1)
-
-  m[i, 'n.samples'] = length(unique(segGL$variable))
-  m[i,'ratio.samples.P'] = length(unique(segGL$variable))/length(unique(subset(globalCNdf, Status == 'Progressor')$variable))
-  m[i, 'ratio.NP'] = table(segGL$Status)[1]/sum(table(segGL$Status))
-  m[i, 'ratio.P'] = table(segGL$Status)[2]/sum(table(segGL$Status))
-
-  segpath = table(subset(patient.info, Samplename %in% segGL$variable)$Pathology)
-
-  m[i,'BE'] = segpath['BE']/sum(segpath)
-  m[i,'HGD'] = sum(segpath[c('HGD','IMC')])/sum(segpath)
-}
-rownames(m) = features$label
-
-range(m$ratio.samples.P)
-summary(m$ratio.samples.P)
-subset(m, ratio.samples.P == max(ratio.samples.P))
-subset(features, label == rownames(subset(m, ratio.samples.P == max(ratio.samples.P))))
-subset(chh, label == rownames(subset(m, ratio.samples.P == min(ratio.samples.P))))
-#table(cdkGain[c('Patient', 'Pathology')])
-
-cor.test(features$hazards, m$n.samples)
-
-# Samples with a gain vs all samples
-
-```
 
@@
   separate(`Genome Location`, c('chr','start','end'),':|-', remove=F) %>%
+  dplyr::filter(start != '' | end != '') %>%
   dplyr::select(`Gene Symbol`, `Genome Location`, `Role in Cancer`, `Mutation Types`, chr, start, end)
 
-genes = ccgenes %>% dplyr::filter(grepl('TSG|oncogene', `Role in Cancer`) & `Mutation Types` != 'T') %>% 
+genes = ccgenes %>% dplyr::filter(grepl('TSG|oncogene', `Role in Cancer`) | `Mutation Types` != 'T') %>% 
   dplyr::select(`Gene Symbol`, `Genome Location`, chr, start, end) %>%
-  dplyr::mutate(value = min(globalCNdf$value), Patient = '')
-
-#genesGR = makeGRangesFromDataFrame(genes, keep.extra.columns = T)
-#chGR = makeGRangesFromDataFrame(chh, keep.extra.columns = T)
-
-#ov = findOverlaps(chGR, genesGR)
+  dplyr::mutate(value = min(globalCNdf$value), Patient = '', chr = factor(chr,levels=levels(chr.lengths$chr),ordered=T), start = as.numeric(start), end=as.numeric(end)) %>% arrange(chr, start)
+
+genesGR = makeGRangesFromDataFrame(genes, keep.extra.columns = T)
+chGR = makeGRangesFromDataFrame(features, keep.extra.columns = T)
+
 #genes = as.data.frame(genesGR[unique(subjectHits(ov))])
@@
 
-g17 = 
-  plot.cn.chr(globalCNdf, chrom='17', chr.lengths, genes=genes, label=labeller(chr=label_both, Status=n.status)) + 
-  labs(title="All Samples") + theme(legend.position = 'none')
-g17
-
-g17 = g17 + theme(text=element_text(size=11), strip.text = element_text(size=11))
-ggsave(filename='plots/hazards/chr17.png', plot=g17, width=10, height=5, units='in', dpi=300)
+g17 = plot.cn.chr(globalCNdf, chrom='17', chr.lengths, haz=features, genes=genes, label=labeller(chr=label_both, Status=n.status)) + 
+  labs(title="Chromosome 17") + scale_y_continuous(breaks = seq(-30,25,10), labels = c('-30', '','',0,'',20)) +
+  theme(legend.position = 'none', strip.text = element_blank(), text=element_text(size=14))
+
+ggsave(filename='plots/hazards/chr17.png', plot=g17, width=8, height=4, units='in', dpi=300)
 
@@
 ```{r echo=F, warning=F, message=F, fig.height=5, fig.width=10, eval=F}
+ov = findOverlaps(chGR, genesGR)
 chGR$Genes = ''
@@
 
-for (chr in unique(chh$chr)) {  
-  p = plot.cn.chr(subset(globalCNdf, seg.type != 'arms') , chrom=chr, chr.lengths, genes=genes, haz=chh)
+for (chr in unique(features$chr)) {  
+  print(chr)
+  p = plot.cn.chr(globalCNdf,  chrom=chr, info=chr.lengths, haz=features, genes=genes, label=labeller(chr=label_both, Status=n.status)) +
+    scale_y_continuous(breaks = seq(-30,25,10), labels = c('-30', '','',0,'',20)) +
+    theme(strip.text.x = element_blank(), text=element_text(size=14)) + labs(title=paste('Chromosome',chr))
   f = paste('chr',chr,'.png',sep='')
-  ggsave(paste('plots/hazards/',f,sep=''), plot=p, width=10, height=5, units='in', dpi=300)
-  print(p)
-}
-
-annoG = arrange(subset(ccgenes, grepl('CDKN2A$|TP53|MYC$|SMAD4|VEG|ERBB2|EGFR|SMYD|ARID1A|FHIT|RUNX1', Gene.Symbol)), chr)
-annoG = merge(annoG, chr.lengths[c('chrom','chr.length')], by.x='chr',by.y='chrom')
-annoG$ymin = round(range(globalCN.plot$data$mean.value))[1]/2
-annoG$ymax = round(range(globalCN.plot$data$mean.value))[2]/2
+  ggsave(paste('plots/hazards/',f,sep=''), plot=p, width=8, height=4, units='in', dpi=300)
+}
+
+annoG = ccgenes %>% filter(grepl('CDKN2A$|TP53|MYC$|SMAD4|VEG|ERBB2|EGFR|SMYD|ARID1A|FHIT|RUNX1', `Gene Symbol`)) %>% arrange(chr) %>%
+  left_join(chr.lengths %>% dplyr::select(chrom,chr.length), by=c('chr'='chrom')) %>%
+  mutate(ymin = round(range(globalCN.plot$data$mean.value))[1]/2, ymax = round(range(globalCN.plot$data$mean.value))[2]/2) %>%
+  mutate(chr = factor(chr, levels=levels(chr.lengths$chr)), start = as.numeric(start), end = as.numeric(end))
 
@@
   geom_point(data=annoG, aes(x=start, y=0), size=1.5, color='red') +
-  geom_text_repel(data=annoG, aes(x=start, y=ymax/2, label=Gene.Symbol), color='red', size=1.5 ) + theme(legend.position = 'none')
+  geom_text_repel(data=annoG, aes(x=start, y=ymax/2, label=`Gene Symbol`), color='red', size=1.5 ) + theme(legend.position = 'none')
 
@@
 
-(genes[,c(1:3,6,7)]) %>% kable() %>% kable_styling(full_width=F)
+#(genes[,c(1:3,6,7)]) %>% kable() %>% kable_styling(full_width=F)
 ```
@@
 
-#cn.mtn.plot(filter(valCNdf, seg.type == '5MB'), labeller(Status=n.status, chr=label_value))
+#cn.mtn.plot(filter(valCNdf, seg.type == '5MB'), labeller(Status=n.status, chr=label_value)) + ylim(-30,20)
 
@@
 roc.plot(roc) + labs(title='Validation cohort')
-ggsave('plots/validation/validatation_roc.png', plot=roc.plot(roc) + labs(title='Validation cohort'), width=6, height=4, units='in',dpi=300)
+ggsave('plots/validation/validatation_roc.png', plot=roc.plot(roc) + labs(title='Validation cohort'), width=5, height=5, units='in',dpi=300)
 
@@
 rp
-ggsave(filename='plots/validation/val_pred_by_path.png', plot=rp, width=9, height=7, units='in', dpi=300)
+ggsave(filename='plots/validation/val_pred_by_path.png', plot=rp, width=5, height=6, units='in', dpi=300)
 
@@
 
-dir.create('plots/validation/heatmaps/P',recursive = T, showWarnings = F)
-dir.create('plots/validation/heatmaps/NP',recursive = T, showWarnings = F)
+dir.create('plots/validation/heatmaps/P/tiny',recursive = T, showWarnings = F)
+dir.create('plots/validation/heatmaps/NP/tiny',recursive = T, showWarnings = F)
 
 for (pt in unique(val.preds$Patient)) {
-print(pt)
+#print(pt)
+  patient = val.preds %>% filter(Patient == pt) %>% 
+    dplyr::select(Patient, Status, Endoscopy, months.before.final, Block,Risk) %>% 
+    dplyr::rename(GEJ.Distance = Block)
+
+    p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final','rev') + theme(legend.position = 'none') + labs(x='',y='',title='') + theme(axis.ticks = element_blank(), axis.text = element_blank(), plot.margin=grid::unit(c(0,0,0,0), "mm"))
+
+    ht = ifelse(length(unique(patient$GEJ.Distance)) > 1, length(unique(patient$GEJ.Distance)), length(unique(patient$GEJ.Distance))+0.5)
+    wd = ifelse(length(unique(patient$months.before.final)) > 1, length(unique(patient$months.before.final)), length(unique(patient$months.before.final))+0.5 )
+
+    ggsave(paste0('plots/validation/heatmaps/',unique(patient$Status),'/tiny/pt_',pt,'.png'), plot = p, height=ht, width=length(unique(patient$months.before.final)), units='in', dpi=150, limitsize=F)
+
     patient = val.preds %>% filter(Patient == pt) %>% 
@@
 
+sq = c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12)
+val.preds = val.preds %>% mutate(months.before.final = ifelse(months.before.final > sq[length(sq)], sq[length(sq)], months.before.final))
+
 v.prog = val.preds %>% filter(Status == 'P') %>% mutate(
   brks = cut(months.before.final, breaks=sq, 
-                 labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-c(length(sq)-1,length(sq))], '>120'),include.lowest = T, ordered_result = T),
+                 labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-c(length(sq)-1,length(sq))], '>120'), include.lowest = T, ordered_result = T),
 ) %>% dplyr::mutate(brks = factor(brks, rev(levels(brks))))
 
-sq = c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12)
 endos = v.prog %>% dplyr::group_by(Status, months.before.final, brks, PID) %>% 
@@
               ) %>% ungroup %>% 
-  dplyr::mutate(brks = cut(months.before.final, include.lowest = T, ordered_result = T,
-     breaks=sq,  labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))) %>% 
+  dplyr::mutate(brks = cut(months.before.final, include.lowest = T, ordered_result = T, breaks=sq,  labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))) %>%
   dplyr::mutate(brks = factor(brks, rev(levels(brks))))
@@
 rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) +
-  geom_bar(stat='identity', fill='darkred', alpha=0.8, show.legend = F) + 
+  geom_bar(stat='identity', fill=BarrettsProgressionRisk:::riskColors()[['High']], show.legend = F) + 
   geom_errorbar(color='grey21', width=0.2) +
@@
 rp
-
+ggsave(filename='plots/validation/pred_endo.png', plot=rp, width=6, height=5, units='in', dpi=300)
 ```
@@
 
-for (pt in c(34,56,20,86,3,23,33,49,51,6,25,68)) {
+pt = 34
+p34 = vig.plot(df=subset(pg.samp, Patient == pt),info= subset(patient.info, Patient == pt),cols=riskCols) + 
+    theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
+    theme(legend.position = 'none', text = element_text(size=12))
+ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p34, width=6.35, height=4.5,dpi=72)
+
+pt = 3
+p3 = vig.plot(df=subset(pg.samp, Patient == pt),info= subset(patient.info, Patient == pt),cols=riskCols) + 
+    theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
+    theme(legend.position = 'none', text = element_text(size=12))
+ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p3, width=6.35, height=4,dpi=72)
+
+pt = 68
+p68 = vig.plot(df=subset(pg.samp, Patient == pt),info= subset(patient.info, Patient == pt),cols=riskCols) + 
+    theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
+    theme(legend.position = 'none', text = element_text(size=12))
+ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p68, width=6.35, height=5,dpi=72)
+
+for (pt in c(56,20,86,23,33,49,51,6,25)) {
   length = sum.patient.data %>% ungroup %>% filter(Patient == pt) %>% dplyr::summarise(years = final.endo-initial.endo) %>% pull
   height = as.numeric(max( pg.samp %>% ungroup %>% filter(Patient == pt) %>% dplyr::select(Block) %>% pull ) ) + 1
+
+  #length = 6
+  #height = 5
 
@@
   vigLeg = get.legend(p)
-  p = p + theme(legend.position = 'none')
+  p = p + theme(legend.position = 'none', text = element_text(size=12))
   ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p, width=length, height=height,dpi=300)
@@
 
+penult %>% filter(Status == 'P' & rule %in% c(3,4)) %>% kable(caption='Patients that would be missed if we moved Moderate to Low risk') %>% kable_styling(full_width = F)
+
 penult %>% group_by(Status, rule) %>% tally %>% spread(rule,n) %>% 
@@
 second %>% group_by(Status, rule) %>% tally %>% spread(rule,n) %>% 
+  kable(caption='Patient recommendation by PATH alone in penult endoscopy' ) %>% kable_styling(full_width = F)
+
+second %>% group_by(Status, rule) %>% tally %>% spread(rule,n) %>% 
   left_join(second %>% group_by(Status) %>% tally(name='total'), by='Status') %>%
   mutate_if(is.numeric, funs(  round((./total)*100,1)  )) %>% dplyr::select(-total) %>%
-  kable(caption='Final endoscopy: % patients by recommendation rule based on the max (PATH) per endoscopy') %>% kable_styling(full_width=F)
-
-```
+  kable(caption='Final endoscopy: % patients by recommendation rule based on the max (PATH) per endoscopy, not (RISK)') %>% kable_styling(full_width=F)
+
+```

### 2019-11-20 snp/SNP_Prediction_Analysis.Rmd
bug fix, forgot to output sample names in the tiled files

@@
 library(BarrettsProgressionRisk)
+library(kableExtra)
 
@@
 
+patient.info = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/Analysis/SNP/metadata_T1T2.xlsx') %>% 
+  mutate( UniqueSampleID = paste(PatientID, `Timepoint Code`, sep='_'), Path.Status = Status, PatientID = as.character(PatientID)) %>% 
+  mutate(Path.Status = as.character(ifelse(Pathology %in% c('IMC','HGD'), 'P', Path.Status)))
+
+patient.info %>% filter(PatientID == 524)
+
 rm(plots,performance.at.1se,models,cvs,labels)
@@
 
-qcdata = qcdata %>% as_tibble %>% 
+qcdata = qcdata %>% as_tibble %>%  mutate(Samplename = sub('_$','',Samplename)) %>%
   separate(Samplename, c('PatientID','SampleID','EndoID','Level'), sep='_', remove=F)
+
+qcdata[749, 'Level'] = 'BLD' # Making an assumption here but this is what it looks like
 
@@
 
-qcdata[1,] -> s
-
-qcdata$ASCAT.SCA.ratio = apply(qcdata,1,function(s) {
-  smp = filter( segments.list[[ s$PatientID ]], sample == s$Samplename & chr %in% c(1:22) )
-  smp = smp %>% #rowwise %>% 
-    dplyr::mutate(
+calc.ratio<-function(PatientID, Samplename,Ploidy) {
+  smp = filter( segments.list[[ PatientID ]], sample == Samplename & chr %in% c(1:22) )
+  smp = smp %>% dplyr::mutate(
       #'Total' = nAraw + nBraw,
       'Total' = nMajor + nMinor,
-      'CNV' = round(Total) - round(as.numeric(s$Ploidy)) )
+      'CNV' = round(Total) - round(as.numeric(Ploidy)) )
+  x = filter(smp, CNV != 0 & chr %in% c(1:22))
+  return(sum(x$endpos - x$startpos)/as.numeric(chr.info[22,'genome.length']))
+}
 
-  x = filter(smp, CNV != 0 & chr %in% c(1:22))
-  sum(x$endpos - x$startpos)/as.numeric(chr.info[22,'genome.length'])
-})
+qcdata = qcdata %>% mutate(
+  ASCAT.SCA.ratio = purrr::pmap_dbl(list(PatientID, Samplename, Ploidy), .f=calc.ratio),
+  Level = toupper(Level),
+  SampleType = case_when(
+    grepl('BLD', Level, ignore.case = T) ~ 'Blood Normal',
+    grepl('gastric', Level, ignore.case = T) ~ 'Gastric Normal',
+    TRUE ~ 'BE'
+))
 
-qcdata$SampleType = 'BE'
-qcdata$SampleType[grep('BLD',qcdata$Level, ignore.case=T)] = 'Blood Normal'
-qcdata$SampleType[grep('gastric',qcdata$Level, ignore.case=T)] = 'Gastric Normal'
+sample.list = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/Analysis/SNP/20180604_Reid_1M_SampleList.xlsx', sheet = 2) %>%
+  dplyr::rename('Total.SCA' = 3) %>% mutate(SCA.Ratio = Total.SCA/max(chr.info$genome.length))
 
-patient.info = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/Analysis/SNP/metadata_T1T2.xlsx') %>% 
-  mutate( UniqueSampleID = paste(PatientID, `Timepoint Code`, sep='_'), Path.Status = Status) %>% 
-  mutate(Path.Status = as.character(ifelse(Pathology %in% c('IMC','HGD'), 'P', Path.Status)))
-
-sample.list = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/Analysis/SNP/20180604_Reid_1M_SampleList.xlsx', sheet = 2)
-colnames(sample.list)[3] = 'Total.SCA'
-
-sample.list = sample.list %>% mutate(SCA.Ratio = Total.SCA/max(chr.info$genome.length))
-
-sample.info = qcdata[,c('PatientID','SampleID','EndoID','Level','Samplename','ASCAT.SCA.ratio')]
+sample.info = dplyr::select(qcdata, c('PatientID','SampleID','EndoID','Level','Samplename','ASCAT.SCA.ratio'))
 
@@
 
-sample.info = base::merge(sample.info, patient.info[,c('PatientID','Timepoint','Timepoint Code','Status','Pathology')], by.x=c('PatientID','EndoID'), by.y=c('PatientID','Timepoint Code'))
+sample.info = left_join(sample.info, patient.info[,c('PatientID','Timepoint','Timepoint Code','Status','Pathology')], by=c('PatientID','EndoID' = 'Timepoint Code')) %>%
+  mutate(Status = factor(as.character( ifelse(is.na(Status) & Level %in% c('BLD','GASTRIC'), 'Normal', Status))))
 head(sample.info)
@@
 
-# ASCAT and Xiahong's estimates correlate in 86% of samples, that suggests that the 14% missing is due to the LOH I'm not trying to call
-cor.test(sample.info$ASCAT.SCA.ratio, sample.info$SCA.Ratio)
+ct = cor.test(sample.info$ASCAT.SCA.ratio, sample.info$SCA.Ratio)
+ct$estimate
+```
 
-```
+ASCAT and Xiahong's estimates show a `r round(ct$estimate,2)` correlation. That suggests that the remainder is due to the LOH I'm not trying to call.
 

### 2019-11-20 snp/process_snp.R

@@ -38,7 +38,7 @@ sampletype<-function(level) {
   return(type)
 }
 
-qcdata = qcdata %>% separate(Samplename, c('PatientID','SampleID','EndoID','Level'), remove=F) %>%
+qcdata = as_tibble(qcdata) %>% separate(Samplename, c('PatientID','SampleID','EndoID','Level'), remove=F) %>%
   mutate( Samplename = sub('_$', '', Samplename),  SampleType = purrr::map(Level, sampletype))
 
 segments.list = lapply(segments.list, function(pt) {
@@ -46,16 +46,25 @@ segments.list = lapply(segments.list, function(pt) {
   pt
 })
 
-qcdata$`ASCAT SCA Ratio` = apply(qcdata,1,function(s) {
-  smp = subset(segments.list[[ s[['PatientID']] ]], sample == s[['Samplename']] & chr %in% c(1:22))
-  smp = smp %>% rowwise %>% dplyr::mutate(
+calc.ratio<-function(PatientID, Samplename,Ploidy) {
+  smp = filter( segments.list[[ PatientID ]], sample == Samplename & chr %in% c(1:22) )
+  smp = smp %>% dplyr::mutate(
     #'Total' = nAraw + nBraw,
     'Total' = nMajor + nMinor,
-    'CNV' = round(Total) - round(as.numeric(s[['Ploidy']])) )
-  
-  x = subset(smp, CNV != 0 & chr %in% c(1:22))
-  sum(x$endpos - x$startpos) / chr.info[22,'genome.length',drop=T]
-})
+    'CNV' = round(Total) - round(as.numeric(Ploidy)) )
+  x = filter(smp, CNV != 0 & chr %in% c(1:22))
+  return(sum(x$endpos - x$startpos)/as.numeric(chr.info[22,'genome.length']))
+}
+
+qcdata = qcdata %>% mutate(
+  `ASCAT SCA Ratio` = purrr::pmap_dbl(list(PatientID, Samplename, Ploidy), .f=calc.ratio),
+  Level = toupper(Level),
+#  SampleType = case_when(
+#    grepl('BLD', Level, ignore.case = T) ~ 'Blood Normal',
+#    grepl('gastric', Level, ignore.case = T) ~ 'Gastric Normal',
+#    TRUE ~ 'BE'
+#  )
+) 
 
 lowSCACutoff = median(qcdata %>% filter(SampleType != 'BE') %>% dplyr::select(`ASCAT SCA Ratio`) %>% pull)
 
[…]

### 2019-11-21 Paper_resubmit.Rmd
...

@@
 
-p11Loss %>% filter(value < -1) %>% group_by(Status) %>% tally(name='p11 Loss') %>% full_join(
-  p11 %>% group_by(Status) %>% tally(name='Total Samples') ) %>% 
+p11Loss %>% filter(value <= -1) %>% group_by(Status) %>% tally(name='p11 Loss') %>% 
+  full_join( p11 %>% group_by(Status) %>% tally(name='Total Samples') ) %>% 
   kable(caption='Chr17 p11 loss') %>% kable_styling(full_width = F)
 
-p11Loss %>% group_by(Status, Hospital.Research.ID) %>% tally
+p11Loss %>% filter(value <= -1) %>% dplyr::select(Patient, Status) %>% distinct %>%
+  group_by(Status) %>%  tally(name='p11 Loss') %>%
+  kable(caption='Patients showing p53 loss & p53 IHC') %>% kable_styling(full_width = F)
 
@@
 patient.info %>% filter(Samplename %in% p11Loss$variable) %>% group_by(forcats::fct_explicit_na(p53.Status)) %>% tally %>% 
-  set_names(c('p53 Status', 'n samples')) %>% kable(caption='p53 status for samples with p11 loss') %>% kable_styling(full_width = F)
+  set_names(c('p53 Status', 'n samples')) %>% kable(caption='p53 IHC status for samples with p11 loss') %>% kable_styling(full_width = F)
 
@@
 
-cor.test(pi$p11Loss, (as.numeric(pi$p53.Status)-1))
-```
+ct = cor.test(pi$p11Loss, (as.numeric(pi$p53.Status)-1))
+```
+
+17p11 loss and p53 IHC correlation  R=`r round(ct$estimate,2)`.
 
@@
 
-## Chr 17 Predictive TODO
+## Chr 17 Predictive
 
 One way of looking at the p53 prediction (e.g. Adam Bass) is to see how predictive the p-arm loss in chr17 is.
-```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=10, eval=F}
+```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=10, eval=T}
 ccgenes = read_tsv('~/Data/CosmicCensusGenes.tsv', col_types = cols(.default = col_character())) %>% 
@@
 
-genes = ccgenes %>% dplyr::filter(grepl('TSG|oncogene', `Role in Cancer`) | `Mutation Types` != 'T') %>% 
+# leave out the translocations
+genes = ccgenes %>% dplyr::filter(grepl('TSG|oncogene', `Role in Cancer`) & `Mutation Types` != 'T') %>% 
   dplyr::select(`Gene Symbol`, `Genome Location`, chr, start, end) %>%
@@
   theme(legend.position = 'none', strip.text = element_blank(), text=element_text(size=14))
+g17
 
@@
 for (chr in unique(features$chr)) {  
-  print(chr)
+  #print(chr)
   p = plot.cn.chr(globalCNdf,  chrom=chr, info=chr.lengths, haz=features, genes=genes, label=labeller(chr=label_both, Status=n.status)) +
@@
   dplyr::select(-matches('SLX|Index'), -`Path Notes`) %>%
-  mutate(Status = factor(Status, levels=levels(pg.samp$Status))) %>% ungroup %>% group_by(`Hospital Research ID`) %>% 
+  mutate(Pathology = recode(Pathology, 'GM'='NDBE'), Status = factor(Status, levels=levels(pg.samp$Status))) %>% 
+  ungroup %>% group_by(`Hospital Research ID`) %>% 
   dplyr::mutate(`Weeks Follow Up` = as.double(difftime(`Last Endoscopy`, `Endoscopy`, units='weeks')))
@@
 
+oac.preds %>% group_by(`Hospital Research ID`) %>% 
+  dplyr::summarise( Probability = max(Probability), `Relative Risk` = max(`Relative Risk`), Risk = max(Risk)) %>%
+  group_by(Risk) %>% tally %>% kable(caption='OAC samples used as a positive control') %>% kable_styling(full_width = F)
+
 all.val.preds = all.val.preds %>% filter( !Sample %in% failedQC$samplename &  
@@
 
-val.preds = left_join(all.val.preds, val.info, by=c('Hospital Research ID', 'Sample'='Samplename')) %>% dplyr::select(Patient, everything())
-
-val.preds = val.preds %>% mutate(Pathology = factor(Pathology, levels=c('NDBE','GM','ID','LGD'), ordered=T)) %>% 
+val.preds = left_join(all.val.preds, val.info, by=c('Hospital Research ID', 'Sample'='Samplename')) %>% 
+  dplyr::select(Patient, everything()) %>% mutate(Pathology = factor(Pathology, levels=c('NDBE','ID','LGD'), ordered=T)) %>% 
   mutate(Final.Endoscopy = `Last Endoscopy`, Endoscopy.Year = Endoscopy, Block = factor(Block, ordered=T))
@@
 
+# Risk Categories KM
+
+```{r, fig.width=6, fig.height=8}
+require(survival)
+require(survminer)
+require(GGally)
+
+ps = pg.samp %>% dplyr::select(Patient, Status,months.before.final, Risk) %>%
+  group_by(Patient) %>% mutate(Diagnosis = as.integer(Status)-1)
+
+sfit = survfit( Surv(months.before.final, Diagnosis, type='right') ~ Risk, data=ps, conf.int=0.95, conf.type='log', na.action=na.omit)
+names(sfit$strata) = sub('.*=','', names(sfit$strata))
+
+kmp = survminer::ggsurvplot(sfit, size=1, linetype="strata",risk.table=T, risk.table.col='strata',
+                      conf.int=T, censor=T, legend='none', title='Progression risk over time', xlab='Time (months prior to progression)',
+                      ylab = 'Progression risk',
+                      ggtheme=theme_minimal(), palette=as.vector(BarrettsProgressionRisk::riskColors()), font.legend=c(14,'bold'))
+
+kmp
+
+#ggsave('plots/risk_km.png', plot=print(kmp), height=9, width=8, units = 'in', dpi = 300)
+
+```
+
+```{r}
+vp = val.preds %>% dplyr::select(Patient, Status, months.before.final, Risk) %>% mutate(Patient = as.factor(Patient)) %>% 
+  group_by(Patient) %>% mutate(Diagnosis = as.integer(Status)-1)
+
+merged = bind_rows(ps,vp)
+
+sfit = survfit( Surv(months.before.final, Diagnosis, type='right') ~ Risk, data=merged, conf.int=0.95, conf.type='log', na.action=na.omit)
+names(sfit$strata) = sub('.*=','', names(sfit$strata))
+
+kmp = survminer::ggsurvplot(sfit, size=1, linetype="strata",risk.table=T, risk.table.col='strata',
+                      conf.int=T, censor=T, legend='none', title='Progression Risk', 
+                      subtitle='Training & Validation predictions', xlab='Time (months prior to progression)',
+                      ylab = 'Progression risk',
+                      ggtheme=theme_minimal(), palette=as.vector(BarrettsProgressionRisk::riskColors()), font.legend=c(14,'bold'))
+
+kmp
+
+ggsave('plots/risk_km.png', plot=print(kmp), height=9, width=8, units = 'in', dpi = 300)
+
+```
+
 # Clinical Story
@@
 pt = 34
-p34 = vig.plot(df=subset(pg.samp, Patient == pt),info= subset(patient.info, Patient == pt),cols=riskCols) + 
+p34 = vig.plot(df=filter(pg.samp, Patient == pt),info= filter(patient.info, Patient == pt),cols=riskCols) + 
     theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
@@
 pt = 3
-p3 = vig.plot(df=subset(pg.samp, Patient == pt),info= subset(patient.info, Patient == pt),cols=riskCols) + 
+p3 = vig.plot(df=filter(pg.samp, Patient == pt),info= filter(patient.info, Patient == pt),cols=riskCols) + 
     theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
@@
 pt = 68
-p68 = vig.plot(df=subset(pg.samp, Patient == pt),info= subset(patient.info, Patient == pt),cols=riskCols) + 
+p68 = vig.plot(df=filter(pg.samp, Patient == pt),info= filter(patient.info, Patient == pt),cols=riskCols) + 
     theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
@@
 
-rules = do.call(rbind, lapply(unique(maxP$Patient), function(pt) {
+pt.counts = maxP %>% group_by(Patient, Status) %>% dplyr::filter( length(which(grepl('Endpoint', brks))) > 0  & length(brks) > 1) %>% 
+  ungroup %>% dplyr::select(Patient, Status) %>% distinct %>% group_by(Status) %>% tally
+
+rules = do.call(bind_rows, purrr::map(unique(maxP$Patient), function(pt) {
   r = with(arrange(subset(maxP, Patient == pt), brks), rx(pt, brks,maxP,Pathology,Risk,p53))
   r$Status = unique(subset(maxP, Patient == pt)$Status)
-  r
-}))
-
-rules$t1 = factor(rules$t1, levels=levels(pg.samp$brks), ordered=T)
-rules$t2 = factor(rules$t2, levels=levels(pg.samp$brks), ordered=T)
-
-penult = filter(rules, t2 != 'Endpoint' & (t1 == rev(levels(pg.samp$brks))[2] | t2 == rev(levels(pg.samp$brks))[2] ))
+  return(as_tibble(r))
+})) %>% dplyr::mutate(t1 = factor(t1, levels=levels(pg.samp$brks), ordered=T), t2 = factor(t2, levels=levels(pg.samp$brks), ordered=T)) %>% arrange(Patient, t2)
+
+rules %>% group_by(Patient, Status) %>% dplyr::filter(t2 == 'Endpoint' & length(t1) > 2) %>% 
+  dplyr::select(Patient, Status) %>% distinct %>%  group_by(Status) %>% tally
+
+ids = rules %>% group_by(Patient, Status) %>% dplyr::filter(t2 == 'Endpoint' & length(t1) > 2) %>% dplyr::select(Patient) %>% pull
+
+penult = rules %>% dplyr::filter(Patient %in% ids & t2 != 'Endpoint') %>% arrange(Patient, desc(t1)) %>% group_by(Patient, Status) %>% 
+  dplyr::mutate(gid = 1:n() ) %>% dplyr::filter(gid == 1)
+
+#penult = filter(rules, t2 != 'Endpoint' & (t1 %in% rev(levels(pg.samp$brks))[2] | t2 == rev(levels(pg.samp$brks))[2] ))
+
+penult %>% dplyr::select(Patient, Status) %>% distinct %>% group_by(Status) %>% tally
 
@@
 
+penult %>% filter(Status == 'NP' & rule == 1) %>% kable(caption='NP patients for RFA') %>% kable_styling(full_width = F)
+
+penult %>% group_by(Status) %>% tally(name='total')
+
 penult %>% group_by(Status, rule) %>% tally %>% spread(rule,n) %>% 
-  left_join(penult %>% group_by(Status) %>% tally(name='total'), by='Status') %>%
-  mutate_if(is.numeric, funs(  round((./total)*100,1)  )) %>% dplyr::select(-total) %>%
-  kable(caption='Penult Endoscopy: % patients by rx rules based on the max (P) per endoscopy') %>% kable_styling(full_width=F)
-
-endpoint = subset(rules,  t2 != 'Endpoint') %>% 
-  dplyr::group_by(Patient, Status) %>% 
-  dplyr::summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )
-
-endpoint %>% group_by(Status, rule) %>% tally %>% spread(rule,n) %>% 
-  left_join(endpoint %>% group_by(Status) %>% tally(name='total'), by='Status') %>%
-  mutate_if(is.numeric, funs(  round((./total)*100,1)  )) %>% dplyr::select(-total) %>%
-  kable(caption='Final endoscopy: % patients by recommendation rule based on the max (P) per endoscopy') %>% kable_styling(full_width=F)
+  kable(caption='Penult Endoscopy: # patients by rx rules based on the max (P) per endoscopy') %>% kable_styling(full_width=F)
+
+penult = rules %>% dplyr::filter(Patient %in% ids & t2 != 'Endpoint') %>% arrange(Patient, desc(t1)) %>% group_by(Patient, Status) %>% 
+  dplyr::mutate(gid = 1:n() ) %>% dplyr::filter(gid == 1)
 

### 2019-11-21 snp/SNP_Prediction_Analysis.Rmd
...

@@
 library(BarrettsProgressionRisk)
+library(pander)
 library(kableExtra)
+library(reshape2)
 
@@
 ct = cor.test(sample.info$ASCAT.SCA.ratio, sample.info$SCA.Ratio)
-ct$estimate
 ```
@@
 
-`r nrow(patient.info)` total endoscopies, `r pander(table(patient.info$Status), caption="Split between P and NP")`
+`r nrow(patient.info)` total endoscopies, `r patient.info %>% group_by(Status) %>% tally %>% spread(Status,n) %>% kable(caption="Split between P and NP") %>% kable_styling(full_width=F)`
 
@@
 
-ggplot(sm, aes(Pathology, freq, group=Endoscopy, fill=factor(Endoscopy))) + facet_grid(~Status) + 
+ggplot(sm %>% filter(!Status == 'Normal'), aes(Pathology, freq, group=Endoscopy, fill=factor(Endoscopy))) + facet_grid(~Status) + 
   geom_col(position = position_dodge()) + ylim(0,1) + labs(title='Pathology frequency per timepoint')
@@
 # Quick QC
-In the sWGS data progressor samples show a variance of `r round(pv, 3)` and non-prog variance is `r round(npv,3)`
 
@@
 ```{r, echo=F, warning=F, message=F}
-qcdata = base::merge(qcdata, sample.info[,c('Samplename','Status','Timepoint','SCA.Ratio')], by='Samplename')
+qcdata = left_join(qcdata, sample.info[,c('Samplename','Status','Timepoint','SCA.Ratio')], by='Samplename')
 
@@
   geom_text(data=anno, aes(x, y, label=paste(round(Ratio,2)*100,'% samples',sep='')), nudge_y = -0.0001)
-
-# ggplot(subset(qcdata,SampleType == 'BE'), aes(Timepoint, SCA.Ratio, color=PatientID, group=PatientID)) + geom_point() + geom_line() + 
-#   facet_wrap(~Status, scales='free_x', ncol=2) + labs(title='SCA Ratio by Timepoint') + theme(legend.position = 'none')
 
@@
 } else {
-
-  mergedSegs = NULL
-  mergedArms = NULL
+  ptdirs = list.dirs('~/Data/Reid_SNP/PerPatient', recursive = F)
+  mergedSegs = NULL;  mergedArms = NULL
   length(ptdirs)
@@
 
-    segvals = as.data.frame(data.table::fread(list.files(pt, '*wins_tiled.txt', full.names=T)))
-    armvals = as.data.frame(data.table::fread(list.files(pt, '*arms_tiled.txt', full.names=T)))
+    segvals = read_tsv(list.files(pt, '*wins_tiled.txt', full.names=T), col_types = cols(.default = col_double(), 'X1'= col_character())) %>% dplyr::rename(Samplename = X1)
+    armvals = read_tsv(list.files(pt, '*arms_tiled.txt', full.names=T), col_types = cols(.default = col_double(), 'X1'= col_character())) %>% dplyr::rename(Samplename = X1)
+    #samplenames = segvals$Samplename
 
-    segvals = segment.matrix(segvals)
-    segvals[is.na(segvals)] = mean(segvals,na.rm=T)
+    segvals = segvals %>% set_rownames(.$Samplename) %>% dplyr::select(-Samplename) %>% as.matrix 
+    armvals = armvals %>% set_rownames(.$Samplename) %>% dplyr::select(-Samplename) %>% as.matrix
 
-    armvals = segment.matrix(armvals)
-    armvals[is.na(armvals)] = mean(armvals,na.rm=T)
+    #segvals[is.na(segvals)] = mean(segvals,na.rm=T)
+    #armvals[is.na(armvals)] = mean(armvals,na.rm=T)
 
@@
 
-  save(mergedSegs, mergedArms, file='tmp_seg_pt.Rdata')
+  save(mergedSegs, mergedArms, file='~/Data/Reid_SNP/PerPatient/tmp_seg_pt.Rdata')
 }
@@
 
-# Adjust for ploidy
-for (i in 1:nrow(mergedSegs)) {
-  sample = rownames(mergedSegs)[i]
-  x = unlist(strsplit(sample, '_'))
-  #ploidy = mean(subset(qcdata, SampleType == 'BE' & PatientID == x[1] & EndoID == x[2])$Ploidy)
-  ploidy = subset(qcdata.endo, PatientID == x[1] & EndoID == x[2] & SampleType == 'BE')$Ploidy_mean
-  if (grepl('BLD|gastric', sample, ignore.case = T)) ploidy = 2
-  mergedSegs[i,] = mergedSegs[i,]-(ploidy-1)
-  mergedArms[i,] = mergedArms[i,]-(ploidy-1)
-}
+# Adjust for ploidy -- this was done in the processing
+# for (i in 1:nrow(mergedSegs)) {
+#   sample = rownames(mergedSegs)[i]
+#   x = unlist(strsplit(sample, '_'))
+#   #ploidy = mean(subset(qcdata, SampleType == 'BE' & PatientID == x[1] & EndoID == x[2])$Ploidy)
+#   ploidy = subset(qcdata.endo, PatientID == x[1] & EndoID == x[2] & SampleType == 'BE')$Ploidy_mean
+#   if (grepl('BLD|gastric', sample, ignore.case = T)) ploidy = 2
+#   mergedSegs[i,] = mergedSegs[i,]-(ploidy-1)
+#   mergedArms[i,] = mergedArms[i,]-(ploidy-1)
+# }
 
@@
 ```{r rnormals, echo=F, warning=F, message=F, fig.width=10, fig.height=8, eval=T}
+
+adjust.cols<-function(df, means, sds) {
+  nms = colnames(df)
+
+  for (n in nms)
+    df[,n] = BarrettsProgressionRisk:::unit.var(df[,n]+0.90, means[[n]], sds[[n]])
+
+  return(df)  
+}
+
 randNormal = sample( grep('BLD',rownames(mergedSegs)[nm.rows], value=T), 1)
@@
 lims = range(segs$medLRR, mergedSegs[randNormal,])
-# grid.arrange(
-#   ggplot(segs) + facet_grid(~chr, space='free_x', scales='free') + ylim(lims) +
-#     geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), size=3) + 
-#     theme(axis.text.x=element_blank() ) + labs(title='ASCAT median LRR', subtitle=pt),
-#   ggplot(melt(as.matrix(normalSegs[randNormal,])), aes(Var1, value)) + geom_point() + ylim(lims) +
-#     labs(title='CN Adjusted & winsorized') + theme(axis.text.x=element_blank()),
-# top='Random BLD or gastric normal', bottom=randNormal)
+
+grid.arrange(
+  ggplot(segs) + facet_grid(~chr, space='free_x', scales='free')  +
+    geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), size=3) +
+    theme(axis.text.x=element_blank() ) + labs(title='Median LRR, ASCAT segments', subtitle=pt),
+
+  ggplot(melt(as.matrix(mergedSegs[sub('\\.LogR','',randNormal),])), aes(Var1, value)) + geom_point() +
+        theme(axis.text.x=element_blank()) + labs(title='Ploidy adjusted CN, binned'),
+
+   ggplot(melt(t(adjust.cols(mergedSegs, z.mean, z.sd)[randNormal,])), aes(x=Var2, y=value)) + geom_point() + 
+    theme(axis.text.x=element_blank()) + labs(title='Col adjusted CN, binned'),
+ top='Random BLD or gastric normal', bottom=randNormal)
 
@@
 
-#lims = range(segs$medLRR, mergedSegs[sub('\\.LogR','',randBE),])
-# grid.arrange(
-#   ggplot(segs) + facet_grid(~chr, space='free_x', scales='free')  +
-#     geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), size=3) + 
-#     theme(axis.text.x=element_blank() ) + labs(title='Median LRR, ASCAT segments', subtitle=pt), 
-#   ggplot(melt(as.matrix(mergedSegs[sub('\\.LogR','',randBE),])), aes(Var1, value)) + geom_point() +
-#     theme(axis.text.x=element_blank()) + labs(title='Winsorized LRR, binned'),  
-#  ggplot(melt(adjust.cols(mergedSegs[sub('\\.LogR','',randBE),,drop=F], z.mean, z.sd)), aes(Var2, value)) + geom_point() + theme(axis.text.x=element_blank()) + labs(title='Col adjusted LRR, binned'),
-#top='BE', bottom=randBE)
+lims = range(segs$medLRR, mergedSegs[sub('\\.LogR','',randBE),])
+grid.arrange(
+  ggplot(segs) + facet_grid(~chr, space='free_x', scales='free')  +
+    geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), size=3) +
+    theme(axis.text.x=element_blank() ) + labs(title='Median LRR, ASCAT segments', subtitle=pt),
+
+  ggplot(melt(as.matrix(mergedSegs[sub('\\.LogR','',randBE),])), aes(Var1, value)) + geom_point() +
+    theme(axis.text.x=element_blank()) + labs(title='Ploidy adjusted CN, binned'),
+
+   ggplot(melt(t(adjust.cols(mergedSegs, z.mean, z.sd)[randBE,])), aes(x=Var2, y=value)) + geom_point() + 
+    theme(axis.text.x=element_blank()) + labs(title='Col adjusted CN, binned'),
+top='BE', bottom=randBE)
 
@@
 training.arms = as.matrix(training[,-1]); rownames(training.arms) = training$X1
+
+median(training.segs)
+median(mergedSegs)
+median(mergedSegs + 0.98)
+
+median(training.arms)
+median(mergedArms)
+median(mergedArms + 0.98)
+
+mergedSegs = mergedSegs + 0.98
+mergedArms = mergedArms + 0.98
 
@@
 
-preds = cbind.data.frame(rownames(prob),prob, rr, stringsAsFactors=F)
-colnames(preds) = c('Samplename','Prob', 'RR')
+preds = bind_cols('Samplename'=rownames(prob),'Prob'=prob[,1],'RR'=rr[,1]) %>%
+  mutate(Risk = case_when(
+    Prob <= 0.3 ~ 'Low',
+    Prob > 0.3 & Prob < 0.5 ~ 'Moderate',
+    Prob >= 0.5 ~ 'High' 
+))
 
-preds = preds %>% mutate(Status = as.character(ifelse(Pathology %in% c('HGD','IMD'), 'P', Status)))
+normals = filter(preds, grepl('BLD|gastric', Samplename, ignore.case=T))
+table(normals$Risk)
 
-preds %>% filter(grepl('BLD|gastric',Level,ignore.case=T)) %>% group_by(Risk, Level) %>% tally %>% spread(Risk,n)
+preds = filter(preds, !grepl('BLD|gastric', Samplename, ignore.case=T))
 
-preds %>% filter(!grepl('BLD|gastric',Level,ignore.case=T)) %>% group_by(Status, Risk, Pathology) %>% tally %>% spread(Risk,n)
+preds = preds %>% left_join(patient.info %>% dplyr::select(UniqueSampleID, Pathology, Status), by=c('Samplename'='UniqueSampleID')) %>%
+   mutate(Status = as.character(ifelse(Pathology %in% c('HGD','IMD'), 'P', Status)))
 
-# offset = log(0.0225)
-# preds = preds %>% dplyr::mutate(
-#   'Adj. RR'=RR+offset,
-#   'Adj. Prob'=1/(1+exp(-RR+abs(offset)) )
-# )
+preds %>% group_by(Status, Risk) %>% tally %>% spread(Risk,n)
 
-preds.normal = preds %>% filter(grepl('BLD|gastric', Level, ignore.case=T))
+hist(preds$Prob)
 
-preds.normal %>% group_by(Risk) %>% tally
-
-length(which(preds.normal$Prob > 0.5))/nrow(preds.normal) # FPR
-length(which(preds.normal$Prob < 0.5))/nrow(preds.normal)
+length(which(normals$Prob > 0.5))/nrow(normals) # FPR
+length(which(normals$Prob < 0.5))/nrow(normals)
 
@@
 
-preds = sample.info %>% mutate(Sample = paste(PatientID,EndoID,sep='_')) %>% left_join(preds, by=c('Sample'='Samplename'))
+#preds = sample.info %>% mutate(Sample = paste(PatientID,EndoID,sep='_')) %>% left_join(preds, by=c('Sample'='Samplename'))
 
@@
 
-preds = preds %>% 
-  mutate(Risk = ifelse(Prob > 0.5, 'High', 'Low')) %>% 
-  mutate(Risk = factor(Risk, levels=c('Low','High'), ordered=T))
-
 preds %>% group_by(Status, Risk) %>% tally %>% spread(Risk, n) %>%
   left_join(preds %>% group_by(Status) %>% tally, by='Status') %>%
-  mutate_if(is.numeric, funs( round(./n,2))) %>% 
+  mutate_if(is.numeric, list( round(./n,2))) %>% 
   dplyr::select(-n)
@@
 
-rocR = pROC::roc(Status ~ Risk, data=preds, auc=T, ci=T, of='thresholds',transpose=T)
+preds = preds %>%
+  mutate(LH = ifelse(Prob > 0.5, 'High', 'Low')) %>%
+  mutate(LH = factor(LH, levels=c('Low','High'), ordered=T))
+
+rocR = pROC::roc(Status ~ LH, data=preds, auc=T, ci=T, of='thresholds',transpose=T)
 roc.plot(rocR) + labs(title='SNP AUC by discretized risks (low/high)')
@@
 A quick look at the two random samples from earlier.  
-`r pander(subset(preds.normal, Samplename == randNormal), caption='Random normal')`
+`r pander(subset(normals, Samplename == randNormal), caption='Random normal')`
 
-The random BE should show a very high probability, but for some reason is missed entirely:
+The random BE should show a very high probability, and it does...
 `r pander(subset(preds, Samplename == randBE), caption='Random BE')`
@@
 
-pred.qc = merge(preds,qcdata.endo, by=c('PatientID','EndoID'))
+#pred.qc = left_join(preds, qcdata.endo %>% dplyr::select(-SampleType), by=c('PatientID', 'EndoID', 'Status')) %>% data.frame
 
-#dim(pred.qc)
-#table(pred.qc$Status)
+#grid.arrange(
+#  ggplot(pred.qc, aes(ASCAT.SCA.ratio_mean, Prob, color=Status)) + geom_point(),
+ # ggplot(pred.qc, aes(Ploidy_max, Prob, color=Status)) + geom_point(),
+#ncol=2)
 
-grid.arrange(
-  ggplot(pred.qc, aes(ASCAT.SCA.ratio_mean, Prob, color=Status)) + geom_point(),
-  ggplot(pred.qc, aes(Ploidy_max, Prob, color=Status)) + geom_point(),
-ncol=2)
-
-pander(cor.test(pred.qc$ASCAT.SCA.ratio_mean, pred.qc$Prob))
+#pander(cor.test(pred.qc$ASCAT.SCA.ratio_mean, pred.qc$Prob))
 ```
 
-```{r, warning=F, fig.height=8}
+```{r, warning=F, fig.height=8, eval=F}
 prog = subset(pred.qc, Status == 'P')
@@
 
-```{r roc, echo=T}
+```{r roc, echo=T, eval=F}
 pred.qc = base::merge( pred.qc, patient.info[,c('PatientID', 'Timepoint Code','Pathology')], by.x=c('PatientID', 'EndoID'), by.y=c('PatientID','Timepoint Code') )
@@
 
-```{r echo=F}
+```{r echo=F, eval=F}
 pred.qc.ex = subset(pred.qc, SampleType == 'BE' & !exclude & !wgd & !lowsca)
@@
 
-```{r regression, echo=F, message=F, eval=T}
+```{r regression, echo=F, message=F, eval=F}
 folds = 10; splits = 5

### 2019-12-09 Compare Models.Rmd
generating all new plots
cleaning things up

@@
 kb.models = list.files(data)
-
-#list.files(f,full.names = T)
 
@@
 
-# Test Data
-
-Using the alpha=0.9 model for each and evaluate the risks based on the LOO performance.
-
-```{r fig.height=10,fig.width=10, eval=F}
-## 0.5
-test.preds.per.kb05 = purrr::map(list.files(data,full.names = T), function(f) {
-  read_tsv(list.files(f, 'test_patients_preds_0.5.tsv', full.names = T), col_types = 'cdccdccclddccccccddccccdddccdddcdd') %>% dplyr::select(Hospital.Research.ID, Status, Endoscopy.Year, Pathology, Prediction, RR) %>% mutate(kb = basename(f))
-}) 
-names(test.preds.per.kb05) = basename(list.files(data))
-
-test.preds.per.kb05 = lapply(test.preds.per.kb05, function(kb) {
-  kb %>% mutate(quants = cut(Prediction, breaks=seq(0,1,0.1), include.lowest = T)) %>% left_join(loo.pred.confidence[[unique(kb$kb)]] %>% dplyr::select(quants, Risk), by='quants') %>%
-      mutate(Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered=T)) 
-})
-
-test.samples = all.patient.info %>% filter(Set == 'Test') %>% group_by(Status) %>% tally
-
-test.counts05 = lapply(test.preds.per.kb05, function(kb) {
-  kb %>% group_by(Status, Risk) %>% tally %>% spread(Risk,n) %>% ungroup %>%
-    mutate(kb = unique(kb$kb)) %>% left_join(test.samples, by='Status') %>% 
-    mutate_if(is.numeric, funs(./n))
-})
-
-plist = lapply(test.counts05, function(kb) {
-  tt = unique(kb$kb)
-  kb = kb  %>% ungroup %>% dplyr::select(-kb,-n)
-
-  ggplot(melt(kb, id.vars=c('Status')), aes(variable, value, group=Status, fill=Status)) + ylim(0,1) + 
-    geom_bar(stat='identity', position='dodge') + geom_text(aes(label=round(value,2)), position = position_dodge(width=1)) +
-    labs(title=tt, x='', y='ratio') + theme_minimal()
-})
-
-do.call(grid.arrange, c(plist, top='Alpha=0.5'))
-```
-
 ## Per-sample alpha=0.9
@@
 ```
-
-# Downsampled Data
-
-## Per-sample alpha=0.9
-```{r, fig.width=8, fig.height=12, eval=F}
-
-ds.info = read_excel('~/Data/BarrettsProgressionRisk/QDNAseq/all_downsampled/downsampled_ids.xlsx')
-
-ddir = '~/Data/BarrettsProgressionRisk/Analysis/downsampled/'
-
-ds.preds =   purrr::map(list.files(ddir, recursive = F, full.names = T), function(dir) {
-    read_tsv(list.files(paste0(dir,'/predictions/0.9'), 'predictions.tsv', full.names = T), col_types = cols(
-  Sample = col_character(), Probability = col_double(), `Relative Risk` = col_double(),
-  Risk = col_character(), PatientID = col_character(), Status = col_character(),
-  Tissue = col_character(), `Expected Risk` = col_character(), Grade_patient = col_character(),
-  Grade_biopsy = col_character(), Study = col_character(), Type = col_character(), Endoscopy = col_date(format = "")
-)) %>% dplyr::select(-Risk, -Status) %>% mutate(quants = cut(Probability, breaks=seq(0,1,0.1), include.lowest = T)) %>% 
-  left_join(loo.pred.confidence[['50kb']] %>% dplyr::select(quants, Risk), by='quants') %>%
-    mutate(Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered=T), kb = basename(dir)) %>% left_join(ds.info %>% dplyr::select(`Illumina ID`, Status), by=c('Sample' = 'Illumina ID'))
-})
-names(ds.preds) = list.files(ddir)
-
-ds.counts = lapply(ds.preds, function(kb) {
-  totals = kb %>% group_by(Status) %>% tally
-
-  kb %>% group_by(Status, Risk) %>% tally %>% spread(Risk,n) %>% ungroup %>%
-    mutate(kb = unique(kb$kb)) %>% left_join(totals, by='Status') %>% 
-    mutate_if(is.numeric, funs(./n)) %>% dplyr::select(-n) %>% 
-    mutate_if(is.numeric, funs(ifelse(is.na(.), 0, .)))
-})
-
-plist = lapply(ds.counts, function(kb) {
-  tt = unique(kb$kb)
-  kb = kb  %>% ungroup %>% dplyr::select(-kb)
-
-  ggplot(melt(kb, id.vars=c('Status')), aes(variable, value, group=Status, fill=Status)) + ylim(0,1) + 
-    geom_bar(stat='identity', position='dodge') + geom_text(aes(label=round(value,2)), position = position_dodge(width=1)) +
-    labs(title=tt, x='', y='ratio') + theme_minimal()
-})
-
-do.call(grid.arrange, c(plist[c('50kb','100kb','15kb','500kb')], top='Downsampled'))
-
-```
-
-```{r}
-
-ds.preds$`50kb` %>% group_by(Type, Risk) %>% tally %>% spread(Risk,n) %>% mutate_if(is.numeric, funs(ifelse(is.na(.), 0, .))) %>% 
-  kable(caption = 'Downsampled types') %>% kable_styling(full_width = F)
-
-```

### 2019-12-09 Paper_resubmit.Rmd
cleaning things up

@@
 ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)
+
+cvRR %>% mutate(`Gain/Loss` = ifelse(coef > 0 ,'Gain', 'Loss')) %>% 
+  dplyr::select(label, hazards, `Gain/Loss`) %>% arrange(desc(hazards)) %>%
+  write_tsv('plots/coef_effect.tsv')
+
 ```
@@
 ```{r leaveoneout, echo=F, message=F, warning=F, fig.width=5, fig.height=5}
+# loo_residuals = do.call(bind_rows, purrr::map(list.files(paste0(dir, '/Analysis/pcf_perPatient/50kb'), 'residuals.tsv', recursive = T, full.names = T), function(f) {
+#   read_tsv(f, col_types = 'ccddddl')
+# }))
+# 
+# failedQC = loo_residuals %>% filter(varMAD_median > 0.008)
+
 file = paste(cache.dir, 'loo_0.9.Rdata', sep='/')
@@
 roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds',transpose=T)
+
+pROC::ci.auc(roc$auc)
+
 ggsave(filename='plots/auc/training_auc.png', plot=roc.plot(roc), width=6.5, height=4.5, units='in', dpi=300)
@@
 
+pROC::ci.auc(p53roc$auc)
+
 p53roc$model = 'p53 IHC'
@@
 pathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('LGD','HGD','IMC'), 1, 0) )
+pROC::ci.auc(pathroc$auc)
+
 pathroc$model = 'HGD & LGD'
@@
 
+#### By pathology?
 ```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
@@
 
-rp = plot.risk.by.path(pg.samp)
+rp = plot.risk.by.path(pg.samp %>% ungroup)
 rp
@@
 
-Using the discretized risk what is the AUC?
+#### Using the discretized risk what is the AUC?
 
@@
 
+One way of looking at the p53 prediction (e.g. Adam Bass) is to see how predictive the p-arm loss in chr17 is.
+
 ## p arm
@@
 
+pi %>%  mutate(p53.Status = recode_factor(p53.Status, '1'='aberrant','0'='normal')) %>% dplyr::group_by(p11Loss, p53.Status) %>% tally %>% spread(p53.Status, n)
+
 ct = cor.test(pi$p11Loss, (as.numeric(pi$p53.Status)-1))
@@
 17p11 loss and p53 IHC correlation  R=`r round(ct$estimate,2)`.
+
+## q arm
+```{r, echo=T}
+
+q11 = filter(globalCNdf, chr == 17 & start > 1 & seg.type == 'arms')
+q11gain = q11 %>% filter(value > 1)
+
+q11gain %>% filter(value >= 1) %>% group_by(Status) %>% tally(name='q11 Gain') %>% 
+  full_join( q11 %>% group_by(Status) %>% tally(name='Total Samples') ) %>% 
+  kable(caption='Chr17 q11 gain') %>% kable_styling(full_width = F)
+
+q11gain %>% filter(value >= 1) %>% dplyr::select(Patient, Status) %>% distinct %>%
+  group_by(Status) %>%  tally(name='q11 Gain') %>%
+  kable(caption='Patients showing q gain') %>% kable_styling(full_width = F)
+
+q11Path = patient.info %>% filter(Samplename %in% q11gain$variable) %>% dplyr::group_by(Pathology) %>% tally(name='samples')
+
+q11Path %>% dplyr::mutate('ratio'=round(samples/sum(q11Path$samples),2)) %>% 
+kable(caption='Ratio of all Chr17 q11 gain samples by pathology' ) %>% kable_styling(full_width = F)
+```
 
@@
 roc.plot(roc) + labs(title='Validation cohort')
+pROC::ci.auc(roc$auc)
+
 ggsave('plots/validation/validatation_roc.png', plot=roc.plot(roc) + labs(title='Validation cohort'), width=5, height=5, units='in',dpi=300)
@@
 ```{r}
-rp = plot.risk.by.path(val.preds) + labs(title='Validation samples predicted by pathology')  
+rp = plot.risk.by.path(val.preds) + labs(title='Validation samples classified by pathology')  
 rp
@@
 kmp = survminer::ggsurvplot(sfit, size=1, linetype="strata",risk.table=T, risk.table.col='strata',
-                      conf.int=T, censor=T, legend='none', title='Progression risk over time', xlab='Time (months prior to progression)',
+                      conf.int=T, censor=T, legend='none', title='Risks over time', xlab='Time (months prior to final sample)',
                       ylab = 'Progression risk',
@@
 kmp = survminer::ggsurvplot(sfit, size=1, linetype="strata",risk.table=T, risk.table.col='strata',
-                      conf.int=T, censor=T, legend='none', title='Progression Risk', 
-                      subtitle='Training & Validation predictions', xlab='Time (months prior to progression)',
+                      conf.int=T, censor=T, legend='none', 
+                      title='Risks over time', 
+                      subtitle='Training & Validation predictions',
+                      xlab='Time (months prior to final sample)',
                       ylab = 'Progression risk',
@@
 
-ggsave('plots/risk_km.png', plot=print(kmp), height=9, width=8, units = 'in', dpi = 300)
+ggsave('plots/risk_km.png', plot=print(kmp), height=8, width=7, units = 'in', dpi = 300)
 
@@
 
+#### From the penultimate endoscopy
 ```{r, rx, echo=F, warning=F, message=T, fig.width=8, fig.height=5}
-
 rx<-function(Patient, brks, Prediction, Pathology, Risk, p53.Status) {
@@
 
-pg.samp = pg.samp %>% mutate(p53.Status = factor(p53.Status, levels=c(0,1), ordered=T))
-
-maxP = pg.samp %>% dplyr::group_by(Patient, Status, brks) %>% 
-  dplyr::summarise( 'maxP'=max(Prediction), 'Risk'=max(Risk), 'Pathology'=max(Pathology), 'p53'=max(p53.Status,na.rm=T) )
-
-pt.counts = maxP %>% group_by(Patient, Status) %>% dplyr::filter( length(which(grepl('Endpoint', brks))) > 0  & length(brks) > 1) %>% 
-  ungroup %>% dplyr::select(Patient, Status) %>% distinct %>% group_by(Status) %>% tally
-
-rules = do.call(bind_rows, purrr::map(unique(maxP$Patient), function(pt) {
-  r = with(arrange(subset(maxP, Patient == pt), brks), rx(pt, brks,maxP,Pathology,Risk,p53))
-  r$Status = unique(subset(maxP, Patient == pt)$Status)
-  return(as_tibble(r))
-})) %>% dplyr::mutate(t1 = factor(t1, levels=levels(pg.samp$brks), ordered=T), t2 = factor(t2, levels=levels(pg.samp$brks), ordered=T)) %>% arrange(Patient, t2)
-
-rules %>% group_by(Patient, Status) %>% dplyr::filter(t2 == 'Endpoint' & length(t1) > 2) %>% 
-  dplyr::select(Patient, Status) %>% distinct %>%  group_by(Status) %>% tally
-
-ids = rules %>% group_by(Patient, Status) %>% dplyr::filter(t2 == 'Endpoint' & length(t1) > 2) %>% dplyr::select(Patient) %>% pull
-
-penult = rules %>% dplyr::filter(Patient %in% ids & t2 != 'Endpoint') %>% arrange(Patient, desc(t1)) %>% group_by(Patient, Status) %>% 
-  dplyr::mutate(gid = 1:n() ) %>% dplyr::filter(gid == 1)
-
-#penult = filter(rules, t2 != 'Endpoint' & (t1 %in% rev(levels(pg.samp$brks))[2] | t2 == rev(levels(pg.samp$brks))[2] ))
-
-penult %>% dplyr::select(Patient, Status) %>% distinct %>% group_by(Status) %>% tally
-
-penult %>% filter(Status == 'P' & rule %in% c(3,4)) %>% kable(caption='Patients that would be missed if we moved Moderate to Low risk') %>% kable_styling(full_width = F)
-
-penult %>% filter(Status == 'NP' & rule == 1) %>% kable(caption='NP patients for RFA') %>% kable_styling(full_width = F)
-
-penult %>% group_by(Status) %>% tally(name='total')
-
-penult %>% group_by(Status, rule) %>% tally %>% spread(rule,n) %>% 
+pg.samp = pg.samp %>% mutate(p53.Status = factor(p53.Status, levels=c(0,1), ordered=T), Probability = Prediction)
+
+max.rules = purrr::map(unique(pg.samp$Patient), function(pt) {
+    preds = pg.samp %>% filter(Patient == pt) %>% arrange(Endoscopy.Year, months.before.final)
+    endo = BarrettsProgressionRisk:::.setUpRxTablePerEndo(preds %>% dplyr::mutate(Endoscopy = months.before.final, Sample = 'Samplename'), 'max', verbose=F) %>% mutate(Endoscopy = Endoscopy*-1)
+    rx = BarrettsProgressionRisk:::.apply.rules(endo, 'Endoscopy', pred.confidence )
+
+    left_join(rx, endo %>% dplyr::select(Endoscopy, months.before.final, Endoscopy.Year, Probability), by=c('Time 1'='Endoscopy')) %>%
+      mutate(`Time 2` = ifelse(months.before.final <= 0,'Endpoint',`Time 2`)) %>% 
+      mutate(Patient = pt) %>% mutate_at( vars(matches('Rule|Time')), funs(as.character) ) %>%
+      dplyr::select(Patient, months.before.final, Endoscopy.Year, Probability, everything()) 
+})
+names(max.rules) = unique(pg.samp$Patient)
+
+progressors = max.rules[as.character(sum.patient.data %>% filter(Status == 'P') %>% dplyr::select(Patient) %>% pull)]
+nonprog = max.rules[as.character(sum.patient.data %>% filter(Status == 'NP') %>% dplyr::select(Patient) %>% pull)]
+
+pids = names(which(sapply(progressors, function(x) {
+  nrow(x) >= 1 & grepl('Endpoint', x[nrow(x),'Time 2'])
+})))
+
+nids = names(which(sapply(nonprog, function(x) {
+  nrow(x) >= 1 & grepl('Endpoint', x[nrow(x),'Time 2'])
+})))
+
+penult = bind_rows( do.call(bind_rows, progressors[pids]) %>% mutate(Status = 'P'),
+                    do.call(bind_rows, nonprog[nids]) %>% mutate(Status = 'NP') )
+
+penult = penult %>% filter(`Time 2` != 'Endpoint') %>% arrange(Patient, months.before.final) %>% group_by(Patient, Status) %>% dplyr::mutate(gid = 1:n() ) %>% dplyr::filter(gid == 1)
+
+penult %>% dplyr::select(Patient, Status) %>% distinct %>% group_by(Status) %>% tally %>% kable(caption='n patients with at least two consecutive endoscopies prior to endpoints') %>%
+  kable_styling(full_width = F)
+
+penult %>% filter(Status == 'P' & Rule %in% c(3,4)) %>% kable(caption='Prog patients that would be missed if we moved Moderate to Low risk') %>% kable_styling(full_width = F)
+
+penult %>% filter(Status == 'NP' & Rule == 1) %>% kable(caption='NP patients for RFA') %>% kable_styling(full_width = F)
+
+penult %>% group_by(Status, Rule) %>% tally %>% spread(Rule,n) %>% 
   kable(caption='Penult Endoscopy: # patients by rx rules based on the max (P) per endoscopy') %>% kable_styling(full_width=F)
 
-penult = rules %>% dplyr::filter(Patient %in% ids & t2 != 'Endpoint') %>% arrange(Patient, desc(t1)) %>% group_by(Patient, Status) %>% 
-  dplyr::mutate(gid = 1:n() ) %>% dplyr::filter(gid == 1)
-
-path.rx<-function(Patient, brks, Pathology) {
-  rules = as.data.frame(matrix(ncol=0, nrow=0))
-  # Consecutive
-  for (i in 1:length(brks)) {
-    #risks = table(Pathology[i:(i+1)])
-    rule = 'None'
-    if ( length(which(grepl('HGD|IMC', Pathology[i:(i+1)]))) > 0  |
-         length(which(grepl('LGD', Pathology[i:(i+1)]))) > 1 ) {
-      rule = 1
-    } else if ( length(which(grepl('LGD', Pathology[i:(i+1)]))) > 0 ) {
-      rule = 2
-    } else if ( length(which(grepl('BE|NDBE|ID', Pathology[i:(i+1)]))) > 1  ) {
-      rule = 3
-    }
-  rules = rbind(rules, cbind( as.character(brks[i]), as.character(brks[(i+1)]), rule))
-  }  
-  colnames(rules) = c('t1','t2', 'rule')
-  rules$Patient = Patient
-  rules
-}
-
-rules2 = do.call(rbind, lapply(unique(maxP$Patient), function(pt) {
-  r = with(arrange(subset(maxP, Patient == pt), brks), path.rx(pt, brks,Pathology))
-  r$Status = unique(subset(maxP, Patient == pt)$Status)
-  r
-}))
-rules2$t1 = factor(rules2$t1, levels=levels(pg.samp$brks), ordered=T)
-rules2$t2 = factor(rules2$t2, levels=levels(pg.samp$brks), ordered=T)
-
-second = subset(rules2,  t2 != 'Endpoint') %>% 
-  dplyr::group_by(Patient, Status) %>% 
-  dplyr::summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )
-
-second %>% group_by(Status, rule) %>% tally %>% spread(rule,n) %>% 
-  kable(caption='Patient recommendation by PATH alone in penult endoscopy' ) %>% kable_styling(full_width = F)
-
-second %>% group_by(Status, rule) %>% tally %>% spread(rule,n) %>% 
-  left_join(second %>% group_by(Status) %>% tally(name='total'), by='Status') %>%
-  mutate_if(is.numeric, funs(  round((./total)*100,1)  )) %>% dplyr::select(-total) %>%
-  kable(caption='Final endoscopy: % patients by recommendation rule based on the max (PATH) per endoscopy, not (RISK)') %>% kable_styling(full_width=F)
-
-```
+```
+
+#### From the second endoscopy
+
+```{r}
+pids = names(which(sapply(progressors, function(x) {
+  nrow(x) > 1 & !grepl('Endpoint', x[2,'Time 2']) & !grepl('None', x[2,'Rule'])
+})))
+
+nids = names(which(sapply(nonprog, function(x) {
+  nrow(x) > 1 & !grepl('Endpoint', x[2,'Time 2']) & !grepl('None', x[2,'Rule'])
+})))
+
+second = bind_rows( do.call(bind_rows, progressors[pids]) %>% mutate(Status = 'P'),
+                    do.call(bind_rows, nonprog[nids]) %>% mutate(Status = 'NP') )
+
+second = second %>% group_by(Patient, Status) %>% arrange(`Time 2`) %>% 
+  dplyr::mutate(gid = 1:n() ) %>% dplyr::filter(gid == 2)
+
+second %>% dplyr::select(Patient, Status) %>% distinct %>% group_by(Status) %>% tally %>% kable(caption='n patients with at least two years follow-up') %>%
+  kable_styling(full_width = F)
+
+second %>% group_by(Status, Rule) %>% tally %>% spread(Rule,n) %>% 
+  kable(caption='Second Endoscopy: # patients by rx rules based on the max (P) per endoscopy') %>% kable_styling(full_width=F)
+
+second %>% filter(Status == 'P' & Rule == '1') %>% ungroup %>% filter(months.before.final < 120) %>%
+  dplyr::summarise_at(vars(months.before.final), list(~min(.), ~max(.), ~mean(.), ~median(.)))
+
+second %>% filter(Status == 'NP' & Rule == '1') %>% ungroup  %>%
+  dplyr::summarise_at(vars(months.before.final), list(~min(.), ~max(.), ~mean(.), ~median(.)))
+
+```

### 2019-12-09 snp/SNP_Prediction_Analysis.Rmd
cleaning things up

@@
 
+dir.create('plots/snp',recursive = T, showWarnings = F)
+
+ggplot2::theme_set(theme_minimal())
+
 chr.info = BarrettsProgressionRisk:::chrInfo(build='hg19')
 
-model.dir = '~/Data/Ellie/Analysis/models_5e6_all/50kb'
-load(paste0(model.dir,'/model_data.Rdata'), verbose = F)
+model.dir = '~/Data/BarrettsProgressionRisk/Analysis/models_5e6_all/50kb'
+load(paste0(model.dir,'/model_data.Rdata'), verbose = T)
 load(paste0(model.dir,'/all.pt.alpha.Rdata'), verbose = F)
 
+train.labels = labels
 alpha = '0.9'
@@
 patient.info = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/Analysis/SNP/metadata_T1T2.xlsx') %>% 
-  mutate( UniqueSampleID = paste(PatientID, `Timepoint Code`, sep='_'), Path.Status = Status, PatientID = as.character(PatientID)) %>% 
-  mutate(Path.Status = as.character(ifelse(Pathology %in% c('IMC','HGD'), 'P', Path.Status)))
+  mutate( UniqueSampleID = paste(PatientID, `Timepoint Code`, sep='_'), Path.Status = Status, PatientID = as.character(PatientID))
+#%>%  group_by(PatientID) %>% mutate(Path.Status = as.character(ifelse(length(which(grepl('IMC|HGD',Pathology))) > 0, 'P', Path.Status)))
+
+path.np.ids = patient.info %>% filter(Pathology %in% c('HGD', 'IMC') & Status == 'NP') %>% dplyr::select(PatientID) %>% distinct %>% pull
+patient.info = patient.info %>% mutate(Path.Status = ifelse(PatientID %in% path.np.ids, 'P', Path.Status))
 
@@
 
+# swgs.seg.count = read_tsv('~/Data/BarrettsProgressionRisk/Analysis/pcf_perPatient/50kb/segment_counts.tsv', col_types = 'dc')
+# 
+# # All segments
+# seg.counts = do.call(bind_rows, purrr::map(segments.list, function(sl) {
+#   sl %>% dplyr::mutate(Total = nMajor+nMinor) %>% group_by(sample, Total) %>% 
+#     tally %>% #filter(Total != 2) %>% 
+#     ungroup %>% group_by(sample) %>% dplyr::summarise(n = sum(n))
+# }))
+
+#plot(seg.counts$n)
+#points(swgs.seg.count$`n segs`, col='red')
+
 calc.ratio<-function(PatientID, Samplename,Ploidy) {
@@
   SampleType = case_when(
-    grepl('BLD', Level, ignore.case = T) ~ 'Blood Normal',
-    grepl('gastric', Level, ignore.case = T) ~ 'Gastric Normal',
+    grepl('BLD|gastric', Level, ignore.case = T) ~ 'Blood/Gastric Normal',
     TRUE ~ 'BE'
@@
 
-sample.info = left_join(sample.info, patient.info[,c('PatientID','Timepoint','Timepoint Code','Status','Pathology')], by=c('PatientID','EndoID' = 'Timepoint Code')) %>%
+sample.info = left_join(sample.info, dplyr::select(patient.info,c('PatientID','Timepoint','Timepoint Code','Status','Pathology')), by=c('PatientID','EndoID' = 'Timepoint Code')) %>%
   mutate(Status = factor(as.character( ifelse(is.na(Status) & Level %in% c('BLD','GASTRIC'), 'Normal', Status))))
@@
 
+#seg.counts %>% dplyr::summarise_at(vars(-`2`), )
+#seg.counts = left_join(seg.counts, dplyr::select(qcdata,c(Samplename, Ploidy, Purity, ASCAT.SCA.ratio, SampleType)), by=c('sample'='Samplename'))
+# seg.counts %>% filter(ASCAT.SCA.ratio < 0.01 & Purity > 0.98)
+# ggplot(filter(seg.counts, !is.na(Ploidy)), aes(n,ASCAT.SCA.ratio)) + facet_grid(~SampleType) + ylim(0,0.05) + xlim(0,200) +
+#  geom_point(aes(color = ASCAT.SCA.ratio < 0.01 & Purity > 0.98)) + theme(legend.position = 'none')
+
 ct = cor.test(sample.info$ASCAT.SCA.ratio, sample.info$SCA.Ratio)
@@
 
-ASCAT and Xiahong's estimates show a `r round(ct$estimate,2)` correlation. That suggests that the remainder is due to the LOH I'm not trying to call.
+# Brief processing notes
+
+ASCAT was run on each SNP sample to obtain SCA calls. ASCAT and Xiahong's estimates show a `r round(ct$estimate,2)` correlation.
+
+Prior to predicting the samples, in the step where I create 5Mb bins across the genome I merge the samples per endoscopy due to differences in the sampling.
+The sWGS had multiple biopsies aggregated per-level, while the SNP data was a single 'research' biopsy per level.
 
@@
 ```{r, echo=F}
-sm = sample.info %>% group_by(Status, Endoscopy, Pathology) %>% dplyr::tally() %>% group_by(Status, Endoscopy) %>% dplyr::mutate(freq=n/sum(n) )
+sm = sample.info %>% group_by(Status, Endoscopy, Pathology) %>% dplyr::tally() %>% 
+  group_by(Status, Endoscopy) %>% dplyr::mutate(freq=n/sum(n) ) 
 sm$Pathology = factor(sm$Pathology, levels=c('NDBE','ID','LGD','HGD','IMC'), ordered = T)
@@
 
-There's a significant difference in the numbers of NDBE or HGD samples between the groups
+There's a significant difference in the numbers of NDBE or HGD samples between the groups at T1.
 ```{r, echo=F}
-pander(chisq.test(table(patient.info$Pathology, patient.info$Status)['HGD',]), caption='HGD')
-pander(chisq.test(table(patient.info$Pathology, patient.info$Status)['NDBE',]), caption='NDBE')
+#pander(chisq.test(table(patient.info$Pathology, patient.info$Status)['HGD',]), caption='HGD')
+#pander(chisq.test(table(patient.info$Pathology, patient.info$Status)['NDBE',]), caption='NDBE')
 
 # Most progressors are HGD at timepoint 1
-pander( round(with(subset(sample.info, Endoscopy == 1), table(Pathology, Status))['HGD',]/table(subset(sample.info, Endoscopy == 1)$Status), 3)*100, caption='% HGD at T1')
+sample.info %>% filter(Timepoint == 'T1') %>% group_by(Status, Pathology) %>% tally %>% spread(Pathology,n) %>% kable(caption='Number of samples by pathology at T1') %>% kable_styling(full_width = F)
+
 ```
@@
 
+`r qcdata %>% group_by(SampleType) %>% tally %>% spread(SampleType,n) %>% kable(caption='Sample numbers') %>% kable_styling(full_width=F)`
+
 ```{r, echo=F, warning=F, message=F}
-qcdata = left_join(qcdata, sample.info[,c('Samplename','Status','Timepoint','SCA.Ratio')], by='Samplename')
+
+qcdata = left_join(qcdata, dplyr::select(sample.info,c('Samplename','Status','Timepoint','SCA.Ratio')), by='Samplename')
 
@@
 
-low = subset(qcdata, SampleType == 'BE' & ASCAT.SCA.ratio <= median(subset(qcdata, SampleType != 'BE')$ASCAT.SCA.ratio))
+low = filter(qcdata, SampleType == 'BE' & ASCAT.SCA.ratio <= median(filter(qcdata, SampleType != 'BE')$ASCAT.SCA.ratio))
 
-anno = cbind.data.frame(table(subset(qcdata, SampleType == 'BE' & ASCAT.SCA.ratio <= median(subset(qcdata, SampleType != 'BE')$ASCAT.SCA.ratio))$Status)/table(subset(qcdata, SampleType == 'BE')$Status))
-colnames(anno) = c('Status','Ratio')
-anno$label='median(normal SCA)'
-anno$x = 0.4
-anno$y = median(subset(qcdata, SampleType != 'BE')$ASCAT.SCA.ratio)
+low %>% group_by(Status, Timepoint) %>% tally %>% spread(Timepoint,n) %>% kable(caption='BE Samples with an ASCAT SCA ratio below the median ratio of the blood/gastric normal samples') %>%
+  kable_styling(full_width = F)
 
-ggplot(subset(qcdata,SampleType == 'BE'), aes(Purity, ASCAT.SCA.ratio, color=Status, group=Status)) + geom_point() + facet_grid(~Status, scales = 'free') + ylim(0,0.01) + 
-  geom_hline(yintercept = anno$y[1]) + geom_text(data=anno, aes(x, y, label=label), nudge_y = 0.0001) +
-  geom_text(data=anno, aes(x, y, label=paste(round(Ratio,2)*100,'% samples',sep='')), nudge_y = -0.0001)
+qcdata = qcdata %>% mutate(SampleType = case_when(grepl('Normal',SampleType) ~ 'Blood/Gastric Normal', TRUE ~ 'BE'),)
 
-ggplot(qcdata, aes(Ploidy, ASCAT.SCA.ratio, color=Status)) + geom_point() + facet_grid(~SampleType)
+anno = qcdata %>% group_by(SampleType) %>% dplyr::summarise(Ratio = median(ASCAT.SCA.ratio) ) %>% rowwise %>% 
+  mutate(label = paste0('median(',SampleType,')','=',round(Ratio,3)))
 
-ggplot(qcdata, aes(Ploidy, ASCAT.SCA.ratio, color=Status)) + geom_point() + facet_grid(~Status)
+ggplot(qcdata, aes(Status, ASCAT.SCA.ratio)) + ylim(0,0.05) + 
+  facet_grid(~SampleType, scales='free_x') +
+  geom_jitter(width=0.1, aes(color=Status)) + 
+  geom_hline(data=dplyr::rename(anno, Status=SampleType), aes(yintercept = Ratio, linetype=Status), color='grey42') +
+  geom_label(data=dplyr::rename(anno, Status=SampleType), aes(x='NP', y=Ratio, label=label), alpha=0.5)
 
-ggplot(subset(qcdata,SampleType == 'BE'), aes(Timepoint, ASCAT.SCA.ratio, color=PatientID, group=Timepoint)) + geom_boxplot() + 
-  facet_wrap(~Status, scales='free_x', ncol=2) + labs(title='SCA Ratio by Timepoint') + theme(legend.position = 'none')
+ggplot(filter(qcdata, !is.na(Ploidy)), aes(factor(round(Ploidy)), ASCAT.SCA.ratio, color=Status)) + 
+  facet_grid(~SampleType, scales = 'free_x', space='free') + geom_boxplot() + labs(x='Ploidy')
 
-ggplot(qcdata, aes(round(Ploidy,2))) + geom_histogram() + 
-  facet_wrap(~SampleType, ncol=1) + labs(x='Ploidy', title='Ploidy, normal vs BE')
+ggplot(filter(qcdata,SampleType == 'BE'), aes(Timepoint, ASCAT.SCA.ratio, fill=Timepoint, group=Timepoint)) + geom_jitter(width=0.2) +
+  geom_boxplot(alpha=0.5) + facet_wrap(~Status, scales='free_x', ncol=2) + labs(title='SCA Ratio by Timepoint') + theme(legend.position = 'none')
 
@@
 
-#x=median(qcdata$SCA.Ratio,na.rm=T)+sd(qcdata$SCA.Ratio,na.rm=T)
-x=0.05
-y=median(qcdata$Purity,na.rm=T)-sd(qcdata$Purity,na.rm=T)/2
+x = 0.01
+y = round(mean(filter(qcdata, SampleType != 'BE')$Purity) - sd(filter(qcdata, SampleType != 'BE')$Purity),2)
 
-ggplot(qcdata, aes(Purity, ASCAT.SCA.ratio, color=ASCAT.SCA.ratio < 0.2)) + geom_point() + 
-         facet_wrap(~Status, ncol=1) + labs(title='Purity~SCA.Ratio') + geom_vline(xintercept=y, color='grey')
-
-ggplot(qcdata, aes(ASCAT.SCA.ratio, Purity, color=Status)) + geom_point() + 
-  facet_wrap(~Status) + labs(x='SCA', title='SCA vs Purity') + 
-  geom_vline(xintercept = x, color='grey') + geom_hline(yintercept = y, color='grey')
-
-table(unique(subset(qcdata, Purity > y & ASCAT.SCA.ratio < x & SampleType == 'BE', select=c('Status','PatientID')))$Status)
-
-ggplot(qcdata, aes(Purity)) + geom_histogram() + geom_vline(xintercept = y, col='red') + 
-  facet_wrap(~SampleType, scales = 'free') + labs(title='Purity by sampletype')
+p = ggplot(qcdata, aes(Purity, ASCAT.SCA.ratio)) + facet_wrap(~SampleType, nrow=1) + 
+  geom_point(aes(color=ASCAT.SCA.ratio < x | Purity > y), show.legend = F) + 
+  geom_vline(xintercept=y, color='grey') + 
+  geom_hline(yintercept=x, color='grey') +
+  labs(title='Purity~SCA.Ratio') + theme(text = element_text(size=12))
+p
+ggsave('plots/snp/purity_v_sca.png', plot=p, height=4, width=4, units = 'in', dpi=300)
 
@@
 
-ggplot(subset(qcdata, SampleType == 'BE'), aes(round(Ploidy,2), Purity, color=Status)) + geom_point() + facet_grid(~Status) + 
-  labs(title='BE only, purity vs ploidy') + geom_hline(yintercept=y, color='grey')
-
-ggplot(qcdata, aes(ASCAT.SCA.ratio, color=Status, fill=Status)) + facet_wrap(~Status, scales='free_y',nrow=1) +  geom_density(alpha=0.5) + labs(x='SCA Ratio', title='SCA Ratio, NP vs P')
-
-ggplot(qcdata, aes(ASCAT.SCA.ratio, color=Status, fill=Status)) + facet_wrap(Status~SampleType, scales='free_y',nrow=2) +  geom_density(alpha=0.5) + labs(title='SCA Ratio, NP vs P')
-
-summary(subset(qcdata, SampleType == 'BE')$ASCAT.SCA.ratio)
-summary(subset(qcdata, SampleType != 'BE')$ASCAT.SCA.ratio)
-
-ggplot(subset(qcdata, SampleType == 'BE'), aes(Purity, ASCAT.SCA.ratio, color=Status)) + facet_grid(~Status) +
-  geom_point() + geom_vline(xintercept = y, color='darkgreen') + geom_hline(yintercept = x, color='red') + 
-  labs(title='Barretts only')
-
 ggplot(subset(qcdata, SampleType == 'BE'), aes(Ploidy, ASCAT.SCA.ratio, color=Status)) + facet_grid(~Status) + geom_point() + 
-  geom_hline(yintercept = 0.02, color='red') + labs(title='Barretts only')
+  geom_hline(yintercept = x, color='red') + labs(title='Barretts only')
 
 ## PER ENDOSCOPY
-qcdata.endo = qcdata %>% group_by(PatientID, EndoID, Status, Timepoint, SampleType) %>% dplyr::summarise_if(is.numeric, c('mean','max','min','sd')) %>% mutate_if(is.numeric, round, digits=3)
-qcdata %>% dplyr::group_by( Status, SampleType) %>% dplyr::summarise( median(ASCAT.SCA.ratio), sd(ASCAT.SCA.ratio) )
+qcdata.endo = qcdata %>% group_by(PatientID, EndoID, Status, Timepoint, SampleType) %>% 
+  dplyr::summarise_if(is.numeric, c('mean','max','min','sd')) %>% mutate_if(is.numeric, round, digits=3)
 
-#ggplot(subset(qcdata.endo, SampleType == 'BE'), aes(ASCAT.SCA.ratio_max, Purity_max)) + geom_point() + facet_grid(~Status) + geom_hline(yintercept = 0.95)  + geom_vline(xintercept = 0.001)
+#qcdata %>% dplyr::group_by( Status, SampleType) %>% dplyr::summarise( median(ASCAT.SCA.ratio), sd(ASCAT.SCA.ratio) )
+# lowsca = filter(qcdata.endo, SampleType == 'BE' & ASCAT.SCA.ratio_mean < 0.01 & Purity_max > 0.98 )
+# nrow(lowsca)
+# nrow(filter(qcdata.endo, SampleType == 'BE' & ASCAT.SCA.ratio_mean >= 0.01 & Purity_max <= 0.98))
 
-lowsca = subset(qcdata.endo, SampleType == 'BE' & ASCAT.SCA.ratio_max <= 0.01 & Purity_max > 0.98 )
-
-#table(unique(lowsca[,c('PatientID','Status')])$Status)/table(unique(qcdata.endo[,c('PatientID','Status')])$Status)
-
-#table(unique(subset(qcdata.endo, SampleType == 'BE' & ASCAT.SCA.ratio_max <= 0.001, select=c('PatientID','Status')))$Status)
-
-ggplot(subset(qcdata.endo,SampleType == 'BE'), aes(Status, ASCAT.SCA.ratio_max, group=Status)) + geom_boxplot() + geom_hline(yintercept = 0.002, color='red')
-
-qcdata %>% dplyr::group_by( Status, SampleType) %>% dplyr::summarise( median(ASCAT.SCA.ratio))
+tmp = qcdata.endo %>% filter(!is.na(Ploidy_mean)) %>% mutate(Ploidy_mean = factor(round(Ploidy_mean)))
+p = ggplot(tmp, aes(Ploidy_mean, ASCAT.SCA.ratio_mean, group=Ploidy_mean)) + 
+  facet_grid(~SampleType) +
+  geom_jitter(width=0.2, aes(color=Ploidy_mean)) + 
+  geom_boxplot(outlier.colour = NA, alpha=0.5) + 
+  geom_hline(yintercept = 0.01, color='grey39') + labs(title='Merged per endoscopy', x='Mean ploidy') + 
+  theme(text=element_text(size=12), legend.position = 'none')
+ggsave('plots/snp/sca_v_ploidy.png', plot=p, height=4, width=4, units = 'in', dpi=300)
 
@@
 
-CN Adjustment uses Annalise's whole-genome non-progressor samples to generate a set of mean and SD values per copy-number state.
-The median LogR values are adjusted using the mean and SD based on the CN state.
-The adjusted LogR values are then winsorize per sample, and tiled in 5Mb bins.
+I use the raw ASCAT total CN values, adjusted for the sample ploidy, to create the weighted average values in each 5Mb and arm bin of the genome.
+Note again that at this point the tiled values are per *endoscopy* not per sample.
 
 ```{r, warning=F, echo=F, message=F, fig.height=10, fig.width=8}
-
+# Adjust for ploidy -- this was done in the processing script
 if (file.exists('~/Data/Reid_SNP/PerPatient/tmp_seg_pt.Rdata')) {
@@
 
-# Adjust for ploidy -- this was done in the processing
-# for (i in 1:nrow(mergedSegs)) {
-#   sample = rownames(mergedSegs)[i]
-#   x = unlist(strsplit(sample, '_'))
-#   #ploidy = mean(subset(qcdata, SampleType == 'BE' & PatientID == x[1] & EndoID == x[2])$Ploidy)
-#   ploidy = subset(qcdata.endo, PatientID == x[1] & EndoID == x[2] & SampleType == 'BE')$Ploidy_mean
-#   if (grepl('BLD|gastric', sample, ignore.case = T)) ploidy = 2
-#   mergedSegs[i,] = mergedSegs[i,]-(ploidy-1)
-#   mergedArms[i,] = mergedArms[i,]-(ploidy-1)
-# }
-
 segmentVariance = apply(mergedSegs, 1, var)
@@
 nm.rows = grep('BLD|GASTRIC', rownames(mergedSegs))
-#normalSegs = mergedSegs[nm.rows,]
-#normalArms = mergedArms[nm.rows,]
 
-# nrow(normalSegs) == nrow(normalArms)
-# 
-# mergedSegs = mergedSegs[-nm.rows,]
-# mergedArms = mergedArms[-nm.rows,]
-# 
-# nrow(mergedSegs) == nrow(mergedArms)
+lowSeg = mergedSegs[grep('^1001', rownames(mergedSegs), value=T),,drop=F]
+p1 = ggplot(melt(lowSeg), aes(Var2, value)) + ylim(-1,1) + facet_grid(~Var1) + geom_point() + 
+  theme(axis.text.x=element_blank() ) + labs(title='Low SCA samples', subtitle='patient 1001 NP', y='5Mb bins')
 
 lowSeg = mergedSegs[grep('^403', rownames(mergedSegs), value=T),,drop=F]
-p1 = ggplot(melt(lowSeg), aes(Var2, value)) + facet_grid(~Var1, space='free_x', scales='free') + geom_point() +
-  theme(axis.text.x=element_blank() ) + labs(title='Low SCA samples', subtitle='patient 403')
+p2 = ggplot(melt(lowSeg), aes(Var2, value)) +  ylim(-1,1) + facet_grid(~Var1) + geom_point() + 
+  theme(axis.text.x=element_blank() ) + labs(title='Low SCA samples', subtitle='patient 403 P', y='5Mb bins')
+
+grid.arrange(p1,p2,ncol=1)
+
+highSeg = mergedSegs[grep('^1114', rownames(mergedSegs), value=T),,drop=F]
+p3 = ggplot(melt(highSeg), aes(Var2, value)) +  ylim(-1,1) + facet_grid(~Var1) + geom_point() + 
+  theme(axis.text.x=element_blank() ) + labs(title='Mixed SCA samples', subtitle='patient 1114 NP', y='5Mb bins')
 
 highSeg = mergedSegs[grep('^512', rownames(mergedSegs), value=T),,drop=F]
-p2 = ggplot(melt(highSeg), aes(Var2, value)) + facet_grid(~Var1, space='free_x', scales='free') + geom_point() +
-  theme(axis.text.x=element_blank() ) + labs(title='Mixed SCA samples', subtitle='patient 512')
+p4 = ggplot(melt(highSeg), aes(Var2, value)) +  ylim(-1,1) + facet_grid(~Var1) + geom_point() + 
+  theme(axis.text.x=element_blank() ) + labs(title='Mixed SCA samples', subtitle='patient 512 P', y='5Mb bins')
 
-grid.arrange(p1,p2,ncol=1)
-```
-
-# Values in the binned segments
-
-Blood/gastric normals have very little variance compared to all of the BE samples (no idea which is P vs NP)
-
-```{r normals, message=F, fig.height=10, fig.width=8}
-#nm = melt(normalSegs)
-nm = melt(mergedSegs[nm.rows,])
-be = melt(mergedSegs[-nm.rows,])
-
-lims = range(nm$value, be$value, na.rm=T)
-
-grid.arrange(
-  ggplot(nm, aes(Var2, value, group=Var2)) + geom_point() + labs(title='Blood/gastric',subtitle=paste('variance=',var(nm$value),sep=''),x='') + 
-    theme(axis.text.x = element_blank()) + ylim(lims),
-  ggplot(be, aes(Var2, value, group=Var2)) + geom_point() + labs(title='BE samples',subtitle=paste('variance=',var(be$value,na.rm=T),sep=''),x='') + 
-    theme(axis.text.x = element_blank()) + ylim(lims)
-)
+grid.arrange(p1,p2,p3,p4,ncol=1)
 
@@
 
-   ggplot(melt(t(adjust.cols(mergedSegs, z.mean, z.sd)[randNormal,])), aes(x=Var2, y=value)) + geom_point() + 
-    theme(axis.text.x=element_blank()) + labs(title='Col adjusted CN, binned'),
  top='Random BLD or gastric normal', bottom=randNormal)
@@
 
-   ggplot(melt(t(adjust.cols(mergedSegs, z.mean, z.sd)[randBE,])), aes(x=Var2, y=value)) + geom_point() + 
-    theme(axis.text.x=element_blank()) + labs(title='Col adjusted CN, binned'),
 top='BE', bottom=randBE)
@@
 
-Adjust using the RAW values...
-
-```{r raw.adjust, fig.height=16, fig.width=4, echo=T}
-q.norm.by<-function(df, target) {
-  dfq = do.call(cbind.data.frame,lapply(1:ncol(df), function(i) {
-    as.data.frame(normalize.quantiles.use.target(df[,i,drop=F], as.vector(target[,i])))
-  }))
-
-  colnames(dfq) = colnames(df)
-  rownames(dfq) = rownames(df)
-
-  return(as.matrix(dfq))
-}
-
-#copySegs = q.norm.by(mergedSegs, raw.segs)
-#copyArms = q.norm.by(mergedArms, raw.arms)
-
+The ploidy adjusted values now need to be mean normalized to match the means of the training data.
+The training data mean is equivalent to the diploid state and the same is true of the adjusted CN data.
+So the mean normalization is simply adding the difference of the means to the SNP bins.
+```{r}
 training = do.call(bind_rows, purrr::map(list.files('~/Data/BarrettsProgressionRisk/Analysis/pcf_perPatient/50kb', '5e06_tiled_segvals.txt', full.names = T, recursive = T), function(f) {
@@
 training.arms = as.matrix(training[,-1]); rownames(training.arms) = training$X1
+```
 
+```{r raw.adjust, fig.height=16, fig.width=4, echo=T}
 median(training.segs)
 median(mergedSegs)
-median(mergedSegs + 0.98)
 
@@
 median(mergedArms)
-median(mergedArms + 0.98)
 
+# difference between the two is 0.98 so...
 mergedSegs = mergedSegs + 0.98
@@
 
+# Also, for the z-adjust values get a mean+sd from both the training and SNP cohort jointly, due to the difference in the variance
 merged = rbind(mergedSegs, training.segs)
@@
 
-# grid.arrange(
-#   ggplot( melt(raw.segs), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, sWGS bins (no arms)'),
-#   ggplot( melt(mergedSegs), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, SNP bins (no arms)')
-#   ggplot( melt(copySegs), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, SNP bins q-normed')
-# )
-
-normal.ids = sample.info %>% filter(grepl('BLD|gastric', Level, ignore.case=T)) %>% dplyr::summarise(rows = paste0(PatientID, '_', EndoID)) %>% pull
-
-nm.rows = which(rownames(mergedSegs) %in% normal.ids)
-
+#normal.ids = sample.info %>% filter(grepl('BLD|gastric', Level, ignore.case=T)) %>% dplyr::summarise(rows = paste0(PatientID, '_', EndoID)) %>% pull
+#nm.rows = which(rownames(mergedSegs) %in% normal.ids)
 ```
@@
 ```{r echo=F,message=F, fig.width=10, fig.height=8}
-#cxN = score.cx(ns, 1)
-#arrayDfNorm = subtract.arms(ns, na)
-#arrayDfNorm = cbind(ns, na)
-#arrayDfNorm = cbind(arrayDfNorm, 'cx'=unit.var(cxN, mn.cx, sd.cx))
+cx = BarrettsProgressionRisk:::scoreCX(mergedSegs, 1)
 
-#probN = predict(fitV, newx=arrayDfNorm, s=l, type='response')
-#length(which(probN > 0.5))/nrow(probN) # FPR
-#length(which(probN <= 0.5))/nrow(probN)
+arrayDf = BarrettsProgressionRisk:::subtractArms(mergedSegs, mergedArms)
+arrayDf = cbind(arrayDf, 'cx'=BarrettsProgressionRisk:::unit.var(cx, mn.cx, sd.cx))
 
-#preds.normal = cbind.data.frame(rownames(probN),probN, predict(fitV, newx=arrayDfNorm, s=l, type='link'))
-#colnames(preds.normal) = c('Samplename','Prob', 'RR')
+melted = melt(arrayDf) %>% left_join(patient.info %>% dplyr::select(UniqueSampleID, Status), by=c('Var1'='UniqueSampleID'))
 
-cx = BarrettsProgressionRisk:::scoreCX(mergedSegs, 1)
-#arrayDf = cbind(ms,ma)
-arrayDf = BarrettsProgressionRisk:::subtractArms(mergedSegs, mergedArms)
-#arrayDf = cbind(arrayDf, 'cx'=unit.var(cx, mn.cx, sd.cx))
-arrayDf = cbind(arrayDf, 'cx'=BarrettsProgressionRisk:::unit.var(cx, mn.cx, sd.cx))
+train.melt = melt(dysplasia.df) %>% 
+  left_join(as_tibble(train.labels) %>% mutate(label = names(train.labels)), by=c('Var1'='label')) %>%
+  mutate(value.y = as.factor(value.y))
 
@@
 grid.arrange(
-  ggplot(melt(arrayDf), aes(Var2, value)) + geom_point() + ylim(lims) +
+  ggplot(melted, aes(Var2, value,color=Status)) + geom_point() + ylim(lims) +
     geom_point(data=melt(arrayDf[,589:634]), col='red') + labs(title='SNP'),
-  ggplot(melt(dysplasia.df), aes(Var2, value)) + geom_point() + ylim(lims) +
-      geom_point(data=melt(dysplasia.df[,589:634]), col='red') + labs(title='sWGS'),
+  ggplot() + geom_point(data=train.melt, aes(Var2, value.x, color=value.y)) + ylim(lims) +
+      geom_point(data=melt(dysplasia.df[,589:634]), aes(Var2, value), col='red') + labs(title='sWGS'),
 nrow=1)
@@
 
+The SNP and sWGS data is comparable, but it's also clear that the variance is fairly different.  Given our experience with other data this is expected.
+
 # Predictions
 
+For simplicity here I'm not bothering with the 'moderate' category, so the cutoff for high ~ low is a probability of  low<0.5>=high
+
+It's worth noting here that the relative risks are much higher, again I'd attribute this to the variance.
+
 ```{r, warning=F, fig.height=8}
-
 prob = predict(fitV, newx=arrayDf, s=l, type='response')
@@
   mutate(Risk = case_when(
-    Prob <= 0.3 ~ 'Low',
-    Prob > 0.3 & Prob < 0.5 ~ 'Moderate',
+    Prob < 0.5 ~ 'Low',
     Prob >= 0.5 ~ 'High' 
@@
 normals = filter(preds, grepl('BLD|gastric', Samplename, ignore.case=T))
-table(normals$Risk)
 
@@
 
-preds = preds %>% left_join(patient.info %>% dplyr::select(UniqueSampleID, Pathology, Status), by=c('Samplename'='UniqueSampleID')) %>%
-   mutate(Status = as.character(ifelse(Pathology %in% c('HGD','IMD'), 'P', Status)))
+preds = preds %>% left_join(patient.info %>% dplyr::select(PatientID, UniqueSampleID, Pathology, Status, Path.Status), by=c('Samplename'='UniqueSampleID'))
+
+# NP patients who are likely P
+np.wgd.pid = qcdata.endo %>% filter(Ploidy_mean > 2.7 & Status == 'NP') %>% ungroup %>% dplyr::select(PatientID) %>% pull
+preds = preds %>% dplyr::mutate(Path.Status = as.character(ifelse(PatientID %in% np.wgd.pid, 'P', Path.Status)))
 
@@
 
-hist(preds$Prob)
-
-length(which(normals$Prob > 0.5))/nrow(normals) # FPR
-length(which(normals$Prob < 0.5))/nrow(normals)
+preds %>% group_by(Path.Status, Risk) %>% tally %>% spread(Risk,n)
 
 myPal = rev(RColorBrewer::brewer.pal(11, 'RdYlBu'))
-grid.arrange(
-  ggplot(preds, aes(RR)) + geom_histogram(aes(fill=..x..), bins=15, show.legend = F) +
+ggplot(bind_rows(preds,normals), aes(RR)) + geom_histogram(aes(fill=..x..), bins=15, show.legend = F) +
   scale_fill_gradientn(colors = myPal,  name='') + 
-    labs(y='n Samples', x='Relative Risk', title='Unadjusted relative risk'),
-  # ggplot(preds, aes(`Adj. RR`)) + geom_histogram(aes(fill=..x..), bins=15, show.legend = F) +
-  #   scale_fill_gradientn(colors = myPal,  name='') +
-  #   labs(y='n Samples', x='Relative Risk', title='Adjusted relative risk'),
-top='Relative risk')
+  labs(y='n Samples', x='Relative Risk', title='Unadjusted relative risk', subtitle='All SNP samples')
 
-grid.arrange(
-  ggplot(preds, aes(Prob)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
+ggplot(bind_rows(preds,normals), aes(Prob)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
   scale_fill_gradientn(colors = myPal,  name='') + 
-    labs(title='Predictions, all samples model', y='n Samples', x='Unadjusted Probability'),
-  # ggplot(preds, aes(`Adj. Prob`)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
-  #   scale_fill_gradientn(colors = myPal,  name='') +
-  #   labs(title='Predictions, all samples model', y='n Samples', x='Adjusted Probability'),
-top='Probability')
+  labs(title='Predictions, all samples model', y='n Samples', x='Unadjusted Probability', subtitle='All SNP samples')
+```
 
-#preds = sample.info %>% mutate(Sample = paste(PatientID,EndoID,sep='_')) %>% left_join(preds, by=c('Sample'='Samplename'))
+The best false positive rate can be assessed from the blood/gastric normals.  So in this case ~17%
+`r normals %>% group_by(Risk) %>% tally %>% mutate(ratio = round(n/nrow(normals),2))  %>% kable(caption='Blood & gastric normals') %>% kable_styling(full_width = F)`
 
-summary(preds$Prob)
-
+We see a higher FPR rate in the full dataset. 
+```{r echo=F}
 preds %>% group_by(Status, Risk) %>% tally %>% spread(Risk, n) %>%
   left_join(preds %>% group_by(Status) %>% tally, by='Status') %>%
-  mutate_if(is.numeric, list( round(./n,2))) %>% 
-  dplyr::select(-n)
+  mutate_if(is.numeric, list( ~round(./n,2))) %>% 
+  dplyr::select(-n) %>% kable(caption='Ratio of samples per risk category (excluding normals)') %>% 
+  kable_styling(full_width = F)
+```
 
-roc = pROC::roc(Status ~ Prob, data=preds, auc=T, ci=T, of='thresholds',transpose=T)
-roc.plot(roc) + labs(title='SNP AUC by (P)')
-pROC::coords(roc, 'best')
+Despite that, we still see a fairly consistent high/low risk prediction for NP vs P.  However, HGD in NP patients predicts at a rate consistent with progression.
+```{r echo=F}
+preds = preds %>% mutate(Pathology = factor(Pathology, levels=c('NDBE','ID','LGD','HGD','IMC'), ordered = T))
+rp = plot.risk.by.path(preds) 
+rp
 
-preds = preds %>%
-  mutate(LH = ifelse(Prob > 0.5, 'High', 'Low')) %>%
-  mutate(LH = factor(LH, levels=c('Low','High'), ordered=T))
+ggsave('plots/snp/path_risk.png', plot=rp, width=7, height=6, units='in', dpi=300)
+```
 
-rocR = pROC::roc(Status ~ LH, data=preds, auc=T, ci=T, of='thresholds',transpose=T)
-roc.plot(rocR) + labs(title='SNP AUC by discretized risks (low/high)')
-pROC::coords(rocR, 'best')
+### ROC 
+```{r}
+roc = pROC::roc(Status ~ Prob, data=bind_rows(preds, normals %>% mutate(Status = 'NP')), auc=T, ci=T, of='thresholds',transpose=T)
+roc.plot(roc) + labs(title='SNP ROC', subtitle='All samples')
+pROC::ci.auc(roc$auc)
+pROC::coords(roc, 'best', transpose=F)
+ggsave(filename = 'plots/snp/all_samp_roc.png', plot = roc.plot(roc) + labs(title='SNP ROC', subtitle='All samples') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
 
+rocBE = pROC::roc(Status ~ Prob, data=preds, auc=T, ci=T, of='thresholds',transpose=T)
+roc.plot(rocBE) + labs(title='SNP ROC', subtitle='BE samples only')
+pROC::ci.auc(rocBE$auc)
+pROC::coords(rocBE, 'best', transpose=F)
+ggsave(filename = 'plots/snp/be_roc.png', plot = roc.plot(rocBE) + labs(title='SNP ROC', subtitle='BE samples only') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
 ```
@@
 A quick look at the two random samples from earlier.  
-`r pander(subset(normals, Samplename == randNormal), caption='Random normal')`
+`r filter(normals, Samplename == randNormal) %>% kable(caption='Random normal') %>% kable_styling(full_width=F)`
 
 The random BE should show a very high probability, and it does...
-`r pander(subset(preds, Samplename == randBE), caption='Random BE')`
-
-In the gastric/blood normals `r signif(length(which(preds.normal$Prob < 0.5))/nrow(preds.normal),2)` are predicted to have <50% probability (TPR).
-
-For blood normals only `r signif(nrow(subset(preds.normal, grepl('BLD',Samplename) & Prob < 0.5))/nrow(subset(preds.normal, grepl('BLD',Samplename) )),2)` have a <50% probabilitye (TPR).
+`r filter(preds, Samplename == randBE) %>% dplyr::select(Samplename, Prob, RR, Risk) %>% kable(caption='Random BE') %>% kable_styling(full_width=F)`
 
@@
 ```{r, warning=F, fig.height=8}
-#qcdata.be = subset(qcdata, SampleType == 'BE')
-# get.qcinfo<-function(vec) {
-#   endo = vec[2]
-#   pt = vec[1]
-#   y = subset(qcdata.be, PatientID == pt & EndoID == endo)
-#   y = y %>% group_by(Status, Timepoint) %>%  dplyr::summarise( mean(Ploidy), mean(Purity), mean(Goodnessoffit), mean(SCA.Ratio) )
-#   cbind.data.frame('PatientID'=pt, 'EndoID'=endo, y, stringsAsFactors = F)
-# }
 preds = preds %>% dplyr::group_by(Samplename) %>% dplyr::mutate(
@@
 
-#pred.qc = left_join(preds, qcdata.endo %>% dplyr::select(-SampleType), by=c('PatientID', 'EndoID', 'Status')) %>% data.frame
+pred.qc = left_join(preds, qcdata.endo %>% dplyr::select(-SampleType), by=c('PatientID', 'EndoID', 'Status'))
 
-#grid.arrange(
-#  ggplot(pred.qc, aes(ASCAT.SCA.ratio_mean, Prob, color=Status)) + geom_point(),
- # ggplot(pred.qc, aes(Ploidy_max, Prob, color=Status)) + geom_point(),
-#ncol=2)
+grid.arrange(
+  ggplot(pred.qc, aes(ASCAT.SCA.ratio_mean, Prob, color=Status)) + geom_point(),
+  ggplot(pred.qc, aes(Ploidy_max, Prob, color=Status)) + geom_point(),
+ncol=2)
 
-#pander(cor.test(pred.qc$ASCAT.SCA.ratio_mean, pred.qc$Prob))
+ct = cor.test(pred.qc$ASCAT.SCA.ratio_mean, pred.qc$Prob)
 ```
 
-```{r, warning=F, fig.height=8, eval=F}
-prog = subset(pred.qc, Status == 'P')
+The ASCAT mean SCA ratio across the genome correlates with the predicted probabilites in the BE samples, but not incredibly highly: R=`r round(ct$estimate,2)`,
+
+```{r, warning=F, fig.height=8}
+prog = filter(pred.qc, Status == 'P')
 p1 = ggplot(prog, aes(Prob)) + facet_wrap(~Timepoint, scales='free_y',nrow=1) + 
@@
   scale_fill_gradientn(colors = myPal,  name='') + 
-  labs(title='Predictions on progressors', y='n Samples', x='Unadjusted Probability')
+  labs(title='Predictions on progressors', y='n Samples', x='Unadjusted Probability') + theme(text=element_text(size=12))
 
@@
   scale_fill_gradientn(colors = myPal,  name='') + 
-  labs(title='Predictions on non-progressors', y='n Samples', x='Unadjusted Probability')
+  labs(title='Predictions on non-progressors', y='n Samples', x='Unadjusted Probability') + theme(text=element_text(size=12))
 
-preds.normal = preds.normal %>% dplyr::mutate(Type = ifelse(grepl('BLD',Samplename), 'Blood','Gastric'))
+normals = normals %>% dplyr::mutate(Type = ifelse(grepl('BLD',Samplename), 'Blood','Gastric'))
 
-p3 = ggplot(preds.normal, aes(Prob)) + facet_wrap(~Type, ncol=2, scales = 'free') +
+p3 = ggplot(normals, aes(Prob)) + facet_wrap(~Type, ncol=2, scales = 'free') +
   geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
   scale_fill_gradientn(colors = myPal,  name='') +
-  labs(title='Predictions on BLD/Gastric', y='n Samples', x='Unadjusted Probability')
+  labs(title='Predictions on Blood/Gastric normals', y='n Samples', x='Unadjusted Probability') + theme(text=element_text(size=12))
 
@@
 
+ggsave('plots/snp/probability_dists.png', plot = grid.arrange(p1,p2,p3,ncol=1), width=6, height=10, units='in', dpi=300)
 ```
@@
 # Compare predictions to known status
+```{r roc, echo=T, eval=T}
+pred.qc = pred.qc %>% mutate(Status = factor(Status))
 
-Using my thresholds
-
-```{r roc, echo=T, eval=F}
-pred.qc = base::merge( pred.qc, patient.info[,c('PatientID', 'Timepoint Code','Pathology')], by.x=c('PatientID', 'EndoID'), by.y=c('PatientID','Timepoint Code') )
-
-# Three groups, prog, non prog (max LGD), non prog (max HGD)
-# pred.qc = pred.qc %>% group_by(Status, PatientID) %>% dplyr::mutate( 
-#   exclude=(Status == 'NP' & (length(which(Pathology == 'HGD')) > 0 | `mean(Ploidy)` > 2.7 )) )
-pred.qc$Status = factor(pred.qc$Status)
-
-pred.qc = pred.qc %>% group_by(Status, PatientID) %>% dplyr::mutate(
-  exclude=(Status == 'NP' & (length(which(Pathology == 'HGD')) > 0)),
-  wgd = (Status == 'NP' & Ploidy_max > 2.7 )
+pred.qc = pred.qc %>% group_by(Status, Samplename) %>% 
+  dplyr::mutate(
+    hgd.NP = (Status == 'NP' & (length(which(Pathology == 'HGD')) > 0)),
+    wgd.NP = (Status == 'NP' & Ploidy_max > 2.7 ),
+    lowsca = (ASCAT.SCA.ratio_mean < 0.01 & Purity_mean > 0.95)
 )
 
-# pred.qc = pred.qc %>% group_by(Status, Samplename) %>% dplyr::mutate(
-#   lowsca = ( Status == 'P' & (`mean(SCA.Ratio)` < 0.03 & `mean(Purity)` >= 0.95))
-# )
-
-pred.qc = pred.qc %>% group_by(Status, Samplename) %>% dplyr::mutate(
-  lowsca = ( ASCAT.SCA.ratio_mean < 0.02 & Purity_mean > 0.95)
-)
+ascat.normals = qcdata %>% filter(grepl('Normal',SampleType)) %>% data.frame
+ggplot(ascat.normals, aes(ASCAT.SCA.ratio, Purity, color=Status)) + geom_point()
 
 ggplot(pred.qc, aes(ASCAT.SCA.ratio_mean, Purity_mean, color=lowsca)) + geom_point() + facet_grid(~Status) + 
-  geom_vline(xintercept=0.02) + geom_hline(yintercept = 0.95)
+  geom_vline(xintercept=0.01) + geom_hline(yintercept = 0.95)
 
-pander(
-  rbind(rbind('n.Samples'=table(subset(pred.qc, !lowsca )$Status)),
-  rbind('n.Patients'=table(unique(subset(pred.qc, !lowsca, select=c('PatientID','Status')))$Status))),
-caption='Number of samples and patients that pass the low SCA:high purity thresholds')
+pred.qc %>% group_by(Status, lowsca) %>% tally(name='n Samples') %>% spread(Status, `n Samples`) %>% kable(caption = 'LOWSCA high purity samples') %>% kable_styling(full_width = F)
 
-pander(
-  rbind(rbind('n.Samples'=table(subset(pred.qc, !lowsca & !exclude & !wgd)$Status)),
-  rbind('n.Patients'=table(unique(subset(pred.qc, !lowsca  & !exclude & !wgd, select=c('PatientID','Status')))$Status))),
-caption='Number of samples and patients that pass the low SCA:high purity & NP thresholds')
-
+filter(pred.qc, !lowsca & !wgd.NP) %>% group_by(Status) %>% tally(name='n Samples') %>% kable(caption='Number of samples that pass the low SCA:high purity & NP (wgd) thresholds') %>% kable_styling(full_width = F)
 ```
 
-```{r echo=F, eval=F}
-pred.qc.ex = subset(pred.qc, SampleType == 'BE' & !exclude & !wgd & !lowsca)
-cuts = seq(0,1,0.1)
+### ROC after exclusions
 
-pred.qc.ex$quants = with(pred.qc.ex, cut(Prob, breaks=cuts))
-qt = as.data.frame.matrix(table(pred.qc.ex$quants, pred.qc.ex$Status))
-qt$quant = rownames(qt)
-qt$P = round(qt$P/sum(qt$P),2)
-qt$NP = round(qt$NP/sum(qt$NP), 2)
-pander(qt, caption='Excluding HGD NP, WGD NP, low SCA P')
+In this case the status has changed slightly.
+All patients that had a HGD/IMC path report are considered as progressors to match the sWGS cohort.
+After that I include exclusions for low SCA and WGD in NP.
 
-# These are the cutoffs from the sWGS data
-get.risk<-function(prob) {
-  if (prob < 0.3) return('Low')
-  if (prob >= 0.3 & prob < 0.7) return('Moderate')
-  if (prob >= 0.7) return('High')
-}
+```{r echo=F}
+# Set Path.Status = P for all patient where a sample was WGD
+#wgd.ids = filter(pred.qc, wgd.NP) %>% ungroup %>% dplyr::select(PatientID) %>% distinct %>% pull 
+#pred.qc = pred.qc %>% mutate(wgd.NP = PatientID %in% wgd.ids) %>% mutate(Path.Status = ifelse(wgd.NP, 'P', Path.Status) )
 
-# All data
-pred.qc = pred.qc %>% dplyr::rowwise() %>% dplyr::mutate(
-  Risk = get.risk(Prob)
-)
-pred.qc$Risk = factor(pred.qc$Risk, levels=c('Low','Moderate','High'), ordered = T)
+rocEX = pROC::roc( Path.Status ~ Prob, data=pred.qc, auc=T, ci=T, of='thresholds', transpose=T)
+roc.plot(rocEX) + labs(title='BE samples', subtitle='Path status (HGD/IMC)')
+pROC::ci.auc(rocEX)
+ggsave(filename='plots/snp/path_status_roc.png', plot = roc.plot(rocEX) + labs(title='BE samples', subtitle='Path status (HGD/IMC)') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
 
-norm.pred = subset(pred.qc, SampleType != 'BE')
-
-pred.qc = subset(pred.qc, SampleType == 'BE')
-
-roc = pROC::roc(Status ~ Prob, data=pred.qc, auc=T, ci=T, of='thresholds')
-roc.plot(roc) + labs(title='All, Prob')
-pander(pROC::coords(roc, 'best'), caption='All samples, ROC best')
-
-# Exclude HGD NP patients
-roc = pROC::roc(Status ~ Prob, data=subset(pred.qc, !exclude), auc=T, ci=T, of='thresholds')
-roc.plot(roc) + labs(title='Excluded HGD NP patients')
-
-# Exclude the WGD NP
-rocpp = pROC::roc(Status ~ Prob, data=subset(pred.qc, !wgd), auc=T, ci=T, of='thresholds')
-roc.plot(rocpp) + labs(title='Both endos - excluded WGD NP')
-pROC::coords(rocpp, 'best')
-
-roc = pROC::roc(Status ~ Prob, data=subset(pred.qc, !exclude & !wgd), auc=T, ci=T, of='thresholds')
-roc.plot(roc) + labs(title='Excluded HGD & WGD NP patients')
+notlow = filter(pred.qc, !lowsca)
+pid = notlow %>% ungroup %>%dplyr::select(PatientID) %>% distinct %>% pull 
+length(pid)
+qcdata %>% filter(PatientID %in% pid & SampleType != 'Blood/Gastric Normal')
 
 # Exclude the low SCA patients
-rocpp = pROC::roc(Status ~ Prob, data=subset(pred.qc, !lowsca), auc=T, ci=T, of='thresholds')
-roc.plot(rocpp) + labs(title='Both endos - excluded low SCA')
-pROC::coords(rocpp, 'best')
+rocpp = pROC::roc(Status ~ Prob, data=filter(pred.qc, !lowsca), auc=T, ci=T, of='thresholds', transpose=T)
+roc.plot(rocpp) + labs(title='BE samples', subtitle='excluding low SCA samples')
+ggsave(filename='plots/snp/auc_ex_lowsca_roc.png', plot = roc.plot(rocpp) + labs(title='BE samples', subtitle='excluding low SCA samples') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
 
 # Exclude all of the above
-rocpp = pROC::roc(Status ~ Prob, data=subset(pred.qc, !exclude & !lowsca & !wgd), auc=T, ci=T, of='thresholds')
-roc.plot(rocpp) + labs(title='Both endos - excluded low SCA P & HGD NP & WGD NP')
-pROC::coords(rocpp, 'best')
+rocexall = pROC::roc(Status ~ Prob, data=filter(pred.qc, !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds', transpose=T)
+roc.plot(rocexall) + labs(title='BE samples', subtitle='excluding low SCA & WGD NP')
+ggsave(filename='plots/snp/auc_ex_lowsca_wgs_roc.png', plot =roc.plot(rocexall) + labs(title='BE samples', subtitle='excluding low SCA & WGD NP') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
+
+# Timepoint 1 only
+roct1 = pROC::roc(Status ~ Prob, data=filter(pred.qc, Timepoint == 'T1' & !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds', transpose=T)
+roc.plot(roct1) + labs(title='BE samples', subtitle='Timepoint T1')
+ggsave(filename='plots/snp/auc_t1_roc.png', plot=roc.plot(roct1) + labs(title='BE samples', subtitle='Timepoint T1') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
 
 # Timepoint 2 only
-roct2 = pROC::roc(Status ~ Prob, data=subset(pred.qc, Timepoint != 'T1' & !exclude & !lowsca & !wgd), auc=T, ci=T, of='thresholds')
-roc.plot(roct2) + labs(title='T2, Prob')
+roct2 = pROC::roc(Status ~ Prob, data=subset(pred.qc, Timepoint != 'T1' & !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds', transpose=T)
+roc.plot(roct2) + labs(title='BE samples', subtitle='Timepoint T2')
+ggsave(filename='plots/snp/auc_t2_roc.png', plot=roc.plot(roct2) + labs(title='BE samples', subtitle='Timepoint T2') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
 
-ggplot(subset(pred.qc, !exclude & !wgd & !lowsca), aes(ASCAT.SCA.ratio_mean,Prob, color=Risk, shape=Status)) + 
-  facet_grid(~Status) + geom_point()
+ggplot(filter(pred.qc, !wgd.NP & !lowsca), aes(Prob)) +
+  geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
+  scale_fill_gradientn(colors = myPal,  name='') +
+  labs(title='Excluding WGD NP & low SCA', y='n Samples', x='Unadjusted Probability') + theme(text=element_text(size=12))
 
-ex.pred = subset(pred.qc, !exclude & !wgd & !lowsca)
-
-write.table(subset(pred.qc, lowsca, select=c('PatientID','EndoID','Status','ASCAT.SCA.ratio_mean','SCA.Ratio_mean','Purity_mean')),
-            sep='\t', quote=F, row.names=F, file='lowsca_pts.txt')
-
-head(qcdata)
-
-x = subset(qcdata, ASCAT.SCA.ratio < 0.02 & Purity > 0.95)
-table(x$SampleType)/table(qcdata$SampleType)
-
-write.table(subset(x, SampleType == 'BE', select=c('PatientID','EndoID','Samplename','ASCAT.SCA.ratio','SCA.Ratio','Purity')), quote=F, sep='\t', row.names = F, file='lowsca_samples.txt')
-
-#x = subset(qcdata, ASCAT.SCA.ratio < 0.02 & Purity > 0.95 & SampleType == 'BE')
-
-# Quick check
-# nrow(subset(ex.pred, Prob > 0.5 & Status == 'P'))/table(ex.pred$Status)['P']
-# nrow(subset(ex.pred, Prob < 0.5 & Status == 'P'))/table(ex.pred$Status)['P'] #False negatives
-# 
-# nrow(subset(ex.pred, Prob > 0.5 & Status == 'NP'))/table(ex.pred$Status)['NP'] # False pos
-# nrow(subset(ex.pred, Prob < 0.5 & Status == 'NP'))/table(ex.pred$Status)['NP'] #
-
-ct = cor.test(pred.qc.ex$Prob, pred.qc.ex$ASCAT.SCA.ratio_mean)
-ct2 = cor.test(pred.qc$Prob, pred.qc$ASCAT.SCA.ratio_mean)
+ct = cor.test(pred.qc$Prob, pred.qc$SCA.Ratio_mean)
+ct2 = cor.test(filter(pred.qc, !wgd.NP & !hgd.NP & !lowsca)$Prob, filter(pred.qc, !wgd.NP & !hgd.NP & !lowsca)$SCA.Ratio_mean)
 ```
 
-The risk probability is correlated with the % of the genome that's been altered as identified by Xiahong, R=`r ct$estimate` and that's true even when no patients are excluded (R=`r ct2$estimate`).
-
-# Regression using SNP Data
+The risk probability is correlated with the % of the genome that's been altered as identified by Xiahong, at about the same rate as the correlation between ASCAT and the probabilites: R=`r ct$estimate`.
+And that's true even when patients are excluded, though at a slightly higher rate (R=`r ct2$estimate`).
 
 ```{r regression, echo=F, message=F, eval=F}
+# Regression using SNP Data
 folds = 10; splits = 5
 ## ----- All ----- ##
-cache.dir = '~/Data/Ellie/Analysis/SNP/no-low/'
+cache.dir = '~/Data/BarrettsProgressionRisk//Analysis/SNP/no-low/'
 file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
@@
  load(file, verbose=T)
- model.performance(performance.at.1se, coefs)$plot + labs(title='SNP CV models')
+ model.performance(performance.at.1se, coefs, folds, splits)$plot + labs(title='SNP CV models')
 } else {
@@
 
-```{r compare.coefs, echo=F, eval=F}
-# Matching
-snp.features = coefs$`0.5`
-
-intersect(rownames(snp.features), rownames(features))
-
-cxisft = grep('cx',rownames(snp.features))
-
-length(which(rowSums(snp.features[,-1])/50 > 0.5))
-
-snp.features = do.call(rbind.data.frame, c(strsplit(rownames(snp.features)[-cxisft], ':|-'), stringsAsFactors=F))
-colnames(snp.features) = c('chr','start','end')
-
-swgs.features = do.call(rbind.data.frame, c(strsplit(rownames(features), ':|-'), stringsAsFactors=F))
-colnames(swgs.features) = c('chr','start','end')
-
-snp.features = makeGRangesFromDataFrame(snp.features)
-snp.features = snp.features[-which(width(snp.features) > 5e6)] # No arms
-
-swgs.features = makeGRangesFromDataFrame(swgs.features)
-
-dist = distanceToNearest(swgs.features,snp.features)
-
-swgs.features[queryHits(dist)]
-snp.features[subjectHits(dist)]
-
-match = dist[elementMetadata(dist)$distance != 0]
-elementMetadata(dist)$bins = round(elementMetadata(dist)$distance/5e6)
-
-x1 = dist[elementMetadata(dist)$bins <= 2]
-
-swgs.features[queryHits(x1)]
-snp.features[subjectHits(x1)]
-```
-
 ```{r loo, eval=F}
-
 file = paste(cache.dir, 'loo.Rdata', sep='/')
@@
 } else {
-  stop("Run snp-regression.R first")
+  snp.preds = do.call(bind_rows, purrr::map(list.files(cache.dir, 'loo.Rdata', recursive = T, full.names = T), function(f) {
+    load(f)
+    as_tibble(preds, rownames = 'UniqueID')
+  }))
+
+  #stop("Run snp-regression.R first")
 }
 
-snp.roc = pROC::roc(Status~Prediction,slabels, auc=T, ci=T, of='thresholds')
+snp.preds = left_join(snp.preds, dplyr::select(slabels, c(UniqueID, Status)), by='UniqueID') %>% filter(!is.na(Prob))
+
+snp.roc = pROC::roc(Status~Prob,snp.preds, auc=T, ci=T, of='thresholds')
 roc.plot(snp.roc, title="SNP LOO Model ROC")

### 2019-12-09 lib/common_plots.R

@@ -356,10 +356,11 @@ mtn.spatial<-function(df, b, limits, title='', pal=NULL) {
 
 
 plot.risk.by.path<-function(preds) {
-  riskPath = preds %>% dplyr::group_by(Status, Pathology, Risk) %>% tally(name='Freq') %>%
+  riskCols = BarrettsProgressionRisk:::riskColors()
+  riskPath = preds %>% dplyr::group_by(Status, Pathology, Risk) %>% tally(name='Freq') %>% ungroup %>%
     mutate(Status = recode_factor(Status, NP = 'Non-Progressor', P = 'Progressor', .ordered = T))
   
-  pathTotal = preds %>% dplyr::group_by(Status, Pathology) %>% tally(name='nPath') %>%
+  pathTotal = preds %>% dplyr::group_by(Status, Pathology) %>% tally(name='nPath') %>% ungroup %>%
     mutate(Status = recode_factor(Status, NP = 'Non-Progressor', P = 'Progressor', .ordered = T))
   
   riskPath = full_join(riskPath, pathTotal, by=c('Status','Pathology')) %>% 

### 2019-12-09 processing_pipeline/run-cv2.R

@@ -17,7 +17,7 @@ suppressPackageStartupMessages(source('~/workspace/shallowwgs_pipeline/lib/cv-pt
 
 
 data = args[1]
-# data = '~/Data/BarrettsProgressionRisk/Analysis/pcf_perPatient/100kb/'
+# data = '~/Data/BarrettsProgressionRisk/Analysis/pcf_perPatient/50kb/'
 outdir = args[2]
 # outdir = '~/Data/BarrettsProgressionRisk/Analysis/models_5e6/100kb/'
 infodir = args[3]
@@ -37,7 +37,6 @@ logT = F
 cache.dir = outdir
 if (logT) cache.dir = paste(cache.dir, '_logR', sep='')
 
-#if (dir.exists(cache.dir)) stop(paste0("Output directory ",cache.dir," exists. Exiting."))
 dir.create(cache.dir, recursive=T, showWarnings=F)
 
 ## Hospital.Research.ID info file
@@ -54,40 +53,12 @@ if (set != 'All') {
 } else {
   patient.info = all.patient.info
 }
-
 length(unique(patient.info$Hospital.Research.ID))
-
-# fncols <- function(data, cname, default=NA) {
-#   add <-cname[!cname%in%names(data)]
-#   
-#   if(length(add)!=0) data[add] <- default
-#   data
-# }
-# pastefun<-function(x) {
-#   if ( !grepl('SLX-', x) ) x = paste0('SLX-',x)
-#   return(x)
-# }
-# val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
-# sheets = readxl::excel_sheets(val.file)[8:13]
-# all.val = do.call(bind_rows, lapply(sheets, function(s) {
-#   readxl::read_xlsx(val.file, s) %>% dplyr::select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA, matches('Collection')) %>% dplyr::filter(!is.na(`SLX-ID`)) %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% fncols('Collection', 'Biopsy')
-# })) %>% rowwise %>% mutate_at(vars(`SLX-ID`), list(pastefun) ) %>% ungroup %>% mutate(
-#   `Hospital Research ID` = str_replace_all( str_remove_all(`Hospital Research ID`, " "), '/', '_'),
[…]

### 2019-12-09 snp/snp-regression.R

@@ -4,7 +4,6 @@ library(glmnet)
 library(readxl)
 library(pROC)
 library(ggfortify)
-library(preprocessCore)
 
 suppressPackageStartupMessages(source('~/workspace/shallowwgs_pipeline/lib/data_func.R'))
 suppressPackageStartupMessages(source('~/workspace/shallowwgs_pipeline/lib/common_plots.R'))
@@ -15,47 +14,50 @@ adjust.cols<-function(mtx, means=NULL, sds=NULL, na.replace=0) {
     if (ncol(mtx) != length(means))
       stop("Vector of means needs to match the columns in the matrix")
     for (i in 1:ncol(mtx)) 
-      mtx[,i] = unit.var(mtx[,i], means[i], sds[i])
+      mtx[,i] = BarrettsProgressionRisk:::unit.var(mtx[,i], means[i], sds[i])
   } else {
     for (i in 1:ncol(mtx)) 
-      mtx[,i] = unit.var(mtx[,i])
+      mtx[,i] = BarrettsProgressionRisk:::unit.var(mtx[,i])
   }
   mtx[is.na(mtx)] = na.replace
   
   return(mtx)
 }
 
-chr.info = get.chr.lengths(file='/tmp/hg19_info.txt')
+chr.info = BarrettsProgressionRisk:::chrInfo(build = 'hg19')
 
-load('~/Data/Ellie/Analysis/5e6_arms_all/model_data.Rdata', verbose = T)
+load('~/Data/BarrettsProgressionRisk//Analysis/5e6_arms_all/model_data.Rdata', verbose = T)
 rm(labels, dysplasia.df,mn.cx,sd.cx,z.mean,z.sd,z.arms.mean,z.arms.sd)
 
 load('~/Data/Reid_SNP/PerPatient/allpts_ascat.Rdata', verbose=T)
-qcdata = qcdata %>% rowwise %>% dplyr::mutate(
-  PatientID = unlist(strsplit(Samplename, '_'))[1],
-  SampleID = unlist(strsplit(Samplename, '_'))[2],
-  EndoID = unlist(strsplit(Samplename, '_'))[3],
-  Level = unlist(strsplit(Samplename, '_'))[4]
-)
+
+qcdata = qcdata %>% as_tibble %>%  mutate(Samplename = sub('_$','',Samplename)) %>%
+  separate(Samplename, c('PatientID','SampleID','EndoID','Level'), sep='_', remove=F)
[…]

### 2019-12-12 Paper_resubmit.Rmd
bug fixing and generating plots

@@
 riskCols = brewer.pal(11, "RdYlBu")[c(10, 5, 1)]
+
+residuals = do.call(bind_rows, purrr::map(list.files(paste0(dir,'/Analysis/pcf_perPatient/50kb'), 'residuals', recursive = T, full.names = T), 
+                              function(f) { read_tsv(f, col_types = 'ccddddl') }
+))
+
 ```
@@
 We have `r table(labels)[1]` samples from non-progressors and `r table(labels)[2]` samples from progressors.
+
+A few samples failed our QC process and were excluded from the model training process.
+`r residuals %>% filter(!Pass) %>% kable(caption = 'Samples that failed post-QDNAseq QC') %>% kable_styling(full_width = F)`
 
@@
 
-```{r echo=F, warning=F, message=F, eval=T, fig.height=12, fig.width=20}
+```{r echo=F, warning=F, message=F, eval=T, fig.height=6, fig.width=10}
 chr.lengths = BarrettsProgressionRisk:::chrInfo() %>% dplyr::filter(chr %in% c(1:22)) %>%
@@
 ## NDBE
-nb.status <- c('Progressor'=paste0('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology == 'BE')$variable)), ')'),
-              'Non-Progressor'=paste0('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology == 'BE')$variable)),')'))
+nb.status <- c('Progressor'=paste0('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology == 'NDBE')$variable)), ')'),
+              'Non-Progressor'=paste0('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology == 'NDBE')$variable)),')'))
 
@@
 
-gp = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples')
-
-gp = globalCN.plot + theme(legend.position = 'none', axis.text = element_blank(), strip.text.y = element_blank(), strip.text.x = element_text(size=8)) + 
+gp = globalCN.plot +  labs(y='mean adj.
+seg.', title='All samples') + theme(legend.position = 'none', axis.text = element_blank(), strip.text.y = element_blank(), strip.text.x = element_text(size=8)) +
   labs(y='',x='', title='All samples')
@@
 
-patient.file = '~/Data/Ellie/QDNAseq/validation/sWGS_validation_batches.xlsx'
+patient.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
 
@@
   kable(caption='Validation cohort') %>% kable_styling(full_width = F)
+
+val.info %>% group_by(Status) %>% dplyr::summarise(
+  `Total Patients` = length(unique(Patient)),
+  `Total Samples` = length(Samplename),
+`max follow-up` = max(`Weeks Follow Up`/52),
+`min follow-up` = min(`Weeks Follow Up`/52),
+  `mean follow-up` = mean(`Weeks Follow Up`/52),
+  `sd follow-up` = sd(`Weeks Follow Up`/52)
+)
+
+val.info %>% group_by(Status, Pathology) %>% tally
+
 ```
@@
 failedQC = residuals %>% filter(varMAD_median > 0.008)
+
+val.info %>% filter(Samplename %in% failedQC$samplename)
 

### 2019-12-12 processing_pipeline/process_training_data2.R

@@ -14,7 +14,7 @@ source('~/workspace/shallowwgs_pipeline/lib/load_patient_metadata.R')
 data = args[1]
 patient.file = args[2]
 outdir = args[3]
-# data = '~/Data/Ellie/QDNAseq/training/merged/100kb'
+# data = '~/Data/Ellie/QDNAseq/training/merged/50kb'
 # patient.file = '~/Data/Ellie/QDNAseq/training/All_patient_info.xlsx'
 # outdir = '~/Data/Ellie/Analysis/'
 

### 2019-12-24 Compare Models.Rmd

@@
 ```{r, fig.width=8,fig.height=7}
+dir.create('plots/parameters')
 kb.models = list.files(data)
@@
 
-ggplot(kb.perf, aes(kb, mean, fill=kb)) + facet_grid(~alpha) + geom_col() + 
-  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme), col='grey39', width=0.2) + 
-  geom_text(aes(label=n.coefs, y = 0.01)) + scale_fill_brewer(palette='Set1') + labs(x='',y='cv performance') +
-  theme_minimal()
-
-ggplot(kb.perf, aes(alpha, mean, fill=kb)) + facet_grid(~kb) + geom_col() + 
-  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme), col='grey39', width=0.2) + 
-  scale_fill_brewer(palette='Set1') + labs(x='',y='cv performance') +
-  geom_text(aes(label=n.coefs, y = 0.01)) + 
-  theme_minimal()
-
+p = ggplot(kb.perf, aes(kb, mean, fill=kb)) + facet_grid(~alpha) + ylim(0,1) + 
+  geom_col() + geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme), col='grey39', width=0.2) + 
+  #geom_text(aes(label=n.coefs, y = 0.01)) + 
+  scale_fill_brewer(palette='Set1') + 
+  labs(x='',y='classification accuracy', title = 'Cross-Validation Classification Accuracy', subtitle='Alpha values:') +
+  theme_minimal() + theme(text = element_text(size=12), axis.text.x  = element_text(hjust = 0.5, angle=90))
+p
+ggsave(filename='plots/parameters/cv_accuracy.png', plot = p, width = 6, height = 6, units = 'in', dpi = 300)
 ```
@@
 
-## LOO Training
+## LOO Discovery
 
@@
 rc9 = rc9 %>% set_names(c('Threshold','FPR','TPR','AUC','kb')) %>% 
-  mutate(kb = factor(kb,levels=c('15kb','50kb','100kb','500kb'), ordered=T), FPR = 1-FPR) %>% arrange(kb)
+  mutate(kb = factor(kb,levels=c('15kb','50kb','100kb','500kb'), ordered=T), FPR = 1-FPR) %>% arrange(kb) %>%
+  dplyr::select(-Threshold)
 
-ggplot(melt(rc9, id.vars='kb'), aes(variable, value, group=kb)) + 
+p = ggplot(melt(rc9, id.vars='kb'), aes(variable, value, group=kb)) + ylim(0,1) + 
   geom_bar(aes(fill=kb), stat='identity', position='dodge') + 
@@
   scale_fill_brewer(palette='Set1') + 
-  labs(title='Training ROC (alpha=0.9)', x='', y='%') + theme_minimal()
+  labs(title='Discovery ROC', subtitle = 'alpha=0.9', x='', y='%') + 
+  theme_minimal() + theme(text=element_text(size=12))
+p
+ggsave(filename='plots/parameters/training_kb_auc.png', plot = p, width = 6, height = 6, units = 'in', dpi = 300)
+
 ```
@@
 val.auc = val.auc %>% set_names(c('Threshold','FPR','TPR','AUC','kb')) %>% 
-  mutate(kb = factor(kb,levels=c('15kb','50kb','100kb','500kb'), ordered=T), FPR = 1-FPR) %>% arrange(kb)
+  mutate(kb = factor(kb,levels=c('15kb','50kb','100kb','500kb'), ordered=T), FPR = 1-FPR) %>% arrange(kb) %>%
+  dplyr::select(-Threshold)
 
-ggplot(melt(val.auc, id.vars='kb'), aes(variable, value, group=kb)) + 
+ggplot(melt(val.auc, id.vars='kb'), aes(variable, value, group=kb)) + ylim(0,1) +
   geom_bar(aes(fill=kb), stat='identity', position='dodge') + 
@@
 ```{r fig.height=7, fig.width=8}
-aucs = bind_rows(rc9 %>% mutate(cohort = 'Training'), val.auc %>% mutate(cohort = 'Validation'))
+aucs = bind_rows(rc9 %>% mutate(Cohort = 'Discovery'), val.auc %>% mutate(Cohort = 'Validation'))
 
-ggplot(aucs, aes(kb, AUC, group=cohort)) + 
-  geom_bar(aes(fill=cohort), stat='identity', position='dodge') + 
+p = ggplot(aucs, aes(kb, AUC, group=Cohort)) + ylim(0,1) + 
+  geom_bar(aes(fill=Cohort), stat='identity', position='dodge') + 
   geom_text(aes(label=round(AUC,2)), position=position_dodge(width=1)) + 
   scale_fill_brewer(palette='Set1') + 
-  labs(title='AUC training & validation', x='', y='AUC') + theme_minimal()
+  labs(title='AUC discovery & validation', x='', y='AUC') + 
+  theme_minimal() + theme(text = element_text(size=12))
+p
+ggsave(filename='plots/parameters/tv_auc_kb_compare.png', plot = p, width = 6, height = 6, units = 'in', dpi = 300)
+
 ```
 
-# Training LOO Performance
+# Discovery LOO Performance
 
@@
 
-## Training LOO alpha=0.9
+## Discovery LOO alpha=0.9
 
@@
 
-do.call(grid.arrange, c(plist, top='Training LOO alpha=0.9'))
+do.call(grid.arrange, c(plist[c('15kb','50kb','100kb','500kb')], top='Discovery LOO alpha=0.9'))
 ```
@@
     dplyr::rename('Total' = 'n') %>%
-    mutate_at(vars(Low,Moderate,High), funs(./Total)) 
+    mutate_at(vars(Low,Moderate,High), funs(./Total)) %>% mutate(kb = kb)
 })
@@
 plist = purrr::map(names(per.samp.train), function(kb) {
-  df = per.samp.train[[kb]]
-
-  ggplot(melt(df, id.vars=c('Status','Total')), aes(variable, value, group=Status, fill=Status)) + 
-    ylim(0,1) + geom_col(position='dodge') + 
-    geom_text(aes(label=round(value,2)), position = position_dodge(width=1)) + 
-    labs(title=unique(kb), x='', y='ratio') + theme_minimal()
+  df = melt(per.samp.train[[kb]], id.vars=c('Status','Total','kb')) %>% as_tibble %>% mutate(Status = recode_factor(Status, NP = 'Non-Progressor', P = 'Progressor'))
+  ggplot(df, aes(Status, value, group=Status, fill=variable)) + 
+    ylim(0,1) + geom_bar(stat = 'identity')  + 
+    geom_text(aes(label=round(value,2)), position = position_stack(vjust=0.5)) + 
+    scale_fill_manual(values =  BarrettsProgressionRisk:::riskColors(), name = 'Class') +
+    labs(title=unique(kb), x='', y='sample ratio') + theme_minimal() + theme(text = element_text(size = 12), legend.position = 'none')
 })
@@
 
-do.call(grid.arrange, c(plist[c('50kb','100kb','15kb','500kb')], top='Training predictions alpha=0.9'))
+p = do.call(grid.arrange, c(plist[c('15kb','50kb','100kb','500kb')], top='Discovery Cohort Classifications (alpha=0.9)'))
+
+ggsave(filename='plots/parameters/kb_classes.png', plot = p, width = 6, height = 12, units = 'in', dpi = 300)
+
+# x = do.call(bind_rows, per.samp.train) 
+# x = melt(x, id.vars=c('Status','Total', 'kb')) %>% as_tibble %>% mutate(Status = recode_factor(Status, NP = 'Non-Progressor', P = 'Progressor'), kb = factor(kb, levels=c('15kb','50kb','100kb','500kb'), ordered=T))
+#   ggplot(x, aes(Status, value, group=Status, fill=variable)) + facet_wrap(kb~Status, scale='free_x', nrow=2) + 
+#     ylim(0,1) + geom_bar(stat = 'identity')  + 
+#     geom_text(aes(label=round(value,2)), position = position_stack(vjust=0.5)) + 
+#     scale_fill_manual(values =  BarrettsProgressionRisk:::riskColors(), name = 'Class') +
+#     labs(title=unique(kb), x='', y='sample ratio') + theme_minimal() + theme(text = element_text(size = 12), legend.position = 'none')
 
@@
 })
+do.call(grid.arrange, c(plist[c('50kb','100kb','15kb','500kb')], top='Validation'))
 
-do.call(grid.arrange, c(plist[c('50kb','100kb','15kb','500kb')], top='Validation'))
+plist = purrr::map(names(val.counts), function(kb) {
+  df = melt(val.counts[[kb]], id.vars=c('Status','kb')) %>% as_tibble %>% mutate(Status = recode_factor(Status, NP = 'Non-Progressor', P = 'Progressor'))
+  ggplot(df, aes(Status, value, group=Status, fill=variable)) + 
+    ylim(0,1) + geom_bar(stat = 'identity')  + 
+    geom_text(aes(label=round(value,2)), position = position_stack(vjust=0.5)) + 
+    scale_fill_manual(values =  BarrettsProgressionRisk:::riskColors(), name = 'Class') +
+    labs(title=unique(kb), x='', y='sample ratio') + theme_minimal() + theme(text = element_text(size = 12), legend.position = 'none')
+})
+names(plist) = names(val.counts)
+
+p = do.call(grid.arrange, c(plist[c('15kb','50kb','100kb','500kb')], top='Validation Cohort Classifications (alpha=0.9)'))
+
+ggsave(filename='plots/parameters/kb_val_classes.png', plot = p, width = 6, height = 12, units = 'in', dpi = 300)
 

### 2019-12-24 Paper_resubmit.Rmd

@@
 
-val.info %>% group_by(Status) %>% dplyr::summarise(
-  `Total Patients` = length(unique(Patient)),
-  `Total Samples` = length(Samplename),
-`max follow-up` = max(`Weeks Follow Up`/52),
-`min follow-up` = min(`Weeks Follow Up`/52),
-  `mean follow-up` = mean(`Weeks Follow Up`/52),
-  `sd follow-up` = sd(`Weeks Follow Up`/52)
-)
-
-val.info %>% group_by(Status, Pathology) %>% tally
+val.info
 
@@
 failedQC = residuals %>% filter(varMAD_median > 0.008)
-
-val.info %>% filter(Samplename %in% failedQC$samplename)
 
@@
   dplyr::summarise_at(vars(months.before.final), list(~min(.), ~max(.), ~mean(.), ~median(.))) 
-
-```
+```
+
+## Check COX
+
+```{r}
+cox.cache.dir = '~/Data/BarrettsProgressionRisk/Analysis/cox_5e6_all/50kb/'
+file = paste(cox.cache.dir, 'model_data.Rdata', sep='/')
+load(file, verbose=T)
+
+file = paste(cox.cache.dir, "all.pt.alpha.Rdata" , sep='/')
+load(file, verbose=T)
+
+do.call(grid.arrange, c(plots, nrow=2))
+
+file = paste0(cox.cache.dir, '/loo_0.9.Rdata')
+load(file, verbose=T)
+
+rocCox = pROC::roc(Status ~ FittedRR, data=pg.samp, auc=T, ci=T, of='thresholds', transpose=T)
+roc.plot(rocCox) + labs(title = 'Cox model ROC')
+
+first = pg.samp %>% group_by(Patient) %>% arrange(Endoscopy.Year) %>% filter(Endoscopy.Year == Endoscopy.Year[1])
+rocCox2 = pROC::roc(Status ~ FittedRR, data = first, auc = T, ci = T, of='thresholds', transpose=T)
+roc.plot(rocCox2) + labs(title = 'Cox model ROC, first endo')
+
+last = filter(pg.samp, Endoscopy.Year == Final.Endoscopy)
+rocCox3 = pROC::roc(Status ~ FittedRR, data = last, auc = T, ci = T, of='thresholds', transpose=T)
+roc.plot(rocCox3) + labs(title = 'Cox model ROC, last endo')
+
+rocPath = pROC::roc(Status ~ Pathology, data=pg.samp, auc=T, ci=T, of='thresholds', transpose=T)
+roc.plot(rocPath) + labs(title = 'Pathology status ROC')
+
+cor.test(pg.samp$FittedRR, as.integer(pg.samp$Pathology))
+
+```

### 2019-12-24 snp/SNP_Prediction_Analysis.Rmd

@@
 ```{r, echo=F}
+sample.info = sample.info %>% mutate(Pathology = recode(Pathology, IMC = 'EAC'))
+
 sm = sample.info %>% group_by(Status, Endoscopy, Pathology) %>% dplyr::tally() %>% 
-  group_by(Status, Endoscopy) %>% dplyr::mutate(freq=n/sum(n) ) 
-sm$Pathology = factor(sm$Pathology, levels=c('NDBE','ID','LGD','HGD','IMC'), ordered = T)
+  group_by(Status, Endoscopy) %>% dplyr::mutate(freq=n/sum(n) ) %>% 
+  mutate(Pathology = factor(Pathology, levels=c('NDBE','ID','LGD','HGD','EAC'), ordered = T))
 
-ggplot(sm %>% filter(!Status == 'Normal'), aes(Pathology, freq, group=Endoscopy, fill=factor(Endoscopy))) + facet_grid(~Status) + 
+ggplot(sm %>% filter(!Status == 'Normal'), aes(Pathology, freq, group=Endoscopy, fill=factor(Endoscopy))) + 
+  facet_grid(~Status) + 
   geom_col(position = position_dodge()) + ylim(0,1) + labs(title='Pathology frequency per timepoint')
@@
 
-ggplot(qcdata, aes(Status, ASCAT.SCA.ratio)) + ylim(0,0.05) + 
-  facet_grid(~SampleType, scales='free_x') +
-  geom_jitter(width=0.1, aes(color=Status)) + 
-  geom_hline(data=dplyr::rename(anno, Status=SampleType), aes(yintercept = Ratio, linetype=Status), color='grey42') +
-  geom_label(data=dplyr::rename(anno, Status=SampleType), aes(x='NP', y=Ratio, label=label), alpha=0.5)
+p = ggplot(qcdata %>% filter(grepl('Normal',SampleType)), aes(Purity,ASCAT.SCA.ratio)) + 
+  ylim(0,0.02) +
+  geom_point(color='steelblue' ) + labs(title='Blood/Gastric Normal ASCAT values')
+p
+ggsave('plots/snp/normal_sca_purity.png', p, height=4, width=4, units='in')
+
+ggplot(qcdata %>% filter(!grepl('Normal',SampleType)), aes(Purity,ASCAT.SCA.ratio)) + 
+  #ylim(0,0.5) +
+  geom_point(color='steelblue' ) + 
+  geom_point(data = qcdata %>% filter(!grepl('Normal',SampleType) & ASCAT.SCA.ratio < 0.01 & Purity < 0.98), color='lightgreen') +
+  labs(title='BE biopsy ASCAT values')
+
+# 
+# ggplot(qcdata, aes(Status, ASCAT.SCA.ratio)) + ylim(0,0.05) + 
+#   facet_grid(~SampleType, scales='free_x') +
+#   geom_jitter(width=0.1, aes(color=Status)) + 
+#   geom_hline(data=dplyr::rename(anno, Status=SampleType), aes(yintercept = Ratio, linetype=Status), color='grey42') +
+#   geom_label(data=dplyr::rename(anno, Status=SampleType), aes(x='NP', y=Ratio, label=label), alpha=0.5)
 
@@
 
-ggplot(filter(qcdata,SampleType == 'BE'), aes(Timepoint, ASCAT.SCA.ratio, fill=Timepoint, group=Timepoint)) + geom_jitter(width=0.2) +
-  geom_boxplot(alpha=0.5) + facet_wrap(~Status, scales='free_x', ncol=2) + labs(title='SCA Ratio by Timepoint') + theme(legend.position = 'none')
+# ggplot(filter(qcdata,SampleType == 'BE'), aes(Timepoint, ASCAT.SCA.ratio, fill=Timepoint, group=Timepoint)) + 
+#   geom_jitter(width=0.2) + geom_boxplot(alpha=0.5) + 
+#   facet_wrap(~Status, scales='free_x', ncol=2) + 
+#   labs(title='SCA Ratio by Timepoint') + theme(legend.position = 'none')
 
@@
 p = ggplot(qcdata, aes(Purity, ASCAT.SCA.ratio)) + facet_wrap(~SampleType, nrow=1) + 
-  geom_point(aes(color=ASCAT.SCA.ratio < x | Purity > y), show.legend = F) + 
+  geom_point(aes(color=ASCAT.SCA.ratio < x & Purity > y), show.legend = F) + 
   geom_vline(xintercept=y, color='grey') + 
@@
 ggplot(subset(qcdata, SampleType == 'BE'), aes(Ploidy, ASCAT.SCA.ratio, color=Status)) + facet_grid(~Status) + geom_point() + 
-  geom_hline(yintercept = x, color='red') + labs(title='Barretts only')
+  geom_hline(yintercept = x, color='grey39', linetype = 'dashed') + labs(title='Barretts biopsies')
 
@@
 
+p = ggplot(qcdata.endo, aes(Purity_mean, ASCAT.SCA.ratio_mean)) + facet_grid(~SampleType) +
+  geom_point(aes(color=ASCAT.SCA.ratio_mean < 0.01 & Purity_mean > 0.95 ), show.legend = F) + 
+  labs(title='Merged biopsies per-timepoint', y='mean(ASCAT.SCA.ratio)', x='mean(ASCAT purity)')
+p
+ggsave('plots/snp/sca_purity_per_tp.png', p, width=4, height=4, units='in')
+
 tmp = qcdata.endo %>% filter(!is.na(Ploidy_mean)) %>% mutate(Ploidy_mean = factor(round(Ploidy_mean)))
@@
   theme(text=element_text(size=12), legend.position = 'none')
+p
 ggsave('plots/snp/sca_v_ploidy.png', plot=p, height=4, width=4, units = 'in', dpi=300)
@@
 
+patient.info = patient.info %>% mutate(Pathology = recode(Pathology, 'IMC' = 'EAC'))
+
 preds = preds %>% left_join(patient.info %>% dplyr::select(PatientID, UniqueSampleID, Pathology, Status, Path.Status), by=c('Samplename'='UniqueSampleID'))
@@
 ```{r echo=F}
-preds = preds %>% mutate(Pathology = factor(Pathology, levels=c('NDBE','ID','LGD','HGD','IMC'), ordered = T))
+preds = preds %>% mutate(Pathology = factor(Pathology, levels=c('NDBE','ID','LGD','HGD','EAC'), ordered = T))
 rp = plot.risk.by.path(preds) 
@@
 ```{r}
-roc = pROC::roc(Status ~ Prob, data=bind_rows(preds, normals %>% mutate(Status = 'NP')), auc=T, ci=T, of='thresholds',transpose=T)
-roc.plot(roc) + labs(title='SNP ROC', subtitle='All samples')
-pROC::ci.auc(roc$auc)
-pROC::coords(roc, 'best', transpose=F)
-ggsave(filename = 'plots/snp/all_samp_roc.png', plot = roc.plot(roc) + labs(title='SNP ROC', subtitle='All samples') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
+# roc = pROC::roc(Status ~ Prob, data=bind_rows(preds, normals %>% mutate(Status = 'NP')), auc=T, ci=T, of='thresholds',transpose=T)
+# roc.plot(roc) + labs(title='SNP ROC', subtitle='All samples')
+# pROC::ci.auc(roc$auc)
+# pROC::coords(roc, 'best', transpose=F)
+# ggsave(filename = 'plots/snp/all_samp_roc.png', plot = roc.plot(roc) + labs(title='SNP ROC', subtitle='All samples') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
 
 rocBE = pROC::roc(Status ~ Prob, data=preds, auc=T, ci=T, of='thresholds',transpose=T)
-roc.plot(rocBE) + labs(title='SNP ROC', subtitle='BE samples only')
+roc.plot(rocBE) + labs(title='SNP ROC', subtitle='BE biopsies')
 pROC::ci.auc(rocBE$auc)
 pROC::coords(rocBE, 'best', transpose=F)
-ggsave(filename = 'plots/snp/be_roc.png', plot = roc.plot(rocBE) + labs(title='SNP ROC', subtitle='BE samples only') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
+ggsave(filename = 'plots/snp/be_roc.png', plot = roc.plot(rocBE) + labs(title='SNP ROC', subtitle='BE biopsies') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
 ```
@@
 
+pred.qc %>% filter(!lowsca | wgd.NP) %>% ungroup %>% dplyr::select(PatientID) %>% distinct
+
 ascat.normals = qcdata %>% filter(grepl('Normal',SampleType)) %>% data.frame
@@
 rocEX = pROC::roc( Path.Status ~ Prob, data=pred.qc, auc=T, ci=T, of='thresholds', transpose=T)
-roc.plot(rocEX) + labs(title='BE samples', subtitle='Path status (HGD/IMC)')
+roc.plot(rocEX) + labs(title='BE biopsies', subtitle='Path status (HGD/IMC)')
 pROC::ci.auc(rocEX)
-ggsave(filename='plots/snp/path_status_roc.png', plot = roc.plot(rocEX) + labs(title='BE samples', subtitle='Path status (HGD/IMC)') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
+ggsave(filename='plots/snp/path_status_roc.png', plot = roc.plot(rocEX) + labs(title='BE biopsies', subtitle='Path status (HGD/IMC)') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
 
@@
 rocpp = pROC::roc(Status ~ Prob, data=filter(pred.qc, !lowsca), auc=T, ci=T, of='thresholds', transpose=T)
-roc.plot(rocpp) + labs(title='BE samples', subtitle='excluding low SCA samples')
-ggsave(filename='plots/snp/auc_ex_lowsca_roc.png', plot = roc.plot(rocpp) + labs(title='BE samples', subtitle='excluding low SCA samples') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
+roc.plot(rocpp) + labs(title='BE biopsies', subtitle='excluding low SCA samples')
+ggsave(filename='plots/snp/auc_ex_lowsca_roc.png', plot = roc.plot(rocpp) + labs(title='BE biopsies', subtitle='excluding low SCA samples') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
 
@@
 rocexall = pROC::roc(Status ~ Prob, data=filter(pred.qc, !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds', transpose=T)
-roc.plot(rocexall) + labs(title='BE samples', subtitle='excluding low SCA & WGD NP')
-ggsave(filename='plots/snp/auc_ex_lowsca_wgs_roc.png', plot =roc.plot(rocexall) + labs(title='BE samples', subtitle='excluding low SCA & WGD NP') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
+roc.plot(rocexall) + labs(title='BE biopsies', subtitle='excluding low SCA & WGD NP')
+ggsave(filename='plots/snp/auc_ex_lowsca_wgs_roc.png', plot =roc.plot(rocexall) + labs(title='BE biopsies', subtitle='excluding low SCA & WGD NP') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
 
@@
 roct1 = pROC::roc(Status ~ Prob, data=filter(pred.qc, Timepoint == 'T1' & !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds', transpose=T)
-roc.plot(roct1) + labs(title='BE samples', subtitle='Timepoint T1')
-ggsave(filename='plots/snp/auc_t1_roc.png', plot=roc.plot(roct1) + labs(title='BE samples', subtitle='Timepoint T1') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
+roc.plot(roct1) + labs(title='BE biopsies', subtitle='Timepoint T1')
+ggsave(filename='plots/snp/auc_t1_roc.png', plot=roc.plot(roct1) + labs(title='BE biopsies', subtitle='Timepoint T1') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
 
@@
 roct2 = pROC::roc(Status ~ Prob, data=subset(pred.qc, Timepoint != 'T1' & !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds', transpose=T)
-roc.plot(roct2) + labs(title='BE samples', subtitle='Timepoint T2')
-ggsave(filename='plots/snp/auc_t2_roc.png', plot=roc.plot(roct2) + labs(title='BE samples', subtitle='Timepoint T2') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
+roc.plot(roct2) + labs(title='BE biopsies', subtitle='Timepoint T2')
+ggsave(filename='plots/snp/auc_t2_roc.png', plot=roc.plot(roct2) + labs(title='BE biopsies', subtitle='Timepoint T2') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
 

### 2019-12-24 Paper_resubmit.Rmd
...

@@
 
-tt = t.test(subset(complexity.score, Status == 'Progressor')$cx, subset(complexity.score, Status == 'Non-Progressor')$cx)
+tt = wilcox.test(subset(complexity.score, Status == 'Progressor')$cx, subset(complexity.score, Status == 'Non-Progressor')$cx)
 pv = ggplot(complexity.score, aes(Status, cx, fill=Status)) + 
@@
 ggsave(filename='plots/auc/training_auc.png', plot=roc.plot(roc), width=6.5, height=4.5, units='in', dpi=300)
+ggsave(filename='plots/auc/training_auc_supp.png', plot=roc.plot(roc) + ggtitle('Per sample'), width=6, height=6, units='in', dpi=300)
 
@@
 rocFirstEndo = pROC::roc(Status ~ avg, data=firstEndo, ci=T, of='thresholds')
-
 rocFirstEndo$model = 'First endoscopy'
@@
 
-endo.thresholds = bind_cols('roc'=c('Max','Avg','Last','First'),
+pendo.rand = pg.samp %>% dplyr::group_by(Patient, Status, PID) %>%
+  dplyr::summarise( r.Prob = sample(Prediction, 1), n = length(PID) ) %>% filter(n > 1)
+rocRandEndo = pROC::roc(Status ~ r.Prob, data=pendo.rand, ci=T, of='thresholds', transpose = T)
+rocRandEndo$model = 'Per Endoscopy Random'
+ggsave('plots/auc/randomendo_auc.png', plot=roc.plot(rocRandEndo, title='Random per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+
+endo.thresholds = bind_cols('roc'=c('Max','Avg','Last','First','Random'),
   bind_rows(
@@
     pROC::coords(rocLastEndo, 'best', transpose=T),
-    pROC::coords(rocFirstEndo, 'best', transpose=T)
+    pROC::coords(rocFirstEndo, 'best', transpose=T),
+    pROC::coords(rocRandEndo, 'best', transpose=T)
   ))
-
 ```
@@
       get.roc.stat(rocPtM), get.roc.stat(rocPtA),
-      get.roc.stat(rocFirstEndo), get.roc.stat(rocLastEndo))
+      get.roc.stat(rocFirstEndo), get.roc.stat(rocLastEndo),
+      get.roc.stat(rocRandEndo))
+
 m$model = factor(m$model)
@@
 
-ggsave('plots/auc/multi_auc.png', plot=p, height=6, width=5, units='in',dpi=300)
+ggsave('plots/auc/multi_auc.png', plot=p, height=6, width=8, units='in',dpi=300)
 ```
@@
 
-ct = cor.test(pi$p11Loss, (as.numeric(pi$p53.Status)-1))
-```
-
-17p11 loss and p53 IHC correlation  R=`r round(ct$estimate,2)`.
+ct = fisher.test(table(pi$p11Loss, pi$p53.Status))
+
+```
+
+17p11 loss and p53 IHC correlation  Fisher's odds `r ct$estimate`, p-value = `r ct$p.value`.
 
@@
     endo = BarrettsProgressionRisk:::.setUpRxTablePerEndo(preds %>% dplyr::mutate(Endoscopy = months.before.final, Sample = 'Samplename'), 'max', verbose=F) %>% mutate(Endoscopy = Endoscopy*-1)
-    rx = BarrettsProgressionRisk:::.apply.rules(endo, 'Endoscopy', pred.confidence )
+    rx = BarrettsProgressionRisk:::.apply.rules(endo, 'Endoscopy', pred.confidence,'numeric' )
 
@@
 ```
+
+# Coefficients by # patients
+
+```{r}
+cache.dir = '~/Data/BarrettsProgressionRisk/Analysis/models_5e6_all/supp_train_n/insilico_splits/'
+
+purrr::map(list.files(cache.dir, 'loo.Rdata',recursive = T, full.names = T), function(f) {
+  load(f,verbose=T)
+
+  length(unique(pg.samp$Patient))
+
+}
+
+```

### 2020-01-06 Paper_resubmit.Rmd
fixed a label mixup that was causing my model training to muck up

@@
 ```{r}
-cache.dir = '~/Data/BarrettsProgressionRisk/Analysis/models_5e6_all/supp_train_n/insilico_splits/'
-
-purrr::map(list.files(cache.dir, 'loo.Rdata',recursive = T, full.names = T), function(f) {
-  load(f,verbose=T)
-
-  length(unique(pg.samp$Patient))
-
-}
-
-```
+model.dir = '~/Data/BarrettsProgressionRisk/Analysis/models_5e6_all/50kb/'
+load(list.files(model.dir, 'loo_0.9.Rdata',recursive = T, full.names = T))
+d.n.patient = length(unique(pg.samp$Patient))
+d.roc = pROC::roc(Status ~ Prediction, data = pg.samp, ci=T, of='thresholds', transpose=T)
+
+load(list.files(model.dir, 'all.pt.alpha',recursive = T, full.names = T))
+d.coefs = coefs[['0.9']]
+d.perf = performance.at.1se[['0.9']]$mean
+rm(dysplasia.df, plots, pg.samp, models, nzcoefs, fits, performance.at.1se)
+
+all.dir = '~/Data/BarrettsProgressionRisk/Analysis/dv_model_5e6_all_dvz/50kb/'
+load(list.files(all.dir, 'loo_0.9.Rdata',recursive = T, full.names = T))
+dv.n.patient = length(unique(pg.samp$Patient))
+dv.n.samples = length(unique(pg.samp$Samplename))
+dv.roc = pROC::roc(Status ~ Probability, data = pg.samp, ci=T, of='thresholds', transpose=T)
+
+load(list.files(all.dir, 'all.pt.alpha',recursive = T, full.names = T))
+dv.coefs = coefs[['0.9']]
+dv.perf = performance.at.1se[['0.9']]$mean
+rm(dysplasia.df, plots, pg.samp, models, nzcoefs, fits, performance.at.1se)
+
+load(list.files(all.dir, 'all.pt.alpha',recursive = T, full.names = T))
+dv.coefs = coefs[['0.9']]
+dv.perf = performance.at.1se[['0.9']]$mean
+rm(dysplasia.df, plots, pg.samp, models, nzcoefs, fits, performance.at.1se)
+
+cache.dir = '~/Data/BarrettsProgressionRisk/Analysis/models_5e6_all/'
+filenames = list.files(cache.dir, 'supp*')
+
+info = do.call(bind_rows, purrr::map(list.files(cache.dir, 'merged.Rdata', recursive = T, full.names = T), function(f) {
+  load(f,verbose=F)
+  model.info
+}))
+
+info = add_row(info, n.patients = d.n.patient, cv.performance = d.perf, n.coefs = nrow(d.coefs), n.samples = 773, auc = d.roc$auc)
+info = add_row(info, n.patients = dv.n.patient, cv.performance = dv.perf, n.coefs = nrow(dv.coefs), n.samples = dv.n.samples, auc = dv.roc$auc)
+
+x = info %>% group_by(n.patients) %>% dplyr::summarise_at(vars(-n.samples), list(~mean(.), ~sd(.))) %>% as_tibble %>% mutate(n.patients = factor(n.patients))
+
+p = ggplot(x, aes(x=n.patients, y=cv.performance_mean, group=n.patients, fill=n.patients)) +
+  geom_col() + geom_errorbar(aes(ymin=cv.performance_mean-cv.performance_sd/2, ymax=cv.performance_mean+cv.performance_sd/2), width=0.2, color='grey39') +
+  geom_text(aes(label=round(cv.performance_mean,2)), y=0.05) + 
+  scale_y_continuous(expand=c(0,0), limits=c(0,0.8)) + theme(legend.position = 'none') +
+  plot.theme + labs(title='Cross validation performance', subtitle='10 random samples', y='mean', x='# patients used to train')
+
+p1 = ggplot(x, aes(x=n.patients, y=n.coefs_mean, group=n.patients, fill=n.patients)) +
+  geom_col() + geom_errorbar(aes(ymin=n.coefs_mean-n.coefs_sd/2, ymax=n.coefs_mean+n.coefs_sd/2), width=0.2, color='grey39') +
+  geom_text(aes(label=round(n.coefs_mean,2)), y=5) +
+  scale_y_continuous(expand=c(0,0), limits=c(0,90)) + theme(legend.position = 'none') +
+  plot.theme + labs(title='# Coefficients selected by model', subtitle='10 random samples', y='mean', x='# patients used to train')
+
+p2 = ggplot(x, aes(x=n.patients, y=auc_mean, group=n.patients, fill=n.patients)) +
+  geom_col() + geom_errorbar(aes(ymin=auc_mean-auc_sd/2, ymax=auc_mean+auc_sd/2), width=0.2, color='grey39') +
+  geom_text(aes(label=round(auc_mean,2)), y=0.05) +
+  scale_y_continuous(expand=c(0,0), limits=c(0,1)) + theme(legend.position = 'none') +
+  plot.theme + labs(title='AUC from leave-one-out analysis', subtitle='10 random samples', y='mean', x='# patients used to train')
+
+ggsave(filename='plots/performance/training_n_pts.png',plot=grid.arrange(p,p1,p2,nrow=1), width = 10, height=8, units='in', dpi=300)
+
+```

### 2020-01-07 Paper_resubmit.Rmd
...

--- 2019-12-24 Paper_resubmit.Rmd
@@
                       ylab = 'Progression risk',
-                      ggtheme=theme_minimal(), palette=as.vector(BarrettsProgressionRisk::riskColors()), font.legend=c(14,'bold'))
+                      ggtheme=theme_minimal(), palette=unlist(BarrettsProgressionRisk::riskColors()), font.legend=c(14,'bold'))
 
@@
 
-## Check COX
+# Check COX
 
@@
 
-```{r}
-cache.dir = '~/Data/BarrettsProgressionRisk/Analysis/models_5e6_all/supp_train_n/insilico_splits/'
-
-purrr::map(list.files(cache.dir, 'loo.Rdata',recursive = T, full.names = T), function(f) {
-  load(f,verbose=T)
-
-  length(unique(pg.samp$Patient))
-
-}
-
-```
+```{r, fig.width=12, fig.height=8}
+model.dir = '~/Data/BarrettsProgressionRisk/Analysis/models_5e6_all/50kb/'
+load(list.files(model.dir, 'loo_0.9.Rdata',recursive = T, full.names = T))
+d.n.patient = length(unique(pg.samp$Patient))
+d.roc = pROC::roc(Status ~ Prediction, data = pg.samp, ci=T, of='thresholds', transpose=T)
+
+load(list.files(model.dir, 'all.pt.alpha',recursive = T, full.names = T))
+d.coefs = coefs[['0.9']]
+d.perf = performance.at.1se[['0.9']]$mean
+rm(dysplasia.df, plots, pg.samp, models, nzcoefs, fits, performance.at.1se)
+
+all.dir = '~/Data/BarrettsProgressionRisk/Analysis/dv_model_5e6_all_dvz/50kb/'
+load(list.files(all.dir, 'loo_0.9.Rdata',recursive = T, full.names = T))
+dv.n.patient = length(unique(pg.samp$Patient))
+dv.n.samples = length(unique(pg.samp$Samplename))
+dv.roc = pROC::roc(Status ~ Probability, data = pg.samp, ci=T, of='thresholds', transpose=T)
+
+load(list.files(all.dir, 'all.pt.alpha',recursive = T, full.names = T))
+dv.coefs = coefs[['0.9']]
+dv.perf = performance.at.1se[['0.9']]$mean
+rm(dysplasia.df, plots, pg.samp, models, nzcoefs, models, performance.at.1se)
+
+cache.dir = '~/Data/BarrettsProgressionRisk/Analysis/models_5e6_all/'
+filenames = list.files(cache.dir, 'supp*')
+
+info = do.call(bind_rows, purrr::map(list.files(cache.dir, 'merged.Rdata', recursive = T, full.names = T), function(f) {
+  load(f,verbose=F)
+  model.info
+})) %>% mutate('Cohort' = 'Discovery (patient subsets)')
+
+info = add_row(info, n.patients = d.n.patient, cv.performance = d.perf, n.coefs = nrow(d.coefs), n.samples = 773, auc = d.roc$auc, Cohort = 'Discovery (all)')
+
+info = add_row(info, n.patients = dv.n.patient, cv.performance = dv.perf, n.coefs = nrow(dv.coefs), n.samples = dv.n.samples, auc = dv.roc$auc, Cohort = 'Discovery+Validation')
+
+x = info %>% group_by(n.patients, Cohort) %>% dplyr::summarise_at(vars(-n.samples), list(~mean(.), ~sd(.))) %>% as_tibble %>% mutate(n.patients = factor(n.patients), Cohort = factor(Cohort, levels=c('Discovery (patient subsets)', 'Discovery (all)', 'Discovery+Validation'), ordered=T))
+
+p = ggplot(x, aes(x=n.patients, y=cv.performance_mean, group=n.patients, fill=Cohort)) + 
+  facet_grid(~Cohort, scales = 'free_x', space='free_x') + 
+  geom_col() + geom_errorbar(aes(ymin=cv.performance_mean-cv.performance_sd/2, ymax=cv.performance_mean+cv.performance_sd/2), width=0.2, color='grey39') +
+  geom_text(aes(label=round(cv.performance_mean,2)), y=0.05) + 
+  scale_y_continuous(expand=c(0,0), limits=c(0,0.8)) + 
+  scale_fill_brewer(palette = 'Set2',name='') + 
+  plot.theme + theme(legend.position = 'bottom', strip.text = element_blank(), panel.spacing = unit(0.5, 'lines')) +
+  labs(title='Cross validation performance', y='mean', x='# patients used to train')
+
+legend = get_legend(p)
+p = p + theme(legend.position = 'none')
+
+p1 = ggplot(x, aes(x=n.patients, y=n.coefs_mean, group=n.patients, fill=Cohort)) + 
+  facet_grid(~Cohort, scales = 'free_x', space='free_x') + 
+  geom_col() + geom_errorbar(aes(ymin=n.coefs_mean-n.coefs_sd/2, ymax=n.coefs_mean+n.coefs_sd/2), width=0.2, color='grey39') +
+  geom_text(aes(label=round(n.coefs_mean,2)), y=5) +
+  scale_y_continuous(expand=c(0,0), limits=c(0,125)) + scale_fill_brewer(palette = 'Set2') + 
+  plot.theme + theme(legend.position = 'none', strip.text = element_blank(), panel.spacing = unit(0.5, 'lines')) +
+  labs(title='# Coefficients',y='mean', x='# patients used to train')
+
+p2 = ggplot(x, aes(x=n.patients, y=auc_mean, group=n.patients, fill=Cohort)) +
+  facet_grid(~Cohort, scales = 'free_x', space='free_x') + 
+  geom_col() + geom_errorbar(aes(ymin=auc_mean-auc_sd/2, ymax=auc_mean+auc_sd/2), width=0.2, color='grey39') +
+  geom_text(aes(label=round(auc_mean,2)), y=0.05) +
+  scale_y_continuous(expand=c(0,0), limits=c(0,1)) + scale_fill_brewer(palette = 'Set2') + 
+  plot.theme + theme(legend.position = 'none', strip.text = element_blank(), panel.spacing = unit(0.5, 'lines')) +
+  labs(title='AUC', y='mean', x='# patients used to train')
+
+grid.arrange(arrangeGrob(p,p1,p2,nrow=1), arrangeGrob(legend), heights=c(4,1))
+
+ggsave(filename='plots/performance/training_n_pts.png',
+       plot=grid.arrange(arrangeGrob(p,p1,p2,nrow=1), arrangeGrob(legend), heights=c(4,1)), width=12, height=8, units='in', dpi=300)
+
+```

### 2020-01-07 processing_pipeline/run_cv_all_snp.R

@@ -2,10 +2,6 @@ args = commandArgs(trailingOnly=TRUE)
 
 set.seed(1234)
 
-if (length(args) < 3)
-  stop("Missing required params: <data dir> <outdir> <info file dir> <Set: All,Training>")
-
-
 suppressPackageStartupMessages( library(BarrettsProgressionRisk) )
 suppressPackageStartupMessages( library(tidyverse) )
 suppressPackageStartupMessages( library(glmnet) )
@@ -30,20 +26,39 @@ get.tiles<-function(files, info, scale=T) {
   prep.matrix(seg.tiles, scale=scale, MARGIN=2)
 }
 
-d.data = args[1]
-# d.data = '~/Data/BarrettsProgressionRisk/Analysis/pcf_perPatient/50kb/'
-v.data = args[2]
-# v.data = '~/Data/BarrettsProgressionRisk/Analysis/validation/pcf_perPatient/50kb/'
-outdir = args[3]
-# outdir = '~/Data/BarrettsProgressionRisk/Analysis/dv_model_5e6_all_dvz/50kb/'
-infodir = args[4]
-# infodir = '~/Data/BarrettsProgressionRisk/QDNAseq'
+d.data = '~/Data/BarrettsProgressionRisk/Analysis/pcf_perPatient/50kb/'
+v.data = '~/Data/BarrettsProgressionRisk/Analysis/validation/pcf_perPatient/50kb/'
+snp.data = '~/Data/Reid_SNP/PerPatient/'
+outdir = '~/Data/BarrettsProgressionRisk/Analysis/dv_snp_model/50kb/'
+infodir = '~/Data/BarrettsProgressionRisk/QDNAseq'
+
+## SNP tiles - not yet adjusted
+load('~/Data/Reid_SNP/PerPatient/tmp_seg_pt.Rdata', verbose=T) 
+rownames(mergedSegs) = sub('\\.LogR','', rownames(mergedSegs))
+rownames(mergedArms) = sub('\\.LogR','', rownames(mergedArms))
+snp.segs = mergedSegs
+snp.arms = mergedArms
+rm(mergedSegs, mergedArms)
+
+nm.rows = grep('BLD|GASTRIC', rownames(snp.segs))
+snp.segs = snp.segs[-nm.rows,]
+snp.arms = snp.arms[-nm.rows,]
+
+# difference between the two is 0.98 so...
[…]

### 2020-01-08 Paper_resubmit.Rmd
...

@@
 
-p11Loss %>% filter(value <= -1) %>% group_by(Status) %>% tally(name='p11 Loss') %>% 
+p11Loss %>% group_by(Status) %>% tally(name='p11 Loss') %>% 
   full_join( p11 %>% group_by(Status) %>% tally(name='Total Samples') ) %>% 
@@
 
-legend = get_legend(p)
+legend = cowplot::get_legend(p)
 p = p + theme(legend.position = 'none')
@@
 
-grid.arrange(arrangeGrob(p,p1,p2,nrow=1), arrangeGrob(legend), heights=c(4,1))
+grid.arrange(arrangeGrob(p,p1,p2,ncol=1), arrangeGrob(legend), heights=c(4,1))
 
 ggsave(filename='plots/performance/training_n_pts.png',
-       plot=grid.arrange(arrangeGrob(p,p1,p2,nrow=1), arrangeGrob(legend), heights=c(4,1)), width=12, height=8, units='in', dpi=300)
-
-```
+       plot=grid.arrange(arrangeGrob(p,p1,p2,ncol=1), arrangeGrob(legend), heights=c(6,1)), width=8, height=12, units='in', dpi=300)
+
+```

### 2020-01-08 snp/SNP_Prediction_Analysis.Rmd
...

@@
 preds = preds %>% mutate(Pathology = factor(Pathology, levels=c('NDBE','ID','LGD','HGD','EAC'), ordered = T))
-rp = plot.risk.by.path(preds) 
+rp = plot.risk.by.path(preds) + labs(y='% Classified', title='Samples classified by pathology')
 rp
@@
   scale_fill_gradientn(colors = myPal,  name='') + 
-  labs(title='Predictions on progressors', y='n Samples', x='Unadjusted Probability') + theme(text=element_text(size=12))
+  labs(title='Progressor BE', y='n Samples', x='Unadjusted Probability') + theme(text=element_text(size=12))
 
@@
   scale_fill_gradientn(colors = myPal,  name='') + 
-  labs(title='Predictions on non-progressors', y='n Samples', x='Unadjusted Probability') + theme(text=element_text(size=12))
+  labs(title='Non-Progressor BE', y='n Samples', x='Unadjusted Probability') + theme(text=element_text(size=12))
 
@@
   scale_fill_gradientn(colors = myPal,  name='') +
-  labs(title='Predictions on Blood/Gastric normals', y='n Samples', x='Unadjusted Probability') + theme(text=element_text(size=12))
+  labs(title='Blood/Gastric Normals', y='n Samples', x='Unadjusted Probability') + theme(text=element_text(size=12))
 

### 2020-01-10 Paper_resubmit.Rmd
...

@@
 
+error = read_tsv('~/Data/BarrettsProgressionRisk/Analysis/pcf_perPatient/50kb/sample_error.tsv')
+
 file = paste(cache.dir, 'loo_0.9.Rdata', sep='/')
@@
 
+error = error %>% left_join(pg.samp %>% dplyr::select(Status, Samplename, RR), by=c('Sample'='Samplename')) %>% 
+  dplyr::mutate(P.High = BarrettsProgressionRisk:::pi.hat(RR+Error), P.Low = BarrettsProgressionRisk:::pi.hat(RR-Error), Abs.Error = P.High-P.Low )
+
+error %>% group_by(Status) %>% dplyr::summarise_at(vars(Abs.Error), list(~mean, ~sd))
+summary(error$Abs.Error)
+sd(error$Abs.Error)
+
 pg.samp = pg.samp %>% mutate(Pathology = recode(Pathology, 'BE' = 'NDBE'))
+
+pg.samp = pg.samp %>% left_join(error %>% dplyr::select(-RR,-Status), by=c('Samplename'='Sample'))
 
@@
 
-```{r, eval=F}
-file = paste(cache.dir, 'model_data.Rdata', sep='/')
-load(file, verbose=T)
-
-vdir ='~/Data/BarrettsProgressionRisk/Analysis/validation/pcf_perPatient/50kb'
-v.tiles = do.call(bind_rows, purrr::map(list.files(vdir, '5e06_tiled_segvals.txt', full.names = T, recursive = T), function(f) {
-  read_tsv(f, col_types = cols(.default=col_double(), 'X1'=col_character()))
-})) %>% dplyr::rename(Samplename = 'X1')
-v.tiles = filter(v.tiles, Samplename %in% val.info$Samplename)
-v.tiles = v.tiles[-which(duplicated(v.tiles$Samplename))[2],]
-
-v.arm.tiles = do.call(bind_rows, purrr::map(list.files(vdir, 'arms_tiled_segvals.txt', full.names = T, recursive = T), function(f) {
-  read_tsv(f, col_types = cols(.default=col_double(), 'X1'=col_character()))
-})) %>% dplyr::rename(Samplename = 'X1')
-v.arm.tiles = filter(v.arm.tiles, Samplename %in% val.info$Samplename)
-v.arm.tiles = v.arm.tiles[-which(duplicated(v.arm.tiles$Samplename))[2],]
-
-for (i in 1:length(z.mean))
-  v.tiles[, i+1] = BarrettsProgressionRisk:::unit.var(v.tiles[,i+1], z.mean[i], z.sd[i])
-
-for (i in 1:length(z.arms.mean))
-  v.arm.tiles[, i+1] = BarrettsProgressionRisk:::unit.var(v.arm.tiles[,i+1], z.arms.mean[i], z.arms.sd[i])
-
-vt = dplyr::select(v.tiles, -Samplename) %>% as.matrix; rownames(vt) = v.tiles$Samplename
-at = dplyr::select(v.arm.tiles, -Samplename) %>% as.matrix; rownames(at) = v.arm.tiles$Samplename
-
-v.df = BarrettsProgressionRisk:::subtractArms(vt, at)
-
-valCNdf = as_tibble(t(v.df), rownames='locs') %>% set_names(c('locs',val.info$Samplename)) %>%
-  separate(locs, c('chr','start','end'), sep=':|-', remove=F) %>%   
-  dplyr::mutate_at(vars(start, end), as.numeric) %>%
-  #dplyr::mutate( seg.type = '5MB' ) %>%
-  dplyr::mutate( seg.type = ifelse( (start %in% arm.loc$start & end %in% arm.loc$end), 'arms','5MB'  ) ) %>%
-  melt(id.vars=c('locs','chr','start','end', 'seg.type')) %>% as_tibble() %>%
-  left_join(chr.lengths %>% dplyr::select(chrom,chr.length), by=c('chr'='chrom')) %>%
-  left_join(val.info %>% dplyr::select(matches('Hospital|Patient|Samplename|Pathology|Status')), by=c('variable'='Samplename')) %>%
-  dplyr::mutate(Status = factor(recode(Status, NP='Non-Progressor',P='Progressor'), ordered=T)) %>% 
-  dplyr::mutate( 
-    cn = factor(unlist(purrr::map(.x=value, .f=get.cn.state)),  levels=c('loss','neutral','gain'), ordered = T),
-    chr = factor(chr, levels = chr.lengths$chrom, ordered=T))
-
-n.status <- c('Progressor'=paste0('Progressor (n=',length(unique(subset(valCNdf, Status == 'Progressor')$variable)), ')'),
-              'Non-Progressor'=paste0('Non-Progressor (n=',length(unique(subset(valCNdf, Status == 'Non-Progressor')$variable)),')'))
-
-#cn.mtn.plot(filter(valCNdf, seg.type == '5MB'), labeller(Status=n.status, chr=label_value)) + ylim(-30,20)
-
-```
-
 ## Per sample (alpha = 0.9)
@@
 vdir ='~/Data/BarrettsProgressionRisk/Analysis/validation/pcf_perPatient/50kb/predictions_5e6_all'
-
-residuals = do.call(bind_rows, purrr::map(list.files( dirname(vdir), 'residuals.tsv', full.names = T, recursive = T ), function(f) {
-  read_tsv(f, col_types = 'ccddddl')
-}))
-
-failedQC = residuals %>% filter(varMAD_median > 0.008)
 
@@
 
-```{r, fig.width=12, fig.height=8}
+```{r, echo=F}
+# SNP
+snp.info = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/Analysis/SNP/metadata_T1T2.xlsx') %>% 
+  mutate( PID = paste(PatientID, `Timepoint Code`, sep='_'), Patient = as.numeric(PatientID), Hospital.Research.ID = as.character(Patient), Samplename = PID )
+
+model.dir = '~/Data/BarrettsProgressionRisk/Analysis/dv_snp_model/50kb/'
+load(list.files(model.dir, 'all.pt.alpha',recursive = T, full.names = T))
+
+dvsnp.coefs = coefs[['0.9']]
+dvsnp.perf = performance.at.1se[['0.9']]$mean
+rm(dysplasia.df, plots, pg.samp, models, nzcoefs, fits, performance.at.1se)
+
+dvsnp.preds = purrr::map(list.files('~/Data/BarrettsProgressionRisk/Analysis/dv_snp_model/50kb/loo', full.names = T), function(f) read_tsv(f, col_types = 'ccccccdd') )
+
+dvsnp.preds = do.call(bind_rows, dvsnp.preds)
+dvsnp.patient = length(unique(dvsnp.preds$Patient))
+dvsnp.roc = pROC::roc(Status ~ Probability, data=dvsnp.preds, ci=T, of='thresholds', transpose=T)
+
+dvsnp.dv.auc = pROC::auc(Status ~ Probability, data = dvsnp.preds %>% filter(cohort != 'snp') )
+dvsnp.snp.auc = pROC::auc(Status ~ Probability, data = dvsnp.preds %>% filter(cohort == 'snp') )
+
+dvsnp.preds = dvsnp.preds %>% mutate(type = ifelse(cohort == 'snp', 'snp', 'sWGS'))
+
+dvsnp.aucs = left_join(dvsnp.preds %>% dplyr::select(Patient, type) %>% distinct %>% group_by(type) %>% tally(name='# patients'),
+            dvsnp.preds %>% dplyr::select(Patient, type) %>% group_by(type) %>% tally(name='# timepoint samples'), by='type') %>% 
+  add_row( type = 'All (snp+sWGS)', `# patients` = dvsnp.preds %>% dplyr::select(Patient) %>% distinct %>% tally %>% pull, `# timepoint samples` = dvsnp.preds %>% dplyr::select(Patient) %>% tally %>% pull) %>%
+  arrange(type) %>% add_column( auc = c(dvsnp.roc$auc, dvsnp.snp.auc, dvsnp.dv.auc) ) %>%
+  mutate_at(vars(-auc), as.factor)
+
+dv.snp.p = ggplot(dvsnp.aucs , aes(`# patients`, auc, group=type, fill=type)) + geom_col() + 
+  geom_text(aes(label=round(auc,2), y=0.05)) + 
+  scale_y_continuous(expand=c(0,0), limits=c(0,1)) + scale_fill_brewer(palette = 'Set2', name='') + 
+  labs(title='sWGS (all) + SNP Model', y='AUC') + theme(legend.position = 'bottom') +
+  plot.theme
+dv.snp.p
+
+ggsave(filename='plots/performance/training_snpdv.png', plot=dv.snp.p, width=6, height=10, units='in', dpi=300)
+
+# discovery + validation
 model.dir = '~/Data/BarrettsProgressionRisk/Analysis/models_5e6_all/50kb/'
@@
 dv.perf = performance.at.1se[['0.9']]$mean
-rm(dysplasia.df, plots, pg.samp, models, nzcoefs, models, performance.at.1se)
-
+rm(dysplasia.df, plots, pg.samp, models, nzcoefs, cvs, performance.at.1se)
+```
+
+```{r, fig.width=12, fig.height=8}
 cache.dir = '~/Data/BarrettsProgressionRisk/Analysis/models_5e6_all/'
@@
   model.info
-})) %>% mutate('Cohort' = 'Discovery (patient subsets)')
-
-info = add_row(info, n.patients = d.n.patient, cv.performance = d.perf, n.coefs = nrow(d.coefs), n.samples = 773, auc = d.roc$auc, Cohort = 'Discovery (all)')
-
-info = add_row(info, n.patients = dv.n.patient, cv.performance = dv.perf, n.coefs = nrow(dv.coefs), n.samples = dv.n.samples, auc = dv.roc$auc, Cohort = 'Discovery+Validation')
-
-x = info %>% group_by(n.patients, Cohort) %>% dplyr::summarise_at(vars(-n.samples), list(~mean(.), ~sd(.))) %>% as_tibble %>% mutate(n.patients = factor(n.patients), Cohort = factor(Cohort, levels=c('Discovery (patient subsets)', 'Discovery (all)', 'Discovery+Validation'), ordered=T))
-
-p = ggplot(x, aes(x=n.patients, y=cv.performance_mean, group=n.patients, fill=Cohort)) + 
+})) %>% mutate('Cohort' = 'Discovery (patient subsets)', Type = 'sWGS')
+
+info = add_row(info, n.patients = d.n.patient, cv.performance = d.perf, n.coefs = nrow(d.coefs), n.samples = 773, auc = d.roc$auc, Cohort = 'Discovery (all)', Type = 'sWGS')
+
+info = add_row(info, n.patients = dv.n.patient, cv.performance = dv.perf, n.coefs = nrow(dv.coefs), n.samples = dv.n.samples, auc = dv.roc$auc, Cohort = 'Discovery+Validation', Type = 'sWGS')
+
+info = add_row(info, n.patients = dvsnp.patient, cv.performance = dvsnp.perf, n.coefs = nrow(dvsnp.coefs), n.samples = nrow(dvsnp.preds), auc = dvsnp.roc$auc, Cohort = 'sWGS (all) + SNP', Type = 'All')
+
+info = add_row(info, n.patients = dvsnp.aucs %>% filter(type == 'snp') %>% dplyr::select(`# patients`) %>% pull, cv.performance = NA, n.coefs = NA, n.samples = NA, auc = dvsnp.aucs %>% filter(type == 'snp') %>% dplyr::select(auc) %>% pull, Cohort = 'sWGS (all) + SNP', Type = 'SNP')
+
+info = add_row(info, n.patients = dvsnp.aucs %>% filter(type == 'sWGS') %>% dplyr::select(`# patients`) %>% pull, cv.performance = NA, n.coefs = NA, n.samples = NA, auc = dvsnp.aucs %>% filter(type == 'sWGS') %>% dplyr::select(auc) %>% pull, Cohort = 'sWGS (all) + SNP', Type = 'sWGS')
+
+x = info %>% group_by(n.patients, Cohort, Type) %>% dplyr::summarise_at(vars(-n.samples), list(~mean(.), ~sd(.))) %>% as_tibble %>% mutate(n.patients = factor(n.patients,levels=c( seq(28,88,10), 164, 413, 165, 248  ), ordered=T), Cohort = factor(Cohort, levels=c('Discovery (patient subsets)', 'Discovery (all)', 'Discovery+Validation', 'sWGS (all) + SNP'), ordered=T))
+
+p = ggplot( x %>% filter(!is.na(cv.performance_mean)), aes(x=n.patients, y=cv.performance_mean, group=n.patients, fill=Cohort)) + 
   facet_grid(~Cohort, scales = 'free_x', space='free_x') + 
@@
   plot.theme + theme(legend.position = 'bottom', strip.text = element_blank(), panel.spacing = unit(0.5, 'lines')) +
-  labs(title='Cross validation performance', y='mean', x='# patients used to train')
+  labs(title='Cross validation performance', y='', x='# patients used to train')
 
@@
 
-p1 = ggplot(x, aes(x=n.patients, y=n.coefs_mean, group=n.patients, fill=Cohort)) + 
+p1 = ggplot(x %>% filter(!is.na(n.coefs_mean)), aes(x=n.patients, y=n.coefs_mean, group=n.patients, fill=Cohort)) + 
   facet_grid(~Cohort, scales = 'free_x', space='free_x') + 
@@
   geom_text(aes(label=round(n.coefs_mean,2)), y=5) +
-  scale_y_continuous(expand=c(0,0), limits=c(0,125)) + scale_fill_brewer(palette = 'Set2') + 
+  scale_y_continuous(expand=c(0,0)) + scale_fill_brewer(palette = 'Set2') + 
   plot.theme + theme(legend.position = 'none', strip.text = element_blank(), panel.spacing = unit(0.5, 'lines')) +
-  labs(title='# Coefficients',y='mean', x='# patients used to train')
+  labs(title='# Coefficients',y='', x='# patients used to train')
 
@@
   facet_grid(~Cohort, scales = 'free_x', space='free_x') + 
-  geom_col() + geom_errorbar(aes(ymin=auc_mean-auc_sd/2, ymax=auc_mean+auc_sd/2), width=0.2, color='grey39') +
+  geom_col() + 
+  geom_errorbar(aes(ymin=auc_mean-auc_sd/2, ymax=auc_mean+auc_sd/2), width=0.2, color='grey39') +
   geom_text(aes(label=round(auc_mean,2)), y=0.05) +
+  geom_text( data=filter(x,Cohort == 'sWGS (all) + SNP'), aes(label=Type, y=0.1) ) +
   scale_y_continuous(expand=c(0,0), limits=c(0,1)) + scale_fill_brewer(palette = 'Set2') + 
   plot.theme + theme(legend.position = 'none', strip.text = element_blank(), panel.spacing = unit(0.5, 'lines')) +
-  labs(title='AUC', y='mean', x='# patients used to train')
+  labs(title='AUC', y='', x='# patients used to train')
 

### 2020-01-10 processing_pipeline/process_validation.R

@@ -94,7 +94,7 @@ for (ra in levels(all.val$RA) ) {
 
     tryCatch({
       info = loadSampleInformation(si %>% filter(Samplename == samples))
-      segmented = BarrettsProgressionRisk::segmentRawData(info,rd,fd,cutoff=0.011,kb=kb,multipcf=kb<100,verbose=T)
+      segmented = BarrettsProgressionRisk::segmentRawData(info,rd,fd,kb=kb,multipcf=kb<100,verbose=T)
       
       raw_dist = tibble(
         patient = segmented$sample.info$Hospital.Research.ID,
@@ -141,6 +141,10 @@ for (ra in levels(all.val$RA) ) {
 
 				write.table(tiles$tiles, sep='\t', quote=F, col.names=NA, row.names=T, file=paste0(dirname(plot.dir), '/5e06_cleaned_tiled.tsv'))
 				write.table(arms$tiles, sep='\t', quote=F, col.names=NA, row.names=T, file=paste0(dirname(plot.dir), '/arms_cleaned_tiled.tsv'))
+				
+				write.table(tiles$tiles, sep='\t', quote=F, col.names=NA, row.names=T, file=paste0(dirname(plot.dir), '/5e06_cleaned_tiled.tsv'))
+				write.table(arms$tiles, sep='\t', quote=F, col.names=NA, row.names=T, file=paste0(dirname(plot.dir), '/arms_cleaned_tiled.tsv'))
+				
   		} 
   }, error = function(e) {
     message(paste("Error in segmentation for patient",pid,'from RA:', ra, ', skipping:\n\t',e))

### 2020-01-14 Paper_resubmit.Rmd
last bits and pieces for paper

@@
 featureStability = rowSums(coefs[[select.alpha]][,-1])/(splits*folds)
-hazards = apply( exp(t(dysplasia.df[, rownames(coefs[[select.alpha]])]) *  coefs[[select.alpha]][,1]), 1, function(x) {
+hazards = apply( exp( t(dysplasia.df[,rownames(coefs[[select.alpha]])]) * coefs[[select.alpha]][,1]), 1, function(x) {
   sd(x)/mean(x)
@@
 ```{r}
-pg.samp = do.call(bind_rows,lapply(unique(pg.samp$Patient) %>% map( function(p) pg.samp %>% filter(Patient == p)), transform.by.time, info=patient.info)) 
+pg.samp = do.call(bind_rows,lapply(unique(pg.samp$Patient) %>% 
+                                     map( function(p) pg.samp %>% filter(Patient == p)), transform.by.time, info=patient.info)) 
 sq = c(0, seq(1,132,12),max(pg.samp$months.before.final))
@@
   scale_fill_gradientn(colors = myPal,  name='') + 
-  labs(y='n Samples', x='Relative Risk', title='Unadjusted relative risk') + plot.theme
+  labs(y='n Samples', x='log(Relative Risk)', title='Unadjusted relative risk') + plot.theme
 unadjRR
@@
   'ratio'=length(which(seg.type != 'arms' & cn %in% c('loss','gain')))/length(cn)
-  ) %>% mutate(Patient = factor(Patient))
+  ) %>% ungroup %>% mutate(Patient = factor(Patient))
 #as.data.frame(df2)
@@
 el2 = el %>% dplyr::group_by(Status, Patient) %>% dplyr::summarise( diff(arms), diff(CNA) )
-wtA = round(wilcox.test(subset(el2, Status == 'Progressor')$`diff(arms)`, subset(el2, Status == 'Non-Progressor')$`diff(arms)`)$p.value, 3)
-if (wtA < 0.0001) wtA = '< 0.0001'
+wtA = signif(wilcox.test(subset(el2, Status == 'Progressor')$`diff(arms)`, subset(el2, Status == 'Non-Progressor')$`diff(arms)`)$p.value,3)
+#if (wtA < 0.0001) wtA = '< 0.0001'
 
@@
   geom_point() +  geom_line(data=subset(df2, early | late)) + theme(legend.position = 'none', text=element_text(size=11)) + 
-  plot.theme + labs(x='Endoscopy Year', y='CN arm values', title='Initial vs final endoscopy', subtitle=paste('p', wtA))
+  plot.theme + labs(x='Endoscopy Year', y='CN arm values', title='Initial vs final endoscopy', subtitle=paste('p-value=', wtA))
 p
@@
 
-wtA = round(wilcox.test(subset(df2, early &  Status == 'Progressor')$arms, subset(df2, early & Status == 'Non-Progressor')$arms)$p.value, 3)
-wtCN = round(wilcox.test(subset(df2, early & Status == 'Progressor')$CNA, subset(df2, early & Status == 'Non-Progressor')$CNA)$p.value, 2)
+wtA = signif(wilcox.test(subset(df2, early &  Status == 'Progressor')$arms, subset(df2, early & Status == 'Non-Progressor')$arms)$p.value, 3)
+wtCN = signif(wilcox.test(subset(df2, early & Status == 'Progressor')$CNA, subset(df2, early & Status == 'Non-Progressor')$CNA)$p.value, 2)
 
@@
   plot.theme + theme(legend.position = 'none', text=element_text(size=11)) +
-  labs(y='', x='', title='CN values, initial endoscopy', subtitle=paste('Arms p=',wtA, '   CNA p=',wtCN, sep=''))
+  labs( y='Segment count', title='CN values, initial endoscopy', subtitle=paste('Arms p=',wtA, '   CNA p=',wtCN, sep=''))
 p
@@
 
-rp = plot.risk.by.path(pg.samp %>% ungroup)
-rp
+rp = plot.risk.by.path(pg.samp %>% ungroup) + labs(title='Samples classified by pathology', y='% Classified')
+
 ggsave(filename='plots/pred_by_path.png', plot=(rp + theme(legend.position = 'none')), width=7, height=6, units='in', dpi=300)
@@
 
-```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=6}
+```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=6, eval=F}
 all.preds = pg.samp %>% dplyr::group_by(Status, Patient, Risk) %>% tally %>% spread(Risk, n) %>% ungroup %>%
@@
   geom_text(aes(y=5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white', nudge_y=2) + 
-  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') + coord_cartesian(ylim = c(0,100)) + 
+  labs(x='Months prior to endpoint', y='Percent classified', title='High risk classifications prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') + coord_cartesian(ylim = c(0,100)) +
   plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12) ) 
@@
 
+vdir ='~/Data/BarrettsProgressionRisk/Analysis/validation/pcf_perPatient/50kb'
+residuals = do.call(bind_rows, purrr::map(list.files(vdir,'residuals.tsv',recursive = T, full.names = T), function(f) read_tsv(f, col_types = 'ccddddl')))
+failedQC = filter(residuals, varMAD_median > 0.008)
 ```
@@
 
-all.val.preds = all.val.preds %>% filter( !Sample %in% failedQC$samplename &  
-                                            !`Hospital Research ID` %in% unique(oac.info$`Hospital Research ID`) )
+all.val.preds = all.val.preds %>% 
+  filter( !Sample %in% failedQC$samplename & !`Hospital Research ID` %in% unique(oac.info$`Hospital Research ID`) )
 
 val.preds = left_join(all.val.preds, val.info, by=c('Hospital Research ID', 'Sample'='Samplename')) %>% 
-  dplyr::select(Patient, everything()) %>% mutate(Pathology = factor(Pathology, levels=c('NDBE','ID','LGD'), ordered=T)) %>% 
+  dplyr::select(Patient, everything()) %>% 
+  mutate(Pathology = factor(Pathology, levels=c('NDBE','ID','LGD'), ordered=T)) %>% 
   mutate(Final.Endoscopy = `Last Endoscopy`, Endoscopy.Year = Endoscopy, Block = factor(Block, ordered=T))
@@
 ```{r}
-rp = plot.risk.by.path(val.preds) + labs(title='Validation samples classified by pathology')  
+rp = plot.risk.by.path(val.preds) + labs(title='Validation samples classified by pathology', y='% Classified')  
 rp
@@
 
-val.preds = do.call(bind_rows, lapply(unique(val.preds$Patient) %>% 
+val.preds = do.call(bind_rows, 
+                    lapply(unique(val.preds$Patient) %>% 
                             map( function(p) val.preds %>% filter(Patient == p) %>% 
@@
 ```{r}
-vp = val.preds %>% dplyr::select(Patient, Status, months.before.final, Risk) %>% mutate(Patient = as.factor(Patient)) %>% 
+vp = val.preds %>% dplyr::select(Patient, Status, months.before.final, Risk) %>% 
+  mutate(Patient = as.factor(Patient)) %>% 
   group_by(Patient) %>% mutate(Diagnosis = as.integer(Status)-1)
@@
 
-sfit = survfit( Surv(months.before.final, Diagnosis, type='right') ~ Risk, data=merged, conf.int=0.95, conf.type='log', na.action=na.omit)
+#merged = merged %>% mutate(Diagnosis = ifelse(Status == 'P' & months.before.final == 0, 1, 0))
+
+merged %>%
+
+sfit = survfit( Surv(months.before.final, Diagnosis) ~ Risk, data=merged, conf.int=0.95, conf.type='log', na.action=na.omit)
 names(sfit$strata) = sub('.*=','', names(sfit$strata))
 
-kmp = survminer::ggsurvplot(sfit, size=1, linetype="strata",risk.table=T, risk.table.col='strata',
-                      conf.int=T, censor=T, legend='none', 
-                      title='Risks over time', 
-                      subtitle='Training & Validation predictions',
-                      xlab='Time (months prior to final sample)',
-                      ylab = 'Progression risk',
+kmp = 
+  survminer::ggsurvplot(sfit, size=1, linetype="strata",risk.table=F, risk.table.col='strata',
+                      conf.int=T, censor=T, legend='none',# xlim=c(0,150),
+                      title='Risk Classifications', 
+                      subtitle='Training & Validation cohorts',
+                      xlab='Months prior to final sample',
+                      ylab = 'Progression free fraction',
                       ggtheme=theme_minimal(), palette=unlist(BarrettsProgressionRisk::riskColors()), font.legend=c(14,'bold'))
+
+kmp$plot = kmp$plot + scale_x_reverse() 
+kmp$table = kmp$table + scale_x_reverse()
 
@@
   plot.theme + theme(legend.position = 'bottom', strip.text = element_blank(), panel.spacing = unit(0.5, 'lines')) +
-  labs(title='Cross validation performance', y='', x='# patients used to train')
+  labs(title='Cross validation performance', y='', x='# patients')
 
@@
   plot.theme + theme(legend.position = 'none', strip.text = element_blank(), panel.spacing = unit(0.5, 'lines')) +
-  labs(title='# Coefficients',y='', x='# patients used to train')
+  labs(title='# Coefficients',y='', x='# patients')
 
@@
   plot.theme + theme(legend.position = 'none', strip.text = element_blank(), panel.spacing = unit(0.5, 'lines')) +
-  labs(title='AUC', y='', x='# patients used to train')
+  labs(title='AUC', y='', x='# patients')
 

### 2020-01-14 Resegmented_validation.Rmd
last bits and pieces for paper

--- 2019-09-06 Validation Report.Rmd
@@
 ---
-title: "sWGS validation report"
+title: "Validation resegmented"
 author: "Sarah Killcoyne"
-date: "11 June 2019"
+date: "Oct 9 2019"
 output: 
@@
 val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
-analysis.dir = '~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf'
-preds.dir = paste0(analysis.dir, '/predictions/0.9')
+analysis.dir = '~/Data/BarrettsProgressionRisk/Analysis/validation/resegment/'
+
+n.sheets = c(8:14)
 
@@
 
-sheets = readxl::excel_sheets(val.file)[1:13]
+sheets = readxl::excel_sheets(val.file)[n.sheets]
 all.val = do.call(bind_rows, lapply(sheets, function(s) {
-  readxl::read_xlsx(val.file, s) %>% dplyr::select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA, matches('Collection')) %>% dplyr::filter(!is.na(`SLX-ID`)) %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% fncols('Collection', 'Biopsy')
-}))
+  readxl::read_xlsx(val.file, s) %>% 
+    dplyr::select(-matches('Notes|Qbit|Extraction')) %>% 
+    dplyr::filter(!is.na(`SLX-ID`)) %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% fncols('Collection', 'Biopsy')
+})) %>% mutate(Endoscopy = as.Date(Endoscopy))
 
@@
 
-get.png<-function(n,dir) {
-  list.files(path=dir, pattern=n, full.names=T, recursive=T)
-}
+val = do.call(bind_rows, lapply(list.files(analysis.dir, 'residuals', recursive = T, full.names = T), function(f) {
+  names = F
+  cls = cols(samplename = col_character(), varMAD_median = col_double(), varMAD_sd = col_double(), varMAD_Q1 = col_double(), varMAD_Q3 = col_double(), Pass = col_logical(), kb = col_integer())
+  cls = 'cddddl'
+  line1 = readr::read_tsv(f, col_names = names, n_max = 1)
+  if (!is.double(line1[,2])) names = T
+  if (ncol(line1) == 7) cls = 'cddddld'
 
-val = do.call(bind_rows, lapply(list.files(analysis.dir, 'residuals', recursive = T, full.names = T), function(f) {
-  readr::read_tsv(f, col_types = 'cdddddl', col_names = F) %>% set_names('sample', 'varMAD_median', 'varMAD_sd', 'varMAD_Q1', 'varMAD_Q3', 'n.segs', 'Pass')  %>% 
+  tsv = readr::read_tsv(f, col_names = names, col_types = cls) 
+  if (ncol(line1) < 7) tsv = tsv %>% mutate(X7=NA)
+
+  tsv %>% set_names(c('sample', 'varMAD_median', 'varMAD_sd', 'varMAD_Q1', 'varMAD_Q3','Pass','kb')) %>%
     mutate(`Patient ID` = as.character(basename(dirname(f))))
-})) %>% dplyr::select(`Patient ID`, sample, varMAD_median, n.segs) %>% mutate(varMAD_median = round(varMAD_median, 3)) %>% mutate(sample = str_replace(sample, '_', '-'))
+})) %>% dplyr::select(`Patient ID`, sample, varMAD_median) %>% mutate(varMAD_median = round(varMAD_median, 3)) %>% mutate(sample = str_replace(sample, '_', '-'))
 
@@
 pass.fail = tibble()
-```
-
-# Quick QC Summary
-
-Effectively every batch C.K.
-ran failed at a rate of 70-90% of samples.
-None of the tests we ran showed any effect, with the possible exception of the extraction kit.
-Frozen samples always do well, FFPE is clearly a little tricker.
-
-Even the old progressor scrolls that had been sitting for a year in tubes worked well enough to pass QC and provide predictions.
-To date those suggest a false negative rate of about 38% (by sample).
-We have more progressors extracted and will sequence these shortly.
-
-We are now requesting and extracting all of the NP cohort that we had previously identified.
-
-## By Batch
-
-```{r}
-
-pass.fail = bind_rows(pass.fail, 
-  val %>% group_by(Batch, Cohort) %>% dplyr::summarise( Pass= length(which(varMAD_median<=qc.cutoff))/nrow(val), Fail=length(which(varMAD_median>qc.cutoff))/nrow(val)))
-
-sval = val %>% group_by(Batch) %>% dplyr::mutate(Cohort = paste(unique(Cohort), collapse=', ' )) %>% group_by(Batch, Cohort) %>% 
-  dplyr::summarise_at(vars(varMAD_median), list(~ min(., na.rm=T), ~ mean(., na.rm=T) ,~ median(., na.rm=T),~ max(., na.rm=T)))
-
-vt = val %>% group_by(Batch) %>% dplyr::mutate(Cohort =  paste(unique(Cohort), collapse=', ' )) %>% group_by(Batch, `Sample Type`, Cohort) %>% dplyr::summarise(
-  `Total Samples` = length(Batch),
-  Pass=length(which(varMAD_median<=qc.cutoff)), Fail=length(which(varMAD_median>qc.cutoff))
-)
-
-vt %>% kable(caption='Pass/Fail by batch, red rows failed > 10% of samples') %>% kable_styling('striped') %>% row_spec(which(vt$Fail/vt$`Total Samples` >  0.1), color='red')
-
-sval %>% kable(caption='Summary of variance per batch.
-Red rows are where the median value of the variance for the entire batch was greater than the cutoff') %>% kable_styling('striped') %>% row_spec(which(sval$median > qc.cutoff), color='red')
-
-val %>% dplyr::group_by(Batch) %>% dplyr::summarise( Pass = length(which(varMAD_median <= qc.cutoff)), Fail = length(which(varMAD_median > qc.cutoff))  ) %>% kable(caption = 'Number of samples per batch that pass/fail') %>% kable_styling('striped', full_width = F)
-```
-
-## By Sample Type
-
-Frozen samples are typically very good, and our few Cytosponge samples pass as well.
-
-```{r}
-val$RA = factor(val$RA)
-
-val %>% group_by(as.integer(RA),`Sample Type`, Collection) %>% dplyr::summarise(
-  `Total Samples` = length(Batch),
-  Pass=length(which(varMAD_median<=qc.cutoff)), Fail=length(which(varMAD_median>qc.cutoff))
-) %>% dplyr::rename('Processing RA' = 'as.integer(RA)' ) %>% kable(caption='Pass/Fail by sample type') %>% kable_styling('striped')
-
-```
-
-# Initial Pilots
-
-The initial pilot results didn't look too bad.  But the problem was that the ones that failed were all FFPE and we only did 10.
-
-```{r}
-
-val %>% ungroup %>% filter(Batch <= 2) %>% dplyr::group_by(Batch, `Sample Type`) %>% dplyr::summarise(
-  `Total Samples` = n(),
-  `Passed QC` = length(which(varMAD_median<=qc.cutoff))
-) %>% kable() %>% kable_styling('striped')
-
-```
-
-## Initial NP Validation
-
-Here we should've done another, slightly larger pilot, before diving in.  We didn't, and more than 90% failed QC.  The question was why.
-
-```{r}
-val %>% ungroup %>% filter(Batch %in% c(3,4)) %>% dplyr::group_by(Batch, `Sample Type`, Collection) %>% dplyr::summarise(
-  Total = n(),
-  `Passed QC` = paste0(length(which(varMAD_median<=qc.cutoff)), ' (', round((length(which(varMAD_median<=qc.cutoff))/n())*100,1), '%)')
-) %>% kable(caption='Total samples and QC results') %>% kable_styling('striped')
-```
-
-## Speedvac test
-
-* 16 samples, 12 samples from the training dataset and 4 from the validation set
-* 8 Speedvac'd
-* 8 were not
-
-There was some concern initially that the Speedvac itself was causing the difference as Ellie had used the one at the CI, while all of the validation dataset were run in the MRC.
-On both manual inspection and via the computational cutoff only the 4 samples that were not in the initial training dataset failed QC.
-
-```{r}
-
-speedvac = left_join(val %>% ungroup %>% filter(Batch %in% c(5)), readxl::read_xlsx(val.file, sheet='Batch 5') %>% dplyr::select(`Hospital Research ID`, `Block ID`, `Original Cohort`, `Speedvac`))
-
-speedvac %>% dplyr::group_by(Batch, `Sample Type`, `Original Cohort`, Speedvac) %>% dplyr::summarise(
-  Total = n(),
-  `Passed QC` = paste0(length(which(varMAD_median<=qc.cutoff)), ' (', round((length(which(varMAD_median<=qc.cutoff))/n())*100,1), '%)')
-) %>% kable() %>% kable_styling('striped')
-
-```
-
-All 4 that were not from Ellie's cohort failed QC.
-The difference in these data were only that the samples used from the training dataset (n=12) were already extracted while the 4 that were new had to be extracted from scratch.
-As the training samples had been extracted using GeneRead and the new samples used AllPrep we considered that the kit could be an issue.
-
-## Kit Test
-
-*	12 cell line FFPE samples 
-  + 10 cell lines
-  + 2 NP patients from an earlier set
-* 7 AllPrep
-* 5 GeneRead
-
-```{r}
-
-kit = left_join(val %>% ungroup %>% filter(Batch %in% c(6)), 
-          readxl::read_xlsx(val.file, sheet='Batch 6') %>% dplyr::select(`Block ID`, `Kit Extraction`, `Sample`) %>% mutate_at(vars(`Block ID`), list(as.character)))
-
-kit %>% dplyr::group_by(Batch, `Kit Extraction`) %>% dplyr::summarise(
-  Total = n(),
-  `Passed QC` = paste0(length(which(varMAD_median<=qc.cutoff)), ' (', round((length(which(varMAD_median<=qc.cutoff))/n())*100,1), '%)')
-) %>% kable() %>% kable_styling('striped')
-
-```
-
-In the Generead vs Allprep kit test we were testing to see whether the specific kit used made the difference.
-Ellie had used GeneRead for all of her samples, while we had started to use Allprep.
-While it is not conclusive, the GeneRead's did perform better.
-However, even here the failure rate was too high and most of the GeneRead's passed manual QC but would not have passed computational QC.
-
-I would not call this definitive simply because there were so few and that’s still a high rate of failure.  However, GeneRead appears to have a better rate overall.
-
-## CI Protocol
-
-At this point we were guessing that the extraction method itself may be at issue, and that there could be small protocol errors that we were unaware of.
-We selected a set of samples that had been previously sequenced by Ellie and Cassandra, and then a number of new samples as well.
-
-Cassandra worked with Anna Piskorz from the Brenton lab at the CI to do the extractions using their protocol and their kits, and Anna then did the library prep as well.
-However, again a very high percentage failed.
-Most concerning is that the samples from Ellie's cohort failed and the orginal ones done in 2017 by Ellie all looked fine.
-
-```{r}
-ci = left_join(val %>% ungroup %>% filter(Batch %in% c(7)),
-          readxl::read_xlsx(val.file, sheet='Batch 7') %>% dplyr::select(`Hospital Research ID`, `Block ID`, `Original cohort`, `Sample Type`, Cohort))
-
-ci %>% dplyr::group_by(Batch, `Original cohort`) %>% dplyr::summarise(
-  Total = n(),
-  `Passed QC` = paste0(length(which(varMAD_median<=qc.cutoff)), ' (', round((length(which(varMAD_median<=qc.cutoff))/n())*100,1), '%)')
-) %>% kable() %>% kable_styling('striped')
 ```
@@
 on = left_join(val %>% ungroup %>% filter(Batch == 8),
-               readxl::read_xlsx(val.file, sheet='Batch 8') %>% dplyr::select(`Block ID`, `Endoscopy`, `Sample Type`, `Index Sequence`),
-               by=c('Block ID','Sample Type', 'Index Sequence')) %>% 
+               readxl::read_xlsx(val.file, sheet='Batch 8') %>% dplyr::select(`Block ID`, `Sample Type`, `Index Sequence`),by=c('Block ID','Sample Type', 'Index Sequence')) %>% 
   arrange(`Patient ID`, sample)
@@
 
-val.info = do.call(bind_rows, lapply(sheets[8:13], function(s) {
+val.info = do.call(bind_rows, lapply(sheets, function(s) {
   readxl::read_xlsx(val.file, s) %>% dplyr::select(-matches('Extraction|Qbit'))%>% dplyr::filter(!is.na(`SLX-ID`)) %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% fncols('Collection', 'Biopsy')
@@
 batch8_preds = do.call(bind_rows,
-  purrr::map(grep( paste(unique(on$`Patient ID`), collapse='|'), list.files(preds.dir, 'predictions.tsv', full.names = T, recursive = T), value=T), function(f) {
+  purrr::map(grep( paste(unique(on$`Patient ID`), collapse='|'), list.files(analysis.dir, 'preds.tsv', full.names = T, recursive = T), value=T), function(f) {
              read_tsv(f, col_types = cols(
@@
                   `Hospital Research ID` = col_character(),
-                `Block ID` = col_character(),
-                Endoscopy = col_date(format = ""),
-                Pathology = col_character(),
+                  Status = col_character(),
+                  `Sample Type` = col_character(),
                   `SLX-ID` = col_character(),
                   `Index Sequence` = col_character(),
-                `Path Notes` = col_character()
+                  Cohort = col_character(),
+                  Batch = col_double(),
+                  RA = col_character(),
+                  Samplename = col_character(),
+                  Endoscopy = col_date(format = "")
       ))
@@
 
-batch8_preds = left_join(batch8_preds,on %>% dplyr::select(sample, `Sample Type`, Age), by=c('Sample'='sample'))
+batch8_preds = left_join(batch8_preds,on %>% dplyr::select(sample,  Age), by=c('Sample'='sample'))
 
@@
 
-oldp = val %>% ungroup %>% filter(Batch %in% c(8:13) & Cohort == 'Old progressor scrolls')
+oldp = val %>% ungroup %>% filter(Batch %in% c(n.sheets) & grepl('Old progressor',Cohort) )
 
-oldp %>% dplyr::group_by(Batch, `Sample Type`, `Cohort`) %>% dplyr::summarise(
+oldp %>% dplyr::group_by(Batch, `Cohort`) %>% dplyr::summarise(
   `n Samples` = n(),
@@
 oldp.preds = do.call(bind_rows,
-  purrr::map(grep( paste(unique(oldp$`Patient ID`), collapse='|'), list.files(preds.dir, 'predictions.tsv', full.names = T, recursive = T), value=T), function(f) {
-             read_tsv(f,col_types = cols(
+  purrr::map(grep( paste(unique(oldp$`Patient ID`), collapse='|'), list.files(analysis.dir, 'preds.tsv', full.names = T, recursive = T), value=T), function(f) {
+             read_tsv(f, col_types = cols(
                   Sample = col_character(),
@@
 
-oldp.preds = left_join(oldp.preds, val %>% dplyr::select(sample, varMAD_median, n.segs), by=c('Sample'='sample'))
-oldp.preds = left_join(oldp.preds, val.info %>% dplyr::select(Samplename, Endoscopy), by=c('Sample'='Samplename'))
+oldp.preds = left_join(oldp.preds, val %>% dplyr::select(`Patient ID`, sample, `Block ID`, varMAD_median), by=c('Sample'='sample', 'Hospital Research ID'='Patient ID'))
+oldp.preds = left_join(oldp.preds, val.info %>% dplyr::select(`Hospital Research ID`, Samplename, Endoscopy), by=c('Sample'='Samplename', 'Hospital Research ID' ))
+
 oldp.preds = oldp.preds %>% filter(Sample %in% (oldp %>% filter(!grepl('OCCAMS',`Hospital Research ID`)) %>% dplyr::select(sample) %>% pull))
@@
 
-oldp.preds %>% group_by(`Risk`) %>% dplyr::summarise( 'n samples'=n() ) %>% spread(Risk, `n samples`) %>%
+oldp.preds %>% group_by(Cohort,`Risk`) %>% dplyr::summarise( 'n samples'=n() ) %>% spread(Risk, `n samples`) %>%
   kable(caption='Progressor sample predicted risks') %>% kable_styling(full_width = F)
@@
   'Low Risk' =   length(which(Risk == 'Low')),
-  'Pathology Notes' = paste(unique(`Path Notes`), collapse='; ')
+#  'Pathology Notes' = paste(unique(`Path Notes`), collapse='; ')
 ) %>% dplyr::arrange(desc(`High Risk`), desc(`Mod. Risk`), desc(`Low Risk`)) %>% 
@@
 
+oldp.preds %>% filter(Batch == 14) %>% dplyr::select(Risk) %>% pull
+
 oldp.preds = oldp.preds %>% separate(`Block ID`, c('PID','Block'), '\\s?(-|\\s)', remove = F) 
@@
 
-However, it looks very likely that the age of the sample is related to the predictability.
-Roughly speaking, samples over 3.5 years old appear more likely to predict as low risk.
-This wasn't so true of Ellie's samples.
-These samples did sit as parafiin scrolls for a year after being cut as well and this could have affected the samples.
-
-```{r}
+```{r, eval=T}
 ggplot(oldp.preds, aes(`Weeks Follow Up`, Probability, color=Probability)) + geom_point(size=2) +
@@
 
-fnr.patients = oldp.preds %>% 
-  group_by(`Hospital Research ID`) %>% dplyr::summarise(
-  FNR = length(which(Risk == 'Low'))/n() ) %>% 
-  dplyr::summarise( 'Misclassified patients' = length(which(FNR > 0.5))/n() ) %>% pull %>% round(2)
+oldp.preds = oldp.preds %>% mutate(Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered=T))
+
+# Max risk per endoscopy
+max.per.endo = oldp.preds %>% group_by(`Hospital Research ID`, PID) %>% 
+  dplyr::mutate('n.samples'=length(PID),'Samples'=paste(Sample,collapse=',')) %>% 
+  dplyr::summarise_at(vars(matches('Prob|Relative|Endo|n.samples|Risk')), list(max)) %>% dplyr::select(-matches('^Sample$'))
+
+max.per.endo %>% arrange(`Hospital Research ID`, PID) %>% group_by(`Hospital Research ID`, Risk) %>% tally() %>% spread(Risk,n) %>% 
+  kable(caption='Max risk per endoscopy') %>% kable_styling('striped',full_width = F)
+
+max.per.endo %>% group_by(Risk) %>% tally() %>% spread(Risk,n) %>% kable(caption='Total max risk per endoscopy') %>% kable_styling(full_width = F)
 
 ```
-
-For the old progressor samples to-date, the model failed to classify any samples as high or moderate risk for *`r fnr.patients*100`%* of patients.
-None of these were the samples I might have said were poor quality via manual QC so taken at face value these patients/samples were incorrectly predicted.
 
@@
 ```{r}
-new.np = val %>% ungroup %>% filter(Batch %in% c(10:13) & Status == 'NP')
+new.np = val %>% ungroup %>% filter(Batch %in% c(10:14) & Status == 'NP')
 
 new.np.preds = do.call(bind_rows,
-  purrr::map(grep( paste(unique(new.np$`Patient ID`), collapse='|'), list.files(preds.dir, 'predictions.tsv', full.names = T, recursive = T), value=T), function(f) {
-             read_tsv(f,col_types = cols(
+  purrr::map(grep( paste(unique(new.np$`Patient ID`), collapse='|'), list.files(analysis.dir, 'preds.tsv', full.names = T, recursive = T), value=T), function(f) {
+                         read_tsv(f, col_types = cols(
                   Sample = col_character(),
@@
                   `Hospital Research ID` = col_character(),
-                `Block ID` = col_character(),
-                Endoscopy = col_date(format = ""),
-                Pathology = col_character(),
+                  Status = col_character(),
+                  `Sample Type` = col_character(),
                   `SLX-ID` = col_character(),
                   `Index Sequence` = col_character(),
-                `Path Notes` = col_character()
+                  Cohort = col_character(),
+                  Batch = col_double(),
+                  RA = col_character(),
+                  Samplename = col_character(),
+                  Endoscopy = col_date(format = "")
       ))
-})) %>% dplyr::mutate_if(is.numeric, list(round), 2) %>% dplyr::select(-Endoscopy)
+})) %>% dplyr::mutate_if(is.numeric, list(round), 2) #%>% dplyr::select(-Endoscopy)
 
-new.np.preds = left_join(new.np.preds, val %>% dplyr::select(sample, varMAD_median, n.segs), by=c('Sample'='sample'))
-new.np.preds = left_join(new.np.preds, val.info %>% dplyr::select(Samplename, Endoscopy), by=c('Sample'='Samplename'))
+new.np.preds = left_join(new.np.preds, val %>% dplyr::select(`Hospital Research ID`, `Block ID`, sample, varMAD_median), by=c('Sample'='sample', 'Hospital Research ID'))
+new.np.preds = left_join(new.np.preds, val.info %>% dplyr::select(`Hospital Research ID`, Samplename), by=c('Hospital Research ID', 'Sample'='Samplename'))
 
@@
 
-val.new.np.preds = left_join(val %>% filter(sample %in% new.np.preds$Sample) %>% dplyr::select(`Hospital Research ID`, sample, varMAD_median, n.segs, Status),
+new.np.preds %>% group_by(`Risk`) %>% dplyr::summarise( 'n samples'=n() ) %>% spread(Risk, `n samples`) %>% mutate_all(funs(./nrow(new.np.preds)))
+
+val.new.np.preds = left_join(val %>% filter(sample %in% new.np.preds$Sample) %>% dplyr::select(`Hospital Research ID`, sample, varMAD_median, Status),
           new.np.preds %>% dplyr::select(Sample, Probability, matches('Risk')), by=c('sample'='Sample'))
 
-new.np.preds = new.np.preds %>% dplyr::mutate( 'Weeks Since Endoscopy' = as.double(difftime( as.Date('2019/06/01'), Endoscopy, units='weeks' )))
+#new.np.preds = new.np.preds %>% dplyr::mutate( 'Weeks Since Endoscopy' = as.double(difftime( as.Date('2019/06/01'), Endoscopy, units='weeks' )))
 
@@
 
-ggplot(new.np.preds, aes(`Hospital Research ID`, `Relative Risk`, color=Probability)) + geom_point(size=2) +
-  scale_color_distiller(palette = 'RdYlBu', name='RR') + plot.theme + theme(axis.text.x = element_text(angle=45,hjust=1))
+# ggplot(new.np.preds, aes(`Hospital Research ID`, `Relative Risk`, color=Probability)) + geom_point(size=2) +
+#   scale_color_distiller(palette = 'RdYlBu', name='RR') + plot.theme + theme(axis.text.x = element_text(angle=45,hjust=1))
 
@@
 
-Time since the endoscopy was performed doesn't appear to have any effect on the non-progressor predictions
-
 ```{r}
-ggplot(new.np.preds, aes(`Weeks Since Endoscopy`, Probability, color=Probability)) + geom_point(size=2) +
-  scale_color_distiller(palette = 'RdYlBu', name='P(P)') + plot.theme + labs(title='Non-progressors', subtitle='Predictions vs time since the endoscopy')
+# ggplot(new.np.preds, aes(`Weeks Since Endoscopy`, Probability, color=Probability)) + geom_point(size=2) +
+#   scale_color_distiller(palette = 'RdYlBu', name='P(P)') + plot.theme + labs(title='Non-progressors', subtitle='Predictions vs time since the endoscopy')
 
@@
 
+new.np.preds = new.np.preds %>% mutate(Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered=T))
+
+# max prediction
 np.per.endo = 
@@
   dplyr::mutate('n.samples'=length(PID),'Samples'=paste(Sample,collapse=',')) %>% 
-  dplyr::summarise_at(vars(matches('Prob|Relative|Endo|n.samples')), list(max)) %>% dplyr::select(-matches('^Sample$')) %>% 
-  arrange(`Hospital Research ID`, Probability, PID)
+  dplyr::summarise_at(vars(matches('Prob|Relative|Endo|n.samples|Risk')), list(max)) %>% dplyr::select(-matches('^Sample$')) %>% 
+  arrange(desc(Probability), `Hospital Research ID`, PID)
 
-fpr.patients = new.np.preds %>% 
-  group_by(`Hospital Research ID`) %>% dplyr::summarise(
-  FPR = length(which(Risk == 'High'))/n() ) %>% 
+# Using the more conservative any risk other than Low
+fpr.per.patient = np.per.endo %>% group_by(`Hospital Research ID`) %>% 
+  #dplyr::summarise( FPR = length(which(Risk == 'High'))/n() ) %>% 
+  dplyr::summarise( FPR = length(which(Risk != 'Low'))/n() ) %>% 
   dplyr::summarise( 'Misclassified patients' = length(which(FPR > 0.5))/n() ) %>% pull %>% round(2)
+
+fpr.per.sample = new.np.preds %>% group_by(Risk) %>% tally %>% filter(Risk != 'Low') %>% summarise(n = sum(n)/nrow(new.np.preds))
+
+new.np.preds %>% group_by(Risk) %>% tally() %>% mutate(n = round(n/nrow(new.np.preds),2)) %>% spread(Risk,n) %>% kable(caption='Ratio of per-sample NP risks') %>% kable_styling(full_width=F)
 
@@
 
-```{r}
+```{r, eval=T}
 
@@
 
-```{r, fig.height=8, eval=F}
+```{r, fig.height=8, eval=T}
 

### 2020-01-14 snp/SNP_Prediction_Analysis.Rmd
last bits and pieces for paper

@@
 roc.plot(rocpp) + labs(title='BE biopsies', subtitle='excluding low SCA samples')
-ggsave(filename='plots/snp/auc_ex_lowsca_roc.png', plot = roc.plot(rocpp) + labs(title='BE biopsies', subtitle='excluding low SCA samples') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
+ggsave(filename='plots/snp/auc_ex_lowsca_roc.png', plot = roc.plot(rocpp) + labs(title='BE biopsies', subtitle='excluding low SCA samples') + theme(text=element_text(size=12)), width=5, height=5, units='in', dpi=300)
 
@@
 roc.plot(roct1) + labs(title='BE biopsies', subtitle='Timepoint T1')
-ggsave(filename='plots/snp/auc_t1_roc.png', plot=roc.plot(roct1) + labs(title='BE biopsies', subtitle='Timepoint T1') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
+ggsave(filename='plots/snp/auc_t1_roc.png', plot=roc.plot(roct1) + labs(title='BE biopsies', subtitle='Timepoint T1') + theme(text=element_text(size=12)), width=5, height=5, units='in', dpi=300)
 
@@
 roc.plot(roct2) + labs(title='BE biopsies', subtitle='Timepoint T2')
-ggsave(filename='plots/snp/auc_t2_roc.png', plot=roc.plot(roct2) + labs(title='BE biopsies', subtitle='Timepoint T2') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
+ggsave(filename='plots/snp/auc_t2_roc.png', plot=roc.plot(roct2) + labs(title='BE biopsies', subtitle='Timepoint T2') + theme(text=element_text(size=12)), width=5, height=5, units='in', dpi=300)
 

### 2020-01-21 Paper.Rmd
replacing the original Rmd file to avoid (my) confusion

@@
 author: "Sarah Killcoyne"
-date: "2 October 2017"
+date: "2019-11-13"
 output: 
@@
     toc_depth: 4
-
+editor_options: 
+  chunk_output_type: console
 ---
@@
 ```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = F, message = F, warning = F)
 library(tidyverse)
 library(BarrettsProgressionRisk)
-
 library(ggrepel)
-#library(GGally)
 library(kableExtra)
@@
 library(gridExtra)
-#library(CoxHD)
 library(RColorBrewer)
@@
 all.patient.info = read.patient.info(patient.file, demo.file, set='All')$info
-
-patient.info = all.patient.info %>% dplyr::filter(Set == 'Training')
+patient.info = all.patient.info
 
@@
 
-```
-
-# Segment size: 5e06 with arms
-
-## Quick look at our demographic information
-
-```{r echo=F}
+patient.info = patient.info %>% mutate( Pathology = recode(Pathology, 'BE'='NDBE'))
+
+```
+
+# Demographics
+
+```{r}
 
@@
 
-```{r echo=F, warning=F, message=F, eval=T, fig.height=12, fig.width=20}
-
+```{r echo=F, warning=F, message=F, eval=T, fig.height=6, fig.width=10}
 chr.lengths = BarrettsProgressionRisk:::chrInfo() %>% dplyr::filter(chr %in% c(1:22)) %>%
@@
 ## NDBE
-nb.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology == 'BE')$variable)), ')', sep=''),
-              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology == 'BE')$variable)),')', sep=''))
-globalBECN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology == 'BE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')
-
-gp = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples')# + theme(text=element_text(size=12))
-ggsave(filename='plots/globalCN.png', plot=gp, width=16, height=4.6, units='in', dpi=300)
-
-gbe = globalBECN.plot + theme(legend.position = 'none') + labs(y='mean adj.seg.', title="Non-Dysplastic Barrett's Samples") 
-ggsave(filename='plots/globalBECN.png', plot=gbe, width=16, height=4.6, units='in', dpi=300)
-
-```
+nb.status <- c('Progressor'=paste0('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology == 'NDBE')$variable)), ')'),
+              'Non-Progressor'=paste0('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology == 'NDBE')$variable)),')'))
+
+globalBECN.plot = cn.mtn.plot(filter(globalCNdf, seg.type != 'arms' & Pathology == 'NDBE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')
+
+gp = globalCN.plot +  labs(y='mean adj.
+seg.', title='All samples') + theme(legend.position = 'none', axis.text = element_blank(), strip.text.y = element_blank(), strip.text.x = element_text(size=8)) +
+  labs(y='',x='', title='All samples')
+ggsave(filename='plots/globalCN.png', plot=gp, width=8, height=2.3, units='in', dpi=300)
+
+gbe = globalBECN.plot +theme(legend.position = 'none', axis.text = element_blank(), strip.text.y = element_blank(), strip.text.x = element_text(size=8)) + 
+  labs(y='',x='', title="Non-Dysplastic Barrett's Samples") 
+ggsave(filename='plots/globalBECN.png', plot=gbe, width=8, height=2.3, units='in', dpi=300)
+
+```
+
+## *Complexity Scores*
 
 ```{r loaddata, message=F, warning=F, echo=F}
-
 file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
@@
 cx = dysplasia.df[,'cx']
-
-x = cbind.data.frame('Status'=labels, 'cx'=cx[names(labels)])
-x$Status = factor(x$Status, labels = c('Non-Progressor','Progressor'))
-x = merge(x,patient.info[,c('Samplename','Pathology','Patient')],by.x='row.names',by.y='Samplename')
-
-tt = t.test(subset(x, Status == 'Progressor')$cx, subset(x, Status == 'Non-Progressor')$cx)
-pv = ggplot(x, aes(Status, cx, fill=Status)) + 
+complexity.score = tibble('Status'=labels, 'cx'=cx[names(labels)], 'Samplename' = names(labels)) %>%
+  mutate(Status = factor(Status, labels = c('Non-Progressor','Progressor'))) %>%
+  left_join(patient.info[,c('Samplename','Pathology','Patient')],by=c('Samplename'))
+
+tt = wilcox.test(subset(complexity.score, Status == 'Progressor')$cx, subset(complexity.score, Status == 'Non-Progressor')$cx)
+pv = ggplot(complexity.score, aes(Status, cx, fill=Status)) + 
   geom_jitter(width=0.2, color='grey39') + 
@@
 
-rocX = pROC::roc(Status ~ cx, data=x, auc=T, ci=T, of='thresholds')
+rocX = pROC::roc(Status ~ cx, data=complexity.score, auc=T, ci=T, of='thresholds')
 roc.plot(rocX, title='Per sample cx') + plot.theme
@@
 
+rocX = pROC::roc(Status ~ cx, data=complexity.score, auc=T, ci=T, of='thresholds')
+roc.plot(rocX, title='Per sample cx') + plot.theme
+
+ggsave('plots/per_sample_cx_roc.png', plot=roc.plot(rocX, title='Per sample cx') + plot.theme + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
 ```
@@
 featureStability = rowSums(coefs[[select.alpha]][,-1])/(splits*folds)
-
-hazards = apply( exp(t(dysplasia.df[, rownames(coefs[[select.alpha]])]) *  coefs[[select.alpha]][,1]), 1, function(x) {
+hazards = apply( exp( t(dysplasia.df[,rownames(coefs[[select.alpha]])]) * coefs[[select.alpha]][,1]), 1, function(x) {
   sd(x)/mean(x)
@@
 
-features = as_tibble(hazards, rownames = 'label') %>% dplyr::rename(hazards = value) %>% mutate(coef = coefs[[select.alpha]][,1], stability = featureStability) %>% arrange(-hazards)
+features = as_tibble(hazards, rownames = 'label') %>% dplyr::rename(hazards = value) %>% 
+  dplyr::mutate(coef = coefs[[select.alpha]][,1], stability = featureStability) %>% arrange(-hazards)
+
+cvRR = features
 
@@
   scale_color_manual(values=c('firebrick3','darkblue')) +
-  labs(title='Top features by hazard ratio', x='Coefficient value', y='Hazard') + plot.theme 
+  labs(title='Top features by cv(RR)', x='Coefficient value', y='cvRR') + plot.theme 
 p
-
 ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)
 
+cvRR %>% mutate(`Gain/Loss` = ifelse(coef > 0 ,'Gain', 'Loss')) %>% 
+  dplyr::select(label, hazards, `Gain/Loss`) %>% arrange(desc(hazards)) %>%
+  write_tsv('plots/coef_effect.tsv')
+
 ```
@@
 ```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
-features = features %>% separate(label, c('chr','start','end'), sep=':|-', remove=F) %>% 
-  mutate_at(vars(start, end), list(as.numeric)) %>% mutate(chr = factor(chr,levels=c(1:22)))
-
-features = features %>% mutate(ymax = ifelse(coef > 0, max(globalCN.plot$data$mean.value), min(globalCN.plot$data$mean.value)), 
-                    gl = factor(ifelse(coef > 0, 'gain','loss')))
-
-features = left_join(features, chr.lengths %>% dplyr::select(chr, chr.length), by=c('chr')) %>% mutate(chr = factor(chr, levels=c(1:22), ordered=T))
+features = features %>% filter(label != 'cx') %>% separate(label, c('chr','start','end'), sep=':|-', remove=F) %>% 
+  dplyr::mutate_at(vars(start, end), list(as.numeric)) %>% 
+  mutate(chr = factor(chr,levels=c(1:22))) %>% 
+  dplyr::mutate(ymax = ifelse(coef > 0, max(globalCN.plot$data$mean.value), min(globalCN.plot$data$mean.value)), gl = factor(ifelse(coef > 0, 'gain','loss'))) %>%
+  left_join(chr.lengths %>% dplyr::select(chr, chr.length), by=c('chr')) %>% mutate(chr = factor(chr, levels=c(1:22), ordered=T))
 
@@
 ```{r leaveoneout, echo=F, message=F, warning=F, fig.width=5, fig.height=5}
-
-file = paste(cache.dir, 'loo.Rdata', sep='/')
+# loo_residuals = do.call(bind_rows, purrr::map(list.files(paste0(dir, '/Analysis/pcf_perPatient/50kb'), 'residuals.tsv', recursive = T, full.names = T), function(f) {
+#   read_tsv(f, col_types = 'ccddddl')
+# }))
+# 
+# failedQC = loo_residuals %>% filter(varMAD_median > 0.008)
+
+error = read_tsv('~/Data/BarrettsProgressionRisk/Analysis/pcf_perPatient/50kb/sample_error.tsv')
+
+file = paste(cache.dir, 'loo_0.9.Rdata', sep='/')
 if (file.exists(file)) {
-  load(file, verbose=T)
+  load(file, verbose=F)
 } else {
@@
 loo.coefs = nzcoefs
+
+error = error %>% left_join(pg.samp %>% dplyr::select(Status, Samplename, RR), by=c('Sample'='Samplename')) %>% 
+  dplyr::mutate(P.High = BarrettsProgressionRisk:::pi.hat(RR+Error), P.Low = BarrettsProgressionRisk:::pi.hat(RR-Error), Abs.Error = P.High-P.Low )
+
+error %>% group_by(Status) %>% dplyr::summarise_at(vars(Abs.Error), list(~mean, ~sd))
+summary(error$Abs.Error)
+sd(error$Abs.Error)
+
+pg.samp = pg.samp %>% mutate(Pathology = recode(Pathology, 'BE' = 'NDBE'))
+
+pg.samp = pg.samp %>% left_join(error %>% dplyr::select(-RR,-Status), by=c('Samplename'='Sample'))
 
@@
 
-Based on these curves we get our best AUC at ~0.36.  Currently we are using 0.5 by default.
+#### LOO minus HGD
+
+```{r loonohgd, echo=F, message=F, warning=T}
+file = paste(cache.dir, 'loonohgd.Rdata', sep='/')
+if (file.exists(file)) {
+  message(paste("loading file", file))
+  load(file, verbose=T)
+} else {
+  stop(paste("No file", file, "found"))
+}
+
+perf = cbind.data.frame(perf=performance.at.1se, pt=unique(pg.sampNOHGD$Patient))
+
+p = ggplot(melt(perf, id.vars='pt'), aes(x=variable,y=value, group=variable, fill=variable)) +
+    geom_boxplot(fill='lightcoral', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) +
+    geom_jitter(color='grey39', width=0.3) + plot.theme + 
+    geom_label(data=melt(subset(perf, perf %in% range(perf) ), id.vars='pt'), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
+    theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')
+ggsave('plots/performance/loo_perf_nohgd.png', plot=p, width=5, height=5, units='in', dpi=300)
+
+pg.sampNOHGD = pg.sampNOHGD %>% filter(!is.na(Prediction))
+
+rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')
+
+```
 
 ```{r insilico_rocs, echo=F, message=F, warning=F, fig.height=8, fig.width=8}
-
-load(file=paste(cache.dir, 'insilico_splits.Rdata', sep='/'), verbose=T)
-
+load(file=paste(cache.dir, 'insilico_splits.Rdata', sep='/'), verbose=F)
 ```
@@
 preds = pg.samp %>% dplyr::select(Status, Prediction)
-roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds',transpose=T)
+roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
+
+pROC::ci.auc(roc$auc)
+
+ggsave(filename='plots/auc/training_auc.png', plot=roc.plot(roc), width=6.5, height=4.5, units='in', dpi=300)
+ggsave(filename='plots/auc/training_auc_supp.png', plot=roc.plot(roc) + ggtitle('Per sample'), width=6, height=6, units='in', dpi=300)
+
 roc$model = 'CN Model'
@@
 insilico.rocs[['discovery']] = roc
-
 df = do.call(rbind, lapply(insilico.rocs, function(r) 
-    cbind.data.frame('Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'model'=r$model, 'is'=as.character(r$is) 
+    cbind.data.frame('AUC' = as.numeric(r$auc), 'Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'model'=r$model, 'is'=as.character(r$is) 
 )))
 
-p = ggplot(df, aes(Specificity, Sensitivity,color=is)) + geom_line(size=1, color='lightblue') + 
-    geom_line(data=subset(df, is == 'discovery'), color='red', size=1) +
+i.thresholds = do.call(bind_rows, purrr::map(insilico.rocs, function(r) pROC::coords(r, 'best') %>% as_tibble %>% mutate(AUC = r$auc, model=r$model, is=as.character(r$is)))) %>%
+  rename_at(vars(matches('thresh|spec|sens')), funs(str_to_title))
+
+#summary(filter(df, model == 'in-silico')$AUC)
+
+getPalette = colorRampPalette(brewer.pal(5, "Blues"))
+p = ggplot(df, aes(Specificity, Sensitivity,color=is)) + 
+  geom_line(linetype='solid', size=0.5) + 
+  geom_line(data=subset(df, is == 'discovery'), color='red', size=1.5) +
   scale_x_reverse() +
-    #scale_color_brewer(palette='Blues') +
+  scale_color_manual(values=getPalette(length(unique(df$is)))) +
   labs(title='ROC, all splits', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
@@
 p
-ggsave(filename='plots/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)
+
+blues = RColorBrewer::brewer.pal(5,'Blues')
+p = ggplot(df, aes(Specificity, Sensitivity)) + 
+    geom_line(color=blues[2]) + 
+    geom_line(data=filter(df, is == 'discovery'), color=blues[5], size=1) +
+    geom_line(data=df %>% filter( (AUC == max(AUC) ) & is != 'discovery'), color=blues[4], linetype = 'dotted', size=0.75) +
+    geom_line(data=df %>% filter( (AUC == min(AUC) ) & is != 'discovery'), color=blues[4], linetype = 'dotted', size=0.75) +
+    scale_x_reverse() +
+    geom_label(data=filter(i.thresholds, AUC %in% range(AUC) | is == 'discovery'), aes(label=paste0('AUC\n',round(AUC,2))), color='black') +
+    #scale_color_brewer(palette='Blues') +
+    labs(title='ROC AUC', subtitle='in-silico splits', x='Specificity (1-FPR)', y='Sensitivity (TPR)')  + plot.theme +
+    theme(legend.position = 'none', text = element_text(size=14))
+p
+ggsave(filename='plots/auc/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)
 
@@
   rbind.data.frame(
-    est=t(round(pROC::coords(roc, "best", transpose=T),2)),
-    max=t(round(pROC::coords(roc, 0.8, transpose=T),2)),
-    'No HGD'=t(round(pROC::coords(rocNOHGD, 0.6, transpose=T),2)),
-    'All samples'=t(round(pROC::coords(roc, 0.6, transpose=T),2))
+    est=t(round(pROC::coords(roc, "best"),2)),
+    max=t(round(pROC::coords(roc, 0.8),2)),
+    'No HGD'=t(round(pROC::coords(rocNOHGD, 0.6),2)),
+    'All samples'=t(round(pROC::coords(roc, 0.6),2))
 )
@@
 
-# roc.plot(rocNOHGD, 'Model excluding HGD/IMC')
-# pROC::coords(rocNOHGD, "best")
-# pROC::coords(rocNOHGD, 0.6)
-# pROC::auc(rocNOHGD, partial.auc=c(0.89,0.9), partial.auc.focus='spec', partial.auc.correct=T)
-
-cutoff = round(pROC::coords(roc, 'best', ret = 'threshold', transpose=T), 2)
+cutoff = round(pROC::coords(roc, 'best', ret = 'threshold'), 2)
 ```
@@
 
+pROC::ci.auc(p53roc$auc)
+
 p53roc$model = 'p53 IHC'
 roc.plot(p53roc, 'p53 IHC')
+ggsave('plots/auc/p53_roc.png', plot=roc.plot(p53roc, 'p53 IHC'), width=6, height=6, units="in", limitsize=F, dpi = 300)
 ```
@@
 
-endo.thresholds = bind_cols('roc'=c('Max','Avg','Last','First'),
+pendo.rand = pg.samp %>% dplyr::group_by(Patient, Status, PID) %>%
+  dplyr::summarise( r.Prob = sample(Prediction, 1), n = length(PID) ) %>% filter(n > 1)
+rocRandEndo = pROC::roc(Status ~ r.Prob, data=pendo.rand, ci=T, of='thresholds')
+rocRandEndo$model = 'Per Endoscopy Random'
+ggsave('plots/auc/randomendo_auc.png', plot=roc.plot(rocRandEndo, title='Random per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+
+endo.thresholds = bind_cols('roc'=c('Max','Avg','Last','First','Random'),
   bind_rows(
-    pROC::coords(rocEndoM, 'best', transpose=T),
-    pROC::coords(rocEndoA, 'best', transpose=T),
-    pROC::coords(rocLastEndo, 'best', transpose=T),
-    pROC::coords(rocFirstEndo, 'best', transpose=T)
+    pROC::coords(rocEndoM, 'best'),
+    pROC::coords(rocEndoA, 'best'),
+    pROC::coords(rocLastEndo, 'best'),
+    pROC::coords(rocFirstEndo, 'best'),
+    pROC::coords(rocRandEndo, 'best')
   ))
-
 ```
@@
   geom_text(aes(y=(value*100)-4, label=paste(round(value,2)*100,'%',sep='')), position=dodge, color='white')  +
+  scale_y_continuous(limits=c(0,100)) +
   scale_fill_manual(values=cols, name='') + plot.theme + 
-  theme(legend.position = 'none', axis.text.x = element_text(angle=45, hjust=1)) + labs(x='', y='AUC', title='ROC Comparison') 
+  theme(legend.position = 'none', text = element_text(size=14), axis.text.x = element_text(angle=45, hjust=1)) + 
+  labs(x='', y='AUC', title='ROC Comparison') 
 comp
 
-ggsave('plots/path_p53_cn_auc.png', plot=comp, width=6, height=8, units="in", limitsize=F, dpi = 300)
-
-```
-
-```{r echo=F, fig.height=8, fig.width=6}
+ggsave('plots/auc/path_p53_cn_auc.png', plot=comp, width=3, height=4, units="in", limitsize=F, dpi = 300)
+
+```
+
+```{r}
 roc$model = "Per Sample"
@@
 
-ggsave('plots/auc/multi_auc.png', plot=p, height=6, width=5, units='in',dpi=300)
-
-#multi.roc.plot(list(roc, rocEndoM, rocEndoA, rocPtM, rocPtA,rocFirstEndo,rocLastEndo), 'ROC Comparisons', palette='Dark2')
+ggsave('plots/auc/multi_auc.png', plot=p, height=6, width=8, units='in',dpi=300)
 ```
@@
 
-```{r back, echo=F, message=F, warning=F}
-transform.by.time<-function(pt, info) {
-  pt = pt %>% dplyr::group_by(PID, Samplename) %>% dplyr::mutate( 'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12) %>% arrange(Pathology, PID, Endoscopy.Year) %>% ungroup
-
-  tb = with(pt, table(Endoscopy.Year, PID))
-  tb[tb>0] = 1
-  for (yr in names(which(rowSums(tb) > 1))) {
-      tmp = pt[which(pt$Endoscopy.Year == yr),]
-      tmp = tmp %>% dplyr::group_by(PID) %>% summarise(max(Pathology))
-
-      match = which(pt$Endoscopy.Year == yr & pt$PID == as.character(tmp[which.min(tmp$`max(Pathology)`), 'PID']))
-      pt[match,'months.before.final'] = pt[match,'months.before.final'] + 6
-  }
-  pt$nl = (pt$Block-1)*2
-
-  tb = table(pt[,c('months.before.final', 'PID')])
-  for (i in 1:nrow(tb)) {
-    pids = tb[i,]
-    yrs = which(pids > 0)
-    if (length(yrs) > 1) {
-    for (j in 2:length(yrs)) 
-      pt[which(pt$PID == names(yrs)[j]), 'nl'] = pt[which(pt$PID == names(yrs)[j]), 'nl'] + 1
-    }
-  }
-  arrange(pt, -Endoscopy.Year, PID)
-}
-
-back = do.call(bind_rows,lapply(unique(pg.samp$Patient) %>% map( function(p) pg.samp %>% filter(Patient == p)  ), transform.by.time, info=patient.info)) %>% mutate(Pathology = factor(Pathology, levels=c('BE','ID','LGD','HGD','IMC'), ordered = T))
-
-sq = c(0, seq(1,132,12),max(back$months.before.final))
-
-back = back %>% mutate(
+```{r}
+pg.samp = do.call(bind_rows,lapply(unique(pg.samp$Patient) %>% 
+                                     map( function(p) pg.samp %>% filter(Patient == p)), transform.by.time, info=patient.info)) 
+sq = c(0, seq(1,132,12),max(pg.samp$months.before.final))
+
+pg.samp = pg.samp %>% mutate(
   brks = cut(months.before.final, breaks=sq, 
                  labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-c(length(sq)-1,length(sq))], '>120'),include.lowest = T, ordered_result = T),
-) %>% mutate(brks = factor(brks, rev(levels(brks))))
-
-ggplot(back, aes(Status, Prediction, group=Status)) + 
+) %>% dplyr::mutate(brks = factor(brks, rev(levels(brks))))
+
+ggplot(pg.samp, aes(Status, Prediction, group=Status)) + 
+  facet_grid(~Pathology) + 
+  #scale_fill_manual(values=RColorBrewer::brewer.pal(4,"Paired")[c(1,3)]) + 
+  scale_color_gradientn(colours = myPal) +
+  geom_boxplot(outlier.shape = NA, fill='#BDBDBD') + 
+  geom_jitter(width=0.2, aes(color=Prediction)) + 
+  plot.theme  + theme(legend.position = 'none') + labs(y='Probability')
+
+ggplot(pg.samp, aes(Status, RR, group=Status)) + 
   facet_grid(~Pathology) + 
@@
 
-back.per.endo = back %>% group_by(Hospital.Research.ID, Status, Patient, PID) %>% 
+mini = pcal + labs(x='P(P)', y='P:NP', title='') + theme(legend.position = 'none') + 
+  scale_x_continuous(expand = c(0, 0), limits = c(-0.5, 1.5), breaks = cuts, labels = c(0,'','',0.3,'',0.5,'','','','',1)) + 
+  scale_y_continuous(expand = c(0, 0), limits = c(-0.5, 1.5), breaks = seq(0,1,0.25), labels = seq(0,1,0.25)) + 
+  theme(axis.text = element_text(size=14))
+ggsave(filename='plots/pred_calib_mini.png', plot=mini, width=4, height=4, units='in', dpi=300)
+
+pred.confidence %>% dplyr::mutate_if(is.double, list(signif), 2) %>% kable() %>% kable_styling('striped')
+
+per.endo = pg.samp %>% group_by(Hospital.Research.ID, Status, Patient, PID) %>% 
   dplyr::mutate('n.samples'=length(PID),'Samples'=paste(Samplename,collapse=',')) %>% 
@@
 ```{r relrisk, echo=F, fig.width=6, fig.height=6}
-unadjRR = ggplot(back, aes(RR)) + geom_histogram(aes(fill=..x..), bins=10, show.legend = F) +
+unadjRR = ggplot(pg.samp, aes(RR)) + geom_histogram(aes(fill=..x..), bins=10, show.legend = F) +
   scale_fill_gradientn(colors = myPal,  name='') + 
-  labs(y='n Samples', x='Relative Risk', title='Unadjusted relative risk') + plot.theme
+  labs(y='n Samples', x='log(Relative Risk)', title='Unadjusted relative risk') + plot.theme
 unadjRR
-
 ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR, width=6, height=5, units='in', dpi=300)
 
-predhist = ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) +
+predhist = ggplot(pg.samp, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) +
     scale_fill_gradientn(colors = myPal,  name='') + 
     plot.theme + labs(title='Predictions, all samples model', y='n Samples', x='Probability') 
-predhist
-
 ggsave(filename='plots/pred_hist.png', plot=predhist, width=5, height=5, units='in', dpi=300)
@@
 
-mtn.spatial<-function(df, b, limits, title='', pal=NULL) {
-  if (is.null(pal))
-    pal = RColorBrewer::brewer.pal(length(limits), 'Set2')
-
-  df = merge(df, b[,c('Samplename','Endoscopy.Year','months.before.final','Block','ogj')], by.x='variable',by.y='Samplename')
-  df$ogj = factor(df$ogj, ordered = T)
-  df$ogj = factor(df$ogj, levels=rev(levels(df$ogj)), ordered = T)
-  #pal = pal[which(limits %in% levels(df$nl))]
-
-  altcolor = c(RColorBrewer::brewer.pal(9, 'Greys')[2],RColorBrewer::brewer.pal(9, 'Greys')[4])
-  fills = c()
-  for (yr in sort(unique(df$Endoscopy.Year))) {
-    ac = (which(sort(unique(df$Endoscopy.Year)) == yr) %% 2) + 1
-    fills = c(fills, rep(altcolor[ac], length(which(table(subset(df, Endoscopy.Year == yr)$ogj) > 0))))
-  }
-
-  blanklabel <- function(s) {
-    return('')
-  }
-
-  p = ggplot(df, aes(x=chr.length)) + facet_grid(Endoscopy.Year+ogj~chr, scales='free', space='free_x', labeller = labeller(.multi_line = F)) +
-      geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=ogj), show.legend=T) + 
-      scale_fill_manual(values=pal, limits=limits,  name='OGJ Dist.') +
-      labs(title=title, x='') + min.theme
-
-  legend = get.legend(p)
-  p = p + theme(legend.position = 'none')
-
-  g = ggplot_gtable(ggplot_build(p))
-  strips = grep('strip-r', g$layout$name)
-
-  for (n in 1:length(strips)) {
-    i = strips[n]
-    j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
-    g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[n]
-  }
-
-  #grid.arrange(g)
-  return(list('grob'=g, 'legend'=legend))
-  #grid::grid.draw(g)  
-  #return(g)
-}
-
-back$ogj = back$nl
-back$ogj[which(back$nl > 0 & back$nl < 3)] = 1 
-back$ogj[which(back$nl > 3 & back$nl < 7)] = 2 
-back$ogj[which(back$nl >= 7)] = 3
-
-ogjLimits = levels(factor(back$ogj, ordered = T))
-
-df = subset(globalCNdf, Patient == 20 & seg.type != 'arms')
-#df = subset(globalCNdf, Patient == 51 & seg.type != 'arms')
+setOgj <- function(nl) {
+  if (nl >= 0 & nl < 2) return(1)
+  if (nl >= 2 & nl < 4) return(2)
+  if (nl >= 4 & nl < 7) return(3)
+  if (nl >= 7) return(4)
+}
+pg.samp = pg.samp %>% mutate(ogj = factor(unlist(purrr::map(nl, setOgj)), ordered=T))
+
+#df = globalCNdf %>% filter(Patient == 20 & seg.type != 'arms')
+#df = globalCNdf %>% filter(Patient == 51 & seg.type != 'arms')
 
@@
 
-#ggplot(el2, aes(Status,`diff(arms)`)) + geom_boxplot() + geom_jitter(width=0.2)
-#ggplot(el2, aes(Status,`diff(CNA)`)) + geom_boxplot() + geom_jitter(width=0.2)
-
-p2 = mtn.spatial(subset(globalCNdf, Patient == 20 & seg.type != 'arms'), back, ogjLimits,'Patient 20 (P)' )
-p6 = mtn.spatial(subset(globalCNdf, Patient == 51 & seg.type != 'arms'), back,ogjLimits,'Patient 51 (NP)' )
-
-for (i in 1:nrow(sum.patient.data)) {
-  print(i)
-  pt = sum.patient.data[i, 'Patient']
-  status = sum.patient.data[i, 'Status']
-
-  pms = mtn.spatial(subset(globalCNdf, Patient == pt & seg.type != 'arms'), back, ogjLimits, paste('Patient', pt, '(', status, ')' ))
-
-  fl = paste('patient_', pt, '.png', sep='')
-  ggsave(paste('plots/spatial/', fl, sep=''), plot=grid.arrange(pms$grob), width=14, height=6, units='in', dpi=300)
-}
-
-back$ogj = NULL
-
-```
-
-```{r, echo=F, fig.width=2, fig.height=4}
-
-grid.arrange(p2$legend)
-
-ggsave(filename='plots/spatial_mtn_pt20.png', plot=grid.arrange(p2$grob), width=14, height=6, units='in', dpi=300)
-ggsave(filename='plots/spatial_mtn_pt51.png', plot=grid.arrange(p6$grob), width=14, height=6, units='in', dpi=300)
-
-```
-
-At each quintile are the percentage of progressor samples that were predicted within that range, and the hypergeometric p-value (Fishers) for the significance within that quintile.
-Then our predictions are at best a coin flip in the middle, but highly significant at the outer edges.
-This would also push the prediction for "high" risk up to `r min(subset(pred.confidence, labels == 'High' )$r1)` while "low" risk is below `r max(subset(pred.confidence, labels == 'Low' )$r2)`.
-
-`r pred.confidence %>% mutate_if(is.double, list(signif), 2) %>% kable() %>% kable_styling('striped')`
-
-## Jacknife coefficients
-
-The idea here is to get some sort of CI for the coefficients, I'm not entirely sure this is correct though.
-
-Could just as easily get a jacknife CI per patient
-
-```{r jacknife, echo=T, warning=F, message=F, fig.height=10, fig.width=12}
-
-load(paste0(cache.dir, '/all.pt.alpha.Rdata'), verbose=T)
-rm(plots, performance.at.1se, dysplasia.df, models, cvs, labels)
-
-coefs.per.pt = data.frame(matrix(ncol=length(unique(pg.samp$Patient)), nrow=nrow(features), dimnames=list( features$label, unique(pg.samp$Hospital.Research.ID))))
-
-for (pt in unique(pg.samp$Hospital.Research.ID)) {
-  lcf = loo.coefs[[pt]][,1]
-  names(lcf) = rownames( loo.coefs[[pt]] )
-  sharedcfs = intersect(names(lcf), features$label)
-  coefs.per.pt[sharedcfs, pt] = lcf[sharedcfs]
-}
-coefs.per.pt[is.na(coefs.per.pt)] = 0
-
-nzcoefs = purrr::map(nzcoefs, as_tibble, rownames='coef')
-
-cfs = unique(unlist(sapply(nzcoefs, function(x) x[['coef']])))
-loo.coefs = data.frame(matrix(ncol=0,nrow=length(cfs)))
-loo.coefs$coef = cfs
-for (pt in names(nzcoefs)) {
-  loo.coefs = dplyr::full_join(loo.coefs, nzcoefs[[pt]], by='coef')
-}
-colnames(loo.coefs)[-1] = names(nzcoefs)
-loo.coefs[is.na(loo.coefs)] = 0
-
-features = as_tibble(coefs$`0.9`[,1,drop=F], rownames='coef')
-
-loo.features = dplyr::left_join(features, loo.coefs, by='coef')
-
-library(bootstrap)
-
-features$jack.se = apply(loo.features, 1, function(x) {
-  jk = bootstrap::jackknife(x[-1], var)
-  jk$jack.se
-})
-
-# per patient
-# jack.se = apply(coefs.per.pt, 2, function(x) {
-#   jk = jackknife(x, var)
-#   jk$jack.se
-# })
-# js = cbind.data.frame(performance.at.1se, jack.se)
-# ggplot(js, aes(rownames(js), performance.at.1se, color=(performance.at.1se > mean(performance.at.1se)+sd(performance.at.1se) | performance.at.1se < mean(performance.at.1se)-sd(performance.at.1se)  ))) + ylim(0.6,1) + geom_point() +
-#   geom_errorbar(aes(ymin=performance.at.1se-jack.se, ymax=performance.at.1se+jack.se)) + theme(legend.position = 'none', axis.text.x = element_text(angle=90, hjust=1))
-
-# in this case jack.se is the SE of the variance 
-# per coefficient
-features$jack.se = apply(coefs.per.pt, 1, function(x) {
-  jk = jackknife(x, var)
-  jk$jack.se
-})
-colnames(features) = c('label','coef','jack.se')
-ggplot(features, aes(reorder(label, coef), coef)) + geom_point() + 
-  geom_errorbar(aes(ymin=coef-jack.se, ymax=coef+jack.se)) + 
-  coord_flip() +
-  plot.theme + labs(x='', title='Jackknifed coefficient CI')
-```
-
-Coefficients have fairly tight CI if the SE of the jackknife variance is used.
+ogjLimits = levels(factor(pg.samp$ogj, ordered = T))
+p2 = mtn.spatial(globalCNdf %>% filter(Patient == 20 & seg.type != 'arms'), pg.samp, ogjLimits,'Patient 20 (P)' )
+p6 = mtn.spatial(globalCNdf %>% filter(Patient == 51 & seg.type != 'arms'), pg.samp, ogjLimits,'Patient 51 (NP)' )
+ggsave(filename='plots/spatial/patient_20_P.png', plot=grid.arrange(p2$grob), width=10.5, height=4.5, units='in', dpi=300)
+ggsave(filename='plots/spatial/patient_51_NP.png', plot=grid.arrange(p6$grob), width=10.5, height=4.5, units='in', dpi=300)
+
+grid.arrange(p2$grob, p6$grob, ncol=1)
+
+# for (i in 1:nrow(sum.patient.data)) {
+#   print(i)
+#   pt = sum.patient.data[[i, 'Patient']]
+#   status = as.character(sum.patient.data[[i, 'Status']])
+# 
+#   pms = mtn.spatial(subset(globalCNdf, Patient == pt & seg.type != 'arms'), pg.samp, ogjLimits, paste('Patient', pt, '(', status, ')' ))
+# 
+#   fl = paste0('patient_', pt, '_',status,'.png')
+#   ggsave(paste('plots/spatial/', fl, sep=''), plot=grid.arrange(pms$grob), width=10.5, height=4.5, units='in', dpi=300)
+# }
+
+pg.samp = pg.samp %>% dplyr::select(-ogj)
+
+ggsave('plots/spatial/legend.png', plot=grid.arrange(p2$legend), width=2,height=2, units='in', dpi=300)
+```
 
@@
 
+#### By pathology?
 ```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
-
-riskPath = back %>% dplyr::group_by(Status, Pathology, Risk) %>% 
-  dplyr::summarise(Freq = length(Risk)) %>% 
-  mutate(Status = recode_factor(Status, NP = 'Non-Progressor', P = 'Progressor', .ordered = T))
-
-# riskPath = rbind( cbind.data.frame( table(subset(back, Status == 'P', select=c('Pathology','Risk'))), 'Status'='Progressor' ), 
-#                   cbind.data.frame( table(subset(back, Status == 'NP', select=c('Pathology','Risk'))), 'Status'='Non-Progressor' ))
-
-pathTotal = back %>% dplyr::group_by(Status, Pathology) %>% 
-  dplyr::summarise( 'nPath'=length(Pathology)) %>%
-  mutate(Status = recode(Status, NP = 'Non-Progressor', P = 'Progressor'))
-
-riskPath = full_join(riskPath, pathTotal, by=c('Status','Pathology')) %>% 
-  mutate(ratio = round(Freq/nPath, 3), Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered = T))
-
-p53Path = back %>% dplyr::filter(Status == 'P' & !is.na(p53.Status)) %>% 
+p53Path = pg.samp %>% dplyr::filter(Status == 'P' & !is.na(p53.Status)) %>% 
   dplyr::group_by(Pathology,p53.Status) %>% 
@@
 
-p53Path = full_join( back %>% dplyr::filter(Status == 'P' & !is.na(p53.Status)) %>% dplyr::group_by(Pathology) %>% dplyr::summarise(nPath = length(Pathology)), p53Path, by='Pathology') %>% 
+p53Path = full_join( pg.samp %>% dplyr::filter(Status == 'P' & !is.na(p53.Status)) %>% 
+                       dplyr::group_by(Pathology) %>% 
+                       dplyr::summarise(nPath = length(Pathology)), p53Path, by='Pathology') %>% 
   dplyr::mutate(ratio = round(Freq/nPath,2))
 
-pathTotal = riskPath %>% dplyr::group_by(Status, Pathology) %>% dplyr::summarise('nPath'=unique(nPath))
-
-rp = ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
-  scale_fill_manual(values=riskCols) + scale_color_manual(values=c('white','black','white')) +
-  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), position=position_stack(), size=4) +
-  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep=''), color=Risk), position=position_stack(vjust = 0.5), size=5, show.legend = F) +
-  facet_grid(~Status) + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom', text=element_text(size=10))
-rp
-
-ggsave(filename='plots/pred_by_path.png', plot=rp, width=9, height=7, units='in', dpi=300)
+rp = plot.risk.by.path(pg.samp %>% ungroup) + labs(title='Samples classified by pathology', y='% Classified')
+
+ggsave(filename='plots/pred_by_path.png', plot=(rp + theme(legend.position = 'none')), width=7, height=6, units='in', dpi=300)
 ```
@@
   scale_fill_manual(values=col, name='p53 IHC', labels=c('Normal','Aberrant')) + 
-  geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101)) +
+  geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), nudge_y=1) +
   geom_text(aes(label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5,color='white') +
-  plot.theme + labs(title='Samples predicted by aberrant p53 IHC', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
+  plot.theme + labs(title='Samples predicted by aberrant p53 IHC', y='% Predicted', x='Pathology') + 
+  theme(legend.position = 'bottom', text = element_text(size=14))
 rp
-ggsave(filename='plots/p53_pred.png', plot=rp, width=7, height=6, units='in', dpi=300)
-
-```
-
-```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=6}
-all.preds = back %>% dplyr::group_by(Status, Patient) %>% dplyr::summarise( 'Low'=sum(length(which(Risk == "Low"))),
-                                          'Moderate'=sum(length(which(Risk == "Moderate"))),
-                                          'High'=sum(length(which(Risk == "High"))),
-                                          'n.samples'=length(Patient) ) %>% dplyr::mutate(Patient = factor(Patient))
+ggsave(filename='plots/p53_pred.png', plot=rp, width=6, height=5, units='in', dpi=300)
+```
+
+```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=6, eval=F}
+all.preds = pg.samp %>% dplyr::group_by(Status, Patient, Risk) %>% tally %>% spread(Risk, n) %>% ungroup %>%
+  mutate_if(is.numeric, funs( ifelse(is.na(.), 0, .) )) %>% 
+  left_join(pg.samp %>% dplyr::group_by(Status, Patient) %>% tally, by=c('Status','Patient')) %>%
+  dplyr::rename('n.samples'='n') %>% 
+  dplyr::mutate(Patient = factor(Patient))
 
@@
 
-```{r enods, warning=F, message=F, echo=F, fig.height=8, fig.width=6}
+```{r endos, warning=F, message=F, echo=F, fig.height=8, fig.width=6}
 sq = c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12)
 
-endos = back %>% dplyr::group_by(Status, months.before.final, brks, PID) %>% dplyr::summarise(
-              'pred.endo'=ifelse(length(which(Risk == 'High')) > 0, 1, 0), 
+endos = pg.samp %>% dplyr::group_by(Status, months.before.final, brks, PID) %>% 
+  dplyr::summarise('pred.endo'=ifelse(length(which(Risk == 'High')) > 0, 1, 0), 
               'n.endos'=length(unique(PID)), 
@@
               'pred.samples'=length(which(Risk == 'High')),
-              'max.path'=max(Pathology)) %>% ungroup %>% 
+              'max.path'=max(Pathology),
+              'max.prediction' = max(Prediction)
+              ) %>% ungroup %>% 
   dplyr::mutate(brks = cut(months.before.final, include.lowest = T, ordered_result = T,
      breaks=sq,  labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))) %>% 
-  dplyr::mutate(brks = factor(brks, rev(levels(brks)))) # invert order
+  dplyr::mutate(brks = factor(brks, rev(levels(brks))))
+
+endos = endos %>% mutate(brks = recode_factor(brks, '96-120' = '>96', '>120'='>96'))
 
 endos.m = endos %>% dplyr::filter(Status == 'P') %>% dplyr::group_by(brks) %>% 
-  dplyr::summarise( 'pred.r'=sum(pred.endo)/sum(n.endos), 'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos))
-
-tmp = back %>% dplyr::filter(Status == 'P') %>% 
-  dplyr::mutate(brks = cut(months.before.final, include.lowest = T, ordered_result = T,breaks=sq,
-     labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))) %>% dplyr::group_by(brks) %>% 
-       dplyr::summarise( 'mean(P)'=mean(Prediction),'sem(P)'=sd(Prediction)/sqrt(length(brks)))
-
-endos.m = arrange(merge(endos.m,tmp), brks)
+  dplyr::summarise( 'pred.r'=sum(pred.endo)/sum(n.endos), 'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos),'mean(P)'=mean(max.prediction),'sem(P)'=sd(max.prediction)/sqrt(length(brks)))
 
 rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) +
-  geom_bar(stat='identity', fill='darkred', alpha=0.8, show.legend = F) + 
-  geom_errorbar(color='grey21', width=0.2) +
-  geom_text(aes(y=3, label=paste(round(pred.r*100), '%', sep='')), color='white') +
-  geom_text(aes(y=pred.r*100-5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
-  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') + coord_cartesian(ylim = c(0,100)) + 
-  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12)) 
+  geom_bar(stat='identity', fill=BarrettsProgressionRisk:::riskColors()[['High']], show.legend = F) + 
+  geom_errorbar(color='grey55', width=0.2) +
+  geom_text(aes(label=paste(round(pred.r*100), '%', sep='')), color='black', nudge_y=2) +
+  geom_text(aes(y=5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white', nudge_y=2) + 
+  labs(x='Months prior to endpoint', y='Percent classified', title='High risk classifications prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') + coord_cartesian(ylim = c(0,100)) +
+  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12) ) 
 rp
-ggsave(filename='plots/pred_endo.png', plot=rp, width=9, height=7, units='in', dpi=300)
-
-#endos$months.before.final = endos$months.before.final*-1
-#me = endos %>% dplyr::group_by(Patient) %>% dplyr::summarise('max.endo' = max(i), 'months' = min(months.before.final))
-```
-
-The first two quintiles and the last two together account for 80% of our samples.  Only 20% fall into the category "Moderate".
+ggsave(filename='plots/pred_endo.png', plot=rp, width=6, height=5, units='in', dpi=300)
+
+```
+
+#### Using the discretized risk what is the AUC?
+
+```{r}
+risk_df = pg.samp %>% dplyr::select(Status, Risk) %>% mutate( Risk  = factor(droplevels(recode_factor(Risk, 'Moderate' = 'Low')), levels=c('Low','High'), ordered=T) )
+
+rocRisk = pROC::roc(Status ~ Risk, data=risk_df, ci=T, of='thresholds')
+p1 = roc.plot(rocRisk) + labs(title='ROC based on discretized risk', subtitle=paste(levels(risk_df$Risk), collapse=', '))
+p1
+ggsave(filename='plots/risk_auc.png', plot=p1, width=4, height=4, units='in', dpi=300)
+
+#rocRisk2 = pROC::roc(Status ~ Risk, data=pg.samp, ci=T, of='thresholds', transpose = T)
+#p2 = roc.plot(rocRisk2) + labs(title='ROC based on discretized risk', subtitle=paste(levels(pg.samp$Risk), collapse=', '))
+#p2
+
+#grid.arrange(p1,p2, ncol=1)
+```
 
 ```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=5 }
-back = back %>% dplyr::mutate(nl = factor(nl, ordered=T), Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered = T), Patient = factor(Patient, ordered=T), Block = factor(Block, ordered=T), Pathology = recode(Pathology, BE = 'NDBE'))
-
-blank.axis = theme(axis.text = element_blank(), legend.key=element_rect(fill='grey39'), panel.background=element_rect(colour = 'black'), panel.grid.major=element_blank(), panel.spacing = unit(0.2, 'lines'), panel.border = element_rect(color="black", fill=NA, size=0.5)  )
-
-p = BarrettsProgressionRisk::patientRiskTilesPlot( back %>% filter(Patient == 6) %>% dplyr::select(Endoscopy.Year,Pathology,Block,Risk) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) ) + blank.axis
+pg.samp = pg.samp %>% dplyr::mutate(nl = factor(nl, ordered=T), 
+                                    Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered = T), 
+                                    Patient = factor(Patient, ordered=T), 
+                                    Block = factor(Block, ordered=T))
+
+#blank.axis = theme(axis.text = element_blank(), legend.key=element_rect(fill='grey39'), panel.background=element_rect(colour = 'black'), panel.grid.major=element_blank(), panel.spacing = unit(0.2, 'lines'), panel.border = element_rect(color="black", fill=NA, size=0.5)  )
+
+p = BarrettsProgressionRisk::patientRiskTilesPlot(pg.samp %>% filter(Patient == 6) %>% dplyr::select(Endoscopy.Year,months.before.final, Pathology,Block,Risk) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block), 'months.before.final' )
 tilesLegend = get.legend(p)
@@
 
-#pred.tiles(subset(back, Patient %in% c(6)), ncol=1, cols=myPal, probs=T, OR=F) + labs(title='') + theme(legend.position = 'none')
-#pred.tiles(subset(back, Patient %in% c(48)), ncol=1, cols=myPal, probs=T) + labs(title='') + theme(legend.position = 'none')
-#pred.tiles(subset(back, Patient %in% c(33)), ncol=1, cols=myPal, probs=T) + labs(title='')+ theme(legend.position = 'none')
-#pred.tiles(subset(back, Patient %in% c(89)), ncol=1, cols=myPal, probs=T) + labs(title='')+ theme(legend.position = 'none')
-```
-
-```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=15 }
-plist = purrr::map(c(13,20,34,56,60,48), function(pt) { 
-  BarrettsProgressionRisk::patientRiskTilesPlot( back %>% filter(Patient == pt) %>% dplyr::select(Endoscopy.Year,Pathology,Block,Risk) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) ) + blank.axis + labs(title=pt) + theme(legend.position = 'none')
+ggsave('plots/heatmaps/legend_long.png',plot=grid.arrange(tilesLegend), width=10, height=2, units='in', dpi=300)
+
+dir.create('plots/heatmaps/tiny/P',recursive = T, showWarnings = F)
+dir.create('plots/heatmaps/tiny/NP',recursive = T, showWarnings = F)
+# Tiny heatmaps
+plist = purrr::map(sum.patient.data$Patient, function(pt) {
+  patient = pg.samp %>% filter(Patient == pt) %>% dplyr::select(Patient,Status,Endoscopy.Year,Block,Risk,months.before.final) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) %>% ungroup
+
+  p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final','rev') + theme(legend.position = 'none') + labs(x='',y='',title='') + theme(axis.ticks = element_blank(), axis.text = element_blank(), plot.margin=grid::unit(c(0,0,0,0), "mm"))
+
+  ht = ifelse(length(unique(patient$GEJ.Distance)) > 1, length(unique(patient$GEJ.Distance)), length(unique(patient$GEJ.Distance))+0.5)
+  wd = ifelse(length(unique(patient$months.before.final)) > 1, length(unique(patient$months.before.final)), length(unique(patient$months.before.final))+0.5 )
+
+  ggsave(paste0('plots/heatmaps/tiny/',unique(patient$Status),'/pt_',pt,'.png'), plot = p, height=ht, width=length(unique(patient$months.before.final)), units='in', dpi=150, limitsize=F)
+
+  patient = pg.samp %>% filter(Patient == pt) %>% dplyr::select(Patient,Status,Pathology,Endoscopy.Year,Block,Risk,months.before.final) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) %>% ungroup
+
+    p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final','rev') + theme(legend.position = 'none') + labs(x='months',y='location') + theme(plot.title = element_text(face = 'bold', size=16))
+
+    ggsave(paste0('plots/heatmaps/',unique(patient$Status),'/pt_',pt,'.png'), plot = p, height=length(unique(patient$GEJ.Distance))+1, width=length(unique(patient$months.before.final))+1, units='in', dpi=300, limitsize=F)
+
+    return(p)
 })
-a = do.call('arrangeGrob', c(plist, top='Progressors'))
-
-plist = purrr::map(c(15,16,21,3,45,19) , function(pt) { 
-  BarrettsProgressionRisk::patientRiskTilesPlot( back %>% filter(Patient == pt) %>% dplyr::select(Endoscopy.Year,Pathology,Block,Risk) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) ) + blank.axis + labs(title=pt) + theme(legend.position = 'none')
-})
-b = do.call('arrangeGrob', c(plist, top='Non-progressors'))
-
-ggsave('plots/ex_pt_heatmap.png', plot=grid.arrange(a,b,ncol=2), width=6, height=15, units='in', dpi=300)
+names(plist) = sum.patient.data$Patient
+
+```
+
+```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=15}
+a = do.call('arrangeGrob', c(plist[c('13','20','34','56','60','48')], top='Progressors'))
+b = do.call('arrangeGrob', c(plist[c('15','16','21','3','45','19')], top='Non-progressors'))
+
+ggsave('plots/ex_pt_heatmap.png', plot=grid.arrange(a,b,ncol=2), width=12, height=15, units='in', dpi=300)
 
@@
 # Chr17
+
+One way of looking at the p53 prediction (e.g. Adam Bass) is to see how predictive the p-arm loss in chr17 is.
 
@@
 
-p11 = subset(globalCNdf, chr == 17 &  start == 1 & seg.type == 'arms')
-p11Loss = subset(p11, value < -1)
-table(p11Loss$Status)
-
-p11path = table(subset(patient.info, Samplename %in% p11Loss$variable)$Pathology)
-
-p11path['BE']/sum(p11path)
-sum(p11path[c('HGD','IMC')])/sum(p11path)
-
-table(p11Loss[c('Patient', 'Pathology')])
-
-# Samples with a loss vs all Progressor samples
-length(unique(p11Loss$variable))/length(unique(subset(globalCNdf, Status == 'Progressor')$variable))
-
-p11Loss = unique(p11Loss[,c('variable', 'Status','Hospital.Research.ID', 'Patient', 'Pathology')])
-length(unique(p11Loss$Patient))/length(subset(sum.patient.data, Status == 'P')$Patient)
-# Of those where we see loss of p arm, p53 IHC shows an "abnormal" result 33% of the time
-
-table(subset(patient.info, Samplename %in% p11Loss$variable, select='p53.Status'))/length(p11Loss$variable)
-
-(table(subset(patient.info, Samplename %in% p11Loss$variable, select=c('p53.Status', 'Pathology'))))
+p11 = filter(globalCNdf, chr == 17 &  start == 1 & seg.type == 'arms')
+p11Loss = p11 %>% filter(value < -1)
+
+p11Loss %>% group_by(Status) %>% tally(name='p11 Loss') %>% 
+  full_join( p11 %>% group_by(Status) %>% tally(name='Total Samples') ) %>% 
+  kable(caption='Chr17 p11 loss') %>% kable_styling(full_width = F)
+
+p11Loss %>% filter(value <= -1) %>% dplyr::select(Patient, Status) %>% distinct %>%
+  group_by(Status) %>%  tally(name='p11 Loss') %>%
+  kable(caption='Patients showing p53 loss & p53 IHC') %>% kable_styling(full_width = F)
+
+p11Path = patient.info %>% filter(Samplename %in% p11Loss$variable) %>% dplyr::group_by(Pathology) %>% tally(name='samples')
+
+p11Path %>% dplyr::mutate('ratio'=round(samples/sum(p11Path$samples),2)) %>% 
+kable(caption='Ratio of all Chr17 p11 loss samples by pathology' ) %>% kable_styling(full_width = F)
+```
+
+Samples with a loss vs all Progressor samples
+Ratio of samples with p11 loss vs all progressor samples: `r length(unique(p11Loss$variable))/length(unique(filter(globalCNdf, Status == 'Progressor')$variable))`
+
+Ratio of progressor patients with a p11 loss in at least 1 sample: `r length(unique(p11Loss$Patient))/length(subset(sum.patient.data, Status == 'P')$Patient)`
+
+```{r}
+patient.info %>% filter(Samplename %in% p11Loss$variable) %>% group_by(forcats::fct_explicit_na(p53.Status)) %>% tally %>% 
+  set_names(c('p53 Status', 'n samples')) %>% kable(caption='p53 IHC status for samples with p11 loss') %>% kable_styling(full_width = F)
 
@@
 pi$p11Loss = as.numeric(pi$Samplename %in% p11Loss$variable)
-
-plot(table(p11=pi$p11Loss, ihc=pi$p53.Status))
-
-cor.test(pi$p11Loss, (as.numeric(pi$p53.Status)-1))
-
-# So why might there be loss but no corresponding aberrant staining?
-table(subset(pi, p11Loss == 1 & p53.Status == 0)$Pathology)
-table(subset(pi, p11Loss == 1 & p53.Status == 1)$Pathology)
-
+#plot(table(p11=pi$p11Loss, ihc=pi$p53.Status))
+
+pi %>%  mutate(p53.Status = recode_factor(p53.Status, '1'='aberrant','0'='normal')) %>% dplyr::group_by(p11Loss, p53.Status) %>% tally %>% spread(p53.Status, n)
+
+ct = fisher.test(table(pi$p11Loss, pi$p53.Status))
+
+```
+
+17p11 loss and p53 IHC correlation  Fisher's odds `r ct$estimate`, p-value = `r ct$p.value`.
+
+## q arm
+```{r, echo=T}
+
+q11 = filter(globalCNdf, chr == 17 & start > 1 & seg.type == 'arms')
+q11gain = q11 %>% filter(value > 1)
+
+q11gain %>% filter(value >= 1) %>% group_by(Status) %>% tally(name='q11 Gain') %>% 
+  full_join( q11 %>% group_by(Status) %>% tally(name='Total Samples') ) %>% 
+  kable(caption='Chr17 q11 gain') %>% kable_styling(full_width = F)
+
+q11gain %>% filter(value >= 1) %>% dplyr::select(Patient, Status) %>% distinct %>%
+  group_by(Status) %>%  tally(name='q11 Gain') %>%
+  kable(caption='Patients showing q gain') %>% kable_styling(full_width = F)
+
+q11Path = patient.info %>% filter(Samplename %in% q11gain$variable) %>% dplyr::group_by(Pathology) %>% tally(name='samples')
+
+q11Path %>% dplyr::mutate('ratio'=round(samples/sum(q11Path$samples),2)) %>% 
+kable(caption='Ratio of all Chr17 q11 gain samples by pathology' ) %>% kable_styling(full_width = F)
 ```
@@
 ## Top Features
-```{r, echo=T}
-m = data.frame(matrix(nrow=nrow(features), ncol=0))
-
-for (i in 1:nrow(features)) {
-  feat = features[i,]
-
-  seg = subset(globalCNdf, chr == as.character(feat$chr) & start == feat$start & end == feat$end & seg.type != 'arms')  
-  if(feat$end-feat$start > 5e6)
-    seg = subset(globalCNdf, chr == as.character(feat$chr) & start == feat$start & end == feat$end & seg.type == 'arms')
-
-  segGL = subset(seg, value < -1)
-  if(feat$coef > 0)
-    segGL = subset(seg, value > 1)
-
-  m[i, 'n.samples'] = length(unique(segGL$variable))
-  m[i,'ratio.samples.P'] = length(unique(segGL$variable))/length(unique(subset(globalCNdf, Status == 'Progressor')$variable))
-  m[i, 'ratio.NP'] = table(segGL$Status)[1]/sum(table(segGL$Status))
-  m[i, 'ratio.P'] = table(segGL$Status)[2]/sum(table(segGL$Status))
-
-  segpath = table(subset(patient.info, Samplename %in% segGL$variable)$Pathology)
-
-  m[i,'BE'] = segpath['BE']/sum(segpath)
-  m[i,'HGD'] = sum(segpath[c('HGD','IMC')])/sum(segpath)
-}
-rownames(m) = features$label
-
-range(m$ratio.samples.P)
-summary(m$ratio.samples.P)
-subset(m, ratio.samples.P == max(ratio.samples.P))
-subset(features, label == rownames(subset(m, ratio.samples.P == max(ratio.samples.P))))
-subset(chh, label == rownames(subset(m, ratio.samples.P == min(ratio.samples.P))))
-#table(cdkGain[c('Patient', 'Pathology')])
-
-cor.test(features$hazards, m$n.samples)
-
-# Samples with a gain vs all samples
-
-```
 
@@
 One way of looking at the p53 prediction (e.g. Adam Bass) is to see how predictive the p-arm loss in chr17 is.
-```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=10}
-ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
-ccgenes = ccgenes[-grep('^X|Y', ccgenes$Genome.Location),]
-#ccgenes = ccgenes[-which(ccgenes$Mutation.Types == 'T'),] # drop the translocations
-#ccgenes = subset(ccgenes, !grepl('^X|Y', Genome.Location))
-#loc = unlist(strsplit(ccgenes$Genome.Location, ':|-'))
-loc = do.call(rbind.data.frame, strsplit(ccgenes$Genome.Location, ':|-'))
-colnames(loc) = c('chr','start','end')
-loc[] = lapply(loc[], function(x) as.numeric(as.character(x)))
-
-ccgenes = cbind(ccgenes, loc)
-ccgenes = ccgenes[c('Gene.Symbol','Genome.Location', 'Role.in.Cancer', 'Mutation.Types',colnames(loc))]
-#ccgenes = subset(ccgenes, grepl('\\w+',Role.in.Cancer))
-
-genes = cbind(subset(ccgenes, grepl('TSG|oncogene', Role.in.Cancer) & Mutation.Types != 'T', select=c('Gene.Symbol','Genome.Location', colnames(loc))), 'value'=min(globalCNdf$mean.value), 'Row.names'='', 'Patient'='', 'selected'=as.factor(0))
+```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=10, eval=T}
+ccgenes = read_tsv('~/Data/CosmicCensusGenes.tsv', col_types = cols(.default = col_character())) %>% 
+  filter(!grepl('^X|Y',`Genome Location`)) %>% 
+  separate(`Genome Location`, c('chr','start','end'),':|-', remove=F) %>%
+  dplyr::filter(start != '' | end != '') %>%
+  dplyr::select(`Gene Symbol`, `Genome Location`, `Role in Cancer`, `Mutation Types`, chr, start, end)
+
+# leave out the translocations
+genes = ccgenes %>% dplyr::filter(grepl('TSG|oncogene', `Role in Cancer`) & `Mutation Types` != 'T') %>% 
+  dplyr::select(`Gene Symbol`, `Genome Location`, chr, start, end) %>%
+  dplyr::mutate(value = min(globalCNdf$value), Patient = '', chr = factor(chr,levels=levels(chr.lengths$chr),ordered=T), start = as.numeric(start), end=as.numeric(end)) %>% arrange(chr, start)
 
 genesGR = makeGRangesFromDataFrame(genes, keep.extra.columns = T)
-chGR = makeGRangesFromDataFrame(chh, keep.extra.columns = T)
-
-ov = findOverlaps(chGR, genesGR)
+chGR = makeGRangesFromDataFrame(features, keep.extra.columns = T)
+
 #genes = as.data.frame(genesGR[unique(subjectHits(ov))])
@@
 
-g17 = plot.cn.chr(globalCNdf, chrom='17', chr.lengths, genes=genes, haz=chh, label=labeller(chr=label_both, Status=n.status)) + labs(
-title="All Samples") + theme(legend.position = 'none')
+g17 = plot.cn.chr(globalCNdf, chrom='17', chr.lengths, haz=features, genes=genes, label=labeller(chr=label_both, Status=n.status)) + 
+  labs(title="Chromosome 17") + scale_y_continuous(breaks = seq(-30,25,10), labels = c('-30', '','',0,'',20)) +
+  theme(legend.position = 'none', strip.text = element_blank(), text=element_text(size=14))
 g17
 
-g17 = g17 + theme(text=element_text(size=11), strip.text = element_text(size=11))
-ggsave(filename='plots/hazards/chr17.png', plot=g17, width=10, height=5, units='in', dpi=300)
-
-```
-
-```{r echo=F, warning=F, message=F, fig.height=5, fig.width=10}
+ggsave(filename='plots/hazards/chr17.png', plot=g17, width=8, height=4, units='in', dpi=300)
+
+```
+
+```{r echo=F, warning=F, message=F, fig.height=5, fig.width=10, eval=F}
+ov = findOverlaps(chGR, genesGR)
 chGR$Genes = ''
@@
 
-for (chr in unique(chh$chr)) {  
-  p = plot.cn.chr(subset(globalCNdf, seg.type != 'arms') , chrom=chr, chr.lengths, genes=genes, haz=chh)
+for (chr in unique(features$chr)) {  
+  #print(chr)
+  p = plot.cn.chr(globalCNdf,  chrom=chr, info=chr.lengths, haz=features, genes=genes, label=labeller(chr=label_both, Status=n.status)) +
+    scale_y_continuous(breaks = seq(-30,25,10), labels = c('-30', '','',0,'',20)) +
+    theme(strip.text.x = element_blank(), text=element_text(size=14)) + labs(title=paste('Chromosome',chr))
   f = paste('chr',chr,'.png',sep='')
-  ggsave(paste('plots/hazards/',f,sep=''), plot=p, width=10, height=5, units='in', dpi=300)
-  print(p)
-}
-
-annoG = arrange(subset(ccgenes, grepl('CDKN2A$|TP53|MYC$|SMAD4|VEG|ERBB2|EGFR|SMYD|ARID1A|FHIT|RUNX1', Gene.Symbol)), chr)
-annoG = merge(annoG, chr.lengths[c('chrom','chr.length')], by.x='chr',by.y='chrom')
-annoG$ymin = round(range(globalCN.plot$data$mean.value))[1]/2
-annoG$ymax = round(range(globalCN.plot$data$mean.value))[2]/2
+  ggsave(paste('plots/hazards/',f,sep=''), plot=p, width=8, height=4, units='in', dpi=300)
+}
+
+annoG = ccgenes %>% filter(grepl('CDKN2A$|TP53|MYC$|SMAD4|VEG|ERBB2|EGFR|SMYD|ARID1A|FHIT|RUNX1', `Gene Symbol`)) %>% arrange(chr) %>%
+  left_join(chr.lengths %>% dplyr::select(chrom,chr.length), by=c('chr'='chrom')) %>%
+  mutate(ymin = round(range(globalCN.plot$data$mean.value))[1]/2, ymax = round(range(globalCN.plot$data$mean.value))[2]/2) %>%
+  mutate(chr = factor(chr, levels=levels(chr.lengths$chr)), start = as.numeric(start), end = as.numeric(end))
 
@@
 
-```
-
-`r (genes[,c(1:3,6,7)]) %>% kable() %>% kable_styling(full_width=F)`
+#(genes[,c(1:3,6,7)]) %>% kable() %>% kable_styling(full_width=F)
+```
 
@@
 
-Individually none of these show very good classification rates, all hovering around 65%.
-However, more importantly is that Chr17 is only marginally better than any of the others (unless you exclude HGD) and all are below the 72% classification rate we get with all of the data
-
-# Validation Data
-
-## Normals sWGS sequenced
-
-These normal tissues came from some samples that had been mistakenly sequenced from a timepoint (path ID) that Barrett's samples were also sequenced.  These include:
+# Validation Cohort
 
 ```{r, echo=F}
-normal.info = readxl::read_xlsx(patient.file, sheet='Normals')
-
-normal.info %>% dplyr::select(`Patient ID`, `Path ID`, `Pathology notes`) %>% kable(caption='Normals') %>% kable_styling('striped', full_width = F)
-```
-
-These were excluded from the model building procedure, but were then predicted using the resulting model.
-We would expect them to all be low risk and all but one are.
-All have a relative risk below 0 as well.
+#p.diag = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/validation/progressor_diagnosis.xlsx') %>% 
+#  dplyr::mutate(`Hospital Research ID` = gsub('/', '_', `Hospital Research ID`), `Weeks Follow Up` = as.double(difftime(`Diagnostic Endoscopy`, `Sequenced Endoscopy`, units='weeks')))
+pastefun<-function(x) {
+  if ( !grepl('SLX-', x) ) x = paste0('SLX-',x)
+  return(x)
+}
+
+patient.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
+
+oac.info = readxl::read_xlsx(patient.file, 'OAC Test Samples') %>% 
+  dplyr::mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% dplyr::filter(!is.na(`SLX-ID`)) %>%
+  mutate(Samplename = paste(`SLX-ID`, `Index Sequence`, sep='.')) %>%
+  dplyr::select(-matches('SLX|Index')) %>%
+  mutate(Status = factor(Status, levels=levels(pg.samp$Status)), Block = factor(Block, levels=c('E','1','2'), ordered=T))
+
+val.info = readxl::read_xlsx(patient.file, 'Final Validation Samples') %>% 
+  dplyr::mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% dplyr::filter(!is.na(`SLX-ID`)) %>%
+  mutate(Samplename = paste(`SLX-ID`, `Index Sequence`, sep='.')) %>%
+  separate(`Block ID`, c('PID','Block'), sep='-|[:blank:]', remove=F) %>% dplyr::rename(PathID = `Block ID`) %>%
+  dplyr::select(-matches('SLX|Index'), -`Path Notes`) %>%
+  mutate(Pathology = recode(Pathology, 'GM'='NDBE'), Status = factor(Status, levels=levels(pg.samp$Status))) %>% 
+  ungroup %>% group_by(`Hospital Research ID`) %>% 
+  dplyr::mutate(`Weeks Follow Up` = as.double(difftime(`Last Endoscopy`, `Endoscopy`, units='weeks')))
+
+val.info$Block = sapply(strsplit(val.info$Block, ''), function(x) {
+  if (is.na(x)) return(1)
+  if (length(x) == 1) {
+    if (!grepl('\\d', x)) return(recode(x, 'A'=1, 'B'=2, 'C'=3, 'D'=4, 'E'=5, 'F'=6,'G'=7 ))
+    return(as.numeric(x))
+  }
+  if (length(x) > 1) {
+    i = which(grepl('\\d', x))
+    return(as.numeric(x[i]))
+  }
+})
+
+bind_rows(val.info %>% dplyr::select(`Hospital Research ID`, Status) %>% distinct %>% group_by(Status) %>% tally %>% 
+  spread(Status,n) %>% add_column(' ' = 'n Patients', .before=1),
+val.info %>% group_by(Status) %>% tally %>% spread(Status,n) %>% add_column(' ' = 'n Samples', .before=1)) %>% 
+  kable(caption='Validation cohort') %>% kable_styling(full_width = F)
+
+val.info
+
+vdir ='~/Data/BarrettsProgressionRisk/Analysis/validation/pcf_perPatient/50kb'
+residuals = do.call(bind_rows, purrr::map(list.files(vdir,'residuals.tsv',recursive = T, full.names = T), function(f) read_tsv(f, col_types = 'ccddddl')))
+failedQC = filter(residuals, varMAD_median > 0.008)
+```
+
+## Per sample (alpha = 0.9)
+
+```{r, fig.width=8, fig.height=12}
+vdir ='~/Data/BarrettsProgressionRisk/Analysis/validation/pcf_perPatient/50kb/predictions_5e6_all'
+
+all.val.preds = do.call(bind_rows, purrr::map(list.files(vdir, 'predictions.tsv', full.names = T, recursive = T), function(f) {
+    pd = read_tsv(f, col_types = 'cddcccDcccc') %>% dplyr::select(`Hospital Research ID`, Sample, Probability, `Relative Risk`)
+    return(pd)
+})) %>% mutate(quants = cut(Probability, breaks=cuts, include.lowest = T)) %>%
+  left_join(pred.confidence %>% dplyr::select(quants, Risk), by='quants')
+
+oac.preds = all.val.preds %>% filter(`Hospital Research ID` %in% unique(oac.info$`Hospital Research ID`))
+
+oac.preds %>% group_by(`Hospital Research ID`) %>% 
+  dplyr::summarise( Probability = max(Probability), `Relative Risk` = max(`Relative Risk`), Risk = max(Risk)) %>%
+  group_by(Risk) %>% tally %>% kable(caption='OAC samples used as a positive control') %>% kable_styling(full_width = F)
+
+all.val.preds = all.val.preds %>% 
+  filter( !Sample %in% failedQC$samplename & !`Hospital Research ID` %in% unique(oac.info$`Hospital Research ID`) )
+
+val.preds = left_join(all.val.preds, val.info, by=c('Hospital Research ID', 'Sample'='Samplename')) %>% 
+  dplyr::select(Patient, everything()) %>% 
+  mutate(Pathology = factor(Pathology, levels=c('NDBE','ID','LGD'), ordered=T)) %>% 
+  mutate(Final.Endoscopy = `Last Endoscopy`, Endoscopy.Year = Endoscopy, Block = factor(Block, ordered=T))
+
+```
 
 ```{r}
-
-normal.preds = read_tsv('~/Data/BarrettsProgressionRisk/Analysis/multipcf_plots_normals/normal_predictions.tsv', col_types = cols(
-  Sample = col_character(),
-  Probability = col_double(),
-  `Relative Risk` = col_double(),
-  Risk = col_character(),
-  Patient = col_character(),
-  Hospital.Research.ID = col_character(),
-  Path.ID = col_character(),
-  Status = col_character(),
-  Endoscopy = col_double(),
-  Pathology = col_logical(),
-  Plate.Index = col_character(),
-  SLX.ID = col_double(),
-  Barretts.Cellularity = col_logical(),
-  P53.Status = col_character(),
-  Total.Reads = col_double(),
-  Batch.Name = col_character(),
-  Set = col_character(),
-  Block = col_double()
-))
-
-ggplot(normal.preds, aes(Patient, Probability, color=Risk)) + ylim(0,1) + geom_point(size=2) +
-  scale_color_manual(values = BarrettsProgressionRisk::riskColors() ) + plot.theme
-
-```
-
-## Test 'holdout' Set
-
-20 Patients (10P, 10NP) set aside for training/test purposes
-
-```{r val_setup, echo=F, message=F, warning=F, include=F}
-validation.patient.info =  all.patient.info %>% filter(Set == 'Test') %>% arrange(Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
-sum.val.info = summarise.patient.info(validation.patient.info)
-
-sum.val.info %>% kable() %>% kable_styling()
-```
-
-```{r, echo=F}
-
-file = paste(cache.dir, 'model_data.Rdata', sep='/')
-if (!file.exists(file))
-  stop(paste("Missing data file", file))
-load(file, verbose=T)
-
-predicted = '~/Data/BarrettsProgressionRisk/Analysis/training/val_predictions/0.9'
-
-vpd = do.call(bind_rows, lapply(list.files(predicted, full.names = T), function(f) {
-  read_tsv(f, col_types = 'cddcccdccdc')
-})) %>% dplyr::select(Hospital.Research.ID, everything())
-
-vpd = left_join(vpd %>% dplyr::select(-Pathology, -matches('P53')), validation.patient.info %>% 
-                  dplyr::select(Samplename, Patient, Pathology, p53.Status, Block, Initial.Endoscopy, Final.Endoscopy, Status ), by=c('Sample'='Samplename')) %>% 
-  dplyr::rename(Samplename = Sample, Endoscopy.Year = Endoscopy, RR = `Relative Risk`, Prediction = Probability) %>% 
-  dplyr::mutate(Risk = factor(Risk, levels=levels(back$Risk), ordered = T), GEJ.Distance = factor(GEJ.Distance, ordered = T),Pathology = recode(Pathology, BE='NDBE'))
-
-vpd = do.call(bind_rows,lapply(unique(vpd$Patient) %>% map( function(p) vpd %>% filter(Patient == p)  ), transform.by.time, info=validation.patient.info))
-
-## Need to add months to each endoscopy if multiple occur in the same year
-vpd = vpd %>% dplyr::mutate(PID = factor(PID)) %>% 
-  dplyr::group_by(Patient, PID) %>% dplyr::arrange(PID) %>% 
-  dplyr::mutate( Endoscopy = as.Date(paste0(Endoscopy.Year,'-01-01')) + as.numeric(PID)  ) %>% ungroup
-
-test.roc = pROC::roc(Status~Prediction, data=vpd,auc=T, ci=T, of='thresholds',transpose=T)
-roc.plot(test.roc, title='Test Set ROC')
-
-vpd.per.endo = vpd %>% group_by(Hospital.Research.ID, Status, Patient, PID) %>% 
-  dplyr::mutate('n.samples'=length(PID),'Samples'=paste(Samplename,collapse=',')) %>% 
-  dplyr::summarise_at(vars(matches('Pred|RR|Year')), list(max)) %>% dplyr::select(-matches('^Sample$')) %>% 
-  arrange(Patient, Endoscopy.Year, PID) 
-vpd.per.endo = vpd.per.endo %>% mutate(quants = cut(Prediction, breaks=cuts, include.lowest = T)) %>% left_join(pred.confidence %>% dplyr::select(quants, Risk), by='quants') %>% dplyr::select(-matches('quant')) %>% dplyr::mutate(Risk = as.character(Risk))
-
-vpd %>% group_by(Status, Risk)  %>% dplyr::summarise(n = length(Risk)) %>% spread(Risk, n) %>% rowwise %>% 
-  dplyr::mutate(n=sum(High,Low,Moderate), Low = Low/n, Moderate = Moderate/n, High = High/n) %>%
-  mutate_if(is.double, list(round),2) %>%
-  kableExtra::kable(caption='Test set per-sample predictions') %>% kableExtra::kable_styling(full_width = F)
-
-vpd.per.endo %>%  group_by(Status, Risk)  %>% dplyr::summarise(n = length(Risk)) %>% spread(Risk, n) %>% rowwise %>% 
-  dplyr::mutate(n = sum(High,Low,Moderate), Low = Low/n, Moderate = Moderate/n, High = High/n) %>% mutate_if(is.double, list(round),2) %>%
-  kableExtra::kable(caption='Test set per-endoscopy predictions') %>% kableExtra::kable_styling(full_width = F)
-
-```
-
-```{r validation, message=F, warning=F, echo=F, fig.height=8, fig.width=8}
-
-# for (s in rownames(pred)) {
-#   vpd[which(vpd$Samplename == s), 'Prediction'] = pred[s,]
-#   vpd[which(vpd$Samplename == s), 'OR'] = or[s,]
-# }
-
-sq = c(0, seq(1,132,12),max(back$months.before.final))
-vpd = vpd %>% mutate(brks = cut(months.before.final, 
-                  breaks=sq, labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x,collapse='-')))[-c(length(sq)-1,length(sq))], '>120'), include.lowest = T, ordered_result = T)) %>%
-  mutate(brks = factor(brks, rev(levels(brks))))
-```
-
-```{r echo=F, message=F, warning=F, fig.height=5, fig.width=8}
-vpd = vpd %>% ungroup %>% dplyr::mutate(Risk = factor(Risk, levels=c('High','Moderate','Low')), nl = factor(nl, ordered = T)) %>%
-  dplyr::mutate(Patient = factor(Patient))
-
-m = melt(vpd %>% dplyr::group_by(Patient, Status) %>% dplyr::summarise( 'Low'=length(which(Risk == 'Low'))/length(Patient),
-                                            'Moderate'=length(which(Risk == 'Moderate'))/length(Patient),
-                                            'High'=length(which(Risk == 'High'))/length(Patient),
-                                            'n.samples'=length(Patient)), id.vars=c('Patient','Status','n.samples'))
-
-ggplot(m, aes(Patient,  value*100, group=variable, fill=variable)) + geom_bar(stat='identity', position=position_stack()) +
-  geom_text(aes(label=n.samples, y=101)) + scale_fill_manual(values=riskCols) +
-  facet_grid(~Status, scales='free_x') + labs(y='Percent of samples', title='Validation patient set') + plot.theme
-
-```
-
-```{r echo=F, warning=F, message=F, fig.height=10,fig.width=25}
-plist = purrr::map((sum.val.info %>% filter(Status == 'P'))$Patient, function(pt) { 
-  BarrettsProgressionRisk::patientRiskTilesPlot( vpd %>% filter(Patient == pt) %>% dplyr::select(Endoscopy,Pathology,GEJ.Distance,Risk) ) + blank.axis + labs(title=pt) + theme(legend.position = 'none')
-})
-p1 = do.call('arrangeGrob', c(plist, top='Progressors'))
-
-plist = purrr::map((sum.val.info %>% filter(Status == 'NP'))$Patient, function(pt) { 
-  BarrettsProgressionRisk::patientRiskTilesPlot( vpd %>% filter(Patient == pt) %>% dplyr::select(Endoscopy,Pathology,GEJ.Distance,Risk) ) + blank.axis + labs(title=pt) + theme(legend.position = 'none')
-})
-p2 = do.call('arrangeGrob', c(plist, top='Non-Progressors'))
-grid.arrange(p1, p2, nrow=2)
-
-ggsave(filename='plots/val_heatmap.png', plot=grid.arrange(p1, p2, nrow=2), width=18, height=6, units='in', dpi=300)
-
-#ggsave(filename='plots/val_leg.png', plot=valLeg, width=4, height=2, units='in', dpi=300)
-
-```
-
-```{r echo=F, warning=F, message=F, fig.height=5,fig.width=8}
-BarrettsProgressionRisk::patientRiskTilesPlot( vpd %>% filter(Patient == 68) %>% 
-                                                 dplyr::select(Endoscopy,Pathology,GEJ.Distance,Risk) ) + 
-  theme(legend.position = 'right')
-```
-
-```{r, echo=F, message=F, warning=F, fig.width=8, fig.height=8}
-riskPath = rbind( cbind.data.frame( table(subset(back.v, Status == 'P', select=c('Pathology','Risk'))), 'Status'='Progressor' ), 
-                  cbind.data.frame( table(subset(back.v, Status == 'NP', select=c('Pathology','Risk'))), 'Status'='Non-Progressor' ))
-
-pathTotal = back.v %>% dplyr::group_by(Status, Pathology) %>% dplyr::summarise( 'nPath'=length(Pathology))
-pathTotal$Status = ifelse(pathTotal$Status == 'P', 'Progressor','Non-Progressor')
-
-riskPath = merge(riskPath, pathTotal)
-riskPath$ratio = with(riskPath, round(Freq/nPath, 3))
-riskPath$Risk = factor(riskPath$Risk, levels=c('Low','Moderate','High'), ordered = T)
-
-hgdimc = subset(riskPath, Pathology %in% c('HGD','IMC')) %>% dplyr::group_by(Risk) %>% summarize (
-  Freq=sum(Freq),  nPath=sum(nPath), ratio=sum(Freq)/sum(nPath))
-
-riskPath = riskPath[-which(riskPath$Pathology %in% c('HGD','IMC')),] 
-riskPath = rbind(riskPath , cbind('Pathology'='HGD/IMC', 'Status'='Progressor', hgdimc))
-
-riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor'), ordered = T)
-
-pathTotal = riskPath %>% dplyr::group_by(Status, Pathology) %>% dplyr::summarise('nPath'=unique(nPath))
-
-pr = ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) +   scale_fill_manual(values=riskCols) + scale_color_manual(values=c('white','black','white')) +
-  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), position=position_stack(), size=4, show.legend = F) +
-  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep=''), color=Risk), position=position_stack(vjust = 0.5), size=5, show.legend = F) +
-  facet_grid(~Status) + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
-pr
-
-ggsave(filename='plots/val_riskpath.png', plot=pr, width=7, height=7, units='in', dpi=300)
-
-```
-
-## Downsampled Datasets
-
-These come from Annalise, adjacent Barrett's tissues, and normal blood that had all been initially WGS at 30-80x and downsampled to an average 0.4x
-
-```{r echo=F, eval=T}
-dir = '~/Data/BarrettsProgressionRisk/Analysis/downsampled/predictions/0.9'
-info = '~/Data/BarrettsProgressionRisk/QDNAseq/all_downsampled/downsampled_ids.xlsx'
-
-info = readxl::read_xlsx(info)
-
-ds.preds = do.call(bind_rows, purrr::map(list.files(dir, 'predictions', recursive=T, full.names=T), function(f) {
-  read_tsv(f, col_types = cols(
-  Sample = col_character(),
-  Probability = col_double(),
-  `Relative Risk` = col_double(),
-  Risk = col_character(),
-  PatientID = col_character(),
-  Status = col_character(),
-  Tissue = col_character(),
-  `Expected Risk` = col_character(),
-  Grade_patient = col_character(),
-  Grade_biopsy = col_character(),
-  Study = col_character(),
-  Type = col_character(),
-  Endoscopy = col_date(format = "") ))
-}))
-
-## TODO merge in RR from 'back'
-
-ds.preds = ds.preds %>% dplyr::select(-Status) %>% left_join(info %>% dplyr::select(`Illumina ID`, Status), by=c('Sample' = 'Illumina ID')) %>% dplyr::mutate(Type = factor(Type, levels=c('Normal','NDBE', 'LGD', 'Tumor Adjacent BE'))) %>% arrange(Probability)
-
-m = reshape2::melt(ds.preds, measure.vars = c('Relative Risk'), id.vars = c('Type', 'Status', 'Risk'))
-
-pr = ggplot(m, aes(variable, value)) + 
-  facet_grid(~Type, scales = 'free_x') + 
-  geom_boxplot(aes(group=variable, fill=Type), outlier.colour = NA, show.legend = F) + 
-  geom_jitter(width=0.2, aes(shape=Status, color=Risk), size=2, show.legend = F) + 
-  scale_color_manual(values =  BarrettsProgressionRisk::riskColors()) +
-  #scale_color_gradientn(colors = myPal,  name='') +
-  scale_fill_manual(values=rev(RColorBrewer::brewer.pal(5,'RdYlBu')[c(1,2,4,5)])) +
-  labs(x='', y='Relative Risk', title='Predictions on downsampled patients') + plot.theme +
-  theme(axis.text.x = element_text(angle=45, hjust=1))
-
-pr
-
-ds.preds %>% dplyr::group_by(Status, Risk) %>% dplyr::summarise(n=length(Risk)) %>% spread(Risk, n) %>% kableExtra::kable(caption = 'Sample by risk and known status') %>% kableExtra::kable_styling(full_width = F)
-
-ggsave('plots/adjacentBE.png', plot=pr, width=8, height=8,dpi=300)
-
-```
-
-## Validation Cohort
-
-We sequenced a new set of non-progressors, progressors, and some OACs from the OCCAMS cohort where we have high-depth WGS.
+roc = pROC::roc(Status ~ Probability, data=val.preds, auc=T, ci=T, of='thresholds')
+roc.plot(roc) + labs(title='Validation cohort')
+pROC::ci.auc(roc$auc)
+
+ggsave('plots/validation/validatation_roc.png', plot=roc.plot(roc) + labs(title='Validation cohort'), width=5, height=5, units='in',dpi=300)
+
+totals = val.preds %>% group_by(Status) %>% tally
+val.counts = val.preds %>% 
+    group_by(Status, Risk) %>% tally %>% spread(Risk,n) %>% ungroup %>%
+    left_join(totals, by='Status') %>% mutate_if(is.numeric, funs(./n)) %>% dplyr::select(-n)
+
+ggplot(melt(val.counts, id.vars=c('Status')), aes(variable, value, group=Status, fill=Status)) + ylim(0,1) + 
+    geom_bar(stat='identity', position='dodge') + geom_text(aes(label=round(value,2)), position = position_dodge(width=1)) +
+    labs(title='Validation risk (sample ratios per NP:P)', x='', y='ratio') + theme_minimal()
+```
+
+## By path
 
 ```{r}
-val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
-analysis.dir = '~/Data/BarrettsProgressionRisk/Analysis/validation'
-
-oac.cohort = readxl::read_xlsx(val.file, 8, trim_ws = T) %>% 
-    dplyr::filter(!is.na(`SLX-ID`)) %>% dplyr::select(`Hospital Research ID`, `Block ID`, `Sample Date`, `SLX-ID`, `Index Sequence`, Status) %>% mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`)) %>% dplyr::mutate(Pathology = 'OAC') %>% dplyr::rename('Endoscopy' = `Sample Date`) %>% dplyr::rename('PID' = `Block ID`) %>% dplyr::mutate(GEJ.Distance = 1)
-
-sheets = readxl::excel_sheets(val.file)[9:12]
-info = do.call(bind_rows, lapply(sheets, function(s) {
-  readxl::read_xlsx(val.file, s, trim_ws = T) %>% 
-    dplyr::filter(!is.na(`SLX-ID`)) %>% dplyr::select(`Hospital Research ID`, `Block ID`, Status, Endoscopy, Pathology, `SLX-ID`, `Index Sequence`, `Path Notes`) %>% mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`))
-}))
-
-info = info %>% separate(`Block ID`, c('PID','GEJ.Distance'), '( |-)', remove = F, fill='right') %>% 
-  dplyr::mutate(GEJ.Distance = ifelse(is.na(GEJ.Distance), '1', GEJ.Distance)) %>%
-  dplyr::mutate(GEJ.Distance = recode(GEJ.Distance, A1 = '1', A = '1', A2 = '2', A3 = '3', A4 = '4', `4A` = '4', B = '2', C = '3', D = '4')) %>%
-  dplyr::mutate(GEJ.Distance = as.numeric(GEJ.Distance))
-
-bind_rows(
-  add_column(info %>% dplyr::select(`Hospital Research ID`, Status) %>% distinct() %>% dplyr::filter(Status %in% c('P','NP')) %>% dplyr::group_by(Status) %>% dplyr::count() %>% spread(Status, n), ' ' = 'BE Patients', .before = 1),
-  add_column(info %>% dplyr::filter(Status %in% c('P','NP')) %>% dplyr::group_by(Status) %>% dplyr::count() %>% spread(Status, n), ' ' = 'BE Samples', .before=1),
-  bind_cols(' ' = 'OAC Samples','NP'=0,'P'=nrow(oac.cohort))
-  ) %>% 
-kable(caption='Validation cohort counts') %>% kable_styling(full_width = F)
+rp = plot.risk.by.path(val.preds) + labs(title='Validation samples classified by pathology', y='% Classified')  
+rp
+ggsave(filename='plots/validation/val_pred_by_path.png', plot=rp, width=5, height=6, units='in', dpi=300)
+
+get.year<-function(x) {
+  as.numeric(format(as.Date(x), '%Y'))
+}
+
+val.preds = dplyr::rename(val.preds, 'Samplename'='Sample')
+
+val.preds = do.call(bind_rows, 
+                    lapply(unique(val.preds$Patient) %>% 
+                            map( function(p) val.preds %>% filter(Patient == p) %>% 
+                                   mutate_at(vars(Endoscopy.Year, Final.Endoscopy), list(~get.year(.)) )
+                                 ), transform.by.time, info=val.info) )
+
+dir.create('plots/validation/heatmaps/P/tiny',recursive = T, showWarnings = F)
+dir.create('plots/validation/heatmaps/NP/tiny',recursive = T, showWarnings = F)
+
+for (pt in unique(val.preds$Patient)) {
+#print(pt)
+  patient = val.preds %>% filter(Patient == pt) %>% 
+    dplyr::select(Patient, Status, Endoscopy, months.before.final, Block,Risk) %>% 
+    dplyr::rename(GEJ.Distance = Block)
+
+    p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final','rev') + theme(legend.position = 'none') + labs(x='',y='',title='') + theme(axis.ticks = element_blank(), axis.text = element_blank(), plot.margin=grid::unit(c(0,0,0,0), "mm"))
+
+    ht = ifelse(length(unique(patient$GEJ.Distance)) > 1, length(unique(patient$GEJ.Distance)), length(unique(patient$GEJ.Distance))+0.5)
+    wd = ifelse(length(unique(patient$months.before.final)) > 1, length(unique(patient$months.before.final)), length(unique(patient$months.before.final))+0.5 )
+
+    ggsave(paste0('plots/validation/heatmaps/',unique(patient$Status),'/tiny/pt_',pt,'.png'), plot = p, height=ht, width=length(unique(patient$months.before.final)), units='in', dpi=150, limitsize=F)
+
+    patient = val.preds %>% filter(Patient == pt) %>% 
+      dplyr::select(Patient, Status, Endoscopy, months.before.final, Pathology,Block,Risk) %>% 
+      dplyr::rename(GEJ.Distance = Block)
+
+  p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final', 'rev') + labs(x='months', y='location') + theme(legend.position = 'none',plot.title = element_text(face = 'bold', size=16))
+
+    ggsave(paste0('plots/validation/heatmaps/',unique(patient$Status),'/pt_',pt,'.png'), plot = p, height=length(unique(patient$GEJ.Distance))+1, width=length(unique(patient$months.before.final))+1, units='in', dpi=300, limitsize=F)
+}
+
+sq = c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12)
+val.preds = val.preds %>% mutate(months.before.final = ifelse(months.before.final > sq[length(sq)], sq[length(sq)], months.before.final))
+
+v.prog = val.preds %>% filter(Status == 'P') %>% mutate(
+  brks = cut(months.before.final, breaks=sq, 
+                 labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-c(length(sq)-1,length(sq))], '>120'), include.lowest = T, ordered_result = T),
+) %>% dplyr::mutate(brks = factor(brks, rev(levels(brks))))
+
+endos = v.prog %>% dplyr::group_by(Status, months.before.final, brks, PID) %>% 
+  dplyr::summarise('pred.endo'=ifelse(length(which(Risk == 'High')) > 0, 1, 0), 
+              'n.endos'=length(unique(PID)), 
+              'n.samples'=length(PID),
+              'pred.samples'=length(which(Risk == 'High')),
+              'max.path'=max(Pathology),
+              'max.prediction' = max(Probability)
+              ) %>% ungroup %>% 
+  dplyr::mutate(brks = cut(months.before.final, include.lowest = T, ordered_result = T, breaks=sq,  labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))) %>%
+  dplyr::mutate(brks = factor(brks, rev(levels(brks))))
+
+endos = endos %>% mutate(brks = recode_factor(brks, '96-120' = '>96', '>120'='>96'))
+
+endos.m = endos %>% dplyr::filter(Status == 'P') %>% dplyr::group_by(brks) %>% 
+  dplyr::summarise( 'pred.r'=sum(pred.endo)/sum(n.endos), 'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos),'mean(P)'=mean(max.prediction),'sem(P)'=sd(max.prediction)/sqrt(length(brks)))
+
+rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) +
+  geom_bar(stat='identity', fill=BarrettsProgressionRisk:::riskColors()[['High']], show.legend = F) + 
+  geom_errorbar(color='grey21', width=0.2) +
+  geom_text(aes(y=3, label=paste(round(pred.r*100), '%', sep='')), color='white') +
+  geom_text(aes(y=pred.r*100-5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
+  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') + coord_cartesian(ylim = c(0,100)) + 
+  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12)) 
+rp
+ggsave(filename='plots/validation/pred_endo.png', plot=rp, width=6, height=5, units='in', dpi=300)
+```
+
+# Risk Categories KM
+
+```{r, fig.width=6, fig.height=8}
+require(survival)
+require(survminer)
+require(GGally)
+
+ps = pg.samp %>% dplyr::select(Patient, Status,months.before.final, Risk) %>%
+  group_by(Patient) %>% mutate(Diagnosis = as.integer(Status)-1)
+
+sfit = survfit( Surv(months.before.final, Diagnosis, type='right') ~ Risk, data=ps, conf.int=0.95, conf.type='log', na.action=na.omit)
+names(sfit$strata) = sub('.*=','', names(sfit$strata))
+
+kmp = survminer::ggsurvplot(sfit, size=1, linetype="strata",risk.table=T, risk.table.col='strata',
+                      conf.int=T, censor=T, legend='none', title='Risks over time', xlab='Time (months prior to final sample)',
+                      ylab = 'Progression risk',
+                      ggtheme=theme_minimal(), palette=unlist(BarrettsProgressionRisk::riskColors()), font.legend=c(14,'bold'))
+
+kmp
+
+#ggsave('plots/risk_km.png', plot=print(kmp), height=9, width=8, units = 'in', dpi = 300)
 
@@
 ```{r}
-pred.dir = paste0(analysis.dir, '/predictions')
-
-batch8_preds = read_tsv(paste0(pred.dir, '/batch8_preds.txt')) %>% mutate(`Hospital Research ID` = gsub('_','/',`Hospital Research ID`)) %>% dplyr::select(`Hospital Research ID`, Sample, Probability, `Relative Risk`, Risk) %>%
-  left_join(oac.cohort, by=c('Hospital Research ID','Sample'))
-
-oldp.preds = read_tsv(paste0(pred.dir, '/oldp_preds.txt')) %>% mutate(`Hospital Research ID` = gsub('_','/',`Hospital Research ID`)) %>% dplyr::select(`Hospital Research ID`, Sample, Probability, `Relative Risk`, Risk) %>% left_join(info, by=c('Hospital Research ID','Sample'))
-
-new.np.preds = read_tsv(paste0(pred.dir, '/val_np.txt')) %>% mutate(`Hospital Research ID` = gsub('_','/',`Hospital Research ID`)) %>% dplyr::select(`Hospital Research ID`, Sample, Probability, `Relative Risk`, Risk) %>% left_join(info, by=c('Hospital Research ID','Sample'))
-
-val.preds = bind_rows(
-  batch8_preds %>% dplyr::select(`Hospital Research ID`, Probability, `Relative Risk`, Endoscopy, Status, PID) %>% mutate(Pathology = 'OAC', GEJ.Distance = 1) %>% dplyr::rename(ID = 'Hospital Research ID'),
-  oldp.preds %>% dplyr::select(`Hospital Research ID`, Probability, `Relative Risk`, Endoscopy, GEJ.Distance, Pathology, PID) %>% mutate(Status = 'P') %>% dplyr::rename(ID = 'Hospital Research ID'),
-  new.np.preds %>% dplyr::select(`Hospital Research ID`, Probability, `Relative Risk`, Endoscopy, GEJ.Distance, Pathology, PID) %>% mutate(Status = 'NP') %>% dplyr::rename(ID = 'Hospital Research ID') )
-
-val.preds = val.preds %>% mutate(quants = cut(Probability, breaks=cuts, include.lowest = T))
-
-val.preds = left_join(val.preds, pred.confidence %>% dplyr::select(quants, Risk), by='quants')
-
-ggplot(val.preds, aes(Status, Probability)) + geom_boxplot(aes(group=Status, fill=Status), outlier.shape = NA) + geom_jitter(width = 0.2) + plot.theme
-
-ggplot(val.preds, aes(Pathology, Probability)) + facet_grid(~Status,scales = 'free_x') + geom_boxplot(aes(group=Pathology), outlier.shape = NA) + geom_jitter(width = 0.3, aes(color=Probability, shape=Risk), size=3) + scale_color_gradientn(colors=myPal) + plot.theme
-
-```
-### Validation AUC
-
-```{r, echo=F, warning=F, message=F, fig.width=4, fig.height=3}
-
-val.preds %>% dplyr::select(Status, Probability)
-
-preds = do.call(rbind.data.frame, lapply(vpd, function(df) df[c('Status','Prediction')]))
-vroc = pROC::roc(Status ~ Probability, data=val.preds, ci=T, of='thresholds', transpose=T)
-vroc$model = 'Validation'
-
-roc$model = 'Discovery'
-multi.roc.plot(list(roc, vroc), title='ROC', palette='Dark2') + theme(text=element_text(size=12))
-ggsave('plots/auc_dv.png', plot=p, width=5, height=4, units='in', dpi=300, scale=2)
-```
-
-```{r, echo=F, warning=F, message=F, fig.width=4, fig.height=4}
-aucs = rbind.data.frame(pROC::ci.auc(roc), pROC::ci.auc(vroc))
-colnames(aucs) = c('ci.min', 'auc','ci.max')
-aucs$model = c('Discovery','Validation')
-
-ggplot(aucs, aes(model, auc)) + geom_point(size=2) + ylim(0,1) +
-  geom_errorbar(aes(ymin=ci.min, ymax=ci.max), width=0.1) + coord_flip() + 
-  geom_text(aes(y=ci.min, label=round(ci.min, 2)), vjust=4) +
-  geom_text(aes(y=ci.max, label=round(ci.max, 2)), vjust=-4) +
-#  geom_text(aes(y=auc, label=round(auc, 2)), vjust=-1) +
-  labs(x='', y='AUC', title='AUC with CI') + plot.theme 
-```
-
-```{r}
-dir.create('plots/validation')
-
-val.preds$Patient =  group_indices(val.preds, ID)
-
-val.preds = val.preds %>% group_by(ID) %>% mutate(PID = as.factor(PID)) %>% ungroup %>% group_by(ID, PID) %>% arrange(PID) %>% dplyr::mutate( Endoscopy = Endoscopy + as.integer(PID)  ) %>% ungroup
-
-plist = purrr::map(val.preds %>% filter(Pathology == 'OAC') %>% dplyr::select(ID) %>% distinct %>% pull, function(pt) { 
-  BarrettsProgressionRisk::patientRiskTilesPlot(
-    val.preds %>% filter(ID == pt) %>% mutate(GEJ.Distance = factor(GEJ.Distance)) ) + 
-    labs(title=pt) +
-    theme(legend.position = 'none')
-})
-a = do.call('arrangeGrob', c(plist, top='OAC'))
-ggsave('plots/validation/oac.png', plot=grid.arrange(a), width=3*3, height=4*2,dpi=300)
-
-plist = purrr::map(val.preds %>% filter(Status == 'P' & Pathology != 'OAC') %>% dplyr::select(ID) %>% distinct %>% pull, function(pt) { 
-  BarrettsProgressionRisk::patientRiskTilesPlot(
-    val.preds %>% filter(ID == pt) %>% mutate(GEJ.Distance = factor(GEJ.Distance)) ) + 
-    labs(title=pt) +
-    theme(legend.position = 'none')
-})
-b = do.call('arrangeGrob', c(plist, top='Progressors'))
-ggsave('plots/validation/progessors.png', plot=grid.arrange(b), width=4*4, height=4*2,dpi=300)
-
-plist = purrr::map(val.preds %>% filter(Status == 'NP') %>% dplyr::select(ID) %>% distinct %>% pull, function(pt) { 
-  BarrettsProgressionRisk::patientRiskTilesPlot(
-    val.preds %>% filter(ID == pt) %>% mutate(GEJ.Distance = factor(GEJ.Distance)) ) + 
-    labs(title=pt) +
-    theme(legend.position = 'none')
-})
-c = do.call('arrangeGrob', c(plist, top='Non Progressors', ncol=4))
-grid.arrange(c)
-ggsave('plots/validation/non-progessors.png', plot=grid.arrange(c), width=4*4, height=11*2,dpi=300, limitsize = F)
-
-val.preds %>% group_by(Status, Risk) %>% dplyr::count(Risk) %>% spread(Risk, n) %>% kable() %>% kable_styling(full_width = F)
+vp = val.preds %>% dplyr::select(Patient, Status, months.before.final, Risk) %>% 
+  mutate(Patient = as.factor(Patient)) %>% 
+  group_by(Patient) %>% mutate(Diagnosis = as.integer(Status)-1)
+
+merged = bind_rows(ps,vp)
+
+#merged = merged %>% mutate(Diagnosis = ifelse(Status == 'P' & months.before.final == 0, 1, 0))
+
+merged %>%
+
+sfit = survfit( Surv(months.before.final, Diagnosis) ~ Risk, data=merged, conf.int=0.95, conf.type='log', na.action=na.omit)
+names(sfit$strata) = sub('.*=','', names(sfit$strata))
+
+kmp = 
+  survminer::ggsurvplot(sfit, size=1, linetype="strata",risk.table=F, risk.table.col='strata',
+                      conf.int=T, censor=T, legend='none',# xlim=c(0,150),
+                      title='Risk Classifications', 
+                      subtitle='Training & Validation cohorts',
+                      xlab='Months prior to final sample',
+                      ylab = 'Progression free fraction',
+                      ggtheme=theme_minimal(), palette=unlist(BarrettsProgressionRisk::riskColors()), font.legend=c(14,'bold'))
+
+kmp$plot = kmp$plot + scale_x_reverse() 
+kmp$table = kmp$table + scale_x_reverse()
+
+kmp
+
+ggsave('plots/risk_km.png', plot=print(kmp), height=8, width=7, units = 'in', dpi = 300)
+
 ```
@@
 
-First note here is that our probability distribution for these 20 patients is somewhat flatter than in the training set.
-Likely a consequence of sample size.
-The labels show the proprotions of progressors that fell into each quintile in the training set.
-
-```{r, echo=F, warning=F, message=F}
-
-summary.v = table.path.risk(back.v)
-
-as.data.frame(summary.v) %>% kable() %>% kable_styling(full_width = F)
-
-write.table(summary.v, sep='\t', quote=F, row.names = F)
-```
-
 ## Vignettes
@@
 ```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=5}
-vign.pts = as_tibble(subset(sum.patient.data, Patient %in% c(3, 33, 34, 49, 56, 20), select=c('Status','Patient','age.diagnosis', 'smoking', 'C','M', 'initial.endo','final.endo', 'total.endos','total.samples')))
-colnames(vign.pts) = c('Status','Patient','Age at Diagnosis','Smoker','Circumference', 'Maximal length', 'Intial Endoscopy','Final Endoscopy', 'Total Endoscopies','Total Samples')
-
-## P
-p = vig.plot(df=subset(back, Patient == 34),info= subset(patient.info, Patient == 34),cols=riskCols) + theme(legend.position = 'bottom')
+sum.patient.data = sum.patient.data %>% mutate(smoking_str = recode(smoking, 'Y'='smoker', 'N'=''))
+sum.patient.data$smoking_str = replace_na(sum.patient.data$smoking_str, '')
+
+create.desc<-function(pt,age,sex,smk,len,date) {
+  str =  paste0('Patient ', pt, ': ', age, sex)
+  if (smk != '' & !is.na(smk)) str = paste0(str, ' ,', smk)
+  str = paste0(str,'\n')
+
+  if (len != '' & !is.na(len)) str = paste0(str, len,'cm length ')
+
+  str = paste0(str, 'BE diagnosed ', date)
+
+  return(str)
+}
+sum.patient.data$desc = sum.patient.data %>% group_by(Patient) %>% rowwise() %>% dplyr::summarise(desc = create.desc(Patient, age.diagnosis, gender, smoking_str, M, initial.endo))
+
+pt = 34
+p34 = vig.plot(df=filter(pg.samp, Patient == pt),info= filter(patient.info, Patient == pt),cols=riskCols) + 
+    theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
+    theme(legend.position = 'none', text = element_text(size=12))
+ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p34, width=6.35, height=4.5,dpi=72)
+
+pt = 3
+p3 = vig.plot(df=filter(pg.samp, Patient == pt),info= filter(patient.info, Patient == pt),cols=riskCols) + 
+    theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
+    theme(legend.position = 'none', text = element_text(size=12))
+ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p3, width=6.35, height=4,dpi=72)
+
+pt = 68
+p68 = vig.plot(df=filter(pg.samp, Patient == pt),info= filter(patient.info, Patient == pt),cols=riskCols) + 
+    theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
+    theme(legend.position = 'none', text = element_text(size=12))
+ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p68, width=6.35, height=5,dpi=72)
+
+for (pt in c(56,20,86,23,33,49,51,6,25)) {
+  length = sum.patient.data %>% ungroup %>% filter(Patient == pt) %>% dplyr::summarise(years = final.endo-initial.endo) %>% pull
+  height = as.numeric(max( pg.samp %>% ungroup %>% filter(Patient == pt) %>% dplyr::select(Block) %>% pull ) ) + 1
+
+  #length = 6
+  #height = 5
+
+  p = vig.plot(df=subset(pg.samp, Patient == pt),info= subset(patient.info, Patient == pt),cols=riskCols) + 
+    theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) 
   vigLeg = get.legend(p)
-p = p + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_34.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(subset(back, Patient == 56), subset(patient.info, Patient == 56),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_56.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(subset(back, Patient == 20), subset(patient.info, Patient == 20),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_20.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(subset(back, Patient == 86), subset(patient.info, Patient == 86),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_86.png', plot=p, width=6, height=6,dpi=300)
-
-print(p)
-
-## NP
-p = vig.plot(subset(back, Patient == 3), subset(patient.info, Patient == 3),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_3.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(df=subset(back, Patient == 23),info= subset(patient.info, Patient == 23),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_23.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(subset(back, Patient == 33), subset(patient.info, Patient == 33),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_33.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(subset(back, Patient == 49), subset(patient.info, Patient == 49),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_49.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(subset(back, Patient == 51), subset(patient.info, Patient == 51),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_51.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(subset(back, Patient == 6), subset(patient.info, Patient == 6),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_6.png', plot=p, width=6, height=6,dpi=300)
-
-p = vig.plot(subset(back, Patient == 54), subset(patient.info, Patient == 54),cols=riskCols) + theme(legend.position = 'none')
-ggsave('plots/vignette/vig_54.png', plot=p, width=6, height=6,dpi=300)
-
-print(p)
+  p = p + theme(legend.position = 'none', text = element_text(size=12))
+  ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p, width=length, height=height,dpi=300)
+}
+ggsave('plots/vignette/legend.png', vigLeg, width=8, height=1, dpi=300)
 
@@
 
-cols = intersect(colnames(back), colnames(back.v))
-#cols = cols[-grep('quants|pred\\.ratio|se\\(Risk\\)', cols)]
-allB = rbind.data.frame(back[,cols], back.v[,cols])
-allB$p53.Status = factor(allB$p53.Status, levels=c(0,1), ordered = T)
-
-maxP = allB %>% dplyr::group_by(Patient, Status, brks) %>% dplyr::summarise( 'maxP'=max(Prediction), 'Risk'=Risk[Prediction == max(Prediction)], 'Pathology'=max(Pathology), 'p53'=max(p53.Status,na.rm=T) )
-
-rules = do.call(rbind, lapply(unique(maxP$Patient), function(pt) {
-  r = with(arrange(subset(maxP, Patient == pt), brks), rx(pt, brks,maxP,Pathology,Risk,p53))
-  r$Status = unique(subset(maxP, Patient == pt)$Status)
-  r
-}))
-
-rules$t1 = factor(rules$t1, levels=levels(back$brks), ordered=T)
-rules$t2 = factor(rules$t2, levels=levels(back$brks), ordered=T)
-
-penult = subset(rules, t2 != 'Endpoint' & (t1 == rev(levels(back$brks))[2] | t2 == rev(levels(back$brks))[2] ))
-round(table(subset(penult, Status == 'P')$rule)/nrow(subset(penult, Status == 'P')), 2)
-round(table(subset(penult, Status == 'NP')$rule)/nrow(subset(penult, Status == 'NP')), 2)
-
-second = subset(rules,  t2 != 'Endpoint') %>% 
-  dplyr::group_by(Patient, Status) %>% 
-  dplyr::summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )
-
-table(subset(second, Status == 'P')$rule)
-table(subset(second, Status == 'NP')$rule)
-
-path.rx<-function(Patient, brks, Pathology) {
-  rules = as.data.frame(matrix(ncol=0, nrow=0))
-  # Consecutive
-  for (i in 1:length(brks)) {
-    #risks = table(Pathology[i:(i+1)])
-    rule = 'None'
-    if ( length(which(grepl('HGD|IMC', Pathology[i:(i+1)]))) > 0  |
-         length(which(grepl('LGD', Pathology[i:(i+1)]))) > 1 ) {
-      rule = 1
-    } else if ( length(which(grepl('LGD', Pathology[i:(i+1)]))) > 0 ) {
-      rule = 2
-    } else if ( length(which(grepl('BE|ID', Pathology[i:(i+1)]))) > 1  ) {
-      rule = 3
-    }
-  rules = rbind(rules, cbind( as.character(brks[i]), as.character(brks[(i+1)]), rule))
-  }  
-  colnames(rules) = c('t1','t2', 'rule')
-  rules$Patient = Patient
-  rules
-}
-
-rules2 = do.call(rbind, lapply(unique(maxP$Patient), function(pt) {
-  r = with(arrange(subset(maxP, Patient == pt), brks), path.rx(pt, brks,Pathology))
-  r$Status = unique(subset(maxP, Patient == pt)$Status)
-  r
-}))
-rules2$t1 = factor(rules2$t1, levels=levels(back$brks), ordered=T)
-rules2$t2 = factor(rules2$t2, levels=levels(back$brks), ordered=T)
-
-second = subset(rules2,  t2 != 'Endpoint') %>% 
-  dplyr::group_by(Patient, Status) %>% 
-  dplyr::summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )
-
-table(subset(second, Status == 'P')$rule)
-table(subset(second, Status == 'NP')$rule)
-
-```
-
-## Incidence
-
-Median follow up for our discovery cohort was `r median(sum.patient.data$end.year-sum.patient.data$start.year)`.
-
-Adjust the RR of the orginal model with 1-3.5% offset to get an absolute risk (AR) with the understanding that the cohort was age matched and unbalanced.
-
-```{r echo=F}
-file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
+pg.samp = pg.samp %>% mutate(p53.Status = factor(p53.Status, levels=c(0,1), ordered=T), Probability = Prediction)
+
+max.rules = purrr::map(unique(pg.samp$Patient), function(pt) {
+    preds = pg.samp %>% filter(Patient == pt) %>% arrange(Endoscopy.Year, months.before.final)
+    endo = BarrettsProgressionRisk:::.setUpRxTablePerEndo(preds %>% dplyr::mutate(Endoscopy = months.before.final, Sample = 'Samplename'), 'max', verbose=F) %>% mutate(Endoscopy = Endoscopy*-1)
+    rx = BarrettsProgressionRisk:::.apply.rules(endo, 'Endoscopy', pred.confidence,'numeric' )
+
+    left_join(rx, endo %>% dplyr::select(Endoscopy, months.before.final, Endoscopy.Year, Probability), by=c('Time 1'='Endoscopy')) %>%
+      mutate(`Time 2` = ifelse(months.before.final <= 0,'Endpoint',`Time 2`)) %>% 
+      mutate(Patient = pt) %>% mutate_at( vars(matches('Rule|Time')), funs(as.character) ) %>%
+      dplyr::select(Patient, months.before.final, Endoscopy.Year, Probability, everything()) 
+})
+names(max.rules) = unique(pg.samp$Patient)
+
+progressors = max.rules[as.character(sum.patient.data %>% filter(Status == 'P') %>% dplyr::select(Patient) %>% pull)]
+nonprog = max.rules[as.character(sum.patient.data %>% filter(Status == 'NP') %>% dplyr::select(Patient) %>% pull)]
+
+pids = names(which(sapply(progressors, function(x) {
+  nrow(x) >= 1 & grepl('Endpoint', x[nrow(x),'Time 2'])
+})))
+
+nids = names(which(sapply(nonprog, function(x) {
+  nrow(x) >= 1 & grepl('Endpoint', x[nrow(x),'Time 2'])
+})))
+
+penult = bind_rows( do.call(bind_rows, progressors[pids]) %>% mutate(Status = 'P'),
+                    do.call(bind_rows, nonprog[nids]) %>% mutate(Status = 'NP') )
+
+penult = penult %>% filter(`Time 2` != 'Endpoint') %>% arrange(Patient, months.before.final) %>% group_by(Patient, Status) %>% dplyr::mutate(gid = 1:n() ) %>% dplyr::filter(gid == 1)
+
+penult %>% dplyr::select(Patient, Status) %>% distinct %>% group_by(Status) %>% tally %>% kable(caption='n patients with at least two consecutive endoscopies prior to endpoints') %>%
+  kable_styling(full_width = F)
+
+penult %>% filter(Status == 'P' & Rule %in% c(3,4)) %>% kable(caption='Prog patients that would be missed if we moved Moderate to Low risk') %>% kable_styling(full_width = F)
+
+penult %>% filter(Status == 'NP' & Rule == 1) %>% kable(caption='NP patients for RFA') %>% kable_styling(full_width = F)
+
+penult %>% group_by(Status, Rule) %>% tally %>% spread(Rule,n) %>% 
+  kable(caption='Penult Endoscopy: # patients by rx rules based on the max (P) per endoscopy') %>% kable_styling(full_width=F)
+
+```
+
+#### From the second endoscopy
+
+```{r}
+pids = names(which(sapply(progressors, function(x) {
+  nrow(x) > 1 & !grepl('Endpoint', x[2,'Time 2']) & !grepl('None', x[2,'Rule'])
+})))
+
+nids = names(which(sapply(nonprog, function(x) {
+  nrow(x) > 1 & !grepl('Endpoint', x[2,'Time 2']) & !grepl('None', x[2,'Rule'])
+})))
+
+second = bind_rows( do.call(bind_rows, progressors[pids]) %>% mutate(Status = 'P'),
+                    do.call(bind_rows, nonprog[nids]) %>% mutate(Status = 'NP') )
+
+second = second %>% group_by(Patient, Status) %>% arrange(`Time 2`) %>% 
+  dplyr::mutate(gid = 1:n() ) %>% dplyr::filter(gid == 2)
+
+second %>% dplyr::select(Patient, Status) %>% distinct %>% group_by(Status) %>% tally %>% kable(caption='n patients with at least two years follow-up') %>%
+  kable_styling(full_width = F)
+
+second %>% group_by(Status, Rule) %>% tally %>% spread(Rule,n) %>% 
+  kable(caption='Second Endoscopy: # patients by rx rules based on the max (P) per endoscopy') %>% kable_styling(full_width=F)
+
+second %>% filter(Status == 'P' & Rule == '1') %>% ungroup %>% filter(months.before.final < 120) %>%
+  dplyr::summarise_at(vars(months.before.final), list(~min(.), ~max(.), ~mean(.), ~median(.)))
+
+second %>% filter(Status == 'NP' & Rule == '1') %>% ungroup  %>%
+  dplyr::summarise_at(vars(months.before.final), list(~min(.), ~max(.), ~mean(.), ~median(.))) 
+```
+
+# Check COX
+
+```{r}
+cox.cache.dir = '~/Data/BarrettsProgressionRisk/Analysis/cox_5e6_all/50kb/'
+file = paste(cox.cache.dir, 'model_data.Rdata', sep='/')
 load(file, verbose=T)
-rm(plots, coefs, cv.patient)
-
-fit = models[[select.alpha]]
-lambda = performance.at.1se[[select.alpha]]$lambda
-
-# number of patients we'd needed to have included to get 45 progressors from a representative population (I think)
-all = rbind(sum.patient.data, sum.validation.info)
-
-# Per year
-# This isn't quite right...we'd want to do something more like a 5 year 
-#pop = round((table(all$Status)['P']/(0.07*100))*100)
-
-```
-
-```{r estimateRisk, echo=F, warning=F, message=F, eval=F}
-# Probability that an individual will progress in a lifetime (20 years) given a rate of progression of 0.07 is about 77%, the probability for a specific individual is less
-1-(0.93^20)
-
-(1-(0.99^3))*15
-(1-(0.99^3))/10
-(1-(0.99^3))/10
-
-# Or probability an individual will progress given a rate of 1%-3.5%
-poft = sapply(seq(0.96,0.99, .01), function(x) {
- sapply(1:5, function(y) 1-(x^y) )
-})
-
-poft = as.data.frame(poft)
-colnames(poft) = paste((1-seq(0.96,0.99, .01))*100, '%', sep='')
-
-rp = apply(poft,1,function(x) cbind(mean(x),min(x), max(x)) )
-
-rownames(rp) = c('mean','min','max')
-colnames(rp) = paste('Year',1:5)
-
-rr = range(as.data.frame(subset(back, Pathology %in% c('BE'), select=c('Patient','Pathology','months.before.final','OR')))$OR)
-
-p = subset(back, Status == 'P' & months.before.final > 6 & OR > 10)$Patient
-subset(sum.patient.data, Patient %in% p) %>% dplyr::group_by(Patient) %>% dplyr::summarise(final.endo-initial.endo)
-
-rrPath = back %>% dplyr::group_by(Status, Pathology) %>% dplyr::summarise( 'mean.RR'=mean(OR),'ci.min'=mean(OR)-(1.96*sd(OR)), 'ci.max'=mean(OR)+(1.96*sd(OR)) )
-
-rrp = ggplot(rrPath, aes(Status, mean.RR)) + #scale_color_gradientn(colors=myPal, limits=c(min(rrPath[,3:5]), max(rrPath[,3:5])),  name='') +
-  facet_wrap(~Pathology, ncol=1) + 
-  geom_point() +  geom_errorbar(aes(ymin=ci.min, ymax=ci.max), width=0.1) + 
-  geom_text(aes(y=ci.min, label=round(ci.min,2)), vjust=2) + geom_text(aes(y=ci.max, label=round(ci.max,2)), vjust=2) +
-  coord_flip() + plot.theme + labs(y='Mean RR', x='')
-rrp
-
-ggsave('plots/relrisk_path.png', plot=rrp, width=4, height=6,dpi=300)
-
-as.data.frame(arrange(subset(back, Status == 'P' & months.before.final > 6 & Pathology %in% c('BE','ID','LGD') & OR > 1), Patient, months.before.final))
-
-ggplot(t(rp), aes(1:5, mean)) +
-  geom_ribbon(aes(ymin=mean/abs(rr[1]),ymax=mean*rr[2]), fill='rosybrown1') +
-  geom_ribbon(aes(ymin=min,ymax=max), fill='grey88') +
-  geom_line() + 
-  geom_text(data=data.frame(t(rp[,5])), aes(x=5,y=min, label='1%')) + geom_text(data=data.frame(t(rp[,5])), aes(x=5,y=max, label='4%')) +
-  geom_line(aes(y=mean*rr[2]), color='red') +
-  geom_line(aes(y=mean/abs(rr[1])), color='blue') +
-  labs(y='P(P)', x='Year', title='Prob. of individual patient progression', subtitle='1-4% over 5 years') + plot.theme
-
-subset(back, OR > 10)
-cor.test(back$OR, as.integer(back$Pathology))
-as.data.frame(subset(back, Pathology %in% c('HGD','IMC'), select=c('Patient','Pathology','months.before.final','OR')))
-
-# as.data.frame(subset(back, Status == 'P' & OR < 2 & OR > -2, select=c('Status','Patient','Pathology','months.before.final','OR')))
-# 
-# hist(subset(back, Pathology %in% c('BE'))$OR)
-
-```
-
-```{r 5yrrisk, echo=F, fig.height=8, fig.width=12}
-cases = table(sum.patient.data$Status)['P']
-
-mn = round((table(sum.patient.data$Status)['P']/(0.01*100))*100)
-m = round((table(sum.patient.data$Status)['P']/(0.0225*100))*100)
-mx = round((table(sum.patient.data$Status)['P']/(0.035*100))*100)
-
-offsetMin = log(cases/mn)
-offsetMean = log(cases/m)
-offsetMax = log(cases/mx)
-
-#binom.test(cases, popL, 0.1)
-# Offset for the linear predictor would then be -log(pop/44)
-# Effectively this is my relative risk for the population
-#offset = log(cases/popL)
-
-# link function provides odds
-#linPred = predict(fit, newx=dysplasia.df, s=lambda, type='link')
-#quantile(linPred)
-
-#df = rbind(dysplasia.df, validation.df)
-#df = dysplasia.df
-
-#p0 = predict(fit, newx=df, s=lambda, type='response')
-#p0 = 1/(1+exp(-linPred)) # logit to get probabilities, same as type='response'
-#p1 = 1/(1+exp(-linPred+abs(offset))) # calculates probs from risk (offset)
-
-cols = c('Hospital.Research.ID','Patient','Status','PID', 'Samplename','Endoscopy.Year','Pathology','Prediction','OR')
-#loopreds = rbind(loopreds[cols], do.call(rbind, vpd)[cols])
-loopreds = do.call(rbind, pg.samp)[cols]
-
-loopreds$P0 = 1/(1+exp(-loopreds$OR))
-
-loopreds$Mean = 1/(1+exp(-loopreds$OR+abs(offsetMean))) # Adjust OR for RR
-loopreds$Min = 1/(1+exp(-loopreds$OR+abs(offsetMin))) # Adjust OR for RR
-loopreds$Max = 1/(1+exp(-loopreds$OR+abs(offsetMax))) # Adjust OR for RR
-
-#ql = cbind.data.frame('y'=quantile(loopreds$OR)[c(2,4)], 'labels'=paste(c(25,75), '%', sep=''))
-
-# From my nested-LOO predictions
-
-ggplot(melt(loopreds[c('Min','Mean','Max')], measure.vars='Mean'), aes(value, fill=variable)) + 
-  geom_density(alpha=0.5) + plot.theme
-
-p = ggplot(melt(loopreds[c('Min','Max')]), aes(value)) + facet_wrap(~variable, ncol=2) +
-  geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + 
-    scale_fill_gradientn(colors = myPal,  name='') + 
-    plot.theme + theme(legend.position = 'none') + 
-    labs(y='n Samples', x='Absolute Risk',
-         title='5-year adjusted risk', subtitle='Adjusted for upper & lower bound: 1-3.5% risk')
-p
-
-ggsave('plots/adj_RR.png', plot=p, width=6, height=6,dpi=300)
-
-# ggplot(loopreds, aes(Min)) +
-#     geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + 
-#     scale_fill_gradientn(colors = myPal,  name='') + 
-#     plot.theme + theme(legend.position = 'none') + 
-#     labs(y='n Samples', x='Absolute Risk',
-#          title='Sample Predictions: 5 year adjusted risk', subtitle='Adjusted for lower bound, 1% risk')
-# 
-# ggplot(loopreds, aes(Max)) +
-#     geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + 
-#     scale_fill_gradientn(colors = myPal,  name='') + 
-#     plot.theme + theme(legend.position = 'none') + 
-#     labs(y='n Samples', x='Absolute Risk',
-#          title='5-year adjusted risk', subtitle='Adjusted for upper bound, 3.5% risk')
-
-```
-
-```{r lifetimeFPR, echo=F, fig.height=6, fig.width=6, eval=F}
-## FPR estimates based on lifetime risks
-# We could consider any NP patient with a minimum of 2 High risk calls?  or High/Moderate Risk?
-# More correctly it should be any patients with concurrent High risk, but that's been pretty stable so two high is good enough for now
-
-bd = bind_rows(back, back.v)
-table(bd$Status)
-
-risksL = subset(bd, Status == 'NP') %>% dplyr::group_by(Patient) %>% dplyr::summarise( 'n.samp'=length(Patient), 'high.risk'=length(which(Risk == 'High')), 'mod.risk'=length(which(Risk == 'Moderate')) )
-
-risksL = subset(risksL, high.risk > 0 | mod.risk > 0)
-
-high = length(subset(risksL, high.risk >= 2 )$Patient)
-#mod = length(subset(risksL, mod.risk >= 2 )$Patient)
-
-# How would I tune it so that the FPR is a lot closer to the real incidence ~1%
-# Current false positive rate is 1.5x higher than the highest incidence ~0.7
-fpr=high*10/44
-x = subset(bd, Status == 'NP' & Patient %in% subset(risksL, high.risk >= 2)$Patient & Risk == 'High')
-(length(unique(subset(x, Prediction > 0.8)$Patient))*10)/44
-
-# How would that affect TPR?
-#pROC::coords(roc, 0.6)
-#pROC::coords(roc, 0.8)
-
-## Per year risk
-## Evaluate over 5 years? How?
-pyr = c()
-for (yr in names(which(table(bd$Endoscopy.Year) > 10))) {
-  risks = subset(bd, Status == 'NP' & Endoscopy.Year == yr) %>% dplyr::group_by(Patient) %>% dplyr::summarise( 'n.samp'=length(Patient), 'high.risk'=length(which(Risk == 'High')), 'mod.risk'=length(which(Risk == 'Moderate')) )
-
-  risks = subset(risks, high.risk > 0 | mod.risk > 0)
-  high = length(subset(risks, high.risk >= 2 )$Patient)
-  mod = length(subset(risks, mod.risk >= 2 )$Patient)
-
-  print(paste(yr,high*15, sep=':'))
-  pyr = c(pyr, high*15)
-}
-
-summary(pyr)
-
-pop*.07
-pop*.02
-pop*.05
-
-```
-
-```{r clonality, echo=F, message=F, warning=F, eval=F}
-## Clonal Heterogenity
-patient.info %>% dplyr::group_by (Hospital.Research.ID, Patient, PID) %>% dplyr::summarise(
-  length(PID)
-)
-
-arms = c(grep(with(chr.lengths, paste(paste(chrom, ':', '1-\\d+000', sep=''), collapse='|')), colnames(dysplasia.df)),
-         grep(with(chr.lengths, paste(paste(chrom, ':', '\\d+000-', chr.length, sep=''), collapse='|')), colnames(dysplasia.df)))
-
-x = do.call(rbind, lapply(unique(patient.info$PID), function(pid) {
-    df = as.matrix(dysplasia.df[subset(info, PID == pid)$Samplename, -grep('cx',colnames(dysplasia.df))])
-    if (ncol(df) > 1) {
-      cbind('arms'=sum(rowSums(df[,arms],na.rm=T))/nrow(df),
-            'seg'=sum(rowSums(df[,-arms],na.rm=T))/nrow(df))
-    } else {
-      cbind('arms'=sum(df[arms,],na.rm=T),
-            'seg'=sum(df[-arms,],na.rm=T))
-    }
-}))
-rownames(x) = unique(patient.info$PID)
-x = x[complete.cases(x),]
-
-hist(x[,'seg'])
-hist(x[,'arms'])
-
-p = pnorm(x[,'seg'], mean(x[,'seg']), sd(x[,'seg']))
-summary(p)
-
-pa = pnorm(x[,'arms'], mean(x[,'arms']), sd(x[,'arms']))
-summary(pa)
-
-x[,'arms'] = scale(x[,'arms'])
-x[,'seg'] = scale(x[,'seg'])
-
-plot(x[unique(subset(patient.info, Patient == 48)$PID), 'arms'])
-plot(x[unique(subset(patient.info, Patient == 48)$PID), 'seg'])
-
-plot(x[unique(subset(patient.info, Patient == 34)$PID), 'arms'])
-plot(x[unique(subset(patient.info, Patient == 34)$PID), 'seg'])
-
-pp = back %>% dplyr::group_by(Patient, Hospital.Research.ID, PID) %>% dplyr::summarise(
-  'max'=max(Prediction), 'avg'=mean(Prediction)
-)
-
-pp = merge(pp, x, by.x='PID', by.y='row.names')
-
-cor.test(pp$seg, pp$max)
-
-```
-
-## Cox model - Time to progress?
-
-```{r cox, warning=F, message=F, echo=F, eval=T}
-library(survival)
-
-cox.pvalue<-function(cox, test='logrank') {
-  pvalue = NULL  
-  csum = summary(cox)
-
-  if (test == 'logrank') {
-    pvalue = csum$sctest[['pvalue']]  
-  } else if (test == 'wald') {
-    pvalue = csum$waldtest[['pvalue']]
-  } else if (test == 'loglik') {
-    pvalue = csum$logtest[['pvalue']]
-  }
-  return(pvalue)
-}
-
-prog.pred = lapply(pg.samp, function(df) {
-  df = df[,c('Hospital.Research.ID', 'Status', 'Patient', 'Age.at.diagnosis', 'PID', 'Status', 'Endoscopy.Year','Initial.Endoscopy', 'Final.Endoscopy', 'Pathology', 'Samplename','Prediction','Prediction.Dev.Resid')] %>% rowwise() %>% dplyr::mutate(
-      'Time.to.Final'=Final.Endoscopy-Endoscopy.Year)
-  df = arrange(df, Endoscopy.Year, Pathology)
-
-  df$Dup.PID = duplicated(df$PID)
-  x = df[!df$Dup.PID,]
-
-  for (i in 1:nrow(x)) {
-    if (i == 1) {
-      x[i,'tstart'] = 0
-      x[i, 'tstop'] = x[(i+1), 'Endoscopy.Year'] - x[i, 'Endoscopy.Year']
-      if (is.na(x[i,'tstop']) || x[i,'tstop'] == 0) x[i,'tstop'] = 0.5
-      next
-    }
-
-    x[i,'tstart'] = x[(i-1), 'tstop']
-
-    # Final sample is different -- will leave it out later
-    if (i == nrow(x)) {
-      x[i, 'tstop'] = x[i,'tstart'] + 0.5
-      break
-    }
-
-    x[i, 'tstop'] =  x[i,'tstart'] + x[(i+1), 'Endoscopy.Year'] - x[i, 'Endoscopy.Year']
-
-    # Samples taken in the same year
-    if ( x[i, 'Endoscopy.Year'] == x[(i+1), 'Endoscopy.Year'] )
-      x[i,'tstop'] = x[i,'tstop'] + 0.5 
-  }
-
-  x$Diagnosis = as.integer(x$Pathology %in% c('HGD','IMC'))
-
-  if (sum(x$Diagnosis) >= 1)
-    x = x[-which(x$Diagnosis == 1),]
-
-  if (nrow(x) > 0 && sum(x$Diagnosis) <= 0 && x$Status == 'P')
-    x[nrow(x), 'Diagnosis'] = 1
-
-  if (nrow(x) > 0) {
-    x$Dup.Samples = ''
-    for (j in 1:nrow(df[df$Dup.PID,]))  {
-      row = df[df$Dup.PID,][j,]
-      x[which(x$PID == row[['PID']]), 'Dup.Samples'] = paste(x[which(x$PID == row[['PID']]), 'Dup.Samples'], row[['Samplename']], sep=',')
-    }
-    x$Dup.Samples = sub('^,','',x$Dup.Samples)
-  }
-
-  return(x)
-})
-## By pathology?
-pp = do.call(rbind, prog.pred)
-
-pp[is.na(pp$tstop), 'tstop'] = 0.5
-pp$Pathology = as.character(pp$Pathology)
-pp$AAD = cut(pp$Age.at.diagnosis, breaks = 3)
-
-pp = merge(pp, back[c('Samplename','Risk', 'quants')], by='Samplename')
-pp$Risk = as.character(pp$Risk)
-
-sfit = survfit( Surv(tstart, tstop, Diagnosis, type='interval') ~ Risk, data=pp, conf.int=0.95, conf.type='log', na.action=na.omit)
-ggsurv(sfit,cens.shape=3,lty.est=1,back.white=F) + xlim(0,5) + labs(y='Diagnosed', title='~Risk')
-
-cox = coxph( Surv(tstart, tstop, Diagnosis) ~ Risk, data=pp) 
-ggsurv(survfit(cox),cens.shape=3,lty.est=1,back.white=F,
-      main="coxph(Survival, Diagnosis) ~ Risk" ) + 
-      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),2),"\n", nrow(pp), 'samples'))
-#summary(cox)
-
-sfit = survfit( Surv(tstart, tstop, Diagnosis, type='interval') ~ AAD, data=pp, conf.int=0.95, conf.type='log', na.action=na.omit)
-ggsurv(sfit,cens.shape=3,lty.est=1,back.white=F) + xlim(0,5) + labs(y='Diagnosed', title='~Age') 
-cox = coxph( Surv(tstart, tstop, Diagnosis) ~ AAD, data=pp) 
-ggsurv(survfit(cox),cens.shape=3,lty.est=1,back.white=F,
-      main="coxph(Survival, Diagnosis) ~ Age at diagnosis" ) + 
-      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),2),"\n", nrow(pp), 'samples'))
-#summary(cox)
-
-sfit = survfit( Surv(tstart, tstop, Diagnosis, type='interval') ~ Pathology, data=pp, conf.int=0.95, conf.type='log', na.action=na.omit)
-ggsurv(sfit,cens.shape=3,lty.est=1,back.white=F) + xlim(0,5) + labs(y='Diagnosed', title='~Pathology')
-cox = coxph( Surv(tstart, tstop, Diagnosis) ~ Pathology, data=pp) 
-ggsurv(survfit(cox),cens.shape=3,lty.est=1,back.white=F,
-      main="coxph(Survival, Diagnosis) ~ Pathology" ) + 
-      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),2),"\n", nrow(pp), 'samples')) + xlim(0,5)
-#summary(cox)
-
-## All genomic regions
-sample.pred = merge( do.call(rbind.data.frame, prog.pred), dysplasia.df, by.x='Samplename', by.y='row.names' )
-sample.pred$Pathology = as.character(sample.pred$Pathology)
-
-gcols = grep('\\d+:', colnames(sample.pred), value=T)
-sample.pred = sample.pred[c('tstart', 'tstop','Diagnosis','Pathology','Age.at.diagnosis',gcols)]
-
-sample.pred$Age.bracket = as.character(cut(sample.pred$Age.at.diagnosis, breaks=3))
-
-# all
-#cox = coxph( Surv(tstart, tstop, Diagnosis) ~., data=sample.pred)
-#ggsurv(survfit(cox))
-#summary(cox)
-```
-### Cox GLM selected features
-
-```{r echo=F, warning=F, echo=F}
-# glm selected features + Path
-cox2 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', 'Pathology',  grep('^\\d+:', features$label, value=T))]) 
-ggsurv(survfit(cox2),cens.shape=3,lty.est=1,back.white=F,
-      main="coxph(Survival, Diagnosis) ~ GLM features + Pathology" ) + 
-      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox2),2),"\n", nrow(pp), 'samples')) 
-#summary(cox2)
-
-cox2 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis','Age.bracket', grep('^\\d+:', features$label, value=T))]) 
-ggsurv(survfit(cox2),cens.shape=3,lty.est=1,back.white=F,
-      main="coxph(Survival, Diagnosis) ~ GLM features + Age" ) + 
-      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox2),2),"\n", nrow(pp), 'samples')) 
-#summary(cox2)
-
-cox2 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis',grep('^\\d+:', features$label, value=T))]) 
-ggsurv(survfit(cox2),cens.shape=3,lty.est=1,back.white=F,
-      main="coxph(Survival, Diagnosis) ~ GLM features" ) + 
-      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox2),2),"\n", nrow(pp), 'samples')) 
-#summary(cox2)
-
-# chr arms only
-options(scipen=999) # turn sci notation off
-coxArms = coxph(Surv(tstart, tstop, Diagnosis) ~., data=sample.pred[,c('tstart', 'tstop','Diagnosis', with(arm.loc, paste(chrom, ':', start, '-', end, sep='')))])
-ggsurv(survfit(coxArms),cens.shape=3,lty.est=1,back.white=F,
-      main="coxph(Survival, Diagnosis) ~ chr arms" ) + 
-      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(coxArms),2),"\n", nrow(pp), 'samples')) 
-#summary(coxArms)
-
-# chr 17 only
-cox17 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', grep('^17:',colnames(sample.pred), value=T))]) 
-ggsurv(survfit(cox17),cens.shape=3,lty.est=1,back.white=F,
-      main="coxph(Survival, Diagnosis) ~ chr17" ) + 
-      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox17),2),"\n", nrow(pp), 'samples')) 
-#summary(cox17)
-
-cox11 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', grep('^11:',colnames(sample.pred), value=T))]) 
-ggsurv(survfit(cox11))
-#summary(cox11)
-
-cox13 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', grep('^13:',colnames(sample.pred), value=T))]) 
-ggsurv(survfit(cox13))
-#summary(cox13)
-
-```
-
-## CoxRFX
-
-```{r echo=F, message=F, warning=F, eval=T}
-plot.waldtest<-function(wt) {
-  p = ggplot(wt, aes(coef.x, coef.y, color=p.value<0.05))
-  if (length(grep('hazards', colnames(wt))) > 0) {
-      p = p + geom_point(alpha=0.5, aes(size=hazards)) + 
-        geom_point(aes(coef.x, coef.y, color=p.value<0.05, size=hazards), show.legend = F) +
-        #scale_shape_manual(values=c(17,18), na.value=21) +
-        scale_size(range=c(2, 10), breaks=c(0, 5, 10, 15)) +
-        #geom_text_repel( aes( label=label), show.legend=F) + 
-        geom_text_repel(data=subset(wt, p.value < 0.05), aes(coef.x,coef.y, label=label), show.legend=F) + 
-        geom_text_repel(data=subset(wt, is.na(stability)), aes(coef.x, coef.y, label=label, color=p.value<0.05), show.legend=F, color='grey39') 
-  } else {
-    p = p + geom_point(alpha=0.5) + 
-      geom_text_repel(data=subset(wt, p.value < 0.05), aes(coef.x,coef.y, label=label), show.legend=F)  
-  }
-  return(p + labs(title='Wald test'))
-}
-```
-
-```{r echo=T, message=F, warning=F, fig.height=8, fig.width=8, eval=F}
-
-library(CoxHD)
-
-coxDF = sample.pred
-
-pathology = do.call(cbind.data.frame, lapply(unique(coxDF$Pathology), function(p) {
-  cbind( as.integer(coxDF$Pathology == p))
-}))
-colnames(pathology) = unique(coxDF$Pathology)
-
-# age = do.call(cbind.data.frame, lapply(unique(coxDF$Age.at.diagnosis), function(p) {
-#   cbind( as.integer(coxDF$Age.at.diagnosis == p))
-# }))
-# colnames(age) = unique(coxDF$Age.at.diagnosis)
-age = coxDF$Age.at.diagnosis
-
-## GLM regions only
-coxDF = cbind(coxDF[,1:3],pathology[,c('BE','ID')], age, coxDF[,grep('^\\d+:', features$label, value=T)])
-coxDF$Pathology = NULL
-
-gcols = colnames(coxDF)
-cols = grep('^\\d+', colnames(coxDF))
-colnames(coxDF)[cols] = paste('chr', gsub(':|-','.', grep('^\\d+', colnames(coxDF), value=T)), sep='')
-
-coxSM = cbind(coxDF[,c(1:3)], StandardizeMagnitude(coxDF[,-c(1:3)]))
-
-#surv = with(coxDF, Surv(tstart, tstop, Diagnosis))
-surv = with(coxSM, Surv(tstart, tstop, Diagnosis))
-groups = factor(c(rep('clinical',3),c(rep('genomic', length(grep('\\d+', colnames(coxDF)))))))
-rfx = CoxRFX(coxSM[,-c(1:3)], surv, groups, verbose=T, nu=1)
-summary(rfx)
-
-plot(rfx)
-
-wt = WaldTest(rfx)
-survConcordance(surv ~ predict(rfx))
-
-wt$label = sub('\\.', '-', sub('\\.', ':', sub('chr', '', rownames(wt))))
-wt = merge(wt, features[,c('label','hazards', 'coef','stability')], by='label', all=T)
-
-wt[is.na(wt$hazards), c('hazards', 'coef.y')] = 0
-plot.waldtest(wt) + labs(x='RFX coef',y='GLM coef', title='RFX vs GLM coefficients', subtitle='Labeled points have a significant p-value in RFX (except BE/ID)')
-
-options(scipen = 999)
-# arms only
-arms = with(arm.loc, paste(chrom, ':', start, '-', end, sep=''))
-coxDfArms = coxDF[,arms]
-colnames(coxDfArms) = paste('chr', gsub(':|-','.', grep('^\\d+', colnames(coxDfArms), value=T)), sep='')
-
-surv = with(coxDF, Surv(tstart, tstop, Diagnosis))
-groups = factor(c(rep('clinical',3), rep('genomic', ncol(coxDfArms))))
-
-rfx = CoxRFX( cbind.data.frame(coxSM[,c('ID','BE','age_10')], coxDfArms), surv, groups, verbose=T, nu=1)
-summary(rfx)
-
-plot(rfx)
-wt = WaldTest(rfx)
-survConcordance(surv ~ predict(rfx))
-
-```
+
+file = paste(cox.cache.dir, "all.pt.alpha.Rdata" , sep='/')
+load(file, verbose=T)
+
+do.call(grid.arrange, c(plots, nrow=2))
+
+file = paste0(cox.cache.dir, '/loo_0.9.Rdata')
+load(file, verbose=T)
+
+rocCox = pROC::roc(Status ~ FittedRR, data=pg.samp, auc=T, ci=T, of='thresholds')
+roc.plot(rocCox) + labs(title = 'Cox model ROC')
+
+first = pg.samp %>% group_by(Patient) %>% arrange(Endoscopy.Year) %>% filter(Endoscopy.Year == Endoscopy.Year[1])
+rocCox2 = pROC::roc(Status ~ FittedRR, data = first, auc = T, ci = T, of='thresholds')
+roc.plot(rocCox2) + labs(title = 'Cox model ROC, first endo')
+
+last = filter(pg.samp, Endoscopy.Year == Final.Endoscopy)
+rocCox3 = pROC::roc(Status ~ FittedRR, data = last, auc = T, ci = T, of='thresholds')
+roc.plot(rocCox3) + labs(title = 'Cox model ROC, last endo')
+
+rocPath = pROC::roc(Status ~ Pathology, data=pg.samp, auc=T, ci=T, of='thresholds')
+roc.plot(rocPath) + labs(title = 'Pathology status ROC')
+
+cor.test(pg.samp$FittedRR, as.integer(pg.samp$Pathology))
+
+```
+
+# Coefficients by # patients
+
+```{r, echo=F}
+# SNP
+snp.info = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/Analysis/SNP/metadata_T1T2.xlsx') %>% 
+  mutate( PID = paste(PatientID, `Timepoint Code`, sep='_'), Patient = as.numeric(PatientID), Hospital.Research.ID = as.character(Patient), Samplename = PID )
+
+model.dir = '~/Data/BarrettsProgressionRisk/Analysis/dv_snp_model/50kb/'
+load(list.files(model.dir, 'all.pt.alpha',recursive = T, full.names = T))
+
+dvsnp.coefs = coefs[['0.9']]
+dvsnp.perf = performance.at.1se[['0.9']]$mean
+rm(dysplasia.df, plots, pg.samp, models, nzcoefs, fits, performance.at.1se)
+
+dvsnp.preds = purrr::map(list.files('~/Data/BarrettsProgressionRisk/Analysis/dv_snp_model/50kb/loo', full.names = T), function(f) read_tsv(f, col_types = 'ccccccdd') )
+
+dvsnp.preds = do.call(bind_rows, dvsnp.preds)
+dvsnp.patient = length(unique(dvsnp.preds$Patient))
+dvsnp.roc = pROC::roc(Status ~ Probability, data=dvsnp.preds, ci=T, of='thresholds')
+
+dvsnp.dv.auc = pROC::auc(Status ~ Probability, data = dvsnp.preds %>% filter(cohort != 'snp') )
+dvsnp.snp.auc = pROC::auc(Status ~ Probability, data = dvsnp.preds %>% filter(cohort == 'snp') )
+
+dvsnp.preds = dvsnp.preds %>% mutate(type = ifelse(cohort == 'snp', 'snp', 'sWGS'))
+
+dvsnp.aucs = left_join(dvsnp.preds %>% dplyr::select(Patient, type) %>% distinct %>% group_by(type) %>% tally(name='# patients'),
+            dvsnp.preds %>% dplyr::select(Patient, type) %>% group_by(type) %>% tally(name='# timepoint samples'), by='type') %>% 
+  add_row( type = 'All (snp+sWGS)', `# patients` = dvsnp.preds %>% dplyr::select(Patient) %>% distinct %>% tally %>% pull, `# timepoint samples` = dvsnp.preds %>% dplyr::select(Patient) %>% tally %>% pull) %>%
+  arrange(type) %>% add_column( auc = c(dvsnp.roc$auc, dvsnp.snp.auc, dvsnp.dv.auc) ) %>%
+  mutate_at(vars(-auc), as.factor)
+
+dv.snp.p = ggplot(dvsnp.aucs , aes(`# patients`, auc, group=type, fill=type)) + geom_col() + 
+  geom_text(aes(label=round(auc,2), y=0.05)) + 
+  scale_y_continuous(expand=c(0,0), limits=c(0,1)) + scale_fill_brewer(palette = 'Set2', name='') + 
+  labs(title='sWGS (all) + SNP Model', y='AUC') + theme(legend.position = 'bottom') +
+  plot.theme
+dv.snp.p
+
+ggsave(filename='plots/performance/training_snpdv.png', plot=dv.snp.p, width=6, height=10, units='in', dpi=300)
+
+# discovery + validation
+model.dir = '~/Data/BarrettsProgressionRisk/Analysis/models_5e6_all/50kb/'
+load(list.files(model.dir, 'loo_0.9.Rdata',recursive = T, full.names = T))
+d.n.patient = length(unique(pg.samp$Patient))
+d.roc = pROC::roc(Status ~ Prediction, data = pg.samp, ci=T, of='thresholds')
+
+load(list.files(model.dir, 'all.pt.alpha',recursive = T, full.names = T))
+d.coefs = coefs[['0.9']]
+d.perf = performance.at.1se[['0.9']]$mean
+rm(dysplasia.df, plots, pg.samp, models, nzcoefs, fits, performance.at.1se)
+
+all.dir = '~/Data/BarrettsProgressionRisk/Analysis/dv_model_5e6_all_dvz/50kb/'
+load(list.files(all.dir, 'loo_0.9.Rdata',recursive = T, full.names = T))
+dv.n.patient = length(unique(pg.samp$Patient))
+dv.n.samples = length(unique(pg.samp$Samplename))
+dv.roc = pROC::roc(Status ~ Probability, data = pg.samp, ci=T, of='thresholds')
+
+load(list.files(all.dir, 'all.pt.alpha',recursive = T, full.names = T))
+dv.coefs = coefs[['0.9']]
+dv.perf = performance.at.1se[['0.9']]$mean
+rm(dysplasia.df, plots, pg.samp, models, nzcoefs, cvs, performance.at.1se)
+```
+
+```{r, fig.width=12, fig.height=8}
+cache.dir = '~/Data/BarrettsProgressionRisk/Analysis/models_5e6_all/'
+filenames = list.files(cache.dir, 'supp*')
+
+info = do.call(bind_rows, purrr::map(list.files(cache.dir, 'merged.Rdata', recursive = T, full.names = T), function(f) {
+  load(f,verbose=F)
+  model.info
+})) %>% mutate('Cohort' = 'Discovery (patient subsets)', Type = 'sWGS')
+
+info = add_row(info, n.patients = d.n.patient, cv.performance = d.perf, n.coefs = nrow(d.coefs), n.samples = 773, auc = d.roc$auc, Cohort = 'Discovery (all)', Type = 'sWGS')
+
+info = add_row(info, n.patients = dv.n.patient, cv.performance = dv.perf, n.coefs = nrow(dv.coefs), n.samples = dv.n.samples, auc = dv.roc$auc, Cohort = 'Discovery+Validation', Type = 'sWGS')
+
+info = add_row(info, n.patients = dvsnp.patient, cv.performance = dvsnp.perf, n.coefs = nrow(dvsnp.coefs), n.samples = nrow(dvsnp.preds), auc = dvsnp.roc$auc, Cohort = 'sWGS (all) + SNP', Type = 'All')
+
+info = add_row(info, n.patients = dvsnp.aucs %>% filter(type == 'snp') %>% dplyr::select(`# patients`) %>% pull, cv.performance = NA, n.coefs = NA, n.samples = NA, auc = dvsnp.aucs %>% filter(type == 'snp') %>% dplyr::select(auc) %>% pull, Cohort = 'sWGS (all) + SNP', Type = 'SNP')
+
+info = add_row(info, n.patients = dvsnp.aucs %>% filter(type == 'sWGS') %>% dplyr::select(`# patients`) %>% pull, cv.performance = NA, n.coefs = NA, n.samples = NA, auc = dvsnp.aucs %>% filter(type == 'sWGS') %>% dplyr::select(auc) %>% pull, Cohort = 'sWGS (all) + SNP', Type = 'sWGS')
+
+x = info %>% group_by(n.patients, Cohort, Type) %>% dplyr::summarise_at(vars(-n.samples), list(~mean(.), ~sd(.))) %>% as_tibble %>% mutate(n.patients = factor(n.patients,levels=c( seq(28,88,10), 164, 413, 165, 248  ), ordered=T), Cohort = factor(Cohort, levels=c('Discovery (patient subsets)', 'Discovery (all)', 'Discovery+Validation', 'sWGS (all) + SNP'), ordered=T))
+
+p = ggplot( x %>% filter(!is.na(cv.performance_mean)), aes(x=n.patients, y=cv.performance_mean, group=n.patients, fill=Cohort)) + 
+  facet_grid(~Cohort, scales = 'free_x', space='free_x') + 
+  geom_col() + geom_errorbar(aes(ymin=cv.performance_mean-cv.performance_sd/2, ymax=cv.performance_mean+cv.performance_sd/2), width=0.2, color='grey39') +
+  geom_text(aes(label=round(cv.performance_mean,2)), y=0.05) + 
+  scale_y_continuous(expand=c(0,0), limits=c(0,0.8)) + 
+  scale_fill_brewer(palette = 'Set2',name='') + 
+  plot.theme + theme(legend.position = 'bottom', strip.text = element_blank(), panel.spacing = unit(0.5, 'lines')) +
+  labs(title='Cross validation performance', y='', x='# patients')
+
+legend = cowplot::get_legend(p)
+p = p + theme(legend.position = 'none')
+
+p1 = ggplot(x %>% filter(!is.na(n.coefs_mean)), aes(x=n.patients, y=n.coefs_mean, group=n.patients, fill=Cohort)) + 
+  facet_grid(~Cohort, scales = 'free_x', space='free_x') + 
+  geom_col() + geom_errorbar(aes(ymin=n.coefs_mean-n.coefs_sd/2, ymax=n.coefs_mean+n.coefs_sd/2), width=0.2, color='grey39') +
+  geom_text(aes(label=round(n.coefs_mean,2)), y=5) +
+  scale_y_continuous(expand=c(0,0)) + scale_fill_brewer(palette = 'Set2') + 
+  plot.theme + theme(legend.position = 'none', strip.text = element_blank(), panel.spacing = unit(0.5, 'lines')) +
+  labs(title='# Coefficients',y='', x='# patients')
+
+p2 = ggplot(x, aes(x=n.patients, y=auc_mean, group=n.patients, fill=Cohort)) +
+  facet_grid(~Cohort, scales = 'free_x', space='free_x') + 
+  geom_col() + 
+  geom_errorbar(aes(ymin=auc_mean-auc_sd/2, ymax=auc_mean+auc_sd/2), width=0.2, color='grey39') +
+  geom_text(aes(label=round(auc_mean,2)), y=0.05) +
+  geom_text( data=filter(x,Cohort == 'sWGS (all) + SNP'), aes(label=Type, y=0.1) ) +
+  scale_y_continuous(expand=c(0,0), limits=c(0,1)) + scale_fill_brewer(palette = 'Set2') + 
+  plot.theme + theme(legend.position = 'none', strip.text = element_blank(), panel.spacing = unit(0.5, 'lines')) +
+  labs(title='AUC', y='', x='# patients')
+
+grid.arrange(arrangeGrob(p,p1,p2,ncol=1), arrangeGrob(legend), heights=c(4,1))
+
+ggsave(filename='plots/performance/training_n_pts.png',
+       plot=grid.arrange(arrangeGrob(p,p1,p2,ncol=1), arrangeGrob(legend), heights=c(6,1)), width=8, height=12, units='in', dpi=300)
+
+```

### 2020-01-21 lib/common_plots.R

@@ -130,14 +130,17 @@ multi.roc.plot <- function(rocList, title="ROC", palette=NULL, colors=NULL) {
   }
   
   aucs$label=with(aucs, paste(model, '\nAUC ', round(AUC,2), '\nTPR=', round(Sensitivity,2), ' FPR=', round((1-Specificity),2), sep=''))
+
+  aucs$label=with(aucs, paste(model, '\nAUC ', round(AUC,2), sep=''))
   
+    
   df = do.call(rbind, lapply(rocList, function(r) 
     cbind.data.frame('Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'model'=r$model) 
   ))
 
   p = ggplot(df, aes(Specificity, Sensitivity,color=model,fill=model)) +
         geom_line(show.legend = F, size=1.5) + scale_x_reverse() +
-        geom_label(data=aucs, aes(x=x, y=y, label=label), color='white', show.legend = F) +
+        geom_label_repel(data=aucs, aes(x=x, y=y, label=label), color='white', show.legend = F) +
         labs(title=title, x='Specificity (1-FPR)', y='Sensitivity (TPR)')  + plot.theme
   if (is.null(colors) & !is.null(palette)) {
     colors = RColorBrewer::brewer.pal(length(rocList)+1, palette)
@@ -385,7 +388,7 @@ plot.risk.by.path<-function(preds) {
 }
 
 
-transform.by.time<-function(pt, info) {
+transform.by.time<-function(pt) {
   pt = pt %>% dplyr::group_by(PID, Samplename) %>%  
     dplyr::mutate( 'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12) %>% 
     arrange(Pathology, PID, Endoscopy.Year) %>% ungroup 

### 2020-02-14 Paper.Rmd
updating for paper

@@
 library(RColorBrewer)
+library(ggfortify)
 
@@
 cohortSummary %>% t %>% kable(caption='Training set') %>% kable_styling()
+
+# fit = glm(Status ~ age.diagnosis+M+gender, data=sum.patient.data, family = binomial(link='logit'))
+# summary(fit)
+# 
+# sum.patient.data = sum.patient.data %>% mutate(total.followup = end.year - start.year)
+# 
+# surv =  Surv(sum.patient.data$total.followup,sum.patient.data$Status)
+# terms = as.formula(paste("Surv(total.followup, status) ~", paste(c('age.diagnosis','M','gender'),collapse=" + "), sep=""))
+# 
+# cox = coxph(terms, data=sum.patient.data, na.action=na.omit,ties='breslow')
+# summary(cox)
+# 
+# ggsurv(survfit(cox),cens.shape=3,lty.est=1,back.white=F, CI=T ) +
+#       annotate("text", x=5, y=0.25,label=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),3),"\n", nrow(data), 'patients')) + theme_minimal()
+# 
+# 
+# t.test(filter(sum.patient.data, Status == 'NP')$age.diagnosis,
+#        filter(sum.patient.data, Status == 'P')$age.diagnosis)
+# t.test(filter(sum.patient.data, Status == 'NP')$M, filter(sum.patient.data, Status == 'P')$M)
+# fisher.test(table(sum.patient.data$Status, sum.patient.data$smoking))
+# fisher.test(table(sum.patient.data$Status, sum.patient.data$gender))
+
 ```
@@
 load(file, verbose=F)
-#dim(dysplasia.df)
-#length(labels)
+melted = melt(dysplasia.df)
+
+#melted$brks = as.integer(factor(cut(melted$value, breaks = quantile(melted$value, probs=c(seq(0,1,0.05))), include.lowest = T)))
+p = ggplot(melted, aes(Var2, Var1, fill = value)) + geom_tile() +
+  scale_fill_gradient2() + 
+  theme(axis.text = element_blank()) + labs(x='Feature', y='sWGS Sample')
+leg = get_legend(p)
+ggsave('plots/samples.png', plot = p + theme(legend.position = 'none'), width=3, height=4, units='in')
 
@@
 ```{r echo=F, fig.height=5, fig.width=5}
-library(ggfortify)
+
 autoplot(models$`0.9`) + theme(legend.position = 'none') + plot.theme
@@
 features = as_tibble(hazards, rownames = 'label') %>% dplyr::rename(hazards = value) %>% 
-  dplyr::mutate(coef = coefs[[select.alpha]][,1], stability = featureStability) %>% arrange(-hazards)
+  dplyr::mutate(coef = coefs[[select.alpha]][,1], stability = featureStability) #%>% arrange(-hazards)
+
+p = ggplot(features, aes(label,hazards, fill=coef)) + geom_col() + 
+  scale_fill_gradient2() + coord_flip() + theme_classic() + 
+  theme(axis.ticks.y = element_blank(), panel.grid = element_blank(), axis.text = element_blank(), legend.position = 'none') + labs(x='', y='', title='Effects') 
+ggsave('plots/effects.png', plot=p, width = 3, height=1.5, units='in')
 
@@
 
-error = read_tsv('~/Data/BarrettsProgressionRisk/Analysis/pcf_perPatient/50kb/sample_error.tsv')
+error = read_tsv('~/Data/BarrettsProgressionRisk/Analysis/pcf_perPatient/50kb/sample_error.tsv', col_types = 'cdD')
 
@@
 
-error %>% group_by(Status) %>% dplyr::summarise_at(vars(Abs.Error), list(~mean, ~sd))
-summary(error$Abs.Error)
-sd(error$Abs.Error)
+#error %>% group_by(Status) %>% dplyr::summarise_at(vars(Abs.Error), list(~mean(.,na.rm=T), ~sd(.,na.rm=T)))
+#summary(error$Abs.Error)
+#sd(error$Abs.Error)
+
+p = ggplot(dplyr::select(pg.samp, Status, Samplename, RR), aes(reorder(Samplename, -RR), 1, fill=RR)) + geom_tile() + 
+  #scale_fill_gradientn(colors=myPal) +
+  scale_fill_gradient2(midpoint = 5, low = scales::muted('blue'), high = scales::muted('red')) + 
+  scale_y_continuous(expand=c(0,0)) + 
+  coord_flip() + labs(y='', x='') + theme_bw() + 
+  theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), legend.position = 'none') 
+ggsave('plots/per_sample_RR.png', plot = p, width=0.5, height=3, units='in')
 
@@
 rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')
-
+roc.plot(rocNOHGD) + ggtitle('Model excluding HGD/IMC')
+
+rocNOHGD$auc
+pROC::ci.auc(rocNOHGD$auc)
+
+ggsave(filename='plots/auc/roc_nohgd_model.png', plot=roc.plot(rocNOHGD) + ggtitle('Model excluding HGD/IMC'), width=6, height=6, units='in', dpi=300)
 ```
@@
 droc
-ggsave(filename='plots/auc/discovery_roc.png', plot=droc, width=6, height=4, units='in', dpi=300)
+#ggsave(filename='plots/auc/discovery_roc.png', plot=droc, width=6, height=4, units='in', dpi=300)
 
@@
 #summary(filter(df, model == 'in-silico')$AUC)
-
-getPalette = colorRampPalette(brewer.pal(5, "Blues"))
-p = ggplot(df, aes(Specificity, Sensitivity,color=is)) + 
-  geom_line(linetype='solid', size=0.5) + 
-  geom_line(data=subset(df, is == 'discovery'), color='red', size=1.5) +
-  scale_x_reverse() +
-  scale_color_manual(values=getPalette(length(unique(df$is)))) +
-  labs(title='ROC, all splits', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
-  theme(legend.position = 'none')
-p
 
@@
 # Not sure why I did this
-df = cbind.data.frame('specificity'=rev(roc$specificities), 'sensitivity'=rev(roc$sensitivities))
-thresholds = 
-  rbind.data.frame(
-    est=t(round(pROC::coords(roc, "best"),2)),
-    max=t(round(pROC::coords(roc, 0.8),2)),
-    'No HGD'=t(round(pROC::coords(rocNOHGD, 0.6),2)),
-    'All samples'=t(round(pROC::coords(roc, 0.6),2))
-)
-thresholds$label=rownames(thresholds)
-
-ggplot() + 
-  geom_line(data=df, aes(specificity, sensitivity), color='darkblue', size=1.5) + scale_x_reverse() +
-  geom_point(data=thresholds, aes(x=specificity,y=sensitivity,color=label), size=3,shape=8, show.legend = T) +
-  geom_label_repel(data=thresholds, aes(x=specificity,y=sensitivity,fill=label,label=
-    paste('Threshold ', round(threshold,2), '\n(FPR ', round((1-specificity),2)*100, '%, TPR ', round(sensitivity,2)*100,'%)\nAUC:',round(roc$auc, 2)*100, '%', sep='')),show.legend = F, color='white',nudge_x=0.05) + scale_fill_manual(values=brewer.pal(4,'Dark2')) + scale_color_manual(values=brewer.pal(4,'Dark2')) +
-  labs(title='CN Model ROC', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme
+# df = cbind.data.frame('specificity'=rev(roc$specificities), 'sensitivity'=rev(roc$sensitivities))
+# thresholds = 
+#   rbind.data.frame(
+#     est=t(round(pROC::coords(roc, "best", transpose = T),2)),
+#     max=t(round(pROC::coords(roc, 0.8, transpose = T),2)),
+#     'No HGD'=t(round(pROC::coords(rocNOHGD, 0.6, transpose = T),2)),
+#     'All samples'=t(round(pROC::coords(roc, 0.6, transpose = T),2))
+# )
+# thresholds$label=rownames(thresholds)
+# 
+# ggplot() + 
+#   geom_line(data=df, aes(specificity, sensitivity), color='darkblue', size=1.5) + scale_x_reverse() +
+#   geom_point(data=thresholds, aes(x=specificity,y=sensitivity,color=label), size=3,shape=8, show.legend = T) +
+#   geom_label_repel(data=thresholds, aes(x=specificity,y=sensitivity,fill=label,label=
+#     paste('Threshold ', round(threshold,2), '\n(FPR ', round((1-specificity),2)*100, '%, TPR ', round(sensitivity,2)*100,'%)\nAUC:',round(roc$auc, 2)*100, '%', sep='')),show.legend = F, color='white',nudge_x=0.05) + scale_fill_manual(values=brewer.pal(4,'Dark2')) + scale_color_manual(values=brewer.pal(4,'Dark2')) +
+#   labs(title='CN Model ROC', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme
 
@@
   bind_rows(
-    pROC::coords(rocEndoM, 'best'),
-    pROC::coords(rocEndoA, 'best'),
-    pROC::coords(rocLastEndo, 'best'),
-    pROC::coords(rocFirstEndo, 'best'),
-    pROC::coords(rocRandEndo, 'best')
+    pROC::coords(rocEndoM, 'best', transpose = T),
+    pROC::coords(rocEndoA, 'best', transpose = T),
+    pROC::coords(rocLastEndo, 'best', transpose = T),
+    pROC::coords(rocFirstEndo, 'best', transpose = T),
+    pROC::coords(rocRandEndo, 'best', transpose = T)
   ))
@@
 ```{r}
-pg.samp = do.call(bind_rows,lapply(unique(pg.samp$Patient) %>% 
-                                     map( function(p) pg.samp %>% filter(Patient == p)), transform.by.time, info=patient.info)) 
+pg.samp = do.call(bind_rows,lapply(unique(pg.samp$Patient) %>% map( function(p) pg.samp %>% filter(Patient == p)), transform.by.time)) 
 sq = c(0, seq(1,132,12),max(pg.samp$months.before.final))
@@
 rp = plot.risk.by.path(pg.samp %>% ungroup) + labs(title='Samples classified by pathology', y='% Classified')
-
+rp
 ggsave(filename='plots/pred_by_path.png', plot=(rp + theme(legend.position = 'none')), width=7, height=6, units='in', dpi=300)
@@
 
-```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=6, eval=F}
-all.preds = pg.samp %>% dplyr::group_by(Status, Patient, Risk) %>% tally %>% spread(Risk, n) %>% ungroup %>%
-  mutate_if(is.numeric, funs( ifelse(is.na(.), 0, .) )) %>% 
-  left_join(pg.samp %>% dplyr::group_by(Status, Patient) %>% tally, by=c('Status','Patient')) %>%
-  dplyr::rename('n.samples'='n') %>% 
-  dplyr::mutate(Patient = factor(Patient))
-
-all.preds = melt(all.preds, id.vars=c('Status','Patient', 'n.samples'))
-
-ggplot(all.preds, aes(Patient, value/n.samples, group=variable, fill=variable)) + 
-  geom_bar(stat='identity', position = position_stack()) +
-  scale_fill_manual(values=riskCols) + geom_text(aes(y=1.01, label=n.samples)) +
-  facet_wrap(~Status, scales='free_x') + labs(y='Sample ratios', title='Per patient ratio of predictions') + 
-  plot.theme
-```
-
 ## Backwards from HGD
@@
 
+dir.create('plots/heatmaps/P',recursive = T, showWarnings = F)
+dir.create('plots/heatmaps/NP',recursive = T, showWarnings = F)
 dir.create('plots/heatmaps/tiny/P',recursive = T, showWarnings = F)
@@
 
+val.demo = readxl::read_xlsx(patient.file, 'Demographics') %>% 
+  mutate_at(vars(Status,Sex), list(~factor(.))) %>%
+  mutate(Followup = as.double(difftime(`Last Endoscopy`, `Initial Diagnosis`, units='weeks'))/52)
+
+no_followup = filter(val.demo, Followup == 0 & Status == 'NP')
+val.info = val.info %>% filter(Patient !=  no_followup$Patient)
+
+val.demo = val.demo %>% mutate_at(vars(Sex,Smoking), list(~forcats::fct_explicit_na(.)))
+
+val.demo %>% filter(Patient != no_followup$Patient) %>% group_by(Status) %>% tally %>% spread(Status, n)
+val.demo %>% filter(Patient != no_followup$Patient) %>% group_by(Status,Sex) %>% tally %>% spread(Status, n)
+
+val.demo %>% filter(Patient != no_followup$Patient) %>% group_by(Status,Smoking) %>% tally %>% spread(Status, n)
+
+val.demo %>% filter(Patient != no_followup$Patient) %>% group_by(Status) %>% dplyr::summarise_at(vars(Maximal), list(~mean(.,na.rm=T), ~sd(.,na.rm=T)))
+
+val.demo %>% filter(Patient != no_followup$Patient) %>% group_by(Status) %>% dplyr::summarise_at(vars(`Age at diagnosis`), list(~mean(.,na.rm=T), ~sd(.,na.rm=T)))
+
+val.demo %>% filter(Patient != no_followup$Patient) %>% group_by(Status) %>% dplyr::summarise_at(vars(Followup), list(~mean(.,na.rm=T), ~sd(.,na.rm=T))) 
+val.demo %>% filter(Patient != no_followup$Patient) %>% group_by(Status) %>% dplyr::summarise_at(vars(Followup), list(~min(.,na.rm=T), ~max(.,na.rm=T)))
+
+wilcox.test(filter(val.demo, Status == 'NP')$`Age at diagnosis`, filter(val.demo, Status == 'P')$`Age at diagnosis`)
+wilcox.test(filter(val.demo, Status == 'NP')$Maximal, filter(val.demo, Status == 'P')$Maximal)
+
+fisher.test(table(val.demo$Status, val.demo$Sex))
+fisher.test(table(val.demo$Status, val.demo$Smoking))
+
+# samples
+val.info %>% group_by(Status) %>% tally
+
 val.info$Block = sapply(strsplit(val.info$Block, ''), function(x) {
@@
 
-val.info
-
 vdir ='~/Data/BarrettsProgressionRisk/Analysis/validation/pcf_perPatient/50kb'
@@
 failedQC = filter(residuals, varMAD_median > 0.008)
+
 ```
@@
 })) %>% mutate(quants = cut(Probability, breaks=cuts, include.lowest = T)) %>%
-  left_join(pred.confidence %>% dplyr::select(quants, Risk), by='quants')
+  left_join(pred.confidence %>% dplyr::select(quants, Risk), by='quants') %>%
+  filter(`Hospital Research ID` != no_followup$`Hospital Research ID`)
 
@@
 
+# samples passed QC
+val.preds %>% group_by(Status) %>% tally
+
 ```
@@
 pROC::ci.auc(roc$auc)
-
 ggsave('plots/validation/validatation_roc.png', plot=roc.plot(roc) + labs(title='Validation cohort'), width=5, height=5, units='in',dpi=300)
@@
     labs(title='Validation risk (sample ratios per NP:P)', x='', y='ratio') + theme_minimal()
+
 ```
@@
 val.preds = do.call(bind_rows, 
-                    lapply(unique(val.preds$Patient) %>% 
-                            map( function(p) val.preds %>% filter(Patient == p) %>% 
+                    lapply(unique(val.preds$Patient) %>% map( function(p) val.preds %>% filter(Patient == p) %>% 
                                    mutate_at(vars(Endoscopy.Year, Final.Endoscopy), list(~get.year(.)) )
-                                 ), transform.by.time, info=val.info) )
+                                 ), transform.by.time) )
 
@@
 
-```{r, fig.width=6, fig.height=8}
+```{r, fig.width=6, fig.height=8, eval=F}
 require(survival)
@@
 
-```{r}
+```{r, eval=F}
 vp = val.preds %>% dplyr::select(Patient, Status, months.before.final, Risk) %>% 
@@
 
-merged %>%
+#merged %>%
 
@@
 
-kmp = 
-  survminer::ggsurvplot(sfit, size=1, linetype="strata",risk.table=F, risk.table.col='strata',
+kmp = survminer::ggsurvplot(sfit, size=1, linetype="strata",risk.table=F, risk.table.col='strata',
                       conf.int=T, censor=T, legend='none',# xlim=c(0,150),
@@
 kmp
-
 ggsave('plots/risk_km.png', plot=print(kmp), height=8, width=7, units = 'in', dpi = 300)
@@
 
+#pts = as.character(pg.samp %>% group_by(Patient, Endoscopy.Year) %>% tally %>% filter(n > 1) %>% dplyr::select(Patient) %>% distinct %>% pull)
+
 max.rules = purrr::map(unique(pg.samp$Patient), function(pt) {
+
     preds = pg.samp %>% filter(Patient == pt) %>% arrange(Endoscopy.Year, months.before.final)
-    endo = BarrettsProgressionRisk:::.setUpRxTablePerEndo(preds %>% dplyr::mutate(Endoscopy = months.before.final, Sample = 'Samplename'), 'max', verbose=F) %>% mutate(Endoscopy = Endoscopy*-1)
+
+    endo = BarrettsProgressionRisk:::.setUpRxTablePerEndo(preds %>% dplyr::mutate(Endoscopy = months.before.final, Sample = 'Samplename'), 'max', verbose=F) %>%
+      mutate(Endoscopy = as.integer(factor(Endoscopy*-1)))
+
+    if (nrow(endo) == 1) return(NA)
+
     rx = BarrettsProgressionRisk:::.apply.rules(endo, 'Endoscopy', pred.confidence,'numeric' )
@@
 names(max.rules) = unique(pg.samp$Patient)
+max.rules = max.rules[which(sapply(max.rules, is.data.frame))]
 
@@
 
-pids = names(which(sapply(progressors, function(x) {
-  nrow(x) >= 1 & grepl('Endpoint', x[nrow(x),'Time 2'])
-})))
-
-nids = names(which(sapply(nonprog, function(x) {
-  nrow(x) >= 1 & grepl('Endpoint', x[nrow(x),'Time 2'])
-})))
+pids = names(which(unlist(  sapply(progressors, function(x) { nrow(x) >= 1 & grepl('Endpoint', x[nrow(x),'Time 2']) }) )))
+nids = names(which(unlist(  sapply(nonprog, function(x) { nrow(x) >= 1 & grepl('Endpoint', x[nrow(x),'Time 2']) }) )))
 
@@
 ```{r}
-pids = names(which(sapply(progressors, function(x) {
+pids = names(which(unlist(sapply(progressors, function(x) {
   nrow(x) > 1 & !grepl('Endpoint', x[2,'Time 2']) & !grepl('None', x[2,'Rule'])
-})))
-
-nids = names(which(sapply(nonprog, function(x) {
+}))))
+
+nids = names(which(unlist(sapply(nonprog, function(x) {
   nrow(x) > 1 & !grepl('Endpoint', x[2,'Time 2']) & !grepl('None', x[2,'Rule'])
-})))
+}))))
 
@@
 
-second %>% dplyr::select(Patient, Status) %>% distinct %>% group_by(Status) %>% tally %>% kable(caption='n patients with at least two years follow-up') %>%
+second %>% dplyr::select(Patient, Status) %>% distinct %>% group_by(Status) %>% tally %>% 
+  kable(caption='n patients with at least two years follow-up') %>%
   kable_styling(full_width = F)
@@
   kable(caption='Second Endoscopy: # patients by rx rules based on the max (P) per endoscopy') %>% kable_styling(full_width=F)
+
+p = second %>% group_by(Status, Rule) %>% tally %>% spread(Rule,n) %>% dplyr::rename(RFA = `1`, `6-12 month recheck` = `2`, `12-24 month recheck` = `3`, `3-5 year surveillance` = `4`) %>% melt(id.vars = 'Status') %>% mutate(Status = recode(Status, NP = 'Non-Progressor' ,P = 'Progressor')) %>%
+  ggplot(aes(variable, value, group = Status)) + 
+  geom_col(aes(fill=Status), position = position_dodge()) + geom_text(aes(label=paste0('n=',value)), position = position_dodge(width = 1)) + coord_flip() +
+  scale_fill_manual(values = c('#4575B4', '#A50026'), name='' ) + plot.theme + labs(x='', y='# Patients', title='Second Endoscopy Rx') + theme(legend.position = 'bottom')
+p
+ggsave('plots/second_endo_rx.png', plot=p, width=6, height=4, units = 'in', dpi = 300)
 

### 2020-02-24 Paper.Rmd
improving plots

@@
 author: "Sarah Killcoyne"
-date: "2019-11-13"
+date: "`r format(Sys.time(), '%d %B %Y')`"
 output: 
@@
 ```{r p53roc, echo=F, fig.height=2, fig.width=2}
-p53samples = subset(patient.info, !is.na(p53.Status), select=c('Samplename','p53.Status','Status'))
+p53samples = subset(patient.info, !is.na(p53.Status), select=c('Samplename','p53.Status','Status', 'Pathology'))
 p53roc = pROC::roc(as.numeric(p53samples$Status)-1, as.integer(p53samples$p53.Status)-1)
@@
 ```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
-p53Path = pg.samp %>% dplyr::filter(Status == 'P' & !is.na(p53.Status)) %>% 
-  dplyr::group_by(Pathology,p53.Status) %>% 
-  dplyr::summarise(Freq = length(Pathology)) %>% dplyr::mutate(Status = 'p53 IHC')
-
-p53Path = full_join( pg.samp %>% dplyr::filter(Status == 'P' & !is.na(p53.Status)) %>% 
-                       dplyr::group_by(Pathology) %>% 
-                       dplyr::summarise(nPath = length(Pathology)), p53Path, by='Pathology') %>% 
-  dplyr::mutate(ratio = round(Freq/nPath,2))
-
 rp = plot.risk.by.path(pg.samp %>% ungroup) + labs(title='Samples classified by pathology', y='% Classified')
@@
 ```{r p53IHC, echo=F, warning=F, message=F, fig.height=8, fig.width=8}
-#p53Path$Risk[p53Path$Risk == 'High'] = 'Aberrant'
-#p53Path$Risk[p53Path$Risk == 'Low'] = 'Normal'
-
-#p53Path$p53.Status = factor(p53Path$p53.Status, levels=c('Aberrant', 'Normal'), ordered = F)
+p53Path = pg.samp %>% dplyr::filter(Status == 'P' & !is.na(p53.Status)) %>% 
+  dplyr::group_by(Pathology,p53.Status) %>% tally(name='Freq') %>% ungroup %>% 
+  group_by(Pathology) %>% dplyr::mutate(nPath = sum(Freq), Status = 'p53 IHC', ratio = round(Freq/nPath,2))
 
 p53Path$p53.Status = factor(p53Path$p53.Status, ordered = T)
-p53Path =arrange(p53Path, Pathology, p53.Status)
+p53Path = arrange(p53Path, Pathology, p53.Status)
 col = brewer.pal(11, "PRGn")[c(10,2)]
@@
   geom_text(aes(label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5,color='white') +
-  plot.theme + labs(title='Samples predicted by aberrant p53 IHC', y='% Predicted', x='Pathology') + 
+  plot.theme + labs(title='p53 IHC in samples from progressive patients', y='% samples', x='Pathology') + 
   theme(legend.position = 'bottom', text = element_text(size=14))
@@
 
-    p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final','rev') + theme(legend.position = 'none') + labs(x='months',y='location') + theme(plot.title = element_text(face = 'bold', size=16))
-
-    ggsave(paste0('plots/heatmaps/',unique(patient$Status),'/pt_',pt,'.png'), plot = p, height=length(unique(patient$GEJ.Distance))+1, width=length(unique(patient$months.before.final))+1, units='in', dpi=300, limitsize=F)
+    p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final','rev', 0.5) + theme(legend.position = 'none') + labs(x='months',y='location') + theme(text = element_text(size = 10), plot.title = element_text(face = 'bold', size=16), axis.text = element_text(size = 10))
+
+      h = (length(unique(patient$GEJ.Distance))+1)*0.5
+  if (length(unique(patient$GEJ.Distance)) <= 1) h = (length(unique(patient$GEJ.Distance))+1)*0.8
+  w = (length(unique(patient$months.before.final))+1)*0.5
+  if (length(unique(patient$months.before.final)) <= 1) w = (length(unique(patient$months.before.final))+1)*0.8
+
+    ggsave(paste0('plots/heatmaps/',unique(patient$Status),'/pt_',pt,'.png'), plot = p, height=h, width=w, units='in', dpi=300, limitsize=F)
 
@@
 
-  p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final', 'rev') + labs(x='months', y='location') + theme(legend.position = 'none',plot.title = element_text(face = 'bold', size=16))
-
-    ggsave(paste0('plots/validation/heatmaps/',unique(patient$Status),'/pt_',pt,'.png'), plot = p, height=length(unique(patient$GEJ.Distance))+1, width=length(unique(patient$months.before.final))+1, units='in', dpi=300, limitsize=F)
+  p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final', 'rev', 0.5) + labs(x='months', y='location') + theme(legend.position = 'none',plot.title = element_text(face = 'bold', size=16), text = element_text(size=12), axis.text = element_text(size=12))
+
+    h = (length(unique(patient$GEJ.Distance))+1)*0.8
+  if (length(unique(patient$GEJ.Distance)) <= 1) h = (length(unique(patient$GEJ.Distance))+1)
+  w = (length(unique(patient$months.before.final))+1)*0.8
+  if (length(unique(patient$months.before.final)) <= 1) w = (length(unique(patient$months.before.final))+1)
+
+    ggsave(paste0('plots/validation/heatmaps/',unique(patient$Status),'/pt_',pt,'.png'), plot = p, height=h, width=w, units='in', dpi=300, limitsize=F) 
 }
@@
 
+# CoxHD with clinical factors
+```{r, eval=F}
+library(CoxHD)
+
+time<-function(x) {
+  for (i in 1:nrow(x)) {
+    if (i == 1) {
+      x[i,'tstart'] = 0
+      x[i, 'tstop'] = x[(i+1), 'Endoscopy.Year'] - x[i, 'Endoscopy.Year']
+      if (is.na(x[i,'tstop']) || x[i,'tstop'] == 0) x[i,'tstop'] = 0.5
+      next
+    }
+
+    x[i,'tstart'] = x[(i-1), 'tstop']
+
+    # Final sample is different -- will leave it out later
+    if (i == nrow(x)) {
+      x[i, 'tstop'] = x[i,'tstart'] + 0.5
+      break
+    }
+
+    x[i, 'tstop'] =  x[i,'tstart'] + x[(i+1), 'Endoscopy.Year'] - x[i, 'Endoscopy.Year']
+
+    # Samples taken in the same year
+    if ( x[i, 'Endoscopy.Year'] == x[(i+1), 'Endoscopy.Year'] )
+      x[i,'tstop'] = x[i,'tstop'] + 0.5 
+  }
+  return(x[,c('tstart','tstop')])
+}
+
+coxDF = do.call(bind_rows, purrr::map(unique(pg.samp$Patient), function(pt) {
+  df = pg.samp %>% filter(Patient == pt) %>% arrange(Endoscopy.Year)
+  df %>% bind_cols(time(df))
+})) %>%  dplyr::select(Samplename, Patient, Status, Risk, Prediction, Pathology, tstart, tstop, Age.at.diagnosis, Smoking, Sex, Maximal) %>% 
+  mutate(Diagnosis = as.integer(Pathology %in% c('HGD','IMC')) ) %>%  dplyr::select(Samplename, Patient, Status, tstart, tstop, Diagnosis, everything() )
+
+coxDF = bind_cols(coxDF, as_tibble(dysplasia.df[coxDF$Samplename,features$label])) %>% 
+  rename_at( vars(matches('\\d+:')), list(~paste0('chr',gsub(':|-','.', .))) ) %>% 
+  mutate(Smoking = as.integer(Smoking)-1, Sex = as.integer(Sex)-1, Pathology = as.integer(Pathology)-1) %>% ungroup
+
+coxSM = CoxHD::StandardizeMagnitude(dplyr::select(coxDF, -matches('Samplename|Patient|Status|Risk|Prediction|Pathology|tstart|tstop|Diagnosis')))
+coxSM = bind_cols(dplyr::select(coxDF, Samplename, tstart, tstop, Diagnosis), as_tibble(coxSM))
+
+surv = with(coxSM, Surv(tstart, tstop, Diagnosis))
+
+groups = factor(c(rep('clinical',4),c(rep('genomic', nrow(features)))))
+rfx = CoxRFX(as.data.frame(coxSM[,-c(1:4)]), surv, groups, verbose=T, nu=1)
+summary(rfx)
+
+plot(rfx)
+
+wt = WaldTest(rfx)
+survConcordance(surv ~ predict(rfx))
+
+wt$label = sub('\\.', '-', sub('\\.', ':', sub('chr', '', rownames(wt))))
+wt = merge(wt, features[,c('label','hazards', 'coef','stability')], by='label', all=T)
+
+wt[is.na(wt$hazards), c('hazards', 'coef.y')] = 0
+plot.waldtest(wt) + labs(x='RFX coef',y='GLM coef', title='RFX vs GLM coefficients', subtitle='Labeled points have a significant p-value in RFX (except BE/ID)')
+
+#gcols = colnames(coxDF)
+#cols = features$label
+
+coxSM = cbind(coxDF[,c(1:3)], StandardizeMagnitude(coxDF[,-c(1:3)]))
+
+```
+
 # Check COX

### 2020-04-14 Paper.Rmd
Bug fixes for paper figures

@@
 
-all.patient.info = read.patient.info(patient.file, demo.file, set='All')$info
-patient.info = all.patient.info
+all.patient.info = read.patient.info(patient.file, demo.file, set='All')
+patient.info = bind_rows(all.patient.info$info, all.patient.info$replicates) # for the moment...
 
@@
 
-```{r}
-
-cohortSummary = sum.patient.data %>% dplyr::group_by(Status) %>% dplyr::summarise(
+```{r demo}
+
+cohortSummary = sum.patient.data %>% dplyr::group_by(Status) %>% 
+  dplyr::summarise(
   'Total Patients'=length(unique(Patient)),
   'Total Samples'=sum(total.samples),
+  'Range follow-up (dates)'=paste(min(start.year), max(end.year), sep='-'),
   'Range follow-up (years)'=paste(range(end.year-start.year), collapse='-'),
-  'Mean follow-up (years)'=paste(round(mean(end.year-start.year), 1),'+/-',round(sd(end.year-start.year),1)),
-  'Range endoscopies'=paste(range(total.endos), collapse='-'),
-  'Mean Endoscopies'=paste(round(mean(total.endos), 1),'+/-',round(sd(total.endos),1)),
-  'Mean age'=paste(round(mean(age.diagnosis), 1),'+/-',round(sd(age.diagnosis),1)),
-  'gender F:M'=paste(table(gender),collapse=':'),
-  'Mean Length'=paste(round(mean(C,na.rm=T), 1),'+/-',round(sd(C,na.rm=T),1)),
-  'naC'=length(which(is.na(C)))
+  #'Mean follow-up (years)'=paste(round(mean(end.year-start.year), 1),'+/-',round(sd(end.year-start.year),1)),
+  #'Total Range endoscopies'=paste(range(total.endos), collapse='-'),
+  #'Mean Endoscopies'=paste(round(mean(total.endos), 1),'+/-',round(sd(total.endos),1)),
+  'Median total endoscopies (IQR)'=paste0(median(total.endos), ' (', paste(quantile(total.endos, prob=c(0.25,0.75)), collapse='-'), ')'),
+  'Median age at diagnosis (IQR)'=paste0(median(age.diagnosis), ' (', paste(quantile(age.diagnosis, prob=c(0.25,0.75)), collapse='-'), ')'),
+  'Mean Length (not recorded)'=paste0(paste(round(mean(M,na.rm=T), 1),'+/-',round(sd(M,na.rm=T),1)), ' (', length( which(is.na(M) )),')'),
+  'Gender F:M'=paste(table(gender),collapse=':'),
+  'Smoking N:Y (not recorded)'=paste0(paste(table(smoking), collapse=':'), ' (', length(which(is.na(smoking) )), ')')
+) %>% left_join(patient.info %>% group_by(Status, Pathology) %>% tally %>% spread(Pathology,n), by='Status')
+
+cohortSummary = cohortSummary %>% add_row(
+  Status = 'p-value (test)', 'Total Patients'= '', 'Total Samples' = '', 'Range follow-up (dates)' = '', 
+  'Range follow-up (years)' = paste0(signif(do(sum.patient.data, tidy(wilcox.test(end.year-start.year ~ Status, data=., exact=F)))$p.value,3), ' (Wilcox)'),
+  'Median age at diagnosis (IQR)' = paste0(signif(do(sum.patient.data, tidy(wilcox.test(age.diagnosis ~ Status, data = ., exact=F)))$p.value,3), ' (Wilcox)'),
+  'Mean Length (not recorded)' = paste0(signif(do(sum.patient.data, tidy(wilcox.test(M ~ Status, data=., exact=F)))$p.value,3), ' (Wilcox)'),
+  'Gender F:M' = paste0(signif(dplyr::summarise(sum.patient.data, pvalue = chisq.test(gender, Status)$p.value)$pvalue,3), ' (ChiSq)'),
+  'Smoking N:Y (not recorded)' = paste0(signif(dplyr::summarise(sum.patient.data, pvalue = chisq.test(smoking, Status)$p.value)$pvalue,3), ' (ChiSq)')
 )
 
-cohortSummary = full_join(cohortSummary, cbind.data.frame('Status' = c('NP','P'), rbind(
-      table(subset(patient.info, Status == 'NP')$Pathology),
-      table(subset(patient.info, Status == 'P')$Pathology))), by='Status')
-
-cohortSummary %>% t %>% kable(caption='Training set') %>% kable_styling()
-
-# fit = glm(Status ~ age.diagnosis+M+gender, data=sum.patient.data, family = binomial(link='logit'))
-# summary(fit)
-# 
-# sum.patient.data = sum.patient.data %>% mutate(total.followup = end.year - start.year)
-# 
-# surv =  Surv(sum.patient.data$total.followup,sum.patient.data$Status)
-# terms = as.formula(paste("Surv(total.followup, status) ~", paste(c('age.diagnosis','M','gender'),collapse=" + "), sep=""))
-# 
-# cox = coxph(terms, data=sum.patient.data, na.action=na.omit,ties='breslow')
-# summary(cox)
-# 
-# ggsurv(survfit(cox),cens.shape=3,lty.est=1,back.white=F, CI=T ) +
-#       annotate("text", x=5, y=0.25,label=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),3),"\n", nrow(data), 'patients')) + theme_minimal()
-# 
-# 
-# t.test(filter(sum.patient.data, Status == 'NP')$age.diagnosis,
-#        filter(sum.patient.data, Status == 'P')$age.diagnosis)
-# t.test(filter(sum.patient.data, Status == 'NP')$M, filter(sum.patient.data, Status == 'P')$M)
-# fisher.test(table(sum.patient.data$Status, sum.patient.data$smoking))
-# fisher.test(table(sum.patient.data$Status, sum.patient.data$gender))
-
-```
-
-```{r, echo=F}
+cohortSummary %>% t %>% kable(caption='Training Cohort') %>% kable_styling()
+
+fit = glm(Status ~ age.diagnosis+M+gender+initial.path, data=mutate(sum.patient.data, initial.path = factor(initial.path,ordered=F)), family = binomial(link='logit'))
+tidy(fit) %>% kable(caption = 'Logistic model to test effects of (matched) demographic factors on progression') %>% 
+  kable_styling(full_width = F)
+
+sample.age = patient.info %>% group_by(Samplename, Status, Pathology) %>% dplyr::summarise(sample.age = 2020-Endoscopy.Year)
+
+ggplot(sample.age, aes(Status, sample.age, group=Status, fill=Status)) + geom_jitter(width=0.2) + geom_boxplot(alpha =0.5, width=0.5, outlier.color = NA) + 
+  scale_y_continuous(breaks = seq(0,28,5)) + labs(y='Years', title='Sample age') + plot.theme
+
+ggplot(sample.age, aes(Pathology, sample.age, group=Pathology, fill=Pathology, color=Status)) + 
+  geom_jitter(width=0.2) + geom_boxplot(width=0.5, outlier.color = NA, color='black') + 
+  scale_y_continuous(breaks = seq(0,28,5)) + labs(y='Years', title='Sample age') + plot.theme
+
+```
+
+```{r, echo=F, eval=F}
 grid.arrange(
@@
   tableGrob( 
-    rbind('Non-progressor'=summary(subset(patient.info, Status == 'NP')$p53.Status), 
-           'Progressor'=summary(subset(patient.info, Status == 'P')$p53.Status))), 
+    patient.info %>% mutate(p53.Status = forcats::fct_explicit_na(p53.Status)) %>% group_by(Status, p53.Status) %>% tally() %>% spread(p53.Status, n)),
   nrow=2, as.table=T, heights=c(4,1) )  
@@
 
-#ddf = rbind(dysplasia.df, 'mean.value'=apply(dysplasia.df, 2, mean))
 globalCNdf = as_tibble( t(dysplasia.df[,grep('^\\d+:',colnames(dysplasia.df))]), rownames = 'locs') %>% 
@@
 
+globalCNdf %<>% filter(!is.na(Status))
+
 n.status <- c('Progressor'=paste0('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor')$variable)), ')'),
@@
 ggsave('plots/per_sample_cx_roc.png', plot=roc.plot(rocX, title='Per sample cx') + plot.theme + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
-
-rocX = pROC::roc(Status ~ cx, data=complexity.score, auc=T, ci=T, of='thresholds')
-roc.plot(rocX, title='Per sample cx') + plot.theme
-
-ggsave('plots/per_sample_cx_roc.png', plot=roc.plot(rocX, title='Per sample cx') + plot.theme + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
 ```
@@
 do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits'))
-
 ggsave('plots/eyebrows.png', plot=do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits')), width=10, height=10, units='in', dpi=300)
-
 ```
@@
 ```{r echo=F, fig.height=5, fig.width=5}
-
 autoplot(models$`0.9`) + theme(legend.position = 'none') + plot.theme
-
 ggsave('plots/glm_0.9.png', plot=autoplot(models$`0.9`) + theme(legend.position = 'none') + plot.theme, width=5, height=5, units='in', dpi=300)
@@
 rm(performance.at.1se, coefs, mp)
-
 file = paste(cache.dir, 'nohgd.Rdata', sep='/')
@@
 nohgd.coefs = coefs[[select.alpha]]
-#do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2))
 ggsave('plots/nohgd_eyebrow.png', plot=do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2)), height=10, width=10, units='in', dpi=300)
@@
                               'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))
-
 rm(performance.at.1se, coefs, mp)
@@
 }
-
-#do.call(grid.arrange, c(nolgd.plots, top="No LGD/HGD/IMC samples", ncol=2))
 ```
@@
 ```{r, echo=F, warning=F, message=F, fig.width=6, fig.height=6}
-#classification.rate$x = 1:nrow(classification.rate)
-
 p = ggplot(classification.rate, aes(model, mean, group=model, fill=model)) + geom_col() +
@@
 p
-
 ggsave('plots/performance/compare_classification.png', plot=p, height=5, width=3, units='in', dpi=300 )
@@
 cvRR = features
-
 qq = quantile(features$hazards, seq(0,1,.15))
@@
   write_tsv('plots/coef_effect.tsv')
-
-```
-
-`r features %>% kable(caption='Selected coefficients') %>% kable_styling(full_width=F)`
+```
+
+`r features %>% arrange(desc(hazards)) %>% kable(caption='Selected coefficients') %>% kable_styling(full_width=F)`
 
@@
 ```{r leaveoneout, echo=F, message=F, warning=F, fig.width=5, fig.height=5}
-# loo_residuals = do.call(bind_rows, purrr::map(list.files(paste0(dir, '/Analysis/pcf_perPatient/50kb'), 'residuals.tsv', recursive = T, full.names = T), function(f) {
-#   read_tsv(f, col_types = 'ccddddl')
-# }))
-# 
-# failedQC = loo_residuals %>% filter(varMAD_median > 0.008)
-
 error = read_tsv('~/Data/BarrettsProgressionRisk/Analysis/pcf_perPatient/50kb/sample_error.tsv', col_types = 'cdD')
@@
   dplyr::mutate(P.High = BarrettsProgressionRisk:::pi.hat(RR+Error), P.Low = BarrettsProgressionRisk:::pi.hat(RR-Error), Abs.Error = P.High-P.Low )
-
-#error %>% group_by(Status) %>% dplyr::summarise_at(vars(Abs.Error), list(~mean(.,na.rm=T), ~sd(.,na.rm=T)))
-#summary(error$Abs.Error)
-#sd(error$Abs.Error)
 
@@
 pg.samp = pg.samp %>% mutate(Pathology = recode(Pathology, 'BE' = 'NDBE'))
-
 pg.samp = pg.samp %>% left_join(error %>% dplyr::select(-RR,-Status), by=c('Samplename'='Sample'))
@@
 ggsave('plots/performance_loo.png', plot=p, width=5, height=5, units='in', dpi=300)
-
-#oddPt = names(pg.samp)[which.max(performance.at.1se)]
-#subset(sum.patient.data, Hospital.Research.ID == oddPt)
-
-#pg.samp[[oddPt]]$Prediction
-
-# Not the pt with the most samples or endos
-#sum.patient.data[which.max(sum.patient.data$total.samples),]
 ```
@@
 rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')
-roc.plot(rocNOHGD) + ggtitle('Model excluding HGD/IMC')
-
-rocNOHGD$auc
-pROC::ci.auc(rocNOHGD$auc)
-
-ggsave(filename='plots/auc/roc_nohgd_model.png', plot=roc.plot(rocNOHGD) + ggtitle('Model excluding HGD/IMC'), width=6, height=6, units='in', dpi=300)
-```
+roc.plot(rocNOHGD, 'Model excluding HGD/IMC',incS = T)
+
+pROC::coords(rocNOHGD, 'best', transpose=F) %>% add_column('AUC'=pROC::auc(rocNOHGD)) %>% mutate_all(list(~round(.,2))) %>% 
+  kable(caption = 'LOO excluding HGD/IMC') %>% kable_styling(full_width = F)
+
+ggsave(filename='plots/auc/roc_nohgd_model.png', plot=roc.plot(rocNOHGD, 'Model excluding HGD/IMC',incS = T) , width=6, height=6, units='in', dpi=300)
+```
+
+#### LOO minus HGD/LGD
+
+```{r loonohgd, echo=F, message=F, warning=T}
+file = paste(cache.dir, 'loonolgd.Rdata', sep='/')
+if (file.exists(file)) {
+  message(paste("loading file", file))
+  load(file, verbose=T)
+} else {
+  stop(paste("No file", file, "found"))
+}
+
+#length(performance.at.1se)
+#length(unique(pg.sampNOLGD$Patient))
+
+#perf = cbind.data.frame(perf=performance.at.1se, pt=unique(pg.sampNOLGD$Patient))
+
+pg.sampNOLGD = pg.sampNOLGD %>% filter(!is.na(Prediction))
+
+rocNOLGD = pROC::roc(Status ~ Prediction, data=pg.sampNOLGD[c('Status', 'Prediction')], ci=T, of='thresholds')
+roc.plot(rocNOLGD, 'Model excluding LGD/HGD/IMC',incS = T)
+
+pROC::coords(rocNOLGD, 'best', transpose=F) %>% add_column('AUC'=pROC::auc(rocNOLGD)) %>% mutate_all(list(~round(.,2))) %>% 
+  kable(caption = 'LOO excluding LGD/HGD/IMC') %>% kable_styling(full_width = F)
+
+ggsave(filename='plots/auc/roc_nolgd_model.png', plot=roc.plot(rocNOLGD, 'Model excluding LGD/HGD/IMC',incS = T) , width=6, height=6, units='in', dpi=300)
+```
+
+```{r}
+roc$model = 'All samples model'
+rocNOHGD$model = 'Excluding HGD/IMC'
+rocNOLGD$model = 'Excluding LGD/HGD/IMC'
+
+m = bind_rows(
+      get.roc.stat(roc),
+      get.roc.stat(rocNOHGD)
+#      get.roc.stat(rocNOLGD)
+      ) %>% dplyr::mutate(model = factor(model)) %>% dplyr::mutate_if(is.numeric, list(~round(.,2)))
+
+p = ggplot(m, aes(model, value, ymin=CI.min, ymax=CI.max, group=Measure)) + 
+  geom_bar(aes(fill=Measure), stat='identity',position=dodge) + 
+  geom_errorbar(position=dodge, width=0.3, color='black' ) +
+  geom_text(aes(label=paste(value,sep=''), y=CI.max), vjust=-0.5, position=dodge) + plot.theme + 
+  scale_y_continuous(limits=c(0,1)) +
+  scale_fill_brewer(palette='Set2') +
+  theme(legend.position = 'right',text = element_text(size=14), axis.text.x = element_text(angle=45, hjust=1)) + 
+  labs(x='', y='ROC', title='Model comparisons')
+
+auc.leg = get.legend(p)
+
+ggsave('plots/auc/loo_model_roc.png', plot=p + theme(legend.position = 'none'), height=6, width=4.5, units="in", limitsize=F, dpi = 300)
+
+#ggsave('plots/auc/multi_auc.png', plot=p+theme(legend.position = 'none'), height=6, width=8, units='in',dpi=300)
+```
+
+#### Insilico splits
 
@@
 
-pROC::ci.auc(roc$auc)
-
-ggsave(filename='plots/auc/training_auc.png', plot=roc.plot(roc), width=6.5, height=4.5, units='in', dpi=300)
-ggsave(filename='plots/auc/training_auc_supp.png', plot=roc.plot(roc) + ggtitle('Per sample'), width=6, height=6, units='in', dpi=300)
+roc.plot(roc,'Discovery Cohort all samples ROC',incS=T)
+round(pROC::coords(roc, 'best', transpose=F),2) %>% kable(caption = 'All samples ROC') %>% kable_styling(full_width = F)
+
+#pROC::coords(roc, 0.3, input='threshold', transpose=T)
+
+ggsave(filename='plots/auc/training_auc.png', plot=roc.plot(roc,'Discovery Cohort all samples ROC',incS = T),width=6, height=6, units='in', dpi=300)
+ggsave(filename='plots/auc/training_auc_supp.png', plot=roc.plot(roc,'Discovery Cohort all samples ROC'), width=6, height=6, units='in', dpi=300)
 
@@
 roc$is = 'discovery'
-
-droc = multi.roc.plot(list(roc), colors=brewer.pal(3,'Greys')[3]) + labs(title='Discovery cohort ROC')
-droc
-#ggsave(filename='plots/auc/discovery_roc.png', plot=droc, width=6, height=4, units='in', dpi=300)
 
@@
 
-i.thresholds = do.call(bind_rows, purrr::map(insilico.rocs, function(r) pROC::coords(r, 'best') %>% as_tibble %>% mutate(AUC = r$auc, model=r$model, is=as.character(r$is)))) %>%
+i.thresholds = do.call(bind_rows, purrr::map(insilico.rocs, function(r) pROC::coords(r, 'best') %>% as_tibble %>% 
+                                               mutate(AUC = r$auc, model=r$model, is=as.character(r$is)))) %>%
   rename_at(vars(matches('thresh|spec|sens')), funs(str_to_title))
 
-#summary(filter(df, model == 'in-silico')$AUC)
-
-blues = RColorBrewer::brewer.pal(5,'Blues')
+blues = RColorBrewer::brewer.pal(5,'Greys')
 p = ggplot(df, aes(Specificity, Sensitivity)) + 
@@
     scale_x_reverse() +
-    geom_label(data=filter(i.thresholds, AUC %in% range(AUC) | is == 'discovery'), aes(label=paste0('AUC\n',round(AUC,2))), color='black') +
+    geom_label_repel(data=filter(i.thresholds, AUC %in% range(AUC) | is == 'discovery'), aes(label=paste0('FPR ', round(1-Specificity,2), ' TPR ', round(Sensitivity,2), '\nAUC ',round(AUC,2))), color='black') +
     #scale_color_brewer(palette='Blues') +
-    labs(title='ROC AUC', subtitle='in-silico splits', x='Specificity (1-FPR)', y='Sensitivity (TPR)')  + plot.theme +
+    labs(title='Discovery cohort: 50 models', subtitle='18:70 random patient splits', x='Specificity (1-FPR)', y='Sensitivity (TPR)')  + plot.theme +
     theme(legend.position = 'none', text = element_text(size=14))
@@
 ggsave(filename='plots/auc/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)
-
-# Not sure why I did this
-# df = cbind.data.frame('specificity'=rev(roc$specificities), 'sensitivity'=rev(roc$sensitivities))
-# thresholds = 
-#   rbind.data.frame(
-#     est=t(round(pROC::coords(roc, "best", transpose = T),2)),
-#     max=t(round(pROC::coords(roc, 0.8, transpose = T),2)),
-#     'No HGD'=t(round(pROC::coords(rocNOHGD, 0.6, transpose = T),2)),
-#     'All samples'=t(round(pROC::coords(roc, 0.6, transpose = T),2))
-# )
-# thresholds$label=rownames(thresholds)
-# 
-# ggplot() + 
-#   geom_line(data=df, aes(specificity, sensitivity), color='darkblue', size=1.5) + scale_x_reverse() +
-#   geom_point(data=thresholds, aes(x=specificity,y=sensitivity,color=label), size=3,shape=8, show.legend = T) +
-#   geom_label_repel(data=thresholds, aes(x=specificity,y=sensitivity,fill=label,label=
-#     paste('Threshold ', round(threshold,2), '\n(FPR ', round((1-specificity),2)*100, '%, TPR ', round(sensitivity,2)*100,'%)\nAUC:',round(roc$auc, 2)*100, '%', sep='')),show.legend = F, color='white',nudge_x=0.05) + scale_fill_manual(values=brewer.pal(4,'Dark2')) + scale_color_manual(values=brewer.pal(4,'Dark2')) +
-#   labs(title='CN Model ROC', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme
-
-cutoff = round(pROC::coords(roc, 'best', ret = 'threshold'), 2)
 ```
@@
 
-pROC::ci.auc(p53roc$auc)
+#pROC::ci.auc(p53roc$auc)
+#pROC::coords(p53roc, 'best',transpose=T)
 
 p53roc$model = 'p53 IHC'
-roc.plot(p53roc, 'p53 IHC')
-ggsave('plots/auc/p53_roc.png', plot=roc.plot(p53roc, 'p53 IHC'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+roc.plot(p53roc, 'p53 IHC', incS = T) 
+ggsave('plots/auc/p53_roc.png', plot=roc.plot(p53roc, 'p53 IHC', incS = T) , width=6, height=6, units="in", limitsize=F, dpi = 300)
 ```
@@
 ```{r pathroc, echo=F, fig.height=4, fig.width=12 }
-#sum(table(subset(patient.info, Status == 'P')$Pathology)[c('IMC','HGD')])/sum(table(subset(patient.info, Status == 'P')$Pathology))
-
 pathsamples = patient.info[,c('Pathology','Status')]
 pathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('LGD','HGD','IMC'), 1, 0) )
-pROC::ci.auc(pathroc$auc)
+#pROC::ci.auc(pathroc$auc)
+#pROC::coords(pathroc, 'best')
 
 pathroc$model = 'HGD & LGD'
-ggsave('plots/auc/dysplasia_auc.png', plot=roc.plot(pathroc, title='Dysplasia (HGD & LGD)'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+ggsave('plots/auc/dysplasia_auc.png', plot=roc.plot(pathroc, title='Dysplasia (HGD & LGD)', incS = T), width=6, height=6, units="in", limitsize=F, dpi = 300)
 
 hgdpathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('HGD','IMC'), 1, 0) )
+
 hgdpathroc$model = 'HGD Only'
-ggsave('plots/auc/hgd_auc.png', plot=roc.plot(hgdpathroc, title='HGD Only'), width=6, height=6, units="in", limitsize=F, dpi = 300)
-
-grid.arrange(roc.plot(pathroc, 'Pathology (LGD,HGD,IMC)'), roc.plot(hgdpathroc, 'Pathology (HGD,IMC)'), ncol=2)
+ggsave('plots/auc/hgd_auc.png', plot=roc.plot(hgdpathroc, title='HGD only', incS = T), width=6, height=6, units="in", limitsize=F, dpi = 300)
+
+grid.arrange(roc.plot(pathroc, 'Pathology (LGD,HGD,IMC)'), roc.plot(hgdpathroc, 'Pathology (HGD,IMC)', incS = T), ncol=2)
 ```
@@
 rocEndoM = pROC::roc(Status ~ max, data=pendo, ci=T, of='thresholds')
-rocEndoM$model = 'Per Endoscopy Max'
+rocEndoM$model = 'Max per Endoscopy'
 ggsave('plots/auc/maxendo_auc.png', plot=roc.plot(rocEndoM, title='Max per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
@@
 rocEndoA = pROC::roc(Status ~ avg, data=pendo, ci=T, of='thresholds')
-rocEndoA$model = 'Per Endoscopy Avg'
-ggsave('plots/auc/avgendo_auc.png', plot=roc.plot(rocEndoA, title='Avg per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+rocEndoA$model = 'Mean per Endoscopy'
+ggsave('plots/auc/avgendo_auc.png', plot=roc.plot(rocEndoA, title='Mean per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 
-firstEndo = pg.samp %>% group_by(Patient, Hospital.Research.ID, Status,Endoscopy.Year, PID) %>% arrange(Endoscopy.Year, PID)  %>% dplyr::summarise(avg = mean(Prediction) ) %>% ungroup %>% group_by(Patient) %>% dplyr::filter( Endoscopy.Year == min(Endoscopy.Year) ) %>% dplyr::select(-Endoscopy.Year, -PID)
+firstEndo = pg.samp %>% group_by(Patient, Hospital.Research.ID, Status,Endoscopy.Year, PID) %>% 
+  arrange(Endoscopy.Year, PID)  %>% dplyr::summarise(avg = mean(Prediction) ) %>% ungroup %>% 
+  group_by(Patient) %>% dplyr::filter( Endoscopy.Year == min(Endoscopy.Year) ) %>% dplyr::select(-Endoscopy.Year, -PID)
 
@@
 
-endo.thresholds = bind_cols('roc'=c('Max','Avg','Last','First','Random'),
+endo.thresholds = bind_cols('roc'=c('Max','Mean','Last','First','Random'),
   bind_rows(
@@
   ))
+
+multi.roc.plot(list(rocEndoM, rocEndoA, rocLastEndo, rocFirstEndo, rocRandEndo)) + labs(title='Endoscopy prediction ROCs')
 ```
@@
 #pROC::coords(rocPtM, 'best', transpose=T)
-rocPtM$model = 'Per Patient Max'
+rocPtM$model = 'Max per Patient'
 ggsave('plots/auc/maxpt_auc.png', plot=roc.plot(rocPtM, title='Max per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)
@@
 #pROC::coords(rocPtA, 'best', transpose=T))
-rocPtA$model = 'Per Patient Avg'
-ggsave('plots/auc/avgpt_auc.png', plot=roc.plot(rocPtA, title='Avg per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)
-
-grid.arrange(roc.plot(rocPtM, title='Max') , roc.plot(rocPtA, title='Avg'), nrow=1, top='Per Patient')
+rocPtA$model = 'Mean per Patient'
+ggsave('plots/auc/avgpt_auc.png', plot=roc.plot(rocPtA, title='Mean per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+
+grid.arrange(roc.plot(rocPtM, title='Max') , roc.plot(rocPtA, title='Mean'), nrow=1, top='Per Patient')
+
+multi.roc.plot(list(rocPtM, rocPtA)) + labs(title='Per-Patient ROCs')
+
 ```
@@
 ```{r echo=F, fig.height=8, fig.width=6}
-get.roc.stat<-function(roc, x='best') {
-  auc = c( pROC::auc(roc), range(pROC::ci.auc(roc))  )
-  sens = c(pROC::coords(roc, x, tranpose=T)[['sensitivity']], pROC::ci.se(roc,pROC::coords(roc, x, tranpose=T)[['specificity']], conf.level=0.95))
-  spec = c(pROC::coords(roc, x, tranpose=T)[['specificity']], pROC::ci.sp(roc,pROC::coords(roc, x, tranpose=T)[['sensitivity']], conf.level=0.95))
-
-  stat = rbind('AUC'=auc, 'Sensitivity'=sens[c(1,2,4)], 'Specificity'=spec[c(1,2,4)])
-  colnames(stat) = c ('value','CI.min','CI.max')
-  stat = as.data.frame(stat)
-  stat$model = roc$model
-  stat$Measure = rownames(stat)
-  stat
-}
-
-get.auc.at<-function(roc, foc='spec', p=c(1,0.99)) {
-# To find the auc at a given spec or sens
-  pROC::auc(roc, partial.auc=p, partial.auc.focus=foc, partial.auc.correct=T)
-}
-
+#get.auc.at(roc, p=c(1,0.8))
 # get.auc.at(p53roc)
@@
 roc$model = "CN Model"
-m = rbind.data.frame(
+m = bind_rows(
       get.roc.stat(p53roc),
@@
       get.roc.stat(pathroc),
-      get.roc.stat(roc))
-m$model = factor(m$model, levels=c('p53 IHC','HGD Only','HGD & LGD','CN Model'), ordered = T)
+      get.roc.stat(roc)) %>%
+  dplyr::mutate(model = factor(model, levels=c('p53 IHC','HGD Only','HGD & LGD','CN Model'), ordered = T)) %>% 
+  dplyr::mutate_if(is.numeric, list(~round(.,2)))
 
@@
 
-comp = ggplot(melt(subset(m, Measure == 'AUC'), measure.vars='value'), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Measure)) + 
-  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
-  geom_text(aes(y=(value*100)-4, label=paste(round(value,2)*100,'%',sep='')), position=dodge, color='white')  +
-  scale_y_continuous(limits=c(0,100)) +
-  scale_fill_manual(values=cols, name='') + plot.theme + 
+ggplot(filter(m, Measure == 'AUC'), aes(model, value)) +
+  geom_bar(stat='identity',position=dodge, fill='#7B3294') + geom_errorbar(aes(ymin=CI.min, ymax=CI.max), position=dodge, width=0.3, color='black' ) +
+  geom_text(aes(label=paste(value,sep=''), y=CI.max), vjust=-0.5, position=dodge) + plot.theme + 
+  scale_y_continuous(limits=c(0,1)) +
   theme(legend.position = 'none', text = element_text(size=14), axis.text.x = element_text(angle=45, hjust=1)) + 
   labs(x='', y='AUC', title='ROC Comparison')
-comp
-
-ggsave('plots/auc/path_p53_cn_auc.png', plot=comp, width=3, height=4, units="in", limitsize=F, dpi = 300)
-
+
+comp = ggplot(m, aes(model, value, ymin=CI.min, ymax=CI.max, group=Measure)) + #facet_wrap(~Measure) +
+  geom_bar(aes(fill=Measure), stat='identity',position=dodge) + 
+  geom_errorbar(position=dodge, width=0.3, color='black' ) +
+  geom_text(aes(label=paste(value,sep=''), y=CI.max), vjust=-0.5, position=dodge) + plot.theme + 
+  scale_y_continuous(limits=c(0,1)) +
+  scale_fill_brewer(palette='Set2') +
+  theme(text = element_text(size=14), axis.text.x = element_text(angle=45, hjust=1)) + 
+  labs(x='', y='AUC', title='ROC Comparison')
+
+ggsave('plots/auc/path_p53_cn_auc.png', plot=comp, width=6, height=5, units="in", limitsize=F, dpi = 300)
 ```
@@
 ```{r}
-roc$model = "Per Sample"
-
-m = rbind.data.frame(
+roc$model = "per Sample"
+
+m = bind_rows(
       get.roc.stat(roc),
       get.roc.stat(rocEndoM),get.roc.stat(rocEndoA),
-      get.roc.stat(rocPtM), get.roc.stat(rocPtA),
-      get.roc.stat(rocFirstEndo), get.roc.stat(rocLastEndo),
-      get.roc.stat(rocRandEndo))
-
-m$model = factor(m$model)
-
-cols = brewer.pal(11, 'Paired')
-dodge = position_dodge(width=0.9)
-mm = melt(subset(m, Measure == 'AUC'), measure.vars='value')
-p = ggplot( arrange(mm, CI.min), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=model)) + 
-  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='grey39' ) + 
-  geom_text(aes(y=CI.min*100, label=paste(round(value,2)*100,'%',sep='')), position=dodge, vjust=3)  +
-  scale_fill_manual(values=cols, name='') + plot.theme + 
-  theme(legend.position='none', axis.text.x=element_text(angle=45, hjust=1)) + labs(x='', y='AUC', title='ROC Comparison') 
-p
-
-ggsave('plots/auc/multi_auc.png', plot=p, height=6, width=8, units='in',dpi=300)
+      get.roc.stat(rocPtM), get.roc.stat(rocPtA)
+      #get.roc.stat(rocFirstEndo), get.roc.stat(rocLastEndo),
+      #get.roc.stat(rocRandEndo)
+      ) %>% 
+  dplyr::mutate_if(is.numeric, list(~round(.,2))) %>%
+  dplyr::mutate(model = factor(model, levels=c('per Sample','Mean per Endoscopy','Max per Endoscopy','Mean per Patient','Max per Patient')))
+
+p = ggplot(m, aes(model, value, ymin=CI.min, ymax=CI.max, group=Measure)) + 
+  geom_bar(aes(fill=Measure), stat='identity',position=dodge) + 
+  geom_errorbar(position=dodge, width=0.3, color='black' ) +
+  geom_text(aes(label=paste(value,sep=''), y=CI.max), vjust=-0.5, position=dodge) + plot.theme + 
+  scale_y_continuous(limits=c(0,1)) +
+  scale_fill_brewer(palette='Set2') +
+  theme(legend.position = 'bottom',text = element_text(size=14), axis.text.x = element_text(angle=45, hjust=1)) + 
+  labs(x='', y='AUC', title='Aggregated ROC Comparison')
+
+# cols = brewer.pal(11, 'Paired')
+# dodge = position_dodge(width=0.9)
+# mm = melt(subset(m, Measure == 'AUC'), measure.vars='value') %>% 
+#   mutate(Group = case_when( grepl('Endoscopy', model, ignore.case=T) ~ 'Endoscopy', grepl('Patient',model,ignore.case=T) ~ 'Patient', TRUE ~ 'Sample')) %>%
+#   mutate(order = row_number(), Group = factor(Group, levels=c('Sample','Endoscopy','Patient'), ordered=T))
+# p = ggplot(mm, aes(x=reorder(model, order), y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Group)) + 
+#   facet_grid(~Group, scales = 'free_x', space='free_x') +
+#   geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='grey39' ) + 
+#   geom_text(aes(y=CI.min*100, label=paste(round(value,2)*100,'%',sep='')), position=dodge, vjust=3)  +
+#   scale_fill_manual(values=cols, name='') + plot.theme + 
+#   theme(legend.position='none', axis.text.x=element_text(angle=45, hjust=1)) + labs(x='', y='AUC', title='ROC Comparison') 
+auc.leg = get.legend(p)
+
+ggsave('plots/auc/multi_auc.png', plot=p+theme(legend.position = 'none'), height=6, width=8, units='in',dpi=300)
+ggsave('plots/auc/roc_leg.png', plot=grid.arrange(auc.leg), height=1, width=4, unit='in', dpi=300)
 ```
@@
 ```{r}
-pg.samp = do.call(bind_rows,lapply(unique(pg.samp$Patient) %>% map( function(p) pg.samp %>% filter(Patient == p)), transform.by.time)) 
+pg.samp = do.call(bind_rows,lapply(unique(pg.samp$Patient) %>% purrr::map( function(p) pg.samp %>% filter(Patient == p)), transform.by.time)) 
 sq = c(0, seq(1,132,12),max(pg.samp$months.before.final))
@@
 ) %>% dplyr::mutate(brks = factor(brks, rev(levels(brks))))
-
-ggplot(pg.samp, aes(Status, Prediction, group=Status)) + 
-  facet_grid(~Pathology) + 
-  #scale_fill_manual(values=RColorBrewer::brewer.pal(4,"Paired")[c(1,3)]) + 
-  scale_color_gradientn(colours = myPal) +
-  geom_boxplot(outlier.shape = NA, fill='#BDBDBD') + 
-  geom_jitter(width=0.2, aes(color=Prediction)) + 
-  plot.theme  + theme(legend.position = 'none') + labs(y='Probability')
-
-ggplot(pg.samp, aes(Status, RR, group=Status)) + 
-  facet_grid(~Pathology) + 
-  #scale_fill_manual(values=RColorBrewer::brewer.pal(4,"Paired")[c(1,3)]) + 
-  scale_color_gradientn(colours = myPal) +
-  geom_boxplot(outlier.shape = NA, fill='#BDBDBD') + 
-  geom_jitter(width=0.2, aes(color=Prediction)) + 
-  plot.theme  + theme(legend.position = 'none')
-
 ```
@@
 maxP = 0.5; minP = 0.3
+
+bind_rows(round(as_tibble(pROC::coords(roc, maxP, transpose=F)),2) %>% mutate(Risk = 'High'),
+          round(as_tibble(pROC::coords(roc, minP, transpose=F)),2) %>% mutate(Risk = 'Low')) %>% 
+  kable(caption='Sens/Spec at high/low thresholds') %>%
+  kable_styling(full_width = F)
 
@@
 ```{r probhist, echo=F, warning=F, message=F, fig.width=5, fig.height=5}
-ggplot(pg.samp, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') +
+ggplot(pg.samp, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + 
+  scale_fill_gradientn( colors=myPal) +
+  #scale_fill_distiller(palette = 'RdYlBu', name='P(P)') +
+  #scale_fill_distiller(palette = myPal) +
     geom_vline(xintercept=annotation.table$max, linetype='longdash', color='grey39') +
@@
 ```{r relrisk, echo=F, fig.width=6, fig.height=6}
-unadjRR = ggplot(pg.samp, aes(RR)) + geom_histogram(aes(fill=..x..), bins=10, show.legend = F) +
+unadjRR = ggplot(pg.samp, aes(RR, fill=Prediction)) + 
+  geom_histogram(aes(fill=..x..), bins=10, show.legend = F) +
   scale_fill_gradientn(colors = myPal,  name='') + 
@@
 
-#df = globalCNdf %>% filter(Patient == 20 & seg.type != 'arms')
-#df = globalCNdf %>% filter(Patient == 51 & seg.type != 'arms')
-
 # early/late samples
-#df = subset(globalCNdf, Patient == pt)
 df = left_join(globalCNdf, pg.samp[,c('Samplename','Endoscopy.Year', 'ogj', 'months.before.final', 'RR','Prediction')], by=c('variable'='Samplename')) %>% arrange(chr, start, Endoscopy.Year)
@@
 el2 = el %>% dplyr::group_by(Status, Patient) %>% dplyr::summarise( diff(arms), diff(CNA) )
-wtA = signif(wilcox.test(subset(el2, Status == 'Progressor')$`diff(arms)`, subset(el2, Status == 'Non-Progressor')$`diff(arms)`)$p.value,3)
+wtA = signif(wilcox.test(filter(el2, Status == 'Progressor')$`diff(arms)`, filter(el2, Status == 'Non-Progressor')$`diff(arms)`)$p.value,3)
 #if (wtA < 0.0001) wtA = '< 0.0001'
@@
 ogjLimits = levels(factor(pg.samp$ogj, ordered = T))
+
+# for (i in sort(unique(sum.patient.data$Patient))) {
+#   print(i)
+#   pt = filter(sum.patient.data, Patient == i)
+#   status = as.character(filter(sum.patient.data, Patient == i)$Status)
+# 
+#   pms = mtn.spatial(filter(globalCNdf, Patient == i & seg.type != 'arms'), pg.samp, ogjLimits, paste('Patient', pt, '(', status, ')' ))
+# 
+#   fl = paste0('patient_', i, '_',status,'.png')
+#   ggsave(paste('plots/spatial/', fl, sep=''), plot=grid.arrange(pms$grob), width=10.5, height=4.5, units='in', dpi=300)
+# }
 p2 = mtn.spatial(globalCNdf %>% filter(Patient == 20 & seg.type != 'arms'), pg.samp, ogjLimits,'Patient 20 (P)' )
@@
 
-# for (i in 1:nrow(sum.patient.data)) {
-#   print(i)
-#   pt = sum.patient.data[[i, 'Patient']]
-#   status = as.character(sum.patient.data[[i, 'Status']])
-# 
-#   pms = mtn.spatial(subset(globalCNdf, Patient == pt & seg.type != 'arms'), pg.samp, ogjLimits, paste('Patient', pt, '(', status, ')' ))
-# 
-#   fl = paste0('patient_', pt, '_',status,'.png')
-#   ggsave(paste('plots/spatial/', fl, sep=''), plot=grid.arrange(pms$grob), width=10.5, height=4.5, units='in', dpi=300)
-# }
-
 pg.samp = pg.samp %>% dplyr::select(-ogj)
@@
 ```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
-rp = plot.risk.by.path(pg.samp %>% ungroup) + labs(title='Samples classified by pathology', y='% Classified')
+rp = plot.risk.by.path(pg.samp %>% ungroup) + labs(title='Discovery cohort samples', y='% Classified')
 rp
@@
 
-endos = pg.samp %>% dplyr::group_by(Status, months.before.final, brks, PID) %>% 
+endos = pg.samp %>% filter(Status == 'P') %>% 
+  dplyr::group_by(Status, months.before.final, brks, PID) %>% 
   dplyr::summarise('pred.endo'=ifelse(length(which(Risk == 'High')) > 0, 1, 0), 
@@
 
-#### Using the discretized risk what is the AUC?
-
-```{r}
-risk_df = pg.samp %>% dplyr::select(Status, Risk) %>% mutate( Risk  = factor(droplevels(recode_factor(Risk, 'Moderate' = 'Low')), levels=c('Low','High'), ordered=T) )
-
-rocRisk = pROC::roc(Status ~ Risk, data=risk_df, ci=T, of='thresholds')
-p1 = roc.plot(rocRisk) + labs(title='ROC based on discretized risk', subtitle=paste(levels(risk_df$Risk), collapse=', '))
-p1
-ggsave(filename='plots/risk_auc.png', plot=p1, width=4, height=4, units='in', dpi=300)
-
-#rocRisk2 = pROC::roc(Status ~ Risk, data=pg.samp, ci=T, of='thresholds', transpose = T)
-#p2 = roc.plot(rocRisk2) + labs(title='ROC based on discretized risk', subtitle=paste(levels(pg.samp$Risk), collapse=', '))
-#p2
-
-#grid.arrange(p1,p2, ncol=1)
+```{r pred_by_time}
+pts = pg.samp %>% dplyr::group_by(Status, Patient, months.before.final, brks, PID)%>% dplyr::select(Patient, Status, months.before.final, brks, PID, Block, Endoscopy.Year, Pathology, Prediction, RR, Risk) %>% ungroup
+
+pts %<>% mutate(Patient = factor(Patient)) %>% group_by(Patient) %>% arrange(Endoscopy.Year)
+
+time.pred = pts %>% dplyr::group_by(Status, Patient, months.before.final, brks, Endoscopy.Year) %>% 
+  dplyr::summarise(
+              'endo.risk'= max(Risk),
+              'n.endos'=length(unique(PID)), 
+              'n.samples'=length(PID),
+              'endo.path'= max(Pathology),
+              'endo.prediction' = max(Prediction),
+              'Endo.Year' = unique(Endoscopy.Year),
+              ) %>% ungroup %>% 
+  dplyr::mutate(brks = cut(months.before.final, include.lowest = T, ordered_result = T,
+     breaks=sq,  labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))) %>% 
+  dplyr::mutate(brks = factor(brks, rev(levels(brks)))) %>% 
+  group_by(Patient) %>% filter(months.before.final == min(months.before.final) | months.before.final == max(months.before.final))
+
+time.pred %<>% filter(Patient %in% (time.pred %>% group_by(Patient) %>% tally %>% filter(n>1) %>% dplyr::select(Patient) %>% pull))
+
+#time.pred = left_join(time.pred,time.pred %>% group_by(Status, Patient) %>% dplyr::summarise(max.path = max(endo.path)), by=c('Status','Patient'))
+
+time.diffs = time.pred %>% group_by(Status, Patient) %>% 
+  arrange(Patient, Endoscopy.Year) %>%
+  dplyr::summarise(
+    min.prob = min(endo.prediction),
+    max.prob = max(endo.prediction),
+    prob.diff = abs(diff(endo.prediction)),
+    time.diff = max(months.before.final),
+    risk.diff = diff((as.integer(endo.risk))),
+    Risk = case_when( risk.diff < 0 ~ min(endo.risk), TRUE ~ max(endo.risk) )
+    )
+
+risk.over.time = pg.samp %>% dplyr::mutate(Patient = factor(Patient)) %>%
+  dplyr::group_by(Status, Patient, months.before.final, Endoscopy.Year) %>% 
+  dplyr::summarise( 'n.samples'=length(PID),
+              'max.Path'= max(Pathology),
+              'max.Prob' = max(Prediction),
+              'max.Risk'= max(Risk) ) %>%
+  arrange(Patient, desc(months.before.final)) %>% dplyr::group_by(Patient) %>% 
+  dplyr::mutate(months.since.initial = (Endoscopy.Year-min(Endoscopy.Year))*12,
+                risk.diff = as.integer(max.Risk)-as.integer(max.Risk[which.min(months.since.initial)]),
+                ie.risk.diff = case_when(
+                  risk.diff[which.max(months.since.initial)] < 0 ~ max.Risk[which.min(months.since.initial)],
+                  TRUE ~ max.Risk[which.max(months.since.initial)] ) ) %>% 
+  ungroup %>% mutate_at(vars(risk.diff, ie.risk.diff), list(~factor(., ordered=T))) %>%
+  mutate(risk.diff = dplyr::recode_factor(risk.diff,'-1' = 'Decreased Risk', '-2'='Decreased Risk','0'='No change', '1'='Increased Risk', '2'='Increased Risk', .ordered = T), Status = dplyr::recode_factor(Status, 'P'='Progressor', 'NP'='Non-Progressor'))
+
+tb = risk.over.time %>% group_by(Patient, Endoscopy.Year) %>% tally() %>% filter(n > 1)
+
+for (i in 1:nrow(tb)) {
+  rows = which(risk.over.time$Patient == tb[i,]$Patient & risk.over.time$Endoscopy.Year == tb[i,]$Endoscopy.Year)
+  risk.over.time[max(rows), 'months.since.initial'] = risk.over.time[max(rows), 'months.since.initial'] + 6
+}
+
+p = ggplot(risk.over.time) + facet_wrap(~Status, scales='free', ncol = 1) +
+  geom_line(aes(x=months.since.initial, y=reorder(Patient, desc(ie.risk.diff)), color=ie.risk.diff)) + 
+  scale_color_manual(values=unlist(riskColors()), name='Initial:Final Risk') + 
+  geom_point(aes(x=months.since.initial, y=reorder(Patient, desc(ie.risk.diff)), fill=max.Risk, shape=max.Path), color='white', size=4 ) +  
+  #geom_point(data=filter(risk.over.time, max.Path %in% c('HGD','IMC')), aes(x=months.since.initial, y=reorder(Patient, desc(ie.risk.diff))), shape=8, color='white', size=4 ) +  
+  scale_fill_manual(values=unlist(riskColors()), name='Max Risk',guide=guide_legend(override.aes=list(fill=unlist(riskColors()), color=unlist(riskColors())))) + 
+  scale_shape_manual(values=c(21,22,23,24,25), name='Max Pathology',guide=guide_legend(override.aes=list(fill='grey39', color='white'))) +
+  labs(x='Months since initial endoscopy', y='Patient', title='Risks over time per endoscopy') + 
+  plot.theme 
+p
+ggsave(filename='plots/risks_over_time.png', plot = p, height=12, width=7, units='in')
+#ggsave(filename='plots/risks_over_time.png', plot = p, width=12, height=7, units='in')
 ```
@@
                                     Block = factor(Block, ordered=T))
-
-#blank.axis = theme(axis.text = element_blank(), legend.key=element_rect(fill='grey39'), panel.background=element_rect(colour = 'black'), panel.grid.major=element_blank(), panel.spacing = unit(0.2, 'lines'), panel.border = element_rect(color="black", fill=NA, size=0.5)  )
 
@@
 
-  p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final','rev') + theme(legend.position = 'none') + labs(x='',y='',title='') + theme(axis.ticks = element_blank(), axis.text = element_blank(), plot.margin=grid::unit(c(0,0,0,0), "mm"))
+  p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final','rev') + 
+    theme(legend.position = 'none',panel.border = element_rect(color="black", fill=NA, size=3),  axis.ticks = element_blank(), axis.text = element_blank(), plot.margin=grid::unit(c(0,0,0,0), "mm")) +
+    labs(x='',y='',title='')
 
@@
 ```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=15}
-a = do.call('arrangeGrob', c(plist[c('13','20','34','56','60','48')], top='Progressors'))
-b = do.call('arrangeGrob', c(plist[c('15','16','21','3','45','19')], top='Non-progressors'))
-
-ggsave('plots/ex_pt_heatmap.png', plot=grid.arrange(a,b,ncol=2), width=12, height=15, units='in', dpi=300)
-
 table.path.risk<-function(df) {
@@
 summary.t = table.path.risk(pg.samp)
-#summary.t %>% kableExtra::kable() %>% kableExtra::kable_styling('striped', full_width = F)
+summary.t %>% kableExtra::kable() %>% kableExtra::kable_styling('striped', full_width = F)
 write_tsv(summary.t, path = 'plots/risk_summary.tsv')
@@
 ```{r, echo=T}
-
 p11 = filter(globalCNdf, chr == 17 &  start == 1 & seg.type == 'arms')
@@
 kable(caption='Ratio of all Chr17 p11 loss samples by pathology' ) %>% kable_styling(full_width = F)
-```
-
-Samples with a loss vs all Progressor samples
-Ratio of samples with p11 loss vs all progressor samples: `r length(unique(p11Loss$variable))/length(unique(filter(globalCNdf, Status == 'Progressor')$variable))`
-
-Ratio of progressor patients with a p11 loss in at least 1 sample: `r length(unique(p11Loss$Patient))/length(subset(sum.patient.data, Status == 'P')$Patient)`
+
+p11Loss %>% group_by(variable) %>% tally
+
+```
+
+* Samples with a loss vs all samples
+`r length(unique(p11Loss$variable))`:`r `length(unique(globalCNdf$variable))`,   `r round(length(unique(p11Loss$variable))/length(unique(globalCNdf$variable)),2)*100`%
+
+* Samples with a loss vs all Progressor samples
+Ratio of samples with p11 loss vs all progressor samples: `r length(unique(p11Loss$variable))`:`r `length(unique(filter(globalCNdf, Status == 'Progressor')$variable))`,   `r round(length(unique(p11Loss$variable))/length(unique(filter(globalCNdf, Status == 'Progressor')$variable)),2)*100`%
+
+Ratio of progressor patients with a p11 loss in at least 1 sample: `r length(unique(p11Loss$Patient))`:`r length(subset(sum.patient.data, Status == 'P')$Patient)`r round(length(unique(p11Loss$Patient))/length(subset(sum.patient.data, Status == 'P')$Patient),2)*100`%
 
@@
 
-17p11 loss and p53 IHC correlation  Fisher's odds `r ct$estimate`, p-value = `r ct$p.value`.
+17p11 loss and p53 IHC correlation  Fisher's `r broom::tidy(ct) %>% kable() %>% kable_styling(full_width=F)`
 
@@
 
+## TODO clean this demo stuff up
 val.demo = val.demo %>% mutate_at(vars(Sex,Smoking), list(~forcats::fct_explicit_na(.)))
@@
 
+val.info %>% mutate(sample.age = as.double(difftime(Sys.Date(), Endoscopy, units='weeks'))/52) %>% 
+  group_by(Status) %>% dplyr::summarise_at(vars(sample.age), list(~min(.), ~max(.), ~median(.)))
+
 wilcox.test(filter(val.demo, Status == 'NP')$`Age at diagnosis`, filter(val.demo, Status == 'P')$`Age at diagnosis`)
@@
 # samples
+val.info %>% ungroup %>% dplyr::select(Patient, Status) %>% distinct %>% group_by(Status) %>% tally
 val.info %>% group_by(Status) %>% tally
@@
 val.preds %>% group_by(Status) %>% tally
-
 ```
@@
 
+#pROC::coords(roc, 'best')
+
 totals = val.preds %>% group_by(Status) %>% tally
@@
 ```{r}
-rp = plot.risk.by.path(val.preds) + labs(title='Validation samples classified by pathology', y='% Classified')  
+rp = plot.risk.by.path(val.preds) + labs(title='Validation cohort samples', y='% Classified')  
 rp
-ggsave(filename='plots/validation/val_pred_by_path.png', plot=rp, width=5, height=6, units='in', dpi=300)
+ggsave(filename='plots/validation/val_pred_by_path.png', plot=rp + theme(legend.position = 'none'), width=5, height=6, units='in', dpi=300)
 
@@
 
+```{r, fig.width=6, fig.height=8, eval=F}
 # Risk Categories KM
-
-```{r, fig.width=6, fig.height=8, eval=F}
 require(survival)
@@
 
-pg.samp = pg.samp %>% mutate(p53.Status = factor(p53.Status, levels=c(0,1), ordered=T), Probability = Prediction)
+pg.samp = pg.samp %>% dplyr::mutate(p53.Status = factor(p53.Status, levels=c(0,1), ordered=T), Probability = Prediction)
 
@@
 
-second = second %>% group_by(Patient, Status) %>% arrange(`Time 2`) %>% 
+second = second %>% dplyr::group_by(Patient, Status) %>% arrange(`Time 2`) %>% 
   dplyr::mutate(gid = 1:n() ) %>% dplyr::filter(gid == 2)
@@
 p
-ggsave('plots/second_endo_rx.png', plot=p, width=6, height=4, units = 'in', dpi = 300)
 
@@
 model.dir = '~/Data/BarrettsProgressionRisk/Analysis/dv_snp_model/50kb/'
-load(list.files(model.dir, 'all.pt.alpha',recursive = T, full.names = T))
+load(list.files(model.dir, 'all.pt.alpha',recursive = T, full.names = T), verbose=F)
 
@@
 ```{r, fig.width=12, fig.height=8}
-cache.dir = '~/Data/BarrettsProgressionRisk/Analysis/models_5e6_all/'
+cache.dir = '~/Data/BarrettsProgressionRisk/Analysis/models_5e6_all'
+
 filenames = list.files(cache.dir, 'supp*')
@@
   load(f,verbose=F)
-  model.info
+  left_join(model.info, do.call(bind_rows, purrr::map(rocs, function(roc) {
+    add_column(auc=as.numeric(pROC::auc(roc)), pROC::coords(roc, 'best', transpose=F))  
+  })), by='auc')
 })) %>% mutate('Cohort' = 'Discovery (patient subsets)', Type = 'sWGS')
 
-info = add_row(info, n.patients = d.n.patient, cv.performance = d.perf, n.coefs = nrow(d.coefs), n.samples = 773, auc = d.roc$auc, Cohort = 'Discovery (all)', Type = 'sWGS')
-
-info = add_row(info, n.patients = dv.n.patient, cv.performance = dv.perf, n.coefs = nrow(dv.coefs), n.samples = dv.n.samples, auc = dv.roc$auc, Cohort = 'Discovery+Validation', Type = 'sWGS')
-
-info = add_row(info, n.patients = dvsnp.patient, cv.performance = dvsnp.perf, n.coefs = nrow(dvsnp.coefs), n.samples = nrow(dvsnp.preds), auc = dvsnp.roc$auc, Cohort = 'sWGS (all) + SNP', Type = 'All')
-
-info = add_row(info, n.patients = dvsnp.aucs %>% filter(type == 'snp') %>% dplyr::select(`# patients`) %>% pull, cv.performance = NA, n.coefs = NA, n.samples = NA, auc = dvsnp.aucs %>% filter(type == 'snp') %>% dplyr::select(auc) %>% pull, Cohort = 'sWGS (all) + SNP', Type = 'SNP')
-
-info = add_row(info, n.patients = dvsnp.aucs %>% filter(type == 'sWGS') %>% dplyr::select(`# patients`) %>% pull, cv.performance = NA, n.coefs = NA, n.samples = NA, auc = dvsnp.aucs %>% filter(type == 'sWGS') %>% dplyr::select(auc) %>% pull, Cohort = 'sWGS (all) + SNP', Type = 'sWGS')
+info = add_row(info, n.patients = d.n.patient, cv.performance = d.perf, n.coefs = nrow(d.coefs), n.samples = 773, auc = d.roc$auc,
+               threshold = pROC::coords(d.roc, 'best', transpose=F)$threshold,
+               specificity = pROC::coords(d.roc, 'best', transpose=F)$specificity,
+               sensitivity = pROC::coords(d.roc, 'best', transpose=F)$sensitivity,
+               Cohort = 'Discovery (all)', Type = 'sWGS')
+
+info = add_row(info, n.patients = dv.n.patient, cv.performance = dv.perf, n.coefs = nrow(dv.coefs), n.samples = dv.n.samples, auc = dv.roc$auc,
+               threshold = pROC::coords(dv.roc, 'best', transpose=F)$threshold,
+               specificity = pROC::coords(dv.roc, 'best', transpose=F)$specificity,
+               sensitivity = pROC::coords(dv.roc, 'best', transpose=F)$sensitivity,
+               Cohort = 'Discovery+Validation', Type = 'sWGS')
+
+info = add_row(info, n.patients = dvsnp.patient, cv.performance = dvsnp.perf, n.coefs = nrow(dvsnp.coefs), n.samples = nrow(dvsnp.preds), auc = dvsnp.roc$auc,
+               threshold = pROC::coords(dvsnp.roc, 'best', transpose=F)$threshold,
+               specificity = pROC::coords(dvsnp.roc, 'best', transpose=F)$specificity,
+               sensitivity = pROC::coords(dvsnp.roc, 'best', transpose=F)$sensitivity,
+               Cohort = 'sWGS (all) + SNP', Type = 'All')
+
+info = add_row(info, n.patients = filter(dvsnp.aucs, type == 'snp')$`# patients`, cv.performance = NA, n.coefs = NA, n.samples = NA, 
+               auc = filter(dvsnp.aucs, type== 'snp')$auc,
+               threshold = NA, specificity = NA, sensitivity = NA,
+               Cohort = 'sWGS (all) + SNP', Type = 'SNP')
+
+info = add_row(info, n.patients = filter(dvsnp.aucs, type == 'sWGS')$`# patients`, cv.performance = NA, n.coefs = NA, n.samples = NA, 
+               auc = filter(dvsnp.aucs, type== 'sWGS')$auc,
+               threshold = NA, specificity = NA, sensitivity = NA,
+               Cohort = 'sWGS (all) + SNP', Type = 'sWGS')
 
@@
 ```
+
+```{r,eval=F}
+dv.cache.dir = '~/Data/BarrettsProgressionRisk/Analysis/dv_model_5e6_all_dvz/'
+load(paste0(dv.cache.dir, '/50kb/loo_0.9.Rdata'), verbose=T)
+dv.preds = pg.samp
+rm(plots, performance.at.1se, coefs, nzcoefs, fits, pg.samp)
+
+dv.preds %<>% dplyr::mutate(Pathology = recode(Pathology, BE = 'NDBE'), Risk = case_when(Probability <= 0.3 ~ 'Low', Probability >= 0.5 ~ 'High', TRUE ~ 'Moderate')) %>%
+  mutate(Pathology = factor(Pathology, levels = c('NDBE','ID','LGD','HGD','IMC'), ordered=T))
+
+p = plot.risk.by.path(dv.preds) + labs(subtitle='164 patient model (d+v)')
+
+ggsave(filename='plots/pred_by_path_all.png', plot=p, width=7, height=6, units='in', dpi=300)
+
+dv.preds %>% group_by(Pathology, Risk) %>% tally %>% spread(Pathology, n)
+
+```
+
+# Estimate BE population
+
+From the d only model, patient estimates at the second endoscopy
+
+```{r}
+pg.samp %>% filter(Patient == 3 & Endoscopy.Year == 1998) %>% dplyr::select(Pathology)
+second %>%
+
+get.path<-function(p,y) {
+  max( filter(pg.samp, Patient == p & Endoscopy.Year == y)$Pathology )
+}
+
+nps = second %>% group_by(Patient, Endoscopy.Year) %>% 
+  mutate( max.Path = get.path(Patient,Endoscopy.Year) ) %>% ungroup %>% 
+  dplyr::select(Status, max.Path) %>% filter(Status == 'NP') %>% mutate(max.Path = droplevels(max.Path))
+nps = nps$max.Path  
+#nps = rep(nps$max.Path,24)
+
+# Estimate false positive of LGD recommendations
+table(nps)['LGD']/length(nps)
+
+second %>% dplyr::mutate(Risk = case_when(Probability <= 0.3 ~ 'Low', Probability >= 0.5 ~ 'High', TRUE ~ 'Moderate')) %>%
+  group_by(Status, Risk) %>% tally %>% spread(Status,n) %>%
+  mutate(Total = sum(NP)+sum(P)) %>%
+  dplyr::mutate_at(vars(NP,P), list(~./Total)) %>% dplyr::select(-Total)
+
+## Get estimates at multiple timepoints during followup?
+
+per.endo.max = pg.samp %>% group_by(Patient, Status, PID, months.before.final, Endoscopy.Year) %>% 
+  dplyr::summarise(max.Path = max(Pathology), max.Risk = max(Risk), max.P = max(Prediction))
+
+pt = per.endo.max %>% dplyr::filter(Status == 'NP' & max.Risk == 'High') %>% dplyr::select(Patient) %>% pull
+
+lgd = per.endo.max %>% filter(Patient %in% pt) %>% filter(max.Path == 'LGD')
+per.endo.max %>% filter(Patient %in% lgd$Patient) %>% group_by(Patient) %>% filter(months.before.final == max(months.before.final)) 
+```
+
+Ross-Innes, C.
+S.
+et al.
+Evaluation of a Minimally Invasive Cell Sampling Device Coupled with Assessment of Trefoil Factor 3 Expression for Diagnosing Barrett’s Esophagus: A Multi-Center Case–Control Study.
+PLOS Med.
+12, e1001780 (2015).
+```{r }
+
+ndbe = 372
+id = 46
+lgd = 77
+hgd = 101
+
+est = tibble(
+  Pathology = c(rep('NDBE',372), rep('ID',46), rep('LGD',77),rep('HGD',101)),
+  Risk = c(rep('Low',372), rep('Low',46), rep('High',77),rep('High',101)),
+  Status = c(rep('NP',372+46), rep('P',101+77))
+)
+
+est %>% group_by(Status,Risk) %>% tally %>% spread(Status,n)
+
+x %>% mutate_at(vars(NP,P), list(~round(.*nrow(est))))
+
+n = ndbe + id + lgd + hgd
+
+i = 1000
+
+# lgd
+round(i*(lgd/n))
+
+# Estimated starting numbers from the BEST2 BE cases
+est.lgd = round(i*(lgd/n))
+est.hgd = round(i*(hgd/n))
+est.be = i - est.lgd - est.hgd
+
+# BE->HGD rate per person is 0.006
+beHGD = round(est.be*0.006)
+beLGD = round(est.be*0.03)
+
+# HGD or high risk triggers RFA so the est.hgd cases are immediately at RFA
+
+# From the all model 68% of all LGDs are HIGH risk
+round(est.lgd * 0.68) 
+# but 23% are false positives
+round(est.lgd * 0.68) * 0.23
+
+```

### 2020-04-14 sWGS Normals.Rmd
Bug fixes for paper figures

@@
 library(gridExtra)
-library(pander)
+library(kableExtra)
 source('lib/load_patient_metadata.R')
@@
 
-modeldir = '~/Data/BarrettsProgressionRisk/Analysis/5e6_arms_all'
+modeldir = paste0(analysis.dir,'/models_5e6_all/50kb')
 
@@
 rpi = read.patient.info(patient.file, demo.file)
-normals = rpi$normal
+normals = rpi$normal %>% mutate(Samplename = sub('_','-',Samplename))
 
@@
 resids = do.call(bind_rows, purrr::map(list.files(paste0(analysis.dir, '/normals'), 'residuals.tsv', recursive = T, full.names = T), function(f) read_tsv(f)))
+resids = left_join( dplyr::select(normals, Patient, Hospital.Research.ID, Samplename, Plate.Index), resids, by=c('Samplename'='samplename'))
 
-resids = sample.residual.variance(seg.residuals)
-resids = merge(normals[,c('Patient','Hospital.Research.ID','Samplename')], resids, by.x='Samplename', by.y='sample')
+failedQC = filter(resids, !Pass)
 
-failedQC = arrange(subset(resids, round(var1sd_median,3) > 0.012, select=c(Hospital.Research.ID,Samplename, var1sd_median,var1sd_sd, var1sd_Q1,var1sd_Q3)),  var1sd_median)
-
-pander(normals[,c('Patient','Hospital.Research.ID','Status','Endoscopy.Year','Pathology notes')], caption='Normal samples', justify='left')
+normals %>% dplyr::select(-matches('Cellularity|p53|Reads|Batch|Set')) %>% kable(caption = 'Normal samples') %>% kable_styling(full_width = F)
 ```
@@
 
-```{r, echo=F, warning=F, fig.height=30, fig.width=8}
+```{r, echo=F, warning=F, fig.height=32, fig.width=12}
+plist = lapply(data.files, function(f) {
+  load(f)
+  BarrettsProgressionRisk::plotSegmentData(segmented)
+})
 do.call(grid.arrange, c(plist, ncol=1))
@@
 ```{r, echo=F}
-
 file = paste(modeldir, 'model_data.Rdata', sep='/')
@@
 load(file, verbose=F)
+fit = models[['0.9']]
+lambda = performance.at.1se[['0.9']]$lambda  
+cvRR = BarrettsProgressionRisk:::cvRR(dysplasia.df, coefs[['0.9']])
 rm(dysplasia.df,coefs,plots,labels)
+file = paste(modeldir, 'loo_0.9.Rdata', sep='/')
+if (!file.exists(file))
+  stop(paste("Missing data file", file))
+load(file, verbose=F)
+rm(plots, performance.at.1se, coefs, fits, pg.samp)
 
-select.alpha = '0.9'
-fitV = models[[select.alpha]]
-lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']
+be.model = BarrettsProgressionRisk:::be.model.fit(fit, lambda, 50, 5e6, z.mean, z.arms.mean, z.sd, z.arms.sd, mn.cx, sd.cx, nzcoefs, cvRR)
 
-tiled.segs = segment.matrix(tiled.normals)
-tiled.arms = segment.matrix(tiled.normal.arms)
+preds = tibble()
+plist = list()
+for (f in data.files) { 
+  load(f)
 
-cx.score = score.cx(tiled.segs,1)
+  tiles = BarrettsProgressionRisk:::tileSamples(segmented, be.model = be.model, scale=T, MARGIN=2, verbose=F)
+
+  prr = BarrettsProgressionRisk:::predictRisk(segmented, tiles, be.model, verbose=F)
+
+  plist[f] = grid.arrange(BarrettsProgressionRisk::copyNumberMountainPlot(prr))
+
+  preds = bind_rows(preds, predictions(prr))
+}
+
+paste0(analysis.dir, '/normals')
+files = list.files(paste0(analysis.dir, '/normals'), '5e06_.*tiled', recursive=T, full.names = T)
+tiled.segs = do.call(bind_rows, purrr::map(files, function(f) {
+  read_tsv(f,col_types=c(.default=col_double()))
+})) %>% dplyr::rename('Sample' = 'X1')
+
+files = list.files(paste0(analysis.dir, '/normals'), 'arms_.*tiled', recursive=T, full.names = T)
+tiled.arms = do.call(bind_rows, purrr::map(files, function(f) {
+  read_tsv(f,col_types=c(.default=col_double()))
+})) %>% dplyr::rename('Sample' = 'X1')
+
+tiles.segs[,-1]
+
+cx.score = BarrettsProgressionRisk::scoreCX(tiled.segs,1)
 
 for (i in 1:ncol(tiled.segs)) 
-  tiled.segs[,i] = unit.var(tiled.segs[,i], z.mean[i], z.sd[i])
+  tiled.segs[,i] = BarrettsProgressionRisk:::unit.var(tiled.segs[,i], z.mean[i], z.sd[i])
 for (i in 1:ncol(tiled.arms)) 
-  tiled.arms[,i] = unit.var(tiled.arms[,i], z.arms.mean[i], z.arms.sd[i])
+  tiled.arms[,i] = BarrettsProgressionRisk:::unit.var(tiled.arms[,i], z.arms.mean[i], z.arms.sd[i])
 

### 2020-04-14 snp/SNP_Prediction_Analysis.Rmd
Bug fixes for paper figures

@@
 
-rocBE = pROC::roc(Status ~ Prob, data=preds, auc=T, ci=T, of='thresholds',transpose=T)
-roc.plot(rocBE) + labs(title='SNP ROC', subtitle='BE biopsies')
-pROC::ci.auc(rocBE$auc)
-pROC::coords(rocBE, 'best', transpose=F)
-ggsave(filename = 'plots/snp/be_roc.png', plot = roc.plot(rocBE) + labs(title='SNP ROC', subtitle='BE biopsies') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
+rocBE = pROC::roc(Status ~ Prob, data=preds, auc=T, ci=T, of='thresholds')
+roc.plot(rocBE) + labs(title='SNP ROC, BE biopsies')
+
+rocBE$model = 'per Biopsy'
+
+ggsave(filename = 'plots/snp/be_roc.png', plot = roc.plot(rocBE, 'SNP ROC: BE biopsies') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
 ```
@@
 
-rocEX = pROC::roc( Path.Status ~ Prob, data=pred.qc, auc=T, ci=T, of='thresholds', transpose=T)
-roc.plot(rocEX) + labs(title='BE biopsies', subtitle='Path status (HGD/IMC)')
-pROC::ci.auc(rocEX)
+rocEX = pROC::roc( Path.Status ~ Prob, data=pred.qc, auc=T, ci=T, of='thresholds')
+roc.plot(rocEX) + labs(title='BE biopsies: Path status (HGD/IMC)')
 ggsave(filename='plots/snp/path_status_roc.png', plot = roc.plot(rocEX) + labs(title='BE biopsies', subtitle='Path status (HGD/IMC)') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
+
+rocEX$model = 'HGD/IMC progressed'
 
@@
 # Exclude the low SCA patients
-rocpp = pROC::roc(Status ~ Prob, data=filter(pred.qc, !lowsca), auc=T, ci=T, of='thresholds', transpose=T)
-roc.plot(rocpp) + labs(title='BE biopsies', subtitle='excluding low SCA samples')
-ggsave(filename='plots/snp/auc_ex_lowsca_roc.png', plot = roc.plot(rocpp) + labs(title='BE biopsies', subtitle='excluding low SCA samples') + theme(text=element_text(size=12)), width=5, height=5, units='in', dpi=300)
+rocpp = pROC::roc(Status ~ Prob, data=filter(pred.qc, !lowsca), auc=T, ci=T, of='thresholds')
+roc.plot(rocpp, 'BE biopsies: excl. low SCA samples')
+ggsave(filename='plots/snp/auc_ex_lowsca_roc.png', plot = roc.plot(rocpp, 'BE biopsies: excl. low SCA samples') + theme(text=element_text(size=12)), width=5, height=5, units='in', dpi=300)
+
+rocpp$model = 'All excl. low SCA'
 
 # Exclude all of the above
-rocexall = pROC::roc(Status ~ Prob, data=filter(pred.qc, !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds', transpose=T)
-roc.plot(rocexall) + labs(title='BE biopsies', subtitle='excluding low SCA & WGD NP')
-ggsave(filename='plots/snp/auc_ex_lowsca_wgs_roc.png', plot =roc.plot(rocexall) + labs(title='BE biopsies', subtitle='excluding low SCA & WGD NP') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
+rocexall = pROC::roc(Status ~ Prob, data=filter(pred.qc, !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds')
+roc.plot(rocexall, 'BE biopsies: excl. low SCA & WGD NP')
+ggsave(filename='plots/snp/auc_ex_lowsca_wgs_roc.png', plot =roc.plot(rocexall, 'BE biopsies: excl. low SCA & WGD NP') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
+
+rocexall$model = 'excl. low SCA & (NP) WGD'
 
 # Timepoint 1 only
-roct1 = pROC::roc(Status ~ Prob, data=filter(pred.qc, Timepoint == 'T1' & !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds', transpose=T)
-roc.plot(roct1) + labs(title='BE biopsies', subtitle='Timepoint T1')
-ggsave(filename='plots/snp/auc_t1_roc.png', plot=roc.plot(roct1) + labs(title='BE biopsies', subtitle='Timepoint T1') + theme(text=element_text(size=12)), width=5, height=5, units='in', dpi=300)
+roct1 = pROC::roc(Status ~ Prob, data=filter(pred.qc, Timepoint == 'T1' & !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds')
+roc.plot(roct1, 'BE biopsies: Timepoint T1')
+ggsave(filename='plots/snp/auc_t1_roc.png', plot=roc.plot(roct1, 'BE biopsies: Timepoint T1') + theme(text=element_text(size=12)), width=5, height=5, units='in', dpi=300)
+
+roct1$model = 'Timepoint 1 (with exclusions)'
 
 # Timepoint 2 only
-roct2 = pROC::roc(Status ~ Prob, data=subset(pred.qc, Timepoint != 'T1' & !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds', transpose=T)
-roc.plot(roct2) + labs(title='BE biopsies', subtitle='Timepoint T2')
-ggsave(filename='plots/snp/auc_t2_roc.png', plot=roc.plot(roct2) + labs(title='BE biopsies', subtitle='Timepoint T2') + theme(text=element_text(size=12)), width=5, height=5, units='in', dpi=300)
+roct2 = pROC::roc(Status ~ Prob, data=subset(pred.qc, Timepoint != 'T1' & !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds')
+roc.plot(roct2, 'BE biopsies: Timepoint T2')
+ggsave(filename='plots/snp/auc_t2_roc.png', plot=roc.plot(roct2, 'BE biopsies: Timepoint T2') + theme(text=element_text(size=12)), width=5, height=5, units='in', dpi=300)
+
+roct2$model = 'Timepoint 2  (with exclusions)'
+
+m = bind_rows(
+      get.roc.stat(rocBE),
+      get.roc.stat(rocEX),
+      get.roc.stat(rocexall),
+      get.roc.stat(rocpp),
+      get.roc.stat(roct1),
+      get.roc.stat(roct2)
+) %>% dplyr::mutate_if(is.numeric, list(~round(.,2))) %>%
+  dplyr::mutate(model = factor(model, levels = c(rocBE$model, rocEX$model, rocexall$model, rocpp$model, roct1$model, roct2$model), ordered = T))
+
+dodge = position_dodge(width=0.9)
+p = ggplot(m, aes(model, value, ymin=CI.min, ymax=CI.max, group=Measure)) + 
+  geom_bar(aes(fill=Measure), stat='identity',position=dodge) + 
+  geom_errorbar(position=dodge, width=0.3, color='black' ) +
+  geom_text(aes(label=paste(value,sep=''), y=CI.max), vjust=-0.5, position=dodge) + plot.theme + 
+  scale_y_continuous(limits=c(0,1)) +
+  scale_fill_brewer(palette='Set2') +
+  theme(legend.position = 'right',text = element_text(size=14), axis.text.x = element_text(angle=45, hjust=1)) + 
+  labs(x='', y='ROC', title='Model comparisons') 
+p 
+ggsave(filename='plots/snp/roc_comparisons.png', plot=p, width=7, height=5, units='in', dpi=300)
 

### 2020-04-14 lib/load_patient_metadata.R

@@ -4,7 +4,7 @@ strip.whitespace <- function (x) gsub("\\s+|\\s+", "", x)
 summarise.patient.info<-function(df) {
   require(plyr)
   
-  sum.pt = df %>% dplyr::group_by(Patient, Hospital.Research.ID, Status, Set) %>% 
+  sum.pt = df %>% dplyr::group_by(Patient, Hospital.Research.ID, Status) %>% 
     dplyr::summarise( 
         age.diagnosis = unique(Age.at.diagnosis),
         wt = unique(`Weight.(kg)`),
@@ -36,11 +36,11 @@ read.patient.info<-function(file, file2=NULL, set='All', sheet=NULL) {
     library(readxl)
     
     patient.info = NULL;  
-      if (is.null(sheet)) sheet = 'All combined'
-      ws = readxl::read_xlsx(file, sheet = sheet)
+    if (is.null(sheet)) sheet = 'All combined'
+    ws = readxl::read_xlsx(file, sheet = sheet)
 
-      slx.cols = grep('SLX', colnames(ws), value=T)
-      slx.cols = slx.cols[order(slx.cols)]
+    slx.cols = grep('^SLX', colnames(ws), value=T)
+    slx.cols = slx.cols[order(slx.cols)]
       
       if (length(slx.cols) > 1) {
         slx.rows = ws[,slx.cols]
@@ -49,22 +49,15 @@ read.patient.info<-function(file, file2=NULL, set='All', sheet=NULL) {
       }
       
       ## Order columns
-      ws = ws[,c(grep('Patient',colnames(ws), ignore.case=T),
-                 grep('^Hospital', colnames(ws), ignore.case=T),
-                 grep('Path ID',colnames(ws), ignore.case=T),
-                 grep('Progressor',colnames(ws), ignore.case=T),
-                 grep('Endoscopy year',colnames(ws), ignore.case=T),
-                 grep('Pathology$',colnames(ws),ignore.case=T),
-                 grep('Plate Index',colnames(ws),ignore.case=T),
-                 grep('SLX',colnames(ws),ignore.case=T),
-                 grep('cellularity',colnames(ws),ignore.case=T),
-                 grep('p53 status', colnames(ws),ignore.case=F),
-                 grep('Number of', colnames(ws), ignore.case=F),
[…]

### 2020-04-14 processing_pipeline/merge_qdnaseq_data.R

@@ -16,8 +16,8 @@ val.file = args[2]
 normals = F
 if (length(args) == 3) normals = as.logical(args[3])
 
- data = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/'
-  val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
+# data = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/'
+# val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
 #val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/training/All_patient_info.xlsx'
 
 pastefun<-function(x) {

### 2020-04-14 processing_pipeline/run-cv2.R

@@ -391,7 +391,7 @@ if (lnhgd) {
       
       if (nrow(test) <= 0) next
       
-      patient.samples = pg.sampNOHGD %>% filter(Hospital.Research.ID == pt) %>% select(Samplename)
+      patient.samples = pg.sampNOHGD %>% filter(Hospital.Research.ID == pt) %>% dplyr::select(Samplename)
       # Predict function giving me difficulty when I have only a single sample, this ensures the dimensions are the same
       sparsed_test_data <- Matrix(data=0, nrow=ifelse(nrow(patient.samples) > 1, nrow(test), 1),  ncol=ncol(training),
                                   dimnames=list(rownames(test),colnames(training)), sparse=T)
@@ -445,6 +445,101 @@ if (lnhgd) {
       }
     }
     save(plots, performance.at.1se, coefs, nzcoefs, fits, pg.sampNOHGD, file=file)
+  }
+}
+
+
+## --------- LOO NO HGD/LGD--------- ##
+lnlgd = T 
+if (lnlgd) {
+  message("LOO NO HGD/LGD")
+  
+  pg.sampNOLGD = patient.info %>% rowwise %>% dplyr::mutate(
+    PID = sub('_$', '', unlist(strsplit(Path.ID, 'B'))[1]),
+    Prediction = NA,
+    RR = NA,
+    Prediction.Dev.Resid = NA
+  ) %>% filter(Samplename %in% rownames(dysplasia.df) & Pathology %nin% c('LGD','HGD','IMC'))
+  
+  select.alpha = 0.9
+  
+  file = paste(cache.dir, 'loonolgd.Rdata', sep='/')
+  if (file.exists(file)) {
+    load(file, verbose=T)
+  } else {
+    dysplasia.dfNOLGD = dysplasia.df[intersect(rownames(dysplasia.df),subset(patient.info, Pathology %nin% c('LGD','HGD','IMC'))$Samplename),]
+    
+    secf = all.coefs[[select.alpha]]
+    a = select.alpha
+    
+    performance.at.1se = c(); coefs = list(); plots = list(); fits = list(); nzcoefs = list()
+    # Remove each patient (LOO)
+    for (pt in unique(pg.sampNOLGD$Hospital.Research.ID)) {
+      #for (pt in names(pg.sampNOHGD)) {
+      print(pt)
+      
+      samples = subset(pg.sampNOLGD, Hospital.Research.ID != pt)$Samplename
+      
+      train.rows = which(rownames(dysplasia.dfNOLGD) %in% samples)
[…]

### 2020-04-15 sWGS QC.Rmd
trying to re-process the qc dataset using the Barretts package

@@
 dir1 = '~/Data/BarrettsProgressionRisk/Analysis/pcf_perPatient/50kb/'
-dir2 = '~/Data/BarrettsProgressionRisk/Analysis/validation/pcf_perPatient/50kb/'
+dir2 = '~/Data/BarrettsProgressionRisk/Analysis/qc_batches/pcf_perPatient/50kb/'
 
@@
 res.variance = arrange(res.variance,varMAD_median)
-cutoff = round(median(res.variance$varMAD_median)+sd(res.variance$varMAD_median)*2, 3)
-#cutoff = 0.015
+cutoff = median(res.variance$varMAD_median)+sd(res.variance$varMAD_median)*2
+cutoff = 0.008
 
@@
 ```{r, echo=F}
-val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
-sheets = readxl::excel_sheets(val.file)[1:4]
+val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/qc_batches/qc_batches.xlsx'
+sheets = readxl::excel_sheets(val.file)
 all.val = do.call(bind_rows, lapply(sheets, function(s) {
-  readxl::read_xlsx(val.file, s) %>% dplyr::select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA, matches('Collection')) %>% 
-    dplyr::filter(!is.na(`SLX-ID`) & RA == 'C.Kosmidou') %>% 
+  readxl::read_xlsx(val.file, s) %>% dplyr::select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`,  Batch, matches('Collection')) %>% 
     mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) 
-}))
-all.val = all.val %>% dplyr::select(-Collection)
+})) %>% dplyr::select(-Collection)
 
@@
   `Index Sequence` = str_replace_all(`Index Sequence`, 'tp', ''),
-  Sample = paste(`SLX-ID`,`Index Sequence`,sep='.')
+  Sample = paste(`SLX-ID`,sub('-','_',`Index Sequence`),sep='.')
   )

### 2020-04-21 Paper.Rmd
Updating figures

@@
 library(tidyverse)
+library(broom)
 library(BarrettsProgressionRisk)
@@
 rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')
-roc.plot(rocNOHGD, 'Model excluding HGD/IMC',incS = T)
-
-pROC::coords(rocNOHGD, 'best', transpose=F) %>% add_column('AUC'=pROC::auc(rocNOHGD)) %>% mutate_all(list(~round(.,2))) %>% 
+roc.plot(rocNOHGD, 'Model excluding HGD/IMC',incS = T, 0.3)
+
+pROC::coords(rocNOHGD, 0.3, transpose=F) %>% add_column('AUC'=pROC::auc(rocNOHGD)) %>% mutate_all(list(~round(.,2))) %>% 
   kable(caption = 'LOO excluding HGD/IMC') %>% kable_styling(full_width = F)
 
-ggsave(filename='plots/auc/roc_nohgd_model.png', plot=roc.plot(rocNOHGD, 'Model excluding HGD/IMC',incS = T) , width=6, height=6, units='in', dpi=300)
+ggsave(filename='plots/auc/roc_nohgd_model.png', plot=roc.plot(rocNOHGD, 'Model excluding HGD/IMC',T, 0.3) , width=6, height=6, units='in', dpi=300)
 ```
@@
 
-```{r loonohgd, echo=F, message=F, warning=T}
+```{r loonolgd, echo=F, message=F, warning=T}
 file = paste(cache.dir, 'loonolgd.Rdata', sep='/')
@@
 rocNOLGD = pROC::roc(Status ~ Prediction, data=pg.sampNOLGD[c('Status', 'Prediction')], ci=T, of='thresholds')
-roc.plot(rocNOLGD, 'Model excluding LGD/HGD/IMC',incS = T)
-
-pROC::coords(rocNOLGD, 'best', transpose=F) %>% add_column('AUC'=pROC::auc(rocNOLGD)) %>% mutate_all(list(~round(.,2))) %>% 
+roc.plot(rocNOLGD, 'Model excluding LGD/HGD/IMC',T,0.3)
+
+pROC::coords(rocNOLGD, 0.3, transpose=F) %>% add_column('AUC'=pROC::auc(rocNOLGD)) %>% mutate_all(list(~round(.,2))) %>% 
   kable(caption = 'LOO excluding LGD/HGD/IMC') %>% kable_styling(full_width = F)
 
-ggsave(filename='plots/auc/roc_nolgd_model.png', plot=roc.plot(rocNOLGD, 'Model excluding LGD/HGD/IMC',incS = T) , width=6, height=6, units='in', dpi=300)
-```
-
-```{r}
-roc$model = 'All samples model'
-rocNOHGD$model = 'Excluding HGD/IMC'
-rocNOLGD$model = 'Excluding LGD/HGD/IMC'
-
-m = bind_rows(
-      get.roc.stat(roc),
-      get.roc.stat(rocNOHGD)
-#      get.roc.stat(rocNOLGD)
-      ) %>% dplyr::mutate(model = factor(model)) %>% dplyr::mutate_if(is.numeric, list(~round(.,2)))
-
-p = ggplot(m, aes(model, value, ymin=CI.min, ymax=CI.max, group=Measure)) + 
-  geom_bar(aes(fill=Measure), stat='identity',position=dodge) + 
-  geom_errorbar(position=dodge, width=0.3, color='black' ) +
-  geom_text(aes(label=paste(value,sep=''), y=CI.max), vjust=-0.5, position=dodge) + plot.theme + 
-  scale_y_continuous(limits=c(0,1)) +
-  scale_fill_brewer(palette='Set2') +
-  theme(legend.position = 'right',text = element_text(size=14), axis.text.x = element_text(angle=45, hjust=1)) + 
-  labs(x='', y='ROC', title='Model comparisons')
-
-auc.leg = get.legend(p)
-
-ggsave('plots/auc/loo_model_roc.png', plot=p + theme(legend.position = 'none'), height=6, width=4.5, units="in", limitsize=F, dpi = 300)
-
-#ggsave('plots/auc/multi_auc.png', plot=p+theme(legend.position = 'none'), height=6, width=8, units='in',dpi=300)
+ggsave(filename='plots/auc/roc_nolgd_model.png', plot=roc.plot(rocNOLGD, 'Model excluding LGD/HGD/IMC',T, 0.3) , width=6, height=6, units='in', dpi=300)
 ```
@@
 rocEndoM$model = 'Max per Endoscopy'
-ggsave('plots/auc/maxendo_auc.png', plot=roc.plot(rocEndoM, title='Max per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+ggsave('plots/auc/maxendo_auc.png', plot=roc.plot(rocEndoM, title='Max per Endoscopy', T, 0.3), width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 rocEndoA$model = 'Mean per Endoscopy'
-ggsave('plots/auc/avgendo_auc.png', plot=roc.plot(rocEndoA, title='Mean per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+ggsave('plots/auc/avgendo_auc.png', plot=roc.plot(rocEndoA, title='Mean per Endoscopy', T, 0.3), width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 rocLastEndo$model = 'Last endoscopy'
-ggsave('plots/auc/lastendo_auc.png', plot=roc.plot(rocLastEndo, title='Last endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+ggsave('plots/auc/lastendo_auc.png', plot=roc.plot(rocLastEndo, title='Last endoscopy', T, 0.3), width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 rocFirstEndo$model = 'First endoscopy'
-ggsave('plots/auc/firstendo_auc.png', plot=roc.plot(rocFirstEndo, title='First Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+ggsave('plots/auc/firstendo_auc.png', plot=roc.plot(rocFirstEndo, title='First Endoscopy', T, 0.3), width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 rocRandEndo$model = 'Per Endoscopy Random'
-ggsave('plots/auc/randomendo_auc.png', plot=roc.plot(rocRandEndo, title='Random per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)
-
-endo.thresholds = bind_cols('roc'=c('Max','Mean','Last','First','Random'),
-  bind_rows(
-    pROC::coords(rocEndoM, 'best', transpose = T),
-    pROC::coords(rocEndoA, 'best', transpose = T),
-    pROC::coords(rocLastEndo, 'best', transpose = T),
-    pROC::coords(rocFirstEndo, 'best', transpose = T),
-    pROC::coords(rocRandEndo, 'best', transpose = T)
-  ))
-
-multi.roc.plot(list(rocEndoM, rocEndoA, rocLastEndo, rocFirstEndo, rocRandEndo)) + labs(title='Endoscopy prediction ROCs')
+ggsave('plots/auc/randomendo_auc.png', plot=roc.plot(rocRandEndo, title='Random per Endoscopy', T, 0.3), width=6, height=6, units="in", limitsize=F, dpi = 300)
+
+# endo.thresholds = bind_cols('roc'=c('Max','Mean','Last','First','Random'),
+#   bind_rows(
+#     pROC::coords(rocEndoM, 'best', transpose = T),
+#     pROC::coords(rocEndoA, 'best', transpose = T),
+#     pROC::coords(rocLastEndo, 'best', transpose = T),
+#     pROC::coords(rocFirstEndo, 'best', transpose = T),
+#     pROC::coords(rocRandEndo, 'best', transpose = T)
+#   ))
+
+multi.roc.plot(list(rocEndoM, rocEndoA, rocLastEndo, rocFirstEndo, rocRandEndo), threshold = 0.3) + labs(title='Endoscopy prediction ROCs')
 ```
@@
 rocPtM$model = 'Max per Patient'
-ggsave('plots/auc/maxpt_auc.png', plot=roc.plot(rocPtM, title='Max per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)
+ggsave('plots/auc/maxpt_auc.png', plot=roc.plot(rocPtM, title='Max per Patient (excl. final)', T, 0.3), width=6, height=6, units="in", limitsize=F, dpi = 300)
 
@@
 rocPtA$model = 'Mean per Patient'
-ggsave('plots/auc/avgpt_auc.png', plot=roc.plot(rocPtA, title='Mean per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)
-
-grid.arrange(roc.plot(rocPtM, title='Max') , roc.plot(rocPtA, title='Mean'), nrow=1, top='Per Patient')
-
-multi.roc.plot(list(rocPtM, rocPtA)) + labs(title='Per-Patient ROCs')
-
+ggsave('plots/auc/avgpt_auc.png', plot=roc.plot(rocPtA, title='Mean per Patient (excl. final)', T, 0.3), width=6, height=6, units="in", limitsize=F, dpi = 300)
+
+multi.roc.plot(list(rocPtM, rocPtA), threshold = 0.3) + labs(title='Per-Patient ROCs', subtitle = 'threshold = 0.3')
 ```
@@
       get.roc.stat(pathroc),
-      get.roc.stat(roc)) %>%
+      get.roc.stat(roc, 0.3)) %>%
   dplyr::mutate(model = factor(model, levels=c('p53 IHC','HGD Only','HGD & LGD','CN Model'), ordered = T)) %>% 
   dplyr::mutate_if(is.numeric, list(~round(.,2)))
-
-rocList = list(roc,pathroc,hgdpathroc,p53roc)
 
@@
 ```{r}
+roc$model = 'All samples model'
+rocNOHGD$model = 'Excluding HGD/IMC'
+rocNOLGD$model = 'Excluding LGD/HGD/IMC'
+
+m = bind_rows(
+      get.roc.stat(roc,0.3),
+      get.roc.stat(rocNOHGD,0.3)
+#      get.roc.stat(rocNOLGD)
+      ) %>% dplyr::mutate(model = factor(model)) %>% dplyr::mutate_if(is.numeric, list(~round(.,2)))
+
+p = ggplot(m, aes(model, value, ymin=CI.min, ymax=CI.max, group=Measure)) + 
+  geom_bar(aes(fill=Measure), stat='identity',position=dodge) + 
+  geom_errorbar(position=dodge, width=0.3, color='black' ) +
+  geom_text(aes(label=paste(value,sep=''), y=CI.max), vjust=-0.5, position=dodge) + plot.theme + 
+  scale_y_continuous(limits=c(0,1)) +
+  scale_fill_brewer(palette='Set2') +
+  theme(legend.position = 'right',text = element_text(size=12)) + 
+  labs(x='', y='ROC', title='LOO model comparisons', subtitle = 'threshold = 0.3') 
+auc.leg = get.legend(p)
+
+#p = multi.roc.plot(list(roc,rocNOHGD), threshold = 0.3, palette = 'Dark2' ) + labs(title='LOO model comparisons', subtitle = 'threshold = 0.3')
+p
+
+#ggsave('plots/auc/loo_model_roc.png', plot=p + theme(legend.position = 'none'), height=4, width=4, units="in", limitsize=F, dpi = 300)
+
+ggsave('plots/auc/loo_model_roc.png', plot=p + theme(legend.position = 'none'), height=4, width=4, units="in", limitsize=F, dpi = 300)
+
+#ggsave('plots/auc/multi_auc.png', plot=p+theme(legend.position = 'none'), height=6, width=8, units='in',dpi=300)
+```
+
+```{r}
 roc$model = "per Sample"
 
+multi.roc.plot(list(roc,rocEndoM, rocEndoA), threshold = 0.3, palette = 'Dark2' ) + labs(title='Aggregated per Endoscopy ', subtitle = 'threshold = 0.3')
+
 m = bind_rows(
-      get.roc.stat(roc),
-      get.roc.stat(rocEndoM),get.roc.stat(rocEndoA),
-      get.roc.stat(rocPtM), get.roc.stat(rocPtA)
+      get.roc.stat(roc,0.3),
+      get.roc.stat(rocEndoM,0.3),get.roc.stat(rocEndoA,0.3),
+      get.roc.stat(rocPtM,0.3), get.roc.stat(rocPtA,0.3)
       #get.roc.stat(rocFirstEndo), get.roc.stat(rocLastEndo),
@@
   theme(legend.position = 'bottom',text = element_text(size=14), axis.text.x = element_text(angle=45, hjust=1)) + 
-  labs(x='', y='AUC', title='Aggregated ROC Comparison')
+  labs(x='', y='ROC ', title='Aggregated ROC Comparison', subtitle='threshold = 0.3')
 
@@
 
+p
+
 ggsave('plots/auc/multi_auc.png', plot=p+theme(legend.position = 'none'), height=6, width=8, units='in',dpi=300)
@@
 ogjLimits = levels(factor(pg.samp$ogj, ordered = T))
-
+```
+
+```{r}
 # for (i in sort(unique(sum.patient.data$Patient))) {
@@
 # }
-p2 = mtn.spatial(globalCNdf %>% filter(Patient == 20 & seg.type != 'arms'), pg.samp, ogjLimits,'Patient 20 (P)' )
-p6 = mtn.spatial(globalCNdf %>% filter(Patient == 51 & seg.type != 'arms'), pg.samp, ogjLimits,'Patient 51 (NP)' )
-ggsave(filename='plots/spatial/patient_20_P.png', plot=grid.arrange(p2$grob), width=10.5, height=4.5, units='in', dpi=300)
-ggsave(filename='plots/spatial/patient_51_NP.png', plot=grid.arrange(p6$grob), width=10.5, height=4.5, units='in', dpi=300)
-
-grid.arrange(p2$grob, p6$grob, ncol=1)
+p20 = mtn.spatial(globalCNdf %>% filter(Patient == 20 & seg.type != 'arms'), pg.samp, ogjLimits )
+ggsave(filename='plots/spatial/patient_20_P.png', plot=grid.arrange(p20$grob), width=16, height=6, units='in', dpi=300)
+
+p51 = mtn.spatial(globalCNdf %>% filter(Patient == 51 & seg.type != 'arms'), pg.samp, ogjLimits,'Patient 51 (NP)' )
+ggsave(filename='plots/spatial/patient_51_NP.png', plot=grid.arrange(p51$grob), width=10.5, height=4.5, units='in', dpi=300)
+
+grid.arrange(p20$grob, p51$grob, ncol=1)
 
@@
 
-ggsave('plots/spatial/legend.png', plot=grid.arrange(p2$legend), width=2,height=2, units='in', dpi=300)
+ggsave('plots/spatial/legend.png', plot=grid.arrange(p20$legend), width=2,height=2, units='in', dpi=300)
 ```
@@
   geom_text(aes(y=5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white', nudge_y=2) + 
-  labs(x='Months prior to endpoint', y='Percent classified', title='High risk classifications prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') + coord_cartesian(ylim = c(0,100)) +
+  labs(x='Months prior to endpoint', y='Sensitivity', title='High risk classifications prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') + coord_cartesian(ylim = c(0,100)) + 
   plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12) ) 
@@
 val.demo %>% filter(Patient != no_followup$Patient) %>% group_by(Status,Sex) %>% tally %>% spread(Status, n)
+chisq.test(table(val.demo$Sex, val.demo$Status))
 
 val.demo %>% filter(Patient != no_followup$Patient) %>% group_by(Status,Smoking) %>% tally %>% spread(Status, n) 
+chisq.test(table(val.demo$Smoking, val.demo$Status)[-3,])
 
@@
   str =  paste0('Patient ', pt, ': ', age, sex)
-  if (smk != '' & !is.na(smk)) str = paste0(str, ' ,', smk)
+  if (smk != '' & !is.na(smk)) str = paste0(str, ', ', smk)
   str = paste0(str,'\n')
@@
 pt = 34
-p34 = vig.plot(df=filter(pg.samp, Patient == pt),info= filter(patient.info, Patient == pt),cols=riskCols) + 
-    theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
+
+p34 = vig.plot(df=filter(pg.samp, Patient == pt),cols=riskCols) + 
+    theme(legend.position = 'bottom', text = element_text(size=12)) + labs(title = filter(sum.patient.data, Patient == pt)$desc)
+
+#vig.legend = get.legend(p34)
+#ggsave('plots/vignette/legend_long.png', plot=grid.arrange(vig.legend), width=8, height=1)
+ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p34 + theme(legend.position = 'none'), width=6.35, height=4.5,dpi=300)
+
+pt = 3
+p3 = vig.plot(df=filter(pg.samp, Patient == pt),cols=riskCols) + 
+    labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
     theme(legend.position = 'none', text = element_text(size=12))
-ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p34, width=6.35, height=4.5,dpi=72)
-
-pt = 3
-p3 = vig.plot(df=filter(pg.samp, Patient == pt),info= filter(patient.info, Patient == pt),cols=riskCols) + 
-    theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
+ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p3, width=6.35, height=4,dpi=300)
+
+pt = 68
+p68 = vig.plot(df=filter(pg.samp, Patient == pt),cols=riskCols) + 
+    labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
     theme(legend.position = 'none', text = element_text(size=12))
-ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p3, width=6.35, height=4,dpi=72)
-
-pt = 68
-p68 = vig.plot(df=filter(pg.samp, Patient == pt),info= filter(patient.info, Patient == pt),cols=riskCols) + 
-    theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
-    theme(legend.position = 'none', text = element_text(size=12))
-ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p68, width=6.35, height=5,dpi=72)
+ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p68, width=6.35, height=6,dpi=300)
 
@@
 p
+ggsave('plots/second_endo_rx.png', plot=p, width=6, height=6, units='in', dpi=300)
 

### 2020-04-21 sWGS QC.Rmd
Updating figures

@@
 res.variance = arrange(res.variance,varMAD_median)
-cutoff = median(res.variance$varMAD_median)+sd(res.variance$varMAD_median)*2
-cutoff = 0.008
+cutoff = round(median(res.variance$varMAD_median)+sd(res.variance$varMAD_median)*3,3)
+#cutoff = 0.010
 
@@
 
-## Validation dataset
+## QC dataset
 
@@
   `Index Sequence` = str_replace_all(`Index Sequence`, 'tp', ''),
-  Sample = paste(`SLX-ID`,sub('-','_',`Index Sequence`),sep='.')
+  Sample = paste(`SLX-ID`,sub('-','_',gsub('tp','',`Index Sequence`)),sep='.')
   )
 
-resVal = do.call(bind_rows, lapply(list.files(dir2, 'residuals', recursive = T, full.names = T), function(f) {
+qc.res.val = do.call(bind_rows, lapply(list.files(dir2, 'residuals', recursive = T, full.names = T), function(f) {
   readr::read_tsv(f, col_types = 'ccddddl') #%>% set_names('sample', 'varMAD_median', 'varMAD_sd', 'varMAD_Q1', 'varMAD_Q3', 'n.segs', 'Pass') %>%  
    # mutate(`Patient ID` = as.character(basename(dirname(f))))
-})) %>% dplyr::select(patient, samplename, matches('varMad'), Pass)
+})) %>% dplyr::select(patient, samplename, matches('varMad'), -Pass) %>% filter(samplename %in% all.val$Sample)
 
-resVal = resVal %>% filter(samplename %in% all.val$Sample)
+qc.res.val %>% kable(caption='Residuals across segments per sample') %>% kable_styling('striped')
+qc.res.val = arrange(qc.res.val,varMAD_median)
 
-resVal %>% kable(caption='Residuals across segments per sample') %>% kable_styling('striped')
-resVal = arrange(resVal,varMAD_median)
-
-ggplot(resVal,aes(sample=varMAD_median)) + stat_qq() + geom_hline(yintercept=cutoff, linetype='dashed',color='grey') + 
+ggplot(qc.res.val,aes(sample=varMAD_median)) + stat_qq() + geom_hline(yintercept=cutoff, linetype='dashed',color='grey') + 
   labs(title='Normal Q-Q plot (all)', x='Theoretical quantiles',y='Sample quantiles') + theme_bw()
 
-ggplot(resVal, aes(varMAD_median)) + geom_histogram(bins=50,color='lightblue') + 
+ggplot(qc.res.val, aes(varMAD_median)) + geom_histogram(bins=50,color='lightblue') + 
   geom_vline(xintercept = cutoff, linetype='dashed', color='grey') + labs(x='var(median(segments)+1sd(segments))') +
@@
 
-failedQCVal = resVal %>% filter(varMAD_median > cutoff) %>% arrange(`Patient ID`, varMAD_median)
+failedQCVal = qc.res.val %>% filter(varMAD_median > cutoff) %>% arrange(patient, varMAD_median)
 
@@
 
-The result from this is that `r nrow(failedQCVal)` samples (`r signif(nrow(failedQCVal)/nrow(resVal)*100,3)`%) would have failed due to high variance across all the segments.
+The result from this is that `r nrow(failedQCVal)` samples (`r signif(nrow(failedQCVal)/nrow(qc.res.val)*100,3)`%) would have failed due to high variance across all the segments.
 A manual check of these agrees with all of them.
-
-## If we learned it from all residuals
-
-```{r}
-res.variance = res.variance %>% mutate_at(vars(patient, sample), as.character)
-resVal = resVal %>% dplyr::rename( patient = `Patient ID` ) %>% mutate_at(vars(patient, sample), as.character)
-
-res.variance$cohort = 'discovery'
-resVal$cohort = 'validation'
-
-ar = bind_rows( res.variance, resVal ) %>% mutate(cohort = factor(cohort))
-
-#newcutoff = round(median(ar$varMAD_median)+sd(ar$varMAD_median)*2, 3)
-newcutoff = 0.015
-
-ggplot(ar,aes(sample=varMAD_median)) + stat_qq() + geom_hline(yintercept=newcutoff, linetype='dashed',color='grey') + 
-    labs(title='Normal Q-Q plot (all)', x='Theoretical quantiles',y='Sample quantiles') + theme_bw()
-
-ggplot(ar, aes(varMAD_median)) + geom_histogram(bins=50,color='lightblue') + 
-  geom_vline(xintercept = newcutoff, linetype='dashed', color='grey') + labs(x='var(MAD)') +
-  theme_bw()
-
-fc = ar %>% filter(round(varMAD_median,3) > newcutoff) %>% select(patient, sample, varMAD_median,varMAD_sd, varMAD_Q1,varMAD_Q3) %>% arrange(patient, varMAD_median)
-```
-
-Using the cutoff `r newcutoff` learned from all of them we would fail `r length(which(resVal$varMAD_median > newcutoff))` samples from the training dataset (Ellie's) and `r length(which(resVal$varMAD_median > newcutoff))` from the initial validation dataset.
-So while it's less stringent it still fails `r length(which(failedQCVal$varMAD_median > newcutoff))/nrow(resVal)*100`% versus `r length(which(failedQC$varMAD_median > newcutoff))/nrow(res.variance)*100`% of the training dataset.

### 2020-04-21 snp/SNP_Prediction_Analysis.Rmd
Updating figures

@@
 rocBE = pROC::roc(Status ~ Prob, data=preds, auc=T, ci=T, of='thresholds')
-roc.plot(rocBE) + labs(title='SNP ROC, BE biopsies')
-
+roc.plot(rocBE,title='SNP ROC, BE biopsies',incS = T, 0.3)
 rocBE$model = 'per Biopsy'
 
-ggsave(filename = 'plots/snp/be_roc.png', plot = roc.plot(rocBE, 'SNP ROC: BE biopsies') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
+ggsave(filename = 'plots/snp/be_roc.png', plot = roc.plot(rocBE, 'SNP ROC: BE biopsies', T, 0.3) + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
 ```
@@
 ggsave(filename='plots/snp/path_status_roc.png', plot = roc.plot(rocEX) + labs(title='BE biopsies', subtitle='Path status (HGD/IMC)') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
-
 rocEX$model = 'HGD/IMC progressed'
@@
 ggsave(filename='plots/snp/auc_ex_lowsca_roc.png', plot = roc.plot(rocpp, 'BE biopsies: excl. low SCA samples') + theme(text=element_text(size=12)), width=5, height=5, units='in', dpi=300)
-
 rocpp$model = 'All excl. low SCA'
@@
 ggsave(filename='plots/snp/auc_ex_lowsca_wgs_roc.png', plot =roc.plot(rocexall, 'BE biopsies: excl. low SCA & WGD NP') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
-
 rocexall$model = 'excl. low SCA & (NP) WGD'
@@
 ggsave(filename='plots/snp/auc_t1_roc.png', plot=roc.plot(roct1, 'BE biopsies: Timepoint T1') + theme(text=element_text(size=12)), width=5, height=5, units='in', dpi=300)
-
 roct1$model = 'Timepoint 1 (with exclusions)'
@@
 ggsave(filename='plots/snp/auc_t2_roc.png', plot=roc.plot(roct2, 'BE biopsies: Timepoint T2') + theme(text=element_text(size=12)), width=5, height=5, units='in', dpi=300)
-
 roct2$model = 'Timepoint 2  (with exclusions)'
@@
 m = bind_rows(
-      get.roc.stat(rocBE),
-      get.roc.stat(rocEX),
-      get.roc.stat(rocexall),
-      get.roc.stat(rocpp),
-      get.roc.stat(roct1),
-      get.roc.stat(roct2)
+      get.roc.stat(rocBE,0.3),
+      get.roc.stat(rocEX,0.3),
+      get.roc.stat(rocexall,0.3),
+      get.roc.stat(rocpp,0.3),
+      get.roc.stat(roct1,0.3),
+      get.roc.stat(roct2,0.3)
 ) %>% dplyr::mutate_if(is.numeric, list(~round(.,2))) %>%
@@
   theme(legend.position = 'right',text = element_text(size=14), axis.text.x = element_text(angle=45, hjust=1)) + 
-  labs(x='', y='ROC', title='Model comparisons') 
+  labs(x='', y='ROC', title='Model comparisons', subtitle = 'threshold = 0.3') 
 p 

### 2020-04-21 processing_pipeline/merge_qdnaseq_data.R

@@ -26,7 +26,7 @@ pastefun<-function(x) {
 }
 
 sheets = readxl::excel_sheets(val.file)
-
+all = grep('All',sheets,value=T)
 
 if (length(all) == 1) {
   all.info = read.patient.info(val.file, sheet=all)
@@ -44,9 +44,9 @@ if (length(all) == 1) {
       mutate_at(vars(`SLX-ID`), list(as.character))
   }))
   all.info = all.info %>% rowwise %>% mutate_at(vars(`SLX-ID`), list(pastefun) ) %>% ungroup
-  all.info = all.info %>% mutate(`Hospital Research ID` = str_replace_all( str_remove_all(`Hospital Research ID`, " "), '/', '_'), `Index Sequence` = str_replace_all(`Index Sequence`, 'tp', ''))  
+  all.info = all.info %>% 
+    mutate(`Hospital Research ID` = str_replace_all( str_remove_all(`Hospital Research ID`, " "), '/', '_'), `Index Sequence` = str_replace_all(`Index Sequence`, 'tp', ''))  
 }
-
 
 print(unique(all.info$`SLX-ID`))
 

### 2020-07-14 Compare Models.Rmd
Updates to the paper, and generating a standardized patient/sample database for metadata

@@
 
-kb.perf = do.call(bind_rows, lapply(list.files(data,full.names = T), function(f) {
+kb.dirnames = grep('kb', list.files(data,full.names = T), value=T)
+
+kb.perf = do.call(bind_rows, lapply(kb.dirnames, function(f) {
   load(paste0(f,'/all.pt.alpha.Rdata'),verbose=F)
-
   nonz = names(which(sapply(performance.at.1se, nrow) > 0))
@@
 # })) %>% mutate(alpha=names(coefs), kb=basename(f))
-
   do.call(bind_rows, performance.at.1se) %>% 
@@
 kb.perf = kb.perf %>% mutate(kb = factor(kb, levels=c('15kb','50kb','100kb','500kb'), ordered=T)) %>% arrange(kb, alpha)
+
+kb.perf %>% dplyr::select(-`lambda-1se`) %>% write_tsv('plots/exfig9a.tsv')
 
@@
 ```{r fig.height=7, fig.width=8}
-rc9 = do.call(bind_rows, lapply(list.files(data,full.names = T), function(f) {
+threshold = 'best'
+rc9 = do.call(bind_rows, lapply(kb.dirnames, function(f) {
   load(paste0(f,'/loo_0.9.Rdata'),verbose=F)
-  roc = pROC::roc(Status ~  Prediction, data=pg.samp, auc=T, ci=T, of='thresholds',transpose=T, quiet=T)
-  pROC::coords(roc, "best", transpose=F) %>% mutate( auc = as.numeric(pROC::auc(roc)), kb = basename(f) )
+  roc = pROC::roc(Status ~  Prediction, data=pg.samp, auc=T, ci=T, of='thresholds',quiet=T)
+
+  bind_cols(
+    as_tibble(pROC::ci.auc(roc)) %>% mutate(m = c('auc.low', 'auc', 'auc.high')) %>% spread(m,value),
+    as_tibble(data.frame(pROC::ci.coords(roc, threshold, transpose=F))) %>% dplyr::select(-matches('threshold')) %>% 
+      set_names(c('specificity.low', 'specificity', 'specificity.high', 'sensitivity.low', 'sensitivity', 'sensitivity.high')) ) %>%
+    mutate(kb = basename(f))
 }))
 
-rc9 = rc9 %>% set_names(c('Threshold','FPR','TPR','AUC','kb')) %>% 
-  mutate(kb = factor(kb,levels=c('15kb','50kb','100kb','500kb'), ordered=T), FPR = 1-FPR) %>% arrange(kb) %>%
-  dplyr::select(-Threshold)
+rc9 = rc9 %>%  mutate(kb = factor(kb,levels=c('15kb','50kb','100kb','500kb'), ordered=T)) %>% arrange(kb)
 
-p = ggplot(melt(rc9, id.vars='kb'), aes(variable, value, group=kb)) + ylim(0,1) + 
-  geom_bar(aes(fill=kb), stat='identity', position='dodge') + 
-  geom_text(aes(label=round(value,2)), position=position_dodge(width=1)) + 
+df = rc9 %>% dplyr::select(matches('auc|kb'))
+pos = position_dodge(width=0.9)
+p = ggplot() + 
+  geom_bar(data=df, aes(x=kb, y=auc, group=kb, fill=kb), stat='identity', position = pos, show.legend = F) + 
+  geom_errorbar(data=df, aes(x=kb, ymin=auc.low, ymax=auc.high, group=kb), position = pos, width=0.2, color='grey39') +
   scale_fill_brewer(palette='Set1') + 
-  labs(title='Discovery ROC', subtitle = 'alpha=0.9', x='', y='%') + 
-  theme_minimal() + theme(text=element_text(size=12))
-p
-ggsave(filename='plots/parameters/training_kb_auc.png', plot = p, width = 6, height = 6, units = 'in', dpi = 300)
+  scale_y_continuous(expand=c(0,0), limits=c(0,1)) + 
+  labs(title='', subtitle = 'alpha=0.9', x='AUC', y='%') + 
+  theme_minimal() + theme(text=element_text(size=12), axis.text.x = element_text(angle=45, hjust=1))
 
+df = rc9 %>% dplyr::select(matches('sensitivity|kb'))
+p1 = ggplot() + 
+  geom_bar(data=df, aes(x=kb, y=sensitivity, group=kb, fill=kb), stat='identity', position = pos, show.legend = F) + 
+  geom_errorbar(data=df, aes(x=kb, ymin=sensitivity.low, ymax=sensitivity.high, group=kb), position = pos, width=0.2, color='grey39') +
+  scale_fill_brewer(palette='Set1') + 
+  scale_y_continuous(expand=c(0,0), limits=c(0,1)) + 
+  labs(title='', subtitle = '', x='TPR', y='') + 
+  theme_minimal() + theme(text=element_text(size=12), axis.text.x = element_text(angle=45, hjust=1), axis.text.y = element_blank())
+
+df = rc9 %>% dplyr::select(matches('specificity|kb'))
+p2 = ggplot() + 
+  geom_bar(data=df, aes(x=kb, y=1-specificity, group=kb, fill=kb), stat='identity', position = pos, show.legend = F) + 
+  geom_errorbar(data=df, aes(x=kb, ymin=1-specificity.low, ymax=1-specificity.high, group=kb), position = pos, width=0.2, color='grey39') +
+  scale_fill_brewer(palette='Set1') + 
+  scale_y_continuous(expand=c(0,0), limits=c(0,1)) + 
+  labs(title='', subtitle = '', x='FPR', y='') + 
+  theme_minimal() + theme(text=element_text(size=12), axis.text.x = element_text(angle=45, hjust=1), axis.text.y = element_blank())
+
+rc9 %>% dplyr::rename('auc.ci.high'='auc.high', 'auc.ci.low'='auc.low') %>% write_tsv('plots/exfig9b.tsv')
+
+grid.arrange(p,p1,p2,nrow=1)
+ggsave(filename='plots/parameters/training_kb_auc.png', plot = grid.arrange(p,p1,p2,nrow=1,top='Discovery ROC'), width = 6, height = 6, units = 'in', dpi = 300)
+# rc9 = rc9 %>% set_names(c('Threshold','FPR','TPR','AUC','AUC.CI.Low','AUC.CI.High','kb')) %>% 
+#   mutate(kb = factor(kb,levels=c('15kb','50kb','100kb','500kb'), ordered=T), FPR = 1-FPR) %>% arrange(kb) %>%
+#   dplyr::select(-Threshold)
 ```
@@
 val.preds = purrr::map(list.files(vdir, full.names=T), function(dir) {
+  print(dir)
   do.call(bind_rows, purrr::map(list.files(paste0(dir,'/predictions_5e6_all/0.9'), 'predictions.tsv', full.names = T, recursive = T), function(f) {
@@
 val.auc = do.call(bind_rows, purrr::map(val.preds, function(kb) {
-  roc = pROC::roc(Status ~ Probability, data=kb, auc=T, ci=T, of='thresholds',transpose=T, quiet=T)
-  pROC::coords(roc, "best", transpose=F) %>% mutate( auc = as.numeric(pROC::auc(roc)), kb = unique(kb$kb) )
+  roc = pROC::roc(Status ~ Probability, data=kb, auc=T, ci=T, of='thresholds', quiet=T)
+  #pROC::coords(roc, "best", transpose=F) %>% mutate( auc = as.numeric(pROC::auc(roc)), kb = unique(kb$kb) )
+
+  bind_cols(
+    as_tibble(pROC::ci.auc(roc)) %>% mutate(m = c('auc.low', 'auc', 'auc.high')) %>% spread(m,value),
+    as_tibble(data.frame(pROC::ci.coords(roc, threshold, transpose=F))) %>% dplyr::select(-matches('threshold')) %>% 
+      set_names(c('specificity.low', 'specificity', 'specificity.high', 'sensitivity.low', 'sensitivity', 'sensitivity.high')) ) %>%
+    mutate(kb = unique(kb$kb))
 }))
 
-val.auc = val.auc %>% set_names(c('Threshold','FPR','TPR','AUC','kb')) %>% 
-  mutate(kb = factor(kb,levels=c('15kb','50kb','100kb','500kb'), ordered=T), FPR = 1-FPR) %>% arrange(kb) %>%
-  dplyr::select(-Threshold)
-
-ggplot(melt(val.auc, id.vars='kb'), aes(variable, value, group=kb)) + ylim(0,1) +
-  geom_bar(aes(fill=kb), stat='identity', position='dodge') + 
-  geom_text(aes(label=round(value,2)), position=position_dodge(width=1)) + 
-  scale_fill_brewer(palette='Set1') + 
-  labs(title='Validation ROC (alpha=0.9)', x='', y='%') + theme_minimal()
-
+# val.auc = val.auc %>% set_names(c('Threshold','FPR','TPR','AUC','kb')) %>% 
+#   mutate(kb = factor(kb,levels=c('15kb','50kb','100kb','500kb'), ordered=T), FPR = 1-FPR) %>% arrange(kb) %>%
+#   dplyr::select(-Threshold)
+# 
+# ggplot(melt(val.auc, id.vars='kb'), aes(variable, value, group=kb)) + ylim(0,1) +
+#   geom_bar(aes(fill=kb), stat='identity', position='dodge') + 
+#   geom_text(aes(label=round(value,2)), position=position_dodge(width=1)) + 
+#   scale_fill_brewer(palette='Set1') + 
+#   labs(title='Validation ROC (alpha=0.9)', x='', y='%') + theme_minimal()
 ```
@@
 aucs = bind_rows(rc9 %>% mutate(Cohort = 'Discovery'), val.auc %>% mutate(Cohort = 'Validation'))
+aucs = aucs %>% dplyr::select(matches('auc|kb|Cohort')) %>% dplyr::rename('AUC'='auc') %>% 
+  mutate(kb = factor(kb, levels=c('15kb','50kb','100kb','500kb'), ordered=T))
+
+aucs %>% dplyr::rename('AUC.CI.high'= 'auc.high', 'AUC.CI.low'='auc.low')  %>% write_tsv('plots/exfig9c.tsv')
 
 p = ggplot(aucs, aes(kb, AUC, group=Cohort)) + ylim(0,1) + 
-  geom_bar(aes(fill=Cohort), stat='identity', position='dodge') + 
-  geom_text(aes(label=round(AUC,2)), position=position_dodge(width=1)) + 
+  geom_bar(aes(fill=Cohort), stat='identity', position=pos) + 
+  geom_errorbar(aes(ymin=auc.low, ymax=auc.high), width=0.2, color='grey39', stat='identity', position=pos) +
+  geom_text(aes(label=round(AUC,2), y=auc.low-0.05), position=position_dodge(width=1)) + 
   scale_fill_brewer(palette='Set1') + 
@@
 loo.pred.confidence = purrr::map( list.files(data,full.names = T), function(f) {
-  #print(f)
+  print(f)
   load(paste0(f,'/loo_0.5.Rdata'),verbose=F)
@@
 ```{r, fig.width=8, fig.height=12}
+files = grep('supp', list.files(data,full.names = T), invert = T, value = T)
 ## ALL LOO is at alpha=0.9
-loo.pred.confidence = purrr::map( list.files(data,full.names = T), function(f) {
+loo.pred.confidence = purrr::map( files, function(f) {
   #print(f)
@@
 }) 
-names(loo.pred.confidence) = list.files(data)
+names(loo.pred.confidence) = basename(files)
 
@@
 
-per.samp.train = lapply(list.files(data,full.names = T), function(f) {
+lapply(files, function(f) {
+  load(paste0(f,'/loo_0.9.Rdata'),verbose=F)
+
+  kb = basename(f)
+
+  pg.samp = pg.samp %>% mutate(quants = cut(Prediction, breaks=seq(0,1,0.1), include.lowest = T)) %>%
+    left_join(loo.pred.confidence[[kb]] %>% dplyr::select(quants, Risk), by='quants') %>%
+    mutate(Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered=T))
+
+  pg.samp %>% group_by(Status, Risk) %>% tally %>% spread(Risk,n) %>% 
+    left_join(pg.samp %>% group_by(Status) %>% tally, by='Status') %>%
+    dplyr::rename('Total' = 'n') %>% mutate(kb = kb)
+
+}) %>% bind_rows() %>% dplyr::rename('Total samples' = 'Total') %>% write_tsv('plots/exfig10a.tsv')
+
+per.samp.train = lapply(files, function(f) {
   load(paste0(f,'/loo_0.9.Rdata'),verbose=F)
@@
 })
-names(per.samp.train) = list.files(data)
+names(per.samp.train) = basename(files)
 
@@
 
+grid.arrange(p)
+
 ggsave(filename='plots/parameters/kb_classes.png', plot = p, width = 6, height = 12, units = 'in', dpi = 300)
@@
 
-```
-
-## Per-sample alpha=0.9
-
-```{r, fig.width=8, fig.height=12, eval=F}
-test.preds.per.kb09 = purrr::map(list.files(data,full.names = T), function(f) {
-  read_tsv(list.files(f, 'test_patients_preds_0.9.tsv', full.names = T), col_types = 'cdccdccclddccccccddccccdddccdddcdd') %>% dplyr::select(Hospital.Research.ID, Status, Endoscopy.Year, Pathology, Prediction, RR, Samplename, Path.ID) %>% mutate(kb = basename(f))
-}) 
-names(test.preds.per.kb09) = basename(list.files(data))
-
-test.preds.per.kb09 = lapply(test.preds.per.kb09, function(kb) {
-  kb %>% mutate(quants = cut(Prediction, breaks=seq(0,1,0.1), include.lowest = T)) %>% 
-    left_join(loo.pred.confidence[[unique(kb$kb)]] %>% dplyr::select(quants, Risk), by='quants') %>%
-    mutate(Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered=T)) 
-})
-
-test.samples = all.patient.info %>% filter(Set == 'Test') %>% group_by(Status) %>% tally
-
-test.counts09 = lapply(test.preds.per.kb09, function(kb) {
-  kb %>% group_by(Status, Risk) %>% tally %>% spread(Risk,n) %>% ungroup %>%
-    mutate(kb = unique(kb$kb)) %>% left_join(test.samples, by='Status') %>% 
-    mutate_if(is.numeric, funs(./n))
-})
-
-plist = lapply(test.counts09, function(kb) {
-  tt = unique(kb$kb)
-  kb = kb  %>% ungroup %>% dplyr::select(-kb,-n)
-
-  ggplot(melt(kb, id.vars=c('Status')), aes(variable, value, group=Status, fill=Status)) + ylim(0,1) + 
-    geom_bar(stat='identity', position='dodge') + geom_text(aes(label=round(value,2)), position = position_dodge(width=1)) +
-    #scale_fill_manual(values=c('#4575B4','#A50026'),labels=c('NP','P')) + 
-    labs(title=tt, x='', y='sample ratio') + theme_minimal()
-})
-
-do.call(grid.arrange, c(plist[c('50kb','100kb','15kb','500kb')], top='Test set predictions, alpha=0.9'))
 ```
@@
 #  pROC::coords(roc, "best", transpose=F) %>% mutate( auc = as.numeric(pROC::auc(roc)), kb = basename(f) )
+
+lapply(val.preds, function(kb) {
+  totals = kb %>% dplyr::filter(Cohort != 'Old progressor scrolls' & Pathology != 'OAC' ) %>% group_by(Status) %>% tally
+
+  kb %>% dplyr::filter(Cohort != 'Old progressor scrolls' & Pathology != 'OAC')  %>% 
+    group_by(Status, Risk) %>% tally %>% spread(Risk,n) %>% ungroup %>%
+    mutate(kb = unique(kb$kb)) %>% left_join(totals, by='Status')
+}) %>% bind_rows() %>% dplyr::rename('Total samples' = 'n') %>% dplyr::select(kb, Status, everything()) %>% write_tsv('plots/exfig10b.tsv')
 

### 2020-07-14 Paper.Rmd
Updates to the paper, and generating a standardized patient/sample database for metadata

@@
 ```{r demo}
-
 cohortSummary = sum.patient.data %>% dplyr::group_by(Status) %>% 
@@
 
+rm(cohortSummary, sample.age, fit)
 ```
@@
 cache.dir = paste(dir, 'Analysis/models_5e6_all/50kb', sep='/') 
-
 file = paste(cache.dir, 'model_data.Rdata', sep='/')
@@
 
-#melted$brks = as.integer(factor(cut(melted$value, breaks = quantile(melted$value, probs=c(seq(0,1,0.05))), include.lowest = T)))
 p = ggplot(melted, aes(Var2, Var1, fill = value)) + geom_tile() +
@@
 
-riskCols = brewer.pal(11, "RdYlBu")[c(10, 5, 1)]
-
 residuals = do.call(bind_rows, purrr::map(list.files(paste0(dir,'/Analysis/pcf_perPatient/50kb'), 'residuals', recursive = T, full.names = T), 
@@
 
+rm(p,leg, melted)
 ```
@@
 ```{r echo=F, warning=F, message=F, eval=T, fig.height=6, fig.width=10}
+rm(residuals)
+
 chr.lengths = BarrettsProgressionRisk:::chrInfo() %>% dplyr::filter(chr %in% c(1:22)) %>%
@@
 
-n.status <- c('Progressor'=paste0('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor')$variable)), ')'),
-              'Non-Progressor'=paste0('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor')$variable)),')'))
+n.status <- c('Progressor'=paste0('P\n(n=',length(unique(subset(globalCNdf, Status == 'Progressor')$variable)), ')'),
+              'Non-Progressor'=paste0('NP\n(n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor')$variable)),')'))
 
@@
 
+gp = globalCN.plot + labs(x='Chromosome', y='Relative CN', title='') + 
+  theme(legend.position='none', axis.text.y=element_text(size=8), strip.text.x=element_text(size=8), strip.text.y=element_text(size=9), title=element_text(face = 'bold'), axis.title=element_text(face="plain")) +
+  scale_y_continuous(breaks = c(-20,0,20) )
+ggsave(filename='plots/globalCN.png', plot=gp, width=10, height=2.2, units='in', dpi=300)
+
 ## NDBE
-nb.status <- c('Progressor'=paste0('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology == 'NDBE')$variable)), ')'),
-              'Non-Progressor'=paste0('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology == 'NDBE')$variable)),')'))
-
-globalBECN.plot = cn.mtn.plot(filter(globalCNdf, seg.type != 'arms' & Pathology == 'NDBE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')
-
-gp = globalCN.plot +  labs(y='mean adj.
-seg.', title='All samples') + theme(legend.position = 'none', axis.text = element_blank(), strip.text.y = element_blank(), strip.text.x = element_text(size=8)) +
-  labs(y='',x='', title='All samples')
-ggsave(filename='plots/globalCN.png', plot=gp, width=8, height=2.3, units='in', dpi=300)
-
-gbe = globalBECN.plot +theme(legend.position = 'none', axis.text = element_blank(), strip.text.y = element_blank(), strip.text.x = element_text(size=8)) + 
-  labs(y='',x='', title="Non-Dysplastic Barrett's Samples") 
-ggsave(filename='plots/globalBECN.png', plot=gbe, width=8, height=2.3, units='in', dpi=300)
-
+nb.status <- c('Progressor'=paste0('P\n(n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology == 'NDBE')$variable)), ')'),
+              'Non-Progressor'=paste0('NP\n(n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology == 'NDBE')$variable)),')'))
+
+globalBECN.plot = cn.mtn.plot(filter(globalCNdf, seg.type != 'arms' & Pathology == 'NDBE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples')
+
+gbe = globalBECN.plot + labs(y='Relative CN',x='Chromosome', title="") +
+  theme(legend.position='none', axis.text.y=element_text(size=8), strip.text.x=element_text(size=8), strip.text.y=element_text(size=9), title=element_text(face = 'bold'), axis.title=element_text(face="plain")) +
+  scale_y_continuous(breaks = c(-10,0,10)) 
+ggsave(filename='plots/globalBECN.png', plot=gbe, width=10, height=2.2, units='in', dpi=300)
+
+rm(arm.loc, cnLegend)
 ```
@@
 tt = wilcox.test(subset(complexity.score, Status == 'Progressor')$cx, subset(complexity.score, Status == 'Non-Progressor')$cx)
+
+cx.labels = complexity.score %>% group_by(Status) %>% tally() %>% spread(Status, n)
+
+complexity.score %>% dplyr::rename('Complexity Score' = 'cx') %>% dplyr::select(-Patient, -Pathology) %>% write_tsv('plots/cx_score.tsv')
+
 pv = ggplot(complexity.score, aes(Status, cx, fill=Status)) + 
   geom_jitter(width=0.2, color='grey39') + 
-  geom_boxplot(outlier.colour = NA, alpha=0.8) +
+  geom_boxplot(outlier.colour = NA, alpha=0.8, width=0.3) +
+  scale_x_discrete(breaks = unique(complexity.score$Status), 
+                   labels = c('Non-Progressor' = paste0('Non-Progressor\n(n=',cx.labels$`Non-Progressor`,')'),'Progressor' = paste0('Progressor\n(n=',cx.labels$`Progressor`,')'))) + 
   labs(x='', y='cx', title='Per sample cx scores', subtitle=paste('p-value=', format.pval(tt$p.value, digits=2))) + theme(legend.position = 'none') +
@@
 ggsave('plots/per_sample_cx_roc.png', plot=roc.plot(rocX, title='Per sample cx') + plot.theme + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
+
+rm(cx,complexity.score, tt, cx.labels, pv, rocX)
 ```
@@
 mp$plot
-
 ggsave('plots/performance/ridge_v_lasso.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)
@@
 mp$performance[,c('n.Coef','25%','50%','75%', '100%')] %>% kable(caption='Number of features stable in >=n% of folds') %>% kable_styling(full_width = F)
+
+rm(mp,most.stable)
 ```
@@
                               'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))
-rm(performance.at.1se, coefs, mp)
+rm(no.hgd.plots,performance.at.1se,coefs,mp)
 ```
@@
                               'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))
-rm(performance.at.1se, coefs, mp)
+rm(nolgd.plots,performance.at.1se,coefs,mp)
 ```
@@
 ggsave('plots/performance/compare_classification.png', plot=p, height=5, width=3, units='in', dpi=300 )
+rm(p)
 ```
@@
 load(file, verbose=F)
-rm(plots, performance.at.1se)
+rm(plots,performance.at.1se,models,cvs,labels)
 
@@
 ggsave('plots/effects.png', plot=p, width = 3, height=1.5, units='in')
+rm(p)
 
@@
 ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)
+rm(qq,p)
 
@@
 ggsave('plots/global_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)
-
-p = globalBECN.plot +  geom_rect(data=features %>% filter(gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=features %>% filter(gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')
+rm(p)
+
+p = globalBECN.plot +  geom_rect(data=features %>% filter(gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + 
+  geom_rect(data=features %>% filter(gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')
 ggsave('plots/ndbe_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)
+rm(p)
 ```
@@
 loo.coefs = nzcoefs
+rm(nzcoefs)
 
@@
 p = ggplot(dplyr::select(pg.samp, Status, Samplename, RR), aes(reorder(Samplename, -RR), 1, fill=RR)) + geom_tile() + 
-  #scale_fill_gradientn(colors=myPal) +
   scale_fill_gradient2(midpoint = 5, low = scales::muted('blue'), high = scales::muted('red')) + 
@@
 ggsave('plots/per_sample_RR.png', plot = p, width=0.5, height=3, units='in')
-
-pg.samp = pg.samp %>% mutate(Pathology = recode(Pathology, 'BE' = 'NDBE'))
-pg.samp = pg.samp %>% left_join(error %>% dplyr::select(-RR,-Status), by=c('Samplename'='Sample'))
-
-names(performance.at.1se) = names(pg.samp)
-df = as.data.frame(performance.at.1se)
-
-p = ggplot(df, aes(y=performance.at.1se, x='LOO')) + 
+rm(p)
+
+pg.samp = pg.samp %>% mutate(Pathology = recode(Pathology, 'BE' = 'NDBE')) %>%
+  left_join(error %>% dplyr::select(-RR,-Status), by=c('Samplename'='Sample'))
+
+pg.samp %>% dplyr::select(Status, -Pathology, Prediction) %>% dplyr::rename('Probability' = 'Prediction') %>%
+  dplyr::mutate(Status = recode_factor(Status, 'NP' = 'Non-Progressor', 'P' = 'Progressor')) %>%
+  write_tsv('plots/loo_all_samples.tsv')
+
+performance.at.1se = enframe(performance.at.1se, value='performance.at.1se')#names(performance.at.1se) = pg.samp$Samplename
+performance.at.1se$name = unique(pg.samp$Patient)
+p = ggplot(performance.at.1se, aes(y=performance.at.1se, x='LOO')) + 
   geom_boxplot(fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) + geom_jitter(color='grey39', width=0.3) +
-  geom_label(data=subset(df, performance.at.1se %in% range(performance.at.1se)), aes(y=performance.at.1se, label=round(performance.at.1se, 3)), fontface='bold', fill='lightblue') +
+  #geom_label(data=filter(performance.at.1se %in% range(performance.at.1se)), aes(y=performance.at.1se, label=round(performance.at.1se, 3)), fontface='bold', fill='lightblue') +
   labs(y='Classification rate', x='LOO model', title='Performance for LOO', subtitle='All samples model') +
@@
 ggsave('plots/performance_loo.png', plot=p, width=5, height=5, units='in', dpi=300)
+rm(p,performance.at.1se,plots,coefs,fits)
 ```
@@
   load(file, verbose=T)
+  rm(plots,coefs,nzcoefs,fits)
 } else {
@@
 ggsave('plots/performance/loo_perf_nohgd.png', plot=p, width=5, height=5, units='in', dpi=300)
+rm(p, perf,performance.at.1se)
 
@@
 
-rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')
+pg.sampNOHGD %>% dplyr::select(Status, -Pathology, Prediction) %>% dplyr::rename('Probability' = 'Prediction') %>%
+  dplyr::mutate(Status = recode_factor(Status, 'NP' = 'Non-Progressor', 'P' = 'Progressor')) %>%
+  write_tsv('plots/loo_noHGD_samples.tsv')
+
+rocNOHGD = pROC::roc(Status ~ Prediction, data = pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')
 roc.plot(rocNOHGD, 'Model excluding HGD/IMC',incS = T, 0.3)
@@
 ggsave(filename='plots/auc/roc_nohgd_model.png', plot=roc.plot(rocNOHGD, 'Model excluding HGD/IMC',T, 0.3) , width=6, height=6, units='in', dpi=300)
+
+rm(pg.sampNOHGD)
 ```
@@
   load(file, verbose=T)
+  rm(plots,performance.at.1se,coefs,nzcoefs,fits)
 } else {
@@
 
-#length(performance.at.1se)
-#length(unique(pg.sampNOLGD$Patient))
-
-#perf = cbind.data.frame(perf=performance.at.1se, pt=unique(pg.sampNOLGD$Patient))
-
 pg.sampNOLGD = pg.sampNOLGD %>% filter(!is.na(Prediction))
@@
 ggsave(filename='plots/auc/roc_nolgd_model.png', plot=roc.plot(rocNOLGD, 'Model excluding LGD/HGD/IMC',T, 0.3) , width=6, height=6, units='in', dpi=300)
+
+rm(pg.sampNOLGD)
 ```
@@
 round(pROC::coords(roc, 'best', transpose=F),2) %>% kable(caption = 'All samples ROC') %>% kable_styling(full_width = F)
-
-#pROC::coords(roc, 0.3, input='threshold', transpose=T)
 
@@
 ggsave(filename='plots/auc/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)
+rm(insilico.rocs, insilico.perf, blues, i.thresholds, df, p)
 ```
@@
 hgdpathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('HGD','IMC'), 1, 0) )
-
 hgdpathroc$model = 'HGD Only'
@@
 ```{r, echo=F, warning=F, message=F,fig.height=4, fig.width=4}
+
+tmp = pg.samp %>% dplyr::select(Patient, Status, Endoscopy.Year, PID, Prediction) %>%
+  arrange(Patient, Endoscopy.Year, PID)
+
+pout = tibble()
+for (pt in unique(pg.samp$Patient)) {
+  pout = pout %>% bind_rows( tmp %>% filter(Patient == pt) %>% group_by(Endoscopy.Year, PID) %>% dplyr::mutate(Endoscopy = dplyr::group_indices()) )
+}
+
+pout %>% dplyr::select(-Endoscopy.Year, -PID) %>% dplyr::rename('Probability' = 'Prediction') %>% 
+  dplyr::mutate(Status = recode_factor(Status, 'NP'='Non-Progressor', 'P'='Progressor')) %>% 
+  write_tsv('plots/aggregate_rocs.tsv')
+
 pendo = pg.samp %>% dplyr::group_by(Patient, Status, Hospital.Research.ID, PID ) %>% 
@@
 
-lastEndo = pg.samp %>% group_by(Patient, Hospital.Research.ID, Status,Endoscopy.Year, PID) %>% arrange(desc(Endoscopy.Year), PID)  %>% dplyr::summarise(avg = mean(Prediction) ) %>% ungroup %>% group_by(Patient) %>% dplyr::filter( Endoscopy.Year == max(Endoscopy.Year) ) %>% dplyr::select(-Endoscopy.Year, -PID)
+lastEndo = pg.samp %>% group_by(Patient, Hospital.Research.ID, Status,Endoscopy.Year, PID) %>% 
+  arrange(desc(Endoscopy.Year), PID) %>% dplyr::summarise(avg = mean(Prediction) ) %>% ungroup %>% 
+  group_by(Patient) %>% dplyr::filter( Endoscopy.Year == max(Endoscopy.Year) ) %>% dplyr::select(-Endoscopy.Year, -PID)
 
@@
 multi.roc.plot(list(rocEndoM, rocEndoA, rocLastEndo, rocFirstEndo, rocRandEndo), threshold = 0.3) + labs(title='Endoscopy prediction ROCs')
+
+rm(pendo, pendo.rand)
 ```
@@
 rocPtM = pROC::roc(Status ~ max, data=ptendo, ci=T, of='thresholds')
+rocPtM$model = 'Max per Patient'
 #pROC::coords(rocPtM, 'best', transpose=T)
-rocPtM$model = 'Max per Patient'
 ggsave('plots/auc/maxpt_auc.png', plot=roc.plot(rocPtM, title='Max per Patient (excl. final)', T, 0.3), width=6, height=6, units="in", limitsize=F, dpi = 300)
@@
 rocPtA = pROC::roc(Status ~ mean, data=ptendo, ci=T, of='thresholds')
+rocPtA$model = 'Mean per Patient'
 #pROC::coords(rocPtA, 'best', transpose=T))
-rocPtA$model = 'Mean per Patient'
 ggsave('plots/auc/avgpt_auc.png', plot=roc.plot(rocPtA, title='Mean per Patient (excl. final)', T, 0.3), width=6, height=6, units="in", limitsize=F, dpi = 300)
@@
 ggsave('plots/auc/path_p53_cn_auc.png', plot=comp, width=6, height=5, units="in", limitsize=F, dpi = 300)
+rm(m,cols,dodge,comp)
 ```
@@
 
+dodge = position_dodge(width=0.9)
 p = ggplot(m, aes(model, value, ymin=CI.min, ymax=CI.max, group=Measure)) + 
@@
 auc.leg = get.legend(p)
-
-#p = multi.roc.plot(list(roc,rocNOHGD), threshold = 0.3, palette = 'Dark2' ) + labs(title='LOO model comparisons', subtitle = 'threshold = 0.3')
 p
-
-#ggsave('plots/auc/loo_model_roc.png', plot=p + theme(legend.position = 'none'), height=4, width=4, units="in", limitsize=F, dpi = 300)
-
-ggsave('plots/auc/loo_model_roc.png', plot=p + theme(legend.position = 'none'), height=4, width=4, units="in", limitsize=F, dpi = 300)
-
-#ggsave('plots/auc/multi_auc.png', plot=p+theme(legend.position = 'none'), height=6, width=8, units='in',dpi=300)
+ggsave('plots/auc/loo_model_roc.png', plot=p+theme(legend.position = 'none'), height=4, width=4, units="in", limitsize=F, dpi = 300)
+
+rm(p,dodge)
 ```
@@
 roc$model = "per Sample"
-
-multi.roc.plot(list(roc,rocEndoM, rocEndoA), threshold = 0.3, palette = 'Dark2' ) + labs(title='Aggregated per Endoscopy ', subtitle = 'threshold = 0.3')
 
@@
   labs(x='', y='ROC ', title='Aggregated ROC Comparison', subtitle='threshold = 0.3') 
-
-# cols = brewer.pal(11, 'Paired')
-# dodge = position_dodge(width=0.9)
-# mm = melt(subset(m, Measure == 'AUC'), measure.vars='value') %>% 
-#   mutate(Group = case_when( grepl('Endoscopy', model, ignore.case=T) ~ 'Endoscopy', grepl('Patient',model,ignore.case=T) ~ 'Patient', TRUE ~ 'Sample')) %>%
-#   mutate(order = row_number(), Group = factor(Group, levels=c('Sample','Endoscopy','Patient'), ordered=T))
-# p = ggplot(mm, aes(x=reorder(model, order), y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Group)) + 
-#   facet_grid(~Group, scales = 'free_x', space='free_x') +
-#   geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='grey39' ) + 
-#   geom_text(aes(y=CI.min*100, label=paste(round(value,2)*100,'%',sep='')), position=dodge, vjust=3)  +
-#   scale_fill_manual(values=cols, name='') + plot.theme + 
-#   theme(legend.position='none', axis.text.x=element_text(angle=45, hjust=1)) + labs(x='', y='AUC', title='ROC Comparison') 
 auc.leg = get.legend(p)
-
 p
@@
 ggsave('plots/auc/roc_leg.png', plot=grid.arrange(auc.leg), height=1, width=4, unit='in', dpi=300)
+
+rm(rocEndoM, rocEndoA, rocLastEndo, rocFirstEndo, rocRandEndo, rocNOHGD, rocNOLGD, rocPtM, rocPtA)
 ```
@@
 cuts = seq(0,1,0.1)
-
 pg.samp = pg.samp %>% mutate(quants = cut(Prediction, breaks=cuts, include.lowest = T))
@@
 
-tt3 <- ttheme_minimal(
-  core=list(bg_params = list(fill = riskCols, col=NA, alpha=0.2),
-            fg_params=list(fontface=3)),
-  colhead=list(fg_params=list(col="black", fontface=4L)),
-  rowhead=list(fg_params=list(col=c('red3','green3', 'orange2'), fontface=3L)),
-  padding=unit(c(5,5),'mm'))
-#grid.arrange( tableGrob(annotation.table, rows=NULL, theme=tt3) )
 pred.confidence$Risk = factor(pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)
@@
 pg.samp = left_join(pg.samp, pred.confidence %>% dplyr::select(quants, Risk), by='quants')
+
+rm(pcfit, rsid)
 ```
@@
 pcal
+
+pred.confidence %>% dplyr::select(quants, NP,P,P.ratio,ci.low,ci.high,p.value,mn,Risk) %>% 
+  set_names(c('Decile', 'NP samples', 'P samples', 'P:NP', 'CI Low','CI High', 'chi p-value', 'mean(Pr)', 'Risk class')) %>%
+  write_tsv('plots/calibration_table.tsv')
 ggsave(filename='plots/pred_calib.png', plot=pcal, width=4, height=4, units='in', dpi=300)
 
-mini = pcal + labs(x='P(P)', y='P:NP', title='') + theme(legend.position = 'none') + 
+mini = pcal + labs(x='Predicted risk', y='Observed risk', title='') + theme(legend.position = 'none') + 
   scale_x_continuous(expand = c(0, 0), limits = c(-0.5, 1.5), breaks = cuts, labels = c(0,'','',0.3,'',0.5,'','','','',1)) + 
@@
   theme(axis.text = element_text(size=14))
+mini
 ggsave(filename='plots/pred_calib_mini.png', plot=mini, width=4, height=4, units='in', dpi=300)
@@
 
+rm(per.endo)
 ```
@@
 ```{r probhist, echo=F, warning=F, message=F, fig.width=5, fig.height=5}
-ggplot(pg.samp, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + 
+myPal = BarrettsProgressionRisk::riskColors()[as.character(pred.confidence$Risk)]
+
+ggplot(pg.samp, aes(Prediction)) + geom_histogram(aes(fill=..x..), color='white', breaks=cuts, show.legend = F) + 
   scale_fill_gradientn( colors=myPal) +
-  #scale_fill_distiller(palette = 'RdYlBu', name='P(P)') +
-  #scale_fill_distiller(palette = myPal) +
     geom_vline(xintercept=annotation.table$max, linetype='longdash', color='grey39') +
@@
     plot.theme + labs(title='Sample predictions', y='n Samples', x='Probability') + theme(legend.position = 'none')
+
+rm(annotation.table)
 ```
@@
 ```{r relrisk, echo=F, fig.width=6, fig.height=6}
-unadjRR = ggplot(pg.samp, aes(RR, fill=Prediction)) + 
-  geom_histogram(aes(fill=..x..), bins=10, show.legend = F) +
-  scale_fill_gradientn(colors = myPal,  name='') + 
-  labs(y='n Samples', x='log(Relative Risk)', title='Unadjusted relative risk') + plot.theme
+pg.samp %>% dplyr::select(Prediction,RR, quants, Risk)
+
+#myPal = BarrettsProgressionRisk::riskColors()[levels(pred.confidence$Risk)[sort(c(pred.confidence$Risk, pred.confidence$Risk))]]
+
+unadjRR = ggplot(pg.samp, aes(RR)) + 
+  geom_histogram(aes(fill=..x..), color='white', bins=10, show.legend = F) +
+  scale_fill_gradientn(colors = myPal,  name='log(RR)') + 
+  labs(y='n Samples', x='log(Relative risk of cancer progression)', title='Unadjusted relative risk') + plot.theme + 
+  theme(axis.text = element_text(size=12))
 unadjRR
-ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR, width=6, height=5, units='in', dpi=300)
-
-predhist = ggplot(pg.samp, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) +
-    scale_fill_gradientn(colors = myPal,  name='') + 
-    plot.theme + labs(title='Predictions, all samples model', y='n Samples', x='Probability') 
-ggsave(filename='plots/pred_hist.png', plot=predhist, width=5, height=5, units='in', dpi=300)
+
+pg.samp %>% dplyr::select(Samplename,Prediction, RR, Risk) %>% arrange(RR) %>% 
+  dplyr::rename('Probability' = 'Prediction', 'log(Relative Risk)' = 'RR', 'Risk class' = 'Risk') %>%
+  write_tsv('plots/unadjusted_RR.tsv')
+
+ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR + ggtitle(''), width=6, height=5, units='in', dpi=300)
+rm(unadjRR, myPal)
 ```
@@
 ## count arms differently than segs??
-df2 = df %>% dplyr::group_by(Status, Patient, Endoscopy.Year) %>% dplyr::summarise( 
+df2 = df %>% dplyr::group_by(Status, Patient, Endoscopy.Year, months.before.final) %>% dplyr::summarise( 
   'max.Pred'=max(Prediction),
@@
   'max.Path'=max(Pathology),
-  'CNA'=length(which(seg.type != 'arms' & cn %in% c('loss','gain'))),
-  'arms'=length(which(seg.type == 'arms' & cn %in% c('loss','gain'))),
-  'ratio'=length(which(seg.type != 'arms' & cn %in% c('loss','gain')))/length(cn)
+  '5MB windows'=length(which(seg.type != 'arms' & cn %in% c('loss','gain')))/length(unique(variable)),
+  'Chromosome arms'=length(which(seg.type == 'arms' & cn %in% c('loss','gain')))/length(unique(variable)),
+  '5MB ratio'=`5MB windows`/length(cn)
   ) %>% ungroup %>% mutate(Patient = factor(Patient))
-#as.data.frame(df2)
-#df2 = subset(df2, max.Path %in% c('BE','ID','LGD'))
 
 df2 = df2 %>% dplyr::group_by(Status, Patient) %>% dplyr::mutate(
-  'early'= Endoscopy.Year == min(Endoscopy.Year),
-  'late' = Endoscopy.Year == max(Endoscopy.Year)
+  'early'= months.before.final == max(months.before.final),   
+  'late' = months.before.final == min(months.before.final)  
+) %>% filter(n() > 1)
+
+el = df2 %>% filter(early | late)
+el2 = el %>% dplyr::group_by(Status, Patient) %>% dplyr::summarise( arms = abs(diff(`Chromosome arms`)), CNA = abs(diff(`5MB windows`)) )
+wtA = signif(wilcox.test(filter(el2, Status == 'Progressor')$arms, filter(el2, Status == 'Non-Progressor')$arms)$p.value,1)
+
+sl = df2 %>% dplyr::select(Status, Patient) %>% distinct() %>% group_by(Status) %>% tally() %>% spread(Status,n)
+status.labels <- c(
+  'Non-Progressor' = paste0('Non-Progressor\n(n=',sl$`Non-Progressor`,' patients)'),
+  'Progressor' = paste0('Progressor\n(n=',sl$Progressor,' patients)')
 )
-df2 = subset(df2, Patient %in% names(which(table(df2$Patient) > 1)))
-
-el = df2 %>% filter(early | late)
-el2 = el %>% dplyr::group_by(Status, Patient) %>% dplyr::summarise( diff(arms), diff(CNA) )
-wtA = signif(wilcox.test(filter(el2, Status == 'Progressor')$`diff(arms)`, filter(el2, Status == 'Non-Progressor')$`diff(arms)`)$p.value,3)
-#if (wtA < 0.0001) wtA = '< 0.0001'
-
-p = ggplot(df2 %>% filter(early|late), aes(Endoscopy.Year, arms, color=Patient)) + facet_grid(~Status) +
-  geom_point() +  geom_line(data=subset(df2, early | late)) + theme(legend.position = 'none', text=element_text(size=11)) + 
-  plot.theme + labs(x='Endoscopy Year', y='CN arm values', title='Initial vs final endoscopy', subtitle=paste('p-value=', wtA))
+
+filter(df2, early|late) %>% dplyr::mutate('Endoscopy' = case_when( early ~ 'Initial', TRUE ~ 'Final'  )) %>% 
+  dplyr::select(Patient, Status, everything(), -early, -late, -Endoscopy.Year, -matches('5MB'), -months.before.final) %>%
+  dplyr::rename('max(Pathology)' = 'max.Path', 'max(Pr)' = 'max.Pred', 'max(RR)' = 'max.RR') %>%
+  write_tsv('plots/arms_initial_to_final.tsv')
+
+p = ggplot(df2 %>% filter(early|late), aes(Endoscopy.Year, `Chromosome arms`, color=Patient)) + 
+  facet_grid(~Status, labeller = labeller(Status=status.labels)) +
+  geom_point() + geom_line(data=subset(df2, early | late)) + 
+  theme(legend.position = 'none', text=element_text(size=11)) + 
+  plot.theme + labs(x='Endoscopy Year', y='CN altered arms', title='Chromosome arms, initial vs final endoscopy', subtitle=paste('p-value=', wtA))
 p
-ggsave(filename='plots/status_hetero.png', plot=p, width=8, height=6, units='in', dpi=300)
-
-wtA = signif(wilcox.test(subset(df2, early &  Status == 'Progressor')$arms, subset(df2, early & Status == 'Non-Progressor')$arms)$p.value, 3)
-wtCN = signif(wilcox.test(subset(df2, early & Status == 'Progressor')$CNA, subset(df2, early & Status == 'Non-Progressor')$CNA)$p.value, 2)
-
-m = melt(subset(df2, early), measure.vars=c('CNA','arms'), id.vars=c('Status'))
-p = ggplot(m, aes(Status, value, fill=Status)) + facet_wrap(~variable, scales='free_y') + 
-  geom_boxplot(outlier.colour = NA) + geom_jitter(width=0.2) + 
-  scale_fill_brewer(palette = 'Dark2') + 
+ggsave(filename='plots/status_hetero.png', plot=p, width=6.5, height=4, units='in', dpi=300)
+rm(p,wtA)
+
+wtA = signif(wilcox.test(subset(df2, early &  Status == 'Progressor')$`Chromosome arms`, subset(df2, early & Status == 'Non-Progressor')$`Chromosome arms`)$p.value, 3)
+wtCN = signif(wilcox.test(subset(df2, early & Status == 'Progressor')$`5MB windows`, subset(df2, early & Status == 'Non-Progressor')$`5MB windows`)$p.value, 2)
+
+wt.labels <- c(
+  '5MB windows' = paste0('5MB windows\np-value=', wtCN),
+  'Chromosome arms' = paste0('Chromosome arms\np-value=',wtA)
+)
+
+m = melt(filter(df2, early), measure.vars=c('5MB windows','Chromosome arms'), id.vars=c('Status'))
+m$Status = factor(m$Status, ordered=F)
+
+m %>% as_tibble() %>% dplyr::rename('CN count' = 'value', 'Segment type' = 'variable') %>% write_tsv('plots/initial_endo.tsv')
+
+p = ggplot(m, aes(Status, value, fill=Status)) + 
+  facet_wrap(~variable, scales='free_y', labeller = labeller(variable=wt.labels)) + 
+  geom_jitter(width=0.2, color='grey39') + 
+  geom_boxplot(outlier.colour = NA, alpha=0.8, width=0.3) +
+  scale_x_discrete(breaks = as.character(unique(m$Status)), 
+                   labels = c('Non-Progressor' = paste0('Non-Progressor\n(n=',sl$`Non-Progressor`,')'),'Progressor' = paste0('Progressor\n(n=',sl$`Progressor`,')'))) + 
   plot.theme + theme(legend.position = 'none', text=element_text(size=11)) +
-  labs( y='Segment count', title='CN values, initial endoscopy', subtitle=paste('Arms p=',wtA, '   CNA p=',wtCN, sep=''))
+  labs( y='CN altered count', title='Initial endoscopy')
 p
@@
 ogjLimits = levels(factor(pg.samp$ogj, ordered = T))
+
+rm(m,p,wtA,wtCN,wt.labels,el,el2,df,df2)
+```
+
+```{r, eval=F}
+# Output every patient, only need to do this once when something changes
+for (i in sort(unique(sum.patient.data$Patient))) {
+  print(i)
+  pt = filter(sum.patient.data, Patient == i)
+  status = as.character(filter(sum.patient.data, Patient == i)$Status)
+
+  pms = mtn.spatial(filter(globalCNdf, Patient == i & seg.type != 'arms'), pg.samp, ogjLimits, paste('Patient', pt, '(', status, ')' ))
+
+  fl = paste0('patient_', i, '_',status,'.png')
+  ggsave(paste('plots/spatial/', fl, sep=''), plot=grid.arrange(pms$grob), width=10.5, height=4.5, units='in', dpi=300)
+}
 ```
@@
 ```{r}
-# for (i in sort(unique(sum.patient.data$Patient))) {
-#   print(i)
-#   pt = filter(sum.patient.data, Patient == i)
-#   status = as.character(filter(sum.patient.data, Patient == i)$Status)
-# 
-#   pms = mtn.spatial(filter(globalCNdf, Patient == i & seg.type != 'arms'), pg.samp, ogjLimits, paste('Patient', pt, '(', status, ')' ))
-# 
-#   fl = paste0('patient_', i, '_',status,'.png')
-#   ggsave(paste('plots/spatial/', fl, sep=''), plot=grid.arrange(pms$grob), width=10.5, height=4.5, units='in', dpi=300)
-# }
-p20 = mtn.spatial(globalCNdf %>% filter(Patient == 20 & seg.type != 'arms'), pg.samp, ogjLimits )
+globalCNdf %>% filter(Patient %in% c(20, 51) & seg.type != 'arms') %>% 
+  left_join(dplyr::select(pg.samp, 'Samplename','Endoscopy.Year','months.before.final','Block','ogj'), by=c('variable' = 'Samplename')) %>%
+  dplyr::select(Patient, everything(), -chr.length, -seg.type, -Hospital.Research.ID, -Pathology, -p53.Status, -variable, -cn, -Endoscopy.Year, -Block) %>%
+  dplyr::rename('Location'='locs', 'CN value'='value', 'Relative esophageal sample location (1-4)' = 'ogj', 'Months before final endoscopy' = 'months.before.final') %>%
+  write_tsv('plots/spatial_pts_20_51.tsv')
+
+p20 = mtn.spatial(globalCNdf %>% filter(Patient == 20 & seg.type != 'arms'), pg.samp, ogjLimits,axis.text.y=T )
 ggsave(filename='plots/spatial/patient_20_P.png', plot=grid.arrange(p20$grob), width=16, height=6, units='in', dpi=300)
 
-p51 = mtn.spatial(globalCNdf %>% filter(Patient == 51 & seg.type != 'arms'), pg.samp, ogjLimits,'Patient 51 (NP)' )
+p51 = mtn.spatial(globalCNdf %>% filter(Patient == 51 & seg.type != 'arms'), pg.samp, ogjLimits)
 ggsave(filename='plots/spatial/patient_51_NP.png', plot=grid.arrange(p51$grob), width=10.5, height=4.5, units='in', dpi=300)
@@
 ggsave('plots/spatial/legend.png', plot=grid.arrange(p20$legend), width=2,height=2, units='in', dpi=300)
+
+rm(p20, p51)
 ```
@@
 ```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
-rp = plot.risk.by.path(pg.samp %>% ungroup) + labs(title='Discovery cohort samples', y='% Classified')
+pg.samp %>% dplyr::group_by(Status, Pathology, Risk) %>% tally(name='Sample count') %>% ungroup %>% dplyr::rename('Risk class' = 'Risk') %>%
+  write_tsv('plots/discovery_cohort_by_path.tsv')
+
+rp = plot.risk.by.path(pg.samp %>% ungroup) + labs(title='Discovery cohort samples', y='% samples classified', x='Sample pathology') + theme(strip.text = element_text(size=14))
 rp
-ggsave(filename='plots/pred_by_path.png', plot=(rp + theme(legend.position = 'none')), width=7, height=6, units='in', dpi=300)
+ggsave(filename='plots/pred_by_path.png', plot=(rp + theme(legend.position = 'none') + ggtitle('')), width=7.5, height=6.5, units='in', dpi=300)
+rm(rp)
 ```
@@
 col = brewer.pal(11, "PRGn")[c(10,2)]
+
+p53Path %>% dplyr::select(-Freq,) %>% dplyr::rename('Sample count' = 'nPath', 'Sample ratio'='ratio') %>%
+  dplyr::mutate(p53.Status = recode_factor(p53.Status,'0'='Normal', '1'='Aberrant'), Status = 'Progressor')  %>% write_tsv('plots/p53IHC.tsv')
+
 rp = ggplot(p53Path, aes(Pathology, ratio*100, group=Pathology, fill=p53.Status)) + geom_bar(stat='identity', position = position_stack()) +
@@
 ggsave(filename='plots/p53_pred.png', plot=rp, width=6, height=5, units='in', dpi=300)
+
+rm(p53Path, col, rp)
 ```
@@
 
+endos.m %>% set_names(c('Months', 'Ratio predicted endoscopies','Sample count', 'Endoscopy count', 'mean(Pr)', 'SEM(Pr)')) %>% write_tsv('plots/pred_endo.tsv')
+
 rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) +
@@
   geom_errorbar(color='grey55', width=0.2) +
-  geom_text(aes(label=paste(round(pred.r*100), '%', sep='')), color='black', nudge_y=2) +
+  geom_text(aes(label=paste(round(pred.r*100), '%', sep=''), y=(pred.r-`sem(P)`)*100), color='white', nudge_y=-3) +
   geom_text(aes(y=5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white', nudge_y=2) + 
-  labs(x='Months prior to endpoint', y='Sensitivity', title='High risk classifications prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') + coord_cartesian(ylim = c(0,100)) + 
+  labs(x='Months prior to endpoint', y='Sensitivity for high risk samples', title='High risk classifications prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') + coord_cartesian(ylim = c(0,100)) +
   plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12) ) 
 rp
-ggsave(filename='plots/pred_endo.png', plot=rp, width=6, height=5, units='in', dpi=300)
-
+
+ggsave(filename='plots/pred_endo.png', plot=rp + ggtitle(''), width=6, height=5, units='in', dpi=300)
+rm(endos,endos.m,rp)
 ```
@@
 ```{r pred_by_time}
-pts = pg.samp %>% dplyr::group_by(Status, Patient, months.before.final, brks, PID)%>% dplyr::select(Patient, Status, months.before.final, brks, PID, Block, Endoscopy.Year, Pathology, Prediction, RR, Risk) %>% ungroup
-
-pts %<>% mutate(Patient = factor(Patient)) %>% group_by(Patient) %>% arrange(Endoscopy.Year)
+pts = pg.samp %>% dplyr::group_by(Status, Patient, months.before.final, brks, PID) %>% 
+  dplyr::select(Patient, Status, months.before.final, brks, PID, Block, Endoscopy.Year, Pathology, Prediction, RR, Risk) %>% ungroup %>% 
+  mutate(Patient = factor(Patient)) %>% group_by(Patient) %>% arrange(Endoscopy.Year)
 
@@
 time.pred %<>% filter(Patient %in% (time.pred %>% group_by(Patient) %>% tally %>% filter(n>1) %>% dplyr::select(Patient) %>% pull))
-
-#time.pred = left_join(time.pred,time.pred %>% group_by(Status, Patient) %>% dplyr::summarise(max.path = max(endo.path)), by=c('Status','Patient'))
 
@@
 
+risk.over.time %>% dplyr::select(Patient, Status, max.Prob,max.Risk, max.Path, months.since.initial, -risk.diff, -ie.risk.diff, -Endoscopy.Year, -n.samples) %>% 
+  set_names(c('Patient', 'Status', 'max(Pr)', 'max(Risk class)', 'max(Pathology)', 'Months since initial endoscopy')) %>%
+  write_tsv('plots/risk_over_time.tsv')
+
 p = ggplot(risk.over.time) + facet_wrap(~Status, scales='free', ncol = 1) +
@@
   labs(x='Months since initial endoscopy', y='Patient', title='Risks over time per endoscopy') + 
-  plot.theme 
+  plot.theme + theme(text = element_text(size=16))
 p
-ggsave(filename='plots/risks_over_time.png', plot = p, height=12, width=7, units='in')
-#ggsave(filename='plots/risks_over_time.png', plot = p, width=12, height=7, units='in')
-```
-
-```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=5 }
+ggsave(filename='plots/risks_over_time.png', plot = p + ggtitle(''), height=12, width=7, units='in')
+
+rm(p,sq,tb,pts,risk.over.time,time.pred)
+```
+
+```{r heatmaps, echo=F, warning=F, message=F, fig.width=5, fig.height=5 }
 pg.samp = pg.samp %>% dplyr::mutate(nl = factor(nl, ordered=T), 
@@
 
-p = BarrettsProgressionRisk::patientRiskTilesPlot(pg.samp %>% filter(Patient == 6) %>% dplyr::select(Endoscopy.Year,months.before.final, Pathology,Block,Risk) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block), 'months.before.final' )
+pg.samp %>% filter(Patient == 69) %>% dplyr::select(Patient, Status, Prediction, Risk, Pathology, nl, months.before.final) %>%
+  dplyr::rename('Esophageal location'='nl', 'Months before final'='months.before.final', 'Probability'='Prediction', 'Risk class'='Risk') %>%
+  write_tsv('plots/heatmaps/P/pt_69.tsv')
+
+p = BarrettsProgressionRisk::patientRiskTilesPlot(pg.samp %>% filter(Patient == 6) %>% dplyr::select(Endoscopy.Year,months.before.final, Pathology,Block,Risk) %>% 
+                                                    dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block), 'months.before.final' ) 
 tilesLegend = get.legend(p)
 p + theme(legend.position = 'none')
-
 ggsave('plots/heatmaps/legend_long.png',plot=grid.arrange(tilesLegend), width=10, height=2, units='in', dpi=300)
+rm(p,tilesLegend)
 
@@
 dir.create('plots/heatmaps/tiny/NP',recursive = T, showWarnings = F)
+
+pg.samp %>% dplyr::select(Patient,Status,Block,Risk,months.before.final) %>% arrange(Patient, months.before.final) %>%
+  dplyr::rename('Risk class' = 'Risk', 'Esophageal location' = 'Block', 'Months before final' = 'months.before.final') %>%
+  write_tsv('plots/pt_heatmaps.tsv')
+
+pg.samp %>% dplyr::select(Patient,Status,Block,Risk,Pathology, months.before.final) %>% arrange(Patient, months.before.final) %>%
+  dplyr::rename('Risk class' = 'Risk', 'Esophageal location' = 'Block', 'Months before final' = 'months.before.final') %>% arrange(Status) %>%
+  write_tsv('plots/pt_heatmaps_with_path.tsv')
+
 # Tiny heatmaps
 plist = purrr::map(sum.patient.data$Patient, function(pt) {
-  patient = pg.samp %>% filter(Patient == pt) %>% dplyr::select(Patient,Status,Endoscopy.Year,Block,Risk,months.before.final) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) %>% ungroup
+  patient = pg.samp %>% filter(Patient == pt) %>% dplyr::select(Patient,Status,Endoscopy.Year,Block,Risk,months.before.final) %>% 
+    dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) %>% ungroup
 
@@
 
-  patient = pg.samp %>% filter(Patient == pt) %>% dplyr::select(Patient,Status,Pathology,Endoscopy.Year,Block,Risk,months.before.final) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) %>% ungroup
-
-    p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final','rev', 0.5) + theme(legend.position = 'none') + labs(x='months',y='location') + theme(text = element_text(size = 10), plot.title = element_text(face = 'bold', size=16), axis.text = element_text(size = 10))
+  patient = pg.samp %>% filter(Patient == pt) %>% dplyr::select(Patient,Status,Pathology,Endoscopy.Year,Block,Risk,months.before.final) %>% 
+    dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) %>% ungroup
+
+  if (pt == 69)
+      patient = pg.samp %>% filter(Patient == pt) %>% dplyr::select(Patient,Status,Endoscopy.Year,Block,Risk,months.before.final) %>% 
+    dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) %>% ungroup
+
+    p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final','rev', 0.5) + theme(legend.position = 'none') + 
+      labs(x='Months to endpoint',y='Location: EGJ->Mouth') + 
+      theme(text = element_text(size = 12), plot.title = element_text(face = 'bold', size=16), axis.text=element_text(size = 10))
 
@@
 write_tsv(summary.t, path = 'plots/risk_summary.tsv')
+rm(summary.t)
 ```
@@
 
+rm(p1,p2, plist)
 #ggsave('plots/risk_legend.png', plot=grid.arrange(riskLegend), width=2,height=4, dpi=300)
 #ggsave('plots/risk_legend_long.png', plot=grid.arrange(riskLegendLong), width=8,height=2, dpi=300)
-
 ```
@@
 pi$p11Loss = as.numeric(pi$Samplename %in% p11Loss$variable)
-#plot(table(p11=pi$p11Loss, ihc=pi$p53.Status))
 
@@
 
-ct = fisher.test(table(pi$p11Loss, pi$p53.Status))
-
-```
-
-17p11 loss and p53 IHC correlation  Fisher's `r broom::tidy(ct) %>% kable() %>% kable_styling(full_width=F)`
+```
+
+17p11 loss and p53 IHC correlation  Fisher's `r broom::tidy(fisher.test(table(pi$p11Loss, pi$p53.Status))) %>% kable() %>% kable_styling(full_width=F)`
 
 ## q arm
+
 ```{r, echo=T}
@@
 kable(caption='Ratio of all Chr17 q11 gain samples by pathology' ) %>% kable_styling(full_width = F)
+
+rm(q11, q11gain, q11Path)
+rm(p11, p11Loss, p11Path, pi)
 ```
@@
 
-#genes = as.data.frame(genesGR[unique(subjectHits(ov))])
-#colnames(genes)[1] = c('chr')
+globalCNdf %>% filter(chr == '17')
 
 g17 = plot.cn.chr(globalCNdf, chrom='17', chr.lengths, haz=features, genes=genes, label=labeller(chr=label_both, Status=n.status)) + 
-  labs(title="Chromosome 17") + scale_y_continuous(breaks = seq(-30,25,10), labels = c('-30', '','',0,'',20)) +
-  theme(legend.position = 'none', strip.text = element_blank(), text=element_text(size=14))
+  labs(title="Chromosome 17", y='Relative CN') + scale_y_continuous(breaks = seq(-30,25,10), labels = c('-30', '','',0,'',20)) +
+  theme(legend.position = 'none',  text=element_text(size=14))
 g17
@@
 
+rm(g17)
 ```
@@
 
-#(genes[,c(1:3,6,7)]) %>% kable() %>% kable_styling(full_width=F)
+rm(ccgenes, annoG)
 ```
@@
 ```{r, echo=F}
-#p.diag = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/validation/progressor_diagnosis.xlsx') %>% 
-#  dplyr::mutate(`Hospital Research ID` = gsub('/', '_', `Hospital Research ID`), `Weeks Follow Up` = as.double(difftime(`Diagnostic Endoscopy`, `Sequenced Endoscopy`, units='weeks')))
 pastefun<-function(x) {
@@
 
+rm(residuals)
 ```
@@
 ggsave('plots/validation/validatation_roc.png', plot=roc.plot(roc) + labs(title='Validation cohort'), width=5, height=5, units='in',dpi=300)
-
-#pROC::coords(roc, 'best')
+rm(roc)
 
@@
 
+rm(totals,val.counts)
 ```
@@
 ```{r}
-rp = plot.risk.by.path(val.preds) + labs(title='Validation cohort samples', y='% Classified')  
+val.preds %>% dplyr::group_by(Status, Pathology, Risk) %>% tally(name='Sample count') %>% dplyr::rename('Risk class' = 'Risk') %>% write_tsv('plots/val_cohort_by_path.tsv')
+
+rp = plot.risk.by.path(val.preds) + labs(title='Validation cohort samples', y='% samples classified', x='Sample pathology')  + theme(strip.text = element_text(size=12))
 rp
-ggsave(filename='plots/validation/val_pred_by_path.png', plot=rp + theme(legend.position = 'none'), width=5, height=6, units='in', dpi=300)
+ggsave(filename='plots/validation/val_pred_by_path.png', plot=rp + theme(legend.position = 'none') + ggtitle(''), width=6, height=7, units='in', dpi=300)
+rm(rp)
 
@@
 val.preds = dplyr::rename(val.preds, 'Samplename'='Sample')
-
 val.preds = do.call(bind_rows, 
@@
 for (pt in unique(val.preds$Patient)) {
-#print(pt)
   patient = val.preds %>% filter(Patient == pt) %>% 
@@
 ggsave(filename='plots/validation/pred_endo.png', plot=rp, width=6, height=5, units='in', dpi=300)
+
+rm(rp,endos,endos.m,v.prog,sq)
 ```
@@
 
+riskCols = BarrettsProgressionRisk::riskColors()
 pt = 34
-
 p34 = vig.plot(df=filter(pg.samp, Patient == pt),cols=riskCols) + 
-    theme(legend.position = 'bottom', text = element_text(size=12)) + labs(title = filter(sum.patient.data, Patient == pt)$desc)
+  labs(title = filter(sum.patient.data, Patient == pt)$desc) +
+  theme(legend.position = 'bottom', text = element_text(size=12))
 
@@
 #ggsave('plots/vignette/legend_long.png', plot=grid.arrange(vig.legend), width=8, height=1)
-ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p34 + theme(legend.position = 'none'), width=6.35, height=4.5,dpi=300)
+ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p34+theme(legend.position = 'none'), width=6.35, height=4.5,dpi=300)
 
@@
 
+rm(p3, p34, p68)
+
 for (pt in c(56,20,86,23,33,49,51,6,25)) {
@@
 
-  #length = 6
-  #height = 5
-
-  p = vig.plot(df=subset(pg.samp, Patient == pt),info= subset(patient.info, Patient == pt),cols=riskCols) + 
+  p = vig.plot(df=subset(pg.samp, Patient == pt),cols=riskCols) + 
     theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) 
@@
   ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p, width=length, height=height,dpi=300)
+  rm(p,length,height)
 }
 ggsave('plots/vignette/legend.png', vigLeg, width=8, height=1, dpi=300)
+rm(vigLeg)
 
@@
 
-#pts = as.character(pg.samp %>% group_by(Patient, Endoscopy.Year) %>% tally %>% filter(n > 1) %>% dplyr::select(Patient) %>% distinct %>% pull)
-
 max.rules = purrr::map(unique(pg.samp$Patient), function(pt) {
-
     preds = pg.samp %>% filter(Patient == pt) %>% arrange(Endoscopy.Year, months.before.final)
@@
 
+    #preds = preds %>% group_by(Endoscopy.Year, months.before.final) %>% dplyr::summarise(Pathology = max(Pathology)) %>% 
+    #  left_join(endo, by=c('Endoscopy.Year','months.before.final')) %>% dplyr::rename('Endoscopy' = 'Endoscopy.Year') %>% 
+    #  dplyr::select(-Initial.Endoscopy, -Final.Endoscopy)
+
     if (nrow(endo) == 1) return(NA)
 
-    rx = BarrettsProgressionRisk:::.apply.rules(endo, 'Endoscopy', pred.confidence,'numeric' )
+    rx = BarrettsProgressionRisk:::.apply.rules(endo, 'Endoscopy', pred.confidence,'numeric')
 
@@
       mutate(Patient = pt) %>% mutate_at( vars(matches('Rule|Time')), funs(as.character) ) %>%
+    left_join((preds %>% group_by(Endoscopy.Year, months.before.final) %>% dplyr::summarise(Pathology = max(Pathology))), by=c('Endoscopy.Year','months.before.final')) %>%
     dplyr::select(Patient, months.before.final, Endoscopy.Year, Probability, everything()) 
@@
 
-penult = penult %>% filter(`Time 2` != 'Endpoint') %>% arrange(Patient, months.before.final) %>% group_by(Patient, Status) %>% dplyr::mutate(gid = 1:n() ) %>% dplyr::filter(gid == 1)
-
-penult %>% dplyr::select(Patient, Status) %>% distinct %>% group_by(Status) %>% tally %>% kable(caption='n patients with at least two consecutive endoscopies prior to endpoints') %>%
+penult = penult %>% filter(`Time 2` != 'Endpoint') %>% arrange(Patient, months.before.final) %>% 
+  group_by(Patient, Status) %>% dplyr::mutate(gid = 1:n() ) %>% dplyr::filter(gid == 1)
+
+penult %>% dplyr::select(Patient, Status) %>% distinct %>% group_by(Status) %>% tally %>% 
+  kable(caption='n patients with at least two consecutive endoscopies prior to endpoints') %>%
   kable_styling(full_width = F)
@@
 
+penult = penult %>% group_by(Status, Rule) %>% tally %>% spread(Rule,n) %>% dplyr::rename(RFA = `1`, `6-12 month recheck` = `2`, `12-24 month recheck` = `3`, `3-5 year surveillance` = `4`)
+p = melt(penult, id.vars = 'Status') %>% mutate(Status = recode(Status, NP = 'Non-Progressor' ,P = 'Progressor')) %>%
+  ggplot(aes(variable, value, group = Status)) + 
+  geom_col(aes(fill=Status), position = position_dodge()) + geom_text(aes(label=paste0('n=',value)), position = position_dodge(width = 1)) + coord_flip() +
+  scale_fill_manual(values = c('#4575B4', '#A50026'), name='' ) + plot.theme + labs(x='', y='# Patients', title='Penultimate Endoscopy Rx') + theme(legend.position = 'bottom')
+p
+ggsave('plots/penult_endo_rx.png', plot=p, width=6, height=6, units='in', dpi=300)
+rm(p)
 ```
@@
 
+second %>% arrange(Status, Rx) %>% dplyr::select(Patient, Status, Pathology, everything(), -Endoscopy.Year, -matches('Time|Rule'), -gid) %>%
+  dplyr::rename('max(Pr)' = Probability, 'Recommendation' = 'Rx', 'max(Pathology)' = 'Pathology', 'Months before endpoint' = 'months.before.final') %>% 
+  write_tsv('plots/second_endo_rx.tsv')
+
 second %>% dplyr::select(Patient, Status) %>% distinct %>% group_by(Status) %>% tally %>% 
@@
 
-second %>% group_by(Status, Rule) %>% tally %>% spread(Rule,n) %>% 
+second %>% group_by(Status, Rx) %>% tally %>% spread(Status,n) %>% 
   kable(caption='Second Endoscopy: # patients by rx rules based on the max (P) per endoscopy') %>% kable_styling(full_width=F)
 
-p = second %>% group_by(Status, Rule) %>% tally %>% spread(Rule,n) %>% dplyr::rename(RFA = `1`, `6-12 month recheck` = `2`, `12-24 month recheck` = `3`, `3-5 year surveillance` = `4`) %>% melt(id.vars = 'Status') %>% mutate(Status = recode(Status, NP = 'Non-Progressor' ,P = 'Progressor')) %>%
+p = second %>% group_by(Status, Rule) %>% tally %>% spread(Rule,n) %>% dplyr::rename(RFA = `1`, `6-12 month recheck` = `2`, `12-24 month recheck` = `3`, `3-5 year surveillance` = `4`)  %>%
+  melt(id.vars = 'Status') %>% mutate(Status = recode(Status, NP = 'Non-Progressor' ,P = 'Progressor')) %>%
   ggplot(aes(variable, value, group = Status)) + 
   geom_col(aes(fill=Status), position = position_dodge()) + geom_text(aes(label=paste0('n=',value)), position = position_dodge(width = 1)) + coord_flip() +
-  scale_fill_manual(values = c('#4575B4', '#A50026'), name='' ) + plot.theme + labs(x='', y='# Patients', title='Second Endoscopy Rx') + theme(legend.position = 'bottom')
+  scale_fill_manual(values = c('#4575B4', '#A50026'), name='' ) + plot.theme + labs(x='', y='Number of Patients', title='Second Endoscopy Rx') + theme(legend.position = 'bottom')
 p
 ggsave('plots/second_endo_rx.png', plot=p, width=6, height=6, units='in', dpi=300)
+rm(p)
 
@@
   dplyr::summarise_at(vars(months.before.final), list(~min(.), ~max(.), ~mean(.), ~median(.)))
+
 ```
@@
 dvsnp.perf = performance.at.1se[['0.9']]$mean
-rm(dysplasia.df, plots, pg.samp, models, nzcoefs, fits, performance.at.1se)
+rm(dysplasia.df, plots, models,  performance.at.1se)
 
@@
 
-dvsnp.aucs = left_join(dvsnp.preds %>% dplyr::select(Patient, type) %>% distinct %>% group_by(type) %>% tally(name='# patients'),
+dvsnp.aucs = 
+  left_join(
+  dvsnp.preds %>% dplyr::select(Patient, type) %>% distinct %>% group_by(type) %>% tally(name='# patients'),
   dvsnp.preds %>% dplyr::select(Patient, type) %>% group_by(type) %>% tally(name='# timepoint samples'), by='type') %>% 
   add_row( type = 'All (snp+sWGS)', `# patients` = dvsnp.preds %>% dplyr::select(Patient) %>% distinct %>% tally %>% pull, `# timepoint samples` = dvsnp.preds %>% dplyr::select(Patient) %>% tally %>% pull) %>%
-  arrange(type) %>% add_column( auc = c(dvsnp.roc$auc, dvsnp.snp.auc, dvsnp.dv.auc) ) %>%
-  mutate_at(vars(-auc), as.factor)
-
-dv.snp.p = ggplot(dvsnp.aucs , aes(`# patients`, auc, group=type, fill=type)) + geom_col() + 
+  arrange(type) %>% 
+  add_column( auc = c(dvsnp.roc$auc, dvsnp.snp.auc, dvsnp.dv.auc) ) %>%
+  add_column( ci.min.auc = c(min(pROC::ci.auc(dvsnp.roc$auc)), min(pROC::ci.auc(dvsnp.snp.auc)), min(pROC::ci.auc(dvsnp.dv.auc))) ) %>%
+  add_column( ci.max.auc = c(max(pROC::ci.auc(dvsnp.roc$auc)), max(pROC::ci.auc(dvsnp.snp.auc)), max(pROC::ci.auc(dvsnp.dv.auc))) ) %>%
+  mutate_at(vars(-matches('auc')), as.factor)
+
+dv.snp.p = ggplot(dvsnp.aucs , aes(`# patients`, auc, group=type, fill=type)) + 
+  geom_col() + 
+  geom_errorbar(aes(ymin=ci.min.auc, ymax=ci.max.auc), width=0.2, color='grey39') +
   geom_text(aes(label=round(auc,2), y=0.05)) + 
@@
 
+threshold = 0.3
 info = do.call(bind_rows, purrr::map(list.files(cache.dir, 'merged.Rdata', recursive = T, full.names = T), function(f) {
@@
   left_join(model.info, do.call(bind_rows, purrr::map(rocs, function(roc) {
-    add_column(auc=as.numeric(pROC::auc(roc)), pROC::coords(roc, 'best', transpose=F))  
+    add_column(auc=as.numeric(pROC::auc(roc)), pROC::coords(roc, threshold, transpose=F))  
   })), by='auc')
@@
 info = add_row(info, n.patients = d.n.patient, cv.performance = d.perf, n.coefs = nrow(d.coefs), n.samples = 773, auc = d.roc$auc,
-               threshold = pROC::coords(d.roc, 'best', transpose=F)$threshold,
-               specificity = pROC::coords(d.roc, 'best', transpose=F)$specificity,
-               sensitivity = pROC::coords(d.roc, 'best', transpose=F)$sensitivity,
+               threshold = pROC::coords(d.roc, threshold, transpose=F)$threshold,
+               specificity = pROC::coords(d.roc, threshold, transpose=F)$specificity,
+               sensitivity = pROC::coords(d.roc, threshold, transpose=F)$sensitivity,
                Cohort = 'Discovery (all)', Type = 'sWGS')
@@
 info = add_row(info, n.patients = dv.n.patient, cv.performance = dv.perf, n.coefs = nrow(dv.coefs), n.samples = dv.n.samples, auc = dv.roc$auc,
-               threshold = pROC::coords(dv.roc, 'best', transpose=F)$threshold,
-               specificity = pROC::coords(dv.roc, 'best', transpose=F)$specificity,
-               sensitivity = pROC::coords(dv.roc, 'best', transpose=F)$sensitivity,
+               threshold = pROC::coords(dv.roc, threshold, transpose=F)$threshold,
+               specificity = pROC::coords(dv.roc, threshold, transpose=F)$specificity,
+               sensitivity = pROC::coords(dv.roc, threshold, transpose=F)$sensitivity,
                Cohort = 'Discovery+Validation', Type = 'sWGS')
@@
 info = add_row(info, n.patients = dvsnp.patient, cv.performance = dvsnp.perf, n.coefs = nrow(dvsnp.coefs), n.samples = nrow(dvsnp.preds), auc = dvsnp.roc$auc,
-               threshold = pROC::coords(dvsnp.roc, 'best', transpose=F)$threshold,
-               specificity = pROC::coords(dvsnp.roc, 'best', transpose=F)$specificity,
-               sensitivity = pROC::coords(dvsnp.roc, 'best', transpose=F)$sensitivity,
+               threshold = pROC::coords(dvsnp.roc, threshold, transpose=F)$threshold,
+               specificity = pROC::coords(dvsnp.roc, threshold, transpose=F)$specificity,
+               sensitivity = pROC::coords(dvsnp.roc, threshold, transpose=F)$sensitivity,
                Cohort = 'sWGS (all) + SNP', Type = 'All')
@@
 
-x = info %>% group_by(n.patients, Cohort, Type) %>% dplyr::summarise_at(vars(-n.samples), list(~mean(.), ~sd(.))) %>% as_tibble %>% mutate(n.patients = factor(n.patients,levels=c( seq(28,88,10), 164, 413, 165, 248  ), ordered=T), Cohort = factor(Cohort, levels=c('Discovery (patient subsets)', 'Discovery (all)', 'Discovery+Validation', 'sWGS (all) + SNP'), ordered=T))
+sem<-function(x) {
+  sd(x)/sqrt(10)
+}
+# 
+# est = info  %>% dplyr::filter(!grepl('SNP', Cohort)) %>% 
+#   dplyr::mutate(n.patients = as.numeric(n.patients)) %>%
+#   group_by(n.patients, Cohort, Type) %>% 
+#   dplyr::summarise_at(vars(-n.samples,-threshold), list(~mean(.,na.rm=T))) %>% 
+#   arrange(n.patients) 
+# 
+# ggplot(est, aes(n.patients, auc)) + geom_point() + geom_smooth(method = 'lm') + xlim(20, 300) + ylim(0.6,1)
+# ggplot(est, aes(n.patients, specificity)) + geom_point() + geom_smooth(method = 'lm') + xlim(20, 300) + ylim(0.6,1)
+# ggplot(est, aes(n.patients, sensitivity)) + geom_point() + geom_smooth(method = 'lm') + xlim(20, 300) + ylim(0.6,1)
+# 
+# fit = lm(sensitivity~n.patients, data=est)
+# predict(fit, data.frame(n.patients = c(164, 190, 200, 250, 300)), interval = 'confidence' )
+# 
+# fit = lm((1-specificity)~n.patients, data=est)
+# predict(fit, data.frame(n.patients = c(190, 200, 250, 300)), interval = 'confidence' )
+# 
+# fit = lm(auc~n.patients, data=est)
+# predict(fit, data.frame(n.patients = c(190, 200, 250, 300, 325)), interval = 'confidence' )
+# 
+# fit = lm(n.patients~auc, data=est)
+# predict(fit, data.frame(auc = c(0.7, 0.89, 0.92, 0.95)), interval = 'confidence' )
+
+x = info %>% group_by(n.patients, Cohort, Type) %>% dplyr::summarise_at(vars(-n.samples,-threshold), list(~mean(.), ~sem(.))) %>% as_tibble %>% 
+  mutate(n.patients = factor(n.patients,levels=c( seq(28,88,10), 164, 413, 165, 248  ), ordered=T), Cohort = factor(Cohort, levels=c('Discovery (patient subsets)', 'Discovery (all)', 'Discovery+Validation', 'sWGS (all) + SNP'), ordered=T)) %>% arrange(n.patients) %>% dplyr::mutate(n.patients = paste0('n=',n.patients))
+
+x %>% write_tsv('plots/exdatafig6.tsv')
 
@@
   facet_grid(~Cohort, scales = 'free_x', space='free_x') + 
-  geom_col() + geom_errorbar(aes(ymin=cv.performance_mean-cv.performance_sd/2, ymax=cv.performance_mean+cv.performance_sd/2), width=0.2, color='grey39') +
+  geom_col() + geom_errorbar(aes(ymin=cv.performance_mean-cv.performance_sem, ymax=cv.performance_mean+cv.performance_sem), width=0.2, color='grey39') +
   geom_text(aes(label=round(cv.performance_mean,2)), y=0.05) + 
@@
   plot.theme + theme(legend.position = 'bottom', strip.text = element_blank(), panel.spacing = unit(0.5, 'lines')) +
-  labs(title='Cross validation performance', y='', x='# patients')
+  labs(title='Cross validation performance', y='', x='')
 
@@
   facet_grid(~Cohort, scales = 'free_x', space='free_x') + 
-  geom_col() + geom_errorbar(aes(ymin=n.coefs_mean-n.coefs_sd/2, ymax=n.coefs_mean+n.coefs_sd/2), width=0.2, color='grey39') +
+  geom_col() + geom_errorbar(aes(ymin=n.coefs_mean-n.coefs_sem, ymax=n.coefs_mean+n.coefs_sem), width=0.2, color='grey39') +
   geom_text(aes(label=round(n.coefs_mean,2)), y=5) +
@@
   plot.theme + theme(legend.position = 'none', strip.text = element_blank(), panel.spacing = unit(0.5, 'lines')) +
-  labs(title='# Coefficients',y='', x='# patients')
+  labs(title='# Coefficients',y='', x='')
 
@@
   geom_col() + 
-  geom_errorbar(aes(ymin=auc_mean-auc_sd/2, ymax=auc_mean+auc_sd/2), width=0.2, color='grey39') +
+  geom_errorbar(aes(ymin=auc_mean-auc_sem, ymax=auc_mean+auc_sem), width=0.2, color='grey39') +
   geom_text(aes(label=round(auc_mean,2)), y=0.05) +
@@
 dv.preds %>% group_by(Pathology, Risk) %>% tally %>% spread(Pathology, n) 
-
-```
-
-# Estimate BE population
-
-From the d only model, patient estimates at the second endoscopy
-
-```{r}
-pg.samp %>% filter(Patient == 3 & Endoscopy.Year == 1998) %>% dplyr::select(Pathology)
-second %>%
-
-get.path<-function(p,y) {
-  max( filter(pg.samp, Patient == p & Endoscopy.Year == y)$Pathology )
-}
-
-nps = second %>% group_by(Patient, Endoscopy.Year) %>% 
-  mutate( max.Path = get.path(Patient,Endoscopy.Year) ) %>% ungroup %>% 
-  dplyr::select(Status, max.Path) %>% filter(Status == 'NP') %>% mutate(max.Path = droplevels(max.Path))
-nps = nps$max.Path  
-#nps = rep(nps$max.Path,24)
-
-# Estimate false positive of LGD recommendations
-table(nps)['LGD']/length(nps)
-
-second %>% dplyr::mutate(Risk = case_when(Probability <= 0.3 ~ 'Low', Probability >= 0.5 ~ 'High', TRUE ~ 'Moderate')) %>%
-  group_by(Status, Risk) %>% tally %>% spread(Status,n) %>%
-  mutate(Total = sum(NP)+sum(P)) %>%
-  dplyr::mutate_at(vars(NP,P), list(~./Total)) %>% dplyr::select(-Total)
-
-## Get estimates at multiple timepoints during followup?
-
-per.endo.max = pg.samp %>% group_by(Patient, Status, PID, months.before.final, Endoscopy.Year) %>% 
-  dplyr::summarise(max.Path = max(Pathology), max.Risk = max(Risk), max.P = max(Prediction))
-
-pt = per.endo.max %>% dplyr::filter(Status == 'NP' & max.Risk == 'High') %>% dplyr::select(Patient) %>% pull
-
-lgd = per.endo.max %>% filter(Patient %in% pt) %>% filter(max.Path == 'LGD')
-per.endo.max %>% filter(Patient %in% lgd$Patient) %>% group_by(Patient) %>% filter(months.before.final == max(months.before.final)) 
-```
-
-Ross-Innes, C.
-S.
-et al.
-Evaluation of a Minimally Invasive Cell Sampling Device Coupled with Assessment of Trefoil Factor 3 Expression for Diagnosing Barrett’s Esophagus: A Multi-Center Case–Control Study.
-PLOS Med.
-12, e1001780 (2015).
-```{r }
-
-ndbe = 372
-id = 46
-lgd = 77
-hgd = 101
-
-est = tibble(
-  Pathology = c(rep('NDBE',372), rep('ID',46), rep('LGD',77),rep('HGD',101)),
-  Risk = c(rep('Low',372), rep('Low',46), rep('High',77),rep('High',101)),
-  Status = c(rep('NP',372+46), rep('P',101+77))
-)
-
-est %>% group_by(Status,Risk) %>% tally %>% spread(Status,n)
-
-x %>% mutate_at(vars(NP,P), list(~round(.*nrow(est))))
-
-n = ndbe + id + lgd + hgd
-
-i = 1000
-
-# lgd
-round(i*(lgd/n))
-
-# Estimated starting numbers from the BEST2 BE cases
-est.lgd = round(i*(lgd/n))
-est.hgd = round(i*(hgd/n))
-est.be = i - est.lgd - est.hgd
-
-# BE->HGD rate per person is 0.006
-beHGD = round(est.be*0.006)
-beLGD = round(est.be*0.03)
-
-# HGD or high risk triggers RFA so the est.hgd cases are immediately at RFA
-
-# From the all model 68% of all LGDs are HIGH risk
-round(est.lgd * 0.68) 
-# but 23% are false positives
-round(est.lgd * 0.68) * 0.23
-
-```
+```

### 2020-07-14 snp/SNP_Prediction_Analysis.Rmd
Updates to the paper, and generating a standardized patient/sample database for metadata

@@
 
-p = ggplot(qcdata.endo, aes(Purity_mean, ASCAT.SCA.ratio_mean)) + facet_grid(~SampleType) +
+plotL <- c('BE' = paste0('BE\n(n=',nrow(filter(qcdata.endo, SampleType == 'BE')),')' ),
+           'Blood/Gastric Normal' = paste0('Blood/Gastric Normal\n(n=', nrow(filter(qcdata.endo, SampleType == 'Blood/Gastric Normal')), ')' ))
+
+qcdata.endo %>% ungroup %>% dplyr::select(Status, Purity_mean, ASCAT.SCA.ratio_mean, SampleType) %>% 
+  dplyr::mutate(Status = recode(Status, 'NP' = 'Non-Progressor', 'P'='Progressor')) %>% dplyr::arrange(Status, SampleType) %>% write_tsv('plots/exfig3d.tsv')
+
+p = ggplot(qcdata.endo, aes(Purity_mean, ASCAT.SCA.ratio_mean)) + facet_grid(~SampleType, labeller = labeller(SampleType = plotL)) +
   geom_point(aes(color=ASCAT.SCA.ratio_mean < 0.01 & Purity_mean > 0.95 ), show.legend = F) + 
@@
 preds = preds %>% mutate(Pathology = factor(Pathology, levels=c('NDBE','ID','LGD','HGD','EAC'), ordered = T))
+
+preds %>% dplyr::select(Prob, RR, Risk, Pathology, Status) %>% dplyr::rename('Probability' = 'Prob', 'Relative Risk' = 'RR') %>%
+  dplyr::mutate(Status = recode(Status, 'NP'='Non-Progressor', 'P' = 'Progressor') ) %>% dplyr::arrange(Status, Pathology) %>%
+  write_tsv('plots/exfig3b.tsv')
+
 rp = plot.risk.by.path(preds) + labs(y='% Classified', title='Samples classified by pathology')
@@
 rocBE$model = 'per Biopsy'
+rocBE$n = nrow(preds)
+
+preds %>% dplyr::select(Status, Prob, RR) %>% dplyr::rename('Probability' = 'Prob', 'Relative Risk' = 'RR') %>%
+  dplyr::mutate(Status = recode(Status, 'NP'='Non-Progressor', 'P' = 'Progressor') ) %>% dplyr::arrange(Status, Probability) %>%
+  write_tsv('plots/exfig3a.tsv')
 
@@
 rocEX$model = 'HGD/IMC progressed'
+rocEX$n = nrow(pred.qc)
 
@@
 rocpp$model = 'All excl. low SCA'
+rocpp$n = nrow(filter(pred.qc, !lowsca))
 
@@
 rocexall$model = 'excl. low SCA & (NP) WGD'
+rocexall$n = nrow(filter(pred.qc, !lowsca & !wgd.NP))
 
@@
 roct1$model = 'Timepoint 1 (with exclusions)'
+roct1$n = nrow(filter(pred.qc, Timepoint == 'T1' & !lowsca & !wgd.NP))
 
 # Timepoint 2 only
-roct2 = pROC::roc(Status ~ Prob, data=subset(pred.qc, Timepoint != 'T1' & !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds')
+roct2 = pROC::roc(Status ~ Prob, data=filter(pred.qc, Timepoint != 'T1' & !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds')
 roc.plot(roct2, 'BE biopsies: Timepoint T2')
@@
 roct2$model = 'Timepoint 2  (with exclusions)'
+roct2$n = nrow(filter(pred.qc, Timepoint != 'T1' & !lowsca & !wgd.NP))
 
@@
 ) %>% dplyr::mutate_if(is.numeric, list(~round(.,2))) %>%
-  dplyr::mutate(model = factor(model, levels = c(rocBE$model, rocEX$model, rocexall$model, rocpp$model, roct1$model, roct2$model), ordered = T))
+  dplyr::mutate(model = factor(model, levels = c(rocBE$model, rocEX$model, rocexall$model, rocpp$model, roct1$model, roct2$model), ordered = T)) %>%
+  dplyr::mutate(model = paste0(model,'\n(n=',n,')'))
+
+m %>% dplyr::mutate(model = sub('\n.*', '', model)) %>% write_tsv('plots/exfig3c.tsv')
 
@@
  load(file, verbose=T)
- model.performance(performance.at.1se, coefs, folds, splits)$plot + labs(title='SNP CV models')
+
+ performance.at.1se %>% bind_rows() %>% dplyr::mutate(penalty = names(performance.at.1se)[-1], 'non-zero coefs'=sapply(coefs, nrow)[-1]) %>% write_tsv('plots/exfig4a.tsv')
+ p = model.performance(performance.at.1se, coefs, folds, splits)$plot + labs(title='SNP CV models') + labs(subtitle = paste0('(n=',nrow(df),')'))
+ p
+ ggsave('plots/snp/model_perf.png', plot=p, width=4, height=4, dpi=300, units='in')
+ rm(p)
 } else {

### 2020-07-14 lib/common_plots.R

@@ -8,13 +8,16 @@ plot.theme = theme(text=element_text(size=12), panel.background=element_blank(),
 
 get.roc.stat<-function(roc, x='best') {
   auc = c( pROC::auc(roc), range(pROC::ci.auc(roc))  )
-  sens = c(pROC::coords(roc, x, tranpose=T)[['sensitivity']], pROC::ci.se(roc,pROC::coords(roc, x, tranpose=T)[['specificity']], conf.level=0.95))
-  spec = c(pROC::coords(roc, x, tranpose=T)[['specificity']], pROC::ci.sp(roc,pROC::coords(roc, x, tranpose=T)[['sensitivity']], conf.level=0.95))
+  sens = c(pROC::coords(roc, x, transpose=T)[['sensitivity']], pROC::ci.se(roc,pROC::coords(roc, x, transpose=T)[['specificity']], conf.level=0.95))
+  spec = c(pROC::coords(roc, x, transpose=T)[['specificity']], pROC::ci.sp(roc,pROC::coords(roc, x, transpose=T)[['sensitivity']], conf.level=0.95))
   
   stat = rbind('AUC'=auc, 'Sensitivity'=sens[c(1,2,4)], 'Specificity'=spec[c(1,2,4)])
   colnames(stat) = c ('value','CI.min','CI.max')
   stat = as.data.frame(stat)
   stat$model = roc$model
+  
+  if (!is.null(roc$n) & is.numeric(roc$n)) stat$n = roc$n
+  
   stat$Measure = rownames(stat)
   stat
 }
@@ -56,7 +59,7 @@ cn.mtn.plot<-function(df, label, type='bar') {
     lowCol = rev(brewer.pal(6, "Greens"))[c(3,1)]
     # grey = brewer.pal(3,'Greys')[1]
     
-    colors = c('#74C476', 'darkblue','#9E9AC8')
+    colors = c('#74C476', 'grey44','#9E9AC8')
     
     p = p + 
       geom_rect(data=rangeDf, aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=cuts), show.legend=T ) + 
@@ -261,7 +264,7 @@ plot.cn.chr<-function(df, chrom='17', info=NULL, haz=NULL, genes=NULL, label=NUL
   }
   
   p + plot.theme + 
-    labs(x='', y='Mean adj. segmentation value') +
+    labs(x='', y='Mean adj. segmentation value', title='') +
[…]

### 2020-07-14 populate_tables_existing_cohort.R

@@ -1,0 +1,250 @@
+
+library(tidyverse)
+library(RSQLite)
+
+
+conn <- dbConnect(RSQLite::SQLite(), "~/OneDrive - MRC Cancer Unit at University of Cambridge/Progressor Study/Patient Database/BE_Progression_Project.db")
+
+delete_all<-function(conn) {
+  tables = dbListTables(conn)
+  for (t in tables) {
+    print(t)
+    dbExecute(conn, paste0("DELETE FROM ", t))
+  }
+}
+
+row_counts<-function(conn, tables) {
+  for (t in tables)
+    message(paste0(t, ': ',dbGetQuery(conn, paste0("SELECT COUNT(*) FROM ", t)) ))
+}
+
+get_patient_id<-function(conn, id) {
+  query = paste0("SELECT DISTINCT PatientID FROM AlternatePatientID WHERE AlternateID = ",shQuote(id)," AND AltIDStudy == 'excel sheet id'")
+  dbGetQuery(conn, query) %>% pull
+}
+
+
+tables = c('SequenceMetadata','EndoscopySample', 'Endoscopy','AlternatePatientID','Patient', 'sqlite_sequence')
+
+train = T
+val = T
+
+if (train) {
+  ### Training Cohort
+  all_pt = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/training/All_patient_info.xlsx', sheet = 'All combined') %>% arrange(`Patient ID`) %>%
+    separate(`Path ID`, c('PathCaseID', 'Block', 'Notes'), sep = '(_|-)', remove = F, fill='right', extra='merge') %>% 
+    dplyr::mutate_at(vars(PathCaseID, Block), list(~str_trim(., 'both'))) %>%
+    dplyr::mutate_at(vars(matches('SLX|Index')), list(~gsub(' ', '', .)))  %>%
+    rowwise() %>%
+    dplyr::mutate( Samplename = paste( paste(na.omit(c(`SLX ID 1`, `SLX ID 2`)), collapse='_'), sub('-', '_',`Plate Index`), sep='.' ) )
+  
+  demo = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/training/Demographics_full.xlsx') %>% 
[…]

### 2020-07-14 snp/snp-regression.R

@@ -251,6 +251,9 @@ features = as_tibble(hazards, rownames = 'label') %>% dplyr::rename(hazards = va
   dplyr::mutate(coef = coefs[['0.9']][,1], stability = featureStability) %>% arrange(-hazards)
 
 qq = quantile(features$hazards, seq(0,1,.15))
+
+features %>% dplyr::arrange(desc(hazards)) %>% write_tsv('plots/exfig4b.tsv')
+
 p = ggplot(features, aes(coef, hazards, col=hazards >= qq[length(qq)])) + 
   geom_point(alpha=0.8, show.legend = F) + 
   ggrepel::geom_text_repel(aes(x=coef, label=ifelse(hazards >= qq[length(qq)], label, '')),show.legend=F) +

### 2021-01-18 Paper.Rmd
...

@@
 
+globalCNdf %>% filter(seg.type != 'arms') %>% dplyr::select(Patient, Status, Pathology, variable, chr, start, end, value) %>% 
+  dplyr::rename_all(list(~str_to_title(.))) %>% dplyr::rename('Chromosome' = 'Chr', 'Sample' = 'Variable', 'CN Value' = 'Value') %>%
+  write_tsv('plots/global_cn_values.tsv')
+
 globalCN.plot = cn.mtn.plot(globalCNdf %>% filter(seg.type != 'arms'), labeller(Status=n.status, chr=label_value))
@@
 
-ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR + ggtitle(''), width=6, height=5, units='in', dpi=300)
+ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR + labs(title='', x='', y=''), width=6, height=5, units='in', dpi=300)
 rm(unadjRR, myPal)
@@
 ```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
+#tmp = pg.samp %>% dplyr::mutate(Pathology = recode(Pathology, 'HGD' = 'HGD/IMC', 'IMC' = 'HGD/IMC'))
 pg.samp %>% dplyr::group_by(Status, Pathology, Risk) %>% tally(name='Sample count') %>% ungroup %>% dplyr::rename('Risk class' = 'Risk') %>%
@@
 
-rp = plot.risk.by.path(pg.samp %>% ungroup) + labs(title='Discovery cohort samples', y='% samples classified', x='Sample pathology') + theme(strip.text = element_text(size=14))
-rp
-ggsave(filename='plots/pred_by_path.png', plot=(rp + theme(legend.position = 'none') + ggtitle('')), width=7.5, height=6.5, units='in', dpi=300)
+rp = plot.risk.by.path(pg.samp %>% ungroup, label_bars = F, ret_df = F)
+
+rp + labs(title='Discovery cohort samples', y='% samples classified', x='Sample pathology') + theme(strip.text = element_text(size=14))
+
+ggsave(filename='plots/pred_by_path.png', plot=(rp + theme(legend.position = 'none') + labs(title='', x='',y='')), width=7.5, height=6.5, units='in', dpi=300)
 rm(rp)
@@
   geom_errorbar(color='grey55', width=0.2) +
-  geom_text(aes(label=paste(round(pred.r*100), '%', sep=''), y=(pred.r-`sem(P)`)*100), color='white', nudge_y=-3) +
-  geom_text(aes(y=5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white', nudge_y=2) + 
-  labs(x='Months prior to endpoint', y='Sensitivity for high risk samples', title='High risk classifications prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') + coord_cartesian(ylim = c(0,100)) +
+  labs(x='Months prior to endpoint', y='Sensitivity for high risk samples', title='High risk classifications prior to final endoscopy') + coord_cartesian(ylim = c(0,100)) + 
   plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12) )
-rp
-
-ggsave(filename='plots/pred_endo.png', plot=rp + ggtitle(''), width=6, height=5, units='in', dpi=300)
+
+rp + geom_text(aes(label=paste(round(pred.r*100), '%', sep=''), y=(pred.r-`sem(P)`)*100), color='white', nudge_y=-3) +
+  geom_text(aes(y=5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white', nudge_y=2) + labs(subtitle='n = #endoscopies, s = #samples')
+
+ggsave(filename='plots/pred_endo.png', plot=rp + labs(title='', x='', y=''), width=6, height=5, units='in', dpi=300)
 rm(endos,endos.m,rp)
@@
 
-p = ggplot(risk.over.time) + facet_wrap(~Status, scales='free', ncol = 1) +
+p = ggplot(risk.over.time) + facet_wrap(~Status, scales='free', ncol = 2) +
   geom_line(aes(x=months.since.initial, y=reorder(Patient, desc(ie.risk.diff)), color=ie.risk.diff)) + 
@@
   labs(x='Months since initial endoscopy', y='Patient', title='Risks over time per endoscopy') + 
-  plot.theme + theme(text = element_text(size=16))
+  plot.theme + theme(text=element_text(size=16), panel.spacing.x=unit(1,'lines')  )
 p
-ggsave(filename='plots/risks_over_time.png', plot = p + ggtitle(''), height=12, width=7, units='in')
+
+legend = get.legend(p)
+
+ggsave(filename='plots/risks_over_time.png', plot = p + labs(title='',x='',y='') + theme(legend.position='none'), height=8, width=12, units='in')
+
+ggsave(filename='plots/risks_over_time_legend.png', plot = grid.arrange(legend), height=5, width=2, units='in')
 
@@
 #ggsave('plots/risk_legend_long.png', plot=grid.arrange(riskLegendLong), width=8,height=2, dpi=300)
+```
+
+# How predictive are the strongest coefs?
+```{r}
+features = features %>% arrange(desc(hazards))
+
+ftsamples = tibble(); ftpatients = tibble()
+for (i in 1:nrow(features)) {
+  ft = features %>% dplyr::slice(i)
+  ft_cn = globalCNdf %>% filter(chr == ft$chr & start >= ft$start & end <= ft$end & 
+                      seg.type == case_when((ft$end-ft$start) > 5e6 ~ 'arms', TRUE ~ '5MB') &
+                      as.character(cn) == as.character(ft$gl))
+
+  ftsamples = bind_rows(ftsamples, ft_cn %>% dplyr::group_by(Status) %>% tally() %>% spread(Status, n) %>% 
+    dplyr::mutate('Sample counts for:' = paste0(ft$gl, ' of ', ft$label)) %>% 
+    dplyr::select('Sample counts for:', everything()))
+
+  ftpatients = bind_rows(ftpatients, ft_cn %>% dplyr::select(Patient, Status) %>% distinct %>% 
+    group_by(Status) %>%  tally() %>% spread(Status, n) %>%
+    dplyr::mutate('Patient counts for:' = paste0(ft$gl, ' of ', ft$label)) %>% 
+    dplyr::select('Patient counts for:', everything()))
+}
+
+ftsamples = ftsamples %>% dplyr::mutate_all(list(~replace_na(.,0)))
+ftpatients = ftpatients %>% dplyr::mutate_all(list(~replace_na(.,0)))
+
+ftpatients %>% arrange(desc(Progressor))
+ftsamples %>% arrange(desc(Progressor))
+
 ```
@@
 
-rp = plot.risk.by.path(val.preds) + labs(title='Validation cohort samples', y='% samples classified', x='Sample pathology')  + theme(strip.text = element_text(size=12))
+plot.risk.by.path(val.preds, label_bars = T) + labs(title='Validation cohort samples', y='% samples classified', x='Sample pathology')  + theme(strip.text = element_text(size=12))
+
+rp = plot.risk.by.path(val.preds, label_bars = F) + labs(title='Validation cohort samples', y='% samples classified', x='Sample pathology')  + theme(strip.text = element_text(size=12))
 rp
-ggsave(filename='plots/validation/val_pred_by_path.png', plot=rp + theme(legend.position = 'none') + ggtitle(''), width=6, height=7, units='in', dpi=300)
+ggsave(filename='plots/validation/val_pred_by_path.png', plot=rp + theme(legend.position = 'none') + labs(title='',x='',y=''), width=6, height=7, units='in', dpi=300)
 rm(rp)
@@
   melt(id.vars = 'Status') %>% mutate(Status = recode(Status, NP = 'Non-Progressor' ,P = 'Progressor')) %>%
-  ggplot(aes(variable, value, group = Status)) + 
-  geom_col(aes(fill=Status), position = position_dodge()) + geom_text(aes(label=paste0('n=',value)), position = position_dodge(width = 1)) + coord_flip() +
-  scale_fill_manual(values = c('#4575B4', '#A50026'), name='' ) + plot.theme + labs(x='', y='Number of Patients', title='Second Endoscopy Rx') + theme(legend.position = 'bottom')
-p
+  ggplot(aes(variable, value, group = Status)) + scale_y_continuous(expand=c(0,0), limits=c(0,25)) +
+  geom_col(aes(fill=Status), position = position_dodge()) + 
+  coord_flip() +
+  scale_fill_manual(values = c('#4575B4', '#A50026'), name='' ) + plot.theme + 
+  labs(x='', y='Number of Patients', title='Second Endoscopy Rx') + theme(legend.position = 'bottom')
+p + geom_text(aes(label=paste0('n=',value)), position = position_dodge(width = 1))
+
 ggsave('plots/second_endo_rx.png', plot=p, width=6, height=6, units='in', dpi=300)

### 2022-06-09 Paper.Rmd
minor updates  for tidy

@@
   dplyr::mutate_at(vars(start, end), list(as.numeric)) %>% 
-  mutate(chr = factor(chr,levels=c(1:22))) %>% 
+  mutate(chr = factor(chr, levels=c(1:22), ordered=T)) %>% 
   dplyr::mutate(ymax = ifelse(coef > 0, max(globalCN.plot$data$mean.value), min(globalCN.plot$data$mean.value)), gl = factor(ifelse(coef > 0, 'gain','loss'))) %>%
-  left_join(chr.lengths %>% dplyr::select(chr, chr.length), by=c('chr')) %>% mutate(chr = factor(chr, levels=c(1:22), ordered=T))
+  left_join(chr.lengths %>% dplyr::select(chrom, chr.length), by=c('chr'='chrom')) %>% mutate(chr = factor(chr, levels=c(1:22), ordered=T))
 
@@
 i.thresholds = do.call(bind_rows, purrr::map(insilico.rocs, function(r) pROC::coords(r, 'best') %>% as_tibble %>% 
-                                               mutate(AUC = r$auc, model=r$model, is=as.character(r$is)))) %>%
+                                               mutate(AUC = as.numeric(r$auc), model=r$model, is=as.character(r$is))) ) %>%
   rename_at(vars(matches('thresh|spec|sens')), funs(str_to_title))
@@
     geom_label_repel(data=filter(i.thresholds, AUC %in% range(AUC) | is == 'discovery'), aes(label=paste0('FPR ', round(1-Specificity,2), ' TPR ', round(Sensitivity,2), '\nAUC ',round(AUC,2))), color='black') +
-    #scale_color_brewer(palette='Blues') +
+    scale_color_brewer(palette='Blues') +
     labs(title='Discovery cohort: 50 models', subtitle='18:70 random patient splits', x='Specificity (1-FPR)', y='Sensitivity (TPR)')  + plot.theme +
@@
 ```{r, echo=F, warning=F, message=F,fig.height=4, fig.width=4}
-
 tmp = pg.samp %>% dplyr::select(Patient, Status, Endoscopy.Year, PID, Prediction) %>%
@@
       get.roc.stat(roc,0.3),
-      get.roc.stat(rocNOHGD,0.3)
+#      get.roc.stat(rocNOHGD,0.3)
 #      get.roc.stat(rocNOLGD)
@@
 
-rm(rocEndoM, rocEndoA, rocLastEndo, rocFirstEndo, rocRandEndo, rocNOHGD, rocNOLGD, rocPtM, rocPtA)
+rm(rocEndoM, rocEndoA, rocLastEndo, rocFirstEndo, rocRandEndo, rocPtM, rocPtA)
 ```
@@
 model.dir = '~/Data/BarrettsProgressionRisk/Analysis/models_5e6_all/50kb/'
-load(list.files(model.dir, 'loo_0.9.Rdata',recursive = T, full.names = T))
+load(list.files(model.dir, 'loo_0.9.Rdata',recursive = T, full.names = T)[2], verbose=T)
+#load(list.files(model.dir, 'loo_0.9.Rdata',recursive = T, full.names = T))
 d.n.patient = length(unique(pg.samp$Patient))
@@
 
-load(list.files(model.dir, 'all.pt.alpha',recursive = T, full.names = T))
+load(list.files(model.dir, 'all.pt.alpha',recursive = T, full.names = T), verbose=T)
 d.coefs = coefs[['0.9']]
@@
 # 
-# est = info  %>% dplyr::filter(!grepl('SNP', Cohort)) %>% 
-#   dplyr::mutate(n.patients = as.numeric(n.patients)) %>%
-#   group_by(n.patients, Cohort, Type) %>% 
-#   dplyr::summarise_at(vars(-n.samples,-threshold), list(~mean(.,na.rm=T))) %>% 
-#   arrange(n.patients) 
+est = info  %>% dplyr::filter(!grepl('SNP', Cohort)) %>%
+  dplyr::mutate(n.patients = as.numeric(n.patients)) %>%
+  group_by(n.patients, Cohort, Type) %>%
+  dplyr::summarise_at(vars(-n.samples,-threshold), list(~mean(.,na.rm=T))) %>%
+  arrange(n.patients) %>% ungroup %>%
+  dplyr::mutate(idx = row_number())
 # 
 # ggplot(est, aes(n.patients, auc)) + geom_point() + geom_smooth(method = 'lm') + xlim(20, 300) + ylim(0.6,1)
-# ggplot(est, aes(n.patients, specificity)) + geom_point() + geom_smooth(method = 'lm') + xlim(20, 300) + ylim(0.6,1)
+# ggplot(est, aes(n.patients, specificity)) + geom_point() + geom_smooth(method = 'lm')  + ylim(0.6,1)
 # ggplot(est, aes(n.patients, sensitivity)) + geom_point() + geom_smooth(method = 'lm') + xlim(20, 300) + ylim(0.6,1)
 # 
-# fit = lm(sensitivity~n.patients, data=est)
-# predict(fit, data.frame(n.patients = c(164, 190, 200, 250, 300)), interval = 'confidence' )
+fit = lm(sensitivity~n.patients, data=est)
+predict(fit, data.frame(n.patients = c(180, 190, 200, 220, 250, 275, 300)), interval = 'confidence', type = 'response' )
 # 
-# fit = lm((1-specificity)~n.patients, data=est)
-# predict(fit, data.frame(n.patients = c(190, 200, 250, 300)), interval = 'confidence' )
+fit = lm((specificity)~n.patients, data=est)
+predict(fit, data.frame(n.patients = c(180, 190, 200, 220, 250, 275, 300)), interval = 'confidence', type = 'response' )
 # 
-# fit = lm(auc~n.patients, data=est)
-# predict(fit, data.frame(n.patients = c(190, 200, 250, 300, 325)), interval = 'confidence' )
+fit = lm(auc~n.patients, data=est)
+predict(fit, data.frame(n.patients = c(180, 190, 200, 220, 250, 275, 300)), interval = 'confidence', type = 'response' )
+
+fit = lm(n.coefs~n.patients, data=est)
+autoplot(fit)
+predict(fit, data.frame(n.patients = c(180, 190, 200, 220, 250, 275, 300)), interval = 'confidence', type = 'response' )
+
 # 
-# fit = lm(n.patients~auc, data=est)
-# predict(fit, data.frame(auc = c(0.7, 0.89, 0.92, 0.95)), interval = 'confidence' )
+#fit = lm(n.patients~auc, data=est)
+#predict(fit, data.frame(auc = c(0.7, 0.89, 0.92, 0.95, 0.97)), interval = 'confidence' )
 

### 2022-06-09 populate_tables_existing_cohort.R

@@ -3,7 +3,8 @@ library(tidyverse)
 library(RSQLite)
 
 
-conn <- dbConnect(RSQLite::SQLite(), "~/OneDrive - MRC Cancer Unit at University of Cambridge/Progressor Study/Patient Database/BE_Progression_Project.db")
+#conn <- dbConnect(RSQLite::SQLite(), "~/OneDrive - MRC Cancer Unit at University of Cambridge/Progressor Study/Patient Database/BE_Progression_Project.db")
+conn <- dbConnect(RSQLite::SQLite(), "~/workspace/BE_Patient_SampleDB/BE_Progression_Project.db")
 
 delete_all<-function(conn) {
   tables = dbListTables(conn)
@@ -31,10 +32,12 @@ last_inserted_rows<-function(conn, table, nRows) {
 
 tables = c('HistoPathSlides', 'SequenceMetadata','EndoscopySample', 'Endoscopy','AlternatePatientID','Patient', 'sqlite_sequence')
 
+delete_all(conn)
+
 train = T
 val = T
 
-slides = readxl::read_xlsx('~/OneDrive - MRC Cancer Unit at University of Cambridge/slide_matching.xlsx')
+slides = readxl::read_xlsx('~/OneDrive - MRC Cancer Unit at University of Cambridge/slide_matching2.xlsx')
 
 if (train) {
   ### Training Cohort
@@ -50,7 +53,7 @@ if (train) {
   
   ## Patient table
   for (id in unique(all_pt$`Patient ID`)) {
-    print(id)  
+    #print(id)  
     pt = all_pt %>% filter(`Patient ID` == id) 
     did = pt %>% dplyr::select(`Hospital Research ID`) %>% distinct() %>% pull
     dem = demo %>% filter(`Study Number` == did | `Alternate Study Number` == did) %>% dplyr::select(`P/NP`, Sex, `Age at diagnosis`, `Smoking Status`,`further smoking info`, `Height (cm)`, `Weight (kg)`, `Year of 'Endpoint'`, `Date of First Endoscopy`, Circumference, Maximal)
@@ -92,8 +95,7 @@ if (train) {
   
   ## Endoscopy & EndoscopySample
   for (id in unique(all_pt$`Patient ID`)) {
-    print(id)  
-    
+    #print(id)  
     pid = get_patient_id(conn, id)
     
[…]
